### 2010-12-09 paper/paper.tex
add sketch of paper describing functional (and put plots into it).
Tested-on: collins (Linux 2.6.32-5-amd64)

% TODO:

% Write ``Liquid-vapor interface'' section A.

% Look up question-mark references (citations).

\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Definitions
\def\re{\text{Re}}
\def\im{\text{Im}}
\def\ket#1{\vert #1 \rangle}
\def\cU{{\cal{U}}}
\def\cD{{\cal{D}}}
\def\re{\text{Re}}
\def\im{\text{Im}}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\refnote}{\red{[ref]}}

\newcommand{\fixme}[1]{\red{[#1]}}

% needsworklater is used to annotate bits that need work, but that we
% can postpone for a while.
\newcommand{\needsworklater}[1]{\emph{[#1]}}
% needsworknow is intended to prioritize stuff that needs fixing.
\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{A Classical Density-Functional Theory for Describing Water Interfaces}

\author{Dennis L. Jackson}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\needsworklater{ We develop a classical density functional theory for
  water by combining the fundamental-measure theory functional--which
  describes hard sphere fluids--with attractive interactions.  Using
  this approximation, we are able to describe various properties of
  water near liquid-vapor and liquid-solid interfaces.  We plan to use
  this functional as the foundation for a efficient method for
  predicting the interactions of chemicals in aqueous solution on an
  atomic level.}
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\subsection{Water---the universal solvent}

A large fraction of interesting chemistry---including all of molecular
biology---takes place in aqueous solution.  However, while quantum chemistry
now enables us to calculate the ground state energies of large molecules in
vacuum, prediction of the free energy of even the smallest molecules in the
presence of a solvent poses a continuing challenge due to the complex
structure of a liquid and the computational cost of \emph{ab initio}
molecular dynamics~\cite{car1985, grossman2004}.  The current state-of-the
art in \emph{ab initio} molecular dynamics is limited to around 64 water
molecules per unit cell, which places the minimum solute concentration at
0.84 molar and allows contact between the second solvation shell of
neighboring solutes~\cite{izvekov2005, choe2007}.  A more efficient
approach is needed in order to study nanoscale and hydrophobic solutes.

Water stands out among liquids for its extraordinary properties.  Water is
the only natural substance that is found in solid, liquid and gaseous
states at temperatures and pressures normally found on Earth.  Water has a
density maximum at 4$^\circ$C, and expands upon freezing.  The heat
capacity of water is the highest of any liquid, and its latent heat of
fusion and vaporization are also the highest.  Water's thermal conductivity
and surface tensions are the highest of any non-metallic liquid.  Water has
an unusually high dielectric constant.
Much of the unusual behavior of water is explained by the ability of each
water molecule to form \emph{four} strong, directional hydrogen bonds---two
acceptors and two donors.  This directional bonding dictates the open
structure of ice, leading to thermal expansion upon freezing.

Due to its large dielectric constant, small molecular size and strong
hydrogen-bonding network, water as ``the universal solvent'' are as
interesting as its anomalous thermodynamic properties.

\begin{figure}
\includegraphics[width=6cm, angle=270]{figs/correlation}
\caption{ Experimental atom-atom radial distribution functions for water at
room temperature and pressure from reference~\cite{Soper2000}.  }
\label{correlation}
\end{figure}

\subsection{Classical density-functional theory}

Numerous approaches have been developed to approximate the effect of water
as a solvent in atomistic calculations.  Each of these approaches gives
adequate description of some aspect of interactions with water, but none of
them is adequate for describing \emph{all} these interactions with an
accuracy close to that attained by \emph{ab initio} calculations.  The
theory of Lum, Chandler and Weeks (LCW)~\cite{LCW}, for instance, can
accurately describe the free energy cost of creating a cavity by placing a
solute in water, but does not lend itself to extensions treating the strong
interaction of water with hydrophilic solutes.  Treatment of water as a
continuum dielectric with a cavity surrounding each solute can give
accurate predictions of the energy of solvation of ions~\cite{latimer1939,
rashin1985, zhan1998, hsu1999, hildebrandt2004, hildebrandt2007}, but
provides no information about the size of this cavity.  In a physically
correct approach, the size of the cavity will naturally arise from a
balance between the free energy required to create the cavity, the
attraction between the water and the solute, and the steric repulsion which
opens up the cavity in the first place.

A very promising approach for an efficient continuum description of water
is that of classical density-functional theory (DFT), which is an approach
for evaluating the free energy and thermally averaged density of fluids in
an arbitrary external potential~\cite{ebner1976}.  Classical DFT is a
natural framework for creating a more flexible theory of hydrophobicity
that can describe interaction of water with arbitrary external
potentials---including potentials describing a \emph{strong} interaction
with a solute.

A number of exact properties are easily achieved in the density-functional
framework, such as the relationship between the pressure on a hard wall and
the contact density---which ensures a correct solvation free energy for
small hard solutes.  Much of the research on classical density-functional
theory has focussed on the hard-sphere fluid~\cite{curtin1985,
rosenfeld1989, rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.
This has led to a large number of very sophisticated functionals, such as
the fundamental-measure theory (FMT) functional~\cite{rosenfeld1989,
rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.  This functional
\needsworklater{has many nice properties}.

Several classical density functionals have already been developed for
water~\cite{ding1987, Yang1992, Jaqaman2004}, capturing some of the
qualitative behavior of water.  However, many of these are less
sophisticated than those describing the hard-sphere fluid, and a
quantitatively accurate functional describing water---which reproduces both
the macroscopic surface tension and its decrease as curvature
increases---has not yet been published.  For instance, Jaqaman \emph{et al}
developed a classical density-functional for water including orientational
effects, but their functional resulted in a surface tension twice the
experimental value, with the result that it can really only be used to make
very qualitative predictions~\cite{Jaqaman2004}.

\needsworknow{Add discussion and citations for more recent papers}

One of the most interesting functionals for water is that of Ding \emph{et
al}, which was developed in order to explain the freezing of
water~\cite{ding1987}.  This functional uses the approach of Chandler
\emph{et al}, which uses separate atomic densities for the constituent
atoms of a molecular fluid~\cite{chandler1986a, chandler1986b}.  Although
this is a compelling approach, the actual functional developed cannot
describe the vapor state or liquid-vapor interface, both of which are
crucial to correct predictions of hydrophobic behavior at large length
scales.  In addition, the functional of Ding \emph{et al} also fails to
rigorously enforce intramolecular correlations, which casts doubt on its
suitability for describing interactions with ionic solutes.

\needsworklater{
Density-Functional Theory is most commonly used for calculating the
densities of electrons in quantum mechanical systems.  We take the
classical limit of this theory to construct a functional that will predict
the density of water molecules in a classical system.  The model used here
starts with a treatment of each individual water molecule as a purely
classical hard sphere.  The only constraint in this model is
the minimum distance between any two molecules is twice the hard sphere
radius.}

\needsworklater{ I still need to rewrite the entire introduction... }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Theory and Methods}
We introduce a new classical density functional for water that
\needsworklater{predicts the vapor density, liquid density, bulk
  modulus, and bulk surface tension, as well as allows for
  liquid-vapor coexistence}.  The Helmholtz free energy functional is
based on SAFT, and is composed of the usual terms:
\begin{equation}
  F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
\end{equation}
where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
a hard-sphere free energy, $F_\textit{assoc}$ is the free energy of
association, and $F_\textit{disp}$ is the dispersion energy.  In
addition, a chemical potential term is used, so we actually work in
the grand canonical ensemble.  In the following sections, we will
introduce the terms of this functional.

\subsection{Ideal gas functional}
The first term is the ideal gas free energy functional,
which incorporates effects of kinetic energy.  The ideal gas
functional is given by
\begin{equation}\label{idealgas}
  F_{id}[n] = k_B T \int n(\xx)\left( \ln{n(\xx)} - 1\right) d\xx
\end{equation}
where n(\xx) is the density of water molecules.  The ideal gas free
energy functional has the property that the contact density at a hard
surface $n_c$ is proportional to the pressure on that surface,
according to the equation
\begin{equation}
  p = n_c k_BT \:,
\end{equation}
which also leads to the property that the free energy of solvation of
a hard solute is proportional to its volume in the limit of small
solutes.  These properties are retained by the total functional,
provided the remaining terms are \emph{purely
  nonlocal}~\cite{ashcroft?}.

\subsection{Hard-sphere repulsion}
The repulsive intermolecular interactions are modeled using a
Fundamental\nobreakdash-Measure Theory~\cite{rosenfeld1997}(FMT) functional,
which combines scaled-particle theory and the Percus-Yevick theory,
and an attractive interaction.  The FMT functional treats
each water molecule as a hard sphere with a radius approximately that of
a water molecule $(R~=~2.3~bohr)$.\\

We model the repulsive interactions using the White Bear version of
the Fundamental-Measure Theory~(FMT) functional published in
reference~\cite{roth2002whitebear}.  \needsworklater{The FMT
  functional describes the excess free energy of a hard-sphere
  fluid. Fundamental-Measure Theory has several advantages over other
  approaches, one of the most important being the simplicity of
  calculations.  This particular FMT also reliably predicts bulk
  properties of fluids in one, two and three dimensions, as well as
  exactly determines the excess free energy in the zero dimensional
  limit.}

The hard sphere functional is adapted from that of
Rosenfeld~\cite{rosenfeld1997},
and is given by
\begin{equation}
F_{hs}[n] = k_B T \int (\Phi_1 + \Phi_2 + \Phi_3) d\xx \; ,
\end{equation}
with integrands
\begin{align}
\Phi_1 &= -n\circ w_2 \ln\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)\\
\Phi_2 &= \frac{4\pi R^3 \lbrack (n\circ w_2)^2 - (n\circ \mathbf{w}_3)^2\rbrack}
          {\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)}\\
\Phi_3 &= \frac{8\pi^2 R^6\lbrack \frac{1}{3}(n\circ w_2)^3 - (n\circ
          w_2)(n\circ \mathbf{w}_3)^2\rbrack}
          {\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)} \; ,
\end{align}
using the normalized weighting functions 
\begin{align}
w_1(\xx) &= \frac{3}{4\pi R^3}\Theta(R-|\xx|)\\
w_2(\xx) &= \frac{1}{4\pi R^2}\delta(R-|\xx|)\\
\mathbf{w}_3(\xx) &= \frac{-1}{4\pi R^2}\vec{\nabla}\Theta(R-|\xx|) \; .
\end{align}

\needsworknow{Think about the HS functional in
  Reference~\cite{tarazona2000}, which claims to have fixed nasty
  singularities!}

\subsection{Association free energy}
To the purely repulsive hard-sphere free energy, we add two attractive
energy terms.  The first is the free energy of \emph{association},
which describes the effects of hydrogen bonds.  These are modeled as
four attractive patches (``association sites'') on the surface of the
hard sphere.  These four sites represent the two hydrogens and two
electron lone pairs.  There is an attractive energy
$\epsilon_\textit{assoc}$ when two molecules are arranged such that
the hydrogen of one interacts with the lone pair of the other.  The
volume of this interaction is $\kappa_\textit{assoc}$.  Thus the
association term in the free energy has two free parameters that must
be fit based on experimental data.

\subsection{Dispersion free energy}
The final term in the free energy is the dispersion term.  We use a
dispersion term based on the SAFT-VR approach, which has two free
parameters, an interaction energy $\epsilon_\textit{disp}$ and a
length scale $\lambda_\textit{disp}$.

\begin{figure}
\begin{center}
%\includegraphics[width=\columnwidth]{free-energy/free-energy}
\end{center}
\caption{By plotting the free energy vs. density it is possible to
  determine a liquid or vapor state by the energy minima.  We can adjust the
  available free parameters to ensure that liquid and vapor coexist with
  the same minimum free energy.  }
\label{fig:homogeneous}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/equation-of-state}
\end{center}
\caption{The theoretical versus experimental phase diagram, vapor
  pressure versus temperature.  }
\label{fig:equation-of-state}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/saturated-liquid-density}
\end{center}
\caption{The theoretical versus experimental saturated liquid density
  versus temperature.  }
\label{fig:saturated-liquid-density}
\end{figure}

\subsection{Determining the empirical parameters}\label{sec:empirical}

\begin{table}
\begin{tabular}{ccc}
  Fitting Parameters& \hspace{2em} & Experimental Values\\
\begin{tabular}{cc}
  $R$ & 2.3 $a_o$ \\
  $A$ & -0.0100909 Hartree \\
  $\gamma$ & 5\\
  $V$ & 515.792 $a_o^3$  \\
  $b$ & 0.75 $a_o$ \\
  $\mu$ & -0.014063 Hartree 
\end{tabular}
&& 
\begin{tabular}{cc}
 $n_l$ & 4.939 $\times 10^{-3} a_o^{-3}$\\
 $n_v$ & 1.141 $\times 10^{-7} a_o^{-3}$\\
 $\sigma$ & ??? \\
 $B$ & $7.435\times 10^{-5}$ Hartree$/a_o^3$\\
 $T$ & 298.15 K \\
\end{tabular}
\end{tabular}

\caption{Empirical parameters of optimized functional and experimental
  values matched, where $a_o$ is the Bohr radius, and $\sigma$ is the
  surface tension.  This table is not yet complete, but it should
  be soon...  We'll also need to think about the best units to use
  here.  Atom units might be best, but it also might be better to
  consistently use eV and angstrom or something.  I'm not sure about
  putting $k_BT$ data in the table.  Maybe it belongs in the caption?}
\label{tab:parameters}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and discussion}
\subsection{Liquid-vapor interface}

We compute the density profile and surface tension by relaxing a slab
of water in a periodic unit cell.  This slab has two surfaces, so the
surface tension is just half the free energy of the cell, divided by
its cross-sectional area.  As mentioned in
section~\ref{sec:empirical}, the $b$ parameter is determined by
setting this surface tension to match the experimental value.  The
associated density profile is plotted in
Figure~\ref{fig:cartesianDensity}, which displays two peaks in
density.

The fluid density at a liquid-solid interface commonly displays
oscillatory behavior, but the presence of density oscillations at
liquid-vapor interfaces is more debated\cite{penfold2001structure}.
In fact, the \emph{meaning} of the density profile at a liquid-vapor
interface is clouded by the existence of long-wavelength perturbations
known as \emph{capillary waves}.  These waves are not accounted for in
density-functional theory (although the \emph{exact}
density-functionaly theory \emph{would} include them), with the result
that the density profile will be smoothed out by CW effects.

\fixme{It'd be worth another search for water intrinsic profile
  papers... but the review I cite is pretty thorough.}

\subsection{Test-particle approach}

More interesting applications are possible when we move into three
dimensions using spherical coordinates.  The first (but not simplest)
thing to try is looking at the response of liquid water to a single
fixed water molecule.  This is referred to as the \emph{test particle}
method\cite{FIXME}, and is as simple as constraining the density to
include a delta function at the origin---meaning that a single
molecule is fixed in location.  The free energy is then minimized
under this constraint, giving a density profile and an associated free
energy.  The density profile is just the self-correlation function
multiplied by the bulk liquid density.  \fixme{We should be
  distinguishing between the properties of the \emph{exact functional}
  and those of an approximate functional...}  Thus, a test-particle
calculation provides a stringent test of the correctness of any
approximate functional.

In order to perform the test-particle test, we introduce a sharp
gaussian density spike at the origin and relax the density under a
constraint that this spike remain.  Figure~\ref{fig:test-particle}
shows the resulting density profile, compared with the experimental
$g_{\textrm{OO}}$.  Note that this data was \emph{not} used in the
fitting of the empirical parameters of the functional.  The agreement
between experiment and theory is both qualitatively and quantitatively
poor.  The integral of the difference between this density profile
(excluding the delta function) and the bulk liquid density should be
0.06~molecules \fixme{look up and show derivation... or cut this},
but we find instead to be $-0.15$~molecules.  In addition, the associated
free energy should be positive: there is an entropic cost in fixing
one molecule in position, but for our functional the free energy of
this density configuration is in fact negative, with a value of
-3.5~eV, meaning that the homogeneous fluid is unstable.  This is a
serious flaw that must be addressed in future work.

\subsection{Hydration energy of hard-sphere solutes}

A more widely studied system that is simpler in many ways than the
test-particle problem is that of a \emph{hard sphere solute} dissolved
in water.  This is a common model for hydrophobic interactions.  The
effect of this solute is merely to create a cavity within which the
water may not penetrate.  This is the simples possible solute, and
will serve as a test case, which if passed could lead to the ability
to insert different chemicals into aqueous solution, instead of a
vapor cavity, in order to observe biologically and chemically
important aqueous solvation interactions.

The most obvious output of such a calculation is the free energy of
solvation.  In Figure~\ref{fig:surfaceTension} we plot the ratio of the
energy of the cavity system to the surface area of the cavity.  This
is an \emph{effective surface tension}, which approaches the bulk
surface tension as the cavity radius increases.  On the same plot, we
include the result from an SPC/E molecular dynamics
simulation\cite{huang2001shs}, showing very good agreement.  At very
small solute radius the effective surface tension is linear, meaning
that the free energy is proportional to the volume---this limiting
case is an exact result, which is exactly reproduced by any functional
that reduces to the ideal gas functional at small distances.

\begin{figure}
\begin{center}
%\includegraphics[width=\columnwidth]{radial-plots/surface-tension}
\end{center}
\caption{Free energy per surface area as a function of cavity
  radius. SPC/E results from~\cite{huang2001shs}}
\label{fig:surfaceTension}
\end{figure}

This exact result for small radius solutes relates to a general property of
fluids interacting with hard surfaces, which is that the pressure on a hard
surface is determined by the contact density of the fluid at the surface.
Thus, a functional that correctly predicts the free energy as a function of
radius will also give the proper contact density.
Figure~\ref{fig:cavities} shows the density profile for several hard sphere
radii, plotted together with with the results of the same SPC/E molecular
dynamics simulation shown in
Figure~\ref{fig:surfaceTension}\cite{huang2001shs}.  As expected---since
the free energies agree so well---the contact densities show good
agreement.  For small spheres the entire curve shows reasonable agreement,
with the agreement becoming worse for larger radii.

\begin{figure}[b]
\begin{center}
%\includegraphics[width=\columnwidth]{radial-plots/cavity-density}
\end{center}
\caption{Density profiles for multiple sizes of hard sphere solute.
  Results from a simulation of SPC/E water~\cite{huang2001shs} are
  shown as a dotted lines.}
\label{fig:cavities}
\end{figure}

In order to understand the worsening agreement in the density profile
for larger solutes in Figure~\ref{fig:cavities}, it is instructive to
consider the limit as the radius becomes very large.  In this limit,
the surface becomes essentially flat, and the liquid should develop a
\emph{vapor layer}, so that as the effective surface tension
approaches the bulk surface tension, the density profile should
approach the profile of the liquid-vapor interface as plotted in
Figure~\ref{fig:cartesianDensity}.  We demonstrate this effect in
Figure~\ref{fig:LargeR_RadialDensity}, which shows superimposed the
density profiles for several large solutes with that of the
liquid-vapor interface.  It is clear that as the radius is increased,
the profile is approaching that of the liquid-vapor interface.  Thus
we can understand the deviations from the molecular dynamics
simulations in Figure~\ref{fig:cavities} at larger radii as a direct
result of the oscillations observed in the planar liquid-vapor
interface.

\fixme{Not sure if we want to keep this paragraph or work it into the
  discussion somewhere?} For \emph{very} small cavities \fixme{not
  shown... should we add one?}, the contact density is equal to the
bulk liquid density, as the solute is too small to really perturb the
water (see the blue curve in Fig.~\ref{fig:cavities}).  As the cavity
size increases, the contact density (and pressure) first rise, and
then fall again, as seen in Figure~\ref{fig:cavities}.  This is due to
short range attractive interactions that are able to pass through the
cavities of small radii.  For larger radii, these interactions are
less prominent and there is less attraction to molecules across the
cavity.  Thus, the contact density is decreased.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
By adding an attractive term to the fundamental-measure theory
functional, we developed a classical density functional for water
which is able reproduce expected results for hydrophobic
interactions.  We hope to extend this functional to allow for
hydrophilic interactions by coupling with electrostatic fields.

This method is computationally efficient, which will be useful when
introducing new functionals that include intramolecular and electromagnetic
interactions, that are not yet included in this model.

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2010-12-09 tests/eos.cpp

@@ -18,7 +18,6 @@
 #include <time.h>
 #include "Functionals.h"
 #include "equation-of-state.h"
-#include "steam-table.h"
 
 int retval = 0;
 double kT = water_prop.kT;
@@ -125,10 +124,22 @@ int main(int, char **argv) {
                nl/water_prop.liquid_density - 1);
         retval++;
       }
+
+      double nv = coexisting_vapor_density(f, water_prop.kT, nl);
+      printf("predicted saturated vapor density: %g\n", nv);
+      printf("actual vapor density:    %g\n", water_prop.vapor_density);
       //double mu = f.derive(-kT*log(nl))*kT/nl; // convert from derivative w.r.t. V
       //o = fopen("saft-fluid-saturated.dat", "w");
       //equation_of_state(o, f + ChemicalPotential(mu)(n), kT, nmin, 1.1*nl);
       //fclose(o);
+
+      double pv = pressure(f, water_prop.kT, nv);
+      printf("vapor pressure is %g\n", pv);
+      if (fabs(pv/water_prop.kT/nv - 1) > 1e-4) {
+        printf("FAIL: error in vapor pressure, steam isn't ideal gas? %g\n",
+               pv/water_prop.kT/nv - 1);
+        retval++;
+      }
     }
 
     {
@@ -156,36 +167,6 @@ int main(int, char **argv) {
       }
       fclose(o);
       printf("Finished plotting room-temperature.dat...\n");
-    }
-
-    {
-      o = fopen("equation-of-state.dat", "w");
-      FILE *experiment = fopen("experimental-equation-of-state.dat", "w");
-      for (int i=0; temperatures_kelvin[i]; i++) {
-        clock_t start = clock();
-        const double kB = 3.16681539628059e-6; // Boltzmann's constant in Hartree/Kelvin
-        double kT = kB*temperatures_kelvin[i];
-        Functional f = SaftFluidSlow(water_prop.lengthscale, kT,
-                                     water_prop.epsilonAB, water_prop.kappaAB,
-                                     water_prop.epsilon_dispersion,
-                                     water_prop.lambda_dispersion, 0);
-        double nl = saturated_liquid(f, kT);
-        double nv = coexisting_vapor_density(f, kT, nl);
-        double p = pressure(f, kT, nl);
-        double pv = pressure(f, kT, nv);
-        printf("EOS:  %g\t%g\t%g\t%g\t%g\t(%g seconds)\n",
[…]

### 2010-12-14 tests/comparisons.sh
Fix bug in a2 from comparison with Jackson.
Tested-on: collins (Linux 2.6.32-5-amd64)

@@ -2,10 +2,10 @@
 
 set -ev
 
-diff -u comparisons/eta_effective*.dat
-
-tests/print-iter.test > tests/print-iter.sh.log.log
-
-diff -u tests/print-iter.sh.correct tests/print-iter.sh.log.log
+for i in eta_effective gHSetaeff a1_sw a2_sw; do
+    echo working on $i
+    echo diff --side-by-side comparisons/${i}_vrpack.dat comparisons/${i}_deft.dat
+    diff --side-by-side comparisons/${i}_deft.dat comparisons/${i}_vrpack.dat
+done
 
 echo this test passes

### 2010-12-14 paper/paper.tex
Fix bug in a2 from comparison with Jackson.
Tested-on: collins (Linux 2.6.32-5-amd64)
Clean up code in SAFT.
Tested-on: collins (Linux 2.6.32-5-amd64)

@@
 \begin{center}
-%\includegraphics[width=\columnwidth]{free-energy/free-energy}
-\end{center}
-\caption{By plotting the free energy vs. density it is possible to
-  determine a liquid or vapor state by the energy minima.  We can adjust the
-  available free parameters to ensure that liquid and vapor coexist with
-  the same minimum free energy.  }
-\label{fig:homogeneous}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[width=\columnwidth]{figs/equation-of-state}
@@
 \subsection{Determining the empirical parameters}\label{sec:empirical}
+
+In order to ensure that the empirical parameters are correct, we test
+the vapor pressure (see Figure~\ref{fig:equation-of-state}) and
+saturated liquid density (see
+Figure~\ref{fig:saturated-liquid-density}) against experimental
+values.  To find these densities, we use the ``common tangent'' rule,
+as illustrated in Figure~\ref{fig:homogeneous}.
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}
+\end{center}
+\caption{By plotting the free energy vs. density it is possible to
+  determine a saturated liquid or vapor state by the common tangent
+  method.  }
+\label{fig:homogeneous}
+\end{figure}
 

### 2010-12-17 src/steam-table.h
Add (ugly) plot of density profile at interface.
Tested-on: collins (Linux 2.6.32-5-amd64)

@@ -73,6 +73,19 @@ static double temperatures_kelvin[] = {
   580,
   590,
   600,
+  610,
+  613,
+  616,
+  619,
+  622,
+  625,
+  628,
+  631,
+  634,
+  637,
+  640,
+  643,
+  646,
   0 // the zero temperature marks the end of the data!
 };
 
@@ -133,7 +146,20 @@ static double water_saturation_liquid[] = {
   0.0035592504,
   0.0034557607,
   0.0033425412,
-  0.0032168704
+  0.0032168704,
+  0.0030744097,
+  0.003027367,
+  0.0029778219,
+  0.0029253456,
+  0.0028693641,
+  0.0028090919,
+  0.0027434354,
+  0.0026708613,
+  0.0025892341,
+  0.0024955994,
+  0.0023852479,
+  0.0022446572,
+  0.0019960588
 };
 
 static double water_vapor_density[] = {
@@ -193,7 +219,20 @@ static double water_vapor_density[] = {
   0.0002170712,
   0.0002562904,
   0.0003033609,
-  0.0003608257
+  0.0003608257,
+  0.000432785,
+  0.0004581583,
+  0.0004857395,
+  0.0005158935,
+  0.0005490903,
+  0.0005859529,
+  0.0006273367,
+  0.0006744725,
+  0.0007292496,
+  0.0007948539,
+  0.000877493,
+  0.0009925767,
+  0.0012059924
 };
 
 static double water_saturated_surface_tension[] = {
@@ -253,7 +292,20 @@ static double water_saturated_surface_tension[] = {
   9.68754999148971E-006,
   8.22132028242856E-006,
   6.78218536110614E-006,
-  5.37969657427005E-006
+  5.37969657427005E-006,
+  4.02658737143302E-006,
+  3.63279609783118E-006,
+  3.2456501071719E-006,
+  2.86587816374686E-006,
+  2.49433666816908E-006,
+  2.13205063207587E-006,
+  1.78027524981818E-006,
+  1.44059335158634E-006,
+  1.11508044251213E-006,
+  8.06611215717945E-007,
+  5.19510947846267E-007,
+  2.61282872845247E-007,
+  0.00000005
 };
 
 static double water_vapor_pressure[] = {
@@ -313,5 +365,18 @@ static double water_vapor_pressure[] = {
   2.79162E-007,
   3.21130E-007,
[…]

### 2011-01-18 paper/paper.tex
Add (ugly) plot of density profile at interface.
Tested-on: collins (Linux 2.6.32-5-amd64)
work with revtex 4.1
add some text about the functionals and what not

@@
 
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 
+The association free energy has the form
+\begin{align}
+  F_\text{assoc} &=
+\end{align}
+\textcolor{red}{Look up paper where I got the association functional!}
+
 \subsection{Dispersion free energy}
@@
 length scale $\lambda_\textit{disp}$.
+
+We use a free square-well dispersion function, which has been
+demonstrated (CITE) to give a reasonable fit to the equation of
+state.  Thus our model interaction has the form:
+\begin{equation}
+  V(r) = \Theta(r-2 \lambda_\textit{disp} R)
+\end{equation}
+where $\Theta$ is the Heaviside step function.
+
+The association free energy has the form
+\begin{align}
+  F_\text{disp} &=
+\end{align}
+\textcolor{red}{Cite the paper where I got the dispersion function!}
 
@@
 \end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
+\end{center}
+\caption{The theoretical versus experimental saturated liquid density
+  versus temperature.  }
+\label{fig:saturated-liquid-density}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature.  }
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-298}
+\end{center}
+\caption{Liquid-vapor interface profile at room temperature.  }
+\label{fig:liquid-vapor-profile}
+\end{figure}
+
+\subsection{Other thermodynamic functions}
+
+From the Helmholtz free energy functional, we may obtain any other
+thermodynamic functions we should choose.  The grand free energy
+$\Phi$, which we ordinarily use in our computations, is obtained by
+simply adding a chemical potential term:
+\begin{equation}
+  \Phi[n] = F[n] + \mu \int n(\xx) d\xx
+\end{equation}
+
+The pressure would naturally be obtained by taking a partial
+derivative with respect to volume at fixed temperature.  Since
+\emph{volume} isn't a parameter in density-functional theory, we
+consider the \emph{free energy per unit volume}.
+\begin{align}
+  F[n] &= f[n]V \\
+  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
+  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
+  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
+   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
+  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n]
+\end{align}
+
+The entropy is naturally obtained by taking a partial derivative with
+respect to temperature:
+\begin{align}
+  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V}
+\end{align}
+This derivative is trivial for the ideal gas and hard sphere free
+energies, which are simply proportional to the temperature.  For the
+association and dispersion free energies, it is a little more work,
+but still not hard.
+
+Once we have the entropy, finding the internal energy is a simple
+matter of subtraction:
+\begin{align}
+  U[n] &= F[n] + TS[n]
+\end{align}
+and the enthalpy is not much more work:
+\begin{align}
+  H[n] &= F[n] + TS[n] - pV
+\end{align}
 

### 2011-01-19 paper/paper.tex
revamp dispersion term yet again...
Tested-on: bingley (Linux 2.6.32-5-amd64)

@@
 
-\needsworknow{Think about the HS functional in
-  Reference~\cite{tarazona2000}, which claims to have fixed nasty
-  singularities!}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\end{center}
+\caption{The theoretical versus experimental phase diagram, vapor
+  pressure versus temperature.  }
+\label{fig:equation-of-state}
+\end{figure}
 
@@
 \end{align}
-\textcolor{red}{Look up paper where I got the association functional!}
 
-\subsection{Dispersion free energy}
-The final term in the free energy is the dispersion term.  We use a
-dispersion term based on the SAFT-VR approach, which has two free
-parameters, an interaction energy $\epsilon_\textit{disp}$ and a
-length scale $\lambda_\textit{disp}$.
+The following contact density for the hard-sphere fluid is in the code
+called \texttt{gHS}, and comes
+from\cite{yu2002fmt-dft-inhomogeneous-associating,
+  fu2005vapor-liquid-dft}, but I don't use it.
+\begin{align}
+  \Delta &= 4\pi \kappa_\textit{assoc} g^\textit{HS}(\sigma)e^{-\beta
+    \epsilon_\textit{assoc}} \\
+  g^\textit{HS}(\sigma) &= \frac1{1-n_3}\left(1+\frac14\frac{\zeta n_2}{1-n_3}
+  \left(1 + \frac{n_2}{14 (1-n_3)}\right)\right)
+\end{align}
+\textcolor{red}{Here is paper where I got the association functional:}
 
-We use a free square-well dispersion function, which has been
-demonstrated (CITE) to give a reasonable fit to the equation of
-state.  Thus our model interaction has the form:
-\begin{equation}
-  V(r) = \Theta(r-2 \lambda_\textit{disp} R)
-\end{equation}
-where $\Theta$ is the Heaviside step function.
-
-The association free energy has the form
+Instead, I use
 \begin{align}
-  F_\text{disp} &=
+  g^\textit{SW}(\sigma) &= g^\textit{HS}_\textit{Carnahan}(\sigma) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial n_3} -
+  \frac{lambda}{3 n3}\frac{\partial a_1}{\partial \lambda}\right)
 \end{align}
-\textcolor{red}{Cite the paper where I got the dispersion function!}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
-\end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
-\end{figure}
+% This is equation 77 of gil-villegas-1997-SAFT-VR
+which I got from reference\cite{gil-villegas-1997-SAFT-VR}.
 
@@
 
+\subsection{Dispersion free energy}
+The final term in the free energy is the dispersion term.  We use a
+dispersion term based on the SAFT-VR
+approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
+parameters, an interaction energy $\epsilon_\textit{disp}$ and a
+length scale $\lambda_\textit{disp}$.
+
+The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+\begin{align}
+  F_\text{disp} &= A_1 + \beta A_2
+\end{align}
+where $A_1$ and $A_2$ are the first two terms in a high-temperature
+perturbation expansion.  $A_1$ is the mean-field dispersion
+interaction, which for a homogeneous fluid is given by
+\begin{align}
+  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
+  g_{HS}(\left|\xx\right|) d\xx
+\end{align}
+The second dispersion term in the free energy $A_2$ describes the
+effect of fluctuations resulting from the compression of the fluid due
+to the dispersion interaction itself, and is commonly approximated
+using the local compressibility approximation (LCA).
+
+We use a free square-well function for the dispersion $\varphi$, which
+has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
+to the equation of state.  Thus our model interaction has the form:
+\begin{equation}
+  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
+\end{equation}
+where $\Theta$ is the Heaviside step function.
+
+\textcolor{red}{TODO:  Add the actual functions that we use and the
+  actual form.  The above is really just an introduction.}
+
 \begin{figure}
@@
 \label{fig:saturated-liquid-density}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature.  }
-\label{fig:surface-tension}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{Liquid-vapor interface profile at room temperature.  }
-\label{fig:liquid-vapor-profile}
 \end{figure}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature.  }
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-298}
+\end{center}
+\caption{Liquid-vapor interface profile at room temperature.  }
+\label{fig:liquid-vapor-profile}
+\end{figure}
+
 \subsection{Determining the empirical parameters}\label{sec:empirical}
@@
 as illustrated in Figure~\ref{fig:homogeneous}.
+
 \begin{figure}

### 2011-01-19 paper/paper.tex
untested commit of fix for gHS -> gSW

@@
 \begin{align}
-  F_\text{assoc} &=
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
 \end{align}
+where the value of $4$ comes from the four association sites per
+molecule, and the functional $X$ is the fraction of assotiation sites
+\emph{not} hydrogen-bonded.  The fraction $X$ is determined by the
+quadratic equation
+\begin{align}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)\Delta(\xx)} - 1}
+  {4 n_0(\xx)\zeta(\xx)\Delta(\xx)}
+\end{align}
+where $\zeta$ comes from
+reference\cite{yu2002fmt-dft-inhomogeneous-associating,
+  fu2005vapor-liquid-dft} and describes the inhomogeneity of the fluid
+density:
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n_{2V}}\cdot\mathbf{n_{2V}}}{n_2^2}
+\end{align}
+The function $\Delta$ is given... below?
 
@@
 \begin{align}
-  g^\textit{SW}(\sigma) &= g^\textit{HS}_\textit{Carnahan}(\sigma) +
-  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial n_3} -
-  \frac{lambda}{3 n3}\frac{\partial a_1}{\partial \lambda}\right)
+  \Delta(\xx) &= \kappa_\textit{assoc}
+  e^{-\beta\epsilon_\textit{assoc}} g^\textit{SW}_\sigma(\xx) \\
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1(\xx)}{\partial n_3(\xx)} -
+  \frac{lambda}{3 n3(\xx)}\frac{\partial a_1(\xx)}{\partial \lambda}\right)
 \end{align}

### 2011-03-04 paper/paper.tex
avoid redundant include
Tested-on: bingley (Linux 2.6.32-5-amd64)
make temperature a field rather than a scalar variable
Tested-on: bingley (Linux 2.6.32-5-amd64)

@@
 A large fraction of interesting chemistry---including all of molecular
-biology---takes place in aqueous solution.  However, while quantum chemistry
-now enables us to calculate the ground state energies of large molecules in
-vacuum, prediction of the free energy of even the smallest molecules in the
-presence of a solvent poses a continuing challenge due to the complex
-structure of a liquid and the computational cost of \emph{ab initio}
-molecular dynamics~\cite{car1985, grossman2004}.  The current state-of-the
-art in \emph{ab initio} molecular dynamics is limited to around 64 water
-molecules per unit cell, which places the minimum solute concentration at
-0.84 molar and allows contact between the second solvation shell of
-neighboring solutes~\cite{izvekov2005, choe2007}.  A more efficient
-approach is needed in order to study nanoscale and hydrophobic solutes.
+biology---takes place in aqueous solution.  However, while quantum
+chemistry now enables us to calculate the ground state energies of
+large molecules in vacuum, prediction of the free energy of even the
+smallest molecules in the presence of a solvent poses a continuing
+challenge due to the complex structure of a liquid and the
+computational cost of \emph{ab initio} molecular
+dynamics~\cite{car1985, grossman2004}.  The current state-of-the art
+in \emph{ab initio} molecular dynamics is limited to around 64 water
+molecules per unit cell, which places the minimum solute concentration
+at 0.84 molar and allows contact between the second solvation shell of
+neighboring solutes~\cite{izvekov2005, choe2007}.  On top of this,
+\emph{ab initio} approaches using classical molecular dynamics have
+\textcolor{red}{vague problems}\cite{weber2010ab-initio-water}.  A
+more efficient approach is needed in order to study nanoscale and
+hydrophobic solutes.
 
@@
 
+\subsection{Crossover formalism for handling critical behavior}
+
+I am currently working on implementing the crossover approach of
+Kiselev and Ely which is designed to make the critical behavior come
+out right\cite{kiselev2006new}.  This method works by defining a
+\emph{classical} free energy $f_0$---which where classical really is
+closer to \emph{mean field} in meaning, and indicates a theory that
+doesn't include long-range fluctuations.  This free energy is then
+used to construct a corrected free energy by feeding portions of it a
+``fake'' density and temperature that are constructed so as to
+reconstruct the correct universal scaling laws near the critical
+point.
+
 \subsection{Determining the empirical parameters}\label{sec:empirical}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}
+\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
+\includegraphics[width=\columnwidth]{figs/near-critical-point}
 \end{center}

### 2011-03-04 src/Functional.h

@@ -13,19 +13,21 @@ public:
   virtual ~FunctionalInterface() {}
 
   // A functional mapping one field onto another...
-  virtual double integral(double kT, const GridDescription &gd, const VectorXd &data) const;
-  virtual VectorXd transform(double kT, const GridDescription &gd, const VectorXd &data) const = 0;
+  virtual double integral(const GridDescription &gd, const VectorXd &kT, const VectorXd &data) const;
+  virtual VectorXd transform(const GridDescription &gd, const VectorXd &kT, const VectorXd &data) const = 0;
   virtual double transform(double kT, double x) const = 0;
 
-  virtual void pgrad(double kT, const GridDescription &gd, const VectorXd &x, const VectorXd &ingrad,
-                     VectorXd *outpgrad) const;
+  virtual void pgrad(const GridDescription &gd, const VectorXd &kT, const VectorXd &x,
+                     const VectorXd &ingrad, VectorXd *outpgrad) const;
 
   // This computes the gradient of the functional, given a gradient of
   // its output field (i.e. it applies the chain rule).
-  virtual void grad(double kT, const GridDescription &gd, const VectorXd &data, const VectorXd &ingrad,
-                    VectorXd *outgrad, VectorXd *outpgrad) const = 0;
+  virtual void grad(const GridDescription &gd, const VectorXd &kT, const VectorXd &data,
+                    const VectorXd &ingrad, VectorXd *outgrad, VectorXd *outpgrad) const = 0;
   virtual double derive(double kT, double data) const = 0;
+  virtual double d_by_dT(double kT, double data) const = 0;
   virtual Functional grad(const Functional &ingrad, const Functional &x, bool ispgrad) const = 0;
+  virtual Functional grad_T(const Functional &ingradT) const = 0;
   virtual Expression printme(const Expression &) const = 0;
   virtual Expression cwiseprintme(const Expression &) const;
 
@@ -89,29 +91,36 @@ public:
   Functional operator-(const Functional &) const;
   Functional operator/(const Functional &) const;
   Functional operator*(const Functional &) const;
-  VectorXd justMe(double kT, const GridDescription &gd, const VectorXd &data) const {
[…]

### 2011-03-04 src/Pow.cpp

@@ -21,7 +21,7 @@ class PowType : public FunctionalInterface {
 public:
   PowType(int nn) : n(nn) {}
 
-  VectorXd transform(double, const GridDescription &gd, const VectorXd &data) const {
+  VectorXd transform(const GridDescription &gd, const VectorXd &, const VectorXd &data) const {
     switch (n) {
     case 0: return VectorXd::Ones(data.rows());
     case 1: return data;
@@ -47,6 +47,9 @@ public:
     for (int p=1; p < n; p++) v *= x;
     return v;
   }
+  double d_by_dT(double, double) const {
+    return 0;
+  }
   Functional grad(const Functional &ingrad, const Functional &x, bool) const {
     switch (n) {
     case 0: return 0;
@@ -54,7 +57,10 @@ public:
     }
     return Pow(n-1)(x)*n*ingrad;
   }
-  void grad(double, const GridDescription &gd, const VectorXd &data, const VectorXd &ingrad,
+  Functional grad_T(const Functional &) const {
+    return 0;
+  }
+  void grad(const GridDescription &gd, const VectorXd &, const VectorXd &data, const VectorXd &ingrad,
             VectorXd *outgrad, VectorXd *outpgrad) const {
     switch (n) {
     case 0: return; // zero gradient!
@@ -94,7 +100,7 @@ class PowAndHalfType : public FunctionalInterface {
 public:
   PowAndHalfType(int nn) : n(nn) {}
 
-  VectorXd transform(double, const GridDescription &, const VectorXd &data) const {
+  VectorXd transform(const GridDescription &, const VectorXd &, const VectorXd &data) const {
     VectorXd out(data.cwise().sqrt());
     if (n < 0) {
       for (int p=0; p > n; p--) out = out.cwise() / data;
@@ -121,10 +127,16 @@ public:
     }
     return out;
   }
+  double d_by_dT(double, double) const {
+    return 0;
+  }
   Functional grad(const Functional &ingrad, const Functional &x, bool) const {
     return PowAndHalf(n-1)(x)*(n+0.5)*ingrad;
   }
-  void grad(double, const GridDescription &gd, const VectorXd &data, const VectorXd &ingrad,
+  Functional grad_T(const Functional &) const {
+    return 0;
+  }
+  void grad(const GridDescription &gd, const VectorXd &, const VectorXd &data, const VectorXd &ingrad,
             VectorXd *outgrad, VectorXd *outpgrad) const {
     if (n > 0) {
       for (int i=0; i<gd.NxNyNz; i++) {

### 2011-06-06 paper/mini-presentation.tex
Changes NativeSlice from red to blue, adds more data points on temperature vs density near critical temp, and more work with constrained water 1D and hydrophobic rods.

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{color}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{bm}
\usetheme{Frankfurt}

\title{A Classical Density-Functional Theory for Describing Confined Water in One, Two and Three(?) Dimensions}
\author{Jessica Gallagher}
\date{May 25, 2011}

\setbeamertemplate{navigation symbols}{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{Theory}
\subsection*{}

\begin{frame}[fragile]{The Functional Terms}
The Helmholtz free energy functional is
based on SAFT, and is composed of:
\begin{equation}
  F[n] = F_\text{id}[n] + F_\text{hs}[n] + F_\text{assoc}[n] + F_\text{disp}
\end{equation}
where $F_\text{id}$ is the ideal gas free energy, $F_\text{hs}$ is
a hard-sphere free energy, $F_\text{assoc}$ is the free energy of
association, and $F_\text{disp}$ is the dispersion energy. 
\end{frame}

\begin{frame}[fragile]{Ideal Gas Free Energy}
The ideal gas functional is given by
\begin{equation}
  F_\text{id}[n] = k_B T \int n(\textbf{x})\left( \ln{n(\textbf{x})} - 1\right) d\textbf{x}
\end{equation}
where $n(\textbf{x})$ is the density of water molecules.  
\end{frame}

\begin{frame}[fragile]{Hard-Sphere Free Energy}
Uses a Fundamental\nobreakdash-Measure Theory (FMT) functional,
which combines scaled-particle theory and the Percus-Yevick theory,
and an attractive interaction.

The FMT functional treats
each water molecule as a hard sphere with a radius approximately that of
a water molecule $(R~=~2.3~bohr)$.
\end{frame}

\begin{frame}[fragile]{Free Energy of Association}
The free energy of association describes the effects of hydrogen bonds. These are modeled as
four association sites on the surface of the
hard sphere. There is an attractive energy
$\epsilon_\text{assoc}$ when two molecules are arranged such that
the hydrogen of one interacts with the lone pair of the other.  The
volume of this interaction is $\kappa_\text{assoc}$ and the free energy is given by
\begin{align}
  F_\text{assoc}[n] &= 4 k_BT \int n_0(\textbf{x})
  \left(\ln X(\textbf{x}) - \frac{X(\textbf{x})}{2} + \frac12\right) d\textbf{x}
\end{align}
where the value of $4$ comes from the four association sites per
molecule, and the functional $X$ is the fraction of association sites
\emph{not} hydrogen-bonded. 
\end{frame}

\begin{frame}[fragile]{Dispersion Energy}
The dispersion free energy is based on the SAFT-VR
approach, which has two free
parameters, an interaction energy $\epsilon_\text{disp}$ and a
length scale $\lambda_\text{disp}$. It has the form
\begin{align}
  F_\text{disp} &= A_1 + \beta A_2
\end{align}
where $A_1$ and $A_2$ are the first two terms in a high-temperature
perturbation expansion.  $A_1$ is the mean-field dispersion
interaction and $A_2$ describes the
effect of fluctuations resulting from the compression of the fluid due
to the dispersion interaction itself, and is commonly approximated
using the local compressibility approximation (LCA).
\end{frame}

\begin{frame}[fragile]{Other Thermodynamic Functions}
\begin{itemize}
 \item Grand free energy
 \item Internal energy
 \item Entropy
 \item Pressure
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Pressure Calculation}
The pressure is obtained by taking a partial
derivative with respect to volume at fixed temperature.  Since
\emph{volume} isn't a parameter in density-functional theory, we
consider the \emph{free energy per unit volume}.
\begin{align}
  F[n] &= f[n]V \\
  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n]
\end{align}
\end{frame}

\section{Results}
\subsection*{}

\begin{frame}[fragile]{Pressure - Water with no Constraints}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
Initial constraint: density equal to saturated liquid density inside cavity, 1000 times vapor density outside
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-1D-030}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-1D-060}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-1D-120}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-1D-030}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-1D-060}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-1D-120}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-1D-030}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-1D-060}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-1D-120}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{1D Slit - Cavity size versus Energy}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
Initial constraint: density equal to saturated liquid density outside rods, 1000 times vapor density inside
\begin{figure}
\begin{center}
%\includegraphics[width=\columnwidth]{figs/rods-in-water-constraint}
\end{center}
\end{figure} 
Diameter of rods = 1 nm (Distance apart = 2nm)
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
\end{center}
\end{figure} 
\end{frame}

\end{document}

### 2011-06-06 paper/project.tex
Changes NativeSlice from red to blue, adds more data points on temperature vs density near critical temp, and more work with constrained water 1D and hydrophobic rods.

--- 2011-01-19 paper/paper.tex
@@
-% TODO:
-
-% Write ``Liquid-vapor interface'' section A.
-
-% Look up question-mark references (citations).
-
 \documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
@@
 \begin{document}
-\title{A Classical Density-Functional Theory for Describing Water Interfaces}
+\title{A Classical Density-Functional Theory for Describing Confined Water in One, Two and Three(?) Dimensions}
 
-\author{Dennis L. Jackson}
+\author{Jessica Gallagher}
 \affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
@@
 
-A large fraction of interesting chemistry---including all of molecular
-biology---takes place in aqueous solution.  However, while quantum chemistry
-now enables us to calculate the ground state energies of large molecules in
-vacuum, prediction of the free energy of even the smallest molecules in the
-presence of a solvent poses a continuing challenge due to the complex
-structure of a liquid and the computational cost of \emph{ab initio}
-molecular dynamics~\cite{car1985, grossman2004}.  The current state-of-the
-art in \emph{ab initio} molecular dynamics is limited to around 64 water
-molecules per unit cell, which places the minimum solute concentration at
-0.84 molar and allows contact between the second solvation shell of
-neighboring solutes~\cite{izvekov2005, choe2007}.  A more efficient
-approach is needed in order to study nanoscale and hydrophobic solutes.
-
-Water stands out among liquids for its extraordinary properties.  Water is
-the only natural substance that is found in solid, liquid and gaseous
-states at temperatures and pressures normally found on Earth.  Water has a
-density maximum at 4$^\circ$C, and expands upon freezing.  The heat
-capacity of water is the highest of any liquid, and its latent heat of
-fusion and vaporization are also the highest.  Water's thermal conductivity
-and surface tensions are the highest of any non-metallic liquid.  Water has
-an unusually high dielectric constant.
-Much of the unusual behavior of water is explained by the ability of each
-water molecule to form \emph{four} strong, directional hydrogen bonds---two
-acceptors and two donors.  This directional bonding dictates the open
-structure of ice, leading to thermal expansion upon freezing.
-
-Due to its large dielectric constant, small molecular size and strong
-hydrogen-bonding network, water as ``the universal solvent'' are as
-interesting as its anomalous thermodynamic properties.
-
-\begin{figure}
-\includegraphics[width=6cm, angle=270]{figs/correlation}
-\caption{ Experimental atom-atom radial distribution functions for water at
-room temperature and pressure from reference~\cite{Soper2000}.  }
-\label{correlation}
-\end{figure}
-
 \subsection{Classical density-functional theory}
-
-Numerous approaches have been developed to approximate the effect of water
-as a solvent in atomistic calculations.  Each of these approaches gives
-adequate description of some aspect of interactions with water, but none of
-them is adequate for describing \emph{all} these interactions with an
-accuracy close to that attained by \emph{ab initio} calculations.  The
-theory of Lum, Chandler and Weeks (LCW)~\cite{LCW}, for instance, can
-accurately describe the free energy cost of creating a cavity by placing a
-solute in water, but does not lend itself to extensions treating the strong
-interaction of water with hydrophilic solutes.  Treatment of water as a
-continuum dielectric with a cavity surrounding each solute can give
-accurate predictions of the energy of solvation of ions~\cite{latimer1939,
-rashin1985, zhan1998, hsu1999, hildebrandt2004, hildebrandt2007}, but
-provides no information about the size of this cavity.  In a physically
-correct approach, the size of the cavity will naturally arise from a
-balance between the free energy required to create the cavity, the
-attraction between the water and the solute, and the steric repulsion which
-opens up the cavity in the first place.
-
-A very promising approach for an efficient continuum description of water
-is that of classical density-functional theory (DFT), which is an approach
-for evaluating the free energy and thermally averaged density of fluids in
-an arbitrary external potential~\cite{ebner1976}.  Classical DFT is a
-natural framework for creating a more flexible theory of hydrophobicity
-that can describe interaction of water with arbitrary external
-potentials---including potentials describing a \emph{strong} interaction
-with a solute.
-
-A number of exact properties are easily achieved in the density-functional
-framework, such as the relationship between the pressure on a hard wall and
-the contact density---which ensures a correct solvation free energy for
-small hard solutes.  Much of the research on classical density-functional
-theory has focussed on the hard-sphere fluid~\cite{curtin1985,
-rosenfeld1989, rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.
-This has led to a large number of very sophisticated functionals, such as
-the fundamental-measure theory (FMT) functional~\cite{rosenfeld1989,
-rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.  This functional
-\needsworklater{has many nice properties}.
-
-Several classical density functionals have already been developed for
-water~\cite{ding1987, Yang1992, Jaqaman2004}, capturing some of the
-qualitative behavior of water.  However, many of these are less
-sophisticated than those describing the hard-sphere fluid, and a
-quantitatively accurate functional describing water---which reproduces both
-the macroscopic surface tension and its decrease as curvature
-increases---has not yet been published.  For instance, Jaqaman \emph{et al}
-developed a classical density-functional for water including orientational
-effects, but their functional resulted in a surface tension twice the
-experimental value, with the result that it can really only be used to make
-very qualitative predictions~\cite{Jaqaman2004}.
-
-\needsworknow{Add discussion and citations for more recent papers}
-
-One of the most interesting functionals for water is that of Ding \emph{et
-al}, which was developed in order to explain the freezing of
-water~\cite{ding1987}.  This functional uses the approach of Chandler
-\emph{et al}, which uses separate atomic densities for the constituent
-atoms of a molecular fluid~\cite{chandler1986a, chandler1986b}.  Although
-this is a compelling approach, the actual functional developed cannot
-describe the vapor state or liquid-vapor interface, both of which are
-crucial to correct predictions of hydrophobic behavior at large length
-scales.  In addition, the functional of Ding \emph{et al} also fails to
-rigorously enforce intramolecular correlations, which casts doubt on its
-suitability for describing interactions with ionic solutes.
-
-\needsworklater{
-Density-Functional Theory is most commonly used for calculating the
-densities of electrons in quantum mechanical systems.  We take the
-classical limit of this theory to construct a functional that will predict
-the density of water molecules in a classical system.  The model used here
-starts with a treatment of each individual water molecule as a purely
-classical hard sphere.  The only constraint in this model is
-the minimum distance between any two molecules is twice the hard sphere
-radius.}
-
-\needsworklater{ I still need to rewrite the entire introduction... }
 
@@
 \section{Theory and Methods}
+
 We introduce a new classical density functional for water that
@@
 where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of assotiation sites
+molecule, and the functional $X$ is the fraction of association sites
 \emph{not} hydrogen-bonded.  The fraction $X$ is determined by the
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/entropy}
+\end{center}
+\caption{The free energy, internal energy, and temperature*entropy versus density at room temperature.  }
+\label{fig:energy-room-temp}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/entropy-at-690K}
+\end{center}
+\caption{The free energy, internal energy, and temperature*entropy versus density near the critical temperature.  }
+\label{fig:energy-near-critical-temp}
+\end{figure}
+
 \subsection{Other thermodynamic functions}
@@
 \end{align}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including experimental pressure data from NIST.  }
+\label{fig:pressure-with-isotherms}
+\end{figure}
 
@@
 
-\fixme{It'd be worth another search for water intrinsic profile
-  papers... but the review I cite is pretty thorough.}
-
-\subsection{Test-particle approach}
-
-More interesting applications are possible when we move into three
-dimensions using spherical coordinates.  The first (but not simplest)
-thing to try is looking at the response of liquid water to a single
-fixed water molecule.  This is referred to as the \emph{test particle}
-method\cite{FIXME}, and is as simple as constraining the density to
-include a delta function at the origin---meaning that a single
-molecule is fixed in location.  The free energy is then minimized
-under this constraint, giving a density profile and an associated free
-energy.  The density profile is just the self-correlation function
-multiplied by the bulk liquid density.  \fixme{We should be
-  distinguishing between the properties of the \emph{exact functional}
-  and those of an approximate functional...}  Thus, a test-particle
-calculation provides a stringent test of the correctness of any
-approximate functional.
-
-In order to perform the test-particle test, we introduce a sharp
-gaussian density spike at the origin and relax the density under a
-constraint that this spike remain.  Figure~\ref{fig:test-particle}
-shows the resulting density profile, compared with the experimental
-$g_{\textrm{OO}}$.  Note that this data was \emph{not} used in the
-fitting of the empirical parameters of the functional.  The agreement
-between experiment and theory is both qualitatively and quantitatively
-poor.  The integral of the difference between this density profile
-(excluding the delta function) and the bulk liquid density should be
-0.06~molecules \fixme{look up and show derivation... or cut this},
-but we find instead to be $-0.15$~molecules.  In addition, the associated
-free energy should be positive: there is an entropic cost in fixing
-one molecule in position, but for our functional the free energy of
-this density configuration is in fact negative, with a value of
--3.5~eV, meaning that the homogeneous fluid is unstable.  This is a
-serious flaw that must be addressed in future work.
-
-\subsection{Hydration energy of hard-sphere solutes}
-
-A more widely studied system that is simpler in many ways than the
-test-particle problem is that of a \emph{hard sphere solute} dissolved
-in water.  This is a common model for hydrophobic interactions.  The
-effect of this solute is merely to create a cavity within which the
-water may not penetrate.  This is the simples possible solute, and
-will serve as a test case, which if passed could lead to the ability
-to insert different chemicals into aqueous solution, instead of a
-vapor cavity, in order to observe biologically and chemically
-important aqueous solvation interactions.
-
-The most obvious output of such a calculation is the free energy of
-solvation.  In Figure~\ref{fig:surfaceTension} we plot the ratio of the
-energy of the cavity system to the surface area of the cavity.  This
-is an \emph{effective surface tension}, which approaches the bulk
-surface tension as the cavity radius increases.  On the same plot, we
-include the result from an SPC/E molecular dynamics
-simulation\cite{huang2001shs}, showing very good agreement.  At very
-small solute radius the effective surface tension is linear, meaning
-that the free energy is proportional to the volume---this limiting
-case is an exact result, which is exactly reproduced by any functional
-that reduces to the ideal gas functional at small distances.
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=\columnwidth]{radial-plots/surface-tension}
-\end{center}
-\caption{Free energy per surface area as a function of cavity
-  radius. SPC/E results from~\cite{huang2001shs}}
-\label{fig:surfaceTension}
-\end{figure}
-
-This exact result for small radius solutes relates to a general property of
-fluids interacting with hard surfaces, which is that the pressure on a hard
-surface is determined by the contact density of the fluid at the surface.
-Thus, a functional that correctly predicts the free energy as a function of
-radius will also give the proper contact density.
-Figure~\ref{fig:cavities} shows the density profile for several hard sphere
-radii, plotted together with with the results of the same SPC/E molecular
-dynamics simulation shown in
-Figure~\ref{fig:surfaceTension}\cite{huang2001shs}.  As expected---since
-the free energies agree so well---the contact densities show good
-agreement.  For small spheres the entire curve shows reasonable agreement,
-with the agreement becoming worse for larger radii.
-
-\begin{figure}[b]
-\begin{center}
-%\includegraphics[width=\columnwidth]{radial-plots/cavity-density}
-\end{center}
-\caption{Density profiles for multiple sizes of hard sphere solute.
-  Results from a simulation of SPC/E water~\cite{huang2001shs} are
-  shown as a dotted lines.}
-\label{fig:cavities}
-\end{figure}
-
-In order to understand the worsening agreement in the density profile
-for larger solutes in Figure~\ref{fig:cavities}, it is instructive to
-consider the limit as the radius becomes very large.  In this limit,
-the surface becomes essentially flat, and the liquid should develop a
-\emph{vapor layer}, so that as the effective surface tension
-approaches the bulk surface tension, the density profile should
-approach the profile of the liquid-vapor interface as plotted in
-Figure~\ref{fig:cartesianDensity}.  We demonstrate this effect in
-Figure~\ref{fig:LargeR_RadialDensity}, which shows superimposed the
-density profiles for several large solutes with that of the
-liquid-vapor interface.  It is clear that as the radius is increased,
-the profile is approaching that of the liquid-vapor interface.  Thus
-we can understand the deviations from the molecular dynamics
-simulations in Figure~\ref{fig:cavities} at larger radii as a direct
-result of the oscillations observed in the planar liquid-vapor
-interface.
-
-\fixme{Not sure if we want to keep this paragraph or work it into the
-  discussion somewhere?} For \emph{very} small cavities \fixme{not
-  shown... should we add one?}, the contact density is equal to the
-bulk liquid density, as the solute is too small to really perturb the
-water (see the blue curve in Fig.~\ref{fig:cavities}).  As the cavity
-size increases, the contact density (and pressure) first rise, and
-then fall again, as seen in Figure~\ref{fig:cavities}.  This is due to
-short range attractive interactions that are able to pass through the
-cavities of small radii.  For larger radii, these interactions are
-less prominent and there is less attraction to molecules across the
-cavity.  Thus, the contact density is decreased.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
-By adding an attractive term to the fundamental-measure theory
-functional, we developed a classical density functional for water
-which is able reproduce expected results for hydrophobic
-interactions.  We hope to extend this functional to allow for
-hydrophilic interactions by coupling with electrostatic fields.
-
-This method is computationally efficient, which will be useful when
-introducing new functionals that include intramolecular and electromagnetic
-interactions, that are not yet included in this model.
 

### 2011-06-06 paper/figs/equation-of-state.cpp

@@ -42,7 +42,7 @@ int main(int, char **) {
                                                 water_prop.epsilon_dispersion,
                                                 water_prop.lambda_dispersion,
                                                 water_prop.length_scaling, 0));
-  for (int i=0; i<imax; i+=3) {
+  for (int i=0; i<imax; i+=1) {
     //printf("Working on equation of state at %g Kelvin...\n", temperatures_kelvin[i]);
     double kT = kB*temperatures_kelvin[i];
     saturated_liquid_vapor(f, kT, 1e-14, 0.0017, 0.0055, &nl, &nv, &mu, 1e-6);
@@ -59,10 +59,10 @@ int main(int, char **) {
             water_saturated_surface_tension[i]);
   }
 
-  for (double T=660; T<=693; T += 5) {
+  for (double T=650; T<=695; T += 1) {
     //printf("Working on bonus equation of state at %g Kelvin...\n", T);
     double kT = kB*T;
-    saturated_liquid_vapor(f, kT, 1e-14, 0.0017, 0.0055, &nl, &nv, &mu, 1e-6);
+    saturated_liquid_vapor(f, kT, 0.0005, 0.0019, 0.003, &nl, &nv, &mu, 1e-6);
     took("Finding coesisting liquid and vapor densities");
     double pv = pressure(f, kT, nv);
     took("Finding pressure");

### 2011-06-06 paper/paper.tex
clarify the FMT functional we actually use

@@
   limit.}
-
-The hard sphere functional is adapted from that of
-Rosenfeld~\cite{rosenfeld1997},
-and is given by
 \begin{equation}
-F_{hs}[n] = k_B T \int (\Phi_1 + \Phi_2 + \Phi_3) d\xx \; ,
+F_{hs}[n] = k_B T \int (\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)) d\xx \; ,
 \end{equation}
@@
 \begin{align}
-\Phi_1 &= -n\circ w_2 \ln\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)\\
-\Phi_2 &= \frac{4\pi R^3 \lbrack (n\circ w_2)^2 - (n\circ \mathbf{w}_3)^2\rbrack}
-          {\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)}\\
-\Phi_3 &= \frac{8\pi^2 R^6\lbrack \frac{1}{3}(n\circ w_2)^3 - (n\circ
-          w_2)(n\circ \mathbf{w}_3)^2\rbrack}
-          {\left( 1-\frac{4}{3}\pi R^3 n\circ w_1\right)} \; ,
+\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
+\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
+\Phi_3 &= \frac{
+  \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+}{
+  8\pi \left( 1 - n_3 \right)^2
+} ,
 \end{align}
-using the normalized weighting functions 
+using the weighted densities
 \begin{align}
-w_1(\xx) &= \frac{3}{4\pi R^3}\Theta(R-|\xx|)\\
-w_2(\xx) &= \frac{1}{4\pi R^2}\delta(R-|\xx|)\\
-\mathbf{w}_3(\xx) &= \frac{-1}{4\pi R^2}\vec{\nabla}\Theta(R-|\xx|) \; .
+  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
+  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
+  \\
+  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 \\
+  n_1 &= \frac{n_2}{4\pi R}\\
+  \mathbf{n}_{V1} &= \frac{\mathbf{n}_{V2}}{4\pi R}\\
+  n_0 &= \frac{n_2}{4\pi R^2}
 \end{align}

### 2011-06-06 contactpaper/paper.tex

% TODO:

% Write ``Liquid-vapor interface'' section A.

% Look up question-mark references (citations).

\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{bm}% bold math
\usepackage{color}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\xx}{\textbf{x}}
\newcommand{\refnote}{\red{[ref]}}

\newcommand{\fixme}[1]{\red{[#1]}}

% needsworklater is used to annotate bits that need work, but that we
% can postpone for a while.
\newcommand{\needsworklater}[1]{\emph{[#1]}}
% needsworknow is intended to prioritize stuff that needs fixing.
\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}

\begin{document}
\title{A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
\needsworklater{ We develop a functional based on FMT to ... for SAFT}
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

\cite{roth2002whitebear}

The hard-sphere contact density---which we will define as the density
of hard spheres touching a given hard sphere---is important for the
association term in Statistical Associating Fluid Theory (SAFT).  For
a homogeneous hard-sphere fluid, the contact density---which is to
say, $n g_{HS}(\sigma)$---is easily computed from the
Carnahan-Starling equation of state
\begin{align}
  g_{HS}(\sigma) &= \frac{1}{1-\eta} + \cdots
\end{align}
The contact density may be easily computed from a knowledge of the
contact theorem, which states that the pressure on any hard surface is
given by
\begin{align}
  p &= k_BT n_\textit{contact}
\end{align}
Since we are interested in the contact density at the hard-sphere
surface, all we need compute is the pressure on that surface, and
we'll have our answer.  The pressure on a hard sphere can be readily
computed from the dependence of the free energy on hard sphere
radius.
\begin{align}
  A_{HS} &= Nk_BT \frac{4\eta - 3\eta^2}{(1-\eta)^2}
\end{align}
Since $\eta \equiv \frac{4\pi}{3} R^3$, we can work out that
\begin{align}
  \frac{dA_{HS}}{dR} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{dR} \\
  &= Nk_BT \left( \frac{4 - 6\eta}{(1-\eta)^2} + 2 \frac{4\eta - 3\eta^2}{(1-\eta)^3} \right) \frac{d\eta}{dR}
  \\
  &= Nk_BT \frac{4 - 4\eta - 6\eta + 6\eta^2 + 8\eta - 6\eta^2}{(1-\eta)^3} \frac{d\eta}{dR}
  \\
  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{d\eta}{dR}
  \\
  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R}
\end{align}
This derivative gives us the force on \emph{all} the hard
spheres---since we're changing all their radii at once.  To compute
the pressure on the spheres, we just need to divide by the total area,
which means dividing by $N 4\pi (2R)^2$.  The area which is relevant
is the area over which the molecules can make contact.
\begin{align}
  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
  &= \frac14 \frac{1}{N 4\pi R^2} Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \\
  &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
  &= n k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
  n_\textit{contact} &= n \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
  g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3}
\end{align}
As it turns out, this is the correct answer! $\ddot\smile$

\section{Contact density at a given hard sphere}

When we consider an inhomogeneous system, we may ask for \emph{which}
sphere we are interested in computing the contact density.  One way to
answer this is to specify a sphere at a fixed position $\mathbf{r}$.
By making the hard-sphere radius dependent on position, we can compute
the pressure on the hard spheres located at $\mathbf{r}$.
\begin{align}
  p_{HS}(\mathbf{r}) &= \frac{1}{4\pi (2R)^2} \frac{\delta A_{HS}}{\delta R(\mathbf{r})}
\end{align}
So to compute this we need
\begin{align}
  \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
  \int \left(
  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
  \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})}
  +
  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
  \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})}
  + \cdots
  \right) d\mathbf{r}'
\end{align}
There are \emph{many} such terms, but none of them are particularly
challenging.

\begin{figure}
\includegraphics[width=3cm]{figs/n0}
\caption{Set of hard spheres that included in $n_0(\mathbf{r})$.}
\label{fig:n0}
\end{figure}

\section{Contact density at a given contact position}

Unfortunately, knowing the average contact density on each hard sphere
doesn't really give us what we want for SAFT.  The trouble is that in
order to work out the fraction of association sites occupied, we need
to know the contact density \emph{at the point of contact}.
Fortunately, we don't need to know the contact density for a
particular sphere at each point of contact.  Just as $n_0(\mathbf{r})$
gives us the density of spheres that are contacting point $\mathbf{r}$
(see Figure~\ref{fig:n0}, we can use similar reasoning to work out the
contact density for these spheres.

To find the contact density at a given point of contact, we can't
simply change the radius for spheres touching that point, since that
would give us the pressure averaged over the entire surface of those
spheres.  Instead, we'll have to imagine extending a patch of those
spheres which is located at the point of interest.

\begin{figure}
\includegraphics[width=3cm]{figs/contact}
\caption{Construction for finding the contact density at a given point
  of contact.}
\label{fig:contact}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}
We get a nice spatial dependence for SAFT.

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2011-06-08 contactpaper/paper.tex
various changes in contact density paper
Tested-on: bingley (Linux 2.6.32-5-amd64)

@@
 \begin{abstract}
-\needsworklater{ We develop a functional based on FMT to ... for SAFT}
+\needsworklater{ We develop a functional based on FMT~\cite{roth2002whitebear}
+ to ... for SAFT}
 \end{abstract}
@@
 \section{Introduction}
-
-\cite{roth2002whitebear}
 
@@
   &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
-  &= n k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
+  &= n k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3}
+\end{align}
+\begin{align}
   n_\textit{contact} &= n \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
@@
 \end{align}
-As it turns out, this is the correct answer! $\ddot\smile$
+As it turns out, this is the standard answer for the contact value of
+the autocorrelation function! $\ddot\smile$
+
+\section{Fundamental-Measure Theory}
+
+We use the White Bear version of the Fundamental-Measure Theory~(FMT)
+functional published in reference~\cite{roth2002whitebear}.  The FMT
+functional describes the excess free energy of a hard-sphere fluid.
+This particular FMT reduces to the Carnahan-Starling equation of state
+for homogeneous systems.
+\begin{equation}
+A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)\right) d\xx \; ,
+\end{equation}
+with integrands
+\begin{align}
+\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
+\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
+\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+  n_3 + (1-n_3)^2 \ln(1-n_3)
+}{
+  36\pi n_3^2\left( 1 - n_3 \right)^2
+} ,
+\end{align}
+using the weighted densities
+\begin{align}
+  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
+  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
+\end{align}
+\begin{align}
+  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 , \quad
+  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{4\pi R} \\
+  n_1 &= \frac{n_2}{4\pi R} , \quad
+  n_0 = \frac{n_2}{4\pi R^2}
+\end{align}
+
+\begin{widetext}
+
+\section{Mean Contact density}
+
+One reasonably simple question to ask is what the \emph{mean} contact
+density of an inhomogeneous system is.  This comes down to simply
+reproducing the derivation for a homogeneous system using the
+FMT functional.
+\begin{align}
+  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR}
+\end{align}
+So to compute this we need
+\begin{align}
+  \frac{d A_{HS}}{d R} &=
+  \frac{\partial A_{HS}}{\partial R} \\
+  &+
+  \int \left(
+  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
+  \frac{d n_3(\mathbf{r}')}{d R}
+  +
+  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
+  \frac{d n_2(\mathbf{r}')}{d R}
+  + \cdots
+  \right) d\mathbf{r}'
+\end{align}
+
+\begin{align}
+  \frac{dn_3(\mathbf{r})}{dR} &= -n_2(\mathbf{r})\\
+  \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
+  \delta'(|\mathbf{r'}| - R)\mathbf{dr}'\\
+  &= \int n(\mathbf{r}+\mathbf{r}')
+  \delta'(r' - R) d\Omega' r'^2 dr'\\
+  &= \int 
+  \delta(r' - R) d\Omega' \left(\frac{dn(\mathbf{r}+\mathbf{r}')}{dr'}r'^2 + 2n(\mathbf{r}+\mathbf{r}')r' \right) dr'\\
+  &= \frac1{R} \int 
+  \delta(|\mathbf{r}'| - R)
+  \left(\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') +
+  2n(\mathbf{r}+\mathbf{r}') \right) \mathbf{dr}' \\
+  &= \frac{2}{R}n_2(\mathbf{r}) + \frac{1}{R} \int 
+  \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+  \frac{dn_0(\mathbf{r})}{dR} &= \frac{1}{R} \int 
+  \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+  \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
+\end{align}
+
+The total force due to $\Phi_1$ is actually very easy to work out:
+\begin{align}
+  \frac{d A_{HS}^{(1)}}{d R} &=
+  \int \mathbf{dr}' \left(-\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
+\end{align}
+where you can see that $\frac{dn_0(\mathbf{r})}{dR}$ is zero for a
+homogeneous system, and small for reasonably smooth density
+distributions.  The next term from $\Phi_2$ is as follows:
+\begin{align}
+  \frac{d A_{HS}^{(2)}}{d R} &=
+  \int \mathbf{dr}' \left(
+    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    n_2
+    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2} - 2 n_1
+      \frac{dn_2}{dR} + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot
+    }{1-n_3}
+  \right) \\
+  &= \int \mathbf{dr}' \left(
+    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    n_2
+    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+      - 2 n_1 \left(\frac2{R}n_2 \frac{dn_0}{dR}\right)
+      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+  \right) \\
+  &= \int \mathbf{dr}' \left(
+    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    n_2
+    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+      - 4 n_0n_2 - 2 n_1\frac{dn_0}{dR}
+      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+  \right) \\
+  &= \int \mathbf{dr}' \left(
+    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    n_2
+    + \frac{3n_0n_2 + \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+      + 2 n_1\frac{dn_0}{dR}
+      - 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+  \right)
+\end{align}
+
+And the last term looks like this.  I should perhaps note that this
+term should be replaced by a tensor version if want to examine
+seriously localized density distributions.
+\begin{align}
+  A_{HS}^{(3)} &=
+  \int \mathbf{dr}'\left(
+     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
+     \frac{
+       n_3 + (1-n_3)^2 \ln(1-n_3)
+     }{
+       36\pi n_3^2\left( 1 - n_3 \right)^2
+     }
+  \right) \\
+  \frac{d A_{HS}^{(3)}}{d R} &=
+  \int \mathbf{dr}'
+     \left[
+       3(n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
+       \left(\frac2{R}n_2 + \frac{dn_0}{dR}\right)
+       - 6n_2 \mathbf{n}_{V2} \cdot \nabla n_2
+     \right]
+     \frac{
+       n_3 + (1-n_3)^2 \ln(1-n_3)
+     }{
+       36\pi n_3^2\left( 1 - n_3 \right)^2
+     }
+   \\
+  &+
+     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) (-n_2)
+     \frac{
+       (1 - 2(1-n_3)\ln(1-n_3) - (1-n_3))n_3^2( 1 - n_3)^2
+       -
+       (n_3 + (1-n_3)^2 \ln(1-n_3))(-2 n_3^2( 1 - n_3) + 2n_3(1-n_3)^2)
+     }{
+       36\pi (n_3^2\left( 1 - n_3 \right)^2)^2
+     }
+     \\
+     &=
+     \int \mathbf{dr}'
+     \left[
+       (n_1n_2 - \mathbf{n}_{V1} \cdot \mathbf{n}_{V2})
+       \left(2n_2 + \frac{1}{4\pi}\frac{dn_0}{dR}\right)
+       - \frac{1}{2\pi}n_2 \mathbf{n}_{V2} \cdot \nabla n_2
+     \right]
+     \frac{
+       n_3 + (1-n_3)^2 \ln(1-n_3)
+     }{
+       3 n_3^2\left( 1 - n_3 \right)^2
+     }
+   \\
+  &-
+     \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
+   \left( \frac{1}{(1-n_3)^2} + 2 \frac{\ln(1-n_3)}{n_3^2(1-n_3)} \right)
+\end{align}
+
+\begin{align}
+  \frac{d A_{HS}}{d R} &=
+  \int \mathbf{dr}' \left(
+    -\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3)
+    + \cdots
+  \right)
+\end{align}
 
@@
 \end{align}
-There are \emph{many} such terms, but none of them are particularly
-challenging.
+There are \emph{many} such terms, and this could get expensive.
+
+  First, let's rewrite the free energy in terms of a smaller set of
+  weighted densities:
+  \begin{align}
+    A_{HS} &= \int \left\{ -n_0 \ln\left( 1 - n_3\right)
+    + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+    + \frac{
+      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+    }{
+      8\pi \left( 1 - n_3 \right)^2
+    }\right\}
+    \\
+    &= \int \left\{ -\frac{n_2}{4\pi R^2} \ln\left( 1 - n_3\right)
+    + \frac{1}{4\pi R} \frac{n_2^2 - \mathbf{n}_{V2} \cdot\mathbf{n}_{V2}}{1-n_3}
+    + \frac{
+      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+    }{
+      8\pi \left( 1 - n_3 \right)^2
+    }\right\}
+\end{align}
+
+\begin{align}
+    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
+    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
+      n_3(\mathbf{r}'))^2}
+    + \frac{1}{4\pi}\frac{
+      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+    }{
+      (1-n_3(\mathbf{r}'))^3
+    }
+    \\
+    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
+    -\frac{1}{4\pi R^2} \ln\left( 1 - n_3\right)
+    +  \frac{2n_1}{1-n_3}
+    + \frac{n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{8\pi \left( 1 - n_3 \right)^2}
+    \\
+    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
+    \frac{2\mathbf{n}_{V1}}{1-n_3}
+    + \frac{n_2 \mathbf{n}_{V2}}{4\pi \left( 1 - n_3 \right)^2}
+    \\
+    \frac{\delta A_{HS}}{\delta R(\mathbf{r}')} &=
+    \frac{n_2}{2\pi R^3} \ln\left( 1 - n_3\right)
+    - \frac1{R} \frac{n_1n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+  \end{align}
+
+\begin{align}
+  \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \\
+  \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  n(\mathbf{r}) \delta'(|\mathbf{r}-\mathbf{r}'| - R) \\
+  \frac{\delta \mathbf{n}_{V2}(\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  -\frac{\delta}{\delta R(\mathbf{r})} \mathbf{\nabla}' n_3(\mathbf{r}')
+  \\
+  &= -\mathbf{\nabla}'
+  \left(n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \right)
+\end{align}
+These give us some basic tools with which to perform derivatives.
+
+Finally, we get an answer like
+\begin{align}
+  \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
+  \int \mathbf{dr}' \left\{
+  \left(
+  \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+  + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
+    n_3(\mathbf{r}'))^2}
+  + \frac{1}{4\pi}\frac{
+    \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+  }{
+    (1-n_3(\mathbf{r}'))^3
+  }
+  \right) n(\mathbf{r}') \delta(|\mathbf{r}-\mathbf{r}'| - R)
+  + \cdots
+  \right\}
+\end{align}
+This comes out to quite a mess.  It wouldn't be infeasible to compute
+this, but it wouldn't be efficient or pretty either.
 
 \begin{figure}
-\includegraphics[width=3cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$.}
+\includegraphics[width=5cm]{figs/n0}
+\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+  consist of those which just touch the point $\mathbf{r}$.}
 \label{fig:n0}
@@
 gives us the density of spheres that are contacting point $\mathbf{r}$
-(see Figure~\ref{fig:n0}, we can use similar reasoning to work out the
+(see Figure~\ref{fig:n0}), we can use similar reasoning to work out the
 contact density for these spheres.
@@
 spheres.  Instead, we'll have to imagine extending a patch of those
-spheres which is located at the point of interest.
+spheres which is located at the point of interest, as illustrated in
+Figure~\ref{fig:contact}.
 
 \begin{figure}
-\includegraphics[width=3cm]{figs/contact}
+\includegraphics[width=5cm]{figs/contact}
 \caption{Construction for finding the contact density at a given point
-  of contact.}
+  of contact.  We envision a small area $dA$ of each sphere that touches
+the point $\mathbf{r}$ extended by a distance $dR$.}
 \label{fig:contact}
 \end{figure}
+
+So we need to find the small change in free energy that is caused by
+these small changes in the shape and size of the hard spheres that are
+in contact with this point.  This changes the weighted densities in
+the following way---assuming that our point of interest is
+$\mathbf{r}$.
+\begin{align}
+  \Delta n_3(\mathbf{r}') &= n(\mathbf{r})\Delta A \Delta R
+  \delta(|\mathbf{r}-\mathbf{r}'| - R)
+   \\
+  \Delta n_2(\mathbf{r}') &=
+  \left(n\left(\mathbf{r} + \Delta R
+  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
+  R}{R}\right)- n(\mathbf{r})\right) \Delta A
+  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
+  &=
+  \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
+    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
+    \right) \Delta A \Delta R
+  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
+  \Delta \mathbf{n}_{V2}(\mathbf{r}') &=
+  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
+  \left(n\left(\mathbf{r} + \Delta R
+  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
+  R}{R}\right)- n(\mathbf{r})\right) \Delta A
+  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
+  &=
+  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
+  \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
+    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
+    \right) \Delta A \Delta R
+  \delta(|\mathbf{r}-\mathbf{r}'|-R) 
+\end{align}
+\end{widetext}
 

### 2011-06-09 contactpaper/paper.tex
a few more changes to contact-density paper
Tested-on: bingley (Linux 2.6.32-5-amd64)

@@
   \begin{align}
-    A_{HS} &= \int \left\{ -n_0 \ln\left( 1 - n_3\right)
+    A_{HS} &= \int \left\{
+      -n_0 \ln\left( 1 - n_3\right)
       + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + \frac{
-      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+      + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+        n_3 + (1-n_3)^2 \ln(1-n_3)
       }{
-      8\pi \left( 1 - n_3 \right)^2
-    }\right\}
+        36\pi n_3^2(1-n_3)^2
+      }
+    \right\}
     \\
-    &= \int \left\{ -\frac{n_2}{4\pi R^2} \ln\left( 1 - n_3\right)
+    &= \int \left\{
+      -\frac{n_2}{4\pi R^2} \ln\left( 1 - n_3\right)
       + \frac{1}{4\pi R} \frac{n_2^2 - \mathbf{n}_{V2} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + \frac{
-      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+      + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+        n_3 + (1-n_3)^2 \ln(1-n_3)
       }{
-      8\pi \left( 1 - n_3 \right)^2
-    }\right\}
+        36\pi n_3^2(1-n_3)^2
+      }
+    \right\}
 \end{align}
@@
     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
     + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
       n_3(\mathbf{r}'))^2}
-    + \frac{1}{4\pi}\frac{
-      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
-    }{
-      (1-n_3(\mathbf{r}'))^3
-    }
+    + \frac{n_2^3 -
+      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
+      \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
+    \right)
     \\
@@
     +  \frac{2n_1}{1-n_3}
-    + \frac{n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{8\pi \left( 1 - n_3 \right)^2}
+    + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
+      (1-n_3)^2\ln(1-n_3)}{
+      36\pi n_3^2(1-n_3)^2
+    }
     \\
@@
     \frac{2\mathbf{n}_{V1}}{1-n_3}
-    + \frac{n_2 \mathbf{n}_{V2}}{4\pi \left( 1 - n_3 \right)^2}
-    \\
+    - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
+      (1-n_3)^2\ln(1-n_3)}{
+      36\pi n_3^2(1-n_3)^2
+    }    \\
     \frac{\delta A_{HS}}{\delta R(\mathbf{r}')} &=
@@
 these small changes in the shape and size of the hard spheres that are
-in contact with this point.  This changes the weighted densities in
-the following way---assuming that our point of interest is
-$\mathbf{r}$.
+in contact with this point.  In the following section, we will work
+out the changes in the weighted densities at position $\mathbf{r}'$
+when point of contact we are considering is $\mathbf{r}$.
+
+The packing fraction $n_3$ is changed at the point where the sphere is
+extended, since a few more spheres will overlap this point.
 \begin{align}
@@
   \delta(|\mathbf{r}-\mathbf{r}'| - R)
-   \\
+\end{align}
+The density $n_0$ is the average density, averaged over a spherical
+shell.  As such, shifting the position of a shell just shifts which
+location we are averaging over.
+\begin{align}
+  \Delta n_0(\mathbf{r}') &=
+    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
+      \mathbf{\nabla}n(\mathbf{r}) \Delta A \Delta R
+    \delta(|\mathbf{r}-\mathbf{r}'|-R)
+\end{align}
+The change in $n_2$ comes from two terms: a change in the density value, as
+we saw in $n_0$, and a change in the area of the segment.
+\begin{align}
   \Delta n_2(\mathbf{r}') &=
@@
     \right) \Delta A \Delta R
-  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
+  \delta(|\mathbf{r}-\mathbf{r}'|-R)
+\end{align}
+The vector density $\mathbf{n}_{V2}$ is similar to the density
+$n_2$: both changes in the density value and changes in the surface
+area play a role.
+\begin{align}
   \Delta \mathbf{n}_{V2}(\mathbf{r}') &=
@@
 \end{align}
+The density $n_1$ (and similarly $\mathbf{n}_{V1}$) can be roughly
+interpreted as the density weighted by the area-averaged mean
+curvature~\cite{oversteegen2005general, roth2010review}.  When we
+extend the sphere at the contact point, we will introduce regions of
+both positive and negative mean curvature.  In principle, we can do
+this in such a way that the two portions cancel, and hence we will
+assume that
+\begin{align}
+  \Delta n_1(\mathbf{r}') &= 0 \\
+  \Delta \mathbf{n}_{V1}(\mathbf{r}') &= 0
+\end{align}
+In fact, the negative and positive portions are spatially separated,
+so these terms should probably include a derivative term, which I
+should work out.  Particularly as it may even happen to cancel a
+different derivative term that we then won't have to drop!
+
+We need to then find the change in Helmholtz free energy due to this
+change at position $\mathbf{r}$:
+\begin{align}
+  \Delta A_{HS} &= \int \mathbf{dr}'
+  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}\Delta n_3(\mathbf{r}')
+  +
+  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}\Delta n_2(\mathbf{r}')
+  +
+  \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} \cdot \Delta \mathbf{n}_{V2}(\mathbf{r}')
+\end{align}
+
 \end{widetext}
@@
 
+\begin{widetext}
+  \appendix
+
+  \section{Partial derivatives of the White Bear free energy functional}
+
+  Here is the White Bear free energy functional for the hard-sphere
+  fluid:~\cite{roth2002whitebear}
+  \begin{align}
+    A_{HS} &= \int \left\{
+    -n_0 \ln\left( 1 - n_3\right)
+    + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+    + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+      n_3 + (1-n_3)^2 \ln(1-n_3)
+    }{
+      36\pi n_3^2(1-n_3)^2
+    }
+    \right\}
+\end{align}
+
+\begin{align}
+    \beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
+    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
+      n_3(\mathbf{r}'))^2}
+    + \frac{n_2^3 -
+      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
+      \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
+    \right)
+\end{align}
+\end{widetext}
+
+\begin{align}
+    \beta\frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &= -\ln(1-n_3)
+    \\
+    \beta\frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
+    \\
+    \beta\frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
+      \frac{n_1}{1-n_3}
+      + (n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
+        (1-n_3)^2\ln(1-n_3)}{
+        12\pi n_3^2(1-n_3)^2
+      }
+\end{align}
+\begin{align}
+    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
+      \frac{\mathbf{n}_{V2}}{1-n_3}
+    \\
+    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
+      \frac{\mathbf{n}_{V1}}{1-n_3}
+      - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
+        (1-n_3)^2\ln(1-n_3)}{
+        36\pi n_3^2(1-n_3)^2
+      }
+  \end{align}
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2011-06-11 paper/exampletalk.tex

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{color}
\usepackage{verbatim}
\usetheme{Frankfurt}

\title{Energy and Entropy and More}

\author{David Roundy}
\institute{Oregon State University}
\date{}

\setbeamertemplate{navigation symbols}{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%\section*{Outline}
% \begin{frame}
%   \frametitle{Outline}
%   \tableofcontents
% \end{frame}

\section{Energy and Entropy}
\subsection*{}

\begin{frame}
  \frametitle{Energy and Entropy paradigm}
  \begin{block}{The course}
    \begin{itemize}
    \item Originally developed by Allen Wassermann.
    \item I cotaught Energy and Entropy in 2009.
    \item I taught it on my own in 2010.
    \end{itemize}
  \end{block}
  \vfill
  \begin{block}{Course challenges}
    \begin{itemize}
    \item Introduce students to thermodynamics for the first time.
    \item Tricky thermodynamic differentials.
    \item Information-theory approach to the statistical approach.
    \end{itemize}
  \end{block}
\end{frame}

\newcommand\myderiv[3]{%
  \ensuremath{\left(\frac{\partial #1}{\partial #2}\right)_{#3}}}

\begin{frame}
  \frametitle{Energy and Entropy paradigm}
  \begin{block}{Thermodynamic derivatives $\longrightarrow$ experiments}
    \emph{``Name the experiment!''} For each of the derivatives I
    write on the board, describe the experiment that you would perform
    in order to measure it.
    \begin{align}
      \myderiv{p}{V}{T} \quad
      \myderiv{p}{V}{S} \quad
      \myderiv{p}{S}{V}
    \end{align}
    \begin{itemize}
    \item Maxwell relations help for confusing cases
    \item We have labs so students get to measure thermodynamic
      quantities themselves
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Energy and Entropy paradigm}
  \begin{block}{Rubber band experiment and measuring entropy}
    Measuring tension versus temperature and length gives $\Delta U$,
    $\Delta F$ and $\Delta S$ for an isothermal stretch.

    \begin{columns}
      \begin{column}{2.4in}
        \begin{center}
          %\includegraphics[height=2in]{figs/hot-rubber}
        \end{center}
      \end{column}
      \begin{column}{2in}
        \begin{center}
          %\includegraphics[height=2in]{figs/ice-rubber}
        \end{center}
      \end{column}
    \end{columns}
  \end{block}
\end{frame}

\newcommand\fairness{\ensuremath{\mathcal{F}}}

\begin{frame}
  \frametitle{Energy and Entropy paradigm}
  \begin{block}{Information theory approach}
    \begin{itemize}
    \item Maximizing the (constrained) fairness gives the Boltzmann
      factor.
      \[
      \fairness = -k_B\sum_i^\text{all states} P_i \ln P_i
      \]
    \item Entropy is thus $\fairness_\text{max}$ rather than
      \[
      S = k_B \ln \Omega
      \]
    \item The Second Law is obvious, given this derivation of
      probabilities.
    \item Is it better?
    \end{itemize}
  \end{block}
\end{frame}

\section{Joining the Paradigms team}
\subsection*{}

\begin{frame}
  \frametitle{Teaching a new Paradigm for the first time}
  \begin{block}{}
    \begin{itemize}
    \item I taught \emph{Symmetries and Idealizations} in Fall 2009
    \item This is the \emph{first} paradigm
      \begin{itemize}
      \item Electrostatic potential
      \item Physics integrals
      \item Taylor expansions
      \item Working in groups
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Teaching a new Paradigm for the first time}
  \begin{block}{The Paradigms Portfolios wiki}
    \begin{itemize}
    \item \verb!http://physics.oregonstate.edu/portfolioswiki!
    \item Homework
    \item Descriptions of activities and discussions
    \item Time estimates---shockingly accurate!
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Teaching a new Paradigm for the first time}
  \begin{block}{Class every day, and how that changes things}
    \begin{itemize}
    \item It's easy to pick things up the next day: \\this changes how
      you think about ``finishing'' an activity
    \item Students who get behind are very behind
    \item Homework \emph{needs} to be very well synchronized
    \end{itemize}
  \end{block}
\end{frame}

\end{document}

### 2011-06-11 paper/paper.tex
fix error in description of WB functional
Tested-on: bingley (Linux 2.6.32-5-amd64)

@@
 \begin{equation}
-F_{hs}[n] = k_B T \int (\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)) d\xx \; ,
+F_\textit{hs}[n] = k_B T \int \left(\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)\right) d\xx \; ,
 \end{equation}
@@
 \Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= \frac{
-  \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+  n_3 + (1-n_3)^2 \ln(1-n_3)
 }{
-  8\pi \left( 1 - n_3 \right)^2
+  36\pi n_3^2\left( 1 - n_3 \right)^2
 } ,
@@
 from\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft}, but I don't use it.
+  fu2005vapor-liquid-dft}.  Actually,
+reference~\cite{fu2005vapor-liquid-dft} has what appear to be a number
+of mistakes in this formula, so
+reference~\cite{yu2002fmt-dft-inhomogeneous-associating} is more
+useful.  In particular, the former has an incorrect formula for
+$\zeta_2$ (wrong dimensions) and the formula for $\zeta_3$ disagrees
+with the latter, which it cites.  Its formula for $\zeta_3$ also
+doesn't match how they use it in ordinary FMT.
 \begin{align}
@@
   g^\textit{HS}(\sigma) &= \frac1{1-n_3}\left(1+\frac14\frac{\zeta n_2}{1-n_3}
-  \left(1 + \frac{n_2}{14 (1-n_3)}\right)\right)
+  \left(1 + \frac{n_2}{14 (1-n_3)}\right)\right) \\
+  &= \frac1{1 - \zeta_3}
+  + \frac32 \frac{\zeta \zeta_2}{\left(1-\zeta_3\right)^2}
+  + \frac12 \frac{\zeta \zeta_2^2}{\left(1-\zeta_3\right)^3} \\
+  \zeta_2 &= \frac{n_2 R}{3} \\
+  \zeta_3 &= n_3
 \end{align}

### 2011-06-15 contactpaper/figs/free-energy.cpp
implement plot of contact density vs filling fraction and free energy
Tested-on: collins (Linux 2.6.32-5-amd64)

@@ -1,0 +1,38 @@
+// Deft is a density functional package developed by the research
+// group of Professor David Roundy
+//
+// Copyright 2010 The Deft Authors
+//
+// Deft is free software; you can redistribute it and/or modify it
+// under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+//
+// You should have received a copy of the GNU General Public License
+// along with deft; if not, write to the Free Software Foundation,
+// Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
+//
+// Please see the file AUTHORS for a list of authors.
+
+#include <stdio.h>
+#include <time.h>
+#include "ContactDensity.h"
+
+
+int main(int, char **) { 
+  FILE *o = fopen("contactpaper/figs/free-energy.dat", "w");
+
+  Functional f = HardSpheresWBnotensor(1.0);
+  Functional eta = Identity();
+  Functional fcar = (4*eta - 3*Pow(3)(eta))/sqr(1-eta);
+  double mykT = 1.0; // so we needn't divide by kT
+
+  for (double eta=0.0001; eta<=0.3; eta *= 1.01) {
+    double fc = fcar(mykT, eta);
+    double n = eta/(4*M_PI/3);
+    double nice = f(mykT, n)/n;
+    fprintf(o, "%g\t%g\t%g\n", eta, fc, nice);
+    fflush(o);
+  }
+  fclose(o);
+}

### 2011-06-16 paper/poster.tex
add mostly-empty paper.

\documentclass[landscape,a0paper,final,showframe]{baposter}

\usepackage{times}
\usepackage{calc}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{relsize}
\usepackage{multirow}
\usepackage{bm}

\usepackage{graphicx}
\usepackage{multicol}

\usepackage{pgfbaselayers}
\pgfdeclarelayer{background}
\pgfdeclarelayer{foreground}
\pgfsetlayers{background,main,foreground}

\usepackage{helvet}
%\usepackage{bookman}
\usepackage{palatino}

\newcommand{\captionfont}{\footnotesize}

\selectcolormodel{cmyk}

\graphicspath{{images/}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% Some math symbols used in the text
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Format 
\newcommand{\Matrix}[1]{\begin{bmatrix} #1 \end{bmatrix}}
\newcommand{\Vector}[1]{\Matrix{#1}}
\newcommand*{\SET}[1]  {\ensuremath{\mathcal{#1}}}
\newcommand*{\MAT}[1]  {\ensuremath{\mathbf{#1}}}
\newcommand*{\VEC}[1]  {\ensuremath{\bm{#1}}}
\newcommand*{\CONST}[1]{\ensuremath{\mathit{#1}}}
\newcommand*{\norm}[1]{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand*{\abs}[1]{\mathopen| #1 \mathclose|}% use instead of $\|x\|$
\newcommand*{\absLR}[1]{\left| #1 \right|}% use instead of $\|x\|$

\def\norm#1{\mathopen\| #1 \mathclose\|}% use instead of $\|x\|$
\newcommand{\normLR}[1]{\left\| #1 \right\|}% use instead of $\|x\|$

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Multicol Settings
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\setlength{\columnsep}{0.7em}
\setlength{\columnseprule}{0mm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Save space in lists. Use this after the opening of the list
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand{\compresslist}{%
\setlength{\itemsep}{1pt}%
\setlength{\parskip}{0pt}%
\setlength{\parsep}{0pt}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Begin of Document
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Here starts the poster
%%%---------------------------------------------------------------------------
%%% Format it to your taste with the options
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\typeout{Poster Starts}
\definecolor{silver}{cmyk}{0,0,0,0.3}
\definecolor{yellow}{cmyk}{0,0,0.9,0.0}
\definecolor{reddishyellow}{cmyk}{0,0.22,1.0,0.0}
\definecolor{black}{cmyk}{0,0,0.0,1.0}
\definecolor{darkYellow}{cmyk}{0,0,1.0,0.5}
\definecolor{darkSilver}{cmyk}{0,0,0,0.1}

\definecolor{lightyellow}{cmyk}{0,0,0.3,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lighteryellow}{cmyk}{0,0,0.1,0.0}
\definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}
\begin{poster}{
  % Show grid to help with alignment
  grid=yes,
  % Column spacing
  colspacing=1em,
  % Color style
  bgColorOne=lighteryellow,
  bgColorTwo=lightestyellow,
  borderColor=reddishyellow,
  headerColorOne=yellow,
  headerColorTwo=reddishyellow,
  headerFontColor=black,
  boxColorOne=lightyellow,
  boxColorTwo=lighteryellow,
  % Format of textbox
  textborder=roundedleft,
  % Format of text header
  eyecatcher=yes,
  headerborder=open,
  headerheight=0.08\textheight,
  headershape=roundedright,
  headershade=plain,
  headerfont=\Large\textsf, %Sans Serif
  boxshade=plain,
%  background=shade-tb,
  background=plain,
  linewidth=2pt
  }
  % Eye Catcher
  { % No eye catcher for this poster. If an eye catcher is present,
    % the title is centered between eye-catcher and logo.
    \includegraphics[height=5.5em]{figs/osu-logo}
  }
  % Title
  {\sf %Sans Serif
  %\bf% Serif
  A Classical Density Functional for Water}
  % Authors
  {\sf %Sans Serif
  % Serif
  David Roundy\hspace{3em}
  Jessica Hughes\hspace{3em}
  Oregon State University, USA
  }
  % University logo
  {
    \includegraphics[height=5.5em]{figs/osu-logo}
  }

  \tikzstyle{light shaded}=[top color=baposterBGtwo!30!white,bottom color=baposterBGone!30!white,shading=axis,shading angle=30]

  % Width of left inset image
     \newlength{\leftimgwidth}
     \setlength{\leftimgwidth}{0.78em+8.0em}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%% Now define the boxes that make up the poster
%%%---------------------------------------------------------------------------
%%% Each box has a name and can be placed absolutely or relatively.
%%% The only inconvenience is that you can only specify a relative position 
%%% towards an already declared box. So if you have a box attached to the 
%%% bottom, one to the top and a third one which should be in between, you 
%%% have to specify the top and bottom boxes before you specify the middle 
%%% box.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %
    % A coloured circle useful as a bullet with an adjustably strong filling
    \newcommand{\colouredcircle}[1]{%
      \tikz{\useasboundingbox (-0.2em,-0.32em) rectangle(0.2em,0.32em); \draw[draw=black,fill=baposterBGone!80!black!#1!white,line width=0.03em] (0,0) circle(0.18em);}}

  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{
    We will present preliminary results for a classical density
   functional for water based on the Statistical Associating Fluid
   Theory (SAFT), designed to be suitable for treatment of either
   solutes or the solid-liquid interface.  The functional is based on
   a combination of the Fundamental Measure Theory functional
   describing hard-sphere fluids with the SAFT-VR equation of state.
   We will present bulk thermodynamic results together with results on
   the hydrophobic interaction of planes, rods and spheres.  }

  \headerbox{Statistical Associating Fluid Teory}{name=saft,column=0,below=cdft}{
    Hello world
  }

  \headerbox{The
    Functional}{name=fitting,column=0,below=saft,above=bottom}{ We
    introduce a new classical density functional for water.  The
    Helmholtz free energy functional is based on SAFT, and is composed
    of the usual terms:
    \begin{equation}
      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
    \end{equation}
    We model the repulsive interactions using the White Bear version of
    the Fundamental-Measure Theory~(FMT) functional published in
    reference~\cite{roth2002whitebear}.
  }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Water Confined in a Slit}{name=1D,column=1,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{A Cylinder in Water}{name=rod,column=1,below=1D,span=1,above=bottom}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

  }
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,row=0}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
      \begin{multicols}{2}
        This text will be in two separate columns.
        This text will be in two separate columns.
        This text will be in two separate columns.
        This text will be in two separate columns.
        This text will be in two separate columns.
        This text will be in two separate columns.
        This text will be in two separate columns.
      \end{multicols}\vspace{-1em}
      \mbox{\hspace{0.3\linewidth}\rule{0.4\linewidth}{1pt}\hspace{0.3\linewidth}}\\

%      \begin{multicols}{2}
        {
          This text will span the whole box.
          This text will span the whole box.
          This text will span the whole box.
          This text will span the whole box.
          This text will span the whole box.
          This text will span the whole box.
        %Note the different scales for the two graphs.
        %Our approach has a high accuracy on the neutral (UND) dataset.
        }
%      \end{multicols}\vspace{-1em}
  }%
  \headerbox{Future Work}{name=future work,column=2,span=1,above=bottom,below=two rods}{

  }
  \headerbox{References}{name=references,column=3,above=bottom,below=two
  rods}{
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \bibliographystyle{plain}	% (uses file "plain.bst")
    \bibliography{paper}
    %% \smaller
    %% \vspace{-0.4em}
    %% \bibliographystyle{ieee}
    %% \renewcommand{\section}[2]{\vskip 0.05em}
    %%   \begin{thebibliography}{1}\itemsep=-0.01em
    %%   \setlength{\baselineskip}{0.4em}
    %%   \bibitem{amberg07:nonrigid}
    %%     B.~Amberg, S.~Romdhani, T. Vetter.
    %%     \newblock {O}ptimal {S}tep {N}onrigid {ICP} {A}lgorithms for {S}urface {R}egistration
    %%     \newblock In {\em CVPR 2007}
    %%   \bibitem{amberg08:recognition}
    %%     B.~Amberg, R.~Knothe, T. Vetter.
    %%     \newblock Expression Invariant Face Recognition with a 3D Morphable Model
    %%     \newblock In {\em AFGR 2008}
    %%   \end{thebibliography}
  }%
\end{poster}%
%
\end{document}

### 2011-06-16 paper/paper.tex
Fixed application of constraint, added more plots, added sphere calculation.

--- 2011-06-06 paper/paper.tex
@@
   \frac{1}{4}\beta\left(\frac{\partial a_1(\xx)}{\partial n_3(\xx)} -
-  \frac{lambda}{3 n3(\xx)}\frac{\partial a_1(\xx)}{\partial \lambda}\right)
+  \frac{\lambda}{3 n3(\xx)}\frac{\partial a_1(\xx)}{\partial \lambda}\right)
 \end{align}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-1D}
+\end{center}
+\caption{ Density profile for a slab of water with molecules constrained to move in one dimension. }
+\label{fig:density-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-1D}
+\end{center}
+\caption{ Energy density profile for a slab of water with molecules constrained to move in one dimension. }
+\label{fig:energy-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/entropy-1D}
+\end{center}
+\caption{ Entropy for a slab of water with molecules constrained to move in one 
+dimension. \textcolor{red}{FIX UNITS}}
+\label{fig:entropy-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
+\end{center}
+\caption{ Xassociation value (fraction of association sites
+\emph{not} hydrogen-bonded) for a slab of water with molecules constrained to move in one 
+dimension. }
+\label{fig:xassoc-1D}
+\end{figure}
+
 \subsection{Test-particle approach}
@@
 
+\subsection{Two Hydrophobic Rods}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\end{center}
+\caption{ Density profile for two hydrophobic rods.}
+\label{fig:density-rods}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-06-16 paper/project.tex

@@
 \begin{document}
-\title{A Classical Density-Functional Theory for Describing Confined Water in One, Two and Three(?) Dimensions}
+\title{A Classical Density-Functional Theory for Describing Water Interfaces}
 
-\author{Jessica Gallagher}
+\author{Jessica Hughes}
 \affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
@@
   water by combining the fundamental-measure theory functional--which
-  describes hard sphere fluids--with attractive interactions.  Using
+  describes hard sphere fluids--with attractive interactions based on the 
+  Statistical Associating Fluid Theory (SAFT).  Using
   this approximation, we are able to describe various properties of
-  water near liquid-vapor and liquid-solid interfaces.  We plan to use
-  this functional as the foundation for a efficient method for
+  water near liquid-vapor and liquid-solid interfaces including water confined 
+  to a slit and two hydrophobic rods immersed in water.  This functional will be the 
+  foundation for a efficient method for future work
   predicting the interactions of chemicals in aqueous solution on an
   atomic level.}
+
+\tableofcontents
 \end{abstract}
@@
 \section{Theory and Methods}
-
 We introduce a new classical density functional for water that
@@
 where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of association sites
+molecule, and the functional $X$ is the fraction of assotiation sites
 \emph{not} hydrogen-bonded.  The fraction $X$ is determined by the
@@
 
-\subsection{Determining the empirical parameters}\label{sec:empirical}
+\subsection{Determining the empirical parameters (?)}\label{sec:empirical}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-1D}
+\end{center}
+\caption{ Density profile for a slab of water with molecules constrained to move in one dimension. }
+\label{fig:density-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-1D}
+\end{center}
+\caption{ Energy density profile for a slab of water with molecules constrained to move in one dimension. }
+\label{fig:energy-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/entropy-1D}
+\end{center}
+\caption{ Entropy for a slab of water with molecules constrained to move in one 
+dimension. \textcolor{red}{FIX UNITS}}
+\label{fig:entropy-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
+\end{center}
+\caption{ Xassociation value (fraction of association sites
+\emph{not} hydrogen-bonded) for a slab of water with molecules constrained to move in one 
+dimension. }
+\label{fig:xassoc-1D}
+\end{figure}
+
 The fluid density at a liquid-solid interface commonly displays
@@
 
+\subsection{Hydrophobic rods}
+
+\subsubsection{Resolution and convergence tests}
+
+\subsubsection{One rod - diameter effects}
+
+\subsubsection{Two rods - interaction effects}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\end{center}
+\caption{ Density profile for two hydrophobic rods.}
+\label{fig:density-rods}
+\end{figure}
+
+\subsection{Hard Sphere}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-06-16 paper/poster.tex
start poster and add missing file

@@
     % the title is centered between eye-catcher and logo.
-    \includegraphics[height=5.5em]{figs/osu-logo}
+    \includegraphics[height=5.5em]{figs/cavitysize-060}
   }
@@
 
-  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{
-    We will present preliminary results for a classical density
-   functional for water based on the Statistical Associating Fluid
-   Theory (SAFT), designed to be suitable for treatment of either
-   solutes or the solid-liquid interface.  The functional is based on
-   a combination of the Fundamental Measure Theory functional
-   describing hard-sphere fluids with the SAFT-VR equation of state.
-   We will present bulk thermodynamic results together with results on
-   the hydrophobic interaction of planes, rods and spheres.  }
+  \headerbox{Classical DFT}{name=cdft,column=0,row=0}{ Classical
+    Density-Functional Theory is a continuum empirical approach to
+    modelling fluid behavior when the density is inhomogeneous---as is
+    the case at interfaces.  We have developed a classical density
+    functional for water, which provides a description of hydrophobic
+    interactions.  }
 
-  \headerbox{Statistical Associating Fluid Teory}{name=saft,column=0,below=cdft}{
-    Hello world
+  \headerbox{Statistical Associating Fluid
+    Teory}{name=saft,column=0,below=cdft}{ The statistical associating
+    fluid theory (SAFT) is a well-established approach for the
+    description of the thermodynamics and phase equilibria of fluid
+    systems, with particular applicability to hydrogen-bonding fluids.
+    Our functional reduces in the homogeneous limit to the SAFT model
+    for water developed in reference~\cite{clark2006developing}.
+    \begin{center}
+      \includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+    \end{center}
+    This functional accurately describes the liquid-vapor phase
+    diagram away from the critical point, but gives a poor description
+    of the compressibility.
   }
@@
     \end{equation}
-    We model the repulsive interactions using the White Bear version of
-    the Fundamental-Measure Theory~(FMT) functional published in
-    reference~\cite{roth2002whitebear}.
+    We model the repulsive interactions using the White Bear version
+    of the Fundamental-Measure Theory~(FMT)
+    functional~\cite{roth2002whitebear}.  The association term is
+    taken from~\cite{yu2002fmt-dft-inhomogeneous-associating}, and we
+    used a simple and efficient functional for the dispersion
+    interaction with a single tunable parameter.
   }
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-  \headerbox{Water Confined in a Slit}{name=1D,column=1,row=0}{
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+  \headerbox{Surface Tension}{name=1D,column=1,row=0}{
 
   }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \headerbox{A Cylinder in Water}{name=rod,column=1,below=1D,span=1,above=bottom}{
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+    \begin{center}
+      \includegraphics[width=\columnwidth]{figs/density-vs-radius}
+    \end{center}
+    \begin{center}
+      \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+    \end{center}
   }
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
   \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,row=0}{
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
     \begin{multicols}{2}
@@
     \end{multicols}\vspace{-1em}
-      \mbox{\hspace{0.3\linewidth}\rule{0.4\linewidth}{1pt}\hspace{0.3\linewidth}}\\
+    \begin{center}
+      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
+      \hspace{1em}
+      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
+    \end{center}
 

### 2011-06-17 paper/poster.tex
add another plot to poster
fancy figures for poster
brighten paper colors

@@
 \definecolor{lightestyellow}{cmyk}{0,0,0.05,0.0}
+
+\definecolor{lightblue}{rgb}{0.8,0.8,1}
+\definecolor{lighterblue}{rgb}{0.9,0.9,1}
+\definecolor{lightestblue}{rgb}{0.98,0.98,1}
+
+\definecolor{cyan}{cmyk}{0.4,0,0.3,0}
+\definecolor{greenblue}{rgb}{0.7,1,.7}
+\definecolor{mediumblue}{rgb}{0.7,0.7,1}
 \begin{poster}{
   % Show grid to help with alignment
-  grid=yes,
+  grid=no,
   % Column spacing
@@
   % Color style
-  bgColorOne=lighteryellow,
+  bgColorOne=lightyellow,
   bgColorTwo=lightestyellow,
-  borderColor=reddishyellow,
-  headerColorOne=yellow,
-  headerColorTwo=reddishyellow,
+  borderColor=mediumblue,
+  headerColorOne=lighterblue,
+  headerColorTwo=mediumblue,
   headerFontColor=black,
-  boxColorOne=lightyellow,
-  boxColorTwo=lighteryellow,
+  boxColorOne=lighterblue,
+  boxColorTwo=lightblue,
   % Format of textbox
@@
   headershape=roundedright,
-  headershade=plain,
+  headershade=shade-tb,
   headerfont=\Large\textsf, %Sans Serif
@@
 %  background=shade-tb,
-  background=plain,
+  background=shade-tb,
   linewidth=2pt
@@
   % Eye Catcher
-  { % No eye catcher for this poster. If an eye catcher is present,
-    % the title is centered between eye-catcher and logo.
+  { % The title is centered between eye-catcher and logo.
     \includegraphics[height=5.5em]{figs/cavitysize-060}
@@
   \headerbox{Surface Tension}{name=1D,column=1,row=0}{
-
+    \begin{center}
+      \includegraphics[width=\columnwidth]{figs/surface-tension}
+    \end{center}
   }
@@
       \hspace{1em}
-      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
+      \includegraphics[width=0.45\columnwidth]{figs/xassoc-rods-in-water}
     \end{center}
@@
     %      \end{multicols}\vspace{-1em}
+    \begin{center}
+      \includegraphics[width=0.45\columnwidth]{figs/rods-energy-vs-distance}
+    \end{center}
   }%

### 2011-06-26 contactpaper/paper.tex
implement plot of contact density vs filling fraction and free energy
Tested-on: collins (Linux 2.6.32-5-amd64)
derive and discuss homogeneous limit of White Bear free energy

@@
 \section{Introduction}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/gHS-vs-n}
+\caption{A comparison of the various approximations to the contact
+  density.}
+\label{fig:gHS-vs-n}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/free-energy}
+\caption{A comparison of the various approximations to the free
+  energy.  Note that the White Bear functional uses a modified form of
+  the Carnahan equation of state.}
+\label{fig:free-energy}
+\end{figure}
 
@@
 \begin{align}
-  \frac{dn_3(\mathbf{r})}{dR} &= -n_2(\mathbf{r})\\
+  \frac{dn_3(\mathbf{r})}{dR} &= n_2(\mathbf{r})\\
   \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
@@
   \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+  \frac{dn_1(\mathbf{r})}{dR} &=
+    \frac1{4\pi R}\frac{dn_2(\mathbf{r})}{dR} - n_0(\mathbf{r}) \\
+    &= n_0(\mathbf{r}) + \frac{1}{4\pi R^2} \int 
+    \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
   \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
@@
   \frac{d A_{HS}^{(1)}}{d R} &=
-  \int \mathbf{dr}' \left(-\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
+  \int \mathbf{dr}' \left(\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
 \end{align}
@@
   \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
     n_2
@@
   &= \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
     n_2
@@
   &+
-     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) (-n_2)
+     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) n_2
      \frac{
@@
    \\
-  &-
+  &+
      \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
@@
     \beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
     + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
@@
     + \frac{n_2^3 -
-      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-      \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
-    \right)
+      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}
+    %\left(
+    %  \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
+    %\right)
+    \\
+    & \times \left(\frac{1}{n_3(1-n_3)^3} -\frac1{n_3^2(1-n_3)^2}  -
+      \frac{1}{n_3^2(1-n_3)} - 2\frac{\ln(1-n_3)}{n_3^3}\right)
 \end{align}
@@
 
+\begin{widetext}
+\section{The White Bear free energy for homogeneous systems}
+
+Here is the White Bear free energy functional for the hard-sphere
+fluid:~\cite{roth2002whitebear}
+\begin{align}
+  \frac{A_{HS}}{k_BT} &= \int \left\{
+  -n_0 \ln\left( 1 - n_3\right)
+  + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+  + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+    n_3 + (1-n_3)^2 \ln(1-n_3)
+  }{
+    36\pi n_3^2(1-n_3)^2
+  }
+  \right\} \\
+  &= \int \left\{
+  -n \ln\left( 1 - \eta\right)
+  + \frac{n \eta}{1-\eta}
+  + n^3 (4\pi R^2)^3 \frac{
+    \eta + (1-\eta)^2 \ln(1-\eta)
+  }{
+    36\pi \eta^2(1-\eta)^2
+  }
+  \right\} \\
+  &= \int \left\{
+  -n \ln\left( 1 - \eta\right)
+  + \frac{n \eta}{1-\eta}
+  + n^3 4\pi (4\pi R^3)^2 \frac{
+    \eta + (1-\eta)^2 \ln(1-\eta)
+  }{
+    36\pi \eta^2(1-\eta)^2
+  }
+  \right\} \\
+  &= \int \left\{
+  -n \ln\left( 1 - \eta\right)
+  + \frac{n \eta}{1-\eta}
+  + n\eta^2 4\pi 9 \frac{
+    \eta + (1-\eta)^2 \ln(1-\eta)
+  }{
+    36\pi \eta^2(1-\eta)^2
+  }
+  \right\} \\
+  &= N \left\{
+  -\ln\left( 1 - \eta\right)
+  + \frac{\eta}{1-\eta}
+  + \frac{
+    \eta + (1-\eta)^2 \ln(1-\eta)
+  }{
+    (1-\eta)^2
+  }
+  \right\} \\
+  &= N \left\{
+  -\ln\left( 1 - \eta\right)
+  + \frac{\eta}{1-\eta}
+  + \frac{
+    \eta
+  }{
+    (1-\eta)^2
+  }
+  + \ln(1-\eta)
+  \right\} \\
+  &= N \left\{
+  \frac{\eta}{1-\eta}
+  + \frac{
+    \eta
+  }{
+    (1-\eta)^2
+  }
+  \right\} \\
+  &= N
+  \frac{\eta(1-\eta) + \eta
+  }{
+    (1-\eta)^2
+  } \\
+  &= N
+  \frac{2\eta -\eta^2}{(1-\eta)^2}
+\end{align}
+This is odd, because Wikipedia says that for the Carnahan-Starling
+equation of state the excess free energy is
+\begin{align}
+  A = Nk_BT\frac{4\eta -3\eta^2}{(1-\eta)^2}
+\end{align}
+\end{widetext}
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2011-06-27 paper/project.tex
Project outline and adding high and low resolution versions of single-rod

@@
 
-We compute the density profile and surface tension by relaxing a slab
-of water in a periodic unit cell.  This slab has two surfaces, so the
-surface tension is just half the free energy of the cell, divided by
-its cross-sectional area.  As mentioned in
-section~\ref{sec:empirical}, the $b$ parameter is determined by
-setting this surface tension to match the experimental value.  The
-associated density profile is plotted in
-Figure~\ref{fig:cartesianDensity}, which displays two peaks in
-density.
+\subsection{Water confined to a slit}
 
@@
 
-\subsubsection{Two rods - interaction effects}
+\subsubsection{Two rods - hydrophobic interaction}
 

### 2011-06-27 paper/poster.tex
add nice little cartoon for the poster.

@@
   { % The title is centered between eye-catcher and logo.
-    \includegraphics[height=5.5em]{figs/cavitysize-060}
+    \includegraphics[height=5.5em]{figs/cavity-cartoon}
   }
@@
     functional for water, which provides a description of hydrophobic
-    interactions.  }
+    interactions.
+
+    \vspace{1em}
+  }
 
   \headerbox{Statistical Associating Fluid
-    Teory}{name=saft,column=0,below=cdft}{ The statistical associating
-    fluid theory (SAFT) is a well-established approach for the
+    Teory}{name=saft,column=0,below=cdft}{ Statistical Associating
+    Fluid Theory (SAFT) is a well-established approach for the
     description of the thermodynamics and phase equilibria of fluid
     systems, with particular applicability to hydrogen-bonding fluids.
-    Our functional reduces in the homogeneous limit to the SAFT model
-    for water developed in reference~\cite{clark2006developing}.
+    Our functional is constructed so as to reduce in the homogeneous
+    limit to the 4-site SAFT model for water developed in
+    reference~\cite{clark2006developing}.
     \begin{center}
@@
     This functional accurately describes the liquid-vapor phase
-    diagram away from the critical point, but gives a poor description
-    of the compressibility.
+    diagram away from the critical point, but gives less accurately
+    describes the compressibility.
+
+    \vspace{1em}
   }
 
-  \headerbox{The
-    Functional}{name=fitting,column=0,below=saft,above=bottom}{ We
-    introduce a new classical density functional for water.  The
-    Helmholtz free energy functional is based on SAFT, and is composed
-    of the usual terms:
-    \begin{equation}
-      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
-    \end{equation}
+  \headerbox{The Functional}{name=functional,column=1,row=0}{
+    The Helmholtz free energy of the new functional is based on SAFT,
+    and is composed of the usual terms:
+    \begin{equation*}
+      F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp}[n],
+    \end{equation*}
     We model the repulsive interactions using the White Bear version
     of the Fundamental-Measure Theory~(FMT)
-    functional~\cite{roth2002whitebear}.  The association term is
-    taken from~\cite{yu2002fmt-dft-inhomogeneous-associating}, and we
-    used a simple and efficient functional for the dispersion
-    interaction with a single tunable parameter.
+    functional~\cite{roth2002whitebear} for the hard sphere fluid,
+    which is expressed in terms of efficiently-computable
+    convolutions.
+    \vspace{0.5em}
+
+    The SAFT free energy of association for a 4-site model of water
+    has the form
+    \begin{align*}
+      f_\text{assoc} &= 4 k_BT n \left(\ln X - \frac{X}{2} + \frac12\right)
+    \end{align*}
+    where $X$ is the fraction of hydrogens \emph{not} hydrogen bonded,
+    and is found using a mass action equation involving the density of
+    hard spheres in contact with one another.  We extend this to the
+    inhomogeneous case using the approach described by Yu and
+    Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which expresses
+    $X$ in terms of the same weighted densities used in FMT.
+    \vspace{0.5em}
+
+    Finally, we add a simple and efficient functional for the
+    dispersion interaction based on
+    SAFT-VR~\cite{gil-villegas-1997-SAFT-VR}, with an additional
+    tunable parameter to adjust the surface tension.
+
+    \vspace{0.8em}
   }
 
-  \headerbox{Surface Tension}{name=1D,column=1,row=0}{
+  \headerbox{Surface
+    Tension}{name=surfacetension,column=1,below=functional,above=bottom}{
+    We compute the surface tension by finding the free energy of a
+    liquid-vapor interface.  The surface tension is fit to experiment
+    by adjusting a length-scale parameter in the dispersion term of
+    the free energy.
     \begin{center}
@@
   }
-  \headerbox{A Cylinder in Water}{name=rod,column=1,below=1D,span=1,above=bottom}{
+  \headerbox{A Single Cylinder in Water}{name=rod,column=2,span=2,row=0}{
+    \begin{multicols}{2}
+      One of the simplest multi-dimensional inhomogeneous systems is a
+      hard cylinder in water.
       \begin{center}
-      \includegraphics[width=\columnwidth]{figs/density-vs-radius}
+        \includegraphics[width=0.9\columnwidth]{figs/density-vs-radius}
       \end{center}
+      \begin{quote}
+        The density of water around a
+        hard-walled cylinder with 1~nm diameter.
+      \end{quote}
+
       \begin{center}
@@
       \end{center}
+      \begin{quote}
+        The hydration energy per unit area of a hard
+        cylinder.  As expected, this asymptotes to the same value as
+        the surface tension.
+      \end{quote}
+    \end{multicols}
   }
-  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,row=0}{
+  \headerbox{Two Cylinders in Water}{name=two rods,column=2,span=2,below=rod,above=bottom}{
+    %% \begin{multicols}{2}
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %%   This text will be in two separate columns.
+    %% \end{multicols}\vspace{-1em}
     \begin{multicols}{2}
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-      This text will be in two separate columns.
-    \end{multicols}\vspace{-1em}
+      \includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+
+      \vspace{-1em}
+      \begin{quote}
+        The density of water in the vicinity of two hard-walled
+        cylinders, each 1~nm in diameter, and separated by 1~nm.
+      \end{quote}
+
+      \includegraphics[width=\columnwidth]{figs/xassoc-rods-in-water}
+
+      \vspace{-1em}
+      \begin{quote}
+      Here we plot the number of hydrogen bonds per molecule for the
+      same system of two hard-walled cylinders.
+      \end{quote}
+    \end{multicols}
+
+    \begin{multicols}{2}
+      One application of our model is the hydrophobic attraction
+      between two cylindrical molecules in water, such as two
+      nanotubes.  We study the interactions between two hard cylinders
+      with 1~nm diameter.
       \begin{center}
-      \includegraphics[width=0.45\columnwidth]{figs/density-rods-in-water}
-      \hspace{1em}
-      \includegraphics[width=0.45\columnwidth]{figs/xassoc-rods-in-water}
+        \includegraphics[width=0.95\columnwidth]{figs/two-rods-diagram}
       \end{center}
-
-    %      \begin{multicols}{2}
-    {
-      This text will span the whole box.
-      This text will span the whole box.
-      This text will span the whole box.
-      This text will span the whole box.
-      This text will span the whole box.
-      This text will span the whole box.
-      %Note the different scales for the two graphs.
-      %Our approach has a high accuracy on the neutral (UND) dataset.
-    }
-    %      \end{multicols}\vspace{-1em}
       \begin{center}
-      \includegraphics[width=0.45\columnwidth]{figs/rods-energy-vs-distance}
+        \includegraphics[width=0.9\columnwidth]{figs/rods-energy-vs-distance}
       \end{center}
+    \end{multicols}
   }%
-  \headerbox{Future Work}{name=future work,column=2,span=1,above=bottom,below=two rods}{
-
-  }
-  \headerbox{References}{name=references,column=3,above=bottom,below=two
-  rods}{
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-    \bibliographystyle{plain}	% (uses file "plain.bst")
+  %% \headerbox{Future Work}{name=future work,column=3,span=1,above=two rods,row=0}{
+  %%   \begin{itemize}
+  %%   \item Extend functional to depend on hydrogen and oxygen densities
+  %%     separately, including intramolecular interactions in the
+  %%     harmonic approximation.
+  %%   \item Introduce elecrtrostatic interactions.
+  %%   \item Couple with.
+  %%   \end{itemize}
+  %% }
+  \headerbox{References}{name=references,column=0,above=bottom,below=saft}{
+    \renewcommand\refname{}
+    \bibliographystyle{unsrt}
+    \tiny
     \bibliography{paper}
-    %% \smaller
-    %% \vspace{-0.4em}
-    %% \bibliographystyle{ieee}
-    %% \renewcommand{\section}[2]{\vskip 0.05em}
-    %%   \begin{thebibliography}{1}\itemsep=-0.01em
-    %%   \setlength{\baselineskip}{0.4em}
-    %%   \bibitem{amberg07:nonrigid}
-    %%     B.~Amberg, S.~Romdhani, T. Vetter.
-    %%     \newblock {O}ptimal {S}tep {N}onrigid {ICP} {A}lgorithms for {S}urface {R}egistration
-    %%     \newblock In {\em CVPR 2007}
-    %%   \bibitem{amberg08:recognition}
-    %%     B.~Amberg, R.~Knothe, T. Vetter.
-    %%     \newblock Expression Invariant Face Recognition with a 3D Morphable Model
-    %%     \newblock In {\em AFGR 2008}
-    %%   \end{thebibliography}
   }%

### 2011-06-27 paper/project.tex

@@
 
+\textcolor{red}{I found a nice paper on the hard-slit
+  problem... actually a couple of papers that address
+  it~\cite{forsman1996computer, lum1999hydrophobicity}}
+
 \subsection{Hydrophobic rods}
@@
 \subsubsection{Two rods - hydrophobic interaction}
+
+Here we will want to cite~\cite{walther2004hydrodynamic}, which gives
+simulations for two interacting nanotubes in water.  It's not
+precisely the same, but if we can't find something more similar, this
+would be worth citing.
 

### 2011-06-29 paper/project.tex

--- 2011-06-16 paper/project.tex
@@
 \subsubsection{Two rods - interaction effects}
+
+Here we will want to cite~\cite{walther2004hydrodynamic}, which gives
+simulations for two interacting nanotubes in water.  It's not
+precisely the same, but if we can't find something more similar, this
+would be worth citing.
 

### 2011-06-29 papers/contact/figs/gHS-vs-n.cpp
fix bug in simplification of white bear and fix bug in build

@@ -16,11 +16,13 @@
 
 #include <stdio.h>
 #include <time.h>
+#include <cassert>
 #include "ContactDensity.h"
 
 
 int main(int, char **) { 
-  FILE *o = fopen("contactpaper/figs/gHS-vs-n.dat", "w");
+  FILE *o = fopen("papers/contact/figs/gHS-vs-n.dat", "w");
+  assert(o);
 
   Functional cd = ContactDensitySimplest(1.0);
   Functional ghs = gHS(Identity(), 1.0);

### 2011-07-01 papers/contact/paper.tex
fix bug in simplification of white bear and fix bug in build
fix bugs in contact density (boy that was a pain)

--- 2011-06-26 contactpaper/paper.tex
@@
   \\
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R}
+  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \label{eq:dAhsdR}
 \end{align}
@@
   \begin{align}
-    A_{HS} &= \int \left\{
+    \beta A_{HS} &= \int \left\{
     -n_0 \ln\left( 1 - n_3\right)
@@
     \\
-    & \times \left(\frac{1}{n_3(1-n_3)^3} -\frac1{n_3^2(1-n_3)^2}  -
-      \frac{1}{n_3^2(1-n_3)} - 2\frac{\ln(1-n_3)}{n_3^3}\right)
+    & \times \left(\frac{2}{n_3(1-n_3)^3} -\frac1{n_3^2(1-n_3)^2}  -
+      \frac{1}{n_3^2(1-n_3)} - 2\frac{\ln(1-n_3)}{n_3^3}\right) \\
+    \beta\frac{\partial A_{HS}}{\partial n_3}
+    &\sim \frac{n}{1-\eta} + \frac{3n\eta}{(1-\eta)^2} +
+      \frac{2n\eta}{(1-\eta)^3} -
+      \frac{n}{(1-\eta)^2} -
+      \frac{n}{1-\eta} -
+      2n \frac{ \ln(1-\eta) }{ \eta } \\
+    &= \frac{3n\eta}{(1-\eta)^2} +
+      \frac{2n\eta}{(1-\eta)^3} -
+      \frac{n}{(1-\eta)^2} -
+      2n \frac{ \ln(1-\eta) }{ \eta } \\
+    &= n\left(
+      \frac{3\eta - 3 \eta^2}{(1-\eta)^3} +
+      \frac{2\eta}{(1-\eta)^3} -
+      \frac{1 - \eta}{(1-\eta)^3} -
+      2 \frac{ \ln(1-\eta) }{ \eta }
+      \right) \\
+    &= n\left(
+      \frac{6\eta - 3 \eta^2 - 1}{(1-\eta)^3} -
+      2 \frac{ \ln(1-\eta) }{ \eta }
+      \right)
 \end{align}
@@
     \\
+    \beta\frac{\partial A_{HS}}{\partial n_1}
+    &\sim n \frac{4\pi R^2}{1-\eta}
+\end{align}
+\begin{align}
     \beta\frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
@@
       }
+\end{align}
+\begin{align}
+    \beta\frac{\partial A_{HS}}{\partial n_2}
+    &\sim n\frac{R}{1-\eta} +
+    \frac{4\pi}{3} R^4 n^2 \frac{\eta + (1-\eta)^2\ln(1-\eta)}{\eta^2(1-\eta)^2}
+    \\
+    &= n\frac{R}{1-\eta} +
+    \frac{R n}{(1-\eta)^2}
+    + Rn \frac{\ln(1-\eta)}{\eta}
+    \\
+    &= nR\left( \frac{1 + \eta^2  - 2\eta}{(1-\eta)^3} +
+      \frac{1 - \eta}{(1-\eta)^3}
+    + \frac{\ln(1-\eta)}{\eta}
+    \right)
+    \\
+    &= nR\left( \frac{2 + \eta^2  - 3\eta}{(1-\eta)^3}
+    + \frac{\ln(1-\eta)}{\eta}
+    \right)
 \end{align}
@@
   \end{align}
+
+\begin{widetext}
+\begin{align}
+  \beta \frac{\partial A_{HS}}{\partial R} &\sim
+    \beta \frac{\partial A_{HS}}{\partial n_3} n_2 +
+    \beta \frac{\partial A_{HS}}{\partial n_2} 2 \frac{n_2}{R} +
+    \beta \frac{\partial A_{HS}}{\partial n_1} n
+  \\
+  &=
+    n_2n \left(
+      \frac{6\eta - 3\eta^2 - 1}{(1-\eta)^3} -
+      2\frac{\ln(1-\eta)}{\eta}
+    \right) +
+    n_2n \left( \frac{4+2\eta^2-6\eta}{(1-\eta)^3} + 2\frac{\ln(1-\eta)}{\eta} \right) +
+    n_2n\frac{1}{1-\eta}
+  \\
+  &=
+    n_2n \left(
+      \frac{6\eta - 3\eta^2 - 1}{(1-\eta)^3} +
+    \frac{4+2\eta^2-6\eta}{(1-\eta)^3} +
+    \frac{1 - 2\eta + \eta^2}{(1-\eta)^3}
+  \right)
+  \\
+  &=
+    n_2n \frac{4 - 2\eta}{(1-\eta)^3}
+\end{align}
+\end{widetext}
+Recall from Equation~\ref{eq:dAhsdR}, the standard Carnahan result is
+\begin{align}
+  \frac{dA_{HS}}{dR}
+  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R}
+  \\ &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} n_2
+\end{align}
+which means that we miraculously come out with a correct answer!
 
@@
   -n \ln\left( 1 - \eta\right)
-  + \frac{n \eta}{1-\eta}
+  + \frac{n^2 (4\pi R^2)  (R)}{1-\eta}
   + n^3 (4\pi R^2)^3 \frac{
@@
   -n \ln\left( 1 - \eta\right)
-  + \frac{n \eta}{1-\eta}
+  + \frac{3 n \eta}{1-\eta}
+  + n^3 (4\pi R^2)^3 \frac{
+    \eta + (1-\eta)^2 \ln(1-\eta)
+  }{
+    36\pi \eta^2(1-\eta)^2
+  }
+  \right\} \\
+  &= \int \left\{
+  -n \ln\left( 1 - \eta\right)
+  + \frac{3 n \eta}{1-\eta}
   + n^3 4\pi (4\pi R^3)^2 \frac{
@@
   -n \ln\left( 1 - \eta\right)
-  + \frac{n \eta}{1-\eta}
+  + \frac{3 n \eta}{1-\eta}
   + n\eta^2 4\pi 9 \frac{
@@
   -\ln\left( 1 - \eta\right)
-  + \frac{\eta}{1-\eta}
+  + \frac{3 \eta}{1-\eta}
   + \frac{
@@
   &= N \left\{
-  \frac{\eta}{1-\eta}
+  \frac{3\eta}{1-\eta}
   + \frac{
@@
   &= N
-  \frac{\eta(1-\eta) + \eta
+  \frac{3\eta(1-\eta) + \eta
   }{
@@
   &= N
-  \frac{2\eta -\eta^2}{(1-\eta)^2}
+  \frac{4\eta -3\eta^2}{(1-\eta)^2}
 \end{align}
-This is odd, because Wikipedia says that for the Carnahan-Starling
-equation of state the excess free energy is
+This matches the usual expression for the Carnahan-Starling equation
+of state the excess free energy, which is
 \begin{align}

### 2011-07-07 paper/project.tex

--- 2011-06-27 paper/project.tex
@@
 \subsection{Hard-sphere repulsion}
-The repulsive intermolecular interactions are modeled using a
-Fundamental\nobreakdash-Measure Theory~\cite{rosenfeld1997}(FMT) functional,
-which combines scaled-particle theory and the Percus-Yevick theory,
-and an attractive interaction.  The FMT functional treats
-each water molecule as a hard sphere with a radius approximately that of
-a water molecule $(R~=~2.3~bohr)$.\\
 
 We model the repulsive interactions using the White Bear version of
-the Fundamental-Measure Theory~(FMT) functional published in
-reference~\cite{roth2002whitebear}.  \needsworklater{The FMT
-  functional describes the excess free energy of a hard-sphere
-  fluid. Fundamental-Measure Theory has several advantages over other
-  approaches, one of the most important being the simplicity of
-  calculations.  This particular FMT also reliably predicts bulk
-  properties of fluids in one, two and three dimensions, as well as
-  exactly determines the excess free energy in the zero dimensional
-  limit.}
+the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
+fluid~\cite{roth2002whitebear}.  In the homogeneous limit, the White
+Bear functional reduces to the Mansoori-Carnahan-Starling-Leland
+equation of state.  FMT functionals are expressed as the integral of
+the \emph{fundamental measures} of a fluid, which provide local
+measures of quantities such as the packing fraction, density of
+spheres touching a given point and mean curvature.  The excess free
+energy is written as:
 \begin{equation}
@@
 \Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= \frac{
-  \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
+\Phi_3 &= (n_2^3 - 3n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
+  \frac{
+    n_3 + (1-n_3)^2\ln(1-n_3)
   }{
-  8\pi \left( 1 - n_3 \right)^2
+    36\pi n_3^2\left( 1 - n_3 \right)^2
   } ,
 \end{align}
-using the weighted densities
+where the fundamental measure densities are given by:
 \begin{align}
@@
 \end{align}
+For our water functional, we use the hard-sphere radius of
+3.03420~\AA, which was found to be optimal by Clark
+\emph{et al}.\cite{clark2006developing}
 

### 2011-07-07 papers/electrostatics/paper.tex

--- 2011-06-09 contactpaper/paper.tex
@@
 \begin{document}
-\title{A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities}
+\title{A Classical Density Functional Electrostatic Interactions in Water }
 
@@
 \section{Introduction}
+We would like to create a classical density functional for water that
+correctly describes its interaction with charged systems.  There are a
+couple of effects that seem particularly important to get accurately
+in order to provide a reasonable description for the solvation of
+ions.
 
-The hard-sphere contact density---which we will define as the density
-of hard spheres touching a given hard sphere---is important for the
-association term in Statistical Associating Fluid Theory (SAFT).  For
-a homogeneous hard-sphere fluid, the contact density---which is to
-say, $n g_{HS}(\sigma)$---is easily computed from the
-Carnahan-Starling equation of state
-\begin{align}
-  g_{HS}(\sigma) &= \frac{1}{1-\eta} + \cdots
-\end{align}
-The contact density may be easily computed from a knowledge of the
-contact theorem, which states that the pressure on any hard surface is
-given by
-\begin{align}
-  p &= k_BT n_\textit{contact}
-\end{align}
-Since we are interested in the contact density at the hard-sphere
-surface, all we need compute is the pressure on that surface, and
-we'll have our answer.  The pressure on a hard sphere can be readily
-computed from the dependence of the free energy on hard sphere
-radius.
-\begin{align}
-  A_{HS} &= Nk_BT \frac{4\eta - 3\eta^2}{(1-\eta)^2}
-\end{align}
-Since $\eta \equiv \frac{4\pi}{3} R^3$, we can work out that
-\begin{align}
-  \frac{dA_{HS}}{dR} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{dR} \\
-  &= Nk_BT \left( \frac{4 - 6\eta}{(1-\eta)^2} + 2 \frac{4\eta - 3\eta^2}{(1-\eta)^3} \right) \frac{d\eta}{dR}
-  \\
-  &= Nk_BT \frac{4 - 4\eta - 6\eta + 6\eta^2 + 8\eta - 6\eta^2}{(1-\eta)^3} \frac{d\eta}{dR}
-  \\
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{d\eta}{dR}
-  \\
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R}
-\end{align}
-This derivative gives us the force on \emph{all} the hard
-spheres---since we're changing all their radii at once.  To compute
-the pressure on the spheres, we just need to divide by the total area,
-which means dividing by $N 4\pi (2R)^2$.  The area which is relevant
-is the area over which the molecules can make contact.
-\begin{align}
-  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
-  &= \frac14 \frac{1}{N 4\pi R^2} Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \\
-  &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
-  &= n k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3}
-\end{align}
-\begin{align}
-  n_\textit{contact} &= n \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
-  g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3}
-\end{align}
-As it turns out, this is the standard answer for the contact value of
-the autocorrelation function! $\ddot\smile$
+First, we should include the nonlocality of the dielectric response:
+water is unable to polarize at length scales small compared to its
+molecular diameter.  This is important because many ions are in fact
+of approximately the same size as water.
 
-\section{Fundamental-Measure Theory}
+Secondly, the polarization of water saturates.  There are only so many
+water molecules in a given volume, which means the polarization will
+slowly asymptote at large electric fields.  This is relevant because
+the  first solvation shell is likely to be close to fully polarized
+around a small ion.
 
-We use the White Bear version of the Fundamental-Measure Theory~(FMT)
-functional published in reference~\cite{roth2002whitebear}.  The FMT
-functional describes the excess free energy of a hard-sphere fluid.
-This particular FMT reduces to the Carnahan-Starling equation of state
-for homogeneous systems.
-\begin{equation}
-A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)\right) d\xx \; ,
-\end{equation}
-with integrands
-\begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
-\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-  n_3 + (1-n_3)^2 \ln(1-n_3)
-}{
-  36\pi n_3^2\left( 1 - n_3 \right)^2
-} ,
-\end{align}
-using the weighted densities
-\begin{align}
-  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
-  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
-\end{align}
-\begin{align}
-  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 , \quad
-  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{4\pi R} \\
-  n_1 &= \frac{n_2}{4\pi R} , \quad
-  n_0 = \frac{n_2}{4\pi R^2}
-\end{align}
-
-\begin{widetext}
-
-\section{Mean Contact density}
-
-One reasonably simple question to ask is what the \emph{mean} contact
-density of an inhomogeneous system is.  This comes down to simply
-reproducing the derivation for a homogeneous system using the
-FMT functional.
-\begin{align}
-  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR}
-\end{align}
-So to compute this we need
-\begin{align}
-  \frac{d A_{HS}}{d R} &=
-  \frac{\partial A_{HS}}{\partial R} \\
-  &+
-  \int \left(
-  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-  \frac{d n_3(\mathbf{r}')}{d R}
-  +
-  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-  \frac{d n_2(\mathbf{r}')}{d R}
-  + \cdots
-  \right) d\mathbf{r}'
-\end{align}
-
-\begin{align}
-  \frac{dn_3(\mathbf{r})}{dR} &= -n_2(\mathbf{r})\\
-  \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
-  \delta'(|\mathbf{r'}| - R)\mathbf{dr}'\\
-  &= \int n(\mathbf{r}+\mathbf{r}')
-  \delta'(r' - R) d\Omega' r'^2 dr'\\
-  &= \int 
-  \delta(r' - R) d\Omega' \left(\frac{dn(\mathbf{r}+\mathbf{r}')}{dr'}r'^2 + 2n(\mathbf{r}+\mathbf{r}')r' \right) dr'\\
-  &= \frac1{R} \int 
-  \delta(|\mathbf{r}'| - R)
-  \left(\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') +
-  2n(\mathbf{r}+\mathbf{r}') \right) \mathbf{dr}' \\
-  &= \frac{2}{R}n_2(\mathbf{r}) + \frac{1}{R} \int 
-  \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-  \frac{dn_0(\mathbf{r})}{dR} &= \frac{1}{R} \int 
-  \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-  \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
-\end{align}
-
-The total force due to $\Phi_1$ is actually very easy to work out:
-\begin{align}
-  \frac{d A_{HS}^{(1)}}{d R} &=
-  \int \mathbf{dr}' \left(-\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
-\end{align}
-where you can see that $\frac{dn_0(\mathbf{r})}{dR}$ is zero for a
-homogeneous system, and small for reasonably smooth density
-distributions.  The next term from $\Phi_2$ is as follows:
-\begin{align}
-  \frac{d A_{HS}^{(2)}}{d R} &=
-  \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2} - 2 n_1
-      \frac{dn_2}{dR} + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot
-    }{1-n_3}
-  \right) \\
-  &= \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      - 2 n_1 \left(\frac2{R}n_2 \frac{dn_0}{dR}\right)
-      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-  \right) \\
-  &= \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      - 4 n_0n_2 - 2 n_1\frac{dn_0}{dR}
-      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-  \right) \\
-  &= \int \mathbf{dr}' \left(
-    -\frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    + \frac{3n_0n_2 + \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      + 2 n_1\frac{dn_0}{dR}
-      - 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-  \right)
-\end{align}
-
-And the last term looks like this.  I should perhaps note that this
-term should be replaced by a tensor version if want to examine
-seriously localized density distributions.
-\begin{align}
-  A_{HS}^{(3)} &=
-  \int \mathbf{dr}'\left(
-     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-     \frac{
-       n_3 + (1-n_3)^2 \ln(1-n_3)
-     }{
-       36\pi n_3^2\left( 1 - n_3 \right)^2
-     }
-  \right) \\
-  \frac{d A_{HS}^{(3)}}{d R} &=
-  \int \mathbf{dr}'
-     \left[
-       3(n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-       \left(\frac2{R}n_2 + \frac{dn_0}{dR}\right)
-       - 6n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-     \right]
-     \frac{
-       n_3 + (1-n_3)^2 \ln(1-n_3)
-     }{
-       36\pi n_3^2\left( 1 - n_3 \right)^2
-     }
-   \\
-  &+
-     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) (-n_2)
-     \frac{
-       (1 - 2(1-n_3)\ln(1-n_3) - (1-n_3))n_3^2( 1 - n_3)^2
-       -
-       (n_3 + (1-n_3)^2 \ln(1-n_3))(-2 n_3^2( 1 - n_3) + 2n_3(1-n_3)^2)
-     }{
-       36\pi (n_3^2\left( 1 - n_3 \right)^2)^2
-     }
-     \\
-     &=
-     \int \mathbf{dr}'
-     \left[
-       (n_1n_2 - \mathbf{n}_{V1} \cdot \mathbf{n}_{V2})
-       \left(2n_2 + \frac{1}{4\pi}\frac{dn_0}{dR}\right)
-       - \frac{1}{2\pi}n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-     \right]
-     \frac{
-       n_3 + (1-n_3)^2 \ln(1-n_3)
-     }{
-       3 n_3^2\left( 1 - n_3 \right)^2
-     }
-   \\
-  &-
-     \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
-   \left( \frac{1}{(1-n_3)^2} + 2 \frac{\ln(1-n_3)}{n_3^2(1-n_3)} \right)
-\end{align}
-
-\begin{align}
-  \frac{d A_{HS}}{d R} &=
-  \int \mathbf{dr}' \left(
-    -\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3)
-    + \cdots
-  \right)
-\end{align}
-
-\section{Contact density at a given hard sphere}
-
-When we consider an inhomogeneous system, we may ask for \emph{which}
-sphere we are interested in computing the contact density.  One way to
-answer this is to specify a sphere at a fixed position $\mathbf{r}$.
-By making the hard-sphere radius dependent on position, we can compute
-the pressure on the hard spheres located at $\mathbf{r}$.
-\begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{4\pi (2R)^2} \frac{\delta A_{HS}}{\delta R(\mathbf{r})}
-\end{align}
-So to compute this we need
-\begin{align}
-  \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
-  \int \left(
-  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-  \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})}
-  +
-  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-  \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})}
-  + \cdots
-  \right) d\mathbf{r}'
-\end{align}
-There are \emph{many} such terms, and this could get expensive.
-
-  First, let's rewrite the free energy in terms of a smaller set of
-  weighted densities:
-  \begin{align}
-    A_{HS} &= \int \left\{
-      -n_0 \ln\left( 1 - n_3\right)
-      + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-      + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-        n_3 + (1-n_3)^2 \ln(1-n_3)
-      }{
-        36\pi n_3^2(1-n_3)^2
-      }
-    \right\}
-    \\
-    &= \int \left\{
-      -\frac{n_2}{4\pi R^2} \ln\left( 1 - n_3\right)
-      + \frac{1}{4\pi R} \frac{n_2^2 - \mathbf{n}_{V2} \cdot\mathbf{n}_{V2}}{1-n_3}
-      + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-        n_3 + (1-n_3)^2 \ln(1-n_3)
-      }{
-        36\pi n_3^2(1-n_3)^2
-      }
-    \right\}
-\end{align}
-
-\begin{align}
-    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-      n_3(\mathbf{r}'))^2}
-    + \frac{n_2^3 -
-      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-      \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
-    \right)
-    \\
-    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-    -\frac{1}{4\pi R^2} \ln\left( 1 - n_3\right)
-    +  \frac{2n_1}{1-n_3}
-    + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
-      (1-n_3)^2\ln(1-n_3)}{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \\
-    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-    \frac{2\mathbf{n}_{V1}}{1-n_3}
-    - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
-      (1-n_3)^2\ln(1-n_3)}{
-      36\pi n_3^2(1-n_3)^2
-    }    \\
-    \frac{\delta A_{HS}}{\delta R(\mathbf{r}')} &=
-    \frac{n_2}{2\pi R^3} \ln\left( 1 - n_3\right)
-    - \frac1{R} \frac{n_1n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-  \end{align}
-
-\begin{align}
-  \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \\
-  \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  n(\mathbf{r}) \delta'(|\mathbf{r}-\mathbf{r}'| - R) \\
-  \frac{\delta \mathbf{n}_{V2}(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  -\frac{\delta}{\delta R(\mathbf{r})} \mathbf{\nabla}' n_3(\mathbf{r}')
-  \\
-  &= -\mathbf{\nabla}'
-  \left(n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \right)
-\end{align}
-These give us some basic tools with which to perform derivatives.
-
-Finally, we get an answer like
-\begin{align}
-  \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
-  \int \mathbf{dr}' \left\{
-  \left(
-  \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-  + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-    n_3(\mathbf{r}'))^2}
-  + \frac{1}{4\pi}\frac{
-    \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
-  }{
-    (1-n_3(\mathbf{r}'))^3
-  }
-  \right) n(\mathbf{r}') \delta(|\mathbf{r}-\mathbf{r}'| - R)
-  + \cdots
-  \right\}
-\end{align}
-This comes out to quite a mess.  It wouldn't be infeasible to compute
-this, but it wouldn't be efficient or pretty either.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:n0}
-\end{figure}
-
-\section{Contact density at a given contact position}
-
-Unfortunately, knowing the average contact density on each hard sphere
-doesn't really give us what we want for SAFT.  The trouble is that in
-order to work out the fraction of association sites occupied, we need
-to know the contact density \emph{at the point of contact}.
-Fortunately, we don't need to know the contact density for a
-particular sphere at each point of contact.  Just as $n_0(\mathbf{r})$
-gives us the density of spheres that are contacting point $\mathbf{r}$
-(see Figure~\ref{fig:n0}), we can use similar reasoning to work out the
-contact density for these spheres.
-
-To find the contact density at a given point of contact, we can't
-simply change the radius for spheres touching that point, since that
-would give us the pressure averaged over the entire surface of those
-spheres.  Instead, we'll have to imagine extending a patch of those
-spheres which is located at the point of interest, as illustrated in
-Figure~\ref{fig:contact}.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/contact}
-\caption{Construction for finding the contact density at a given point
-  of contact.  We envision a small area $dA$ of each sphere that touches
-the point $\mathbf{r}$ extended by a distance $dR$.}
-\label{fig:contact}
-\end{figure}
-
-So we need to find the small change in free energy that is caused by
-these small changes in the shape and size of the hard spheres that are
-in contact with this point.  In the following section, we will work
-out the changes in the weighted densities at position $\mathbf{r}'$
-when point of contact we are considering is $\mathbf{r}$.
-
-The packing fraction $n_3$ is changed at the point where the sphere is
-extended, since a few more spheres will overlap this point.
-\begin{align}
-  \Delta n_3(\mathbf{r}') &= n(\mathbf{r})\Delta A \Delta R
-  \delta(|\mathbf{r}-\mathbf{r}'| - R)
-\end{align}
-The density $n_0$ is the average density, averaged over a spherical
-shell.  As such, shifting the position of a shell just shifts which
-location we are averaging over.
-\begin{align}
-  \Delta n_0(\mathbf{r}') &=
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-      \mathbf{\nabla}n(\mathbf{r}) \Delta A \Delta R
-    \delta(|\mathbf{r}-\mathbf{r}'|-R)
-\end{align}
-The change in $n_2$ comes from two terms: a change in the density value, as
-we saw in $n_0$, and a change in the area of the segment.
-\begin{align}
-  \Delta n_2(\mathbf{r}') &=
-  \left(n\left(\mathbf{r} + \Delta R
-  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
-  R}{R}\right)- n(\mathbf{r})\right) \Delta A
-  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
-  &=
-  \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
-    \right) \Delta A \Delta R
-  \delta(|\mathbf{r}-\mathbf{r}'|-R)
-\end{align}
-The vector density $\mathbf{n}_{V2}$ is similar to the density
-$n_2$: both changes in the density value and changes in the surface
-area play a role.
-\begin{align}
-  \Delta \mathbf{n}_{V2}(\mathbf{r}') &=
-  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
-  \left(n\left(\mathbf{r} + \Delta R
-  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
-  R}{R}\right)- n(\mathbf{r})\right) \Delta A
-  \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
-  &=
-  \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
-  \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
-    \right) \Delta A \Delta R
-  \delta(|\mathbf{r}-\mathbf{r}'|-R) 
-\end{align}
-The density $n_1$ (and similarly $\mathbf{n}_{V1}$) can be roughly
-interpreted as the density weighted by the area-averaged mean
-curvature~\cite{oversteegen2005general, roth2010review}.  When we
-extend the sphere at the contact point, we will introduce regions of
-both positive and negative mean curvature.  In principle, we can do
-this in such a way that the two portions cancel, and hence we will
-assume that
-\begin{align}
-  \Delta n_1(\mathbf{r}') &= 0 \\
-  \Delta \mathbf{n}_{V1}(\mathbf{r}') &= 0
-\end{align}
-In fact, the negative and positive portions are spatially separated,
-so these terms should probably include a derivative term, which I
-should work out.  Particularly as it may even happen to cancel a
-different derivative term that we then won't have to drop!
-
-We need to then find the change in Helmholtz free energy due to this
-change at position $\mathbf{r}$:
-\begin{align}
-  \Delta A_{HS} &= \int \mathbf{dr}'
-  \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}\Delta n_3(\mathbf{r}')
-  +
-  \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}\Delta n_2(\mathbf{r}')
-  +
-  \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} \cdot \Delta \mathbf{n}_{V2}(\mathbf{r}')
-\end{align}
-
-\end{widetext}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-We get a nice spatial dependence for SAFT.
-
-\begin{widetext}
-  \appendix
-
-  \section{Partial derivatives of the White Bear free energy functional}
-
-  Here is the White Bear free energy functional for the hard-sphere
-  fluid:~\cite{roth2002whitebear}
-  \begin{align}
-    A_{HS} &= \int \left\{
-    -n_0 \ln\left( 1 - n_3\right)
-    + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \right\}
-\end{align}
-
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-      n_3(\mathbf{r}'))^2}
-    + \frac{n_2^3 -
-      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-      \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
-    \right)
-\end{align}
-\end{widetext}
-
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &= -\ln(1-n_3)
-    \\
-    \beta\frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
-    \\
-    \beta\frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-      \frac{n_1}{1-n_3}
-      + (n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
-        (1-n_3)^2\ln(1-n_3)}{
-        12\pi n_3^2(1-n_3)^2
-      }
-\end{align}
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
-      \frac{\mathbf{n}_{V2}}{1-n_3}
-    \\
-    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-      \frac{\mathbf{n}_{V1}}{1-n_3}
-      - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
-        (1-n_3)^2\ln(1-n_3)}{
-        36\pi n_3^2(1-n_3)^2
-      }
-  \end{align}
+Either of these two  effects could be fudged using the other, but a
+correct description should include both effects.
 

### 2011-07-12 papers/contact/figs/sphere.cpp
add a few figures for contact density

@@ -1,0 +1,235 @@
+// Deft is a density functional package developed by the research
+// group of Professor David Roundy
+//
+// Copyright 2010 The Deft Authors
+//
+// Deft is free software; you can redistribute it and/or modify it
+// under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+//
+// You should have received a copy of the GNU General Public License
+// along with deft; if not, write to the Free Software Foundation,
+// Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
+//
+// Please see the file AUTHORS for a list of authors.
+
+#include <stdio.h>
+#include <time.h>
+#include "Functionals.h"
+#include "equation-of-state.h"
+#include "LineMinimizer.h"
+#include "ContactDensity.h"
+#include "utilities.h"
+#include "handymath.h"
+
+const double nm = 18.8972613;
+// Here we set up the lattice.
+double diameter = 6;
+double N = 4;
+bool using_default_diameter = true;
+
+double notinwall(Cartesian r) {
+  const double z = r.z();
+  const double y = r.y();
+  const double x = r.x();
+  if (sqrt(sqr(z)+sqr(y)+sqr(x)) < diameter/2) {
+      return 1; 
+  }
+  return 0;
+}
+
+static void took(const char *name) {
+  static clock_t last_time = clock();
+  clock_t t = clock();
+  assert(name); // so it'll count as being used...
+  printf("\t\t%s took %g seconds\n", name, (t-last_time)/double(CLOCKS_PER_SEC));
+  last_time = t;
+}
+
+void radial_plot(const char *fname, const Grid &a, const Grid &b, const Grid &c, const Grid &d) {
+  FILE *out = fopen(fname, "w");
+  if (!out) {
+    fprintf(stderr, "Unable to create file %s!\n", fname);
+    // don't just abort?
+    return;
+  }
+  const GridDescription gd = a.description();
+  const int x = 0;
+  const int z = 0;
+  for (int y=0; y<gd.Ny/2; y++) {
+    Cartesian here = gd.fineLat.toCartesian(Relative(x,y,z));
+    double ahere = a(x,y,z);
+    double bhere = b(x,y,z);
[…]

### 2011-07-12 paper/project.tex

@@
 
-\subsection{Water confined to a slit}
+We compute the density profile and surface tension by relaxing a slab
+of water in a periodic unit cell.  This slab has two surfaces, so the
+surface tension is just half the free energy of the cell, divided by
+its cross-sectional area.  As mentioned in
+section~\ref{sec:empirical}, the $b$ parameter is determined by
+setting this surface tension to match the experimental value.  The
+associated density profile is plotted in
+Figure~\ref{fig:cartesianDensity}, which displays two peaks in
+density.
 
@@
 
-\subsubsection{Two rods - hydrophobic interaction}
+\subsubsection{Two rods - interaction effects}
 

### 2011-07-13 paper/project.tex
Fixed low res version of single-rod and work on project text

--- 2011-07-07 paper/project.tex
@@
 
+One of the most computationally simple ways to study confined water is by
+constraining it to a slit where the water molecules can only move in one
+dimension. We apply a constraining potential of the form
+
+\begin{align}
+V = n_l*constraint + 100*n_v (FIXME)
+\end{align}
+
+such that the density is 100 times the vapor density outside the slit 
+and approximately equal to the saturated liquid density inside. The slits
+range in size from 5 bohrs to 150 bohrs (FIX - redo in nm?). We 
+minimize the energy and calculate density, energy, and $X$ on a lattice
+with a spacing of 0.2 bohrs between data points.
+
+The density profile for a slit of size 60 bohrs is shown in Figure 
+\ref{fig:density-1D}. Near the interface, there are oscillations around the 
+liquid density which decrease near the center of the slit. We expect the
+center of the slit to act approximately like bulk water, and the absence 
+of oscillations supports this. For larger slits (~120 bohrs or more) the
+majority of the water molecules in the slit act like the bulk. We also 
+determine the smallest slit that can contain liquid water, approximately 
+30 bohrs. For smaller slits, the initial liquid water inside will vaporize.
+
 \begin{figure}
@@
 \end{center}
-\caption{ Density profile for a slab of water with molecules constrained to move in one dimension. }
+\caption{ Density profile for a slab of water with molecules constrained 
+to move in one dimension. }
 \label{fig:density-1D}
@@
 \end{center}
-\caption{ Energy density profile for a slab of water with molecules constrained to move in one dimension. }
+\caption{ Energy density profile for a slab of water with molecules 
+constrained to move in one dimension. }
 \label{fig:energy-1D}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy-1D}
-\end{center}
-\caption{ Entropy for a slab of water with molecules constrained to move in one 
-dimension. \textcolor{red}{FIX UNITS}}
-\label{fig:entropy-1D}
-\end{figure}
+The value of $X$ behaves as expected for water constrained to a slit. We
+show in Figure \ref{fig:xassoc-1D} the plot of $X$ for a slit of 60 bohrs.
+Outside the slit, there is no hydrogen bonding, and inside $X$ is 
+approximately equal to
+that of saturated liquid water.
 
@@
 \caption{ Xassociation value (fraction of association sites
-\emph{not} hydrogen-bonded) for a slab of water with molecules constrained to move in one 
-dimension. }
+\emph{not} hydrogen-bonded) for a slab of water with molecules constrained 
+to move in one dimension. }
 \label{fig:xassoc-1D}
@@
 
+\subsubsection{One rod - diameter effects}
+
+We move into two dimensions by studying a single hydrophobic rod
+immersed in water.
+
 \subsubsection{Resolution and convergence tests}
 
-\subsubsection{One rod - diameter effects}
+(Plot with low (0.4 bohrs), regular (0.2), and high (0.05) resolution 
+for single rod.)
 

### 2011-07-14 paper/project.tex

@@
 \subsection{Association free energy}
-To the purely repulsive hard-sphere free energy, we add two attractive
-energy terms.  The first is the free energy of \emph{association},
-which describes the effects of hydrogen bonds.  These are modeled as
-four attractive patches (``association sites'') on the surface of the
-hard sphere.  These four sites represent the two hydrogens and two
-electron lone pairs.  There is an attractive energy
-$\epsilon_\textit{assoc}$ when two molecules are arranged such that
-the hydrogen of one interacts with the lone pair of the other.  The
-volume of this interaction is $\kappa_\textit{assoc}$.  Thus the
-association term in the free energy has two free parameters that must
-be fit based on experimental data.
+The first attractive energy term is the association term, which
+accounts for hydrogen bonding.  Hydrogen bonds are modeled as four
+attractive patches (``association sites'') on the surface of the hard
+sphere.  These four sites represent two protons and two electron lone
+pairs.  There is an attractive energy $\epsilon_\textit{assoc}$ when
+two molecules are arranged such that the proton of one overlaps
+with the lone pair of the other.  The volume over which this
+interaction occurs is $\kappa_\textit{assoc}$, giving the association
+term in the free energy has two empirical parameters that are fit to
+experimental data.
 
-The association free energy has the form
+The association functional we use is a modified version of that of Yu
+and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}.\footnote{We
+  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
+  the same functional, but their paper contains errors in the
+  association term and is not useful as a reference.}  The functional
+of Fu and We includes the effects of density inhomogeneities on the
+\emph{contact density} of hard spheres, which conventionally appears
+in the form of the density multiplied by the contact value of the
+correlation function $g^{HS}_\sigma$, but is based on the SAFT-HS
+equation of state\cite{FIXME}, rather than the SAFT-VR equation of
+state\cite{gil-villegas-1997-SAFT-VR} which is used in the optimal
+SAFT parametrization for water found by Clark \emph{et
+  al}\cite{clark2006developing}.
+
+The association free energy for our four-site model has the form
 \begin{align}
@@
 \end{align}
-where $\zeta$ comes from
-reference\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft} and describes the inhomogeneity of the fluid
-density:
+where the factor $\zeta$ from
+Yu and We\cite{yu2002fmt-dft-inhomogeneous-associating,
+  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
+density, and reduces to unity in the homogeneous limit.
 \begin{align}
@@
 \end{align}
-The function $\Delta$ is given... below?
-
-The following contact density for the hard-sphere fluid is in the code
-called \texttt{gHS}, and comes
-from\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft}, but I don't use it.
+The functional $\Delta$ is a measure of the hydrogen-bonding
+probability.  In the SAFT-VR model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
 \begin{align}
-  \Delta &= 4\pi \kappa_\textit{assoc} g^\textit{HS}(\sigma)e^{-\beta
-    \epsilon_\textit{assoc}} \\
-  g^\textit{HS}(\sigma) &= \frac1{1-n_3}\left(1+\frac14\frac{\zeta n_2}{1-n_3}
-  \left(1 + \frac{n_2}{14 (1-n_3)}\right)\right)
-\end{align}
-\textcolor{red}{Here is paper where I got the association functional:}
-
-Instead, I use
-\begin{align}
-  \Delta(\xx) &= \kappa_\textit{assoc}
-  e^{-\beta\epsilon_\textit{assoc}} g^\textit{SW}_\sigma(\xx) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
-  \frac{1}{4}\beta\left(\frac{\partial a_1(\xx)}{\partial n_3(\xx)} -
-  \frac{\lambda}{3 n3(\xx)}\frac{\partial a_1(\xx)}{\partial \lambda}\right)
+  \Delta &= \kappa_\textit{assoc} g^\textit{SW}_\sigma
+  \left(e^{-\beta\epsilon_\textit{assoc}} - 1\right) \\
+  g^\textit{SW}_\sigma &= g^\textit{HS}_\sigma +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial n_3} -
+  \frac{\lambda}{3 n3}\frac{\partial a_1}{\partial \lambda}\right)
 \end{align}
 % This is equation 77 of gil-villegas-1997-SAFT-VR
-which I got from reference\cite{gil-villegas-1997-SAFT-VR}.
+where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+contact for a hard-sphere fluid with a square-well attractive
+potential, and $a_1$ and $a_2$ are two terms in the dispersive term of
+the free energy.  The correlation function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere correlation
+function, for which we use the functional of Yu and
+Wu.\cite{yu2002fmt-dft-inhomogeneous-associating}
+\begin{align}
+  g_\sigma^{HS} &= \frac{1}{1-n_3}
+  +\frac32\frac{\zeta}{(1-n_3)^2}
+  + \frac12\frac{n_3^2\zeta}{(1-n_3)^3}
+\end{align}
 

### 2011-07-15 papers/contact/paper.tex
add a few figures for contact density
add a bunch more contact-density visualization and new c.d. functional

@@
 
+\section{Comparison with simulation}
+
+\subsection{Hard spheres confined in a spherical cavity}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/sphere-radial-06-04}
+  \caption{Density and contact densities of four hard spheres in a
+    spherical cavity with radius 6.}
+  \label{fig:sphere-density}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/sphere-radial-08-13}
+  \caption{Density and contact densities of fourteen hard spheres in a
+    spherical cavity with radius 8.}
+  \label{fig:sphere-density2}
+\end{figure}
+
+\subsection{Hard spheres confined in a cubical cavity}
+
+\begin{figure*}
+\includegraphics[height=5cm]{figs/box-100c--05,05,05-13}
+\includegraphics[height=5cm]{figs/box-100s--05,05,05-13}
+\includegraphics[height=5cm]{figs/box-110c--05,05,05-13}
+\caption{Density of hard spheres in a $5\times5\times5$ box.}
+\label{fig:box-density}
+\end{figure*}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
+\section{Derivatives of the fundamental-measure weighted densities}
+
+The derivative of the filling fraction is very easy:
+\begin{align}
+  n_3(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
+  \Theta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
+  \frac{\delta n_3(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= n(\mathbf{r'})\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+\end{align}
+
+The derivative of the density of surface is harder:
+\begin{align}
+  n_2(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
+  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
+  \frac{\delta n_2(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+\end{align}
+where the trouble here is that the derivative of the delta function,
+$\delta'$, isn't something we really like to deal with.  Particularly
+as we're looking  at a 3D integral involving a 1D delta function,
+which means there are Jacobian factors to be considered.  On the other
+hand, there's no reason why we couldn't convolve with this kernel as
+efficiently as we do with any other kernel.
+
+The derivatives oof the remaining scalar densities can be reduced to
+sums of the terms above:
+\begin{align}
+  n_1(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
+  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
+  \frac{\delta n_1(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= \frac{n(\mathbf{r'})}{4\pi
+    R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+  -
+  \frac{n(\mathbf{r'})}{4\pi
+    R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+\end{align}
+\begin{align}
+  n_0(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}
+  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
+  \frac{\delta n_0(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= \frac{n(\mathbf{r'})}{4\pi
+    R(\mathbf{r'})^2}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+  -2
+  \frac{n(\mathbf{r'})}{4\pi
+    R(\mathbf{r'})^3}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+\end{align}
+
+Since the vector-weighted densities have no additional occurrences of
+$R$, their derivatives come out similarly.
+\begin{align}
+  \mathbf{n}_{V2}(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
+  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
+  \frac{\delta \mathbf{n}_{V2}(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
+\end{align}
+and analogously for $\mathbf{n}_{V1}$.
+\begin{align}
+  \mathbf{n}_{V1}(\mathbf{r}) &= \int \mathbf{dr'}
+  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
+  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
+  \frac{\delta \mathbf{n}_{V1}(\mathbf{r})}{\delta R(\mathbf{r}')}
+  &= \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} -
+  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
+  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
+\end{align}
+
+So, what is this convolution with a $\delta'$? Let's look at the
+scalar version:
+\begin{align}
+  \bar{f}(\mathbf{r})
+  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
+  \\
+  &= \int
+        f(\mathbf{r'})
+        \lim_{\Delta R \rightarrow 0}
+          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+            -
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &= \lim_{\Delta R \rightarrow 0}
+          \frac{\int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+            -
+            \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &\approx\lim_{\Delta R \rightarrow 0}
+          \frac{\int
+            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+            -
+            \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &\approx\lim_{\Delta R \rightarrow 0}
+          \frac{2\frac{\Delta R}{R} \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &= \frac{2}{R} \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+\end{align}
+There is a sign ambiguity here with regard to whether we are asking
+about the derivative of this delta function with regard to $R$ or
+something else.  We just need to get it right.
+
 \begin{widetext}

### 2011-07-22 paper/project.tex
Changed bohrs to nm in several 1D slit files and in project

@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy}
-\end{center}
-\caption{The free energy, internal energy, and temperature*entropy versus density at room temperature.  }
-\label{fig:energy-room-temp}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy-at-690K}
-\end{center}
-\caption{The free energy, internal energy, and temperature*entropy versus density near the critical temperature.  }
-\label{fig:energy-near-critical-temp}
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/entropy}
+% \end{center}
+% \caption{The free energy, internal energy, and temperature*entropy versus density at room temperature.  }
+% \label{fig:energy-room-temp}
+% \end{figure}
+% 
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/entropy-at-690K}
+% \end{center}
+% \caption{The free energy, internal energy, and temperature*entropy versus density near the critical temperature.  }
+% \label{fig:energy-near-critical-temp}
+% \end{figure}
 
@@
 
+The fluid density at a liquid-solid interface commonly displays
+oscillatory behavior, but the presence of density oscillations at
+liquid-vapor interfaces is more debated\cite{penfold2001structure}.
+In fact, the \emph{meaning} of the density profile at a liquid-vapor
+interface is clouded by the existence of long-wavelength perturbations
+known as \emph{capillary waves}.  These waves are not accounted for in
+density-functional theory (although the \emph{exact}
+density-functionaly theory \emph{would} include them), with the result
+that the density profile will be smoothed out by CW effects.
+
+\textcolor{red}{I found a nice paper on the hard-slit
+  problem... actually a couple of papers that address
+  it~\cite{forsman1996computer, lum1999hydrophobicity}}
+
 \subsection{Water confined to a slit}
@@
 constraining it to a slit where the water molecules can only move in one
-dimension. We apply a constraining potential of the form
+dimension. Experimentally, this could represent two different situations.
+The first, water between two perfectly aligned and smooth planes, is not common 
+in experiment due to the difficulty of creating such an ideal situation. The second
+possibility is a large sphere brought close to a plane, which eliminates the 
+need for perfect alignment.
 
-\begin{align}
-V = n_l*constraint + 100*n_v (FIXME)
-\end{align}
+We choose a chemical potential corresponding to a pressure of 1 atm and we apply 
+a constraining potential where the density is equal to the saturated 
+liquid density inside the slit. Outside the slit, we choose a small non-zero
+density for numerical reasons. The slits studied range in size from 0.1 nm to 
+8.0 nm. We minimize the energy and calculate density, energy, and $X$ on a lattice
+with a spacing of approximately 0.01 nm (0.2 bohrs) between lattice points.
 
-such that the density is 100 times the vapor density outside the slit 
-and approximately equal to the saturated liquid density inside. The slits
-range in size from 5 bohrs to 150 bohrs (FIX - redo in nm?). We 
-minimize the energy and calculate density, energy, and $X$ on a lattice
-with a spacing of 0.2 bohrs between data points.
-
-The density profile for a slit of size 60 bohrs is shown in Figure 
+The density profile for a slit of size 4.0 nm is shown in Figure 
 \ref{fig:density-1D}. Near the interface, there are oscillations around the 
@@
 center of the slit to act approximately like bulk water, and the absence 
-of oscillations supports this. For larger slits (~120 bohrs or more) the
+of oscillations supports this. For larger slits (6.0 nm or more) the
 majority of the water molecules in the slit act like the bulk. We also 
 determine the smallest slit that can contain liquid water, approximately 
-30 bohrs. For smaller slits, the initial liquid water inside will vaporize.
+0.6 nm (\textcolor{red}{Check this again}. For smaller slits, the initial 
+liquid water inside will vaporize.
 
@@
 \end{center}
-\caption{ Energy density profile for a slab of water with molecules 
-constrained to move in one dimension. }
+\caption{ Energy densities for a slab of water with molecules 
+constrained to move in one dimension. \textcolor{red}{FIXME}}
 \label{fig:energy-1D}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
+\end{center}
+\caption{ Free energy versus slit size for a slab of water with molecules 
+constrained to move in one dimension. \textcolor{red}{Not sure if this needs
+to have a bulk energy amount removed to be accurate.}}
+\label{fig:cavitysize-vs-energy-1D}
+\end{figure}
+
 The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure \ref{fig:xassoc-1D} the plot of $X$ for a slit of 60 bohrs.
-Outside the slit, there is no hydrogen bonding, and inside $X$ is 
-approximately equal to
-that of saturated liquid water.
+show in Figure \ref{fig:xassoc-1D} the plot of $X$ for a slit of 4.0 nm.
+Outside the slit, there is no hydrogen bonding ($X$ = 1), and inside $X$ is 
+approximately equal to 0.13, or that of saturated liquid water.
 
@@
 
-The fluid density at a liquid-solid interface commonly displays
-oscillatory behavior, but the presence of density oscillations at
-liquid-vapor interfaces is more debated\cite{penfold2001structure}.
-In fact, the \emph{meaning} of the density profile at a liquid-vapor
-interface is clouded by the existence of long-wavelength perturbations
-known as \emph{capillary waves}.  These waves are not accounted for in
-density-functional theory (although the \emph{exact}
-density-functionaly theory \emph{would} include them), with the result
-that the density profile will be smoothed out by CW effects.
-
-\textcolor{red}{I found a nice paper on the hard-slit
-  problem... actually a couple of papers that address
-  it~\cite{forsman1996computer, lum1999hydrophobicity}}
-
 \subsection{Hydrophobic rods}
@@
 immersed in water.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+\end{center}
+\caption{ Energy versus diameter for a single hydrophobic rod
+immersed in water. This should have an asymptote equal to the surface
+tension. (FIXME - Data points are not correct yet) }
+\label{fig:energy-vs-diameter}
+\end{figure}
 

### 2011-08-14 paper/project.tex
Some intro text added to project.tex

@@
 \subsection{Water---the universal solvent}
+
+The simple yet complex water molecule is one of the most important 
+substances for life and industry. The structure and geometry of the 
+water molecule leads to a strong polarity and complex intermolecular
+forces. The hydrogen bonds formed by water are very strong, creating
+tetragonal clusters (CITE?) that lead to many of water's unusual properties,
+such as the comparatively high melting, boiling, and critical temperatures.
+The large dielectric constant (D~80) (CITE) is partially responsible for
+water as a universal solvent, along with the ease of availability and use.
+
+It is challenging to describe all the relevant properties of water using
+a model, especially if a low computational cost is also desired.
+
+(Previous/popular models of water summary/review??)
 

### 2011-08-15 papers/contact/paper.tex
add test for convolution of convolution
Tested-on: bingley (Linux 2.6.32-5-amd64)
add a few more tests and plot a larger sphere

@@
   \caption{Density and contact densities of fourteen hard spheres in a
-    spherical cavity with radius 8.}
+    spherical cavity with radius 8.  Also similar densities of spheres
+    in even larger cavities.}
   \label{fig:sphere-density2}
@@
 
-The derivatives oof the remaining scalar densities can be reduced to
+The derivatives of the remaining scalar densities can be reduced to
 sums of the terms above:

### 2011-08-23 paper/project.tex
Edits to some energy plots for 1D slit and metastability discussion

@@
 \author{Jessica Hughes}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR
+97331}
 
 \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR
+97331}
 
@@
   water near liquid-vapor and liquid-solid interfaces including water confined 
-  to a slit and two hydrophobic rods immersed in water.  This functional will be the 
+  to a slit and hydrophobic rods immersed in water.  This functional will be
+  the 
   foundation for a efficient method for future work
@@
 such as the comparatively high melting, boiling, and critical temperatures.
-The large dielectric constant (D~80) (CITE) is partially responsible for
+The large dielectric constant (D$\sim$80) (CITE) is partially responsible for
 water as a universal solvent, along with the ease of availability and use.
@@
 
-(Previous/popular models of water summary/review??)
+(Types and models of water summary/review)
 
@@
 \begin{equation}
-  F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
+  F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] +
+F_\textit{disp},
 \end{equation}
@@
 The functional $\Delta$ is a measure of the hydrogen-bonding
-probability.  In the SAFT-VR model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
+probability.  In the SAFT-VR model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is
+given by
 \begin{align}
@@
 % \end{center}
-% \caption{The free energy, internal energy, and temperature*entropy versus density at room temperature.  }
+% \caption{The free energy, internal energy, and temperature*entropy 
+% versus density at room temperature.  }
 % \label{fig:energy-room-temp}
-% \end{figure}
-% 
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/entropy-at-690K}
-% \end{center}
-% \caption{The free energy, internal energy, and temperature*entropy versus density near the critical temperature.  }
-% \label{fig:energy-near-critical-temp}
 % \end{figure}
@@
 \end{center}
-\caption{The pressure versus density for various temperatures, including experimental pressure data from NIST.  }
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST.  }
 \label{fig:pressure-with-isotherms}
@@
 \begin{center}
+\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
+\includegraphics[width=\columnwidth]{figs/near-critical-point}
+\end{center}
+\caption{By plotting the free energy vs. density it is possible to
+  determine a saturated liquid or vapor state by the common tangent
+  method.  }
+\label{fig:homogeneous}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=\columnwidth]{figs/surface-tension}
@@
 
-\subsection{Determining the empirical parameters (?)}\label{sec:empirical}
+\subsection{Determining the empirical parameters}\label{sec:empirical}
 
@@
 as illustrated in Figure~\ref{fig:homogeneous}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
-\includegraphics[width=\columnwidth]{figs/near-critical-point}
-\end{center}
-\caption{By plotting the free energy vs. density it is possible to
-  determine a saturated liquid or vapor state by the common tangent
-  method.  }
-\label{fig:homogeneous}
-\end{figure}
 
@@
 density-functional theory (although the \emph{exact}
-density-functionaly theory \emph{would} include them), with the result
+density-functional theory \emph{would} include them), with the result
 that the density profile will be smoothed out by CW effects.
 
-\textcolor{red}{I found a nice paper on the hard-slit
-  problem... actually a couple of papers that address
-  it~\cite{forsman1996computer, lum1999hydrophobicity}}
-
 \subsection{Water confined to a slit}
-
-One of the most computationally simple ways to study confined water is by
-constraining it to a slit where the water molecules can only move in one
-dimension. Experimentally, this could represent two different situations.
-The first, water between two perfectly aligned and smooth planes, is not common 
-in experiment due to the difficulty of creating such an ideal situation. The second
-possibility is a large sphere brought close to a plane, which eliminates the 
-need for perfect alignment.
-
-We choose a chemical potential corresponding to a pressure of 1 atm and we apply 
-a constraining potential where the density is equal to the saturated 
-liquid density inside the slit. Outside the slit, we choose a small non-zero
-density for numerical reasons. The slits studied range in size from 0.1 nm to 
-8.0 nm. We minimize the energy and calculate density, energy, and $X$ on a lattice
-with a spacing of approximately 0.01 nm (0.2 bohrs) between lattice points.
-
-The density profile for a slit of size 4.0 nm is shown in Figure 
-\ref{fig:density-1D}. Near the interface, there are oscillations around the 
-liquid density which decrease near the center of the slit. We expect the
-center of the slit to act approximately like bulk water, and the absence 
-of oscillations supports this. For larger slits (6.0 nm or more) the
-majority of the water molecules in the slit act like the bulk. We also 
-determine the smallest slit that can contain liquid water, approximately 
-0.6 nm (\textcolor{red}{Check this again}. For smaller slits, the initial 
-liquid water inside will vaporize.
 
@@
 
+One of the most computationally simple ways to study confined water is by
+constraining it to a slit where the water molecules can only move in one
+dimension. Experimentally, this could represent two different situations.
+The first, water between two perfectly aligned and smooth planes, is not common 
+in experiment due to the difficulty of creating such an ideal situation. The
+second
+possibility is a large sphere brought close to a plane, which eliminates the 
+need for perfect alignment.
+
+We choose a chemical potential corresponding to a pressure of 1 atm and set the
+temperature to a typical room temperature of 298 K. We apply 
+a constraining potential outside the slit, where we choose a small non-zero
+density for numerical reasons. The slits studied range in size from 0.1 nm to 
+8.0 nm. We minimize the energy and calculate density, energy, and $X$ on a
+lattice
+with a spacing of approximately 0.01 nm (0.2 bohrs) between lattice points.
+
+The density profile for a slit of size 4.0 nm is shown in Figure
+\ref{fig:density-1D}. Near the interface, there are oscillations around the
+liquid density which decrease near the center of the slit. We expect the center
+of the slit to act approximately like bulk water, and the absence of
+oscillations supports this. For larger slits (6.0 nm or more) the majority of
+the water molecules in the slit act like the bulk. Based on our computations, we
+also determine the smallest slit that can contain liquid water. This result is
+approximately 0.6 nm. For smaller slits, the initial liquid water inside will
+vaporize. The results for larger liquid-filled slits are only metastable,
+however,as a vapor-filled slit is thermodynamically favorable. Based on our
+calculations for a pressure of 1 atm, the work needed to pull apart
+the two hard walls comprising the slit is one to two orders of magnitude less
+thanthe free energy of the liquid-filled slit. For the range of slit sizes we
+studied (0.1 to 8.0 nm), all the computational results of liquid-filled are
+metastable. Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
+stability of water between large hydrophobic plates and the critical separation
+necessary for stability. For water at room temperature and atmospheric pressure,
+plates with a separation smallerthan about 100 nm will not contain stable liquid
+water. They go on to say that confined water has a region of metastability down
+to separations of about 5 nm. This is comparable to the slit separation where
+our calculations show water inside the slit acting like the bulk liquid, which
+we know to be a metastable solution. Forsman et al.\cite{forsman1996computer}
+also study water between two hard walls at constant chemical potential. They 
+also conclude that the liquid state is likely metastable, and there is a free
+energy barrier to reach the stable vapor phase.
+
 \begin{figure}
@@
 \caption{ Energy densities for a slab of water with molecules 
-constrained to move in one dimension. \textcolor{red}{FIXME}}
+constrained to move in one dimension. The free energy, internal energy
+and entropic energy densities are shown for a slit size of 4.0 nm.}
 \label{fig:energy-1D}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
-\end{center}
-\caption{ Free energy versus slit size for a slab of water with molecules 
-constrained to move in one dimension. \textcolor{red}{Not sure if this needs
-to have a bulk energy amount removed to be accurate.}}
-\label{fig:cavitysize-vs-energy-1D}
 \end{figure}
@@
 \label{fig:xassoc-1D}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
+\end{center}
+\caption{ Energy per length versus slit size for a slab of water with molecules 
+constrained to move in one dimension. The last few data points are likely inaccurate
+due to the slit separation being very close to the size of the system. \textcolor{red}{
+Not sure if this needs to have a bulk energy amount removed to be accurate.}}
+\label{fig:cavitysize-vs-energy-1D}
 \end{figure}

### 2011-08-25 paper/project.tex
Added enthalpy calculation, changes to figures, single rod resolutions

@@
 tetragonal clusters (CITE?) that lead to many of water's unusual properties,
-such as the comparatively high melting, boiling, and critical temperatures.
+such as the high melting, boiling, and critical temperatures.
 The large dielectric constant (D$\sim$80) (CITE) is partially responsible for
 water as a universal solvent, along with the ease of availability and use.
+
+\subsection{Classical density-functional theory}
 
@@
 
-(Types and models of water summary/review)
+Some models, including SAFT, are designed for use with
+homogeneous systems only. Although these models are useful, much of the interesting
+chemistry that occurs in water cannot be described with a homogeneous only model. To
+describe solvation and other processes in aqueous solution, a more robust theory is 
+needed.
 
-\subsection{Classical density-functional theory}
+Classical density-functional theory (DFT) provides an efficient way to evaluate the
+free energy using the density of the fluid. The free energy is a functional of the 
+density function and is independent of any external potentials\cite{ebner1976}. DFT 
+can be used with a variety of external potentials, from the simple interaction with a 
+hydrophobic wall to the complex strong interaction with hydrophilic solutes.
 
@@
 We introduce a new classical density functional for water that
-\needsworklater{predicts the vapor density, liquid density, bulk
+predicts the \needsworklater{vapor density, liquid density, bulk
   modulus, and bulk surface tension, as well as allows for
@@
 \end{center}
-\caption{The theoretical versus experimental saturated liquid density
-  versus temperature.  }
+\caption{The theoretical versus experimental temperature versus density. }
 \label{fig:saturated-liquid-density}
@@
 \caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST.  }
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
 \label{fig:pressure-with-isotherms}
 \end{figure}
+
+Figure \ref{fig:pressure-with-isotherms} shows the calculated pressures for
+isotherms ranging from room temperature to past the critical temperature. The solid 
+line is the theoretical coexistence curve calculated from the SAFT functional. 
+For a given isotherm, the pressure curve should intersect this coexistence curve 
+twice, at the values of the vapor and liquid densities. The functional matches well with
+the experimental data from NIST away from the critical point. The slope of the pressure
+versus density curve is proportional to the compressibility, which is close but slightly too
+large at room temperature.
 
@@
 \begin{align}
-  H[n] &= F[n] + TS[n] - pV
+  H[n] &= F[n] + TS[n] + pV
 \end{align}
+
+The calculated heat of vaporization at atmospheric pressure is
+$\Delta H_{vap}$= 39.41 kJ/mol at the normal boiling point of water,
+373 K. This compares very well to the experimental
+value measured by NIST ($\Delta H_{vap}$= 40.65 kJ/mol\cite{nistwater}), with 
+an error of only a few percent.
 
@@
 \section{Results and discussion}
-\subsection{Liquid-vapor interface}
+
+\subsection{Water confined to a slit}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-1D}
+\includegraphics[width=\columnwidth]{figs/energy-1D}
+\end{center}
+\caption{ Density profile (top) and 
+energy densities (bottom) for a slab of water with molecules 
+constrained to move in one dimension. The free energy, internal energy
+and entropic energy densities are shown for a slit size of 4.0 nm. The dashed
+line in the top figure represents the saturated liquid density.}
+\label{fig:energy-1D}
+\end{figure}
 
@@
 density-functional theory \emph{would} include them), with the result
-that the density profile will be smoothed out by CW effects.
-
-\subsection{Water confined to a slit}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D}
-\end{center}
-\caption{ Density profile for a slab of water with molecules constrained 
-to move in one dimension. }
-\label{fig:density-1D}
-\end{figure}
+that the density profile will be smoothed out by CW effects. The liquid-vapor
+interface is plotted in Figure \ref{fig:liquid-vapor-profile} and shows
+some small oscillations.
 
@@
 The density profile for a slit of size 4.0 nm is shown in Figure
-\ref{fig:density-1D}. Near the interface, there are oscillations around the
+\ref{fig:energy-1D}. Near the interface, there are oscillations around the
 liquid density which decrease near the center of the slit. We expect the center
@@
 the two hard walls comprising the slit is one to two orders of magnitude less
-thanthe free energy of the liquid-filled slit. For the range of slit sizes we
+than the free energy of the liquid-filled slit. For the range of slit sizes we
 studied (0.1 to 8.0 nm), all the computational results of liquid-filled are
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-1D}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
 \end{center}
-\caption{ Energy densities for a slab of water with molecules 
-constrained to move in one dimension. The free energy, internal energy
-and entropic energy densities are shown for a slit size of 4.0 nm.}
-\label{fig:energy-1D}
+\caption{The number of hydrogen bonds per molecule for a slab of water 
+with molecules constrained to move in one dimension. Each molecule may contribute to
+four bonds, although each bond involves two molecules so to calculate the actual 
+number of bonds there would be another factor of 2.}
+\label{fig:xassoc-1D}
 \end{figure}
@@
 The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure \ref{fig:xassoc-1D} the plot of $X$ for a slit of 4.0 nm.
-Outside the slit, there is no hydrogen bonding ($X$ = 1), and inside $X$ is 
-approximately equal to 0.13, or that of saturated liquid water.
+show in Figure \ref{fig:xassoc-1D} the number of hydrogen bonds per water 
+molecule, equal to $4(1-X)$, for a slit of 4.0 nm.
+Outside the slit, there is no hydrogen bonding and inside the number of hydrogen
+bonds is approximately equal to 3.5, which is expected for saturated liquid water.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D}
-\end{center}
-\caption{ Xassociation value (fraction of association sites
-\emph{not} hydrogen-bonded) for a slab of water with molecules constrained 
-to move in one dimension. }
-\label{fig:xassoc-1D}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
-\end{center}
-\caption{ Energy per length versus slit size for a slab of water with molecules 
-constrained to move in one dimension. The last few data points are likely inaccurate
-due to the slit separation being very close to the size of the system. \textcolor{red}{
-Not sure if this needs to have a bulk energy amount removed to be accurate.}}
-\label{fig:cavitysize-vs-energy-1D}
-\end{figure}
-
-\subsection{Hydrophobic rods}
-
-\subsubsection{One rod - diameter effects}
+\subsection{One hydrophobic rod}
 
@@
 immersed in water. This should have an asymptote equal to the surface
-tension. (FIXME - Data points are not correct yet) }
+tension. (FIX - 0.4 nm?) }
 \label{fig:energy-vs-diameter}
@@
 
-(Plot with low (0.4 bohrs), regular (0.2), and high (0.05) resolution 
-for single rod.)
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-vs-radius}
+\end{center}
+\caption{ Density profiles for 
+a single hydrophobic rod with a radius of 0.5 nm using several different 
+lattice spacing resolutions. high = 0.05 bohrs, 
+med = 0.2 bohrs, low = 0.5 bohrs }
+\label{fig:densityresolution}
+\end{figure}
 
-\subsubsection{Two rods - hydrophobic interaction}
+\subsection{Hydrophobic interaction of two rods}
 
@@
 
-\subsection{Hard Sphere}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
+
+[conclusion text]
 

### 2011-08-26 paper/project.tex
revamp empirical parameters section
Tested-on: bingley (Linux 2.6.32-5-amd64)

--- 2011-08-23 paper/project.tex
@@
 
-In order to ensure that the empirical parameters are correct, we test
-the vapor pressure (see Figure~\ref{fig:equation-of-state}) and
+The majority of the empirical parameters used in our functional are
+taken from the paper of Clark \emph{et al} on developing an optimal
+SAFT model for water\cite{clark2006developing}.  This SAFT model
+contains five empirical parameters: the hard-sphere radius, an energy
+and length scale for the dispersion interaction, and an energy and
+length scale for the association interaction.  In addition to the five
+empirical parameters of Clark \emph{et al}, we add a single additional
+parameter defining the length scale over which the density is averaged
+when computing the dispersion free energy.  We use this final
+parameter to fit the surface tension, with the result shown in
+Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
+\emph{et al} in overestimates the critical temperature---which is a
+common feature of models which do not explicitly treat the critical
+point---we cannot correctly describe the surface tension at all
+temperatures, and chose to fit it for the temperature range at which
+water is liquid at one atmosphere of pressure.
+
+\fixme{I'm not sure if this paragraph belongs here.  Somewhere I think
+  it's worth discussing any of these homogeneous plots we include, and
+  this might be the right place.  We also might want to include one or
+  more of them earlier in the text.  A general rule is that any figure
+  that is included should be referenced in the text.  If there's
+  nowhere in the text that it fits, then it probably doesn't belong in
+  the paper....  So I'm not removing the follwoing bits since it's up
+  to you, Jess, whether to remove the figures.  We should
+  \emph{definitely} remove some of them.}  For reference, we plot the
+predicted vapor pressure (see Figure~\ref{fig:equation-of-state}) and
 saturated liquid density (see

### 2011-08-26 paper/project.tex
A few edits to single rod and project

--- 2011-08-25 paper/project.tex
@@
   this approximation, we are able to describe various properties of
-  water near liquid-vapor and liquid-solid interfaces including water confined 
-  to a slit and hydrophobic rods immersed in water.  This functional will be
-  the 
-  foundation for a efficient method for future work
+  water near liquid-vapor and liquid-solid interfaces, including water confined 
+  to a hard-walled slit and hydrophobic rods immersed in water.  This functional 
+  will be the foundation for a efficient method for future work
   predicting the interactions of chemicals in aqueous solution on an
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\end{center}
+\caption{ Energy versus separation for two identical hydrophobic rods with diameters
+of 1 nm. }
+\label{fig:rods-energy-vs-distance}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2011-08-29 paper/project.tex
Comparing resolutions, improving resolution density plot

@@
 Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
-\emph{et al} in overestimates the critical temperature---which is a
+\emph{et al} overestimates the critical temperature---which is a
 common feature of models which do not explicitly treat the critical
@@
 
+Figure \ref{fig:densityresolution} shows the density profiles for three different
+values of the lattice spacing resolution. The density oscillations outside a single
+rod of radius 0.5 nm look nearly identical for all resolutions except for the first peak. 
+This contact density is important to get correct as it \fixme{relates to energy}. The 
+contact density calculated using the medium resolution is within 3.5\% of the value at
+high resolution, while the low resolution contact density is only within 5.1\%. The total
+free energy calculations compare much more favorably. The medium resolution is within 
+0.5\% and the low resolution is within 1.0\% of the high resolution free energy.
+
 \begin{figure}
@@
 a single hydrophobic rod with a radius of 0.5 nm using several different 
-lattice spacing resolutions. high = 0.05 bohrs, 
-med = 0.2 bohrs, low = 0.5 bohrs }
+lattice spacing resolutions. We used a high resolution of 0.05 bohrs ($\approx$ 0.003 nm), 
+medium 0.2 bohrs ($\approx$ 0.001 nm), and low 0.5 bohrs ($\approx$ 0.03 nm).}
 \label{fig:densityresolution}

### 2011-08-29 paper/project.tex
Edits to figures for 1 and 2 rods, some intro work

@@
 free energy using the density of the fluid. The free energy is a functional of the 
-density function and is independent of any external potentials\cite{ebner1976}. DFT 
-can be used with a variety of external potentials, from the simple interaction with a 
+density function and is independent of any external potentials\cite{ebner1976}. 
+Mermin\cite{Mermin} gives a theorem extending that of Hohenberg and Kohn\cite{HK} 
+to nonzero temperatures:
+\begin{equation}
+  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r) d^3r\right\}
+\end{equation}
+This is the basis for classical DFT, which can be used with a variety of external potentials, 
+from the simple interaction with a 
 hydrophobic wall to the complex strong interaction with hydrophilic solutes.
@@
 We move into two dimensions by studying a single hydrophobic rod
-immersed in water.
+immersed in water. Figure \ref{fig:density-single-rod} shows density
+profiles for three different radii, 0.3 nm, 0.5 nm, and 1 nm. There are the
+expected oscillations outside the rod and as the distance away from the rod increases
+the density approaches that of the bulk saturated liquid density. The small radius
+rod has comparatively larger oscillations and a higher contact density.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\end{center}
+\caption{ Density profiles for a single rod of different radii. The dotted line
+represents the saturated liquid density.  }
+\label{fig:density-single-rod}
+\end{figure}
 
@@
 Figure \ref{fig:densityresolution} shows the density profiles for three different
-values of the lattice spacing resolution. The density oscillations outside a single
+values of the lattice spacing resolution. The density oscillations around a single
 rod of radius 0.5 nm look nearly identical for all resolutions except for the first peak. 
 This contact density is important to get correct as it \fixme{relates to energy}. The 
-contact density calculated using the medium resolution is within 3.5\% of the value at
-high resolution, while the low resolution contact density is only within 5.1\%. The total
+contact density calculated using the medium resolution is within less than 4\% of the value at
+high resolution, while the low resolution contact density is off by about 5\%. There is some
+uncertainty in the value of contact density due to the different lattice spacings. The total
 free energy calculations compare much more favorably. The medium resolution is within 
-0.5\% and the low resolution is within 1.0\% of the high resolution free energy.
+0.5\% and the low resolution is within 1\% of the high resolution free energy.
 
@@
 \end{center}
-\caption{ Density profile for two hydrophobic rods.}
+\caption{ Density profile for two hydrophobic rods (diameter 1.0 nm) at a 
+separation of 0.8 nm. The scale on the right is density in g/mL.}
 \label{fig:density-rods}
@@
 \caption{ Energy versus separation for two identical hydrophobic rods with diameters
-of 1 nm. }
+of 1.0 nm. }
 \label{fig:rods-energy-vs-distance}

### 2011-08-30 paper/project.tex
Changing units etc in figures, fixes to rods, text

@@
 density function and is independent of any external potentials\cite{ebner1976}. 
-Mermin\cite{Mermin} gives a theorem extending that of Hohenberg and Kohn\cite{HK} 
-to nonzero temperatures:
+Mermin\cite{mermin1965thermal} gives a theorem extending that of Hohenberg and 
+Kohn\cite{hohenberg1964inhomogeneous} to nonzero temperatures:
 \begin{equation}
@@
 \begin{equation}
-  F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] +
-F_\textit{disp},
+  F[n] = F_\textit{id}[n] + F_\textit{hs}[n]  +
+F_\textit{disp}[n]+ F_\textit{assoc}[n],
 \end{equation}
@@
 surface $n_c$ is proportional to the pressure on that surface,
-according to the equation
-\begin{equation}
+according to the contact value theorem
+\begin{equation}\label{contactvaluethm}
   p = n_c k_BT \:,
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/saturated-liquid-density}
-\end{center}
-\caption{The theoretical versus experimental saturated liquid density
-  versus temperature.  }
-\label{fig:saturated-liquid-density}
-\end{figure}
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/saturated-liquid-density}
+% \end{center}
+% \caption{The theoretical versus experimental saturated liquid density
+%   versus temperature.  }
+% \label{fig:saturated-liquid-density}
+% \end{figure}
 
@@
 \caption{The theoretical versus experimental temperature versus density. }
-\label{fig:saturated-liquid-density}
+\label{fig:temperature-vs-density}
 \end{figure}
@@
 $\Delta H_{vap}$= 39.41 kJ/mol at the normal boiling point of water,
-373 K. This compares very well to the experimental
+373 K. This compares well to the experimental
 value measured by NIST ($\Delta H_{vap}$= 40.65 kJ/mol\cite{nistwater}), with 
@@
 \caption{The theoretical versus experimental surface tension
-  versus temperature.  }
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater} }
 \label{fig:surface-tension}
@@
 predicted vapor pressure (see Figure~\ref{fig:equation-of-state}) and
-saturated liquid density (see
-Figure~\ref{fig:saturated-liquid-density}) against experimental
-values.  To find these densities, we use the ``common tangent'' rule,
+coexistence curve (see
+Figure~\ref{fig:temperature-vs-density}) against experimental
+values.  To find the densities, we use the ``common tangent'' rule,
 as illustrated in Figure~\ref{fig:homogeneous}.
 
-\begin{table}
-\begin{tabular}{ccc}
-  Fitting Parameters& \hspace{2em} & Experimental Values\\
-\begin{tabular}{cc}
-  $R$ & 2.3 $a_o$ \\
-  $A$ & -0.0100909 Hartree \\
-  $\gamma$ & 5\\
-  $V$ & 515.792 $a_o^3$  \\
-  $b$ & 0.75 $a_o$ \\
-  $\mu$ & -0.014063 Hartree 
-\end{tabular}
-&& 
-\begin{tabular}{cc}
- $n_l$ & 4.939 $\times 10^{-3} a_o^{-3}$\\
- $n_v$ & 1.141 $\times 10^{-7} a_o^{-3}$\\
- $\sigma$ & ??? \\
- $B$ & $7.435\times 10^{-5}$ Hartree$/a_o^3$\\
- $T$ & 298.15 K \\
-\end{tabular}
-\end{tabular}
-
-\caption{Empirical parameters of optimized functional and experimental
-  values matched, where $a_o$ is the Bohr radius, and $\sigma$ is the
-  surface tension.  This table is not yet complete, but it should
-  be soon...  We'll also need to think about the best units to use
-  here.  Atom units might be best, but it also might be better to
-  consistently use eV and angstrom or something.  I'm not sure about
-  putting $k_BT$ data in the table.  Maybe it belongs in the caption?}
-\label{tab:parameters}
-\end{table}
+% \begin{table}
+% \begin{tabular}{ccc}
+%   Fitting Parameters& \hspace{2em} & Experimental Values\\
+% \begin{tabular}{cc}
+%   $R$ & 2.3 $a_o$ \\
+%   $A$ & -0.0100909 Hartree \\
+%   $\gamma$ & 5\\
+%   $V$ & 515.792 $a_o^3$  \\
+%   $b$ & 0.75 $a_o$ \\
+%   $\mu$ & -0.014063 Hartree 
+% \end{tabular}
+% && 
+% \begin{tabular}{cc}
+%  $n_l$ & 4.939 $\times 10^{-3} a_o^{-3}$\\
+%  $n_v$ & 1.141 $\times 10^{-7} a_o^{-3}$\\
+%  $\sigma$ & ??? \\
+%  $B$ & $7.435\times 10^{-5}$ Hartree$/a_o^3$\\
+%  $T$ & 298.15 K \\
+% \end{tabular}
+% \end{tabular}
+% 
+% \caption{Empirical parameters of optimized functional and experimental
+%   values matched, where $a_o$ is the Bohr radius, and $\sigma$ is the
+%   surface tension.  This table is not yet complete, but it should
+%   be soon...  We'll also need to think about the best units to use
+%   here.  Atom units might be best, but it also might be better to
+%   consistently use eV and angstrom or something.  I'm not sure about
+%   putting $k_BT$ data in the table.  Maybe it belongs in the caption?}
+% \label{tab:parameters}
+% \end{table}
 
@@
 immersed in water. This should have an asymptote equal to the surface
-tension. (FIX - 0.4 nm?) }
+tension at room temperature, and it agrees well with the surface tension in Figure 
+\ref{fig:surface-tension}. (FIX - 0.4 nm?) }
 \label{fig:energy-vs-diameter}
@@
 rod of radius 0.5 nm look nearly identical for all resolutions except for the first peak. 
-This contact density is important to get correct as it \fixme{relates to energy}. The 
+This contact density is important as it will be proportional to the pressure on the 
+rod\ref{contactvaluethm}. The 
 contact density calculated using the medium resolution is within less than 4\% of the value at
@@
 
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+% \end{center}
+% \caption{ Density profile for two hydrophobic rods (diameter 1.0 nm) at a 
+% separation of 0.8 nm. The scale on the right is density in g/mL.}
+% \label{fig:density-rods}
+% \end{figure}
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\includegraphics[width=\columnwidth]{figs/rods-diagram}
 \end{center}
-\caption{ Density profile for two hydrophobic rods (diameter 1.0 nm) at a 
-separation of 0.8 nm. The scale on the right is density in g/mL.}
-\label{fig:density-rods}
+\caption{ Cross section schematic for two hydrophobic rods immersed in water.
+The radius and separation are shown.}
+\label{fig:rods}
 \end{figure}

### 2011-09-01 paper/project.tex
Single rod updated data files, edits to figures and text

@@
   water near liquid-vapor and liquid-solid interfaces, including water confined 
-  to a hard-walled slit and hydrophobic rods immersed in water.  This functional 
+  to a hard-walled slit and hydrophobic rods immersed in water.  This
+functional 
   will be the foundation for a efficient method for future work
@@
 forces. The hydrogen bonds formed by water are very strong, creating
-tetragonal clusters (CITE?) that lead to many of water's unusual properties,
-such as the high melting, boiling, and critical temperatures.
-The large dielectric constant (D$\sim$80) (CITE) is partially responsible for
-water as a universal solvent, along with the ease of availability and use.
+tetragonal clusters that lead to many of water's unusual properties,
+such as the high melting, boiling, and critical temperatures as well as the
+density maximum above the melting point.
+The large dielectric constant is partially responsible for
+water acting as a universal solvent, along with the ease of availability and
+use.
+There are many models used to describe water since so many different
+processes occur in aqueous solution. It is challenging to describe all the 
+relevant properties of water at once, especially if a low computational cost is
+also desired.
+
+Some models are primarily designed for use with homogeneous systems. Although
+these 
+models are useful in predicting some bulk properties, the interesting
+chemistry that occurs in water cannot be described with a homogeneous only
+model. To
+describe solvation and other processes in aqueous solution, a more robust theory
+that 
+can handle varying densities is needed.
 
@@
 
-It is challenging to describe all the relevant properties of water using
-a model, especially if a low computational cost is also desired.
+Classical density-functional theory (DFT) provides an efficient way to evaluate
+the free energy using the density of the fluid. The free energy is a functional
+of the density function and is independent of any external
+potentials\cite{ebner1976}. The most common use of DFT is calculating ground
+state electron densities in quantum mechanical systems. The Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} uses DFT to write the ground state
+energy in terms of a universal functional of the electron density and a separate
+term including the external potential. Mermin\cite{mermin1965thermal} gives a
+theorem extending that of Hohenberg and Kohn to nonzero temperatures:
+\begin{equation}
+  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
+d^3r\right\}
+\end{equation}
+This is the basis for classical DFT, where the relevant density is not the
+electron density,
+but the density of molecules. This can be used with a variety of 
+external potentials, from the simple interaction with a 
+hydrophobic wall to the more complex, stronger interaction with hydrophilic
+solutes.
 
-Some models, including SAFT, are designed for use with
-homogeneous systems only. Although these models are useful, much of the interesting
-chemistry that occurs in water cannot be described with a homogeneous only model. To
-describe solvation and other processes in aqueous solution, a more robust theory is 
-needed.
+\subsection{Statistical Associating Fluid Theory}
 
-Classical density-functional theory (DFT) provides an efficient way to evaluate the
-free energy using the density of the fluid. The free energy is a functional of the 
-density function and is independent of any external potentials\cite{ebner1976}. 
-Mermin\cite{mermin1965thermal} gives a theorem extending that of Hohenberg and 
-Kohn\cite{hohenberg1964inhomogeneous} to nonzero temperatures:
-\begin{equation}
-  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r) d^3r\right\}
-\end{equation}
-This is the basis for classical DFT, which can be used with a variety of external potentials, 
-from the simple interaction with a 
-hydrophobic wall to the complex strong interaction with hydrophilic solutes.
+The high hydrogen bonding and strong polar interactions in water characterize it
+as an associating fluid, and thus it requires a different treatment than simple
+fluids which are mainly defined by van der Waals forces and weak electrostatic
+interations. The Statistical Associating Fluid Theory (SAFT) was designed for
+use with fluids such as water, to account for the important association forces
+(like hydrogen bonds) as well as any effects due to clusters or chains forming
+in the liquid\cite{muller2001molecular}. It is based on Wertheim's first-order
+thermodynamic perturbation theory (TPT1)
+\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
+wertheim1986fluidsIV} which provides a useful way to account for strong
+associative interations between molecules.
 
@@
 We introduce a new classical density functional for water that
-predicts the \needsworklater{vapor density, liquid density, bulk
-  modulus, and bulk surface tension, as well as allows for
-  liquid-vapor coexistence}.  The Helmholtz free energy functional is
+predicts the vapor density, liquid density, and bulk surface tension, as well 
+as allows for liquid-vapor coexistence.  The Helmholtz free energy functional is
 based on SAFT, and is composed of the usual terms:
@@
 where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
-a hard-sphere free energy, $F_\textit{assoc}$ is the free energy of
-association, and $F_\textit{disp}$ is the dispersion energy.  In
+a hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
+and $F_\textit{assoc}$ is the free energy of association.  In
 addition, a chemical potential term is used, so we actually work in
@@
 provided the remaining terms are \emph{purely
-  nonlocal}~\cite{ashcroft?}.
+  nonlocal}~\cite{ashcroft?}\needsworknow{I can't find this citation anywhere!}.
 
@@
 
+\subsection{Dispersion free energy}
+The first term in the attractive energy is the dispersion term.  We use a
+dispersion term based on the SAFT-VR
+approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
+parameters, an interaction energy $\epsilon_\textit{disp}$ and a
+length scale $\lambda_\textit{disp}$.
+
+The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+\begin{align}
+  F_\text{disp} &= A_1 + \beta A_2
+\end{align}
+where $A_1$ and $A_2$ are the first two terms in a high-temperature
+perturbation expansion.  $A_1$ is the mean-field dispersion
+interaction, which for a homogeneous fluid is given by
+\begin{align}
+  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
+  g_{HS}(\left|\xx\right|) d\xx
+\end{align}
+The second dispersion term in the free energy $A_2$ describes the
+effect of fluctuations resulting from the compression of the fluid due
+to the dispersion interaction itself, and is commonly approximated
+using the local compressibility approximation (LCA).
+
+We use a free square-well function for the dispersion $\varphi$, which
+has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
+to the equation of state.  Thus our model interaction has the form:
+\begin{equation}
+  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
+\end{equation}
+where $\Theta$ is the Heaviside step function.
+
+\textcolor{red}{TODO:  Add the actual functions that we use and the
+  actual form.  The above is really just an introduction.}
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
 \end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
+\caption{The theoretical versus experimental coexistence curve. Away from
+the critical temperature, the agreement is good. }
+\label{fig:temperature-vs-density}
 \end{figure}
@@
 \subsection{Association free energy}
-The first attractive energy term is the association term, which
+The final attractive energy term is the association term, which
 accounts for hydrogen bonding.  Hydrogen bonds are modeled as four
@@
 experimental data.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\end{center}
+\caption{The theoretical versus experimental phase diagram, vapor
+  pressure versus temperature.  }
+\label{fig:equation-of-state}
+\end{figure}
 
@@
 
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/saturated-liquid-density}
-% \end{center}
-% \caption{The theoretical versus experimental saturated liquid density
-%   versus temperature.  }
-% \label{fig:saturated-liquid-density}
-% \end{figure}
-
-\subsection{Dispersion free energy}
-The final term in the free energy is the dispersion term.  We use a
-dispersion term based on the SAFT-VR
-approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters, an interaction energy $\epsilon_\textit{disp}$ and a
-length scale $\lambda_\textit{disp}$.
-
-The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
-\begin{align}
-  F_\text{disp} &= A_1 + \beta A_2
-\end{align}
-where $A_1$ and $A_2$ are the first two terms in a high-temperature
-perturbation expansion.  $A_1$ is the mean-field dispersion
-interaction, which for a homogeneous fluid is given by
-\begin{align}
-  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
-  g_{HS}(\left|\xx\right|) d\xx
-\end{align}
-The second dispersion term in the free energy $A_2$ describes the
-effect of fluctuations resulting from the compression of the fluid due
-to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA).
-
-We use a free square-well function for the dispersion $\varphi$, which
-has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
-to the equation of state.  Thus our model interaction has the form:
-\begin{equation}
-  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
-\end{equation}
-where $\Theta$ is the Heaviside step function.
-
-\textcolor{red}{TODO:  Add the actual functions that we use and the
-  actual form.  The above is really just an introduction.}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental temperature versus density. }
-\label{fig:temperature-vs-density}
-\end{figure}
-
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/entropy}
-% \end{center}
-% \caption{The free energy, internal energy, and temperature*entropy 
-% versus density at room temperature.  }
-% \label{fig:energy-room-temp}
-% \end{figure}
-
 \subsection{Other thermodynamic functions}
@@
 Figure \ref{fig:pressure-with-isotherms} shows the calculated pressures for
-isotherms ranging from room temperature to past the critical temperature. The solid 
-line is the theoretical coexistence curve calculated from the SAFT functional. 
-For a given isotherm, the pressure curve should intersect this coexistence curve 
-twice, at the values of the vapor and liquid densities. The functional matches well with
-the experimental data from NIST away from the critical point. The slope of the pressure
-versus density curve is proportional to the compressibility, which is close but slightly too
-large at room temperature.
+isotherms ranging from room temperature to past the critical temperature. The
+solid line is the theoretical coexistence curve calculated from the SAFT
+functional. For a given isotherm, the pressure curve should intersect this
+coexistence curve twice, at the values of the vapor and liquid densities. The
+functional matches well with the experimental data from NIST away from the
+critical point. The slope of the pressure versus density curve is proportional
+to the compressibility, which is close but slightly too large at room
+temperature.
 
@@
 \caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater} }
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+}
 \label{fig:surface-tension}
@@
 
-\fixme{I'm not sure if this paragraph belongs here.  Somewhere I think
-  it's worth discussing any of these homogeneous plots we include, and
-  this might be the right place.  We also might want to include one or
-  more of them earlier in the text.  A general rule is that any figure
-  that is included should be referenced in the text.  If there's
-  nowhere in the text that it fits, then it probably doesn't belong in
-  the paper....  So I'm not removing the follwoing bits since it's up
-  to you, Jess, whether to remove the figures.  We should
-  \emph{definitely} remove some of them.}  For reference, we plot the
-predicted vapor pressure (see Figure~\ref{fig:equation-of-state}) and
-coexistence curve (see
-Figure~\ref{fig:temperature-vs-density}) against experimental
-values.  To find the densities, we use the ``common tangent'' rule,
-as illustrated in Figure~\ref{fig:homogeneous}.
-
-% \begin{table}
-% \begin{tabular}{ccc}
-%   Fitting Parameters& \hspace{2em} & Experimental Values\\
-% \begin{tabular}{cc}
-%   $R$ & 2.3 $a_o$ \\
-%   $A$ & -0.0100909 Hartree \\
-%   $\gamma$ & 5\\
-%   $V$ & 515.792 $a_o^3$  \\
-%   $b$ & 0.75 $a_o$ \\
-%   $\mu$ & -0.014063 Hartree 
-% \end{tabular}
-% && 
-% \begin{tabular}{cc}
-%  $n_l$ & 4.939 $\times 10^{-3} a_o^{-3}$\\
-%  $n_v$ & 1.141 $\times 10^{-7} a_o^{-3}$\\
-%  $\sigma$ & ??? \\
-%  $B$ & $7.435\times 10^{-5}$ Hartree$/a_o^3$\\
-%  $T$ & 298.15 K \\
-% \end{tabular}
-% \end{tabular}
-% 
-% \caption{Empirical parameters of optimized functional and experimental
-%   values matched, where $a_o$ is the Bohr radius, and $\sigma$ is the
-%   surface tension.  This table is not yet complete, but it should
-%   be soon...  We'll also need to think about the best units to use
-%   here.  Atom units might be best, but it also might be better to
-%   consistently use eV and angstrom or something.  I'm not sure about
-%   putting $k_BT$ data in the table.  Maybe it belongs in the caption?}
-% \label{tab:parameters}
-% \end{table}
+For reference, we plot the predicted vapor pressure (see
+Figure~\ref{fig:equation-of-state}) and coexistence curve (see
+Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
+match those calculated by Clark \emph{et al}. To find the densities, we use the
+``common tangent'' rule, as illustrated in Figure~\ref{fig:homogeneous}.
+\needsworknow{Expand the previous paragraph}
 
@@
 \caption{The number of hydrogen bonds per molecule for a slab of water 
-with molecules constrained to move in one dimension. Each molecule may contribute to
-four bonds, although each bond involves two molecules so to calculate the actual 
+with molecules constrained to move in one dimension. Each molecule may
+contribute to
+four bonds, although each bond involves two molecules so to calculate the
+actual 
 number of bonds there would be another factor of 2.}
@@
 \end{figure}
-
-The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure \ref{fig:xassoc-1D} the number of hydrogen bonds per water 
-molecule, equal to $4(1-X)$, for a slit of 4.0 nm.
-Outside the slit, there is no hydrogen bonding and inside the number of hydrogen
-bonds is approximately equal to 3.5, which is expected for saturated liquid water.
-
-\subsection{One hydrophobic rod}
-
-We move into two dimensions by studying a single hydrophobic rod
-immersed in water. Figure \ref{fig:density-single-rod} shows density
-profiles for three different radii, 0.3 nm, 0.5 nm, and 1 nm. There are the
-expected oscillations outside the rod and as the distance away from the rod increases
-the density approaches that of the bulk saturated liquid density. The small radius
-rod has comparatively larger oscillations and a higher contact density.
 
@@
 
+The value of $X$ behaves as expected for water constrained to a slit. We
+show in Figure \ref{fig:xassoc-1D} the number of hydrogen bonds per water 
+molecule, equal to $4(1-X)$, for a slit of 4.0 nm.
+Outside the slit, there is no hydrogen bonding and inside the number of hydrogen
+bonds is approximately equal to 3.5, which is expected for saturated liquid
+water.
+
 \begin{figure}
@@
 immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in Figure 
-\ref{fig:surface-tension}. (FIX - 0.4 nm?) }
+tension at room temperature, and it agrees well with the surface tension in
+Figure \ref{fig:surface-tension}. }
 \label{fig:energy-vs-diameter}
 \end{figure}
-
-\subsubsection{Resolution and convergence tests}
-
-Figure \ref{fig:densityresolution} shows the density profiles for three different
-values of the lattice spacing resolution. The density oscillations around a single
-rod of radius 0.5 nm look nearly identical for all resolutions except for the first peak. 
-This contact density is important as it will be proportional to the pressure on the 
-rod\ref{contactvaluethm}. The 
-contact density calculated using the medium resolution is within less than 4\% of the value at
-high resolution, while the low resolution contact density is off by about 5\%. There is some
-uncertainty in the value of contact density due to the different lattice spacings. The total
-free energy calculations compare much more favorably. The medium resolution is within 
-0.5\% and the low resolution is within 1\% of the high resolution free energy.
 
@@
 a single hydrophobic rod with a radius of 0.5 nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05 bohrs ($\approx$ 0.003 nm), 
+lattice spacing resolutions. We used a high resolution of 0.05 bohrs ($\approx$
+0.003 nm), 
 medium 0.2 bohrs ($\approx$ 0.001 nm), and low 0.5 bohrs ($\approx$ 0.03 nm).}
@@
 
+\subsection{One hydrophobic rod}
+We move into two dimensions by studying a single hydrophobic rod
+immersed in water. Figure \ref{fig:density-single-rod} shows density
+profiles for three different radii, 0.3 nm, 0.5 nm, and 1 nm. There are the
+expected oscillations outside the rod and as the distance away from the rod
+increases
+the density approaches that of the bulk saturated liquid density. The small
+radius
+rod has comparatively larger oscillations and a higher contact density.
+
+Figure \ref{fig:energy-vs-diameter} shows the free energy per area versus
+radius for the range of rods studied, with radii from 0.05 nm to 1.5 nm. After 
+about 0.5 nm or so, this reaches
+an asymptote equal to the surface tension. From the inset in Figure 
+\ref{fig:surface-tension}, it is clear that the surface tension at room 
+temperature is approximately 72 mN/m. This is very close to the value from
+the larger diameter single rod energy per area, which is about 75~mN/m. 
+\needsworknow{Here it will be beneficial to add some discussion of any
+relevant experimental work or other simulations that deal with one rod-like 
+thing in water. Do we have anything currently matching this?}
+
+\subsubsection{Resolution and convergence tests}
+Figure \ref{fig:densityresolution} shows the density profiles for three
+different values of the lattice spacing resolution. The density oscillations
+around a single rod of radius 0.5 nm look nearly identical for all resolutions
+except for the first peak. This contact density is important as it will be
+proportional to the pressure on the rod (see equation \ref{contactvaluethm}). The contact
+density calculated using the medium resolution is within less than 4\% of the
+value at high resolution, while the low resolution contact density is off by
+about 5\%. There is some uncertainty in the value of contact density due to the
+different lattice spacings. The total free energy calculations compare much more
+favorably. The medium resolution is within 0.5\% and the low resolution is
+within 1\% of the high resolution free energy.
+
 \subsection{Hydrophobic interaction of two rods}
 
-Here we will want to cite~\cite{walther2004hydrodynamic}, which gives
-simulations for two interacting nanotubes in water.  It's not
-precisely the same, but if we can't find something more similar, this
-would be worth citing.
+Here we will want to cite~\cite{walther2004hydrodynamic}, which gives simulations for 
+two interacting
+nanotubes in water.  It's not precisely the same, but if we can't find something
+more similar, this would be worth citing - they use CNTS with diameter 1.25 nm
+and find drying occurs for separations less than 0.9-1nm.
 
@@
 \end{center}
-\caption{ Energy versus separation for two identical hydrophobic rods with diameters
-of 1.0 nm. }
+\caption{ Energy versus separation for two hydrophobic rods. Radii 0.3nm, 0.5nm, 1.0nm
+ }
 \label{fig:rods-energy-vs-distance}

### 2011-09-06 paper/project.tex
Fixed entropy in rods, edits to paper, adding xassoc for single rod

@@
 solutes.  These properties are retained by the total functional,
-provided the remaining terms are \emph{purely
-  nonlocal}~\cite{ashcroft?}\needsworknow{I can't find this citation anywhere!}.
+provided the remaining terms are \emph{purely nonlocal}. This requires
+that
+\begin{equation}
+ \frac{\delta^2F_{\textit{excess}}[n]}{\delta n(\xx_1)~\delta n(\xx_1)}\neq\infty
+\end{equation}
+where $F_{\textit{excess}}[n]$ describes all the remaining terms in the functional.
+This ensures that any higher order terms in the Taylor expansion
+of $F[n+\Delta n]$ are very small.
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
+\end{center}
+\caption{ Number of bonds per molecule near the interface with a single
+hydrophobic rod. Compare with Walther\cite{walther2001carbon} fig 9. }
+\label{fig:xassoc-single-rod}
+\end{figure}
+
 Figure \ref{fig:energy-vs-diameter} shows the free energy per area versus
@@
 the larger diameter single rod energy per area, which is about 75~mN/m. 
-\needsworknow{Here it will be beneficial to add some discussion of any
-relevant experimental work or other simulations that deal with one rod-like 
-thing in water. Do we have anything currently matching this?}
+\needsworknow{Discussion of Walther 2001 paper\cite{walther2001carbon}}
 
@@
 
+\bibliographystyle{unsrt}
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2011-09-07 paper/project.tex
improve the dispersion section just a tad
Tested-on: bingley (Linux 2.6.32-5-amd64)

--- 2011-09-01 paper/project.tex
@@
 perturbation expansion.  $A_1$ is the mean-field dispersion
-interaction, which for a homogeneous fluid is given by
+interaction, which reduces in the homogeneous limit to
 \begin{align}
-  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
+  \frac{A_1}{V} &= \frac12 n_b \int \varphi(\left|\xx\right|)
   g_{HS}(\left|\xx\right|) d\xx
@@
 We use a free square-well function for the dispersion $\varphi$, which
-has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
-to the equation of state.  Thus our model interaction has the form:
+is the choice used by Clark \emph{et al}~\cite{clark2006developing},
+and allows a reasonable fit to the equation of state.  Thus our model
+interaction has the form:
 \begin{equation}
@@
 
-\textcolor{red}{TODO:  Add the actual functions that we use and the
-  actual form.  The above is really just an introduction.}
+The form of $A_1$ is given in
+reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
+of the filling fraction.  In order to apply this form to an
+inhomogeneous density distribution, we construct an effective local
+filling fraction $\eta{\textit{eff}}$, given by
+\begin{align}
+  \eta_{\textit{eff}}(\mathbf{r}) &= \frac{4\pi R^3}{3}
+  \int \left(\frac{\pi}{\sigma}\right)^{\frac32}
+  e^{-\frac{|\mathbf{r}-\mathbf{r}'|^2}{2\sigma^2}} n(\mathbf{r}') d\mathbf{r}'
+\end{align}
+This effective filling fraction is used throughout the dispersion
+calculations.  This functional introduces an additional empirical
+parameter $\sigma$ \fixme{(The above formula still needs fixing)}
+which adjusts the length scale over which the dispersion interaction
+is correlated.
+
+\fixme{Do we want to include all the gory details? It's pretty
+  complicated, which could be used to argue in either direction.  On
+  the one hand, a complete summary that doesn't require users to look
+  up the 1997 paper would be helpful.  On the other hand, these
+  precise details are lengthy and a pain to write, and aren't all that
+  interesting for most readers to read.}
+
+The second term, $A_2$, which describes the contribution to the free
+energy associated with fluctuations is given by
+\begin{align}
+  A_2 &= \int \frac12 \epsilon_\textit{disp}
+              K_{HS}(\eta_{\textit{eff}}(\mathbf{r}))
+              \eta_{\textit{eff}}(\mathbf{r})
+              \frac{\delta A_1}{\delta \eta_\textit{eff}(\mathbf{r})}
+                d\mathbf{r}
+\end{align}
+where $K_{HS}$ is the Percus-Yevick hard-sphere compressibility, given
+by
+\begin{align}
+  K_{HS} &=
+    \frac{(1 - \eta_\textit{eff})^4}{1 + 4(\eta_\textit{eff} + \eta_\textit{eff}^2)}
+\end{align}
 

### 2011-09-08 paper/project.tex
rewrite association section and extend dispersion section Tested everything.

@@
 
+\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
+\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
+\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
+\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
 \subsection{Dispersion free energy}
@@
 approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters, an interaction energy $\epsilon_\textit{disp}$ and a
-length scale $\lambda_\textit{disp}$.
+parameters, an interaction energy $\epsilondisp$ and a
+length scale $\lambdadisp$.
 
@@
 interaction, which reduces in the homogeneous limit to
-\begin{align}
+\begin{align}\label{eq:A1-simple}
   \frac{A_1}{V} &= \frac12 n_b \int \varphi(\left|\xx\right|)
@@
 \begin{equation}
-  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
+  \varphi(r) = \Theta(r-2 \lambdadisp R)
 \end{equation}
@@
 inhomogeneous density distribution, we construct an effective local
-filling fraction $\eta{\textit{eff}}$, given by
+filling fraction for dispersion $\etadisp$, given by
 \begin{align}
-  \eta_{\textit{eff}}(\mathbf{r}) &= \frac{4\pi R^3}{3}
+  \etadisp(\xx) &= \frac{4\pi R^3}{3}
   \int \left(\frac{\pi}{\sigma}\right)^{\frac32}
-  e^{-\frac{|\mathbf{r}-\mathbf{r}'|^2}{2\sigma^2}} n(\mathbf{r}') d\mathbf{r}'
+  e^{-\frac{|\xx-\xx'|^2}{2\sigma^2}} n(\xx') d\xx'
 \end{align}
@@
 calculations.  This functional introduces an additional empirical
-parameter $\sigma$ \fixme{(The above formula still needs fixing)}
+parameter $\sigma$ \fixme{(The above formula still needs fixing, and
+  is affected by one of Eric's recent changes.)}
 which adjusts the length scale over which the dispersion interaction
@@
 
-\fixme{Do we want to include all the gory details? It's pretty
-  complicated, which could be used to argue in either direction.  On
-  the one hand, a complete summary that doesn't require users to look
-  up the 1997 paper would be helpful.  On the other hand, these
-  precise details are lengthy and a pain to write, and aren't all that
-  interesting for most readers to read.}
+The first term in the dispersion functional $A_1$ when written using
+the above filling fraction for dispersion has the form
+\begin{align}
+  A_1 &= \int
+   -4*(\lambdadisp^3 - 1)*\epsilondisp \etadisp(\xx)
+    g^{HS}_\sigma(\eta_\textit{eff}(\xx))
+  d\xx \\
+  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
+  \lambdadisp^j
+  \\
+  g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
+  +\frac32\frac{\zeta(\xx)}{(1-\eta)^2}
+  + \frac12\frac{\eta^2\zeta(\xx)}{(1-\eta)^3}
+  \\
+  \zeta(\xx) &= 1 - \frac{\mathbf{n_{2V}}\cdot\mathbf{n_{2V}}}{n_2^2}
+\end{align}
+where $C_{ij}$ are numerical constants taken from
+reference\cite{gil-villegas-1997-SAFT-VR}, which come from a numerical
+fit to the integral in Equation~\ref{eq:A1-simple} over a range of
+values for $\lambdadisp$ and filling fractions.  The $\zeta$ factor
+(also seen in the association energy, below) is a dimensionless
+measure of the local density inhomogeneity.
 
@@
 \begin{align}
-  A_2 &= \int \frac12 \epsilon_\textit{disp}
-              K_{HS}(\eta_{\textit{eff}}(\mathbf{r}))
-              \eta_{\textit{eff}}(\mathbf{r})
-              \frac{\delta A_1}{\delta \eta_\textit{eff}(\mathbf{r})}
-                d\mathbf{r}
+  A_2 &= \int \frac12 \epsilondisp
+              K_{HS}(\etadisp(\xx))
+              \etadisp(\xx)
+              \frac{\delta A_1}{\delta \etadisp(\xx)}
+                d\xx
 \end{align}
@@
   K_{HS} &=
-    \frac{(1 - \eta_\textit{eff})^4}{1 + 4(\eta_\textit{eff} + \eta_\textit{eff}^2)}
+    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}
 \end{align}
@@
 sphere.  These four sites represent two protons and two electron lone
-pairs.  There is an attractive energy $\epsilon_\textit{assoc}$ when
-two molecules are arranged such that the proton of one overlaps
+pairs.  There is an attractive energy $\epsilonassoc$ when
+two molecules are oriented such that the proton of one overlaps
 with the lone pair of the other.  The volume over which this
@@
   the same functional, but their paper contains errors in the
-  association term and is not useful as a reference.}  The functional
-of Fu and We includes the effects of density inhomogeneities on the
-\emph{contact density} of hard spheres, which conventionally appears
-in the form of the density multiplied by the contact value of the
-correlation function $g^{HS}_\sigma$, but is based on the SAFT-HS
-equation of state\cite{FIXME}, rather than the SAFT-VR equation of
-state\cite{gil-villegas-1997-SAFT-VR} which is used in the optimal
-SAFT parametrization for water found by Clark \emph{et
+  association term and is not useful as a reference.}  The association
+functional of Fu and Wu includes the effects of density
+inhomogeneities on the \emph{contact density} of hard spheres, which
+conventionally appears in the form of the density multiplied by the
+contact value of the correlation function $g^{HS}_\sigma$, but is
+based on the SAFT-HS equation of state, rather than the SAFT-VR
+equation of state\cite{gil-villegas-1997-SAFT-VR} which is used in the
+optimal SAFT parametrization for water found by Clark \emph{et
   al}\cite{clark2006developing}.
@@
 \begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
   \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
@@
 molecule, and the functional $X$ is the fraction of assotiation sites
-\emph{not} hydrogen-bonded.  The fraction $X$ is determined by the
-quadratic equation
+\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
+measure of the density inhomogeneity.  The fraction $X$ is determined
+by the quadratic equation
 \begin{align}
@@
   fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
-density, and reduces to unity in the homogeneous limit.
+density, and reduces to unity in the homogeneous limit:
 \begin{align}
@@
 \end{align}
-The functional $\Delta$ is a measure of the hydrogen-bonding
-probability.  In the SAFT-VR model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is
-given by
+The functional $\Delta$ which appears in the formula for $X$ is a
+measure of hydrogen-bonding probability.  In the SAFT-VR
+model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
 \begin{align}
-  \Delta &= \kappa_\textit{assoc} g^\textit{SW}_\sigma
-  \left(e^{-\beta\epsilon_\textit{assoc}} - 1\right) \\
-  g^\textit{SW}_\sigma &= g^\textit{HS}_\sigma +
-  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial n_3} -
-  \frac{\lambda}{3 n3}\frac{\partial a_1}{\partial \lambda}\right)
+  \Delta(\xx) &= \kappa_\textit{assoc} g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right) \\
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\etadisp(\xx)) +
+  \frac{1}{4}\beta\left(\frac{\delta A_1}{\delta \etadisp(\xx)} -
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)
 \end{align}
@@
 \begin{align}
-  g_\sigma^{HS} &= \frac{1}{1-n_3}
-  +\frac32\frac{\zeta}{(1-n_3)^2}
-  + \frac12\frac{n_3^2\zeta}{(1-n_3)^3}
+  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
+  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
+  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}
 \end{align}

### 2011-09-15 paper/project.tex
Edits to two rods section, figures and text Tested everything.

@@
 \newcommand{\rr}{\textbf{r}}
-\newcommand{\xx}{\textbf{x}}
+\newcommand{\xx}{\textbf{r}}
 \newcommand{\refnote}{\red{[ref]}}
@@
 \begin{abstract}
-\needsworklater{ We develop a classical density functional theory for
+ {We develop a classical density functional theory for
   water by combining the fundamental-measure theory functional--which
@@
 models are useful in predicting some bulk properties, the interesting
-chemistry that occurs in water cannot be described with a homogeneous only
+chemistry that occurs in water cannot be described with a homogeneous-only
 model. To
@@
 Classical density-functional theory (DFT) provides an efficient way to evaluate
-the free energy using the density of the fluid. The free energy is a functional
+the free energy using the probability distribution of the fluid. The free energy 
+is a functional
 of the density function and is independent of any external
@@
 state electron densities in quantum mechanical systems. The Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} uses DFT to write the ground state
+theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
 energy in terms of a universal functional of the electron density and a separate
@@
 \begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n_{2V}}\cdot\mathbf{n_{2V}}}{n_2^2}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}
 \end{align}
@@
 \end{center}
-\caption{ Energy versus diameter for a single hydrophobic rod
+\caption{ Energy versus radius for a single hydrophobic rod
 immersed in water. This should have an asymptote equal to the surface
@@
 
-Here we will want to cite~\cite{walther2004hydrodynamic}, which gives simulations for 
-two interacting
-nanotubes in water.  It's not precisely the same, but if we can't find something
-more similar, this would be worth citing - they use CNTS with diameter 1.25 nm
-and find drying occurs for separations less than 0.9-1nm.
-
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/density-rods-in-water}
-% \end{center}
-% \caption{ Density profile for two hydrophobic rods (diameter 1.0 nm) at a 
-% separation of 0.8 nm. The scale on the right is density in g/mL.}
-% \label{fig:density-rods}
-% \end{figure}
-
 \begin{figure}
@@
 \begin{center}
+\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\end{center}
+\caption{ Density profiles illustrating the transition from vacuum 
+to liquid water between the rods. The top figure is at a separation of 0.2nm and the
+bottom is 0.4nm. }
+\label{fig:density-rods}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
 \end{center}
-\caption{ Energy versus separation for two hydrophobic rods. Radii 0.3nm, 0.5nm, 1.0nm
- }
+\caption{ Energy versus separation for two hydrophobic rods ranging in radius from
+0.3nm to 1.2nm. All were arbitrarily set to go to zero at large separations. The
+discontinuous transition corresponds to the phase change from
+vacuum to liquid between the rods. }
 \label{fig:rods-energy-vs-distance}
 \end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-rods}
+\end{center}
+\caption{ Free energy, internal energy, and the entropic energy contribution versus
+separation for two hydrophobic rods of radii 0.5nm. All energies are arbitrarily
+set to go to zero at large separations. }
+\label{fig:energy-rods}
+\end{figure}
+
+We now look at the more interesting problem of two rods in water. We continue to
+use hydrophobic rods of various radii at separation $d$ (see Figure
+\ref{fig:rods}), ranging from $d=0$ (touching rods) to about 1nm. At small
+separations ($d<$2nm-4nm) there is only vacuum between the rods. As the rods are
+pulled apart, the vacuum region expands until some critical separation is
+reached and liquid water fills the region between the rods. Figure
+\ref{fig:density-rods} shows density profiles before and after this transition
+for rods of radius 0.5nm. Since we are looking at only hard-walled
+non-interacting rods, the critical separation for this transition will be
+smaller than any system where there is an attractive force between rods. Walther
+\emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
+carbon nanotubes. They use molecular dynamics with Lennard-Jones potentials for
+nanotubes of diameter 1.25nm and separations ranging from about 0.3nm to 1.5nm.
+They also find a drying transition occurs, and find it to be at separations less
+than 0.9-1.0nm\cite{walther2004hydrodynamic}. This is several times larger than
+what we find for hydrophobic rods, as expected due to the carbon-carbon
+interations between two nanotubes. The density oscillations are very similar to
+those for the single hydrophobic rod in water, as seen in Figure
+\ref{fig:density-single-rod}.
+
+The free energy per length is calculated for several different radii and is plotted
+in Figure \ref{fig:rods-energy-vs-distance}. Each data set is adjusted such that the
+free energy per length goes to zero at large separations. Once a certain separation is
+reached, the rods will simply act like the single rod in water and the energy will
+reach an asymptote proportional to the surface tension (as in Figure \ref{fig:energy-vs-diameter}).
+
+For one set of rods (radii of 0.5nm) the free energy is plotted with the internal energy
+and the entropic energy contribution, where all are in units of energy per length (see Figure
+\ref{fig:energy-rods}).
+
+Here we will want to cite~\cite{walther2004hydrodynamic}, which gives simulations for 
+two interacting
+nanotubes in water.  It's not precisely the same, but if we can't find something
+more similar, this would be worth citing - they use CNTS with diameter 1.25 nm
+and find drying occurs for separations less than 0.9-1nm.
 

### 2011-09-20 paper/project.tex
fix the discussion of \etadisp Tested everything.

@@
 \newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
+\newcommand\lscale{\ensuremath{s_d}}
 \subsection{Dispersion free energy}
@@
 \begin{align}
-  \etadisp(\xx) &= \frac{4\pi R^3}{3}
-  \int \left(\frac{\pi}{\sigma}\right)^{\frac32}
-  e^{-\frac{|\xx-\xx'|^2}{2\sigma^2}} n(\xx') d\xx'
+  \etadisp(\xx) &=
+  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
+             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'
 \end{align}
 This effective filling fraction is used throughout the dispersion
-calculations.  This functional introduces an additional empirical
-parameter $\sigma$ \fixme{(The above formula still needs fixing, and
-  is affected by one of Eric's recent changes.)}
-which adjusts the length scale over which the dispersion interaction
-is correlated.
+calculations, and represents a filling fraction averaged over the
+effective range of the dispersive interaction.  This functional
+introduces an additional empirical parameter $\lscale$ which adjusts
+the length scale over which the dispersion interaction is correlated.
 
@@
 empirical parameters of Clark \emph{et al}, we add a single additional
-parameter defining the length scale over which the density is averaged
-when computing the dispersion free energy.  We use this final
-parameter to fit the surface tension, with the result shown in
+parameter $\lscale$---with a fitted value of 0.72---which determines
+the length scale over which the density is averaged when computing the
+dispersion free energy.  We use this final parameter to fit the
+surface tension, with the result shown in
 Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark

### 2011-09-26 paper/project.tex
Changed all energy units, added arrows in rods figs, fixed tests to work with new ideal gas term, edits to project text Tested everything.

--- 2011-09-15 paper/project.tex
@@
 \begin{abstract}
- {We develop a classical density functional theory for
-  water by combining the fundamental-measure theory functional--which
-  describes hard sphere fluids--with attractive interactions based on the 
-  Statistical Associating Fluid Theory (SAFT).  Using
-  this approximation, we are able to describe various properties of
-  water near liquid-vapor and liquid-solid interfaces, including water confined 
-  to a hard-walled slit and hydrophobic rods immersed in water.  This
-functional 
-  will be the foundation for a efficient method for future work
-  predicting the interactions of chemicals in aqueous solution on an
+{We develop a classical density functional theory for water by combining the fundamental-measure theory functional--which describes hard sphere fluids--with attractive interactions based on the Statistical Associating Fluid Theory (SAFT).
+Using this approximation, we are able to describe the density, free energy, and
+entropy of water, as well as the number of hydrogen bonds formed near liquid-vapor and liquid-solid interfaces.
+These interfaces include water confined to a hard-walled slit, a single hydrophobic rod immersed in water, and two hydrophobic rods
+at various separations. We found \needsworknow{conclusion-like stuff about two rods, vapor-liquid transition between rods, constant force of twice surface tension, more?}.
+This functional provides a foundation for future work predicting the interactions of chemicals in aqueous solution on an
 atomic level.}
 
-\tableofcontents
+%\tableofcontents
 \end{abstract}
@@
 \begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{n(\xx)} - 1\right) d\xx
+  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx
 \end{equation}
-where n(\xx) is the density of water molecules.  The ideal gas free
+where n(\xx) is the density of water molecules and $n_Q$ is the quantum 
+concentration. This is the particle concentration where quantum effects
+become appreciable and is given by
+\begin{equation}\label{quantumconcentration}
+ n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}
+\end{equation}
+where $m$ is the molecular mass of water. The ideal gas free
 energy functional has the property that the contact density at a hard
@@
 where the factor $\zeta$ from
-Yu and We\cite{yu2002fmt-dft-inhomogeneous-associating,
+Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
   fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
@@
 functional. For a given isotherm, the pressure curve should intersect this
-coexistence curve twice, at the values of the vapor and liquid densities. The
+coexistence curve twice, at the values of the vapor and liquid densities. Even
+though the theoretical pressure isotherms drop below the coexistence curve,
+this is non-physical and will not actually occur. The experimental data shows
+how the pressure will jump to the other side of the coexistence curve (for
+temperatures less than the critical temperature), corresponding to a phase 
+change in the material. The
 functional matches well with the experimental data from NIST away from the
@@
 
-Once we have the entropy, finding the internal energy is a simple
-matter of subtraction:
+Once we have the entropy, finding the internal energy is simple:
 \begin{align}
@@
 The calculated heat of vaporization at atmospheric pressure is
-$\Delta H_{vap}$= 39.41 kJ/mol at the normal boiling point of water,
+$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
 373 K. This compares well to the experimental
-value measured by NIST ($\Delta H_{vap}$= 40.65 kJ/mol\cite{nistwater}), with 
+value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
 an error of only a few percent.
@@
 \end{center}
-\caption{By plotting the free energy vs. density it is possible to
+\caption{By plotting the free energy versus density it is possible to
   determine a saturated liquid or vapor state by the common tangent
-  method.  }
+  method. The top plot illustrates the common tangents near room 
+  temperature, while the bottom plot shows the same thing close to the 
+  critical temperature. }
 \label{fig:homogeneous}
@@
   versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\sigma$ is fit so that the theoretical surface tension
+  will match the experimental surface tension near room temperature.
 }
@@
 empirical parameters of Clark \emph{et al}, we add a single additional
-parameter defining the length scale over which the density is averaged
+parameter $\sigma$ defining the length scale over which the density is averaged
 when computing the dispersion free energy.  We use this final
@@
 constrained to move in one dimension. The free energy, internal energy
-and entropic energy densities are shown for a slit size of 4.0 nm. The dashed
+and entropic energy densities are shown for a slit size of 4.0~nm. The dashed
 line in the top figure represents the saturated liquid density.}
@@
 that the density profile will be smoothed out by CW effects. The liquid-vapor
-interface is plotted in Figure \ref{fig:liquid-vapor-profile} and shows
+interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
 some small oscillations.
@@
 
-We choose a chemical potential corresponding to a pressure of 1 atm and set the
-temperature to a typical room temperature of 298 K. We apply 
+We choose a chemical potential corresponding to a pressure of 1~atm and set the
+temperature to a typical room temperature of 298~K. We apply 
 a constraining potential outside the slit, where we choose a small non-zero
-density for numerical reasons. The slits studied range in size from 0.1 nm to 
-8.0 nm. We minimize the energy and calculate density, energy, and $X$ on a
+density for numerical reasons. The slits studied range in size from 0.1~nm to 
+8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
 lattice
-with a spacing of approximately 0.01 nm (0.2 bohrs) between lattice points.
+with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
 
-The density profile for a slit of size 4.0 nm is shown in Figure
-\ref{fig:energy-1D}. Near the interface, there are oscillations around the
+The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:energy-1D}. Near the interface, there are oscillations around the
 liquid density which decrease near the center of the slit. We expect the center
 of the slit to act approximately like bulk water, and the absence of
-oscillations supports this. For larger slits (6.0 nm or more) the majority of
+oscillations supports this. For larger slits (6.0~nm or more) the majority of
 the water molecules in the slit act like the bulk. Based on our computations, we
 also determine the smallest slit that can contain liquid water. This result is
-approximately 0.6 nm. For smaller slits, the initial liquid water inside will
+approximately 0.6~nm. For smaller slits, the initial liquid water inside will
 vaporize. The results for larger liquid-filled slits are only metastable,
 however,as a vapor-filled slit is thermodynamically favorable. Based on our
-calculations for a pressure of 1 atm, the work needed to pull apart
+calculations for a pressure of 1~atm, the work needed to pull apart
 the two hard walls comprising the slit is one to two orders of magnitude less
 than the free energy of the liquid-filled slit. For the range of slit sizes we
-studied (0.1 to 8.0 nm), all the computational results of liquid-filled are
+studied (0.1 to 8.0~nm), all the computational results of liquid-filled are
 metastable. Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
@@
 necessary for stability. For water at room temperature and atmospheric pressure,
-plates with a separation smallerthan about 100 nm will not contain stable liquid
+plates with a separation smaller than about 100~nm will not contain stable liquid
 water. They go on to say that confined water has a region of metastability down
-to separations of about 5 nm. This is comparable to the slit separation where
+to separations of about 5~nm. This is comparable to the slit separation where
 our calculations show water inside the slit acting like the bulk liquid, which
-we know to be a metastable solution. Forsman et al.\cite{forsman1996computer}
+we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
 also study water between two hard walls at constant chemical potential. They 
@@
 The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure \ref{fig:xassoc-1D} the number of hydrogen bonds per water 
-molecule, equal to $4(1-X)$, for a slit of 4.0 nm.
+show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
+molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
 Outside the slit, there is no hydrogen bonding and inside the number of hydrogen
@@
 tension at room temperature, and it agrees well with the surface tension in
-Figure \ref{fig:surface-tension}. }
+Figure~\ref{fig:surface-tension}. }
 \label{fig:energy-vs-diameter}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
+\end{center}
+\caption{ Number of bonds per molecule near the interface with a single
+hydrophobic rod. There is a decrease in the hydrogen bonding near the surface 
+of the rod, as expected at the interface of any hydrophobic wall. }
+\label{fig:xassoc-single-rod}
 \end{figure}
@@
 \caption{ Density profiles for 
-a single hydrophobic rod with a radius of 0.5 nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05 bohrs ($\approx$
-0.003 nm), 
-medium 0.2 bohrs ($\approx$ 0.001 nm), and low 0.5 bohrs ($\approx$ 0.03 nm).}
+a single hydrophobic rod with a radius of 0.5~nm using several different 
+lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$
+0.003~nm), 
+medium 0.2~bohrs ($\approx$ 0.001~nm), and low 0.5~bohrs ($\approx$ 0.03~nm).}
 \label{fig:densityresolution}
@@
 We move into two dimensions by studying a single hydrophobic rod
-immersed in water. Figure \ref{fig:density-single-rod} shows density
-profiles for three different radii, 0.3 nm, 0.5 nm, and 1 nm. There are the
+immersed in water. Figure~\ref{fig:density-single-rod} shows density
+profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
 expected oscillations outside the rod and as the distance away from the rod
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
-\end{center}
-\caption{ Number of bonds per molecule near the interface with a single
-hydrophobic rod. Compare with Walther\cite{walther2001carbon} fig 9. }
-\label{fig:xassoc-single-rod}
-\end{figure}
+Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
+radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
+about 0.5~nm or so, this reaches
+an asymptote equal to the surface tension. From the inset in Figure~\ref{fig:surface-tension}, the surface tension at room 
+temperature is approximately 72~mN/m. This is very close to the value from
+the larger diameter single rod energy per area, which is about 75~mN/m.
 
-Figure \ref{fig:energy-vs-diameter} shows the free energy per area versus
-radius for the range of rods studied, with radii from 0.05 nm to 1.5 nm. After 
-about 0.5 nm or so, this reaches
-an asymptote equal to the surface tension. From the inset in Figure 
-\ref{fig:surface-tension}, it is clear that the surface tension at room 
-temperature is approximately 72 mN/m. This is very close to the value from
-the larger diameter single rod energy per area, which is about 75~mN/m. 
-\needsworknow{Discussion of Walther 2001 paper\cite{walther2001carbon}}
+The value of $X$ is calculated and converted to number of bonds per molecule and 
+shown for a rod with radius 0.2~nm in Figure~\ref{fig:xassoc-single-rod}. Inside the
+rod (less than 0.2~nm) there are no molecules to form hydrogen bonds although the value
+of $X$ is still calculated there. Walther \emph{et al.}\cite{walther2001carbon} 
+calculate the number of hydrogen bonds for a carbon nanotube immersed in water.
+They show a decrease in the bonding from a bulk value of 3.75 to 2.89 near the 
+interface with the nanotube. This is a much larger change than for our hydrophobic
+rod of a similar radius, where the number of bonds decreases from a bulk value of 3.5 
+to about 3.25 at the interface (see Figure~\ref{fig:xassoc-single-rod}).
 
 \subsubsection{Resolution and convergence tests}
-Figure \ref{fig:densityresolution} shows the density profiles for three
+Figure~\ref{fig:densityresolution} shows the density profiles for three
 different values of the lattice spacing resolution. The density oscillations
-around a single rod of radius 0.5 nm look nearly identical for all resolutions
+around a single rod of radius 0.5~nm look nearly identical for all resolutions
 except for the first peak. This contact density is important as it will be
-proportional to the pressure on the rod (see equation \ref{contactvaluethm}). The contact
+proportional to the pressure on the rod (see Equation~\ref{contactvaluethm}). The contact
 density calculated using the medium resolution is within less than 4\% of the
@@
 \end{center}
-\caption{ Density profiles illustrating the transition from vacuum 
-to liquid water between the rods. The top figure is at a separation of 0.2nm and the
-bottom is 0.4nm. }
+\caption{ Density profiles illustrating the transition from vapor 
+to liquid water between the rods. The radius is 0.5~nm, the top figure is at a separation of 0.2~nm and the
+bottom is 0.4~nm. }
 \label{fig:density-rods}
@@
 \caption{ Energy versus separation for two hydrophobic rods ranging in radius from
-0.3nm to 1.2nm. All were arbitrarily set to go to zero at large separations. The
+0.3~nm to 0.9~nm. All were arbitrarily set to go to zero at large separations. The
 discontinuous transition corresponds to the phase change from
-vacuum to liquid between the rods. }
+vapor to liquid between the rods. }
 \label{fig:rods-energy-vs-distance}
@@
 \caption{ Free energy, internal energy, and the entropic energy contribution versus
-separation for two hydrophobic rods of radii 0.5nm. All energies are arbitrarily
+separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
 set to go to zero at large separations. }
@@
 We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of various radii at separation $d$ (see Figure
-\ref{fig:rods}), ranging from $d=0$ (touching rods) to about 1nm. At small
-separations ($d<$2nm-4nm) there is only vacuum between the rods. As the rods are
-pulled apart, the vacuum region expands until some critical separation is
-reached and liquid water fills the region between the rods. Figure
-\ref{fig:density-rods} shows density profiles before and after this transition
-for rods of radius 0.5nm. Since we are looking at only hard-walled
+use hydrophobic rods of various radii at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+(touching rods) to about 1~nm. At small
+separations ($d<2$~-~4~nm) there is only vapor between the rods. As the rods are
+pulled apart, the vapor region expands until some critical separation is
+reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
+shows density profiles before and after this transition
+for rods of radius 0.5~nm. Since we are looking at only hard-walled
 non-interacting rods, the critical separation for this transition will be
-smaller than any system where there is an attractive force between rods. Walther
+different than any system where there is an appreciable force between rods.
+
+The free energy per length is calculated for several different radii and is plotted
+in Figure~\ref{fig:rods-energy-vs-distance}. Each data set is adjusted such that the
+free energy per length goes to zero at large separations. Once a certain separation is
+reached, the rods will simply act like the single rod in water and the energy will
+reach an asymptote proportional to the surface tension (as in Figure~\ref{fig:energy-vs-diameter}).
+
+For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
+and the entropic energy contribution, where all are in units of energy per length 
+(see Figure~\ref{fig:energy-rods}). \needsworknow{More discussion of Figure~\ref{fig:energy-rods}}
+
+Walther
 \emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
 carbon nanotubes. They use molecular dynamics with Lennard-Jones potentials for
-nanotubes of diameter 1.25nm and separations ranging from about 0.3nm to 1.5nm.
+nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
 They also find a drying transition occurs, and find it to be at separations less
-than 0.9-1.0nm\cite{walther2004hydrodynamic}. This is several times larger than
+than 0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is several times larger than
 what we find for hydrophobic rods, as expected due to the carbon-carbon
-interations between two nanotubes. The density oscillations are very similar to
-those for the single hydrophobic rod in water, as seen in Figure
-\ref{fig:density-single-rod}.
+interations between two nanotubes.
 
-The free energy per length is calculated for several different radii and is plotted
-in Figure \ref{fig:rods-energy-vs-distance}. Each data set is adjusted such that the
-free energy per length goes to zero at large separations. Once a certain separation is
-reached, the rods will simply act like the single rod in water and the energy will
-reach an asymptote proportional to the surface tension (as in Figure \ref{fig:energy-vs-diameter}).
-
-For one set of rods (radii of 0.5nm) the free energy is plotted with the internal energy
-and the entropic energy contribution, where all are in units of energy per length (see Figure
-\ref{fig:energy-rods}).
-
-Here we will want to cite~\cite{walther2004hydrodynamic}, which gives simulations for 
-two interacting
-nanotubes in water.  It's not precisely the same, but if we can't find something
-more similar, this would be worth citing - they use CNTS with diameter 1.25 nm
-and find drying occurs for separations less than 0.9-1nm.
+The density oscillations are very similar to
+those for the single hydrophobic rod in water, as seen in Figure~\ref{fig:density-single-rod}.
 

### 2011-09-26 paper/project.tex
Edits to project figures and text. Tests skipped!

@@
 \end{center}
-\caption{Liquid-vapor interface profile at room temperature.  }
+\caption{The liquid-vapor density profile at room temperature shows
+some small oscillations close to the interface.  }
 \label{fig:liquid-vapor-profile}
@@
 
-For reference, we plot the predicted vapor pressure (see
-Figure~\ref{fig:equation-of-state}) and coexistence curve (see
-Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
-match those calculated by Clark \emph{et al}. To find the densities, we use the
+For reference, we plot the coexistence curve (see
+Figure~\ref{fig:temperature-vs-density}) and the predicted vapor pressure (see
+Figure~\ref{fig:equation-of-state}) against experimental values.  These
+match those calculated by Clark \emph{et al}. The functional provides a good description of
+the vapor pressure, which has an error of less than 
+1\% of the experimental data. The coexistence curve is calculated by 
+\fixme{explanation of coexistence curve calculation?}. Near the critical temperature,
+the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
+experiment. To find the densities, we use the
 ``common tangent'' rule, as illustrated in Figure~\ref{fig:homogeneous}.
-\needsworknow{Expand the previous paragraph}
+\needsworknow{Expand the previous sentence - common tangent rule}
 
@@
 
-The fluid density at a liquid-solid interface commonly displays
-oscillatory behavior, but the presence of density oscillations at
-liquid-vapor interfaces is more debated\cite{penfold2001structure}.
-In fact, the \emph{meaning} of the density profile at a liquid-vapor
-interface is clouded by the existence of long-wavelength perturbations
-known as \emph{capillary waves}.  These waves are not accounted for in
-density-functional theory (although the \emph{exact}
-density-functional theory \emph{would} include them), with the result
-that the density profile will be smoothed out by CW effects. The liquid-vapor
-interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
-some small oscillations.
-
 One of the most computationally simple ways to study confined water is by
@@
 
+The fluid density at a liquid-solid interface commonly displays
+oscillatory behavior, but the presence of density oscillations at
+liquid-vapor interfaces is more debated\cite{penfold2001structure}.
+In fact, the \emph{meaning} of the density profile at a liquid-vapor
+interface is clouded by the existence of long-wavelength perturbations
+known as \emph{capillary waves}.  These waves are not accounted for in
+density-functional theory (although the exact
+density-functional theory would include them), with the result
+that the density profile will be smoothed out by CW effects. The liquid-vapor
+interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
+some small oscillations.
+
 We choose a chemical potential corresponding to a pressure of 1~atm and set the
@@
 
-The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:energy-1D}. Near the interface, there are oscillations around the
+The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:energy-1D}. 
+Near the interface, there are oscillations around the
 liquid density which decrease near the center of the slit. We expect the center
@@
 to liquid water between the rods. The radius is 0.5~nm, the top figure is at a separation of 0.2~nm and the
-bottom is 0.4~nm. }
+bottom is 0.4~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
+the energy for these and other separations.}
 \label{fig:density-rods}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
-\end{center}
-\caption{ Energy versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 0.9~nm. All were arbitrarily set to go to zero at large separations. The
-discontinuous transition corresponds to the phase change from
-vapor to liquid between the rods. }
-\label{fig:rods-energy-vs-distance}
 \end{figure}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\end{center}
+\caption{ Energy versus separation for two hydrophobic rods ranging in radius from
+0.3~nm to 0.9~nm. All were arbitrarily set to go to zero at large separations. The
+transition corresponds to the phase change from
+vapor to liquid between the rods as pictured in the density profiles in 
+Figure~\ref{fig:density-rods}. }
+\label{fig:rods-energy-vs-distance}
+\end{figure}
+
 We now look at the more interesting problem of two rods in water. We continue to
@@
 shows density profiles before and after this transition
-for rods of radius 0.5~nm. Since we are looking at only hard-walled
+for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
+somewhat on the radii of the rods, and is about 0.25~nm for the rods in Figure~\ref{fig:density-rods}.
+Since we are looking at only hard-walled
 non-interacting rods, the critical separation for this transition will be
 different than any system where there is an appreciable force between rods. 
-
-The free energy per length is calculated for several different radii and is plotted
-in Figure~\ref{fig:rods-energy-vs-distance}. Each data set is adjusted such that the
-free energy per length goes to zero at large separations. Once a certain separation is
-reached, the rods will simply act like the single rod in water and the energy will
-reach an asymptote proportional to the surface tension (as in Figure~\ref{fig:energy-vs-diameter}).
+The density oscillations  around the rods are very similar to
+those for the single hydrophobic rod in water, as seen in Figure~\ref{fig:density-single-rod}.
+The oscillations around the two rods at small separation looks as if the two rods were one
+solid object shaped \fixme{racetrack shaped?}.
 
@@
 and the entropic energy contribution, where all are in units of energy per length 
-(see Figure~\ref{fig:energy-rods}). \needsworknow{More discussion of Figure~\ref{fig:energy-rods}}
+(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
+goes from 0 to the critical separation (0.25~nm) and then is approximately constant at
+larger separations. Since there is no repulsive interaction between the rods, there is no
+large energy increase when the rods touch. The entropy has a minimum just before the rods
+transition to let liquid between them. As the rods separate starting from zero, the area 
+between them containing vapor increases and there is less overall volume for the 
+surrounding liquid water to occupy and thus the entropy decreases. 
+\needsworknow{More discussion of Figure~\ref{fig:energy-rods}, internal energy, how they add
+together, etc.}
 
-Walther
-\emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
-carbon nanotubes. They use molecular dynamics with Lennard-Jones potentials for
+The free energy per length is compared for rods of several different radii
+in Figure~\ref{fig:rods-energy-vs-distance}. For comparison, we arbitrarily adjust each data set such that the
+free energy per length goes to zero at large separations. At large separations, the rods will 
+simply act like two separate single rods in water and the total energy will
+reach an asymptote proportional to the surface tension (similar to that shown in 
+Figure~\ref{fig:energy-vs-diameter}).
+The change in free energy per length at the point of transition increases as the radii of
+the rods increases. The smallest radius we show is 0.3~nm, which shows only a slight increase in
+free energy at the critical separation, while the largest radius has a more significant jump in free energy. 
+We see that all the rods studied have the same behavior, with a positive linear relation between free energy
+per length and the separation during the initial state where the rods have vapor between them. 
+The slope of this line is equal to the force per length on the rods, which we calculate to be 144~mN/m, 
+or twice the surface tension
+found by the functional (see Figure~\ref{fig:surface-tension}).
+
+Walther \emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
+carbon nanotubes, which are very geometrically similar to our hydrophobic rods. 
+They use molecular dynamics with Lennard-Jones potentials for
 nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
-They also find a drying transition occurs, and find it to be at separations less
-than 0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is several times larger than
+They also find a drying transition occurs, and find it to be at separations near
+0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is several times larger than
 what we find for hydrophobic rods, as expected due to the carbon-carbon
 interations between two nanotubes.
-
-The density oscillations are very similar to
-those for the single hydrophobic rod in water, as seen in Figure~\ref{fig:density-single-rod}.
 
@@
 
-[conclusion text]
+\needsworknow{Conclusions about slit, single rod, two hydrophobic rods (mostly). future uses}
 

### 2011-09-26 paper/project.tex
Project text and bib edits. Tests skipped!

@@
 \begin{abstract}
-{We develop a classical density functional theory for water by combining the fundamental-measure theory functional--which describes hard sphere fluids--with attractive interactions based on the Statistical Associating Fluid Theory (SAFT).
-Using this approximation, we are able to describe the density, free energy, and
-entropy of water, as well as the number of hydrogen bonds formed near liquid-vapor and liquid-solid interfaces.
-These interfaces include water confined to a hard-walled slit, a single hydrophobic rod immersed in water, and two hydrophobic rods
-at various separations. We found \needsworknow{conclusion-like stuff about two rods, vapor-liquid transition between rods, constant force of twice surface tension, more?}.
-This functional provides a foundation for future work predicting the interactions of chemicals in aqueous solution on an
+{We develop a classical density functional theory for water by combining the
+fundamental-measure theory functional--which describes hard sphere fluids--with
+attractive interactions based on the Statistical Associating Fluid Theory
+(SAFT). Using this approximation, we are able to describe the vapor pressure, 
+various densities, surface tension, enthalpy and
+entropy of water, as well as the number of hydrogen bonds formed near
+liquid-vapor and liquid-solid interfaces. These interfaces include water
+confined to a hard-walled slit, a single hydrophobic rod immersed in water, and
+two hydrophobic rods 
+at various separations. We found \needsworknow{conclusion-like stuff about two
+rods, vapor-liquid transition between rods, constant force of twice surface
+tension, more?}.
+This functional provides a foundation for future work predicting the
+interactions of chemicals in aqueous solution on an
 atomic level.}
@@
 of the density function and is independent of any external
-potentials\cite{ebner1976}. The most common use of DFT is calculating ground
+potentials\cite{ebner1976density}. The most common use of DFT is calculating ground
 state electron densities in quantum mechanical systems. The Hohenberg-Kohn
@@
 the vapor pressure, which has an error of less than 
-1\% of the experimental data. The coexistence curve is calculated by 
-\fixme{explanation of coexistence curve calculation?}. Near the critical temperature,
+1\% of the experimental data. Near the critical temperature,
 the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
-experiment. To find the densities, we use the
-``common tangent'' rule, as illustrated in Figure~\ref{fig:homogeneous}.
-\needsworknow{Expand the previous sentence - common tangent rule}
+experiment. To find the densities for the coexistence curve, we use the
+``common tangent'' rule. The two densities 
+will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
+As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
+(see the bottom of Figure~\ref{fig:homogeneous}. 
+\needsworknow{Expand the previous common tangent rule stuff?}
 
@@
 a single hydrophobic rod with a radius of 0.5~nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$
-0.003~nm), 
-medium 0.2~bohrs ($\approx$ 0.001~nm), and low 0.5~bohrs ($\approx$ 0.03~nm).}
+lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
+medium 0.2~bohrs ($\approx$~0.001~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
 \label{fig:densityresolution}

### 2011-09-26 paper/project.tex

@@
 empirical parameters of Clark \emph{et al}, we add a single additional
-parameter $\sigma$ defining the length scale over which the density is averaged
-when computing the dispersion free energy.  We use this final
-parameter to fit the surface tension, with the result shown in
+parameter $\lscale$---with a fitted value of 0.72---which determines
+the length scale over which the density is averaged when computing the
+dispersion free energy.  We use this final parameter to fit the
+surface tension, with the result shown in
+>>>>>>> 92fcd98014a991adc9e2804a16040e70e2b2bfcb
 Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark

### 2011-09-26 papers/contact/paper.tex

@@
 
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-08-013}
+  \caption{Density and contact densities of thirteen hard spheres in a
+    spherical cavity with diameter 8.}
+  \label{fig:sphere-8}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-12-112}
+  \caption{Density and contact densities of 112 hard spheres in a
+    spherical cavity with diameter 12.}
+  \label{fig:sphere-12}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-16-265}
+  \caption{Density and contact densities of 265 hard spheres in a
+    spherical cavity with diameter 16.}
+  \label{fig:sphere-16}
+\end{figure}
+
 \subsection{Hard spheres confined in a cubical cavity}

### 2011-09-28 paper/project.tex
Deleted nonsense line Tests skipped!
New rods data, edits to project figures and text. Tests skipped!

@@
 (SAFT). Using this approximation, we are able to describe the vapor pressure, 
-various densities, surface tension, enthalpy and
-entropy of water, as well as the number of hydrogen bonds formed near
+vapor density, liquid density, surface tension, enthalpy and
+entropy of water, as well as the number of hydrogen bonds and density oscillations
+formed near
 liquid-vapor and liquid-solid interfaces. These interfaces include water
@@
 two hydrophobic rods 
-at various separations. We found \needsworknow{conclusion-like stuff about two
-rods, vapor-liquid transition between rods, constant force of twice surface
-tension, more?}.
-This functional provides a foundation for future work predicting the
-interactions of chemicals in aqueous solution on an
+at various separations. For two rods being pulled apart, a transition occurs 
+\fixme{at $(\pi-2)r$?} from a vapor state to a liquid state in the area between
+the rods. Even for rods of a small radius (0.3~-~0.9~nm), we find that for the
+vapor state, the force can be approximated
+classically and is equal to 144~mN/m, or twice the bulk surface tension.
+This functional is computationally efficient and provides a foundation for 
+future work describing interactions in aqueous solution on an
 atomic level.}
@@
 We introduce a new classical density functional for water that
-predicts the vapor density, liquid density, and bulk surface tension, as well 
+predicts the vapor pressure, vapor density, liquid density, bulk surface tension, 
+and enthalpy, as well 
 as allows for liquid-vapor coexistence.  The Helmholtz free energy functional is
@@
 where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
-a hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
+the hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
 and $F_\textit{assoc}$ is the free energy of association.  In
@@
 This ensures that any higher order terms in the Taylor expansion
-of $F[n+\Delta n]$ are very small.
+of $F[n+\Delta n]$ are small and do not contribute to the small
+solute free energy of solvation.
 
@@
   \mathbf{n}_{V1} &= \frac{\mathbf{n}_{V2}}{4\pi R}\\
-  n_0 &= \frac{n_2}{4\pi R^2}
+  n_0 &= \frac{n_2}{4\pi R^2}.
 \end{align}
@@
 \subsection{Dispersion free energy}
-The first term in the attractive energy is the dispersion term.  We use a
+The first term in the attractive energy is the dispersion term.  This
+will include van der Waals and other non-associative interations. We use a
 dispersion term based on the SAFT-VR
@@
 \end{align}
-where $A_1$ and $A_2$ are the first two terms in a high-temperature
-perturbation expansion.  $A_1$ is the mean-field dispersion
+where $A_1$ and $A_2$ \fixme{change all to $a_1$ and $a_2$} are the first 
+two terms in a high-temperature
+perturbation expansion and $\beta=\frac{1}{k_BT}$.  $A_1$ is the mean-field dispersion
 interaction, which reduces in the homogeneous limit to
@@
 \end{align}
+where $n_b$ is the bulk density and $g_{HS}$ is the homogeneous
+hard-sphere fluid correlation function.
 The second dispersion term in the free energy $A_2$ describes the
@@
 to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA).
+using the local compressibility approximation (LCA). In the LCA,
+we assume the energy fluctuation is directly related to the
+compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
 
@@
   \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
-             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'
+             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
 \end{align}
@@
   A_1 &= \int
-   -4*(\lambdadisp^3 - 1)*\epsilondisp \etadisp(\xx)
+   -4(\lambdadisp^3 - 1)\epsilondisp \etadisp(\xx)
     g^{HS}_\sigma(\eta_\textit{eff}(\xx))
@@
   \\
-  \zeta(\xx) &= 1 - \frac{\mathbf{n_{2V}}\cdot\mathbf{n_{2V}}}{n_2^2}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}
 \end{align}
@@
 fit to the integral in Equation~\ref{eq:A1-simple} over a range of
-values for $\lambdadisp$ and filling fractions.  The $\zeta$ factor
+values for $\lambdadisp$ and filling fractions.  Here $g_\sigma^{HS}$
+is the hard-sphere fluid correlation function evaluated at contact and uses the
+effective filling fraction for inhomogeneous density distributions.
+The $\zeta$ factor
 (also seen in the association energy, below) is a dimensionless
@@
   K_{HS} &=
-    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}
+    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}.
 \end{align}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental coexistence curve. Away from
-the critical temperature, the agreement is good. }
-\label{fig:temperature-vs-density}
-\end{figure}
 
@@
 experimental data.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
-\end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
-\end{figure}
 
@@
 \begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
 \end{align}
@@
 the free energy.  The correlation function $g^\textit{SW}_\sigma$ is
-written as a perturbative correction to the hard-sphere correlation
-function, for which we use the functional of Yu and
-Wu.\cite{yu2002fmt-dft-inhomogeneous-associating}
+written as a perturbative correction to the hard-sphere fluid correlation
+function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
 \begin{align}
@@
   +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
-  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}
-\end{align}
-
-\subsection{Other thermodynamic functions}
-
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx
-\end{equation}
-
-The pressure would naturally be obtained by taking a partial
-derivative with respect to volume at fixed temperature.  Since
-\emph{volume} isn't a parameter in density-functional theory, we
-consider the \emph{free energy per unit volume}.
-\begin{align}
-  F[n] &= f[n]V \\
-  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
-  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
-  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
-   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
-  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n]
+  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
 \end{align}
@@
 
-Figure \ref{fig:pressure-with-isotherms} shows the calculated pressures for
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
+  will match the experimental surface tension near room temperature.
+}
+\label{fig:surface-tension}
+\end{figure}
+
+\subsection{Other thermodynamic functions}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\end{center}
+\caption{The theoretical versus experimental phase diagram, vapor
+  pressure versus temperature.  }
+\label{fig:equation-of-state}
+\end{figure}
+
+From the Helmholtz free energy functional, we may obtain any other
+thermodynamic functions we should choose.  The grand free energy
+$\Phi$, which we ordinarily use in our computations, is obtained by
+simply adding a chemical potential term:
+\begin{equation}
+  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
+\end{equation}
+
+The pressure would naturally be obtained by taking a partial
+derivative with respect to volume at fixed temperature.  Since
+\emph{volume} isn't a parameter in density-functional theory, we
+consider the \emph{free energy per unit volume}:
+\begin{align}
+  F[n] &= f[n]V \\
+  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
+  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
+  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
+   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
+  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n].
+\end{align}
+
+Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
 isotherms ranging from room temperature to past the critical temperature. The
@@
 functional. For a given isotherm, the pressure curve should intersect this
-coexistence curve twice, at the values of the vapor and liquid densities. Even
+coexistence curve twice, at the vapor and liquid coexistence 
+pressure corresponding to the vapor and liquid densities. Even
 though the theoretical pressure isotherms drop below the coexistence curve,
 this is non-physical and will not actually occur. The experimental data shows
-how the pressure will jump to the other side of the coexistence curve (for
+how the actual density value will ``jump'' to the other side of 
+the coexistence curve at that pressure value (for
 temperatures less than the critical temperature), corresponding to a phase 
-change in the material. The
+change in the material. For example, follow the room temperature isotherm
+starting from high pressure and high density (top right of 
+Figure~\ref{fig:pressure-with-isotherms}).
+As the pressure decreases, the density decreases \emph{slightly} until reaching
+the coexistence curve, at which point a phase change from liquid to vapor
+will occur. The
 functional matches well with the experimental data from NIST away from the
@@
 \begin{align}
-  H[n] &= F[n] + TS[n] + pV
+  H[n] &= F[n] + TS[n] + pV.
 \end{align}
-
 The calculated heat of vaporization at atmospheric pressure is
@@
 an error of only a few percent.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
+\end{center}
+\caption{The theoretical versus experimental coexistence curve. Away from
+the critical temperature, the agreement is good. }
+\label{fig:temperature-vs-density}
+\end{figure}
 
@@
 \label{fig:homogeneous}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\sigma$ is fit so that the theoretical surface tension
-  will match the experimental surface tension near room temperature.
-}
-\label{fig:surface-tension}
 \end{figure}
@@
 surface tension, with the result shown in
->>>>>>> 92fcd98014a991adc9e2804a16040e70e2b2bfcb
 Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
@@
 water is liquid at one atmosphere of pressure.
-
-For reference, we plot the coexistence curve (see
-Figure~\ref{fig:temperature-vs-density}) and the predicted vapor pressure (see
-Figure~\ref{fig:equation-of-state}) against experimental values.  These
-match those calculated by Clark \emph{et al}. The functional provides a good description of
-the vapor pressure, which has an error of less than 
-1\% of the experimental data. Near the critical temperature,
-the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
-experiment. To find the densities for the coexistence curve, we use the
-``common tangent'' rule. The two densities 
-will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
-As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
-(see the bottom of Figure~\ref{fig:homogeneous}. 
-\needsworknow{Expand the previous common tangent rule stuff?}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and discussion}
-
-\subsection{Water confined to a slit}
 
@@
 
+For reference, we plot the predicted vapor pressure (see
+Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
+Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
+match those calculated by Clark \emph{et al}. The functional provides a good description of
+the vapor pressure, which has an error of less than 
+1\% compared with the experimental data. The coexistence curve in 
+Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
+curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
+temperature coexistence values.
+Near the critical temperature,
+the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
+experiment. 
+To find the densities for the coexistence curve, we use the
+``common tangent'' rule. The two densities 
+will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
+As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
+(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
+in correctly treating the critical temperature. As the vapor and liquid densities approach
+each other, the free energy barrier between the two states gets smaller and it becomes
+challenging to computationally find the correct minimum. This is the reason 
+why the coexistence curves
+in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
+are not smooth around the critical point.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results and discussion}
+
+\subsection{Water confined to a slit}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
+\end{center}
+\caption{The number of hydrogen bonds per molecule for a slab of water 
+with molecules constrained to move in one dimension. Each molecule may
+contribute to
+four bonds, although each bond involves two molecules so the
+actual number of bonds would be halved.}
+\label{fig:xassoc-1D}
+\end{figure}
+
 One of the most computationally simple ways to study confined water is by
@@
 
+Before looking at the liquid-solid one-dimensional slit, we set up a liquid-vapor
+interface and plot the density profile.
 The fluid density at a liquid-solid interface commonly displays
@@
 density-functional theory would include them), with the result
-that the density profile will be smoothed out by CW effects. The liquid-vapor
+that the density profile will be smoothed out by capillary wave effects. The liquid-vapor
 interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
@@
 
-We choose a chemical potential corresponding to a pressure of 1~atm and set the
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\end{center}
+\caption{ Density profiles for a single rod of different radii. The dotted line
+represents the saturated liquid density.  }
+\label{fig:density-single-rod}
+\end{figure}
+
+To set up the calculations for the slit, we choose a chemical potential 
+corresponding to a pressure of 1~atm and set the
 temperature to a typical room temperature of 298~K. We apply 
@@
 with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+\end{center}
+\caption{ Energy versus radius for a single hydrophobic rod
+immersed in water. This should have an asymptote equal to the surface
+tension at room temperature, and it agrees well with the surface tension in
+Figure~\ref{fig:surface-tension}. }
+\label{fig:energy-vs-diameter}
+\end{figure}
 
@@
 vaporize. The results for larger liquid-filled slits are only metastable,
-however,as a vapor-filled slit is thermodynamically favorable. Based on our
+however, as a vapor-filled slit is thermodynamically favorable. Based on our
 calculations for a pressure of 1~atm, the work needed to pull apart
@@
 than the free energy of the liquid-filled slit. For the range of slit sizes we
-studied (0.1 to 8.0~nm), all the computational results of liquid-filled are
+studied, all the computational results of liquid-filled slits are
 metastable. Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D}
-\end{center}
-\caption{The number of hydrogen bonds per molecule for a slab of water 
-with molecules constrained to move in one dimension. Each molecule may
-contribute to
-four bonds, although each bond involves two molecules so to calculate the
-actual 
-number of bonds there would be another factor of 2.}
-\label{fig:xassoc-1D}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-single-rod}
-\end{center}
-\caption{ Density profiles for a single rod of different radii. The dotted line
-represents the saturated liquid density.  }
-\label{fig:density-single-rod}
-\end{figure}
-
 The value of $X$ behaves as expected for water constrained to a slit. We
@@
 molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
-Outside the slit, there is no hydrogen bonding and inside the number of hydrogen
+Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
 bonds is approximately equal to 3.5, which is expected for saturated liquid
 water.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
-\end{center}
-\caption{ Energy versus radius for a single hydrophobic rod
-immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. }
-\label{fig:energy-vs-diameter}
-\end{figure}
 
@@
 about 0.5~nm or so, this reaches
-an asymptote equal to the surface tension. From the inset in Figure~\ref{fig:surface-tension}, the surface tension at room 
-temperature is approximately 72~mN/m. This is very close to the value from
-the larger diameter single rod energy per area, which is about 75~mN/m.
+an asymptote equal to the surface tension. From the inset in 
+Figure~\ref{fig:surface-tension}, the surface tension at room 
+temperature is approximately 72~mN/m. This is very close to the 
+asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
 
@@
 rod (less than 0.2~nm) there are no molecules to form hydrogen bonds although the value
-of $X$ is still calculated there. Walther \emph{et al.}\cite{walther2001carbon} 
+of $X$ is still calculated there. The number of bonds decreases from a bulk value of 3.5 
+to about 3.25 at the interface. Walther \emph{et al.}\cite{walther2001carbon} 
 calculate the number of hydrogen bonds for a carbon nanotube immersed in water.
 They show a decrease in the bonding from a bulk value of 3.75 to 2.89 near the 
-interface with the nanotube. This is a much larger change than for our hydrophobic
-rod of a similar radius, where the number of bonds decreases from a bulk value of 3.5 
-to about 3.25 at the interface (see Figure~\ref{fig:xassoc-single-rod}).
+interface with the nanotube. This is a much larger change than for our 
+rod of a similar radius, indicating that carbon nanotubes disrupt hydrogen
+bonds in water more than hydrophobic rods.
 
@@
 about 5\%. There is some uncertainty in the value of contact density due to the
-different lattice spacings. The total free energy calculations compare much more
+different lattice spacings which cause the contact density to be calculated
+at a slightly different position. The total free energy calculations compare much more
 favorably. The medium resolution is within 0.5\% and the low resolution is
@@
 \caption{ Density profiles illustrating the transition from vapor 
-to liquid water between the rods. The radius is 0.5~nm, the top figure is at a separation of 0.2~nm and the
-bottom is 0.4~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
+to liquid water between the rods. The radius is 0.5~nm, the top figure is 
+at a separation of 0.2~nm and the
+bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
 the energy for these and other separations.}
@@
 \end{figure}
+
+We now look at the more interesting problem of two rods in water. We continue to
+use hydrophobic rods of various radii at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+(touching rods) to $d=1$~nm. At small
+separations there is only vapor between the rods. As the rods are
+pulled apart, the vapor region expands until some critical separation is
+reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
+shows density profiles before and after this transition
+for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
+on the radii of the rods, and is about 0.55~nm for the rods shown in 
+Figure~\ref{fig:density-rods}.
+\fixme{add analysis of circumferences of stadium shape vs two rods and
+calculating the expected critical separation}
+Since we are looking at only hard-walled
+non-interacting rods, the critical separation will be
+different for any system where there is an appreciable force between rods. 
+The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
+are very similar to
+those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
+The oscillations around the two rods at small separation looks as if the two rods were one
+solid stadium shaped object (approximately like a rectangle with semi-circles on the ends).
 
@@
 
+For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
+and the entropic energy contribution, where all are in units of energy per length 
+(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
+goes from zero to the critical separation and then is approximately constant at
+larger separations. The entropy has a minimum just before the rods
+transition to allow liquid between them. As the rods separate starting from zero, the area 
+between them containing vapor increases and there is less overall space for the 
+surrounding liquid water to occupy and thus the entropy decreases. When that area fills suddenly
+with liquid, the entropy increases due to the increased space for the water molecules. The
+internal energy also has a minimum just before the transition to the liquid filled state.
+
 \begin{figure}
@@
 \caption{ Energy versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 0.9~nm. All were arbitrarily set to go to zero at large separations. The
+0.3~nm to 0.9~nm. \fixme{r=0.9 to be added when it finishes!}
+All were arbitrarily set to go to zero at large separations. The
 transition corresponds to the phase change from
@@
 
-We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of various radii at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
-(touching rods) to about 1~nm. At small
-separations ($d<2$~-~4~nm) there is only vapor between the rods. As the rods are
-pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
-shows density profiles before and after this transition
-for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
-somewhat on the radii of the rods, and is about 0.25~nm for the rods in Figure~\ref{fig:density-rods}.
-Since we are looking at only hard-walled
-non-interacting rods, the critical separation for this transition will be
-different than any system where there is an appreciable force between rods. 
-The density oscillations  around the rods are very similar to
-those for the single hydrophobic rod in water, as seen in Figure~\ref{fig:density-single-rod}.
-The oscillations around the two rods at small separation looks as if the two rods were one
-solid object shaped \fixme{racetrack shaped?}.
-
-For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
-and the entropic energy contribution, where all are in units of energy per length 
-(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
-goes from 0 to the critical separation (0.25~nm) and then is approximately constant at
-larger separations. Since there is no repulsive interaction between the rods, there is no
-large energy increase when the rods touch. The entropy has a minimum just before the rods
-transition to let liquid between them. As the rods separate starting from zero, the area 
-between them containing vapor increases and there is less overall volume for the 
-surrounding liquid water to occupy and thus the entropy decreases. 
-\needsworknow{More discussion of Figure~\ref{fig:energy-rods}, internal energy, how they add
-together, etc.}
-
 The free energy per length is compared for rods of several different radii
-in Figure~\ref{fig:rods-energy-vs-distance}. For comparison, we arbitrarily adjust each data set such that the
+in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for a radius 0.5~nm
+is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
+For comparison, we arbitrarily adjust each data set such that the
 free energy per length goes to zero at large separations. At large separations, the rods will 
@@
 Figure~\ref{fig:energy-vs-diameter}).
-The change in free energy per length at the point of transition increases as the radii of
-the rods increases. The smallest radius we show is 0.3~nm, which shows only a slight increase in
-free energy at the critical separation, while the largest radius has a more significant jump in free energy. 
-We see that all the rods studied have the same behavior, with a positive linear relation between free energy
-per length and the separation during the initial state where the rods have vapor between them. 
-The slope of this line is equal to the force per length on the rods, which we calculate to be 144~mN/m, 
-or twice the surface tension
-found by the functional (see Figure~\ref{fig:surface-tension}).
+We see that all the rods studied have the same behavior, with a positive linear relation 
+between free energy
+per length and separation during the initial state where the rods have vapor between them. 
+The slope of this line is equal to the force per length on the rods, which we calculate to 
+be 144~mN/m, or twice the surface tension
+found by the functional (see Figure~\ref{fig:surface-tension}). This indicates that
+even though the rods are nanoscale, to the water molecules they can be treated classically
+\fixme{reference above equations/explanation with circumference, critical separation, etc.}
 
@@
 They also find a drying transition occurs, and find it to be at separations near
-0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is several times larger than
-what we find for hydrophobic rods, as expected due to the carbon-carbon
+0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than
+what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
 interations between two nanotubes.
@@
 
-\needsworknow{Conclusions about slit, single rod, two hydrophobic rods (mostly). future uses}
+\needsworknow{Conclusions about slit, single rod, two hydrophobic rods (mostly). 
+computationally efficient. future uses}
 

### 2011-09-30 paper/project.tex
Fixing rods-energy dat files and figure, adding conclusion text, other text edits. Tested everything.

@@
 (SAFT). Using this approximation, we are able to describe the vapor pressure, 
-vapor density, liquid density, surface tension, enthalpy and
+saturated liquid density, surface tension, enthalpy and
 entropy of water, as well as the number of hydrogen bonds and density oscillations
@@
 at various separations. For two rods being pulled apart, a transition occurs 
-\fixme{at $(\pi-2)r$?} from a vapor state to a liquid state in the area between
-the rods. Even for rods of a small radius (0.3~-~0.9~nm), we find that for the
+from a vapor state to a liquid state in the area between
+the rods at approximately $(\pi-2)$ multipled by the radius. Even 
+for these nanoscale rods (0.3~-~1.0~nm), we find that in the
 vapor state, the force can be approximated
@@
 \end{align}
-where $A_1$ and $A_2$ \fixme{change all to $a_1$ and $a_2$} are the first 
+where $a_1$ and $a_2$ \fixme{change all to $a_1$ and $a_2$} are the first 
 two terms in a high-temperature
-perturbation expansion and $\beta=\frac{1}{k_BT}$.  $A_1$ is the mean-field dispersion
+perturbation expansion and $\beta=1/k_BT$.  The first term $a_1$ is 
+the mean-field dispersion
 interaction, which reduces in the homogeneous limit to
@@
 hard-sphere fluid correlation function.
-The second dispersion term in the free energy $A_2$ describes the
+The second dispersion term in the free energy $a_2$ describes the
 effect of fluctuations resulting from the compression of the fluid due
@@
 
-The form of $A_1$ is given in
+The form of $a_1$ is given in
 reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
@@
 
-The first term in the dispersion functional $A_1$ when written using
+The first term in the dispersion functional $a_1$ when written using
 the above filling fraction for dispersion has the form
@@
 
-The second term, $A_2$, which describes the contribution to the free
+The second term, $a_2$, which describes the contribution to the free
 energy associated with fluctuations is given by
@@
 oscillations supports this. For larger slits (6.0~nm or more) the majority of
-the water molecules in the slit act like the bulk. Based on our computations, we
+the water molecules in the slit act like the bulk. \fixme{Either add brief
+description of bottom half of Figure~\ref{fig:energy-1D} or delete it}
+Based on our computations, we
 also determine the smallest slit that can contain liquid water. This result is
@@
 We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of various radii at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
 (touching rods) to $d=1$~nm. At small
@@
 Figure~\ref{fig:density-rods}. 
-\fixme{add analysis of circumferences of stadium shape vs two rods and
-calculating the expected critical separation}
 Since we are looking at only hard-walled
@@
 \caption{ Energy versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 0.9~nm. \fixme{r=0.9 to be added when it finishes!}
+0.3~nm to 1.0~nm.
 All were arbitrarily set to go to zero at large separations. The
@@
 found by the functional (see Figure~\ref{fig:surface-tension}). This indicates that
-even though the rods are nanoscale, to the water molecules they can be treated classically
-\fixme{reference above equations/explanation with circumference, critical separation, etc.}
+even though the rods are nanoscale, to the water molecules they can be seen classically.
+To estimate how the critical transition depends on the radius,
+we assume a classical approximation where the energy on the rod is equal to the
+surface area multiplied by the bulk surface tension. The surface area is just the
+circumference multiplied by the width, and this width has the same value for both the
+vapor-filled stadium shape and the liquid-filled separate rods. We propose that 
+the critical separation
+for the transition, $d_c$, will occur when the circumference of the stadium shape (see 
+top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom
+of Figure~\ref{fig:density-rods}) are equal:
+\begin{align}
+C_{stadium} &= C_{rods} \\
+2\pi r+4r+2d_c &= 4\pi r \\
+d_c &= (\pi-2)r.\label{criticalseparation}
+\end{align}
+
+\begin{table}
+\begin{tabular} {|c|c|c|c|}
+\hline
+$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
+\hline
+0.3 & 0.34 & 0.25 & 26\% \\
+\hline
+0.5 & 0.57 & 0.55 & 3.5\% \\
+\hline
+0.7 & 0.80 & 0.83 & 3.8\% \\
+\hline
+0.9 & 1.02 & 1.07 & 4.9\% \\
+\hline
+1.0 & 1.14 & 1.18 & 3.5\% \\
+\hline
+\end{tabular}
+\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
+for the transition
+from liquid to vapor between two hydrophobic rods.}
+\label{criticalseparationtable}
+\end{table}
+The predicted and actual computed values for $d_c$ are shown in 
+Table~\ref{criticalseparationtable},
+and they match fairly well for the 0.5~-~1.0~nm rods. We hypothesize that
+larger radii rods
+would also follow this trend, but smaller rods are likely to have a different
+relationship with the surrounding water molecules. This may explain why the critical
+separation result for a radius of 0.3~nm does not match the prediction well.
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 \section{Conclusion}
 
-\needsworknow{Conclusions about slit, single rod, two hydrophobic rods (mostly). 
-computationally efficient. future uses}
+The classical density functional we developed by combining fundamental-measure 
+theory with SAFT is able to produce accurate results for hydrophobic interations.
+\fixme{decribe any results for slit?}
+We added the parameter \lscale to the functional and fit it to the surface tension
+at room temperature, which we find to be consistent with results for hydrophobic
+rods. The free energy per area for water surrounding a hydrophobic rod shows the
+expected behavior with an asymptote equal to the surface tension.
+Based on this result, we expect to be
+able to describe classically rods with a radius of about 0.5~nm or larger. Even
+though this is still nanoscale, from the perspective of the water molecules
+the interaction with a hydrophobic rod is easily approximated classically.
+For the interactions between water and two hydrophobic rods, we see that a
+classical treatment of the phase transition critical separation works well
+for rods larger than 0.5~nm, within 5\%. This transition from a vapor-filled
+stadium shape to two rods with liquid between them (see Figure~\ref{fig:density-rods})
+is commonly seen for objects of a similar geometry. Further evidence for the
+accuracy of this classical treatment is the calculation of the force on the
+stadium shape. The force per length is constant and equal to twice the surface tension.
+
+This functional produces many of the results of \emph{ab initio} calculations
+while being much more computationally efficient. This allows for future work
+involving hydrophilic interactions with more complex electrostatic potentials.
+\fixme{more about efficiency or future work?}
 

### 2011-10-01 paper/project.tex
Fixed dispersion, two rods, and conclusion. Tested everything.

@@
 the rods at approximately $(\pi-2)$ multipled by the radius. Even 
-for these nanoscale rods (0.3~-~1.0~nm), we find that in the
+for these nanoscale rods (radius 0.5~-~1.0~nm), we find that in the
 vapor state, the force can be approximated
@@
 the \emph{fundamental measures} of a fluid, which provide local
-measures of quantities such as the packing fraction, density of
+measures of quantities such as the filling fraction, density of
 spheres touching a given point and mean curvature.  The excess free
@@
 \end{align}
-For our water functional, we use the hard-sphere radius of
+The density $n_3$ is the filling fraction and $n_2$ describes the number
+of spheres touching a given point. For our water functional, we use the 
+hard-sphere radius of
 3.03420~\AA, which was found to be optimal by Clark
@@
 \begin{align}
-  F_\text{disp} &= A_1 + \beta A_2
+  F_\text{disp}[n] &= \int (a_1(\xx) + \beta a_2(\xx))d\xx
 \end{align}
-where $a_1$ and $a_2$ \fixme{change all to $a_1$ and $a_2$} are the first 
-two terms in a high-temperature
+where $a_1$ and $a_2$ are the first two terms in a high-temperature
 perturbation expansion and $\beta=1/k_BT$.  The first term $a_1$ is 
@@
 \begin{align}\label{eq:A1-simple}
-  \frac{A_1}{V} &= \frac12 n_b \int \varphi(\left|\xx\right|)
+  a_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
   g_{HS}(\left|\xx\right|) d\xx
@@
 \begin{align}
-  A_1 &= \int
+  a_1 &= 
    -4(\lambdadisp^3 - 1)\epsilondisp \etadisp(\xx)
-    g^{HS}_\sigma(\eta_\textit{eff}(\xx))
-  d\xx \\
-  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
-  \lambdadisp^j
-  \\
+    g^{HS}_\sigma(\eta_\textit{eff}(\xx)) \\
   g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
@@
   \\
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}\\
+  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
+  \lambdadisp^j
 \end{align}
-where $C_{ij}$ are numerical constants taken from
+where $g_\sigma^{HS}$
+is the hard-sphere fluid correlation function evaluated at contact.
+The $\zeta$ factor
+(also seen in the association energy, below) is a dimensionless
+measure of the local density inhomogeneity. The $C_{ij}$ values
+ are numerical constants taken from
 reference\cite{gil-villegas-1997-SAFT-VR}, which come from a numerical
 fit to the integral in Equation~\ref{eq:A1-simple} over a range of
-values for $\lambdadisp$ and filling fractions.  Here $g_\sigma^{HS}$
-is the hard-sphere fluid correlation function evaluated at contact and uses the
-effective filling fraction for inhomogeneous density distributions.
-The $\zeta$ factor
-(also seen in the association energy, below) is a dimensionless
-measure of the local density inhomogeneity.
+values for $\lambdadisp$ and filling fractions.
 
@@
 \begin{align}
-  A_2 &= \int \frac12 \epsilondisp
-              K_{HS}(\etadisp(\xx))
-              \etadisp(\xx)
-              \frac{\delta A_1}{\delta \etadisp(\xx)}
-                d\xx
+  a_2 &= \frac12 \epsilondisp
+              K_{HS}(\etadisp(\xx)) \etadisp(\xx)
+              \frac{\partial a_1}{\partial \etadisp(\xx)}
 \end{align}
@@
   g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\etadisp(\xx)) +
-  \frac{1}{4}\beta\left(\frac{\delta A_1}{\delta \etadisp(\xx)} -
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
   \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)
@@
 
+\subsection{Other thermodynamic functions}
+
+From the Helmholtz free energy functional, we may obtain any other
+thermodynamic functions we should choose.  The grand free energy
+$\Phi$, which we ordinarily use in our computations, is obtained by
+simply adding a chemical potential term:
+\begin{equation}
+  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
+\end{equation}
+
 \begin{figure}
@@
 \end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
-  will match the experimental surface tension near room temperature.
-}
-\label{fig:surface-tension}
-\end{figure}
-
-\subsection{Other thermodynamic functions}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
-\end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
-\end{figure}
-
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
-\end{equation}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
+  will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\end{center}
+\caption{The theoretical versus experimental phase diagram, vapor
+  pressure versus temperature.  }
+\label{fig:equation-of-state}
+\end{figure}
+
 The entropy is naturally obtained by taking a partial derivative with
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental coexistence curve. Away from
-the critical temperature, the agreement is good. }
-\label{fig:temperature-vs-density}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
-\includegraphics[width=\columnwidth]{figs/near-critical-point}
-\end{center}
-\caption{By plotting the free energy versus density it is possible to
-  determine a saturated liquid or vapor state by the common tangent
-  method. The top plot illustrates the common tangents near room 
-  temperature, while the bottom plot shows the same thing close to the 
-  critical temperature. }
-\label{fig:homogeneous}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{The liquid-vapor density profile at room temperature shows
-some small oscillations close to the interface.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
-
 \subsection{Determining the empirical parameters}\label{sec:empirical}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D}
-\includegraphics[width=\columnwidth]{figs/energy-1D}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
 \end{center}
-\caption{ Density profile (top) and 
-energy densities (bottom) for a slab of water with molecules 
-constrained to move in one dimension. The free energy, internal energy
-and entropic energy densities are shown for a slit size of 4.0~nm. The dashed
-line in the top figure represents the saturated liquid density.}
-\label{fig:energy-1D}
+\caption{The theoretical versus experimental coexistence curve. Away from
+the critical temperature, the agreement is good. }
+\label{fig:temperature-vs-density}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
+\includegraphics[width=\columnwidth]{figs/near-critical-point}
+\end{center}
+\caption{By plotting the free energy versus density it is possible to
+  determine a saturated liquid or vapor state by the common tangent
+  method. The top plot illustrates the common tangent near room 
+  temperature, while the bottom plot shows the same thing close to the 
+  critical temperature. }
+\label{fig:homogeneous}
 \end{figure}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D}
-\end{center}
-\caption{The number of hydrogen bonds per molecule for a slab of water 
-with molecules constrained to move in one dimension. Each molecule may
-contribute to
-four bonds, although each bond involves two molecules so the
-actual number of bonds would be halved.}
-\label{fig:xassoc-1D}
-\end{figure}
-
 One of the most computationally simple ways to study confined water is by
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\includegraphics[width=\columnwidth]{figs/surface-298}
 \end{center}
-\caption{ Density profiles for a single rod of different radii. The dotted line
-represents the saturated liquid density.  }
-\label{fig:density-single-rod}
+\caption{The liquid-vapor density profile at room temperature shows
+some small oscillations close to the interface.  }
+\label{fig:liquid-vapor-profile}
 \end{figure}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+\includegraphics[width=\columnwidth]{figs/density-1D}
 \end{center}
-\caption{ Energy versus radius for a single hydrophobic rod
-immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. }
-\label{fig:energy-vs-diameter}
+\caption{ Density profile for a slab of water with molecules 
+constrained to move in one dimension. The slit size is 4.0~nm and the dashed
+line represents the saturated liquid density.}
+\label{fig:density-1D}
 \end{figure}
 
-The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:energy-1D}. 
+The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
 Near the interface, there are oscillations around the
@@
 oscillations supports this. For larger slits (6.0~nm or more) the majority of
-the water molecules in the slit act like the bulk. \fixme{Either add brief
-description of bottom half of Figure~\ref{fig:energy-1D} or delete it}
-Based on our computations, we
-also determine the smallest slit that can contain liquid water. This result is
+the water molecules in the slit act like the bulk.
+
+Based on our computations, we determine the smallest slit that can contain 
+liquid water. This result is
 approximately 0.6~nm. For smaller slits, the initial liquid water inside will
@@
 studied, all the computational results of liquid-filled slits are
-metastable. Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
+metastable.
+
+Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
 stability of water between large hydrophobic plates and the critical separation
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
+\end{center}
+\caption{The number of hydrogen bonds per molecule for a slab of water 
+with molecules constrained to move in one dimension. Each molecule may
+contribute to
+four bonds, although each bond involves two molecules so the
+actual number of bonds would be halved.}
+\label{fig:xassoc-1D}
+\end{figure}
+
 The value of $X$ behaves as expected for water constrained to a slit. We
@@
 water.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\end{center}
+\caption{ Density profiles for a single rod of different radii. The dotted line
+represents the saturated liquid density.  }
+\label{fig:density-single-rod}
+\end{figure}
+
+\subsection{One hydrophobic rod}
+
+We move into two dimensions by studying a single hydrophobic rod
+immersed in water. Figure~\ref{fig:density-single-rod} shows density
+profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
+expected oscillations outside the rod and as the distance away from the rod
+increases
+the density approaches that of the bulk saturated liquid density. The small
+radius
+rod has comparatively larger oscillations and a higher contact density.
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-vs-radius}
+\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
 \end{center}
-\caption{ Density profiles for 
-a single hydrophobic rod with a radius of 0.5~nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
-medium 0.2~bohrs ($\approx$~0.001~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
-\label{fig:densityresolution}
+\caption{ Free nergy versus radius for a single hydrophobic rod
+immersed in water. This should have an asymptote equal to the surface
+tension at room temperature, and it agrees well with the surface tension in
+Figure~\ref{fig:surface-tension}. }
+\label{fig:energy-vs-diameter}
 \end{figure}
-
-\subsection{One hydrophobic rod}
-We move into two dimensions by studying a single hydrophobic rod
-immersed in water. Figure~\ref{fig:density-single-rod} shows density
-profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
-expected oscillations outside the rod and as the distance away from the rod
-increases
-the density approaches that of the bulk saturated liquid density. The small
-radius
-rod has comparatively larger oscillations and a higher contact density.
-
-Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
-radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
-about 0.5~nm or so, this reaches
-an asymptote equal to the surface tension. From the inset in 
-Figure~\ref{fig:surface-tension}, the surface tension at room 
-temperature is approximately 72~mN/m. This is very close to the 
-asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
 
@@
 bonds in water more than hydrophobic rods.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-vs-radius}
+\end{center}
+\caption{ Density profiles for 
+a single hydrophobic rod with a radius of 0.5~nm using several different 
+lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
+medium 0.2~bohrs ($\approx$~0.001~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
+\label{fig:densityresolution}
+\end{figure}
+
+Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
+radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
+about 0.5~nm or so, this reaches
+an asymptote equal to the surface tension. From the inset in 
+Figure~\ref{fig:surface-tension}, the surface tension at room 
+temperature is approximately 72~mN/m. This is very close to the 
+asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
 
@@
 
+We now look at the more interesting problem of two rods in water. We continue to
+use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+(touching rods) to $d=1.3$~nm. At small
+separations there is only vapor between the rods. As the rods are
+pulled apart, the vapor region expands until some critical separation is
+reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
+shows density profiles before and after this transition
+for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
+on the radii of the rods, and is about 0.54~nm for the rods shown in 
+Figure~\ref{fig:density-rods}. 
+Since we are looking at only hard-walled
+non-interacting rods, the critical separation will be
+different for any system where there is an appreciable force between rods. 
+The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
+are very similar to
+those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
+At small separations, the oscillations around the two rods look as if the two rods were one
+solid ``stadium'' shaped object (a rectangle with semi-circles on the ends).
+
 \begin{figure}
@@
 \label{fig:density-rods}
-\end{figure}
-
-We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
-(touching rods) to $d=1$~nm. At small
-separations there is only vapor between the rods. As the rods are
-pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
-shows density profiles before and after this transition
-for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
-on the radii of the rods, and is about 0.55~nm for the rods shown in 
-Figure~\ref{fig:density-rods}. 
-Since we are looking at only hard-walled
-non-interacting rods, the critical separation will be
-different for any system where there is an appreciable force between rods. 
-The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
-are very similar to
-those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
-The oscillations around the two rods at small separation looks as if the two rods were one
-solid stadium shaped object (approximately like a rectangle with semi-circles on the ends).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-rods}
-\end{center}
-\caption{ Free energy, internal energy, and the entropic energy contribution versus
-separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
-set to go to zero at large separations. }
-\label{fig:energy-rods}
 \end{figure}
@@
 \begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-rods}
+\end{center}
+\caption{ Free energy, internal energy, and the entropic energy contribution versus
+separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
+offset to zero at large separations. }
+\label{fig:energy-rods}
+\end{figure}
+
+The free energy per length is compared for rods of several different radii
+in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for radius 0.5~nm
+is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
+For comparison, we arbitrarily offset each free energy such that they all go 
+to zero at large separations. For large $d$, the rods will 
+simply act like two separate single rods in water and the energy will
+reach an asymptote determined by the effective surface tension (shown in 
+Figure~\ref{fig:energy-vs-diameter}). For small $d$, we see that all the rods 
+have the same behavior, with a positive linear relationship
+between free energy per length and separation during the initial state 
+(where the rods have vapor between them). This indicates that the force 
+per length on the rods is constant, and surprisingly not dependent on the separation. 
+By finding the slope, we calculate this force per length to be 144~mN/m.
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
 \end{center}
-\caption{ Energy versus separation for two hydrophobic rods ranging in radius from
+\caption{ Free energy (also known as the potential of mean force) 
+versus separation for two hydrophobic rods ranging in radius from
 0.3~nm to 1.0~nm.
-All were arbitrarily set to go to zero at large separations. The
+All were arbitrarily offset to zero at large separations. The
 transition corresponds to the phase change from
@@
 
-The free energy per length is compared for rods of several different radii
-in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for a radius 0.5~nm
-is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
-For comparison, we arbitrarily adjust each data set such that the
-free energy per length goes to zero at large separations. At large separations, the rods will 
-simply act like two separate single rods in water and the total energy will
-reach an asymptote proportional to the surface tension (similar to that shown in 
-Figure~\ref{fig:energy-vs-diameter}).
-We see that all the rods studied have the same behavior, with a positive linear relation 
-between free energy
-per length and separation during the initial state where the rods have vapor between them. 
-The slope of this line is equal to the force per length on the rods, which we calculate to 
-be 144~mN/m, or twice the surface tension
-found by the functional (see Figure~\ref{fig:surface-tension}). This indicates that
-even though the rods are nanoscale, to the water molecules they can be seen classically.
-To estimate how the critical transition depends on the radius,
-we assume a classical approximation where the energy on the rod is equal to the
-surface area multiplied by the bulk surface tension. The surface area is just the
-circumference multiplied by the width, and this width has the same value for both the
-vapor-filled stadium shape and the liquid-filled separate rods. We propose that 
-the critical separation
-for the transition, $d_c$, will occur when the circumference of the stadium shape (see 
+If we use a classical approximation where the energy on the rod is equal to 
+the surface area multiplied by the bulk surface tension, then the energy per 
+length is equal to the circumference multiplied by the surface tenson. The force
+per length is the derivative of this with respect to the separation. We approximate 
+the circumference of the stadium-shape by
+\begin{equation}
+C_{s} = 2\pi r +4r+2d
+\end{equation}
+and so the force per length is equal to $2\gamma$. This 
+is consistent with the calculation of the force per length from 
+Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
+in Figure~\ref{fig:surface-tension} (72~mN/m). This indicates that
+even though the rods are nanoscale, the water molecules will ``see'' them classically.
+
+To estimate how the transition from vapor to liquid between the rods depends on their 
+radius, we will use this classical approximation and propose that 
+the critical separation, $d_c$, will occur when the free energies are equal. This 
+corresponds to equating the circumference of the stadium shape (see 
 top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom
-of Figure~\ref{fig:density-rods}) are equal:
-\begin{align}
-C_{stadium} &= C_{rods} \\
-2\pi r+4r+2d_c &= 4\pi r \\
-d_c &= (\pi-2)r.\label{criticalseparation}
-\end{align}
-
-\begin{table}
-\begin{tabular} {|c|c|c|c|}
-\hline
-$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
-\hline
-0.3 & 0.34 & 0.25 & 26\% \\
-\hline
-0.5 & 0.57 & 0.55 & 3.5\% \\
-\hline
-0.7 & 0.80 & 0.83 & 3.8\% \\
-\hline
-0.9 & 1.02 & 1.07 & 4.9\% \\
-\hline
-1.0 & 1.14 & 1.18 & 3.5\% \\
-\hline
-\end{tabular}
-\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
-for the transition
-from liquid to vapor between two hydrophobic rods.}
-\label{criticalseparationtable}
-\end{table}
+of Figure~\ref{fig:density-rods}), which results in
+\begin{equation}
+d_c = (\pi-2)r.\label{criticalseparation}
+\end{equation}
 The predicted and actual computed values for $d_c$ are shown in 
-Table~\ref{criticalseparationtable},
-and they match fairly well for the 0.5~-~1.0~nm rods. We hypothesize that
-larger radii rods
+Table~\ref{criticalseparationtable}, and they match fairly well for the 
+0.5~-~1.0~nm rods. We hypothesize that larger radii rods
 would also follow this trend, but smaller rods are likely to have a different
@@
 what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
-interations between two nanotubes.
+and carbon-water interations.
+
+\begin{table}
+\begin{tabular} {|c|c|c|c|}
+\hline
+$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
+\hline
+0.3 & 0.34 & 0.23 & 32\% \\
+\hline
+0.5 & 0.57 & 0.54 & 5.3\% \\
+\hline
+0.7 & 0.80 & 0.81 & 1.3\% \\
+\hline
+0.9 & 1.03 & 1.06 & 2.9\% \\
+\hline
+1.0 & 1.14 & 1.19 & 4.4\% \\
+\hline
+\end{tabular}
+\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
+for the transition
+from liquid to vapor between two hydrophobic rods.}
+\label{criticalseparationtable}
+\end{table}
 
@@
 The classical density functional we developed by combining fundamental-measure 
-theory with SAFT is able to produce accurate results for hydrophobic interations.
-\fixme{decribe any results for slit?}
-We added the parameter \lscale to the functional and fit it to the surface tension
-at room temperature, which we find to be consistent with results for hydrophobic
-rods. The free energy per area for water surrounding a hydrophobic rod shows the
+theory with SAFT is able to produce reasonable results for hydrophobic interations.
+We added the parameter $\lscale$ to the functional and fit it to the surface tension
+at room temperature, which we find to be consistent with results for larger hydrophobic
+rods.
+
+The free energy per area for water surrounding a hydrophobic rod shows the
 expected behavior with an asymptote equal to the surface tension.
 Based on this result, we expect to be
-able to describe classically rods with a radius of about 0.5~nm or larger. Even
-though this is still nanoscale, from the perspective of the water molecules
-the interaction with a hydrophobic rod is easily approximated classically.
-For the interactions between water and two hydrophobic rods, we see that a
-classical treatment of the phase transition critical separation works well
-for rods larger than 0.5~nm, within 5\%. This transition from a vapor-filled
-stadium shape to two rods with liquid between them (see Figure~\ref{fig:density-rods})
-is commonly seen for objects of a similar geometry. Further evidence for the
+able to classically describe rods with a radius of about 0.5~nm or larger. 
+For two hydrophobic rods surrounded by water, we see a transition from a vapor-filled
+stadium shape to two rods with liquid between them. A
+classical treatment of the critical separation for this transition
+works well for rods larger than 0.5~nm. Further evidence for the
 accuracy of this classical treatment is the calculation of the force on the
@@
 
-This functional produces many of the results of \emph{ab initio} calculations
-while being much more computationally efficient. This allows for future work
+For the hydrophobic systems studied, this functional is more computationally 
+efficient than molecular dynamics. This allows for future work
 involving hydrophilic interactions with more complex electrostatic potentials.
-\fixme{more about efficiency or future work?}
 

### 2011-10-04 paper/presentation.tex
A few final edits to project and creation of presentation. Tested everything.

--- 2011-06-06 paper/mini-presentation.tex
@@
 
-\title{A Classical Density-Functional Theory for Describing Confined Water in One, Two and Three(?) Dimensions}
-\author{Jessica Gallagher}
-\date{May 25, 2011}
+\title{A Classical Density-Functional Theory for Describing Water Interfaces}
+\author{Jessica Hughes}
+\date{October 11, 2011}
 
@@
 \subsection*{}
+
+\begin{frame}{Why Water is Cool}
+\end{frame}
+
+\begin{frame}{Density Functional Theory}
+\end{frame}
+
+\begin{frame}{Statistical Associating Fluid Theory}
+\end{frame}
 
@@
 \begin{equation}
-  F_\text{id}[n] = k_B T \int n(\textbf{x})\left( \ln{n(\textbf{x})} - 1\right) d\textbf{x}
+  F_\text{id}[n] = k_B T \int n(\textbf{r})\left( \ln{n(\textbf{r})} - 1\right) d\textbf{r}
 \end{equation}
@@
 each water molecule as a hard sphere with a radius approximately that of
-a water molecule $(R~=~2.3~bohr)$.
+a water molecule $(R~=~3~\AA)$.
+\end{frame}
+
+\begin{frame}[fragile]{Dispersion Energy}
+The dispersion free energy is based on the SAFT-VR
+approach, which has two free
+parameters, an interaction energy $\epsilon_\text{disp}$ and a
+length scale $\lambda_\text{disp}$. It has the form
+\begin{align}
+  F_\text{disp} &= 
+\end{align}
+where $a_1$ and $a_2$ are the first two terms in a high-temperature
+perturbation expansion.  $a_1$ is the mean-field dispersion
+interaction and $a_2$ describes the
+effect of fluctuations resulting from the compression of the fluid due
+to the dispersion interaction itself, and is commonly approximated
+using the local compressibility approximation (LCA).
 \end{frame}
@@
 \begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\textbf{x})
-  \left(\ln X(\textbf{x}) - \frac{X(\textbf{x})}{2} + \frac12\right) d\textbf{x}
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\textbf{r})
+  \left(\ln X(\textbf{r}) - \frac{X(\textbf{r})}{2} + \frac12\right) d\textbf{r}
 \end{align}
@@
 
-\begin{frame}[fragile]{Dispersion Energy}
-The dispersion free energy is based on the SAFT-VR
-approach, which has two free
-parameters, an interaction energy $\epsilon_\text{disp}$ and a
-length scale $\lambda_\text{disp}$. It has the form
-\begin{align}
-  F_\text{disp} &= A_1 + \beta A_2
-\end{align}
-where $A_1$ and $A_2$ are the first two terms in a high-temperature
-perturbation expansion.  $A_1$ is the mean-field dispersion
-interaction and $A_2$ describes the
-effect of fluctuations resulting from the compression of the fluid due
-to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA).
-\end{frame}
-
 \begin{frame}[fragile]{Other Thermodynamic Functions}
@@
  \item Grand free energy
+ \item Pressure
  \item Internal energy
  \item Entropy
- \item Pressure
+ \item Enthalpy
 \end{itemize}
@@
 
-\section{Results}
+\section{Testing the Functional}
 \subsection*{}
@@
 
-\begin{frame}
+\begin{frame}[fragile]{Surface Tension}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surace-tension}
+\end{center}
+\end{figure} 
+\end{frame}
+
+\begin{frame}[fragile]{Vapor Pressure}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/equation-of-state}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}[fragile]{Entropy, Internal Energy, Enthalpy}
+[Add equations, calculations?]
+\end{frame}
+
+\begin{frame}[fragile]{Temperature Coexistence Curve}
 \begin{figure}
@@
 \includegraphics[width=\columnwidth]{figs/temperature-versus-density}
+\end{center}
+\end{figure}
+\end{frame}
+
+\begin{frame}[fragile]{Determining Densities - Common Tangent Rule}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}
+%\includegraphics[width=\columnwidth]{figs/near-critical-point}
+\end{center}
+\end{figure}
+\end{frame}
+
+\section{Results}
+\subsection*{}
+
+\begin{frame}[fragile]{Liquid-Vapor Interface}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-298}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D-030}
+\includegraphics[width=\columnwidth]{figs/density-1D}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D-060}
+\includegraphics[width=\columnwidth]{figs/energy-1D}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D-120}
-\end{center}
-\end{figure} 
-\end{frame}
-
-\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-1D-030}
-\end{center}
-\end{figure} 
-\end{frame}
-
-\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-1D-060}
-\end{center}
-\end{figure} 
-\end{frame}
-
-\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-1D-120}
-\end{center}
-\end{figure} 
-\end{frame}
-
-\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D-030}
+\includegraphics[width=\columnwidth]{figs/xassoc-1D}
 \end{center}
@@
 
-\begin{frame}[fragile]{Water Constrained to a Slit (1D)}
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D-120}
-\end{center}
-\end{figure} 
-\end{frame}
-
-\begin{frame}[fragile]{1D Slit - Cavity size versus Energy}
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/cavitysize-vs-energy-1D}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
 \end{center}
@@
 \begin{frame}[fragile]{Hydrophobic Rods in Water}
-Initial constraint: density equal to saturated liquid density outside rods, 1000 times vapor density inside
 \begin{figure}
 \begin{center}
-%\includegraphics[width=\columnwidth]{figs/rods-in-water-constraint}
+\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
 \end{center}
 \end{figure} 
-Diameter of rods = 1 nm (Distance apart = 2nm)
+\end{frame}
+
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
+\end{center}
+\end{figure} 
+\end{frame}
+
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
+Resolution and Convergence Tests
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-vs-radius}
+\end{center}
+\end{figure} 
 \end{frame}
@@
 
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-rods}
+\end{center}
+\end{figure} 
+\end{frame}
+
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\end{center}
+\end{figure} 
+\end{frame}
+
+\begin{frame}[fragile]{Hydrophobic Rods in Water}
+The critical transition
+\end{frame}
+
 \end{document}

### 2011-10-10 paper/figs/constrained-water-1D.cpp
compute tensor white bear densities in contact paper Tested everything.

@@ -16,7 +16,7 @@
 
 #include <stdio.h>
 #include <time.h>
-#include "Functionals.h"
+#include "OptimizedFunctionals.h"
 #include "Testables.h"
 #include "equation-of-state.h"
 #include "LineMinimizer.h"

### 2011-10-18 paper/project.tex

@@
 The first term in the attractive energy is the dispersion term.  This
-will include van der Waals and other non-associative interations. We use a
+will include van der Waals attraction and any orientation-independent
+interactions. We use a
 dispersion term based on the SAFT-VR
@@
 non-interacting rods, the critical separation will be
-different for any system where there is an appreciable force between rods. 
+different for any system where there is an appreciable direct force between rods. 
 The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
@@
 At small separations, the oscillations around the two rods look as if the two rods were one
-solid ``stadium'' shaped object (a rectangle with semi-circles on the ends).
+solid ``stadium''-shaped object (a rectangle with semi-circles on the ends).
 
@@
 is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
-For comparison, we arbitrarily offset each free energy such that they all go 
-to zero at large separations. For large $d$, the rods will 
+For ease of comparison, we arbitrarily offset each free energy such that they all go 
+to zero at large separations. For large $d$, the rods 
 simply act like two separate single rods in water and the energy will
@@
 (where the rods have vapor between them). This indicates that the force 
-per length on the rods is constant, and surprisingly not dependent on the separation. 
-By finding the slope, we calculate this force per length to be 144~mN/m.
+per length on the rods is constant, and surprisingly independent of separation. 
+By finding the slope of this curve, we calculate this force per length to be 144~mN/m.
 
@@
 
-If we use a classical approximation where the energy on the rod is equal to 
-the surface area multiplied by the bulk surface tension, then the energy per 
+If we use a classical approximation where the energy of the rods is equal to 
+the water's surface area multiplied by the bulk surface tension, then the energy per 
 length is equal to the circumference multiplied by the surface tenson. The force
@@
 To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will use this classical approximation and propose that 
+radius, we will again use this classical approximation and propose that 
 the critical separation, $d_c$, will occur when the free energies are equal. This 
@@
 
-Walther \emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
-carbon nanotubes, which are very geometrically similar to our hydrophobic rods. 
+Walther \emph{et al.}\cite{walther2004hydrodynamic} studied the interactions between two
+carbon nanotubes, which are geometrically similar to our hydrophobic rods. 
 They use molecular dynamics with Lennard-Jones potentials for
 nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
-They also find a drying transition occurs, and find it to be at separations near
+They also find that a drying transition occurs, and find it to be at separations near
 0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than
@@
 efficient than molecular dynamics. This allows for future work
-involving hydrophilic interactions with more complex electrostatic potentials.
+involving hydrophilic interactions with more complex electrostatic interactions.
 

### 2011-10-19 papers/thesis-hughes/presentation.tex
Archive project in separate directory and add the sections to paper Tests skipped!

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{color}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{bm}
\usetheme{Frankfurt}
\usecolortheme{lily}

\newcommand{\xx}{\textbf{r}}

\title{A Classical Density-Functional Theory for Describing Water Interfaces}
\author{Jessica Hughes}
\date{October 11, 2011}

\setbeamertemplate{navigation symbols}{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\section{Introduction}
\subsection*{}

\begin{frame}{Why Water is Cool}
\framesubtitle{Some interesting and unusual properties}
\begin{itemize}
 \item <1-> Simple structure/geometry
 \vspace{0.2cm}
 \item <2-> Strong hydrogen bonds
 \vspace{0.2cm}
 \item <3-> High melting, boiling, and critical temperature 
 \vspace{0.2cm}
 \item <4-> Density maximum above the melting point
 \vspace{0.2cm}
 \item <5-> Large dielectric constant
 \vspace{0.2cm}
 \item <6-> Cheap, abundant, nontoxic
\end{itemize}
\end{frame}

\begin{frame}{Density Functional Theory}
The Hohenberg-Kohn
theorem allows us to write the ground state
energy in terms of a universal functional of the electron density and a separate
term including the external potential. \\
\vspace{0.5cm}
\pause
Mermin gives a
theorem extending that of Hohenberg and Kohn to nonzero temperatures:
\begin{equation}
  A(T) = \underset{n(\xx)}{min}\left\{ F[n(\xx),T] + \int V_{ext}(\xx) n(\xx)
d^3r\right\}
\end{equation}
This is the basis for classical DFT, where $n(\xx)$ is the density of molecules
instead of electrons.
\end{frame}

\begin{frame}{Statistical Associating Fluid Theory}
\framesubtitle{Why water needs special treatment}
 \begin{itemize}
  \item <1-> Simple fluids: characterized by van der Waals and weak electrostatic interations
  \item <2-> Water: strong hydrogen bonding and polar interactions
 \end{itemize}
\begin{columns}
 \column <3-> {.5\textwidth}
\begin{figure}
 \includegraphics[width=4.2cm]{figs/hydrogen-bonding}
\end{figure}
\column <4-> {.5\textwidth}
Other examples of hydrogen-bonding fluids/molecules:\\
\begin{itemize}
 \item <4-> Alcohols
 \item <4-> Ammonia
 \item <4-> DNA
\end{itemize}
\end{columns}
\end{frame}

\begin{frame}{Statistical Associating Fluid Theory}
\vspace{0.7cm}
SAFT was designed for
use with fluids such as water, to account for the important association forces
as well as any effects due to clusters or chains forming
in the liquid.\\
\vspace{0.7cm}
\pause
Clark \emph{et al.}$^1$ have developed
a useful SAFT model for water that will be the basis for our functional in the
homogeneous limit.\\
\vspace{2cm}
\scriptsize{$^1$G.N.I. Clark, A.J. Haslam, A. Galindo, and 
G. Jackson.\\ \emph{Molecular Physics}, 104(22):3561-3581, 2006.}
\end{frame}

\section{The Functional}
\subsection*{}

\begin{frame}[fragile]{The Functional Terms}
The Helmholtz free energy functional is composed of:
\begin{equation}
  F[n] = F_\text{id}[n] + F_\text{hs}[n] + F_\text{disp}[n] + F_\text{assoc}[n] 
\end{equation}
where $F_\text{id}$ is the ideal gas free energy, $F_\text{hs}$ is
a hard-sphere free energy, $F_\text{disp}$ is the dispersion energy, and 
$F_\text{assoc}$ is the free energy of
association.
\end{frame}

\begin{frame}[fragile]{Ideal Gas Free Energy}
The ideal gas functional is given by
\begin{equation}
  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx
\end{equation}
where $n(\xx)$ is the density of water molecules and $n_Q$ is the
quantum concentration.
\begin{equation}
 n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}
\end{equation}
\end{frame}

\begin{frame}[fragile]{Hard-Sphere Free Energy}
\begin{itemize}
 \item <1-> For repulsive interactions, we use the White Bear version of
the Fundamental-Measure Theory~(FMT) functional.
 \vspace{0.5cm}
 \item <2-> This describes a hard-sphere fluid with a radius of $R=3.03420~\AA$
  (approximate size of a water molecule).
 \vspace{0.5cm}
 \item <3-> FMT functionals are expressed as the integral of
the \emph{fundamental measures} of a fluid.
  \begin{itemize}
    \item <4-> Filling fraction ($n_3$)
    \item <5-> Density of spheres touching a given point  ($n_2$)
  \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Dispersion Free Energy}
The dispersion free energy includes van der Waals attraction and 
any orientation-independent interactions. We use a
dispersion term based on the SAFT-VR approach, which has two free
parameters, an interaction energy $\epsilon_\text{d}$ and a 
length scale $\lambda_\text{d}$. It has the form
\begin{align}
  F_\text{disp}[n] &= \int (a_1(\xx) + \beta a_2(\xx))d\xx 
\end{align}
where $a_1$ and $a_2$ are the first two terms in a high-temperature
perturbation expansion.
\end{frame}

\begin{frame}[fragile]{Association Free Energy}
The association term models the hydrogen bonds as
four association sites on the surface of the
hard sphere. There is an attractive energy
$\epsilon_\text{assoc}$ when two molecules are arranged such that
the hydrogen of one interacts with the lone pair of the other.  The
volume of this interaction is $\kappa_\text{assoc}$ and the free 
energy is given by
\begin{align}  
  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
\end{align}
where the value of $4$ comes from the four association sites per
molecule, and the functional $X$ is the fraction of association sites
\emph{not} hydrogen-bonded. 
\end{frame}

\begin{frame}[fragile]{Empirical Parameters}
\framesubtitle{Fitting the functional to experiment}
There are five optimized parameters we take from Clark \emph{et al.}:
\begin{itemize}
 \item <2-> $R$ : hard sphere radius
 \item <3-> $\epsilon_\text{d}$ : dispersion interaction energy
 \item <4-> $\lambda_\text{d}$ : dispersion length scale
 \item <5-> $\epsilon_\text{assoc}$ : association attractive energy
 \item <6-> $\kappa_\text{assoc}$ : association interaction volume
\end{itemize}
\vspace{0.4cm}
\pause\pause\pause\pause\pause\pause
And one additional parameter we add:
\begin{itemize}
 \item <7-> $s_\text{d}$ : length scale for density averaging
\end{itemize}
\vspace{0.4cm}
\pause
These are optimized to fit experimental values such as vapor pressure, saturated 
liquid density, degree of association, and surface tension.
\end{frame}

\section{Thermodynamic Calculations}
\subsection*{}

% \begin{frame}[fragile]{Pressure Calculation}
% The pressure is obtained by taking a partial
% derivative with respect to volume at fixed temperature.  Since
% \emph{volume} isn't a parameter in density-functional theory, we
% consider the \emph{free energy per unit volume}.
% \begin{align}
%   F[n] &= f[n]V \\
%   p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
%   &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
%   &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
%    - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
%   &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n]
% \end{align}
% \end{frame}

\begin{frame}[fragile]{Additional Thermodynamic Quantities}
In addition to Helmholtz free energy, we can easily calculate
\begin{align}
  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} 
    = -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \nonumber \\
  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n] \\
  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V} \\
  U[n] &= F[n] + TS[n] \\
  H[n] &= F[n] + TS[n] + pV
\end{align}
\pause
One simple calculation: 
\begin{itemize}
 \item Heat of vaporization at atmospheric pressure is
$\Delta H_{vap}$=~39.41~kJ/mol (NIST: 40.65~kJ/mol).
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Temperature Coexistence Curve}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Determining Densities with the Common Tangent Rule}
\framesubtitle{At room temperature (298 K)}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Determining Densities with the Common Tangent Rule}
\framesubtitle{Close to the critical temperature}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/near-critical-point}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Pressure Isotherms and Coexistence Curve}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
\end{center}
\end{figure} 
\end{frame}

% \begin{frame}[fragile]{Vapor Pressure - Cut this slide?}
% \begin{figure}
% \begin{center}
% \includegraphics[width=\columnwidth]{figs/equation-of-state}
% \end{center}
% \end{figure}
% \end{frame}

\begin{frame}[fragile]{Surface Tension}
\framesubtitle{Fit for liquid water near room temperature and 
atmospheric pressure using $s_\text{d}$}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/surface-tension}
\end{center}
\end{figure} 
\end{frame}

\section{Results: Slit}
\subsection*{}

% \begin{frame}[fragile]{Liquid-Vapor Interface}
% \begin{figure}
% \begin{center}
% \includegraphics[width=\columnwidth]{figs/surface-298}
% \end{center}
% \end{figure} 
% \end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit}
\framesubtitle{One dimensional problem}
\begin{columns}
 \column {0.5\textwidth}
  \begin{figure}
  \begin{center}
  \includegraphics[width=4.5cm]{figs/slit-diagram}
  \end{center}
  \end{figure}
  Diagram of the water-filled slit.
 \column {0.5\textwidth}
  Setting up calculations for the slit:
  \begin{itemize}
  \item Pressure = 1~atm
  \item Temperature = 298~K
  \item Lattice spacing $\approx$~0.01~nm 
  \item Small non-zero density constraint outside the slit
  \item Slit size = 0.1~nm to 8~nm
  \end{itemize}
\end{columns}
\vspace{0.8cm}
We minimize the energy and calculate density, free energy, entropy, 
internal energy, and $X$.
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit}
\framesubtitle{Density profile for 4~nm slit}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-1D}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit}
\framesubtitle{Hydrogen bonding in a 4~nm slit}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-1D}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Water Constrained to a Slit}
\framesubtitle{Metastability}
Is it really possible to have liquid water in these nanoscale slits?
\vspace{0.5cm}
\begin{itemize}
 \item <2-> For slits smaller than 0.6~nm, the initial liquid water inside will
vaporize.
 \vspace{0.5cm}
 \item <3-> Will the larger slits containing liquid water be stable?
  \begin{itemize}
    \item <4-> Work to pull apart two hard walls = $10^{-7} - 10^{-8}$ Htr
    \item <4-> Free energy of liquid-filled slit = $10^{-5} - 10^{-6}$ Htr
  \end{itemize}
 \vspace{0.5cm}
 \item <5-> For the range of slit sizes we
studied, all the computational results of liquid-filled slits are
metastable.  
\end{itemize}
\end{frame}

\section{Results: Rods}
\subsection*{}

\begin{frame}{Hydrophobic Rods in Water}
\framesubtitle{Single rod diagram}
\begin{columns}
 \column {0.5\textwidth}
  \begin{figure}
  \begin{center}
  \includegraphics[width=4.5cm]{figs/single-rod-diagram}
  \end{center}
  \end{figure}
  Cross section diagram of a single hydrophobic rod in water.
 \column {0.5\textwidth}
  Setting up calculations for the rod:
  \begin{itemize}
  \item Pressure = 1~atm
  \item Temperature = 298~K
  \item Lattice spacing $\approx$~0.01~nm 
  \item Small non-zero density constraint outside the rod
  \item Radius = 0~nm to 1.5~nm
  \end{itemize}
\end{columns}
\vspace{0.5cm}
We minimize the energy and calculate density, free energy, entropy, 
internal energy, and $X$.
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Density profiles for single rods of various radius}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/density-single-rod}
\end{center}
\end{figure}
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Hydrogen bonding outside a slit of radius 0.2~nm}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Total free energy per area for all radii}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Resolution and convergence testing for a rod with radius 0.5~nm}
\begin{figure}
\begin{center}
\includegraphics[width=7.7cm]{figs/density-vs-radius}
\end{center}
\end{figure} 
\footnotesize{Resolutions:}
\begin{itemize}
 \item \footnotesize{High $\approx$ 0.003~nm (0.05 bohrs)}
 \item \footnotesize{Medium $\approx$ 0.01~nm (0.2 bohrs)}
 \item \footnotesize{Low $\approx$ 0.03~nm (0.5 bohrs)}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Two rods diagram}
Cross section diagram of two rods in water.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/rods-diagram}
\end{center}
\end{figure}
The radius $r$ ranges from 0.1~nm to 1.0~nm and the separation
$d$ ranges from 0~nm to about 1.2~nm.
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Density profiles for two rods of radius 0.5~nm}
\begin{columns}
 \column{0.7\textwidth}
  \begin{figure}
  \includegraphics[width=7.3cm]{figs/density-rods-in-water}
  \end{figure} 
 \column{0.3\textwidth}
 \small{``Stadium''-shaped vapor-filled area between rods}\\
 \vspace{0.3cm}
 \small{$d=$~0.2~nm}\\
 \vspace{1.4cm}
 \small{Liquid-filled area between two rods}\\
 \vspace{0.3cm}
 \small{$d=$~0.6~nm}
\end{columns}
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Energies for rods of radius 0.5~nm}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/energy-rods}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{Free energy per length for all radii}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
\end{center}
\end{figure} 
\end{frame}

\begin{frame}[fragile]{Hydrophobic Rods in Water}
\framesubtitle{The force per length}
  \begin{figure}
  \includegraphics[width=6.7cm]{figs/rods-energy-vs-distance}
  \end{figure}  
The force per length is constant, and surprisingly 
independent of separation. By finding the slope of this curve, we 
calculate this force per length to be 144~mN/m.
\end{frame}

\begin{frame}[fragile]{A Macroscopic Surface Tension Model}
\framesubtitle{Explaining the constant force per length for small separation}
\begin{itemize}
 \item <1-> Free energy of the rods = surface area times surface tension
 \item <2-> Free energy per length = circumference times surface tension
 \item <3-> Force per length = derivative of free energy per length with respect to the separation 
 \item <4-> We approximate the circumference of the stadium-shape by
  \begin{equation}
  C_{s} = 2\pi r +4r+2d
  \end{equation}
 \item <5-> Force per length = 2 times the surface tension
 \begin{itemize}
  \item <6-> Fitted surface tension for bulk water = 72 mN/m
  \item <6-> Free energy per area asymptote for single rods = 75 mN/m
  \item <6-> Force per length from slope = 144~mN/m
 \end{itemize}
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Vapor-Liquid Transition}
\framesubtitle{Estimating the critical separation} 
Using the same macroscopic surface tension model, we propose that 
the critical separation, $d_c$, will occur when the free energies are equal. This 
corresponds to equating the circumference of the stadium-shape and the circumference of the two rods, which results in
\pause
\begin{align}
C_\text{stadium} &= C_\text{rods} \\
2\pi r +4r+2d_c &= 4\pi r \\
d_c &= (\pi-2)r
\end{align}
\end{frame}

\begin{frame}[fragile]{Vapor-Liquid Transition}
\framesubtitle{The critical separation results}
The results for estimating the critical separation for the vapor-liquid 
transition.
\begin{table}
\begin{tabular} {|c|c|c|c|}
\hline
$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
\hline
0.3 & 0.34 & 0.23 & 32\% \\
\hline
0.5 & 0.57 & 0.54 & 5.3\% \\
\hline
0.7 & 0.80 & 0.81 & 1.3\% \\
\hline
0.9 & 1.03 & 1.06 & 2.9\% \\
\hline
1.0 & 1.14 & 1.19 & 4.4\% \\
\hline
\end{tabular}
\end{table}
\pause 
The diameter of a water molecule is approximately 0.6~nm.\\
\vspace{0.5cm}
\pause
This relation will not hold for all radii rods, it will 
likely become a gradual transition for very large rods.
\end{frame}

\section{Conclusions}
\subsection*{}

\begin{frame}[fragile]{Hydrophobic Interactions}
\begin{itemize}
 \item <1-> Two hydrophobic rods in water will transition from vapor to liquid between the rods.
 \vspace{0.5cm}
 \item <2-> This transition can be approximated with a macroscopic surface tension model
  for rods larger than 0.5~nm.
 \vspace{0.5cm}
 \item <3-> The separation necessary for the transition is $d_c = (\pi-2)r$.
 \vspace{0.5cm}
 \item <4-> The computationally efficient functional produces reasonable results for
hydrophobic interations.
\end{itemize}
\end{frame}

\begin{frame}[fragile]{Future Work}
\framesubtitle{Using the functional for more complex interactions}
\begin{itemize}
 \item <1-> More configurations of hydrophobic rods
 \vspace{0.5cm}
 \item <2-> Hydrophobic spheres
 \vspace{0.5cm}
 \item <3-> Hydrophilic solutes
\end{itemize}
\end{frame}

\begin{frame}{Thanks for Listening...}
\begin{center} 
\huge{Questions?}
\end{center}
\end{frame}

\end{document}

### 2011-10-19 paper/paper.tex

--- 2011-06-16 paper/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Introduction}
+
+\subsection{Water---the universal solvent}
+
+The simple yet complex water molecule is one of the most important 
+substances for life and industry. The structure and geometry of the 
+water molecule leads to a strong polarity and complex intermolecular
+forces. The hydrogen bonds formed by water are very strong, creating
+tetragonal clusters that lead to many of water's unusual properties,
+such as the high melting, boiling, and critical temperatures as well as the
+density maximum above the melting point.
+The large dielectric constant is partially responsible for
+water acting as a universal solvent, along with the ease of availability and
+use.
+There are many models used to describe water since so many different
+processes occur in aqueous solution. It is challenging to describe all the 
+relevant properties of water at once, especially if a low computational cost is
+also desired.
+
+Some models are primarily designed for use with homogeneous systems. Although
+these 
+models are useful in predicting some bulk properties, the interesting
+chemistry that occurs in water cannot be described with a homogeneous-only
+model. To
+describe solvation and other processes in aqueous solution, a more robust theory
+that 
+can handle varying densities is needed.
+
+\subsection{Classical density-functional theory}
+
+Classical density-functional theory (DFT) provides an efficient way to evaluate
+the free energy using the probability distribution of the fluid. The free energy 
+is a functional
+of the density function and is independent of any external
+potentials\cite{ebner1976density}. The most common use of DFT is calculating ground
+state electron densities in quantum mechanical systems. The Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
+energy in terms of a universal functional of the electron density and a separate
+term including the external potential. Mermin\cite{mermin1965thermal} gives a
+theorem extending that of Hohenberg and Kohn to nonzero temperatures:
+\begin{equation}
+  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
+d^3r\right\}
+\end{equation}
+This is the basis for classical DFT, where the relevant density is not the
+electron density,
+but the density of molecules. This can be used with a variety of 
+external potentials, from the simple interaction with a 
+hydrophobic wall to the more complex, stronger interaction with hydrophilic
+solutes.
+
+\subsection{Statistical Associating Fluid Theory}
+
+The high hydrogen bonding and strong polar interactions in water characterize it
+as an associating fluid, and thus it requires a different treatment than simple
+fluids which are mainly defined by van der Waals forces and weak electrostatic
+interations. The Statistical Associating Fluid Theory (SAFT) was designed for
+use with fluids such as water, to account for the important association forces
+(like hydrogen bonds) as well as any effects due to clusters or chains forming
+in the liquid\cite{muller2001molecular}. It is based on Wertheim's first-order
+thermodynamic perturbation theory (TPT1)
+\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
+wertheim1986fluidsIV} which provides a useful way to account for strong
+associative interations between molecules.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Theory and Methods}
 We introduce a new classical density functional for water that
-\needsworklater{predicts the vapor density, liquid density, bulk
-  modulus, and bulk surface tension, as well as allows for
-  liquid-vapor coexistence}.  The Helmholtz free energy functional is
+predicts the vapor pressure, vapor density, liquid density, bulk surface tension, 
+and enthalpy, as well 
+as allows for liquid-vapor coexistence.  The Helmholtz free energy functional is
 based on SAFT, and is composed of the usual terms:
 \begin{equation}
-  F[n] = F_\textit{id}[n] + F_\textit{hs}[n] + F_\textit{assoc}[n] + F_\textit{disp},
+  F[n] = F_\textit{id}[n] + F_\textit{hs}[n]  +
+F_\textit{disp}[n]+ F_\textit{assoc}[n],
 \end{equation}
 where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
-a hard-sphere free energy, $F_\textit{assoc}$ is the free energy of
-association, and $F_\textit{disp}$ is the dispersion energy.  In
+the hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
+and $F_\textit{assoc}$ is the free energy of association.  In
 addition, a chemical potential term is used, so we actually work in
@@
 \begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{n(\xx)} - 1\right) d\xx
+  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx
 \end{equation}
-where n(\xx) is the density of water molecules.  The ideal gas free
+where n(\xx) is the density of water molecules and $n_Q$ is the quantum 
+concentration. This is the particle concentration where quantum effects
+become appreciable and is given by
+\begin{equation}\label{quantumconcentration}
+ n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}
+\end{equation}
+where $m$ is the molecular mass of water. The ideal gas free
 energy functional has the property that the contact density at a hard
 surface $n_c$ is proportional to the pressure on that surface,
-according to the equation
-\begin{equation}
+according to the contact value theorem
+\begin{equation}\label{contactvaluethm}
   p = n_c k_BT \:,
@@
 solutes.  These properties are retained by the total functional,
-provided the remaining terms are \emph{purely
-  nonlocal}~\cite{ashcroft?}.
+provided the remaining terms are \emph{purely nonlocal}. This requires
+that
+\begin{equation}
+ \frac{\delta^2F_{\textit{excess}}[n]}{\delta n(\xx_1)~\delta n(\xx_1)}\neq\infty
+\end{equation}
+where $F_{\textit{excess}}[n]$ describes all the remaining terms in the functional.
+This ensures that any higher order terms in the Taylor expansion
+of $F[n+\Delta n]$ are small and do not contribute to the small
+solute free energy of solvation.
 
 \subsection{Hard-sphere repulsion}
-The repulsive intermolecular interactions are modeled using a
-Fundamental\nobreakdash-Measure Theory~\cite{rosenfeld1997}(FMT) functional,
-which combines scaled-particle theory and the Percus-Yevick theory,
-and an attractive interaction.  The FMT functional treats
-each water molecule as a hard sphere with a radius approximately that of
-a water molecule $(R~=~2.3~bohr)$.\\
 
 We model the repulsive interactions using the White Bear version of
-the Fundamental-Measure Theory~(FMT) functional published in
-reference~\cite{roth2002whitebear}.  \needsworklater{The FMT
-  functional describes the excess free energy of a hard-sphere
-  fluid. Fundamental-Measure Theory has several advantages over other
-  approaches, one of the most important being the simplicity of
-  calculations.  This particular FMT also reliably predicts bulk
-  properties of fluids in one, two and three dimensions, as well as
-  exactly determines the excess free energy in the zero dimensional
-  limit.}
+the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
+fluid~\cite{roth2002whitebear}.  In the homogeneous limit, the White
+Bear functional reduces to the Mansoori-Carnahan-Starling-Leland
+equation of state.  FMT functionals are expressed as the integral of
+the \emph{fundamental measures} of a fluid, which provide local
+measures of quantities such as the filling fraction, density of
+spheres touching a given point and mean curvature.  The excess free
+energy is written as:
 \begin{equation}
@@
   \mathbf{n}_{V1} &= \frac{\mathbf{n}_{V2}}{4\pi R}\\
-  n_0 &= \frac{n_2}{4\pi R^2}
+  n_0 &= \frac{n_2}{4\pi R^2}.
 \end{align}
+The density $n_3$ is the filling fraction and $n_2$ describes the number
+of spheres touching a given point. For our water functional, we use the 
+hard-sphere radius of
+3.03420~\AA, which was found to be optimal by Clark
+\emph{et al}.\cite{clark2006developing}
+
+\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
+\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
+\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
+\newcommand\kappaassoc{\ensuremath{\kappa_\textit{a}}}
+\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
+\newcommand\lscale{\ensuremath{s_d}}
+\subsection{Dispersion free energy}
+The first term in the attractive energy is the dispersion term.  This
+will include van der Waals attraction and any orientation-independent
+interactions. We use a
+dispersion term based on the SAFT-VR
+approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
+parameters, an interaction energy $\epsilondisp$ and a
+length scale $\lambdadisp$.
+
+The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+\begin{align}
+  F_\text{disp}[n] &= \int (a_1(\xx) + \beta a_2(\xx))d\xx
+\end{align}
+where $a_1$ and $a_2$ are the first two terms in a high-temperature
+perturbation expansion and $\beta=1/k_BT$.  The first term $a_1$ is 
+the mean-field dispersion
+interaction, which reduces in the homogeneous limit to
+\begin{align}\label{eq:A1-simple}
+  a_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
+  g_{HS}(\left|\xx\right|) d\xx
+\end{align}
+where $n_b$ is the bulk density and $g_{HS}$ is the homogeneous
+hard-sphere fluid correlation function.
+The second dispersion term in the free energy $a_2$ describes the
+effect of fluctuations resulting from the compression of the fluid due
+to the dispersion interaction itself, and is commonly approximated
+using the local compressibility approximation (LCA). In the LCA,
+we assume the energy fluctuation is directly related to the
+compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
+
+We use a free square-well function for the dispersion $\varphi$, which
+is the choice used by Clark \emph{et al}~\cite{clark2006developing},
+and allows a reasonable fit to the equation of state.  Thus our model
+interaction has the form:
+\begin{equation}
+  \varphi(r) = \Theta(r-2 \lambdadisp R)
+\end{equation}
+where $\Theta$ is the Heaviside step function.
+
+The form of $a_1$ is given in
+reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
+of the filling fraction.  In order to apply this form to an
+inhomogeneous density distribution, we construct an effective local
+filling fraction for dispersion $\etadisp$, given by
+\begin{align}
+  \etadisp(\xx) &=
+  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
+             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
+\end{align}
+This effective filling fraction is used throughout the dispersion
+calculations, and represents a filling fraction averaged over the
+effective range of the dispersive interaction.  This functional
+introduces an additional empirical parameter $\lscale$ which adjusts
+the length scale over which the dispersion interaction is correlated.
+
+The first term in the dispersion functional $a_1$ when written using
+the above filling fraction for dispersion has the form
+\begin{align}
+  a_1 &= 
+   -4(\lambdadisp^3 - 1)\epsilondisp \etadisp(\xx)
+    g^{HS}_\sigma(\eta_\textit{eff}(\xx)) \\
+  g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
+  +\frac32\frac{\zeta(\xx)}{(1-\eta)^2}
+  + \frac12\frac{\eta^2\zeta(\xx)}{(1-\eta)^3}
+  \\
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}\\
+  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
+  \lambdadisp^j
+\end{align}
+where $g_\sigma^{HS}$
+is the hard-sphere fluid correlation function evaluated at contact.
+The $\zeta$ factor
+(also seen in the association energy, below) is a dimensionless
+measure of the local density inhomogeneity. The $C_{ij}$ values
+ are numerical constants taken from
+reference\cite{gil-villegas-1997-SAFT-VR}, which come from a numerical
+fit to the integral in Equation~\ref{eq:A1-simple} over a range of
+values for $\lambdadisp$ and filling fractions.
+
+The second term, $a_2$, which describes the contribution to the free
+energy associated with fluctuations is given by
+\begin{align}
+  a_2 &= \frac12 \epsilondisp
+              K_{HS}(\etadisp(\xx)) \etadisp(\xx)
+              \frac{\partial a_1}{\partial \etadisp(\xx)}
+\end{align}
+where $K_{HS}$ is the Percus-Yevick hard-sphere compressibility, given
+by
+\begin{align}
+  K_{HS} &=
+    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}.
+\end{align}
+
+\subsection{Association free energy}
+The final attractive energy term is the association term, which
+accounts for hydrogen bonding.  Hydrogen bonds are modeled as four
+attractive patches (``association sites'') on the surface of the hard
+sphere.  These four sites represent two protons and two electron lone
+pairs.  There is an attractive energy $\epsilonassoc$ when
+two molecules are oriented such that the proton of one overlaps
+with the lone pair of the other.  The volume over which this
+interaction occurs is $\kappaassoc$, giving the association
+term in the free energy has two empirical parameters that are fit to
+experimental data.
+
+The association functional we use is a modified version of that of Yu
+and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}.\footnote{We
+  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
+  the same functional, but their paper contains errors in the
+  association term and is not useful as a reference.}  The association
+functional of Fu and Wu includes the effects of density
+inhomogeneities on the \emph{contact density} of hard spheres, which
+conventionally appears in the form of the density multiplied by the
+contact value of the correlation function $g^{HS}_\sigma$, but is
+based on the SAFT-HS equation of state, rather than the SAFT-VR
+equation of state\cite{gil-villegas-1997-SAFT-VR} which is used in the
+optimal SAFT parametrization for water found by Clark \emph{et
+  al}\cite{clark2006developing}.
+
+The association free energy for our four-site model has the form
+\begin{align}
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
+\end{align}
+where the value of $4$ comes from the four association sites per
+molecule, and the functional $X$ is the fraction of assotiation sites
+\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
+measure of the density inhomogeneity.  The fraction $X$ is determined
+by the quadratic equation
+\begin{align}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)\Delta(\xx)} - 1}
+  {4 n_0(\xx)\zeta(\xx)\Delta(\xx)}
+\end{align}
+where the factor $\zeta$ from
+Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
+  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
+density, and reduces to unity in the homogeneous limit:
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
+\end{align}
+The functional $\Delta$ which appears in the formula for $X$ is a
+measure of hydrogen-bonding probability.  In the SAFT-VR
+model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
+\begin{align}
+  \Delta(\xx) &= \kappaassoc g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right) \\
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\etadisp(\xx)) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)
+\end{align}
+% This is equation 77 of gil-villegas-1997-SAFT-VR
+where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+contact for a hard-sphere fluid with a square-well attractive
+potential, and $a_1$ and $a_2$ are two terms in the dispersive term of
+the free energy.  The correlation function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere fluid correlation
+function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
+\begin{align}
+  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
+  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
+  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
+\end{align}
+
+\subsection{Other thermodynamic functions}
+
+From the Helmholtz free energy functional, we may obtain any other
+thermodynamic functions we should choose.  The grand free energy
+$\Phi$, which we ordinarily use in our computations, is obtained by
+simply adding a chemical potential term:
+\begin{equation}
+  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
+\end{equation}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
+\label{fig:pressure-with-isotherms}
+\end{figure}
+
+The pressure would naturally be obtained by taking a partial
+derivative with respect to volume at fixed temperature.  Since
+\emph{volume} isn't a parameter in density-functional theory, we
+consider the \emph{free energy per unit volume}:
+\begin{align}
+  F[n] &= f[n]V \\
+  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
+  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
+  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
+   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
+  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n].
+\end{align}
+
+Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
+isotherms ranging from room temperature to past the critical temperature. The
+solid line is the theoretical coexistence curve calculated from the SAFT
+functional. For a given isotherm, the pressure curve should intersect this
+coexistence curve twice, at the vapor and liquid coexistence 
+pressure corresponding to the vapor and liquid densities. Even
+though the theoretical pressure isotherms drop below the coexistence curve,
+this is non-physical and will not actually occur. The experimental data shows
+how the actual density value will ``jump'' to the other side of 
+the coexistence curve at that pressure value (for
+temperatures less than the critical temperature), corresponding to a phase 
+change in the material. For example, follow the room temperature isotherm
+starting from high pressure and high density (top right of 
+Figure~\ref{fig:pressure-with-isotherms}).
+As the pressure decreases, the density decreases \emph{slightly} until reaching
+the coexistence curve, at which point a phase change from liquid to vapor
+will occur. The
+functional matches well with the experimental data from NIST away from the
+critical point. The slope of the pressure versus density curve is proportional
+to the compressibility, which is close but slightly too large at room
+temperature.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
+  will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
 
@@
 \label{fig:equation-of-state}
-\end{figure}
-
-\subsection{Association free energy}
-To the purely repulsive hard-sphere free energy, we add two attractive
-energy terms.  The first is the free energy of \emph{association},
-which describes the effects of hydrogen bonds.  These are modeled as
-four attractive patches (``association sites'') on the surface of the
-hard sphere.  These four sites represent the two hydrogens and two
-electron lone pairs.  There is an attractive energy
-$\epsilon_\textit{assoc}$ when two molecules are arranged such that
-the hydrogen of one interacts with the lone pair of the other.  The
-volume of this interaction is $\kappa_\textit{assoc}$.  Thus the
-association term in the free energy has two free parameters that must
-be fit based on experimental data.
-
-The association free energy has the form
-\begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)
-  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
-\end{align}
-where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of assotiation sites
-\emph{not} hydrogen-bonded.  The fraction $X$ is determined by the
-quadratic equation
-\begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)\Delta(\xx)} - 1}
-  {4 n_0(\xx)\zeta(\xx)\Delta(\xx)}
-\end{align}
-where $\zeta$ comes from
-reference\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft} and describes the inhomogeneity of the fluid
-density:
-\begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n_{2V}}\cdot\mathbf{n_{2V}}}{n_2^2}
-\end{align}
-The function $\Delta$ is given... below?
-
-The following contact density for the hard-sphere fluid is in the code
-called \texttt{gHS}, and comes
-from\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft}, but I don't use it.
-\begin{align}
-  \Delta &= 4\pi \kappa_\textit{assoc} g^\textit{HS}(\sigma)e^{-\beta
-    \epsilon_\textit{assoc}} \\
-  g^\textit{HS}(\sigma) &= \frac1{1-n_3}\left(1+\frac14\frac{\zeta n_2}{1-n_3}
-  \left(1 + \frac{n_2}{14 (1-n_3)}\right)\right)
-\end{align}
-\textcolor{red}{Here is paper where I got the association functional:}
-
-Instead, I use
-\begin{align}
-  \Delta(\xx) &= \kappa_\textit{assoc}
-  e^{-\beta\epsilon_\textit{assoc}} g^\textit{SW}_\sigma(\xx) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
-  \frac{1}{4}\beta\left(\frac{\partial a_1(\xx)}{\partial n_3(\xx)} -
-  \frac{\lambda}{3 n3(\xx)}\frac{\partial a_1(\xx)}{\partial \lambda}\right)
-\end{align}
-% This is equation 77 of gil-villegas-1997-SAFT-VR
-which I got from reference\cite{gil-villegas-1997-SAFT-VR}.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/saturated-liquid-density}
-\end{center}
-\caption{The theoretical versus experimental saturated liquid density
-  versus temperature.  }
-\label{fig:saturated-liquid-density}
-\end{figure}
-
-\subsection{Dispersion free energy}
-The final term in the free energy is the dispersion term.  We use a
-dispersion term based on the SAFT-VR
-approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters, an interaction energy $\epsilon_\textit{disp}$ and a
-length scale $\lambda_\textit{disp}$.
-
-The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
-\begin{align}
-  F_\text{disp} &= A_1 + \beta A_2
-\end{align}
-where $A_1$ and $A_2$ are the first two terms in a high-temperature
-perturbation expansion.  $A_1$ is the mean-field dispersion
-interaction, which for a homogeneous fluid is given by
-\begin{align}
-  A_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
-  g_{HS}(\left|\xx\right|) d\xx
-\end{align}
-The second dispersion term in the free energy $A_2$ describes the
-effect of fluctuations resulting from the compression of the fluid due
-to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA).
-
-We use a free square-well function for the dispersion $\varphi$, which
-has been demonstrated \textcolor{red}{(CITE)} to give a reasonable fit
-to the equation of state.  Thus our model interaction has the form:
-\begin{equation}
-  \varphi(r) = \Theta(r-2 \lambda_\textit{disp} R)
-\end{equation}
-where $\Theta$ is the Heaviside step function.
-
-\textcolor{red}{TODO:  Add the actual functions that we use and the
-  actual form.  The above is really just an introduction.}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental saturated liquid density
-  versus temperature.  }
-\label{fig:saturated-liquid-density}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy}
-\end{center}
-\caption{The free energy, internal energy, and temperature*entropy versus density at room temperature.  }
-\label{fig:energy-room-temp}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy-at-690K}
-\end{center}
-\caption{The free energy, internal energy, and temperature*entropy versus density near the critical temperature.  }
-\label{fig:energy-near-critical-temp}
-\end{figure}
-
-\subsection{Other thermodynamic functions}
-
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx
-\end{equation}
-
-The pressure would naturally be obtained by taking a partial
-derivative with respect to volume at fixed temperature.  Since
-\emph{volume} isn't a parameter in density-functional theory, we
-consider the \emph{free energy per unit volume}.
-\begin{align}
-  F[n] &= f[n]V \\
-  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
-  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
-  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
-   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
-  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n]
-\end{align}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including experimental pressure data from NIST.  }
-\label{fig:pressure-with-isotherms}
 \end{figure}
@@
 \begin{align}
-  H[n] &= F[n] + TS[n] - pV
+  H[n] &= F[n] + TS[n] + pV.
 \end{align}
+The calculated heat of vaporization at atmospheric pressure is
+$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
+373 K. This compares well to the experimental
+value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
+an error of only a few percent.
+
+\subsection{Determining the empirical parameters}\label{sec:empirical}
+
+The majority of the empirical parameters used in our functional are
+taken from the paper of Clark \emph{et al} on developing an optimal
+SAFT model for water\cite{clark2006developing}.  This SAFT model
+contains five empirical parameters: the hard-sphere radius, an energy
+and length scale for the dispersion interaction, and an energy and
+length scale for the association interaction.  In addition to the five
+empirical parameters of Clark \emph{et al}, we add a single additional
+parameter $\lscale$---with a fitted value of 0.72---which determines
+the length scale over which the density is averaged when computing the
+dispersion free energy.  We use this final parameter to fit the
+surface tension, with the result shown in
+Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
+\emph{et al} overestimates the critical temperature---which is a
+common feature of models which do not explicitly treat the critical
+point---we cannot correctly describe the surface tension at all
+temperatures, and chose to fit it for the temperature range at which
+water is liquid at one atmosphere of pressure.
 
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
 \end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature.  }
-\label{fig:surface-tension}
+\caption{The theoretical versus experimental coexistence curve. Away from
+the critical temperature, the agreement is good. }
+\label{fig:temperature-vs-density}
 \end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{Liquid-vapor interface profile at room temperature.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
-
-\subsection{Crossover formalism for handling critical behavior}
-
-I am currently working on implementing the crossover approach of
-Kiselev and Ely which is designed to make the critical behavior come
-out right\cite{kiselev2006new}.  This method works by defining a
-\emph{classical} free energy $f_0$---which where classical really is
-closer to \emph{mean field} in meaning, and indicates a theory that
-doesn't include long-range fluctuations.  This free energy is then
-used to construct a corrected free energy by feeding portions of it a
-``fake'' density and temperature that are constructed so as to
-reconstruct the correct universal scaling laws near the critical
-point.
-
-\subsection{Determining the empirical parameters}\label{sec:empirical}
-
-In order to ensure that the empirical parameters are correct, we test
-the vapor pressure (see Figure~\ref{fig:equation-of-state}) and
-saturated liquid density (see
-Figure~\ref{fig:saturated-liquid-density}) against experimental
-values.  To find these densities, we use the ``common tangent'' rule,
-as illustrated in Figure~\ref{fig:homogeneous}.
 
@@
 \end{center}
-\caption{By plotting the free energy vs. density it is possible to
+\caption{By plotting the free energy versus density it is possible to
   determine a saturated liquid or vapor state by the common tangent
-  method.  }
+  method. The top plot illustrates the common tangent near room 
+  temperature, while the bottom plot shows the same thing close to the 
+  critical temperature. }
 \label{fig:homogeneous}
@@
 
-\begin{table}
-\begin{tabular}{ccc}
-  Fitting Parameters& \hspace{2em} & Experimental Values\\
-\begin{tabular}{cc}
-  $R$ & 2.3 $a_o$ \\
-  $A$ & -0.0100909 Hartree \\
-  $\gamma$ & 5\\
-  $V$ & 515.792 $a_o^3$  \\
-  $b$ & 0.75 $a_o$ \\
-  $\mu$ & -0.014063 Hartree 
-\end{tabular}
-&& 
-\begin{tabular}{cc}
- $n_l$ & 4.939 $\times 10^{-3} a_o^{-3}$\\
- $n_v$ & 1.141 $\times 10^{-7} a_o^{-3}$\\
- $\sigma$ & ??? \\
- $B$ & $7.435\times 10^{-5}$ Hartree$/a_o^3$\\
- $T$ & 298.15 K \\
-\end{tabular}
-\end{tabular}
-
-\caption{Empirical parameters of optimized functional and experimental
-  values matched, where $a_o$ is the Bohr radius, and $\sigma$ is the
-  surface tension.  This table is not yet complete, but it should
-  be soon...  We'll also need to think about the best units to use
-  here.  Atom units might be best, but it also might be better to
-  consistently use eV and angstrom or something.  I'm not sure about
-  putting $k_BT$ data in the table.  Maybe it belongs in the caption?}
-\label{tab:parameters}
-\end{table}
+For reference, we plot the predicted vapor pressure (see
+Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
+Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
+match those calculated by Clark \emph{et al}. The functional provides a good description of
+the vapor pressure, which has an error of less than 
+1\% compared with the experimental data. The coexistence curve in 
+Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
+curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
+temperature coexistence values.
+Near the critical temperature,
+the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
+experiment. 
+To find the densities for the coexistence curve, we use the
+``common tangent'' rule. The two densities 
+will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
+As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
+(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
+in correctly treating the critical temperature. As the vapor and liquid densities approach
+each other, the free energy barrier between the two states gets smaller and it becomes
+challenging to computationally find the correct minimum. This is the reason 
+why the coexistence curves
+in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
+are not smooth around the critical point.
 
@@
 \section{Results and discussion}
-\subsection{Liquid-vapor interface}
 
-We compute the density profile and surface tension by relaxing a slab
-of water in a periodic unit cell.  This slab has two surfaces, so the
-surface tension is just half the free energy of the cell, divided by
-its cross-sectional area.  As mentioned in
-section~\ref{sec:empirical}, the $b$ parameter is determined by
-setting this surface tension to match the experimental value.  The
-associated density profile is plotted in
-Figure~\ref{fig:cartesianDensity}, which displays two peaks in
-density.
+\subsection{Water confined to a slit}
 
+One of the most computationally simple ways to study confined water is by
+constraining it to a slit where the water molecules can only move in one
+dimension. Experimentally, this could represent two different situations.
+The first, water between two perfectly aligned and smooth planes, is not common 
+in experiment due to the difficulty of creating such an ideal situation. The
+second
+possibility is a large sphere brought close to a plane, which eliminates the 
+need for perfect alignment.
+
+Before looking at the liquid-solid one-dimensional slit, we set up a liquid-vapor
+interface and plot the density profile.
 The fluid density at a liquid-solid interface commonly displays
@@
 known as \emph{capillary waves}.  These waves are not accounted for in
-density-functional theory (although the \emph{exact}
-density-functionaly theory \emph{would} include them), with the result
-that the density profile will be smoothed out by CW effects.
+density-functional theory (although the exact
+density-functional theory would include them), with the result
+that the density profile will be smoothed out by capillary wave effects. The liquid-vapor
+interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
+some small oscillations.
 
-\fixme{It'd be worth another search for water intrinsic profile
-  papers... but the review I cite is pretty thorough.}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-298}
+\end{center}
+\caption{The liquid-vapor density profile at room temperature shows
+some small oscillations close to the interface.  }
+\label{fig:liquid-vapor-profile}
+\end{figure}
+
+To set up the calculations for the slit, we choose a chemical potential 
+corresponding to a pressure of 1~atm and set the
+temperature to a typical room temperature of 298~K. We apply 
+a constraining potential outside the slit, where we choose a small non-zero
+density for numerical reasons. The slits studied range in size from 0.1~nm to 
+8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
+lattice
+with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
 
@@
 \end{center}
-\caption{ Density profile for a slab of water with molecules constrained to move in one dimension. }
+\caption{ Density profile for a slab of water with molecules 
+constrained to move in one dimension. The slit size is 4.0~nm and the dashed
+line represents the saturated liquid density.}
 \label{fig:density-1D}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-1D}
-\end{center}
-\caption{ Energy density profile for a slab of water with molecules constrained to move in one dimension. }
-\label{fig:energy-1D}
-\end{figure}
+The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
+Near the interface, there are oscillations around the
+liquid density which decrease near the center of the slit. We expect the center
+of the slit to act approximately like bulk water, and the absence of
+oscillations supports this. For larger slits (6.0~nm or more) the majority of
+the water molecules in the slit act like the bulk.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/entropy-1D}
-\end{center}
-\caption{ Entropy for a slab of water with molecules constrained to move in one 
-dimension. \textcolor{red}{FIX UNITS}}
-\label{fig:entropy-1D}
-\end{figure}
+Based on our computations, we determine the smallest slit that can contain 
+liquid water. This result is
+approximately 0.6~nm. For smaller slits, the initial liquid water inside will
+vaporize. The results for larger liquid-filled slits are only metastable,
+however, as a vapor-filled slit is thermodynamically favorable. Based on our
+calculations for a pressure of 1~atm, the work needed to pull apart
+the two hard walls comprising the slit is one to two orders of magnitude less
+than the free energy of the liquid-filled slit. For the range of slit sizes we
+studied, all the computational results of liquid-filled slits are
+metastable.
+
+Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
+stability of water between large hydrophobic plates and the critical separation
+necessary for stability. For water at room temperature and atmospheric pressure,
+plates with a separation smaller than about 100~nm will not contain stable liquid
+water. They go on to say that confined water has a region of metastability down
+to separations of about 5~nm. This is comparable to the slit separation where
+our calculations show water inside the slit acting like the bulk liquid, which
+we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
+also study water between two hard walls at constant chemical potential. They 
+also conclude that the liquid state is likely metastable, and there is a free
+energy barrier to reach the stable vapor phase.
 
@@
 \end{center}
-\caption{ Xassociation value (fraction of association sites
-\emph{not} hydrogen-bonded) for a slab of water with molecules constrained to move in one 
-dimension. }
+\caption{The number of hydrogen bonds per molecule for a slab of water 
+with molecules constrained to move in one dimension. Each molecule may
+contribute to
+four bonds, although each bond involves two molecules so the
+actual number of bonds would be halved.}
 \label{fig:xassoc-1D}
 \end{figure}
+
+The value of $X$ behaves as expected for water constrained to a slit. We
+show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
+molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
+Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
+bonds is approximately equal to 3.5, which is expected for saturated liquid
+water.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\end{center}
+\caption{ Density profiles for a single rod of different radii. The dotted line
+represents the saturated liquid density.  }
+\label{fig:density-single-rod}
+\end{figure}
+
+\subsection{One hydrophobic rod}
+
+We move into two dimensions by studying a single hydrophobic rod
+immersed in water. Figure~\ref{fig:density-single-rod} shows density
+profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
+expected oscillations outside the rod and as the distance away from the rod
+increases
+the density approaches that of the bulk saturated liquid density. The small
+radius
+rod has comparatively larger oscillations and a higher contact density.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
+\end{center}
+\caption{ Number of bonds per molecule near the interface with a single
+hydrophobic rod. There is a decrease in the hydrogen bonding near the surface 
+of the rod, as expected at the interface of any hydrophobic wall. }
+\label{fig:xassoc-single-rod}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+\end{center}
+\caption{ Free energy versus radius for a single hydrophobic rod
+immersed in water. This should have an asymptote equal to the surface
+tension at room temperature, and it agrees well with the surface tension in
+Figure~\ref{fig:surface-tension}. }
+\label{fig:energy-vs-diameter}
+\end{figure}
+
+The value of $X$ is calculated and converted to number of bonds per molecule and 
+shown for a rod with radius 0.2~nm in Figure~\ref{fig:xassoc-single-rod}. Inside the
+rod (less than 0.2~nm) there are no molecules to form hydrogen bonds although the value
+of $X$ is still calculated there. The number of bonds decreases from a bulk value of 3.5 
+to about 3.25 at the interface. Walther \emph{et al.}\cite{walther2001carbon} 
+calculate the number of hydrogen bonds for a carbon nanotube immersed in water.
+They show a decrease in the bonding from a bulk value of 3.75 to 2.89 near the 
+interface with the nanotube. This is a much larger change than for our 
+rod of a similar radius, indicating that carbon nanotubes disrupt hydrogen
+bonds in water more than hydrophobic rods.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-vs-radius}
+\end{center}
+\caption{ Density profiles for 
+a single hydrophobic rod with a radius of 0.5~nm using several different 
+lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
+medium 0.2~bohrs ($\approx$~0.01~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
+\label{fig:densityresolution}
+\end{figure}
+
+Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
+radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
+about 0.5~nm or so, this reaches
+an asymptote equal to the surface tension. From the inset in 
+Figure~\ref{fig:surface-tension}, the surface tension at room 
+temperature is approximately 72~mN/m. This is very close to the 
+asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
+
+\subsubsection{Resolution and convergence tests}
+Figure~\ref{fig:densityresolution} shows the density profiles for three
+different values of the lattice spacing resolution. The density oscillations
+around a single rod of radius 0.5~nm look nearly identical for all resolutions
+except for the first peak. This contact density is important as it will be
+proportional to the pressure on the rod (see Equation~\ref{contactvaluethm}). The contact
+density calculated using the medium resolution is within less than 4\% of the
+value at high resolution, while the low resolution contact density is off by
+about 5\%. There is some uncertainty in the value of contact density due to the
+different lattice spacings which cause the contact density to be calculated
+at a slightly different position. The total free energy calculations compare much more
+favorably. The medium resolution is within 0.5\% and the low resolution is
+within 1\% of the high resolution free energy.
+
+\subsection{Hydrophobic interaction of two rods}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-diagram}
+\end{center}
+\caption{ Cross section schematic for two hydrophobic rods immersed in water.
+The radius and separation are shown.}
+\label{fig:rods}
+\end{figure}
+
+We now look at the more interesting problem of two rods in water. We continue to
+use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+(touching rods) to $d=1.3$~nm. At small
+separations there is only vapor between the rods. As the rods are
+pulled apart, the vapor region expands until some critical separation is
+reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
+shows density profiles before and after this transition
+for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
+on the radii of the rods, and is about 0.54~nm for the rods shown in 
+Figure~\ref{fig:density-rods}. 
+Since we are looking at only hard-walled
+non-interacting rods, the critical separation will be
+different for any system where there is an appreciable direct force between rods. 
+The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
+are very similar to
+those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
+At small separations, the oscillations around the two rods look as if the two rods were one
+solid ``stadium''-shaped object (a rectangle with semi-circles on the ends).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\end{center}
+\caption{ Density profiles illustrating the transition from vapor 
+to liquid water between the rods. The radius is 0.5~nm, the top figure is 
+at a separation of 0.2~nm and the
+bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
+the energy for these and other separations.}
+\label{fig:density-rods}
+\end{figure}
+
+For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
+and the entropic energy contribution, where all are in units of energy per length 
+(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
+goes from zero to the critical separation and then is approximately constant at
+larger separations. The entropy has a minimum just before the rods
+transition to allow liquid between them. As the rods separate starting from zero, the area 
+between them containing vapor increases and there is less overall space for the 
+surrounding liquid water to occupy and thus the entropy decreases. When that area fills suddenly
+with liquid, the entropy increases due to the increased space for the water molecules. The
+internal energy also has a minimum just before the transition to the liquid filled state.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/energy-rods}
+\end{center}
+\caption{ Free energy, internal energy, and the entropic energy contribution versus
+separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
+offset to zero at large separations. }
+\label{fig:energy-rods}
+\end{figure}
+
+The free energy per length is compared for rods of several different radii
+in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for radius 0.5~nm
+is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
+For ease of comparison, we arbitrarily offset each free energy such that they all go 
+to zero at large separations. For large $d$, the rods 
+simply act like two separate single rods in water and the energy will
+reach an asymptote determined by the effective surface tension (shown in 
+Figure~\ref{fig:energy-vs-diameter}). For small $d$, we see that all the rods 
+have the same behavior, with a positive linear relationship
+between free energy per length and separation during the initial state 
+(where the rods have vapor between them). This indicates that the force 
+per length on the rods is constant, and surprisingly independent of separation. 
+By finding the slope of this curve, we calculate this force per length to be 144~mN/m.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\end{center}
+\caption{ Free energy (also known as the potential of mean force) 
+versus separation for two hydrophobic rods ranging in radius from
+0.3~nm to 1.0~nm.
+All were arbitrarily offset to zero at large separations. The
+transition corresponds to the phase change from
+vapor to liquid between the rods as pictured in the density profiles in 
+Figure~\ref{fig:density-rods}. }
+\label{fig:rods-energy-vs-distance}
+\end{figure}
+
+If we use a classical approximation where the energy of the rods is equal to 
+the water's surface area multiplied by the bulk surface tension, then the energy per 
+length is equal to the circumference multiplied by the surface tenson. The force
+per length is the derivative of this with respect to the separation. We approximate 
+the circumference of the stadium-shape by
+\begin{equation}
+C_{s} = 2\pi r +4r+2d
+\end{equation}
+and so the force per length is equal to $2\gamma$. This 
+is consistent with the calculation of the force per length from 
+Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
+in Figure~\ref{fig:surface-tension} (72~mN/m). This indicates that
+even though the rods are nanoscale, the water molecules will ``see'' them classically.
+
+To estimate how the transition from vapor to liquid between the rods depends on their 
+radius, we will again use this classical approximation and propose that 
+the critical separation, $d_c$, will occur when the free energies are equal. This 
+corresponds to equating the circumference of the stadium-shape (see 
+top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom
+of Figure~\ref{fig:density-rods}), which results in
+\begin{equation}
+d_c = (\pi-2)r.\label{criticalseparation}
+\end{equation}
+The predicted and actual computed values for $d_c$ are shown in 
+Table~\ref{criticalseparationtable}, and they match fairly well for the 
+0.5~-~1.0~nm rods. We hypothesize that larger radii rods
+would also follow this trend, but smaller rods are likely to have a different
+relationship with the surrounding water molecules. This may explain why the critical
+separation result for a radius of 0.3~nm does not match the prediction well.
+
+Walther \emph{et al.}\cite{walther2004hydrodynamic} studied the interactions between two
+carbon nanotubes, which are geometrically similar to our hydrophobic rods. 
+They use molecular dynamics with Lennard-Jones potentials for
+nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
+They also find that a drying transition occurs, and find it to be at separations near
+0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than
+what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
+and carbon-water interations.
+
+\begin{table}
+\begin{tabular} {|c|c|c|c|}
+\hline
+$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
+\hline
+0.3 & 0.34 & 0.23 & 32\% \\
+\hline
+0.5 & 0.57 & 0.54 & 5.3\% \\
+\hline
+0.7 & 0.80 & 0.81 & 1.3\% \\
+\hline
+0.9 & 1.03 & 1.06 & 2.9\% \\
+\hline
+1.0 & 1.14 & 1.19 & 4.4\% \\
+\hline
+\end{tabular}
+\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
+for the transition
+from liquid to vapor between two hydrophobic rods.}
+\label{criticalseparationtable}
+\end{table}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\subsection{Two Hydrophobic Rods}
+\section{Conclusion}
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
-\end{center}
-\caption{ Density profile for two hydrophobic rods.}
-\label{fig:density-rods}
-\end{figure}
+The classical density functional we developed by combining fundamental-measure 
+theory with SAFT is able to produce reasonable results for hydrophobic interations.
+We added the parameter $\lscale$ to the functional and fit it to the surface tension
+at room temperature, which we find to be consistent with results for larger hydrophobic
+rods.
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-By adding an attractive term to the fundamental-measure theory
-functional, we developed a classical density functional for water
-which is able reproduce expected results for hydrophobic
-interactions.  We hope to extend this functional to allow for
-hydrophilic interactions by coupling with electrostatic fields.
+The free energy per area for water surrounding a hydrophobic rod shows the
+expected behavior with an asymptote equal to the surface tension.
+Based on this result, we expect to be
+able to classically describe rods with a radius of about 0.5~nm or larger. 
+For two hydrophobic rods surrounded by water, we see a transition from a vapor-filled
+stadium shape to two rods with liquid between them. A
+classical treatment of the critical separation for this transition
+works well for rods larger than 0.5~nm. Further evidence for the
+accuracy of this classical treatment is the calculation of the force on the
+stadium shape. The force per length is constant and equal to twice the surface tension.
 
-This method is computationally efficient, which will be useful when
-introducing new functionals that include intramolecular and electromagnetic
-interactions, that are not yet included in this model.
+For the hydrophobic systems studied, this functional is more computationally 
+efficient than molecular dynamics. This allows for future work
+involving hydrophilic interactions with more complex electrostatic interactions.
 

### 2011-10-24 papers/fuzzy-fmt/figs/density.py
add new empty paper Tested everything.

@@ -1,0 +1,26 @@
+#!/usr/bin/python
+
+# We need the following two lines in order for matplotlib to work
+# without access to an X server.
+import matplotlib
+matplotlib.use('Agg')
+
+import pylab, numpy, sys
+
+if len(sys.argv) != 5:
+    print("Usage:  " + sys.argv[0] + " mc-filename.dat wb-filename.dat wbt-filename.dat out-filename.pdf")
+    exit(1)
+
+mcdata = numpy.loadtxt(sys.argv[1])
+dftdata = numpy.loadtxt(sys.argv[2])
+wbtdata = numpy.loadtxt(sys.argv[3])
+
+pylab.plot(mcdata[:,0],mcdata[:,1]*4*numpy.pi/3,"b-",label='MC density')
+pylab.plot(dftdata[:,0],dftdata[:,1]*4*numpy.pi/3,"b--",label='WB density')
+pylab.plot(wbtdata[:,0],wbtdata[:,1]*4*numpy.pi/3,"m-.",label='WBT density')
+
+pylab.xlabel("radius")
+pylab.ylabel("filling fraction")
+pylab.legend(loc='upper left', ncol=2).get_frame().set_alpha(0.5)
+
+pylab.savefig(sys.argv[4])

### 2011-11-09 paper/paper.tex
add nice sphere pictures Tested everything.
recompute some of the sphere data Tested everything.
new sphere data Tested everything.

@@
 
+\section{Hard sphere solute}
+
+Figure~\ref{fig:sphere-energy-vs-diameter} shows the hydration free
+energy per unit surface area of a hard sphere.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/sphere-energy-vs-diameter}
+\end{center}
+\caption{ Free energy versus radius for a single hydrophobic sphere
+immersed in water. This should have an asymptote equal to the surface
+tension at room temperature, and it agrees well with the surface tension in
+Figure~\ref{fig:surface-tension}. }
+\label{fig:sphere-energy-vs-diameter}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/sphere-density}
+\end{center}
+\caption{ Density profiles illustrating the transition from vapor 
+to liquid water between the rods. The radius is 0.5~nm, the top figure is 
+at a separation of 0.2~nm and the
+bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
+the energy for these and other separations.}
+\label{fig:sphere-density}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-sphere}
+\end{center}
+\caption{ Density profiles for spheres of different radii. The dotted line
+represents the saturated liquid density.  }
+\label{fig:density-sphere}
+\end{figure}
+
 \section{Conclusion}

### 2011-11-10 paper/paper.tex
make a few changes in the theory section Tested everything.

@@
 We introduce a new classical density functional for water that
-predicts the vapor pressure, vapor density, liquid density, bulk surface tension, 
-and enthalpy, as well 
-as allows for liquid-vapor coexistence.  The Helmholtz free energy functional is
-based on SAFT, and is composed of the usual terms:
+predicts the vapor pressure, vapor density, liquid density, bulk
+surface tension, and enthalpy, as well as allows for liquid-vapor
+coexistence.  The Helmholtz free energy functional is based on SAFT,
+and is composed of the usual terms:
 \begin{equation}
@@
 \subsection{Ideal gas functional}
-The first term is the ideal gas free energy functional,
-which incorporates effects of kinetic energy.  The ideal gas
-functional is given by
+The first term is the ideal gas free energy functional, which
+incorporates effects of kinetic energy.  The ideal gas functional is
+given by
 \begin{equation}\label{idealgas}
@@
 \end{equation}
-where n(\xx) is the density of water molecules and $n_Q$ is the quantum 
-concentration. This is the particle concentration where quantum effects
-become appreciable and is given by
+where n(\xx) is the density of water molecules and $n_Q$ is the
+quantum concentration
 \begin{equation}\label{quantumconcentration}
- n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}
+ n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}.
 \end{equation}
-where $m$ is the molecular mass of water. The ideal gas free
-energy functional has the property that the contact density at a hard
-surface $n_c$ is proportional to the pressure on that surface,
-according to the contact value theorem
+The ideal gas free energy functional satisfies the contact density
+theorem, which states that the contact density at a hard surface $n_c$
+is proportional to the pressure on that surface:
 \begin{equation}\label{contactvaluethm}
@@
 solutes.  These properties are retained by the total functional,
-provided the remaining terms are \emph{purely nonlocal}. This requires
-that
-\begin{equation}
- \frac{\delta^2F_{\textit{excess}}[n]}{\delta n(\xx_1)~\delta n(\xx_1)}\neq\infty
-\end{equation}
-where $F_{\textit{excess}}[n]$ describes all the remaining terms in the functional.
-This ensures that any higher order terms in the Taylor expansion
-of $F[n+\Delta n]$ are small and do not contribute to the small
-solute free energy of solvation.
+provided the remaining terms are \emph{purely nonlocal}.  A common
+failing of SAFT-based functionals (see, for instance
+references\cite{felipe2001examination, gloor2002saft,
+  gloor2004accurate, clark2006developing, gloor2007prediction,
+  kahl2008modified, gross2009density}) is to treat other terms in the
+functional locally, resulting in a functional that is suitable only
+for studying the liquid-vapor interface.
 

### 2011-11-11 paper/paper.tex
TODO list added to paper, abstract updated. Tested everything.

--- 2011-11-09 paper/paper.tex
@@
 
-\author{Dennis L. Jackson}
+\author{Jessica Hughes}
 \affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
-
+\author{Eric Krebs}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
 \author{David Roundy}
@@
 \begin{abstract}
-\needsworklater{ We develop a classical density functional theory for
-  water by combining the fundamental-measure theory functional--which
-  describes hard sphere fluids--with attractive interactions.  Using
-  this approximation, we are able to describe various properties of
-  water near liquid-vapor and liquid-solid interfaces.  We plan to use
-  this functional as the foundation for a efficient method for
-  predicting the interactions of chemicals in aqueous solution on an
+{We develop a classical density functional theory for water by combining the
+fundamental-measure theory functional--which describes hard sphere fluids--with
+attractive interactions based on the Statistical Associating Fluid Theory
+(SAFT-VR).
+Using this approximation with an additional length scale parameter, we are able to describe the vapor pressure, saturated liquid density, surface tension, and enthalpy of water, as well as the number of hydrogen bonds and density oscillations formed near hydrophobic rods and spheres immersed in water.
+\needsworknow{For two rods being pulled apart, a transition occurs 
+from a vapor state to a liquid state in the area between
+the rods at approximately $(\pi-2)$ multipled by the radius. Even 
+for nanoscale rods (radius 0.5~-~1.0~nm), we find that in the
+vapor state, the force can be approximated
+classically and is equal to 144~mN/m, or twice the bulk surface tension.
+(Keep this? Add spheres stuff?)}
+This functional is computationally efficient and provides a foundation for 
+future work describing interactions in aqueous solution on an
 atomic level.}
 \end{abstract}
-
 \maketitle
+
+{\begin{itemize}
+\color{blue}
+\item Put finished things up here
+\item[JRH] Rewrite abstract (copying from project)
+\color{red}
+\item[JRH] Rewrite and join introductions
+\item[JRH] Work on abstract
+\item[DJR] Read and edit changes to abstract
+\item[DJR] Look through theory section for stuff to cut
+\item[DJR,JRH] Find figures to cut
+\item[JRH] Cut the test-particle section
+\item[JRH] Decide what to do with slit (cut?)
+\item[DJR] Write text (or clean up existing) for hard sphere solutes
+\item[EJK] Add 4-rods figures
+\item[EJK] Add 4-rods text
+\end{itemize}
+}
 

### 2011-11-16 paper/paper.tex
updates to paper and turn off zeta (as an experiment). Tested everything.

@@
 
-We model the repulsive interactions using the White Bear version of
+We treat the repulsive interactions using the White Bear version of
 the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
@@
 will include van der Waals attraction and any orientation-independent
-interactions. We use a
-dispersion term based on the SAFT-VR
+interactions. We use a dispersion term based on the SAFT-VR
 approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
 parameters, an interaction energy $\epsilondisp$ and a
-length scale $\lambdadisp$.
+length scale $\lambdadisp R$.
 
@@
 \begin{align}
-  F_\text{disp}[n] &= \int (a_1(\xx) + \beta a_2(\xx))d\xx
+  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx
 \end{align}
@@
 \begin{equation}
-  \varphi(r) = \Theta(r-2 \lambdadisp R)
+  \varphi(r) = \color{red}\Theta(r-2 \lambdadisp R)
 \end{equation}

### 2011-11-17 papers/contact/paper.tex
compute tensor white bear densities in contact paper Tested everything.
clean up contact paper a tad Tested everything.

--- 2011-08-15 papers/contact/paper.tex
@@
 \begin{abstract}
-\needsworklater{ We develop a functional based on FMT~\cite{roth2002whitebear}
- to ... for SAFT}
+  We investigate the use of the White Bear functional,
+  based on Fundamental Measure Theory~\cite{roth2002whitebear} to
+  predict contact densities for the inhomogeneous hard-sphere fluid.
 \end{abstract}
@@
 \caption{A comparison of the various approximations to the contact
-  density.}
+  density in the homogeneous limit.  As expected, they are all
+  identical, and we won't include this plot in the paper (but it's
+  here to verify that our code is working).}
 \label{fig:gHS-vs-n}
@@
 
-\begin{figure}
-\includegraphics[width=5cm]{figs/free-energy}
-\caption{A comparison of the various approximations to the free
-  energy.  Note that the White Bear functional uses a modified form of
-  the Carnahan equation of state.}
-\label{fig:free-energy}
-\end{figure}
+%% \begin{figure}
+%% \includegraphics[width=5cm]{figs/free-energy}
+%% \caption{A comparison of the various approximations to the free energy
+%%   in the homogeneous limit.  Note that the White Bear functional uses
+%%   a modified form of the Carnahan equation of state.  As expected,
+%%   these are all identical, and we won't include this plot in the paper
+%%   (but it's here to verify that our code is working).}
+%% \label{fig:free-energy}
+%% \end{figure}
 
-The hard-sphere contact density---which we will define as the density
-of hard spheres touching a given hard sphere---is important for the
+The hard-sphere contact density---that is to say, the density of hard
+spheres touching a given hard sphere---is important for the
 association term in Statistical Associating Fluid Theory (SAFT).  For
 a homogeneous hard-sphere fluid, the contact density---which is to
-say, $n g_{HS}(\sigma)$---is easily computed from the
+say, $n g_{HS}(\sigma)$, where $g_{HS}(\sigma)$ is the contact value
+of the correlation function---is easily computed from the
 Carnahan-Starling equation of state
@@
 
-\section{Mean Contact density}
+\section{Mean Contact density}\label{simple-contact}
 
@@
 
-\section{Contact density at a given hard sphere}
+\section{Contact density at a given hard
+  sphere}\label{contact-at-sphere}
 
@@
 
+We performed a Monte-Carlo simulation of the hard sphere fluid to
+measure the contact density as a function of position for a few simple
+geometries.  For each scenario, we compute and compare several
+quantities.  We compare the Monte Carlo density with the DFT
+prediction.  We also plot Monte Carlo results for the two contact
+densities: the contact density evaluated at contact, and the contact
+density evaluated at the sphere location.  Alongside these, we plot
+the contact density of Fu and Wu~\cite{fu2005vapor-liquid-dft} (or
+should it be Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}?), the ``simple
+contact density'' from Section~\ref{simple-contact}, and the ``contact
+density at sphere'' from Section~\ref{contact-at-sphere}.  In
+principle, the functional of Wu (FIXME) should match the
+CenConDensity, while the DFT at sphere should match the ConDensity...
+
 \subsection{Hard spheres confined in a spherical cavity}
 
+The first geometry we consider is a number of hard spheres confined
+within a spherical cavity.  In
+Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show
+the densities and contact densities for several cavity sizes.  In each
+case, the number of hard spheres was chosen to set the mean filling
+fraction to around 0.3?.  This value was chosen because it roughly
+corresponds to the filling fraction of water at standard atmospheric
+temperature and pressure in the optimal SAFT model developed by Clark
+\emph{et al}\cite{clark2006developing}.
+
+\textcolor{red}{We should also try looking at a few other densities.}
+
+\newcommand\sphereExplanation{ Blue curves describe the density of
+  hard spheres.  Green curves describe the contact density averaged
+  according to the point of contact.  Red curves describe the contact
+  density as averaged according to the centers of the spheres that are
+  touching.  In each case, solid lines correspond to the Monte Carlo
+  simulation, and dashed lines of various sorts represent DFT model
+  predictions.  }
+
 \begin{figure}
-  \includegraphics[height=5cm]{figs/sphere-radial-06-04}
-  \caption{Density and contact densities of four hard spheres in a
-    spherical cavity with radius 6.}
-  \label{fig:sphere-density}
+  \includegraphics[height=5cm]{figs/contact-08-013}
+  \caption{Density and contact densities of thirteen hard spheres in a
+    spherical cavity with diameter 8. \sphereExplanation }
+  \label{fig:sphere-8}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/sphere-radial-08-13}
-  \caption{Density and contact densities of fourteen hard spheres in a
-    spherical cavity with radius 8.  Also similar densities of spheres
-    in even larger cavities.}
-  \label{fig:sphere-density2}
+  \includegraphics[height=5cm]{figs/contact-12-112}
+  \caption{Density and contact densities of 112 hard spheres in a
+    spherical cavity with diameter 12.  \sphereExplanation}
+  \label{fig:sphere-12}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-16-265}
+  \caption{Density and contact densities of 265 hard spheres in a
+    spherical cavity with diameter 16. \sphereExplanation}
+  \label{fig:sphere-16}
 \end{figure}
@@
 \subsection{Hard spheres confined in a cubical cavity}
+
+\textcolor{red}{
+  I am thinking of also performing simulations for a cubical cavity, and
+  examining the result near the center of a face, which (for a large
+  cavity) will give us a prediction for the hard-sphere fluid near a
+  single hard wall.  We've got some simulations in
+  Figure~\ref{fig:box-density}, but these aren't very nice plots.
+  We'll want to do a much larger box, and plot the density vs $x$ near
+  the wall.  But I'm waiting on simulation data for this...
+}
 
@@
 \section{Conclusion}
-We get a nice spatial dependence for SAFT.
+We (will) have compared several functionals for the contact density of
+a hard-sphere fluid with simulations.  This quantity is of particular
+interest, as it plays a critical role in Statistical Associating Fluid
+Theory (SAFT), which is the basis of a number of recent classical
+density functionals.  We compute the contact density directly from the
+White Bear version of the FMT hard-sphere functional, and find
+reasonable agreement?
 

### 2011-11-18 papers/contact/paper.tex
rewrite a bit of the contact paper Tested everything.

@@
 
+%\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
+\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
+
 % needsworklater is used to annotate bits that need work, but that we
@@
 \begin{abstract}
-  We investigate the use of the White Bear functional,
-  based on Fundamental Measure Theory~\cite{roth2002whitebear} to
-  predict contact densities for the inhomogeneous hard-sphere fluid.
+  We investigate the contact density of an inhomogeneous hard-sphere
+  fluid, which is of particular interest, as it plays a critical role
+  in Statistical Associating Fluid Theory (SAFT), which is the basis
+  of a number of recent classical density functionals.  We derive a
+  formula for the contact density from the White Bear version of the
+  Fundamental Measure Theory functional~\cite{roth2002whitebear}, and
+  test this functional agains Monte Carlo simulations.
+  \textcolor{red}{Insert summary of results here...}
 \end{abstract}
@@
 
-\begin{figure}
-\includegraphics[width=5cm]{figs/gHS-vs-n}
-\caption{A comparison of the various approximations to the contact
-  density in the homogeneous limit.  As expected, they are all
-  identical, and we won't include this plot in the paper (but it's
-  here to verify that our code is working).}
-\label{fig:gHS-vs-n}
-\end{figure}
-
-%% \begin{figure}
-%% \includegraphics[width=5cm]{figs/free-energy}
-%% \caption{A comparison of the various approximations to the free energy
-%%   in the homogeneous limit.  Note that the White Bear functional uses
-%%   a modified form of the Carnahan equation of state.  As expected,
-%%   these are all identical, and we won't include this plot in the paper
-%%   (but it's here to verify that our code is working).}
-%% \label{fig:free-energy}
-%% \end{figure}
-
 The hard-sphere contact density---that is to say, the density of hard
 spheres touching a given hard sphere---is important for the
-association term in Statistical Associating Fluid Theory (SAFT).  For
-a homogeneous hard-sphere fluid, the contact density---which is to
+association term in Statistical Associating Fluid Theory (SAFT).
+For a homogeneous hard-sphere fluid, the contact density---which is to
 say, $n g_{HS}(\sigma)$, where $g_{HS}(\sigma)$ is the contact value
-of the correlation function---is easily computed from the
-Carnahan-Starling equation of state
-\begin{align}
-  g_{HS}(\sigma) &= \frac{1}{1-\eta} + \cdots
-\end{align}
-The contact density may be easily computed from a knowledge of the
-contact theorem, which states that the pressure on any hard surface is
-given by
-\begin{align}
-  p &= k_BT n_\textit{contact}
-\end{align}
-Since we are interested in the contact density at the hard-sphere
-surface, all we need compute is the pressure on that surface, and
+of the correlation function---is well-known.  However, for
+inhomogeneous fluids, the contact density has not been well studied.
+
+\subsection{Association free energy of SAFT}
+\newcommand\epsilonassoc{\ensuremath{\varepsilon_\textit{AB}}}
+\newcommand\kappaassoc{\ensuremath{\kappa_\textit{AB}}}
+\newcommand\ncontact{\ensuremath{n_\textit{contact}}}
+
+The free energy term in Statistical Associating Fluid Teory (SAFT)
+from which it derives its name is the association term, which accounts
+for hydrogen bonding.  Hydrogen bonds are modeled as attractive
+patches (``association sites'') on the surface of hard spheres.  These
+sites represent protons or electron lone pairs, and have an attractive
+energy $\epsilonassoc$ when two molecules are oriented such that the
+proton of one overlaps with the lone pair of the other.  The volume
+over which this interaction occurs is $\kappaassoc$, giving the
+association term in the free energy has two empirical parameters fit
+to experimental data.
+
+The association free energy per unit volume has the form
+\begin{align}
+  f_\text{assoc} &= k_BT n\sum_A 
+                  \left(\ln X^A - \frac{X^A}{2} + \frac12\right)
+\end{align}
+where the summation is over the association sites, and $X^A$ is the
+fraction of association sites \emph{not} hydrogen-bonded.  The
+fractions $X^A$ are determined by the self-consistent equations
+\begin{align}
+  X^A &= \frac{1}
+  {1 + \sum_B n X^B \Delta^{AB}}\label{eq:X}
+  \\
+  \Delta^{AB} &= g(\sigma) \kappaassoc\left( e^{\epsilonassoc/kT} -
+  1 \right)\label{eq:delta}
+\end{align}
+where $g(\sigma)$ is the correlation function evaluated at contact.
+The product of density with correlation function evaluated at contact
+gives the contact density, when we combine Equations~\ref{eq:X} and
+\ref{eq:delta}:
+\begin{align}
+  X^A &= \frac{1}
+  {1 + \sum_B \ncontact\kappaassoc X^B\left( e^{\epsilonassoc/kT} -
+  1 \right)}
+\end{align}
+where the product $\ncontact\kappaassoc$ is the expected number of
+molecules present in the association volume in the absence of the
+association interaction.  In the SAFT-VR
+model\cite{gil-villegas-1997-SAFT-VR}, this contact density is found
+by adding a perturbative correction to the contact value for a hard
+sphere fluid.
+
+\subsection{Homogeneous limit}
+
+In order to motivate the derivation of our functional for contact
+density, we begin by deriving the well-known contact density for a
+homogeneous hard-sphere fluid from the Carnahan-Starling free energy.
+The contact density is most easily computed from the contact density
+theorem, which states that the pressure on any hard surface is
+determined by the density at contact:
+\begin{align}
+  p &= k_BT n_\textit{contact}.
+\end{align}
+Since we are interested in the contact density at the surface of the
+hard spheres, all we need compute is the pressure on that surface, and
 we'll have our answer.  The pressure on a hard sphere can be readily
-computed from the dependence of the free energy on hard sphere
-radius.
+computed from the dependence of the Carnahan-Starling free energy on
+hard sphere radius.
 \begin{align}
@@
 \end{align}
-Since $\eta \equiv \frac{4\pi}{3} R^3$, we can work out that
+where $\eta \equiv \frac{4\pi}{3} R^3 n$ is the filling fraction.  We
+can thus readily compute the derivative of free energy with respect to
+hard-sphere radius:
 \begin{align}
   \frac{dA_{HS}}{dR} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{dR} \\
+  \derivation{
     &= Nk_BT \left( \frac{4 - 6\eta}{(1-\eta)^2} + 2 \frac{4\eta - 3\eta^2}{(1-\eta)^3} \right) \frac{d\eta}{dR}
@@
     \\
+  }
   &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \label{eq:dAhsdR}
@@
 the pressure on the spheres, we just need to divide by the total area,
-which means dividing by $N 4\pi (2R)^2$.  The area which is relevant
+which means dividing by $N 4\pi (2R)^2$, since the relevant area
 is the area over which the molecules can make contact.
@@
   p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
+  \derivation{
     &= \frac14 \frac{1}{N 4\pi R^2} Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \\
-  &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
-  &= n k_BT \frac{1 - \frac{\eta}2}{(1-\eta)^3}
-\end{align}
+    &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 -
+      \frac{\eta}2}{(1-\eta)^3} \\
+  }
+  &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
+\end{align}
+Using the contact-value theorem, we can thus find the contact density
+and the well-known correlation function evaluated at contact.
 \begin{align}
@@
 \end{align}
-As it turns out, this is the standard answer for the contact value of
-the autocorrelation function! $\ddot\smile$
+Thus we can readily derive the contact density for the homogeneous
+hard-sphere fluid.  Extending this to the inhomogeneous fluid will
+require that we use density-functional theory.
 
@@
 \end{align}
+
+\section{Direct correlation function}
+
+\textcolor{red}{FIXME WRITE THIS SECTION!} The direct correlation
+function is readily accessible from density-functional theory, as the
+second functional derivative of the free energy:
+\textcolor{red}{(Something like...)}
+\begin{align}
+  -c^{(2)}(\xx_1, \xx_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\xx_1) \delta n(\xx_2)}
+\end{align}
+The direct correlation function is related to the ordinary correlation
+function by the Ornstein-Zernike relation
+\begin{align}
+  h(\xx_1, \xx'_2) &= g(\xx_1, \xx_2) - 1 \\
+  h(\xx_1, \xx_2) &= c(\xx_1, \xx_2)
+    + \int n(\xx_3)c(\xx_1, \xx_3) h(\xx_2, \xx_3)d\xx''
+\end{align}
+From these two equation, we could in principle extract the correlation
+function evaluated at contact from the DFT free energy functional.
+The Ornstein-Zernike equation is usually solved by taking a Fourier
+transform of all the relevant quantities.  In the inhomogeneous case,
+however, this approach isn't workable, since the Fourier transforms
+don't speed things up like you'd hope, and we don't have a periodic
+system in general.
+
+\textcolor{red}{It seems worth looking into this, to see what (if
+  anything) we can do! I don't see a solution just now, but that
+  doesn't mean there isn't one around somewhere.  In any case, just
+  looking at the direct correlation function might give us some
+  hints as to a different and interesting approximation for the
+  contact density.}
 
@@
 
+\appendix
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/gHS-vs-n}
+\caption{A comparison of the various approximations to the contact
+  density in the homogeneous limit.  As expected, they are all
+  identical, and we won't include this plot in the paper (but it's
+  here to verify that our code is working).}
+\label{fig:gHS-vs-n}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/free-energy}
+\caption{A comparison of the various approximations to the free energy
+  in the homogeneous limit.  Note that the White Bear functional uses
+  a modified form of the Carnahan equation of state.  As expected,
+  these are all identical, and we won't include this plot in the paper
+  (but it's here to verify that our code is working).}
+\label{fig:free-energy}
+\end{figure}
+
 \begin{widetext}
-  \appendix
-
   \section{Partial derivatives of the White Bear free energy functional}

### 2011-11-22 papers/contact/paper.tex
a bit of work on contact paper Tested everything.

@@
 
+{
+  \color{blue}We need to figure out a few bits of terminology.
+
+  \begin{enumerate}
+  \item What should we call the contact density between two hard
+    spheres, to distinguish it from the contact density at any other
+    hard surface.  One option would be to focus on the contact value
+    of the correlation function, which would make it clear what we're
+    talking about.
+  \item How should we distinguish between the various definitions.  In
+    particular, ``at the sphere'' vs. ``at contact''.  It'd be helpful
+    to have a short way of describing these.
+  \end{enumerate}
+}
+
 \subsection{Association free energy of SAFT}
@@
 
+\derivation{
   \begin{widetext}
+}
 
@@
 \begin{align}
-  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR}
-\end{align}
-So to compute this we need
+  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
+  \ncontact &= \frac{1}{N k_BT 4\pi (2R)^2} \frac{dA_{HS}}{dR}
+\end{align}
+While this should give the true mean contact density for any given
+system, this is not particularly useful, as we have no way of knowing
+the contact density on any given sphere.  However, the mean contact
+density comes out as an integral, so we can plot the integrand, and
+treat it as a somewhat ill-defined local contact density.
+
+\derivation{
+
+  To compute the mean contact density of a system, we need
   \begin{align}
@@
   \end{align}
+}
 
@@
 
-When we consider an inhomogeneous system, we may ask for \emph{which}
-sphere we are interested in computing the contact density.  One way to
-answer this is to specify a sphere at a fixed position $\mathbf{r}$.
-By making the hard-sphere radius dependent on position, we can compute
-the pressure on the hard spheres located at $\mathbf{r}$.
-\begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{4\pi (2R)^2} \frac{\delta A_{HS}}{\delta R(\mathbf{r})}
-\end{align}
+When we consider an inhomogeneous system, we may prefer to ask for
+\emph{which} sphere we are interested in computing the contact
+density.  One way to answer this is to specify a sphere at a fixed
+position $\mathbf{r}$.  By making the hard-sphere radius dependent on
+position, we can compute the pressure on the hard spheres located at
+$\mathbf{r}$.
+\begin{align}
+  p_{HS}(\mathbf{r}) &= \frac{1}{4\pi (2R)^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \\
+  \ncontact(\xx) &= \frac{1}{k_BT 4\pi (2R)^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})}
+\end{align}
+After considerable math to do the above functional derivative, we find
+a functional giving the contact density for spheres at any given
+position.  This functional unavoidably involves more convolutions than
+are present in the free energy itself, which makes it relatively
+expensive.
+
+\derivation{
   So to compute this we need
@@
   this, but it wouldn't be efficient or pretty either.
+}
 
@@
 
-Unfortunately, knowing the average contact density on each hard sphere
-doesn't really give us what we want for SAFT.  The trouble is that in
-order to work out the fraction of association sites occupied, we need
-to know the contact density \emph{at the point of contact}.
-Fortunately, we don't need to know the contact density for a
-particular sphere at each point of contact.  Just as $n_0(\mathbf{r})$
-gives us the density of spheres that are contacting point $\mathbf{r}$
-(see Figure~\ref{fig:n0}), we can use similar reasoning to work out the
-contact density for these spheres.
+A final definition of the inhomogeneous contact density would be the
+contact density for spheres \emph{touching at a given point}.  This
+would be a symmetric measure, since both spheres involved are treated
+in the same way.  In fact, the density of such spheres is just the
+density $n_0(\xx)$ from FMT.  See Figure~\ref{fig:n0} for an
+illustration $n_0$.  Since $n_0$ is the fundamental measure of sphere
+surface contact, it would make sense to define a contact density that
+would correspond.
 
@@
 would give us the pressure averaged over the entire surface of those
-spheres.  Instead, we'll have to imagine extending a patch of those
+spheres.  Instead, we would have to imagine extending a patch of those
 spheres which is located at the point of interest, as illustrated in
-Figure~\ref{fig:contact}.
+Figure~\ref{fig:contact}.  \textcolor{red}{I'm not sure how to go
+  about doing this.  In the derivation section, I have some ideas
+  jotted down, but they don't really work.}
+
+Yu and Wu developed a functional that should correspond to the contact
+density at a given contact
+position~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We will
+compare both their functional as well as ours with simulation data.
 
@@
 \end{figure}
+
+\derivation{
 
@@
   \end{align}
-
+}
+
+\derivation{
   \end{widetext}
+}
 
@@
   the wall.  But I'm waiting on simulation data for this...
+}
+
+\subsection{Hard spheres around a hard sphere (or rod)}
+
+\textcolor{red}{
+  It would also be interesting to study a system with opposite
+  curvature.  Put a (larger) hard sphere in the hard-sphere fluid, and
+  see what the contact density is around that hard sphere.  This
+  touches on an important limiting case...
 }

### 2011-11-22 papers/contact/figs/contact-density.py

@@ -23,7 +23,7 @@ pylab.plot(mcdata[:,0],mcdata[:,2]*4*numpy.pi/3,"r-",label="ConDensity")
 pylab.plot(mcdata[:,0],mcdata[:,3]*4*numpy.pi/3,"g-",label="CenConDensity")
 
 pylab.plot(dftdata[:,0],dftdata[:,3]*4*numpy.pi/3,"g+--",label="simple contact")
-pylab.plot(dftdata[:,0],dftdata[:,4]*4*numpy.pi/3,"gx--",label="Wu and Fu")
+pylab.plot(dftdata[:,0],dftdata[:,4]*4*numpy.pi/3,"gx--",label="Yu and Wu")
 pylab.plot(dftdata[:,0],dftdata[:,5]*4*numpy.pi/3,"ro--",label="DFT at sphere")
 
 pylab.xlabel("radius")

### 2011-12-02 paper/paper.tex
Work on paper, combining intro, deleting unnecessary figures and sections, cleaning up 1 rod section. Tested everything.

@@
 (SAFT-VR).
-Using this approximation with an additional length scale parameter, we are able to describe the vapor pressure, saturated liquid density, surface tension, and enthalpy of water, as well as the number of hydrogen bonds and density oscillations formed near hydrophobic rods and spheres immersed in water.
+Using this approximation with an additional length scale parameter, we are able to calculate the vapor pressure, saturated liquid density, surface tension, and enthalpy of water, as well as the number of hydrogen bonds and density oscillations formed near hydrophobic rods and spheres immersed in water.
 \needsworknow{For two rods being pulled apart, a transition occurs 
@@
 classically and is equal to 144~mN/m, or twice the bulk surface tension.
-(Keep this? Add spheres stuff?)}
+(Edit this and add spheres stuff)}
 This functional is computationally efficient and provides a foundation for 
-future work describing interactions in aqueous solution on an
+future work describing complex interactions in aqueous solution on an
 atomic level.}
+\tableofcontents
 \end{abstract}
@@
 \item[JRH] Rewrite abstract (copying from project)
+\item[JRH] Cut the test-particle section
+\item[JRH] Rewrite and join introductions
+\item[JRH] Decide what to do with slit (cut)
+\item[JRH] Combine and edit Other Thermo and Empirical Parameters sections
 \color{red}
-\item[JRH] Rewrite and join introductions
 \item[JRH] Work on abstract
-\item[DJR] Read and edit changes to abstract
+\item[DJR] Read and edit changes to abstract and introduction
 \item[DJR] Look through theory section for stuff to cut
-\item[DJR,JRH] Find figures to cut
-\item[JRH] Cut the test-particle section
-\item[JRH] Decide what to do with slit (cut?)
+\item[DJR,JRH] Finalize figures in beginning
+\item[JRH] Better/expanded description of ``macroscopic surface tension'' model in 2 rods section?
+\item[JRH] Recalculate and clean up 2 rods figures and data - new code?
 \item[DJR] Write text (or clean up existing) for hard sphere solutes
@@
 \item[EJK] Add 4-rods text
+\item[JRH,DJR] Rewrite conclusion
 \end{itemize}
@@
 Due to its large dielectric constant, small molecular size and strong
-hydrogen-bonding network, water as ``the universal solvent'' are as
+hydrogen-bonding network, water as ``the universal solvent'' is as
 interesting as its anomalous thermodynamic properties.
 
-\begin{figure}
-\includegraphics[width=6cm, angle=270]{figs/correlation}
-\caption{ Experimental atom-atom radial distribution functions for water at
-room temperature and pressure from reference~\cite{Soper2000}.  }
-\label{correlation}
-\end{figure}
-
 \subsection{Classical density-functional theory}
+
+%\begin{figure}
+%\includegraphics[width=6cm, angle=270]{figs/correlation}
+%\caption{ Experimental atom-atom radial distribution functions for water at
+%room temperature and pressure from reference~\cite{Soper2000}.\needsworknow{Either 
+%mention this figure in the text or delete it.}  }
+%\label{correlation}
+%\end{figure}
 
@@
 accuracy close to that attained by \emph{ab initio} calculations.  The
-theory of Lum, Chandler and Weeks (LCW)~\cite{LCW}, for instance, can
+theory of Lum, Chandler and Weeks (LCW)~\cite{lum1999hydrophobicity}, for 
+instance, can
 accurately describe the free energy cost of creating a cavity by placing a
@@
 for evaluating the free energy and thermally averaged density of fluids in
-an arbitrary external potential~\cite{ebner1976}.  Classical DFT is a
+an arbitrary external potential~\cite{ebner1976density}. The Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
+energy in terms of a universal functional of the electron density and a separate
+term including the external potential. Mermin\cite{mermin1965thermal} gives a
+theorem extending that of Hohenberg and Kohn to nonzero temperatures:
+\begin{equation}
+  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
+d^3r\right\}.
+\end{equation}
+This is the basis for classical DFT, where the relevant density is not the
+electron density, but the density of molecules. Classical DFT is a
 natural framework for creating a more flexible theory of hydrophobicity
@@
 small hard solutes.  Much of the research on classical density-functional
-theory has focussed on the hard-sphere fluid~\cite{curtin1985,
+theory has focused on the hard-sphere fluid~\cite{curtin1985,
 rosenfeld1989, rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.
@@
 rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.  This functional
-\needsworklater{has many nice properties}.
+\needsworknow{has many nice properties}.
 
@@
 
-\needsworklater{
-Density-Functional Theory is most commonly used for calculating the
-densities of electrons in quantum mechanical systems.  We take the
-classical limit of this theory to construct a functional that will predict
-the density of water molecules in a classical system.  The model used here
+\needsworknow{Add smooth transition from others' work to ours}
+
+We use classical DFT to construct a new functional that will predict
+the density of water molecules in a classical system.  The model we use
 starts with a treatment of each individual water molecule as a purely
-classical hard sphere.  The only constraint in this model is
-the minimum distance between any two molecules is twice the hard sphere
-radius.}
-
-\needsworklater{ I still need to rewrite the entire introduction... }
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Introduction}
-
-\subsection{Water---the universal solvent}
-
-The simple yet complex water molecule is one of the most important 
-substances for life and industry. The structure and geometry of the 
-water molecule leads to a strong polarity and complex intermolecular
-forces. The hydrogen bonds formed by water are very strong, creating
-tetragonal clusters that lead to many of water's unusual properties,
-such as the high melting, boiling, and critical temperatures as well as the
-density maximum above the melting point.
-The large dielectric constant is partially responsible for
-water acting as a universal solvent, along with the ease of availability and
-use.
-There are many models used to describe water since so many different
-processes occur in aqueous solution. It is challenging to describe all the 
-relevant properties of water at once, especially if a low computational cost is
-also desired.
-
-Some models are primarily designed for use with homogeneous systems. Although
-these 
-models are useful in predicting some bulk properties, the interesting
-chemistry that occurs in water cannot be described with a homogeneous-only
-model. To
-describe solvation and other processes in aqueous solution, a more robust theory
-that 
-can handle varying densities is needed.
-
-\subsection{Classical density-functional theory}
-
-Classical density-functional theory (DFT) provides an efficient way to evaluate
-the free energy using the probability distribution of the fluid. The free energy 
-is a functional
-of the density function and is independent of any external
-potentials\cite{ebner1976density}. The most common use of DFT is calculating ground
-state electron densities in quantum mechanical systems. The Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
-energy in terms of a universal functional of the electron density and a separate
-term including the external potential. Mermin\cite{mermin1965thermal} gives a
-theorem extending that of Hohenberg and Kohn to nonzero temperatures:
-\begin{equation}
-  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
-d^3r\right\}
-\end{equation}
-This is the basis for classical DFT, where the relevant density is not the
-electron density,
-but the density of molecules. This can be used with a variety of 
-external potentials, from the simple interaction with a 
-hydrophobic wall to the more complex, stronger interaction with hydrophilic
-solutes.
-
-\subsection{Statistical Associating Fluid Theory}
-
-The high hydrogen bonding and strong polar interactions in water characterize it
-as an associating fluid, and thus it requires a different treatment than simple
+classical hard sphere and adds association terms based on Statistical 
+Associating Fluid Theory (SAFT). Water, unlike simple
 fluids which are mainly defined by van der Waals forces and weak electrostatic
-interations. The Statistical Associating Fluid Theory (SAFT) was designed for
-use with fluids such as water, to account for the important association forces
-(like hydrogen bonds) as well as any effects due to clusters or chains forming
-in the liquid\cite{muller2001molecular}. It is based on Wertheim's first-order
+interations, requires direct treatment of association effects. The strong 
+hydrogen bonds and polar
+interactions in water are ideal for describing with SAFT, which was designed for
+use with such fluids. It is based on Wertheim's first-order
 thermodynamic perturbation theory (TPT1)
@@
 wertheim1986fluidsIV} which provides a useful way to account for strong
-associative interations between molecules.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+associative interations between molecules. SAFT can account for the association energies
+as well as any effects due to clusters or chains forming
+in the liquid\cite{muller2001molecular}. \needsworknow{Expand or move this 
+introduction to SAFT?}
+
 \section{Theory and Methods}
@@
 
-\subsection{Other thermodynamic functions}
-
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
-\end{equation}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
-\end{figure}
-
-The pressure would naturally be obtained by taking a partial
-derivative with respect to volume at fixed temperature.  Since
-\emph{volume} isn't a parameter in density-functional theory, we
-consider the \emph{free energy per unit volume}:
-\begin{align}
-  F[n] &= f[n]V \\
-  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
-  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
-  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
-   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
-  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n].
-\end{align}
-
-Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
-isotherms ranging from room temperature to past the critical temperature. The
-solid line is the theoretical coexistence curve calculated from the SAFT
-functional. For a given isotherm, the pressure curve should intersect this
-coexistence curve twice, at the vapor and liquid coexistence 
-pressure corresponding to the vapor and liquid densities. Even
-though the theoretical pressure isotherms drop below the coexistence curve,
-this is non-physical and will not actually occur. The experimental data shows
-how the actual density value will ``jump'' to the other side of 
-the coexistence curve at that pressure value (for
-temperatures less than the critical temperature), corresponding to a phase 
-change in the material. For example, follow the room temperature isotherm
-starting from high pressure and high density (top right of 
-Figure~\ref{fig:pressure-with-isotherms}).
-As the pressure decreases, the density decreases \emph{slightly} until reaching
-the coexistence curve, at which point a phase change from liquid to vapor
-will occur. The
-functional matches well with the experimental data from NIST away from the
-critical point. The slope of the pressure versus density curve is proportional
-to the compressibility, which is close but slightly too large at room
-temperature.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
-  will match the experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
-\end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
-\end{figure}
-
-The entropy is naturally obtained by taking a partial derivative with
-respect to temperature:
-\begin{align}
-  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V}
-\end{align}
-This derivative is trivial for the ideal gas and hard sphere free
-energies, which are simply proportional to the temperature.  For the
-association and dispersion free energies, it is a little more work,
-but still not hard.
-
-Once we have the entropy, finding the internal energy is simple:
-\begin{align}
-  U[n] &= F[n] + TS[n]
-\end{align}
-and the enthalpy is not much more work:
-\begin{align}
-  H[n] &= F[n] + TS[n] + pV.
-\end{align}
-The calculated heat of vaporization at atmospheric pressure is
-$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
-373 K. This compares well to the experimental
-value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
-an error of only a few percent.
-
 \subsection{Determining the empirical parameters}\label{sec:empirical}
@@
 
+From the Helmholtz free energy functional, we may obtain any other
+thermodynamic functions we should choose.  The grand free energy
+$\Phi$, which we ordinarily use in our computations, is obtained by
+simply adding a chemical potential term:
+\begin{equation}
+  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
+\end{equation}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
+\label{fig:pressure-with-isotherms}
+\end{figure}
+
+The pressure would naturally be obtained by taking a partial
+derivative with respect to volume at fixed temperature. 
+Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
+isotherms ranging from room temperature to past the critical temperature. The
+solid line is the theoretical coexistence curve calculated from the SAFT
+functional. For a given isotherm, the pressure curve should intersect this
+coexistence curve twice, at the vapor and liquid coexistence 
+pressure corresponding to the vapor and liquid densities. Even
+though the theoretical pressure isotherms drop below the coexistence curve,
+this is non-physical and will not actually occur. The experimental data shows
+how the actual density value will ``jump'' to the other side of 
+the coexistence curve at that pressure value (for
+temperatures less than the critical temperature), corresponding to a phase 
+change in the material. The
+functional matches well with the experimental data from NIST away from the
+critical point. The slope of the pressure versus density curve is proportional
+to the compressibility, which is close but slightly too large at room
+temperature.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
+  tension will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\columnwidth]{figs/equation-of-state}
+%\end{center}
+%\caption{The theoretical versus experimental phase diagram, vapor
+%  pressure versus temperature.  }
+%\label{fig:equation-of-state}
+%\end{figure}
+
+%The entropy is naturally obtained by taking a partial derivative with
+%respect to temperature:
+%\begin{align}
+%  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V}
+%\end{align}
+%This derivative is trivial for the ideal gas and hard sphere free
+%energies, which are simply proportional to the temperature.  For the
+%association and dispersion free energies, it is a little more work,
+%but still not hard.
+
+%Once we have the entropy, finding the internal energy is simple:
+%\begin{align}
+%  U[n] &= F[n] + TS[n]
+%\end{align}
+%and the enthalpy is not much more work:
+%\begin{align}
+%  H[n] &= F[n] + TS[n] + pV.
+%\end{align}
+
+The calculated heat of vaporization at atmospheric pressure is
+$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
+373 K. This compares well to the experimental
+value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
+an error of only a few percent.
+
 \begin{figure}
@@
 \caption{The theoretical versus experimental coexistence curve. Away from
-the critical temperature, the agreement is good. }
+the critical temperature, the agreement is good.\needsworknow{(keep figure?
+its the same as Clark paper)} }
 \label{fig:temperature-vs-density}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
-\includegraphics[width=\columnwidth]{figs/near-critical-point}
-\end{center}
-\caption{By plotting the free energy versus density it is possible to
-  determine a saturated liquid or vapor state by the common tangent
-  method. The top plot illustrates the common tangent near room 
-  temperature, while the bottom plot shows the same thing close to the 
-  critical temperature. }
-\label{fig:homogeneous}
-\end{figure}
-
-For reference, we plot the predicted vapor pressure (see
-Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
-Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
-match those calculated by Clark \emph{et al}. The functional provides a good description of
-the vapor pressure, which has an error of less than 
-1\% compared with the experimental data. The coexistence curve in 
-Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
-curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
-temperature coexistence values.
-Near the critical temperature,
-the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
-experiment. 
-To find the densities for the coexistence curve, we use the
-``common tangent'' rule. The two densities 
-will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
-As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
-(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
-in correctly treating the critical temperature. As the vapor and liquid densities approach
-each other, the free energy barrier between the two states gets smaller and it becomes
-challenging to computationally find the correct minimum. This is the reason 
-why the coexistence curves
-in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
-are not smooth around the critical point.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and discussion}
-
-\subsection{Water confined to a slit}
-
-One of the most computationally simple ways to study confined water is by
-constraining it to a slit where the water molecules can only move in one
-dimension. Experimentally, this could represent two different situations.
-The first, water between two perfectly aligned and smooth planes, is not common 
-in experiment due to the difficulty of creating such an ideal situation. The
-second
-possibility is a large sphere brought close to a plane, which eliminates the 
-need for perfect alignment.
-
-Before looking at the liquid-solid one-dimensional slit, we set up a liquid-vapor
+\subsection{Liquid-vapor interface}
+
+We set up a liquid-vapor
 interface and plot the density profile.
@@
 density-functional theory would include them), with the result
-that the density profile will be smoothed out by capillary wave effects. The liquid-vapor
+that the density profile will be smoothed out by capillary wave effects. 
+The liquid-vapor
 interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
@@
 
-To set up the calculations for the slit, we choose a chemical potential 
-corresponding to a pressure of 1~atm and set the
-temperature to a typical room temperature of 298~K. We apply 
-a constraining potential outside the slit, where we choose a small non-zero
-density for numerical reasons. The slits studied range in size from 0.1~nm to 
-8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
-lattice
-with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D}
-\end{center}
-\caption{ Density profile for a slab of water with molecules 
-constrained to move in one dimension. The slit size is 4.0~nm and the dashed
-line represents the saturated liquid density.}
-\label{fig:density-1D}
-\end{figure}
-
-The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
-Near the interface, there are oscillations around the
-liquid density which decrease near the center of the slit. We expect the center
-of the slit to act approximately like bulk water, and the absence of
-oscillations supports this. For larger slits (6.0~nm or more) the majority of
-the water molecules in the slit act like the bulk.
-
-Based on our computations, we determine the smallest slit that can contain 
-liquid water. This result is
-approximately 0.6~nm. For smaller slits, the initial liquid water inside will
-vaporize. The results for larger liquid-filled slits are only metastable,
-however, as a vapor-filled slit is thermodynamically favorable. Based on our
-calculations for a pressure of 1~atm, the work needed to pull apart
-the two hard walls comprising the slit is one to two orders of magnitude less
-than the free energy of the liquid-filled slit. For the range of slit sizes we
-studied, all the computational results of liquid-filled slits are
-metastable.
-
-Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
-stability of water between large hydrophobic plates and the critical separation
-necessary for stability. For water at room temperature and atmospheric pressure,
-plates with a separation smaller than about 100~nm will not contain stable liquid
-water. They go on to say that confined water has a region of metastability down
-to separations of about 5~nm. This is comparable to the slit separation where
-our calculations show water inside the slit acting like the bulk liquid, which
-we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
-also study water between two hard walls at constant chemical potential. They 
-also conclude that the liquid state is likely metastable, and there is a free
-energy barrier to reach the stable vapor phase.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D}
-\end{center}
-\caption{The number of hydrogen bonds per molecule for a slab of water 
-with molecules constrained to move in one dimension. Each molecule may
-contribute to
-four bonds, although each bond involves two molecules so the
-actual number of bonds would be halved.}
-\label{fig:xassoc-1D}
-\end{figure}
-
-The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
-molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
-Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
-bonds is approximately equal to 3.5, which is expected for saturated liquid
-water.
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
+%\includegraphics[width=\columnwidth]{figs/near-critical-point}
+%\end{center}
+%\caption{By plotting the free energy versus density it is possible to
+%  determine a saturated liquid or vapor state by the common tangent
+%  method. The top plot illustrates the common tangent near room 
+%  temperature, while the bottom plot shows the same thing close to the 
+%  critical temperature. }
+%\label{fig:homogeneous}
+%\end{figure}
+
+%For reference, we plot the predicted vapor pressure (see
+%Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
+%Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
+%match those calculated by Clark \emph{et al}. 
+%The functional provides a good description of
+%the vapor pressure, which has an error of less than 
+%1\% compared with the experimental data. The coexistence curve in 
+%Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
+%curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
+%temperature coexistence values.
+%Near the critical temperature,
+%the agreement is off, and the theoretical prediction  of the temperature 
+%is quite a bit higher than
+%experiment. 
+%To find the densities for the coexistence curve, we use the
+%``common tangent'' rule. The two densities 
+%will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
+%As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
+%(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
+%in correctly treating the critical temperature. As the vapor and liquid densities approach
+%each other, the free energy barrier between the two states gets smaller and it becomes
+%challenging to computationally find the correct minimum. This is the reason 
+%why the coexistence curves
+%in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
+%are not smooth around the critical point.
+% 
+% To set up the calculations for the slit, we choose a chemical potential 
+% corresponding to a pressure of 1~atm and set the
+% temperature to a typical room temperature of 298~K. We apply 
+% a constraining potential outside the slit, where we choose a small non-zero
+% density for numerical reasons. The slits studied range in size from 0.1~nm to 
+% 8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
+% lattice
+% with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
+% 
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/density-1D}
+% \end{center}
+% \caption{ Density profile for a slab of water with molecules 
+% constrained to move in one dimension. The slit size is 4.0~nm and the dashed
+% line represents the saturated liquid density.}
+% \label{fig:density-1D}
+% \end{figure}
+% 
+% The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
+% Near the interface, there are oscillations around the
+% liquid density which decrease near the center of the slit. We expect the center
+% of the slit to act approximately like bulk water, and the absence of
+% oscillations supports this. For larger slits (6.0~nm or more) the majority of
+% the water molecules in the slit act like the bulk. 
+% 
+% Based on our computations, we determine the smallest slit that can contain 
+% liquid water. This result is
+% approximately 0.6~nm. For smaller slits, the initial liquid water inside will
+% vaporize. The results for larger liquid-filled slits are only metastable,
+% however, as a vapor-filled slit is thermodynamically favorable. Based on our
+% calculations for a pressure of 1~atm, the work needed to pull apart
+% the two hard walls comprising the slit is one to two orders of magnitude less
+% than the free energy of the liquid-filled slit. For the range of slit sizes we
+% studied, all the computational results of liquid-filled slits are
+% metastable. 
+% 
+% Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
+% stability of water between large hydrophobic plates and the critical separation
+% necessary for stability. For water at room temperature and atmospheric pressure,
+% plates with a separation smaller than about 100~nm will not contain stable liquid
+% water. They go on to say that confined water has a region of metastability down
+% to separations of about 5~nm. This is comparable to the slit separation where
+% our calculations show water inside the slit acting like the bulk liquid, which
+% we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
+% also study water between two hard walls at constant chemical potential. They 
+% also conclude that the liquid state is likely metastable, and there is a free
+% energy barrier to reach the stable vapor phase. 
+% 
+% \begin{figure}
+% \begin{center}
+% \includegraphics[width=\columnwidth]{figs/xassoc-1D}
+% \end{center}
+% \caption{The number of hydrogen bonds per molecule for a slab of water 
+% with molecules constrained to move in one dimension. Each molecule may
+% contribute to
+% four bonds, although each bond involves two molecules so the
+% actual number of bonds would be halved.}
+% \label{fig:xassoc-1D}
+% \end{figure}
+% 
+% The value of $X$ behaves as expected for water constrained to a slit. We
+% show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
+% molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
+% Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
+% bonds is approximately equal to 3.5, which is expected for saturated liquid
+% water.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Results and discussion}
+
+\subsection{One hydrophobic rod}
 
@@
 
-\subsection{One hydrophobic rod}
-
 We move into two dimensions by studying a single hydrophobic rod
-immersed in water. Figure~\ref{fig:density-single-rod} shows density
+immersed in water. We use a lattice spacing of about 0.01~nm (0.2 bohrs), 
+which is a reasonable resolution based on several lattice spacing calculations.
+The results are within 0.5\% of the free energy and a few percent of the contact
+density for the highest resolution tested.
+Figure~\ref{fig:density-single-rod} shows density
 profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
-\end{center}
-\caption{ Number of bonds per molecule near the interface with a single
-hydrophobic rod. There is a decrease in the hydrogen bonding near the surface 
-of the rod, as expected at the interface of any hydrophobic wall. }
-\label{fig:xassoc-single-rod}
-\end{figure}
-
-\begin{figure}
-\begin{center}
 \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
@@
 tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. }
+Figure~\ref{fig:surface-tension}.\needsworknow{Add more data at small radius 
+and include an inset.} }
 \label{fig:energy-vs-diameter}
-\end{figure}
-
-The value of $X$ is calculated and converted to number of bonds per molecule and 
-shown for a rod with radius 0.2~nm in Figure~\ref{fig:xassoc-single-rod}. Inside the
-rod (less than 0.2~nm) there are no molecules to form hydrogen bonds although the value
-of $X$ is still calculated there. The number of bonds decreases from a bulk value of 3.5 
-to about 3.25 at the interface. Walther \emph{et al.}\cite{walther2001carbon} 
-calculate the number of hydrogen bonds for a carbon nanotube immersed in water.
-They show a decrease in the bonding from a bulk value of 3.75 to 2.89 near the 
-interface with the nanotube. This is a much larger change than for our 
-rod of a similar radius, indicating that carbon nanotubes disrupt hydrogen
-bonds in water more than hydrophobic rods.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-vs-radius}
-\end{center}
-\caption{ Density profiles for 
-a single hydrophobic rod with a radius of 0.5~nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
-medium 0.2~bohrs ($\approx$~0.01~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
-\label{fig:densityresolution}
 \end{figure}
@@
 
-\subsubsection{Resolution and convergence tests}
-Figure~\ref{fig:densityresolution} shows the density profiles for three
-different values of the lattice spacing resolution. The density oscillations
-around a single rod of radius 0.5~nm look nearly identical for all resolutions
-except for the first peak. This contact density is important as it will be
-proportional to the pressure on the rod (see Equation~\ref{contactvaluethm}). The contact
-density calculated using the medium resolution is within less than 4\% of the
-value at high resolution, while the low resolution contact density is off by
-about 5\%. There is some uncertainty in the value of contact density due to the
-different lattice spacings which cause the contact density to be calculated
-at a slightly different position. The total free energy calculations compare much more
-favorably. The medium resolution is within 0.5\% and the low resolution is
-within 1\% of the high resolution free energy.
-
 \subsection{Hydrophobic interaction of two rods}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-diagram}
-\end{center}
-\caption{ Cross section schematic for two hydrophobic rods immersed in water.
-The radius and separation are shown.}
-\label{fig:rods}
-\end{figure}
 
@@
 \begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-diagram}
+\end{center}
+\caption{ Cross section schematic for two hydrophobic rods immersed in water.
+The radius and separation are shown.}
+\label{fig:rods}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=\columnwidth]{figs/density-rods-in-water}
@@
 
-For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
-and the entropic energy contribution, where all are in units of energy per length 
-(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
-goes from zero to the critical separation and then is approximately constant at
-larger separations. The entropy has a minimum just before the rods
-transition to allow liquid between them. As the rods separate starting from zero, the area 
-between them containing vapor increases and there is less overall space for the 
-surrounding liquid water to occupy and thus the entropy decreases. When that area fills suddenly
-with liquid, the entropy increases due to the increased space for the water molecules. The
-internal energy also has a minimum just before the transition to the liquid filled state.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-rods}
-\end{center}
-\caption{ Free energy, internal energy, and the entropic energy contribution versus
-separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
-offset to zero at large separations. }
-\label{fig:energy-rods}
-\end{figure}
-
 The free energy per length is compared for rods of several different radii
-in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for radius 0.5~nm
-is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
-For ease of comparison, we arbitrarily offset each free energy such that they all go 
+in Figure~\ref{fig:rods-energy-vs-distance}. For ease of comparison, we 
+arbitrarily offset each free energy such that they all go 
 to zero at large separations. For large $d$, the rods 
@@
 
+\subsection{Hydrophobic interactions of four rods}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Test-particle approach}
-
-More interesting applications are possible when we move into three
-dimensions using spherical coordinates.  The first (but not simplest)
-thing to try is looking at the response of liquid water to a single
-fixed water molecule.  This is referred to as the \emph{test particle}
-method\cite{FIXME}, and is as simple as constraining the density to
-include a delta function at the origin---meaning that a single
-molecule is fixed in location.  The free energy is then minimized
-under this constraint, giving a density profile and an associated free
-energy.  The density profile is just the self-correlation function
-multiplied by the bulk liquid density.  \fixme{We should be
-  distinguishing between the properties of the \emph{exact functional}
-  and those of an approximate functional...}  Thus, a test-particle
-calculation provides a stringent test of the correctness of any
-approximate functional.
-
-In order to perform the test-particle test, we introduce a sharp
-gaussian density spike at the origin and relax the density under a
-constraint that this spike remain.  Figure~\ref{fig:test-particle}
-shows the resulting density profile, compared with the experimental
-$g_{\textrm{OO}}$.  Note that this data was \emph{not} used in the
-fitting of the empirical parameters of the functional.  The agreement
-between experiment and theory is both qualitatively and quantitatively
-poor.  The integral of the difference between this density profile
-(excluding the delta function) and the bulk liquid density should be
-0.06~molecules \fixme{look up and show derivation... or cut this},
-but we find instead to be $-0.15$~molecules.  In addition, the associated
-free energy should be positive: there is an entropic cost in fixing
-one molecule in position, but for our functional the free energy of
-this density configuration is in fact negative, with a value of
--3.5~eV, meaning that the homogeneous fluid is unstable.  This is a
-serious flaw that must be addressed in future work.
+% 
+% \subsection{Test-particle approach}
+% 
+% More interesting applications are possible when we move into three
+% dimensions using spherical coordinates.  The first (but not simplest)
+% thing to try is looking at the response of liquid water to a single
+% fixed water molecule.  This is referred to as the \emph{test particle}
+% method, and is as simple as constraining the density to
+% include a delta function at the origin---meaning that a single
+% molecule is fixed in location.  The free energy is then minimized
+% under this constraint, giving a density profile and an associated free
+% energy.  The density profile is just the self-correlation function
+% multiplied by the bulk liquid density.  \fixme{We should be
+%   distinguishing between the properties of the \emph{exact functional}
+%   and those of an approximate functional...}  Thus, a test-particle
+% calculation provides a stringent test of the correctness of any
+% approximate functional.
+% 
+% In order to perform the test-particle test, we introduce a sharp
+% gaussian density spike at the origin and relax the density under a
+% constraint that this spike remain.  Figure~\ref{fig:test-particle}
+% shows the resulting density profile, compared with the experimental
+% $g_{\textrm{OO}}$.  Note that this data was \emph{not} used in the
+% fitting of the empirical parameters of the functional.  The agreement
+% between experiment and theory is both qualitatively and quantitatively
+% poor.  The integral of the difference between this density profile
+% (excluding the delta function) and the bulk liquid density should be
+% 0.06~molecules \fixme{look up and show derivation... or cut this},
+% but we find instead to be $-0.15$~molecules.  In addition, the associated
+% free energy should be positive: there is an entropic cost in fixing
+% one molecule in position, but for our functional the free energy of
+% this density configuration is in fact negative, with a value of
+% -3.5~eV, meaning that the homogeneous fluid is unstable.  This is a
+% serious flaw that must be addressed in future work.
+%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\section{Hard sphere solute}
-
 Figure~\ref{fig:sphere-energy-vs-diameter} shows the hydration free

### 2011-12-07 paper/paper.tex
various changes on sphere in water Tested everything.

@@
 \begin{equation}
-  \varphi(r) = \color{red}\Theta(r-2 \lambdadisp R)
+  \varphi(r) = \Theta(r-2 \lambdadisp R)
 \end{equation}
@@
 \begin{align}
-  \etadisp(\xx) &=
-  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
-             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
+%    ((4/3.0*M_PI)*Pow(3)(R)*GaussianConvolve(2*lambdainput*lscale*radius,
+%                                             2*lambdaE*length_scalingE*Rexpr)
+%  \etadisp(\xx) &= \color{red}
+%  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
+%             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
+  \\
+  \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
+  \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
+      \lscale R)^2}}\right)d\xx'
 \end{align}
@@
 \caption{Density profiles for multiple sizes of hard sphere solute.
-  Results from a simulation of SPC/E water~\cite{huang2001shs} are
-  shown as a dotted lines.}
+  }
 \label{fig:cavities}
@@
 \caption{ Free energy versus radius for a single hydrophobic sphere
-immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. }
+  immersed in water. This should have an asymptote equal to the
+  surface tension at room temperature, and it agrees well with the
+  surface tension in Figure~\ref{fig:surface-tension}. Results from a
+  simulation of SPC/E water~\cite{huang2001shs} are shown as
+  circles. }
 \label{fig:sphere-energy-vs-diameter}
@@
 \end{center}
-\caption{ Density profiles for spheres of different radii. The dotted line
-represents the saturated liquid density.  }
+\caption{ Density profiles around hard-sphere solutes of different radii. Predictions
+  from our classical density-functional theory are in solid red, while
+  the dotted line shows the result of a molecular dynamics simulation
+  of SPC/E water~\cite{huang2001shs}.  }
 \label{fig:density-sphere}

### 2011-12-13 paper/paper.tex
try new functional in contact Tested everything.

@@
     g^{HS}_\sigma(\eta_\textit{eff}(\xx)) \\
+  g_\sigma^{HS}(\eta) &= \frac{1 - \frac12 \eta}{(1 - \eta)^3} \label{eq:ghs}
+  \\
   g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
-  +\frac32\frac{\zeta(\xx)}{(1-\eta)^2}
-  + \frac12\frac{\eta^2\zeta(\xx)}{(1-\eta)^3}
+  +\frac32\frac{1}{(1-\eta)^2}
+  + \frac12\frac{\eta^2}{(1-\eta)^3}
   \\
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}\\
   \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
@@
 where $g_\sigma^{HS}$
-is the hard-sphere fluid correlation function evaluated at contact.
-The $\zeta$ factor
-(also seen in the association energy, below) is a dimensionless
-measure of the local density inhomogeneity. The $C_{ij}$ values
+is the Carnahan-Starling value for the hard-sphere fluid correlation
+function evaluated at contact.  \textcolor{red}{Jess, could you verify
+that the two formulas for gHS above are equivalent? Thanks!}
+The $C_{ij}$ values
  are numerical constants taken from
@@
 fit to the integral in Equation~\ref{eq:A1-simple} over a range of
-values for $\lambdadisp$ and filling fractions.
+values for filling fraction and $\lambdadisp$.
 
@@
 
+\textcolor{red}{
 The association functional we use is a modified version of that of Yu
@@
   al}\cite{clark2006developing}.
-
+}
+
+The association functional we use is constructed by using the density
+$n_0(\xx)$, which is the density of hard spheres touching a given
+point, in the standard SAFT-VR association
+energy\cite{gil-villegas-1997-SAFT-VR}.
 The association free energy for our four-site model has the form
 \begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) % \zeta(\xx)
   \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
@@
 molecule, and the functional $X$ is the fraction of assotiation sites
-\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
-measure of the density inhomogeneity.  The fraction $X$ is determined
+\emph{not} hydrogen-bonded. %, and $\zeta(\xx)$ is a dimensionless
+%measure of the density inhomogeneity.
+The fraction $X$ is determined
 by the quadratic equation
 \begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)\Delta(\xx)} - 1}
-  {4 n_0(\xx)\zeta(\xx)\Delta(\xx)}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)%\zeta(\xx)
+      \Delta(\xx)} - 1}
+  {4 n_0(\xx)%\zeta(\xx)
+    \Delta(\xx)}
 \end{align}
-where the factor $\zeta$ from
-Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
-density, and reduces to unity in the homogeneous limit:
-\begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
-\end{align}
+%where the factor $\zeta$ from
+%Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
+%  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
+%density, and reduces to unity in the homogeneous limit:
+%\begin{align}
+%  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
+%\end{align}
 The functional $\Delta$ which appears in the formula for $X$ is a
@@
   \left(e^{-\beta\epsilonassoc} - 1\right) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\etadisp(\xx)) +
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma\left(\frac{4\pi R^3}{3}n_0(\xx)\right) +
   \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
@@
 written as a perturbative correction to the hard-sphere fluid correlation
-function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
-\begin{align}
-  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
-  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
-  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
-\end{align}
+function $g^\textit{HS}_\sigma$ given in Equation~\ref{eq:ghs},
+evaluated using the filling fraction computed from $n_0(\xx)$.
+%, for which we use the functional of Yu and
+%Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
+%\begin{align}
+%  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
+%  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
+%  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
+%\end{align}
 

### 2012-01-11 paper/paper.tex
Figures changed and deleted in paper. Tests skipped!

--- 2011-12-02 paper/paper.tex
@@
 theorem, which states that the contact density at a hard surface $n_c$
-is proportional to the pressure on that surface:
+is proportional to the pressure on that surface. This
+also leads to the property that the free energy of solvation of
+a hard solute is proportional to its volume in the limit of small
+solutes: 
 \begin{equation}\label{contactvaluethm}
-  p = n_c k_BT \:,
+  F = n_c k_BT V\:.
 \end{equation}
-which also leads to the property that the free energy of solvation of
-a hard solute is proportional to its volume in the limit of small
-solutes.  These properties are retained by the total functional,
+These properties are retained by the total functional,
 provided the remaining terms are \emph{purely nonlocal}.  A common
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
+  tension will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
+\label{fig:pressure-with-isotherms}
+\end{figure}
+
 The majority of the empirical parameters used in our functional are
@@
 \end{equation}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
-\end{figure}
 
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
-  tension will match the experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
-
 %\begin{figure}
@@
 
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
+%\end{center}
+%\caption{The theoretical versus experimental coexistence curve. Away from
+%the critical temperature, the agreement is good.\needsworknow{(keep figure?
+%its the same as Clark paper)} }
+%\label{fig:temperature-vs-density}
+%x\end{figure}
+
+\subsection{Liquid-vapor interface}
+
 \begin{figure}
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental coexistence curve. Away from
-the critical temperature, the agreement is good.\needsworknow{(keep figure?
-its the same as Clark paper)} }
-\label{fig:temperature-vs-density}
-\end{figure}
-
-\subsection{Liquid-vapor interface}
+\includegraphics[width=\columnwidth]{figs/surface-298}
+\end{center}
+\caption{The liquid-vapor density profile at room temperature shows
+some small oscillations close to the interface.  }
+\label{fig:liquid-vapor-profile}
+\end{figure}
 
@@
 some small oscillations.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{The liquid-vapor density profile at room temperature shows
-some small oscillations close to the interface.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
 
@@
 which is a reasonable resolution based on several lattice spacing calculations.
-The results are within 0.5\% of the free energy and a few percent of the contact
+The results for this spacing of 0.01~nm are within 0.5\% of the free energy 
+and a few percent of the contact
 density for the highest resolution tested.
@@
 rod has comparatively larger oscillations and a higher contact density.
+\needsworknow{Expand or add to this description of density profiles and
+oscillations for single rod?}
 
@@
 \end{center}
-\caption{ Free energy versus radius for a single hydrophobic rod
+\caption{ Free energy of hydration versus radius for a single hydrophobic rod
 immersed in water. This should have an asymptote equal to the surface
 tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}.\needsworknow{Add more data at small radius 
-and include an inset.} }
+Figure~\ref{fig:surface-tension}. The inset for rods with a very small 
+radius shows the linear relationship expected based on 
+Equation~\ref{contactvaluethm}.}
 \label{fig:energy-vs-diameter}
@@
 asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m. 
+For very small radius (less than about 0.1~nm) the
+relationship between the free energy per area and radius is linear (see inset
+of Figure~\ref{fig:energy-vs-diameter}). Based on the contact value theorem
+stated in Equation~\ref{contactvaluethm}, we expect this linear relationship.
+For slightly larger yet still small rods (with a radius about 0.1~nm to 
+0.3~nm), the relationship between free energy per area and radius shows
+quadratic behavior.
 
@@
 We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
+use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ 
+(see Figure~\ref{fig:rods}), ranging from $d=0$ 
 (touching rods) to $d=1.3$~nm. At small
@@
 pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
+reached and liquid water fills the region between the rods. 
+Figure~\ref{fig:density-rods} 
 shows density profiles before and after this transition
@@
 at a separation of 0.2~nm and the
-bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
+bottom is 0.6~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
 the energy for these and other separations.}
@@
 \end{center}
-\caption{ Free energy (also known as the potential of mean force) 
+\caption{ Free energy of hydration (also known as the potential of mean force) 
 versus separation for two hydrophobic rods ranging in radius from
@@
 
-If we use a classical approximation where the energy of the rods is equal to 
-the water's surface area multiplied by the bulk surface tension, then the energy per 
-length is equal to the circumference multiplied by the surface tenson. The force
-per length is the derivative of this with respect to the separation. We approximate 
-the circumference of the stadium-shape by
+To explain this, we consider a macroscopic surface
+tension model where the free energy in the macroscopic limit is given by 
+\begin{equation}
+F = \gamma A + pV
+\end{equation}
+The first term describes the surface energy and the second term
+is the work needed to create a cavity of volume $V$. Since this pressure term
+scales with volume, it can be neglected relative to the surface term provided
+the length scale is small compared with $\frac{\gamma} p$. At room temperature,
+the value of $\frac{\gamma} p$ is equal to about $20~\mu m$ which is
+much larger than any of the solutes we study. We expect that for larger 
+rods ... \needsworknow{Explain ``bowing'' out of vapor area for larger rods.}
+
+Starting from the surface energy term, we can calculate the 
+free energy per 
+length, which is equal to the circumference multiplied by the surface tenson. 
+The force
+per length is the derivative of this with respect to the separation. 
+We approximate the circumference of the stadium-shape by
 \begin{equation}
@@
 \end{equation}
-and so the force per length is equal to $2\gamma$. This 
-is consistent with the calculation of the force per length from 
+and so the force per length is equal to $2\gamma$.
+
+This is consistent with the calculation of the force per length from 
 Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
@@
 To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will again use this classical approximation and propose that 
+radius, we will again use this model and propose that 
 the critical separation, $d_c$, will occur when the free energies are equal. This 
@@
 \end{tabular}
-\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
+\caption{Calculations of the critical separation 
+(see Equation~\ref{criticalseparation})
 for the transition
-from liquid to vapor between two hydrophobic rods.}
+from liquid to vapor between two hydrophobic rods.\needsworknow{Delete
+this table and move some of the information into the text along with
+prediction model and explanation of Figure~\ref{fig:rods-energy-vs-distance}}.}
 \label{criticalseparationtable}

### 2012-01-11 paper/paper.tex
clean up text in sphere section of paper. Tested everything.

@@
 
-A more widely studied system that is simpler in many ways than the
-test-particle problem is that of a \emph{hard sphere solute} dissolved
-in water.  This is a common model for hydrophobic interactions.  The
-effect of this solute is merely to create a cavity within which the
-water may not penetrate.  This is the simples possible solute, and
-will serve as a test case, which if passed could lead to the ability
-to insert different chemicals into aqueous solution, instead of a
-vapor cavity, in order to observe biologically and chemically
-important aqueous solvation interactions.
-
-The most obvious output of such a calculation is the free energy of
-solvation.  In Figure~\ref{fig:surfaceTension} we plot the ratio of the
-energy of the cavity system to the surface area of the cavity.  This
-is an \emph{effective surface tension}, which approaches the bulk
-surface tension as the cavity radius increases.  On the same plot, we
-include the result from an SPC/E molecular dynamics
-simulation\cite{huang2001shs}, showing very good agreement.  At very
-small solute radius the effective surface tension is linear, meaning
-that the free energy is proportional to the volume---this limiting
-case is an exact result, which is exactly reproduced by any functional
-that reduces to the ideal gas functional at small distances.
-
-\begin{figure}
-\begin{center}
-%\includegraphics[width=\columnwidth]{radial-plots/surface-tension}
-\end{center}
-\caption{Free energy per surface area as a function of cavity
-  radius. SPC/E results from~\cite{huang2001shs}}
-\label{fig:surfaceTension}
-\end{figure}
-
-This exact result for small radius solutes relates to a general property of
-fluids interacting with hard surfaces, which is that the pressure on a hard
-surface is determined by the contact density of the fluid at the surface.
-Thus, a functional that correctly predicts the free energy as a function of
-radius will also give the proper contact density.
-Figure~\ref{fig:cavities} shows the density profile for several hard sphere
-radii, plotted together with with the results of the same SPC/E molecular
-dynamics simulation shown in
-Figure~\ref{fig:surfaceTension}\cite{huang2001shs}.  As expected---since
-the free energies agree so well---the contact densities show good
-agreement.  For small spheres the entire curve shows reasonable agreement,
-with the agreement becoming worse for larger radii.
-
-\begin{figure}[b]
-\begin{center}
-%\includegraphics[width=\columnwidth]{radial-plots/cavity-density}
-\end{center}
-\caption{Density profiles for multiple sizes of hard sphere solute.
-  }
-\label{fig:cavities}
-\end{figure}
-
-In order to understand the worsening agreement in the density profile
-for larger solutes in Figure~\ref{fig:cavities}, it is instructive to
-consider the limit as the radius becomes very large.  In this limit,
-the surface becomes essentially flat, and the liquid should develop a
-\emph{vapor layer}, so that as the effective surface tension
-approaches the bulk surface tension, the density profile should
-approach the profile of the liquid-vapor interface as plotted in
-Figure~\ref{fig:cartesianDensity}.  We demonstrate this effect in
-Figure~\ref{fig:LargeR_RadialDensity}, which shows superimposed the
-density profiles for several large solutes with that of the
-liquid-vapor interface.  It is clear that as the radius is increased,
-the profile is approaching that of the liquid-vapor interface.  Thus
-we can understand the deviations from the molecular dynamics
-simulations in Figure~\ref{fig:cavities} at larger radii as a direct
-result of the oscillations observed in the planar liquid-vapor
-interface.
-
-\fixme{Not sure if we want to keep this paragraph or work it into the
-  discussion somewhere?} For \emph{very} small cavities \fixme{not
-  shown... should we add one?}, the contact density is equal to the
-bulk liquid density, as the solute is too small to really perturb the
-water (see the blue curve in Fig.~\ref{fig:cavities}).  As the cavity
-size increases, the contact density (and pressure) first rise, and
-then fall again, as seen in Figure~\ref{fig:cavities}.  This is due to
-short range attractive interactions that are able to pass through the
-cavities of small radii.  For larger radii, these interactions are
-less prominent and there is less attraction to molecules across the
-cavity.  Thus, the contact density is decreased.
-
-Figure~\ref{fig:sphere-energy-vs-diameter} shows the hydration free
-energy per unit surface area of a hard sphere.
-
 \begin{figure}
@@
   surface tension in Figure~\ref{fig:surface-tension}. Results from a
-  simulation of SPC/E water~\cite{huang2001shs} are shown as
-  circles. }
+  simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
+  The horizonal lines show the experimental and SPC/E bulk surface
+  tension for water at standard atmospheric temperature and
+  pressure. }
 \label{fig:sphere-energy-vs-diameter}
 \end{figure}
+
+A common model of hydrophobic solutes is the \emph{hard sphere solute}
+\fixme{add citations?}.  The effect of this solute is merely to create
+a cavity within which the water may not penetrate.  This is the
+simples possible solute, and will serve as a test case, which if
+passed could lead to the ability to insert different chemicals into
+aqueous solution, instead of a vapor cavity, in order to observe
+biologically and chemically important aqueous solvation interactions.
+
+As in the single rod, we begin by examining the free energy of
+hydration, expressed in Figure~\ref{fig:sphere-energy-vs-diameter} as
+the ratio of the energy of the cavity system to the surface area of
+the cavity.  This \emph{effective surface tension}, approaches the
+bulk surface tension as the cavity radius increases.  As with the
+single rod, we see the analytically correct behavior in the limit of
+small solutes.  We also plot the surface tension calculated using a
+molecular dynamics simulation of SPC/E water\cite{huang2001shs}.  The
+agreement is quite excellent, apart from the issue that the SPC/E
+model for water significantly underestimates the surface tension of
+water at room temperature\cite{vega2007surface}.
+
+Figure~\ref{fig:density-sphere} shows the density profile for several
+hard sphere radii, plotted together with with the results of the same
+SPC/E molecular dynamics simulation shown in
+Figure~\ref{fig:sphere-energy-vs-diameter}\cite{huang2001shs}.  The
+agreement with simulation is quite reasonable.  The largest
+disagreement involves the density at contact, which according to the
+contact value theorem cannot agree, since the free energies do not
+agree.
 

### 2012-01-11 paper/figs/all-spheres.pl

@@ -6,8 +6,8 @@ my $dd;
 # We do the largest diameters first, so the small calculations won't
 # be scheduled on big-memory nodes leaving the big ones to wait for
 # them to finish.
-#foreach $dd (2.0, 1.6, 1.4, 1.2, 1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.15, 0.1, 0.05, 0) {
-foreach $dd (2.0) {
+foreach $dd (2.0, 1.6, 1.4, 1.2, 1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.15, 0.1, 0.05, 0) {
+#foreach $dd (2.0) {
   my $d = sprintf("%04.2f", $dd);
 
   print "Submitting sphere for $d nm\n";

### 2012-01-18 paper/paper.tex
added four rods energy vs distance plot. Tests skipped!

--- 2012-01-11 paper/paper.tex
@@
 theorem, which states that the contact density at a hard surface $n_c$
-is proportional to the pressure on that surface. This
-also leads to the property that the free energy of solvation of
+is proportional to the pressure on that surface:
+\begin{equation}\label{contactvaluethm}
+  p = n_c k_BT \:,
+\end{equation}
+which also leads to the property that the free energy of solvation of
 a hard solute is proportional to its volume in the limit of small
-solutes: 
-\begin{equation}\label{contactvaluethm}
-  F = n_c k_BT V\:.
-\end{equation}
-These properties are retained by the total functional,
+solutes.  These properties are retained by the total functional,
 provided the remaining terms are \emph{purely nonlocal}.  A common
@@
 \subsection{Determining the empirical parameters}\label{sec:empirical}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
-  tension will match the experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
-\end{figure}
 
@@
 \end{equation}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
+\label{fig:pressure-with-isotherms}
+\end{figure}
 
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-tension}
+\end{center}
+\caption{The theoretical versus experimental surface tension
+  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
+  tension will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
 %\begin{figure}
@@
 
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-%\end{center}
-%\caption{The theoretical versus experimental coexistence curve. Away from
-%the critical temperature, the agreement is good.\needsworknow{(keep figure?
-%its the same as Clark paper)} }
-%\label{fig:temperature-vs-density}
-%x\end{figure}
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
+\end{center}
+\caption{The theoretical versus experimental coexistence curve. Away from
+the critical temperature, the agreement is good.\needsworknow{(keep figure?
+its the same as Clark paper)} }
+\label{fig:temperature-vs-density}
+\end{figure}
 
 \subsection{Liquid-vapor interface}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{The liquid-vapor density profile at room temperature shows
-some small oscillations close to the interface.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
 
@@
 some small oscillations.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/surface-298}
+\end{center}
+\caption{The liquid-vapor density profile at room temperature shows
+some small oscillations close to the interface.  }
+\label{fig:liquid-vapor-profile}
+\end{figure}
 
@@
 which is a reasonable resolution based on several lattice spacing calculations.
-The results for this spacing of 0.01~nm are within 0.5\% of the free energy 
-and a few percent of the contact
+The results are within 0.5\% of the free energy and a few percent of the contact
 density for the highest resolution tested.
@@
 rod has comparatively larger oscillations and a higher contact density.
-\needsworknow{Expand or add to this description of density profiles and
-oscillations for single rod?}
 
@@
 \end{center}
-\caption{ Free energy of hydration versus radius for a single hydrophobic rod
+\caption{ Free energy versus radius for a single hydrophobic rod
 immersed in water. This should have an asymptote equal to the surface
 tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. The inset for rods with a very small 
-radius shows the linear relationship expected based on 
-Equation~\ref{contactvaluethm}.}
+Figure~\ref{fig:surface-tension}.\needsworknow{Add more data at small radius 
+and include an inset.} }
 \label{fig:energy-vs-diameter}
@@
 asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
-For very small radius (less than about 0.1~nm) the
-relationship between the free energy per area and radius is linear (see inset
-of Figure~\ref{fig:energy-vs-diameter}). Based on the contact value theorem
-stated in Equation~\ref{contactvaluethm}, we expect this linear relationship.
-For slightly larger yet still small rods (with a radius about 0.1~nm to 
-0.3~nm), the relationship between free energy per area and radius shows
-quadratic behavior.
 
@@
 We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ 
-(see Figure~\ref{fig:rods}), ranging from $d=0$ 
+use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
 (touching rods) to $d=1.3$~nm. At small
@@
 pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. 
-Figure~\ref{fig:density-rods} 
+reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
 shows density profiles before and after this transition
@@
 at a separation of 0.2~nm and the
-bottom is 0.6~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
+bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
 the energy for these and other separations.}
@@
 \end{center}
-\caption{ Free energy of hydration (also known as the potential of mean force) 
+\caption{ Free energy (also known as the potential of mean force) 
 versus separation for two hydrophobic rods ranging in radius from
@@
 
-To explain this, we consider a macroscopic surface
-tension model where the free energy in the macroscopic limit is given by 
-\begin{equation}
-F = \gamma A + pV
-\end{equation}
-The first term describes the surface energy and the second term
-is the work needed to create a cavity of volume $V$. Since this pressure term
-scales with volume, it can be neglected relative to the surface term provided
-the length scale is small compared with $\frac{\gamma} p$. At room temperature,
-the value of $\frac{\gamma} p$ is equal to about $20~\mu m$ which is
-much larger than any of the solutes we study. We expect that for larger 
-rods ... \needsworknow{Explain ``bowing'' out of vapor area for larger rods.}
-
-Starting from the surface energy term, we can calculate the 
-free energy per 
-length, which is equal to the circumference multiplied by the surface tenson. 
-The force
-per length is the derivative of this with respect to the separation. 
-We approximate the circumference of the stadium-shape by
+If we use a classical approximation where the energy of the rods is equal to 
+the water's surface area multiplied by the bulk surface tension, then the energy per 
+length is equal to the circumference multiplied by the surface tenson. The force
+per length is the derivative of this with respect to the separation. We approximate 
+the circumference of the stadium-shape by
 \begin{equation}
@@
 \end{equation}
-and so the force per length is equal to $2\gamma$.
-
-This is consistent with the calculation of the force per length from 
+and so the force per length is equal to $2\gamma$. This 
+is consistent with the calculation of the force per length from 
 Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
@@
 To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will again use this model and propose that 
+radius, we will again use this classical approximation and propose that 
 the critical separation, $d_c$, will occur when the free energies are equal. This 
@@
 \end{tabular}
-\caption{Calculations of the critical separation 
-(see Equation~\ref{criticalseparation})
+\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
 for the transition
-from liquid to vapor between two hydrophobic rods.\needsworknow{Delete
-this table and move some of the information into the text along with
-prediction model and explanation of Figure~\ref{fig:rods-energy-vs-distance}}.}
+from liquid to vapor between two hydrophobic rods.}
 \label{criticalseparationtable}
@@
 \subsection{Hydrophobic interactions of four rods}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/four-rods-energy-vs-distance}
+\end{center}
+\caption{ Free energy (also known as the potential of mean force) 
+versus separation for four hydrophobic rods ranging in radius from
+0.3~nm to 1.0~nm.
+All were arbitrarily offset to zero at large separations. The
+transition corresponds to the phase change from
+vapor to liquid between the rods as pictured in the density profiles in 
+Figure~\ref{fig:density-four-rods}. }
+\label{fig:four-rods-energy-vs-distance}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-four-rods}
+\end{center}
+\caption{ Density profiles illustrating the transition from vapor 
+to liquid water between four rods. The radius is 0.5~nm, the top figure is 
+at a separation of 1.1~nm and the
+bottom is 1.2~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
+the energy for these and other separations.}
+
+\label{fig:density-four-rods}
+\end{figure}
+
+We extend the study to four rods in water using the same radii as the two rods and
+separation distance $d = 0$~nm to $d = 2$~nm for nearest neighbor rods. These exhibit
+a similar behavior to the two rods where for small $d$ only vapor is present in the 
+middle of the rods until a critical separation between where water occupies the middle.
+Before the transition, the four rods and vapor form a square with rounded courners 
+where each rod forms a corner (shown in Fig. \ref{fig:density-four-rods} as 'before').
+For large $d$, the rods behave like separate single rods.  Free energy
+is plotted against rod separation in Fig. \ref{fig:four-rods-energy-vs-distance}.
+From the slope we calculate the force per length to be [number].
+
+The rounded square's circumference is estimated by
+\begin{equation}
+C_{rs} = 2 \pi r + 8r + 4d.
+\end{equation}
+We again equate this circumference with the total circumference of the four rods, which
+results in a critical separation
+\begin{equation}
+d_{c4} = \frac{3 \pi - 4}{2} r \label{criticalfourrods}
+\end{equation}
+
+\begin{table}
+\begin{tabular} {|c|c|c|c|}
+\hline
+$r$ (nm) & $d_{rs}$ prediction (nm) & $d_{c4}$ result (nm) & \% error \\
+\hline
+0.3 & 0.81 & 0.40 & 51\% \\
+\hline
+0.5 & 1.36 & 1.15 & 15\% \\
+\hline
+0.7 & 1.90 & 1.70 & 11\% \\
+\hline
+0.9 & 2.44 & 2.30 & 06\% \\
+\hline
+1.0 & 2.70 & 0.00 & 00\% \\
+\hline
+\end{tabular}
+\caption{Calculations of the critical separation (see Equation~\ref{criticalfourrods})
+for the transition
+from liquid to vapor between four hydrophobic rods.}
+\label{criticalfourrodstable}
+\end{table}
+
+The predicted and computed valued for $d_{c4}$ are show in Table \ref{criticalfourrodstable}. 
+[talk about table here]
 

### 2012-01-27 papers/contact/paper.tex

@@
 
-\subsection{Hard spheres confined in a cubical cavity}
-
-\textcolor{red}{
-  I am thinking of also performing simulations for a cubical cavity, and
-  examining the result near the center of a face, which (for a large
-  cavity) will give us a prediction for the hard-sphere fluid near a
-  single hard wall.  We've got some simulations in
-  Figure~\ref{fig:box-density}, but these aren't very nice plots.
-  We'll want to do a much larger box, and plot the density vs $x$ near
-  the wall.  But I'm waiting on simulation data for this...
-}
-
-\subsection{Hard spheres around a hard sphere (or rod)}
-
-\textcolor{red}{
-  It would also be interesting to study a system with opposite
-  curvature.  Put a (larger) hard sphere in the hard-sphere fluid, and
-  see what the contact density is around that hard sphere.  This
-  touches on an important limiting case...
-}
-
-\begin{figure*}
-\includegraphics[height=5cm]{figs/box-100c--05,05,05-13}
-\includegraphics[height=5cm]{figs/box-100s--05,05,05-13}
-\includegraphics[height=5cm]{figs/box-110c--05,05,05-13}
-\caption{Density of hard spheres in a $5\times5\times5$ box.}
-\label{fig:box-density}
-\end{figure*}
+\subsection{Hard spheres at a hard wall}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-10}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:walls-10}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-30}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.3.}
+  \label{fig:walls-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-40}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.4.}
+  \label{fig:walls-40}
+\end{figure}
+
+We study the case of hard spheres near a single hard wall.  This is
+shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.  Actually, we
+should begin with this.
+
+\subsection{Hard spheres around a hard sphere}
+
+It is also interesting to study a system with opposite curvature.  Put
+a (larger) hard sphere in the hard-sphere fluid, and see what the
+contact density is around that hard sphere.  This is particularly
+relevant when we consider using classical density-functional theory to
+model solvation.
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-4-30}
+  \caption{Density and contact densities around a hard sphere with
+    radius 1.}
+  \label{fig:inner-4-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-4-40}
+  \caption{Density and contact densities around a hard sphere with
+    radius 1.}
+  \label{fig:inner-4-40}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-8-30}
+  \caption{Density and contact densities around a hard sphere with
+    radius 3.}
+  \label{fig:inner-8-30}
+\end{figure}
 

### 2012-01-27 paper/paper.tex

--- 2012-01-11 paper/paper.tex
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/four-rods-energy-vs-distance}
+\end{center}
+\caption{ Free energy (also known as the potential of mean force) 
+versus separation for four hydrophobic rods ranging in radius from
+0.3~nm to 1.0~nm.
+All were arbitrarily offset to zero at large separations. The
+transition corresponds to the phase change from
+vapor to liquid between the rods as pictured in the density profiles in 
+Figure~\ref{fig:density-four-rods}. }
+\label{fig:four-rods-energy-vs-distance}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-four-rods}
+\end{center}
+\caption{ Density profiles illustrating the transition from vapor 
+to liquid water between four rods. The radius is 0.5~nm, the top figure is 
+at a separation of 1.1~nm and the
+bottom is 1.2~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
+the energy for these and other separations.}
+
+\label{fig:density-four-rods}
+\end{figure}
+
+We extend the study to four rods in water using the same radii as the two rods and
+separation distance $d = 0$~nm to $d = 2$~nm for nearest neighbor rods. These exhibit
+a similar behavior to the two rods where for small $d$ only vapor is present in the 
+middle of the rods until a critical separation between where water occupies the middle.
+Before the transition, the four rods and vapor form a square with rounded courners 
+where each rod forms a corner (shown in Fig. \ref{fig:density-four-rods} as 'before').
+For large $d$, the rods behave like separate single rods.  Free energy
+is plotted against rod separation in Fig. \ref{fig:four-rods-energy-vs-distance}.
+From the slope we calculate the force per length to be [number].
+
+The rounded square's circumference is estimated by
+\begin{equation}
+C_{rs} = 2 \pi r + 8r + 4d.
+\end{equation}
+We again equate this circumference with the total circumference of the four rods, which
+results in a critical separation
+\begin{equation}
+d_{c4} = \frac{3 \pi - 4}{2} r \label{criticalfourrods}
+\end{equation}
+
+\begin{table}
+\begin{tabular} {|c|c|c|c|}
+\hline
+$r$ (nm) & $d_{rs}$ prediction (nm) & $d_{c4}$ result (nm) & \% error \\
+\hline
+0.3 & 0.81 & 0.40 & 51\% \\
+\hline
+0.5 & 1.36 & 1.15 & 15\% \\
+\hline
+0.7 & 1.90 & 1.70 & 11\% \\
+\hline
+0.9 & 2.44 & 2.30 & 06\% \\
+\hline
+1.0 & 2.70 & 0.00 & 00\% \\
+\hline
+\end{tabular}
+\caption{Calculations of the critical separation (see Equation~\ref{criticalfourrods})
+for the transition
+from liquid to vapor between four hydrophobic rods.}
+\label{criticalfourrodstable}
+\end{table}
+
+The predicted and computed valued for $d_{c4}$ are show in Table \ref{criticalfourrodstable}. 
+[talk about table here]
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-02-01 papers/contact/paper.tex
a bit of work on contact paper Tested everything.

@@
 \section{Introduction}
+
+% The following are papers that use a SAFT-based classical DFT with at
+% least some of the terms purely local
+\newcommand\saftlocaldft{felipe2001examination, gloor2002saft,%
+  gloor2004accurate, clark2006developing, gloor2007prediction,%
+  kahl2008modified, gross2009density}
+% The following are papers that use a SAFT-based classical DFT with
+% all the terms that should be non-local being non-local.
+\newcommand\saftnonlocaldft{yu2002fmt-dft-inhomogeneous-associating,%
+  fu2005vapor-liquid-dft,bryk2006density}
+
+There has been considerable recent interest in using Statistical
+Associating Fluid Theory (SAFT) to construct classical density
+functionals to describe associating
+fluids\cite{\saftlocaldft,\saftnonlocaldft}.  This approach has been
+successful in qualitatively describing the dependence of surface
+tension on temperature.  Unfortunately, most of these constructed
+functionals\cite{\saftlocaldft} cannot be used to study arbitrary
+inhomogeneous density distributions, because they use local
+density approximations that are only valid for densities that vary
+slowly over molecular length scales.
+
+Constructing a classical density functional based on SAFT involves
+rewriting each term in the free energy as a functional of an
+inhomogeneous density.  The SAFT free energy consists of four terms:
+\begin{align}
+  A_\textit{SAFT} &= A_\textit{ideal} + A_\textit{HS} + A_\textit{chain} + A_\textit{disp} + A_\textit{assoc}
+\end{align}
+where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
+is the excess free energy of a hard-sphere fluid of monomers,
+$A_\textit{chain}$ is the contribution to the free energy from the
+chaining effect in a polymeric fluid, $A_\textit{disp}$ is the
+dispersion contribution to the free energy, and $A_\textit{assoc}$ is
+the free energy contribution from \emph{association}, which is to say,
+hydrogen bonding.  Each of these terms has some dependence on the
+inhomogeneity in the density.  The ideal gas is purely local, and in
+fact must be the only purely local term in the free energy in order
+for the functional to satisfy the contact-value theorem.  The
+hard-sphere contribution contribution to the free energy has been
+thoroughly studied, and is well approximated by the White Bear version
+of the Fundamental-Measure Theory (FMT)
+functional\cite{roth2002whitebear}.  A functional for the chain
+contribution is has recently been introduced\cite{bryk2006density,
+  gross2009density}, but is not needed for small associating molecules
+such as water.  The dispersion term is long-range and thus has
+significant dependence on the density distribution, but because its
+relatively weak position dependence, is amenable to mean-field
+approximations.  A functional for the association term of the free
+energy was introduced by Wu and
+Yu\cite{yu2002fmt-dft-inhomogeneous-associating}, and was used in
+several subsequent functionals\cite{fu2005vapor-liquid-dft,
+  bryk2006density}.  A key input (FIXME: term, factor?) to this
+functional is the contact value of the correlation function of a
+hard-sphere fluid.
+
+The contact value of the correlation function---also known as the
+cavity correlation function---is the the value of the
+correlation function evaluated a distance equal to the hard-sphere
+diameter.  Yu and Wu introduce an approximation based on the
+fundamental measures of Fundamental Measure Theory (see
+Section~\ref{sec:yuwu}).  In this paper, we will report on the
+accuracy of this approximation by comparing with Monte Carlo
+simulations of the inhomogeneous hard-sphere fluid.  In addition, we
+will introduce and test differing approximations for the cavity
+correlation function.
+
+\textcolor{red}{Cavity correlation function!}
 
@@
 
+\section{Gross' functional}\label{sec:gross}
+We compare with the functional of Gross \emph{et
+  al}\cite{gross2009density}, which is of the first kind, where the
+contact density is evaluated at the sphere location.
+\begin{align}
+  n_\textit{contact} &= n_{0'} \frac{1 - \frac12 n_{3'}}{\left(1 - n_{3'}\right)^3}
+\end{align}
+where $n_{0'}$ and $n_{3'}$ are defined in a manner analogous to the
+$n_0$ and $n_3$ of FMT, but with the hard-sphere radius replaced by
+the hard-sphere diameter, so the integrals range over spheres that are
+twice as large.
+
+\section{Yu and Wu's functional}\label{sec:yuwu}
+
 Yu and Wu developed a functional that should correspond to the contact
@@
 the contact density of Fu and Wu~\cite{fu2005vapor-liquid-dft} (or
-should it be Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}?), the ``simple
-contact density'' from Section~\ref{simple-contact}, and the ``contact
-density at sphere'' from Section~\ref{contact-at-sphere}.  In
-principle, the functional of Wu (FIXME) should match the
-CenConDensity, while the DFT at sphere should match the ConDensity...
+should it be Yu and Wu\cite{yu2002structures,
+  yu2002fmt-dft-inhomogeneous-associating}?), the ``simple contact
+density'' from Section~\ref{simple-contact}, and the ``contact density
+at sphere'' from Section~\ref{contact-at-sphere}.  In principle, the
+functional of Wu (FIXME) should match the CenConDensity, while the DFT
+at sphere should match the ConDensity...
 

### 2012-02-04 papers/contact/paper.tex

@@
 several subsequent functionals\cite{fu2005vapor-liquid-dft,
-  bryk2006density}.  A key input (FIXME: term, factor?) to this
-functional is the contact value of the correlation function of a
-hard-sphere fluid.
-
-The contact value of the correlation function---also known as the
-cavity correlation function---is the the value of the
-correlation function evaluated a distance equal to the hard-sphere
-diameter.  Yu and Wu introduce an approximation based on the
-fundamental measures of Fundamental Measure Theory (see
-Section~\ref{sec:yuwu}).  In this paper, we will report on the
-accuracy of this approximation by comparing with Monte Carlo
-simulations of the inhomogeneous hard-sphere fluid.  In addition, we
-will introduce and test differing approximations for the cavity
-correlation function.
+  bryk2006density}.  A key input (FIXME: term, factor?) to both the
+association and chain free energies is the contact value of the
+correlation function of a hard-sphere fluid.
+
+The contact value of the correlation function---which is the same as
+the contact value of the cavity correlation function for hard
+spheres---is the the value of the correlation function evaluated a
+distance equal to the hard-sphere diameter.  In an inhomogeneous
+system treated using fundamental measures, there is some ambiguity in
+the meaning of the \emph{local} contact value of the correlation
+function which appears in the equations for the homogeneous system.
+Clearly the contact value of the correlation function will depend on
+position, but how should that position be defined? It will depend on
+how the correlation function is used.  There are two reasonable
+choices for functionals based on fundamental measure theory, which
+will depend on which densities are used...
+\begin{align}
+  F[n(\rr)] = \int n(\rr) f(\bar{n}(\rr), g_\sigma(\rr)) d\rr
+\end{align}
+\begin{align}
+  F[n(\rr)] = \int f(n_2(\rr), g_\sigma(\rr)) d\rr
+\end{align}
+
+Yu and Wu introduce an approximation based on the fundamental measures
+of Fundamental Measure Theory (see Section~\ref{sec:yuwu}).  Gross
+also does\cite{gross2009density}.
+
+In this paper, we will report on the accuracy of this approximation by
+comparing with Monte Carlo simulations of the inhomogeneous
+hard-sphere fluid.  In addition, we will introduce and test differing
+approximations for the cavity correlation function.
 

### 2012-02-06 papers/contact/figs/contact-density.py
add nicer plots for contact monte-carlo data Tested everything on collins.

@@ -21,8 +21,9 @@ pylab.plot(dftdata[:,0],dftdata[:,1]*4*numpy.pi/3,"b--",label='DFT density')
 pylab.plot(wbtdata[:,0],wbtdata[:,1]*4*numpy.pi/3,"m-.",label='WBT density')
 pylab.plot(wbm2data[:,0],wbm2data[:,1]*4*numpy.pi/3,"c--",label='WB mark II density')
 
-pylab.plot(mcdata[:,0],mcdata[:,2]*4*numpy.pi/3,"r-",label="ConDensity")
-pylab.plot(mcdata[:,0],mcdata[:,3]*4*numpy.pi/3,"g-",label="CenConDensity")
+off = 2
+pylab.plot(mcdata[:,0],mcdata[:,2+2*off]*4*numpy.pi/3,"r-",label="ConDensity")
+pylab.plot(mcdata[:,0],mcdata[:,3+2*off]*4*numpy.pi/3,"g-",label="CenConDensity")
 
 me = 3
 pylab.plot(dftdata[:,0],dftdata[:,3]*4*numpy.pi/3,"g+--",label="simple contact")

### 2012-02-06 papers/contact/paper.tex
add nicer plots for contact monte-carlo data Tested everything on collins.
minor edit on contact paper Tested everything on bingley.

@@
 
-%\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
-\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
+\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
+%\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
 
@@
 \maketitle
+
+%\input{formulas}
 
@@
 
-A final definition of the inhomogeneous contact density would be the
-contact density for spheres \emph{touching at a given point}.  This
-would be a symmetric measure, since both spheres involved are treated
-in the same way.  In fact, the density of such spheres is just the
-density $n_0(\xx)$ from FMT.  See Figure~\ref{fig:n0} for an
+An alternate definition of the inhomogeneous contact density would be
+the contact density for spheres \emph{touching at a given point}.
+This would be a symmetric measure, since both spheres involved are
+treated in the same way.  In fact, the density of such spheres is just
+the density $n_0(\xx)$ from FMT.  See Figure~\ref{fig:n0} for an
 illustration $n_0$.  Since $n_0$ is the fundamental measure of sphere
 surface contact, it would make sense to define a contact density that
-would correspond.
-
-To find the contact density at a given point of contact, we can't
-simply change the radius for spheres touching that point, since that
-would give us the pressure averaged over the entire surface of those
-spheres.  Instead, we would have to imagine extending a patch of those
-spheres which is located at the point of interest, as illustrated in
-Figure~\ref{fig:contact}.  \textcolor{red}{I'm not sure how to go
-  about doing this.  In the derivation section, I have some ideas
-  jotted down, but they don't really work.}
+would correspond.  This is what is done by Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as discussed in
+Section~\ref{sec:yuwu}.
 
@@
 compare both their functional as well as ours with simulation data.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/contact}
-\caption{Construction for finding the contact density at a given point
-  of contact.  We envision a small area $dA$ of each sphere that touches
-the point $\mathbf{r}$ extended by a distance $dR$.}
-\label{fig:contact}
-\end{figure}
 
@@
 
+%% \begin{widetext}
+%% \section{Formulas generated from Haskell}
+
+%% \begin{align}
+%%   \Phi_1 &= \phione \\
+%%   \Phi_2 &= \phitwo \\
+%%   \Phi_3 &= \phithree \\
+%%   \frac{d F}{dn_3} &= \dwbdnthree \\
+%%   \frac{d F}{dn_2} &= \dwbdntwo \\
+%%   \frac{d F}{dn_1} &= \dwbdnone \\
+%%   \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
+%%   \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev \\
+%% \end{align}
+%% \end{widetext}
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2012-02-07 papers/contact/paper.tex
various minor contact-related improvements Tested everything on bingley.

@@
     }
-    \right) n(\mathbf{r}') \delta(|\mathbf{r}-\mathbf{r}'| - R)
-    + \cdots
+    \right) n(\mathbf{r}') \delta(|\mathbf{r}-\mathbf{r}'| - R) \right.
+    \\
+    & \left.
+    + \frac{\delta A}{\delta n_2(\rr')} n(\rr) \delta'(|\rr-\rr'|-R) \cdots
     \right\}
   \end{align}
-  This comes out to quite a mess.  It wouldn't be infeasible to compute
-  this, but it wouldn't be efficient or pretty either.
+  This comes out to quite a mess.  It isn't be infeasible to compute
+  this, but it is challenging.
+}
+
+\derivation{
+  \end{widetext}
 }
@@
 compare both their functional as well as ours with simulation data.
-
-\derivation{
-
-  So we need to find the small change in free energy that is caused by
-  these small changes in the shape and size of the hard spheres that are
-  in contact with this point.  In the following section, we will work
-  out the changes in the weighted densities at position $\mathbf{r}'$
-  when point of contact we are considering is $\mathbf{r}$.
-
-  The packing fraction $n_3$ is changed at the point where the sphere is
-  extended, since a few more spheres will overlap this point.
-  \begin{align}
-    \Delta n_3(\mathbf{r}') &= n(\mathbf{r})\Delta A \Delta R
-    \delta(|\mathbf{r}-\mathbf{r}'| - R)
-  \end{align}
-  The density $n_0$ is the average density, averaged over a spherical
-  shell.  As such, shifting the position of a shell just shifts which
-  location we are averaging over.
-  \begin{align}
-    \Delta n_0(\mathbf{r}') &=
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-    \mathbf{\nabla}n(\mathbf{r}) \Delta A \Delta R
-    \delta(|\mathbf{r}-\mathbf{r}'|-R)
-  \end{align}
-  The change in $n_2$ comes from two terms: a change in the density value, as
-  we saw in $n_0$, and a change in the area of the segment.
-  \begin{align}
-    \Delta n_2(\mathbf{r}') &=
-    \left(n\left(\mathbf{r} + \Delta R
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
-      R}{R}\right)- n(\mathbf{r})\right) \Delta A
-    \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
-    &=
-    \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
-    \right) \Delta A \Delta R
-    \delta(|\mathbf{r}-\mathbf{r}'|-R)
-  \end{align}
-  The vector density $\mathbf{n}_{V2}$ is similar to the density
-  $n_2$: both changes in the density value and changes in the surface
-  area play a role.
-  \begin{align}
-    \Delta \mathbf{n}_{V2}(\mathbf{r}') &=
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
-    \left(n\left(\mathbf{r} + \Delta R
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\right)\left(1+2\frac{\Delta
-      R}{R}\right)- n(\mathbf{r})\right) \Delta A
-    \delta(|\mathbf{r}-\mathbf{r}'|-R) \\
-    &=
-    \frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}
-    \left(\frac{\mathbf{r}-\mathbf{r}'}{|\mathbf{r}-\mathbf{r}'|}\cdot
-    \mathbf{\nabla}n(\mathbf{r}) + 2 \frac{n(\mathbf{r})}{R}
-    \right) \Delta A \Delta R
-    \delta(|\mathbf{r}-\mathbf{r}'|-R) 
-  \end{align}
-  The density $n_1$ (and similarly $\mathbf{n}_{V1}$) can be roughly
-  interpreted as the density weighted by the area-averaged mean
-  curvature~\cite{oversteegen2005general, roth2010review}.  When we
-  extend the sphere at the contact point, we will introduce regions of
-  both positive and negative mean curvature.  In principle, we can do
-  this in such a way that the two portions cancel, and hence we will
-  assume that
-  \begin{align}
-    \Delta n_1(\mathbf{r}') &= 0 \\
-    \Delta \mathbf{n}_{V1}(\mathbf{r}') &= 0
-  \end{align}
-  In fact, the negative and positive portions are spatially separated,
-  so these terms should probably include a derivative term, which I
-  should work out.  Particularly as it may even happen to cancel a
-  different derivative term that we then won't have to drop!
-
-  We need to then find the change in Helmholtz free energy due to this
-  change at position $\mathbf{r}$:
-  \begin{align}
-    \Delta A_{HS} &= \int \mathbf{dr}'
-    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}\Delta n_3(\mathbf{r}')
-    +
-    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}\Delta n_2(\mathbf{r}')
-    +
-    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} \cdot \Delta \mathbf{n}_{V2}(\mathbf{r}')
-  \end{align}
-}
-
-\derivation{
-  \end{widetext}
-}
 

### 2012-02-07 papers/contact/figs/all-spheres.pl

@@ -17,7 +17,7 @@ foreach $dd ([16,265], [12,112], [8,13]) {
     my $resolution = 0.1; # spacing of grid points in hard-sphere radius.
 
     # Here I estimate the amount of memory that will be needed...
-    my $memuse = sprintf "%.0f", 2000*((($d + $padding)/20.0)*(0.1/$resolution))**3;
+    my $memuse = sprintf "%.0f", 2200*((($d + $padding)/20.0)*(0.1/$resolution))**3;
 
     my $scriptname = "papers/contact/figs/sphere-$d-$N-$F.tmp.sh";
     open SCRIPT, ">$scriptname" or die $!;

### 2012-02-07 papers/contact/figs/contact-density.py

@@ -29,8 +29,8 @@ me = 3
 pylab.plot(dftdata[:,0],dftdata[:,3]*4*numpy.pi/3,"g+--",label="simple contact")
 pylab.plot(dftdata[:,0],dftdata[:,4]*4*numpy.pi/3,"gx--",label="Yu and Wu")
 pylab.plot(dftdata[:,0],dftdata[:,5]*4*numpy.pi/3,"ro--",label="DFT at sphere")
-pylab.plot(wbm2data[:,0],wbm2data[:,5]*4*numpy.pi/3,"r+--",markevery=me,label="DFT at sphere (mark II)",
-           markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
+#pylab.plot(wbm2data[:,0],wbm2data[:,5]*4*numpy.pi/3,"r+--",markevery=me,label="DFT at sphere (mark II)",
+#           markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
 pylab.plot(dftdata[:,0],dftdata[:,7]*4*numpy.pi/3,"rx--",markevery=me,label="Gross",
            markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
 pylab.plot(dftdata[:,0],dftdata[:,6]*4*numpy.pi/3,"g*--",label="n2-only")

### 2012-02-07 papers/contact/figs/plot-inner.py

@@ -30,8 +30,8 @@ me = 3
 pylab.plot(dftdata[:,0]-rmin,dftdata[:,3]*4*numpy.pi/3,"g+--",markevery=me,label="simple contact")
 pylab.plot(dftdata[:,0]-rmin,dftdata[:,4]*4*numpy.pi/3,"gx--",markevery=me,label="Yu and Wu")
 pylab.plot(dftdata[:,0]-rmin,dftdata[:,5]*4*numpy.pi/3,"ro--",markevery=me,label="DFT at sphere")
-pylab.plot(wbm2data[:,0]-rmin,wbm2data[:,5]*4*numpy.pi/3,"r+--",markevery=me,label="DFT at sphere (mark II)",
-           markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
+#pylab.plot(wbm2data[:,0]-rmin,wbm2data[:,5]*4*numpy.pi/3,"r+--",markevery=me,label="DFT at sphere (mark II)",
+#           markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
 pylab.plot(dftdata[:,0]-rmin,dftdata[:,7]*4*numpy.pi/3,"rx--",markevery=me,label="Gross",
            markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
 pylab.plot(dftdata[:,0]-rmin,dftdata[:,6]*4*numpy.pi/3,"g*--",markevery=me,label="n2-only")

### 2012-02-07 src/ContactDensity.cpp

@@ -147,15 +147,15 @@ static Functional phi2_n3(const Functional &n3) {
 }
 
 static Functional phi3_n3(const Functional &n3) {
-  return (1 - (1.0/9)*(6 - 9*n3 + 6*sqr(n3) + 6*sqr(1-n3)*log(1-n3)/n3)/n3)/sqr(1 - n3);
+  return (1 - (1.0/9)*(6 - 9*n3 + 6*sqr(n3) + 6*sqr(1-n3)*log(1-n3)/n3)/sqr(n3))/sqr(1 - n3);
 }
 
 Functional dWBm2_dn1v_over_n2v(double radius) {
-  return Functional(1)*phi2_n3(n3(radius));
+  return Functional(-1)*phi2_n3(n3(radius));
 }
 
 Functional dWBm2_dn1(double radius) {
-  return n2(radius)*dWBm2_dn1v_over_n2v(radius);
+  return -n2(radius)*dWBm2_dn1v_over_n2v(radius);
 }
 
 Functional dWBm2_dn2(double radius) {
@@ -169,12 +169,11 @@ Functional dWBm2_dn2(double radius) {
 
 Functional dWBm2_dn2v_over_n2v(double radius) {
   Functional R(radius, "R");
-  return (Functional(1)/(4*M_PI*R))*phi2_n3(n3(radius)) -
+  return (Functional(-1)/(4*M_PI*R))*phi2_n3(n3(radius)) -
     6*n2(radius)*phi3_n3(n3(radius))/Functional(36*M_PI);
 }
 
 Functional dWBm2_dn3(double radius) {
-  // FIXME
   Functional R(radius, "R");
   Functional vtt = VectorThirdTerm(radius);
   Functional n3_ = n3(radius);
@@ -184,17 +183,25 @@ Functional dWBm2_dn3(double radius) {
   Functional n2z = zShellConvolve(radius);
   Functional nV22 = sqr(n2x) + sqr(n2y) + sqr(n2z);
   Functional n22mnV22 = sqr(n2(radius)) - nV22;
-  return n0(radius)/omn3 +
-
-    n22mnV22/(4*M_PI*R)*((Functional(1) + (1/9.0)*(6-3*n3_+6*omn3*log(omn3)/n3_))/sqr(omn3) +
-                         (1/9.0)*(Functional(-3) +
-                                  6*(-log(omn3)/n3_ +
[…]

### 2012-02-10 paper/paper.tex
remove another pair of unused figures Tested everything.
various changes to papers Tested everything on bingley.

@@
 
-\textcolor{red}{
 The association functional we use is a modified version of that of Yu
@@
   association term and is not useful as a reference.}  The association
-functional of Fu and Wu includes the effects of density
+functional of Yu and Wu includes the effects of density
 inhomogeneities on the \emph{contact density} of hard spheres, which
@@
 optimal SAFT parametrization for water found by Clark \emph{et
-  al}\cite{clark2006developing}.
-}
+  al}\cite{clark2006developing}.  The modification to use the SAFT-VR
+equation of state involves adding a correction term, see
+Equation~\ref{eq:gSW}.  \fixme{needs work}
 
@@
 \begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) % \zeta(\xx)
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) \zeta(\xx)
   \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
@@
 where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of assotiation sites
-\emph{not} hydrogen-bonded. %, and $\zeta(\xx)$ is a dimensionless
-%measure of the density inhomogeneity.
-The fraction $X$ is determined
-by the quadratic equation
+molecule, and the functional $X$ is the fraction of association sites
+\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
+measure of the density inhomogeneity.
+%
+The fraction $X$ is determined by the quadratic equation
 \begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)%\zeta(\xx)
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
       \Delta(\xx)} - 1}
-  {4 n_0(\xx)%\zeta(\xx)
+  {4 n_0(\xx)\zeta(\xx)
     \Delta(\xx)}
 \end{align}
-%where the factor $\zeta$ from
-%Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
-%  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
-%density, and reduces to unity in the homogeneous limit:
-%\begin{align}
-%  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
-%\end{align}
+where the factor $\zeta$ from Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating} describes the
+inhomogeneity of the fluid density, and reduces to unity in the
+homogeneous limit:
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
+\end{align}
 The functional $\Delta$ which appears in the formula for $X$ is a
@@
   \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
-  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW}
 \end{align}
@@
 written as a perturbative correction to the hard-sphere fluid correlation
-function $g^\textit{HS}_\sigma$ given in Equation~\ref{eq:ghs},
-evaluated using the filling fraction computed from $n_0(\xx)$.
-%, for which we use the functional of Yu and
-%Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
-%\begin{align}
-%  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
-%  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
-%  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
-%\end{align}
+function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
+\begin{align}
+  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
+  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
+  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
+\end{align}
 
@@
 some small oscillations.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
-%\includegraphics[width=\columnwidth]{figs/near-critical-point}
-%\end{center}
-%\caption{By plotting the free energy versus density it is possible to
-%  determine a saturated liquid or vapor state by the common tangent
-%  method. The top plot illustrates the common tangent near room 
-%  temperature, while the bottom plot shows the same thing close to the 
-%  critical temperature. }
-%\label{fig:homogeneous}
-%\end{figure}
 

### 2012-02-10 papers/contact/paper.tex
clean up contact paper, and avoid re-running mc calculations Tested everything on bingley.
various changes to papers Tested everything on bingley.

@@
 
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\author{Jeff Schulte}
+\author{Chris Haglund}
+\author{Patrick Kreitzburg}
 \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
 
@@
 functionals\cite{\saftlocaldft} cannot be used to study arbitrary
-inhomogeneous density distributions, because they use local
-density approximations that are only valid for densities that vary
-slowly over molecular length scales.
+inhomogeneous density distributions, because they use local density
+approximations that are only valid for densities that vary slowly over
+molecular length scales.  A key input to those SAFT-based functionals
+that can handle sharply varying densities is the correlation function
+evaluated at contact. \fixme{I need to figure out how best to phrase
+  this... do we want to compute correlation at contact, or contact
+  density itself? I prefer the latter, but the former is what is
+  commonly included in formulas.}
 
@@
 the free energy contribution from \emph{association}, which is to say,
-hydrogen bonding.  Each of these terms has some dependence on the
-inhomogeneity in the density.  The ideal gas is purely local, and in
-fact must be the only purely local term in the free energy in order
-for the functional to satisfy the contact-value theorem.  The
-hard-sphere contribution contribution to the free energy has been
-thoroughly studied, and is well approximated by the White Bear version
-of the Fundamental-Measure Theory (FMT)
-functional\cite{roth2002whitebear}.  A functional for the chain
-contribution is has recently been introduced\cite{bryk2006density,
-  gross2009density}, but is not needed for small associating molecules
-such as water.  The dispersion term is long-range and thus has
-significant dependence on the density distribution, but because its
-relatively weak position dependence, is amenable to mean-field
-approximations.  A functional for the association term of the free
-energy was introduced by Wu and
-Yu\cite{yu2002fmt-dft-inhomogeneous-associating}, and was used in
-several subsequent functionals\cite{fu2005vapor-liquid-dft,
-  bryk2006density}.  A key input (FIXME: term, factor?) to both the
-association and chain free energies is the contact value of the
-correlation function of a hard-sphere fluid.
-
-The contact value of the correlation function---which is the same as
-the contact value of the cavity correlation function for hard
-spheres---is the the value of the correlation function evaluated a
-distance equal to the hard-sphere diameter.  In an inhomogeneous
-system treated using fundamental measures, there is some ambiguity in
-the meaning of the \emph{local} contact value of the correlation
-function which appears in the equations for the homogeneous system.
-Clearly the contact value of the correlation function will depend on
-position, but how should that position be defined? It will depend on
-how the correlation function is used.  There are two reasonable
-choices for functionals based on fundamental measure theory, which
-will depend on which densities are used...
-\begin{align}
-  F[n(\rr)] = \int n(\rr) f(\bar{n}(\rr), g_\sigma(\rr)) d\rr
-\end{align}
-\begin{align}
-  F[n(\rr)] = \int f(n_2(\rr), g_\sigma(\rr)) d\rr
-\end{align}
-
-Yu and Wu introduce an approximation based on the fundamental measures
-of Fundamental Measure Theory (see Section~\ref{sec:yuwu}).  Gross
-also does\cite{gross2009density}.
-
-In this paper, we will report on the accuracy of this approximation by
-comparing with Monte Carlo simulations of the inhomogeneous
-hard-sphere fluid.  In addition, we will introduce and test differing
-approximations for the cavity correlation function.
-
-\textcolor{red}{Cavity correlation function!}
-
-The hard-sphere contact density---that is to say, the density of hard
-spheres touching a given hard sphere---is important for the
-association term in Statistical Associating Fluid Theory (SAFT).
-For a homogeneous hard-sphere fluid, the contact density---which is to
-say, $n g_{HS}(\sigma)$, where $g_{HS}(\sigma)$ is the contact value
-of the correlation function---is well-known.  However, for
-inhomogeneous fluids, the contact density has not been well studied.
-
-{
-  \color{blue}We need to figure out a few bits of terminology.
-
-  \begin{enumerate}
-  \item What should we call the contact density between two hard
-    spheres, to distinguish it from the contact density at any other
-    hard surface.  One option would be to focus on the contact value
-    of the correlation function, which would make it clear what we're
-    talking about.
-  \item How should we distinguish between the various definitions.  In
-    particular, ``at the sphere'' vs. ``at contact''.  It'd be helpful
-    to have a short way of describing these.
-  \end{enumerate}
-}
+hydrogen bonding.  The chain and association contributions to the SAFT
+free energy are of particular interest for this paper, since each uses
+as input the contact value of the correlation function.
+%
+Yu and Wu introduced in 2002 a functional for the association term of
+the free energy, which included a functional for the contact value of
+the correlation function (see
+Section~\ref{sec:yuwu})\cite{yu2002fmt-dft-inhomogeneous-associating},
+which has subsequently been used in other
+functionals\cite{fu2005vapor-liquid-dft, bryk2006density}.
+%% Each of these terms has some dependence on the
+%% inhomogeneity in the density.  The ideal gas is purely local, and in
+%% fact must be the only purely local term in the free energy in order
+%% for the functional to satisfy the contact-value theorem.  The
+%% hard-sphere contribution contribution to the free energy has been
+%% thoroughly studied, and is well approximated by the White Bear version
+%% of the Fundamental-Measure Theory (FMT)
+%% functional\cite{roth2002whitebear}.
+Two functionals for the chain contribution have recently been
+introduced\cite{bryk2006density, gross2009density}, one which uses the
+contact value of the correlation function of Yu and
+Wu\cite{bryk2006density}, while the other introduces a new functional
+for the contact value of the correlation
+function (see Section~\ref{sec:gross})\cite{gross2009density}.
+%% The dispersion term is long-range and thus has significant
+%% dependence on the density distribution, but because its relatively
+%% weak position dependence, is amenable to mean-field approximations.
+Given these different approaches, it seems worthwhile to examine this
+property of the hard-sphere fluid through direct simulation, in order
+to establish the advantages and disadvantages of each approach.
+
+\fixme{Should we include these equations? Maybe just later?}
+\begin{align}
+  \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(g_\sigma\right)-1\right)
+\end{align}
+\begin{align}
+  \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
+  X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
+\end{align}
+
+\section{Contact density and correlation function}
+
+\fixme{We should decide on good names for these two densities.  In
+  this section I define them formally, and we should keep using the
+  same names (and mathematical symbols) for them throughout.}
+
+We will define our terms by making use of the two-particle density
+$n^{(2)}(\rr_1,\rr_2)$, which gives the probability per unit volume of
+finding one particle at position $\rr_1$ and the other at position
+$\rr_2$.  The pair correlation function is defined by
+\begin{align}
+  g(\rr_1,\rr_2) &\equiv \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
+\end{align}
+In a homogeneous fluid, the pair correlation only depends on the
+distance $|\rr_1-\rr_2|$, and can be expressed as a function of a
+single variable, with the contact value being when that distance is
+$2R$.  It is desirable for reasons of efficiency to limit CDFT
+functionals to one-center convolutions, which leads us to seek a
+simplified expression for the contact value of the correlation
+function---which is the same as the contact value of the cavity
+correlation function for hard spheres.
+%
+\fixme{The following bit is awkward.}
+%
+In a system with an inhomogeneous density, we seek a \emph{local}
+contact value of the correlation function.  There are two reasonable
+options for defining such a local function:  a symmetric and an
+asymmetric formulation.
+
+The first option treats the two particles that are in contact
+symmetrically:
+\begin{align}
+  g^a_\sigma(\rr) &= \frac{n^a_\textit{contact}(\rr)}{n_a(\rr)}
+\end{align}
+\begin{align}
+  n^a_\textit{contact}(\rr) &= \frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
+  R)d\rr'
+\end{align}
+\begin{align}
+  n_a(\rr) &= \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
+\end{align}
+This functional gives a value averaged over spheres that touch at the
+position $\rr$.  The density $n_a$ is the same as the density $n_0$
+defined in Fundamental Measure Theory (FMT), and seems ideal for
+treating touching spheres.
+
+The second option is to treat one of the two touching spheres
+specially:
+\begin{align}
+  g^b_\sigma(\rr) &= \frac{n^b_\textit{contact}(\rr)}{n_b(\rr)} \\
+  n^b_\textit{contact}(\rr) &= \frac1{16\pi R^2}\int n^{(2)}(\rr, \rr + \rr')
+  \delta(|\rr'| - 2R)d\rr' \\
+  n_b(\rr) &= \frac1{16\pi R^2}\int n(\rr') \delta(|\rr-\rr'|-2R) d\rr'
+\end{align}
+In this case, we find the average number of contacts for a sphere at a
+given position.
+
+In the process of defining these two averaged correlation functions,
+we also defined four averaged densities, two \emph{contact
+  densities}, $n^a_\textit{contact}(\rr)$ and $n^b_\textit{contact}(\rr)$,
+and two \emph{contacting densities}, $n_a(\rr)$ and $n_b(\rr)$.  The
+former are the density of spheres that are in contact with each other,
+while the latter are the density of spheres available to be in
+contact.  In the homogeneous limit, $n_a = n_b = n$ and
+$n^a_\textit{contact} = n^b_\textit{contact} = g(\sigma)n$.  It is an
+open question, which of these averages will be more useful in any
+particular functional.  Suffice to say, either average is a
+\emph{possible} way to convert a function that is defined for a
+homogeneous system to a functional that is applicable to inhomogeneous
+systems.
+
+Yu and Wu introduce an approximation for
+$g^a$\cite{yu2002fmt-dft-inhomogeneous-associating}, which will be
+discussed in Section~\ref{sec:yuwu}.  Gross introduced an
+approximation for $g^b$\cite{gross2009density}, which will be
+introduced in Section~\ref{sec:gross}.  We will also introduce our own
+approximation to $g^b$ which is based on the assumption of
+thermodynamic consistency within FMT, in
+Section~\ref{contact-at-sphere}.  We will report on the accuracy of
+these approximations by comparing with Monte Carlo simulations of the
+inhomogeneous hard-sphere fluid.  We will focus on the contact
+densities, rather than the correlation function itself, because these
+are simpler, and because they are the quantity that you want to
+``average''.  i.e. if you were to average $g$ itself, you'd want to
+weight that average based on the density, which would mean averaging
+$n^{(2)}$.
 
@@
 
-\section{Direct correlation function}
-
-\textcolor{red}{FIXME WRITE THIS SECTION!} The direct correlation
-function is readily accessible from density-functional theory, as the
-second functional derivative of the free energy:
-\textcolor{red}{(Something like...)}
-\begin{align}
-  -c^{(2)}(\xx_1, \xx_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\xx_1) \delta n(\xx_2)}
-\end{align}
-The direct correlation function is related to the ordinary correlation
-function by the Ornstein-Zernike relation
-\begin{align}
-  h(\xx_1, \xx'_2) &= g(\xx_1, \xx_2) - 1 \\
-  h(\xx_1, \xx_2) &= c(\xx_1, \xx_2)
-    + \int n(\xx_3)c(\xx_1, \xx_3) h(\xx_2, \xx_3)d\xx''
-\end{align}
-From these two equation, we could in principle extract the correlation
-function evaluated at contact from the DFT free energy functional.
-The Ornstein-Zernike equation is usually solved by taking a Fourier
-transform of all the relevant quantities.  In the inhomogeneous case,
-however, this approach isn't workable, since the Fourier transforms
-don't speed things up like you'd hope, and we don't have a periodic
-system in general.
-
-\textcolor{red}{It seems worth looking into this, to see what (if
-  anything) we can do! I don't see a solution just now, but that
-  doesn't mean there isn't one around somewhere.  In any case, just
-  looking at the direct correlation function might give us some
-  hints as to a different and interesting approximation for the
-  contact density.}
+%% \section{Direct correlation function}
+
+%% \textcolor{red}{FIXME WRITE THIS SECTION!} The direct correlation
+%% function is readily accessible from density-functional theory, as the
+%% second functional derivative of the free energy:
+%% \textcolor{red}{(Something like...)}
+%% \begin{align}
+%%   -c^{(2)}(\xx_1, \xx_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\xx_1) \delta n(\xx_2)}
+%% \end{align}
+%% The direct correlation function is related to the ordinary correlation
+%% function by the Ornstein-Zernike relation
+%% \begin{align}
+%%   h(\xx_1, \xx'_2) &= g(\xx_1, \xx_2) - 1 \\
+%%   h(\xx_1, \xx_2) &= c(\xx_1, \xx_2)
+%%     + \int n(\xx_3)c(\xx_1, \xx_3) h(\xx_2, \xx_3)d\xx''
+%% \end{align}
+%% From these two equation, we could in principle extract the correlation
+%% function evaluated at contact from the DFT free energy functional.
+%% The Ornstein-Zernike equation is usually solved by taking a Fourier
+%% transform of all the relevant quantities.  In the inhomogeneous case,
+%% however, this approach isn't workable, since the Fourier transforms
+%% don't speed things up like you'd hope, and we don't have a periodic
+%% system in general.
+
+%% \textcolor{red}{It seems worth looking into this, to see what (if
+%%   anything) we can do! I don't see a solution just now, but that
+%%   doesn't mean there isn't one around somewhere.  In any case, just
+%%   looking at the direct correlation function might give us some
+%%   hints as to a different and interesting approximation for the
+%%   contact density.}
 
@@
 
-\section{Mean Contact density}\label{simple-contact}
-
-One reasonably simple question to ask is what the \emph{mean} contact
-density of an inhomogeneous system is.  This comes down to simply
-reproducing the derivation for a homogeneous system using the
-FMT functional.
-\begin{align}
-  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
-  \ncontact &= \frac{1}{N k_BT 4\pi (2R)^2} \frac{dA_{HS}}{dR}
-\end{align}
-While this should give the true mean contact density for any given
-system, this is not particularly useful, as we have no way of knowing
-the contact density on any given sphere.  However, the mean contact
-density comes out as an integral, so we can plot the integrand, and
-treat it as a somewhat ill-defined local contact density.
-
-\derivation{
-
-  To compute the mean contact density of a system, we need
-  \begin{align}
-    \frac{d A_{HS}}{d R} &=
-    \frac{\partial A_{HS}}{\partial R} \\
-    &+
-    \int \left(
-    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-    \frac{d n_3(\mathbf{r}')}{d R}
-    +
-    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-    \frac{d n_2(\mathbf{r}')}{d R}
-    + \cdots
-    \right) d\mathbf{r}'
-  \end{align}
-
-  \begin{align}
-    \frac{dn_3(\mathbf{r})}{dR} &= n_2(\mathbf{r})\\
-    \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
-    \delta'(|\mathbf{r'}| - R)\mathbf{dr}'\\
-    &= \int n(\mathbf{r}+\mathbf{r}')
-    \delta'(r' - R) d\Omega' r'^2 dr'\\
-    &= \int 
-    \delta(r' - R) d\Omega' \left(\frac{dn(\mathbf{r}+\mathbf{r}')}{dr'}r'^2 + 2n(\mathbf{r}+\mathbf{r}')r' \right) dr'\\
-    &= \frac1{R} \int 
-    \delta(|\mathbf{r}'| - R)
-    \left(\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') +
-    2n(\mathbf{r}+\mathbf{r}') \right) \mathbf{dr}' \\
-    &= \frac{2}{R}n_2(\mathbf{r}) + \frac{1}{R} \int 
-    \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-    \frac{dn_0(\mathbf{r})}{dR} &= \frac{1}{R} \int 
-    \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-    \frac{dn_1(\mathbf{r})}{dR} &=
-    \frac1{4\pi R}\frac{dn_2(\mathbf{r})}{dR} - n_0(\mathbf{r}) \\
-    &= n_0(\mathbf{r}) + \frac{1}{4\pi R^2} \int 
-    \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-    \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
-  \end{align}
-
-  The total force due to $\Phi_1$ is actually very easy to work out:
-  \begin{align}
-    \frac{d A_{HS}^{(1)}}{d R} &=
-    \int \mathbf{dr}' \left(\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
-  \end{align}
-  where you can see that $\frac{dn_0(\mathbf{r})}{dR}$ is zero for a
-  homogeneous system, and small for reasonably smooth density
-  distributions.  The next term from $\Phi_2$ is as follows:
-  \begin{align}
-    \frac{d A_{HS}^{(2)}}{d R} &=
-    \int \mathbf{dr}' \left(
-    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2} - 2 n_1
-      \frac{dn_2}{dR} + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot
-    }{1-n_3}
-    \right) \\
-    &= \int \mathbf{dr}' \left(
-    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      - 2 n_1 \left(\frac2{R}n_2 \frac{dn_0}{dR}\right)
-      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-    \right) \\
-    &= \int \mathbf{dr}' \left(
-    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      - 4 n_0n_2 - 2 n_1\frac{dn_0}{dR}
-      + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-    \right) \\
-    &= \int \mathbf{dr}' \left(
-    \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-    n_2
-    + \frac{3n_0n_2 + \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-      + 2 n_1\frac{dn_0}{dR}
-      - 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-    \right)
-  \end{align}
-
-  And the last term looks like this.  I should perhaps note that this
-  term should be replaced by a tensor version if want to examine
-  seriously localized density distributions.
-  \begin{align}
-    A_{HS}^{(3)} &=
-    \int \mathbf{dr}'\left(
-    (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-    \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2\left( 1 - n_3 \right)^2
-    }
-    \right) \\
-    \frac{d A_{HS}^{(3)}}{d R} &=
-    \int \mathbf{dr}'
-    \left[
-      3(n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-      \left(\frac2{R}n_2 + \frac{dn_0}{dR}\right)
-      - 6n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-      \right]
-    \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2\left( 1 - n_3 \right)^2
-    }
-    \\
-    &+
-    (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) n_2
-    \frac{
-      (1 - 2(1-n_3)\ln(1-n_3) - (1-n_3))n_3^2( 1 - n_3)^2
-      -
-      (n_3 + (1-n_3)^2 \ln(1-n_3))(-2 n_3^2( 1 - n_3) + 2n_3(1-n_3)^2)
-    }{
-      36\pi (n_3^2\left( 1 - n_3 \right)^2)^2
-    }
-    \\
-    &=
-    \int \mathbf{dr}'
-    \left[
-      (n_1n_2 - \mathbf{n}_{V1} \cdot \mathbf{n}_{V2})
-      \left(2n_2 + \frac{1}{4\pi}\frac{dn_0}{dR}\right)
-      - \frac{1}{2\pi}n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-      \right]
-    \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      3 n_3^2\left( 1 - n_3 \right)^2
-    }
-    \\
-    &+
-    \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
-    \left( \frac{1}{(1-n_3)^2} + 2 \frac{\ln(1-n_3)}{n_3^2(1-n_3)} \right)
-  \end{align}
-
-  \begin{align}
-    \frac{d A_{HS}}{d R} &=
-    \int \mathbf{dr}' \left(
-    -\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3)
-    + \cdots
-    \right)
-  \end{align}
-}
+%% \section{Mean Contact density}\label{simple-contact}
+
+%% One reasonably simple question to ask is what the \emph{mean} contact
+%% density of an inhomogeneous system is.  This comes down to simply
+%% reproducing the derivation for a homogeneous system using the
+%% FMT functional.
+%% \begin{align}
+%%   p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
+%%   \ncontact &= \frac{1}{N k_BT 4\pi (2R)^2} \frac{dA_{HS}}{dR}
+%% \end{align}
+%% While this should give the true mean contact density for any given
+%% system, this is not particularly useful, as we have no way of knowing
+%% the contact density on any given sphere.  However, the mean contact
+%% density comes out as an integral, so we can plot the integrand, and
+%% treat it as a somewhat ill-defined local contact density.
+
+%% \derivation{
+
+%%   To compute the mean contact density of a system, we need
+%%   \begin{align}
+%%     \frac{d A_{HS}}{d R} &=
+%%     \frac{\partial A_{HS}}{\partial R} \\
+%%     &+
+%%     \int \left(
+%%     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
+%%     \frac{d n_3(\mathbf{r}')}{d R}
+%%     +
+%%     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
+%%     \frac{d n_2(\mathbf{r}')}{d R}
+%%     + \cdots
+%%     \right) d\mathbf{r}'
+%%   \end{align}
+
+%%   \begin{align}
+%%     \frac{dn_3(\mathbf{r})}{dR} &= n_2(\mathbf{r})\\
+%%     \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
+%%     \delta'(|\mathbf{r'}| - R)\mathbf{dr}'\\
+%%     &= \int n(\mathbf{r}+\mathbf{r}')
+%%     \delta'(r' - R) d\Omega' r'^2 dr'\\
+%%     &= \int 
+%%     \delta(r' - R) d\Omega' \left(\frac{dn(\mathbf{r}+\mathbf{r}')}{dr'}r'^2 + 2n(\mathbf{r}+\mathbf{r}')r' \right) dr'\\
+%%     &= \frac1{R} \int 
+%%     \delta(|\mathbf{r}'| - R)
+%%     \left(\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') +
+%%     2n(\mathbf{r}+\mathbf{r}') \right) \mathbf{dr}' \\
+%%     &= \frac{2}{R}n_2(\mathbf{r}) + \frac{1}{R} \int 
+%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+%%     \frac{dn_0(\mathbf{r})}{dR} &= \frac{1}{R} \int 
+%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+%%     \frac{dn_1(\mathbf{r})}{dR} &=
+%%     \frac1{4\pi R}\frac{dn_2(\mathbf{r})}{dR} - n_0(\mathbf{r}) \\
+%%     &= n_0(\mathbf{r}) + \frac{1}{4\pi R^2} \int 
+%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
+%%     \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
+%%   \end{align}
+
+%%   The total force due to $\Phi_1$ is actually very easy to work out:
+%%   \begin{align}
+%%     \frac{d A_{HS}^{(1)}}{d R} &=
+%%     \int \mathbf{dr}' \left(\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
+%%   \end{align}
+%%   where you can see that $\frac{dn_0(\mathbf{r})}{dR}$ is zero for a
+%%   homogeneous system, and small for reasonably smooth density
+%%   distributions.  The next term from $\Phi_2$ is as follows:
+%%   \begin{align}
+%%     \frac{d A_{HS}^{(2)}}{d R} &=
+%%     \int \mathbf{dr}' \left(
+%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+%%     n_2
+%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2} - 2 n_1
+%%       \frac{dn_2}{dR} + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot
+%%     }{1-n_3}
+%%     \right) \\
+%%     &= \int \mathbf{dr}' \left(
+%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+%%     n_2
+%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+%%       - 2 n_1 \left(\frac2{R}n_2 \frac{dn_0}{dR}\right)
+%%       + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+%%     \right) \\
+%%     &= \int \mathbf{dr}' \left(
+%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+%%     n_2
+%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+%%       - 4 n_0n_2 - 2 n_1\frac{dn_0}{dR}
+%%       + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+%%     \right) \\
+%%     &= \int \mathbf{dr}' \left(
+%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
+%%     n_2
+%%     + \frac{3n_0n_2 + \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
+%%       + 2 n_1\frac{dn_0}{dR}
+%%       - 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
+%%     \right)
+%%   \end{align}
+
+%%   And the last term looks like this.  I should perhaps note that this
+%%   term should be replaced by a tensor version if want to examine
+%%   seriously localized density distributions.
+%%   \begin{align}
+%%     A_{HS}^{(3)} &=
+%%     \int \mathbf{dr}'\left(
+%%     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
+%%     \frac{
+%%       n_3 + (1-n_3)^2 \ln(1-n_3)
+%%     }{
+%%       36\pi n_3^2\left( 1 - n_3 \right)^2
+%%     }
+%%     \right) \\
+%%     \frac{d A_{HS}^{(3)}}{d R} &=
+%%     \int \mathbf{dr}'
+%%     \left[
+%%       3(n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
+%%       \left(\frac2{R}n_2 + \frac{dn_0}{dR}\right)
+%%       - 6n_2 \mathbf{n}_{V2} \cdot \nabla n_2
+%%       \right]
+%%     \frac{
+%%       n_3 + (1-n_3)^2 \ln(1-n_3)
+%%     }{
+%%       36\pi n_3^2\left( 1 - n_3 \right)^2
+%%     }
+%%     \\
+%%     &+
+%%     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) n_2
+%%     \frac{
+%%       (1 - 2(1-n_3)\ln(1-n_3) - (1-n_3))n_3^2( 1 - n_3)^2
+%%       -
+%%       (n_3 + (1-n_3)^2 \ln(1-n_3))(-2 n_3^2( 1 - n_3) + 2n_3(1-n_3)^2)
+%%     }{
+%%       36\pi (n_3^2\left( 1 - n_3 \right)^2)^2
+%%     }
+%%     \\
+%%     &=
+%%     \int \mathbf{dr}'
+%%     \left[
+%%       (n_1n_2 - \mathbf{n}_{V1} \cdot \mathbf{n}_{V2})
+%%       \left(2n_2 + \frac{1}{4\pi}\frac{dn_0}{dR}\right)
+%%       - \frac{1}{2\pi}n_2 \mathbf{n}_{V2} \cdot \nabla n_2
+%%       \right]
+%%     \frac{
+%%       n_3 + (1-n_3)^2 \ln(1-n_3)
+%%     }{
+%%       3 n_3^2\left( 1 - n_3 \right)^2
+%%     }
+%%     \\
+%%     &+
+%%     \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
+%%     \left( \frac{1}{(1-n_3)^2} + 2 \frac{\ln(1-n_3)}{n_3^2(1-n_3)} \right)
+%%   \end{align}
+
+%%   \begin{align}
+%%     \frac{d A_{HS}}{d R} &=
+%%     \int \mathbf{dr}' \left(
+%%     -\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3)
+%%     + \cdots
+%%     \right)
+%%   \end{align}
+%% }
 
@@
     \right\}
-    \\
-    &= \int \left\{
-    -\frac{n_2}{4\pi R^2} \ln\left( 1 - n_3\right)
-    + \frac{1}{4\pi R} \frac{n_2^2 - \mathbf{n}_{V2} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \right\}
   \end{align}
@@
     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-    -\frac{1}{4\pi R^2} \ln\left( 1 - n_3\right)
-    +  \frac{2n_1}{1-n_3}
+    \frac{n_1}{1-n_3}
     + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
@@
     \\
+    \frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
+    \\
+    \frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &=
+    -\frac{1}{4\pi R^2} \ln\left( 1 - n_3\right) \\
     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-    \frac{2\mathbf{n}_{V1}}{1-n_3}
+    \frac{\mathbf{n}_{V1}}{1-n_3}
     - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
@@
     }    \\
+    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
+    \frac{\mathbf{n}_{V2}}{1-n_3} \\
     \frac{\delta A_{HS}}{\delta R(\mathbf{r}')} &=
@@
 We compare with the functional of Gross \emph{et
-  al}\cite{gross2009density}, which is of the first kind, where the
+  al}\cite{gross2009density}, which is of the second kind ($g^b$), where the
 contact density is evaluated at the sphere location.
 \begin{align}
-  n_\textit{contact} &= n_{0'} \frac{1 - \frac12 n_{3'}}{\left(1 - n_{3'}\right)^3}
-\end{align}
-where $n_{0'}$ and $n_{3'}$ are defined in a manner analogous to the
-$n_0$ and $n_3$ of FMT, but with the hard-sphere radius replaced by
-the hard-sphere diameter, so the integrals range over spheres that are
-twice as large.
+  n_\textit{contact} &= n_b \frac{1 - \frac12 n_{3'}}{\left(1 -
+    n_{3'}\right)^3} \\
+  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
+\end{align}
+where $n_b$ is defined in the introduction and $n_{3'}$ is an averaged
+filling fraction defined in a manner analogous to the $n_3$ of FMT,
+but with the hard-sphere radius replaced by the hard-sphere diameter,
+so the integrals range over spheres that are twice as large.
 
@@
 density at a given contact
-position~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We will
+position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We will
 compare both their functional as well as ours with simulation data.
@@
 
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-10}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:walls-10}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-30}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.3.}
+  \label{fig:walls-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-40}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.4.}
+  \label{fig:walls-40}
+\end{figure}
+
+\subsection{Hard spheres at a hard wall}
+
+We begin by examining the simplest situation with an inhomogeneous
+density:  the case of hard spheres near a single hard wall.  This is
+shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
+
+\newcommand\sphereExplanation{ Blue curves describe the density of
+  hard spheres.  Green curves describe the contact density averaged
+  according to the point of contact.  Red curves describe the contact
+  density as averaged according to the centers of the spheres that are
+  touching.  In each case, solid lines correspond to the Monte Carlo
+  simulation, and dashed lines of various sorts represent DFT model
+  predictions.  }
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-08-013}
+  \caption{Density and contact densities of thirteen hard spheres in a
+    spherical cavity with diameter 8. \sphereExplanation }
+  \label{fig:sphere-8}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-12-112}
+  \caption{Density and contact densities of 112 hard spheres in a
+    spherical cavity with diameter 12.  \sphereExplanation}
+  \label{fig:sphere-12}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-16-265}
+  \caption{Density and contact densities of 265 hard spheres in a
+    spherical cavity with diameter 16. \sphereExplanation}
+  \label{fig:sphere-16}
+\end{figure}
+
 \subsection{Hard spheres confined in a spherical cavity}
@@
 
-\textcolor{red}{We should also try looking at a few other densities.}
-
-\newcommand\sphereExplanation{ Blue curves describe the density of
-  hard spheres.  Green curves describe the contact density averaged
-  according to the point of contact.  Red curves describe the contact
-  density as averaged according to the centers of the spheres that are
-  touching.  In each case, solid lines correspond to the Monte Carlo
-  simulation, and dashed lines of various sorts represent DFT model
-  predictions.  }
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-08-013}
-  \caption{Density and contact densities of thirteen hard spheres in a
-    spherical cavity with diameter 8. \sphereExplanation }
-  \label{fig:sphere-8}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-12-112}
-  \caption{Density and contact densities of 112 hard spheres in a
-    spherical cavity with diameter 12.  \sphereExplanation}
-  \label{fig:sphere-12}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-16-265}
-  \caption{Density and contact densities of 265 hard spheres in a
-    spherical cavity with diameter 16. \sphereExplanation}
-  \label{fig:sphere-16}
-\end{figure}
-
-\subsection{Hard spheres at a hard wall}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-10}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.1.}
-  \label{fig:walls-10}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-30}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.3.}
-  \label{fig:walls-30}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-40}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.4.}
-  \label{fig:walls-40}
-\end{figure}
-
-We study the case of hard spheres near a single hard wall.  This is
-shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.  Actually, we
-should begin with this.
-
-\subsection{Hard spheres around a hard sphere}
-
-It is also interesting to study a system with opposite curvature.  Put
-a (larger) hard sphere in the hard-sphere fluid, and see what the
-contact density is around that hard sphere.  This is particularly
-relevant when we consider using classical density-functional theory to
-model solvation.
-
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Hard spheres around a hard sphere}
+
+It is also interesting to study a system with opposite curvature.  Put
+a (larger) hard sphere in the hard-sphere fluid, and see what the
+contact density is around that hard sphere.  This is particularly
+relevant when we consider using classical density-functional theory to
+model solvation.
 

### 2012-02-20 paper/paper.tex

@@
 
-Water stands out among liquids for its extraordinary properties.  Water is
-the only natural substance that is found in solid, liquid and gaseous
-states at temperatures and pressures normally found on Earth.  Water has a
-density maximum at 4$^\circ$C, and expands upon freezing.  The heat
-capacity of water is the highest of any liquid, and its latent heat of
-fusion and vaporization are also the highest.  Water's thermal conductivity
-and surface tensions are the highest of any non-metallic liquid.  Water has
-an unusually high dielectric constant.
-Much of the unusual behavior of water is explained by the ability of each
-water molecule to form \emph{four} strong, directional hydrogen bonds---two
-acceptors and two donors.  This directional bonding dictates the open
-structure of ice, leading to thermal expansion upon freezing.
+Water stands out among liquids for its extraordinary properties.
+Water is the only natural substance that is found in solid, liquid and
+gaseous states at temperatures and pressures normally found on Earth.
+Water has a density maximum at 4$^\circ$C, and expands upon freezing.
+Water has the highest heat capacity, latent heat of fusion and
+vaporization, thermal conductivity and surface tension of any
+non-metallic liquid.  Water has an unusually high dielectric constant.
+Much of the unusual behavior of water is explained by the ability of
+each water molecule to form \emph{four} strong, directional hydrogen
+bonds---two acceptors and two donors.  This directional bonding
+dictates the open structure of ice, leading to thermal expansion upon
+freezing.
 
@@
 sophisticated than those describing the hard-sphere fluid, and a
-quantitatively accurate functional describing water---which reproduces both
-the macroscopic surface tension and its decrease as curvature
-increases---has not yet been published.  For instance, Jaqaman \emph{et al}
-developed a classical density-functional for water including orientational
-effects, but their functional resulted in a surface tension twice the
-experimental value, with the result that it can really only be used to make
-very qualitative predictions~\cite{Jaqaman2004}.
+quantitatively accurate functional describing water---which reproduces
+both the macroscopic surface tension and its decrease as curvature
+increases---has not yet been published.  For instance, Jaqaman
+\emph{et al} developed a classical density-functional for water
+including orientational effects, but their functional resulted in a
+surface tension twice the experimental value, limiting it to purely
+qualitative predictions~\cite{Jaqaman2004}.
 
@@
 
+\section{Theory and Methods}
 We use classical DFT to construct a new functional that will predict
@@
 interactions in water are ideal for describing with SAFT, which was designed for
-use with such fluids. It is based on Wertheim's first-order
+use with hydrogen-bonding fluids. It is based on Wertheim's first-order
 thermodynamic perturbation theory (TPT1)
@@
 associative interations between molecules. SAFT can account for the association energies
-as well as any effects due to clusters or chains forming
-in the liquid\cite{muller2001molecular}. \needsworknow{Expand or move this 
-introduction to SAFT?}
-
-\section{Theory and Methods}
-We introduce a new classical density functional for water that
-predicts the vapor pressure, vapor density, liquid density, bulk
-surface tension, and enthalpy, as well as allows for liquid-vapor
-coexistence.  The Helmholtz free energy functional is based on SAFT,
-and is composed of the usual terms:
+as well as any effects due to clusters or chains forming \fixme{not true}
+in the liquid\cite{muller2001molecular}.  \needsworknow{Expand this a
+  bit to accurately describe SAFT and its advantages.}
+
+The Helmholtz free energy functional is based on SAFT,
+and is composed of the terms:
 \begin{equation}
@@
 approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters, an interaction energy $\epsilondisp$ and a
+parameters: an interaction energy $\epsilondisp$ and a
 length scale $\lambdadisp R$.
@@
 a cavity within which the water may not penetrate.  This is the
-simples possible solute, and will serve as a test case, which if
+simplest possible solute, and will serve as a test case, which if
 passed could lead to the ability to insert different chemicals into

### 2012-02-21 papers/contact/mod-paper.tex
modified paper changed white bear deriv a bit and added work in progress paper.tex Tested everything on crystal.

--- 2012-02-10 papers/contact/paper.tex
@@
 
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp,10pt,aip]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
 \author{Jeff Schulte}
@@
 \author{David Roundy}
+\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
 
@@
 \maketitle
-
-%\input{formulas}
 
@@
     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    -\frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
     + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
       n_3(\mathbf{r}'))^2}
-    + \frac{n_2^3 -
+    - \frac{n_2^3 -
       3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-    \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
+    \frac{n_3^2-5n_3+2}{n_3^3(1-n_3)^3} + 2\frac{\ln(1-n_3)}{n_3^3}
     \right)
@@
     \frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &=
-    -\frac{1}{4\pi R^2} \ln\left( 1 - n_3\right) \\
+    -\ln\left( 1 - n_3\right) \\
     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
@@
   \end{align}
+
+  ??? Derivatives of the density functions with respect to R.
+  n_3(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(?)) d\mathbf{r}''
+
+  For the derivation of n_3 in the symmetric case:
+  \frac{\delta n_3 (\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  \int n (\mthbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}''
+
+  R(\mathbf{r}) refers to all the spheres that just touch the point \mathbf{r}
+
+  some stuff times p(\mathbf{r}) &= \int \frac{\deltaA_{HS}}{\delta n_3(\methbf{r}')} 
+  \frac{\delta n_3(\mathbf{r}'}{\deltaR(\mthbf{r})} d\mathbf{r}'
+
+  &= \int \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}'} \int n (\mthbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}'' d\mathbf{r}'
+
+  &= \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}'} \int n (\mthbf{r}'') \delta(|\mathbf{r} - \mathbf{r}''| - R(\mathbf{r})) d\mathbf{r}''
+
+  &= \frac{\delta A_{HS}}{\deltan_3(\mathbf{r}''} n_2(\mathbf{r})
+
+  So only one n_3(\methbf{r}' is affected, and that's when \mathbf{r}' = \mathbf{r}
+  It's affected by all the spheres that \emph{touch} \mathbf{r}'.  To say that R(\mathbf{r})
+  changes means that all the spheres touching \methbf{r}'.  And keep in mind these are possible spheres,
+  given by densities, not real spheres.  (Write this more good like grammar and stuff.  Look at how to do superscripts)
+
+  And for Assymmetric case:
+
+  \frac{\delta n_3 (\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  \int n (\mthbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}''
+
+  &= n (\mthbf{r}) \delta(|\mathbf{r} - \mathbf{r}'| - R(\mathbf{r}))
+
+  here it's a lot of n_3(\mathbf{r}') points that are affected by the sphere changing at R(\mathbf{r}) 
+  so sphere is centered at \mathbf{r} here
+
+  Also, here's n_2(\mathbf{r}') for Symmetric case:
+
+  n_2(\mathbf{r}')
+
+  eh.  I'm done for tonight.  Keep going with this.
+
+  \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})} &=
+    n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \\
+
+n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
+  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
+
   These give us some basic tools with which to perform derivatives.
@@
 density evaluated at the sphere location.  Alongside these, we plot
-the contact density of Fu and Wu~\cite{fu2005vapor-liquid-dft} (or
-should it be Yu and Wu\cite{yu2002structures,
-  yu2002fmt-dft-inhomogeneous-associating}?), the ``simple contact
-density'' from Section~\ref{simple-contact}, and the ``contact density
-at sphere'' from Section~\ref{contact-at-sphere}.  In principle, the
-functional of Wu (FIXME) should match the CenConDensity, while the DFT
+the contact density of Yu and
+Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, the contact density
+of Gross~\cite{gross2009density}, and the ``contact density at
+sphere'' from Section~\ref{contact-at-sphere}.  In principle, the
+functional of Yu (FIXME) should match the CenConDensity, while the DFT
 at sphere should match the ConDensity...
@@
 
-%% \begin{widetext}
-%% \section{Formulas generated from Haskell}
-
-%% \begin{align}
-%%   \Phi_1 &= \phione \\
-%%   \Phi_2 &= \phitwo \\
-%%   \Phi_3 &= \phithree \\
-%%   \frac{d F}{dn_3} &= \dwbdnthree \\
-%%   \frac{d F}{dn_2} &= \dwbdntwo \\
-%%   \frac{d F}{dn_1} &= \dwbdnone \\
-%%   \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
-%%   \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev \\
-%% \end{align}
-%% \end{widetext}
+\begin{widetext}
+\section{Formulas generated from Haskell}
+
+\input{formulas}
+\begin{align}
+  \Phi_1 &= \phione \\
+  \Phi_2 &= \phitwo \\
+  \Phi_3 &= \phithree \\
+  \frac{d F}{dn_3} &= \dwbdnthree \\
+  \frac{d F}{dn_2} &= \dwbdntwo \\
+  \frac{d F}{dn_1} &= \dwbdnone \\
+  \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
+  \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev
+\end{align}
+\end{widetext}
 

### 2012-02-24 paper/march-2012.tex
add march meeting talk Tested everything on bingley.

\documentclass[compress]{beamer}
\usepackage{graphicx}
\usepackage{color}
\usepackage{verbatim}
\usepackage{amsmath}
\usepackage{bm}

%\usetheme{Frankfurt}
%\usecolortheme{lily}

%\useoutertheme[subsection=true]{smoothbars}
\useoutertheme[subsection=true]{miniframes}
\usecolortheme{whale}
\usecolortheme{orchid}
\useinnertheme{rounded}

\newcommand{\xx}{\textbf{r}}

\title{A Classical Density-Functional Theory for Describing Water
  Interfaces}

\author{Jessica Hughes, Eric Krebs, David Roundy}

\date{February 27, 2012}

\setbeamertemplate{navigation symbols}{}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

\begin{frame}{Outline}
  \tableofcontents
\end{frame}

\section{Methods}

\subsection{Statistical Associating Fluid Theory}

\begin{frame}{}
  SAFT models fluids that have strong association interactions
  (i.e. hydrogen bonding).

  \begin{block}{The SAFT Helmholtz free energy}
    \begin{equation*}
      A = A_\text{id} + A_\text{hs} + A_\text{assoc} + A_\text{disp}
    \end{equation*}
  \end{block}

  \vfill

  We use the optimized SAFT-VR model for water from Clark \emph{et
    al.}$^1$ in the homogeneous limit.  This reproduces water's
  equation of state moderately well away from the critical point.

  \vfill

  \scriptsize{$^1$G.N.I. Clark, A.J. Haslam, A. Galindo, and 
    G. Jackson.\\ \emph{Molecular Physics}, 104(22):3561-3581, 2006.}
\end{frame}

%% \begin{frame}[fragile]{}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figs/temperature-versus-density}
%% \end{center}
%% \end{figure}
%% \end{frame}

\begin{frame}[fragile]{}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms-truncated}
\end{center}
\end{figure} 
\end{frame}

\newcommand\rr{\mathbf{r}}

\subsection{Classical density-functional theory}

\begin{frame}{Classical density-functional theory}

  \begin{block}{}
    Once we construct a classical density functional, we can find the
    free energy and equilibrium density of a fluid from its external
    potential:
    \begin{equation*}
      A(T) = \underset{n(\xx)}{min}\left\{ F[n(\xx),T] + \int V_{ext}(\xx) n(\xx)
      d^3r\right\}
    \end{equation*}
  \end{block}
  This allows us to address arbitrary \emph{inhomogeneous} density
  distributions.  The challenge is to construct a good functional.

  \vfill

  We have constructed a functional the reproduces the bulk properties
  of the optimized SAFT-VR model of Clark \emph{et al}.

\end{frame}

\begin{frame}[fragile]{}
  \begin{block}{Free energy functional}
    \begin{equation*}
      F[n] = F_\text{id}[n] + F_\text{hs}[n] + F_\text{assoc}[n]  + F_\text{disp}[n]
    \end{equation*}
  \end{block}
  \begin{itemize}
  \item $F_\text{id}[n]$: Ideal gas free energy (the only \emph{local}
    term)
  \item $F_\text{hs}[n]$: Hard sphere excess free energy\\

    We use the White Bear fundamental-measure theory functional, which
    is using convolved densities such as
    \begin{align*}
      n_3(\rr) &= \int n(\rr') \Theta(|\rr-\rr'|-R) d\rr'
    \end{align*}
    $n_2(\rr)$, $n_1(\rr)$, etc. are defined similarly.
  \item $F_\text{assoc}[n]$: Association contribution adapted from Yu
    and Wu$^1$
  \item $F_\text{disp}[n]$: Dispersion contribution (see next slide)
  \end{itemize}

  \vfill

  \scriptsize{$^1$Y.X. Yu and J. Wu., J. Chem. Phys., \textbf{116},
    7094 (2002).}
\end{frame}

%% \begin{frame}[fragile]{Association Free Energy}
%%   The association term models the hydrogen bonds as four association
%%   sites on the surface of the hard sphere.  We use the association
%%   functional of Yu and Wu$^1$, modified to use a SAFT-VR model.

%%   \vfill

%%   \scriptsize{$^1$Y.X. Yu and J. Wu., J. Chem. Phys., \textbf{116},
%%     7094 (2002).}

%%   %% \begin{block}{}
%%   %%   \vspace{-1em}
%%   %%   \begin{align*}
%%   %%     F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
%%   %%     \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx \\
%%   %%     X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
%%   %%         \Delta(\xx)} - 1}
%%   %%     {4 n_0(\xx)\zeta(\xx)
%%   %%       \Delta(\xx)} \\
%%   %%     \Delta(\xx) &= \kappa g^{\text{SW}}_\sigma(\xx)\left(e^{-\beta\epsilon} - 1\right)
%%   %%   \end{align*}
%%   %% \end{block}
%%   %% Here $g^{\text{SW}}_\sigma$ is the correlation function evaluated at contact

%%   %% The functional $X$ is the fraction of association sites \emph{not}
%%   %% hydrogen-bonded.  We use the functional of Yu and Wu$^1$ for the
%%   %% correlation function at contact
%% \end{frame}

\begin{frame}[fragile]{Dispersion Free Energy}
  The dispersion free energy includes any orientation-independent
  interactions. We use a dispersion term based on the SAFT-VR
  approach, with the filling fraction $\eta$ replaced by a convolved
  density
  \begin{align*}
    \eta(\rr) &= \frac{4\pi R^3}{3} \int n(\rr')
    \frac{e^{-\frac{|\rr-\rr'|^2}{2\sigma^2}}}{8\pi^\frac32\sigma^3} d\rr'
  \end{align*}
  where $\sigma$ is an additional empirical parameter used to fit the
  bulk surface tension.
\end{frame}

\begin{frame}[fragile]{}
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/surface-tension}
\end{center}
\end{figure} 
\end{frame}

\section{Results}
%% \subsection{Single hydrophobic rod}

%% \begin{frame}{}
%% \begin{columns}
%%  \column {0.5\textwidth}
%%   \begin{figure}
%%   \begin{center}
%%   \includegraphics[width=4.5cm]{../papers/thesis-hughes/figs/single-rod-diagram}
%%   \end{center}
%%   \end{figure}
%%   Cross section diagram of a single hydrophobic rod in water.
%%  \column {0.5\textwidth}
%%   \begin{itemize}
%%   \item Pressure = 1~atm
%%   \item Temperature = 298~K
%%   \item Lattice spacing $\approx$~0.01~nm 
%%   \item Radius = 0~nm to 1.5~nm
%%   \end{itemize}
%% \end{columns}
%% \end{frame}

%% \begin{frame}[fragile]{}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figs/density-single-rod}
%% \end{center}
%% \end{figure}
%% \end{frame}

%% \begin{frame}[fragile]{}
%% \begin{figure}
%% \begin{center}
%% \includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
%% \end{center}
%% \end{figure} 
%% \end{frame}

\subsection{Two hydrophobic rods}

\begin{frame}[fragile]{}
Two hard purely hydrophobic rods in water.
\begin{figure}
\begin{center}
\includegraphics[width=\columnwidth]{figs/rods-diagram}
\end{center}
\end{figure}
We vary the radius $r$ from 0.1~nm to 1.0~nm and the separation $d$
from 0~nm to about 1.2~nm.
\end{frame}

\begin{frame}[fragile]{1 nm  rod diameter}
\begin{columns}
 \column{0.7\textwidth}
  \begin{figure}
  \includegraphics[width=7.3cm]{figs/density-rods-in-water}
  \end{figure} 
 \column{0.3\textwidth}
 \small{``Stadium''-shaped vapor-filled area between rods}\\
 \vspace{0.3cm}
 \small{$d=$~0.5~nm}\\
 \vspace{1.4cm}
 \small{Liquid-filled area between two rods}\\
 \vspace{0.3cm}
 \small{$d=$~0.6~nm}
\end{columns}
\end{frame}

\begin{frame}[fragile]{}
  \begin{columns}
    \begin{column}{0.5\columnwidth}
      \begin{block}{A simple surface-tension model}
        \begin{itemize}
        \item Free energy of the rods $\approx$ surface area times surface tension
        \item Surface tension of bulk water = 72 mN/m
        \item Circumference of the stadium shape:
          \begin{equation*}
            C_{s} = 2\pi r +4r+2d
          \end{equation*}
        \item Circumference of two circles:
          \begin{equation*}
            C_{c} = 4\pi r
          \end{equation*}
        \end{itemize}
      \end{block}
    \end{column}
    \begin{column}{0.5\columnwidth}
      \includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}

      \vfill

      \includegraphics[width=\columnwidth]{figs/stadium-vs-circles}
    \end{column}
  \end{columns}
\end{frame}

\subsection{Four hydrophobic rods}

\begin{frame}[fragile]{}
  \begin{columns}
    \begin{column}{0.5\columnwidth}
      \includegraphics[width=\columnwidth]{figs/density-four-rods}
    \end{column}
    \begin{column}{0.5\columnwidth}
    \includegraphics[width=\columnwidth]{figs/four-rods-energy-vs-distance}
    \end{column}
  \end{columns}
\end{frame}

\subsection{A hydrophobic spherical solute}
\subsection*{}

\begin{frame}[fragile]{Hydrophobic spheres in water}
  \begin{columns}
    \begin{column}{0.5\columnwidth}
      \includegraphics[width=\columnwidth]{figs/density-sphere}
    \end{column}
    \begin{column}{0.5\columnwidth}
      \includegraphics[width=\columnwidth]{figs/sphere-energy-vs-diameter}
    \end{column}
  \end{columns}
\end{frame}

\section{Conclusions}
\subsection*{}

\begin{frame}[fragile]{}
  \begin{block}{}
    \begin{itemize}
    \item Created a classical density functional for water
    \item Expected to be valid over wide range of temperatures and pressures
    \item Models hydrophobic interactions reasonably
    \item Handles both long and short length scales
    \end{itemize}
  \end{block}
  \begin{block}{To do...}
    \begin{itemize}
    \item Realistic dispersive attraction to hydrophobic surfaces
    \item Handle electrostatics
    \item Pseudopotential to couple with electrons in solute
    \end{itemize}
  \end{block}
\end{frame}

\end{document}

### 2012-03-03 papers/contact/paper.tex

@@
 to establish the advantages and disadvantages of each approach.
+\fixme{
+
+Jeff's Notes:
+A novice (me) is confused by the use of 'contact value of correlation function' throuhout this introduction while throughout the rest of the paper we refer to the 'contact density'.
+Pehaps the audince will understand that they're interchangable (if they are) but this was confusing to me.
+Also, I would think I'd like to see something like:
+
+We first discuss two possibilities for the definition of contact density.
+The first involves a more symmetric interpretation and so we call it the 'S case'.
+The second involves a sphere-centered, assymetrical interpretation and so we call it the 'A case'.
+We present where in the SAFT free energy formulation these contact densities will contribute.
+We next present a thermodynamic method that we will use to calculate the above contact densities.
+This method involves the use of surface pressure and the minimization of the excess free energy refered to above.
+We use the White Bear and White Bear Mark II formulations of these free energy terms and take neccesary derivitives in the concpetual spirit of the seperately defined S an A case contact densities.
+Finally we compare the resulting contact densities with those of Gross, those of Yu and Wu, and as well with our wn monte carlo simulations.}
 
@@
 contact value of the correlation function.  There are two reasonable
-options for defining such a local function:  a symmetric and an
-asymmetric formulation.
-
-The first option treats the two particles that are in contact
-symmetrically:
-\begin{align}
-  g^a_\sigma(\rr) &= \frac{n^a_\textit{contact}(\rr)}{n_a(\rr)}
-\end{align}
-\begin{align}
-  n^a_\textit{contact}(\rr) &= \frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
+options for defining such a local function:  a symmetric (which we refer to as S) and an
+asymmetric (which we refer to as A) formulation.
+
+For the S case:
+\begin{align}
+  g^S_\sigma(\rr) &= \frac{n^S_\textit{contact}(\rr)}{n_S(\rr)}
+\end{align}
+\begin{align}\label{eq:n-contactS}
+  n^S_\textit{contact}(\rr) &= \frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
   R)d\rr'
@@
 \begin{align}
-  n_a(\rr) &= \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
+  n_S(\rr) &= \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
 \end{align}
 This functional gives a value averaged over spheres that touch at the
-position $\rr$.  The density $n_a$ is the same as the density $n_0$
+position $\rr$.  The density $n_S$ is the same as the density $n_0$
 defined in Fundamental Measure Theory (FMT), and seems ideal for
@@
 
-The second option is to treat one of the two touching spheres
-specially:
-\begin{align}
-  g^b_\sigma(\rr) &= \frac{n^b_\textit{contact}(\rr)}{n_b(\rr)} \\
-  n^b_\textit{contact}(\rr) &= \frac1{16\pi R^2}\int n^{(2)}(\rr, \rr + \rr')
+And for the A case:
+\begin{align}
+  g^A_\sigma(\rr) &= \frac{n^A_\textit{contact}(\rr)}{n_A(\rr)} \\
+  n^A_\textit{contact}(\rr) &= \frac1{16\pi R^2}\int n^{(2)}(\rr, \rr + \rr')
   \delta(|\rr'| - 2R)d\rr' \\
-  n_b(\rr) &= \frac1{16\pi R^2}\int n(\rr') \delta(|\rr-\rr'|-2R) d\rr'
+  n_A(\rr) &= \frac1{16\pi R^2}\int n(\rr') \delta(|\rr-\rr'|-2R) d\rr'
 \end{align}
@@
 we also defined four averaged densities, two \emph{contact
-  densities}, $n^a_\textit{contact}(\rr)$ and $n^b_\textit{contact}(\rr)$,
-and two \emph{contacting densities}, $n_a(\rr)$ and $n_b(\rr)$.  The
+  densities}, $n^S_\textit{contact}(\rr)$ and $n^A_\textit{contact}(\rr)$,
+and two \emph{contacting densities}, $n_S(\rr)$ and $n_A(\rr)$.  The
 former are the density of spheres that are in contact with each other,
 while the latter are the density of spheres available to be in
-contact.  In the homogeneous limit, $n_a = n_b = n$ and
-$n^a_\textit{contact} = n^b_\textit{contact} = g(\sigma)n$.  It is an
+contact.  In the homogeneous limit, $n_S = n_A = n$ and
+$n^S_\textit{contact} = n^A_\textit{contact} = g(\sigma)n$.  It is an
 open question, which of these averages will be more useful in any
@@
 Yu and Wu introduce an approximation for
-$g^a$\cite{yu2002fmt-dft-inhomogeneous-associating}, which will be
+$g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which will be
 discussed in Section~\ref{sec:yuwu}.  Gross introduced an
-approximation for $g^b$\cite{gross2009density}, which will be
+approximation for $g^A$\cite{gross2009density}, which will be
 introduced in Section~\ref{sec:gross}.  We will also introduce our own
-approximation to $g^b$ which is based on the assumption of
+approximation to $g^A$ which is based on the assumption of
 thermodynamic consistency within FMT, in
@@
 
-  \begin{align}
-    \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-    n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \\
-    \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-    n(\mathbf{r}) \delta'(|\mathbf{r}-\mathbf{r}'| - R) \\
-    \frac{\delta \mathbf{n}_{V2}(\mathbf{r}')}{\delta R(\mathbf{r})} &=
-    -\frac{\delta}{\delta R(\mathbf{r})} \mathbf{\nabla}' n_3(\mathbf{r}')
-    \\
-    &= -\mathbf{\nabla}'
-    \left(n(\mathbf{r})\delta(|\mathbf{r}-\mathbf{r}'| - R) \right)
-  \end{align}
+*******************************************************************************8
+\fixme{The following is the same calculation for WBm2.  Not sure if you want these in the appendix or in the paper at all?}
+
+\begin{align} 
+A_{HS} &= \int \left\{
+    -n_0 \ln\left(1 - n_3\right)
+    + (1 + \frac{1}{9}n_3^2 \phi_2(n_3))   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+    + (1 - \frac{4}{9}n_3\phi_3(n_3))\frac{(n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{24\pi (1-n_3)^2}
+    \right\}
+\end{align}
+ where
+ \begin{align}
+   \phi_2 &= (6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3))\frac{1}{n_3^3}\\
+   \phi_3 &= (6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3))\frac{1}{4n_3^3}
+ \end{align}
+For the following derivitives we have
+\begin{align}
+   \frac{\delta \phi_2}{\delta n_3} &= ((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2]\frac{3}{n_3^4}\\
+   \frac{\delta \phi_3}{\delta n_3} &= (-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3))\frac{3}{4n_3^4}
+ \end{align}
+And using these
+\begin{align}
+  \frac{\delta A_{HS}}{\delta n_3} &= \psi _1 + \psi _2 + \psi _3
+\end{align}
+where
+\begin{align}
+  \psi_1 &= \frac{n_0}{1-n_3}\\
+  \psi_2 &= (\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3})(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}) + (1 + \frac{1}{9}n_3^2\phi_2)(\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2})\\
+  \psi_3 &= (\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2})(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}) + 2(1 - \frac{4}{9}n_3\phi_3)(\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3})
+\end{align}
+We also have
+\begin{align}
+  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2(n_3))\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)(\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2})\\
+  \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
+  \frac{\delta A_{HS}}{\delta n_0} &= -\ln(1 - n_3)\\
+  \frac{\delta A_{HS}}{\delta \mathbf{n}_1} &= -(1 + \frac{1}{9}n_3^2 \phi_2)\frac{\mathbf{n}_2}{1 - n_3}\\
+  \frac{\delta A_{HS}}{\delta \mathbf{n}_2} &= -(1 + \frac{1}{9}n_3^2\phi_2)\frac{\mathbf{n}_1}{1 - n_3} 
+  - (1 - \frac{4}{9}n_3\phi_3)\frac{6\mathbf{n}_2}{24\pi(1 - n_3)^2}
+\end{align}
+
+****************************************************************************************\\
+Jeff's understanding as of 02/22/2012
+
+  In order to motivate our particular forms for $\frac{\delta n_3^{S} (\mathbf{r}')}{\delta R(\mathbf{r})}$, $\frac{\delta n_2^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc.
+we interpret the contact density in the $S$ case as follows:
+  Given that there is a sphere just touching the point $\mathbf{r}'$, $n_{Contact}^{S}(\mathbf{r}')$ is the density
+  of spheres that are also touching that point so that they are touching the given sphere.
+We can see this in the integral in eq \ref{eq:n-contactS},where we sum over the double densities of spheres that are positioned opposite eachother and touching at the point $\mathbf{r}$ and then divide
+  by $n_0$, the average density of a single sphere touching the point.  Thus we have the density of contact at $\mathbf{r}$, where the
+  contact between the two spheres actually takes place at $\mathbf{r}$.
+
+Involved in the terms in our free energy functional are derivatives of density functions with respect to $R$.
+We take these integrals differently when dealing with the S and the A cases.
+In both cases we wish to calculate how the density function changes when a 'certain amount of sphere', at some point, is changed.
+The A case is conceptually simple.
+We say that a changing spherical radius at a point $\mathbf{r}$ represents the outward expansion, by a distance $\delta R$, of the shell of any sphere that is centered at $\mathbf{r}$.
+In the case of S we instead expand a multitude of spheres, all of them positioned around and just touching the point $\mathbf{r}$.
+When their radii expand each does not expand outward in all directions as in the case of A, rather they each expand outward at only the point $\mathbf{r}$.
+One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains at radius $R$.
+When we expand each sphere that touches $\mathbf{r}$ in this peculiar fashion, we have in the end the same amount of shell area protruding outward as we would if we had simply expanded one sphere in fashion of case A.
+Also the protrusions, once added up for all spheres sourrounding $\mathbf{r}$, will yield a symmetry in their direction of protrusion.
+Each protruding sphere bit will expand outward from its center and all directions will be covered uniformly.
+
+  For the derivation of $n_{3}$ in the symmetric case:
+
+\begin{align}  
+  n_3^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}')) d\mathbf{r}''\\
+  \frac{\delta n_3^{S} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}''
+\end{align}
+
+  $\delta R(\mathbf{r})$ in this equation refers to an expansion, in the fashion described above, of all the spheres that just touch $\mathbf{r}$.
+We can see that if the radii change in this fashion, the only point at which $n_3(\mathbf{r}')$ will be affected is at $\mathbf{r}' = \mathbf{r}$, so that $\delta n_3(\mathbf{r}')$ is zero everywhere except where $\mathbf{r}' = \mathbf{r}$.
+Thus we have the 2nd $\delta$ function in the lower integral.
+
+This term in the free energy functional, once integrated, takes an interesting form:
+
+\begin{align}
+  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} 
+  \frac{\delta n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}' 
+  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}'' d\mathbf{r}' \\
+  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r} - \mathbf{r}''| - R(\mathbf{r})) d\mathbf{r}''\\
+  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}'')} n_2(\mathbf{r})\\
+\end{align}
+
+ $n_2(\mathbf{r}')$ in the case of S in calculated similarly:
+\begin{align}
+  n_2^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
+    \frac{\delta n_2^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int n(\mathbf{r}'') 
+  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}''
+\end{align}
+
+For $\mathbf{n}_{V2}(\mathbf{r}')$ in the case of S we take the gradient of $n_3(\mathbf{r}')$ with respect to $\mathbf{r}'$: 
+\begin{align}
+  \mathbf{n}_{V2}^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'))
+  \frac{(\mathbf{r}' - \mathbf{r}'')}{|\mathbf{r}' - \mathbf{r}''|} d \mathbf{r}''\\
+  \frac{\delta \mathbf{n}_{V2}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= \delta(\mathbf{r} - \mathbf{r}')
+  \int n(\mathbf{r}'') \delta'(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) 
+  \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} d\mathbf{r}''
+\end{align}
+
+  In the case of A, the changing radius we refer to is the radius of a sphere that is centered at $\mathbf{r}$ and it expands in a more conventional fashion.
+The $n_3$ values that will be affected are those which are a distance $R$ fom $\mathbf{r}$.
+
+\begin{align}
+  n_3^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}'')) d\mathbf{r}''\\ 
+  \frac{\delta n_3^{A} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}'' \\
+   &= n (\mathbf{r}) \delta(|\mathbf{r} - \mathbf{r}'| - R(\mathbf{r}))
+\end{align}
+
+And likewise:
+\begin{align}
+  n_2^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) d \mathbf{r}''\\
+  \frac{\delta n_2^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R)\\
+ \end{align}
+
+ \fixme {
+And then there's a lot with taking the derivative of the delta for the case of $n_{2}^{S}(\mathbf{r}')$.
+Not sure if we'll use any of this, but wrote down some equations that Roundy wrote out.
+So can use the terms at least if going to write about this - I'm just copying it in for now (probably a lotof mistakes)
+
+\begin{align}
+  p(\mathbf{r}) &= \frac{\delta A}{\delta n_2(\mathbf{r}')} \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}'\\
+  &= \frac{\delta A}{\delta n_2(\mathbf{r})} - \int n(\mathbf{r}'') \delta' (|\mathbf{r} - \mathbf{r}''| - R) d\mathbf{r}''
+  \int n (\mathbf{r} +\mathbf{r}'') \delta'(|\mathbf{r}'''| - R) d \mathbf{r}'''\\
+  &= \int n(\mathbf{r} + \mathbf{r}''') \delta'(|\mathbf{r}'''| - R) r'''^2 cos\theta d \theta''' dr'''\\
+  du = \delta'(\mathbf{r}''' - R) d \mathbf{r}'''     v = n(\mathbf{r} - \mathbf{r}'')r'''\\
+   u = \delta(\mathbf{r}''' - R)                       
+   dv = \mathbf{\nabla} n(\mathbf{r} + \mathbf{r}'') \cdot \frac{\mathbf{r}'''}{r'''} + 2n(\mathbf{r}' - \mathbf{r}'')r'''\\
+  &= -\int(\mathbf{\nabla}n(\mathbf{r}' + \mathbf{r}'') \cdot \mathbf r'''(hat)r'''^2 dr''' 
+  + 2n(\mathbf{r} + \mathbf{r}''')r''') dr''' \delta(\mathbf{r} - R) dcos\theta''' d\theta''\\
+  &= - \int \delta(\mathbf{r}''' - R) d\mathbf{r}''' 
+  (\mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) + \frac{2n(\mathbf{r} + \mathbf{r}'')}{\mathbf{r}'''})\\
+  &= \frac{2n_2(\mathbf{r})}{R} + 
+  \int \delta(|\mathbf{r}'''| - R) \mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) d \mathbf{r}'''\\
+\end{align} 
+}
+********************************************************************************\\
+
 These give us some basic tools with which to perform derivatives.

### 2012-03-05 src/ContactDensity.cpp
Fixed bugs to WBm2 derivs in paper and code Tested everything on crystal.

@@ -172,14 +172,14 @@ Functional dWBm2_dn2(double radius) {
   Functional n2z = zShellConvolve(radius);
   return (Functional(1) + (1/9.0)*sqr(n3(radius))*phi2(n3(radius)))*n1(radius) / (1-n3(radius)) +
     (1 - (4.0/9.0)*n3(radius)*phi3(n3(radius)))*
-    (n2(radius) - (sqr(n2x) + sqr(n2y) + sqr(n2z))) / (8*M_PI*sqr(1-n3(radius)));
+    (sqr(n2(radius)) - (sqr(n2x) + sqr(n2y) + sqr(n2z))) / (8*M_PI*sqr(1-n3(radius)));
 }
 
 Functional dWBm2_dn2v_over_n2v(double radius) {
   Functional R(radius, "R");
-  return -(Functional(1) + (1.0/9.0)*sqr(n3(radius))*phi2(n3(radius))) / (4*M_PI*R*(1-n3(radius))) -
-    (1 - (4.0/9.0)*n3(radius)*phi3(n3(radius)))*
-    Functional(6) / (24*M_PI*sqr(1-n3(radius)));
+  return -(Functional(1) + (1.0/9.0)*sqr(n3(radius))*phi2(n3(radius))) / (4*M_PI*R*(1-n3(radius))) 
+    -(1 - (4.0/9.0)*n3(radius)*phi3(n3(radius)))*
+    Functional(6)*n2(radius) / (24*M_PI*sqr(1-n3(radius)));
 }
 
 Functional dWBm2_dn3(double radius) {

### 2012-03-06 papers/contact/paper.tex
Fixed bugs to WBm2 derivs in paper and code Tested everything on crystal.
clean up bits of contact paper Tested everything on bingley.

@@
 single variable, with the contact value being when that distance is
-$2R$.  It is desirable for reasons of efficiency to limit CDFT
+the diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
 functionals to one-center convolutions, which leads us to seek a
@@
 correlation function for hard spheres.
-%
-\fixme{The following bit is awkward.}
-%
 In a system with an inhomogeneous density, we seek a \emph{local}
-contact value of the correlation function.  There are two reasonable
-options for defining such a local function:  a symmetric (which we refer to as S) and an
-asymmetric (which we refer to as A) formulation.
-
-For the S case:
-\begin{align}
-  g^S_\sigma(\rr) &= \frac{n^S_\textit{contact}(\rr)}{n_S(\rr)}
-\end{align}
-\begin{align}\label{eq:n-contactS}
-  n^S_\textit{contact}(\rr) &= \frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
-  R)d\rr'
-\end{align}
-\begin{align}
-  n_S(\rr) &= \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
+value for $g_\sigma$.  There are two reasonable options for defining
+such a local function: a symmetric formulation (which we refer to as $S$) and an
+asymmetric formulation (which we refer to as $A$). \fixme{Should we
+  introduce a third one, for Yu and Wu?}
+
+For the symmetric $S$ case:
+\begin{align}
+  g^S_\sigma(\rr) &= \frac{1}{n_S(\rr)^2}\int n^{(2)}(\rr - \rr', \rr
+  + \rr')
+  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \\
+  n_S(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
 \end{align}
@@
 
-And for the A case:
-\begin{align}
-  g^A_\sigma(\rr) &= \frac{n^A_\textit{contact}(\rr)}{n_A(\rr)} \\
-  n^A_\textit{contact}(\rr) &= \frac1{16\pi R^2}\int n^{(2)}(\rr, \rr + \rr')
-  \delta(|\rr'| - 2R)d\rr' \\
-  n_A(\rr) &= \frac1{16\pi R^2}\int n(\rr') \delta(|\rr-\rr'|-2R) d\rr'
-\end{align}
-In this case, we find the average number of contacts for a sphere at a
-given position.
+And for the asymmetric $A$ case:
+\begin{align}
+  g^A_\sigma(\rr) &= \frac{1}{n(\rr)n_A(\rr)}
+  \int n^{(2)}(\rr, \rr + \rr')
+  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \\
+  n_A(\rr) &= \int n(\rr')
+  \frac{\delta(|\rr-\rr'|-\sigma)}{4\pi\sigma^2} d\rr'
+\end{align}
+In this case, the two-particle density is averaged over spheres
+touching a sphere that is aat the given position $\rr$.
 
 In the process of defining these two averaged correlation functions,
-we also defined four averaged densities, two \emph{contact
-  densities}, $n^S_\textit{contact}(\rr)$ and $n^A_\textit{contact}(\rr)$,
-and two \emph{contacting densities}, $n_S(\rr)$ and $n_A(\rr)$.  The
-former are the density of spheres that are in contact with each other,
-while the latter are the density of spheres available to be in
-contact.  In the homogeneous limit, $n_S = n_A = n$ and
-$n^S_\textit{contact} = n^A_\textit{contact} = g(\sigma)n$.  It is an
+we also defined two averaged densities, $n_S(\rr)$ and $n_A(\rr)$,
+which are the density of spheres available to be in
+contact.  In the homogeneous limit, $n_S = n_A = n$.  It is an
 open question, which of these averages will be more useful in any
@@
     -n_0 \ln\left(1 - n_3\right)
-    + (1 + \frac{1}{9}n_3^2 \phi_2(n_3))   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+    + \left(1 + \frac{1}{9}n_3^2 \phi_2(n_3)\right)   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
     + (1 - \frac{4}{9}n_3\phi_3(n_3))\frac{(n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{24\pi (1-n_3)^2}
@@
 \begin{align}
-  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2(n_3))\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)(\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2})\\
+  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)(\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2})\\
   \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
@@
   \frac{\delta A_{HS}}{\delta \mathbf{n}_2} &= -(1 + \frac{1}{9}n_3^2\phi_2)\frac{\mathbf{n}_1}{1 - n_3} 
-  - (1 - \frac{4}{9}n_3\phi_3)\frac{6\mathbf{n}_2}{24\pi(1 - n_3)^2}
+  - (1 - \frac{4}{9}n_3\phi_3)\frac{6n_2\mathbf{n}_2}{24\pi(1 - n_3)^2}
 \end{align}
@@
 \begin{align}
-  n_\textit{contact} &= n_b \frac{1 - \frac12 n_{3'}}{\left(1 -
+  n_\textit{contact} &= n_{A} \frac{1 - \frac12 n_{3'}}{\left(1 -
     n_{3'}\right)^3} \\
@@
 \end{align}
-where $n_b$ is defined in the introduction and $n_{3'}$ is an averaged
+where $n_A$ is defined in the introduction and $n_{3'}$ is an averaged
 filling fraction defined in a manner analogous to the $n_3$ of FMT,
@@
 compare both their functional as well as ours with simulation data.
+
+The functional of Yu and Wu for $g_\sigma$ is a symmetric formulation,
+but instead of using $n_0$ as the associated density, it uses a
+density given by
+\begin{align}
+  n_{YW} &= n_0 \zeta \\
+  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
+\end{align}
+where the function $\zeta$ is a measure of local inhomogeneity at the
+point of contact.  Given this definition of density, $g_\sigma$ is
+given by:
+\begin{align}
+  g_\sigma^{YW} &= \frac{1}{1-n_3}
+    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
+    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
+    \\
+  g_\sigma^{YW} &= \frac{1}{1-n_3}
+    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
+    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
+\end{align}
+We should write a bit more about this functional.  I think we can view
+it as a third alternative, which we could also predict:
+\begin{align}
+  g_\sigma^{S'}(\rr) &= \frac1{n_{S'}(\rr)^2}\frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
+  R)d\rr'
+\end{align}
+\begin{align}
+  n_{S'}(\rr) &= \zeta(\rr) \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
+\end{align}
+This is closely related to $g_\sigma^{S}$, but is ever so slightly
+different.  In order to be completely fair, we should see about
+computing this from our MC calculations.  Unfortunately, it will
+require that we compute $\mathbf{n}_2 \cdot \mathbf{n}_2$ which will
+be a bit of a pain.  $\ddot\frown$
 

### 2012-03-06 papers/contact/paper.tex
update text and implementation on Yu and Wu's ghs Tested everything on bingley.

@@
 
-\section{Contact density at a given hard
-  sphere}\label{contact-at-sphere}
+\section{Asymmetrically averaged correlation function}\label{contact-at-sphere}
 
@@
 
+\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+We compare with the averaged correlation function of
+Gross\cite{gross2009density}, which is of the asymmetrical variety
+($g_\sigma^A$), where the correlation function is averaged at location
+of one of the spheres in concact.
+\begin{align}
+  g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
+    n_{3'}\right)^3} \\
+  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
+\end{align}
+where $n_{3'}$ is an averaged
+filling fraction defined in a manner analogous to the $n_3$ of FMT,
+but with the hard-sphere radius replaced by the hard-sphere diameter,
+so the integrals range over spheres that are twice as large.
+
 \begin{figure}
@@
 
-\section{Contact density at a given contact position}
-
-An alternate definition of the inhomogeneous contact density would be
-the contact density for spheres \emph{touching at a given point}.
-This would be a symmetric measure, since both spheres involved are
-treated in the same way.  In fact, the density of such spheres is just
-the density $n_0(\xx)$ from FMT.  See Figure~\ref{fig:n0} for an
-illustration $n_0$.  Since $n_0$ is the fundamental measure of sphere
-surface contact, it would make sense to define a contact density that
-would correspond.  This is what is done by Yu and
+\section{Symmetrically averaged correlation function}
+
+An alternate definition of the averaged correlation function for an
+inhomogeneous system would be the correlation function averaged for
+spheres \emph{touching at a given point}.  This would be a symmetric
+measure, since both spheres involved are treated in the same way.  In
+fact, the density of such spheres is just the density $n_0(\xx)$ from
+FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
+is the fundamental measure of sphere surface contact, it makes sense
+to define a corresponding correlation function $g_\sigma^S$.  This is
+\emph{almost} what is done by Yu and
 Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as discussed in
-Section~\ref{sec:yuwu}.
-
-\section{Gross' functional}\label{sec:gross}
-We compare with the functional of Gross \emph{et
-  al}\cite{gross2009density}, which is of the second kind ($g^b$), where the
-contact density is evaluated at the sphere location.
-\begin{align}
-  n_\textit{contact} &= n_{A} \frac{1 - \frac12 n_{3'}}{\left(1 -
-    n_{3'}\right)^3} \\
-  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
-\end{align}
-where $n_A$ is defined in the introduction and $n_{3'}$ is an averaged
-filling fraction defined in a manner analogous to the $n_3$ of FMT,
-but with the hard-sphere radius replaced by the hard-sphere diameter,
-so the integrals range over spheres that are twice as large.
+Section~\ref{sec:yuwu}, although they use a modified density $\zeta
+n_0$ to account for density gradients.
 
@@
 density at a given contact
-position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We will
-compare both their functional as well as ours with simulation data.
-
-The functional of Yu and Wu for $g_\sigma$ is a symmetric formulation,
-but instead of using $n_0$ as the associated density, it uses a
-density given by
-\begin{align}
-  n_{YW} &= n_0 \zeta \\
+position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We
+will compare both their functional as well as ours with simulation
+data.  Yu and Wu use a symmetric formulation for $g_\sigma$, but
+instead of using $n_0$ as the associated density, it uses a density
+given by
+\begin{align}
+  n^\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
   \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
@@
 where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact.  Given this definition of density, $g_\sigma$ is
+point of contact.  Given this definition of density, their $g_\sigma$ is
 given by:
 \begin{align}
-  g_\sigma^{YW} &= \frac{1}{1-n_3}
+  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
     + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
@@
     \\
-  g_\sigma^{YW} &= \frac{1}{1-n_3}
+  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
     + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
     + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
-\end{align}
-We should write a bit more about this functional.  I think we can view
-it as a third alternative, which we could also predict:
-\begin{align}
-  g_\sigma^{S'}(\rr) &= \frac1{n_{S'}(\rr)^2}\frac1{4\pi R^2}\int n^{(2)}(\rr - \rr', \rr + \rr')\delta(|\rr'| -
-  R)d\rr'
-\end{align}
-\begin{align}
-  n_{S'}(\rr) &= \zeta(\rr) \frac{1}{4\pi R^2}\int n(\rr')\delta(|\rr-\rr'|-R) d\rr'
-\end{align}
-This is closely related to $g_\sigma^{S}$, but is ever so slightly
-different.  In order to be completely fair, we should see about
-computing this from our MC calculations.  Unfortunately, it will
-require that we compute $\mathbf{n}_2 \cdot \mathbf{n}_2$ which will
-be a bit of a pain.  $\ddot\frown$
+    \\
+  g_\sigma^\text{Yu,S} &= \frac{\frac{1}{1-n_3}
+    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
+    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}}{\zeta^2}
+\end{align}
+I think the adjusted functional $g_\sigma^{\text{Yu,}S}$ can be directly
+compared with our $g_\sigma^S$.  It is implemented in our code.
 

### 2012-03-07 papers/contact/paper.tex
a few small changes to paper Tested everything on crystal.
various improvements to contact paper Tested everything on bingley.

@@
 \newcommand{\rr}{\textbf{r}}
-\newcommand{\xx}{\textbf{x}}
 \newcommand{\refnote}{\red{[ref]}}
@@
 
-\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
-%\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
+%\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
+\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
 
@@
 that can handle sharply varying densities is the correlation function
-evaluated at contact. \fixme{I need to figure out how best to phrase
-  this... do we want to compute correlation at contact, or contact
-  density itself? I prefer the latter, but the former is what is
-  commonly included in formulas.}
+evaluated at contact.
 
@@
 to establish the advantages and disadvantages of each approach.
-\fixme{
-
-Jeff's Notes:
-A novice (me) is confused by the use of 'contact value of correlation function' throuhout this introduction while throughout the rest of the paper we refer to the 'contact density'.
-Pehaps the audince will understand that they're interchangable (if they are) but this was confusing to me.
-Also, I would think I'd like to see something like:
-
-We first discuss two possibilities for the definition of contact density.
-The first involves a more symmetric interpretation and so we call it the 'S case'.
-The second involves a sphere-centered, assymetrical interpretation and so we call it the 'A case'.
-We present where in the SAFT free energy formulation these contact densities will contribute.
-We next present a thermodynamic method that we will use to calculate the above contact densities.
-This method involves the use of surface pressure and the minimization of the excess free energy refered to above.
-We use the White Bear and White Bear Mark II formulations of these free energy terms and take neccesary derivitives in the concpetual spirit of the seperately defined S an A case contact densities.
-Finally we compare the resulting contact densities with those of Gross, those of Yu and Wu, and as well with our wn monte carlo simulations.}
-
-\fixme{Should we include these equations? Maybe just later?}
-\begin{align}
-  \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(g_\sigma\right)-1\right)
-\end{align}
-\begin{align}
-  \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
-  X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
-\end{align}
-
-\section{Contact density and correlation function}
-
-\fixme{We should decide on good names for these two densities.  In
-  this section I define them formally, and we should keep using the
-  same names (and mathematical symbols) for them throughout.}
+
+We first discuss two possibilities for the definition of contact
+density.  The first involves a more symmetric interpretation and so we
+call it the '$S$ case'.  The second involves a sphere-centered,
+assymetrical interpretation and so we call it the '$A$ case'.  We
+present where in the SAFT free energy formulation these contact
+densities will contribute.  We next present a thermodynamic method
+that we will use to calculate the above contact densities.  This
+method involves the use of surface pressure and the minimization of
+the excess free energy refered to above.  We use the White Bear and
+White Bear Mark II formulations of these free energy terms and take
+neccesary derivitives in the concpetual spirit of the seperately
+defined $S$ an $A$ case contact densities.  Finally we compare the
+resulting contact densities with those of Gross, those of Yu and Wu,
+and as well with our own Monte Carlo simulations.
+
+\section{Correlation function with inhomogeneity}
 
@@
 such a local function: a symmetric formulation (which we refer to as $S$) and an
-asymmetric formulation (which we refer to as $A$). \fixme{Should we
-  introduce a third one, for Yu and Wu?}
+asymmetric formulation (which we refer to as $A$).
 
@@
 \begin{align}
-  g^S_\sigma(\rr) &= \frac{1}{n_S(\rr)^2}\int n^{(2)}(\rr - \rr', \rr
+  g^S_\sigma(\rr) &= \frac{1}{n_0(\rr)^2}\int n^{(2)}(\rr - \rr', \rr
   + \rr')
   \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \\
-  n_S(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
+  n_0(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
 \end{align}
 This functional gives a value averaged over spheres that touch at the
-position $\rr$.  The density $n_S$ is the same as the density $n_0$
-defined in Fundamental Measure Theory (FMT), and seems ideal for
-treating touching spheres.
+position $\rr$.  The density $n_0$ is one of the fundamental measures
+of Fundamental Measure Theory (FMT), and seems ideal for treating
+touching spheres.
 
@@
 In the process of defining these two averaged correlation functions,
-we also defined two averaged densities, $n_S(\rr)$ and $n_A(\rr)$,
+we also defined two averaged densities, $n_0(\rr)$ and $n_A(\rr)$,
 which are the density of spheres available to be in
-contact.  In the homogeneous limit, $n_S = n_A = n$.  It is an
+contact.  In the homogeneous limit, $n_0 = n_A = n$.  It is an
 open question, which of these averages will be more useful in any
@@
 Yu and Wu introduce an approximation for
-$g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which will be
-discussed in Section~\ref{sec:yuwu}.  Gross introduced an
-approximation for $g^A$\cite{gross2009density}, which will be
-introduced in Section~\ref{sec:gross}.  We will also introduce our own
-approximation to $g^A$ which is based on the assumption of
-thermodynamic consistency within FMT, in
-Section~\ref{contact-at-sphere}.  We will report on the accuracy of
+$g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which we will
+discus in Section~\ref{sec:yuwu}.  Gross introduced an
+approximation for $g^A$\cite{gross2009density}, which we will
+introduce in Section~\ref{sec:gross}.  We will also introduce our own
+approximations for both $g^A$ and $g^S$ which are based on the
+assumption of thermodynamic consistency within FMT, in
+Sections~\ref{sec:g-S} and~\ref{sec:g-A}.  We will report on the accuracy of
 these approximations by comparing with Monte Carlo simulations of the
@@
 \subsection{Association free energy of SAFT}
+
+\fixme{Integrate the following equations into this section...}
+\begin{align}
+  \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(n g_\sigma \right)-1\right)
+\end{align}
+\begin{align}
+  \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
+  X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
+\end{align}
+
 \newcommand\epsilonassoc{\ensuremath{\varepsilon_\textit{AB}}}
@@
 the pressure on the spheres, we just need to divide by the total area,
-which means dividing by $N 4\pi (2R)^2$, since the relevant area
+which means dividing by $N 4\pi \sigma^2$, since the relevant area
 is the area over which the molecules can make contact.
 \begin{align}
-  p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
+  p_{HS} &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
   \derivation{
@@
 \begin{equation}
-A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)\right) d\xx \; ,
+A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
 \end{equation}
@@
 \begin{align}
-  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
-  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
+  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - R) d\rr' \\
+  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - R) d\rr'
 \end{align}
@@
 %% \begin{align}
-%%   -c^{(2)}(\xx_1, \xx_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\xx_1) \delta n(\xx_2)}
+%%   -c^{(2)}(\rr_1, \rr_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\rr_1) \delta n(\rr_2)}
 %% \end{align}
@@
 %% \begin{align}
-%%   h(\xx_1, \xx'_2) &= g(\xx_1, \xx_2) - 1 \\
-%%   h(\xx_1, \xx_2) &= c(\xx_1, \xx_2)
-%%     + \int n(\xx_3)c(\xx_1, \xx_3) h(\xx_2, \xx_3)d\xx''
+%%   h(\rr_1, \rr'_2) &= g(\rr_1, \rr_2) - 1 \\
+%%   h(\rr_1, \rr_2) &= c(\rr_1, \rr_2)
+%%     + \int n(\rr_3)c(\rr_1, \rr_3) h(\rr_2, \rr_3)d\rr''
 %% \end{align}
@@
 %% \begin{align}
-%%   p_{HS} &= \frac{1}{N 4\pi (2R)^2} \frac{dA_{HS}}{dR} \\
-%%   \ncontact &= \frac{1}{N k_BT 4\pi (2R)^2} \frac{dA_{HS}}{dR}
+%%   p_{HS} &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
+%%   \ncontact &= \frac{1}{N k_BT 4\pi \sigma^2} \frac{dA_{HS}}{dR}
 %% \end{align}
@@
 
-\section{Asymmetrically averaged correlation function}\label{contact-at-sphere}
+\section{Asymmetrically averaged correlation function}\label{sec:g-A}
 
@@
 \begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{4\pi (2R)^2} \frac{\delta
+  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2} \frac{\delta
     A_{HS}}{\delta R(\mathbf{r})} \\
-  \ncontact(\xx) &= \frac{1}{k_BT 4\pi (2R)^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})}
-\end{align}
-After considerable math to do the above functional derivative, we find
-a functional giving the contact density for spheres at any given
-position.  This functional unavoidably involves more convolutions than
-are present in the free energy itself, which makes it relatively
-expensive.
+  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \\
+  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
+  g_\sigma^A(\rr) &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
+\end{align}
+Equation~\ref{eq:g-A-exact} is an exact expression for $g_\sigma^A$
+given the exact functional $A_{HS}$.  In practice, we use the
+approximate White Bear functional, and find an approximate expression
+for $g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
+expression will involve more convolutions than are present in the free
+energy itself, making this formulation relatively expensive.
 
@@
 
-\section{Symmetrically averaged correlation function}
+\section{Symmetrically averaged correlation function}\label{sec:g-S}
 
@@
 measure, since both spheres involved are treated in the same way.  In
-fact, the density of such spheres is just the density $n_0(\xx)$ from
+fact, the density of such spheres is just the density $n_0(\rr)$ from
 FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
@@
 \emph{almost} what is done by Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as discussed in
-Section~\ref{sec:yuwu}, although they use a modified density $\zeta
-n_0$ to account for density gradients.
+
+\fixme{Add qualitative description of the derivation needed to get
+  $g_\sigma^S$.}
+
+Complete expressions for $g_\sigma^S$ are included in
+Appendix~\ref{appendix:g-S}.
 
@@
 \begin{align}
-  n^\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
+  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
   \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
@@
     \\
-  g_\sigma^\text{Yu,S} &= \frac{\frac{1}{1-n_3}
+  g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
     + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
-    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}}{\zeta^2}
+    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}\right)
 \end{align}
@@
 of Gross~\cite{gross2009density}, and the ``contact density at
-sphere'' from Section~\ref{contact-at-sphere}.  In principle, the
+sphere'' from Section~\ref{sec:g-A}.  In principle, the
 functional of Yu (FIXME) should match the CenConDensity, while the DFT
@@
 
+\derivation{
 \begin{widetext}
@@
 \end{widetext}
+}
+
+\section{Expression for asymmetrically averaged correlation
+  function}\label{appendix:g-A}
+
+\section{Expression for symmetrically averaged correlation
+  function}\label{appendix:g-S}
+
+\section{Expression for asymmetrically averaged correlation
+  function based on White Brear Mark II free energy}\label{appendix:g-A-ii}
+
+\section{Expression for symmetrically averaged correlation
+  function based on White Brear Mark II free energy}\label{appendix:g-S-ii}
 

### 2012-03-07 paper/paper.tex
add march meeting talk Tested everything on bingley.

@@
 to liquid water between the rods. The radius is 0.5~nm, the top figure is 
-at a separation of 0.2~nm and the
+at a separation of 0.5~nm and the
 bottom is 0.6~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
@@
 
-We extend the study to four rods in water using the same radii as the two rods and
-separation distance $d = 0$~nm to $d = 2$~nm for nearest neighbor rods. These exhibit
-a similar behavior to the two rods where for small $d$ only vapor is present in the 
-middle of the rods until a critical separation between where water occupies the middle.
-Before the transition, the four rods and vapor form a square with rounded courners 
-where each rod forms a corner (shown in Fig. \ref{fig:density-four-rods} as 'before').
-For large $d$, the rods behave like separate single rods.  Free energy
+We simulate four hydrophobic rods in water using the same radii as the two rods and
+separation distances $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
+the rods in a 'square' arrangement and observed a similar behavior to the two rods
+where vapor is present in between the rods(shown in Fig. \ref{fig:density-four-rods}
+as 'before') until a critical separation where it becomes filled with water
+(shown in Fig. \ref{fig:density-four-rods} as 'after'). For large $d$,
+the rods behave like separate single rods.  Free energy per length
 is plotted against rod separation in Fig. \ref{fig:four-rods-energy-vs-distance}.
@@
 
-The rounded square's circumference is estimated by
+The rounded square formed by the rods has a circumference estimated by
 \begin{equation}
@@
 \end{equation}
-
-\begin{table}
-\begin{tabular} {|c|c|c|c|}
-\hline
-$r$ (nm) & $d_{rs}$ prediction (nm) & $d_{c4}$ result (nm) & \% error \\
-\hline
-0.3 & 0.81 & 0.40 & 51\% \\
-\hline
-0.5 & 1.36 & 1.15 & 15\% \\
-\hline
-0.7 & 1.90 & 1.70 & 11\% \\
-\hline
-0.9 & 2.44 & 2.30 & 06\% \\
-\hline
-1.0 & 2.70 & 0.00 & 00\% \\
-\hline
-\end{tabular}
-\caption{Calculations of the critical separation (see Equation~\ref{criticalfourrods})
-for the transition
-from liquid to vapor between four hydrophobic rods.}
-\label{criticalfourrodstable}
-\end{table}
-
-The predicted and computed valued for $d_{c4}$ are show in Table \ref{criticalfourrodstable}. 
-[talk about table here]
 
@@
 Figure~\ref{fig:density-sphere} shows the density profile for several
-hard sphere radii, plotted together with with the results of the same
+hard sphere radii, plotted together with the results of the same
 SPC/E molecular dynamics simulation shown in

### 2012-03-15 paper/paper.tex

@@
 vaporization, thermal conductivity and surface tension of any
-non-metallic liquid.  Water has an unusually high dielectric constant.
+non-metallic liquid;  water also has an unusually high dielectric constant.
 Much of the unusual behavior of water is explained by the ability of
@@
 continuum dielectric with a cavity surrounding each solute can give
-accurate predictions of the energy of solvation of ions~\cite{latimer1939,
+accurate predictions for the energy of solvation of ions~\cite{latimer1939,
 rashin1985, zhan1998, hsu1999, hildebrandt2004, hildebrandt2007}, but
@@
 
-The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+The SAFT-VR dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
 \begin{align}
@@
 where $a_1$ and $a_2$ are the first two terms in a high-temperature
-perturbation expansion and $\beta=1/k_BT$.  The first term $a_1$ is 
-the mean-field dispersion
-interaction, which reduces in the homogeneous limit to
-\begin{align}\label{eq:A1-simple}
-  a_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
-  g_{HS}(\left|\xx\right|) d\xx
-\end{align}
-where $n_b$ is the bulk density and $g_{HS}$ is the homogeneous
-hard-sphere fluid correlation function.
-The second dispersion term in the free energy $a_2$ describes the
-effect of fluctuations resulting from the compression of the fluid due
-to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA). In the LCA,
-we assume the energy fluctuation is directly related to the
+perturbation expansion and $\beta=1/k_BT$.  The first term ,$a_1$, is 
+the mean-field dispersion interaction and the second term, $a_2$, describes the
+effect of fluctuations resulting from compression of the fluid due
+to the dispersion interaction itself. $a_2$ is approximated
+using the local compressibility approximation (LCA), which
+assumes the energy fluctuation is directly related to the
 compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
 
-We use a free square-well function for the dispersion $\varphi$, which
-is the choice used by Clark \emph{et al}~\cite{clark2006developing},
-and allows a reasonable fit to the equation of state.  Thus our model
-interaction has the form:
-\begin{equation}
-  \varphi(r) = \Theta(r-2 \lambdadisp R)
-\end{equation}
-where $\Theta$ is the Heaviside step function.
-
-The form of $a_1$ is given in
-reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
+The form of $a_1$ and $a_2$ for SAFT-VR is given in
+reference~\cite{gil-villegas-1997-SAFT-VR} but is expressed in terms
 of the filling fraction.  In order to apply this form to an
@@
 
-The first term in the dispersion functional $a_1$ when written using
-the above filling fraction for dispersion has the form
-\begin{align}
-  a_1 &= 
-   -4(\lambdadisp^3 - 1)\epsilondisp \etadisp(\xx)
-    g^{HS}_\sigma(\eta_\textit{eff}(\xx)) \\
-  g_\sigma^{HS}(\eta) &= \frac{1 - \frac12 \eta}{(1 - \eta)^3} \label{eq:ghs}
-  \\
-  g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
-  +\frac32\frac{1}{(1-\eta)^2}
-  + \frac12\frac{\eta^2}{(1-\eta)^3}
-  \\
-  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
-  \lambdadisp^j
-\end{align}
-where $g_\sigma^{HS}$
-is the Carnahan-Starling value for the hard-sphere fluid correlation
-function evaluated at contact.  \textcolor{red}{Jess, could you verify
-that the two formulas for gHS above are equivalent? Thanks!}
-The $C_{ij}$ values
- are numerical constants taken from
-reference\cite{gil-villegas-1997-SAFT-VR}, which come from a numerical
-fit to the integral in Equation~\ref{eq:A1-simple} over a range of
-values for filling fraction and $\lambdadisp$.
-
 The second term, $a_2$, which describes the contribution to the free
@@
 interaction occurs is $\kappaassoc$, giving the association
-term in the free energy has two empirical parameters that are fit to
+term in the free energy two empirical parameters that are fit to
 experimental data.
@@
   the same functional, but their paper contains errors in the
-  association term and is not useful as a reference.}  The association
-functional of Yu and Wu includes the effects of density
-inhomogeneities on the \emph{contact density} of hard spheres, which
-conventionally appears in the form of the density multiplied by the
-contact value of the correlation function $g^{HS}_\sigma$, but is
+  association term and is not useful as a reference.} 
+which includes the effects of density inhomogeneities in the
+contact value of the correlation function $g^{HS}_\sigma$. This, however, is
 based on the SAFT-HS equation of state, rather than the SAFT-VR
@@
 \end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
+\caption{Comparison of Surface tension versus temperature for theoretical and
+  experimental data. The experimental data is taken from NIST.\cite{nistwater}
   The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
@@
 dispersion free energy.  We use this final parameter to fit the
-surface tension, with the result shown in
+surface tension with the result shown in
 Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
@@
 point---we cannot correctly describe the surface tension at all
-temperatures, and chose to fit it for the temperature range at which
+temperatures, and choose to fit it for the temperature range at which
 water is liquid at one atmosphere of pressure.

### 2012-03-28 paper/paper.tex
Changes in paper Tested everything on quipu.

@@
 where $a_1$ and $a_2$ are the first two terms in a high-temperature
-perturbation expansion and $\beta=1/k_BT$.  The first term ,$a_1$, is 
+perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
 the mean-field dispersion interaction and the second term, $a_2$, describes the
@@
 the length scale over which the dispersion interaction is correlated.
-
-The second term, $a_2$, which describes the contribution to the free
-energy associated with fluctuations is given by
-\begin{align}
-  a_2 &= \frac12 \epsilondisp
-              K_{HS}(\etadisp(\xx)) \etadisp(\xx)
-              \frac{\partial a_1}{\partial \etadisp(\xx)}
-\end{align}
-where $K_{HS}$ is the Percus-Yevick hard-sphere compressibility, given
-by
-\begin{align}
-  K_{HS} &=
-    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}.
-\end{align}
 
@@
 thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
+$\Omega$, which we ordinarily use in our computations, is obtained by
 simply adding a chemical potential term:
 \begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
+  \Omega[n] = F[n] + \mu \int n(\xx) d\xx.
 \end{equation}
@@
 \end{equation}
-The predicted and actual computed values for $d_c$ are shown in 
-Table~\ref{criticalseparationtable}, and they match fairly well for the 
-0.5~-~1.0~nm rods. We hypothesize that larger radii rods
+The predicted and actual computed values for $d_c$ are shown graphically in 
+Figure~\ref{fig:rods-energy-vs-distance}, and they match fairly well for the 
+0.5~-~1.0~nm rods with differences between 1.3~-~5.5\%. We hypothesize that larger radii rods
 would also follow this trend, but smaller rods are likely to have a different
 relationship with the surrounding water molecules. This may explain why the critical
-separation result for a radius of 0.3~nm does not match the prediction well.
+separation result for a radius of 0.3~nm does not match the prediction
+well(32\% difference).
 
@@
 
-\begin{table}
-\begin{tabular} {|c|c|c|c|}
-\hline
-$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
-\hline
-0.3 & 0.34 & 0.23 & 32\% \\
-\hline
-0.5 & 0.57 & 0.54 & 5.3\% \\
-\hline
-0.7 & 0.80 & 0.81 & 1.3\% \\
-\hline
-0.9 & 1.03 & 1.06 & 2.9\% \\
-\hline
-1.0 & 1.14 & 1.19 & 4.4\% \\
-\hline
-\end{tabular}
-\caption{Calculations of the critical separation 
-(see Equation~\ref{criticalseparation})
-for the transition
-from liquid to vapor between two hydrophobic rods.\needsworknow{Delete
-this table and move some of the information into the text along with
-prediction model and explanation of Figure~\ref{fig:rods-energy-vs-distance}}.}
-\label{criticalseparationtable}
-\end{table}
-
 \subsection{Hydrophobic interactions of four rods}
@@
 
-We simulate four hydrophobic rods in water using the same radii as the two rods and
-separation distances $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
-the rods in a 'square' arrangement and observed a similar behavior to the two rods
-where vapor is present in between the rods(shown in Fig. \ref{fig:density-four-rods}
-as 'before') until a critical separation where it becomes filled with water
-(shown in Fig. \ref{fig:density-four-rods} as 'after'). For large $d$,
+We simulate four hydrophobic rods in water using the same radii as our
+two rods ($r = 0.3$~nm~-~$1.0$~nm)and
+separation distances between $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
+the rods in a 'square' arrangement and observed a similar behavior as the two rods
+where vapor is present in the space between the rods until a critical separation
+where it becomes filled with water (shown in Fig. \ref{fig:density-four-rods}). 
+For large $d$,
 the rods behave like separate single rods.  Free energy per length
-is plotted against rod separation in Fig. \ref{fig:four-rods-energy-vs-distance}.
+is plotted against rod separation in Figure \ref{fig:four-rods-energy-vs-distance}.
 From the slope we calculate the force per length to be [number].
 
-The rounded square formed by the rods has a circumference estimated by
+The rounded square formed by the area of the four rods and the space
+inbetween them has a circumference estimated by
 \begin{equation}
@@
 \end{equation}
-We again equate this circumference with the total circumference of the four rods, which
-results in a critical separation
+We again equate this circumference with the total circumference of
+only the four rods, which results in a critical separation of
 \begin{equation}
-d_{c4} = \frac{3 \pi - 4}{2} r \label{criticalfourrods}
-\end{equation}
+d_{c4} = \frac{3 \pi - 4}{2} r. \label{criticalfourrods}
+\end{equation}
+We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 3.0$~nm
+rod has the largest difference between simulation and prediction with
+a 23\% difference, and better agreement for larger radii (between 0.5\% and 5.6\%).
 
@@
 contact value theorem cannot agree, since the free energies do not
-agree.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/sphere-density}
-\end{center}
-\caption{ Density profiles illustrating the transition from vapor 
-to liquid water between the rods. The radius is 0.5~nm, the top figure is 
-at a separation of 0.2~nm and the
-bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
-the energy for these and other separations.}
-\label{fig:sphere-density}
-\end{figure}
+agree.c
 

### 2012-03-28 paper/figs/sphere.cpp

@@ -58,28 +58,6 @@ void plot_grids_y_direction(const char *fname, const Grid &a) {
   fclose(out);
 }
 
-void plot_grids_yz_directions(const char *fname, const Grid &a) {
-  FILE *out = fopen(fname, "w");
-  if (!out) {
-    fprintf(stderr, "Unable to create file %s!\n", fname);
-    // don't just abort?
-    return;
-  }
-  const GridDescription gd = a.description();
-  const int x = 0;
-  const int stepsize = 2;
-  //const int y = gd.Ny/2;
-  for (int y=-gd.Ny/2; y<=gd.Ny/2; y+=stepsize) {
-    for (int z=-gd.Nz/2; z<=gd.Nz/2; z+=stepsize) {
-      Cartesian here = gd.fineLat.toCartesian(Relative(x,y,z));
-      double ahere = a(here);
-      fprintf(out, "%g\t%g\t%g\t%g\n", here[0], here[1], here[2], ahere);
-    }
-    fprintf(out,"\n");
- }  
-  fclose(out);
-}
-
 int main(int argc, char *argv[]) {
   clock_t start_time = clock();
   if (argc > 1) {
@@ -205,9 +183,6 @@ int main(int argc, char *argv[]) {
     Grid density(gd, EffectivePotentialToDensity()(water_prop.kT, gd, potential));
 
     char *plotname = (char *)malloc(1024);
-
-    sprintf(plotname, "paper/figs/sphere-%04.2f-slice.dat", diameter/nm);
-    plot_grids_yz_directions(plotname, density);
 
     sprintf(plotname, "paper/figs/sphere-%04.2f.dat", diameter/nm);
     plot_grids_y_direction(plotname, density);

### 2012-04-04 papers/contact/paper.tex

@@
 
-%% \section{Direct correlation function}
-
-%% \textcolor{red}{FIXME WRITE THIS SECTION!} The direct correlation
-%% function is readily accessible from density-functional theory, as the
-%% second functional derivative of the free energy:
-%% \textcolor{red}{(Something like...)}
-%% \begin{align}
-%%   -c^{(2)}(\rr_1, \rr_2) &= \frac{1}{k_BT}\frac{\delta^2 F_{ex}}{\delta n(\rr_1) \delta n(\rr_2)}
-%% \end{align}
-%% The direct correlation function is related to the ordinary correlation
-%% function by the Ornstein-Zernike relation
-%% \begin{align}
-%%   h(\rr_1, \rr'_2) &= g(\rr_1, \rr_2) - 1 \\
-%%   h(\rr_1, \rr_2) &= c(\rr_1, \rr_2)
-%%     + \int n(\rr_3)c(\rr_1, \rr_3) h(\rr_2, \rr_3)d\rr''
-%% \end{align}
-%% From these two equation, we could in principle extract the correlation
-%% function evaluated at contact from the DFT free energy functional.
-%% The Ornstein-Zernike equation is usually solved by taking a Fourier
-%% transform of all the relevant quantities.  In the inhomogeneous case,
-%% however, this approach isn't workable, since the Fourier transforms
-%% don't speed things up like you'd hope, and we don't have a periodic
-%% system in general.
-
-%% \textcolor{red}{It seems worth looking into this, to see what (if
-%%   anything) we can do! I don't see a solution just now, but that
-%%   doesn't mean there isn't one around somewhere.  In any case, just
-%%   looking at the direct correlation function might give us some
-%%   hints as to a different and interesting approximation for the
-%%   contact density.}
-
 \derivation{
@@
 }
-
-%% \section{Mean Contact density}\label{simple-contact}
-
-%% One reasonably simple question to ask is what the \emph{mean} contact
-%% density of an inhomogeneous system is.  This comes down to simply
-%% reproducing the derivation for a homogeneous system using the
-%% FMT functional.
-%% \begin{align}
-%%   p_{HS} &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
-%%   \ncontact &= \frac{1}{N k_BT 4\pi \sigma^2} \frac{dA_{HS}}{dR}
-%% \end{align}
-%% While this should give the true mean contact density for any given
-%% system, this is not particularly useful, as we have no way of knowing
-%% the contact density on any given sphere.  However, the mean contact
-%% density comes out as an integral, so we can plot the integrand, and
-%% treat it as a somewhat ill-defined local contact density.
-
-%% \derivation{
-
-%%   To compute the mean contact density of a system, we need
-%%   \begin{align}
-%%     \frac{d A_{HS}}{d R} &=
-%%     \frac{\partial A_{HS}}{\partial R} \\
-%%     &+
-%%     \int \left(
-%%     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-%%     \frac{d n_3(\mathbf{r}')}{d R}
-%%     +
-%%     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-%%     \frac{d n_2(\mathbf{r}')}{d R}
-%%     + \cdots
-%%     \right) d\mathbf{r}'
-%%   \end{align}
-
-%%   \begin{align}
-%%     \frac{dn_3(\mathbf{r})}{dR} &= n_2(\mathbf{r})\\
-%%     \frac{dn_2(\mathbf{r})}{dR} &= \int n(\mathbf{r}+\mathbf{r}')
-%%     \delta'(|\mathbf{r'}| - R)\mathbf{dr}'\\
-%%     &= \int n(\mathbf{r}+\mathbf{r}')
-%%     \delta'(r' - R) d\Omega' r'^2 dr'\\
-%%     &= \int 
-%%     \delta(r' - R) d\Omega' \left(\frac{dn(\mathbf{r}+\mathbf{r}')}{dr'}r'^2 + 2n(\mathbf{r}+\mathbf{r}')r' \right) dr'\\
-%%     &= \frac1{R} \int 
-%%     \delta(|\mathbf{r}'| - R)
-%%     \left(\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') +
-%%     2n(\mathbf{r}+\mathbf{r}') \right) \mathbf{dr}' \\
-%%     &= \frac{2}{R}n_2(\mathbf{r}) + \frac{1}{R} \int 
-%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-%%     \frac{dn_0(\mathbf{r})}{dR} &= \frac{1}{R} \int 
-%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-%%     \frac{dn_1(\mathbf{r})}{dR} &=
-%%     \frac1{4\pi R}\frac{dn_2(\mathbf{r})}{dR} - n_0(\mathbf{r}) \\
-%%     &= n_0(\mathbf{r}) + \frac{1}{4\pi R^2} \int 
-%%     \delta(|\mathbf{r}'| - R)\mathbf{r}'\cdot \nabla n(\mathbf{r}+\mathbf{r}') \mathbf{dr}' \\
-%%     \frac{d\mathbf{n}_{V2}}{dR} &= \nabla n_2
-%%   \end{align}
-
-%%   The total force due to $\Phi_1$ is actually very easy to work out:
-%%   \begin{align}
-%%     \frac{d A_{HS}^{(1)}}{d R} &=
-%%     \int \mathbf{dr}' \left(\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3) \right)
-%%   \end{align}
-%%   where you can see that $\frac{dn_0(\mathbf{r})}{dR}$ is zero for a
-%%   homogeneous system, and small for reasonably smooth density
-%%   distributions.  The next term from $\Phi_2$ is as follows:
-%%   \begin{align}
-%%     \frac{d A_{HS}^{(2)}}{d R} &=
-%%     \int \mathbf{dr}' \left(
-%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-%%     n_2
-%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2} - 2 n_1
-%%       \frac{dn_2}{dR} + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot
-%%     }{1-n_3}
-%%     \right) \\
-%%     &= \int \mathbf{dr}' \left(
-%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-%%     n_2
-%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-%%       - 2 n_1 \left(\frac2{R}n_2 \frac{dn_0}{dR}\right)
-%%       + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-%%     \right) \\
-%%     &= \int \mathbf{dr}' \left(
-%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-%%     n_2
-%%     - \frac{n_0n_2 - \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-%%       - 4 n_0n_2 - 2 n_1\frac{dn_0}{dR}
-%%       + 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-%%     \right) \\
-%%     &= \int \mathbf{dr}' \left(
-%%     \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{(1-n_3)^2}
-%%     n_2
-%%     + \frac{3n_0n_2 + \mathbf{n}_{V0} \cdot\mathbf{n}_{V2}
-%%       + 2 n_1\frac{dn_0}{dR}
-%%       - 2 \mathbf{n}_{V1} \cdot \nabla n_2 \cdot }{1-n_3}
-%%     \right)
-%%   \end{align}
-
-%%   And the last term looks like this.  I should perhaps note that this
-%%   term should be replaced by a tensor version if want to examine
-%%   seriously localized density distributions.
-%%   \begin{align}
-%%     A_{HS}^{(3)} &=
-%%     \int \mathbf{dr}'\left(
-%%     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-%%     \frac{
-%%       n_3 + (1-n_3)^2 \ln(1-n_3)
-%%     }{
-%%       36\pi n_3^2\left( 1 - n_3 \right)^2
-%%     }
-%%     \right) \\
-%%     \frac{d A_{HS}^{(3)}}{d R} &=
-%%     \int \mathbf{dr}'
-%%     \left[
-%%       3(n_2^2 - \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-%%       \left(\frac2{R}n_2 + \frac{dn_0}{dR}\right)
-%%       - 6n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-%%       \right]
-%%     \frac{
-%%       n_3 + (1-n_3)^2 \ln(1-n_3)
-%%     }{
-%%       36\pi n_3^2\left( 1 - n_3 \right)^2
-%%     }
-%%     \\
-%%     &+
-%%     (n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) n_2
-%%     \frac{
-%%       (1 - 2(1-n_3)\ln(1-n_3) - (1-n_3))n_3^2( 1 - n_3)^2
-%%       -
-%%       (n_3 + (1-n_3)^2 \ln(1-n_3))(-2 n_3^2( 1 - n_3) + 2n_3(1-n_3)^2)
-%%     }{
-%%       36\pi (n_3^2\left( 1 - n_3 \right)^2)^2
-%%     }
-%%     \\
-%%     &=
-%%     \int \mathbf{dr}'
-%%     \left[
-%%       (n_1n_2 - \mathbf{n}_{V1} \cdot \mathbf{n}_{V2})
-%%       \left(2n_2 + \frac{1}{4\pi}\frac{dn_0}{dR}\right)
-%%       - \frac{1}{2\pi}n_2 \mathbf{n}_{V2} \cdot \nabla n_2
-%%       \right]
-%%     \frac{
-%%       n_3 + (1-n_3)^2 \ln(1-n_3)
-%%     }{
-%%       3 n_3^2\left( 1 - n_3 \right)^2
-%%     }
-%%     \\
-%%     &+
-%%     \frac{n_2(n_2^3 - 3n_2\mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{36\pi}
-%%     \left( \frac{1}{(1-n_3)^2} + 2 \frac{\ln(1-n_3)}{n_3^2(1-n_3)} \right)
-%%   \end{align}
-
-%%   \begin{align}
-%%     \frac{d A_{HS}}{d R} &=
-%%     \int \mathbf{dr}' \left(
-%%     -\frac{n_0n_2}{1-n_3} - \frac{dn_0(\mathbf{r})}{dR}\ln(1-n_3)
-%%     + \cdots
-%%     \right)
-%%   \end{align}
-%% }
 
@@
   p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \\
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^A}\\
   \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2} \frac{\delta
@@
 \derivation{
-  So to compute this we need
+  \end{widetext}
+}
+
+\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+We compare with the averaged correlation function of
+Gross\cite{gross2009density}, which is of the asymmetrical variety
+($g_\sigma^A$), where the correlation function is averaged at location
+of one of the spheres in concact.
+\begin{align}
+  g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
+    n_{3'}\right)^3} \\
+  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
+\end{align}
+where $n_{3'}$ is an averaged
+filling fraction defined in a manner analogous to the $n_3$ of FMT,
+but with the hard-sphere radius replaced by the hard-sphere diameter,
+so the integrals range over spheres that are twice as large.
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/n0}
+\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+  consist of those which just touch the point $\mathbf{r}$.}
+\label{fig:n0}
+\end{figure}
+
+\section{Symmetrically averaged correlation function}\label{sec:g-S}
+
+An alternate definition of the averaged correlation function for an
+inhomogeneous system would be the correlation function averaged for
+spheres \emph{touching at a given point}.  This would be a symmetric
+measure, since both spheres involved are treated in the same way.  In
+fact, the density of such spheres is just the density $n_0(\rr)$ from
+FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
+is the fundamental measure of sphere surface contact, it makes sense
+to define a corresponding correlation function $g_\sigma^S$.  This is
+\emph{almost} what is done by Yu and  Wu
+
+\fixme{Analogous to the assymetric case is the symmetric, for which 
+we are no longer able to simply pick a hypothetical sphere centered 
+at position $r$ and expand its radius.  Instead we must formulate a way 
+to expand the radii of spheres that just touch a specific point of 
+contact and only expand them at that point of contact, in a symmetric 
+fashion.  Therefore, when we compute the neccasary derivitives wih 
+respect to $R$', we expand a multitude of spheres, all of them positioned 
+around and just touching the point $\mathbf{r}$.  When their radii 
+expand, each does not expand outward in all directions as in the 
+case of A, rather they each expand outward at only the point $\mathbf{r}$.
+One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains
+at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
+in this peculiar fashion, we have in the end the same amount of shell 
+area protruding outward as we would if we had simply expanded one 
+sphere in fashion of case A.  Also the protrusions, once added up 
+for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
+in their direction of protrusion.  Each protruding sphere bit will 
+expand outward from its center and all directions will be covered uniformly.
+
+These qualitative concepts lead us to derivations for the $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ term in our derivation of the pressure, eq \ref{eq:p_{HS}}.
+The rest of the derivation results when we recognize that instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ as in the assymmetric case, we will instead divide by $n_0(\mathbf{r})^2$ since we have symmetry around the point $\mathbf{r}$.
+This will yield our symmetric correlation, $g_\sigma^S$:
+
+\begin{align}
+  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
+  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \\
+  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
+  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
+\end{align}}Complete expressions for $g_\sigma^S$ are included in
+Appendix~\ref{appendix:g-S}.\\
+
+\section{Yu and Wu's functional}\label{sec:yuwu}
+
+Yu and Wu developed a functional that should correspond to the contact
+density at a given contact
+position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We
+will compare both their functional as well as ours with simulation
+data.  Yu and Wu use a symmetric formulation for $g_\sigma$, but
+instead of using $n_0$ as the associated density, it uses a density
+given by
+\begin{align}
+  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
+  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
+\end{align}
+where the function $\zeta$ is a measure of local inhomogeneity at the
+point of contact.  Given this definition of density, their $g_\sigma$ is
+given by:
+\begin{align}
+  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
+    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
+    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
+    \\
+  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
+    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
+    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
+    \\
+  g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
+    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
+    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}\right)
+\end{align}
+I think the adjusted functional $g_\sigma^{\text{Yu,}S}$ can be directly
+compared with our $g_\sigma^S$.  It is implemented in our code.
+
+\section{Comparison with simulation}
+
+We performed a Monte-Carlo simulation of the hard sphere fluid to
+measure the contact density as a function of position for a few simple
+geometries.  For each scenario, we compute and compare several
+quantities.  We compare the Monte Carlo density with the DFT
+prediction.  We also plot Monte Carlo results for the two contact
+densities: the contact density evaluated at contact, and the contact
+density evaluated at the sphere location.  Alongside these, we plot
+the contact density of Yu and
+Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, the contact density
+of Gross~\cite{gross2009density}, and the ``contact density at
+sphere'' from Section~\ref{sec:g-A}.  In principle, the
+functional of Yu (FIXME) should match the CenConDensity, while the DFT
+at sphere should match the ConDensity...
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-10}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:walls-10}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-30}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.3.}
+  \label{fig:walls-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/walls-40}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.4.}
+  \label{fig:walls-40}
+\end{figure}
+
+\subsection{Hard spheres at a hard wall}
+
+We begin by examining the simplest situation with an inhomogeneous
+density:  the case of hard spheres near a single hard wall.  This is
+shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
+
+\newcommand\sphereExplanation{ Blue curves describe the density of
+  hard spheres.  Green curves describe the contact density averaged
+  according to the point of contact.  Red curves describe the contact
+  density as averaged according to the centers of the spheres that are
+  touching.  In each case, solid lines correspond to the Monte Carlo
+  simulation, and dashed lines of various sorts represent DFT model
+  predictions.  }
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-08-013}
+  \caption{Density and contact densities of thirteen hard spheres in a
+    spherical cavity with diameter 8. \sphereExplanation }
+  \label{fig:sphere-8}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-12-112}
+  \caption{Density and contact densities of 112 hard spheres in a
+    spherical cavity with diameter 12.  \sphereExplanation}
+  \label{fig:sphere-12}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/contact-16-265}
+  \caption{Density and contact densities of 265 hard spheres in a
+    spherical cavity with diameter 16. \sphereExplanation}
+  \label{fig:sphere-16}
+\end{figure}
+
+\subsection{Hard spheres confined in a spherical cavity}
+
+The first geometry we consider is a number of hard spheres confined
+within a spherical cavity.  In
+Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show
+the densities and contact densities for several cavity sizes.  In each
+case, the number of hard spheres was chosen to set the mean filling
+fraction to around 0.3?.  This value was chosen because it roughly
+corresponds to the filling fraction of water at standard atmospheric
+temperature and pressure in the optimal SAFT model developed by Clark
+\emph{et al}\cite{clark2006developing}.
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-4-30}
+  \caption{Density and contact densities around a hard sphere with
+    radius 1.}
+  \label{fig:inner-4-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-4-40}
+  \caption{Density and contact densities around a hard sphere with
+    radius 1.}
+  \label{fig:inner-4-40}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=5cm]{figs/inner-8-30}
+  \caption{Density and contact densities around a hard sphere with
+    radius 3.}
+  \label{fig:inner-8-30}
+\end{figure}
+
+\subsection{Hard spheres around a hard sphere}
+
+It is also interesting to study a system with opposite curvature.  Put
+a (larger) hard sphere in the hard-sphere fluid, and see what the
+contact density is around that hard sphere.  This is particularly
+relevant when we consider using classical density-functional theory to
+model solvation.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}
+We (will) have compared several functionals for the contact density of
+a hard-sphere fluid with simulations.  This quantity is of particular
+interest, as it plays a critical role in Statistical Associating Fluid
+Theory (SAFT), which is the basis of a number of recent classical
+density functionals.  We compute the contact density directly from the
+White Bear version of the FMT hard-sphere functional, and find
+reasonable agreement?
+
+\appendix
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/gHS-vs-n}
+\caption{A comparison of the various approximations to the contact
+  density in the homogeneous limit.  As expected, they are all
+  identical, and we won't include this plot in the paper (but it's
+  here to verify that our code is working).}
+\label{fig:gHS-vs-n}
+\end{figure}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/free-energy}
+\caption{A comparison of the various approximations to the free energy
+  in the homogeneous limit.  Note that the White Bear functional uses
+  a modified form of the Carnahan equation of state.  As expected,
+  these are all identical, and we won't include this plot in the paper
+  (but it's here to verify that our code is working).}
+\label{fig:free-energy}
+\end{figure}
+
+\begin{widetext}
+
+\section*{Introduction to Appendices}
+
+In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ terms that must be derived for both the S and A cases.
+We break these terms up as follows:
+
   \begin{align}
@@
   \end{align}
-  There are \emph{many} such terms, and this could get expensive.
-
-  First, let's rewrite the free energy in terms of a smaller set of
-  weighted densities:
+
+We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc., terms.
+We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.
+Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc., terms for both the A and S cases.
+These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
+
+\subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Bear Mark I free energy $A_{HS}$}
+
+  First, let's rewrite the free energy in terms of a smaller set of weighted densities
+
   \begin{align}
@@
   \end{align}
+
+Taking derivatives we have
 
@@
     \frac{\mathbf{n}_{V2}}{1-n_3} \\
-    \frac{\delta A_{HS}}{\delta R(\mathbf{r}')} &=
-    \frac{n_2}{2\pi R^3} \ln\left( 1 - n_3\right)
-    - \frac1{R} \frac{n_1n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-  \end{align}
-
-*******************************************************************************8
-\fixme{The following is the same calculation for WBm2.  Not sure if you want these in the appendix or in the paper at all?}
+  \end{align}
+
+\subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Brear Mark II free energy $A_{HS}$}
+
+Similarly to the White Bear I case we first write
 
@@
     + \left(1 + \frac{1}{9}n_3^2 \phi_2(n_3)\right)   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + (1 - \frac{4}{9}n_3\phi_3(n_3))\frac{(n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})}{24\pi (1-n_3)^2}
+    + \left(1 - \frac{4}{9}n_3\phi_3(n_3)\right)\frac{n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{24\pi (1-n_3)^2}
     \right\}
@@
  \begin{align}
-   \phi_2 &= (6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3))\frac{1}{n_3^3}\\
-   \phi_3 &= (6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3))\frac{1}{4n_3^3}
+   \phi_2 &= \left(6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3)\right)\frac{1}{n_3^3}\\
+   \phi_3 &= \left(6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3)\right)\frac{1}{4n_3^3}
  \end{align}
@@
 \begin{align}
-   \frac{\delta \phi_2}{\delta n_3} &= ((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2]\frac{3}{n_3^4}\\
-   \frac{\delta \phi_3}{\delta n_3} &= (-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3))\frac{3}{4n_3^4}
+   \frac{\delta \phi_2}{\delta n_3} &= \left((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2\right)\frac{3}{n_3^4}\\
+   \frac{\delta \phi_3}{\delta n_3} &= \left(-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3)\right)\frac{3}{4n_3^4}
  \end{align}
@@
   \psi_1 &= \frac{n_0}{1-n_3}\\
-  \psi_2 &= (\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3})(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}) + (1 + \frac{1}{9}n_3^2\phi_2)(\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2})\\
-  \psi_3 &= (\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2})(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}) + 2(1 - \frac{4}{9}n_3\phi_3)(\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3})
+  \psi_2 &= \frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3}\left(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}\right) + \left(1 + \frac{1}{9}n_3^2\phi_2\right)\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2}\\
+  \psi_3 &= \frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2}\left(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}\right) + 2\left(1 - \frac{4}{9}n_3\phi_3\right)\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3}
 \end{align}
@@
 \begin{align}
-  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)(\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2})\\
+  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2}\\
   \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
@@
 
-****************************************************************************************\\
-Jeff's understanding as of 02/22/2012
-
-  In order to motivate our particular forms for $\frac{\delta n_3^{S} (\mathbf{r}')}{\delta R(\mathbf{r})}$, $\frac{\delta n_2^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc.
-we interpret the contact density in the $S$ case as follows:
-  Given that there is a sphere just touching the point $\mathbf{r}'$, $n_{Contact}^{S}(\mathbf{r}')$ is the density
-  of spheres that are also touching that point so that they are touching the given sphere.
-We can see this in the integral in eq \ref{eq:n-contactS},where we sum over the double densities of spheres that are positioned opposite eachother and touching at the point $\mathbf{r}$ and then divide
-  by $n_0$, the average density of a single sphere touching the point.  Thus we have the density of contact at $\mathbf{r}$, where the
-  contact between the two spheres actually takes place at $\mathbf{r}$.
-
-Involved in the terms in our free energy functional are derivatives of density functions with respect to $R$.
-We take these integrals differently when dealing with the S and the A cases.
-In both cases we wish to calculate how the density function changes when a 'certain amount of sphere', at some point, is changed.
-The A case is conceptually simple.
-We say that a changing spherical radius at a point $\mathbf{r}$ represents the outward expansion, by a distance $\delta R$, of the shell of any sphere that is centered at $\mathbf{r}$.
-In the case of S we instead expand a multitude of spheres, all of them positioned around and just touching the point $\mathbf{r}$.
-When their radii expand each does not expand outward in all directions as in the case of A, rather they each expand outward at only the point $\mathbf{r}$.
-One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains at radius $R$.
-When we expand each sphere that touches $\mathbf{r}$ in this peculiar fashion, we have in the end the same amount of shell area protruding outward as we would if we had simply expanded one sphere in fashion of case A.
-Also the protrusions, once added up for all spheres sourrounding $\mathbf{r}$, will yield a symmetry in their direction of protrusion.
-Each protruding sphere bit will expand outward from its center and all directions will be covered uniformly.
+\subsection{Expressions for $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the assymetric case}\label{appendix:g-A}
+
+In the case of A, the changing radius we refer to is the radius of a sphere that is centered at $\mathbf{r}$.
+The $n_3$ values that will be affected are those which are a distance $R$ from $\mathbf{r}$.
+
+\begin{align}
+  n_3^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}'')) d\mathbf{r}''\\ 
+  \frac{\delta n_3^{A} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
+  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}'' \\
+   &= n (\mathbf{r}) \delta(|\mathbf{r'} - \mathbf{r}| - R(\mathbf{r}))
+\end{align}
+
+And likewise:
+\begin{align}
+  n_2^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) d \mathbf{r}''\\
+  \frac{\delta n_2^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))\\
+ \end{align}
+
+where the trouble here is that the derivative of the delta function,
+$\delta'$, isn't something we really like to deal with.  Particularly
+as we're looking  at a 3D integral involving a 1D delta function,
+which means there are Jacobian factors to be considered.  On the other
+hand, there's no reason why we couldn't convolve with this kernel as
+efficiently as we do with any other kernel.
+\\
+\\
+The derivatives of the remaining scalar densities can be reduced to
+sums of the terms above:
+\begin{align}
+  n_1(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
+  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
+  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})}
+  &= -\frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+  -
+  \frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+\end{align}
+\begin{align}
+  n_0(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})^2}
+  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
+  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})^2}
+  &= -\frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})^2}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+  -
+  2\frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})^3}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+\end{align}
+
+Since the vector-weighted densities have no additional occurrences of
+$R$, their derivatives come out similarly.
+\begin{align}
+  \mathbf{n}_{V2}^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) 
+    \frac{\mathbf{r'}-\mathbf{r''}}{|\mathbf{r'}-\mathbf{r''}|} d \mathbf{r}''\\
+  \frac{\delta \mathbf{n}_{V2}^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= - n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))
+    \frac{\mathbf{r}'-\mathbf{r}}{|\mathbf{r}'-\mathbf{r}|}\\
+\end{align}
+
+and analogously for $\mathbf{n}_{V1}$:
+
+\begin{align}
+  \mathbf{n}_{V1}(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
+  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}\\
+  \frac{\delta \mathbf{n}_{V1}(\mathbf{r}')}{\delta R(\mathbf{r})}
+  &= -\frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}
+  -
+  \frac{n(\mathbf{r})}{4\pi
+    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}  
+\end{align}
+
+So, what is this convolution with a $\delta'$? Let's look at the
+scalar version:
+\fixme{With following the primes are certainly not in the right places}
+\begin{align}
+  \bar{f}(\mathbf{r})
+  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
+  \\
+  &= \int
+        f(\mathbf{r'})
+        \lim_{\Delta R \rightarrow 0}
+          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+            -
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &= \lim_{\Delta R \rightarrow 0}
+          \frac{\int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+            -
+            \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &\approx\lim_{\Delta R \rightarrow 0}
+          \frac{\int
+            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+            -
+            \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &\approx\lim_{\Delta R \rightarrow 0}
+          \frac{2\frac{\Delta R}{R} \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+      }{\Delta R}
+  \\
+  &= \frac{2}{R} \int
+            f(\mathbf{r'})
+            \delta(|\mathbf{r}-\mathbf{r'}| - R)
+\end{align}
+There is a sign ambiguity here with regard to whether we are asking
+about the derivative of this delta function with regard to $R$ or
+something else.  We just need to get it right.
+
+\subsection{Expressions for $\beta\frac{\delta n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the symetric case}\label{appendix:g-S}
 
@@
 
-  $\delta R(\mathbf{r})$ in this equation refers to an expansion, in the fashion described above, of all the spheres that just touch $\mathbf{r}$.
+\fixme{Not sure if we need any more explanation here (probably want better than this) -  $\delta R(\mathbf{r})$ in this equation refers to an expansion, in the fashion described above, of all the spheres that just touch $\mathbf{r}$.
 We can see that if the radii change in this fashion, the only point at which $n_3(\mathbf{r}')$ will be affected is at $\mathbf{r}' = \mathbf{r}$, so that $\delta n_3(\mathbf{r}')$ is zero everywhere except where $\mathbf{r}' = \mathbf{r}$.
-Thus we have the 2nd $\delta$ function in the lower integral.
-
+Thus we have the 2nd $\delta$ function in the lower integral.}
+\\
+\\
 This term in the free energy functional, once integrated, takes an interesting form:
@@
 
- $n_2(\mathbf{r}')$ in the case of S in calculated similarly:
+ $n_2(\mathbf{r}')$ in the case of S in calculated similarly
 \begin{align}
@@
 
-For $\mathbf{n}_{V2}(\mathbf{r}')$ in the case of S we take the gradient of $n_3(\mathbf{r}')$ with respect to $\mathbf{r}'$: 
+And similarly to the A case we compute
+\begin{align}
+  n_1^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
+    \frac{\delta n_1^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
+  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
+  -
+  \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
+\end{align}
+
+and 
+\begin{align}
+  n_0^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
+    \frac{\delta n_0^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} 
+  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
+  -
+  2\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^3} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
+\end{align}
+
 \begin{align}
@@
 
-  In the case of A, the changing radius we refer to is the radius of a sphere that is centered at $\mathbf{r}$ and it expands in a more conventional fashion.
-The $n_3$ values that will be affected are those which are a distance $R$ fom $\mathbf{r}$.
-
-\begin{align}
-  n_3^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}'')) d\mathbf{r}''\\ 
-  \frac{\delta n_3^{A} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}'' \\
-   &= n (\mathbf{r}) \delta(|\mathbf{r} - \mathbf{r}'| - R(\mathbf{r}))
-\end{align}
-
-And likewise:
-\begin{align}
-  n_2^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) d \mathbf{r}''\\
-  \frac{\delta n_2^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R)\\
- \end{align}
+\begin{align}
+ \mathbf{n}_{V1}^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
+   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}\\
+ \frac{\delta \mathbf{n}_{V1}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
+   \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} 
+   -
+   \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r'' 
+   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}
+\end{align}
+
+\begin{widetext}
+\section{Formulas generated from Haskell}
+
+\input{formulas}
+\begin{align}
+  \Phi_1 &= \phione \\
+  \Phi_2 &= \phitwo \\
+  \Phi_3 &= \phithree \\
+  \frac{d F}{dn_3} &= \dwbdnthree \\
+  \frac{d F}{dn_2} &= \dwbdntwo \\
+  \frac{d F}{dn_1} &= \dwbdnone \\
+  \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
+  \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev
+\end{align}
+
+\fixme{This is the hypthetical end of the paper}
 
@@
 Not sure if we'll use any of this, but wrote down some equations that Roundy wrote out.
-So can use the terms at least if going to write about this - I'm just copying it in for now (probably a lotof mistakes)
+So can use the terms at least if going to write about this - I'm just copying it in for now (probably a lotof mistakes)}
 
@@
 \end{align}
-}
-********************************************************************************\\
-
-These give us some basic tools with which to perform derivatives.
-
-  Finally, we get an answer like
-  \begin{align}
-    \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
-    \int \mathbf{dr}' \left\{
-    \left(
-    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-      n_3(\mathbf{r}'))^2}
-    + \frac{1}{4\pi}\frac{
-      \frac13 n_2^3 - n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}
-    }{
-      (1-n_3(\mathbf{r}'))^3
-    }
-    \right) n(\mathbf{r}') \delta(|\mathbf{r}-\mathbf{r}'| - R) \right.
-    \\
-    & \left.
-    + \frac{\delta A}{\delta n_2(\rr')} n(\rr) \delta'(|\rr-\rr'|-R) \cdots
-    \right\}
-  \end{align}
-  This comes out to quite a mess.  It isn't be infeasible to compute
-  this, but it is challenging.
-}
-
-\derivation{
+
 \end{widetext}
-}
-
-\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
-We compare with the averaged correlation function of
-Gross\cite{gross2009density}, which is of the asymmetrical variety
-($g_\sigma^A$), where the correlation function is averaged at location
-of one of the spheres in concact.
-\begin{align}
-  g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
-    n_{3'}\right)^3} \\
-  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
-\end{align}
-where $n_{3'}$ is an averaged
-filling fraction defined in a manner analogous to the $n_3$ of FMT,
-but with the hard-sphere radius replaced by the hard-sphere diameter,
-so the integrals range over spheres that are twice as large.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:n0}
-\end{figure}
-
-\section{Symmetrically averaged correlation function}\label{sec:g-S}
-
-An alternate definition of the averaged correlation function for an
-inhomogeneous system would be the correlation function averaged for
-spheres \emph{touching at a given point}.  This would be a symmetric
-measure, since both spheres involved are treated in the same way.  In
-fact, the density of such spheres is just the density $n_0(\rr)$ from
-FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
-is the fundamental measure of sphere surface contact, it makes sense
-to define a corresponding correlation function $g_\sigma^S$.  This is
-\emph{almost} what is done by Yu and
-
-\fixme{Add qualitative description of the derivation needed to get
-  $g_\sigma^S$.}
-
-Complete expressions for $g_\sigma^S$ are included in
-Appendix~\ref{appendix:g-S}.
-
-\section{Yu and Wu's functional}\label{sec:yuwu}
-
-Yu and Wu developed a functional that should correspond to the contact
-density at a given contact
-position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We
-will compare both their functional as well as ours with simulation
-data.  Yu and Wu use a symmetric formulation for $g_\sigma$, but
-instead of using $n_0$ as the associated density, it uses a density
-given by
-\begin{align}
-  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
-  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
-\end{align}
-where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact.  Given this definition of density, their $g_\sigma$ is
-given by:
-\begin{align}
-  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
-    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
-    \\
-  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
-    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
-    \\
-  g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
-    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
-    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}\right)
-\end{align}
-I think the adjusted functional $g_\sigma^{\text{Yu,}S}$ can be directly
-compared with our $g_\sigma^S$.  It is implemented in our code.
-
-\section{Comparison with simulation}
-
-We performed a Monte-Carlo simulation of the hard sphere fluid to
-measure the contact density as a function of position for a few simple
-geometries.  For each scenario, we compute and compare several
-quantities.  We compare the Monte Carlo density with the DFT
-prediction.  We also plot Monte Carlo results for the two contact
-densities: the contact density evaluated at contact, and the contact
-density evaluated at the sphere location.  Alongside these, we plot
-the contact density of Yu and
-Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, the contact density
-of Gross~\cite{gross2009density}, and the ``contact density at
-sphere'' from Section~\ref{sec:g-A}.  In principle, the
-functional of Yu (FIXME) should match the CenConDensity, while the DFT
-at sphere should match the ConDensity...
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-10}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.1.}
-  \label{fig:walls-10}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-30}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.3.}
-  \label{fig:walls-30}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/walls-40}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.4.}
-  \label{fig:walls-40}
-\end{figure}
-
-\subsection{Hard spheres at a hard wall}
-
-We begin by examining the simplest situation with an inhomogeneous
-density:  the case of hard spheres near a single hard wall.  This is
-shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
-
-\newcommand\sphereExplanation{ Blue curves describe the density of
-  hard spheres.  Green curves describe the contact density averaged
-  according to the point of contact.  Red curves describe the contact
-  density as averaged according to the centers of the spheres that are
-  touching.  In each case, solid lines correspond to the Monte Carlo
-  simulation, and dashed lines of various sorts represent DFT model
-  predictions.  }
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-08-013}
-  \caption{Density and contact densities of thirteen hard spheres in a
-    spherical cavity with diameter 8. \sphereExplanation }
-  \label{fig:sphere-8}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-12-112}
-  \caption{Density and contact densities of 112 hard spheres in a
-    spherical cavity with diameter 12.  \sphereExplanation}
-  \label{fig:sphere-12}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/contact-16-265}
-  \caption{Density and contact densities of 265 hard spheres in a
-    spherical cavity with diameter 16. \sphereExplanation}
-  \label{fig:sphere-16}
-\end{figure}
-
-\subsection{Hard spheres confined in a spherical cavity}
-
-The first geometry we consider is a number of hard spheres confined
-within a spherical cavity.  In
-Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show
-the densities and contact densities for several cavity sizes.  In each
-case, the number of hard spheres was chosen to set the mean filling
-fraction to around 0.3?.  This value was chosen because it roughly
-corresponds to the filling fraction of water at standard atmospheric
-temperature and pressure in the optimal SAFT model developed by Clark
-\emph{et al}\cite{clark2006developing}.
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/inner-4-30}
-  \caption{Density and contact densities around a hard sphere with
-    radius 1.}
-  \label{fig:inner-4-30}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/inner-4-40}
-  \caption{Density and contact densities around a hard sphere with
-    radius 1.}
-  \label{fig:inner-4-40}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[height=5cm]{figs/inner-8-30}
-  \caption{Density and contact densities around a hard sphere with
-    radius 3.}
-  \label{fig:inner-8-30}
-\end{figure}
-
-\subsection{Hard spheres around a hard sphere}
-
-It is also interesting to study a system with opposite curvature.  Put
-a (larger) hard sphere in the hard-sphere fluid, and see what the
-contact density is around that hard sphere.  This is particularly
-relevant when we consider using classical density-functional theory to
-model solvation.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Conclusion}
-We (will) have compared several functionals for the contact density of
-a hard-sphere fluid with simulations.  This quantity is of particular
-interest, as it plays a critical role in Statistical Associating Fluid
-Theory (SAFT), which is the basis of a number of recent classical
-density functionals.  We compute the contact density directly from the
-White Bear version of the FMT hard-sphere functional, and find
-reasonable agreement?
-
-\appendix
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/gHS-vs-n}
-\caption{A comparison of the various approximations to the contact
-  density in the homogeneous limit.  As expected, they are all
-  identical, and we won't include this plot in the paper (but it's
-  here to verify that our code is working).}
-\label{fig:gHS-vs-n}
-\end{figure}
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/free-energy}
-\caption{A comparison of the various approximations to the free energy
-  in the homogeneous limit.  Note that the White Bear functional uses
-  a modified form of the Carnahan equation of state.  As expected,
-  these are all identical, and we won't include this plot in the paper
-  (but it's here to verify that our code is working).}
-\label{fig:free-energy}
-\end{figure}
-
-\begin{widetext}
-  \section{Partial derivatives of the White Bear free energy functional}
-
-  Here is the White Bear free energy functional for the hard-sphere
-  fluid:~\cite{roth2002whitebear}
-  \begin{align}
-    \beta A_{HS} &= \int \left\{
-    -n_0 \ln\left( 1 - n_3\right)
-    + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \right\}
-\end{align}
-
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-      n_3(\mathbf{r}'))^2}
-    + \frac{n_2^3 -
-      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}
-    %\left(
-    %  \frac{1}{(1-n_3)^2} + 2\frac{\ln(1-n_3)}{n_2^2(1-n_3)}
-    %\right)
-    \\
-    & \times \left(\frac{2}{n_3(1-n_3)^3} -\frac1{n_3^2(1-n_3)^2}  -
-      \frac{1}{n_3^2(1-n_3)} - 2\frac{\ln(1-n_3)}{n_3^3}\right) \\
-    \beta\frac{\partial A_{HS}}{\partial n_3}
-    &\sim \frac{n}{1-\eta} + \frac{3n\eta}{(1-\eta)^2} +
-      \frac{2n\eta}{(1-\eta)^3} -
-      \frac{n}{(1-\eta)^2} -
-      \frac{n}{1-\eta} -
-      2n \frac{ \ln(1-\eta) }{ \eta } \\
-    &= \frac{3n\eta}{(1-\eta)^2} +
-      \frac{2n\eta}{(1-\eta)^3} -
-      \frac{n}{(1-\eta)^2} -
-      2n \frac{ \ln(1-\eta) }{ \eta } \\
-    &= n\left(
-      \frac{3\eta - 3 \eta^2}{(1-\eta)^3} +
-      \frac{2\eta}{(1-\eta)^3} -
-      \frac{1 - \eta}{(1-\eta)^3} -
-      2 \frac{ \ln(1-\eta) }{ \eta }
-      \right) \\
-    &= n\left(
-      \frac{6\eta - 3 \eta^2 - 1}{(1-\eta)^3} -
-      2 \frac{ \ln(1-\eta) }{ \eta }
-      \right)
-\end{align}
 \end{widetext}
 
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &= -\ln(1-n_3)
-    \\
-    \beta\frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
-    \\
-    \beta\frac{\partial A_{HS}}{\partial n_1}
-    &\sim n \frac{4\pi R^2}{1-\eta}
-\end{align}
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-      \frac{n_1}{1-n_3}
-      + (n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
-        (1-n_3)^2\ln(1-n_3)}{
-        12\pi n_3^2(1-n_3)^2
-      }
-\end{align}
-\begin{align}
-    \beta\frac{\partial A_{HS}}{\partial n_2}
-    &\sim n\frac{R}{1-\eta} +
-    \frac{4\pi}{3} R^4 n^2 \frac{\eta + (1-\eta)^2\ln(1-\eta)}{\eta^2(1-\eta)^2}
-    \\
-    &= n\frac{R}{1-\eta} +
-    \frac{R n}{(1-\eta)^2}
-    + Rn \frac{\ln(1-\eta)}{\eta}
-    \\
-    &= nR\left( \frac{1 + \eta^2  - 2\eta}{(1-\eta)^3} +
-      \frac{1 - \eta}{(1-\eta)^3}
-    + \frac{\ln(1-\eta)}{\eta}
-    \right)
-    \\
-    &= nR\left( \frac{2 + \eta^2  - 3\eta}{(1-\eta)^3}
-    + \frac{\ln(1-\eta)}{\eta}
-    \right)
-\end{align}
-\begin{align}
-    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
-      \frac{\mathbf{n}_{V2}}{1-n_3}
-    \\
-    \beta\frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-      \frac{\mathbf{n}_{V1}}{1-n_3}
-      - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
-        (1-n_3)^2\ln(1-n_3)}{
-        36\pi n_3^2(1-n_3)^2
-      }
-  \end{align}
-
-\begin{widetext}
-\begin{align}
-  \beta \frac{\partial A_{HS}}{\partial R} &\sim
-    \beta \frac{\partial A_{HS}}{\partial n_3} n_2 +
-    \beta \frac{\partial A_{HS}}{\partial n_2} 2 \frac{n_2}{R} +
-    \beta \frac{\partial A_{HS}}{\partial n_1} n
-  \\
-  &=
-    n_2n \left(
-      \frac{6\eta - 3\eta^2 - 1}{(1-\eta)^3} -
-      2\frac{\ln(1-\eta)}{\eta}
-    \right) +
-    n_2n \left( \frac{4+2\eta^2-6\eta}{(1-\eta)^3} + 2\frac{\ln(1-\eta)}{\eta} \right) +
-    n_2n\frac{1}{1-\eta}
-  \\
-  &=
-    n_2n \left(
-      \frac{6\eta - 3\eta^2 - 1}{(1-\eta)^3} +
-    \frac{4+2\eta^2-6\eta}{(1-\eta)^3} +
-    \frac{1 - 2\eta + \eta^2}{(1-\eta)^3}
-  \right)
-  \\
-  &=
-    n_2n \frac{4 - 2\eta}{(1-\eta)^3}
-\end{align}
-\end{widetext}
-Recall from Equation~\ref{eq:dAhsdR}, the standard Carnahan result is
-\begin{align}
-  \frac{dA_{HS}}{dR}
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R}
-  \\ &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} n_2
-\end{align}
-which means that we miraculously come out with a correct answer!
-
-\section{Derivatives of the fundamental-measure weighted densities}
-
-The derivative of the filling fraction is very easy:
-\begin{align}
-  n_3(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
-  \Theta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
-  \frac{\delta n_3(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= n(\mathbf{r'})\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-\end{align}
-
-The derivative of the density of surface is harder:
-\begin{align}
-  n_2(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
-  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
-  \frac{\delta n_2(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-\end{align}
-where the trouble here is that the derivative of the delta function,
-$\delta'$, isn't something we really like to deal with.  Particularly
-as we're looking  at a 3D integral involving a 1D delta function,
-which means there are Jacobian factors to be considered.  On the other
-hand, there's no reason why we couldn't convolve with this kernel as
-efficiently as we do with any other kernel.
-
-The derivatives of the remaining scalar densities can be reduced to
-sums of the terms above:
-\begin{align}
-  n_1(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
-  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
-  \frac{\delta n_1(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= \frac{n(\mathbf{r'})}{4\pi
-    R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-  -
-  \frac{n(\mathbf{r'})}{4\pi
-    R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-\end{align}
-\begin{align}
-  n_0(\mathbf{r}) &= \int \mathbf{dr'} \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}
-  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'})) \\
-  \frac{\delta n_0(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= \frac{n(\mathbf{r'})}{4\pi
-    R(\mathbf{r'})^2}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-  -2
-  \frac{n(\mathbf{r'})}{4\pi
-    R(\mathbf{r'})^3}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-\end{align}
-
-Since the vector-weighted densities have no additional occurrences of
-$R$, their derivatives come out similarly.
-\begin{align}
-  \mathbf{n}_{V2}(\mathbf{r}) &= \int \mathbf{dr'} n(\mathbf{r'})
-  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
-  \frac{\delta \mathbf{n}_{V2}(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= n(\mathbf{r'})\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
-\end{align}
-and analogously for $\mathbf{n}_{V1}$.
-\begin{align}
-  \mathbf{n}_{V1}(\mathbf{r}) &= \int \mathbf{dr'}
-  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}
-  \delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))\frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} \\
-  \frac{\delta \mathbf{n}_{V1}(\mathbf{r})}{\delta R(\mathbf{r}')}
-  &= \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})}\delta'(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|} -
-  \frac{n(\mathbf{r'})}{4\pi R(\mathbf{r'})^2}\delta(|\mathbf{r}-\mathbf{r'}| - R(\mathbf{r'}))
-  \frac{\mathbf{r}-\mathbf{r'}}{|\mathbf{r}-\mathbf{r'}|}
-\end{align}
-
-So, what is this convolution with a $\delta'$? Let's look at the
-scalar version:
-\begin{align}
-  \bar{f}(\mathbf{r})
-  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
-  \\
-  &= \int
-        f(\mathbf{r'})
-        \lim_{\Delta R \rightarrow 0}
-          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-            -
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &= \lim_{\Delta R \rightarrow 0}
-          \frac{\int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-            -
-            \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &\approx\lim_{\Delta R \rightarrow 0}
-          \frac{\int
-            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-            -
-            \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &\approx\lim_{\Delta R \rightarrow 0}
-          \frac{2\frac{\Delta R}{R} \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &= \frac{2}{R} \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-\end{align}
-There is a sign ambiguity here with regard to whether we are asking
-about the derivative of this delta function with regard to $R$ or
-something else.  We just need to get it right.
-
-\derivation{
-\begin{widetext}
-\section{The White Bear free energy for homogeneous systems}
-
-Here is the White Bear free energy functional for the hard-sphere
-fluid:~\cite{roth2002whitebear}
-\begin{align}
-  \frac{A_{HS}}{k_BT} &= \int \left\{
-  -n_0 \ln\left( 1 - n_3\right)
-  + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-  + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-    n_3 + (1-n_3)^2 \ln(1-n_3)
-  }{
-    36\pi n_3^2(1-n_3)^2
-  }
-  \right\} \\
-  &= \int \left\{
-  -n \ln\left( 1 - \eta\right)
-  + \frac{n^2 (4\pi R^2)  (R)}{1-\eta}
-  + n^3 (4\pi R^2)^3 \frac{
-    \eta + (1-\eta)^2 \ln(1-\eta)
-  }{
-    36\pi \eta^2(1-\eta)^2
-  }
-  \right\} \\
-  &= \int \left\{
-  -n \ln\left( 1 - \eta\right)
-  + \frac{3 n \eta}{1-\eta}
-  + n^3 (4\pi R^2)^3 \frac{
-    \eta + (1-\eta)^2 \ln(1-\eta)
-  }{
-    36\pi \eta^2(1-\eta)^2
-  }
-  \right\} \\
-  &= \int \left\{
-  -n \ln\left( 1 - \eta\right)
-  + \frac{3 n \eta}{1-\eta}
-  + n^3 4\pi (4\pi R^3)^2 \frac{
-    \eta + (1-\eta)^2 \ln(1-\eta)
-  }{
-    36\pi \eta^2(1-\eta)^2
-  }
-  \right\} \\
-  &= \int \left\{
-  -n \ln\left( 1 - \eta\right)
-  + \frac{3 n \eta}{1-\eta}
-  + n\eta^2 4\pi 9 \frac{
-    \eta + (1-\eta)^2 \ln(1-\eta)
-  }{
-    36\pi \eta^2(1-\eta)^2
-  }
-  \right\} \\
-  &= N \left\{
-  -\ln\left( 1 - \eta\right)
-  + \frac{3 \eta}{1-\eta}
-  + \frac{
-    \eta + (1-\eta)^2 \ln(1-\eta)
-  }{
-    (1-\eta)^2
-  }
-  \right\} \\
-  &= N \left\{
-  -\ln\left( 1 - \eta\right)
-  + \frac{3 \eta}{1-\eta}
-  + \frac{
-    \eta
-  }{
-    (1-\eta)^2
-  }
-  + \ln(1-\eta)
-  \right\} \\
-  &= N \left\{
-  \frac{3\eta}{1-\eta}
-  + \frac{
-    \eta
-  }{
-    (1-\eta)^2
-  }
-  \right\} \\
-  &= N
-  \frac{3\eta(1-\eta) + \eta
-  }{
-    (1-\eta)^2
-  } \\
-  &= N
-  \frac{4\eta -3\eta^2}{(1-\eta)^2}
-\end{align}
-This matches the usual expression for the Carnahan-Starling equation
-of state the excess free energy, which is
-\begin{align}
-  A = Nk_BT\frac{4\eta -3\eta^2}{(1-\eta)^2}
-\end{align}
-\end{widetext}
-}
-
-\section{Expression for asymmetrically averaged correlation
-  function}\label{appendix:g-A}
-
-\section{Expression for symmetrically averaged correlation
-  function}\label{appendix:g-S}
-
-\section{Expression for asymmetrically averaged correlation
-  function based on White Brear Mark II free energy}\label{appendix:g-A-ii}
-
-\section{Expression for symmetrically averaged correlation
-  function based on White Brear Mark II free energy}\label{appendix:g-S-ii}
-
-\begin{widetext}
-\section{Formulas generated from Haskell}
-
-\input{formulas}
-\begin{align}
-  \Phi_1 &= \phione \\
-  \Phi_2 &= \phitwo \\
-  \Phi_3 &= \phithree \\
-  \frac{d F}{dn_3} &= \dwbdnthree \\
-  \frac{d F}{dn_2} &= \dwbdntwo \\
-  \frac{d F}{dn_1} &= \dwbdnone \\
-  \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
-  \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev
-\end{align}
-\end{widetext}
-
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2012-04-23 papers/water-SAFT/paper.tex
Made a few changes to paper.  Moved paper directory to papers/water-SAFT. Tested everything on quipu.

--- 2012-03-28 paper/paper.tex
@@
 vaporization, thermal conductivity and surface tension of any
-non-metallic liquid;  water also has an unusually high dielectric constant.
+non-metallic liquid.  Water also has an unusually high dielectric constant.
 Much of the unusual behavior of water is explained by the ability of
@@
 \needsworknow{Add discussion and citations for more recent papers}
+Test
 
@@
 classical hard sphere and adds association terms based on Statistical 
-Associating Fluid Theory (SAFT). Water, unlike simple
-fluids which are mainly defined by van der Waals forces and weak electrostatic
-interations, requires direct treatment of association effects. The strong 
+Associating Fluid Theory (SAFT). Water is unlike simple
+fluids (which are mainly defined by van der Waals forces and weak electrostatic
+interations) because it requires direct treatment of association effects. The strong 
 hydrogen bonds and polar
@@
 The association functional we use is a modified version of that of Yu
-and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}.\footnote{We
+and Wu\cite{yu2002fmt-dft-inhomogeneous-associating} \footnote{We
   should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
@@
 Figure~\ref{fig:density-single-rod} shows density
-profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
-expected oscillations outside the rod and as the distance away from the rod
-increases
-the density approaches that of the bulk saturated liquid density. The small
-radius
+profiles for three different radii (0.3~nm, 0.5~nm, and 1~nm) which have the
+expected oscillations outside the rod that reduce to the bulk saturated liquid
+density as the distance away from the rods increases. The small radius
 rod has comparatively larger oscillations and a higher contact density.
@@
 for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
-on the radii of the rods, and is about 0.54~nm for the rods shown in 
+on the radii of the rods, which is about 0.54~nm for the rods shown in 
 Figure~\ref{fig:density-rods}.
@@
 the value of $\frac{\gamma} p$ is equal to about $20~\mu m$ which is
-much larger than any of the solutes we study. We expect that for larger 
-rods ... \needsworknow{Explain ``bowing'' out of vapor area for larger rods.}
+much larger than any of the solutes we study. We expect that for those 
+rods, the water on the sides of the 'stadium' configuration will bow inward 
+between the rods and the densty will still reduce to vapor where the rods 
+are closest to each other.
 
@@
 separation result for a radius of 0.3~nm does not match the prediction
-well(32\% difference).
+well (32\% difference).
 
@@
 We simulate four hydrophobic rods in water using the same radii as our
-two rods ($r = 0.3$~nm~-~$1.0$~nm)and
+two rods ($r = 0.3$~nm~-~$1.0$~nm) and
 separation distances between $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
@@
 where vapor is present in the space between the rods until a critical separation
-where it becomes filled with water (shown in Fig. \ref{fig:density-four-rods}). 
-For large $d$,
-the rods behave like separate single rods.  Free energy per length
-is plotted against rod separation in Figure \ref{fig:four-rods-energy-vs-distance}.
-From the slope we calculate the force per length to be [number].
-
-The rounded square formed by the area of the four rods and the space
+where it becomes filled with water and the rods behave like separate single rods
+(shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
+is plotted against rod separation in Figure \ref{fig:four-rods-energy-vs-distance} and
+from the slope we calculate the force per length to be [number].
+
+The 'rounded square' formed by the area of the four rods and the space
 inbetween them has a circumference estimated by
@@
 \end{equation}
-We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 3.0$~nm
+We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 0.3$~nm
 rod has the largest difference between simulation and prediction with
-a 23\% difference, and better agreement for larger radii (between 0.5\% and 5.6\%).
+a 23\% difference, and better agreement for larger radii rods (between 0.5\% and 5.6\%).
 
@@
 contact value theorem cannot agree, since the free energies do not
-agree.c
+agree.
 
@@
 accuracy of this classical treatment is the calculation of the force on the
-stadium shape. The force per length is constant and equal to twice the surface tension.
+stadium shape. The force per length is constant and equal to twice the
+surface tension. A simulations of hydrophobic spheres agrees with
+SPC/E calculations and a system with four hydrophobic rods exhibited
+similar behavior as the two rod.
 

### 2012-04-24 papers/water-SAFT/paper.tex
Minor modifications to paper and GNUmakefile.  Removed lines added to configure.ac when I moved /paper Tested everything on quipu.

@@
 Using this approximation with an additional length scale parameter, we are able to calculate the vapor pressure, saturated liquid density, surface tension, and enthalpy of water, as well as the number of hydrogen bonds and density oscillations formed near hydrophobic rods and spheres immersed in water.
-\needsworknow{For two rods being pulled apart, a transition occurs 
-from a vapor state to a liquid state in the area between
-the rods at approximately $(\pi-2)$ multipled by the radius. Even 
-for nanoscale rods (radius 0.5~-~1.0~nm), we find that in the
-vapor state, the force can be approximated
-classically and is equal to 144~mN/m, or twice the bulk surface tension.
-(Edit this and add spheres stuff)}
+For two rods and four rods(arranged as the corners of a square) in water, our functional predicts
+a transition from a vapor state to a liquid state in the area between
+the rods that agrees with clasical theory.   We find that in the vapor state, the force per length can be
+approximated classically and is equal to 144~mN/m, or twice the bulk surface 
+tension even for nm scale rods, and  our hydrophobic spheres agree
+reasonably well with SPC/E simulations.
 This functional is computationally efficient and provides a foundation for 
@@
 \emph{ab initio} approaches using classical molecular dynamics have
-\textcolor{red}{vague problems}\cite{weber2010ab-initio-water}.  A
+\textcolor{red}{vague problems [DJR]}\cite{weber2010ab-initio-water}.  A
 more efficient approach is needed in order to study nanoscale and
@@
 rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.  This functional
-\needsworknow{has many nice properties}.
+\needsworknow{has many nice properties DJR}.
 
@@
 \needsworknow{Add discussion and citations for more recent papers}
-Test
+
+A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
+describes water as non-interacting molecules in a an
+effective external potential.  This functional contains the exact
+entropy for no-interacting molecules and, among other things, can 
+predict the nonlinear dielectric response in water but must be paramaterized
+for different temperatures.
 
@@
 (shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
-is plotted against rod separation in Figure \ref{fig:four-rods-energy-vs-distance} and
-from the slope we calculate the force per length to be [number].
+is plotted against rod separation in Figure
+\ref{fig:four-rods-energy-vs-distance} with a slight disagreement of
+the predicted$2\gamma$ force per unit length.
 

### 2012-04-30 papers/water-SAFT/paper.tex
Rearranged/rewrote Fig. 8 portion of two rods section so the references are less spread out.  Other minor changes. Tested everything on quipu.

@@
 for different temperatures.
-
-One of the most interesting functionals for water is that of Ding \emph{et
-al}, which was developed in order to explain the freezing of
-water~\cite{ding1987}.  This functional uses the approach of Chandler
-\emph{et al}, which uses separate atomic densities for the constituent
-atoms of a molecular fluid~\cite{chandler1986a, chandler1986b}.  Although
-this is a compelling approach, the actual functional developed cannot
-describe the vapor state or liquid-vapor interface, both of which are
-crucial to correct predictions of hydrophobic behavior at large length
-scales.  In addition, the functional of Ding \emph{et al} also fails to
-rigorously enforce intramolecular correlations, which casts doubt on its
-suitability for describing interactions with ionic solutes.
 
@@
 
-The free energy per length is compared for rods of several different radii
-in Figure~\ref{fig:rods-energy-vs-distance}. For ease of comparison, we 
-arbitrarily offset each free energy such that they all go 
-to zero at large separations. For large $d$, the rods 
-simply act like two separate single rods in water and the energy will
-reach an asymptote determined by the effective surface tension (shown in 
-Figure~\ref{fig:energy-vs-diameter}). For small $d$, we see that all the rods 
-have the same behavior, with a positive linear relationship
-between free energy per length and separation during the initial state 
-(where the rods have vapor between them). This indicates that the force 
-per length on the rods is constant, and surprisingly independent of separation. 
-By finding the slope of this curve, we calculate this force per length to be 144~mN/m.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
-\end{center}
-\caption{ Free energy of hydration (also known as the potential of mean force) 
-versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 1.0~nm.
-All were arbitrarily offset to zero at large separations. The
-transition corresponds to the phase change from
-vapor to liquid between the rods as pictured in the density profiles in 
-Figure~\ref{fig:density-rods}. }
-\label{fig:rods-energy-vs-distance}
-\end{figure}
-
-To explain this, we consider a macroscopic surface
+To determine the critical separation, we consider a macroscopic surface
 tension model where the free energy in the macroscopic limit is given by 
@@
 much larger than any of the solutes we study. We expect that for those
+larger radius 
 rods, the water on the sides of the 'stadium' configuration will bow inward 
@@
 Starting from the surface energy term, we can calculate the 
-free energy per 
-length, which is equal to the circumference multiplied by the surface tenson. 
-The force
-per length is the derivative of this with respect to the separation. 
+free energy per length, which is equal to the circumference multiplied 
+by the surface tenson. The force per length is the derivative of this
+with respect to the separation. 
 We approximate the circumference of the stadium-shape by
@@
 
-This is consistent with the calculation of the force per length from 
-Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
-in Figure~\ref{fig:surface-tension} (72~mN/m). This indicates that
-even though the rods are nanoscale, the water molecules will ``see'' them classically.
+This is consistent with the calculation of the surface tension in
+Figure~\ref{figs:surface-tension} (72~mN/m) and the force per length
+from Figure ~\ref{fig:rods-energy-vs-distance} (144~mN/m) which
+compares the free energy per length for rods of several different
+radii. This agreement indicates that even though the rods are nanoscale, the
+water molecules will ``see'' them classically.  For large $d$, the rods 
+simply act like two separate single rods in water and the energy will
+reach an asymptote determined by the effective surface tension (shown in 
+Figure~\ref{fig:energy-vs-diameter}). In figure
+~\ref{fig:rods-energy-vs-distance}, we see that for small $d$ all the rods 
+have the same behavior with a positive linear relationship
+between free energy per length and separation during the initial state 
+(where the rods have vapor between them). This indicates that the force 
+per length on the rods is constant, and surprisingly independent of separation. 
+By finding the slope of this curve, we calculate this force per length to be 144~mN/m.
 
 To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will again use this model and propose that 
+radius, we will again use the aforementioned model and propose that 
 the critical separation, $d_c$, will occur when the free energies are equal. This 
@@
 The predicted and actual computed values for $d_c$ are shown graphically in 
-Figure~\ref{fig:rods-energy-vs-distance}, and they match fairly well for the 
-0.5~-~1.0~nm rods with differences between 1.3~-~5.5\%. We hypothesize that larger radii rods
+Figure~\ref{fig:rods-energy-vs-distance} where there is a transition
+betwen the sloped and flat portion of each plot. They match fairly well for the 
+0.5~-~1.0~nm rods with differences between 1.3~-~5.5\% and we hypothesize that larger radii rods
 would also follow this trend, but smaller rods are likely to have a different
@@
 well (32\% difference).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\end{center}
+\caption{ Free energy of hydration (also known as the potential of mean force) 
+versus separation for two hydrophobic rods ranging in radius from
+0.3~nm to 1.0~nm.
+All were arbitrarily offset to zero at large separations for ease of comparison. The
+transition corresponds to the phase change from
+vapor to liquid between the rods as pictured in the density profiles in 
+Figure~\ref{fig:density-rods}. }
+\label{fig:rods-energy-vs-distance}
+\end{figure}
 

### 2012-05-05 papers/water-SAFT/paper.tex
various saft paper edits Tested everything on bingley.

@@
 
-\newcommand{\fixme}[1]{\red{[#1]}}
-
-% needsworklater is used to annotate bits that need work, but that we
-% can postpone for a while.
-\newcommand{\needsworklater}[1]{\emph{[#1]}}
 % needsworknow is intended to prioritize stuff that needs fixing.
@@
 
-{\begin{itemize}
-\color{blue}
-\item Put finished things up here
-\item[JRH] Rewrite abstract (copying from project)
-\item[JRH] Cut the test-particle section
-\item[JRH] Rewrite and join introductions
-\item[JRH] Decide what to do with slit (cut)
-\item[JRH] Combine and edit Other Thermo and Empirical Parameters sections
-\color{red}
-\item[JRH] Work on abstract
-\item[DJR] Read and edit changes to abstract and introduction
-\item[DJR] Look through theory section for stuff to cut
-\item[DJR,JRH] Finalize figures in beginning
-\item[JRH] Better/expanded description of ``macroscopic surface tension'' model in 2 rods section?
-\item[JRH] Recalculate and clean up 2 rods figures and data - new code?
-\item[DJR] Write text (or clean up existing) for hard sphere solutes
-\item[EJK] Add 4-rods figures
-\item[EJK] Add 4-rods text
-\item[JRH,DJR] Rewrite conclusion
-\end{itemize}
-}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 neighboring solutes~\cite{izvekov2005, choe2007}.  On top of this,
-\emph{ab initio} approaches using classical molecular dynamics have
-\textcolor{red}{vague problems [DJR]}\cite{weber2010ab-initio-water}.  A
-more efficient approach is needed in order to study nanoscale and
+standard \emph{ab initio} methods using classical molecular dynamics
+without van der Waals corrections strongly over-structure water, to
+the point that ice melts at over 400K\cite{yoo2009phase}!  There has
+been a recent flurry of publications implicating van der Waals effects
+as significant in reducing this
+over-structuring\cite{lin2009importance, wang2011density,
+  mogelhoj2011ab, jonchiere2011van}.  It has also been bound that
+inclusion of nuclear quantum effects can have a similar
+effect\cite{morrone2008nuclear}.  Each of these corrections imposes an
+additional computational burden on an approach that is already
+feasible only for very small number of water molecules. A more
+efficient approach is needed in order to study nanoscale and
 hydrophobic solutes.
@@
 \subsection{Classical density-functional theory}
-
-%\begin{figure}
-%\includegraphics[width=6cm, angle=270]{figs/correlation}
-%\caption{ Experimental atom-atom radial distribution functions for water at
-%room temperature and pressure from reference~\cite{Soper2000}.\needsworknow{Either 
-%mention this figure in the text or delete it.}  }
-%\label{correlation}
-%\end{figure}
 
@@
 
-A number of exact properties are easily achieved in the density-functional
-framework, such as the relationship between the pressure on a hard wall and
-the contact density---which ensures a correct solvation free energy for
-small hard solutes.  Much of the research on classical density-functional
-theory has focused on the hard-sphere fluid~\cite{curtin1985,
-rosenfeld1989, rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.
-This has led to a large number of very sophisticated functionals, such as
-the fundamental-measure theory (FMT) functional~\cite{rosenfeld1989,
-rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}.  This functional
-\needsworknow{has many nice properties DJR}.
-
-Several classical density functionals have already been developed for
-water~\cite{ding1987, Yang1992, Jaqaman2004}, capturing some of the
-qualitative behavior of water.  However, many of these are less
-sophisticated than those describing the hard-sphere fluid, and a
-quantitatively accurate functional describing water---which reproduces
-both the macroscopic surface tension and its decrease as curvature
-increases---has not yet been published.  For instance, Jaqaman
-\emph{et al} developed a classical density-functional for water
-including orientational effects, but their functional resulted in a
-surface tension twice the experimental value, limiting it to purely
-qualitative predictions~\cite{Jaqaman2004}.
-
-\needsworknow{Add discussion and citations for more recent papers}
-
-A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
-describes water as non-interacting molecules in a an
-effective external potential.  This functional contains the exact
-entropy for no-interacting molecules and, among other things, can 
-predict the nonlinear dielectric response in water but must be paramaterized
-for different temperatures.
-
-\needsworknow{Add smooth transition from others' work to ours}
+A number of exact properties are easily achieved in the
+density-functional framework, such as the relationship between the
+pressure on a hard wall and the contact density---which ensures a
+correct solvation free energy for small hard solutes.  Much of the
+research on classical density-functional theory has focused on the
+hard-sphere fluid~\cite{curtin1985, rosenfeld1989, rosenfeld1993,
+  rosenfeld1997, tarazona1997, tarazona2000}.  This has led to a large
+number of very sophisticated functionals, such as the
+fundamental-measure theory (FMT) functional~\cite{rosenfeld1989,
+  rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000,
+  roth2002whitebear}.  The White Bear version of the FMT
+functional\cite{roth2002whitebear}  has a number of useful
+properties.  Like all FMT functionals, it is expressed as a function
+of a few convolutions of the density that can be efficiently
+computed.  In the homogeneous limit, it reduces to the
+Carnahan-Starling equation of state, and in the strongly-confined
+limit of a zero-dimensional cavity, it gives the exact free energy.
+
+A number of classical density functionals have been developed for
+water~\cite{ding1987, Yang1992, gloor2002saft, Jaqaman2004,
+  clark2006developing, lischner2010classical}\needsworknow{ADD REST OF
+  LIST HERE!}, capturing some of the qualitative behavior of water.
+However, many of these are less sophisticated than those describing
+the hard-sphere fluid, and a quantitatively accurate functional
+describing water---which reproduces the temperature dependence of the
+macroscopic surface tension, its decrease as curvature increases, and
+its quantitative value---has not yet been published.  For instance,
+the functional of Lischner \emph{et al}\cite{lischner2010classical}
+treats the surface tension correctly, but can only be used at room
+temperature.  A numer of classical density functionals have recently
+been produced that are based on Statistical Associating Fluid Theory
+(SAFT)\cite{gloor2002saft, clark2006developing}\needsworknow{add
+  more}.  These functionals are based on a perturbative thermodynamic
+expansion, and give a reasonable prediction for the temperature
+dependence of the surface tension.  However, they also overestimate
+the surface tension by \needsworknow{how much? significant factor, I
+  think}, and are unsuited for studying systems with molecular-scale
+inhomogeneities.
+
+%% A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
+%% describe water as non-interacting molecules in a an
+%% effective external potential.  This functional contains the exact
+%% entropy for no-interacting molecules and, among other things, can 
+%% predict the nonlinear dielectric response in water, but it must be paramaterized
+%% for different temperatures.
+
+\subsection{Statistical associating fluid theory}
+
+Statistical Associating Fluid Theory (SAFT) is a set of models for
+complex fluids in which hydrogren bonding plays a significant
+role\cite{muller2001molecular}.  SAFT is used to accurately model the
+equations of state of pure fluids and predict the equations of state
+for mixtures over a wide range of temperatures and pressures.  SAFT is
+based on Wertheim's first-order thermodynamic perturbation theory
+(TPT1) \cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV} which allows it to
+account for strong associative interations between molecules.
+
+The SAFT Helmholtz free energy is composed of five terms:
+\begin{equation}
+  F[n] = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
+  F_\textit{assoc} + F_\textit{chain},
+\end{equation}
+where the first three terms encompass the \emph{monomer} contribution
+to the free energy, the fourth is the \emph{association} free energy,
+describing hydrogen bonds, and the final term is the chain formation
+energy of for fluids that are chain polymers.  The final term is not
+relevant for water.  While a number of SAFT formulations have been
+published, we will focus on SAFT-VR\cite{gil-villegas-1997-SAFT-VR},
+which has been used by Clark \emph{et al} to construct an optimal SAFT
+model for water\cite{clark2006developing}.
+
+SAFT has also been used to construct classical density functionals,
+which have been applied to study the surface tension as a function of
+temperature\cite{whotocite}.  Such functionals have qualitatively
+predicted the dependence of surface tension on temperature, but have
+in general provided poor quantitative agreement with experiment.
+Moreover, the vast majority of SAFT-based classical density
+functionals have made local-density approximations that make them
+unsuitable for application to systems that are inhomogeneous at the
+molecular length scale\cite{felipe2001examination, gloor2002saft,
+  gloor2004accurate, clark2006developing, gloor2007prediction,
+  kahl2008modified, gross2009density}.
 
 \section{Theory and Methods}
-We use classical DFT to construct a new functional that will predict
-the density of water molecules in a classical system.  The model we use
-starts with a treatment of each individual water molecule as a purely
-classical hard sphere and adds association terms based on Statistical 
-Associating Fluid Theory (SAFT). Water is unlike simple
-fluids (which are mainly defined by van der Waals forces and weak electrostatic
-interations) because it requires direct treatment of association effects. The strong 
-hydrogen bonds and polar
-interactions in water are ideal for describing with SAFT, which was designed for
-use with hydrogen-bonding fluids. It is based on Wertheim's first-order
-thermodynamic perturbation theory (TPT1)
-\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
-wertheim1986fluidsIV} which provides a useful way to account for strong
-associative interations between molecules. SAFT can account for the association energies
-as well as any effects due to clusters or chains forming \fixme{not true}
-in the liquid\cite{muller2001molecular}.  \needsworknow{Expand this a
-  bit to accurately describe SAFT and its advantages.}
+We construct a new classical density functional for water.  We
+start with a treatment of each individual water molecule as a purely
 
@@
 
-The association functional we use is a modified version of that of Yu
-and Wu\cite{yu2002fmt-dft-inhomogeneous-associating} \footnote{We
+The association functional we use is a modified version of Yu
+and Wu.\cite{yu2002fmt-dft-inhomogeneous-associating} \footnote{We
   should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
@@
   association term and is not useful as a reference.} 
-which includes the effects of density inhomogeneities in the
-contact value of the correlation function $g^{HS}_\sigma$. This, however, is
-based on the SAFT-HS equation of state, rather than the SAFT-VR
-equation of state\cite{gil-villegas-1997-SAFT-VR} which is used in the
+This functional includes the effects of density inhomogeneities in the
+contact value of the correlation function $g^{HS}_\sigma$, but is
+based on the SAFT-HS equation of state rather than the SAFT-VR
+equation of state\cite{gil-villegas-1997-SAFT-VR}. SAFT-VR is used in the
 optimal SAFT parametrization for water found by Clark \emph{et
-  al}\cite{clark2006developing}.  The modification to use the SAFT-VR
-equation of state involves adding a correction term, see
-Equation~\ref{eq:gSW}.  \fixme{needs work}
+  al}\cite{clark2006developing}, but to use it's
+equation of state involves adding a correction term (see
+Equation~\ref{eq:gSW}).
 
@@
 
-To determine the critical separation, we consider a macroscopic surface
+To determine the critical separation, we start by consider a macroscopic surface
 tension model where the free energy in the macroscopic limit is given by 
@@
 rods, the water on the sides of the 'stadium' configuration will bow inward 
-between the rods and the densty will still reduce to vapor where the rods 
-are closest to each other.
+between the rods and the density will reduce to vapor near the center
+point where the rods are closest to each other.
 
@@
 This is consistent with the calculation of the surface tension in
-Figure~\ref{figs:surface-tension} (72~mN/m) and the force per length
+Figure~\ref{fig:surface-tension} (72~mN/m) and the force per length
 from Figure ~\ref{fig:rods-energy-vs-distance} (144~mN/m) which
@@
 reach an asymptote determined by the effective surface tension (shown in 
-Figure~\ref{fig:energy-vs-diameter}). In figure
+Figure~\ref{fig:energy-vs-diameter}). In Figure
 ~\ref{fig:rods-energy-vs-distance}, we see that for small $d$ all the rods 
@@
 \end{center}
-\caption{ Free energy (also known as the potential of mean force) 
+\caption{ Free energy of hydradon 
 versus separation for four hydrophobic rods ranging in radius from
@@
 \ref{fig:four-rods-energy-vs-distance} with a slight disagreement of
-the predicted$2\gamma$ force per unit length.
+the predicted force per unit length.
 
 The 'rounded square' formed by the area of the four rods and the space
-inbetween them has a circumference estimated by
+between them has a circumference estimated by
 \begin{equation}
@@
 We again equate this circumference with the total circumference of
-only the four rods, which results in a critical separation of
+the four separate rods, which results in a critical separation of
 \begin{equation}
@@
 a 23\% difference, and better agreement for larger radii rods (between 0.5\% and 5.6\%).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% 
-% \subsection{Test-particle approach}
-% 
-% More interesting applications are possible when we move into three
-% dimensions using spherical coordinates.  The first (but not simplest)
-% thing to try is looking at the response of liquid water to a single
-% fixed water molecule.  This is referred to as the \emph{test particle}
-% method, and is as simple as constraining the density to
-% include a delta function at the origin---meaning that a single
-% molecule is fixed in location.  The free energy is then minimized
-% under this constraint, giving a density profile and an associated free
-% energy.  The density profile is just the self-correlation function
-% multiplied by the bulk liquid density.  \fixme{We should be
-%   distinguishing between the properties of the \emph{exact functional}
-%   and those of an approximate functional...}  Thus, a test-particle
-% calculation provides a stringent test of the correctness of any
-% approximate functional.
-% 
-% In order to perform the test-particle test, we introduce a sharp
-% gaussian density spike at the origin and relax the density under a
-% constraint that this spike remain.  Figure~\ref{fig:test-particle}
-% shows the resulting density profile, compared with the experimental
-% $g_{\textrm{OO}}$.  Note that this data was \emph{not} used in the
-% fitting of the empirical parameters of the functional.  The agreement
-% between experiment and theory is both qualitatively and quantitatively
-% poor.  The integral of the difference between this density profile
-% (excluding the delta function) and the bulk liquid density should be
-% 0.06~molecules \fixme{look up and show derivation... or cut this},
-% but we find instead to be $-0.15$~molecules.  In addition, the associated
-% free energy should be positive: there is an entropic cost in fixing
-% one molecule in position, but for our functional the free energy of
-% this density configuration is in fact negative, with a value of
-% -3.5~eV, meaning that the homogeneous fluid is unstable.  This is a
-% serious flaw that must be addressed in future work.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 A common model of hydrophobic solutes is the \emph{hard sphere solute}
-\fixme{add citations?}.  The effect of this solute is merely to create
+\cite{sedlmeier2011entropy}.  The effect of this solute is merely to create
 a cavity within which the water may not penetrate.  This is the

### 2012-05-05 papers/water-SAFT/paper.tex

--- 2012-04-30 papers/water-SAFT/paper.tex
@@
 A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
-describes water as non-interacting molecules in a an
+describe water as non-interacting molecules in a an
 effective external potential.  This functional contains the exact
 entropy for no-interacting molecules and, among other things, can 
-predict the nonlinear dielectric response in water but must be paramaterized
+predict the nonlinear dielectric response in water, but it must be paramaterized
 for different temperatures.
@@
 \needsworknow{Add smooth transition from others' work to ours}
+Using classical DFT, we aim to find an equation of state for water
+that describes interactions qualitatively and gives accurate
+quantitative results within a reasonable temperature range without
+needing extensive paramaterization.
 
@@
 as well as any effects due to clusters or chains forming \fixme{not true}
-in the liquid\cite{muller2001molecular}.  \needsworknow{Expand this a
+in the liquid\cite{muller2001molecular}. SAFT assumes a repulsive hard
+sphere core for each atom with a number of short-ranged off center
+attractive sites. This model applied to give equation of state for
+various systmes and can be applied to inhomogeneous fluids.
+\needsworknow{Expand this a
   bit to accurately describe SAFT and its advantages.}

### 2012-05-09 papers/water-SAFT/paper.tex
Removed water-vapor density plot and moved written section to empirical parameters. Fixed directory paths in papers/water-SAFT/figs/ Tested everything on quipu.

--- 2012-05-05 papers/water-SAFT/paper.tex
@@
 
-% needsworknow is intended to prioritize stuff that needs fixing.
-\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}
+% fixme is intended to prioritize stuff that needs fixing.
+\newcommand{\fixme}[1]{\textcolor{red}{[\emph{#1}]}}
 
@@
 water~\cite{ding1987, Yang1992, gloor2002saft, Jaqaman2004,
-  clark2006developing, lischner2010classical}\needsworknow{ADD REST OF
+  clark2006developing, lischner2010classical}\fixme{ADD REST OF
   LIST HERE!}, capturing some of the qualitative behavior of water.
@@
 been produced that are based on Statistical Associating Fluid Theory
-(SAFT)\cite{gloor2002saft, clark2006developing}\needsworknow{add
+(SAFT)\cite{gloor2002saft, clark2006developing}\fixme{add
   more}.  These functionals are based on a perturbative thermodynamic
@@
 dependence of the surface tension.  However, they also overestimate
-the surface tension by \needsworknow{how much? significant factor, I
+the surface tension by \fixme{how much? significant factor, I
   think}, and are unsuited for studying systems with molecular-scale
@@
 start with a treatment of each individual water molecule as a purely
+classical hard sphere and add association terms based on Statistical 
+Associating Fluid Theory (SAFT). Water is unlike simple
+fluids (which are mainly defined by van der Waals forces and weak electrostatic
+interations) because it requires direct treatment of association effects. The strong 
+hydrogen bonds and polar
+interactions in water are ideal for describing with SAFT, which was designed for
+use with hydrogen-bonding fluids. It is based on Wertheim's first-order
+thermodynamic perturbation theory (TPT1)
+\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
+wertheim1986fluidsIV} which provides a useful way to account for strong
+associative interations between molecules. SAFT can account for the association energies
+as well as any effects due to clusters or chains forming \fixme{not true}
+in the liquid\cite{muller2001molecular}. SAFT assumes a repulsive hard
+sphere core for each atom with a number of short-ranged off center
+attractive sites. This model can be applied to give an equation of state for
+various systems even inhomogeneous fluids.
+\fixme{Expand this a
+  bit to accurately describe SAFT and its advantages.}
 
@@
 \begin{equation}\label{contactvaluethm}
-  F = n_c k_BT V\:.
-\end{equation}
+  F = n k_BT V\:.
+\end{equation}\fixme{may be ambiguous about n}
 These properties are retained by the total functional,
@@
 %\caption{The theoretical versus experimental coexistence curve. Away from
-%the critical temperature, the agreement is good.\needsworknow{(keep figure?
+%the critical temperature, the agreement is good.\fixme{(keep figure?
 %its the same as Clark paper)} }
@@
 
-\subsection{Liquid-vapor interface}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{The liquid-vapor density profile at room temperature shows
-some small oscillations close to the interface.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
-
-We set up a liquid-vapor
-interface and plot the density profile.
 The fluid density at a liquid-solid interface commonly displays
@@
 density-functional theory would include them), with the result
-that the density profile will be smoothed out by capillary wave effects. 
-The liquid-vapor
-interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
-some small oscillations.
+that the density profile will be smoothed out by capillary wave
+effects. We set up a liquid-vapor interface and observed small density
+oscillations in the water near the surface.
 
@@
 rod has comparatively larger oscillations and a higher contact density.
-\needsworknow{Expand or add to this description of density profiles and
+\fixme{Expand or add to this description of density profiles and
 oscillations for single rod?}

### 2012-05-12 papers/contact/paper.tex
Fix bug in ContactDensity (the "Mysterious Four") and a few minor paper changes Tests skipped!
clean up appendices of contact paper a bit Tested everything on bingley.

@@
 expand outward from its center and all directions will be covered uniformly.
+See Figure \ref{fig:contact}.
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/contact}
+\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+  consist of those which just touch the point $\mathbf{r}$.}
+\label{fig:contact}
+\end{figure}
 
@@
 
-In the case of A, the changing radius we refer to is the radius of a sphere that is centered at $\mathbf{r}$.
-The $n_3$ values that will be affected are those which are a distance $R$ from $\mathbf{r}$.
-
+The derivatation of the asymmetrically averaged correlation function
+$g_\sigma^A(\rr)$ requires that we express the free energy in terms of
+a set of weighted densities in which the radius of the weighting
+kernel is determined according to the position of the sphere in
+question---which is to say, the location where the density itself is
+evaluated.  This is the most natural way to construct such a
+functional, in which the radius of spheres is a function of position.
+Thus we define $n_3^A$ as
 \begin{align}
-  n_3^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}'')) d\mathbf{r}''\\ 
-  \frac{\delta n_3^{A} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) \delta(\mathbf{r}-\mathbf{r}'') d\mathbf{r}'' \\
-   &= n (\mathbf{r}) \delta(|\mathbf{r'} - \mathbf{r}| - R(\mathbf{r}))
+  n_3^{A}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr')) d\rr'
 \end{align}
-
-And likewise:
+and we define the other weighted densities similarly.  In order to
+compute derivatives of the free energy with respect to $R(\rr)$, we
+need to find $g_\sigma^A$.  In the case of $n_3^A$, we needS
 \begin{align}
-  n_2^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) d \mathbf{r}''\\
-  \frac{\delta n_2^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))\\
+  \frac{\delta n_3^{A} (\rr')}{\delta R(\rr)} &=
+  \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'')) \delta(\rr-\rr'') d\rr'' \\
+   &= n (\rr) \delta(|\rr' - \rr| - R(\rr))
 \end{align}
-
-where the trouble here is that the derivative of the delta function,
-$\delta'$, isn't something we really like to deal with.  Particularly
-as we're looking  at a 3D integral involving a 1D delta function,
-which means there are Jacobian factors to be considered.  On the other
-hand, there's no reason why we couldn't convolve with this kernel as
-efficiently as we do with any other kernel.
-\\
-\\
+Similarly, for $n_2^A$, we find
+\begin{align}
+  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - R(\rr'')) d \rr''\\
+  \frac{\delta n_2^{A}(\rr')}{\delta R(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - R(\rr))\\
+\end{align}
+This is somewhat more challenging, since it involves an unfamiliar
+function, the derivative the Dirac delta function $\delta'$.
+%
 The derivatives of the remaining scalar densities can be reduced to
@@
 \begin{align}
-  n_1(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
-  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
-  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})}
-  &= -\frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')}
+  \delta(|\rr'-\rr''| - R(\rr'')) \\
+  \frac{\delta n_1(\rr')}{\delta R(\rr)}
+  &= -\frac{n(\rr)}{4\pi
+    R(\rr)}\delta'(|\rr'-\rr| - R(\rr))
   -
-  \frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+  \frac{n(\rr)}{4\pi
+    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr))
+\end{align}
+and
+\begin{align}
+  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')^2}
+  \delta(|\rr'-\rr''| - R(\rr'')) \\
+  \frac{\delta n_1(\rr')}{\delta R(\rr)^2}
+  &= -\frac{n(\rr)}{4\pi
+    R(\rr)^2}\delta'(|\rr'-\rr| - R(\rr))
+  -
+  2\frac{n(\rr)}{4\pi
+    R(\rr)^3}\delta(|\rr'-\rr| - R(\rr))
+\end{align}
+The vector-weighted densities have no additional occurrences of $R$,
+their functional derivatives come out similarly:
+\begin{align}
+  \mathbf{n}_{V2}^{A}(\rr) &= \int n(\rr') \delta(|\rr' - \rr''| - R(\rr''))
+    \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
+  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta R(\rr)} &= - n(\rr) \delta'(|\rr' - \rr| - R(\rr))
+    \frac{\rr'-\rr}{|\rr'-\rr|}\\
+\end{align}
+and analogously for $\mathbf{n}_{V1}^A$:
+\begin{align}
+  \mathbf{n}_{V1}^A(\rr) &= \int d\rr' \frac{n(\rr')}{4\pi R(\rr')}
+  \delta(|\mathbf{r}-\rr'| - R(\rr')) \frac{\rr-\rr'}{|\rr-\rr'|}\\
+  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta R(\rr)}
+  &= -\frac{n(\rr)}{4\pi
+    R(\rr)}\delta'(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
+  -
+  \frac{n(\rr)}{4\pi
+    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
+\end{align}
+\fixme{double-check the directions of all the vector quantities...}
+
+%% So, what is this convolution with a $\delta'$? Let's look at the
+%% scalar version:
+%% \fixme{With following the primes are certainly not in the right places}
+%% \begin{align}
+%%   \bar{f}(\mathbf{r})
+%%   &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
+%%   \\
+%%   &= \int
+%%         f(\mathbf{r'})
+%%         \lim_{\Delta R \rightarrow 0}
+%%           \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+%%             -
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%%       }{\Delta R}
+%%   \\
+%%   &= \lim_{\Delta R \rightarrow 0}
+%%           \frac{\int
+%%             f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
+%%             -
+%%             \int
+%%             f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%%       }{\Delta R}
+%%   \\
+%%   &\approx\lim_{\Delta R \rightarrow 0}
+%%           \frac{\int
+%%             \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%%             -
+%%             \int
+%%             f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%%       }{\Delta R}
+%%   \\
+%%   &\approx\lim_{\Delta R \rightarrow 0}
+%%           \frac{2\frac{\Delta R}{R} \int
+%%             f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%%       }{\Delta R}
+%%   \\
+%%   &= \frac{2}{R} \int
+%%             f(\mathbf{r'})
+%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
+%% \end{align}
+%% There is a sign ambiguity here with regard to whether we are asking
+%% about the derivative of this delta function with regard to $R$ or
+%% something else.  We just need to get it right.
+
+This completes our exploration of the partial derivatives of weighted
+densities of the $A$ variety.
+
+\subsection{Expressions for $\beta\frac{\delta
+    n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the
+  symetric case}\label{appendix:g-S}
+
+The derivatation of the asymmetrically averaged correlation function
+$g_\sigma^S(\rr)$ requires that we express the free energy in terms of
+a set of weighted densities in which the radius of the weighting
+kernel is determined according to the position at which the weighted
+density is being evaluated.  This is less intuitive than the $A$ case,
+since \emph{nothing is there}!  However, in practice it corresponds to
+the picture shown in Figure~\ref{fig:contact}, which involves
+extending the radius for spheres just at the position touching a
+given spot, where we are evaluating the correlation function.  We
+define the weighted density $n_3^S$ as:
+\begin{align}
+  n_3^{S}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr)) d\rr'
+\end{align}
+In the symmetric case, the functional derivative is strikingly simple:
+\begin{align}
+  \frac{\delta n_3^{S} (\rr')}{\delta R(\rr)} &=
+   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'))
+   \delta(\rr'-\rr) d\rr''
+   \\
+   &= \delta(\rr'-\rr) n_2^S(\rr')
+\end{align}
+The functional derivative does \emph{not} eliminate the integral,
+which simply changes from one weighted density to another.  The
+corresponding term in the free energy functional, once integrated,
+takes a simple form:
+\begin{align}
+  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
+  \frac{\delta n_3^{S}(\rr')}{\delta R(\rr)} d\rr'
+  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
+  n_2^S(\rr') d\rr' \\
+  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
+\end{align}
+which means that we can compute $g_\sigma^S(\rr)$ using only
+convolutions of the density, in strong contrast with
+$g_\sigma^A(\rr)$, which requires convolutions of derivatives of the
+free energy density.
+
+The functional derivative of $n_2^S(\rr')$ is quite similar:
+\begin{align}
+  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| + R(\rr))d\mathbf
+  r'\\
+  \frac{\delta n_2^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
+  \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+\end{align}
+where we recognize that the convolution---while not one of the standard
+FMT weighted density---is a simple convolution of the density.  We can
+go on to compute the derivatives of the remaining weighted densities:
+\begin{align}
+  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
+    \frac{\delta n_1^{S}(\rr')}{\delta R(\rr)} &=
+    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)}
+    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+    + n_0^s(\rr') \right)
 \end{align}
 \begin{align}
-  n_0(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})^2}
-  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \\
-  \frac{\delta n_1(\mathbf{r}')}{\delta R(\mathbf{r})^2}
-  &= -\frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})^2}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
-  -
-  2\frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})^3}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r}))
+  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
+    \frac{\delta n_0^{S}(\rr')}{\delta R(\rr)}
+    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)^2}
+    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+    +
+    \frac{2}{R} n_0^S(\rr') \right)
+\end{align}
+\begin{align}
+  \mathbf{n}_{V2}^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))
+  \frac{(\rr - \rr')}{|\rr - \rr'|} d \rr'\\
+  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= \delta(\rr - \rr')
+  \int n(\rr'') \delta'(|\rr' - \rr''| - R(\rr'))
+  \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
+\end{align}
+\begin{align}
+ \mathbf{n}_{V1}^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'
+   \frac{(\rr - \rr')}{|\rr - \rr'|}\\
+ \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta R(\rr)}
+ &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi R(\rr)}
+   \delta'(|\rr'-\rr''| - R(\rr')) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
+   + \mathbf{n}_{V2}^S(\rr') \right)
 \end{align}
 
-Since the vector-weighted densities have no additional occurrences of
-$R$, their derivatives come out similarly.
-\begin{align}
-  \mathbf{n}_{V2}^{A}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'')) 
-    \frac{\mathbf{r'}-\mathbf{r''}}{|\mathbf{r'}-\mathbf{r''}|} d \mathbf{r}''\\
-  \frac{\delta \mathbf{n}_{V2}^{A}(\mathbf{r}')}{\delta R(\mathbf{r})} &= - n(\mathbf{r}) \delta'(|\mathbf{r}' - \mathbf{r}| - R(\mathbf{r}))
-    \frac{\mathbf{r}'-\mathbf{r}}{|\mathbf{r}'-\mathbf{r}|}\\
-\end{align}
+The above equations involve just two additional weighted densities
+beyond those required for FMT: one scalar density computed with a
+derivative of the Dirac delta function, and one analogous vector
+density.
 
-and analogously for $\mathbf{n}_{V1}$:
-
-\begin{align}
-  \mathbf{n}_{V1}(\mathbf{r}') &= \int \mathbf{dr''} \frac{n(\mathbf{r''})}{4\pi R(\mathbf{r''})}
-  \delta(|\mathbf{r'}-\mathbf{r''}| - R(\mathbf{r''})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}\\
-  \frac{\delta \mathbf{n}_{V1}(\mathbf{r}')}{\delta R(\mathbf{r})}
-  &= -\frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})}\delta'(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}
-  -
-  \frac{n(\mathbf{r})}{4\pi
-    R(\mathbf{r})^2}\delta(|\mathbf{r'}-\mathbf{r}| - R(\mathbf{r})) \frac{\mathbf{r'}-\mathbf{r}}{|\mathbf{r'}-\mathbf{r}|}  
-\end{align}
-
-So, what is this convolution with a $\delta'$? Let's look at the
-scalar version:
-\fixme{With following the primes are certainly not in the right places}
-\begin{align}
-  \bar{f}(\mathbf{r})
-  &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
-  \\
-  &= \int
-        f(\mathbf{r'})
-        \lim_{\Delta R \rightarrow 0}
-          \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-            -
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &= \lim_{\Delta R \rightarrow 0}
-          \frac{\int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-            -
-            \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &\approx\lim_{\Delta R \rightarrow 0}
-          \frac{\int
-            \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-            -
-            \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &\approx\lim_{\Delta R \rightarrow 0}
-          \frac{2\frac{\Delta R}{R} \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-      }{\Delta R}
-  \\
-  &= \frac{2}{R} \int
-            f(\mathbf{r'})
-            \delta(|\mathbf{r}-\mathbf{r'}| - R)
-\end{align}
-There is a sign ambiguity here with regard to whether we are asking
-about the derivative of this delta function with regard to $R$ or
-something else.  We just need to get it right.
-
-\subsection{Expressions for $\beta\frac{\delta n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the symetric case}\label{appendix:g-S}
-
-For the derivation of $n_{3}$ in the symmetric case:
-
-\begin{align}  
-  n_3^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \Theta(\left|\mathbf{r}' - \mathbf{r}''\right| -R(\mathbf{r}')) d\mathbf{r}''\\
-  \frac{\delta n_3^{S} (\mathbf{r}')}{\delta R(\mathbf{r})} &=
-  \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}''
-\end{align}
-
-\fixme{Not sure if we need any more explanation here (probably want better than this) -  $\delta R(\mathbf{r})$ in this equation refers to an expansion, in the fashion described above, of all the spheres that just touch $\mathbf{r}$.
-We can see that if the radii change in this fashion, the only point at which $n_3(\mathbf{r}')$ will be affected is at $\mathbf{r}' = \mathbf{r}$, so that $\delta n_3(\mathbf{r}')$ is zero everywhere except where $\mathbf{r}' = \mathbf{r}$.
-Thus we have the 2nd $\delta$ function in the lower integral.}
-\\
-\\
-This term in the free energy functional, once integrated, takes an interesting form:
-
-\begin{align}
-  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} 
-  \frac{\delta n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}' 
-  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}'-\mathbf{r}) d\mathbf{r}'' d\mathbf{r}' \\
-  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}')} \int n (\mathbf{r}'') \delta(|\mathbf{r} - \mathbf{r}''| - R(\mathbf{r})) d\mathbf{r}''\\
-  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\mathbf{r}'')} n_2(\mathbf{r})\\
-\end{align}
-
- $n_2(\mathbf{r}')$ in the case of S in calculated similarly
-\begin{align}
-  n_2^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
-    \frac{\delta n_2^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int n(\mathbf{r}'') 
-  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}''
-\end{align}
-
-And similarly to the A case we compute
-\begin{align}
-  n_1^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
-    \frac{\delta n_1^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
-  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
-  -
-  \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
-\end{align}
-
-and 
-\begin{align}
-  n_0^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''\\
-    \frac{\delta n_0^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} 
-  \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' 
-  -
-  2\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^3} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
-\end{align}
-
-\begin{align}
-  \mathbf{n}_{V2}^{S}(\mathbf{r}') &= \int n(\mathbf{r}'') \delta(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}'))
-  \frac{(\mathbf{r}' - \mathbf{r}'')}{|\mathbf{r}' - \mathbf{r}''|} d \mathbf{r}''\\
-  \frac{\delta \mathbf{n}_{V2}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= \delta(\mathbf{r} - \mathbf{r}')
-  \int n(\mathbf{r}'') \delta'(|\mathbf{r}' - \mathbf{r}''| - R(\mathbf{r}')) 
-  \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} d\mathbf{r}''
-\end{align}
-
-\begin{align}
- \mathbf{n}_{V1}^{S}(\mathbf{r}') &= \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r''
-   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}\\
- \frac{\delta \mathbf{n}_{V1}^{S}(\mathbf{r}')}{\delta R(\mathbf{r})} &= -\int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})} 
-   \delta'(|\mathbf{r}'-\mathbf{r}''| - R(\mathbf{r}')) \delta(\mathbf{r}-\mathbf{r}') d\mathbf{r}'' \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|} 
-   -
-   \int \frac{n(\mathbf{r}'')}{4\pi R(\mathbf{r})^2} \delta(|\mathbf{r}' - \mathbf{r}''| + R(\mathbf{r}'))d\mathbf r'' 
-   \frac{(\mathbf{r} - \mathbf{r}'')}{|\mathbf{r} - \mathbf{r}''|}
-\end{align}
-
-\begin{widetext}
 \section{Formulas generated from Haskell}
@@
 
-\fixme{This is the hypthetical end of the paper}
-
- \fixme {
-And then there's a lot with taking the derivative of the delta for the case of $n_{2}^{S}(\mathbf{r}')$.
-Not sure if we'll use any of this, but wrote down some equations that Roundy wrote out.
-So can use the terms at least if going to write about this - I'm just copying it in for now (probably a lotof mistakes)}
-
-\begin{align}
-  p(\mathbf{r}) &= \frac{\delta A}{\delta n_2(\mathbf{r}')} \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})} d\mathbf{r}'\\
-  &= \frac{\delta A}{\delta n_2(\mathbf{r})} - \int n(\mathbf{r}'') \delta' (|\mathbf{r} - \mathbf{r}''| - R) d\mathbf{r}''
-  \int n (\mathbf{r} +\mathbf{r}'') \delta'(|\mathbf{r}'''| - R) d \mathbf{r}'''\\
-  &= \int n(\mathbf{r} + \mathbf{r}''') \delta'(|\mathbf{r}'''| - R) r'''^2 cos\theta d \theta''' dr'''\\
-  du = \delta'(\mathbf{r}''' - R) d \mathbf{r}'''     v = n(\mathbf{r} - \mathbf{r}'')r'''\\
-   u = \delta(\mathbf{r}''' - R)                       
-   dv = \mathbf{\nabla} n(\mathbf{r} + \mathbf{r}'') \cdot \frac{\mathbf{r}'''}{r'''} + 2n(\mathbf{r}' - \mathbf{r}'')r'''\\
-  &= -\int(\mathbf{\nabla}n(\mathbf{r}' + \mathbf{r}'') \cdot \mathbf r'''(hat)r'''^2 dr''' 
-  + 2n(\mathbf{r} + \mathbf{r}''')r''') dr''' \delta(\mathbf{r} - R) dcos\theta''' d\theta''\\
-  &= - \int \delta(\mathbf{r}''' - R) d\mathbf{r}''' 
-  (\mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) + \frac{2n(\mathbf{r} + \mathbf{r}'')}{\mathbf{r}'''})\\
-  &= \frac{2n_2(\mathbf{r})}{R} + 
-  \int \delta(|\mathbf{r}'''| - R) \mathbf{\nabla}n(\mathbf{r} + \mathbf{r}''') \cdot \mathbf{r}''' (hat) d \mathbf{r}'''\\
-\end{align}
-
-\end{widetext}
 \end{widetext}

### 2012-05-15 papers/water-SAFT/paper.tex
Added density_calc.py and data file.  Made minor changes to paper. Fixed error due to assert in src/lattice.cpp by declaring vol0. Tested everything on quipu.

@@
 over-structuring\cite{lin2009importance, wang2011density,
-  mogelhoj2011ab, jonchiere2011van}.  It has also been bound that
-inclusion of nuclear quantum effects can have a similar
-effect\cite{morrone2008nuclear}.  Each of these corrections imposes an
+  mogelhoj2011ab, jonchiere2011van}.  It has also been found that
+inclusion of nuclear quantum effects can provide similar improvements
+\cite{morrone2008nuclear}.  Each of these corrections imposes an
 additional computational burden on an approach that is already
-feasible only for very small number of water molecules. A more
+feasible only for a very small number of water molecules. A more
 efficient approach is needed in order to study nanoscale and
@@
 hydrogen-bonding network, water as ``the universal solvent'' is as
-interesting as its anomalous thermodynamic properties.
+interesting as its anomalous thermodynamic properties.  \fixme{what
+  the heck?}
 
@@
 \section{Theory and Methods}
-We construct a new classical density functional for water.  We
-start with a treatment of each individual water molecule as a purely
-classical hard sphere and add association terms based on Statistical 
-Associating Fluid Theory (SAFT). Water is unlike simple
-fluids (which are mainly defined by van der Waals forces and weak electrostatic
-interations) because it requires direct treatment of association effects. The strong 
+We construct a new classical density functional for water by treating
+each individual water molecule as a purely
+classical hard sphere and adding association terms based on SAFT. Water
+is unlike simple fluids (which are mainly defined by van der Waals
+forces and weak electrostatic interations) because it requires direct
+treatment of association effects. The strong 
 hydrogen bonds and polar
 interactions in water are ideal for describing with SAFT, which was designed for
-use with hydrogen-bonding fluids. It is based on Wertheim's first-order
-thermodynamic perturbation theory (TPT1)
-\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
-wertheim1986fluidsIV} which provides a useful way to account for strong
-associative interations between molecules. SAFT can account for the association energies
-as well as any effects due to clusters or chains forming \fixme{not true}
-in the liquid\cite{muller2001molecular}. SAFT assumes a repulsive hard
-sphere core for each atom with a number of short-ranged off center
-attractive sites. This model can be applied to give an equation of state for
-various systems even inhomogeneous fluids.
-\fixme{Expand this a
-  bit to accurately describe SAFT and its advantages.}
+use with hydrogen-bonding fluids.
 

### 2012-05-19 papers/contact/paper.tex
clean up description of contact functionals Tested everything on bingley.

@@
   + \rr')
-  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \\
+  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \label{eq:gS} \\
   n_0(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
@@
 of Fundamental Measure Theory (FMT), and seems ideal for treating
-touching spheres.
+touching spheres, as illustrated in Figure~\ref{fig:n0}.
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/n0}
+\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+  consist of those which just touch the point $\mathbf{r}$.}
+\label{fig:n0}
+\end{figure}
 
@@
   \int n^{(2)}(\rr, \rr + \rr')
-  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \\
+  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \label{eq:gA} \\
   n_A(\rr) &= \int n(\rr')
-  \frac{\delta(|\rr-\rr'|-\sigma)}{4\pi\sigma^2} d\rr'
+  \frac{\delta(|\rr-\rr'|-\sigma)}{4\pi\sigma^2} d\rr' \label{eq:nA}
 \end{align}
@@
   n_\textit{contact} &= n \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
-  g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3}
+  g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3} \label{eq:cs-g}
 \end{align}
@@
 position, we can compute the pressure on the hard spheres located at
-$\mathbf{r}$.
+$\mathbf{r}$.  Since the hard-sphere radius enters the FMT free energy
+by means of the fundamental measures, this requires that we adjust the
+convolutions to have a radius $R$ which depends on the location of the
+sphere:
+\begin{align}
+  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr'))d\rr' \label{eq:n3A}
+\end{align}
+Once we have done this---and similarly for the other fundamental
+measures---we derive a formula for the correlation function
+$g_\sigma^A(\rr)$ by finding the pressure on the surface of any hard
+spheres located at position $\rr$, and using the contact-density
+theorem to extract the contact density, from which we can establish
+the averaged correlation function at contact.
 \begin{align}
@@
   g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
-  g_\sigma^A(\rr) &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
     A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
 \end{align}
-Equation~\ref{eq:g-A-exact} is an exact expression for $g_\sigma^A$
-given the exact functional $A_{HS}$.  In practice, we use the
-approximate White Bear functional, and find an approximate expression
-for $g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
-expression will involve more convolutions than are present in the free
-energy itself, making this formulation relatively expensive.
+The functional derivative $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$
+in the above formula must be computed with the asymmetric formulation
+for the fundamental measures.  In practice, we use the approximate
+White Bear functional to find an approximate expression for
+$g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
+expression will involve convolutions of derivatives of the free
+energy, making this formulation computationally more expensive than
+the free energy itself.
 
@@
 
-\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
-We compare with the averaged correlation function of
-Gross\cite{gross2009density}, which is of the asymmetrical variety
-($g_\sigma^A$), where the correlation function is averaged at location
-of one of the spheres in concact.
-\begin{align}
-  g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
-    n_{3'}\right)^3} \\
-  n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
-\end{align}
-where $n_{3'}$ is an averaged
-filling fraction defined in a manner analogous to the $n_3$ of FMT,
-but with the hard-sphere radius replaced by the hard-sphere diameter,
-so the integrals range over spheres that are twice as large.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:n0}
-\end{figure}
-
 \section{Symmetrically averaged correlation function}\label{sec:g-S}
 
-An alternate definition of the averaged correlation function for an
-inhomogeneous system would be the correlation function averaged for
-spheres \emph{touching at a given point}.  This would be a symmetric
+We now address the symmetrically averaged correlation function, which
+is defined in Equation~\ref{eq:gS}.  This corresponds to the
+correlation function averaged for
+spheres \emph{touching at a given point}.  It is a symmetric
 measure, since both spheres involved are treated in the same way.  In
@@
 to define a corresponding correlation function $g_\sigma^S$.  This is
-\emph{almost} what is done by Yu and  Wu
+\emph{almost} what is done by Yu and  Wu in
+reference\cite{yu2002fmt-dft-inhomogeneous-associating}, as will be
+discussed in Section~\ref{sec:yuwu}.
 
-\fixme{Analogous to the assymetric case is the symmetric, for which 
-we are no longer able to simply pick a hypothetical sphere centered 
-at position $r$ and expand its radius.  Instead we must formulate a way 
-to expand the radii of spheres that just touch a specific point of 
-contact and only expand them at that point of contact, in a symmetric 
-fashion.  Therefore, when we compute the neccasary derivitives wih 
-respect to $R$', we expand a multitude of spheres, all of them positioned 
-around and just touching the point $\mathbf{r}$.  When their radii 
-expand, each does not expand outward in all directions as in the 
-case of A, rather they each expand outward at only the point $\mathbf{r}$.
-One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains
-at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
-in this peculiar fashion, we have in the end the same amount of shell 
-area protruding outward as we would if we had simply expanded one 
-sphere in fashion of case A.  Also the protrusions, once added up 
-for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
-in their direction of protrusion.  Each protruding sphere bit will 
-expand outward from its center and all directions will be covered uniformly.
-See Figure \ref{fig:contact}.
-
+The procedure for finding the symmetrically averaged correlation
+function $g_\sigma^S(\rr)$ is less straight-forward than 
+\begin{align}
+  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr))d\rr'
+\end{align}
+The hard-sphere radius $R$ in this weighted density has an unusual
+position dependence:  it depends on the position at which we are
+evaluating the weighted density, rather than the position of the hard
+spheres themselves.  This has the effect of expanding spheres as they
+touch the point $\rr$, a concept illustrated in
+Figure~\ref{fig:contact}.  This enables a similar derivation to that
+which gave us $g_\sigma^A(\rr)$:
 \begin{figure}
@@
 \end{figure}
-
-These qualitative concepts lead us to derivations for the $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ term in our derivation of the pressure, eq \ref{eq:p_{HS}}.
-The rest of the derivation results when we recognize that instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ as in the assymmetric case, we will instead divide by $n_0(\mathbf{r})^2$ since we have symmetry around the point $\mathbf{r}$.
-This will yield our symmetric correlation, $g_\sigma^S$:
-
 \begin{align}
@@
   g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
     A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
-\end{align}}Complete expressions for $g_\sigma^S$ are included in
-Appendix~\ref{appendix:g-S}.\\
+\end{align}
+where the functional derivative $\frac{\delta A_{HS}}{\delta
+  R(\mathbf{r})}$ in this case is computed with the symmetric
+formulation of the fundamental measures.  Complete expressions for
+$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\
 
-\section{Yu and Wu's functional}\label{sec:yuwu}
+\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+One existing approximation for the ccorrelation function is that of
+Gross\cite{gross2009density}, which is of the asymmetrically averaged
+variety ($g_\sigma^A$):
+\begin{align}
+  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{2\pi}{3}R^3n_A(\rr)}{\left(1 -
+    \frac{4\pi}{3}R^3n_A(\rr)\right)^3}
+\end{align}
+where $n_A$ is the averaged density defined in Equation~\ref{eq:nA}.
+This formula is arrived at by using the density averaged over all
+spheres that could be touching a sphere at point $\rr$ in the
+Carnahan-Starling equation for the correlation function at contact,
+given in Equation~\ref{eq:cs-g}.
 
+\section{Yu and Wu's symmetrically averaged functional}\label{sec:yuwu}
+
+\fixme{I (DJR) haven't yet edited this section!}
 Yu and Wu developed a functional that should correspond to the contact

### 2012-05-21 papers/water-SAFT/paper.tex

@@
 over-structuring\cite{lin2009importance, wang2011density,
-  mogelhoj2011ab, jonchiere2011van}.  It has also been found that
+  mogelhoj2011ab, jonchiere2011van}.  It has also been found that the
 inclusion of nuclear quantum effects can provide similar improvements
@@
 additional computational burden on an approach that is already
-feasible only for a very small number of water molecules. A more
+feasible for only a very small number of water molecules. A more
 efficient approach is needed in order to study nanoscale and
@@
 Numerous approaches have been developed to approximate the effect of water
-as a solvent in atomistic calculations.  Each of these approaches gives
+as a solvent in atomistic calculations.  Each of these approaches gives an
 adequate description of some aspect of interactions with water, but none of
@@
 \section{Theory and Methods}
-We construct a new classical density functional for water by treating
-each individual water molecule as a purely
-classical hard sphere and adding association terms based on SAFT. Water
-is unlike simple fluids (which are mainly defined by van der Waals
-forces and weak electrostatic interations) because it requires direct
-treatment of association effects. The strong 
-hydrogen bonds and polar
-interactions in water are ideal for describing with SAFT, which was designed for
-use with hydrogen-bonding fluids.
-
-The Helmholtz free energy functional is based on SAFT,
-and is composed of the terms:
-\begin{equation}
-  F[n] = F_\textit{id}[n] + F_\textit{hs}[n]  +
-F_\textit{disp}[n]+ F_\textit{assoc}[n],
-\end{equation}
-where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
-the hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
-and $F_\textit{assoc}$ is the free energy of association.  In
+We construct a new classical density functional for water. We treat
+each individual water molecule as a purely classical hard sphere and
+add association terms based on SAFT. For our free energy, we use the
+first four terms from equation (number): $F_\textit{id}$,
+$F_\textit{hs}$, $F_\textit{disp}$ and $F_\textit{assoc}$. In
 addition, a chemical potential term is used, so we actually work in
@@
 \begin{equation}\label{contactvaluethm}
-  F = n k_BT V\:.
-\end{equation}\fixme{may be ambiguous about n}
+  F = n_c k_BT V\:
+\end{equation}
 These properties are retained by the total functional,
@@
 \caption{ Density profiles for a single rod of different radii. The dotted line
-represents the saturated liquid density.  }
+represents the saturated liquid density and the points represent the
+expected contact density derived from the contact value theorem and
+data used in Figure \ref{fig:energy-vs-diameter}.}
 \label{fig:density-single-rod}
@@
 expected oscillations outside the rod that reduce to the bulk saturated liquid
-density as the distance away from the rods increases. The small radius
-rod has comparatively larger oscillations and a higher contact density.
-\fixme{Expand or add to this description of density profiles and
-oscillations for single rod?}
+density as the distance away from the rods increases. We use the
+contact value theorem to derive the contact density at the water-rod
+interface.  We start with the thermodyamic expression relating
+pressure to the partial derivative of free energy with respect the
+radius of a cylinder. Using the contact value theorem and sustituting in
+the free energy gradient for the pressure, we can derive values for
+the contact density from data used in Figure
+\ref{fig:energy-vs-diameter}.  As the radius of the rods approch zero we
+expect the contact density to approach the bulk density, which our
+functional does.  This self consistency shows that our functional is
+thermodynamically correct.
 
@@
 Figure~\ref{fig:surface-tension} (72~mN/m) and the force per length
-from Figure ~\ref{fig:rods-energy-vs-distance} (144~mN/m) which
+found from the slopes of the curves in Figure
+~\ref{fig:rods-energy-vs-distance} (144~mN/m)  which
 compares the free energy per length for rods of several different
@@
 per length on the rods is constant, and surprisingly independent of separation.
-By finding the slope of this curve, we calculate this force per length to be 144~mN/m.
 
@@
 where vapor is present in the space between the rods until a critical separation
-where it becomes filled with water and the rods behave like separate single rods
+where it becomes filled with water and the rods behave as separate single rods
 (shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
@@
 \end{equation}
-We again equate this circumference with the total circumference of
-the four separate rods, which results in a critical separation of
+As with the two rods, we equate this circumference with the total circumference of
+the four separate rods. This gives a critical separation of
 \begin{equation}

### 2012-05-21 papers/contact/paper.tex
Some changes to paper, some data for walls .1 fillfrac that has been computed since bug fix Tested everything on bennet.

--- 2012-05-12 papers/contact/paper.tex
@@
     A_{HS}}{\delta R(\mathbf{r})} \\
-  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
+  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \label{eq:g-A-exact-a}\\
   g_\sigma^A(\rr) &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
@@
 approximate White Bear functional, and find an approximate expression
-for $g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
+for $g_\sigma^A$, which is given in the Appendix.  This
 expression will involve more convolutions than are present in the free
 energy itself, making this formulation relatively expensive.
+
+In addition to the different densities in the denominators of Equations 
+\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
+between the treatment of the A and S cases in the derivatives 
+$\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
+and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
+that we use involve a number of weighted densities and for the derivatives 
+of these we use two different conceptual approaches.  In the assymetric case, 
+a change in $R(\mathbf{r})$ will refer to a change in the radius of a sphere 
+that is located at the position $\mathbf{r}$.  The weighted densities that are 
+affected by such a change will be those that are located a distance $R$ away 
+from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.
 
@@
 }
+
+\section{Symmetrically averaged correlation function}\label{sec:g-S}
+
+For the computation of the symmetrically averaged correlation function 
+we must formulate derivatives with respect to $R$ in a symmetric fashion.  
+Conceptually we expand the radii of spheres that just touch a specific 
+point of contact and only expand them outward at that point of contact.  
+Therefore $R(\mathbf{r})$ in our calculations refers to the radii of 
+spheres that touch $\mathbf{r}$ and $\delta R(\mathbf{r})$ refers to a 
+change of these radii at only the point $\mathbf{r}$.  One can imagine 
+a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell 
+area protruding outward at $\mathbf{r}$.  Figure~\ref{fig:contact} 
+illustrates such a change in $R$.  The full derivations are found in Appendix~\ref{appendix:g-S}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/n0}
+\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+  consist of those which just touch the point $\mathbf{r}$.}
+\label{fig:n0}
+\end{figure}
+
+These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
+which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the 
+derivation of the correlatation function results when we recognize that 
+instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ 
+as in the assymmetric case we will instead divide by $n_0(\mathbf{r})^2$ 
+since we have symmetry around the point $\mathbf{r}$.
+
+\begin{align}
+  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
+  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \\
+  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \label{eq:g-S-exact-a}\\
+  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-S-exact}
+\end{align}Complete expressions for $g_\sigma^S$ are included in
+Appendix~\ref{appendix:g-S}.\\
 
@@
 \begin{figure}
-\includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:n0}
-\end{figure}
-
-\section{Symmetrically averaged correlation function}\label{sec:g-S}
-
-An alternate definition of the averaged correlation function for an
-inhomogeneous system would be the correlation function averaged for
-spheres \emph{touching at a given point}.  This would be a symmetric
-measure, since both spheres involved are treated in the same way.  In
-fact, the density of such spheres is just the density $n_0(\rr)$ from
-FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
-is the fundamental measure of sphere surface contact, it makes sense
-to define a corresponding correlation function $g_\sigma^S$.  This is
-\emph{almost} what is done by Yu and  Wu
-
-\fixme{Analogous to the assymetric case is the symmetric, for which 
-we are no longer able to simply pick a hypothetical sphere centered 
-at position $r$ and expand its radius.  Instead we must formulate a way 
-to expand the radii of spheres that just touch a specific point of 
-contact and only expand them at that point of contact, in a symmetric 
-fashion.  Therefore, when we compute the neccasary derivitives wih 
-respect to $R$', we expand a multitude of spheres, all of them positioned 
-around and just touching the point $\mathbf{r}$.  When their radii 
-expand, each does not expand outward in all directions as in the 
-case of A, rather they each expand outward at only the point $\mathbf{r}$.
-One can imagine a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell area protruding outward at $\mathbf{r}$ while the rest of its shell remains
-at radius $R$.  When we expand each sphere that touches $\mathbf{r}$ 
-in this peculiar fashion, we have in the end the same amount of shell 
-area protruding outward as we would if we had simply expanded one 
-sphere in fashion of case A.  Also the protrusions, once added up 
-for all spheres sourrounding $\mathbf{r}$, will yield a symmetry 
-in their direction of protrusion.  Each protruding sphere bit will 
-expand outward from its center and all directions will be covered uniformly.
-See Figure \ref{fig:contact}.
-
-\begin{figure}
 \includegraphics[width=5cm]{figs/contact}
@@
 \end{figure}
-
-These qualitative concepts lead us to derivations for the $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ term in our derivation of the pressure, eq \ref{eq:p_{HS}}.
-The rest of the derivation results when we recognize that instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ as in the assymmetric case, we will instead divide by $n_0(\mathbf{r})^2$ since we have symmetry around the point $\mathbf{r}$.
-This will yield our symmetric correlation, $g_\sigma^S$:
-
-\begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
-  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \\
-  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
-\end{align}}Complete expressions for $g_\sigma^S$ are included in
-Appendix~\ref{appendix:g-S}.\\
 

### 2012-05-22 papers/contact/paper.tex

--- 2012-05-19 papers/contact/paper.tex
@@
 
+In addition to the different densities in the denominators of Equations 
+\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
+between the treatment of the A and S cases in the derivatives 
+$\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
+and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
+that we use involve a number of weighted densities and for the derivatives 
+of these we use two different conceptual approaches.  In the assymetric case, 
+a change in $R(\mathbf{r})$ will refer to a change in the radius of a sphere 
+that is located at the position $\mathbf{r}$.  The weighted densities that are 
+affected by such a change will be those that are located a distance $R$ away 
+from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.
+
 \derivation{
@@
 which gave us $g_\sigma^A(\rr)$:
+%% =======
+%% For the computation of the symmetrically averaged correlation function 
+%% we must formulate derivatives with respect to $R$ in a symmetric fashion.  
+%% Conceptually we expand the radii of spheres that just touch a specific 
+%% point of contact and only expand them outward at that point of contact.  
+%% Therefore $R(\mathbf{r})$ in our calculations refers to the radii of 
+%% spheres that touch $\mathbf{r}$ and $\delta R(\mathbf{r})$ refers to a 
+%% change of these radii at only the point $\mathbf{r}$.  One can imagine 
+%% a shell that touches $\mathbf{r}$ that has an infinitesimal bit of shell 
+%% area protruding outward at $\mathbf{r}$.  Figure~\ref{fig:contact} 
+%% illustrates such a change in $R$.  The full derivations are found in Appendix~\ref{appendix:g-S}
+
+%% \begin{figure}
+%% \includegraphics[width=5cm]{figs/n0}
+%% \caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+%%   consist of those which just touch the point $\mathbf{r}$.}
+%% \label{fig:n0}
+%% \end{figure}
+
+%% These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
+%% which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the 
+%% derivation of the correlatation function results when we recognize that 
+%% instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ 
+%% as in the assymmetric case we will instead divide by $n_0(\mathbf{r})^2$ 
+%% since we have symmetry around the point $\mathbf{r}$.
+
+%% \begin{align}
+%%   p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
+%%     A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
+%%   \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
+%%     A_{HS}}{\delta R(\mathbf{r})} \\
+%%   g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \label{eq:g-S-exact-a}\\
+%%   g_\sigma^A(\rr) &= \frac{1}{n_0(\rr) n_0(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
+%%     A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-S-exact}
+%% \end{align}Complete expressions for $g_\sigma^S$ are included in
+%% Appendix~\ref{appendix:g-S}.\\
+
+%% \section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+%% We compare with the averaged correlation function of
+%% Gross\cite{gross2009density}, which is of the asymmetrical variety
+%% ($g_\sigma^A$), where the correlation function is averaged at location
+%% of one of the spheres in concact.
+%% \begin{align}
+%%   g_\sigma^\text{Gross,A} &= \frac{1 - \frac12 n_{3'}}{\left(1 -
+%%     n_{3'}\right)^3} \\
+%%   n_{3'}(\rr) &= \int n(\rr')\Theta(|\rr-\rr'| - 2R) d\rr'
+%% \end{align}
+%% where $n_{3'}$ is an averaged
+%% filling fraction defined in a manner analogous to the $n_3$ of FMT,
+%% but with the hard-sphere radius replaced by the hard-sphere diameter,
+%% so the integrals range over spheres that are twice as large.
+
+%% >>>>>>> a989df3e529c6d7afbecef041c2a6afb3cb30c08
 \begin{figure}

### 2012-05-29 papers/water-SAFT/paper.tex

@@
 expected contact density derived from the contact value theorem and
-data used in Figure \ref{fig:energy-vs-diameter}.}
+the free energy data
+data plotted in Figure \ref{fig:energy-vs-diameter}.}
 \label{fig:density-single-rod}

### 2012-05-29 papers/water-SAFT/paper.tex
Reran two rods simulations. Added files for single rod density plot.  Changed some wording in conclusion and introduction. Added small section for contact-value theorem. Tested everything on quipu.

@@
 atomic level.}
-\tableofcontents
 \end{abstract}
@@
 
-Due to its large dielectric constant, small molecular size and strong
-hydrogen-bonding network, water as ``the universal solvent'' is as
-interesting as its anomalous thermodynamic properties.  \fixme{what
-  the heck?}
+As common as water is, its large dielectric constant, small
+molecular size, strong hydrogen-bonding network and anomolous
+thermodynamic properties make ``the universal solvent'' an interesting
+subject of study.
 
@@
 dependence of the surface tension.  However, they also overestimate
-the surface tension by \fixme{how much? significant factor, I
-  think}, and are unsuited for studying systems with molecular-scale
-inhomogeneities.
+the surface tension by a factor of about 1.5, and are unsuited for studying
+systems with molecular-scale inhomogeneities.
 
-%% A recent functional presented by Lischner and Arias~\cite{lischner2010classical}
-%% describe water as non-interacting molecules in a an
-%% effective external potential.  This functional contains the exact
-%% entropy for no-interacting molecules and, among other things, can 
-%% predict the nonlinear dielectric response in water, but it must be paramaterized
-%% for different temperatures.
+\subsection{Contact-value theorem}
+
+The contact value theorem states that the pressure of a fluid at a
+hard wall interface is proportional to the contact density of the fluid,
+$n_c$.  Specifically,
+\begin{equation}\label{eq:contact}
+  P = n_ck_BT
+\end{equation}
+where $k_B$ is the Boltzmann constant and $T$ is the temperature of the
+fluid. This leads to the property that the free energy of solvation of
+a hard solute is proportional to its volume in the limit of small
+solutes: 
+\begin{equation}\label{contactvaluethm}
+  F = n_c k_BT V\:
+\end{equation}
 
@@
 model for water\cite{clark2006developing}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\end{center}
+\caption{The pressure versus density for various temperatures, including
+experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
+indicate the computationally calculated pressure and the open circles are 
+NIST data points, plotted for every other isotherm. The solid and dotted lines
+represent the theoretical and experimental coexistence curves.}
+\label{fig:pressure-with-isotherms}
+\end{figure}
 
@@
 \end{equation}
-The ideal gas free energy functional satisfies the contact density
-theorem, which states that the contact density at a hard surface $n_c$
-is proportional to the pressure on that surface. This
-also leads to the property that the free energy of solvation of
-a hard solute is proportional to its volume in the limit of small
-solutes: 
-\begin{equation}\label{contactvaluethm}
-  F = n_c k_BT V\:
-\end{equation}
-These properties are retained by the total functional,
-provided the remaining terms are \emph{purely nonlocal}.  A common
-failing of SAFT-based functionals (see, for instance
-references\cite{felipe2001examination, gloor2002saft,
-  gloor2004accurate, clark2006developing, gloor2007prediction,
-  kahl2008modified, gross2009density}) is to treat other terms in the
-functional locally, resulting in a functional that is suitable only
-for studying the liquid-vapor interface.
+The ideal gas free energy functional satisfies the contact value
+theorem and its limiting case of small solutes (Equation
+\ref{contactvaluethm}). These properties are retained by the total
+functional, provided the remaining terms are \emph{purely nonlocal}.
 
@@
 \label{fig:surface-tension}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
 \end{figure}
@@
 The results for this spacing of 0.01~nm are within 0.5\% of the free energy 
-and a few percent of the contact
-density for the highest resolution tested.
-Figure~\ref{fig:density-single-rod} shows density
-profiles for three different radii (0.3~nm, 0.5~nm, and 1~nm) which have the
+and a few percent of the contact density for the highest resolution tested.
+Figure~\ref{fig:density-single-rod} shows density profiles for
+different radii rods which have the
 expected oscillations outside the rod that reduce to the bulk saturated liquid
@@
 Based on this result, we expect to be
-able to classically describe rods with a radius of about 0.5~nm or larger. 
+able to classically describe a single rod with a radius of about 0.5~nm or larger. 
 For two hydrophobic rods surrounded by water, we see a transition from a vapor-filled
@@
 classical treatment of the critical separation for this transition
-works well for rods larger than 0.5~nm.  Further evidence for the
+works well for pairs of rods with radii larger than 0.5~nm.  Further evidence for the
 accuracy of this classical treatment is the calculation of the force on the
 stadium shape. The force per length is constant and equal to twice the
-surface tension. A simulations of hydrophobic spheres agrees with
-SPC/E calculations and a system with four hydrophobic rods exhibited
-similar behavior as the two rod.
+surface tension. A similar system with four hydrophobic rods exhibited the same
+behavior while simulations of hydrophobic spheres agrees with
+SPC/E calculations.
 

### 2012-06-01 papers/water-SAFT/paper.tex
Reran canculations for four-rods 1.8nm in diameter. Moved contact value section of paper (still needs work). Changed plot for pressure vs. density and added more experimental data. Tested everything on quipu.

@@
 A number of classical density functionals have been developed for
-water~\cite{ding1987, Yang1992, gloor2002saft, Jaqaman2004,
-  clark2006developing, lischner2010classical}\fixme{ADD REST OF
-  LIST HERE!}, capturing some of the qualitative behavior of water.
+water~\cite{ding1987, Yang1992, gloor2002saft,gloor2004accurate,
+  gloor2007prediction, Jaqaman2004, clark2006developing,
+  lischner2010classical, fu2005vapor-liquid-dft,felipe2001examination}
+, capturing some of the qualitative behavior of water.
 However, many of these are less sophisticated than those describing
@@
 treats the surface tension correctly, but can only be used at room
-temperature.  A numer of classical density functionals have recently
+temperature.  A number of classical density functionals have recently
 been produced that are based on Statistical Associating Fluid Theory
-(SAFT)\cite{gloor2002saft, clark2006developing}\fixme{add
-  more}.  These functionals are based on a perturbative thermodynamic
+(SAFT)\cite{gloor2002saft, clark2006developing, gloor2007prediction,
+  gloor2004accurate, gross2009density, kahl2008modified, felipe2001examination}.
+These functionals are based on a perturbative thermodynamic
 expansion, and give a reasonable prediction for the temperature
 dependence of the surface tension.  However, they also overestimate
-the surface tension by a factor of about 1.5, and are unsuited for studying
-systems with molecular-scale inhomogeneities.
-
-\subsection{Contact-value theorem}
-
-The contact value theorem states that the pressure of a fluid at a
-hard wall interface is proportional to the contact density of the fluid,
-$n_c$.  Specifically,
-\begin{equation}\label{eq:contact}
-  P = n_ck_BT
-\end{equation}
-where $k_B$ is the Boltzmann constant and $T$ is the temperature of the
-fluid. This leads to the property that the free energy of solvation of
-a hard solute is proportional to its volume in the limit of small
-solutes: 
-\begin{equation}\label{contactvaluethm}
-  F = n_c k_BT V\:
-\end{equation}
+the surface tension by a factor of about 1.5, and \fixme{most?} are unsuited
+for studying systems that have density variations on a molecular
+length scale due to the use of a local density approximation.
 
@@
 The SAFT Helmholtz free energy is composed of five terms:
-\begin{equation}
+\begin{equation} \label{eq:SAFT-free-energy}
   F[n] = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
+\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms-truncated}
 \end{center}
@@
 indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
+NIST data points. The solid and dotted lines
 represent the theoretical and experimental coexistence curves.}
@@
 functionals have made local-density approximations that make them
-unsuitable for application to systems that are inhomogeneous at the
-molecular length scale\cite{felipe2001examination, gloor2002saft,
+unsuitable for application to systems that obey the contact-value
+theorem. \cite{felipe2001examination, gloor2002saft,
   gloor2004accurate, clark2006developing, gloor2007prediction,
-  kahl2008modified, gross2009density}.
+  kahl2008modified, gross2009density}.\
+
+ However, they also overestimate
+the surface tension by a factor of about 1.5, and \fixme{most?} are unsuited
+for studying systems that have density variations on a molecular
+length scale due to the use of a local density approximation.
+
+Functionals constructed using a local density approximation
+  will always fail to satisfy the contact-value theorem, and will
+  therefore incorrectly model small hard solutes.
+The contact-value theorem states that the pressure of any fluid at a
+hard wall interface is proportional to the contact density of the
+fluid\cite{henderson1979exact}, according to the following relation:
+\begin{equation}\label{eq:contact}
+  p = n_ck_BT
+\end{equation}
+where $n_c$ is the contact density, $k_B$ is the Boltzmann constant
+and $T$ is the temperature of
+the fluid. This leads to the property that the free energy of
+solvation of a small hard solute is proportional to the
+solvent-excluded volume:
+\begin{equation}\label{contactvaluethm}
+  F = n k_BT V
+\end{equation}
+\fixme{add this here? :} The contact-value theorem is satisfied for
+functionals for with the only purely local term is the free energy of
+an ideal gas.
 
@@
 add association terms based on SAFT. For our free energy, we use the
-first four terms from equation (number): $F_\textit{id}$,
+first four terms from Equation \ref{eq:SAFT-free-energy}: $F_\textit{id}$,
 $F_\textit{hs}$, $F_\textit{disp}$ and $F_\textit{assoc}$. In

### 2012-06-05 papers/water-SAFT/paper.tex
Changed coloring/size on single rod density plot.  Made edits to paper. Tested everything on node07.

@@
 molecular size, strong hydrogen-bonding network and anomolous
-thermodynamic properties make ``the universal solvent'' an interesting
-subject of study.
+thermodynamic properties make ``the universal solvent'' a unique
+and interesting material.
 
@@
 expansion, and give a reasonable prediction for the temperature
-dependence of the surface tension.  However, they also overestimate
-the surface tension by a factor of about 1.5, and \fixme{most?} are unsuited
-for studying systems that have density variations on a molecular
-length scale due to the use of a local density approximation.
+dependence of the surface tension.
 
@@
 which has been used by Clark \emph{et al} to construct an optimal SAFT
-model for water\cite{clark2006developing}.
+model for water\cite{clark2006developing}. A majority of the empirical
+parameters used in our functional are taken from the Clark\emph{et al}
+SAFT functional which contains five empirical parameters and give the
+correct pressure at liquid-vapor coexistince curve shown in Figure
+\ref{fig:pressure-with-isotherms}.
 
@@
 indicate the computationally calculated pressure and the open circles are 
-NIST data points. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
+NIST data points. The solid and dotted lines represent the theoretical
+and experimental coexistence curves.}
 \label{fig:pressure-with-isotherms}
@@
 temperature\cite{whotocite}.  Such functionals have qualitatively
-predicted the dependence of surface tension on temperature, but have
-in general provided poor quantitative agreement with experiment.
-Moreover, the vast majority of SAFT-based classical density
-functionals have made local-density approximations that make them
-unsuitable for application to systems that obey the contact-value
-theorem. \cite{felipe2001examination, gloor2002saft,
-  gloor2004accurate, clark2006developing, gloor2007prediction,
-  kahl2008modified, gross2009density}.\
-
- However, they also overestimate
+predicted the dependence of surface tension on temperature, however, they also overestimate
 the surface tension by a factor of about 1.5, and \fixme{most?} are unsuited
@@
 \end{equation}
-\fixme{add this here? :} The contact-value theorem is satisfied for
-functionals for with the only purely local term is the free energy of
-an ideal gas.
+The contact-value theorem is satisfied for functionals where the
+only purely local term is the free energy of an ideal gas.
 
@@
 between free energy per length and separation during the initial state 
-(where the rods have vapor between them). This indicates that the force 
+with vapor between them. This indicates that the force 
 per length on the rods is constant, and surprisingly independent of separation.

### 2012-06-08 papers/contact/paper.tex

--- 2012-05-19 papers/contact/paper.tex
@@
 which gave us $g_\sigma^A(\rr)$:
+
 \begin{figure}
@@
 \end{figure}
+
+These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
+which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the 
+derivation of the correlatation function results when we recognize that 
+instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$ 
+as in the assymmetric case we will divide by $n_0(\mathbf{r})^2$
+since we have symmetry around the point $\mathbf{r}$.
+
 \begin{align}
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/walls-10}
+  \includegraphics[height=10cm]{figs/walls-10}
   \caption{Density and contact densities at a wall with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/walls-30}
+  \includegraphics[height=10cm]{figs/walls-30}
   \caption{Density and contact densities at a wall with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/walls-40}
+  \includegraphics[height=10cm]{figs/walls-40}
   \caption{Density and contact densities at a wall with bulk filling
@@
   \label{fig:walls-40}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-walls-10}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:gnn-walls-10}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-walls-30}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:gnn-walls-30}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-walls-40}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.1.}
+  \label{fig:gnn-walls-10}
 \end{figure}

### 2012-06-08 papers/thesis-hughes/project.tex

--- 2011-10-18 paper/project.tex
@@
 \newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
+\newcommand\kappaassoc{\ensuremath{\kappa_\textit{a}}}
 \newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
@@
 with the lone pair of the other.  The volume over which this
-interaction occurs is $\kappa_\textit{assoc}$, giving the association
+interaction occurs is $\kappaassoc$, giving the association
 term in the free energy has two empirical parameters that are fit to
@@
 \begin{align}
-  \Delta(\xx) &= \kappa_\textit{assoc} g^\textit{SW}_\sigma(\xx)
+  \Delta(\xx) &= \kappaassoc g^\textit{SW}_\sigma(\xx)
   \left(e^{-\beta\epsilonassoc} - 1\right) \\
@@
 \end{center}
-\caption{ Free nergy versus radius for a single hydrophobic rod
+\caption{ Free energy versus radius for a single hydrophobic rod
 immersed in water. This should have an asymptote equal to the surface
@@
 lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
-medium 0.2~bohrs ($\approx$~0.001~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
+medium 0.2~bohrs ($\approx$~0.01~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
 \label{fig:densityresolution}
@@
 the critical separation, $d_c$, will occur when the free energies are equal. This 
-corresponds to equating the circumference of the stadium shape (see 
+corresponds to equating the circumference of the stadium-shape (see 
 top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom

### 2012-06-08 papers/water-SAFT/paper.tex

@@
 \begin{abstract}
-{We develop a classical density functional theory for water by combining the
-fundamental-measure theory functional--which describes hard sphere fluids--with
-attractive interactions based on the Statistical Associating Fluid Theory
-(SAFT-VR).
-Using this approximation with an additional length scale parameter, we are able to calculate the vapor pressure, saturated liquid density, surface tension, and enthalpy of water, as well as the number of hydrogen bonds and density oscillations formed near hydrophobic rods and spheres immersed in water.
-For two rods and four rods(arranged as the corners of a square) in water, our functional predicts
-a transition from a vapor state to a liquid state in the area between
-the rods that agrees with clasical theory.   We find that in the vapor state, the force per length can be
-approximated classically and is equal to 144~mN/m, or twice the bulk surface 
-tension even for nm scale rods, and  our hydrophobic spheres agree
-reasonably well with SPC/E simulations.
-This functional is computationally efficient and provides a foundation for 
-future work describing complex interactions in aqueous solution on an
-atomic level.}
+We develop a classical density functional for water which combines the
+White Bear fundamental-measure theory (FMT)
+functional\cite{roth2002whitebear} for the hard sphere fluid with
+attractive interactions based on the Statistical Associating Fluid
+Theory (SAFT-VR)\cite{clark2006developing}.
+This functional reproduces the properties of water at both long and
+short length scales over a wide range of temperatures, and is
+efficient to evaluate computationally, comparable to the cost of FMT
+itself.  We demonstrate our functional by applying it to systems
+composed of of two hard rods, four hard rods arranged in a square and
+hard spheres in water.
 \end{abstract}
@@
 \section{Introduction}
-
-\subsection{Water---the universal solvent}
 
@@
 at 0.84 molar and allows contact between the second solvation shell of
-neighboring solutes~\cite{izvekov2005, choe2007}.  On top of this,
-standard \emph{ab initio} methods using classical molecular dynamics
-without van der Waals corrections strongly over-structure water, to
-the point that ice melts at over 400K\cite{yoo2009phase}!  There has
-been a recent flurry of publications implicating van der Waals effects
-as significant in reducing this
-over-structuring\cite{lin2009importance, wang2011density,
-  mogelhoj2011ab, jonchiere2011van}.  It has also been found that the
-inclusion of nuclear quantum effects can provide similar improvements
-\cite{morrone2008nuclear}.  Each of these corrections imposes an
-additional computational burden on an approach that is already
-feasible for only a very small number of water molecules. A more
-efficient approach is needed in order to study nanoscale and
-hydrophobic solutes.
-
-Water stands out among liquids for its extraordinary properties.
-Water is the only natural substance that is found in solid, liquid and
-gaseous states at temperatures and pressures normally found on Earth.
-Water has a density maximum at 4$^\circ$C, and expands upon freezing.
-Water has the highest heat capacity, latent heat of fusion and
-vaporization, thermal conductivity and surface tension of any
-non-metallic liquid.  Water also has an unusually high dielectric constant.
-Much of the unusual behavior of water is explained by the ability of
-each water molecule to form \emph{four} strong, directional hydrogen
-bonds---two acceptors and two donors.  This directional bonding
-dictates the open structure of ice, leading to thermal expansion upon
-freezing.
-
-As common as water is, its large dielectric constant, small
-molecular size, strong hydrogen-bonding network and anomolous
-thermodynamic properties make ``the universal solvent'' a unique
-and interesting material.
+neighboring solutes~\cite{izvekov2005, choe2007}\fixme{Look up more
+  recent numbers on this!}.  On top of this, standard \emph{ab initio}
+methods using classical molecular dynamics without van der Waals
+corrections strongly over-structure water, to the point that ice melts
+at over 400K\cite{yoo2009phase}!  There has been a recent flurry of
+publications implicating van der Waals effects as significant in
+reducing this over-structuring\cite{lin2009importance,
+  wang2011density, mogelhoj2011ab, jonchiere2011van}.  It has also
+been found that the inclusion of nuclear quantum effects can provide
+similar improvements \cite{morrone2008nuclear}.  Each of these
+corrections imposes an additional computational burden on an approach
+that is already feasible for only a very small number of water
+molecules. A more efficient approach is needed in order to study
+nanoscale and hydrophobic solutes.
 
@@
 provides no information about the size of this cavity.  In a physically
-correct approach, the size of the cavity will naturally arise from a
+consistent approach, the size of the cavity will naturally arise from a
 balance between the free energy required to create the cavity, the
@@
 
-A very promising approach for an efficient continuum description of water
-is that of classical density-functional theory (DFT), which is an approach
-for evaluating the free energy and thermally averaged density of fluids in
-an arbitrary external potential~\cite{ebner1976density}. The Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
-energy in terms of a universal functional of the electron density and a separate
-term including the external potential. Mermin\cite{mermin1965thermal} gives a
-theorem extending that of Hohenberg and Kohn to nonzero temperatures:
+One promising approach for an efficient continuum description of water
+is that of classical density-functional theory (DFT), which is an
+approach for evaluating the free energy and thermally averaged density
+of fluids in an arbitrary external potential~\cite{ebner1976density}.
+The foundation of classical DFT is the Mermin
+theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature,
+stating that
 \begin{equation}
-  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
-d^3r\right\}.
+  A(T) = \underset{n(\rr)}{min}\left\{ F[n(r),T] + \int V_\textit{ext}(\rr) n(\rr)
+d\rr\right\}.
 \end{equation}
-This is the basis for classical DFT, where the relevant density is not the
-electron density, but the density of molecules. Classical DFT is a
-natural framework for creating a more flexible theory of hydrophobicity
-that can describe interaction of water with arbitrary external
-potentials---including potentials describing a \emph{strong} interaction
-with a solute.
+where $A(T)$ is the Helmholtz free energy of a system in the external
+potential $V_\textit{ext}$ at temperature $T$, $F[n(r),T]$ is a
+universal free-energy functional, independent of external potential,
+and $n(\rr)$ is a density of atoms or molecules.  Classical DFT is a
+natural framework for creating a more flexible theory of
+hydrophobicity that can describe interaction of water with arbitrary
+external potentials---including potentials describing strong
+interactions with solutes or surfaces.
 
 A number of exact properties are easily achieved in the
-density-functional framework, such as the relationship between the
-pressure on a hard wall and the contact density---which ensures a
-correct solvation free energy for small hard solutes.  Much of the
-research on classical density-functional theory has focused on the
-hard-sphere fluid~\cite{curtin1985, rosenfeld1989, rosenfeld1993,
-  rosenfeld1997, tarazona1997, tarazona2000}.  This has led to a large
-number of very sophisticated functionals, such as the
-fundamental-measure theory (FMT) functional~\cite{rosenfeld1989,
+density-functional framework, such as the contact-value theorem, which
+ensures a correct free energy of hydration for small hard solutes.
+Much of the research on classical density-functional theory has
+focused on the hard-sphere fluid~\cite{curtin1985, rosenfeld1989,
+  rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}, which has
+led to a number of sophisticated functionals, such as the
+fundamental-measure theory (FMT) functionals\cite{rosenfeld1989,
   rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000,
-  roth2002whitebear}.  The White Bear version of the FMT
-functional\cite{roth2002whitebear}  has a number of useful
-properties.  Like all FMT functionals, it is expressed as a function
-of a few convolutions of the density that can be efficiently
-computed.  In the homogeneous limit, it reduces to the
-Carnahan-Starling equation of state, and in the strongly-confined
-limit of a zero-dimensional cavity, it gives the exact free energy.
+  roth2002whitebear}, which are entirely expressed as an integral of
+local functions of a few convolutions of the density that can be
+efficiently computed.  We will use the White Bear version of the FMT
+functional\cite{roth2002whitebear}.  In the homogeneous limit, this
+functional reduces to the Carnahan-Starling equation of state, and in
+the strongly-confined limit of a highly confined cavity, it reproduces
+the exact free energy.
 
@@
   gloor2007prediction, Jaqaman2004, clark2006developing,
-  lischner2010classical, fu2005vapor-liquid-dft,felipe2001examination}
-, capturing some of the qualitative behavior of water.
-However, many of these are less sophisticated than those describing
-the hard-sphere fluid, and a quantitatively accurate functional
-describing water---which reproduces the temperature dependence of the
-macroscopic surface tension, its decrease as curvature increases, and
-its quantitative value---has not yet been published.  For instance,
-the functional of Lischner \emph{et al}\cite{lischner2010classical}
-treats the surface tension correctly, but can only be used at room
-temperature.  A number of classical density functionals have recently
-been produced that are based on Statistical Associating Fluid Theory
-(SAFT)\cite{gloor2002saft, clark2006developing, gloor2007prediction,
-  gloor2004accurate, gross2009density, kahl2008modified, felipe2001examination}.
-These functionals are based on a perturbative thermodynamic
-expansion, and give a reasonable prediction for the temperature
-dependence of the surface tension.
+  lischner2010classical,
+  fu2005vapor-liquid-dft,felipe2001examination}, each of which
+captures some of the qualitative behavior of water.  However, each of
+these functionals also fail to capture some of water's unique
+properties.  For instance, the functional of Lischner \emph{et
+  al}\cite{lischner2010classical} treats the surface tension
+correctly, but can only be used at room temperature, and thus captures
+none of the temperature-dependence of water.  A number of classical
+density functionals have recently been produced that are based on
+Statistical Associating Fluid Theory (SAFT)\cite{gloor2002saft,
+  clark2006developing, gloor2007prediction, gloor2004accurate,
+  gross2009density, kahl2008modified, felipe2001examination}.  These
+functionals are based on a perturbative thermodynamic expansion, and
+do reproduce the temperature-dependence of water's properties.
 

### 2012-06-12 papers/water-SAFT/paper.tex
Added more papers to table (for organized citing) and improved pressure-density plot in paper.tex. Fixed an error in paper.bib. Tested everything on quipu.

@@
 A number of classical density functionals have been developed for
-water~\cite{ding1987, Yang1992, gloor2002saft,gloor2004accurate,
+water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,gloor2004accurate,
   gloor2007prediction, Jaqaman2004, clark2006developing,
-  lischner2010classical,
-  fu2005vapor-liquid-dft,felipe2001examination}, each of which
+  lischner2010classical, fu2005vapor-liquid-dft,
+  blas2001examination}, each of which
 captures some of the qualitative behavior of water.  However, each of
@@
 density functionals have recently been produced that are based on
-Statistical Associating Fluid Theory (SAFT)\cite{gloor2002saft,
+Statistical Associating Fluid Theory (SAFT)\cite{ 
+  yu2002fmt-dft-inhomogeneous-associating,
+  fu2005vapor-liquid-dft,gloor2002saft,
   clark2006developing, gloor2007prediction, gloor2004accurate,
-  gross2009density, kahl2008modified, felipe2001examination}.  These
+  gross2009density, kahl2008modified, blas2001examination}.  These
 functionals are based on a perturbative thermodynamic expansion, and
@@
 Statistical Associating Fluid Theory (SAFT) is a set of models for
-complex fluids in which hydrogren bonding plays a significant
+complex fluids in which hydrogen bonding plays a significant
 role\cite{muller2001molecular}.  SAFT is used to accurately model the
@@
   wertheim1986fluidsIII, wertheim1986fluidsIV} which allows it to
-account for strong associative interations between molecules.
+account for strong associative interactions between molecules.
 
@@
 model for water\cite{clark2006developing}. A majority of the empirical
-parameters used in our functional are taken from the Clark\emph{et al}
-SAFT functional which contains five empirical parameters and give the
-correct pressure at liquid-vapor coexistince curve shown in Figure
+parameters used in our functional are taken from the Clark \emph{et al}
+paper which contains five empirical parameters and gives
+correct pressures at the liquid-vapor coexistence curve shown in Figure
 \ref{fig:pressure-with-isotherms}.
@@
 \caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points. The solid and dotted lines represent the theoretical
-and experimental coexistence curves.}
+experimental pressure data from NIST\cite{nistwater}. The solid colored lines
+indicate the computationally calculated pressure and the dotted
+colored lines are NIST data points. The solid and dotted black lines
+represent the theoretical and experimental coexistence curves.}
 \label{fig:pressure-with-isotherms}
@@
 which have been applied to study the surface tension as a function of
-temperature\cite{whotocite}.  Such functionals have qualitatively
-predicted the dependence of surface tension on temperature, however, they also overestimate
-the surface tension by a factor of about 1.5, and \fixme{most?} are unsuited
+temperature\cite{clark2006developing, gloor2004accurate,
+  gloor2007prediction, blas2001examination, gloor2002saft}.  Such
+functionals have qualitatively predicted the dependence of surface
+tension on temperature, however, they also overestimate
+the surface tension by a factor of about 1.5, and most are unsuited
 for studying systems that have density variations on a molecular
-length scale due to the use of a local density approximation.
+length scale due to the use of a local density
+approximation\cite{gloor2002saft,clark2006developing, gloor2007prediction,
+gloor2004accurate, gross2009density, kahl2008modified, blas2001examination}.
+
+\begin{table}
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+paper & SAFT functional & water & LDA & tension \\
+\hline
+yuwu2002dft & X & & & \\
+\hline
+fuwu2005 & X & X & & X \\
+\hline
+kiselev2006 &  & X & X & X \\
+\hline
+clark2006developing & X & X & X & X \\
+\hline
+gloor2002 & X & X & X & X \\
+\hline
+gloor2004 & X & X & X & X \\
+\hline
+gloor2007 & X & X & X & X \\
+\hline
+gross2009 & X &  & X & X \\
+\hline
+Lischner2010 &  & X &  &  \\
+\hline
+Blas2001 & X & X & X & X \\
+\hline
+Kahl2008 & X &  & X & X \\
+\hline
+Muller2001 & X &  &  &  \\
+\hline
+Yang1994 &  & X &  & X \\
+\hline
+ding1987 &  & X & ? &  \\
+\hline
+next & ? & ? & ? & ? \\
+\hline
+\end{tabular}
+\caption{Which papers do what}
+\end{table}
 
@@
 The ideal gas free energy functional satisfies the contact value
-theorem and its limiting case of small solutes (Equation
-\ref{contactvaluethm}). These properties are retained by the total
-functional, provided the remaining terms are \emph{purely nonlocal}.
+theorem and its limiting case of small solutes (Equations
+\ref{eq:contact} and \ref{contactvaluethm}). These properties are
+retained by the total functional, provided the remaining terms are
+\emph{purely nonlocal}.
 
@@
 The association functional we use is a modified version of Yu
-and Wu.\cite{yu2002fmt-dft-inhomogeneous-associating} \footnote{We
+and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}. \footnote{We
   should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
@@
 \end{center}
-\caption{ Density profiles for a single rod of different radii. The dotted line
+\caption{ Density profiles for single rods of different radii. The dotted line
 represents the saturated liquid density and the points represent the
 expected contact density derived from the contact value theorem and
-the free energy data
-data plotted in Figure \ref{fig:energy-vs-diameter}.}
+calculated free energy data.}
 \label{fig:density-single-rod}
@@
 contact value theorem to derive the contact density at the water-rod
-interface.  We start with the thermodyamic expression relating
+interface.  We start with the thermodynamic expression relating
 pressure to the partial derivative of free energy with respect the
-radius of a cylinder. Using the contact value theorem and sustituting in
+radius of a cylinder. Using the contact value theorem and substituting in
 the free energy gradient for the pressure, we can derive values for
 the contact density from data used in Figure
-\ref{fig:energy-vs-diameter}.  As the radius of the rods approch zero we
+\ref{fig:energy-vs-diameter}.  As the radius of the rods approach zero we
 expect the contact density to approach the bulk density, which our
@@
 
-To determine the critical separation, we start by consider a macroscopic surface
+To determine the critical separation, we start by considering a macroscopic surface
 tension model where the free energy in the macroscopic limit is given by 
@@
 free energy per length, which is equal to the circumference multiplied 
-by the surface tenson. The force per length is the derivative of this
+by the surface tension. The force per length is the derivative of this
 with respect to the separation. 
@@
 Figure~\ref{fig:rods-energy-vs-distance} where there is a transition
-betwen the sloped and flat portion of each plot. They match fairly well for the 
+between the sloped and flat portion of each plot. They match fairly well for the 
 0.5~-~1.0~nm rods with differences between 1.3~-~5.5\% and we hypothesize that larger radii rods
@@
 what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
-and carbon-water interations.
+and carbon-water interactions.
 
@@
 \end{center}
-\caption{ Free energy of hydradon 
+\caption{ Free energy of hydration 
 versus separation for four hydrophobic rods ranging in radius from
@@
   simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
-  The horizonal lines show the experimental and SPC/E bulk surface
+  The horizontal lines show the experimental and SPC/E bulk surface
   tension for water at standard atmospheric temperature and
@@
 The classical density functional we developed by combining fundamental-measure 
-theory with SAFT is able to produce reasonable results for hydrophobic interations.
+theory with SAFT is able to produce reasonable results for hydrophobic interactions.
 We added the parameter $\lscale$ to the functional and fit it to the surface tension

### 2012-06-20 papers/contact/paper.tex
changed plotting and latex for contact Tested everything on node03.

@@
   \includegraphics[height=9cm]{figs/gnn-walls-10}
-  \caption{Density and contact densities at a wall with bulk filling
+  \caption{Correlation multiplied twice by density with bulk filling
     fraction of 0.1.}
@@
   \includegraphics[height=9cm]{figs/gnn-walls-30}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.1.}
+  \caption{Correlation multiplied twice by density with bulk filling
+    fraction of 0.3.}
   \label{fig:gnn-walls-30}
@@
   \includegraphics[height=9cm]{figs/gnn-walls-40}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.1.}
+  \caption{Correlation multiplied twice by density with bulk filling
+    fraction of 0.4.}
   \label{fig:gnn-walls-10}
@@
   \label{fig:sphere-16}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-sphere-08-013}
+  \caption{Correlation multiplied twice by density of 13 hard spheres in a
+    spherical cavity with diameter 8.}
+  \label{fig:gnn-sphere-08-013}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-sphere-12-112}
+  \caption{Correlation multiplied twice by density of 112 hard spheres in a
+    spherical cavity with diameter 12.}
+  \label{fig:gnn-sphere-12-112}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[height=9cm]{figs/gnn-sphere-16-265}
+  \caption{Correlation multiplied twice by density of 265 hard spheres in a
+    spherical cavity with diameter 16.}
+  \label{fig:gnn-sphere-16-265}
 \end{figure}

### 2012-06-20 papers/water-SAFT/paper.tex

@@
 
-Statistical Associating Fluid Theory (SAFT) is a set of models for
+Statistical Associating Fluid Theory (SAFT) is a theory describing
 complex fluids in which hydrogen bonding plays a significant
 role\cite{muller2001molecular}.  SAFT is used to accurately model the
-equations of state of pure fluids and predict the equations of state
-for mixtures over a wide range of temperatures and pressures.  SAFT is
+equations of state of both pure fluids and mixtures over a wide range
+of temperatures and pressures.  SAFT is
 based on Wertheim's first-order thermodynamic perturbation theory
-(TPT1) \cite{wertheim1984fluidsI, wertheim1984fluidsII,
-  wertheim1986fluidsIII, wertheim1986fluidsIV} which allows it to
+(TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
 account for strong associative interactions between molecules.
@@
 describing hydrogen bonds, and the final term is the chain formation
-energy of for fluids that are chain polymers.  The final term is not
-relevant for water.  While a number of SAFT formulations have been
+energy of for fluids that (unlike water) are chain polymers.
+While a number of forumlations of SAFT have been
 published, we will focus on SAFT-VR\cite{gil-villegas-1997-SAFT-VR},
-which has been used by Clark \emph{et al} to construct an optimal SAFT
-model for water\cite{clark2006developing}. A majority of the empirical
-parameters used in our functional are taken from the Clark \emph{et al}
-paper which contains five empirical parameters and gives
-correct pressures at the liquid-vapor coexistence curve shown in Figure
-\ref{fig:pressure-with-isotherms}.
+which was used by Clark \emph{et al} to construct an optimal SAFT
+model for water\cite{clark2006developing}. All but one of the six
+empirical parameters used in the functional introduced in this paper
+are taken directly from this Clark \emph{et al} paper.  The phase
+diagram for this optimal SAFT model for water is shown in
+Figure~\ref{fig:pressure-with-isotherms}, which demonstrates that its
+vapor pressure as a function of temperature is very accurate, while
+the liquid density shows larger discrepancies.  The critical point is
+very poorly described, which is a common failing of mean-field
+expansions.
 
@@
 SAFT has also been used to construct classical density functionals,
-which have been applied to study the surface tension as a function of
+which are commonly used to study the surface tension as a function of
 temperature\cite{clark2006developing, gloor2004accurate,
@@
 functionals have qualitatively predicted the dependence of surface
-tension on temperature, however, they also overestimate
-the surface tension by a factor of about 1.5, and most are unsuited
-for studying systems that have density variations on a molecular
-length scale due to the use of a local density
+tension on temperature, but they also overestimate
+the surface tension by about 50\%, and most SAFT-based functionals are
+unsuited for studying systems that have density variations on a
+molecular length scale due to the use of a local density
 approximation\cite{gloor2002saft,clark2006developing, gloor2007prediction,
@@
 Functionals constructed using a local density approximation
-will always fail to satisfy the contact-value theorem, and will
+fail to satisfy the contact-value theorem, and
 therefore incorrectly model small hard solutes.
@@
 hard wall interface is proportional to the contact density of the
-fluid\cite{henderson1979exact}, according to the following relation:
+fluid\cite{henderson1979exact}:
 \begin{equation}\label{eq:contact}
@@
 The contact-value theorem is satisfied for functionals where the
-only purely local term is the free energy of an ideal gas.
+only purely local term is the free energy of an ideal gas, and is not
+satisfied for functionals built using a local density approximation.
 
 \section{Theory and Methods}
-We construct a new classical density functional for water. We treat
-each individual water molecule as a purely classical hard sphere and
-add association terms based on SAFT. For our free energy, we use the
+We construct a classical density functional for water, which reduces
+in the homogeneous limit to the optimal SAFT model for water found by
+Clark~\emph{et al}.  The Helmholtz free energy is constructed using the
 first four terms from Equation \ref{eq:SAFT-free-energy}: $F_\textit{id}$,
-$F_\textit{hs}$, $F_\textit{disp}$ and $F_\textit{assoc}$. In
-addition, a chemical potential term is used, so we actually work in
-the grand canonical ensemble.  In the following sections, we will
-introduce the terms of this functional.
+$F_\textit{hs}$, $F_\textit{disp}$ and $F_\textit{assoc}$.  In the
+following sections, we will introduce the terms of this functional.
 
 \subsection{Ideal gas functional}
-The first term is the ideal gas free energy functional, which
-incorporates effects of kinetic energy.  The ideal gas functional is
-given by
+The first term is the ideal gas free energy functional, which is
+purely local:
 \begin{equation}\label{idealgas}
@@
 \end{equation}
-The ideal gas free energy functional satisfies the contact value
-theorem and its limiting case of small solutes (Equations
+The ideal gas free energy functional on its own satisfies the contact
+value theorem and its limiting case of small solutes (Equations
 \ref{eq:contact} and \ref{contactvaluethm}). These properties are
-retained by the total functional, provided the remaining terms are
+retained by the total functional, provided all the remaining terms are
 \emph{purely nonlocal}.
@@
 
-We treat the repulsive interactions using the White Bear version of
+We treat the hard-sphere repulsive interactions using the White Bear version of
 the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
 fluid~\cite{roth2002whitebear}.  In the homogeneous limit, the White
-Bear functional reduces to the Mansoori-Carnahan-Starling-Leland
+Bear functional reduces to the Carnahan-Starling
 equation of state.  FMT functionals are expressed as the integral of
@@
 measures of quantities such as the filling fraction, density of
-spheres touching a given point and mean curvature.  The excess free
-energy is written as:
+spheres touching a given point and mean curvature.  The hard-sphere
+excess free energy is written as:
 \begin{equation}
@@
 \subsection{Dispersion free energy}
-The first term in the attractive energy is the dispersion term.  This
-will include van der Waals attraction and any orientation-independent
-interactions. We use a dispersion term based on the SAFT-VR
-approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters: an interaction energy $\epsilondisp$ and a
-length scale $\lambdadisp R$.
+The dispersion free energy includes the van der Waals attraction and
+any orientation-independent interactions. We use a dispersion term
+based on the SAFT-VR approach\cite{gil-villegas-1997-SAFT-VR}, which
+has two free parameters (taken from Clark~\emph{et
+  al}\cite{clark2006developing}): an interaction energy $\epsilondisp$
+and a length scale $\lambdadisp R$.
 
@@
 using the local compressibility approximation (LCA), which
-assumes the energy fluctuation is directly related to the
+assumes the energy fluctuation is simply related to the
 compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
@@
 The form of $a_1$ and $a_2$ for SAFT-VR is given in
-reference~\cite{gil-villegas-1997-SAFT-VR} but is expressed in terms
+reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
 of the filling fraction.  In order to apply this form to an
-inhomogeneous density distribution, we construct an effective local
-filling fraction for dispersion $\etadisp$, given by
+\emph{inhomogeneous} density distribution, we construct an effective local
+filling fraction for dispersion $\etadisp$, given by a Gaussian
+convolution of the density:
 \begin{align}
@@
 %             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
-  \\
   \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
@@
 This effective filling fraction is used throughout the dispersion
-calculations, and represents a filling fraction averaged over the
+functional, and represents a filling fraction averaged over the
 effective range of the dispersive interaction.  This functional
@@
 term in the free energy two empirical parameters that are fit to
-experimental data.
+experimental data (again, taken from Clark~\emph{et
+  al}\cite{clark2006developing}).
 
 The association functional we use is a modified version of Yu
-and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}. \footnote{We
-  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
-  the same functional, but their paper contains errors in the
-  association term and is not useful as a reference.}
-This functional includes the effects of density inhomogeneities in the
-contact value of the correlation function $g^{HS}_\sigma$, but is
-based on the SAFT-HS equation of state rather than the SAFT-VR
-equation of state\cite{gil-villegas-1997-SAFT-VR}. SAFT-VR is used in the
-optimal SAFT parametrization for water found by Clark \emph{et
-  al}\cite{clark2006developing}, but to use it's
-equation of state involves adding a correction term (see
-Equation~\ref{eq:gSW}).
+and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, which
+% \footnote{We
+%  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
+%  the same functional, but their paper contains errors in the
+%  association term and is not useful as a reference.}
+includes the effects of density inhomogeneities in the contact value
+of the correlation function $g^{HS}_\sigma$, but is based on the
+SAFT-HS model, rather than the SAFT-VR
+model\cite{gil-villegas-1997-SAFT-VR}, which is used in the optimal
+SAFT parametrization for water of Clark \emph{et
+  al}\cite{clark2006developing}.  Adapting Yu and Wu's association
+free energy to SAFT-VR, simply involves the addition of a correction
+term in the correlation function (see Equation~\ref{eq:gSW}).
 
@@
 measure of the density inhomogeneity.
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
+\end{align}
 %
@@
 \end{align}
-where the factor $\zeta$ from Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating} describes the
-inhomogeneity of the fluid density, and reduces to unity in the
-homogeneous limit:
-\begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
-\end{align}
-The functional $\Delta$ which appears in the formula for $X$ is a
-measure of hydrogen-bonding probability.  In the SAFT-VR
-model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
+where the functional $\Delta$ which is a measure of hydrogen-bonding
+probability, given by
 \begin{align}
@@
   \left(e^{-\beta\epsilonassoc} - 1\right) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma\left(\frac{4\pi R^3}{3}n_0(\xx)\right) +
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
   \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
@@
 contact for a hard-sphere fluid with a square-well attractive
-potential, and $a_1$ and $a_2$ are two terms in the dispersive term of
-the free energy.  The correlation function $g^\textit{SW}_\sigma$ is
-written as a perturbative correction to the hard-sphere fluid correlation
-function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
+potential, and $a_1$ and $a_2$ are the two terms in the dispersion
+free energy.  The correlation function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere fluid
+correlation function $g^\textit{HS}_\sigma$, for which we use the
+functional of Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
 \begin{align}
-  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
-  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
-  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
+  g_\sigma^{HS} &= \frac{1}{1-n_3}
+  +\frac32\frac{\zeta\zeta_2}{(1-n_3)^2}
+  + \frac12\frac{\zeta_2^2\zeta}{(1-n_3)^3} \\
+  \zeta_2 &= \frac{R}{3}n_2
 \end{align}

### 2012-06-23 papers/water-SAFT/paper.tex
edit rest of water-SAFT paper Tested everything on kipu.

@@
 describing hydrogen bonds, and the final term is the chain formation
-energy of for fluids that (unlike water) are chain polymers.
-While a number of forumlations of SAFT have been
-published, we will focus on SAFT-VR\cite{gil-villegas-1997-SAFT-VR},
-which was used by Clark \emph{et al} to construct an optimal SAFT
-model for water\cite{clark2006developing}. All but one of the six
-empirical parameters used in the functional introduced in this paper
-are taken directly from this Clark \emph{et al} paper.  The phase
-diagram for this optimal SAFT model for water is shown in
+energy of for fluids that (unlike water) are chain polymers.  While a
+number of forumlations of SAFT have been published, we will focus on
+SAFT-VR\cite{gil-villegas-1997-SAFT-VR}, which was used by Clark
+\emph{et al} to construct an optimal SAFT model for
+water\cite{clark2006developing}. All but one of the six empirical
+parameters used in the functional introduced in this paper are taken
+directly from this Clark \emph{et al} paper.  As an example of the
+power of this model, it predicts an enthalpy of vaporization at
+100$^\circ$C of $\Delta H_{vap}$= 39.41~kJ/mol, compared with the
+experimental value $\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater},
+with an error of only a few percent.  We show the phase diagram for
+this optimal SAFT model for water in
 Figure~\ref{fig:pressure-with-isotherms}, which demonstrates that its
@@
 the liquid density shows larger discrepancies.  The critical point is
-very poorly described, which is a common failing of mean-field
-expansions.
+very poorly described, which is a common failing of models based on a
+mean-field expansion.
 
@@
 empirical parameters of Clark \emph{et al}, we add a single additional
-parameter $\lscale$---with a fitted value of 0.72---which determines
+parameter $\lscale$---with a fitted value of 0.353---which determines
 the length scale over which the density is averaged when computing the
-dispersion free energy.  We use this final parameter to fit the
-surface tension with the result shown in
-Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
-\emph{et al} overestimates the critical temperature---which is a
-common feature of models which do not explicitly treat the critical
-point---we cannot correctly describe the surface tension at all
-temperatures, and choose to fit it for the temperature range at which
-water is liquid at one atmosphere of pressure.
+dispersion free energy and its derivative.  We determine this final
+parameter by fitting the to the experimental surface tension with the
+result shown in Figure~\ref{fig:surface-tension}.  Because the SAFT
+model of Clark \emph{et al} overestimates the critical
+temperature---which is a common feature of SAFT-based functionals that
+do not explicitly treat the critical point---we cannot reasonably
+describe the surface tension at all temperatures, and choose to fit
+the surface tension at and around room temperature.
 
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Omega$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Omega[n] = F[n] + \mu \int n(\xx) d\xx.
-\end{equation}
-
-The pressure would naturally be obtained by taking a partial
-derivative with respect to volume at fixed temperature. 
-Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
-isotherms ranging from room temperature to past the critical temperature. The
-solid line is the theoretical coexistence curve calculated from the SAFT
-functional. For a given isotherm, the pressure curve should intersect this
-coexistence curve twice, at the vapor and liquid coexistence 
-pressure corresponding to the vapor and liquid densities. Even
-though the theoretical pressure isotherms drop below the coexistence curve,
-this is non-physical and will not actually occur. The experimental data shows
-how the actual density value will ``jump'' to the other side of 
-the coexistence curve at that pressure value (for
-temperatures less than the critical temperature), corresponding to a phase 
-change in the material. The
-functional matches well with the experimental data from NIST away from the
-critical point. The slope of the pressure versus density curve is proportional
-to the compressibility, which is close but slightly too large at room
-temperature.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\columnwidth]{figs/equation-of-state}
-%\end{center}
-%\caption{The theoretical versus experimental phase diagram, vapor
-%  pressure versus temperature.  }
-%\label{fig:equation-of-state}
-%\end{figure}
-
-%The entropy is naturally obtained by taking a partial derivative with
-%respect to temperature:
-%\begin{align}
-%  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V}
-%\end{align}
-%This derivative is trivial for the ideal gas and hard sphere free
-%energies, which are simply proportional to the temperature.  For the
-%association and dispersion free energies, it is a little more work,
-%but still not hard.
-
-%Once we have the entropy, finding the internal energy is simple:
-%\begin{align}
-%  U[n] &= F[n] + TS[n]
-%\end{align}
-%and the enthalpy is not much more work:
-%\begin{align}
-%  H[n] &= F[n] + TS[n] + pV.
-%\end{align}
-
-The calculated heat of vaporization at atmospheric pressure is
-$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
-373 K. This compares well to the experimental
-value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
-an error of only a few percent.
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-%\end{center}
-%\caption{The theoretical versus experimental coexistence curve. Away from
-%the critical temperature, the agreement is good.\fixme{(keep figure?
-%its the same as Clark paper)} }
-%\label{fig:temperature-vs-density}
-%x\end{figure}
-
-The fluid density at a liquid-solid interface commonly displays
-oscillatory behavior, but the presence of density oscillations at
-liquid-vapor interfaces is more debated\cite{penfold2001structure}.
-In fact, the \emph{meaning} of the density profile at a liquid-vapor
-interface is clouded by the existence of long-wavelength perturbations
-known as \emph{capillary waves}.  These waves are not accounted for in
-density-functional theory (although the exact
-density-functional theory would include them), with the result
-that the density profile will be smoothed out by capillary wave
-effects. We set up a liquid-vapor interface and observed small density
-oscillations in the water near the surface.
-
-%For reference, we plot the predicted vapor pressure (see
-%Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
-%Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
-%match those calculated by Clark \emph{et al}. 
-%The functional provides a good description of
-%the vapor pressure, which has an error of less than 
-%1\% compared with the experimental data. The coexistence curve in 
-%Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
-%curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
-%temperature coexistence values.
-%Near the critical temperature,
-%the agreement is off, and the theoretical prediction  of the temperature 
-%is quite a bit higher than
-%experiment. 
-%To find the densities for the coexistence curve, we use the
-%``common tangent'' rule. The two densities 
-%will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
-%As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
-%(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
-%in correctly treating the critical temperature. As the vapor and liquid densities approach
-%each other, the free energy barrier between the two states gets smaller and it becomes
-%challenging to computationally find the correct minimum. This is the reason 
-%why the coexistence curves
-%in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
-%are not smooth around the critical point.
-% 
-% To set up the calculations for the slit, we choose a chemical potential 
-% corresponding to a pressure of 1~atm and set the
-% temperature to a typical room temperature of 298~K. We apply 
-% a constraining potential outside the slit, where we choose a small non-zero
-% density for numerical reasons. The slits studied range in size from 0.1~nm to 
-% 8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
-% lattice
-% with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
-% 
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/density-1D}
-% \end{center}
-% \caption{ Density profile for a slab of water with molecules 
-% constrained to move in one dimension. The slit size is 4.0~nm and the dashed
-% line represents the saturated liquid density.}
-% \label{fig:density-1D}
-% \end{figure}
-% 
-% The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
-% Near the interface, there are oscillations around the
-% liquid density which decrease near the center of the slit. We expect the center
-% of the slit to act approximately like bulk water, and the absence of
-% oscillations supports this. For larger slits (6.0~nm or more) the majority of
-% the water molecules in the slit act like the bulk. 
-% 
-% Based on our computations, we determine the smallest slit that can contain 
-% liquid water. This result is
-% approximately 0.6~nm. For smaller slits, the initial liquid water inside will
-% vaporize. The results for larger liquid-filled slits are only metastable,
-% however, as a vapor-filled slit is thermodynamically favorable. Based on our
-% calculations for a pressure of 1~atm, the work needed to pull apart
-% the two hard walls comprising the slit is one to two orders of magnitude less
-% than the free energy of the liquid-filled slit. For the range of slit sizes we
-% studied, all the computational results of liquid-filled slits are
-% metastable. 
-% 
-% Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
-% stability of water between large hydrophobic plates and the critical separation
-% necessary for stability. For water at room temperature and atmospheric pressure,
-% plates with a separation smaller than about 100~nm will not contain stable liquid
-% water. They go on to say that confined water has a region of metastability down
-% to separations of about 5~nm. This is comparable to the slit separation where
-% our calculations show water inside the slit acting like the bulk liquid, which
-% we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
-% also study water between two hard walls at constant chemical potential. They 
-% also conclude that the liquid state is likely metastable, and there is a free
-% energy barrier to reach the stable vapor phase. 
-% 
-% \begin{figure}
-% \begin{center}
-% \includegraphics[width=\columnwidth]{figs/xassoc-1D}
-% \end{center}
-% \caption{The number of hydrogen bonds per molecule for a slab of water 
-% with molecules constrained to move in one dimension. Each molecule may
-% contribute to
-% four bonds, although each bond involves two molecules so the
-% actual number of bonds would be halved.}
-% \label{fig:xassoc-1D}
-% \end{figure}
-% 
-% The value of $X$ behaves as expected for water constrained to a slit. We
-% show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
-% molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
-% Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
-% bonds is approximately equal to 3.5, which is expected for saturated liquid
-% water.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Results and discussion}
@@
 \subsection{One hydrophobic rod}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-single-rod}
-\end{center}
-\caption{ Density profiles for single rods of different radii. The dotted line
-represents the saturated liquid density and the points represent the
-expected contact density derived from the contact value theorem and
-calculated free energy data.}
-\label{fig:density-single-rod}
-\end{figure}
-
-We move into two dimensions by studying a single hydrophobic rod
-immersed in water. We use a lattice spacing of about 0.01~nm (0.2 bohrs), 
-which is a reasonable resolution based on several lattice spacing calculations.
-The results for this spacing of 0.01~nm are within 0.5\% of the free energy 
-and a few percent of the contact density for the highest resolution tested.
-Figure~\ref{fig:density-single-rod} shows density profiles for
-different radii rods which have the
-expected oscillations outside the rod that reduce to the bulk saturated liquid
-density as the distance away from the rods increases. We use the
-contact value theorem to derive the contact density at the water-rod
-interface.  We start with the thermodynamic expression relating
-pressure to the partial derivative of free energy with respect the
-radius of a cylinder. Using the contact value theorem and substituting in
-the free energy gradient for the pressure, we can derive values for
-the contact density from data used in Figure
-\ref{fig:energy-vs-diameter}.  As the radius of the rods approach zero we
-expect the contact density to approach the bulk density, which our
-functional does.  This self consistency shows that our functional is
-thermodynamically correct.
 
@@
 
-Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
-radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
-about 0.5~nm or so, this reaches
-an asymptote equal to the surface tension. From the inset in 
-Figure~\ref{fig:surface-tension}, the surface tension at room 
-temperature is approximately 72~mN/m. This is very close to the 
-asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m. 
-For very small radius (less than about 0.1~nm) the
-relationship between the free energy per area and radius is linear (see inset
-of Figure~\ref{fig:energy-vs-diameter}). Based on the contact value theorem
-stated in Equation~\ref{contactvaluethm}, we expect this linear relationship.
-For slightly larger yet still small rods (with a radius about 0.1~nm to 
-0.3~nm), the relationship between free energy per area and radius shows
-quadratic behavior.
+We begin by studying a single hydrophobic rod immersed in water.
+% We
+% model the density on a grid with spacing of about 0.01~nm, which gives
+% the free energy of hydration to a precision of better than a percent,
+% and the density at contact to a few percent precision.
+We show in Figure~\ref{fig:energy-vs-diameter} the free energy of
+hydration at room temperature, scaled by the solvent-accessible
+surface area of the hard rod, plotted as a function of hard-rod
+radius, which we define as the radius from which water is excluded.
+For rods with radius larger than 0.5~nm or so, this reaches a value of
+75~mN/m, which is slightly higher than the bulk surface, which will be
+approached from above in the limit of very large rods.  For very small
+radius (less than about 0.1~nm) the relationship between the free
+energy per area and radius is linear, with the slope required by the
+contact-value theorem (see Equation~\ref{contactvaluethm}).
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\end{center}
+\caption{ Density profiles for single rods of different radii. The dotted line
+represents the saturated liquid density and the points represent the
+expected contact density derived from the contact value theorem and
+calculated free energy data.}
+\label{fig:density-single-rod}
+\end{figure}
+
+We show in Figure~\ref{fig:density-single-rod} density profiles for
+different radii rods, as well as the prediction for the contact value
+of the density as a function of rod radius, as computed from the free
+energies plotted in Figure~\ref{fig:energy-vs-diameter}.  Thee
+agreement between these curves confirms that our functional satisfies
+the contact-value theorem and that our minimization is well converged.
+As expected, as the radius of the rods approach zero the
+contact density approaches the bulk density, and as the radius becomes
+large, the contact density will approach the vapor density.
 
@@
 
-We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ 
-(see Figure~\ref{fig:rods}), ranging from $d=0$ 
-(touching rods) to $d=1.3$~nm. At small
-separations there is only vapor between the rods. As the rods are
-pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. 
-Figure~\ref{fig:density-rods} 
-shows density profiles before and after this transition
-for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
-on the radii of the rods, which is about 0.54~nm for the rods shown in 
-Figure~\ref{fig:density-rods}.
-Since we are looking at only hard-walled
-non-interacting rods, the critical separation will be
-different for any system where there is an appreciable direct force between rods. 
-The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
-are very similar to
-those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
-At small separations, the oscillations around the two rods look as if the two rods were one
-solid ``stadium''-shaped object (a rectangle with semi-circles on the ends).
+We now look at the more interesting problem of two parallel hard rods
+in water, separated by a distance $d$, as shown in
+Figure~\ref{fig:rods}.  At small separations there is only vapor
+between the rods, but as the rods are pulled apart, the vapor region
+expands until a critical separation is reached at which point liquid
+water fills the region between the rods.
+Figure~\ref{fig:density-rods} shows density profiles before and after
+this transition for rods of radius 0.5~nm. This critical separation
+for the transition to liquid depends on the radii of the rods, and is
+about 0.54~nm for the rods shown in Figure~\ref{fig:density-rods}.
+The critical separation will be different for a system where there is
+a significant direct force between rods, or attraction between the
+rods and water.  At small separations, the water around the two
+rods look as if the two rods were one solid ``stadium''-shaped object
+(a rectangle with semi-circles on both ends).
 
@@
 
-To determine the critical separation, we start by considering a macroscopic surface
-tension model where the free energy in the macroscopic limit is given by 
+To understand this critical separation, we consider the free energy in
+the macroscopic limit, which is given by
 \begin{equation}
@@
 \end{equation}
-The first term describes the surface energy and the second term
-is the work needed to create a cavity of volume $V$. Since this pressure term
-scales with volume, it can be neglected relative to the surface term provided
-the length scale is small compared with $\frac{\gamma} p$. At room temperature,
-the value of $\frac{\gamma} p$ is equal to about $20~\mu m$ which is
-much larger than any of the solutes we study. We expect that for those
-larger radius 
-rods, the water on the sides of the 'stadium' configuration will bow inward 
-between the rods and the density will reduce to vapor near the center
-point where the rods are closest to each other.
+The first term describes the surface energy and the second term is the
+work needed to create a cavity of volume $V$. Since the pressure term
+scales with volume, it can be neglected relative to the surface term
+provided the length scale is small compared with $\gamma / p$, which
+is around $20~\mu m$, which is much larger than any of the solutes we
+study. For micron-scale rods, the water on the sides of the `stadium'
+configuration will bow inward between the rods and the density will
+reduce to vapor near the center point where the rods are closest to
+each other.
 
@@
 \end{equation}
-and so the force per length is equal to $2\gamma$.
+and so the force per length is equal to twice the surface tension.
 
-This is consistent with the calculation of the surface tension in
-Figure~\ref{fig:surface-tension} (72~mN/m) and the force per length
-found from the slopes of the curves in Figure
-~\ref{fig:rods-energy-vs-distance} (144~mN/m)  which
-compares the free energy per length for rods of several different
-radii. This agreement indicates that even though the rods are nanoscale, the
-water molecules will ``see'' them classically.  For large $d$, the rods 
-simply act like two separate single rods in water and the energy will
-reach an asymptote determined by the effective surface tension (shown in 
-Figure~\ref{fig:energy-vs-diameter}). In Figure
-~\ref{fig:rods-energy-vs-distance}, we see that for small $d$ all the rods 
-have the same behavior with a positive linear relationship
-between free energy per length and separation during the initial state 
-with vapor between them. This indicates that the force 
-per length on the rods is constant, and surprisingly independent of separation.
-
-To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will again use the aforementioned model and propose that 
-the critical separation, $d_c$, will occur when the free energies are equal. This 
-corresponds to equating the circumference of the stadium-shape (see 
-top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom
-of Figure~\ref{fig:density-rods}), which results in
-\begin{equation}
-d_c = (\pi-2)r.\label{criticalseparation}
-\end{equation}
-The predicted and actual computed values for $d_c$ are shown graphically in 
-Figure~\ref{fig:rods-energy-vs-distance} where there is a transition
-between the sloped and flat portion of each plot. They match fairly well for the 
-0.5~-~1.0~nm rods with differences between 1.3~-~5.5\% and we hypothesize that larger radii rods
-would also follow this trend, but smaller rods are likely to have a different
-relationship with the surrounding water molecules. This may explain why the critical
-separation result for a radius of 0.3~nm does not match the prediction
-well (32\% difference).
+We plot in Figure~\ref{fig:rods-energy-vs-distance} the computed free energy
+per unit length from our classical density functional (solid lines),
+as a function of the separation $d$, along with the free energy
+predicted by our simple macroscopic model (dashed lines).  The models
+agree very well on the force between the two rods at close
+separations, and have reasonable agreement as to the critical
+separation for rods greater than 1~nm in diameter.
 
@@
 
-Walther \emph{et al.}\cite{walther2004hydrodynamic} studied the interactions between two
-carbon nanotubes, which are geometrically similar to our hydrophobic rods. 
-They use molecular dynamics with Lennard-Jones potentials for
-nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
-They also find that a drying transition occurs, and find it to be at separations near
-0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than
-what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
-and carbon-water interactions.
+Walther \emph{et al.}\cite{walther2004hydrodynamic} studied the
+interactions between two carbon nanotubes, which are geometrically
+similar to our hydrophobic rods.  They use molecular dynamics with
+Lennard-Jones potentials for nanotubes of diameter 1.25~nm and
+separations ranging from about 0.3~nm to 1.5~nm, and find that a
+drying transition occurs, and find it to be at separations near
+0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than what
+we find for hydrophobic rods of a similar size, as expected due to the
+carbon-water interactions.
 
@@
 
-We simulate four hydrophobic rods in water using the same radii as our
-two rods ($r = 0.3$~nm~-~$1.0$~nm) and
-separation distances between $d = 0$~nm to $d = 3$~nm for nearest neighbor rods. We placed
-the rods in a 'square' arrangement and observed a similar behavior as the two rods
-where vapor is present in the space between the rods until a critical separation
-where it becomes filled with water and the rods behave as separate single rods
-(shown in Fig. \ref{fig:density-four-rods}).  Free energy per length
-is plotted against rod separation in Figure
-\ref{fig:four-rods-energy-vs-distance} with a slight disagreement of
-the predicted force per unit length.
-
-The 'rounded square' formed by the area of the four rods and the space
-between them has a circumference estimated by
-\begin{equation}
-C_{rs} = 2 \pi r + 8r + 4d.
-\end{equation}
-As with the two rods, we equate this circumference with the total circumference of
-the four separate rods. This gives a critical separation of
-\begin{equation}
-d_{c4} = \frac{3 \pi - 4}{2} r. \label{criticalfourrods}
-\end{equation}
-We see in Figure \ref{fig:four-rods-energy-vs-distance} that the $r = 0.3$~nm
-rod has the largest difference between simulation and prediction with
-a 23\% difference, and better agreement for larger radii rods (between 0.5\% and 5.6\%).
+We go on to study four parallel hard rods, as examined by Lum,
+Chandler and Weeks in their classic paper on hydrophobicity at small
+and large length scales~\cite{lum1999hydrophobicity}.  As in the case
+of two rods---and as predicted by Lum~\emph{et al}---we observe a
+drying transition, as seen the density plot shown in
+Figure~\ref{fig:density-four-rods}.  In
+Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the free energy
+of hydration together with the macroscopic approximation, and find
+good agreement for rods larger than 1~nm in diameter.  \fixme{ Add
+  data from LCW paper to the plot? I ended up cutting what felt like
+  needless verbiage (and excessive explanation of the macroscopic
+  model), but now it seems too short. }
 
@@
 
-A common model of hydrophobic solutes is the \emph{hard sphere solute}
-\cite{sedlmeier2011entropy}.  The effect of this solute is merely to create
-a cavity within which the water may not penetrate.  This is the
-simplest possible solute, and will serve as a test case, which if
-passed could lead to the ability to insert different chemicals into
-aqueous solution, instead of a vapor cavity, in order to observe
-biologically and chemically important aqueous solvation interactions.
-
-As in the single rod, we begin by examining the free energy of
-hydration, expressed in Figure~\ref{fig:sphere-energy-vs-diameter} as
-the ratio of the energy of the cavity system to the surface area of
-the cavity.  This \emph{effective surface tension}, approaches the
-bulk surface tension as the cavity radius increases.  As with the
-single rod, we see the analytically correct behavior in the limit of
-small solutes.  We also plot the surface tension calculated using a
-molecular dynamics simulation of SPC/E water\cite{huang2001shs}.  The
-agreement is quite excellent, apart from the issue that the SPC/E
-model for water significantly underestimates the surface tension of
-water at room temperature\cite{vega2007surface}.
+A common model of hydrophobic solutes is the hard-sphere solute, which
+is the simplest possible solute, and serves as a test case for
+understanding of hydrophobic solutes in
+water\cite{sedlmeier2011entropy}.  As in the single rod, we begin by
+examining the free energy of hydration, expressed in
+Figure~\ref{fig:sphere-energy-vs-diameter} as the ratio of the energy
+of the cavity system to the solvent-accessible surface area.  This
+effective surface tension, surpasses the bulk surface tension at a
+radius of almost 1~nm, and will eventually drop to the bulk value.  As
+with the single rod, we see the analytically correct behavior in the
+limit of small solutes.  For comparison, we plot the free energy
+calculated using a molecular dynamics simulation of SPC/E
+water\cite{huang2001shs}.  The agreement is quite good, apart from the
+issue that the SPC/E model for water significantly underestimates the
+surface bulk tension of water at room
+temperature\cite{vega2007surface}.
 
@@
 
-The classical density functional we developed by combining fundamental-measure 
-theory with SAFT is able to produce reasonable results for hydrophobic interactions.
-We added the parameter $\lscale$ to the functional and fit it to the surface tension
-at room temperature, which we find to be consistent with results for larger hydrophobic
-rods.
+We have developed a classical density functional for water that
+combines SAFT with the fundamental-measure theory for hard spheres,
+using one additional empirical parameter beyond those in the SAFT
+equation of state, which is used to match the experimental surface
+tension.  This functional does \emph{not} make a local density
+approximation, and therefore correctly models water at both small and
+large length scales.  In addition, like all FMT functionals, this
+functional is expressed entirely in terms of convolutions of the
+density, which makes it efficient to compute and minimize.
 
-The free energy per area for water surrounding a hydrophobic rod shows the
-expected behavior with an asymptote equal to the surface tension.
-Based on this result, we expect to be
-able to classically describe a single rod with a radius of about 0.5~nm or larger. 
-For two hydrophobic rods surrounded by water, we see a transition from a vapor-filled
-stadium shape to two rods with liquid between them. A
-classical treatment of the critical separation for this transition
-works well for pairs of rods with radii larger than 0.5~nm.  Further evidence for the
-accuracy of this classical treatment is the calculation of the force on the
-stadium shape. The force per length is constant and equal to twice the
-surface tension. A similar system with four hydrophobic rods exhibited the same
-behavior while simulations of hydrophobic spheres agrees with
-SPC/E calculations.
-
-For the hydrophobic systems studied, this functional is more computationally 
-efficient than molecular dynamics. This allows for future work
-involving hydrophilic interactions with more complex electrostatic interactions.
+We apply this functional to the case of hard hydrophobic rods and
+spheres in water.  For systems of two or four hydrophobic rods
+surrounded by water, we see a transition from a vapor-filled state a
+liquid-filled state.  A simple model treatment for the critical
+separation for this transition works well for rods with diameters
+larger than 1~nm.  In the case of spherical solutes, we find good
+agreement with SPC/E simulations.
 

### 2012-06-26 papers/water-SAFT/paper.tex
add fixme to water paper Tested everything on node08.

@@
 density-functional framework, such as the contact-value theorem, which
-ensures a correct free energy of hydration for small hard solutes.
+ensures a correct free energy of hydration \fixme{change this to
+  excess chemical potential?} for small hard solutes.
 Much of the research on classical density-functional theory has

### 2012-06-29 papers/water-SAFT/paper.tex
rename four-rods density files (and edit paper) Tested everything on kipu.

@@
 \begin{equation}
-  A(T) = \underset{n(\rr)}{min}\left\{ F[n(r),T] + \int V_\textit{ext}(\rr) n(\rr)
+  A(T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int V_\textit{ext}(\rr) n(\rr)
 d\rr\right\}.
@@
 where $A(T)$ is the Helmholtz free energy of a system in the external
-potential $V_\textit{ext}$ at temperature $T$, $F[n(r),T]$ is a
+potential $V_\textit{ext}$ at temperature $T$, $F[n(\rr),T]$ is a
 universal free-energy functional, independent of external potential,
@@
 A number of classical density functionals have been developed for
-water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,gloor2004accurate,
-  gloor2007prediction, Jaqaman2004, clark2006developing,
-  lischner2010classical, fu2005vapor-liquid-dft,
+water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft, 
+  gloor2004accurate, gloor2007prediction, Jaqaman2004, clark2006developing,
+  lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
   blas2001examination}, each of which
@@
   yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft,gloor2002saft,
+  fu2005vapor-liquid-dft,gloor2002saft,muller2001molecular,
   clark2006developing, gloor2007prediction, gloor2004accurate,
@@
 which are commonly used to study the surface tension as a function of
-temperature\cite{clark2006developing, gloor2004accurate,
-  gloor2007prediction, blas2001examination, gloor2002saft}.  Such
+temperature\cite{clark2006developing, gloor2004accurate, kahl2008modified,
+  gloor2007prediction, blas2001examination, kiselev2006new,
+  gloor2002saft,fu2005vapor-liquid-dft, gross2009density,
+  yang1994density}.  Such
 functionals have qualitatively predicted the dependence of surface
@@
 approximation\cite{gloor2002saft,clark2006developing, gloor2007prediction,
-gloor2004accurate, gross2009density, kahl2008modified, blas2001examination}.
+gloor2004accurate, gross2009density, kahl2008modified,
+blas2001examination, kiselev2006new}.
 
@@
   g_\sigma^{HS} &= \frac{1}{1-n_3}
-  +\frac32\frac{\zeta\zeta_2}{(1-n_3)^2}
-  + \frac12\frac{\zeta_2^2\zeta}{(1-n_3)^3} \\
-  \zeta_2 &= \frac{R}{3}n_2
+  +\frac{R}{2}\frac{\zeta n_2}{(1-n_3)^2}
+  + \frac{R^2}{18}\frac{\zeta n_2^2}{(1-n_3)^3}
 \end{align}

### 2012-06-29 src/haskell/SomeFunctionals.hs

@@ -100,9 +100,8 @@ yuwu_zeta = var "zeta_yuwu" "{\\zeta}" $ (n2**2 - sqr_n2v)/n2**2
 
 yuwu_correlation :: Expression RealSpace
 yuwu_correlation = var "ghsyuwu" "g_{HS}^{\\textit{YuWu}}" ghs
-  where ghs = invdiff*(1 + 0.5*(invdiff*zeta2)*yuwu_zeta*(3 + invdiff*zeta2))
-        zeta2 = var "zeta2" "\\zeta_2" $ rad*n2/3
-        invdiff = 1/(1-n3)
+    where ghs = 1/(1-n3) + rad/2*n2*yuwu_zeta/(1-n3)**2 +
+                rad**2/18*n2**2*yuwu_zeta/(1-n3)**3
 
 lambda_dispersion, epsilon_dispersion :: Type a => Expression a
 lambda_dispersion = s_tex "lambda_dispersion" "\\lambda_d"

### 2012-07-06 papers/contact/paper.tex

@@
 \section{Introduction}
+
+\fixme{Figure out whether to stick with both $R$ and $\sigma$, or move
+to only $\sigma$ in the paper, which would have a certain appeal.}
 
@@
 position $\rr$.  The density $n_0$ is one of the fundamental measures
-of Fundamental Measure Theory (FMT), and seems ideal for treating
+of Fundamental Measure Theory (FMT), and is ideal for treating
 touching spheres, as illustrated in Figure~\ref{fig:n0}.
@@
 In this case, the two-particle density is averaged over spheres
-touching a sphere that is aat the given position $\rr$.
+touching a sphere that is at the given position~$\rr$.
 
@@
 \section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
-One existing approximation for the ccorrelation function is that of
+One approximation for the correlation function is that of
 Gross\cite{gross2009density}, which is of the asymmetrically averaged
@@
 
-\fixme{I (DJR) haven't yet edited this section!}
-Yu and Wu developed a functional that should correspond to the contact
-density at a given contact
-position~$g^a$~\cite{yu2002fmt-dft-inhomogeneous-associating}.  We
-will compare both their functional as well as ours with simulation
-data.  Yu and Wu use a symmetric formulation for $g_\sigma$, but
-instead of using $n_0$ as the associated density, it uses a density
-given by
+Yu and Wu developed a functional for the correlation function
+evaluated at contac, which should correspond to the symmetrically
+averaged correlation
+function~$g_\sigma^S$~\cite{yu2002fmt-dft-inhomogeneous-associating}.
+Although Yu and Wu use a symmetric formulation for $g_\sigma$, instead
+of using $n_0$ as the associated density, they use a density given by
 \begin{align}
@@
 where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact.  Given this definition of density, their $g_\sigma$ is
-given by:
+point of contact.  Given this definition of density, their $g_\sigma$
+is given by: \fixme{Which of the following two equations is more clear
+  to you? We only want one.}
 \begin{align}
@@
     + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
-    \\
+\end{align}
+From this function, we extract a symmetric correlation function
+\begin{align}
   g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
@@
 \end{align}
-I think the adjusted functional $g_\sigma^{\text{Yu,}S}$ can be directly
-compared with our $g_\sigma^S$.  It is implemented in our code.
+which corresponds to our $g_\sigma^S$ and can be directly compared
+with it.
 
@@
 \begin{figure}
-  \includegraphics[height=10cm]{figs/walls-10}
+  \includegraphics[width=\columnwidth]{figs/walls-10}
   \caption{Density and contact densities at a wall with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=10cm]{figs/walls-30}
+  \includegraphics[width=\columnwidth]{figs/walls-30}
   \caption{Density and contact densities at a wall with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=10cm]{figs/walls-40}
+  \includegraphics[width=\columnwidth]{figs/walls-40}
   \caption{Density and contact densities at a wall with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-walls-10}
+  \includegraphics[width=\columnwidth]{figs/gnn-walls-10}
   \caption{Correlation multiplied twice by density with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-walls-30}
+  \includegraphics[width=\columnwidth]{figs/gnn-walls-30}
   \caption{Correlation multiplied twice by density with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-walls-40}
+  \includegraphics[width=\columnwidth]{figs/gnn-walls-40}
   \caption{Correlation multiplied twice by density with bulk filling
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/contact-08-013}
+  \includegraphics[width=\columnwidth]{figs/contact-08-013}
   \caption{Density and contact densities of thirteen hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/contact-12-112}
+  \includegraphics[width=\columnwidth]{figs/contact-12-112}
   \caption{Density and contact densities of 112 hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/contact-16-265}
+  \includegraphics[width=\columnwidth]{figs/contact-16-265}
   \caption{Density and contact densities of 265 hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-sphere-08-013}
+  \includegraphics[width=\columnwidth]{figs/gnn-sphere-08-013}
   \caption{Correlation multiplied twice by density of 13 hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-sphere-12-112}
+  \includegraphics[width=\columnwidth]{figs/gnn-sphere-12-112}
   \caption{Correlation multiplied twice by density of 112 hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=9cm]{figs/gnn-sphere-16-265}
+  \includegraphics[width=\columnwidth]{figs/gnn-sphere-16-265}
   \caption{Correlation multiplied twice by density of 265 hard spheres in a
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/inner-4-30}
+  \includegraphics[width=\columnwidth]{figs/inner-4-30}
   \caption{Density and contact densities around a hard sphere with
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/inner-4-40}
+  \includegraphics[width=\columnwidth]{figs/inner-4-40}
   \caption{Density and contact densities around a hard sphere with
@@
 \begin{figure}
-  \includegraphics[height=5cm]{figs/inner-8-30}
+  \includegraphics[width=\columnwidth]{figs/inner-8-30}
   \caption{Density and contact densities around a hard sphere with
@@
 \begin{figure}
-\includegraphics[width=5cm]{figs/gHS-vs-n}
+\includegraphics[width=\columnwidth]{figs/gHS-vs-n}
 \caption{A comparison of the various approximations to the contact
@@
 \begin{figure}
-\includegraphics[width=5cm]{figs/free-energy}
+\includegraphics[width=\columnwidth]{figs/free-energy}
 \caption{A comparison of the various approximations to the free energy

### 2012-07-08 papers/water-SAFT/paper.tex

@@
   rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000,
-  roth2002whitebear}, which are entirely expressed as an integral of
+  roth2002whitebear}.  These functionals are entirely expressed as an integral of
 local functions of a few convolutions of the density that can be
@@
 describing hydrogen bonds, and the final term is the chain formation
-energy of for fluids that (unlike water) are chain polymers.  While a
-number of forumlations of SAFT have been published, we will focus on
+energy for fluids that (unlike water) are chain polymers.  While a
+number of formulations of SAFT have been published, we will focus on
 SAFT-VR\cite{gil-villegas-1997-SAFT-VR}, which was used by Clark
@@
 perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
-the mean-field dispersion interaction and the second term, $a_2$, describes the
+the mean-field dispersion interaction. The second term, $a_2$, describes the
 effect of fluctuations resulting from compression of the fluid due
-to the dispersion interaction itself. $a_2$ is approximated
+to the dispersion interaction itself; it is approximated
 using the local compressibility approximation (LCA), which
@@
   al}\cite{clark2006developing}.  Adapting Yu and Wu's association
-free energy to SAFT-VR, simply involves the addition of a correction
+free energy to SAFT-VR simply involves the addition of a correction
 term in the correlation function (see Equation~\ref{eq:gSW}).
@@
 \end{align}
-where the functional $\Delta$ which is a measure of hydrogen-bonding
+where the functional $\Delta$ is a measure of hydrogen-bonding
 probability, given by
@@
 75~mN/m, which is slightly higher than the bulk surface, which will be
-approached from above in the limit of very large rods.  For very small
+approached from above in the limit of very large rods \fixme{this is a
+terrible sentence, want to say something more concrete about free
+energy / area dropping (getting data) DJR}.  For very small
 radius (less than about 0.1~nm) the relationship between the free
@@
 of the density as a function of rod radius, as computed from the free
-energies plotted in Figure~\ref{fig:energy-vs-diameter}.  Thee
+energies plotted in Figure~\ref{fig:energy-vs-diameter}.  The
 agreement between these curves confirms that our functional satisfies
 the contact-value theorem and that our minimization is well converged.
-As expected, as the radius of the rods approach zero the
+As expected, as the radius of the rods becomes zero the
 contact density approaches the bulk density, and as the radius becomes
@@
 rods and water.  At small separations, the water around the two
-rods look as if the two rods were one solid ``stadium''-shaped object
+rods looks as if the two rods were one solid ``stadium''-shaped object
 (a rectangle with semi-circles on both ends).
@@
 scales with volume, it can be neglected relative to the surface term
-provided the length scale is small compared with $\gamma / p$, which
-is around $20~\mu m$, which is much larger than any of the solutes we
+provided the length scale is small compared with $\gamma / p$. The
+value of $\gamma / p$ is around $20~\mu m$, which is much larger than
+any of the solutes we
 study. For micron-scale rods, the water on the sides of the `stadium'
@@
 separations, and have reasonable agreement as to the critical
-separation for rods greater than 1~nm in diameter.
+separation for rods greater than 0.5~nm in radius.
 
@@
 \end{center}
-\caption{ Free energy of hydration (also known as the potential of mean force) 
+\caption{ \fixme{Free energy of hydration} (also known as the potential of mean force) 
 versus separation for two hydrophobic rods ranging in radius from
@@
 of the cavity system to the solvent-accessible surface area.  This
-effective surface tension, surpasses the bulk surface tension at a
+effective surface tension surpasses the bulk surface tension at a
 radius of almost 1~nm, and will eventually drop to the bulk value.  As

### 2012-07-12 papers/water-SAFT/paper.tex
Changes to paper figures (some need fixing still).  Updated two rod energy-vs-distance. Removed energy-rods.gp since it is not being used. Tested everything on node04.

@@
 density-functional framework, such as the contact-value theorem, which
-ensures a correct free energy of hydration \fixme{change this to
-  excess chemical potential?} for small hard solutes.
+ensures a correct excess chemical potential for small hard solutes.
 Much of the research on classical density-functional theory has
@@
 \end{center}
-\caption{ Free energy of hydration versus radius for a single hydrophobic rod
+\caption{ \fixme{Free energy of hydration} versus radius for a single hydrophobic rod
 immersed in water. This should have an asymptote equal to the surface
@@
 % and the density at contact to a few percent precision.
-We show in Figure~\ref{fig:energy-vs-diameter} the free energy of
-hydration at room temperature, scaled by the solvent-accessible
+We show in Figure~\ref{fig:energy-vs-diameter} the excess chemical
+potential \fixme{does this work?} at room temperature, scaled by the solvent-accessible
 surface area of the hard rod, plotted as a function of hard-rod
@@
 in water, separated by a distance $d$, as shown in
-Figure~\ref{fig:rods}.  At small separations there is only vapor
+Figure~\ref{fig:density-rods}.  At small separations there is only vapor
 between the rods, but as the rods are pulled apart, the vapor region
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-diagram}
-\end{center}
-\caption{ Cross section schematic for two hydrophobic rods immersed in water.
-The radius and separation are shown.}
-\label{fig:rods}
-\end{figure}
+%\begin{figure}
+%\begin{center}
+%\includegraphics[width=\columnwidth]{figs/rods-diagram}
+%\end{center}
+%\caption{ Cross section schematic for two hydrophobic rods immersed in water.
+%The radius and separation are shown.}
+%\label{fig:rods}
+%\end{figure}
 
@@
 \end{center}
-\caption{ Free energy of hydration 
+\caption{ \fixme{Free energy of hydration} 
 versus separation for four hydrophobic rods ranging in radius from
@@
 Figure~\ref{fig:density-four-rods}.  In
-Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the free energy
-of hydration together with the macroscopic approximation, and find
+Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the \fixme{free energy
+of hydration} together with the macroscopic approximation, and find
 good agreement for rods larger than 1~nm in diameter.  \fixme{ Add
@@
 water\cite{sedlmeier2011entropy}.  As in the single rod, we begin by
-examining the free energy of hydration, expressed in
+examining the \fixme{free energy of hydration}, expressed in
 Figure~\ref{fig:sphere-energy-vs-diameter} as the ratio of the energy

### 2012-07-12 papers/contact/paper.tex
some very minor cleanups in contact paper Tested everything on node06.

@@
 
-The first geometry we consider is a number of hard spheres confined
+The next geometry we consider is a number of hard spheres confined
 within a spherical cavity.  In
-Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show
-the densities and contact densities for several cavity sizes.  In each
-case, the number of hard spheres was chosen to set the mean filling
-fraction to around 0.3?.  This value was chosen because it roughly
-corresponds to the filling fraction of water at standard atmospheric
-temperature and pressure in the optimal SAFT model developed by Clark
-\emph{et al}\cite{clark2006developing}.
+Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show the densities
+and contact densities for several cavity sizes \fixme{at a variety of
+  filling fractions}.  We chose cavities that are large enough that
+the center of the sphere is bulk-like, so that we can interpret the
+behavior near the interface as behavior that should be typical of the
+behavior of the hard sphere fluid near a concave surface.
 
@@
 
-It is also interesting to study a system with opposite curvature.  Put
-a (larger) hard sphere in the hard-sphere fluid, and see what the
-contact density is around that hard sphere.  This is particularly
-relevant when we consider using classical density-functional theory to
-model solvation.
+It is also interesting to study the hard sphere fluid at a convex
+surface.  We did so by placing a (larger) hard sphere in the
+hard-sphere fluid, to see how the correlation function behaves near
+the surface of that hard sphere.  This is particularly relevant when
+we consider using classical density-functional theory to model
+solvation.
 
@@
 
-\section{Formulas generated from Haskell}
-
-\input{formulas}
-\begin{align}
-  \Phi_1 &= \phione \\
-  \Phi_2 &= \phitwo \\
-  \Phi_3 &= \phithree \\
-  \frac{d F}{dn_3} &= \dwbdnthree \\
-  \frac{d F}{dn_2} &= \dwbdntwo \\
-  \frac{d F}{dn_1} &= \dwbdnone \\
-  \frac{d F}{d\vec{n}_{2v}} &= \dwbdntwov \\
-  \frac{d F}{d\vec{n}_{1v}} &= \dwbdnonev
-\end{align}
-
 \end{widetext}

### 2012-07-17 papers/contact/paper.tex
Added correlation in SomeFunctionals.hs small changes to Expresions.lhs, fixed a few things in Reciprical Shell prime functions Tested everything on node02.

--- 2012-06-20 papers/contact/paper.tex
@@
     A_{HS}}{\delta R(\mathbf{r})} \\
-  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
+  g_\sigma^S(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
   &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
@@
   simulation, and dashed lines of various sorts represent DFT model
-  predictions.  }
+  predictions.}
 
@@
   \delta(|\rr'-\rr''| - R(\rr'')) \\
-  \frac{\delta n_1(\rr')}{\delta R(\rr)^2}
+  \frac{\delta n_0(\rr')}{\delta R(\rr)}
   &= -\frac{n(\rr)}{4\pi
@@
 \begin{align}
-  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| + R(\rr))d\mathbf
+  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))d\mathbf
   r'\\
@@
 \begin{align}
-  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
+  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
     \frac{\delta n_1^{S}(\rr')}{\delta R(\rr)} &=
@@
 \begin{align}
-  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'\\
+  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
     \frac{\delta n_0^{S}(\rr')}{\delta R(\rr)}
@@
   \frac{(\rr - \rr')}{|\rr - \rr'|} d \rr'\\
-  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= \delta(\rr - \rr')
+  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr - \rr')
   \int n(\rr'') \delta'(|\rr' - \rr''| - R(\rr'))
@@
    \delta'(|\rr'-\rr''| - R(\rr')) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
-   + \mathbf{n}_{V2}^S(\rr') \right)
+   + \frac{\mathbf{n}_{V2}^S(\rr')}{4\pi R(\rr)^2} \right)
 \end{align}

### 2012-07-17 papers/contact/figs/walls.cpp

@@ -144,11 +144,11 @@ void run_walls(double eta, const char *name, Functional fhs) {
   char *plotname = (char *)malloc(1024);
   sprintf(plotname, "papers/contact/figs/walls%s-%04.1f-%04.2f.dat", name, width, eta);
   Grid energy_density(gd, f(1, gd, potential));
-  Grid correlation_S(gd, Correlation_S(1.0)(1, gd, density));
-  Grid correlation_A(gd, Correlation_A(1.0)(1, gd, density));
+  Grid correlation_S(gd, Correlation_S2(1.0)(1, gd, density));
+  Grid correlation_A(gd, Correlation_A2(1.0)(1, gd, density));
   if (strlen(name) == 4) { 
-    correlation_S = Correlation_S_WBm2(1.0)(1, gd, density);    
-    correlation_A = Correlation_A_WBm2(1.0)(1, gd, density);
+    correlation_S = Correlation_S(1.0)(1, gd, density);    
+    correlation_A = Correlation_A(1.0)(1, gd, density);
   }
   Grid gross_correlation(gd, GrossCorrelation(1.0)(1, gd, density));
   Grid n0(gd, ShellConvolve(1)(1, density)/(4*M_PI));

### 2012-07-17 src/haskell/SomeFunctionals.hs

@@ -1,5 +1,5 @@
 module SomeFunctionals 
-       ( whitebear, wb_contact_at_sphere, idealgas, mu, n,
+       ( whitebear, correlation_S_WB, correlation_A_WB, idealgas, mu, n,
          phi1, phi2, phi3,
          of_effective_potential,
          xshell, yshell, zshell,
@@ -26,24 +26,40 @@ kR = k * rad
 smear :: Expression KSpace
 smear = exp (-6.0*kdr*kdr)
 
-n, n3, n2, n1, n0, n2x, n2y, n2z, n1v_dot_n2v, sqr_n2v :: Expression RealSpace
+n, n3, n2, n1, n0, n1x, n1y, n1z, n2x, n2y, n2z, n2p, n2px, n2py, n2pz, n1v_dot_n2v, sqr_n2v :: Expression RealSpace
 n = "n" === r_var "x"
 n3 = "n3" === step n
 n2 = "n2" === shell n
-n1 = "n1" === n2 / (4*pi*rad)
-n0 = "n0" === n2 / (4*pi*rad**2)
+n2p = "n2p" === shellPrime n
+n1 = "n1" === shell n / (4*pi*rad)
+n0 = "n0" === shell n / (4*pi*rad**2)
 n2x = "n2x" === xshell n
 n2y = "n2y" === yshell n
 n2z = "n2z" === zshell n
-sqr_n2v = var "n2vsqr" "{\\vec{n}_{1v}\\cdot\\vec{n}_{2v}}" (n2x**2 + n2y**2 + n2z**2)
-n1v_dot_n2v = {- var "n1vdn2v" "{\\left|\\vec{n}_{2v}\\right|}" -} (sqr_n2v/(4*pi*rad))
+n1x = "n1x" === xshell n / (4*pi*rad)
+n1y = "n1y" === yshell n / (4*pi*rad)
+n1z = "n1z" === zshell n / (4*pi*rad)
+n2px = "n2px" === xshellPrime n
+n2py = "n2py" === yshellPrime n
+n2pz = "n2pz" === zshellPrime n
 
-shell, step, xshell, yshell, zshell :: Expression RealSpace -> Expression RealSpace
+sqr_n2v = var "n2vsqr" "{\\left|\\vec{n}_{2v}\\right|^2}" (n2x**2 + n2y**2 + n2z**2)
[…]

### 2012-07-17 src/haskell/functionals.hs

@@ -11,8 +11,10 @@ main =
                    else return ()
      gen "src/HardSpheresNoTensor2Fast.cpp" $
        defineFunctional whitebear ["R"] "HardSpheresNoTensor2"
-     gen "src/ContactAtSphereFast.cpp" $
-       generateHeader wb_contact_at_sphere ["R"] "ContactAtSphere"
+     gen "src/Correlation_S_Fast.cpp" $
+       generateHeader correlation_S_WB ["R"] "Correlation_S2"
+     gen "src/Correlation_A_Fast.cpp" $
+       generateHeader correlation_A_WB ["R"] "Correlation_A2"
      gen "src/YuWuCorrelationFast.cpp" $
        generateHeader yuwu_correlation ["R"] "YuWuCorrelationFast"
      gen "src/SaftFluid2Fast.cpp" $

### 2012-07-17 src/haskell/latex-functionals.hs

@@ -22,6 +22,7 @@ main =
      pdf "doc/Association.pdf" $ latexEasy $ "Fassoc" === saft_association
      pdf "doc/Dispersion.pdf" $ latexEasy $ "Fdisp" === saft_dispersion
      pdf "doc/SimpDispersion.pdf" $ latexSimp $ "Fdisp" === saft_dispersion
+     --pdf "doc/CorrelationS.pdf" $ latexEasy $ correlation_S_WB
      pdf "doc/SaftFluid.pdf" $ latexEasy $ "Fw" === saft_fluid
      pdf "doc/GradDispersion.pdf" $ latexEasy $ "grad" === (grad "x" saft_dispersion)
      pdf "doc/JoinedGradDispersion.pdf" $ latexEasy $ "grad" === (joinFFTs $ grad "x" saft_dispersion)

### 2012-07-19 papers/contact/paper.tex
work on sphere.cpp, and make contact plots a tad nicer Tested everything on kipu.

@@
 
-\newcommand\sphereExplanation{ Blue curves describe the density of
-  hard spheres.  Green curves describe the contact density averaged
-  according to the point of contact.  Red curves describe the contact
-  density as averaged according to the centers of the spheres that are
-  touching.  In each case, solid lines correspond to the Monte Carlo
-  simulation, and dashed lines of various sorts represent DFT model
-  predictions.}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/contact-08-013}
-  \caption{Density and contact densities of thirteen hard spheres in a
-    spherical cavity with diameter 8. \sphereExplanation }
-  \label{fig:sphere-8}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/contact-12-112}
-  \caption{Density and contact densities of 112 hard spheres in a
-    spherical cavity with diameter 12.  \sphereExplanation}
-  \label{fig:sphere-12}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/contact-16-265}
-  \caption{Density and contact densities of 265 hard spheres in a
-    spherical cavity with diameter 16. \sphereExplanation}
-  \label{fig:sphere-16}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-sphere-08-013}
-  \caption{Correlation multiplied twice by density of 13 hard spheres in a
-    spherical cavity with diameter 8.}
-  \label{fig:gnn-sphere-08-013}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-sphere-12-112}
-  \caption{Correlation multiplied twice by density of 112 hard spheres in a
-    spherical cavity with diameter 12.}
-  \label{fig:gnn-sphere-12-112}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-sphere-16-265}
-  \caption{Correlation multiplied twice by density of 265 hard spheres in a
-    spherical cavity with diameter 16.}
-  \label{fig:gnn-sphere-16-265}
-\end{figure}
-
 \subsection{Hard spheres confined in a spherical cavity}
@@
 behavior of the hard sphere fluid near a concave surface.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-30}
-  \caption{Density and contact densities around a hard sphere with
-    radius 1.}
-  \label{fig:inner-4-30}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-40}
-  \caption{Density and contact densities around a hard sphere with
-    radius 1.}
-  \label{fig:inner-4-40}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-8-30}
-  \caption{Density and contact densities around a hard sphere with
-    radius 3.}
-  \label{fig:inner-8-30}
-\end{figure}
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/outer-16-05}
+\end{center}
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/outer-16-04}
+\end{center}
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/outer-16-03}
+\end{center}
 
@@
 solvation.
+
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/inner-4-30}
+\end{center}
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/inner-4-40}
+\end{center}
+\begin{center}
+  \includegraphics[width=\columnwidth]{figs/inner-8-30}
+\end{center}
 

### 2012-07-26 papers/water-SAFT/paper.tex

@@
 \end{center}
-\caption{ \fixme{Free energy of hydration} versus radius for a single hydrophobic rod
-immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. The inset for rods with a very small 
-radius shows the linear relationship expected based on 
+\caption{ Excess chemical potential per area versus radius for a single
+hydrophobic rod immersed in water. This should have an asymptote equal
+to the surface tension at room temperature, and it agrees well with
+the surface tension in Figure~\ref{fig:surface-tension}. The inset for
+rods with a very small radius shows the linear relationship expected based on
 Equation~\ref{contactvaluethm}.}
@@
 We show in Figure~\ref{fig:energy-vs-diameter} the excess chemical
-potential \fixme{does this work?} at room temperature, scaled by the solvent-accessible
+potential at room temperature, scaled by the solvent-accessible
 surface area of the hard rod, plotted as a function of hard-rod
@@
 water\cite{sedlmeier2011entropy}.  As in the single rod, we begin by
-examining the \fixme{free energy of hydration}, expressed in
-Figure~\ref{fig:sphere-energy-vs-diameter} as the ratio of the energy
-of the cavity system to the solvent-accessible surface area.  This
+examining the ratio of the excess chemical potential of the cavity
+system to the solvent-accessible surface area
+(Figure~\ref{fig:sphere-energy-vs-diameter}).  This
 effective surface tension surpasses the bulk surface tension at a

### 2012-07-31 papers/water-SAFT/paper.tex
updated ab inition MD citation Tested everything on node10.

@@
 dynamics~\cite{car1985, grossman2004}.  The current state-of-the art
-in \emph{ab initio} molecular dynamics is limited to around 64 water
-molecules per unit cell, which places the minimum solute concentration
-at 0.84 molar and allows contact between the second solvation shell of
-neighboring solutes~\cite{izvekov2005, choe2007}\fixme{Look up more
-  recent numbers on this!}.  On top of this, standard \emph{ab initio}
+in \emph{ab initio} molecular dynamics is limited to a few hundred water
+molecules per unit cell~\cite{lewis2011doesnitric}.  On top of this,
+standard \emph{ab initio}
 methods using classical molecular dynamics without van der Waals
@@
 radius, which we define as the radius from which water is excluded.
-For rods with radius larger than 0.5~nm or so, this reaches a value of
-75~mN/m, which is slightly higher than the bulk surface, which will be
-approached from above in the limit of very large rods \fixme{this is a
-terrible sentence, want to say something more concrete about free
-energy / area dropping (getting data) DJR}.  For very small
-radius (less than about 0.1~nm) the relationship between the free
-energy per area and radius is linear, with the slope required by the
-contact-value theorem (see Equation~\ref{contactvaluethm}).
+For rods with radius larger than 0.5~nm or so, this reaches a maximum value of
+75~mN/m, which is slightly higher than the bulk surface.  In the limit
+of very large rods, this value will decrease and approach the bulk
+surface value.  For very small radius (less than about 0.1~nm) the
+relationship between the free energy per area and radius is linear,
+with the slope required by the contact-value theorem (see Equation~\ref{contactvaluethm}).
 

### 2012-08-02 papers/water-SAFT/paper.tex
Fixed all /fixme's in paper and updated figures Tested everything on node10.

@@
   gloor2002saft,fu2005vapor-liquid-dft, gross2009density,
-  yang1994density}.  Such
+  yang1994density, Jaqaman2004}.  Such
 functionals have qualitatively predicted the dependence of surface
@@
 ding1987 &  & X & ? &  \\
+\hline
+Jaqaman2004 &  & X & ? & X \\
 \hline
@@
 of very large rods, this value will decrease and approach the bulk
-surface value.  For very small radius (less than about 0.1~nm) the
-relationship between the free energy per area and radius is linear,
-with the slope required by the contact-value theorem (see Equation~\ref{contactvaluethm}).
+surface value.  As seen in the inset of
+Fig.~\ref{fig:energy-vs-diameter}, for rods with very small radius (less than
+about 0.5~\AA) the excess chemical potential is proportional to volume,
+as required by the contact-value theorem (see Equation~\ref{contactvaluethm}).
 
@@
 Figure~\ref{fig:density-rods} shows density profiles before and after
-this transition for rods of radius 0.5~nm. This critical separation
+this transition for rods of radius 0.6~nm. This critical separation
 for the transition to liquid depends on the radii of the rods, and is
-about 0.54~nm for the rods shown in Figure~\ref{fig:density-rods}.
+about 0.65~nm for the rods shown in Figure~\ref{fig:density-rods}.
 The critical separation will be different for a system where there is
@@
 \caption{ Density profiles illustrating the transition from vapor 
-to liquid water between the rods. The radius is 0.5~nm, the top figure is 
-at a separation of 0.5~nm and the
-bottom is 0.6~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
+to liquid water between the rods. The radius is 0.6~nm, the top figure is
+at a separation of 0.6~nm and the
+bottom is 0.7~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
 the energy for these and other separations.}
@@
 
-We plot in Figure~\ref{fig:rods-energy-vs-distance} the computed free energy
+We plot in Figure~\ref{fig:rods-energy-vs-distance} the computed free
+energy of interaction
 per unit length from our classical density functional (solid lines),
@@
 \end{center}
-\caption{ \fixme{Free energy of hydration} (also known as the potential of mean force) 
+\caption{ Free energy of interaction (also known as the potential of mean force)
 versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 1.0~nm.
+0.2~nm to 1.2~nm.
 All were arbitrarily offset to zero at large separations for ease of comparison. The
@@
 \end{center}
-\caption{ \fixme{Free energy of hydration} 
+\caption{ Free energy of interaction
 versus separation for four hydrophobic rods ranging in radius from
-0.3~nm to 1.0~nm.
+0.2~nm to 1.2~nm.
 All were arbitrarily offset to zero at large separations. The
@@
 \caption{ Density profiles illustrating the transition from vapor 
-to liquid water between four rods. The radius is 0.5~nm, the top figure is 
-at a separation of 1.1~nm and the
-bottom is 1.2~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
+to liquid water between four rods. The radius is 0.6~nm, the top figure is
+at a separation of 1.53~nm and the
+bottom is 1.56~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
 the energy for these and other separations.}
-
 \label{fig:density-four-rods}
@@
 Figure~\ref{fig:density-four-rods}.  In
-Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the \fixme{free energy
-of hydration} together with the macroscopic approximation, and find
-good agreement for rods larger than 1~nm in diameter.  \fixme{ Add
-  data from LCW paper to the plot? I ended up cutting what felt like
-  needless verbiage (and excessive explanation of the macroscopic
-  model), but now it seems too short. }
+Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the free energy
+of interaction together with the macroscopic approximation, and find
+good agreement for rods larger than 0.5~nm in radius.  This free
+energy plot is qualitatively similar to that predicted by the LCW
+theory~\cite{lum1999hydrophobicity}, with the minor difference that
+we find no barrier to the association of four rods.
 
@@
 \end{center}
-\caption{ Free energy versus radius for a single hydrophobic sphere
+\caption{ Excess chemical potential per area versus radius for
+  a single hydrophobic sphere
   immersed in water. This should have an asymptote equal to the

### 2012-08-06 papers/water-SAFT/paper.tex
edits on water paper (approaching submission!) Tested everything on bingley.

@@
 short length scales over a wide range of temperatures, and is
-efficient to evaluate computationally, comparable to the cost of FMT
+computationally efficient, comparable to the cost of FMT
 itself.  We demonstrate our functional by applying it to systems
@@
 biology---takes place in aqueous solution.  However, while quantum
-chemistry now enables us to calculate the ground state energies of
+chemistry enables us to calculate the ground state energies of
 large molecules in vacuum, prediction of the free energy of even the
@@
 corrections strongly over-structure water, to the point that ice melts
-at over 400K\cite{yoo2009phase}!  There has been a recent flurry of
+at over 400K\cite{yoo2009phase}!  There has been a flurry of recent
 publications implicating van der Waals effects as significant in
@@
 molecules. A more efficient approach is needed in order to study
-nanoscale and hydrophobic solutes.
+nanoscale and larger solutes.
 
@@
 where $A(T)$ is the Helmholtz free energy of a system in the external
-potential $V_\textit{ext}$ at temperature $T$, $F[n(\rr),T]$ is a
-universal free-energy functional, independent of external potential,
-and $n(\rr)$ is a density of atoms or molecules.  Classical DFT is a
+potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
+of atoms or molecules, and $F[n(\rr),T]$ is a
+universal free-energy functional for the fluid, which is independent
+of the external potential $V_\textit{ext}$.  Classical DFT is a
 natural framework for creating a more flexible theory of
-hydrophobicity that can describe interaction of water with arbitrary
-external potentials---including potentials describing strong
+hydrophobicity that can readily describe interaction of water with arbitrary
+external potentials---such as potentials describing strong
 interactions with solutes or surfaces.
@@
   roth2002whitebear}.  These functionals are entirely expressed as an integral of
-local functions of a few convolutions of the density that can be
-efficiently computed.  We will use the White Bear version of the FMT
-functional\cite{roth2002whitebear}.  In the homogeneous limit, this
-functional reduces to the Carnahan-Starling equation of state, and in
-the strongly-confined limit of a highly confined cavity, it reproduces
-the exact free energy.
+local functions of a few convolutions of the density (fundamental
+measures) that can be efficiently computed.  We will use the White
+Bear version of the FMT functional\cite{roth2002whitebear}.  This
+functional reduces to the Carnahan-Starling equation of state in the
+homogeneous limit, and it reproduces the exact free energy in the
+strongly-confined limit of a small cavity.
 
@@
 \begin{equation} \label{eq:SAFT-free-energy}
-  F[n] = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
+  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
   F_\textit{assoc} + F_\textit{chain},
 \end{equation}
-where the first three terms encompass the \emph{monomer} contribution
+where the first three terms---ideal gas, hard-sphere repulsion and
+dispersion---encompass the \emph{monomer} contribution
 to the free energy, the fourth is the \emph{association} free energy,
@@
 experimental value $\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater},
-with an error of only a few percent.  We show the phase diagram for
+with an error of only a few percent.  We show a phase diagram for
 this optimal SAFT model for water in
@@
 the liquid density shows larger discrepancies.  The critical point is
-very poorly described, which is a common failing of models based on a
-mean-field expansion.
+very poorly described, which is a common failing of models that are
+based on a mean-field expansion.
 
@@
 
-SAFT has also been used to construct classical density functionals,
-which are commonly used to study the surface tension as a function of
+SAFT has been used to construct classical density functionals,
+which are often used to study the surface tension as a function of
 temperature\cite{clark2006developing, gloor2004accurate, kahl2008modified,
@@
 
-\begin{table}
-\begin{tabular}{|c|c|c|c|c|}
-\hline
-paper & SAFT functional & water & LDA & tension \\
-\hline
-yuwu2002dft & X & & & \\
-\hline
-fuwu2005 & X & X & & X \\
-\hline
-kiselev2006 &  & X & X & X \\
-\hline
-clark2006developing & X & X & X & X \\
-\hline
-gloor2002 & X & X & X & X \\
-\hline
-gloor2004 & X & X & X & X \\
-\hline
-gloor2007 & X & X & X & X \\
-\hline
-gross2009 & X &  & X & X \\
-\hline
-Lischner2010 &  & X &  &  \\
-\hline
-Blas2001 & X & X & X & X \\
-\hline
-Kahl2008 & X &  & X & X \\
-\hline
-Muller2001 & X &  &  &  \\
-\hline
-Yang1994 &  & X &  & X \\
-\hline
-ding1987 &  & X & ? &  \\
-\hline
-Jaqaman2004 &  & X & ? & X \\
-\hline
-next & ? & ? & ? & ? \\
-\hline
-\end{tabular}
-\caption{Which papers do what}
-\end{table}
+%% \begin{table}
+%% \begin{tabular}{|c|c|c|c|c|}
+%% \hline
+%% paper & SAFT functional & water & LDA & tension \\
+%% \hline
+%% yuwu2002dft & X & & & \\
+%% \hline
+%% fuwu2005 & X & X & & X \\
+%% \hline
+%% kiselev2006 &  & X & X & X \\
+%% \hline
+%% clark2006developing & X & X & X & X \\
+%% \hline
+%% gloor2002 & X & X & X & X \\
+%% \hline
+%% gloor2004 & X & X & X & X \\
+%% \hline
+%% gloor2007 & X & X & X & X \\
+%% \hline
+%% gross2009 & X &  & X & X \\
+%% \hline
+%% Lischner2010 &  & X &  &  \\
+%% \hline
+%% Blas2001 & X & X & X & X \\
+%% \hline
+%% Kahl2008 & X &  & X & X \\
+%% \hline
+%% Muller2001 & X &  &  &  \\
+%% \hline
+%% Yang1994 &  & X &  & X \\
+%% \hline
+%% ding1987 &  & X & ? &  \\
+%% \hline
+%% Jaqaman2004 &  & X & ? & X \\
+%% \hline
+%% next & ? & ? & ? & ? \\
+%% \hline
+%% \end{tabular}
+%% \caption{Which papers do what}
+%% \end{table}
 
@@
 therefore incorrectly model small hard solutes.
-The contact-value theorem states that the pressure of any fluid at a
-hard wall interface is proportional to the contact density of the
+The contact-value theorem states that the pressure any fluid exerts
+on a hard wall interface is proportional to the contact density of the
 fluid\cite{henderson1979exact}:
@@
 and $T$ is the temperature of
-the fluid. This leads to the property that the free energy of
-solvation of a small hard solute is proportional to the
+the fluid. This leads to the property that the excess chemical
+potential of a small hard solute is proportional to the
 solvent-excluded volume:
@@
 \end{equation}
-The contact-value theorem is satisfied for functionals where the
-only purely local term is the free energy of an ideal gas, and is not
-satisfied for functionals built using a local density approximation.
+The contact-value theorem is satisfied by classical density
+functionals in which the only purely local term is the ideal gas
+contribution to the free energy, and conversely, this theorem is not
+satisfied by functionals built using a local density approximation.
 
@@
 \ref{eq:contact} and \ref{contactvaluethm}). These properties are
-retained by the total functional, provided all the remaining terms are
-\emph{purely nonlocal}.
+retained by our total functional, since all the remaining terms are
+purely nonlocal.
 
@@
 the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
-fluid~\cite{roth2002whitebear}.  In the homogeneous limit, the White
-Bear functional reduces to the Carnahan-Starling
-equation of state.  FMT functionals are expressed as the integral of
+fluid~\cite{roth2002whitebear}.  FMT functionals are expressed as the integral of
 the \emph{fundamental measures} of a fluid, which provide local
@@
 The density $n_3$ is the filling fraction and $n_2$ describes the number
-of spheres touching a given point. For our water functional, we use the 
+of spheres touching a given point. For our functional for water, we use the 
 hard-sphere radius of
@@
 effect of fluctuations resulting from compression of the fluid due
-to the dispersion interaction itself; it is approximated
+to the dispersion interaction itself, and is approximated
 using the local compressibility approximation (LCA), which
@@
 functional, and represents a filling fraction averaged over the
-effective range of the dispersive interaction.  This functional
-introduces an additional empirical parameter $\lscale$ which adjusts
+effective range of the dispersive interaction.  Here we have
+introduced an additional empirical parameter $\lscale$ which modifies
 the length scale over which the dispersion interaction is correlated.
@@
 term in the free energy two empirical parameters that are fit to
-experimental data (again, taken from Clark~\emph{et
-  al}\cite{clark2006developing}).
+the experimental equation of state of water (again, taken from
+Clark~\emph{et al}\cite{clark2006developing}).
 
@@
 \end{align}
-where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of association sites
+where the factor of $4$ comes from the four association sites per
+molecule, the functional $X$ is the fraction of association sites
 \emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
@@
 where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
-contact for a hard-sphere fluid with a square-well attractive
+contact for a hard-sphere fluid with a square-well dispersion
 potential, and $a_1$ and $a_2$ are the two terms in the dispersion
@@
 length scale for the association interaction.  In addition to the five
-empirical parameters of Clark \emph{et al}, we add a single additional
+empirical parameters of Clark \emph{et al}, we add a single additional dimensionless
 parameter $\lscale$---with a fitted value of 0.353---which determines
@@
 The critical separation will be different for a system where there is
-a significant direct force between rods, or attraction between the
+attraction between the
 rods and water.  At small separations, the water around the two
@@
 scales with volume, it can be neglected relative to the surface term
-provided the length scale is small compared with $\gamma / p$. The
-value of $\gamma / p$ is around $20~\mu m$, which is much larger than
-any of the solutes we
+provided the length scale is small compared with $\gamma / p$, which
+is around $20~\mu m$, and is much larger than any of the systems we
 study. For micron-scale rods, the water on the sides of the `stadium'
@@
 with respect to the separation. 
-We approximate the circumference of the stadium-shape by
+The circumference of the stadium-shape is
 \begin{equation}
@@
 we find for hydrophobic rods of a similar size, as expected due to the
-carbon-water interactions.
+carbon-water interactions.  \fixme{Can you (Jess or Eric) double-check
+  what the excluded-volume diameter is for them? The discrepancy we
+  describe in the text is actually the opposite of what I would
+  expect, but looking at the plot, I see something that is the
+  opposite of what is described in the text.  We should really figure
+  this out, and probably spend another phrase or two clarifying why it
+  is the direction we expect.}
 
@@
 energy plot is qualitatively similar to that predicted by the LCW
-theory~\cite{lum1999hydrophobicity}, with the minor difference that
-we find no barrier to the association of four rods.
+theory~\cite{lum1999hydrophobicity}, with the difference that we find
+no significant barrier to the association of four rods.
 
@@
 effective surface tension surpasses the bulk surface tension at a
-radius of almost 1~nm, and will eventually drop to the bulk value.  As
+radius of almost 1~nm, and at large radius will drop to the bulk value.  As
 with the single rod, we see the analytically correct behavior in the

### 2012-08-10 papers/contact/paper.tex
add MC data to outer-sphere plots Tested everything on node02.
new contact data and cleaned-up plots Tested everything on kipu.
select specific figures to include in paper Tested everything on kipu.

@@
 to only $\sigma$ in the paper, which would have a certain appeal.}
+
+\fixme{Consider eliminating $n_0$ and/or $n_A$ from the plots to make
+  things cleaner.}
 
@@
 
+\subsection{Hard spheres at a hard wall}
+
+We begin by examining the simplest situation with an inhomogeneous
+density:  the case of hard spheres near a single hard wall.  This is
+shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
+
 \begin{figure}
@@
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-30}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.3.}
-  \label{fig:walls-30}
+  \includegraphics[width=\columnwidth]{figs/inner-4-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-4-10}
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-40}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.4.}
-  \label{fig:walls-40}
+  \includegraphics[width=\columnwidth]{figs/inner-12-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-12-10}
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-walls-10}
-  \caption{Correlation multiplied twice by density with bulk filling
+  \includegraphics[width=\columnwidth]{figs/outer-12-01}
+  \caption{Density and contact densities at a xxx with bulk filling
     fraction of 0.1.}
-  \label{fig:gnn-walls-10}
+  \label{fig:outer-10}
 \end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-walls-30}
-  \caption{Correlation multiplied twice by density with bulk filling
-    fraction of 0.3.}
-  \label{fig:gnn-walls-30}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gnn-walls-40}
-  \caption{Correlation multiplied twice by density with bulk filling
-    fraction of 0.4.}
-  \label{fig:gnn-walls-10}
-\end{figure}
-
-\subsection{Hard spheres at a hard wall}
-
-We begin by examining the simplest situation with an inhomogeneous
-density:  the case of hard spheres near a single hard wall.  This is
-shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
 
@@
 behavior of the hard sphere fluid near a concave surface.
-\begin{center}
-  \includegraphics[width=\columnwidth]{figs/outer-16-05}
-\end{center}
-\begin{center}
-  \includegraphics[width=\columnwidth]{figs/outer-16-04}
-\end{center}
-\begin{center}
-  \includegraphics[width=\columnwidth]{figs/outer-16-03}
-\end{center}
 
@@
 
-\begin{center}
-  \includegraphics[width=\columnwidth]{figs/inner-4-30}
-\end{center}
-\begin{center}
+\section{Results}
+
+\subsection{Test-particle result}
+
+Here we discuss comparison of the correlation function calculated via
+both Monte Carlo and the test-particle approach (see
+Figures~\ref{fig:inner-4-10} and~\ref{fig:inner-4-40}).
+
+\subsection{Low density}
+
+Here we discuss the low-density results (filling fraction of 0.1),
+which are shown in Figures~\ref{fig:walls-10}-\ref{fig:outer-10}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/walls-40}
+  \caption{Density and contact densities at a wall with bulk filling
+    fraction of 0.4.}
+  \label{fig:walls-40}
+\end{figure}
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{figs/inner-4-40}
-\end{center}
-\begin{center}
-  \includegraphics[width=\columnwidth]{figs/inner-8-30}
-\end{center}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.4.}
+  \label{fig:inner-4-40}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/inner-12-40}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.4.}
+  \label{fig:inner-12-40}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/outer-16-04}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.4.}
+  \label{fig:outer-40}
+\end{figure}
+
+\subsection{High density}
+
+Here we discuss the high-density results (filling fraction of 0.4),
+which are shown in Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.
 
@@
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{figs/gHS-vs-n}
-\caption{A comparison of the various approximations to the contact
-  density in the homogeneous limit.  As expected, they are all
-  identical, and we won't include this plot in the paper (but it's
-  here to verify that our code is working).}
-\label{fig:gHS-vs-n}
-\end{figure}
+%% \begin{figure}
+%% \includegraphics[width=\columnwidth]{figs/gHS-vs-n}
+%% \caption{A comparison of the various approximations to the contact
+%%   density in the homogeneous limit.  As expected, they are all
+%%   identical, and we won't include this plot in the paper (but it's
+%%   here to verify that our code is working).}
+%% \label{fig:gHS-vs-n}
+%% \end{figure}
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{figs/free-energy}
-\caption{A comparison of the various approximations to the free energy
-  in the homogeneous limit.  Note that the White Bear functional uses
-  a modified form of the Carnahan equation of state.  As expected,
-  these are all identical, and we won't include this plot in the paper
-  (but it's here to verify that our code is working).}
-\label{fig:free-energy}
-\end{figure}
+%% \begin{figure}
+%% \includegraphics[width=\columnwidth]{figs/free-energy}
+%% \caption{A comparison of the various approximations to the free energy
+%%   in the homogeneous limit.  Note that the White Bear functional uses
+%%   a modified form of the Carnahan equation of state.  As expected,
+%%   these are all identical, and we won't include this plot in the paper
+%%   (but it's here to verify that our code is working).}
+%% \label{fig:free-energy}
+%% \end{figure}
 

### 2012-08-10 papers/water-SAFT/paper.tex
changes to paper and fixed a citation Tested everything on kipu.

@@
 describing hydrogen bonds, and the final term is the chain formation
-energy for fluids that (unlike water) are chain polymers.  While a
+energy for fluids that are chain polymers.  While a
 number of formulations of SAFT have been published, we will focus on
@@
 % and the density at contact to a few percent precision.
-We show in Figure~\ref{fig:energy-vs-diameter} the excess chemical
+In Figure~\ref{fig:energy-vs-diameter} we show the excess chemical
 potential at room temperature, scaled by the solvent-accessible
 surface area of the hard rod, plotted as a function of hard-rod
-radius, which we define as the radius from which water is excluded.
+radius.  We define the hard-rods radius as the radius from which water
+is excluded.
 For rods with radius larger than 0.5~nm or so, this reaches a maximum value of
@@
 attraction between the
-rods and water.  At small separations, the water around the two
-rods looks as if the two rods were one solid ``stadium''-shaped object
+rods and water.  At small separations, the shape of the water around the two
+rods makes them appear as one solid ``stadium''-shaped object
 (a rectangle with semi-circles on both ends).
@@
 
-Walther \emph{et al.}\cite{walther2004hydrodynamic} studied the
+Walther \emph{et al.}\cite{walther2004hydrophobic} studied the
 interactions between two carbon nanotubes, which are geometrically
-similar to our hydrophobic rods.  They use molecular dynamics with
+similar to our hydrophobic rods, by using molecular dynamics with
 Lennard-Jones potentials for nanotubes of diameter 1.25~nm and
-separations ranging from about 0.3~nm to 1.5~nm, and find that a
-drying transition occurs, and find it to be at separations near
-0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than what
-we find for hydrophobic rods of a similar size, as expected due to the
-carbon-water interactions.  \fixme{Can you (Jess or Eric) double-check
-  what the excluded-volume diameter is for them? The discrepancy we
-  describe in the text is actually the opposite of what I would
-  expect, but looking at the plot, I see something that is the
-  opposite of what is described in the text.  We should really figure
-  this out, and probably spend another phrase or two clarifying why it
-  is the direction we expect.}
+separations ranging from about 0.3~nm to 1.5~nm. They compute the
+force to constrain the nanotubes at different sepearation distances
+and find that a drying transition begins at separations near 0.9~-~1.0~nm where two
+layers of water molecules can exist in a metastable state between the
+rods. This reduces to one layer at a spacing of 0.55~-~0.58~nm, and at
+0.35~nm water is complete expelled and is in
+equilibrium\cite{walther2004hydrophobic}. The final drying state is
+smaller than what we find for hydrophobic rods of a similar size, as
+expected due to the carbon-water interaction and the
+van der Waals interaction between the rods that causes more attraction
+than our rods.  Our model does not predict any of the transitional
+states seen here\fixme{why?}.
 

### 2012-08-11 papers/water-SAFT/paper.tex
next edit of walther-referencing paragraph Tested everything on kipu.

@@
 
-Walther \emph{et al.}\cite{walther2004hydrophobic} studied the
+Walther \emph{et al}\cite{walther2004hydrophobic} studied the
 interactions between two carbon nanotubes, which are geometrically
-similar to our hydrophobic rods, by using molecular dynamics with
-Lennard-Jones potentials for nanotubes of diameter 1.25~nm and
-separations ranging from about 0.3~nm to 1.5~nm. They compute the
-force to constrain the nanotubes at different sepearation distances
-and find that a drying transition begins at separations near 0.9~-~1.0~nm where two
-layers of water molecules can exist in a metastable state between the
-rods. This reduces to one layer at a spacing of 0.55~-~0.58~nm, and at
-0.35~nm water is complete expelled and is in
-equilibrium\cite{walther2004hydrophobic}. The final drying state is
-smaller than what we find for hydrophobic rods of a similar size, as
-expected due to the carbon-water interaction and the
-van der Waals interaction between the rods that causes more attraction
-than our rods.  Our model does not predict any of the transitional
-states seen here\fixme{why?}.
+similar to our hydrophobic rods, using molecular dynamics with the SPC
+model for water, and a Lennard-Jones potential for the interaction of
+carbon with water, for nanotubes of diameter 1.25~nm and separations
+ranging from about 0.3~nm to 1.5~nm.  The SPC model underestimates the
+surface tension of water by about 24\%\cite{vega2007surface}, so we
+cannot expect this work to provide quantitative agreement with real
+water.  Walther \emph{et al} observe a drying transition between the
+two nanotubes, which occurs at a smaller radius than our results
+suggest.  However, when Walther \emph{et al} disable the attractive
+interaction between nanotube and water, the drying effect occurs at
+much longer range, in agreement with our results.  \fixme{Please read
+  this carefully, as I did some major changes.  I took a look at the
+  paper, and it seems that in Figure 3c (and the paragraph that ends
+  at the top of that page) they finally do a truly hydrophobic
+  interaction, which is a better comparison with what we did.  The
+  drying transition is inhibited by the attraction between tube and
+  water. }
 

### 2012-08-14 papers/contact/figs/plot-outer.py
fix minus-sign bug in A correlation, and use new (fixed) data (mark II correlation function still buggy) Tested everything on node06.

@@ -39,7 +39,7 @@ dft_len = len(r)
 dft_dr = r[2] - r[1]
 padding = 4 # amount of extra space in cell
 radius = round(r[dft_len-1]) - padding/2
-showrmax = radius + 2
+showrmax = radius + 1
 showrmin = radius - 6
 
 pylab.figure(figsize=(8, 7))
@@ -55,6 +55,7 @@ n_plt.plot(r,n*4*numpy.pi/3,"b--",label='DFT $n$')
 n_plt.plot(r,n0*4*numpy.pi/3,"c--",label="$n_0$")
 n_plt.plot(r,nA*4*numpy.pi/3,"m-.",label="$n_A$")
 n_plt.legend(loc='upper left', ncol=2).get_frame().set_alpha(0.5)
+pylab.ylim(0, n.max()*4*numpy.pi/3*1.1)
 pylab.xlim(showrmin,showrmax)
 pylab.xlabel("$r/R$")
 pylab.ylabel("filling fraction")

### 2012-08-14 src/ContactDensity.h

@@ -35,6 +35,8 @@ Functional Correlation_A(double radius);
 Functional Correlation_S(double radius);
 Functional Correlation_S2(double radius);
 Functional Correlation_A2(double radius);
+Functional Correlation_Sm2(double radius); // from Haskell
+Functional Correlation_Am2(double radius); // from Haskell
 Functional Correlation_A_WBm2(double radius);
 Functional Correlation_S_WBm2(double radius);
 

### 2012-08-14 papers/contact/paper.tex
changed appendix equations in paper so that they sem to make more sense and are easier to match up with code Tested everything on node02.
fix minus-sign bug in A correlation, and use new (fixed) data (mark II correlation function still buggy) Tested everything on node06.
new title and abstract for contact paper (and new reference in water paper) Tests skipped!

@@
 \begin{document}
-\title{A Fundamental Measure Theory Functional for Hard-Sphere Contact Densities}
+\title{Using Fundamental Measure Theory to Treat the Correlation
+  Function of the Inhomogeneous Hard-Sphere Fluid}
 
 \author{Jeff Schulte}
+\author{Patrick Kreitzburg}
 \author{Chris Haglund}
-\author{Patrick Kreitzburg}
 \author{David Roundy}
@@
 \begin{abstract}
-  We investigate the contact density of an inhomogeneous hard-sphere
-  fluid, which is of particular interest, as it plays a critical role
-  in Statistical Associating Fluid Theory (SAFT), which is the basis
-  of a number of recent classical density functionals.  We derive a
-  formula for the contact density from the White Bear version of the
-  Fundamental Measure Theory functional~\cite{roth2002whitebear}, and
-  test this functional agains Monte Carlo simulations.
-  \textcolor{red}{Insert summary of results here...}
+  We investigate the value of the correlation function an
+  inhomogeneous hard-sphere fluid at contact $g_\sigma$.  This
+  quantity plays a critical role in Statistical Associating Fluid
+  Theory (SAFT), which is the basis of a number of recent classical
+  density functionals.  We define two averaged values for the
+  correlation function at contact, and derive formulas for each of
+  them White Bear version of the Fundamental Measure Theory
+  functional~\cite{roth2002whitebear} based on an assumption of
+  thermodynamic consistency. We have tested these functionals and two
+  existing functionals~\cite{yu2002fmt-dft-inhomogeneous-associating,
+    gross2009density} against Monte Carlo simulations.  We find that
+  excellent agreement with the Monte Carlo results for our
+  asymmetrically averaged correlation function, and less impressive
+  results for the two existing functionals and our symmetrically
+  averaged correlation function.
 \end{abstract}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-12-01}
+  \includegraphics[width=\columnwidth]{figs/outer-16-01}
   \caption{Density and contact densities at a xxx with bulk filling
@@
 \begin{align}
-  n_3^{A}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr')) d\rr'
+  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -R(\rr'')) d\rr''
 \end{align}
@@
 \begin{align}
-  \mathbf{n}_{V2}^{A}(\rr) &= \int n(\rr') \delta(|\rr' - \rr''| - R(\rr''))
+  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(|\rr' - \rr''| - R(\rr''))
     \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta R(\rr)} &= - n(\rr) \delta'(|\rr' - \rr| - R(\rr))
-    \frac{\rr'-\rr}{|\rr'-\rr|}\\
+  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta R(\rr)} &= n(\rr) \delta'(|\rr' - \rr| - R(\rr))
+    \frac{\rr-\rr'}{|\rr-\rr'|}\\
 \end{align}
@@
 \begin{align}
-  \mathbf{n}_{V1}^A(\rr) &= \int d\rr' \frac{n(\rr')}{4\pi R(\rr')}
-  \delta(|\mathbf{r}-\rr'| - R(\rr')) \frac{\rr-\rr'}{|\rr-\rr'|}\\
+  \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{4\pi R(\rr'')}
+  \delta(|\mathbf{r'}-\rr''| - R(\rr'')) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
   \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta R(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    R(\rr)}\delta'(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
-  -
+  &= \frac{n(\rr)}{4\pi
+    R(\rr)}\delta'(|\rr'-\rr| - R(\rr)) \frac{\rr-\rr'}{|\rr-\rr'|}
+  +
   \frac{n(\rr)}{4\pi
-    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
+    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr)) \frac{\rr-\rr'}{|\rr-\rr'|}
 \end{align}
@@
 \begin{align}
-  n_3^{S}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr)) d\rr'
+  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -R(\rr')) d\rr''
 \end{align}
@@
 \begin{align}
-  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))d\mathbf
-  r'\\
+  n_2^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - R(\rr'))d\mathbf
+  r''\\
   \frac{\delta n_2^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
@@
 \begin{align}
-  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
+  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{4\pi R(\rr')} \delta(|\rr' - \rr''| - R(\rr'))d\mathbf r''\\
     \frac{\delta n_1^{S}(\rr')}{\delta R(\rr)} &=
@@
 \begin{align}
-  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
+  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{4\pi R(\rr')^2} \delta(|\rr' - \rr''| - R(\rr'))d\mathbf r''\\
     \frac{\delta n_0^{S}(\rr')}{\delta R(\rr)}
@@
 \begin{align}
-  \mathbf{n}_{V2}^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))
-  \frac{(\rr - \rr')}{|\rr - \rr'|} d \rr'\\
+  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - R(\rr'))
+  \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
   \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr - \rr')
@@
 \begin{align}
- \mathbf{n}_{V1}^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'
-   \frac{(\rr - \rr')}{|\rr - \rr'|}\\
+ \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{4\pi R(\rr')} \delta(|\rr' - \rr''| - R(\rr'))d\mathbf r''
+   \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
  \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta R(\rr)}

### 2012-08-14 papers/contact/paper.tex
changed all the Rs in the paper to sigmas, but not in the text yet Tested everything on node06.

--- 2012-08-10 papers/contact/paper.tex
@@
 \end{align}
-where $\eta \equiv \frac{4\pi}{3} R^3 n$ is the filling fraction.  We
+where $\eta \equiv \frac{4\pi}{24} \sigma^3 n$ is the filling fraction.  We
 can thus readily compute the derivative of free energy with respect to
@@
 \begin{align}
-  \frac{dA_{HS}}{dR} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{dR} \\
+  \frac{dA_{HS}}{dR} &= 2\frac{dA_{HS}}{d\eta} \frac{d\eta}{d\sigma} \\
   \derivation{
@@
   }
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \label{eq:dAhsdR}
+  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{6 \eta}{\sigma} \label{eq:dAhsdR}
 \end{align}
@@
 \begin{align}
-  p_{HS} &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
+  p_{HS} &= \frac{1}{N 2\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
   \derivation{
@@
 \begin{align}
-  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - R) d\rr' \\
-  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - R) d\rr'
+  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
+  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) d\rr'
 \end{align}
 \begin{align}
-  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 , \quad
-  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{4\pi R} \\
-  n_1 &= \frac{n_2}{4\pi R} , \quad
-  n_0 = \frac{n_2}{4\pi R^2}
+  \mathbf{n}_{V2} &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr - \rr'}{\left|\rr - \rr'\right|} d\rr' \quad \\
+  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{2\pi \sigma} , \quad
+  n_1 &= \frac{n_2}{2\pi \sigma} , \quad
+  n_0 = \frac{n_2}{\pi \sigma^2}
 \end{align}
@@
 \begin{align}
-  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr'))d\rr' \label{eq:n3A}
+  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr')/2)d\rr' \label{eq:n3A}
 \end{align}
@@
 \begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^A}\\
-  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \\
+  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
+  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \\
   g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
-  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
+  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
 \end{align}
-The functional derivative $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$
+The functional derivative $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
 in the above formula must be computed with the asymmetric formulation
@@
 between the treatment of the A and S cases in the derivatives 
-$\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
+$2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
 and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
@@
 of these we use two different conceptual approaches.  In the assymetric case, 
-a change in $R(\mathbf{r})$ will refer to a change in the radius of a sphere 
+a change in $\sigma(\mathbf{r})/2$ will refer to a change in the radius of a sphere 
 that is located at the position $\mathbf{r}$.  The weighted densities that are 
-affected by such a change will be those that are located a distance $R$ away 
+affected by such a change will be those that are located a distance $\sigma/2$ away 
 from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.
@@
 \begin{align}
-  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-R(\rr))d\rr'
+  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr)/2)d\rr'
 \end{align}
@@
 
-These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$, 
+These qualitative concepts lead us to the derivative terms in $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
 which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the
@@
 \begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:p_{HS}^S}\\
-  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \\
+  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^S}\\
+  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \\
   g_\sigma^S(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 4\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta R(\mathbf{r})} \label{eq:g-A-exact}
+  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
+    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
 \end{align}
-where the functional derivative $\frac{\delta A_{HS}}{\delta
-  R(\mathbf{r})}$ in this case is computed with the symmetric
+where the functional derivative $2 \frac{\delta A_{HS}}{\delta
+  \sigma(\mathbf{r})}$ in this case is computed with the symmetric
 formulation of the fundamental measures.  Complete expressions for
@@
 \begin{align}
-  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{2\pi}{3}R^3n_A(\rr)}{\left(1 -
-    \frac{4\pi}{3}R^3n_A(\rr)\right)^3}
+  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{\pi}{12}\sigma^3n_A(\rr)}{\left(1 -
+    \frac{\pi}{6}\sigma^3n_A(\rr)\right)^3}
 \end{align}
@@
   g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac32 \frac{R n_2}{3}\frac{\zeta}{(1-n_3)^2}
-    + \frac12 \frac{\left(\frac{R n_2}{3}\right)^2 \zeta}{(1-n_3)^3}
+    + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
+    + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
     \\
   g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
-    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}
+    + \frac14 \frac{\sigma n_2\zeta}{(1-n_3)^2}
+    + \frac1{72} \frac{\sigma^2 n_2^2 \zeta}{(1-n_3)^3}
 \end{align}
@@
   g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
-    + \frac12 \frac{R n_2\zeta}{(1-n_3)^2}
-    + \frac1{18} \frac{R^2 n_2^2 \zeta}{(1-n_3)^3}\right)
+    + \frac14 \frac{\sigma n_2\zeta}{(1-n_3)^2}
+    + \frac1{72} \frac{\sigma^2 n_2^2 \zeta}{(1-n_3)^3}\right)
 \end{align}
@@
 
-In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta R(\mathbf{r})}$ terms that must be derived for both the S and A cases.
+In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.
 We break these terms up as follows:
@@
   \begin{align}
-    \frac{\delta A_{HS}}{\delta R(\mathbf{r})} &=
+    \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})} &=
     \int \left(
     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-    \frac{\delta n_3(\mathbf{r}')}{\delta R(\mathbf{r})}
+    \frac{\delta n_3(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
     +
     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-    \frac{\delta n_2(\mathbf{r}')}{\delta R(\mathbf{r})}
+    \frac{\delta n_2(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
     + \cdots
@@
 
-We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc., terms.
+We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.
 We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.
-Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc., terms for both the A and S cases.
+Following this we will compute the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.
 These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
@@
 
-\subsection{Expressions for $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the assymetric case}\label{appendix:g-A}
+\subsection{Expressions for $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc. for the assymetric case}\label{appendix:g-A}
 
@@
 \begin{align}
-  n_3^{A}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr')) d\rr'
+  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr'')/2) d\rr''
 \end{align}
@@
 \begin{align}
-  \frac{\delta n_3^{A} (\rr')}{\delta R(\rr)} &=
-  \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'')) \delta(\rr-\rr'') d\rr'' \\
-   &= n (\rr) \delta(|\rr' - \rr| - R(\rr))
+  2 \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
+  \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
+   &= n (\rr) \delta(|\rr' - \rr| - \sigma(\rr)/2)
 \end{align}
@@
 \begin{align}
-  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - R(\rr'')) d \rr''\\
-  \frac{\delta n_2^{A}(\rr')}{\delta R(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - R(\rr))\\
+  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) d \rr''\\
+  2 \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
 \end{align}
@@
 \begin{align}
-  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')}
-  \delta(|\rr'-\rr''| - R(\rr'')) \\
-  \frac{\delta n_1(\rr')}{\delta R(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    R(\rr)}\delta'(|\rr'-\rr| - R(\rr))
+  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
+  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
+  2 \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
+  &= -\frac{n(\rr)}{2\pi
+    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
   -
-  \frac{n(\rr)}{4\pi
-    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr))
+  \frac{n(\rr)}{\pi
+    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2)
 \end{align}
@@
 \begin{align}
-  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{4\pi R(\rr'')^2}
-  \delta(|\rr'-\rr''| - R(\rr'')) \\
-  \frac{\delta n_0(\rr')}{\delta R(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    R(\rr)^2}\delta'(|\rr'-\rr| - R(\rr))
+  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
+  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
+  2 \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
+  &= -\frac{n(\rr)}{\pi
+    \sigma(\rr)^2}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
   -
-  2\frac{n(\rr)}{4\pi
-    R(\rr)^3}\delta(|\rr'-\rr| - R(\rr))
+  4 \frac{n(\rr)}{\pi
+    \sigma(\rr)^3}\delta(|\rr'-\rr| - \sigma(\rr)/2)
 \end{align}
@@
 \begin{align}
-  \mathbf{n}_{V2}^{A}(\rr) &= \int n(\rr') \delta(|\rr' - \rr''| - R(\rr''))
+  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(|\rr' - \rr''| - \sigma(\rr'')/2)
     \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta R(\rr)} &= - n(\rr) \delta'(|\rr' - \rr| - R(\rr))
-    \frac{\rr'-\rr}{|\rr'-\rr|}\\
+  2 \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
+    \frac{\rr-\rr'}{|\rr-\rr'|}\\
 \end{align}
@@
 \begin{align}
-  \mathbf{n}_{V1}^A(\rr) &= \int d\rr' \frac{n(\rr')}{4\pi R(\rr')}
-  \delta(|\mathbf{r}-\rr'| - R(\rr')) \frac{\rr-\rr'}{|\rr-\rr'|}\\
-  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta R(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    R(\rr)}\delta'(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
-  -
-  \frac{n(\rr)}{4\pi
-    R(\rr)^2}\delta(|\rr'-\rr| - R(\rr)) \frac{\rr'-\rr}{|\rr'-\rr|}
+  \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
+  \delta(|\mathbf{r'}-\rr''| - \sigma(\rr'')/2) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
+  2 \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
+  &= \frac{n(\rr)}{2\pi
+    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
+  +
+  \frac{n(\rr)}{\pi
+    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
 \end{align}
 \fixme{double-check the directions of all the vector quantities...}
-
-%% So, what is this convolution with a $\delta'$? Let's look at the
-%% scalar version:
-%% \fixme{With following the primes are certainly not in the right places}
-%% \begin{align}
-%%   \bar{f}(\mathbf{r})
-%%   &= \int f(\mathbf{r'}) \delta'(|\mathbf{r}-\mathbf{r'}| - R)
-%%   \\
-%%   &= \int
-%%         f(\mathbf{r'})
-%%         \lim_{\Delta R \rightarrow 0}
-%%           \frac{\delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-%%             -
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%%       }{\Delta R}
-%%   \\
-%%   &= \lim_{\Delta R \rightarrow 0}
-%%           \frac{\int
-%%             f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R - \Delta R)
-%%             -
-%%             \int
-%%             f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%%       }{\Delta R}
-%%   \\
-%%   &\approx\lim_{\Delta R \rightarrow 0}
-%%           \frac{\int
-%%             \frac{(R+\Delta R)^2}{R^2}f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%%             -
-%%             \int
-%%             f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%%       }{\Delta R}
-%%   \\
-%%   &\approx\lim_{\Delta R \rightarrow 0}
-%%           \frac{2\frac{\Delta R}{R} \int
-%%             f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%%       }{\Delta R}
-%%   \\
-%%   &= \frac{2}{R} \int
-%%             f(\mathbf{r'})
-%%             \delta(|\mathbf{r}-\mathbf{r'}| - R)
-%% \end{align}
-%% There is a sign ambiguity here with regard to whether we are asking
-%% about the derivative of this delta function with regard to $R$ or
-%% something else.  We just need to get it right.
 
@@
 \subsection{Expressions for $\beta\frac{\delta
-    n_3^{S}(\mathbf{r}')}{\delta R(\mathbf{r})}$, etc. for the
+    n_3^{S}(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc. for the
   symetric case}\label{appendix:g-S}
@@
 \begin{align}
-  n_3^{S}(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| -R(\rr)) d\rr'
+  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr')/2) d\rr''
 \end{align}
@@
 \begin{align}
-  \frac{\delta n_3^{S} (\rr')}{\delta R(\rr)} &=
-   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - R(\rr'))
+  2 \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
+   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
    \delta(\rr'-\rr) d\rr''
@@
 \begin{align}
-  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
-  \frac{\delta n_3^{S}(\rr')}{\delta R(\rr)} d\rr'
-  &= \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
+  \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
+  \frac{\delta n_3^{S}(\rr')}{\delta \sigma(\rr)} d\rr'
+  &= \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
   n_2^S(\rr') d\rr' \\
-  &= \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
+  &= 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
 \end{align}
@@
 \begin{align}
-  n_2^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))d\mathbf
-  r'\\
-  \frac{\delta n_2^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
-  \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+  n_2^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf
+  r''\\
+  2 \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
+  \delta'(|\rr'-\rr''| - \sigma(\rr')) d\rr''
 \end{align}
@@
 \begin{align}
-  n_1^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
-    \frac{\delta n_1^{S}(\rr')}{\delta R(\rr)} &=
-    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)}
-    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
+    2\frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
+    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{2\pi \sigma(\rr)}
+    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
     + n_0^s(\rr') \right)
@@
 \begin{align}
-  n_0^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)^2} \delta(|\rr - \rr'| - R(\rr))d\mathbf r'\\
-    \frac{\delta n_0^{S}(\rr')}{\delta R(\rr)}
-    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi R(\rr)^2}
-    \delta'(|\rr'-\rr''| - R(\rr')) d\rr''
+  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
+    2 \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
+    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
+    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
     +
-    \frac{2}{R} n_0^S(\rr') \right)
+    \frac{4}{\sigma} n_0^S(\rr') \right)
 \end{align}
 \begin{align}
-  \mathbf{n}_{V2}^{S}(\rr) &= \int n(\rr') \delta(|\rr - \rr'| - R(\rr))
-  \frac{(\rr - \rr')}{|\rr - \rr'|} d \rr'\\
-  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta R(\rr)} &= -\delta(\rr - \rr')
-  \int n(\rr'') \delta'(|\rr' - \rr''| - R(\rr'))
+  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
+  \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
+  2 \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
+  \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
   \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
@@
 \begin{align}
- \mathbf{n}_{V1}^{S}(\rr) &= \int \frac{n(\rr')}{4\pi R(\rr)} \delta(|\rr - \rr'| + R(\rr))d\mathbf r'
-   \frac{(\rr - \rr')}{|\rr - \rr'|}\\
- \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta R(\rr)}
- &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi R(\rr)}
-   \delta'(|\rr'-\rr''| - R(\rr')) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
-   + \frac{\mathbf{n}_{V2}^S(\rr')}{4\pi R(\rr)^2} \right)
+ \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''
+   \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
+ 2 \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
+ &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{2\pi \sigma(\rr)}
+   \delta'(|\rr'-\rr''| - \sigma(\rr')/2) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
+   + \frac{\mathbf{n}_{V2}^S(\rr')}{\pi \sigma(\rr)^2} \right)
 \end{align}

### 2012-08-14 papers/water-SAFT/paper.tex

@@
 water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft, 
-  gloor2004accurate, gloor2007prediction, Jaqaman2004, clark2006developing,
+  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006, clark2006developing,
   lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
@@
 %% paper & SAFT functional & water & LDA & tension \\
+%% \hline
+%% chuev2006 &  & X &  & X \\
 %% \hline

### 2012-08-14 papers/contact/paper.tex
content edit of beginning of contact paper Tested everything on node06.

@@
   correlation function at contact, and derive formulas for each of
-  them White Bear version of the Fundamental Measure Theory
-  functional~\cite{roth2002whitebear} based on an assumption of
+  them from the White Bear version of the Fundamental Measure Theory
+  functional~\cite{roth2002whitebear}, based on an assumption of
   thermodynamic consistency. We have tested these functionals and two
   existing functionals~\cite{yu2002fmt-dft-inhomogeneous-associating,
-    gross2009density} against Monte Carlo simulations.  We find that
-  excellent agreement with the Monte Carlo results for our
+    gross2009density} against Monte Carlo simulations.  We find
+  excellent agreement between the Monte Carlo data and our
   asymmetrically averaged correlation function, and less impressive
-  results for the two existing functionals and our symmetrically
-  averaged correlation function.
+  agreement with the other three correlation functionals.
 \end{abstract}
@@
 successful in qualitatively describing the dependence of surface
-tension on temperature.  Unfortunately, most of these constructed
-functionals\cite{\saftlocaldft} cannot be used to study arbitrary
-inhomogeneous density distributions, because they use local density
-approximations that are only valid for densities that vary slowly over
-molecular length scales.  A key input to those SAFT-based functionals
-that can handle sharply varying densities is the correlation function
-evaluated at contact.
+tension on temperature.
+%
+%% Unfortunately, most of these constructed
+%% functionals\cite{\saftlocaldft} cannot be used to study arbitrary
+%% inhomogeneous density distributions, because they use local density
+%% approximations that are only valid for densities that vary slowly over
+%% molecular length scales.
+%
+A key input to SAFT-based functionals is the correlation function
+evaluated at contact.  Although two recent works have introduced
+approximate functionals for this
+quantity\cite{yu2002fmt-dft-inhomogeneous-associating,
+  gross2009density}, there has yet to be a careful study of the
+correlation function at contact in the inhomogeneous hard-sphere
+fluid.
 
@@
 where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
-is the excess free energy of a hard-sphere fluid of monomers,
-$A_\textit{chain}$ is the contribution to the free energy from the
-chaining effect in a polymeric fluid, $A_\textit{disp}$ is the
-dispersion contribution to the free energy, and $A_\textit{assoc}$ is
-the free energy contribution from \emph{association}, which is to say,
-hydrogen bonding.  The chain and association contributions to the SAFT
-free energy are of particular interest for this paper, since each uses
-as input the contact value of the correlation function.
+is the excess free energy of a hard-sphere fluid, $A_\textit{chain}$
+is the free energy of formation for polymeric fluids,
+$A_\textit{disp}$ is the dispersion contribution to the free energy,
+and $A_\textit{assoc}$ is the free energy contribution from
+\emph{association}, which is to say, hydrogen bonding.  The chain and
+association contributions to the SAFT free energy are of particular
+interest for this paper, since each requires as input the contact
+value of the correlation function.
 %
@@
 Section~\ref{sec:yuwu})\cite{yu2002fmt-dft-inhomogeneous-associating},
-which has subsequently been used in other
-functionals\cite{fu2005vapor-liquid-dft, bryk2006density}.
+which has subsequently been used in the development of other
+SAFT-based functionals\cite{fu2005vapor-liquid-dft, bryk2006density}.
 %% Each of these terms has some dependence on the
@@
 %% weak position dependence, is amenable to mean-field approximations.
-Given these different approaches, it seems worthwhile to examine this
+Given these different approaches, it seems valuable to examine this
 property of the hard-sphere fluid through direct simulation, in order
@@
 
-We first discuss two possibilities for the definition of contact
-density.  The first involves a more symmetric interpretation and so we
-call it the '$S$ case'.  The second involves a sphere-centered,
-assymetrical interpretation and so we call it the '$A$ case'.  We
-present where in the SAFT free energy formulation these contact
-densities will contribute.  We next present a thermodynamic method
-that we will use to calculate the above contact densities.  This
-method involves the use of surface pressure and the minimization of
-the excess free energy refered to above.  We use the White Bear and
-White Bear Mark II formulations of these free energy terms and take
-neccesary derivitives in the concpetual spirit of the seperately
-defined $S$ an $A$ case contact densities.  Finally we compare the
-resulting contact densities with those of Gross, those of Yu and Wu,
-and as well with our own Monte Carlo simulations.
+In this paper we introduce two definitions for the locally averaged
+correlation function of an inhomogeneous system.  The first is more
+symmetric (which we call the $S$ case), while the second involves a
+sphere-centered, assymetrical interpretation (the $A$ case).  Given
+these definitions, we will present a thermodynamic approach for
+evaluating these these correlation functions directly from the free
+energy functional itself.  We will then explain how the correlation
+function of Yu and Wu and that of Gross compare with our
+definitions---the former is a variant of the $S$ correlation function,
+while the latter is an $A$ function.  Finally, we will compare these
+four approximations with exact Monte-Carlo simulations for the
+hard-sphere fluid in at a variety of hard-wall surfaces.
 
@@
 We will define our terms by making use of the two-particle density
-$n^{(2)}(\rr_1,\rr_2)$, which gives the probability per unit volume of
-finding one particle at position $\rr_1$ and the other at position
-$\rr_2$.  The pair correlation function is defined by
+$n^{(2)}(\rr_1,\rr_2)$, which gives the probability per unit volume
+squared of finding one particle at position $\rr_1$ and the other at
+position $\rr_2$.  The pair correlation function is defined by
 \begin{align}
@@
 
-For the symmetric $S$ case:
+For the symmetric $S$ case, the correlation function at contact is
+given by:
 \begin{align}
@@
   + \rr')
-  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \label{eq:gS} \\
+  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \label{eq:gS}
+\end{align}
+where density $n_0$ is one of the fundamental measures of Fundamental
+Measure Theory (FMT), and is ideal for treating touching spheres, as
+illustrated in Figure~\ref{fig:n0}.
+\begin{align}
   n_0(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
 \end{align}
-This functional gives a value averaged over spheres that touch at the
-position $\rr$.  The density $n_0$ is one of the fundamental measures
-of Fundamental Measure Theory (FMT), and is ideal for treating
-touching spheres, as illustrated in Figure~\ref{fig:n0}.
+This functional gives a value averaged over all spheres that touch at
+the position $\rr$.
 
@@
 
-And for the asymmetric $A$ case:
+In contrast, the asymmetrically averaged $A$ correlation function is
+given by
 \begin{align}
@@
   \int n^{(2)}(\rr, \rr + \rr')
-  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \label{eq:gA} \\
+  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \label{eq:gA}
+\end{align}
+where the density $n_A(\rr)$ is similar to $n_0$, but measures the
+density of spheres that could be touching a sphere that is located at
+point $\rr$.
+\begin{align}
   n_A(\rr) &= \int n(\rr')
@@
 \end{align}
-In this case, the two-particle density is averaged over spheres
-touching a sphere that is at the given position~$\rr$.
+Thus $g_\sigma^A$ corresponds to an average of the two-particle
+density over spheres touching a sphere that is located at the
+position~$\rr$.
 
-In the process of defining these two averaged correlation functions,
-we also defined two averaged densities, $n_0(\rr)$ and $n_A(\rr)$,
-which are the density of spheres available to be in
-contact.  In the homogeneous limit, $n_0 = n_A = n$.  It is an
-open question, which of these averages will be more useful in any
-particular functional.  Suffice to say, either average is a
-\emph{possible} way to convert a function that is defined for a
-homogeneous system to a functional that is applicable to inhomogeneous
-systems.
+%% In the process of defining these two averaged correlation functions,
+%% we also defined two averaged densities, $n_0(\rr)$ and $n_A(\rr)$,
+%% which are the density of spheres available to be in
+%% contact.  In the homogeneous limit, $n_0 = n_A = n$.  It is an
+%% open question, which of these averages will be more useful in any
+%% particular functional.  Suffice to say, either average is a
+%% \emph{possible} way to convert a function that is defined for a
+%% homogeneous system to a functional that is applicable to inhomogeneous
+%% systems.
 
-Yu and Wu introduce an approximation for
-$g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which we will
-discus in Section~\ref{sec:yuwu}.  Gross introduced an
-approximation for $g^A$\cite{gross2009density}, which we will
-introduce in Section~\ref{sec:gross}.  We will also introduce our own
-approximations for both $g^A$ and $g^S$ which are based on the
-assumption of thermodynamic consistency within FMT, in
-Sections~\ref{sec:g-S} and~\ref{sec:g-A}.  We will report on the accuracy of
-these approximations by comparing with Monte Carlo simulations of the
-inhomogeneous hard-sphere fluid.  We will focus on the contact
-densities, rather than the correlation function itself, because these
-are simpler, and because they are the quantity that you want to
-``average''.  i.e. if you were to average $g$ itself, you'd want to
-weight that average based on the density, which would mean averaging
-$n^{(2)}$.
+%% Yu and Wu introduce an approximation for
+%% $g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which we will
+%% discus in Section~\ref{sec:yuwu}.  Gross introduced an
+%% approximation for $g^A$\cite{gross2009density}, which we will
+%% introduce in Section~\ref{sec:gross}.  We will also introduce our own
+%% approximations for both $g^A$ and $g^S$ which are based on the
+%% assumption of thermodynamic consistency within FMT, in
+%% Sections~\ref{sec:g-S} and~\ref{sec:g-A}.  We will report on the accuracy of
+%% these approximations by comparing with Monte Carlo simulations of the
+%% inhomogeneous hard-sphere fluid.  We will focus on the contact
+%% densities, rather than the correlation function itself, because these
+%% are simpler, and because they are the quantity that you want to
+%% ``average''.  i.e. if you were to average $g$ itself, you'd want to
+%% weight that average based on the density, which would mean averaging
+%% $n^{(2)}$.
 
-\subsection{Association free energy of SAFT}
+%% \subsection{Association free energy of SAFT}
 
-\fixme{Integrate the following equations into this section...}
+%% \fixme{Integrate the following equations into this section...}
+%% \begin{align}
+%%   \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(n g_\sigma \right)-1\right)
+%% \end{align}
+%% \begin{align}
+%%   \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
+%%   X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
+%% \end{align}
+
+%% \newcommand\epsilonassoc{\ensuremath{\varepsilon_\textit{AB}}}
+%% \newcommand\kappaassoc{\ensuremath{\kappa_\textit{AB}}}
+\newcommand\ncontact{\ensuremath{n_\textit{contact}}}
+
+%% The free energy term in Statistical Associating Fluid Teory (SAFT)
+%% from which it derives its name is the association term, which accounts
+%% for hydrogen bonding.  Hydrogen bonds are modeled as attractive
+%% patches (``association sites'') on the surface of hard spheres.  These
+%% sites represent protons or electron lone pairs, and have an attractive
+%% energy $\epsilonassoc$ when two molecules are oriented such that the
+%% proton of one overlaps with the lone pair of the other.  The volume
+%% over which this interaction occurs is $\kappaassoc$, giving the
+%% association term in the free energy has two empirical parameters fit
+%% to experimental data.
+
+%% The association free energy per unit volume has the form
+%% \begin{align}
+%%   f_\text{assoc} &= k_BT n\sum_A 
+%%                   \left(\ln X^A - \frac{X^A}{2} + \frac12\right)
+%% \end{align}
+%% where the summation is over the association sites, and $X^A$ is the
+%% fraction of association sites \emph{not} hydrogen-bonded.  The
+%% fractions $X^A$ are determined by the self-consistent equations
+%% \begin{align}
+%%   X^A &= \frac{1}
+%%   {1 + \sum_B n X^B \Delta^{AB}}\label{eq:X}
+%%   \\
+%%   \Delta^{AB} &= g(\sigma) \kappaassoc\left( e^{\epsilonassoc/kT} -
+%%   1 \right)\label{eq:delta}
+%% \end{align}
+%% where $g(\sigma)$ is the correlation function evaluated at contact.
+%% The product of density with correlation function evaluated at contact
+%% gives the contact density, when we combine Equations~\ref{eq:X} and
+%% \ref{eq:delta}:
+%% \begin{align}
+%%   X^A &= \frac{1}
+%%   {1 + \sum_B \ncontact\kappaassoc X^B\left( e^{\epsilonassoc/kT} -
+%%   1 \right)}
+%% \end{align}
+%% where the product $\ncontact\kappaassoc$ is the expected number of
+%% molecules present in the association volume in the absence of the
+%% association interaction.  In the SAFT-VR
+%% model\cite{gil-villegas-1997-SAFT-VR}, this contact density is found
+%% by adding a perturbative correction to the contact value for a hard
+%% sphere fluid.
+
+\subsection{Fundamental-Measure Theory}
+
+We use the White Bear version of the Fundamental-Measure Theory~(FMT)
+functional published in reference~\cite{roth2002whitebear}.  The FMT
+functional describes the excess free energy of a hard-sphere fluid.
+This particular FMT reduces to the Carnahan-Starling equation of state
+for homogeneous systems.
+\begin{equation}
+A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
+\end{equation}
+with integrands
 \begin{align}
-  \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(n g_\sigma \right)-1\right)
+\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
+\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
+\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+  n_3 + (1-n_3)^2 \ln(1-n_3)
+}{
+  36\pi n_3^2\left( 1 - n_3 \right)^2
+} ,
+\end{align}
+using the weighted densities
+\begin{align}
+  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
+  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
+  \mathbf{n}_{2V}(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
 \end{align}
 \begin{align}
-  \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
-  X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
+  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{2\pi \sigma}, \quad
+  n_1 &= \frac{n_2}{2\pi \sigma} , \quad
+  n_0 = \frac{n_2}{\pi \sigma^2}
 \end{align}
-
-\newcommand\epsilonassoc{\ensuremath{\varepsilon_\textit{AB}}}
-\newcommand\kappaassoc{\ensuremath{\kappa_\textit{AB}}}
-\newcommand\ncontact{\ensuremath{n_\textit{contact}}}
-
-The free energy term in Statistical Associating Fluid Teory (SAFT)
-from which it derives its name is the association term, which accounts
-for hydrogen bonding.  Hydrogen bonds are modeled as attractive
-patches (``association sites'') on the surface of hard spheres.  These
-sites represent protons or electron lone pairs, and have an attractive
-energy $\epsilonassoc$ when two molecules are oriented such that the
-proton of one overlaps with the lone pair of the other.  The volume
-over which this interaction occurs is $\kappaassoc$, giving the
-association term in the free energy has two empirical parameters fit
-to experimental data.
-
-The association free energy per unit volume has the form
-\begin{align}
-  f_\text{assoc} &= k_BT n\sum_A 
-                  \left(\ln X^A - \frac{X^A}{2} + \frac12\right)
-\end{align}
-where the summation is over the association sites, and $X^A$ is the
-fraction of association sites \emph{not} hydrogen-bonded.  The
-fractions $X^A$ are determined by the self-consistent equations
-\begin{align}
-  X^A &= \frac{1}
-  {1 + \sum_B n X^B \Delta^{AB}}\label{eq:X}
-  \\
-  \Delta^{AB} &= g(\sigma) \kappaassoc\left( e^{\epsilonassoc/kT} -
-  1 \right)\label{eq:delta}
-\end{align}
-where $g(\sigma)$ is the correlation function evaluated at contact.
-The product of density with correlation function evaluated at contact
-gives the contact density, when we combine Equations~\ref{eq:X} and
-\ref{eq:delta}:
-\begin{align}
-  X^A &= \frac{1}
-  {1 + \sum_B \ncontact\kappaassoc X^B\left( e^{\epsilonassoc/kT} -
-  1 \right)}
-\end{align}
-where the product $\ncontact\kappaassoc$ is the expected number of
-molecules present in the association volume in the absence of the
-association interaction.  In the SAFT-VR
-model\cite{gil-villegas-1997-SAFT-VR}, this contact density is found
-by adding a perturbative correction to the contact value for a hard
-sphere fluid.
 
@@
 
-In order to motivate the derivation of our functional for contact
-density, we begin by deriving the well-known contact density for a
-homogeneous hard-sphere fluid from the Carnahan-Starling free energy.
-The contact density is most easily computed from the contact density
-theorem, which states that the pressure on any hard surface is
-determined by the density at contact:
+In order to motivate the derivation of our functional for the value of
+the correlation function at contact, we begin by deriving the
+well-known value of $g_\sigma$ for a homogeneous hard-sphere fluid
+from the Carnahan-Starling free energy.  The contact value of the
+correlation function density is most easily found by using the
+contact-value theorem, which states that the pressure on any hard
+surface is determined by the density at contact:
 \begin{align}
-  p &= k_BT n_\textit{contact}.
+  p &= k_BT n_\textit{contact} \\
+  &= k_BT n g_\sigma
 \end{align}
-Since we are interested in the contact density at the surface of the
-hard spheres, all we need compute is the pressure on that surface, and
-we'll have our answer.  The pressure on a hard sphere can be readily
-computed from the dependence of the Carnahan-Starling free energy on
-hard sphere radius.
+Since we are interested in the correlation function at the surface of
+the hard spheres, all we need compute is the pressure on that surface,
+and we'll have our answer.  The pressure on a hard sphere can be
+readily computed from the dependence of the Carnahan-Starling free
+energy on hard sphere radius.
 \begin{align}
@@
 by means of the fundamental measures, this requires that we adjust the
-convolutions to have a radius $R$ which depends on the location of the
-sphere:
+convolutions to have a diameter $\sigma$ which depends on the location
+of the sphere:
 \begin{align}
@@
 \end{align}
-The hard-sphere radius $R$ in this weighted density has an unusual
+The hard-sphere radius $\sigma/2$ in this weighted density has an unusual
 position dependence:  it depends on the position at which we are

### 2012-08-15 papers/contact/paper.tex
made some minor changes to appendices of paper - brought down twos from derives Tested everything on node06.

--- 2012-08-14 papers/contact/paper.tex
@@
 \end{align}
-The functional derivative $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
+The functional derivative $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
 in the above formula must be computed with the asymmetric formulation
@@
 between the treatment of the A and S cases in the derivatives 
-$2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
+$\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
 and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
@@
 
-These qualitative concepts lead us to the derivative terms in $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
+These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
 which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the
@@
 \end{align}
-where the functional derivative $2 \frac{\delta A_{HS}}{\delta
+where the functional derivative $\frac{\delta A_{HS}}{\delta
   \sigma(\mathbf{r})}$ in this case is computed with the symmetric
@@
 
-In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $2 \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.
+In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.
 We break these terms up as follows:
@@
 
-We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.
+We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.
 We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.
-Following this we will compute the $2 \frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.
+Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.
 These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
@@
 compute derivatives of the free energy with respect to $R(\rr)$, we
-need to find $g_\sigma^A$.  In the case of $n_3^A$, we needS
+need to find $g_\sigma^A$.  In the case of $n_3^A$, we need
 \begin{align}
-  2 \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
-  \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
+  \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
+  \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
    &= n (\rr) \delta(|\rr' - \rr| - \sigma(\rr)/2)
@@
   n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) d \rr''\\
-  2 \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
+  \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
 \end{align}
@@
   \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
-  2 \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{2\pi
+  \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
+  &= -\frac{n(\rr)}{4\pi
     \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
   -
-  \frac{n(\rr)}{\pi
+  \frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2)
@@
   \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
-  2 \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{\pi
+  \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
+  &= -\frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
   -
-  4 \frac{n(\rr)}{\pi
+  2 \frac{n(\rr)}{\pi
     \sigma(\rr)^3}\delta(|\rr'-\rr| - \sigma(\rr)/2)
@@
     \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  2 \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
+  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
     \frac{\rr-\rr'}{|\rr-\rr'|}\\
@@
   \delta(|\mathbf{r'}-\rr''| - \sigma(\rr'')/2) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
-  2 \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{2\pi
+  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
+  &= \frac{n(\rr)}{4\pi
     \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
   +
-  \frac{n(\rr)}{\pi
+  \frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
@@
 \begin{align}
-  2 \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
-   \delta(\rr'-\rr) \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
+   \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
+   \delta(\rr'-\rr) \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
    \delta(\rr'-\rr) d\rr''
    \\
-   &= \delta(\rr'-\rr) n_2^S(\rr')
+   &= \frac 12 \delta(\rr'-\rr) n_2^S(\rr')
 \end{align}
@@
 \begin{align}
-  \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
+  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
   \frac{\delta n_3^{S}(\rr')}{\delta \sigma(\rr)} d\rr'
-  &= \int 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
+  &= \frac 12 \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
   n_2^S(\rr') d\rr' \\
-  &= 2 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
+  &= \frac 12 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
 \end{align}
@@
   r''\\
-  2 \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \int n(\rr'')
-  \delta'(|\rr'-\rr''| - \sigma(\rr')) d\rr''
+  \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \frac 12 \int n(\rr'')
+  \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
 \end{align}
@@
   n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
-    2\frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
-    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{2\pi \sigma(\rr)}
+    \frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
+    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi \sigma(\rr)}
     \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
-    + n_0^s(\rr') \right)
+    + \frac 12 n_0^s(\rr') \right)
 \end{align}
@@
   n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
-    2 \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
-    &= -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
+    \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
+    &= -\delta(\rr-\rr') \left( \frac 12 \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
     \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
     +
-    \frac{4}{\sigma} n_0^S(\rr') \right)
+    \frac{2}{\sigma} n_0^S(\rr') \right)
 \end{align}
@@
   \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
-  2 \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
-  \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
+  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
+  \frac 12 \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
   \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
@@
    \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
- 2 \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
- &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{2\pi \sigma(\rr)}
+ \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
+ &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi \sigma(\rr)}
    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
-   + \frac{\mathbf{n}_{V2}^S(\rr')}{\pi \sigma(\rr)^2} \right)
+   + \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
 \end{align}

### 2012-08-15 papers/contact/paper.tex
edit of paper and new contact data for mark II WB Tested everything on kipu.

--- 2012-08-14 papers/contact/paper.tex
@@
 \end{align}
-where $\eta \equiv \frac{4\pi}{24} \sigma^3 n$ is the filling fraction.  We
+where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.  We
 can thus readily compute the derivative of free energy with respect to
-hard-sphere radius:
+hard-sphere diameter:
 \begin{align}
-  \frac{dA_{HS}}{dR} &= 2\frac{dA_{HS}}{d\eta} \frac{d\eta}{d\sigma} \\
-  \derivation{
-    &= Nk_BT \left( \frac{4 - 6\eta}{(1-\eta)^2} + 2 \frac{4\eta - 3\eta^2}{(1-\eta)^3} \right) \frac{d\eta}{dR}
-    \\
-    &= Nk_BT \frac{4 - 4\eta - 6\eta + 6\eta^2 + 8\eta - 6\eta^2}{(1-\eta)^3} \frac{d\eta}{dR}
-    \\
-    &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{d\eta}{dR}
-    \\
-  }
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{6 \eta}{\sigma} \label{eq:dAhsdR}
+  \frac{dA_{HS}}{d\sigma} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{d\sigma} \\
+  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{\sigma} \label{eq:dAhsdR}
 \end{align}
-This derivative gives us the force on \emph{all} the hard
+This derivative gives us half the force on \emph{all} the hard
 spheres---since we're changing all their radii at once.  To compute
-the pressure on the spheres, we just need to divide by the total area,
-which means dividing by $N 4\pi \sigma^2$, since the relevant area
-is the area over which the molecules can make contact.
+the pressure on the spheres, we just need to divide by half the total
+area, which means dividing by $N 8\pi \sigma^2$, since the relevant
+area is the area over which the molecules can make contact.
 \begin{align}
-  p_{HS} &= \frac{1}{N 2\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
-  \derivation{
-    &= \frac14 \frac{1}{N 4\pi R^2} Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{R} \\
-    &= \frac{3}{4\pi R^3} \eta k_BT \frac{1 -
-      \frac{\eta}2}{(1-\eta)^3} \\
-  }
+  p_{HS} &= \frac{1}{N 8\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
   &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
 \end{align}
-Using the contact-value theorem, we can thus find the contact density
-and the well-known correlation function evaluated at contact.
+Using the contact-value theorem, we can thus find the well-known
+correlation function evaluated at contact.
 \begin{align}
-  n_\textit{contact} &= n \frac{1 - \frac{\eta}2}{(1-\eta)^3} \\
   g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3} \label{eq:cs-g}
@@
 Thus we can readily derive the contact density for the homogeneous
-hard-sphere fluid.  Extending this to the inhomogeneous fluid will
-require that we use density-functional theory.
-
-\section{Fundamental-Measure Theory}
-
-We use the White Bear version of the Fundamental-Measure Theory~(FMT)
-functional published in reference~\cite{roth2002whitebear}.  The FMT
-functional describes the excess free energy of a hard-sphere fluid.
-This particular FMT reduces to the Carnahan-Starling equation of state
-for homogeneous systems.
-\begin{equation}
-A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
-\end{equation}
-with integrands
-\begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
-\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-  n_3 + (1-n_3)^2 \ln(1-n_3)
-}{
-  36\pi n_3^2\left( 1 - n_3 \right)^2
-} ,
-\end{align}
-using the weighted densities
-\begin{align}
-  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
-  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) d\rr'
-\end{align}
-\begin{align}
-  \mathbf{n}_{V2} &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr - \rr'}{\left|\rr - \rr'\right|} d\rr' \quad \\
-  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{2\pi \sigma} , \quad
-  n_1 &= \frac{n_2}{2\pi \sigma} , \quad
-  n_0 = \frac{n_2}{\pi \sigma^2}
-\end{align}
-
-\derivation{
-  \begin{widetext}
-}
+hard-sphere fluid.  Extending this derivation to the inhomogeneous
+fluid requires that we find the pressure on particular spheres---or
+potentially particular parts of particular spheres.
 

### 2012-08-15 papers/contact/figs/plot-test-particle-cont-dens.py
added test-particle data and plotting, put data in paper.  Somethings not right scaling wise but not sure what yet Tested everything on node06.

@@ -1,0 +1,103 @@
+#!/usr/bin/python
+
+# We need the following two lines in order for matplotlib to work
+# without access to an X server.
+import matplotlib
+matplotlib.use('Agg')
+
+import pylab, numpy, sys, math, scipy
+
+from scipy.interpolate import interp1d
+
+if len(sys.argv) != 7:
+    print("Usage:  " + sys.argv[0] + " mc-filename.dat wb-filename.dat wbt-filename.dat wb-m2.dat test-particle-filename.dat out-filename.pdf")
+    exit(1)
+
+mcdata = numpy.loadtxt(sys.argv[1])
+dftdata = numpy.loadtxt(sys.argv[2])
+wbtdata = numpy.loadtxt(sys.argv[3])
+wbm2data = numpy.loadtxt(sys.argv[4])
+testdata = numpy.loadtxt(sys.argv[5])
+
+dft_len = len(dftdata[:,0])
+dft_dr = dftdata[2,0] - dftdata[1,0]
+test_len = len(testdata[:,0])
+
+# for x in a[:]: # make a slice copy of the entire list
+# ...    if len(x) > 6: a.insert(0, x)
+# ...
+# >>> a
+# ['defenestrate', 'cat', 'window', 'defenestrate']
+
+pos = 0
+minpos = [0]*test_len
+minposmc = [0]*test_len
+new_gA = [0]*test_len
+new_gmc = [0]*test_len
+new_nAmc = [0]*test_len
+new_nA = [0]*test_len
+mcoffset = 13
+
+for i in range(test_len):
+    new_nA[i] = numpy.interp(testdata[i,0], dftdata[:,0], dftdata[:,8])
+    new_gA[i] = numpy.interp(testdata[i,0], dftdata[:,0], dftdata[:,5])
+    new_nAmc[i] = numpy.interp(testdata[i,0], mcdata[:,0], mcdata[:,11])
+    new_gmc[i] = numpy.interp(testdata[i,0], mcdata[:,0], mcdata[:,4])/new_nAmc[i]
+
+
+n0 = dftdata[:,6]
+nA = dftdata[:,8]
+nAmc = mcdata[:,11]
+n0mc = mcdata[:,10]
+
[…]

### 2012-08-17 papers/contact/paper.tex
added test-particle data and plotting, put data in paper.  Somethings not right scaling wise but not sure what yet Tested everything on node06.
new conclusions and remove two fixmes Tests skipped!

@@
 \section{Introduction}
-
-\fixme{Figure out whether to stick with both $R$ and $\sigma$, or move
-to only $\sigma$ in the paper, which would have a certain appeal.}
-
-\fixme{Consider eliminating $n_0$ and/or $n_A$ from the plots to make
-  things cleaner.}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/test-particle-10}
+  \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+    fraction of 0.1.}
+  \label{fig:test-particle-10}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/test-particle-40}
+  \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+    fraction of 0.4.}
+  \label{fig:test-particle-40}
+\end{figure}
+
 \subsection{High density}
@@
 \section{Conclusion}
-We (will) have compared several functionals for the contact density of
-a hard-sphere fluid with simulations.  This quantity is of particular
-interest, as it plays a critical role in Statistical Associating Fluid
-Theory (SAFT), which is the basis of a number of recent classical
-density functionals.  We compute the contact density directly from the
-White Bear version of the FMT hard-sphere functional, and find
-reasonable agreement?
+We investigated several approximations to the contact value of the
+correlation function for inhomogeneous fluid distributions
+corresponding to flat walls, and both concave and convex walls.  We
+defined and simulated two averages of the correlation function, an
+asymmetric $A$ average centered at the location of one of the two
+spheres that is in contact, and a symmetric $S$ average centered at
+the point of contact of touching spheres.  For each average, we
+derived a functional form from FMT, and also found an approximation
+that has been used in the literature.  When compared with essentially
+exact Monte Carlo simulations, the $A$ correlation function derived
+from Fundamental Measure Theory in Section~\ref{sec:g-A} gives
+excellent results for each surface, and at both high density and low
+density.  The other three approximations that we studied all showed
+significant and systematic deviations under some circumstances.  Thus,
+we recommend that creators of SAFT-based classical density functionals
+consider using the $g_\sigma^A$ functional defined in
+Section~\ref{sec:g-A}.
 

### 2012-08-17 papers/contact/paper.tex
plots look much prettier now!  number of small changes made to them.  Got rid of test particle section in paper and first to appendix subsections Tested everything on node06.

--- 2012-08-15 papers/contact/paper.tex
@@
 
-\subsection{Test-particle result}
+%% \subsection{Test-particle result}
 
-Here we discuss comparison of the correlation function calculated via
-both Monte Carlo and the test-particle approach (see
-Figures~\ref{fig:inner-4-10} and~\ref{fig:inner-4-40}).
+%% Here we discuss comparison of the correlation function calculated via
+%% both Monte Carlo and the test-particle approach (see
+%% Figures~\ref{fig:inner-4-10} and~\ref{fig:inner-4-40}).
 
@@
 \end{figure}
+
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/test-particle-10}
+%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+%%     fraction of 0.1.}
+%%   \label{fig:test-particle-10}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/test-particle-40}
+%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+%%     fraction of 0.4.}
+%%   \label{fig:test-particle-40}
+%% \end{figure}
 
@@
 These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
-
-\subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Bear Mark I free energy $A_{HS}$}
-
-  First, let's rewrite the free energy in terms of a smaller set of weighted densities
-
-  \begin{align}
-    A_{HS} &= \int \left\{
-    -n_0 \ln\left( 1 - n_3\right)
-    + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-      n_3 + (1-n_3)^2 \ln(1-n_3)
-    }{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \right\}
-  \end{align}
-
-Taking derivatives we have
-
-  \begin{align}
-    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-    \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-    + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-      n_3(\mathbf{r}'))^2}
-    - \frac{n_2^3 -
-      3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-    \frac{n_3^2-5n_3+2}{n_3^3(1-n_3)^3} + 2\frac{\ln(1-n_3)}{n_3^3}
-    \right)
-    \\
-    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-    \frac{n_1}{1-n_3}
-    + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
-      (1-n_3)^2\ln(1-n_3)}{
-      36\pi n_3^2(1-n_3)^2
-    }
-    \\
-    \frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
-    \\
-    \frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &=
-    -\ln\left( 1 - n_3\right) \\
-    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-    \frac{\mathbf{n}_{V1}}{1-n_3}
-    - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
-      (1-n_3)^2\ln(1-n_3)}{
-      36\pi n_3^2(1-n_3)^2
-    }    \\
-    \frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
-    \frac{\mathbf{n}_{V2}}{1-n_3} \\
-  \end{align}
-
-\subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Brear Mark II free energy $A_{HS}$}
-
-Similarly to the White Bear I case we first write
-
-\begin{align} 
-A_{HS} &= \int \left\{
-    -n_0 \ln\left(1 - n_3\right)
-    + \left(1 + \frac{1}{9}n_3^2 \phi_2(n_3)\right)   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-    + \left(1 - \frac{4}{9}n_3\phi_3(n_3)\right)\frac{n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{24\pi (1-n_3)^2}
-    \right\}
-\end{align}
- where
- \begin{align}
-   \phi_2 &= \left(6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3)\right)\frac{1}{n_3^3}\\
-   \phi_3 &= \left(6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3)\right)\frac{1}{4n_3^3}
- \end{align}
-For the following derivitives we have
-\begin{align}
-   \frac{\delta \phi_2}{\delta n_3} &= \left((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2\right)\frac{3}{n_3^4}\\
-   \frac{\delta \phi_3}{\delta n_3} &= \left(-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3)\right)\frac{3}{4n_3^4}
- \end{align}
-And using these
-\begin{align}
-  \frac{\delta A_{HS}}{\delta n_3} &= \psi _1 + \psi _2 + \psi _3
-\end{align}
-where
-\begin{align}
-  \psi_1 &= \frac{n_0}{1-n_3}\\
-  \psi_2 &= \frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3}\left(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}\right) + \left(1 + \frac{1}{9}n_3^2\phi_2\right)\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2}\\
-  \psi_3 &= \frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2}\left(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}\right) + 2\left(1 - \frac{4}{9}n_3\phi_3\right)\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3}
-\end{align}
-We also have
-\begin{align}
-  \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2}\\
-  \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
-  \frac{\delta A_{HS}}{\delta n_0} &= -\ln(1 - n_3)\\
-  \frac{\delta A_{HS}}{\delta \mathbf{n}_1} &= -(1 + \frac{1}{9}n_3^2 \phi_2)\frac{\mathbf{n}_2}{1 - n_3}\\
-  \frac{\delta A_{HS}}{\delta \mathbf{n}_2} &= -(1 + \frac{1}{9}n_3^2\phi_2)\frac{\mathbf{n}_1}{1 - n_3} 
-  - (1 - \frac{4}{9}n_3\phi_3)\frac{6n_2\mathbf{n}_2}{24\pi(1 - n_3)^2}
-\end{align}
 
@@
 \end{document}
+
+%% \subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Bear Mark I free energy $A_{HS}$}
+
+%%   First, let's rewrite the free energy in terms of a smaller set of weighted densities
+
+%%   \begin{align}
+%%     A_{HS} &= \int \left\{
+%%     -n_0 \ln\left( 1 - n_3\right)
+%%     + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+%%     + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
+%%       n_3 + (1-n_3)^2 \ln(1-n_3)
+%%     }{
+%%       36\pi n_3^2(1-n_3)^2
+%%     }
+%%     \right\}
+%%   \end{align}
+
+%% Taking derivatives we have
+
+%%   \begin{align}
+%%     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
+%%     \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
+%%     + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
+%%       n_3(\mathbf{r}'))^2}
+%%     - \frac{n_2^3 -
+%%       3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
+%%     \frac{n_3^2-5n_3+2}{n_3^3(1-n_3)^3} + 2\frac{\ln(1-n_3)}{n_3^3}
+%%     \right)
+%%     \\
+%%     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
+%%     \frac{n_1}{1-n_3}
+%%     + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
+%%       (1-n_3)^2\ln(1-n_3)}{
+%%       36\pi n_3^2(1-n_3)^2
+%%     }
+%%     \\
+%%     \frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
+%%     \\
+%%     \frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &=
+%%     -\ln\left( 1 - n_3\right) \\
+%%     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
+%%     \frac{\mathbf{n}_{V1}}{1-n_3}
+%%     - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
+%%       (1-n_3)^2\ln(1-n_3)}{
+%%       36\pi n_3^2(1-n_3)^2
+%%     }    \\
+%%     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
+%%     \frac{\mathbf{n}_{V2}}{1-n_3} \\
+%%   \end{align}
+
+%% \subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Brear Mark II free energy $A_{HS}$}
+
+%% Similarly to the White Bear I case we first write
+
+%% \begin{align} 
+%% A_{HS} &= \int \left\{
+%%     -n_0 \ln\left(1 - n_3\right)
+%%     + \left(1 + \frac{1}{9}n_3^2 \phi_2(n_3)\right)   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
+%%     + \left(1 - \frac{4}{9}n_3\phi_3(n_3)\right)\frac{n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{24\pi (1-n_3)^2}
+%%     \right\}
+%% \end{align}
+%%  where
+%%  \begin{align}
+%%    \phi_2 &= \left(6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3)\right)\frac{1}{n_3^3}\\
+%%    \phi_3 &= \left(6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3)\right)\frac{1}{4n_3^3}
+%%  \end{align}
+%% For the following derivitives we have
+%% \begin{align}
+%%    \frac{\delta \phi_2}{\delta n_3} &= \left((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2\right)\frac{3}{n_3^4}\\
+%%    \frac{\delta \phi_3}{\delta n_3} &= \left(-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3)\right)\frac{3}{4n_3^4}
+%%  \end{align}
+%% And using these
+%% \begin{align}
+%%   \frac{\delta A_{HS}}{\delta n_3} &= \psi _1 + \psi _2 + \psi _3
+%% \end{align}
+%% where
+%% \begin{align}
+%%   \psi_1 &= \frac{n_0}{1-n_3}\\
+%%   \psi_2 &= \frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3}\left(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}\right) + \left(1 + \frac{1}{9}n_3^2\phi_2\right)\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2}\\
+%%   \psi_3 &= \frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2}\left(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}\right) + 2\left(1 - \frac{4}{9}n_3\phi_3\right)\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3}
+%% \end{align}
+%% We also have
+%% \begin{align}
+%%   \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2}\\
+%%   \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
+%%   \frac{\delta A_{HS}}{\delta n_0} &= -\ln(1 - n_3)\\
+%%   \frac{\delta A_{HS}}{\delta \mathbf{n}_1} &= -(1 + \frac{1}{9}n_3^2 \phi_2)\frac{\mathbf{n}_2}{1 - n_3}\\
+%%   \frac{\delta A_{HS}}{\delta \mathbf{n}_2} &= -(1 + \frac{1}{9}n_3^2\phi_2)\frac{\mathbf{n}_1}{1 - n_3} 
+%%   - (1 - \frac{4}{9}n_3\phi_3)\frac{6n_2\mathbf{n}_2}{24\pi(1 - n_3)^2}
+%% \end{align}

### 2012-08-20 papers/water-SAFT/paper.tex

@@
 water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,
-  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006, clark2006developing,
+  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
+  clark2006developing, chuev2006,
   lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
@@
 correctly, but can only be used at room temperature, and thus captures
-none of the temperature-dependence of water.  A number of classical
-density functionals have recently been produced that are based on
+none of the temperature-dependence of water.  A functional by Chuev
+and Skolov\cite{chuev2006} uses an ad hoc modification of FMT that can
+predict hydrophobic hydration near temperatures of 298~K, but does not produce a
+correct equation of state due to their method producting a high value
+for pressure. A number of classical density functionals have recently been produced that are based on
 Statistical Associating Fluid Theory (SAFT)\cite{ 
@@
 interaction between nanotube and water, the drying effect occurs at
-much longer range, in agreement with our results.  \fixme{Please read
-  this carefully, as I did some major changes.  I took a look at the
-  paper, and it seems that in Figure 3c (and the paragraph that ends
-  at the top of that page) they finally do a truly hydrophobic
-  interaction, which is a better comparison with what we did.  The
-  drying transition is inhibited by the attraction between tube and
-  water. }
+much longer range, in agreement with our results.
 

### 2012-08-20 papers/water-SAFT/paper.tex
cut redundant citation of same paper Tested everything on node05.

@@
 water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,
-  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
+  gloor2004accurate, gloor2007prediction, Jaqaman2004,
   clark2006developing, chuev2006,

### 2012-08-20 papers/contact/paper.tex
minor cleanups in contact paper Tested everything on node05.

@@
 point of contact.  Given this definition of density, their $g_\sigma$
-is given by: \fixme{Which of the following two equations is more clear
-  to you? We only want one.}
+is given by:
+% \fixme{Which of the following two equations is more clear to you? We
+%  only want one.}
 \begin{align}
-  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
-    + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
-    \\
+  %% g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
+  %%   + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
+  %%   + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
+  %%   \\
   g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
@@
 
+\bibliography{paper}% Produces the bibliography via BibTeX.
+
 \appendix
@@
 
-In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.
-We break these terms up as follows:
+In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq
+\ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta
+  \sigma(\mathbf{r})}$ terms that must be derived for both the S and A
+cases.  We break these terms up as follows:
 
@@
 
-We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.
-We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.
-Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.
-These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
+We must therefore compute both the $\frac{\delta A_{HS}}{\delta
+  n(\mathbf{r}')}$, etc., terms and the $\frac{\delta
+  n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.  We will
+compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms
+for both the functional derived in the White Bear I formulation and
+for that in the White Bear II formulation.  Following this we will
+compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$,
+etc., terms for both the A and S cases.  These latter derivitives will
+differ for the A and S cases according to the qualitive concepts
+described above.
 
@@
 \end{align}
-\fixme{double-check the directions of all the vector quantities...}
-
 This completes our exploration of the partial derivatives of weighted
@@
 
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
 \end{document}

### 2012-08-20 papers/contact/paper.tex
rewrite most of the theory section of contact paper Tested everything on node05.

@@
 \begin{align}
-  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
+  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2)
+  d\rr' \label{eq:FMn3} \\
   n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
@@
   n_1 &= \frac{n_2}{2\pi \sigma} , \quad
-  n_0 = \frac{n_2}{\pi \sigma^2}
+  n_0 = \frac{n_2}{\pi \sigma^2} \label{eq:FMrest}
 \end{align}
+
+\section{Theoretical Approaches}
 
@@
 
-\section{Asymmetrically averaged correlation function}\label{sec:g-A}
+\subsection{Asymmetrically averaged correlation function}\label{sec:g-A}
 
-When we consider an inhomogeneous system, we may prefer to ask for
-\emph{which} sphere we are interested in computing the contact
-density.  One way to answer this is to specify a sphere at a fixed
-position $\mathbf{r}$.  By making the hard-sphere radius dependent on
-position, we can compute the pressure on the hard spheres located at
-$\mathbf{r}$.  Since the hard-sphere radius enters the FMT free energy
-by means of the fundamental measures, this requires that we adjust the
-convolutions to have a diameter $\sigma$ which depends on the location
-of the sphere:
+We will begin our derivation of the locally averaged correlation
+function with the asymmetric definition of $g_\sigma^A(\rr)$ given in
+Equation~\ref{eq:gA}, which is averaged over contacts in which one of
+the two spheres is located at position~$\rr$.  This correlation
+function is related to the contact density averaged over the surface
+of a sphere located at $\rr$, and can thus be determined by finding
+the pressure on that sphere.  We find this pressure by treating the
+hard-sphere diameter as a function of the location of the sphere in
+question, when defining the fundamental measured (in
+Equations~\ref{eq:FMn3}-\ref{eq:FMrest}):
 \begin{align}
@@
 \end{align}
-Once we have done this---and similarly for the other fundamental
-measures---we derive a formula for the correlation function
-$g_\sigma^A(\rr)$ by finding the pressure on the surface of any hard
-spheres located at position $\rr$, and using the contact-density
-theorem to extract the contact density, from which we can establish
-the averaged correlation function at contact.
+Once we have done this, we derive a formula for the correlation
+function $g_\sigma^A(\rr)$ by finding the pressure on the surface of
+any hard spheres located at position $\rr$, and using the
+contact-density theorem to extract the contact density, from which we
+can establish the averaged correlation function at contact.
 \begin{align}
@@
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
-  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \\
-  g_\sigma^A(\rr) &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
+%  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
+%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
+  g_\sigma^A(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
   &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
@@
 \end{align}
-The functional derivative $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
-in the above formula must be computed with the asymmetric formulation
-for the fundamental measures.  In practice, we use the approximate
-White Bear functional to find an approximate expression for
-$g_\sigma^A$, which is given in Appendix~\ref{appendix:g-A}.  This
-expression will involve convolutions of derivatives of the free
-energy, making this formulation computationally more expensive than
-the free energy itself.
-
-In addition to the different densities in the denominators of Equations 
-\ref{eq:g-A-exact-a} and \ref{eq:g-S-exact-a} there is also a difference 
-between the treatment of the A and S cases in the derivatives 
-$\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ in Equations \ref{eq:g-A-exact} 
-and \ref{eq:g-S-exact}. The White Bear approximations for $A_{HS}$ 
-that we use involve a number of weighted densities and for the derivatives 
-of these we use two different conceptual approaches.  In the assymetric case, 
-a change in $\sigma(\mathbf{r})/2$ will refer to a change in the radius of a sphere 
-that is located at the position $\mathbf{r}$.  The weighted densities that are 
-affected by such a change will be those that are located a distance $\sigma/2$ away 
-from such a sphere.  We compute the assymetric derivatives In Appendix~\ref{appendix:g-A}.
+The functional derivative $\frac{\delta A_{HS}}{\delta
+  \sigma(\mathbf{r})}$ in the above formula is computed with the
+asymmetric formulation of the fundamental measures, as in
+Equation~\ref{eq:n3A}.  Details of some aspects of finding this
+derivative using FMT are given in Appendix~\ref{appendix:g-A}.  The
+equation for $g_\sigma^A$ involves convolutions of local derivatives
+of the free energy, making this formulation somewhat computationally
+more expensive than the free energy itself.
 
@@
 
-\section{Symmetrically averaged correlation function}\label{sec:g-S}
+\subsection{Symmetrically averaged correlation function}\label{sec:g-S}
+
+\begin{figure}
+\includegraphics[width=5cm]{figs/contact}
+\caption{An illustration of the prospect of changing the radius of
+  hard spheres only where they contact a given point, in order to find
+  the pressure on those spheres at just that point. }
+\label{fig:contact}
+\end{figure}
 
@@
 is defined in Equation~\ref{eq:gS}.  This corresponds to the
-correlation function averaged for
-spheres \emph{touching at a given point}.  It is a symmetric
-measure, since both spheres involved are treated in the same way.  In
-fact, the density of such spheres is just the density $n_0(\rr)$ from
-FMT.  See Figure~\ref{fig:n0} for an illustration $n_0$.  Since $n_0$
-is the fundamental measure of sphere surface contact, it makes sense
-to define a corresponding correlation function $g_\sigma^S$.  This is
-\emph{almost} what is done by Yu and  Wu in
-reference\cite{yu2002fmt-dft-inhomogeneous-associating}, as will be
-discussed in Section~\ref{sec:yuwu}.
+correlation function averaged for spheres \emph{touching at a given
+  point}.  It is a symmetric measure, since the two spheres involved
+are both treated in the same way.  The density of spheres that could
+be touching at the point $\rr$ is precisely the density $n_0(\rr)$
+from FMT (see Figure~\ref{fig:n0} for an illustration $n_0$).  Since
+$n_0$ is the fundamental measure of sphere surface contact, it makes
+sense to define a corresponding correlation function $g_\sigma^S$.
+This is closely related what is done by Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as will be discussed
+in Section~\ref{sec:yuwu}.
 
 The procedure for finding the symmetrically averaged correlation
-function $g_\sigma^S(\rr)$ is less straight-forward than 
+function $g_\sigma^S(\rr)$ is less straight-forward than that of the
+previous section, since we must find the average pressure exerted on
+the portion of the surface of spheres which is at position $\rr$.  We
+achieve this, as in the previous section, by defining a
+position-dependent hard-sphere diameter $\sigma(\rr)$, but this time
+we define the diameter to be dependent on the position at which the
+fundamental measure itself is evaluated:
 \begin{align}
   n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr)/2)d\rr'
+  \label{eq:n3S}
 \end{align}
-The hard-sphere radius $\sigma/2$ in this weighted density has an unusual
-position dependence:  it depends on the position at which we are
-evaluating the weighted density, rather than the position of the hard
-spheres themselves.  This has the effect of expanding spheres as they
-touch the point $\rr$, a concept illustrated in
-Figure~\ref{fig:contact}.  This enables a similar derivation to that
-which gave us $g_\sigma^A(\rr)$:
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/contact}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:contact}
-\end{figure}
-
-These qualitative concepts lead us to the derivative terms in $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$,
-which is in turn found in Equation~\ref{eq:p_{HS}^S}.  The rest of the
-derivation of the correlatation function results when we recognize that
-instead of dividing the pressure by $n(\mathbf{r})$ and $n_A(\mathbf{r})$
-as in the assymmetric case we will divide by $n_0(\mathbf{r})^2$
-since we have symmetry around the point $\mathbf{r}$.
-
+This has the effect of expanding spheres as they touch the point
+$\rr$, a concept illustrated in Figure~\ref{fig:contact}.This
+enables a similar derivation to that which gave us $g_\sigma^A(\rr)$:
 \begin{align}
@@
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^S}\\
-  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \\
-  g_\sigma^S(\rr) &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
+%  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
+%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
+  g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
   &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
@@
 where the functional derivative $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ in this case is computed with the symmetric
-formulation of the fundamental measures.  Complete expressions for
-$g_\sigma^S$ are included in Appendix~\ref{appendix:g-S}.\\
+  \sigma(\mathbf{r})}$ in the above formula is computed with the
+symmetric formulation of the fundamental measures, as in
+Equation~\ref{eq:n3S}.  Details of some aspects of finding this
+derivative using FMT are given in Appendix~\ref{appendix:g-S}.  The
+equation for $g_\sigma^S$ is computationally simpler than that for
+$g_\sigma^A$, as it only involves convolutions of the density itself,
+although it does require two new fundamental measures
+\begin{align}
+  n_{2'}(\rr) &= \int n(\rr') \delta'(\left|\rr - \rr'\right| -
+  \sigma/2) d\rr' \\
+  \mathbf{n}_{2V'}(\rr) &=
+  \int n(\rr') \delta'(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
+\end{align}
+where $\delta'$ is the derivative of the Dirac delta function.  These
+convolutions are easily expressed in reciprocal space, and can be
+easily and efficiently computed.
 
-\section{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+\subsection{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
 One approximation for the correlation function is that of
@@
 
-\section{Yu and Wu's symmetrically averaged functional}\label{sec:yuwu}
+\subsection{Yu and Wu's symmetrically averaged functional}\label{sec:yuwu}
 
 Yu and Wu developed a functional for the correlation function
-evaluated at contac, which should correspond to the symmetrically
-averaged correlation
-function~$g_\sigma^S$~\cite{yu2002fmt-dft-inhomogeneous-associating}.
-Although Yu and Wu use a symmetric formulation for $g_\sigma$, instead
-of using $n_0$ as the associated density, they use a density given by
+evaluated at contact, which is symmetrically
+averaged~\cite{yu2002fmt-dft-inhomogeneous-associating}.  Although Yu
+and Wu use a symmetric formulation for $g_\sigma$, instead of using
+$n_0$ as the associated density, they use a density given by
 \begin{align}
@@
 where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact.  Given this definition of density, their $g_\sigma$
-is given by:
+point of contact, and has the effect of reducing the density at
+interfaces.  Given this definition of density, their $g_\sigma$ is
+given by:
 % \fixme{Which of the following two equations is more clear to you? We
@@
 \end{align}
-which corresponds to our $g_\sigma^S$ and can be directly compared
-with it.
+which corresponds to our $g_\sigma^S$ so that they can be directly
+compared.
 

### 2012-08-21 papers/contact/paper.tex
rewrite one of the contact function appendices Tested everything on node08.

@@
   n_{2'}(\rr) &= \int n(\rr') \delta'(\left|\rr - \rr'\right| -
-  \sigma/2) d\rr' \\
+  \sigma/2) d\rr' \label{n2p} \\
   \mathbf{n}_{2V'}(\rr) &=
-  \int n(\rr') \delta'(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
+  \int n(\rr') \delta'(\left|\rr - \rr'\right| - \sigma/2)
+  \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
+  \label{n2vp}
 \end{align}
@@
   \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
+  \\
+  &= -\delta(\rr-\rr') \frac12 n_{2'}(\rr')
 \end{align}
 where we recognize that the convolution---while not one of the standard
-FMT weighted density---is a simple convolution of the density.  We can
+FMT weighted density---is a simple convolution of the density, which
+we define in Equation~\ref{n2p} as $n_{2'}$.  We can
 go on to compute the derivatives of the remaining weighted densities:
@@
     + \frac 12 n_0^s(\rr') \right)
+    \\
+    &= -\delta(\rr-\rr')\left(
+    \frac{n_{2'}(\rr')}{4\pi\sigma} + \frac12n_0(\rr')
+    \right)
 \end{align}
@@
     \frac{2}{\sigma} n_0^S(\rr') \right)
+    \\
+    &= -\delta(\rr-\rr')\left(
+    \frac{n_{2'}(\rr')}{2\pi\sigma^2}
+    + \frac{2}{\sigma} n_0(\rr')
+    \right)
 \end{align}
@@
   \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
+  \\
+  &= -\delta(\rr-\rr')
+   \frac12\mathbf{n}_{V2'}(\rr')
 \end{align}
@@
    + \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
+   \\
+   &= -\delta(\rr-\rr')\left(
+   \frac{\mathbf{n}_{V2'}(\rr')}{4\pi \sigma}
+   + \frac{\mathbf{n}_{V2}(\rr')}{2 \pi \sigma^2}
+   \right)
 \end{align}
@@
 The above equations involve just two additional weighted densities
-beyond those required for FMT: one scalar density computed with a
+beyond those required for FMT: one scalar density $n_{2'}$ computed with a
 derivative of the Dirac delta function, and one analogous vector
-density.
+density $\mathbf{n}_{2V'}$, defined in Equations~\ref{n2p} and \ref{n2vp}.
 

### 2012-08-21 papers/contact/paper.tex
added first attempt at results writing Tested everything on node04.

--- 2012-08-20 papers/contact/paper.tex
@@
 point of contact.  Given this definition of density, their $g_\sigma$
-is given by:
-% \fixme{Which of the following two equations is more clear to you? We
-%  only want one.}
+is given by: \fixme{Which of the following two equations is more clear
+  to you? We only want one.}
 \begin{align}
-  %% g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-  %%   + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
-  %%   + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
-  %%   \\
+  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
+    + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
+    + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
+    \\
   g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
@@
 
+We can see that the White Bear DFT calulations for the densities at varying distances from the respective surfaces match up well with the monte-carlo simulated densities.
+The densities slope up towards and reach a maximum at the surfaces as the diffusion cutoff in that direction causes hard spheres to collect.
+We also see that $n_0$ slopes sharply downward towards the different surfaces, beginning the trend at about one radius away.
+Part of the average that is taken for spheres in these locations will be taken within a volume for which the sphere density is uniformily zero.
+A similar trend is followed by $n_A$ beginning at about two radii away.
+A sphere that is in contact with another at any position must be two radii away from the center of the first.
+Thus, within two radii of the surface, one enters into a region where some of the contact volume is past the zero density boundary.
+
+The S case correlation functions do not match the monte carlo simulated correlation as well as the functionals of Yu and Wu.
+For the walls and the sperical cavity, the monte carlo simulations show this function going to zero at these surfaces, roughly beginning its downward trend at about one sphere radius out.
+This stands to reason since that because there is an increased density up against the wall, there is little chance that points within a radius of the wall and the center of those spheres will be a contact point between two spheres, since the hard spheres cannot overlap.
+
+The A correlation White Bear calculations match up very well with monte-carlo simulation.
+Although they seem to be both very close, the Gross functional deos not have the near perfect overlap with simulation that the White Bear calculation deos.
+
 \begin{figure}
@@
 
+Although the general density attributes look similar to those of the low density calculations as one moves away from the surfaces, the range of difference is altered.
+The spikes at the surfaces reach a height of roughly five times the homogeneous average, whereas with the low densities the difference is closer to a factor of $0.5$.
+With a higher density of spheres to pack in those that are up against the wall, it is more difficult for those spheres to ever find their way out and away.
+We also notice that the downward trend in both $n_0$ and $n_A$ have similar characteristics but are over all much more flat.
+(Don't see why!)
+
+We once again see that in the S case, although neither functionals tested fit perfectly, Yu and Wu's functional is a closer fit than the White Bear's.
+We can see in the monte carlo's oscillations over increased distance from the surfaces that there are layers of spheres that are formed.
+If one picks a point in one of these layters, that point is more likely to be a contact point between spheres, whereas if one picks a sphere a little over half a radius away, the likelyhood is at a minimum.
+There is then another maximum at roughly one radius away from the surface.
+(This suggests close packed layers?)
+
+As with the lower densities, the A case White Bear fits to the monte carlo simulations very well.  The Gross functional seems to be accurate until it falls off close to surfaces.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Section~\ref{sec:g-A}.
-
-\bibliography{paper}% Produces the bibliography via BibTeX.
 
@@
 
-In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq
-\ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ terms that must be derived for both the S and A
-cases.  We break these terms up as follows:
+In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$ terms that must be derived for both the S and A cases.
+We break these terms up as follows:
 
@@
 
-We must therefore compute both the $\frac{\delta A_{HS}}{\delta
-  n(\mathbf{r}')}$, etc., terms and the $\frac{\delta
-  n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.  We will
-compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms
-for both the functional derived in the White Bear I formulation and
-for that in the White Bear II formulation.  Following this we will
-compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$,
-etc., terms for both the A and S cases.  These latter derivitives will
-differ for the A and S cases according to the qualitive concepts
-described above.
+We must therefore compute both the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms and the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.
+We will compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms for both the functional derived in the White Bear I formulation and for that in the White Bear II formulation.
+Following this we will compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms for both the A and S cases.
+These latter derivitives will differ for the A and S cases according to the qualitive concepts described above.
 
@@
 \end{align}
+\fixme{double-check the directions of all the vector quantities...}
+
 This completes our exploration of the partial derivatives of weighted
@@
 
+\bibliography{paper}% Produces the bibliography via BibTeX.
+
 \end{document}

### 2012-08-21 papers/contact/paper.tex

--- 2012-08-21 papers/contact/paper.tex
@@
 which are shown in Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.
+
+Although the general density attributes look similar to those of the low density calculations as one moves away from the surfaces, the range of difference is altered.
+The spikes at the surfaces reach a height of roughly five times the homogeneous average, whereas with the low densities the difference is closer to a factor of $0.5$.
+With a higher density of spheres to pack in those that are up against the wall, it is more difficult for those spheres to ever find their way out and away.
+We also notice that the downward trend in both $n_0$ and $n_A$ have similar characteristics but are over all much more flat.
+(Don't see why!)
+
+We once again see that in the S case, although neither functionals tested fit perfectly, Yu and Wu's functional is a closer fit than the White Bear's.
+We can see in the monte carlo's oscillations over increased distance from the surfaces that there are layers of spheres that are formed.
+If one picks a point in one of these layters, that point is more likely to be a contact point between spheres, whereas if one picks a sphere a little over half a radius away, the likelyhood is at a minimum.
+There is then another maximum at roughly one radius away from the surface.
+(This suggests close packed layers?)
+
+As with the lower densities, the A case White Bear fits to the monte carlo simulations very well.  The Gross functional seems to be accurate until it falls off close to surfaces.
 

### 2012-08-21 papers/contact/paper.tex
switch A and S to $A$ and $S$ in contact paper Tested everything on node05.

@@
 
-The S case correlation functions do not match the monte carlo simulated correlation as well as the functionals of Yu and Wu.
-For the walls and the sperical cavity, the monte carlo simulations show this function going to zero at these surfaces, roughly beginning its downward trend at about one sphere radius out.
-This stands to reason since that because there is an increased density up against the wall, there is little chance that points within a radius of the wall and the center of those spheres will be a contact point between two spheres, since the hard spheres cannot overlap.
-
-The A correlation White Bear calculations match up very well with monte-carlo simulation.
-Although they seem to be both very close, the Gross functional deos not have the near perfect overlap with simulation that the White Bear calculation deos.
+The $S$ case correlation functions do not match the monte carlo
+simulated correlation as well as the functionals of Yu and Wu.  For
+the walls and the sperical cavity, the monte carlo simulations show
+this function going to zero at these surfaces, roughly beginning its
+downward trend at about one sphere radius out.  This stands to reason
+since that because there is an increased density up against the wall,
+there is little chance that points within a radius of the wall and the
+center of those spheres will be a contact point between two spheres,
+since the hard spheres cannot overlap.
+
+The $A$ correlation White Bear calculations match up very well with
+monte-carlo simulation.  Although they seem to be both very close, the
+Gross functional deos not have the near perfect overlap with
+simulation that the White Bear calculation deos.
 
@@
 
-We once again see that in the S case, although neither functionals tested fit perfectly, Yu and Wu's functional is a closer fit than the White Bear's.
-We can see in the monte carlo's oscillations over increased distance from the surfaces that there are layers of spheres that are formed.
-If one picks a point in one of these layters, that point is more likely to be a contact point between spheres, whereas if one picks a sphere a little over half a radius away, the likelyhood is at a minimum.
-There is then another maximum at roughly one radius away from the surface.
-(This suggests close packed layers?)
-
-As with the lower densities, the A case White Bear fits to the monte carlo simulations very well.  The Gross functional seems to be accurate until it falls off close to surfaces.
+We once again see that in the $S$ case, although neither functionals
+tested fit perfectly, Yu and Wu's functional is a closer fit than the
+White Bear's.  We can see in the monte carlo's oscillations over
+increased distance from the surfaces that there are layers of spheres
+that are formed.  If one picks a point in one of these layers, that
+point is more likely to be a contact point between spheres, whereas if
+one picks a sphere a little over half a radius away, the likelyhood is
+at a minimum.  There is then another maximum at roughly one radius
+away from the surface.  (This suggests close packed layers?)
+
+As with the lower densities, the $A$ case White Bear fits to the monte
+carlo simulations very well.  The Gross functional seems to be
+accurate until it falls off close to surfaces.
 
@@
 \ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ terms that must be derived for both the S and A
+  \sigma(\mathbf{r})}$ terms that must be derived for both the $S$ and A
 cases.  We break these terms up as follows:
@@
 compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$,
-etc., terms for both the A and S cases.  These latter derivitives will
-differ for the A and S cases according to the qualitive concepts
+etc., terms for both the $A$ and $S$ cases.  These latter derivitives will
+differ for the $A$ and $S$ cases according to the qualitive concepts
 described above.

### 2012-08-29 papers/contact/paper.tex
edits on paper Tests skipped!

@@
 
-\begin{figure}
-\includegraphics[width=5cm]{figs/contact}
-\caption{An illustration of the prospect of changing the radius of
-  hard spheres only where they contact a given point, in order to find
-  the pressure on those spheres at just that point. }
-\label{fig:contact}
-\end{figure}
-
 We now address the symmetrically averaged correlation function, which
@@
 correlation function averaged for spheres \emph{touching at a given
-  point}.  It is a symmetric measure, since the two spheres involved
-are both treated in the same way.  The density of spheres that could
-be touching at the point $\rr$ is precisely the density $n_0(\rr)$
-from FMT (see Figure~\ref{fig:n0} for an illustration $n_0$).  Since
-$n_0$ is the fundamental measure of sphere surface contact, it makes
-sense to define a corresponding correlation function $g_\sigma^S$.
-This is closely related what is done by Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, as will be discussed
-in Section~\ref{sec:yuwu}.
-
-The procedure for finding the symmetrically averaged correlation
-function $g_\sigma^S(\rr)$ is less straight-forward than that of the
-previous section, since we must find the average pressure exerted on
-the portion of the surface of spheres which is at position $\rr$.  We
-achieve this, as in the previous section, by defining a
-position-dependent hard-sphere diameter $\sigma(\rr)$, but this time
-we define the diameter to be dependent on the position at which the
-fundamental measure itself is evaluated:
-\begin{align}
-  n_3^S(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr)/2)d\rr'
-  \label{eq:n3S}
-\end{align}
-This has the effect of expanding spheres as they touch the point
-$\rr$, a concept illustrated in Figure~\ref{fig:contact}.This
-enables a similar derivation to that which gave us $g_\sigma^A(\rr)$:
-\begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^S}\\
+  point}.  In this case, we conceptually would like to evaluate the
+pressure felt by the surface of spheres where that surface is located
+at point~$\rr$.  We can approximate this value by assuming that this
+pressure will be simply related to the free energy density at
+point~$\rr$.  Through math similar to the previous derivations, this
+leads to the expression
+\begin{align}
+  %% p_{HS}^S(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\partial
+  %%   a_{HS}(\rr)}{\partial \sigma} \label{eq:p_{HS}^S}\\
 %  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
@@
   g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  &= \frac{1}{n_0(\rr)^2}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
-\end{align}
-where the functional derivative $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ in the above formula is computed with the
-symmetric formulation of the fundamental measures, as in
-Equation~\ref{eq:n3S}.  Details of some aspects of finding this
-derivative using FMT are given in Appendix~\ref{appendix:g-S}.  The
-equation for $g_\sigma^S$ is computationally simpler than that for
-$g_\sigma^A$, as it only involves convolutions of the density itself,
-although it does require two new fundamental measures
-\begin{align}
-  n_{2'}(\rr) &= \int n(\rr') \delta'(\left|\rr - \rr'\right| -
-  \sigma/2) d\rr' \label{n2p} \\
-  \mathbf{n}_{2V'}(\rr) &=
-  \int n(\rr') \delta'(\left|\rr - \rr'\right| - \sigma/2)
-  \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
-  \label{n2vp}
-\end{align}
-where $\delta'$ is the derivative of the Dirac delta function.  These
-convolutions are easily expressed in reciprocal space, and can be
-easily and efficiently computed.
+  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 2\pi \sigma^2}
+  \frac{\partial \Phi(\rr)}{\partial \sigma} \label{eq:g-S}
+\end{align}
+where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
+dimensionless free energy density.  This requires that we evaluate the
+derivatives of the fundamental measures $n_\alpha(\rr)$ with respect
+to diameter.  These derivatives, when performed exact, lead to
+derivatives of the $\delta$~function, resulting in a functional gives
+poor results.  We find that our $g_\sigma^S$ is considerably improved
+by using integration by parts to eliminate derivatives of the
+$\delta$~function by approximating the density to be smoothly varying,
+which leads to derivatives such as:
+\begin{align}
+  \frac{\partial n_2(\rr)}{\partial \sigma} &\approx \frac{2}{\sigma}n_2(\rr)
+\end{align}
+We note that the density at a hard surface is \emph{not} smoothly
+varying, but that the removal of these terms improves the functional,
+while at the same time reducing its computational cost by avoiding the
+need to calculate any additional weighted densities or convolutions.
+
+The final formula for the symmetrically averaged correlation function
+under these approximations is:
+\begin{align}
+  g_\sigma^S &= \text{\textcolor{red}{put formula in here, if it is
+      short enough}}
+\end{align}
 
@@
 This formula is arrived at by using the density averaged over all
-spheres that could be touching a sphere at point $\rr$ in the
+spheres that could be touching a sphere at point~$\rr$ in the
 Carnahan-Starling equation for the correlation function at contact,
@@
 
-\section{Comparison with simulation}
-
-We performed a Monte-Carlo simulation of the hard sphere fluid to
-measure the contact density as a function of position for a few simple
-geometries.  For each scenario, we compute and compare several
-quantities.  We compare the Monte Carlo density with the DFT
-prediction.  We also plot Monte Carlo results for the two contact
-densities: the contact density evaluated at contact, and the contact
-density evaluated at the sphere location.  Alongside these, we plot
-the contact density of Yu and
-Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, the contact density
-of Gross~\cite{gross2009density}, and the ``contact density at
-sphere'' from Section~\ref{sec:g-A}.  In principle, the
-functional of Yu (FIXME) should match the CenConDensity, while the DFT
-at sphere should match the ConDensity...
-
-\subsection{Hard spheres at a hard wall}
-
-We begin by examining the simplest situation with an inhomogeneous
-density:  the case of hard spheres near a single hard wall.  This is
-shown in Figures~\ref{fig:walls-10}-\ref{fig:walls-40}.
+%% \section{Method}
+
+%% We implemented the above functionals as part of the Deft software
+%% package~\cite{deft}, which uses a regular Cartesian grid with a grid
+%% spacing of $0.05\sigma$.  Convolutions are implemented with fast
+%% fourier transforms.  For comparison, we performed Monte Carlo
+%% simulations of the hard-sphere fluid.  We performed these simulations
+%% in a fixed-number ensemble.
 
@@
 \end{figure}
+
+\section{Comparison with simulation}
+
+We performed a Monte-Carlo simulation of the hard sphere fluid to
+measure the contact value of the correlation function for several
+simple inhomogeneous configurations.  For each configuration, we
+compute the mean density, and the contact values of the correlation
+function, averaged as defined in Equations~\ref{eq:gA} and
+\ref{eq:gS}.  We compare these with the functionals presented in
+sections~\ref{sec:g-A} to~\ref{sec:yuwu}.  We constructed our
+functionals using both the original White Bear
+functional~\cite{roth2002whitebear} as well as the mark II version of
+the White Bear functional~\cite{hansen2006density}, but the results
+were essentially indistinguishable on our plots, so we exclusively
+show the results due to the original White Bear functional.
+
+We simulate the inhomogeneous hard sphere fluid at four hard-wall
+interfaces.  The first and simplest is a simple flat hard wall.  We
+then study two convex hard surfaces.  One is an excluded sphere with
+diameter $2\sigma$, which corresponds to a ``test particle''
+simulation with one of a hard sphere at the origin with diameter
+$\sigma$.  The second is an excluded sphere with diameter $6\sigma$,
+which demonstrates behavior typical of mildly convex hard surfaces.
+Finally, we study a concave surface given by a hard cavity in which
+our fluid is free to move up to a diameter of $16\sigma$, which
+demonstrates behavior typical of mildly concave surfaces.  In each
+case, we performed a low-density (filling fraction 0.1) and high-density
+(filling fraction 0.4) simulation.  We performed additional
+computations at over a wider range of curvatures and densities, but
+chose these as typical examples.
+
 \begin{figure}
@@
 \end{figure}
+
+\subsection{Low density}
+
+We begin by presenting our low-density results, corresponding to a
+filling fraction of 0.1, which are shown in
+Figures~\ref{fig:walls-10}-\ref{fig:outer-10}.  At this low density,
+the contact value of the correlation function in the bulk is only 1.3,
+indicating that correlations are indeed small, the fluid should be
+relatively easy to model.  Indeed, the contact density at the hard
+surface is only around 50\% higher than the bulk, and the FMT
+predicted density is close to indistinguishable from the true
+density for each of the four configurations.
+
+The $g_\sigma^A$ correlation function in each configuration is very
+flat, with only small, smooth changes as the surface is approached.
+Our functional $g_\sigma^A$ very closely matches the Monte Carlo
+predictions in each case, while that of Gross consistently
+underestimates the correlation at the interface by about
+\textcolor{red}{few percent}.
+
+The $g_\sigma^S$ correlation function shows considerably more
+structure, as well as additional variation due to the curvature of the
+hard surface.
+
 \begin{figure}
@@
 
-\subsection{Hard spheres confined in a spherical cavity}
-
-The next geometry we consider is a number of hard spheres confined
-within a spherical cavity.  In
-Figures~\ref{fig:sphere-8}-\ref{fig:sphere-16}, we show the densities
-and contact densities for several cavity sizes \fixme{at a variety of
-  filling fractions}.  We chose cavities that are large enough that
-the center of the sphere is bulk-like, so that we can interpret the
-behavior near the interface as behavior that should be typical of the
-behavior of the hard sphere fluid near a concave surface.
-
-\subsection{Hard spheres around a hard sphere}
-
-It is also interesting to study the hard sphere fluid at a convex
-surface.  We did so by placing a (larger) hard sphere in the
-hard-sphere fluid, to see how the correlation function behaves near
-the surface of that hard sphere.  This is particularly relevant when
-we consider using classical density-functional theory to model
-solvation.
-
-\section{Results}
-
-%% \subsection{Test-particle result}
-
-%% Here we discuss comparison of the correlation function calculated via
-%% both Monte Carlo and the test-particle approach (see
-%% Figures~\ref{fig:inner-4-10} and~\ref{fig:inner-4-40}).
-
-\subsection{Low density}
-
-Here we discuss the low-density results (filling fraction of 0.1),
-which are shown in Figures~\ref{fig:walls-10}-\ref{fig:outer-10}.
-
-We can see that the White Bear DFT calulations for the densities at varying distances from the respective surfaces match up well with the monte-carlo simulated densities.
-The densities slope up towards and reach a maximum at the surfaces as the diffusion cutoff in that direction causes hard spheres to collect.
-We also see that $n_0$ slopes sharply downward towards the different surfaces, beginning the trend at about one radius away.
-Part of the average that is taken for spheres in these locations will be taken within a volume for which the sphere density is uniformily zero.
-A similar trend is followed by $n_A$ beginning at about two radii away.
-A sphere that is in contact with another at any position must be two radii away from the center of the first.
-Thus, within two radii of the surface, one enters into a region where some of the contact volume is past the zero density boundary.
+We can see that the White Bear DFT calulations for the densities at
+varying distances from the respective surfaces match up well with the
+monte-carlo simulated densities.  The densities slope up towards and
+reach a maximum at the surfaces as the diffusion cutoff in that
+direction causes hard spheres to collect. We also see that $n_0$
+slopes sharply downward towards the different surfaces, beginning the
+trend at about one radius away.  Part of the average that is taken for
+spheres in these locations will be taken within a volume for which the
+sphere density is uniformily zero. A similar trend is followed by
+$n_A$ beginning at about two radii away.  A sphere that is in contact
+with another at any position must be two radii away from the center of
+the first.  Thus, within two radii of the surface, one enters into a
+region where some of the contact volume is past the zero density
+boundary.
 
@@
 compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms
-for both the functional derived in the White Bear I formulation and
-for that in the White Bear II formulation.  Following this we will
-compute the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$,
-etc., terms for both the $A$ and $S$ cases.  These latter derivitives will
+for both the functional derived from the White Bear
+functional\cite{roth2002whitebear} and from the White Bear mark
+II functional\cite{hansen2006density}.  Following this we will compute
+the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc.,
+terms for both the $A$ and $S$ cases.  These latter derivitives will
 differ for the $A$ and $S$ cases according to the qualitive concepts
@@
 \begin{align}
-  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr'')/2) d\rr''
+  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
 \end{align}
@@
   \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
-  \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) \delta(\rr-\rr'') d\rr'' \\
-   &= n (\rr) \delta(|\rr' - \rr| - \sigma(\rr)/2)
+  \frac 12 \int n (\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \delta(\rr-\rr'') d\rr'' \\
+   &= n (\rr) \delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
 \end{align}
@@
 \begin{align}
-  n_2^{A}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr'')/2) d \rr''\\
-  \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)\\
+  n_2^{A}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d \rr''\\
+  \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)\\
 \end{align}
@@
   n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
+  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
   \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
+  &= \frac{n(\rr)}{4\pi
+    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
   -
   \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2)
+    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
 \end{align}
@@
   n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
-  \delta(|\rr'-\rr''| - \sigma(\rr'')/2) \\
+  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
   \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta'(|\rr'-\rr| - \sigma(\rr)/2)
+  &= \frac{n(\rr)}{2\pi
+    \sigma(\rr)^2}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
   -
   2 \frac{n(\rr)}{\pi
-    \sigma(\rr)^3}\delta(|\rr'-\rr| - \sigma(\rr)/2)
+    \sigma(\rr)^3}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
 \end{align}
@@
 \begin{align}
-  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(|\rr' - \rr''| - \sigma(\rr'')/2)
+  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
     \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(|\rr' - \rr| - \sigma(\rr)/2)
+  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
     \frac{\rr-\rr'}{|\rr-\rr'|}\\
@@
   \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  \delta(|\mathbf{r'}-\rr''| - \sigma(\rr'')/2) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
+  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
   \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
+  &= -\frac{n(\rr)}{4\pi
+    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
   +
   \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(|\rr'-\rr| - \sigma(\rr)/2) \frac{\rr-\rr'}{|\rr-\rr'|}
+    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
 \end{align}
@@
 densities of the $A$ variety.
+
+\end{widetext}
+\end{document}
 
@@
 \begin{align}
-  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\left|\rr' - \rr''\right| -\sigma(\rr')/2) d\rr''
+  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
 \end{align}
@@
    \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
-   \delta(\rr'-\rr) \frac 12 \int n (\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
+   \delta(\rr'-\rr) \frac 12 \int n (\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
    \delta(\rr'-\rr) d\rr''
@@
 \begin{align}
-  n_2^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf
+  n_2^{S}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf
   r''\\
-  \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr-\rr') \frac 12 \int n(\rr'')
-  \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
+  \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= \delta(\rr-\rr') \frac 12 \int n(\rr'')
+  \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
   \\
-  &= -\delta(\rr-\rr') \frac12 n_{2'}(\rr')
+  &= \delta(\rr-\rr') \frac12 n_{2'}(\rr')
 \end{align}
@@
 \begin{align}
-  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
+  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''\\
     \frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
-    -\delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi \sigma(\rr)}
-    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
-    + \frac 12 n_0^s(\rr') \right)
+    \delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi \sigma(\rr)}
+    \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
+    - \frac 12 n_0^s(\rr') \right)
     \\
-    &= -\delta(\rr-\rr')\left(
-    \frac{n_{2'}(\rr')}{4\pi\sigma} + \frac12n_0(\rr')
+    &= \delta(\rr-\rr')\left(
+    \frac{n_{2'}(\rr')}{4\pi\sigma} - \frac12n_0(\rr')
     \right)
@@
 \begin{align}
-  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''\\
+  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''\\
     \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
-    &= -\delta(\rr-\rr') \left( \frac 12 \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
-    \delta'(|\rr'-\rr''| - \sigma(\rr')/2) d\rr''
-    +
+    &= \delta(\rr-\rr') \left( \frac 12 \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
+    \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
+    -
     \frac{2}{\sigma} n_0^S(\rr') \right)
     \\
-    &= -\delta(\rr-\rr')\left(
+    &= \delta(\rr-\rr')\left(
     \frac{n_{2'}(\rr')}{2\pi\sigma^2}
-    + \frac{2}{\sigma} n_0(\rr')
+    - \frac{2}{\sigma} n_0(\rr')
     \right)
@@
 \begin{align}
-  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(|\rr' - \rr''| - \sigma(\rr')/2)
+  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
   \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= -\delta(\rr - \rr')
-  \frac 12 \int n(\rr'') \delta'(|\rr' - \rr''| - \sigma(\rr')/2)
+  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= \delta(\rr - \rr')
+  \frac 12 \int n(\rr'') \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
   \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
   \\
-  &= -\delta(\rr-\rr')
+  &= \delta(\rr-\rr')
    \frac12\mathbf{n}_{V2'}(\rr')
@@
 \begin{align}
- \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(|\rr' - \rr''| - \sigma(\rr')/2)d\mathbf r''
+ \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''
    \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
  \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
- &= -\delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi \sigma(\rr)}
-   \delta'(|\rr'-\rr''| - \sigma(\rr')/2) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
-   + \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
+ &= \delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi \sigma(\rr)}
+   \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
+   - \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
    \\
-   &= -\delta(\rr-\rr')\left(
+   &= \delta(\rr-\rr')\left(
    \frac{\mathbf{n}_{V2'}(\rr')}{4\pi \sigma}
-   + \frac{\mathbf{n}_{V2}(\rr')}{2 \pi \sigma^2}
+   - \frac{\mathbf{n}_{V2}(\rr')}{2 \pi \sigma^2}
    \right)

### 2012-08-29 papers/contact/paper.tex
a few edits to results and discussion of contact paper Tested everything on darcy.

@@
 
+\section{Comparison with simulation}
+
 \begin{figure}
@@
 \end{figure}
-
-\section{Comparison with simulation}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-10}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-4-10}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-10}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-12-10}
-\end{figure}
-
 \subsection{Low density}
@@
 predictions in each case, while that of Gross consistently
-underestimates the correlation at the interface by about
-\textcolor{red}{few percent}.
+underestimates the correlation at the interface by
+\textcolor{red}{several percent}.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/inner-4-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-4-10}
+\end{figure}
 
@@
 structure, as well as additional variation due to the curvature of the
-hard surface.
+hard surface.  The symmetric correlation function is nonzero at
+locations where spheres may touch, which for a convex hard surface
+means that $g_\sigma^S$ may be nonzero in the volume in which hard
+spheres are excluded.  In every configuration studied, the agreement
+between the theoretical predictions and the Monte Carlo simulation in
+each case is very poor in the region where there should be no contacts
+at all.  Because $n_0$ is comparable to its bulk value in this region,
+this means that these functionals predict a significant number of
+contacts in the region where there should be none.  The correlation
+function of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
+and ours described in Section~\ref{sec:g-S} give similar results, with
+slightly larger errors in our prediction.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/inner-12-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-12-10}
+\end{figure}
 
@@
 
-We can see that the White Bear DFT calulations for the densities at
-varying distances from the respective surfaces match up well with the
-monte-carlo simulated densities.  The densities slope up towards and
-reach a maximum at the surfaces as the diffusion cutoff in that
-direction causes hard spheres to collect. We also see that $n_0$
-slopes sharply downward towards the different surfaces, beginning the
-trend at about one radius away.  Part of the average that is taken for
-spheres in these locations will be taken within a volume for which the
-sphere density is uniformily zero. A similar trend is followed by
-$n_A$ beginning at about two radii away.  A sphere that is in contact
-with another at any position must be two radii away from the center of
-the first.  Thus, within two radii of the surface, one enters into a
-region where some of the contact volume is past the zero density
-boundary.
-
-The $S$ case correlation functions do not match the monte carlo
-simulated correlation as well as the functionals of Yu and Wu.  For
-the walls and the sperical cavity, the monte carlo simulations show
-this function going to zero at these surfaces, roughly beginning its
-downward trend at about one sphere radius out.  This stands to reason
-since that because there is an increased density up against the wall,
-there is little chance that points within a radius of the wall and the
-center of those spheres will be a contact point between two spheres,
-since the hard spheres cannot overlap.
-
-The $A$ correlation White Bear calculations match up very well with
-monte-carlo simulation.  Although they seem to be both very close, the
-Gross functional deos not have the near perfect overlap with
-simulation that the White Bear calculation deos.
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/test-particle-10}
+%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+%%     fraction of 0.1.}
+%%   \label{fig:test-particle-10}
+%% \end{figure}
+
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/test-particle-40}
+%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
+%%     fraction of 0.4.}
+%%   \label{fig:test-particle-40}
+%% \end{figure}
+
+\subsection{High density}
 
@@
 \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-10}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.1.}
-%%   \label{fig:test-particle-10}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-40}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.4.}
-%%   \label{fig:test-particle-40}
-%% \end{figure}
-
-\subsection{High density}
 

### 2012-08-29 papers/contact/paper.tex
a few more edits to contact paper Tested everything on darcy.

@@
 
-\section{Comparison with simulation}
-
 \begin{figure}
@@
 \end{figure}
+
+\section{Comparison with simulation}
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/inner-4-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-4-10}
+\end{figure}
+
 \subsection{Low density}
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/inner-12-10}
+  \caption{Density and contact densities at a xxx with bulk filling
+    fraction of 0.1.}
+  \label{fig:inner-12-10}
+\end{figure}
+
 The $g_\sigma^A$ correlation function in each configuration is very
@@
 underestimates the correlation at the interface by
-\textcolor{red}{several percent}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-10}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-4-10}
-\end{figure}
+\textcolor{red}{several percent}.  We note that the theoretical curves
+extend into the region from which the fluid is excluded.  This value
+corresponds to the correlation function that would be observed in the
+vanishingly unlikely scenario in which there was a sphere present at
+that location.  Naturally, we are unable to observe this quantity in
+our Monte Carlo simulations.
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-10}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-12-10}
-\end{figure}
-
-\begin{figure}
   \includegraphics[width=\columnwidth]{figs/outer-16-01}
-  \caption{Density and contact densities at a xxx with bulk filling
+  \caption{Density and correlation functions near the surface of a
+    spherical cavity with diameter $16\sigma$ at a bulk filling
     fraction of 0.1.}
@@
 
+At a higher density corresponding to a filling fraction of 0.4,
+correlations are much stronger, with the bulk contact value of the
+correlation function of 3.7, as seen in
+Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.  This results in larger
+oscillations in the density at the hard surfaces, and correspondingly
+more interesting behavior in the correlation function near the
+interface.  The density predicted by the White Bear functional agrees
+reasonably well with the simulation results, although not so well as
+it did at lower density.  The discrepancies are largest in the case of
+the spherical cavity, in which the DFT considerably underestimates the
+range of the density oscillations.
+
 \begin{figure}
   \includegraphics[width=\columnwidth]{figs/outer-16-04}
-  \caption{Density and contact densities at a xxx with bulk filling
+  \caption{Density and correlation functions near the surface of a
+    spherical cavity with diameter $16\sigma$ at a bulk filling
     fraction of 0.4.}
@@
 \end{figure}
+
+The asymmetric version of the correlation function once again displays
+relatively smooth behavior near the interface, with a few small
+oscillations near the interface, and a somewhat elevated value within
+a diameter of the hard surface, with the magnitude of this elevation
+somewhat different in each configuration.  As was the case at low
+density, our correlation function $g_\sigma^A$ matches very closely
+the Monte Carlo data, reproducing quite well the structure near the
+interface in each configuration, although in the spherical cavity
+(Figure~\ref{fig:outer-40}), there is a small, but significant
+discrepancy, comparable to the discrepancy found in the density
+itself.  In each case, the correlation of Gross dramatically
+underestimates the correlation at the interface, at one extreme by
+\fixme{40\%} in the case of the spherical cavity
+(Figure~\ref{fig:outer-40}), and at the other extreme by \fixme{10\%}
+in the test-particle scenario (Figure~\ref{fig:inner-4-40}).
+
+The symmetrically averaged correlation function shows considerably
+more structure near the interface at high density, and this structure
+varies considerably depending on the curvature of the hard surface.
+In each case, this structure is not reflected in the theoretical
+predictions, neither that of this paper, nor that of Yu and
+Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  As was the case at
+low density, both functionals give significant and finite values in
+the region in which there are no contacts, but at high density they
+also miss the large oscillations that are present near the flat wall
+and the concave surface (Figures~\ref{fig:walls-40}
+and~\ref{fig:outer-40}).  As was the case at low density, the
+functional of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
+gives slightly better agreement with the simulation results than that
+which we derive in Section~\ref{sec:g-S}.
 

### 2012-08-29 papers/contact/paper.tex
a few more minor edits to contact paper Tested everything on node08.

@@
 
-Although the general density attributes look similar to those of the low density calculations as one moves away from the surfaces, the range of difference is altered.
-The spikes at the surfaces reach a height of roughly five times the homogeneous average, whereas with the low densities the difference is closer to a factor of $0.5$.
-With a higher density of spheres to pack in those that are up against the wall, it is more difficult for those spheres to ever find their way out and away.
-We also notice that the downward trend in both $n_0$ and $n_A$ have similar characteristics but are over all much more flat.
-(Don't see why!)
-
-We once again see that in the $S$ case, although neither functionals
-tested fit perfectly, Yu and Wu's functional is a closer fit than the
-White Bear's.  We can see in the monte carlo's oscillations over
-increased distance from the surfaces that there are layers of spheres
-that are formed.  If one picks a point in one of these layers, that
-point is more likely to be a contact point between spheres, whereas if
-one picks a sphere a little over half a radius away, the likelyhood is
-at a minimum.  There is then another maximum at roughly one radius
-away from the surface.  (This suggests close packed layers?)
-
-As with the lower densities, the $A$ case White Bear fits to the monte
-carlo simulations very well.  The Gross functional seems to be
-accurate until it falls off close to surfaces.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 correlation function for inhomogeneous fluid distributions
-corresponding to flat walls, and both concave and convex walls.  We
+corresponding to flat, concave, and convex walls.  We
 defined and simulated two averages of the correlation function, an
@@
 from Fundamental Measure Theory in Section~\ref{sec:g-A} gives
-excellent results for each surface, and at both high density and low
+excellent results for each surface, at both high density and low
 density.  The other three approximations that we studied all showed

### 2012-08-29 papers/contact/paper.tex
more changes to paper Tests skipped!

--- 2012-08-29 papers/contact/paper.tex
@@
 \begin{abstract}
-  We investigate the value of the correlation function an
-  inhomogeneous hard-sphere fluid at contact $g_\sigma$.  This
+  We investigate the value of the correlation function of an
+  inhomogeneous hard-sphere fluid at contact.  This
   quantity plays a critical role in Statistical Associating Fluid
@@
 these definitions, we will present a thermodynamic approach for
-evaluating these these correlation functions directly from the free
+evaluating these correlation functions directly from the free
 energy functional itself.  We will then explain how the correlation
@@
 four approximations with exact Monte-Carlo simulations for the
-hard-sphere fluid in at a variety of hard-wall surfaces.
+hard-sphere fluid at a variety of hard-wall surfaces.
 
@@
 In a homogeneous fluid, the pair correlation only depends on the
-distance $|\rr_1-\rr_2|$, and can be expressed as a function of a
-single variable, with the contact value being when that distance is
-the diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
+distance $|\rr_1-\rr_2|$ and can be expressed as a function of a
+single variable. The contact value is this correlation evaluated at a distance of the
+diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
 functionals to one-center convolutions, which leads us to seek a
@@
   + \rr')
-  \frac{\delta(|\rr'| -\sigma/2)}{\pi\sigma^2}d\rr' \label{eq:gS}
+  \frac{\delta(\sigma/2 -|\rr'|)}{\pi\sigma^2}d\rr' \label{eq:gS}
 \end{align}
@@
 \begin{align}
-  n_0(\rr) &= \int n(\rr')\frac{\delta(|\rr-\rr'|-\sigma/2)}{\pi\sigma^2} d\rr'
+  n_0(\rr) &= \int n(\rr')\frac{\delta(\sigma/2 -|\rr-\rr'|)}{\pi\sigma^2} d\rr'
 \end{align}
@@
   \int n^{(2)}(\rr, \rr + \rr')
-  \frac{\delta(|\rr'| - \sigma)}{4\pi\sigma^2}d\rr' \label{eq:gA}
+  \frac{\delta(\sigma/2 -|\rr'|)}{4\pi\sigma^2}d\rr' \label{eq:gA}
 \end{align}
 where the density $n_A(\rr)$ is similar to $n_0$, but measures the
-density of spheres that could be touching a sphere that is located at
+density of spheres that are touching a sphere that is located at
 point $\rr$.
@@
   n_A(\rr) &= \int n(\rr')
-  \frac{\delta(|\rr-\rr'|-\sigma)}{4\pi\sigma^2} d\rr' \label{eq:nA}
+  \frac{\delta(\sigma/2 -|\rr-\rr'|)}{4\pi\sigma^2} d\rr' \label{eq:nA}
 \end{align}
@@
 \begin{align}
-  n_3(\rr) &= \int n(\rr') \Theta(\left|\rr - \rr'\right| - \sigma/2)
+  n_3(\rr) &= \int n(\rr') \Theta(\sigma/2 -\left|\rr - \rr'\right|)
   d\rr' \label{eq:FMn3} \\
-  n_2(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) d\rr' \\
-  \mathbf{n}_{2V}(\rr) &= \int n(\rr') \delta(\left|\rr - \rr'\right| - \sigma/2) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
+  n_2(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) d\rr' \\
+  \mathbf{n}_{2V}(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
 \end{align}
@@
 hard-sphere diameter as a function of the location of the sphere in
-question, when defining the fundamental measured (in
-Equations~\ref{eq:FMn3}-\ref{eq:FMrest}):
-\begin{align}
-  n_3^A(\rr) &= \int n(\rr')\Theta(|\rr-\rr'|-\sigma(\rr')/2)d\rr' \label{eq:n3A}
-\end{align}
-Once we have done this, we derive a formula for the correlation
+question when defining the fundamental measures.  Once we have done this, we derive a formula for the correlation
 function $g_\sigma^A(\rr)$ by finding the pressure on the surface of
@@
 (filling fraction 0.4) simulation.  We performed additional
-computations at over a wider range of curvatures and densities, but
+computations over a wider range of curvatures and densities, but
 chose these as typical examples.
@@
 the contact value of the correlation function in the bulk is only 1.3,
-indicating that correlations are indeed small, the fluid should be
+indicating that correlations are indeed small and that the fluid should be
 relatively easy to model.  Indeed, the contact density at the hard
@@
 The asymmetric version of the correlation function once again displays
-relatively smooth behavior near the interface, with a few small
+relatively smooth behavior with a few small
 oscillations near the interface, and a somewhat elevated value within

### 2012-08-29 papers/contact/paper.tex
edit "S" section of theory section Tested everything on node08.

--- 2012-08-29 papers/contact/paper.tex
@@
 where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
-dimensionless free energy density.  This requires that we evaluate the
-derivatives of the fundamental measures $n_\alpha(\rr)$ with respect
-to diameter.  These derivatives, when performed exact, lead to
-derivatives of the $\delta$~function, resulting in a functional gives
-poor results.  We find that our $g_\sigma^S$ is considerably improved
-by using integration by parts to eliminate derivatives of the
-$\delta$~function by approximating the density to be smoothly varying,
-which leads to derivatives such as:
-\begin{align}
-  \frac{\partial n_2(\rr)}{\partial \sigma} &\approx \frac{2}{\sigma}n_2(\rr)
-\end{align}
-We note that the density at a hard surface is \emph{not} smoothly
-varying, but that the removal of these terms improves the functional,
-while at the same time reducing its computational cost by avoiding the
-need to calculate any additional weighted densities or convolutions.
-
-The final formula for the symmetrically averaged correlation function
-under these approximations is:
+dimensionless free energy density.  This step is an
+approximation---unlike the analogous
+Equation~\ref{eq:g-A-exact}---because it assumes that we have
+available a local functional $\Phi(\rr)$ whose derivative provides the
+pressure needed to compute $g_\sigma(\rr)$.  Equation~\ref{eq:g-S}
+requires that we evaluate the derivatives of the fundamental measures
+$n_\alpha(\rr)$ with respect to diameter, which leads us to
+derivatives of the $\delta$~function, which we can simplify and
+approximate using an assumption of a reasonably smooth density:
+\begin{align}
+  \frac{\partial n_2(\rr)}{\partial \sigma}
+  &= \int \delta'(\sigma - |\rr-\rr'|) n(\rr')d\rr' \\
+  &= \frac2{\sigma}n_2(\rr) - \int \delta\left(\frac\sigma2 - |\rr-\rr'|\right)
+  \frac{\rr-\rr'}{|\rr-\rr'|}\cdot\mathbf{\nabla}n(\rr')d\rr' \\
+  &\approx \frac{2}{\sigma}n_2(\rr)
+\end{align}
+In the systems that we study, the density is \emph{not} reasonably
+smooth, but we can state empirically making this approximation
+nevertheless improves the predictions of our functional, while at the
+same time reducing its computational cost by avoiding the need to
+calculate any additional weighted densities or convolutions.
+
+The final formula for our approximation to the symmetrically averaged
+correlation function under these approximations is:
 \begin{align}
@@
 \end{align}
+We note that there is a flaw in this functional, which tracks back to
+the assumption in Equation~\ref{eq:g-S} that the pressure due to
+spheres touching at point~$\rr$ can be written as a derivative of
+$\Phi(\rr)$.  This assumption cannot be valid in regions of space
+(such as within a hard wall) where there are no contacts, but the
+local energy density, as defined within FMT, is nonzero.  As we shall
+see in Section~\ref{sec:comparison}, this leads to a significant and
+systematic error.
 
@@
 
-\section{Comparison with simulation}
+\section{Comparison with simulation}\label{sec:simulation}
 

### 2012-08-29 papers/contact/paper.tex
another edit Tests skipped!

@@
 \end{align}
-The functional derivative $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ in the above formula is computed with the
-asymmetric formulation of the fundamental measures, as in
-Equation~\ref{eq:n3A}.  Details of some aspects of finding this
-derivative using FMT are given in Appendix~\ref{appendix:g-A}.  The
-equation for $g_\sigma^A$ involves convolutions of local derivatives
-of the free energy, making this formulation somewhat computationally
-more expensive than the free energy itself.
+Details of finding the functional derivative in
+Equation~\ref{eq:g-A-exact} using FMT are given in
+Appendix~\ref{appendix:g-A}.  The equation for $g_\sigma^A$ involves
+convolutions of local derivatives of the free energy, making this
+formulation somewhat computationally more expensive than the free
+energy itself.
 

### 2012-08-29 papers/contact/paper.tex
fix a bad reference Tests skipped!
major edit of beginning of the paper Tested everything on node09.

@@
   quantity plays a critical role in Statistical Associating Fluid
-  Theory (SAFT), which is the basis of a number of recent classical
+  Theory (SAFT), which is the basis of a number of recently developed classical
   density functionals.  We define two averaged values for the
@@
   functional~\cite{roth2002whitebear}, based on an assumption of
-  thermodynamic consistency. We have tested these functionals and two
-  existing functionals~\cite{yu2002fmt-dft-inhomogeneous-associating,
-    gross2009density} against Monte Carlo simulations.  We find
-  excellent agreement between the Monte Carlo data and our
-  asymmetrically averaged correlation function, and less impressive
-  agreement with the other three correlation functionals.
+  thermodynamic consistency. We test these formulas, as well as two
+  existing formulas~\cite{yu2002fmt-dft-inhomogeneous-associating,
+    gross2009density} against Monte Carlo simulations, and find
+  excellent agreement between the Monte Carlo data and one of our
+  averaged correlation functions.
 \end{abstract}
@@
 A key input to SAFT-based functionals is the correlation function
-evaluated at contact.  Although two recent works have introduced
-approximate functionals for this
-quantity\cite{yu2002fmt-dft-inhomogeneous-associating,
-  gross2009density}, there has yet to be a careful study of the
-correlation function at contact in the inhomogeneous hard-sphere
-fluid.
-
-Constructing a classical density functional based on SAFT involves
-rewriting each term in the free energy as a functional of an
-inhomogeneous density.  The SAFT free energy consists of four terms:
-\begin{align}
-  A_\textit{SAFT} &= A_\textit{ideal} + A_\textit{HS} + A_\textit{chain} + A_\textit{disp} + A_\textit{assoc}
-\end{align}
-where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
-is the excess free energy of a hard-sphere fluid, $A_\textit{chain}$
-is the free energy of formation for polymeric fluids,
-$A_\textit{disp}$ is the dispersion contribution to the free energy,
-and $A_\textit{assoc}$ is the free energy contribution from
-\emph{association}, which is to say, hydrogen bonding.  The chain and
-association contributions to the SAFT free energy are of particular
-interest for this paper, since each requires as input the contact
-value of the correlation function.
+evaluated at contact.
+
+%% Constructing a classical density functional based on SAFT involves
+%% rewriting each term in the free energy as a functional of an
+%% inhomogeneous density.  The SAFT free energy consists of four terms:
+%% \begin{align}
+%%   A_\textit{SAFT} &= A_\textit{ideal} + A_\textit{HS} + A_\textit{chain} + A_\textit{disp} + A_\textit{assoc}
+%% \end{align}
+%% where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
+%% is the excess free energy of a hard-sphere fluid, $A_\textit{chain}$
+%% is the free energy of formation for polymeric fluids,
+%% $A_\textit{disp}$ is the dispersion contribution to the free energy,
+%% and $A_\textit{assoc}$ is the free energy contribution from
+%% \emph{association}, which is to say, hydrogen bonding.
+
+The chain and association terms in the SAFT free energy are of
+particular interest for this paper, since each requires as input the
+contact value of the correlation function.  The chain term describes
+the chain formation energy in polymeric fluids, while the association
+term describes the effects of hydrogen bonding, both of which can be
+large.
 %
@@
 the free energy, which included a functional for the contact value of
-the correlation function (see
+the correlation function (described in
 Section~\ref{sec:yuwu})\cite{yu2002fmt-dft-inhomogeneous-associating},
@@
 introduced\cite{bryk2006density, gross2009density}, one which uses the
-contact value of the correlation function of Yu and
-Wu\cite{bryk2006density}, while the other introduces a new functional
-for the contact value of the correlation
-function (see Section~\ref{sec:gross})\cite{gross2009density}.
+correlation function of Yu and Wu\cite{bryk2006density}, while the
+other introduces a new approximation for the contact value of the
+correlation function (described in
+Section~\ref{sec:gross})\cite{gross2009density}.
 %% The dispersion term is long-range and thus has significant
@@
 
+Although these recent works have introduced approximate functionals
+for the contact value of the correlation
+function\cite{yu2002fmt-dft-inhomogeneous-associating,
+  gross2009density}, there has not been a study that specifically
+addresses the correlation function at contact of an inhomogeneous
+hard-sphere fluid.
+%
 In this paper we introduce two definitions for the locally averaged
-correlation function of an inhomogeneous system.  The first is more
-symmetric (which we call the $S$ case), while the second involves a
-sphere-centered, assymetrical interpretation (the $A$ case).  Given
-these definitions, we will present a thermodynamic approach for
-evaluating these correlation functions directly from the free
-energy functional itself.  We will then explain how the correlation
-function of Yu and Wu and that of Gross compare with our
-definitions---the former is a variant of the $S$ correlation function,
-while the latter is an $A$ function.  Finally, we will compare these
-four approximations with exact Monte-Carlo simulations for the
-hard-sphere fluid at a variety of hard-wall surfaces.
+correlation function of an inhomogeneous system.
+%
+%% The first is more symmetric (which we call the $S$ case), while the
+%% second involves a sphere-centered, assymetrical interpretation (the
+%% $A$ case).
+%
+Given these definitions, we will present a thermodynamic derivation
+for each correlation functions from the free energy functional.  We
+will then discuss the correlation functions of Yu and Wu of Gross, and
+will end by comparing all four approximations with Monte-Carlo
+simulations of the hard-sphere fluid at a variety of hard-wall
+surfaces.
 
@@
 
-We will define our terms by making use of the two-particle density
+We define our terms using the two-particle density
 $n^{(2)}(\rr_1,\rr_2)$, which gives the probability per unit volume
@@
   \int n^{(2)}(\rr, \rr + \rr')
-  \frac{\delta(\sigma/2 -|\rr'|)}{4\pi\sigma^2}d\rr' \label{eq:gA}
+  \frac{\delta(\sigma -|\rr'|)}{4\pi\sigma^2}d\rr' \label{eq:gA}
 \end{align}
@@
   n_A(\rr) &= \int n(\rr')
-  \frac{\delta(\sigma/2 -|\rr-\rr'|)}{4\pi\sigma^2} d\rr' \label{eq:nA}
+  \frac{\delta(\sigma -|\rr-\rr'|)}{4\pi\sigma^2} d\rr' \label{eq:nA}
 \end{align}
@@
 
-\subsection{Fundamental-Measure Theory}
+\subsection*{Fundamental-Measure Theory}
 
 We use the White Bear version of the Fundamental-Measure Theory~(FMT)
-functional published in reference~\cite{roth2002whitebear}.  The FMT
-functional describes the excess free energy of a hard-sphere fluid.
-This particular FMT reduces to the Carnahan-Starling equation of state
-for homogeneous systems.
+functional~\cite{roth2002whitebear}, which describes the excess free
+energy of a hard-sphere fluid.  The White Bear functional reduces to
+the Carnahan-Starling equation of state for homogeneous systems.  It
+is written as an integral over all space of a local function of a set
+of ``fundamental measures'' $n_\alpha(\rr)$, each of which is written
+as a one-center convolution of the density.  The White Bear free
+energy is thus
 \begin{equation}
@@
 \end{align}
-using the weighted densities
+using the fundamental measures
 \begin{align}
@@
 
-In order to motivate the derivation of our functional for the value of
-the correlation function at contact, we begin by deriving the
-well-known value of $g_\sigma$ for a homogeneous hard-sphere fluid
-from the Carnahan-Starling free energy.  The contact value of the
-correlation function density is most easily found by using the
-contact-value theorem, which states that the pressure on any hard
-surface is determined by the density at contact:
+In order to motivate our derivation of the correlation function at
+contact for the \emph{inhomogeneous} hard-sphere fluid, we begin by
+deriving the well-known formula for $g_\sigma$ for the
+\emph{homogeneous} fluid that comes from the Carnahan-Starling free
+energy.  The contact value of the correlation function density can be
+found by using the contact-value theorem, which states that the
+pressure on any hard surface is determined by the density at contact:
 \begin{align}
@@
 \end{align}
-where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.  We
-can thus readily compute the derivative of free energy with respect to
-hard-sphere diameter:
+where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.
+We can thus readily compute the derivative of free energy with respect
+to hard-sphere diameter:
 \begin{align}
@@
 \begin{align}
-  p_{HS} &= \frac{1}{N 8\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
+  p_\sigma &= \frac{1}{N 8\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
   &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
@@
 \begin{align}
-  g_{HS}(\sigma) &= \frac{1 - \frac{\eta}2}{(1-\eta)^3} \label{eq:cs-g}
-\end{align}
-Thus we can readily derive the contact density for the homogeneous
-hard-sphere fluid.  Extending this derivation to the inhomogeneous
-fluid requires that we find the pressure on particular spheres---or
-potentially particular parts of particular spheres.
+  g_\sigma &= \frac{1 - \frac{\eta}2}{(1-\eta)^3} \label{eq:cs-g}
+\end{align}
+Extending this derivation to the inhomogeneous fluid requires that we
+find the pressure felt by the surface of particular spheres.
 
@@
 of a sphere located at $\rr$, and can thus be determined by finding
-the pressure on that sphere.  We find this pressure by treating the
-hard-sphere diameter as a function of the location of the sphere in
-question when defining the fundamental measures.  Once we have done this, we derive a formula for the correlation
-function $g_\sigma^A(\rr)$ by finding the pressure on the surface of
-any hard spheres located at position $\rr$, and using the
-contact-density theorem to extract the contact density, from which we
-can establish the averaged correlation function at contact.
-\begin{align}
-  p_{HS}(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
+the pressure on that sphere.  We find this pressure from the change in
+free energy changes resulting from an infinitesimal expansion of any
+spheres located at position~$\rr$.  From this pressure, we derive a
+formula for the correlation function $g_\sigma^A(\rr)$ as was done in
+the previous section:
+\begin{align}
+  p_\sigma(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
@@
 \end{align}
-Details of finding the functional derivative in
+Details of evaluating the functional derivative in
 Equation~\ref{eq:g-A-exact} using FMT are given in
-Appendix~\ref{appendix:g-A}.  The equation for $g_\sigma^A$ involves
-convolutions of local derivatives of the free energy, making this
-formulation somewhat computationally more expensive than the free
+Appendix~\ref{appendix:g-A}.  The equation for $g_\sigma^A$ requires
+finding convolutions of local derivatives of the free energy, making
+this formulation computationally somewhat more expensive than the free
 energy itself.
@@
 
-\section{Comparison with simulation}\label{sec:simulation}
+\section{Comparison with simulation}\label{sec:comparison}
 

### 2012-08-30 papers/contact/paper.tex
improve figure captions and change to diameter units (partly) Tests skipped!

@@
   \includegraphics[width=\columnwidth]{figs/walls-10}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.1.}
+  \caption{Density and correlation function at a hard wall with bulk
+    filling fraction of 0.1.}
   \label{fig:walls-10}
@@
   \includegraphics[width=\columnwidth]{figs/inner-4-10}
-  \caption{Density and contact densities at a xxx with bulk filling
+  \caption{Density and correlation function around a hard sphere with
+    diameter the same as those in the fluid, with a bulk filling
     fraction of 0.1.}
@@
   \includegraphics[width=\columnwidth]{figs/inner-12-10}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.1.}
+  \caption{Density and correlation function around a hard sphere with
+    an excluded diameter of $6\sigma$, with a bulk filling fraction of
+    0.1.}
   \label{fig:inner-12-10}
@@
   \includegraphics[width=\columnwidth]{figs/walls-40}
-  \caption{Density and contact densities at a wall with bulk filling
-    fraction of 0.4.}
+  \caption{Density and correlation function at a hard wall with bulk
+    filling fraction of 0.4.}
   \label{fig:walls-40}
@@
   \includegraphics[width=\columnwidth]{figs/inner-4-40}
-  \caption{Density and contact densities at a xxx with bulk filling
+  \caption{Density and correlation function around a hard sphere with
+    diameter the same as those in the fluid, with a bulk filling
     fraction of 0.4.}
@@
   \includegraphics[width=\columnwidth]{figs/inner-12-40}
-  \caption{Density and contact densities at a xxx with bulk filling
-    fraction of 0.4.}
+  \caption{Density and correlation function around a hard sphere with
+    an excluded diameter of $6\sigma$, with a bulk filling fraction of
+    0.4.}
   \label{fig:inner-12-40}

### 2012-08-30 papers/contact/paper.tex
small edits on paper Tested everything on node08.

--- 2012-08-29 papers/contact/paper.tex
@@
   \frac{\partial n_2(\rr)}{\partial \sigma}
-  &= \int \delta'(\sigma - |\rr-\rr'|) n(\rr')d\rr' \\
-  &= \frac2{\sigma}n_2(\rr) - \int \delta\left(\frac\sigma2 - |\rr-\rr'|\right)
+  &= \frac12 \int \delta'(\frac\sigma2 - |\rr-\rr'|) n(\rr')d\rr' \\
+  &= \frac2{\sigma}n_2(\rr) - \frac12 \int \delta\left(\frac\sigma2 - |\rr-\rr'|\right)
   \frac{\rr-\rr'}{|\rr-\rr'|}\cdot\mathbf{\nabla}n(\rr')d\rr' \\
@@
 calculate any additional weighted densities or convolutions.
-
-The final formula for our approximation to the symmetrically averaged
-correlation function under these approximations is:
-\begin{align}
-  g_\sigma^S &= \text{\textcolor{red}{put formula in here, if it is
-      short enough}}
-\end{align}
-We note that there is a flaw in this functional, which tracks back to
-the assumption in Equation~\ref{eq:g-S} that the pressure due to
-spheres touching at point~$\rr$ can be written as a derivative of
-$\Phi(\rr)$.  This assumption cannot be valid in regions of space
-(such as within a hard wall) where there are no contacts, but the
-local energy density, as defined within FMT, is nonzero.  As we shall
-see in Section~\ref{sec:comparison}, this leads to a significant and
-systematic error.
-
-\subsection{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
-One approximation for the correlation function is that of
-Gross\cite{gross2009density}, which is of the asymmetrically averaged
-variety ($g_\sigma^A$):
-\begin{align}
-  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{\pi}{12}\sigma^3n_A(\rr)}{\left(1 -
-    \frac{\pi}{6}\sigma^3n_A(\rr)\right)^3}
-\end{align}
-where $n_A$ is the averaged density defined in Equation~\ref{eq:nA}.
-This formula is arrived at by using the density averaged over all
-spheres that could be touching a sphere at point~$\rr$ in the
-Carnahan-Starling equation for the correlation function at contact,
-given in Equation~\ref{eq:cs-g}.
 

### 2012-08-30 papers/contact/paper.tex
yet a few more minor edits to contact paper Tests skipped!

@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-16-04}
-  \caption{Density and correlation functions near the surface of a
-    spherical cavity with diameter $16\sigma$ at a bulk filling
-    fraction of 0.4.}
-  \label{fig:outer-40}
-\end{figure}
-
 The asymmetric version of the correlation function once again displays
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/outer-16-04}
+  \caption{Density and correlation functions near the surface of a
+    spherical cavity with diameter $16\sigma$ at a bulk filling
+    fraction of 0.4.}
+  \label{fig:outer-40}
+\end{figure}
+
 The symmetrically averaged correlation function shows considerably
@@
 
-Here we discuss the high-density results (filling fraction of 0.4),
-which are shown in Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.
-
-Although the general density attributes look similar to those of the low density calculations as one moves away from the surfaces, the range of difference is altered.
-The spikes at the surfaces reach a height of roughly five times the homogeneous average, whereas with the low densities the difference is closer to a factor of $0.5$.
-With a higher density of spheres to pack in those that are up against the wall, it is more difficult for those spheres to ever find their way out and away.
-We also notice that the downward trend in both $n_0$ and $n_A$ have similar characteristics but are over all much more flat.
-(Don't see why!)
-
-We once again see that in the $S$ case, although neither functionals
-tested fit perfectly, Yu and Wu's functional is a closer fit than the
-White Bear's.  We can see in the monte carlo's oscillations over
-increased distance from the surfaces that there are layers of spheres
-that are formed.  If one picks a point in one of these layers, that
-point is more likely to be a contact point between spheres, whereas if
-one picks a sphere a little over half a radius away, the likelyhood is
-at a minimum.  There is then another maximum at roughly one radius
-away from the surface.  (This suggests close packed layers?)
-
-As with the lower densities, the $A$ case White Bear fits to the monte
-carlo simulations very well.  The Gross functional seems to be
-accurate until it falls off close to surfaces.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2012-08-30 papers/contact/paper.tex
add Gross section back in Tests skipped!

@@
 
+\subsection{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
+One approximation for the correlation function is that of
+Gross\cite{gross2009density}, which is of the asymmetrically averaged
+variety ($g_\sigma^A$):
+\begin{align}
+  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{\pi}{12}\sigma^3n_A(\rr)}{\left(1 -
+    \frac{\pi}{6}\sigma^3n_A(\rr)\right)^3}
+\end{align}
+where $n_A$ is the averaged density defined in Equation~\ref{eq:nA}.
+This formula is arrived at by using the density averaged over all
+spheres that could be touching a sphere at point~$\rr$ in the
+Carnahan-Starling equation for the correlation function at contact,
+given in Equation~\ref{eq:cs-g}.
+
 \subsection{Yu and Wu's symmetrically averaged functional}\label{sec:yuwu}

### 2012-08-30 papers/contact/paper.tex
more changes, appendix and some in beginning Tests skipped!

@@
   them from the White Bear version of the Fundamental Measure Theory
-  functional~\cite{roth2002whitebear}, based on an assumption of
+  functional~\cite{roth2002whitebear}, using an assumption of
   thermodynamic consistency. We test these formulas, as well as two
@@
 %
-A key input to SAFT-based functionals is the correlation function
-evaluated at contact.
-
-%% Constructing a classical density functional based on SAFT involves
-%% rewriting each term in the free energy as a functional of an
-%% inhomogeneous density.  The SAFT free energy consists of four terms:
-%% \begin{align}
-%%   A_\textit{SAFT} &= A_\textit{ideal} + A_\textit{HS} + A_\textit{chain} + A_\textit{disp} + A_\textit{assoc}
-%% \end{align}
-%% where $A_\textit{ideal}$ is the ideal-gas free energy, $A_\textit{HS}$
-%% is the excess free energy of a hard-sphere fluid, $A_\textit{chain}$
-%% is the free energy of formation for polymeric fluids,
-%% $A_\textit{disp}$ is the dispersion contribution to the free energy,
-%% and $A_\textit{assoc}$ is the free energy contribution from
-%% \emph{association}, which is to say, hydrogen bonding.
-
-The chain and association terms in the SAFT free energy are of
-particular interest for this paper, since each requires as input the
-contact value of the correlation function.  The chain term describes
+A key input to these functionals is the correlation function
+evaluated at contact, which is required for both the chain and
+association terms in the SAFT free energy.  The chain term describes
 the chain formation energy in polymeric fluids, while the association
@@
   gross2009density}, there has not been a study that specifically
-addresses the correlation function at contact of an inhomogeneous
-hard-sphere fluid.
+addresses this contact value for an inhomogeneous hard-sphere fluid.
 %
@@
 Given these definitions, we will present a thermodynamic derivation
-for each correlation functions from the free energy functional.  We
-will then discuss the correlation functions of Yu and Wu of Gross, and
+for each correlation function from the free energy functional.  We
+will then discuss the correlation functions of Yu and Wu and of Gross, and
 will end by comparing all four approximations with Monte-Carlo
@@
 distance $|\rr_1-\rr_2|$ and can be expressed as a function of a
-single variable. The contact value is this correlation evaluated at a distance of the
+single variable. The contact value of the correlation function is this
+correlation function's value when evaluated at a distance of the
 diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
@@
 \end{align}
-where density $n_0$ is one of the fundamental measures of Fundamental
+where $\sigma$ is the hard sphere diameter and density $n_0$ is one of the fundamental measures of Fundamental
 Measure Theory (FMT), and is ideal for treating touching spheres, as
@@
 \includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that included in $n_0(\mathbf{r})$, which
+\caption{Set of hard spheres that are included in $n_0(\mathbf{r})$, which
   consist of those which just touch the point $\mathbf{r}$.}
@@
 function is related to the contact density averaged over the surface
-of a sphere located at $\rr$, and can thus be determined by finding
+of a sphere located at~$\rr$, and can thus be determined by finding
 the pressure on that sphere.  We find this pressure from the change in
-free energy changes resulting from an infinitesimal expansion of any
+free energy resulting from an infinitesimal expansion of any
 spheres located at position~$\rr$.  From this pressure, we derive a
@@
 
-\section*{Introduction to Appendices}
-
-In our expressions for pressure (eq \ref{eq:p_{HS}^A} and eq
-\ref{eq:p_{HS}^S}) we have $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ terms that must be derived for both the $S$ and A
-cases.  We break these terms up as follows:
+\section*{Appendix}
+
+In the expression for our A correlation function $g_\sigma^A(\rr)$ (eq \ref{eq:g-A-exact}) we have $\frac{\delta A_{HS}}{\delta
+  \sigma(\mathbf{r})}$ terms that must be derived.  We break these terms up as follows:
 
@@
 
-We must therefore compute both the $\frac{\delta A_{HS}}{\delta
-  n(\mathbf{r}')}$, etc., terms and the $\frac{\delta
-  n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc., terms.  We will
-compute the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms
-for both the functional derived from the White Bear
-functional\cite{roth2002whitebear} and from the White Bear mark
-II functional\cite{hansen2006density}.  Following this we will compute
-the $\frac{\delta n(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc.,
-terms for both the $A$ and $S$ cases.  These latter derivitives will
-differ for the $A$ and $S$ cases according to the qualitive concepts
-described above.
-
-\subsection{Expressions for $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc. for the assymetric case}\label{appendix:g-A}
-
-The derivatation of the asymmetrically averaged correlation function
-$g_\sigma^A(\rr)$ requires that we express the free energy in terms of
-a set of weighted densities in which the radius of the weighting
-kernel is determined according to the position of the sphere in
-question---which is to say, the location where the density itself is
-evaluated.  This is the most natural way to construct such a
-functional, in which the radius of spheres is a function of position.
-Thus we define $n_3^A$ as
-\begin{align}
-  n_3^{A}(\rr') &= \int n(\rr'') \Theta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
-\end{align}
-and we define the other weighted densities similarly.  In order to
-compute derivatives of the free energy with respect to $R(\rr)$, we
-need to find $g_\sigma^A$.  In the case of $n_3^A$, we need
+We will show here the derivation of the $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc., terms for the assymetric case.
+The derivation of the S correlation follows the same format using the symmetric conceptual approach introduced in \ref{sec:g-S} and the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms are straightforward (if tedious) derivatives.
+
+The derivation of $g_\sigma^A(\rr)$ requires that we express our weighted
+densities so that the radius of the weighting kernel is determined according
+to the position of the sphere in question, the location where the density
+itself is evaluated.  Thus for the derivative of $n_3^A$, we have
 \begin{align}
@@
 \end{align}
-This is somewhat more challenging, since it involves an unfamiliar
-function, the derivative the Dirac delta function $\delta'$.
-%
+
 The derivatives of the remaining scalar densities can be reduced to
@@
 \end{align}
-and
+
 \begin{align}
@@
 \end{align}
-The vector-weighted densities have no additional occurrences of $R$,
-their functional derivatives come out similarly:
+The vector-weighted densities have no additional occurrences of $\sigma$.  Their functional derivatives come out similarly:
 \begin{align}
@@
 \end{align}
-and analogously for $\mathbf{n}_{V1}^A$:
 \begin{align}
@@
 \end{align}
-This completes our exploration of the partial derivatives of weighted
-densities of the $A$ variety.
 

### 2012-08-30 papers/contact/paper.tex
fixed percentages that Gross was wrong by, at end Tests skipped!

@@
 %% The first is more symmetric (which we call the $S$ case), while the
-%% second involves a sphere-centered, assymetrical interpretation (the
+%% second involves a sphere-centered, asymmetrical interpretation (the
 %% $A$ case).
@@
 underestimates the correlation at the interface, at one extreme by
-\fixme{40\%} in the case of the spherical cavity
-(Figure~\ref{fig:outer-40}), and at the other extreme by \fixme{10\%}
+40\% in the case of the spherical cavity
+(Figure~\ref{fig:outer-40}), and at the other extreme by 15\%
 in the test-particle scenario (Figure~\ref{fig:inner-4-40}).
@@
 
-We will show here the derivation of the $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc., terms for the assymetric case.
+We will show here the derivation of the $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc., terms for the asymmetric case.
 The derivation of the S correlation follows the same format using the symmetric conceptual approach introduced in \ref{sec:g-S} and the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms are straightforward (if tedious) derivatives.

### 2012-08-30 papers/contact/paper.tex
changed one word!!! Tests skipped!

@@
 predictions in each case, while that of Gross consistently
-underestimates the correlation at the interface by
-\textcolor{red}{several percent}.  We note that the theoretical curves
+underestimates the correlation at the interface by a significant margin.  We note that the theoretical curves
 extend into the region from which the fluid is excluded.  This value

### 2012-08-30 papers/contact/figs/plot-outer.py
changed plotting so legends dont interfere and small edit to paper Tested everything on node09.

@@ -59,7 +59,10 @@ n_plt.plot(r/2,n0*4*numpy.pi/3,"c-.",label="$n_0$")
 n_plt.plot(r/2,nA*4*numpy.pi/3,"m--",label="$n_A$")
 #n_plt.plot(r/2,n0*4*numpy.pi/3,"c--",label="$n_0$")
 #n_plt.plot(r/2,nA*4*numpy.pi/3,"m-.",label="$n_A$")
-n_plt.legend(loc='upper left', ncol=2).get_frame().set_alpha(0.5)
+if (n[2]*4*numpy.pi/3 < 0.15 and n[2]*4*numpy.pi/3 > 0.05):
+    n_plt.legend(loc='lower left', ncol=2).get_frame().set_alpha(0.5)
+else:
+    n_plt.legend(loc='upper left', ncol=2).get_frame().set_alpha(0.5)
 n_plt.yaxis.set_major_locator(pylab.MaxNLocator(4,steps=[1,5,10],prune='upper'))
 pylab.ylim(0, n.max()*4*numpy.pi/3*1.1)
 pylab.xlim(showrmin/2,showrmax/2)
@@ -93,7 +96,10 @@ S_plt.axvline(x=radius/2, color='k', linestyle=':')
 S_plt.plot(r_mc/2,gS_mc,"g-",label="$g_\sigma^S$ (MC)")
 S_plt.plot(r[0:sphere_end]/2,gS[0:sphere_end], "go--",markevery=me, label="$g_\sigma^S$ (White Bear)")
 S_plt.plot(r/2,gyuwu,"gx--",markevery=me, label="Yu and Wu")
-S_plt.legend(loc='lower left', ncol=1).get_frame().set_alpha(0.5)
+if (n[2]*4*numpy.pi/3 < 0.15 and n[2]*4*numpy.pi/3 > 0.05):
+    S_plt.legend(loc='lower left', ncol=1).get_frame().set_alpha(0.5)
+else:
+    S_plt.legend(loc='upper left', ncol=1).get_frame().set_alpha(0.5)
 S_plt.yaxis.set_major_locator(pylab.MaxNLocator(integer=True,prune='upper'))
 pylab.xlim(showrmin/2,showrmax/2)
 

### 2012-08-30 papers/contact/paper.tex
changed plotting so legends dont interfere and small edit to paper Tested everything on node09.
start editing appendix of contact paper Tests skipped!

@@
 \author{Jeff Schulte}
-\author{Patrick Kreitzburg}
+\author{Patrick Kreitzberg}
 \author{Chris Haglund}
@@
 Equation~\ref{eq:g-A-exact} using FMT are given in
-Appendix~\ref{appendix:g-A}.  The equation for $g_\sigma^A$ requires
+the appendix.  The equation for $g_\sigma^A$ requires
 finding convolutions of local derivatives of the free energy, making
@@
 
-In the expression for our A correlation function $g_\sigma^A(\rr)$ (eq \ref{eq:g-A-exact}) we have $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$ terms that must be derived.  We break these terms up as follows:
-
+The expression for the asymmetric correlation function
+$g_\sigma^A(\rr)$ (Equation~\ref{eq:g-A-exact}) involves the
+functional derivative $\frac{\delta A_{HS}}{\delta
+  \sigma(\mathbf{r})}$.  In this appendix we will explain how this
+derivative is evaluated.  We begin by applying the chain rule in the
+following way:
   \begin{align}
@@
     \int \left(
-    \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}
-    \frac{\delta n_3(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
-    +
-    \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')}
-    \frac{\delta n_2(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
-    + \cdots
+    \sum_\alpha
+    \frac{\delta A_{HS}}{\delta n_\alpha(\mathbf{r}')}
+    \frac{\delta n_\alpha(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
     \right) d\mathbf{r}'
   \end{align}
+This expression requires us to evaluate $\frac{\delta A_{HS}}{\delta
+  n_\alpha(\mathbf{r}')}$ and $\frac{\delta
+  n_\alpha(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$.  The former is
+straightforward, if tedious, and we will write no more about it.  The
+functional derivatives of the fundamental measures, however, require a
+bit more subtlety, and we will address them here.  We note that
+\textcolor{red}{still working on this...}
 

### 2012-08-30 papers/contact/paper.tex
changed which mc data we use in plots and made some of Dannys suggested changes to paper Tests skipped!

@@
 Two functionals for the chain contribution have recently been
-introduced\cite{bryk2006density, gross2009density}, one which uses the
-correlation function of Yu and Wu\cite{bryk2006density}, while the
-other introduces a new approximation for the contact value of the
+introduced, one which uses the
+correlation function of Yu and Wu\cite{bryk2006density} and another that
+introduces a new approximation for the contact value of the
 correlation function (described in
@@
 Since we are interested in the correlation function at the surface of
-the hard spheres, all we need compute is the pressure on that surface,
-and we'll have our answer.  The pressure on a hard sphere can be
+the hard spheres, we need to compute the pressure on that surface.  The pressure on a hard sphere can be
 readily computed from the dependence of the Carnahan-Starling free
@@
 spheres---since we're changing all their radii at once.  To compute
-the pressure on the spheres, we just need to divide by half the total
+the pressure on the spheres, we need to divide by half the total
 area, which means dividing by $N 8\pi \sigma^2$, since the relevant
@@
 pressure will be simply related to the free energy density at
-point~$\rr$.  Through math similar to the previous derivations, this
+point~$\rr$.  Through a process similar to the previous derivations, this
 leads to the expression
@@
 We simulate the inhomogeneous hard sphere fluid at four hard-wall
-interfaces.  The first and simplest is a simple flat hard wall.  We
+interfaces.  The first and simplest is a flat hard wall.  We
 then study two convex hard surfaces.  One is an excluded sphere with

### 2012-08-31 papers/contact/paper.tex
rewrite appendix in contact paper Tests skipped!

@@
 \begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
+\Phi_1 &= -n_0 \ln\left( 1 - n_3\right) \label{eq:Phi1}\\
 \Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
@@
   36\pi n_3^2\left( 1 - n_3 \right)^2
-} ,
+} , \label{eq:Phi3}
 \end{align}
@@
 
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
 \appendix
@@
 
-\begin{widetext}
-
 \section*{Appendix}
@@
   n_\alpha(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$.  The former is
-straightforward, if tedious, and we will write no more about it.  The
-functional derivatives of the fundamental measures, however, require a
-bit more subtlety, and we will address them here.  We note that
-\textcolor{red}{still working on this...}
+straightforward, given Equations~\ref{eq:Phi1}-\ref{eq:Phi3}, and we
+will write no more about it.  The functional derivatives of the
+fundamental measures, however, require a bit more subtlety, and we
+will address them here.
+
+We begin with the derivative of $n_3$, the filling fraction, which we
+will discuss in somewhat more detail than the remainder, which are
+similar in nature.  Because the diameter $\sigma(\rr)$ is the diameter
+of a sphere \emph{at position~$\rr$}, we write the fundamental measure
+$n_3(\rr')$ as
+\begin{align}
+  n_3(\rr') &= \int n(\rr'') \Theta\left(\frac{\sigma(\rr'')}{2}
+  -\left|\rr' - \rr''\right|\right)
+  d\rr''
+\end{align}
+where we note that the integration variable~$\rr''$ is the position of
+the sphere in question.  Thus the derivative we seek is
+\begin{align}
+  \frac{\delta n_3(\rr')}{\delta \sigma(\rr)} &= \frac 12 \int n
+  (\rr'') \delta\left(\frac{\sigma(\rr'')}{2} -\left|\rr' - \rr''\right|\right)
+  \delta(\rr-\rr'') d\rr'' \\ &= n (\rr) \delta(\sigma(\rr)/2
+  -\left|\rr' - \rr\right|)
+\end{align}
+This pattern will hold for each fundamental measure: because we are
+seeking the change in free energy when spheres at point~$\rr$ are
+expanded, the integral over density is eliminated.  To compute the
+correlation funtion $g_\sigma^A$, we convolve this delta function with
+the product of the density and a local derivative of $\Phi(\rr)$:
+\begin{align}
+  \frac{\delta A_{HS}}{\delta \sigma(\rr)} &= \int \frac{\partial \Phi(\rr')}{\partial
+    n_3(\rr')}n(\rr') \delta(\sigma/2-|\rr'-\rr|)d\rr'
+  + \cdots
+\end{align}
+As we shall see, there are only four convolution kernels, leading to
+four additional convolutions beyond those required for FMT.
+
+The functional derivative of $n_2$ introduces our second convolution
+kernel, which is a derivative of the delta function.
+\begin{align}
+  %n_2(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d \rr''\\
+  \frac{\delta n_2(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+\end{align}
+The derivatives of the remaining scalar densities $n_1$ and $n_0$ reduce to
+sums of the terms above:
+\begin{align}
+  %n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
+  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
+  \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
+  &= \frac{n(\rr)}{4\pi
+    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \\
+  &\quad-
+  \frac{n(\rr)}{2\pi
+    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+\end{align}
+and
+\begin{align}
+  %n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
+  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
+  \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
+  &= \frac{n(\rr)}{2\pi
+    \sigma(\rr)^2}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+  \\&\quad-
+  2 \frac{n(\rr)}{\pi
+    \sigma(\rr)^3}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+\end{align}
+
+The vector-weighted densities $\mathbf{n}_{V1}$ and $\mathbf{n}_{V2}$
+give terms analogous to those of $n_1$ and $n_2$:
+\begin{align}
+  %\mathbf{n}_{V2}(\rr') &= \int n(\rr') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
+  %  \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
+  \frac{\delta \mathbf{n}_{V2}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+    \frac{\rr-\rr'}{|\rr-\rr'|}\\
+\end{align}
+\begin{align}
+  %\mathbf{n}_{V1}(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
+  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
+  \frac{\delta \mathbf{n}_{V1}(\rr')}{\delta \sigma(\rr)}
+  &= -\frac{n(\rr)}{4\pi
+    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
+  \\ &\quad+
+  \frac{n(\rr)}{2\pi
+    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
+\end{align}
+Thus there are four convolution kernels used in computing $g_\sigma^A$
+one scalar and one vector delta function, and one scalar and one
+vector derivative of the delta function.
+
+\bibliography{paper}% Produces the bibliography via BibTeX.
+
+\end{document}
+
+\begin{widetext}
 

### 2012-08-31 papers/contact/paper.tex
more edits to paper, Dannys cahnges, and resize figures Tests skipped!

@@
 We can thus readily compute the derivative of free energy with respect
-to hard-sphere diameter:
-\begin{align}
-  \frac{dA_{HS}}{d\sigma} &= \frac{dA_{HS}}{d\eta} \frac{d\eta}{d\sigma} \\
-  &= Nk_BT \frac{4 - 2\eta}{(1-\eta)^3} \frac{3 \eta}{\sigma} \label{eq:dAhsdR}
-\end{align}
-This derivative gives us half the force on \emph{all} the hard
-spheres---since we're changing all their radii at once.  To compute
-the pressure on the spheres, we need to divide by half the total
-area, which means dividing by $N 8\pi \sigma^2$, since the relevant
-area is the area over which the molecules can make contact.
-\begin{align}
-  p_\sigma &= \frac{1}{N 8\pi \sigma^2} \frac{dA_{HS}}{d\sigma} \\
+to hard-sphere diameter, which is half the force on \emph{all} the
+hard spheres---since we're changing all their diameters at once.  To
+compute the pressure on the spheres, we need to divide the total force
+by the total area of contact, which means dividing by $N 4\pi
+\sigma^2$.
+\begin{align}
+  p_\sigma &= \frac{1}{N 4\pi \sigma^2} \frac12 \frac{dA_{HS}}{d\sigma} \\
   &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
@@
 Yu and Wu developed a functional for the correlation function
-evaluated at contact, which is symmetrically
-averaged~\cite{yu2002fmt-dft-inhomogeneous-associating}.  Although Yu
-and Wu use a symmetric formulation for $g_\sigma$, instead of using
-$n_0$ as the associated density, they use a density given by
-\begin{align}
-  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \\
+evaluated at contact which is symmetrically
+averaged~\cite{yu2002fmt-dft-inhomogeneous-associating}.  However,
+instead of using $n_0$ as the corresponding density, they use a
+density given by
+\begin{align}
+  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \label{eq:nYu} \\
   \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
@@
 where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact, and has the effect of reducing the density at
-interfaces.  Given this definition of density, their $g_\sigma$ is
-given by:
-% \fixme{Which of the following two equations is more clear to you? We
-%  only want one.}
-\begin{align}
-  %% g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-  %%   + \frac32 \frac{\sigma n_2}{6}\frac{\zeta}{(1-n_3)^2}
-  %%   + \frac12 \frac{\left(\frac{\sigma n_2}{6}\right)^2 \zeta}{(1-n_3)^3}
-  %%   \\
-  g_\sigma^\text{Yu} &= \frac{1}{1-n_3}
-    + \frac14 \frac{\sigma n_2\zeta}{(1-n_3)^2}
-    + \frac1{72} \frac{\sigma^2 n_2^2 \zeta}{(1-n_3)^3}
-\end{align}
-From this function, we extract a symmetric correlation function
-\begin{align}
-  g_\sigma^\text{Yu,S} &= \zeta^2\left(\frac{1}{1-n_3}
+point of contact, and has the effect of reducing this density at
+interfaces.  Because of this difference, the correlation function of
+Yu and Wu cannot be directly compared with $g_\sigma^S$ as defined in
+Equation~\ref{eq:gS}.  Therefore in order to make a comparison we move
+the factors of $\zeta$ in Equation~\ref{eq:nYu} from the density into
+the correlation function itself.
+\begin{align}
+  g_\sigma^\text{Yu,S} &= \zeta^2 g_\sigma^\text{Yu} \\
+   &= \zeta^2\left(\frac{1}{1-n_3}
     + \frac14 \frac{\sigma n_2\zeta}{(1-n_3)^2}
@@
 \end{align}
-which corresponds to our $g_\sigma^S$ so that they can be directly
-compared.
+where $g_\sigma^\text{Yu}$ is the correlation function as defined in
+reference~\cite{yu2002fmt-dft-inhomogeneous-associating}, and
+$g_\sigma^\text{Yu,S}$ is the function we will examine in this paper.
 
@@
 \end{align}
-where we note that the integration variable~$\rr''$ is the position of
-the sphere in question.  Thus the derivative we seek is
+where we note that $\sigma(\rr'')$ and $n(\rr'')$ are the diameter and
+density, respectively, of spheres centered at position~$\rr''$.  Thus the
+derivative with respect to the diameter of spheres at position
+$\rr$ is
 \begin{align}
@@
 sums of the terms above:
-\begin{align}
+\begin{multline}
   %n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
@@
   \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{4\pi
+  = \frac{n(\rr)}{4\pi
     \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \\
-  &\quad-
+  -
   \frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
+\end{multline}
 and
-\begin{align}
+\begin{multline}
   %n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
@@
   \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{2\pi
+  = \frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-  \\&\quad-
+  \\-
   2 \frac{n(\rr)}{\pi
     \sigma(\rr)^3}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
+\end{multline}
 
@@
 give terms analogous to those of $n_1$ and $n_2$:
-\begin{align}
+\begin{multline}
   %\mathbf{n}_{V2}(\rr') &= \int n(\rr') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
   %  \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-    \frac{\rr-\rr'}{|\rr-\rr'|}\\
-\end{align}
-\begin{align}
+  \frac{\delta \mathbf{n}_{V2}(\rr')}{\delta \sigma(\rr)} = -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
+    \frac{\rr-\rr'}{|\rr-\rr'|}
+\end{multline}
+\begin{multline}
   %\mathbf{n}_{V1}(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
@@
   \frac{\delta \mathbf{n}_{V1}(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{4\pi
+  = -\frac{n(\rr)}{4\pi
     \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-  \\ &\quad+
+  \\ +
   \frac{n(\rr)}{2\pi
     \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-\end{align}
-Thus there are four convolution kernels used in computing $g_\sigma^A$
+\end{multline}
+Thus there are four convolution kernels used in computing $g_\sigma^A$:
 one scalar and one vector delta function, and one scalar and one

### 2012-08-31 papers/water-SAFT/paper.tex
format water paper for submission Tests skipped!

@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
+\documentclass[a4paper,12pt,amsmath,amssymb,prb]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 \author{Jessica Hughes}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\affiliation{Department of Physics, Oregon State University,
+  Corvallis, OR 97331, USA}
 \author{Eric Krebs}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\affiliation{Department of Physics, Oregon State University,
+  Corvallis, OR 97331, USA}
 \author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\affiliation{Department of Physics, Oregon State University,
+  Corvallis, OR 97331, USA}
 
@@
 White Bear fundamental-measure theory (FMT)
-functional\cite{roth2002whitebear} for the hard sphere fluid with
+functional for the hard sphere fluid with
 attractive interactions based on the Statistical Associating Fluid
-Theory (SAFT-VR)\cite{clark2006developing}.
+Theory (SAFT-VR).
 This functional reproduces the properties of water at both long and
@@
 itself.  We demonstrate our functional by applying it to systems
-composed of of two hard rods, four hard rods arranged in a square and
+composed of two hard rods, four hard rods arranged in a square and
 hard spheres in water.
@@
   A(T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int V_\textit{ext}(\rr) n(\rr)
-d\rr\right\}.
+d\rr\right\},
 \end{equation}
@@
 \begin{equation}\label{eq:contact}
-  p = n_ck_BT
+  p = n_ck_BT,
 \end{equation}
@@
 \begin{equation}\label{contactvaluethm}
-  F = n k_BT V
+  F = n k_BT V.
 \end{equation}
@@
 \begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx
+  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx,
 \end{equation}
@@
 \begin{align}
-  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx
+  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
 \end{align}
@@
   \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
-      \lscale R)^2}}\right)d\xx'
+      \lscale R)^2}}\right)d\xx'.
 \end{align}
@@
   F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) \zeta(\xx)
-  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx,
 \end{align}
@@
   {4 n_0(\xx)\zeta(\xx)
-    \Delta(\xx)}
+    \Delta(\xx)},
 \end{align}
@@
   \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
-  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW}
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
 \end{align}
@@
   +\frac{R}{2}\frac{\zeta n_2}{(1-n_3)^2}
-  + \frac{R^2}{18}\frac{\zeta n_2^2}{(1-n_3)^3}
+  + \frac{R^2}{18}\frac{\zeta n_2^2}{(1-n_3)^3}.
 \end{align}
@@
 \begin{equation}
-F = \gamma A + pV
+F = \gamma A + pV.
 \end{equation}

### 2012-10-26 papers/contact/paper.tex
add arxiv-generating output for contact paper Tests skipped!
changes for referees Tested everything on node06.

@@
 
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp,10pt,aip]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,10pt,pre]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 \end{align}
-In a homogeneous fluid, the pair correlation only depends on the
+The contact value of the correlation function is this
+correlation function's value when evaluated at a distance of the
+diameter $\sigma$.  In a homogeneous fluid, the pair correlation only depends on the
 distance $|\rr_1-\rr_2|$ and can be expressed as a function of a
-single variable. The contact value of the correlation function is this
-correlation function's value when evaluated at a distance of the
-diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
-functionals to one-center convolutions, which leads us to seek a
+single variable. The pair correlations for an inhomogeneous fluid are of course
+not as simple but it is desirable for reasons of computational efficiency and theoretical elegance
+to limit these classical density functionals
+to one-center convolutions, which leads us to seek a
 simplified expression for the contact value of the correlation
-function---which is the same as the contact value of the cavity
+function--- which is the same as the contact value of the cavity
 correlation function for hard spheres.
 In a system with an inhomogeneous density, we seek a \emph{local}
-value for $g_\sigma$.  There are two reasonable options for defining
+value for $g_\sigma$ that is dependent upon only one position variable $\rr$.
+There are two reasonable options for defining
 such a local function: a symmetric formulation (which we refer to as $S$) and an
@@
 where $\sigma$ is the hard sphere diameter and density $n_0$ is one of the fundamental measures of Fundamental
-Measure Theory (FMT), and is ideal for treating touching spheres, as
+Measure Theory (FMT).\ \ $g^S_\sigma(\rr)$ is defined to treat the geometrically
+symmetric possibility of spheres touching at the position $\rr$ as
 illustrated in Figure~\ref{fig:n0}.
@@
 where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.
-We can thus readily compute the derivative of free energy with respect
+Each sphere will be in contact with spheres that are one diameter away
+from its center so we expand a number of contact area spheres
+of radius $\sigma$ and divide by two.  We can thus readily compute the derivative of free energy with respect
 to hard-sphere diameter, which is half the force on \emph{all} the
@@
 compute the pressure on the spheres, we need to divide the total force
-by the total area of contact, which means dividing by $N 4\pi
-\sigma^2$.
+by the total area of contact which means we divide by $N 4\pi \sigma^2$.
 \begin{align}

### 2012-11-02 papers/contact/paper.tex
edit contact paper more in response to referees Tested everything on bingley.

--- 2012-08-31 papers/contact/paper.tex
@@
 
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,jcp,10pt,aip]{revtex4-1}
+\documentclass[letterpaper,amsmath,amssymb,10pt,pre]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 A key input to these functionals is the correlation function
-evaluated at contact, which is required for both the chain and
+evaluated at contact, which is identical to the contact value of the cavity
+correlation function for hard spheres, and is required for both the chain and
 association terms in the SAFT free energy.  The chain term describes
@@
 introduces a new approximation for the contact value of the
-correlation function (described in
-Section~\ref{sec:gross})\cite{gross2009density}.
+correlation function\cite{gross2009density}.
 %% The dispersion term is long-range and thus has significant
@@
 to establish the advantages and disadvantages of each approach.
+
+Here we will briefly describe how the contact value of the correlation
+function has been used in two recent papers introducing SAFT-based
+classical density functionals.  In order to aide in understanding, we
+will use our own notation to describe the work of these authors.  In
+his paper presenting a density functional based on the PCP-SAFT
+equation of state\cite{gross2009density}, Gross introduces the chain
+free energy in SAFT as
+\begin{align}
+  \frac{A_\textit{chain}}{kT} &= -(m-1) \int n(\rr) \left(\ln\left(n^A(\rr)
+  g_\sigma^A(\rr) \right)-1\right) d\rr \label{eq:Achain}
+\end{align}
+where $n^A(\rr)$ is a weighted density defined in Equation~\ref{eq:nA}
+and $g_\sigma^A(\rr)$ the local value of the correlation function at
+contact, which we define in Equation~\ref{eq:gA}, and in
+Section~\ref{sec:gross} we describe Gross's approximation for this
+function. In a paper describing a classical
+density functional for inhomogeneous associating
+fluids\cite{yu2002fmt-dft-inhomogeneous-associating}, Yu and Wu define
+the association free energy as
+\begin{align}
+  \frac{A_\textit{assoc}}{kT} &= \sum_i \int \zeta(\rr)n_0(\rr) \left(\ln X_i(\rr) - \frac12
+  X_i(\rr) + \frac12\right) d\rr \label{eq:Aassoc} \\
+  X_i(\rr) &= \frac{1}{1 + \sum_j \frac{n_0(\rr) g_\sigma^{S}(\rr)}{\zeta(\rr)}
+                                 X_j(\rr)\kappa_{ij} \left(e^{\beta
+                                   \epsilon_{ij}}-1\right)}
+  \label{eq:X}
+\end{align}
+where $X_i(\rr)$ is the fraction of interaction sites of type $i$ at
+position $\rr$ that are unoccupied, $n_0(\rr)$ is a weighted density
+defined in Equation~\ref{eq:n0}, $\zeta(\rr)$ is a nonlocal measure of
+the local density gradient defined in Equation~\ref{eq:zeta}, and
+$g_\sigma^S(\rr)$ is a different local value of the correlation
+function at contact, which we define in Equation~\ref{eq:gS}.  In
+Section~\ref{sec:yuwu} we present the approximation to
+$g_\sigma^S(\rr)$ used by Yu and Wu.
 
@@
 distance $|\rr_1-\rr_2|$ and can be expressed as a function of a
-single variable. The contact value of the correlation function is this
-correlation function's value when evaluated at a distance of the
-diameter $\sigma$.  It is desirable for reasons of efficiency to limit CDFT
-functionals to one-center convolutions, which leads us to seek a
-simplified expression for the contact value of the correlation
-function---which is the same as the contact value of the cavity
-correlation function for hard spheres.
-In a system with an inhomogeneous density, we seek a \emph{local}
-value for $g_\sigma$.  There are two reasonable options for defining
-such a local function: a symmetric formulation (which we refer to as $S$) and an
-asymmetric formulation (which we refer to as $A$).
+single variable, and the contact value of the correlation function is
+its value when evaluated at a distance of the diameter $\sigma$.  The
+pair correlation function of an \emph{inhomogeneous} fluid is not as
+simple, but it is desirable for reasons of computational efficiency to
+construct classical density functionals using just one-center
+convolutions.  Moreover, a local function is helpful when defining
+functionals such as Equations~\ref{eq:Achain}-\ref{eq:X}.  This leads
+us to seek a \emph{local} value for $g_\sigma$ that is dependent upon
+only one position variable $\rr$.  There are two reasonable options
+for defining such a local function: a symmetric formulation such as
+that used in Equations~\ref{eq:Aassoc}-\ref{eq:X} (which we refer to
+as $S$) and an asymmetric formulation such as that used in
+Equation~\ref{eq:Achain} (which we refer to as $A$).
 
@@
 where $\sigma$ is the hard sphere diameter and density $n_0$ is one of the fundamental measures of Fundamental
-Measure Theory (FMT), and is ideal for treating touching spheres, as
+Measure Theory (FMT).  $g^S_\sigma(\rr)$ is defined to treat the geometrically
+symmetric possibility of spheres touching at the position $\rr$ as
 illustrated in Figure~\ref{fig:n0}.
@@
   n_0(\rr) &= \int n(\rr')\frac{\delta(\sigma/2 -|\rr-\rr'|)}{\pi\sigma^2} d\rr'
-\end{align}
-This functional gives a value averaged over all spheres that touch at
-the position $\rr$.
+  \label{eq:n0}
+\end{align}
+This functional $n_0(\rr)$ gives a density averaged over all spheres that touch at
+the position $\rr$.  Together, $n_0(\rr)$ and $g_\sigma^S(\rr)$ are
+used in the association free energy given in Equations~\ref{eq:Aassoc}-\ref{eq:X}.
 
@@
 
+\begin{figure}
+\includegraphics[width=7cm]{figs/nA}
+\caption{Set of hard spheres that are included in $n_A(\mathbf{r})$,
+  which consist of those which just touch a sphere centered at
+  $\mathbf{r}$.  The dashed line illustrates the surface over which
+  contact is possible.}
+\label{fig:nA}
+\end{figure}
+
 In contrast, the asymmetrically averaged $A$ correlation function is
@@
 density of spheres that are touching a sphere that is located at
-point $\rr$.
+point $\rr$, as illustrated in Figure~\ref{fig:nA}.
 \begin{align}
@@
 density over spheres touching a sphere that is located at the
-position~$\rr$.
+position~$\rr$.  The functionals $n_A(\rr)$ and $g_\sigma^A(\rr)$ are
+used in the chain free energy given in Equation~\ref{eq:Achain}.
 
@@
 where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.
-We can thus readily compute the derivative of free energy with respect
-to hard-sphere diameter, which is half the force on \emph{all} the
-hard spheres---since we're changing all their diameters at once.  To
-compute the pressure on the spheres, we need to divide the total force
-by the total area of contact, which means dividing by $N 4\pi
-\sigma^2$.
-\begin{align}
-  p_\sigma &= \frac{1}{N 4\pi \sigma^2} \frac12 \frac{dA_{HS}}{d\sigma} \\
+We compute the pressure on each sphere using the total force with
+respect to a changing radius of all the spheres.
+To find the pressure, we divide by $4\pi \sigma^2$, which is the area
+over which contact is possible, as defined by the centers of the
+contacting spheres, as illustrated in Figure~\ref{fig:nA}.
+Finally, we divide by $N$ to account for the total area of all the
+spheres in the fluid.
+\begin{align}
+  p_\sigma &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
+  &= \frac{1}{N 4\pi \sigma^2} \frac12 \frac{dA_{HS}}{d\sigma} \\
   &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
@@
 \end{align}
-Details of evaluating the functional derivative in
-Equation~\ref{eq:g-A-exact} using FMT are given in
-the appendix.  The equation for $g_\sigma^A$ requires
-finding convolutions of local derivatives of the free energy, making
-this formulation computationally somewhat more expensive than the free
-energy itself.
+Equation~\ref{eq:g-A-exact} is an exact expression for
+$g_\sigma^A(\rr)$ as defined in Equation~\ref{eq:gA}.  However, since
+we do not know the exact hard-sphere free energy, we approximate
+$A_{HS}$ using FMT.  Details of evaluating the functional derivative
+in Equation~\ref{eq:g-A-exact} using FMT are given in the appendix.
+The equation for $g_\sigma^A$ requires finding convolutions of local
+derivatives of the free energy, making this formulation
+computationally somewhat more expensive than the free energy itself.
 
@@
 where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
-dimensionless free energy density.  This step is an
+dimensionless free energy density.  This expression is an
 approximation---unlike the analogous
@@
   n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \label{eq:nYu} \\
-  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
+  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2} \label{eq:zeta}
 \end{align}

### 2012-11-06 papers/contact/paper.tex
final edits for referees Tested everything on kipu.

@@
-% TODO:
-
-% Write ``Liquid-vapor interface'' section A.
-
-% Look up question-mark references (citations).
-
-\documentclass[letterpaper,amsmath,amssymb,10pt,pre]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 
-\author{Jeff Schulte}
-\author{Patrick Kreitzberg}
-\author{Chris Haglund}
+\author{Jeff B. Schulte}
+\author{Patrick A. Kreitzberg}
+\author{Chris V. Haglund}
 \author{David Roundy}
@@
 
+%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 introduced, one which uses the
-correlation function of Yu and Wu\cite{bryk2006density} and another that
+correlation function of Yu and Wu\cite{bryk2006density} and another which
 introduces a new approximation for the contact value of the
@@
 %% weak position dependence, is amenable to mean-field approximations.
-Given these different approaches, it seems valuable to examine this
-property of the hard-sphere fluid through direct simulation, in order
-to establish the advantages and disadvantages of each approach.
 
 Here we will briefly describe how the contact value of the correlation
-function has been used in two recent papers introducing SAFT-based
-classical density functionals.  In order to aide in understanding, we
+function has been used in two of these recent papers introducing SAFT-based
+classical density functionals.  For simplicity, we
 will use our own notation to describe the work of these authors.  In
@@
 and $g_\sigma^A(\rr)$ the local value of the correlation function at
-contact, which we define in Equation~\ref{eq:gA}, and in
+contact, which we define in Equation~\ref{eq:gA}.  In
 Section~\ref{sec:gross} we describe Gross's approximation for this
@@
 defined in Equation~\ref{eq:n0}, $\zeta(\rr)$ is a nonlocal measure of
-the local density gradient defined in Equation~\ref{eq:zeta}, and
-$g_\sigma^S(\rr)$ is a different local value of the correlation
+the density gradient defined in Equation~\ref{eq:zeta}, and
+$g_\sigma^S(\rr)$ is a form of the correlation
 function at contact, which we define in Equation~\ref{eq:gS}.  In
 Section~\ref{sec:yuwu} we present the approximation to
-$g_\sigma^S(\rr)$ used by Yu and Wu.
+$g_\sigma^S(\rr)$ introduced by Yu and Wu.
+Given these differing approaches, it seems valuable to examine this
+property of the hard-sphere fluid through direct simulation, in order
+to establish the advantages and disadvantages of each approach.
 
 Although these recent works have introduced approximate functionals
-for the contact value of the correlation
-function\cite{yu2002fmt-dft-inhomogeneous-associating,
+for the contact value of the correlation function for an inhomogeneous
+hard-sphere fluid\cite{yu2002fmt-dft-inhomogeneous-associating,
   gross2009density}, there has not been a study that specifically
-addresses this contact value for an inhomogeneous hard-sphere fluid.
+addresses this functional.
 %
@@
 simple, but it is desirable for reasons of computational efficiency to
-construct classical density functionals using just one-center
+construct classical density functionals using only one-center
 convolutions.  Moreover, a local function is helpful when defining
-functionals such as Equations~\ref{eq:Achain}-\ref{eq:X}.  This leads
+functionals based on perturbation theory, such as those in
+Equations~\ref{eq:Achain}-\ref{eq:X}.  This leads
 us to seek a \emph{local} value for $g_\sigma$ that is dependent upon
@@
 \end{align}
-where $\sigma$ is the hard sphere diameter and density $n_0$ is one of the fundamental measures of Fundamental
-Measure Theory (FMT).  $g^S_\sigma(\rr)$ is defined to treat the geometrically
+where $\sigma$ is the hard sphere diameter and the density $n_0$ is one of
+the fundamental measures of Fundamental Measure Theory (FMT).  The
+functional $g^S_\sigma(\rr)$ is defined to treat the geometrically
 symmetric possibility of spheres touching at the position $\rr$ as
@@
 \end{align}
-where the density $n_A(\rr)$ is similar to $n_0$, but measures the
+where the density $n_A(\rr)$ is analogous to $n_0(\rr)$, but measures the
 density of spheres that are touching a sphere that is located at
@@
 \end{align}
-Since we are interested in the correlation function at the surface of
-the hard spheres, we need to compute the pressure on that surface.  The pressure on a hard sphere can be
+Since we are interested in the correlation function at the surface at
+which hard spheres contact, we need to compute the pressure on that
+surface.   This pressure can be
 readily computed from the dependence of the Carnahan-Starling free
@@
 \begin{align}
-  p_\sigma(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
+  p_\sigma(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2}\frac12 \frac{\delta
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
-%  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
+%  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 4\pi \sigma^2}\frac12 \frac{\delta
 %    A_{HS}}{\delta \sigma(\mathbf{r})} \\
   g_\sigma^A(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
-  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
+  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 4\pi \sigma^2}\frac12 \frac{\delta
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
@@
   g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 2\pi \sigma^2}
+  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 4\pi \sigma^2}\frac12
   \frac{\partial \Phi(\rr)}{\partial \sigma} \label{eq:g-S}

### 2012-11-16 papers/pair-correlation/paper.tex
begin pair-correlation paper Tested everything on node08.

--- 2012-10-26 papers/contact/paper.tex
@@
-% TODO:
-
-% Write ``Liquid-vapor interface'' section A.
-
-% Look up question-mark references (citations).
-
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,10pt,pre]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 \begin{document}
-\title{Using Fundamental Measure Theory to Treat the Correlation
-  Function of the Inhomogeneous Hard-Sphere Fluid}
+\title{Using Fundamental Measure Theory to treat the two-point pair
+  distribution function of the Inhomogeneous Hard-Sphere Fluid}
 
-\author{Jeff Schulte}
-\author{Patrick Kreitzberg}
-\author{Chris Haglund}
+\author{Paho}
+\author{???}
 \author{David Roundy}
@@
 
+%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
-  We investigate the value of the correlation function of an
-  inhomogeneous hard-sphere fluid at contact.  This
-  quantity plays a critical role in Statistical Associating Fluid
-  Theory (SAFT), which is the basis of a number of recently developed classical
-  density functionals.  We define two averaged values for the
-  correlation function at contact, and derive formulas for each of
-  them from the White Bear version of the Fundamental Measure Theory
-  functional~\cite{roth2002whitebear}, using an assumption of
-  thermodynamic consistency. We test these formulas, as well as two
-  existing formulas~\cite{yu2002fmt-dft-inhomogeneous-associating,
-    gross2009density} against Monte Carlo simulations, and find
-  excellent agreement between the Monte Carlo data and one of our
-  averaged correlation functions.
+  We develop and test an efficient approximation for the two-point
+  pair distribution function of an inhomogeneous hard-sphere fluid.
 \end{abstract}
@@
 
-% The following are papers that use a SAFT-based classical DFT with at
-% least some of the terms purely local
-\newcommand\saftlocaldft{felipe2001examination, gloor2002saft,%
-  gloor2004accurate, clark2006developing, gloor2007prediction,%
-  kahl2008modified, gross2009density}
-% The following are papers that use a SAFT-based classical DFT with
-% all the terms that should be non-local being non-local.
-\newcommand\saftnonlocaldft{yu2002fmt-dft-inhomogeneous-associating,%
-  fu2005vapor-liquid-dft,bryk2006density}
+Attard worked out the triplet correlation function, which is something
+that we can access from the pair distribution function, if we consider
+the pair correlation near a single hard sphere
+solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al} wrote an
+interesting paper using DFT and Monte Carlo to compute the
+three-particle distribution function (same as the triplet correlation
+function)\cite{gonzalez1999test}.  Their theory compares favorably
+with the ``superposition approximation'', and we could probably do
+something similar.
 
-There has been considerable recent interest in using Statistical
-Associating Fluid Theory (SAFT) to construct classical density
-functionals to describe associating
-fluids\cite{\saftlocaldft,\saftnonlocaldft}.  This approach has been
-successful in qualitatively describing the dependence of surface
-tension on temperature.
-%
-%% Unfortunately, most of these constructed
-%% functionals\cite{\saftlocaldft} cannot be used to study arbitrary
-%% inhomogeneous density distributions, because they use local density
-%% approximations that are only valid for densities that vary slowly over
-%% molecular length scales.
-%
-A key input to these functionals is the correlation function
-evaluated at contact, which is required for both the chain and
-association terms in the SAFT free energy.  The chain term describes
-the chain formation energy in polymeric fluids, while the association
-term describes the effects of hydrogen bonding, both of which can be
-large.
-%
-Yu and Wu introduced in 2002 a functional for the association term of
-the free energy, which included a functional for the contact value of
-the correlation function (described in
-Section~\ref{sec:yuwu})\cite{yu2002fmt-dft-inhomogeneous-associating},
-which has subsequently been used in the development of other
-SAFT-based functionals\cite{fu2005vapor-liquid-dft, bryk2006density}.
-%% Each of these terms has some dependence on the
-%% inhomogeneity in the density.  The ideal gas is purely local, and in
-%% fact must be the only purely local term in the free energy in order
-%% for the functional to satisfy the contact-value theorem.  The
-%% hard-sphere contribution contribution to the free energy has been
-%% thoroughly studied, and is well approximated by the White Bear version
-%% of the Fundamental-Measure Theory (FMT)
-%% functional\cite{roth2002whitebear}.
-Two functionals for the chain contribution have recently been
-introduced, one which uses the
-correlation function of Yu and Wu\cite{bryk2006density} and another that
-introduces a new approximation for the contact value of the
-correlation function (described in
-Section~\ref{sec:gross})\cite{gross2009density}.
-%% The dispersion term is long-range and thus has significant
-%% dependence on the density distribution, but because its relatively
-%% weak position dependence, is amenable to mean-field approximations.
-Given these different approaches, it seems valuable to examine this
-property of the hard-sphere fluid through direct simulation, in order
-to establish the advantages and disadvantages of each approach.
+Plischke and Henderson worked out the pair correlation function near a
+hard wall using Percus-Yevick, and compare with Monte Carlo
+results\cite{plischke1986pair}.  They plot the pair correlation
+function along interesting paths.
 
-Although these recent works have introduced approximate functionals
-for the contact value of the correlation
-function\cite{yu2002fmt-dft-inhomogeneous-associating,
-  gross2009density}, there has not been a study that specifically
-addresses this contact value for an inhomogeneous hard-sphere fluid.
-%
-In this paper we introduce two definitions for the locally averaged
-correlation function of an inhomogeneous system.
-%
-%% The first is more symmetric (which we call the $S$ case), while the
-%% second involves a sphere-centered, asymmetrical interpretation (the
-%% $A$ case).
-%
-Given these definitions, we will present a thermodynamic derivation
-for each correlation function from the free energy functional.  We
-will then discuss the correlation functions of Yu and Wu and of Gross, and
-will end by comparing all four approximations with Monte-Carlo
-simulations of the hard-sphere fluid at a variety of hard-wall
-surfaces.
-
-\section{Correlation function with inhomogeneity}
-
-We define our terms using the two-particle density
-$n^{(2)}(\rr_1,\rr_2)$, which gives the probability per unit volume
-squared of finding one particle at position $\rr_1$ and the other at
-position $\rr_2$.  The pair correlation function is defined by
-\begin{align}
-  g(\rr_1,\rr_2) &\equiv \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
-\end{align}
-The contact value of the correlation function is this
-correlation function's value when evaluated at a distance of the
-diameter $\sigma$.  In a homogeneous fluid, the pair correlation only depends on the
-distance $|\rr_1-\rr_2|$ and can be expressed as a function of a
-single variable. The pair correlations for an inhomogeneous fluid are of course
-not as simple but it is desirable for reasons of computational efficiency and theoretical elegance
-to limit these classical density functionals
-to one-center convolutions, which leads us to seek a
-simplified expression for the contact value of the correlation
-function--- which is the same as the contact value of the cavity
-correlation function for hard spheres.
-In a system with an inhomogeneous density, we seek a \emph{local}
-value for $g_\sigma$ that is dependent upon only one position variable $\rr$.
-There are two reasonable options for defining
-such a local function: a symmetric formulation (which we refer to as $S$) and an
-asymmetric formulation (which we refer to as $A$).
-
-For the symmetric $S$ case, the correlation function at contact is
-given by:
-\begin{align}
-  g^S_\sigma(\rr) &= \frac{1}{n_0(\rr)^2}\int n^{(2)}(\rr - \rr', \rr
-  + \rr')
-  \frac{\delta(\sigma/2 -|\rr'|)}{\pi\sigma^2}d\rr' \label{eq:gS}
-\end{align}
-where $\sigma$ is the hard sphere diameter and density $n_0$ is one of the fundamental measures of Fundamental
-Measure Theory (FMT).\ \ $g^S_\sigma(\rr)$ is defined to treat the geometrically
-symmetric possibility of spheres touching at the position $\rr$ as
-illustrated in Figure~\ref{fig:n0}.
-\begin{align}
-  n_0(\rr) &= \int n(\rr')\frac{\delta(\sigma/2 -|\rr-\rr'|)}{\pi\sigma^2} d\rr'
-\end{align}
-This functional gives a value averaged over all spheres that touch at
-the position $\rr$.
-
-\begin{figure}
-\includegraphics[width=5cm]{figs/n0}
-\caption{Set of hard spheres that are included in $n_0(\mathbf{r})$, which
-  consist of those which just touch the point $\mathbf{r}$.}
-\label{fig:n0}
-\end{figure}
-
-In contrast, the asymmetrically averaged $A$ correlation function is
-given by
-\begin{align}
-  g^A_\sigma(\rr) &= \frac{1}{n(\rr)n_A(\rr)}
-  \int n^{(2)}(\rr, \rr + \rr')
-  \frac{\delta(\sigma -|\rr'|)}{4\pi\sigma^2}d\rr' \label{eq:gA}
-\end{align}
-where the density $n_A(\rr)$ is similar to $n_0$, but measures the
-density of spheres that are touching a sphere that is located at
-point $\rr$.
-\begin{align}
-  n_A(\rr) &= \int n(\rr')
-  \frac{\delta(\sigma -|\rr-\rr'|)}{4\pi\sigma^2} d\rr' \label{eq:nA}
-\end{align}
-Thus $g_\sigma^A$ corresponds to an average of the two-particle
-density over spheres touching a sphere that is located at the
-position~$\rr$.
-
-%% In the process of defining these two averaged correlation functions,
-%% we also defined two averaged densities, $n_0(\rr)$ and $n_A(\rr)$,
-%% which are the density of spheres available to be in
-%% contact.  In the homogeneous limit, $n_0 = n_A = n$.  It is an
-%% open question, which of these averages will be more useful in any
-%% particular functional.  Suffice to say, either average is a
-%% \emph{possible} way to convert a function that is defined for a
-%% homogeneous system to a functional that is applicable to inhomogeneous
-%% systems.
-
-%% Yu and Wu introduce an approximation for
-%% $g^S$\cite{yu2002fmt-dft-inhomogeneous-associating}, which we will
-%% discus in Section~\ref{sec:yuwu}.  Gross introduced an
-%% approximation for $g^A$\cite{gross2009density}, which we will
-%% introduce in Section~\ref{sec:gross}.  We will also introduce our own
-%% approximations for both $g^A$ and $g^S$ which are based on the
-%% assumption of thermodynamic consistency within FMT, in
-%% Sections~\ref{sec:g-S} and~\ref{sec:g-A}.  We will report on the accuracy of
-%% these approximations by comparing with Monte Carlo simulations of the
-%% inhomogeneous hard-sphere fluid.  We will focus on the contact
-%% densities, rather than the correlation function itself, because these
-%% are simpler, and because they are the quantity that you want to
-%% ``average''.  i.e. if you were to average $g$ itself, you'd want to
-%% weight that average based on the density, which would mean averaging
-%% $n^{(2)}$.
-
-%% \subsection{Association free energy of SAFT}
-
-%% \fixme{Integrate the following equations into this section...}
-%% \begin{align}
-%%   \frac{A_\textit{chain}}{kT} &= -(m-1) n \left(\ln\left(n g_\sigma \right)-1\right)
-%% \end{align}
-%% \begin{align}
-%%   \frac{A_\textit{assoc}}{kT} &= \sum_i n \left(\ln X_i - \frac12 X_i + \frac12\right) \\
-%%   X_i &= \frac{1}{1 + \sum_j n g_\sigma X_j\kappa_{ij} \left(e^{\beta \epsilon_{ij}}-1\right)}
-%% \end{align}
-
-%% \newcommand\epsilonassoc{\ensuremath{\varepsilon_\textit{AB}}}
-%% \newcommand\kappaassoc{\ensuremath{\kappa_\textit{AB}}}
-\newcommand\ncontact{\ensuremath{n_\textit{contact}}}
-
-%% The free energy term in Statistical Associating Fluid Teory (SAFT)
-%% from which it derives its name is the association term, which accounts
-%% for hydrogen bonding.  Hydrogen bonds are modeled as attractive
-%% patches (``association sites'') on the surface of hard spheres.  These
-%% sites represent protons or electron lone pairs, and have an attractive
-%% energy $\epsilonassoc$ when two molecules are oriented such that the
-%% proton of one overlaps with the lone pair of the other.  The volume
-%% over which this interaction occurs is $\kappaassoc$, giving the
-%% association term in the free energy has two empirical parameters fit
-%% to experimental data.
-
-%% The association free energy per unit volume has the form
-%% \begin{align}
-%%   f_\text{assoc} &= k_BT n\sum_A 
-%%                   \left(\ln X^A - \frac{X^A}{2} + \frac12\right)
-%% \end{align}
-%% where the summation is over the association sites, and $X^A$ is the
-%% fraction of association sites \emph{not} hydrogen-bonded.  The
-%% fractions $X^A$ are determined by the self-consistent equations
-%% \begin{align}
-%%   X^A &= \frac{1}
-%%   {1 + \sum_B n X^B \Delta^{AB}}\label{eq:X}
-%%   \\
-%%   \Delta^{AB} &= g(\sigma) \kappaassoc\left( e^{\epsilonassoc/kT} -
-%%   1 \right)\label{eq:delta}
-%% \end{align}
-%% where $g(\sigma)$ is the correlation function evaluated at contact.
-%% The product of density with correlation function evaluated at contact
-%% gives the contact density, when we combine Equations~\ref{eq:X} and
-%% \ref{eq:delta}:
-%% \begin{align}
-%%   X^A &= \frac{1}
-%%   {1 + \sum_B \ncontact\kappaassoc X^B\left( e^{\epsilonassoc/kT} -
-%%   1 \right)}
-%% \end{align}
-%% where the product $\ncontact\kappaassoc$ is the expected number of
-%% molecules present in the association volume in the absence of the
-%% association interaction.  In the SAFT-VR
-%% model\cite{gil-villegas-1997-SAFT-VR}, this contact density is found
-%% by adding a perturbative correction to the contact value for a hard
-%% sphere fluid.
+\section{Pair distribution function with inhomogeneity}
 
@@
 
-\subsection{Homogeneous limit}
-
-In order to motivate our derivation of the correlation function at
-contact for the \emph{inhomogeneous} hard-sphere fluid, we begin by
-deriving the well-known formula for $g_\sigma$ for the
-\emph{homogeneous} fluid that comes from the Carnahan-Starling free
-energy.  The contact value of the correlation function density can be
-found by using the contact-value theorem, which states that the
-pressure on any hard surface is determined by the density at contact:
-\begin{align}
-  p &= k_BT n_\textit{contact} \\
-  &= k_BT n g_\sigma
-\end{align}
-Since we are interested in the correlation function at the surface of
-the hard spheres, we need to compute the pressure on that surface.  The pressure on a hard sphere can be
-readily computed from the dependence of the Carnahan-Starling free
-energy on hard sphere radius.
-\begin{align}
-  A_{HS} &= Nk_BT \frac{4\eta - 3\eta^2}{(1-\eta)^2}
-\end{align}
-where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.
-Each sphere will be in contact with spheres that are one diameter away
-from its center so we expand a number of contact area spheres
-of radius $\sigma$ and divide by two.  We can thus readily compute the derivative of free energy with respect
-to hard-sphere diameter, which is half the force on \emph{all} the
-hard spheres---since we're changing all their diameters at once.  To
-compute the pressure on the spheres, we need to divide the total force
-by the total area of contact which means we divide by $N 4\pi \sigma^2$.
-\begin{align}
-  p_\sigma &= \frac{1}{N 4\pi \sigma^2} \frac12 \frac{dA_{HS}}{d\sigma} \\
-  &= k_BT n \frac{1 - \frac{\eta}2}{(1-\eta)^3}
-\end{align}
-Using the contact-value theorem, we can thus find the well-known
-correlation function evaluated at contact.
-\begin{align}
-  g_\sigma &= \frac{1 - \frac{\eta}2}{(1-\eta)^3} \label{eq:cs-g}
-\end{align}
-Extending this derivation to the inhomogeneous fluid requires that we
-find the pressure felt by the surface of particular spheres.
-
-\subsection{Asymmetrically averaged correlation function}\label{sec:g-A}
-
-We will begin our derivation of the locally averaged correlation
-function with the asymmetric definition of $g_\sigma^A(\rr)$ given in
-Equation~\ref{eq:gA}, which is averaged over contacts in which one of
-the two spheres is located at position~$\rr$.  This correlation
-function is related to the contact density averaged over the surface
-of a sphere located at~$\rr$, and can thus be determined by finding
-the pressure on that sphere.  We find this pressure from the change in
-free energy resulting from an infinitesimal expansion of any
-spheres located at position~$\rr$.  From this pressure, we derive a
-formula for the correlation function $g_\sigma^A(\rr)$ as was done in
-the previous section:
-\begin{align}
-  p_\sigma(\mathbf{r}) &= \frac{1}{n(\rr) 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
-%  \ncontact(\rr) &= \frac{1}{n(\rr) k_BT 2\pi \sigma^2} \frac{\delta
-%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
-  g_\sigma^A(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
-  &= \frac{1}{n(\rr) n_A(\rr)}\frac{1}{ k_BT 2\pi \sigma^2} \frac{\delta
-    A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:g-A-exact}
-\end{align}
-Details of evaluating the functional derivative in
-Equation~\ref{eq:g-A-exact} using FMT are given in
-the appendix.  The equation for $g_\sigma^A$ requires
-finding convolutions of local derivatives of the free energy, making
-this formulation computationally somewhat more expensive than the free
-energy itself.
-
-\derivation{
-  \end{widetext}
-}
-
-\subsection{Symmetrically averaged correlation function}\label{sec:g-S}
-
-We now address the symmetrically averaged correlation function, which
-is defined in Equation~\ref{eq:gS}.  This corresponds to the
-correlation function averaged for spheres \emph{touching at a given
-  point}.  In this case, we conceptually would like to evaluate the
-pressure felt by the surface of spheres where that surface is located
-at point~$\rr$.  We can approximate this value by assuming that this
-pressure will be simply related to the free energy density at
-point~$\rr$.  Through a process similar to the previous derivations, this
-leads to the expression
-\begin{align}
-  %% p_{HS}^S(\mathbf{r}) &= \frac{1}{n_0(\rr) 2\pi \sigma^2} \frac{\partial
-  %%   a_{HS}(\rr)}{\partial \sigma} \label{eq:p_{HS}^S}\\
-%  \ncontact(\rr) &= \frac{1}{n_0(\rr) k_BT 2\pi \sigma^2} \frac{\delta
-%    A_{HS}}{\delta \sigma(\mathbf{r})} \\
-  g_\sigma^S(\rr)% &= \frac{\ncontact(\rr)}{n_0(\rr)} \\
-  &= \frac{1}{n_0(\rr)^2}\frac{1}{ 2\pi \sigma^2}
-  \frac{\partial \Phi(\rr)}{\partial \sigma} \label{eq:g-S}
-\end{align}
-where $\Phi(\rr) = \Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)$ is the
-dimensionless free energy density.  This step is an
-approximation---unlike the analogous
-Equation~\ref{eq:g-A-exact}---because it assumes that we have
-available a local functional $\Phi(\rr)$ whose derivative provides the
-pressure needed to compute $g_\sigma(\rr)$.  Equation~\ref{eq:g-S}
-requires that we evaluate the derivatives of the fundamental measures
-$n_\alpha(\rr)$ with respect to diameter, which leads us to
-derivatives of the $\delta$~function, which we can simplify and
-approximate using an assumption of a reasonably smooth density:
-\begin{align}
-  \frac{\partial n_2(\rr)}{\partial \sigma}
-  &= \frac12 \int \delta'(\frac\sigma2 - |\rr-\rr'|) n(\rr')d\rr' \\
-  &= \frac2{\sigma}n_2(\rr) - \frac12 \int \delta\left(\frac\sigma2 - |\rr-\rr'|\right)
-  \frac{\rr-\rr'}{|\rr-\rr'|}\cdot\mathbf{\nabla}n(\rr')d\rr' \\
-  &\approx \frac{2}{\sigma}n_2(\rr)
-\end{align}
-In the systems that we study, the density is \emph{not} reasonably
-smooth, but we can state empirically making this approximation
-nevertheless improves the predictions of our functional, while at the
-same time reducing its computational cost by avoiding the need to
-calculate any additional weighted densities or convolutions.
-
-\subsection{Gross's asymmetrically averaged correlation functional}\label{sec:gross}
-One approximation for the correlation function is that of
-Gross\cite{gross2009density}, which is of the asymmetrically averaged
-variety ($g_\sigma^A$):
-\begin{align}
-  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{\pi}{12}\sigma^3n_A(\rr)}{\left(1 -
-    \frac{\pi}{6}\sigma^3n_A(\rr)\right)^3}
-\end{align}
-where $n_A$ is the averaged density defined in Equation~\ref{eq:nA}.
-This formula is arrived at by using the density averaged over all
-spheres that could be touching a sphere at point~$\rr$ in the
-Carnahan-Starling equation for the correlation function at contact,
-given in Equation~\ref{eq:cs-g}.
-
-\subsection{Yu and Wu's symmetrically averaged functional}\label{sec:yuwu}
-
-Yu and Wu developed a functional for the correlation function
-evaluated at contact which is symmetrically
-averaged~\cite{yu2002fmt-dft-inhomogeneous-associating}.  However,
-instead of using $n_0$ as the corresponding density, they use a
-density given by
-\begin{align}
-  n_\text{Yu}(\rr) &= n_0(\rr) \zeta(\rr) \label{eq:nYu} \\
-  \zeta &= 1 - \frac{\mathbf{n_2}\cdot\mathbf{n_2}}{n_2^2}
-\end{align}
-where the function $\zeta$ is a measure of local inhomogeneity at the
-point of contact, and has the effect of reducing this density at
-interfaces.  Because of this difference, the correlation function of
-Yu and Wu cannot be directly compared with $g_\sigma^S$ as defined in
-Equation~\ref{eq:gS}.  Therefore in order to make a comparison we move
-the factors of $\zeta$ in Equation~\ref{eq:nYu} from the density into
-the correlation function itself.
-\begin{align}
-  g_\sigma^\text{Yu,S} &= \zeta^2 g_\sigma^\text{Yu} \\
-   &= \zeta^2\left(\frac{1}{1-n_3}
-    + \frac14 \frac{\sigma n_2\zeta}{(1-n_3)^2}
-    + \frac1{72} \frac{\sigma^2 n_2^2 \zeta}{(1-n_3)^3}\right)
-\end{align}
-where $g_\sigma^\text{Yu}$ is the correlation function as defined in
-reference~\cite{yu2002fmt-dft-inhomogeneous-associating}, and
-$g_\sigma^\text{Yu,S}$ is the function we will examine in this paper.
-
-%% \section{Method}
-
-%% We implemented the above functionals as part of the Deft software
-%% package~\cite{deft}, which uses a regular Cartesian grid with a grid
-%% spacing of $0.05\sigma$.  Convolutions are implemented with fast
-%% fourier transforms.  For comparison, we performed Monte Carlo
-%% simulations of the hard-sphere fluid.  We performed these simulations
-%% in a fixed-number ensemble.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-10}
-  \caption{Density and correlation function at a hard wall with bulk
-    filling fraction of 0.1.}
-  \label{fig:walls-10}
-\end{figure}
-
 \section{Comparison with simulation}\label{sec:comparison}
 
-We performed a Monte-Carlo simulation of the hard sphere fluid to
-measure the contact value of the correlation function for several
-simple inhomogeneous configurations.  For each configuration, we
-compute the mean density, and the contact values of the correlation
-function, averaged as defined in Equations~\ref{eq:gA} and
-\ref{eq:gS}.  We compare these with the functionals presented in
-sections~\ref{sec:g-A} to~\ref{sec:yuwu}.  We constructed our
-functionals using both the original White Bear
-functional~\cite{roth2002whitebear} as well as the mark II version of
-the White Bear functional~\cite{hansen2006density}, but the results
-were essentially indistinguishable on our plots, so we exclusively
-show the results due to the original White Bear functional.
-
-We simulate the inhomogeneous hard sphere fluid at four hard-wall
-interfaces.  The first and simplest is a flat hard wall.  We
-then study two convex hard surfaces.  One is an excluded sphere with
-diameter $2\sigma$, which corresponds to a ``test particle''
-simulation with one of a hard sphere at the origin with diameter
-$\sigma$.  The second is an excluded sphere with diameter $6\sigma$,
-which demonstrates behavior typical of mildly convex hard surfaces.
-Finally, we study a concave surface given by a hard cavity in which
-our fluid is free to move up to a diameter of $16\sigma$, which
-demonstrates behavior typical of mildly concave surfaces.  In each
-case, we performed a low-density (filling fraction 0.1) and high-density
-(filling fraction 0.4) simulation.  We performed additional
-computations over a wider range of curvatures and densities, but
-chose these as typical examples.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-10}
-  \caption{Density and correlation function around a hard sphere with
-    diameter the same as those in the fluid, with a bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-4-10}
-\end{figure}
-
-\subsection{Low density}
-
-We begin by presenting our low-density results, corresponding to a
-filling fraction of 0.1, which are shown in
-Figures~\ref{fig:walls-10}-\ref{fig:outer-10}.  At this low density,
-the contact value of the correlation function in the bulk is only 1.3,
-indicating that correlations are indeed small and that the fluid should be
-relatively easy to model.  Indeed, the contact density at the hard
-surface is only around 50\% higher than the bulk, and the FMT
-predicted density is close to indistinguishable from the true
-density for each of the four configurations.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-10}
-  \caption{Density and correlation function around a hard sphere with
-    an excluded diameter of $6\sigma$, with a bulk filling fraction of
-    0.1.}
-  \label{fig:inner-12-10}
-\end{figure}
-
-The $g_\sigma^A$ correlation function in each configuration is very
-flat, with only small, smooth changes as the surface is approached.
-Our functional $g_\sigma^A$ very closely matches the Monte Carlo
-predictions in each case, while that of Gross consistently
-underestimates the correlation at the interface by a significant margin.  We note that the theoretical curves
-extend into the region from which the fluid is excluded.  This value
-corresponds to the correlation function that would be observed in the
-vanishingly unlikely scenario in which there was a sphere present at
-that location.  Naturally, we are unable to observe this quantity in
-our Monte Carlo simulations.
-
-The $g_\sigma^S$ correlation function shows considerably more
-structure, as well as additional variation due to the curvature of the
-hard surface.  The symmetric correlation function is nonzero at
-locations where spheres may touch, which for a convex hard surface
-means that $g_\sigma^S$ may be nonzero in the volume in which hard
-spheres are excluded.  In every configuration studied, the agreement
-between the theoretical predictions and the Monte Carlo simulation in
-each case is very poor in the region where there should be no contacts
-at all.  Because $n_0$ is comparable to its bulk value in this region,
-this means that these functionals predict a significant number of
-contacts in the region where there should be none.  The correlation
-function of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
-and ours described in Section~\ref{sec:g-S} give similar results, with
-slightly larger errors in our prediction.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-16-01}
-  \caption{Density and correlation functions near the surface of a
-    spherical cavity with diameter $16\sigma$ at a bulk filling
-    fraction of 0.1.}
-  \label{fig:outer-10}
-\end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-10}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.1.}
-%%   \label{fig:test-particle-10}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-40}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.4.}
-%%   \label{fig:test-particle-40}
-%% \end{figure}
-
-\subsection{High density}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-40}
-  \caption{Density and correlation function at a hard wall with bulk
-    filling fraction of 0.4.}
-  \label{fig:walls-40}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-40}
-  \caption{Density and correlation function around a hard sphere with
-    diameter the same as those in the fluid, with a bulk filling
-    fraction of 0.4.}
-  \label{fig:inner-4-40}
-\end{figure}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-40}
-  \caption{Density and correlation function around a hard sphere with
-    an excluded diameter of $6\sigma$, with a bulk filling fraction of
-    0.4.}
-  \label{fig:inner-12-40}
-\end{figure}
-
-At a higher density corresponding to a filling fraction of 0.4,
-correlations are much stronger, with the bulk contact value of the
-correlation function of 3.7, as seen in
-Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.  This results in larger
-oscillations in the density at the hard surfaces, and correspondingly
-more interesting behavior in the correlation function near the
-interface.  The density predicted by the White Bear functional agrees
-reasonably well with the simulation results, although not so well as
-it did at lower density.  The discrepancies are largest in the case of
-the spherical cavity, in which the DFT considerably underestimates the
-range of the density oscillations.
-
-The asymmetric version of the correlation function once again displays
-relatively smooth behavior with a few small
-oscillations near the interface, and a somewhat elevated value within
-a diameter of the hard surface, with the magnitude of this elevation
-somewhat different in each configuration.  As was the case at low
-density, our correlation function $g_\sigma^A$ matches very closely
-the Monte Carlo data, reproducing quite well the structure near the
-interface in each configuration, although in the spherical cavity
-(Figure~\ref{fig:outer-40}), there is a small, but significant
-discrepancy, comparable to the discrepancy found in the density
-itself.  In each case, the correlation of Gross dramatically
-underestimates the correlation at the interface, at one extreme by
-40\% in the case of the spherical cavity
-(Figure~\ref{fig:outer-40}), and at the other extreme by 15\%
-in the test-particle scenario (Figure~\ref{fig:inner-4-40}).
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-16-04}
-  \caption{Density and correlation functions near the surface of a
-    spherical cavity with diameter $16\sigma$ at a bulk filling
-    fraction of 0.4.}
-  \label{fig:outer-40}
-\end{figure}
-
-The symmetrically averaged correlation function shows considerably
-more structure near the interface at high density, and this structure
-varies considerably depending on the curvature of the hard surface.
-In each case, this structure is not reflected in the theoretical
-predictions, neither that of this paper, nor that of Yu and
-Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  As was the case at
-low density, both functionals give significant and finite values in
-the region in which there are no contacts, but at high density they
-also miss the large oscillations that are present near the flat wall
-and the concave surface (Figures~\ref{fig:walls-40}
-and~\ref{fig:outer-40}).  As was the case at low density, the
-functional of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
-gives slightly better agreement with the simulation results than that
-which we derive in Section~\ref{sec:g-S}.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Conclusion}
-We investigated several approximations to the contact value of the
-correlation function for inhomogeneous fluid distributions
-corresponding to flat, concave, and convex walls.  We
-defined and simulated two averages of the correlation function, an
-asymmetric $A$ average centered at the location of one of the two
-spheres that is in contact, and a symmetric $S$ average centered at
-the point of contact of touching spheres.  For each average, we
-derived a functional form from FMT, and also found an approximation
-that has been used in the literature.  When compared with essentially
-exact Monte Carlo simulations, the $A$ correlation function derived
-from Fundamental Measure Theory in Section~\ref{sec:g-A} gives
-excellent results for each surface, at both high density and low
-density.  The other three approximations that we studied all showed
-significant and systematic deviations under some circumstances.  Thus,
-we recommend that creators of SAFT-based classical density functionals
-consider using the $g_\sigma^A$ functional defined in
-Section~\ref{sec:g-A}.
 
@@
 
-%% \begin{figure}
-%% \includegraphics[width=\columnwidth]{figs/gHS-vs-n}
-%% \caption{A comparison of the various approximations to the contact
-%%   density in the homogeneous limit.  As expected, they are all
-%%   identical, and we won't include this plot in the paper (but it's
-%%   here to verify that our code is working).}
-%% \label{fig:gHS-vs-n}
-%% \end{figure}
-
-%% \begin{figure}
-%% \includegraphics[width=\columnwidth]{figs/free-energy}
-%% \caption{A comparison of the various approximations to the free energy
-%%   in the homogeneous limit.  Note that the White Bear functional uses
-%%   a modified form of the Carnahan equation of state.  As expected,
-%%   these are all identical, and we won't include this plot in the paper
-%%   (but it's here to verify that our code is working).}
-%% \label{fig:free-energy}
-%% \end{figure}
-
 \section*{Appendix}
-
-The expression for the asymmetric correlation function
-$g_\sigma^A(\rr)$ (Equation~\ref{eq:g-A-exact}) involves the
-functional derivative $\frac{\delta A_{HS}}{\delta
-  \sigma(\mathbf{r})}$.  In this appendix we will explain how this
-derivative is evaluated.  We begin by applying the chain rule in the
-following way:
-  \begin{align}
-    \frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})} &=
-    \int \left(
-    \sum_\alpha
-    \frac{\delta A_{HS}}{\delta n_\alpha(\mathbf{r}')}
-    \frac{\delta n_\alpha(\mathbf{r}')}{\delta \sigma(\mathbf{r})}
-    \right) d\mathbf{r}'
-  \end{align}
-This expression requires us to evaluate $\frac{\delta A_{HS}}{\delta
-  n_\alpha(\mathbf{r}')}$ and $\frac{\delta
-  n_\alpha(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$.  The former is
-straightforward, given Equations~\ref{eq:Phi1}-\ref{eq:Phi3}, and we
-will write no more about it.  The functional derivatives of the
-fundamental measures, however, require a bit more subtlety, and we
-will address them here.
-
-We begin with the derivative of $n_3$, the filling fraction, which we
-will discuss in somewhat more detail than the remainder, which are
-similar in nature.  Because the diameter $\sigma(\rr)$ is the diameter
-of a sphere \emph{at position~$\rr$}, we write the fundamental measure
-$n_3(\rr')$ as
-\begin{align}
-  n_3(\rr') &= \int n(\rr'') \Theta\left(\frac{\sigma(\rr'')}{2}
-  -\left|\rr' - \rr''\right|\right)
-  d\rr''
-\end{align}
-where we note that $\sigma(\rr'')$ and $n(\rr'')$ are the diameter and
-density, respectively, of spheres centered at position~$\rr''$.  Thus the
-derivative with respect to the diameter of spheres at position
-$\rr$ is
-\begin{align}
-  \frac{\delta n_3(\rr')}{\delta \sigma(\rr)} &= \frac 12 \int n
-  (\rr'') \delta\left(\frac{\sigma(\rr'')}{2} -\left|\rr' - \rr''\right|\right)
-  \delta(\rr-\rr'') d\rr'' \\ &= n (\rr) \delta(\sigma(\rr)/2
-  -\left|\rr' - \rr\right|)
-\end{align}
-This pattern will hold for each fundamental measure: because we are
-seeking the change in free energy when spheres at point~$\rr$ are
-expanded, the integral over density is eliminated.  To compute the
-correlation funtion $g_\sigma^A$, we convolve this delta function with
-the product of the density and a local derivative of $\Phi(\rr)$:
-\begin{align}
-  \frac{\delta A_{HS}}{\delta \sigma(\rr)} &= \int \frac{\partial \Phi(\rr')}{\partial
-    n_3(\rr')}n(\rr') \delta(\sigma/2-|\rr'-\rr|)d\rr'
-  + \cdots
-\end{align}
-As we shall see, there are only four convolution kernels, leading to
-four additional convolutions beyond those required for FMT.
-
-The functional derivative of $n_2$ introduces our second convolution
-kernel, which is a derivative of the delta function.
-\begin{align}
-  %n_2(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d \rr''\\
-  \frac{\delta n_2(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
-The derivatives of the remaining scalar densities $n_1$ and $n_0$ reduce to
-sums of the terms above:
-\begin{multline}
-  %n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
-  \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
-  = \frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \\
-  -
-  \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{multline}
-and
-\begin{multline}
-  %n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
-  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
-  \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
-  = \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-  \\-
-  2 \frac{n(\rr)}{\pi
-    \sigma(\rr)^3}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{multline}
-
-The vector-weighted densities $\mathbf{n}_{V1}$ and $\mathbf{n}_{V2}$
-give terms analogous to those of $n_1$ and $n_2$:
-\begin{multline}
-  %\mathbf{n}_{V2}(\rr') &= \int n(\rr') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
-  %  \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}(\rr')}{\delta \sigma(\rr)} = -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-    \frac{\rr-\rr'}{|\rr-\rr'|}
-\end{multline}
-\begin{multline}
-  %\mathbf{n}_{V1}(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  %\delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
-  \frac{\delta \mathbf{n}_{V1}(\rr')}{\delta \sigma(\rr)}
-  = -\frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-  \\ +
-  \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-\end{multline}
-Thus there are four convolution kernels used in computing $g_\sigma^A$:
-one scalar and one vector delta function, and one scalar and one
-vector derivative of the delta function.
 
@@
 \end{document}
-
-\begin{widetext}
-
-We will show here the derivation of the $\beta\frac{\delta n_3^{A}(\mathbf{r}')}{\delta \sigma(\mathbf{r})/2}$, etc., terms for the asymmetric case.
-The derivation of the S correlation follows the same format using the symmetric conceptual approach introduced in \ref{sec:g-S} and the $\frac{\delta A_{HS}}{\delta n(\mathbf{r}')}$, etc., terms are straightforward (if tedious) derivatives.
-
-The derivation of $g_\sigma^A(\rr)$ requires that we express our weighted
-densities so that the radius of the weighting kernel is determined according
-to the position of the sphere in question, the location where the density
-itself is evaluated.  Thus for the derivative of $n_3^A$, we have
-\begin{align}
-  \frac{\delta n_3^{A} (\rr')}{\delta \sigma(\rr)} &=
-  \frac 12 \int n (\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \delta(\rr-\rr'') d\rr'' \\
-   &= n (\rr) \delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
-Similarly, for $n_2^A$, we find
-\begin{align}
-  n_2^{A}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d \rr''\\
-  \frac{\delta n_2^{A}(\rr')}{\delta \sigma(\rr)} &= \frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)\\
-\end{align}
-
-The derivatives of the remaining scalar densities can be reduced to
-sums of the terms above:
-\begin{align}
-  n_1(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
-  \frac{\delta n_1(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-  -
-  \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
-
-\begin{align}
-  n_0(\rr') &= \int \mathbf{dr''} \frac{n(\rr'')}{\pi \sigma(\rr'')^2}
-  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \\
-  \frac{\delta n_0(\rr')}{\delta \sigma(\rr)}
-  &= \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-  -
-  2 \frac{n(\rr)}{\pi
-    \sigma(\rr)^3}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-\end{align}
-The vector-weighted densities have no additional occurrences of $\sigma$.  Their functional derivatives come out similarly:
-\begin{align}
-  \mathbf{n}_{V2}^{A}(\rr') &= \int n(\rr') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
-    \frac{\rr'-\rr''}{|\rr'-\rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{A}(\rr')}{\delta \sigma(\rr)} &= -\frac 12 n(\rr) \delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|)
-    \frac{\rr-\rr'}{|\rr-\rr'|}\\
-\end{align}
-\begin{align}
-  \mathbf{n}_{V1}^A(\rr') &= \int d\rr'' \frac{n(\rr'')}{2\pi \sigma(\rr'')}
-  \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{\rr'-\rr''}{|\rr'-\rr''|}\\
-  \frac{\delta \mathbf{n}_{V1}^A(\rr')}{\delta \sigma(\rr)}
-  &= -\frac{n(\rr)}{4\pi
-    \sigma(\rr)}\delta'(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-  +
-  \frac{n(\rr)}{2\pi
-    \sigma(\rr)^2}\delta(\sigma(\rr)/2 -\left|\rr' - \rr\right|) \frac{\rr-\rr'}{|\rr-\rr'|}
-\end{align}
-
-\end{widetext}
-\end{document}
-
-\subsection{Expressions for $\beta\frac{\delta
-    n_3^{S}(\mathbf{r}')}{\delta \sigma(\mathbf{r})}$, etc. for the
-  symetric case}\label{appendix:g-S}
-
-The derivatation of the asymmetrically averaged correlation function
-$g_\sigma^S(\rr)$ requires that we express the free energy in terms of
-a set of weighted densities in which the radius of the weighting
-kernel is determined according to the position at which the weighted
-density is being evaluated.  This is less intuitive than the $A$ case,
-since \emph{nothing is there}!  However, in practice it corresponds to
-the picture shown in Figure~\ref{fig:contact}, which involves
-extending the radius for spheres just at the position touching a
-given spot, where we are evaluating the correlation function.  We
-define the weighted density $n_3^S$ as:
-\begin{align}
-  n_3^{S}(\rr') &= \int n(\rr'') \Theta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
-\end{align}
-In the symmetric case, the functional derivative is strikingly simple:
-\begin{align}
-   \frac{\delta n_3^{S} (\rr')}{\delta \sigma(\rr)} &=
-   \delta(\rr'-\rr) \frac 12 \int n (\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
-   \delta(\rr'-\rr) d\rr''
-   \\
-   &= \frac 12 \delta(\rr'-\rr) n_2^S(\rr')
-\end{align}
-The functional derivative does \emph{not} eliminate the integral,
-which simply changes from one weighted density to another.  The
-corresponding term in the free energy functional, once integrated,
-takes a simple form:
-\begin{align}
-  \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')}
-  \frac{\delta n_3^{S}(\rr')}{\delta \sigma(\rr)} d\rr'
-  &= \frac 12 \int \frac{\delta A_{HS}}{\delta n_3^{S}(\rr')} \delta(\rr'-\rr)
-  n_2^S(\rr') d\rr' \\
-  &= \frac 12 \frac{\delta A_{HS}}{\delta n_3^{S}(\rr)} n_2(\rr)
-\end{align}
-which means that we can compute $g_\sigma^S(\rr)$ using only
-convolutions of the density, in strong contrast with
-$g_\sigma^A(\rr)$, which requires convolutions of derivatives of the
-free energy density.
-
-The functional derivative of $n_2^S(\rr')$ is quite similar:
-\begin{align}
-  n_2^{S}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf
-  r''\\
-  \frac{\delta n_2^{S}(\rr')}{\delta \sigma(\rr)} &= \delta(\rr-\rr') \frac 12 \int n(\rr'')
-  \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
-  \\
-  &= \delta(\rr-\rr') \frac12 n_{2'}(\rr')
-\end{align}
-where we recognize that the convolution---while not one of the standard
-FMT weighted density---is a simple convolution of the density, which
-we define in Equation~\ref{n2p} as $n_{2'}$.  We can
-go on to compute the derivatives of the remaining weighted densities:
-\begin{align}
-  n_1^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''\\
-    \frac{\delta n_1^{S}(\rr')}{\delta \sigma(\rr)} &=
-    \delta(\rr-\rr') \left( \int \frac{n(\rr'')}{4\pi \sigma(\rr)}
-    \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
-    - \frac 12 n_0^s(\rr') \right)
-    \\
-    &= \delta(\rr-\rr')\left(
-    \frac{n_{2'}(\rr')}{4\pi\sigma} - \frac12n_0(\rr')
-    \right)
-\end{align}
-\begin{align}
-  n_0^{S}(\rr') &= \int \frac{n(\rr'')}{\pi \sigma(\rr')^2} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''\\
-    \frac{\delta n_0^{S}(\rr')}{\delta \sigma(\rr)}
-    &= \delta(\rr-\rr') \left( \frac 12 \int \frac{n(\rr'')}{\pi \sigma(\rr)^2}
-    \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) d\rr''
-    -
-    \frac{2}{\sigma} n_0^S(\rr') \right)
-    \\
-    &= \delta(\rr-\rr')\left(
-    \frac{n_{2'}(\rr')}{2\pi\sigma^2}
-    - \frac{2}{\sigma} n_0(\rr')
-    \right)
-\end{align}
-\begin{align}
-  \mathbf{n}_{V2}^{S}(\rr') &= \int n(\rr'') \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
-  \frac{(\rr' - \rr'')}{|\rr' - \rr''|} d \rr''\\
-  \frac{\delta \mathbf{n}_{V2}^{S}(\rr')}{\delta \sigma(\rr)} &= \delta(\rr - \rr')
-  \frac 12 \int n(\rr'') \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)
-  \frac{(\rr - \rr'')}{|\rr - \rr''|} d\rr''
-  \\
-  &= \delta(\rr-\rr')
-   \frac12\mathbf{n}_{V2'}(\rr')
-\end{align}
-\begin{align}
- \mathbf{n}_{V1}^{S}(\rr') &= \int \frac{n(\rr'')}{2\pi \sigma(\rr')} \delta(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|)d\mathbf r''
-   \frac{(\rr' - \rr'')}{|\rr' - \rr''|}\\
- \frac{\delta \mathbf{n}_{V1}^{S}(\rr')}{\delta \sigma(\rr)}
- &= \delta(\rr-\rr')\left(\int \frac{n(\rr'')}{4\pi \sigma(\rr)}
-   \delta'(\sigma(\rr'')/2 -\left|\rr' - \rr''\right|) \frac{(\rr - \rr'')}{|\rr - \rr''|}  d\rr''
-   - \frac{\mathbf{n}_{V2}^S(\rr')}{2 \pi \sigma(\rr)^2} \right)
-   \\
-   &= \delta(\rr-\rr')\left(
-   \frac{\mathbf{n}_{V2'}(\rr')}{4\pi \sigma}
-   - \frac{\mathbf{n}_{V2}(\rr')}{2 \pi \sigma^2}
-   \right)
-\end{align}
-
-The above equations involve just two additional weighted densities
-beyond those required for FMT: one scalar density $n_{2'}$ computed with a
-derivative of the Dirac delta function, and one analogous vector
-density $\mathbf{n}_{2V'}$, defined in Equations~\ref{n2p} and \ref{n2vp}.
-
-\end{widetext}
-
-\end{document}
-
-%% \subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Bear Mark I free energy $A_{HS}$}
-
-%%   First, let's rewrite the free energy in terms of a smaller set of weighted densities
-
-%%   \begin{align}
-%%     A_{HS} &= \int \left\{
-%%     -n_0 \ln\left( 1 - n_3\right)
-%%     + \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-%%     + (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-%%       n_3 + (1-n_3)^2 \ln(1-n_3)
-%%     }{
-%%       36\pi n_3^2(1-n_3)^2
-%%     }
-%%     \right\}
-%%   \end{align}
-
-%% Taking derivatives we have
-
-%%   \begin{align}
-%%     \frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')} &=
-%%     \frac{n_0(\mathbf{r}')}{1 - n_3(\mathbf{r}')}
-%%     + \frac{n_1n_2 - \mathbf{n}_{V1}\cdot\mathbf{n}_{V2}}{(1 -
-%%       n_3(\mathbf{r}'))^2}
-%%     - \frac{n_2^3 -
-%%       3n_2\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{36\pi}\left(
-%%     \frac{n_3^2-5n_3+2}{n_3^3(1-n_3)^3} + 2\frac{\ln(1-n_3)}{n_3^3}
-%%     \right)
-%%     \\
-%%     \frac{\delta A_{HS}}{\delta n_2(\mathbf{r}')} &=
-%%     \frac{n_1}{1-n_3}
-%%     + 3(n_2^2 - \mathbf{n}_{V2}\cdot\mathbf{n}_{V2})\frac{n_3 +
-%%       (1-n_3)^2\ln(1-n_3)}{
-%%       36\pi n_3^2(1-n_3)^2
-%%     }
-%%     \\
-%%     \frac{\delta A_{HS}}{\delta n_1(\mathbf{r}')} &= \frac{n_2}{1-n_3}
-%%     \\
-%%     \frac{\delta A_{HS}}{\delta n_0(\mathbf{r}')} &=
-%%     -\ln\left( 1 - n_3\right) \\
-%%     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V2}(\mathbf{r}')} &=
-%%     \frac{\mathbf{n}_{V1}}{1-n_3}
-%%     - 6 n_2 \mathbf{n}_{V2} \frac{n_3 +
-%%       (1-n_3)^2\ln(1-n_3)}{
-%%       36\pi n_3^2(1-n_3)^2
-%%     }    \\
-%%     \frac{\delta A_{HS}}{\delta \mathbf{n}_{V1}(\mathbf{r}')} &=
-%%     \frac{\mathbf{n}_{V2}}{1-n_3} \\
-%%   \end{align}
-
-%% \subsection{Expressions for $\beta\frac{\delta A_{HS}}{\delta n_3(\mathbf{r}')}$, etc. for the White Brear Mark II free energy $A_{HS}$}
-
-%% Similarly to the White Bear I case we first write
-
-%% \begin{align} 
-%% A_{HS} &= \int \left\{
-%%     -n_0 \ln\left(1 - n_3\right)
-%%     + \left(1 + \frac{1}{9}n_3^2 \phi_2(n_3)\right)   \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3}
-%%     + \left(1 - \frac{4}{9}n_3\phi_3(n_3)\right)\frac{n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}}{24\pi (1-n_3)^2}
-%%     \right\}
-%% \end{align}
-%%  where
-%%  \begin{align}
-%%    \phi_2 &= \left(6n_3 - 3n_3^2 + 6(1-n_3)\ln(1-n_3)\right)\frac{1}{n_3^3}\\
-%%    \phi_3 &= \left(6n_3 - 9n_3^2 + 6n_3^3 + 6(1-n_3)^2 \ln(1-n_3)\right)\frac{1}{4n_3^3}
-%%  \end{align}
-%% For the following derivitives we have
-%% \begin{align}
-%%    \frac{\delta \phi_2}{\delta n_3} &= \left((-6+4n_3)\ln(1-n_3) - 6n_3 + n_3^2\right)\frac{3}{n_3^4}\\
-%%    \frac{\delta \phi_3}{\delta n_3} &= \left(-6n_3 + 5n_3^2 + (-2n_3^2 + 8n_3 -6)\ln(1-n_3)\right)\frac{3}{4n_3^4}
-%%  \end{align}
-%% And using these
-%% \begin{align}
-%%   \frac{\delta A_{HS}}{\delta n_3} &= \psi _1 + \psi _2 + \psi _3
-%% \end{align}
-%% where
-%% \begin{align}
-%%   \psi_1 &= \frac{n_0}{1-n_3}\\
-%%   \psi_2 &= \frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{1-n_3}\left(\frac{2}{9}n_3\phi_2 + \frac{1}{9}n_3^2\frac{\delta \phi_2}{\delta n_3}\right) + \left(1 + \frac{1}{9}n_3^2\phi_2\right)\frac{n_1n_2 - \mathbf{n}_1\cdot \mathbf{n}_2}{(1-n_3)^2}\\
-%%   \psi_3 &= \frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^2}\left(-\frac{4}{9}\phi_3 - \frac{4}{9}n_3\frac{\delta \phi_3}{\delta n_3}\right) + 2\left(1 - \frac{4}{9}n_3\phi_3\right)\frac{n_2^3 - 3n_2\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1-n_3)^3}
-%% \end{align}
-%% We also have
-%% \begin{align}
-%%   \frac{\delta A_{HS}}{\delta n_2} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_1}{1-n_3} + (1 - \frac{4}{9}n_3\phi_3)\frac{3n_2^2 - 3\mathbf{n}_2 \cdot \mathbf{n}_2}{24\pi(1 - n_3)^2}\\
-%%   \frac{\delta A_{HS}}{\delta n_1} &= (1 + \frac{1}{9}n_3^2\phi_2)\frac{n_2}{1 - n_3}\\
-%%   \frac{\delta A_{HS}}{\delta n_0} &= -\ln(1 - n_3)\\
-%%   \frac{\delta A_{HS}}{\delta \mathbf{n}_1} &= -(1 + \frac{1}{9}n_3^2 \phi_2)\frac{\mathbf{n}_2}{1 - n_3}\\
-%%   \frac{\delta A_{HS}}{\delta \mathbf{n}_2} &= -(1 + \frac{1}{9}n_3^2\phi_2)\frac{\mathbf{n}_1}{1 - n_3} 
-%%   - (1 - \frac{4}{9}n_3\phi_3)\frac{6n_2\mathbf{n}_2}{24\pi(1 - n_3)^2}
-%% \end{align}

### 2012-12-07 papers/water-SAFT/paper.tex
some changes in response to referee on water paper Tests skipped!

@@
-\documentclass[a4paper,12pt,amsmath,amssymb,prb]{revtex4-1}
+\documentclass[twocolumn,amsmath,amssymb,prb]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 biology---takes place in aqueous solution.  However, while quantum
-chemistry enables us to calculate the ground state energies of
-large molecules in vacuum, prediction of the free energy of even the
+chemistry enables us to calculate the ground state energies of large
+molecules in vacuum, prediction of the free energy of even the
 smallest molecules in the presence of a solvent poses a continuing
@@
 dynamics~\cite{car1985, grossman2004}.  The current state-of-the art
-in \emph{ab initio} molecular dynamics is limited to a few hundred water
-molecules per unit cell~\cite{lewis2011doesnitric}.  On top of this,
-standard \emph{ab initio}
-methods using classical molecular dynamics without van der Waals
-corrections strongly over-structure water, to the point that ice melts
-at over 400K\cite{yoo2009phase}!  There has been a flurry of recent
-publications implicating van der Waals effects as significant in
+in \emph{ab initio} molecular dynamics is limited to a few hundred
+water molecules per unit cell~\cite{lewis2011doesnitric}.
+%
+On top of this, traditional density-functional theory (DFT) methods
+without the use of dispersion corrections strongly over-structure
+water, to the point that ice melts at over 120$^\circ$C\cite{yoo2009phase}!
+%
+There has been a flurry of recent
+publications implicating van der Waals effects (i.e. dispersion
+corrections) as significant in
 reducing this over-structuring\cite{lin2009importance,
-  wang2011density, mogelhoj2011ab, jonchiere2011van}.  It has also
+  wang2011density, mogelhoj2011ab, jonchiere2011van}.
+%
+However, one particular study found that water modeled using a
+hybrid functional with dispersion corrections \emph{still} has a
+melting point over 80$^\circ$C\cite{yoo2011}.
+%
+It has also
 been found that the inclusion of nuclear quantum effects can provide
@@
 \begin{equation}
-  A(T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int V_\textit{ext}(\rr) n(\rr)
+  \Omega (T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int (V_\textit{ext}(\rr) +\mu)n(\rr)
 d\rr\right\},
 \end{equation}
-where $A(T)$ is the Helmholtz free energy of a system in the external
-potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
-of atoms or molecules, and $F[n(\rr),T]$ is a
-universal free-energy functional for the fluid, which is independent
-of the external potential $V_\textit{ext}$.  Classical DFT is a
-natural framework for creating a more flexible theory of
-hydrophobicity that can readily describe interaction of water with arbitrary
-external potentials---such as potentials describing strong
-interactions with solutes or surfaces.
+where $\Omega (T)$ is the grand potential of a system in the
+external potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is
+the density of atoms or molecules, and $F[n(\rr),T]$ is a universal
+free-energy functional for the fluid, which is independent of the
+external potential $V_\textit{ext}$.  Classical DFT is a natural
+framework for creating a more flexible theory of hydrophobicity that
+can readily describe interaction of water with arbitrary external
+potentials---such as potentials describing strong interactions with
+solutes or surfaces.
 
@@
   rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000,
-  roth2002whitebear}.  These functionals are entirely expressed as an integral of
-local functions of a few convolutions of the density (fundamental
-measures) that can be efficiently computed.  We will use the White
-Bear version of the FMT functional\cite{roth2002whitebear}.  This
-functional reduces to the Carnahan-Starling equation of state in the
-homogeneous limit, and it reproduces the exact free energy in the
+  roth2002whitebear, yu2002structures}.  These functionals are
+entirely expressed as an integral of local functions of a few
+convolutions of the density (fundamental measures) that can be
+efficiently computed.  We will use the White Bear version of the FMT
+functional\cite{roth2002whitebear, yu2002structures}.  This functional
+reduces to the Carnahan-Starling equation of state in the homogeneous
+limit, and it reproduces the exact free energy in the
 strongly-confined limit of a small cavity.
@@
   gloor2004accurate, gloor2007prediction, Jaqaman2004,
-  clark2006developing, chuev2006,
-  lischner2010classical, fu2005vapor-liquid-dft,kiselev2006new,
-  blas2001examination}, each of which
-captures some of the qualitative behavior of water.  However, each of
-these functionals also fail to capture some of water's unique
+  clark2006developing, chuev2006, lischner2010classical,
+  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination}, each of
+which captures some of the qualitative behavior of water.  However,
+each of these functionals also fail to capture some of water's unique
 properties.  For instance, the functional of Lischner \emph{et
@@
 and Skolov\cite{chuev2006} uses an ad hoc modification of FMT that can
-predict hydrophobic hydration near temperatures of 298~K, but does not produce a
-correct equation of state due to their method producting a high value
-for pressure. A number of classical density functionals have recently been produced that are based on
-Statistical Associating Fluid Theory (SAFT)\cite{ 
+predict hydrophobic hydration near temperatures of 298~K, but does not
+produce a correct equation of state due to their method producting a
+high value for pressure. A number of classical density functionals
+have recently been produced that are based on Statistical Associating
+Fluid Theory (SAFT)\cite{ segura1997associating, segura1998comparison,
   yu2002fmt-dft-inhomogeneous-associating,
@@
 complex fluids in which hydrogen bonding plays a significant
-role\cite{muller2001molecular}.  SAFT is used to accurately model the
-equations of state of both pure fluids and mixtures over a wide range
-of temperatures and pressures.  SAFT is
+role\cite{chapman1989saft, muller2001molecular}.  SAFT is used to
+accurately model the equations of state of both pure fluids and
+mixtures over a wide range of temperatures and pressures.  SAFT is
 based on Wertheim's first-order thermodynamic perturbation theory
@@
 approximation\cite{gloor2002saft,clark2006developing, gloor2007prediction,
-gloor2004accurate, gross2009density, kahl2008modified,
+gloor2004accurate, kahl2008modified,
 blas2001examination, kiselev2006new}.
@@
 \begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx,
+  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{{l_Q}^3}} - 1\right) d\xx,
 \end{equation}
-where n(\xx) is the density of water molecules and $n_Q$ is the
-quantum concentration
+where n(\xx) is the density of water molecules and $l_Q$ is the
+thermal wavelength
 \begin{equation}\label{quantumconcentration}
- n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}.
+ l_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{1/2}.
 \end{equation}
@@
 the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
-fluid~\cite{roth2002whitebear}.  FMT functionals are expressed as the integral of
+fluid~\cite{roth2002whitebear, yu2002structures}.  FMT functionals are expressed as the integral of
 the \emph{fundamental measures} of a fluid, which provide local
-measures of quantities such as the filling fraction, density of
+measures of quantities such as the packing fraction, density of
 spheres touching a given point and mean curvature.  The hard-sphere
@@
 \end{align}
-The density $n_3$ is the filling fraction and $n_2$ describes the number
-of spheres touching a given point. For our functional for water, we use the 
-hard-sphere radius of
-3.03420~\AA, which was found to be optimal by Clark
-\emph{et al}.\cite{clark2006developing}
+The density $n_3$ is the packing fraction and $n_0$ is the
+average density at contact distance. For our functional for water, we
+use the hard-sphere diameter of 3.0342~\AA, which was found to be
+optimal by Clark \emph{et al}.\cite{clark2006developing}
 
@@
 reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
-of the filling fraction.  In order to apply this form to an
+of the packing fraction.  In order to apply this form to an
 \emph{inhomogeneous} density distribution, we construct an effective local
-filling fraction for dispersion $\etadisp$, given by a Gaussian
+packing fraction for dispersion $\etadisp$, given by a Gaussian
 convolution of the density:
@@
 \end{align}
-This effective filling fraction is used throughout the dispersion
-functional, and represents a filling fraction averaged over the
+This effective packing fraction is used throughout the dispersion
+functional, and represents a packing fraction averaged over the
 effective range of the dispersive interaction.  Here we have
@@
 attractive patches (``association sites'') on the surface of the hard
-sphere.  These four sites represent two protons and two electron lone
-pairs.  There is an attractive energy $\epsilonassoc$ when
-two molecules are oriented such that the proton of one overlaps
-with the lone pair of the other.  The volume over which this
-interaction occurs is $\kappaassoc$, giving the association
-term in the free energy two empirical parameters that are fit to
-the experimental equation of state of water (again, taken from
+sphere.  These four sites represent the two hydrogen bond donor sites,
+and two hydrogen bond acceptor sites.  There is an attractive energy
+$\epsilonassoc$ when two molecules are oriented such that the donor site
+of one overlaps with the acceptor site of the other.  The volume over
+which this interaction occurs is $\kappaassoc$, giving the association
+term in the free energy two empirical parameters that are fit to the
+experimental equation of state of water (again, taken from
 Clark~\emph{et al}\cite{clark2006developing}).
@@
 length scale for the association interaction.  In addition to the five
-empirical parameters of Clark \emph{et al}, we add a single additional dimensionless
-parameter $\lscale$---with a fitted value of 0.353---which determines
-the length scale over which the density is averaged when computing the
-dispersion free energy and its derivative.  We determine this final
-parameter by fitting the to the experimental surface tension with the
-result shown in Figure~\ref{fig:surface-tension}.  Because the SAFT
-model of Clark \emph{et al} overestimates the critical
-temperature---which is a common feature of SAFT-based functionals that
-do not explicitly treat the critical point---we cannot reasonably
-describe the surface tension at all temperatures, and choose to fit
-the surface tension at and around room temperature.
+empirical parameters of Clark \emph{et al}, we add a single additional
+dimensionless parameter $\lscale$---with a fitted value of
+0.353---which determines the length scale over which the density is
+averaged when computing the dispersion free energy and its derivative.
+We determine this final parameter by fitting the computed surface
+tension to the experimental surface tension with the result shown in
+Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
+\emph{et al} overestimates the critical temperature---which is a
+common feature of SAFT-based functionals that do not explicitly treat
+the critical point---we cannot reasonably describe the surface tension
+at all temperatures, and choose to fit the surface tension at and
+around room temperature.
 
@@
 different radii rods, as well as the prediction for the contact value
-of the density as a function of rod radius, as computed from the free
-energies plotted in Figure~\ref{fig:energy-vs-diameter}.  The
-agreement between these curves confirms that our functional satisfies
-the contact-value theorem and that our minimization is well converged.
-As expected, as the radius of the rods becomes zero the
-contact density approaches the bulk density, and as the radius becomes
-large, the contact density will approach the vapor density.
+of the density as a function of rod radius, as computed from the
+excess chemical potentials plotted in
+Figure~\ref{fig:energy-vs-diameter}.  The agreement between these
+curves confirms that our functional satisfies the contact-value
+theorem and that our minimization is well converged.  As expected, as
+the radius of the rods becomes zero the contact density approaches the
+bulk density, and as the radius becomes large, the contact density
+will approach the vapor density.
 

### 2012-12-10 papers/water-SAFT/paper.tex
changes to water paper Tests skipped!

@@
   clark2006developing, chuev2006, lischner2010classical,
-  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination}, each of
+  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination,
+  sundararaman2012computationally}, each of
 which captures some of the qualitative behavior of water.  However,
@@
 predict hydrophobic hydration near temperatures of 298~K, but does not
-produce a correct equation of state due to their method producting a
-high value for pressure. A number of classical density functionals
+produce a correct equation of state due to their method producing a
+high value for pressure.
+A number of classical density functionals
 have recently been produced that are based on Statistical Associating
@@
 do reproduce the temperature-dependence of water's properties.
+%
+We should give special mention to Sundararaman \emph{et al} recently
+introduced a classical density functional for water using an model in
+which a water molecule is seen as a hard sphere attached to two
+tetrahedrally oriented hard spheres representing voids, or
+orientations in which a hydrogen bond may not be formed, with all
+attractive interactions being lumped into a single pair potential
+treated in a mean field
+approximation\cite{sundararaman2012computationally}.
 
@@
 
-Functionals constructed using a local density approximation
-fail to satisfy the contact-value theorem, and
-therefore incorrectly model small hard solutes.
-The contact-value theorem states that the pressure any fluid exerts
-on a hard wall interface is proportional to the contact density of the
-fluid\cite{henderson1979exact}:
+Functionals constructed using a local density approximation fail to
+satisfy the contact-value theorem, and therefore incorrectly model
+small hard solutes.  The contact-value theorem relates the pressure on
+a hard wall interface to the contact density of the fluid at that
+interface:
 \begin{equation}\label{eq:contact}
@@
 where $n_c$ is the contact density, $k_B$ is the Boltzmann constant
-and $T$ is the temperature of
-the fluid. This leads to the property that the excess chemical
-potential of a small hard solute is proportional to the
+and $T$ is the temperature of the fluid.  The pressure in this formula
+is defined as the force per unit \emph{solvent accessible area}, which
+differs from the actual pressure when the surface is curved.  For a
+cavity in which the solvent is excluded from an arbitrarily small
+volume, the contact density will be the same as the bulk density, and
+therefore we can integrate the above pressure to find that the excess
+chemical potential of a small hard solute is proportional to the
 solvent-excluded volume:
@@
 \end{equation}
-The contact-value theorem is satisfied by classical density
-functionals in which the only purely local term is the ideal gas
-contribution to the free energy, and conversely, this theorem is not
-satisfied by functionals built using a local density approximation.
+The contact-value theorem is violated by local classical density
+functionals such as those using a local density approximation or a
+square-gradient term, but is satisfied by non-local classical density
+functionals, such as those using a weighted-density approach.
 
@@
 at all temperatures, and choose to fit the surface tension at and
-around room temperature.
+around room temperature.  We note here that although we could have
+chosen to fit the surface tension with a square-gradient term in the
+free energy, instead of adjusting the length scale for the dispersive
+attraction.  However, would have resulted in a functional that
+violates the contact-value theorem, which among other problems would
+fail to satisfy Equation~\ref{contactvaluethm} for the excess chemical
+potential of small solutes.
 
@@
 radius.  We define the hard-rods radius as the radius from which water
-is excluded.
-For rods with radius larger than 0.5~nm or so, this reaches a maximum value of
-75~mN/m, which is slightly higher than the bulk surface.  In the limit
-of very large rods, this value will decrease and approach the bulk
-surface value.  As seen in the inset of
-Fig.~\ref{fig:energy-vs-diameter}, for rods with very small radius (less than
-about 0.5~\AA) the excess chemical potential is proportional to volume,
-as required by the contact-value theorem (see Equation~\ref{contactvaluethm}).
+is excluded.  For rods with radius larger than 0.5~nm or so, this
+reaches a maximum value of 75~mN/m, which is slightly higher than
+macroscopic surface tension.  In the limit of very large rods, this
+value will decrease and approach the macroscopic value.  As seen in
+the inset of Fig.~\ref{fig:energy-vs-diameter}, for rods with very
+small radius (less than about 0.5~\AA) the excess chemical potential
+is proportional to volume, satisfying Equation~\ref{contactvaluethm},
+which results from the contact-value theorem.
 
@@
 free energy per length, which is equal to the circumference multiplied 
-by the surface tension. The force per length is the derivative of this
-with respect to the separation. 
-The circumference of the stadium-shape is
+by the surface tension. 
+Working out the circumference of the stadium-shape leads us to
 \begin{equation}
-C_{s} = 2\pi r +4r+2d
+F = (2\pi r +4r+2d)\gamma
 \end{equation}
-and so the force per length is equal to twice the surface tension.
+where $\gamma$ is the surface tension, $r$ is the radius of the rods,
+and $d$ is the separation between rods illustrated in
+Figure~\ref{fig:density-rods}.  The force per length is the derivative
+of the free energy with respect to the separation $d$, from which we
+conclude that the force per length is twice the
+surface tension.
 
@@
 similar to our hydrophobic rods, using molecular dynamics with the SPC
-model for water, and a Lennard-Jones potential for the interaction of
-carbon with water, for nanotubes of diameter 1.25~nm and separations
-ranging from about 0.3~nm to 1.5~nm.  The SPC model underestimates the
-surface tension of water by about 24\%\cite{vega2007surface}, so we
-cannot expect this work to provide quantitative agreement with real
-water.  Walther \emph{et al} observe a drying transition between the
-two nanotubes, which occurs at a smaller radius than our results
-suggest.  However, when Walther \emph{et al} disable the attractive
-interaction between nanotube and water, the drying effect occurs at
-much longer range, in agreement with our results.
+model for water, for nanotubes of diameter 1.25~nm and separations
+ranging from about 0.3~nm to 1.5~nm.  In agreement with our findings
+for two purely hydrophobic rods, Walther \emph{et al} find that in the
+absence of Lennard-Jones attraction between carbon and oxygen, there
+is a drying transition at a distance comparable to the diameter of the
+nanotube.  In contrast to this, when the attraction between nanotubes
+and water is turned on, they find that the drying transition occurs at
+much shorter distances, comparable to the diameter of water.
 
@@
   simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
-  The horizontal lines show the experimental and SPC/E bulk surface
+  The horizontal lines show the experimental and SPC/E macroscopic surface
   tension for water at standard atmospheric temperature and
@@
 (Figure~\ref{fig:sphere-energy-vs-diameter}).  This
-effective surface tension surpasses the bulk surface tension at a
-radius of almost 1~nm, and at large radius will drop to the bulk value.  As
+effective surface tension surpasses the macroscopic surface tension at a
+radius of almost 1~nm, and at large radius will drop to the macroscopic value.  As
 with the single rod, we see the analytically correct behavior in the
@@
 issue that the SPC/E model for water significantly underestimates the
-surface bulk tension of water at room
+macroscopic surface tension of water at room
 temperature\cite{vega2007surface}.

### 2012-12-10 papers/contact/paper.tex
a few edits on defining contact value theorem Tests skipped!

@@
 \begin{align}
-  p &= k_BT n_\textit{contact} \\
-  &= k_BT n g_\sigma
+  p(\rr_c) &= k_BT n(\rr_c)
+\end{align}
+where $\rr_c$ is a position of a sphere that would be in contact with
+the hard surface, and $p(\rr_c)$ is the pressure that the spheres
+exert on that surface at the point $\rr_c$.  This pressure is
+understood as the ratio of force to an infinitesimal element of
+\emph{solvent accessible surface} area.
+%
+In the homogeneous fluid, the contact-value theorem implies that
+\begin{align}
+  p &= k_BT n g_\sigma.
 \end{align}

### 2012-12-11 papers/water-SAFT/paper.tex
more edits to water paper Tests skipped!

@@
 We develop a classical density functional for water which combines the
-White Bear fundamental-measure theory (FMT)
-functional for the hard sphere fluid with
-attractive interactions based on the Statistical Associating Fluid
-Theory (SAFT-VR).
-This functional reproduces the properties of water at both long and
-short length scales over a wide range of temperatures, and is
-computationally efficient, comparable to the cost of FMT
-itself.  We demonstrate our functional by applying it to systems
-composed of two hard rods, four hard rods arranged in a square and
-hard spheres in water.
+White Bear fundamental-measure theory (FMT) functional for the hard
+sphere fluid with attractive interactions based on the Statistical
+Associating Fluid Theory (SAFT-VR).  This functional reproduces the
+properties of water at both long and short length scales over a wide
+range of temperatures, and is computationally efficient, comparable to
+the cost of FMT itself.  We demonstrate our functional by applying it
+to systems composed of two hard rods, four hard rods arranged in a
+square and hard spheres in water.
 \end{abstract}
@@
 \end{equation}
-where $\Omega (T)$ is the grand potential of a system in the
-external potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is
-the density of atoms or molecules, and $F[n(\rr),T]$ is a universal
-free-energy functional for the fluid, which is independent of the
-external potential $V_\textit{ext}$.  Classical DFT is a natural
-framework for creating a more flexible theory of hydrophobicity that
-can readily describe interaction of water with arbitrary external
-potentials---such as potentials describing strong interactions with
-solutes or surfaces.
+where $\Omega (T)$ is the grand potential of a system in the external
+potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
+of atoms or molecules, $\mu$ is the chemical potential and
+$F[n(\rr),T]$ is a universal free-energy functional for the fluid,
+which is independent of the external potential $V_\textit{ext}$.
+Classical DFT is a natural framework for creating a more flexible
+theory of hydrophobicity that can readily describe interaction of
+water with arbitrary external potentials---such as potentials
+describing strong interactions with solutes or surfaces.
 
@@
 % 
-We should give special mention to Sundararaman \emph{et al} recently
-introduced a classical density functional for water using an model in
-which a water molecule is seen as a hard sphere attached to two
-tetrahedrally oriented hard spheres representing voids, or
+We should give special mention to Sundararaman \emph{et al} who
+recently introduced a classical density functional for water using a
+model in which a water molecule is treated as a hard sphere attached to
+two tetrahedrally oriented hard spheres representing voids, or
 orientations in which a hydrogen bond may not be formed, with all
@@
 small hard solutes.  The contact-value theorem relates the pressure on
-a hard wall interface to the contact density of the fluid at that
-interface:
+a hard surface to the contact density of the fluid at that
+surface:
 \begin{equation}\label{eq:contact}
-  p = n_ck_BT,
+  p(\rr_c) = n(\rr_c)k_BT,
 \end{equation}
-where $n_c$ is the contact density, $k_B$ is the Boltzmann constant
-and $T$ is the temperature of the fluid.  The pressure in this formula
-is defined as the force per unit \emph{solvent accessible area}, which
-differs from the actual pressure when the surface is curved.  For a
-cavity in which the solvent is excluded from an arbitrarily small
+where $\rr_c$ is the position at which a molecule is in contact with
+the hard surface, $n(\rr_c)$ is the density at that point of contact,
+and $p(\rr_c)$ is the pressure that the fluid exerts on the surface at
+the same point.  This pressure is defined as a ratio of force to
+\emph{solvent accessible surface} area.
+%
+For a solute which excludes the solvent from an arbitrarily small
 volume, the contact density will be the same as the bulk density, and
@@
 at all temperatures, and choose to fit the surface tension at and
-around room temperature.  We note here that although we could have
-chosen to fit the surface tension with a square-gradient term in the
-free energy, instead of adjusting the length scale for the dispersive
-attraction.  However, would have resulted in a functional that
-violates the contact-value theorem, which among other problems would
-fail to satisfy Equation~\ref{contactvaluethm} for the excess chemical
-potential of small solutes.
+around room temperature.  We note here that we could have chosen to
+fit the surface tension with a square-gradient term in the free energy
+instead of adjusting the length scale for the dispersive attraction.
+This would result in a functional that violates the contact-value
+theorem which, among other problems, would fail to satisfy
+Equation~\ref{contactvaluethm} for the excess chemical potential of
+small solutes.
 
@@
 In Figure~\ref{fig:energy-vs-diameter} we show the excess chemical
-potential at room temperature, scaled by the solvent-accessible
+potential at room temperature, scaled by the solvent accessible
 surface area of the hard rod, plotted as a function of hard-rod
@@
 similar to our hydrophobic rods, using molecular dynamics with the SPC
-model for water, for nanotubes of diameter 1.25~nm and separations
-ranging from about 0.3~nm to 1.5~nm.  In agreement with our findings
-for two purely hydrophobic rods, Walther \emph{et al} find that in the
-absence of Lennard-Jones attraction between carbon and oxygen, there
-is a drying transition at a distance comparable to the diameter of the
-nanotube.  In contrast to this, when the attraction between nanotubes
-and water is turned on, they find that the drying transition occurs at
-much shorter distances, comparable to the diameter of water.
+model for water.  Their simulations used nanotubes of diameter 1.25~nm
+and separations ranging from about 0.3~nm to 1.5~nm.  In agreement
+with our findings for two purely hydrophobic rods, Walther \emph{et
+  al} find that in the absence of Lennard-Jones attraction between
+carbon and oxygen, there is a drying transition at a distance
+comparable to the diameter of the nanotube.  In contrast to this, when
+the attraction between nanotubes and water is turned on, they find
+that the drying transition occurs at much shorter distances,
+comparable to the diameter of water.
 
@@
 examining the ratio of the excess chemical potential of the cavity
-system to the solvent-accessible surface area
+system to the solvent accessible surface area
 (Figure~\ref{fig:sphere-energy-vs-diameter}).  This

### 2012-12-12 papers/contact/paper.tex
Another big edit after third referee response.  Added svg figures and changed some text in paper Eq 17,18,24 Tested everything on kipu.

@@
   them from the White Bear version of the Fundamental Measure Theory
-  functional~\cite{roth2002whitebear}, using an assumption of
+  functional, using an assumption of
   thermodynamic consistency. We test these formulas, as well as two
-  existing formulas~\cite{yu2002fmt-dft-inhomogeneous-associating,
-    gross2009density} against Monte Carlo simulations, and find
+  existing formulas against Monte Carlo simulations, and find
   excellent agreement between the Monte Carlo data and one of our
@@
 where $\rr_c$ is a position of a sphere that would be in contact with
-the hard surface, and $p(\rr_c)$ is the pressure that the spheres
-exert on that surface at the point $\rr_c$.  This pressure is
+the hard surface, $n(\rr_c)$ is the density at this point $\rr_c$, and $p(\rr_c)$ is the pressure that the spheres
+exert on the surface at the same point.  This pressure is
 understood as the ratio of force to an infinitesimal element of
-\emph{solvent accessible surface} area.
+\emph{solvent accessible surface} (SAS) area.
 %
@@
 \begin{align}
-  p &= k_BT n g_\sigma.
+  p_{\textit{SAS}} &= k_BT n g_\sigma
 \end{align}
-Since we are interested in the correlation function at the surface at
-which hard spheres contact, we need to compute the pressure on that
-surface.   This pressure can be
+where $p_{SAS}$ is the pressure on a hard sphere's solvent accessible surface.
+This pressure can be
 readily computed from the dependence of the Carnahan-Starling free
-energy on hard sphere radius.
+energy on hard sphere radius:
 \begin{align}
@@
 where $\eta \equiv \frac{\pi}{6} \sigma^3 n$ is the filling fraction.
-We compute the pressure on each sphere using the total force with
+We compute the pressure using the total force with
 respect to a changing radius of all the spheres.
-To find the pressure, we divide by $4\pi \sigma^2$, which is the area
-over which contact is possible, as defined by the centers of the
-contacting spheres, as illustrated in Figure~\ref{fig:nA}.
+To find the pressure, we divide this force by $4\pi \sigma^2$, which
+is the SAS area of a single hard sphere, illustrated in Figure~\ref{fig:nA}.
 Finally, we divide by $N$ to account for the total area of all the
@@
 \begin{align}
-  p_\sigma &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
+  p_{\textit{SAS}} &= \frac{1}{N 4\pi \sigma^2} \frac{dA_{HS}}{dR} \\
   &= \frac{1}{N 4\pi \sigma^2} \frac12 \frac{dA_{HS}}{d\sigma} \\
@@
 \end{align}
-Using the contact-value theorem, we can thus find the well-known
+Using the contact-value theorem, we thus find the well-known
 correlation function evaluated at contact.
@@
 the two spheres is located at position~$\rr$.  This correlation
-function is related to the contact density averaged over the surface
+function is related to the contact density averaged over the solvent accessible surface
 of a sphere located at~$\rr$, and can thus be determined by finding
-the pressure on that sphere.  We find this pressure from the change in
-free energy resulting from an infinitesimal expansion of any
-spheres located at position~$\rr$.  From this pressure, we derive a
+the pressure on that surface.  We find this pressure from the change in
+free energy resulting from an infinitesimal expansion of spheres
+located at position~$\rr$.  From this pressure, we derive a
 formula for the correlation function $g_\sigma^A(\rr)$ as was done in
@@
 \begin{align}
-  p_\sigma(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2}\frac12 \frac{\delta
+  p_{\textit{SAS}}(\mathbf{r}) &= \frac{1}{n(\rr) 4\pi \sigma^2}\frac12 \frac{\delta
     A_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:p_{HS}^A}\\
@@
 \end{align}
+where $\sigma(\mathbf{r})$ is the diameter of spheres located at position $\rr$.
+Details regarding the evaluation of the functional derivative $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
+are discussed in the Appendix.
 Equation~\ref{eq:g-A-exact} is an exact expression for
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-10}
-  \caption{Density and correlation function at a hard wall with bulk
-    filling fraction of 0.1.}
-  \label{fig:walls-10}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/walls-10}
+%%   \caption{Density and correlation function at a hard wall with bulk
+%%     filling fraction of 0.1.  In the top and middle panels are plotted
+%%     the asymmetrically averaged correlation function $g_\sigma^A$
+%%     (defined in Equation~\ref{eq:gA}) and the symmetrically averaged
+%%     correlation function $g_\sigma^S$ (defined in
+%%     Equation~\ref{eq:gS}).  The results of Monte Carlo, our functional,
+%%     and one previously published functional\cite{gross2009density,
+%%       yu2002fmt-dft-inhomogeneous-associating} are compared in each
+%%     case.  In the bottom panel we plot the density, computed with
+%%     Monte Carlo and with DFT.}
+%%   \label{fig:walls-10}
+%% \end{figure}
 
@@
 simulation with one of a hard sphere at the origin with diameter
-$\sigma$.  The second is an excluded sphere with diameter $6\sigma$,
+$\sigma$.  The
+second is an excluded sphere with diameter $6\sigma$,
 which demonstrates behavior typical of mildly convex hard surfaces.
 Finally, we study a concave surface given by a hard cavity in which
-our fluid is free to move up to a diameter of $16\sigma$, which
+our fluid is free to move up to a diameter of $8\sigma$, which
 demonstrates behavior typical of mildly concave surfaces.  In each
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-10}
-  \caption{Density and correlation function around a hard sphere with
-    diameter the same as those in the fluid, with a bulk filling
-    fraction of 0.1.}
-  \label{fig:inner-4-10}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/inner-4-10}
+%%   \caption{(Color online) Density and correlation function around a hard sphere with
+%%     diameter the same as those in the fluid, with a bulk filling
+%%     fraction of 0.1.}
+%%   \label{fig:inner-4-10}
+%% \end{figure}
 
@@
 filling fraction of 0.1, which are shown in
-Figures~\ref{fig:walls-10}-\ref{fig:outer-10}.  At this low density,
+Figure~\ref{fig:low-density}.  At this low density,
 the contact value of the correlation function in the bulk is only 1.3,
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-10}
-  \caption{Density and correlation function around a hard sphere with
-    an excluded diameter of $6\sigma$, with a bulk filling fraction of
-    0.1.}
-  \label{fig:inner-12-10}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/inner-12-10}
+%%   \caption{Density and correlation function around a hard sphere with
+%%     an excluded diameter of $6\sigma$, with a bulk filling fraction of
+%%     0.1.}
+%%   \label{fig:inner-12-10}
+%% \end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-16-01}
-  \caption{Density and correlation functions near the surface of a
-    spherical cavity with diameter $16\sigma$ at a bulk filling
-    fraction of 0.1.}
-  \label{fig:outer-10}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/outer-16-01}
+%%   \caption{Density and correlation functions near the surface of a
+%%     spherical cavity with diameter $16\sigma$ at a bulk filling
+%%     fraction of 0.1.}
+%%   \label{fig:outer-10}
+%% \end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/walls-40}
-  \caption{Density and correlation function at a hard wall with bulk
-    filling fraction of 0.4.}
-  \label{fig:walls-40}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/walls-40}
+%%   \caption{Density and correlation function at a hard wall with bulk
+%%     filling fraction of 0.4.}
+%%   \label{fig:walls-40}
+%% \end{figure}
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-4-40}
-  \caption{Density and correlation function around a hard sphere with
-    diameter the same as those in the fluid, with a bulk filling
-    fraction of 0.4.}
-  \label{fig:inner-4-40}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/inner-4-40}
+%%   \caption{Density and correlation function around a hard sphere with
+%%     diameter the same as those in the fluid, with a bulk filling
+%%     fraction of 0.4.}
+%%   \label{fig:inner-4-40}
+%% \end{figure}
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/inner-12-40}
-  \caption{Density and correlation function around a hard sphere with
-    an excluded diameter of $6\sigma$, with a bulk filling fraction of
-    0.4.}
-  \label{fig:inner-12-40}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/inner-12-40}
+%%   \caption{Density and correlation function around a hard sphere with
+%%     an excluded diameter of $6\sigma$, with a bulk filling fraction of
+%%     0.4.}
+%%   \label{fig:inner-12-40}
+%% \end{figure}
+
+\begin{figure*}
+  \noindent\includegraphics[width=0.8\textwidth]{figs/low-density}
+  %% \noindent
+  %% \begin{tabular}{cc}
+  %%   (a)\includegraphics[width=8cm]{figs/walls-10} &
+  %%   (c)\includegraphics[width=8cm]{figs/inner-4-10}\\
+  %%   (b)\includegraphics[width=8cm]{figs/inner-12-10} &
+  %%   (d)\includegraphics[width=8cm]{figs/outer-16-01}
+  %% \end{tabular}
+  \caption{ (Color online) Density and correlation function in systems
+    with a ``low density'' bulk filling fraction of 0.1.  The
+    subplots each show a different system: (a) next to a flat hard wall, (b)
+    around a hard sphere with an excluded diameter of $6\sigma$, (c)
+    around a hard sphere with an excluded diameter of $2\sigma$, and
+    (d) within a spherical cavity with an included diameter of $8\sigma$.
+    In the top and middle panels of each subfigure are the
+    asymmetrically averaged correlation function $g_\sigma^A$ (defined
+    in Equation~\ref{eq:gA}) and the symmetrically averaged
+    correlation function $g_\sigma^S$ (defined in
+    Equation~\ref{eq:gS}).  The results of Monte Carlo, our
+    functional, and one previously published
+    functional\cite{gross2009density,
+      yu2002fmt-dft-inhomogeneous-associating} are compared in each
+    case.  The bottom panels show the density computed with
+    Monte Carlo and with DFT.}
+  \label{fig:low-density}
+\end{figure*}
+
+\begin{figure*}
+  \noindent\includegraphics[width=0.8\textwidth]{figs/high-density}
+  %% \noindent
+  %% \begin{tabular}{cc}
+  %%   (a)\includegraphics[width=8cm]{figs/walls-40} &
+  %%   (c)\includegraphics[width=8cm]{figs/inner-4-40}\\
+  %%   (b)\includegraphics[width=8cm]{figs/inner-12-40} &
+  %%   (d)\includegraphics[width=8cm]{figs/outer-16-04}
+  %% \end{tabular}
+  \caption{ (Color online) Density and correlation function in systems
+    with a ``high density'' bulk filling fraction of 0.4.  The
+    subplots each show a different system: (a) next to a flat hard wall, (b)
+    around a hard sphere with an excluded diameter of $6\sigma$, (c)
+    around a hard sphere with an excluded diameter of $2\sigma$, and
+    (d) within a spherical cavity with an included diameter of $8\sigma$.
+    In the top and middle panels of each subfigure are the
+    asymmetrically averaged correlation function $g_\sigma^A$ (defined
+    in Equation~\ref{eq:gA}) and the symmetrically averaged
+    correlation function $g_\sigma^S$ (defined in
+    Equation~\ref{eq:gS}).  The results of Monte Carlo, our
+    functional, and one previously published
+    functional\cite{gross2009density,
+      yu2002fmt-dft-inhomogeneous-associating} are compared in each
+    case.  The bottom panels show the density computed with
+    Monte Carlo and with DFT.}
+  \label{fig:high-density}
+\end{figure*}
 
@@
 correlation function of 3.7, as seen in
-Figures~\ref{fig:walls-40}-\ref{fig:outer-40}.  This results in larger
+Figure~\ref{fig:high-density}.  This results in larger
 oscillations in the density at the hard surfaces, and correspondingly
@@
 it did at lower density.  The discrepancies are largest in the case of
-the spherical cavity, in which the DFT considerably underestimates the
+the spherical cavity (Figure~\ref{fig:high-density}d), in which the DFT considerably underestimates the
 range of the density oscillations.
@@
 the Monte Carlo data, reproducing quite well the structure near the
-interface in each configuration, although in the spherical cavity
-(Figure~\ref{fig:outer-40}), there is a small, but significant
+interface in each configuration, although in the spherical cavity there is a small, but significant
 discrepancy, comparable to the discrepancy found in the density
@@
 underestimates the correlation at the interface, at one extreme by
-40\% in the case of the spherical cavity
-(Figure~\ref{fig:outer-40}), and at the other extreme by 15\%
-in the test-particle scenario (Figure~\ref{fig:inner-4-40}).
+40\% in the case of the spherical cavity, and at the other extreme by 15\%
+in the test-particle scenario (Figure~\ref{fig:high-density}c).
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/outer-16-04}
-  \caption{Density and correlation functions near the surface of a
-    spherical cavity with diameter $16\sigma$ at a bulk filling
-    fraction of 0.4.}
-  \label{fig:outer-40}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/outer-16-04}
+%%   \caption{Density and correlation functions near the surface of a
+%%     spherical cavity with diameter $16\sigma$ at a bulk filling
+%%     fraction of 0.4.}
+%%   \label{fig:outer-40}
+%% \end{figure}
 
@@
 low density, both functionals give significant and finite values in
-the region in which there are no contacts, but at high density they
+the region in which there are no contacts (Figures~\ref{fig:low-density}a), but at high density they
 also miss the large oscillations that are present near the flat wall
-and the concave surface (Figures~\ref{fig:walls-40}
-and~\ref{fig:outer-40}).  As was the case at low density, the
+and the concave surface (Figures~\ref{fig:high-density}a).  As was the case at low density, the
 functional of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}

### 2012-12-12 papers/contact/paper.tex
move contact figures around Tested everything on kipu.
more changes for resubmission of contact Tested everything on kipu.

@@
 where $n^A(\rr)$ is a weighted density defined in Equation~\ref{eq:nA}
-and $g_\sigma^A(\rr)$ the local value of the correlation function at
+and $g_\sigma^A(\rr)$ is the local value of the correlation function at
 contact, which we define in Equation~\ref{eq:gA}.  In
@@
 \begin{align}
-  p(\rr_c) &= k_BT n(\rr_c)
+  p_{\textit{SAS}}(\rr_c) &= k_BT n(\rr_c)
 \end{align}
 where $\rr_c$ is a position of a sphere that would be in contact with
-the hard surface, $n(\rr_c)$ is the density at this point $\rr_c$, and $p(\rr_c)$ is the pressure that the spheres
+the hard surface, $n(\rr_c)$ is the density at this point $\rr_c$, and
+$p_{\textit{SAS}}(\rr_c)$ is the pressure that the spheres
 exert on the surface at the same point.  This pressure is
@@
 where $\sigma(\mathbf{r})$ is the diameter of spheres located at position $\rr$.
-Details regarding the evaluation of the functional derivative $\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
+Details regarding the evaluation of the functional derivative
+$\frac{\delta A_{HS}}{\delta \sigma(\mathbf{r})}$
 are discussed in the Appendix.
@@
 we do not know the exact hard-sphere free energy, we approximate
-$A_{HS}$ using FMT.  Details of evaluating the functional derivative
-in Equation~\ref{eq:g-A-exact} using FMT are given in the appendix.
-The equation for $g_\sigma^A$ requires finding convolutions of local
-derivatives of the free energy, making this formulation
-computationally somewhat more expensive than the free energy itself.
+$A_{HS}$ using FMT.
+The equation for $g_\sigma^A$ found using FMT requires finding
+convolutions of local derivatives of the free energy, making this
+formulation computationally somewhat more expensive than the free
+energy itself.
 
@@
 smooth, but we can state empirically making this approximation
-nevertheless improves the predictions of our functional, while at the
+nevertheless improves the predictions of our functional $g_\sigma^S$, while at the
 same time reducing its computational cost by avoiding the need to
@@
 Finally, we study a concave surface given by a hard cavity in which
-our fluid is free to move up to a diameter of $8\sigma$, which
+our fluid is free to move up to a diameter of $16\sigma$, which
 demonstrates behavior typical of mildly concave surfaces.  In each
@@
 %%   \label{fig:inner-4-10}
-%% \end{figure}
-
-\subsection{Low density}
-
-We begin by presenting our low-density results, corresponding to a
-filling fraction of 0.1, which are shown in
-Figure~\ref{fig:low-density}.  At this low density,
-the contact value of the correlation function in the bulk is only 1.3,
-indicating that correlations are indeed small and that the fluid should be
-relatively easy to model.  Indeed, the contact density at the hard
-surface is only around 50\% higher than the bulk, and the FMT
-predicted density is close to indistinguishable from the true
-density for each of the four configurations.
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/inner-12-10}
-%%   \caption{Density and correlation function around a hard sphere with
-%%     an excluded diameter of $6\sigma$, with a bulk filling fraction of
-%%     0.1.}
-%%   \label{fig:inner-12-10}
-%% \end{figure}
-
-The $g_\sigma^A$ correlation function in each configuration is very
-flat, with only small, smooth changes as the surface is approached.
-Our functional $g_\sigma^A$ very closely matches the Monte Carlo
-predictions in each case, while that of Gross consistently
-underestimates the correlation at the interface by a significant margin.  We note that the theoretical curves
-extend into the region from which the fluid is excluded.  This value
-corresponds to the correlation function that would be observed in the
-vanishingly unlikely scenario in which there was a sphere present at
-that location.  Naturally, we are unable to observe this quantity in
-our Monte Carlo simulations.
-
-The $g_\sigma^S$ correlation function shows considerably more
-structure, as well as additional variation due to the curvature of the
-hard surface.  The symmetric correlation function is nonzero at
-locations where spheres may touch, which for a convex hard surface
-means that $g_\sigma^S$ may be nonzero in the volume in which hard
-spheres are excluded.  In every configuration studied, the agreement
-between the theoretical predictions and the Monte Carlo simulation in
-each case is very poor in the region where there should be no contacts
-at all.  Because $n_0$ is comparable to its bulk value in this region,
-this means that these functionals predict a significant number of
-contacts in the region where there should be none.  The correlation
-function of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
-and ours described in Section~\ref{sec:g-S} give similar results, with
-slightly larger errors in our prediction.
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/outer-16-01}
-%%   \caption{Density and correlation functions near the surface of a
-%%     spherical cavity with diameter $16\sigma$ at a bulk filling
-%%     fraction of 0.1.}
-%%   \label{fig:outer-10}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-10}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.1.}
-%%   \label{fig:test-particle-10}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/test-particle-40}
-%%   \caption{DFT and monte-carlo nA compared with averaged test particle data with bulk filling
-%%     fraction of 0.4.}
-%%   \label{fig:test-particle-40}
-%% \end{figure}
-
-\subsection{High density}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/walls-40}
-%%   \caption{Density and correlation function at a hard wall with bulk
-%%     filling fraction of 0.4.}
-%%   \label{fig:walls-40}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/inner-4-40}
-%%   \caption{Density and correlation function around a hard sphere with
-%%     diameter the same as those in the fluid, with a bulk filling
-%%     fraction of 0.4.}
-%%   \label{fig:inner-4-40}
-%% \end{figure}
-
-%% \begin{figure}
-%%   \includegraphics[width=\columnwidth]{figs/inner-12-40}
-%%   \caption{Density and correlation function around a hard sphere with
-%%     an excluded diameter of $6\sigma$, with a bulk filling fraction of
-%%     0.4.}
-%%   \label{fig:inner-12-40}
 %% \end{figure}
@@
     around a hard sphere with an excluded diameter of $2\sigma$, and
-    (d) within a spherical cavity with an included diameter of $8\sigma$.
-    In the top and middle panels of each subfigure are the
+    (d) within a spherical cavity with an included diameter of $16\sigma$.
+    In the top and middle panels of each subfigure respectively are the
     asymmetrically averaged correlation function $g_\sigma^A$ (defined
@@
 \end{figure*}
+
+\subsection{Low density}
+
+We begin by presenting our low-density results, corresponding to a
+filling fraction of 0.1, which are shown in
+Figure~\ref{fig:low-density}.  At this low density,
+the contact value of the correlation function in the bulk is only 1.3,
+indicating that correlations are indeed small and that the fluid should be
+relatively easy to model.  Indeed, the contact density at the hard
+surface is only around 50\% higher than the bulk, and the FMT
+predicted density is close to indistinguishable from the true
+density for each of the four configurations, as seen in the bottom
+subpanel of each subfigure within Figure~\ref{fig:low-density}.
+
+The $g_\sigma^A$ correlation function in each configuration (plotted
+in the top panel of each subfigure within Figure~\ref{fig:low-density}) is
+very flat, with only small, smooth changes as the surface is approached.
+Our functional $g_\sigma^A$ very closely matches the Monte Carlo
+predictions in each case, while that of Gross consistently
+underestimates the correlation at the interface by a significant margin.  We note that the theoretical curves
+extend into the region from which the fluid is excluded.  This value
+corresponds to the correlation function that would be observed in the
+vanishingly unlikely scenario in which there was a sphere present at
+that location.  Naturally, we are unable to observe this quantity in
+our Monte Carlo simulations.
+
+The $g_\sigma^S$ correlation function (plotted
+in the middle panels of Figure~\ref{fig:low-density}) shows considerably more
+structure, as well as additional variation due to the curvature of the
+hard surface.  The symmetric correlation function is nonzero at
+locations where spheres may touch, which for a convex hard surface
+means that $g_\sigma^S$ may be nonzero in the volume in which hard
+spheres are excluded.  In every configuration studied, the agreement
+between the theoretical predictions and the Monte Carlo simulation in
+each case is very poor in the region where there should be no contacts
+at all.  Because $n_0$ is comparable to its bulk value in this region,
+this means that these functionals predict a significant number of
+contacts in the region where there should be none.  The correlation
+function of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}
+and ours described in Section~\ref{sec:g-S} give similar results, with
+slightly larger errors in our prediction.
 
@@
 
+\subsection{High density}
+
 At a higher density corresponding to a filling fraction of 0.4,
 correlations are much stronger, with the bulk contact value of the
-correlation function of 3.7, as seen in
-Figure~\ref{fig:high-density}.  This results in larger
-oscillations in the density at the hard surfaces, and correspondingly
-more interesting behavior in the correlation function near the
-interface.  The density predicted by the White Bear functional agrees
-reasonably well with the simulation results, although not so well as
-it did at lower density.  The discrepancies are largest in the case of
-the spherical cavity (Figure~\ref{fig:high-density}d), in which the DFT considerably underestimates the
-range of the density oscillations.
+correlation function of 3.7, as seen in Figure~\ref{fig:high-density}.
+This results in larger oscillations in the density at the hard
+surfaces, and correspondingly more interesting behavior in the
+correlation function near the interface, as shown in the bottom panels
+of the plots in Figure~\ref{fig:high-density}.  The density predicted
+by the White Bear functional agrees reasonably well with the
+simulation results, although not so well as it did at lower density.
+The discrepancies are largest in the case of the spherical cavity
+(Figure~\ref{fig:high-density}d), in which the DFT considerably
+underestimates the range of the density oscillations.
 
-The asymmetric version of the correlation function once again displays
+The asymmetric version of the correlation function (plotted
+in the top panels of Figure~\ref{fig:high-density}) once again displays
 relatively smooth behavior with a few small
@@
 underestimates the correlation at the interface, at one extreme by
-40\% in the case of the spherical cavity, and at the other extreme by 15\%
+40\% in the case of the spherical cavity
+(Figure~\ref{fig:high-density}d), and at the other extreme by 15\%
 in the test-particle scenario (Figure~\ref{fig:high-density}c).
@@
 
-The symmetrically averaged correlation function shows considerably
+The symmetrically averaged correlation function (plotted
+in the middle panels of Figure~\ref{fig:high-density}) shows considerably
 more structure near the interface at high density, and this structure
@@
 low density, both functionals give significant and finite values in
-the region in which there are no contacts (Figures~\ref{fig:low-density}a), but at high density they
+the region in which there are no contacts, but at high density they
 also miss the large oscillations that are present near the flat wall
-and the concave surface (Figures~\ref{fig:high-density}a).  As was the case at low density, the
+and the concave surface (Figures~\ref{fig:high-density}a
+and~\ref{fig:high-density}d).  As was the case at low density, the
 functional of Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}

### 2012-12-13 papers/water-SAFT/paper.tex

@@
   Corvallis, OR 97331, USA}
-\author{Eric Krebs}
+\author{Eric J. Krebs}
 \affiliation{Department of Physics, Oregon State University,
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms-truncated}
+\includegraphics[width=3.5in]{figs/pressure-with-isotherms-truncated}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
+\includegraphics[width=3.5in]{figs/surface-tension}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
+\includegraphics[width=3.5in]{figs/energy-vs-diameter}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-single-rod}
+\includegraphics[width=3.5in]{figs/density-single-rod}
 \end{center}
@@
 %\begin{center}
-%\includegraphics[width=\columnwidth]{figs/rods-diagram}
+%\includegraphics[width=3.5in]{figs/rods-diagram}
 %\end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
+\includegraphics[width=3.5in]{figs/density-rods-in-water}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
+\includegraphics[width=3.5in]{figs/rods-energy-vs-distance}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/four-rods-energy-vs-distance}
+\includegraphics[width=3.5in]{figs/four-rods-energy-vs-distance}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-four-rods}
+\includegraphics[width=3.5in]{figs/density-four-rods}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/sphere-energy-vs-diameter}
+\includegraphics[width=3.5in]{figs/sphere-energy-vs-diameter}
 \end{center}
@@
 \begin{center}
-\includegraphics[width=\columnwidth]{figs/density-sphere}
+\includegraphics[width=3.5in]{figs/density-sphere}
 \end{center}

### 2012-12-21 papers/water-SAFT/paper.tex

@@
 \begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{{l_Q}^3}} - 1\right) d\xx,
+  F_{id}[n] = k_B T \int n(\xx)\left( \ln({n(\xx){\Lambda}^3}) - 1\right) d\xx,
 \end{equation}
-where n(\xx) is the density of water molecules and $l_Q$ is the
+where n(\xx) is the density of water molecules and $\Lambda$ is the
 thermal wavelength
-\begin{equation}\label{quantumconcentration}
- l_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{1/2}.
-\end{equation}
+%\begin{equation}\label{quantumconcentration}
+$ \Lambda = \left(\frac{2\pi\hbar^2}{mk_BT}\right )^{1/2}$.
+%\end{equation}
 The ideal gas free energy functional on its own satisfies the contact

### 2013-01-18 papers/polyhedra/paper.tex
add sketchy polyhedra paper Tested everything on darcy.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
\usepackage{graphicx}% Include figure files
%\usepackage{breqn}
\usepackage{color}

\newcommand{\fixme}[1]{\red{[#1]}}

\begin{document}
\title{Fundamental measure theory for hard polyhedra.}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}
%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We develop a classical density functional for hard polyhedra.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Why study hard polyhedra? Hard spheres are the standard reference
system for essentially all fluids, but always have 12-fold
coordination.  A few very interesting liquids have tetrahedral
coordination, which is particularly significant when we consider the
freezing transition.  By studying hard polyhedra, we potentially can
develop a reference functional that can incorporate orientational
effects.

There have been several studies of non-spherical hard shapes.  At
least one omitted the rotational degrees of freedom (parallel hard
cubes, add citation).  At least one included the rotational degrees of
freedom, treating each orientation as a distinct species, resulting in
a density that exists in six dimensions rather than three, which was
applied to systems with a high aspect ratio, which undergo a nematic
transition.  It seems a good thing to have a density functional that
is more efficient, with fewer degrees of freedom to optimize.  In
particular, we are optimistic that one can do better for ``almost
spherical'' polyhedra.

\section{Main ideas}

\newcommand\rr{\mathbf{r}}
\newcommand\rhat{\mathbf{\hat{r}}}
\newcommand\rot{\mathbf{\varpi}}

We will begin with the extension of fundamental measure theory (FMT)
to non-spherical particles by (so and so: add citation).  This theory
is based around a density that is a function of both position $\rr$
and orientation $\rot$.  The weighted densities are defined for a
given orientation, and then rotated appropriately when computing the
fundamental measures:
\begin{align}
  n_\alpha(\rr) &= \int n(\rr',\rot) w_\alpha(R(\rot)(\rr-\rr')) d\rr' d\rot
\end{align}
where $R(\rot)$ is the rotation operator corresponding to the
orientation $\rot$, and $w_\alpha(\rr)$ is the fundamental measure of
a single particle in the reference orientation at the origin.

Given these fundamental measures, the free energy is computed with
the standard Rosenfeld formula, plus a single correction, which we
will discuss later.
\begin{align}
  A_{HS} &= k_BT \int \Phi(\{n_\alpha(\rr)\}) d\rr
\end{align}

We propose to treat the orientational dependence of the fundamental
measures using an expansion in spherical harmonics:
\begin{align}
  w_\alpha(\rr) &= \sum_{lm} \tilde{w}_\alpha^{lm}(|\rr|) Y_{lm}(\rhat)
\end{align}
Given this expansion, we can simplify the weighted densities:
\begin{align}
  n_\alpha(\rr) &= \int \sum_{lm} n(\rr',\rot)
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm}(\widehat{R(\rot)(\rr-\rr')}) d\rr' d\rot
  \\
  &= \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
  Y_{lm'}(\widehat{(\rr-\rr')})
  n(\rr',\rot)
       D_{lmm'}(\rot) d\rot d\rr'
       \\
  &=
    \sum_{lmm'} \int \tilde{w}_\alpha^{lm}(|\rr-\rr'|)
    Y_{lm'}(\widehat{(\rr-\rr')})
    n_{lmm'}(\rr') d\rr'
\end{align}
This reduces the challenge of handling the rotation to that of solving
for a set of densities $n_{lmm'}(\rr)$.
\subsection{Ideal gas free energy}
To address this, we will
first consider the free energy of an ideal gas.
\begin{align}
  A_{\textit{ideal}} &=
  k_BT \int
  \left(n(\rr,\rot)\ln(\Lambda^3\Theta^3 n(\rr,\rot)) - n(\rr,\rot)
  \right) d\rr d\rot
  \\
  &=
  k_BT \sum_{lmm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)\left(
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right) - 1
  \right) d\rr d\rot
  \\
  &=
  k_BT \int n_{000}(\rr)
  \left(\ln\left(\Lambda^3\Theta^3 \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)\right) - 1
  \right) d\rr d\rot
  \\&+
  k_BT \sum_{l>1,mm'} \int
  n_{lmm'}(\rr)D_{lmm'}(\rot)
  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &= A_{000} + \sum_{l>1,mm'} A_{lmm'} - k_BT\int n_{000}(\rr)d\rr
\end{align}
Here I have broken the $000$ term out of the outer summation over
$lmm'$, since this one is special, since it is independent of $\rot$.
We'll tackle that integral first, since it is a tad simpler.
\begin{align}
  A_{000} &= k_BT\int
  n_{000}(\rr)\ln\left(
  \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)
  \right)
  d\rr d\rot
\end{align}
We can simplify this integral a bit if we write the log as a power
series...
\begin{align}
  A_{000}
  &= k_BT\int
  n_{000}(\rr)\ln\left(
  n_{000}(\rr)\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  \right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\ln\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &= k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)
  \\&+k_BT\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\&-\frac{k_BT}{2!}\int
  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)^2
  d\rr d\rot
  \\&+\cdots
\end{align}
Now the linear term in the power series will vanish, since the
summation is over $l>0$ only.  In the quadratic term, only a few terms
will survive, and we'll end up with
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} n_{lmm'}(\rr)n_{l,-m,-m'}}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad-\frac{k_BT}{2!}\int
  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} |n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr
  \\&\quad+\cdots
\end{align}
Now an interesting question is whether we can mimic this power series
with a log, such as
\begin{align}
  A_{000}
  &=
  k_BT\int
  n_{000}(\rr)\ln \left( \sqrt{n_{000}(\rr)^2 + \sum_{lmm'}
    c_ln_{lmm'}(\rr)^2} \right)d\rr
  \\
  &= 
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int n_{000}(\rr)\ln\left( \sqrt{1 + \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)^2}} \right)d\rr
  \\
  &=
  k_BT\int
  n_{000}(\rr)\ln n_{000}(\rr)d\rr
  \\&\quad+
  k_BT\int \frac12 \frac{\sum_{lmm'}
    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)} d\rr
  \\&\quad+\cdots
\end{align}
from which we can conclude that to get an answer correct to the
quadratic order, we need
\begin{align}
  c_l &= -\frac{2l+1}{8\pi^2}
\end{align}
which seems all right except for the sign:  I expect that the free
energy should increase as we increase $n_{lmm'}$, since it means the
system is becoming more ordered, and it looks like to this quadratic
term, the free energy decreases... except that we're only looking at
$A_{000}$ right now, and thus are neglecting a possibly significant
increase due to the other terms...  so perhaps we should call this all
right, and move on for now.

The $A_{lmm'}$ terms are a bit different, but not too different:
\begin{align}
  A_{lmm'} &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \ln\left(\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \ln\left(1 + \frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
  \left(\frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}+\cdots\right)
  d\rr d\rot
  \\
  &=
  k_BT \int n_{lmm'}(\rr) \frac{8\pi^2}{2l+1} \frac{n_{l,-m,-m'}(\rr)}{n_{000}(\rr)}
  d\rr + \cdots
  \\
  &=
  k_BT \int \frac{8\pi^2}{2l+1} \frac{|n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
  d\rr + \cdots
\end{align}
It is interesting how similar the $A_{lmm'}$ terms are to the
quadratic terms in $A_{000}$, off by a factor of $-\frac12$.  This
leads me to wonder about the possibility of writing
\begin{align}
  A_{\textit{ideal}} &\approx
  k_BT \int n_{000}(\rr)\left( \ln\left(
  \sqrt{\sum_{lmm'}\frac{2l+1}{16\pi^2}|n_{lmm'}|^2}
  \right) - 1\right)d\rr
\end{align}
This would be correct up to the quadratic terms, and doesn't require
any rotational integrals.  It also seems liable to be reasonable up to
higher order terms, although I have no guarantee of that.  An
interesting excersize would be to take this out to the fourth order,
and also take the exact expression out to the fourth order and see how
they compare.  They will actually differ in the third order terms, I
expect, since the exact expression will indeed have nonzero third
order terms.

So we have two plausible options that I can see: we could either go
with this logarithmic approximation, or we could just go out to second
order.  The latter is probably more defensible in terms of
perturbation theory, and would give us something like
\begin{align}
  A_{\textit{ideal}} &\approx
  k_BT \int \left(n(\rr)\ln n(\rr) - n(\rr) +
  \sum_{lmm'}\frac{2l+1}{16\pi^2}\frac{|n_{lmm'}|^2}{n_{000}(\rr)}
  \right)d\rr
\end{align}
where I have substituted the notation $n(\rr)$ for $n_{000}(\rr)$ to
save space.

\subsection{Finding the orientational moments}
We find the orientational moments $n_{lmm'}(\rr)$ by minimizing the
free energy.  Let's begin by optimizing the free energy with respect
to nonzero angular momenta:
\begin{align}
  0&=\frac{\delta A}{\delta n_{lmm'}(\rr)} \\
  &= \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  + k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \frac{\delta n_\alpha(\rr')}{\delta n_{lmm'}(\rr)} d\rr'
  \\
  &=
   \frac{2l+1}{8\pi^2}n_{lmm'}(\rr)
  \\&\quad+ k_BT \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
Solving for $n_{lmm'}(\rr)$, we find that
\begin{align}\label{eq:nlmm'}
  n_{lmm'}(\rr) &= \frac{8\pi^2k_BT}{2l+1}
  \int\sum_\alpha
  \frac{\partial \Phi(\rr')}{\partial n_\alpha(\rr')}
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|)Y_{lm'}(\widehat{\rr-\rr'}) d\rr'
\end{align}
I note that this equation for $n_{lmm'}$ is approximate, since we made
an approximation when we neglected higher order terms in the expansion
of the log in the ideal gas free energy.  Thus in cases where there is
relatively little orientational ordering, it ought to be an accurate
equation.  In any case I imagine it will be plausible.

\subsection{The Procedure}
Our input when computing the free energy is the ordinary density
$n(\rr)$, which is precisely equal to $n_{000}(\rr)$ above, because it
is just the orientationally averaged density, which is also what
$n_{000}(\rr)$ is.  From this density, we can compute a first set of
fundamental measures
\begin{align}
  n_\alpha(\rr) &= \int \tilde{w}_\alpha^{000}(|\rr-\rr'|)n(\rr')d\rr'
\end{align}
where we will discuss elsewhere how we find the weighting functions
$\tilde{w}_\alpha^{000}$.  Thus we have our zeroth-order approximation
to the fundamental measures.  At this point we can compute a
zeroth-order free energy just by evaluating the free energy functional
with these fundamental measures, which would be the true free energy
if there were no orientational ordering.

From the zeroth-order fundamental measures, we can
compute the orientational densities using Eq.~\ref{eq:nlmm'}.  Once we
have these densities, we can find an improved set of fundamental
measures:
\begin{align}
  n_\alpha(\rr) &= \sum_{lmm'} \int
  \tilde{w}_\alpha^{lm}(|\rr-\rr'|) Y_{lm'}(\rr-\rr')n_{lmm'}(\rr')
\end{align}
With this improved set of fundamental measures, we can now just
evaluate the free energy.  How could it be simpler? Well, it could be
more complicated if we were to go back and try to compute an improved
set of $n_{lmm'}(\rr)$ using our improved fundamental measures.  I
suspect this would be silly, since we have already dropped out cubic
terms in $A_{\textit{ideal}}$.  No benefit from precisely computing a
higher order correction to a low-order approximation.  It \emph{is}
worth checking (particularly while debugging) that our two
approximations for the fundamental measures aren't \emph{too} wildly
different, and that the two approximations to the free energy also
aren't too wildly different.

\subsection{Finding the weighted densities}

\newcommand\xhat{\mathbf{\hat{x}}}
\newcommand\yhat{\mathbf{\hat{y}}}
\newcommand\zhat{\mathbf{\hat{z}}}
\newcommand\facehat{\mathbf{\hat{f}}}
\newcommand\Rface{\ensuremath{R_{\text{face}}}}
\newcommand\Rvertex{\ensuremath{R_{\text{vertex}}}}

Here I will discuss how to find the weighted densities we need in
order to study polyhedra.  We begin with $n_3(\rr)$, the packing
fraction.
\begin{align}
  w_3(\rr) &= 1 \text{ inside the polyhedron}\\
  &= 0 \text{ outside the polyhedron} \\
  &= \prod_i^{\text{faces}} \Theta(\Rface - \rr\cdot\facehat_i)
\end{align}
This is pretty easy, although the integrals will be less so.  The
$000$ integral is always going to give us something that is
essentially the same, but I won't go into this yet.

$n_2(\rr)$ is the area, which we can easily get from the above
weighted density.
\begin{align}
  w_2(\rr) &= |\nabla w_3(\rr)| \\
  &= \sum_i^{\text{faces}} \delta(\Rface-\rr\cdot\facehat_i)
  \prod_{j\ne i} \Theta(\Rface-\rr\cdot\facehat_j)
\end{align}
We could also just ask what the area is at a given position.

$n_1(\rr)$ is the first fundamental measure that is interestingly
different for polyhedra than it is for spheres.  This is a measure of
the mean curvature, and polyhedra have no curvature except at the
edges.  Thus $n_1(\rr)$ is going to be nonzero only on the edges, and
there it will be a delta function.
\begin{align}\label{eq:w1}
  w_1(\rr) &= \sum_{ij}^{\text{NN faces}}
  \Theta\left(\Rvertex - r\right)
  \delta(\Rface - \rr\cdot \facehat_i)
  \delta(\Rface - \rr\cdot \facehat_j)
\end{align}
where $\facehat_i$ is a unit vector in the direction of the $i$ face
of the polyhedron, $\Rvertex$ is the distance between the center of
the polyhedron and the vertex, and $\Rface$ is the distance of the
center of the face from the origin.  WARNING: I may have something off
in these formulas, but the general idea is there.  To find a constant
factor, we'd consider the

$n_0(\rr)$ is the fundamental measure of gaussian curvature, which is
the product of the curvature in the two eigendirections.  This is zero
(for polyhedra) except at the corners, so $n_0(\rr)$ itself is going
to be zero except at the corners.  Also, the gaussian curvature
must satisfy a sum rule, which tells us that
\begin{align}
  w_0(\rr) &= \frac1{N}\sum_{i}^{N\text{ vertices}} \delta(\rr-\rr_i)
\end{align}
Interestingly, this meas that
\begin{align}
  w_0^{000}(\rr) &= \frac{\delta(\Rvertex - |\rr|)}{4\pi \Rvertex^2}
\end{align}
which means that as far as $w_0^{000}$ is concerned, the polyhedron is a
sphere with a radius equal to the vertex distance (i.e. the distance
from the center to the vertex of the polyhedron).

\subsection{Tetrahedron}

\subsubsection{$w_3^{000}$}

In order to solve for $w_3^{000}(r)$ of a tetrahedron, we need to
consider where $r$ lies.  If $r<\frac{\sqrt{3}}{2}R$ (which is
\Rface), then the sphere lies entirely within the tetrahedron, and
$w_3^{000}(r)=1$.  Conversely, if $r<\sqrt{3}R$ (which is \Rvertex),
then $w_3^{000}(r)=0$.  There are two in-between cases:  either
$R<r<\Rvertex$ (in which case each vertex forms an isolated triangle
poking out of the sphere), or $\Rface<r<R$, in which case isolated
circles of spherical surface poke out of the faces of the
tetrahedron.  The latter case is pretty easy, if we just integrate
over the sticking-out bits, and subtract that from the total surface
area of the sphere:
\begin{align}
  w_3^{000}(r) &= 1 - \frac1{\pi}\int d\Omega \\
  &= 1 - \frac1{\pi}2\pi \int_0^{\cos^{-1}\left(\frac{\sqrt{3}R}{2r}\right)} \sin\theta d\theta
  \\
  &= 1 - 2 \int^1_{\frac{\sqrt{3}R}{2r}} du
  \\
  &= \frac{\sqrt{3}R}{r} - 1
  \\
  \frac{\sqrt{3}R}{2}&<r<R
\end{align}

The final consideration (for $w_3^{000}(r)$ of a tetrahedron) is when
$R < r < \sqrt{3}R$, when we will want to just integrate over the
curvy triangles formed by the portion of the tetrahedron that stick
through the sphere.  We just need to integrate through one of these
points, and multiply by four... which might be easiest to do in a
Cartesian coordinates rotated by 45$^\circ$ around the ordinary $z$
axis.
\begin{align}
  w_3^{000}(r)
  &=
  \frac4{4\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_0^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_0^{\frac{R+z}{\sqrt{2}}} dy
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \delta(r-\sqrt{x^2+y^2+z^2})
  \\
  u&= \sqrt{x^2+y^2+z^2}\\
  du &= \frac{ydy}{\sqrt{x^2+y^2+z^2}}
  \\
  w_3^{000}(r)
  &=
  \frac1{\pi r^2}
  \int_{0}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  &=
  \frac1{\pi r^2}
  \int_{\sqrt{r^2-2R^2}}^R dz
  \int_{\frac{z-R}{\sqrt{2}}}^{\frac{R-z}{\sqrt{2}}} dx
  \int_{\sqrt{z^2+x^2}}^{\sqrt{x^2+z^2+\frac{(R+z)^2}{2}}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2-x^2}}
  \\
  z^2+x^2+y^2 &= r^2 \\
  z_{min}^2 + (R-z_{min})^2/2 &= r^2
\end{align}
Instead I'm going now to think of using cylindrical coordinates...
\begin{align}
  w_3^{000}(r) &= 3\frac{4}{4\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_0^{2\frac{\sqrt{3}R-z}{\cos\phi}} ds
  \delta(r-\sqrt{z^2+s^2})
  \\
  u &= z^2+s^2 \\
  du &= sds \\
  &= \frac{\sqrt{u^2-z^2}}{u}ds
  \\
  w_3^{000}(r)
  &= \frac{3}{\pi r^2}
  \int_0^{\sqrt{3}R}dz\int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi\int_{z}^{\sqrt{z^2+\left(2\frac{\sqrt{3}R-z}{\cos\phi}\right)^2}} du
  \delta(r-u)\frac{u}{\sqrt{u^2-z^2}}
  \\
  R &= \frac{2}{\sqrt{3}}s\cos\phi + \frac{1}{\sqrt{3}}z \\
  s &= 2\frac{\sqrt{3}R-z}{\cos\phi}
\end{align}
\newcommand\thetamax{\theta_{\text{max}}}
\begin{widetext}
Maybe spherical coordinates? Rotated $\pi/4$? Let's look first at the
plane that is our limit for $\theta$:
\begin{align}
  \sqrt{3}R &= x + y + z \\
  &= \sqrt{2}r\sin\thetamax\cos\phi +
  r\cos\thetamax
  \\
  \sqrt{3}\frac{R}{r} &=
  \sqrt{2}\sqrt{1-\cos^2\thetamax}\cos\phi
  + \cos\thetamax
  \\
  \sqrt{1-\cos^2\thetamax} &=
  \frac{\sqrt{3}\frac{R}{r}-\cos\thetamax}{\sqrt{2}\cos\phi}
  \\
  1-\cos^2\thetamax &=
  \frac{3\frac{R^2}{r^2}+\cos^2\thetamax -
    2\sqrt{3}\frac{R}{r}\cos\thetamax}{2\cos^2\phi}
  \\
  0 &=
  (1+2\cos^2\phi)\cos^2\thetamax
  - 2\sqrt{3}\frac{R}{r}\cos\thetamax
  + \left(3\frac{R^2}{r^2} - 2\cos^2\phi\right)
  \\
  \cos\thetamax &= \frac{2\sqrt{3}\frac{R}{r}
  - \sqrt{12\frac{R^2}{r^2} - 4(1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{2(1+2\cos^2\phi)}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
  \\
  &= \frac{\sqrt{3}\frac{R}{r}
  - \sqrt{3\frac{R^2}{r^2} - (2-\cos2\phi)\left(3\frac{R^2}{r^2}-1+\cos2\phi\right)}}{2-\cos2\phi}
\end{align}
\end{widetext}
This is sort of a mess, but does look useable.
\begin{align}
  w_3^{000}(r) &=
  \frac{12}{4\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}d\phi
  \int_0^{\thetamax}d\theta
  \sin\theta
  \\
  &=
  \frac{3}{\pi}
  \int_{-\frac{\pi}{3}}^{\frac{\pi}{3}}
  (1 - \cos\thetamax)d\phi
\end{align}
Surely we can either use Euler's relation or a geometrical
construction to solve for $\thetamax$?!

\subsubsection{$w_2^{000}$}
The easiest way to get $w_2^{000}(r)$ is just to take a derivative of
$w_3^{000}(r)$ we should have just computed:
\begin{align}
  w_2^{000}(r) &= \frac{dw_3^{000}(r)}{dr}
\end{align}

\subsubsection{$w_1^{000}$}
As mentioned before, $w_1^{000}(r)$ is the average of an integral
around the edges of the tetrahedron.  This is a little bit tricky, as
seen in Eq.~\ref{eq:w1}.  $w_1^{000}$ is zero except when $\sqrt{2}R <
r < \sqrt{3}R$.
\begin{align}
  \rr &= r\left(
  \sin\theta\cos\phi\xhat +
  \sin\theta\sin\phi\yhat +
  \cos\theta\zhat \right)
  \\
  \facehat_1 &= \frac{\xhat+\yhat+\zhat}{\sqrt{3}} \\
  \facehat_2 &= \frac{-\xhat-\yhat+\zhat}{\sqrt{3}} \\
  \rr\cdot\facehat_1 &=
  \frac{r}{\sqrt{3}}\left(
  \cos\theta+\sin\theta\cos\phi + \sin\theta\sin\phi
  \right)
  \\
  &= \frac{r}{\sqrt{3}}\left(
  \cos\theta+\sqrt{2}\sin\theta\cos(\phi+\pi/4)
  \right)
  \\
  \rr\cdot\facehat_2 &=
  \frac{r}{\sqrt{3}}\left(
  \cos\theta-\sqrt{2}\sin\theta\cos(\phi+\pi/4)
  \right)
\end{align}
\begin{widetext}
\begin{align}
  w_1^{000}(r) &= \frac{6}{4\pi}\int
  \delta(\sqrt{2}R - \rr\cdot \facehat_1)
  \delta(\sqrt{2}R - \rr\cdot \facehat_2)
  \sin\theta d\theta d\phi
  \\
  &= \frac{6}{4\pi}\int
  \delta\left(\sqrt{2}R
  - \frac{r}{\sqrt{3}}(\cos\theta-\sin\theta\cos(\phi+\pi/4))\right)
  \delta\left(\sqrt{2}R
  - \frac{r}{\sqrt{3}}(\cos\theta+\sin\theta\cos(\phi+\pi/4))\right)
  \sin\theta d\theta d\phi
\end{align}
\end{widetext}
\begin{align}
  u &= \sqrt{2}R -\frac{r}{\sqrt{3}}(\cos\theta + \sin\theta\cos(\phi+\pi/4)) \\
  du &= \frac{r}{\sqrt{3}}\sin\theta\sin(\phi+\pi/4)d\phi \\
  d\phi &= \frac{\sqrt{3}}{r\sin\theta}\frac{1}{\sin(\phi+\pi/4)}
  \\
  w_1^{000}(r)
  &= \frac{6}{4\pi}\int
  \delta\left(-2\frac{r}{\sqrt{3}}\cos\theta+u\right)
  \delta(u)
   \frac{\sqrt{3}}{r}\frac{ d\theta du}{\sqrt{1-u^2}OOPS!}
  \\
  &= \frac{6}{4\pi}\int
  \delta\left(-2\frac{r}{\sqrt{3}}\cos\theta\right)
   \frac{\sqrt{3}}{r}d\theta
   \\
  v &= -2\frac{r}{\sqrt{3}}\cos\theta \\
  dv &= 2\frac{r}{\sqrt{3}}\sin\theta d\theta \\
  &= 2\frac{r}{\sqrt{3}}\sqrt{1-v^2}d\theta \\
  w_1^{000}(r)
  &= \frac{6}{4\pi}\int
  \delta\left(v\right)
   \frac{6}{r^2}\frac{dv}{\sqrt{1-v^2}} \\
   &= \frac{9}{\pi r^2}
\end{align}
I'm a little doubtful of the above angular integral.  I don't see how
it will be able to integrate to anything very reasonable when
integrating over $r$.  In particular, I am pretty certain that
\emph{has} to depend on the distance of closest approach.
I know that
\begin{align}
  \int_{r_{min}}^{r} w_1^{000}(r)4\pi r^2dr
  &\propto \sqrt{r^2 - r_{min}^2}
\end{align}

\subsection{Cube}

In calculating $w_3^{000}$ we only need to integrate over one octant.
\begin{align}
  w_3^{000}(r) &= \int \Theta(R-x)\Theta(R-y)\Theta(R-z) d\Omega
  \\
  &= 8\int_0^{\frac{\pi}{2}} \int_0^{\frac{\pi}{4}}
  \Theta(R-r\sin\theta\cos\phi)
  \Theta(R-r\sin\theta\sin\phi)
  \Theta(R-r\cos\theta)
  r^2\sin\theta d\theta d\phi
  \\
  &=
  x
\end{align}
There are a couple of possibilities.  If $r<R$ then $w_3^{000}(r) =
1$.  If $r>\sqrt{3}R$ the $w_3^{000}(r) = 0$.  That leaves two
interesting possibilities.  If $R<r<\sqrt{2}R$ then there are six
edges to the integral, and we find that
\begin{align}
  w_3^{000}(r) &= \frac1{4\pi r^2}\int \Theta(R-x)\Theta(R-y)\Theta(R-z) d\Omega
\end{align}
If $\sqrt{2}R<r<\sqrt{3}R$ then there are just three sides to the
integral, as illustrated in Fig.~\ref{fig:rbig}.  This integral is
easiest to do by integrating over just one of the eight protruding
corners.  This seems easiest if we define the corner as the $z$ axis.
\begin{align}
  w_3^{000}(r) &= \frac{6}{4\pi}
  \int_{0}^{\frac{\pi}{3}}
  \int_0^{\theta_{m}} \sin\theta d\theta d\phi \\
  r\cos\theta_{m} + r\sin\theta_{m}\cos\phi  &= R
  \\
  \cos\theta_{m} + \sin\theta_{m}\cos\phi  &= \frac{R}{r} \\
  w_3^{000}(r) &= \frac{6}{4\pi}
  \int_{0}^{\frac{\pi}{3}}
  \int_0^{\theta_{m}} \sin\theta d\theta d\phi
\end{align}
\begin{align}
  w_3^{000}(r) &= \frac{8}{4\pi r^2}\iint_0^R \delta(r-\sqrt{x^2+y^2+z^2})
  dxdydz \\
  u &= \sqrt{x^2+y^2+z^2} \\
  du &= \frac{zdz}{u} \\
  w_3^{000}(r) &= \frac{8}{4\pi r^2}\iint_0^R \delta(r-u)
  dxdy\frac{udu}{\sqrt{u^2 -x^2-y^2}} \\
  &= \frac{2}{\pi r}\int_{r_0}^R\int_{r_0}^R
  dxdy\frac{1}{\sqrt{r^2 -x^2-y^2}} \\
  &= \frac{2}{\pi r}\int_{\sqrt{r^2-2R^2}}^R\int_{\sqrt{r^2-R^2-x^2}}^R
  \frac{dxdy}{\sqrt{r^2 -x^2-y^2}} \\
\end{align}
\begin{figure}
\includegraphics[width=2in]{figs/w3cube}
\caption{Picture of the sphere with $\sqrt{2}R<r<\sqrt{3}R$.}\label{fig:rbig}
\end{figure}

%\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2013-01-18 papers/polyhedra/paper.tex
add polyhedra citations Tested everything on node24.

@@
 develop a reference functional that can incorporate orientational
-effects.
+effects.  (For reference, an excellent review of hard-particle systems
+is the following book chapter~\cite{tarazona2008chapter}.)
 
 There have been several studies of non-spherical hard shapes.  At
-least one omitted the rotational degrees of freedom (parallel hard
-cubes, add citation).  At least one included the rotational degrees of
-freedom, treating each orientation as a distinct species, resulting in
-a density that exists in six dimensions rather than three, which was
-applied to systems with a high aspect ratio, which undergo a nematic
-transition.  It seems a good thing to have a density functional that
-is more efficient, with fewer degrees of freedom to optimize.  In
-particular, we are optimistic that one can do better for ``almost
-spherical'' polyhedra.
+least one omitted the rotational degrees of
+freedom~\cite{cuesta1996cubes, cuesta1997dimensional,
+  cuesta1997fundamental, buhot1998numerical, martinez1999fundamental,
+  capitan2007hexagons,
+  capitan2008phase, martinez2008fundamental}.  At least one included
+the rotational degrees of freedom, treating each orientation as a
+distinct species, resulting in a density that exists in more
+dimensions than three, which was applied to systems with a high aspect
+ratio, which undergo a nematic transition~\cite{schmidt2001density,
+  esztermann2006density, harnau2008structure, hansen2009fundamental,
+  hansen2010tensorial}.  It seems a good thing to have a density
+functional that is more efficient, with fewer degrees of freedom to
+optimize.  In particular, we are optimistic that one can do better for
+``almost spherical'' polyhedra.
+
+There has been simulation of the melting of hard cubes, demonstrating
+that allowing rotational disorder leads to a first-order phase
+transition~\cite{jagla1998melting}, which is not seen in parallel hard
+cubes.  There has been another study of melting of parallel hard
+cubes~\ref{groh2001closer}.
 
@@
 
-%\bibliography{paper}% Produces the bibliography via BibTeX.
+\bibliography{paper}% Produces the bibliography via BibTeX.
 

### 2013-01-22 papers/pair-correlation/paper.tex

@@
 \usepackage{color}
+\usepackage{breqn}
 
@@
 
-\maketitle
+%\maketitle
 
@@
 
+Lado recently introduced a new and improved efficient algorithm for
+implementing integral equation theory for inhomogeneous fluids, which
+computes $g^{(2)}(\rr_1,\rr_2)$ among other
+things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
+use, and their example is limited to 1D systems.  And the scaling is
+still (so far as I can tell) $O(N^2)$, since their improvement is in
+the prefactor, by developing a nice basis set.
+
 \section{Pair distribution function with inhomogeneity}
+
+The pair distribution function can be defined using the two particle
+density as
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) &\equiv \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
+\end{align}
+
+We may also want to consider the \emph{triplet} distribution function
+\begin{align}
+  g^{(3)}(\rr_1,\rr_2,\rr_3) &\equiv \frac{n^{(3)}(\rr_1,\rr_2,\rr_3)}{n(\rr_1)n(\rr_2)n(\rr_3)}
+\end{align}
+The triplet distribution is relevant, because we can compute the
+triplet distribution of the homogeneous system by considering the
+\emph{pair} distribution of the inhomogeneous ``test-particle
+scenario.''  The process is that we solve the inhomogeneous situation
+in which we have one hard sphere located at the origin.  We solve for
+the density around that hard sphere, and this gives us the pair
+distribution function of the homogeneous system $g^{(2)}(0, \rr)$.
+When we solve for the pair distribution $g^{(2)}(\rr_1,\rr_2)$ of the
+inhomogeneous system, this will give us the \emph{triplet}
+distribution function of the homogeneous system
+$g^{(3)}(0,\rr_1,\rr_2)$.
+
+\section{Goals of this paper (to be integrated into rest of paper)}
+
+In this paper, we seek an efficient approximation for the pair
+correlation function of an inhomogeneous hard-sphere fluid, where we
+define efficient as an algorithm that scales as $O(N\log N)$ where $N$
+scales as the volume of the system under consideration.
+Integral equation theory (look this up) scales as $O(N^2)$, which can
+be very expensive for large three-dimensional systems.
+
+Fundamental-Measure Theory (FMT) is constructed using only single-point
+convolutions (check terminology) such as
+\begin{align}
+  n_\alpha(\rr) &= \int n(\rr')w_\alpha(|\rr-\rr'|) d\rr'
+\end{align}
+These convolutions can be computed in Fourier space in $O(N\log N)$
+time, with the result that the FMT free energy---and its
+gradient---can be computed in $O(N\log N)$ time.
+
+The pair distribution function is used in computing the first term in
+the high-temperature perturbation expansion, which is used in SAFT
+(for instance) to treat dispersion interactions.  This term is written
+as
+\begin{align}\label{eq:a1}
+  a_1 &\propto \iint n(\rr)n(\rr') \Phi(|\rr-\rr'|)
+  g^{(2)}(\rr,\rr')d\rr d\rr'
+\end{align}
+where $\Phi(r)$ is the pair potential.  If $g^{(2)}(\rr,\rr')$ could
+be written as $g(|\rr-\rr'|)$, this term could be computed in $O(N\log
+N)$ time.  However, this is not possible when the density is
+inhomogeneous, because the pair distribution function depends on the
+density.
+
+In this paper, we consider a superposition approximation, in which
+construct an approximate representation for the pair distribution
+function, which allows the first term in the high temperature
+expansion given in Equation~\ref{eq:a1} to be computed in $O(N\log N)$
+time.  One input to this approach is our recently published expression
+for an average of the value of the pair distribution function at
+contact $g_\sigma(\rr)$, which itself can be computed in $O(N\log N)$
+time.
+
+The general idea is to write
+\begin{dmath}
+  g^{(2)}(\rr_1,\rr_2) = \Theta(r_{12}-\sigma)
+  \left(1 + \sum_{\alpha} \frac{f_{\alpha}(g_\sigma(\rr_1)) + f_{\alpha}(g_\sigma(\rr_2))}{2}g_{\alpha}(r_{12})\right)\label{eq:gpair-approx}
+\end{dmath}
+where $f_{\alpha}$ and $g_{\alpha}$ are a set of functions that
+satisfy the following constraint:
+\begin{dmath}\label{eq:definealpha}
+  \sum_{\alpha} f_{\alpha}(g_\sigma(\eta)) g_{\alpha}(r) \approx g(r; \eta)
+\end{dmath}
+which ensures that our pair distribution function reduces
+(approximately) to the radial distribution function in the homogeneous
+limit.  Equation~\ref{eq:gpair-approx} is constructed such that
+Equation~\ref{eq:a1} can be evaluated in $O(N\log N)$ time, since all
+the required convolutions may be performed with a fixed kernel.
+
+Our proposed pair distribution function has several good properties.
+When evaluated at contact ($r_{12}=\sigma$), it gives the same good
+result as our previously-published functional for the averaged pair
+distribution function at contact, which is constructed to give an
+accurate prediction for the contribution to Equation~\ref{eq:a1} from
+spheres in contact.  In the homogeneous limit,
+Equation~\ref{eq:gpqair-approx} should give a prediction that is as
+accurate as the approximate equality in
+Equation~\ref{eq:definealpha}.  In inhomogeneous distributions and at
+intermediate distances, we expect our pair distribution function to be
+least accurate, but one hopes that it will be reasonable, simply
+because it is an average of two reasonable results.
 
@@
 \begin{equation}
-A_\textit{HS}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
+A_{\textit{HS}}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
 \end{equation}
@@
 
+The traditional approach for solving for the pair distribution
+function of an inhomogeneous density distribution is to use an
+integral equation theory...
+
 \section{Comparison with simulation}\label{sec:comparison}

### 2013-01-22 papers/polyhedra/wigner-properties.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}

\begin{document}
\title{Properties of the Wigner D Function}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}
%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  This is just some notes to myself regarding the Wigner $D$ function,
  which I hope to find useful in my research.
\end{abstract}

\newcommand\rot{\ensuremath{\mathbf{\varpi}}}
\newcommand\rhat{\ensuremath{\mathbf{\hat{r}}}}
\newcommand\solidangle{\ensuremath{\Omega}}

\maketitle

\subsection{Properties of the Wigner $D$ matrix}

There are various ways one can understand the Wigner $D$ function, but
I think the simplest is to state that it defines how spherical
harmonics behave under a rotation that I will call $\rot$, which we
can think of as being defined by a set of Euler angles.  I will
describe the rotation matrix as $R(\rot)$.  Without further ado, I
will state that the $D$ functions (or matrices) are defined by:
\begin{align}\label{eq:Ylm-rotation}
  Y_{lm}(R(\rot)\rhat) &= \sum_{|m'|<l} D^{l}_{mm'}(\rot) Y_{lm'}(\rhat)
\end{align}
which is to say that the $D$ matrix defines how spherical harmonics
transform under rotation.  $D$ functions also form an orthogonal basis
set in the space of rotations (or orientations).
\begin{align}
  \int D^{l}_{mn}(\rot)^{*}D^{l'}_{m'n'}(\rot) d\rot
  &= \frac{8\pi^2}{2l+1}\delta_{ll'}\delta_{mm'}\delta_{nn'}
\end{align}
Sadly, the $D$ matrices aren't normalized... but if they were, then
that normalization would have popped up in
Equation~\ref{eq:Ylm-rotation}, so you can't have everything.

The complex conjugate of a $D$ matrix just corresponds to an inverse
rotation:
\begin{align}
  D^{l}_{mm'}(\rot)^{*} &= (-1)^{m'-m}D^{l}_{-m-m'}(\rot)
\end{align}

\subsection{Properties of spherical harmonics}

Spherical harmonics are properly orthonormal
\begin{align}
  \int Y_{lm}(\rhat)^{*}Y_{l'm'}(\rhat) d\solidangle
  &= \delta_{ll'}\delta_{mm'}\delta_{nn'}
\end{align}

The complex conjugate of a $Y_{lm}$ corresponds to time reversal, or
an opposite rotation (up to a sign):
\begin{align}
  Y_{lm}^{*}(\rhat) &= (-1)^{m} Y_{l,-m}(\rhat) \\
  Y_{lm}(\rhat) &= (-1)^{m} Y_{lm}(-\rhat)
\end{align}

\end{document}

### 2013-01-31 papers/polyhedra/paper.tex
a few edits to polyhedron paper Tests skipped!

@@
 \newcommand{\fixme}[1]{\red{[#1]}}
+
+\newcommand\thetamax{\theta_{\text{max}}}
 
@@
 for a set of densities $n_{lmm'}(\rr)$.
+
 \subsection{Ideal gas free energy}
@@
 first consider the free energy of an ideal gas.
+\begin{widetext}
 \begin{align}
@@
 
+\end{widetext}
 \subsection{Finding the orientational moments}
@@
 axis.
+\begin{widetext}
 \begin{align}
@@
 \end{align}
-\newcommand\thetamax{\theta_{\text{max}}}
-\begin{widetext}
 Maybe spherical coordinates? Rotated $\pi/4$? Let's look first at the
@@
   &= \frac{\sqrt{3}\frac{R}{r}
-  - \sqrt{3\frac{R^2}{r^2} - (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
+  - \sqrt{3\frac{R^2}{r^2} -
+    (1+2\cos^2\phi)\left(3\frac{R^2}{r^2}-2\cos^2\phi\right)}}{1+2\cos^2\phi}
   \\

### 2013-01-31 papers/polyhedra/wigner-properties.tex
a few edits to polyhedron paper Tests skipped!

@@
 \documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
+
+\usepackage{color}
 
@@
 
+We have a final relationship that allows us to convert products of $D$
+matrices into a linear combination of other $D$ matrices:
+\begin{align}
+  D^{l}_{mk}(\rot)D^{l'}_{m'k'} &=
+  \sum_{l'', m'', k''} \left<lml'm' | l''m''\right>
+  \left<lkl'k' | l''k''\right> D^{l''}_{m''k''}(\rot)
+\end{align}
+where the coefficients are Clebsch-Gordon coefficients.  Relating to
+this, we have a special case that seems to be sort of handy for
+certain Clebsch-Gordon coefficients:
+\begin{align}
+  \left< lm l'm' | 00 \right> &= \delta_{ll'}\delta{m,-m'}\frac{(-1)^{l-m}}{\sqrt{2l+1}}
+\end{align}
+
 \subsection{Properties of spherical harmonics}
@@
 
+\subsection{A simple integral}
+Let's consider the following integral, where $n>1$:
+\begin{align}
+  X_{lmkn} &\equiv \int \left(\Delta^{l}_{mk}(\rot)\right)^{2n} d\rot
+\end{align}
+This is useful in the above.  Of course, odd powers vanish.  I expect
+that this integral is:
+\begin{align}
+  X_{lmkn} &=  8\pi^2 \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} i}
+\end{align}
+based on the assumption that it is independent of the particular value
+of $m$ and $k$.  If this is true, then we can simply evaluate
+\begin{align}
+  X_{ll0n} &= \int \left(\Delta^{l}_{l0}(\rot)\right)^{2n} d\rot \\
+  &= \int \left( \sin(l\phi) \right)^{2n} \sin\theta d\theta d\phi
+  d\chi \\
+  &= 4\pi \int_0^{2\pi} \left( \sin(l\phi) \right)^{2n} d\phi \\
+  &=  8\pi^2 \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} 2i}
+\end{align}
+So if this particular integral, which could be gotten from appropriate
+sums of Clebsch-Gordon coefficients, is independent of $m$ and
+$k$---which seems likely, somehow, since it shouldn't be affected by a
+rotation---then we have a relatively nice closed-form solution for its
+value.
+
+\begin{widetext}
+\subsection{An interesting integral}
+
+Let's consider the following quantity, which is the simplest
+nontrivial real and positive function of orientation that I can
+imagine:
+\begin{align}
+  f(\rot) &= a + b\frac{D^{l}_{mk}(\rot) +
+    (-1)^{m}D^{l}_{-m,-k}(\rot)}{\sqrt 2}
+  \\
+  &= a + b \Delta^{l}_{mk}(\rot)
+\end{align}
+where I've taken a constant plus a single $lmk$ value, but made it
+real.  If $a > |b|$ by enough, then this will always be real, which
+will turn out to be important for a density.  I have defined the
+function $\Delta^{l}_{mk}(\rot)$, which is a real version of the
+Wigner $D$ function, and has the interesting properties that
+\begin{align}
+  \int \Delta^{l}_{mk}(\rot) \Delta^{l'}_{m'k'}(\rot)d\rot
+  &= \frac{8\pi^2}{2l+1}\delta_{ll'}\delta_{mm'}\delta_{nn'}
+  \\
+  \Delta^{l}_{mk}(\rot)^2 &= \frac{\Delta^{0}_{00}(\rot)}{2l+1} + \text{h.o.t.}
+\end{align}
+where the higher-order terms are terms with non-zero $l$ value.
+
+We are interested in evaluating the integral:
+\begin{align}
+  I &= \int f(\rot)\ln(f(\rot)) d\rot
+  \\
+  &=
+  \int f(\rot)\left(\ln a
+  + \ln\left(1 + \frac{b}{a}\Delta^{l}_{mk}(\rot)\right)\right) d\rot
+  \\
+  &=
+  \int f(\rot)\left(\ln a
+  + \frac{b}{a}\Delta^{l}_{mk}(\rot)
+  - \frac12 \frac{b^2}{a^2}(\Delta^{l}_{mk}(\rot))^2
+  + \frac13 \frac{b^3}{a^3}(\Delta^{l}_{mk}(\rot))^3
+  - \frac14 \frac{b^4}{a^4}(\Delta^{l}_{mk}(\rot))^4
+  + \cdots\right) d\rot \\
+  &=
+  \int f(\rot)\left(\ln a
+  - \sum_{n=1}^{\infty} \frac{(-1)^{n}}{n} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^n
+  \right) d\rot
+  \\
+  &=
+  \int (a + b \Delta^{l}_{mk}(\rot))\left(\ln a
+  - \sum_{n=1}^{\infty} \frac{(-1)^{n}}{n} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^n
+  \right) d\rot
+  \\
+  &=
+  a\int \left(\ln a
+  - \sum_{n=1}^{\infty} \frac{(-1)^{n}}{n} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^n
+  \right) d\rot
+  -
+  b\int \left(\sum_{n=1}^{\infty} \frac{(-1)^{n}}{n} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^{n+1}
+  \right) d\rot
+  \\
+  &=
+  a\int \left(\ln a
+  - \sum_{n=1}^{\infty} \frac{1}{2n} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^{2n}
+  \right) d\rot
+  +
+  b\int \left(\sum_{n=1}^{\infty} \frac{1}{2n-1} \left(\frac{b}{a}\Delta^{l}_{mk}\right)^{2n}
+  \right) d\rot
+  \\
+  &=
+  8\pi^2 a \ln a
+  - 8\pi^2a\left(
+  \sum_{n=1}^{\infty} \frac{1}{2n} \left(\frac{b}{a}\right)^{2n}
+   \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  \right)
+  +
+  8\pi^2 b \left(
+   \sum_{n=1}^{\infty} \frac{1}{2n-1} \left(\frac{b}{a}\right)^{2n}
+   \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  \right)
+  \\
+  &=
+  8\pi^2 a \ln a
+  - 8\pi^2a\left(
+  \sum_{n=1}^{\infty} \frac{1}{2n} \left(\frac{b}{a}\right)^{2n}
+   \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  \right)
+  +
+  8\pi^2 b \left(
+   \sum_{n=1}^{\infty} \left(\frac{b}{a}\right)^{2n}
+   \frac{\prod_{i=1}^{n-1} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  \right)
+  \\
+  &=
+  8\pi^2 a \left\{
+  \ln a
+  -
+  \sum_{n=1}^{\infty} \frac{1}{2n} \left(\frac{b}{a}\right)^{2n}
+   \frac{\prod_{i=1}^{n} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  +
+   \sum_{n=1}^{\infty} \left(\frac{b}{a}\right)^{2n+1}
+   \frac{\prod_{i=1}^{n-1} (2i-1) }{ \prod_{i=1}^{n} 2i}
+  \right\}
+\end{align}
+\end{widetext}
+The next idea would be to take these power series sums, and see if I
+could turn them into an analytic form.  As it is, I don't recognize
+them, but it's quite possible that they have a relatively simple
+form.  If they don't, I'll come up with a reasonable approximation
+that has the correct behavior in the limit as $b \ll a$, and also has
+the correct divergent behavior as $b$ approaches $a$.
+
 \end{document}

### 2013-02-01 papers/fuzzy-fmt/figs/functions_plot.py
updated fuzzy paper and added files for making w2*w2 plots Tested everything on bennet.

@@ -1,0 +1,52 @@
+#!/usr/bin/python
+
+from __future__ import division
+
+# We need the following two lines in order for matplotlib to work
+# without access to an X server.
+import matplotlib
+matplotlib.use('Agg')
+
+from pylab import *
+from scipy.special import erf
+
+r0 = 2
+R = r0/2
+dr = r0/200
+rsmall = arange(0, r0/2 + dr/2, dr)
+rlarge = arange(r0/2, r0 + dr/2, dr)
+r = arange(0,r0 + dr/2, dr)
+
+betaV0 = 100
+
+gamma = betaV0/2.7
+alpha = betaV0/r0
+A = 4*sqrt(betaV0/r0**2*exp(-betaV0))
+
+#plot(rsmall, A**2*rsmall*exp(alpha*rsmall), 'r-', label="exponential w2 convolved with itself")
+#plot(rlarge, A**2*(r0-rlarge)*exp(alpha*rlarge), 'r-')
+
+#plot(r, betaV0/r0*exp(-betaV0*(1-r/r0)), 'b-', label="exponential f'")
+gaussianfprime = -2*betaV0*(r-r0)/r0*exp(-betaV0*(1-r/r0)**2)
+#gaussianfprime = gaussianfprime/gaussianfprime.max()
+plot(r, gaussianfprime, label="gaussian f'")
+
+#plot(rsmall, R*exp(-gamma*(2-2*R/rsmall-(R/rsmall)**2/2))*
+#                (2*rsmall*R*sqrt(gamma)*exp(-gamma*rsmall**2/(2*R**2)) +
+#                 sqrt(2*pi)*(gamma*rsmall**2-R**2)*erf(sqrt(gamma)/(2*R)*rsmall)))
+
+B = sqrt(1-exp(-betaV0))*2*gamma/(sqrt(pi*gamma)-1)/R**2
+alphag = B**2*exp(-gamma*2*(1-rlarge/(2*R))**2)
+root2gam = sqrt(2*gamma)
+w2w2 = 2*B**2*R/root2gam*(sqrt(pi)/4*(rlarge**2-R**2/gamma)*exp(-2*gamma*(1-rlarge/(2*R))**2)*erf((1-rlarge/(2*R))*root2gam)
+                        +R/root2gam*(R-rlarge/2)*exp(-4*gamma*(1-rlarge/(2*R))**2))
+#w2w2 = w2w2/w2w2.max()
[…]

### 2013-02-06 papers/polyhedra/paper.tex
work on generic solution for weighting functions of polyhedra Tested everything on node03.

@@
 
-\newcommand{\fixme}[1]{\red{[#1]}}
+\newcommand{\todo}[1]{\textcolor{red}{[TODO: #1]}}
 
@@
   \left(n(\rr,\rot)\ln(\Lambda^3\Theta^3 n(\rr,\rot)) - n(\rr,\rot)
-  \right) d\rr d\rot
+  \right) d\rr d\rot \label{eq:Aideal}
   \\
@@
 relatively little orientational ordering, it ought to be an accurate
-equation.  In any case I imagine it will be plausible.
+equation.
+
+In cases where $n_{lmm'} \approx n_{000}$, we could reasonbly become
+concerned that $n(\rr, \rot)$ could become negative for some
+orientations, which would be unphysical.  If this becomes a problem,
+we would want to address it by including nonlinear effects of the
+$\ln$ in Equation~\ref{eq:Aideal}.  For now, let's stick with the
+linear approximation, and hope for the best.  If we find that
+$n_{lmm'} > n_{000}$ in practice, we would want to revisit this decision.
 
@@
 
-\subsection{Finding the weighted densities}
+\subsection{Finding the $000$ weighting functions for any polyhedron}
 
@@
 \newcommand\facehat{\mathbf{\hat{f}}}
-\newcommand\Rface{\ensuremath{R_{\text{face}}}}
-\newcommand\Rvertex{\ensuremath{R_{\text{vertex}}}}
+\newcommand\Rface{\ensuremath{R_f}}
+\newcommand\Rvertex{\ensuremath{R_v}}
 
 Here I will discuss how to find the weighted densities we need in
-order to study polyhedra.  We begin with $n_3(\rr)$, the packing
-fraction.
+order to study polyhedra.  I will attempt in this section to evaluate
+the $000$ weighting function for an arbitrary polyhedron, specified in
+terms of the locations of its vertices and face centers (and possibly
+also edge centers, although they should be computable from the vertex
+and face radii).
+
+\subsubsection{$w_3$}
+We begin with the packing fraction $n_3(\rr)$.
 \begin{align}
@@
 $000$ integral is always going to give us something that is
-essentially the same, but I won't go into this yet.
+essentially the same.  This integral is simply:
+\begin{align}
+  w_3^{000}(r) = \int w_3(\rr') \delta(|\rr'|-r) d^3r'\label{eq:w3000}
+\end{align}
+or we could write it as an integral over solid angle:
+\begin{align}
+  w_3^{000}(r) = \int w_3(\rr) d\Omega
+\end{align}
+Let's suppose we have $N$ faces and $M$ vertices per face.  If the
+face center is at radius $\Rface$ and the vertex is at radius $\Rvertex$, then
+we can compute $w_3^{000}$ by integrating over the irregular
+tetrahedron enclosed by the for radii given by one vertex, one edge
+center and one face center, as depicted in Figure~\ref{fig:r3000}, and
+multiplying by $NM$ to account for the entire volume.  By doing this,
+we can solve once for $w_3^{000}(r)$ for any regular polyhedron.
 
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/w3-000}
+\caption{How we integrate for arbitrary polyhedron.}\label{fig:r3000}
+\end{figure}
+
+To begin, we solve for the positions of the vertices, which are shown
+in Figure~\ref{fig:r3000}.  To perform our integration, as in
+Equation~\ref{eq:w3000}, we will choose a coordinate system in which
+the center of the face is in the $\xhat$ direction, and the line from the
+center of the face to the center of the edge is in the $\yhat$
+direction.  Our integral then becomes:
+\begin{widetext}
+\begin{align}
+  w_3^{000}(r) &= \int w_3(\rr') \delta(|\rr'|-r) d^3r'
+  \\
+  &= NM \int_0^{\Rface} dx
+  \int_0^{\sqrt{\Rvertex^2-\Rface^2}\cos\left(\frac{2\pi}{M}\right)} dy
+  \int_0^{y\tan\left(\frac{2\pi}{M}\right)} dz \delta(r - \sqrt{x^2 +
+    y^2 + z^2})
+\end{align}
+\end{widetext}
+This integral gives a piecewise smooth result, with different forms
+when $r<\Rface$, $\Rface<r<R_e$, $R_e<r<\Rvertex$ and $\Rvertex<r$ (when it is zero).
+
+\todo{Do the above integral}
+
+\subsubsection{$w_2$}
 $n_2(\rr)$ is the area, which we can easily get from the above
@@
 \end{align}
-We could also just ask what the area is at a given position.
+Using the same notation as in the previous section, the integral
+for $w_2^{000}$ becomes:
+\begin{widetext}
+\begin{align}
+  w_2^{000}(r) &= \int w_2(\rr') \delta(|\rr'|-r) d^3r'
+  \\
+  &= NM
+  \int_0^{\sqrt{\Rvertex^2-\Rface^2}\cos\left(\frac{2\pi}{M}\right)} dy
+  \int_0^{y\tan\left(\frac{2\pi}{M}\right)} dz \delta(r - \sqrt{\Rface^2 +
+    y^2 + z^2})
+\end{align}
+\end{widetext}
+This integral gives a piecewise smooth result, with different forms
+when $r<\Rface$ (it is zero), $\Rface<r<R_e$, $R_e<r<\Rvertex$ and $\Rvertex<r$ (it is
+zero).
+
+\todo{Do the above integral}
+
+\subsubsection{$w_1$}
 
@@
 the polyhedron and the vertex, and $\Rface$ is the distance of the
-center of the face from the origin.  WARNING: I may have something off
-in these formulas, but the general idea is there.  To find a constant
-factor, we'd consider the
+center of the face from the origin.  \todo{WARNING: I may have something off
+in these formulas, but the general idea is there.  To check on a
+constant factor that may be off, we could check that the result
+approaches the spherical value for a dodecahedron.  If it's off, it's
+likely off by a largish factor.}
+
+As before, we should be able to do this integral for $w_1^{000}$
+independently of the specific polyhedron in question.  Our choice of
+coordinates will be different, as we only want to integrate over the
+edge, so we may as well put the edge center on the $\xhat$ axis, and
+the edge itself along the $\yhat$ axis.
+\begin{align}
+  w_1^{000}(r) &= \int w_1(\rr') \delta(|\rr'|-r) d^3r'
+  \\
+  &= NM \text{constant factor}
+  \int_0^{\sqrt{\Rvertex^2-\Rface^2}\sin\left(\frac{2\pi}{M}\right)} dy
+  \delta(r - \sqrt{R_e^2 + y^2})
+\end{align}
+The only tricky bit is to get this constant factor right, which will
+relate to the angle between adjacent faces.
+
+\todo{Do this integral}
+
+\todo{Find the constant factor}
+
+\subsubsection{$w_0$}
 
@@
 \end{align}
-Interestingly, this meas that
+Interestingly, this means that
 \begin{align}
@@
 from the center to the vertex of the polyhedron).
+
+\subsubsection{$\mathbf{w_{2V}}$}
+
+\todo{TODO}
+
+\subsubsection{$\mathbf{w_{1V}}$}
+
+\todo{TODO}
 

### 2013-02-07 papers/polyhedra/harmonics.tex
add empty paper about harmonics Tested everything on node08.

--- 2013-01-22 papers/polyhedra/wigner-properties.tex
@@
 
+\usepackage{color}
+
 \begin{document}
-\title{Properties of the Wigner D Function}
+\title{Harmonics with symmetries of platonic solids}
 
@@
 \begin{abstract}
-  This is just some notes to myself regarding the Wigner $D$ function,
-  which I hope to find useful in my research.
+  This is just some notes to myself regarding harmonics.
 \end{abstract}
@@
 
-\subsection{Properties of the Wigner $D$ matrix}
-
-There are various ways one can understand the Wigner $D$ function, but
-I think the simplest is to state that it defines how spherical
-harmonics behave under a rotation that I will call $\rot$, which we
-can think of as being defined by a set of Euler angles.  I will
-describe the rotation matrix as $R(\rot)$.  Without further ado, I
-will state that the $D$ functions (or matrices) are defined by:
-\begin{align}\label{eq:Ylm-rotation}
-  Y_{lm}(R(\rot)\rhat) &= \sum_{|m'|<l} D^{l}_{mm'}(\rot) Y_{lm'}(\rhat)
-\end{align}
-which is to say that the $D$ matrix defines how spherical harmonics
-transform under rotation.  $D$ functions also form an orthogonal basis
-set in the space of rotations (or orientations).
-\begin{align}
-  \int D^{l}_{mn}(\rot)^{*}D^{l'}_{m'n'}(\rot) d\rot
-  &= \frac{8\pi^2}{2l+1}\delta_{ll'}\delta_{mm'}\delta_{nn'}
-\end{align}
-Sadly, the $D$ matrices aren't normalized... but if they were, then
-that normalization would have popped up in
-Equation~\ref{eq:Ylm-rotation}, so you can't have everything.
-
-The complex conjugate of a $D$ matrix just corresponds to an inverse
-rotation:
-\begin{align}
-  D^{l}_{mm'}(\rot)^{*} &= (-1)^{m'-m}D^{l}_{-m-m'}(\rot)
-\end{align}
-
-\subsection{Properties of spherical harmonics}
-
-Spherical harmonics are properly orthonormal
-\begin{align}
-  \int Y_{lm}(\rhat)^{*}Y_{l'm'}(\rhat) d\solidangle
-  &= \delta_{ll'}\delta_{mm'}\delta_{nn'}
-\end{align}
-
-The complex conjugate of a $Y_{lm}$ corresponds to time reversal, or
-an opposite rotation (up to a sign):
-\begin{align}
-  Y_{lm}^{*}(\rhat) &= (-1)^{m} Y_{l,-m}(\rhat) \\
-  Y_{lm}(\rhat) &= (-1)^{m} Y_{lm}(-\rhat)
-\end{align}
+I plan on adding here an introduction to symmetrized linear
+combinations of spherical harmonics with the symmetries of platonic
+solids.
 

### 2013-02-12 papers/contact/paper.tex
added a new haskell gross correlation function and corrected gross section of paper first try Tested everything on node03.

--- 2012-12-12 papers/contact/paper.tex
@@
 \begin{align}
-  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - \frac{\pi}{12}\sigma^3n_A(\rr)}{\left(1 -
-    \frac{\pi}{6}\sigma^3n_A(\rr)\right)^3}
+  g_\sigma^\text{Gross,A}(\rr) &= \frac{1 - 0.5\bar{\eta}(\mathbf{r})}{\left(1 -
+    \bar{\eta}(\mathbf{r})\right)^3}
 \end{align}
-where $n_A$ is the averaged density defined in Equation~\ref{eq:nA}.
+where $\bar{\eta}$ is defined as
+\begin{align}
+\bar{\eta} &= \frac{1}{8} \int n(\rr') \Theta(\sigma -\left|\rr - \rr'\right|)
+  d\rr'
+\end{align}
 This formula is arrived at by using the density averaged over all
-spheres that could be touching a sphere at point~$\rr$ in the
+spheres that overlap a point~$\rr$ in the
 Carnahan-Starling equation for the correlation function at contact,

### 2013-02-12 papers/polyhedra/paper.tex

@@
   &=
-  k_BT \sum_{lmm'} \int
-  n_{lmm'}(\rr)D_{lmm'}(\rot)\left(
-  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right) - 1
+  k_BT \sum_{lmk} \int
+  n_{lmk}(\rr)D_{lmk}(\rot)\left(
+  \ln\left(\Lambda^3\Theta^3 \sum_{l'm'k'} n_{l'm'k'}(\rr)D_{l'm'k'}(\rot)\right) - 1
   \right) d\rr d\rot
@@
   &=
-  k_BT \int n_{000}(\rr)
-  \left(\ln\left(\Lambda^3\Theta^3 \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)\right) - 1
-  \right) d\rr d\rot
-  \\&+
-  k_BT \sum_{l>1,mm'} \int
-  n_{lmm'}(\rr)D_{lmm'}(\rot)
-  \ln\left(\Lambda^3\Theta^3 \sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)D_{l_1m_1m'_1}(\rot)\right)
-  d\rr d\rot
-  \\
-  &= A_{000} + \sum_{l>1,mm'} A_{lmm'} - k_BT\int n_{000}(\rr)d\rr
-\end{align}
-Here I have broken the $000$ term out of the outer summation over
-$lmm'$, since this one is special, since it is independent of $\rot$.
-We'll tackle that integral first, since it is a tad simpler.
-\begin{align}
-  A_{000} &= k_BT\int
-  n_{000}(\rr)\ln\left(
-  \sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)
-  \right)
-  d\rr d\rot
-\end{align}
-We can simplify this integral a bit if we write the log as a power
-series...
-\begin{align}
-  A_{000}
-  &= k_BT\int
-  n_{000}(\rr)\ln\left(
-  n_{000}(\rr)\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
-  \right)
-  d\rr d\rot
-  \\
-  &= k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)
-  \\&+k_BT\int
-  n_{000}(\rr)\ln\left(1 + \frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
-  d\rr d\rot
-  \\
-  &= k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)
-  \\&+k_BT\int
-  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)
-  d\rr d\rot
-  \\&-\frac{k_BT}{2!}\int
-  n_{000}(\rr)\left(\frac{\sum_{lmm'} n_{lmm'}(\rr)D_{lmm'}(\rot)}{n_{000}(\rr)}\right)^2
-  d\rr d\rot
-  \\&+\cdots
-\end{align}
-Now the linear term in the power series will vanish, since the
-summation is over $l>0$ only.  In the quadratic term, only a few terms
-will survive, and we'll end up with
-\begin{align}
-  A_{000}
-  &=
-  k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)d\rr
-  \\&\quad-\frac{k_BT}{2!}\int
-  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} n_{lmm'}(\rr)n_{l,-m,-m'}}{n_{000}(\rr)}
-  d\rr
-  \\&\quad+\cdots
-  \\
-  &=
-  k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)d\rr
-  \\&\quad-\frac{k_BT}{2!}\int
-  \frac{\sum_{lmm'} \frac{2l+1}{8\pi^2} |n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
-  d\rr
-  \\&\quad+\cdots
-\end{align}
-Now an interesting question is whether we can mimic this power series
-with a log, such as
-\begin{align}
-  A_{000}
-  &=
-  k_BT\int
-  n_{000}(\rr)\ln \left( \sqrt{n_{000}(\rr)^2 + \sum_{lmm'}
-    c_ln_{lmm'}(\rr)^2} \right)d\rr
-  \\
-  &= 
-  k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)d\rr
-  \\&\quad+
-  k_BT\int n_{000}(\rr)\ln\left( \sqrt{1 + \frac{\sum_{lmm'}
-    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)^2}} \right)d\rr
-  \\
-  &=
-  k_BT\int
-  n_{000}(\rr)\ln n_{000}(\rr)d\rr
-  \\&\quad+
-  k_BT\int \frac12 \frac{\sum_{lmm'}
-    c_ln_{lmm'}(\rr)^2}{n_{000}(\rr)} d\rr
-  \\&\quad+\cdots
-\end{align}
-from which we can conclude that to get an answer correct to the
-quadratic order, we need
-\begin{align}
-  c_l &= -\frac{2l+1}{8\pi^2}
-\end{align}
-which seems all right except for the sign:  I expect that the free
-energy should increase as we increase $n_{lmm'}$, since it means the
-system is becoming more ordered, and it looks like to this quadratic
-term, the free energy decreases... except that we're only looking at
-$A_{000}$ right now, and thus are neglecting a possibly significant
-increase due to the other terms...  so perhaps we should call this all
-right, and move on for now.
-
-The $A_{lmm'}$ terms are a bit different, but not too different:
-\begin{align}
-  A_{lmm'} &=
-  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
-  \ln\left(\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
-  D_{l_1m_1m'_1}(\rot)\right)
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3) - 1) +
+  k_BT \sum_{lmk} \int
+  n_{lmk}(\rr)D_{lmk}(\rot)
+  \ln\left(n_{000}(\rr) + \sum_{\substack{l'>0\\m'k'}} n_{l'm'k'}(\rr)D_{l'm'k'}(\rot)\right)
   d\rr d\rot
@@
   &=
-  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
-  \ln\left(1 + \frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
-  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}\right)
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3n_{000}(\rr)) - 1) +
+  k_BT \sum_{lmk} \int
+  n_{lmk}(\rr)D_{lmk}(\rot)
+  \ln\left(1 + \sum_{\substack{l'>0\\m'k'}} \frac{n_{l'm'k'}}{n_{000}(\rr)}(\rr)D_{l'm'k'}(\rot)\right)
   d\rr d\rot
@@
   &=
-  k_BT \int n_{lmm'}(\rr) D_{lmm'}(\rot)
-  \left(\frac{\sum_{l_1m_1m'_1} n_{l_1m_1m'_1}(\rr)
-  D_{l_1m_1m'_1}(\rot)}{n_{000}(\rr)}+\cdots\right)
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3n_{000}(\rr)) - 1)
+  \\&\quad+
+  k_BT \sum_{lmk} \int
+  n_{lmk}(\rr)D_{lmk}(\rot)
+  \left(\sum_{\substack{l'>0\\m'k'}}
+  \frac{n_{l'm'k'}(\rr)}{n_{000}(\rr)}D_{l'm'k'}(\rot)
+  - \frac12 \left(\sum_{\substack{l'>0\\m'k'}}
+    \frac{n_{l'm'k'}(\rr)}{n_{000}(\rr)}D_{l'm'k'}(\rot)
+  \right)^2
+  + \cdots \right)
   d\rr d\rot
@@
   &=
-  k_BT \int n_{lmm'}(\rr) \frac{8\pi^2}{2l+1} \frac{n_{l,-m,-m'}(\rr)}{n_{000}(\rr)}
-  d\rr + \cdots
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3n_{000}(\rr)) - 1)
+  %\\&\quad
+  +
+  k_BT \sum_{l>0mk}
+  \frac{8\pi^2}{2l+1}\frac{|n_{l'm'k'}(\rr)|^2}{n_{000}(\rr)}
+  - \frac12 \frac{8\pi^2}{2l+1}\frac{|n_{l'm'k'}(\rr)|^2}{n_{000}(\rr)}
+  + \cdots \label{eq:integrated}
   \\
   &=
-  k_BT \int \frac{8\pi^2}{2l+1} \frac{|n_{lmm'}(\rr)|^2}{n_{000}(\rr)}
-  d\rr + \cdots
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3n_{000}(\rr)) - 1)
+  +
+  k_BT \sum_{l>0mk}
+  \frac{4\pi^2}{2l+1}\frac{|n_{l'm'k'}(\rr)|^2}{n_{000}(\rr)}
+  + \cdots \\
+  &\approx
+  k_BT 8\pi^2 n_{000}(\rr)( \ln(\Lambda^3\Theta^3n_{000}(\rr)) - 1)
+  -
+  \frac{n_{000}(\rr)}{2}k_BT \ln\left(1 - \sum_{l>0mk}
+  \frac{8\pi^2}{2l+1}\frac{|n_{l'm'k'}(\rr)|^2}{n_{000}(\rr)}
+  \right) \label{eq:approxideal}
 \end{align}
-It is interesting how similar the $A_{lmm'}$ terms are to the
-quadratic terms in $A_{000}$, off by a factor of $-\frac12$.  This
-leads me to wonder about the possibility of writing
-\begin{align}
-  A_{\textit{ideal}} &\approx
-  k_BT \int n_{000}(\rr)\left( \ln\left(
-  \sqrt{\sum_{lmm'}\frac{2l+1}{16\pi^2}|n_{lmm'}|^2}
-  \right) - 1\right)d\rr
-\end{align}
-This would be correct up to the quadratic terms, and doesn't require
-any rotational integrals.  It also seems liable to be reasonable up to
-higher order terms, although I have no guarantee of that.  An
-interesting excersize would be to take this out to the fourth order,
-and also take the exact expression out to the fourth order and see how
-they compare.  They will actually differ in the third order terms, I
-expect, since the exact expression will indeed have nonzero third
-order terms.
+\end{widetext}
+In Equation~\ref{eq:integrated} we have used the fact that
+$D_{000}(\rot)=1$, and because $n(\rr,\rot)$ is real, $n_{l,-m,-k} =
+n_{lmk}^*$.  In Equation~\ref{eq:approxideal}, we have constructed an
+approxmation to the ideal gas free energy which has the correct
+first-order response in the orientational order $n_{l>0,mk}(\rr)$ and
+diverges (as it should) when $n_{l>0,mk}(\rr) \gg n_{000}(\rr)$, which
+would lead to the nonsense result that $n(\rr,\rot) < 0$.  We can tune
+the divergence a bit by moving constant factors (e.g. $8\pi^2$ into
+our out of the logarithm.
 
-So we have two plausible options that I can see: we could either go
-with this logarithmic approximation, or we could just go out to second
-order.  The latter is probably more defensible in terms of
-perturbation theory, and would give us something like
-\begin{align}
-  A_{\textit{ideal}} &\approx
-  k_BT \int \left(n(\rr)\ln n(\rr) - n(\rr) +
-  \sum_{lmm'}\frac{2l+1}{16\pi^2}\frac{|n_{lmm'}|^2}{n_{000}(\rr)}
-  \right)d\rr
-\end{align}
-where I have substituted the notation $n(\rr)$ for $n_{000}(\rr)$ to
-save space.
-
-\end{widetext}
 \subsection{Finding the orientational moments}
@@
 linear approximation, and hope for the best.  If we find that
-$n_{lmm'} > n_{000}$ in practice, we would want to revisit this decision.
+$n_{lmm'} > n_{000}$ in practice, we would want to revisit this
+decision, and make use of something like
+Equation~\ref{eq:approxideal}.
 

### 2013-03-08 papers/fuzzy-fmt/march-2013.tex

\documentclass{beamer}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{color}
\usepackage{cancel}
\graphicspath{ {/home/krebse/deft/papers/water-SAFT/figs/} }
\usetheme{Warsaw}
\newcommand{\rr}{\textbf{r}}
\title[Soft FMT]{Testing a ``soft'' Fundamental Measure Theory}
\author{Eric Krebs, Patrick Kreitzberg, David Roundy}
\institute{Oregon State University}
\date{}
\begin{document}

\begin{frame}
 \titlepage
\end{frame}

\begin{frame}{Motivation}
  \begin{itemize}
    \item We currently model water as a hard sphere fluid
    \item Hard spheres not physical for water
    \item We would like to ``soften'' spheres
  \end{itemize}
\end{frame}

\begin{frame}{Fundamental measure theory}
  \begin{itemize}
    \item Introduced by Rosenfeld in 1989
    \item Hard sphere fluid can be represented by fundamental measures
  \end{itemize}
  \begin{block}{weight functions for sphere of radius R}
    \begin{align}
      w_3(\rr) = \theta (|\rr| - R) \\
      w_2(\rr) = \delta (|\rr| - R)
    \end{align}
  \end{block}
\end{frame}

\begin{frame}{Schmidt paper (2001)}
  \begin{block}{Mayer Function}
    $f(r) = e^{-\beta V(r)} - 1$
    \begin{itemize}
      \item $\beta = (k_BT)^{-1}$
      \item $V(r)$ is a pair potential
    \end{itemize}
  \end{block}
  Schmidt shows that:
  \begin{align}
    \frac{\partial f(r)}{\partial r} = \int_{-\infty}^{\infty}w_2(r')
    \ast w_2(r - r') dr'
  \end{align}
\end{frame}

\begin{frame}{Choosing a $w_2(r)$}
  \begin{itemize}
    \item We considered linear and quadratic pair potentials, $V(r)$
    \item
  \end{itemize}
\end{frame}

\begin{frame}{theory results}
\end{frame}

\begin{frame}{Monte Carlo simulation results}
\end{frame}

\begin{frame}{Comparison}
\end{frame}

\begin{frame}{Finish}
  Thank you!
\end{frame}

\end{document}

### 2013-03-21 papers/fuzzy-fmt/march-2013.tex

@@
 \newcommand{\rr}{\textbf{r}}
-\title[Soft FMT]{Testing a ``soft'' Fundamental Measure Theory}
+\title[Testing Soft FMT]{Testing a ``soft'' Fundamental Measure Theory}
 \author{Eric Krebs, Patrick Kreitzberg, David Roundy}
 \institute{Oregon State University}
-\date{}
+\date{March, 20 2013}
 \begin{document}
@@
   \begin{itemize}
-    \item We currently model water as a hard sphere fluid
-    \item Hard spheres not physical for water
+    \item Want theory that works for inhomogeneous liquids
+    \item The hard sphere fluid is well studied
+    \item Hard spheres used as reference for real liquids
+      \begin{itemize}
+        \item Hard spheres are not very physical
+        \item Not computationally convenient (delta functions)
+      \end{itemize}
     \item We would like to ``soften'' spheres
   \end{itemize}
+  \begin{block}{Classical density functional theory}
+    \begin{align}
+    \Omega(T) = \underset{n(\rr)}{min} \left\{F[n(\rr),T] + \int
+                \left(V_{ext}(\rr) + \mu \right)n(\rr)d\rr \right\}
+    \end{align}
+    \begin{itemize}
+      \item $n(\rr)$ can be an inhomogeneous particle density
+    \end{itemize}
+  \end{block}
 \end{frame}
@@
   \begin{itemize}
+    \item Classical DFT for hard sphere fluid
     \item Introduced by Rosenfeld in 1989
-    \item Hard sphere fluid can be represented by fundamental measures
+    \item Hard sphere fluid represented by fundamental measures
+    \begin{itemize}
+      \item Volume, surface area, mean curvature
     \end{itemize}
-  \begin{block}{weight functions for sphere of radius R}
+    \item $n_{\alpha}(\rr) = \int n(\rr')w_{\alpha}(\rr - \rr')d\rr'$
+  \end{itemize}
+  \begin{block}{Some weight functions for sphere of radius R}
     \begin{align}
-      w_3(\rr) = \theta (|\rr| - R) \\
-      w_2(\rr) = \delta (|\rr| - R)
+      w_3(\rr) = \theta (R - |\rr|) \\
+      w_2(\rr) = \delta (R - |\rr|) \\
+      \notag
     \end{align}
@@
 
-\begin{frame}{Schmidt paper (2001)}
+\begin{frame}{Soft fundamental measure theory}
+  \begin{itemize}
+    \item Modification of FMT to treat not-so-hard spheres
+    \item Introduced by Schmidt in 1999
+    \item Weight functions are modified
+  \end{itemize}
+  Schmidt showed that:
+  \begin{align}
+    \int_{-\infty}^{\infty}w_2(r')w_2(r - r') dr' = \frac{\partial f(r)}{\partial r}
+  \end{align}
   \begin{block}{Mayer Function}
-    $f(r) = e^{-\beta V(r)} - 1$
+    \begin{center}
+      $f(r) = e^{-\beta V(r)} - 1$ \\
+    \end{center}
+\end{block}
   \begin{itemize}
-      \item $\beta = (k_BT)^{-1}$
-      \item $V(r)$ is a pair potential
+    \item If $w_2$ is known, all weight functions, $w_{\alpha}$'s, can
+      be derived
   \end{itemize}
-  \end{block}
-  Schmidt shows that:
-  \begin{align}
-    \frac{\partial f(r)}{\partial r} = \int_{-\infty}^{\infty}w_2(r')
-    \ast w_2(r - r') dr'
-  \end{align}
 \end{frame}
@@
   \begin{itemize}
-    \item We considered linear and quadratic pair potentials, $V(r)$
-    \item
+    \item Want spheres to overlap a little at ambient temperatures
+    \begin{itemize}
+      \item We use a harmonic potential in Mayer function\\
+      \includegraphics[width = 4cm]{figs/harmonic}
+      \includegraphics[width = 4cm]{figs/fprime}
+      \item Could not find simple analytic solution for $w_2$
+    \end{itemize}
+  \end{itemize}
+ %% \begin{align}
+ %%    w_2 &= \frac{2 \gamma r}{(\sqrt{\pi \gamma}-1)R^2} e^{-\gamma
+ %%       \left(1 - \frac{r}{2r_0} \right)^2} \Theta(r) \Theta(2r_0-r)\\
+ %%    \gamma &= \frac{1}{32} \left( \sqrt{\pi \beta V_0} + \sqrt{\pi \beta
+ %%        V_0 - 16\sqrt{\beta V_0}} \right)^2
+ %%  \end{align}
+\end{frame}
+
+\begin{frame}{Comparing $w_2$ with $f'(r)$}
+  \begin{itemize}
+    \item Recall: $\int_{-\infty}^{\infty}w_2(r')w_2(r - r') dr' = \frac{\partial f(r)}{\partial r}$
+    \item Chose a simple form of $w_2(r)$ with fitting parameters
+  \end{itemize}
+  \begin{center}
+    \includegraphics[trim = 0 13 0 37, clip, scale = 0.40]{figs/w2convolves}  \end{center}
+\end{frame}
+
+\begin{frame}{Weight functions comparison with FMT}
+  \begin{center}
+  \begin{tabular}{ccc}
+  %\includegraphics[trim = 0 13 0 37, clip, scale = 0.20]{figs/w_3} &
+  \includegraphics[width=4cm]{figs/w_3} \hspace{-1.5em} &
+  \includegraphics[width=4cm]{figs/w_2} \hspace{-1.5em} &
+  \includegraphics[width=4cm]{figs/w_1} \\
+  \includegraphics[width=4cm]{figs/step} \hspace{-1.5em} &
+  \includegraphics[width=4cm]{figs/delta} \hspace{-1.5em} &
+  \includegraphics[width=4cm]{figs/delta}  
+  \end{tabular}
+  \end{center}
+  %% \begin{tabular}{cc}
+  %% %\includegraphics[trim = 0 13 0 37, clip, scale = 0.20]{figs/w_3} &
+  %% \includegraphics[width=4cm]{figs/w_3} &
+  %% \includegraphics[width=4cm]{figs/step} \\
+  %% \includegraphics[width=4cm]{figs/w_2}
+  %% \vspace{-.7em} & \\
+  %% \includegraphics[width=4cm]{figs/w_1} &
+  %% \end{tabular}
+\end{frame}
+
+\begin{frame}{Homogeneous fluid}
+  \begin{center}
+    \includegraphics[trim = 0 13 0 35, clip, scale = 0.35]{figs/p-vs-packing}
+  \end{center}
+  \begin{itemize}
+    \item At low temps, should reduce to hard spheres
+    \item At low packing fraction, agrees with ideal gas
   \end{itemize}
@@
 
-\begin{frame}{theory results}
-\end{frame}
-
-\begin{frame}{Monte Carlo simulation results}
-\end{frame}
-
-\begin{frame}{Comparison}
-\end{frame}
-
-\begin{frame}{Finish}
-  Thank you!
+\begin{frame}{Conclusion}
+  \begin{itemize}
+    \item Soft FMT might be a possible alternative to hard spheres
+    \item Our approximate $w_2$ works well for homogeneous case
+    \item Reduces to hard spheres at low temps
+  \end{itemize}
+  \begin{block}{Future work}
+    \begin{itemize}
+      \item Compare with Monte Carlo simulation
+      \item Simulate soft spheres near a hard wall
+    \end{itemize}
+  \end{block}
 \end{frame}

### 2013-03-21 papers/polyhedra/harmonics.tex

@@
 
+\section{Cubic}
+
+Von der Lage and Bethe in 1947 give the first few cubic harmonics in
+Cartesian coordinates.  They refer to these as ``Kubic Harmonics.''
+We will want to convert these into spherical harmonics, using $x/\rho
+= \sin\theta\cos\phi$, etc.
+\begin{align}
+  f_0 &= 1 \\
+  f_4 &= \frac{5\sqrt{3\cdot 7}}{4}\left(
+  x^4 + y^4 + z^4 - \frac35 \rho^4
+  \right) \\
+  f_6 &= \frac{3\cdot 7 \cdot 11 \sqrt{2\cdot 13}}{8}\left(
+  x^2y^2z^2 + \frac1{22} f_4 - \frac1{105}\rho^6
+  \right)\\
+  f_8 &= \frac{5\cdot 13  \sqrt{3\cdot 11 \cdot 17}}{16}\left(
+  x^8 + y^8 - \frac{28}{f} f_6 \rho^2 - \frac{210}{143} f_4 \rho^4 - \frac16\rho^8
+  \right)
+\end{align}
+
+\section{Icosohedral}
+
+Zheng and Doerschuk give the icosohedral harmonics in terms of
+$\theta$ and $\phi$.  They also give formulas for these in terms of
+spherical harmonics, but these are hard to read.
+\begin{align}
+  I_{00}(\theta,\phi) &= 1 \\
+  I_{60}(\theta,\phi) &= 2^3 \cdot 3^2 \cdot 5 \cdot 11
+  P_{60}(\cos\theta) - P_{65}(\cos\theta)\cos(5\phi)
+  \\
+  P_{10,0} &= 2^8 \cdot 3^2 \cdot 5^2 \cdot 7 \cdot 13
+  \cdot 19 P_{10,10}
+\end{align}
+
+\section{Notes}
+
+The following are defined on page 5 of Zheng and Doerschuk:
+\begin{align}
+  \phi_k &\equiv \phi_0 + k \frac{2\pi}{5} \\
+  \cos \gamma_k &\equiv \cos\beta\cos\theta_0 + \sin\beta
+  \sin\theta_0\cos \phi_k \\
+  \sin \alpha_k &\equiv - \frac{\sin\theta_0 \sin\phi_k}{\sin\gamma_k}
+\end{align}
+
 \end{document}

### 2013-03-29 papers/pair-correlation/paper.tex
some previous approach text for pair-correlation Tested everything on darcy.
couple of equations on pair-correlation Tested everything on kipu.
added few changes to intro of paper.  Spelled some things wrong Tested everything on node07.

@@
 
+\section{Previous approaches}
+
+Gloor \emph{et al} explain that little is known of the pair
+distribution function of the hard-sphere fluid, and therefore use the
+formula
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
+\end{align}
+They go on to further approximate the radial distribution function,
+but we can just stop here.
+
+Gloor et al \cite{gloor2007prediction} uses average density of two points and the uses
+bulk radial distribution for the reference system (hard spheres) at
+this density, in order to approximate pair correlation.
+
+Gross \cite{gross2009density} does the same as Gloor et al except that
+he uses a hard chain reference fluid.
+
+Sokolowski and Fischer \cite{sokolowski1992role} do the same as Gloor et al except that instead of
+taking the exact densities $\rho(\mathbf{r_1})$ and $\rho(\mathbf{r_2})$ to
+average he takes an integral $\int_{V_{sphere}} \rho(\mathbf{r})d\mathbf{r}$
+over a sphere of some radius centered at each point.  It's
+these two averaged densities that he takes the average of.
+
+Fischer and Methfessel \cite{fischer1980born}, in a paper that is cited by Sokolowski and Fischer (above) do the same thing (the originate the idea of the spherical average density.
+Their Introduction has a very good discussion of the different averaging methods to get to $g(\mathbf{r})$ tried up until then (paper is written in 1980).
+
+Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but maybe you (Dr. Roundy)
+should take a look at it.  They reference Tang \cite{tang2008accurate} which I've
+mentioned below in the Appendix.
+
+McCabe and Kiselev \cite{mccabe2004application} model a square well potential.  One of their terms uses a radial
+distribution function that once again is taken to be the bulk function evaluated at an
+effective density.  They reference a source that calculates this effective density (or filling fraction) to be
+$\eta_{effective} = c_1\eta + c_2\eta^2 + c_3\eta^3$ where the coeffecients are gotten
+from three coupled polynomial equations with $1,\lambda$, and $\lambda^2$, where $\lambda$
+is a measure of the model's potential range.  They reference another source who calculates
+the effective density in a similar (but more complicated) way.  I can look more at these
+if you'd like.  They also use a different first order expansion formula for the ditribution
+at contact in the chain term.
+
+Toxvaerd \cite{toxvaerd1976hydrostatic}approximates the radial distribution as the radial distribution
+of the bulk at the same density.  Uses the first two terms in a tempurature
+expansion.  First term is just Carnahan Starling derived, second is empirical.
+
+>>>>>>> Stashed changes
+
 \section{Pair distribution function with inhomogeneity}
@@
 time.
+
+\begin{dmath}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, \eta(g_\sigma(\rr_1))) + g(|r_{12}|, \eta(g_\sigma(\rr_2)))}{2}
+\end{dmath}
+Bad:
+\begin{dmath}
+  g^{(2)}(\rr_1,\rr_2) = g\left(|r_{12}|, \frac{\eta(g_\sigma(\rr_1)) + \eta(g_\sigma(\rr_2))}{2}\right)
+\end{dmath}
 
@@
 \section*{Appendix}
+\cite{jin2011perturbative} Combines hard sphere (white bear) with attraction that is not mean field but rather uses two other correltation expressions given in  -\cite{tang2008accurate} and \cite{hlushak2009direct}
+
+Tang \cite{tang2008accurate}Derives the direct correltation for a SW and Yukawa and LJ potentials by first getting the radial distribution function through a method I don't really understand.
+Composes fourier transforms $c(k)$ and $h(k)$ of these complicated functions that depend on filling fraction and the potential in an integral.
+Uses first order mean spherical approximation.
+This is just square potential well around particle but the radial distribution function takes only the first order in an expansion in...
+(high temp?).
+
+\cite{hlushak2009direct}
+Cites \cite{tang2008accurate} (above) and does largely the same thing but then adds a complication for $2 < \lambda \leq 3$
+
+\cite{levesque2012scalar}
+Uses direct correlation funtion that's gotten from a radial distribution function of a homogeneous density.
+The direct correlation itself is only dependent on distance.
+The $\rho_0$ that they use is the same $\rho_0$ that the used in the reference HS term in their functional.
+The use the Carnahan Starling or the Perkis Yevick for these hard sphere free energies.
+The other terms involve $\delta\rho$ so the change is a variation from the reference density.
+They further modify it by looking at molecular dynamic simulations of a certain model of water molecules and looking at the oxygen-oxygen correlation.
 

### 2013-05-08 papers/water-saft/paper.tex
add paper for new water functional (not yet modified from old one) Tested everything on bennet.

--- 2012-12-21 papers/water-SAFT/paper.tex
@@
 
-\author{Jessica Hughes}
+\author{Eric J. Krebs}
 \affiliation{Department of Physics, Oregon State University,
   Corvallis, OR 97331, USA}
-\author{Eric J. Krebs}
+\author{Jeff B. Schulte}
 \affiliation{Department of Physics, Oregon State University,
@@
 \begin{abstract}
-We develop a classical density functional for water which combines the
-White Bear fundamental-measure theory (FMT) functional for the hard
-sphere fluid with attractive interactions based on the Statistical
-Associating Fluid Theory (SAFT-VR).  This functional reproduces the
-properties of water at both long and short length scales over a wide
-range of temperatures, and is computationally efficient, comparable to
-the cost of FMT itself.  We demonstrate our functional by applying it
-to systems composed of two hard rods, four hard rods arranged in a
-square and hard spheres in water.
+We develop an improved version of our old functional.
 \end{abstract}
@@
 \begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{n_2^2}.
 \end{align}

### 2013-05-08 papers/hughes-saft/figs/pressure-with-isotherms.cpp

@@ -25,7 +25,7 @@ const double kB = 3.16681539628059e-6; // This is Boltzmann's constant in Hartre
 
 int main(int, char **) {
   
-  FILE *o = fopen("papers/water-SAFT/figs/pressure-with-isotherms.dat", "w");
+  FILE *o = fopen("papers/hughes-saft/figs/pressure-with-isotherms.dat", "w");
   
   Functional f = OfEffectivePotential(SaftFluid2(water_prop.lengthscale,
                                                 water_prop.epsilonAB, water_prop.kappaAB,

### 2013-05-08 papers/hughes-saft/figs/sphere.cpp

@@ -1,0 +1,231 @@
+// Deft is a density functional package developed by the research
+// group of Professor David Roundy
+//
+// Copyright 2010 The Deft Authors
+//
+// Deft is free software; you can redistribute it and/or modify it
+// under the terms of the GNU General Public License as published by
+// the Free Software Foundation; either version 2 of the License, or
+// (at your option) any later version.
+//
+// You should have received a copy of the GNU General Public License
+// along with deft; if not, write to the Free Software Foundation,
+// Inc., 59 Temple Place, Suite 330, Boston, MA 02111-1307 USA
+//
+// Please see the file AUTHORS for a list of authors.
+
+#include <stdio.h>
+#include <time.h>
+#include "OptimizedFunctionals.h"
+#include "equation-of-state.h"
+#include "LineMinimizer.h"
+#include "utilities.h"
+#include "handymath.h"
+
+const double nm = 18.8972613;
+// Here we set up the lattice.
+double zmax = 2.5*nm;
+double ymax = 2.5*nm;
+double xmax = 2.5*nm;
+double diameter = 1*nm;
+bool using_default_diameter = true;
+
+double notinwall(Cartesian r) {
+  const double z = r.z();
+  const double y = r.y();
+  const double x = r.x();
+  if (sqrt(sqr(z)+sqr(y)+sqr(x)) < diameter/2) {
+      return 0; 
+  }
+  return 1;
+}
+
+void plot_grids_y_direction(const char *fname, const Grid &a) {
+  FILE *out = fopen(fname, "w");
+  if (!out) {
+    fprintf(stderr, "Unable to create file %s!\n", fname);
+    // don't just abort?
+    return;
+  }
+  const GridDescription gd = a.description();
+  const int x = 0;
+  const int z = 0;
+  for (int y=0; y<gd.Ny/2; y++) {
+    Cartesian here = gd.fineLat.toCartesian(Relative(x,y,z));
+    double ahere = a(x,y,z);
+    fprintf(out, "%g\t%g\t%g\t%g\n", here[0], here[1], here[2], ahere);
+  }
+  fclose(out);
+}
+
+int main(int argc, char *argv[]) {
+  clock_t start_time = clock();
+  if (argc > 1) {
+    if (sscanf(argv[1], "%lg", &diameter) != 1) {
+      printf("Got bad argument: %s\n", argv[1]);
+      return 1;
+    }
+    diameter *= nm;
+    using_default_diameter = false;
+  }
[…]

### 2013-05-08 papers/water-saft/figs/all-spheres.pl

@@ -1,0 +1,42 @@
+#!/usr/bin/perl -w
+
+use strict;
+
+my $dd;
+# We do the largest diameters first, so the small calculations won't
+# be scheduled on big-memory nodes leaving the big ones to wait for
+# them to finish.
+foreach $dd (0.3, 0.2, 0.15, 0.1, 0.05) {
+#foreach $dd (2.0, 1.6, 1.4, 1.2, 1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.15, 0.1, 0.05, 0) {
+#foreach $dd (0.2) {
+  my $d = sprintf("%04.2f", $dd);
+
+  print "Submitting sphere for $d nm\n";
+  my $padding = 1; # amount of extra space in nm that is added...
+  my $resolution = 0.2; # spacing of grid points in nm.
+
+  # Here I estimate the amount of memory that will be needed...
+  my $memuse = sprintf "%.0f", 1500*((($d + 2*$padding)/1.9)*(0.2/$resolution))**3;
+
+  my $scriptname = "papers/water-saft/figs/sphere-$d.tmp.sh";
+  open SCRIPT, ">$scriptname" or die $!;
+  print SCRIPT "#!/bin/sh
+#SBATCH --mem-per-cpu=$memuse
+##SBATCH --mail-type ALL
+##SBATCH --mail-user daveroundy\@gmail.com
+#SBATCH --output sphere-$d.out
+
+set -ev
+
+hostname
+date
+
+echo I think this will take $memuse megs of memory
+time nice -19 papers/water-saft/figs/sphere.mkdat $d
+
+date
+
+";
+  close(SCRIPT);
+  system("sbatch", $scriptname);
+}

### 2013-06-03 papers/fuzzy-fmt/figs/homogeneous.py
Updated w_2 in functions_plot.py Tested everything on node02.

@@ -30,9 +30,14 @@ P_cs = density*(1+eta+eta**2)/(1-eta)**3
 colors = { 0.1: 'r', 0.01: 'b', 0.001: 'g', 0.0001: 'k', 0.00001: 'm' }
 
 Temp = 0.00001
+eta = density*4*pi/3
+#P_cs = density*.001*(1+eta+eta**2-eta**3)/(1-eta)**3
+P_cs = density*.001*(1+eta+eta**2)/(1-eta)**3
+plot(eta,P_cs/.001, 'k',linewidth=2, label = 'Hard spheres')
 while Temp <= .011:
   betaV0 = V0/Temp
-  gamma = 2*((sqrt(pi*betaV0)+sqrt(pi*betaV0-16*sqrt(betaV0)))/8)**2
+  gamma = betaV0*(4+sqrt((4+4*sqrt(pi/betaV0)-2*pi*sqrt(pi/betaV0)+pi/betaV0)**2-4*pi**3/betaV0)+4*sqrt(pi/betaV0)-2*pi*sqrt(pi/betaV0)+pi/betaV0)/(2*pi**2)
+  #gamma = 2*((sqrt(pi*betaV0)+sqrt(pi*betaV0-16*sqrt(betaV0)))/8)**2
   sg = sqrt(gamma)
   #Integrals for the different weighted densities
   W3 = (-pi*R/(3*gamma**(3/2)*(sqrt(pi*gamma) -1)))*(2*sg*(8*(1 + gamma) - exp(-gamma)*(2*gamma+5))-sqrt(pi)*(4*gamma**2+12*gamma+3)*erf(sg))
@@ -79,7 +84,6 @@ for ff in arange(0.1,0.8, 0.1):
       print 'could not find', fname
 
 #plot(density*(4*pi/3), density, label = 'ideal gas')
-
 #mcdata = loadtxt('figs/mc-soft-homogenous-20-382-1.00000.dat.prs')
 #plot(mcdata[:,1],mcdata[:,0],'*')
 xlabel('packing fraction')

### 2013-06-19 papers/pair-correlation/paper.tex
cut bad text from tex and clean up diagram examples Tested everything on node37.
updated pair-dist paper.tex with an outline of sections and updated plot-ghs.py with a leastsq function that jeffs version of python has Tested everything on node09.

@@
 
-%\maketitle
+\section{Why does pair distribution matter?}
+\subsection{Interfaces are important}
+\subsection{cDFT theory enables treatment of inhomogeneous systems}
+\subsection{Pair distribution function is key input in thermodynamic perturbation theory}
+The pair distribution function is a key input into thermodynamic
+perturbation theory, which is widely used for treating dispersion
+interactions in fluids.  In most cases, the hard-sphere fluid is used
+as the reference system, making the pair distribution function of the
+hard-sphere system an important quantity.
+\subsection{Explain what the $a_1$ in thermodynamic perturbation theory is}
+
+\section{What has been done before?}
+\subsection{Gloor simple method, problematic}
+\subsection{Fischer accurate but slow}
+
+\section{Triplet correlation function?}
+
+\section{What is our new idea?}
+
+\section{What does it look like (graphs that show the function itself)?}
+\subsection{Fit to radial distribution function}
+%% Figure~\ref{fig:radial-distribution} shows...
+%% \begin{figure}
+%%   \Huge{Need figure here}
+%%   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
+%% \end{figure}
+%% \subsection{Pair distribution function}
+%% Figure~\ref{fig:pair-distribution} shows...
+%% \begin{figure}
+%%   \Huge{Need figure here}
+%%   \caption{Plot of the pair distribution function near a hard wall.
+%%     How do we plot this?}\label{fig:pair-distribution}
+%% \end{figure}
+
+%% \section{How well does it work?}
+%% Figure~\ref{fig:dadz} shows...
+%% \begin{figure}
+%%   \includegraphics[width=5cm]{figs/dadz}
+%%   \caption{Plot of $\frac{da_1}{dz}$ near a hard
+%%     wall. For what $\Phi$?}\label{fig:dadz}
+%% \end{figure}
+
+\section{THE END OF THE PAPER}
 
@@
 expansion.  First term is just Carnahan Starling derived, second is empirical.
-
->>>>>>> Stashed changes
 

### 2013-06-20 papers/water-saft/paper.tex
Added calculation and plot for X near a hydrophobic rod. Sarted on new water paper. Tested everything on node09.

@@
-\documentclass[twocolumn,amsmath,amssymb,prb]{revtex4-1}
+\documentclass[twocolumn,amsmath,amssymb,prl]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+
+\fixme{Some stuff about water}.  An important aspects of water is its
+ability to form hydrogen bonds.  Each water molecule contains two
+hydrogen bond donor sites and two acceptor sites.
+
+One promising approach for an efficient continuum description of water
+is that of classical density-functional theory (DFT), which is an
+approach for evaluating the free energy and thermally averaged density
+of fluids in an arbitrary external potential~\cite{ebner1976density}.
+The foundation of classical DFT is the Mermin
+theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature,
+stating that
+\begin{equation}
+  \Omega (T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int (V_\textit{ext}(\rr) +\mu)n(\rr)
+d\rr\right\},
+\end{equation}
+where $\Omega (T)$ is the grand potential of a system in the external
+potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
+of atoms or molecules, $\mu$ is the chemical potential and
+$F[n(\rr),T]$ is a universal free-energy functional for the fluid,
+which is independent of the external potential $V_\textit{ext}$.
+Classical DFT is a natural framework for creating a more flexible
+theory of hydrophobicity that can readily describe interaction of
+water with arbitrary external potentials---such as potentials
+describing strong interactions with solutes or surfaces.
+
+\subsection{Statistical Associating Fluid Theory}
+
+Statistical Associating Fluid Theory (SAFT) is a theory describing
+complex fluids in which hydrogen bonding plays a significant
+role\cite{chapman1989saft, muller2001molecular}.  SAFT is used to
+accurately model the equations of state of both pure fluids and
+mixtures over a wide range of temperatures and pressures.  SAFT is
+based on Wertheim's first-order thermodynamic perturbation theory
+(TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
+account for strong associative interactions between molecules.
+
+The SAFT Helmholtz free energy is composed of five terms:
+\begin{equation} \label{eq:SAFT-free-energy}
+  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
+  F_\textit{assoc} + F_\textit{chain},
+\end{equation}
+where the first three terms---ideal gas, hard-sphere repulsion and
+dispersion---encompass the \emph{monomer} contribution
+to the free energy, the fourth is the \emph{association} free energy,
+describing hydrogen bonds, and the final term is the chain formation
+energy for fluids that are chain polymers. In our model, we use a hard
+sphere reference fluid which renders $F_\textit{chain}$ irrelivant for
+our functional.
+
+In a recent paper \cite{hughes2013classical}, Hughes \emph{et al}
+constructed a classical density functional theory for water which used
+Clark \emph{et al}'s \cite{clark2006developing} five fitting
+paramenters for an optimal cDFT for water.  In addition, they
+introduced a parameter for the length scale over which the dispersion
+interaction is correlated
+
+\section{Method}
+
+Schulte \emph{et al}\cite{schulte2012using} derived expressions for
+the correlation fucntoin at contact based on the White Bear version of
+FMT. They found their $g_{\sigma}^A$ to agree with Monte Carlo
+simulation than Yu and Wu's which we was used in Hughes \emph{et al}'s
+paper.  We use Hughs \emph{et al}'s functional and replace the
+correlation with Schulte \emph{et al}'s, which we will call
+$g_{\sigma}$. \fixme{expand this}.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/surface-tension}
+\end{center}
+\caption{Comparison of Surface tension versus temperature for theoretical and
+  experimental data. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $s_d$ is fit so that the theoretical surface 
+  tension will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/single-rod-X-plot}
+\end{center}
+\caption{ The average number of hydrogen bonding per molecule at
+  distances from the center of a hydrophobic rod for various
+  radii. The solid lines are the results from using the older
+  $g_{\sigma}$, while the dotted lines are from Schulte \emph{et al}'s
+  asymmetric $g_{\sigma}$.}
+\label{fig:single-rod-X}
+\end{figure}
+
+\clearpage
+
 \section{Introduction}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/surface-tension}
-\end{center}
-\caption{Comparison of Surface tension versus temperature for theoretical and
-  experimental data. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
-  tension will match the experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
+%% \begin{figure}
+%% \begin{center}
+%% \includegraphics[width=3.5in]{figs/surface-tension}
+%% \end{center}
+%% \caption{Comparison of Surface tension versus temperature for theoretical and
+%%   experimental data. The experimental data is taken from NIST.\cite{nistwater}
+%%   The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
+%%   tension will match the experimental surface tension near room temperature.}
+%% \label{fig:surface-tension}
+%% \end{figure}
 

### 2013-06-20 papers/water-saft/figs/all-spheres.pl

@@ -6,9 +6,7 @@ my $dd;
 # We do the largest diameters first, so the small calculations won't
 # be scheduled on big-memory nodes leaving the big ones to wait for
 # them to finish.
-#foreach $dd (0.3, 0.2, 0.15, 0.1, 0.05) {
 foreach $dd (2.0, 1.6, 1.4, 1.2, 1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.15, 0.1, 0.05, 0) {
-#foreach $dd (0.2) {
   my $d = sprintf("%04.2f", $dd);
 
   print "Submitting sphere for $d nm\n";
@@ -16,7 +14,7 @@ foreach $dd (2.0, 1.6, 1.4, 1.2, 1.0, 0.9, 0.8, 0.7, 0.6, 0.5, 0.4, 0.3, 0.2, 0.
   my $resolution = 0.2; # spacing of grid points in nm.
 
   # Here I estimate the amount of memory that will be needed...
-  my $memuse = sprintf "%.0f", 3000*((($d + 2*$padding)/1.9)*(0.2/$resolution))**3;
+  my $memuse = sprintf "%.0f", 2750*((($d + 2*$padding)/1.9)*(0.2/$resolution))**3;
 
   my $scriptname = "papers/water-saft/figs/sphere-$d.tmp.sh";
   open SCRIPT, ">$scriptname" or die $!;

### 2013-06-20 papers/water-saft/figs/hughes-single-rod.cpp

@@ -101,59 +101,50 @@ int main(int argc, char **argv) {
   const double ymax = diameter+2*padding;
   const double zmax = diameter+2*padding;
 
-  Functional f = OfEffectivePotential(WaterSaft(new_water_prop.lengthscale,
-						new_water_prop.epsilonAB, new_water_prop.kappaAB,
-						new_water_prop.epsilon_dispersion,
-						new_water_prop.lambda_dispersion,
-						new_water_prop.length_scaling, 0));
-
-  double n_1atm = pressure_to_density(f, new_water_prop.kT, atmospheric_pressure,
+  Functional f = OfEffectivePotential(SaftFluid2(hughes_water_prop.lengthscale,
+						hughes_water_prop.epsilonAB, hughes_water_prop.kappaAB,
+						hughes_water_prop.epsilon_dispersion,
+						hughes_water_prop.lambda_dispersion,
+						hughes_water_prop.length_scaling, 0));
+  double n_1atm = pressure_to_density(f, hughes_water_prop.kT, atmospheric_pressure,
 					      0.001, 0.01);
 
-  double mu_satp = find_chemical_potential(f, new_water_prop.kT, n_1atm);
+  double mu_satp = find_chemical_potential(f, hughes_water_prop.kT, n_1atm);
 
-    
-  f = OfEffectivePotential(WaterSaft(new_water_prop.lengthscale,
-                                     new_water_prop.epsilonAB, new_water_prop.kappaAB,
-                                     new_water_prop.epsilon_dispersion,
-                                     new_water_prop.lambda_dispersion,
-                                     new_water_prop.length_scaling, mu_satp));
-
-  Functional X = WaterX(new_water_prop.lengthscale,
-                        new_water_prop.epsilonAB, new_water_prop.kappaAB,
-                        new_water_prop.epsilon_dispersion,
-                        new_water_prop.lambda_dispersion,
-                        new_water_prop.length_scaling, mu_satp);
+  f = OfEffectivePotential(SaftFluid2(hughes_water_prop.lengthscale,
+				     hughes_water_prop.epsilonAB, hughes_water_prop.kappaAB,
+				     hughes_water_prop.epsilon_dispersion,
+				     hughes_water_prop.lambda_dispersion,
+				     hughes_water_prop.length_scaling, mu_satp));
   
-  const double EperVolume = f(new_water_prop.kT, -new_water_prop.kT*log(n_1atm));
[…]

### 2013-06-20 src/haskell/HughesSaft.hs

@@ -9,7 +9,7 @@
 module HughesSaft
        ( yuwu_zeta, yuwu_correlation,
          eta_for_dispersion, lambda_dispersion, a1, a2, eta_effective,
-         saft_dispersion, saft_association, saft_fluid, saft_entropy, mu )
+         saft_dispersion, saft_association, saft_fluid, saft_entropy, hughes_X, mu )
        where
 
 import FMT ( rad, n, n0, n2, n3, sqr_n2v )
@@ -60,14 +60,14 @@ da1_detad = derive eta_for_dispersion 1 a1
 a2 = "a2" === 0.5*khs*epsilon_dispersion*eta_for_dispersion*da1_detad
      where khs = var "KHS" "{\\kappa_{HS}}" $ (1 - eta_for_dispersion)**4/(1 + 4*(eta_for_dispersion + eta_for_dispersion**2))
 
-saft_association = "Fassoc" === integrate (4*kT*n0*yuwu_zeta*(log xsaft - xsaft/2 + 1/2))
+saft_association = "Fassoc" === integrate (4*kT*n0*yuwu_zeta*(log hughes_X - hughes_X/2 + 1/2))
 
 kappa_association, epsilon_association :: Type a => Expression a
 kappa_association = s_tex "kappa_association" "\\kappa_a"
 epsilon_association = s_tex "epsilon_association" "\\epsilon_a"
 
-xsaft, deltasaft, a1, a2 :: Expression RealSpace
-xsaft = "X" === (sqrt(1 + 8*n0*yuwu_zeta*deltasaft) - 1) / (4*n0*yuwu_zeta*deltasaft)
+hughes_X, deltasaft, a1, a2 :: Expression RealSpace
+hughes_X = "X" === (sqrt(1 + 8*n0*yuwu_zeta*deltasaft) - 1) / (4*n0*yuwu_zeta*deltasaft)
 
 deltasaft = var "deltasaft" "{\\Delta}" $ gSW*kappa_association*boltz
   where boltz = "boltz" === exp(epsilon_association/kT)-1

### 2013-06-21 papers/pair-correlation/paper.tex

@@
 \subsection{Fit to radial distribution function}
-%% Figure~\ref{fig:radial-distribution} shows...
-%% \begin{figure}
-%%   \Huge{Need figure here}
-%%   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
-%% \end{figure}
-%% \subsection{Pair distribution function}
-%% Figure~\ref{fig:pair-distribution} shows...
-%% \begin{figure}
-%%   \Huge{Need figure here}
-%%   \caption{Plot of the pair distribution function near a hard wall.
-%%     How do we plot this?}\label{fig:pair-distribution}
-%% \end{figure}
+Figure~\ref{fig:radial-distribution} shows...
+\begin{figure}
+  \includegraphics[width=3cm]{figs/ghs-g}
+  \includegraphics[width=3cm]{figs/ghs-g-ghs}
+  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
+\end{figure}
+\subsection{Pair distribution function}
+Figure~\ref{fig:pair-distribution} shows...
+\begin{figure}
+  \includegraphics[width=6cm]{figs/pair-correlation-2-395-5}
+  \caption{Plot of the pair distribution function near a hard wall.
+    How do we plot this?}\label{fig:pair-distribution}
+\end{figure}
 
-%% \section{How well does it work?}
-%% Figure~\ref{fig:dadz} shows...
-%% \begin{figure}
-%%   \includegraphics[width=5cm]{figs/dadz}
-%%   \caption{Plot of $\frac{da_1}{dz}$ near a hard
-%%     wall. For what $\Phi$?}\label{fig:dadz}
-%% \end{figure}
+\section{How well does it work?}
+Figure~\ref{fig:dadz} shows...
+\begin{figure}
+  \includegraphics[width=3cm]{figs/dadz-3-2}
+  \includegraphics[width=3cm]{figs/dadz-3-3}
+  %%\includegraphics[width=3cm]{figs/dadz-4-3}
+  \caption{Plot of $\frac{da_1}{dz}$ near a hard
+    wall. For what $\Phi$?}\label{fig:dadz}
+\end{figure}
 

### 2013-06-25 papers/pair-correlation/paper.tex
added system to dadz.py so that if file is not there it does not exit just gives warning in the plot Tested everything on node39.

@@
 \subsection{Gloor simple method, problematic}
+Gil-Villegas\cite{gil1997statistical} approximates the first order inverse tenpurature Helmholtz Free energy term, which involves an integral of a potential multiplied by a radial distribution function $g^{HS}(r,\eta)$, by first taking the distribution function out of the integral and evaluating it at the correct mean distance $\zeta$, according to the mean value theorem.
+They then approximate the pair correlation function $g^{HS}(\zeta,\eta)$ with the correlation at contact and at a effective $\eta_{eff}$.
+The $\eta_{eff}$ is found by fitting a general function form to correct energy solutions $a_1$ with depenedent variable $\eta$ and
+
 \subsection{Fischer accurate but slow}
@@
 \begin{figure}
-  \includegraphics[width=6cm]{figs/pair-correlation-2-395-5}
+  %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-0.pdf}
+  \includegraphics[width=6cm]{figs/pair-correlation-2-395-5.pdf}
+  %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-10.pdf}
   \caption{Plot of the pair distribution function near a hard wall.
-    How do we plot this?}\label{fig:pair-distribution}
+    In the interest of space and memeory, we should look at the function and decided which data to add?}\label{fig:pair-distribution}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=3cm]{figs/dadz-3-2}
-  \includegraphics[width=3cm]{figs/dadz-3-3}
+  \includegraphics[width=3cm]{figs/dadz-3-2.pdf}
+  \includegraphics[width=3cm]{figs/dadz-3-3.pdf}
+  \includegraphics[width=3cm]{figs/dadz-2-2.pdf}
+  \includegraphics[width=3cm]{figs/dadz-2-3.pdf}
   %%\includegraphics[width=3cm]{figs/dadz-4-3}
@@
 results\cite{plischke1986pair}.  They plot the pair correlation
-function along interesting paths.
+function along interesting paths.  \fixme{Jeff read this paper for a
+  while and does not understand what they do in oder to find the pair
+  correlation function as compared to the radial distribution function}
 
@@
 
-Gloor et al \cite{gloor2007prediction} uses average density of two points and the uses
-bulk radial distribution for the reference system (hard spheres) at
-this density, in order to approximate pair correlation.
+Gloor et al \cite{gloor2007prediction} uses average density of two points:
+\begin{align}
+\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
+\end{align}
+and the uses bulk radial distribution for the reference system (hard
+spheres) at this density, in order to approximate pair correlation.
 
-Gross \cite{gross2009density} does the same as Gloor et al except that
-he uses a hard chain reference fluid.
+Evans \cite{evans1992density}perscribes to use the same approximation but with an unspecified 'mean denisty':
+
+\begin{align}
+g^{(2)}(\rr_1,\rr_2) &\equiv g_r(\rho^m,\rr_{12})
+\end{align}
+
+Gross \cite{gross2009density} uses a hard chain reference fluid.  In
+his functional he needs a pair correlation beween the different
+segments of chains.  He approximates this function with the radial
+distribution given by the Carnahan Starling at a density that is the
+average density of the segment positions $\rr$ and $\rr'$:
+
+\begin{align}
+\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
+\end{align}
+
+Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but maybe you (Dr. Roundy)
+should take a look at it.  They reference Tang \cite{tang2008accurate} which I've
+mentioned below in the Appendix.
 

### 2013-06-25 papers/pair-correlation/figs/plot-mc.py

@@ -45,10 +45,6 @@ def plot1d():
     gslice[i].set_data(numpy.linspace(0, rlen, num), zi)
     i += 1
 
-colors = ['k', 'b', 'r', 'g']
-plots = ['mc', 'this-work', 'fischer', 'gross']
-numplots = 4
-
 def plot():
   global ax, CS
   i = 0
@@ -187,7 +183,8 @@ z0_slider.on_changed(update)
 # angle slider
 angle_valinit = numpy.pi/3
 angax = pylab.axes([0.25, 0.035, 0.5, 0.025], axisbg='slategray')
-angslider = Slider(angax, 'theta', 0, numpy.pi/2, valinit = angle_valinit)
+angslider = Slider(angax, 'theta', 0, numpy.pi, valinit = 0)
+
 
 def upangle(val):
   global theta

### 2013-06-26 papers/pair-correlation/paper.tex

@@
 \section{What is our new idea?}
+We approximate the pair correlation with the average of two radial
+distribution functions, evaluated at the distance between the two
+points, that are themselves functions of our published pair
+correlation at contact $g_{sigma}$ evaluated at the two points.
 
+\begin{dmath}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
+\end{dmath}
+
+In the previous approaches mentioned above, function forms of radial
+distributions (used to approximate pair correlation) were dependent
+upon independent variables $\eta$, the density, wether at single
+points or averaged over sourrounding volumes.  We can achieve more
+accuracy by using instead a function dependent upon our $g_{\sigma}$,
+which holds more information about an inhomogeneous system
+then an average $\eta$ does.  We therefore need a function form for
+our radial distribution that is dependent upon $g_{\sigma}$ and that
+will approach the correct solution in the homogeneous limit and also
+in the limit of contact distance between the two points.
+
+The demand that our solution is correct in the homogeneous limit forms
+the basis of the derivation of our radial distribtuion's function
+form. We use the well known and very successful Carnahan Starling
+equation to calculated $g_{\sigma}$ and then simple monte-carlo
+simulations to calculate values of the homogeneous radial distribution
+function at varying distance perameter $r$ (Our previously published
+$g_{\sigma}$ approaches the Carnahan Starling version in the
+homogeneous limit and it's simplest to use an analytical expression
+for this calculation).  We then construct our radial distribution
+function that has the form of an expansion of powers of $h_{\sigma} =
+g_{\sigma}-1$ and perform a least squares fit to the monte-carlo
+solution to find the needed constant perameters.  The function is
+constructed in such a way as to appoach $g_{\sigma}$ in the limit of
+contact distance between the particles.  After this process we have a
+radial distribution function that is dependent only on $g_{\sigma}$
+and that gives the correct solution in the homogeneous and contact
+distance limit.
+
+The function has the form:
+
+\begin{align}
+  g(|\rr|) = 1 + g_{sigma}e^{-x_0r} + x_1g_{sigma}sin(x_2r)e^{-x_3r} -x_4g_{sigma}^2sin(x_5r)e^(-x_6r)
+\end{align}
+
+The finalalized form of our pair correlation function is:
+
+\begin{dmath}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
+\end{dmath}
+
+%% gsigma = gsigconcatenated[i] - 1
+%%         h0 = gsigma # was x[0]*gsig
+%%         f0 = numpy.exp(-x[0]*rconcatenated[i])
+%%         h1 = x[1]*gsigma
+%%         f1 = numpy.sin(x[2]*rconcatenated[i]) * numpy.exp(-x[3]*rconcatenated[i])
+%%         h2 = -x[4]*gsigma**(2)
+%%         f2 = numpy.sin(x[5]*rconcatenated[i]) * numpy.exp(-x[6]*rconcatenated[i])
+%%         g[i] = 1 + h0*f0 + h1*f1 + h2*f2
+%%     return g
+
+%% The homogeneoues limit $g_{sigm}$ can be calculated using the well
+%% known and very successful Carnahan Starling equation while the
+%% homogeneous limit radial distribution function can be calculated using
+%% Monte-Carlo methods.  We used these calculations in a least squares
+%% fit along the data points at increasing radial distances in order to
+%% fit the perameters of our more general pair correlation function.  Our
+%% function will therefore approach the most accurate solution possible
+%% in the homogeneous limit.  Created a function that has the form of an
+%% expansion in powers of $g_{sigma}$.  $g_{sigm}$ in the homogeneous
+%% limit can be calculated using the well known and very successful
+%% Carnahan Starling equation.  has a well known and Perameters in the
+%% expansion where fit
 \section{What does it look like (graphs that show the function itself)?}

### 2013-06-26 papers/pair-correlation/figs/plot-mc-modified.py
made alterations to plot-ghs and plot-mc-mod that allow for data not being loaded properly Tested everything on node10.

@@ -18,6 +18,7 @@ numplots = 4
 dx = 0.1
 ############################
 
+able_to_read_file = True
 
 z0 = 0.95
 theta = numpy.pi/2
@@ -68,6 +69,9 @@ def plot():
   while i < numplots:
     ax[i].collections = []
     g2[i] = read_walls(ff, z0, plots[i])
+    global able_to_read_file
+    if able_to_read_file == False:
+        break
 
     rmax = len(g2[i][:,0])*dx
     zmax = len(g2[i][0,:])*dx
@@ -89,17 +93,26 @@ def plot():
 zmax = 20
 rmax = 10
 
+
 def read_walls(ff, z0, fun):
-  if fun == 'mc':
-    filename = "figs/mc/wallsMC-pair-%1.1f-%1.2f.dat" % (ff, z0)
-  else:
-    filename = "figs/walls/wallsWB-%s-pair-%1.2f-%1.2f.dat" %(fun, ff, z0)
-  print 'Using', filename
-  if (os.path.isfile(filename) == False):
-    print "File does not exist."
-    sys.exit(1)
-  data = numpy.loadtxt(filename)
-  return data
+    if fun == 'mc':
+        filename = "figs/mc/wallsMC-pair-%1.1f-%1.2f.dat" % (ff, z0)
+        try:
+            data = numpy.loadtxt(filename)
+        except IOError:
+            global able_to_read_file
+            able_to_read_file = False
+            return 0
+    else:
+        filename = "figs/walls/wallsWB-%s-pair-%1.2f-%1.2f.dat" %(fun, ff, z0)
+        try:
+            data = numpy.loadtxt(filename)
+        except IOError:
+            global able_to_read_file
+            able_to_read_file = False
+            return 0
+    print 'Using', filename
+    return data
 
 # COLORS
 cdict = {'red':  ((0.0, 0.0, 0.0),
@@ -128,6 +141,13 @@ levels = numpy.linspace(0, 4, 49)
 
 g2 = [0]*numplots
 g2[0] = read_walls(ff, z0, 'mc')
+
+if able_to_read_file == False:
+    matplotlib.pyplot.plot(numpy.arange(0,10,1), [0]*10, 'k')
+    matplotlib.pyplot.suptitle('!!!!WARNING!!!!! There is data missing from this plot!', fontsize=25)
[…]

### 2013-06-27 papers/water-saft/paper.tex
Started work on new water paper.  Started using python to generate some figures. Tested everything on node10.

@@
 \newcommand{\refnote}{\red{[ref]}}
+
+\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
+\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
+\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
+\newcommand\kappaassoc{\ensuremath{\kappa_\textit{a}}}
+\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
+\newcommand\lscale{\ensuremath{s_d}}
 
@@
 
-\subsection{Statistical Associating Fluid Theory}
-
 Statistical Associating Fluid Theory (SAFT) is a theory describing
@@
 
-The SAFT Helmholtz free energy is composed of five terms:
-\begin{equation} \label{eq:SAFT-free-energy}
-  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
-  F_\textit{assoc} + F_\textit{chain},
-\end{equation}
-where the first three terms---ideal gas, hard-sphere repulsion and
-dispersion---encompass the \emph{monomer} contribution
-to the free energy, the fourth is the \emph{association} free energy,
-describing hydrogen bonds, and the final term is the chain formation
-energy for fluids that are chain polymers. In our model, we use a hard
-sphere reference fluid which renders $F_\textit{chain}$ irrelivant for
-our functional.
+\fixme{Talk about Wertheim's TPT}
 
 In a recent paper \cite{hughes2013classical}, Hughes \emph{et al}
-constructed a classical density functional theory for water which used
-Clark \emph{et al}'s \cite{clark2006developing} five fitting
-paramenters for an optimal cDFT for water.  In addition, they
-introduced a parameter for the length scale over which the dispersion
-interaction is correlated
+constructed a classical density functional theory for water which uses
+SAFT as a basis for their free energy functional.  Hughes employed Clark
+\emph{et al}'s \cite{clark2006developing} five fitting parameters for
+an optimal cDFT for water and, in addition, introduces a parameter for
+the length scale over which the dispersion interaction is correlated.
+Clark found that the four association site model works best to
+describe water, which is physically accurate since each water molecule
+has four hydrogen bonding sites.
+
+Schulte \emph{et al}\cite{schulte2012using} derived two expressions
+for the correlation function at contact for inhomogeneous fluids based
+on the White Bear version of fundamental measure theory (FMT). One
+expression was labeled as ``symmetric'' and found the average number
+of spheres touching a point $\rr$. The other expression, labeled
+``assymetric,'' sets $\rr$ at the center of a sphere, and finds the
+average spheres touching the solvent accessable surface.  They found
+the asymmetric correlation function, $g_{\sigma}^A$, to agree
+closely with Monte Carlo simulation.
 
@@
 
-Schulte \emph{et al}\cite{schulte2012using} derived expressions for
-the correlation fucntoin at contact based on the White Bear version of
-FMT. They found their $g_{\sigma}^A$ to agree with Monte Carlo
-simulation than Yu and Wu's which we was used in Hughes \emph{et al}'s
-paper.  We use Hughs \emph{et al}'s functional and replace the
-correlation with Schulte \emph{et al}'s, which we will call
-$g_{\sigma}$. \fixme{expand this}.
+We modify Hughes \emph{et al}'s method of creating a classical DFT
+based on SAFT using Clark's four site model with five fitting
+parameters and the added fitting parameter for the correlation
+distance of the dispersion interaction.  Hughe's association free
+energy term is modified to use Schulte's $g_{\sigma}^A$ rather than Yu
+and Wu's.  The association free energy for our four-site model has the
+form \fixme{I copy/pasted and modified these equations, but they still
+  require a number of other equations to come before.  Maybe only show
+  the $g_\sigma^A$?}
+\begin{align}
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) \zeta(\xx)
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx,
+\end{align}
+where the factor of $4$ comes from the four association sites per
+molecule, the functional $X$ is the fraction of association sites
+\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
+measure of the density inhomogeneity.
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{n_2^2}.
+\end{align}
+%
+The fraction $X$ is determined by the quadratic equation
+\begin{align}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
+      \Delta(\xx)} - 1}
+  {4 n_0(\xx)\zeta(\xx)
+    \Delta(\xx)}, \label{eq:X}
+\end{align}a
+where the functional $\Delta$ is a measure of hydrogen-bonding
+probability, given by
+\begin{align}
+  \Delta(\xx) &= \kappaassoc g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right) \\
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{A}_\sigma(\xx) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
+\end{align}
+% This is equation 77 of gil-villegas-1997-SAFT-VR
+where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+contact for a hard-sphere fluid with a square-well dispersion
+potential, and $a_1$ and $a_2$ are the two terms in the dispersion
+free energy.  The correlation function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere fluid
+correlation function $g^\textit{A}_\sigma$, for which we use the
+functional of Schulte \emph{et al}\cite{schulte2012using}:
+\begin{align}
+  g_\sigma^{A}(\rr) &= \frac{1}{n(\rr)n_A(\rr)}\frac{1}{k_B T 4 \pi
+    (2r)^2}\frac{\delta F_{hs}}{\delta R(\rr)} 
+\end{align}
+where the $n_A(\rr)$ is the density of spheres that are touching a
+sphere that is located at point $\rr$.\fixme{The equation for
+$g_\sigma^A$ requires finidng convolutions of local derivative of
+the free energy, aking this formulation computationally more
+expensive than the free energy itself.}
+
+We used Clark's five empirical parameters, and fit the calculated
+surface tension to experimental surface tension at ambient
+conditions.  This fit was found by choosing the value of the
+$\lscale$, which we found to be 0.454 for this new functional.
 
@@
 \begin{center}
+\includegraphics[width=3.5in]{figs/density-compare}
+\end{center}
+\caption{ Density profiles for single rods of different
+  radii from old and new functionals.}
+\label{fig:density-single-rod}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=3.5in]{figs/single-rod-X-plot}
 \end{center}
-\caption{ The average number of hydrogen bonding per molecule at
-  distances from the center of a hydrophobic rod for various
-  radii. The solid lines are the results from using the older
-  $g_{\sigma}$, while the dotted lines are from Schulte \emph{et al}'s
-  asymmetric $g_{\sigma}$.}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of hydrophobic rods of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}$.}
 \label{fig:single-rod-X}
 \end{figure}
+
+\section{Hydrogen bonding near a hydrophobic rod}
+
+We use X from eq. \ref{eq:X} above to find the average number of
+hydrogen bonds per molecule as a function of distance for different
+radius hydrophobic rods.  Figure \ref{fig:single-rod-X} compares our
+new functional with Hughes's functional, \fixme{Physical description
+  for why it looks that way}
+
+\section{Conclusion}
 
@@
 
-\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
-\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
-\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
-\newcommand\kappaassoc{\ensuremath{\kappa_\textit{a}}}
-\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
-\newcommand\lscale{\ensuremath{s_d}}
 \subsection{Dispersion free energy}
@@
 which results from the contact-value theorem.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/density-single-rod}
-\end{center}
-\caption{ Density profiles for single rods of different radii. The dotted line
-represents the saturated liquid density and the points represent the
-expected contact density derived from the contact value theorem and
-calculated free energy data.}
-\label{fig:density-single-rod}
-\end{figure}
 

### 2013-07-02 papers/pair-correlation/paper.tex
made alterations to plot-ghs and plot-mc-mod that allow for data not being loaded properly Tested everything on node10.

@@
 \begin{align}
-  g(|\rr|) = 1 + g_{sigma}e^{-x_0r} + x_1g_{sigma}sin(x_2r)e^{-x_3r} -x_4g_{sigma}^2sin(x_5r)e^(-x_6r)
+  g(|\rr|) &= 1 + g_{sigma}e^{-x_0(r-\sigma)} + x_1g_{sigma}sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
+  &~~-x_4g_{\sigma}^2sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
 \end{align}
@@
     In the interest of space and memeory, we should look at the function and decided which data to add?}\label{fig:pair-distribution}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:pair-distribution-1}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:pair-distribution-2}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:pair-distribution-3}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:pair-distribution-4}
 \end{figure}

### 2013-07-08 papers/water-saft/paper.tex
Added and improved water-saft plots.  Fixed truncation bug in hughes-single-rod Tested everything on node10.

--- 2013-05-08 papers/water-saft/paper.tex
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Introduction}
+
+\fixme{Some stuff about water}.  An important aspects of water is its
+ability to form hydrogen bonds.  Each water molecule contains two
+hydrogen bond donor sites and two acceptor sites.
+
+One promising approach for an efficient continuum description of water
+is that of classical density-functional theory (DFT), which is an
+approach for evaluating the free energy and thermally averaged density
+of fluids in an arbitrary external potential~\cite{ebner1976density}.
+The foundation of classical DFT is the Mermin
+theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature,
+stating that
+\begin{equation}
+  \Omega (T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int (V_\textit{ext}(\rr) +\mu)n(\rr)
+d\rr\right\},
+\end{equation}
+where $\Omega (T)$ is the grand potential of a system in the external
+potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
+of atoms or molecules, $\mu$ is the chemical potential and
+$F[n(\rr),T]$ is a universal free-energy functional for the fluid,
+which is independent of the external potential $V_\textit{ext}$.
+Classical DFT is a natural framework for creating a more flexible
+theory of hydrophobicity that can readily describe interaction of
+water with arbitrary external potentials---such as potentials
+describing strong interactions with solutes or surfaces.
+
+Statistical Associating Fluid Theory (SAFT) is a theory describing
+complex fluids in which hydrogen bonding plays a significant
+role\cite{chapman1989saft, muller2001molecular}.  SAFT is used to
+accurately model the equations of state of both pure fluids and
+mixtures over a wide range of temperatures and pressures.  SAFT is
+based on Wertheim's first-order thermodynamic perturbation theory
+(TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
+account for strong associative interactions between molecules.
+
+Wertheim's thermodynamic perturbation theory \fixme{expand/correct this}
+
+In a recent paper \cite{hughes2013classical}, Hughes \emph{et al}
+constructed a classical DFT for water which uses SAFT as a basis for
+their free energy functional.  Hughes employed Clark \emph{et al}'s
+\cite{clark2006developing} five fitting parameters for an optimal cDFT
+for water and, in addition, introduces a parameter for the length
+scale over which the dispersion interaction is correlated.  Clark
+found that the four association site model works best to describe
+water, which is physically accurate since each water molecule has four
+hydrogen bonding sites.
+
+This functional was successful in \fixme{list stuff}. However, the
+correlation function for hard spheres from Yu and
+Wu\cite{yu2002fmt-dft-inhomogeneous-associating} that is used in the
+association free energy term was \fixme{untested}.
+
+Schulte \emph{et al}\cite{schulte2012using} derived two approximate
+expressions for the correlation function at contact for and
+inhomogeneous hard sphere fluid based on the White Bear version of
+fundamental measure theory (FMT). One of Schulte's expression found
+the average number of spheres touching a point $\rr$ and is referred
+to as a ``symmetric'' correlation functoin, $g_\sigma^S$ . The other
+expression sets $\rr$ at the center of a sphere and find the average
+spheres touching the solvent accessable surface, which could be
+considered an ``asymmetric'' correlation function, $g_{\sigma}^A$. The
+asymmetric function agreed closely with Monte Carlo simulation, and
+much better than Yu and Wu's.
+
+\section{Method}
+
+We modify Hughes \emph{et al}'s method of creating a classical DFT
+based on SAFT using Clark's four site model with five fitting
+parameters and the added fitting parameter for the correlation
+distance of the dispersion interaction.  We modify Hughes's
+association free energy term to use Schulte's $g_{\sigma}^A$ rather
+than Yu and Wu's.  The association free energy for our four-site model
+has the form \fixme{I copy/pasted and modified these equations, but
+  they still require a number of other equations to come before.
+  Maybe only show the $g_\sigma^A$?}
+\begin{align}
+  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) \zeta(\xx)
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx,
+\end{align}
+where the factor of $4$ comes from the four association sites per
+molecule, the functional $X$ is the fraction of association sites
+\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
+measure of the density inhomogeneity.
+\begin{align}
+  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{n_2^2}.
+\end{align}
+%
+The fraction $X$ is determined by the quadratic equation
+\begin{align}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
+      \Delta(\xx)} - 1}
+  {4 n_0(\xx)\zeta(\xx)
+    \Delta(\xx)}, \label{eq:X}
+\end{align}a
+where the functional $\Delta$ is a measure of hydrogen-bonding
+probability, given by
+\begin{align}
+  \Delta(\xx) &= \kappaassoc g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right) \\
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{A}_\sigma(\xx) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
+\end{align}
+% This is equation 77 of gil-villegas-1997-SAFT-VR
+where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+contact for a hard-sphere fluid with a square-well dispersion
+potential, and $a_1$ and $a_2$ are the two terms in the dispersion
+free energy.  The correlation function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere fluid
+correlation function $g^\textit{A}_\sigma$, for which we use the
+functional of Schulte \emph{et al}\cite{schulte2012using}:
+\begin{align}
+  g_\sigma^{A}(\rr) &= \frac{1}{n(\rr)n_A(\rr)}\frac{1}{k_B T 4 \pi
+    (2R)^2}\frac{\delta F_{hs}}{\delta R(\rr)} 
+\end{align}
+where the $n_A(\rr)$ is the density of spheres that are touching a
+sphere that is located at point $\rr$. The modification to Hughes's
+functional appers here where we have replaced $g_\sigma^{HS}$ with
+$g_\sigma^A$. \fixme{The equation for $g_\sigma^A$ requires finding
+  convolutions of local derivative of the free energy, aking this
+  formulation computationally more expensive than the free energy
+  itself.}
+
+We use Clark's five empirical parameters, and fit the calculated
+surface tension to experimental surface tension at ambient conditions.
+We find these agree when $\lscale$ is 0.454.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/surface-tension}
+\end{center}
+\caption{Comparison of Surface tension versus temperature for theoretical and
+  experimental data. The experimental data is taken from NIST.\cite{nistwater}
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
+  tension will match the experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/density-compare}
+\end{center}
+\caption{ Density profiles for single rods of different
+  radii for the new functional (top) and the old functional
+  (bottom). Each separately colored line represents a density profile
+  for a rod; the black dotted line is the bulk density.}
+\label{fig:density-single-rod}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/single-rod-X-plot}
+\end{center}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of hydrophobic rods of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}^{HS}$.}
+\label{fig:single-rod-X}
+\end{figure}
+
+\section{Hydrogen bonding near a hydrophobic rod}
+
+We first simulate a single hydrophobic rod immersed in water. The
+density profile of the water with our functional is compared with that
+of Hughes in Fig. \ref{fig:density-single-rod}.  For small rods, the
+water densities are attenuated near the rods for our functional as
+compared to Hughes's.   Since we have changed the association term in
+the functional, a noticeable change might occur in the aveerage number
+of hydrogen bonds in a given system.  We compare the hydrogen bonds
+with the new and old functional for rods of a few different radii
+(Fig. \ref{fig:single-rod-X}) and find that there is a very noticeable
+difference.  This is not suprising since $g_\sigma^A$ is antisymmetric
+and $g\sigma^{HS}$ is symmetric.  \fixme{Physical description for why
+  it looks that way}
+
+\section{Hydrogen bonding near hard-sphere solutes}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/spheres-X-plot}
+\end{center}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of spherical cavities of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}$.}
+\label{fig:spheres-X}
+\end{figure}
+
+\fixme{More stuff here for spheres}
+
+\section{Conclusion}
+
+\clearpage
+
 \section{Introduction}

### 2013-07-08 papers/water-saft/figs/single-rod-X-plot.py

@@ -7,55 +7,44 @@ matplotlib.use('Agg')
 import math
 import matplotlib.pyplot as pyplot
 import numpy
+import pylab
 
-def plot_X(fin, col, lin):
-    nm = 18.8972613
-    x = []
-    y = []
-    for line in fin:
-        current = str(line)
-        pieces = current.split('\t')
-        x.append(float(pieces[0])/nm)
-        y.append(4.0*(1-float(pieces[2])))
-    pyplot.plot(x, y, color = col, linestyle = lin)
+# def plot_X(fin, radius, col, lin):
+#     # data = pylab.loadtxt('filename')
+#     nm = 18.8972613
+#     x = []
+#     y = []
+#     for line in fin:
+#         current = str(line)
+#         pieces = current.split('\t')
+#         x.append(float(pieces[0])/nm)
+#         y.append(4.0*(1-float(pieces[2])))
+#     pyplot.plot(x, y, color = col, linestyle = lin)
+#     pyplot.vlines(x = radius, ymin = 0, ymax = 4, color=col, linestyles=':')
 
 #colors for plot
-col1 = "#99ff88"
-col2 = "#44dd55"
-col3 = "#33bb88"
-col4 = "#337799"
-col5 = "#1133aa"
-col6 = "#000044"
+#col1 = "#441155"
+colors = ["#990022", "#dd6622", "#eecc22", "#22b544", "#3333bb"]
+radii = [0.2, 0.6, 1.0]
 
-oldline = "-"
-newline = ":"
+# oldline = "--"
+# newline = "-"
 
-#make plot from files
-rod1 = open('figs/single-rod-slice-00.1.dat', 'r')
-plot_X(rod1, col1, newline)
-rod2 = open('figs/single-rod-slice-00.3.dat', 'r')
-plot_X(rod2, col2, newline)
-rod3 = open('figs/single-rod-slice-00.6.dat', 'r')
-plot_X(rod3, col3, newline)
-rod4 = open('figs/single-rod-slice-01.0.dat', 'r')
-plot_X(rod4, col4, newline)
-rod5 = open('figs/single-rod-slice-01.6.dat', 'r')
-plot_X(rod5, col5, newline)
-rod6 = open('figs/single-rod-slice-02.0.dat', 'r')
-plot_X(rod6, col6, newline)
+nm = 18.8972613
[…]

### 2013-07-31 papers/water-saft/paper.tex
Updated water-saft stuff for new paper Tested everything on node09.

--- 2013-06-20 papers/water-saft/paper.tex
@@
 
-\fixme{Some stuff about water}.  An important aspects of water is its
-ability to form hydrogen bonds.  Each water molecule contains two
-hydrogen bond donor sites and two acceptor sites.
+Water \fixme{stuff about water}.  An important aspects of water
+is its ability to form hydrogen bonds.  Each water molecule contains
+two hydrogen bond donor sites and two acceptor sites.
 
@@
 theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature,
-stating that
-\begin{equation}
-  \Omega (T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int (V_\textit{ext}(\rr) +\mu)n(\rr)
-d\rr\right\},
-\end{equation}
-where $\Omega (T)$ is the grand potential of a system in the external
-potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
-of atoms or molecules, $\mu$ is the chemical potential and
-$F[n(\rr),T]$ is a universal free-energy functional for the fluid,
-which is independent of the external potential $V_\textit{ext}$.
+theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
 Classical DFT is a natural framework for creating a more flexible
@@
 accurately model the equations of state of both pure fluids and
-mixtures over a wide range of temperatures and pressures.  SAFT is
-based on Wertheim's first-order thermodynamic perturbation theory
-(TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+mixtures over a wide range of temperatures and pressures. In a recent
+paper \cite{hughes2013classical}, we constructed a classical DFT for
+water which uses SAFT as a basis for a free energy functional. We
+used Clark \emph{et al}'s \cite{clark2006developing} five fitting
+parameters for an optimal cDFT for water and, in addition, introduced
+a parameter for the length scale over which the dispersion interaction
+is correlated.
+
+SAFT is based on Wertheim's first-order thermodynamic perturbation
+theory (TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
   wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
-account for strong associative interactions between molecules.
-
-The SAFT Helmholtz free energy is composed of five terms:
-\begin{equation} \label{eq:SAFT-free-energy}
-  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
-  F_\textit{assoc} + F_\textit{chain},
-\end{equation}
-where the first three terms---ideal gas, hard-sphere repulsion and
-dispersion---encompass the \emph{monomer} contribution
-to the free energy, the fourth is the \emph{association} free energy,
-describing hydrogen bonds, and the final term is the chain formation
-energy for fluids that are chain polymers. In our model, we use a hard
-sphere reference fluid which renders $F_\textit{chain}$ irrelivant for
-our functional.
-
-In a recent paper \cite{hughes2013classical}, Hughes \emph{et al}
-constructed a classical density functional theory for water which used
-Clark \emph{et al}'s \cite{clark2006developing} five fitting
-paramenters for an optimal cDFT for water.  In addition, they
-introduced a parameter for the length scale over which the dispersion
-interaction is correlated
+account for strong associative interactions between
+molecules. Wertheim's thermodynamic perturbation theory uses a
+hard-sphere reference fluid where each sphere has a number of
+association sites on the surface. The association sites have an
+effective range and a small energy contribution when a site is within
+range of another hard-sphere's site. The free energy due to this
+association depends on the radial distribution function
+$g_\sigma^\text{HS}(\rr)$. Our previous water functional used Yu and
+Wu's radial distribution function
+\cite{yu2002fmt-dft-inhomogeneous-associating}, which was found to
+agree less with hard-sphere Monte-Carlo simulation than a distribution
+function developed by us in a recent paper.\cite{schulte2012using} In
+our paper, we developed an approximation for correlation at
+contact to describe the hard-sphere fluid.
 
@@
 
-Schulte \emph{et al}\cite{schulte2012using} derived expressions for
-the correlation fucntoin at contact based on the White Bear version of
-FMT. They found their $g_{\sigma}^A$ to agree with Monte Carlo
-simulation than Yu and Wu's which we was used in Hughes \emph{et al}'s
-paper.  We use Hughs \emph{et al}'s functional and replace the
-correlation with Schulte \emph{et al}'s, which we will call
-$g_{\sigma}$. \fixme{expand this}.
+We modify our previous the association free energy term to use our
+$g_{\sigma}^A$ rather than Yu and Wu's $g_\sigma^{YuWu}$.  The
+association free energy for our four-site model has the form:
+\begin{align}
+  F_\text{assoc}[n] &= k_BT \int n_\text{site}(\xx)
+  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
+\end{align}
+where
+\begin{align}
+  n_\text{site}(\rr) &=
+  \begin{cases}
+    4 n(\rr) & \text{this work}\\
+    4 n_0(\rr) \zeta(\rr) & \text{previous paper}.
+  \end{cases} 
+\end{align}
+The four in $n_{site}(\rr)$ comes from the four hydrogen bond sites
+and, $n_o(\rr)$ is the average density at contact, and $\zeta(\xx)$ is
+a dimensionless measure of the density inhomogeneity from Yu and
+Wu. \cite{yu2002fmt-dft-inhomogeneous-associating} The functional
+$X(\rr)$ is the fraction of association sites \emph{not}
+hydrogen-bonded, which is determined by the quadratic equation
+\begin{align}
+  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
+      \kappaassoc g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right)} - 1}
+  {4 n_0(\xx)\zeta(\xx)
+    \kappaassoc g^\textit{SW}_\sigma(\xx)
+  \left(e^{-\beta\epsilonassoc} - 1\right)}, \label{eq:X}
+\end{align}
+where
+\begin{align}
+  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
+  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
+  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
+\end{align}
+% This is equation 77 of gil-villegas-1997-SAFT-VR
+where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+contact for a hard-sphere fluid with a square-well dispersion
+potential, and $a_1$ and $a_2$ are the two terms in the dispersion
+free energy.  The radial distribution function $g^\textit{SW}_\sigma$ is
+written as a perturbative correction to the hard-sphere fluid
+correlation function $g^\textit{HS}_\sigma$. Four our old functional
+we used Yu and Wu's radial distribution function, and for this paper we
+use our $g^\textit{A}_\sigma$.
+
+We use Clark's five empirical parameters, and fit the calculated
+surface tension to experimental surface tension at ambient conditions.
+We find these agree when $\lscale$ is 0.454.
 
@@
   experimental data. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $s_d$ is fit so that the theoretical surface 
+  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
   tension will match the experimental surface tension near room temperature.}
@@
 \begin{center}
+\includegraphics[width=3.5in]{figs/density-compare}
+\end{center}
+\caption{ Density profiles for single rods of different
+  radii for the new functional (top) and the old functional
+  (bottom). Each separately colored line represents a density profile
+  for a rod; the black dotted line is the bulk density.}
+\label{fig:density-single-rod}
+\end{figure}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=3.5in]{figs/single-rod-X-plot}
 \end{center}
-\caption{ The average number of hydrogen bonding per molecule at
-  distances from the center of a hydrophobic rod for various
-  radii. The solid lines are the results from using the older
-  $g_{\sigma}$, while the dotted lines are from Schulte \emph{et al}'s
-  asymmetric $g_{\sigma}$.}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of hydrophobic rods of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}^{HS}$.}
 \label{fig:single-rod-X}
 \end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/single-rod-broken-HB}
+\end{center}
+\caption{ Broken hydrogen bonds broken when rods are placed in
+  homogeneous water.  \fixme{Our old and new functional are both plotted here}}
+\label{fig:single-rod-broken-HB}
+\end{figure}
+
+\section{Hydrogen bonding near a hydrophobic rod}
+
+We first simulate a single hydrophobic rod immersed in water. The
+density profile of the water with our functional is compared with that
+of Hughes in Fig. \ref{fig:density-single-rod}. The density at contact
+with the rods are nearly identical, even more so for larger radius
+rods. Since we have changed the association term in the functional, a
+difference will occur in the average number of hydrogen bonds in a
+given system. We compare the hydrogen bonds with the new and old
+functional for rods of a few different radii
+(Fig. \ref{fig:single-rod-X}) and find that there is a notable
+difference.
+
+\section{Hydrogen bonding near hard-sphere solutes}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/spheres-X-plot}
+\end{center}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of spherical cavities of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}$.}
+\label{fig:spheres-X}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/sphere-broken-HB}
+\end{center}
+\caption{ \fixme{spheres!}}
+\label{fig:spheres-broken-HB}
+\end{figure}
+
+We continue our study into three dimensions with a hard sphere solute
+in water.\fixme{stuff}.
+
+\section{Conclusion}
+
+We have modified a classical DFT for water with a more accurate radial
+distribution function which changes the way that the molecules assosciate.
 

### 2013-07-31 papers/water-saft/figs/hughes-sphere.cpp

@@ -40,7 +40,7 @@ double notinwall(Cartesian r) {
   return 1;
 }
 
-void plot_grids_y_direction(const char *fname, const Grid &a, const Grid &b) {
+void plot_grids_y_direction(const char *fname, const Grid &a, const Grid &b, const Grid &c) {
   FILE *out = fopen(fname, "w");
   if (!out) {
     fprintf(stderr, "Unable to create file %s!\n", fname);
@@ -54,7 +54,8 @@ void plot_grids_y_direction(const char *fname, const Grid &a, const Grid &b) {
     Cartesian here = gd.fineLat.toCartesian(Relative(x,y,z));
     double ahere = a(x,y,z);
     double bhere = b(x,y,z);
-    fprintf(out, "%g\t%g\t%g\t%g\t%g\n", here[0], here[1], here[2], ahere, bhere);
+    double chere = c(x,y,z);
+    fprintf(out, "%g\t%g\t%g\t%g\t%g\t%g\n", here[0], here[1], here[2], ahere, bhere, chere);
   }
   fclose(out);
 }
@@ -93,6 +94,12 @@ int main(int argc, char *argv[]) {
 				     hughes_water_prop.length_scaling, mu_satp));
 
   Functional X = WaterX(hughes_water_prop.lengthscale,
+                        hughes_water_prop.epsilonAB, hughes_water_prop.kappaAB,
+                        hughes_water_prop.epsilon_dispersion,
+                        hughes_water_prop.lambda_dispersion,
+                        hughes_water_prop.length_scaling, mu_satp);
+  
+  Functional HB = HughesHB(hughes_water_prop.lengthscale,
                         hughes_water_prop.epsilonAB, hughes_water_prop.kappaAB,
                         hughes_water_prop.epsilon_dispersion,
                         hughes_water_prop.lambda_dispersion,
@@ -144,18 +151,18 @@ int main(int argc, char *argv[]) {
                                 PreconditionedConjugateGradient(f, gd, hughes_water_prop.kT, 
                                                                 &potential,
                                                                 QuadraticLineMinimizer));
-      
+
       printf("\nDiameter of sphere = %g bohr (%g nm)\n", diameter, diameter/nm);
-      
+
       const int numiters = 200;
       for (int i=0;i<numiters && min.improve_energy(true);i++) {
[…]

### 2013-07-31 papers/water-saft/figs/single-rod.cpp

@@ -125,6 +125,12 @@ int main(int argc, char **argv) {
                         new_water_prop.lambda_dispersion,
                         new_water_prop.length_scaling, mu_satp);
   
+  Functional HB = HughesHB(hughes_water_prop.lengthscale,
+                        hughes_water_prop.epsilonAB, hughes_water_prop.kappaAB,
+                        hughes_water_prop.epsilon_dispersion,
+                        hughes_water_prop.lambda_dispersion,
+                        hughes_water_prop.length_scaling, mu_satp);
+  
   const double EperVolume = f(new_water_prop.kT, -new_water_prop.kT*log(n_1atm));
   const double EperNumber = EperVolume/n_1atm;
   const double EperCell = EperVolume*(zmax*ymax - 0.25*M_PI*diameter*diameter)*width;
@@ -178,18 +184,22 @@ int main(int argc, char **argv) {
   }
 
   Grid density(gd, EffectivePotentialToDensity()(new_water_prop.kT, gd, potential));
+  Grid zeroed_out_density(gd, density.cwise()*constraint); // this is zero inside the rod!  :;
   //printf("The bulk energy per cell should be %g\n", EperCell);
   //printf("The bulk energy based on number should be %g\n", EperNumber*density.integrate());
   //printf("Number of water molecules is %g\n", density.integrate());
   double energy = (min.energy() - EperCell)/width;
   energy = (min.energy() - EperNumber*density.integrate())/width;
   //printf("Energy is %.15g\n", energy);
+  Grid X_values(gd, X(new_water_prop.kT, gd, density));
+  Grid H_bonds_grid(gd, zeroed_out_density.cwise()*(4*(VectorXd::Ones(gd.NxNyNz)-X_values)));
+  const double broken_H_bonds = (HB(new_water_prop.kT, n_1atm)/n_1atm)*zeroed_out_density.integrate() - H_bonds_grid.integrate();
 
   char *datname = new char[1024];
   sprintf(datname, "papers/water-saft/figs/single-rod-%04.2fnm-energy.dat", diameter/nm);
   FILE *o = fopen(datname, "w");
   delete[] datname;
-  fprintf(o, "%g\t%.15g\n", diameter/nm, energy);
+  fprintf(o, "%g\t%.15g\t%g\n", diameter/nm, energy, broken_H_bonds);
   fclose(o);
 
   {
@@ -197,16 +207,14 @@ int main(int argc, char **argv) {
[…]

### 2013-07-31 papers/water-saft/paper.tex
edit beginning of introduction Tested everything on node09.

@@
 
-Water \fixme{stuff about water}.  An important aspects of water
-is its ability to form hydrogen bonds.  Each water molecule contains
-two hydrogen bond donor sites and two acceptor sites.
-
-One promising approach for an efficient continuum description of water
-is that of classical density-functional theory (DFT), which is an
-approach for evaluating the free energy and thermally averaged density
-of fluids in an arbitrary external potential~\cite{ebner1976density}.
-The foundation of classical DFT is the Mermin
-theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
-Classical DFT is a natural framework for creating a more flexible
-theory of hydrophobicity that can readily describe interaction of
-water with arbitrary external potentials---such as potentials
-describing strong interactions with solutes or surfaces.
-
-Statistical Associating Fluid Theory (SAFT) is a theory describing
-complex fluids in which hydrogen bonding plays a significant
+Water, the universal solvent, is of critical practical importance, and
+a continuum description of water is in high demand for a solvation
+model.  A number of recent attempts to develop improved solvation
+models for water have built on the approach of classical
+density-functional theory (DFT)~\fixme{add citations here,
+  particularly Jeanmairet 2013 and his citations 35-44, possibly
+  cutting non-acqueous papers from that list}.
+Classical DFT is based on a description of a fluid written as a free
+energy functional of the density distribution.  There are two
+approaches used to construct a classical DFT for water.  The first is
+to choose a functional form that is fit to properties of the bulk
+liquid at a given temperature and pressure~\fixme{cite Jeanmairet and
+  similar}.  Using this approach, it is possible to construct a
+functional that reproduces the exact second-order response function of
+the liquid under the fitted conditions.  This class of functional,
+however, will be less accurate at other temperatures or
+pressures---and in the inhomogeneous scenarios in which solvation
+models are applied.  The second approach is construct a functional by
+applying liquid-state theory to a model system, and then fit the model
+to experimental data such as the equation of state.
+
+%% \fixme{stuff about water}.  An important aspects of water
+%% is its ability to form hydrogen bonds.  Each water molecule contains
+%% two hydrogen bond donor sites and two acceptor sites.
+%% %
+%% One promising approach for an efficient continuum description of water
+%% is that of classical density-functional theory (DFT), which is an
+%% approach for evaluating the free energy and thermally averaged density
+%% of fluids in an arbitrary external potential~\cite{ebner1976density}.
+%% The foundation of classical DFT is the Mermin
+%% theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
+%% theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
+%% Classical DFT is a natural framework for creating a more flexible
+%% theory of hydrophobicity that can readily describe interaction of
+%% water with arbitrary external potentials---such as potentials
+%% describing strong interactions with solutes or surfaces.
+
+A widely used family of models used in the development of classical
+density functionals is based on Statistical Associating Fluid Theory
+(SAFT).
+%
+SAFT is a theory for fluids in which hydrogen bonding plays a significant
 role\cite{chapman1989saft, muller2001molecular}. SAFT is used to
 accurately model the equations of state of both pure fluids and
-mixtures over a wide range of temperatures and pressures. In a recent
-paper \cite{hughes2013classical}, we constructed a classical DFT for
-water which uses SAFT as a basis for a free energy functional. We
-used Clark \emph{et al}'s \cite{clark2006developing} five fitting
-parameters for an optimal cDFT for water and, in addition, introduced
-a parameter for the length scale over which the dispersion interaction
-is correlated.
+mixtures over a wide range of temperatures and pressures.
+\fixme{Talk about and cite papers using SAFT to build DFTs.}
+%
+In a recent paper \cite{hughes2013classical}, we constructed a
+classical DFT for water which uses SAFT as a basis for a free energy
+functional. We used Clark \emph{et al}'s \cite{clark2006developing}
+five fitting parameters for an optimal cDFT for water and, in
+addition, introduced a parameter for the length scale over which the
+dispersion interaction is correlated.
 

### 2013-08-01 papers/water-saft/paper.tex
modified plots, added citations, corrected hydrogen bond breaking calculations in water-saft Tested everything on node09.

@@
 models for water have built on the approach of classical
-density-functional theory (DFT)~\fixme{add citations here,
-  particularly Jeanmairet 2013 and his citations 35-44, possibly
-  cutting non-acqueous papers from that list}.
-Classical DFT is based on a description of a fluid written as a free
-energy functional of the density distribution.  There are two
-approaches used to construct a classical DFT for water.  The first is
-to choose a functional form that is fit to properties of the bulk
-liquid at a given temperature and pressure~\fixme{cite Jeanmairet and
-  similar}.  Using this approach, it is possible to construct a
-functional that reproduces the exact second-order response function of
-the liquid under the fitted conditions.  This class of functional,
-however, will be less accurate at other temperatures or
-pressures---and in the inhomogeneous scenarios in which solvation
-models are applied.  The second approach is construct a functional by
-applying liquid-state theory to a model system, and then fit the model
-to experimental data such as the equation of state.
+density-functional theory (DFT)\cite{jeanmairet2013molecular,
+  zhao2011molecular, zhao2011new, ramirez2005direct,
+  ramirez2005density, levesque2012solvation, levesque2012scalar}
+~\fixme{add citations here, particularly Jeanmairet 2013 and his
+  citations 35-44, possibly cutting non-acqueous papers from that
+  list}.  Classical DFT is based on a description of a fluid written
+as a free energy functional of the density distribution.  There are
+two approaches used to construct a classical DFT for water.  The first
+is to choose a functional form that is fit to properties of the bulk
+liquid at a given temperature and pressure
+\cite{jeanmairet2013molecular} ~\fixme{cite Jeanmairet and similar}.
+Using this approach, it is possible to construct a functional that
+reproduces the exact second-order response function of the liquid
+under the fitted conditions.  This class of functional, however, will
+be less accurate at other temperatures or pressures---and in the
+inhomogeneous scenarios in which solvation models are applied.  The
+second approach is construct a functional by applying liquid-state
+theory to a model system, and then fit the model to experimental data
+such as the equation of state.
 
@@
     4 n(\rr) & \text{this work}\\
-    4 n_0(\rr) \zeta(\rr) & \text{previous paper}.
+    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
   \end{cases} 
 \end{align}
-The four in $n_{site}(\rr)$ comes from the four hydrogen bond sites
-and, $n_o(\rr)$ is the average density at contact, and $\zeta(\xx)$ is
-a dimensionless measure of the density inhomogeneity from Yu and
-Wu. \cite{yu2002fmt-dft-inhomogeneous-associating} The functional
-$X(\rr)$ is the fraction of association sites \emph{not}
+is the density of bonding sites at position~$\rr$.  The four in
+$n_\text{site}(\rr)$ comes from the four hydrogen bond sites,
+$n_0(\rr)$ is the average density contacting point $\rr$, and
+$\zeta(\xx)$ is a dimensionless measure of the density inhomogeneity
+from Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  The
+functional $X(\rr)$ is the fraction of association sites \emph{not}
 hydrogen-bonded, which is determined by the quadratic equation
 \begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
+  X(\xx) &= \frac{\sqrt{1 + 2n_\text{site}'(\rr)
       \kappaassoc g^\textit{SW}_\sigma(\xx)
   \left(e^{-\beta\epsilonassoc} - 1\right)} - 1}
-  {4 n_0(\xx)\zeta(\xx)
+  {n_\text{site}'(\rr)
     \kappaassoc g^\textit{SW}_\sigma(\xx)
@@
 where
+\begin{align}
+  n_\text{site}'(\rr) &=
+  \begin{cases}
+    \frac{4}{\pi\sigma^2} \int n(\rr')\delta(\sigma - |\rr-\rr'|) d\rr' & \text{this work}\\
+    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
+  \end{cases} 
+\end{align}
+is the density of bonding sites that could bond to the sites~$n_\text{site}(\rr)$, and
 \begin{align}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/single-rod-X-plot}
-\end{center}
-\caption{ The average number of hydrogen bonds per molecule at
-  distances from the center of hydrophobic rods of various
-  radii. The solid lines are the results from using the assymetric
-  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}^{HS}$.}
-\label{fig:single-rod-X}
-\end{figure}
+\section{Hydrogen bonding near a hydrophobic rod}
 
@@
 \end{figure}
-
-\section{Hydrogen bonding near a hydrophobic rod}
 
@@
 \begin{center}
+\includegraphics[width=3.5in]{figs/sphere-broken-HB}
+\end{center}
+\caption{ \fixme{spheres!}}
+\label{fig:spheres-broken-HB}
+\end{figure}
+
+We continue our study into three dimensions with a hard sphere solute
+in water.\fixme{stuff}.
+
+\section{Conclusion}
+
+We have modified a classical DFT for water with a more accurate radial
+distribution function which changes the way that the molecules assosciate.
+
+\clearpage
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/single-rod-X-plot}
+\end{center}
+\caption{ The average number of hydrogen bonds per molecule at
+  distances from the center of hydrophobic rods of various
+  radii. The solid lines are the results from using the assymetric
+  $g_{\sigma}^A$, while the dotted lines are from Yu and Wu's $g_{\sigma}^{HS}$.}
+\label{fig:single-rod-X}
+\end{figure}
+\section{Introduction}
+
+\begin{figure}
+\begin{center}
 \includegraphics[width=3.5in]{figs/spheres-X-plot}
@@
 \end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/sphere-broken-HB}
-\end{center}
-\caption{ \fixme{spheres!}}
-\label{fig:spheres-broken-HB}
-\end{figure}
-
-We continue our study into three dimensions with a hard sphere solute
-in water.\fixme{stuff}.
-
-\section{Conclusion}
-
-We have modified a classical DFT for water with a more accurate radial
-distribution function which changes the way that the molecules assosciate.
-
-\clearpage
-
-\section{Introduction}
 

### 2013-08-01 papers/water-saft/figs/hughes-single-rod.cpp

@@ -204,7 +204,8 @@ int main(int argc, char **argv) {
   Grid zeroed_out_density(gd, density.cwise()*constraint); // this is zero inside the rod!  :)
   Grid X_values(gd, X(new_water_prop.kT, gd, density));
   Grid H_bonds(gd, HB(new_water_prop.kT, gd, zeroed_out_density));
-  const double broken_H_bonds = (HB(new_water_prop.kT, n_1atm)/n_1atm)*zeroed_out_density.integrate() - H_bonds.integrate();
+  const double broken_H_bonds_per_cell = (HB(new_water_prop.kT, n_1atm)/n_1atm)*zeroed_out_density.integrate() - H_bonds.integrate();
+  const double broken_H_bonds_per_nm = broken_H_bonds_per_cell*(1*nm/width);
   char *plotname = (char *)malloc(1024);
   sprintf(plotname, "papers/water-saft/figs/hughes-single-rod-slice-%04.2f.dat", diameter/nm);
   plot_grids_y_direction(plotname, density, X_values, H_bonds, constraint);
@@ -214,7 +215,7 @@ int main(int argc, char **argv) {
   sprintf(datname, "papers/water-saft/figs/hughes-single-rod-%04.2fnm-energy.dat", diameter/nm);
   FILE *o = fopen(datname, "w");
   delete[] datname;
-  fprintf(o, "%g\t%.15g\t%g\n", diameter/nm, energy, broken_H_bonds);
+  fprintf(o, "%g\t%.15g\t%g\n", diameter/nm, energy, broken_H_bonds_per_nm);
   fclose(o);
 
   {

### 2013-08-02 papers/water-saft/paper.tex
Merged files.  Some changes to plots and plot descriptions in water. Tested everything on node36.

@@
 \end{center}
-\caption{ Density profiles for single rods of different
-  radii for the new functional (top) and the old functional
-  (bottom). Each separately colored line represents a density profile
-  for a rod; the black dotted line is the bulk density.}
+\caption{ Density profiles for a water around a single hydrophobic
+  rod. The solid red profile is the density using our new functional
+  and the dashed blue profile the result of our old functional.  For
+  comparison, under the profiles is a cartoon of a string of hard
+  spheres touching in one dimension. The horizontal black dotted line
+  is the bulk density.}
 \label{fig:density-single-rod}
@@
 \end{center}
-\caption{ Broken hydrogen bonds broken when rods are placed in
-  homogeneous water.  \fixme{Our old and new functional are both plotted here}}
+\caption{Broken hydrogen bonds per nanometer for various hydrophobic rods
+  immeresed in water.  The solid red line uses our new functional
+  while the dashed blue line uses the old one. For large enough rods,
+  the graph increases linearly for both functionals.}
 \label{fig:single-rod-broken-HB}
@@
 \end{center}
-\caption{ \fixme{spheres!}}
+\caption{Broken hydrogen bonds per for various hard spheres
+  immeresed in water.  The solid red line uses our new functional
+  while the dashed blue line uses the old one.}
 \label{fig:spheres-broken-HB}

### 2013-08-02 papers/pair-correlation/paper.tex
start working on a Matplotlib scons command that will parse python scripts Tested everything on darcy.

@@
 \subsection{Pair distribution function}
-Figure~\ref{fig:pair-distribution} shows...
-\begin{figure}
-  %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-0.pdf}
-  \includegraphics[width=6cm]{figs/pair-correlation-2-395-5.pdf}
-  %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-10.pdf}
-  \caption{Plot of the pair distribution function near a hard wall.
-    In the interest of space and memeory, we should look at the function and decided which data to add?}\label{fig:pair-distribution}
-\end{figure}
+%% Figure~\ref{fig:pair-distribution} shows...
+%% \begin{figure}
+%%   %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-0.pdf}
+%%   \includegraphics[width=6cm]{figs/pair-correlation-2-395-5.pdf}
+%%   %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-10.pdf}
+%%   \caption{Plot of the pair distribution function near a hard wall.
+%%     In the interest of space and memeory, we should look at the function and decided which data to add?}\label{fig:pair-distribution}
+%% \end{figure}
 

### 2013-08-03 papers/water-saft/paper.tex

@@
 We first simulate a single hydrophobic rod immersed in water. The
-density profile of the water with our functional is compared with that
-of Hughes in Fig. \ref{fig:density-single-rod}. The density at contact
-with the rods are nearly identical, even more so for larger radius
-rods. Since we have changed the association term in the functional, a
-difference will occur in the average number of hydrogen bonds in a
-given system. We compare the hydrogen bonds with the new and old
-functional for rods of a few different radii
-(Fig. \ref{fig:single-rod-X}) and find that there is a notable
-difference.
+density profile of the water with our new functional is compared
+agains our previous results in Fig. \ref{fig:density-single-rod}. The
+density at contact with the rods are similar and there are slight
+difference in density, particularly near the first peak after contact;
+all rods we tested exhibited this sharp peak near one diamater form
+contact. Since we have changed the association term in the functional,
+a difference will occur in the average number of hydrogen bonds. We
+compare the hydrogen bonds per length with the new and old functionals
+for various rods (shown in Fig. \ref{fig:single-rod-broken-HB}) and
+find that there is a noteable difference. The number of broken bonds
+both increase linearly with the rod radius, but where this starts
+differs between the two methods.
 
@@
 We continue our study into three dimensions with a hard sphere solute
-in water.\fixme{stuff}.
+in water. Unlike the single rod, the number of hydrogen bonds does not appear to increase
+linearly with the sphere radii we used. \fixme{When the size of
+each rodis increase, the surface area increases proportional to radius.
+However, when we increase the radius of each sphere, the surface area
+increases as $R^2$.}
 

### 2013-08-06 papers/pair-correlation/paper.tex
major edit of pair distribution paper Tested everything on bingley.

@@
 \usepackage{color}
-\usepackage{breqn}
 
 \newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
 \newcommand{\rr}{\textbf{r}}
@@
 
-% needsworklater is used to annotate bits that need work, but that we
-% can postpone for a while.
-\newcommand{\needsworklater}[1]{\emph{[#1]}}
-% needsworknow is intended to prioritize stuff that needs fixing.
-\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}
+\begin{document}
+\title{Using Fundamental Measure Theory to treat the Pair
+  Distribution Function of the Inhomogeneous Hard-Sphere Fluid}
 
-\begin{document}
-\title{Using Fundamental Measure Theory to treat the two-point pair
-  distribution function of the Inhomogeneous Hard-Sphere Fluid}
-
-\author{Paho}
-\author{???}
+\author{Paho Lurie-Gregg}
+\author{Jeff B. Schulte}
 \author{David Roundy}
@@
 \begin{abstract}
-  We develop and test an efficient approximation for the two-point
-  pair distribution function of an inhomogeneous hard-sphere fluid.
+  We have developed an approximation for the pair distribution
+  function of hard spheres, which is based on the average value of the
+  pair distribution function at contact described in
+  Section~\ref{sec:contact}.  Our pair distribution is constructed so
+  as to be computable with only single-site convolutions and to
+  accurately reproduce the averaged distribution function at contact.
+  This results in a functional that is efficient to compute, and
+  predicts the pair distribution function as well as the most accurate
+  existing analytical approximation.
 \end{abstract}
 
-\section{Why does pair distribution matter?}
-\subsection{Interfaces are important}
-\subsection{cDFT theory enables treatment of inhomogeneous systems}
-\subsection{Pair distribution function is key input in thermodynamic perturbation theory}
-The pair distribution function is a key input into thermodynamic
-perturbation theory, which is widely used for treating dispersion
-interactions in fluids.  In most cases, the hard-sphere fluid is used
-as the reference system, making the pair distribution function of the
-hard-sphere system an important quantity.
-\subsection{Explain what the $a_1$ in thermodynamic perturbation theory is}
+\maketitle
 
-\section{What has been done before?}
-\subsection{Gloor simple method, problematic}
-Gil-Villegas\cite{gil1997statistical} approximates the first order inverse tenpurature Helmholtz Free energy term, which involves an integral of a potential multiplied by a radial distribution function $g^{HS}(r,\eta)$, by first taking the distribution function out of the integral and evaluating it at the correct mean distance $\zeta$, according to the mean value theorem.
-They then approximate the pair correlation function $g^{HS}(\zeta,\eta)$ with the correlation at contact and at a effective $\eta_{eff}$.
-The $\eta_{eff}$ is found by fitting a general function form to correct energy solutions $a_1$ with depenedent variable $\eta$ and
+\newcommand\saftlocaldft{}
+% The following are papers that use a SAFT-based classical DFT with
+% all the terms that should be non-local being non-local.
+\newcommand\saftnonlocaldft{}
 
-\subsection{Fischer accurate but slow}
+The standard approach in liquid state theory is to treat a liquid as a
+hard-sphere reference fluid with attractive interactions that are
+treated perturbatively~\cite{hansen2006theory}.  Recently there have
+been considerable advances in the development of classical density
+functionals \fixme{introduce DFT first} have been based on the
+application of these theoretical approaches to inhomogeneous
+densities~\cite{ jain2007modified, gloor2007prediction, gross2009density,
+  %felipe2001examination, gloor2002saft,
+  %gloor2004accurate, clark2006developing,
+  kahl2008modified,
+  % yu2002fmt-dft-inhomogeneous-associating,
+  %fu2005vapor-liquid-dft, bryk2006density,
+  hughes2013classical,
+  %segura1998comparison, felipe2001examination,
+  %gloor2002saft, gloor2004accurate, fu2005vapor-liquid-dft,
+  bryk2006density, clark2006developing,
+  kahl2008modified, gross2009density,
+  %yu2002fmt-dft-inhomogeneous-associating,
+  sundararaman2012computationally, marshall2013density}.  The
+perturbation theory treatment of intermolecular interactions relies on
+the pair distribution function of the reference fluid:
+$g^{(2)}(\rr_1,\rr_2)$.  Unlike the radial distribution function of a
+homogeneous fluid, there does not currently exist a tractable form for
+the pair distribution function of an inhomogeneous hard-sphere fluid,
+suitable for use in constructing a density
+functional~\cite{gloor2007prediction, jain2007modified}.
 
-\section{Triplet correlation function?}
+\fixme{explain a bit of perturbation theory here?}
+
+\section{Theoretical approaches}
+
+The classic (and earliest) approach for computing the pair
+distribution function is to use Percus' trick of treating one sphere
+as an external field, and using the resultant equilibrium density to
+find the pair distribution function~\cite{hansen2006theory}.  This
+elegant approach lends itself to DFT, and can be used to compute and
+plot the pair distribution function, but requires a full free-energy
+minimization \emph{for each position} $\rr_1$ in
+$g^{(2)}(\rr_1,\rr_2)$, and hence would be prohibitively expensive as
+a tool in constructing a free energy functional.
+
+The first canonical inhomogeneous configuration for the hard-sphere
+fluid is the system consisting of a hard spheres at a hard wall.  In
+1986, Plischke and Henderson solved the pair distribution function of
+this system using integral equation theory under the Percus-Yevick
+approximation~\cite{plischke1986pair}.  Ten years later,
+G{\"o}tzelmann \emph{et al.} found the pair distribution function near
+a hard wall using classical DFT by finding the direct correlation
+function by taking a functional derivative of the free energy
+functional and then solving the Ornstein-Zernicke
+equation~\cite{gotzelmann1996structure}.  The approach of solving the
+Ornstein-Zernicke equation for an inhomogeneous system is
+computationally challenging~\cite{paul2003variational}.
+
+A second inhomogeneous configuration that is of interest is the
+test-particle configuration, in which one hard sphere is fixed in the
+otherwise homogeneous fluid.  Where the hard-wall is a surface with no
+curvature, the test-particle configuration has a surface with
+curvature at the molecular length scale.  In this case, the density
+gives the pair distribution function---this is just Percus'
+trick---and the pair distribution function gives the triplet
+distribution function.  The triplet distribution function was computed
+by Gonz\'alez \emph{et al.} using the test-particle approach with
+\emph{two} spheres fixed~\cite{gonzalez1999test}.
+
+Lado recently introduced a new and improved efficient algorithm for
+implementing integral equation theory for inhomogeneous fluids, which
+computes $g^{(2)}(\rr_1,\rr_2)$~\cite{lado2009efficient}.  While this
+approach is two orders of magnitude more efficient than previous
+implementations, it remains a computationally expensive approach, and
+unsuitable for repeated evaluation within a free-energy minimization.
+
+\section{Analytic approximations}
+There are few analytic approximations for the inhomogeneous pair
+distribution function, which extend the radial distribution function
+to inhomogeneous scenarios.  These works differ in what density they
+use to evaluate $g(r)$.
+
+Several recent works have cited a
+mean-density approximation:
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \frac{\pi\sigma^3}{6}\tfrac12
+  (n(\rr_1)+n(\rr_2))\right)
+\end{align}
+where $g(r,\eta)$ is the radial distribution function as a function of
+radius and packing fraction $\eta$~\cite{gloor2007prediction,
+  gross2009density}.  These works did not however use this approach,
+which fails dramatically at higher densities, as the packing fraction
+found from the mean density exceeds unity.
+
+Approaches that do not diverge use an average of the density over some
+volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
+approximation
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, n_3\left(\tfrac12
+  (\rr_1+\rr_2)\right)\right)
+  \label{eq:fischer}
+\end{align}
+where $n_3$ is one of the fundamental measures defined in Fundamental
+Measure Theory~\cite{rosenfeld1989free}, which is an integral of the
+density over a spherical volume:
+\begin{align}
+  n_3(\rr) = \int n(\rr')\Theta(\tfrac12 \sigma - |\rr-\rr'|) d\rr'
+\end{align}
+where $\sigma$ is the hard-sphere diameter.  Equation~\ref{eq:fischer}
+is computationally awkward, because it relies on the midpoint
+$\tfrac12(\rr_1+\rr_2)$.  For this reason Sokolowski and Fischer
+modified this approach to use averages centered on the two points
+$\rr_1$ and $\rr_2$:
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
+  \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
+  \label{eq:sokolowski}
+\end{align}
+where
+\begin{align}
+  \bar{n}(\rr) = \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
+\end{align}
+is the density averaged over a sphere with diameter
+$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
+was arrived at by fitting to Monte Carlo simulation.  Although
+Eq.~\ref{eq:sokolowski} has the advantages of only involving density
+averages at the points at which the pair distribution function is
+desired, cannot be written as a single-site convolution, since the
+convolution kernel depends on both points.
 
@@
 
-\begin{dmath}
+\begin{align}
   g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
-\end{dmath}
+\end{align}
 
@@
 %% expansion where fit
-\section{What does it look like (graphs that show the function itself)?}
-\subsection{Fit to radial distribution function}
-Figure~\ref{fig:radial-distribution} shows...
 \begin{figure}
-  \includegraphics[width=3cm]{figs/ghs-g}
-  \includegraphics[width=3cm]{figs/ghs-g-ghs}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/ghs-g}\\
+  \includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
 \end{figure}
+\section{What does it look like (graphs that show the function itself)?}
+\subsection{Separable fit to the radial distribution function}
+Figure~\ref{fig:radial-distribution} shows...
+
 \subsection{Pair distribution function}
@@
 
-Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but maybe you (Dr. Roundy)
-should take a look at it.  They reference Tang \cite{tang2008accurate} which I've
-mentioned below in the Appendix.
+Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but
+maybe you (Dr. Roundy) should take a look at it.  They reference Tang
+\cite{tang2008accurate} which I've mentioned below in the Appendix.
+\fixme{DJR: Not same as Gloor, they do a WDA approximation.}
 
-Sokolowski and Fischer \cite{sokolowski1992role} do the same as Gloor et al except that instead of
-taking the exact densities $\rho(\mathbf{r_1})$ and $\rho(\mathbf{r_2})$ to
-average he takes an integral $\int_{V_{sphere}} \rho(\mathbf{r})d\mathbf{r}$
-over a sphere of some radius centered at each point.  It's
-these two averaged densities that he takes the average of.
+Sokolowski and Fischer \cite{sokolowski1992role} do the same as Gloor
+et al except that instead of taking the exact densities
+$\rho(\mathbf{r_1})$ and $\rho(\mathbf{r_2})$ to average he takes an
+integral $\int_{V_{sphere}} \rho(\mathbf{r})d\mathbf{r}$ over a sphere
+of some radius centered at each point.  It's these two averaged
+densities that he takes the average of.
 
-Fischer and Methfessel \cite{fischer1980born}, in a paper that is cited by Sokolowski and Fischer (above) do the same thing (the originate the idea of the spherical average density.
-Their Introduction has a very good discussion of the different averaging methods to get to $g(\mathbf{r})$ tried up until then (paper is written in 1980).
+Fischer and Methfessel \cite{fischer1980born}, in a paper that is
+cited by Sokolowski and Fischer (above) do the same thing (the
+originate the idea of the spherical average density.  Their
+Introduction has a very good discussion of the different averaging
+methods to get to $g(\mathbf{r})$ tried up until then (paper is
+written in 1980).
 
-Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but maybe you (Dr. Roundy)
-should take a look at it.  They reference Tang \cite{tang2008accurate} which I've
-mentioned below in the Appendix.
+McCabe and Kiselev \cite{mccabe2004application} model a square well
+potential.  One of their terms uses a radial distribution function
+that once again is taken to be the bulk function evaluated at an
+effective density.  They reference a source that calculates this
+effective density (or filling fraction) to be $\eta_{effective} =
+c_1\eta + c_2\eta^2 + c_3\eta^3$ where the coeffecients are gotten
+from three coupled polynomial equations with $1,\lambda$, and
+$\lambda^2$, where $\lambda$ is a measure of the model's potential
+range.  They reference another source who calculates the effective
+density in a similar (but more complicated) way.  I can look more at
+these if you'd like.  They also use a different first order expansion
+formula for the ditribution at contact in the chain term.
 
-McCabe and Kiselev \cite{mccabe2004application} model a square well potential.  One of their terms uses a radial
-distribution function that once again is taken to be the bulk function evaluated at an
-effective density.  They reference a source that calculates this effective density (or filling fraction) to be
-$\eta_{effective} = c_1\eta + c_2\eta^2 + c_3\eta^3$ where the coeffecients are gotten
-from three coupled polynomial equations with $1,\lambda$, and $\lambda^2$, where $\lambda$
-is a measure of the model's potential range.  They reference another source who calculates
-the effective density in a similar (but more complicated) way.  I can look more at these
-if you'd like.  They also use a different first order expansion formula for the ditribution
-at contact in the chain term.
-
-Toxvaerd \cite{toxvaerd1976hydrostatic}approximates the radial distribution as the radial distribution
-of the bulk at the same density.  Uses the first two terms in a tempurature
-expansion.  First term is just Carnahan Starling derived, second is empirical.
+Toxvaerd \cite{toxvaerd1976hydrostatic}approximates the radial
+distribution as the radial distribution of the bulk at the same
+density.  Uses the first two terms in a tempurature expansion.  First
+term is just Carnahan Starling derived, second is empirical.
 
@@
 
-\begin{dmath}
+\begin{align}
   g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, \eta(g_\sigma(\rr_1))) + g(|r_{12}|, \eta(g_\sigma(\rr_2)))}{2}
-\end{dmath}
+\end{align}
 Bad:
-\begin{dmath}
+\begin{align}
   g^{(2)}(\rr_1,\rr_2) = g\left(|r_{12}|, \frac{\eta(g_\sigma(\rr_1)) + \eta(g_\sigma(\rr_2))}{2}\right)
-\end{dmath}
+\end{align}
 
 The general idea is to write
-\begin{dmath}
+\begin{align}
   g^{(2)}(\rr_1,\rr_2) = \Theta(r_{12}-\sigma)
   \left(1 + \sum_{\alpha} \frac{f_{\alpha}(g_\sigma(\rr_1)) + f_{\alpha}(g_\sigma(\rr_2))}{2}g_{\alpha}(r_{12})\right)\label{eq:gpair-approx}
-\end{dmath}
+\end{align}
 where $f_{\alpha}$ and $g_{\alpha}$ are a set of functions that
 satisfy the following constraint:
-\begin{dmath}\label{eq:definealpha}
+\begin{align}\label{eq:definealpha}
   \sum_{\alpha} f_{\alpha}(g_\sigma(\eta)) g_{\alpha}(r) \approx g(r; \eta)
-\end{dmath}
+\end{align}
 which ensures that our pair distribution function reduces
@@
 \section*{Appendix}
-\cite{jin2011perturbative} Combines hard sphere (white bear) with attraction that is not mean field but rather uses two other correltation expressions given in  -\cite{tang2008accurate} and \cite{hlushak2009direct}
+\cite{jin2011perturbative} Combines hard sphere (white bear) with
+attraction that is not mean field but rather uses two other
+correltation expressions given in -\cite{tang2008accurate} and
+\cite{hlushak2009direct}
 
-Tang \cite{tang2008accurate}Derives the direct correltation for a SW and Yukawa and LJ potentials by first getting the radial distribution function through a method I don't really understand.
-Composes fourier transforms $c(k)$ and $h(k)$ of these complicated functions that depend on filling fraction and the potential in an integral.
-Uses first order mean spherical approximation.
-This is just square potential well around particle but the radial distribution function takes only the first order in an expansion in...
-(high temp?).
+Tang \cite{tang2008accurate}Derives the direct correltation for a SW
+and Yukawa and LJ potentials by first getting the radial distribution
+function through a method I don't really understand.  Composes fourier
+transforms $c(k)$ and $h(k)$ of these complicated functions that
+depend on filling fraction and the potential in an integral. Uses
+first order mean spherical approximation.  This is just square
+potential well around particle but the radial distribution function
+takes only the first order in an expansion in... (high temp?).
 
-\cite{hlushak2009direct}
-Cites \cite{tang2008accurate} (above) and does largely the same thing but then adds a complication for $2 < \lambda \leq 3$
+\cite{hlushak2009direct} Cites \cite{tang2008accurate} (above) and
+does largely the same thing but then adds a complication for $2 <
+\lambda \leq 3$
 
-\cite{levesque2012scalar}
-Uses direct correlation funtion that's gotten from a radial distribution function of a homogeneous density.
-The direct correlation itself is only dependent on distance.
-The $\rho_0$ that they use is the same $\rho_0$ that the used in the reference HS term in their functional.
-The use the Carnahan Starling or the Perkis Yevick for these hard sphere free energies.
-The other terms involve $\delta\rho$ so the change is a variation from the reference density.
-They further modify it by looking at molecular dynamic simulations of a certain model of water molecules and looking at the oxygen-oxygen correlation.
+\cite{levesque2012scalar} Uses direct correlation funtion that's
+gotten from a radial distribution function of a homogeneous density.
+The direct correlation itself is only dependent on distance.  The
+$\rho_0$ that they use is the same $\rho_0$ that the used in the
+reference HS term in their functional.  The use the Carnahan Starling
+or the Perkis Yevick for these hard sphere free energies.  The other
+terms involve $\delta\rho$ so the change is a variation from the
+reference density.  They further modify it by looking at molecular
+dynamic simulations of a certain model of water molecules and looking
+at the oxygen-oxygen correlation.
 
@@
 \end{document}
+
+%% G{\"o}tzelmann \emph{et al.} worked out the pair distribution function near a hard
+%% wall by finding the direct correlation function by taking a functional
+%% derivative of the free energy functional and then solving the
+%% Ornstein-Zernicke equation~\cite{gotzelmann1996structure}.  They have plots like our pretty plots,
+%% but with the fixed sphere one radius away from touching.
+%% Interestingly, they plot the second bit as $z$, which enables them to
+%% have just two axis values.  I think I prefer our ``total distance''
+%% method, but there is a certain elegance to just plotting $z$ for the
+%% curvy part as well as the straight-away part.  They found poor
+%% quantitative agreement, particularly at contact.  They also studied
+%% the pair correlation function in a hard slit~\cite{gotzelmann1997density}.
+
+%% Integrating the Ornstein-Zernicke equation is challenging, requiring a
+%% variational minimization itself~\cite{paul2003variational}.
+
+%% Attard worked out the triplet correlation function, which is something
+%% that we can access from the pair distribution function, if we consider
+%% the pair correlation near a single hard sphere
+%% solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al.} wrote an
+%% interesting paper using DFT (test-particle approach of Percus) and
+%% Monte Carlo to compute the three-particle distribution function (same
+%% as the triplet correlation function)\cite{gonzalez1999test}.  Their
+%% theory compares favorably with the ``superposition approximation'',
+%% and we could probably do something similar.
+
+%% Plischke and Henderson worked out the pair correlation function near a
+%% hard wall using Percus-Yevick, and compare with Monte Carlo
+%% results\cite{plischke1986pair}.  They plot the pair correlation
+%% function along interesting paths.  \fixme{Jeff read this paper for a
+%%   while and does not understand what they do in oder to find the pair
+%%   correlation function as compared to the radial distribution function}
+
+%% Lado recently introduced a new and improved efficient algorithm for
+%% implementing integral equation theory for inhomogeneous fluids, which
+%% computes $g^{(2)}(\rr_1,\rr_2)$ among other
+%% things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
+%% use, and their example is limited to 1D systems.  And the scaling is
+%% still (so far as I can tell) $O(N^2)$, since their improvement is in
+%% the prefactor, by developing a nice basis set.
+
+%% Gloor \emph{et al} explain that little is known of the pair
+%% distribution function of the hard-sphere fluid, and therefore use the
+%% formula
+%% \begin{align}
+%%   g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
+%% \end{align}
+%% They go on to further approximate the radial distribution function,
+%% but we can just stop here.
+
+%% Gloor et al \cite{gloor2007prediction} uses average density of two points:
+%% \begin{align}
+%% \overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
+%% \end{align}
+%% and the uses bulk radial distribution for the reference system (hard
+%% spheres) at this density, in order to approximate pair correlation.
+
+%% \section{What has been done before?}
+%% \subsection{Gloor simple method, problematic}
+%% Gil-Villegas\cite{gil1997statistical} approximates the first order
+%% inverse tenpurature Helmholtz Free energy term, which involves an
+%% integral of a potential multiplied by a radial distribution function
+%% $g^{HS}(r,\eta)$, by first taking the distribution function out of the
+%% integral and evaluating it at the correct mean distance $\zeta$,
+%% according to the mean value theorem.  They then approximate the pair
+%% correlation function $g^{HS}(\zeta,\eta)$ with the correlation at
+%% contact and at a effective $\eta_{eff}$.  The $\eta_{eff}$ is found by
+%% fitting a general function form to correct energy solutions $a_1$ with
+%% depenedent variable $\eta$ and
+
+%% \subsection{Fischer accurate but slow}

### 2013-08-07 papers/water-saft/paper.tex
Some small changes to water paper Tested everything on node25.

@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
@@
   ramirez2005density, levesque2012solvation, levesque2012scalar}
-~\fixme{add citations here, particularly Jeanmairet 2013 and his
-  citations 35-44, possibly cutting non-acqueous papers from that
-  list}.  Classical DFT is based on a description of a fluid written
+~\fixme{These are from Jeanmairet 2013 citations 35-44; only the ones
+  that discuss water}.  Classical DFT is based on a description of a fluid written
 as a free energy functional of the density distribution.  There are
@@
 
-%% \fixme{stuff about water}.  An important aspects of water
-%% is its ability to form hydrogen bonds.  Each water molecule contains
-%% two hydrogen bond donor sites and two acceptor sites.
-%% %
-%% One promising approach for an efficient continuum description of water
-%% is that of classical density-functional theory (DFT), which is an
-%% approach for evaluating the free energy and thermally averaged density
-%% of fluids in an arbitrary external potential~\cite{ebner1976density}.
-%% The foundation of classical DFT is the Mermin
-%% theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
-%% theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
-%% Classical DFT is a natural framework for creating a more flexible
-%% theory of hydrophobicity that can readily describe interaction of
-%% water with arbitrary external potentials---such as potentials
-%% describing strong interactions with solutes or surfaces.
+\fixme{stuff about water}.  An important aspects of water
+is its ability to form hydrogen bonds.  Each water molecule contains
+two hydrogen bond donor sites and two acceptor sites.
+%
+One promising approach for an efficient continuum description of water
+is that of classical density-functional theory (DFT), which is an
+approach for evaluating the free energy and thermally averaged density
+of fluids in an arbitrary external potential~\cite{ebner1976density}.
+The foundation of classical DFT is the Mermin
+theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
+theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
+Classical DFT is a natural framework for creating a more flexible
+theory of hydrophobicity that can readily describe interaction of
+water with arbitrary external potentials---such as potentials
+describing strong interactions with solutes or surfaces.
 
@@
 (SAFT).
-%
+
 SAFT is a theory for fluids in which hydrogen bonding plays a significant
@@
 \fixme{Talk about and cite papers using SAFT to build DFTs.}
-%
+
 In a recent paper \cite{hughes2013classical}, we constructed a
@@
 agree less with hard-sphere Monte-Carlo simulation than a distribution
-function developed by us in a recent paper.\cite{schulte2012using} In
+function developed by us in a recent paper \cite{schulte2012using}. In
 our paper, we developed an approximation for correlation at
@@
 
-We modify our previous the association free energy term to use our
-$g_{\sigma}^A$ rather than Yu and Wu's $g_\sigma^{YuWu}$.  The
+We modify our previous association free energy term to use our
+$g_{\sigma}^\textit{A}$ rather than Yu and Wu's $g_\sigma$.  The
 association free energy for our four-site model has the form:
@@
 \end{align}
-% This is equation 77 of gil-villegas-1997-SAFT-VR
+This is equation 77 of gil-villegas-1997-SAFT-VR
 where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
@@
 \caption{ Density profiles for a water around a single hydrophobic
-  rod. The solid red profile is the density using our new functional
+  rod of radius 0.2 nm. The solid red profile is the density using our new functional
   and the dashed blue profile the result of our old functional.  For
@@
   spheres touching in one dimension. The horizontal black dotted line
-  is the bulk density.}
+  is the bulk density for water.}
 \label{fig:density-single-rod}
@@
 
-We continue our study into three dimensions with a hard sphere solute
-in water. Unlike the single rod, the number of hydrogen bonds does not appear to increase
-linearly with the sphere radii we used. \fixme{When the size of
-each rodis increase, the surface area increases proportional to radius.
-However, when we increase the radius of each sphere, the surface area
-increases as $R^2$.}
+We continue our study in three dimensions with a hard sphere solute in
+water. Similar to the single rod, we see a decrease in the number of
+broken hydrogen bonds for a given rod radius. However, with the single
+rods the number of hydrogen bonds appears to increase linearly with
+rod radius where the sphere radii we used. \fixme{When the size of
+  each rod is increased, the surface area increases proportional to
+  radius.  However, when we increase the radius of each sphere, the
+  surface area increases as $R^2$.}
 
@@
 We have modified a classical DFT for water with a more accurate radial
-distribution function which changes the way that the molecules assosciate.
+distribution function which modifies assosciation between water
+molecules. This modification predicts less broken hydrogen bonds due
+to a hard rod, and solutes placed in water than the previous theory.
 
@@
 water molecules per unit cell~\cite{lewis2011doesnitric}.
-%
+
 On top of this, traditional density-functional theory (DFT) methods
@@
 water, to the point that ice melts at over 120$^\circ$C\cite{yoo2009phase}!
-%
+
 There has been a flurry of recent
@@
   wang2011density, mogelhoj2011ab, jonchiere2011van}.
-%
+
 However, one particular study found that water modeled using a
@@
 melting point over 80$^\circ$C\cite{yoo2011}.
-%
+
 It has also
@@
 do reproduce the temperature-dependence of water's properties.
-% 
+
 We should give special mention to Sundararaman \emph{et al} who

### 2013-08-08 papers/water-saft/paper.tex
Small changes to water paper Tested everything on node25.

@@
 models for water have built on the approach of classical
-density-functional theory (DFT)\cite{jeanmairet2013molecular,
+density-functional theory (DFT)~\cite{jeanmairet2013molecular,
   zhao2011molecular, zhao2011new, ramirez2005direct,
-  ramirez2005density, levesque2012solvation, levesque2012scalar}
-~\fixme{These are from Jeanmairet 2013 citations 35-44; only the ones
-  that discuss water}.  Classical DFT is based on a description of a fluid written
-as a free energy functional of the density distribution.  There are
-two approaches used to construct a classical DFT for water.  The first
-is to choose a functional form that is fit to properties of the bulk
+  ramirez2005density, levesque2012solvation, levesque2012scalar}.
+Classical DFT is based on a description of a fluid written as a free
+energy functional of the density distribution.  There are two
+approaches used to construct a classical DFT for water.  The first is
+to choose a functional form that is fit to properties of the bulk
 liquid at a given temperature and pressure
@@
 \end{align}
-This is equation 77 of gil-villegas-1997-SAFT-VR
+%This is equation 77 of gil-villegas-1997-SAFT-VR
 where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
@@
 difference in density, particularly near the first peak after contact;
-all rods we tested exhibited this sharp peak near one diamater form
+all rods we tested exhibited this sharp peak near one diamater from
 contact. Since we have changed the association term in the functional,
@@
 
-We continue our study in three dimensions with a hard sphere solute in
-water. Similar to the single rod, we see a decrease in the number of
-broken hydrogen bonds for a given rod radius. However, with the single
-rods the number of hydrogen bonds appears to increase linearly with
-rod radius where the sphere radii we used. \fixme{When the size of
-  each rod is increased, the surface area increases proportional to
-  radius.  However, when we increase the radius of each sphere, the
+A common test case for studying hydrophobic solutes in water is the
+hard-sphere solute. We continue our study in three dimensions by
+creating a hydrophobic spherical cavity in water. Similar to the
+single rods, we see a decrease in the number of broken hydrogen bonds
+for spheres of the same radius compared to our previous
+functional. The number of broken hydrogen bonds increases linearly
+with rod radius for large enough rods, where for spheres we don't see
+this linearity (Fig. \ref{fig:spheres-broken-HB}). \fixme{When the
+  size of each rod is increased, the surface area increases
+  proportional to radius.  However, when sphere radius increase, the
   surface area increases as $R^2$.}

### 2013-08-08 papers/pair-correlation/paper.tex
Modified the alternate path plot and the paper to use it and to annotate the x-axis nicely in pair-correlation. This is a temporary change, as the annotation is done in the paper and will need to be separate. Also fixed a small bug in triplet-monte-carlo.cpp. Tested everything on paho-deb.

@@
 \usepackage{color}
+\usepackage{tikz}
+\usetikzlibrary{arrows,decorations.pathreplacing}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-1.pdf}};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (4.75, .75) node(a){} -- (5.95, .75) node(b){};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (5.99, .75) node(a){} -- (7.83, .75) node(b){};
+  \end{tikzpicture}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-2.pdf}};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (4.75, .75) node(a){} -- (5.95, .75) node(b){};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (5.99, .75) node(a){} -- (7.83, .75) node(b){};
+  \end{tikzpicture}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-3.pdf}};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (4.75, .75) node(a){} -- (5.95, .75) node(b){};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (5.99, .75) node(a){} -- (7.83, .75) node(b){};
+  \end{tikzpicture}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
+  \begin{tikzpicture}
+    \node[anchor=south west,inner sep=0] at (0,0) {\includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-4.pdf}};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (4.75, .75) node(a){} -- (5.95, .75) node(b){};
+    \draw [decorate, decoration = {brace, mirror, raise=0.5cm}]
+    (5.99, .75) node(a){} -- (7.83, .75) node(b){};
+  \end{tikzpicture}
   \caption{Pretty plot of the pair distribution function near a hard

### 2013-08-08 src/Monte-Carlo/triplet-monte-carlo.cpp

@@ -555,7 +555,7 @@ int main(int argc, char *argv[]){
                   const double z_cent = z1/2.0;
                   const double theta = acos(r12_v.dot(zhat)/r12);
                   if (z2 < lenx/2.0 && x2 < path_dx && z2 > 3*R) {
-                    const int index = path_zbins - (z2-z1-2*R)/path_dr;
+                    const int index = path_zbins - ceil((z2-z1-2*R)/path_dr);
                     if (index < 0 || index >= path_zbins)
                       fprintf(stderr,"Index out of bounds: %i, z1: %.2f, z2: %.2f, x2: %.2f\n",index,z1,z2,x2);
                     path_histogram[index] ++;

### 2013-08-09 papers/pair-correlation/paper.tex
make curly braces in matplotlib Tested everything on bingley.

--- 2013-08-06 papers/pair-correlation/paper.tex
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-1.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-2.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-3.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \end{figure}
-
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-4.pdf}
   \caption{Pretty plot of the pair distribution function near a hard

### 2013-08-12 papers/water-saft/paper.tex
Merged and added concatination of energy files in correct order for water. Added data and small changes to paper for commit Tested everything on node09.

@@
 theory to a model system, and then fit the model to experimental data
-such as the equation of state.
-
-\fixme{stuff about water}.  An important aspects of water
+such as the equation of state.  Classical DFT is a natural framework
+for creating a more flexible theory of hydrophobicity that can readily
+describe interaction of water with arbitrary external
+potentials---such as potentials describing strong interactions with
+solutes or surfaces.
+
+An important aspects of water
 is its ability to form hydrogen bonds.  Each water molecule contains
 two hydrogen bond donor sites and two acceptor sites.
-%
-One promising approach for an efficient continuum description of water
-is that of classical density-functional theory (DFT), which is an
-approach for evaluating the free energy and thermally averaged density
-of fluids in an arbitrary external potential~\cite{ebner1976density}.
-The foundation of classical DFT is the Mermin
-theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature.
-Classical DFT is a natural framework for creating a more flexible
-theory of hydrophobicity that can readily describe interaction of
-water with arbitrary external potentials---such as potentials
-describing strong interactions with solutes or surfaces.
 
@@
 density functionals is based on Statistical Associating Fluid Theory
-(SAFT).
-
-SAFT is a theory for fluids in which hydrogen bonding plays a significant
-role\cite{chapman1989saft, muller2001molecular}. SAFT is used to
-accurately model the equations of state of both pure fluids and
-mixtures over a wide range of temperatures and pressures.
+(SAFT). SAFT is a theory for fluids in which hydrogen bonding plays a
+significant role\cite{chapman1989saft, muller2001molecular}. SAFT is
+used to accurately model the equations of state of both pure fluids
+and mixtures over a wide range of temperatures and pressures.
 \fixme{Talk about and cite papers using SAFT to build DFTs.}
@@
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/energy-vs-diameter}
-\end{center}
-\caption{ Excess chemical potential per area versus radius for a single
-hydrophobic rod immersed in water. This should have an asymptote equal
-to the surface tension at room temperature, and it agrees well with
-the surface tension in Figure~\ref{fig:surface-tension}. The inset for
-rods with a very small radius shows the linear relationship expected based on
-Equation~\ref{contactvaluethm}.}
-\label{fig:energy-vs-diameter}
-\end{figure}
-
 We begin by studying a single hydrophobic rod immersed in water.
@@
 % and the density at contact to a few percent precision.
-In Figure~\ref{fig:energy-vs-diameter} we show the excess chemical
+In Figure\fixme{deleted} we show the excess chemical
 potential at room temperature, scaled by the solvent accessible
@@
 value will decrease and approach the macroscopic value.  As seen in
-the inset of Fig.~\ref{fig:energy-vs-diameter}, for rods with very
+the inset of Fig.~\fixme{deleted}, for rods with very
 small radius (less than about 0.5~\AA) the excess chemical potential
@@
 excess chemical potentials plotted in
-Figure~\ref{fig:energy-vs-diameter}.  The agreement between these
+Figure~\fixme{deleted}.  The agreement between these
 curves confirms that our functional satisfies the contact-value

### 2013-08-13 papers/pair-correlation/paper.tex
rework intro to pair distribution paper a bit Tested everything on bingley.

@@
 been considerable advances in the development of classical density
-functionals \fixme{introduce DFT first} have been based on the
+functionals have been based on the
 application of these theoretical approaches to inhomogeneous
@@
 
-\fixme{explain a bit of perturbation theory here?}
+At its core, thermodynamic perturbation theory---sometimes referred to
+as the cluster, or high-temperature expansion---is an expansion of the
+free energy in powers of a small parameter, which typically is a
+pairwise attractive interaction:
+\begin{align}
+  F &= F_0 + F_1 + \beta F_2 + \mathcal{O}(\beta^2)
+\end{align}
+where the terms $F_n$ are corrections to the free energy at order $n$
+in the small interaction.  The first and largest term in this
+expansion is
+\begin{align}
+  F_1 &= \tfrac12 \int g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|) d\rr_1d\rr_2
+\end{align}
+where $g^{(2)}_{HS}(\rr_1,\rr_2)$ is the pair distribution function of
+the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
+In this paper, we will introduce an approximation for the hard-sphere
+pair distribution function which is suitable for use in the creation
+of classical density functionals based on a cluster expansion.
 

### 2013-08-13 papers/water-saft/paper.tex

@@
 
-An important aspects of water
-is its ability to form hydrogen bonds.  Each water molecule contains
-two hydrogen bond donor sites and two acceptor sites.
-
 A widely used family of models used in the development of classical
@@
 
+SAFT is based on Wertheim's first-order thermodynamic perturbation
+theory (TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV}, which models
+\fixme{fluids} as hard-spheres with strong associative interactions at
+sites on the surfaces of the spheres. This is ideal for water, since
+its four hydrogen bond sites can be represented by these association
+sites. The association sites have an effective range and a small
+energy contribution when a site is within range of another
+hard-sphere's site. The free energy due to this association depends on
+the radial distribution function $g_\sigma^\text{HS}(\rr)$. Our
+previous water functional used Yu and Wu's radial distribution
+function \cite{yu2002fmt-dft-inhomogeneous-associating}, which was
+found to agree less with hard-sphere Monte-Carlo simulation than a
+distribution function developed by us in a recent paper
+\cite{schulte2012using}. In our paper, we developed an approximation
+for correlation at contact to describe the hard-sphere fluid.
+
 In a recent paper \cite{hughes2013classical}, we constructed a
@@
 dispersion interaction is correlated.
-
-SAFT is based on Wertheim's first-order thermodynamic perturbation
-theory (TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
-  wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
-account for strong associative interactions between
-molecules. Wertheim's thermodynamic perturbation theory uses a
-hard-sphere reference fluid where each sphere has a number of
-association sites on the surface. The association sites have an
-effective range and a small energy contribution when a site is within
-range of another hard-sphere's site. The free energy due to this
-association depends on the radial distribution function
-$g_\sigma^\text{HS}(\rr)$. Our previous water functional used Yu and
-Wu's radial distribution function
-\cite{yu2002fmt-dft-inhomogeneous-associating}, which was found to
-agree less with hard-sphere Monte-Carlo simulation than a distribution
-function developed by us in a recent paper \cite{schulte2012using}. In
-our paper, we developed an approximation for correlation at
-contact to describe the hard-sphere fluid.
 

### 2013-08-13 papers/water-saft/paper.tex
various edits of the water-saft paper Tested everything on bingley.

@@
 
-We use Clark's five empirical parameters, and fit the calculated
-surface tension to experimental surface tension at ambient conditions.
-We find these agree when $\lscale$ is 0.454.
-
 \begin{figure}
@@
 \end{center}
-\caption{Comparison of Surface tension versus temperature for theoretical and
-  experimental data. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
-  tension will match the experimental surface tension near room temperature.}
+\caption{Comparison of Surface tension versus temperature for
+  theoretical and experimental data. The experimental data is taken
+  from NIST~\cite{nistwater}.  The length-scaling parameter $\lscale$
+  is fit so that the theoretical surface tension will match the
+  experimental surface tension near room temperature.}
 \label{fig:surface-tension}
 \end{figure}
+
+As in Ref.~\citenum{hughes2013classical}, we use Clark's five
+empirical parameters, and fit the calculated surface tension to
+experimental surface tension at ambient conditions by tuning the
+parameter $\lscale$, which adjusts the length-scale of the average
+density used for the dispersion interaction.  With the improved
+association term, we find these agree when $\lscale$ is 0.454, which
+is an increase from the value of 0.353 found in
+Ref.~\citenum{hughes2013classical}.  In order to explore further the
+change made by the improved association term, we compared the new
+functional with that of Ref.~\citenum{hughes2013classical} for the two
+hydrophobic cases of hard rods and the hard spherical solutes.
 
@@
   and the dashed blue profile the result of our old functional.  For
-  comparison, under the profiles is a cartoon of a string of hard
+  scale, under the profiles is a cartoon of a string of hard
   spheres touching in one dimension. The horizontal black dotted line
@@
 
-\section{Hydrogen bonding near a hydrophobic rod}
+\section{Results}
+
+We will first discuss the case of a single hydrophobic rod immersed in
+water. Figure~\ref{fig:density-single-rod} shows the density profile
+of water near a rod with radius~0.1\AA\fixme{Eric, please figure this
+  radius out}.  The density computed using the functional of this
+paper is qualitatively similar to that from
+Ref.~\citenum{hughes2013classical}, with a comparable density at
+contact---consistent with having made only a moderate change in the
+free energy.  The first density peak near the surface is consistently
+higher than that from Ref.~\citenum{hughes2013classical}, and the peak
+is not smooth.  This results from the improved accurately of the
+$g_\sigma$ from Ref.~\citenum{schulte2012using}, since beyond the
+first peak water molecules are unable to touch---or hydrogen bond
+to---molecules at the surface of the hard rod.
 
@@
 
-We first simulate a single hydrophobic rod immersed in water. The
-density profile of the water with our new functional is compared
-agains our previous results in Fig. \ref{fig:density-single-rod}. The
-density at contact with the rods are similar and there are slight
-difference in density, particularly near the first peak after contact;
-all rods we tested exhibited this sharp peak near one diamater from
-contact. Since we have changed the association term in the functional,
-a difference will occur in the average number of hydrogen bonds. We
-compare the hydrogen bonds per length with the new and old functionals
-for various rods (shown in Fig. \ref{fig:single-rod-broken-HB}) and
-find that there is a noteable difference. The number of broken bonds
-both increase linearly with the rod radius, but where this starts
-differs between the two methods.
-
-\section{Hydrogen bonding near hard-sphere solutes}
+A common test case for studying hydrophobic solutes in water is the
+hard-sphere solute.  Figure~\ref{fig:spheres-broken-HB} shows results
+for the number of broken hydrogen bonds caused by a hard-sphere soute,
+as a function of the solute radius.  As in
+Fig.~\ref{fig:single-rod-broken-HB}, the number of broken bonds scales
+with surface area for large solutes, and the number of broken bonds is
+about four times smaller than the number from the functional of
+Ref.~\citenum{hughes2013classical}.  For solutes smaller than
+3~\AA\ in radius, there is less than a tenth of a hydrogen bond
+broken. This is consistent with the well-known fact that small solutes
+(unlike large solutes) do not disrupt the hydrogen-bonding network of
+water~\cite{chandler2005}.
 
@@
 
-A common test case for studying hydrophobic solutes in water is the
-hard-sphere solute. We continue our study in three dimensions by
-creating a hydrophobic spherical cavity in water. Similar to the
-single rods, we see a decrease in the number of broken hydrogen bonds
-for spheres of the same radius compared to our previous
-functional. The number of broken hydrogen bonds increases linearly
-with rod radius for large enough rods, where for spheres we don't see
-this linearity (Fig. \ref{fig:spheres-broken-HB}). \fixme{When the
-  size of each rod is increased, the surface area increases
-  proportional to radius.  However, when sphere radius increase, the
-  surface area increases as $R^2$.}
-
 \section{Conclusion}
 
-We have modified a classical DFT for water with a more accurate radial
-distribution function which modifies assosciation between water
-molecules. This modification predicts less broken hydrogen bonds due
-to a hard rod, and solutes placed in water than the previous theory.
+We have modified the classical DFT for water developed by
+Hughes~\emph{et al.}~\cite{hughes2013classical} with the more accurate
+radial distribution function at contact developed by Schulte~\emph{et
+  al.}~\cite{schulte2012using}, which affects the predicted hydrogen
+bonding between water molecules.  We found that while this
+modification has a relatively mild effect on the free energy and
+density profiles, it predicts fewer broken hydrogen bonds around
+hydrophobic solutes and at aqueous interfaces.
+
+\bibliography{paper}% Produces the bibliography via BibTeX.
 
 \clearpage
+
+\section{Figures that are not included in the paper (but could be)}
 
@@
 \end{figure}
-\section{Introduction}
 
@@
 
-A large fraction of interesting chemistry---including all of molecular
-biology---takes place in aqueous solution.  However, while quantum
-chemistry enables us to calculate the ground state energies of large
-molecules in vacuum, prediction of the free energy of even the
-smallest molecules in the presence of a solvent poses a continuing
-challenge due to the complex structure of a liquid and the
-computational cost of \emph{ab initio} molecular
-dynamics~\cite{car1985, grossman2004}.  The current state-of-the art
-in \emph{ab initio} molecular dynamics is limited to a few hundred
-water molecules per unit cell~\cite{lewis2011doesnitric}.
-
-On top of this, traditional density-functional theory (DFT) methods
-without the use of dispersion corrections strongly over-structure
-water, to the point that ice melts at over 120$^\circ$C\cite{yoo2009phase}!
-
-There has been a flurry of recent
-publications implicating van der Waals effects (i.e. dispersion
-corrections) as significant in
-reducing this over-structuring\cite{lin2009importance,
-  wang2011density, mogelhoj2011ab, jonchiere2011van}.
-
-However, one particular study found that water modeled using a
-hybrid functional with dispersion corrections \emph{still} has a
-melting point over 80$^\circ$C\cite{yoo2011}.
-
-It has also
-been found that the inclusion of nuclear quantum effects can provide
-similar improvements \cite{morrone2008nuclear}.  Each of these
-corrections imposes an additional computational burden on an approach
-that is already feasible for only a very small number of water
-molecules. A more efficient approach is needed in order to study
-nanoscale and larger solutes.
-
-\subsection{Classical density-functional theory}
-
-Numerous approaches have been developed to approximate the effect of water
-as a solvent in atomistic calculations.  Each of these approaches gives an
-adequate description of some aspect of interactions with water, but none of
-them is adequate for describing \emph{all} these interactions with an
-accuracy close to that attained by \emph{ab initio} calculations.  The
-theory of Lum, Chandler and Weeks (LCW)~\cite{lum1999hydrophobicity}, for 
-instance, can
-accurately describe the free energy cost of creating a cavity by placing a
-solute in water, but does not lend itself to extensions treating the strong
-interaction of water with hydrophilic solutes.  Treatment of water as a
-continuum dielectric with a cavity surrounding each solute can give
-accurate predictions for the energy of solvation of ions~\cite{latimer1939,
-rashin1985, zhan1998, hsu1999, hildebrandt2004, hildebrandt2007}, but
-provides no information about the size of this cavity.  In a physically
-consistent approach, the size of the cavity will naturally arise from a
-balance between the free energy required to create the cavity, the
-attraction between the water and the solute, and the steric repulsion which
-opens up the cavity in the first place.
-
-One promising approach for an efficient continuum description of water
-is that of classical density-functional theory (DFT), which is an
-approach for evaluating the free energy and thermally averaged density
-of fluids in an arbitrary external potential~\cite{ebner1976density}.
-The foundation of classical DFT is the Mermin
-theorem\cite{mermin1965thermal}, which extends the Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} to non-zero temperature,
-stating that
-\begin{equation}
-  \Omega (T) = \underset{n(\rr)}{min}\left\{ F[n(\rr),T] + \int (V_\textit{ext}(\rr) +\mu)n(\rr)
-d\rr\right\},
-\end{equation}
-where $\Omega (T)$ is the grand potential of a system in the external
-potential $V_\textit{ext}$ at temperature $T$, $n(\rr)$ is the density
-of atoms or molecules, $\mu$ is the chemical potential and
-$F[n(\rr),T]$ is a universal free-energy functional for the fluid,
-which is independent of the external potential $V_\textit{ext}$.
-Classical DFT is a natural framework for creating a more flexible
-theory of hydrophobicity that can readily describe interaction of
-water with arbitrary external potentials---such as potentials
-describing strong interactions with solutes or surfaces.
-
-A number of exact properties are easily achieved in the
-density-functional framework, such as the contact-value theorem, which
-ensures a correct excess chemical potential for small hard solutes.
-Much of the research on classical density-functional theory has
-focused on the hard-sphere fluid~\cite{curtin1985, rosenfeld1989,
-  rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000}, which has
-led to a number of sophisticated functionals, such as the
-fundamental-measure theory (FMT) functionals\cite{rosenfeld1989,
-  rosenfeld1993, rosenfeld1997, tarazona1997, tarazona2000,
-  roth2002whitebear, yu2002structures}.  These functionals are
-entirely expressed as an integral of local functions of a few
-convolutions of the density (fundamental measures) that can be
-efficiently computed.  We will use the White Bear version of the FMT
-functional\cite{roth2002whitebear, yu2002structures}.  This functional
-reduces to the Carnahan-Starling equation of state in the homogeneous
-limit, and it reproduces the exact free energy in the
-strongly-confined limit of a small cavity.
-
-A number of classical density functionals have been developed for
-water~\cite{ding1987, Yang1992, yang1994density, gloor2002saft,
-  gloor2004accurate, gloor2007prediction, Jaqaman2004,
-  clark2006developing, chuev2006, lischner2010classical,
-  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination,
-  sundararaman2012computationally}, each of
-which captures some of the qualitative behavior of water.  However,
-each of these functionals also fail to capture some of water's unique
-properties.  For instance, the functional of Lischner \emph{et
-  al}\cite{lischner2010classical} treats the surface tension
-correctly, but can only be used at room temperature, and thus captures
-none of the temperature-dependence of water.  A functional by Chuev
-and Skolov\cite{chuev2006} uses an ad hoc modification of FMT that can
-predict hydrophobic hydration near temperatures of 298~K, but does not
-produce a correct equation of state due to their method producing a
-high value for pressure.
-A number of classical density functionals
-have recently been produced that are based on Statistical Associating
-Fluid Theory (SAFT)\cite{ segura1997associating, segura1998comparison,
-  yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft,gloor2002saft,muller2001molecular,
- clark2006developing, gloor2007prediction, gloor2004accurate,
-  gross2009density, kahl2008modified, blas2001examination}.  These
-functionals are based on a perturbative thermodynamic expansion, and
-do reproduce the temperature-dependence of water's properties.
-
-We should give special mention to Sundararaman \emph{et al} who
-recently introduced a classical density functional for water using a
-model in which a water molecule is treated as a hard sphere attached to
-two tetrahedrally oriented hard spheres representing voids, or
-orientations in which a hydrogen bond may not be formed, with all
-attractive interactions being lumped into a single pair potential
-treated in a mean field
-approximation\cite{sundararaman2012computationally}.
-
-\subsection{Statistical associating fluid theory}
-
-Statistical Associating Fluid Theory (SAFT) is a theory describing
-complex fluids in which hydrogen bonding plays a significant
-role\cite{chapman1989saft, muller2001molecular}.  SAFT is used to
-accurately model the equations of state of both pure fluids and
-mixtures over a wide range of temperatures and pressures.  SAFT is
-based on Wertheim's first-order thermodynamic perturbation theory
-(TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
-  wertheim1986fluidsIII, wertheim1986fluidsIV}, which allows it to
-account for strong associative interactions between molecules.
-
-The SAFT Helmholtz free energy is composed of five terms:
-\begin{equation} \label{eq:SAFT-free-energy}
-  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
-  F_\textit{assoc} + F_\textit{chain},
-\end{equation}
-where the first three terms---ideal gas, hard-sphere repulsion and
-dispersion---encompass the \emph{monomer} contribution
-to the free energy, the fourth is the \emph{association} free energy,
-describing hydrogen bonds, and the final term is the chain formation
-energy for fluids that are chain polymers.  While a
-number of formulations of SAFT have been published, we will focus on
-SAFT-VR\cite{gil-villegas-1997-SAFT-VR}, which was used by Clark
-\emph{et al} to construct an optimal SAFT model for
-water\cite{clark2006developing}. All but one of the six empirical
-parameters used in the functional introduced in this paper are taken
-directly from this Clark \emph{et al} paper.  As an example of the
-power of this model, it predicts an enthalpy of vaporization at
-100$^\circ$C of $\Delta H_{vap}$= 39.41~kJ/mol, compared with the
-experimental value $\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater},
-with an error of only a few percent.  We show a phase diagram for
-this optimal SAFT model for water in
-Figure~\ref{fig:pressure-with-isotherms}, which demonstrates that its
-vapor pressure as a function of temperature is very accurate, while
-the liquid density shows larger discrepancies.  The critical point is
-very poorly described, which is a common failing of models that are
-based on a mean-field expansion.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/pressure-with-isotherms-truncated}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The solid colored lines
-indicate the computationally calculated pressure and the dotted
-colored lines are NIST data points. The solid and dotted black lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
-\end{figure}
-
-SAFT has been used to construct classical density functionals,
-which are often used to study the surface tension as a function of
-temperature\cite{clark2006developing, gloor2004accurate, kahl2008modified,
-  gloor2007prediction, blas2001examination, kiselev2006new,
-  gloor2002saft,fu2005vapor-liquid-dft, gross2009density,
-  yang1994density, Jaqaman2004}.  Such
-functionals have qualitatively predicted the dependence of surface
-tension on temperature, but they also overestimate
-the surface tension by about 50\%, and most SAFT-based functionals are
-unsuited for studying systems that have density variations on a
-molecular length scale due to the use of a local density
-approximation\cite{gloor2002saft,clark2006developing, gloor2007prediction,
-gloor2004accurate, kahl2008modified,
-blas2001examination, kiselev2006new}.
-
-%% \begin{table}
-%% \begin{tabular}{|c|c|c|c|c|}
-%% \hline
-%% paper & SAFT functional & water & LDA & tension \\
-%% \hline
-%% chuev2006 &  & X &  & X \\
-%% \hline
-%% yuwu2002dft & X & & & \\
-%% \hline
-%% fuwu2005 & X & X & & X \\
-%% \hline
-%% kiselev2006 &  & X & X & X \\
-%% \hline
-%% clark2006developing & X & X & X & X \\
-%% \hline
-%% gloor2002 & X & X & X & X \\
-%% \hline
-%% gloor2004 & X & X & X & X \\
-%% \hline
-%% gloor2007 & X & X & X & X \\
-%% \hline
-%% gross2009 & X &  & X & X \\
-%% \hline
-%% Lischner2010 &  & X &  &  \\
-%% \hline
-%% Blas2001 & X & X & X & X \\
-%% \hline
-%% Kahl2008 & X &  & X & X \\
-%% \hline
-%% Muller2001 & X &  &  &  \\
-%% \hline
-%% Yang1994 &  & X &  & X \\
-%% \hline
-%% ding1987 &  & X & ? &  \\
-%% \hline
-%% Jaqaman2004 &  & X & ? & X \\
-%% \hline
-%% next & ? & ? & ? & ? \\
-%% \hline
-%% \end{tabular}
-%% \caption{Which papers do what}
-%% \end{table}
-
-Functionals constructed using a local density approximation fail to
-satisfy the contact-value theorem, and therefore incorrectly model
-small hard solutes.  The contact-value theorem relates the pressure on
-a hard surface to the contact density of the fluid at that
-surface:
-\begin{equation}\label{eq:contact}
-  p(\rr_c) = n(\rr_c)k_BT,
-\end{equation}
-where $\rr_c$ is the position at which a molecule is in contact with
-the hard surface, $n(\rr_c)$ is the density at that point of contact,
-and $p(\rr_c)$ is the pressure that the fluid exerts on the surface at
-the same point.  This pressure is defined as a ratio of force to
-\emph{solvent accessible surface} area.
-%
-For a solute which excludes the solvent from an arbitrarily small
-volume, the contact density will be the same as the bulk density, and
-therefore we can integrate the above pressure to find that the excess
-chemical potential of a small hard solute is proportional to the
-solvent-excluded volume:
-\begin{equation}\label{contactvaluethm}
-  F = n k_BT V.
-\end{equation}
-The contact-value theorem is violated by local classical density
-functionals such as those using a local density approximation or a
-square-gradient term, but is satisfied by non-local classical density
-functionals, such as those using a weighted-density approach.
-
-\section{Theory and Methods}
-We construct a classical density functional for water, which reduces
-in the homogeneous limit to the optimal SAFT model for water found by
-Clark~\emph{et al}.  The Helmholtz free energy is constructed using the
-first four terms from Equation \ref{eq:SAFT-free-energy}: $F_\textit{id}$,
-$F_\textit{hs}$, $F_\textit{disp}$ and $F_\textit{assoc}$.  In the
-following sections, we will introduce the terms of this functional.
-
-\subsection{Ideal gas functional}
-The first term is the ideal gas free energy functional, which is
-purely local:
-\begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln({n(\xx){\Lambda}^3}) - 1\right) d\xx,
-\end{equation}
-where n(\xx) is the density of water molecules and $\Lambda$ is the
-thermal wavelength
-%\begin{equation}\label{quantumconcentration}
-$ \Lambda = \left(\frac{2\pi\hbar^2}{mk_BT}\right )^{1/2}$.
-%\end{equation}
-The ideal gas free energy functional on its own satisfies the contact
-value theorem and its limiting case of small solutes (Equations
-\ref{eq:contact} and \ref{contactvaluethm}). These properties are
-retained by our total functional, since all the remaining terms are
-purely nonlocal.
-
-\subsection{Hard-sphere repulsion}
-
-We treat the hard-sphere repulsive interactions using the White Bear version of
-the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
-fluid~\cite{roth2002whitebear, yu2002structures}.  FMT functionals are expressed as the integral of
-the \emph{fundamental measures} of a fluid, which provide local
-measures of quantities such as the packing fraction, density of
-spheres touching a given point and mean curvature.  The hard-sphere
-excess free energy is written as:
-\begin{equation}
-F_{hs}[n] = k_B T \int (\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)) d\xx \; ,
-\end{equation}
-with integrands
-\begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
-\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= (n_2^3 - 3n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-  \frac{
-    n_3 + (1-n_3)^2\ln(1-n_3)
-  }{
-    36\pi n_3^2\left( 1 - n_3 \right)^2
-  } ,
-\end{align}
-where the fundamental measure densities are given by:
-\begin{align}
-  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
-  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
-  \\
-  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 \\
-  n_1 &= \frac{n_2}{4\pi R}\\
-  \mathbf{n}_{V1} &= \frac{\mathbf{n}_{V2}}{4\pi R}\\
-  n_0 &= \frac{n_2}{4\pi R^2}.
-\end{align}
-The density $n_3$ is the packing fraction and $n_0$ is the
-average density at contact distance. For our functional for water, we
-use the hard-sphere diameter of 3.0342~\AA, which was found to be
-optimal by Clark \emph{et al}.\cite{clark2006developing}
-
-\subsection{Dispersion free energy}
-The dispersion free energy includes the van der Waals attraction and
-any orientation-independent interactions. We use a dispersion term
-based on the SAFT-VR approach\cite{gil-villegas-1997-SAFT-VR}, which
-has two free parameters (taken from Clark~\emph{et
-  al}\cite{clark2006developing}): an interaction energy $\epsilondisp$
-and a length scale $\lambdadisp R$.
-
-The SAFT-VR dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
-\begin{align}
-  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
-\end{align}
-where $a_1$ and $a_2$ are the first two terms in a high-temperature
-perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
-the mean-field dispersion interaction. The second term, $a_2$, describes the
-effect of fluctuations resulting from compression of the fluid due
-to the dispersion interaction itself, and is approximated
-using the local compressibility approximation (LCA), which
-assumes the energy fluctuation is simply related to the
-compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
-
-The form of $a_1$ and $a_2$ for SAFT-VR is given in
-reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
-of the packing fraction.  In order to apply this form to an
-\emph{inhomogeneous} density distribution, we construct an effective local
-packing fraction for dispersion $\etadisp$, given by a Gaussian
-convolution of the density:
-\begin{align}
-%    ((4/3.0*M_PI)*Pow(3)(R)*GaussianConvolve(2*lambdainput*lscale*radius,
-%                                             2*lambdaE*length_scalingE*Rexpr)
-%  \etadisp(\xx) &= \color{red}
-%  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
-%             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
-  \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
-  \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
-      \lscale R)^2}}\right)d\xx'.
-\end{align}
-This effective packing fraction is used throughout the dispersion
-functional, and represents a packing fraction averaged over the
-effective range of the dispersive interaction.  Here we have
-introduced an additional empirical parameter $\lscale$ which modifies
-the length scale over which the dispersion interaction is correlated.
-
-\subsection{Association free energy}
-The final attractive energy term is the association term, which
-accounts for hydrogen bonding.  Hydrogen bonds are modeled as four
-attractive patches (``association sites'') on the surface of the hard
-sphere.  These four sites represent the two hydrogen bond donor sites,
-and two hydrogen bond acceptor sites.  There is an attractive energy
-$\epsilonassoc$ when two molecules are oriented such that the donor site
-of one overlaps with the acceptor site of the other.  The volume over
-which this interaction occurs is $\kappaassoc$, giving the association
-term in the free energy two empirical parameters that are fit to the
-experimental equation of state of water (again, taken from
-Clark~\emph{et al}\cite{clark2006developing}).
-
-The association functional we use is a modified version of Yu
-and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}, which
-% \footnote{We
-%  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
-%  the same functional, but their paper contains errors in the
-%  association term and is not useful as a reference.}
-includes the effects of density inhomogeneities in the contact value
-of the correlation function $g^{HS}_\sigma$, but is based on the
-SAFT-HS model, rather than the SAFT-VR
-model\cite{gil-villegas-1997-SAFT-VR}, which is used in the optimal
-SAFT parametrization for water of Clark \emph{et
-  al}\cite{clark2006developing}.  Adapting Yu and Wu's association
-free energy to SAFT-VR simply involves the addition of a correction
-term in the correlation function (see Equation~\ref{eq:gSW}).
-
-The association functional we use is constructed by using the density
-$n_0(\xx)$, which is the density of hard spheres touching a given
-point, in the standard SAFT-VR association
-energy\cite{gil-villegas-1997-SAFT-VR}.
-The association free energy for our four-site model has the form
-\begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx) \zeta(\xx)
-  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx,
-\end{align}
-where the factor of $4$ comes from the four association sites per
-molecule, the functional $X$ is the fraction of association sites
-\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
-measure of the density inhomogeneity.
-\begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{V2}\cdot\mathbf{n}_{V2}}{n_2^2}.
-\end{align}
-%
-The fraction $X$ is determined by the quadratic equation
-\begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)
-      \Delta(\xx)} - 1}
-  {4 n_0(\xx)\zeta(\xx)
-    \Delta(\xx)},
-\end{align}
-where the functional $\Delta$ is a measure of hydrogen-bonding
-probability, given by
-\begin{align}
-  \Delta(\xx) &= \kappaassoc g^\textit{SW}_\sigma(\xx)
-  \left(e^{-\beta\epsilonassoc} - 1\right) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\xx) +
-  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
-  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)\label{eq:gSW},
-\end{align}
-% This is equation 77 of gil-villegas-1997-SAFT-VR
-where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
-contact for a hard-sphere fluid with a square-well dispersion
-potential, and $a_1$ and $a_2$ are the two terms in the dispersion
-free energy.  The correlation function $g^\textit{SW}_\sigma$ is
-written as a perturbative correction to the hard-sphere fluid
-correlation function $g^\textit{HS}_\sigma$, for which we use the
-functional of Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
-\begin{align}
-  g_\sigma^{HS} &= \frac{1}{1-n_3}
-  +\frac{R}{2}\frac{\zeta n_2}{(1-n_3)^2}
-  + \frac{R^2}{18}\frac{\zeta n_2^2}{(1-n_3)^3}.
-\end{align}
-
-\subsection{Determining the empirical parameters}\label{sec:empirical}
-
-%% \begin{figure}
-%% \begin{center}
-%% \includegraphics[width=3.5in]{figs/surface-tension}
-%% \end{center}
-%% \caption{Comparison of Surface tension versus temperature for theoretical and
-%%   experimental data. The experimental data is taken from NIST.\cite{nistwater}
-%%   The length-scaling parameter $\lscale$ is fit so that the theoretical surface 
-%%   tension will match the experimental surface tension near room temperature.}
-%% \label{fig:surface-tension}
-%% \end{figure}
-
-The majority of the empirical parameters used in our functional are
-taken from the paper of Clark \emph{et al} on developing an optimal
-SAFT model for water\cite{clark2006developing}.  This SAFT model
-contains five empirical parameters: the hard-sphere radius, an energy
-and length scale for the dispersion interaction, and an energy and
-length scale for the association interaction.  In addition to the five
-empirical parameters of Clark \emph{et al}, we add a single additional
-dimensionless parameter $\lscale$---with a fitted value of
-0.353---which determines the length scale over which the density is
-averaged when computing the dispersion free energy and its derivative.
-We determine this final parameter by fitting the computed surface
-tension to the experimental surface tension with the result shown in
-Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
-\emph{et al} overestimates the critical temperature---which is a
-common feature of SAFT-based functionals that do not explicitly treat
-the critical point---we cannot reasonably describe the surface tension
-at all temperatures, and choose to fit the surface tension at and
-around room temperature.  We note here that we could have chosen to
-fit the surface tension with a square-gradient term in the free energy
-instead of adjusting the length scale for the dispersive attraction.
-This would result in a functional that violates the contact-value
-theorem which, among other problems, would fail to satisfy
-Equation~\ref{contactvaluethm} for the excess chemical potential of
-small solutes.
-
-\section{Results and discussion}
-
-\subsection{One hydrophobic rod}
-
-We begin by studying a single hydrophobic rod immersed in water.
-% We
-% model the density on a grid with spacing of about 0.01~nm, which gives
-% the free energy of hydration to a precision of better than a percent,
-% and the density at contact to a few percent precision.
-In Figure\fixme{deleted} we show the excess chemical
-potential at room temperature, scaled by the solvent accessible
-surface area of the hard rod, plotted as a function of hard-rod
-radius.  We define the hard-rods radius as the radius from which water
-is excluded.  For rods with radius larger than 0.5~nm or so, this
-reaches a maximum value of 75~mN/m, which is slightly higher than
-macroscopic surface tension.  In the limit of very large rods, this
-value will decrease and approach the macroscopic value.  As seen in
-the inset of Fig.~\fixme{deleted}, for rods with very
-small radius (less than about 0.5~\AA) the excess chemical potential
-is proportional to volume, satisfying Equation~\ref{contactvaluethm},
-which results from the contact-value theorem.
-
-We show in Figure~\ref{fig:density-single-rod} density profiles for
-different radii rods, as well as the prediction for the contact value
-of the density as a function of rod radius, as computed from the
-excess chemical potentials plotted in
-Figure~\fixme{deleted}.  The agreement between these
-curves confirms that our functional satisfies the contact-value
-theorem and that our minimization is well converged.  As expected, as
-the radius of the rods becomes zero the contact density approaches the
-bulk density, and as the radius becomes large, the contact density
-will approach the vapor density.
-
-\subsection{Hydrophobic interaction of two rods}
-
-We now look at the more interesting problem of two parallel hard rods
-in water, separated by a distance $d$, as shown in
-Figure~\ref{fig:density-rods}.  At small separations there is only vapor
-between the rods, but as the rods are pulled apart, the vapor region
-expands until a critical separation is reached at which point liquid
-water fills the region between the rods.
-Figure~\ref{fig:density-rods} shows density profiles before and after
-this transition for rods of radius 0.6~nm. This critical separation
-for the transition to liquid depends on the radii of the rods, and is
-about 0.65~nm for the rods shown in Figure~\ref{fig:density-rods}.
-The critical separation will be different for a system where there is
-attraction between the
-rods and water.  At small separations, the shape of the water around the two
-rods makes them appear as one solid ``stadium''-shaped object
-(a rectangle with semi-circles on both ends).
-
-%\begin{figure}
-%\begin{center}
-%\includegraphics[width=3.5in]{figs/rods-diagram}
-%\end{center}
-%\caption{ Cross section schematic for two hydrophobic rods immersed in water.
-%The radius and separation are shown.}
-%\label{fig:rods}
-%\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/density-rods-in-water}
-\end{center}
-\caption{ Density profiles illustrating the transition from vapor 
-to liquid water between the rods. The radius is 0.6~nm, the top figure is
-at a separation of 0.6~nm and the
-bottom is 0.7~nm. Figure~\ref{fig:rods-energy-vs-distance} shows
-the energy for these and other separations.}
-\label{fig:density-rods}
-\end{figure}
-
-To understand this critical separation, we consider the free energy in
-the macroscopic limit, which is given by
-\begin{equation}
-F = \gamma A + pV.
-\end{equation}
-The first term describes the surface energy and the second term is the
-work needed to create a cavity of volume $V$. Since the pressure term
-scales with volume, it can be neglected relative to the surface term
-provided the length scale is small compared with $\gamma / p$, which
-is around $20~\mu m$, and is much larger than any of the systems we
-study. For micron-scale rods, the water on the sides of the `stadium'
-configuration will bow inward between the rods and the density will
-reduce to vapor near the center point where the rods are closest to
-each other.
-
-Starting from the surface energy term, we can calculate the 
-free energy per length, which is equal to the circumference multiplied 
-by the surface tension. 
-Working out the circumference of the stadium-shape leads us to
-\begin{equation}
-F = (2\pi r +4r+2d)\gamma
-\end{equation}
-where $\gamma$ is the surface tension, $r$ is the radius of the rods,
-and $d$ is the separation between rods illustrated in
-Figure~\ref{fig:density-rods}.  The force per length is the derivative
-of the free energy with respect to the separation $d$, from which we
-conclude that the force per length is twice the
-surface tension.
-
-We plot in Figure~\ref{fig:rods-energy-vs-distance} the computed free
-energy of interaction
-per unit length from our classical density functional (solid lines),
-as a function of the separation $d$, along with the free energy
-predicted by our simple macroscopic model (dashed lines).  The models
-agree very well on the force between the two rods at close
-separations, and have reasonable agreement as to the critical
-separation for rods greater than 0.5~nm in radius.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/rods-energy-vs-distance}
-\end{center}
-\caption{ Free energy of interaction (also known as the potential of mean force)
-versus separation for two hydrophobic rods ranging in radius from
-0.2~nm to 1.2~nm.
-All were arbitrarily offset to zero at large separations for ease of comparison. The
-transition corresponds to the phase change from
-vapor to liquid between the rods as pictured in the density profiles in 
-Figure~\ref{fig:density-rods}. }
-\label{fig:rods-energy-vs-distance}
-\end{figure}
-
-Walther \emph{et al}\cite{walther2004hydrophobic} studied the
-interactions between two carbon nanotubes, which are geometrically
-similar to our hydrophobic rods, using molecular dynamics with the SPC
-model for water.  Their simulations used nanotubes of diameter 1.25~nm
-and separations ranging from about 0.3~nm to 1.5~nm.  In agreement
-with our findings for two purely hydrophobic rods, Walther \emph{et
-  al} find that in the absence of Lennard-Jones attraction between
-carbon and oxygen, there is a drying transition at a distance
-comparable to the diameter of the nanotube.  In contrast to this, when
-the attraction between nanotubes and water is turned on, they find
-that the drying transition occurs at much shorter distances,
-comparable to the diameter of water.
-
-\subsection{Hydrophobic interactions of four rods}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/four-rods-energy-vs-distance}
-\end{center}
-\caption{ Free energy of interaction
-versus separation for four hydrophobic rods ranging in radius from
-0.2~nm to 1.2~nm.
-All were arbitrarily offset to zero at large separations. The
-transition corresponds to the phase change from
-vapor to liquid between the rods as pictured in the density profiles in 
-Figure~\ref{fig:density-four-rods}. }
-\label{fig:four-rods-energy-vs-distance}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/density-four-rods}
-\end{center}
-\caption{ Density profiles illustrating the transition from vapor 
-to liquid water between four rods. The radius is 0.6~nm, the top figure is
-at a separation of 1.53~nm and the
-bottom is 1.56~nm. Figure~\ref{fig:four-rods-energy-vs-distance} shows
-the energy for these and other separations.}
-\label{fig:density-four-rods}
-\end{figure}
-
-We go on to study four parallel hard rods, as examined by Lum,
-Chandler and Weeks in their classic paper on hydrophobicity at small
-and large length scales~\cite{lum1999hydrophobicity}.  As in the case
-of two rods---and as predicted by Lum~\emph{et al}---we observe a
-drying transition, as seen the density plot shown in
-Figure~\ref{fig:density-four-rods}.  In
-Figure~\ref{fig:four-rods-energy-vs-distance}, we plot the free energy
-of interaction together with the macroscopic approximation, and find
-good agreement for rods larger than 0.5~nm in radius.  This free
-energy plot is qualitatively similar to that predicted by the LCW
-theory~\cite{lum1999hydrophobicity}, with the difference that we find
-no significant barrier to the association of four rods.
-
-\subsection{Hydration energy of hard-sphere solutes}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/sphere-energy-vs-diameter}
-\end{center}
-\caption{ Excess chemical potential per area versus radius for
-  a single hydrophobic sphere
-  immersed in water. This should have an asymptote equal to the
-  surface tension at room temperature, and it agrees well with the
-  surface tension in Figure~\ref{fig:surface-tension}. Results from a
-  simulation of SPC/E water~\cite{huang2001shs} are shown as circles.
-  The horizontal lines show the experimental and SPC/E macroscopic surface
-  tension for water at standard atmospheric temperature and
-  pressure. }
-\label{fig:sphere-energy-vs-diameter}
-\end{figure}
-
-A common model of hydrophobic solutes is the hard-sphere solute, which
-is the simplest possible solute, and serves as a test case for
-understanding of hydrophobic solutes in
-water\cite{sedlmeier2011entropy}.  As in the single rod, we begin by
-examining the ratio of the excess chemical potential of the cavity
-system to the solvent accessible surface area
-(Figure~\ref{fig:sphere-energy-vs-diameter}).  This
-effective surface tension surpasses the macroscopic surface tension at a
-radius of almost 1~nm, and at large radius will drop to the macroscopic value.  As
-with the single rod, we see the analytically correct behavior in the
-limit of small solutes.  For comparison, we plot the free energy
-calculated using a molecular dynamics simulation of SPC/E
-water\cite{huang2001shs}.  The agreement is quite good, apart from the
-issue that the SPC/E model for water significantly underestimates the
-macroscopic surface tension of water at room
-temperature\cite{vega2007surface}.
-
-Figure~\ref{fig:density-sphere} shows the density profile for several
-hard sphere radii, plotted together with the results of the same
-SPC/E molecular dynamics simulation shown in
-Figure~\ref{fig:sphere-energy-vs-diameter}\cite{huang2001shs}.  The
-agreement with simulation is quite reasonable.  The largest
-disagreement involves the density at contact, which according to the
-contact value theorem cannot agree, since the free energies do not
-agree.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/density-sphere}
-\end{center}
-\caption{ Density profiles around hard-sphere solutes of different radii. Predictions
-  from our classical density-functional theory are in solid red, while
-  the dotted line shows the result of a molecular dynamics simulation
-  of SPC/E water~\cite{huang2001shs}.  }
-\label{fig:density-sphere}
-\end{figure}
-
-\section{Conclusion}
-
-We have developed a classical density functional for water that
-combines SAFT with the fundamental-measure theory for hard spheres,
-using one additional empirical parameter beyond those in the SAFT
-equation of state, which is used to match the experimental surface
-tension.  This functional does \emph{not} make a local density
-approximation, and therefore correctly models water at both small and
-large length scales.  In addition, like all FMT functionals, this
-functional is expressed entirely in terms of convolutions of the
-density, which makes it efficient to compute and minimize.
-
-We apply this functional to the case of hard hydrophobic rods and
-spheres in water.  For systems of two or four hydrophobic rods
-surrounded by water, we see a transition from a vapor-filled state a
-liquid-filled state.  A simple model treatment for the critical
-separation for this transition works well for rods with diameters
-larger than 1~nm.  In the case of spherical solutes, we find good
-agreement with SPC/E simulations.
-
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
 \end{document}

### 2013-08-14 papers/water-saft/paper.tex
edit water paper some more Tested everything on bingley.

@@
 Classical DFT is based on a description of a fluid written as a free
-energy functional of the density distribution.  There are two
+energy functional of the density distribution.  There are two general
 approaches used to construct a classical DFT for water.  The first is
-to choose a functional form that is fit to properties of the bulk
-liquid at a given temperature and pressure
-\cite{jeanmairet2013molecular} ~\fixme{cite Jeanmairet and similar}.
-Using this approach, it is possible to construct a functional that
-reproduces the exact second-order response function of the liquid
-under the fitted conditions.  This class of functional, however, will
-be less accurate at other temperatures or pressures---and in the
-inhomogeneous scenarios in which solvation models are applied.  The
-second approach is construct a functional by applying liquid-state
-theory to a model system, and then fit the model to experimental data
-such as the equation of state.  Classical DFT is a natural framework
-for creating a more flexible theory of hydrophobicity that can readily
-describe interaction of water with arbitrary external
+to choose a convenient functional form which is then fit to properties
+of the bulk liquid at a given temperature and pressure
+\cite{jeanmairet2013molecular, lischner2010classical} ~\fixme{cite
+  Jeanmairet and similar}.  Using this approach, it is possible to
+construct a functional that reproduces the exact second-order response
+function of the liquid under the fitted conditions.  However, this
+class of functional will be less accurate at other temperatures or
+pressures---and in the inhomogeneous scenarios in which solvation
+models are applied.  The second approach is construct a functional by
+applying liquid-state theory to a model system, and then fit the model
+to experimental data such as the equation of
+state~\cite{hughes2013classical, clark2006developing, gloor2002saft,
+  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
+  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination,
+  sundararaman2012computationally}.  Classical DFT is a natural
+framework for creating a more flexible theory of hydrophobicity that
+can readily describe interaction of water with arbitrary external
 potentials---such as potentials describing strong interactions with
@@
 density functionals is based on Statistical Associating Fluid Theory
-(SAFT). SAFT is a theory for fluids in which hydrogen bonding plays a
-significant role\cite{chapman1989saft, muller2001molecular}. SAFT is
-used to accurately model the equations of state of both pure fluids
-and mixtures over a wide range of temperatures and pressures.
-\fixme{Talk about and cite papers using SAFT to build DFTs.}
+(SAFT)~\cite{chapman1989saft}.  SAFT is a theory based on a model of
+hard sphere with weak dispersion interactions and hydrogen-bonding
+association sites, which has been used to accurately model the
+equations of state of both pure fluids and mixtures over a wide range
+of temperatures and pressures~\cite{muller2001molecular,
+  tan2008recent}.  The association term in SAFT is based on Wertheim's
+first-order thermodynamic perturbation theory, which models fluids as
+hard-spheres with strong associative interactions at specific sites on
+the surfaces of the spheres~\cite{wertheim1984fluidsI,
+  wertheim1984fluidsII, wertheim1986fluidsIII, wertheim1986fluidsIV}.
+These association sites have an attractive interaction at contact, and
+rely on the hard-sphere radial distribution function at contact
+$g_\sigma^\text{HS}$ in order to determine the extent of association.
+While this function is well-known for the homogeneous hard-sphere
+fluid, it is less well-known for inhomogeneous systems, such as occur
+at liquid interfaces.  In a recent paper, we examined this function at
+a variety of hard-sphere interfaces~\cite{schulte2012using}.  We
+tested the accuracy of existing approximations for the distribution
+function at contact~\cite{yu2002fmt-dft-inhomogeneous-associating,
+  gross2009density}, and derived a significantly improved
+approximation for the averaged distribution function at contact.
 
-SAFT is based on Wertheim's first-order thermodynamic perturbation
-theory (TPT1)\cite{wertheim1984fluidsI, wertheim1984fluidsII,
-  wertheim1986fluidsIII, wertheim1986fluidsIV}, which models
-\fixme{fluids} as hard-spheres with strong associative interactions at
-sites on the surfaces of the spheres. This is ideal for water, since
-its four hydrogen bond sites can be represented by these association
-sites. The association sites have an effective range and a small
-energy contribution when a site is within range of another
-hard-sphere's site. The free energy due to this association depends on
-the radial distribution function $g_\sigma^\text{HS}(\rr)$. Our
-previous water functional used Yu and Wu's radial distribution
-function \cite{yu2002fmt-dft-inhomogeneous-associating}, which was
-found to agree less with hard-sphere Monte-Carlo simulation than a
-distribution function developed by us in a recent paper
-\cite{schulte2012using}. In our paper, we developed an approximation
-for correlation at contact to describe the hard-sphere fluid.
-
-In a recent paper \cite{hughes2013classical}, we constructed a
-classical DFT for water which uses SAFT as a basis for a free energy
-functional. We used Clark \emph{et al}'s \cite{clark2006developing}
-five fitting parameters for an optimal cDFT for water and, in
-addition, introduced a parameter for the length scale over which the
-dispersion interaction is correlated.
+We will apply this improved $g_\sigma^{HS}(\rr)$ to a SAFT-based
+classical density functional for water~\cite{hughes2013classical}.
+This functional was constructed so as to reproduce in the homogeneous
+limit the 4-site optimal SAFT model for water developed by Clark
+\emph{et al.}~\cite{clark2006developing}.  The DFT of
+Ref.~\citenum{hughes2013classical} uses the association free energy
+functional of Yu and
+Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is based on a
+$g_\sigma^{HS}$ that we have since found to be
+inaccurate~\cite{schulte2012using}.  In this paper, we will examine
+the result of using the improved functional for $g_\sigma^{HS}$
+developed in Ref.~\citenum{schulte2012using} to construct an
+association free energy functional.
 
@@
 \end{align}
-where
+where $n_\text{site}(\rr)$ is the density of bonding sites at
+position~$\rr$:
 \begin{align}
@@
 \end{align}
-is the density of bonding sites at position~$\rr$.  The four in
-$n_\text{site}(\rr)$ comes from the four hydrogen bond sites,
+where the factor of four comes from the four hydrogen bond sites,
 $n_0(\rr)$ is the average density contacting point $\rr$, and
@@
 functional $X(\rr)$ is the fraction of association sites \emph{not}
-hydrogen-bonded, which is determined by the quadratic equation
+hydrogen-bonded, which is determined for our 4-site model by the
+quadratic equation
 \begin{align}
@@
 potential, and $a_1$ and $a_2$ are the two terms in the dispersion
-free energy.  The radial distribution function $g^\textit{SW}_\sigma$ is
-written as a perturbative correction to the hard-sphere fluid
-correlation function $g^\textit{HS}_\sigma$. Four our old functional
-we used Yu and Wu's radial distribution function, and for this paper we
-use our $g^\textit{A}_\sigma$.
+free energy.  The radial distribution function of the square-well
+fluid $g^\textit{SW}_\sigma$ is written as a perturbative correction
+to the hard-sphere fluid correlation function
+$g^\textit{HS}_\sigma$. For \fixme{our} old functional we used Yu and Wu's
+radial distribution function, and for this paper we use our
+$g^\textit{A}_\sigma$.
 
@@
 water. Figure~\ref{fig:density-single-rod} shows the density profile
-of water near a rod with radius~0.1\AA\fixme{Eric, please figure this
+of water near a rod with radius~0.1~\AA\fixme{Eric, please figure this
   radius out}.  The density computed using the functional of this

### 2013-08-14 papers/pair-correlation/paper.tex
minor notational fixes in pair paper Tested everything on bingley.

@@
 points, that are themselves functions of our published pair
-correlation at contact $g_{sigma}$ evaluated at the two points.
+correlation at contact $g_{\sigma}$ evaluated at the two points.
 
@@
 \begin{align}
-  g(|\rr|) &= 1 + g_{sigma}e^{-x_0(r-\sigma)} + x_1g_{sigma}sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
-  &~~-x_4g_{\sigma}^2sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
+  g(|\rr|) &= 1 + g_{\sigma}e^{-x_0(r-\sigma)} + x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
+  &~~-x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
 \end{align}
@@
 %% in the homogeneous limit.  Created a function that has the form of an
-%% expansion in powers of $g_{sigma}$.  $g_{sigm}$ in the homogeneous
+%% expansion in powers of $g_{\sigma}$.  $g_{sigm}$ in the homogeneous
 %% limit can be calculated using the well known and very successful

### 2013-08-16 papers/pair-correlation/paper.tex
move references to end of paper in pair paper Tested everything on bingley.
remove unwanted paragraphs around equations Tested everything on darcy.
add sketchy derivation of g_sigma' to pair paper Tested everything on darcy.

@@
-\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
+\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 The function has the form:
-
 \begin{align}
@@
 \end{align}
-
 The finalalized form of our pair correlation function is:
-
 \begin{align}
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g}\\
-  \includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
+  \includegraphics[width=\columnwidth]{figs/ghs-g}% \\
+  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
@@
 Figure~\ref{fig:dadz} shows...
+
+\appendix
+
+\section{Properties of $g_\sigma$}
+We begin with some definitions of basic quantities, looking at the
+homogeneous hard-sphere fluid.  I am ignoring here the distinction
+between $N$ and $N-1$, which is unimportant in the thermodynamic limit
+$N\gg 1$.
+\begin{align}
+  Z(\sigma) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
+  \\
+  n^{(2)}(\rr_1,\rr_2) &= \frac{N^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
+\end{align}
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) &=
+  \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
+  \\
+  &= \frac{V^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j}
+  \Theta(r_{ij}-\sigma)
+  \label{eq:g2-integral-defn}
+\end{align}
+
 \begin{figure}
@@
 \end{figure}
+
+At this point we consider the effect of increasing the diameter
+$\sigma$ of all our spheres.  This will reduce the number of available
+states, thus reducing the partition function $Z$, which we will
+ultimately relate to a change in free energy.
+\begin{align}
+  Z(\sigma+\delta) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma-\delta)
+  \\
+  &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)(1-\Theta(\sigma+\delta-r_{ij}))
+\end{align}
+I have written $Z(\sigma+\delta)$ in such a way that it is clear that
+it is equal to $Z(\sigma)$ plus a small quantity, provided $\delta$ is
+small.  We will now investigate that small quantity, which we will
+write as a power series in $\delta$.  We begin by expanding the
+product of $(1-\Theta(\sigma+\delta-r_{ij}))$:
+\begin{multline}
+    \prod_{i<j}(1-\Theta(\sigma+\delta-r_{ij})) =
+    1 - \sum_{i<j} \Theta(\sigma+\delta-r_{ij}) \\
+    + \mathop{\mathop{\sum_{i<j}}_{k<}}_{kl \ne ij}
+    \Theta(\sigma+\delta-r_{ij}) \Theta(\sigma+\delta-r_{kl})
+    \\ + \cdots.
+\end{multline}
+Given this expansion, we can readily expand the partition function:
+\begin{multline}
+  Z(\sigma+\delta) = Z(\sigma) \\ -
+  \sum_{k<l} \iiiint \Theta(\sigma+\delta-r_{kl}) \prod_{i<j}
+  \Theta(r_{ij}-\sigma) \\
+  + \mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
+  \iiiint \Theta(\sigma+\delta-r_{kl}) \Theta(\sigma+\delta-r_{mn}) \prod_{i<j}
+  \Theta(r_{ij}-\sigma) \\
+  + \cdots.\label{eq:z-power-series}
+\end{multline}
+Our challenge now is to manipulate this into forms involving $g(r)$,
+which we can then simplify using a Taylor expansion around
+$r=\sigma$.  We will begin by taking advantage of the fact that the
+summations in Eq.~\ref{eq:z-power-series} are mostly of identical
+terms, since the integrals are over all coordinates.
+\begin{multline}
+  Z(\sigma)-Z(\sigma+\delta) =
+  N^2\iiiint \Theta(\sigma+\delta-r_{12})
+  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
+  - \frac{N^3}{2}\iiiint
+  \Theta(\sigma+\delta-r_{12})
+  \Theta(\sigma+\delta-r_{34})
+  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
+  + \cdots \label{eq:z-expansion}
+\end{multline}
+At this point we are ready to connect with $g(r)$, which we can do by
+considering an integral of the pair distribution function, from
+Eq.~\ref{eq:g2-integral-defn}.
+\begin{multline}
+  \int_{\sigma}^{\sigma+\delta} g(r) d\rr\\
+  \shoveleft{\quad= \int g^{(2)}(\rr_1,\rr_2)
+  \Theta(\sigma+\delta-r_{12})d\rr_2} \\
+  \shoveleft{\quad= \frac{1}{V}\int g^{(2)}(\rr_1,\rr_2)
+  \Theta(\sigma+\delta-r_{12})d\rr_1d\rr_2} \\
+  \shoveleft{\quad= \frac{V}{Z}\iiiint d\rr_1\cdots d\rr_N
+  \Theta(\sigma+\delta-r_{12})}\\
+  \prod_{i<j} \Theta(r_{ij}-\sigma)
+  \label{eq:g-shell-integral}
+\end{multline}
+Looking back at Eq.~\ref{eq:z-expansion}, we can see that the first
+term is easily related to Eq.~\ref{eq:g-shell-integral}.  The second
+term has two shell $\delta$ functions, and is thus a bit trickier.
+Fortunately, in a very large system, $\rr_3$ and $\rr_4$ are
+vanishingly seldom near $\rr_1$ or $\rr_2$, so we can write the second
+term as a product.
+\begin{multline}
+  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)}\\
+  \shoveleft{\quad= nN \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
+  - \frac{n^2N}{2} \left(\int_{\sigma}^{\sigma+\delta} g(r)
+  d\rr\right)^2
+  \\
+  + \frac{n^3N}{6} \left(\int_{\sigma}^{\sigma+\delta} g(r)
+  d\rr\right)^3 + \cdots
+  \label{eq:zdiff-g}
+\end{multline}
+Now let us look at how this integral relates to $g(r)$ evaluated at
+contact.  I'll skip a few steps doing this integral...
+\begin{equation}
+   \int_{\sigma}^{\sigma+\delta} g(r)
+  d\rr = 4\pi\sigma^2\left(g_\sigma\delta + \left(\tfrac12g_\sigma' +
+  \tfrac1{\sigma}\right) + \cdots \right)
+\end{equation}
+Putting this together with the above we find
+\begin{multline}
+  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
+  4\pi\sigma^2 n g_\sigma \delta +
+  4\pi\sigma^2 n \left(\tfrac{g_\sigma'}{2} + \tfrac{2}{\sigma}\right)
+  \delta^2 \\
+  - \tfrac{1}{2}\left(4\pi\sigma^2ng_\sigma\right)^2\delta^2 + \mathcal{O}(\delta^3)
+\end{multline}
+But by considering a Taylor expansion, we can also see that
+\begin{multline}
+  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
+  -\frac{\partial Z}{\partial \sigma} \delta
+  -\frac12 \frac{\partial^2 Z}{\partial \sigma^2} \delta^2  + \mathcal{O}(\delta^3)
+\end{multline}
+from which we can extract relationships for the derivatives of $Z$
+with respect to $\sigma$.  Which if we knew what those derivatives
+were, would tell us about $g_\sigma'$ and $g_\sigma$, which is what we
+are after here.  We find this out by considering the Helmholtz free
+energy.
+\begin{equation}
+  F = -kT \ln Z
+\end{equation}
+The derivatives are then given by
+\begin{equation}
+  \frac{\partial F}{\partial \sigma} = -\frac{kT}{Z} \frac{\partial
+    Z}{\partial \sigma}
+\end{equation}
+and
+\begin{equation}
+  \frac{\partial^2 F}{\partial \sigma^2} =
+  \frac{1}{kT}\left(\frac{\partial F}{\partial \sigma}\right)^2
+  -\frac{kT}{Z} \frac{\partial^2
+    Z}{\partial \sigma^2}.
+\end{equation}
+With a bunch more work, we find out that
+\begin{equation}
+  g_\sigma' = \frac{\frac{\partial^2 F}{\partial
+      \sigma^2}}{4\pi\sigma^2nkT}
+  - \frac{2g_\sigma}{\sigma}
+\end{equation}
+except that there is a mistake in here.  I can tell because $F$ is
+extensive and $g_\sigma'$ is not.  This tracks back to
+Eq.~\ref{eq:zdiff-g}, which is correct in this \LaTeX, but is wrong on
+my scratch paper.
+
+\bibliography{paper}% Produces the bibliography via BibTeX.
 

### 2013-08-20 papers/water-saft/paper.tex
another water paragraph from yesterday Tested everything on bingley.

@@
 \section{Method}
+
+The classical density functional for water from
+Reference~\citenum{hughes2013classical} consists of four terms:
+\begin{align}
+  F[n(\rr)] &= F_{\text{ideal}}[n(\rr)] + F_{\text{HS}}[n(\rr)]
+  + F_{\text{disp}}[n(\rr)] + F_{\text{assoc}}[n(\rr)]
+\end{align}
+where $F_{\text{ideal}}$ is the ideal gas free energy and
+$F_{\text{HS}}$ is the hard-sphere excess free energy, for which we
+use the White Bear functional~\cite{roth2002whitebear}.
+$F_{\text{disp}}$ is the free energy contribution due to the
+square-well dispersion interaction.  The dispersion term of the free
+energy contains one empirical parameter, which is used to fit the
+surface tension of water near one atmosphere.  Finally,
+$F_{\text{assoc}}$ is the free energy contribution due to association,
+which is the term that we examine in this paper.
 

### 2013-08-21 papers/pair-correlation/paper.tex

@@
 
-\newcommand{\fixme}[1]{\red{[#1]}}
+\newcommand{\fixme}[1]{\text{\red{[#1]}}}
 
@@
 Figure~\ref{fig:radial-distribution} shows...
+
+We here construct a separable fit to the radial distribution function
+$g(r, g_\sigma)$, which is a function of the radius $r$ and its value
+at contact $g_\sigma$.  We could convert this to a more conventional
+function of the packing fraction $\eta$ by inverting the
+Carnahan-Starling formula for $g_\sigma$
+\begin{equation}
+  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}
+\end{equation}
+but that would add an unnecessary step in the use of $g(r)$, since the
+averaged value of $g_\sigma$ is what we can directly compute for the
+inhomogeneous system.
+
+We construct $g(r, g_\sigma)$ using several constraints.  Naturally,
+\begin{equation}
+  g(\sigma, g_\sigma) = g_\sigma.
+\end{equation}
+In addition, we know that
+\begin{equation}
+  \begin{split}
+  \int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
+  &= \eta \eta \cdots \fixme{fill in, double-check}
+  \end{split}
+\end{equation}
+where $\chi_T$ is the isothermal compressibility, for which we use the
+Carnahan-Starling equation of state.
+
+An additional constraint that we use sets the slope of $g(r,
+g_\sigma)$ at $r=\sigma$.  \fixme{explain this}
+
+Given these constraints, we use a functional form of
+\begin{equation}
+  g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
+  + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
+\end{equation}
+where $A$, $B$, and $C$ are amplitudes determined by the above
+constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
+simulations of the hard-sphere radial distribution function.
+
+\fixme{add paragraphs and equations with $A$, $B$ and $C$}
+
+\fixme{add table with parameters}
 

### 2013-08-22 papers/pair-correlation/paper.tex
add triplet plots to paper, fix triplet plot title, and make it build without necessary triplet MC data Tested everything on darcy.

@@
     wall.}\label{fig:pair-distribution-4}
+\end{figure}
+
+\subsection{Triplet distribution function}
+
+In Fig.~\ref{fig:triplet-distribution-3} and
+Fig.~\ref{fig:triplet-distribution-4} we can see the triplet
+distribution computed using Monte Carlo.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-3.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    in contact.}\label{fig:triplet-distribution-3}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-4.pdf}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:triplet-distribution-4}
 \end{figure}

### 2013-08-22 papers/water-saft/paper.tex

@@
 pressures---and in the inhomogeneous scenarios in which solvation
-models are applied.  The second approach is construct a functional by
+models are applied.  The second approach is to construct a functional by
 applying liquid-state theory to a model system, and then fit the model
@@
 (SAFT)~\cite{chapman1989saft}.  SAFT is a theory based on a model of
-hard sphere with weak dispersion interactions and hydrogen-bonding
+hard spheres with weak dispersion interactions and hydrogen-bonding
 association sites, which has been used to accurately model the
@@
 
-We will apply this improved $g_\sigma^{HS}(\rr)$ to a SAFT-based
+We will apply this improved $g_\sigma^{HS}$ to a SAFT-based
 classical density functional for water~\cite{hughes2013classical}.
@@
 
-We modify our previous association free energy term to use our
-$g_{\sigma}^\textit{A}$ rather than Yu and Wu's $g_\sigma$.  The
-association free energy for our four-site model has the form:
+The association free energy for our four-site model has the form:
 \begin{align}
@@
 fluid $g^\textit{SW}_\sigma$ is written as a perturbative correction
-to the hard-sphere fluid correlation function
-$g^\textit{HS}_\sigma$. For \fixme{our} old functional we used Yu and Wu's
-radial distribution function, and for this paper we use our
-$g^\textit{A}_\sigma$.
+to the hard-sphere fluid correlation function $g^\textit{HS}_\sigma$.
 
@@
 higher than that from Ref.~\citenum{hughes2013classical}, and the peak
-is not smooth.  This results from the improved accurately of the
-$g_\sigma$ from Ref.~\citenum{schulte2012using}, since beyond the
+is not smooth.  This results from the improved accuracy of the
+$g_\sigma^\textit{HS}$ from Ref.~\citenum{schulte2012using}, since beyond the
 first peak water molecules are unable to touch---or hydrogen bond

### 2013-08-22 papers/pair-correlation/paper.tex
debugged sphere-with-wall.cpp and triplet-dft.cpp and modified some equations in paper Tested everything on node22.
Added 2d plot data for the path plots. Changed the names of the path plot scripts, they are now:    plot-path.py    plot-path-triplet-contact.py    plot-path-triplet-inbetween.py Also changed the name of their outputs, and adjusted the paper to use the new ones. Tested everything on paho-deb.

@@
   \begin{split}
-  \int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
-  &= \eta \eta \cdots \fixme{fill in, double-check}
+    1 + n\int\left[g(r)-1\right]d\rr = nkT\chi_T\\
+    \chi_T  = -\frac{1}{V}\left(\frac{\partial V}{\partial p}\right)_T
+    = - \left(V\left(\frac{\partial p}{\partial V}\right)_T\right)^{-1}\\
+    p = \frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\Longrightarrow \left(\frac{\partial p}{\partial V}\right)_T = \frac{dp}{d\eta}\frac{d\eta}{dV}\\
+    \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}\\
+    \left(\frac{\partial p}{\partial V}\right)_T = -\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
+  %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
+  %&= \eta \eta \cdots \fixme{fill in, double-check}
   \end{split}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-1.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-2.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-alt-4.pdf}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
   \caption{Pretty plot of the pair distribution function near a hard
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
   \caption{The triplet distribution function
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-4.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:triplet-distribution-4}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    in contact.}\label{fig:triplet-distribution-4}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-distribution-3}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-distribution-4}
 \end{figure}

### 2013-08-22 papers/pair-correlation/figs/fit-ghs.py

@@ -1,177 +1,0 @@
-#!/usr/bin/python
-
-from __future__ import division
-# We need the following two lines in order for matplotlib to work
-# without access to an X server.
-import matplotlib, sys, os.path
-from pylab import *
-from scipy.optimize import leastsq
-
-sigma = 2
-
-x = rand(5)*5
-
-colors = ['r', 'g', 'b', 'c', 'm', 'k', 'y']*2
-ff = array([.1, .2, .3, .4])
-able_to_read_file = True
-
-
-def read_ghs(base, ff):
-  mcdatafilename = "%s-%4.2f.dat" % (base, ff)
-  if (os.path.isfile(mcdatafilename) == False):
-    print "File does not exist: ", mcdatafilename
-    global able_to_read_file
-    able_to_read_file = False
-    return 0, 0
-
-  mcdata = loadtxt(mcdatafilename)
-  print 'Using', mcdatafilename, 'for filling fraction', ff
-  r_mc = mcdata[:,0]
-  n_mc = mcdata[:,1]
-  ghs = n_mc/ff
-  return r_mc, ghs
-
-
-# READ DATA
-ghs = [0]*len(ff)
-#ghslores = [0]*len(ff)
-eta = [0]*len(ff)
-
-
-for i in range(len(ff)):
-    r_mc, ghs[i] = read_ghs("figs/gr", ff[i])
-    if able_to_read_file == False:
-        break
-    #r_mclores, ghslores[i] = read_ghs("grlores", ff[i])
-
-    # The following is the Monte Carlo approximation of the
-    # distribution function at contact.  This gives us an answer with
-    # no systematic error (well, very little, and what we have is due
-    # to the finite bin size), and if we wait long enough the
-    # statistical error should become negligible.
-    #gsig[i] = ghs[i].max()
-    # The following is the Carnahan Starling formula for the
-    # distribution function at contact.  This is approximate, but it
-    # is a good approximation.
-    eta[i] = ff[i]
-    r = r_mc # shift to x = 0 at contact
-
[…]

### 2013-08-22 papers/pair-correlation/figs/plot-path-triplet-contact.py

@@ -25,9 +25,9 @@ able_to_read_file = True
 z0 = 0.05
 
 # Set the max parameters for plotting.
-zmax = 9.0
-zmin = 1.0
-rmax = 4.1
+zmax = 8
+zmin = 0
+rmax = 4
 ############################
 
 if len(sys.argv) < 2:
@@ -38,13 +38,13 @@ ff = float(sys.argv[1])
 
 def read_triplet_path(ff,z0,fun):
   if fun == 'mc':
-    # input:  "figs/mc/triplet/tripletMC-%03.1f-path2-trimmed.dat" % (ff)
-    filename = "figs/mc/triplet/tripletMC-%03.1f-path2-trimmed.dat" % (ff)
+    # input:  "figs/mc/triplet/tripletMC-%03.1f-path-trimmed.dat" % (ff)
+    filename = "figs/mc/triplet/tripletMC-%03.1f-path-trimmed.dat" % (ff)
   # elif fun == 'sphere-dft':
   #   filename = "figs/wallsWB-with-sphere-path-%1.2f.dat" % ff
   # else:
   #   # input: "figs/walls.dat" % ()
-  #   # input: "figs/walls/wallsWB-path-*-pair-%1.2f-*.dat" %(ff)
+  #   # inp ut: "figs/walls/wallsWB-path-*-pair-%1.2f-*.dat" %(ff)
   #   filename = "figs/walls/wallsWB-path-%s-pair-%1.2f-%1.2f.dat" %(fun, ff, z0)
   # if (os.path.isfile(filename) == False):
   #   # Just use walls data if we do not have the MC (need to be careful!)
@@ -59,19 +59,13 @@ def read_triplet_path(ff,z0,fun):
 
 def read_triplet(ff, z0, fun):
   if fun == 'mc':
-    # input: "figs/mc/triplet/tripletMC-%3.1f-4.05.dat" % (ff)
-    filename = "figs/mc/triplet/tripletMC-%3.1f-04.05.dat" % (ff)
+    # input: "figs/mc/triplet/tripletMC-%3.1f-02.05-trimmed.dat" % (ff)
+    filename = "figs/mc/triplet/tripletMC-%3.1f-02.05-trimmed.dat" % (ff)
   # else:
-  #   # input: "figs/walls/wallsWB-*-pair-%1.2f-*.dat" %(ff)
[…]

### 2013-08-22 papers/pair-correlation/figs/trim-path-plots.py

@@ -5,31 +5,62 @@ import os
 # Trims the path data to be used for git storage
 
 for eta in [0.1, 0.2, 0.3, 0.4, 0.5]:
+  # pair
   filename = "figs/mc/wallsMC-pair-%03.1f-path.dat" %eta
   if (os.path.isfile(filename) == False):
     print("File does not exist: ", filename)
     continue
   data = loadtxt(filename)
   trimmed_data = data[800:1900,0:4]
-  savetxt("figs/mc/wallsMC-pair-%03.1f-path-trimmed.dat" %eta, trimmed_data, fmt='%g')
-  print("%s trimmed." %filename)
+  savetxt("figs/mc/wallsMC-pair-%03.1f-path-trimmed.dat" %eta, trimmed_data, fmt='%.3f')
+  print("Trimmed: %s." %filename)
 
-  # now triplet:
+  # pair 2d:
+  filename = "figs/mc/wallsMC-pair-%03.1f-0.05.dat" %eta
+  if (os.path.isfile(filename) == False):
+    print("File does not exist: ", filename)
+    continue
+  data = loadtxt(filename)
+  trimmed_data = data[:50, :70]
+  savetxt("figs/mc/wallsMC-pair-%03.1f-0.05-trimmed.dat" %eta, trimmed_data, fmt='%.3f')
+  print("Trimmed: %s." %filename)
+
+  # triplet @ contact:
   filename = "figs/mc/triplet/tripletMC-%03.1f-path.dat" %eta
   if (os.path.isfile(filename) == False):
     print("File does not exist: ", filename)
     continue
   data = loadtxt(filename)
   trimmed_data = data[800:1900, 0:4]
-  savetxt("figs/mc/triplet/tripletMC-%3.1f-path-trimmed.dat" %eta, trimmed_data, fmt='%g')
-  print("%s trimmed." %filename)
+  savetxt("figs/mc/triplet/tripletMC-%3.1f-path-trimmed.dat" %eta, trimmed_data, fmt='%.3f')
+  print("Trimmed: %s." %filename)
 
-  # now triplet 2:
+  # triplet @ contact 2d:
+  filename = "figs/mc/triplet/tripletMC-%03.1f-02.05.dat" %eta
+  if (os.path.isfile(filename) == False):
+    print("File does not exist: ", filename)
+    continue
+  data = loadtxt(filename)
+  trimmed_data = data[:50, :90]
[…]

### 2013-08-22 papers/pair-correlation/paper.tex
beautify triplet plots Tested everything on bingley.

--- 2013-08-22 papers/pair-correlation/paper.tex
@@
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    in contact.}\label{fig:triplet-distribution-3}
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+    shows the triplet distribution function, with white corresponding
+    to the contact value of the \emph{pair} distribution function.
+    The plot on the right shows the value of $g^{(3)}$ along the path
+    shown as a dotted lin on the plot to the
+    left.}\label{fig:triplet-distribution-3}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-4.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:triplet-distribution-4}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+    shows the triplet distribution function, with white corresponding
+    to the contact value of the \emph{pair} distribution function.
+    The plot on the right shows the value of $g^{(3)}$ along the path
+    shown as a dotted lin on the plot to the
+    left.}\label{fig:triplet-distribution-4}
 \end{figure}
@@
     1 - \sum_{i<j} \Theta(\sigma+\delta-r_{ij}) \\
-    + \mathop{\mathop{\sum_{i<j}}_{k<}}_{kl \ne ij}
+    + \frac{1}{2}\mathop{\mathop{\sum_{i<j}}_{k<l}}_{kl \ne ij}
     \Theta(\sigma+\delta-r_{ij}) \Theta(\sigma+\delta-r_{kl})
@@
   \Theta(r_{ij}-\sigma) \\
-  + \mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
+  + \frac{1}{2}\mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
   \iiiint \Theta(\sigma+\delta-r_{kl}) \Theta(\sigma+\delta-r_{mn}) \prod_{i<j}
@@
   Z(\sigma)-Z(\sigma+\delta) =
-  N^2\iiiint \Theta(\sigma+\delta-r_{12})
+  \frac{N^2}{2}\iiiint \Theta(\sigma+\delta-r_{12})
   \prod_{i<j} \Theta(r_{ij}-\sigma) \\
-  - \frac{N^3}{2}\iiiint
+  - \frac{N^4}{8}\iiiint
   \Theta(\sigma+\delta-r_{12})
@@
   \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)}\\
-  \shoveleft{\quad= nN \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
-  - \frac{n^2N}{2} \left(\int_{\sigma}^{\sigma+\delta} g(r)
+  \shoveleft{\quad= \frac{nN}{2} \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
+  - \frac{n^2N^2}{8} \left(\int_{\sigma}^{\sigma+\delta} g(r)
   d\rr\right)^2
-  \\
-  + \frac{n^3N}{6} \left(\int_{\sigma}^{\sigma+\delta} g(r)
-  d\rr\right)^3 + \cdots
+  + \cdots
   \label{eq:zdiff-g}

### 2013-08-22 papers/pair-correlation/paper.tex

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-4.pdf}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
   \caption{The triplet distribution function
@@
     left.}\label{fig:triplet-distribution-4}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-3}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-4}
 \end{figure}

### 2013-08-22 papers/pair-correlation/paper.tex
modified math in paper and merged some pulled python stuff Tested everything on node22.

--- 2013-08-22 papers/pair-correlation/paper.tex
@@
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-distribution-3}
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    in contact.}\label{fig:triplet-distribution-3}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/triplet-correlation-alt-4.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-distribution-4}
+  \caption{Pretty plot of the pair distribution function near a hard
+    wall.}\label{fig:triplet-distribution-4}
 \end{figure}
@@
   d\rr = 4\pi\sigma^2\left(g_\sigma\delta + \left(\tfrac12g_\sigma' +
-  \tfrac1{\sigma}\right) + \cdots \right)
+  \frac{g_\sigma}{\sigma}\right)\delta^2 + \cdots \right)
 \end{equation}
@@
   \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
-  4\pi\sigma^2 n g_\sigma \delta +
-  4\pi\sigma^2 n \left(\tfrac{g_\sigma'}{2} + \tfrac{2}{\sigma}\right)
+  2\pi\sigma^2 n N g_\sigma \delta +
+  \pi\sigma n N  \left(g_\sigma'\sigma + 2g\right)
   \delta^2 \\
-  - \tfrac{1}{2}\left(4\pi\sigma^2ng_\sigma\right)^2\delta^2 + \mathcal{O}(\delta^3)
+  - 2\pi^2\sigma^4n^2N^2\sigma^4g_\sigma^2\delta^2 + \mathcal{O}(\delta^3)
 \end{multline}
@@
 \begin{multline}
-  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
+  Z(\sigma)-Z(\sigma+\delta) =
   -\frac{\partial Z}{\partial \sigma} \delta
   -\frac12 \frac{\partial^2 Z}{\partial \sigma^2} \delta^2  + \mathcal{O}(\delta^3)
+\end{multline}
+So that
+\begin{multline}
+  \frac{\partial Z}{\partial \sigma} =
+  -2ZnN\pi\sigma^2g_\sigma\\
+  \frac{\partial^2 Z}{\partial \sigma^2} =
+  -2ZnN\pi\sigma\left(g_\sigma'\sigma + 2g\sigma \right)
+  +4Zn^2N^2\sigma^4g^2
 \end{multline}
@@
   g_\sigma' = \frac{\frac{\partial^2 F}{\partial
-      \sigma^2}}{4\pi\sigma^2nkT}
+      \sigma^2}}{2\pi\sigma^2nkT}
   - \frac{2g_\sigma}{\sigma}

### 2013-08-22 papers/pair-correlation/paper.tex
Completed derivation of chi_T in paper.

--- 2013-08-22 papers/pair-correlation/paper.tex
@@
 In addition, we know that
+\begin{align}
+  1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
+\end{align}
 \begin{equation}
   \begin{split}
-    1 + n\int\left[g(r)-1\right]d\rr = nkT\chi_T\\
-    \chi_T  = -\frac{1}{V}\left(\frac{\partial V}{\partial p}\right)_T
+    \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial p}\right)_T
     = - \left(V\left(\frac{\partial p}{\partial V}\right)_T\right)^{-1}\\
-    p = \frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\Longrightarrow \left(\frac{\partial p}{\partial V}\right)_T = \frac{dp}{d\eta}\frac{d\eta}{dV}\\
-    \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}\\
-    \left(\frac{\partial p}{\partial V}\right)_T = -\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
+    p &= nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\\
+    \left(\frac{\partial p}{\partial V}\right)_T &= \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}\\
+    n &= \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}\\
+    \eta &= \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}\\
+    \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
+    &  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
+    &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
+    &  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
+    &= -\frac{kT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
+    nkT\chi_T &= \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
@@
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-distribution-4}
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    in contact.}\label{fig:triplet-distribution-4}
 \end{figure}
@@
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-3}
+    separated by the diameter of one sphere.}\label{fig:triplet-distribution-3}
 \end{figure}
@@
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-4}
+    separated by the diameter of one sphere.}\label{fig:triplet-distribution-4}
 \end{figure}
@@
     1 - \sum_{i<j} \Theta(\sigma+\delta-r_{ij}) \\
-    + \frac{1}{2}\mathop{\mathop{\sum_{i<j}}_{k<l}}_{kl \ne ij}
+    + \mathop{\mathop{\sum_{i<j}}_{k<}}_{kl \ne ij}
     \Theta(\sigma+\delta-r_{ij}) \Theta(\sigma+\delta-r_{kl})
@@
   \Theta(r_{ij}-\sigma) \\
-  + \frac{1}{2}\mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
+  + \mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
   \iiiint \Theta(\sigma+\delta-r_{kl}) \Theta(\sigma+\delta-r_{mn}) \prod_{i<j}
@@
   Z(\sigma)-Z(\sigma+\delta) =
-  \frac{N^2}{2}\iiiint \Theta(\sigma+\delta-r_{12})
+  N^2\iiiint \Theta(\sigma+\delta-r_{12})
   \prod_{i<j} \Theta(r_{ij}-\sigma) \\
-  - \frac{N^4}{8}\iiiint
+  - \frac{N^3}{2}\iiiint
   \Theta(\sigma+\delta-r_{12})
@@
   \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)}\\
-  \shoveleft{\quad= \frac{nN}{2} \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
-  - \frac{n^2N^2}{8} \left(\int_{\sigma}^{\sigma+\delta} g(r)
+  \shoveleft{\quad= nN \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
+  - \frac{n^2N}{2} \left(\int_{\sigma}^{\sigma+\delta} g(r)
   d\rr\right)^2
-  + \cdots
+  \\
+  + \frac{n^3N}{6} \left(\int_{\sigma}^{\sigma+\delta} g(r)
+  d\rr\right)^3 + \cdots
   \label{eq:zdiff-g}

### 2013-08-23 papers/water-saft/paper.tex

@@
 
-We will apply this improved $g_\sigma^{HS}$ to a SAFT-based
-classical density functional for water~\cite{hughes2013classical}.
-This functional was constructed so as to reproduce in the homogeneous
-limit the 4-site optimal SAFT model for water developed by Clark
-\emph{et al.}~\cite{clark2006developing}.  The DFT of
+We will apply this improved $g_\sigma^{HS}$ to a SAFT-based classical
+density functional for water~\cite{hughes2013classical}.  This
+functional was constructed to reduce in the homogeneous limit
+to the 4-site optimal SAFT model for water developed by Clark \emph{et
+  al.}~\cite{clark2006developing}.  The DFT of
 Ref.~\citenum{hughes2013classical} uses the association free energy
@@
 \end{center}
-\caption{ Density profiles for a water around a single hydrophobic
-  rod of radius 0.2 nm. The solid red profile is the density using our new functional
-  and the dashed blue profile the result of our old functional.  For
-  scale, under the profiles is a cartoon of a string of hard
-  spheres touching in one dimension. The horizontal black dotted line
-  is the bulk density for water.}
+\caption{ Density profiles for a water around a single hydrophobic rod
+  of radius 0.1 nm. The solid red profile is from the functional
+  developed in this paper and the dashed blue profile is the result
+  from Ref. \cite{hughes2013classical}.  For scale, under the profiles
+  is a cartoon of a string of hard spheres touching in one
+  dimension. The horizontal black dotted line is the bulk density for
+  water and the vertical line on the left represents the wall of the rod.}
 \label{fig:density-single-rod}
@@
 water. Figure~\ref{fig:density-single-rod} shows the density profile
-of water near a rod with radius~0.1~\AA\fixme{Eric, please figure this
-  radius out}.  The density computed using the functional of this
+of water near a rod with radius~1~\AA.  The density computed using the functional of this
 paper is qualitatively similar to that from
@@
 \end{center}
-\caption{Broken hydrogen bonds per nanometer for various hydrophobic rods
-  immeresed in water.  The solid red line uses our new functional
-  while the dashed blue line uses the old one. For large enough rods,
-  the graph increases linearly for both functionals.}
+\caption{Broken hydrogen bonds per nanometer for hydrophobic rods
+  immeresed in water.  The solid red line uses the functional developd
+  in this paper while the dashed blue line uses the functional from
+  Ref. \cite{hughes2013classical}. For large enough rods, the graph
+  increases linearly for both functionals.}
 \label{fig:single-rod-broken-HB}
@@
 \end{center}
-\caption{Broken hydrogen bonds per for various hard spheres
-  immeresed in water.  The solid red line uses our new functional
-  while the dashed blue line uses the old one.}
+\caption{Broken hydrogen bonds for hard spheres immeresed in water.
+  The solid red line uses our the functional developd in this paper
+  while the dashed blue line is from
+  Ref. \cite{hughes2013classical}l.}
 \label{fig:spheres-broken-HB}

### 2013-08-27 papers/water-saft/paper.tex
Some changes to water paper Tested everything on node25.

@@
 hard-spheres with strong associative interactions at specific sites on
-the surfaces of the spheres~\cite{wertheim1984fluidsI,
+the surface of each sphere~\cite{wertheim1984fluidsI,
   wertheim1984fluidsII, wertheim1986fluidsIII, wertheim1986fluidsIV}.
@@
 $F_{\text{disp}}$ is the free energy contribution due to the
-square-well dispersion interaction.  The dispersion term of the free
-energy contains one empirical parameter, which is used to fit the
-surface tension of water near one atmosphere.  Finally,
-$F_{\text{assoc}}$ is the free energy contribution due to association,
-which is the term that we examine in this paper.
+square-well dispersion interaction; this term contains one empirical
+parameter, $\lscale$, which is used to fit the surface tension of water near one
+atmosphere.  Finally, $F_{\text{assoc}}$ is the free energy
+contribution due to association, which is the term that we examine in
+this paper.
 
@@
 fluid $g^\textit{SW}_\sigma$ is written as a perturbative correction
-to the hard-sphere fluid correlation function $g^\textit{HS}_\sigma$.
+to the hard-sphere radial distribution function
+$g^\textit{HS}_\sigma$, which comes from Yu and Wu
+\cite{yu2002fmt-dft-inhomogeneous-associating} for
+Ref.~\citenum{hughes2013classical} and we have previously
+derived \cite{schulte2012using} for this work.
 
@@
 \caption{ Density profiles for a water around a single hydrophobic rod
-  of radius 0.1 nm. The solid red profile is from the functional
+  of radius 0.1~nm. The solid red profile is from the functional
   developed in this paper and the dashed blue profile is the result
@@
   dimension. The horizontal black dotted line is the bulk density for
-  water and the vertical line on the left represents the wall of the rod.}
+  water and the vertical line on the left at 0.1~nm represents the
+  rod wall.}
 \label{fig:density-single-rod}
@@
   while the dashed blue line is from
-  Ref. \cite{hughes2013classical}l.}
+  Ref. \cite{hughes2013classical}.}
 \label{fig:spheres-broken-HB}

### 2013-08-31 papers/water-saft/paper.tex
fix typo noticed by paho Tested everything on darcy.
add a few citations in water-saft Tested everything on quipu.

@@
 of the bulk liquid at a given temperature and pressure
-\cite{jeanmairet2013molecular, lischner2010classical} ~\fixme{cite
+\cite{jeanmairet2013molecular, zhao2011molecular, zhao2011new, lischner2010classical} ~\fixme{cite
   Jeanmairet and similar}.  Using this approach, it is possible to
@@
 \caption{Broken hydrogen bonds per nanometer for hydrophobic rods
-  immeresed in water.  The solid red line uses the functional developd
+  immeresed in water.  The solid red line uses the functional developed
   in this paper while the dashed blue line uses the functional from

### 2013-08-31 papers/pair-correlation/paper.tex
fix bug in triplet-dft path, improve plots, and add text Tested everything on darcy.

@@
 
+Discuss the pair distribution function here, explaining plots
+  and discussing results and comparison.
+
 \subsection{Triplet distribution function}
-
-In Fig.~\ref{fig:triplet-distribution-3} and
-Fig.~\ref{fig:triplet-distribution-4} we can see the triplet
-distribution computed using Monte Carlo.
 
@@
     shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-distribution-3}
+    left.}\label{fig:triplet-contact-distribution-3}
 \end{figure}
@@
     shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-distribution-4}
+    left.}\label{fig:triplet-contact-distribution-4}
 \end{figure}
+
+Just as the radial distribution function of a homogeneous fluid may be
+computed from the density of an inhomogeneous one using Percus'
+test-particle trick, the triplet distribution function of a
+homogeneous system can be computed using an approximation of the pair
+distribution for an inhomogeneous fluid, such as we have developed.
+\begin{align}
+  g^{(3)}(\rr_1,\rr_2,\rr_3) &= \frac{n_{\textrm{TP}(\rr_1)}(\rr_2)
+  n_{\textrm{TP}(\rr_1)}(\rr_3)}{n^2}
+  g^{(2)}_{\textrm{TP}(\rr_1)}(\rr_2,\rr_3)
+\end{align}
+where $g^{(3)}(\rr_1,\rr_2,\rr_3)$ is the triplet distribution
+function of a homogeneous fluid with density $n$, and the
+$\textrm{TP}(\rr_1)$ subscript indicates quantities computed for the
+inhomogeneous density configuration in which one sphere is fixed at
+position $\rr_1$.  This method treats one of the three positions---the
+location of the test particle---differently from the others, which
+means that a poor approximation to the pair distribution function may
+beak the symmetry which is present in the true triplet distribution
+function.
+
+Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
+compare the triplet distribution function computed using our
+approximation to $g^{(2)}$ with results from Monte Carlo simulation.
+In each figure $\rr_1$ and $\rr_2$ are held fixed, while the third
+position $\rr_3$ is varied.  The test-particle position in each case
+is $\rr_1$, which is on the left-hand side of the figure.  In the
+color plot on the left, the first and second spheres are displayed in
+grey.  On the bottom half of each left-hand plot is the triplet
+correlation function as computed using the approximation presented in
+this work.  The top half displays the triplet correlation function
+from Monte Carlo.  On the right, we display
 
@@
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-3}
+    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
 \end{figure}
@@
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-gap-distribution-4}
+    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
 \end{figure}

### 2013-09-01 papers/pair-correlation/presentation/presentation.tex
Added a presentation folder in papers/pair-correlation. It contains a python program that generates a pdfs of a sample monte carlo simulation, and a latex file that puts them together. Run presentation/mc-circle-slow.py to get the frames and then compile presentation/presentation.tex for the finished animation. Tested everything on paho-deb.

\documentclass{beamer}
%\usepackage{beamerthemelined}
\usepackage{pstricks}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{graphicx}
\usepackage[draft]{animate}
%\setbeamertemplate{navigation symbols}{}
\beamertemplatenavigationsymbolsempty

\usetheme{Boadilla}
\usecolortheme{whale}
\setbeamertemplate{itemize items}[triangle]

\title{}
\author{}
\date{}

\begin{document}

\begin{frame}
  \frametitle{Monte Carlo Example}
  \begin{figure}[h]
    \centering
    \animategraphics[width=100mm,autoplay,final]{.6}{anim/mc-slow-}{000}{100}
  \end{figure}
\end{frame}

\end{document}

### 2013-09-02 papers/water-saft/paper.tex
add citations to remove fixme from water-saft paper Tested everything on bingley.
add edits with fixmes Tested everything on bingley.

--- 2013-08-27 papers/water-saft/paper.tex
@@
 \newcommand{\fixme}[1]{\textcolor{red}{[\emph{#1}]}}
+%\newcommand\hughesetal{Ref.~\citenum{hughes2013classical}}
+\newcommand\hughesetal{Hughes \emph{et al.}}
+
+\newcommand\hughesetalcite{Hughes \emph{et al.}~\cite{hughes2013classical}}
 
@@
 of the bulk liquid at a given temperature and pressure
-\cite{jeanmairet2013molecular, lischner2010classical} ~\fixme{cite
-  Jeanmairet and similar}.  Using this approach, it is possible to
-construct a functional that reproduces the exact second-order response
-function of the liquid under the fitted conditions.  However, this
-class of functional will be less accurate at other temperatures or
-pressures---and in the inhomogeneous scenarios in which solvation
-models are applied.  The second approach is to construct a functional by
-applying liquid-state theory to a model system, and then fit the model
-to experimental data such as the equation of
-state~\cite{hughes2013classical, clark2006developing, gloor2002saft,
-  gloor2004accurate, gloor2007prediction, Jaqaman2004, chuev2006,
-  fu2005vapor-liquid-dft,kiselev2006new, blas2001examination,
-  sundararaman2012computationally}.  Classical DFT is a natural
-framework for creating a more flexible theory of hydrophobicity that
-can readily describe interaction of water with arbitrary external
-potentials---such as potentials describing strong interactions with
-solutes or surfaces.
+\cite{jeanmairet2013molecular, zhao2011molecular, zhao2011new,
+  ramirez2005direct, ramirez2005density, levesque2012solvation,
+  levesque2012scalar, lischner2010classical}.  Using this approach, it
+is possible to construct a functional that reproduces the exact
+second-order response function of the liquid under the fitted
+conditions.  However, this class of functional will be less accurate
+at other temperatures or pressures---and in the inhomogeneous
+scenarios in which solvation models are applied.  The second approach
+is to construct a functional by applying liquid-state theory to a
+model system, and then fit the model to experimental data such as the
+equation of state~\cite{hughes2013classical, clark2006developing,
+  gloor2002saft, gloor2004accurate, gloor2007prediction, Jaqaman2004,
+  chuev2006, fu2005vapor-liquid-dft,kiselev2006new,
+  blas2001examination, sundararaman2012computationally}.
+%%  Classical
+%% DFT is a natural framework for creating a more flexible theory of
+%% hydrophobicity that can readily describe interaction of water with
+%% arbitrary external potentials---such as potentials describing strong
+%% interactions with solutes or surfaces.
 
@@
 of temperatures and pressures~\cite{muller2001molecular,
-  tan2008recent}.  The association term in SAFT is based on Wertheim's
-first-order thermodynamic perturbation theory, which models fluids as
-hard-spheres with strong associative interactions at specific sites on
-the surface of each sphere~\cite{wertheim1984fluidsI,
+  tan2008recent}.  The association term in SAFT uses Wertheim's
+first-order thermodynamic perturbation theory do describe an
+associating fluid as hard-spheres with strong associative interactions
+at specific sites on the surface of each sphere~\cite{wertheim1984fluidsI,
   wertheim1984fluidsII, wertheim1986fluidsIII, wertheim1986fluidsIV}.
@@
 $g_\sigma^\text{HS}$ in order to determine the extent of association.
-While this function is well-known for the homogeneous hard-sphere
-fluid, it is less well-known for inhomogeneous systems, such as occur
-at liquid interfaces.  In a recent paper, we examined this function at
-a variety of hard-sphere interfaces~\cite{schulte2012using}.  We
-tested the accuracy of existing approximations for the distribution
-function at contact~\cite{yu2002fmt-dft-inhomogeneous-associating,
+While this function is known for the homogeneous hard-sphere
+fluid, it must be approximated for inhomogeneous systems, such as occur
+at liquid interfaces.
+
+In a recent paper, we examined the pair distribution function at
+contact at a variety of hard-sphere
+interfaces~\cite{schulte2012using}.  We tested the accuracy of
+existing approximations for the distribution function at
+contact~\cite{yu2002fmt-dft-inhomogeneous-associating,
   gross2009density}, and derived a significantly improved
-approximation for the averaged distribution function at contact.
-
-We will apply this improved $g_\sigma^{HS}$ to a SAFT-based classical
-density functional for water~\cite{hughes2013classical}.  This
-functional was constructed to reduce in the homogeneous limit
-to the 4-site optimal SAFT model for water developed by Clark \emph{et
-  al.}~\cite{clark2006developing}.  The DFT of
-Ref.~\citenum{hughes2013classical} uses the association free energy
-functional of Yu and
-Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is based on a
-$g_\sigma^{HS}$ that we have since found to be
+approximation for the averaged distribution function at contact.  In
+this paper we apply this improved $g_\sigma^{HS}$ to the SAFT-based
+classical density functional for water developed by Hughes \emph{et
+  al.}~\cite{hughes2013classical}.  This functional was constructed to
+reduce in the homogeneous limit to the 4-site optimal SAFT model for
+water developed by Clark \emph{et al.}~\cite{clark2006developing}.
+The DFT of \hughesetal\ uses the association free energy functional of
+Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}, which is
+based on a $g_\sigma^{HS}$ that we have since found to be
 inaccurate~\cite{schulte2012using}.  In this paper, we will examine
 the result of using the improved functional for $g_\sigma^{HS}$
-developed in Ref.~\citenum{schulte2012using} to construct an
-association free energy functional.
+developed in \hughesetal\ to construct an association free energy
+functional.
 
@@
 
-The classical density functional for water from
-Reference~\citenum{hughes2013classical} consists of four terms:
+The classical density functional for water of \hughesetal\ consists of
+four terms:
 \begin{align}
@@
 
+\subsection{Dispersion}
+
+\fixme{Explain the dispersion term here.}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/surface-tension}
+\end{center}
+\caption{Comparison of Surface tension versus temperature for
+  theoretical and experimental data. The experimental data is taken
+  from NIST~\cite{nistwater}.  The length-scaling parameter $\lscale$
+  is fit so that the theoretical surface tension will match the
+  experimental surface tension near room temperature.}
+\label{fig:surface-tension}
+\end{figure}
+
+\subsection{Association}
+
 The association free energy for our four-site model has the form:
@@
     4 n(\rr) & \text{this work}\\
-    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
+    4 n_0(\rr) \zeta(\rr) & \text{\hughesetalcite}
   \end{cases}
@@
     \frac{4}{\pi\sigma^2} \int n(\rr')\delta(\sigma - |\rr-\rr'|) d\rr' & \text{this work}\\
-    4 n_0(\rr) \zeta(\rr) & \text{Ref.~\citenum{hughes2013classical}}
+    4 n_0(\rr) \zeta(\rr) & \text{\hughesetal}
   \end{cases} 
@@
 %This is equation 77 of gil-villegas-1997-SAFT-VR
-where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
+where $g^\textit{HS}_\sigma$ is the correlation function evaluated at
 contact for a hard-sphere fluid with a square-well dispersion
@@
 to the hard-sphere radial distribution function
-$g^\textit{HS}_\sigma$, which comes from Yu and Wu
-\cite{yu2002fmt-dft-inhomogeneous-associating} for
-Ref.~\citenum{hughes2013classical} and we have previously
-derived \cite{schulte2012using} for this work.
+$g^\textit{HS}_\sigma$.  The functional of \hughesetal\ uses the
+$g_\sigma^\textit{HS}$ from Yu and Wu
+\cite{yu2002fmt-dft-inhomogeneous-associating}.  In this work, we use
+the $g_\sigma^\textit{HS}$ derived by Schulte~\emph{et~al.}~\cite{schulte2012using}.
 
-\begin{figure}
-\begin{center}
-\includegraphics[width=3.5in]{figs/surface-tension}
-\end{center}
-\caption{Comparison of Surface tension versus temperature for
-  theoretical and experimental data. The experimental data is taken
-  from NIST~\cite{nistwater}.  The length-scaling parameter $\lscale$
-  is fit so that the theoretical surface tension will match the
-  experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
-
-As in Ref.~\citenum{hughes2013classical}, we use Clark's five
+As in \hughesetal, we use Clark's five
 empirical parameters, and fit the calculated surface tension to
@@
 association term, we find these agree when $\lscale$ is 0.454, which
-is an increase from the value of 0.353 found in
-Ref.~\citenum{hughes2013classical}.  In order to explore further the
+is an increase from the value of 0.353 found by
+\hughesetal.  In order to explore further the
 change made by the improved association term, we compared the new
-functional with that of Ref.~\citenum{hughes2013classical} for the two
+functional with that of \hughesetal\ for the two
 hydrophobic cases of hard rods and the hard spherical solutes.
@@
   developed in this paper and the dashed blue profile is the result
-  from Ref. \cite{hughes2013classical}.  For scale, under the profiles
+  from \hughesetal.  For scale, under the profiles
   is a cartoon of a string of hard spheres touching in one
@@
 paper is qualitatively similar to that from
-Ref.~\citenum{hughes2013classical}, with a comparable density at
+\hughesetal, with a comparable density at
 contact---consistent with having made only a moderate change in the
-free energy.  The first density peak near the surface is consistently
-higher than that from Ref.~\citenum{hughes2013classical}, and the peak
-is not smooth.  This results from the improved accuracy of the
-$g_\sigma^\textit{HS}$ from Ref.~\citenum{schulte2012using}, since beyond the
+free energy.  The first density peak near the surface is
+higher than that from \hughesetal, and the peak
+is has a kink at the top.  This reflects the improved accuracy of the
+$g_\sigma^\textit{HS}$ from \hughesetal, since beyond the
 first peak water molecules are unable to touch---or hydrogen bond
@@
 \caption{Broken hydrogen bonds per nanometer for hydrophobic rods
-  immeresed in water.  The solid red line uses the functional developd
+  immeresed in water.  The solid red line uses the functional developed
   in this paper while the dashed blue line uses the functional from
-  Ref. \cite{hughes2013classical}. For large enough rods, the graph
+  \hughesetal. For large enough rods, the graph
   increases linearly for both functionals.}
@@
 
+\fixme{Talk about Fig.~\ref{fig:single-rod-broken-HB}}
+
 A common test case for studying hydrophobic solutes in water is the
 hard-sphere solute.  Figure~\ref{fig:spheres-broken-HB} shows results
-for the number of broken hydrogen bonds caused by a hard-sphere soute,
+for the number of broken hydrogen bonds caused by a hard-sphere solute,
 as a function of the solute radius.  As in
@@
 about four times smaller than the number from the functional of
-Ref.~\citenum{hughes2013classical}.  For solutes smaller than
+\hughesetal.  For solutes smaller than
 3~\AA\ in radius, there is less than a tenth of a hydrogen bond
@@
   while the dashed blue line is from
-  Ref. \cite{hughes2013classical}.}
+  \hughesetal.}
 \label{fig:spheres-broken-HB}

### 2013-09-02 papers/water-saft/paper.tex
address fixmes Tested everything on bingley.

@@
 \subsection{Dispersion}
+The dispersion free energy includes the van der Waals attraction and
+any orientation-independent interactions. Following \hughesetal, we
+use a dispersion term based on the SAFT-VR
+approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
+parameters (taken from Clark~\emph{et al}\cite{clark2006developing}):
+an interaction energy $\epsilondisp$ and a length scale $\lambdadisp
+R$.
 
-\fixme{Explain the dispersion term here.}
+The SAFT-VR dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+\begin{align}
+  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
+\end{align}
+where $a_1$ and $a_2$ are the first two terms in a high-temperature
+perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
+the mean-field dispersion interaction. The second term, $a_2$, describes the
+effect of fluctuations resulting from compression of the fluid due
+to the dispersion interaction itself, and is approximated
+using the local compressibility approximation (LCA), which
+assumes the energy fluctuation is simply related to the
+compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
+
+The form of $a_1$ and $a_2$ for SAFT-VR is given in
+reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
+of the packing fraction.  In order to apply this form to an
+\emph{inhomogeneous} density distribution, we construct an effective local
+packing fraction for dispersion $\etadisp$, given by a Gaussian
+convolution of the density:
+\begin{align}
+  \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
+  \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
+      \lscale R)^2}}\right)d\xx'.\label{eq:packing-fraction}
+\end{align}
+This effective packing fraction is used throughout the dispersion
+functional, and represents a packing fraction averaged over the
+effective range of the dispersive interaction.
+Eq.~\ref{eq:packing-fraction} contains an additional empirical
+parameter $\lscale$ introduced by \hughesetal, which modifies the
+length scale over which the dispersion interaction is correlated.
 
@@
 
-As in \hughesetal, we use Clark's five
-empirical parameters, and fit the calculated surface tension to
-experimental surface tension at ambient conditions by tuning the
-parameter $\lscale$, which adjusts the length-scale of the average
-density used for the dispersion interaction.  With the improved
-association term, we find these agree when $\lscale$ is 0.454, which
-is an increase from the value of 0.353 found by
-\hughesetal.  In order to explore further the
-change made by the improved association term, we compared the new
-functional with that of \hughesetal\ for the two
-hydrophobic cases of hard rods and the hard spherical solutes.
+As in \hughesetal, we use Clark's five empirical parameters, and fit
+the calculated surface tension to experimental surface tension at
+ambient conditions by tuning the parameter $\lscale$, which adjusts
+the length-scale of the average density used for the dispersion
+interaction.  With the improved association term, we find these agree
+when $\lscale$ is 0.454, which is an increase from the value of 0.353
+found by \hughesetal.  In order to explore further the change made by
+the improved association term, we compared the new functional with
+that of \hughesetal\ for the two hydrophobic cases of hard rods and
+the hard spherical solutes.
 
@@
 
-\fixme{Talk about Fig.~\ref{fig:single-rod-broken-HB}}
+In addition to the density, we examine the number of hydrogen bonds
+which are broken due to the presence of a hard rod.  We define this
+quantity as
+\begin{align}
+  N_{\text{broken HB}} &= 2 \int (X(\rr) - X_{\text{bulk}})n_{\text{site}}(\rr) d\rr
+\end{align}
+where $X_{\text{bulk}} = 0.1?$ \fixme{get this number} is the fraction
+of unbonded association sites in the bulk.  The factor of 2 is chosen
+to account for the four association sites per molecule, and the fact
+that each hydrogen bond is twice---once for each of the molecules
+involved.  In Fig.~\ref{fig:single-rod-broken-HB} we show the number
+of hydrogen bonds broken by a hard rod per nanometer length, as
+predicted by the functional of \hughesetal\ (dashed line) and this
+work (solid line), as a function of the radius of the hard rod.  In
+each case in the limit of large rods, the number of broken bonds is
+proportional to the surface area, which is what one would expect.  At
+every radius, the functional of \hughesetal\ predicts approximately
+four times as many broken hydrogen bonds as the improved functional.
 
@@
 
+\end{document}
+
 \clearpage

### 2013-09-03 papers/water-saft/paper.tex

@@
   tan2008recent}.  The association term in SAFT uses Wertheim's
-first-order thermodynamic perturbation theory do describe an
+first-order thermodynamic perturbation theory to describe an
 associating fluid as hard-spheres with strong associative interactions
@@
 These association sites have an attractive interaction at contact, and
-rely on the hard-sphere radial distribution function at contact
+rely on the hard-sphere pair distribution function at contact
 $g_\sigma^\text{HS}$ in order to determine the extent of association.
@@
 In a recent paper, we examined the pair distribution function at
-contact at a variety of hard-sphere
-interfaces~\cite{schulte2012using}.  We tested the accuracy of
-existing approximations for the distribution function at
+contact in various inhomogeneous
+configurations~\cite{schulte2012using}.  We tested the accuracy of
+existing approximations for the pair distribution function at
 contact~\cite{yu2002fmt-dft-inhomogeneous-associating,
@@
 \subsection{Dispersion}
-The dispersion free energy includes the van der Waals attraction and
-any orientation-independent interactions. Following \hughesetal, we
-use a dispersion term based on the SAFT-VR
-approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters (taken from Clark~\emph{et al}\cite{clark2006developing}):
-an interaction energy $\epsilondisp$ and a length scale $\lambdadisp
-R$.
+The dispersion term in the free energy includes the van
+der Waals attraction and any orientation-independent
+interactions. Following \hughesetal, we use a dispersion term based on
+the SAFT-VR approach\cite{gil-villegas-1997-SAFT-VR}, which has two
+free parameters (taken from Clark~\emph{et
+  al}\cite{clark2006developing}): an interaction energy $\epsilondisp$
+and a length scale $\lambdadisp R$.
 
@@
 \begin{align}
-  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
+  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx
+  \label{eq:dispersion-term}
 \end{align}
@@
 \end{align}
-where the factor of four comes from the four hydrogen bond sites,
-$n_0(\rr)$ is the average density contacting point $\rr$, and
-$\zeta(\xx)$ is a dimensionless measure of the density inhomogeneity
-from Yu and Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  The
-functional $X(\rr)$ is the fraction of association sites \emph{not}
+where the factor of four comes from the four hydrogen bond sites, the
+fundamental measure $n_0(\rr)$ is the average density contacting point
+$\rr$, and $\zeta(\xx)$ is a dimensionless measure of the density
+inhomogeneity from Yu and
+Wu~\cite{yu2002fmt-dft-inhomogeneous-associating}.  The functional
+$X(\rr)$ is the fraction of association sites \emph{not}
 hydrogen-bonded, which is determined for our 4-site model by the
@@
 potential, and $a_1$ and $a_2$ are the two terms in the dispersion
-free energy.  The radial distribution function of the square-well
+free energy defined below Eq.~\ref{eq:dispersion-term}.  The radial distribution function of the square-well
 fluid $g^\textit{SW}_\sigma$ is written as a perturbative correction
@@
 water. Figure~\ref{fig:density-single-rod} shows the density profile
-of water near a rod with radius~1~\AA.  The density computed using the functional of this
-paper is qualitatively similar to that from
-\hughesetal, with a comparable density at
-contact---consistent with having made only a moderate change in the
-free energy.  The first density peak near the surface is
-higher than that from \hughesetal, and the peak
-is has a kink at the top.  This reflects the improved accuracy of the
-$g_\sigma^\textit{HS}$ from \hughesetal, since beyond the
-first peak water molecules are unable to touch---or hydrogen bond
-to---molecules at the surface of the hard rod.
+of water near a rod with radius~1~\AA.  The density computed using the
+functional of this paper is qualitatively similar to that from
+\hughesetal, with a comparable density at contact---consistent with
+having made only a moderate change in the free energy.  The first
+density peak near the surface is higher than that from \hughesetal,
+and the peak has a kink at the top.  This reflects the improved
+accuracy of the $g_\sigma^\textit{HS}$ from \hughesetal, since beyond
+the first peak water molecules are unable to touch---or hydrogen bond
+to---molecules at the surface of the hard rod. This is illustrated
+under the profiles in Figure~\ref{fig:density-single-rod} by a cartoon
+of adjacent hard spheres that are increasingly distant from the hard
+rod surface.
 
@@
 to account for the four association sites per molecule, and the fact
-that each hydrogen bond is twice---once for each of the molecules
-involved.  In Fig.~\ref{fig:single-rod-broken-HB} we show the number
-of hydrogen bonds broken by a hard rod per nanometer length, as
-predicted by the functional of \hughesetal\ (dashed line) and this
-work (solid line), as a function of the radius of the hard rod.  In
-each case in the limit of large rods, the number of broken bonds is
-proportional to the surface area, which is what one would expect.  At
-every radius, the functional of \hughesetal\ predicts approximately
-four times as many broken hydrogen bonds as the improved functional.
+that each broken hydrogen bond must be represented twice---once for
+each of the molecules involved.  In
+Fig.~\ref{fig:single-rod-broken-HB} we show the number of hydrogen
+bonds broken by a hard rod per nanometer length, as predicted by the
+functional of \hughesetal\ (dashed line) and this work (solid line),
+as a function of the radius of the hard rod.  In each case in the
+limit of large rods, the number of broken bonds is proportional to the
+surface area.  At every radius, the functional of
+\hughesetal\ predicts approximately four times as many broken hydrogen
+bonds as the improved functional.
 

### 2013-09-03 talks/colloquium/theory-of-liquids.tex
add some content to colloquium slides Tested everything on darcy.

\section{Theory of Liquids}
\subsection*{What is a liquid?}
\begin{frame}
  \frametitle{What is a liquid?}
  \framesubtitle{Elementary school definition}
  \begin{center}
    \includegraphics[width=7cm]{figs/solid-liquid-gas}
  \end{center}
  \begin{description}
  \item[Solid] Does not take the shape of its container
  \item[Gas] Takes the shape of its container and fills it
  \item[Liquid] Takes the shape of a container, but doesn't fill it
  \end{description}
\end{frame}

\begin{frame}
  \frametitle{What is a liquid?}
  %\framesubtitle{Phase diagram}
  \begin{center}
    \includegraphics[width=8.8cm]{figs/water-phase-diagram}
  \end{center}
  \vspace{-1em}
  \hfill \tiny figure from Wikipedia
\end{frame}

\begin{frame}
  \frametitle{What is a liquid?}
  \framesubtitle{Energy versus entropy}
  \begin{center}
    \includegraphics<1>[width=7cm]{figs/energy-solid}
    \includegraphics<2>[width=7cm]{figs/energy-solid-gas}
    \includegraphics<3->[width=7cm]{figs/energy-solid-gas-liquid}
  \end{center}
  \begin{description}
  \item[Solid] Energy dominates, entropy is a small perturbation
  \item<2->[Gas] Entropy dominates, energy is a small perturbation
  \item<3->[Liquid] Energy and entropy are balanced
  \end{description}
  \begin{block}{Perturbation theory}<4->
    \begin{itemize}
    \item<5-> An ``easy'' problem we know how to solve
    \item<6-> A ``hard'' correction that is small
    \item<7-> We construct a power series to solve combined problem
    \end{itemize}
  \end{block}
  %% \begin{enumerate}
  %% \item Solid and gas perturbation expansion
  %% \item No easy perturbation for liquid
  %% \item Ice floats
  %% \item Critical point
  %% \item Below the triple point
  %% \end{enumerate}
\end{frame}

%% \section[SAFT]{Understanding water using Statistical Associating
%%   Fluid Theory}

%% \subsection*{Simple liquids}
%% \begin{frame}
%%   \frametitle{Dense fluids}
%%   In order to use a perturbative approach, we need a ``solved''
%%   problem that is similar to a liquid.

%%   \begin{block}{Correlation function}<2->
%%     \vspace{-3em}
%%     \hfill \includegraphics[width=5cm, angle=270]{figs/correlation}\\
%%     %\vspace{-1em}
%%     \hfill\mycite{Soper}{J. Chem. Phys.}{2000}\\
%%     The correlation function $g(r)$ tells us how likely we are to find
%%     an atom at a given distance from any other atom.
%%   \end{block}
%% \end{frame}

\begin{frame}
  \frametitle{Dense fluids}
  In order to use a perturbative approach, we need a ``solved''
  problem that is similar to a liquid.
  \begin{block}{Hard-sphere fluid}<2->
    \begin{itemize}
    \item Captures the repulsive portion of the structure
    \item Still no energy!
    \item Not actually a liquid
    \end{itemize}
    \begin{center}
      \includegraphics[width=8cm]{figs/hard-spheres}
    \end{center}
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Simple liquids}
  \begin{itemize}
  \item Perturbative attraction added to hard spheres
  \item ``High T expansion''
    \[
    F_\text{disp} = \left(a_1(n) + a_2(n)\beta + \mathcal{O}(\beta^2)\right)n,
    \quad \quad \beta \equiv \frac{1}{k_BT}
    \]

    \vspace{1em} The temperature-independent $a_1$ term corresponds to
    a mean-field approximation, and incorporates correlations present
    in the hard-sphere fluid.
    \vspace{1em}

    The $a_2$ term incorporates---to first order---correlations due to
    the attraction itself.
  \end{itemize}
\end{frame}

%\subsection{Hydrogen bonding (Association)}
\begin{frame}
  \frametitle{Hydrogen bonds (Association)}
  \vspace{-2em}
  \hfill\includegraphics[width=7cm]{figs/hydrogen-bonds}\hspace{-3em}
  \vspace{-2em}
  \begin{itemize}
  \item Highly directional bonds
  \item Not a ``small'' interaction (around $4k_BT$)
  \item Small volume of interaction
  \item Wertheim's Thermodynamic Perturbation Theory\\ \hfill
    \mycite{Wertheim}{J. Stat. Phys.}{1984,1986}
  \item How often do hard spheres touch?
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Statistical Associating Fluid Theory (SAFT)}
  The above terms combined describe a well-established
  theory for homogeneous liquids such as water.
  \begin{center}
    \vspace{-1em}
    \includegraphics[height=6cm]{../../papers/water-saft/figs/pressure-with-isotherms-truncated}
  \end{center}
  \hfill Clark \emph{et al}, \emph{Mol. Phys.}
  (2006)
\end{frame}

\begin{frame}
  \frametitle{Interfaces:  Classical Density Functional Theory}
  In order to describe liquid \emph{interfaces} we need a theory that
  can handle \emph{inhomogeneous} configurations.
  \begin{block}{Classical Density Functional Theory (cDFT)}
    \[\Omega = \min_{n(\rr)} \left\{ F[n(\rr), T] + \int (V(\rr)-\mu) n(\rr)d\rr \right\}
    \]
    \begin{itemize}
    \item $\Omega$ is grand potential
    \item $n(\rr)$ is number density of atoms or molecules
    \item $V(\rr)$ is an arbitrary external potential energy
    \item $\mu$ is the chemical potential
    \item $F[n(\rr),T]$ is a functional that depends only on the
      liquid in question
    \end{itemize}
  \end{block}
\end{frame}

### 2013-09-03 papers/pair-correlation/paper.tex
nicer formatting in pair paper Tested everything on bingley.

@@
   \begin{split}
-    \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial p}\right)_T
-    = - \left(V\left(\frac{\partial p}{\partial V}\right)_T\right)^{-1}\\
-    p &= nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\\
-    \left(\frac{\partial p}{\partial V}\right)_T &= \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}\\
-    n &= \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}\\
-    \eta &= \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}\\
+    \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
+      p}\right)_T \\
+    &= - \left(V\left(\frac{\partial p}{\partial
+      V}\right)_T\right)^{-1}
+  \end{split}
+\end{equation}
+\begin{equation}
+    p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
+\end{equation}
+\begin{equation}
+    \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
+\end{equation}
+\begin{equation}
+    n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
+\end{equation}
+\begin{equation}
+    \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
+\end{equation}
+\begin{equation}
+  \begin{split}
     \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
-    &  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
+    &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
     &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
-    &  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
-    &= -\frac{kT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
-    nkT\chi_T &= \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
+    &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
+    &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
+  \end{split}
+\end{equation}
+\begin{equation}
+    nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
   %&= \eta \eta \cdots \fixme{fill in, double-check}
-  \end{split}
 \end{equation}

### 2013-09-04 papers/water-saft/paper.tex
Small changes to water paper Tested everything on node38.

@@
 of temperatures and pressures~\cite{muller2001molecular,
-  tan2008recent}.  The association term in SAFT uses Wertheim's
-first-order thermodynamic perturbation theory to describe an
-associating fluid as hard-spheres with strong associative interactions
-at specific sites on the surface of each sphere~\cite{wertheim1984fluidsI,
-  wertheim1984fluidsII, wertheim1986fluidsIII, wertheim1986fluidsIV}.
-These association sites have an attractive interaction at contact, and
-rely on the hard-sphere pair distribution function at contact
-$g_\sigma^\text{HS}$ in order to determine the extent of association.
-While this function is known for the homogeneous hard-sphere
-fluid, it must be approximated for inhomogeneous systems, such as occur
-at liquid interfaces.
+  tan2008recent}. The association contribution to the free energy uses
+Wertheim's first-order thermodynamic perturbation theory to describe
+an associating fluid as hard-spheres with strong associative
+interactions at specific sites on the surface of each
+sphere~\cite{wertheim1984fluidsI, wertheim1984fluidsII,
+  wertheim1986fluidsIII, wertheim1986fluidsIV}.  These association
+sites have an attractive interaction at contact, and rely on the
+hard-sphere pair distribution function at contact $g_\sigma^\text{HS}$
+in order to determine the extent of association.  While this function
+is known for the homogeneous hard-sphere fluid, it must be
+approximated for inhomogeneous systems, such as occur at liquid
+interfaces.
 
@@
 the improved association term, we compared the new functional with
-that of \hughesetal\ for the two hydrophobic cases of hard rods and
-the hard spherical solutes.
+that of \hughesetal\ for the two hydrophobic cases of the hard rod and
+the hard spherical solute.
 
@@
 \end{align}
-where $X_{\text{bulk}} = 0.1?$ \fixme{get this number} is the fraction
-of unbonded association sites in the bulk.  The factor of 2 is chosen
-to account for the four association sites per molecule, and the fact
-that each broken hydrogen bond must be represented twice---once for
-each of the molecules involved.  In
-Fig.~\ref{fig:single-rod-broken-HB} we show the number of hydrogen
-bonds broken by a hard rod per nanometer length, as predicted by the
-functional of \hughesetal\ (dashed line) and this work (solid line),
-as a function of the radius of the hard rod.  In each case in the
-limit of large rods, the number of broken bonds is proportional to the
-surface area.  At every radius, the functional of
-\hughesetal\ predicts approximately four times as many broken hydrogen
-bonds as the improved functional.
+where $X_{\text{bulk}} = 0.13$ is the fraction of unbonded
+association sites in the bulk.  The factor of 2 is chosen to account
+for the four association sites per molecule, and the fact that each
+broken hydrogen bond must be represented twice---once for each of the
+molecules involved.  In Fig.~\ref{fig:single-rod-broken-HB} we show
+the number of hydrogen bonds broken by a hard rod per nanometer
+length, as predicted by the functional of \hughesetal\ (dashed line)
+and this work (solid line), as a function of the radius of the hard
+rod.  In each case in the limit of large rods, the number of broken
+bonds is proportional to the surface area.  At every radius, the
+functional of \hughesetal\ predicts approximately four times as many
+broken hydrogen bonds as the improved functional.
 

### 2013-09-04 talks/colloquium/water.tex
update talk a bit Tested everything on bingley.

\section{Putting it all together: water}

\subsection{}

\begin{frame}
  \frametitle{Putting it all together}
  \begin{center}
    \includegraphics[height=3cm]{figs/HughesJessica}
    \includegraphics[height=3cm]{figs/KrebsEric}
  \end{center}
  \begin{block}{Statistical Associating Fluid Theory}
    \[
    F = F_{\textit{ideal gas}} + F_{\textit{hard spheres}} + F_{\textit{dispersion}} + F_{\textit{association}}
    \]
    We combine these terms to produce a classical density functional
    describing water.
  \end{block}
\end{frame}

\begin{frame}
  \frametitle{Homogeneous limit}
  \begin{center}
    \vspace{-1em}
    \includegraphics[height=6cm]{../../papers/water-saft/figs/pressure-with-isotherms-truncated}
   \end{center}
  Five SAFT empirical parameters taken from\\\hfill Clark \emph{et al.}, \emph{Mol. Phys.}
  (2006)
\end{frame}

\begin{frame}
  \frametitle{Homogeneous limit}
  \begin{center}
    \vspace{-1em}
    \includegraphics[height=7cm]{../../papers/water-saft/figs/surface-tension}
  \end{center}
  \vspace{-1.5em}
  One additional empirical parameter (in dispersion) used to fit
  surface tension.
\end{frame}

%% \begin{frame}
%%   \frametitle{Two hard rods}
%%   \begin{center}
%%     \vspace{-1em}
%%     \includegraphics[height=8cm]{figs/density-rods-in-water}
%%   \end{center}
%%   \vspace{-1.5em}
%% \end{frame}

\begin{frame}
  \frametitle{Hydration of a hard sphere}
  \begin{center}
    \vspace{-1em}
    %\includegraphics[height=7cm]{figs/sphere-energy-vs-diameter}
  \end{center}
  \vspace{-1.5em}
\end{frame}

\begin{frame}
  \frametitle{Conclusion \#2}
  \vspace{2em}
  \begin{block}{Hard sphere contact}
    \vspace{-6em}
    \hfill
    \includegraphics[height=2cm]{figs/walls-10}
    \includegraphics[height=2cm]{figs/walls-40}
    \begin{itemize}
    \item Found and tested an accurate functional for the correlation
      function at contact for hard spheres
    \item Reasonably efficient: scales like the hard sphere free
      energy
    \end{itemize}
  \end{block}
  \begin{block}{Water}
    \begin{itemize}
    \item Developed and tested a SAFT-based functional for water
    \end{itemize}
  \end{block}
  \begin{block}{Future work}
    \begin{columns}
      \begin{column}{0.7\columnwidth}
        \begin{itemize}
        \item Test dispersion versus MC
        \item Renormalization group for critical point
        \item Hard polyhedra for directional correlation
        \end{itemize}
      \end{column}
      \begin{column}{0.5\columnwidth}
        \begin{itemize}
        \item Dielectric interactions
        \item Softening the spheres
        \item Coupling with solute
        \end{itemize}
      \end{column}
    \end{columns}
  \end{block}
\end{frame}

### 2013-09-04 papers/water-saft/paper.tex
Added abstract, changed title for water paper. Tested on node32. Took 0:21.72.

@@
 \begin{document}
-\title{A Classical Density-Functional Theory for Describing Water Interfaces}
+\title{Improved Association in a Classical Density-Functional Theory
+  for Water}
 
@@
 \begin{abstract}
-We develop an improved version of our old functional.
+We present a modification to the classical density-functional theory
+(DFT) for water by Hughes \emph{et al.} which uses Statistical
+Associating Fluid Theory(SAFT-VR) for attractive interactions. We
+modify the association contribution to the free energy to use an
+improved radial distribution function $g_\sigma^{HS}$. We study the
+difference between the two functionals density profile, number
+of broken hydrogen bonds for hard rods in water, and the number of
+broked hydrogen bonds for a hard sphere in water.
 \end{abstract}

### 2013-09-05 papers/water-saft/paper.tex
improve abstract and density-functional -> density functional Tested on bingley. Took 0:35.01.

--- 2013-09-04 papers/water-saft/paper.tex
@@
-\documentclass[twocolumn,amsmath,amssymb,prl]{revtex4-1}
+\documentclass[twocolumn,amsmath,amssymb,jcp]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 \begin{document}
-\title{A Classical Density-Functional Theory for Describing Water Interfaces}
+\title{Improved association in a classical density functional theory
+  for water}
 
@@
 \begin{abstract}
-We develop an improved version of our old functional.
+We present a modification to our recently published classical
+density functional theory for water.  Both the modified and original
+functionals reduces to the statistical associating fluid theory with
+variable attractive range (SAFT-VR) equation of state in the
+homogeneous limit.
+%
+We have recently developed and tested a functional for the averaged
+radial distribution function at contact of the hard-sphere fluid,
+which is dramatically more accurate at interfaces than earlier
+approximations.
+%
+We now incorporate this improved functional into the association term
+of our free energy functional for water, which improves the
+description of hydrogen bonding.
+%
+We examine the effect of this improvement by studying the cases of a
+hard hydrophobic rods in water.  Our results show that there is a
+moderate change in the density profile near a hard rot, and the number
+of broken hydrogen bonds is decreased by approximately a factor of
+four.  Examination of a hard sphere solute reveals a similarly large
+chage in the number of broken hydrogen bonds.
 \end{abstract}
@@
 models for water have built on the approach of classical
-density-functional theory (DFT)~\cite{jeanmairet2013molecular,
+density functional theory (DFT)~\cite{jeanmairet2013molecular,
   zhao2011molecular, zhao2011new, ramirez2005direct,

### 2013-09-05 papers/pair-correlation/paper.tex
Fixed a couple typos in gsigma' derivation. Tested everything on paho-deb.

@@
 
-\section{Analytic approximations}
+\section{Approximations to the pair distribution function}
 There are few analytic approximations for the inhomogeneous pair
@@
 to inhomogeneous scenarios.  These works differ in what density they
-use to evaluate $g(r)$.
+use when evaluating the radial distribution function $g(r,\eta)$.
 
@@
 where $g(r,\eta)$ is the radial distribution function as a function of
-radius and packing fraction $\eta$~\cite{gloor2007prediction,
-  gross2009density}.  These works did not however use this approach,
-which fails dramatically at higher densities, as the packing fraction
-found from the mean density exceeds unity.
+radius and packing fraction $\eta$, and $\sigma$ is the hard-sphere
+diameter~\cite{gloor2007prediction, gross2009density}.  These works
+did not however use this approach, which fails dramatically at higher
+densities, as the packing fraction found from the mean density exceeds
+unity.
 
@@
 is computationally awkward, because it relies on the midpoint
-$\tfrac12(\rr_1+\rr_2)$.  For this reason Sokolowski and Fischer
-modified this approach to use averages centered on the two points
-$\rr_1$ and $\rr_2$:
+$\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
+Methfessel predicts the pair distribution function only at contact,
+when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
+diameter.
+
+For these reasons Sokolowski and Fischer modified this approach to use
+averages centered on the two points $\rr_1$ and $\rr_2$:
 \begin{align}
@@
 desired, cannot be written as a single-site convolution, since the
-convolution kernel depends on both points.
+convolution kernel depends on both points.  This limits the accuracy.
 
-\section{What is our new idea?}
+\section{Our approach}
 We approximate the pair correlation with the average of two radial
@@
 points, that are themselves functions of our published pair
-correlation at contact $g_{\sigma}$ evaluated at the two points.
-
+correlation at contact $g_{\sigma}(\rr)$ evaluated at the two
+points~\cite{schulte2012using}.
 \begin{align}
@@
 distributions (used to approximate pair correlation) were dependent
-upon independent variables $\eta$, the density, wether at single
+upon independent variables $\eta$, the density, whether at single
 points or averaged over sourrounding volumes.  We can achieve more
@@
 \end{figure}
-\section{What does it look like (graphs that show the function itself)?}
 \subsection{Separable fit to the radial distribution function}
@@
 
+\section{Results}
 \subsection{Pair distribution function}
-%% Figure~\ref{fig:pair-distribution} shows...
-%% \begin{figure}
-%%   %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-0.pdf}
-%%   \includegraphics[width=6cm]{figs/pair-correlation-2-395-5.pdf}
-%%   %%\includegraphics[width=6cm]{figs/pair-correlation-2-395-10.pdf}
-%%   \caption{Plot of the pair distribution function near a hard wall.
-%%     In the interest of space and memeory, we should look at the function and decided which data to add?}\label{fig:pair-distribution}
-%% \end{figure}
+
+\newcommand\paircaption[1]{ The pair distribution function near a hard
+  wall, with a packing fraction of #1, with $\rr_1$ in contact with
+  the hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$
+  as $\rr_2$ varies.  The top half of this figure shows the result of
+  Monte Carlo simulation, while the bottom half shows the result of
+  the appproximation developed in this work.  On the right is a plot
+  of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the figure to
+  the left.  This plot compares the Monte Carlo results (black) with
+  those of this work (color?), Sokolowski and Fischer
+  (color?)~\cite{sokolowski1992role}, and Fischer and Methfessel
+  (color?)~\cite{fischer1980born}.  The latter is only plotted at
+  contact, where it is defined.  }
 
@@
   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:pair-distribution-1}
+  \caption{ \paircaption{0.1}
+  }\label{fig:pair-distribution-1}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:pair-distribution-2}
+  \caption{\paircaption{0.2}}\label{fig:pair-distribution-2}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:pair-distribution-3}
+  \caption{\paircaption{0.3}}\label{fig:pair-distribution-3}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
-  \caption{Pretty plot of the pair distribution function near a hard
-    wall.}\label{fig:pair-distribution-4}
+  \caption{\paircaption{0.4}}\label{fig:pair-distribution-4}
 \end{figure}
@@
   2\pi\sigma^2 n N g_\sigma \delta +
-  \pi\sigma n N  \left(g_\sigma'\sigma + 2g\right)
+  \pi\sigma n N  \left(g_\sigma'\sigma + 2g_\sigma\right)
   \delta^2 \\
-  - 2\pi^2\sigma^4n^2N^2\sigma^4g_\sigma^2\delta^2 + \mathcal{O}(\delta^3)
+  - 2\pi^2\sigma^4n^2N^2g_\sigma^2\delta^2 + \mathcal{O}(\delta^3)
 \end{multline}
@@
   \frac{\partial^2 Z}{\partial \sigma^2} =
-  -2ZnN\pi\sigma\left(g_\sigma'\sigma + 2g\sigma \right)
-  +4Zn^2N^2\sigma^4g^2
+  -2ZnN\pi\sigma\left(g_\sigma'\sigma + 2g_\sigma \right)
+  +4Zn^2N^2\pi^2\sigma^4{g_\sigma}^2
 \end{multline}
@@
   g_\sigma' = \frac{\frac{\partial^2 F}{\partial
-      \sigma^2}}{2\pi\sigma^2nkT}
+      \sigma^2}}{2\pi\sigma^2nNkT}
   - \frac{2g_\sigma}{\sigma}

### 2013-09-06 papers/water-saft/paper.tex
Abstract changes to water paper Tested on node33. Took 0:22.52.

--- 2013-09-04 papers/water-saft/paper.tex
@@
-\documentclass[twocolumn,amsmath,amssymb,prl]{revtex4-1}
+\documentclass[preprint, endfloats,amsmath,amssymb,jcp]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 \begin{abstract}
-We develop an improved version of our old functional.
+We present a modification to our recently published SAFT-based classical
+density functional theory for water.
+%
+We have recently developed and tested a functional for the averaged
+radial distribution function at contact of the hard-sphere fluid
+that is dramatically more accurate at interfaces than earlier
+approximations.
+%
+We now incorporate this improved functional into the association term
+of our free energy functional for water, improving its description of
+hydrogen bonding.
+%
+We examine the effect of this improvement by studying two hard solutes: a
+hard hydrophobic rod and a hard sphere.
+%
+The improved functional leads to a moderate change in the density
+profile and a large decrease in the number of hydrogen bonds broken in
+the vicinity of the solutes.
 \end{abstract}

### 2013-09-06 talks/colloquium/theory-of-liquids.tex
update talk a bit Tested everything on bingley.
more edits on talk for next week Tested on darcy. Took 0:19.24.

@@
     \[
-    F_\text{disp} = \left(a_1(n) + a_2(n)\beta + \mathcal{O}(\beta^2)\right)n,
+    F_\text{disp} = N\left(a_1(n) + a_2(n)\beta + \mathcal{O}(\beta^2)\right),
     \quad \quad \beta \equiv \frac{1}{k_BT}
@@
     \mycite{Wertheim}{J. Stat. Phys.}{1984,1986}
-  \item How often do hard spheres touch?
   \end{itemize}
@@
 \begin{frame}
-  \frametitle{Statistical Associating Fluid Theory (SAFT)}
-  The above terms combined describe a well-established
-  theory for homogeneous liquids such as water.
+  \frametitle{Putting it all together}
+  All these terms combined describe SAFT, a well-established theory
+  for \emph{homogeneous} liquids such as water.
   \begin{center}
@@
 \end{frame}
+
+\begin{frame}
+  \frametitle{Testing the theory:  Monte Carlo simulation}
+  \vspace{-0.8em}
+  \begin{center}
+    \animategraphics[height=20em]{.6}{anim/mc-slow-}{000}{30}\\
+    \vspace{-0.5em}
+    hard circles between two walls
+  \end{center}
+\end{frame}
+
+\begin{frame}
+  \frametitle{Testing the theory:  Monte Carlo simulation for $n(\rr)$}
+  \vspace{-0.8em}
+  \begin{center}
+    \animategraphics[height=20em,autoplay]{.6}{anim/mc-density-}{000}{30}\\
+    \vspace{-0.5em}
+    hard circles between two walls
+  \end{center}
+\end{frame}

### 2013-09-06 talks/colloquium/figs/plot-walls-mc.py
make talk figures transparent Tested on darcy. Took 0:13.34.

@@ -73,7 +73,7 @@ mcdata40 = numpy.loadtxt('../../papers/contact/figs/mc-walls-20-817.dat')
 dftdata40 = numpy.loadtxt('../../papers/contact/figs/wallsWB-0.40.dat')
 
 plotit(dftdata10, mcdata10)
-pylab.savefig('figs/walls-mc-10.pdf')
+pylab.savefig('figs/walls-mc-10.pdf', transparent=True)
 
 plotit(dftdata40, mcdata40)
-pylab.savefig('figs/walls-mc-40.pdf')
+pylab.savefig('figs/walls-mc-40.pdf', transparent=True)

### 2013-09-07 talks/colloquium/figs/plot-pair-mc.py
clean up brackets for pair path plots in talk Tested on darcy. Took 0:20.65.

@@ -84,10 +84,63 @@ zplot.set_xlim(-4, 6)
 zplot.set_xticks([])
 #xplot.set_ylim(0)
 
-figtext(.613, .04, r"$\underbrace{\hspace{9em}}$", horizontalalignment='center')
-figtext(.613, .01, r"$x$", horizontalalignment='center')
-figtext(.796, .04, r"$\underbrace{\hspace{13.1em}}$", horizontalalignment='center')
-figtext(.796, .01, r"$z$", horizontalalignment='center')
+# The following is a tedious way to make a bracket
+def bracket(ax, x1, x2, y, em, text):
+  xmiddle = 0.5*(x1+x2)
+  xleft = 0.5*(x1+xmiddle)
+  xight = 0.5*(xmiddle+x2)
+  bracket_height = 0.6*em
+  ax.text(xmiddle, y-(em+bracket_height), text, horizontalalignment='center',
+          transform=ax.transAxes)
+  shrink = 0
+  rad = 2
+  ax.annotate("",
+              xy=(x1,y), xycoords='axes fraction',
+              xytext=(xleft,y-0.5*bracket_height), textcoords='axes fraction',
+              arrowprops=dict(arrowstyle="-",
+                               color="0.0",
+                               shrinkA=0, shrinkB=shrink,
+                               patchA=None,
+                               patchB=None,
+                               connectionstyle="angle,angleA=0,angleB=-45,rad=%g" % rad,
+                               ),
+               )
+  ax.annotate("",
+              xy=(xleft,y-0.5*bracket_height), xycoords='axes fraction',
+              xytext=(xmiddle,y-bracket_height), textcoords='axes fraction',
+              arrowprops=dict(arrowstyle="-",
+                               color="0.0",
+                               shrinkA=shrink, shrinkB=0,
+                               patchA=None,
+                               patchB=None,
+                               connectionstyle="angle,angleA=-45,angleB=0,rad=%g" % rad,
+                               ),
+               )
+  ax.annotate("",
+              xy=(x2,y), xycoords='axes fraction',
+              xytext=(xight,y-0.5*bracket_height), textcoords='axes fraction',
+              arrowprops=dict(arrowstyle="-",
+                               color="0.0",
+                               shrinkA=0, shrinkB=shrink,
+                               patchA=None,
+                               patchB=None,
+                               connectionstyle="angle,angleA=180,angleB=45,rad=%g" % rad,
[…]

### 2013-09-09 talks/colloquium/theory-of-liquids.tex
make talk figures transparent Tested on darcy. Took 0:13.34.
various fixes for the talk (and disable parallel commit tests) Tested on bingley. Took 0:24.18.
more work on talk Tested on bingley. Took 0:49.32.

@@
-\section{Theory of Liquids}
+\section{Theory of Liquids (Statistical Associating Fluid Theory)}
 \subsection*{What is a liquid?}
@@
   \item[Solid] Does not take the shape of its container
+  \item[Liquid] Takes the shape of a container, but doesn't fill it
   \item[Gas] Takes the shape of its container and fills it
-  \item[Liquid] Takes the shape of a container, but doesn't fill it
   \end{description}
@@
   \begin{center}
-    \includegraphics[width=8.8cm]{figs/water-phase-diagram}
+    \includegraphics[height=7.8cm]{figs/water-phase-diagram}
   \end{center}
-  \vspace{-1em}
-  \hfill \tiny figure from Wikipedia
 \end{frame}
@@
   \item[Solid] Energy dominates, entropy is a small perturbation
+  \item<3->[Liquid] Energy and entropy are balanced
   \item<2->[Gas] Entropy dominates, energy is a small perturbation
-  \item<3->[Liquid] Energy and entropy are balanced
   \end{description}
@@
     \begin{itemize}
-    \item<5-> An ``easy'' problem we know how to solve
-    \item<6-> A ``hard'' correction that is small
-    \item<7-> We construct a power series to solve combined problem
+    \item An ``easy'' problem we know how to solve
+    \item A ``hard'' correction that is small
+    \item We construct a power series to solve combined problem
     \end{itemize}
@@
 \begin{frame}
-  \frametitle{Dense fluids}
+  \frametitle{Dense fluids (Hard spheres)}
   In order to use a perturbative approach, we need a ``solved''
   problem that is similar to a liquid.
-  \begin{block}{Hard-sphere fluid}<2->
+  \begin{block}{Hard-sphere fluid}
     \begin{itemize}
     \item Captures the repulsive portion of the structure
-    \item Still no energy!
+    \item All entropy, no energy!
     \item Not actually a liquid
@@
 \begin{frame}
-  \frametitle{Simple liquids}
+  \frametitle{Simple liquids (Dispersion)}
   \begin{itemize}
@@
 \begin{frame}
-  \frametitle{Putting it all together}
+  \frametitle{Putting it all together:  SAFT for water}
   All these terms combined describe SAFT, a well-established theory
-  for \emph{homogeneous} liquids such as water.
+  for \emph{homogeneous} liquids.
   \begin{center}
@@
     \begin{itemize}
-    \item $\Omega$ is grand potential
-    \item $n(\rr)$ is number density of atoms or molecules
-    \item $V(\rr)$ is an arbitrary external potential energy
-    \item $\mu$ is the chemical potential
-    \item $F[n(\rr),T]$ is a functional that depends only on the
-      liquid in question
+    %% \item $\Omega$ is grand potential
+    %% \item $n(\rr)$ is number density of atoms or molecules
+    %% \item $V(\rr)$ is an arbitrary external potential energy
+    %% \item $\mu$ is the chemical potential
+    \item minimize the grand potential $\Omega$
+    \item $F[n(\rr),T]$ is a universal functional for a given fluid
     \end{itemize}
@@
     \animategraphics[height=20em]{.6}{anim/mc-slow-}{000}{30}\\
-    \vspace{-0.5em}
+    \vspace{-2.0em}
     hard circles between two walls
@@
     \animategraphics[height=20em,autoplay]{.6}{anim/mc-density-}{000}{30}\\
-    \vspace{-0.5em}
+    \vspace{-2.0em}
     hard circles between two walls

### 2013-09-09 talks/colloquium/figs/plot-walls.py

@@ -8,6 +8,9 @@ matplotlib.use('Agg')
 
 import pylab, numpy, sys
 
+xmax = 2.5
+xmin = -0.4
+
 def plotit(dftdata, mcdata):
     dft_len = len(dftdata[:,0])
     dft_dr = dftdata[2,0] - dftdata[1,0]
@@ -22,16 +25,16 @@ def plotit(dftdata, mcdata):
     nAmc = mcdata[:,11]
     n0mc = mcdata[:,10]
 
-    pylab.figure(figsize=(6, 4))
+    pylab.figure(figsize=(6, 6))
     pylab.subplots_adjust(hspace=0.001)
 
     n_plt = pylab.subplot(3,1,3)
     n_plt.plot(mcdata[:,0]/2+mcoffset,mcdata[:,1]*4*numpy.pi/3,"b-",label='$n$ Monte Carlo')
     n_plt.plot(dftdata[:,0]/2+offset,dftdata[:,1]*4*numpy.pi/3,"b--",label='$n$ DFT')
-    n_plt.legend(loc='best', ncol=2).draw_frame(False) #.get_frame().set_alpha(0.5)
+    n_plt.legend(loc='best', ncol=1).draw_frame(False) #.get_frame().set_alpha(0.5)
     n_plt.yaxis.set_major_locator(pylab.MaxNLocator(6,steps=[1,5,10],prune='upper'))
     pylab.ylim(ymin=0)
-    pylab.xlim(-1/2,9/2)
+    pylab.xlim(xmin, xmax)
     pylab.xlabel("$z/\sigma$")
     pylab.ylabel("$n(\mathbf{r})$")
     n_plt.axvline(x=0, color='k', linestyle=':')
@@ -53,11 +56,11 @@ def plotit(dftdata, mcdata):
     A_plt.plot(dftdata[start_here:stop_here,0]/2+offset,dftdata[start_here:stop_here,5],"ro",markevery=me,label="$g_\sigma^A$ this work")
     A_plt.plot(dftdata[:,0]/2+offset,dftdata[:,7],"rx",markevery=me,label="Gross",
                markerfacecolor='none',markeredgecolor='red', markeredgewidth=1)
-    A_plt.legend(loc='best', ncol=2).draw_frame(False) #.get_frame().set_alpha(0.5)
+    A_plt.legend(loc='best', ncol=1).draw_frame(False) #.get_frame().set_alpha(0.5)
[…]

### 2013-09-09 talks/colloquium/water.tex
add dadz plot to talk Tested on darcy. Took 0:11.35.
fix conclusion Tested on bingley. Took 0:20.06.

@@
   \frametitle{Putting it all together}
-  \begin{center}
-    \includegraphics[height=3cm]{figs/HughesJessica}
-    \includegraphics[height=3cm]{figs/KrebsEric}
-  \end{center}
   \begin{block}{Statistical Associating Fluid Theory}
@@
   \end{block}
+  \begin{center}
+    \includegraphics[height=3cm]{figs/HughesJessica}
+    \includegraphics[height=3cm]{figs/KrebsEric}
+  \end{center}
 \end{frame}
@@
     \vspace{-1em}
-    %\includegraphics[height=7cm]{figs/sphere-energy-vs-diameter}
+    \includegraphics[height=7cm]{figs/sphere-energy}
   \end{center}
@@
 \begin{frame}
-  \frametitle{Conclusion \#2}
+  \frametitle{\conclude}
   \vspace{2em}
-  \begin{block}{Hard sphere contact}
-    \vspace{-6em}
-    \hfill
-    \includegraphics[height=2cm]{figs/walls-10}
-    \includegraphics[height=2cm]{figs/walls-40}
+  \begin{block}{Hard sphere distribution function}
     \begin{itemize}
-    \item Found and tested an accurate functional for the correlation
-      function at contact for hard spheres
-    \item Reasonably efficient: scales like the hard sphere free
+    \item Developed and tested an accurate functional for contact
+    \item Developed and tested functional pair distribution
+    \item Reasonably efficient: scale like the hard sphere free
       energy
@@
     \begin{columns}
-      \begin{column}{0.7\columnwidth}
+      \begin{column}{0.6\columnwidth}
         \begin{itemize}
-        \item Test dispersion versus MC
         \item Renormalization group for critical point
-        \item Hard polyhedra for directional correlation
+        \item Hard polyhedra for orientation
         \end{itemize}

### 2013-09-10 talks/polyhedra/slides.tex
Added polyhedra talk Tested on paho-deb. Took 1:58.57.

\documentclass[handout]{beamer}
%\usepackage{beamerthemelined}
\usepackage{pstricks}
\usepackage{amsfonts,amssymb,amsmath,amsthm}
\usepackage{graphicx}
\usepackage[draft]{animate}
\usepackage{wallpaper}
%\setbeamertemplate{navigation symbols}{}
\beamertemplatenavigationsymbolsempty

\usetheme{Boadilla}
\usecolortheme{whale}
\setbeamertemplate{itemize items}[triangle]

\title{Hard Polyhedra Fluids}
\author{Paho Lurie-Gregg}
\date{}

\newcommand{\f}[2]{\dfrac{#1}{#2}}
\newcommand{\p}[1]{\left(#1\right)}
\renewcommand{\t}[1]{\text{#1}}
\newcommand{\abs}[1]{\left|#1\right|}

\begin{document}
{
\usebackgroundtemplate{\includegraphics[width=\paperwidth]{figs/background}}
\begin{frame}
  \maketitle
 \begin{center}
   Dr. David Roundy, Department of Physics
 \end{center}

\end{frame}
}
\begin{frame}
  \frametitle{Water}
  \begin{itemize}
  \item<1-> Water is arguably the most important fluid in existence, and yet it has some very strange properties
  \item<2-> There currently exist no accurate continuum models for water at melting
  \end{itemize}
  \begin{figure}[h]
    \includegraphics[width=80mm]{figs/ice-water.png}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Background}
  Why Hard Polyhedra?
  \begin{itemize}
  \item<1-> Hard sphere fluids have been studied extensively
  \item<2-> Hard polyhedra have little known theory
  \item<3-> We expect to be able to match the crystal structure of ice with truncated tetrahedra
  \item<4-> Other polyhedra may show interesting characteristics
  \end{itemize}
  \begin{figure}[h]
    \raggedleft
    \includegraphics[width=60mm]{figs/Cryst_struct_ice}
  \end{figure}
\end{frame}
\begin{frame}
  \frametitle{Classical Density Functional Theory}
  \begin{itemize}
  \item<1-> A functional is simply a function of a function
  \item<2-> Classical DFT provides a continuum approach for fluids
  \item<3-> Helmholtz Free Energy is minimized at equilibrium
    \begin{align*}
      F = U - TS
    \end{align*}
  \end{itemize}
\end{frame}

\begin{frame}
  \frametitle{Monte Carlo Example}
  \begin{figure}[h]
    \centering
    \animategraphics[width=100mm,autoplay]{.6}{anim/mc-slow-}{000}{200}
  \end{figure}

\end{frame}

\begin{frame}
  \frametitle{Monte Carlo Example}
  \begin{figure}[h]
    \centering
    \animategraphics[width=120mm,autoplay]{3}{anim/mc100-0.50-}{000}{999}
  \end{figure}

\end{frame}
\end{document}

### 2013-09-18 papers/pair-correlation/figs/plot-path-triplet-contact.py
make pair-correlation figures more consistent and readable in b/w Tested on bingley. Took 2:09.61.

@@ -9,15 +9,15 @@ import scipy.ndimage
 import os.path
 import math
 import matplotlib.patheffects
+import bracket # our handy bracket function
+import styles # our preferred line styles
 from matplotlib import rc
 rc('text', usetex=True)
 
 from matplotlib.colors import NoNorm
 
 # these are the things to set
-colors = ['k', 'b', 'g', 'r']
 plots = ['mc', 'this-work', 'fischer', 'sokolowski']#, 'sphere-dft'] # , 'gloor'
-titles = ['Monte Carlo', 'this work', 'Fischer et al.', 'Sokolowski et al.'] # , 'gloor'
 dx = 0.1
 sigma = 2.0
 contact_delta = 0.1
@@ -89,31 +89,26 @@ xplot.set_xlim(xlow, xhigh)
 zplot.set_xlim(rpath/2 - xlow + xmin, rpath/2 - xhigh + xmin)
 
 xticks = [6, 4, 2]
-zticks = [2, 4, 6]
+zticks = [0, 2, 4, 6]
 
-plotticks = xticks + [-z+rpath/2+xmin for z in zticks]
+plotticks = xticks + [-z+xmin for z in zticks]
 xplot.set_xticks(plotticks)
-xplot.set_xticklabels(xticks + zticks)
+xplot.set_xticklabels(xticks[:-1] + ["20"] + [""] + zticks[1:])
 zplot.set_xticks([])
 
 xplot.axvline(x=xmin, color='k')
 zplot.axvline(x=2*rpath, color='k')
 
-xloc = .621
-zloc = .802
-figtext(xloc, .04, r"$\underbrace{\hspace{9.3em}}$", horizontalalignment='center')
-figtext(xloc, .01, r"$x$", horizontalalignment='center')
-figtext(zloc, .04, r"$\underbrace{\hspace{12.3em}}$", horizontalalignment='center')
-figtext(zloc, .01, r"$z$", horizontalalignment='center')
-
+bracket.bracket(xplot, 0, (xlow - 2)/(xlow - xhigh), -.06, .04, r'$x$')
+bracket.bracket(xplot, (xlow - xmin)/(xlow - xhigh), 1.0, -.06, .04, r'$z$')
 
 twod_plot.set_xlim(zmin, zmax)
 twod_plot.set_ylim(-rmax, rmax)
 
 fig.subplots_adjust(hspace=0.001)
 
-for i in range(len(plots)):
-    g3_path = read_triplet_path(ff, plots[i])
+for name in plots:
[…]

### 2013-09-18 papers/pair-correlation/figs/plot-path.py

@@ -9,6 +9,8 @@ import scipy.ndimage
 import matplotlib.patheffects
 import os.path
 import math
+import bracket # our handy bracket function
+import styles # our preferred line styles
 
 from matplotlib import rc
 rc('text', usetex=True)
@@ -16,9 +18,7 @@ rc('text', usetex=True)
 from matplotlib.colors import NoNorm
 
 # these are the things to set
-colors = ['k', 'b', 'g', 'r', 'm']
-plots = ['mc', 'this-work', 'fischer', 'sokolowski'] # , 'gloor' sphere-dft
-titles = ['Monte Carlo', 'this work', 'Fischer et al.', 'Sokolowski'] # , 'gloor' test particle
+plots = ['mc', 'this-work', 'sokolowski', 'fischer'] # , 'gloor' sphere-dft
 
 dx = 0.1
 ############################
@@ -77,25 +77,25 @@ zplot = xplot.twiny()
 #zplot = fig.add_subplot(1,3,3, sharey=xplot)
 twod_plot = fig.add_subplot(1,2,1)
 
-xplot.set_xlim(6, -4)
+zmax_lineplot = 6.
+xmax_lineplot = 4.
+xplot.set_xlim(zmax_lineplot, -xmax_lineplot)
 xplot.set_xticks([6, 4, 2, 0, -2, -4])
 xplot.set_xticklabels([6, 4, "2 0", 2, 4, 6])
-zplot.set_xlim(-4, 6)
+zplot.set_xlim(-xmax_lineplot, zmax_lineplot)
 zplot.set_xticks([])
 #xplot.set_ylim(0)
 
-figtext(.613, .04, r"$\underbrace{\hspace{9em}}$", horizontalalignment='center')
-figtext(.613, .01, r"$x$", horizontalalignment='center')
-figtext(.796, .04, r"$\underbrace{\hspace{13.1em}}$", horizontalalignment='center')
-figtext(.796, .01, r"$z$", horizontalalignment='center')
+bracket.bracket(xplot, 0, xmax_lineplot/(xmax_lineplot+zmax_lineplot), -.06, .04, r'$x$')
+bracket.bracket(zplot, xmax_lineplot/(xmax_lineplot+zmax_lineplot), 1.0, -.06, .04, r'$z$')
 
 twod_plot.set_xlim(-0.5, 1.5*ymax)
 twod_plot.set_ylim(-ymax, ymax)
 
 fig.subplots_adjust(hspace=0.001)
 
[…]

### 2013-09-21 papers/pair-correlation/figs/plot-dadz-well.py
beautify dadz plots a bit more Tested on darcy. Took 0:28.35.

@@ -26,12 +26,21 @@ tw_z0, tw_dadz = data[:,0], data[:,1]
 data = loadtxt("figs/walls/square-well-dadz-sokolowski-%04.2f-1.790.dat" % eta)
 s_z0, s_dadz = data[:,0], data[:,1]
 
-plt.plot(tw_z0, tw_dadz, styles.plot['this-work'], label=styles.title['this-work'])
-plt.plot(s_z0, s_dadz, styles.plot['sokolowski'], label=styles.title['sokolowski'])
+plt.figure(figsize=(5,3))
 
-plt.plot(mc_z0, mc_da_dz, styles.plot['mc'], label=styles.title['mc'])
-plt.xlim([-.5,4])
+plt.plot(tw_z0/2, tw_dadz, styles.plot['this-work'], label=styles.title['this-work'])
+plt.plot(s_z0/2, s_dadz, styles.plot['sokolowski'], label=styles.title['sokolowski'])
+
+plt.plot(mc_z0/2, mc_da_dz, styles.plot['mc'], label=styles.title['mc'])
+plt.xlim([-.5/2,6/2.])
 plt.legend(loc='best').draw_frame(False)
+
+plt.xlabel('$z/\sigma$')
+plt.ylabel(r'$\frac{da_1}{dz}$')
+plt.yticks([])
+plt.xticks([0,1,2,3])
+
+plt.tight_layout()
 
 savefig("figs/dadz-square-well-%d.pdf" % (int(eta*10)))
 

### 2013-10-19 papers/pair-correlation/paper.tex
make pair-correlation figures more consistent and readable in b/w Tested on bingley. Took 2:09.61.
add square-well da1dz plot (currently broken) Tested on darcy. Took 0:22.33.
beautify dadz plots a bit more Tested on darcy. Took 0:28.35.
add text to results section of pair paper Tested on collins. Took 0:35.50.

@@
 
-\newcommand{\fixme}[1]{\text{\red{[#1]}}}
+\newcommand{\fixme}[1]{\red{[#1]}}
 
@@
 \begin{document}
-\title{Using Fundamental Measure Theory to treat the Pair
-  Distribution Function of the Inhomogeneous Hard-Sphere Fluid}
+\title{The Pair Distribution Function of the Inhomogeneous Hard-Sphere
+  Fluid}
 
@@
 
-\section{Results}
-\subsection{Pair distribution function}
-
 \newcommand\paircaption[1]{ The pair distribution function near a hard
@@
 
+\section{Results}
+
 \begin{figure}
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
-  \caption{\paircaption{0.2}}\label{fig:pair-distribution-2}
-\end{figure}
+\subsection{Pair distribution function}
+
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-2.pdf}
+%%   \caption{\paircaption{0.2}}\label{fig:pair-distribution-2}
+%% \end{figure}
 \begin{figure}
@@
 \end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
-  \caption{\paircaption{0.4}}\label{fig:pair-distribution-4}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-4.pdf}
+%%   \caption{\paircaption{0.4}}\label{fig:pair-distribution-4}
+%% \end{figure}
 
-Discuss the pair distribution function here, explaining plots
-  and discussing results and comparison.
+We begin by examining the pair distribution function near a hard wall,
+with a focus on the case where one of the two spheres is in contact
+with the hard wall.  Figure~\ref{fig:pair-distribution-1}a
+and~\ref{fig:pair-distribution-3}a compare the results of our
+approximate functional with Monte Carlo simulation at packing
+fractions of 0.1 and 0.3 respectively.  We see reasonable agreement at
+the lower density, with some reduced oscillations at larger distances,
+and a flatter angular dependence when the two spheres are in contact.
+At the higher density, we see significant structure developing in the
+simulation results that is not reflected in our approximation.
+
+Figure~\ref{fig:pair-distribution-1}b
+and~\ref{fig:pair-distribution-3}b show the pair distribution function
+as plotted along a path illustrated in
+Fig.~\ref{fig:pair-distribution-1}a.  These plots compare this work
+with Monte Carlo results, as well as the approximations of Sokolowski
+and Fischer~\cite{sokolowski1992role} and that of Fischer and
+Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
+The approach of Fischer and Methfessel is only valid when the two
+spheres are in contact, and is therefore only plotted on that segment
+of the path.  As an input to both previous approximations we use the
+hard sphere radial distribution function found using Monte Carlo
+simulation, interpolated as necessary.  We find that both previous
+approximations to the pair distribution function predict the
+qualtitative angular dependence of the pair distribution function at
+contact better than this work.  However, in each case the pair
+distribution function has a systematic error at contact---either too
+high or too low.  Thus while our approximation is smoother than either
+of the existing approaches, its errors will have a tendancy to cancel
+when used in a perturbation expansion.  At higher densities, the
+approximation of Fischer and Methfessel requires evaluating the radial
+distribution function at densities significantly higher than the
+freezing density, which poses numerical difficulties when using the
+radial distribution function from simulation.  When the two points are
+more than a radius away from contact, we find that any of these
+approaches gives a reasonable prediction, with our approach
+underestimating the oscillations in $g^(2)$, as expected based on our
+fit in Fig.~\ref{fig:radial-distribution}.
 
 \subsection{Triplet distribution function}
+
+Just as the radial distribution function of a homogeneous fluid may be
+computed from the density of an inhomogeneous one using Percus'
+test-particle trick, the triplet distribution function of a
+homogeneous system can be computed using an approximation of the pair
+distribution for an inhomogeneous fluid, such as we have developed.
+\begin{align}
+  g^{(3)}(\rr_1,\rr_2,\rr_3) &= \frac{n_{\textrm{TP}(\rr_1)}(\rr_2)
+  n_{\textrm{TP}(\rr_1)}(\rr_3)}{n^2}
+  g^{(2)}_{\textrm{TP}(\rr_1)}(\rr_2,\rr_3)
+\end{align}
+where $g^{(3)}(\rr_1,\rr_2,\rr_3)$ is the triplet distribution
+function of a homogeneous fluid with density $n$, and the
+$\textrm{TP}(\rr_1)$ subscript indicates quantities computed for the
+inhomogeneous density configuration in which one sphere is fixed at
+position $\rr_1$.  This method treats one of the three positions---the
+location of the test particle---differently from the others, which
+means that a poor approximation to the pair distribution function may
+beak the symmetry which is present in the true triplet distribution
+function.
 
@@
 \end{figure}
-
-Just as the radial distribution function of a homogeneous fluid may be
-computed from the density of an inhomogeneous one using Percus'
-test-particle trick, the triplet distribution function of a
-homogeneous system can be computed using an approximation of the pair
-distribution for an inhomogeneous fluid, such as we have developed.
-\begin{align}
-  g^{(3)}(\rr_1,\rr_2,\rr_3) &= \frac{n_{\textrm{TP}(\rr_1)}(\rr_2)
-  n_{\textrm{TP}(\rr_1)}(\rr_3)}{n^2}
-  g^{(2)}_{\textrm{TP}(\rr_1)}(\rr_2,\rr_3)
-\end{align}
-where $g^{(3)}(\rr_1,\rr_2,\rr_3)$ is the triplet distribution
-function of a homogeneous fluid with density $n$, and the
-$\textrm{TP}(\rr_1)$ subscript indicates quantities computed for the
-inhomogeneous density configuration in which one sphere is fixed at
-position $\rr_1$.  This method treats one of the three positions---the
-location of the test particle---differently from the others, which
-means that a poor approximation to the pair distribution function may
-beak the symmetry which is present in the true triplet distribution
-function.
 
@@
 \begin{figure}
-  \includegraphics[width=3cm]{figs/dadz-3-2.pdf}
-  \includegraphics[width=3cm]{figs/dadz-3-3.pdf}
-  \includegraphics[width=3cm]{figs/dadz-2-2.pdf}
-  \includegraphics[width=3cm]{figs/dadz-2-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
+  %\includegraphics[width=\columnwidth]{figs/dadz-3-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/dadz-2-2.pdf}
+  \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
   %%\includegraphics[width=3cm]{figs/dadz-4-3}

### 2013-10-19 papers/pair-correlation/paper.tex
add text to results section of pair paper Tested on collins. Took 0:22.07.
format long equation more nicely Tested on collins. Took 0:24.47.

@@
 distribution for an inhomogeneous fluid, such as we have developed.
-\begin{align}
-  g^{(3)}(\rr_1,\rr_2,\rr_3) &= \frac{n_{\textrm{TP}(\rr_1)}(\rr_2)
+\begin{multline}
+    g^{(3)}(\rr_1,\rr_2,\rr_3) =\\
+    \frac{n_{\textrm{TP}(\rr_1)}(\rr_2)
       n_{\textrm{TP}(\rr_1)}(\rr_3)}{n^2}
     g^{(2)}_{\textrm{TP}(\rr_1)}(\rr_2,\rr_3)
-\end{align}
+\end{multline}
 where $g^{(3)}(\rr_1,\rr_2,\rr_3)$ is the triplet distribution

### 2013-11-12 papers/pair-correlation/paper.tex
a few edits to pair correlation paper Tested on collins. Took 1:10.41.

@@
 \begin{align}
-  F_1 &= \tfrac12 \int g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|) d\rr_1d\rr_2
+  F_1 &= \tfrac12 \iint g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|) d\rr_1d\rr_2
 \end{align}
@@
 $\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
-Methfessel predicts the pair distribution function only at contact,
+Methfessel is intended to the pair distribution function only at contact,
 when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
@@
 desired, cannot be written as a single-site convolution, since the
-convolution kernel depends on both points.  This limits the accuracy.
+convolution kernel depends on both points.
 
@@
 \subsection{Separable fit to the radial distribution function}
-Figure~\ref{fig:radial-distribution} shows...
-
-We here construct a separable fit to the radial distribution function
-$g(r, g_\sigma)$, which is a function of the radius $r$ and its value
-at contact $g_\sigma$.  We could convert this to a more conventional
-function of the packing fraction $\eta$ by inverting the
+
+We have constructed a separable fit to the radial distribution
+function $g(r, g_\sigma)$, which is a function of the radius $r$ and
+its value at contact $g_\sigma$.  We could convert this to a more
+conventional function of the packing fraction $\eta$ by inverting the
 Carnahan-Starling formula for $g_\sigma$
@@
 averaged value of $g_\sigma$ is what we can directly compute for the
-inhomogeneous system.
+inhomogeneous system~\cite{schulte2012using}.
 
@@
 \fixme{add table with parameters}
+
+Figure~\ref{fig:radial-distribution} shows our function
+$g(r,g_\sigma)$ plotted against the $g(r)$ found from simulation.
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+    shows the triplet distribution function, with white corresponding
+    to the contact value of the \emph{pair} distribution function.
+    The plot on the right shows the value of $g^{(3)}$ along the path
+    shown as a dotted line on the plot to the
+    left.}\label{fig:triplet-contact-distribution-3}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+    shows the triplet distribution function, with white corresponding
+    to the contact value of the \emph{pair} distribution function.
+    The plot on the right shows the value of $g^{(3)}$ along the path
+    shown as a dotted line on the plot to the
+    left.}\label{fig:triplet-contact-distribution-4}
+\end{figure}
+
 Just as the radial distribution function of a homogeneous fluid may be
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-contact-distribution-3}
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
 \end{figure}
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
+  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted lin on the plot to the
-    left.}\label{fig:triplet-contact-distribution-4}
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
 \end{figure}
@@
 
+\section{Performance in thermodynamic perturbation theory}
+
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
+  \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
+  %\includegraphics[width=\columnwidth]{figs/dadz-3-3.pdf}
+  %\includegraphics[width=\columnwidth]{figs/dadz-2-2.pdf}
+  \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
+  %%\includegraphics[width=3cm]{figs/dadz-4-3}
+  \caption{Plot of $\frac{da_1}{dz}$ near a hard
+    wall. For what $\Phi$?}\label{fig:dadz}
 \end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
-\end{figure}
-
-\section{How well does it work?}
-Figure~\ref{fig:dadz} shows...
+
+The most relevant quantitative test of a pair distribution function is
+how well it predicts the interaction energy due to a pair potential.
+To this end, we have computed the error in the first term in a
+high-temperature perturbation expansion $a_1$ for several pair
+potentials.  In order to focus on effects at the interface, we have
+defined a position-dependent pair interaction energy as
+\begin{align}
+  \frac{dF_1}{dz} &=
+  \tfrac12 \int g^{(2)}_{HS}(\rr,\rr')n(\rr)n(\rr')\Phi(|\rr-\rr'|)
+  d\rr'\, dxdy
+\end{align}
+which gives the contribution to the mean-field free energy due to
+molecules located at position $z$.  We plot in Figure~\ref{fig:dadz}
+this quantity for two pair potentials.
+
+\section{Conclusion}
+
+\bibliography{paper}% Produces the bibliography via BibTeX.
 
@@
 \end{align}
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
-  %\includegraphics[width=\columnwidth]{figs/dadz-3-3.pdf}
-  \includegraphics[width=\columnwidth]{figs/dadz-2-2.pdf}
-  \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
-  %%\includegraphics[width=3cm]{figs/dadz-4-3}
-  \caption{Plot of $\frac{da_1}{dz}$ near a hard
-    wall. For what $\Phi$?}\label{fig:dadz}
-\end{figure}
 
@@
 
-\bibliography{paper}% Produces the bibliography via BibTeX.
+\end{document}
 

### 2013-11-12 papers/pair-correlation/figs/plot-dadz-well.py

@@ -40,6 +40,8 @@ plt.ylabel(r'$\frac{da_1}{dz}$')
 plt.yticks([])
 plt.xticks([0,1,2,3])
 
+plt.title('$\Phi(r) = \Theta(1.79\sigma - r)$     $\eta = %g$' % eta)
+
 plt.tight_layout()
 
 savefig("figs/dadz-square-well-%d.pdf" % (int(eta*10)))

### 2013-11-18 papers/pair-correlation/paper.tex
modified paper Tested on node33. Took .

@@
 \newcommand{\red}[1]{{\bf \color{red} #1}}
+\newcommand{\green}[1]{{\bf \color{green} #1}}
 \newcommand{\rr}{\textbf{r}}
@@
 \newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...
+\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
+\newcommand{\jeffsays}[1]{{\color{red} [\green{Jeff:} \emph{#1}]}}
 
@@
 
-The demand that our solution is correct in the homogeneous limit forms
-the basis of the derivation of our radial distribtuion's function
-form. We use the well known and very successful Carnahan Starling
-equation to calculated $g_{\sigma}$ and then simple monte-carlo
-simulations to calculate values of the homogeneous radial distribution
-function at varying distance perameter $r$ (Our previously published
-$g_{\sigma}$ approaches the Carnahan Starling version in the
-homogeneous limit and it's simplest to use an analytical expression
-for this calculation).  We then construct our radial distribution
-function that has the form of an expansion of powers of $h_{\sigma} =
-g_{\sigma}-1$ and perform a least squares fit to the monte-carlo
-solution to find the needed constant perameters.  The function is
-constructed in such a way as to appoach $g_{\sigma}$ in the limit of
-contact distance between the particles.  After this process we have a
-radial distribution function that is dependent only on $g_{\sigma}$
-and that gives the correct solution in the homogeneous and contact
-distance limit.
+We construct our function's form in terms of an expansion of powers of
+$h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
+in the contact distance limit, leaving our function with tunable
+constant perameters.  We run a simple monte-carlo simulations to
+calculate values of the homogeneous radial distribution function at
+varying distance perameter $r$.  We then tune our function's constant
+perameters by matching our function's results against the monte-carlo
+data, using a least squares fit method.  During this process it is
+simplest to use an analytic expression for $g_{\sigma}$ in our
+function and so we use the well known and very successful Carnahan
+Starling equation to calculate $g_{\sigma}$.  This is appropriate
+since that our published $g_{\sigma}$ approaches the Carnahan Starling
+version in the homogeneous limit and it's this limit, along with the
+contact limit, that we would like our constructed radial distribution
+function to get right.  After this process we have an appropriate
+radial distribution function that is dependent only on $r$ and
+$g_{\sigma}$.
 
@@
 \subsection{Separable fit to the radial distribution function}
+
+\jeffsays{I'm confused by this next part.  Are we just saying that
+  instead of using our published $g_{\sigma}$ we could have used a
+  Carnahan Starling derived $g_{\sigma}$? Don't understand why that
+  would add an unnecessary step in using g(r).}
 

### 2013-11-18 papers/pair-correlation/paper.tex
edit fit section of pair-correlation paper Tested on collins. Took .

@@
 The function has the form:
-\begin{align}
-  g(|\rr|) &= 1 + g_{\sigma}e^{-x_0(r-\sigma)} + x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
-  &~~-x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
-\end{align}
+\begin{multline}
+  g(|\rr|) = 1 + g_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
+  -x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
+\end{multline}
 The finalalized form of our pair correlation function is:
@@
 
-%% gsigma = gsigconcatenated[i] - 1
-%%         h0 = gsigma # was x[0]*gsig
-%%         f0 = numpy.exp(-x[0]*rconcatenated[i])
-%%         h1 = x[1]*gsigma
-%%         f1 = numpy.sin(x[2]*rconcatenated[i]) * numpy.exp(-x[3]*rconcatenated[i])
-%%         h2 = -x[4]*gsigma**(2)
-%%         f2 = numpy.sin(x[5]*rconcatenated[i]) * numpy.exp(-x[6]*rconcatenated[i])
-%%         g[i] = 1 + h0*f0 + h1*f1 + h2*f2
-%%     return g
-
-%% The homogeneoues limit $g_{sigm}$ can be calculated using the well
-%% known and very successful Carnahan Starling equation while the
-%% homogeneous limit radial distribution function can be calculated using
-%% Monte-Carlo methods.  We used these calculations in a least squares
-%% fit along the data points at increasing radial distances in order to
-%% fit the perameters of our more general pair correlation function.  Our
-%% function will therefore approach the most accurate solution possible
-%% in the homogeneous limit.  Created a function that has the form of an
-%% expansion in powers of $g_{\sigma}$.  $g_{sigm}$ in the homogeneous
-%% limit can be calculated using the well known and very successful
-%% Carnahan Starling equation.  has a well known and Perameters in the
-%% expansion where fit
 \begin{figure}
   \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g}% \\
+  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
   %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
@@
 
-\jeffsays{I'm confused by this next part.  Are we just saying that
-  instead of using our published $g_{\sigma}$ we could have used a
-  Carnahan Starling derived $g_{\sigma}$? Don't understand why that
-  would add an unnecessary step in using g(r).}
-
-We have constructed a separable fit to the radial distribution
-function $g(r, g_\sigma)$, which is a function of the radius $r$ and
-its value at contact $g_\sigma$.  We could convert this to a more
-conventional function of the packing fraction $\eta$ by inverting the
-Carnahan-Starling formula for $g_\sigma$
-\begin{equation}
-  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}
-\end{equation}
-but that would add an unnecessary step in the use of $g(r)$, since the
-averaged value of $g_\sigma$ is what we can directly compute for the
-inhomogeneous system~\cite{schulte2012using}.
-
-We construct $g(r, g_\sigma)$ using several constraints.  Naturally,
+In contrast to the previous work, which are based on the homogeneous
+radial distribution function $g(r,\eta)$ expressed as a function of
+distance and packing fraction, we have constructed our separable
+approximation for the radial distribution function as $g(r,
+g_\sigma)$, which is a function of the radius $r$ and its value at
+contact $g_\sigma$.  One can convert between these two representation
+using the Carnahan-Starling formula for $g_\sigma(\eta)$
+\begin{equation}
+  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
+\end{equation}
+The form $g(r, g_\sigma)$ is more convenient for treating the
+inhomogeneous case, because we can directly compute the averaged value
+of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.
+
+We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
 \begin{equation}
@@
 \end{equation}
-In addition, we know that
-\begin{align}
-  1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
-\end{align}
-\begin{equation}
-  \begin{split}
-    \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
-      p}\right)_T \\
-    &= - \left(V\left(\frac{\partial p}{\partial
-      V}\right)_T\right)^{-1}
-  \end{split}
-\end{equation}
-\begin{equation}
-    p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
-\end{equation}
-\begin{equation}
-    \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
-\end{equation}
-\begin{equation}
-    n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
-\end{equation}
-\begin{equation}
-    \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
-\end{equation}
-\begin{equation}
-  \begin{split}
-    \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
-    &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
-    &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
-    &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
-    &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
-  \end{split}
-\end{equation}
-\begin{equation}
-    nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
-  %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
-  %&= \eta \eta \cdots \fixme{fill in, double-check}
-\end{equation}
-where $\chi_T$ is the isothermal compressibility, for which we use the
-Carnahan-Starling equation of state.
-
-An additional constraint that we use sets the slope of $g(r,
-g_\sigma)$ at $r=\sigma$.  \fixme{explain this}
-
-Given these constraints, we use a functional form of
-\begin{equation}
-  g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
-  + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
-\end{equation}
-where $A$, $B$, and $C$ are amplitudes determined by the above
-constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
-simulations of the hard-sphere radial distribution function.
-
-\fixme{add paragraphs and equations with $A$, $B$ and $C$}
-
-\fixme{add table with parameters}
-
-Figure~\ref{fig:radial-distribution} shows our function
-$g(r,g_\sigma)$ plotted against the $g(r)$ found from simulation.
+A second constraint applies to the integral of the total correlation
+function
+\begin{align}
+  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
+\end{align}
+where $h(r) = g(r) - 1$ is the total correlation function.
+A final constraint is that the slope of the radial distribution
+function at contact is given to a very good approximation by
+\begin{align}
+  g'(\sigma) \approx - h_\sigma g_\sigma
+\end{align}
+where $h_\sigma = g_\sigma - 1$ is the total correlation function.
+Under these two constraints, we applied a least-squares fit to
+simulation data for the radial distribution function at packing
+fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
+distribution function is given by
+\newcommand\kappaa{\kappa_0}
+\newcommand\kappab{\kappa_1}
+\newcommand\kappac{\kappa_2}
+\begin{multline}
+  h(r,g_\sigma) = h_\sigma e^{-\kappaa z} \\
+  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
+  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
+      \eta^4}\frac{1}{n}
+    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
+\end{multline}
+\begin{equation}
+  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
+  - \frac{4\pi}{3} \sigma^3
+\end{equation}
+\begin{equation}
+  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
+\end{equation}
+\begin{equation}
+  I_2 = \frac{8\pi h_\sigma (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
+\end{equation}
+\begin{equation}
+  z = r - \sigma
+\end{equation}
+\davidsays{Can someone check the above equations for correctness
+  against the python code, in plot-ghs2.py.  We'll want to
+  ensure the C++ code agrees with this as well.}
+This function includes three fitted parameters, $\kappaa$, $\kappab$,
+and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
+The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
+figure, it is apparent how the on the integral of $h(r)$ from
+Eq.~\ref{eq:total-constraint} results in cancelling errors at large
+distances.  The maximum error in $g(r)$ at packing fractions of less
+than 0.4 is \fixme{get a number}.
+
+\begin{table}
+  \begin{tabular}{cc}
+    $\kappaa$ & 1 \\
+    $\kappab$ & 1 \\
+    $\kappac$ & 1 \\
+  \end{tabular}
+  \caption{Fit parameters \fixme{... put them here and format this nicely}}
+  \label{tab:fit-parameters}
+\end{table}
+
+%% In addition, we know that
+%% \begin{align}
+%%   1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
+%% \end{align}
+%% \begin{equation}
+%%   \begin{split}
+%%     \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
+%%       p}\right)_T \\
+%%     &= - \left(V\left(\frac{\partial p}{\partial
+%%       V}\right)_T\right)^{-1}
+%%   \end{split}
+%% \end{equation}
+%% \begin{equation}
+%%     p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
+%% \end{equation}
+%% \begin{equation}
+%%     \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
+%% \end{equation}
+%% \begin{equation}
+%%     n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
+%% \end{equation}
+%% \begin{equation}
+%%     \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
+%% \end{equation}
+%% \begin{equation}
+%%   \begin{split}
+%%     \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
+%%     &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
+%%     &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
+%%     &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
+%%     &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
+%%   \end{split}
+%% \end{equation}
+%% \begin{equation}
+%%     nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
+%%   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
+%%   %&= \eta \eta \cdots \fixme{fill in, double-check}
+%% \end{equation}
+%% where $\chi_T$ is the isothermal compressibility, for which we use the
+%% Carnahan-Starling equation of state.
+
+%% An additional constraint that we use sets the slope of $g(r,
+%% g_\sigma)$ at $r=\sigma$.  \fixme{explain this}
+
+%% Given these constraints, we use a functional form of
+%% \begin{equation}
+%%   g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
+%%   + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
+%% \end{equation}
+%% where $A$, $B$, and $C$ are amplitudes determined by the above
+%% constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
+%% simulations of the hard-sphere radial distribution function.
 

### 2013-11-19 papers/renormalization/paper.tex
add very sketcy draft of renormalization paper Tested on collins. Took .

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{color}

\newcommand{\fixme}[1]{{\bf\color{red}{[#1]}}}
\newcommand{\rr}{\mathbf{r}}

\begin{document}
\title{Renormalization}

\author{Daniel ?. Roth}
\author{Michael ?. Perlin}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

\begin{abstract}
  We present an analysis of the convergence properties of
  renormalization group theory  when applied to the square-well
  liquid.
\end{abstract}

\maketitle

Long-wavelength fluctuations lead to significant changes in fluid
behavior in two scenarios: near the critical point, and at
liquid-vapor interfaces.  The failing of perturbation theory near the
critical point is well-understood, and is commonly treated in the
homogeneous case using either renormalization group
theory~\cite{white2000global, white2001global, del2002vapour,
  kiselev2002computer, reiner2002hierarchical, mi2004renormalization,
  mi2004improved, fu2006study, giacometti2009liquid, jiuxun2005simple,
  forte2011application, el2008integral, ramana2012generalized}, or
crossover theory~\cite{kiselev1999crossover, kiselev2000crossover,
  kiselev2001crossover, kiselev2000simplified, hu2003crossover,
  hu2003back, mccabe2004crossover, llovell2006global}.  The
liquid-vapor interface is beset by long-wavelength fluctuations in the
form of capillary waves~\cite{buff1965interfacial,
  weeks1989consistency}, which are handled---when treated at all---as
an analytic correction to the surface tension computed using density
functional theory~\cite{wadewitz2000application, winkelmann2001liquid,
  gross2009density}.  These corrections to the surface tension involve
an ad hoc cutoff wavelength, and do not correct the prediction of the
interfacial density profile.  The effect of capillary waves on surface
tension is largest near the critical point.  Together, these effects
suggest that a density functional theory directly incorporating
long-wavelength fluctuations would be a major gain for the study of
liquid interfaces.

I plan to develop such a functional using renormalization group theory
(RGT).  Implementation of RGT in an inhomogeneous system will be
computationally challenging, because it is an inherently recursive
theory.  This results in long-range fluctuations (and correlations)
near the critical point, which will be reflected in a long-range
density functional.  It will be more computationally
intensive than existing density functionals, but at the same time will
enable far greater predictive power, over a wide range of
temperatures, pressures, and density distributions.

There has been one study of the effect of critical behavior on surface
tension using a renormalization approach, which used the Density
Gradient Theory (DGT) rather than DFT due to its easier integration
with renormalization group theory~\cite{fu2006study}.  The use of DGT,
which is not self-consistent, puts a major limitation the power of the
method, particularly when applied to diverse interfaces.

We will study a hard-sphere fluid with a square well attractive
potential.  This is a widely used model fluid within the SAFT
community~\cite{mi2004renormalization, mi2004improved, fu2006study,
  forte2011application} as well as in the study of the critical
behavior of fluids~\cite{white2000global, white2001global,
  del2002vapour, kiselev2002computer, reiner2002hierarchical,
  mi2004renormalization, mi2004improved, fu2006study,
  giacometti2009liquid, jiuxun2005simple, forte2011application,
  el2008integral, ramana2012generalized}. This model is particularly
suitable for testing over a wide range of temperatures and pressures
due to its ease of computation.  Because the set of energy levels is
discrete, this system is suitable for the use of histogram-based Monte
Carlo methods~\cite{ferrenberg1988histogram, lee1993new, de1998broad}.
These approaches accelerate the computation of thermodynamic
properties over a range of temperatures, and are particularly helpful
near phase transitions and critical points.  There has been a recent
Monte Carlo study of phase behavior of the the square-well
fluid~\cite{liu2005direct}, which did not take advantage of the
histogram family of Monte Carlo methods, although it did use the
aggregation volume bias sampling method to accelerate sampling of
bonded and non-bonded configurations~\cite{chen2000novel}.

We will begin by studying the simple square-well liquid already
introduced in a previous paragraph.  This liquid is a standard test
case for theories of critical behavior, although its critical behavior
at the liquid-vapor interface has not yet been examined.  We will
closely follow the approach recently developed by Forte \emph{et al.}
who constructed a SAFT-VR equation of state using
RGT~\cite{forte2011application}.  Their approach to RGT is expressed
in terms of the pair distribution function
$g^{(2)}_{HS}(\rr_1,\rr_2)$, which will enable us to leverage our
earlier work (from the previous paragraph) to construct a
renormalization group classical density functional theory for the
inhomogeneous fluid.  Once we have developed a RGT density functional,
we will test it against Monte Carlo simulations of the surface
tension, as well as a fluid near the critical point at a hard wall.
%
Future work will include applying this approach to associating fluids.

\section{Methods}

\subsection{Renormalization Group Theory}

\fixme{Here we will explain the RGT that we use, and how it works
  out.}

\subsection{Monte Carlo}

\fixme{Here we will explain our Monte Carlo simulations.}

\section{Results}

\section{Conclusion}

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2013-11-20 papers/pair-correlation/notes.py
add some notes from g-sigma-prime derivation (probably no longer relevant) Tested on collins. Took 0:29.96.

@@ -1,0 +1,115 @@
+from __future__ import division
+from sympy import *
+
+sigma, delta, N, V, kT = symbols('sigma delta N V kT')
+r1, r2, r3, r4, rN = symbols('r_1 r_2 r_3 r_4 r_N')
+Theta = symbols('\\Theta', cls=Function)
+
+allthetas = symbols(r'{\prod_{i<j}}\Theta')
+
+
+def multint(expr, rs):
+    for r in rs:
+        expr = Integral(expr, r)
+    return expr
+
+Z = multint(allthetas, [r1, r2, r3, r4, rN])
+Zsym = symbols('Z')
+
+
+def n2(ra, rb):
+    integrals = allthetas
+    for r in [r1, r2, r3, r4, rN]:
+        if r != ra and r != rb:
+            integrals = Integral(integrals, r)
+    return N**2 * integrals/Zsym
+
+def n4(ra, rb, rc, rd):
+    integrals = allthetas
+    for r in [r1, r2, r3, r4, rN]:
+        if r != ra and r != rb and r != rc and r != rd:
+            integrals = Integral(integrals, r)
+    return N**4 * integrals/Zsym
+
+f = open('notes-content.tex', 'w')
+
+f.write(r'''
+To begin, let's define $Z$ as
+\begin{equation}
+%s
+\end{equation}
+''' % latex(Eq(Zsym, Z)))
+
+n2sym = symbols('n_2',cls=Function)(r1,r2)
+f.write(r'''
+And now we define $n_2$ and $n_4$ (which we normally write as $n^{(2)}$
+and $n^{(4)}$, but I will abbreviate in these notes):
+\begin{equation}
+%s
+\end{equation}
+''' % latex(Eq(n2sym, n2(r1, r2))))
+
+n4sym = symbols('n_4',cls=Function)(r1,r2,r3,r4)
+f.write(r'''\begin{equation}
+%s
+\end{equation}
+''' % latex(Eq(n4sym, n4(r1, r2, r3, r4))))
+
+Zplusdelta = Zsym + Rational(1,2)*multint(Zsym*n2(r1,r2)*Theta(r1,r2), [r1,r2]) + Rational(1,8)*multint(Zsym*n4(r1,r2,r3,r4)*Theta(r1,r2)*Theta(r3,r4), [r1,r2,r3,r4])
+
+f.write(r'''
[…]

### 2013-11-21 papers/pair-correlation/paper.tex
make plot-ghs2.py generate numbers to include (or input) directly into *.cpp (or paper.tex) Tested on collins. Took 0:30.35.

@@
 \end{figure}
+
+\input{figs/fit-parameters}
+
 \subsection{Separable fit to the radial distribution function}
@@
 distances.  The maximum error in $g(r)$ at packing fractions of less
-than 0.4 is \fixme{get a number}.
+than 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+\rmaxerr$.
 
@@
   \begin{tabular}{cc}
-    $\kappaa$ & 1 \\
-    $\kappab$ & 1 \\
-    $\kappac$ & 1 \\
+    $\kappaa$ & \kappazero \\
+    $\kappab$ & \kappaone \\
+    $\kappac$ & \kappatwo
   \end{tabular}
@@
 
-\appendix
-
-\section{Properties of $g_\sigma$}
-We begin with some definitions of basic quantities, looking at the
-homogeneous hard-sphere fluid.  I am ignoring here the distinction
-between $N$ and $N-1$, which is unimportant in the thermodynamic limit
-$N\gg 1$.
-\begin{align}
-  Z(\sigma) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
-  \\
-  n^{(2)}(\rr_1,\rr_2) &= \frac{N^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)
-\end{align}
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) &=
-  \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
-  \\
-  &= \frac{V^2}{Z}\iiiint d\rr_3\cdots d\rr_N \prod_{i<j}
-  \Theta(r_{ij}-\sigma)
-  \label{eq:g2-integral-defn}
-\end{align}
-
-At this point we consider the effect of increasing the diameter
-$\sigma$ of all our spheres.  This will reduce the number of available
-states, thus reducing the partition function $Z$, which we will
-ultimately relate to a change in free energy.
-\begin{align}
-  Z(\sigma+\delta) &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma-\delta)
-  \\
-  &= \iiiint d\rr_1\cdots d\rr_N \prod_{i<j} \Theta(r_{ij}-\sigma)(1-\Theta(\sigma+\delta-r_{ij}))
-\end{align}
-I have written $Z(\sigma+\delta)$ in such a way that it is clear that
-it is equal to $Z(\sigma)$ plus a small quantity, provided $\delta$ is
-small.  We will now investigate that small quantity, which we will
-write as a power series in $\delta$.  We begin by expanding the
-product of $(1-\Theta(\sigma+\delta-r_{ij}))$:
-\begin{multline}
-    \prod_{i<j}(1-\Theta(\sigma+\delta-r_{ij})) =
-    1 - \sum_{i<j} \Theta(\sigma+\delta-r_{ij}) \\
-    + \frac{1}{2}\mathop{\mathop{\sum_{i<j}}_{k<l}}_{kl \ne ij}
-    \Theta(\sigma+\delta-r_{ij}) \Theta(\sigma+\delta-r_{kl})
-    \\ + \cdots.
-\end{multline}
-Given this expansion, we can readily expand the partition function:
-\begin{multline}
-  Z(\sigma+\delta) = Z(\sigma) \\ -
-  \sum_{k<l} \iiiint \Theta(\sigma+\delta-r_{kl}) \prod_{i<j}
-  \Theta(r_{ij}-\sigma) \\
-  + \frac{1}{2}\mathop{\mathop{\sum_{k<l}}_{m<n}}_{kl\ne mn}
-  \iiiint \Theta(\sigma+\delta-r_{kl}) \Theta(\sigma+\delta-r_{mn}) \prod_{i<j}
-  \Theta(r_{ij}-\sigma) \\
-  + \cdots.\label{eq:z-power-series}
-\end{multline}
-Our challenge now is to manipulate this into forms involving $g(r)$,
-which we can then simplify using a Taylor expansion around
-$r=\sigma$.  We will begin by taking advantage of the fact that the
-summations in Eq.~\ref{eq:z-power-series} are mostly of identical
-terms, since the integrals are over all coordinates.
-\begin{multline}
-  Z(\sigma)-Z(\sigma+\delta) =
-  \frac{N^2}{2}\iiiint \Theta(\sigma+\delta-r_{12})
-  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
-  - \frac{N^4}{8}\iiiint
-  \Theta(\sigma+\delta-r_{12})
-  \Theta(\sigma+\delta-r_{34})
-  \prod_{i<j} \Theta(r_{ij}-\sigma) \\
-  + \cdots \label{eq:z-expansion}
-\end{multline}
-At this point we are ready to connect with $g(r)$, which we can do by
-considering an integral of the pair distribution function, from
-Eq.~\ref{eq:g2-integral-defn}.
-\begin{multline}
-  \int_{\sigma}^{\sigma+\delta} g(r) d\rr\\
-  \shoveleft{\quad= \int g^{(2)}(\rr_1,\rr_2)
-  \Theta(\sigma+\delta-r_{12})d\rr_2} \\
-  \shoveleft{\quad= \frac{1}{V}\int g^{(2)}(\rr_1,\rr_2)
-  \Theta(\sigma+\delta-r_{12})d\rr_1d\rr_2} \\
-  \shoveleft{\quad= \frac{V}{Z}\iiiint d\rr_1\cdots d\rr_N
-  \Theta(\sigma+\delta-r_{12})}\\
-  \prod_{i<j} \Theta(r_{ij}-\sigma)
-  \label{eq:g-shell-integral}
-\end{multline}
-Looking back at Eq.~\ref{eq:z-expansion}, we can see that the first
-term is easily related to Eq.~\ref{eq:g-shell-integral}.  The second
-term has two shell $\delta$ functions, and is thus a bit trickier.
-Fortunately, in a very large system, $\rr_3$ and $\rr_4$ are
-vanishingly seldom near $\rr_1$ or $\rr_2$, so we can write the second
-term as a product.
-\begin{multline}
-  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)}\\
-  \shoveleft{\quad= \frac{nN}{2} \int_{\sigma}^{\sigma+\delta} g(r) d\rr }
-  - \frac{n^2N^2}{8} \left(\int_{\sigma}^{\sigma+\delta} g(r)
-  d\rr\right)^2
-  + \cdots
-  \label{eq:zdiff-g}
-\end{multline}
-Now let us look at how this integral relates to $g(r)$ evaluated at
-contact.  I'll skip a few steps doing this integral...
-\begin{equation}
-   \int_{\sigma}^{\sigma+\delta} g(r)
-  d\rr = 4\pi\sigma^2\left(g_\sigma\delta + \left(\tfrac12g_\sigma' +
-  \frac{g_\sigma}{\sigma}\right)\delta^2 + \cdots \right)
-\end{equation}
-Putting this together with the above we find
-\begin{multline}
-  \frac{Z(\sigma)-Z(\sigma+\delta)}{Z(\sigma)} =
-  2\pi\sigma^2 n N g_\sigma \delta +
-  \pi\sigma n N  \left(g_\sigma'\sigma + 2g_\sigma\right)
-  \delta^2 \\
-  - 2\pi^2\sigma^4n^2N^2g_\sigma^2\delta^2 + \mathcal{O}(\delta^3)
-\end{multline}
-But by considering a Taylor expansion, we can also see that
-\begin{multline}
-  Z(\sigma)-Z(\sigma+\delta) =
-  -\frac{\partial Z}{\partial \sigma} \delta
-  -\frac12 \frac{\partial^2 Z}{\partial \sigma^2} \delta^2  + \mathcal{O}(\delta^3)
-\end{multline}
-So that
-\begin{multline}
-  \frac{\partial Z}{\partial \sigma} =
-  -2ZnN\pi\sigma^2g_\sigma\\
-  \frac{\partial^2 Z}{\partial \sigma^2} =
-  -2ZnN\pi\sigma\left(g_\sigma'\sigma + 2g_\sigma \right)
-  +4Zn^2N^2\pi^2\sigma^4{g_\sigma}^2
-\end{multline}
-from which we can extract relationships for the derivatives of $Z$
-with respect to $\sigma$.  Which if we knew what those derivatives
-were, would tell us about $g_\sigma'$ and $g_\sigma$, which is what we
-are after here.  We find this out by considering the Helmholtz free
-energy.
-\begin{equation}
-  F = -kT \ln Z
-\end{equation}
-The derivatives are then given by
-\begin{equation}
-  \frac{\partial F}{\partial \sigma} = -\frac{kT}{Z} \frac{\partial
-    Z}{\partial \sigma}
-\end{equation}
-and
-\begin{equation}
-  \frac{\partial^2 F}{\partial \sigma^2} =
-  \frac{1}{kT}\left(\frac{\partial F}{\partial \sigma}\right)^2
-  -\frac{kT}{Z} \frac{\partial^2
-    Z}{\partial \sigma^2}.
-\end{equation}
-With a bunch more work, we find out that
-\begin{equation}
-  g_\sigma' = \frac{\frac{\partial^2 F}{\partial
-      \sigma^2}}{2\pi\sigma^2nNkT}
-  - \frac{2g_\sigma}{\sigma}
-\end{equation}
-except that there is a mistake in here.  I can tell because $F$ is
-extensive and $g_\sigma'$ is not.  This tracks back to
-Eq.~\ref{eq:zdiff-g}, which is correct in this \LaTeX, but is wrong on
-my scratch paper.
-
 \end{document}
-
-\section{THE END OF THE PAPER}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Introduction}
-
-Attard worked out the triplet correlation function, which is something
-that we can access from the pair distribution function, if we consider
-the pair correlation near a single hard sphere
-solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al} wrote an
-interesting paper using DFT and Monte Carlo to compute the
-three-particle distribution function (same as the triplet correlation
-function)\cite{gonzalez1999test}.  Their theory compares favorably
-with the ``superposition approximation'', and we could probably do
-something similar.
-
-Plischke and Henderson worked out the pair correlation function near a
-hard wall using Percus-Yevick, and compare with Monte Carlo
-results\cite{plischke1986pair}.  They plot the pair correlation
-function along interesting paths.  \fixme{Jeff read this paper for a
-  while and does not understand what they do in oder to find the pair
-  correlation function as compared to the radial distribution function}
-
-Lado recently introduced a new and improved efficient algorithm for
-implementing integral equation theory for inhomogeneous fluids, which
-computes $g^{(2)}(\rr_1,\rr_2)$ among other
-things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
-use, and their example is limited to 1D systems.  And the scaling is
-still (so far as I can tell) $O(N^2)$, since their improvement is in
-the prefactor, by developing a nice basis set.
-
-\section{Previous approaches}
-
-Gloor \emph{et al} explain that little is known of the pair
-distribution function of the hard-sphere fluid, and therefore use the
-formula
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
-\end{align}
-They go on to further approximate the radial distribution function,
-but we can just stop here.
-
-Gloor et al \cite{gloor2007prediction} uses average density of two points:
-\begin{align}
-\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
-\end{align}
-and the uses bulk radial distribution for the reference system (hard
-spheres) at this density, in order to approximate pair correlation.
-
-Evans \cite{evans1992density}perscribes to use the same approximation but with an unspecified 'mean denisty':
-
-\begin{align}
-g^{(2)}(\rr_1,\rr_2) &\equiv g_r(\rho^m,\rr_{12})
-\end{align}
-
-Gross \cite{gross2009density} uses a hard chain reference fluid.  In
-his functional he needs a pair correlation beween the different
-segments of chains.  He approximates this function with the radial
-distribution given by the Carnahan Starling at a density that is the
-average density of the segment positions $\rr$ and $\rr'$:
-
-\begin{align}
-\overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
-\end{align}
-
-Fu and Wu \cite{fu2005vapor} do I think the same as Gloor et al but
-maybe you (Dr. Roundy) should take a look at it.  They reference Tang
-\cite{tang2008accurate} which I've mentioned below in the Appendix.
-\fixme{DJR: Not same as Gloor, they do a WDA approximation.}
-
-Sokolowski and Fischer \cite{sokolowski1992role} do the same as Gloor
-et al except that instead of taking the exact densities
-$\rho(\mathbf{r_1})$ and $\rho(\mathbf{r_2})$ to average he takes an
-integral $\int_{V_{sphere}} \rho(\mathbf{r})d\mathbf{r}$ over a sphere
-of some radius centered at each point.  It's these two averaged
-densities that he takes the average of.
-
-Fischer and Methfessel \cite{fischer1980born}, in a paper that is
-cited by Sokolowski and Fischer (above) do the same thing (the
-originate the idea of the spherical average density.  Their
-Introduction has a very good discussion of the different averaging
-methods to get to $g(\mathbf{r})$ tried up until then (paper is
-written in 1980).
-
-McCabe and Kiselev \cite{mccabe2004application} model a square well
-potential.  One of their terms uses a radial distribution function
-that once again is taken to be the bulk function evaluated at an
-effective density.  They reference a source that calculates this
-effective density (or filling fraction) to be $\eta_{effective} =
-c_1\eta + c_2\eta^2 + c_3\eta^3$ where the coeffecients are gotten
-from three coupled polynomial equations with $1,\lambda$, and
-$\lambda^2$, where $\lambda$ is a measure of the model's potential
-range.  They reference another source who calculates the effective
-density in a similar (but more complicated) way.  I can look more at
-these if you'd like.  They also use a different first order expansion
-formula for the ditribution at contact in the chain term.
-
-Toxvaerd \cite{toxvaerd1976hydrostatic}approximates the radial
-distribution as the radial distribution of the bulk at the same
-density.  Uses the first two terms in a tempurature expansion.  First
-term is just Carnahan Starling derived, second is empirical.
-
-\section{Pair distribution function with inhomogeneity}
-
-The pair distribution function can be defined using the two particle
-density as
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) &\equiv \frac{n^{(2)}(\rr_1,\rr_2)}{n(\rr_1)n(\rr_2)}
-\end{align}
-
-We may also want to consider the \emph{triplet} distribution function
-\begin{align}
-  g^{(3)}(\rr_1,\rr_2,\rr_3) &\equiv \frac{n^{(3)}(\rr_1,\rr_2,\rr_3)}{n(\rr_1)n(\rr_2)n(\rr_3)}
-\end{align}
-The triplet distribution is relevant, because we can compute the
-triplet distribution of the homogeneous system by considering the
-\emph{pair} distribution of the inhomogeneous ``test-particle
-scenario.''  The process is that we solve the inhomogeneous situation
-in which we have one hard sphere located at the origin.  We solve for
-the density around that hard sphere, and this gives us the pair
-distribution function of the homogeneous system $g^{(2)}(0, \rr)$.
-When we solve for the pair distribution $g^{(2)}(\rr_1,\rr_2)$ of the
-inhomogeneous system, this will give us the \emph{triplet}
-distribution function of the homogeneous system
-$g^{(3)}(0,\rr_1,\rr_2)$.
-
-\section{Goals of this paper (to be integrated into rest of paper)}
-
-In this paper, we seek an efficient approximation for the pair
-correlation function of an inhomogeneous hard-sphere fluid, where we
-define efficient as an algorithm that scales as $O(N\log N)$ where $N$
-scales as the volume of the system under consideration.
-Integral equation theory (look this up) scales as $O(N^2)$, which can
-be very expensive for large three-dimensional systems.
-
-Fundamental-Measure Theory (FMT) is constructed using only single-point
-convolutions (check terminology) such as
-\begin{align}
-  n_\alpha(\rr) &= \int n(\rr')w_\alpha(|\rr-\rr'|) d\rr'
-\end{align}
-These convolutions can be computed in Fourier space in $O(N\log N)$
-time, with the result that the FMT free energy---and its
-gradient---can be computed in $O(N\log N)$ time.
-
-The pair distribution function is used in computing the first term in
-the high-temperature perturbation expansion, which is used in SAFT
-(for instance) to treat dispersion interactions.  This term is written
-as
-\begin{align}\label{eq:a1}
-  a_1 &\propto \iint n(\rr)n(\rr') \Phi(|\rr-\rr'|)
-  g^{(2)}(\rr,\rr')d\rr d\rr'
-\end{align}
-where $\Phi(r)$ is the pair potential.  If $g^{(2)}(\rr,\rr')$ could
-be written as $g(|\rr-\rr'|)$, this term could be computed in $O(N\log
-N)$ time.  However, this is not possible when the density is
-inhomogeneous, because the pair distribution function depends on the
-density.
-
-In this paper, we consider a superposition approximation, in which
-construct an approximate representation for the pair distribution
-function, which allows the first term in the high temperature
-expansion given in Equation~\ref{eq:a1} to be computed in $O(N\log N)$
-time.  One input to this approach is our recently published expression
-for an average of the value of the pair distribution function at
-contact $g_\sigma(\rr)$, which itself can be computed in $O(N\log N)$
-time.
-
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, \eta(g_\sigma(\rr_1))) + g(|r_{12}|, \eta(g_\sigma(\rr_2)))}{2}
-\end{align}
-Bad:
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) = g\left(|r_{12}|, \frac{\eta(g_\sigma(\rr_1)) + \eta(g_\sigma(\rr_2))}{2}\right)
-\end{align}
-
-The general idea is to write
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \Theta(r_{12}-\sigma)
-  \left(1 + \sum_{\alpha} \frac{f_{\alpha}(g_\sigma(\rr_1)) + f_{\alpha}(g_\sigma(\rr_2))}{2}g_{\alpha}(r_{12})\right)\label{eq:gpair-approx}
-\end{align}
-where $f_{\alpha}$ and $g_{\alpha}$ are a set of functions that
-satisfy the following constraint:
-\begin{align}\label{eq:definealpha}
-  \sum_{\alpha} f_{\alpha}(g_\sigma(\eta)) g_{\alpha}(r) \approx g(r; \eta)
-\end{align}
-which ensures that our pair distribution function reduces
-(approximately) to the radial distribution function in the homogeneous
-limit.  Equation~\ref{eq:gpair-approx} is constructed such that
-Equation~\ref{eq:a1} can be evaluated in $O(N\log N)$ time, since all
-the required convolutions may be performed with a fixed kernel.
-
-Our proposed pair distribution function has several good properties.
-When evaluated at contact ($r_{12}=\sigma$), it gives the same good
-result as our previously-published functional for the averaged pair
-distribution function at contact, which is constructed to give an
-accurate prediction for the contribution to Equation~\ref{eq:a1} from
-spheres in contact.  In the homogeneous limit,
-Equation~\ref{eq:gpqair-approx} should give a prediction that is as
-accurate as the approximate equality in
-Equation~\ref{eq:definealpha}.  In inhomogeneous distributions and at
-intermediate distances, we expect our pair distribution function to be
-least accurate, but one hopes that it will be reasonable, simply
-because it is an average of two reasonable results.
-
-\subsection*{Fundamental-Measure Theory}
-
-We use the White Bear version of the Fundamental-Measure Theory~(FMT)
-functional~\cite{roth2002whitebear}, which describes the excess free
-energy of a hard-sphere fluid.  The White Bear functional reduces to
-the Carnahan-Starling equation of state for homogeneous systems.  It
-is written as an integral over all space of a local function of a set
-of ``fundamental measures'' $n_\alpha(\rr)$, each of which is written
-as a one-center convolution of the density.  The White Bear free
-energy is thus
-\begin{equation}
-A_{\textit{HS}}[n] = k_B T \int \left(\Phi_1(\rr) + \Phi_2(\rr) + \Phi_3(\rr)\right) d\rr \; ,
-\end{equation}
-with integrands
-\begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right) \label{eq:Phi1}\\
-\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= (n_2^3 - 3 n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2}) \frac{
-  n_3 + (1-n_3)^2 \ln(1-n_3)
-}{
-  36\pi n_3^2\left( 1 - n_3 \right)^2
-} , \label{eq:Phi3}
-\end{align}
-using the fundamental measures
-\begin{align}
-  n_3(\rr) &= \int n(\rr') \Theta(\sigma/2 -\left|\rr - \rr'\right|)
-  d\rr' \label{eq:FMn3} \\
-  n_2(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) d\rr' \\
-  \mathbf{n}_{2V}(\rr) &= \int n(\rr') \delta(\sigma/2 -\left|\rr - \rr'\right|) \frac{\rr-\rr'}{|\rr-\rr'|}d\rr'
-\end{align}
-\begin{align}
-  \mathbf{n}_{V1} = \frac{\mathbf{n}_{V2}}{2\pi \sigma}, \quad
-  n_1 &= \frac{n_2}{2\pi \sigma} , \quad
-  n_0 = \frac{n_2}{\pi \sigma^2} \label{eq:FMrest}
-\end{align}
-
-\section{Theoretical Approaches}
-
-The traditional approach for solving for the pair distribution
-function of an inhomogeneous density distribution is to use an
-integral equation theory...
-
-\section{Comparison with simulation}\label{sec:comparison}
-
-\section{Conclusion}
-
-\appendix
-
-\section*{Appendix}
-\cite{jin2011perturbative} Combines hard sphere (white bear) with
-attraction that is not mean field but rather uses two other
-correltation expressions given in -\cite{tang2008accurate} and
-\cite{hlushak2009direct}
-
-Tang \cite{tang2008accurate}Derives the direct correltation for a SW
-and Yukawa and LJ potentials by first getting the radial distribution
-function through a method I don't really understand.  Composes fourier
-transforms $c(k)$ and $h(k)$ of these complicated functions that
-depend on filling fraction and the potential in an integral. Uses
-first order mean spherical approximation.  This is just square
-potential well around particle but the radial distribution function
-takes only the first order in an expansion in... (high temp?).
-
-\cite{hlushak2009direct} Cites \cite{tang2008accurate} (above) and
-does largely the same thing but then adds a complication for $2 <
-\lambda \leq 3$
-
-\cite{levesque2012scalar} Uses direct correlation funtion that's
-gotten from a radial distribution function of a homogeneous density.
-The direct correlation itself is only dependent on distance.  The
-$\rho_0$ that they use is the same $\rho_0$ that the used in the
-reference HS term in their functional.  The use the Carnahan Starling
-or the Perkis Yevick for these hard sphere free energies.  The other
-terms involve $\delta\rho$ so the change is a variation from the
-reference density.  They further modify it by looking at molecular
-dynamic simulations of a certain model of water molecules and looking
-at the oxygen-oxygen correlation.
-
-\end{document}
-
-%% G{\"o}tzelmann \emph{et al.} worked out the pair distribution function near a hard
-%% wall by finding the direct correlation function by taking a functional
-%% derivative of the free energy functional and then solving the
-%% Ornstein-Zernicke equation~\cite{gotzelmann1996structure}.  They have plots like our pretty plots,
-%% but with the fixed sphere one radius away from touching.
-%% Interestingly, they plot the second bit as $z$, which enables them to
-%% have just two axis values.  I think I prefer our ``total distance''
-%% method, but there is a certain elegance to just plotting $z$ for the
-%% curvy part as well as the straight-away part.  They found poor
-%% quantitative agreement, particularly at contact.  They also studied
-%% the pair correlation function in a hard slit~\cite{gotzelmann1997density}.
-
-%% Integrating the Ornstein-Zernicke equation is challenging, requiring a
-%% variational minimization itself~\cite{paul2003variational}.
-
-%% Attard worked out the triplet correlation function, which is something
-%% that we can access from the pair distribution function, if we consider
-%% the pair correlation near a single hard sphere
-%% solute\cite{attard1989spherically}.  Gonz\'alez \emph{et al.} wrote an
-%% interesting paper using DFT (test-particle approach of Percus) and
-%% Monte Carlo to compute the three-particle distribution function (same
-%% as the triplet correlation function)\cite{gonzalez1999test}.  Their
-%% theory compares favorably with the ``superposition approximation'',
-%% and we could probably do something similar.
-
-%% Plischke and Henderson worked out the pair correlation function near a
-%% hard wall using Percus-Yevick, and compare with Monte Carlo
-%% results\cite{plischke1986pair}.  They plot the pair correlation
-%% function along interesting paths.  \fixme{Jeff read this paper for a
-%%   while and does not understand what they do in oder to find the pair
-%%   correlation function as compared to the radial distribution function}
-
-%% Lado recently introduced a new and improved efficient algorithm for
-%% implementing integral equation theory for inhomogeneous fluids, which
-%% computes $g^{(2)}(\rr_1,\rr_2)$ among other
-%% things\cite{lado2009efficient}.  Lado's work looks pretty awkward to
-%% use, and their example is limited to 1D systems.  And the scaling is
-%% still (so far as I can tell) $O(N^2)$, since their improvement is in
-%% the prefactor, by developing a nice basis set.
-
-%% Gloor \emph{et al} explain that little is known of the pair
-%% distribution function of the hard-sphere fluid, and therefore use the
-%% formula
-%% \begin{align}
-%%   g^{(2)}(\rr_1,\rr_2) \approx g\left(|\rr_1-\rr_2|, \frac{\rho(\rr_1)+\rho(\rr_2)}2\right)
-%% \end{align}
-%% They go on to further approximate the radial distribution function,
-%% but we can just stop here.
-
-%% Gloor et al \cite{gloor2007prediction} uses average density of two points:
-%% \begin{align}
-%% \overline{\rho} &\equiv \frac{\rho(\rr)+\rho(\rr')}{2}
-%% \end{align}
-%% and the uses bulk radial distribution for the reference system (hard
-%% spheres) at this density, in order to approximate pair correlation.
-
-%% \section{What has been done before?}
-%% \subsection{Gloor simple method, problematic}
-%% Gil-Villegas\cite{gil1997statistical} approximates the first order
-%% inverse tenpurature Helmholtz Free energy term, which involves an
-%% integral of a potential multiplied by a radial distribution function
-%% $g^{HS}(r,\eta)$, by first taking the distribution function out of the
-%% integral and evaluating it at the correct mean distance $\zeta$,
-%% according to the mean value theorem.  They then approximate the pair
-%% correlation function $g^{HS}(\zeta,\eta)$ with the correlation at
-%% contact and at a effective $\eta_{eff}$.  The $\eta_{eff}$ is found by
-%% fitting a general function form to correct energy solutions $a_1$ with
-%% depenedent variable $\eta$ and
-
-%% \subsection{Fischer accurate but slow}

### 2013-11-22 papers/pair-correlation/paper.tex
changed fit to g(r) in paper so it fits with plot-ghs.py code, on lines 94-124 in python Tested on node09. Took 1:12.31.

@@
 where $h_\sigma = g_\sigma - 1$ is the total correlation function.
-Under these two constraints, we applied a least-squares fit to
+Under these three constraints, we applied a least-squares fit to
 simulation data for the radial distribution function at packing
@@
 \begin{multline}
-  h(r,g_\sigma) = h_\sigma e^{-\kappaa z} \\
+  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
   + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
@@
 \begin{equation}
-  I_2 = \frac{8\pi h_\sigma (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
+  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
 \end{equation}

### 2013-11-27 papers/pair-correlation/paper.tex
rewrite last section and conclusion of pair paper Tested on collins. Took 0:47.12.

@@
 \usepackage{color}
+\usepackage{subcaption} % provides subfigure
 
@@
 \begin{figure}
+  \begin{subfigure}{\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
-  %\includegraphics[width=\columnwidth]{figs/dadz-3-3.pdf}
-  %\includegraphics[width=\columnwidth]{figs/dadz-2-2.pdf}
+    \vspace{-0.8cm}
+    \caption{Sticky hard-sphere fluid}\label{fig:dadz-delta}
+  \end{subfigure}
+  \begin{subfigure}{\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
-  %%\includegraphics[width=3cm]{figs/dadz-4-3}
+    \vspace{-0.8cm}
+    \caption{Hard-core square well fluid}\label{fig:dadz-square-well}
+  \end{subfigure}
   \caption{Plot of $\frac{da_1}{dz}$ near a hard
-    wall. For what $\Phi$?}\label{fig:dadz}
+    wall.}\label{fig:dadz}
 \end{figure}
 
-The most relevant quantitative test of a pair distribution function is
-how well it predicts the interaction energy due to a pair potential.
-To this end, we have computed the error in the first term in a
-high-temperature perturbation expansion $a_1$ for several pair
+A particularly relevant quantitative test of a pair distribution
+function is how well it predicts the interaction energy due to a pair
+potential.  To this end, we have computed the error in the first term
+in a high-temperature perturbation expansion $a_1$ for several pair
 potentials.  In order to focus on effects at the interface, we have
@@
   \tfrac12 \int g^{(2)}_{HS}(\rr,\rr')n(\rr)n(\rr')\Phi(|\rr-\rr'|)
-  d\rr'\, dxdy
+  d\rr'\, dxdy\label{eq:da1}
 \end{align}
 which gives the contribution to the mean-field free energy due to
-molecules located at position $z$.  We plot in Figure~\ref{fig:dadz}
-this quantity for two pair potentials.
+molecules located at position $z$.
+
+We plot in Fig.~\ref{fig:dadz} this quantity for two pair potentials
+near a hard wall.  Figure~\ref{fig:dadz-delta} shows the results for a
+delta-function interaction at contact, which represents a so-called
+``sticky hard-sphere fluid.''  Our pair distribution function is
+constructed to produce this result exactly, provided the averaged pair
+distribution function at contact from Ref.~\citenum{schulte2012using}
+is exact.  As expected, we see excellent agreement with the Monte
+Carlo simulation results, while the approximations of Fischer and
+Sokolowski each show deviations near the interface.
+Figure~\ref{fig:dadz-square-well} shows the same curve from
+Eq.~\ref{eq:da1} for a square-well fluid.  We have chosen a square
+well fluid with a potential range of 1.79, which is the optimal value
+found for a SAFT-based model of water~\cite{clark2006developing}.
+This potential stresses our approximation to the radial distribution
+function, which results in less satisfactory results near a
+hard-sphere diameter from the wall.  We see that over these length
+scales, Sokolowski's approximation is very accurate.
 
 \section{Conclusion}
+
+We have introduced and tested an approximation for the pair
+distribution function of the inhomogeneous hard-sphere fluid.  Our
+approximation is suitable for use in the development of classical
+density functionals based constructed using perturbation theory, and
+may be computed using exclusively single-center convolutions.  We have
+tested this function at a hard wall and near a single fixed hard
+sphere, and find that it gives qualitatively reasonable results.
+Tests of the pair distribution function in integrals that arise in
+thermodynamic perturbation theory suggest that our approximation gives
+good results for both short-range and long-range attractions.
 

### 2013-12-06 papers/pair-correlation/paper.tex
add 1/r^6 potential to pairs paper (and update some MC data) Tested on collins. Took 1:29.90.
clean up da1dz text a bit, and use higher resolution in plots Tested on collins. Took 2:19.60.

@@
   \end{subfigure}
+  \begin{subfigure}{\columnwidth}
+    \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}
+    \vspace{-0.8cm}
+    \caption{Hard-core inverse-sixth fluid}\label{fig:dadz-inverse-sixth}
+  \end{subfigure}
   \caption{Plot of $\frac{da_1}{dz}$ near a hard
@@
 
-We plot in Fig.~\ref{fig:dadz} this quantity for two pair potentials
-near a hard wall.  Figure~\ref{fig:dadz-delta} shows the results for a
-delta-function interaction at contact, which represents a so-called
-``sticky hard-sphere fluid.''  Our pair distribution function is
-constructed to produce this result exactly, provided the averaged pair
-distribution function at contact from Ref.~\citenum{schulte2012using}
-is exact.  As expected, we see excellent agreement with the Monte
-Carlo simulation results, while the approximations of Fischer and
-Sokolowski each show deviations near the interface.
-Figure~\ref{fig:dadz-square-well} shows the same curve from
-Eq.~\ref{eq:da1} for a square-well fluid.  We have chosen a square
-well fluid with a potential range of 1.79, which is the optimal value
-found for a SAFT-based model of water~\cite{clark2006developing}.
+We plot in Fig.~\ref{fig:dadz} this quantity for hard-core fluids with
+three representative pair potentials near a hard wall.  We have chosen
+to illustrate a delta-function ineraction at contact (i.e. ``sticky
+hard spheres''), a hard-core square-well fluid, with the length-scale
+of interaction taken from the optimal SAFT model for water found by
+Clark \emph{et al.}~\cite{clark2006developing}, and a $1/r^6$
+power-law attraction.  These pair potentials represent a short-range
+interaction, a medium-range discontinuous interaction, and a
+long-range smooth interaction.
+
+Figure~\ref{fig:dadz-delta} shows the results for the sticky
+hard-sphere fluid.  Our pair distribution function is constructed to
+produce this result exactly, provided the averaged pair distribution
+function at contact from Ref.~\citenum{schulte2012using} is exact.  As
+expected, we see excellent agreement with the Monte Carlo simulation
+results, while the approximations of Fischer and Sokolowski each show
+deviations near the interface.  Figure~\ref{fig:dadz-square-well}
+shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
 This potential stresses our approximation to the radial distribution
-function, which results in less satisfactory results near a
-hard-sphere diameter from the wall.  We see that over these length
-scales, Sokolowski's approximation is very accurate.
+function, since the edge of the attraction occurs at the dip in the
+radial distribution function at this density, which poor results for
+the bulk fluid.  We see that over these length scales, Sokolowski's
+approximation is very accurate.  Finally, in the case of the power-law
+attraction both approximations work very well.
 

### 2014-01-15 papers/water-saft/paper.tex
Fixed an error in the association term "boltz" in water papers Tested on node25. Took 0:13.18.
add new data to water paper about Kr-water g Tested on wentworth. Took 1:45.92.
small changes to water paper Tested on node34. Took 0:22.83.

@@
-\documentclass[preprint, endfloats,amsmath,amssymb,jcp]{revtex4-1}
+\documentclass[preprint,amsmath,amssymb]{revtex4-1}
 \usepackage{graphicx}% Include figure files
@@
 profile and a large decrease in the number of hydrogen bonds broken in
-the vicinity of the solutes.
+the vicinity of the solutes.  \fixme{It improves agreement with experiment...}
 \end{abstract}
@@
       \kappaassoc g^\textit{SW}_\sigma(\xx)
-  \left(e^{-\beta\epsilonassoc} - 1\right)} - 1}
+  \left(e^{\beta\epsilonassoc} - 1\right)} - 1}
   {n_\text{site}'(\rr)
     \kappaassoc g^\textit{SW}_\sigma(\xx)
-  \left(e^{-\beta\epsilonassoc} - 1\right)}, \label{eq:X}
+  \left(e^{\beta\epsilonassoc} - 1\right)}, \label{eq:X}
 \end{align}
@@
 \end{center}
-\caption{ Density profiles for a water around a single hydrophobic rod
+\caption{ Density profiles for a water around a single hard rod
   of radius 0.1~nm. The solid red profile is from the functional
@@
 
-We will first discuss the case of a single hydrophobic rod immersed in
+We will first discuss the case of a single hard rod immersed in
 water. Figure~\ref{fig:density-single-rod} shows the density profile
@@
 \end{center}
-\caption{Broken hydrogen bonds per nanometer for hydrophobic rods
+\caption{Broken hydrogen bonds per nanometer for hard rods
   immeresed in water.  The solid red line uses the functional developed
@@
 
+Finally, in order to compare with experimental results, we examined
+the hydration of Krypton.  Figure~\ref{fig:g-Kr} shows... demonstrate
+an improved agreement with experiment.  \fixme{describe this}
+
+%% \section{New stuff}
+
+%% \begin{figure}
+%%   \begin{center}
+%%     \includegraphics[width=3.5in]{figs/lj-Xe-energy}
+%%   \end{center}
+%%   \caption{Here is xenon.}
+%%   \label{fig:xe-energy}
+%% \end{figure}
+
+%% Excess free energy for single Lennard-Jones atoms in water \fixme{are
+%%   shown} and compared with simulation of other water theories
+%% \cite{paschek2004temperature}and experiment in \fixme{a figure}.
+
+%% \begin{center}
+%%   \begin{tabular}{|l|l|l|l|}
+%%     \hline
+%%     & $\epsilon ~\textrm{(kJ/mol)}$ & $\sigma ~\textrm{(\AA)}$ & OK
+%%     with reference? \\ \hline
+%%     Ne & 0.317098 & 3.1003 & \\ \hline
+%%     Ar & 0.822039 & 3.2903 & \\ \hline
+%%     Kr & 0.955831 & 3.4203 & \\ \hline
+%%     Xe & 1.077342 & 3.5703 & \\ \hline
+%%   \end{tabular}
+%% \end{center}
+
 \section{Conclusion}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=3.5in]{figs/Kr-278K-densities}
+\end{center}
+\caption{ $g$ for Kr and water at low temperature high pressure...}
+\label{fig:g-Kr}
+\end{figure}
 
@@
 density profiles, it predicts fewer broken hydrogen bonds around
-hydrophobic solutes and at aqueous interfaces.
+hard hydrophobic solutes and at aqueous interfaces.  \fixme{add in
+  improvement demonstrated...}
 
@@
 \caption{ The average number of hydrogen bonds per molecule at
-  distances from the center of hydrophobic rods of various
+  distances from the center of hard rods of various
   radii. The solid lines are the results from using the assymetric

### 2014-01-16 papers/water-saft/paper.tex
Small changes to water paper. Tested on knightley. Took 0:39.65.

@@
 %
-We examine the effect of this improvement by studying two hard solutes: a
-hard hydrophobic rod and a hard sphere.
+We examine the effect of this improvement by studying two hard solutes(a
+hard hydrophobic rod and a hard sphere) \fixme{and a Lennard-Jones solute.}
 %
@@
 profile and a large decrease in the number of hydrogen bonds broken in
-the vicinity of the solutes.  \fixme{It improves agreement with experiment...}
+the vicinity of the hard solutes.  
+%
+We find an improvement of the radial distribution for a Lennard-Jones
+Krypton atom in water.
 \end{abstract}
@@
 \end{center}
-\caption{ $g$ for Kr and water at low temperature high pressure...}
+\caption{ $g$ for Kr and water at low temperature ($5^\circ$ C) and high
+  pressure (110 bar).}
 \label{fig:g-Kr}

### 2014-01-16 papers/water-saft/figs/all-hughes-lj-atoms.pl

@@ -8,7 +8,7 @@ my $a;
 # be scheduled on big-memory nodes leaving the big ones to wait for
 # them to finish.
 foreach $a ( "Kr" ) { #"Ne", "Ar", "Kr",  "Xe" ) {
-  foreach $t ( 278, 320 ) {
+  foreach $t ( 278 ) {
     my $atom = sprintf("%s", $a);
     my $temperature = sprintf("%04.2f", $t);
 
@@ -25,7 +25,7 @@ foreach $a ( "Kr" ) { #"Ne", "Ar", "Kr",  "Xe" ) {
 #SBATCH --mem-per-cpu=$memuse
 ##SBATCH --mail-type ALL
 ##SBATCH --mail-user daveroundy\@gmail.com
-#SBATCH --output hughes-lj-$a-$t-hires.out
+#SBATCH --output hughes-lj-$a-$t.out
 
 set -ev
 

### 2014-01-16 papers/water-saft/paper.tex
Changes to water paper Tested on knightley. Took 0:14.83.

@@
 %
-We examine the effect of this improvement by studying two hard solutes(a
-hard hydrophobic rod and a hard sphere) \fixme{and a Lennard-Jones solute.}
+We examine the effect of this improvement by studying two hard solutes
+(a
+hard hydrophobic rod and a hard sphere) and a Lennard-Jones
+approximation of a krypton atom solute.
 %
@@
 %
-We find an improvement of the radial distribution for a Lennard-Jones
-Krypton atom in water.
+We find an improvement of the partial radial distribution for a
+krypton atom in water when compared with experiment.
 \end{abstract}
@@
 Finally, in order to compare with experimental results, we examined
-the hydration of Krypton.  Figure~\ref{fig:g-Kr} shows... demonstrate
-an improved agreement with experiment.  \fixme{describe this}
+the hydration of Krypton.  We use a Lennard-Jones potential with
+values $\epsilon = .9518$\fixme{units} and $\sigma = 3.42 \AA$ with mixing rules
+and values from\cite{paschek2004temperature} where the water
+parameters are based on SPC calculations. Figure~\ref{fig:g-Kr} shows a
+comparison of the krypton-oxygen partial radial
+distribution($g_{Kr-O}$) from the functional in this work and on
+created from the functional of \hughesetal\ with experiment
+\cite{bowron1998hydrophobic}.  The new functional shows an improvement
+in the height of the first peak, and the general shape of $g_{O-Kr}$
+compare to that of \hughesetal \
 
@@
 \end{center}
-\caption{ $g$ for Kr and water at low temperature ($5^\circ$ C) and high
-  pressure (110 bar).}
+\caption{ The Kr-O partial radial distribution function at low
+  temperature ($5^\circ$ C) and high pressure (110 bar) in the limit
+  of low concentration of krypton in water. The dashed blue line is
+  computed using using the functional from \hughesetal, the solid red
+  line is this work, and the black dotted line is from experiment
+  \cite{bowron1998hydrophobic}. }
 \label{fig:g-Kr}
@@
 density profiles, it predicts fewer broken hydrogen bonds around
-hard hydrophobic solutes and at aqueous interfaces.  \fixme{add in
-  improvement demonstrated...}
+hard hydrophobic solutes and at aqueous interfaces.
+%
+The improved functional does indeed show better agreement with
+experiment when used to compute the partial radial distribution
+function of a krypton atom dissolved in water.
 

### 2014-01-17 papers/water-saft/paper.tex
more (final?) changes to water paper Tested on knightley. Took 1:00.12.

@@
 Finally, in order to compare with experimental results, we examined
-the hydration of Krypton.  We use a Lennard-Jones potential with
-values $\epsilon = .9518$\fixme{units} and $\sigma = 3.42 \AA$ with mixing rules
-and values from\cite{paschek2004temperature} where the water
-parameters are based on SPC calculations. Figure~\ref{fig:g-Kr} shows a
-comparison of the krypton-oxygen partial radial
-distribution($g_{Kr-O}$) from the functional in this work and on
-created from the functional of \hughesetal\ with experiment
-\cite{bowron1998hydrophobic}.  The new functional shows an improvement
-in the height of the first peak, and the general shape of $g_{O-Kr}$
-compare to that of \hughesetal \
+the hydration of Krypton.  To describe the interaction of water with
+krypton, we use a Lennard-Jones potential with values $\epsilon =
+.9518$~kJ/mol and $\sigma = 3.42$~\AA\ calculated using the
+Lorentz-Berthelot mixing rules and the Lennard-Jones parameters for
+water from SPC/E
+calculations~\cite{paschek2004temperature}. Figure~\ref{fig:g-Kr}
+shows the krypton-oxygen partial radial distribution function
+$g_{Kr-O}(r)$, which gives the relative probability density that an
+oxygen atom resides at a distance $r$ from a krypton atom centered at
+the origin.  We present theoretical curves computed using both this
+work and the functional of \hughesetal, which we compare with
+experimental data from extended x-ray absorption fine structure
+spectroscopy~(EXAFS)~\cite{bowron1998hydrophobic}.  The new functional
+shows improved agreement with experiment in the height and position
+of the first maximum as well as the hight and position of the first
+minimum in $g_{Kr-O}(r)$ when compared with that of \hughesetal \
 

### 2014-01-17 papers/water-saft/paper.tex

--- 2014-01-16 papers/water-saft/paper.tex
@@
 values $\epsilon = .9518$\fixme{units} and $\sigma = 3.42 \AA$ with mixing rules
-and values from\cite{paschek2004temperature} where the water
+and values from \cite{paschek2004temperature} where the water
 parameters are based on SPC calculations. Figure~\ref{fig:g-Kr} shows a

### 2014-01-18 papers/pair-correlation/paper.tex

@@
   pair distribution function at contact described in
-  Section~\ref{sec:contact}.  Our pair distribution is constructed so
+  Section~\ref{sec:contact}.  Our pair distribution function is constructed so
   as to be computable with only single-site convolutions and to
@@
 been considerable advances in the development of classical density
-functionals have been based on the
+functionals based on the
 application of these theoretical approaches to inhomogeneous
@@
 The first canonical inhomogeneous configuration for the hard-sphere
-fluid is the system consisting of a hard spheres at a hard wall.  In
+fluid is the system consisting of a hard sphere at a hard wall.  In
 1986, Plischke and Henderson solved the pair distribution function of
@@
 $\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
-Methfessel is intended to the pair distribution function only at contact,
-when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
+Methfessel is intended to approximate the pair distribution function only
+at contact, when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
 diameter.
@@
 averages at the points at which the pair distribution function is
-desired, cannot be written as a single-site convolution, since the
+desired, it cannot be written as a single-site convolution, since the
 convolution kernel depends on both points.
@@
 in the contact distance limit, leaving our function with tunable
-constant perameters.  We run a simple monte-carlo simulations to
+constant perameters.  We run a simple monte-carlo simulation to
 calculate values of the homogeneous radial distribution function at
 varying distance perameter $r$.  We then tune our function's constant
-perameters by matching our function's results against the monte-carlo
+parameters by matching our function's results against the monte-carlo
 data, using a least squares fit method.  During this process it is
@@
 Starling equation to calculate $g_{\sigma}$.  This is appropriate
-since that our published $g_{\sigma}$ approaches the Carnahan Starling
+since our published $g_{\sigma}$ approaches the Carnahan Starling
 version in the homogeneous limit and it's this limit, along with the
@@
 \begin{multline}
-  g(|\rr|) = 1 + g_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1g_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
-  -x_4g_{\sigma}^2\sin(x_5(r-\sigma))e^(-x_6(r-\sigma))
+  g(|\rr|) = 1 + h_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1h_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
+  -x_4h_{\sigma}^2\sin(x_5(r-\sigma))e^{-x_6(r-\sigma)}
 \end{multline}
-The finalalized form of our pair correlation function is:
+The finalized form of our pair correlation function is:
 \begin{align}
@@
 
-In contrast to the previous work, which are based on the homogeneous
+In contrast to the previous works, which are based on the homogeneous
 radial distribution function $g(r,\eta)$ expressed as a function of
@@
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
-figure, it is apparent how the on the integral of $h(r)$ from
+figure, it is apparent that the integral of $h(r)$ from
 Eq.~\ref{eq:total-constraint} results in cancelling errors at large
-distances.  The maximum error in $g(r)$ at packing fractions of less
-than 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+distances.  The maximum error in $g(r)$ at packing fractions up to
+0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
 \rmaxerr$.
@@
 Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
-The approach of Fischer and Methfessel is only valid when the two
+The approach of Fischer and Methfessel is only defined when the two
 spheres are in contact, and is therefore only plotted on that segment
 of the path.  As an input to both previous approximations we use the
-hard sphere radial distribution function found using Monte Carlo
+hard sphere radial distribution function found with Monte Carlo
 simulation, interpolated as necessary.  We find that both previous
@@
 approaches gives a reasonable prediction, with our approach
-underestimating the oscillations in $g^(2)$, as expected based on our
+underestimating the oscillations in $g^{(2)}$, as expected based on our
 fit in Fig.~\ref{fig:radial-distribution}.
@@
   \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
     $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
@@
 approximation to $g^{(2)}$ with results from Monte Carlo simulation.
-In each figure $\rr_1$ and $\rr_2$ are held fixed, while the third
+In each figure, $\rr_1$ and $\rr_2$ are held fixed, while the third
 position $\rr_3$ is varied.  The test-particle position in each case
@@
 function, since the edge of the attraction occurs at the dip in the
-radial distribution function at this density, which poor results for
+radial distribution function at this density, with poor results for
 the bulk fluid.  We see that over these length scales, Sokolowski's

### 2014-01-21 papers/pair-correlation/paper.tex
edits in pair paper Tested on bennet. Took 2:40.42.

@@
 A final constraint is that the slope of the radial distribution
-function at contact is given to a very good approximation by
+function at contact is given to a very good approximation by \fixme{fix dimensions below}
 \begin{align}
@@
   ensure the C++ code agrees with this as well.}
+\fixme{Clean up the above equations, if possible.}
 This function includes three fitted parameters, $\kappaa$, $\kappab$,
@@
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
-figure, it is apparent that the integral of $h(r)$ from
+figure, it is apparent that the integral f $h(r)$ from
 Eq.~\ref{eq:total-constraint} results in cancelling errors at large
@@
   \end{tabular}
-  \caption{Fit parameters \fixme{... put them here and format this nicely}}
+  \caption{Fit parameters \fixme{... put them here and format this nicely}\fixme{add dimensions here}}
   \label{tab:fit-parameters}
@@
 
-\begin{figure}
+\begin{figure*}
   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
-  \caption{ \paircaption{0.1}
+  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+  \caption{ \paircaption{0.1} \fixme{Fix *ALL* plots to specify units of distance $x/R$.}
   }\label{fig:pair-distribution-1}
-\end{figure}
+\end{figure*}
 
@@
 
-\begin{figure}
+\begin{figure*}
   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
-  \caption{The triplet distribution function
+  \caption{\fixme{Put all triplets together here.  Tweak spacing, add a, b, c, etc.} The triplet distribution function
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
@@
     left.}\label{fig:triplet-contact-distribution-3}
-\end{figure}
+\end{figure*}
 \begin{figure}
@@
 this work.  The top half displays the triplet correlation function
-from Monte Carlo.  On the right, we display
+from Monte Carlo.  On the right, we display \fixme{this needs to be finished...}
 

### 2014-01-21 papers/pair-correlation/paper.tex
edit of paper (very quick) Tested on bennet. Took 0:16.12.

@@
 suitable for use in constructing a density
-functional~\cite{gloor2007prediction, jain2007modified}.
+functional~\cite{gloor2007prediction, jain2007modified}. \fixme{We need to say what a density functional is}
 
@@
 find the pair distribution function~\cite{hansen2006theory}.  This
-elegant approach lends itself to DFT, and can be used to compute and
+elegant approach lends itself to DFT \fixme{Define DFT earlier!}, and can be used to compute and
 plot the pair distribution function, but requires a full free-energy
@@
 
-The first canonical inhomogeneous configuration for the hard-sphere
+The canonical inhomogeneous configuration for the hard-sphere
 fluid is the system consisting of a hard sphere at a hard wall.  In
@@
 
-A second inhomogeneous configuration that is of interest is the
+Another inhomogeneous configuration that is of interest is the
 test-particle configuration, in which one hard sphere is fixed in the
@@
 trick---and the pair distribution function gives the triplet
-distribution function.  The triplet distribution function was computed
+distribution function.  The triplet distribution function has been computed
 by Gonz\'alez \emph{et al.} using the test-particle approach with
@@
 unsuitable for repeated evaluation within a free-energy minimization.
+\fixme{When we introduce DFT, we should talk about minimization.}
 
@@
 Methfessel is intended to approximate the pair distribution function only
-at contact, when the distance between $\rr_1$ and $\rr_2$ is a hard-sphere
+at contact, when the distance between $\rr_1$ and $\rr_2$ is the hard-sphere
 diameter.
@@
 
+\fixme{Add a paragraph here explaining the work of Schulte {\em et
+    al.}  In particular, we want to emphasize that we have an accurate
+representation for an \emph{average} of the pair distribution function
+\emph{at contact}.  Then we can cite this in the ``Our Approach''
+section.}
+
 \section{Our approach}
@@
 
-In the previous approaches mentioned above, function forms of radial
+In the approaches mentioned above, function forms of radial
 distributions (used to approximate pair correlation) were dependent
 upon independent variables $\eta$, the density, whether at single
-points or averaged over sourrounding volumes.  We can achieve more
+points or averaged over sourrounding volumes.  We can achieve greater
 accuracy by using instead a function dependent upon our $g_{\sigma}$,
 which holds more information about an inhomogeneous system
-then an average $\eta$ does.  We therefore need a function form for
+then a simple convolution of the density does.  We therefore seek a function form for
 our radial distribution that is dependent upon $g_{\sigma}$ and that
@@
 approaches gives a reasonable prediction, with our approach
-underestimating the oscillations in $g^{(2)}$, as expected based on our
+underestimating the oscillations in $g^{(2)}$, as expected based on the
 fit in Fig.~\ref{fig:radial-distribution}.
@@
 
+\fixme{The following paragraph needs to be rewritten when we've
+  reorganized the figures. (Obviously.)}
+
 Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
@@
 color plot on the left, the first and second spheres are displayed in
-grey.  On the bottom half of each left-hand plot is the triplet
+grey.  On the bottom half of figures \fixme{b,d,etc?} is the triplet
 correlation function as computed using the approximation presented in
-this work.  The top half displays the triplet correlation function
-from Monte Carlo.  On the right, we display \fixme{this needs to be finished...}
+this work.  The top half of those figures displays the triplet
+correlation function from Monte Carlo.  Figures~\fixme{a,c,e?} show
+the triplet distribution function along the paths illustrated in the
+corresponding 2D plots.
+
+Figures~\fixme{which} and~\fixme{which} show the triplet distribution
+function when two of the positions $\rr_1$ and $\rr_2$ are in contact.
+The triplet distribution function predicted by this work is slightly
+asymmetric, indicating a failure in thermodynamic consistency.  Once
+again, we observe the lack of oscillations at large distances.
+Figures~\fixme{which} and~\fixme{which} show the triplet distribution
+function when two of the positions $\rr_1$ and $\rr_2$ are spaced so
+that the third could just fit between them.  Again, the triplet
+distribution function predicted by this work is slightly asymmetric,
+and lacks the oscillations seen in the Monte Carlo simulations at
+large distances.  In both cases, our agreement with simulation is
+comparable to that of previous approximations.
 

### 2014-01-22 papers/pair-correlation/paper.tex
Added new 2d pair data, and updated the pair plots. I have changed the orientation so that the 2d plot is on the left, rather than above the 1d plot. Waiting on consensus for this layout to make final tweaks. Tested on paho-arch. Took 1:22.36.
More pair-correlation plot tweaks. Tested on paho-arch. Took 0:51.95.

@@
 \begin{figure*}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-1.pdf}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+  \centering
+  \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-1.pdf}\\
+  \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-3.pdf}
   \caption{ \paircaption{0.1} \fixme{Fix *ALL* plots to specify units of distance $x/R$.}
@@
 %% \end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
-  \caption{\paircaption{0.3}}\label{fig:pair-distribution-3}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{figs/pair-correlation-pretty-3.pdf}
+%   \caption{\paircaption{0.3}}\label{fig:pair-distribution-3}
+% \end{figure}
 %% \begin{figure}
@@
 \begin{figure*}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
+  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}\\
+  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
   \caption{\fixme{Put all triplets together here.  Tweak spacing, add a, b, c, etc.} The triplet distribution function
@@
 \end{figure*}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted line on the plot to the
-    left.}\label{fig:triplet-contact-distribution-4}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
+%   \caption{The triplet distribution function
+%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
+%     $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+%     shows the triplet distribution function, with white corresponding
+%     to the contact value of the \emph{pair} distribution function.
+%     The plot on the right shows the value of $g^{(3)}$ along the path
+%     shown as a dotted line on the plot to the
+%     left.}\label{fig:triplet-contact-distribution-4}
+% \end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
-  \caption{The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-    separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
-\end{figure}
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
+%   \caption{The triplet distribution function
+%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
+% \end{figure}
+% \begin{figure}
+%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
+%   \caption{The triplet distribution function
+%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
+%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
+% \end{figure}
 

### 2014-01-22 papers/pair-correlation/paper.tex
Tweaks to the dadz plots for pair-correlation. Tested on paho-arch. Took 0:30.27.

@@
   \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-3.pdf}
-  \caption{ \paircaption{0.1} \fixme{Fix *ALL* plots to specify units of distance $x/R$.}
+  \caption{ \paircaption{0.1} \fixme{Do we want to label the y-axis for the 1d plots?}
   }\label{fig:pair-distribution-1}
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Put all triplets together here.  Tweak spacing, add a, b, c, etc.} The triplet distribution function
+  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still have two curves for this work, do we want to remove one?} The triplet distribution function
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
@@
   \caption{Plot of $\frac{da_1}{dz}$ near a hard
-    wall.}\label{fig:dadz}
+    wall.\fixme{Do we want y-axis ticks or no? Also, why does the data start
+      at $z = 0$?}}
+  \label{fig:dadz}
 \end{figure}

### 2014-01-23 papers/pair-correlation/paper.tex
More pair plot tweaks Tested on paho-arch. Took 1:04.43.

@@
 
-\newcommand\paircaption[1]{ The pair distribution function near a hard
-  wall, with a packing fraction of #1, with $\rr_1$ in contact with
+\newcommand\paircaption{The pair distribution function near a hard
+  wall, with packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with
   the hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$
-  as $\rr_2$ varies.  The top half of this figure shows the result of
+  as $\rr_2$ varies.  The top half of this figure shows the result of the
   Monte Carlo simulation, while the bottom half shows the result of
@@
   the left.  This plot compares the Monte Carlo results (black) with
-  those of this work (color?), Sokolowski and Fischer
-  (color?)~\cite{sokolowski1992role}, and Fischer and Methfessel
-  (color?)~\cite{fischer1980born}.  The latter is only plotted at
+  those of this work (blue), Sokolowski and Fischer
+  (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
+  (green)~\cite{fischer1980born}.  The latter is only plotted at
   contact, where it is defined.  }
@@
 \begin{figure*}
-  \centering
-  \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-1.pdf}\\
+  \begin{subfigure}{\textwidth}
+    \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-1.pdf}
+    \vspace{-0.6cm}
+    \caption{Pair distribution function near a hard wall with $\eta = 0.1$.}
+    \label{fig:pair-distribution-1}
+  \end{subfigure}
+  \begin{subfigure}{\textwidth}
     \includegraphics[width=\linewidth]{figs/pair-correlation-pretty-3.pdf}
-  \caption{ \paircaption{0.1} \fixme{Do we want to label the y-axis for the 1d plots?}
-  }\label{fig:pair-distribution-1}
+    \vspace{-0.6cm}
+    \caption{Pair distribution function near a hard wall with $\eta = 0.3$.}
+    \label{fig:pair-distribution-3}
+  \end{subfigure}
+  \caption{ \paircaption \fixme{Do we want to label the y-axis for the 1d plots?}
+  }\label{fig:pair-distribution}
 \end{figure*}

### 2014-01-24 papers/pair-correlation/paper.tex
respond to Paho questions about figures Tested on bennet. Took 5:08.12.

@@
   \end{subfigure}
-  \caption{ \paircaption \fixme{Do we want to label the y-axis for the 1d plots?}
+  \caption{ \paircaption \fixme{Do we want to label the y-axis for the
+      1d plots? } \davidsays{I think so...}
   }\label{fig:pair-distribution}
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still have two curves for this work, do we want to remove one?} The triplet distribution function
+  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still
+      have two curves for this work, do we want to remove one?
+      \davidsays{I'm not sure, I might be tempted to keep both,
+        although if we do we should explain the difference between
+        them. (left-right mirror image)}} The triplet distribution function
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
@@
   \end{subfigure}
-  \caption{Plot of $\frac{da_1}{dz}$ near a hard
-    wall.\fixme{Do we want y-axis ticks or no? Also, why does the data start
-      at $z = 0$?}}
+  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.\fixme{Do we
+      want y-axis ticks or no? Also, why does the data start at $z =
+      0$? \davidsays{I think ticks on the vertical axis aren't needed
+        because we are not including the scale (which depends on
+        prefactors).  I think data starting at $z=0$ is a result of a
+        (poor?) choice of where to define $z=0$.  It's the same
+        definition we use in Fig. 2, so I think the two should stay
+        consistent...}}}
   \label{fig:dadz}

### 2014-01-26 talks/polyhedra/slides.tex
Some old, minor changes to polyhedra files that I had been neglecting. Tested on paho-arch. Took 0:27.56.

@@
-\documentclass[handout]{beamer}
-%\usepackage{beamerthemelined}
+\documentclass[]{beamer}
+% \usepackage{beamerthemelined}
 \usepackage{pstricks}
@@
 \usepackage{graphicx}
-\usepackage[draft]{animate}
+\usepackage[]{animate}
 \usepackage{wallpaper}
-%\setbeamertemplate{navigation symbols}{}
+% \setbeamertemplate{navigation symbols}{}
 \beamertemplatenavigationsymbolsempty
@@
 \begin{document}
+% slide 1: title
 {
@@
       Dr. David Roundy, Department of Physics
+      \vspace{1.5in}
     \end{center}
@@
 }
+
+% slides 2,3: background / application
+
 \begin{frame}
-  \frametitle{Water}
+  \frametitle{Long Term Goals}
+  In addition to wanting a better understanding of water, many important things happen in contact with water
   \begin{itemize}
-  \item<1-> Water is arguably the most important fluid in existence, and yet it has some very strange properties
-  \item<2-> There currently exist no accurate continuum models for water at melting
+  \item Rust
+  \item Surfactants, e.g. soap
+  \item Acid-base reactions
+  \item Cells
+  \end{itemize}\vspace{.5in}
+  Final goal: to efficiently compute properties of systems involving water
+\end{frame}
+
+\begin{frame}
+  \frametitle{Goals of this Project}
+  \begin{itemize}\itemsep2ex
+  \item Theory for liquids not well understood
+  \item Almost exclusively based on hard spheres
+    \begin{itemize}
+    \item Hard sphere fluid: Spherical particles with no interaction except that they cannot overlap
     \end{itemize}
+  \item Spheres want 12 neighbors, water molecules want 4
+  \end{itemize}\vspace{.25in}
+  Explore the properties of various hard polyhedra fluids, in particular truncated tetrahedra
+
+  \begin{figure}[b]
+    \includegraphics[width=60mm]{figs/ice-water.png}
+  \end{figure}
+
+\end{frame}
+
+% slide 4: ice structure image
+\begin{frame}
+  \frametitle{Why Truncated Tetrahedra}
+  \begin{columns}[T,totalwidth=\textwidth]
+    \column{.2\linewidth}
     \begin{figure}[h]
-    \includegraphics[width=80mm]{figs/ice-water.png}
+      \includegraphics[width=30mm]{figs/tet-0}\\
+      \includegraphics[width=30mm]{figs/tet-1}\\
+      \includegraphics[width=30mm]{figs/tet-2}
+    \end{figure}
+    \column{.8\linewidth}
+    \begin{overprint}
+      \begin{figure}[t]
+        \centering
+        \includegraphics<1>[width=90mm]{figs/ice-structure-0}
+        \includegraphics<2>[width=90mm]{figs/ice-structure-1}
+        \includegraphics<3>[width=90mm]{figs/ice-structure-2}
+        \includegraphics<4>[width=90mm]{figs/ice-structure-3}
+        \includegraphics<5>[width=90mm]{figs/ice-structure-4}
+      \end{figure}
+    \end{overprint}
+  \end{columns}
+\end{frame}
+
+% slide 5: mc, explanation
+
+\begin{frame}
+  \frametitle{Simple 2d Fluid Between Parallel Walls}
+  \begin{columns}[T,totalwidth=\textwidth]
+    \column{.65\linewidth}
+  \begin{figure}[h]
+    \centering
+    \animategraphics[width=\columnwidth]{.6}{anim/mc-slow-}{000}{029}
+  \end{figure}
+  \column{.35\linewidth}
+  Monte Carlo Method:
+  \begin{itemize}
+  \item Average of many configurations
+  \item Move and rotate to get from one configuration to the next
+  \item Reject any moves that overlap
+  \end{itemize}
+  \end{columns}
+\end{frame}
+
+% slide 6: mc example with density / angles
+\begin{frame}
+  \frametitle{Hard Triangle Fluid Between Parallel Walls}
+  \begin{figure}[h]
+    \centering
+    \animategraphics[width=100mm,final]{.6}{anim/mc100-0.50-}{000}{029}
+  \end{figure}
+\end{frame}
+
+% slides 7-8: similar to slide 6, but with 3d cubes, fluid and then solid
+\begin{frame}
+  \frametitle{Hard Cube Fluid Between Parallel Walls}
+  \begin{figure}[h]
+    \centering
+    \includegraphics[width=100mm]{figs/cube-density-30}
   \end{figure}
@@
 \begin{frame}
-  \frametitle{Background}
-  Why Hard Polyhedra?
-  \begin{itemize}
-  \item<1-> Hard sphere fluids have been studied extensively
-  \item<2-> Hard polyhedra have little known theory
-  \item<3-> We expect to be able to match the crystal structure of ice with truncated tetrahedra
-  \item<4-> Other polyhedra may show interesting characteristics
-  \end{itemize}
+  \frametitle{Hard Cube Solid Between Parallel Walls}
   \begin{figure}[h]
-    \raggedleft
-    \includegraphics[width=60mm]{figs/Cryst_struct_ice}
+    \centering
+    \includegraphics[width=100mm]{figs/cube-density-60}
   \end{figure}
 \end{frame}
+
 \begin{frame}
-  \frametitle{Classical Density Functional Theory}
+  \frametitle{Fluid and Solid Truncated Tetrahedra}
+  \begin{figure}[h]
+    \centering
+\animategraphics[width=55mm,autoplay,loop,final]{1}{../../papers/polyhedra/figs/anim/periodic-0.42-truncated_tetrahedron-216-}{0}{9} \animategraphics[width=55mm,autoplay,loop,final]{1}{../../papers/polyhedra/figs/anim/periodic-0.71-truncated_tetrahedron-216-}{0}{9}
+  \end{figure}
+\end{frame}
+
+% slide 10: conclusion
+\begin{frame}
+  \frametitle{Conclusions}
+  \begin{itemize}\itemsep2ex
+  \item Wrote a program for simulating hard polyhedra
+  \item Constructed ice with truncated tetrahedra
+  \item Observed the solid and fluid phases of truncated tetrahedra
+  \item Future work:
     \begin{itemize}
-  \item<1-> A functional is simply a function of a function
-  \item<2-> Classical DFT provides a continuum approach for fluids
-  \item<3-> Helmholtz Free Energy is minimized at equilibrium
-    \begin{align*}
-      F = U - TS
-    \end{align*}
+    \item Explore correlations between truncated tetrahedra
+    \item Examine angular dependence in truncated tetrahedra
+    \item Observe phase transitions in truncated tetrahedra
+    \end{itemize}
+
   \end{itemize}
@@
 \begin{frame}
-  \frametitle{Monte Carlo Example}
-  \begin{figure}[h]
-    \centering
-    \animategraphics[width=100mm,autoplay]{.6}{anim/mc-slow-}{000}{200}
-  \end{figure}
-
+  \frametitle{Acknowledgements}
+  \begin{itemize}\itemsep2ex
+  \item Undergraduate Research, Innovation, Scholarship, \& Creativity (URISC)
+  \item Dr. David Roundy
+  \item Research Group: Jeff Schulte, Eric Krebs, Rene Zeto, Sam Loomis, Dan Roth, Daniel Gluck
+  \end{itemize}
 \end{frame}
 
-\begin{frame}
-  \frametitle{Monte Carlo Example}
-  \begin{figure}[h]
-    \centering
-    \animategraphics[width=120mm,autoplay]{3}{anim/mc100-0.50-}{000}{999}
-  \end{figure}
-
-\end{frame}
 \end{document}

### 2014-01-26 talks/polyhedra/mc-tri-slow.py

@@ -28,7 +28,7 @@ area = (sqrt(3)/2*l)*(l/2)/2*6
 lenx = 20
 leny = 10
 xedge = 1.5
-yedge = 4.5
+yedge = 6
 
 n = int(ff*lenx*leny/area)
 

### 2014-01-30 papers/square-well-liquid/paper.tex

--- 2013-11-19 papers/renormalization/paper.tex
@@
 \usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
 \usepackage{color}
 
-\newcommand{\fixme}[1]{{\bf\color{red}{[#1]}}}
-\newcommand{\rr}{\mathbf{r}}
+\begin{document}
+\title{Classical density functional theory for the square-well liquid}
 
-\begin{document}
-\title{Renormalization}
-
-\author{Daniel ?. Roth}
-\author{Michael ?. Perlin}
+\author{Student One}
+\author{Student Two}
 \author{David Roundy}
@@
 \begin{abstract}
-  We present an analysis of the convergence properties of
-  renormalization group theory  when applied to the square-well
-  liquid.
+  We have developed a functional.
 \end{abstract}
@@
 
-Long-wavelength fluctuations lead to significant changes in fluid
-behavior in two scenarios: near the critical point, and at
-liquid-vapor interfaces.  The failing of perturbation theory near the
-critical point is well-understood, and is commonly treated in the
-homogeneous case using either renormalization group
-theory~\cite{white2000global, white2001global, del2002vapour,
-  kiselev2002computer, reiner2002hierarchical, mi2004renormalization,
-  mi2004improved, fu2006study, giacometti2009liquid, jiuxun2005simple,
-  forte2011application, el2008integral, ramana2012generalized}, or
-crossover theory~\cite{kiselev1999crossover, kiselev2000crossover,
-  kiselev2001crossover, kiselev2000simplified, hu2003crossover,
-  hu2003back, mccabe2004crossover, llovell2006global}.  The
-liquid-vapor interface is beset by long-wavelength fluctuations in the
-form of capillary waves~\cite{buff1965interfacial,
-  weeks1989consistency}, which are handled---when treated at all---as
-an analytic correction to the surface tension computed using density
-functional theory~\cite{wadewitz2000application, winkelmann2001liquid,
-  gross2009density}.  These corrections to the surface tension involve
-an ad hoc cutoff wavelength, and do not correct the prediction of the
-interfacial density profile.  The effect of capillary waves on surface
-tension is largest near the critical point.  Together, these effects
-suggest that a density functional theory directly incorporating
-long-wavelength fluctuations would be a major gain for the study of
-liquid interfaces.
-
-I plan to develop such a functional using renormalization group theory
-(RGT).  Implementation of RGT in an inhomogeneous system will be
-computationally challenging, because it is an inherently recursive
-theory.  This results in long-range fluctuations (and correlations)
-near the critical point, which will be reflected in a long-range
-density functional.  It will be more computationally
-intensive than existing density functionals, but at the same time will
-enable far greater predictive power, over a wide range of
-temperatures, pressures, and density distributions.
-
-There has been one study of the effect of critical behavior on surface
-tension using a renormalization approach, which used the Density
-Gradient Theory (DGT) rather than DFT due to its easier integration
-with renormalization group theory~\cite{fu2006study}.  The use of DGT,
-which is not self-consistent, puts a major limitation the power of the
-method, particularly when applied to diverse interfaces.
-
-We will study a hard-sphere fluid with a square well attractive
-potential.  This is a widely used model fluid within the SAFT
-community~\cite{mi2004renormalization, mi2004improved, fu2006study,
-  forte2011application} as well as in the study of the critical
-behavior of fluids~\cite{white2000global, white2001global,
-  del2002vapour, kiselev2002computer, reiner2002hierarchical,
-  mi2004renormalization, mi2004improved, fu2006study,
-  giacometti2009liquid, jiuxun2005simple, forte2011application,
-  el2008integral, ramana2012generalized}. This model is particularly
-suitable for testing over a wide range of temperatures and pressures
-due to its ease of computation.  Because the set of energy levels is
-discrete, this system is suitable for the use of histogram-based Monte
-Carlo methods~\cite{ferrenberg1988histogram, lee1993new, de1998broad}.
-These approaches accelerate the computation of thermodynamic
-properties over a range of temperatures, and are particularly helpful
-near phase transitions and critical points.  There has been a recent
-Monte Carlo study of phase behavior of the the square-well
-fluid~\cite{liu2005direct}, which did not take advantage of the
-histogram family of Monte Carlo methods, although it did use the
-aggregation volume bias sampling method to accelerate sampling of
-bonded and non-bonded configurations~\cite{chen2000novel}.
-
-We will begin by studying the simple square-well liquid already
-introduced in a previous paragraph.  This liquid is a standard test
-case for theories of critical behavior, although its critical behavior
-at the liquid-vapor interface has not yet been examined.  We will
-closely follow the approach recently developed by Forte \emph{et al.}
-who constructed a SAFT-VR equation of state using
-RGT~\cite{forte2011application}.  Their approach to RGT is expressed
-in terms of the pair distribution function
-$g^{(2)}_{HS}(\rr_1,\rr_2)$, which will enable us to leverage our
-earlier work (from the previous paragraph) to construct a
-renormalization group classical density functional theory for the
-inhomogeneous fluid.  Once we have developed a RGT density functional,
-we will test it against Monte Carlo simulations of the surface
-tension, as well as a fluid near the critical point at a hard wall.
-%
-Future work will include applying this approach to associating fluids.
-
-\section{Methods}
-
-\subsection{Renormalization Group Theory}
-
-\fixme{Here we will explain the RGT that we use, and how it works
-  out.}
-
-\subsection{Monte Carlo}
-
-\fixme{Here we will explain our Monte Carlo simulations.}
-
-\section{Results}
-
-\section{Conclusion}
-
-\bibliography{paper}% Produces the bibliography via BibTeX.
-
 \end{document}

### 2014-02-11 papers/pair-correlation/paper.tex
incomplete changes from meeting with Jeff (see paper for reasoning) Tested on node23. Took 29:35.78.

@@
 distribution function is given by
-\newcommand\kappaa{\kappa_0}
-\newcommand\kappab{\kappa_1}
-\newcommand\kappac{\kappa_2}
+\newcommand\kappaO{c_0}
+\newcommand\kappaI{c_1}
+\newcommand\kappaZ{c_2}
 \begin{multline}
-  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
-  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
+  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  + h_\sigma(\kappaO - g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+  + B  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
+\end{multline}
+\begin{equation}
+  B = \kappaZ^5 \frac{
+    \frac{\chi-1}{24\eta} - h_\sigma \frac{(\kappaO-2
+      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
+    - h_\sigma\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
+  }{2 \kappaZ^2 + 12 \kappaZ + 24}
+\end{equation}
+\begin{equation}
+  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
+\end{equation}
+\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
+    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
+\begin{equation}
+  \eta = 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}
+\end{equation}
+\davidsays{The above is a sketch of how we might want to write our
+  equations.  I think it matches up with the python now.  Below is the
+old (presumably wrong) version.}
+
+\begin{multline}
+  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO z} \\
+  + h_\sigma(\kappaO - 2 g_\sigma) \xi e^{-\kappaI \xi} \\
   + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
       \eta^4}\frac{1}{n}
-    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
+    - I_0  -I_1(\kappaO - g_\sigma)}{I_2} z^2e^{-\kappaZ z}
 \end{multline}
 \begin{equation}
-  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
+  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaO (2 + \sigma \kappaO))}{\kappaO^3}
   - \frac{4\pi}{3} \sigma^3
@@
 \begin{equation}
-  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
+  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappaI (4 + \sigma \kappaI))}{\kappaI^4}
 \end{equation}
 \begin{equation}
-  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
+  I_2 = \frac{8\pi (12 + \sigma \kappaZ (6 + \sigma \kappaZ))}{\kappaZ^5}
 \end{equation}
@@
 \fixme{Clean up the above equations, if possible.}
-This function includes three fitted parameters, $\kappaa$, $\kappab$,
-and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
+This function includes three fitted parameters, $\kappaO$, $\kappaI$,
+and $\kappaZ$, which are given in Table~\ref{tab:fit-parameters}.
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
@@
 \begin{table}
-  \begin{tabular}{cc}
-    $\kappaa$ & \kappazero \\
-    $\kappab$ & \kappaone \\
-    $\kappac$ & \kappatwo
-  \end{tabular}
+  \begin{align*}
+    \kappaO &= \kappazero \\
+    \kappaI &= \kappaone \\
+    \kappaZ &= \kappatwo
+  \end{align*}
   \caption{Fit parameters \fixme{... put them here and format this nicely}\fixme{add dimensions here}}

### 2014-02-25 papers/pair-correlation/paper.tex
clean up pairs text Tested on bennet. Took 5:27.21.

--- 2014-01-24 papers/pair-correlation/paper.tex
@@
 A final constraint is that the slope of the radial distribution
-function at contact is given to a very good approximation by \fixme{fix dimensions below}
+function at contact is given to a very good approximation by
 \begin{align}
@@
 distribution function is given by
-\newcommand\kappaa{\kappa_0}
-\newcommand\kappab{\kappa_1}
-\newcommand\kappac{\kappa_2}
+\newcommand\kappaO{\kappa_0}
+\newcommand\kappaI{\kappa_1}
+\newcommand\kappaZ{\kappa_2}
 \begin{multline}
-  h(r,g_\sigma) = 1 + h_\sigma e^{-\kappaa z} \\
-  + h_\sigma(\kappaa - g_\sigma) ze^{-\kappab z} \\
-  + \frac{\frac{(1-\eta)^4}{1+4\eta + 4\eta^2 - 4\eta^3 +
-      \eta^4}\frac{1}{n}
-    - I_0  -I_1(\kappaa - g_\sigma)}{I_2} z^2e^{-\kappac z}
+  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
 \end{multline}
 \begin{equation}
-  I_0 = \frac{4\pi h_\sigma (2 + \sigma \kappaa (2 + \sigma \kappaa))}{\kappaa^3}
-  - \frac{4\pi}{3} \sigma^3
+  I(g_\sigma) = \kappaZ^5 \frac{
+    \frac{\chi-1}{24\eta} - h_\sigma \frac{(\kappaO-2
+      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
+    - h_\sigma\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
+  }{2 \kappaZ^2 + 12 \kappaZ + 24}
 \end{equation}
 \begin{equation}
-  I_1 = \frac{4\pi h_\sigma (6 + \sigma \kappab (4 + \sigma \kappab))}{\kappab^4}
+  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
 \end{equation}
+\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
+    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
 \begin{equation}
-  I_2 = \frac{8\pi (12 + \sigma \kappac (6 + \sigma \kappac))}{\kappac^5}
+  \eta = 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}
 \end{equation}
-\begin{equation}
-  z = r - \sigma
-\end{equation}
-\davidsays{Can someone check the above equations for correctness
-  against the python code, in plot-ghs2.py.  We'll want to
-  ensure the C++ code agrees with this as well.}
-\fixme{Clean up the above equations, if possible.}
-This function includes three fitted parameters, $\kappaa$, $\kappab$,
-and $\kappac$, which are given in Table~\ref{tab:fit-parameters}.
+This function includes three fitted parameters, $\kappaO =
+\kappazero$, $\kappaI = \kappaone$,
+and $\kappaZ = \kappatwo$.
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
@@
 \rmaxerr$.
-
-\begin{table}
-  \begin{tabular}{cc}
-    $\kappaa$ & \kappazero \\
-    $\kappab$ & \kappaone \\
-    $\kappac$ & \kappatwo
-  \end{tabular}
-  \caption{Fit parameters \fixme{... put them here and format this nicely}\fixme{add dimensions here}}
-  \label{tab:fit-parameters}
-\end{table}
-
-%% In addition, we know that
-%% \begin{align}
-%%   1 + n\int\left[g(r)-1\right]d\rr &= nkT\chi_T
-%% \end{align}
-%% \begin{equation}
-%%   \begin{split}
-%%     \chi_T  &= -\frac{1}{V}\left(\frac{\partial V}{\partial
-%%       p}\right)_T \\
-%%     &= - \left(V\left(\frac{\partial p}{\partial
-%%       V}\right)_T\right)^{-1}
-%%   \end{split}
-%% \end{equation}
-%% \begin{equation}
-%%     p = nkT\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}
-%% \end{equation}
-%% \begin{equation}
-%%     \left(\frac{\partial p}{\partial V}\right)_T = \frac{\partial p}{\partial\eta}\frac{\partial\eta}{\partial V} + \frac{\partial p}{\partial n}\frac{\partial n}{\partial V}
-%% \end{equation}
-%% \begin{equation}
-%%     n = \frac{N}{V}\Longrightarrow \frac{\partial n}{\partial V} = -\frac{N}{V^2} = -\frac{n}{V}
-%% \end{equation}
-%% \begin{equation}
-%%     \eta = \frac{4\pi R^3N}{3V}\Longrightarrow \frac{d\eta}{dV} = -\frac{\eta}{V}
-%% \end{equation}
-%% \begin{equation}
-%%   \begin{split}
-%%     \left(\frac{\partial p}{\partial V}\right)_T &= -nkT\frac{\eta}{V}\left[\frac{1 + 2\eta - 3\eta^2}{(1-\eta)^3} + 3\frac{1 + \eta + \eta^2 - \eta^3}{(1-\eta)^4}\right]\\
-%%     &\quad  - \frac{nkT}{V}\left[\frac{1+\eta+\eta^2-\eta^3}{(1-\eta)^3}\right]\\
-%%     &= -nkT\frac{\eta}{V}\left[\frac{(1+2\eta-3\eta^2)(1-\eta) + 3 + 3\eta + 3\eta^2 - 3\eta^3}{(1-\eta)^4}\right]\\
-%%     &\quad  - \frac{nkT}{V}\left[\frac{(1+\eta+\eta^2-\eta^3)(1-\eta)}{(1-\eta)^4}\right]\\
-%%     &= -\frac{nkT}{V}\left[\frac{4\eta + 4\eta^2 - 2\eta^3}{(1-\eta)^4} + \frac{1 - 2\eta^3 + \eta^4}{(1-\eta)^4}\right]\\
-%%   \end{split}
-%% \end{equation}
-%% \begin{equation}
-%%     nkT\chi_T = \frac{(1-\eta)^4}{1 + 4\eta + 4\eta^2 - 4\eta^3 + \eta^4}
-%%   %\int (g(r)-1)d\rr &= \frac{\chi_T}{kT} \\
-%%   %&= \eta \eta \cdots \fixme{fill in, double-check}
-%% \end{equation}
-%% where $\chi_T$ is the isothermal compressibility, for which we use the
-%% Carnahan-Starling equation of state.
-
-%% An additional constraint that we use sets the slope of $g(r,
-%% g_\sigma)$ at $r=\sigma$.  \fixme{explain this}
-
-%% Given these constraints, we use a functional form of
-%% \begin{equation}
-%%   g(r,g_\sigma) = 1 + A e^{-a(r-\sigma)}
-%%   + B (r-\sigma) e^{-b(r-\sigma)} + C (r-\sigma)^2 e^{-c(r-\sigma)}
-%% \end{equation}
-%% where $A$, $B$, and $C$ are amplitudes determined by the above
-%% constraints, and $a$, $b$, and $c$ are parameters fit to Monte Carlo
-%% simulations of the hard-sphere radial distribution function.
 
@@
   \caption{ \paircaption \fixme{Do we want to label the y-axis for the
-      1d plots? } \davidsays{I think so...}
+      1d plots? } \davidsays{I think so... also add a) b) c) d)}
   }\label{fig:pair-distribution}
@@
     shown as a dotted line on the plot to the
-    left.}\label{fig:triplet-contact-distribution-3}
+    left.}\label{fig:triplet-contact-distribution}
 \end{figure*}
-% \begin{figure}
-%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-contact-4.pdf}
-%   \caption{The triplet distribution function
-%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.4, plotted when
-%     $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-%     shows the triplet distribution function, with white corresponding
-%     to the contact value of the \emph{pair} distribution function.
-%     The plot on the right shows the value of $g^{(3)}$ along the path
-%     shown as a dotted line on the plot to the
-%     left.}\label{fig:triplet-contact-distribution-4}
-% \end{figure}
 
@@
 
-% \begin{figure}
-%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-%   \caption{The triplet distribution function
-%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-3}
-% \end{figure}
-% \begin{figure}
-%   \includegraphics[width=\columnwidth]{figs/triplet-correlation-pretty-inbetween-4.pdf}
-%   \caption{The triplet distribution function
-%     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ plotted when $\rr_1$ and $\rr_2$ are
-%     separated by the diameter of one sphere.}\label{fig:triplet-inbetween-distribution-4}
-% \end{figure}
-
-\fixme{The following paragraph needs to be rewritten when we've
-  reorganized the figures. (Obviously.)}
-
-Figures~\ref{fig:triplet-contact-distribution-3}-\ref{fig:triplet-inbetween-distribution-4}
-compare the triplet distribution function computed using our
+Figure~\ref{fig:triplet-contact-distribution} compares the triplet
+distribution function at a packing fraction of 0.3 computed using our
 approximation to $g^{(2)}$ with results from Monte Carlo simulation.
-In each figure, $\rr_1$ and $\rr_2$ are held fixed, while the third
+In each subfigure, $\rr_1$ and $\rr_2$ are held fixed, while the third
 position $\rr_3$ is varied.  The test-particle position in each case
@@
 color plot on the left, the first and second spheres are displayed in
-grey.  On the bottom half of figures \fixme{b,d,etc?} is the triplet
-correlation function as computed using the approximation presented in
-this work.  The top half of those figures displays the triplet
-correlation function from Monte Carlo.  Figures~\fixme{a,c,e?} show
-the triplet distribution function along the paths illustrated in the
-corresponding 2D plots.
+grey.  On the bottom half of
+figures~\ref{fig:triplet-contact-distribution}a and
+\ref{fig:triplet-contact-distribution}c is the triplet distribution
+function as computed using the approximation presented in this work.
+The top half of those figures displays the triplet distribution
+function from Monte Carlo.
+Figures~\ref{fig:triplet-contact-distribution}b
+and~\ref{fig:triplet-contact-distribution}d show the triplet
+distribution function along the paths illustrated in the corresponding
+2D plots.
 
-Figures~\fixme{which} and~\fixme{which} show the triplet distribution
+Figures~\ref{fig:triplet-contact-distribution}a
+and~\ref{fig:triplet-contact-distribution}b show the triplet
+distribution
 function when two of the positions $\rr_1$ and $\rr_2$ are in contact.
@@
 again, we observe the lack of oscillations at large distances.
-Figures~\fixme{which} and~\fixme{which} show the triplet distribution
+Figures~\ref{fig:triplet-contact-distribution}c
+and~\ref{fig:triplet-contact-distribution}d show the triplet distribution
 function when two of the positions $\rr_1$ and $\rr_2$ are spaced so
@@
 
-\section{Performance in thermodynamic perturbation theory}
-
 \begin{figure}
-  \begin{subfigure}{\columnwidth}
+  \begin{subfigure}{0.5\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}{\columnwidth}
+  \begin{subfigure}{0.5\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}{\columnwidth}
+  \begin{subfigure}{0.5\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}
@@
   \end{subfigure}
-  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.\fixme{Do we
-      want y-axis ticks or no? Also, why does the data start at $z =
-      0$? \davidsays{I think ticks on the vertical axis aren't needed
-        because we are not including the scale (which depends on
-        prefactors).  I think data starting at $z=0$ is a result of a
-        (poor?) choice of where to define $z=0$.  It's the same
-        definition we use in Fig. 2, so I think the two should stay
-        consistent...}}}
+  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.  (a) shows a
+    sticky hard-sphere fluid defined by a pair potential
+    $\delta(\sigma-r+\delta)$ where $\sigma$ is the hard-sphere
+    diameter, and $\delta$ is an infinitesimal distance; (b) shows a
+    square well fluid defined by a pair potential $\Theta(1.79
+    \sigma-r)$; and (c) shows a hard-core inverse-sixth potential
+    fluid with an attractive pair potential proportional to $r^{-6}$.
+  }
   \label{fig:dadz}
 \end{figure}
+
+\section{Performance in thermodynamic perturbation theory}
 

### 2014-02-25 papers/pair-correlation/paper.tex
edit pair paper Tested on bennet. Took 0:14.15.

@@
 hard-sphere reference fluid with attractive interactions that are
-treated perturbatively~\cite{hansen2006theory}.  Recently there have
-been considerable advances in the development of classical density
-functionals based on the
-application of these theoretical approaches to inhomogeneous
-densities~\cite{ jain2007modified, gloor2007prediction, gross2009density,
+treated perturbatively~\cite{hansen2006theory}.  Recent advances have
+extended these perturbative approaches to inhomogeneous density
+distributions---that is, liquid interfaces---through the use of
+classical density functional theory (DFT), in which the grand free
+energy is found by minimizing a free energy functional of the
+density~\cite{ jain2007modified, gloor2007prediction,
+  gross2009density,
   %felipe2001examination, gloor2002saft,
@@
 suitable for use in constructing a density
-functional~\cite{gloor2007prediction, jain2007modified}. \fixme{We need to say what a density functional is}
+functional~\cite{gloor2007prediction, jain2007modified}.
 
@@
 find the pair distribution function~\cite{hansen2006theory}.  This
-elegant approach lends itself to DFT \fixme{Define DFT earlier!}, and can be used to compute and
+elegant approach lends itself to DFT, and can be used to compute and
 plot the pair distribution function, but requires a full free-energy
@@
 unsuitable for repeated evaluation within a free-energy minimization.
-\fixme{When we introduce DFT, we should talk about minimization.}
 
@@
 
-Approaches that do not diverge use an average of the density over some
+Non-divergent approaches use an average of the density over some
 volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
@@
 
-For these reasons Sokolowski and Fischer modified this approach to use
-averages centered on the two points $\rr_1$ and $\rr_2$:
+Sokolowski and Fischer addressed the shortcomings of the theory of
+Fischer and Methfessel by modifying this approach to use
+averages, centered on the two points $\rr_1$ and $\rr_2$:
 \begin{align}
@@
 
-\fixme{Add a paragraph here explaining the work of Schulte {\em et
-    al.}  In particular, we want to emphasize that we have an accurate
-representation for an \emph{average} of the pair distribution function
-\emph{at contact}.  Then we can cite this in the ``Our Approach''
-section.}
+In a previous paper~\cite{schulte2012using}, we have introduced a
+functional that gives a good approximation for the pair distribution
+function averaged over positions $\rr_2$ that are in contact with
+$\rr_1$, defined as:
+\begin{align}
+  g_\sigma(\rr_1) &\equiv \frac{ \int g^{(2)}(\rr_1,\rr_2) \delta(\sigma -|\rr_1-\rr_2|)n(\rr_2)
+    d\rr_2 }{ \tilde{n}(\rr_1)  }
+\end{align}
+where the averaged density $\tilde{n}(\rr_1)$ is defined by
+\begin{align}
+  \tilde{n}(\rr_1) &\equiv \int \delta(\sigma -|\rr_1 - \rr_2|)n(\rr_2) d\rr_2.
+\end{align}
+We use the contact-value theorem to derive the exact formula:
+\begin{align}
+  g_\sigma(\rr)% &= \frac{\ncontact(\rr)}{n_A(\rr)} \\
+  &= \frac12 \frac{1}{k_BT n(\rr) \tilde{n}(\rr)} \frac{\delta
+    A}{\delta \sigma(\mathbf{r})} \label{eq:gsigma-exact}
+\end{align}
+where $\sigma(\rr)$ is the diameter of hard spheres located at
+position $\rr$, and $A$ is the Helmholtz free energy of the
+hard-sphere fluid.  Our approximation is obtained by using the White
+Bear functional for Helmholtz free energy of the hard-sphere
+fluid~\cite{roth2002whitebear} in Eq.~\ref{eq:gsigma-exact}.  This
+provides an excellent approximation for this averaged value of the
+pair distribution function for a variety of interfaces, and over a
+wide range of densities.
 
@@
 distribution functions, evaluated at the distance between the two
-points, that are themselves functions of our published pair
-correlation at contact $g_{\sigma}(\rr)$ evaluated at the two
-points~\cite{schulte2012using}.
+points, that are themselves functions of the averaged pair correlation
+at contact $g_{\sigma}(\rr)$ discussed above~\cite{schulte2012using},
+evaluated at the two points.
 \begin{align}
@@
 
-In the approaches mentioned above, function forms of radial
-distributions (used to approximate pair correlation) were dependent
-upon independent variables $\eta$, the density, whether at single
-points or averaged over sourrounding volumes.  We can achieve greater
-accuracy by using instead a function dependent upon our $g_{\sigma}$,
-which holds more information about an inhomogeneous system
-then a simple convolution of the density does.  We therefore seek a function form for
-our radial distribution that is dependent upon $g_{\sigma}$ and that
-will approach the correct solution in the homogeneous limit and also
-in the limit of contact distance between the two points.
+In the approaches for the pair distribution function mentioned above,
+function forms of radial distributions (used to approximate pair
+correlation) were dependent upon independent variables $\eta$, the
+density, whether at single points or averaged over sourrounding
+volumes.  We can achieve greater accuracy by using instead a function
+dependent upon our $g_{\sigma}$, which holds more information about an
+inhomogeneous system then a simple convolution of the density does.
+We therefore seek a function form for our radial distribution that is
+dependent upon $g_{\sigma}$ and that will approach the correct
+solution in the homogeneous limit and also in the limit of contact
+distance between the two points.
 
@@
 The function has the form:
+\davidsays{fix below to match next section... or should we just delete it,
+and express the concept in words? It seems maybe redundant.}
 \begin{multline}
-  g(|\rr|) = 1 + h_{\sigma}e^{-x_0(r-\sigma)} \\+ x_1h_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
+  g(|\rr|) = 1 + h_{\sigma}e^{-\kappa_0\left(\frac{r}{\sigma}-1\right)} \\+ x_1h_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
   -x_4h_{\sigma}^2\sin(x_5(r-\sigma))e^{-x_6(r-\sigma)}

### 2014-03-04 papers/pair-correlation/paper.tex
Added abcd, etc. to figures in pair paper. Tested on paho-arch. Took 0:49.00.

@@
     \vspace{-0.6cm}
-    \caption{Pair distribution function near a hard wall with $\eta = 0.1$.}
-    \label{fig:pair-distribution-1}
   \end{subfigure}
@@
     \vspace{-0.6cm}
-    \caption{Pair distribution function near a hard wall with $\eta = 0.3$.}
-    \label{fig:pair-distribution-3}
   \end{subfigure}
-  \caption{ \paircaption \fixme{Do we want to label the y-axis for the
-      1d plots? } \davidsays{I think so... also add a) b) c) d)}
-  }\label{fig:pair-distribution}
+  \caption{\paircaption}
+  \label{fig:pair-distribution}
 \end{figure*}
@@
 with a focus on the case where one of the two spheres is in contact
-with the hard wall.  Figure~\ref{fig:pair-distribution-1}a
-and~\ref{fig:pair-distribution-3}a compare the results of our
+with the hard wall.  Figure~\ref{fig:pair-distribution}a
+and~\ref{fig:pair-distribution}a compare the results of our
 approximate functional with Monte Carlo simulation at packing
@@
 
-Figure~\ref{fig:pair-distribution-1}b
-and~\ref{fig:pair-distribution-3}b show the pair distribution function
+Figure~\ref{fig:pair-distribution}b
+and~\ref{fig:pair-distribution}b show the pair distribution function
 as plotted along a path illustrated in
-Fig.~\ref{fig:pair-distribution-1}a.  These plots compare this work
+Fig.~\ref{fig:pair-distribution}a.  These plots compare this work
 with Monte Carlo results, as well as the approximations of Sokolowski
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Tweak spacing, add a, b, c, etc. The 1d plots still
-      have two curves for this work, do we want to remove one?
-      \davidsays{I'm not sure, I might be tempted to keep both,
-        although if we do we should explain the difference between
-        them. (left-right mirror image)}} The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
-    $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding
-    to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path
-    shown as a dotted line on the plot to the
+  \caption{\fixme{Remove one of, or make a note of, the two curves for this
+      work} The triplet distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at
+    packing fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.  The
+    color plot on the left shows the triplet distribution function, with white
+    corresponding to the contact value of the \emph{pair} distribution function.
+    The plot on the right shows the value of $g^{(3)}$ along the path shown as a
+    dotted line on the plot to the
     left.}\label{fig:triplet-contact-distribution}

### 2014-03-04 papers/pair-correlation/paper.tex
fix offsets Tested on bennet. Took 0:45.50.

@@
 
-\newcommand\paircaption{The pair distribution function near a hard
-  wall, with packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with
-  the hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$
-  as $\rr_2$ varies.  The top half of this figure shows the result of the
+\newcommand\paircaption{\fixme{Look into smoothing the top plot on
+    right.} The pair distribution function near a hard wall, with
+  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the
+  hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as
+  $\rr_2$ varies.  The top half of this figure shows the result of the
   Monte Carlo simulation, while the bottom half shows the result of
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Remove one of, or make a note of, the two curves for this
-      work} The triplet distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at
-    packing fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.  The
-    color plot on the left shows the triplet distribution function, with white
-    corresponding to the contact value of the \emph{pair} distribution function.
-    The plot on the right shows the value of $g^{(3)}$ along the path shown as a
-    dotted line on the plot to the
-    left.}\label{fig:triplet-contact-distribution}
+  \caption{\fixme{Remove one of, or make a note of, the two curves for
+      this work}\fixme{Fix MC issue here (volume wrong?)} The triplet
+    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing
+    fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.
+    The color plot on the left shows the triplet distribution
+    function, with white corresponding to the contact value of the
+    \emph{pair} distribution function.  The plot on the right shows
+    the value of $g^{(3)}$ along the path shown as a dotted line on
+    the plot to the left.}\label{fig:triplet-contact-distribution}
 \end{figure*}
@@
 \begin{figure}
-  \begin{subfigure}{0.5\columnwidth}
+  \begin{subfigure}{1.0\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-3-2.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}{0.5\columnwidth}
+  \begin{subfigure}{1.0\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-square-well-3.pdf}
@@
   \end{subfigure}
-  \begin{subfigure}{0.5\columnwidth}
+  \begin{subfigure}{1.0\columnwidth}
     \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}

### 2014-03-06 papers/pair-correlation/paper.tex

@@
   function of hard spheres, which is based on the average value of the
-  pair distribution function at contact described in
-  Section~\ref{sec:contact}.  Our pair distribution function is constructed so
-  as to be computable with only single-site convolutions and to
-  accurately reproduce the averaged distribution function at contact.
-  This results in a functional that is efficient to compute, and
-  predicts the pair distribution function as well as the most accurate
-  existing analytical approximation.
+  pair distribution function at contact we recently developed.  Our
+  pair distribution function is constructed so as to be computable
+  with only single-site convolutions and to accurately reproduce the
+  averaged distribution function at contact.  This results in a
+  functional that is efficient to compute, and predicts the pair
+  distribution function as well as the most accurate existing
+  analytical approximation.  \fixme{needs special attention.}
 \end{abstract}
@@
 
-The canonical inhomogeneous configuration for the hard-sphere
-fluid is the system consisting of a hard sphere at a hard wall.  In
-1986, Plischke and Henderson solved the pair distribution function of
-this system using integral equation theory under the Percus-Yevick
+The canonical inhomogeneous configuration for the hard-sphere fluid is
+the system consisting of a hard sphere at a hard wall.  In 1986,
+Plischke and Henderson solved the pair distribution function of this
+system using integral equation theory under the Percus-Yevick
 approximation~\cite{plischke1986pair}.  Ten years later,
 G{\"o}tzelmann \emph{et al.} found the pair distribution function near
-a hard wall using classical DFT by finding the direct correlation
-function by taking a functional derivative of the free energy
-functional and then solving the Ornstein-Zernicke
-equation~\cite{gotzelmann1996structure}.  The approach of solving the
-Ornstein-Zernicke equation for an inhomogeneous system is
-computationally challenging~\cite{paul2003variational}.
+a hard wall by solving the Ornstein-Zernicke equation using the direct
+correlation function found from a functional derivative of a classical
+DFT free energy functional~\cite{gotzelmann1996structure}.
+Unfortunately, solving the inhomogeneous Ornstein-Zernicke equation is
+computationally challenging, although more efficient approximate
+algorithms have been developed~\cite{paul2003variational}.
 
@@
 \section{Approximations to the pair distribution function}
-There are few analytic approximations for the inhomogeneous pair
+There are several analytic approximations for the inhomogeneous pair
 distribution function, which extend the radial distribution function
-to inhomogeneous scenarios.  These works differ in what density they
-use when evaluating the radial distribution function $g(r,\eta)$.
+to inhomogeneous scenarios.  These works differ in what density is
+used when evaluating the radial distribution function $g(r,\eta)$.
 
-Several recent works have cited a
-mean-density approximation:
+Several recent works have assumed a mean-density approximation for
+the pair distribution function:
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \frac{\pi\sigma^3}{6}\tfrac12
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \tfrac{\pi\sigma^3}{6}\tfrac12
   (n(\rr_1)+n(\rr_2))\right)
@@
 radius and packing fraction $\eta$, and $\sigma$ is the hard-sphere
-diameter~\cite{gloor2007prediction, gross2009density}.  These works
-did not however use this approach, which fails dramatically at higher
-densities, as the packing fraction found from the mean density exceeds
-unity.
+diameter~\cite{gloor2007prediction, gross2009density}.  This approach,
+however, cannot be applied to highly inhomogeneous systems such as a
+dense fluid at a solid surface.  These systems exhibit large density
+oscillations that lead to mean packing fractions greater than unity,
+which cannot occur in the bulk reference system that defines $g(r,
+\eta)$.  The above papers restricted themselves to the liquid-vapor
+interface, which does not exhibit this pathology.
 
-Non-divergent approaches use an average of the density over some
+Non-pathological approaches use an average of the density over some
 volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
@@
 where $\sigma$ is the hard-sphere diameter.  Equation~\ref{eq:fischer}
-is computationally awkward, because it relies on the midpoint
+is computationally awkward, because it treats as special the midpoint
 $\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
@@
 \section{Our approach}
+%% We approximate the pair correlation with the average of two radial
+%% distribution functions, evaluated at the distance between the two
+%% points, that are themselves functions of the averaged pair correlation
+%% at contact $g_{\sigma}(\rr)$ discussed above~\cite{schulte2012using},
+%% evaluated at the two points.
+In the approaches for the pair distribution function mentioned above,
+the radial distribution function used to approximate pair correlation
+was dependent upon the density averaged over some volume.  We seek to
+achieve greater accuracy by using instead a function dependent upon
+our averaged $g_{\sigma}(\rr)$ discussed above, which holds more
+information about an inhomogeneous system than does a simple
+convolution of the density.
+%
 We approximate the pair correlation with the average of two radial
@@
 points, that are themselves functions of the averaged pair correlation
-at contact $g_{\sigma}(\rr)$ discussed above~\cite{schulte2012using},
-evaluated at the two points.
+at contact $g_{\sigma}(\rr)$ evaluated at the two points:
+%
+%% We therefore seek a function form
+%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
+%% and that will approach the correct solution in the homogeneous limit
+%% and also in the limit of contact distance between the two points.
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}.
 \end{align}
-
-In the approaches for the pair distribution function mentioned above,
-function forms of radial distributions (used to approximate pair
-correlation) were dependent upon independent variables $\eta$, the
-density, whether at single points or averaged over sourrounding
-volumes.  We can achieve greater accuracy by using instead a function
-dependent upon our $g_{\sigma}$, which holds more information about an
-inhomogeneous system then a simple convolution of the density does.
-We therefore seek a function form for our radial distribution that is
-dependent upon $g_{\sigma}$ and that will approach the correct
-solution in the homogeneous limit and also in the limit of contact
-distance between the two points.
 
@@
 
+\newcommand\kappaO{\kappa_0}
+\newcommand\kappaI{\kappa_1}
+\newcommand\kappaZ{\kappa_2}
+
 The function has the form:
-\davidsays{fix below to match next section... or should we just delete it,
-and express the concept in words? It seems maybe redundant.}
 \begin{multline}
-  g(|\rr|) = 1 + h_{\sigma}e^{-\kappa_0\left(\frac{r}{\sigma}-1\right)} \\+ x_1h_{\sigma}\sin(x_2(r-\sigma))e^{-x_3(r-\sigma)}\\
-  -x_4h_{\sigma}^2\sin(x_5(r-\sigma))e^{-x_6(r-\sigma)}
+  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
 \end{multline}
+where $\kappaO$,$\kappaI$, and $\kappaZ$ are fitted constants and
+$I(g_\sigma)$ is a function of $g_\sigma$.
+
 The finalized form of our pair correlation function is:
@@
 \end{align}
-
 \begin{figure}
@@
 distribution function is given by
-\newcommand\kappaO{\kappa_0}
-\newcommand\kappaI{\kappa_1}
-\newcommand\kappaZ{\kappa_2}
 \begin{multline}

### 2014-03-07 papers/pair-correlation/paper.tex
more edits on paper plus reenable this-work-mc Tested on bennet. Took 28:53.21.

@@
 \begin{align}
-  F_1 &= \tfrac12 \iint g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|) d\rr_1d\rr_2
+  F_1 &= \tfrac12 \iint
+  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
+  d\rr_1d\rr_2
+  \label{eq:mean-field}
 \end{align}
@@
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}.
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) +
+    g(|r_{12}|, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
 \end{align}
+This approximation is constructed to reproduce the exact value for the
+integral$ \int n(\rr_1)n(\rr_2)
+g^{(2)}(\rr_1,\rr_2)\delta(|\rr_1-\rr_2|-\sigma)d\rr_1d\rr_2$.  This
+approximation (described in Eq.~\ref{eq:g2-our-mean}) requires the
+radial distribution function as a function of $r$ and $g_\sigma$.  One
+approach is to compute $\eta$ from $g_\sigma$ and then find
+$g(r,\eta)$ either using some approximate functional form
+(e.g. Percus-Yevick) or using Monte Carlo simulation results.  In this
+paper we use the latter (also for evaluating the approximations of
+earlier works), and refer to this approximation (with an exact
+$g(r,g_\sigma)$) as ``this work MC.''
 
-We construct our function's form in terms of an expansion of powers of
-$h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
-in the contact distance limit, leaving our function with tunable
-constant perameters.  We run a simple monte-carlo simulation to
-calculate values of the homogeneous radial distribution function at
-varying distance perameter $r$.  We then tune our function's constant
-parameters by matching our function's results against the monte-carlo
-data, using a least squares fit method.  During this process it is
-simplest to use an analytic expression for $g_{\sigma}$ in our
-function and so we use the well known and very successful Carnahan
-Starling equation to calculate $g_{\sigma}$.  This is appropriate
-since our published $g_{\sigma}$ approaches the Carnahan Starling
-version in the homogeneous limit and it's this limit, along with the
-contact limit, that we would like our constructed radial distribution
-function to get right.  After this process we have an appropriate
-radial distribution function that is dependent only on $r$ and
-$g_{\sigma}$.
+Equation~\ref{eq:g2-our-mean} is a reasonable approximation, but
+cannot be computed efficiently with single-center convolutions as
+written.  In order to ease computational application of this pair
+distribution function in classical density functionals, we have
+developed a separable appoximation for the function $g(r,g_\sigma)$.
+\begin{align}
+  g(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+\end{align}
+This separable approximation allows the mean field contribution to
+the free energy (Eq.~\ref{eq:mean-field}) to be computed with only
+single-center convolutions, which can be efficiently computed using
+fast Fourier transforms.
+
+%% We construct our function's form in terms of an expansion of powers of
+%% $h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
+%% in the contact distance limit, leaving our function with tunable
+%% constant perameters.  We run a simple monte-carlo simulation to
+%% calculate values of the homogeneous radial distribution function at
+%% varying distance perameter $r$.  We then tune our function's constant
+%% parameters by matching our function's results against the monte-carlo
+%% data, using a least squares fit method.  During this process it is
+%% simplest to use an analytic expression for $g_{\sigma}$ in our
+%% function and so we use the well known and very successful Carnahan
+%% Starling equation to calculate $g_{\sigma}$.  This is appropriate
+%% since our published $g_{\sigma}$ approaches the Carnahan Starling
+%% version in the homogeneous limit and it's this limit, along with the
+%% contact limit, that we would like our constructed radial distribution
+%% function to get right.  After this process we have an appropriate
+%% radial distribution function that is dependent only on $r$ and
+%% $g_{\sigma}$.
+
+%% The function has the form:
+%% \begin{multline}
+%%   g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+%%   + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+%%   + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
+%% \end{multline}
+%% where $\kappaO$,$\kappaI$, and $\kappaZ$ are fitted constants and
+%% $I(g_\sigma)$ is a function of $g_\sigma$.
+
+%% The finalized form of our pair correlation function is:
+%% \begin{align}
+%%   g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
+%% \end{align}
+%% \begin{figure}
+%%   \centering
+%%   \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
+%%   %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
+%%   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
+%% \end{figure}
 
@@
 \newcommand\kappaZ{\kappa_2}
-
-The function has the form:
-\begin{multline}
-  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
-  + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
-  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
-\end{multline}
-where $\kappaO$,$\kappaI$, and $\kappaZ$ are fitted constants and
-$I(g_\sigma)$ is a function of $g_\sigma$.
-
-The finalized form of our pair correlation function is:
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
-\end{align}
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
-  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
-  \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
-\end{figure}
 

### 2014-03-07 papers/pair-correlation/paper.tex

@@
 earlier works), and refer to this approximation (with an exact
-$g(r,g_\sigma)$) as ``this work MC.''
+$g(r,g_\sigma)$) as ``this work MC.''  \fixme{We should develop
+  separate acroyms for each method.}
 

### 2014-03-10 papers/pair-correlation/paper.tex
Made path plots nicer in pair-correlation paper. More changes to come. Tested on paho-arch. Took 1:17.44.

@@
 \newcommand{\jeffsays}[1]{{\color{red} [\green{Jeff:} \emph{#1}]}}
+\newcommand{\pahosays}[1]{{\color{red} [\green{Paho:} \emph{#1}]}}
 
@@
 
-\newcommand\paircaption{\fixme{Look into smoothing the top plot on
-    right.} The pair distribution function near a hard wall, with
-  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the
-  hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as
-  $\rr_2$ varies.  The top half of this figure shows the result of the
-  Monte Carlo simulation, while the bottom half shows the result of
-  the appproximation developed in this work.  On the right is a plot
-  of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the figure to
-  the left.  This plot compares the Monte Carlo results (black) with
-  those of this work (blue), Sokolowski and Fischer
+\newcommand\paircaption{\pahosays{The MC has one point that jumps up at the
+    beginning of the incontact region.} The pair distribution function near a
+  hard wall, with packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with
+  the hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$
+  varies.  The top half of this figure shows the result of the Monte Carlo
+  simulation, while the bottom half shows the result of the appproximation
+  developed in this work.  On the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on
+  the path illustrated in the figure to the left.  This plot compares the Monte
+  Carlo results (black) with those of this work (blue), Sokolowski and Fischer
   (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
-  (green)~\cite{fischer1980born}.  The latter is only plotted at
-  contact, where it is defined.  }
+  (green)~\cite{fischer1980born}.  The latter is only plotted at contact, where
+  it is defined.  }
 
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Remove one of, or make a note of, the two curves for
-      this work}\fixme{Fix MC issue here (volume wrong?)} The triplet
-    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing
-    fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.
-    The color plot on the left shows the triplet distribution
-    function, with white corresponding to the contact value of the
-    \emph{pair} distribution function.  The plot on the right shows
-    the value of $g^{(3)}$ along the path shown as a dotted line on
+  \caption{\fixme{Remove one of, or make a note of, the two curves for this
+      work}\fixme{Fix MC issue here (volume wrong?)} \fixme{Add backward curve
+      for this work mc?}\fixme{z-axis wrong}The triplet
+    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3,
+    plotted when $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
+    shows the triplet distribution function, with white corresponding to the
+    contact value of the \emph{pair} distribution function.  The plot on the
+    right shows the value of $g^{(3)}$ along the path shown as a dotted line on
     the plot to the left.}\label{fig:triplet-contact-distribution}

### 2014-03-10 papers/pair-correlation/paper.tex
Fixed a couple plot bugs Tested on paho-arch. Took 0:48.83.
Fixed plot axis labels Tested on paho-arch. Took 0:55.48.

@@
 %% \end{align}
+\pahosays{Was plotting the radial distribution function meant to be commented?}
 %% \begin{figure}
@@
       work}\fixme{Fix MC issue here (volume wrong?)} \fixme{Add backward curve
-      for this work mc?}\fixme{z-axis wrong}The triplet
-    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3,
-    plotted when $\rr_1$ and $\rr_2$ are in contact.  The color plot on the left
-    shows the triplet distribution function, with white corresponding to the
-    contact value of the \emph{pair} distribution function.  The plot on the
-    right shows the value of $g^{(3)}$ along the path shown as a dotted line on
-    the plot to the left.}\label{fig:triplet-contact-distribution}
+      for this work mc?}The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$
+    and $\rr_2$ are in contact.  The color plot on the left shows the triplet
+    distribution function, with white corresponding to the contact value of the
+    \emph{pair} distribution function.  The plot on the right shows the value of
+    $g^{(3)}$ along the path shown as a dotted line on the plot to the
+    left.}\label{fig:triplet-contact-distribution}
 \end{figure*}

### 2014-03-10 papers/pair-correlation/figs/plot-path-triplet-inbetween.py

@@ -99,10 +99,9 @@ zticks = [4, 6, 8, 10]
 
 plotticks = xticks + [z-rpath for z in zticks]
 xplot.set_xticks(plotticks)
-xplot.set_xticklabels(['$%i$' %tick for tick in xticks[:-1]] + ['$0$ $0$'] +
+xplot.set_xticklabels(['$%i$' %tick for tick in xticks[:-1]] + ['$0$ $2$'] +
                       ['$%i$' %tick for tick in zticks])
 zplot.set_xticks([])
-
 
 zplot.axvline(x=rpath, color='k')
 zplot.axvline(x=3*rpath, color='k')

### 2014-03-10 papers/pair-correlation/paper.tex
More plot tweaks. All plots now use latex fonts, and all garbage mc data is gone. Tested on paho-arch. Took 0:49.16.

@@
 
-\newcommand\paircaption{\pahosays{The MC has one point that jumps up at the
-    beginning of the incontact region.} The pair distribution function near a
-  hard wall, with packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with
-  the hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$
-  varies.  The top half of this figure shows the result of the Monte Carlo
-  simulation, while the bottom half shows the result of the appproximation
-  developed in this work.  On the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on
-  the path illustrated in the figure to the left.  This plot compares the Monte
-  Carlo results (black) with those of this work (blue), Sokolowski and Fischer
+\newcommand\paircaption{The pair distribution function near a hard wall, with
+  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the hard wall.
+  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$ varies.  The top
+  half of this figure shows the result of the Monte Carlo simulation, while the
+  bottom half shows the result of the appproximation developed in this work.  On
+  the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the
+  figure to the left.  This plot compares the Monte Carlo results (black) with
+  those of this work (blue), Sokolowski and Fischer
   (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Remove one of, or make a note of, the two curves for this
-      work}\fixme{Fix MC issue here (volume wrong?)} \fixme{Add backward curve
-      for this work mc?}The triplet distribution function
+  \caption{\fixme{Make note of the double curves}\fixme{Add backward curve for
+      this work mc?}\pahosays{would annotations be clearer if they were on the
+      x/z ticks instead of arrows?}The triplet distribution function
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$

### 2014-03-10 papers/pair-correlation/paper.tex
modified text in paper.  Passing hot potato back Tested on crystal. Took 1:11.02.

@@
 \newcommand{\jeffsays}[1]{{\color{red} [\green{Jeff:} \emph{#1}]}}
-\newcommand{\pahosays}[1]{{\color{red} [\green{Paho:} \emph{#1}]}}
 
@@
 \end{align}
-This approximation is constructed to reproduce the exact value for the
-integral$ \int n(\rr_1)n(\rr_2)
-g^{(2)}(\rr_1,\rr_2)\delta(|\rr_1-\rr_2|-\sigma)d\rr_1d\rr_2$.  This
-approximation (described in Eq.~\ref{eq:g2-our-mean}) requires the
+$g^{(2)}(\rr_1,\rr_2)$ above is constructed to reproduce the exact
+value for the integral$ \int n(\rr_1)n(\rr_2)
+g^{(2)}(\rr_1,\rr_2)\delta(|\rr_1-\rr_2|-\sigma)d\rr_1d\rr_2$.  \par
+The approximation described in Eq.~\ref{eq:g2-our-mean} requires the
 radial distribution function as a function of $r$ and $g_\sigma$.  One
-approach is to compute $\eta$ from $g_\sigma$ and then find
-$g(r,\eta)$ either using some approximate functional form
-(e.g. Percus-Yevick) or using Monte Carlo simulation results.  In this
-paper we use the latter (also for evaluating the approximations of
-earlier works), and refer to this approximation (with an exact
-$g(r,g_\sigma)$) as ``this work MC.''  \fixme{We should develop
-  separate acroyms for each method.}
+approach to constructing $g(r,g_\sigma)$ is to compute $\eta$ from
+$g_\sigma$ and then find $g(r,\eta)$ either using some approximate
+functional form (e.g. Percus-Yevick) or using Monte Carlo simulation
+results.  In this paper we first use the Monte Carlo results (which
+are also used in evaluating the approximations of earlier works), and
+refer to this approximation (which has an exact $g(r,g_\sigma)$) as
+``MCRD.''  \jeffsays{I thought maybe for Monte Carlo Radial
+  Distribution} We will then construct our own, more computationally
+efficient radial distribution approximation, which is addressed in the
+following section.
+
+\newcommand\kappaO{\kappa_0}
+\newcommand\kappaI{\kappa_1}
+\newcommand\kappaZ{\kappa_2}
+
+\input{figs/fit-parameters}
+
+\subsection{Separable fit to the radial distribution function}
 
@@
 %% \end{align}
-\pahosays{Was plotting the radial distribution function meant to be commented?}
 %% \begin{figure}
@@
 
-\newcommand\kappaO{\kappa_0}
-\newcommand\kappaI{\kappa_1}
-\newcommand\kappaZ{\kappa_2}
-
-\input{figs/fit-parameters}
-
-\subsection{Separable fit to the radial distribution function}
-
-In contrast to the previous works, which are based on the homogeneous
-radial distribution function $g(r,\eta)$ expressed as a function of
-distance and packing fraction, we have constructed our separable
-approximation for the radial distribution function as $g(r,
-g_\sigma)$, which is a function of the radius $r$ and its value at
-contact $g_\sigma$.  One can convert between these two representation
-using the Carnahan-Starling formula for $g_\sigma(\eta)$
-\begin{equation}
-  g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
-\end{equation}
-The form $g(r, g_\sigma)$ is more convenient for treating the
-inhomogeneous case, because we can directly compute the averaged value
-of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.
+%% In contrast to the previous works, which are based on the homogeneous
+%% radial distribution function $g(r,\eta)$ expressed as a function of
+%% distance and packing fraction, we have constructed our separable
+%% approximation for the radial distribution function as $g(r,
+%% g_\sigma)$, which is a function of the radius $r$ and its value at
+%% contact $g_\sigma$.  One can convert between these two representation
+%% using the Carnahan-Starling formula for $g_\sigma(\eta)$
+%% \begin{equation}
+%%   g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
+%% \end{equation}
+%% The form $g(r, g_\sigma)$ is more convenient for treating the
+%% inhomogeneous case, because we can directly compute the averaged value
+%% of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.
+\jeffsays{I just commented out a whole paragraph that it seems to me
+  has become reduntant.  In it was the Carnahan-Starling
+  approximation.  Do we need to put this back in?}
 
@@
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
-figure, it is apparent that the integral f $h(r)$ from
+figure, it is apparent that the integral of $h(r)$ from
 Eq.~\ref{eq:total-constraint} results in cancelling errors at large
@@
 
-\newcommand\paircaption{The pair distribution function near a hard wall, with
-  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the hard wall.
-  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$ varies.  The top
-  half of this figure shows the result of the Monte Carlo simulation, while the
-  bottom half shows the result of the appproximation developed in this work.  On
-  the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the
-  figure to the left.  This plot compares the Monte Carlo results (black) with
+\newcommand\paircaption{\fixme{Look into smoothing the top plot on
+    right.} The pair distribution function near a hard wall, with
+  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the
+  hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as
+  $\rr_2$ varies.  The top half of this figure shows the result of the
+  Monte Carlo simulation, while the bottom half shows the result of
+  the appproximation developed in this work.  On the right is a plot
+  of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the figure to
+  the left.  This plot compares the Monte Carlo results (black) with
   those of this work (blue), Sokolowski and Fischer
   (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
-  (green)~\cite{fischer1980born}.  The latter is only plotted at contact, where
-  it is defined.  }
+  (green)~\cite{fischer1980born}.  The latter is only plotted at
+  contact, where it is defined.  }
 
@@
 with a focus on the case where one of the two spheres is in contact
-with the hard wall.  Figure~\ref{fig:pair-distribution}a
-and~\ref{fig:pair-distribution}a compare the results of our
+with the hard wall.  Figures~\ref{fig:pair-distribution}a,\ref{fig:pair-distribution}b
+and Figures~\ref{fig:pair-distribution}c,\ref{fig:pair-distribution}d compare the results of our
 approximate functional with Monte Carlo simulation at packing
@@
 
-Figure~\ref{fig:pair-distribution}b
-and~\ref{fig:pair-distribution}b show the pair distribution function
-as plotted along a path illustrated in
-Fig.~\ref{fig:pair-distribution}a.  These plots compare this work
-with Monte Carlo results, as well as the approximations of Sokolowski
-and Fischer~\cite{sokolowski1992role} and that of Fischer and
+Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
+show the pair distribution function as plotted along paths
+illustrated in Figures~\ref{fig:pair-distribution}a and
+\ref{fig:pair-distribution}c.  These plots compare this work with
+Monte Carlo results, as well as the approximations of Sokolowski and
+Fischer~\cite{sokolowski1992role} and that of Fischer and
 Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
@@
 approximations to the pair distribution function predict the
-qualtitative angular dependence of the pair distribution function at
+qualitative angular dependence of the pair distribution function at
 contact better than this work.  However, in each case the pair
@@
 approaches gives a reasonable prediction, with our approach
-underestimating the oscillations in $g^{(2)}$, as expected based on the
-fit in Fig.~\ref{fig:radial-distribution}.
+underestimating the oscillations in $g^{(2)}$, as expected based on
+the fit in Fig.~\ref{fig:radial-distribution}.
 
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Make note of the double curves}\fixme{Add backward curve for
-      this work mc?}\pahosays{would annotations be clearer if they were on the
-      x/z ticks instead of arrows?}The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$
-    and $\rr_2$ are in contact.  The color plot on the left shows the triplet
-    distribution function, with white corresponding to the contact value of the
-    \emph{pair} distribution function.  The plot on the right shows the value of
-    $g^{(3)}$ along the path shown as a dotted line on the plot to the
-    left.}\label{fig:triplet-contact-distribution}
+  \caption{\fixme{Remove one of, or make a note of, the two curves for
+      this work}\fixme{Fix MC issue here (volume wrong?)} The triplet
+    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing
+    fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.
+    The color plot on the left shows the triplet distribution
+    function, with white corresponding to the contact value of the
+    \emph{pair} distribution function.  The plot on the right shows
+    the value of $g^{(3)}$ along the path shown as a dotted line on
+    the plot to the left.}\label{fig:triplet-contact-distribution}
 \end{figure*}
@@
 which gives the contribution to the mean-field free energy due to
-molecules located at position $z$.
+molecules located at position $z$.  \jeffsays{About this last
+  sentence: this is molecules in the plane $z$, right?  Wanted to make
+  sure that's the case before changing to something like 'in the
+  plane'.  Saying the position z makes it seem that there's only one.}
 

### 2014-03-10 papers/pair-correlation/paper.tex

@@
 
-\newcommand\paircaption{\fixme{Look into smoothing the top plot on
-    right.} The pair distribution function near a hard wall, with
-  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the
-  hard wall.  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as
-  $\rr_2$ varies.  The top half of this figure shows the result of the
-  Monte Carlo simulation, while the bottom half shows the result of
-  the appproximation developed in this work.  On the right is a plot
-  of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the figure to
-  the left.  This plot compares the Monte Carlo results (black) with
+\newcommand\paircaption{The pair distribution function near a hard wall, with
+  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the hard wall.
+  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$ varies.  The top
+  half of this figure shows the result of the Monte Carlo simulation, while the
+  bottom half shows the result of the appproximation developed in this work.  On
+  the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the
+  figure to the left.  This plot compares the Monte Carlo results (black) with
   those of this work (blue), Sokolowski and Fischer
   (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
-  (green)~\cite{fischer1980born}.  The latter is only plotted at
-  contact, where it is defined.  }
+  (green)~\cite{fischer1980born}.  The latter is only plotted at contact, where
+  it is defined.  }
 
@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Remove one of, or make a note of, the two curves for
-      this work}\fixme{Fix MC issue here (volume wrong?)} The triplet
-    distribution function $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing
-    fraction 0.3, plotted when $\rr_1$ and $\rr_2$ are in contact.
-    The color plot on the left shows the triplet distribution
-    function, with white corresponding to the contact value of the
-    \emph{pair} distribution function.  The plot on the right shows
-    the value of $g^{(3)}$ along the path shown as a dotted line on
-    the plot to the left.}\label{fig:triplet-contact-distribution}
+  \caption{\fixme{Make note of the double curves}\fixme{Add backward curve for
+      this work mc?}\pahosays{would annotations be clearer if they were on the
+      x/z ticks instead of arrows?}The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$
+    and $\rr_2$ are in contact.  The color plot on the left shows the triplet
+    distribution function, with white corresponding to the contact value of the
+    \emph{pair} distribution function.  The plot on the right shows the value of
+    $g^{(3)}$ along the path shown as a dotted line on the plot to the
+    left.}\label{fig:triplet-contact-distribution}
 \end{figure*}

### 2014-03-10 papers/pair-correlation/paper.tex
address a couple of fixmes Tested on bennet. Took 0:50.88.

@@
 
-Equation~\ref{eq:g2-our-mean} is a reasonable approximation, but
-cannot be computed efficiently with single-center convolutions as
-written.  In order to ease computational application of this pair
-distribution function in classical density functionals, we have
-developed a separable appoximation for the function $g(r,g_\sigma)$.
+The approximation described in Eq.~\ref{eq:g2-our-mean} is reasonably
+accurate, but cannot be computed efficiently with single-center
+convolutions as written.  In order to ease computational application
+of this pair distribution function in classical density functionals,
+we have developed a separable appoximation for the function
+$g(r,g_\sigma)$.
 \begin{align}
@@
 %% \end{align}
-\pahosays{Was plotting the radial distribution function meant to be commented?}
-%% \begin{figure}
-%%   \centering
-%%   \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
-%%   %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
-%%   \caption{Plot of the radial distribution function, with our fit.}\label{fig:radial-distribution}
-%% \end{figure}
-
-%% In contrast to the previous works, which are based on the homogeneous
-%% radial distribution function $g(r,\eta)$ expressed as a function of
-%% distance and packing fraction, we have constructed our separable
-%% approximation for the radial distribution function as $g(r,
-%% g_\sigma)$, which is a function of the radius $r$ and its value at
-%% contact $g_\sigma$.  One can convert between these two representation
-%% using the Carnahan-Starling formula for $g_\sigma(\eta)$
-%% \begin{equation}
-%%   g_\sigma = \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}.
-%% \end{equation}
-%% The form $g(r, g_\sigma)$ is more convenient for treating the
-%% inhomogeneous case, because we can directly compute the averaged value
-%% of $g_\sigma$ for an inhomogeneous system~\cite{schulte2012using}.
-\jeffsays{I just commented out a whole paragraph that it seems to me
-  has become reduntant.  In it was the Carnahan-Starling
-  approximation.  Do we need to put this back in?}
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
+  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
+  \caption{Plot of the hard-sphere radial distribution function, with
+    our fit.}\label{fig:radial-distribution}
+\end{figure}
 
@@
 \end{equation}
+Finally, we can convert between the packing fraction $\eta$ and the
+radial distribution function at contact $g_\sigma$ using the
+Carnahan-Starling approximation for $g_\sigma$:
 \newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
     6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
-\begin{equation}
-  \eta = 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}
-\end{equation}
+\begin{align}
+  g_\sigma &= \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}\text{, and its inverse} \\
+  \eta &= 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}.
+\end{align}
 This function includes three fitted parameters, $\kappaO =

### 2014-03-11 papers/pair-correlation/paper.tex
minor fix in pair paper (and remove unused file) Tested on bennet. Took 1:04.61.

@@
 \begin{align}
-  g'(\sigma) \approx - h_\sigma g_\sigma
+  g'(\sigma) \approx - g_\sigma (g_\sigma - 1).
 \end{align}
-where $h_\sigma = g_\sigma - 1$ is the total correlation function.
 Under these three constraints, we applied a least-squares fit to

### 2014-03-11 papers/thesis-roth/Project-Report/Project-notes.tex

%Notes on MS Project

\documentclass{article}

\usepackage[superscript]{cite}
\usepackage{amsmath}

\renewcommand{\thesection}{}

\begin{document}
\begin{sffamily}

\title{Notes on MS Project}
\author{D. Roth}
\date{}
\maketitle

\section{c. 26 Aug 13}
Started with group. My project will generally be based on using
Renormalization Group (RG) theory to get a better model of liquids
near the critical point. The theory has been developed by Forte
\textit{et al.}\cite{Forte11}. My project will mostly focus on the
numerics of this method---can it be efficiently implemented in a
computational model?

As an introduction, Roundy set me working on coding the recent Hughes
paper\cite{Hughes13} in Python, for the homogeneous (i.e. number
density is independent of position) case.

\section{11 Sept 13}
Using $a_1$ and $a_2$ from Gil-Villegas\cite{Gil-Villegas97}. In
particular, I am using the square-well model: $a_1^{SW}$ and
$a_2^{SW}$.

\section{13 Sept 13 (Friday the 13$^{th}$)}
Using the constants from Clark \textit{et al.}\cite{Clark06}. In
particular, I'm using the W1 set from Wertheim's four-cite model
(Table 1 in the paper). Reasons for using W1 are described in the
Clark paper, and this is the set that is defined as ``optimal'' in
Table 6.

\section{16 Sept 13}
Dr. Roundy took a look at my Python code for the 1$^{st}$ time. He had
suggestions on code readability, and we did some debugging. We got
some `nan' (not a number) results in my function $X$, and we narrowed
it down to trouble with my $g_\sigma^{SW}$ function, which seemed to
be far too large. It was on the order of 10$^3$, which is certainly
\textit{not} a small perturbation! The likely culprit is the
derivatives, which I shall re-evaluate and check against the code that
I have written. Hopefully it was a simple mistake on my part.

I plan on having plots of pressure vs.\ density in the time frame of a few weeks.

\section{17 Sept 13}
From ref \citenum{Gil-Villegas97}, with $g^{HS}(1;\eta_{eff}) \equiv g_{eff}^{HS}$\ :
\begin{equation*}
  a_1^{SW}(\eta,\lambda) = a_1^{VDW}(\eta, \lambda)g_{eff}^{HS}(\eta_{eff}(\eta,\lambda))
\end{equation*}
\renewcommand{\arraystretch}{2}
\begin{tabular}{| c | c |}
\hline
  Correctly Coded? & Equation \\
\hline
  Y & $ \partial_\eta a_1^{SW} = \partial_\eta a_1^{VDW}g_{eff}^{HS} + a_1^{VDW}\partial_\eta g_{eff}^{HS} $ \\
\hline
  Y & $\partial_\eta a_1^{VDW} = -4\epsilon(\lambda^3 - 1)$ \\
\hline
  N & $\begin{aligned}
        \partial_\eta g_{eff}^{HS} &= \partial_{\eta_{eff}}g_{\eta_{eff}}^{HS}\partial_\eta\eta_{eff} \\
        &= \frac{\partial}{\partial\eta_{eff}}\left[\frac{1-\frac{1}{2}\eta_{eff}}{(1-\eta_{eff})^3}\right]\frac{\partial\eta_{eff}}{\partial\eta} \\
        &= \left[\frac{1}{2(1-\eta_{eff})^3} + \frac{3(1 - \frac{1}{2}\eta_{eff})}{(1 - \eta_{eff})^4} \right]\left[c_1(\lambda) + 2c_2(\lambda)\eta + 3c_3(\lambda)\eta^2\right]
      \end{aligned}$ \\
\hline
  Probably not & $\partial_\lambda a_1^{SW} = \partial_\lambda a_1^{VDW}g_{eff}^{HS} + a_1^{VDW}\partial_\lambda g_{eff}^{HS}$ \\
\hline
  Y & $\partial_\lambda a_1^{VDW} = -12\eta\epsilon\lambda^2$ \\
\hline
  N & $\partial_\lambda g_{eff}^{HS} = \partial_{\eta_{eff}}g_{eff}^{HS}\partial_\lambda\eta_{eff}$ \\
\hline
  Y & $\partial_\eta\eta_{eff} = dc_1(\lambda)\eta + dc_2(\lambda)\eta^2 + dc_3(\lambda)\eta^3$ \\
\hline
\end{tabular}

After making appropriate changes, $g_\sigma^{SW}$ is on the order of 10$^0$, and no more `nan' results.

\section*{18 Sept 13}
Roundy suggested that $g_\sigma^{SW}$ should be larger than
$g_\sigma^{HS}$, which is not the case.

\section*{19 Sept 13}
I leared a trick to debugging: just because two expressions (in your
code and a reference code, for example) are mathematically the same
\textit{and} coded the same/correctly does not mean that your code is
correct. The trick is to actually have the program print what it is
doing.

\section*{24 Sept 13}
Jess Hughes' project report has the derivation for $p(T,n)$. It's a
bit different than ``normal'' because our functional gives free energy
per volume---so it's not straightforward to take a derivative wrt
volume. The formula ends up being
\begin{equation*}
  p(T,n) = \frac{\partial f}{\partial n}\bigg|_T + f(n)
\end{equation*}
where $f(n) = \tfrac{F(n)}{V}$.

\section*{Coexistance curve algorith using Hughes}
See comments in prefac.py

\bibliographystyle{ieeetr} %ieeetr is similar to unsrt combined with abbr
\bibliography{refs}

\end{sffamily}
\end{document}

### 2014-03-11 papers/pair-correlation/paper.tex

@@
   \caption{\fixme{Make note of the double curves}\fixme{Add backward curve for
-      this work mc?}\pahosays{would annotations be clearer if they were on the
-      x/z ticks instead of arrows?}The triplet distribution function
+      this work mc?}The triplet distribution function
     $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$

### 2014-03-12 papers/pair-correlation/paper.tex

@@
 \begin{document}
-\title{The Pair Distribution Function of the Inhomogeneous Hard-Sphere
-  Fluid}
+\title{The contact value approximation for the pair distribution function of the inhomogeneous hard-sphere
+  fluid}
 
@@
   analytical approximation.  \fixme{needs special attention.}
+  \fixme{add acronym, and discuss separable vs. non}
 \end{abstract}
@@
 
-\section{Our approach}
+\section{Contact value approximation}
 %% We approximate the pair correlation with the average of two radial
@@
 refer to this approximation (which has an exact $g(r,g_\sigma)$) as
-``MCRD.''  \jeffsays{I thought maybe for Monte Carlo Radial
-  Distribution} We will then construct our own, more computationally
-efficient radial distribution approximation, which is addressed in the
-following section.
+the ``contact value approximation'' (CVA).  \fixme{Should we mention
+  this earlier?}
 
@@
 
-\subsection{Separable fit to the radial distribution function}
+\section{Separable radial distribution function for the CVA}
 
-The approximation described in Eq.~\ref{eq:g2-our-mean} is reasonably
-accurate, but cannot be computed efficiently with single-center
-convolutions as written.  In order to ease computational application
-of this pair distribution function in classical density functionals,
-we have developed a separable appoximation for the function
-$g(r,g_\sigma)$.
+The contact value approximation (described in
+Eq.~\ref{eq:g2-our-mean}) is reasonably accurate, but cannot be
+computed efficiently with single-center convolutions as written.  In
+order to ease computational application of this pair distribution
+function in classical density functionals, we have developed a
+separable appoximation for the function $g(r,g_\sigma)$.
 \begin{align}
@@
 \end{align}
-This separable approximation allows the mean field contribution to
-the free energy (Eq.~\ref{eq:mean-field}) to be computed with only
-single-center convolutions, which can be efficiently computed using
-fast Fourier transforms.
+This separable contact value approximation (CVA-S) allows the mean
+field contribution to the free energy (Eq.~\ref{eq:mean-field}) to be
+computed with only single-center convolutions, which can be
+efficiently computed using fast Fourier transforms.  \fixme{Comb paper
+  for places to say CVA-S and CVA.}
 
@@
 which gives the contribution to the mean-field free energy due to
-molecules located at position $z$.  \jeffsays{About this last
-  sentence: this is molecules in the plane $z$, right?  Wanted to make
-  sure that's the case before changing to something like 'in the
-  plane'.  Saying the position z makes it seem that there's only one.}
+molecules located in a plane of fixed $z$.
 

### 2014-03-15 papers/thesis-roth/project.tex

--- 2011-10-01 paper/project.tex
@@
 %% Definitions
-\def\re{\text{Re}}
-\def\im{\text{Im}}
-\def\ket#1{\vert #1 \rangle}
-\def\cU{{\cal{U}}}
-\def\cD{{\cal{D}}}
-\def\re{\text{Re}}
-\def\im{\text{Im}}
-\newcommand{\red}[1]{{\bf \color{red} #1}}
-\newcommand{\blue}[1]{{\bf \color{blue} #1}}
-\newcommand{\green}[1]{{\bf \color{green} #1}}
-\newcommand{\rr}{\textbf{r}}
-\newcommand{\xx}{\textbf{r}}
-\newcommand{\refnote}{\red{[ref]}}
-
-\newcommand{\fixme}[1]{\red{[#1]}}
-
-% needsworklater is used to annotate bits that need work, but that we
-% can postpone for a while.
-\newcommand{\needsworklater}[1]{\emph{[#1]}}
-% needsworknow is intended to prioritize stuff that needs fixing.
-\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}
 
@@
 \begin{document}
-\title{A Classical Density-Functional Theory for Describing Water Interfaces}
+\title{Applying Renormalization Group Theory to the Square Well Liquid}
 
-\author{Jessica Hughes}
+\author{Dan Roth}
 \affiliation{Department of Physics, Oregon State University, Corvallis, OR
@@
 \begin{abstract}
-{We develop a classical density functional theory for water by combining the
-fundamental-measure theory functional--which describes hard sphere fluids--with
-attractive interactions based on the Statistical Associating Fluid Theory
-(SAFT). Using this approximation, we are able to describe the vapor pressure, 
-saturated liquid density, surface tension, enthalpy and
-entropy of water, as well as the number of hydrogen bonds and density oscillations
-formed near
-liquid-vapor and liquid-solid interfaces. These interfaces include water
-confined to a hard-walled slit, a single hydrophobic rod immersed in water, and
-two hydrophobic rods 
-at various separations. For two rods being pulled apart, a transition occurs 
-from a vapor state to a liquid state in the area between
-the rods at approximately $(\pi-2)$ multipled by the radius. Even 
-for these nanoscale rods (radius 0.5~-~1.0~nm), we find that in the
-vapor state, the force can be approximated
-classically and is equal to 144~mN/m, or twice the bulk surface tension.
-This functional is computationally efficient and provides a foundation for 
-future work describing interactions in aqueous solution on an
-atomic level.}
+{}
 
@@
 
-\subsection{Water---the universal solvent}
-
-The simple yet complex water molecule is one of the most important 
-substances for life and industry. The structure and geometry of the 
-water molecule leads to a strong polarity and complex intermolecular
-forces. The hydrogen bonds formed by water are very strong, creating
-tetragonal clusters that lead to many of water's unusual properties,
-such as the high melting, boiling, and critical temperatures as well as the
-density maximum above the melting point.
-The large dielectric constant is partially responsible for
-water acting as a universal solvent, along with the ease of availability and
-use.
-There are many models used to describe water since so many different
-processes occur in aqueous solution. It is challenging to describe all the 
-relevant properties of water at once, especially if a low computational cost is
-also desired.
-
-Some models are primarily designed for use with homogeneous systems. Although
-these 
-models are useful in predicting some bulk properties, the interesting
-chemistry that occurs in water cannot be described with a homogeneous-only
-model. To
-describe solvation and other processes in aqueous solution, a more robust theory
-that 
-can handle varying densities is needed.
-
-\subsection{Classical density-functional theory}
-
-Classical density-functional theory (DFT) provides an efficient way to evaluate
-the free energy using the probability distribution of the fluid. The free energy 
-is a functional
-of the density function and is independent of any external
-potentials\cite{ebner1976density}. The most common use of DFT is calculating ground
-state electron densities in quantum mechanical systems. The Hohenberg-Kohn
-theorem\cite{hohenberg1964inhomogeneous} allows us to write the ground state
-energy in terms of a universal functional of the electron density and a separate
-term including the external potential. Mermin\cite{mermin1965thermal} gives a
-theorem extending that of Hohenberg and Kohn to nonzero temperatures:
-\begin{equation}
-  A(T) = \underset{n(r)}{min}\left\{ F[n(r),T] + \int V_{ext}(r) n(r)
-d^3r\right\}
-\end{equation}
-This is the basis for classical DFT, where the relevant density is not the
-electron density,
-but the density of molecules. This can be used with a variety of 
-external potentials, from the simple interaction with a 
-hydrophobic wall to the more complex, stronger interaction with hydrophilic
-solutes.
-
-\subsection{Statistical Associating Fluid Theory}
-
-The high hydrogen bonding and strong polar interactions in water characterize it
-as an associating fluid, and thus it requires a different treatment than simple
-fluids which are mainly defined by van der Waals forces and weak electrostatic
-interations. The Statistical Associating Fluid Theory (SAFT) was designed for
-use with fluids such as water, to account for the important association forces
-(like hydrogen bonds) as well as any effects due to clusters or chains forming
-in the liquid\cite{muller2001molecular}. It is based on Wertheim's first-order
-thermodynamic perturbation theory (TPT1)
-\cite{wertheim1984fluidsI,wertheim1984fluidsII,wertheim1986fluidsIII,
-wertheim1986fluidsIV} which provides a useful way to account for strong
-associative interations between molecules.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Theory and Methods}
-We introduce a new classical density functional for water that
-predicts the vapor pressure, vapor density, liquid density, bulk surface tension, 
-and enthalpy, as well 
-as allows for liquid-vapor coexistence.  The Helmholtz free energy functional is
-based on SAFT, and is composed of the usual terms:
-\begin{equation}
-  F[n] = F_\textit{id}[n] + F_\textit{hs}[n]  +
-F_\textit{disp}[n]+ F_\textit{assoc}[n],
-\end{equation}
-where $F_\textit{id}$ is the ideal gas free energy, $F_\textit{hs}$ is
-the hard-sphere free energy, $F_\textit{disp}$ is the dispersion energy,
-and $F_\textit{assoc}$ is the free energy of association.  In
-addition, a chemical potential term is used, so we actually work in
-the grand canonical ensemble.  In the following sections, we will
-introduce the terms of this functional.
 
-\subsection{Ideal gas functional}
-The first term is the ideal gas free energy functional,
-which incorporates effects of kinetic energy.  The ideal gas
-functional is given by
-\begin{equation}\label{idealgas}
-  F_{id}[n] = k_B T \int n(\xx)\left( \ln{\frac{n(\xx)}{n_Q}} - 1\right) d\xx
-\end{equation}
-where n(\xx) is the density of water molecules and $n_Q$ is the quantum 
-concentration. This is the particle concentration where quantum effects
-become appreciable and is given by
-\begin{equation}\label{quantumconcentration}
- n_Q =\left(\frac{mk_BT}{2\pi\hbar^2}\right)^{3/2}
-\end{equation}
-where $m$ is the molecular mass of water. The ideal gas free
-energy functional has the property that the contact density at a hard
-surface $n_c$ is proportional to the pressure on that surface,
-according to the contact value theorem
-\begin{equation}\label{contactvaluethm}
-  p = n_c k_BT \:,
-\end{equation}
-which also leads to the property that the free energy of solvation of
-a hard solute is proportional to its volume in the limit of small
-solutes.  These properties are retained by the total functional,
-provided the remaining terms are \emph{purely nonlocal}. This requires
-that
-\begin{equation}
- \frac{\delta^2F_{\textit{excess}}[n]}{\delta n(\xx_1)~\delta n(\xx_1)}\neq\infty
-\end{equation}
-where $F_{\textit{excess}}[n]$ describes all the remaining terms in the functional.
-This ensures that any higher order terms in the Taylor expansion
-of $F[n+\Delta n]$ are small and do not contribute to the small
-solute free energy of solvation.
-
-\subsection{Hard-sphere repulsion}
-
-We model the repulsive interactions using the White Bear version of
-the Fundamental-Measure Theory~(FMT) functional for the hard-sphere
-fluid~\cite{roth2002whitebear}.  In the homogeneous limit, the White
-Bear functional reduces to the Mansoori-Carnahan-Starling-Leland
-equation of state.  FMT functionals are expressed as the integral of
-the \emph{fundamental measures} of a fluid, which provide local
-measures of quantities such as the filling fraction, density of
-spheres touching a given point and mean curvature.  The excess free
-energy is written as:
-\begin{equation}
-F_{hs}[n] = k_B T \int (\Phi_1(\xx) + \Phi_2(\xx) + \Phi_3(\xx)) d\xx \; ,
-\end{equation}
-with integrands
-\begin{align}
-\Phi_1 &= -n_0 \ln\left( 1 - n_3\right)\\
-\Phi_2 &= \frac{n_1 n_2 - \mathbf{n}_{V1} \cdot\mathbf{n}_{V2}}{1-n_3} \\
-\Phi_3 &= (n_2^3 - 3n_2 \mathbf{n}_{V2} \cdot \mathbf{n}_{V2})
-  \frac{
-    n_3 + (1-n_3)^2\ln(1-n_3)
-  }{
-    36\pi n_3^2\left( 1 - n_3 \right)^2
-  } ,
-\end{align}
-where the fundamental measure densities are given by:
-\begin{align}
-  n_3(\xx) &= \int n(\xx') \Theta(\left|\xx - \xx'\right| - R) d\xx' \\
-  n_2(\xx) &= \int n(\xx') \delta(\left|\xx - \xx'\right| - R) d\xx'
-  \\
-  \mathbf{n}_{V2} &= \mathbf{\nabla} n_3 \\
-  n_1 &= \frac{n_2}{4\pi R}\\
-  \mathbf{n}_{V1} &= \frac{\mathbf{n}_{V2}}{4\pi R}\\
-  n_0 &= \frac{n_2}{4\pi R^2}.
-\end{align}
-The density $n_3$ is the filling fraction and $n_2$ describes the number
-of spheres touching a given point. For our water functional, we use the 
-hard-sphere radius of
-3.03420~\AA, which was found to be optimal by Clark
-\emph{et al}.\cite{clark2006developing}
-
-\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
-\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
-\newcommand\epsilonassoc{\ensuremath{\epsilon_\textit{a}}}
-\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
-\newcommand\lscale{\ensuremath{s_d}}
-\subsection{Dispersion free energy}
-The first term in the attractive energy is the dispersion term.  This
-will include van der Waals and other non-associative interations. We use a
-dispersion term based on the SAFT-VR
-approach\cite{gil-villegas-1997-SAFT-VR}, which has two free
-parameters, an interaction energy $\epsilondisp$ and a
-length scale $\lambdadisp$.
-
-The dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
-\begin{align}
-  F_\text{disp}[n] &= \int (a_1(\xx) + \beta a_2(\xx))d\xx
-\end{align}
-where $a_1$ and $a_2$ are the first two terms in a high-temperature
-perturbation expansion and $\beta=1/k_BT$.  The first term $a_1$ is 
-the mean-field dispersion
-interaction, which reduces in the homogeneous limit to
-\begin{align}\label{eq:A1-simple}
-  a_1 &= \frac12 n_b \int \varphi(\left|\xx\right|)
-  g_{HS}(\left|\xx\right|) d\xx
-\end{align}
-where $n_b$ is the bulk density and $g_{HS}$ is the homogeneous
-hard-sphere fluid correlation function.
-The second dispersion term in the free energy $a_2$ describes the
-effect of fluctuations resulting from the compression of the fluid due
-to the dispersion interaction itself, and is commonly approximated
-using the local compressibility approximation (LCA). In the LCA,
-we assume the energy fluctuation is directly related to the
-compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
-
-We use a free square-well function for the dispersion $\varphi$, which
-is the choice used by Clark \emph{et al}~\cite{clark2006developing},
-and allows a reasonable fit to the equation of state.  Thus our model
-interaction has the form:
-\begin{equation}
-  \varphi(r) = \Theta(r-2 \lambdadisp R)
-\end{equation}
-where $\Theta$ is the Heaviside step function.
-
-The form of $a_1$ is given in
-reference~\cite{gil-villegas-1997-SAFT-VR}, but is expressed in terms
-of the filling fraction.  In order to apply this form to an
-inhomogeneous density distribution, we construct an effective local
-filling fraction for dispersion $\etadisp$, given by
-\begin{align}
-  \etadisp(\xx) &=
-  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
-             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
-\end{align}
-This effective filling fraction is used throughout the dispersion
-calculations, and represents a filling fraction averaged over the
-effective range of the dispersive interaction.  This functional
-introduces an additional empirical parameter $\lscale$ which adjusts
-the length scale over which the dispersion interaction is correlated.
-
-The first term in the dispersion functional $a_1$ when written using
-the above filling fraction for dispersion has the form
-\begin{align}
-  a_1 &= 
-   -4(\lambdadisp^3 - 1)\epsilondisp \etadisp(\xx)
-    g^{HS}_\sigma(\eta_\textit{eff}(\xx)) \\
-  g_\sigma^{HS}(\eta) &= \frac{1}{1-\eta}
-  +\frac32\frac{\zeta(\xx)}{(1-\eta)^2}
-  + \frac12\frac{\eta^2\zeta(\xx)}{(1-\eta)^3}
-  \\
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}\\
-  \eta_\textit{eff}(\xx) &= \sum_{ij=0}^3 C_{ij}  \etadisp^i(\xx)
-  \lambdadisp^j
-\end{align}
-where $g_\sigma^{HS}$
-is the hard-sphere fluid correlation function evaluated at contact.
-The $\zeta$ factor
-(also seen in the association energy, below) is a dimensionless
-measure of the local density inhomogeneity. The $C_{ij}$ values
- are numerical constants taken from
-reference\cite{gil-villegas-1997-SAFT-VR}, which come from a numerical
-fit to the integral in Equation~\ref{eq:A1-simple} over a range of
-values for $\lambdadisp$ and filling fractions.
-
-The second term, $a_2$, which describes the contribution to the free
-energy associated with fluctuations is given by
-\begin{align}
-  a_2 &= \frac12 \epsilondisp
-              K_{HS}(\etadisp(\xx)) \etadisp(\xx)
-              \frac{\partial a_1}{\partial \etadisp(\xx)}
-\end{align}
-where $K_{HS}$ is the Percus-Yevick hard-sphere compressibility, given
-by
-\begin{align}
-  K_{HS} &=
-    \frac{(1 - \etadisp)^4}{1 + 4(\etadisp + \etadisp^2)}.
-\end{align}
-
-\subsection{Association free energy}
-The final attractive energy term is the association term, which
-accounts for hydrogen bonding.  Hydrogen bonds are modeled as four
-attractive patches (``association sites'') on the surface of the hard
-sphere.  These four sites represent two protons and two electron lone
-pairs.  There is an attractive energy $\epsilonassoc$ when
-two molecules are oriented such that the proton of one overlaps
-with the lone pair of the other.  The volume over which this
-interaction occurs is $\kappa_\textit{assoc}$, giving the association
-term in the free energy has two empirical parameters that are fit to
-experimental data.
-
-The association functional we use is a modified version of that of Yu
-and Wu\cite{yu2002fmt-dft-inhomogeneous-associating}.\footnote{We
-  should note that Fu and Wu\cite{fu2005vapor-liquid-dft} use almost
-  the same functional, but their paper contains errors in the
-  association term and is not useful as a reference.}  The association
-functional of Fu and Wu includes the effects of density
-inhomogeneities on the \emph{contact density} of hard spheres, which
-conventionally appears in the form of the density multiplied by the
-contact value of the correlation function $g^{HS}_\sigma$, but is
-based on the SAFT-HS equation of state, rather than the SAFT-VR
-equation of state\cite{gil-villegas-1997-SAFT-VR} which is used in the
-optimal SAFT parametrization for water found by Clark \emph{et
-  al}\cite{clark2006developing}.
-
-The association free energy for our four-site model has the form
-\begin{align}
-  F_\text{assoc}[n] &= 4 k_BT \int n_0(\xx)\zeta(\xx)
-  \left(\ln X(\xx) - \frac{X(\xx)}{2} + \frac12\right) d\xx
-\end{align}
-where the value of $4$ comes from the four association sites per
-molecule, and the functional $X$ is the fraction of assotiation sites
-\emph{not} hydrogen-bonded, and $\zeta(\xx)$ is a dimensionless
-measure of the density inhomogeneity.  The fraction $X$ is determined
-by the quadratic equation
-\begin{align}
-  X(\xx) &= \frac{\sqrt{1 + 8n_0(\xx)\zeta(\xx)\Delta(\xx)} - 1}
-  {4 n_0(\xx)\zeta(\xx)\Delta(\xx)}
-\end{align}
-where the factor $\zeta$ from
-Yu and Wu\cite{yu2002fmt-dft-inhomogeneous-associating,
-  fu2005vapor-liquid-dft} describes the inhomogeneity of the fluid
-density, and reduces to unity in the homogeneous limit:
-\begin{align}
-  \zeta(\xx) &= 1 - \frac{\mathbf{n}_{2V}\cdot\mathbf{n}_{2V}}{n_2^2}.
-\end{align}
-The functional $\Delta$ which appears in the formula for $X$ is a
-measure of hydrogen-bonding probability.  In the SAFT-VR
-model\cite{gil-villegas-1997-SAFT-VR}, $\Delta$ is given by
-\begin{align}
-  \Delta(\xx) &= \kappa_\textit{assoc} g^\textit{SW}_\sigma(\xx)
-  \left(e^{-\beta\epsilonassoc} - 1\right) \\
-  g^\textit{SW}_\sigma(\xx) &= g^\textit{HS}_\sigma(\etadisp(\xx)) +
-  \frac{1}{4}\beta\left(\frac{\partial a_1}{\partial \etadisp(\xx)} -
-  \frac{\lambdadisp}{3 \etadisp}\frac{\partial a_1}{\partial \lambdadisp}\right)
-\end{align}
-% This is equation 77 of gil-villegas-1997-SAFT-VR
-where $g^\textit{SW}_\sigma$ is the correlation function evaluated at
-contact for a hard-sphere fluid with a square-well attractive
-potential, and $a_1$ and $a_2$ are two terms in the dispersive term of
-the free energy.  The correlation function $g^\textit{SW}_\sigma$ is
-written as a perturbative correction to the hard-sphere fluid correlation
-function $g^\textit{HS}_\sigma$, for which we use the functional of Yu and
-Wu\cite{yu2002fmt-dft-inhomogeneous-associating}:
-\begin{align}
-  g_\sigma^{HS}(\xx) &= \frac{1}{1-\etadisp(\xx)}
-  +\frac32\frac{\zeta(\xx)}{(1-\etadisp(\xx))^2}
-  + \frac12\frac{\etadisp(\xx)^2\zeta(\xx)}{(1-\etadisp(\xx))^3}.
-\end{align}
-
-\subsection{Other thermodynamic functions}
-
-From the Helmholtz free energy functional, we may obtain any other
-thermodynamic functions we should choose.  The grand free energy
-$\Phi$, which we ordinarily use in our computations, is obtained by
-simply adding a chemical potential term:
-\begin{equation}
-  \Phi[n] = F[n] + \mu \int n(\xx) d\xx.
-\end{equation}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/pressure-with-isotherms}
-\end{center}
-\caption{The pressure versus density for various temperatures, including
-experimental pressure data from NIST\cite{nistwater}. The dash-dotted lines
-indicate the computationally calculated pressure and the open circles are 
-NIST data points, plotted for every other isotherm. The solid and dotted lines
-represent the theoretical and experimental coexistence curves.}
-\label{fig:pressure-with-isotherms}
-\end{figure}
-
-The pressure would naturally be obtained by taking a partial
-derivative with respect to volume at fixed temperature.  Since
-\emph{volume} isn't a parameter in density-functional theory, we
-consider the \emph{free energy per unit volume}:
-\begin{align}
-  F[n] &= f[n]V \\
-  p[n] &= -\left(\frac{\partial F[n]}{\partial V}\right)_{T} \\
-  &= -\left(\frac{\partial f[n]V}{\partial V}\right)_{T} \\
-  &= -V\left(\frac{\partial f[n]}{\partial V}\right)_{T}
-   - f[n]\left(\frac{\partial V}{\partial V}\right)_{T} \\
-  &= n \left(\frac{\partial f[n]}{\partial n}\right)_{T} - f[n].
-\end{align}
-
-Figure~\ref{fig:pressure-with-isotherms} shows the calculated pressures for
-isotherms ranging from room temperature to past the critical temperature. The
-solid line is the theoretical coexistence curve calculated from the SAFT
-functional. For a given isotherm, the pressure curve should intersect this
-coexistence curve twice, at the vapor and liquid coexistence 
-pressure corresponding to the vapor and liquid densities. Even
-though the theoretical pressure isotherms drop below the coexistence curve,
-this is non-physical and will not actually occur. The experimental data shows
-how the actual density value will ``jump'' to the other side of 
-the coexistence curve at that pressure value (for
-temperatures less than the critical temperature), corresponding to a phase 
-change in the material. For example, follow the room temperature isotherm
-starting from high pressure and high density (top right of 
-Figure~\ref{fig:pressure-with-isotherms}).
-As the pressure decreases, the density decreases \emph{slightly} until reaching
-the coexistence curve, at which point a phase change from liquid to vapor
-will occur. The
-functional matches well with the experimental data from NIST away from the
-critical point. The slope of the pressure versus density curve is proportional
-to the compressibility, which is close but slightly too large at room
-temperature.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-tension}
-\end{center}
-\caption{The theoretical versus experimental surface tension
-  versus temperature. The experimental data is taken from NIST.\cite{nistwater}
-  The length-scaling parameter $\lscale$ is fit so that the theoretical surface tension
-  will match the experimental surface tension near room temperature.}
-\label{fig:surface-tension}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/equation-of-state}
-\end{center}
-\caption{The theoretical versus experimental phase diagram, vapor
-  pressure versus temperature.  }
-\label{fig:equation-of-state}
-\end{figure}
-
-The entropy is naturally obtained by taking a partial derivative with
-respect to temperature:
-\begin{align}
-  S[n] &= \left(\frac{\partial F[n]}{\partial T}\right)_{V}
-\end{align}
-This derivative is trivial for the ideal gas and hard sphere free
-energies, which are simply proportional to the temperature.  For the
-association and dispersion free energies, it is a little more work,
-but still not hard.
-
-Once we have the entropy, finding the internal energy is simple:
-\begin{align}
-  U[n] &= F[n] + TS[n]
-\end{align}
-and the enthalpy is not much more work:
-\begin{align}
-  H[n] &= F[n] + TS[n] + pV.
-\end{align}
-The calculated heat of vaporization at atmospheric pressure is
-$\Delta H_{vap}$= 39.41~kJ/mol at the normal boiling point of water,
-373 K. This compares well to the experimental
-value measured by NIST ($\Delta H_{vap}$= 40.65~kJ/mol\cite{nistwater}), with 
-an error of only a few percent.
-
-\subsection{Determining the empirical parameters}\label{sec:empirical}
-
-The majority of the empirical parameters used in our functional are
-taken from the paper of Clark \emph{et al} on developing an optimal
-SAFT model for water\cite{clark2006developing}.  This SAFT model
-contains five empirical parameters: the hard-sphere radius, an energy
-and length scale for the dispersion interaction, and an energy and
-length scale for the association interaction.  In addition to the five
-empirical parameters of Clark \emph{et al}, we add a single additional
-parameter $\lscale$---with a fitted value of 0.72---which determines
-the length scale over which the density is averaged when computing the
-dispersion free energy.  We use this final parameter to fit the
-surface tension, with the result shown in
-Figure~\ref{fig:surface-tension}.  Because the SAFT model of Clark
-\emph{et al} overestimates the critical temperature---which is a
-common feature of models which do not explicitly treat the critical
-point---we cannot correctly describe the surface tension at all
-temperatures, and chose to fit it for the temperature range at which
-water is liquid at one atmosphere of pressure.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/temperature-versus-density}
-\end{center}
-\caption{The theoretical versus experimental coexistence curve. Away from
-the critical temperature, the agreement is good. }
-\label{fig:temperature-vs-density}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/finding-vapor-pressure}\\
-\includegraphics[width=\columnwidth]{figs/near-critical-point}
-\end{center}
-\caption{By plotting the free energy versus density it is possible to
-  determine a saturated liquid or vapor state by the common tangent
-  method. The top plot illustrates the common tangent near room 
-  temperature, while the bottom plot shows the same thing close to the 
-  critical temperature. }
-\label{fig:homogeneous}
-\end{figure}
-
-For reference, we plot the predicted vapor pressure (see
-Figure~\ref{fig:equation-of-state}) and the coexistence curve (see
-Figure~\ref{fig:temperature-vs-density}) against experimental values.  These
-match those calculated by Clark \emph{et al}. The functional provides a good description of
-the vapor pressure, which has an error of less than 
-1\% compared with the experimental data. The coexistence curve in 
-Figure~\ref{fig:temperature-vs-density} is similar to the pressure coexistence
-curve in Figure~\ref{fig:pressure-with-isotherms} (solid line) but gives 
-temperature coexistence values.
-Near the critical temperature,
-the agreement is off, and the theoretical prediction  of the temperature is quite a bit higher than
-experiment. 
-To find the densities for the coexistence curve, we use the
-``common tangent'' rule. The two densities 
-will be on the same tangent line to the free energy as illustrated in the top of Figure~\ref{fig:homogeneous}.
-As the critical point approaches, the vapor and liquid coexistence densities get closer and closer together 
-(see the bottom of Figure~\ref{fig:homogeneous}. This illustrates the difficulty
-in correctly treating the critical temperature. As the vapor and liquid densities approach
-each other, the free energy barrier between the two states gets smaller and it becomes
-challenging to computationally find the correct minimum. This is the reason 
-why the coexistence curves
-in Figures~\ref{fig:pressure-with-isotherms}~and~\ref{fig:temperature-vs-density}
-are not smooth around the critical point.
+\subsection{Renormalization Group Theory}
+Renormalization group (RG) theory is an iterative procedure that was
+developed to deal with large-scale fluctuations of the order parameter
+near the critical point of a system.
 
@@
 
-\subsection{Water confined to a slit}
-
-One of the most computationally simple ways to study confined water is by
-constraining it to a slit where the water molecules can only move in one
-dimension. Experimentally, this could represent two different situations.
-The first, water between two perfectly aligned and smooth planes, is not common 
-in experiment due to the difficulty of creating such an ideal situation. The
-second
-possibility is a large sphere brought close to a plane, which eliminates the 
-need for perfect alignment.
-
-Before looking at the liquid-solid one-dimensional slit, we set up a liquid-vapor
-interface and plot the density profile.
-The fluid density at a liquid-solid interface commonly displays
-oscillatory behavior, but the presence of density oscillations at
-liquid-vapor interfaces is more debated\cite{penfold2001structure}.
-In fact, the \emph{meaning} of the density profile at a liquid-vapor
-interface is clouded by the existence of long-wavelength perturbations
-known as \emph{capillary waves}.  These waves are not accounted for in
-density-functional theory (although the exact
-density-functional theory would include them), with the result
-that the density profile will be smoothed out by capillary wave effects. The liquid-vapor
-interface is plotted in Figure~\ref{fig:liquid-vapor-profile} and shows
-some small oscillations.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/surface-298}
-\end{center}
-\caption{The liquid-vapor density profile at room temperature shows
-some small oscillations close to the interface.  }
-\label{fig:liquid-vapor-profile}
-\end{figure}
-
-To set up the calculations for the slit, we choose a chemical potential 
-corresponding to a pressure of 1~atm and set the
-temperature to a typical room temperature of 298~K. We apply 
-a constraining potential outside the slit, where we choose a small non-zero
-density for numerical reasons. The slits studied range in size from 0.1~nm to 
-8.0~nm. We minimize the energy and calculate density, energy, and $X$ on a
-lattice
-with a spacing of approximately 0.01~nm (0.2~bohrs) between lattice points.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-1D}
-\end{center}
-\caption{ Density profile for a slab of water with molecules 
-constrained to move in one dimension. The slit size is 4.0~nm and the dashed
-line represents the saturated liquid density.}
-\label{fig:density-1D}
-\end{figure}
-
-The density profile for a slit of size 4.0~nm is shown in Figure~\ref{fig:density-1D}. 
-Near the interface, there are oscillations around the
-liquid density which decrease near the center of the slit. We expect the center
-of the slit to act approximately like bulk water, and the absence of
-oscillations supports this. For larger slits (6.0~nm or more) the majority of
-the water molecules in the slit act like the bulk.
-
-Based on our computations, we determine the smallest slit that can contain 
-liquid water. This result is
-approximately 0.6~nm. For smaller slits, the initial liquid water inside will
-vaporize. The results for larger liquid-filled slits are only metastable,
-however, as a vapor-filled slit is thermodynamically favorable. Based on our
-calculations for a pressure of 1~atm, the work needed to pull apart
-the two hard walls comprising the slit is one to two orders of magnitude less
-than the free energy of the liquid-filled slit. For the range of slit sizes we
-studied, all the computational results of liquid-filled slits are
-metastable.
-
-Lum, Chandler and Weeks \cite{lum1999hydrophobicity} discuss the
-stability of water between large hydrophobic plates and the critical separation
-necessary for stability. For water at room temperature and atmospheric pressure,
-plates with a separation smaller than about 100~nm will not contain stable liquid
-water. They go on to say that confined water has a region of metastability down
-to separations of about 5~nm. This is comparable to the slit separation where
-our calculations show water inside the slit acting like the bulk liquid, which
-we know to be a metastable solution. Forsman \emph{et al.}\cite{forsman1996computer}
-also study water between two hard walls at constant chemical potential. They 
-also conclude that the liquid state is likely metastable, and there is a free
-energy barrier to reach the stable vapor phase.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-1D}
-\end{center}
-\caption{The number of hydrogen bonds per molecule for a slab of water 
-with molecules constrained to move in one dimension. Each molecule may
-contribute to
-four bonds, although each bond involves two molecules so the
-actual number of bonds would be halved.}
-\label{fig:xassoc-1D}
-\end{figure}
-
-The value of $X$ behaves as expected for water constrained to a slit. We
-show in Figure~\ref{fig:xassoc-1D} the number of hydrogen bonds per water 
-molecule, equal to $4(1-X)$, for a slit of 4.0~nm.
-Outside the slit, there is almost no hydrogen bonding and inside the number of hydrogen
-bonds is approximately equal to 3.5, which is expected for saturated liquid
-water.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-single-rod}
-\end{center}
-\caption{ Density profiles for a single rod of different radii. The dotted line
-represents the saturated liquid density.  }
-\label{fig:density-single-rod}
-\end{figure}
-
-\subsection{One hydrophobic rod}
-
-We move into two dimensions by studying a single hydrophobic rod
-immersed in water. Figure~\ref{fig:density-single-rod} shows density
-profiles for three different radii, 0.3~nm, 0.5~nm, and 1~nm. There are the
-expected oscillations outside the rod and as the distance away from the rod
-increases
-the density approaches that of the bulk saturated liquid density. The small
-radius
-rod has comparatively larger oscillations and a higher contact density.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/xassoc-single-rod}
-\end{center}
-\caption{ Number of bonds per molecule near the interface with a single
-hydrophobic rod. There is a decrease in the hydrogen bonding near the surface 
-of the rod, as expected at the interface of any hydrophobic wall. }
-\label{fig:xassoc-single-rod}
-\end{figure}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-vs-diameter}
-\end{center}
-\caption{ Free nergy versus radius for a single hydrophobic rod
-immersed in water. This should have an asymptote equal to the surface
-tension at room temperature, and it agrees well with the surface tension in
-Figure~\ref{fig:surface-tension}. }
-\label{fig:energy-vs-diameter}
-\end{figure}
-
-The value of $X$ is calculated and converted to number of bonds per molecule and 
-shown for a rod with radius 0.2~nm in Figure~\ref{fig:xassoc-single-rod}. Inside the
-rod (less than 0.2~nm) there are no molecules to form hydrogen bonds although the value
-of $X$ is still calculated there. The number of bonds decreases from a bulk value of 3.5 
-to about 3.25 at the interface. Walther \emph{et al.}\cite{walther2001carbon} 
-calculate the number of hydrogen bonds for a carbon nanotube immersed in water.
-They show a decrease in the bonding from a bulk value of 3.75 to 2.89 near the 
-interface with the nanotube. This is a much larger change than for our 
-rod of a similar radius, indicating that carbon nanotubes disrupt hydrogen
-bonds in water more than hydrophobic rods.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-vs-radius}
-\end{center}
-\caption{ Density profiles for 
-a single hydrophobic rod with a radius of 0.5~nm using several different 
-lattice spacing resolutions. We used a high resolution of 0.05~bohrs ($\approx$~0.003~nm), 
-medium 0.2~bohrs ($\approx$~0.001~nm), and low 0.5~bohrs ($\approx$~0.03~nm).}
-\label{fig:densityresolution}
-\end{figure}
-
-Figure~\ref{fig:energy-vs-diameter} shows the free energy per area versus
-radius for the range of rods studied, with radii from 0.05~nm to 1.5~nm. After 
-about 0.5~nm or so, this reaches
-an asymptote equal to the surface tension. From the inset in 
-Figure~\ref{fig:surface-tension}, the surface tension at room 
-temperature is approximately 72~mN/m. This is very close to the 
-asymptote in Figure~\ref{fig:energy-vs-diameter}, which is about 75~mN/m.
-
-\subsubsection{Resolution and convergence tests}
-Figure~\ref{fig:densityresolution} shows the density profiles for three
-different values of the lattice spacing resolution. The density oscillations
-around a single rod of radius 0.5~nm look nearly identical for all resolutions
-except for the first peak. This contact density is important as it will be
-proportional to the pressure on the rod (see Equation~\ref{contactvaluethm}). The contact
-density calculated using the medium resolution is within less than 4\% of the
-value at high resolution, while the low resolution contact density is off by
-about 5\%. There is some uncertainty in the value of contact density due to the
-different lattice spacings which cause the contact density to be calculated
-at a slightly different position. The total free energy calculations compare much more
-favorably. The medium resolution is within 0.5\% and the low resolution is
-within 1\% of the high resolution free energy.
-
-\subsection{Hydrophobic interaction of two rods}
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-diagram}
-\end{center}
-\caption{ Cross section schematic for two hydrophobic rods immersed in water.
-The radius and separation are shown.}
-\label{fig:rods}
-\end{figure}
-
-We now look at the more interesting problem of two rods in water. We continue to
-use hydrophobic rods of radii 0.3~-~1.0~nm at separation $d$ (see Figure~\ref{fig:rods}), ranging from $d=0$ 
-(touching rods) to $d=1.3$~nm. At small
-separations there is only vapor between the rods. As the rods are
-pulled apart, the vapor region expands until some critical separation is
-reached and liquid water fills the region between the rods. Figure~\ref{fig:density-rods} 
-shows density profiles before and after this transition
-for rods of radius 0.5~nm. This critical separation for the transition to liquid depends
-on the radii of the rods, and is about 0.54~nm for the rods shown in 
-Figure~\ref{fig:density-rods}. 
-Since we are looking at only hard-walled
-non-interacting rods, the critical separation will be
-different for any system where there is an appreciable force between rods. 
-The density oscillations seen around the rods in Figure~\ref{fig:density-rods} 
-are very similar to
-those for the single hydrophobic rod in water (see Figure~\ref{fig:density-single-rod}).
-At small separations, the oscillations around the two rods look as if the two rods were one
-solid ``stadium'' shaped object (a rectangle with semi-circles on the ends).
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/density-rods-in-water}
-\end{center}
-\caption{ Density profiles illustrating the transition from vapor 
-to liquid water between the rods. The radius is 0.5~nm, the top figure is 
-at a separation of 0.2~nm and the
-bottom is 0.6~nm. Figures~\ref{fig:rods-energy-vs-distance}~and~\ref{fig:energy-rods} show
-the energy for these and other separations.}
-\label{fig:density-rods}
-\end{figure}
-
-For one set of rods (radii of 0.5~nm) the free energy is plotted with the internal energy
-and the entropic energy contribution, where all are in units of energy per length 
-(see Figure~\ref{fig:energy-rods}). The free energy slightly increases as the separation
-goes from zero to the critical separation and then is approximately constant at
-larger separations. The entropy has a minimum just before the rods
-transition to allow liquid between them. As the rods separate starting from zero, the area 
-between them containing vapor increases and there is less overall space for the 
-surrounding liquid water to occupy and thus the entropy decreases. When that area fills suddenly
-with liquid, the entropy increases due to the increased space for the water molecules. The
-internal energy also has a minimum just before the transition to the liquid filled state.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/energy-rods}
-\end{center}
-\caption{ Free energy, internal energy, and the entropic energy contribution versus
-separation for two hydrophobic rods of radii 0.5~nm. All energies are arbitrarily
-offset to zero at large separations. }
-\label{fig:energy-rods}
-\end{figure}
-
-The free energy per length is compared for rods of several different radii
-in Figure~\ref{fig:rods-energy-vs-distance}. In this figure, the energy for radius 0.5~nm
-is the same as the free energy shown in Figure~\ref{fig:energy-rods}.
-For comparison, we arbitrarily offset each free energy such that they all go 
-to zero at large separations. For large $d$, the rods will 
-simply act like two separate single rods in water and the energy will
-reach an asymptote determined by the effective surface tension (shown in 
-Figure~\ref{fig:energy-vs-diameter}). For small $d$, we see that all the rods 
-have the same behavior, with a positive linear relationship
-between free energy per length and separation during the initial state 
-(where the rods have vapor between them). This indicates that the force 
-per length on the rods is constant, and surprisingly not dependent on the separation. 
-By finding the slope, we calculate this force per length to be 144~mN/m.
-
-\begin{figure}
-\begin{center}
-\includegraphics[width=\columnwidth]{figs/rods-energy-vs-distance}
-\end{center}
-\caption{ Free energy (also known as the potential of mean force) 
-versus separation for two hydrophobic rods ranging in radius from
-0.3~nm to 1.0~nm.
-All were arbitrarily offset to zero at large separations. The
-transition corresponds to the phase change from
-vapor to liquid between the rods as pictured in the density profiles in 
-Figure~\ref{fig:density-rods}. }
-\label{fig:rods-energy-vs-distance}
-\end{figure}
-
-If we use a classical approximation where the energy on the rod is equal to 
-the surface area multiplied by the bulk surface tension, then the energy per 
-length is equal to the circumference multiplied by the surface tenson. The force
-per length is the derivative of this with respect to the separation. We approximate 
-the circumference of the stadium-shape by
-\begin{equation}
-C_{s} = 2\pi r +4r+2d
-\end{equation}
-and so the force per length is equal to $2\gamma$. This 
-is consistent with the calculation of the force per length from 
-Figure~\ref{fig:rods-energy-vs-distance} (144~mN/m) and the surface tension 
-in Figure~\ref{fig:surface-tension} (72~mN/m). This indicates that
-even though the rods are nanoscale, the water molecules will ``see'' them classically.
-
-To estimate how the transition from vapor to liquid between the rods depends on their 
-radius, we will use this classical approximation and propose that 
-the critical separation, $d_c$, will occur when the free energies are equal. This 
-corresponds to equating the circumference of the stadium shape (see 
-top of Figure~\ref{fig:density-rods}) and the circumference of the two rods (see bottom
-of Figure~\ref{fig:density-rods}), which results in
-\begin{equation}
-d_c = (\pi-2)r.\label{criticalseparation}
-\end{equation}
-The predicted and actual computed values for $d_c$ are shown in 
-Table~\ref{criticalseparationtable}, and they match fairly well for the 
-0.5~-~1.0~nm rods. We hypothesize that larger radii rods
-would also follow this trend, but smaller rods are likely to have a different
-relationship with the surrounding water molecules. This may explain why the critical
-separation result for a radius of 0.3~nm does not match the prediction well.
-
-Walther \emph{et al.}\cite{walther2004hydrodynamic} study the interactions between two
-carbon nanotubes, which are very geometrically similar to our hydrophobic rods. 
-They use molecular dynamics with Lennard-Jones potentials for
-nanotubes of diameter 1.25~nm and separations ranging from about 0.3~nm to 1.5~nm.
-They also find a drying transition occurs, and find it to be at separations near
-0.9~-~1.0~nm\cite{walther2004hydrodynamic}. This is larger than
-what we find for hydrophobic rods of a similar size, as expected due to the carbon-carbon
-and carbon-water interations.
-
-\begin{table}
-\begin{tabular} {|c|c|c|c|}
-\hline
-$r$ (nm) & $d_c$ prediction (nm) & $d_c$ result (nm) & \% error \\
-\hline
-0.3 & 0.34 & 0.23 & 32\% \\
-\hline
-0.5 & 0.57 & 0.54 & 5.3\% \\
-\hline
-0.7 & 0.80 & 0.81 & 1.3\% \\
-\hline
-0.9 & 1.03 & 1.06 & 2.9\% \\
-\hline
-1.0 & 1.14 & 1.19 & 4.4\% \\
-\hline
-\end{tabular}
-\caption{Calculations of the critical separation (see Equation~\ref{criticalseparation})
-for the transition
-from liquid to vapor between two hydrophobic rods.}
-\label{criticalseparationtable}
-\end{table}
+\includegraphics[width=\columnwidth]{figs/combined-coexistence-plot}
 
@@
 
-The classical density functional we developed by combining fundamental-measure 
-theory with SAFT is able to produce reasonable results for hydrophobic interations.
-We added the parameter $\lscale$ to the functional and fit it to the surface tension
-at room temperature, which we find to be consistent with results for larger hydrophobic
-rods.
-
-The free energy per area for water surrounding a hydrophobic rod shows the
-expected behavior with an asymptote equal to the surface tension.
-Based on this result, we expect to be
-able to classically describe rods with a radius of about 0.5~nm or larger. 
-For two hydrophobic rods surrounded by water, we see a transition from a vapor-filled
-stadium shape to two rods with liquid between them. A
-classical treatment of the critical separation for this transition
-works well for rods larger than 0.5~nm. Further evidence for the
-accuracy of this classical treatment is the calculation of the force on the
-stadium shape. The force per length is constant and equal to twice the surface tension.
-
-For the hydrophobic systems studied, this functional is more computationally 
-efficient than molecular dynamics. This allows for future work
-involving hydrophilic interactions with more complex electrostatic potentials.
-
 \bibliographystyle{unsrt}
-\bibliography{paper}% Produces the bibliography via BibTeX.
+\bibliography{refs}% Produces the bibliography via BibTeX.
 

### 2014-03-18 papers/thesis-roth/project.tex

@@
 %% Definitions
+
+% 1/k_B/T
+\newcommand{\kt}{k_BT}
+
+% special values for n
+\newcommand{\npart}{n_{particular}}
+\newcommand{\nliq}{n_{liquid}}
+\newcommand{\nvap}{n_{vapor}}
 
@@
 
+The free energy density for each iteration is given by
+\[
+  f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)\ .
+\]
+
+The total free energy density is obtained when one takes the
+iterations to infinity and adds in the attractive component:
+\[
+  f(T,n) = \lim_{i \to \infty} f_i(T,n) + f_{att}(T,n)\ .
+\]
+
+The initial free energy density, $f_0$ depends on the liquid model of study. For this work, I use the square well model:
+\[
+  f_0(T,n) = f_{id}(T,n) + f_{HS}(T,n) + a_2^{SW}(T,n)\ ,
+\]
+where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kt$.~\cite{Gil-Villegas97}
+
+\subsection{Coexistence Curve Algorithm}
+Coexistence curves are found by modifying the chemical potential $\mu$
+in the grand free energy per volume $\Phi/V$. Grand free energy
+density is defined as
+\begin{align*}
+  \frac{\Phi}{V} &= \phi \\
+                 &= f(T,n) - n\mu \\
+                 &= f(T,n) - n\frac{d}{dn}f(T,\npart)\ .
+\end{align*}
+It is the value $\npart$ that we adjust to find coexistence curves.
+
+The grand free energy density generally has two distinct minima when
+plotted versus density. The liquid will be in liquid-vapor equilibrium
+when those minima have the same value for $\phi$. There is a local
+maximum between those minima, and the value of the density at that
+maximum is used for $\npart$. Find this value at a given temperature
+then plot $\phi$ vs $n$ at slightly higher temperature. Find the new
+value for $\npart$ such that the two minima have the same value for
+$\phi$. (It is helpful to note that if $\npart$ is increased,
+$\phi(\nliq)$ increases, and $\phi(\nliq)$ decreases if $\npart$ is
+decresed.) The program I wrote (npart.py) uses the previous
+temperature's $\npart$ as a ``first guess,'' maximizing
+$\phi(T,\npart)$. From there, the program adjusts $\npart$
+until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
+uses this new value of $\npart$ as the ``first guess'' at the next
+higher temperature.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\includegraphics[width=\columnwidth]{figs/combined-coexistence-plot}
+% \includegraphics[width=\columnwidth]{figs/combined-coexistence-plot}
 
@@
 \bibliographystyle{unsrt}
-\bibliography{refs}% Produces the bibliography via BibTeX.
+\bibliography{project} % Produces the bibliography via BibTeX.
 

### 2014-03-18 papers/pair-correlation/paper.tex

@@
 
-The standard approach in liquid state theory is to treat a liquid as a
+The standard approach in liquid state theory is to model a liquid as a
 hard-sphere reference fluid with attractive interactions that are
@@
 as the cluster, or high-temperature expansion---is an expansion of the
-free energy in powers of a small parameter, which typically is a
+free energy in powers of a small parameter, which is typically a
 pairwise attractive interaction:
@@
 \end{align}
-where the terms $F_n$ are corrections to the free energy at order $n$
+where the terms $F_n$ are corrections to the free energy of order $n$
 in the small interaction.  The first and largest term in this
@@
 the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
-In this paper, we will introduce an approximation for the hard-sphere
-pair distribution function which is suitable for use in the creation
-of classical density functionals based on a cluster expansion.
+In this paper, we will introduce the contact value approximation (CVA)
+for the hard-sphere pair distribution function which is suitable for
+use in the creation of classical density functionals based on a
+cluster expansion. We will then introduce the separable contact value
+approximation (CVA-S), which is based on a radial distribution
+function that is separable in terms of position and local
+density.\pahosays{this could be phrased better, not quite sure the
+  best way}
 
 \section{Theoretical approaches}
+\pahosays{retitle this ``Past theoretical approaches'' or ``current
+  theoretical approaches'' or something to show that ours isn't
+  included?}
 
@@
 \end{align}
-where $\sigma$ is the hard-sphere diameter.  Equation~\ref{eq:fischer}
-is computationally awkward, because it treats as special the midpoint
-$\tfrac12(\rr_1+\rr_2)$.  Moreover, the approach of Fischer and
-Methfessel is intended to approximate the pair distribution function only
-at contact, when the distance between $\rr_1$ and $\rr_2$ is the hard-sphere
-diameter.
+Equation~\ref{eq:fischer} is computationally awkward, because it
+treats as special the midpoint $\tfrac12(\rr_1+\rr_2)$.  Moreover, the
+approach of Fischer and Methfessel is intended to approximate the pair
+distribution function only at contact, when the distance between
+$\rr_1$ and $\rr_2$ is the hard-sphere diameter.
 
@@
 position $\rr$, and $A$ is the Helmholtz free energy of the
-hard-sphere fluid.  Our approximation is obtained by using the White
-Bear functional for Helmholtz free energy of the hard-sphere
-fluid~\cite{roth2002whitebear} in Eq.~\ref{eq:gsigma-exact}.  This
-provides an excellent approximation for this averaged value of the
-pair distribution function for a variety of interfaces, and over a
-wide range of densities.
+hard-sphere fluid.  The contact value approximation is obtained by
+using the White Bear functional for Helmholtz free energy of the
+hard-sphere fluid~\cite{roth2002whitebear} in
+Eq.~\ref{eq:gsigma-exact}.  This provides an excellent approximation
+for this averaged value of the pair distribution function at contact
+for a variety of interfaces, and over a wide range of densities.
 
@@
 %
-We approximate the pair correlation with the average of two radial
-distribution functions, evaluated at the distance between the two
-points, that are themselves functions of the averaged pair correlation
-at contact $g_{\sigma}(\rr)$ evaluated at the two points:
+We construct the CVA with the average of two radial distribution
+functions, evaluated at the distance between the two points, that are
+themselves functions of the averaged pair correlation at contact
+$g_{\sigma}(\rr)$ evaluated at the two points:
 %
@@
 $g^{(2)}(\rr_1,\rr_2)$ above is constructed to reproduce the exact
-value for the integral$ \int n(\rr_1)n(\rr_2)
-g^{(2)}(\rr_1,\rr_2)\delta(|\rr_1-\rr_2|-\sigma)d\rr_1d\rr_2$.  \par
-The approximation described in Eq.~\ref{eq:g2-our-mean} requires the
-radial distribution function as a function of $r$ and $g_\sigma$.  One
-approach to constructing $g(r,g_\sigma)$ is to compute $\eta$ from
-$g_\sigma$ and then find $g(r,\eta)$ either using some approximate
-functional form (e.g. Percus-Yevick) or using Monte Carlo simulation
-results.  In this paper we first use the Monte Carlo results (which
-are also used in evaluating the approximations of earlier works), and
-refer to this approximation (which has an exact $g(r,g_\sigma)$) as
-the ``contact value approximation'' (CVA).  \fixme{Should we mention
-  this earlier?}
+value for the integral:
+\begin{align}
+  \label{eq:1}
+  \int n(\rr_1)n(\rr_2)g_{HS}^{(2)}(\rr_1,\rr_2)\Phi(|\rr_1 -
+  \rr_2|-\sigma)d\rr_1d\rr_2
+\end{align}
+\pahosays{give this integral a name so it's an equation?}
+\par
+The CVA requires the radial distribution function as a function of $r$
+and $g_\sigma$.  One approach to constructing $g(r,g_\sigma)$ is to
+compute $\eta$ from $g_\sigma$ and then find $g(r,\eta)$ either using
+some approximate functional form (e.g. Percus-Yevick) or using Monte
+Carlo simulation results.  In this paper we first use the Monte Carlo
+results (which are also used in evaluating the approximations of
+earlier works). We then use an approximation to the radial
+distribution function that is separable in terms of $r$ and
+$g_\sigma$.
 
@@
 
-The contact value approximation (described in
-Eq.~\ref{eq:g2-our-mean}) is reasonably accurate, but cannot be
-computed efficiently with single-center convolutions as written.  In
-order to ease computational application of this pair distribution
-function in classical density functionals, we have developed a
-separable appoximation for the function $g(r,g_\sigma)$.
+The CVA is reasonably accurate, but cannot be computed efficiently
+with single-center convolutions as written.  In order to ease
+computational application of this pair distribution function in
+classical density functionals, we have developed a separable
+approximation for the function $g(r,g_\sigma)$:
+
+\pahosays{would it make sense to call this $g_s(r,g_\sigma)$ or
+  something similar?}
 \begin{align}
@@
 computed with only single-center convolutions, which can be
-efficiently computed using fast Fourier transforms.  \fixme{Comb paper
-  for places to say CVA-S and CVA.}
+efficiently computed using fast Fourier transforms.
 
@@
 
-We construct $g(r, g_\sigma)$ to satisfy several constraints.  Naturally,
+We construct $g(r, g_\sigma)$ to satisfy several constraints.
+Naturally, we match the value at contact:
 \begin{equation}
@@
 A second constraint applies to the integral of the total correlation
-function
+function, $h(r) = g(r) - 1$:
 \begin{align}
@@
 \end{align}
-where $h(r) = g(r) - 1$ is the total correlation function.
-A final constraint is that the slope of the radial distribution
-function at contact is given to a very good approximation by
+We have found that the slope of the radial distribution function at
+contact is given to a very good approximation by:
 \begin{align}
-  g'(\sigma) \approx - g_\sigma (g_\sigma - 1).
+  g'(\sigma) \approx - g_\sigma (g_\sigma - 1)
 \end{align}
+and we apply this as a final constraint.\par
 Under these three constraints, we applied a least-squares fit to
 simulation data for the radial distribution function at packing
-fractions from 0.05 to 0.4 by steps of 0.05.  Our fitted radial
+fractions from 0.05 to 0.4 in steps of 0.05.  Our fitted radial
 distribution function is given by
@@
 
-\newcommand\paircaption{The pair distribution function near a hard wall, with
-  packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the hard wall.
-  On the left is a 2D plot of $g^{(2)}(\rr_1,\rr_2)$ as $\rr_2$ varies.  The top
-  half of this figure shows the result of the Monte Carlo simulation, while the
-  bottom half shows the result of the appproximation developed in this work.  On
-  the right is a plot of $g^{(2)}(\rr_1,\rr_2)$ on the path illustrated in the
-  figure to the left.  This plot compares the Monte Carlo results (black) with
-  those of this work (blue), Sokolowski and Fischer
-  (red)~\cite{sokolowski1992role}, and Fischer and Methfessel
-  (green)~\cite{fischer1980born}.  The latter is only plotted at contact, where
-  it is defined.  }
+\newcommand{\plotcomp}[1]{The top halves of these figures show the
+  results of Monte Carlo simulations, while the bottom halves show the
+  CVA-S.  On the right are plots of #1 on the paths illustrated in the
+  figures to the left.  These plots compare the CVA-S (blue) and CVA
+  (\fixme{color: change from black?}) with Monte Carlo results (black)
+  and the results of Sokolowski and Fischer
+  (red)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
+  (green)~\cite{fischer1980born}.  The latter is only plotted at
+  contact, where it is defined}
 
@@
   \end{subfigure}
-  \caption{\paircaption}
+  \caption{The pair distribution function near a hard wall, with
+    packing fractions of 0.1 and 0.3 and $\rr_1$ in contact with the
+    hard wall.  On the left are 2D plots of $g^{(2)}(\rr_1,\rr_2)$ as
+    $\rr_2$ varies. \plotcomp{$g^{(2)}(\rr_1,\rr_2)$}.}
   \label{fig:pair-distribution}
@@
 with a focus on the case where one of the two spheres is in contact
-with the hard wall.  Figures~\ref{fig:pair-distribution}a,\ref{fig:pair-distribution}b
-and Figures~\ref{fig:pair-distribution}c,\ref{fig:pair-distribution}d compare the results of our
-approximate functional with Monte Carlo simulation at packing
-fractions of 0.1 and 0.3 respectively.  We see reasonable agreement at
-the lower density, with some reduced oscillations at larger distances,
-and a flatter angular dependence when the two spheres are in contact.
-At the higher density, we see significant structure developing in the
-simulation results that is not reflected in our approximation.
+with the hard wall.  Figures~\ref{fig:pair-distribution}a
+and~\ref{fig:pair-distribution}c compare the results of the CVA-S with
+Monte Carlo simulation at packing fractions of 0.1 and 0.3
+respectively. We see reasonable agreement at the lower density, with
+some reduced oscillations at larger distances, and a flatter angular
+dependence when the two spheres are in contact.  At the higher
+density, we see significant structure developing in the simulation
+results that is not reflected in our approximation.
 
 Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
-show the pair distribution function as plotted along paths
-illustrated in Figures~\ref{fig:pair-distribution}a and
-\ref{fig:pair-distribution}c.  These plots compare this work with
-Monte Carlo results, as well as the approximations of Sokolowski and
-Fischer~\cite{sokolowski1992role} and that of Fischer and
+show the pair distribution function as plotted along paths illustrated
+in Figures~\ref{fig:pair-distribution}a
+and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
+CVA-S with Monte Carlo results, as well as the approximations of
+Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
 Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
@@
 spheres are in contact, and is therefore only plotted on that segment
-of the path.  As an input to both previous approximations we use the
-hard sphere radial distribution function found with Monte Carlo
-simulation, interpolated as necessary.  We find that both previous
-approximations to the pair distribution function predict the
+of the path.  As an input to both previous approximations and to the
+CVA, we use the hard sphere radial distribution function found with
+Monte Carlo simulation, interpolated as necessary. We find that both
+previous approximations to the pair distribution function predict the
 qualitative angular dependence of the pair distribution function at
@@
 more than a radius away from contact, we find that any of these
-approaches gives a reasonable prediction, with our approach
+approaches gives a reasonable prediction, with the CVA-S
 underestimating the oscillations in $g^{(2)}$, as expected based on
@@
 \begin{figure*}
-  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}\\
+  \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
+\\
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
-  \caption{\fixme{Make note of the double curves}\fixme{Add backward curve for
-      this work mc?}The triplet distribution function
-    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when $\rr_1$
-    and $\rr_2$ are in contact.  The color plot on the left shows the triplet
-    distribution function, with white corresponding to the contact value of the
-    \emph{pair} distribution function.  The plot on the right shows the value of
-    $g^{(3)}$ along the path shown as a dotted line on the plot to the
-    left.}\label{fig:triplet-contact-distribution}
+  \caption{The triplet distribution function
+    $g^{(3)}(\rr_1,\rr_2,\rr_3)$ at packing fraction 0.3, plotted when
+    $\rr_1$ and $\rr_2$ are in contact (a,b) and when $\rr_1$ and
+    $\rr_2$ are separated by a distance $2.1\sigma$ (c,d). On the left
+    are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
+    varies. \plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}. Note that an
+    asymmetry is introduced in the theoretical results, so their
+    values along a mirrored path are also shown, with arrows along the
+    curve indicating the path
+    direction.} \label{fig:triplet-contact-distribution}
 \end{figure*}
@@
 homogeneous system can be computed using an approximation of the pair
-distribution for an inhomogeneous fluid, such as we have developed.
+distribution for an inhomogeneous fluid, such as we have
+developed. The triplet distribution function of a homogeneous fluid
+with density $n$ is given by:
 \begin{multline}
@@
 \end{multline}
-where $g^{(3)}(\rr_1,\rr_2,\rr_3)$ is the triplet distribution
-function of a homogeneous fluid with density $n$, and the
-$\textrm{TP}(\rr_1)$ subscript indicates quantities computed for the
-inhomogeneous density configuration in which one sphere is fixed at
-position $\rr_1$.  This method treats one of the three positions---the
-location of the test particle---differently from the others, which
-means that a poor approximation to the pair distribution function may
-beak the symmetry which is present in the true triplet distribution
-function.
+where the $\textrm{TP}(\rr_1)$ subscript indicates quantities computed for
+the inhomogeneous density configuration in which one sphere is fixed
+at position $\rr_1$.  This method treats one of the three
+positions---the location of the test particle---differently from the
+others, which means that a poor approximation to the pair distribution
+function may beak the symmetry which is present in the true triplet
+distribution function.
 
 Figure~\ref{fig:triplet-contact-distribution} compares the triplet
-distribution function at a packing fraction of 0.3 computed using our
-approximation to $g^{(2)}$ with results from Monte Carlo simulation.
-In each subfigure, $\rr_1$ and $\rr_2$ are held fixed, while the third
-position $\rr_3$ is varied.  The test-particle position in each case
-is $\rr_1$, which is on the left-hand side of the figure.  In the
-color plot on the left, the first and second spheres are displayed in
-grey.  On the bottom half of
-figures~\ref{fig:triplet-contact-distribution}a and
-\ref{fig:triplet-contact-distribution}c is the triplet distribution
-function as computed using the approximation presented in this work.
-The top half of those figures displays the triplet distribution
-function from Monte Carlo.
-Figures~\ref{fig:triplet-contact-distribution}b
+distribution function at a packing fraction of 0.3 computed using the
+CVA-S with results from Monte Carlo simulation.  In each subfigure,
+$\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
+is varied.  The test-particle position in each case is $\rr_1$, which
+is on the left-hand side of the figure.  In the color plots on the
+left, the first and second spheres are displayed in grey.  On the
+bottom half of figures~\ref{fig:triplet-contact-distribution}a
+and~\ref{fig:triplet-contact-distribution}c is the triplet
+distribution function as computed using the CVA-S.  The top half of
+those figures displays the triplet distribution function from Monte
+Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
 and~\ref{fig:triplet-contact-distribution}d show the triplet
@@
 and~\ref{fig:triplet-contact-distribution}b show the triplet
-distribution
-function when two of the positions $\rr_1$ and $\rr_2$ are in contact.
-The triplet distribution function predicted by this work is slightly
-asymmetric, indicating a failure in thermodynamic consistency.  Once
-again, we observe the lack of oscillations at large distances.
-Figures~\ref{fig:triplet-contact-distribution}c
-and~\ref{fig:triplet-contact-distribution}d show the triplet distribution
-function when two of the positions $\rr_1$ and $\rr_2$ are spaced so
-that the third could just fit between them.  Again, the triplet
-distribution function predicted by this work is slightly asymmetric,
-and lacks the oscillations seen in the Monte Carlo simulations at
-large distances.  In both cases, our agreement with simulation is
-comparable to that of previous approximations.
+distribution function when two of the positions $\rr_1$ and $\rr_2$
+are in contact.  The triplet distribution function predicted by the
+CVA-S is slightly asymmetric, indicating a failure in thermodynamic
+consistency.  Once again, we observe the lack of oscillations at large
+distances.  Figures~\ref{fig:triplet-contact-distribution}c
+and~\ref{fig:triplet-contact-distribution}d show the triplet
+distribution function when two of the positions $\rr_1$ and $\rr_2$
+are spaced so that the third could just fit between them.  Again, the
+triplet distribution function predicted by the CVA-S is slightly
+asymmetric, and lacks the oscillations seen in the Monte Carlo
+simulations at large distances.  In both cases, our agreement with
+simulation is comparable to that of previous approximations.
 
@@
 potential.  To this end, we have computed the error in the first term
-in a high-temperature perturbation expansion $a_1$ for several pair
-potentials.  In order to focus on effects at the interface, we have
-defined a position-dependent pair interaction energy as
+in a high-temperature perturbation expansion $a_1$
+\pahosays{we sometimes call this $a_1$ and sometimes $F_1$. Should we
+  change it to $F_1$ everywhere?}
+for several pair potentials.  In order to focus on effects at the
+interface, we have defined a position-dependent pair interaction
+energy as
 \begin{align}
@@
 
-We plot in Fig.~\ref{fig:dadz} this quantity for hard-core fluids with
-three representative pair potentials near a hard wall.  We have chosen
-to illustrate a delta-function ineraction at contact (i.e. ``sticky
-hard spheres''), a hard-core square-well fluid, with the length-scale
-of interaction taken from the optimal SAFT model for water found by
-Clark \emph{et al.}~\cite{clark2006developing}, and a $1/r^6$
-power-law attraction.  These pair potentials represent a short-range
-interaction, a medium-range discontinuous interaction, and a
-long-range smooth interaction.
+We plot this quantity for hard-core fluids with three representative
+pair potentials near a hard wall in Fig.~\ref{fig:dadz}.  We have
+chosen to illustrate a delta-function ineraction at contact
+(i.e. ``sticky hard spheres''); a hard-core square-well fluid, with
+the length-scale of interaction taken from the optimal SAFT model for
+water found by Clark \emph{et al.}~\cite{clark2006developing}; and a
+$1/r^6$ power-law attraction.  These pair potentials represent a
+short-range interaction, a medium-range discontinuous interaction, and
+a long-range smooth interaction.
 
 Figure~\ref{fig:dadz-delta} shows the results for the sticky
-hard-sphere fluid.  Our pair distribution function is constructed to
-produce this result exactly, provided the averaged pair distribution
-function at contact from Ref.~\citenum{schulte2012using} is exact.  As
-expected, we see excellent agreement with the Monte Carlo simulation
-results, while the approximations of Fischer and Sokolowski each show
+hard-sphere fluid.  The CVA and CVA-S are constructed to produce this
+result exactly, provided the averaged pair distribution function at
+contact from Ref.~\citenum{schulte2012using} is exact.  As expected,
+we see excellent agreement with the Monte Carlo simulation results,
+while the approximations of Fischer and Sokolowski each show
 deviations near the interface.  Figure~\ref{fig:dadz-square-well}
 shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
-This potential stresses our approximation to the radial distribution
-function, since the edge of the attraction occurs at the dip in the
-radial distribution function at this density, with poor results for
-the bulk fluid.  We see that over these length scales, Sokolowski's
-approximation is very accurate.  Finally, in the case of the power-law
-attraction both approximations work very well.
+This potential stresses the CVA-S, since the edge of the
+attraction occurs at the dip in the radial distribution function at
+this density, with poor results for the bulk fluid.  We see that over
+these length scales, Sokolowski's approximation is very accurate.
+Finally, in the case of the power-law attraction both approximations
+work very well.
 
@@
 
-We have introduced and tested an approximation for the pair
+We have introduced and tested the contact value approximation for the pair
 distribution function of the inhomogeneous hard-sphere fluid.  Our
 approximation is suitable for use in the development of classical
-density functionals based constructed using perturbation theory, and
+density functionals constructed using perturbation theory, and
 may be computed using exclusively single-center convolutions.  We have
@@
 
+\pahosays{Sometimes we call the pair distribution function
+  $g^{(2)}_{HS}(\rr_1,\rr_2)$ and sometimes $g^{(2)}(\rr_1,\rr_2)$. I feel
+  like we should stick to one for consistency, but I don't know
+  which.}
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2014-03-19 papers/thesis-roth/project.tex

@@
 The free energy density for each iteration is given by
-\[
+\begin{equation}
   f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)\ .
-\]
-
+\end{equation}
 The total free energy density is obtained when one takes the
 iterations to infinity and adds in the attractive component:
-\[
+\begin{equation}
   f(T,n) = \lim_{i \to \infty} f_i(T,n) + f_{att}(T,n)\ .
-\]
-
+\end{equation}
 The initial free energy density, $f_0$ depends on the liquid model of study. For this work, I use the square well model:
-\[
+\begin{equation}
   f_0(T,n) = f_{id}(T,n) + f_{HS}(T,n) + a_2^{SW}(T,n)\ ,
-\]
+\end{equation}
 where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kt$.~\cite{Gil-Villegas97}
@@
 density is defined as
-\begin{align*}
-  \frac{\Phi}{V} &= \phi \\
-                 &= f(T,n) - n\mu \\
+\begin{align}
+  \frac{\Phi}{V} &= \phi \nonumber \\
+                 &= f(T,n) - n\mu \nonumber \\
                  &= f(T,n) - n\frac{d}{dn}f(T,\npart)\ .
-\end{align*}
+\end{align}
 It is the value $\npart$ that we adjust to find coexistence curves.

### 2014-03-24 papers/square-well-liquid/histogram-paper.tex
add support for canonical simulations to square-well MC, and add (mostly empty) "histogram" paper Tested on darcy. Took 0:19.55.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{color}

\begin{document}
\title{Applying a clever histogram method to the square-well liquid}

\author{Michael MI? Perlin}
\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

\begin{abstract}
  We have applied the clever histogram method to the square-well liquid.
\end{abstract}

\maketitle

\begin{figure}
  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}\
  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
  \caption{A histogram plot demonstrating the difficulty of using
    canonical Monte Carlo to simulate a system.  A simulation at a
    give temperature only provides information fo a small number of
    energy states.\label{fig:histograms}}
\end{figure}

As shown in Fig.~\ref{fig:histograms}, there are several difficulties
encountered when running Monte Carlo simulations at a fixed
temperature.  On the top plot, which shows three simulations run at
$\lambda = 1.3$, it is apparent that each given canonical simulation
only provides statistical information for a small number of energy
states.  The bottom plot shows an even more serious issue:  when
running a simulation at a fixed temperature, it is possible to become
frozen in a state that is far from the ground state, as happened in
the $kT=0.1\epsilon$ simulation.

\end{document}

### 2014-03-24 papers/pair-correlation/paper.tex
switch to cyan for CVA approximation Tested on darcy. Took 0:21.47.

@@
   CVA-S.  On the right are plots of #1 on the paths illustrated in the
-  figures to the left.  These plots compare the CVA-S (blue) and CVA
-  (\fixme{color: change from black?}) with Monte Carlo results (black)
+  figures to the left.  These plots compare the CVA-S (blue solid line) and CVA
+  (cyan dotted line) with Monte Carlo results (black circle)
   and the results of Sokolowski and Fischer
-  (red)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
-  (green)~\cite{fischer1980born}.  The latter is only plotted at
+  (red dashed line)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
+  (green dot-dashed line)~\cite{fischer1980born}.  The latter is only plotted at
   contact, where it is defined}

### 2014-03-24 papers/pair-correlation/paper.tex
respond to Pahosays in pair paper Tested on darcy. Took 0:21.39.

@@
 the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
-In this paper, we will introduce the contact value approximation (CVA)
-for the hard-sphere pair distribution function which is suitable for
-use in the creation of classical density functionals based on a
-cluster expansion. We will then introduce the separable contact value
-approximation (CVA-S), which is based on a radial distribution
-function that is separable in terms of position and local
-density.\pahosays{this could be phrased better, not quite sure the
-  best way}
+In this paper, we introduce a contact value approximation (CVA) for
+the hard-sphere pair distribution function which is suitable for use
+in the creation of classical density functionals based on
+thermodynamic perturbation theory. We then demonstrate a more
+computationally efficient separable version of the contact
+value approximation (CVA-S), which is based on a fit to the radial
+distribution function that is separable in terms of position and local
+density.
 
-\section{Theoretical approaches}
-\pahosays{retitle this ``Past theoretical approaches'' or ``current
-  theoretical approaches'' or something to show that ours isn't
-  included?}
+\section{Previous theoretical approaches}
 
@@
 \end{align}
-$g^{(2)}(\rr_1,\rr_2)$ above is constructed to reproduce the exact
-value for the integral:
+This contact value approximation for $g^{(2)}(\rr_1,\rr_2)$ is
+constructed to reproduce the exact value for the integral:
 \begin{align}
-  \label{eq:1}
-  \int n(\rr_1)n(\rr_2)g_{HS}^{(2)}(\rr_1,\rr_2)\Phi(|\rr_1 -
-  \rr_2|-\sigma)d\rr_1d\rr_2
+  F_1^{\text{contact}} &= \tfrac12 \iint
+  g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\delta(|\rr_1-\rr_2|-\sigma)
+  d\rr_1d\rr_2
+  \label{eq:mean-field-contact}
 \end{align}
-\pahosays{give this integral a name so it's an equation?}
-\par
+which is the mean-field correction to the free energy (see
+Eq.~\ref{eq:mean-field}) for a purely contact interaction.
+
 The CVA requires the radial distribution function as a function of $r$
@@
 Carlo simulation results.  In this paper we first use the Monte Carlo
-results (which are also used in evaluating the approximations of
-earlier works). We then use an approximation to the radial
+results, which are also used in evaluating the approximations of
+earlier works.  This approximation we will refer to as CVA in the
+remainder of the text.  We then use an approximation to the radial
 distribution function that is separable in terms of $r$ and
-$g_\sigma$.
+$g_\sigma$, and additional approximation that we will refer to as
+CVA-S, which we will describe in the next section.
 
@@
 approximation for the function $g(r,g_\sigma)$:
-
-\pahosays{would it make sense to call this $g_s(r,g_\sigma)$ or
-  something similar?}
 \begin{align}
-  g(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
 \end{align}
@@
   \caption{Plot of the hard-sphere radial distribution function, with
-    our fit.}\label{fig:radial-distribution}
+    our separable fit.  The solid lines show the true radial
+    distribution function $g(r)$, and the dashed lines show our
+    separable fit to the radial distribution functions
+    $g_S(r)$.}\label{fig:radial-distribution}
 \end{figure}
 
-We construct $g(r, g_\sigma)$ to satisfy several constraints.
+We construct $g_S(r, g_\sigma)$ to satisfy several constraints.
 Naturally, we match the value at contact:
 \begin{equation}
-  g(\sigma, g_\sigma) = g_\sigma.
+  g_S(\sigma, g_\sigma) = g_\sigma.
 \end{equation}
@@
 \begin{multline}
-  g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  g_S(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
   + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
@@
 Eq.~\ref{eq:total-constraint} results in cancelling errors at large
-distances.  The maximum error in $g(r)$ at packing fractions up to
+distances.  The maximum error in $g_S(r)$ at packing fractions up to
 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
@@
   \end{subfigure}
-  \caption{Plot of $\frac{da_1}{dz}$ near a hard wall.  (a) shows a
+  \caption{Plot of $\frac{dF_1}{dz}$ near a hard wall.  (a) shows a
     sticky hard-sphere fluid defined by a pair potential
@@
 potential.  To this end, we have computed the error in the first term
-in a high-temperature perturbation expansion $a_1$
-\pahosays{we sometimes call this $a_1$ and sometimes $F_1$. Should we
-  change it to $F_1$ everywhere?}
+in a high-temperature perturbation expansion $F_1$
 for several pair potentials.  In order to focus on effects at the
@@
   which.}
+\davidsays{The reason for the distinction is that when we consider a
+  real liquid, then the true $g(r)$ is different from $g_{HS}(r)$, so
+  the mean-field equations ought to specify that they use the
+  hard-sphere pair distribution function.  But when we're talking
+  about the hard-sphere fluid, $g(r) = g_{HS}(r)$ and it's clumsy to
+  always have the HS present. }
 

### 2014-03-24 papers/pair-correlation/paper.tex
updates to pair paper Tested on bennet. Took 0:21.45.

@@
   distribution function as well as the most accurate existing
-  analytical approximation.  \fixme{needs special attention.}
-  \fixme{add acronym, and discuss separable vs. non}
+  analytical approximation.  \davidsays{Jeff will rewrite this
+    abstract.}
 \end{abstract}
@@
 \newcommand\saftnonlocaldft{}
+
+\davidsays{Everyone reads a print copy of the the paper by Tuesday
+  evening.  Then Paho submits!}
+
+\davidsays{Paho volunteered to check over and figure out what to do
+  with the $HS$ in $g$.  Can we get away with $HS$ everywhere?
+  Defining two macros so that we can go back and forth.}
 
@@
 the pair distribution function of the reference fluid:
-$g^{(2)}(\rr_1,\rr_2)$.  Unlike the radial distribution function of a
+$g_{HS}^{(2)}(\rr_1,\rr_2)$.  Unlike the radial distribution function of a
 homogeneous fluid, there does not currently exist a tractable form for

### 2014-03-25 papers/pair-correlation/paper.tex
rewrote abstract and added a jeffsays Tested on crystal. Took 0:37.07.

@@
 \begin{abstract}
-  We have developed an approximation for the pair distribution
-  function of hard spheres, which is based on the average value of the
-  pair distribution function at contact we recently developed.  Our
-  pair distribution function is constructed so as to be computable
-  with only single-site convolutions and to accurately reproduce the
-  averaged distribution function at contact.  This results in a
-  functional that is efficient to compute, and predicts the pair
-  distribution function as well as the most accurate existing
-  analytical approximation.  \davidsays{Jeff will rewrite this
-    abstract.}
+We introduce an approximation for the pair distribution function of
+the inhomogeneous hard sphere fluid. Instead of relying on simple
+convolutions of density, our approximation incorporates a radial
+distribution function that is dependant upon a recently published
+averaged pair distribution function at contact which has been shown to
+give good results. We develop a separable form of this radial
+distribution function that achieves greater computational efficiency
+by allowing for the use of fast Fourier transforms. We compare results
+for this separable version of our pair distribution approximation with
+a Monte-Carlo based version, two previously published works, and
+Monte-Carlo simulation, showing favorable results.  \jeffsays{In this
+  I refer to the 'radial distribution function at contact'.  In the
+  text (and in our previous paper) this is refered to using
+  'correlation' versus 'distribution'.  I'm confused about how we
+  switch back and forth in this paper.  Is it that we always refer to
+  the contact pair function as correlation but the others as
+  distribution?  Otherwise I didn't see any problems with the text.}
 \end{abstract}

### 2014-03-25 papers/pair-correlation/paper.tex
address a Jeffsays Tested on bennet. Took 0:25.37.

@@
 a Monte-Carlo based version, two previously published works, and
-Monte-Carlo simulation, showing favorable results.  \jeffsays{In this
-  I refer to the 'radial distribution function at contact'.  In the
-  text (and in our previous paper) this is refered to using
-  'correlation' versus 'distribution'.  I'm confused about how we
-  switch back and forth in this paper.  Is it that we always refer to
-  the contact pair function as correlation but the others as
-  distribution?  Otherwise I didn't see any problems with the text.}
+Monte-Carlo simulation, showing favorable results.
 \end{abstract}
@@
 \section{Contact value approximation}
-%% We approximate the pair correlation with the average of two radial
-%% distribution functions, evaluated at the distance between the two
-%% points, that are themselves functions of the averaged pair correlation
-%% at contact $g_{\sigma}(\rr)$ discussed above~\cite{schulte2012using},
-%% evaluated at the two points.
 In the approaches for the pair distribution function mentioned above,
-the radial distribution function used to approximate pair correlation
+the radial distribution function used in the approximation
 was dependent upon the density averaged over some volume.  We seek to
@@
 functions, evaluated at the distance between the two points, that are
-themselves functions of the averaged pair correlation at contact
+themselves functions of the averaged pair distribution function at contact
 $g_{\sigma}(\rr)$ evaluated at the two points:
@@
 \input{figs/fit-parameters}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
+  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
+  \caption{Plot of the hard-sphere radial distribution function, with
+    our separable fit.  The solid lines show the true radial
+    distribution function $g(r)$, and the dashed lines show our
+    separable fit to the radial distribution functions
+    $g_S(r)$.}\label{fig:radial-distribution}
+\end{figure}
 
@@
 %% \end{align}
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
-  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
-  \caption{Plot of the hard-sphere radial distribution function, with
-    our separable fit.  The solid lines show the true radial
-    distribution function $g(r)$, and the dashed lines show our
-    separable fit to the radial distribution functions
-    $g_S(r)$.}\label{fig:radial-distribution}
-\end{figure}
 

### 2014-03-25 papers/pair-correlation/paper.tex
edits with Paho Tested on bennet. Took 0:17.18.

@@
 We introduce an approximation for the pair distribution function of
-the inhomogeneous hard sphere fluid. Instead of relying on simple
-convolutions of density, our approximation incorporates a radial
-distribution function that is dependant upon a recently published
-averaged pair distribution function at contact which has been shown to
-give good results. We develop a separable form of this radial
-distribution function that achieves greater computational efficiency
-by allowing for the use of fast Fourier transforms. We compare results
-for this separable version of our pair distribution approximation with
-a Monte-Carlo based version, two previously published works, and
-Monte-Carlo simulation, showing favorable results.
+the inhomogeneous hard sphere fluid. Our approximation makes use of
+our recently published averaged pair distribution function at contact
+which has been shown to accurately reproduce the averaged pair
+distribution function at contact for inhomogeneous density
+distributions. We further develop a separable form of this
+approximation that achieves greater computational efficiency by using
+exclusively fixed-kernel convolutions and thus allowing an
+implementation using fast Fourier transforms. We compare results for
+both versions of our pair distribution approximation with two
+previously published works and Monte-Carlo simulation, showing
+favorable results.
 \end{abstract}
@@
 
-\newcommand\saftlocaldft{}
-% The following are papers that use a SAFT-based classical DFT with
-% all the terms that should be non-local being non-local.
-\newcommand\saftnonlocaldft{}
-
-\davidsays{Everyone reads a print copy of the the paper by Tuesday
-  evening.  Then Paho submits!}
-
-\davidsays{Paho volunteered to check over and figure out what to do
-  with the $HS$ in $g$.  Can we get away with $HS$ everywhere?
-  Defining two macros so that we can go back and forth.}
+\section{Introduction}
 
@@
 value approximation (CVA-S), which is based on a fit to the radial
-distribution function that is separable in terms of position and local
-density.
+distribution function that is separable in a way that will enable
+efficient evaluation of the integral in Eq.~\ref{eq:mean-field}.
 
@@
 distribution function is to use Percus' trick of treating one sphere
-as an external field, and using the resultant equilibrium density to
+as an external field, and to use the resultant equilibrium density to
 find the pair distribution function~\cite{hansen2006theory}.  This
@@
 G{\"o}tzelmann \emph{et al.} found the pair distribution function near
-a hard wall by solving the Ornstein-Zernicke equation using the direct
+a hard wall by solving the Ornstein-Zernike equation using the direct
 correlation function found from a functional derivative of a classical
 DFT free energy functional~\cite{gotzelmann1996structure}.
-Unfortunately, solving the inhomogeneous Ornstein-Zernicke equation is
+Unfortunately, solving the inhomogeneous Ornstein-Zernike equation is
 computationally challenging, although more efficient approximate
@@
 Another inhomogeneous configuration that is of interest is the
-test-particle configuration, in which one hard sphere is fixed in the
-otherwise homogeneous fluid.  Where the hard-wall is a surface with no
-curvature, the test-particle configuration has a surface with
-curvature at the molecular length scale.  In this case, the density
-gives the pair distribution function---this is just Percus'
-trick---and the pair distribution function gives the triplet
-distribution function.  The triplet distribution function has been computed
-by Gonz\'alez \emph{et al.} using the test-particle approach with
+test-particle configuration, in which one hard sphere is fixed.  Where
+the hard-wall is a surface with no curvature, the test-particle
+configuration has a surface with curvature at the molecular length
+scale.  In this case, the density gives the pair distribution
+function---this is just Percus' trick---and the pair distribution
+function of this inhomogeneous test-particle system gives the triplet
+distribution function of the homogeneous fluid.  The triplet
+distribution function of the homogeneous fluid has been computed by
+Gonz\'alez \emph{et al.} using the test-particle approach with
 \emph{two} spheres fixed~\cite{gonzalez1999test}.
@@
 
-\section{Approximations to the pair distribution function}
+\section{Previous approximations to the pair distribution function}
 There are several analytic approximations for the inhomogeneous pair
@@
 
-Several recent works have assumed a mean-density approximation for
-the pair distribution function:
+Several recent works~\cite{gloor2007prediction, gross2009density} have
+assumed a mean-density approximation for the pair distribution
+function:
 \begin{align}
@@
 radius and packing fraction $\eta$, and $\sigma$ is the hard-sphere
-diameter~\cite{gloor2007prediction, gross2009density}.  This approach,
+diameter.  This approach,
 however, cannot be applied to highly inhomogeneous systems such as a
@@
 volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
-approximation
+approximation:
 \begin{align}
@@
 \end{align}
-where
+where their averaged density $\bar{n}(\rr)$ is given by:
 \begin{align}
-  \bar{n}(\rr) = \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
+  \bar{n}(\rr) \equiv \frac{3}{4\pi (0.8\sigma)^3}\int n(\rr')\Theta(0.8\sigma - |\rr-\rr'|) d\rr'
 \end{align}
@@
 
-In a previous paper~\cite{schulte2012using}, we have introduced a
+In a previous paper~\cite{schulte2012using}, we introduced a
 functional that gives a good approximation for the pair distribution
@@
 \end{align}
-where the averaged density $\tilde{n}(\rr_1)$ is defined by
+where the weighted density $\tilde{n}(\rr_1)$ is defined by:
 \begin{align}
-  \tilde{n}(\rr_1) &\equiv \int \delta(\sigma -|\rr_1 - \rr_2|)n(\rr_2) d\rr_2.
+  \tilde{n}(\rr) &\equiv \int n(\rr') \delta(\sigma -|\rr - \rr'|)d\rr'.
 \end{align}
@@
   &= \frac12 \frac{1}{k_BT n(\rr) \tilde{n}(\rr)} \frac{\delta
-    A}{\delta \sigma(\mathbf{r})} \label{eq:gsigma-exact}
+    F_{HS}}{\delta \sigma(\mathbf{r})} \label{eq:gsigma-exact}
 \end{align}
 where $\sigma(\rr)$ is the diameter of hard spheres located at
-position $\rr$, and $A$ is the Helmholtz free energy of the
-hard-sphere fluid.  The contact value approximation is obtained by
-using the White Bear functional for Helmholtz free energy of the
-hard-sphere fluid~\cite{roth2002whitebear} in
-Eq.~\ref{eq:gsigma-exact}.  This provides an excellent approximation
-for this averaged value of the pair distribution function at contact
-for a variety of interfaces, and over a wide range of densities.
+position $\rr$, and $F_{HS}$ is the Helmholtz free energy of the
+hard-sphere fluid.  The approximation for $g_\sigma(\rr)$ is obtained
+by using the White Bear functional~\cite{roth2002whitebear} for the
+Helmholtz free energy $F_{HS}$ in Eq.~\ref{eq:gsigma-exact}.  This
+provides an excellent approximation for this averaged value of the
+pair distribution function at contact for a variety of interfaces, and
+over a wide range of densities.
 
@@
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) +
-    g(|r_{12}|, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
+    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
 \end{align}
@@
 and $g_\sigma$.  One approach to constructing $g(r,g_\sigma)$ is to
-compute $\eta$ from $g_\sigma$ and then find $g(r,\eta)$ either using
-some approximate functional form (e.g. Percus-Yevick) or using Monte
-Carlo simulation results.  In this paper we first use the Monte Carlo
-results, which are also used in evaluating the approximations of
-earlier works.  This approximation we will refer to as CVA in the
-remainder of the text.  We then use an approximation to the radial
-distribution function that is separable in terms of $r$ and
-$g_\sigma$, and additional approximation that we will refer to as
-CVA-S, which we will describe in the next section.
+compute $\eta$ from $g_\sigma$ and then find $g(r,\eta)$ using either
+some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
+simulation results.  In this paper we first use the Monte Carlo
+results for the radial distribution function $g(r)$, which we also use
+in evaluating the approximations of earlier works.
 
@@
 
-\section{Separable radial distribution function for the CVA}
+\section{Separable radial distribution function for the CVA-S}
 
-The CVA is reasonably accurate, but cannot be computed efficiently
-with single-center convolutions as written.  In order to ease
+The CVA is quite accurate, but cannot be computed efficiently
+with fixed-kernel convolutions as written.  In order to ease
 computational application of this pair distribution function in
@@
 field contribution to the free energy (Eq.~\ref{eq:mean-field}) to be
-computed with only single-center convolutions, which can be
+computed with only fixed-kernel convolutions, which can be
 efficiently computed using fast Fourier transforms.
@@
 \end{align}
+where $\chi_T$ is the isothermal compressibility of the hard-sphere fluid.
 We have found that the slope of the radial distribution function at
@@
 \begin{align}
-  g'(\sigma) \approx - g_\sigma (g_\sigma - 1)
+  g'(\sigma) \approx g_\sigma (1 - g_\sigma)
 \end{align}
@@
 \begin{multline}
-  g_S(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
-  + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+  g_S(r,g_\sigma) = 1 + (g_\sigma-1) e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  + (g_\sigma -1)(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
   + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
@@
   I(g_\sigma) = \kappaZ^5 \frac{
-    \frac{\chi-1}{24\eta} - h_\sigma \frac{(\kappaO-2
+    \frac{\chi-1}{24\eta} + (1-g_\sigma) \frac{(\kappaO-2
       g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
-    - h_\sigma\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
+    + (1-g_\sigma)\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
   }{2 \kappaZ^2 + 12 \kappaZ + 24}
@@
 The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
-figure, it is apparent that the integral of $h(r)$ from
-Eq.~\ref{eq:total-constraint} results in cancelling errors at large
+figure, it is apparent that the constraint on the integral of $h(r)$ from
+Eq.~\ref{eq:total-constraint} results in canceling errors at large
 distances.  The maximum error in $g_S(r)$ at packing fractions up to
 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
-\rmaxerr$.
+\rmaxerr R$.
 
@@
   contact, where it is defined}
-
-\section{Results}
 
@@
 \end{figure*}
+
+\section{Results}
 
@@
 and~\ref{fig:pair-distribution}c compare the results of the CVA-S with
-Monte Carlo simulation at packing fractions of 0.1 and 0.3
+Monte Carlo simulations at packing fractions of 0.1 and 0.3
 respectively. We see reasonable agreement at the lower density, with
@@
 results that is not reflected in our approximation.
-
-Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
-show the pair distribution function as plotted along paths illustrated
-in Figures~\ref{fig:pair-distribution}a
-and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
-CVA-S with Monte Carlo results, as well as the approximations of
-Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
-Methfessel~\cite{fischer1980born} at packing fractions of 0.1 and 0.3.
-The approach of Fischer and Methfessel is only defined when the two
-spheres are in contact, and is therefore only plotted on that segment
-of the path.  As an input to both previous approximations and to the
-CVA, we use the hard sphere radial distribution function found with
-Monte Carlo simulation, interpolated as necessary. We find that both
-previous approximations to the pair distribution function predict the
-qualitative angular dependence of the pair distribution function at
-contact better than this work.  However, in each case the pair
-distribution function has a systematic error at contact---either too
-high or too low.  Thus while our approximation is smoother than either
-of the existing approaches, its errors will have a tendancy to cancel
-when used in a perturbation expansion.  At higher densities, the
-approximation of Fischer and Methfessel requires evaluating the radial
-distribution function at densities significantly higher than the
-freezing density, which poses numerical difficulties when using the
-radial distribution function from simulation.  When the two points are
-more than a radius away from contact, we find that any of these
-approaches gives a reasonable prediction, with the CVA-S
-underestimating the oscillations in $g^{(2)}$, as expected based on
-the fit in Fig.~\ref{fig:radial-distribution}.
-
-\subsection{Triplet distribution function}
 
@@
     curve indicating the path
-    direction.} \label{fig:triplet-contact-distribution}
+    direction. \fixme{Explain paths and arrows more.}} \label{fig:triplet-contact-distribution}
 \end{figure*}
+
+Figures~\ref{fig:pair-distribution}b and~\ref{fig:pair-distribution}d
+show the pair distribution function as plotted along paths illustrated
+in Figures~\ref{fig:pair-distribution}a
+and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
+CVA-S with Monte Carlo results, as well as the approximations of
+Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
+Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1 and 0.3.
+The approach of Fischer and Methfessel is only defined when the two
+spheres are in contact, and is therefore only plotted on that segment
+of the path.  As an input to both previous approximations and to the
+CVA, we use the hard sphere radial distribution function found with
+Monte Carlo simulation, interpolated as necessary. We find that both
+previous approximations to the pair distribution function predict the
+qualitative angular dependence of the pair distribution function at
+contact better than this work.  However, in each case the pair
+distribution function has a systematic error at contact---either too
+high or too low.  While our approximation is smoother than either
+of the existing approaches, its errors will have a tendency to cancel
+when used in a perturbation expansion.  At higher densities, the
+approximation of Fischer and Methfessel requires evaluating the radial
+distribution function at densities significantly higher than the
+freezing density, which poses numerical difficulties when using the
+radial distribution function from simulation.  When the two points
+$\rr_1$ and $\rr_2$ are both
+more than a radius away from contact, we find that any of these
+approaches gives a reasonable prediction, with the CVA-S
+underestimating the oscillations in $g_{HS}^{(2)}$, as expected based on
+the fit in Fig.~\ref{fig:radial-distribution}.
+
+\subsection{Triplet distribution function}
 
@@
 where the $\textrm{TP}(\rr_1)$ subscript indicates quantities computed for
-the inhomogeneous density configuration in which one sphere is fixed
+the inhomogeneous density configuration in which one sphere (the
+``test particle'') is fixed
 at position $\rr_1$.  This method treats one of the three
@@
 others, which means that a poor approximation to the pair distribution
-function may beak the symmetry which is present in the true triplet
+function may break the symmetry which is present in the true triplet
 distribution function.
@@
 is on the left-hand side of the figure.  In the color plots on the
-left, the first and second spheres are displayed in grey.  On the
+left, the first and second spheres are displayed in gray.  On the
 bottom half of figures~\ref{fig:triplet-contact-distribution}a
@@
 asymmetric, and lacks the oscillations seen in the Monte Carlo
-simulations at large distances.  In both cases, our agreement with
+simulations at large distances.  \fixme{We need to discuss the plots
+  and the directions and arrows.} In both cases, our agreement with
 simulation is comparable to that of previous approximations.
@@
 
-We plot this quantity for hard-core fluids with three representative
+We plot this quantity for three representative
 pair potentials near a hard wall in Fig.~\ref{fig:dadz}.  We have
-chosen to illustrate a delta-function ineraction at contact
+chosen to illustrate a delta-function interaction at contact
 (i.e. ``sticky hard spheres''); a hard-core square-well fluid, with
@@
 shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
-This potential stresses the CVA-S, since the edge of the
-attraction occurs at the dip in the radial distribution function at
-this density, with poor results for the bulk fluid.  We see that over
-these length scales, Sokolowski's approximation is very accurate.
-Finally, in the case of the power-law attraction both approximations
-work very well.
+This potential stresses the CVA-S, since the edge of the attraction
+occurs at the dip in the radial distribution function at this density,
+with poor results for the bulk fluid.  We see that over these length
+scales, both the CVA and Sokolowski's approximation are very accurate.
+Finally, in the case of the power-law attraction all approximations
+tested work very well.
 
@@
 We have introduced and tested the contact value approximation for the pair
-distribution function of the inhomogeneous hard-sphere fluid.  Our
+distribution function of the inhomogeneous hard-sphere fluid.  The
+separable version of the contact value
 approximation is suitable for use in the development of classical
-density functionals constructed using perturbation theory, and
-may be computed using exclusively single-center convolutions.  We have
+density functionals constructed using perturbation theory, as it
+may be efficiently computed using exclusively fixed-kernel convolutions.  We have
 tested this function at a hard wall and near a single fixed hard
@@
 Tests of the pair distribution function in integrals that arise in
-thermodynamic perturbation theory suggest that our approximation gives
-good results for both short-range and long-range attractions.
-
-\pahosays{Sometimes we call the pair distribution function
-  $g^{(2)}_{HS}(\rr_1,\rr_2)$ and sometimes $g^{(2)}(\rr_1,\rr_2)$. I feel
-  like we should stick to one for consistency, but I don't know
-  which.}
-\davidsays{The reason for the distinction is that when we consider a
-  real liquid, then the true $g(r)$ is different from $g_{HS}(r)$, so
-  the mean-field equations ought to specify that they use the
-  hard-sphere pair distribution function.  But when we're talking
-  about the hard-sphere fluid, $g(r) = g_{HS}(r)$ and it's clumsy to
-  always have the HS present. }
+thermodynamic perturbation theory suggest that our approximation is
+accurate for both short-range and long-range attractions.
 

### 2014-03-25 papers/pair-correlation/figs/plot-path-triplet-inbetween.py

@@ -278,17 +278,18 @@ for name in plots:
   suba = axes([.77, .59, .195, .28])
   suba.plot(z_c, g_c, styles.plot[name], label=styles.title[name])
   sub_ylim = (1.8, 2.5)
+  suba.set_xticks([2, 3, 4, 5, 6])
   suba.set_yticks([2, 2.5])
   suba.set_ylim(sub_ylim)
   sub_xlim = (2.5, z_c.max())
   suba.set_xlim(sub_xlim)
 
   for i in suba.spines.itervalues():
-    i.set_linewidth(2)
+    i.set_linewidth(1)
   if name == 'this-work': # only want to draw rectangle once
     zplot.add_patch(Rectangle((sub_xlim[0], sub_ylim[0]),
                               sub_xlim[1]-sub_xlim[0], sub_ylim[1]-sub_ylim[0],
-                              facecolor='none', linewidth=2))
+                              facecolor='none', edgecolor='#cccccc', linewidth=1))
 
 
 def get_closest(x, y, xref, space):

### 2014-03-26 papers/square-well-liquid/paper.tex
automagically generate two stupid plots for square-well paper (to demonstrate how) Tested on darcy. Took 0:20.07.
some work on square well fluid paper (adding structure and a few citations) Tested on darcy. Took 0:41.21.

@@
 \usepackage{color}
+
+\newcommand\xx{\mathbf{x}}
+\newcommand\rr{\mathbf{x}}
+\newcommand\fixme[1]{\textcolor{red}{\textbf{[#1]}}}
 
@@
 \author{Student One}
-\author{Student Two}
+\author{Michael MI? Perlin}
 \author{David Roundy}
@@
 
+The SAFT Helmholtz free energy is composed of five terms:
+\begin{equation} \label{eq:SAFT-free-energy}
+  F = F_\textit{id} + F_\textit{hs} + F_\textit{disp} +
+  F_\textit{assoc} + F_\textit{chain},
+\end{equation}
+where the first three terms---ideal gas, hard-sphere repulsion and
+dispersion---encompass the \emph{monomer} contribution
+to the free energy, the fourth is the \emph{association} free energy,
+describing hydrogen bonds, and the final term is the chain formation
+energy for fluids that are chain polymers.  While a
+number of formulations of SAFT have been published, we will focus on
+SAFT-VR\cite{gil-villegas-1997-SAFT-VR}.
+
+\newcommand\etadisp{\ensuremath{\eta_\textit{d}}}
+\newcommand\epsilondisp{\ensuremath{\epsilon_\textit{d}}}
+\newcommand\lambdadisp{\ensuremath{\lambda_\textit{d}}}
+\newcommand\lscale{\ensuremath{s_d}}
+
+\section{Theory}
+
+\subsection{Thermodynamic perturbation theory}
+
+\fixme{Add here an introduction to thermodyamic perturbation theory,
+  explaining $a_1$ and $a_2$ and their ``exact'' formulas that we wish
+  to approximate.}
+
+\subsection{Square well contribution in homogeneous case}
+The square well contribution to the free energy is based on
+thermodynamic perturbation theory (sometimes known as the ``high
+temperature expansion'').  We use a dispersion term based on the
+SAFT-VR approach\cite{gil-villegas-1997-SAFT-VR}, which has two
+parameters an interaction energy $\epsilondisp$ and a length scale
+$\lambdadisp R$.
+
+The SAFT-VR dispersion free energy has the form~\cite{gil-villegas-1997-SAFT-VR}
+\begin{align}
+  F_\text{disp}[n] &= \int \left(a_1(\xx) + \beta a_2(\xx)\right)n(\xx)d\xx,
+\end{align}
+where $a_1$ and $a_2$ are the first two terms in a high-temperature
+perturbation expansion and $\beta=1/k_BT$.  The first term, $a_1$, is 
+the mean-field dispersion interaction. The second term, $a_2$, describes the
+effect of fluctuations resulting from compression of the fluid due
+to the dispersion interaction itself, and is approximated
+using the local compressibility approximation (LCA), which
+assumes the energy fluctuation is simply related to the
+compressibility of a hard-sphere reference fluid\cite{barker1976liquid}.
+
+\fixme{Add here a complete exposition of this functional form,
+  reproducing what Gil Villegas says.}
+
+\subsection{A simple weighted density approach}
+
+Hughes \emph{et al.} used a simple weighted density approach to
+approximate the square-well contribution to the free energy, which we
+will describe here~\cite{hughes2013classical}.
+
+The form of $a_1$ and $a_2$ for SAFT-VR is given in
+reference~\cite{gil-villegas-1997-SAFT-VR}, expressed in terms
+of the packing fraction.  In order to apply this form to an
+\emph{inhomogeneous} density distribution, we construct an effective local
+packing fraction for dispersion $\etadisp$, given by a Gaussian
+convolution of the density:
+\begin{align}
+%    ((4/3.0*M_PI)*Pow(3)(R)*GaussianConvolve(2*lambdainput*lscale*radius,
+%                                             2*lambdaE*length_scalingE*Rexpr)
+%  \etadisp(\xx) &= \color{red}
+%  \int \frac{\Theta(2\lambdadisp\lscale R - \left|\xx-\xx'\right|)}{
+%             \left(2\lambdadisp\lscale\right)^3} n(\xx') d\xx'.
+  \etadisp(\xx) &= \frac{1}{6\sqrt{\pi} \lambdadisp^3\lscale^3}
+  \int n(\xx')\exp\left({-\frac{|\xx-\xx'|^2}{2(2 \lambdadisp
+      \lscale R)^2}}\right)d\xx'.
+\end{align}
+This effective packing fraction is used throughout the dispersion
+functional, and represents a packing fraction averaged over the
+effective range of the dispersive interaction.  Here we have
+introduced an additional empirical parameter $\lscale$ which modifies
+the length scale over which the dispersion interaction is correlated.
+
+\subsection{Others}
+
+\fixme{Check out what other groups have done, especially Jackson
+  and/or Gross.  Also Arias.}
+
+\subsection{Weighted density approach}
+
+\cite{peng2008meanfield} used by
+\cite{sundararaman2013efficient} and many others.
+
+See also \cite{yu2009novel} which explicitly addresses the square well
+liquid.
+
+Also look up \cite{shen2013hybrid} which might follow Yu above.
+
+\subsection{Using the pair distribution function}
+
+\fixme{We have introduced an approximation to the pair distribution
+  function...}
+
+\section{Simulation}
+
+\fixme{A short explanation of MC methods here?}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}
+  \caption{A simple DOS plot!}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
+  \caption{A simple DOS plot!}
+\end{figure}
+
+\section{Results}
+
+\subsection{Homogeneous fluid}
+
+\fixme{Here we present the comparison between various theories and in
+  the homogeneous case.  This will mostly be our pair distribution
+  function vs. Gil Villegas.}
+
+\subsection{Hard wall}
+
+\fixme{Here we present the comparison between various theories and
+  simulation for a fluid at a hard wall.}
+
+\subsection{Test particle}
+
+\fixme{Here we present the comparison between various theories and
+  simulation for a test-particle simulation (i.e. the radial
+  distribution function).}
+
+\section{Conclusion}
+
+\fixme{We have an awesome theory.}
+
+\bibliography{paper}% Produces the bibliography via BibTeX.
+
 \end{document}

### 2014-03-26 papers/pair-correlation/paper.tex
make plot a little tighter Tested on bennet. Took 0:17.82.
Rewrite of triplet results. Also edited plots so arrows show on top. Tested on paho-arch. Took 0:54.82.

@@
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-contact-3.pdf}
+  \vspace{-0.7cm}
 \\
   \includegraphics[width=\textwidth]{figs/triplet-correlation-pretty-inbetween-3.pdf}
+  \vspace{-0.7cm}
   \caption{The triplet distribution function
@@
     values along a mirrored path are also shown, with arrows along the
-    curve indicating the path
-    direction. \fixme{Explain paths and arrows more.}} \label{fig:triplet-contact-distribution}
+    curve indicating the path direction, as explained in the
+    text.} \label{fig:triplet-contact-distribution}
 \end{figure*}
@@
 approaches gives a reasonable prediction, with the CVA-S
-underestimating the oscillations in $g_{HS}^{(2)}$, as expected based on
+underestimating the oscillations in $g^{(2)}$, as expected based on
 the fit in Fig.~\ref{fig:radial-distribution}.
@@
 
-Figure~\ref{fig:triplet-contact-distribution} compares the triplet
+Figures~\ref{fig:triplet-contact-distribution}a
+and~\ref{fig:triplet-contact-distribution}c compare the triplet
 distribution function at a packing fraction of 0.3 computed using the
-CVA-S with results from Monte Carlo simulation.  In each subfigure,
-$\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
-is varied.  The test-particle position in each case is $\rr_1$, which
-is on the left-hand side of the figure.  In the color plots on the
-left, the first and second spheres are displayed in gray.  On the
-bottom half of figures~\ref{fig:triplet-contact-distribution}a
-and~\ref{fig:triplet-contact-distribution}c is the triplet
-distribution function as computed using the CVA-S.  The top half of
-those figures displays the triplet distribution function from Monte
-Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
+CVA-S with results from Monte Carlo simulations. In
+Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
+and $\rr_2$ are in contact, in
+Figure~\ref{fig:triplet-contact-distribution}c they are spaced so that
+a third sphere can just fit between them, and in both figures $\rr_3$
+is varied. The test-particle position for the CVA-S in each case is
+$\rr_1$, which is on the left-hand side of the figure. We see
+reasonable agreement, with flatter oscillations in the CVA-S at larger
+distances. Also, the Monte Carlo results have perfect left-right
+symmetry, while the CVA-S has an asymmetry introduced with
+the test particle due to imperfections in the pair distribution
+function. Most notably, the CVA-S has less angular dependence at
+contact with the test particle than it does with the sphere at
+$\rr_2$.
+
+Figures~\ref{fig:triplet-contact-distribution}b
 and~\ref{fig:triplet-contact-distribution}d show the triplet
-distribution function along the paths illustrated in the corresponding
-2D plots.
+distribution function as plotted along the paths illustrated in
+Figures~\ref{fig:triplet-contact-distribution}a
+and~\ref{fig:triplet-contact-distribution}c. As the theoretical
+results are asymmetric, they are also plotted along similar paths that
+travel to the left, along the test particle. The different paths are
+indicated by arrows along the curves, with right-facing arrows
+indicating the paths shown in
+Figures~\ref{fig:triplet-contact-distribution}a and
+\ref{fig:triplet-contact-distribution}c, and left-facing arrows
+indicating paths traveling along the test particle.  As the work of
+Fischer and Methfessel is only defined in contact, it is only plotted
+in Figure~\ref{fig:triplet-contact-distribution}b and only along the
+portion of the path shown that is in contact with $\rr_2$, and arrows
+are omitted. Again, the CVA-S underestimates the oscillations,
+especially in contact with the test particle when the points $\rr_1$
+and $\rr_2$ are separated by a distance of $2.1\sigma$, as seen in
+Figure~\ref{fig:triplet-contact-distribution}d. Otherwise, all methods
+perform similarly, with the method of Fischer and Methfessel giving
+the best results when the spheres at $\rr_1$ and $\rr_2$ are in
+contact.
 
-Figures~\ref{fig:triplet-contact-distribution}a
-and~\ref{fig:triplet-contact-distribution}b show the triplet
-distribution function when two of the positions $\rr_1$ and $\rr_2$
-are in contact.  The triplet distribution function predicted by the
-CVA-S is slightly asymmetric, indicating a failure in thermodynamic
-consistency.  Once again, we observe the lack of oscillations at large
-distances.  Figures~\ref{fig:triplet-contact-distribution}c
-and~\ref{fig:triplet-contact-distribution}d show the triplet
-distribution function when two of the positions $\rr_1$ and $\rr_2$
-are spaced so that the third could just fit between them.  Again, the
-triplet distribution function predicted by the CVA-S is slightly
-asymmetric, and lacks the oscillations seen in the Monte Carlo
-simulations at large distances.  \fixme{We need to discuss the plots
-  and the directions and arrows.} In both cases, our agreement with
-simulation is comparable to that of previous approximations.
+% Figure~\ref{fig:triplet-contact-distribution} compares the triplet
+% distribution function at a packing fraction of 0.3 computed using the
+% CVA-S with results from Monte Carlo simulation.  In each subfigure,
+% $\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
+% is varied.  The test-particle position in each case is $\rr_1$, which
+% is on the left-hand side of the figure.  In the color plots on the
+% left, the first and second spheres are displayed in gray.  On the
+% bottom half of figures~\ref{fig:triplet-contact-distribution}a
+% and~\ref{fig:triplet-contact-distribution}c is the triplet
+% distribution function as computed using the CVA-S.  The top half of
+% those figures displays the triplet distribution function from Monte
+% Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
+% and~\ref{fig:triplet-contact-distribution}d show the triplet
+% distribution function along the paths illustrated in the corresponding
+% 2D plots.
+
+% Figures~\ref{fig:triplet-contact-distribution}a
+% and~\ref{fig:triplet-contact-distribution}b show the triplet
+% distribution function when two of the positions $\rr_1$ and $\rr_2$
+% are in contact.  The triplet distribution function predicted by the
+% CVA-S is slightly asymmetric, indicating a failure in thermodynamic
+% consistency.  Once again, we observe the lack of oscillations at large
+% distances.  Figures~\ref{fig:triplet-contact-distribution}c
+% and~\ref{fig:triplet-contact-distribution}d show the triplet
+% distribution function when two of the positions $\rr_1$ and $\rr_2$
+% are spaced so that the third could just fit between them.  Again, the
+% triplet distribution function predicted by the CVA-S is slightly
+% asymmetric, and lacks the oscillations seen in the Monte Carlo
+% simulations at large distances.  \fixme{We need to discuss the plots
+%   and the directions and arrows.} In both cases, our agreement with
+% simulation is comparable to that of previous approximations.
 

### 2014-03-26 papers/thesis-roth/project.tex
added fig comparing SW and SW+RG Tested on elliot. Took 0:20.38.

@@
 % 1/k_B/T
-\newcommand{\kt}{k_BT}
+\newcommand{\kT}{k_BT}
 
@@
 \begin{equation}
-  f_0(T,n) = f_{id}(T,n) + f_{HS}(T,n) + a_2^{SW}(T,n)\ ,
+  f_0(T,n) = f_{id}(T,n) + f_{HS}(T,n) + \frac{a_2^{SW}(T,n)}{\kT}\ ,
 \end{equation}
-where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kt$.~\cite{Gil-Villegas97}
+where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kT$.~\cite{Gil-Villegas97}
 
@@
 
-% \includegraphics[width=\columnwidth]{figs/combined-coexistence-plot}
+When only including $f^{id}$ and $f^{HS}$ terms, SW and SW+RG match up
+nicely. Including the $a_1^{SW}$ and $a_2$ terms doesn't work so
+well...
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
+  \end{center}
+  \caption{Liquid-vapor coexistance for a square well liquid.}
+  \label{fig:coexistance_SW}
+\end{figure}
+
+\begin{figure}
+  \begin{center}
+%  \includegraphics[width=\columnwidth]{figs/SW-RG-compare}
+  \end{center}
+  \caption{Comparison of free energy density for SW and SW+RG at low temperature}
+  \label{fig:SW-RG-compare-loT}
+\end{figure}
+
+%% \begin{figure}
+%%   \begin{center}
+%%   \includegraphics[width=\columnwidth]{figs/}
+%%   \end{center}  \caption{Comaprison of SW and RG near the critical point}
+%%   \label{fig:SW-RG-compare-hiT}
+%% \end{figure}
 

### 2014-03-26 papers/thesis-roth/figs/RG.py

@@ -46,9 +46,9 @@ def VD(i):
 # Free energy density, defined by an iterative process
 def fiterative(T,n,i):
     # eqn (55) Forte 2011:
-    fnaught = SW.fid(T,n) + SW.fhs(T,n) + SW.a2(n) # SW (Hughes) a2 is the same as Forte's f2
+    fnaught = SW.fid(T,n) + SW.fhs(T,n) + SW.a2(n)/k_B/T*n # SW (and Hughes) a2/kT is the same as Forte's f2
     # eqn (5) from Forte 2011:
-    for j in range(i+1): # The function range(x) only goes up to x-1; range(x+1) will include x
+    for j in range(i+1): # The function range(y) only goes up to y-1; range(y+1) will include y
         if j == 0:
             f = fnaught
         else:
@@ -115,7 +115,7 @@ def u(T,n,lambda_d,i):
 # Total
 # Eqn (32), Forte 2011; f1 = fatt (see paragraph after eqn 55)
 def ftot(T,n,i):
-    return fiterative(T,n,i) + SW.a1SW(n) # SW.a1SW is the same as Forte's f1
+    return fiterative(T,n,i) + SW.a1SW(n)*n # SW.a1SW is the same as Forte's f1
 
 # Grand free energy per volume
 def phi(T,n,nparticular,i):

### 2014-03-26 papers/pair-correlation/paper.tex
edit paper yay Tested on bennet. Took 0:19.85.

@@
     are 2D plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ as $\rr_3$
-    varies. \plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}. Note that an
-    asymmetry is introduced in the theoretical results, so their
-    values along a mirrored path are also shown, with arrows along the
-    curve indicating the path direction, as explained in the
-    text.} \label{fig:triplet-contact-distribution}
+    varies. %\plotcomp{$g^{(3)}(\rr_1,\rr_2,\rr_3)$}.
+%
+    The top halves of these figures show the results of Monte Carlo
+    simulations, while the bottom halves show the CVA-S.  On the right
+    are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
+    in the figures to the left.
+%
+    We also plot these curves along a left-right mirror image of this
+    path.  The data for the right-hand paths (as shown in the 2D
+    images) are marked with right-pointing triangles, while the
+    left-hand paths are marked with left-pointing triangles.
+%
+    %% The line styles are the same as in
+    %% Fig.~\ref{fig:pair-distribution}.
+  }
+  \label{fig:triplet-contact-distribution}
 \end{figure*}
@@
 positions---the location of the test particle---differently from the
-others, which means that a poor approximation to the pair distribution
-function may break the symmetry which is present in the true triplet
-distribution function.
+other two, which means that a poor approximation to the pair distribution
+function may break the symmetry between $\rr_1$ and $\rr_2$ which is
+present in the true triplet distribution function.
 
@@
 Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
-and $\rr_2$ are in contact, in
+and $\rr_2$ are in contact; in
 Figure~\ref{fig:triplet-contact-distribution}c they are spaced so that
-a third sphere can just fit between them, and in both figures $\rr_3$
+a third sphere can just fit between them; and in both figures $\rr_3$
 is varied. The test-particle position for the CVA-S in each case is
 $\rr_1$, which is on the left-hand side of the figure. We see
-reasonable agreement, with flatter oscillations in the CVA-S at larger
-distances. Also, the Monte Carlo results have perfect left-right
+reasonable agreement, with fewer oscillations in the CVA-S at larger
+distances. Also, the Monte Carlo results have the expected left-right
 symmetry, while the CVA-S has an asymmetry introduced with
-the test particle due to imperfections in the pair distribution
+the test particle due to errors in the pair distribution
 function. Most notably, the CVA-S has less angular dependence at
@@
 Figures~\ref{fig:triplet-contact-distribution}a
-and~\ref{fig:triplet-contact-distribution}c. As the theoretical
-results are asymmetric, they are also plotted along similar paths that
-travel to the left, along the test particle. The different paths are
-indicated by arrows along the curves, with right-facing arrows
+and~\ref{fig:triplet-contact-distribution}c.  We also plot the
+results along a left-right mirror image path, corresponding to
+swapping $\rr_1$ and $\rr_2$. The two mirror-image paths are
+distinguished by arrows (triangles) along the curves, with right-facing arrows
 indicating the paths shown in
@@
 \ref{fig:triplet-contact-distribution}c, and left-facing arrows
-indicating paths traveling along the test particle.  As the work of
-Fischer and Methfessel is only defined in contact, it is only plotted
-in Figure~\ref{fig:triplet-contact-distribution}b and only along the
-portion of the path shown that is in contact with $\rr_2$, and arrows
-are omitted. Again, the CVA-S underestimates the oscillations,
-especially in contact with the test particle when the points $\rr_1$
-and $\rr_2$ are separated by a distance of $2.1\sigma$, as seen in
-Figure~\ref{fig:triplet-contact-distribution}d. Otherwise, all methods
-perform similarly, with the method of Fischer and Methfessel giving
-the best results when the spheres at $\rr_1$ and $\rr_2$ are in
-contact.
+indicating the mirror image path.  As the work of
+Fischer and Methfessel is only defined when $\rr_2$ and $\rr_3$ are in
+contact, we only plot it along the
+central portion of the path, which is in contact with $\rr_2$, and arrows
+are omitted.
+%
+All methods perform similarly over this range with one exception.  The
+separable CVA-S method once again underestimates the oscillations on
+the left-hand side of the curve, where $\rr_2$ and $\rr_3$ are
+separated by a large distance.  On the path where $\rr_2$ and $\rr_3$
+are closer, the CVA-S method compares favorably with the other
+approximations.  This is consistent with
+Fig.~\ref{fig:radial-distribution}, which shows that $g_S(r)$ fits
+well the first peak of $g(r)$, but has fewer oscillations at large
+distances.
 

### 2014-03-28 papers/square-well-liquid/paper.tex
add description of g_S fit and a bit more text about pair distribution Tested on darcy. Took 0:23.05.

@@
 
+\subsection{The contact value approximation for the hard-sphere pair distribution function}
+
+We have recently introduced an efficient approximation for the pair
+distribution function of the inhomogeneous hard-sphere fluid.  This
+approximation takes the form:
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g_S(r_{12}, g_\sigma(\rr_1)) +
+    g_S(r_{12}, g_\sigma(\rr_2))}{2} \label{eq:g2-our-mean}
+\end{align}
+where $g_S(r,g_\sigma)$ is a separable fit to the radial distribution
+function of the homogeneus hard-sphere fluid, and $g_\sigma(\rr)$ is
+an approximation for the pair distribution function of an
+inhomogeneous hard-sphere fluid averaged over contact with a sphere
+located at position $\rr$.
+
+\subsubsection{Separable radial distribution function
+  $g_S(r,g_\sigma)$}
+
+The separable radial distribution function $g_S(r,g_\sigma)$ is given
+by
+\newcommand\kappaO{\kappa_0}
+\newcommand\kappaI{\kappa_1}
+\newcommand\kappaZ{\kappa_2}
+\newcommand\kappazero{3.68}
+\newcommand\kappaone{2.16}
+\newcommand\kappatwo{2.79}
+
+\begin{multline}
+  g_S(r,g_\sigma) = 1 + (g_\sigma-1) e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
+  + (g_\sigma -1)(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
+  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
+\end{multline}
+\begin{equation}
+  I(g_\sigma) = \kappaZ^5 \frac{
+    \frac{\chi-1}{24\eta} + (1-g_\sigma) \frac{(\kappaO-2
+      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
+    + (1-g_\sigma)\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
+  }{2 \kappaZ^2 + 12 \kappaZ + 24}
+\end{equation}
+\begin{equation}
+  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
+\end{equation}
+Finally, we can convert between the packing fraction $\eta$ and the
+radial distribution function at contact $g_\sigma$ using the
+Carnahan-Starling approximation for $g_\sigma$:
+\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
+    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
+\begin{align}
+  g_\sigma &= \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}\text{, and its inverse} \\
+  \eta &= 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}.
+\end{align}
+This function includes three fitted parameters, $\kappaO =
+\kappazero$, $\kappaI = \kappaone$,
+and $\kappaZ = \kappatwo$.
+
+\subsubsection{Averaged pair distribution  $g_\sigma(\rr)$}
+
 \subsection{Using the pair distribution function}

### 2014-03-28 papers/square-well-liquid/paper.tex
add a bit about the gsigma (section still needs to be added) Tested on darcy. Took 0:20.28.

@@
 
+The approximation for the averaged pair distribution $g_\sigma(\rr)$
+is from Schulte \emph{et al.}~\cite{schulte2012using}.  We will here
+summarize the key results needed to implement this approximation.
+
 \subsection{Using the pair distribution function}
@@
 \fixme{We have introduced an approximation to the pair distribution
-  function...}
+  function above... here we discuss an implementation of thermodynamic
+  perturbation theory for an inhomogeneous system using this pair
+  distribution function.}
 

### 2014-03-31 papers/square-well-liquid/paper.tex
added some thoughts to section IB of square well liquid paper Tested on elliot. Took 0:16.51.

@@
   reproducing what Gil Villegas says.}
+
+\fixme{Some notes are as follows:}
+In Gil-Villegas' paper, HS and Dispersion are wrapped up into the \textit{monomer} contribution, expressed in terms of energy densities as:
+\[
+  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+\]
+with $\alpha^\text{VDW}$ given by
+\[
+  \alpha^\text{VDW} = 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr
+\]
+Then, the high-temp expansion is an expansion of the monomer term:
+\[
+  f_\text{mono} = f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
+\]
+\fixme{I'm not sure how to express this in terms of $F_\text{disp}[n]$ alone.}
 
@@
 
-\fixme{We have an awesome theory.}
+\fixme{We have an awesome theory. You should agree.}
 

### 2014-04-01 papers/ice/paper.tex
prefer "equation" environment over \[ (and accidentally add embryonic ice paper) Tested on darcy. Took 0:22.46.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre]{revtex4-1}
\usepackage{graphicx}% Include figure files
%\usepackage{breqn}
\usepackage{color}

%% \newcommand{\red}[1]{{\bf \color{red} #1}}
%% \newcommand{\blue}[1]{{\bf \color{blue} #1}}
%% \newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\rr}{\textbf{r}}
\newcommand{\refnote}{\red{[ref]}}

\newcommand{\fixme}[1]{\red{[#1]}}

%\newcommand{\derivation}[1]{#1} % Use this to show all derivations in detail
\newcommand{\derivation}[1]{} % Use this for nice pegagogical paper...

% needsworklater is used to annotate bits that need work, but that we
% can postpone for a while.
\newcommand{\needsworklater}[1]{\emph{[#1]}}
% needsworknow is intended to prioritize stuff that needs fixing.
\newcommand{\needsworknow}[1]{\textcolor{red}{[\emph{#1}]}}

\begin{document}
\title{A Simple Theory for the Melting of Ice}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University}
%\pacs{61.20.Ne, 61.20.Gy, 61.20.Ja}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
  We develop a theory for ice.
\end{abstract}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

Wertheim's Thermodynamic Pertubation Theory (TPT) provides a formalism
for treating systems with short-range interactions combined with a
hard-core repulsion~\cite{wertheim1984fluidsI, *wertheim1984fluidsII,
  *wertheim1986fluidsIII}.  Various groups have
applied Wertheim's TPT to four-site systems such as water (see
\cite{peery2003association} for references).  Peery and Evans (and
others) applied Wertheim's methods to a four-site model for a
water-like fluid~\cite{peery2003association}.  Alas, all such models
studied to date fail to enforce tetrahedral coordination, instead
allowing the four association sites to float freely on the surface of
the hard sphere.

\subsection{Equations of state for water}

An interesting paper in which they use a single hydrogen bond to
``explain'' water, and claim that tetrahedral coordination is not so
important~\cite{truskett1999single}.

A molecular dynamics paper, in which they  model water using
Stillinger-Weber, and also talk about Si and
diamond\cite{molinero2008water, *hujo2011rise}.

A lattice model for water that makes impressive claims, but is a bit
hard to read at first glance\cite{hakem2007temperature}.  Here is a
second lattice model that also predicts thermodynamic anomalies, and
gives what looks like a good discussion of the low-temperature
properties of water\cite{pretti2004thermodynamic}.

A recent paper that introduces a new model for water to explain
low-temperature behavior of supercooled water\cite{tu2009anomalies}.
Has some nice references and explanation of supercooled behavior.

A key early paper in the ``chemical theory'' of associating liquids,
which is closely aligned to the theory I am
proposing\cite{heidemann1976van}.

An excellent review article, which outlines and explains the primary
approaches to modelling associating fluids, in particular giving
background on non-SAFT approaches such as the ``chemical
theory''\cite{muller2001molecular}.

\section{Theory}

\subsection{A look at dimers}

We will begin by outlining some aspects Wertheim's Thermodynamic
Pertubation Theory (TPT)~\cite{wertheim1984fluidsI,
  *wertheim1984fluidsII, *wertheim1986fluidsIII},
with a view towards extending it towards tetrahedrally coordinated
molecules such as water.

\newcommand\nI{\ensuremath{{n^{(1)}}}}
\newcommand\nII{\ensuremath{{n^{(2)}}}}

\newcommand\PI{\ensuremath{{P^{(1)}}}}
\newcommand\PII{\ensuremath{{P^{(2)}}}}

The idea of Wertheim's is to construct multiple densities for
different bonding states of molecules.  One then constructs
self-consistent equations which are solved for those densities, which
then appear in the free energy itself.  In Wertheim's approach, the
density is a ``bonding-state'' density.  We will use a related
approach in which we consider ``cluster'' densities.  Here, we will
describe Wertheim's approach with our ``cluster'' language, which will
be simplest to describe in the case where there is a single
bonding site on each molecule, so that at most dimers may be formed.
In this case, we can envision the fluid as composed of two densities,
a density of unbonded monomers $\nI$ and a density of dimers $\nII$.
Together they make up the total density:
\begin{align}
  n = \nI + 2\nII
\end{align}
We find the self-consistent equation for $\nII$ by considering the
relative probability of finding a dimer, versus finding two monomers
that are touching in a volume equal to the bonding volume, but in a
particular wrong orientation.  This ratio is given by a Boltzmann
factor
\begin{align}
  \frac{\PII}{\PI_{\textit{touching}}} &= e^{\beta\epsilon}
\end{align}
where $\epsilon$ is the energy difference between the two
configurations, which by construction have the same entropy.  The
probability of two spheres touching within a volume $\kappa$ is
determined by the monomer-monomer distribution function
\begin{align}
  \PI_{\textit{touching}} &= \PI \nI\kappa g_{11}(\sigma)
\end{align}
Thus, with a little algebra, we find
\begin{align}
  \PII &= \PI \nI\kappa g_{11}(\sigma) e^{\beta\epsilon} \\
  \nII &= \nI^2\kappa g_{11}(\sigma) e^{\beta\epsilon} \\
  n - \nI &= \nI^2\kappa g_{11}(\sigma) e^{\beta\epsilon} \\
  \nI\left(1 + \nI\kappa g_{11}(\sigma) e^{\beta\epsilon}\right) &= n \\
  \nI &= \frac{n}{\left(1 + \nI\kappa g_{11}(\sigma) e^{\beta\epsilon}\right)} \\
  \PI &= \frac{1}{\left(1 + \nI\kappa g_{11}(\sigma) e^{\beta\epsilon}\right)}
\end{align}
This quantity is normally called $X$ in the SAFT literature.  Once we
have these densities (and related probabilities), we are most of the
way towards finding the free energy and equation of state.

\subsection{Energies of clumps}

We will treat tetrahedral coordination by examining larger clusters
than just two, and taking into account the geometry of coordination.
There are different conformations, but we will begin by focusing on
just the lowest energy conformations, which are roughly spherical.
Later, we will consider the highest energy conformations, which are
linear chains.  The energy is equal to $\epsilon B_s$, where $B_s$ is
the number of bonds in a cluster of $s$ molecules.  We will
approximate $B_s$ by assuming that molecules within a specified
distance of the surface of the sphere have no bonds apart from those
to molecules on the interior, while the molecules on the interior have
four bonds each, keeping in mind that each bond is shared by two
molecules.
\begin{align}
  B_s &= 2(s^{\frac13} - 0.45)^3
\end{align}
The constant $0.45$ in this formula is reached by fitting the
thickness of the ``exterior'' in order to produce approximately
correct results for small values of $s$, as shown in
Table~\ref{tab:Bs}.
\begin{table}
  \begin{tabular}{c|cc}
    $s$ & $B_s$ & true $B_s$ \\
    \hline
    1 & 0.33 & 0 \\
    2 & 1.06 & 1 \\
    3 & 1.95 & 2 \\
    4 & 2.94 & 3 \\
    5 & 4.00 & 4 \\
    6 & 5.11 & 5 \\
    7 & 6.26 & 6 \\
    8 & 7.45 & 7 \\
  \end{tabular}
  \caption{Number of bonds $B_s$ as a function of cluster
    size.}\label{tab:Bs}
\end{table}

\section{Theory, take II}

I have another approach that sounds interesting, with a different way
of describing clusters.  The idea is to describe a cluster (as defined
aboce) in terms of connected \emph{rigid clusters}.  A rigid cluster
is a set of molecules any two of which are connected by more than one
bond.  A rigid cluster should be unable to be rotated at all, beyond
the amount of flex that comes about because each hydrogen bond has a
certain volume of give.  The smallest possible rigid cluster will be
composed of a single molecule, and I believe the next largest will
have six molecules in it.

The rigid clusters that make up a given cluster will be connected by a
single hydrogen bond.  Since this bond rotates freely, it will have a
certain amount of rotational entropy.  Moreover, there can be many
rigid clusters in a particular cluster, and there may be many ways of
putting together these rigid clusters (which could really do with a
better name).

This gives us a way to characterize clusters under several
parameters.  Hopefully only a smallish number of cluster formations
will be common, so we don't need an excessive number of terms in our
summations.

The main idea is that we have two problems to solve in order to
understand clusters: the rigid cluster problem (relatively easy to
approximate, I hope), and how to combine rigid clusters together to
get flexible clusters.  The rigid clusters will have very little
entropy, and quite a high energy (relatively speaking, not counting
the single-molecule rigid cluster), while connecting them together
will give a looser higher-entropy configuration.  Once we understand
the energy and entropy of each cluster configuration, we are most of
the way towards an equation of state for the tetrahedrally-coordinated
fluid.

The missing term will be the interaction between different clusters.
This interaction is purely entropic (since the only energy term shows
up within clusters, by their definition), and we will need a plausible
way to approximate it.  A start for this is to work out the volume of
each cluster.  Then we could use scaled-particle theory to handle
their interactions, assuming each cluster is a sphere, which seems
like a reasonable starting point.  It does require that we come up
with a reasonable way of defining the volume of a cluster.  We would
also need to define their area and mean curvature, if we wanted to go
beyond assuming these things are spheres.

\subsection{Free energy of a rigid cluster}

We can find the entropy of a given cluster configuration pretty easily
by computing its free energy.  The process is as follows.  The free
energy can be written as
\newcommand\dr{d^3r}
\begin{align}
  F &= -k_BT \ln Z \\
  Z &=
  \frac{1}{N!}
  \left(\frac{2m}{\beta\hbar^2}\right)^N
  \left(\frac{2I}{\beta\hbar^2}\right)^N
  \int \dr_1d\Omega_1\dr_2d\Omega_2\cdots e^{-\beta
    \Phi(\rr_1,\Omega_1,\cdots)}
\end{align}
where $\Omega_i$ is the orientation of molecule $i$, and $\Phi$ is the
potential energy.  If the molecules had internal degrees of freedom,
we would need to integrate over those as well.

\subsubsection{Rotational and motional free energy}

We consider a single cluster by setting $N$ to the number of molecules
in the cluster, and enforcing a constraint that they be attached in
the desired topology.  We can envision this as a potential that is
infinite when the cluster is broken.

The first integral $\dr_1$ is unconstrained, and corresponds to the
translational motion of the cluster.  For a rigid cluster, each of the
remaining integrals is constrained to a smallish volume.  This volume
is related to, but not identical to, the ``mutual volume of
interaction'' $\kappa$, which determines the probability at low
density of a molecule being bound on a particular site:
\begin{align}
  P &= \kappa n
\end{align}
The volume in the integrals above may be smaller than this, because
the location of a given molecule is constrained by that of all the
molecules it is bound to.

Anyhow, the final result is that we get something that looks
approximately like (leaving out the $N!$, which we'll include when we
combine together a bunch of clusters:
\begin{align}
  Z &\approx
  \left(\frac{2m}{\beta\hbar^2}\right)^{3N}
  \left(\frac{2I}{\beta\hbar^2}\right)^{3N}
  (V 8\pi^2) (\delta V \delta \Omega)^{N-1}
  \\
  &= \frac{V}{\Lambda^3}\frac{8\pi^2}{\Xi}
  \left(\frac{\delta V \delta \Omega}{\Lambda^3\Xi}\right)^{N-1}
\end{align}
where I have defined $\Xi \equiv
\left(\frac{2I}{\beta\hbar^2}\right)^3$ as an $SO(3)$ ``themal solid
solid angle''\footnote{I've been very sloppy here, and this $\Xi$
  needs to be carefully derived.}, which defines the angular
constriction needed to reduce the entropy of orientation to zero at a
given temperature.  This partition function does not include the
energy of interaction, so we can easily relate it to the entropy of
the cluster:
\begin{align}
  F_N &= -k_BT \ln Z_N \\
  &= -k_BT \ln\left(
     \frac{V}{\Lambda^3}\frac{8\pi^2}{\Xi}
     \left(\frac{\delta V \delta \Omega}{\Lambda^3\Xi}\right)^{N-1}
  \right)
  \\
  &= -k_BT \left(
    \ln\left(
     \frac{V}{\Lambda^3}\frac{8\pi^2}{\Xi}
    \right)
    +
    (N-1)
    \ln\left(
     \frac{\delta V \delta \Omega}{\Lambda^3\Xi}
    \right)
  \right)
\end{align}
where we can see that the first log is the ideal gas free energy
corresponding to center-of-mass motion and rotation of the cluster,
and the second is the free energy of the cluster itself.  We have made
a simplifying approximation with regard to the amount of volume
available for each molecule added to the cluster, but this seems like
a reasonably small approximation (which we can also test).

I again note here that I have omitted the fixed energy of bonding from
this free energy.

\subsubsection{Configurational free energy}
To the above free energy, we will need to add the free energy from
configurational entropy, corresponding to the number of ways we can
arrange the molecules with a given number of bonds (and thus bonding
energy) and a given number of molecules.

\subsection{Free energy of a complete connected cluster}

We will now consider a complete connected cluster, which is composed
of a number of connected rigid clusters, which could themselves be
composed of a single molecule.  There are bascially two sources of
entropy for such a cluster.  The first is the configurational entropy,
which describes the number of distinct ways we could put together such
a cluster.  The second is the motional and rotational entropy
corresponding to the wiggle room the rigid clusters have, and their
ability to rotate around the bonds connecting them.  I note that there
is an interaction between the two contributions, since the
configuration (as in, which cluster is bonded to which other cluster)
will affect how much wiggle room there is, since there could be some
steric hindrance (i.e. things could bump).  As a first approximation,
we will neglect interactions between the two contributions, and assume
that if there is no steric hindrance preventing a given configuration
from forming, then there will be no steric hindrance preventing each
bond from rotating freely.

\subsubsection{Rotational and motional entropy}

I will begin with the rotational and motional entropy, since it is
more closely related to the previously computed entropy (which is
purely motional) of the rigid cluster.  We will imagine building up a
complete cluster by connecting together rigid clusters one at a time.
Each time a cluster is connected, it loses most of its ``center of mass and
rotation'' degrees of freedom.  Its motion will be restricted to
essentially the same volume $\delta V$ that the molecules within a
cluster are restricted to.  The second cluster, however, may freely
rotate about the bond between the two clusters, which means that it
will have considerably greater rotational entropy than is present in
the case of a rigid cluster.  We just need to figure out the amount of
``solid solid angle'' that is allowed, which will replace the $\delta
\Omega$ in our formula.  The final result should look something like
\begin{widetext}
\begin{align}
  F &= -k_BT\left(
    \ln\left(
     \frac{V}{\Lambda^3}\frac{8\pi^2}{\Xi}
    \right)
    +
    (N_1-1)
    \ln\left(
     \frac{\delta V \delta \Omega}{\Lambda^3\Xi}
    \right)
    +
    \sum_{i=2}^{M} \left(
        \ln\left(\frac{\delta V
          \delta\Omega_{\textit{rot}}}{\Lambda^3\Xi}\right)
        +
        (N_i-1)
        \ln\left(
         \frac{\delta V \delta \Omega}{\Lambda^3\Xi}
        \right)
    \right)
  \right)
\end{align}
\end{widetext}
where it is immediately clear that there is considerable
simplification to be performed.  I have introduced a new variable
$\delta\Omega_{\textit{rot}}$, which is the amount of ``solid solid
angle'' that is available to a rigid cluster that is rotating freely
about a bond.  When we do the obvious simplification ($N = \sum N_i$),
we get something that looks like:
\begin{widetext}
\begin{align}
  F &= -k_BT\left(
    \ln\left(
     \frac{V}{\Lambda^3}\frac{8\pi^2}{\Xi}
    \right)
    +
    (M-1)
     \ln\left(\frac{\delta V
       \delta\Omega_{\textit{rot}}}{\Lambda^3\Xi}\right)
    +
    (N-M-1)
    \ln\left(
      \frac{\delta V \delta \Omega}{\Lambda^3\Xi}
    \right)
  \right)
\end{align}
\end{widetext}
which itself could be simplified a bit more, in that the motional and
rotational bits could be treated separately.  Anyhow, it is
interesting that when we ignore steric hindrances, the entropy of a
connected cluster is independent of how it is connected, and in fact
only depends on the total number of molecules $N$ and the total number
of rigid clusters $M$.

\subsubsection{Configurational entropy}

The configurational multiplicity will always have a factor of $N!$
that will come about because of the various permutations that could
describe the same configuration.  We won't discuss this $N!$ below,
but must include it in our final answer.

When looking at the configuration entropy, we cannot avoid considering
steric hindrances, since they prevent many potential configurations
from forming.  One effect of this steric hindrance is that the number
of molecules in a cluster $N$ places a constraint on the number of
rigid sub-clusters in that cluster $M$.For instance, if $N = 6$, we
can conclude that it must be made of either one rigid cluster (of 6)
or 6 rigid clusters (each composed of a single molecule).
Unfortunately, the constraint isn't very limiting, since (almost?) any
given cluster size (greater than $N=5$) could always be made of either
a single rigid cluster (e.g. a loop) or a maximum of $M=N$
single-molecule rigid clusters, arranged in an open-ended chain.

Let's consider the linear chain clusters.  In this case $M=N$, and
this is a very high entropy configuration, primarily because of its
maximum value of rotational entropy.  However, the linear chain has no
configurational entropy.  But there is also a whole family of branched
chain clusters, still composed of single-molecule rigid clusters, so
we can view the $M=N$ family as having a resultant configurational
entropy, which we will estimate here.  When we add a molecule to a
branched chain made of $N-1$ molecules, we have roughly $2N$ places
we could put it.  Thus
\begin{align}
  g_N &\approx 2N g_{N-1} \\
  \ln g_N &\approx \ln g_{N-1} + \ln 2 + \ln N \\
  \ln g_N &\approx N\ln 2 + N\ln N - N \\
  S_N &\approx k_B\left(N\ln N - (1-\ln 2)N \right)
\end{align}
I note that this is a very crude approximation.  It ignores steric
constraints, which are admittedly smaller for more flexible
configurations.  In the limit of large $N$, the assumption that the
number of binding sites is proportional to $N$ is wrong.  In this
limit there are guaranteed to be major steric effects, since the
random walk of even a linear chain is likely to overlap with itself.
Since the mean radius of a random walk in three dimensions scales as
$\sqrt{N}$, the volume occupied by most of the sites in the random
walk scales as $N^{3/2}$.

As it turns out, there is a large literature on self-avoiding random
walks.  According to one paper studying a 3D cubic
lattice\cite{grassberger1993monte},
\begin{align}
  g_N^{\text{cubic}} &= 4.68^N N^{0.16}
\end{align}
Actually, it looks like for a diamond lattice it should
be\cite{chen2002universal}
\begin{align}
  g_N^{\text{cubic}} &= 1.24 2.88^N N^{0.16}
\end{align}
which gives us a slightly reasonable starting point.  This is the
degeneracy of a linear chain with length $N$.  The total degeneracy
including branching configurations will of course be smaller.  The
radius of gyration is given by
\begin{align}
  G_N &= C N^{2\nu} \\
  &= 0.48 N^{0.59}
\end{align}
from which you can see that self-avoidance has some effect, but not a
dramatic effect.  I'm thinking that we might like to have a
differently defined self-avoiding random walk, in which the path can't
come adjascent to another point on the path, which would keep the
cluster ``flexible'' to a greater degree.

\section{Various other notes to self}

TODO: look at paper from segura et al 1987? for MD and interesting
results on just the 4-site associating hard spheres.

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2014-04-04 papers/square-well-liquid/paper.tex
prefer "equation" environment over \[ (and accidentally add embryonic ice paper) Tested on darcy. Took 0:22.46.
More work on SW theory section Tested on elliot. Took 0:19.28.

--- 2014-03-28 papers/square-well-liquid/paper.tex
@@
   reproducing what Gil Villegas says.}
+
+\fixme{DR -- Some notes are as follows:}
+
+In Gil-Villegas' paper, HS and Dispersion are wrapped up into the
+\textit{monomer} contribution, expressed in terms of energy densities
+as:
+\begin{align}
+  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+\end{align}
+with $\alpha^\text{VDW}$ given by
+\begin{align}
+  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
+  \intertext{using reduced units of $x = r/\sigma$, we have}
+  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
+  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
+\end{align}
+where $b^\text{VDW}$ is the van der Waals size parameter. It corresponds to the volume excluded by two spherical particles of volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
+Then, the high-temp expansion is an expansion of the monomer term:
+\begin{align}
+  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
+\end{align}
+\fixme{I'm not sure how to express this in terms of $F_\text{disp}[n]$ alone.}
+
+The term $a_1$ is given by
+\begin{align}
+  a_1 &= -2\pi\rho_s\epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3\rho_sb^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+\end{align}
+If assume a random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
+\begin{align}
+  a_1^\text{VDW} = -\rho_s\alpha^\text{VDW},
+\end{align}
+the van der Waals mean-field energy.
+
+The second-order term is quite a bit more complicated; one must have
+knowledge of higher-order correlation functions to determine
+$a_2$. This term describes the response of the attractive energy due
+to the compression of the fluid from the attractive well. Gil-Villegas
+\emph{et al.} base their expression on an approximation by Barker and
+Henderson\cite{barker1967-SW-perturbation}. \fixme{Gil-Villegas also
+  cites 2 other Barker and Henderson papers, but I can't find one and
+  the other doesn't seem relevant}. Barker and Henderson's
+approximation considers fluctuations in the number of particles in the
+potential well. In this way, the fluctuations in $a_2$ are related to the pressure and compressibility
+of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial\rho_s/\partial P^\text{HS}\right)_T$, we have:
+\begin{align}
+  a_2 &= -\pi\rho_s\epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial\rho_s g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  &= -\pi\rho_s\epsilon^2K^\text{HS}\frac{\partial}{\partial\rho_s}\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2\rho_s g^\text{HS}\,dr\right] \\
+  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial\rho_s}\left[-3\rho_sb^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+\end{align}
+
+We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion\cite{barker1976liquid}:
+\begin{align}
+  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
+\end{align}
+
+\fixme{End of my comments --DR}
 

### 2014-04-12 papers/square-well-liquid/histogram-paper.tex

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}\
-  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}\
+  \includegraphics[width=\columnwidth]{figs/periodic-ww300-ff30-N200-E.pdf}
   \caption{A histogram plot demonstrating the difficulty of using
+    canonical Monte Carlo to simulate a system.  A simulation at a
+    give temperature only provides information fo a small number of
+    energy states.\label{fig:histograms}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos.pdf}\
+  \includegraphics[width=\columnwidth]{figs/periodic-ww300-ff30-N200-dos.pdf}
+  \caption{A density of states plot demonstrating the difficulty of using
     canonical Monte Carlo to simulate a system.  A simulation at a

### 2014-04-12 src/Monte-Carlo/square-well.cpp
make flat histogram mostly work (involved fixing bug in weight-handling in move_one_ball) Tested on darcy. Took 3:47.08.

@@ -184,7 +184,7 @@ ball random_move(const ball &p, double move_scale, const double len[3]){
 void move_one_ball(int id, ball *p, int N, double len[3], int walls,
                    double neighbor_R, double translation_distance,
                    double interaction_distance, int max_neighbors, double dr,
-                   move_info *moves, double *ln_energy_weights){
+                   move_info *moves, int interactions, double *ln_energy_weights){
   moves->total++;
   moves->old_count =
     count_interactions(id, p, interaction_distance, len, walls);
@@ -222,7 +222,7 @@ void move_one_ball(int id, ball *p, int N, double len[3], int walls,
   p[id] = pid;
   // Now we can check if we actually want to do this move based on the
   // new energy.
-  const double lnPmove = ln_energy_weights[moves->new_count] - ln_energy_weights[moves->old_count];
+  const double lnPmove = ln_energy_weights[moves->new_count - moves->old_count + interactions] - ln_energy_weights[interactions];
   if (lnPmove < 0) {
     const double Pmove = exp(lnPmove);
     if (random::ran() > Pmove) {
@@ -252,6 +252,23 @@ int count_interactions(int id, ball *p, double interaction_distance,
                      len, walls).normsquared()
        <= uipow(interaction_distance,2))
       interactions++;
+  }
+  return interactions;
+}
+
+int count_all_interactions(ball *balls, int N, double interaction_distance, double len[3], int walls) {
+  // Count initial number of interactions
+  // Sum over i < k for all |ball[i].pos - ball[k].pos| < interaction_distance
+  int interactions = 0;
+  for(int i = 0; i < N; i++) {
+    for(int j = 0; j < balls[i].num_neighbors; j++) {
+      if(i < balls[i].neighbors[j]
+         && periodic_diff(balls[i].pos,
+                          balls[balls[i].neighbors[j]].pos,
+                          len, walls).norm()
+         <= interaction_distance)
+        interactions++;
+    }
   }
   return interactions;
 }

### 2014-04-13 papers/square-well-liquid/histogram-paper.tex
make flat histogram mostly work (involved fixing bug in weight-handling in move_one_ball) Tested on darcy. Took 3:47.08.
added RDF plotting script Tested on MAPHost. Took 1:24.70.
add figures to histogram-paper, tweak iterations and cut an unneeded printf Tested on darcy. Took 0:37.15.
fixed up some of the plotting scripts for SW Tested on MAPHost. Took 1:06.85.
added internal energy plot Tested on MAPHost. Took 1:09.95.

--- 2014-03-24 papers/square-well-liquid/histogram-paper.tex
@@
 
-\author{Michael MI? Perlin}
+\author{Michael A. Perlin}
 \author{David Roundy}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}\
-  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
   \caption{A histogram plot demonstrating the difficulty of using
@@
     energy states.\label{fig:histograms}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos.pdf}
+  \caption{A density of states plot demonstrating the difficulty of using
+    canonical Monte Carlo to simulate a system.  A simulation at a
+    give temperature only provides information fo a small number of
+    energy states.\label{fig:histograms}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u.pdf}
+  \caption{Specific internal energy.\label{fig:u}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc.pdf}
+  \caption{Specific heat capacity.\label{fig:hc}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rd.pdf}
+  \caption{Radial distribution function.\label{fig:rdf}}
 \end{figure}

### 2014-04-13 papers/square-well-liquid/figs/plot-rd.py

@@ -33,7 +33,8 @@ def g(gs,kte,counts,DS):
         output[i] += sum(gs[:,i]*DS*numpy.exp(-counts/kte)) / Z
     return output
 
-plt.title('Radial distribution function for $ww=%g$, $ff=%g$, and $N=%i$' % (ww, ff, N))
+plt.title('Radial distribution function for $\lambda=%g$, $\eta=%g$, and $N=%i$'
+          % (ww, ff, N))
 
 # FIXME: would it be better to use data from simulations with fixed kT?
 # input: "data/periodic-ww%04.2f-ff%04.2f-N%i-nw-g.dat" % (ww, ff, N)

### 2014-04-14 papers/square-well-liquid/paper.tex
make flat histogram mostly work (involved fixing bug in weight-handling in move_one_ball) Tested on darcy. Took 3:47.08.
added RDF plotting script Tested on MAPHost. Took 1:24.70.
minor changes to SW paper Tested on elliot. Took 0:19.75.

@@
 
-\fixme{DR -- Some notes are as follows:}
+\fixme{DRoth -- Some notes are as follows:}
 
@@
 where $b^\text{VDW}$ is the van der Waals size parameter. It corresponds to the volume excluded by two spherical particles of volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
+
 Then, the high-temp expansion is an expansion of the monomer term:
@@
 \begin{align}
-  a_1 &= -2\pi\rho_s\epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
-  &= -3\rho_sb^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+  a_1 &= -2\pi\rho\epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
 \end{align}
@@
 \begin{align}
-  a_1^\text{VDW} = -\rho_s\alpha^\text{VDW},
+  a_1^\text{VDW} = -\rho\alpha^\text{VDW},
 \end{align}
@@
 potential well. In this way, the fluctuations in $a_2$ are related to the pressure and compressibility
-of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial\rho_s/\partial P^\text{HS}\right)_T$, we have:
+of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial\rho /\partial P^\text{HS}\right)_T$, we have:
 \begin{align}
-  a_2 &= -\pi\rho_s\epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial\rho_s g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
-  &= -\pi\rho_s\epsilon^2K^\text{HS}\frac{\partial}{\partial\rho_s}\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2\rho_s g^\text{HS}\,dr\right] \\
-  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial\rho_s}\left[-3\rho_sb^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+  a_2 &= -\pi\rho\epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial\rho g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  &= -\pi\rho\epsilon^2K^\text{HS}\frac{\partial}{\partial\rho}\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2\rho g^\text{HS}\,dr\right] \\
+  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial\rho}\left[-3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
 \end{align}
@@
 
-\fixme{End of my comments --DR}
+\fixme{End of my comments --DRoth}
 

### 2014-04-14 papers/thesis-roth/figs/plot_SW.py
need to commit before I can pull Tested on elliot. Took 6:31.78.

@@ -6,10 +6,10 @@ import SW
 import numpy as np
 import matplotlib.pyplot as plt
 
-# Plot f vs n
+# Plot vs n
 R = 1 # HS radius
 eta_conv = SW.sigma**3*np.pi/6
-nmax = 0.4/eta_conv
+nmax = 0.9/eta_conv
 n = nmax*np.exp(np.arange(-15, 0, 1e-3))
 
 def plotPhi(T,npart):
@@ -30,6 +30,6 @@ def plotF(T):
     plt.ylabel(r'$\frac{df}{dn}$ (SW units)')
 
 T = 0.1
-plotPhi(T, 0.06)#/eta_conv)
+plotPhi(T, 0.025)#/eta_conv)
 #plt.savefig('figs/SW-T%0.2f.pdf'%T)
 plt.show()

### 2014-04-30 papers/fuzzy-fmt/NWAPS-2014.tex
Added draft of NWAPS-2014 fuzzy talk Tested on quipu. Took 0:20.65.

\documentclass{beamer}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{color}
\usepackage{cancel}
\graphicspath{ {/home/krebse/deft-soft/papers/fuzzy-fmt/figs/} }
\usetheme{Warsaw}
\newcommand{\rr}{\textbf{r}}
\title{Softening the Hard-Sphere Fluid}
\author{Eric Krebs, Samuel Loomis, Patrick Kreitzberg, David Roundy}
\institute{Oregon State University}
\date{May 3, 2014}
\begin{document}

\begin{frame}
 \titlepage
\end{frame}

\begin{frame}{Motivation}
  \begin{itemize}
    \item Want a theory that can deal with an inhomogeneous liquid
    \item The hard-sphere fluid is well studied and understood
    \item Hard spheres are used as a reference for real liquids
      \begin{itemize}
        \item Hard spheres are not very physical
        \item Computationally difficult (deta and step functions)
      \end{itemize}
  \end{itemize}
  \begin{block}{Classical density functional theory (cDFT)}
    \begin{align}
    \Omega(T) = \underset{n(\rr)}{min} \left\{F[n(\rr),T] + \int
                \left(V_{ext}(\rr) - \mu \right)n(\rr)d\rr \right\}
    \end{align}
    \begin{itemize}
      \item $n(\rr)$ can be an inhomogeneous particle density
    \end{itemize}
  \end{block}
\end{frame}

\begin{frame}{Fundamental measure theory (FMT)}
  \begin{itemize}
    \item Standard cDFT for hard-sphere
      fluid
    \item Hard spheres represented by ``fundamental measures''
      \begin{itemize}
        \item Volume, surface area, mean curvature
      \end{itemize}
    \item Weighted densities calculated using weight functions
      \begin{itemize}
        \item $n_{\alpha}(\rr) = \int n(\rr')w_{\alpha}(\rr -
        \rr')d\rr'$
      \end{itemize}
      \begin{block}{Some weight functions for sphere of radius R}
        \begin{align}
          w_3(\rr) = \theta (R - |\rr|) \\
          w_2(\rr) = \delta (R - |\rr|) \\
          \notag
        \end{align}
      \end{block}
    \item Other weight functions are simply related:
      \begin{itemize}
        \item $w_1(\rr) = \frac{w_2(\rr)}{4\pi R}$, $w_0(\rr) =
          \frac{w_2(\rr)}{4\pi R^2}$,  etc.
      \end{itemize}
  \end{itemize}
\end{frame}

\begin{frame}{Soft fundamental measure theory}
  \begin{itemize}
    \item A modification of FMT to treat penetrating spheres
    \item Weight functions are modified to be determined by the Mayer
      Function
    \item Reduces to hard spheres as $T \to 0$
      \begin{align}
        \int_{-\infty}^{\infty}w_2(r')w_2(r - r') dr' = \frac{\partial f(r)}{\partial r}
      \end{align}
      \begin{block}{Mayer Function}
        \begin{center}
          $f(r) = e^{-\beta V(r)} - 1$ \\
        \end{center}
      \end{block}
    \item If $w_2$ is known, all weight functions, $w_{\alpha}$'s, can
      be derived
  \end{itemize}
\end{frame}

\begin{frame}{Choosing a $w_2(\rr)$}
 \begin{itemize}
   \item Want spheres to overlap slightly at ambient temperatures
   \item We use a harmonic potential for $r < 2R$
     \begin{center}
       \includegraphics[trim = 0 0 0 20, clip, scale = 0.40]{figs/harmonic}
     \end{center}
   \item However, we could not find a simple analytic solution for $w_2(\rr)$
 \end{itemize}
\end{frame}

\begin{frame}{Error function method}
  \begin{itemize}
    \item Necessary to discuss this?
  \end{itemize}
\end{frame}

\begin{frame}{The homogeneous fluid}
  \begin{itemize}
    \item We compare homogeneous equation of state against Monte-Carlo
      (MC) simulation
    \item Good agreement for moderate temperatures and lower
      packing fraction
  \end{itemize}
  \begin{center}
    \includegraphics[trim = 0 13 0 37, clip, scale = 0.40]{figs/p-vs-packing}
  \end{center}
\end{frame}

\begin{frame}{Radial distribution}
  \begin{itemize}
    \item Compare radial distribution of homogeneous fluid with MC
      simulation
      \item Good agreement, especially in our range of interest
  \end{itemize}
  \begin{center}
    \includegraphics[trim = 35 17 35 25, clip, scale = .32]{figs/radial-distribution-10}
    \includegraphics[trim = 35 17 -1000 25 , clip, scale = .32]{figs/radial-distribution-30}
  \end{center}
\end{frame}

\begin{frame}{Soft-sphere fluid near a hard wall}
  \begin{itemize}
    \item We compare filling fraction near a hard wall against MC simulation
    \item Again we find good agreement.
  \end{itemize}
  \begin{center}
    \includegraphics[trim = 0 13 0 25, clip, scale = .40]{figs/walls-30}
  \end{center}
\end{frame}

\end{document}

### 2014-04-30 papers/fuzzy-fmt/NWAPS-2014.tex
Updated soft spheres NWAPS-2014 talk and plotting Tested on igor-stravinsky. Took 0:40.81.

@@
     \item Want a theory that can deal with an inhomogeneous liquid
-    \item The hard-sphere fluid is well studied and understood
-    \item Hard spheres are used as a reference for real liquids
-      \begin{itemize}
-        \item Hard spheres are not very physical
-        \item Computationally difficult (deta and step functions)
-      \end{itemize}
-  \end{itemize}
       \begin{block}{Classical density functional theory (cDFT)}
@@
       \end{block}
+    \item Hard spheres are used as a reference for real liquids
+      \begin{itemize}
+        \item Thermodynamic Perturbation Theory (TPT)
+      \end{itemize}
+    \item The hard-sphere fluid is well studied and understood
+      \begin{itemize}
+        \item Hard spheres are not very physical
+        \item Computationally difficult (deta and step functions)
+      \end{itemize}
+  \end{itemize}
 \end{frame}
@@
     \item A modification of FMT to treat penetrating spheres
-    \item Weight functions are modified to be determined by the Mayer
-      Function
+    \item Modified Weight functions are determined by Mayer function
+      \begin{block}{Mayer function}
+        \begin{center}
+          $f(r) = e^{-\beta V(r)} - 1$ \\
+        \end{center}
+      \end{block}
     \item Reduces to hard spheres as $T \to 0$
@@
       \end{align}
-      \begin{block}{Mayer Function}
-        \begin{center}
-          $f(r) = e^{-\beta V(r)} - 1$ \\
-        \end{center}
-      \end{block}
     \item If $w_2$ is known, all weight functions, $w_{\alpha}$'s, can
@@
 
-\begin{frame}{Choosing a $w_2(\rr)$}
+\begin{frame}{Choosing a Potential}
  \begin{itemize}
@@
      \end{center}
-   \item However, we could not find a simple analytic solution for $w_2(\rr)$
- \end{itemize}
-\end{frame}
-
-\begin{frame}{Error function method}
-  \begin{itemize}
-    \item Necessary to discuss this?
+   \item However, we could not find an analytic solution for $w_2(\rr)$
+   \item Approximated solution for low density and temperature
  \end{itemize}
@@
 
-\begin{frame}{Radial distribution}
+\begin{frame}{Radial distribution: theory vs. MC}
   \begin{itemize}
-    \item Compare radial distribution of homogeneous fluid with MC
-      simulation
   \item Good agreement, especially in our range of interest
@@
   \begin{center}
-    \includegraphics[trim = 35 17 35 25, clip, scale = .32]{figs/radial-distribution-10}
-    \includegraphics[trim = 35 17 -1000 25 , clip, scale = .32]{figs/radial-distribution-30}
+    \includegraphics[trim = 35 14 35 25, clip, scale = .32]{figs/radial-distribution-30}
+    \includegraphics[trim = 35 14 35 25 , clip, scale = .32]{figs/radial-distribution-50}
   \end{center}
@@
 \begin{frame}{Soft-sphere fluid near a hard wall}
-  \begin{itemize}
-    \item We compare filling fraction near a hard wall against MC simulation
-    \item Again we find good agreement.
-  \end{itemize}
   \begin{center}
-    \includegraphics[trim = 0 13 0 25, clip, scale = .40]{figs/walls-30}
+    \includegraphics[trim = 0 13 0 25, clip, scale = .45]{figs/walls-30}
   \end{center}
@@
 
+\begin{frame}{Conclusion}
+  \begin{itemize}
+    \item Successfully modeled soft-spheres in area of interest
+    \item Need to apply to a real liquid functional: water
+  \end{itemize}
+\end{frame}
+
 \end{document}

### 2014-05-04 papers/square-well-liquid/histogram-paper.tex
renamed all rd to rdf Tested on MAPHost. Took 0:45.43.
rewrote some of plot-uhc, including adding an analytic computation of heat capacity. also, I had to comment out all figures in histogram-paper.tex due to an unknown error I get when committing, but not when building it with pdflatex manually or just by running scons... Tested on quipu. Took 0:21.81.
Tested on quipu. Took 0:20.95.

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
+%  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
   \caption{A histogram plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos.pdf}
+%  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos.pdf}
   \caption{A density of states plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u.pdf}
+%  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u.pdf}
   \caption{Specific internal energy.\label{fig:u}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc.pdf}
+%  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc.pdf}
   \caption{Specific heat capacity.\label{fig:hc}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rd.pdf}
+%  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rdf.pdf}
   \caption{Radial distribution function.\label{fig:rdf}}

### 2014-05-04 papers/square-well-liquid/histogram-paper.tex
truncate walker optimization weight histogram at max entropy point, so as to not waste time at high energy states. also fixed problem with histogram-paper not finding a figure Tested on quipu. Took 0:23.34.

--- 2014-04-13 papers/square-well-liquid/histogram-paper.tex
@@
 \usepackage{color}
+
+\graphicspath{ {figs/} }
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
+  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-E.pdf}
   \caption{A histogram plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos.pdf}
+  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-dos.pdf}
   \caption{A density of states plot demonstrating the difficulty of using
@@
     give temperature only provides information fo a small number of
-    energy states.\label{fig:histograms}}
+    energy states.\label{fig:dos}}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u.pdf}
+  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-u.pdf}
   \caption{Specific internal energy.\label{fig:u}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc.pdf}
+  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-hc.pdf}
   \caption{Specific heat capacity.\label{fig:hc}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rd.pdf}
+  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-rdf.pdf}
   \caption{Radial distribution function.\label{fig:rdf}}

### 2014-05-05 papers/square-well-liquid/histogram-paper.tex
fix scons build issue in square-well Tested on crystal. Took 1:35.16.

@@
 \usepackage{color}
-
-\graphicspath{ {figs/} }
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-E.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E}
   \caption{A histogram plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-dos.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos}
   \caption{A density of states plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-u.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u}
   \caption{Specific internal energy.\label{fig:u}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-hc.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc}
   \caption{Specific heat capacity.\label{fig:hc}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{periodic-ww130-ff30-N200-rdf.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rdf}
   \caption{Radial distribution function.\label{fig:rdf}}

### 2014-05-08 papers/square-well-liquid/histogram-paper.tex
switch to using even fewer spheres in quick computation Tested on bennet. Took 0:17.92.

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-E}
   \caption{A histogram plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-dos}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-dos}
   \caption{A density of states plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-u}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-u}
   \caption{Specific internal energy.\label{fig:u}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-hc}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-hc}
   \caption{Specific heat capacity.\label{fig:hc}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-rdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-rdf}
   \caption{Radial distribution function.\label{fig:rdf}}

### 2014-05-08 papers/square-well-liquid/paper.tex
switch to using even fewer spheres in quick computation Tested on bennet. Took 0:17.92.
started TPT section for SW paper Tested on elliot. Took 1:14.07.

@@
 \author{Student One}
-\author{Michael MI? Perlin}
+\author{Michael A. Perlin}
 \author{David Roundy}
@@
 
+\fixme{And introduction to TPT---why do we use it?}
+
+\fixme{For now, the discussion follows entirely from the Simple Liquids book~\cite{simple-liquids}}
+
+The pair potential $V(\rr)$ describes the interaction between two hard spheres. For the square-well model,
+\begin{align}
+  v_\text{SW}(\rr) &=
+    \begin{cases}
+      \infty & r < \sigma \\
+      -\epsilon & \sigma < r < \lambda\sigma \\
+      0 & \lambda\sigma < r
+    \end{cases}
+\end{align}
+
+Now, consider this to be the reference, and add a small perturbation:
+\begin{align}
+  v_\lambda(\rr) &= v_\text{SW}(\rr) + \lambda w(\rr)
+\end{align}
+
+The total potential energy of our system is
+\begin{align}
+  V_N(\lambda) = \sum_\rr v_\lambda(\rr)
+\end{align}
+
 \subsection{Square well contribution in homogeneous case}
@@
 
-\fixme{Add here a complete exposition of this functional form,
+\fixme{The following hopes to be a complete exposition of this functional form,
   reproducing what Gil Villegas says.}
-
-\fixme{DRoth -- Some notes are as follows:}
 
@@
 \end{align}
-\fixme{I'm not sure how to express this in terms of $F_\text{disp}[n]$ alone.}
 
@@
 \begin{align}
-  a_1 &= -2\pi\rho\epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
-  &= -3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
 \end{align}
-If assume a random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
+If we assume random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
 \begin{align}
-  a_1^\text{VDW} = -\rho\alpha^\text{VDW},
+  a_1^\text{VDW} = - n \alpha^\text{VDW},
 \end{align}
@@
 \emph{et al.} base their expression on an approximation by Barker and
-Henderson\cite{barker1967-SW-perturbation}. \fixme{Gil-Villegas also
-  cites 2 other Barker and Henderson papers, but I can't find one and
-  the other doesn't seem relevant}. Barker and Henderson's
+Henderson\cite{barker1967-SW-perturbation}. Barker and Henderson's
 approximation considers fluctuations in the number of particles in the
-potential well. In this way, the fluctuations in $a_2$ are related to the pressure and compressibility
-of the liquid. Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial\rho /\partial P^\text{HS}\right)_T$, we have:
+potential well. In this way, the fluctuations in $a_2$ are related to
+the pressure and compressibility of the liquid. Given pressure
+$P^\text{HS}$ and isothermal compressibility $K^\text{HS} =
+kT\left(\partial n  /\partial P^\text{HS}\right)_T$, we have:
 \begin{align}
-  a_2 &= -\pi\rho\epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial\rho g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
-  &= -\pi\rho\epsilon^2K^\text{HS}\frac{\partial}{\partial\rho}\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2\rho g^\text{HS}\,dr\right] \\
-  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial\rho}\left[-3\rho b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
+  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
 \end{align}
@@
 \end{align}
-
-\fixme{End of my comments --DRoth}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww13-ff30-N200-E.pdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
   \caption{A simple DOS plot!}
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww30-ff30-N200-E.pdf}
-  \caption{A simple DOS plot!}
-\end{figure}
+%% \begin{figure}
+%%   \in cludegraphics[width=\columnwidth]{figs/periodic-ww300-ff30-N200-E.pdf}
+%%   \caption{A simple DOS plot!}
+%% \end{figure}
 

### 2014-05-08 papers/thesis-roth/project.tex
need to commit before I can pull Tested on elliot. Took 6:31.78.
minor changes. committing so I can get some work done at home Tested on quipu. Took 0:26.46.
updates to RG stuff. Not a whole lot of writing, but I fixed some code...I hope Tested on elliot. Took 0:19.56.

@@
 \begin{abstract}
-{}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory and Methods}
+\section{Theory and Methods}\label{sec:theory}
 
-\subsection{Renormalization Group Theory}
+\subsection{Renormalization Group Theory}\label{subsec:RGT}
 Renormalization group (RG) theory is an iterative procedure that was
@@
 
-\subsection{Coexistence Curve Algorithm}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}\label{sec:methods}
+
+\subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
 Coexistence curves are found by modifying the chemical potential $\mu$
-in the grand free energy per volume $\Phi/V$. Grand free energy
+in the grand free energy per volume $\phi(T,n)/V$. Grand free energy
 density is defined as
 \begin{align}
-  \frac{\Phi}{V} &= \phi \nonumber \\
+  \frac{\Phi(T,n)}{V} &= \phi(T,n) \nonumber \\
                  &= f(T,n) - n\mu \nonumber \\
-                 &= f(T,n) - n\frac{d}{dn}f(T,\npart)\ .
+                 &= f(T,n) - n\frac{df(T,n)}{dn}\bigg|_{\npart}\ .
 \end{align}
@@
 The grand free energy density generally has two distinct minima when
-plotted versus density. The liquid will be in liquid-vapor equilibrium
-when those minima have the same value for $\phi$. There is a local
+plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an example.) The liquid will be in liquid-vapor equilibrium
+when those minima have the same value for $\phi(T,n)$. There is a local
 maximum between those minima, and the value of the density at that
 maximum is used for $\npart$. Find this value at a given temperature
-then plot $\phi$ vs $n$ at slightly higher temperature. Find the new
+then plot $\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new
 value for $\npart$ such that the two minima have the same value for
-$\phi$. (It is helpful to note that if $\npart$ is increased,
-$\phi(\nliq)$ increases, and $\phi(\nliq)$ decreases if $\npart$ is
-decresed.) The program I wrote (npart.py) uses the previous
+$\phi(T,n)$. (It is helpful to note that if $\npart$ is increased,
+$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$ is
+decresed.) The program I wrote uses the previous
 temperature's $\npart$ as a ``first guess,'' maximizing
@@
 until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
-uses this new value of $\npart$ as the ``first guess'' at the next
+uses this new value of $\npart$ as the ``first guess'' at $\npart$ for the next
 higher temperature.
+
+The algorithm is as follows; start at some temperature $T=T_{start}$, with an intial guess for $\npart$.
+\begin{enumerate}
+  \item Calculate $\nvap$ and $\nliq$ by minimizing $\phi(T_{start},n)$ over an appropriate region
+  \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
+  \item Calculate \[\delta\mu = \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\nliq - \nvap}\]
+  \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + \delta\mu$
+  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
+  \item Calculate $\phi(T_{start},\npart^*)$
+  \item Re-caluclate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
+  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some tolerance (I used $tol=10^{-5}$)
+  \item Increase the temperature to $T = T + \delta T$ and repeat
+\end{enumerate}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
+  \caption{Grand free energy per unit volume, SW, low temp}
+  \label{fig:SW-phi-lowT}
+\end{figure}
 
@@
 \section{Results and discussion}
-
-When only including $f^{id}$ and $f^{HS}$ terms, SW and SW+RG match up
-nicely. Including the $a_1^{SW}$ and $a_2$ terms doesn't work so
-well...
 
@@
   \begin{center}
-%  \includegraphics[width=\columnwidth]{figs/SW-RG-compare}
+  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-lowT}
   \end{center}
-  \caption{Comparison of free energy density for SW and SW+RG at low temperature}
-  \label{fig:SW-RG-compare-loT}
+  \caption{Free energy density for RG at low temperature}
+  \label{fig:SW-RG-compare-lowT}
 \end{figure}
 
-%% \begin{figure}
-%%   \begin{center}
-%%   \includegraphics[width=\columnwidth]{figs/}
-%%   \end{center}  \caption{Comaprison of SW and RG near the critical point}
-%%   \label{fig:SW-RG-compare-hiT}
-%% \end{figure}
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-highT}
+  \end{center}
+  \caption{Free energy density for RG near critical temp}
+  \label{fig:SW-RG-compare-highT}
+\end{figure}
 

### 2014-05-09 papers/thesis-roth/project.tex
more updates to project report; slight changes to SW coexistence plot Tested on elliot. Took 0:32.19.

@@
 \newcommand{\nvap}{n_{vapor}}
+
+% Free energy terms
+\newcommand{\fid}{f_\text{ID}(T,n)}
+\newcommand{\fhs}{f_\text{HS}(T,n)}
+\newcommand{\fdisp}{f_\text{disp}(T,n)}
+\newcommand{\fattr}{f_\text{attr}(T,n)}
+
+\newcommand{\fixme}[1]{\textcolor{red}{\textbf{[#1]}}}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Theory and Methods}\label{sec:theory}
+\section{Theory}\label{sec:theory}
+
+This work applies entirely to the homogeneous case---that is, there are no walls containing our liquid, nor are there any sort of solutes in the liquid.
+This being the case, the free energy is infinite.
+Therefore, I work with free energy density:
+\begin{align}
+  f(T,n) &\equiv \frac{F(T,n)}{V}
+\end{align}
 
@@
 developed to deal with large-scale fluctuations of the order parameter
-near the critical point of a system.
+near the critical point of a system. The formulation used here is provided by Forte~\textit{et al}\cite{Forte11}.
 
-The free energy density for each iteration is given by
-\begin{equation}
-  f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)\ .
-\end{equation}
-The total free energy density is obtained when one takes the
-iterations to infinity and adds in the attractive component:
-\begin{equation}
-  f(T,n) = \lim_{i \to \infty} f_i(T,n) + f_{att}(T,n)\ .
-\end{equation}
-The initial free energy density, $f_0$ depends on the liquid model of study. For this work, I use the square well model:
-\begin{equation}
-  f_0(T,n) = f_{id}(T,n) + f_{HS}(T,n) + \frac{a_2^{SW}(T,n)}{\kT}\ ,
-\end{equation}
-where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kT$.~\cite{Gil-Villegas97}
+\fixme{Discuss Wilson phase cell}
+
+We begin with a baseline free enrgy $f_0(T,n)$.
+We assume this baseline accounts for all fluctuations of wavelength $\lambda < \lambda_0$.
+We then double the size of our phase cell and add a correction to the free energy $\delta f_1(T,n)$.
+This correction accounts for fluctuations with a wavelength $\lambda = 2\lambda_0$.
+The new free energy is $f_1(T,n) = f_0(T,n) + \delta f_1(T,n)$.
+Then double the length scale again; $\delta f_2(T,n)$ accounts for fluctionations with $\lambda = 4\lambda_0$ and $f_2(T,n) = f_1(T,n) + \delta f_2(T,n)$.
+As we continue this process, the $i^{th}$ correction accounts for fluctuations with $\lambda = 2^i\lambda_0$, and $f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)$.
+In the end, your free energy is
+\begin{align}
+  f(T,n) = f_0(T,n) + \sum_{i=1}^\infty\delta f_i(T,n) + \fattr
+\end{align}
+
+Where $\fattr$ is the contribution due to the longest-wavelenght density fluctuations. At this stage, I do not entirely understand why Forte splits up $\fattr$.
+
+%% \fixme{The following comes from Forte~\cite{Forte11} and doesn't make a whole lot of sense to me}
+
+%% The free energy for each iteration is given by
+%% \begin{equation}
+%%   f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)\ .
+%% \end{equation}
+%% The total free is obtained when one takes the
+%% iterations to infinity and adds in the attractive component:
+%% \begin{equation}
+%%   f(T,n) = \lim_{i \to \infty} f_i(T,n) + \fattr\ .
+%% \end{equation}
+%% The initial free energy density, $f_0$ depends on the liquid model of study. For this work, I use the square well model:
+%% \begin{equation}
+%%   f_0(T,n) = \fid + \fhs + \frac{a_2^{SW}(T,n)}{\kT}\ ,
+%% \end{equation}
+%% where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kT$. (See ref~\cite{Gil-Villegas97}.)
+%% \fixme{End of Forte confusion}
+
+\subsection{Square well liquid free enrgy}\label{subsec:SW}
+For the baseline, this work is based on the square well liquid free energy\cite{Hughes13}:
+\begin{align}
+  f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
+\end{align}
+The long-range attractive term, $\fattr$ is given by
+\begin{align}
+  \fattr &= n a_1(n) \ .
+\end{align}
+$\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere repulsion, and $\fdisp$ is the attraction.
+The functions $a_j(n)$ are the first two terms from a high-temperature perturbation expansion.
+Each function is outlined below.
+
+\subsubsection{Ideal gas}\label{sub2sec:ID}
+\begin{align}
+  \fid &= n\kT\left(\log(n) - 1\right)
+\end{align}
+
+\subsubsection{Hard sphere repulsion}\label{sub2sec:HS}
+
+\subsubsection{Attraction}\label{sub2sec:disp}
+The attractive free energy includes van der Waals attraction between hard spheres and is founded in Thermodynamic Perturbation Theory (TPT).
+\begin{align}
+  \fdisp &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
+\end{align}
+The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}. I give their forms here. \fixme{Pull this from the Square well contribution in homogeneous case section in the SW paper}
+
+\subsection{Thermodynamic perturbation theory}\label{subsec:TPT}
+\fixme{Pull TPT section from SW paper}
 
@@
 \begin{enumerate}
-  \item Calculate $\nvap$ and $\nliq$ by minimizing $\phi(T_{start},n)$ over an appropriate region
+  \item Calculate $\nvap$ and $\nliq$ by calculating the minima in $\phi(T_{start},n)$ over appropriate regions
   \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
@@
   \item Re-caluclate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
-  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some tolerance (I used $tol=10^{-5}$)
+  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some computational tolerance (I used $tol=10^{-5}$)
   \item Increase the temperature to $T = T + \delta T$ and repeat

### 2014-05-09 papers/thesis-roth/figs/coexistance_combined.py

@@ -3,29 +3,46 @@
 import matplotlib, sys
 if 'show' not in sys.argv:
   matplotlib.use('Agg')
-import pylab
+import matplotlib.pyplot as plt
+import numpy as np
+import SW
+import RG
 
-# Read in data
-data_Hughes = pylab.loadtxt('figs/npart_Hughes-out.dat')
-data_SW = pylab.loadtxt('figs/npart_SW-out.dat')
+# ### SW ###
 
-T_Hughes = data_Hughes[:,0]
-nvapor_Hughes = data_Hughes[:,1]
-nliquid_Hughes = data_Hughes[:,2]
+# # Read in data
+# data = np.loadtxt('figs/npart_SW-out.dat')
 
-T_SW = data_SW[:,0]
-nvapor_SW = data_SW[:,1]
-nliquid_SW = data_SW[:,2]
+# T = data[:,0]
+# etavapor = data[:,1]*np.pi*SW.sigma**3/6
+# etaliquid = data[:,2]*np.pi*SW.sigma**3/6
 
-# Plot the curve
-pylab.plot(nvapor_Hughes/0.374, T_Hughes/698, 'b-')
-pylab.plot(nliquid_Hughes/0.374, T_Hughes/698, 'b-')
+# # Plot the curve
+# plt.plot(etavapor, T, 'bx',label='SW')
+# plt.plot(etaliquid, T, 'bx')
 
-pylab.plot(nvapor_SW/0.0737, T_SW/8.6, 'r-')
-pylab.plot(nliquid_SW/0.0737, T_SW/8.6, 'r-')
 
-pylab.xlabel(r'$n/n_{crit}$')
-pylab.ylabel(r'$T/T_{crit}$')
+### RG ###
 
-pylab.savefig('figs/combined-coexistence-plot.pdf')
-pylab.show()
+# define the colors/symbols
+colors = np.array(['b-','g-','r-'])
+
+for i in range(2):
+  # Read in data
+  data = np.loadtxt('figs/npart_RG-i%d-out.dat'%i)
+
+  T = data[:,0]
+  etavapor = data[:,1]*np.pi*RG.sigma**3/6
+  etaliquid = data[:,2]*np.pi*RG.sigma**3/6
+
+  # Plot the curve
+  plt.plot(etavapor, T, colors[i],label='RG '+r'$i=$ '+'%d'%i)
+  plt.plot(etaliquid, T, colors[i])
+
[…]

### 2014-05-14 papers/thesis-roth/project.tex
updates to RG and project report Tested on elliot. Took 0:22.17.
project report writing Tested on elliot. Took 0:18.35.

@@
 % 1/k_B/T
-\newcommand{\kT}{k_BT}
+\newcommand{\kT}{\ensuremath{k_BT}}
+
+% vector
+\newcommand{\rr}{\ensuremath{\bf{r}}}
 
 % special values for n
-\newcommand{\npart}{n_{particular}}
-\newcommand{\nliq}{n_{liquid}}
-\newcommand{\nvap}{n_{vapor}}
+\newcommand{\npart}{\ensuremath{n_{particular}}}
+\newcommand{\nliq}{\ensuremath{n_{liquid}}}
+\newcommand{\nvap}{\ensuremath{n_{vapor}}}
 
 % Free energy terms
-\newcommand{\fid}{f_\text{ID}(T,n)}
-\newcommand{\fhs}{f_\text{HS}(T,n)}
-\newcommand{\fdisp}{f_\text{disp}(T,n)}
-\newcommand{\fattr}{f_\text{attr}(T,n)}
+\newcommand{\fid}{\ensuremath{f_\text{ID}(T,n)}}
+\newcommand{\fhs}{\ensuremath{f_\text{HS}(T,n)}}
+\newcommand{\fdisp}{\ensuremath{f_\text{disp}(T,n)}}
+\newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}
 
+% pair potential
+\newcommand{\1}{\ensuremath{\textbf{r}_1}}
+\newcommand{\2}{\ensuremath{\textbf{r}_2}}
+
+% fixme
 \newcommand{\fixme}[1]{\textcolor{red}{\textbf{[#1]}}}
@@
 
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR
-97331}
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
+
+Near the critical point of a liquid, there are density \fixme{is saying $g(\rr)$ more accurate?} fluctuations at all length scales.
+In this regime, mean field theories break down, for obvious reasons.
+As one approaches the critical point, there becomes less and less of a difference in free energy of the liquid and vapor states; this causes difficulty in modeling liquid-vapor coexistence.
+This work develops a method for calculating liquid-vapor coexistence, and applies renormalization group theory to the square well liquid.
 
@@
 This work applies entirely to the homogeneous case---that is, there are no walls containing our liquid, nor are there any sort of solutes in the liquid.
-This being the case, the free energy is infinite.
+Since our fluid is of infinite extent, the free energy is infinite.
 Therefore, I work with free energy density:
@@
 \end{align}
-The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}. I give their forms here. \fixme{Pull this from the Square well contribution in homogeneous case section in the SW paper}
+The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}.
+
+The first term, $a_1$, is the mean-field dispersion interaction.
+The second term, $a_2$, describes the effect of fluctuations resulting from compression of the fluid due to the dispersion interaction itself, and is approximated using the local compressibility approximation (LCA), which assumes the energy fluctuation is simply related to the compressibility of a hard-sphere reference fluid\cite{Barker76}.
+
+In Gil-Villegas' paper, HS and Dispersion are wrapped up into the
+\textit{monomer} contribution, expressed in terms of energy densities
+as:
+\begin{align}
+  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+\end{align}
+with $\alpha^\text{VDW}$ given by
+\begin{align}
+  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
+  \intertext{using reduced units of $x = r/\sigma$, we have}
+  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
+  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
+\end{align}
+where $b^\text{VDW}$ is the van der Waals size parameter. It corresponds to the volume excluded by two spherical particles of volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
+
+Then, the high-temp expansion is an expansion of the monomer term:
+\begin{align}
+  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
+\end{align}
+
+The term $a_1$ is given by
+\begin{align}
+  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+\end{align}
+If we assume random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
+\begin{align}
+  a_1^\text{VDW} = - n \alpha^\text{VDW},
+\end{align}
+the van der Waals mean-field energy.
+
+The second-order term is quite a bit more complicated; one must have knowledge of higher-order correlation functions to determine $a_2$.
+This term describes the response of the attractive energy due to the compression of the fluid from the attractive well.
+Gil-Villegas \emph{et al.} base their expression on an approximation by Barker and Henderson\cite{Barker67}.
+Barker and Henderson's approximation considers fluctuations in the number of particles in the potential well.
+In this way, the fluctuations in $a_2$ are related to the pressure and compressibility of the liquid.
+Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial n  /\partial P^\text{HS}\right)_T$, we have:
+\begin{align}
+  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
+  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+\end{align}
+
+We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion\cite{Barker76}:
+\begin{align}
+  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
+\end{align}
 
 \subsection{Thermodynamic perturbation theory}\label{subsec:TPT}
-\fixme{Pull TPT section from SW paper}
+\fixme{Add here an introduction to thermodyamic perturbation theory,
+  explaining $a_1$ and $a_2$ and their ``exact'' formulas that we wish
+  to approximate.}
+
+\fixme{And introduction to TPT---why do we use it?}
+
+\fixme{For now, the discussion follows entirely from the Simple Liquids book~\cite{Hansen06}}
+
+The pair potential $v(\1,\2)$ describes the interaction between two hard spheres. For the square-well model,
+\begin{align}
+  v_\text{SW}(\1,\2) &=
+    \begin{cases}
+      \infty & r < \sigma \\
+      -\epsilon & \sigma < r < \lambda\sigma \\
+      0 & \lambda\sigma < r
+    \end{cases}
+\end{align}
+
+Now, consider this to be the reference, and add a small perturbation:
+\begin{align}
+  v_\lambda(\1,\2) &= v_\text{SW}(\1,\2) + \lambda w(\1,\2) \label{eqn:small-perturbation}
+\end{align}
+
+The total potential energy of our system is
+\begin{align}
+  V_N(\lambda) = \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j)
+\end{align}
+
+Given the \emph{configuration integral} (the potenial energy component of the partition function)
+\begin{align}
+  Z_N &= \int d\mathbf{r}^N\, e^{-\beta V_N}
+\end{align}
+and the excess free energy (i.e.~the non-ideal part; $F = F^\text{id} + F^\text{ex}$)
+\begin{align}
+  F^\text{ex} &= -\kT\ln\left( \frac{Z_N}{V^N} \right)
+\end{align}
+where $V^N$ is the total volume \fixme{I think?}, then the derivative of the excess free energy with respect to the coupling parameter $\lambda$ is
+\begin{align}
+  \frac{\partial F^\text{ex}}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
+  &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda} \\
+  \hookrightarrow \beta\frac{\partial F^\text{ex}}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
+  \hookrightarrow \beta F^\text{ex} &= \beta F^\text{ex}_0 + \beta\int_0^1 d\lambda\, \left\langle V_N^{'}(\lambda) \right\rangle_\lambda
+\end{align}
+where $\left\langle V_N^{'}(\lambda) \right\rangle_\lambda$ is the canonical ensemble average. We can take a series expansion of this average:
+\begin{align}\begin{split} % too long; use split environment, and align to a space (&{})
+  \left\langle V_N^{'}(\lambda) \right\rangle_\lambda = &{} \left\langle V_N^{'}(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda)^2
+\end{split}\end{align}
+The derivative term is
+\begin{align}
+  \frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda} &= \left\langle V_N^{''} \right\rangle_\lambda - \beta\sigma_\lambda^2
+\end{align}
+where
+\begin{align}
+  \sigma_\lambda^2 &\equiv \left\langle \left[ V_N^{'}(\lambda) \right]^2 \right\rangle_\lambda - \left\langle V_N^{'}(\lambda) \right\rangle^2_\lambda
+\end{align}
+Putting this together, we have
+\begin{align}
+  \begin{split}
+    \beta F^\text{ex} = &{} \beta F^\text{ex}_0 + \beta(1 - 0)\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
+  \end{split} \\
+  \begin{split}
+    =&{}  \beta F^\text{ex}_0 + \beta\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
+  \end{split}
+\end{align}
+Now, we can define the total perturbation energy as
+\begin{align}
+  W_N &= \sum_{i=1}^N\sum_{j>i}^N w_\lambda(\mathbf{r}_i,\mathbf{r}_j)
+\end{align}
+Then, following from equation~\ref{eqn:small-perturbation},
+\begin{align}
+  V_N^{'} &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \\
+  &= W_N \\
+  \therefore V_N^{''} &= 0
+\end{align}
+With this in mind, we can simplify equation~\ref{eqn:complicated-HTE} to the \emph{high-temperature expansion}, first derived by Zwanzig~\cite{Zwanzig54}:
+\begin{align}
+  \begin{split}
+    \beta F^\text{ex} = &{} \beta F^\text{ex}_0 + \beta\left\langle W_N \right\rangle_0  \\ &{} - \frac{1}{2}\beta^2\left( \left\langle W_N^2 \right\rangle_0 - \left\langle W_N \right\rangle_0^2 \right) + \mathcal{O}(\beta)^3
+  \end{split}
+\end{align}
 
@@
 It is the value $\npart$ that we adjust to find coexistence curves.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
+  \caption{Grand free energy per unit volume, SW, low temp}
+  \label{fig:SW-phi-lowT}
+\end{figure}
 
@@
 \end{enumerate}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
-  \caption{Grand free energy per unit volume, SW, low temp}
-  \label{fig:SW-phi-lowT}
-\end{figure}
 
@@
 \section{Conclusion}
+I have implemented a the use of renormalization group theory in calculations for the square well liquid. The program is not entirely de-bugged.
 

### 2014-05-15 papers/pair-correlation/paper.tex

@@
 distribution function, which extend the radial distribution function
-to inhomogeneous scenarios.  These works differ in what density is
-used when evaluating the radial distribution function $g(r,\eta)$.
+to inhomogeneous scenarios.  Work on these approximations date back to
+Toxvaerd \cite{toxvaerd1973statistical} and have been tested
+rigorously for both functional and non-functional based theories
+\cite{kalos1977structure,osborn1980monotonic,mccoy1981comparison}.
+The approximations differ in what density is used when evaluating the
+radial distribution function $g(r,\eta)$ or in how radial
+distributions evaluated at different densities are combined to yield
+an effective function \cite{toxvaerd1973statistical}.
 
-Several recent works~\cite{gloor2007prediction, gross2009density} have
-assumed a mean-density approximation for the pair distribution
+One common approximation \cite{gloor2007prediction, gross2009density}
+assumes a mean-density approximation for the pair distribution
 function:
@@
 Non-pathological approaches use an average of the density over some
-volume. Fischer and Methfessel~\cite{fischer1980born} introduce the
+volume. Fischer and
+Methfessel~\cite{fischer1980born,harris1987comment} introduce the
 approximation:
@@
 \end{align}
-where $n_3$ is one of the fundamental measures defined in Fundamental
-Measure Theory~\cite{rosenfeld1989free}, which is an integral of the
-density over a spherical volume:
+where $n_3$ is an integral of the density over a spherical volume that
+is later defined as one of the fundamental measures in Fundamental
+Measure Theory~\cite{rosenfeld1989free}:
 \begin{align}

### 2014-05-15 papers/pair-correlation/paper.tex
minor wording tweak in change to pair paper Tested on bennet. Took 0:19.35.

@@
 where $n_3$ is an integral of the density over a spherical volume that
-is later defined as one of the fundamental measures in Fundamental
+is now used as one of the fundamental measures in Fundamental
 Measure Theory~\cite{rosenfeld1989free}:

### 2014-05-16 papers/thesis-roth/project.tex

@@
 developed to deal with large-scale fluctuations of the order parameter
-near the critical point of a system. The formulation used here is provided by Forte~\textit{et al}\cite{Forte11}.
+near the critical point of a system. The formulation used here is provided by Forte~\textit{et al}.\cite{Forte11}
 
@@
 \subsection{Square well liquid free enrgy}\label{subsec:SW}
-For the baseline, this work is based on the square well liquid free energy\cite{Hughes13}:
+For the baseline, this work is based on the square well liquid free energy:\cite{Hughes13}
 \begin{align}
@@
 $\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere repulsion, and $\fdisp$ is the attraction.
-The functions $a_j(n)$ are the first two terms from a high-temperature perturbation expansion.
+The function $a_1(n)$ is the first term from a high-temperature perturbation expansion, described in further detail in sections~\ref{sub2sec:disp} and~\ref{subsec:TPT}.
 Each function is outlined below.
@@
 \subsubsection{Ideal gas}\label{sub2sec:ID}
+The ideal gas term is purely local \fixme{whatever that means}:
 \begin{align}
@@
 \subsubsection{Hard sphere repulsion}\label{sub2sec:HS}
+Hard sphere repulsive forces are dealt with by the White Bear version of Fundamental-Mearuse Theory.\cite{Roth02} The hard sphere excess free energy is given in the homogeneous case as
+\begin{align}
+  \fhs &= \kT\left( \Phi_1(n) + \Phi_2(n) + \Phi_3(n) \right)
+\end{align}
+The $\Phi_j(n)$ terms are given as
+\begin{align}
+  \Phi_1(n) &= -n_0\ln(1 - n_3) \\
+  \Phi_2(n) &= \frac{n_1n_2}{1 - n_3} \\
+  \Phi_3(n) &= n_2^3\left( \frac{n_3 + (1 - n_3)^2\ln(1 - n_3)}{36\pi n_3^2(1 - n_3)^2} \right)
+\end{align}
+The \emph{fundamental measure densities} are
+\begin{align}
+  n_0(n) &= \frac{n_2}{4\pi R^2} \nonumber \\
+  &= n \\
+  n_1(n) &= \frac{n_2}{4\pi R} \nonumber \\
+  &= nR \\
+  n_2(n) &= n4\pi R^2 \\
+  n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
+  &= \eta
+\end{align}
+where $\eta$ is the packing fraction.
 
@@
 The first term, $a_1$, is the mean-field dispersion interaction.
-The second term, $a_2$, describes the effect of fluctuations resulting from compression of the fluid due to the dispersion interaction itself, and is approximated using the local compressibility approximation (LCA), which assumes the energy fluctuation is simply related to the compressibility of a hard-sphere reference fluid\cite{Barker76}.
+The second term, $a_2$, describes the effect of fluctuations resulting from compression of the fluid due to the dispersion interaction itself, and is approximated using the local compressibility approximation (LCA), which assumes the energy fluctuation is simply related to the compressibility of a hard-sphere reference fluid.\cite{Barker76}
 
@@
 This term describes the response of the attractive energy due to the compression of the fluid from the attractive well.
-Gil-Villegas \emph{et al.} base their expression on an approximation by Barker and Henderson\cite{Barker67}.
-Barker and Henderson's approximation considers fluctuations in the number of particles in the potential well.
+Gil-Villegas \emph{et al.} base their expression on an approximation by Barker and Henderson.\cite{Barker67} Barker and Henderson's approximation considers fluctuations in the number of particles in the potential well.
 In this way, the fluctuations in $a_2$ are related to the pressure and compressibility of the liquid.
@@
 
-We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion\cite{Barker76}:
+We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion:\cite{Barker76}
 \begin{align}
@@
 
-\fixme{For now, the discussion follows entirely from the Simple Liquids book~\cite{Hansen06}}
+\fixme{For now, the discussion follows entirely from the Simple Liquids book\cite{Hansen06}}
 
@@
 \end{align}
-and the excess free energy (i.e.~the non-ideal part; $F = F^\text{id} + F^\text{ex}$)
+and the non-ideal part of the free energy (i.e.~the excess free energy) is
 \begin{align}
-  F^\text{ex} &= -\kT\ln\left( \frac{Z_N}{V^N} \right)
+  F &= -\kT\ln\left( \frac{Z_N}{V^N} \right)
 \end{align}
@@
 \begin{align}
-  \frac{\partial F^\text{ex}}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
+  \frac{\partial F}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
   &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda} \\
-  \hookrightarrow \beta\frac{\partial F^\text{ex}}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
-  \hookrightarrow \beta F^\text{ex} &= \beta F^\text{ex}_0 + \beta\int_0^1 d\lambda\, \left\langle V_N^{'}(\lambda) \right\rangle_\lambda
+  \hookrightarrow \beta\frac{\partial F}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
+  \hookrightarrow \beta F &= \beta F_0 + \beta\int_0^1 d\lambda\, \left\langle V_N^{'}(\lambda) \right\rangle_\lambda
 \end{align}
@@
 \begin{align}\begin{split} % too long; use split environment, and align to a space (&{})
-  \left\langle V_N^{'}(\lambda) \right\rangle_\lambda = &{} \left\langle V_N^{'}(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda)^2
+  \left\langle V_N^{'}(\lambda) \right\rangle_\lambda = &{} \left\langle V_N^{'}(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda^2)
 \end{split}\end{align}
@@
   \begin{split}
-    \beta F^\text{ex} = &{} \beta F^\text{ex}_0 + \beta(1 - 0)\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
+    \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
   \end{split} \\
   \begin{split}
-    =&{}  \beta F^\text{ex}_0 + \beta\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
+    =&{}  \beta F_0 + \beta\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
   \end{split}
@@
 \begin{align}
-  V_N^{'} &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \\
+  V_N^{'} &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \nonumber \\
   &= W_N \\
@@
 \end{align}
-With this in mind, we can simplify equation~\ref{eqn:complicated-HTE} to the \emph{high-temperature expansion}, first derived by Zwanzig~\cite{Zwanzig54}:
+With this in mind, we can simplify equation~\ref{eqn:complicated-HTE} to the \emph{high-temperature expansion}, first derived by Zwanzig:\cite{Zwanzig54}
 \begin{align}
   \begin{split}
-    \beta F^\text{ex} = &{} \beta F^\text{ex}_0 + \beta\left\langle W_N \right\rangle_0  \\ &{} - \frac{1}{2}\beta^2\left( \left\langle W_N^2 \right\rangle_0 - \left\langle W_N \right\rangle_0^2 \right) + \mathcal{O}(\beta)^3
+    \beta F = &{} \beta F_0 + \beta\left\langle W_N \right\rangle_0  \\ &{} - \frac{1}{2}\beta^2\left( \left\langle W_N^2 \right\rangle_0 - \left\langle W_N \right\rangle_0^2 \right) + \mathcal{O}(\beta^3)
   \end{split}
+\end{align}
+Zwanzig further showed that the $n^{th}$ term could be written in terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle W_N\right\rangle_0 \right]^\nu \right\rangle_0$ with $\nu \leq n$.
+
+Writing the statistical averages in terms of the pair density $n_\lambda^{(2)}(\1,\2)$ (i.e. the probability of having a sphere at \1 and \2 \fixme{make this more clear?}) gives us
+\begin{align}
+  \frac{\beta F}{N} &= \frac{\beta F_0}{N} + \frac{\beta}{2N}\int_0^1 d\lambda\, \iint d\1 d\2\, n_\lambda^{(2)}(\1,\2)w(\1,\2) \label{eqn:Henderson5-2-12}
+\end{align}
+The pair density itself can be expanded as
+\begin{align}
+  n_\lambda^{(2)}(\1,\2) &= n_0^{(2)}(\1,\2) + \lambda\frac{\partial n_\lambda^{(2)}(\1,\2)}{\partial\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2)
+\end{align}
+Applying this to the second term in equation~\ref{eqn:Henderson5-2-12}, we have
+\begin{align}
+  \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
+  &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)
 \end{align}

### 2014-05-16 papers/pair-correlation/paper.tex

@@
 
-One common approximation \cite{gloor2007prediction, gross2009density}
+One common approximation \cite{gloor2007prediction,
+  gross2009density,osborn1980monotonic,mccoy1981comparison,barrett2006some}
 assumes a mean-density approximation for the pair distribution
@@
 Non-pathological approaches use an average of the density over some
-volume. Fischer and
-Methfessel~\cite{fischer1980born,harris1987comment} introduce the
-approximation:
+volume. Fischer and Methfessel introduce the
+approximation~\cite{fischer1980born,harris1987comment,wadewitz2000application,winkelmann2001liquid}:
 \begin{align}

### 2014-05-16 papers/pair-correlation/paper.tex
added equation equation for previously used pair correlation function Tested on crystal. Took 0:38.82.

--- 2014-05-15 papers/pair-correlation/paper.tex
@@
 rigorously for both functional and non-functional based theories
-\cite{kalos1977structure,osborn1980monotonic,mccoy1981comparison}.
+\cite{kalos1977structure,osborn1980monotonic, mccoy1981comparison}.
 The approximations differ in what density is used when evaluating the
@@
 
-One common approximation \cite{gloor2007prediction, gross2009density}
-assumes a mean-density approximation for the pair distribution
-function:
+One common approximation \cite{gloor2007prediction, gross2009density,
+  osborn1980monotonic, mccoy1981comparison,barrett2006some} assumes a
+mean-density approximation for the pair distribution function:
 \begin{align}
@@
 radius and packing fraction $\eta$, and $\sigma$ is the hard-sphere
-diameter.  This approach,
-however, cannot be applied to highly inhomogeneous systems such as a
-dense fluid at a solid surface.  These systems exhibit large density
-oscillations that lead to mean packing fractions greater than unity,
-which cannot occur in the bulk reference system that defines $g(r,
-\eta)$.  The above papers restricted themselves to the liquid-vapor
-interface, which does not exhibit this pathology.
+diameter.
+
+Another approximation uses $g(r,\eta)$ and assumes the density
+evaulated at the position between the two points:
+\begin{align}
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \tfrac{\pi\sigma^3}{6}
+  n\left(\tfrac12(\rr_1+\rr_2)\right)\right)
+\end{align}
+
+The above two approaches cannot be applied to highly inhomogeneous
+systems such as a dense fluid at a solid surface.  These systems
+exhibit large density oscillations that lead to mean packing fractions
+greater than unity, which cannot occur in the bulk reference system
+that defines $g(r, \eta)$.  The above papers restricted themselves to
+the liquid-vapor interface, which does not exhibit this pathology.
 
 Non-pathological approaches use an average of the density over some
-volume. Fischer and
-Methfessel~\cite{fischer1980born,harris1987comment} introduce the
-approximation:
+volume. Fischer and Methfessel introduce the
+approximation~\cite{fischer1980born,harris1987comment}:
 \begin{align}
@@
 distribution function only at contact, when the distance between
-$\rr_1$ and $\rr_2$ is the hard-sphere diameter.
+$\rr_1$ and $\rr_2$ is the hard-sphere diameter.  \davidsays{Tang does something similar, but maybe different.\cite{tang1991density}}
 
@@
 is the density averaged over a sphere with diameter
-$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
-was arrived at by fitting to Monte Carlo simulation.  Although
-Eq.~\ref{eq:sokolowski} has the advantages of only involving density
-averages at the points at which the pair distribution function is
-desired, it cannot be written as a single-site convolution, since the
-convolution kernel depends on both points.
+$0.8\sigma$~\cite{sokolowski1992role, wadewitz2000application, winkelmann2001liquid}.  The
+value 0.8 in this formula was arrived at by fitting to Monte Carlo
+simulation.  Although Eq.~\ref{eq:sokolowski} has the advantages of
+only involving density averages at the points at which the pair
+distribution function is desired, it cannot be written as a
+single-site convolution, since the convolution kernel depends on both
+points.
 

### 2014-05-16 papers/pair-correlation/paper.tex
update with Jeff on lit background Tested on bennet. Took 0:17.74.

@@
 distribution function, which extend the radial distribution function
-to inhomogeneous scenarios.  Work on these approximations date back to
-Toxvaerd \cite{toxvaerd1973statistical} and have been tested
-rigorously for both functional and non-functional based theories
-\cite{kalos1977structure,osborn1980monotonic, mccoy1981comparison}.
-The approximations differ in what density is used when evaluating the
-radial distribution function $g(r,\eta)$ or in how radial
-distributions evaluated at different densities are combined to yield
-an effective function \cite{toxvaerd1973statistical}.
+to inhomogeneous scenarios.  The approximations differ in what density
+is used when evaluating the radial distribution function $g(r,\eta)$
+or in how radial distributions evaluated at different densities are
+combined to yield an effective function
+\cite{toxvaerd1973statistical}.
 
-One common approximation \cite{gloor2007prediction, gross2009density,
-  osborn1980monotonic, mccoy1981comparison,barrett2006some} assumes a
-mean-density approximation for the pair distribution function:
+Early approximations to the pair distribution function used the
+density at one or two positions to determine the pair distribution
+function.  There are three common approaches:
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \tfrac{\pi\sigma^3}{6}\tfrac12
-  (n(\rr_1)+n(\rr_2))\right)
+  g^{(2)}(\rr_1,\rr_2) &\approx
+  g\left(r_{12};n\left(\frac{\rr_1+\rr_2}{2}\right)\right)
+     & \text{midpoint} \\
+  g^{(2)}(\rr_1,\rr_2) &\approx
+  g\left(r_{12};\frac{n(\rr_1)+n(\rr_2)}{2}\right)
+     & \text{mean density} \\
+  g^{(2)}(\rr_1,\rr_2) &\approx
+  \frac{g(r_{12},n(\rr_1))+g(r_{12},n(\rr_2))}{2}
+     & \text{mean function}
 \end{align}
-where $g(r,\eta)$ is the radial distribution function as a function of
-radius and packing fraction $\eta$, and $\sigma$ is the hard-sphere
-diameter.
-
-Another approximation uses $g(r,\eta)$ and assumes the density
-evaulated at the position between the two points:
-\begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, \tfrac{\pi\sigma^3}{6}
-  n\left(\tfrac12(\rr_1+\rr_2)\right)\right)
-\end{align}
-
-The above two approaches cannot be applied to highly inhomogeneous
-systems such as a dense fluid at a solid surface.  These systems
-exhibit large density oscillations that lead to mean packing fractions
-greater than unity, which cannot occur in the bulk reference system
-that defines $g(r, \eta)$.  The above papers restricted themselves to
-the liquid-vapor interface, which does not exhibit this pathology.
+These approaches have been successfully and widely used in treating
+the surface tension of simple fluids~\cite{pressing1973, salter1975,
+  bongiorno1975, toxvaerd1976, kalos1977structure, carey1978,
+  osborn1980monotonic, mccoy1981comparison, barrett2006some}.  The
+mean density approximation has also been quoted (as a goal) by recent
+papers that proceed to make further
+approximations~\cite{gloor2007prediction, gross2009density}.  While
+these approximations fail dramatically when applied to strongly
+inhomogeneous systems such as a dense fluid at a solid surface.  Such
+systems exhibit a strongly oscillatory density distribution, with
+density peaks that can have local packing fractions greater than
+unity, which cannot occur in the bulk reference system that defines
+$g(r, n)$.  The above papers restrict themselves to the liquid-vapor
+interface, which does not exhibit this pathology.
 
@@
 distribution function only at contact, when the distance between
-$\rr_1$ and $\rr_2$ is the hard-sphere diameter.  \davidsays{Tang does something similar, but maybe different.\cite{tang1991density}}
+$\rr_1$ and $\rr_2$ is the hard-sphere diameter.
+%
+Tang \emph{et. al} employed an approximation for the pair distribution
+function that is similar to that of Fischer and Methfessel, but with a
+self-consistent weighted density computed with weighting function that
+is itself dependent on the weighted density~\cite{tang1991density}.
+This weighted density was computed using the hard-sphere weighted
+density of Tarazona, which was developed using the direct correlation
+function of the homogeneous hard-sphere fluid~\cite{tarazona1985}.
 

### 2014-05-16 papers/pair-correlation/paper.tex
added correct bib for references added earlier Tested on crystal. Took 0:32.27.

@@
 These approaches have been successfully and widely used in treating
-the surface tension of simple fluids~\cite{pressing1973, salter1975,
-  bongiorno1975, toxvaerd1976, kalos1977structure, carey1978,
+the surface tension of simple fluids~\cite{pressing2003surface,
+  salter2008statistical, bongiorno1975modified,
+  toxvaerd1976hydrostatic, kalos1977structure, carey2008gradient,
   osborn1980monotonic, mccoy1981comparison, barrett2006some}.  The
@@
 density of Tarazona, which was developed using the direct correlation
-function of the homogeneous hard-sphere fluid~\cite{tarazona1985}.
+function of the homogeneous hard-sphere fluid~\cite{tarazona1985free}.
 

### 2014-05-19 papers/renormalization/paper.tex
add blurb on RGT to (still mostly empty) RG paper Tested on bennet. Took 0:21.17.

@@
 
-\fixme{Here we will explain the RGT that we use, and how it works
-  out.}
+\subsubsection{Defining our approximation}
+Let's begin by considering the partition function and free energy, as
+computed in the canonical ensemble:
+\begin{align}
+  Z &= \sum_i^{\scriptscriptstyle\begin{array}{c}\text{\tiny all}\\\text{\tiny microstates}\end{array}}
+       e^{-\beta E_i}, & F &= -kT\log Z
+\end{align}
+We now consider a ``smooth'' configuration that contains only
+microstates that are periodic with a given periodicity $2^j\lambda$:
+\begin{align}
+  Z_j &= \sum_i^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny periodic }2^j\lambda\\
+                 \end{array}}
+       e^{-\beta E_i}, & F_j &= -kT\log Z_j
+       \label{eq:Zn}
+\end{align}
+This defines a family of free energies that contain fluctuations with
+a variety of length scales.  For the moment we will assume a simple
+cubic periodicity, although it may be advantageous to select face
+centered cubic (FCC) instead.  So far, we have defined a family of
+free energy functions (each containing longer length-scale
+fluctuations) that we can compute numerically using Monte Carlo, but
+we have not yet constructed a means to approximate these functions.
+Before developing this approximation, let's reorganize this sum over
+periodic states.
+\begin{align}
+  Z_j(N) &= \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+\end{align}
+where here $g_{N_1\cdots N_8}$ is a multiplicity factor
+corresponding to the number of ways the cell can be divided into
+subcells with occupancies of $N_1\cdots N_8$.  This is so far only a
+different way of writing Eq.~\ref{eq:Zn}.  At this point, we can
+multiply and divide by a few factors, with a goal of replacing the
+unknown $E_{i_1\cdots i_8}^{(j)}$ with something that is easier to
+approximate.
+\begin{multline}
+  F_j(n) = -kT\log\Bigg(
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \\
+       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+  \Bigg)
+\end{multline}
+Now we consider the same free energy computed without accounting for
+the energy in the fluctuations:
+\begin{multline}
+  F_{j-1}(n) = -kT\log\Bigg(
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \\
+       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta \left(E_{i_1}^{(j-1)} + \cdots + E_{i_8}^{(j-1)}\right)}
+  \Bigg)
+\end{multline}
+We can now recognize that each of the octants is just a similar smooth
+system.
+\begin{multline}
+  F_{j-1}(n) = -kT\log\Bigg(
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \\
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8) \right)}
+  \Bigg)
+\end{multline}
+Here I assume that the fluctuations will only be important in so far
+as the affect the free energy directly.  Putting these last two
+equations together, we can find a formula for the energy difference as
+we include the larger fluctuations:
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sum_{AAN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+  }{
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+        e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8) \right)}
+  }
+  \right)
+\end{multline}
+At this point, we recognize that we can put anything we like in the
+numerator \emph{and} the denominator, which will help us to prevent
+the exponentials from getting too hairy.  This is really just a shift
+of energy for both the numerator and denominator partition functions.
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sum_{NN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta \left(E_{i_1\cdots i_8}^{(j)}  - F_{j-1}(n) \right)}
+  }{
+    \sum_{BN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
+  }
+  \right)
+\end{multline}
+At this stage we have still not introduced an uncontrolled
+approximation.  Each $F_{j}$ is exactly defined, and the exact free
+energy is found by taking a limit $j\rightarrow \infty$.  \fixme{I
+  need to clarify the scale of each $F$.  Some of the above are about
+  systems $\frac18$ the volume of others.}
+
+The flavor of the approximation used by White and others in GRG is now
+clarified.  We begin by assuming that we already have an adequate (and
+tractable) approximation for $F_{j-1}$.  Thus the denominator
+partition function is readily computable.  We further simplify the
+numerator by multiplying and dividing by the same quantity, and
+simplify the notation by defining
+\begin{align}
+  F_j(N_1\cdots N_8) &= \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny microstates}\\
+                  \text{\tiny within octants}\\
+                 \end{array}}
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+\end{align}
+and thus rewrite the energy difference as
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sum_{CN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       e^{-\beta \big(F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) +
+           \big(F_j(N_1\cdots N_8) - F_{j-1}(N_1) - \cdots - F_{j-1}(N_8)\big) \big)}
+  }{
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
+  }
+  \right)
+\end{multline}
+At this point we can recognize and identify two energy differences as
+\begin{align}
+  \bar{F}(N_1\cdots N_8) &= F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \\
+  \bar{U}(N_1\cdots N_8) &= F_j(N_1\cdots N_8)
+  - \left(F_{j-1}(N_1) + \cdots + F_{j-1}(N_8) \right)
+\end{align}
+and using these definitions can rewrite our energy difference in a
+more compact form:
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sum_{CN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       e^{-\beta \big( \bar{F}(N_1\cdots N_8) + \bar{U}(N_1\cdots N_8)
+         \big)}
+  }{
+    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+                  \text{\tiny occupancies}\\
+                  \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}
+       e^{-\beta \bar{F}(N_1\cdots N_8)}
+  }
+  \right)
+\end{multline}
+At this point, we can recognize $\bar{F}(N_1\cdots N_8)$ as the free
+energy change due to oscillations \emph{if we neglect the actual
+  energy of the oscillations themselves, but instead make a ``local''
+  approximation}.  Meanwhile $\bar{U}(N_1\cdots N_8)$ is the free
+energy change due to the energy of oscillations.
+
+\fixme{I have now ``defined'' the terms used by Forte in his
+  notation of GRG.  On the to-do list: }
+\begin{enumerate}
+\item \fixme{Establish what $g_{N_1\cdots N_8}$ is.}
+\item \fixme{Elaborate on the approximation (ignoring corners) that
+  leads from a $\sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
+    \text{\tiny occupancies}\\ \text{\tiny of octants}\\
+                 \end{array}}
+       g_{N_1\cdots N_8}$ to an integral over density
+       amplitudes.}
+\item \fixme{Develop an approach for approximating $\bar{U}$.  Perhaps
+  is $F_{j-1}$ is expressed as a functional, we can use a non-local
+  DFT to approximate $\bar U_j$?  The standard GRG approach seems to
+  be to use a simple mean-field or thermodynamic perturbation theory
+  approach to approximate $\bar U$, suggesting that a simple
+  (standard) DFT might work just fine for this.}
+\end{enumerate}
 

### 2014-05-20 papers/thesis-roth/project.tex
updates to project report Tested on elliot. Took 0:21.08.

@@
 \subsection{Thermodynamic perturbation theory}\label{subsec:TPT}
-\fixme{Add here an introduction to thermodyamic perturbation theory,
-  explaining $a_1$ and $a_2$ and their ``exact'' formulas that we wish
-  to approximate.}
 
 \fixme{And introduction to TPT---why do we use it?}
+This discussion follows primarily from \textit{Theory of Simple Liquids}, by Hansen and McDonald\cite{Hansen06}.
 
-\fixme{For now, the discussion follows entirely from the Simple Liquids book\cite{Hansen06}}
+Thermodynamic perturbation theory treats the intermolecular potential as separable into two parts, one for short-range repulsion and another for long-range attraction.
 
-The pair potential $v(\1,\2)$ describes the interaction between two hard spheres. For the square-well model,
+The pair potential $v(\1,\2)$ describes the interaction between two hard spheres. For the square-well model, the pair potential is
 \begin{align}
@@
 \section{Conclusion}
-I have implemented a the use of renormalization group theory in calculations for the square well liquid. The program is not entirely de-bugged.
+I have implemented a the use of renormalization group theory in calculations for the square well liquid.
+The program is not entirely de-bugged.
+I have also developed an algorithm that calculates liqiud-vapor coexistence curves, even close to the critical point.
 

### 2014-05-20 papers/renormalization/paper.tex
Small but important change to RG paper Tested on elliot. Took 0:19.43.
clean up summation notation and write about multiplicity Tested on bennet. Took 0:16.41.

@@
 \begin{align}
-  Z &= \sum_i^{\scriptscriptstyle\begin{array}{c}\text{\tiny all}\\\text{\tiny microstates}\end{array}}
+  Z &= \sum_i^{\text{\tiny\begin{minipage}{5em}
+      \begin{center}
+        \tiny all microstates\end{center}\end{minipage}}}
        e^{-\beta E_i}, & F &= -kT\log Z
@@
 microstates that are periodic with a given periodicity $2^j\lambda$:
+\newcommand\sumstatesperiodic[1]{\sum_i^{\text{\tiny\begin{minipage}{6em}
+      \begin{center}
+        \tiny microstates periodic $#1$\end{center}\end{minipage}}}}
 \begin{align}
-  Z_j &= \sum_i^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny periodic }2^j\lambda\\
-                 \end{array}}
+  Z_j &= \sumstatesperiodic{2^j\lambda}
        e^{-\beta E_i}, & F_j &= -kT\log Z_j
@@
 periodic states.
+\newcommand\sumoccupanciesofoctants{
+  \sum_{N_1\cdots N_8}^{\text{\tiny\begin{minipage}{5em}
+      \begin{center}
+      \tiny numbers in octants\end{center}\end{minipage}}}
+}
+\newcommand\summicrostatesofoctants{
+  \sum_{i_1\cdots i_8}^{\text{\tiny\begin{minipage}{4em}
+      \begin{center}
+      \tiny microstates\\within octants\end{center}\end{minipage}}}
+}
 \begin{align}
-  Z_j(N) &= \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
-       g_{N_1\cdots N_8}
-       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+  Z_j(N) &= \sumoccupanciesofoctants g_{N_1\cdots N_8}
+       \summicrostatesofoctants
        e^{-\beta E_{i_1\cdots i_8}^{(j)}}
@@
   F_j(n) = -kT\log\Bigg(
-    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
        \\
-       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+       \summicrostatesofoctants
        e^{-\beta E_{i_1\cdots i_8}^{(j)}}
@@
   F_{j-1}(n) = -kT\log\Bigg(
-    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
        \\
-       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+       \summicrostatesofoctants
        e^{-\beta \left(E_{i_1}^{(j-1)} + \cdots + E_{i_8}^{(j-1)}\right)}
@@
   F_{j-1}(n) = -kT\log\Bigg(
-    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
@@
   -kT\log\left(\frac{
-    \sum_{AAN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
-       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+       \summicrostatesofoctants
        e^{-\beta E_{i_1\cdots i_8}^{(j)}}
   }{
-    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
@@
   -kT\log\left(\frac{
-    \sum_{NN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
-       \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+       \summicrostatesofoctants
        e^{-\beta \left(E_{i_1\cdots i_8}^{(j)}  - F_{j-1}(n) \right)}
   }{
-    \sum_{BN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
@@
 \begin{align}
-  F_j(N_1\cdots N_8) &= \sum_{i_1\cdots i_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny microstates}\\
-                  \text{\tiny within octants}\\
-                 \end{array}}
+  F_j(N_1\cdots N_8) &= \summicrostatesofoctants
        e^{-\beta E_{i_1\cdots i_8}^{(j)}}
@@
   -kT\log\left(\frac{
-    \sum_{CN_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
@@
   }{
-    \sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-                  \text{\tiny occupancies}\\
-                  \text{\tiny of octants}\\
-                 \end{array}}
+    \sumoccupanciesofoctants
        g_{N_1\cdots N_8}
@@
 \item \fixme{Establish what $g_{N_1\cdots N_8}$ is.}
+  \begin{align}
+    g_{N_1\cdots N_8} &=
+    {N \choose N_1} {N - N_1 \choose N_2} {N - N_1 - N_2 \choose N_3}
+    \cdots \\
+    &\propto \sum_{i=1}^{8} N_i!
+    \\
+    &\approx \sum_{i=1}^{8} (N_i-1) e^{N_i}
+  \end{align}
+  This combinatoric term ends up showing up as the ideal gas free
+  energy term.
 \item \fixme{Elaborate on the approximation (ignoring corners) that
-  leads from a $\sum_{N_1\cdots N_8}^{\scriptscriptstyle\begin{array}{c}
-    \text{\tiny occupancies}\\ \text{\tiny of octants}\\
-                 \end{array}}
-       g_{N_1\cdots N_8}$ to an integral over density
-       amplitudes.}
+  leads from a $\sumoccupanciesofoctants g_{N_1\cdots N_8}$ to an
+  integral over density amplitudes.}
 \item \fixme{Develop an approach for approximating $\bar{U}$.  Perhaps

### 2014-05-20 papers/pair-correlation/paper.tex
try to clarify wording of section IV and paragraph vefore Tested on bennet. Took 0:17.17.

@@
 is the density averaged over a sphere with diameter
-$0.8\sigma$~\cite{sokolowski1992role, wadewitz2000application, winkelmann2001liquid}.  The
-value 0.8 in this formula was arrived at by fitting to Monte Carlo
-simulation.  Although Eq.~\ref{eq:sokolowski} has the advantages of
-only involving density averages at the points at which the pair
-distribution function is desired, it cannot be written as a
-single-site convolution, since the convolution kernel depends on both
-points.
+$0.8\sigma$~\cite{sokolowski1992role}.  The value 0.8 in this formula
+was arrived at by fitting to Monte Carlo simulation.  Although
+Eq.~\ref{eq:sokolowski} has the advantages of only involving density
+averages at the points at which the pair distribution function is
+desired, it cannot be written as a single-site convolution, since the
+convolution kernel depends on both points.  However, this approach has
+been used by two papers to do
+\fixme{what?}~\cite{wadewitz2000application, winkelmann2001liquid}.
 
@@
 position $\rr$, and $F_{HS}$ is the Helmholtz free energy of the
-hard-sphere fluid.  The approximation for $g_\sigma(\rr)$ is obtained
-by using the White Bear functional~\cite{roth2002whitebear} for the
-Helmholtz free energy $F_{HS}$ in Eq.~\ref{eq:gsigma-exact}.  This
+hard-sphere fluid.  The functional derivative with of free energy with
+respect to hard-sphere diameter in Eq.~\ref{eq:gsigma-exact} requires
+that we be able to evaluate the change in free energy resulting from a
+change in the diameter of specifically the hard spheres located at
+position $\rr$.  This somewhat unusual construction is mathematically
+straightforward within Fundamental Measure Theory
+(FMT)~\cite{rosenfeld1989free}.  We employ the White Bear variation of
+the FMT free energy functional~\cite{roth2002whitebear}, which
 provides an excellent approximation for this averaged value of the
@@
 was dependent upon the density averaged over some volume.  We seek to
-achieve greater accuracy by using instead a function dependent upon
+achieve greater accuracy by making use of a function dependent upon
 our averaged $g_{\sigma}(\rr)$ discussed above, which holds more

### 2014-05-20 papers/renormalization/paper.tex
handle multiplicity in RG writeup Tested on bennet. Took 0:15.92.

--- 2013-11-19 papers/renormalization/paper.tex
@@
 \documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
+
 \usepackage{graphicx}% Include figure files
-\usepackage{dcolumn}% Align table columns on decimal point
 \usepackage{color}
@@
 
-\fixme{Here we will explain the RGT that we use, and how it works
-  out.}
+\subsubsection{Defining our approximation}
+Let's begin by considering the partition function and free energy, as
+computed in the canonical ensemble:
+\begin{align}
+  Z &= \sum_i^{\text{\tiny\begin{minipage}{5em}
+      \begin{center}
+        \tiny all microstates\end{center}\end{minipage}}}
+       e^{-\beta E_i}, & F &= -kT\log Z
+\end{align}
+We now consider a ``smooth'' configuration that contains only
+microstates that are periodic with a given periodicity $2^j\lambda$:
+\newcommand\sumstatesperiodic[1]{\sum_i^{\text{\tiny\begin{minipage}{6em}
+      \begin{center}
+        \tiny microstates periodic $#1$\end{center}\end{minipage}}}}
+\begin{align}
+  Z_j &= \sumstatesperiodic{2^j\lambda}
+       e^{-\beta E_i}, & F_j &= -kT\log Z_j
+       \label{eq:Zn}
+\end{align}
+This defines a family of free energies that contain fluctuations with
+a variety of length scales.  For the moment we will assume a simple
+cubic periodicity, although it may be advantageous to select face
+centered cubic (FCC) instead.  So far, we have defined a family of
+free energy functions (each containing longer length-scale
+fluctuations) that we can compute numerically using Monte Carlo, but
+we have not yet constructed a means to approximate these functions.
+Before developing this approximation, let's reorganize this sum over
+periodic states.
+\newcommand\sumoccupanciesofoctants{
+  \sum_{N_1\cdots N_8}^{\text{\tiny\begin{minipage}{5em}
+      \begin{center}
+      \tiny numbers in octants\end{center}\end{minipage}}}
+}
+\newcommand\summicrostatesofoctants{
+  \sum_{i_1\cdots i_8}^{\text{\tiny\begin{minipage}{4em}
+      \begin{center}
+      \tiny microstates\\within octants\end{center}\end{minipage}}}
+}
+\begin{align}
+  Z_j(N) &= \sumoccupanciesofoctants g_{N_1\cdots N_8}
+       \summicrostatesofoctants
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+\end{align}
+where here $g_{N_1\cdots N_8}$ is a multiplicity factor
+corresponding to the number of ways the cell can be divided into
+subcells with occupancies of $N_1\cdots N_8$.  This is so far only a
+different way of writing Eq.~\ref{eq:Zn}.  At this point, we can
+multiply and divide by a few factors, with a goal of replacing the
+unknown $E_{i_1\cdots i_8}^{(j)}$ with something that is easier to
+approximate.
+\begin{multline}
+  F_j(n) = -kT\log\Bigg(
+    \sumoccupanciesofoctants
+       g_{N_1\cdots N_8}
+       \\
+       \summicrostatesofoctants
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+  \Bigg)
+\end{multline}
+
+\subsubsection{Free energy of the sub-systems}
+
+Now we consider the same free energy computed without accounting for
+the energy in the fluctuations:
+\begin{multline}
+  F_{j-1}(n) = -kT\log\Bigg(
+    \sumoccupanciesofoctants
+       g_{N_1\cdots N_8}
+       \\
+       \summicrostatesofoctants
+       e^{-\beta \left(E_{i_1}^{(j-1)} + \cdots + E_{i_8}^{(j-1)}\right)}
+  \Bigg)\label{eq:Fjm1}
+\end{multline}
+We can now recognize that each of the octants is just a similar smooth
+system.
+
+The multiplicity term $g_{N_1\cdots N_8}$ that shows up in
+Eq.~\ref{eq:Fjm1} (and elsewhere) originates from the choice made when
+splitting our sum over all microstates into separate sums for
+microstates in each of the octants.  This required us to pick a
+combination of states for each octant, thus the $g_{N_1\cdots N_8}$.
+This multiplicity is given by:
+\begin{align}
+  g_{N_1\cdots N_8} &=
+  {N \choose N_1} {N - N_1 \choose N_2} {N - N_1 - N_2 \choose N_3}
+  \cdots \\
+  &\propto \sum_{i=1}^{8} N_i!
+  \\
+  &\approx \sum_{i=1}^{8} (N_i-1) e^{N_i}
+\end{align}
+Physically, this multiplicity term corresponds to the ideal gas
+contribution to the free energy.  In fact, the ideal gas free energy
+represents \emph{only} the multiplicity and kinetic energy.  Since the
+free energy incorporates the multiplicity, we can simplify
+Eq.~\ref{eq:Fjm1} as:
+\begin{multline}
+  F_{j-1}(n) = -kT\log\Bigg(
+    \sumoccupanciesofoctants
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8) \right)}
+  \Bigg)
+\end{multline}
+where I have introduced the confusing notation where $F_{j-1}(n)$
+represents the free energy computed for any density $n$ representable
+in the $2^j\lambda$ cell by averaging over particle distributions
+$N_1\cdots N_8$, while $F_{j-1}(N)$ represents the canonical free
+energy found for a set of microstates with $N$ particles in the cell
+volume.
+
+Putting these last two equations together, we can find a formula for
+the energy difference as we include the larger fluctuations:
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sumoccupanciesofoctants
+       g_{N_1\cdots N_8}
+       \summicrostatesofoctants
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+  }{
+    \sumoccupanciesofoctants
+        e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8) \right)}
+  }
+  \right)
+\end{multline}
+At this point, we recognize that we can put anything we like in the
+numerator \emph{and} the denominator, which will help us to prevent
+the exponentials from getting too hairy.  This is really just a shift
+of energy for both the numerator and denominator partition functions.
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sumoccupanciesofoctants
+       g_{N_1\cdots N_8}
+       \summicrostatesofoctants
+       e^{-\beta \left(E_{i_1\cdots i_8}^{(j)}  - F_{j-1}(n) \right)}
+  }{
+    \sumoccupanciesofoctants
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
+  }
+  \right)
+\end{multline}
+At this stage we have still not introduced an uncontrolled
+approximation.  Each $F_{j}$ is exactly defined, and the exact free
+energy is found by taking a limit $j\rightarrow \infty$.  \fixme{I
+  need to clarify the scale of each $F$.  Some of the above are about
+  systems $\frac18$ the volume of others.}
+
+The flavor of the approximation used by White and others in GRG is now
+clarified.  We begin by assuming that we already have an adequate (and
+tractable) approximation for $F_{j-1}$.  Thus the denominator
+partition function is readily computable.  We further simplify the
+numerator by multiplying and dividing by the same quantity, and
+simplify the notation by defining
+\begin{align}
+  F_j(N_1\cdots N_8) &= \summicrostatesofoctants
+       e^{-\beta E_{i_1\cdots i_8}^{(j)}}
+\end{align}
+and thus rewrite the energy difference as
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sumoccupanciesofoctants
+       e^{-\beta \big(F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) +
+           \big(F_j(N_1\cdots N_8) - F_{j-1}(N_1) - \cdots - F_{j-1}(N_8)\big) \big)}
+  }{
+    \sumoccupanciesofoctants
+       e^{-\beta \left( F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \right)}
+  }
+  \right)
+\end{multline}
+At this point we can recognize and identify two energy differences as
+\begin{align}
+  \bar{F}(N_1\cdots N_8) &= F_{j-1}(N_1) + \cdots F_{j-1}(N_8)  - F_{j-1}(n) \\
+  \bar{U}(N_1\cdots N_8) &= F_j(N_1\cdots N_8)
+  - \left(F_{j-1}(N_1) + \cdots + F_{j-1}(N_8) \right)
+\end{align}
+and using these definitions can rewrite our energy difference in a
+more compact form:
+\begin{multline}
+  F_j(n) - F_{j-1}(n) =\\
+  -kT\log\left(\frac{
+    \sumoccupanciesofoctants
+       e^{-\beta \big( \bar{F}(N_1\cdots N_8) + \bar{U}(N_1\cdots N_8)
+         \big)}
+  }{
+    \sumoccupanciesofoctants
+       e^{-\beta \bar{F}(N_1\cdots N_8)}
+  }
+  \right)
+\end{multline}
+At this point, we can recognize $\bar{F}(N_1\cdots N_8)$ as the free
+energy change due to oscillations \emph{if we neglect the actual
+  energy of the oscillations themselves, but instead make a ``local''
+  approximation}.  Meanwhile $\bar{U}(N_1\cdots N_8)$ is the free
+energy change due to the energy of oscillations.
+
+\fixme{I have now ``defined'' the terms used by Forte in her
+  notation of GRG.  On the to-do list: }
+\begin{enumerate}
+\item \fixme{Elaborate on the approximation (ignoring corners) that
+  leads from a $\sumoccupanciesofoctants g_{N_1\cdots N_8}$ to an
+  integral over density amplitudes.}
+\item \fixme{Develop an approach for approximating $\bar{U}$.  Perhaps
+  is $F_{j-1}$ is expressed as a functional, we can use a non-local
+  DFT to approximate $\bar U_j$?  The standard GRG approach seems to
+  be to use a simple mean-field or thermodynamic perturbation theory
+  approach to approximate $\bar U$, suggesting that a simple
+  (standard) DFT might work just fine for this.}
+\end{enumerate}
 

### 2014-05-20 papers/thesis-roth/project.tex
Project report writing Tested on elliot. Took 0:32.16.

@@
 \newcommand{\2}{\ensuremath{\textbf{r}_2}}
+\newcommand{\3}{\ensuremath{\textbf{r}_3}}
+\newcommand{\4}{\ensuremath{\textbf{r}_4}}
 
@@
 
-%% \fixme{The following comes from Forte~\cite{Forte11} and doesn't make a whole lot of sense to me}
-
-%% The free energy for each iteration is given by
-%% \begin{equation}
-%%   f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)\ .
-%% \end{equation}
-%% The total free is obtained when one takes the
-%% iterations to infinity and adds in the attractive component:
-%% \begin{equation}
-%%   f(T,n) = \lim_{i \to \infty} f_i(T,n) + \fattr\ .
-%% \end{equation}
-%% The initial free energy density, $f_0$ depends on the liquid model of study. For this work, I use the square well model:
-%% \begin{equation}
-%%   f_0(T,n) = \fid + \fhs + \frac{a_2^{SW}(T,n)}{\kT}\ ,
-%% \end{equation}
-%% where $a_2^{SW}(T,n)$ is the second-order term in a high-temperature expansion around $1/\kT$. (See ref~\cite{Gil-Villegas97}.)
-%% \fixme{End of Forte confusion}
+Forte gives $\delta f_i(T,n)$ as
+\begin{align}
+  \delta f_i(T,n) &= -\frac{\kT}{V_D}\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
+\end{align}
+where
+\begin{align}
+  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \bar{f}_D(T,n) + \bar{u}_D(T,n \right) \right] \\
+  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\bar{f}_D(T,n) \right]
+\end{align}
+and
+\begin{align}
+  \bar{f}_D(T,n) &= \frac{f_{i-1}(T,n+x) + f_{i-1}(T,n-x)}{2} - f_{i-1}(T,n) \\
+  \bar{u}_D(T,n) &= \frac{u(T,n+x,\lambda_D) + u(T,n-x,\lambda_D)}{2} - u(T,n,\lambda_D)
+\end{align}
+Finally, $u(T,n)$ is given by
+\begin{align}
+  \begin{split}
+  u(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \\ &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma
+  \end{split}
+\end{align}
+with
+\begin{align}
+  \alpha &= \frac{2}{3}\pi\epsilon\sigma^3(\lambda^3 - 1) \\
+  \omega^2 &= \frac{1}{5}\sigma^2\frac{\lambda^5 - 1}{\lambda^3 - 1} \\
+  \gamma &= \frac{1}{70}\sigma^4\frac{\lambda^7 - 1}{\lambda^3 - 1}
+\end{align}
 
@@
 
-In Gil-Villegas' paper, HS and Dispersion are wrapped up into the
+In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
 \textit{monomer} contribution, expressed in terms of energy densities
@@
 \end{align}
+For this work I use $\lambda = 1.5$.
 
@@
 \end{align}
+\fixme{I don't really understand how this derivative is \ldots derived}
 Putting this together, we have
@@
 \end{align}
-Applying this to the second term in equation~\ref{eqn:Henderson5-2-12}, we have
+Applying this to the \fixme{0th, 1st or 2nd, depending on how you think of it} term in equation~\ref{eqn:Henderson5-2-12}, we have
 \begin{align}
   \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
-  &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)
+  &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)w(\1,\2)
 \end{align}
+The second-order $\lambda$ term is quite a bit more complicated:
+\begin{widetext}
+ \begin{align}
+   \begin{split}
+   \frac{\beta F_2}{N} =&{} -\frac{\beta^2}{2}\left[ \frac{n}{2}\int d\2\, g_0(\1,\2)w^2(\1,\2) \right.
+\\ &{} + n^2 \iint d\2 d\3\, g_0^{(3)}(\1,\2,\3)w(\1,\2)w(\1,\3) \\ &{} \left.
++ \frac{n^3}{4}\iiint d\2 d\3 d\4\, \left( g_0^{(4)}(\1,\2,\3,\4) - g_0^{(2)}(\1,\2)g_0^{(2)}(\3,\4) \right)w(\1,\2)2(\3,\4) \right] \\ &{} - \frac{1}{4}S_0(0)\left[ \frac{\partial}{\partial n}\left( n^2\int d\2\, g_0(\1,\2)w(\1,\2) \right)^2 \right]
+   \end{split}
+ \end{align}
+\end{widetext}
+This term requires use to know the form of higher-order correlation functions and the structure function $S_0(k)$ of the reference system.
+These things are very difficult.
+Hence, we use approximations, as outlined in section~\ref{sub2sec:disp}.
 
@@
 
+\subsection{Free Energies}\label{subsec:free-energies}
+The $a_j(n)$ terms require even more simplification when ``converting'' from the mathematical formulation to Python code, to rid ourselves of troublesome integrals. Gil-Villegas gives us
+\begin{align}
+  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{eff}) \\
+  &= -4\eta\epsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{eff}/2 \right)}{(1 - \eta_{eff})^3}
+\end{align}
+and
+\begin{align}
+  \eta_{eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
+\end{align}
+with
+\begin{align}
+  \left( \begin{array}{c}
+    c_1 \\
+    c_2 \\
+    c_3
+    \end{array} \right)
+  &= \left( \begin{array}{ccc}
+    2.25855 & -1.50349 & 0.249434 \\
+    -0.669270 & 1.40049 & -0.827739 \\
+    10.1576 & -15.0427 & 5.30827
+    \end{array} \right)
+  \left( \begin{array}{c}
+    1 \\
+    \lambda \\
+    \lambda^2
+    \end{array} \right)
+\end{align}
+
+\subsection{RG partition functions}\label{subsec:fbar-ubar}
+I use the midpoint method for evaluating the integrals of $\bar{f}_D$ and $\bar{u}_D$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-lowT}
+  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
   \end{center}
   \caption{Free energy density for RG at low temperature}
-  \label{fig:SW-RG-compare-lowT}
+  \label{fig:phi-RG-lowT}
 \end{figure}
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/SW-RG-compare-phi-highT}
+  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
   \end{center}
   \caption{Free energy density for RG near critical temp}
-  \label{fig:SW-RG-compare-highT}
+  \label{fig:phi-RG-highT}
 \end{figure}
+
+Figures~\ref{fig:phi-RG-lowT} and~\ref{fig:phi-RG-highT} show the grand free energy per unit volume at a low temperature, and near the critical point. It is clear that the program is not de-bugged.
 

### 2014-05-21 papers/thesis-roth/project.tex
comments on Roth project report Tested on bennet. Took 0:33.33.

@@
 \newcommand{\fixme}[1]{\textcolor{red}{\textbf{[#1]}}}
+\newcommand{\davidsays}[1]{\textcolor{blue}{\textit{[#1]}}}
 
@@
 \subsubsection{Ideal gas}\label{sub2sec:ID}
-The ideal gas term is purely local \fixme{whatever that means}:
+The ideal gas term is purely local \fixme{whatever that
+  means}\davidsays{For your purposes, the purely localness isn't
+  really relevant.  It means that when we consider an inhomogeneous
+  configuration, the correct answer is just this free energy
+  integrated.  i.e. you just add up the local contributions to the
+  free energy.  For a homogeneous fluid, the locality isn't really
+  relevant.  Except in so far as how it relates to
+  fluctuations... since fluctuations are inhomogeneities of a sort.}:
 \begin{align}
@@
 
-\fixme{And introduction to TPT---why do we use it?}
+\fixme{And introduction to TPT---why do we use it?}\davidsays{We use
+  TPT because it gives a good description of simple liquids, as long
+  as you are far from the critical point.  Thus it covers most of our
+  bases.  It is also an elegant theory, since it comes from a rigorous
+  power series expansion in a small parameter... or one that one hopes
+  is small. }
+
 This discussion follows primarily from \textit{Theory of Simple Liquids}, by Hansen and McDonald\cite{Hansen06}.
@@
 \end{align}
-For this work I use $\lambda = 1.5$.
+For this work I use $\lambda = 1.5$.  \davidsays{I think you've gotten
+things a little confused.  When $v_\lambda$ is derived below, the $w$
+should be the attraction portion, and what appears as $v_{SW}$ should
+just be the hard-sphere potential (without attraction).  That's how
+the TPT works, it is the attraction that is being added in with the
+$\lambda$ proportionality.}
 
@@
   V_N(\lambda) = \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j)
+  \label{eq:VN}
 \end{align}
 
-Given the \emph{configuration integral} (the potenial energy component of the partition function)
+Given the \emph{configuration integral} (the potential energy component of the partition function)
 \begin{align}
@@
 \end{align}
-where $V^N$ is the total volume \fixme{I think?}, then the derivative of the excess free energy with respect to the coupling parameter $\lambda$ is
+where $V^N$ is the total volume \fixme{I think?}\davidsays{It is the
+  volume taken to the $N^{th}$ power.  It's unfortunate that we have a
+  $V_N$ and a $V^N$ that are such different quantities here!}, then
+the derivative of the excess free energy with respect to the coupling
+parameter $\lambda$ is
 \begin{align}
   \frac{\partial F}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
-  &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda} \\
+  &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda}
+  \label{eqn:dFdlambda} \\
   \hookrightarrow \beta\frac{\partial F}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
@@
 \end{align}
+\davidsays{In the above you could simplify things a bit by working out
+$dF/d\lambda$ directly (without the $\beta$ factor) which would remove
+all the factors of $\beta$.  I would also mention that the subscript
+$\lambda$ indicates that the average (and derivative) is evaluated at
+$\lambda = \lambda$ (as opposed to $\lambda=0$ which is done later.
+Note that at this stage we have an exact equation.  We just took a
+derivative of $F$ and then integrated that derivative to find $F$
+again.  So far as I can see (and I looked at this derivation a bit
+with Eric, who may also be able to help you), $V'_N = \sum w_{ij}$, which
+means that $\langle V'_N\rangle = \frac12 N^2 \langle w_{12}
+\rangle$, which is potentially a simpler way to look at it.}
 where $\left\langle V_N^{'}(\lambda) \right\rangle_\lambda$ is the canonical ensemble average. We can take a series expansion of this average:
@@
 \end{align}
-\fixme{I don't really understand how this derivative is \ldots derived}
-Putting this together, we have
+\fixme{I don't really understand how this derivative is \ldots
+  derived}\davidsays{This is definitely a spot where you can be more
+  clear (once you understand).  I believe the $V''_N$ term is zero
+  (see above).  Unless I'm severely misunderstanding
+  equations~\ref{eqn:small-perturbation} and~\ref{eq:VN}.  The
+  derivative is (pretty) easy if you write your ensemble average as a
+  multidimensional integral, as in Eq.~\ref{eqn:dFdlambda}.  Write
+  down $dF/d\lambda$ in terms of $Z_N$ and an integral of Boltzmann
+  factors times $V_N'$.  Then when you take a derivative, it hits top
+  and bottom, but each is an integral of an exponential, so those are
+  easy to differentiate.  A little math will give you this variance
+  formula.} Putting this together, we have
 \begin{align}
@@
 \end{align}
-With this in mind, we can simplify equation~\ref{eqn:complicated-HTE} to the \emph{high-temperature expansion}, first derived by Zwanzig:\cite{Zwanzig54}
+\davidsays{Ah ha! Here you do what I mentioned earlier.  I'd tend to
+  put this earlier, so you can throw away the $V_N''$ term sooner.
+  And also talking about $W_N$ with this definition I would find nicer
+  and clearer than all the $V_N'$.}  With this in mind, we can
+simplify equation~\ref{eqn:complicated-HTE} to the
+\emph{high-temperature expansion}, first derived by
+Zwanzig:\cite{Zwanzig54}
 \begin{align}
@@
 
-Writing the statistical averages in terms of the pair density $n_\lambda^{(2)}(\1,\2)$ (i.e. the probability of having a sphere at \1 and \2 \fixme{make this more clear?}) gives us
+Writing the statistical averages in terms of the pair density
+$n_\lambda^{(2)}(\1,\2)$ (i.e. the probability of having a sphere at
+\1 and \2 \fixme{make this more clear?}\davidsays{I'd perhaps phrase
+  it as ``probability per volume squared of having one sphere at \1
+  and another at \2''}) gives us
 \begin{align}
@@
 \end{align}
-The pair density itself can be expanded as
+\davidsays{Here's another place where all the $N$s on the bottom are
+  muddying things up, I think.  (like the $\beta$s were before).  I
+  also think this discussion is pretty redundant with parts of the
+  previous derivation in terms of $V_N$.}  The pair density itself can
+be expanded as
 \begin{align}
@@
 \end{align}
-Applying this to the \fixme{0th, 1st or 2nd, depending on how you think of it} term in equation~\ref{eqn:Henderson5-2-12}, we have
+Applying this to the \fixme{0th, 1st or 2nd, depending on how you
+  think of it}\davidsays{it's ``second'' since ``terms'' are things
+  that are added together, while factors are things that are
+  multiplied together.  And in Eq~\ref{eqn:Henderson5-2-12} there are
+  only two things added together.  (I figured this terminology out on
+  my own some time after becoming a professor... it seems to be used,
+  but not taught).} term in equation~\ref{eqn:Henderson5-2-12}, we
+have
 \begin{align}
@@
 Hence, we use approximations, as outlined in section~\ref{sub2sec:disp}.
+
+\davidsays{I'm thinking this whole section should go \emph{before} the
+  section on the square well free energy, since that section uses this
+  method (in approximation).  Then you can say something to the effect
+  of ``in the next section we will see a simple approximation for this
+  term.''  Hansen's explanation of TPT is not very clear, which makes
+  it all the more valuable to make \emph{this} a clear and lucid
+  explanation.  I'd love to see just a couple more steps at key
+  places---largely where you were confused.  Probably it's best to
+  actually talk about this section in person before rewriting.  This
+  has been an excellent first shot, though! }
 
@@
 \subsection{Free Energies}\label{subsec:free-energies}
-The $a_j(n)$ terms require even more simplification when ``converting'' from the mathematical formulation to Python code, to rid ourselves of troublesome integrals. Gil-Villegas gives us
+The $a_j(n)$ terms require even more simplification when
+``converting'' from the mathematical formulation to Python code, to
+rid ourselves of troublesome integrals. Gil-Villegas gives us
+\newcommand\eff{\textit{eff}}
 \begin{align}
-  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{eff}) \\
-  &= -4\eta\epsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{eff}/2 \right)}{(1 - \eta_{eff})^3}
+  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{\eff}) \\
+  &= -4\eta\epsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{\eff}/2 \right)}{(1 - \eta_{\eff})^3}
 \end{align}
@@
 \begin{align}
-  \eta_{eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
+  \eta_{\eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
 \end{align}

### 2014-05-21 papers/renormalization/paper.tex
Project report writing Tested on elliot. Took 0:32.16.
more renormalization work Tested on bennet. Took 0:18.13.

@@
 periodic states.
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/divisions}
+  \caption{A cell being subdivided with $N_1\cdots N_8$ particles in
+    each subdivision.}
+\end{figure}
 \newcommand\sumoccupanciesofoctants{
@@
 \begin{align}
-  F_j(N_1\cdots N_8) &= \summicrostatesofoctants
+  F_j(N_1\cdots N_8) &= g_{N_1\cdots N_8}
+       \summicrostatesofoctants
        e^{-\beta E_{i_1\cdots i_8}^{(j)}}
 \end{align}
+\begin{widetext}
 and thus rewrite the energy difference as
@@
 \end{multline}
+\end{widetext}
 At this point we can recognize and identify two energy differences as
@@
 
-\fixme{I have now ``defined'' the terms used by Forte in her
-  notation of GRG.  On the to-do list: }
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/interactions}
+  \caption{Interactions between neighboring subdivisions.}\label{fig:interactions}
+\end{figure}
+
+\subsubsection{Interaction between subsystems}
+At this stage, there is just one term that need be approximated
+(assuming, as always, that we have an acceptable approximation for
+$F_{j-1}$), which is $\bar{U}$.  This is the free energy difference
+between two free energies.  The first, $F_j(N_1\cdots N_8)$, is the
+fully interacting system that is periodic with periodicity
+$2^j\lambda$, and which is constrained to have $N_1\cdots N_8$
+particles in each of the subdivisions of the cell.  The second is the
+sum of $F_{j-1}$ for each of those subdivisions.  This difference
+describes the interaction between the octants of the cell, and is at
+the heart of the GRG method.
+
+We desire to simplify the summation over seven independent numbers
+$N_1\cdots N_7$ (with the eighth constrained by $\sum_i N_i = N$) to a
+summation over a single fluctuation amplitude.  As illustrated in
+Fig.~\ref{fig:interactions}, these interactions can be seen as
+a sum of interactions between neighboring subdivisions, combined with
+a many-body correction due to ``corner'' interactions.
+
+\subsection{To do later}
+
 \begin{enumerate}

### 2014-05-21 papers/renormalization/paper.tex
more renormalization work Tested on bennet. Took 0:18.13.

@@
   \includegraphics[width=0.8\columnwidth]{figs/interactions}
-  \caption{Interactions between neighboring subdivisions.}\label{fig:interactions}
+  \caption{Interactions between all subdivisions.}\label{fig:interactions}
 \end{figure}
@@
 summation over a single fluctuation amplitude.  As illustrated in
-Fig.~\ref{fig:interactions}, these interactions can be seen as
-a sum of interactions between neighboring subdivisions, combined with
-a many-body correction due to ``corner'' interactions.
+Fig.~\ref{fig:interactions}, these interactions can be seen as a sum
+of interactions between neighboring subdivisions, combined with a
+many-body corrections due to the free energy not being precisely
+separable into additive pairwise contributions (due to for instance
+``corner'' interactions).
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\columnwidth]{figs/neighbors}
+  \caption{Interactions between only neighboring subdivisions.}\label{fig:neighbors}
+\end{figure}
+
+If we make this assumption, that the free energy can be separated into
+neighboring contributions, we can write
+\begin{align}
+  F_j(N_1\cdots N_8) &= \alpha \sum_{ij}^{\text{neighbors}} F_j^{\text{neighbor}}(N_i, N_j)
+\end{align}
+where $F_j^{\text{neighbor}}(N_i, N_j)$ is the free energy of the
+system illustrated in Fig.~\ref{fig:neighbors}, which has a slab-like
+density oscillation, and $\alpha$ \fixme{is a prefactor that I need to
+  work out}.
+
+I will note that we have a choice of lattice when defining our
+periodic system.  At this point we can see that a simple cubic lattice
+gives us a flat-slab neighbor system.  If we had chosen a close-packed
+FCC lattice, then each subdivision would interact equally with all the
+other subdivisions (and corners would involve four subdivisions rather
+than all eight), but our slab geometry would have a bumpy interface
+(e.g. if we chose a $\langle 111\rangle$ slab it would have an
+equilateral triangular lattice of bumps).  I will continue thinking in
+terms of a simple cubic interface, with the option to revise this to
+switch lattices.
+
+Once we have rewritten $F_j(N_1\cdots N_8)$ in terms of
+$F_j^{\text{neighbor}}(N_i, N_j)$, we can further simplify our free energy
+difference (and definitions of $\bar{U}$ and $\bar{F}$):
 

### 2014-05-23 papers/thesis-roth/project.tex
Project writing Tested on elliot. Took 0:19.69.

--- 2014-05-20 papers/thesis-roth/project.tex
@@
 % vector
-\newcommand{\rr}{\ensuremath{\bf{r}}}
+\newcommand{\rr}{\ensuremath{\mathbf{r}}}
 
@@
 
-Near the critical point of a liquid, there are density \fixme{is saying $g(\rr)$ more accurate?} fluctuations at all length scales.
-In this regime, mean field theories break down, for obvious reasons.
-As one approaches the critical point, there becomes less and less of a difference in free energy of the liquid and vapor states; this causes difficulty in modeling liquid-vapor coexistence.
-This work develops a method for calculating liquid-vapor coexistence, and applies renormalization group theory to the square well liquid.
+Near the critical point of a liquid, there are density \fixme{is
+  saying $g(\rr)$ more accurate?} fluctuations at all length
+scales. In this regime, mean field theories break down, for obvious
+reasons. As one approaches the critical point, there becomes less and
+less of a difference in free energy of the liquid and vapor states;
+this causes difficulty in modeling liquid-vapor coexistence. This work
+develops a method for calculating liquid-vapor coexistence, and
+applies renormalization group theory to the square well liquid.
 
@@
 \section{Introduction}
+\fixme{Write introduction; why do we do this? why is it useful? how is this relevant? what is the history?}
 
@@
 
-This work applies entirely to the homogeneous case---that is, there are no walls containing our liquid, nor are there any sort of solutes in the liquid.
-Since our fluid is of infinite extent, the free energy is infinite.
-Therefore, I work with free energy density:
+This work applies entirely to the homogeneous case---that is, there
+are no walls containing our liquid, nor are there any sort of solutes
+in the liquid. Since our fluid is of infinite extent, the free energy
+is infinite. Therefore, I work with free energy density:
 \begin{align}
@@
 developed to deal with large-scale fluctuations of the order parameter
-near the critical point of a system. The formulation used here is provided by Forte~\textit{et al}.\cite{Forte11}
+near the critical point of a system. The formulation used here is
+provided by Forte~\textit{et al}.\cite{Forte11}
 
@@
 
-We begin with a baseline free enrgy $f_0(T,n)$.
-We assume this baseline accounts for all fluctuations of wavelength $\lambda < \lambda_0$.
-We then double the size of our phase cell and add a correction to the free energy $\delta f_1(T,n)$.
-This correction accounts for fluctuations with a wavelength $\lambda = 2\lambda_0$.
-The new free energy is $f_1(T,n) = f_0(T,n) + \delta f_1(T,n)$.
-Then double the length scale again; $\delta f_2(T,n)$ accounts for fluctionations with $\lambda = 4\lambda_0$ and $f_2(T,n) = f_1(T,n) + \delta f_2(T,n)$.
-As we continue this process, the $i^{th}$ correction accounts for fluctuations with $\lambda = 2^i\lambda_0$, and $f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)$.
-In the end, your free energy is
+We begin with a baseline free enrgy $f_0(T,n)$. We assume this
+baseline accounts for all fluctuations of wavelength $\lambda <
+\lambda_0$. We then double the size of our phase cell and add a
+correction to the free energy $\delta f_1(T,n)$. This correction
+accounts for fluctuations with a wavelength $\lambda =
+2\lambda_0$. The new free energy is $f_1(T,n) = f_0(T,n) + \delta
+f_1(T,n)$. Then double the length scale again; $\delta f_2(T,n)$
+accounts for fluctionations with $\lambda = 4\lambda_0$ and $f_2(T,n)
+= f_1(T,n) + \delta f_2(T,n)$. As we continue this process, the
+$i^{th}$ correction accounts for fluctuations with $\lambda =
+2^i\lambda_0$, and $f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)$. In the
+end, your free energy is
 \begin{align}
@@
 
-Where $\fattr$ is the contribution due to the longest-wavelenght density fluctuations. At this stage, I do not entirely understand why Forte splits up $\fattr$.
+Where $\fattr$ is the contribution due to the longest-wavelength
+density fluctuations. At this stage, I do not entirely understand why
+Forte splits up $\fattr$.
 
@@
 
-\subsection{Square well liquid free enrgy}\label{subsec:SW}
-For the baseline, this work is based on the square well liquid free energy:\cite{Hughes13}
-\begin{align}
-  f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
-\end{align}
-The long-range attractive term, $\fattr$ is given by
-\begin{align}
-  \fattr &= n a_1(n) \ .
-\end{align}
-$\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere repulsion, and $\fdisp$ is the attraction.
-The function $a_1(n)$ is the first term from a high-temperature perturbation expansion, described in further detail in sections~\ref{sub2sec:disp} and~\ref{subsec:TPT}.
-Each function is outlined below.
-
-\subsubsection{Ideal gas}\label{sub2sec:ID}
-The ideal gas term is purely local \fixme{whatever that means}:
-\begin{align}
-  \fid &= n\kT\left(\log(n) - 1\right)
-\end{align}
-
-\subsubsection{Hard sphere repulsion}\label{sub2sec:HS}
-Hard sphere repulsive forces are dealt with by the White Bear version of Fundamental-Mearuse Theory.\cite{Roth02} The hard sphere excess free energy is given in the homogeneous case as
-\begin{align}
-  \fhs &= \kT\left( \Phi_1(n) + \Phi_2(n) + \Phi_3(n) \right)
-\end{align}
-The $\Phi_j(n)$ terms are given as
-\begin{align}
-  \Phi_1(n) &= -n_0\ln(1 - n_3) \\
-  \Phi_2(n) &= \frac{n_1n_2}{1 - n_3} \\
-  \Phi_3(n) &= n_2^3\left( \frac{n_3 + (1 - n_3)^2\ln(1 - n_3)}{36\pi n_3^2(1 - n_3)^2} \right)
-\end{align}
-The \emph{fundamental measure densities} are
-\begin{align}
-  n_0(n) &= \frac{n_2}{4\pi R^2} \nonumber \\
-  &= n \\
-  n_1(n) &= \frac{n_2}{4\pi R} \nonumber \\
-  &= nR \\
-  n_2(n) &= n4\pi R^2 \\
-  n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
-  &= \eta
-\end{align}
-where $\eta$ is the packing fraction.
-
-\subsubsection{Attraction}\label{sub2sec:disp}
-The attractive free energy includes van der Waals attraction between hard spheres and is founded in Thermodynamic Perturbation Theory (TPT).
-\begin{align}
-  \fdisp &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
-\end{align}
-The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}.
-
-The first term, $a_1$, is the mean-field dispersion interaction.
-The second term, $a_2$, describes the effect of fluctuations resulting from compression of the fluid due to the dispersion interaction itself, and is approximated using the local compressibility approximation (LCA), which assumes the energy fluctuation is simply related to the compressibility of a hard-sphere reference fluid.\cite{Barker76}
-
-In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
-\textit{monomer} contribution, expressed in terms of energy densities
-as:
-\begin{align}
-  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
-\end{align}
-with $\alpha^\text{VDW}$ given by
-\begin{align}
-  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
-  \intertext{using reduced units of $x = r/\sigma$, we have}
-  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
-  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
-\end{align}
-where $b^\text{VDW}$ is the van der Waals size parameter. It corresponds to the volume excluded by two spherical particles of volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
-
-Then, the high-temp expansion is an expansion of the monomer term:
-\begin{align}
-  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
-\end{align}
-
-The term $a_1$ is given by
-\begin{align}
-  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
-  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
-\end{align}
-If we assume random correlations between the particles' positions, for all distances, we have $g^\text{HS}(r) = 1$. This yields
-\begin{align}
-  a_1^\text{VDW} = - n \alpha^\text{VDW},
-\end{align}
-the van der Waals mean-field energy.
-
-The second-order term is quite a bit more complicated; one must have knowledge of higher-order correlation functions to determine $a_2$.
-This term describes the response of the attractive energy due to the compression of the fluid from the attractive well.
-Gil-Villegas \emph{et al.} base their expression on an approximation by Barker and Henderson.\cite{Barker67} Barker and Henderson's approximation considers fluctuations in the number of particles in the potential well.
-In this way, the fluctuations in $a_2$ are related to the pressure and compressibility of the liquid.
-Given pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} = kT\left(\partial n  /\partial P^\text{HS}\right)_T$, we have:
-\begin{align}
-  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
-  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
-  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
-\end{align}
-
-We use the isothermal compressibility in terms of packing fraction $\eta$, given by the Percus-Yevick exprssion:\cite{Barker76}
-\begin{align}
-  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
-\end{align}
-
 \subsection{Thermodynamic perturbation theory}\label{subsec:TPT}
 
-\fixme{And introduction to TPT---why do we use it?}
-This discussion follows primarily from \textit{Theory of Simple Liquids}, by Hansen and McDonald\cite{Hansen06}.
+\fixme{And introduction to TPT---why do we use it?}\davidsays{We use
+  TPT because it gives a good description of simple liquids, as long
+  as you are far from the critical point.  Thus it covers most of our
+  bases.  It is also an elegant theory, since it comes from a rigorous
+  power series expansion in a small parameter... or one that one hopes
+  is small. }
 
-Thermodynamic perturbation theory treats the intermolecular potential as separable into two parts, one for short-range repulsion and another for long-range attraction.
+This discussion follows primarily from \textit{Theory of Simple
+  Liquids}, by Hansen and McDonald\cite{Hansen06}.
 
-The pair potential $v(\1,\2)$ describes the interaction between two hard spheres. For the square-well model, the pair potential is
+Thermodynamic perturbation theory treats the intermolecular potential
+as separable into two parts, one for short-range repulsion and another
+for long-range attraction.
+
+The pair potential $v(\1,\2)$ describes the interaction between two
+hard spheres. For the square-well model, the pair potential is
 \begin{align}
-  v_\text{SW}(\1,\2) &=
+  v_\text{HS}(\1,\2) &=
     \begin{cases}
       \infty & r < \sigma \\
-      -\epsilon & \sigma < r < \lambda\sigma \\
-      0 & \lambda\sigma < r
+      0 & \sigma < r
     \end{cases}
 \end{align}
-For this work I use $\lambda = 1.5$.
+\fixme{When $v_\lambda$ is derived below, the $w$
+should be the attraction portion (i.e.~the $-\varepsilon$ between $\sigma$ and $\lambda\sigma$), and $v_{HS}$ should
+just be the hard-sphere potential (without attraction).  That's how
+the TPT works, it is the attraction that is being added in with the
+$\lambda$ proportionality.}
 
@@
 \begin{align}
-  v_\lambda(\1,\2) &= v_\text{SW}(\1,\2) + \lambda w(\1,\2) \label{eqn:small-perturbation}
+  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambda w(\1,\2) \label{eqn:small-perturbation}
 \end{align}
 
-The total potential energy of our system is
+The total potential energy of our system and its derivatives with resepect to $\lambda$---which become important later---are
 \begin{align}
-  V_N(\lambda) = \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j)
+  V_N(\lambda) &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
+  \hookrightarrow \frac{\partial V_N(\lambda)}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
+  &\equiv W_N \\
+  \hookrightarrow \frac{\partial^2 V_N(\lambda)}{\partial\lambda^2} &= 0
 \end{align}
 
-Given the \emph{configuration integral} (the potenial energy component of the partition function)
+The \emph{configuration integral} (the potential energy component of the partition function) and its derivative with respect to $\lambda$ are
 \begin{align}
-  Z_N &= \int d\mathbf{r}^N\, e^{-\beta V_N}
+  Z_N(\lambda) &= \int d\rr^N\, e^{-\beta V_N(\lambda)} \\
+  \hookrightarrow \frac{\partial Z_N(\lambda)}{\partial\lambda} &=  \int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N(\lambda)} \\
+  &= \int d\rr^N\, -\beta\frac{\partial V_N(\lambda)}{\partial\lambda}e^{-\beta V_N(\lambda)} \\
+  &= \int d\rr^N\, -\beta W_N e^{-\beta V_N(\lambda)}
 \end{align}
-and the non-ideal part of the free energy (i.e.~the excess free energy) is
+And the non-ideal part of the free energy (i.e.~the excess free energy) and its derivative are
 \begin{align}
-  F &= -\kT\ln\left( \frac{Z_N}{V^N} \right)
+  F &= -\kT\ln\left( \frac{Z_N}{V^N} \right) \\
+  \hookrightarrow \frac{\partial F}{\partial\lambda} &= \frac{V^N}{Z_N}\frac{\partial}{\partial\lambda}\left( \frac{Z_N}{V^N} \right) \\
+  &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
+  &= \frac{1}{Z_N} \int d\rr^N\, W_N e^{-\beta V_N(\lambda)}\\
+  &= \left\langle W_N \right\rangle_\lambda
 \end{align}
-where $V^N$ is the total volume \fixme{I think?}, then the derivative of the excess free energy with respect to the coupling parameter $\lambda$ is
+(where $-\kT$ and $-\beta$ have divided out to $1$). Here, $V^N$ is
+the total volume taken to the $N^{th}$ power, and
+$\left\langle W_N \right\rangle_\lambda$ is the ensemble
+average of $W_N$, with the subscript indicating some given
+value of $\lambda$. We can simplify this even further by noting
 \begin{align}
-  \frac{\partial F}{\partial\lambda} &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
-  &= -\kT\frac{1}{Z_N}\int d\mathbf{r}^N\, e^{-\beta V_N}(-\beta)\frac{\partial V_N}{\partial\lambda} \\
-  \hookrightarrow \beta\frac{\partial F}{\partial\lambda} &= \beta\left\langle V_N^{'}(\lambda) \right\rangle_\lambda \\
-  \hookrightarrow \beta F &= \beta F_0 + \beta\int_0^1 d\lambda\, \left\langle V_N^{'}(\lambda) \right\rangle_\lambda
+  \left\langle W_N \right\rangle &= \left\langle \sum_i\sum_jw(\mathbf{r}_i,\mathbf{r}_j) \right\rangle \\
+  &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle
+  \intertext{which gives us}
+  \frac{\partial F}{\partial\lambda} &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle_\lambda
 \end{align}
-where $\left\langle V_N^{'}(\lambda) \right\rangle_\lambda$ is the canonical ensemble average. We can take a series expansion of this average:
+
+At this point, note that we have an exact equation with no approximation. Now take a series expansion:
 \begin{align}\begin{split} % too long; use split environment, and align to a space (&{})
-  \left\langle V_N^{'}(\lambda) \right\rangle_\lambda = &{} \left\langle V_N^{'}(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda^2)
+  \left\langle W_N(\lambda) \right\rangle_\lambda = &{} \left\langle W_N(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda^2)
 \end{split}\end{align}
@@
 \begin{align}
-  \frac{\partial}{\partial\lambda}\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda} &= \left\langle V_N^{''} \right\rangle_\lambda - \beta\sigma_\lambda^2
+  \frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda} &= \left\langle V_N^{''} \right\rangle_\lambda - \beta\sigma_\lambda^2
 \end{align}
@@
 \begin{align}
-  \sigma_\lambda^2 &\equiv \left\langle \left[ V_N^{'}(\lambda) \right]^2 \right\rangle_\lambda - \left\langle V_N^{'}(\lambda) \right\rangle^2_\lambda
+  \sigma_\lambda^2 &\equiv \left\langle \left[ W_N(\lambda) \right]^2 \right\rangle_\lambda - \left\langle W_N(\lambda) \right\rangle^2_\lambda
 \end{align}
-\fixme{I don't really understand how this derivative is \ldots derived}
+Now note that $W_N$ has no dependence on $\lambda$, therefore $W_N' = 0$, and
+\begin{align}
+  \frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda} &= - \beta\sigma_\lambda^2 \\
+  \therefore \left\langle W_N(\lambda) \right\rangle_\lambda &= \left\langle W_N(\lambda)\right\rangle_{\lambda = 0} - \lambda\beta\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
+\end{align}
+
+\davidsays{The derivative is (pretty) easy if you write your ensemble
+  average as a multidimensional integral, as in
+  Eq.~\ref{eqn:dFdlambda}.  Write down $dF/d\lambda$ in terms of $Z_N$
+  and an integral of Boltzmann factors times $W_N$.  Then when you
+  take a derivative, it hits top and bottom, but each is an integral
+  of an exponential, so those are easy to differentiate.  A little
+  math will give you this variance formula.}
+
+\fixme{Below this, I have not updated}
+
 Putting this together, we have
@@
   \begin{split}
-    \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
+    \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle W_N(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
   \end{split} \\
   \begin{split}
-    =&{}  \beta F_0 + \beta\left\langle V_N^{'}(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle V_N^{''}(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
+    =&{}  \beta F_0 + \beta\left\langle W_N(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
   \end{split}
@@
 \begin{align}
-  V_N^{'} &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \nonumber \\
+  W_N &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \nonumber \\
   &= W_N \\
-  \therefore V_N^{''} &= 0
+  \therefore W_N' &= 0
 \end{align}
-With this in mind, we can simplify equation~\ref{eqn:complicated-HTE} to the \emph{high-temperature expansion}, first derived by Zwanzig:\cite{Zwanzig54}
+\davidsays{Ah ha! Here you do what I mentioned earlier.  I'd tend to
+  put this earlier, so you can throw away the $V_N''$ term sooner.
+  And also talking about $W_N$ with this definition I would find nicer
+  and clearer than all the $V_N'$.}  With this in mind, we can
+simplify equation~\ref{eqn:complicated-HTE} to the
+\emph{high-temperature expansion}, first derived by
+Zwanzig:\cite{Zwanzig54}
 \begin{align}
@@
 
+\davidsays{ you can say something to the effect
+  of ``in the next section we will see a simple approximation for this
+  term.''  Hansen's explanation of TPT is not very clear, which makes
+  it all the more valuable to make \emph{this} a clear and lucid
+  explanation.  I'd love to see just a couple more steps at key
+  places---largely where you were confused. This
+  has been an excellent first shot, though! }
+
+\subsection{Square well liquid free enrgy}\label{subsec:SW}
+For the baseline, this work is based on the square well liquid free energy:\cite{Hughes13}
+\begin{align}
+  f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
+\end{align}
+The long-range attractive term, $\fattr$ is given by
+\begin{align}
+  \fattr &= n a_1(n) \ .
+\end{align}
+$\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere
+repulsion, and $\fdisp$ is the attraction. The function $a_1(n)$ is
+the first term from a high-temperature perturbation expansion,
+described in further detail in sections~\ref{sub2sec:disp}
+and~\ref{subsec:TPT}. Each function is outlined below.
+
+\subsubsection{Ideal gas}\label{sub2sec:ID}
+The ideal gas term is given as
+\begin{align}
+  \fid &= n\kT\left(\log(n) - 1\right)
+\end{align}
+
+\subsubsection{Hard sphere repulsion}\label{sub2sec:HS}
+Hard sphere repulsive forces are dealt with by the White Bear version
+of Fundamental-Mearuse Theory.\cite{Roth02} The hard sphere excess
+free energy is given in the homogeneous case as
+\begin{align}
+  \fhs &= \kT\left( \Phi_1(n) + \Phi_2(n) + \Phi_3(n) \right)
+\end{align}
+The $\Phi_j(n)$ terms are given as
+\begin{align}
+  \Phi_1(n) &= -n_0(n)\ln(1 - n_3(n)) \\
+  \Phi_2(n) &= \frac{n_1(n)n_2(n)}{1 - n_3(n)} \\
+  \Phi_3(n) &= n_2^3(n)\left( \frac{n_3(n) + (1 - n_3(n))^2\ln(1 - n_3(n))}{36\pi n_3^2(n)(1 - n_3(n))^2} \right)
+\end{align}
+The \emph{fundamental measure densities} are
+\begin{align}
+  n_0(n) &= \frac{n_2}{4\pi R^2} \nonumber \\
+  &= n \\
+  n_1(n) &= \frac{n_2}{4\pi R} \nonumber \\
+  &= nR \\
+  n_2(n) &= n4\pi R^2 \\
+  n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
+  &= \eta
+\end{align}
+where $\eta$ is the packing fraction.
+
+\subsubsection{Attraction}\label{sub2sec:disp}
+The attractive free energy includes van der Waals attraction between
+hard spheres and is founded in Thermodynamic Perturbation Theory
+(TPT).
+\begin{align}
+  \fdisp &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
+\end{align}
+The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}.
+
+The first term, $a_1$, is the mean-field dispersion interaction. The
+second term, $a_2$, describes the effect of fluctuations resulting
+from compression of the fluid due to the dispersion interaction
+itself, and is approximated using the local compressibility
+approximation (LCA), which assumes the energy fluctuation is simply
+related to the compressibility of a hard-sphere reference
+fluid.\cite{Barker76}
+
+In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
+\textit{monomer} contribution, expressed in terms of energy densities
+as:
+\begin{align}
+  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+\end{align}
+with $\alpha^\text{VDW}$ given by
+\begin{align}
+  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
+  \intertext{using reduced units of $x = r/\sigma$, we have}
+  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
+  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
+\end{align}
+where $b^\text{VDW}$ is the van der Waals size parameter. It
+corresponds to the volume excluded by two spherical particles of
+volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
+
+Then, the high-temp expansion is an expansion of the monomer term:
+\begin{align}
+  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
+\end{align}
+
+The term $a_1$ is given by
+\begin{align}
+  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+\end{align}
+If we assume random correlations between the particles' positions, for
+all distances, we have $g^\text{HS}(r) = 1$. This yields
+\begin{align}
+  a_1^\text{VDW} = - n \alpha^\text{VDW},
+\end{align}
+the van der Waals mean-field energy.
+
+The second-order term is quite a bit more complicated; one must have
+knowledge of higher-order correlation functions to determine
+$a_2$. This term describes the response of the attractive energy due
+to the compression of the fluid from the attractive well. Gil-Villegas
+\emph{et al.} base their expression on an approximation by Barker and
+Henderson.\cite{Barker67} Barker and Henderson's approximation
+considers fluctuations in the number of particles in the potential
+well. In this way, the fluctuations in $a_2$ are related to the
+pressure and compressibility of the liquid. Given pressure
+$P^\text{HS}$ and isothermal compressibility $K^\text{HS} =
+kT\left(\partial n /\partial P^\text{HS}\right)_T$, we have:
+\begin{align}
+  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
+  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+\end{align}
+
+We use the isothermal compressibility in terms of packing fraction
+$\eta$, given by the Percus-Yevick exprssion:\cite{Barker76}
+\begin{align}
+  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
+\end{align}
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The grand free energy density generally has two distinct minima when
-plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an example.) The liquid will be in liquid-vapor equilibrium
-when those minima have the same value for $\phi(T,n)$. There is a local
-maximum between those minima, and the value of the density at that
-maximum is used for $\npart$. Find this value at a given temperature
-then plot $\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new
-value for $\npart$ such that the two minima have the same value for
+plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an
+example.) The liquid will be in liquid-vapor equilibrium when those
+minima have the same value for $\phi(T,n)$. There is a local maximum
+between those minima, and the value of the density at that maximum is
+used for $\npart$. Find this value at a given temperature then plot
+$\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new value
+for $\npart$ such that the two minima have the same value for
 $\phi(T,n)$. (It is helpful to note that if $\npart$ is increased,
-$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$ is
-decresed.) The program I wrote uses the previous
-temperature's $\npart$ as a ``first guess,'' maximizing
-$\phi(T,\npart)$. From there, the program adjusts $\npart$
-until $\phi(T,\nvap) = \phi(T,\nliq)$. When that condition is met, it
-uses this new value of $\npart$ as the ``first guess'' at $\npart$ for the next
-higher temperature.
+$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$
+is decresed.) The program I wrote uses the previous temperature's
+$\npart$ as a ``first guess,'' maximizing $\phi(T,\npart)$. From
+there, the program adjusts $\npart$ until $\phi(T,\nvap) =
+\phi(T,\nliq)$. When that condition is met, it uses this new value of
+$\npart$ as the ``first guess'' at $\npart$ for the next higher
+temperature.
 
-The algorithm is as follows; start at some temperature $T=T_{start}$, with an intial guess for $\npart$.
+The algorithm is as follows; start at some temperature $T=T_{start}$,
+with an intial guess for $\npart$.
 \begin{enumerate}
@@
 \subsection{RG partition functions}\label{subsec:fbar-ubar}
-I use the midpoint method for evaluating the integrals of $\bar{f}_D$ and $\bar{u}_D$.
+I use the midpoint method for evaluating the integrals of $\bar{f}_D$
+and $\bar{u}_D$.
 
@@
 \section{Results and discussion}
+
+For this work, I use $\lambda_\text{SW} = 1.5$.
 
@@
 
-Figures~\ref{fig:phi-RG-lowT} and~\ref{fig:phi-RG-highT} show the grand free energy per unit volume at a low temperature, and near the critical point. It is clear that the program is not de-bugged.
+Figures~\ref{fig:phi-RG-lowT} and~\ref{fig:phi-RG-highT} show the
+grand free energy per unit volume at a low temperature, and near the
+critical point. It is clear that the program is not de-bugged.
 
@@
 \section{Conclusion}
-I have implemented a the use of renormalization group theory in calculations for the square well liquid.
-The program is not entirely de-bugged.
-I have also developed an algorithm that calculates liqiud-vapor coexistence curves, even close to the critical point.
+I have implemented a the use of renormalization group theory in
+calculations for the square well liquid. The program is not entirely
+de-bugged. I have also developed an algorithm that calculates
+liqiud-vapor coexistence curves, even close to the critical point.
 

### 2014-05-27 papers/thesis-roth/project.tex

@@
 average of $W_N$, with the subscript indicating some given
-value of $\lambda$. We can simplify this even further by noting
+value of $\lambda$.
+
+\fixme{The following could probably go someplace else}
+We can simplify this even further by noting
 \begin{align}
@@
 \end{align}
+\fixme{(end)}
 
-At this point, note that we have an exact equation with no approximation. Now take a series expansion:
-\begin{align}\begin{split} % too long; use split environment, and align to a space (&{})
-  \left\langle W_N(\lambda) \right\rangle_\lambda = &{} \left\langle W_N(\lambda)\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda}\bigg|_{\lambda = 0} \\ &{} + \mathcal{O}(\lambda^2)
-\end{split}\end{align}
+We can integrate this expression over $F\in[F_0,F]$ and $\lambda\in[0,1]$ to obtain
+\begin{align}
+  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda \nonumber \\
+  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda \label{eqn:F-inTermsW}
+\end{align}
+At this point, note that we have an exact equation with no approximation. All we have done was take a derivative, simplify, and integrate. Now take a series expansion:
+\begin{align}
+  \left\langle W_N \right\rangle_\lambda = &{} \left\langle W_N\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda}\bigg|_{\lambda = 0} \nonumber\\
+  &{} + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion}
+\end{align}
 The derivative term is
 \begin{align}
-  \frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda} &= \left\langle V_N^{''} \right\rangle_\lambda - \beta\sigma_\lambda^2
+  \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} =&{} \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N}\right) \\
+  =&{} \frac{1}{Z_N}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta V_N} \right) \nonumber\\
+  &{} - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\frac{\partial}{\partial\lambda}Z_N \\
+  =&{} \frac{1}{Z_N}\int d\rr^N\, \left( e^{-\beta V_N}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta V_N} \right) \nonumber \\
+  &{} - \frac{1}{Z_N^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N} \\
 \end{align}
-where
+Note that $\partial/\partial\lambda W_N = 0$, so we have
 \begin{align}
-  \sigma_\lambda^2 &\equiv \left\langle \left[ W_N(\lambda) \right]^2 \right\rangle_\lambda - \left\langle W_N(\lambda) \right\rangle^2_\lambda
+   \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} =&{} \frac{1}{Z_N}\int d\rr^N (-\beta) W_N^2 e^{-\beta V_N} \nonumber\\
+   &{} - \frac{1}{Z_N^2}\Bigg(\int d\rr^N\, W_N e^{-\beta V_N} \nonumber\\
+   &{} \times \int d\rr^N\, (-\beta)W_Ne^{-\beta V_N}\Bigg) \\
+   =&{} -\beta\Bigg[ \frac{1}{Z_N}\int d\rr^N\, W_N^2e^{-\beta V_N}  \nonumber\\
+   &{} - \left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N} \right)^2 \Bigg] \\
+   =&{} -\beta\left[ \left\langle W_N^2 \right\rangle_\lambda - \left\langle W_N \right\rangle_\lambda^2 \right] \\
+   \equiv &{} -\beta\sigma_\lambda^2
 \end{align}
-Now note that $W_N$ has no dependence on $\lambda$, therefore $W_N' = 0$, and
+Combining this result with equation~\ref{eqn:Wn-expansion} gives us
 \begin{align}
-  \frac{\partial}{\partial\lambda}\left\langle W_N(\lambda) \right\rangle_{\lambda} &= - \beta\sigma_\lambda^2 \\
-  \therefore \left\langle W_N(\lambda) \right\rangle_\lambda &= \left\langle W_N(\lambda)\right\rangle_{\lambda = 0} - \lambda\beta\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
+  \left\langle W_N \right\rangle_\lambda &= \left\langle W_N\right\rangle_{\lambda = 0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion-simplified}
 \end{align}
-
-\davidsays{The derivative is (pretty) easy if you write your ensemble
-  average as a multidimensional integral, as in
-  Eq.~\ref{eqn:dFdlambda}.  Write down $dF/d\lambda$ in terms of $Z_N$
-  and an integral of Boltzmann factors times $W_N$.  Then when you
-  take a derivative, it hits top and bottom, but each is an integral
-  of an exponential, so those are easy to differentiate.  A little
-  math will give you this variance formula.}
-
-\fixme{Below this, I have not updated}
-
-Putting this together, we have
+Putting together equations~\ref{eqn:F-inTermsW} and~\ref{eqn:Wn-expansion-simplified}, we have
 \begin{align}
-  \begin{split}
-    \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle W_N(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
-  \end{split} \\
-  \begin{split}
-    =&{}  \beta F_0 + \beta\left\langle W_N(\lambda) \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
-  \end{split}
+  F &= F_0 + \left\langle W_N \right\rangle_{\lambda=0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
 \end{align}
-Now, we can define the total perturbation energy as
-\begin{align}
-  W_N &= \sum_{i=1}^N\sum_{j>i}^N w_\lambda(\mathbf{r}_i,\mathbf{r}_j)
-\end{align}
-Then, following from equation~\ref{eqn:small-perturbation},
-\begin{align}
-  W_N &= \frac{\partial}{\partial\lambda}\left(\sum_{i=1}^N\sum_{j>i}^N \lambda w(\mathbf{r}_i,\mathbf{r}_j)\right) \nonumber \\
-  &= W_N \\
-  \therefore W_N' &= 0
-\end{align}
-\davidsays{Ah ha! Here you do what I mentioned earlier.  I'd tend to
-  put this earlier, so you can throw away the $V_N''$ term sooner.
-  And also talking about $W_N$ with this definition I would find nicer
-  and clearer than all the $V_N'$.}  With this in mind, we can
-simplify equation~\ref{eqn:complicated-HTE} to the
-\emph{high-temperature expansion}, first derived by
+%% \begin{align}
+%%   \begin{split}
+%%     \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle W_N \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
+%%   \end{split} \\
+%%   \begin{split}
+%%     =&{}  \beta F_0 + \beta\left\langle W_N \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
+%%   \end{split}
+%% \end{align}
+This is the \emph{high-temperature expansion}, first derived by
 Zwanzig:\cite{Zwanzig54}
+\fixme{Zwanzig has a factor of $1/2$ that I'm missing}
+\fixme{Below this point has not been updated}
 \begin{align}

### 2014-05-27 papers/square-well-liquid/paper.tex
Added Barker and Henderson citations for their square well work.  Added parts of derivations we have done but commented them out pending better organization. Tested on knightley. Took 1:06.49.
Added some perturbation theory stuff to square well paper Tested on knightley. Took 0:28.95.

@@
 
+%%Stuff commented out below was me typing out this section before I
+%%realised we had something like this.  I'm leaving the code here for now.
+
+%% If we take into account
+%% the $\beta$ term then $a_2$ becomes
+%% \begin{align}\label{eq:A2-exact}
+%%   A_2 &= \int \int \textrm{d$\rr_1$
+%%     d$\rr_2$}n(\rr_1)n(\rr_2)g^{(2)}(\rr_1\,\rr_2)\Phi(r_{12})
+%% \end{align}
+%% We want to use our approximation for $g^{(2)}$
+%% written as Eqn. \ref{eq:g2-our-mean}:
+%% \begin{align}
+%%   g^{(2)}(\rr_1,\rr_2) \approx 
+%%   \frac{g_s \left(\rr_{12},g_{\sigma}(\rr_1)\right) + g_s
+%%     \left(\rr_{12},g_{\sigma}(\rr_2)\right)}{2}.
+%% \end{align}
+%% If the look at the homogeneous case then
+%% \begin{align}
+%%   g_s \left(\rr_{12},g_{\sigma}(\rr_1)\right) = g_s
+%%   \left(\rr_{21},g_{\sigma}(\rr_2)\right).
+%% \end{align}
+%% With which we can rewrite equation \ref{eq:A1-exact} to be  
+%% \begin{align}
+%%   A_2 &= \int \int \textrm{d$\rr_1$
+%%     d$\rr_2$}n(\rr_1)n(\rr_2)g_s(\rr_{12}\,g_\sigma(\rr_1))\Phi(r_{12}).
+%% \end{align}
+%% We can write $g_s$ as a sum of functions
+%% \begin{align}\
+%%   g_s(\rr_{12}\,g_\sigma(\rr_1)) = \sum_i f_i(r_{12})G_i(g_\sigma)
+%% \end{align}
+%% where the ith function we assume is a product of two functions that explicitly
+%% depend on $r_{12}$ and $g_\sigma(\rr_1)$ respectively.
+
+\subsection{Local-compressibility approximation}
+
+Barker and Henderson presented two approximations for the second order
+term in the inverse temperature expansion, one of which they name the
+local-compressiblitity
+approximation. \cite{barker2004perturbationSW,barker2004perturbationII}
+\fixme{below is some perturbation theory stuff}
+We start with the definition for the Helmholtz free energy of the
+reference system and add a correction to the energy due to a perturbation correction:
+\begin{align}
+  A &= A_{ref}+A_a \\
+  &=-kT \ln \left( \sum_i e^{-\beta [E_{ref}(i) + \lambda E_a(i)]}
+  \right).
+\end{align}
+Take the derivative at $\lambda = 0$:
+\begin{align}
+  \frac{\partial A}{\partial \lambda} \bigg{|}_{\lambda = 0} &= -kT
+  \frac{\sum_i -\beta E_a(i) e^{-\beta [E_{ref}(i) + \lambda E_a(i)]}}{\sum_i e^{-\beta
+      [E_{ref}(i) + \lambda E_a(i)]}} \\
+  &= \frac{\sum_i E_a(i) e^{-\beta [E_{ref}(i) + \lambda E_a(i)]}}{\sum_i e^{-\beta
+      [E_{ref}(i) + \lambda E_a(i)]}} \\
+  &= \sum_i E_{a,i}{P_i}|_{\lambda = 0}.
+\end{align}
+Then we take the second derivative;
+\begin{align}
+  \frac{\partial^2 A}{\partial \lambda^2} \bigg{|}_{\lambda = 0} &=
+  \sum_i E_{a,i} \frac{\partial P_i}{\partial
+    \lambda}\bigg{|}_{\lambda = 0} \\
+  &= \sum_i E_{a,i}\Bigg[ \frac{-\beta E_a(i)e^{-\beta [E_{ref}(i) +
+        \lambda E_a(i)]}}{\sum_j e^{-\beta [E_{ref}(j) + \lambda E_a(j)]}}
+  \notag \\
+  &- \frac{e^{-\beta [E_{ref}(i) + \lambda
+        E_a(i)]}\sum_k -\beta E_a(k)e^{-\beta [E_{ref}(k) + \lambda
+        E_a(k)]}}{\left( \sum_j e^{-\beta [E_{ref}(j) + \lambda
+        E_a(j)]} \right)^2} \Bigg] \\
+  &= \sum_i E_a(i)\bigg[ -\beta E_a(i)P_i + \beta P_i \sum_j
+  E_a(j)P_j \bigg] \\
+  &= -\beta \sum_i E_a^2(i)P_i + \beta \sum_i P_i E_a(i)\sum_j E_a(j)
+  P_j \\
+  &= -\beta \sum_i E_a^2(i)P_i + \beta \bigg(\sum_i E_a(i)P_i\bigg)^2
+\end{align}
+Since we set $\lambda = 0$, the probabilities are over the reference state so these become
+averages over those states.
+\begin{align}
+  \frac{\partial^2 A}{\partial \lambda^2} \bigg{|}_{\lambda = 0} &=
+  -\beta \left( \left< E_a^2 \right>_0 - {\langle E_a \rangle}_0^2 \right)
+\end{align}
+%% Our derivation for this looks something like this:
+%% \begin{align}
+%%   A_2 &\approx -\left(\frac{\partial n}{\partial p}\right)
+%%   \int\textrm{d$\rr'$} \notag \\
+%%   &\times \frac{\delta}{\delta n(\rr')} \int \int
+%%   \textrm{d$\rr_1$ d$\rr_2$} \left[n(\rr_1)
+%%   n(\rr_2)g^{(2)}(\rr_1\,\rr_2)\Phi(r_{12})^2 \right]
+%% \end{align}
+%% \begin{align}
+%%   A-A_{ref} &= -kT \ln \left(\frac{Z_a}{Z_{ref}}\right) 
+%%   \\ &= -kT \ln \left( \frac{\sum_i e^{-\beta[E_{ref}(i) + \lambda
+%%         E_a(i)]}} {\sum_i e^{-\beta E_{ref}(i)}} \right) \\
+%% \end{align}
+%% Since our perturbed Energy $E_a$ is caused by some small potential
+%% $u_a$, we can expand $\exp(-\beta \lambda u_a)$ in a power series.
+%% This makes
+%% \begin{align}
+%%   A-A_{ref} &= -kT \ln \bigg{\langle} 1 - \beta
+%%   \lambda u_a + \frac{1}{2} \beta^2 \lambda^2 u_a^2 + \cdots \bigg{\rangle}_0
+%% \end{align}
+%% Here, the average is over the reference energies.  Now we can expand
+%% the the natural log in a series
+%% \begin{align}
+%%   A-A_{ref} &= -kT \ln \bigg{\langle} 1 - \beta
+%%   \lambda u_a + \frac{1}{2} \beta^2 \lambda^2 u_a^2 + \cdots \bigg{\rangle}_0
+%% \end{align}
+
 \subsection{Using the pair distribution function}
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
-  \caption{A simple DOS plot!}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N200-E.pdf}
+%%   \caption{A simple DOS plot!}
+%% \end{figure}
 

### 2014-05-28 papers/thesis-roth/project.tex
project report Tested on elliot. Took 0:19.16.

@@
 
-Near the critical point of a liquid, there are density \fixme{is
-  saying $g(\rr)$ more accurate?} fluctuations at all length
-scales. In this regime, mean field theories break down, for obvious
-reasons. As one approaches the critical point, there becomes less and
-less of a difference in free energy of the liquid and vapor states;
-this causes difficulty in modeling liquid-vapor coexistence. This work
-develops a method for calculating liquid-vapor coexistence, and
-applies renormalization group theory to the square well liquid.
+Near the critical point of a liquid, there are density fluctuations at
+all length scales. In this regime, mean field theories break down. As
+the critical point is approached, there becomes less and less of a
+difference in free energy of the liquid and vapor states; this causes
+difficulty in modeling liquid-vapor coexistence. This work develops a
+method for calculating liquid-vapor coexistence, and applies
+renormalization group theory to the square well liquid.
 
@@
 \section{Introduction}
-\fixme{Write introduction; why do we do this? why is it useful? how is this relevant? what is the history?}
+\fixme{Some thoughts:
+  \begin{itemize}
+    \item Liquids are essential for life
+    \item Rust, aqueous environment for biology, etc
+    \item Critical point is important for many industries
+    \item Square well liquid is a good simple model for liquids
+    \item Include lots (or at least plenty) of references
+  \end{itemize}
+}
+
+The square well liquid is an approximation to simple liquids that is widely used.
 
@@
 
-\fixme{Discuss Wilson phase cell}
+We start by dividing the total volume into cells of volume $V_i$. We
+assume that within a cell $n(\rr)$ varies only slightly, and that the
+cells do not interact with one another. This is refered to as Wilson's
+phase cell method.\cite{Ramana12}
 
-We begin with a baseline free enrgy $f_0(T,n)$. We assume this
+Then consider a baseline free enrgy $f_0(T,n)$. We assume this
 baseline accounts for all fluctuations of wavelength $\lambda <
@@
 accounts for fluctuations with a wavelength $\lambda =
-2\lambda_0$. The new free energy is $f_1(T,n) = f_0(T,n) + \delta
-f_1(T,n)$. Then double the length scale again; $\delta f_2(T,n)$
-accounts for fluctionations with $\lambda = 4\lambda_0$ and $f_2(T,n)
-= f_1(T,n) + \delta f_2(T,n)$. As we continue this process, the
-$i^{th}$ correction accounts for fluctuations with $\lambda =
-2^i\lambda_0$, and $f_i(T,n) = f_{i-1}(T,n) + \delta f_i(T,n)$. In the
-end, your free energy is
+2\lambda_0$. The new free energy is
+\begin{align}
+  f_1(T,n) &= f_0(T,n) + \delta f_1(T,n)
+\end{align}
+Then double the length scale again; $\delta f_2(T,n)$ accounts for
+fluctionations with $\lambda = 4\lambda_0$ and
+\begin{align}
+  f_2(T,n) &= f_1(T,n) + \delta f_2(T,n)
+\end{align}
+As we continue this process, the $i^{th}$ correction accounts for
+fluctuations with $\lambda = 2^i\lambda_0$, and
+\begin{align}
+  f_i(T,n) &= f_{i-1}(T,n) + \delta f_i(T,n)
+\end{align}
+In the end, your free energy is
 \begin{align}
@@
 \end{align}
-\fixme{When $v_\lambda$ is derived below, the $w$
-should be the attraction portion (i.e.~the $-\varepsilon$ between $\sigma$ and $\lambda\sigma$), and $v_{HS}$ should
-just be the hard-sphere potential (without attraction).  That's how
-the TPT works, it is the attraction that is being added in with the
-$\lambda$ proportionality.}
 
-Now, consider this to be the reference, and add a small perturbation:
+We consider this as our referecne, and take a square well attraction as a perturbation:
 \begin{align}
-  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambda w(\1,\2) \label{eqn:small-perturbation}
+  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambda w^\text{SW}(\1,\2) \label{eqn:small-perturbation}
+\end{align}
+where
+\begin{align}
+  w^\text{SW}(\1,\2) &= - \varepsilon \text{ for } \sigma < r < \lambda\sigma
 \end{align}
@@
   V_N(\lambda) &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
-  \hookrightarrow \frac{\partial V_N(\lambda)}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
+  \hookrightarrow \frac{\partial V_N(\lambda)}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w^\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
   &\equiv W_N \\
@@
   &= \frac{1}{Z_N} \int d\rr^N\, W_N e^{-\beta V_N(\lambda)}\\
-  &= \left\langle W_N \right\rangle_\lambda
+  &= \left\langle W_N \right\rangle_\lambda \label{eqn:dfdlambda}
 \end{align}
@@
 
-\fixme{The following could probably go someplace else}
+\fixme{The following could probably go someplace else (is it even necessary?):
   We can simplify this even further by noting
@@
     \frac{\partial F}{\partial\lambda} &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle_\lambda
-\end{align}
-\fixme{(end)}
+\end{align}}
 
-We can integrate this expression over $F\in[F_0,F]$ and $\lambda\in[0,1]$ to obtain
+Note that equation~\ref{eqn:dfdlambda} is exact; now take a series expansion:
 \begin{align}
-  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda \nonumber \\
-  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda \label{eqn:F-inTermsW}
-\end{align}
-At this point, note that we have an exact equation with no approximation. All we have done was take a derivative, simplify, and integrate. Now take a series expansion:
-\begin{align}
-  \left\langle W_N \right\rangle_\lambda = &{} \left\langle W_N\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda}\bigg|_{\lambda = 0} \nonumber\\
-  &{} + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion}
+  \left\langle W_N \right\rangle_\lambda &= \left\langle W_N\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion}
 \end{align}
 The derivative term is
+\begin{widetext}
   \begin{align}
-  \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} =&{} \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N}\right) \\
-  =&{} \frac{1}{Z_N}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta V_N} \right) \nonumber\\
-  &{} - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\frac{\partial}{\partial\lambda}Z_N \\
-  =&{} \frac{1}{Z_N}\int d\rr^N\, \left( e^{-\beta V_N}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta V_N} \right) \nonumber \\
-  &{} - \frac{1}{Z_N^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N} \\
+    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N}\right) \\
+    &= \frac{1}{Z_N}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta V_N} \right) - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\frac{\partial}{\partial\lambda}Z_N \\
+    &= \frac{1}{Z_N}\int d\rr^N\, \left( e^{-\beta V_N}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta V_N} \right) - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N} \\
   \end{align}
-Note that $\partial/\partial\lambda W_N = 0$, so we have
+\end{widetext}
+
+Noting that $\partial W_N/\partial\lambda = 0$, so we have
+\begin{widetext}
   \begin{align}
-   \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} =&{} \frac{1}{Z_N}\int d\rr^N (-\beta) W_N^2 e^{-\beta V_N} \nonumber\\
-   &{} - \frac{1}{Z_N^2}\Bigg(\int d\rr^N\, W_N e^{-\beta V_N} \nonumber\\
-   &{} \times \int d\rr^N\, (-\beta)W_Ne^{-\beta V_N}\Bigg) \\
-   =&{} -\beta\Bigg[ \frac{1}{Z_N}\int d\rr^N\, W_N^2e^{-\beta V_N}  \nonumber\\
-   &{} - \left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N} \right)^2 \Bigg] \\
-   =&{} -\beta\left[ \left\langle W_N^2 \right\rangle_\lambda - \left\langle W_N \right\rangle_\lambda^2 \right] \\
-   \equiv &{} -\beta\sigma_\lambda^2
+    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{1}{Z_N}\int d\rr^N (-\beta) W_N^2 e^{-\beta V_N} - \frac{1}{Z_N^2}\left(\int d\rr^N\, W_N e^{-\beta V_N} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta V_N} \right) \\
+    &= -\beta\left[ \frac{1}{Z_N}\int d\rr^N\, W_N^2e^{-\beta V_N} - \left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N} \right)^2 \right] \\
+    &= -\beta\left[ \left\langle W_N^2 \right\rangle_\lambda - \left\langle W_N \right\rangle_\lambda^2 \right] \\
+    &\equiv -\beta\sigma_\lambda^2
   \end{align}
+\end{widetext}
+
 Combining this result with equation~\ref{eqn:Wn-expansion} gives us
@@
 \end{align}
-Putting together equations~\ref{eqn:F-inTermsW} and~\ref{eqn:Wn-expansion-simplified}, we have
+and we can insert this into equation~\ref{eqn:dfdlambda}:
 \begin{align}
-  F &= F_0 + \left\langle W_N \right\rangle_{\lambda=0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
+  \frac{\partial F}{\partial\lambda} &= \left\langle W_N\right\rangle_{\lambda = 0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
 \end{align}
+
+We can integrate this expression over $F\in[F_0,F]$ and $\lambda\in[0,1]$ to obtain
+\begin{align}
+  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambda=0}^2 \nonumber \\
+  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta\sigma_{\lambda=0}^2 \label{eqn:F-inTermsW}
+\end{align}
+Multiply by $\beta$ and we have the \emph{high-temperature expansion}, first derived by
+Zwanzig:\cite{Zwanzig54}
+\begin{align}
+  \beta F &= \beta F_0 + \beta \left\langle W_N \right\rangle_{\lambda=0} - \beta^2\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^3)
+\end{align}
+
+Zwanzig further showed that the $n^{th}$ term could be written in terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle W_N\right\rangle_{\lambda=0} \right]^\nu \right\rangle_{\lambda=0}$ with $\nu \leq n$.
+This requires us to know the form of higher-order correlation functions of the reference system, which can be quite complicated.
+In the next section we will see a simple aproximation for the $n=2$ term.
+
+%% Writing the statistical averages in terms of the pair density
+%% $n_\lambda^{(2)}(\1,\2)$ (i.e. the probability per volume squared of having one sphere at \1
+%%   and another at \2) gives us
+%% \begin{align}
+%%   \beta F &=\beta F_0 + \frac{1}{2}\beta\int_0^1 d\lambda\, \iint d\1 d\2\, n_\lambda^{(2)}(\1,\2)w(\1,\2) \label{eqn:Henderson5-2-12}
+%% \end{align}
+%% \davidsays{I think this discussion is pretty redundant with
+%%   parts of the previous derivation in terms of $V_N$.}  The pair
+%% density itself can be expanded as
+%% \begin{align}
+%%   n_\lambda^{(2)}(\1,\2) &= n_0^{(2)}(\1,\2) + \lambda\frac{\partial n_\lambda^{(2)}(\1,\2)}{\partial\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2)
+%% \end{align}
+%% Applying this to the 2nd term in equation~\ref{eqn:Henderson5-2-12}, we
+%% have
+%% \begin{align}
+%%   \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
+%%   &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)w(\1,\2)
+%% \end{align}
+%% The second-order $\lambda$ term is quite a bit more complicated:
+%% \begin{widetext}
 %%  \begin{align}
 %%    \begin{split}
-%%     \beta F = &{} \beta F_0 + \beta(1 - 0)\left\langle W_N \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}(1 - 0)^2\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right)
-%%   \end{split} \\
-%%   \begin{split}
-%%     =&{}  \beta F_0 + \beta\left\langle W_N \right\rangle_{\lambda = 0} \\ &{} - \beta\frac{1}{2}\left(\left\langle W_N'(\lambda = 0) \right\rangle_{\lambda = 0} - \beta^2\sigma^2_{\lambda = 0} \right) \label{eqn:complicated-HTE}
+%%    \frac{\beta F_2}{N} =&{} -\frac{\beta^2}{2}\left[ \frac{n}{2}\int d\2\, g_0(\1,\2)w^2(\1,\2) \right.
+\\ &{} + n^2 \iint d\2 d\3\, g_0^{(3)}(\1,\2,\3)w(\1,\2)w(\1,\3) \\ &{} \left.
++ \frac{n^3}{4}\iiint d\2 d\3 d\4\, \left( g_0^{(4)}(\1,\2,\3,\4) - g_0^{(2)}(\1,\2)g_0^{(2)}(\3,\4) \right)w(\1,\2)2(\3,\4) \right] \\ &{} - \frac{1}{4}S_0(0)\left[ \frac{\partial}{\partial n}\left( n^2\int d\2\, g_0(\1,\2)w(\1,\2) \right)^2 \right]
 %%    \end{split}
 %%  \end{align}
-This is the \emph{high-temperature expansion}, first derived by
-Zwanzig:\cite{Zwanzig54}
-\fixme{Zwanzig has a factor of $1/2$ that I'm missing}
-\fixme{Below this point has not been updated}
-\begin{align}
-  \begin{split}
-    \beta F = &{} \beta F_0 + \beta\left\langle W_N \right\rangle_0  \\ &{} - \frac{1}{2}\beta^2\left( \left\langle W_N^2 \right\rangle_0 - \left\langle W_N \right\rangle_0^2 \right) + \mathcal{O}(\beta^3)
-  \end{split}
-\end{align}
-Zwanzig further showed that the $n^{th}$ term could be written in terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle W_N\right\rangle_0 \right]^\nu \right\rangle_0$ with $\nu \leq n$.
-
-Writing the statistical averages in terms of the pair density
-$n_\lambda^{(2)}(\1,\2)$ (i.e. the probability of having a sphere at
-\1 and \2 \fixme{make this more clear?}\davidsays{I'd perhaps phrase
-  it as ``probability per volume squared of having one sphere at \1
-  and another at \2''}) gives us
-\begin{align}
-  \frac{\beta F}{N} &= \frac{\beta F_0}{N} + \frac{\beta}{2N}\int_0^1 d\lambda\, \iint d\1 d\2\, n_\lambda^{(2)}(\1,\2)w(\1,\2) \label{eqn:Henderson5-2-12}
-\end{align}
-\davidsays{Here's another place where all the $N$s on the bottom are
-  muddying things up, I think.  (like the $\beta$s were before).  I
-  also think this discussion is pretty redundant with parts of the
-  previous derivation in terms of $V_N$.}  The pair density itself can
-be expanded as
-\begin{align}
-  n_\lambda^{(2)}(\1,\2) &= n_0^{(2)}(\1,\2) + \lambda\frac{\partial n_\lambda^{(2)}(\1,\2)}{\partial\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2)
-\end{align}
-Applying this to the \fixme{0th, 1st or 2nd, depending on how you
-  think of it}\davidsays{it's ``second'' since ``terms'' are things
-  that are added together, while factors are things that are
-  multiplied together.  And in Eq~\ref{eqn:Henderson5-2-12} there are
-  only two things added together.  (I figured this terminology out on
-  my own some time after becoming a professor... it seems to be used,
-  but not taught).} term in equation~\ref{eqn:Henderson5-2-12}, we
-have
-\begin{align}
-  \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
-  &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)w(\1,\2)
-\end{align}
-The second-order $\lambda$ term is quite a bit more complicated:
-\begin{widetext}
- \begin{align}
-   \begin{split}
-   \frac{\beta F_2}{N} =&{} -\frac{\beta^2}{2}\left[ \frac{n}{2}\int d\2\, g_0(\1,\2)w^2(\1,\2) \right.
-\\ &{} + n^2 \iint d\2 d\3\, g_0^{(3)}(\1,\2,\3)w(\1,\2)w(\1,\3) \\ &{} \left.
-+ \frac{n^3}{4}\iiint d\2 d\3 d\4\, \left( g_0^{(4)}(\1,\2,\3,\4) - g_0^{(2)}(\1,\2)g_0^{(2)}(\3,\4) \right)w(\1,\2)2(\3,\4) \right] \\ &{} - \frac{1}{4}S_0(0)\left[ \frac{\partial}{\partial n}\left( n^2\int d\2\, g_0(\1,\2)w(\1,\2) \right)^2 \right]
-   \end{split}
- \end{align}
-\end{widetext}
-This term requires use to know the form of higher-order correlation functions and the structure function $S_0(k)$ of the reference system.
+%% \end{widetext}
+%% This term requires use to know the form of higher-order correlation functions and the structure function $S_0(k)$ of the reference system.
 These things are very difficult.
@@
 
-\davidsays{ you can say something to the effect
-  of ``in the next section we will see a simple approximation for this
-  term.''  Hansen's explanation of TPT is not very clear, which makes
+\davidsays{Hansen's explanation of TPT is not very clear, which makes
   it all the more valuable to make \emph{this} a clear and lucid
@@
 the first term from a high-temperature perturbation expansion,
-described in further detail in sections~\ref{sub2sec:disp}
-and~\ref{subsec:TPT}. Each function is outlined below.
+described in further detail in sections~\ref{subsec:TPT}
+and~\ref{sub2sec:disp}. Each function is outlined below.
 
@@
 \end{align}
+This is called the Local Compressibility Approximation (LCA).
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2014-05-28 papers/square-well-liquid/histogram-paper.tex
decrease # of spheres yet again (and increase iterations) for "quick" SW MC plots Tested on bennet. Took 3:00.68.

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-E}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-E}
   \caption{A histogram plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-dos}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-weights}
+  \caption{The weight functions from different methods.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-dos}
   \caption{A density of states plot demonstrating the difficulty of using
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-u}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-u}
   \caption{Specific internal energy.\label{fig:u}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-hc}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-hc}
   \caption{Specific heat capacity.\label{fig:hc}}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N100-rdf}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rdf}
   \caption{Radial distribution function.\label{fig:rdf}}

### 2014-05-29 papers/thesis-roth/project.tex
project report Tested on elliot. Took 0:18.92.

@@
 % Free energy terms
-\newcommand{\fid}{\ensuremath{f_\text{ID}(T,n)}}
-\newcommand{\fhs}{\ensuremath{f_\text{HS}(T,n)}}
-\newcommand{\fdisp}{\ensuremath{f_\text{disp}(T,n)}}
-\newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}
+\newcommand{\fid}{\ensuremath{F_\text{ID}(T,n)}}
+\newcommand{\fhs}{\ensuremath{F_\text{HS}(T,n)}}
+\newcommand{\fdisp}{\ensuremath{F_\text{disp}(T,n)}}
+\newcommand{\fattr}{\ensuremath{F_\text{attr}(T,n)}}
 
@@
 \section{Introduction}
-\fixme{Some thoughts:
-  \begin{itemize}
-    \item Liquids are essential for life
-    \item Rust, aqueous environment for biology, etc
-    \item Critical point is important for many industries
-    \item Square well liquid is a good simple model for liquids
-    \item Include lots (or at least plenty) of references
-  \end{itemize}
-}
 
-The square well liquid is an approximation to simple liquids that is widely used.
+Gasses have been accurately modeled since the 19$^{th}$
+century\cite{Lederman92} and the kinetic theory of gasses. With the
+advent of quantum mechanics in the early 20$^{th}$ century, we have
+had a pretty good grasp on solids. A robust model of liquids continues
+to be elusive. Liquids are not periodic structures like solids, but
+neither are they disperse, as are gasses.
+
+However, understanding liquids has both philosophical and economic
+desireability. Liquid water is---as far as we know---essential for
+developing and sustaining life. All celluar processes take place in an
+aqueous environment. Understanding liquids is in many ways a step
+towards understanding the development and existence of life. On the
+economic side of the coin, we recognize liquids to be an important
+part of a wide array of industrial applications.
+
+One particularly troublesome regime is around the so-called ``critical
+point'' of a liquid. At the appropriate density and temperature, the
+distinction between liquid and vapor dissapears. Beyond this point,
+you have a ``supercritical fluid.'' Supercritical fluids are used in
+many industrial processes\cite{Perrut00}. For example (and of personal
+interest), supercritical fluids are used in coffee decaffeination and
+the extraction of hops resins for brewing beer.
+
+The simplest model of a liquid is to consider each molecule to be a
+hard sphere (i.e.~spheres cannot overlap) with a square-well potenial
+between any pair of spheres. This is called the ``square well
+liquid,'' and it is a widely used approximation to simple
+liquids. This is one of the foundations of this work.
+
+We use statistical mechanics to study liquids; this requires large
+sample sizes, long simulation times, or some combination of the
+two. Monte Carlo simulations are exact, if you allow them to run
+``long enough.'' Our models are ultimately compared to these
+simulations, but Monte Carlo simulations are otherwise outside the
+scope of this work. The Roundy group uses classical Density-funtional
+Theory (DFT) to calculate free energies and probability distributions
+of bulk liquids. The end goal of this work is to investigate into
+wheather or not implementing renormalization group theory is
+worthwhile to the Roundy research group.
+
+In section~\ref{sec:theory}, I present some of the main theories I
+deal with: renormalization group theory, thermodynamic perturbation
+theory and the square well liquid free energy. In
+section~\ref{sec:methods} I discuss the the methods used. I cover the
+calculation of liquid-vapor coexistence curves in some detail, as
+caluclating coexistence near the critical point is tricky.
 
@@
 
-This work applies entirely to the homogeneous case---that is, there
-are no walls containing our liquid, nor are there any sort of solutes
-in the liquid. Since our fluid is of infinite extent, the free energy
-is infinite. Therefore, I work with free energy density:
-\begin{align}
-  f(T,n) &\equiv \frac{F(T,n)}{V}
-\end{align}
-
 \subsection{Renormalization Group Theory}\label{subsec:RGT}
-Renormalization group (RG) theory is an iterative procedure that was
+Renormalization group theory (RGT) is an iterative procedure that was
 developed to deal with large-scale fluctuations of the order parameter
@@
 
-Then consider a baseline free enrgy $f_0(T,n)$. We assume this
+Then consider a baseline free enrgy $F_0(T,n)$. We assume this
 baseline accounts for all fluctuations of wavelength $\lambda <
 \lambda_0$. We then double the size of our phase cell and add a
-correction to the free energy $\delta f_1(T,n)$. This correction
+correction to the free energy $\delta F_1(T,n)$. This correction
 accounts for fluctuations with a wavelength $\lambda =
@@
 \begin{align}
-  f_1(T,n) &= f_0(T,n) + \delta f_1(T,n)
+  F_1(T,n) &= F_0(T,n) + \delta F_1(T,n)
 \end{align}
-Then double the length scale again; $\delta f_2(T,n)$ accounts for
+Then double the length scale again; $\delta F_2(T,n)$ accounts for
 fluctionations with $\lambda = 4\lambda_0$ and
 \begin{align}
-  f_2(T,n) &= f_1(T,n) + \delta f_2(T,n)
+  F_2(T,n) &= F_1(T,n) + \delta F_2(T,n)
 \end{align}
@@
 \begin{align}
-  f_i(T,n) &= f_{i-1}(T,n) + \delta f_i(T,n)
+  F_i(T,n) &= F_{i-1}(T,n) + \delta F_i(T,n)
 \end{align}
@@
 \begin{align}
-  f(T,n) = f_0(T,n) + \sum_{i=1}^\infty\delta f_i(T,n) + \fattr
+  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \fattr
 \end{align}
@@
 
-Forte gives $\delta f_i(T,n)$ as
+Forte gives $\delta F_i(T,n)$ as
+\fixme{Talk about ratio of partition functions}
 \begin{align}
-  \delta f_i(T,n) &= -\frac{\kT}{V_D}\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
+  \delta F_i(T,n) &= -\frac{\kT}{V_D}\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
 \end{align}
@@
 \begin{align}
-  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \bar{f}_D(T,n) + \bar{u}_D(T,n \right) \right] \\
-  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\bar{f}_D(T,n) \right]
+  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \bar{F}_D(T,n) + \bar{U}_D(T,n \right) \right] \\
+  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\bar{F}_D(T,n) \right]
 \end{align}
@@
 \begin{align}
-  \bar{f}_D(T,n) &= \frac{f_{i-1}(T,n+x) + f_{i-1}(T,n-x)}{2} - f_{i-1}(T,n) \\
-  \bar{u}_D(T,n) &= \frac{u(T,n+x,\lambda_D) + u(T,n-x,\lambda_D)}{2} - u(T,n,\lambda_D)
+  \bar{F}_D(T,n) &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n) \\
+  \bar{U}_D(T,n) &= \frac{U(T,n+x,\lambda_D) + U(T,n-x,\lambda_D)}{2} - U(T,n,\lambda_D)
 \end{align}
-Finally, $u(T,n)$ is given by
+Finally, $U(T,n)$ is given by
 \begin{align}
-  \begin{split}
-  u(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \\ &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma
-  \end{split}
+  U(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \nonumber \\
+  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma
 \end{align}
@@
 
+This work applies entirely to the homogeneous case---that is, there
+are no walls containing our liquid, nor are there any sort of solutes
+in the liquid. Since our fluid is of infinite extent, the free energy
+is infinite. Therefore, in the calculations, I work with free energy density:
+\begin{align}
+  f(T,n) &\equiv \frac{F(T,n)}{V}
+\end{align}
+
 \subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Results and discussion}
+\section{Results and discussion}\label{sec:results}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Conclusion}\label{sec:conclusion}
 
-\section{Conclusion}
 I have implemented a the use of renormalization group theory in
@@
 
+I conclude that it would be worthwhile to further investigate implementing renormalization group theory in the Roundy DFT.
+
 \bibliographystyle{unsrt}

### 2014-05-30 papers/thesis-roth/project.tex

@@
 % Free energy terms
-\newcommand{\fid}{\ensuremath{F_\text{ID}(T,n)}}
-\newcommand{\fhs}{\ensuremath{F_\text{HS}(T,n)}}
-\newcommand{\fdisp}{\ensuremath{F_\text{disp}(T,n)}}
-\newcommand{\fattr}{\ensuremath{F_\text{attr}(T,n)}}
+\newcommand{\Fattr}{\ensuremath{F_\text{attr}(T,n)}}
+
+\newcommand{\fid}{\ensuremath{f_\text{ID}(T,n)}}
+\newcommand{\fhs}{\ensuremath{f_\text{HS}(T,n)}}
+\newcommand{\fdisp}{\ensuremath{f_\text{disp}(T,n)}}
+\newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}
 
@@
 \newcommand{\4}{\ensuremath{\textbf{r}_4}}
+
+% RG stuff
+\newcommand{\FbarD}{\ensuremath{\bar{F}_D(T,n)}}
+\newcommand{\UbarD}{\ensuremath{\bar{U}_D(T,n)}}
 
@@
 \begin{align}
-  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \fattr
+  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \Fattr
 \end{align}
 
-Where $\fattr$ is the contribution due to the longest-wavelength
+Where $\Fattr$ is the contribution due to the longest-wavelength
 density fluctuations. At this stage, I do not entirely understand why
-Forte splits up $\fattr$.
+Forte splits up $\Fattr$.
 
-Forte gives $\delta F_i(T,n)$ as
-\fixme{Talk about ratio of partition functions}
+The $\detla F_i(T,n)$ term is a difference in free energies:
 \begin{align}
-  \delta F_i(T,n) &= -\frac{\kT}{V_D}\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
+  \delta F_i(T,n) &= -\kT\ln Z - \left( -\kT\ln Z^* \right) \\
+  &= -\kT\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
 \end{align}
-where
+Here, $Z$ is the partition function for the entire system, and $Z^*$ is the partition function of a reference system consisting only of the term $\FbarD$. Forte gives these terms as
 \begin{align}
-  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \bar{F}_D(T,n) + \bar{U}_D(T,n \right) \right] \\
-  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\bar{F}_D(T,n) \right]
+  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \FbarD + \bar{U}_D(T,n \right) \right] \\
+  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\FbarD \right]
 \end{align}
-and
+where $x$ is the amplitude of density fluctuations. $\FbarD$ and $\UbarD$ are formulated similar to averages:
 \begin{align}
-  \bar{F}_D(T,n) &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n) \\
-  \bar{U}_D(T,n) &= \frac{U(T,n+x,\lambda_D) + U(T,n-x,\lambda_D)}{2} - U(T,n,\lambda_D)
+  \FbarD &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n) \\
+  \UbarD &= \frac{U(T,n+x) + U(T,n-x)}{2} - U(T,n)
 \end{align}
-Finally, $U(T,n)$ is given by
+Finally, $U(T,n)$ is given by Forte as
 \begin{align}
   U(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \nonumber \\
-  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma
+  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma \label{eqn:URG} % URG! It sounds like it hurt! :P
 \end{align}
@@
 \end{align}
+The pair correlation function $g(n,r)$ tells us how likely we are to find another hard sphere some distance away, given some density.
+In equation~\ref{eqn:URG} we are considering this quantity at contact ($r = \sigma$) with some effective density.
 
@@
 
-\fixme{And introduction to TPT---why do we use it?}\davidsays{We use
-  TPT because it gives a good description of simple liquids, as long
-  as you are far from the critical point.  Thus it covers most of our
-  bases.  It is also an elegant theory, since it comes from a rigorous
-  power series expansion in a small parameter... or one that one hopes
-  is small. }
-
-This discussion follows primarily from \textit{Theory of Simple
-  Liquids}, by Hansen and McDonald\cite{Hansen06}.
+Thermodynamic perturbation theory (TPT) gives a good description of
+simple liquids, as long as you are far from the critical point---which
+is most of where we calculate. TPT is a rigorous power series
+expansion in a small parameter; as long as the parameter is small, the
+theory is valid. This discussion follows primarily from \textit{Theory
+  of Simple Liquids}, by Hansen and McDonald\cite{Hansen06}.
 
@@
 \begin{align}
-  V_N(\lambda) &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
-  \hookrightarrow \frac{\partial V_N(\lambda)}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w^\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
+  \widetilde{V} &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
+  \hookrightarrow \frac{\partial \widetilde{V}}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w^\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
   &\equiv W_N \\
-  \hookrightarrow \frac{\partial^2 V_N(\lambda)}{\partial\lambda^2} &= 0
+  \hookrightarrow \frac{\partial^2 \widetilde{V}}{\partial\lambda^2} &= 0
 \end{align}
@@
 \begin{align}
-  Z_N(\lambda) &= \int d\rr^N\, e^{-\beta V_N(\lambda)} \\
-  \hookrightarrow \frac{\partial Z_N(\lambda)}{\partial\lambda} &=  \int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N(\lambda)} \\
-  &= \int d\rr^N\, -\beta\frac{\partial V_N(\lambda)}{\partial\lambda}e^{-\beta V_N(\lambda)} \\
-  &= \int d\rr^N\, -\beta W_N e^{-\beta V_N(\lambda)}
+  Z_\lambda &= \int d\rr^N\, e^{-\beta \widetilde{V}(\lambda)} \\
+  \hookrightarrow \frac{\partial Z_\lambda}{\partial\lambda} &=  \int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}(\lambda)} \\
+  &= \int d\rr^N\, -\beta\frac{\partial \widetilde{V}}{\partial\lambda}e^{-\beta \widetilde{V}(\lambda)} \\
+  &= \int d\rr^N\, -\beta W_N e^{-\beta \widetilde{V}(\lambda)}
 \end{align}
@@
 \begin{align}
-  F &= -\kT\ln\left( \frac{Z_N}{V^N} \right) \\
-  \hookrightarrow \frac{\partial F}{\partial\lambda} &= \frac{V^N}{Z_N}\frac{\partial}{\partial\lambda}\left( \frac{Z_N}{V^N} \right) \\
-  &= -\kT\left( \frac{1}{Z_N}\frac{\partial Z_N}{\partial\lambda} \right) \\
-  &= \frac{1}{Z_N} \int d\rr^N\, W_N e^{-\beta V_N(\lambda)}\\
+  F &= -\kT\ln\left( \frac{Z_\lambda}{V^N} \right) \\
+  \hookrightarrow \frac{\partial F}{\partial\lambda} &= \frac{V^N}{Z_\lambda}\frac{\partial}{\partial\lambda}\left( \frac{Z_\lambda}{V^N} \right) \\
+  &= -\kT\left( \frac{1}{Z_\lambda}\frac{\partial Z_\lambda}{\partial\lambda} \right) \\
+  &= \frac{1}{Z_\lambda} \int d\rr^N\, W_N e^{-\beta \widetilde{V}(\lambda)}\\
   &= \left\langle W_N \right\rangle_\lambda \label{eqn:dfdlambda}
@@
   \begin{align}
-    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N}\right) \\
-    &= \frac{1}{Z_N}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta V_N} \right) - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\frac{\partial}{\partial\lambda}Z_N \\
-    &= \frac{1}{Z_N}\int d\rr^N\, \left( e^{-\beta V_N}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta V_N} \right) - \left( \int d\rr^N\, W_N e^{-\beta V_N} \right)\frac{1}{Z_N^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta V_N} \\
+    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}}\right) \\
+    &= \frac{1}{Z_\lambda}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\frac{\partial}{\partial\lambda}Z_\lambda \\
+    &= \frac{1}{Z_\lambda}\int d\rr^N\, \left( e^{-\beta \widetilde{V}}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}} \\
   \end{align}
@@
 
-Noting that $\partial W_N/\partial\lambda = 0$, so we have
+Noting that $\partial W_N/\partial\lambda = 0$, we have
 \begin{widetext}
   \begin{align}
-    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{1}{Z_N}\int d\rr^N (-\beta) W_N^2 e^{-\beta V_N} - \frac{1}{Z_N^2}\left(\int d\rr^N\, W_N e^{-\beta V_N} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta V_N} \right) \\
-    &= -\beta\left[ \frac{1}{Z_N}\int d\rr^N\, W_N^2e^{-\beta V_N} - \left( \frac{1}{Z_N}\int d\rr^N\, W_N e^{-\beta V_N} \right)^2 \right] \\
+    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{1}{Z_\lambda}\int d\rr^N (-\beta) W_N^2 e^{-\beta \widetilde{V}} - \frac{1}{Z_\lambda^2}\left(\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta \widetilde{V}} \right) \\
+    &= -\beta\left[ \frac{1}{Z_\lambda}\int d\rr^N\, W_N^2e^{-\beta \widetilde{V}} - \left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)^2 \right] \\
     &= -\beta\left[ \left\langle W_N^2 \right\rangle_\lambda - \left\langle W_N \right\rangle_\lambda^2 \right] \\
@@
 \begin{align}
-  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambda=0}^2 \nonumber \\
-  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta\sigma_{\lambda=0}^2 \label{eqn:F-inTermsW}
+  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^2) \nonumber \\
+  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW}
 \end{align}
@@
 %% \davidsays{I think this discussion is pretty redundant with
-%%   parts of the previous derivation in terms of $V_N$.}  The pair
+%%   parts of the previous derivation in terms of $\widetilde{V}$.}  The pair
 %% density itself can be expanded as
@@
 Hence, we use approximations, as outlined in section~\ref{sub2sec:disp}.
-
-\davidsays{Hansen's explanation of TPT is not very clear, which makes
-  it all the more valuable to make \emph{this} a clear and lucid
-  explanation.  I'd love to see just a couple more steps at key
-  places---largely where you were confused. This
-  has been an excellent first shot, though! }
 
@@
 \begin{align}
-  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}(r)}{\partial P^\text{HS}}\,dr \\
+  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \\
   &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
@@
 \begin{align}
-  f(T,n) &\equiv \frac{F(T,n)}{V}
+  f(T,n) &\equiv \frac{F}{V}(T,n)
 \end{align}
@@
 \begin{align}
-  \frac{\Phi(T,n)}{V} &= \phi(T,n) \nonumber \\
+  \frac{\Phi}{V}(T,n) &= \phi(T,n) \nonumber \\
                  &= f(T,n) - n\mu \nonumber \\
-                 &= f(T,n) - n\frac{df(T,n)}{dn}\bigg|_{\npart}\ .
+                 &= f(T,n) - n\frac{df}{dn}(T,n)\bigg|_{\npart}\ .
 \end{align}
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
+%%  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
   \caption{Grand free energy per unit volume, SW, low temp}
@@
 \subsection{RG partition functions}\label{subsec:fbar-ubar}
-I use the midpoint method for evaluating the integrals of $\bar{f}_D$
-and $\bar{u}_D$.
+I use the midpoint method for evaluating the integrals of $\FbarD$
+and $\UbarD$.
 
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
+%%  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
+%%  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
+%%  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
   \end{center}

### 2014-05-31 papers/thesis-roth/project.tex
fix typo in project.tex Tested on bennet. Took 0:18.45.
project report Tested on elliot. Took 0:19.23.

@@
 \documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
-\usepackage{graphicx}% Include figure files
+\usepackage[demo]{graphicx}% Include figure files. When building at home, use the option [demo] to supress actual need of figures
 \usepackage{dcolumn}% Align table columns on decimal point
@@
 % special values for n
-\newcommand{\npart}{\ensuremath{n_{particular}}}
+\newcommand{\npart}{\ensuremath{n_{part}}}
 \newcommand{\nliq}{\ensuremath{n_{liquid}}}
@@
 \newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}
+
+% lambda
+\newcommand{\lambdaSW}{\ensuremath{\lambda_\text{SW}}}
+\newcommand{\lambdaRG}{\ensuremath{\lambda_\text{RG}}}
 
@@
 Near the critical point of a liquid, there are density fluctuations at
-all length scales. In this regime, mean field theories break down. As
+all length scales. In this regime, mean field theories break down. \fixme{re-word the next sentence} As
 the critical point is approached, there becomes less and less of a
@@
 Gasses have been accurately modeled since the 19$^{th}$
-century\cite{Lederman92} and the kinetic theory of gasses. With the
+century\cite{Lederman92} and formulation of the kinetic theory of gasses. With the
 advent of quantum mechanics in the early 20$^{th}$ century, we have
@@
 
-One particularly troublesome regime is around the so-called ``critical
-point'' of a liquid. At the appropriate density and temperature, the
+One particularly troublesome regime is around the critical
+point of a liquid. At the appropriate density and temperature, the
 distinction between liquid and vapor dissapears. Beyond this point,
 you have a ``supercritical fluid.'' Supercritical fluids are used in
-many industrial processes\cite{Perrut00}. For example (and of personal
+many industrial processes.\cite{Perrut00} For example (and of personal
 interest), supercritical fluids are used in coffee decaffeination and
 the extraction of hops resins for brewing beer.
+
+It is the aim of renormalization group theory to tackle the critical
+point. The reason that calculations are dificult near the critical
+point is that there are large-scale fluctuations of the density. In
+fact, there are density fluctuations at \emph{all} length
+scales, which is why mean field theories fail. Renormalization group theory is an iterative procedure that
+was developed in th 1950's, and Kenneth Wilson's work with the theory
+in the 1970's won him the 1982 Nobel Prize in physics.\cite{Nobel82}
+
+\fixme{Transition here?}
 
@@
 two. Monte Carlo simulations are exact, if you allow them to run
-``long enough.'' Our models are ultimately compared to these
-simulations, but Monte Carlo simulations are otherwise outside the
-scope of this work. The Roundy group uses classical Density-funtional
+``long enough.'' The Roundy group uses classical Density-funtional
 Theory (DFT) to calculate free energies and probability distributions
-of bulk liquids. The end goal of this work is to investigate into
-wheather or not implementing renormalization group theory is
-worthwhile to the Roundy research group.
+of bulk liquids. These calculations are ultimately compared to Monte Carlo simulations.
+
+The end goal of this work is to investigate into whether or not
+implementing renormalization group theory is worthwhile to the Roundy
+Research Group. Many other groups use renormalization group theory,
+but none of them tell the iteration depth used in their
+calculations.\cite{Forte11,Ramana12,Ghobadi13,Wang13,Zhao13} If only
+few iterations are required, then this would certainly be an effective
+tool to use. However, if many iterations are required, the
+computational cost could be unwieldy.
 
@@
 calculation of liquid-vapor coexistence curves in some detail, as
-caluclating coexistence near the critical point is tricky.
+caluclating coexistence near the critical point is tricky. The
+results---such as they are---are covered in section~\ref{sec:results}.
 
@@
 We start by dividing the total volume into cells of volume $V_i$. We
-assume that within a cell $n(\rr)$ varies only slightly, and that the
-cells do not interact with one another. This is refered to as Wilson's
+assume that within a cell the number density $n(\rr)$ varies only slightly, and that the
+cells do not interact with one another. This comes from Wilson's
 phase cell method.\cite{Ramana12}
 
-Then consider a baseline free enrgy $F_0(T,n)$. We assume this
-baseline accounts for all fluctuations of wavelength $\lambda <
-\lambda_0$. We then double the size of our phase cell and add a
-correction to the free energy $\delta F_1(T,n)$. This correction
-accounts for fluctuations with a wavelength $\lambda =
-2\lambda_0$. The new free energy is
+Then consider a baseline free enrgy $F_0(T,n)$, usually based on a
+mean field theory. We assume this baseline accounts for all
+fluctuations of wavelength $\lambdaRG < \lambda_0$. We then
+double the size of our phase cell and add a correction to the free
+energy $\delta F_1(T,n)$. This correction accounts for fluctuations
+with a wavelength $\lambdaRG = 2\lambda_0$. The new free
+energy, $F_1(T,n)$, is
 \begin{align}
@@
 Then double the length scale again; $\delta F_2(T,n)$ accounts for
-fluctionations with $\lambda = 4\lambda_0$ and
+fluctionations with $\lambdaRG = 4\lambda_0$ and
 \begin{align}
@@
 As we continue this process, the $i^{th}$ correction accounts for
-fluctuations with $\lambda = 2^i\lambda_0$, and
+fluctuations with $\lambdaRG = 2^i\lambda_0$, and
 \begin{align}
@@
 density fluctuations. At this stage, I do not entirely understand why
-Forte splits up $\Fattr$.
+Forte separates out $\Fattr$.
 
-The $\detla F_i(T,n)$ term is a difference in free energies:
+The $\delta F_i(T,n)$ term is a difference in free energies:
 \begin{align}
-  \delta F_i(T,n) &= -\kT\ln Z - \left( -\kT\ln Z^* \right) \\
-  &= -\kT\ln\left[ \frac{Z_D(T,n)}{Z_D^*(T,n} \right]
+  \delta F_i(T,n) &= -\kT\ln Z - \left( -\kT\ln Z^* \right) \nonumber \\
+  &= -\kT\ln\left[ \frac{Z(T,n)}{Z^*(T,n)} \right]
 \end{align}
-Here, $Z$ is the partition function for the entire system, and $Z^*$ is the partition function of a reference system consisting only of the term $\FbarD$. Forte gives these terms as
+Here, $Z$ is the partition function for the entire system, and $Z^*$
+is the partition function of a reference system consisting only of the
+term $\FbarD$. Forte gives these terms as
 \begin{align}
-  Z_D(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\left( \FbarD + \bar{U}_D(T,n \right) \right] \\
-  Z_D^*(T,n) &= \int dx\, \exp\left[ -\frac{V_D}{\kT}\FbarD \right]
+  Z(T,n) &= \int dx\, \exp\left[ -\frac{V}{\kT}\left( \FbarD + \bar{U}(T,n \right) \right] \label{eqn:Z} \\
+  Z^*(T,n) &= \int dx\, \exp\left[ -\frac{V}{\kT}\FbarD \right] \label{eqn:Zstar}
 \end{align}
@@
   \FbarD &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n) \\
-  \UbarD &= \frac{U(T,n+x) + U(T,n-x)}{2} - U(T,n)
+  \UbarD &= \frac{U(T,n+x,\lambdaRG) + U(T,n-x,\lambdaRG)}{2} - U(T,n,\lambdaRG)
 \end{align}
-Finally, $U(T,n)$ is given by Forte as
+Finally, $U(T,n,\lambdaRG)$ is given by Forte as
 \begin{align}
-  U(T,n) = &{} g^\text{HS}(n_{eff};\sigma)\alpha - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha\omega^2 \nonumber \\
-  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha\gamma \label{eqn:URG} % URG! It sounds like it hurt! :P
+  U(T,n,\lambdaRG) = &{} g^\text{HS}(n_{eff};\sigma)\alpha(\lambdaRG) - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha(\lambdaRG)\omega^2(\lambdaRG) \nonumber \\
+  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha(\lambdaRG)\gamma(\lambdaRG) \label{eqn:URG} % URG! It sounds like it hurt! :P
 \end{align}
@@
 \begin{align}
-  \alpha &= \frac{2}{3}\pi\epsilon\sigma^3(\lambda^3 - 1) \\
-  \omega^2 &= \frac{1}{5}\sigma^2\frac{\lambda^5 - 1}{\lambda^3 - 1} \\
-  \gamma &= \frac{1}{70}\sigma^4\frac{\lambda^7 - 1}{\lambda^3 - 1}
+  \alpha(\lambdaRG) &= \frac{2}{3}\pi\epsilon\sigma^3(\lambdaRG^3 - 1) \\
+  \omega^2(\lambdaRG) &= \frac{1}{5}\sigma^2\frac{\lambdaRG^5 - 1}{\lambdaRG^3 - 1} \\
+  \gamma(\lambdaRG) &= \frac{1}{70}\sigma^4\frac{\lambdaRG^7 - 1}{\lambdaRG^3 - 1}
 \end{align}
-The pair correlation function $g(n,r)$ tells us how likely we are to find another hard sphere some distance away, given some density.
-In equation~\ref{eqn:URG} we are considering this quantity at contact ($r = \sigma$) with some effective density.
+The pair correlation function $g(n,r)$ tells us how likely we are to
+find another hard sphere some distance away, given some density. In
+equation~\ref{eqn:URG} we are considering this quantity at contact ($r
+= \sigma$) with an effective density $n_{eff}$.
+
+\fixme{Transition?}
 
@@
 simple liquids, as long as you are far from the critical point---which
-is most of where we calculate. TPT is a rigorous power series
+is most of the time. TPT is a rigorous power series
 expansion in a small parameter; as long as the parameter is small, the
@@
 \begin{align}
-  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambda w^\text{SW}(\1,\2) \label{eqn:small-perturbation}
+  w_\text{SW}(\1,\2) &= - \varepsilon \text{ for } \sigma < r < \lambdaSW\sigma
 \end{align}
-where
+so that we have
 \begin{align}
-  w^\text{SW}(\1,\2) &= - \varepsilon \text{ for } \sigma < r < \lambda\sigma
+  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambdaSW w_\text{SW}(\1,\2) \label{eqn:small-perturbation}
 \end{align}
 
-The total potential energy of our system and its derivatives with resepect to $\lambda$---which become important later---are
+The total potential energy of our system and its derivatives with
+resepect to $\lambdaSW$ (which become important later) are
 \begin{align}
   \widetilde{V} &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
-  \hookrightarrow \frac{\partial \widetilde{V}}{\partial\lambda} &= \sum_{i=1}^N\sum_{j>i}^N w^\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
+  \frac{\partial \widetilde{V}}{\partial\lambdaSW} &= \sum_{i=1}^N\sum_{j>i}^N w_\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
   &\equiv W_N \\
-  \hookrightarrow \frac{\partial^2 \widetilde{V}}{\partial\lambda^2} &= 0
+  \frac{\partial^2 \widetilde{V}}{\partial\lambdaSW^2} &= 0
 \end{align}
 
-The \emph{configuration integral} (the potential energy component of the partition function) and its derivative with respect to $\lambda$ are
+The \emph{configuration integral} (the potential energy component of
+the partition function) and its derivative with respect to $\lambdaSW$
+are
 \begin{align}
-  Z_\lambda &= \int d\rr^N\, e^{-\beta \widetilde{V}(\lambda)} \\
-  \hookrightarrow \frac{\partial Z_\lambda}{\partial\lambda} &=  \int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}(\lambda)} \\
-  &= \int d\rr^N\, -\beta\frac{\partial \widetilde{V}}{\partial\lambda}e^{-\beta \widetilde{V}(\lambda)} \\
-  &= \int d\rr^N\, -\beta W_N e^{-\beta \widetilde{V}(\lambda)}
+  Z_\lambda &= \int d\rr^N\, e^{-\beta \widetilde{V}} \\
+  \frac{\partial Z_\lambda}{\partial\lambdaSW} &=  \int d\rr^N\, \frac{\partial}{\partial\lambdaSW}\left( e^{-\beta \widetilde{V}} \right) \nonumber \\
+  &= \int d\rr^N\, -\beta\frac{\partial \widetilde{V}}{\partial\lambda}e^{-\beta \widetilde{V}} \nonumber \\
+  &= \int d\rr^N\, -\beta W_N e^{-\beta \widetilde{V}}
 \end{align}
-And the non-ideal part of the free energy (i.e.~the excess free energy) and its derivative are
+And the non-ideal part of the free energy (i.e.~the excess free
+energy) and its derivative are
 \begin{align}
   F &= -\kT\ln\left( \frac{Z_\lambda}{V^N} \right) \\
-  \hookrightarrow \frac{\partial F}{\partial\lambda} &= \frac{V^N}{Z_\lambda}\frac{\partial}{\partial\lambda}\left( \frac{Z_\lambda}{V^N} \right) \\
-  &= -\kT\left( \frac{1}{Z_\lambda}\frac{\partial Z_\lambda}{\partial\lambda} \right) \\
-  &= \frac{1}{Z_\lambda} \int d\rr^N\, W_N e^{-\beta \widetilde{V}(\lambda)}\\
-  &= \left\langle W_N \right\rangle_\lambda \label{eqn:dfdlambda}
+  \frac{\partial F}{\partial\lambdaSW} &= \frac{V^N}{Z_\lambda}\frac{\partial}{\partial\lambdaSW}\left( \frac{Z_\lambda}{V^N} \right) \nonumber \\
+  &= -\kT\left( \frac{1}{Z_\lambda}\frac{\partial Z_\lambda}{\partial\lambdaSW} \right) \nonumber \\
+  &= \frac{1}{Z_\lambda} \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \label{eqn:dfdlambda_Z}\\
+  &= \left\langle W_N \right\rangle_{\lambdaSW} \label{eqn:dfdlambda}
 \end{align}
-(where $-\kT$ and $-\beta$ have divided out to $1$). Here, $V^N$ is
-the total volume taken to the $N^{th}$ power, and
-$\left\langle W_N \right\rangle_\lambda$ is the ensemble
-average of $W_N$, with the subscript indicating some given
-value of $\lambda$.
+(where $-\kT$ and $-\beta$ have divided out to $1$ before equation~\ref{eqn:dfdlambda_Z}). Here, $V^N$ is
+the total volume taken to the $N^{th}$ power, and $\left\langle W_N
+\right\rangle_{\lambdaSW}$ is the ensemble average of $W_N$, with the
+subscript indicating some value of $\lambdaSW$.
 
-\fixme{The following could probably go someplace else (is it even necessary?):
-  We can simplify this even further by noting
-  \begin{align}
-    \left\langle W_N \right\rangle &= \left\langle \sum_i\sum_jw(\mathbf{r}_i,\mathbf{r}_j) \right\rangle \\
-    &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle
-    \intertext{which gives us}
-    \frac{\partial F}{\partial\lambda} &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle_\lambda
-\end{align}}
+%% \fixme{The following could probably go someplace else (is it even necessary?):
+%%   We can simplify this even further by noting
+%%   \begin{align}
+%%     \left\langle W_N \right\rangle &= \left\langle \sum_i\sum_jw(\mathbf{r}_i,\mathbf{r}_j) \right\rangle \\
+%%     &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle
+%%     \intertext{which gives us}
+%%     \frac{\partial F}{\partial\lambda} &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle_\lambda
+%% \end{align}}
 
 Note that equation~\ref{eqn:dfdlambda} is exact; now take a series expansion:
-\begin{align}
-  \left\langle W_N \right\rangle_\lambda &= \left\langle W_N\right\rangle_{\lambda = 0} + (\lambda)\frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion}
-\end{align}
-The derivative term is
 \begin{widetext}
 \begin{align}
-    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{\partial}{\partial\lambda}\left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}}\right) \\
-    &= \frac{1}{Z_\lambda}\frac{\partial}{\partial\lambda}\left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\frac{\partial}{\partial\lambda}Z_\lambda \\
-    &= \frac{1}{Z_\lambda}\int d\rr^N\, \left( e^{-\beta \widetilde{V}}\frac{\partial}{\partial\lambda}W_N + W_N\frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\int d\rr^N\, \frac{\partial}{\partial\lambda}e^{-\beta \widetilde{V}} \\
+  \left\langle W_N \right\rangle_{\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} + (\lambdaSW)\frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW}\bigg|_{\lambdaSW = 0} + \mathcal{O}(\lambdaSW^2) \label{eqn:Wn-expansion}
 \end{align}
-\end{widetext}
+%% \end{widetext}
+The derivative term is
+%% \begin{widetext}
+  \begin{align}
+    \frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW} &= \frac{\partial}{\partial\lambdaSW}\left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}}\right) \nonumber \\
+    &= \frac{1}{Z_\lambda}\frac{\partial}{\partial\lambdaSW}\left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\frac{\partial}{\partial\lambdaSW}Z_\lambda \nonumber \\
+    &= \frac{1}{Z_\lambda}\int d\rr^N\, \left( e^{-\beta \widetilde{V}}\frac{\partial}{\partial\lambdaSW}W_N + W_N\frac{\partial}{\partial\lambdaSW}e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\int d\rr^N\, \frac{\partial}{\partial\lambdaSW}e^{-\beta \widetilde{V}}
+  \end{align}
+%% \end{widetext}
 
 Noting that $\partial W_N/\partial\lambda = 0$, we have
-\begin{widetext}
+%% \begin{widetext}
   \begin{align}
-    \frac{\partial}{\partial\lambda}\left\langle W_N \right\rangle_{\lambda} &= \frac{1}{Z_\lambda}\int d\rr^N (-\beta) W_N^2 e^{-\beta \widetilde{V}} - \frac{1}{Z_\lambda^2}\left(\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta \widetilde{V}} \right) \\
-    &= -\beta\left[ \frac{1}{Z_\lambda}\int d\rr^N\, W_N^2e^{-\beta \widetilde{V}} - \left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)^2 \right] \\
-    &= -\beta\left[ \left\langle W_N^2 \right\rangle_\lambda - \left\langle W_N \right\rangle_\lambda^2 \right] \\
-    &\equiv -\beta\sigma_\lambda^2
+    \frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW} &= \frac{1}{Z_\lambda}\int d\rr^N (-\beta) W_N^2 e^{-\beta \widetilde{V}} - \frac{1}{Z_\lambda^2}\left(\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta \widetilde{V}} \right) \nonumber \\
+    &= -\beta\left[ \frac{1}{Z_\lambda}\int d\rr^N\, W_N^2e^{-\beta \widetilde{V}} - \left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)^2 \right] \nonumber \\
+    &= -\beta\left[ \left\langle W_N^2 \right\rangle_{\lambdaSW} - \left\langle W_N \right\rangle_{\lambdaSW}^2 \right] \nonumber \\
+    &\equiv -\beta\sigma_{\lambdaSW}^2
   \end{align}
@@
 \begin{align}
-  \left\langle W_N \right\rangle_\lambda &= \left\langle W_N\right\rangle_{\lambda = 0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2) \label{eqn:Wn-expansion-simplified}
+  \left\langle W_N \right\rangle_{\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2) \label{eqn:Wn-expansion-simplified}
 \end{align}
@@
 \begin{align}
-  \frac{\partial F}{\partial\lambda} &= \left\langle W_N\right\rangle_{\lambda = 0} - \beta\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\lambda^2)
+  \frac{\partial F}{\partial\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2)
 \end{align}
 
-We can integrate this expression over $F\in[F_0,F]$ and $\lambda\in[0,1]$ to obtain
+Now integrate this expression over $F\in[F_0,F]$ (\textsc{N.B.}~this is different than the RGT baseline free energy $F_0(T,n)$ in section~\ref{subsec:RGT}) and $\lambda\in[0,1]$ to obtain
 \begin{align}
-  F - F_0 &= (1-0)\left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^2) \nonumber \\
-  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_\lambda - \frac{1}{2}\beta\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW}
+  \int_{F_0}^F\, dF &= \int_0^1 d\lambdaSW\, \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2) \nonumber \\
+  \intertext{\fixme{Do I need that first line?}}
+  \hookrightarrow F - F_0 &= (1-0)\left\langle W_N \right\rangle_{\lambdaSW} - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^2) \nonumber \\
+  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_{\lambdaSW} - \frac{1}{2}\beta\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW}
 \end{align}
@@
 \begin{align}
-  \beta F &= \beta F_0 + \beta \left\langle W_N \right\rangle_{\lambda=0} - \beta^2\lambda\sigma_{\lambda=0}^2 + \mathcal{O}(\beta^3)
+  \beta F &= \beta F_0 + \beta \left\langle W_N \right\rangle_{\lambdaSW=0} - \beta^2\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^3)
 \end{align}
 
-Zwanzig further showed that the $n^{th}$ term could be written in terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle W_N\right\rangle_{\lambda=0} \right]^\nu \right\rangle_{\lambda=0}$ with $\nu \leq n$.
-This requires us to know the form of higher-order correlation functions of the reference system, which can be quite complicated.
-In the next section we will see a simple aproximation for the $n=2$ term.
+Zwanzig further showed that the $n^{th}$ term could be written in
+terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle
+  W_N\right\rangle_{\lambdaSW=0} \right]^\nu \right\rangle_{\lambdaSW=0}$
+with $\nu \leq n$. This requires us to know the form of higher-order
+correlation functions of the reference system, which can be quite
+complicated. In the next section we will see a simple aproximation for
+the $n=2$ term.
 
@@
 \subsection{Square well liquid free enrgy}\label{subsec:SW}
-For the baseline, this work is based on the square well liquid free energy:\cite{Hughes13}
+\fixme{I need to either note here that I use F/V (see start of section~\ref{sec:methods}), or multiply every term in this section by V}
+For the baseline, this work uses the square well liquid free energy:\cite{Hughes13}
 \begin{align}
   f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
-\end{align}
-The long-range attractive term, $\fattr$ is given by
-\begin{align}
-  \fattr &= n a_1(n) \ .
 \end{align}
@@
   n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
-  &= \eta
+  &= \eta \label{eqn:packingFrac}
 \end{align}
@@
 from compression of the fluid due to the dispersion interaction
-itself, and is approximated using the local compressibility
+itself. It is approximated using the local compressibility
 approximation (LCA), which assumes the energy fluctuation is simply
-related to the compressibility of a hard-sphere reference
+related \fixme{what precisely does ``simply related'' mean?} to the compressibility of a hard-sphere reference
 fluid.\cite{Barker76}
@@
 \begin{align}
-  \alpha^\text{VDW} &= 2\pi\epsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
-  \intertext{using reduced units of $x = r/\sigma$, we have}
-  &= 2\pi\sigma^3\int_1^\infty x^2\phi(x)\,dx \\
-  &= 3b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)\,dx
+  \alpha^\text{VDW} &= 2\pi\varepsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
+  \intertext{using reduced units of $x = r/\sigma$ (\textsc{N.B.}~this is unrelated to the RGT amplitude of fluctuations $x$ from section~\ref{subsec:RGT}), we have}
+  &= 2\pi\sigma^3\varepsilon\int_1^\infty x^2\phi(x)\,dx \nonumber \\
+  &= 3b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)\,dx
 \end{align}
@@
 
-Then, the high-temp expansion is an expansion of the monomer term:
+Then, the high temperature expansion is an expansion of the monomer term:
 \begin{align}
-  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \cdots
+  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \mathcal{O}(\beta^3)
 \end{align}
@@
 \begin{align}
-  a_1 &= -2\pi n \epsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
-  &= -3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+  a_1 &= -2\pi n \varepsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+  &= -3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
 \end{align}
 If we assume random correlations between the particles' positions, for
-all distances, we have $g^\text{HS}(r) = 1$. This yields
+all distances, we have $g^\text{HS}(r) = 1$. This yields \fixme{derive this---or at least explain a little bit}
 \begin{align}
@@
 
-The second-order term is quite a bit more complicated; one must have
-knowledge of higher-order correlation functions to determine
-$a_2$. This term describes the response of the attractive energy due
-to the compression of the fluid from the attractive well. Gil-Villegas
-\emph{et al.} base their expression on an approximation by Barker and
-Henderson.\cite{Barker67} Barker and Henderson's approximation
-considers fluctuations in the number of particles in the potential
-well. In this way, the fluctuations in $a_2$ are related to the
-pressure and compressibility of the liquid. Given pressure
-$P^\text{HS}$ and isothermal compressibility $K^\text{HS} =
-kT\left(\partial n /\partial P^\text{HS}\right)_T$, we have:
+The second-order term is quite a bit more complicated. As discussed in
+section~\ref{subsec:TPT}, one must have knowledge of higher-order
+correlation functions to determine $a_2$. This term describes the
+response of the attractive energy due to the compression of the fluid
+from the attractive well. Gil-Villegas \emph{et al.} base their
+expression on an approximation by Barker and Henderson.\cite{Barker67}
+Barker and Henderson's approximation considers fluctuations in the
+number of particles in the potential well. In this way, the
+fluctuations in $a_2$ are related to the pressure and compressibility
+of the liquid. Given pressure $P^\text{HS}$ and isothermal
+compressibility $K^\text{HS} = kT\left(\partial n /\partial
+P^\text{HS}\right)_T$, we have:
 \begin{align}
-  a_2 &= -\pi n \epsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \\
-  &= -\pi n \epsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \\
-  &= \frac{1}{2}\epsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\epsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+  a_2 &= -\pi n \varepsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \nonumber \\
+  &= -\pi n \varepsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \nonumber \\
+  &= \frac{1}{2}\varepsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
 \end{align}
@@
 We use the isothermal compressibility in terms of packing fraction
-$\eta$, given by the Percus-Yevick exprssion:\cite{Barker76}
+$\eta$ (see equation~\ref{eqn:packingFrac}), given by the Percus-Yevick exprssion:\cite{Barker76}
 \begin{align}
@@
 in the liquid. Since our fluid is of infinite extent, the free energy
-is infinite. Therefore, in the calculations, I work with free energy density:
+is infinite. Therefore, in the calculations I work with free energy density:
 \begin{align}
-  f(T,n) &\equiv \frac{F}{V}(T,n)
+  f(T,n) &\equiv \frac{F(T,n)}{V}
 \end{align}
@@
 \end{align}
-It is the value $\npart$ that we adjust to find coexistence curves.
+$\npart$ is some particular density; it is this value that we adjust to find coexistence curves.
 
@@
   \centering
-%%  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
-  \caption{Grand free energy per unit volume, SW, low temp}
+  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
+  \caption{Grand free energy per unit volume, square well, low temperature}
   \label{fig:SW-phi-lowT}
@@
 for $\npart$ such that the two minima have the same value for
-$\phi(T,n)$. (It is helpful to note that if $\npart$ is increased,
-$\phi(T,\nliq)$ increases, and $\phi(T,\nliq)$ decreases if $\npart$
-is decresed.) The program I wrote uses the previous temperature's
+$\phi(T,n)$. [It is helpful to note that if $\npart$ is increased,
+$\phi(T,\nliq)$ increases, and vice-versa.] The program I wrote uses the previous temperature's
 $\npart$ as a ``first guess,'' maximizing $\phi(T,\npart)$. From
@@
 The algorithm is as follows; start at some temperature $T=T_{start}$,
-with an intial guess for $\npart$.
+with an intial guess for $\npart$. This initual guess can be read off
+of a plot, and optimized ``by hand'' with minimal effort.
 \begin{enumerate}
@@
 
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
+  \end{center}
+  \caption{Liquid-vapor coexistance for a square well liquid.}
+  \label{fig:coexistance_SW}
+\end{figure}
+
+Figure~\ref{fig:coexistance_SW} is a demonstration of this algorithm. \fixme{Compare this one to something.}
+
 \subsection{Free Energies}\label{subsec:free-energies}
-The $a_j(n)$ terms require even more simplification when
-``converting'' from the mathematical formulation to Python code, to
-rid ourselves of troublesome integrals. Gil-Villegas gives us
+To minmize calculation of integrals, we follow Gil-Villegas' approximations to the $a_j(n)$ terms.
 \newcommand\eff{\textit{eff}}
 \begin{align}
-  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{\eff}) \\
-  &= -4\eta\epsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{\eff}/2 \right)}{(1 - \eta_{\eff})^3}
+  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{\eff}) \nonumber \\
+  &= -4\eta\varepsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{\eff}/2 \right)}{(1 - \eta_{\eff})^3}
 \end{align}
@@
 \subsection{RG partition functions}\label{subsec:fbar-ubar}
-I use the midpoint method for evaluating the integrals of $\FbarD$
-and $\UbarD$.
+I use the midpoint method for evaluating the integrals $Z$ and
+$Z^*$ (equations~\ref{eqn:Z} and~\ref{eqn:Zstar}). This method was
+chosen for ease of implementation and low comuptational cost. The approximation is
+\begin{align}
+  \int_a^b dx\, f(x) &\approx dx\sum_{i=0}^{N-1} f(x_i) \\
+  x_i &= a + dx(i+1/2)
+\end{align}
 
@@
 
-For this work, I use $\lambda_\text{SW} = 1.5$.
+For this work, I use $\lambdaSW = 1.5$.
 
@@
   \begin{center}
-%%  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
-  \end{center}
-  \caption{Liquid-vapor coexistance for a square well liquid.}
-  \label{fig:coexistance_SW}
-\end{figure}
-
-\begin{figure}
-  \begin{center}
-%%  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
+  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
   \end{center}
@@
   \begin{center}
-%%  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
+  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
   \end{center}
@@
 
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=\columnwidth]{figs/coexistence-RG-forte}
+  \end{center}
+  \caption{Liquid-vapor coexistence, comparted with Forte's 2011 data}
+  \label{fig:coexistence-RG-forte}
+\end{figure}
+
 Figures~\ref{fig:phi-RG-lowT} and~\ref{fig:phi-RG-highT} show the
-grand free energy per unit volume at a low temperature, and near the
-critical point. It is clear that the program is not de-bugged.
+grand free energy per unit volume at a low temperature and near the
+critical point, respectively. Results are not particularly clear at
+this point, as the program has not been fully debugged. This is
+evident in figure~\ref{fig:phi-RG-lowT}; one notes that the curve for
+$i=1$ is neither complete nor
+accurate. Figure~\ref{fig:coexistence-RG-forte} gives an even starker
+picture of the shortcomings of my RG implementation. We see that the
+baseline ($i=0$; identical to SW calculations) is reasonable, with
+expected shortcomings near the critical point. However, calculations
+including even one iteration ($i=1$) flat-out fail.
 
@@
 
-I conclude that it would be worthwhile to further investigate implementing renormalization group theory in the Roundy DFT.
+We do not yet know if RGT is worth implementing. Since the program has
+not been fully debugged, I am unable to say what the computational
+cost would be to obtain an accurate liquid-vapor coexistene curve.
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \bibliographystyle{unsrt}

### 2014-06-05 papers/thesis-roth/project.tex
project report Tested on elliot. Took 0:20.40.

@@
 % special values for n
-\newcommand{\npart}{\ensuremath{n_{part}}}
-\newcommand{\nliq}{\ensuremath{n_{liquid}}}
-\newcommand{\nvap}{\ensuremath{n_{vapor}}}
+\newcommand{\npart}{\ensuremath{n_\textit{part}}}
+\newcommand{\nliq}{\ensuremath{n_\textit{liquid}}}
+\newcommand{\nvap}{\ensuremath{n_\textit{vapor}}}
+\newcommand{\neff}{\ensuremath{n_\textit{eff}}}
 
@@
 \newcommand{\Fattr}{\ensuremath{F_\text{attr}(T,n)}}
+\newcommand{\Fex}{\ensuremath{F_\text{ex}}}
 
@@
 \newcommand{\fattr}{\ensuremath{f_\text{attr}(T,n)}}
+\newcommand{\fsw}{\ensuremath{f_\text{SW}(T,n)}}
 
 % lambda
-\newcommand{\lambdaSW}{\ensuremath{\lambda_\text{SW}}}
-\newcommand{\lambdaRG}{\ensuremath{\lambda_\text{RG}}}
+\newcommand{\lambdaSW}{\ensuremath{\lambda_\text{sw}}}
+\newcommand{\lambdaRG}{\ensuremath{\lambda_\text{rg}}}
 
-% pair potential
-\newcommand{\1}{\ensuremath{\textbf{r}_1}}
-\newcommand{\2}{\ensuremath{\textbf{r}_2}}
-\newcommand{\3}{\ensuremath{\textbf{r}_3}}
-\newcommand{\4}{\ensuremath{\textbf{r}_4}}
+% TPT
+\newcommand{\III}{\ensuremath{\textbf{r}_{12}}} % Roman numerals for \12
+\newcommand{\Vtilde}{\ensuremath{\widetilde{V}}}
+\newcommand{\Wtilde}{\ensuremath{\widetilde{W}}}
 
 % RG stuff
-\newcommand{\FbarD}{\ensuremath{\bar{F}_D(T,n)}}
-\newcommand{\UbarD}{\ensuremath{\bar{U}_D(T,n)}}
+\newcommand{\FbarD}{\ensuremath{\bar{F}_D(T,n,x)}}
+\newcommand{\UbarD}{\ensuremath{\bar{U}_D(T,n,x,\lambdaRG)}}
 
@@
 
-Near the critical point of a liquid, there are density fluctuations at
-all length scales. In this regime, mean field theories break down. \fixme{re-word the next sentence} As
-the critical point is approached, there becomes less and less of a
-difference in free energy of the liquid and vapor states; this causes
-difficulty in modeling liquid-vapor coexistence. This work develops a
-method for calculating liquid-vapor coexistence, and applies
-renormalization group theory to the square well liquid.
+This work develops a method for calculating liquid-vapor coexistence,
+and applies renormalization group theory to the square well
+liquid. Near the critical point of a liquid, there are density
+fluctuations at all length scales. In this regime, mean field theories
+break down. As the liquid approaches the critical point, the
+difference in free energy between the liquid and vapor states
+diminishes; this causes difficulty in modeling liquid-vapor
+coexistence.
 
@@
 \section{Introduction}
-
-Gasses have been accurately modeled since the 19$^{th}$
+Gases have been accurately modeled since the 19$^{th}$
 century\cite{Lederman92} and formulation of the kinetic theory of gasses. With the
@@
 desireability. Liquid water is---as far as we know---essential for
-developing and sustaining life. All celluar processes take place in an
+developing and sustaining life. All cellular processes take place in an
 aqueous environment. Understanding liquids is in many ways a step
@@
 point of a liquid. At the appropriate density and temperature, the
-distinction between liquid and vapor dissapears. Beyond this point,
+distinction between liquid and vapor disappears. Beyond this point,
 you have a ``supercritical fluid.'' Supercritical fluids are used in
@@
 
+In this work, I focus exclusively on what is called the square well
+liquid. The simplest model of a liquid is to consider each molecule to
+be a hard sphere (i.e.~spheres which cannot overlap) with a square-well
+potential between any pair of spheres. This is a widely used
+approximation to simple liquids, and one of the foundations of
+this work.
+
+There are several approaches to study liquids, each with advantages
+and disadvantages. Some methods, such as Monte Carlo and Molecular
+Dynamics, are simulations of systems with a discrete number of
+particles. Other methods are more abstract in nature. Methods such as
+Thermodynamic Perturbation Theory, Density Functional Theory and
+Renormalization Group Theory are mathematical theories that were not
+necessarily developed with computational methods in mind (though in
+practice numerical solutions are required).
+
+Monte Carlo simulations (MC) are one method. These simulations can
+give exact answers, if you run them ``long enough.''  So, MC methods
+are very accurate, but very computationally expensive. Molecular
+dynamics simulations (MD) calculate forces on molecules over a
+particular span of time. These simulations are also accurate but even
+more computationally expensive than MC. Calculations simulating only a
+few hundred molecules for picoseconds to nanoseconds of interactions
+can be expected.\cite{Hughes13}
+
+Thermodynamic Perturbation Theory (TPT) gives a good description of
+liquid systems far from the critical point. Similar to quantum
+perturbation theory, TPT utilizes the power and simplicity of a power
+series expansion in a small parameter. Unfortunately, TPT does break
+down near the critical point.
+
+Density Functional Theory (DFT) implements calculations based on
+ground-state electron density distributions. DFT is good for solid
+state systems, but liquids are too complicated. Simple DFT MD
+simulations ``predict'' a melting point of water at 120$^\circ$C. More
+complicated---and therefore computationally expensive---models are
+slightly more accurate; one model calculates a melting point of ice as
+87$^\circ$C.\cite{Yoo11} Classical DFT (cDFT) describes a fluid by a
+free energy functional of the particle density
+distribution.\cite{Krebs13} cDFT efficiently calculates density
+profiles for inhomogeneous systems, that is, systems where there are
+liquid interfaces such as dissolved solutes or ``hard walls.''
+However, existing cDFTs do not accurately model the critical point of
+a fluid.
+
 It is the aim of renormalization group theory to tackle the critical
-point. The reason that calculations are dificult near the critical
-point is that there are large-scale fluctuations of the density. In
-fact, there are density fluctuations at \emph{all} length
-scales, which is why mean field theories fail. Renormalization group theory is an iterative procedure that
-was developed in th 1950's, and Kenneth Wilson's work with the theory
-in the 1970's won him the 1982 Nobel Prize in physics.\cite{Nobel82}
-
-\fixme{Transition here?}
-
-The simplest model of a liquid is to consider each molecule to be a
-hard sphere (i.e.~spheres cannot overlap) with a square-well potenial
-between any pair of spheres. This is called the ``square well
-liquid,'' and it is a widely used approximation to simple
-liquids. This is one of the foundations of this work.
-
-We use statistical mechanics to study liquids; this requires large
-sample sizes, long simulation times, or some combination of the
-two. Monte Carlo simulations are exact, if you allow them to run
-``long enough.'' The Roundy group uses classical Density-funtional
-Theory (DFT) to calculate free energies and probability distributions
-of bulk liquids. These calculations are ultimately compared to Monte Carlo simulations.
-
-The end goal of this work is to investigate into whether or not
-implementing renormalization group theory is worthwhile to the Roundy
-Research Group. Many other groups use renormalization group theory,
-but none of them tell the iteration depth used in their
+point. Calculations are difficult near the critical point because
+there are large-scale fluctuations of the density. Indeed, there are
+density fluctuations at \emph{all} length scales, which causes mean
+field theories to fail. Renormalization group theory is an iterative
+procedure that was developed in the 1950s, and Kenneth Wilson's work
+with the theory in the 1970s won him the 1982 Nobel Prize in
+physics.\cite{Nobel82} The end goal of this particular work is to
+investigate into whether or not implementing renormalization group
+theory is worthwhile to our research group. Many other groups use
+renormalization group theory, but none of them report the iteration
+depth used in their
 calculations.\cite{Forte11,Ramana12,Ghobadi13,Wang13,Zhao13} If only
-few iterations are required, then this would certainly be an effective
-tool to use. However, if many iterations are required, the
-computational cost could be unwieldy.
+few iterations are required, then this could be an effective tool to
+improve the accuracy of cDFTs. However, if many iterations are
+required, the computational cost could be unwieldy.
 
 In section~\ref{sec:theory}, I present some of the main theories I
-deal with: renormalization group theory, thermodynamic perturbation
-theory and the square well liquid free energy. In
+deal with: thermodynamic perturbation
+theory, the square well liquid free energy and renormalization group theory. In
 section~\ref{sec:methods} I discuss the the methods used. I cover the
 calculation of liquid-vapor coexistence curves in some detail, as
-caluclating coexistence near the critical point is tricky. The
+calculating coexistence near the critical point is tricky. The
 results---such as they are---are covered in section~\ref{sec:results}.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Theory}\label{sec:theory}
-
-\subsection{Renormalization Group Theory}\label{subsec:RGT}
-Renormalization group theory (RGT) is an iterative procedure that was
-developed to deal with large-scale fluctuations of the order parameter
-near the critical point of a system. The formulation used here is
-provided by Forte~\textit{et al}.\cite{Forte11}
-
-We start by dividing the total volume into cells of volume $V_i$. We
-assume that within a cell the number density $n(\rr)$ varies only slightly, and that the
-cells do not interact with one another. This comes from Wilson's
-phase cell method.\cite{Ramana12}
-
-Then consider a baseline free enrgy $F_0(T,n)$, usually based on a
-mean field theory. We assume this baseline accounts for all
-fluctuations of wavelength $\lambdaRG < \lambda_0$. We then
-double the size of our phase cell and add a correction to the free
-energy $\delta F_1(T,n)$. This correction accounts for fluctuations
-with a wavelength $\lambdaRG = 2\lambda_0$. The new free
-energy, $F_1(T,n)$, is
-\begin{align}
-  F_1(T,n) &= F_0(T,n) + \delta F_1(T,n)
-\end{align}
-Then double the length scale again; $\delta F_2(T,n)$ accounts for
-fluctionations with $\lambdaRG = 4\lambda_0$ and
-\begin{align}
-  F_2(T,n) &= F_1(T,n) + \delta F_2(T,n)
-\end{align}
-As we continue this process, the $i^{th}$ correction accounts for
-fluctuations with $\lambdaRG = 2^i\lambda_0$, and
-\begin{align}
-  F_i(T,n) &= F_{i-1}(T,n) + \delta F_i(T,n)
-\end{align}
-In the end, your free energy is
-\begin{align}
-  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \Fattr
-\end{align}
-
-Where $\Fattr$ is the contribution due to the longest-wavelength
-density fluctuations. At this stage, I do not entirely understand why
-Forte separates out $\Fattr$.
-
-The $\delta F_i(T,n)$ term is a difference in free energies:
-\begin{align}
-  \delta F_i(T,n) &= -\kT\ln Z - \left( -\kT\ln Z^* \right) \nonumber \\
-  &= -\kT\ln\left[ \frac{Z(T,n)}{Z^*(T,n)} \right]
-\end{align}
-Here, $Z$ is the partition function for the entire system, and $Z^*$
-is the partition function of a reference system consisting only of the
-term $\FbarD$. Forte gives these terms as
-\begin{align}
-  Z(T,n) &= \int dx\, \exp\left[ -\frac{V}{\kT}\left( \FbarD + \bar{U}(T,n \right) \right] \label{eqn:Z} \\
-  Z^*(T,n) &= \int dx\, \exp\left[ -\frac{V}{\kT}\FbarD \right] \label{eqn:Zstar}
-\end{align}
-where $x$ is the amplitude of density fluctuations. $\FbarD$ and $\UbarD$ are formulated similar to averages:
-\begin{align}
-  \FbarD &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n) \\
-  \UbarD &= \frac{U(T,n+x,\lambdaRG) + U(T,n-x,\lambdaRG)}{2} - U(T,n,\lambdaRG)
-\end{align}
-Finally, $U(T,n,\lambdaRG)$ is given by Forte as
-\begin{align}
-  U(T,n,\lambdaRG) = &{} g^\text{HS}(n_{eff};\sigma)\alpha(\lambdaRG) - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha(\lambdaRG)\omega^2(\lambdaRG) \nonumber \\
-  &{} + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha(\lambdaRG)\gamma(\lambdaRG) \label{eqn:URG} % URG! It sounds like it hurt! :P
-\end{align}
-with
-\begin{align}
-  \alpha(\lambdaRG) &= \frac{2}{3}\pi\epsilon\sigma^3(\lambdaRG^3 - 1) \\
-  \omega^2(\lambdaRG) &= \frac{1}{5}\sigma^2\frac{\lambdaRG^5 - 1}{\lambdaRG^3 - 1} \\
-  \gamma(\lambdaRG) &= \frac{1}{70}\sigma^4\frac{\lambdaRG^7 - 1}{\lambdaRG^3 - 1}
-\end{align}
-The pair correlation function $g(n,r)$ tells us how likely we are to
-find another hard sphere some distance away, given some density. In
-equation~\ref{eqn:URG} we are considering this quantity at contact ($r
-= \sigma$) with an effective density $n_{eff}$.
-
-\fixme{Transition?}
 
@@
 Thermodynamic perturbation theory (TPT) gives a good description of
-simple liquids, as long as you are far from the critical point---which
+simple liquids, as long a system is far from the critical point---which
 is most of the time. TPT is a rigorous power series
@@
 as separable into two parts, one for short-range repulsion and another
-for long-range attraction.
-
-The pair potential $v(\1,\2)$ describes the interaction between two
-hard spheres. For the square-well model, the pair potential is
+for long-range attraction. The short-range part is considered to be a
+reference potential to which we add a perturbation. The following pair
+potential $v(\III)$ describing the interaction between hard spheres is
+our reference:
 \begin{align}
-  v_\text{HS}(\1,\2) &=
+  v_\text{hs}(\III) &=
     \begin{cases}
       \infty & r < \sigma \\
-      0 & \sigma < r
+      0 & \sigma \leq r
     \end{cases}
 \end{align}
-
-We consider this as our referecne, and take a square well attraction as a perturbation:
+To this reference, we take a square well attraction as a perturbation:
 \begin{align}
-  w_\text{SW}(\1,\2) &= - \varepsilon \text{ for } \sigma < r < \lambdaSW\sigma
+  w_\text{sw}(\III) &= - \varepsilon \quad \sigma < r < \lambdaSW\sigma
 \end{align}
@@
 \begin{align}
-  v_\lambda(\1,\2) &= v_\text{HS}(\1,\2) + \lambdaSW w_\text{SW}(\1,\2) \label{eqn:small-perturbation}
+  v_\xi(\III) &= v_\text{hs}(\III) + \xi w_\text{sw}(\III) \label{eqn:small-perturbation}
 \end{align}
-
-The total potential energy of our system and its derivatives with
-resepect to $\lambdaSW$ (which become important later) are
+where $\xi$ is a parameter that allows us to make the perturbation arbitrarily small.
+The total potential energy $\Vtilde$ of our system and its derivatives with
+respect to $\xi$ (which become important later) are
 \begin{align}
-  \widetilde{V} &= \sum_{i=1}^N\sum_{j>i}^N v_\lambda(\mathbf{r}_i,\mathbf{r}_j) \\ % \label{eq:VN} \\
-  \frac{\partial \widetilde{V}}{\partial\lambdaSW} &= \sum_{i=1}^N\sum_{j>i}^N w_\text{SW}(\mathbf{r}_i,\mathbf{r}_j) \nonumber \\
-  &\equiv W_N \\
-  \frac{\partial^2 \widetilde{V}}{\partial\lambdaSW^2} &= 0
+  \Vtilde &= \sum_{i=1}^N\sum_{j>i}^N v_\xi(\mathbf{r}_{ij})
+\end{align}
+\begin{align}
+  \frac{\partial \Vtilde}{\partial\xi} &= \sum_{i=1}^N\sum_{j>i}^N w_\text{sw}(\mathbf{r}_{ij}) \nonumber \\
+  &\equiv \Wtilde \\
+  \frac{\partial^2 \Vtilde}{\partial\xi^2} &= 0
 \end{align}
@@
 The \emph{configuration integral} (the potential energy component of
-the partition function) and its derivative with respect to $\lambdaSW$
+the partition function) and its derivative with respect to $\xi$
 are
 \begin{align}
-  Z_\lambda &= \int d\rr^N\, e^{-\beta \widetilde{V}} \\
-  \frac{\partial Z_\lambda}{\partial\lambdaSW} &=  \int d\rr^N\, \frac{\partial}{\partial\lambdaSW}\left( e^{-\beta \widetilde{V}} \right) \nonumber \\
-  &= \int d\rr^N\, -\beta\frac{\partial \widetilde{V}}{\partial\lambda}e^{-\beta \widetilde{V}} \nonumber \\
-  &= \int d\rr^N\, -\beta W_N e^{-\beta \widetilde{V}}
+  Z_\xi &= \int d\rr^N\, e^{-\beta \Vtilde}
+\end{align}
+\begin{align}
+  \frac{\partial Z_\xi}{\partial\xi} &=  \int d\rr^N\, \frac{\partial}{\partial\xi}\left( e^{-\beta \Vtilde} \right) \nonumber \\
+  &= -\int d\rr^N\, \beta\frac{\partial \Vtilde}{\partial\lambda}e^{-\beta \Vtilde} \nonumber \\
+  &= -\int d\rr^N\, \beta \Wtilde e^{-\beta \Vtilde}
 \end{align}
@@
 \begin{align}
-  F &= -\kT\ln\left( \frac{Z_\lambda}{V^N} \right) \\
-  \frac{\partial F}{\partial\lambdaSW} &= \frac{V^N}{Z_\lambda}\frac{\partial}{\partial\lambdaSW}\left( \frac{Z_\lambda}{V^N} \right) \nonumber \\
-  &= -\kT\left( \frac{1}{Z_\lambda}\frac{\partial Z_\lambda}{\partial\lambdaSW} \right) \nonumber \\
-  &= \frac{1}{Z_\lambda} \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \label{eqn:dfdlambda_Z}\\
-  &= \left\langle W_N \right\rangle_{\lambdaSW} \label{eqn:dfdlambda}
+  \Fex &= -\kT\ln\left( \frac{Z_\xi}{V^N} \right)
+\end{align}
+\begin{align}
+  \frac{\partial \Fex}{\partial\xi} &= -\kT\frac{V^N}{Z_\xi}\frac{\partial}{\partial\xi}\left( \frac{Z_\xi}{V^N} \right) \nonumber \\
+  &= -\kT\left( \frac{1}{Z_\xi}\frac{\partial Z_\xi}{\partial\xi} \right) \nonumber \\
+  &= \frac{1}{Z_\xi} \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \label{eqn:dfdlambda_Z}\\
+  &= \left\langle \Wtilde \right\rangle_{\xi} \label{eqn:dfdlambda}
 \end{align}
 (where $-\kT$ and $-\beta$ have divided out to $1$ before equation~\ref{eqn:dfdlambda_Z}). Here, $V^N$ is
-the total volume taken to the $N^{th}$ power, and $\left\langle W_N
-\right\rangle_{\lambdaSW}$ is the ensemble average of $W_N$, with the
-subscript indicating some value of $\lambdaSW$.
-
-%% \fixme{The following could probably go someplace else (is it even necessary?):
-%%   We can simplify this even further by noting
-%%   \begin{align}
-%%     \left\langle W_N \right\rangle &= \left\langle \sum_i\sum_jw(\mathbf{r}_i,\mathbf{r}_j) \right\rangle \\
-%%     &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle
-%%     \intertext{which gives us}
-%%     \frac{\partial F}{\partial\lambda} &= \frac{1}{2}N^2\left\langle w(\1,\2) \right\rangle_\lambda
-%% \end{align}}
-
-Note that equation~\ref{eqn:dfdlambda} is exact; now take a series expansion:
+the total volume taken to the $N^{th}$ power, and $\left\langle \Wtilde
+\right\rangle_{\xi}$ is the ensemble average of $\Wtilde$, with the
+subscript indicating some value of $\xi$.
+Note that equation~\ref{eqn:dfdlambda} is exact; now take a series expansion around $\xi=0$:
+\begin{align}
+  \left\langle \Wtilde \right\rangle_{\xi} &= \left\langle \Wtilde\right\rangle_{\xi = 0} + (\xi)\frac{\partial\left\langle \Wtilde \right\rangle_{\xi}}{\partial\xi}\bigg|_{\xi = 0} + \mathcal{O}(\xi^2) \label{eqn:Wn-expansion}
+\end{align}
+The derivative term is
 \begin{widetext}
   \begin{align}
-  \left\langle W_N \right\rangle_{\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} + (\lambdaSW)\frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW}\bigg|_{\lambdaSW = 0} + \mathcal{O}(\lambdaSW^2) \label{eqn:Wn-expansion}
-\end{align}
-%% \end{widetext}
-The derivative term is
-%% \begin{widetext}
-  \begin{align}
-    \frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW} &= \frac{\partial}{\partial\lambdaSW}\left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}}\right) \nonumber \\
-    &= \frac{1}{Z_\lambda}\frac{\partial}{\partial\lambdaSW}\left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\frac{\partial}{\partial\lambdaSW}Z_\lambda \nonumber \\
-    &= \frac{1}{Z_\lambda}\int d\rr^N\, \left( e^{-\beta \widetilde{V}}\frac{\partial}{\partial\lambdaSW}W_N + W_N\frac{\partial}{\partial\lambdaSW}e^{-\beta \widetilde{V}} \right) - \left( \int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\frac{1}{Z_\lambda^2}\int d\rr^N\, \frac{\partial}{\partial\lambdaSW}e^{-\beta \widetilde{V}}
-  \end{align}
-%% \end{widetext}
-
-Noting that $\partial W_N/\partial\lambda = 0$, we have
-%% \begin{widetext}
-  \begin{align}
-    \frac{\partial\left\langle W_N \right\rangle_{\lambdaSW}}{\partial\lambdaSW} &= \frac{1}{Z_\lambda}\int d\rr^N (-\beta) W_N^2 e^{-\beta \widetilde{V}} - \frac{1}{Z_\lambda^2}\left(\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)\left( \int d\rr^N\, (-\beta)W_Ne^{-\beta \widetilde{V}} \right) \nonumber \\
-    &= -\beta\left[ \frac{1}{Z_\lambda}\int d\rr^N\, W_N^2e^{-\beta \widetilde{V}} - \left( \frac{1}{Z_\lambda}\int d\rr^N\, W_N e^{-\beta \widetilde{V}} \right)^2 \right] \nonumber \\
-    &= -\beta\left[ \left\langle W_N^2 \right\rangle_{\lambdaSW} - \left\langle W_N \right\rangle_{\lambdaSW}^2 \right] \nonumber \\
-    &\equiv -\beta\sigma_{\lambdaSW}^2
+    \frac{\partial}{\partial\xi}\left\langle \Wtilde \right\rangle_{\xi} &= \frac{\partial}{\partial\xi}\left( \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta \Vtilde}\right) \nonumber \\
+    &= \frac{1}{Z_\xi}\frac{\partial}{\partial\xi}\left( \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right) - \left( \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right)\frac{1}{Z_\xi^2}\frac{\partial}{\partial\xi}Z_\xi \nonumber \\
+    &= -\beta\left[ \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde^2e^{-\beta \Vtilde} - \left( \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right)^2 \right] \nonumber \\
+    &= -\beta\left[ \left\langle \Wtilde^2 \right\rangle_{\xi} - \left\langle \Wtilde \right\rangle_{\xi}^2 \right] \nonumber \\
+    &\equiv -\beta\sigma_{\xi}^2
   \end{align}
 \end{widetext}
-
-Combining this result with equation~\ref{eqn:Wn-expansion} gives us
+where $\sigma_xi^2$ is the variance of the perturbative potential \Wtilde. Combining this result with equation~\ref{eqn:Wn-expansion} gives us
 \begin{align}
-  \left\langle W_N \right\rangle_{\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2) \label{eqn:Wn-expansion-simplified}
+  \left\langle \Wtilde \right\rangle_{\xi} &= \left\langle \Wtilde\right\rangle_{\xi = 0} - \beta\xi\sigma_{\xi=0}^2 + \mathcal{O}(\xi^2) \label{eqn:Wn-expansion-simplified}
 \end{align}
@@
 \begin{align}
-  \frac{\partial F}{\partial\lambdaSW} &= \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2)
+  \frac{\partial \Fex}{\partial\xi} &= \left\langle \Wtilde\right\rangle_{\xi = 0} - \beta\xi\sigma_{\xi=0}^2 + \mathcal{O}(\xi^2)
+\end{align}
+Now integrate this expression over $\Fex\in[F_\text{ref},\Fex]$ and $\xi\in[0,1]$ to obtain
+\begin{align}
+  \int_{F_\text{ref}}^{\Fex}\, d\Fex &= \int_0^1 d\xi\, \left[ \left\langle \Wtilde\right\rangle_{\xi = 0} - \beta\lambda\sigma_{\xi=0}^2 + \mathcal{O}(\xi^2) \right] \nonumber \\
+  \hookrightarrow \Fex - F_\text{ref} &= (1-0)\left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta(1 - 0)^2\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2) \nonumber \\
+  \hookrightarrow \Fex &= F_\text{ref} + \left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW} \\
+  &\equiv F_\text{ref} + a_1(n) + a_2(n)
+\end{align}
+This is the \emph{high-temperature expansion}, first derived by
+Zwanzig\cite{Zwanzig54}. He further showed that the $n^{th}$ term could be written in
+terms of the mean fluctuations $\left\langle \left[ \Wtilde - \left\langle
+  \Wtilde\right\rangle_{\xi=0} \right]^\nu \right\rangle_{\xi=0}$
+with $\nu \leq n$. This requires us to know the form of higher-order
+correlation functions of the reference system, which can be quite
+complicated. In the next section we will see simple approximations for
+the $n=1$ and $n=2$ terms.
+
+\subsection{Square well liquid free energy}\label{subsec:SW}
+
+This work applies entirely to the homogeneous case---that is, there
+are no walls containing our liquid, nor are there any sort of solutes
+in the liquid. Since our fluid is of infinite extent, the free energy
+is infinite. Therefore, in the calculations I work with free energy density:
+\begin{align}
+  f(T,n) &\equiv \frac{F(T,n)}{V}
 \end{align}
 
-Now integrate this expression over $F\in[F_0,F]$ (\textsc{N.B.}~this is different than the RGT baseline free energy $F_0(T,n)$ in section~\ref{subsec:RGT}) and $\lambda\in[0,1]$ to obtain
+As a baseline for RG, I use the square well liquid free energy:\cite{Hughes13}
 \begin{align}
-  \int_{F_0}^F\, dF &= \int_0^1 d\lambdaSW\, \left\langle W_N\right\rangle_{\lambdaSW = 0} - \beta\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\lambdaSW^2) \nonumber \\
-  \intertext{\fixme{Do I need that first line?}}
-  \hookrightarrow F - F_0 &= (1-0)\left\langle W_N \right\rangle_{\lambdaSW} - \frac{1}{2}\beta(1 - 0)^2\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^2) \nonumber \\
-  \hookrightarrow F &= F_0 + \left\langle W_N \right\rangle_{\lambdaSW} - \frac{1}{2}\beta\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW}
+  f_\text{TPT}(T,n) &= \fid + \fhs + \fsw
 \end{align}
-Multiply by $\beta$ and we have the \emph{high-temperature expansion}, first derived by
-Zwanzig:\cite{Zwanzig54}
-\begin{align}
-  \beta F &= \beta F_0 + \beta \left\langle W_N \right\rangle_{\lambdaSW=0} - \beta^2\lambda\sigma_{\lambdaSW=0}^2 + \mathcal{O}(\beta^3)
-\end{align}
-
-Zwanzig further showed that the $n^{th}$ term could be written in
-terms of the mean fluctuations $\left\langle \left[ W_N - \left\langle
-  W_N\right\rangle_{\lambdaSW=0} \right]^\nu \right\rangle_{\lambdaSW=0}$
-with $\nu \leq n$. This requires us to know the form of higher-order
-correlation functions of the reference system, which can be quite
-complicated. In the next section we will see a simple aproximation for
-the $n=2$ term.
-
-%% Writing the statistical averages in terms of the pair density
-%% $n_\lambda^{(2)}(\1,\2)$ (i.e. the probability per volume squared of having one sphere at \1
-%%   and another at \2) gives us
-%% \begin{align}
-%%   \beta F &=\beta F_0 + \frac{1}{2}\beta\int_0^1 d\lambda\, \iint d\1 d\2\, n_\lambda^{(2)}(\1,\2)w(\1,\2) \label{eqn:Henderson5-2-12}
-%% \end{align}
-%% \davidsays{I think this discussion is pretty redundant with
-%%   parts of the previous derivation in terms of $\widetilde{V}$.}  The pair
-%% density itself can be expanded as
-%% \begin{align}
-%%   n_\lambda^{(2)}(\1,\2) &= n_0^{(2)}(\1,\2) + \lambda\frac{\partial n_\lambda^{(2)}(\1,\2)}{\partial\lambda}\bigg|_{\lambda = 0} + \mathcal{O}(\lambda^2)
-%% \end{align}
-%% Applying this to the 2nd term in equation~\ref{eqn:Henderson5-2-12}, we
-%% have
-%% \begin{align}
-%%   \frac{\beta F_1}{N} &= \frac{\beta}{2N}\iint d\1 d\2\, n_0^{(2)}(\1,\2)w(\1,\2) \\
-%%   &= \frac{\beta n}{2}\int d\mathbf{r}_{12}\, g_0(\1,\2)w(\1,\2)
-%% \end{align}
-%% The second-order $\lambda$ term is quite a bit more complicated:
-%% \begin{widetext}
-%%  \begin{align}
-%%    \begin{split}
-%%    \frac{\beta F_2}{N} =&{} -\frac{\beta^2}{2}\left[ \frac{n}{2}\int d\2\, g_0(\1,\2)w^2(\1,\2) \right.
-\\ &{} + n^2 \iint d\2 d\3\, g_0^{(3)}(\1,\2,\3)w(\1,\2)w(\1,\3) \\ &{} \left.
-+ \frac{n^3}{4}\iiint d\2 d\3 d\4\, \left( g_0^{(4)}(\1,\2,\3,\4) - g_0^{(2)}(\1,\2)g_0^{(2)}(\3,\4) \right)w(\1,\2)2(\3,\4) \right] \\ &{} - \frac{1}{4}S_0(0)\left[ \frac{\partial}{\partial n}\left( n^2\int d\2\, g_0(\1,\2)w(\1,\2) \right)^2 \right]
-%%    \end{split}
-%%  \end{align}
-%% \end{widetext}
-%% This term requires use to know the form of higher-order correlation functions and the structure function $S_0(k)$ of the reference system.
-These things are very difficult.
-Hence, we use approximations, as outlined in section~\ref{sub2sec:disp}.
-
-\subsection{Square well liquid free enrgy}\label{subsec:SW}
-\fixme{I need to either note here that I use F/V (see start of section~\ref{sec:methods}), or multiply every term in this section by V}
-For the baseline, this work uses the square well liquid free energy:\cite{Hughes13}
-\begin{align}
-  f_0(T,n) &= \fid + \fhs + \left( \fdisp - n a_1(n) \right) \ .
-\end{align}
-$\fid$ is the ideal gas free energy, $\fhs$ is the hard sphere
-repulsion, and $\fdisp$ is the attraction. The function $a_1(n)$ is
-the first term from a high-temperature perturbation expansion,
-described in further detail in sections~\ref{subsec:TPT}
-and~\ref{sub2sec:disp}. Each function is outlined below.
+$\fid$ is the ideal gas free energy, $\fhs$ is the free energy contribution due to hard sphere
+repulsion, and $\fsw$ is the free energy contribution due to square well attraction. Each function is outlined below.
 
@@
 Hard sphere repulsive forces are dealt with by the White Bear version
-of Fundamental-Mearuse Theory.\cite{Roth02} The hard sphere excess
-free energy is given in the homogeneous case as
+of Fundamental-Measure Theory,\cite{Roth02} which simplifies to the
+Carnahan-Starling free energy\cite{Carnahan69} in the homogeneous limit.
 \begin{align}
-  \fhs &= \kT\left( \Phi_1(n) + \Phi_2(n) + \Phi_3(n) \right)
+  \fhs &= \kT\left( \Phi_1 + \Phi_2 + \Phi_3 \right)
 \end{align}
-The $\Phi_j(n)$ terms are given as
+The $\Phi_j$ terms are given as
 \begin{align}
-  \Phi_1(n) &= -n_0(n)\ln(1 - n_3(n)) \\
-  \Phi_2(n) &= \frac{n_1(n)n_2(n)}{1 - n_3(n)} \\
-  \Phi_3(n) &= n_2^3(n)\left( \frac{n_3(n) + (1 - n_3(n))^2\ln(1 - n_3(n))}{36\pi n_3^2(n)(1 - n_3(n))^2} \right)
+  \Phi_1 &= -n_0\ln(1 - n_3) \\
+  \Phi_2 &= \frac{n_1n_2}{1 - n_3} \\
+  \Phi_3 &= n_2^3\left( \frac{n_3 + (1 - n_3)^2\ln(1 - n_3)}{36\pi n_3^2(1 - n_3)^2} \right)
 \end{align}
-The \emph{fundamental measure densities} are
+The \emph{fundamental measures} are given by
 \begin{align}
-  n_0(n) &= \frac{n_2}{4\pi R^2} \nonumber \\
-  &= n \\
-  n_1(n) &= \frac{n_2}{4\pi R} \nonumber \\
-  &= nR \\
-  n_2(n) &= n4\pi R^2 \\
-  n_3(n) &= n\frac{4}{3}\pi R^3 \nonumber \\
+  n_0 &= \frac{n_2}{4\pi R^2} \nonumber \\
+  &= n
+\end{align}
+\begin{align}
+  n_1 &= \frac{n_2}{4\pi R} \nonumber \\
+  &= nR
+\end{align}
+\begin{align}
+  n_2 &= n4\pi R^2
+\end{align}
+\begin{align}
+  n_3 &= n\frac{4}{3}\pi R^3 \nonumber \\
   &= \eta \label{eqn:packingFrac}
@@
 \subsubsection{Attraction}\label{sub2sec:disp}
-The attractive free energy includes van der Waals attraction between
+The attractive free energy includes square well attraction between
 hard spheres and is founded in Thermodynamic Perturbation Theory
@@
 \begin{align}
-  \fdisp &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
+  \fsw &= n \left( a_1(n) + \frac{1}{\kT}a_2(n) \right)
 \end{align}
-The $a_j(n)$ terms come from TPT, outlined in section~\ref{subsec:TPT}.
+The terms $a_1(n)$ and $a_2(n)$ come from TPT, outlined in section~\ref{subsec:TPT}.
 
-The first term, $a_1$, is the mean-field dispersion interaction. The
+The first term, $a_1$, is the zeroth-order interaction. The
 second term, $a_2$, describes the effect of fluctuations resulting
@@
 approximation (LCA), which assumes the energy fluctuation is simply
-related \fixme{what precisely does ``simply related'' mean?} to the compressibility of a hard-sphere reference
+related to the compressibility of a hard-sphere reference
 fluid.\cite{Barker76}
 
-In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
-\textit{monomer} contribution, expressed in terms of energy densities
-as:
+Gil-Villegas\cite{gil-villegas97} gives an approximation for $a_1$ as
 \begin{align}
-  f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+  a_1^\text{SW}(n) &= -4\eta\varepsilon(\lambda^3 - 1)\frac{1 - \left( \eta_\textit{eff}/2 \right)}{(1 - \eta_\textit{eff})^3}
 \end{align}
-with $\alpha^\text{VDW}$ given by
+where
 \begin{align}
-  \alpha^\text{VDW} &= 2\pi\varepsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
-  \intertext{using reduced units of $x = r/\sigma$ (\textsc{N.B.}~this is unrelated to the RGT amplitude of fluctuations $x$ from section~\ref{subsec:RGT}), we have}
-  &= 2\pi\sigma^3\varepsilon\int_1^\infty x^2\phi(x)\,dx \nonumber \\
-  &= 3b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)\,dx
-\end{align}
-where $b^\text{VDW}$ is the van der Waals size parameter. It
-corresponds to the volume excluded by two spherical particles of
-volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
-
-Then, the high temperature expansion is an expansion of the monomer term:
-\begin{align}
-  f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \mathcal{O}(\beta^3)
-\end{align}
-
-The term $a_1$ is given by
-\begin{align}
-  a_1 &= -2\pi n \varepsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
-  &= -3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
-\end{align}
-If we assume random correlations between the particles' positions, for
-all distances, we have $g^\text{HS}(r) = 1$. This yields \fixme{derive this---or at least explain a little bit}
-\begin{align}
-  a_1^\text{VDW} = - n \alpha^\text{VDW},
-\end{align}
-the van der Waals mean-field energy.
-
-The second-order term is quite a bit more complicated. As discussed in
-section~\ref{subsec:TPT}, one must have knowledge of higher-order
-correlation functions to determine $a_2$. This term describes the
-response of the attractive energy due to the compression of the fluid
-from the attractive well. Gil-Villegas \emph{et al.} base their
-expression on an approximation by Barker and Henderson.\cite{Barker67}
-Barker and Henderson's approximation considers fluctuations in the
-number of particles in the potential well. In this way, the
-fluctuations in $a_2$ are related to the pressure and compressibility
-of the liquid. Given pressure $P^\text{HS}$ and isothermal
-compressibility $K^\text{HS} = kT\left(\partial n /\partial
-P^\text{HS}\right)_T$, we have:
-\begin{align}
-  a_2 &= -\pi n \varepsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \nonumber \\
-  &= -\pi n \varepsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \nonumber \\
-  &= \frac{1}{2}\varepsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
-\end{align}
-
-We use the isothermal compressibility in terms of packing fraction
-$\eta$ (see equation~\ref{eqn:packingFrac}), given by the Percus-Yevick exprssion:\cite{Barker76}
-\begin{align}
-  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
-\end{align}
-This is called the Local Compressibility Approximation (LCA).
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}\label{sec:methods}
-
-This work applies entirely to the homogeneous case---that is, there
-are no walls containing our liquid, nor are there any sort of solutes
-in the liquid. Since our fluid is of infinite extent, the free energy
-is infinite. Therefore, in the calculations I work with free energy density:
-\begin{align}
-  f(T,n) &\equiv \frac{F(T,n)}{V}
-\end{align}
-
-\subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
-Coexistence curves are found by modifying the chemical potential $\mu$
-in the grand free energy per volume $\phi(T,n)/V$. Grand free energy
-density is defined as
-\begin{align}
-  \frac{\Phi}{V}(T,n) &= \phi(T,n) \nonumber \\
-                 &= f(T,n) - n\mu \nonumber \\
-                 &= f(T,n) - n\frac{df}{dn}(T,n)\bigg|_{\npart}\ .
-\end{align}
-$\npart$ is some particular density; it is this value that we adjust to find coexistence curves.
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/SW-phi-lowT}
-  \caption{Grand free energy per unit volume, square well, low temperature}
-  \label{fig:SW-phi-lowT}
-\end{figure}
-
-The grand free energy density generally has two distinct minima when
-plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an
-example.) The liquid will be in liquid-vapor equilibrium when those
-minima have the same value for $\phi(T,n)$. There is a local maximum
-between those minima, and the value of the density at that maximum is
-used for $\npart$. Find this value at a given temperature then plot
-$\phi(T,n)$ vs $n$ at slightly higher temperature. Find the new value
-for $\npart$ such that the two minima have the same value for
-$\phi(T,n)$. [It is helpful to note that if $\npart$ is increased,
-$\phi(T,\nliq)$ increases, and vice-versa.] The program I wrote uses the previous temperature's
-$\npart$ as a ``first guess,'' maximizing $\phi(T,\npart)$. From
-there, the program adjusts $\npart$ until $\phi(T,\nvap) =
-\phi(T,\nliq)$. When that condition is met, it uses this new value of
-$\npart$ as the ``first guess'' at $\npart$ for the next higher
-temperature.
-
-The algorithm is as follows; start at some temperature $T=T_{start}$,
-with an intial guess for $\npart$. This initual guess can be read off
-of a plot, and optimized ``by hand'' with minimal effort.
-\begin{enumerate}
-  \item Calculate $\nvap$ and $\nliq$ by calculating the minima in $\phi(T_{start},n)$ over appropriate regions
-  \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
-  \item Calculate \[\delta\mu = \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\nliq - \nvap}\]
-  \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + \delta\mu$
-  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
-  \item Calculate $\phi(T_{start},\npart^*)$
-  \item Re-caluclate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
-  \item Go back to step~\ref{while-start} and repeat until \[ \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\phi(T_{start},\npart^*)} > tol  \] where $tol$ is some computational tolerance (I used $tol=10^{-5}$)
-  \item Increase the temperature to $T = T + \delta T$ and repeat
-\end{enumerate}
-
-\begin{figure}
-  \begin{center}
-  \includegraphics[width=\columnwidth]{figs/coexistance_SW}
-  \end{center}
-  \caption{Liquid-vapor coexistance for a square well liquid.}
-  \label{fig:coexistance_SW}
-\end{figure}
-
-Figure~\ref{fig:coexistance_SW} is a demonstration of this algorithm. \fixme{Compare this one to something.}
-
-\subsection{Free Energies}\label{subsec:free-energies}
-To minmize calculation of integrals, we follow Gil-Villegas' approximations to the $a_j(n)$ terms.
-\newcommand\eff{\textit{eff}}
-\begin{align}
-  a_1^\text{SW}(n) &= a_1^\text{VDW}(n)g^\text{HS}(1;\eta_{\eff}) \nonumber \\
-  &= -4\eta\varepsilon(\lambda^3 - 1)\frac{1 - \left( \eta_{\eff}/2 \right)}{(1 - \eta_{\eff})^3}
-\end{align}
-and
-\begin{align}
-  \eta_{\eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
+  \eta_\textit{eff} &= c_1\eta + c_2\eta^2 + c_3\eta^3
 \end{align}
@@
     1 \\
-    \lambda \\
-    \lambda^2
+    \lambdaSW \\
+    \lambdaSW^2
     \end{array} \right)
@@
 
+%% In Gil-Villegas' paper,\cite{gil-villegas97} HS and Dispersion are wrapped up into the
+%% \textit{monomer} contribution, expressed in terms of energy densities
+%% as:
+%% \begin{align}
+%%   f_\text{mono} = f_\text{HS} - \frac{\alpha^\text{VDW}n}{kT}
+%% \end{align}
+%% with $\alpha^\text{VDW}$ given by
+%% \begin{align}
+%%   \alpha^\text{VDW} &= 2\pi\varepsilon\int_\sigma^\infty r^2\phi(r)\,dr \\
+%%   \intertext{using reduced units of $x = r/\sigma$ (\textsc{N.B.}~this is unrelated to the RGT amplitude of fluctuations $x$ from section~\ref{subsec:RGT}), we have}
+%%   &= 2\pi\sigma^3\varepsilon\int_1^\infty x^2\phi(x)\,dx \nonumber \\
+%%   &= 3b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)\,dx
+%% \end{align}
+%% where $b^\text{VDW}$ is the van der Waals size parameter. It
+%% corresponds to the volume excluded by two spherical particles of
+%% volume $b$: $b^\text{VDW} = 4b = 4\left(\pi\sigma^3/6\right)$.
+
+%% Then, the high temperature expansion is an expansion of the monomer term:
+%% \begin{align}
+%%   f_\text{mono} &= f_\text{HS} + \beta a_1 + \beta^2 a_2 + \mathcal{O}(\beta^3)
+%% \end{align}
+
+%% The term $a_1$ is given by
+%% \begin{align}
+%%   a_1 &= -2\pi n \varepsilon\int_\sigma^\infty r^2\phi(r)g^\text{HS}(r)\,dr \\
+%%   &= -3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\phi(x)g^\text{HS}(x)\,dx
+%% \end{align}
+%% If we assume random correlations between the particles' positions, for
+%% all distances, we have $g^\text{HS}(r) = 1$. This yields \fixme{derive this---or at least explain a little bit}
+%% \begin{align}
+%%   a_1^\text{VDW} = - n \alpha^\text{VDW},
+%% \end{align}
+%% the van der Waals mean-field energy.
+
+The second-order term is quite a bit more complicated. As discussed in
+section~\ref{subsec:TPT}, one must have knowledge of higher-order
+correlation functions to determine $a_2$. This term describes the
+response of the attractive energy due to the compression of the fluid
+from the attractive well. Gil-Villegas \emph{et al.} base their
+expression on an approximation by Barker and Henderson.\cite{Barker67}
+Barker and Henderson's approximation, called the Local Compressibility
+Approximation (LCA), considers fluctuations in the number of particles
+in the potential well. In this way, the fluctuations in $a_2$ are
+related to the pressure and compressibility of the liquid. Given
+pressure $P^\text{HS}$ and isothermal compressibility $K^\text{HS} =
+kT\left(\partial n /\partial P^\text{HS}\right)_T$, we have:
+\begin{align}
+  a_2 &= -\pi n \varepsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \nonumber \\
+  &= -\pi n \varepsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \nonumber \\
+  &= \frac{1}{2}\varepsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+\end{align}
+
+We use the isothermal compressibility in terms of packing fraction
+$\eta$ (see equation~\ref{eqn:packingFrac}), given by the Percus-Yevick expression:\cite{Barker76}
+\begin{align}
+  K^\text{HS} &= \frac{\left(1 - \eta\right)^4}{1 + 4\eta + 4\eta^2}
+\end{align}
+
+\subsection{Renormalization Group Theory}\label{subsec:RGT}
+Renormalization group theory (RGT) is a recursive procedure that was
+developed to deal with large-scale fluctuations of the density
+near the critical point of a system. The formulation used here is
+provided by Forte~\textit{et al}.\cite{Forte11}
+
+We begin by dividing the total volume into cells of volume $V_i$. We
+assume that within a cell the number density $n(\rr)$ varies only slightly, and that the
+cells do not interact with one another. This approach is referred to as Wilson's
+phase cell method.\cite{Ramana12}
+
+Then consider a baseline free energy $F_0(T,n)$, which we will take
+from TPT. We assume this baseline accounts for all fluctuations of
+wavelength $ < \lambda_0$. We then double the size of our phase cell
+and add a correction to the free energy $\delta F_1(T,n)$. This
+correction accounts for fluctuations with a wavelength up to
+$2\lambda_0$. The new free energy, $F_1(T,n)$, is
+\begin{align}
+  F_1(T,n) &= F_0(T,n) + \delta F_1(T,n)
+\end{align}
+Then double the length scale again; $\delta F_2(T,n)$ accounts for
+fluctuations up to $4\lambda_0$ and
+\begin{align}
+  F_2(T,n) &= F_1(T,n) + \delta F_2(T,n)
+\end{align}
+As we continue this process, the $i^{th}$ correction accounts for
+fluctuations of $2^i\lambda_0$, and
+\begin{align}
+  F_i(T,n) &= F_{i-1}(T,n) + \delta F_i(T,n)
+\end{align}
+In the end, the free energy is
+\begin{align}
+  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \Fattr
+\end{align}
+where $\Fattr$ is the contribution due to the longest-wavelength
+density fluctuations. At this stage, I do not entirely understand why
+Forte separates out $\Fattr$.
+
+The $\delta F_i(T,n)$ term is a difference in free energies:
+\begin{align}
+  \delta F_i(T,n) &= -\kT\ln Z - \left( -\kT\ln Z^* \right) \nonumber \\
+  &= -\kT\ln\left[ \frac{Z(T,n)}{Z^*(T,n)} \right]
+\end{align}
+where $Z$ is the partition function for the entire system, and $Z^*$
+is the partition function of a reference system consisting only of the
+term $\FbarD$. Forte gives these terms as
+\begin{multline}
+  Z(T,n,x) = \\ \int dx\, \exp\left[-\beta V\left( \FbarD + \UbarD \right)\right] \label{eqn:Z} \\
+\end{multline}
+\begin{multline}
+  Z^*(T,n,x) = \int dx\, \exp\left[ -\beta V\FbarD \right] \label{eqn:Zstar}
+\end{multline}
+where $x$ is the amplitude of density fluctuations. $\FbarD$ and $\UbarD$ are formulated similar to averages:
+\begin{align}
+  \FbarD &= \frac{F_{i-1}(T,n+x) + F_{i-1}(T,n-x)}{2} - F_{i-1}(T,n)
+\end{align}
+\begin{multline}
+  \UbarD = \frac{U(T,n+x,\lambdaRG) + U(T,n-x,\lambdaRG)}{2} \\
+  - U(T,n,\lambdaRG)
+\end{multline}
+\FbarD represents the free energy change due to fluctuations if interactions between cells are neglected, and \UbarD is the contrigution of the interactions between cells.
+Finally, $U(T,n,\lambdaRG)$ is given by Forte as
+\begin{multline}
+  U(T,n,\lambdaRG) = g_\sigma^\text{HS}(\neff)\alpha(\lambdaRG) \\
+  - \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)\alpha(\lambdaRG)\omega^2(\lambdaRG) \\
+  + \left( \frac{4\pi^2}{2^{2i+1}L^2} \right)^2\alpha(\lambdaRG)\gamma(\lambdaRG) \label{eqn:URG} % URG! It sounds like it hurt! :P
+\end{multline}
+with
+\begin{align}
+  \alpha(\lambdaRG) &= \frac{2}{3}\pi\epsilon\sigma^3(\lambdaRG^3 - 1) \\
+  \omega^2(\lambdaRG) &= \frac{1}{5}\sigma^2\frac{\lambdaRG^5 - 1}{\lambdaRG^3 - 1} \\
+  \gamma(\lambdaRG) &= \frac{1}{70}\sigma^4\frac{\lambdaRG^7 - 1}{\lambdaRG^3 - 1}
+\end{align}
+The pair correlation function $g_\sigma(n)$ tells us how likely we
+are to find two hard spheres touching, given some density. In
+equation~\ref{eqn:URG} we are considering this quantity with an
+effective density \neff.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}\label{sec:methods}
+
+\subsection{Coexistence Curve Algorithm}\label{subsec:coexis}
+Coexistence curves are found by modifying the chemical potential $\mu$
+in the grand free energy per volume $\phi(T,n)/V$. Grand free energy
+density is defined as
+\begin{align}
+  \frac{\Phi}{V}(T,n) &= \phi(T,n) \nonumber \\
+                 &= f(T,n) - n\mu \nonumber \\
+                 &= f(T,n) - n\frac{df}{dn}(T,n)\bigg|_{\npart}\ .
+\end{align}
+$\npart$ is some particular density, which defines the chemical
+potential; it is this value that we adjust to find coexistence curves.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.5\textwidth]{figs/SW-phi-lowT}
+  \caption{Grand free energy per unit volume, square well, low temperature}
+  \label{fig:SW-phi-lowT}
+\end{figure}
+
+The grand free energy density below the critical temperature has two distinct minima when
+plotted versus density. (See fig.~\ref{fig:SW-phi-lowT} for an
+example.) The fluid will be in liquid-vapor equilibrium when those
+minima have the same grand free energy. I have designed an algorithm
+that adjusts \npart\ (that is, adjusts $\mu$) until the system is in
+liquid-vapor equilibrium.  The algorithm is as follows; start at some
+temperature $T=T_{start}$, with an initial guess for $\npart$. This
+initial guess can be read off of a plot, and optimized ``by hand''
+with minimal effort.
+\begin{enumerate}
+  \item Calculate $\nvap$ and $\nliq$ by calving for the minima in $\phi(T_{start},n)$ over appropriate regions
+  \item Calculate $\phi(T_{start},\nvap)$ and $\phi(T_{start},\nliq)$ \label{while-start}
+  \item Calculate an estimated change in chemical potential using the slope \[\delta\mu = \frac{\phi(T_{start},\nvap) - \phi(T_{start},\nliq)}{\nliq - \nvap}\]
+  \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + n\delta\mu$
+  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
+  \item Re-calculate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
+  \item Go back to step~\ref{while-start} and repeat until the grand free energy at liquid and vapor densityies are sufficiently similar. I used a dimensionless computation tolerance of 10$^{-5}$.
+  \item Increase the temperature to $T = T + \delta T$ and repeat using the final $\npart^*$ as our initial guess for the dividing point between the liquid and vapor free energy minima
+\end{enumerate}
+
+\begin{figure}
+  \begin{center}
+  \includegraphics[width=0.5\textwidth]{figs/coexistance_SW}
+  \end{center}
+  \caption{Liquid-vapor coexistence for a square well liquid.}
+  \label{fig:coexistance_SW}
+\end{figure}
+
+Figure~\ref{fig:coexistance_SW} is a demonstration of this algorithm.
+
 \subsection{RG partition functions}\label{subsec:fbar-ubar}
-I use the midpoint method for evaluating the integrals $Z$ and
-$Z^*$ (equations~\ref{eqn:Z} and~\ref{eqn:Zstar}). This method was
-chosen for ease of implementation and low comuptational cost. The approximation is
+I use the midpoint method for evaluating the integrals $Z$ and $Z^*$
+(equations~\ref{eqn:Z} and~\ref{eqn:Zstar}). This method was chosen
+for ease of implementation and to avoid problematic endpoint
+cases. The approximation is
 \begin{align}
@@
 
-For this work, I use $\lambdaSW = 1.5$.
-
 \begin{figure}
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/phi-RG-lowT}
+  \includegraphics[width=0.5\textwidth]{figs/phi-RG-lowT}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/phi-RG-highT}
+  \includegraphics[width=0.5\textwidth]{figs/phi-RG-highT}
   \end{center}
@@
   \begin{center}
-  \includegraphics[width=\columnwidth]{figs/coexistence-RG-forte}
+  \includegraphics[width=0.5\textwidth]{figs/coexistence-RG-forte}
   \end{center}
-  \caption{Liquid-vapor coexistence, comparted with Forte's 2011 data}
+  \caption{Liquid-vapor coexistence, compared with Forte's 2011 data}
   \label{fig:coexistence-RG-forte}
@@
 grand free energy per unit volume at a low temperature and near the
-critical point, respectively. Results are not particularly clear at
+critical point, respectively, for a value of $\lambdaSW = 1.5$. Results are not particularly clear at
 this point, as the program has not been fully debugged. This is
@@
 calculations for the square well liquid. The program is not entirely
-de-bugged. I have also developed an algorithm that calculates
-liqiud-vapor coexistence curves, even close to the critical point.
+debugged. I have also developed an algorithm that calculates
+liquid-vapor coexistence curves, even close to the critical point.
 
-We do not yet know if RGT is worth implementing. Since the program has
+We do not yet know if RGT is worth implementing in DFT. Since the program has
 not been fully debugged, I am unable to say what the computational
-cost would be to obtain an accurate liquid-vapor coexistene curve.
+cost would be to obtain an accurate liquid-vapor coexistence curve.
 

### 2014-06-06 papers/thesis-roth/project.tex
figs should work now Tested on elliot. Took 0:18.62.
project report; close to final draft Tested on elliot. Took 0:18.53.

@@
 \documentclass[letterpaper,twocolumn,amsmath,amssymb,prb]{revtex4-1}
-\usepackage[demo]{graphicx}% Include figure files. When building at home, use the option [demo] to supress actual need of figures
+\usepackage{graphicx}% Include figure files. When building at home, use the option [demo] to supress actual need of figures
 \usepackage{dcolumn}% Align table columns on decimal point
@@
 % RG stuff
-\newcommand{\FbarD}{\ensuremath{\bar{F}_D(T,n,x)}}
-\newcommand{\UbarD}{\ensuremath{\bar{U}_D(T,n,x,\lambdaRG)}}
+\newcommand{\FbarD}{\ensuremath{\bar{F}(T,n,x)}}
+\newcommand{\UbarD}{\ensuremath{\bar{U}(T,n,x,\lambdaRG)}}
 
@@
 However, understanding liquids has both philosophical and economic
-desireability. Liquid water is---as far as we know---essential for
+desirability. Liquid water is---as far as we know---essential for
 developing and sustaining life. All cellular processes take place in an
@@
   \hookrightarrow \Fex &= F_\text{ref} + \left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2) \label{eqn:F-inTermsW} \\
-  &\equiv F_\text{ref} + a_1(n) + a_2(n)
+  &= F_\text{ref} + a_1(n) + \beta a_2(n) + \mathcal{O}(\beta^2) \label{eqn:hi-T-exp}
 \end{align}
-This is the \emph{high-temperature expansion}, first derived by
+with
+\begin{align}
+  a_1(n) &= \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta\Vtilde} \\
+  a_2(n) &= -\frac{\sigma_{\xi=0}^2}{2}
+\end{align}
+Equation~\ref{eqn:hi-T-exp} is the \emph{high-temperature expansion}, first derived by
 Zwanzig\cite{Zwanzig54}. He further showed that the $n^{th}$ term could be written in
@@
 section~\ref{subsec:TPT}, one must have knowledge of higher-order
-correlation functions to determine $a_2$. This term describes the
+correlation functions to exactly determine $a_2$. This term describes the
 response of the attractive energy due to the compression of the fluid
@@
 \begin{align}
-  a_2 &= -\pi n \varepsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \nonumber \\
-  &= -\pi n \varepsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \nonumber \\
-  &= \frac{1}{2}\varepsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
+  a_2 &\approx \frac{1}{2}\varepsilon n K^\text{HS}\frac{da_1}{dn}
+  %% a_2 &\approx -\pi n \varepsilon^2kT\int_0^\infty r^2\left[\phi(r)\right]^2\frac{\partial n  g^\text{HS}}{\partial P^\text{HS}}(r)\,dr \nonumber \\
+  %% &= -\pi n \varepsilon^2K^\text{HS}\frac{\partial}{\partial n }\left[\int_\sigma^\infty r^2\left[\phi(r)\right]^2 n  g^\text{HS}\,dr\right] \nonumber \\
+  %% &= \frac{1}{2}\varepsilon K^\text{HS}\frac{\partial}{\partial n }\left[-3 n  b^\text{VDW}\varepsilon\int_1^\infty x^2\left[\phi(x)\right]^2 g^\text{HS}(x)\,dx \right]
 \end{align}
@@
 \begin{align}
-  F(T,n) = F_0(T,n) + \sum_{i=1}^\infty\delta F_i(T,n) + \Fattr
+  F(T,n) = F_0(T,n) + \Fattr + \sum_{i=1}^\infty\delta F_i(T,n)
 \end{align}
@@
 \end{multline}
-\FbarD represents the free energy change due to fluctuations if interactions between cells are neglected, and \UbarD is the contrigution of the interactions between cells.
+\FbarD represents the free energy change due to fluctuations if interactions between cells are neglected, and \UbarD is the contribution of the interactions between cells.
 Finally, $U(T,n,\lambdaRG)$ is given by Forte as
@@
   \item Calculate $\phi^*(T_{start},n) = \phi(T_{start},n) + n\delta\mu$
-  \item Calculate a new $\npart$, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
-  \item Re-calculate $\nvap$, $\nliq$, $\phi(T_{start},\nvap)$, and $\phi(T_{start},\nliq)$, using $\npart^*$
-  \item Go back to step~\ref{while-start} and repeat until the grand free energy at liquid and vapor densityies are sufficiently similar. I used a dimensionless computation tolerance of 10$^{-5}$.
+  \item Calculate a new $n$ particular value, $\npart^*$, based on $\phi^*(T_{start},n)$ by finding the local maximum between $\nvap$ and $\nliq$
+  \item Recalculate the minima in the grand free energy using $\npart^*$ to find an improved estimate for \nvap\ and \nliq
+  \item Go back to step~\ref{while-start} and repeat until the grand free energy at liquid and vapor densities are sufficiently similar. I used a dimensionless computation tolerance of 10$^{-5}$.
   \item Increase the temperature to $T = T + \delta T$ and repeat using the final $\npart^*$ as our initial guess for the dividing point between the liquid and vapor free energy minima

### 2014-06-11 papers/pair-correlation/paper.tex
work on pairs paper, add todos Tested on bennet. Took 0:18.96.

@@
 \section{Introduction}
+
+\fixme{replace approximation with approach}
 
@@
 
-\section{Previous theoretical approaches}
+\fixme{Somewhere above introduce the question of whether g2 is
+  expressed in terms of V or n}
+
+\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}
+
+\fixme{introduce this as methods given V}
 
@@
 find the pair distribution function~\cite{hansen2006theory}.  This
-elegant approach lends itself to DFT, and can be used to compute and
+elegant approach lends itself to computation with DFT, and can be used to compute and
 plot the pair distribution function, but requires a full free-energy
@@
 system using integral equation theory under the Percus-Yevick
-approximation~\cite{plischke1986pair}.  Ten years later,
+approximation~\cite{plischke1986pair}.  \fixme{Move the following to
+  the next section} Ten years later,
 G{\"o}tzelmann \emph{et al.} found the pair distribution function near
@@
 DFT free energy functional~\cite{gotzelmann1996structure}.
-Unfortunately, solving the inhomogeneous Ornstein-Zernike equation is
-computationally challenging, although more efficient approximate
-algorithms have been developed~\cite{paul2003variational}.
+Unfortunately, solving the inhomogeneous Ornstein-Zernike equation
+remains computationally challenging, although more efficient
+approximate algorithms have been developed~\cite{paul2003variational}.
+This approach, while appealing, remains unsuitable for use in the
+construction of a classical density functional due to its significant
+computational cost.
 
@@
 
-\section{Previous approximations to the pair distribution function}
+\section{Previous approximations to the pair distribution function
+  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}
+
+\fixme{introduce this as methods given n}
+
 There are several analytic approximations for the inhomogeneous pair
@@
   g^{(2)}(\rr_1,\rr_2) &\approx
-  g\left(r_{12};n\left(\frac{\rr_1+\rr_2}{2}\right)\right)
+  g\left(r_{12}, n\left(\frac{\rr_1+\rr_2}{2}\right)\right)
      & \text{midpoint} \\
   g^{(2)}(\rr_1,\rr_2) &\approx
-  g\left(r_{12};\frac{n(\rr_1)+n(\rr_2)}{2}\right)
+  g\left(r_{12}, \frac{n(\rr_1)+n(\rr_2)}{2}\right)
      & \text{mean density} \\
@@
 
+\davidsays{Please edit!}\fixme{Note that the HS free energy is used
+  for two purposes when applying Eq.~\ref{eq:gsigma-exact} : it is
+  minimized (together with an external potential) to find the density
+  $n(\rr)$.  Then its functional derivative is used to find
+  $g_\sigma(\rr)$.  In application of perturbation theory, the HS
+  functional would only be used alone for the latter purpose, as the
+  density would be determined by the total free energy, which
+  incorporates both the HS free energy and a perturabative
+  contribution.}
+
 \section{Contact value approximation}
@@
     our separable fit.  The solid lines show the true radial
-    distribution function $g(r)$, and the dashed lines show our
-    separable fit to the radial distribution functions
+    distribution function $g(r)$, as found from Monte Carlo
+    simulation, and the dashed lines show our separable fit to the
+    radial distribution functions
     $g_S(r)$.}\label{fig:radial-distribution}
@@
 \end{equation}
+where the dimensionless isothermal compressibility $\chi$ is obtained
+from the Carnahan-Starling equation of state.
 Finally, we can convert between the packing fraction $\eta$ and the
@@
 
+\subsection{Efficiency}
+
+\fixme{Explain why the separable version is efficient.}
+
 \section{Conclusion}

### 2014-06-13 papers/pair-correlation/paper.tex
Added Gil-Villegas approximation to RDF. It needs fixing; there is an m variable, the value of which I could not find. Also, added an alternate version of the RDF plot. Tested on paho-arch. Took 1:13.32.

--- 2014-05-20 papers/pair-correlation/paper.tex
@@
   %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
-  \caption{Plot of the hard-sphere radial distribution function, with
-    our separable fit.  The solid lines show the true radial
-    distribution function $g(r)$, and the dashed lines show our
-    separable fit to the radial distribution functions
+  \caption{Plot of the hard-sphere radial distribution function with our
+    separable fit.  The solid lines show the radial distribution function,
+    $g(r)$, from MC simulations and the dashed lines show our separable fit
+    to the radial distribution function,
     $g_S(r)$.}\label{fig:radial-distribution}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/ghs-alt}
+  \caption{Plot of the hard-sphere radial distribution function. The blue
+    lines show our separable fit, the black dots are Monte Carlo data, and
+    the dashed lines are results of the Gil-Villegas\cite{gil1997statistical}
+    fit.\pahosays{An idea for an alternate version of the radial distribution
+      figure.}}\label{fig:radial-distribution-alt}
 \end{figure}

### 2014-06-13 papers/pair-correlation/paper.tex

--- 2014-06-11 papers/pair-correlation/paper.tex
@@
     $g_S(r)$.}\label{fig:radial-distribution}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/ghs-alt}
+  \caption{Plot of the hard-sphere radial distribution function. The blue
+    lines show our separable fit, the black dots are Monte Carlo data, and
+    the dashed lines are results of the Gil-Villegas\cite{gil1997statistical}
+    fit.\pahosays{An idea for an alternate version of the radial distribution
+      figure.}}\label{fig:radial-distribution-alt}
 \end{figure}

### 2014-06-13 papers/pair-correlation/paper.tex
divy up writing tasks and work on fit Tested on bennet. Took 0:30.70.

@@
 \newcommand{\green}[1]{{\bf \color{green} #1}}
+\newcommand{\blue}[1]{{\bf \color{blue} #1}}
+\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
 \newcommand{\rr}{\textbf{r}}
@@
 \newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
-\newcommand{\jeffsays}[1]{{\color{red} [\green{Jeff:} \emph{#1}]}}
-\newcommand{\pahosays}[1]{{\color{red} [\green{Paho:} \emph{#1}]}}
+\newcommand{\jeffsays}[1]{{\color{red} [\cyan{Jeff:} \emph{#1}]}}
+\newcommand{\pahosays}[1]{{\color{red} [\blue{Paho:} \emph{#1}]}}
 
@@
 
-\fixme{replace approximation with approach}
+\fixme{replace approximation with approach}\jeffsays{I'm stuck with this.}
 
@@
 \fixme{Somewhere above introduce the question of whether g2 is
-  expressed in terms of V or n}
+  expressed in terms of V or n}\davidsays{I'll take a stab at this}
 
@@
 
-\fixme{introduce this as methods given V}
+\fixme{introduce this as methods given V}\davidsays{I'll take a stab
+  at this}
 
@@
 
-\fixme{introduce this as methods given n}
+\fixme{introduce this as methods given n}\davidsays{I'll do this}
 
@@
 convolution kernel depends on both points.  However, this approach has
-been used by two papers to do
-\fixme{what?}~\cite{wadewitz2000application, winkelmann2001liquid}.
+been used by two papers to construct classical density functionals for
+studying the liquid vapor interface~\cite{wadewitz2000application,
+  winkelmann2001liquid}.
 
@@
 
-\davidsays{Please edit!}\fixme{Note that the HS free energy is used
+\davidsays{Please edit!}\jeffsays{I will!}\fixme{Note that the HS free energy is used
   for two purposes when applying Eq.~\ref{eq:gsigma-exact} : it is
@@
 
-\fixme{Explain why the separable version is efficient.}
+\fixme{Explain why the separable version is efficient.}\davidsays{I'll
+  do it.}
 

### 2014-06-14 papers/square-well-liquid/histogram-paper.tex
better tentative title for histogram paper Tested on bennet. Took 0:18.66.
fix wang_landau with Michael's help.  It works.  :) Tested on bennet. Took 10:36.15.

@@
 \begin{document}
-\title{Applying a clever histogram method to the square-well liquid}
+\title{Applying the optimized ensemble histogram method to the square-well liquid}
 
@@
 
+\section{New ideas for algorithms}
+
+\subsection{Finding maximum probability state}
+
+Technically maximum histogram or maximum entropy.
+
+Loop:
+
+run iterations until we see $N$ energy changes, counting the number of
+returns to the initial energy.  If there are sufficient returns, then
+we are at the maximum, where sufficient returns will be determined by
+a random walk (weighted?) where we compute the expected number of
+returns if the walk is not weighted.  We may want to correct to get
+all the way to the top.
+
+If we are not at the top yet, we try again.  Eventually we will be.
+
+\subsection{Finding variance of the histogram}
+
+I assume we first ran the max-entropy/histogram algorithm, which gives
+us a maximum value of $E_0$.  We then zero the histogram.
+
+We run for a short while, and then compute the mean and standard
+deviation of the energy from the histogram: $\bar E$ and $\sigma$.
+While doing this, we track the number of returns to the initial state
+$n$.  Now the uncertainty in the mean is given by
+\begin{equation}
+  \Delta \bar E \approx \frac{\sigma}{\sqrt{n}}
+\end{equation}
+if $\Delta \bar E \ll (E_0 - \bar E)$ then we can conclude that we are
+certain where the maximum is, and how wide it is.
+
+\subsubsection{Counting frequency of returns!}
+If we are at a local maximum probability, then the frequency of return
+should exceed the zero-bias frequency of return.  If we can figure out
+the zero-bias frequency of return, then when we have a statistically
+significant deviation, we should know what the variance is.
+
+Once we have the variance, we can add to the weights an appropriate
+gaussian to eliminate this maximum in probability.
+
+If the rate of return is significantly smaller than expected for zero
+bias, we conclude this was not actually a maximum, and we go back to
+try again for a probability maximum.  If we fail twice, then maybe
+there is no significant maximum.  There are subtleties here!
+
+\subsection{Pushing down the bubbles again and again}
+Once we've gotten a new set of weights (with the bubble pushed down),
+we can find a new max probability, and then a new variance and push
+down yet another bubble.  In each case, the ``probability'' that we
+are maximizing and finding the variance of is the ratio between the
+density of states and the weight function, so its log is the
+difference between the logs of the density of states and the weight
+function.
+
 \end{document}

### 2014-06-14 src/Monte-Carlo/square-well-monte-carlo.cpp

@@ -93,10 +93,11 @@ int main(int argc, const char *argv[]) {
 
   double wl_factor = 1;
   double wl_fmod = 2;
-  double wl_threshold = 0.1;
-  double wl_cutoff = 0.1;
+  double wl_threshold = 0.05;
+  double wl_cutoff = 1e-5;
 
   double len[3] = {1, 1, 1};
+
   int walls = 0;
   int wall_dim = 1;
   unsigned long int seed = 0;
@@ -506,14 +507,7 @@ int main(int argc, const char *argv[]) {
       energy_histogram[interactions]++;
 
       if(wang_landau && iteration > first_weight_update){
-        ln_energy_weights[interactions] -= wl_factor
-          - log(1-1/(energy_histogram[interactions]+1e-10));
-        // the first term is from wang-langau
-        // the second term is to correct for the fact that wang-landau uses
-        //   a ratio of DoS rather than a ratio of weights to calculate the
-        //   probability of accepting a move
-        // the 1e-10 factor is added for the case energy_histogram[interactions] == 1;
-        //   its effect should otherwise be negligibly small
+        ln_energy_weights[interactions] -= wl_factor;
       }
       if(walker_weights){
         walkers_total[interactions]++;
@@ -814,8 +808,7 @@ int main(int argc, const char *argv[]) {
       fprintf(w_out, "\n# interactions   ln(weight)\n");
       for(int i = 0; i < energy_levels; i++)
         if(energy_histogram[i] != 0){
-          fprintf(w_out, "%i  %g\n",i,ln_energy_weights[i]
-                  + (wang_landau ? log(energy_histogram[i]) : 0));
+          fprintf(w_out, "%i  %g\n",i,ln_energy_weights[i]);
         }
       fclose(w_out);
 

### 2014-06-16 papers/thesis-roth/slides-v2.tex

\documentclass[xcolor=dvipsnames]{beamer}
%\usepackage{beamerthemelined}
\usepackage{amsfonts,amssymb,amsmath}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{color}

\setbeamertemplate{caption}{\insertcaption} % remove the word 'figure' before captions

\usetheme{Frankfurt}
%% \useinnertheme{rounded}
%% \useoutertheme{infolines}
\usecolortheme{dolphin}
\usecolortheme[named=OliveGreen]{structure}
\beamertemplatenavigationsymbolsempty
%% \setbeamertemplate{footline}[page number]

%% Citations
\newcommand{\myciteetal}[3]{{\tiny #1 \emph{et al}, \textit{#2} (#3)}}
\newcommand{\mycite}[3]{{\tiny #1, \textit{#2} (#3)}}

%% n values
\newcommand{\npart}{\ensuremath{n_{part}}}
\newcommand{\nvap}{\ensuremath{n_{vap}}}
\newcommand{\nliq}{\ensuremath{n_{liq}}}

%% 1/kT
\newcommand{\kT}{\ensuremath{k_{B}T}}

%% Lambda values
\newcommand{\lambdaSW}{\ensuremath{\lambda_\text{sw}}}
\newcommand{\lambdaRG}{\ensuremath{\lambda_\text{rg}}}

%% Free energy
\newcommand{\Fex}{\ensuremath{F_\text{ex}}}

%% Vectors
\newcommand{\rr}{\ensuremath{\mathbf{r}}}
\newcommand{\pp}{\ensuremath{\mathbf{p}}}

% TPT
\newcommand{\III}{\ensuremath{\textbf{r}_{12}}} % Roman numerals for \12
\newcommand{\Vtilde}{\ensuremath{\widetilde{V}}}
\newcommand{\Wtilde}{\ensuremath{\widetilde{W}}}

%% fixme
\newcommand{\fixme}[1]{\textcolor{red}{\textbf{[#1]}}}

%-----------------------------------------
% TITLE
%-----------------------------------------

\title[Applying RGT to SW liquid]{Applying Renormalization Group Theory to the Square Well Liquid}
\author[Roth \& Roundy]{Dan Roth}
\institute[OSU]{Oregon State University}
\date{12 June 2014}

\begin{document}

\begin{frame}
  \titlepage
\end{frame}

%-----------------------------------------
% TOC
%-----------------------------------------
\section{}
\begin{frame}{Overview}
  \tableofcontents
\end{frame}

%-----------------------------------------
% INTRO
%-----------------------------------------
\section{Introduction}

\subsection{} %% Liquid
\begin{frame}{What is a liquid?}
  \framesubtitle{Simple definition}
  \begin{block}{Liquid}
    Something that takes the shape of its container, but does not necessarily fill the container
  \end{block}
  \begin{itemize}
    \item Pretty good, but a bit too simple $\rightarrow$ some things could be classified as a liquid that should not be
  \end{itemize}
  \onslide<2->{
    \begin{center}
      \includegraphics[width=0.4\textwidth]{figs/cat-liquid-2} \\
%%      \tiny{\url{http://newauthorpublishing.wordpress.com/2013/08/06/895}}
    \end{center}
    }
\end{frame}

\begin{frame}{What is a liquid?}
  \framesubtitle{More advanced definition}
  \begin{block}{Liquid}
    Phase of matter where neither energy nor entropy are dominant
  \end{block}

  We use perturbation theory to investigate liquids:
  \begin{itemize}
    \item Take an ``easy'' problem we know how to solve
    \item Add a ``difficult'' correction that is small
    \item Construct a power series to solve combined problem
  \end{itemize}
\end{frame}

\begin{frame}{Hard Spheres}
  For the ``easily'' solved problem we use a \emph{hard sphere liquid}:
  \begin{itemize}
    \item Not actually easy to solve $\rightarrow$ never been solved exactly, but good approximations have been made
    \item Spheres cannot overlap (they are ``hard'')
    \item No attraction between spheres
      \begin{itemize}
        \item<1->Simulations are simple
        \item<2->Not actually a liquid
      \end{itemize}
  \end{itemize}
  \begin{center}
    \includegraphics[width=0.75\textwidth]{figs/hard-spheres}
  \end{center}
\end{frame}

\begin{frame}{Square well attraction}
  Attraction is what makes this system a liquid
  \begin{columns}
    \column{0.5\textwidth}
    \begin{itemize}
      \item Each molecule is a sphere with diameter $\sigma$
      \item Infinite potential when two spheres are touching
      \item Constant attractive potential $-\varepsilon$ for some distance $\lambdaSW\sigma$
      \item Zero potential to infinity (and beyond)
    \end{itemize}

    \column{0.6\textwidth}
    \begin{center}
      \includegraphics[width=\columnwidth]{figs/SW-schematic-buzz}
    \end{center}
  \end{columns}
%%   There seem to be infinite forces at the boundaries, but forces play no role in thermodynamic averages
\end{frame}

\subsection{} %% Some basic Stat Mech.
\begin{frame}{Statistical Mechanics Basics}
  \begin{block}{Partition Function}
    \begin{itemize}
      \item Sum up the \emph{Boltzmann factor} of every possible microstate of the system
    \end{itemize}
     \[ Z = \sum_i^\text{\tiny{microstates}}e^{-\beta E_i} \quad\quad  \beta \equiv \frac{1}{\kT} \]
  \end{block}

  We can also write this as an integral:
  \begin{align*}
    Z &= \frac{1}{h^{3N}N!}\int d\rr^N d\pp^N\, \exp\left( -\frac{\beta}{2m}\sum p^2 - \beta U(r) \right) \\
    &= \frac{1}{h^{3N}N!}\int d\pp^N\, \exp\left( -\frac{\beta}{2m}\sum p^2 \right)\int d\rr^N\, \exp(-\beta U(r)) \\
    &= Z_\text{id}Z_\text{ex}
  \end{align*}

\end{frame}

\begin{frame}{Statistical Mechanics Basics}
  \begin{block}{Helmholtz free energy}
    \begin{itemize}
      \item Know Helmholtz $\rightarrow$ figure out anything else you want to know
      \item Useful for fixed temperature and particle number
    \end{itemize}
    \[ F = U - TS = -k_BT\ln Z \]
  \end{block}

  \begin{block}{Grand free energy}
    \begin{itemize}
      \item Number of particles not fixed
      \item Calculate from Helmholtz
    \end{itemize}
    \[ \varPhi = F - N\mu \]
  \end{block}

  \begin{block}{Chemical potential $\mu$}
    Energy cost that the environment must pay to add a particle to the system
  \end{block}
\end{frame}

\subsection{} %% Critical point

\begin{frame}{Critical Point}
  \begin{center}
    \includegraphics[width=0.7\textwidth]{figs/phase-diagram-zoom}

    Phase diagram for water
  \end{center}

  \begin{itemize}
    \item Lines are where two phases can exist in equilibrium
  \end{itemize}

\end{frame}

\begin{frame}{Density fluctuations}
  \begin{columns}[T]
  \column{0.6\textwidth}
  \includegraphics<1>[width=\columnwidth]{~/deft/papers/renormalization/figs/fluctuations}
  \includegraphics<2>[width=\columnwidth]{figs/critical-opalescence}

  \onslide<2>{Size of fluctuations comparable with wavelength of light near the critical point}

  \column{0.4\textwidth}
  \begin{itemize}
    \item Take a snapshot in time
    \item At the critical point, there are fluctuations in density at all length scales
    \item Makes perturbation theories fall apart
    \item Visible effect: critical opalescence
  \end{itemize}
  \end{columns}
\end{frame}

\subsection{} %% Motivation for this work
\begin{frame}{Motivation}
  \begin{block}{Importance of critical point}
  \begin{itemize}
  \item Supercritical fluids used in many industrial processes, such as
    \begin{itemize}
    \item Extraction of caffeine from coffee beans \\ (very relevant in PNW)
    \item Extraction of hops resins for beer production \\ (also relevant in PNW)
    \item Decomposing biomass for hydrocarbon fuel production
    \item \ldots and many more
    \end{itemize}
  \end{itemize}
  \end{block}

  \begin{itemize}
    \item Using TPT and RGT should give us a good way of studying the critical point of a liquid
    \item We start with the square well liquid because it is a widely-used approximation to simple liquids
    \item We want to see if implementing RGT would be worthwhile to this research group
    \begin{itemize}
      \item Other groups use RGT, but don't mention computational cost
    \end{itemize}
  \end{itemize}
\end{frame}

%-----------------------------------------
% TPT
%-----------------------------------------
\section[TPT]{Thermodynamic Perturbation Theory}
\subsection{}

\begin{frame}{Quantum Perturbation Theory}
  \begin{block}{Hamiltonian}
    Sum of kinetic and potential energy terms
  \end{block}
  \begin{enumerate}
    \item Start with a Hamiltonian we can solve analytically, $\widetilde{H}_0$
    \item Add a small change, $\xi \Wtilde$
    \item The complete Hamiltonian of the system is then: $ \widetilde{H} = \widetilde{H}_0 + \xi \Wtilde $
    \item Apply to the Schroedinger equation: $ \widetilde{H}|n\rangle = E|n\rangle \Rrightarrow (\widetilde{H}_0 + \xi \Wtilde)|n\rangle = E|n\rangle $
    \item For small values of $\xi$, we can expand $E$ and $|n\rangle$ as power series: \[ E_n = E_n^{(0)} + \xi E_n^{(1)} + \xi^2E_n^{(2)} + \cdots \] and \[ |n\rangle = |n^{(0)}\rangle + \xi|n^{(1)}\rangle + \xi^2|n^{(2)}\rangle + \cdots \]
  \end{enumerate}
\end{frame}

\begin{frame}{Thermodynamic Perturbation Theory}
  \begin{itemize}
    \item For TPT we are perturbing the potential only, not the entire Hamiltonian
  \end{itemize}

  \begin{block}{The reference system}
    \begin{itemize}
      \item Hard sphere potential
    \end{itemize}
    \[
      v_\text{hs}(\III) =
      \begin{cases}
        \infty & r \leq \sigma \\
        0 & \sigma < r
      \end{cases}
    \]
  \end{block}

  \begin{block}{Perturbation}
    \begin{itemize}
      \item Square well attraction
    \end{itemize}
    \[
      w_\text{sw}(\III) = - \varepsilon \quad  \sigma < r < \lambdaSW\sigma
    \]
  \end{block}

  \begin{block}{}
  \[ v(\III) = v_\text{hs}(\III) + w_\text{sw}(\III) \]
  \end{block}
\end{frame}

\begin{frame}{Excess Free Energy}
  The total potential energy of our system is \[ \Vtilde = \sum_{i=1}^N\sum_{j>i}^N \left( v_\text{hs}(\rr_{ij}) + \xi w_\text{sw}(\rr_{ij}) \right) \]
  The \emph{configuration integral} is the potential energy component of the partition function: \[ Z_\xi = \frac{1}{V^N}\int d\rr^N\, e^{-\beta \Vtilde} \] where volume to the $N^{th}$ power has shown up for normalization.
  From the configuration integral we can get the excess (from ideal) free energy:\[ \Fex = -\kT\ln Z_\xi \]
  \mycite{Hansen \& McDonald}{Academic Press}{2006}
\end{frame}

\begin{frame}{Derivative}
  It is easier, and no less exact, to work with the derivative of \Fex:
    \begin{align*}
    \frac{\partial \Fex}{\partial\xi} &=-\kT\frac{\partial}{\partial\xi}\left( \ln Z_\xi \right)  \\
    &\vdots \\
    &= \frac{1}{Z_\xi} \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \\
    &= \left\langle \Wtilde \right\rangle_{\xi}
  \end{align*}
  where $\Wtilde = \partial \Vtilde/\partial \xi$ and $\left\langle \Wtilde\right\rangle_{\xi}$ is the ensemble average of $\Wtilde$ for some value of $\xi$
\end{frame}

\begin{frame}{Expansion}
  Now take a series expansion of $\left\langle W_N\right\rangle_{\xi}$ around $\xi = 0$
  \begin{align*}
    \left\langle \Wtilde \right\rangle_{\xi} &= \left\langle \Wtilde\right\rangle_{\xi = 0} + (\xi)\frac{\partial\left\langle \Wtilde \right\rangle_{\xi}}{\partial\xi}\bigg|_{\xi = 0} + \mathcal{O}(\xi^2)
  \end{align*}
  And the derivative term is
  \begin{align*}
    \frac{\partial}{\partial\xi}\left\langle \Wtilde \right\rangle_{\xi} &= \frac{\partial}{\partial\xi}\left( \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta \Vtilde}\right)  \\
    &\vdots  \\
    &= -\beta\left[ \left\langle \Wtilde^2 \right\rangle_{\xi} - \left\langle \Wtilde \right\rangle_{\xi}^2 \right]  \\
    &\equiv -\beta\sigma_{\xi}^2 \quad \text{Note that $\sigma^2 > 0 \rightarrow$ this term is always negative}
  \end{align*}
  So we have
  \begin{align*}
    \left\langle \Wtilde \right\rangle_{\xi} &= \left\langle \Wtilde\right\rangle_{\xi = 0} - (\xi)\beta\sigma_{\xi=0}^2 + \mathcal{O}(\xi^2)
  \end{align*}
\end{frame}

%% \begin{frame}{Integrate to get \emph{High Temperature Expansion}}
%%   Limits of integration:
%%   \begin{itemize}
%%     \item $ F \in [F_\text{ref},\Fex] $
%%     \item $ \xi \in [0,1] $
%%   \end{itemize}
%%   \begin{align*}
%%     \Fex - F_\text{ref} &= (1-0)\left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta(1 - 0)^2\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2)  \\
%%     \hookrightarrow \Fex &= F_\text{ref} + \left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2) \\
%%   \end{align*}
%%   \mycite{Robert Zwanzig}{J.~Chem.~Phys.}{1954}
%% \end{frame}

\begin{frame}{High Temperature Expansion}
  Integrate to get back to the free energy
  \begin{align*}
    \Fex &= F_\text{ref} + \left\langle \Wtilde \right\rangle_{\xi=0} - \frac{1}{2}\beta\sigma_{\xi=0}^2 + \mathcal{O}(\beta^2) \\
    \Fex &= F_\text{ref} + A_1(n) + \beta A_2(n) + \mathcal{O}(\beta^2)
  \end{align*}
  with
  \begin{align*}
    A_1(n) &= \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta\Vtilde} \qquad A_2(n) = -\frac{\sigma_{\xi=0}^2}{2}
  \end{align*}
  \begin{itemize}
    \item $A_1$ is the ``zeroth'' order interaction between molecules
    \item $A_2$ is a bit more complicated; to exactly determine it, we need to know higher-order correlations
    \item We approximate $A_2$ with the local compressibility approximation
  \end{itemize}
  \mycite{Robert Zwanzig}{J.~Chem.~Phys.}{1954}
\end{frame}

%-----------------------------------------
% RGT
%-----------------------------------------
\section[RGT]{Renormalization Group Theory}
\subsection{}

\begin{frame}{Account for fluctuations at all length scales}
  \begin{block}{Basic idea behind TPT}
    Take a baseline and consider small changes in the attractive potential
  \end{block}

  \begin{block}{Basic idea behind RGT}
    Take a baseline and consider not-so-small changes in the free energy
  \end{block}

  RGT is an iterative procedure:
  \begin{itemize}
    \item Start with a baseline free energy---we use thermodynamic perturbation theory (TPT)
    \item Add corrections
  \end{itemize}

  \begin{block}{Historical note}
    Developed in the 1950s; work by Kenneth Wilson in the '70s won him the 1982 Nobel Prize in physics
  \end{block}
\end{frame}

\begin{frame}{Renormalization Group Theory}
  \begin{itemize}
    \item Consider density fluctuations\ldots
    \item Fluctuations of wavelength $\lambdaRG < \lambda_0$ are assumed to be accounted for in the baseline free energy (TPT for this work)
    \item What if we include fluctuations of wavelength $2\lambda_0$? How about $4\lambda_0$? \ldots and so on
    \item Divide up all space into cells; each cell has a volume \[V_\text{cell,i} = \left(\frac{2^i\lambda_0}{2}\right)^3\] where $i$ is the iteration you are on and $\lambda_0$ is the cutoff wavelength
    \item Within a cell, we assume that the density is only slowly varying
  \end{itemize}
\end{frame}

\begin{frame}{Iterations}
  \begin{align*}
    F_1 &= F_0 + \delta F_0  \\
    F_2 &= F_1 + \delta F_1  \\
%%     F_3 &= F_2 + \delta F_2  \\
    &\vdots  \\
    F_i(T,n) &= F_{i-1}(T,n) + \delta F_{i-1}(T,n) \\
    \hookrightarrow F &= F_0 + \sum_i\delta F_i(T,n) \\
    \delta F_i(T,n) &= -k_BT\ln\left( \frac{Z_i}{Z_i^*} \right)
  \end{align*}

  \begin{itemize}
    \item The ratio in the log is nothing more than a free energy difference
    \item $Z_i$ is the partition function for the entire system
    \item $Z_i^*$ is the partition function of a reference system that does not contain fluctuations
  \end{itemize}
\end{frame}

\begin{frame}{Partition function}
  \begin{itemize}
    \item This formulation comes from work done by Esther Forte
  \end{itemize}

  \begin{align*}
    Z_i &= \int_0^nd\nu\, \exp\left[-\beta\left(\bar{F}(T,n,\nu) + \bar{U}(T,n,\nu,\lambdaRG) \right)\right] \\
    Z_i^* &= \int_0^nd\nu\, \exp\left[-\beta\bar{F}(T,n,\nu)\right]
  \end{align*}
  \begin{itemize}
    \item $\nu$ is the amplitude of the fluctuation
    \item $\bar{F}(T,n,\nu)$ is the free energy without interaction between cells
    \item $\bar{U}(T,n,\nu,\lambdaRG)$ is the energy of interaction (between cells)
  \end{itemize}
  \myciteetal{Esther Forte}{J.~Chem.~Phys.}{2011}
\end{frame}

%-----------------------------------------
% Coexistence
%-----------------------------------------
\section[l-v coexistence]{Liquid-Vapor Coexistence}
\subsection{}

\begin{frame}{What is liquid-vapor coexistence?}
  \begin{block}{Grand Free Energy}
     \[ \varPhi = F - N\mu \]
  \end{block}

  \begin{columns}
  \column{0.4\textwidth}
  \begin{itemize}
    \item Coexistence curve tells us where the critical point is
    \item In general, one density is more energetically favorable than another
    \item At coexistence, both states are equally favorable
  \end{itemize}

  \column{0.6\textwidth}
  \includegraphics<1>[width=\columnwidth]{figs/SW-phi-lowT-vapor-liquid-annotation}
  \includegraphics<2>[width=\columnwidth]{figs/SW-phi-lowT-equilibrium-vapor-liquid-annotation}
  \includegraphics<3>[width=\columnwidth]{figs/SW-phi-highT}
  \end{columns}
\end{frame}

\begin{frame}{Example at $\kT = 0.8\varepsilon$}
  \begin{columns}

  \column{0.6\textwidth}
  \begin{center}
    \includegraphics<1>[width=\columnwidth]{figs/fitting-step0-noslope}
    \includegraphics<2-3>[width=\columnwidth]{figs/fitting-step0}
    \includegraphics<4>[width=\columnwidth]{figs/fitting-step1-unzoomed}
    \includegraphics<5-6>[width=\columnwidth]{figs/fitting-step1}
    \includegraphics<7->[width=\columnwidth]{figs/fitting-step2}
  \end{center}

  \column{0.5\textwidth}
  \begin{enumerate}
    \item<1-> Start with a guess for $\mu$
    \begin{itemize}
      \item<2-> Use the slope to tell us how much to adjust $\mu$
    \end{itemize}
    \item<3-> Adjust $\mu$
    \begin{itemize}
      \item<4-> Looks pretty good, right? Let's zoom in\ldots
    \end{itemize}
    \item<6-> We need to change $\mu$ again
    \item<8-> Excellent! Now use this value for $\mu$ as the ``first guess'' at a slightly higher temperature
  \end{enumerate}

  \end{columns}
\end{frame}

%-----------------------------------------
% Result
%-----------------------------------------
\section{Result}

\subsection{}
\begin{frame}{Liquid-vapor coexistence}
  \begin{center}
    \includegraphics[width=0.8\textwidth]{figs/coexistence-RG-forte}
  \end{center}
\mycite{Esther Forte}{J.~Chem.~Phys.}{2011}

\mycite{J.~Richard Elliot}{J.~Chem.~Phys}{1999}
\end{frame}

%-----------------------------------------
% Conclusion/Future/Acknowledgments
%-----------------------------------------
\section{Conclusion}

\subsection{}
\begin{frame}{Conclusion}
  \begin{block}{Summary}
    \begin{itemize}
      \item Renormalization Group Theory is an iterative procedure useful for investigating the critical point
      \item Square well liquid is a widely used model for simple liquids
      \item Combine the two, and get a good model for simple liquids near the critical point
    \end{itemize}
  \end{block}

  \begin{block}{Future Work}
    \begin{itemize}
      \item Finish debugging
      \item Takes a long time to run---iterative process means computing integrals of integrals of integrals\ldots
      \begin{itemize}
        \item Investigate better methods of integration
      \end{itemize}
      \item Apply to inhomogeneous fluids
      \begin{itemize}
        \item How will the liquid behave near hydrophobic solutes? etc.
      \end{itemize}
    \end{itemize}
  \end{block}
\end{frame}

\subsection{}
\begin{frame}{Acknowledgments}
  Without these peoples' support, encouragement \& patience, this would not have been possible
  \begin{columns}[t]
  \column{0.6\textwidth}
  \begin{itemize}
    \item David Roundy
    \item Eric Krebs
    \item Jeff Schulte
    \item Paho Lurie-Gregg
    \item Mary Roth
  \end{itemize}
  \column{0.4\textwidth}
  \begin{figure}
    \centering
    \includegraphics[width=0.9\columnwidth]{figs/coffee}
    \caption{Coffee helped, too}
  \end{figure}
  \end{columns}
\end{frame}

%-----------------------------------------
% Backup
%-----------------------------------------
%% \appendix
%% \section{TPT derivation}
%% \subsection{}

%% \begin{frame}{Total potential energy}
%%   Total potential energy \& derivatives
%%   \begin{align*}
%%     v_\xi(\III) &= v_\text{hs}(\III) + \xi w_\text{sw}(\III) \\
%%     \Vtilde &= \sum_{i=1}^N\sum_{j>i}^N v_\xi(\mathbf{r}_{ij}) \\ % \label{eq:VN} \\
%%     \frac{\partial \Vtilde}{\partial\xi} &= \sum_{i=1}^N\sum_{j>i}^N w_\text{sw}(\mathbf{r}_{ij})  \\
%%     &\equiv \Wtilde \\
%%     \frac{\partial^2 \Vtilde}{\partial\xi^2} &= 0
%%   \end{align*}
%% \end{frame}

%% \begin{frame}{Configuration integral}
%%   Configuration integral \& derivative
%%   \begin{align*}
%%     Z_\xi &= \int d\rr^N\, e^{-\beta \Vtilde} \\
%%     \frac{\partial Z_\xi}{\partial\xi} &=  \int d\rr^N\, \frac{\partial}{\partial\xi}\left( e^{-\beta \Vtilde} \right)  \\
%%     &= -\int d\rr^N\, \beta\frac{\partial \Vtilde}{\partial\xi}e^{-\beta \Vtilde}  \\
%%     &= -\int d\rr^N\, \beta\Wtilde e^{-\beta \Vtilde}
%%   \end{align*}
%% \end{frame}

%% \begin{frame}{Excess free energy}
%%   Excess free energy \& derivative
%%   \begin{align*}
%%     F &= -\kT\ln\left( \frac{Z_\xi}{V^N} \right) \\
%%     \frac{\partial F}{\partial\xi} &= \frac{V^N}{Z_\xi}\frac{\partial}{\partial\xi}\left( \frac{Z_\xi}{V^N} \right)  \\
%%     &= -\kT\left( \frac{1}{Z_\xi}\frac{\partial Z_\xi}{\partial\xi} \right)  \\
%%     &= \frac{1}{Z_\xi} \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \\ %% \label{eqn:dfdlambda_Z}\\
%%     &= \left\langle \Wtilde \right\rangle_{\xi} \\ %% \label{eqn:dfdlambda}
%%   \end{align*}
%% \end{frame}

%% \begin{frame}{Series expansion}
%%   Series expansion \& derivative
%%   \begin{align*}
%%     \left\langle \Wtilde \right\rangle_{\xi} =&{} \left\langle \Wtilde\right\rangle_{\xi = 0} + (\xi)\frac{\partial\left\langle \Wtilde \right\rangle_{\xi}}{\partial\xi}\bigg|_{\xi = 0} + \mathcal{O}(\xi^2) \\
%%     \frac{\partial}{\partial\xi}\left\langle \Wtilde \right\rangle_{\xi} =&{} \frac{\partial}{\partial\xi}\left( \frac{1}{Z_\xi}\int d\rr^N\, \Wtilde e^{-\beta \Vtilde}\right)  \\
%%     =&{} \frac{1}{Z_\xi}\frac{\partial}{\partial\xi}\left( \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right)  \\
%%     &{}- \left( \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right)\frac{1}{Z_\xi^2}\frac{\partial}{\partial\xi}Z_\xi  \\
%%     =&{} \frac{1}{Z_\xi}\int d\rr^N\, \left( e^{-\beta \Vtilde}\frac{\partial}{\partial\xi}\Wtilde + \Wtilde\frac{\partial}{\partial\xi}e^{-\beta \Vtilde} \right)  \\
%%     &{}- \left( \int d\rr^N\, \Wtilde e^{-\beta \Vtilde} \right)\frac{1}{Z_\xi^2}\int d\rr^N\, \frac{\partial}{\partial\xi}e^{-\beta \Vtilde}
%%   \end{align*}
%% \end{frame}

\end{document}

### 2014-06-16 papers/pair-correlation/figs/plot-ghs.py
help Dan get his slides sconsing Tested on bennet. Took 1:00:04.

@@ -181,13 +181,6 @@ for i in range(len(ff)):
     eta[i] = ff[i]
     r = r_mc
 
-if able_to_read_file == False:
-  pylab.plot(arange(0,10,1), [0]*10, 'k')
-  suptitle('!!!!WARNING!!!!! There is data missing from this plot!', fontsize=25)
-  pylab.savefig("figs/ghs-g2.pdf")
-  pylab.savefig("figs/ghs-g-ghs2.pdf")
-  exit(0)
-
 # now do the least squares fit
 fit_rcutoff = 30.6
 def dist(x):
@@ -506,16 +499,4 @@ fig2.tight_layout()
 fig2.subplots_adjust(hspace=0.01)
 fig2.savefig('figs/ghs-alt.pdf')
 
-
-
-# pylab.figure(2)
-# pylab.xlim(2,6.5)
-# pylab.ylim(-.25, .25)
-# pylab.xlabel(r"$r/R$")
-# pylab.ylabel("|ghs - g|")
-# pylab.savefig("figs/ghs-g-ghs2.pdf")
-
-# pylab.axhline(y=0)
-# pylab.xlim(2,6.5)
-# pylab.legend(loc='best')
 pylab.show()

### 2014-06-16 papers/thesis-roth/project.tex

@@
   \begin{center}
-  \includegraphics[width=0.5\textwidth]{figs/coexistance_SW}
+  \includegraphics[width=0.5\textwidth]{figs/coexistence_SW}
   \end{center}
@@
   \end{center}
-  \caption{Liquid-vapor coexistence, compared with Forte's 2011 data}
+  \caption{Liquid-vapor coexistence, compared with Forte's 2011 data, and with molcular dynamics simulations from White\cite{White00}}
   \label{fig:coexistence-RG-forte}
@@
 $i=1$ is neither complete nor
-accurate. Figure~\ref{fig:coexistence-RG-forte} gives an even starker
+accurate. Figure~\ref{fig:coexistence-RG-forte} gives another
 picture of the shortcomings of my RG implementation. We see that the
@@
 expected shortcomings near the critical point. However, calculations
-including even one iteration ($i=1$) flat-out fail.
+including even one iteration ($i=1$) are not included because they flat-out fail.
 

### 2014-06-17 papers/pair-correlation/paper.tex
help Dan get his slides sconsing Tested on bennet. Took 1:00:04.
edit introduction to refer to g(n) vs g(V) Tested on bennet. Took 0:21.51.

@@
 \begin{align}
-  F_1 &= \tfrac12 \iint
+  F_1[n(\rr)] &= \tfrac12 \iint \!\!
   g^{(2)}_{HS}(\rr_1,\rr_2)n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
@@
 the hard-sphere reference fluid, and $\Phi(r)$ is the pair potential.
-In this paper, we introduce a contact value approximation (CVA) for
+Formally, this requires the pair distribution function as a functional
+of the density $n(\rr)$.  In Section~\ref{sec:gV}, we will introduce
+existing theoretical approaches for computing
+$g^{(2)}_{HS}(\rr_1,\rr_2)$ given the external potential felt by the
+hard spheres.  In Section~\ref{sec:gn}, we will introduce existing
+approximations for the hard-sphere pair distribution that are
+expressed as a functional of the density distribution $n(\rr)$, which
+is a form that is more directly useful in the construction of
+classical density functionals---which are themselves expressed as a
+functional of the density.
+
+In this paper, we introduce a new contact value approximation (CVA) for
 the hard-sphere pair distribution function which is suitable for use
@@
 
-\fixme{Somewhere above introduce the question of whether g2 is
-  expressed in terms of V or n}\davidsays{I'll take a stab at this}
-
-\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}
+\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}\label{sec:gV}
 
@@
 \section{Previous approximations to the pair distribution function
-  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}
+  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}\label{sec:gn}
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g2}% \\
-  %\includegraphics[width=\columnwidth]{figs/ghs-g-ghs}
+  \includegraphics[width=\columnwidth]{figs/ghs-g2}
   \caption{Plot of the hard-sphere radial distribution function, with

### 2014-06-19 papers/pair-correlation/paper.tex

@@
 
-\section{Previous theoretical approaches $g^{(2)}(\rr_1,\rr_2)[V(\rr)]$}\label{sec:gV}
+\section{Pair distribution from the external potential}\label{sec:gV}
 
-\fixme{introduce this as methods given V}\davidsays{I'll take a stab
-  at this}
-
-The classic (and earliest) approach for computing the pair
-distribution function is to use Percus' trick of treating one sphere
-as an external field, and to use the resultant equilibrium density to
-find the pair distribution function~\cite{hansen2006theory}.  This
-elegant approach lends itself to computation with DFT, and can be used to compute and
-plot the pair distribution function, but requires a full free-energy
-minimization \emph{for each position} $\rr_1$ in
-$g^{(2)}(\rr_1,\rr_2)$, and hence would be prohibitively expensive as
-a tool in constructing a free energy functional.
+Given the external potential $V(\rr)$ felt by a hard-sphere fluid,
+there are several approaches that have been used to compute the pair
+distribution function.  We will review these approaches here.  The
+classic (and earliest) approach for computing the pair distribution
+function given the external potential is Percus' trick of treating one
+sphere as an additional contribution to the external potential, and to
+find the pair distribution function from the resultant equilibrium
+density~\cite{hansen2006theory}.  This elegant approach lends itself
+to computation with DFT, and can be used to compute and plot the pair
+distribution function, but requires a full free-energy minimization
+\emph{for each position} $\rr_1$ in $g^{(2)}(\rr_1,\rr_2)$, and hence
+would be prohibitively expensive as a tool in constructing a free
+energy functional.
 
@@
 system using integral equation theory under the Percus-Yevick
-approximation~\cite{plischke1986pair}.  \fixme{Move the following to
-  the next section} Ten years later,
-G{\"o}tzelmann \emph{et al.} found the pair distribution function near
-a hard wall by solving the Ornstein-Zernike equation using the direct
-correlation function found from a functional derivative of a classical
-DFT free energy functional~\cite{gotzelmann1996structure}.
-Unfortunately, solving the inhomogeneous Ornstein-Zernike equation
-remains computationally challenging, although more efficient
-approximate algorithms have been developed~\cite{paul2003variational}.
-This approach, while appealing, remains unsuitable for use in the
-construction of a classical density functional due to its significant
-computational cost.
+approximation~\cite{plischke1986pair}.
 
@@
 configuration has a surface with curvature at the molecular length
-scale.  In this case, the density gives the pair distribution
+scale.  In this case, the density gives the radial distribution
 function---this is just Percus' trick---and the pair distribution
@@
 
-\section{Previous approximations to the pair distribution function
-  $g^{(2)}(\rr_1,\rr_2)[n(\rr)]$}\label{sec:gn}
+\section{Pair distribution from the density}\label{sec:gn}
 
-\fixme{introduce this as methods given n}\davidsays{I'll do this}
+The alternative to specifying the external potential is to specify the
+density distribution $n(\rr)$.  One may move between these
+representations by either computing the external potential
+corresponding to a given density of hard spheres by taking a
+functional derivative of the hard-sphere free energy functional, or by
+minimizing the free energy given an external potential.  However, in
+general it is simplest to use an approach that makes use of the
+natural variables, which in the case of classical density functional
+theory is the density.
 
-There are several analytic approximations for the inhomogeneous pair
-distribution function, which extend the radial distribution function
-to inhomogeneous scenarios.  The approximations differ in what density
+The most direct and rigorous approach to find the pair distribution
+function given the density is to take a second functional derivative
+of the hard-sphere free energy to find the direct correlation
+function.  One can then solve the Ornstein-Zernike equation
+numerically to find the pair distribution function.  This approach was
+used by G{\"o}tzelmann \emph{et al.} to solve for the pair
+distribution function near a hard wall using an early hard-sphere free
+energy functional~\cite{gotzelmann1996structure}.  While this approach
+is rigorous, solving the inhomogeneous Ornstein-Zernike equation
+remains computationally challenging, although more efficient
+approximate algorithms have been developed~\cite{paul2003variational}.
+This approach, while appealing, remains unsuitable for use in the
+construction of a classical density functional due to its significant
+computational cost.
+
+In addition to the above exact approach, there are a number of
+analytic approximations for the inhomogeneous pair distribution
+function, which extend the radial distribution function to
+inhomogeneous scenarios.  These approximations differ in what density
 is used when evaluating the radial distribution function $g(r,\eta)$
@@
 papers that proceed to make further
-approximations~\cite{gloor2007prediction, gross2009density}.  While
+approximations~\cite{gloor2007prediction, gross2009density}.  However,
 these approximations fail dramatically when applied to strongly
@@
 was arrived at by fitting to Monte Carlo simulation.  Although
-Eq.~\ref{eq:sokolowski} has the advantages of only involving density
-averages at the points at which the pair distribution function is
-desired, it cannot be written as a single-site convolution, since the
-convolution kernel depends on both points.  However, this approach has
-been used by two papers to construct classical density functionals for
-studying the liquid vapor interface~\cite{wadewitz2000application,
-  winkelmann2001liquid}.
+Although Eq.~\ref{eq:sokolowski} has the advantages of only involving
+density averages at the points at which the pair distribution function
+is desired, it remains sufficiently computationally cumbersome that it
+has only been used in two papers studying the liquid vapor
+interface~\cite{wadewitz2000application, winkelmann2001liquid}.
+Because it cannot be written as a single-site convolution, this
+approach is particularly computationally demanding when applied to
+systems featuring inhomogeneity in more than one dimension.
 
@@
 
-\section{Performance in thermodynamic perturbation theory}
+\section{Accuracy in thermodynamic perturbation theory}
 
@@
 
-\subsection{Efficiency}
+\section{Efficiency in Thermodynamic Perturbation Theory}
 
-\fixme{Explain why the separable version is efficient.}\davidsays{I'll
-  do it.}
+The key insight of the CVA-S is that we can implement
+Eq.~\ref{eq:mean-field} on $\mathcal{O}(N\log N)$ time using Fast
+Fourier Transforms (FFTs) where $N$ is proportional to the volume of
+the system.  We begin by examining the CVA.  The two terms that are
+averaged in Eq.~\ref{eq:CVA} give equal contributions to the integral:
+\begin{align}
+  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
+  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
+  d\rr_1d\rr_2
+\end{align}
+When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
+further simplify this integral as
+\begin{multline}
+  F_1^{\textit{CVA-S}} =
+  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
+                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
+                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
+  d\rr_1d\rr_2
+\end{multline}
+where the functional is written as a summation of integrals, each of
+which is an integral of simple convolution in three dimensions.  Thus,
+each of these integrals may be computed in $\mathcal{O}(N\log N)$
+time, which is the same scaling behavior that is common to the most
+common and accurate classical density functional theories, most
+notably FMT.
 

### 2014-06-23 papers/pair-correlation/paper.tex

@@
 
-In this paper, we introduce a new contact value approximation (CVA) for
+In this paper, we introduce a new contact value approach (CVA) for
 the hard-sphere pair distribution function which is suitable for use
@@
 computationally efficient separable version of the contact
-value approximation (CVA-S), which is based on a fit to the radial
+value approach (CVA-S), which is based on a fit to the radial
 distribution function that is separable in a way that will enable
@@
 
-\davidsays{Please edit!}\jeffsays{I will!}\fixme{Note that the HS free energy is used
-  for two purposes when applying Eq.~\ref{eq:gsigma-exact} : it is
-  minimized (together with an external potential) to find the density
-  $n(\rr)$.  Then its functional derivative is used to find
-  $g_\sigma(\rr)$.  In application of perturbation theory, the HS
-  functional would only be used alone for the latter purpose, as the
-  density would be determined by the total free energy, which
-  incorporates both the HS free energy and a perturabative
-  contribution.}
+\jeffsays{My attempt at this paragraph:} To be clear we note that for
+our purposes in this work we use the hard sphere free energy twice.
+It is first minimized to find $n(\rr)$ and then its functional
+derivative is used to find $g_\sigma(\rr)$.  In actual application of
+perturbation theory the hard sphere free energy would be used only to
+find $g_\sigma(\rr)$, which would in turn be used in the approximation
+for $g^{(2)}(\rr_1,\rr_2)$ that would appear in a perturbation term.
+The free energy that would be minimized to find $n(\rr)$ in this case
+would be the total free energy and not the hard sphere free energy.
 
-\section{Contact value approximation}
+\section{Contact value approach}
 In the approaches for the pair distribution function mentioned above,
@@
 \end{align}
-This contact value approximation for $g^{(2)}(\rr_1,\rr_2)$ is
+This contact value approach for $g^{(2)}(\rr_1,\rr_2)$ is
 constructed to reproduce the exact value for the integral:
@@
 \end{align}
-This separable contact value approximation (CVA-S) allows the mean
+This separable contact value approach (CVA-S) allows the mean
 field contribution to the free energy (Eq.~\ref{eq:mean-field}) to be
@@
 
-We have introduced and tested the contact value approximation for the pair
+We have introduced and tested the contact value approach for the pair
 distribution function of the inhomogeneous hard-sphere fluid.  The
 separable version of the contact value
-approximation is suitable for use in the development of classical
+approach is suitable for use in the development of classical
 density functionals constructed using perturbation theory, as it

### 2014-06-25 papers/pair-correlation/paper.tex
experiment with a higher-order fit specifically to the same range as Gil Villegas Tested on bennet. Took 0:23.24.

@@
 \begin{document}
-\title{The contact value approximation for the pair distribution function of the inhomogeneous hard-sphere
+\title{The contact value approach for the pair distribution function of the inhomogeneous hard-sphere
   fluid}
@@
     $g_S(r)$.}\label{fig:radial-distribution}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
+  \caption{Plot of the hard-sphere radial distribution function, with
+    our short-range separable
+    fit.}\label{fig:short-range-radial-distribution}
 \end{figure}

### 2014-06-27 papers/pair-correlation/paper.tex

@@
 \begin{document}
-\title{The contact value approach for the pair distribution function of the inhomogeneous hard-sphere
-  fluid}
+\title{An efficient contact value approach for the pair distribution
+  function of the inhomogeneous hard-sphere fluid}
 
@@
 
-\fixme{replace approximation with approach}\jeffsays{I'm stuck with this.}
+\fixme{
+  TO DO:
+  \begin{itemize}
+  \item \pahosays{Only use short-range CVA-S (in plots, and in walls.cpp)}
+  \item \pahosays{Remove CVA from all plots, but leave it easy to
+    re-add (so we can check what we say in the text).}
+  \item \pahosays{Make all plots not show CVA (new name for CVA-short)
+    outside the fitted range.}
+  \item \davidsays{Rewrite CVA-S section to reflect new fit.}
+  \item \jeffsays{Make certain when proof-reading that other prior
+    formulations cannot work in ``separable'' way.}
+  \item \davidsays{Be sure that we explain the fitted data and fitting
+    process well.}
+  \item \pahosays{Add $\eta=?$ labels and legends to new fit plot.}
+  \item \jeffsays{I will diligently read what you guys write!}
+  \end{itemize}
+}
 
@@
 \end{align}
-This contact value approach for $g^{(2)}(\rr_1,\rr_2)$ is
+This Contact Value Approach (CVA) for $g^{(2)}(\rr_1,\rr_2)$ is
 constructed to reproduce the exact value for the integral:
@@
 \section{Separable radial distribution function for the CVA-S}
+\fixme{Section title:  A separable fit for the radial distribution function?}
 
@@
 the system.  We begin by examining the CVA.  The two terms that are
-averaged in Eq.~\ref{eq:CVA} give equal contributions to the integral:
+averaged in Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
 \begin{align}

### 2014-06-27 papers/pair-correlation/paper.tex
start adding new fit equation to pair paper Tested on bennet. Took 0:20.43.

@@
 
-\input{figs/fit-parameters}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-g2}
-  \caption{Plot of the hard-sphere radial distribution function, with
-    our separable fit.  The solid lines show the true radial
-    distribution function $g(r)$, as found from Monte Carlo
-    simulation, and the dashed lines show our separable fit to the
-    radial distribution functions
-    $g_S(r)$.}\label{fig:radial-distribution}
-\end{figure}
+\input{figs/short-range-fit-parameters}
 
@@
   \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
-  \caption{Plot of the hard-sphere radial distribution function, with
-    our short-range separable
-    fit.}\label{fig:short-range-radial-distribution}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=\columnwidth]{figs/ghs-alt}
-  \caption{Plot of the hard-sphere radial distribution function. The blue
-    lines show our separable fit, the black dots are Monte Carlo data, and
-    the dashed lines are results of the Gil-Villegas\cite{gil1997statistical}
-    fit.\pahosays{An idea for an alternate version of the radial distribution
-      figure.}}\label{fig:radial-distribution-alt}
+  \caption{Plot of the hard-sphere radial distribution function. The
+    blue lines show our separable fit, the black dots show the true
+    radial distribution function $g(r)$ as found from Monte Carlo
+    simulation, and the dashed lines are results of the
+    Gil-Villegas\cite{gil1997statistical} fit.  The dotted extension
+    of each fitted curve indicates the value of the function outside
+    of the fitted region.  }\label{fig:radial-distribution}
 \end{figure}
@@
 
-%% We construct our function's form in terms of an expansion of powers of
-%% $h_{\sigma} = g_{\sigma}-1$, ensuring that it approaches $g_{\sigma}$
-%% in the contact distance limit, leaving our function with tunable
-%% constant perameters.  We run a simple monte-carlo simulation to
-%% calculate values of the homogeneous radial distribution function at
-%% varying distance perameter $r$.  We then tune our function's constant
-%% parameters by matching our function's results against the monte-carlo
-%% data, using a least squares fit method.  During this process it is
-%% simplest to use an analytic expression for $g_{\sigma}$ in our
-%% function and so we use the well known and very successful Carnahan
-%% Starling equation to calculate $g_{\sigma}$.  This is appropriate
-%% since our published $g_{\sigma}$ approaches the Carnahan Starling
-%% version in the homogeneous limit and it's this limit, along with the
-%% contact limit, that we would like our constructed radial distribution
-%% function to get right.  After this process we have an appropriate
-%% radial distribution function that is dependent only on $r$ and
-%% $g_{\sigma}$.
-
-%% The function has the form:
-%% \begin{multline}
-%%   g(r,g_\sigma) = 1 + h_\sigma e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
-%%   + h_\sigma(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
-%%   + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
-%% \end{multline}
-%% where $\kappaO$,$\kappaI$, and $\kappaZ$ are fitted constants and
-%% $I(g_\sigma)$ is a function of $g_\sigma$.
-
-%% The finalized form of our pair correlation function is:
-%% \begin{align}
-%%   g^{(2)}(\rr_1,\rr_2) = \frac{g(|r_{12}|, g_\sigma(\rr_1)) + g(|r_{12}|, g_\sigma(\rr_2))}{2}
-%% \end{align}
-
 We construct $g_S(r, g_\sigma)$ to satisfy several constraints.
@@
 \end{equation}
-A second constraint applies to the integral of the total correlation
-function, $h(r) = g(r) - 1$:
+
+\begin{widetext}
 \begin{align}
-  1 + n\int h(r)d\rr &= nkT\chi_T \label{eq:total-constraint}
+  g_S(r,g_\sigma) &=
+  1 + \left(h_\sigma + \sum_{i=2}^{4} \sum_{j=1}^{4} h_\sigma^i
+  \zeta^j\right)e^{-\alpha \zeta}
+  \\
+  &= \kappamatrix
 \end{align}
-where $\chi_T$ is the isothermal compressibility of the hard-sphere fluid.
-We have found that the slope of the radial distribution function at
-contact is given to a very good approximation by:
-\begin{align}
-  g'(\sigma) \approx g_\sigma (1 - g_\sigma)
-\end{align}
-and we apply this as a final constraint.\par
-Under these three constraints, we applied a least-squares fit to
-simulation data for the radial distribution function at packing
-fractions from 0.05 to 0.4 in steps of 0.05.  Our fitted radial
-distribution function is given by
-\begin{multline}
-  g_S(r,g_\sigma) = 1 + (g_\sigma-1) e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
-  + (g_\sigma -1)(\kappaO - 2g_\sigma)  \left(\frac{r}{\sigma}-1\right)e^{-\kappaI  \left(\frac{r}{\sigma}-1\right)} \\
-  + I(g_\sigma)  \left(\frac{r}{\sigma}-1\right)^2e^{-\kappaZ  \left(\frac{r}{\sigma}-1\right)}
-\end{multline}
-\begin{equation}
-  I(g_\sigma) = \kappaZ^5 \frac{
-    \frac{\chi-1}{24\eta} + (1-g_\sigma) \frac{(\kappaO-2
-      g_\sigma)(\kappaI^2 + 4 \kappaI + 6)}{\kappaI^4}
-    + (1-g_\sigma)\frac{\kappaO^2 + 2\kappaO + 2}{\kappaO}
-  }{2 \kappaZ^2 + 12 \kappaZ + 24}
-\end{equation}
-\begin{equation}
-  \chi = \frac{(1-\eta)^4}{\eta^4 - 4\eta^3 + 4\eta^2 + 4\eta + 1}
-\end{equation}
-where the dimensionless isothermal compressibility $\chi$ is obtained
-from the Carnahan-Starling equation of state.
-Finally, we can convert between the packing fraction $\eta$ and the
-radial distribution function at contact $g_\sigma$ using the
-Carnahan-Starling approximation for $g_\sigma$:
-\newcommand\nastyetacuberoot{\sqrt[3]{54 g_\sigma^2 -
-    6\sqrt{81g_\sigma^4 - 6g_\sigma^3}}}
-\begin{align}
-  g_\sigma &= \frac{1-\tfrac{\eta}{2}}{(1-\eta)^3}\text{, and its inverse} \\
-  \eta &= 1 - \frac{1}{\nastyetacuberoot} - \frac{\nastyetacuberoot}{6g_\sigma}.
-\end{align}
-This function includes three fitted parameters, $\kappaO =
-\kappazero$, $\kappaI = \kappaone$,
-and $\kappaZ = \kappatwo$.
-The fit is displayed in Fig.~\ref{fig:radial-distribution}.  In that
-figure, it is apparent that the constraint on the integral of $h(r)$ from
-Eq.~\ref{eq:total-constraint} results in canceling errors at large
-distances.  The maximum error in $g_S(r)$ at packing fractions up to
-0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
-\rmaxerr R$.
+\end{widetext}
+where $h_\sigma = g_\sigma - 1$ and $\zeta \equiv
+\frac{r}{\sigma}-1$.  \fixme{Why start at $i=2$ rather than $i=1$ for
+  the fitted terms? This seems suboptimal.}
+
+\fixme{Change code to round to a fixed (small) number of digits in fit.}
 

### 2014-06-28 papers/pair-correlation/paper.tex
Only use new short range CVA in plots and walls. All plots only show fitted range. Added labels to fit plot. Tested on paho-arch. Took 0:36.79.

@@
 \usepackage{subcaption} % provides subfigure
+
+\usepackage{dcolumn}
+\newcolumntype{d}[1]{D{.}{.}{#1}}
 
@@
   \begin{itemize}
-  \item \pahosays{Only use short-range CVA-S (in plots, and in walls.cpp)}
+  \item \pahosays{Only use short-range CVA-S (in plots, and in walls.cpp)}\pahosays{DONE}
   \item \pahosays{Remove CVA from all plots, but leave it easy to
-    re-add (so we can check what we say in the text).}
+    re-add (so we can check what we say in the text).}\pahosays{DONE}
   \item \pahosays{Make all plots not show CVA (new name for CVA-short)
-    outside the fitted range.}
+    outside the fitted range.}\pahosays{DONE}
   \item \davidsays{Rewrite CVA-S section to reflect new fit.}
@@
     process well.}
-  \item \pahosays{Add $\eta=?$ labels and legends to new fit plot.}
+  \item \pahosays{Add $\eta=?$ labels and legends to new fit plot.}\pahosays{DONE}
   \item \jeffsays{I will diligently read what you guys write!}
@@
 
+The fit is displayed in Fig.~\ref{fig:radial-distribution}.  This
+function is fit to data for $r <= \maxrfit R$.  The maximum error
+within this radius in $g_S(r)$ at packing fractions up to 0.4 is
+\maxerr, which occurs at $\eta = \etamaxerr$ and $r = \rmaxerr R$.
+
 \newcommand{\plotcomp}[1]{The top halves of these figures show the

### 2014-06-28 papers/pair-correlation/figs/styles.py

@@ -1,9 +1,7 @@
-
-
 color = { 'mc': 'k',
           'this-work': 'b',
           'this-work-mc': 'c',
-          'this-work-short': 'm',
+          'this-work-short': 'b',
           'fischer': 'g',
           'sokolowski': 'r'}
 
@@ -30,12 +28,12 @@ back_line = { 'mc': '.',
 
 title = { 'mc': 'Monte Carlo',
           'this-work': 'CVA-S',
-          'this-work-mc': 'CVA',
-          'this-work-short': 'CVA-short',
+          'this-work-mc': 'CVA-old',
+          'this-work-short': 'CVA',
           'fischer': 'Fischer',
           'sokolowski': 'Sokolowski'}
 
-start = { 'this-work' : 0,
+start = { 'this-work-short' : 0,
           'this-work-mc' : 1/6.,
           'sokolowski' : 1/3.}
 
@@ -51,5 +49,7 @@ plot_forward = {}
 for k in color:
     plot_forward[k] = color[k] + forward_line[k]
 
+plots = ['mc', 'sokolowski', 'fischer', 'this-work-short']
+oriented_plots = ['sokolowski', 'this-work-short']
 
 short_range = 4.0

### 2014-07-01 papers/pair-correlation/paper.tex
proofread and added jeffsays notes to current version of pair paper Tested on node22. Took 0:32.38.

@@
 \begin{document}
-\title{An efficient contact value approach for the pair distribution
+\title{An efficient contact value approach to approximating the pair distribution
   function of the inhomogeneous hard-sphere fluid}
@@
 implementation using fast Fourier transforms. We compare results for
-both versions of our pair distribution approximation with two
-previously published works and Monte-Carlo simulation, showing
-favorable results.
+both versions of our pair distribution approximation \jeffsays{Is it
+  still both? Do we want that?} with two previously published works
+and Monte-Carlo simulation, showing favorable results.
 \end{abstract}
@@
 position $\rr$, and $F_{HS}$ is the Helmholtz free energy of the
-hard-sphere fluid.  The functional derivative with of free energy with
+hard-sphere fluid.  The functional derivative of the free energy with
 respect to hard-sphere diameter in Eq.~\ref{eq:gsigma-exact} requires
@@
 pair distribution function at contact for a variety of interfaces, and
-over a wide range of densities.
-
-\jeffsays{My attempt at this paragraph:} To be clear we note that for
-our purposes in this work we use the hard sphere free energy twice.
-It is first minimized to find $n(\rr)$ and then its functional
-derivative is used to find $g_\sigma(\rr)$.  In actual application of
-perturbation theory the hard sphere free energy would be used only to
-find $g_\sigma(\rr)$, which would in turn be used in the approximation
-for $g^{(2)}(\rr_1,\rr_2)$ that would appear in a perturbation term.
-The free energy that would be minimized to find $n(\rr)$ in this case
-would be the total free energy and not the hard sphere free energy.
+over a wide range of densities.  \jeffsays{We had a clarifying
+  paragraph below but I don't think it's needed anymore, at least not
+  here.  The explanation up until this point is moving along fine and
+  I don't think it needs clarification.}
+%% \jeffsays{My attempt at this paragraph:} To be
+%% clear we note that for our purposes in this work we use the hard
+%% sphere free energy twice.  It is first minimized to find $n(\rr)$ and
+%% then its functional derivative is used to find $g_\sigma(\rr)$.  In
+%% actual application of perturbation theory the hard sphere free energy
+%% would be used only to find $g_\sigma(\rr)$, which would in turn be
+%% used in the approximation for $g^{(2)}(\rr_1,\rr_2)$ that would appear
+%% in a perturbation term.  The free energy that would be minimized to
+%% find $n(\rr)$ in this case would be the total free energy and not the
+%% hard sphere free energy.
 
 \section{Contact value approach}
-In the approaches for the pair distribution function mentioned above,
-the radial distribution function used in the approximation
-was dependent upon the density averaged over some volume.  We seek to
-achieve greater accuracy by making use of a function dependent upon
-our averaged $g_{\sigma}(\rr)$ discussed above, which holds more
-information about an inhomogeneous system than does a simple
-convolution of the density.
+\jeffsays{In my current version latex's formatting of this and the
+  next section is very poor so we should keep our eye on it as the
+  text develops} In the approaches for the pair distribution function
+mentioned above, the radial distribution function used in the
+approximation was dependent upon the density averaged over some
+volume.  We seek to achieve greater accuracy by making use of a
+function dependent upon our averaged $g_{\sigma}(\rr)$ discussed
+above, which holds more information about an inhomogeneous system than
+does a simple convolution of the density.
 %
@@
 results for the radial distribution function $g(r)$, which we also use
-in evaluating the approximations of earlier works.
+in consturcting and evaluating the approximations of earlier works.
+\jeffsays{With this last sentence: it states that we 'first' use the
+  Monte Carlo version for our radial distribution but then doesn't say
+  what we use afterward.}
 
@@
 \end{equation}
+
+\jeffsays{The rest of the fitting needs to be explained, and
+  discussing the connection with Gill-Viegas}
 
@@
 \end{figure*}
+
+\jeffsays{I've noted this in the penultimate section as well but I
+  think it might be good to include that section after this one.
+  Where that section is now, I feel like it seems we think it's less
+  important.  This is subjective, though.}
 
@@
 \section{Efficiency in Thermodynamic Perturbation Theory}
+\jeffsays{I think this ought to be included after the Seperable Fit
+  secction.  It may seem more important that way.}
 

### 2014-07-01 papers/pair-correlation/paper.tex
various pair paper edits Tested on bennet. Took 0:27.99.

@@
 implementation using fast Fourier transforms. We compare results for
-both versions of our pair distribution approximation \jeffsays{Is it
-  still both? Do we want that?} with two previously published works
-and Monte-Carlo simulation, showing favorable results.
+our pair distribution approximation with two previously published
+works and Monte-Carlo simulation, showing favorable results.
 \end{abstract}
@@
 pair distribution function at contact for a variety of interfaces, and
-over a wide range of densities.  \jeffsays{We had a clarifying
-  paragraph below but I don't think it's needed anymore, at least not
-  here.  The explanation up until this point is moving along fine and
-  I don't think it needs clarification.}
+over a wide range of densities.
+
+  %% \jeffsays{We had a clarifying paragraph below but I don't think
+  %% it's needed anymore, at least not here.  The explanation up until
+  %% this point is moving along fine and I don't think it needs
+  %% clarification.}
+
 %% \jeffsays{My attempt at this paragraph:} To be
@@
 \section{Contact value approach}
-\jeffsays{In my current version latex's formatting of this and the
-  next section is very poor so we should keep our eye on it as the
-  text develops} In the approaches for the pair distribution function
+In the approaches for the pair distribution function
 mentioned above, the radial distribution function used in the
@@
 some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
-simulation results.  In this paper we first use the Monte Carlo
-results for the radial distribution function $g(r)$, which we also use
-in consturcting and evaluating the approximations of earlier works.
-\jeffsays{With this last sentence: it states that we 'first' use the
-  Monte Carlo version for our radial distribution but then doesn't say
-  what we use afterward.}
-
-\newcommand\kappaO{\kappa_0}
-\newcommand\kappaI{\kappa_1}
-\newcommand\kappaZ{\kappa_2}
+simulation results.  In this paper we use a fit to Monte Carlo results
+for the radial distribution function $g(r)$, which facilitates
+efficient computation.  When evaluating the approximations of earlier
+works described in Section~\ref{sec:gn}, we use interpolated Monte
+Carlo results for $g(r,n)$.
 
 \input{figs/short-range-fit-parameters}
+
+\section{Making the CVA efficient}
+The CVA as described in the previous section is an incremental
+improvment over existing approaches outlined in Section~\ref{sec:gn}.
+However, as discussed above, these approximations have not been widely
+used in the construction of density functionals based on thermodynamic
+perturbation theory, largely due to their computational complexity.
+We address this issue by developing a \emph{separable} fit to the
+radial distribution function of the hard-sphere fluid (see
+Section~\ref{sec:separable-fit} for details).  The separable fit is of
+the form
+\begin{align}
+  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+\end{align}
+where the notable aspect is that the radial distribution function is
+written as a sum of terms that are each a simple product of a function
+of radius with a function of $g_\sigma$.  This enables us to write
+typicaly integrals involving the pair distribution function as a
+summation of fixed-kernel convolutions, which may be efficiently
+computed using Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log
+N)$ time.
+
+We begin by examining the lowest-order correction in TPT, given by
+Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
+Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
+\begin{align}
+  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
+  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
+  d\rr_1d\rr_2
+\end{align}
+When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
+further simplify this integral as
+\begin{multline}
+  F_1^{\textit{CVA-S}} =
+  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
+                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
+                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
+  d\rr_1d\rr_2
+\end{multline}
+where the functional is written as a summation of integrals, each of
+which is an integral of simple convolution in three dimensions.  Thus,
+each of these integrals may be computed in $\mathcal{O}(N\log N)$
+time, which is the same scaling behavior that is common to the most
+common and accurate classical density functional theories, most
+notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
+  roth2002whitebear}.
 
@@
 
-\section{Separable radial distribution function for the CVA-S}
-\fixme{Section title:  A separable fit for the radial distribution function?}
+\section{A separable fit for the radial distribution function}\label{sec:separable-fit}
 
-The CVA is quite accurate, but cannot be computed efficiently
-with fixed-kernel convolutions as written.  In order to ease
-computational application of this pair distribution function in
-classical density functionals, we have developed a separable
-approximation for the function $g(r,g_\sigma)$:
-\begin{align}
-  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
-\end{align}
-This separable contact value approach (CVA-S) allows the mean
-field contribution to the free energy (Eq.~\ref{eq:mean-field}) to be
-computed with only fixed-kernel convolutions, which can be
-efficiently computed using fast Fourier transforms.
+We have performed a separable fit to the radial distribution function,
+which we compute using Monte Carlo simulation.  We chose to focus our
+fit on the range of distances $r_{12} < 2\sigma$.  This range of
+distances is relevant to the widely used \fixme{add citations?}
+Statistical Associating Fluid Theory of Variable Range (SAFT-VR) free
+energy with square-well dispersive attraction developed by
+Gil-Villegas \emph{et al.}~\cite{gil1997statistical}.  For comparison,
+we plot in Fig.~\ref{fig:radial-distribution} alongside our the fit to
+the radial distribution function that of Gil-Villegas \emph{et al.},
+which we have extracted from their approximation for the first term in
+the dispersion free energy given by Eq.~\ref{eq:mean-field}.
 
-We construct $g_S(r, g_\sigma)$ to satisfy several constraints.
-Naturally, we match the value at contact:
+For ease of implementation and future extension to larger radii, we
+chose to express our fit using a polynomial fit.  We constrain our
+functional form for $g_S(r, g_\sigma)$ to match the value at contact:
 \begin{equation}
@@
 \end{equation}
-
-\jeffsays{The rest of the fitting needs to be explained, and
-  discussing the connection with Gill-Viegas}
-
-\begin{widetext}
+Beyond this constraint, we simply included a polynomial multiplied by
+an exponential decay:
 \begin{align}
   g_S(r,g_\sigma) &=
-  1 + \left(h_\sigma + \sum_{i=2}^{4} \sum_{j=1}^{4} h_\sigma^i
-  \zeta^j\right)e^{-\alpha \zeta}
-  \\
-  &= \kappamatrix
+  1 + \left(\sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} h_\sigma^i
+  \zeta^j\right)e^{-\alpha \zeta} \label{eq:fit-form}
 \end{align}
-\end{widetext}
-where $h_\sigma = g_\sigma - 1$ and $\zeta \equiv
-\frac{r}{\sigma}-1$.  \fixme{Why start at $i=2$ rather than $i=1$ for
-  the fitted terms? This seems suboptimal.}
+where $h_\sigma = g_\sigma - 1$ is the total correlation function at
+contact, and $\zeta \equiv \frac{r}{\sigma}-1$ is a dimensionless
+measure of the distance from contact.
 
-\fixme{Change code to round to a fixed (small) number of digits in fit.}
+\begin{table}
+  \begin{align*}
+    \kappa &= \kappatable
+  \end{align*}
+  \caption{The fitted $\kappa_{ij}$ matrix.  In addition to these 12
+    parameters, we also optimized the exponetial decay constant
+    $\alpha=\alphaval$.
+    %% Note that the first column of $\kappa$ corresponds to $j=1$
+    %% while the first row corresponds to $i=1$ in
+    %% Eq.~\ref{eq:fit-form}.
+  }
+\end{table}
 
-The fit is displayed in Fig.~\ref{fig:radial-distribution}.  This
-function is fit to data for $r <= \maxrfit R$.  The maximum error
-within this radius in $g_S(r)$ at packing fractions up to 0.4 is
-\maxerr, which occurs at $\eta = \etamaxerr$ and $r = \rmaxerr R$.
+The fit is displayed in Fig.~\ref{fig:radial-distribution}, along with
+the radial distribution function fit by Gil-Villegas and Monte Carlo
+simulation results.  This function is fit to data for $r <= \maxrfit
+R$.  The maximum error within this radius in $g_S(r)$ at packing
+fractions up to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$
+and $r = \rmaxerr R$.
 
@@
 \end{figure*}
-
-\jeffsays{I've noted this in the penultimate section as well but I
-  think it might be good to include that section after this one.
-  Where that section is now, I feel like it seems we think it's less
-  important.  This is subjective, though.}
-
 \section{Results}
@@
 
-\section{Efficiency in Thermodynamic Perturbation Theory}
-\jeffsays{I think this ought to be included after the Seperable Fit
-  secction.  It may seem more important that way.}
-
-The key insight of the CVA-S is that we can implement
-Eq.~\ref{eq:mean-field} on $\mathcal{O}(N\log N)$ time using Fast
-Fourier Transforms (FFTs) where $N$ is proportional to the volume of
-the system.  We begin by examining the CVA.  The two terms that are
-averaged in Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
-\begin{align}
-  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
-  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
-  d\rr_1d\rr_2
-\end{align}
-When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
-further simplify this integral as
-\begin{multline}
-  F_1^{\textit{CVA-S}} =
-  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
-                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
-                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
-  d\rr_1d\rr_2
-\end{multline}
-where the functional is written as a summation of integrals, each of
-which is an integral of simple convolution in three dimensions.  Thus,
-each of these integrals may be computed in $\mathcal{O}(N\log N)$
-time, which is the same scaling behavior that is common to the most
-common and accurate classical density functional theories, most
-notably FMT.
-
 \section{Conclusion}

### 2014-07-01 papers/square-well-liquid/paper.tex

@@
 
+\newcommand\kk{\mathbf{k}}
 \newcommand\xx{\mathbf{x}}
-\newcommand\rr{\mathbf{x}}
+\newcommand\rr{\mathbf{r}}
 \newcommand\fixme[1]{\textcolor{red}{\textbf{[#1]}}}
@@
 
-\begin{multline}
+\begin{multline}\label{eq:gs}
   g_S(r,g_\sigma) = 1 + (g_\sigma-1) e^{-\kappaO \left(\frac{r}{\sigma}-1\right)} \\
@@
 
+\subsubsection{Fourier transform of radial distribution function}
+
+\newcommand\kapovsig[1]{\frac{\kappa_{#1}}{\sigma}}
+
+For or purposes, we want to Fourier transform $g_s(r,g_\sigma)u_{SW}(\rr)$ and
+the square well potential since it is used in the free energy functional.
+\begin{align}
+  \tilde{g_s}(k, g_\sigma) =\int^\infty_{-\infty} g_s(r, g_\sigma)u_{SW}(\rr)e^{-i\kk \cdot \rr}\mathrm{d}^3\rr
+\end{align}
+Since the radial distribution function and potential are radially symmetric, we can
+express the dot product differently so that the transform is
+\begin{align}
+  \tilde{g_s}(k, g_\sigma) &=\int^\infty_{-\infty} g_s(r,
+  g_\sigma)u_{sw}(r)e^{-ikr \cos \theta}\mathrm{d}^3\rr \\
+  &= 2 \pi \int_\sigma^{\lambda \sigma} \int_0^\pi g_s(r, g_\sigma) e^{-ikr \cos
+    \theta} r^2 \sin \theta \, \mathrm{d}r \mathrm{d}\theta \\
+  &= 2\pi\int_\sigma^{\lambda \sigma} \int_{1}^{-1} g_s(r, g_\sigma) e^{-ikr \cos
+    \theta} r^2 \mathrm{d}r \mathrm{d}\cos \theta \\
+  &= 2\pi\int_\sigma^{\lambda \sigma} g_s(r, g_\sigma) \frac{1}{-ikr}\bigg[e^{ikr} -
+    e^{-ikr} \bigg]r^2 \mathrm{d}r \\
+ &= \frac {4\pi}{k}\int_\sigma^{\lambda \sigma} r g_s(r,g_\sigma) \sin (kr) \mathrm{d}r.
+\end{align}
+Now I'll do the separate integrals for each part on the right hand
+side of Eqn. \ref{eq:gs}. The first term is 1, so we are left with
+a delta function.  We will ignore this term, and
+move onto the next one whose integral is
+\begin{align}
+  \int_\sigma^{\lambda \sigma}r(g_\sigma -1) e^{-\kappa_0
+    \left( \frac{r}{\sigma}-1 \right)} \sin(kr)\mathrm{d}r.
+\end{align}
+If we pull out constants and put this into Mathematica,
+\begin{widetext}
+  \begin{align}
+  \int_\sigma^{\lambda
+    \sigma}re^{-\frac{\kappa_0}{\sigma}r}\sin(kr)\mathrm{d}r
+  &= \frac{1}{(k^2 + (\kapovsig{0})^2)^2} \bigg\{ e^{-\kappaO} \left[k
+      \left(2\kapovsig{0} + \sigma (k^2 + (\kapovsig{0})^2) \right)
+      \cos (\sigma k) + \left((\kapovsig{0})^2(1 + \kappaO) +
+      k^2(\kappaO-1)\right) \sin (\sigma k) \right] \notag \\
+  &- e^{-\lambda
+    \kappaO} \left[ k
+      \left(2\kapovsig{0} + \lambda \sigma (k^2 + (\kapovsig{0})^2) \right)
+      \cos (\lambda \sigma k) + \left((\kapovsig{0})^2(1 + \lambda \kappaO) +
+      k^2(\lambda\kappaO-1)\right) \sin (\lambda \sigma k)\right] \bigg\}
+  \end{align}
+\end{widetext}
 \subsubsection{Averaged pair distribution  $g_\sigma(\rr)$}

### 2014-07-02 papers/pair-correlation/paper.tex
Another round of proof reading.  Made a handful of edits to the text in CVA Efficient section and the A Separable Fit Section Tested on node26. Took 0:41.37.

@@
 \section{Making the CVA efficient}
-The CVA as described in the previous section is an incremental
-improvment over existing approaches outlined in Section~\ref{sec:gn}.
-However, as discussed above, these approximations have not been widely
-used in the construction of density functionals based on thermodynamic
+The existing approaches to approximating the pair correlation function
+outlined in Section~\ref{sec:gn} have not been widely used in the
+construction of density functionals based on thermodynamic
 perturbation theory, largely due to their computational complexity.
-We address this issue by developing a \emph{separable} fit to the
-radial distribution function of the hard-sphere fluid (see
-Section~\ref{sec:separable-fit} for details).  The separable fit is of
-the form
+While our CVA as described in the previous section is, in terms of
+accuracy, an incremental improvement over these existing approaches,
+it makes significant gains in terms of computational efficiency,
+allowing for its more managable use in such a construction.  We
+address the issue by developing a \emph{separable} fit to the radial
+distribution function of the hard-sphere fluid (see
+Section~\ref{sec:separable-fit} for details).  This separable fit is
+of the form
 \begin{align}
@@
 of radius with a function of $g_\sigma$.  This enables us to write
-typicaly integrals involving the pair distribution function as a
-summation of fixed-kernel convolutions, which may be efficiently
-computed using Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log
-N)$ time.
+typicaly \jeffsays{'typical?' maybe just 'integrals'} integrals
+involving the pair distribution function as a summation of
+fixed-kernel convolutions, which may be efficiently computed using
+Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log N)$ time.
 
@@
 
-We have performed a separable fit to the radial distribution function,
-which we compute using Monte Carlo simulation.  We chose to focus our
-fit on the range of distances $r_{12} < 2\sigma$.  This range of
-distances is relevant to the widely used \fixme{add citations?}
-Statistical Associating Fluid Theory of Variable Range (SAFT-VR) free
-energy with square-well dispersive attraction developed by
-Gil-Villegas \emph{et al.}~\cite{gil1997statistical}.  For comparison,
-we plot in Fig.~\ref{fig:radial-distribution} alongside our the fit to
-the radial distribution function that of Gil-Villegas \emph{et al.},
+We have performed a separable fit to the Monte-Carlo simulated radial
+distribution function.  We chose to focus our fit on the range of
+distances $r_{12} < 2\sigma$.  This range of distances is relevant to
+the widely used \fixme{add citations?}  Statistical Associating Fluid
+Theory of Variable Range (SAFT-VR) free energy with square-well
+dispersive attraction developed by Gil-Villegas \emph{et
+  al.}~\cite{gil1997statistical}.  For comparison, in
+Fig.~\ref{fig:radial-distribution} we plot our fit, Monte-Carlo data,
+and the radial distribution function of Gil-Villegas \emph{et al.},
 which we have extracted from their approximation for the first term in
@@
 For ease of implementation and future extension to larger radii, we
-chose to express our fit using a polynomial fit.  We constrain our
-functional form for $g_S(r, g_\sigma)$ to match the value at contact:
+chose to express our fit with a polynomial function, multiplied by an
+exponential decay.  We constrain our functional form for $g_S(r,
+g_\sigma)$ to match the value at contact:
 \begin{equation}
@@
 \end{equation}
-Beyond this constraint, we simply included a polynomial multiplied by
-an exponential decay:
+Beyond this constraint we have:
 \begin{align}
@@
 
-The fit is displayed in Fig.~\ref{fig:radial-distribution}, along with
-the radial distribution function fit by Gil-Villegas and Monte Carlo
-simulation results.  This function is fit to data for $r <= \maxrfit
-R$.  The maximum error within this radius in $g_S(r)$ at packing
-fractions up to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$
-and $r = \rmaxerr R$.
+The Gil-Villegas function is fit to data for $r <= \maxrfit R$.  The
+maximum error within this radius in $g_S(r)$ at packing fractions up
+to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+\rmaxerr R$.
 

### 2014-07-03 papers/pair-correlation/paper.tex
various sloppy edits during group meeting Tested on darcy. Took 0:57.53.

@@
   \begin{itemize}
-  \item \pahosays{Only use short-range CVA-S (in plots, and in walls.cpp)}\pahosays{DONE}
-  \item \pahosays{Remove CVA from all plots, but leave it easy to
-    re-add (so we can check what we say in the text).}\pahosays{DONE}
-  \item \pahosays{Make all plots not show CVA (new name for CVA-short)
-    outside the fitted range.}\pahosays{DONE}
-  \item \davidsays{Rewrite CVA-S section to reflect new fit.}
-  \item \jeffsays{Make certain when proof-reading that other prior
+  \item \davidsays{Rewrite CVA-S section to reflect new fit.  Still working.}
+  \item \pahosays{Make certain when proof-reading that other prior
     formulations cannot work in ``separable'' way.}
   \item \davidsays{Be sure that we explain the fitted data and fitting
-    process well.}
-  \item \pahosays{Add $\eta=?$ labels and legends to new fit plot.}\pahosays{DONE}
-  \item \jeffsays{I will diligently read what you guys write!}
+    process well.  Still working.}
+  \item \jeffsays{I will repeatedly read what you guys write!}
   \end{itemize}
@@
 computationally efficient separable version of the contact
-value approach (CVA-S), which is based on a fit to the radial
+value approach (CVA-S)\fixme{this needs editing}, which is based on a fit to the radial
 distribution function that is separable in a way that will enable
@@
 When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
-further simplify this integral as
+further simplify this integral as\fixme{CVA-S???}
 \begin{multline}
@@
   results of Monte Carlo simulations, while the bottom halves show the
-  CVA-S.  On the right are plots of #1 on the paths illustrated in the
+  \fixme{CVA}CVA-S.  On the right are plots of #1 on the paths illustrated in the
   figures to the left.  These plots compare the CVA-S (blue solid line) and CVA
@@
 with the hard wall.  Figures~\ref{fig:pair-distribution}a
-and~\ref{fig:pair-distribution}c compare the results of the CVA-S with
+and~\ref{fig:pair-distribution}c compare the results of the \fixme{CVA}CVA-S with
 Monte Carlo simulations at packing fractions of 0.1 and 0.3
@@
     The top halves of these figures show the results of Monte Carlo
-    simulations, while the bottom halves show the CVA-S.  On the right
+    simulations, while the bottom halves show the \fixme{CVA}CVA-S.  On the right
     are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
@@
 and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
-CVA-S with Monte Carlo results, as well as the approximations of
+\fixme{CVA}CVA-S with Monte Carlo results, as well as the approximations of
 Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
@@
 more than a radius away from contact, we find that any of these
-approaches gives a reasonable prediction, with the CVA-S
+approaches gives a reasonable prediction, with the \fixme{CVA}CVA-S
 underestimating the oscillations in $g^{(2)}$, as expected based on
@@
 distribution function at a packing fraction of 0.3 computed using the
-CVA-S with results from Monte Carlo simulations. In
+\fixme{CVA}CVA-S with results from Monte Carlo simulations. In
 Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
@@
 All methods perform similarly over this range with one exception.  The
-separable CVA-S method once again underestimates the oscillations on
+separable \fixme{CVA}CVA-S method once again underestimates the oscillations on
 the left-hand side of the curve, where $\rr_2$ and $\rr_3$ are
@@
 Figure~\ref{fig:dadz-delta} shows the results for the sticky
-hard-sphere fluid.  The CVA and CVA-S are constructed to produce this
+hard-sphere fluid.  The CVA and \fixme{NOT}CVA-S are constructed to produce this
 result exactly, provided the averaged pair distribution function at

### 2014-07-06 papers/square-well-liquid/histogram-paper.tex
define "flat" as multiple gaussian adjustments Tested on darcy. Took 0:35.94.

@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rdf}
-  \caption{Radial distribution function.\label{fig:rdf}}
-\end{figure}
+%% \begin{figure}
+%%   \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rdf}
+%%   \caption{Radial distribution function.\label{fig:rdf}}
+%% \end{figure}
 
@@
 
+\textcolor{red}{I think maybe we only want to keep ``pushing down the
+  bubbles'' so long as the variance increases.  I'm afraid that if the
+  variance drops we could possibly create a dangerous hole when we try
+  pushing down a sharp and pointy bubble...}
+
 \end{document}

### 2014-07-07 papers/pair-correlation/paper.tex
plot just the nice values of ff, but fit to all of them. Tested on darcy. Took 0:40.89.

@@
   \end{align*}
-  \caption{The fitted $\kappa_{ij}$ matrix.  In addition to these 12
-    parameters, we also optimized the exponetial decay constant
-    $\alpha=\alphaval$.
+  \caption{The fitted $\kappa_{ij}$ matrix.
+    %% In addition to these 12 parameters, we also optimized the
+    %% exponetial decay constant $\alpha=\alphaval$.za
     %% Note that the first column of $\kappa$ corresponds to $j=1$
@@
 maximum error within this radius in $g_S(r)$ at packing fractions up
-to 0.4 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
-\rmaxerr R$.
+to 0.45 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+\rmaxerr R$.  We used in this fit Monte Carlo data for the radial
+distribution function at values of packing fraction at increments of
+0.05 from 0.05 to 0.45. Fig.~\ref{fig:radial-distribution} displays
+our approximation at just under half of the densities that were
+included in the fit.
 

### 2014-07-07 papers/pair-correlation/paper.tex
clean up labels in radial-distribution plot Tested on bennet. Took 0:18.95.

@@
     simulation, and the dashed lines are results of the
-    Gil-Villegas\cite{gil1997statistical} fit.  The dotted extension
+    Gil-Villegas fit~\cite{gil1997statistical}.  The dotted extension
     of each fitted curve indicates the value of the function outside

### 2014-07-07 papers/pair-correlation/figs/short-range-ghs.py

@@ -531,17 +531,17 @@ for i in indexes:
   pylab.axhline(zeros[i] + maxes[i], color='k')
   pylab.axhline(zeros[i] + 1, color='k', ls=':')
   pylab.plot(r_mc[r_mc<fit_rcutoff], g[i*len(r):(i+1)*len(r)][r_mc<fit_rcutoff] + offset,
-             alt_styles['gS'], label='$g_S(r)$')
+             alt_styles['gS'], label='this work $g_S(r)$')
   pylab.plot(r_mc[r_mc<1.1*fit_rcutoff], g[i*len(r):(i+1)*len(r)][r_mc<1.1*fit_rcutoff] + offset,
              alt_styles['gSunfitted'])
   r_gil = r_mc[r_mc < gil_rmax]
   g_gil = g_gil[r_mc < gil_rmax]
   pylab.plot(r_gil[r_gil<gil_rcutoff], g_gil[r_gil<gil_rcutoff] + offset,
-             alt_styles['gil'], label='FIXME gil-villegas')
+             alt_styles['gil'], label="fit by Gil-Villegas")
   pylab.plot(r_gil[r_gil>gil_rcutoff], g_gil[r_gil>gil_rcutoff] + offset,
              alt_styles['gil_unfitted'])
   pylab.plot(r_mc_avged, g_mc_avged + offset,
-             alt_styles['mc'], label='mc')
+             alt_styles['mc'], label='Monte Carlo')
 
   dg = 0.2
   if maxes[i] > 2:
@@ -583,7 +583,7 @@ pylab.axes().set_yticks(ticks)
 pylab.axes().set_yticklabels(ticklabels)
 
 for i in indexes:
-  pylab.text((xlim[0]+xlim[1])/2, zeros[i]+maxes[i]-0.1, r"$\eta = %.2f$" %(ff[i]), ha='center', va='top')
+  pylab.text((2*xlim[0]+xlim[1])/3, zeros[i]+maxes[i]-0.1, r"$\eta = %.2f$" %(ff[i]), ha='center', va='top')
 
 
 handles, labels = ax2.get_legend_handles_labels()

### 2014-07-07 papers/pair-correlation/paper.tex
fix wording issue pointed out by Jeff Tested on bennet. Took 0:18.77.

@@
 of radius with a function of $g_\sigma$.  This enables us to write
-typicaly \jeffsays{'typical?' maybe just 'integrals'} integrals
-involving the pair distribution function as a summation of
-fixed-kernel convolutions, which may be efficiently computed using
-Fast Fourier Transforms (FFTs), in $\mathcal{O}(N\log N)$ time.
+integrals that are linear in the pair distribution function---such as
+Eq.~\ref{eq:mean-field}---as a summation of fixed-kernel convolutions,
+which may be efficiently computed using Fast Fourier Transforms
+(FFTs), in $\mathcal{O}(N\log N)$ time.
 

### 2014-07-08 papers/pair-correlation/paper.tex
clean up text just a bit more Tested on bennet. Took 0:18.58.

@@
 distribution function.  We chose to focus our fit on the range of
-distances $r_{12} < 2\sigma$.  This range of distances is relevant to
+distances $r_{12} \le \maxrfit R$.  This range of distances is relevant to
 the widely used \fixme{add citations?}  Statistical Associating Fluid
@@
 the dispersion free energy given by Eq.~\ref{eq:mean-field}.
-
-For ease of implementation and future extension to larger radii, we
-chose to express our fit with a polynomial function, multiplied by an
-exponential decay.  We constrain our functional form for $g_S(r,
-g_\sigma)$ to match the value at contact:
-\begin{equation}
-  g_S(\sigma, g_\sigma) = g_\sigma.
-\end{equation}
-Beyond this constraint we have:
-\begin{align}
-  g_S(r,g_\sigma) &=
-  1 + \left(\sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} h_\sigma^i
-  \zeta^j\right)e^{-\alpha \zeta} \label{eq:fit-form}
-\end{align}
-where $h_\sigma = g_\sigma - 1$ is the total correlation function at
-contact, and $\zeta \equiv \frac{r}{\sigma}-1$ is a dimensionless
-measure of the distance from contact.
 
@@
     %% Eq.~\ref{eq:fit-form}.
-  }
+  }\label{tab:kappa}
 \end{table}
 
-The Gil-Villegas function is fit to data for $r <= \maxrfit R$.  The
-maximum error within this radius in $g_S(r)$ at packing fractions up
-to 0.45 is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
-\rmaxerr R$.  We used in this fit Monte Carlo data for the radial
-distribution function at values of packing fraction at increments of
-0.05 from 0.05 to 0.45. Fig.~\ref{fig:radial-distribution} displays
+For ease of implementation and future extension to larger radii, we
+chose to express our fit with a fourth-order polynomial.  We constrain
+our functional form for $g_S(r, g_\sigma)$ to match the value
+$g_\sigma$ at contact and to reproduce $g(r)=1$ in the low-density
+($g_\sigma =1$) limit.  Incorporating these constraints we have the
+functional form:
+\begin{align}
+  g_S(r,g_\sigma) &=
+  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
+  \left(\tfrac{r}{\sigma}-1\right)^j.
+  \label{eq:fit-form}
+\end{align}
+where the matrix $\kappa_{ij}$ is determined from our least-squares
+fit to Monte Carlo data for the radial distribution function, over the
+range $2R \le r \le \maxrfit R$, and for packing fractions $\eta \le
+0.45$.  The resulting parameters are displayed in
+Table~\ref{tab:kappa}.  The maximum error in $g_S(r)$ within this
+range is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+\rmaxerr R$.  Fig.~\ref{fig:radial-distribution} displays
 our approximation at just under half of the densities that were

### 2014-07-11 papers/thesis-roth/project.tex
Copied important files from thesis-roth to renormalization. Added fig comparing grand free energy for various iteration depths Tested on elliot. Took 0:33.68.

@@
 
-One particularly troublesome regime is around the critical
-point of a liquid. At the appropriate density and temperature, the
-distinction between liquid and vapor disappears. Beyond this point,
-you have a ``supercritical fluid.'' Supercritical fluids are used in
-many industrial processes.\cite{Perrut00} For example (and of personal
-interest), supercritical fluids are used in coffee decaffeination and
-the extraction of hops resins for brewing beer.
+One particularly troublesome regime is around the critical point of a
+liquid. At the appropriate density and temperature, the distinction
+between liquid and vapor disappears. Beyond this point, you have a
+``supercritical fluid.'' Supercritical fluids are used in many
+industrial processes.\cite{Perrut00} For example, supercritical fluids
+are used in coffee decaffeination and the extraction of hops resins
+for brewing beer.
 
@@
   \caption{Liquid-vapor coexistence for a square well liquid.}
-  \label{fig:coexistance_SW}
+  \label{fig:coexistence_SW}
 \end{figure}
 
-Figure~\ref{fig:coexistance_SW} is a demonstration of this algorithm.
+Figure~\ref{fig:coexistence_SW} is a demonstration of this algorithm.
 

### 2014-07-14 papers/pair-correlation/paper.tex
added citations after widely used SAFT-VR Tested on knightley. Took 1:09.45.

@@
   \item \jeffsays{I will repeatedly read what you guys write!}
+  \item \jeffsays{still concerned about the use of CVA.  It seems
+    sometimes it's used as an approach and sometimes as the resulting
+    pair-dist function.}
   \end{itemize}
@@
 
-In this paper, we introduce a new contact value approach (CVA) for
-the hard-sphere pair distribution function which is suitable for use
-in the creation of classical density functionals based on
-thermodynamic perturbation theory. We then demonstrate a more
-computationally efficient separable version of the contact
-value approach (CVA-S)\fixme{this needs editing}, which is based on a fit to the radial
-distribution function that is separable in a way that will enable
-efficient evaluation of the integral in Eq.~\ref{eq:mean-field}.
+\jeffsays{I've edited the following paragraph but it should be
+  checked.}  In this paper, we introduce a new contact value approach
+(CVA) to approximating the hard-sphere pair distribution function
+which is suitable for use in the creation of classical density
+functionals based on thermodynamic perturbation theory. The resulting
+function is based on a fit to the radial distribution function that is
+separable in a way that will enable efficient evaluation of the
+integral in Eq.~\ref{eq:mean-field}.
 
@@
 \end{align}
-This Contact Value Approach (CVA) for $g^{(2)}(\rr_1,\rr_2)$ is
+This Contact Value Approach for $g^{(2)}(\rr_1,\rr_2)$ is
 constructed to reproduce the exact value for the integral:
@@
 When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
-further simplify this integral as\fixme{CVA-S???}
+further simplify this integral as
 \begin{multline}
@@
 
-We have performed a separable fit to the Monte-Carlo simulated radial
-distribution function.  We chose to focus our fit on the range of
-distances $r_{12} \le \maxrfit R$.  This range of distances is relevant to
-the widely used \fixme{add citations?}  Statistical Associating Fluid
-Theory of Variable Range (SAFT-VR) free energy with square-well
-dispersive attraction developed by Gil-Villegas \emph{et
+Having settled on the basic structure of our function, we futher
+refine it by performing a separable fit to the Monte-Carlo simulated
+radial distribution function.  We chose to focus our fit on the range
+of distances $r_{12} \le \maxrfit R$.  This range of distances is
+relevant to the widely used
+\cite{chapman1989saft,muller2001molecular,tan2008recent} Statistical
+Associating Fluid Theory of Variable Range (SAFT-VR) free energy with
+square-well dispersive attraction developed by Gil-Villegas \emph{et
   al.}~\cite{gil1997statistical}.  For comparison, in
@@
 our functional form for $g_S(r, g_\sigma)$ to match the value
-$g_\sigma$ at contact and to reproduce $g(r)=1$ in the low-density
-($g_\sigma =1$) limit.  Incorporating these constraints we have the
+$g_\sigma$ at contact and to reproduce the $g(r)=1$ result of the
+low-density limit.  Incorporating these constraints we have the
 functional form:
@@
 
-\newcommand{\plotcomp}[1]{The top halves of these figures show the
-  results of Monte Carlo simulations, while the bottom halves show the
-  \fixme{CVA}CVA-S.  On the right are plots of #1 on the paths illustrated in the
-  figures to the left.  These plots compare the CVA-S (blue solid line) and CVA
-  (cyan dotted line) with Monte Carlo results (black circle)
-  and the results of Sokolowski and Fischer
-  (red dashed line)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
-  (green dot-dashed line)~\cite{fischer1980born}.  The latter is only plotted at
-  contact, where it is defined}
+\newcommand{\plotcomp}[1]{\jeffsays{there's no green dash label in the
+    legend above.  Do we want the green dashes?}The top halves of
+  these figures show the results of Monte Carlo simulations, while the
+  bottom halves show the CVA.  On the right are plots of #1 on the
+  paths illustrated in the figures to the left.  These plots compare
+  the CVA (blue solid line), Monte Carlo results (black circles), the
+  results of Sokolowski and Fischer (red dashed
+  line)~\cite{sokolowski1992role}, and those of Fischer and Methfessel
+  (green dot-dashed line)~\cite{fischer1980born}.  The latter is only
+  plotted at contact, where it is defined}
 
@@
 with the hard wall.  Figures~\ref{fig:pair-distribution}a
-and~\ref{fig:pair-distribution}c compare the results of the \fixme{CVA}CVA-S with
+and~\ref{fig:pair-distribution}c compare the results of the CVA with
 Monte Carlo simulations at packing fractions of 0.1 and 0.3
@@
     The top halves of these figures show the results of Monte Carlo
-    simulations, while the bottom halves show the \fixme{CVA}CVA-S.  On the right
+    simulations, while the bottom halves show the CVA.  On the right
     are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
@@
 in Figures~\ref{fig:pair-distribution}a
-and~\ref{fig:pair-distribution}c.  These plots compare the CVA and
-\fixme{CVA}CVA-S with Monte Carlo results, as well as the approximations of
-Sokolowski and Fischer~\cite{sokolowski1992role} and of Fischer and
-Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1 and 0.3.
-The approach of Fischer and Methfessel is only defined when the two
-spheres are in contact, and is therefore only plotted on that segment
-of the path.  As an input to both previous approximations and to the
-CVA, we use the hard sphere radial distribution function found with
-Monte Carlo simulation, interpolated as necessary. We find that both
-previous approximations to the pair distribution function predict the
-qualitative angular dependence of the pair distribution function at
-contact better than this work.  However, in each case the pair
+and~\ref{fig:pair-distribution}c.  These plots compare the CVA with
+Monte Carlo results, as well as the approximations of Sokolowski and
+Fischer~\cite{sokolowski1992role} and of Fischer and
+Methfessel~\cite{fischer1980born} at the same packing fractions of 0.1
+and 0.3.  The approach of Fischer and Methfessel is only defined when
+the two spheres are in contact, and is therefore only plotted on that
+segment of the path.  As an input to both previous approximations and
+to the CVA, we use the hard sphere radial distribution function found
+with Monte Carlo simulation, interpolated as necessary. We find that
+both previous approximations to the pair distribution function predict
+the qualitative angular dependence of the pair distribution function
+at contact better than this work.  However, in each case the pair
 distribution function has a systematic error at contact---either too
-high or too low.  While our approximation is smoother than either
-of the existing approaches, its errors will have a tendency to cancel
+high or too low.  While our approximation is smoother than either of
+the existing approaches, its errors will have a tendency to cancel
 when used in a perturbation expansion.  At higher densities, the
@@
 radial distribution function from simulation.  When the two points
-$\rr_1$ and $\rr_2$ are both
-more than a radius away from contact, we find that any of these
-approaches gives a reasonable prediction, with the \fixme{CVA}CVA-S
-underestimating the oscillations in $g^{(2)}$, as expected based on
-the fit in Fig.~\ref{fig:radial-distribution}.
+$\rr_1$ and $\rr_2$ are both more than a radius away from contact, we
+find that any of these approaches gives a reasonable prediction, with
+the CVA underestimating the oscillations in $g^{(2)}$, as
+expected based on the fit in Fig.~\ref{fig:radial-distribution}.
 
@@
 distribution function at a packing fraction of 0.3 computed using the
-\fixme{CVA}CVA-S with results from Monte Carlo simulations. In
+CVA with results from Monte Carlo simulations. In
 Figure~\ref{fig:triplet-contact-distribution}a the spheres at $\rr_1$
@@
 a third sphere can just fit between them; and in both figures $\rr_3$
-is varied. The test-particle position for the CVA-S in each case is
+is varied. The test-particle position for the CVA in each case is
 $\rr_1$, which is on the left-hand side of the figure. We see
-reasonable agreement, with fewer oscillations in the CVA-S at larger
+reasonable agreement, with fewer oscillations in the CVA at larger
 distances. Also, the Monte Carlo results have the expected left-right
-symmetry, while the CVA-S has an asymmetry introduced with
-the test particle due to errors in the pair distribution
-function. Most notably, the CVA-S has less angular dependence at
-contact with the test particle than it does with the sphere at
-$\rr_2$.
+symmetry, while the CVA has an asymmetry introduced with the test
+particle due to errors in the pair distribution function. Most
+notably, the CVA has less angular dependence at contact with the
+test particle than it does with the sphere at $\rr_2$.
 
@@
 All methods perform similarly over this range with one exception.  The
-separable \fixme{CVA}CVA-S method once again underestimates the oscillations on
-the left-hand side of the curve, where $\rr_2$ and $\rr_3$ are
-separated by a large distance.  On the path where $\rr_2$ and $\rr_3$
-are closer, the CVA-S method compares favorably with the other
-approximations.  This is consistent with
-Fig.~\ref{fig:radial-distribution}, which shows that $g_S(r)$ fits
-well the first peak of $g(r)$, but has fewer oscillations at large
-distances.
+CVA method once again underestimates the oscillations on the left-hand
+side of the curve, where $\rr_2$ and $\rr_3$ are separated by a large
+distance.  On the path where $\rr_2$ and $\rr_3$ are closer, the CVA
+method compares favorably with the other approximations.  This is
+consistent with Fig.~\ref{fig:radial-distribution}, which shows that
+$g_S(r)$ fits well the first peak of $g(r)$, but has fewer
+oscillations at large distances.
 
@@
 Figure~\ref{fig:dadz-delta} shows the results for the sticky
-hard-sphere fluid.  The CVA and \fixme{NOT}CVA-S are constructed to produce this
+hard-sphere fluid.  The CVA is constructed to produce this
 result exactly, provided the averaged pair distribution function at
@@
 shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
-This potential stresses the CVA-S, since the edge of the attraction
+This potential stresses the CVA, since the edge of the attraction
 occurs at the dip in the radial distribution function at this density,

### 2014-07-14 papers/pair-correlation/paper.tex
fix another CVA-S usage and cut some old todo items Tested on darcy. Took 0:37.79.

@@
   \begin{itemize}
-  \item \davidsays{Rewrite CVA-S section to reflect new fit.  Still working.}
-  \item \pahosays{Make certain when proof-reading that other prior
-    formulations cannot work in ``separable'' way.}
-  \item \davidsays{Be sure that we explain the fitted data and fitting
-    process well.  Still working.}
-  \item \jeffsays{I will repeatedly read what you guys write!}
   \item \jeffsays{still concerned about the use of CVA.  It seems
@@
 \begin{multline}
-  F_1^{\textit{CVA-S}} =
+  F_1^{\textit{CVA}} =
   \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\

### 2014-07-14 papers/pair-correlation/paper.tex
fix bug where we were not including Fischer in the legend Tested on darcy. Took 0:35.14.

@@
 
-\newcommand{\plotcomp}[1]{\jeffsays{there's no green dash label in the
-    legend above.  Do we want the green dashes?}The top halves of
+\newcommand{\plotcomp}[1]{The top halves of
   these figures show the results of Monte Carlo simulations, while the

### 2014-07-15 papers/pair-correlation/paper.tex
Jeff's edit seems fine Tested on bennet. Took 0:20.01.

@@
 
-\jeffsays{I've edited the following paragraph but it should be
-  checked.}  In this paper, we introduce a new contact value approach
+In this paper, we introduce a new contact value approach
 (CVA) to approximating the hard-sphere pair distribution function

### 2014-07-15 papers/pair-correlation/paper.tex
a few "davidsays" Tested on bennet. Took 0:18.83.

@@
 All methods perform similarly over this range with one exception.  The
-CVA method once again underestimates the oscillations on the left-hand
+CVA once again underestimates the oscillations on the left-hand
 side of the curve, where $\rr_2$ and $\rr_3$ are separated by a large
 distance.  On the path where $\rr_2$ and $\rr_3$ are closer, the CVA
-method compares favorably with the other approximations.  This is
+compares favorably with the other approximations.  This is
 consistent with Fig.~\ref{fig:radial-distribution}, which shows that
@@
     sticky hard-sphere fluid defined by a pair potential
+    \davidsays{Probablyt cut out inverse sixth?}
     $\delta(\sigma-r+\delta)$ where $\sigma$ is the hard-sphere
@@
 Figure~\ref{fig:dadz-delta} shows the results for the sticky
-hard-sphere fluid.  The CVA is constructed to produce this
-result exactly, provided the averaged pair distribution function at
-contact from Ref.~\citenum{schulte2012using} is exact.  As expected,
-we see excellent agreement with the Monte Carlo simulation results,
-while the approximations of Fischer and Sokolowski each show
-deviations near the interface.  Figure~\ref{fig:dadz-square-well}
-shows the same curve from Eq.~\ref{eq:da1} for the square-well fluid.
-This potential stresses the CVA, since the edge of the attraction
-occurs at the dip in the radial distribution function at this density,
-with poor results for the bulk fluid.  We see that over these length
-scales, both the CVA and Sokolowski's approximation are very accurate.
-Finally, in the case of the power-law attraction all approximations
-tested work very well.
+hard-sphere fluid.  The CVA is constructed to produce this result
+exactly, provided the averaged pair distribution function at contact
+from Ref.~\citenum{schulte2012using} is exact.  As expected, we see
+excellent agreement with the Monte Carlo simulation results, while the
+approximations of Fischer and Sokolowski each show deviations near the
+interface.  Figure~\ref{fig:dadz-square-well} shows the same curve
+from Eq.~\ref{eq:da1} for the square-well fluid.  This potential
+stresses the CVA, since the edge of the attraction occurs at the dip
+in the radial distribution function at this density, with poor results
+for the bulk fluid.  \davidsays{The last sentence is inaccurate.}  We
+see that over these length scales, both the CVA and Sokolowski's
+approximation are very accurate.  Finally, in the case of the
+power-law attraction all approximations tested work very well.
 
@@
 accurate for both short-range and long-range attractions.
+\davidsays{Probably cut the long-range attractions.}
 

### 2014-07-15 papers/pair-correlation/paper.tex
edits from Jeff and David together, working in harmony Tested on bennet. Took 0:18.74.

@@
 
-\fixme{
-  TO DO:
-  \begin{itemize}
-  \item \jeffsays{still concerned about the use of CVA.  It seems
-    sometimes it's used as an approach and sometimes as the resulting
-    pair-dist function.}
-  \end{itemize}
-}
-
 The standard approach in liquid state theory is to model a liquid as a
@@
 At its core, thermodynamic perturbation theory---sometimes referred to
-as the cluster, or high-temperature expansion---is an expansion of the
+as the high-temperature expansion---is an expansion of the
 free energy in powers of a small parameter, which is typically a
@@
 square-well dispersive attraction developed by Gil-Villegas \emph{et
-  al.}~\cite{gil1997statistical}.  For comparison, in
-Fig.~\ref{fig:radial-distribution} we plot our fit, Monte-Carlo data,
-and the radial distribution function of Gil-Villegas \emph{et al.},
-which we have extracted from their approximation for the first term in
-the dispersion free energy given by Eq.~\ref{eq:mean-field}.
+  al.}~\cite{gil1997statistical}.
+%
+Although we consider this range of radii particularly interesting,
+this is not a fundamental limit on the approach, as one could readily
+extend the fit to larger radii by including additional fitting
+parameters.
+%
+For comparison, in Fig.~\ref{fig:radial-distribution} we plot our fit,
+Monte-Carlo data, and the radial distribution function of Gil-Villegas
+\emph{et al.}, which we have extracted from their approximation for
+the first term in the dispersion free energy given by
+Eq.~\ref{eq:mean-field}.
 
@@
 
-% Figure~\ref{fig:triplet-contact-distribution} compares the triplet
-% distribution function at a packing fraction of 0.3 computed using the
-% CVA-S with results from Monte Carlo simulation.  In each subfigure,
-% $\rr_1$ and $\rr_2$ are held fixed, while the third position $\rr_3$
-% is varied.  The test-particle position in each case is $\rr_1$, which
-% is on the left-hand side of the figure.  In the color plots on the
-% left, the first and second spheres are displayed in gray.  On the
-% bottom half of figures~\ref{fig:triplet-contact-distribution}a
-% and~\ref{fig:triplet-contact-distribution}c is the triplet
-% distribution function as computed using the CVA-S.  The top half of
-% those figures displays the triplet distribution function from Monte
-% Carlo simulations.  Figures~\ref{fig:triplet-contact-distribution}b
-% and~\ref{fig:triplet-contact-distribution}d show the triplet
-% distribution function along the paths illustrated in the corresponding
-% 2D plots.
-
-% Figures~\ref{fig:triplet-contact-distribution}a
-% and~\ref{fig:triplet-contact-distribution}b show the triplet
-% distribution function when two of the positions $\rr_1$ and $\rr_2$
-% are in contact.  The triplet distribution function predicted by the
-% CVA-S is slightly asymmetric, indicating a failure in thermodynamic
-% consistency.  Once again, we observe the lack of oscillations at large
-% distances.  Figures~\ref{fig:triplet-contact-distribution}c
-% and~\ref{fig:triplet-contact-distribution}d show the triplet
-% distribution function when two of the positions $\rr_1$ and $\rr_2$
-% are spaced so that the third could just fit between them.  Again, the
-% triplet distribution function predicted by the CVA-S is slightly
-% asymmetric, and lacks the oscillations seen in the Monte Carlo
-% simulations at large distances.  \fixme{We need to discuss the plots
-%   and the directions and arrows.} In both cases, our agreement with
-% simulation is comparable to that of previous approximations.
+\section{Accuracy in thermodynamic perturbation theory}
 
@@
   \end{subfigure}
-  \begin{subfigure}{1.0\columnwidth}
-    \includegraphics[width=\columnwidth]{figs/dadz-inverse-sixth-3.pdf}
-    \vspace{-0.8cm}
-    \caption{Hard-core inverse-sixth fluid}\label{fig:dadz-inverse-sixth}
-  \end{subfigure}
   \caption{Plot of $\frac{dF_1}{dz}$ near a hard wall.  (a) shows a
     sticky hard-sphere fluid defined by a pair potential
-    \davidsays{Probablyt cut out inverse sixth?}
     $\delta(\sigma-r+\delta)$ where $\sigma$ is the hard-sphere
-    diameter, and $\delta$ is an infinitesimal distance; (b) shows a
+    diameter, and $\delta$ is an infinitesimal distance; and (b) shows a
     square well fluid defined by a pair potential $\Theta(1.79
-    \sigma-r)$; and (c) shows a hard-core inverse-sixth potential
-    fluid with an attractive pair potential proportional to $r^{-6}$.
+    \sigma-r)$.
   }
@@
 \end{figure}
-
-\section{Accuracy in thermodynamic perturbation theory}
 
@@
 
-We plot this quantity for three representative
-pair potentials near a hard wall in Fig.~\ref{fig:dadz}.  We have
-chosen to illustrate a delta-function interaction at contact
-(i.e. ``sticky hard spheres''); a hard-core square-well fluid, with
-the length-scale of interaction taken from the optimal SAFT model for
-water found by Clark \emph{et al.}~\cite{clark2006developing}; and a
-$1/r^6$ power-law attraction.  These pair potentials represent a
-short-range interaction, a medium-range discontinuous interaction, and
-a long-range smooth interaction.
+We plot this quantity for two representative pair potentials near a
+hard wall in Fig.~\ref{fig:dadz}.  We have chosen to illustrate a
+delta-function interaction at contact (i.e. ``sticky hard spheres''),
+and a hard-core square-well fluid, with the length-scale of
+interaction taken from the optimal SAFT model for water found by Clark
+\emph{et al.}~\cite{clark2006developing}.  These pair potentials
+represent both a very short-range interaction and a medium-range
+interaction.
 
@@
 interface.  Figure~\ref{fig:dadz-square-well} shows the same curve
-from Eq.~\ref{eq:da1} for the square-well fluid.  This potential
-stresses the CVA, since the edge of the attraction occurs at the dip
-in the radial distribution function at this density, with poor results
-for the bulk fluid.  \davidsays{The last sentence is inaccurate.}  We
-see that over these length scales, both the CVA and Sokolowski's
-approximation are very accurate.  Finally, in the case of the
-power-law attraction all approximations tested work very well.
+from Eq.~\ref{eq:da1} for the square-well fluid.  For the square well
+fluid both the CVA and Sokolowski's approximation give excellent
+agreement with simulation.
 
@@
 Tests of the pair distribution function in integrals that arise in
-thermodynamic perturbation theory suggest that our approximation is
-accurate for both short-range and long-range attractions.
-\davidsays{Probably cut the long-range attractions.}
+thermodynamic perturbation theory suggest that the CVA is
+accurate for attractions up to the distance to which the radial
+distribution function is fit.
 

### 2014-07-16 papers/pair-correlation/paper.tex
a few edits based on my reading of paper Tested on bennet. Took 0:19.82.

@@
 distribution function at contact for inhomogeneous density
-distributions. We further develop a separable form of this
-approximation that achieves greater computational efficiency by using
-exclusively fixed-kernel convolutions and thus allowing an
-implementation using fast Fourier transforms. We compare results for
-our pair distribution approximation with two previously published
-works and Monte-Carlo simulation, showing favorable results.
+distributions. This approach achieves greater computational efficiency
+that previous approaches by using exclusively fixed-kernel
+convolutions and thus allowing an implementation using fast Fourier
+transforms. We compare results for our pair distribution approximation
+with two previously published works and Monte-Carlo simulation,
+showing favorable results.
 \end{abstract}
@@
 density~\cite{hansen2006theory}.  This elegant approach lends itself
-to computation with DFT, and can be used to compute and plot the pair
+to computation \emph{using} DFT, and can be used to compute and plot the pair
 distribution function, but requires a full free-energy minimization
@@
 approximation~\cite{plischke1986pair}.
+
+Lado recently introduced a new and improved efficient algorithm for
+implementing integral equation theory for inhomogeneous fluids, which
+computes $g^{(2)}(\rr_1,\rr_2)$~\cite{lado2009efficient}.  While this
+approach is two orders of magnitude more efficient than previous
+implementations, it remains a computationally expensive approach, and
+unsuitable for repeated evaluation within a free-energy minimization.
 
@@
 \emph{two} spheres fixed~\cite{gonzalez1999test}.
-
-Lado recently introduced a new and improved efficient algorithm for
-implementing integral equation theory for inhomogeneous fluids, which
-computes $g^{(2)}(\rr_1,\rr_2)$~\cite{lado2009efficient}.  While this
-approach is two orders of magnitude more efficient than previous
-implementations, it remains a computationally expensive approach, and
-unsuitable for repeated evaluation within a free-energy minimization.
 
@@
 inhomogeneous scenarios.  These approximations differ in what density
-is used when evaluating the radial distribution function $g(r,\eta)$
+is used when evaluating the radial distribution function $g(r,n)$
 or in how radial distributions evaluated at different densities are
@@
 \end{align}
-where their averaged density $\bar{n}(\rr)$ is given by:
+where their averaged density $\bar{n}(\rr)$ given by
 \begin{align}
@@
 was arrived at by fitting to Monte Carlo simulation.  Although
-Although Eq.~\ref{eq:sokolowski} has the advantages of only involving
+Eq.~\ref{eq:sokolowski} has the advantage of only involving
 density averages at the points at which the pair distribution function
@@
 \end{align}
-We use the contact-value theorem to derive the exact formula:
+In \cite{schulte2012using} we use the contact-value theorem to derive the exact formula:
 \begin{align}
@@
 
-The CVA requires the radial distribution function as a function of $r$
+The CVA requires the radial distribution function expressed as a function of $r$
 and $g_\sigma$.  One approach to constructing $g(r,g_\sigma)$ is to
-compute $\eta$ from $g_\sigma$ and then find $g(r,\eta)$ using either
+compute $n$ from $g_\sigma$ (using e.g. the Carnahan-Starling formula
+for $g_\sigma$)
+and then find $g(r,n)$ using either
 some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
@@
 \begin{multline}
-  F_1^{\textit{CVA}} =
-  \sum_i \tfrac12 \!\! \iint \!\! \bigg( n(\rr_1)  \bigg) \\
-                             \bigg( a_i(r_{12})\Phi(r_{12}) \bigg) \\
-                             \bigg( b_i(g_\sigma(\rr_2))n(\rr_2) \bigg)
-  d\rr_1d\rr_2
+  F_1^{\textit{CVA}} \! = \!
+  \sum_i \!\tfrac12 \!\! \int \!\! n(\rr_1) \!\!
+                            \int \!\!\!
+                            %\big(
+                            a_i(r_{12})\Phi(r_{12})
+                            %\big)
+                            %\big(
+                            b_i(
+g_\sigma(\rr_2))n(\rr_2)
+                            %\big)
+  d\rr_2d\rr_1
 \end{multline}
@@
 \end{align}
-where the matrix $\kappa_{ij}$ is determined from our least-squares
+where the matrix $\kappa_{ij}$ is determined from a least-squares
 fit to Monte Carlo data for the radial distribution function, over the
@@
   these figures show the results of Monte Carlo simulations, while the
-  bottom halves show the CVA.  On the right are plots of #1 on the
+  bottom halves show the CVA \fixme{over the range of radii for which
+    the fit was performed}.  On the right are plots of #1 on the
   paths illustrated in the figures to the left.  These plots compare
@@
 Monte Carlo simulations at packing fractions of 0.1 and 0.3
-respectively. We see reasonable agreement at the lower density, with
-some reduced oscillations at larger distances, and a flatter angular
+respectively. We see reasonable agreement at the lower density, with a flatter angular
 dependence when the two spheres are in contact.  At the higher
@@
     The top halves of these figures show the results of Monte Carlo
-    simulations, while the bottom halves show the CVA.  On the right
+    simulations, while the bottom halves show the CVA \fixme{over the
+      range of radii for which the fit was performed}.  On the right
     are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
@@
 $\rr_1$ and $\rr_2$ are both more than a radius away from contact, we
-find that any of these approaches gives a reasonable prediction, with
-the CVA underestimating the oscillations in $g^{(2)}$, as
-expected based on the fit in Fig.~\ref{fig:radial-distribution}.
+find that any of these approaches gives a reasonable prediction.
 
@@
 is varied. The test-particle position for the CVA in each case is
-$\rr_1$, which is on the left-hand side of the figure. We see
-reasonable agreement, with fewer oscillations in the CVA at larger
-distances. Also, the Monte Carlo results have the expected left-right
-symmetry, while the CVA has an asymmetry introduced with the test
-particle due to errors in the pair distribution function. Most
-notably, the CVA has less angular dependence at contact with the
-test particle than it does with the sphere at $\rr_2$.
+$\rr_1$, which is on the left-hand side of the figure. As before, we
+see reasonable agreement with simulation. Also, the Monte Carlo
+results have the expected left-right symmetry, while the CVA has a
+small asymmetry introduced with the test particle due to errors in the
+pair distribution function.
 
@@
 %
-All methods perform similarly over this range with one exception.  The
-CVA once again underestimates the oscillations on the left-hand
-side of the curve, where $\rr_2$ and $\rr_3$ are separated by a large
-distance.  On the path where $\rr_2$ and $\rr_3$ are closer, the CVA
-compares favorably with the other approximations.  This is
-consistent with Fig.~\ref{fig:radial-distribution}, which shows that
-$g_S(r)$ fits well the first peak of $g(r)$, but has fewer
-oscillations at large distances.
+All methods tested perform similarly over their range of validity.
 
@@
 interface.  Figure~\ref{fig:dadz-square-well} shows the same curve
-from Eq.~\ref{eq:da1} for the square-well fluid.  For the square well
-fluid both the CVA and Sokolowski's approximation give excellent
-agreement with simulation.
+from Eq.~\ref{eq:da1} for the square-well fluid.  In this case both
+the CVA and Sokolowski's approximation give excellent agreement with
+simulation.
 
@@
 accurate for attractions up to the distance to which the radial
-distribution function is fit.
+distribution function is fit, and is an improvement over existing
+approximations near contact.
 

### 2014-07-16 papers/pair-correlation/paper.tex
Removed an incorrect period in pair paper. Tested on knightley. Took 0:46.24.
edits with Paho and Jeff Tested on bennet. Took 0:22.58.

@@
 distributions. This approach achieves greater computational efficiency
-that previous approaches by using exclusively fixed-kernel
+than previous approaches by enabling the use of exclusively fixed-kernel
 convolutions and thus allowing an implementation using fast Fourier
@@
 as the high-temperature expansion---is an expansion of the
-free energy in powers of a small parameter, which is typically a
-pairwise attractive interaction:
+free energy in powers of a small parameter, which is typically the
+product of a
+pairwise attractive interaction with the inverse temperature $\beta$:
 \begin{align}
@@
 Formally, this requires the pair distribution function as a functional
-of the density $n(\rr)$.  In Section~\ref{sec:gV}, we will introduce
+of the density $n(\rr)$.  In Section~\ref{sec:gV}, we introduce
 existing theoretical approaches for computing
 $g^{(2)}_{HS}(\rr_1,\rr_2)$ given the external potential felt by the
-hard spheres.  In Section~\ref{sec:gn}, we will introduce existing
+hard spheres.  In Section~\ref{sec:gn}, we introduce existing
 approximations for the hard-sphere pair distribution that are
@@
 function is based on a fit to the radial distribution function that is
-separable in a way that will enable efficient evaluation of the
+separable in a way that enables efficient evaluation of the
 integral in Eq.~\ref{eq:mean-field}.
@@
 there are several approaches that have been used to compute the pair
-distribution function.  We will review these approaches here.  The
+distribution function.  We review these approaches here.  The
 classic (and earliest) approach for computing the pair distribution
@@
 approximation~\cite{plischke1986pair}.
-
-Lado recently introduced a new and improved efficient algorithm for
+%
+Lado recently introduced a new and more efficient algorithm for
 implementing integral equation theory for inhomogeneous fluids, which
@@
 implementations, it remains a computationally expensive approach, and
-unsuitable for repeated evaluation within a free-energy minimization.
+unsuitable for repeated evaluation within a free-energy minimization
+as required by DFT.
 
@@
 function, which extend the radial distribution function to
-inhomogeneous scenarios.  These approximations differ in what density
-is used when evaluating the radial distribution function $g(r,n)$
-or in how radial distributions evaluated at different densities are
-combined to yield an effective function
-\cite{toxvaerd1973statistical}.
+inhomogeneous scenarios.  These approximations differ both in what
+density to use when evaluating the radial distribution function
+$g(r;n)$, and in how to combine the radial distribution function
+evaluated at these densities~\cite{toxvaerd1973statistical}.
 
@@
   g^{(2)}(\rr_1,\rr_2) &\approx
-  g\left(r_{12}, n\left(\frac{\rr_1+\rr_2}{2}\right)\right)
+  g\left(r_{12}; n\left(\frac{\rr_1+\rr_2}{2}\right)\right)
      & \text{midpoint} \\
   g^{(2)}(\rr_1,\rr_2) &\approx
-  g\left(r_{12}, \frac{n(\rr_1)+n(\rr_2)}{2}\right)
+  g\left(r_{12}; \frac{n(\rr_1)+n(\rr_2)}{2}\right)
      & \text{mean density} \\
   g^{(2)}(\rr_1,\rr_2) &\approx
-  \frac{g(r_{12},n(\rr_1))+g(r_{12},n(\rr_2))}{2}
+  \frac{g(r_{12};n(\rr_1))+g(r_{12};n(\rr_2))}{2}
      & \text{mean function}
@@
 unity, which cannot occur in the bulk reference system that defines
-$g(r, n)$.  The above papers restrict themselves to the liquid-vapor
+$g(r; n)$.  The above papers restrict themselves to the liquid-vapor
 interface, which does not exhibit this pathology.
@@
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}, n_3\left(\tfrac12
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12}; n_3\left(\tfrac12
   (\rr_1+\rr_2)\right)\right)
@@
 %
-Tang \emph{et. al} employed an approximation for the pair distribution
+Tang \emph{et al.} employed an approximation for the pair distribution
 function that is similar to that of Fischer and Methfessel, but with a
-self-consistent weighted density computed with weighting function that
+self-consistent weighted density computed with a weighting function that
 is itself dependent on the weighted density~\cite{tang1991density}.
@@
 Sokolowski and Fischer addressed the shortcomings of the theory of
-Fischer and Methfessel by modifying this approach to use
-averages, centered on the two points $\rr_1$ and $\rr_2$:
+Fischer and Methfessel by modifying this approach to use density
+averages centered on the two points $\rr_1$ and $\rr_2$:
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12},
+  g^{(2)}(\rr_1,\rr_2) \approx g\left(r_{12};
   \tfrac12(\bar{n}(\rr_1)+\bar{n}(\rr_2))\right)
@@
 is desired, it remains sufficiently computationally cumbersome that it
-has only been used in two papers studying the liquid vapor
+has only been used in two papers studying the one-dimensional liquid vapor
 interface~\cite{wadewitz2000application, winkelmann2001liquid}.
@@
 
-In a previous paper~\cite{schulte2012using}, we introduced a
+In a previous paper~\cite{schulte2012using}, we introduce a
 functional that gives a good approximation for the pair distribution
@@
 $g_{\sigma}(\rr)$ evaluated at the two points:
-%
-%% We therefore seek a function form
-%% for our radial distribution that is dependent upon $g_{\sigma}(\rr)$
-%% and that will approach the correct solution in the homogeneous limit
-%% and also in the limit of contact distance between the two points.
 \begin{align}
-  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}, g_\sigma(\rr_1)) +
-    g(r_{12}, g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
+  g^{(2)}(\rr_1,\rr_2) = \frac{g(r_{12}; g_\sigma(\rr_1)) +
+    g(r_{12}; g_\sigma(\rr_2))}{2}. \label{eq:g2-our-mean}
 \end{align}
@@
 The CVA requires the radial distribution function expressed as a function of $r$
-and $g_\sigma$.  One approach to constructing $g(r,g_\sigma)$ is to
+and $g_\sigma$.  One approach to constructing $g(r;g_\sigma)$ is to
 compute $n$ from $g_\sigma$ (using e.g. the Carnahan-Starling formula
 for $g_\sigma$)
-and then find $g(r,n)$ using either
+and then find $g(r;n)$ using either
 some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
 simulation results.  In this paper we use a fit to Monte Carlo results
-for the radial distribution function $g(r)$, which facilitates
+for the radial distribution function $g(r)$ which facilitates
 efficient computation.  When evaluating the approximations of earlier
 works described in Section~\ref{sec:gn}, we use interpolated Monte
-Carlo results for $g(r,n)$.
+Carlo results for $g(r;n)$.
 
@@
 \section{Making the CVA efficient}
-The existing approaches to approximating the pair correlation function
-outlined in Section~\ref{sec:gn} have not been widely used in the
-construction of density functionals based on thermodynamic
+The existing approaches to approximating the pair distribution
+function outlined in Section~\ref{sec:gn} have not been widely used in
+the construction of density functionals based on thermodynamic
 perturbation theory, largely due to their computational complexity.
-While our CVA as described in the previous section is, in terms of
-accuracy, an incremental improvement over these existing approaches,
-it makes significant gains in terms of computational efficiency,
-allowing for its more managable use in such a construction.  We
-address the issue by developing a \emph{separable} fit to the radial
+While our CVA provides only an incremental improvement in accuracy,
+its construction enables significant gains in computational
+efficiency, allowing for practical application in density functionals.
+We achieve this gain by developing a \emph{separable} fit to the radial
 distribution function of the hard-sphere fluid (see
@@
 \begin{align}
-  g_S(|r_{12}|, g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+  g(|r_{12}|; g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
 \end{align}
@@
 of radius with a function of $g_\sigma$.  This enables us to write
-integrals that are linear in the pair distribution function---such as
-Eq.~\ref{eq:mean-field}---as a summation of fixed-kernel convolutions,
-which may be efficiently computed using Fast Fourier Transforms
-(FFTs), in $\mathcal{O}(N\log N)$ time.
-
-We begin by examining the lowest-order correction in TPT, given by
-Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
-Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral:
-\begin{align}
-  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
-  g(r_{12},g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
-  d\rr_1d\rr_2
-\end{align}
-When we introduce the separable form for $g(r_{12},g_\sigma)$ we can
-further simplify this integral as
-\begin{multline}
-  F_1^{\textit{CVA}} \! = \!
-  \sum_i \!\tfrac12 \!\! \int \!\! n(\rr_1) \!\!
-                            \int \!\!\!
-                            %\big(
-                            a_i(r_{12})\Phi(r_{12})
-                            %\big)
-                            %\big(
-                            b_i(
-g_\sigma(\rr_2))n(\rr_2)
-                            %\big)
-  d\rr_2d\rr_1
-\end{multline}
-where the functional is written as a summation of integrals, each of
-which is an integral of simple convolution in three dimensions.  Thus,
-each of these integrals may be computed in $\mathcal{O}(N\log N)$
-time, which is the same scaling behavior that is common to the most
-common and accurate classical density functional theories, most
-notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
-  roth2002whitebear}.
+integrals---such as Eq.~\ref{eq:mean-field}---that are linear in the
+pair distribution function as a summation of fixed-kernel
+convolutions, which may be efficiently computed using Fast Fourier
+Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.
 
@@
 
+We begin by examining the lowest-order correction in TPT, given by
+Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
+Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral
+\begin{align}
+  F_1^{\textit{CVA}} &= \tfrac12 \!\! \iint \!\!
+  g(r_{12};g_\sigma(\rr_2))n(\rr_1)n(\rr_2)\Phi(|\rr_1-\rr_2|)
+  d\rr_1d\rr_2.
+\end{align}
+When we introduce the separable form for $g(r_{12};g_\sigma)$ we can
+further simplify this integral as
+\begin{align}
+  F_1^{\textit{CVA}} \! = \!
+  \sum_i \!\tfrac12 \!\! \int \!\! n(\rr_1) \!\!
+                            \int \!\!\!
+                            %\big(
+                            a_i(r_{12})\Phi(r_{12})
+                            %\big)
+                            %\big(
+                            b_i(
+g_\sigma(\rr_2))n(\rr_2)
+                            %\big)
+  d\rr_2d\rr_1
+\end{align}
+where the functional is written as a summation of integrals, each of
+which is an integral of a simple convolution in three dimensions.  Thus,
+each of these integrals may be computed in $\mathcal{O}(N\log N)$
+time, which is the same scaling behavior that is common to the most
+common and accurate classical density functional theories, most
+notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
+  roth2002whitebear}.
+
 \section{A separable fit for the radial distribution function}\label{sec:separable-fit}
 
-Having settled on the basic structure of our function, we futher
-refine it by performing a separable fit to the Monte-Carlo simulated
-radial distribution function.  We chose to focus our fit on the range
-of distances $r_{12} \le \maxrfit R$.  This range of distances is
-relevant to the widely used
-\cite{chapman1989saft,muller2001molecular,tan2008recent} Statistical
+Having settled on the basic structure of our function, we further
+refine it by performing a separable fit to the radial distribution
+function from Monte Carlo simulation.  We chose to focus our fit on
+the range of distances $r_{12} \le \maxrfit R$.  This range is
+relevant to the widely used \cite{chapman1989saft,
+  muller2001molecular, tan2008recent} Statistical
 Associating Fluid Theory of Variable Range (SAFT-VR) free energy with
@@
 Although we consider this range of radii particularly interesting,
-this is not a fundamental limit on the approach, as one could readily
+this is not a fundamental limit of the approach, as one could readily
 extend the fit to larger radii by including additional fitting
@@
 Eq.~\ref{eq:mean-field}.
+
+For ease of implementation and future extension to larger radii, we
+chose to express our fit with a fourth-order polynomial.  We constrain
+our functional form such that $g(r; g_\sigma)$ reduces to $g_\sigma$
+at contact and approaches $g(r)=1$ in the low-density limit.
+Incorporating these constraints we have the functional form
+\begin{align}
+  g(r;g_\sigma) &=
+  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
+  \left(\tfrac{r}{\sigma}-1\right)^j,
+  \label{eq:fit-form}
+\end{align}
+where the matrix $\kappa_{ij}$ is determined from a least-squares
+fit to Monte Carlo data for the radial distribution function, over the
+range $2R \le r \le \maxrfit R$, and for packing fractions $\eta \le
+0.45$.  The resulting parameters are displayed in
+Table~\ref{tab:kappa}.  The maximum error in $g(r)$ within this
+range is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
+\rmaxerr R$.  Fig.~\ref{fig:radial-distribution} displays
+our approximation at just under half of the densities that were
+included in the fit.
 
@@
 
-For ease of implementation and future extension to larger radii, we
-chose to express our fit with a fourth-order polynomial.  We constrain
-our functional form for $g_S(r, g_\sigma)$ to match the value
-$g_\sigma$ at contact and to reproduce the $g(r)=1$ result of the
-low-density limit.  Incorporating these constraints we have the
-functional form:
-\begin{align}
-  g_S(r,g_\sigma) &=
-  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
-  \left(\tfrac{r}{\sigma}-1\right)^j.
-  \label{eq:fit-form}
-\end{align}
-where the matrix $\kappa_{ij}$ is determined from a least-squares
-fit to Monte Carlo data for the radial distribution function, over the
-range $2R \le r \le \maxrfit R$, and for packing fractions $\eta \le
-0.45$.  The resulting parameters are displayed in
-Table~\ref{tab:kappa}.  The maximum error in $g_S(r)$ within this
-range is \maxerr, which occurs at $\eta = \etamaxerr$ and $r =
-\rmaxerr R$.  Fig.~\ref{fig:radial-distribution} displays
-our approximation at just under half of the densities that were
-included in the fit.
-
 \newcommand{\plotcomp}[1]{The top halves of
   these figures show the results of Monte Carlo simulations, while the
-  bottom halves show the CVA \fixme{over the range of radii for which
-    the fit was performed}.  On the right are plots of #1 on the
+  bottom halves show the CVA, truncated beyond the range of the fit.
+  On the right are plots of #1 on the
   paths illustrated in the figures to the left.  These plots compare
@@
 \end{figure*}
+
 \section{Results}
@@
     The top halves of these figures show the results of Monte Carlo
-    simulations, while the bottom halves show the CVA \fixme{over the
-      range of radii for which the fit was performed}.  On the right
+    simulations, while the bottom halves show the CVA, truncated
+    beyond the range of the fit.  On the right
     are plots of $g^{(3)}(\rr_1,\rr_2,\rr_3)$ on the paths illustrated
@@
 the two spheres are in contact, and is therefore only plotted on that
-segment of the path.  As an input to both previous approximations and
-to the CVA, we use the hard sphere radial distribution function found
-with Monte Carlo simulation, interpolated as necessary. We find that
-both previous approximations to the pair distribution function predict
-the qualitative angular dependence of the pair distribution function
-at contact better than this work.  However, in each case the pair
-distribution function has a systematic error at contact---either too
-high or too low.  While our approximation is smoother than either of
-the existing approaches, its errors will have a tendency to cancel
-when used in a perturbation expansion.  At higher densities, the
-approximation of Fischer and Methfessel requires evaluating the radial
-distribution function at densities significantly higher than the
-freezing density, which poses numerical difficulties when using the
-radial distribution function from simulation.  When the two points
-$\rr_1$ and $\rr_2$ are both more than a radius away from contact, we
-find that any of these approaches gives a reasonable prediction.
+segment of the path.  As an input to the previous approximations we
+use the hard sphere radial distribution function found with Monte
+Carlo simulation, interpolated as necessary. We find that both
+previous approximations to the pair distribution function predict
+stronger angular dependence of the pair distribution function at
+contact than this work.  The previous approximations each have a
+systematic error at contact---either too high or too low.  In
+contrast, our errors at contact have a tendency to cancel when used in
+a perturbation expansion.  At higher densities, the approximation of
+Fischer and Methfessel requires evaluating the radial distribution
+function at densities significantly higher than the freezing density,
+which poses numerical difficulties when using the radial distribution
+function from simulation.  When the two points $\rr_1$ and $\rr_2$ are
+both more than a radius away from contact, we find that any of these
+approaches gives a reasonable prediction.
 
@@
   \end{subfigure}
-  \caption{Plot of $\frac{dF_1}{dz}$ near a hard wall.  (a) shows a
+  \caption{Plot of $\frac{dF_1}{dz}$ near a hard wall, with arbitrary
+    vertical scale.  (a) shows a
     sticky hard-sphere fluid defined by a pair potential
@@
 in a high-temperature perturbation expansion $F_1$
-for several pair potentials.  In order to focus on effects at the
+for two typical pair potentials.  In order to focus on effects at the
 interface, we have defined a position-dependent pair interaction
@@
 We have introduced and tested the contact value approach for the pair
-distribution function of the inhomogeneous hard-sphere fluid.  The
-separable version of the contact value
-approach is suitable for use in the development of classical
-density functionals constructed using perturbation theory, as it
-may be efficiently computed using exclusively fixed-kernel convolutions.  We have
-tested this function at a hard wall and near a single fixed hard
+distribution function of the inhomogeneous hard-sphere fluid.  The CVA
+is suitable for use in the development of classical density
+functionals constructed using perturbation theory, as it may be
+efficiently computed using exclusively fixed-kernel convolutions.  We
+have tested this function at a hard wall and near a single fixed hard
 sphere, and find that it gives qualitatively reasonable results.
 Tests of the pair distribution function in integrals that arise in
-thermodynamic perturbation theory suggest that the CVA is
-accurate for attractions up to the distance to which the radial
-distribution function is fit, and is an improvement over existing
-approximations near contact.
+thermodynamic perturbation theory suggest that the CVA is accurate for
+attractions up to the distance to which the radial distribution
+function is fit, and is an improvement over existing approximations
+near contact.
 

### 2014-07-16 papers/pair-correlation/figs/plot-path-triplet-contact.py

@@ -278,7 +278,7 @@ for name in styles.plots:
     z_z, g_z = avg_points(z_z, g_z, dpath)
 
   zplot.plot(z_c, g_c, styles.plot[name], label=styles.title[name])
-  if name != 'fisher':
+  if name != 'fischer':
     # Fischer et al only predict pair distribution function in contact
     xplot.plot(x_x, g_x, styles.plot[name], label=styles.title[name])
     zplot.plot(z_z, g_z, styles.plot[name], label=styles.title[name])

### 2014-07-17 papers/pair-correlation/paper.tex
edits from Paho Tested on bennet. Took 0:19.71.

@@
 hard-sphere fluid.  The functional derivative of the free energy with
-respect to hard-sphere diameter in Eq.~\ref{eq:gsigma-exact} requires
+respect to the hard-sphere diameter in Eq.~\ref{eq:gsigma-exact} requires
 that we be able to evaluate the change in free energy resulting from a
@@
 \begin{align}
-  g(|r_{12}|; g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+  g(r_{12}; g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
 \end{align}
@@
 \end{align}
-where the functional is written as a summation of integrals, each of
-which is an integral of a simple convolution in three dimensions.  Thus,
+where the functional is written as a summation of integrals of simple
+convolutions in three dimensions.  Thus,
 each of these integrals may be computed in $\mathcal{O}(N\log N)$
@@
 refine it by performing a separable fit to the radial distribution
-function from Monte Carlo simulation.  We chose to focus our fit on
+function from Monte Carlo simulation.  We focus our fit on
 the range of distances $r_{12} \le \maxrfit R$.  This range is
@@
 For ease of implementation and future extension to larger radii, we
-chose to express our fit with a fourth-order polynomial.  We constrain
+fit the radial distribution function using a fourth-order polynomial.  We constrain
 our functional form such that $g(r; g_\sigma)$ reduces to $g_\sigma$

### 2014-07-23 papers/square-well-liquid/paper.tex
Added Fourier transforms of radial distribution polynomial expansion for square well liquid. Tested on knightley. Took 0:25.41.

@@
 \end{widetext}
+\subsection{Polynomial expansion}
+
+A different way to do this that avoids some of the Fourier transform
+troubles is to formulate $g_S$ as a polynomial expansion.  We do this
+up to fouth order so that
+\begin{align}
+  g_S(r;g_\sigma) &=
+  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
+  \left(\frac{r}{\sigma}-1\right)^j,
+  \label{eq:fit-form}
+\end{align}
+which makes the transforms for $i=1,2,3,4$
+\begin{align}
+ \frac {4\pi}{k}\int_\sigma^{\lambda \sigma} r
+ \left( \frac{r}{\sigma}-1\right)^i \sin (kr) \mathrm{d}r.
+\end{align}
+Let's do a u-substitution where $u= \frac{r}{\sigma}-1$, then the
+integral becomes
+\begin{align}
+ \frac {4\pi}{k}\int_0^{\lambda -1} \sigma^2 (u+1)u^i \sin(k\sigma
+       [u+1]) \mathrm{d}u.
+\end{align}
+Now we calculate the integrals.
+\begin{widetext}
+\begin{align}
+  \int_\sigma^{\lambda \sigma} r
+ \left( \frac{r}{\sigma}-1\right) \sin (kr) \mathrm{d}r =&
+ \frac{1}{k^3 \sigma}\left[(2-\lambda k^2
+ \sigma^2(\lambda-1))\cos(k\lambda\sigma)-k\sigma(\sin(k\sigma)
+ +(1-2\lambda) \sin(k\lambda\sigma)) -\cos(k\sigma) \right] \\
+  \int_\sigma^{\lambda \sigma} r
+ \left( \frac{r}{\sigma}-1\right)^2 \sin (kr) \mathrm{d}r =&
+ \frac{1}{k^4\sigma^2} \big[
+ 6\sin(k\sigma)+(k^2\sigma^2(1-4\lambda+3\lambda^2)
+ -6)\sin(k\lambda\sigma) \notag \\
+ &-2k\sigma\cos(k\sigma) - k\sigma(4 +
+ \lambda(k^2\sigma^2(\lambda-1)^2 - 6))\cos(k\lambda\sigma)\big] \\
+  \int_\sigma^{\lambda \sigma} r
+ \left( \frac{r}{\sigma}-1\right)^3 \sin (kr) \mathrm{d}r =&
+ \frac{1}{k^5\sigma^3} \big[k\sigma(6\sin(k\sigma) + (18-24\lambda
+   +k^2\sigma^2(\lambda-1)^2(4\lambda-1))\sin(k\lambda\sigma)) \notag \\
+   &+24\cos(k\sigma) + (6k^2\sigma^2(\lambda-1)(2\lambda-1)-\lambda
+   k^4\sigma^4(\lambda-1)^3)\cos(k\lambda\sigma)\big] \\
+  \int_\sigma^{\lambda \sigma} r
+ \left( \frac{r}{\sigma}-1\right)^4 \sin (kr) \mathrm{d}r =&
+ \frac{1}{k^6\sigma^4}
+ \big[(k\sigma(\lambda-1)(36-60\lambda+k^2\sigma^2(\lambda-1)^2(5\lambda-1))
+   -1)\sin(k\lambda\sigma) -120\sin(k\sigma) \notag \\ 
+   &+k\sigma(24\cos(k\sigma) -
+   (24(5\lambda-4)-4k^2\sigma^2(\lambda-1)^2 (5\lambda-2) +
+   \lambda k^4\sigma^4(\lambda-1)^4)\cos(k\lambda\sigma) \big]
+\end{align}
+\end{widetext}
 \subsubsection{Averaged pair distribution  $g_\sigma(\rr)$}

### 2014-08-27 papers/fuzzy-fmt/figs/plot-walls.py
Some changes to various plots for soft spheres. Tested on knightley. Took 0:45.54.

@@ -46,7 +46,7 @@ for kT in [0.0, 0.1, 0.01, 0.001, 0.0001]:
     for fname in glob.glob('figs/mcwalls-0.%02d00-%.4f*.dat' % (ff, kT)):
         print 'examining', fname
         d = numpy.loadtxt(fname)
-        d = smooth(d, 1)
+        d = smooth(d, 300)
         pylab.plot(15-abs(d[:,0]), d[:,1]*(4*pi/3), styles.mc[kT], label=fname)
 
 names.append('DFT kT = 0')
@@ -83,8 +83,8 @@ for i in range(len(data)):
 pylab.title('Packing fraction = %f' % (ff/100.0))
 pylab.xlabel('$z/R$')
 pylab.ylabel('Local Filling Fraction')
-pylab.legend(loc = 'best')
-pylab.xlim(-0.1, 5)
+#pylab.legend(loc = 'best')
+pylab.xlim(-0.1, 15)
 #pylab.ylim(.25, .6)
 pylab.savefig('figs/walls-%02d.pdf' % (ff))
 pylab.show()

### 2014-08-29 papers/pair-correlation/paper.tex
edits based on referee Tested on bennet. Took 0:43.15.

@@
 \begin{document}
-\title{An efficient contact value approach to approximating the pair distribution
+\title{An efficient approach to approximating the pair distribution
   function of the inhomogeneous hard-sphere fluid}
@@
 
-At its core, thermodynamic perturbation theory---sometimes referred to
-as the high-temperature expansion---is an expansion of the
-free energy in powers of a small parameter, which is typically the
+At its core, thermodynamic perturbation theory (TPT), sometimes referred to
+as the high-temperature expansion, is an expansion of the
+free energy in powers of a small parameter, which is the
 product of a
@@
   \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
-  \caption{Plot of the hard-sphere radial distribution function. The
+  \caption{Plot of the hard-sphere radial distribution function of the
+    homogeneous fluid at several values for packing fraction $\eta$. The
     blue lines show our separable fit, the black dots show the true

### 2014-08-29 papers/pair-correlation/figs/plot-dadz-sticky.py

@@ -44,7 +44,8 @@ z['sokolowski'], dadz['sokolowski'] = data[:,0], data[:,1]
 data = loadtxt("figs/walls/walls_daWB-fischer-%04.2f-%05.3f.dat" % (eta, delta_r))
 z['fischer'], dadz['fischer'] = data[:,0], data[:,1]
 
-plt.figure(figsize=(6,4))
+scaleme=0.8
+plt.figure(figsize=(6*scaleme,4*scaleme))
 
 for version in versions:
   plt.plot(z[version], dadz[version], styles.plot[version], label=styles.title[version])

### 2014-08-29 papers/pair-correlation/figs/short-range-ghs.py

@@ -438,8 +438,8 @@ def avg_points(x, y, dpath):
       old_i = i
   return (new_x, new_y)
 
-
-fig2 = pylab.figure(2, figsize=(6,8))
+scaleme = 0.8
+fig2 = pylab.figure(2, figsize=(6*scaleme,8*scaleme))
 ax2 = fig2.gca()
 
 # ax.axvline(x=sigma, color='gray', linestyle='-')
@@ -560,7 +560,9 @@ pylab.axes().set_yticks(ticks)
 pylab.axes().set_yticklabels(ticklabels)
 
 for i in indexes:
-  pylab.text((2*xlim[0]+xlim[1])/3, zeros[i]+maxes[i]-0.1, r"$\eta = %.2f$" %(ff[i]), ha='center', va='top')
+  frac = 1.0/3 + 0.03*i - 0.1
+  pylab.text((1-frac)*xlim[0]+frac*xlim[1], zeros[i]+maxes[i]-0.14,
+             r"$\eta = %.1f$" %(ff[i]), ha='center', va='top')
 
 
 handles, labels = ax2.get_legend_handles_labels()

### 2014-08-29 papers/pair-correlation/paper.tex
Changed the end of section IV in pair paper. Tested on knightley. Took 1:24.97.

@@
 \newcommand{\fixme}[1]{\red{[#1]}}
+\newcommand{\newstuff}[1]{\blue{#1}}
 
@@
 
-The CVA requires the radial distribution function expressed as a function of $r$
-and $g_\sigma$.  One approach to constructing $g(r;g_\sigma)$ is to
-compute $n$ from $g_\sigma$ (using e.g. the Carnahan-Starling formula
-for $g_\sigma$)
-and then find $g(r;n)$ using either
-some approximate functional form (e.g. Percus-Yevick) or Monte Carlo
-simulation results.  In this paper we use a fit to Monte Carlo results
-for the radial distribution function $g(r)$ which facilitates
-efficient computation.  When evaluating the approximations of earlier
-works described in Section~\ref{sec:gn}, we use interpolated Monte
-Carlo results for $g(r;n)$.
+The CVA requires the radial distribution function expressed as a
+function of $r$ and $g_\sigma$.  \fixme{One approach to constructing
+  $g(r;g_\sigma)$ is to compute $n$ from $g_\sigma$ (using e.g. the
+  Carnahan-Starling formula for $g_\sigma$) and then find $g(r;n)$
+  using either some approximate functional form (e.g. Percus-Yevick)
+  or Monte Carlo simulation results.}  \newstuff{In this paper we
+  construct a functional form for $g(r, g_\sigma)$ that allows for
+  greater computational efficiency. In section \ref{sec:efficient} we
+  discuss the general form that allows for this efficiency and in
+  section \ref{sec:separable-fit} we detail the fit to Monte Carlo
+  results.}  When evaluating the approximations of earlier works
+described in Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated Monte Carlo
+results for $g(r;n)$.
 
@@
 \section{Making the CVA efficient}
+\label{sec:efficient}
 The existing approaches to approximating the pair distribution
@@
 \begin{align}
-  g(r_{12}; g_\sigma) &= \sum_{i} a_i(r_{12}) b_i(g_\sigma)
+  g(r; g_\sigma) &= \sum_{i} a_i(r) b_i(g_\sigma)
+  \\\text{\newstuff{removed subscript $_{12}$.}}
 \end{align}

### 2014-09-02 papers/pair-correlation/paper.tex
polish off end of Section IV Tested on bennet. Took 0:19.21.

@@
 The CVA requires the radial distribution function expressed as a
-function of $r$ and $g_\sigma$.  \fixme{One approach to constructing
-  $g(r;g_\sigma)$ is to compute $n$ from $g_\sigma$ (using e.g. the
-  Carnahan-Starling formula for $g_\sigma$) and then find $g(r;n)$
-  using either some approximate functional form (e.g. Percus-Yevick)
-  or Monte Carlo simulation results.}  \newstuff{In this paper we
-  construct a functional form for $g(r, g_\sigma)$ that allows for
-  greater computational efficiency. In section \ref{sec:efficient} we
-  discuss the general form that allows for this efficiency and in
-  section \ref{sec:separable-fit} we detail the fit to Monte Carlo
-  results.}  When evaluating the approximations of earlier works
-described in Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated Monte Carlo
-results for $g(r;n)$.
+function of $r$ and $g_\sigma$.
+%
+%% \fixme{One approach to constructing $g(r;g_\sigma)$ is to compute $n$
+%%   from $g_\sigma$ (using e.g. the Carnahan-Starling formula for
+%%   $g_\sigma$) and then find $g(r;n)$ using either some approximate
+%%   functional form (e.g. Percus-Yevick) or Monte Carlo simulation
+%%   results.}
+%
+We construct a functional form for $g(r, g_\sigma)$ that allows for
+improved computational efficiency.  We introduce the general form that
+allows for this efficiency in Section~\ref{sec:efficient}, and we
+detail our specific approximation for $g(r,g_\sigma)$ that uses this
+general form in Section~\ref{sec:separable-fit}.
+%
+%% When evaluating the approximations of earlier works described in
+%% Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated
+%% Monte Carlo results for $g(r;n)$.
 
@@
   g(r; g_\sigma) &= \sum_{i} a_i(r) b_i(g_\sigma)
-  \\\text{\newstuff{removed subscript $_{12}$.}}
 \end{align}

### 2014-09-02 papers/pair-correlation/paper.tex
finish another response to referee Tested on bennet. Took 0:19.68.

@@
 \newcommand{\fixme}[1]{\red{[#1]}}
-\newcommand{\newstuff}[1]{\blue{#1}}
+\newcommand{\newstuff}[1]{{\color{blue} #1}}
 
@@
   \frac{g(r_{12};n(\rr_1))+g(r_{12};n(\rr_2))}{2}
-     & \text{mean function}
+     & \text{mean function}\label{eq:mean-function}
 \end{align}
@@
 %% When evaluating the approximations of earlier works described in
-%% Section~\ref{sec:gn} \newstuff{for comparison}, we use interpolated
+%% Section~\ref{sec:gn} for comparison, we use interpolated
 %% Monte Carlo results for $g(r;n)$.
@@
 convolutions, which may be efficiently computed using Fast Fourier
-Transforms (FFTs) in $\mathcal{O}(N\log N)$ time.
+Transforms (FFTs).
+
+\newcommand\Vcell{V_{\textit{cell}}}
+\newcommand\Vinteraction{V_{\Phi}}
+
+Computation of the free energy correction from Eq.~\ref{eq:mean-field}
+for a periodic system by direct integration requires a nested
+integration over the volume of the system~$\Vcell$, and the volume
+over which the interaction is nonzero~$\Vinteraction$.  Thus the cost
+of computation scales as
+$\mathcal{O}\left(\frac{\Vcell\Vinteraction}{\Delta V^2}\right)$ where
+$\Delta V$ is the volume resolution of the computational grid.  Direct
+integration is the most efficient algorithm when using the existing
+functionals for $g^{(2)}(\rr_1,\rr_2)$ described in
+Section~\ref{sec:gn}.  The one exception is the ``mean-function''
+approximation (Eq.~\ref{eq:mean-function}), which could in principle
+be made more efficient using the same technique we describe here.
+Because the CVA allows the integral in Eq.~\ref{eq:mean-field} to be
+written as a sum of fixed-kernel convolutions, it can be computed
+without a nested integral, at the cost of performing a few FFTs.  This
+approach scales as $\mathcal{O}\left(\frac{\Vcell}{\Delta
+  V}\log\frac{\Vcell}{\Delta V}\right)$, as do most widely used DFT
+functionals such as FMT~\cite{rosenfeld1989free, roth2002whitebear}.
+%
+With this scaling, when examining systems with long interaction
+distances or high resolution---which is often necessary when working
+with hard-sphere functionals---the CVA has the potential to be far
+more efficient than existing methods.
 
@@
 
-We begin by examining the lowest-order correction in TPT, given by
-Eq.~\ref{eq:mean-field}.  The two terms that are averaged in
-Eq.~\ref{eq:g2-our-mean} give equal contributions to the integral
+To see how we obtain this improved scaling, we examine the
+lowest-order correction in TPT, given by Eq.~\ref{eq:mean-field}.  The
+two terms that are averaged in Eq.~\ref{eq:g2-our-mean} give equal
+contributions to the integral
 \begin{align}
@@
 where the functional is written as a summation of integrals of simple
-convolutions in three dimensions.  Thus,
-each of these integrals may be computed in $\mathcal{O}(N\log N)$
-time, which is the same scaling behavior that is common to the most
-common and accurate classical density functional theories, most
-notably Fundamental Measure Theory (FMT)~\cite{rosenfeld1989free,
-  roth2002whitebear}.
+convolutions in three dimensions.  Thus, each of these integrals may
+be computed in $\mathcal{O}(N\log N)$ time, where $N$ is the number of
+grid points in the computational cell.  This is the same scaling as is
+required to compute the fundamental measures such as $n_3$ which are
+used in FMT.
 

### 2014-09-02 papers/pair-correlation/paper.tex
edit conclusion paragraph Tested on bennet. Took 0:19.61.

@@
 
-\section{Accuracy in thermodynamic perturbation theory}
-
 \begin{figure}
@@
 \end{figure}
+
+\section{Accuracy in thermodynamic perturbation theory}
 
@@
 We have introduced and tested the contact value approach for the pair
-distribution function of the inhomogeneous hard-sphere fluid.  The CVA
-is suitable for use in the development of classical density
-functionals constructed using perturbation theory, as it may be
-efficiently computed using exclusively fixed-kernel convolutions.  We
-have tested this function at a hard wall and near a single fixed hard
-sphere, and find that it gives qualitatively reasonable results.
-Tests of the pair distribution function in integrals that arise in
-thermodynamic perturbation theory suggest that the CVA is accurate for
-attractions up to the distance to which the radial distribution
-function is fit, and is an improvement over existing approximations
-near contact.
+distribution function $g^{(2)}(\rr_1,\rr_2)$ of the inhomogeneous
+hard-sphere fluid.  The pair distribution function plays a key role in
+thermodynamic perturbation theory, which is widely used in the
+construction of classical density functionals.  The CVA---unlike
+existing approximations---is suitable for use in classical density
+functionals based on perturbation theory, as it may be efficiently
+computed using exclusively fixed-kernel convolutions.  We have tested
+this function at a hard wall and near a single fixed hard sphere, and
+find that it gives excellent agreement with simulation.  Tests of the
+pair distribution function in integrals that arise in thermodynamic
+perturbation theory suggest that the CVA is accurate for attractions
+up to the distance to which the radial distribution function is fit,
+and is a significant improvement over existing approximations near
+contact.  But most importantly, the computational cost of using the
+CVA in a classical density functional scales much more favorably than
+existing methods in high resolution computations.
 

### 2014-11-04 papers/square-well-liquid/histogram-paper.tex
added round trip rate figure Tested on MAPHost. Took 0:39.62.

@@
 \begin{document}
-\title{Applying the optimized ensemble histogram method to the square-well liquid}
+\title{Applying the optimized ensemble histogram method to the
+  square-well liquid}
 
-\author{Michael A. Perlin}
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\author{Michael A. Perlin} \author{David Roundy}
+\affiliation{Department of Physics, Oregon State University,
+  Corvallis, OR 97331}
 
 \begin{abstract}
-  We have applied the clever histogram method to the square-well liquid.
+  We have applied the clever histogram method to the square-well
+  liquid.
 \end{abstract}
@@
   \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-dos}
-  \caption{A density of states plot demonstrating the difficulty of using
-    canonical Monte Carlo to simulate a system.  A simulation at a
-    give temperature only provides information fo a small number of
+  \caption{A density of states plot demonstrating the difficulty of
+    using canonical Monte Carlo to simulate a system.  A simulation at
+    a give temperature only provides information fo a small number of
     energy states.\label{fig:dos}}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rt}
+  \caption{Round trip counts from different methods.}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rt-rate}
+  \caption{Round trip rates from different methods.}
 \end{figure}
@@
 
+\textcolor{blue}{Michael (14 November 2014): I tried running this
+  method with the end condition that our gaussian has a width 0.5
+  times that of our energy range (maximum entropy to low energy
+  point), and this is precisely what happened: the simulation got
+  stuck pushing down sharp and pointy bubbles. The largest gaussian
+  width ever observed was 0.27 times the energy range ($\lambda=1.3$,
+  $\eta=0.3$, $N=20$). Furthermore, the gaussian width jumped around a
+  good deal during initialization, and I might have to make some
+  figures to see what exactly happens to it during initialization.}
+
 \end{document}

### 2014-11-05 papers/square-well-liquid/histogram-paper.tex

--- 2014-07-06 papers/square-well-liquid/histogram-paper.tex
@@
 
+\newcommand{\red}[1]{{\bf \color{red} #1}}
+\newcommand{\green}[1]{{\bf \color{green} #1}}
+\newcommand{\blue}[1]{{\bf \color{blue} #1}}
+\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}
+
+\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
+\newcommand{\mpsays}[1]{{\color{red} [\blue{Michael:} \emph{#1}]}}
+
 \begin{document}
-\title{Applying the optimized ensemble histogram method to the square-well liquid}
+\title{Applying the optimized ensemble histogram method to the
+  square-well liquid}
 
-\author{Michael A. Perlin}
-\author{David Roundy}
-\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}
+\author{Michael A. Perlin} \author{David Roundy}
+\affiliation{Department of Physics, Oregon State University,
+  Corvallis, OR 97331}
 
 \begin{abstract}
-  We have applied the clever histogram method to the square-well liquid.
+  We have applied the clever histogram method to the square-well
+  liquid.
 \end{abstract}
@@
 
-\textcolor{red}{I think maybe we only want to keep ``pushing down the
+\davidsays{I think maybe we only want to keep ``pushing down the
   bubbles'' so long as the variance increases.  I'm afraid that if the
@@
 
+\mpsays{(14 November 2014): I tried running the bubble pushing method
+  with the end condition that the added log-space weight gaussian has
+  a width 0.5 times that of our energy range (maximum entropy to low
+  energy points), and this is precisely what happened: the simulation
+  got stuck pushing down sharp and pointy bubbles. Other observations:
+  the largest gaussian width ever observed was 0.27 times the energy
+  range ($\lambda=1.3$, $\eta=0.3$, $N=20$). Gaussian width also
+  jumped around a good deal during initialization, and I might have to
+  make some figures to see what exactly happens to it during
+  initialization.}
+
 \end{document}

### 2014-11-17 papers/square-well-liquid/figs/plot-scaling.py
plot round trip rates during simulation instead of initialization Tested on MAPHost. Took 0:45.30.

@@ -19,7 +19,7 @@ ff = float(sys.argv[2])
 versions = eval(sys.argv[3])
 #arg versions = [["wang_landau", "gaussian", "flat", "walkers"]]
 
-# input: ["data/periodic-ww%04.2f-ff%04.2f-N%i-%s-rt.dat" % (ww, ff, N, version) for version in versions for N in [5,6,7,8,9]]
+# input: ["data/periodic-ww%04.2f-ff%04.2f-N%i-%s-lnw.dat" % (ww, ff, N, version) for version in versions for N in [5,6,7,8,9,10,20]]
 
 # input: "data/periodic-ww*-ff*-N*-*-rt.dat"
 
@@ -32,7 +32,7 @@ init_iters = {}
 for version in versions:
     Ns[version] = []
     init_iters[version] = []
-    for filename in glob.glob("data/periodic-ww%04.2f-ff%04.2f-N*-%s-rt.dat" % (ww, ff, version)):
+    for filename in glob.glob("data/periodic-ww%04.2f-ff%04.2f-N*-%s-lnw.dat" % (ww, ff, version)):
         data = numpy.loadtxt(filename)
 
         N = int(N_regex.findall(filename)[0])
@@ -43,15 +43,6 @@ for version in versions:
 
         init_iters[version].append(int(initialization_iters_regex.findall(content)[0]))
 
-        e_hist = numpy.loadtxt(
-            "data/periodic-ww%04.2f-ff%04.2f-N%i-%s-E.dat" % (ww, ff, N, version))
-        lnw = numpy.loadtxt(
-            "data/periodic-ww%04.2f-ff%04.2f-N%i-%s-lnw.dat" % (ww, ff, N, version))
-        energy = -e_hist[:,0]/N
-        log10w = lnw[e_hist[:,0].astype(int),1]*numpy.log10(numpy.exp(1))
-        log10_dos = numpy.log10(e_hist[:,1]) - log10w
-        log10_dos -= log10_dos.max()
-
     # The following sorts the lists according to the number of spheres
     indexes = range(len(Ns[version]))
     indexes.sort(key=Ns[version].__getitem__)
@@ -61,7 +52,7 @@ for version in versions:
     plt.semilogy(Ns[version], init_iters[version], styles.color[version]+'.-', label=version)
 
 plt.xlabel('$N$')
-plt.ylabel('initialization iterations')
[…]

### 2014-11-22 papers/square-well-liquid/histogram-paper.tex
add fake plot for scaling Tested on bennet. Took 0:19.05.
plot round trip rates during simulation instead of initialization Tested on MAPHost. Took 0:45.30.
add table to histogram paper Tested on darcy. Took 27:10.56.
changed a bunch of naming convensions. instead of 'round trips' we have 'samples', as what we were really counting   was the number of times we had independently sampled a given energy   (which is almost, but not quite the same thing). changed 'seeking_energy' to 'energy_observed', which I think is more descriptive. renamed 'flat' to the 'bubble_suppression', which is a mouthful,

@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rt}
-  \caption{Round trip counts from different methods.}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-sample-rate}
+  \caption{Round trip rates from different methods.}
 \end{figure}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-rt-rate}
-  \caption{Round trip rates from different methods.}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-scaling}
+  \caption{Scaling.\label{fig:scaling}}
 \end{figure}
@@
 the $kT=0.1\epsilon$ simulation.
+
+\begin{table*}
+\input{figs/scaling-table-ww130-ff30}
+\caption{Scaling data}
+\end{table*}
 

### 2014-11-22 papers/square-well-liquid/figs/plot-dos.py

@@ -20,7 +20,7 @@ N = int(sys.argv[3])
 #arg N = [10, 20, 100, 200, 1000]
 
 versions = eval(sys.argv[4])
-#arg versions = [["nw", "wang_landau", "robustly-optimistic", "gaussian", "flat", "walkers", "kT2", "kT1"]]
+#arg versions = [["nw", "wang_landau", "robustly_optimistic", "gaussian", "bubble_suppression", "walkers", "kT2", "kT1"]]
 
 # input: ["data/periodic-ww%04.2f-ff%04.2f-N%i-%s-%s.dat" % (ww, ff, N, version, data) for version in versions for data in ["E","lnw"]]
 

### 2014-12-16 src/Monte-Carlo/square-well-monte-carlo.cpp
implemented vanilla wang landau, i.e. wang landau with the parameter values   given in the original WL paper, with one exception: I made the end condition less   stringent, because otherwise initialization took at least an hour on my computer. Tested on quipu. Took 8:07.89.
Tested on quipu. Took 0:30.86.

@@ -89,6 +89,7 @@ int main(int argc, const char *argv[]) {
   int gaussian_fit = false;
   int walker_weights = false;
   int wang_landau = false;
+  int vanilla_wang_landau = false;
 
   // Tuning factors
   double wl_factor = 0.125;
@@ -96,8 +97,17 @@ int main(int argc, const char *argv[]) {
   double wl_threshold = 3;
   double wl_cutoff = 1e-6;
   double gaussian_bubble_cutoff = 0.2;
-  double robust_update_scale = 0.8;
-  double robust_cutoff = 0.2;
+  double robust_update_scale = 0.8; // should be ~1, but must be <= 1
+  double robust_cutoff = 0.2; // should be ~0.1, and must be < 1
+
+  // Do not change these*! They are taken directly from the WL paper.
+  double vanilla_wl_factor = exp(1);
+  double vanilla_wl_fmod = 2;
+  double vanilla_wl_threshold = 0.25;
+  double vanilla_wl_cutoff = 3e-8;
+  // *fixme: the actual cutoff should be 1e-8, but that can make some of the small default
+  //   runs take hours to complete! We should definitely comment on that in the
+  //   histogram paper.
 
   sw_simulation sw;
 
@@ -192,6 +202,8 @@ int main(int argc, const char *argv[]) {
      "Use a walker optimization weight histogram method", "BOOLEAN"},
     {"wang_landau", '\0', POPT_ARG_NONE, &wang_landau, 0,
      "Use Wang-Landau histogram method", "BOOLEAN"},
+    {"vanilla_wang_landau", '\0', POPT_ARG_NONE, &vanilla_wang_landau, 0,
+     "Use Wang-Landau histogram method with vanilla settings", "BOOLEAN"},
     {"wl_factor", '\0', POPT_ARG_DOUBLE | POPT_ARGFLAG_SHOW_DEFAULT, &wl_factor,
      0, "Initial value of Wang-Landau factor", "DOUBLE"},
     {"wl_fmod", '\0', POPT_ARG_DOUBLE | POPT_ARGFLAG_SHOW_DEFAULT, &wl_fmod, 0,
@@ -228,7 +240,8 @@ int main(int argc, const char *argv[]) {
 
   // check that only one method is used
   if(bool(no_weights) + bool(robustly_optimistic) + bool(bubble_suppression)
[…]

### 2014-12-16 src/Monte-Carlo/square-well.cpp

@@ -577,8 +577,8 @@ void sw_simulation::initialize_canonical(double kT) {
 }
 
 // improve the weight array using the Wang-Landau method.
-void sw_simulation::initialize_wang_landau(double wl_factor, double wl_threshold,
-                                           double wl_cutoff) {
+void sw_simulation::initialize_wang_landau(double wl_factor, double wl_fmod,
+                                           double wl_threshold, double wl_cutoff) {
   int weight_updates = 0;
   const int starting_iterations = iteration;
   while (1) {
@@ -598,7 +598,7 @@ void sw_simulation::initialize_wang_landau(double wl_factor, double wl_threshold
     fflush(stdout);
 
     if (variation > 0 && variation < max(wl_threshold, exp(wl_factor))) {
-      wl_factor /= 2;
+      wl_factor /= wl_fmod;
       // for wang-landau, only flush energy histogram; keep weights
       flush_arrays(energy_histogram, ln_energy_weights, energy_levels, true,
                    state_of_max_entropy);

### 2015-01-07 papers/square-well-liquid/histogram-paper.tex
implemented vanilla wang landau, i.e. wang landau with the parameter values   given in the original WL paper, with one exception: I made the end condition less   stringent, because otherwise initialization took at least an hour on my computer. Tested on quipu. Took 8:07.89.
Tested on quipu. Took 0:30.86.

@@
   \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-sample-rate}
-  \caption{Round trip rates from different methods.}
+  \caption{Inverse sampling rates from different methods.}
 \end{figure}
@@
 
+\subsection{Tracking transitions}
+
+We can track the number of attempted transitions from each energy
+state to each other energy state.  This statistic has the advantage of
+being approximately independent of the weighting function used
+(although it is highly dependent on the step size).  Thus we could
+continue to refine this statistic, even as we are updating the weight
+array.
+
+The ratio of density of states at two energies should be proportional
+to the ratio of upgoing and downgoing transition rates between those
+two energies.  Thus these transitions give us information about
+$\Delta D$ (where $D$ is the density of states).
+
+Figure~\ref{fig:transitions} below shows the transition matrix for a
+20-sphere system, computed during initialization using the Wang-Landau
+algorithm.
+
+How can we use this?
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-transitions}
+  \caption{A plot showing the probability of energy transitions (with
+    no weighting) from given energy states.\label{fig:transitions}}
+\end{figure}
+
+\subsubsection{Generating $w$ from the transition matrix}
+
+We should be able to generate a set of weighting functions directly
+from the transition matrix.  We could do this either by optimizing for
+a flat histogram, or by optimizing the round-trip rate.  Both would be
+interesting to implement.
+
+One natural approach would use linear algebra.  If the normalized
+transition matrix is $T$, then the density of states is given by
+\begin{equation}
+  D = TD
+\end{equation}
+By solving this equation for $D$, we could find a good approximation
+for $w$ to obtain a flat histogram.
+
+I expect that we could use similar math to the optimized ensemble
+folks to optimize the round-trip rate.
+
+\subsubsection{Sampling the transition matrix (more) directly}
+
+If our goal (in initialization) is to sample the transition matrix
+(from which we compute the optimal weights), we could use a much more
+direct approach.  When computing a histogram, we are forced to perform
+a complete random walk.  However, the transition matrix can be
+computed at a given energy \emph{without sampling any other energies}!
+
+We can start at a given energy, sample the transition matrix at that
+energy to our satisfaction, i.e. until we have observed transitions in
+both directions a statistically significant number of times.  Then we
+would proceed to the next energy and sample it, and so forth.  There
+is a weakness in that we may not be able to sample all states at a
+given energy without going to other energies.  But we could repeat
+this procedure many times to accomplish this.  In a way, this
+algorithm is reminiscent of Wang-Landau, which ``forces'' the
+simulation to sample a whole slew of different energies, but in our
+case we would be able to put strict bounds on how long it spends
+sampling each energy, and could reasonably keep the code from wasting
+huge amounts of time sampling the high-entropy states.
+
 \end{document}

### 2015-01-29 papers/square-well-liquid/histogram-paper.tex
generate multiple transition matrix figures for histogram-paper Tested on bennet. Took 0:37.40.

@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-wang_landau-transitions}\\
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-tmmc-transitions}\\
+  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-wang_landau-tmmc-compare-transitions}
+  \caption{A plot showing the probability of energy transitions (with
+    no weighting) from given energy states.\label{fig:transitions}}
+\end{figure}
+
 \subsection{Tracking transitions}
@@
 continue to refine this statistic, even as we are updating the weight
-array.
+array.  Here are two papers that seem highly relevant to this
+method~\cite{wang1999transition, wang2002transition}, but I'm not sure
+if they're identical.
 
@@
 How can we use this?
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-transitions}
-  \caption{A plot showing the probability of energy transitions (with
-    no weighting) from given energy states.\label{fig:transitions}}
-\end{figure}
 

### 2015-01-30 papers/square-well-liquid/histogram-paper.tex
various cleanups in TMMC method + some actual text describing it Tested on bennet. Took 12:13.97.

@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-wang_landau-transitions}\\
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-tmmc-transitions}\\
-  \includegraphics[width=\columnwidth]{figs/periodic-ww130-ff30-N20-wang_landau-tmmc-compare-transitions}
+\begin{figure*}
+  \includegraphics[width=0.33\textwidth]{figs/periodic-ww130-ff30-N20-wang_landau-transitions}\hfill%
+\includegraphics[width=0.33\textwidth]{figs/periodic-ww130-ff30-N20-tmmc-transitions}\hfill%
+\includegraphics[width=0.33\textwidth]{figs/periodic-ww130-ff30-N20-wang_landau-tmmc-compare-transitions}
   \caption{A plot showing the probability of energy transitions (with
     no weighting) from given energy states.\label{fig:transitions}}
-\end{figure}
+\end{figure*}
 
@@
 
-\subsubsection{Sampling the transition matrix (more) directly}
+\subsubsection{Initializing using the transition matrix directly}
 
-If our goal (in initialization) is to sample the transition matrix
-(from which we compute the optimal weights), we could use a much more
-direct approach.  When computing a histogram, we are forced to perform
-a complete random walk.  However, the transition matrix can be
-computed at a given energy \emph{without sampling any other energies}!
+An alternative initialization algorithm involves using the transition
+matrix directly during initialization.  This process was proposed by
+Swendsen \emph{et al.} as a modification for their initialization
+procedure~\cite{swendsen1999transition}.  In this case, the transition
+matrix is continually updated, thus modifying the relative weighting
+of different energies as the simulation proceeds.  This strictly
+speaking no longer satisfies detailed balance, but as the simulation
+proceeds this violation will quickly become increasingly small.
 
-We can start at a given energy, sample the transition matrix at that
-energy to our satisfaction, i.e. until we have observed transitions in
-both directions a statistically significant number of times.  Then we
-would proceed to the next energy and sample it, and so forth.  There
-is a weakness in that we may not be able to sample all states at a
-given energy without going to other energies.  But we could repeat
-this procedure many times to accomplish this.  In a way, this
-algorithm is reminiscent of Wang-Landau, which ``forces'' the
-simulation to sample a whole slew of different energies, but in our
-case we would be able to put strict bounds on how long it spends
-sampling each energy, and could reasonably keep the code from wasting
-huge amounts of time sampling the high-entropy states.
+This raises the question of how to use the transition matrix to
+directly determine transition probabilities, since the self-consistent
+eigenvalue solution is unsuitable.  The key is to recognize that we do
+not require weights for each energy but only \emph{differences} of
+weights for on pair of energies at a time.  This is a much smaller
+problem.
+
+Swendsen \emph{et al.} find for a flat distribution, the acceptance
+rate for a given transition is equal to the ratio of that element in
+the transition matrix to the element corresponding to the reverse
+transition~\cite{swendsen1999transition}.  Naturally, for a different
+distribution, we can scale the acceptance ratios according to the
+ratio of occupations desired.  We modify this algorithm to always
+accept transitions to a lower energy, thus employing a broad histogram
+method that does not much sample energies above the maximum-entropy
+energy.
+
+Swendsen only uses this approach after a slightly hokey two-stage
+initialization.  It seems to work pretty well, and asymptotically it
+does have detailed balance.
 

### 2015-02-26 src/Monte-Carlo/square-well-monte-carlo.cpp
rename papers/square-well-liquid to papers/histogram and fix fac up Tested on darcy. Took 15:24.96.

@@ -136,7 +136,7 @@ int main(int argc, const char *argv[]) {
   unsigned long int seed = 0;
 
   char *data_dir = new char[1024];
-  sprintf(data_dir, "papers/square-well-liquid/data");
+  sprintf(data_dir, "papers/histogram/data");
   char *filename = new char[1024];
   sprintf(filename, "default_filename");
   char *filename_suffix = new char[1024];

### 2015-02-26 papers/hard-sphere-free-energy/figs/run-free-energy.py
script reorganization for free-energy Tested on bennet. Took 2:56.75.

@@ -2,15 +2,15 @@
 # this should run and plot the free energy simulation eventually
 
 import numpy as np
+import sys, os, math, matplotlib
+if 'show' not in sys.argv:
+    matplotlib.use('Agg')
 import matplotlib.pyplot as plt
-import sys, os, math
 from subprocess import call
-from ast import literal_eval as make_tuple
 
-FILENAME = "periodic-ww1.00-ff{0}-N10-sf{1:f}"
+# input: "../../free-energy-monte-carlo"
 
-
-def main(args):
+def main():
 	# run iteration
 	# parse meta data
 	# start iteration at "shrunk" filling fraction
@@ -27,16 +27,19 @@ def main(args):
 	success_ratios.append(1) # for now. should be the absolute ratio
 
 	for i in xrange(20):
-		filename = FILENAME.format(ff, sf)
+		filename = "periodic-ww1.00-ff{0}-N10-sf{1:f}".format(ff, sf)
 
-		call(['../free-energy-monte-carlo',
-			'--sf', str(sf),
-			'--ff', str(ff),
-			'--filename', filename,
-			])
+                os.system('../../free-energy-monte-carlo --data_dir data --sf %g --ff %g --filename %s'
+                          % (sf, ff, filename))
+		# call(['../free-energy-monte-carlo',
+		# 	'--sf', str(sf),
+		# 	'--ff', str(ff),
+		# 	'--filename', filename,
+		# 	])
+
 
  		# at some point I need to chdir so that the files aren't down a directory on accident.
-		data = read_data_file_to_dict(os.path.join(os.getcwd(), 'free-energy-data', filename+"-g.dat"))
+		data = read_data_file_to_dict(os.path.join('data', filename+"-g.dat"))
 		next_ff = data['ff_small']
 		total_checks = data['total checks of small cell']
 		valid_checks = data['total valid small checks']
@@ -53,9 +56,10 @@ def main(args):
 	#do plot
 	energy = -np.cumsum(map(math.log, success_ratios))
 	print energy
-	plt.scatter(np.array(ffs), energy)
+	plt.scatter(ffs, energy)
 	plt.ylabel('F/kT')
-	plt.xlabel('ff')
+	plt.xlabel(r'$\eta$')
+        plt.savefig('rename-me-please.pdf')
 	plt.show()
 
 
@@ -67,23 +71,7 @@ def read_data_file_to_dict(filepath):
 		for line in f:
 			if line.startswith('#') and ':' in line:
[…]

### 2015-03-04 papers/square-well-fluid/paper.tex
rename papers/square-well-liquid to papers/histogram and fix fac up Tested on darcy. Took 15:24.96.
Added another Fourier transform, and clarified some math in square well fluid paper. Tested on knightley. Took 0:30.22.

--- 2014-07-23 papers/square-well-liquid/paper.tex
@@
 \newcommand\rr{\mathbf{r}}
+\newcommand\dd{\mathrm{d}}
 \newcommand\fixme[1]{\textcolor{red}{\textbf{[#1]}}}
@@
 \begin{align}
-  g_S(r;g_\sigma) &=
-  g_\sigma + \sum_{i=1}^{4} \sum_{j=1}^{4} \kappa_{ij} (g_\sigma - 1)^i
-  \left(\frac{r}{\sigma}-1\right)^j,
+  g_S(r;g_\sigma) &= g_\sigma + \sum_i \gamma_i(g_\sigma) \xi_i(r) \\
+  &=
+  g_\sigma + \sum_{i=1}^{4} \left(\sum_{j=1}^{4} (g_\sigma
+  - 1)^j \kappa_{ji} \right)
+  \left(\frac{r}{\sigma}-1\right)^i,
   \label{eq:fit-form}
@@
 \begin{align}
- \frac {4\pi}{k}\int_\sigma^{\lambda \sigma} r
- \left( \frac{r}{\sigma}-1\right)^i \sin (kr) \mathrm{d}r.
+ \frac {4\pi}{k}\int r \xi_i(r) \Phi(r) \sin (kr) \dd r =  \frac
+       {-4\pi\epsilon}{k}\int_\sigma^{\lambda \sigma} r \xi_i(r)\sin (kr) \dd r
 \end{align}
-Let's do a u-substitution where $u= \frac{r}{\sigma}-1$, then the
-integral becomes
-\begin{align}
- \frac {4\pi}{k}\int_0^{\lambda -1} \sigma^2 (u+1)u^i \sin(k\sigma
-       [u+1]) \mathrm{d}u.
-\end{align}
-Now we calculate the integrals.
+For the ``zeroth'' term, we treat it like $\xi=1$, so that we're only
+transforming the potential well.
+%% For the other i's, we'll do a
+%% u-substitution where $u= \frac{r}{\sigma}-1$, then the
+%% integral becomes
+%% \begin{align}
+%%  \frac {4\pi}{k}\int_0^{\lambda -1} \sigma^2 (u+1)u^i \sin(k\sigma
+%%        [u+1]) \mathrm{d}u.
+%% \end{align}
+%% Now we calculate the integrals.
 \begin{widetext}
 \begin{align}
+  \int_\sigma^{\lambda \sigma} r \sin (kr) \mathrm{d}r =&
+  \frac{1}{k^2}\left[ \sin (k\lambda\sigma) - k\lambda\sigma
+    \cos (k\lambda\sigma) - \sin (k\sigma) + k\sigma \cos (k-sigma) \right]\\
   \int_\sigma^{\lambda \sigma} r
@@
  \frac{1}{k^6\sigma^4}
- \big[(k\sigma(\lambda-1)(36-60\lambda+k^2\sigma^2(\lambda-1)^2(5\lambda-1))
-   -1)\sin(k\lambda\sigma) -120\sin(k\sigma) \notag \\ 
+ \big[(k^2\sigma^2(\lambda-1)(36-60\lambda+k^2\sigma^2(\lambda - 1)^2(5\lambda-1))
+   + 120)\sin(k\lambda\sigma) - 120\sin(k\sigma) \notag \\
    &+k\sigma(24\cos(k\sigma) -
    (24(5\lambda-4)-4k^2\sigma^2(\lambda-1)^2 (5\lambda-2) +
-   \lambda k^4\sigma^4(\lambda-1)^4)\cos(k\lambda\sigma) \big]
+   \lambda k^4\sigma^4(\lambda-1)^4))\cos(k\lambda\sigma) \big]
 \end{align}
@@
 
+We begin with the definition of $A_1$,
+\begin{align}
+  A_1 = \frac{1}{2}\iint n^{(2)}(\rr_1,\rr_2) \Phi(|\rr_1-\rr_2|)
+  \dd\rr_1 \dd\rr_2
+\end{align}
+
+But we must use the relation
+\begin{align}
+  n^{(2)}(\rr_1,\rr_2) = n(\rr_1)n(\rr_2)g^{(2)}(\rr_1,\rr_2)
+\end{align}
+and equation \ref{eq:g2-our-mean} so that
+\begin{align}
+  A_1 &=\frac{1}{4}\iint n(\rr_1)n(\rr_2)\Big[
+    g_s(r_{12},g_\sigma(\rr_1)) \notag \\
+    &\hspace{2cm} + g_s(r_{12},g_\sigma(\rr_2)) \Big] \Phi(r_{12})
+  \dd\rr_1 \dd\rr_2.
+\end{align}
+$g_s$ is symmetric in the exchange of $\rr_1$ and $\rr_2$, and we'll
+expand $g_s$ into its polynomial form
+\begin{align}
+  A_1 &= \frac{1}{2}\iint n(\rr_1)n(\rr_2)g_s(r_{12},g_\sigma(\rr_1))
+  \Phi(r_{12}) \dd\rr_1 \dd\rr_2 \\
+  &= \frac{1}{2}\iint n(\rr_1)n(\rr_2)\sum_i
+  \gamma_i(g_\sigma(\rr_1))\xi_i(r_{12})\Phi(r_{12}) \dd\rr_1 \dd\rr_2 \\
+  &= \frac{1}{2} \sum_i \iint \dd\rr_1 \dd\rr_2
+  n(\rr_1)n(\rr_2)\gamma_i(g_\sigma(\rr_1)) \xi_i(r_{12})\Phi(r_{12}).
+\end{align}
+Then we can rearrange things a little better
+\begin{align}
+  A_1 &= \frac{1}{2} \sum_i \int \dd\rr_1 n(\rr_1)
+  \gamma_i(g_\sigma(\rr_1)) \int \dd\rr_2 n(\rr_2)
+  \xi_i(r_{12}) \Phi(r_{12})
+\end{align}
 \fixme{We have introduced an approximation to the pair distribution

### 2015-03-12 papers/histogram/figs/plot-uhc.py
added errors in U and C_V Tested on MAPHost. Took 0:55.43.

@@ -8,7 +8,7 @@ import numpy
 import styles
 
 if len(sys.argv) not in [5,6]:
-    print 'useage: %s ww ff N versions show' % sys.argv[0]
+    print 'useage: %s ww ff N methods show' % sys.argv[0]
     exit(1)
 
 ww = float(sys.argv[1])
@@ -21,12 +21,12 @@ ff = float(sys.argv[2])
 N = float(sys.argv[3])
 #arg N = range(5,21)
 
-versions = eval(sys.argv[4])
-#arg versions = [["nw","wang_landau","simple_flat","optimized_ensemble","kT0.4","kT0.5","tmmc","oetmmc"]]
+methods = eval(sys.argv[4])
+#arg methods = [["nw","wang_landau","simple_flat","optimized_ensemble","kT0.4","kT0.5","tmmc","oetmmc"]]
 
-# input: ["data/periodic-ww%04.2f-ff%04.2f-N%i-%s-%s.dat" % (ww, ff, N, version, data) for version in versions for data in ["E","lnw"]]
+# input: ["data/periodic-ww%04.2f-ff%04.2f-N%i-%s-%s.dat" % (ww, ff, N, method, data) for method in methods for data in ["E","lnw"]]
 
-reference_method = "wang_landau"
+reference = "wang_landau"
 
 max_T = 2
 T_bins = 1e3
@@ -44,9 +44,13 @@ plt.title('Specific heat capacity for $\lambda=%g$, $\eta=%g$, and $N=%i$' % (ww
 U = {} # internal energy
 CV = {} # heat capacity
 
-for version in versions:
+# we want to keep our methods distinct from our reference
+if reference in methods:
+    methods.remove(reference)
 
-    with open("data/periodic-ww%04.2f-ff%04.2f-N%i-%s-E.dat" % (ww, ff, N, version)) as file:
+for method in set(methods+[reference]):
+
+    with open("data/periodic-ww%04.2f-ff%04.2f-N%i-%s-E.dat" % (ww, ff, N, method)) as file:
         for line in file:
             if("min_T" in line):
                 this_min_T = float(line.split()[-1])
@@ -56,31 +60,43 @@ for version in versions:
 
     # energy histogram file; indexed by [-energy,counts]
     e_hist = numpy.loadtxt(
[…]

### 2015-03-31 papers/renormalization/interesting-idea.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx}% Include figure files
\usepackage{dcolumn}% Align table columns on decimal point
\usepackage{color}

\newcommand{\fixme}[1]{{\bf\color{red}{[#1]}}}
\newcommand{\rr}{\mathbf{r}}

\begin{document}
\title{Renormalization}

\author{David Roundy}
\affiliation{Department of Physics, Oregon State University, Corvallis, OR 97331}

\begin{abstract}
  Here is an interesting way to look at GRG.
\end{abstract}

\maketitle

Let's talk about how we would solve for the free energy precisely.
\begin{align}
  F(N, V, T) &= -kT\ln Z \\
  Z(N, V, T) &= \sum_{i}^{\text{microstates}} e^{-\beta E_i}
\end{align}
This is our exact free energy, which requires that we specify an
explicit volume $V$.  We typically assume that $N \gg 1$ and $V$ is a
macroscopic volume.  An interesting option is to consider the
possibility where $N$ is of arbitrary size, and $V$ likewise may be
small.  In this case, we absolutely need to ask ourselves what the
boundary conditions will be.  One option is to choose periodic
boundary conditions, and a cubic cell with side length $L$, so that $V
= L^3$.  This enables easy comparison with Monte Carlo simulations,
and we can obtain the thermodynamic limit by taking the limit as
$N\rightarrow \infty$ with the density held fixed.  The appeal of this
approach is that it enables us to define a hierarchy of free energies
$f_i$ that have lengths $L_i=2^i L_0$ which provide a physical meanint
to the intermediate free energies obtained in a GRG computation.

We define the $i^{th}$ free energy:
\begin{align}
  f_i(n) &\equiv -\frac{kT}{2^{3i}L_0^3}\ln \sum_{i}^{L_i} e^{-\beta E_i}
\end{align}
where the summation is over microstates with periodicity $L_i$.

We now ask how $f_i$ relates to $f_{i-1}$.  Mathematically, the
difference between these free energies is
\begin{align}
  f_i(n) - f_{i-1}(n) &\equiv \delta f_i \\
  &= -\frac{kT}{2^{3i}L_0^3}\ln \sum_{i}^{L_i} e^{-\beta E_i}
  +\frac{kT}{2^{3i}L_0^3}\ln \sum_{i}^{L_i} e^{-\beta E_i}
\end{align}

\end{document}

### 2015-04-08 fac/all-papers.py
start work on figure-putting-into renormalization using SW MC data Tested on bennet. Took 2:26.09.

@@ -4,7 +4,7 @@ import os
 
 papers = """
    histogram pair-correlation contact hard-sphere-free-energy
-   fuzzy-fmt water-saft polyhedra electrostatics
+   fuzzy-fmt water-saft polyhedra electrostatics renormalization
 """.split()
 
 for paper in papers:

### 2015-04-17 papers/fuzzy-fmt/figs/run-new-radial.py
extend fuzzy-fmt radial-distribution.py to plot with new data
This causes it to use the new-radial-wca.cpp data and MC data from FCC
starting positions.  Sadly, I accidentally added a bunch of incomplete
data a few commits back.  This commit does not add the correct data
(which is not yet available) but does change the build to use the
incomplete data that was already completed.

Tested on bennet. Took 0:24.98.

@@ -10,8 +10,8 @@ def srun(n_reduced, kT):
     return 'srun --mem=20000 -J name-%06.4f-%04.2f' % (kT, n_reduced)
 
 # always remember to build the executable before running it
-assert not os.system('fac papers/fuzzy-fmt/figs/new-radial-lj.mkdat')
-assert not os.system('fac papers/fuzzy-fmt/figs/new-radial-wca.mkdat')
+assert not os.system('scons papers/fuzzy-fmt/figs/new-radial-lj.mkdat')
+assert not os.system('scons papers/fuzzy-fmt/figs/new-radial-wca.mkdat')
 
 def runme(reduced_density, temperature):
     outfilename = figsdir+'/new-data/radial-lj-%06.4f-%04.2f.out' % (temperature, reduced_density)

### 2015-04-20 papers/histogram/thesis/thesis.tex

\documentclass[11pt]{article}

%%% standard header
\usepackage[margin=1in]{geometry} % one inch margins
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancyplain} % page formatting style
\usepackage{hyperref} % for linking references
\newcommand{\psl}{6pt} % store spacing as command for consistency
\setlength{\parindent}{0cm} % don't indent new paragraphs...
\parskip \psl % ... place a space between paragraphs instead
\usepackage[inline]{enumitem} % include for \setlist{}, use below
\frenchspacing % add a single space after a period
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

%%% physics
\usepackage[boldvectors,braket]{physymb} % physics package
\newcommand{\bk}{\Braket} % shorthand for braket notation

%%% math
\usepackage{amsmath} % math package
\renewcommand{\t}{\text} % for text in math environment
\newcommand{\f}[2]{\dfrac{#1}{#2}} % shorthand for fractions
\newcommand{\p}[1]{\left(#1\right)} % parenthesis
\renewcommand{\sp}[1]{\left[#1\right]} % square parenthesis
\renewcommand{\set}[1]{\left\{#1\right\}} % curly parenthesis

%%% figures
\usepackage{graphicx,float,subcaption} % floats, etc.
\usepackage{footnote} % for footnotes in floats
\usepackage[font=small,labelfont=bf]{caption} % caption text options

%%% bibliography, table of contents
\usepackage[sort&compress,numbers]{natbib} % bibliography options
\bibliographystyle{apsrev4-1}
\usepackage[english]{babel} % allow editing bibliography title
\addto\captionsenglish{ % set bibliography title
  \renewcommand{\contentsname}{Table of Contents}
  \renewcommand\refname{\bf References}}
\usepackage{tocloft}
\renewcommand{\cftsecleader}{\cftdotfill{\cftdotsep}}

%%% algorithm float
\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}
{\hrulefill\begin{enumerate}}
{\end{enumerate}\hrulefill}

\renewcommand{\title}{Optimizing Monte Carlo state space exploration
  of the square-well fluid}
\renewcommand{\author}{Michael A. Perlin}
\newcommand{\institution}{Oregon State University}
\newcommand{\department}{Department of Physics}
\newcommand{\supervisor}{Dr. David Roundy}
\renewcommand{\date}{19 April 2015}

%%% notes, etc.
\usepackage{color}
\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\fixme}[1]{[\red{fixme:} \emph{#1}]}

\fancypagestyle{abstract}{
  \rhead{}
  \lhead{}
}

\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{titlepage}

  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}

  \center

  \textsc{\LARGE \institution}\\[1.5cm]
  \textsc{\Large \department}\\[0.5cm]

  \HRule \\[0.4cm]
  { \huge \bfseries \title}\\[0.4cm]
  \HRule \\[1.5cm]

  \begin{minipage}{0.4\textwidth}
    \begin{flushleft} \large
      \emph{Author:}\\
      \author
    \end{flushleft}
  \end{minipage}
  ~
  \begin{minipage}{0.4\textwidth}
    \begin{flushright} \large
      \emph{Supervisor:} \\
      \supervisor
    \end{flushright}
  \end{minipage}\\[4cm]

  {\large\date}

  \vfill

\end{titlepage}

\thispagestyle{empty}

\newpage

\tableofcontents

\thispagestyle{empty}

\newpage

\listof{figure}{\Large List of Figures}

\listof{algorithm}{\Large List of Algorithms}

\thispagestyle{empty}

\newpage

\setcounter{page}{1}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Abstract}
\label{sec:abstract}

\thispagestyle{abstract}

We wish to identify and develop efficient numerical methods for
determining thermodynamic properties of the square-well fluid, so as
to test square-well density functional theories. The liquid-vapor
phase interface is an interesting regime for testing density
functional theories, but unbiased Monte Carlo simulations are
incapable of sampling the low energy fluid states which dominate the
partition function at low (i.e. liquid-state) temperatures. Several
generic Monte Carlo ``histogram methods'' for collecting statistics on
low energy system states have been developed, but little or no
literature exists on their systematic comparison and their application
to the square-well fluid. The square-well fluid in particular
introduces application challenges not manifest in traditional models
for testing and benchmarking such numerical techniques (e.g. the Ising
model). We propose to implement these methods in Monte Carlo
simulations of the square-well fluid, determine appropriate
performance metrics for their comparison, and identify those methods
which are most efficient and effective at sampling otherwise
improbable system states.

\fixme{Rewrite after the rest of the thesis is completed}

\newpage

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Background}
\label{sec:background}

\subsection{Introduction}
\label{sec:intro}

A major challenge in the field of classical density functional theory
is understanding critical point behavior of fluids, the study of which
via simulations requires exploring the state space boundary between
liquid and gaseous fluid phases. For a generic square-well fluid to be
observed in a liquid state, the fluid must have a relatively localized
distribution of the spheres which make up the fluid. Such a
distribution implies a low energy fluid state; conversely, a low
energy liquid state of a square-well fluid otherwise observable in
both liquid and gaseous phases implies a localized distribution of
spheres. The search for liquid states is thus equivalent to the search
for low energy states. In unbiased Monte Carlo simulations, however,
highly localized distributions of spheres are extremely unlikely to
occur, making such simulations impractical for studying the
square-well fluid.

Several generic computational methods exist for dealing with such a
problem; namely, that interesting regions of a system's state space
cannot be sampled via unbiased Monte Carlo simulations in a reasonable
amount of time. There is no consensus, however, on the relative
efficacy of these methods. Furthermore, some methods, such as the
Wang-Landau method\cite{wang_landau}, require tuning several free
parameters that lack both ``obvious'' values and systematic means for
assigning them. This work will thus attempt to quantify and compare
the effectiveness of some such methods.

\subsection{The square-well fluid}
\label{sec:sw_fluid}

\begin{figure}[!b]
  \centering
  \includegraphics[width=0.45\textwidth]{figs/square-well.pdf}
  \caption[The square-well pair potential]{The square-well pair
    potential can be used to model short-range attractive forces to
    first order. The infinite potential at $r<\sigma$ simply enforces
    the condition that spheres cannot overlap. At $r>\lambda\sigma$,
    the potential is null, and makes no contribution to the fluid's
    internal energy.}
  \label{fig:pair_potential}
\end{figure}

The square-well (SW) fluid is a model used in classical density
functional theories to capture low order effects of short-range
attractive forces, such as the van der Waals force. The fluid is
composed of spheres with diameter $\sigma$ which have a pair potential
\begin{align}
  v_{sw}\p{\vec r}=\left\{
    \begin{array}{ll}
      \infty & \abs{\vec r}<\sigma \\
      -\epsilon & \sigma<\abs{\vec r}<\lambda\sigma \\
      0 & \abs{\vec r}>\lambda\sigma
    \end{array}
  \right., \label{eq:pair_potential}
\end{align}
also shown graphically in Figure \ref{fig:pair_potential}, where the
parameters $\lambda$ and $\epsilon$ are referred to as the well width
and depth, respectively. The first ($\abs{\vec r}<\sigma$) part of
this potential forbids spheres from overlapping, whereas the second
($\sigma<\abs{\vec r}<\lambda\sigma$) associates an energy $-\epsilon$
with each pair of spheres whose centers are within distance of
$\lambda\sigma$ of each other (where typically
$\lambda\in\sp{1.05,3}$). The net potential energy of the square-well
fluid is thus
\begin{align}
  E=\sum_{i<j}v_{sw}\p{\vec r_i-\vec r_j},
  \label{eq:internal_energy}
\end{align}
where $\vec r_i$ is the position of the $i$-th sphere. As the
potential energy $E$ is the primary form of energy concerning us in
this paper, we will refer to $E$ as simply the ``energy'' of the
fluid. An important feature of the square-well fluid is that its
energy is always an integer multiple of the well depth. A homogeneous
square-well fluid is thus uniquely identified by its well width
$\lambda$ and filling fraction $\eta$ (i.e. the proportion of space
filled by spheres; a dimensionless density), as all other properties
can be normalized to the natural energy scale $\epsilon$ and length
scale $\sigma$. In practice, our simulation codes use dimensionless
energies $E/\epsilon$, temperatures $kT/\epsilon$, and distances
$r/\sigma$.

\subsection{Monte Carlo fluid simulations}
\label{sec:MC_simulations}

While model systems are powerful tools for understanding complex
physical systems, they do not themselves exist in the real
world. Consequently, direct experimental tests of model system
theories (e.g. square-well density functional theories) are not
possible. For this reason, model system theories are commonly tested
against Monte Carlo simulations. Proper implementation of Monte Carlo
methods to study completely characterized systems ensures that
statistical results from simulations converge on the exact properties
of the simulated system in the infinite simulation time
limit\fixme{cite?}. Furthermore, uncertainties in quantities computed
via Monte Carlo simulations are typically well-defined, monotonically
decreasing functions of simulation time, allowing one to run
simulations to the desired level of accuracy.

\begin{algorithm}[!b]
  \caption{Unbiased Monte Carlo fluid simulation}
  \label{alg:MC}
  \begin{alg}

  \item Construct an initial ``typical'' fluid configuration.

  \item Attempt to change the position of one sphere (in general, a
    single fluid ``atom'' or ``molecule'') to a random location,
    rejecting the change if it results in a forbidden fluid
    configuration (e.g. two or more spheres overlap) and accepting the
    change otherwise. Attempting to move a single sphere is referred
    to as a move. \label{alg:MC_move}

  \item Repeat step \ref{alg:MC_move} for every other sphere in the
    fluid. Attempting to move every sphere once is referred to as an
    iteration of the simulation.
    \label{alg:MC_iteration}

  \item Repeat step \ref{alg:MC_iteration} indefinitely, or until data
    of sufficient quality has been generated, periodically collecting
    and dumping statistics on fluid states (e.g. energy, pair
    distribution histograms, etc.) to data files.

  \end{alg}
\end{algorithm}

Algorithm \ref{alg:MC} provides a sketch of unbiased Monte Carlo fluid
simulations. Such an algorithm is ``unbiased'' in the sense that it
collects statistics (i.e. data) on all valid system
configurations. Statistics whose collection time scales as $\mathcal
O\p{1}$ with system size can be collected after every move (defined in
the algorithm), while statistics whose collection time scales as
$\mathcal O\p{N}$, where $N$ is the number of simulated elements
(e.g. spheres), can be collected after each iteration. In general,
collection with $\mathcal O\p{\chi}$ time scaling in $N$ should not
occur more often than once every $\chi$ moves \fixme{explain better?},
so as to ensure that scaling up simulations does not cause them to
asymptotically spend all of their time collecting statistics, and no
time actually simulating the fluid. Collected statistics determine
thermodynamic properties of the simulated fluid.

In this work, we are concerned with simulating the homogeneous
square-well fluid. To avoid edge effects resulting from fluid behavior
near a wall, we enforce periodic boundary conditions. These boundary
conditions suppress all density fluctuations on scales larger than the
dimensions of the simulated fluid cell, thereby introducing a source
of error; addressing this error, however, is outside the scope of this
work. \fixme{Anything else to add?}

\subsection{Histogram methods}
\label{sec:histogram_methods}

Due to the fact that unbiased Monte Carlo simulations sample all of
state space randomly and without preference, a histogram $H\p{X}$ of
observations of some system property $X$ is directly proportional to
the density of states $D\p{X}$ in $X$ \fixme{explain?}. It is
sometimes the case, however, that the density of states in some region
$R$ in the range of possible $X$ is so low that it is practically
impossible to sufficiently sample $R$, that is, accumulate a
statistically significant $H\p{R}$, in any reasonable amount of time.

\subsubsection{Biased sampling}
\label{sec:biased_sampling}

Histogram methods provide a means to address unbiased Monte Carlo's
inability to sufficiently sample regions of parameter space with a low
density of states by introducing a bias into the otherwise random
sampling of state space. A weighting function $w$ is introduced, whose
domain is the value of some property $X\p{s}$ which depends on the
system state $s$. An additional condition is then added to step
\ref{alg:MC_move} of Algorithm \ref{alg:MC} in order to accept an
attempted move: the weights $w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of
the initial (pre-move) state $s_i$ and final (post-move) state $s_f$
are used to determine the probability $P\p{s_i\to s_f}$ of accepting
an otherwise valid move via
\begin{align}
  P\p{s_i\to s_f}=\min\set{\f{w\sp{X\p{s_f}}}{w\sp{X\p{s_i}}},1}.
  \label{eq:move_prob}
\end{align}
This formula means that when $w\sp{X\p{s_f}}>w\sp{X\p{s_i}}$, the move
$s_i\to s_f$ is accepted; when $w\sp{X\p{s_f}}<w\sp{X\p{s_i}}$, the
ratio of these weights determines the probability of accepting the
move. Due to the fact that only ratios of weights determine
$P\p{s_i\to s_f}$, the weights $w\sp{X\p{s}}$ are scale-invariant,
meaning that their effect on simulations is unchanged by scale factors
that are constant with respect $X\p{s}$.

\begin{algorithm}
  \caption{Biased Monte Carlo fluid simulation}
  \label{alg:biased_MC}
  \begin{alg}

  \item Construct an appropriate weight function $w\sp{X\p{s}}$, whose
    argument is a system property $X\p{s}$ determined by the
    microstate $s$.

  \item Construct an initial ``typical'' fluid configuration.

  \item Attempt to change the position of one sphere, rejecting the
    change if
    \begin{enumerate*}[label=\roman*)]
    \item it results in a forbidden fluid configuration, or
    \item a newly chosen random number on the interval $\sp{0,1}$ is
      larger than the probability given by (\ref{eq:move_prob}).
    \end{enumerate*}
    \label{alg:biased_MC_move}

  \item Repeat step \ref{alg:biased_MC_move} for every other sphere in
    the fluid.
    \label{alg:biased_MC_iteration}

  \item Repeat step \ref{alg:biased_MC_iteration} until data of
    sufficient quality has been generated.

  \end{alg}
\end{algorithm}

Employing biased Monte Carlo simulations, sketched out in Algorithm
\ref{alg:biased_MC}, allows one to construct weight functions that
favor some region of state space over others, as transitions to states
with higher weights are always accepted, whereas transitions to states
with lower weights may be rejected, artificially preventing the
simulated system from leaving interesting regions of state
space. Crucially, the bias introduced by weights can be reversed when
computing system properties from sampling statistics, as the weight of
a particular state is directly proportional to the probability bias of
that state; that is, a state with a weight of 2 will be sampled twice
as often as it would have been with a weight of 1 (\fixme{include a
  subsection on detailed balance? cite?}). If we wish to convert a
histogram $H\p{X}$ of observations in a biased Monte Carlo simulation
into a numerical unnormalized density of states $\tilde D\p{X}$ in
$X$, we therefore point-wise divide the histogram by the corresponding
weights $w\p{X}$, i.e.
\begin{align}
  \tilde D\p{X}=\f{H\p{X}}{w\p{X}}.
  \label{eq:dos_unnormed}
\end{align}
The normalized density of states $D\p{X}$ in $X$ is then
\begin{align}
  D\p{X}=\f{\tilde D\p{X}}{\sum_X\tilde D\p{X}}
  =\f{H\p{X}/w\p{X}}{\sum_XH\p{X}/w\p{X}}.
  \label{eq:dos}
\end{align}
In general, the domain and shape of the weight function will depend on
the desired yields (e.g.  density of states, heat capacity) of a
simulation. A ``histogram method'' is an algorithm for determining a
weight function appropriate for a particular simulation.

In this paper, we will consider weights $w\p{E}$ which depend only on
the energy $E$ of the system. Due to the fact that the square-well
fluid can only have discrete energies $E=-n\epsilon$, where $n$ is a
non-negative integer and $\epsilon$ is the well depth, in simulation
we store the weight function as an array of values. We may therefore
use the terms ``weight function,'' and ``weight array,'' and
``weights'' interchangeably.

\subsubsection{Broad energy sampling}
\label{sec:broad_energy_sampling}

The sort of Monte Carlo simulations described in this paper are only
capable of collecting statistics on system microstates, and functions
of microstates. For example, in simulation one might periodically
compute both the energy $E\p{s}$ and some system property $X\p{s}$,
both of which are determined by the system microstate $s$, in order to
find the mean value of $X$ at any given energy $E$, i.e. $X\p{E}$. In
the real world, however, information about a system's microstate and
functions of microstates (e.g. the energy $E$) is generally is
inaccessible. One therefore cannot measure $X\p{E}$ directly. Instead,
one typically measures the dependence of thermodynamic properties on
macroscopic state variables such as temperature, i.e. $X\p{T}$.

The temperature dependence $X\p{T}$ can be expressed in terms of the
density of states $D\p{E}$ and energy dependence $X\p{E}$ by
\begin{align}
  X\p{T}=\f1{Z\p{T}}\sum_EX\p{E}D\p{E}e^{-E/kT},
  \label{eq:temp_dep_X_norm}
\end{align}
where the partition function $Z\p{T}$ is
\begin{align}
  Z\p{T}=\sum_ED\p{E}e^{-E/kT}.
\end{align}
To reduce redundant computations and numerical error in
implementation, we will use an unnormalized partition function
\begin{align}
  \tilde Z\p{T}=\sum_E\tilde D\p{E}e^{-E/kT},
\end{align}
in terms of which
\begin{align}
  X\p{T}=\f1{\tilde Z\p{T}}\sum_EX\p{E}\tilde D\p{E}e^{-E/kT}.
  \label{eq:temp_dep_X}
\end{align}
Sufficient accumulation of statistics on $X\p{E}$ at all available
energies thus in principle allows one to determine $X\p{T}$ for any
$T$. In practice, the density of states can fall off so quickly with
energy that some range of allowable energies is practically
inaccessible via Monte Carlo simulations, biased or otherwise. In such
a case, computing $X\p{T}$ to a reasonable degree of accuracy requires
sufficiently sampling the energies at which $\tilde D\p{E}e^{-E/kT}$
dominates the sum in (\ref{eq:temp_dep_X_norm}).

We will employ histogram methods in order to sample as broad of an
energy range as possible in each simulation. Broad energy sampling
will allow us to determine, for various square-well fluids,
\begin{enumerate*}[label=\roman*)]
\item the density of states (over energy), and
\item the temperature dependence of various thermodynamic properties,
  particularly at temperatures near the liquid-gas phase boundary.
\end{enumerate*}

% \subsubsection{Detailed balance}
% \label{sec:detailed_balance}

% \fixme{In this section I will explain why and how weights work as
% advertised. Essentially, we will consider the effect of weights on a
% two-state system (or on the transitions between two particular
% states in an otherwise multi-state system), and convince ourselves
% that the biasing effect of weights is the identical in an ensemble
% of transitions between any number of states.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methods}
\label{sec:methods}

\fixme{Give an overview of the methods section}

\subsection{Canonical (fixed temperature) weights}
\label{sec:canonical_weights}

A standard construction of the weight array $w\p{E}$ for what is
called a canonical Monte Carlo simulation involves \fixme{better word}
choosing a particular temperature $T_0$, and using weights
proportional to the Boltzmann distribution at that temperature, i.e.
\begin{align}
  w\p{E}=e^{-E/kT_0},
\end{align}
where there is no reason to normalize $w\p{E}$ due to the fact that
only ratios of weights, as per (\ref{eq:move_prob}), are ever used in
simulation. The fact that only ratios of weights are used in
simulation makes $w\p{E}$ scale-invariant.

The partition function at $T=T_0$ for simulations with canonical
weights is
\begin{align}
  \tilde Z\p{T_0}=\sum_E\tilde D\p{E}e^{-E/kT}
  =\sum_EH\p{E}/w\p{E}~e^{-E/kT} =\sum_EH\p{E},
  \label{Z_canonical}
\end{align}
and the value of a thermodynamic property $X\p{T_0}$
\begin{align}
  X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_EX\p{E}\tilde D\p{E}e^{-E/kT_0}
  =\f{\sum_EX\p{E}H\p{E}}{\sum_EH\p{E}}.
  \label{eq:X_canonical}
\end{align}
The simplifications in in (\ref{Z_canonical}) and
(\ref{eq:X_canonical}), which have no explicit dependence on $T_0$,
occur because canonical Monte Carlo simulations sample energies in
proportion to the distribution (over energy) of microstates at a
temperature of $T_0$.  Canonical Monte Carlo simulations can therefore
fail to sufficiently sample energies which are important
(i.e. energies with a non-negligible state probability density) at
different temperatures. As a consequence, such simulations should not
be used to determine properties $X\p{T\ne T_0}$, and are thus referred
to as ``fixed temperature'' simulations.

Though canonical Monte Carlo is simple to implement, its inability to
investigate a system at more than one temperature at a time is a
disadvantage for determining the temperature dependence of system
properties. In order to find the behavior of $X\p{T}$, one must run
many simulations at discrete temperature intervals; each such
simulation will yield one sample (e.g. one data point) of $X\p{T}$.

Sampling low energies, however, or understanding system behavior at
low temperatures, is even more problematic with canonical
weights. Using low temperature canonical weights will indeed force a
simulated system down to low energies, but will also likely freeze the
system into a local minimum of its energy landscape. Freezing into a
state means that a simulation will sample only a small portion of the
energy landscape, even though there may (and generally will) be many
other states with the same energy.

Due to these problems \fixme{be more explicit?}, we will not use
canonical weights alone to study the square-well fluid. We will,
however, use canonical weights for part of all weight arrays, as
discussed in Section \ref{sec:min_energy}.

\subsection{The multi-canonical (flat histogram) method}
\label{sec:flat_historam}

The multi-canonical histogram method assumes complete knowledge of the
density of states $D\p E$ of the system in question, and solves
(\ref{eq:dos_unnormed}) for a weight array $w\p{E}$ which should yield
a flat energy histogram $H\p E=H_0$ to get
\begin{align}
  w\p E=\f1{D\p E},
  \label{eq:flat_weights}
\end{align}
where we may neglect the constant scale factor $H_0$ and do not worry
about normalization factor distinguishing $D\p{E}$ from $\tilde
D\p{E}$ due to the scale-invariance of $w\p{E}$.

For all but the most trivial or well-studied systems, however, the
density of states $D\p{E}$ is not known prior to simulation, and is in
fact one of the system properties which a simulation is intended to
determine. Nonetheless, this toy method has pedagogical value by
enabling discussion of ``optimal'' weights for generating a flat
histogram. All algorithms to construct $w\p{E}$ with the aim of
producing a flat histogram should therefore converge on the same
weight array, given by (\ref{eq:flat_weights}).

\fixme{Talk about a simple implementation of a flat histogram method
  (essentially a single iteration of Simple Flat), and why one might
  want to use it.}

\subsection{The simple flat method}
\label{sec:simple_flat}

\fixme{Explain, provide algorithm. This method is our own.}

\begin{algorithm}[H]
  \caption{Simple flat initialization}
  \label{alg:simple_flat}
\end{algorithm}

\subsection{The Wang-Landau (WL) method}
\label{sec:wang_landau}

\fixme{Explain the motivation behind the Wang-Landau method, and
  provide the algorithm. Talk about the free parameters and roughly
  what they control.}

\begin{algorithm}[H]
  \caption{Wang-Landau calculation of weights}
  \label{alg:wang_landau}
\end{algorithm}

% \subsubsection{Modifications}
% \label{sec:wang_landau_mod}

% \fixme{Talk about our modifications to Wang-Landau made to adapt it
% to the square-well fluid, as well as (if I have time) other research
% groups' modifications.}

\subsection{The optimized ensemble (OE) method}
\label{sec:optimized_ensemble}

\fixme{Explain, provide algorithm.}

\begin{algorithm}[H]
  \caption{Finding weights by optimizing energy walker currents}
  \label{alg:optimized_ensemble}
\end{algorithm}

\subsection{The transition matrix Monte Carlo (TMMC) method}
\label{sec:tmmc}

\fixme{Explain, provide algorithm. This method is sort of our own, but
  based on other work.}

\begin{algorithm}[H]
  \caption{Transition matrix Monte Carlo simulation}
  \label{alg:tmmc}
\end{algorithm}

\subsection{The hybrid OE-TMMC method}
\label{sec:oetmmc}

\fixme{Explain, provide algorithm. This method is sort of our own, but
  based on other work.}

\begin{algorithm}[H]
  \caption{Hybrid OE-TMMC simulation}
  \label{alg:oetmmc}
\end{algorithm}

\subsection{Identifying the energy range of interest}
\label{sec:energy_range}

Histogram methods typically rely on knowledge of the minimum and
maximum energies of the system in question \fixme{how?}, which are
unknown for the square-well fluid. Worse still, some square-well fluid
energies have such a low density of states that one cannot reasonably
expect to ever observe them via Monte Carlo, biased or otherwise. As
such, we need to identify an energy range of interest, bounded by
minimum and maximum ``important'' energies. \fixme{Rewrite, expand}

\subsubsection{State of maximal entropy}
\label{sec:max_entropy}

\fixme{The energy at which the density of states is maximal, i.e. the
  state of maximal entropy, sets an upper bound for the energy range
  in which appropriate weights are not obvious. Explain why.}

\subsubsection{Minimum important energy}
\label{sec:min_energy}

\fixme{Determining the minimum important energy is considerably
  trickier. Punchline: we have to choose a minimum temperature of
  interest, $T_{\t{min}}$, and the minimum important energy is the
  energy at which the slope of the density of states is
  $\exp\p{kT_{\t{min}}/\epsilon}/\epsilon$. This condition for the
  minimum important energy is directly related to canonical weights at
  a temperature of $T_{\t{min}}$.}

\subsubsection{Computing the density of states}
\label{sec:dos}

\fixme{This section might just become a part of the section on the
  minimum important energy. (\ref{eq:density_of_states}) gives us an
  explicit formula for the density of states if we are confident in
  our energy histogram. We can also compute the density of states from
  the transition matrix, which is sometimes more appropriate to do.}

% \subsection{End conditions}
% \label{sec:end_conditions}

% \fixme{Talk about how we want method-independent end conditions, if
% possible. Introduce the concept of statistically independent energy
% samples, and explain two measures (one optimistic, and one
% pessimistic) for determining the number of statistically independent
% times we have samples a given energy.}

% \fixme{Talk about two end conditions: a fixed number of times which
% we require to independently sample the minimum important energy, and
% an enforced fractional error in the number of times we have
% independently sampled any energy (at a given minimum temperature).}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Results and Discussion}
\label{sec:results}

\subsection{Implementation and overview}
\label{sec:implementation}

\fixme{Talk about the relative difficulty of implementing each method,
  as well as the relative difficulty of optimizing or tuning each
  method. Free parameters are generally bad, and especially so if they
  come without ``obvious'' or prescribed values. Justify our
  modifications to Wang-Landau by showing that we did not make it
  worse. Include mention of how memory requirements for transition
  matrix Monte Carlo scale poorly with system size, as well as energy
  resolution for systems with continuous energies.}

\subsection{Simulation results and errors}
\label{sec:results_and_errors}

\fixme{Talk about the energy histogram, weights, and density of states
  yielded by the methods. These should all be fairly similar. Talk
  about why this is so.}

\begin{figure}[H]
  \centering
  \caption[Energy histograms]{Energy histograms}
  \label{fig:histograms}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Energy weights]{Energy weights}
  \label{fig:weights}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Normalized densities of states]{Normalized densities of
    states}
  \label{fig:density_of_states}
\end{figure}

\fixme{Talk about initialization times, and how they scale with system
  size}

\begin{figure}[H]
  \centering
  \caption[Initialization iterations vs. system size]{Initialization
    iterations vs. system size}
  \label{fig:scaling}
\end{figure}

\fixme{Discuss energy initialization times, sampling rates, and errors
  in computed thermodynamic properties. Explain that errors are given
  relative to a simulation which ran much longer, and talk about how
  we ran that simulation (which method did we use?). Show that all
  methods converge on the same answer as they simulate for longer.}

\begin{figure}[H]
  \centering
  \caption[Optimistic energy sampling rates]{Optimistic energy
    sampling rates}
  \label{fig:opt_sample_rate}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Pessimistic energy sampling rates]{Pessimistic energy
    sampling rates}
  \label{fig:pes_sample_rate}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Initialization iterations vs. data quality]{Initialization
    iterations vs. data quality}
  \label{fig:quality}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Specific internal energy]{Specific internal energy}
  \label{fig:internal_energy}
\end{figure}

\begin{figure}[H]
  \centering
  \caption[Specific heat capacity]{Specific heat capacity}
  \label{fig:heat_capacity}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusions}
\label{sec:conclusions}

\fixme{What have we learned? What recommendations can we make? When is
  it appropriate to use which histogram methods?}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\nocite{*} \bibliography{thesis}

\end{document}

### 2015-04-23 papers/histogram/thesis/thesis.tex
some updates to histogram thesis

@@
 \newcommand{\supervisor}{Dr. David Roundy}
-\renewcommand{\date}{19 April 2015}
+\renewcommand{\date}{21 April 2015}
 
@@
 We wish to identify and develop efficient numerical methods for
-determining thermodynamic properties of the square-well fluid, so as
+determining thermodynamic properties of the square-well fluid in order
 to test square-well density functional theories. The liquid-vapor
@@
 generic Monte Carlo ``histogram methods'' for collecting statistics on
-low energy system states have been developed, but little or no
-literature exists on their systematic comparison and their application
-to the square-well fluid. The square-well fluid in particular
-introduces application challenges not manifest in traditional models
-for testing and benchmarking such numerical techniques (e.g. the Ising
-model). We propose to implement these methods in Monte Carlo
-simulations of the square-well fluid, determine appropriate
-performance metrics for their comparison, and identify those methods
-which are most efficient and effective at sampling otherwise
-improbable system states.
+low energy system states have been developed \fixme{by whom?}, but
+little or no literature exists on their systematic comparison and
+their application to the square-well fluid. The square-well fluid in
+particular introduces application challenges not manifest in
+traditional models for testing and benchmarking such numerical
+techniques (e.g. the Ising model). We propose to implement these
+methods in Monte Carlo simulations of the square-well fluid, determine
+appropriate performance metrics for their comparison, and identify
+those methods which are most efficient and effective at sampling
+otherwise improbable system states.
 
@@
 
-A major challenge in the field of classical density functional theory
-is understanding critical point behavior of fluids, the study of which
-via simulations requires exploring the state space boundary between
-liquid and gaseous fluid phases. For a generic square-well fluid to be
-observed in a liquid state, the fluid must have a relatively localized
-distribution of the spheres which make up the fluid. Such a
-distribution implies a low energy fluid state; conversely, a low
-energy liquid state of a square-well fluid otherwise observable in
-both liquid and gaseous phases implies a localized distribution of
-spheres. The search for liquid states is thus equivalent to the search
-for low energy states. In unbiased Monte Carlo simulations, however,
-highly localized distributions of spheres are extremely unlikely to
-occur, making such simulations impractical for studying the
-square-well fluid.
+\fixme{Intro is confusing. Move these paragraphs in between sections
+  on SW fluid and biased sampling?}
+
+Understanding the behavior of fluids near the critical point, or the
+point at which the distinction between liquid and gas ceases to become
+well-defined, is a major challenge in the field of classical density
+functional theory. Studying the critical point of a fluid via
+simulations requires exploring the state space \fixme{remove words
+  ``state space''?}  boundary between its liquid and gaseous
+phases. For a generic square-well fluid to be observed in a liquid
+state, the fluid must have a relatively localized distribution of the
+spheres which make up the fluid. Such a distribution implies a low
+energy fluid state; conversely, a low energy liquid state of a
+square-well fluid otherwise observable in both liquid and gaseous
+phases implies a localized distribution of spheres. The search for
+liquid states is thus equivalent to the search for low energy
+states. In unbiased Monte Carlo simulations, however, highly localized
+distributions of spheres are extremely unlikely to occur, making such
+simulations impractical for studying the square-well fluid.
 
@@
   \caption[The square-well pair potential]{The square-well pair
-    potential can be used to model short-range attractive forces to
-    first order. The infinite potential at $r<\sigma$ simply enforces
-    the condition that spheres cannot overlap. At $r>\lambda\sigma$,
-    the potential is null, and makes no contribution to the fluid's
+    potential can be used to model short range forces to first
+    order. The infinite potential at $r<\sigma$ simply enforces the
+    condition that spheres cannot overlap. At $r>\lambda\sigma$, the
+    potential is zero, and makes no contribution to the fluid's
     internal energy.}
@@
 
-The square-well (SW) fluid is a model used in classical density
+The square-well (SW) fluid is a simple model used in classical density
 functional theories to capture low order effects of short-range
@@
 physical systems, they do not themselves exist in the real
-world. Consequently, direct experimental tests of model system
-theories (e.g. square-well density functional theories) are not
+world. Consequently, direct experimental tests of theories for model
+system (e.g. square-well density functional theories) are not
 possible. For this reason, model system theories are commonly tested
@@
 of the simulated system in the infinite simulation time
-limit\fixme{cite?}. Furthermore, uncertainties in quantities computed
-via Monte Carlo simulations are typically well-defined, monotonically
-decreasing functions of simulation time, allowing one to run
-simulations to the desired level of accuracy.
+limit\fixme{explain. cite?}. Furthermore, uncertainties in quantities
+computed via Monte Carlo simulations are typically well-defined,
+monotonically decreasing functions of simulation time, allowing one to
+run simulations to the desired level of accuracy.
 
@@
 simulations. Such an algorithm is ``unbiased'' in the sense that it
-collects statistics (i.e. data) on all valid system
-configurations. Statistics whose collection time scales as $\mathcal
-O\p{1}$ with system size can be collected after every move (defined in
-the algorithm), while statistics whose collection time scales as
-$\mathcal O\p{N}$, where $N$ is the number of simulated elements
-(e.g. spheres), can be collected after each iteration. In general,
-collection with $\mathcal O\p{\chi}$ time scaling in $N$ should not
-occur more often than once every $\chi$ moves \fixme{explain better?},
-so as to ensure that scaling up simulations does not cause them to
-asymptotically spend all of their time collecting statistics, and no
-time actually simulating the fluid. Collected statistics determine
-thermodynamic properties of the simulated fluid.
+collects statistics (i.e. data) on all valid system configurations
+with equal probability. Statistics whose collection time scales as
+$\mathcal O\p{1}$ with system size can be collected after every move
+(defined in the algorithm), while statistics whose collection time
+scales as $\mathcal O\p{N}$, where $N$ is the number of simulated
+elements (e.g. spheres), can be collected after each iteration. In
+general, collection with $\mathcal O\p{\chi}$ time scaling in $N$
+should not occur more often than once every $\chi$ moves
+\fixme{explain better?}, so as to ensure that scaling up simulations
+does not cause them to asymptotically spend all of their time
+collecting statistics, and no time actually simulating the
+fluid. Collected statistics are used to find thermodynamic properties
+of the simulated fluid.
 
-In this work, we are concerned with simulating the homogeneous
+In this work, we are concerned with simulating the {\emph homogeneous}
 square-well fluid. To avoid edge effects resulting from fluid behavior
-near a wall, we enforce periodic boundary conditions. These boundary
-conditions suppress all density fluctuations on scales larger than the
-dimensions of the simulated fluid cell, thereby introducing a source
-of error; addressing this error, however, is outside the scope of this
-work. \fixme{Anything else to add?}
+near a wall, we employ periodic boundary conditions. The use of a
+finite cell with periodic boundary conditions suppresses all density
+fluctuations on scales larger than the dimensions of the simulated
+fluid cell, thereby introducing a source of error; addressing this
+error, however, is outside the scope of this work. \fixme{Add
+  feedback}
 
@@
 state space randomly and without preference, a histogram $H\p{X}$ of
-observations of some system property $X$ is directly proportional to
-the density of states $D\p{X}$ in $X$ \fixme{explain?}. It is
-sometimes the case, however, that the density of states in some region
-$R$ in the range of possible $X$ is so low that it is practically
-impossible to sufficiently sample $R$, that is, accumulate a
-statistically significant $H\p{R}$, in any reasonable amount of time.
+observations of some system property $X$ (e.g. energy) \fixme{other
+  examples} is directly proportional to the density of states $D\p{X}$
+in $X$ \fixme{explain?}. It is sometimes the case, however, that the
+density of states in some region $R$ in the range of possible $X$ is
+so low that it is practically impossible to sufficiently sample $R$,
+that is, accumulate a statistically significant $H\p{R}$, in any
+reasonable amount of time.
 
@@
 meaning that their effect on simulations is unchanged by scale factors
-that are constant with respect $X\p{s}$.
+that are constant with respect to $X\p{s}$.
 
@@
 that state; that is, a state with a weight of 2 will be sampled twice
-as often as it would have been with a weight of 1 (\fixme{include a
-  subsection on detailed balance? cite?}). If we wish to convert a
-histogram $H\p{X}$ of observations in a biased Monte Carlo simulation
-into a numerical unnormalized density of states $\tilde D\p{X}$ in
-$X$, we therefore point-wise divide the histogram by the corresponding
+as often as it would have been with a weight of 1. If we wish to
+convert a histogram $H\p{X}$ of observations in a biased Monte Carlo
+simulation into a numerical unnormalized density of states $\tilde
+D\p{X}$ in $X$, we therefore divide the histogram by the corresponding
 weights $w\p{X}$, i.e.
@@
 
+\fixme{Put canonical before this?}
+
 The sort of Monte Carlo simulations described in this paper are only
@@
 both of which are determined by the system microstate $s$, in order to
-find the mean value of $X$ at any given energy $E$, i.e. $X\p{E}$. In
-the real world, however, information about a system's microstate and
-functions of microstates (e.g. the energy $E$) is generally is
-inaccessible. One therefore cannot measure $X\p{E}$ directly. Instead,
-one typically measures the dependence of thermodynamic properties on
-macroscopic state variables such as temperature, i.e. $X\p{T}$.
+find the mean value of $X$ at any given energy $E$,
+i.e. $\bk{X}_E$. In the real world, however, information about a
+system's microstate and functions of microstates (e.g. the energy $E$)
+is generally is inaccessible. One therefore cannot measure $\bk{X}_E$
+directly. Instead, one typically measures the dependence of
+thermodynamic properties on macroscopic state variables such as
+temperature, i.e. $\bk{X}_T$.
 
-The temperature dependence $X\p{T}$ can be expressed in terms of the
-density of states $D\p{E}$ and energy dependence $X\p{E}$ by
+The temperature dependence of $\bk{X}_T$ can be expressed in terms of
+the density of states $D\p{E}$ and the expectation value $\bk{X}_E$ by
 \begin{align}
-  X\p{T}=\f1{Z\p{T}}\sum_EX\p{E}D\p{E}e^{-E/kT},
+  \bk{X}_T=\f1{Z\p{T}}\sum_E\bk{X}_ED\p{E}e^{-E/kT},
   \label{eq:temp_dep_X_norm}
 \end{align}
-where the partition function $Z\p{T}$ is
+where the partition function $Z\p{T}$ is simply a normalization
+factor, given by
 \begin{align}
@@
 To reduce redundant computations and numerical error in
-implementation, we will use an unnormalized partition function
+implementation, we will use a partition function with an unnormalized
+density of states,
 \begin{align}
@@
 \begin{align}
-  X\p{T}=\f1{\tilde Z\p{T}}\sum_EX\p{E}\tilde D\p{E}e^{-E/kT}.
+  \bk{X}_T=\f1{\tilde Z\p{T}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT}.
   \label{eq:temp_dep_X}
 \end{align}
-Sufficient accumulation of statistics on $X\p{E}$ at all available
-energies thus in principle allows one to determine $X\p{T}$ for any
+Sufficient accumulation of statistics on $\bk{X}_E$ at all available
+energies thus in principle allows one to determine $\bk{X}_T$ for any
 $T$. In practice, the density of states can fall off so quickly with
@@
 inaccessible via Monte Carlo simulations, biased or otherwise. In such
-a case, computing $X\p{T}$ to a reasonable degree of accuracy requires
-sufficiently sampling the energies at which $\tilde D\p{E}e^{-E/kT}$
-dominates the sum in (\ref{eq:temp_dep_X_norm}).
+a case, computing $\bk{X}_T$ to a reasonable degree of accuracy
+requires sufficiently sampling the energies at which $\tilde
+D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:temp_dep_X}).
 
@@
 \end{enumerate*}
-
-% \subsubsection{Detailed balance}
-% \label{sec:detailed_balance}
-
-% \fixme{In this section I will explain why and how weights work as
-% advertised. Essentially, we will consider the effect of weights on a
-% two-state system (or on the transitions between two particular
-% states in an otherwise multi-state system), and convince ourselves
-% that the biasing effect of weights is the identical in an ensemble
-% of transitions between any number of states.}
 
@@
 choosing a particular temperature $T_0$, and using weights
-proportional to the Boltzmann distribution at that temperature, i.e.
+proportional to the Boltzmann factor at that temperature, i.e.
 \begin{align}
@@
 \begin{align}
-  X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_EX\p{E}\tilde D\p{E}e^{-E/kT_0}
-  =\f{\sum_EX\p{E}H\p{E}}{\sum_EH\p{E}}.
+  X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT_0}
+  =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}}.
   \label{eq:X_canonical}
@@
 disadvantage for determining the temperature dependence of system
-properties. In order to find the behavior of $X\p{T}$, one must run
+properties. In order to find the behavior of $\bk{X}_T$, one must run
 many simulations at discrete temperature intervals; each such
-simulation will yield one sample (e.g. one data point) of $X\p{T}$.
+simulation will yield one sample (e.g. one data point) of $\bk{X}_T$.
 
@@
 determine. Nonetheless, this toy method has pedagogical value by
-enabling discussion of ``optimal'' weights for generating a flat
+enabling discussion of ``optimal'' weights generate a flat
 histogram. All algorithms to construct $w\p{E}$ with the aim of
@@
 \end{algorithm}
-
+uu
 \subsection{The transition matrix Monte Carlo (TMMC) method}
@@
 unknown for the square-well fluid. Worse still, some square-well fluid
-energies have such a low density of states that one cannot reasonably
-expect to ever observe them via Monte Carlo, biased or otherwise. As
-such, we need to identify an energy range of interest, bounded by
-minimum and maximum ``important'' energies. \fixme{Rewrite, expand}
+energies have such an incredibly low density of states that one cannot
+reasonably expect to ever observe them via Monte Carlo, biased or
+otherwise. As such, we need to identify an energy range of interest,
+bounded by minimum and maximum ``important'' energies. \fixme{Rewrite,
+  expand}
 

### 2015-04-24 papers/histogram/thesis/thesis.tex
updates to thesis

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Background}
-\label{sec:background}
-
-\subsection{Introduction}
+\section{Introduction}
 \label{sec:intro}
 
-\fixme{Intro is confusing. Move these paragraphs in between sections
-  on SW fluid and biased sampling?}
+\fixme{talk more about phase transitions and the critical point?}
 
@@
 functional theory. Studying the critical point of a fluid via
-simulations requires exploring the state space \fixme{remove words
-  ``state space''?}  boundary between its liquid and gaseous
-phases. For a generic square-well fluid to be observed in a liquid
-state, the fluid must have a relatively localized distribution of the
-spheres which make up the fluid. Such a distribution implies a low
-energy fluid state; conversely, a low energy liquid state of a
-square-well fluid otherwise observable in both liquid and gaseous
-phases implies a localized distribution of spheres. The search for
-liquid states is thus equivalent to the search for low energy
-states. In unbiased Monte Carlo simulations, however, highly localized
-distributions of spheres are extremely unlikely to occur, making such
-simulations impractical for studying the square-well fluid.
+simulations requires exploring the boundary between its liquid and
+gaseous phases. The square-well fluid (Section \ref{sec:sw_fluid}) is
+the simplest system with a liquid-vapor phase transition, and is
+therefore of interest for studying the critical point of fluids. For a
+generic square-well fluid to be observed in a liquid state, it must
+have a localized distribution of the spheres which make up the
+fluid. Such a distribution implies a low energy fluid state, as all
+interactions between spheres are attractive. Conversely, a low energy
+liquid state of a square-well fluid otherwise observable in both
+liquid and gaseous phases implies a localized distribution of
+spheres. The search for liquid states of the square-well fluid is thus
+equivalent to the search for low energy states.
 
-Several generic computational methods exist for dealing with such a
-problem; namely, that interesting regions of a system's state space
-cannot be sampled via unbiased Monte Carlo simulations in a reasonable
-amount of time. There is no consensus, however, on the relative
-efficacy of these methods. Furthermore, some methods, such as the
-Wang-Landau method\cite{wang_landau}, require tuning several free
-parameters that lack both ``obvious'' values and systematic means for
-assigning them. This work will thus attempt to quantify and compare
-the effectiveness of some such methods.
+Monte Carlo simulations (Section \ref{sec:mc_sim}) are a standard
+method for studying the thermodynamic properties of a system. In the
+most straightforward implementation of Monte Carlo simulations,
+however, highly localized distributions of spheres are extremely
+unlikely to occur, making such simulations impractical for studying
+the critical point of the square-well fluid. Several generic
+computational methods, called histogram methods (Section
+\ref{sec:histogram_methods}), exist for dealing with such a problem;
+namely, that interesting regions of a system's state space cannot be
+sampled via standard Monte Carlo simulations in a reasonable amount of
+time. There is no consensus, however, on the relative efficacy of
+these methods. Furthermore, some methods, such as the Wang-Landau
+method\cite{wang_landau}, require tuning several free parameters that
+lack both ``obvious'' values and systematic means for assigning
+them. This work will thus attempt to quantify and compare the
+effectiveness of a few such methods.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Background}
+\label{sec:background}
 
@@
 
-\begin{figure}[!b]
+\begin{figure}[tb]
   \centering
@@
 with each pair of spheres whose centers are within distance of
-$\lambda\sigma$ of each other (where typically
-$\lambda\in\sp{1.05,3}$). The net potential energy of the square-well
-fluid is thus
+$\lambda\sigma$ of each other (where typically, $\lambda\in(1,3]$).
+The net potential energy of the square-well fluid is thus
 \begin{align}
@@
 \subsection{Monte Carlo fluid simulations}
-\label{sec:MC_simulations}
+\label{sec:mc_sim}
 
@@
 of the simulated system in the infinite simulation time
-limit\fixme{explain. cite?}. Furthermore, uncertainties in quantities
-computed via Monte Carlo simulations are typically well-defined,
-monotonically decreasing functions of simulation time, allowing one to
-run simulations to the desired level of accuracy.
+limit. Furthermore, uncertainties in quantities computed via Monte
+Carlo simulations are typically well-defined, monotonically decreasing
+functions of simulation time, allowing one to run simulations to the
+desired level of accuracy.
 
-\begin{algorithm}[!b]
+\subsubsection{Implementation}
+\label{sec:mc_implementation}
+
+\begin{algorithm}[tb]
   \caption{Unbiased Monte Carlo fluid simulation}
@@
     change otherwise. Attempting to move a single sphere is referred
-    to as a move. \label{alg:MC_move}
+    to as a move. \label{alg:mc_move}
 
-  \item Repeat step \ref{alg:MC_move} for every other sphere in the
+  \item Repeat step \ref{alg:mc_move} for every other sphere in the
     fluid. Attempting to move every sphere once is referred to as an
     iteration of the simulation.
-    \label{alg:MC_iteration}
+    \label{alg:mc_iteration}
 
-  \item Repeat step \ref{alg:MC_iteration} indefinitely, or until data
+  \item Repeat step \ref{alg:mc_iteration} indefinitely, or until data
     of sufficient quality has been generated, periodically collecting
@@
 with equal probability. Statistics whose collection time scales as
-$\mathcal O\p{1}$ with system size can be collected after every move
-(defined in the algorithm), while statistics whose collection time
-scales as $\mathcal O\p{N}$, where $N$ is the number of simulated
-elements (e.g. spheres), can be collected after each iteration. In
-general, collection with $\mathcal O\p{\chi}$ time scaling in $N$
-should not occur more often than once every $\chi$ moves
-\fixme{explain better?}, so as to ensure that scaling up simulations
-does not cause them to asymptotically spend all of their time
-collecting statistics, and no time actually simulating the
-fluid. Collected statistics are used to find thermodynamic properties
-of the simulated fluid.
+$\mathcal O\p{1}$ with system size, meaning that increasing the number
+of spheres $N$ in the simulation does not affect collection time, can
+be collected after every move (defined in the algorithm). Statistics
+whose collection time scales as $\mathcal O\p{N}$, meaning that
+doubling $N$ doubles collection time, can be collected after each
+iteration. In general, collection with $\mathcal O\sp{\chi\p{N}}$ time
+scaling should not occur more often than once every $\chi\p{N}$ moves
+(where $\chi\p{N}$ may be an arbitrary function of only $N$ void of
+constant prefactors, e.g. $N\log N$, or $2^N$). These collection rules
+ensure that scaling up simulations does not cause them to
+asymptotically spend all of their time collecting statistics, and no
+time actually simulating the fluid. Collected statistics are used to
+find thermodynamic properties of the simulated fluid.
 
-In this work, we are concerned with simulating the {\emph homogeneous}
+In this work, we are concerned with simulating the \emph{homogeneous}
 square-well fluid. To avoid edge effects resulting from fluid behavior
@@
 fluctuations on scales larger than the dimensions of the simulated
-fluid cell, thereby introducing a source of error; addressing this
-error, however, is outside the scope of this work. \fixme{Add
-  feedback}
+fluid cell, thereby introducing a source of error. Addressing and
+sequestering this error, however, is outside the scope of this work,
+and involves considering the limit of numerical results as the number
+of spheres $N\to\infty$ (with a constant filling fraction).
+
+\subsubsection{Computing observables}
+\label{sec:computing_observables}
+
+The sort of Monte Carlo simulations described above are only capable
+of collecting statistics on functions of system microstates, and
+finding correlations between these functions. For example, in
+simulation one might periodically compute both the energy $E\p{s}$ and
+some system property $X\p{s}$, both of which are determined by the
+microstate $s$, in order to find the mean value of $X$ at any given
+energy $E$, i.e. $\bk{X}_E$. In the real world, however, information
+about a system's microstates, and by extension functions of
+microstates (e.g. the energy $E$), is inaccessible. One therefore
+cannot measure $\bk{X}_E$ directly. Instead, one typically measures
+the dependence of thermodynamic properties on macroscopic state
+variables such as temperature, i.e. $\bk{X}_T$.
+
+To find $\bk{X}_T$, we first need to understand the concept of a
+density of states. Given an arbitrary system property $Y\p{s}$
+(e.g. energy) determined by the microstate $s$, one can define a
+density of states $D\p{Y}$ such that
+\begin{enumerate*}[label=\roman*)]
+\item $D\p{Y_0}$ is proportional to the number of microstates $s$ for
+  which $Y\p{s}=Y_0$, and
+\item $\sum_YD\p{Y}=1$.
+\end{enumerate*}
+The temperature dependence of $\bk{X}_T$ can be expressed in terms of
+the expectation value $\bk{X}_E$ and the density of states $D\p{E}$ by
+\begin{align}
+  \bk{X}_T=\f1{Z\p{T}}\sum_E\bk{X}_ED\p{E}e^{-E/kT},
+  \label{eq:XT_norm}
+\end{align}
+where the partition function $Z\p{T}$ is simply a normalization
+factor, given by
+\begin{align}
+  Z\p{T}=\sum_ED\p{E}e^{-E/kT}.
+\end{align}
+To reduce redundant computations and numerical error in
+implementation, we will use a partition function with an unnormalized
+density of states,
+\begin{align}
+  \tilde Z\p{T}=\sum_E\tilde D\p{E}e^{-E/kT},
+\end{align}
+in terms of which
+\begin{align}
+  \bk{X}_T=\f1{\tilde Z\p{T}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT}.
+  \label{eq:XT}
+\end{align}
 
@@
 state space randomly and without preference, a histogram $H\p{X}$ of
-observations of some system property $X$ (e.g. energy) \fixme{other
-  examples} is directly proportional to the density of states $D\p{X}$
-in $X$ \fixme{explain?}. It is sometimes the case, however, that the
-density of states in some region $R$ in the range of possible $X$ is
-so low that it is practically impossible to sufficiently sample $R$,
-that is, accumulate a statistically significant $H\p{R}$, in any
-reasonable amount of time.
+observations of some system property $X$ is directly proportional to
+the density of states $D\p{X}$; that is, the number of observations
+$H\p{X_0}$ of the system with $X\p{s}=X_0$ is proportional to the
+total number of states $s$ for which $X\p{s}=X_0$. It is sometimes the
+case, however, that the density of states in some region $R$ in the
+range of possible $X$ is so low that it is practically impossible (via
+unbiased Monte Carlo) to sufficiently sample $R$, that is, accumulate
+a statistically significant histogram $H\p{R}$, in any reasonable
+amount of time.
 
@@
 system state $s$. An additional condition is then added to step
-\ref{alg:MC_move} of Algorithm \ref{alg:MC} in order to accept an
+\ref{alg:mc_move} of Algorithm \ref{alg:MC} in order to accept an
 attempted move: the weights $w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of
@@
 
-\begin{algorithm}
+\begin{algorithm}[tb]
   \caption{Biased Monte Carlo fluid simulation}
@@
     \end{enumerate*}
-    \label{alg:biased_MC_move}
+    \label{alg:biased_mc_move}
 
-  \item Repeat step \ref{alg:biased_MC_move} for every other sphere in
+  \item Repeat step \ref{alg:biased_mc_move} for every other sphere in
     the fluid.
-    \label{alg:biased_MC_iteration}
+    \label{alg:biased_mc_iteration}
 
-  \item Repeat step \ref{alg:biased_MC_iteration} until data of
+  \item Repeat step \ref{alg:biased_mc_iteration} until data of
     sufficient quality has been generated.
@@
   \tilde D\p{X}=\f{H\p{X}}{w\p{X}}.
-  \label{eq:dos_unnormed}
+  \label{eq:dos}
 \end{align}
@@
   =\f{H\p{X}/w\p{X}}{\sum_XH\p{X}/w\p{X}}.
-  \label{eq:dos}
+  \label{eq:dos_norm}
 \end{align}
@@
 
-\subsubsection{Broad energy sampling}
-\label{sec:broad_energy_sampling}
-
-\fixme{Put canonical before this?}
-
-The sort of Monte Carlo simulations described in this paper are only
-capable of collecting statistics on system microstates, and functions
-of microstates. For example, in simulation one might periodically
-compute both the energy $E\p{s}$ and some system property $X\p{s}$,
-both of which are determined by the system microstate $s$, in order to
-find the mean value of $X$ at any given energy $E$,
-i.e. $\bk{X}_E$. In the real world, however, information about a
-system's microstate and functions of microstates (e.g. the energy $E$)
-is generally is inaccessible. One therefore cannot measure $\bk{X}_E$
-directly. Instead, one typically measures the dependence of
-thermodynamic properties on macroscopic state variables such as
-temperature, i.e. $\bk{X}_T$.
-
-The temperature dependence of $\bk{X}_T$ can be expressed in terms of
-the density of states $D\p{E}$ and the expectation value $\bk{X}_E$ by
-\begin{align}
-  \bk{X}_T=\f1{Z\p{T}}\sum_E\bk{X}_ED\p{E}e^{-E/kT},
-  \label{eq:temp_dep_X_norm}
-\end{align}
-where the partition function $Z\p{T}$ is simply a normalization
-factor, given by
-\begin{align}
-  Z\p{T}=\sum_ED\p{E}e^{-E/kT}.
-\end{align}
-To reduce redundant computations and numerical error in
-implementation, we will use a partition function with an unnormalized
-density of states,
-\begin{align}
-  \tilde Z\p{T}=\sum_E\tilde D\p{E}e^{-E/kT},
-\end{align}
-in terms of which
-\begin{align}
-  \bk{X}_T=\f1{\tilde Z\p{T}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT}.
-  \label{eq:temp_dep_X}
-\end{align}
-Sufficient accumulation of statistics on $\bk{X}_E$ at all available
-energies thus in principle allows one to determine $\bk{X}_T$ for any
-$T$. In practice, the density of states can fall off so quickly with
-energy that some range of allowable energies is practically
-inaccessible via Monte Carlo simulations, biased or otherwise. In such
-a case, computing $\bk{X}_T$ to a reasonable degree of accuracy
-requires sufficiently sampling the energies at which $\tilde
-D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:temp_dep_X}).
-
-We will employ histogram methods in order to sample as broad of an
-energy range as possible in each simulation. Broad energy sampling
-will allow us to determine, for various square-well fluids,
-\begin{enumerate*}[label=\roman*)]
-\item the density of states (over energy), and
-\item the temperature dependence of various thermodynamic properties,
-  particularly at temperatures near the liquid-gas phase boundary.
-\end{enumerate*}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Methods}
-\label{sec:methods}
-
-\fixme{Give an overview of the methods section}
-
-\subsection{Canonical (fixed temperature) weights}
+\subsubsection{Canonical weights}
 \label{sec:canonical_weights}
 
-A standard construction of the weight array $w\p{E}$ for what is
-called a canonical Monte Carlo simulation involves \fixme{better word}
-choosing a particular temperature $T_0$, and using weights
-proportional to the Boltzmann factor at that temperature, i.e.
+The most common weight array $w\p{E}$ used by physicists for what is
+called a canonical Monte Carlo simulation involves choosing a
+particular temperature $T_0$, and using weights proportional to the
+Boltzmann factor at that temperature, i.e.
 \begin{align}
@@
 \begin{align}
-  \tilde Z\p{T_0}=\sum_E\tilde D\p{E}e^{-E/kT}
-  =\sum_EH\p{E}/w\p{E}~e^{-E/kT} =\sum_EH\p{E},
+  \tilde Z\p{T_0}=\sum_E\tilde D\p{E}e^{-E/kT_0}
+  =\sum_E\f{H\p{E}}{w\p{E}}~e^{-E/kT_0}
+  =\sum_E\f{H\p{E}}{e^{-E/kT_0}}~e^{-E/kT_0}=\sum_EH\p{E},
   \label{Z_canonical}
@@
 
-Due to these problems \fixme{be more explicit?}, we will not use
-canonical weights alone to study the square-well fluid. We will,
-however, use canonical weights for part of all weight arrays, as
-discussed in Section \ref{sec:min_energy}.
+Due to these problems, we will not use canonical weights alone to
+study the square-well fluid. We will, however, use canonical weights
+for part of all weight arrays, as discussed in Section
+\ref{sec:min_energy}.
+
+\subsubsection{Broad energy sampling}
+\label{sec:broad_energy_sampling}
+
+Given the expression for $\bk{X}_T$ in (\ref{eq:XT}), sufficient
+accumulation of statistics on $\bk{X}_E$ at all available energies in
+principle allows one to determine $\bk{X}_T$ for any temperature
+$T$. In practice, the density of states can fall off so quickly with
+energy that some range of allowable energies is practically
+inaccessible via Monte Carlo simulations, biased or otherwise. In such
+a case, computing $\bk{X}_T$ to a reasonable degree of accuracy
+requires sufficiently sampling the energies at which $\tilde
+D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:XT}).
+
+We will employ histogram methods in order to sample as broad of an
+energy range as possible in each simulation. Broad energy sampling
+will allow us to determine, for various square-well fluids,
+\begin{enumerate*}[label=\roman*)]
+\item the density of states $D\p{E}$, and
+\item the temperature dependence $\bk{X}_T$ of various thermodynamic
+  properties $X$, particularly at temperatures near the liquid-gas
+  phase boundary.
+\end{enumerate*}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Methods}
+\label{sec:methods}
+
+\fixme{Give an overview of the methods section}
 
@@
 density of states $D\p E$ of the system in question, and solves
-(\ref{eq:dos_unnormed}) for a weight array $w\p{E}$ which should yield
-a flat energy histogram $H\p E=H_0$ to get
+(\ref{eq:dos}) for a weight array $w\p{E}$ which should yield a flat
+energy histogram $H\p E=H_0$ to get
 \begin{align}

### 2015-04-26 papers/histogram/thesis/thesis.tex

@@
   \centering
-  \includegraphics[width=0.45\textwidth]{figs/square-well.pdf}
+  \includegraphics[width=0.47\textwidth]{figs/square-well.pdf}
   \caption[The square-well pair potential]{The square-well pair
@@
     condition that spheres cannot overlap. At $r>\lambda\sigma$, the
-    potential is zero, and makes no contribution to the fluid's
+    potential is null, and makes no contribution to the fluid's
     internal energy.}
@@
 unbiased Monte Carlo) to sufficiently sample $R$, that is, accumulate
-a statistically significant histogram $H\p{R}$, in any reasonable
+a statistically significant histogram $H\p{X\in R}$, in any reasonable
 amount of time.
@@
 requires sufficiently sampling the energies at which $\tilde
-D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:XT}).
+D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:XT}). Section
+\ref{sec:energy_range} provides more discussion on this subject,
+namely on identifying the energies which are important for computing
+system properties at a given temperature.
 
@@
 
-\subsection{The multi-canonical (flat histogram) method}
+\subsection{The flat histogram (multi-canonical) method}
 \label{sec:flat_historam}
 
-The multi-canonical histogram method assumes complete knowledge of the
-density of states $D\p E$ of the system in question, and solves
-(\ref{eq:dos}) for a weight array $w\p{E}$ which should yield a flat
-energy histogram $H\p E=H_0$ to get
+The flat histogram method, also called the multi-canonical method,
+assumes complete knowledge of the density of states $D\p{E}$ of the
+system in question, and solves (\ref{eq:dos}) for a weight array
+$w\p{E}$ which should yield a flat energy histogram $H\p E=H_0$ to get
 \begin{align}
@@
 fact one of the system properties which a simulation is intended to
-determine. Nonetheless, this toy method has pedagogical value by
-enabling discussion of ``optimal'' weights generate a flat
-histogram. All algorithms to construct $w\p{E}$ with the aim of
-producing a flat histogram should therefore converge on the same
-weight array, given by (\ref{eq:flat_weights}).
+determine. A naive way to implement the flat histogram method without
+knowing a priori the density of states $D\p{E}$ involves first running
+a simulation without weights (or with constant weights $w\p{E}=w_0$)
+while collecting a histogram $H\p{E}$ of the observed energies
+$E$. After some time (loosely defined), the histogram $H\p{E}$ may be
+taken as an approximation of the density of states $\tilde D\p{E}$
+(see Section \ref{sec:histogram_methods}), and used via
+(\ref{eq:flat_weights}) to generate a weight array for an actual
+simulation.
 
-\fixme{Talk about a simple implementation of a flat histogram method
-  (essentially a single iteration of Simple Flat), and why one might
-  want to use it.}
+The problem with this implementation is that histogram methods are
+generally necessary when some range of energies $R$ is inaccessible
+via unbiased Monte Carlo simulations. In such a case, the energy
+histogram $H\p{E\in R}$ after simulating for any reasonable amount of
+time will be statistically insignificant or null, invalidating the
+approximation $H\p{E}\approx\tilde D\p{E}$. This method will therefore
+only ``flatten'' the histogram at energies $E$ which have been
+sufficiently sampled for $H\p{E}$ to be statistically significant. The
+end goal in implementing histogram methods, however, is precisely to
+sample those energies that cannot be sufficiently sampled without
+employing clever schemes to bias Monte Carlo simulations. The flat
+histogram method therefore does not directly help us study the
+square-well fluid. Nonetheless, the this method has pedagogical value
+by enabling discussion of an ``optimal'' weight array for generating a
+flat energy histogram in simulations. All algorithms to construct
+$w\p{E}$ with the aim of producing a flat energy histogram should
+converge on the same weight array, given by (\ref{eq:flat_weights}).
 

### 2015-04-26 papers/histogram/thesis/thesis.tex
updates to histogram thesis

@@
 
+\listof{algorithm}{\Large List of Algorithms}
+
 \listof{figure}{\Large List of Figures}
-
-\listof{algorithm}{\Large List of Algorithms}
 
@@
 phase interface is an interesting regime for testing density
-functional theories, but unbiased Monte Carlo simulations are
-incapable of sampling the low energy fluid states which dominate the
-partition function at low (i.e. liquid-state) temperatures. Several
-generic Monte Carlo ``histogram methods'' for collecting statistics on
-low energy system states have been developed \fixme{by whom?}, but
-little or no literature exists on their systematic comparison and
-their application to the square-well fluid. The square-well fluid in
-particular introduces application challenges not manifest in
-traditional models for testing and benchmarking such numerical
-techniques (e.g. the Ising model). We propose to implement these
-methods in Monte Carlo simulations of the square-well fluid, determine
-appropriate performance metrics for their comparison, and identify
-those methods which are most efficient and effective at sampling
-otherwise improbable system states.
+functional theories, but unbiased (i.e. Metropolis) Monte Carlo
+simulations are incapable of sampling the low energy fluid states
+which dominate the partition function at low (i.e. liquid-state)
+temperatures. Several generic Monte Carlo ``histogram methods'' for
+collecting statistics on low energy system states have been developed
+\fixme{by whom?}, but little or no literature exists on their
+systematic comparison and their application to the square-well
+fluid. The square-well fluid in particular introduces application
+challenges not manifest in traditional models for testing and
+benchmarking such numerical techniques (e.g. the Ising model). We
+propose to implement these methods in Monte Carlo simulations of the
+square-well fluid, determine appropriate performance metrics for their
+comparison, and identify those methods which are most efficient and
+effective at sampling otherwise improbable system states.
 
@@
 
-Monte Carlo simulations (Section \ref{sec:mc_sim}) are a standard
-method for studying the thermodynamic properties of a system. In the
-most straightforward implementation of Monte Carlo simulations,
-however, highly localized distributions of spheres are extremely
-unlikely to occur, making such simulations impractical for studying
-the critical point of the square-well fluid. Several generic
-computational methods, called histogram methods (Section
-\ref{sec:histogram_methods}), exist for dealing with such a problem;
-namely, that interesting regions of a system's state space cannot be
-sampled via standard Monte Carlo simulations in a reasonable amount of
-time. There is no consensus, however, on the relative efficacy of
-these methods. Furthermore, some methods, such as the Wang-Landau
-method\cite{wang_landau}, require tuning several free parameters that
-lack both ``obvious'' values and systematic means for assigning
-them. This work will thus attempt to quantify and compare the
-effectiveness of a few such methods.
+Monte Carlo simulations are a standard means for studying the
+thermodynamic properties of a system. In the most straightforward
+implementation of Monte Carlo fluid simulations (Section
+\ref{sec:mc_sim}), however, highly localized distributions of fluid
+elements are extremely unlikely to occur, making such simulations
+impractical for studying the critical point of the square-well
+fluid. Several generic computational methods, called histogram methods
+(Section \ref{sec:histogram_methods}), exist for dealing with such a
+problem; namely, that interesting regions of a system's state space
+cannot be sampled via standard Monte Carlo methods in a reasonable
+amount of time. There is no consensus, however, on the relative
+efficacy of these methods. Furthermore, some methods, such as the
+Wang-Landau method\cite{wang_landau}, require tuning several free
+parameters that lack both ``obvious'' values and systematic means for
+assigning them. This work will thus attempt to quantify and compare
+the effectiveness of a few such methods.
 
@@
 \begin{algorithm}[tb]
-  \caption{Unbiased Monte Carlo fluid simulation}
+  \caption{Metropolis Monte Carlo fluid simulation}
   \label{alg:MC}
@@
 
-  \item Construct an initial ``typical'' fluid configuration.
+  \item Construct an initial ``typical'' (i.e. non-ordered) fluid
+    configuration. \fixme{explain how}
 
@@
 
-Algorithm \ref{alg:MC} provides a sketch of unbiased Monte Carlo fluid
-simulations. Such an algorithm is ``unbiased'' in the sense that it
-collects statistics (i.e. data) on all valid system configurations
-with equal probability. Statistics whose collection time scales as
-$\mathcal O\p{1}$ with system size, meaning that increasing the number
-of spheres $N$ in the simulation does not affect collection time, can
-be collected after every move (defined in the algorithm). Statistics
-whose collection time scales as $\mathcal O\p{N}$, meaning that
-doubling $N$ doubles collection time, can be collected after each
-iteration. In general, collection with $\mathcal O\sp{\chi\p{N}}$ time
-scaling should not occur more often than once every $\chi\p{N}$ moves
-(where $\chi\p{N}$ may be an arbitrary function of only $N$ void of
-constant prefactors, e.g. $N\log N$, or $2^N$). These collection rules
-ensure that scaling up simulations does not cause them to
-asymptotically spend all of their time collecting statistics, and no
-time actually simulating the fluid. Collected statistics are used to
-find thermodynamic properties of the simulated fluid.
+Algorithm \ref{alg:MC} provides a sketch of unbiased, Metropolis Monte
+Carlo fluid simulations. Such an algorithm is ``unbiased'' in the
+sense that it collects statistics (i.e. data) on all valid system
+configurations with equal probability. Statistics whose collection
+time scales as $\mathcal O\p{1}$ with system size, meaning that
+increasing the number of spheres $N$ in the simulation does not affect
+collection time, can be collected after every move (defined in the
+algorithm). Statistics whose collection time scales as $\mathcal
+O\p{N}$, meaning that doubling $N$ doubles collection time, can be
+collected after each iteration. In general, collection with $\mathcal
+O\sp{\chi\p{N}}$ time scaling should not occur more often than once
+every $\chi\p{N}$ moves (where $\chi\p{N}$ may be an arbitrary
+function of only $N$, void of constant prefactors, e.g. $N\log N$, or
+$2^N$). These collection rules ensure that scaling up simulations does
+not cause them to asymptotically spend all of their time collecting
+statistics, and no time actually simulating the fluid. Collected
+statistics are used to find thermodynamic properties of the simulated
+fluid.
 
@@
 
-Due to the fact that unbiased Monte Carlo simulations sample all of
+Due to the fact that Metropolis Monte Carlo simulations sample all of
 state space randomly and without preference, a histogram $H\p{X}$ of
@@
 range of possible $X$ is so low that it is practically impossible (via
-unbiased Monte Carlo) to sufficiently sample $R$, that is, accumulate
-a statistically significant histogram $H\p{X\in R}$, in any reasonable
-amount of time.
+Metropolis Monte Carlo) to sufficiently sample $R$, that is,
+accumulate a statistically significant histogram $H\p{X\in R}$, in any
+reasonable amount of time.
 
@@
 
-Histogram methods provide a means to address unbiased Monte Carlo's
+Histogram methods provide a means to address Metropolis Monte Carlo's
 inability to sufficiently sample regions of parameter space with a low
@@
 
-  \item Construct an initial ``typical'' fluid configuration.
+  \item Construct an initial typical fluid configuration.
 
@@
 ``weights'' interchangeably.
+
+\fixme{mention that simulating with flat weights $w\p{E}=w_0$ is the
+  same as running a Metropolis MC simulation.}
 
@@
 
+\fixme{Make a note on storing the logarithm of weights and
+  periodically rescaling them (resetting the zero on $\ln w$) to
+  reduce numerical error.}
+
 \subsection{The flat histogram (multi-canonical) method}
-\label{sec:flat_historam}
+\label{sec:flat_histogram}
 
@@
 fact one of the system properties which a simulation is intended to
-determine. A naive way to implement the flat histogram method without
-knowing a priori the density of states $D\p{E}$ involves first running
-a simulation without weights (or with constant weights $w\p{E}=w_0$)
-while collecting a histogram $H\p{E}$ of the observed energies
-$E$. After some time (loosely defined), the histogram $H\p{E}$ may be
-taken as an approximation of the density of states $\tilde D\p{E}$
-(see Section \ref{sec:histogram_methods}), and used via
-(\ref{eq:flat_weights}) to generate a weight array for an actual
-simulation.
-
-The problem with this implementation is that histogram methods are
-generally necessary when some range of energies $R$ is inaccessible
-via unbiased Monte Carlo simulations. In such a case, the energy
-histogram $H\p{E\in R}$ after simulating for any reasonable amount of
-time will be statistically insignificant or null, invalidating the
-approximation $H\p{E}\approx\tilde D\p{E}$. This method will therefore
-only ``flatten'' the histogram at energies $E$ which have been
-sufficiently sampled for $H\p{E}$ to be statistically significant. The
-end goal in implementing histogram methods, however, is precisely to
-sample those energies that cannot be sufficiently sampled without
-employing clever schemes to bias Monte Carlo simulations. The flat
-histogram method therefore does not directly help us study the
-square-well fluid. Nonetheless, the this method has pedagogical value
-by enabling discussion of an ``optimal'' weight array for generating a
+determine. Nonetheless, the this method has pedagogical value by
+enabling discussion of an ``optimal'' weight array for generating a
 flat energy histogram in simulations. All algorithms to construct
@@
 
+\begin{algorithm}[b]
+  \caption{A naive flat histogram method}
+  \label{alg:flat_histogram}
+  \begin{alg}
+
+  \item Construct an initial typical fluid configuration.
+
+  \item Simulate (without weights) for some time $\tau$, which may be
+    real time, computer time, or a number of simulation iterations,
+    collecting a histogram $H\p{E}$ of energy observations after every
+    move.
+
+  \item Set weights $w\p{E}=1/H\p{E}$. At energies with $H\p{E}=0$,
+    set $w\p{E}=1$.
+
+  \end{alg}
+\end{algorithm}
+
+Algorithm \ref{alg:flat_histogram} provides a naive way to implement
+the flat histogram method without knowing a priori the density of
+states $D\p{E}$. This algorithm involves first running a simulation
+without weights (or with constant weights $w\p{E}=w_0$) while
+collecting a histogram $H\p{E}$ of the observed energies $E$. After
+some time (loosely defined), the histogram $H\p{E}$ may be taken as an
+approximation of the density of states $\tilde D\p{E}$ (see Section
+\ref{sec:histogram_methods}), and used via (\ref{eq:flat_weights}) to
+generate a weight array for an actual simulation. Algorithm
+\ref{alg:flat_histogram} has one free parameter, which determines the
+amount of time for which to collect an energy histogram $H\p{E}$.
+
+The problem with this implementation of the flat histogram method is
+that histogram methods are generally necessary when some range of
+energies $R$ is inaccessible via Metropolis Monte Carlo
+simulations. In such a case, the energy histogram $H\p{E\in R}$ after
+simulating for any reasonable amount of time will be statistically
+insignificant or null, invalidating the approximation
+$H\p{E}\approx\tilde D\p{E}$. This method will therefore only
+``flatten'' the histogram at energies $E$ which have been sufficiently
+sampled for $H\p{E}$ to be statistically significant. The end goal in
+implementing histogram methods, however, is precisely to sample those
+energies that cannot be sufficiently sampled without employing clever
+schemes to bias Monte Carlo simulations. The flat histogram method
+therefore does not directly help us study the square-well fluid, but
+it does provide the basis for the method in Section
+\ref{sec:simple_flat}.
+
 \subsection{The simple flat method}
@@
 
-\fixme{Explain, provide algorithm. This method is our own.}
+The ``simple flat'' method, developed by ourselves, extends the
+implementation of the flat histogram method discussed in Section
+\ref{sec:flat_histogram} to an algorithm intended to be the simplest
+viable broad energy histogram method. While we do not expect this
+method to outperform other, more sophisticated methods, the simple
+flat method should provide a lower bar and a standard of comparison
+for the performance of other methods. Any nontrivial histogram method
+should, at a minimum, consistently outperform the simple flat method
+in order to be considered for any applications.
 
-\begin{algorithm}[H]
-  \caption{Simple flat initialization}
+The idea behind the simple flat method is such: after simulating for
+some time, the energy histogram $H\p{E}$ will have a peak, or some
+other uneven distribution of observations, in a region of energies $R$
+for which $H\p{E\in R}$ statistically significant. One can thus use
+the existing histogram to construct weights which should flatten
+$H\p{R}$. Subsequent simulations should then spend less time at
+energies with the highest state densities, and more time at energies
+with relatively low state densities. The energy histogram from such
+simulations should therefore be statistically significant in a larger
+range of energies, improving the available estimate for a density of
+states with which to compute flat histogram weights. One could then
+use the improved estimate to construct new weights, and repeat this
+simulation and updating process until simulations satisfy some end
+condition (in fact, all histograms will require end conditions, which
+is discussed in Section \ref{sec:end_conditions}).
+
+\begin{algorithm}[!b]
+  \caption{The simple flat method}
   \label{alg:simple_flat}
+  \begin{alg}
+
+  \item Construct an initial typical fluid configuration.
+
+  \item Set $k=0$, set initial weights $w_0\p{E}=1$ for all $E$, and
+    choose an initial number of iterations $n_0$ for which to
+    simulate.
+
+  \item Simulate for $n_k$ iterations with weights $w_k\p{E}$,
+    collecting a histogram $H_k\p{E}$ of energy observations after
+    every move.
+    \label{alg:simple_flat_sim}
+
+  \item If some predetermined initialization end condition has
+    \emph{not} been met, then:
+    \begin{enumerate}
+    \item increment $k\leftarrow k+1$,
+    \item set $n_k=u\p{n_{k-1}}$ with a predetermined function
+      $u:u\p{n}\ge n$,
+      \label{alg:simple_flat_f}
+    \item update weights as $w_k\p{E}=w_{k-1}\p{E}/H_{k-1}\p{E}$ for
+      all $E$ at which $H_{k-1}\p{E}\ne0$,
+      \label{alg:simple_flat_weight_update}
+    \item return to step \ref{alg:simple_flat_sim}.
+    \end{enumerate}
+
+  \end{alg}
 \end{algorithm}
+
+Algorithm \ref{alg:simple_flat} provides an implementation of the
+simple flat method. This algorithm has two free ``parameters'': the
+initial number of iterations $n_0$ for which to simulate, and the
+function $u\p{n}$ used in step \ref{alg:simple_flat_f}. With an
+appropriate choice of the function $u\p{n}$, this algorithm should not
+be particularly sensitive to the value of $n_0$.
+
+The most interesting part of Algorithm \ref{alg:simple_flat} is step
+\ref{alg:simple_flat_weight_update}, which updates the weights as
+$w_k\p{E}=w_{k-1}\p{E}/H_{k-1}\p{E}$. The reasoning behind this step
+is such: say that for two energies $E_0$ and $E_b$ we have that
+$H\p{E_0}=\bk{H}_E$ and $H\p{E_b}=b\bk{H}_E$, where $\bk{H}_E$ is the
+mean histogram value over all energies. In this case, we wish to
+decrease the current simulation bias on $E_b$ by a factor of $b$ while
+keeping the bias on $E_0$ the same. As the weights $w\p{E}$ are
+proportional to the simulation bias on $E$, dividing the weights
+$w\p{E}$ by the current histogram $H\p{E}$ achieves the desired
+changes to the simulation biases. The factor $\bk{H}_E$ in this weight
+update does not affect results, and is in practice immediately scaled
+out of the weight array.
+
+In our implementation of the simple flat method, we chose $u\p{n}=2n$
+so that even if the choice of initial $n_0$ is too small, $n_k=2^kn_0$
+should quickly approach the appropriate scale for the system at
+hand. We also used $n_0=L\exp\p{\epsilon/kT_{\t{min}}}$, where $L$ is
+the number of energy levels in the given square-well fluid and
+$T_{\t{min}}$ is a minimum temperature of interest for the current
+simulation. The factor of $L$ appears because initialization time
+should scale with the range of energies of the simulated system. If we
+have twice the number of energies, we should explore energy space for
+twice as long. To estimate $L$, we multiplied the maximum number of
+spheres $M$ which fit within a radius of $\lambda\sigma$ of a single
+sphere on a face-centered cubic lattice (i.e. the maximum number of
+spheres with which a given sphere could interact) by the total number
+of spheres $N$, and divided the result by two so as to not
+double-count interactions between spheres (so $L=MN/2$). The
+exponential $\exp\p{\epsilon/kT_{\t{min}}}$, a Boltzmann factor for
+two adjacent energies, appears because lower minimum temperatures of
+interest warrant sampling lower energies, which in turn requires
+simulating for longer (in step \ref{alg:simple_flat_sim} of Algorithm
+\ref{alg:simple_flat}) because these energies have lower state
+densities. The introduction of a minimum temperature may seem ad-hoc,
+but all simulations in this paper require a choice of minimum
+temperature anyways (see Section \ref{sec:min_energy}). In this paper,
+we will always use $kT_{\t{min}}=0.2\epsilon$.
 
@@
 
+The Wang-Landau method is the first published histogram method we
+discuss, and is the standard method\fixme{cite} to achieve broad
+energy sampling in Monte Carlo simulations\cite{wang_landau,
+  wang_landau_analysis, wang_landau_mod}.
+
 \fixme{Explain the motivation behind the Wang-Landau method, and
-  provide the algorithm. Talk about the free parameters and roughly
-  what they control.}
+  explain/provide the algorithm.}
 
-\begin{algorithm}[H]
-  \caption{Wang-Landau calculation of weights}
+\begin{algorithm}[tb]
+  \caption{Wang-Landau initialization of weights}
   \label{alg:wang_landau}
+  \begin{alg}
+
+  \item Construct an initial typical fluid configuration
+
+  \item Set $k=0$, choose some factor $f_0>0$, and initialize
+    $w\p{E}=1$ for all $E$.
+
+  \item Simulate for $n$ iterations. After each move, increment the
+    histogram $H\p{E}$ of energy observations and update
+    $w\p{E}\leftarrow w\p{E}/f_k$, where $E$ is the current energy of
+    the system.
+    \label{alg:wang_landau_sim}
+
+  \item If the histogram $H\p{E}$ is sufficiently flat, i.e. if it
+    satisfies some predetermined ``flatness'' condition $C\sp{H}$,
+    then
+    \begin{enumerate}[label=(\alph*)]
+    \item increment $k\leftarrow k+1$,
+    \item set $f_k=u\p{f_{k-1}}$ with a predetermined function
+      $u:u\p{f}<f$,
+    \item if $f_k>c$ for some predetermined $c$, return to step
+      \ref{alg:wang_landau_sim}.
+    \end{enumerate}
+
+  \end{alg}
 \end{algorithm}
-
-% \subsubsection{Modifications}
-% \label{sec:wang_landau_mod}
-
-% \fixme{Talk about our modifications to Wang-Landau made to adapt it
-% to the square-well fluid, as well as (if I have time) other research
-% groups' modifications.}
 
@@
 \begin{algorithm}[H]
-  \caption{Finding weights by optimizing energy walker currents}
+  \caption{Modifying weights via the optimized ensemble}
   \label{alg:optimized_ensemble}
 \end{algorithm}
-uu
+
 \subsection{The transition matrix Monte Carlo (TMMC) method}
@@
 \begin{algorithm}[H]
-  \caption{Transition matrix Monte Carlo simulation}
+  \caption{Transition matrix Monte Carlo initialization}
   \label{alg:tmmc}
@@
 
+\fixme{Mention that in all initializations which simulate for some
+  time with unchanging weights (i.e. simple flat), the weight array
+  should be set to canonical weights below the minimum important
+  energy, as soon as it is identified.}
+
+\fixme{Mention that weights should be set to canonical weights below
+  the minimum important energy after all initializations.}
+
 \subsubsection{Computing the density of states}
@@
 
-% \subsection{End conditions}
-% \label{sec:end_conditions}
+\subsection{End conditions}
+\label{sec:end_conditions}
 

### 2015-04-28 papers/histogram/thesis/thesis.tex

@@
 which dominate the partition function at low (i.e. liquid-state)
-temperatures. Several generic Monte Carlo ``histogram methods'' for
-collecting statistics on low energy system states have been developed
-\fixme{by whom?}, but little or no literature exists on their
-systematic comparison and their application to the square-well
+temperatures. Previous works have developed several generic Monte
+Carlo histogram methods for collecting statistics on low energy system
+states, but little or no literature exists on these methods'
+systematic comparison, as well as their application to the square-well
 fluid. The square-well fluid in particular introduces application
@@
 
-\fixme{talk more about phase transitions and the critical point?}
+\fixme{give more intro about phase transitions and the critical
+  point?}
 
@@
 simulations requires exploring the boundary between its liquid and
-gaseous phases. The square-well fluid (Section \ref{sec:sw_fluid}) is
-the simplest system with a liquid-vapor phase transition, and is
-therefore of interest for studying the critical point of fluids. For a
-generic square-well fluid to be observed in a liquid state, it must
-have a localized distribution of the spheres which make up the
-fluid. Such a distribution implies a low energy fluid state, as all
-interactions between spheres are attractive. Conversely, a low energy
-liquid state of a square-well fluid otherwise observable in both
-liquid and gaseous phases implies a localized distribution of
-spheres. The search for liquid states of the square-well fluid is thus
-equivalent to the search for low energy states.
+gaseous phases. The square-well fluid is the simplest system with a
+liquid-vapor phase transition, and is therefore of interest for
+studying the critical point of fluids. For a generic square-well fluid
+to be observed in a liquid state, it must have a localized
+distribution of the spheres which make up the fluid. Such a
+distribution implies a low energy fluid state, as all interactions
+between spheres are attractive. Conversely, a low energy liquid state
+of a square-well fluid otherwise observable in both liquid and gaseous
+phases implies a localized distribution of spheres. The search for
+liquid states of the square-well fluid is thus equivalent to the
+search for low energy states.
 
 Monte Carlo simulations are a standard means for studying the
-thermodynamic properties of a system. In the most straightforward
-implementation of Monte Carlo fluid simulations (Section
-\ref{sec:mc_sim}), however, highly localized distributions of fluid
-elements are extremely unlikely to occur, making such simulations
-impractical for studying the critical point of the square-well
-fluid. Several generic computational methods, called histogram methods
-(Section \ref{sec:histogram_methods}), exist for dealing with such a
-problem; namely, that interesting regions of a system's state space
-cannot be sampled via standard Monte Carlo methods in a reasonable
-amount of time. There is no consensus, however, on the relative
-efficacy of these methods. Furthermore, some methods, such as the
-Wang-Landau method\cite{wang_landau}, require tuning several free
-parameters that lack both ``obvious'' values and systematic means for
-assigning them. This work will thus attempt to quantify and compare
-the effectiveness of a few such methods.
+equilibrium thermodynamic properties of a system. In the most
+straightforward implementation of Monte Carlo fluid simulations,
+however, highly localized distributions of fluid elements are
+extremely unlikely to occur, making such simulations impractical for
+studying the critical point of the square-well fluid. Several generic
+computational methods, called histogram methods, exist for dealing
+with such a problem; namely, that interesting regions of a system's
+state space cannot be sampled via standard Monte Carlo methods in a
+reasonable amount of time. There is no common consensus, however, on
+the relative efficacy of these methods. Furthermore, some methods,
+such as the Wang-Landau method\cite{wang_landau}, require tuning
+several free parameters that lack both ``obvious'' values and
+systematic means for assigning them. This work will thus attempt to
+quantify and compare the effectiveness of a few such methods as
+applied to the square-well fluid.
 
@@
 every $\chi\p{N}$ moves (where $\chi\p{N}$ may be an arbitrary
-function of only $N$, void of constant prefactors, e.g. $N\log N$, or
+function of $N$ void of constant prefactors, e.g. $N\log N$, or
 $2^N$). These collection rules ensure that scaling up simulations does
-not cause them to asymptotically spend all of their time collecting
+not cause them to asymptotically spend all computation time collecting
 statistics, and no time actually simulating the fluid. Collected
@@
 and involves considering the limit of numerical results as the number
-of spheres $N\to\infty$ (with a constant filling fraction).
+of spheres $N\to\infty$ (keeping the filling fraction $\eta$
+constant).
 
@@
 \begin{enumerate*}[label=\roman*)]
-\item $D\p{Y_0}$ is proportional to the number of microstates $s$ for
-  which $Y\p{s}=Y_0$, and
+\item for arbitrary $Y_0$, the value of $D\p{Y_0}$ is proportional to
+  the number of microstates $s$ for which $Y\p{s}=Y_0$, and
 \item $\sum_YD\p{Y}=1$.
 \end{enumerate*}
-The temperature dependence of $\bk{X}_T$ can be expressed in terms of
-the expectation value $\bk{X}_E$ and the density of states $D\p{E}$ by
+The units of $D\p{Y}$ are inverse to the units of $Y$. The temperature
+dependence of $\bk{X}_T$ can be expressed in terms of the expectation
+value $\bk{X}_E$ and the density of states $D\p{E}$ by
 \begin{align}
@@
 To reduce redundant computations and numerical error in
-implementation, we will use a partition function with an unnormalized
-density of states,
+implementation, we will generally use a partition function with an
+unnormalized density of states $\tilde D\p{E}$, i.e.
 \begin{align}
@@
 the desired yields (e.g.  density of states, heat capacity) of a
-simulation. A ``histogram method'' is an algorithm for determining a
-weight function appropriate for a particular simulation.
+simulation. A histogram method is simply an algorithm or procedure for
+determining a weight function appropriate for a particular simulation.
 
@@
 simulation and updating process until simulations satisfy some end
-condition (in fact, all histograms will require end conditions, which
-is discussed in Section \ref{sec:end_conditions}).
-
-\begin{algorithm}[!b]
+condition (in fact, most histogram methods will require end
+conditions, which is discussed in Section \ref{sec:end_conditions}).
+
+\begin{algorithm}[!t]
   \caption{The simple flat method}
@@
 
-  \item Set $k=0$, set initial weights $w_0\p{E}=1$ for all $E$, and
-    choose an initial number of iterations $n_0$ for which to
-    simulate.
-
-  \item Simulate for $n_k$ iterations with weights $w_k\p{E}$,
-    collecting a histogram $H_k\p{E}$ of energy observations after
-    every move.
+  \item Set $k=0$, choose an initial number of iterations $n$ for
+    which to simulate, and initialize weights $w\p{E}=1$ for all $E$.
+
+  \item Simulate for $n$ iterations (with weights), collecting a
+    histogram $H\p{E}$ of energy observations after every move.
     \label{alg:simple_flat_sim}
@@
   \item If some predetermined initialization end condition has
-    \emph{not} been met, then:
+    \emph{not} been met, then
     \begin{enumerate}
-    \item increment $k\leftarrow k+1$,
-    \item set $n_k=u\p{n_{k-1}}$ with a predetermined function
-      $u:u\p{n}\ge n$,
+    \item increment $k\leftarrow k+1$
+    \item set $n_k=un_{k-1}$ with a predetermined factor $u\ge1$,
       \label{alg:simple_flat_f}
-    \item update weights as $w_k\p{E}=w_{k-1}\p{E}/H_{k-1}\p{E}$ for
-      all $E$ at which $H_{k-1}\p{E}\ne0$,
+    \item update weights as $w\p{E}\leftarrow w\p{E}/H\p{E}$ for all
+      $E$ at which $H\p{E}\ne0$,
       \label{alg:simple_flat_weight_update}
@@
 Algorithm \ref{alg:simple_flat} provides an implementation of the
-simple flat method. This algorithm has two free ``parameters'': the
-initial number of iterations $n_0$ for which to simulate, and the
-function $u\p{n}$ used in step \ref{alg:simple_flat_f}. With an
-appropriate choice of the function $u\p{n}$, this algorithm should not
-be particularly sensitive to the value of $n_0$.
+simple flat method. This algorithm has two free parameters: the
+initial number of iterations $n_0$ for which to simulate and the
+factor $u$. Due to the exponential growth of $n_k=u^kn_0$, this
+algorithm should not be sensitive to the value of $n_0$.
 
@@
 \ref{alg:simple_flat_weight_update}, which updates the weights as
-$w_k\p{E}=w_{k-1}\p{E}/H_{k-1}\p{E}$. The reasoning behind this step
-is such: say that for two energies $E_0$ and $E_b$ we have that
+$w\p{E}\leftarrow w\p{E}/H\p{E}$. The reasoning behind this step is
+such: say that for two energies $E_0$ and $E_b$ we have that
 $H\p{E_0}=\bk{H}_E$ and $H\p{E_b}=b\bk{H}_E$, where $\bk{H}_E$ is the
@@
 keeping the bias on $E_0$ the same. As the weights $w\p{E}$ are
-proportional to the simulation bias on $E$, dividing the weights
-$w\p{E}$ by the current histogram $H\p{E}$ achieves the desired
-changes to the simulation biases. The factor $\bk{H}_E$ in this weight
-update does not affect results, and is in practice immediately scaled
-out of the weight array.
-
-In our implementation of the simple flat method, we chose $u\p{n}=2n$
-so that even if the choice of initial $n_0$ is too small, $n_k=2^kn_0$
-should quickly approach the appropriate scale for the system at
-hand. We also used $n_0=L\exp\p{\epsilon/kT_{\t{min}}}$, where $L$ is
-the number of energy levels in the given square-well fluid and
-$T_{\t{min}}$ is a minimum temperature of interest for the current
-simulation. The factor of $L$ appears because initialization time
-should scale with the range of energies of the simulated system. If we
-have twice the number of energies, we should explore energy space for
-twice as long. To estimate $L$, we multiplied the maximum number of
-spheres $M$ which fit within a radius of $\lambda\sigma$ of a single
-sphere on a face-centered cubic lattice (i.e. the maximum number of
-spheres with which a given sphere could interact) by the total number
-of spheres $N$, and divided the result by two so as to not
-double-count interactions between spheres (so $L=MN/2$). The
-exponential $\exp\p{\epsilon/kT_{\t{min}}}$, a Boltzmann factor for
-two adjacent energies, appears because lower minimum temperatures of
-interest warrant sampling lower energies, which in turn requires
-simulating for longer (in step \ref{alg:simple_flat_sim} of Algorithm
+proportional to the simulation bias on each respective energy $E$,
+dividing the weights $w\p{E}$ by the current histogram $H\p{E}$
+achieves the desired changes to the simulation biases. The factor
+$\bk{H}_E$ in this weight update has no effect on the function of the
+weights, and is in practice immediately scaled out of the weight
+array.
+
+In our implementation of the simple flat method, we used $u=2$ and
+$n_0=L\exp\p{\epsilon/kT_{\t{min}}}$, where $L$ is the number of
+energy levels of the given square-well fluid and $T_{\t{min}}$ is a
+minimum temperature of interest for the current simulation. The factor
+of $L$ appears because initialization time should scale with the range
+of energies of the simulated system. If we have twice the number of
+energies, we should explore energy space for twice as long. To
+estimate $L$, we multiplied the maximum number of spheres $M$ which
+fit within a radius of $\lambda\sigma$ of a single sphere on a
+face-centered cubic lattice (i.e. the maximum number of spheres with
+which a given sphere could interact) by the total number of spheres
+$N$, and divided the result by two so as to not double-count
+interactions between spheres (so $L=MN/2$). The exponential
+$\exp\p{\epsilon/kT_{\t{min}}}$, a Boltzmann factor for two adjacent
+energies, appears because lower minimum temperatures of interest
+warrant sampling lower energies, which in turn requires simulating for
+longer (in step \ref{alg:simple_flat_sim} of Algorithm
 \ref{alg:simple_flat}) because these energies have lower state
@@
 but all simulations in this paper require a choice of minimum
-temperature anyways (see Section \ref{sec:min_energy}). In this paper,
-we will always use $kT_{\t{min}}=0.2\epsilon$.
+temperature anyways (addressed in Section \ref{sec:energy_range}). In
+this paper, we will always use $kT_{\t{min}}=0.2\epsilon$.
 
@@
 The Wang-Landau method is the first published histogram method we
-discuss, and is the standard method\fixme{cite} to achieve broad
-energy sampling in Monte Carlo simulations\cite{wang_landau,
-  wang_landau_analysis, wang_landau_mod}.
-
-\fixme{Explain the motivation behind the Wang-Landau method, and
-  explain/provide the algorithm.}
-
-\begin{algorithm}[tb]
+discuss\cite{wang_landau, wang_landau_analysis, wang_landau_mod}. This
+method has been used to study a wide variety of systems\fixme{which
+  systems? cite}, and is the standard method\fixme{cite} to achieve
+broad energy sampling via Monte Carlo simulations. \fixme{more intro
+  on WL}
+
+Unlike the simple flat method, the Wang-Landau method modifies the
+weight array on the fly. After every move during initialization
+Wang-Landau decreases the weight $w\p{E}$ on the current energy $E$ by
+some factor of $f$, so as to decrease weights on the most commonly
+observed energies relative to those of the least commonly observed
+energies. Eventually, such a simulation with constant modification of
+the weights $w\p{E}$ should yield a flat histogram
+$H\p{E}\approx\bk{H}_E$ for all $E$, as the modifications to $w\p{E}$
+increasingly pushes the simulated system away from the most sampled
+energies, and toward the least sampled energies. When the energy
+histogram is sufficiently flat, $H\p{E}$ is reset (i.e. set to
+$H\p{E}=0$ for all $E$), the factor $f$ is decreased (geometrically
+approaching unity), and the entire process is repeated until $f-1$
+falls below some cutoff $c\ll 1$.
+
+Resetting and repeating this initialization process is necessary to
+perform fine tuning and refinement of the weights. A flat histogram
+after a simulation in which $w\p{E}$ has changed does not guarantee
+that simulating with the final weights will likewise result in a flat
+histogram. Eventually, when $f-1<c\ll 1$, i.e. $f\approx 1$, the
+initialization process no longer makes any appreciable modifications
+to the weights $w\p{E}$, so that upon achieving a flat histogram one
+can be confident that a regular simulation which fixes the current
+weights will yield similar results.
+
+Interestingly, the flatness condition $C\sp{H}$ for updating $f$ and
+restarting the process should not be too stringent: if the histogram
+is completely flat, say $H\p{E}=H_0$ for all $E$, then the weights
+$w\p{E}$ have all been modified by the same factor of $f^{H_0}$.
+Remembering the scale-invariance of $w\p{E}$, this result would mean
+that the weights have not actually changed at all! Similarly, too lax
+of a flatness condition will prompt the initialization process to
+\begin{enumerate*}[label=\roman*)]
+\item reset and decrease $f$ too early, i.e. when the weight array
+  could still use some heavy handed adjustment, and
+\item quit before one has any reason to think that simulations might
+  yield a flat histogram.
+\end{enumerate*}
+Tuning the flatness condition is thus crucial to the success of the
+Wang-Landau method.
+
+\begin{algorithm}[!b]
   \caption{Wang-Landau initialization of weights}
@@
 
-  \item Set $k=0$, choose some factor $f_0>0$, and initialize
+  \item Set $k=0$, choose some factor $f_0>1$, and initialize
     $w\p{E}=1$ for all $E$.
@@
 
-  \item If the histogram $H\p{E}$ is sufficiently flat, i.e. if it
-    satisfies some predetermined ``flatness'' condition $C\sp{H}$,
-    then
-    \begin{enumerate}[label=(\alph*)]
+  \item If the histogram $H\p{E}$ is not sufficiently flat, i.e. if it
+    fails a ``flatness'' condition $C\sp{H}$, return to step
+    \ref{alg:wang_landau_sim}, otherwise
+    \begin{enumerate}
+    \item reset $H\p{E}=0$ for all $E$,
     \item increment $k\leftarrow k+1$,
-    \item set $f_k=u\p{f_{k-1}}$ with a predetermined function
-      $u:u\p{f}<f$,
-    \item if $f_k>c$ for some predetermined $c$, return to step
+    \item set $f_k=\sqrt[u]{f_{k-1}}$ for some predetermined $u>1$,
+      \label{alg:wang_landau_update}
+    \item if $f_k\ge c$ for some predetermined $c$, return to step
       \ref{alg:wang_landau_sim}.
@@
 \end{algorithm}
+
+Algorithm \ref{alg:wang_landau} spells out the Wang-Landau
+initialization process, which has five free parameters: the number of
+iterations $n$ to run in step \ref{alg:wang_landau_sim}, the initial
+factor $f_0$, the update factor $u$, the flatness condition $C\sp{H}$,
+and the cutoff $c$. The first of these parameters, $n$, controls how
+often to check the flatness condition $C\sp{H}$ and run updates, if
+necessary. The value of $n$ is not too important, so long as it is
+reasonable. Too small value of $n$ will waste computation time
+determining $C\sp{H}$, while too large of a value will waste time
+simulating when $C\sp{H}$ has already been satisfied. In general, $n$
+should be proportional to the computation time of $C\sp{H}$, which
+scales with energy range of the simulated system. We therefore use
+$n=L$, where $L$ is the number of energy levels for the given
+square-well fluid (as discussed in Section \ref{sec:simple_flat}).
+Papers in the literature typically neglect providing any heuristic or
+formula for $n$, reporting a system-independent $n=10^3$ or
+$10^5$\cite{wang_landau, wang_landau_mod}.
+
+The next parameter, $f_0$, controls the initial amount by which to
+adjust the weight of the current energy after each move. As mentioned
+in Algorithm \ref{alg:wang_landau}, one should always have $f_0>1$,
+but the appropriate value of $f_0$ will generally depend on the system
+at hand. Too large a value of $f_0$ will magnify stochastic error in
+initialization, while too small a value of $f_0$ will cause
+simulations to take an exceedingly long time to converge.  Lacking any
+heuristics for assigning $f_0$, literature typically suggests
+$f_0=e$\cite{wang_landau}; after some experimentation with the free
+parameters in the Wang-Landau method, we used $f_0=e^{1/8}$ (in
+practice, one usually works with $\ln f$, making the commonly
+suggested value $\ln f_0=1$ and our own $\ln f_0=1/8$).
+
+The factor $u$ appearing in step \ref{alg:wang_landau_update} of
+Algorithm \ref{alg:wang_landau} controls the amount by which to adjust
+$f$ when the histogram $H\p{E}$ is sufficiently flat. In principle,
+one need not take a root of $f$ at each reset: the general idea behind
+Algorithm \ref{alg:wang_landau} requires only $f_k<f_{k-1}$ to
+converge (assuming that $C\sp{H}$ can be satisfied) and
+$\lim_{k\to\infty}f_k=1$ to terminate. Taking the $u$-th root of $f$
+at each increment of $k$ is a natural and convenient iterative means
+to satisfy these requirements. Previous papers have used $u=2$, which
+we have adopted as well\cite{wang_landau, wang_landau_mod}.
+
+\fixme{Walk through $C\sp{H}$ and $c$. Finish discussion of WL.}
 

### 2015-05-01 papers/histogram/thesis/thesis.tex
additions to histogram thesis

@@
 
-\renewcommand{\title}{Optimizing Monte Carlo state space exploration
-  of the square-well fluid}
-\renewcommand{\author}{Michael A. Perlin}
+\renewcommand{\title}{Optimizing Monte Carlo simulation of the
+  square-well fluid} \renewcommand{\author}{Michael A. Perlin}
 \newcommand{\institution}{Oregon State University}
 \newcommand{\department}{Department of Physics}
-\newcommand{\supervisor}{Dr. David Roundy}
-\renewcommand{\date}{21 April 2015}
+\newcommand{\supervisor}{Dr. David Roundy} \renewcommand{\date}{21
+  April 2015}
 
@@
 
-We wish to identify and develop efficient numerical methods for
-determining thermodynamic properties of the square-well fluid in order
-to test square-well density functional theories. The liquid-vapor
-phase interface is an interesting regime for testing density
-functional theories, but unbiased (i.e. Metropolis) Monte Carlo
-simulations are incapable of sampling the low energy fluid states
-which dominate the partition function at low (i.e. liquid-state)
-temperatures. Previous works have developed several generic Monte
-Carlo histogram methods for collecting statistics on low energy system
-states, but little or no literature exists on these methods'
-systematic comparison, as well as their application to the square-well
-fluid. The square-well fluid in particular introduces application
-challenges not manifest in traditional models for testing and
-benchmarking such numerical techniques (e.g. the Ising model). We
-propose to implement these methods in Monte Carlo simulations of the
-square-well fluid, determine appropriate performance metrics for their
-comparison, and identify those methods which are most efficient and
-effective at sampling otherwise improbable system states.
+We identify and develop efficient numerical methods for determining
+thermodynamic properties of the square-well fluid in order to test
+square-well density functional theories. The liquid-vapor phase
+interface is an interesting regime for testing density functional
+theories, but unbiased Metropolis Monte Carlo simulations are
+incapable of sampling the low energy fluid states which dominate the
+partition function at low (i.e. liquid-state) temperatures. Previous
+works have developed several generic Monte Carlo histogram methods for
+collecting statistics on low energy system states, but little or no
+literature exists on these methods' systematic comparison, as well as
+their application to the square-well fluid. The square-well fluid in
+particular introduces application challenges not manifest in
+traditional models for testing and benchmarking such numerical
+techniques (e.g. the Ising model). We propose to implement these
+methods in Monte Carlo simulations of the square-well fluid, determine
+appropriate performance metrics for their comparison, and identify
+those methods which are most efficient and effective at sampling
+otherwise improbable system states.
 
@@
 
-  \item Attempt to change the position of one sphere (in general, a
-    single fluid ``atom'' or ``molecule'') to a random location,
-    rejecting the change if it results in a forbidden fluid
+  \item Randomly attempt to change the position of one sphere (in
+    general, a single fluid ``atom'' or ``molecule'') to another
+    location, rejecting the change if it results in a forbidden fluid
     configuration (e.g. two or more spheres overlap) and accepting the
@@
 
-  \item Attempt to change the position of one sphere, rejecting the
-    change if
+  \item Randomly attempt to change the position of one sphere,
+    rejecting the change if
     \begin{enumerate*}[label=\roman*)]
@@
     \item a newly chosen random number on the interval $\sp{0,1}$ is
-      larger than the probability given by (\ref{eq:move_prob}).
+      larger than the probability determined by (\ref{eq:move_prob})
+      for the initial and final states $s_i$ and $s_f$, respectively.
     \end{enumerate*}
@@
 schemes to bias Monte Carlo simulations. The flat histogram method
-therefore does not directly help us study the square-well fluid, but
-it does provide the basis for the method in Section
-\ref{sec:simple_flat}.
+therefore does not directly help us study the square-well fluid. This
+method does, however, motivate the method in Section
+\ref{sec:simple_flat}, and provides some of the theory behind the
+methods in Sections \ref{sec:tmmc} and \ref{sec:oetmmc}.
 
@@
 
+\fixme{find more sources to cite for some of the claims made in this
+  section}
+
 The Wang-Landau method is the first published histogram method we
-discuss\cite{wang_landau, wang_landau_analysis, wang_landau_mod}. This
+discuss\cite{wang_landau, wang_landau_mod, wang_landau_analysis}. This
 method has been used to study a wide variety of systems\fixme{which
@@
 
-Interestingly, the flatness condition $C\sp{H}$ for updating $f$ and
-restarting the process should not be too stringent: if the histogram
-is completely flat, say $H\p{E}=H_0$ for all $E$, then the weights
-$w\p{E}$ have all been modified by the same factor of $f^{H_0}$.
-Remembering the scale-invariance of $w\p{E}$, this result would mean
-that the weights have not actually changed at all! Similarly, too lax
-of a flatness condition will prompt the initialization process to
-\begin{enumerate*}[label=\roman*)]
-\item reset and decrease $f$ too early, i.e. when the weight array
-  could still use some heavy handed adjustment, and
-\item quit before one has any reason to think that simulations might
-  yield a flat histogram.
-\end{enumerate*}
-Tuning the flatness condition is thus crucial to the success of the
-Wang-Landau method.
-
 \begin{algorithm}[!b]
@@
 necessary. The value of $n$ is not too important, so long as it is
-reasonable. Too small value of $n$ will waste computation time
-determining $C\sp{H}$, while too large of a value will waste time
-simulating when $C\sp{H}$ has already been satisfied. In general, $n$
-should be proportional to the computation time of $C\sp{H}$, which
-scales with energy range of the simulated system. We therefore use
-$n=L$, where $L$ is the number of energy levels for the given
-square-well fluid (as discussed in Section \ref{sec:simple_flat}).
-Papers in the literature typically neglect providing any heuristic or
-formula for $n$, reporting a system-independent $n=10^3$ or
-$10^5$\cite{wang_landau, wang_landau_mod}.
+reasonable for the system at hand. Too small value of $n$ will waste
+computation time determining $C\sp{H}$, while too large of a value
+will waste time simulating when $C\sp{H}$ has already been
+satisfied. In general, $n$ should be proportional to the computation
+time of $C\sp{H}$, which scales with energy range of the simulated
+system. We therefore use $n=L$, where $L$ is, as described in Section
+\ref{sec:simple_flat}, the number of energy levels for the given
+square-well fluid. Other works typically neglect providing any
+heuristic or formula for $n$, reporting a system-independent
+$n=10^3$\cite{wang_landau_mod} or $10^5$\cite{wang_landau}.
+\fixme{find more citations for this last claim}
 
@@
 simulations to take an exceedingly long time to converge. Lacking any
-heuristics for assigning $f_0$, literature typically suggests
-$f_0=e$\cite{wang_landau}; after some experimentation with the free
-parameters in the Wang-Landau method, we used $f_0=e^{1/8}$ (in
+heuristics for assigning $f_0$, literature typically suggests $f_0=e$
+\cite{wang_landau}. After some experimentation with the free
+parameters in the Wang-Landau method, we decided on $f_0=e^{1/8}$. In
 practice, one usually works with $\ln f$, making the commonly
-suggested value $\ln f_0=1$ and our own $\ln f_0=1/8$).
+suggested value $\ln f_0=1$ and our own $\ln f_0=1/8$.
 
@@
 at each increment of $k$ is a natural and convenient iterative means
-to satisfy these requirements. Previous papers have used $u=2$, which
-we have adopted as well\cite{wang_landau, wang_landau_mod}.
-
-\fixme{Walk through $C\sp{H}$ and $c$. Finish discussion of WL.}
-
-\subsection{The optimized ensemble (OE) method}
-\label{sec:optimized_ensemble}
-
-\fixme{Explain, provide algorithm.}
-
-\begin{algorithm}[H]
-  \caption{Modifying weights via the optimized ensemble}
-  \label{alg:optimized_ensemble}
-\end{algorithm}
-
-\subsection{The transition matrix Monte Carlo (TMMC) method}
+to satisfy these requirements. Previous works have reported using
+values such as $u=2$ \cite{wang_landau, wang_landau_mod} and $u=10$
+\cite{wang_landau_analysis}; we use the former, conservatively
+favoring more initialization iterations by decreasing $f$ by a smaller
+amount on each reset.
+
+One might guess why the flatness condition $C\sp{H}$ for updating $f$
+and restarting the Wang-Landau initialization process should not be
+made too lax: doing so would prompt the initialization process to
+\begin{enumerate*}[label=\roman*)]
+\item reset and decrease $f$ too early, i.e. when the weight array
+  could still use some heavy handed adjustment, and
+\item quit before one has any reason to think that simulations might
+  yield a flat histogram.
+\end{enumerate*}
+Less obvious, however, is the fact that the flatness condition
+$C\sp{H}$ should not be made too stringent: if the histogram is
+completely flat, say $H\p{E}=H_0$ for all $E$, then the weights
+$w\p{E}$ have all been modified by the same factor of $f^{H_0}$.
+Remembering the scale-invariance of $w\p{E}$, this result would mean
+that the weights have not actually changed at all!
+
+The success of Wang-Landau thus relies on some degree of leniency in
+the flatness condition, which makes tuning $C\sp{H}$ crucial to the
+efficient and effective implementation of this method. Some previous
+works\cite{wang_landau} suggest that $C\sp{H}$ should be made as
+stringent as possible, cautioning only that some simulations might not
+satisfy too strict of a flatness condition in any reasonable amount of
+time. These works fail to recognize the reliance of Wang-Landau on the
+{\it failure} of $C\sp{H}$ to enforce a perfectly flat histogram. A
+commonly suggested flatness condition is that the minimum histogram
+value be at least some fixed proportion of the mean histogram value,
+i.e. $C\sp{H}:\min\sp{H\p{E}}\ge x\bk{H}_E$, and several papers report
+using $x=0.95$. We used the same condition with $x=0.25$.
+
+The final free parameter in the Wang-Landau method is the cutoff $c$,
+which determines how small $f$ can get before the algorithm stops
+modifying the weight array. Too large of a value for $c$ will make
+Wang-Landau quit prematurely, whereas too small a value will cause the
+algorithm to waste time simulating after the weights $w\p{E}$ have
+essentially converged. One might imagine using an alternate end
+condition for Wang-Landau, which involves comparing the current
+weights $w_k\p{E}$ to those at the end of the previous cycle,
+$w_{k-1}\p{E}$, to check whether the initialization process is still
+making appreciable modifications to the weights, but such a check
+would require both a metric and a free parameter to define what an
+appreciable change is. For simplicity, we stuck with the standard end
+condition given in Algorithm \ref{alg:wang_landau}. As with the other
+free parameters in Wang-Landau, we are not aware of any heuristics for
+determining an appropriate value of $c$, but literature cites typical
+values of $\ln c=10^{-8}$; we used $\ln c=10^{-6}$.
+
+\subsection{The transition matrix Monte Carlo method}
 \label{sec:tmmc}
+
+Unlike the previous histogram methods, the transition matrix Monte
+Carlo (TMMC) method does not use weights, and merely compute the
+density of states $D\p{E}$ of a system. The density of states can in
+turn be used via the flat histogram method (Section
+\ref{sec:flat_histogram}) to determine appropriate weights for
+subsequent Monte Carlo simulations. TMMC introduces a new object: the
+energy transition matrix, $T$, whose components $T_{ij}$ are the
+probabilities that a system will transition {\it to} a state with
+energy $E_i$ {\it from} a given state with energy $E_j$,
+i.e. $T_{ij}=P\p{E_j\to E_i}$.
+
+The transition matrix $T$ is trivially square and positive-definite,
+but it is not symmetric. Given that Metropolis Monte Carlo samples all
+of state-space randomly, meaning all possible system states are
+equally likely to occur during simulation, a transition $E_i\to E_f$
+from an energy with a low state density to one with a high state
+density, $D\p{E_i}<D\p{E_j}$, is more probable than the inverse
+transition $E_j\to E_i$ simply due to the fact that there are more
+states with energy $E_i$ than those with energy $E_j$. Thus
+$D\p{E_i}<D\p{E_j}$ implies $P\p{E_i\to E_j}>P\p{E_j\to E_i}$, which
+means $T_{ji}>T_{ij}$. Said otherwise, the asymmetry of the transition
+matrix is a direct consequence of the second law of thermodynamics:
+systems are exceedingly likely to evolve towards macroscopic states
+with higher accessible microstate densities.
+
+In principle, computationally determining the transition matrix of a
+system is straightforward. One need only simulate a system and collect
+a histogram $\tilde T\p{E_i,E_j}$ of the energy transitions after each
+move, directly after deciding whether to reject the move based on
+whether it resulted in an (in)valid system state. That is, transition
+data should be collected (only) on all moves as they would have been
+accepted by Metropolis Monte Carlo, i.e. before deciding whether to
+reject any given move for any reasons involving weights, or other
+biasing schemes.
+
+\vspace{3cm}
+
+\fixme{schemes for initializing to collect transition matrix}
+
+\fixme{normalizing transition matrix}
+
+\fixme{computing density of states via transition matrix}
+
+\fixme{storing transition matrix in the form $T\p{E,\Delta E}$}
+
+\vspace{3cm}
+
+For the square-well fluid, as with many systems, there is a maximum
+amount $\Delta_ME$ by which a single move can change the energy of the
+system. The transition matrix can therefore be stored in the form
+$T\p{E,\Delta E}$
+
+the transition matrix is band-diagonal because . In implementation,
+the transition matrix can therefore be stored in the form
+$T\p{E,\Delta E}$, making it an $L\times\Delta_ME$ matrix, where $L$
+is the number of energy levels in a given system. Given that $L$
+scales as $N\Delta_ME$, where $N$ is the number of spheres in the
+square-well fluid, the memory required to store the transition matrix
+therefore scales as $N\p{\Delta_ME}^2$, an improvement over the $N^2$
+growth of a transition matrix in the form $T\p{E_i,E_j}$.
+
+\newpage
 
@@
 
+\subsection{The optimized ensemble method}
+\label{sec:optimized_ensemble}
+
+\fixme{Explain briefly. No need to provide algorithm.}
+
 \subsection{The hybrid OE-TMMC method}
@@
 
-\fixme{Explain, provide algorithm. This method is sort of our own, but
-  based on other work.}
+\fixme{Explain, provide algorithm. This method is our own synthesis of
+  other work.}
 
@@
 \fixme{This section might just become a part of the section on the
-  minimum important energy. (\ref{eq:density_of_states}) gives us an
-  explicit formula for the density of states if we are confident in
-  our energy histogram. We can also compute the density of states from
-  the transition matrix, which is sometimes more appropriate to do.}
+  minimum important energy. (\ref{eq:dos}) gives us an explicit
+  formula for the density of states if we are confident in our energy
+  histogram. We can also compute the density of states from the
+  transition matrix, which is sometimes more appropriate to do.}
 

### 2015-05-04 papers/histogram/thesis/thesis.tex
added tmmc and optimized ensemble to metheds section

@@
 \newcommand{\department}{Department of Physics}
-\newcommand{\supervisor}{Dr. David Roundy} \renewcommand{\date}{21
-  April 2015}
+\newcommand{\supervisor}{Dr. David Roundy}
+\renewcommand{\date}{08 May 2015}
 
@@
   \caption{Metropolis Monte Carlo fluid simulation}
-  \label{alg:MC}
+  \label{alg:metropolis}
   \begin{alg}
@@
     change otherwise. Attempting to move a single sphere is referred
-    to as a move. \label{alg:mc_move}
-
-  \item Repeat step \ref{alg:mc_move} for every other sphere in the
-    fluid. Attempting to move every sphere once is referred to as an
-    iteration of the simulation.
-    \label{alg:mc_iteration}
-
-  \item Repeat step \ref{alg:mc_iteration} indefinitely, or until data
-    of sufficient quality has been generated, periodically collecting
-    and dumping statistics on fluid states (e.g. energy, pair
-    distribution histograms, etc.) to data files.
+    to as a move. \label{alg:metropolis_move}
+
+  \item Repeat step \ref{alg:metropolis_move} for every other sphere
+    in the fluid. Attempting to move every sphere once is referred to
+    as an iteration of the simulation.
+    \label{alg:metropolis_iteration}
+
+  \item Repeat step \ref{alg:metropolis_iteration} indefinitely, or
+    until data of sufficient quality has been generated, periodically
+    collecting and dumping statistics on fluid states (e.g. energy,
+    pair distribution histograms, etc.) to data files.
 
@@
 
-Algorithm \ref{alg:MC} provides a sketch of unbiased, Metropolis Monte
-Carlo fluid simulations. Such an algorithm is ``unbiased'' in the
-sense that it collects statistics (i.e. data) on all valid system
-configurations with equal probability. Statistics whose collection
-time scales as $\mathcal O\p{1}$ with system size, meaning that
-increasing the number of spheres $N$ in the simulation does not affect
-collection time, can be collected after every move (defined in the
-algorithm). Statistics whose collection time scales as $\mathcal
-O\p{N}$, meaning that doubling $N$ doubles collection time, can be
-collected after each iteration. In general, collection with $\mathcal
-O\sp{\chi\p{N}}$ time scaling should not occur more often than once
-every $\chi\p{N}$ moves (where $\chi\p{N}$ may be an arbitrary
-function of $N$ void of constant prefactors, e.g. $N\log N$, or
-$2^N$). These collection rules ensure that scaling up simulations does
-not cause them to asymptotically spend all computation time collecting
-statistics, and no time actually simulating the fluid. Collected
+Algorithm \ref{alg:metropolis} provides a sketch of unbiased,
+Metropolis Monte Carlo fluid simulations. Such an algorithm is
+``unbiased'' in the sense that it collects statistics (i.e. data) on
+all valid system configurations with equal probability. Statistics
+whose collection time scales as $\mathcal O\p{1}$ with system size,
+meaning that increasing the number of spheres $N$ in the simulation
+does not affect collection time, can be collected after every move
+(defined in the algorithm). Statistics whose collection time scales as
+$\mathcal O\p{N}$, meaning that doubling $N$ doubles collection time,
+can be collected after each iteration. In general, collection with
+$\mathcal O\sp{\chi\p{N}}$ time scaling should not occur more often
+than once every $\chi\p{N}$ moves, where $\chi\p{N}$ may be an
+arbitrary function of $N$ void of constant prefactors, e.g. $N\log N$,
+or $2^N$. These collection rules ensure that scaling up simulations
+does not cause them to asymptotically spend all computation time only
+collecting statistics, or only simulating the fluid. Collected
 statistics are used to find thermodynamic properties of the simulated
@@
 energy $E$, i.e. $\bk{X}_E$. In the real world, however, information
-about a system's microstates, and by extension functions of
-microstates (e.g. the energy $E$), is inaccessible. One therefore
-cannot measure $\bk{X}_E$ directly. Instead, one typically measures
-the dependence of thermodynamic properties on macroscopic state
-variables such as temperature, i.e. $\bk{X}_T$.
+about a system's microstates, and by extension information about
+functions of microstates (e.g. the energy $E$), is inaccessible. One
+therefore cannot measure $\bk{X}_E$ directly. Instead, one typically
+measures the dependence of thermodynamic properties on macroscopic
+state variables such as temperature, i.e. $\bk{X}_T$.
 
@@
 range of possible $X$ is so low that it is practically impossible (via
-Metropolis Monte Carlo) to sufficiently sample $R$, that is,
+Metropolis Monte Carlo) to sufficiently sample $R$, that is, to
 accumulate a statistically significant histogram $H\p{X\in R}$, in any
-reasonable amount of time.
+reasonable amount of time. Given that the entropy $S\p{X}$ can be
+expressed in terms of the number of microstates $\Omega\p{X}$ as
+$S\p{X}=k\ln\Omega\p{X}\propto\ln D\p{X}$, we may refer to regions $R$
+low state densities $D\p{X\in R}$ as low entropy states.
 
@@
 Histogram methods provide a means to address Metropolis Monte Carlo's
-inability to sufficiently sample regions of parameter space with a low
-density of states by introducing a bias into the otherwise random
-sampling of state space. A weighting function $w$ is introduced, whose
-domain is the value of some property $X\p{s}$ which depends on the
-system state $s$. An additional condition is then added to step
-\ref{alg:mc_move} of Algorithm \ref{alg:MC} in order to accept an
-attempted move: the weights $w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of
-the initial (pre-move) state $s_i$ and final (post-move) state $s_f$
-are used to determine the probability $P\p{s_i\to s_f}$ of accepting
-an otherwise valid move via
-\begin{align}
-  P\p{s_i\to s_f}=\min\set{\f{w\sp{X\p{s_f}}}{w\sp{X\p{s_i}}},1}.
+inability to sufficiently sample low entropy states by introducing a
+bias into the otherwise random sampling of state space. A weighting
+function $w$ is introduced, whose domain is the value of some property
+$X\p{s}$ which depends on the system state $s$. An additional
+condition is then added to step \ref{alg:metropolis_move} of Algorithm
+\ref{alg:metropolis} in order to accept an attempted move: the weights
+$w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of the initial (pre-move) state
+$s_i$ and final (post-move) state $s_f$ are used to determine the
+probability $P_m\p{s_i\to s_f}$ of accepting an otherwise valid move
+via
+\begin{align}
+  P_m\p{s_i\to s_f}=\min\set{\f{w\sp{X\p{s_f}}}{w\sp{X\p{s_i}}},1}.
   \label{eq:move_prob}
@@
 move. Due to the fact that only ratios of weights determine
-$P\p{s_i\to s_f}$, the weights $w\sp{X\p{s}}$ are scale-invariant,
+$P_m\p{s_i\to s_f}$, the weights $w\sp{X\p{s}}$ are scale-invariant,
 meaning that their effect on simulations is unchanged by scale factors
@@
 
-  \item Repeat step \ref{alg:biased_mc_move} for every other sphere in
-    the fluid.
-    \label{alg:biased_mc_iteration}
-
-  \item Repeat step \ref{alg:biased_mc_iteration} until data of
-    sufficient quality has been generated.
+  \item Repeat step \ref{alg:biased_mc_move} the simulation produces
+    statistics of sufficient quality.
 
@@
 In this paper, we will consider weights $w\p{E}$ which depend only on
-the energy $E$ of the system. Due to the fact that the square-well
-fluid can only have discrete energies $E=-n\epsilon$, where $n$ is a
-non-negative integer and $\epsilon$ is the well depth, in simulation
-we store the weight function as an array of values. We may therefore
-use the terms ``weight function,'' and ``weight array,'' and
-``weights'' interchangeably.
+the energy $E$ of the square-well fluid. Due to the fact that the
+square-well fluid can only have discrete energies $E=-n\epsilon$,
+where $n$ is a non-negative integer and $\epsilon$ is the well depth,
+in simulation we store the weight function as an array of values. We
+may therefore use the terms ``weight function,'' and ``weight array,''
+and ``weights'' interchangeably.
 
@@
 but all simulations in this paper require a choice of minimum
-temperature anyways (addressed in Section \ref{sec:energy_range}). In
+temperature anyways (discussed in Section \ref{sec:energy_range}). In
 this paper, we will always use $kT_{\t{min}}=0.2\epsilon$.
@@
 
+\fixme{add citations}
+
 Unlike the previous histogram methods, the transition matrix Monte
 Carlo (TMMC) method does not use weights, and merely compute the
-density of states $D\p{E}$ of a system. The density of states can in
-turn be used via the flat histogram method (Section
-\ref{sec:flat_histogram}) to determine appropriate weights for
-subsequent Monte Carlo simulations. TMMC introduces a new object: the
-energy transition matrix, $T$, whose components $T_{ij}$ are the
+density of states $D\p{E}$ of a system\cite{tmmc}. The density of
+states can in turn be used to determine flat histogram weights via
+(\ref{eq:flat_weights}). TMMC introduces a new object: the energy
+transition matrix, $T$, whose components $T_{ij}$ are the
 probabilities that a system will transition {\it to} a state with
@@
 
-The transition matrix $T$ is trivially square and positive-definite,
-but it is not symmetric. Given that Metropolis Monte Carlo samples all
-of state-space randomly, meaning all possible system states are
-equally likely to occur during simulation, a transition $E_i\to E_f$
-from an energy with a low state density to one with a high state
-density, $D\p{E_i}<D\p{E_j}$, is more probable than the inverse
-transition $E_j\to E_i$ simply due to the fact that there are more
-states with energy $E_i$ than those with energy $E_j$. Thus
-$D\p{E_i}<D\p{E_j}$ implies $P\p{E_i\to E_j}>P\p{E_j\to E_i}$, which
-means $T_{ji}>T_{ij}$. Said otherwise, the asymmetry of the transition
-matrix is a direct consequence of the second law of thermodynamics:
-systems are exceedingly likely to evolve towards macroscopic states
-with higher accessible microstate densities.
-
-In principle, computationally determining the transition matrix of a
-system is straightforward. One need only simulate a system and collect
-a histogram $\tilde T\p{E_i,E_j}$ of the energy transitions after each
-move, directly after deciding whether to reject the move based on
-whether it resulted in an (in)valid system state. That is, transition
-data should be collected (only) on all moves as they would have been
-accepted by Metropolis Monte Carlo, i.e. before deciding whether to
-reject any given move for any reasons involving weights, or other
-biasing schemes.
-
-\vspace{3cm}
-
-\fixme{schemes for initializing to collect transition matrix}
-
-\fixme{normalizing transition matrix}
-
-\fixme{computing density of states via transition matrix}
-
-\fixme{storing transition matrix in the form $T\p{E,\Delta E}$}
-
-\vspace{3cm}
-
-For the square-well fluid, as with many systems, there is a maximum
-amount $\Delta_ME$ by which a single move can change the energy of the
-system. The transition matrix can therefore be stored in the form
-$T\p{E,\Delta E}$
-
-the transition matrix is band-diagonal because . In implementation,
-the transition matrix can therefore be stored in the form
-$T\p{E,\Delta E}$, making it an $L\times\Delta_ME$ matrix, where $L$
-is the number of energy levels in a given system. Given that $L$
-scales as $N\Delta_ME$, where $N$ is the number of spheres in the
-square-well fluid, the memory required to store the transition matrix
-therefore scales as $N\p{\Delta_ME}^2$, an improvement over the $N^2$
-growth of a transition matrix in the form $T\p{E_i,E_j}$.
-
-\newpage
-
-\fixme{Explain, provide algorithm. This method is sort of our own, but
-  based on other work.}
-
-\begin{algorithm}[H]
+The transition matrix is trivially square and positive-definite, but
+it is not symmetric. Given that Metropolis Monte Carlo samples all of
+state-space randomly, meaning all possible system states are equally
+likely to occur during simulation, a transition $E_i\to E_f$ from an
+energy with a low state density to one with a high state density,
+$D\p{E_i}<D\p{E_j}$, is more probable than the inverse transition
+$E_j\to E_i$ simply due to the fact that there are more states with
+energy $E_i$ than those with energy $E_j$. Thus $D\p{E_i}<D\p{E_j}$
+implies $P\p{E_i\to E_j}>P\p{E_j\to E_i}$, which means
+$T_{ji}>T_{ij}$. The asymmetry of the transition matrix is precisely
+the origin of the second law of thermodynamics: systems are
+exceedingly likely to evolve towards macroscopic states with higher
+accessible microstate densities.
+
+Knowing the transition matrix of a system determines, among other
+properties, the system's density of states. Consider an ensemble of
+Metropolis Monte Carlo simulations with a distribution of energies at
+a time $t$ given by $\hat D^{t}\p{E}$, such that the probability of
+any given simulation to have an energy $E_i$ at a time $t$ is $\hat
+D_i^{t}=\hat D^{t}\p{E_i}$. The probability $\hat D_i^{t}$ can be
+expressed in terms of the distribution $\hat D^{t-1}\p{E}$ one time
+step (i.e. one Metropolis Monte Carlo move) prior by
+\begin{align}
+  \hat D_i^{t}=\sum_jP\p{E_i\to E_j}\hat D_j^{t-1} =\sum_j T_{ij}\hat
+  D_j^{t-1}.
+  \label{eq:transition_evolution}
+\end{align}
+At equilibrium, the probability distributions $\hat D^{t}=\hat
+D^{t-1}$. Furthermore, the distribution of states in an equilibrium
+ensemble of Metropolis Monte Carlo simulations is precisely the
+density of states $D\p{E}$\fixme{explain why?}, which means
+$D_i=T_{ij}D_j$. Simultaneously expressing this condition for all
+energies,
+\begin{align}
+  D=TD. \label{eq:dos_eigen}
+\end{align}
+Finding the density of states of a system can thus be reduced to
+finding the eigenvector of the transition matrix which has a
+corresponding unit eigenvalue\fixme{why is this eigenvector
+  unique?}. Once one has determined the density of states of a system,
+one can initialize weights via the flat histogram method,
+i.e. (\ref{eq:flat_weights}).
+
+Computationally determining the transition matrix of a system is
+straightforward: one need only simulate the system and collect a
+histogram $\tilde T_{ij}=\tilde T\p{E_i,E_j}$ of the energy
+transitions after each Metropolis Monte Carlo move, i.e. after step
+\ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis}. A
+histogram of transitions $\tilde T\p{E_i,\bar E_j}_{}$ from a fixed
+energy $\bar E_j$ to energies $E_i$ is then proportional to the
+probability distribution of transitions from $\bar E_j$, i.e. the
+transition matrix $T\p{E_i,\bar E_j}$. Crucially, this proportionality
+does not depend on the history of a simulation, or on how the
+simulation got to be in the energy $\bar E_j$, so long as samples of
+the energy $\bar E_j$ are themselves void of systemic
+bias. Furthermore, one does not actually need to transition from $E_j$
+to $E_i$ in order to collect statistics on $T\p{E_i,E_j}$: after
+deciding whether to accept a move based on whether it results in a
+valid system state, one can increment $\tilde T\p{E_i,E_j}$ before
+deciding to reject the move for some other reason, e.g. because
+$w\p{E_j}\gg w\p{E_i}$, or even because one wishes to collect more
+statistics on $T\p{E_i,\bar E_j}$.
+
+The proportionality $\tilde T_{ij}\propto T_{ij}$ can be made explicit
+by enforcing that the probabilities of all transitions from an energy
+$E_j$ sum to unity, i.e.
+\begin{align}
+  \sum_iP\p{E_j\to E_i}=\sum_iT_{ij}=1,
+  \label{eq:transition_norm_condition}
+\end{align}
+which implies
+\begin{align}
+  T_{ij}=\f{\tilde T_{ij}}{\sum_i\tilde T_{ij}}.
+  \label{eq:transition_normalization}
+\end{align}
+To compute $T$ from $\tilde T$, one must therefore normalize each row
+of $\tilde T$ independently.
+
+If a system has $L$ energy levels, the transition histogram $\tilde
+T\p{E_i,E_j}$ is an $L\times L$ matrix. Given that $L$ is proportional
+to the system size $N$, the memory footprint of $\tilde T\p{E_i,E_j}$
+grows as $N^2$. For most systems, however, there will be a maximum
+energy $M$ independent of system size by which a single move can
+change the energy of the system. The transition histogram for such
+systems is therefore band-diagonal, which one can exploit to save
+memory. In particular, it is natural to store the transition histogram
+in the form $\tilde T_d\p{E,\Delta E}$, which gives the probability
+that a system will transition from the energy $E$ to the energy
+$E+\Delta E$ in a single move. One can convert between $\tilde
+T_d\p{E,\Delta E}$ and $\tilde T\p{E_i,E_j}$ using the relation
+\begin{align}
+  \tilde T\p{E_i,E_j}=\tilde T_d\p{E_j,E_i-E_i},
+\end{align}
+which means, from (\ref{eq:transition_normalization}),
+\begin{align}
+  T_{ij}=\f{\tilde T_d\p{E_j,E_i-E_j}}{\sum_i\tilde T_d\p{E_j,E_i-E_j}}
+  =\f{\tilde T_d\p{E_j,E_i-E_j}}
+  {\sum_{\Delta E}\tilde T_d\p{E_j,\Delta E}}.
+  \label{eq:transition_conversion}
+\end{align}
+Expressed in this form, the memory footprint of the transition
+histogram grows linearly with system size $N$ and the maximal energy
+difference $M$.
+
+While one could imagine many initialization routines to determine the
+transition matrix $T$, we use Algorithm \ref{alg:tmmc}, which is a
+modified version of a routine provided in \cite{tmmc_prl}\fixme{is
+  this citation correct?}. The primary motivation behind this
+algorithm is such: as lower energies are always more difficult to
+sample than higher energies, transitions to lower energies should
+always be accepted, while transitions to higher energies should pass a
+probabilistic acceptance test. To understand the probability in step
+\ref{alg:tmmc_prob} of Algorithm \ref{alg:tmmc} of accepting a move
+$s_i\to s_f$ for which $E_f>E_i$, first consider the case of $c=0$ and
+$T_{\t{min}}\to0$, in which case the probability becomes
+\begin{align}
+  P_m\p{s_i\to s_f}=\f{\p{\tilde T_d\p{E_i,\Delta E}}\big/
+    \p{\sum_{\Delta E'}\tilde T_d\p{E_i,\Delta E'}}}
+  {\p{\tilde T_d\p{E_f,-\Delta E}}\big/
+    \p{\sum_{\Delta E'}\tilde T_d\p{E_f,\Delta E'}}}
+  =\f{T\p{E_i,\Delta E}}{T\p{E_f,-\Delta E}}
+  =\f{P\p{E_i\to E_f}}{P\p{E_f\to E_i}},
+  \label{eq:tmmc_prob_simplified}
+\end{align}
+where $\Delta E=E_f-E_i$ and $P\p{E_m\to E_n}$ is the unbiased
+probability of a transition from an energy $E_m$ to an energy
+$E_n$. The result in (\ref{eq:tmmc_prob_simplified}) is the
+probability of accepting moves $s_i\to s_f$ required for simulation
+via Algorithm \ref{alg:tmmc} to yield a flat energy
+histogram\cite{tmmc}. Unlike the previous histogram methods, we desire
+a flat histogram during initialization via Algorithm \ref{alg:tmmc}
+not because we are constructing weights $w\p{E}$, but because we wish
+to sample the transition histogram $\tilde T_d\p{E,\Delta E}$ equally
+at all energies $E$. \fixme{briefly explain
+  (\ref{eq:tmmc_prob_simplified}) by considering a two-level system,
+  and extending to larger systems}.
+
+Aside from a minimum temperature $T_{\t{min}}$ and an end condition,
+Algorithm \ref{alg:tmmc} contains one free parameter, $c$, used in
+step \ref{alg:tmmc_prob} to modify (\ref{eq:tmmc_prob_simplified}).
+This parameter makes the rejection of transitions more conservative
+when there are too few statistics in the histogram $\tilde T_d$ to be
+confident in the current estimate of the transition matrix. In order
+to make the effects of $c$ negligible as $\tilde T_d$ collects more
+statistics throughout initialization, $c$ should be of order unity,
+though its optimal value is generally system-dependent. In our
+implementation of Algorithm \ref{alg:tmmc}, we use $c=16$.
+
+The last interesting part of the probability in step
+\ref{alg:tmmc_prob} of Algorithm \ref{alg:tmmc} is the limit on
+$P_m\p{s_i\to s_f}$ to a minimum value of $\exp\p{-\Delta
+  E/kT_{\t{min}}}$. This limit caps the bias on energies to that
+introduced by canonical (fixed-temperature) weights at a temperature
+of $T_{\t{min}}$. As a result, Algorithm \ref{alg:tmmc} will not waste
+time oversampling energies which are unimportant for determining
+system properties at temperatures $T\ge T_{\t{min}}$.
+
+\begin{algorithm}[tb]
   \caption{Transition matrix Monte Carlo initialization}
   \label{alg:tmmc}
+  \begin{alg}
+
+  \item Construct an initial typical fluid configuration.
+
+  \item Initialize a transition histogram $\tilde T_d\p{E,\Delta E}=0$
+    for all $E$ and $\Delta E\in\sp{-M,M}$, where $M$ is the possible
+    maximum energy transition of a single move.
+
+  \item Randomly attempt to change the position of one sphere,
+    tentatively accepting the transition from state $s_i$ to state
+    $s_f$ if $s_f$ is not a forbidden fluid configuration.
+    \label{alg:tmmc_move}
+
+  \item Let $\Delta E=E_f-E_i$ and increment $\tilde
+    T_d\p{E_i,\Delta{E}}$, where $E_k$ is the energy of state $s_k$.
+
+  \item If $\Delta E<0$, accept the move. Otherwise
+    \begin{enumerate}
+    \item compute the normalization factor
+      \begin{align*}
+        n_k=\sum_{\Delta E'}\tilde T_d\p{E_k,\Delta E'}
+      \end{align*}
+      for $k\in\set{i,f}$, and
+
+    \item accept the move $s_i\to s_f$ with probability
+      \begin{align*}
+        P_m\p{s_i\to s_f}
+        =\max\set{\f{\p{\tilde T_d\p{E_i,\Delta E}+c}/\p{n_i+c}}
+          {\p{\tilde T_d\p{E_f,-\Delta E}+c}/\p{n_f+c}},
+          \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
+      \end{align*}
+      for a predetermined number $c$ and temperature $T_{\t{min}}$.
+      \label{alg:tmmc_prob}
+    \end{enumerate}
+
+  \item If some predetermined initialization end condition has
+    \emph{not} been met, return to step \ref{alg:tmmc_move}.
+
+  \end{alg}
 \end{algorithm}
@@
 
-\fixme{Explain briefly. No need to provide algorithm.}
+All histogram methods introduced in this paper thus far have focused
+on determining a weight array $w\p{E}$ for a flat energy histogram
+$H\p{E}$, in order to sample and collect statistics on all energies
+equally. This motivation considers the quantity of statistics, but not
+their quality. Biased simulations collect statistics on low energy
+states by first getting into these states, and then rejecting moves
+which move the system into higher energy states. Statistics on low
+energy states will therefore generally be highly correlated, as they
+are based on many samples of only a few low energy regions of a
+system's energy landscape. The optimized ensemble
+(OE)\cite{optimized_ensemble} attempts to address the autocorrelation
+between low energy samples by finding weights to maximize the rate at
+which a simulation makes round trips between low and high energy
+states. Maximizing the round trip rate, in turn, maximizes the rate at
+which a simulation makes independent, uncorrelated samples of low
+energies.
+
+The optimized ensemble works by considering an ensemble of
+simulations, each of which defines the position of a ``walker'' in
+energy space. Given that we want walkers to move back and forth
+between two extrema $E_+$ and $E_-$ of some specified energy range, we
+label each walker by which extrema it has visited most recently. We
+can then identify ``down-going'' (``up-going'') walkers by those which
+more recently visited $E_+$ ($E_-$), which means that to make a round
+trip they first needed to get to $E_-$ ($E_+$). Denoting the total
+walker and down-going density at an energy $E$ respectively by
+$n\p{E}$ and $n_+\p{E}$, the ratio $f\p{E}=n_+\p{E}/n\p{E}$ gives the
+proportion of walkers at $E$ which are down-going. We also denote the
+net and down-going walker diffusivity at an energy $E$ respectively by
+$\alpha\p{E}$ and $\alpha_+\p{E}$.
+
+We can meaningfully discuss a down-going walker flux $j\p{E}$, which
+is defined by
+\begin{align}
+  j=\f{dn_+}{dt}=\alpha_+\f{dn_+}{dE}=f\alpha\f{dn_+}{dE}.
+  \label{eq:walker_current_definition}
+\end{align}
+Computing
+\begin{align}
+  n\f{df}{dE}=n\sp{\f1{n}\f{dn_+}{dE}-\f{n_+}{n^2}\f{dn}{dE}}
+  =\f{dn_+}{dE}-f\f{dn}{dE},
+  \label{eq:walker_intermediate}
+\end{align}
+we can say that
+\begin{align}
+  \f{dn_+}{dE}=n\f{df}{dE}+f\f{dn}{dE}.
+  \label{eq:walker_gradient}
+\end{align}
+Substituting (\ref{eq:walker_gradient}) into
+(\ref{eq:walker_current_definition}) yields
+\begin{align}
+  j=f\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
+  \label{eq:walker_current_full}
+\end{align}
+If the energy histogram does vary strongly with energy, e.g. because
+we are simulating with flat histogram weights, then $dn/dE\approx0$,
+which means
+\begin{align}
+  j=f\alpha n\f{df}{dE}.
+  \label{eq:walker_current}
+\end{align}
+To formalize the approximation made between
+(\ref{eq:walker_current_full}) and (\ref{eq:walker_current}), we can
+observe that
+\begin{align}
+  \f{f~dn/dE}{n~df/dE}
+  =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}=\p{\f1f\f{dn_+}{dn}-1}^{-1}
+  =\p{\f{d\ln n_+}{d\ln n}-1}^{-1}.
+  \label{eq:walker_current_intermediate_ratio}
+\end{align}
+For (\ref{eq:walker_current}) to be valid, we want the ratio in
+(\ref{eq:walker_current_intermediate_ratio}) to vanish. If the walker
+density $n$ varies weakly with energy, e.g. because we are simulating
+with flat histogram weights, then $\ln n$ varies even weaker
+still. The down-going walker density $n_+$, meanwhile must run from
+$n\p{E_-}=0$ to $n\p{E_+}=n\p{E_+}$ meaning $\ln n_+$ runs from $\ln
+n\p{E_-}\to-\infty$ to a finite positive value $\ln n_+\p{E_+}=\ln
+n\p{E_+}$. It is therefore reasonable to assume that $d\ln n_+/d\ln
+n\gg 1$, which makes the ratio in
+\ref{eq:walker_current_intermediate_ratio} vanish.
+
+We can rearrange (\ref{eq:walker_current}) and integrate from the low
+energy state $E_-$, at which $f\p{E_-}=0$, to the high energy state
+$E_+$, at which $f\p{E_+}=1$, to say
+\begin{align}
+  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \label{eq:walker_current_setup}
+\end{align}
+At equilibrium, $j=dn_+/dt$ is constant, which means we can evaluate
+one of the integrals in (\ref{eq:walker_current_setup})
+\begin{align}
+  \f1j=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \label{eq:walker_current_integral}
+\end{align}
+
+The optimized ensemble method minimizes the integral in
+(\ref{eq:walker_current_integral}), thereby maximizing the down-going
+walker current $j$, by varying the walker density $n$ with the
+constraint that $n$ must remain normalized. With the additional
+assumption that $\alpha\p{E}$ does not strongly depend on the weights
+$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
+optimal walker density $n$ is
+\begin{align}
+  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
+  \label{eq:optimal_walker_density}
+\end{align}
+where $\lambda$ is the Lagrange multiplier used in the minimization
+process, and here acts merely as a normalization factor.
+
+Equating the set of observations in a single simulation with a single
+observation of an ensemble of simulations, the walker density $n\p{E}$
+is proportional to the energy histogram $H\p{E}=\tilde D\p{E}w\p{E}$,
+which means
+\begin{align}
+  w_{\t{opt}}=\f1{D\sqrt{\alpha}},
+  \label{eq:optimized_weights}
+\end{align}
+where we neglect the distinction between $\tilde D$ and $D$, which has
+no effect on the weights $w$.
+
+The algorithm provided in \cite{optimized_ensemble} to determine and
+construct weights given by \ref{eq:optimized_weights} involves
+simulating with and iteratively modifying flat histogram weights. This
+algorithm is not meant to construct weights from scratch, but rather
+to optimize them. We therefore will not be comparing the optimized
+ensemble to other methods. We do, however, credit this method for
+motivating the hybrid OE-TMMC method in Section \ref{sec:oetmmc}.
 
@@
 \begin{algorithm}[H]
-  \caption{Hybrid OE-TMMC simulation}
+  \caption{Modifying transition weights via the optimized ensemble}
   \label{alg:oetmmc}

### 2015-05-04 papers/histogram/thesis/thesis.tex
fixed error in optimized ensemble section

@@
 We can meaningfully discuss a down-going walker flux $j\p{E}$, which
-is defined by
-\begin{align}
-  j=\f{dn_+}{dt}=\alpha_+\f{dn_+}{dE}=f\alpha\f{dn_+}{dE}.
+is defined in terms of the diffusivity $\alpha_+\p{E}$ by
+\begin{align}
+  j=\alpha_+\f{dn_+}{dE}=f\alpha\f{dn_+}{dE}.
   \label{eq:walker_current_definition}
@@
 \end{align}
-At equilibrium, $j=dn_+/dt$ is constant, which means we can evaluate
-one of the integrals in (\ref{eq:walker_current_setup})
+At equilibrium, the walker flux $j\p{E}$ must be constant with respect
+to energy. If $j\p{E}$ is not constant with respect to energy, some
+energies would have either a net influx or net outflow of walkers,
+which would mean that the ensemble is not at equilibrium. We can
+therefore evaluate one of the integrals in
+(\ref{eq:walker_current_setup}) to get
 \begin{align}
@@
 (\ref{eq:walker_current_integral}), thereby maximizing the down-going
-walker current $j$, by varying the walker density $n$ with the
-constraint that $n$ must remain normalized. With the additional
-assumption that $\alpha\p{E}$ does not strongly depend on the weights
-$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
-optimal walker density $n$ is
+walker flux $j$, by varying the walker density $n$ with the constraint
+that $n$ must remain normalized. With the additional assumption that
+$\alpha\p{E}$ does not strongly depend on the weights $w\p{E}$, the
+authors of \cite{optimized_ensemble} find that the optimal walker
+density $n$ is
 \begin{align}

### 2015-05-04 papers/histogram/thesis/thesis.tex
fixed one more error in optimized ensemble section

@@
 proportion of walkers at $E$ which are down-going. We also denote the
-net and down-going walker diffusivity at an energy $E$ respectively by
-$\alpha\p{E}$ and $\alpha_+\p{E}$.
-
-We can meaningfully discuss a down-going walker flux $j\p{E}$, which
-is defined in terms of the diffusivity $\alpha_+\p{E}$ by
-\begin{align}
-  j=\alpha_+\f{dn_+}{dE}=f\alpha\f{dn_+}{dE}.
+down-going walker diffusivity at an energy $E$ by $\alpha_+\p{E}$.
+
+The down-going walker flux $j\p{E}$ is defined in terms of the
+diffusivity $\alpha_+\p{E}$ by
+\begin{align}
+  j=\alpha_+\f{dn_+}{dE}.
   \label{eq:walker_current_definition}
@@
 \begin{align}
-  j=f\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
+  j=\alpha_+\sp{n\f{df}{dE}+f\f{dn}{dE}}.
   \label{eq:walker_current_full}
@@
 \begin{align}
-  j=f\alpha n\f{df}{dE}.
+  j=\alpha_+n\f{df}{dE}.
   \label{eq:walker_current}
@@
 \begin{align}
-  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha_+n}.
   \label{eq:walker_current_setup}
@@
 \begin{align}
-  \f1j=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha_+n}.
   \label{eq:walker_current_integral}
@@
 \begin{align}
-  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
+  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha_+}},
   \label{eq:optimal_walker_density}
@@
 \begin{align}
-  w_{\t{opt}}=\f1{D\sqrt{\alpha}},
+  w_{\t{opt}}=\f1{D\sqrt{\alpha_+}},
   \label{eq:optimized_weights}

### 2015-05-04 papers/histogram/thesis/thesis.tex
cleared up confusion I had about diffusivity

--- 2015-05-04 papers/histogram/thesis/thesis.tex
@@
 proportion of walkers at $E$ which are down-going. We also denote the
-net and down-going walker diffusivity at an energy $E$ respectively by
-$\alpha\p{E}$ and $\alpha_+\p{E}$.
-
-We can meaningfully discuss a down-going walker flux $j\p{E}$, which
-is defined in terms of the diffusivity $\alpha_+\p{E}$ by
-\begin{align}
-  j=\alpha_+\f{dn_+}{dE}=f\alpha\f{dn_+}{dE}.
+walker diffusivity at an energy $E$ by $\alpha\p{E}$.
+
+The down-going walker flux $j\p{E}$ is defined in terms of the
+diffusivity $\alpha\p{E}$ by
+\begin{align}
+  j=\alpha\f{dn_+}{dE}.
   \label{eq:walker_current_definition}
@@
 \begin{align}
-  j=f\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
+  j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
   \label{eq:walker_current_full}
@@
 \begin{align}
-  j=f\alpha n\f{df}{dE}.
+  j=\alpha n\f{df}{dE}.
   \label{eq:walker_current}
@@
 \end{align}
-For (\ref{eq:walker_current}) to be valid, we want the ratio in
+For (\ref{eq:walker_current}) to be a valid, we want the ratio in
 (\ref{eq:walker_current_intermediate_ratio}) to vanish. If the walker
@@
 still. The down-going walker density $n_+$, meanwhile must run from
-$n\p{E_-}=0$ to $n\p{E_+}=n\p{E_+}$ meaning $\ln n_+$ runs from $\ln
+$n\p{E_-}=0$ to $n_+\p{E_+}=n\p{E_+}$ meaning $\ln n_+$ runs from $\ln
 n\p{E_-}\to-\infty$ to a finite positive value $\ln n_+\p{E_+}=\ln
-n\p{E_+}$. It is therefore reasonable to assume that $d\ln n_+/d\ln
-n\gg 1$, which makes the ratio in
+n\p{E_+}$. It is therefore reasonable to assume that $\abs{d\ln
+  n_+/d\ln n}\gg 1$, which makes the ratio in
 \ref{eq:walker_current_intermediate_ratio} vanish.
@@
 \begin{align}
-  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
   \label{eq:walker_current_setup}
@@
 \begin{align}
-  \f1j=\int_{E=E_-}^{E_+}\f{dE}{f\alpha n}.
+  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
   \label{eq:walker_current_integral}
@@
 \label{sec:oetmmc}
+
+\begin{align}
+  \bk{\Delta E}=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
+    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
+    w\p{E+\Delta E'}/w\p{E}}
+\end{align}
+
+\begin{align}
+  \bk{\Delta E^2}=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
+    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
+    w\p{E+\Delta E'}/w\p{E}}
+\end{align}
+
+\begin{align}
+  \alpha=\abs{\bk{\Delta E^2}-\bk{\Delta E}^2}
+\end{align}
 

### 2015-05-04 papers/histogram/thesis/thesis.tex
modifications to a derivation for walker current

@@
 
-The down-going walker flux $j\p{E}$ is defined in terms of the
+The down-going walker current $j\p{E}$ is defined in terms of the
 diffusivity $\alpha\p{E}$ by
@@
   j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
+  \label{eq:walker_current_intermediate}
+\end{align}
+We can observe that
+\begin{align}
+  \f{f~dn/dE}{n~df/dE} =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}
+  =\p{\f1f\f{dn_+}{dn}-1}^{-1}.
+  \label{eq:walker_current_approx_ratio_setup}
+\end{align}
+Furthermore,
+\begin{align}
+  \p{\f1f\f{dn_+}{dn}-1}^{-1}
+  =\p{\f1f\f{d\p{fn}}{dn}-1}^{-1}
+  =\p{\f nf\f{df}{dn}+\f{dn}{dn}-1}^{-1}
+  =\p{\f nf\f{df}{dn}}^{-1}=\f nf\f{dn}{df}=\f{d\ln n}{d\ln f},
+  \label{eq:walker_current_approx_intermediate}
+\end{align}
+which means
+\begin{align}
+  \f{f~dn/dE}{n~df/dE}=\f{d\ln n}{d\ln f}.
+  \label{eq:walker_current_approx_ratio}
+\end{align}
+The walker current is therefore
+\begin{align}
+  j=\alpha n\f{df}{dE}\p{1+\f{d\ln n}{d\ln f}}.
   \label{eq:walker_current_full}
 \end{align}
-If the energy histogram does vary strongly with energy, e.g. because
-we are simulating with flat histogram weights, then $dn/dE\approx0$,
-which means
+We know that $f$ runs from $f\p{E_-}=0$ to $f\p{E_+}=1$ meaning $\ln
+f$ runs from $\ln f\p{E_-}\to-\infty$ to $\ln f\p{E_+}=0$. If the
+walker density $n$ varies weakly with energy, e.g. because we are
+simulating with flat histogram weights, then $\ln n$ varies even
+weaker still. We may therefore reasonably to assume that $\abs{d\ln
+  n/d\ln f}\ll 1$ and express
 \begin{align}
@@
 \end{align}
-To formalize the approximation made between
-(\ref{eq:walker_current_full}) and (\ref{eq:walker_current}), we can
-observe that
-\begin{align}
-  \f{f~dn/dE}{n~df/dE}
-  =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}=\p{\f1f\f{dn_+}{dn}-1}^{-1}
-  =\p{\f{d\ln n_+}{d\ln n}-1}^{-1}.
-  \label{eq:walker_current_intermediate_ratio}
-\end{align}
-For (\ref{eq:walker_current}) to be a valid, we want the ratio in
-(\ref{eq:walker_current_intermediate_ratio}) to vanish. If the walker
-density $n$ varies weakly with energy, e.g. because we are simulating
-with flat histogram weights, then $\ln n$ varies even weaker
-still. The down-going walker density $n_+$, meanwhile must run from
-$n\p{E_-}=0$ to $n_+\p{E_+}=n\p{E_+}$ meaning $\ln n_+$ runs from $\ln
-n\p{E_-}\to-\infty$ to a finite positive value $\ln n_+\p{E_+}=\ln
-n\p{E_+}$. It is therefore reasonable to assume that $\abs{d\ln
-  n_+/d\ln n}\gg 1$, which makes the ratio in
-\ref{eq:walker_current_intermediate_ratio} vanish.
-
-We can rearrange (\ref{eq:walker_current}) and integrate from the low
-energy state $E_-$, at which $f\p{E_-}=0$, to the high energy state
-$E_+$, at which $f\p{E_+}=1$, to say
+We can rearrange (\ref{eq:walker_current}) and integrate over the
+energy range to say
 \begin{align}
   \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
-  \label{eq:walker_current_setup}
-\end{align}
-At equilibrium, the walker flux $j\p{E}$ must be constant with respect
-to energy. If $j\p{E}$ is not constant with respect to energy, some
-energies would have either a net influx or net outflow of walkers,
-which would mean that the ensemble is not at equilibrium. We can
-therefore evaluate one of the integrals in
-(\ref{eq:walker_current_setup}) to get
+  \label{eq:walker_current_integral_setup}
+\end{align}
+At equilibrium, the walker current $j\p{E}$ must be constant with
+respect to energy. If $j\p{E}$ is not constant with respect to energy,
+some energies would have either a net incurrent or net outflow of
+walkers, which would mean that the ensemble is not at equilibrium. We
+can therefore evaluate one of the integrals in
+(\ref{eq:walker_current_integral_setup}) to get
 \begin{align}
@@
 (\ref{eq:walker_current_integral}), thereby maximizing the down-going
-walker flux $j$, by varying the walker density $n$ with the constraint
-that $n$ must remain normalized. With the additional assumption that
-$\alpha\p{E}$ does not strongly depend on the weights $w\p{E}$, the
-authors of \cite{optimized_ensemble} find that the optimal walker
-density $n$ is
+walker current $j$, by varying the walker density $n$ with the
+constraint that $n$ must remain normalized. With the additional
+assumption that $\alpha\p{E}$ does not strongly depend on the weights
+$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
+optimal walker density $n$ is
 \begin{align}

### 2015-05-04 papers/histogram/thesis/thesis.tex
added comment about walker current derivation

@@
 \end{align}
+The work developing the optimized ensemble in
+\cite{optimized_ensemble} states (\ref{eq:walker_current}) without
+showing how it follows from the definition of walker current expressed
+in (\ref{eq:walker_current_definition}) and making explicit the
+approximation between the exact result (\ref{eq:walker_current_full})
+and the working form (\ref{eq:walker_current}).
+
 We can rearrange (\ref{eq:walker_current}) and integrate over the

### 2015-05-05 papers/histogram/thesis/thesis.tex

--- 2015-05-04 papers/histogram/thesis/thesis.tex
@@
 
-The down-going walker current $j\p{E}$ is defined in terms of the
+At equilibrium with a flat histogram, meaning $n$ is constant, the
+down-going walker current $j\p{E}$ is defined in terms of the
 diffusivity $\alpha\p{E}$ by
 \begin{align}
-  j=\alpha\f{dn_+}{dE}.
+  j=-\alpha\f{dn_+}{dE}.
   \label{eq:walker_current_definition}
 \end{align}
-Computing
-\begin{align}
-  n\f{df}{dE}=n\sp{\f1{n}\f{dn_+}{dE}-\f{n_+}{n^2}\f{dn}{dE}}
-  =\f{dn_+}{dE}-f\f{dn}{dE},
-  \label{eq:walker_intermediate}
-\end{align}
-we can say that
-\begin{align}
-  \f{dn_+}{dE}=n\f{df}{dE}+f\f{dn}{dE}.
-  \label{eq:walker_gradient}
-\end{align}
-Substituting (\ref{eq:walker_gradient}) into
-(\ref{eq:walker_current_definition}) yields
-\begin{align}
-  j=\alpha\sp{n\f{df}{dE}+f\f{dn}{dE}}.
-  \label{eq:walker_current_intermediate}
-\end{align}
-We can observe that
-\begin{align}
-  \f{f~dn/dE}{n~df/dE} =\p{\f{dn_+/dE-fdn/dE}{f~dn/dE}}^{-1}
-  =\p{\f1f\f{dn_+}{dn}-1}^{-1}.
-  \label{eq:walker_current_approx_ratio_setup}
-\end{align}
-Furthermore,
-\begin{align}
-  \p{\f1f\f{dn_+}{dn}-1}^{-1}
-  =\p{\f1f\f{d\p{fn}}{dn}-1}^{-1}
-  =\p{\f nf\f{df}{dn}+\f{dn}{dn}-1}^{-1}
-  =\p{\f nf\f{df}{dn}}^{-1}=\f nf\f{dn}{df}=\f{d\ln n}{d\ln f},
-  \label{eq:walker_current_approx_intermediate}
-\end{align}
-which means
-\begin{align}
-  \f{f~dn/dE}{n~df/dE}=\f{d\ln n}{d\ln f}.
-  \label{eq:walker_current_approx_ratio}
-\end{align}
-The walker current is therefore
-\begin{align}
-  j=\alpha n\f{df}{dE}\p{1+\f{d\ln n}{d\ln f}}.
-  \label{eq:walker_current_full}
-\end{align}
-We know that $f$ runs from $f\p{E_-}=0$ to $f\p{E_+}=1$ meaning $\ln
-f$ runs from $\ln f\p{E_-}\to-\infty$ to $\ln f\p{E_+}=0$. If the
-walker density $n$ varies weakly with energy, e.g. because we are
-simulating with flat histogram weights, then $\ln n$ varies even
-weaker still. We may therefore reasonably to assume that $\abs{d\ln
-  n/d\ln f}\ll 1$ and express
-\begin{align}
-  j=\alpha n\f{df}{dE}.
+This identity can be taken as the definition of the walker diffusivity
+$\alpha$, which is the independent of the label we have assigned any
+given walker. Substituting $n_+=fn$,
+\begin{align}
+  j=-\alpha\p{n\f{df}{dE}+f\f{dn}{dE}}=-\alpha n\f{df}{dE},
   \label{eq:walker_current}
 \end{align}
-We can rearrange (\ref{eq:walker_current}) and integrate over the
-energy range to say
-\begin{align}
-  \int_{f=0}^1\f{df}{j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
+where a constant $n$ means $dn/dE=0$. We can rearrange
+(\ref{eq:walker_current}) and integrate over the energy range as
+\begin{align}
+  \int_{f=f\p{E_-}}^{f\p{E_+}}\f{df}{j}=-\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
   \label{eq:walker_current_integral_setup}
 \end{align}
-At equilibrium, the walker current $j\p{E}$ must be constant with
-respect to energy. If $j\p{E}$ is not constant with respect to energy,
-some energies would have either a net incurrent or net outflow of
-walkers, which would mean that the ensemble is not at equilibrium. We
-can therefore evaluate one of the integrals in
-(\ref{eq:walker_current_integral_setup}) to get
-\begin{align}
-  \f1j=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
+At equilibrium, the number of down-going walkers $n_+\p{E}$ at any
+given energy $E$ is constant with time, which means that the walker
+flux $j\p{E}$ must be constant with energy. We can therefore evaluate
+one of the integrals in (\ref{eq:walker_current_integral_setup}) to
+get
+\begin{align}
+  \f1{\abs j}=\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
   \label{eq:walker_current_integral}
@@
 (\ref{eq:walker_current_integral}), thereby maximizing the down-going
-walker current $j$, by varying the walker density $n$ with the
+walker current $\abs{j}$, by varying the walker density $n$ with the
 constraint that $n$ must remain normalized. With the additional
@@
 \begin{align}
-  n_{\t{opt}}=\f1{\sqrt{\lambda\alpha}},
+  n_{\t{opt}}\propto\f1{\sqrt{\alpha}},
   \label{eq:optimal_walker_density}
 \end{align}
-where $\lambda$ is the Lagrange multiplier used in the minimization
-process, and here acts merely as a normalization factor.
+where the proportionality is determined by enforcing that $n$ is
+normalized.
 
@@
 
-\begin{align}
-  \bk{\Delta E}=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
+By the central limit theorem, the probability distribution $P\p{E\to
+  E+\Delta E}$ of Monte Carlo moves which changes the energy of a
+system by $\Delta E$ is
+\begin{align}
+  P\p{E\to E+\Delta E}=\f1{\sqrt{2\pi\sigma^2}}
+  \exp\p{-\f{\Delta E^2}{2\sigma^2}}
+\end{align}
+for some variance
+\begin{align}
+  \sigma^2=\sigma_E\p{\Delta E}^2=\bk{\Delta E^2}_E-\bk{\Delta E}_E
+\end{align}
+which generally depends on the energy $E$.
+
+\vspace{3cm}
+
+\begin{align}
+  \bk{\Delta E}_E=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
     w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
@@
 \begin{align}
-  \bk{\Delta E^2}=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
+  \bk{\Delta E^2}_E=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
     w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
@@
 \begin{align}
-  \alpha=\abs{\bk{\Delta E^2}-\bk{\Delta E}^2}
+  \alpha=\bk{\Delta E^2}_E-\bk{\Delta E}_E^2=\sigma\p{E}^2
 \end{align}

### 2015-05-05 papers/fuzzy-fmt/figs/plot_radial.py
Made plots prettier for soft spheres paper. Tested on knightley. Took 0:28.67.

@@ -1,143 +1,0 @@
-#!/usr/bin/python
-
-from __future__ import division
-# We need the following two lines in order for matplotlib to work
-# without access to an X server.
-import matplotlib
-matplotlib.use('Agg')
-from pylab import *
-from scipy.special import erf
-import matplotlib.lines as mlines
-import os
-
-import styles
-
-if len(sys.argv) != 3:
-    print("Usage:  " + sys.argv[0] + ' filling-fraction temperatures')
-    exit(1)
-
-reduced_density = float(sys.argv[1])
-#reduced_density = [.3, .1,.2]
-
-all_temperatures = eval(sys.argv[2])
-#all_temperatures = [[4,4,1]]
-print 'type of all temps is',type(all_temperatures)
-
-print 'all_temperatures are', all_temperatures
-
-#Constants and variables
-#k_b = 8.6173324*10**(-5) # in eV
-dT = .001
-#Temp_max = 600 #in Kelvin
-#Temp = arange(.001, .1 + dT/2, dT)
-V0 = 1
-#betaV0 = V0/Temp
-R = 1# in Angstroms
-density = arange(0, .8 - .001/2, .001)/(4*pi/3)
-#gamma = 2*((sqrt(pi*betaV0)+sqrt(pi*betaV0-16*sqrt(betaV0)))/8)**2
-#sg = sqrt(gamma)
-
-eta = density*4*pi/3
-#P_cs = density*.001*(1+eta+eta**2-eta**3)/(1-eta)**3
-P_cs = density*(1+eta+eta**2)/(1-eta)**3
-#plot(eta,P_cs/.001, 'k',linewidth=2, label = 'Hard spheres')
-
-
-density = (reduced_density/100)/(4*pi/3)
-phs = density*(1+(reduced_density/100)+(reduced_density/100)**2)/(1-(reduced_density/100))**3
-
-mysymbol_names = []
-mysymbol_lines = []
-
-mylines = []
-# for temp in all_temperatures:
-#   # input: ['figs/mc-0.%02d00-%.4f.dat.gradial' % (reduced_density, temp) for temp in all_temperatures]
-#   fname = 'figs/mc-0.%02d00-%.4f.dat.gradial' % (reduced_density, temp)
-#   if os.path.exists(fname):
-#     print 'found', fname
-#     g = loadtxt(fname)
[…]

### 2015-05-06 papers/histogram/thesis/thesis.tex
added subsection about end conditions, as well as some helpful figures

@@
 total number of states $s$ for which $X\p{s}=X_0$. It is sometimes the
-case, however, that the density of states in some region $R$ in the
-range of possible $X$ is so low that it is practically impossible (via
+case, however, that the density of states in some range $R$ of
+possible $X$ is so low that it is practically impossible (via
 Metropolis Monte Carlo) to sufficiently sample $R$, that is, to
@@
 expressed in terms of the number of microstates $\Omega\p{X}$ as
-$S\p{X}=k\ln\Omega\p{X}\propto\ln D\p{X}$, we may refer to regions $R$
-low state densities $D\p{X\in R}$ as low entropy states.
+$S\p{X}=k\ln\Omega\p{X}\propto\ln D\p{X}$, we may refer to regions
+with low state densities $D\p{X}$ as low entropy states.
+
+Figure \ref{fig:dos_sample} provides an example of an unnormalized
+density of states $\tilde D\p{E}$ in energy for a particular
+square-well fluid. The density of states has been ``normalized'' to
+have a maximum value of $1$ in order to emphasize the logarithmic span
+of $D\p{E}$. This logarithmic span ($\sim10^{45}$) is proportional to
+the number of moves which one must simulate via Metropolis Monte Carlo
+in order to observe the energies with the lowest state densities. For
+reference, the world's current fastest supercomputers can perform
+$\sim10^{16}$ floating-point operations per second.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figs/dos-sample.pdf}
+  \caption[Sample density of states]{Sample density of states for
+    square-well fluid with a well width $\lambda=1.3$, filling
+    fraction $\eta=0.3$, and $N=20$ spheres, as computed by several
+    histogram methods. The density of states has been normalized to
+    have a maximum value of $1$.}
+  \label{fig:dos_sample}
+\end{figure}
 
@@
 \ref{alg:metropolis} in order to accept an attempted move: the weights
-$w\sp{X\p{s_i}}$ and $w\sp{X\p{s_f}}$ of the initial (pre-move) state
-$s_i$ and final (post-move) state $s_f$ are used to determine the
+$w\p{X_i}$ and $w\p{X_f}$ of the initial (pre-move) state $s_i$ and
+final (post-move) state $s_f$, respectively, are used to determine the
 probability $P_m\p{s_i\to s_f}$ of accepting an otherwise valid move
@@
 \begin{align}
-  P_m\p{s_i\to s_f}=\min\set{\f{w\sp{X\p{s_f}}}{w\sp{X\p{s_i}}},1}.
+  P_m\p{s_i\to s_f}=\min\set{\f{w\p{X_f}}{w\p{X_i}},1}.
   \label{eq:move_prob}
 \end{align}
-This formula means that when $w\sp{X\p{s_f}}>w\sp{X\p{s_i}}$, the move
-$s_i\to s_f$ is accepted; when $w\sp{X\p{s_f}}<w\sp{X\p{s_i}}$, the
-ratio of these weights determines the probability of accepting the
-move. Due to the fact that only ratios of weights determine
-$P_m\p{s_i\to s_f}$, the weights $w\sp{X\p{s}}$ are scale-invariant,
-meaning that their effect on simulations is unchanged by scale factors
-that are constant with respect to $X\p{s}$.
-
-\begin{algorithm}[tb]
-  \caption{Biased Monte Carlo fluid simulation}
-  \label{alg:biased_MC}
-  \begin{alg}
-
-  \item Construct an appropriate weight function $w\sp{X\p{s}}$, whose
-    argument is a system property $X\p{s}$ determined by the
-    microstate $s$.
-
-  \item Construct an initial typical fluid configuration.
-
-  \item Randomly attempt to change the position of one sphere,
-    rejecting the change if
-    \begin{enumerate*}[label=\roman*)]
-    \item it results in a forbidden fluid configuration, or
-    \item a newly chosen random number on the interval $\sp{0,1}$ is
-      larger than the probability determined by (\ref{eq:move_prob})
-      for the initial and final states $s_i$ and $s_f$, respectively.
-    \end{enumerate*}
-    \label{alg:biased_mc_move}
-
-  \item Repeat step \ref{alg:biased_mc_move} the simulation produces
-    statistics of sufficient quality.
-
-  \end{alg}
-\end{algorithm}
+This formula means that when $w\p{X_f}>w\p{X_i}$, the move $s_i\to
+s_f$ is accepted; when $w\p{X_f}<w\p{X_i}$, the ratio of these weights
+determines the probability of accepting the move. Due to the fact that
+only ratios of weights determine $P_m\p{s_i\to s_f}$, the weights
+$w\p{X}$ are scale-invariant, meaning that their effect on simulations
+is unchanged by scale factors that are constant with respect to
+$X\p{s}$. Furthermore, simulating with flat weights $w\p{X}=w_0$ for
+all $X$ is equivalent to simulating without weights, as the
+probability $P_m\p{s_i\to s_f}$ will always equal 1.
 
@@
 
-In this paper, we will consider weights $w\p{E}$ which depend only on
-the energy $E$ of the square-well fluid. Due to the fact that the
-square-well fluid can only have discrete energies $E=-n\epsilon$,
-where $n$ is a non-negative integer and $\epsilon$ is the well depth,
-in simulation we store the weight function as an array of values. We
-may therefore use the terms ``weight function,'' and ``weight array,''
-and ``weights'' interchangeably.
-
-\fixme{mention that simulating with flat weights $w\p{E}=w_0$ is the
-  same as running a Metropolis MC simulation.}
+\begin{algorithm}[tb]
+  \caption{Biased Monte Carlo fluid simulation}
+  \label{alg:biased_MC}
+  \begin{alg}
+
+  \item Construct an appropriate weight function $w\sp{X\p{s}}$, whose
+    argument is a system property $X\p{s}$ determined by the
+    microstate $s$.
+
+  \item Construct an initial typical fluid configuration.
+
+  \item Randomly attempt to change the position of one sphere,
+    rejecting the change if
+    \begin{enumerate*}[label=\roman*)]
+    \item it results in a forbidden fluid configuration, or
+    \item a newly chosen random number on the interval $\sp{0,1}$ is
+      larger than the probability determined by (\ref{eq:move_prob})
+      for the initial and final states $s_i$ and $s_f$, respectively.
+    \end{enumerate*}
+    \label{alg:biased_mc_move}
+
+  \item Repeat step \ref{alg:biased_mc_move} the simulation produces
+    statistics of sufficient quality.
+
+  \end{alg}
+\end{algorithm}
+
+In this paper, we will use a histogram $H\p{E}$, weights $w\p{E}$, and
+density of states $D\p{E}$ which are functions of the square-well
+fluid's energy $E$. Due to the fact that the square-well fluid can
+only have discrete energies $E=-n\epsilon$, where $n$ is a
+non-negative integer and $\epsilon$ is the well depth, in simulation
+we store the weight function as an array of values. We will therefore
+generally refer to the weight function as a ``weight array,'' or
+simply ``weights.''
 
@@
   X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT_0}
-  =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}}.
+  =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}},
   \label{eq:X_canonical}
 \end{align}
-The simplifications in in (\ref{Z_canonical}) and
-(\ref{eq:X_canonical}), which have no explicit dependence on $T_0$,
-occur because canonical Monte Carlo simulations sample energies in
-proportion to the distribution (over energy) of microstates at a
-temperature of $T_0$.  Canonical Monte Carlo simulations can therefore
-fail to sufficiently sample energies which are important
-(i.e. energies with a non-negligible state probability density) at
-different temperatures. As a consequence, such simulations should not
-be used to determine properties $X\p{T\ne T_0}$, and are thus referred
-to as ``fixed temperature'' simulations.
+where $\bk{X}_E$ is the mean value of $X$ at the energy $E$. The
+simplifications in in (\ref{Z_canonical}) and (\ref{eq:X_canonical}),
+which have no explicit dependence on $T_0$, occur because canonical
+Monte Carlo simulations sample energies in proportion to the
+distribution (over energy) of microstates at a temperature of $T_0$.
+Canonical Monte Carlo simulations can therefore fail to sufficiently
+sample energies which are important (i.e. energies with a
+non-negligible state probability density) at different
+temperatures. As a consequence, such simulations should not be used to
+determine properties $X\p{T\ne T_0}$, and are thus referred to as
+``fixed temperature'' simulations.
 
@@
 many simulations at discrete temperature intervals; each such
-simulation will yield one sample (e.g. one data point) of $\bk{X}_T$.
+simulation will yield one data point on $\bk{X}_T$.
 
@@
 system into a local minimum of its energy landscape. Freezing into a
-state means that a simulation will sample only a small portion of the
-energy landscape, even though there may (and generally will) be many
-other states with the same energy.
+state in this manner will limit a simulation to sampling only a small
+portion of the energy landscape, even though there may (and generally
+will) be many other states with the same energy.
 
@@
 study the square-well fluid. We will, however, use canonical weights
-for part of all weight arrays, as discussed in Section
-\ref{sec:min_energy}.
+for portions of all weight arrays, which we discuss in Section
+\ref{sec:energy_range}.
 
@@
 states $D\p{E}$. This algorithm involves first running a simulation
-without weights (or with constant weights $w\p{E}=w_0$) while
-collecting a histogram $H\p{E}$ of the observed energies $E$. After
-some time (loosely defined), the histogram $H\p{E}$ may be taken as an
-approximation of the density of states $\tilde D\p{E}$ (see Section
-\ref{sec:histogram_methods}), and used via (\ref{eq:flat_weights}) to
-generate a weight array for an actual simulation. Algorithm
-\ref{alg:flat_histogram} has one free parameter, which determines the
-amount of time for which to collect an energy histogram $H\p{E}$.
+without weights (or with constant weights $w\p{E}=1$) while collecting
+a histogram $H\p{E}$ of the observed energies $E$. After some time
+(loosely defined), the histogram $H\p{E}$ may be taken as an
+approximation of the density of states $\tilde D\p{E}$, and used via
+(\ref{eq:flat_weights}) to generate a weight array for an actual
+simulation. Algorithm \ref{alg:flat_histogram} has one free parameter,
+which determines the amount of time for which to collect an energy
+histogram $H\p{E}$.
 
@@
 simulating for any reasonable amount of time will be statistically
-insignificant or null, invalidating the approximation
-$H\p{E}\approx\tilde D\p{E}$. This method will therefore only
-``flatten'' the histogram at energies $E$ which have been sufficiently
-sampled for $H\p{E}$ to be statistically significant. The end goal in
-implementing histogram methods, however, is precisely to sample those
-energies that cannot be sufficiently sampled without employing clever
-schemes to bias Monte Carlo simulations. The flat histogram method
-therefore does not directly help us study the square-well fluid. This
-method does, however, motivate the method in Section
-\ref{sec:simple_flat}, and provides some of the theory behind the
-methods in Sections \ref{sec:tmmc} and \ref{sec:oetmmc}.
+insignificant or null, invalidating the proportionality $H\p{E}\propto
+D\p{E}$. This method will therefore only ``flatten'' the histogram at
+energies $E$ which have been sufficiently sampled for $H\p{E}$ to be
+statistically significant. The end goal in implementing histogram
+methods, however, is precisely to sample those energies that cannot be
+sufficiently sampled without employing clever schemes to bias Monte
+Carlo simulations. Therefore, this method does not directly help us
+study the square-well fluid. This method does, however, motivate the
+method in Section \ref{sec:simple_flat}, and provides some of the
+theory behind the methods in Sections \ref{sec:tmmc} and
+\ref{sec:oetmmc}.
 
@@
 
-\begin{algorithm}[!t]
+\begin{algorithm}[!b]
   \caption{The simple flat method}
@@
 
-\begin{algorithm}[!b]
+\begin{algorithm}[tb]
   \caption{Wang-Landau initialization of weights}
@@
       \ref{alg:wang_landau_sim}.
+      \label{alg:wang_landau_end}
     \end{enumerate}
@@
 Algorithm \ref{alg:wang_landau} spells out the Wang-Landau
-initialization process, which has five free parameters: the number of
-iterations $n$ to run in step \ref{alg:wang_landau_sim}, the initial
-factor $f_0$, the update factor $u$, the flatness condition $C\sp{H}$,
-and the cutoff $c$. The first of these parameters, $n$, controls how
-often to check the flatness condition $C\sp{H}$ and run updates, if
-necessary. The value of $n$ is not too important, so long as it is
-reasonable for the system at hand. Too small value of $n$ will waste
-computation time determining $C\sp{H}$, while too large of a value
-will waste time simulating when $C\sp{H}$ has already been
-satisfied. In general, $n$ should be proportional to the computation
-time of $C\sp{H}$, which scales with energy range of the simulated
-system. We therefore use $n=L$, where $L$ is, as described in Section
-\ref{sec:simple_flat}, the number of energy levels for the given
-square-well fluid. Other works typically neglect providing any
-heuristic or formula for $n$, reporting a system-independent
-$n=10^3$\cite{wang_landau_mod} or $10^5$\cite{wang_landau}.
-\fixme{find more citations for this last claim}
+method. This algorithm includes four free parameters which affect
+initialization process: the number of iterations $n$ to run in step
+\ref{alg:wang_landau_sim}, the initial factor $f_0$, the update factor
+$u$, and the flatness condition $C\sp{H}$. The cutoff $c$ determines
+the condition for Algorithm \ref{alg:wang_landau} to terminate.
+
+The first of these parameters, $n$, controls how often to check the
+flatness condition $C\sp{H}$ and run updates, if necessary. The value
+of $n$ is not too important, so long as it is reasonable for the
+system at hand. Too small value of $n$ will waste computation time
+determining $C\sp{H}$, while too large of a value will waste time
+simulating when $C\sp{H}$ has already been satisfied. In general, $n$
+should be proportional to the computation time of $C\sp{H}$, which
+scales with energy range of the simulated system. We therefore use
+$n=L$, where $L$ is, as described in Section \ref{sec:simple_flat},
+the number of energy levels for the given square-well fluid. Other
+works typically neglect providing any heuristic or formula for $n$,
+reporting a system-independent $n=10^3$\cite{wang_landau_mod} or
+$10^5$\cite{wang_landau}.  \fixme{find more citations for this last
+  claim}
 
@@
 determining an appropriate value of $c$, but literature cites typical
-values of $\ln c=10^{-8}$; we used $\ln c=10^{-6}$.
-
-\subsection{The transition matrix Monte Carlo method}
+values of $\ln c=10^{-8}$. In order to compare different histogram
+methods fairly, we used an alternate end condition in step
+\ref{alg:wang_landau_end} of Algorithm \ref{alg:wang_landau}
+(discussed in Section \ref{sec:end_conditions}), and therefore did not
+need to choose a value of $c$.
+
+\subsection{The transition matrix Monte Carlo (TMMC) method}
 \label{sec:tmmc}
 
-\fixme{add citations}
-
 Unlike the previous histogram methods, the transition matrix Monte
-Carlo (TMMC) method does not use weights, and merely compute the
-density of states $D\p{E}$ of a system\cite{tmmc}. The density of
-states can in turn be used to determine flat histogram weights via
+Carlo method does not use weights, and merely compute the density of
+states $D\p{E}$ of a system\cite{tmmc}. The density of states can in
+turn be used to determine flat histogram weights via
 (\ref{eq:flat_weights}). TMMC introduces a new object: the energy
@@
 \end{align}
-At equilibrium, the probability distributions $\hat D^{t}=\hat
-D^{t-1}$. Furthermore, the distribution of states in an equilibrium
-ensemble of Metropolis Monte Carlo simulations is precisely the
-density of states $D\p{E}$\fixme{explain why?}, which means
+At equilibrium, the probability distributions do not vary with time,
+meaning $\hat D^{t}=\hat D^{t-1}$. Furthermore, the distribution of
+states in an equilibrium ensemble of Metropolis Monte Carlo
+simulations is precisely the density of states $D\p{E}$, which means
 $D_i=T_{ij}D_j$.  Simultaneously expressing this condition for all
@@
 \ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis}. A
-histogram of transitions $\tilde T\p{E_i,\bar E_j}_{}$ from a fixed
+histogram of transitions $\tilde T\p{E_i,\bar E_j}$ from a fixed
 energy $\bar E_j$ to energies $E_i$ is then proportional to the
@@
 \begin{align}
-  T_{ij}=\f{\tilde T_d\p{E_j,E_i-E_j}}{\sum_i\tilde T_d\p{E_j,E_i-E_j}}
-  =\f{\tilde T_d\p{E_j,E_i-E_j}}
-  {\sum_{\Delta E}\tilde T_d\p{E_j,\Delta E}}.
+  T_{ij}=\f{\tilde T_d\p{E_j,E_i-E_j}}{\sum_i\tilde
+    T_d\p{E_j,E_i-E_j}} =\f{\tilde T_d\p{E_j,E_i-E_j}} {\sum_{\Delta
+      E}\tilde T_d\p{E_j,\Delta E}}.
   \label{eq:transition_conversion}
@@
 difference $M$.
+
+\begin{algorithm}[!b]
+  \caption{Transition matrix Monte Carlo initialization}
+  \label{alg:tmmc}
+  \begin{alg}
+
+  \item Construct an initial typical fluid configuration.
+
+  \item Initialize a transition histogram $\tilde T_d\p{E,\Delta E}=0$
+    for all $E$ and $\Delta E\in\sp{-M,M}$, where $M$ is the possible
+    maximum energy transition of a single move.
+
+  \item Randomly attempt to change the position of one sphere,
+    tentatively accepting the transition from state $s_i$ to state
+    $s_f$ if $s_f$ is not a forbidden fluid configuration.
+    \label{alg:tmmc_move}
+
+  \item Let $\Delta E=E_f-E_i$ and increment $\tilde
+    T_d\p{E_i,\Delta{E}}$, where $E_k$ is the energy of state $s_k$.
+
+  \item If $\Delta E<0$, accept the move. Otherwise
+    \begin{enumerate}
+    \item compute the normalization factor
+      \begin{align*}
+        n_k=\sum_{\Delta E'}\tilde T_d\p{E_k,\Delta E'}
+      \end{align*}
+      for $k\in\set{i,f}$, and
+
+    \item accept the move $s_i\to s_f$ with probability
+      \begin{align*}
+        P_m\p{s_i\to s_f} =\max\set{\f{\p{\tilde T_d\p{E_i,\Delta
+                E}+c}/\p{n_i+c}} {\p{\tilde T_d\p{E_f,-\Delta
+                E}+c}/\p{n_f+c}}, \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
+      \end{align*}
+      for a predetermined number $c$ and temperature $T_{\t{min}}$.
+      \label{alg:tmmc_prob}
+    \end{enumerate}
+
+  \item If some predetermined initialization end condition has
+    \emph{not} been met, return to step
+    \ref{alg:tmmc_move}. Otherwise, compute the density of states
+    $D\p{E}$ by solving (\ref{eq:dos_eigen}), and set weights
+    $w\p{E}=1/D\p{E}$.
+
+  \end{alg}
+\end{algorithm}
+
+Figure \ref{fig:transitions_sample} provides an example of a
+transition matrix $T_d\p{E,\Delta E}$ for a particular square-well
+fluid. The yellow band at $\Delta E=0$ indicates that most moves
+transition from an energy to itself. Low energies are more likely to
+transition up in energy, with $\Delta E>0$, than down in energy, with
+$\Delta E<0$. The high energies which are more likely to transition
+down in energy than up are those above the state of maximum entropy.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.6\textwidth]{figs/transitions-sample.pdf}
+  \caption[Sample transition matrix]{Sample transition matrix
+    $T_d\p{E,\Delta E}$ for square-well fluid with a well width
+    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=20$ spheres.}
+  \label{fig:transitions_sample}
+\end{figure}
 
@@
   P_m\p{s_i\to s_f}=\f{\p{\tilde T_d\p{E_i,\Delta E}}\big/
-    \p{\sum_{\Delta E'}\tilde T_d\p{E_i,\Delta E'}}}
-  {\p{\tilde T_d\p{E_f,-\Delta E}}\big/
-    \p{\sum_{\Delta E'}\tilde T_d\p{E_f,\Delta E'}}}
-  =\f{T\p{E_i,\Delta E}}{T\p{E_f,-\Delta E}}
-  =\f{P\p{E_i\to E_f}}{P\p{E_f\to E_i}},
+    \p{\sum_{\Delta E'}\tilde T_d\p{E_i,\Delta E'}}} {\p{\tilde
+      T_d\p{E_f,-\Delta E}}\big/ \p{\sum_{\Delta E'}\tilde
+      T_d\p{E_f,\Delta E'}}} =\f{T_d\p{E_i,\Delta
+      E}}{T_d\p{E_f,-\Delta E}} =\f{P\p{E_i\to E_f}}{P\p{E_f\to E_i}},
   \label{eq:tmmc_prob_simplified}
@@
 
-\begin{algorithm}[tb]
-  \caption{Transition matrix Monte Carlo initialization}
-  \label{alg:tmmc}
-  \begin{alg}
-
-  \item Construct an initial typical fluid configuration.
-
-  \item Initialize a transition histogram $\tilde T_d\p{E,\Delta E}=0$
-    for all $E$ and $\Delta E\in\sp{-M,M}$, where $M$ is the possible
-    maximum energy transition of a single move.
-
-  \item Randomly attempt to change the position of one sphere,
-    tentatively accepting the transition from state $s_i$ to state
-    $s_f$ if $s_f$ is not a forbidden fluid configuration.
-    \label{alg:tmmc_move}
-
-  \item Let $\Delta E=E_f-E_i$ and increment $\tilde
-    T_d\p{E_i,\Delta{E}}$, where $E_k$ is the energy of state $s_k$.
-
-  \item If $\Delta E<0$, accept the move. Otherwise
-    \begin{enumerate}
-    \item compute the normalization factor
-      \begin{align*}
-        n_k=\sum_{\Delta E'}\tilde T_d\p{E_k,\Delta E'}
-      \end{align*}
-      for $k\in\set{i,f}$, and
-
-    \item accept the move $s_i\to s_f$ with probability
-      \begin{align*}
-        P_m\p{s_i\to s_f}
-        =\max\set{\f{\p{\tilde T_d\p{E_i,\Delta E}+c}/\p{n_i+c}}
-          {\p{\tilde T_d\p{E_f,-\Delta E}+c}/\p{n_f+c}},
-          \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
-      \end{align*}
-      for a predetermined number $c$ and temperature $T_{\t{min}}$.
-      \label{alg:tmmc_prob}
-    \end{enumerate}
-
-  \item If some predetermined initialization end condition has
-    \emph{not} been met, return to step \ref{alg:tmmc_move}.
-
-  \end{alg}
-\end{algorithm}
-
-\subsection{The optimized ensemble method}
+\subsection{The optimized ensemble (OE) method}
 \label{sec:optimized_ensemble}
@@
 are based on many samples of only a few low energy regions of a
-system's energy landscape. The optimized ensemble
-(OE)\cite{optimized_ensemble} attempts to address the autocorrelation
-between low energy samples by finding weights to maximize the rate at
-which a simulation makes round trips between low and high energy
-states. Maximizing the round trip rate, in turn, maximizes the rate at
-which a simulation makes independent, uncorrelated samples of low
-energies.
+system's energy landscape. The optimized
+ensemble\cite{optimized_ensemble} attempts to address the
+autocorrelation between low energy samples by finding weights to
+maximize the rate at which a simulation makes round trips between low
+and high energy states. Maximizing the round trip rate, in turn,
+maximizes the rate at which a simulation makes independent,
+uncorrelated samples of low energies.
 
@@
 \begin{align}
-  \int_{f=f\p{E_-}}^{f\p{E_+}}\f{df}{j}=-\int_{E=E_-}^{E_+}\f{dE}{\alpha n}.
+  \int_{f=f\p{E_-}}^{f\p{E_+}}\f{df}{j}=-\int_{E=E_-}^{E_+}\f{dE}{\alpha
+    n}.
   \label{eq:walker_current_integral_setup}
@@
 
-The algorithm provided in \cite{optimized_ensemble} to determine and
-construct weights given by \ref{eq:optimized_weights} involves
-simulating with and iteratively modifying flat histogram weights. This
-algorithm is not meant to construct weights from scratch, but rather
-to optimize them. We therefore will not be comparing the optimized
-ensemble to other methods. We do, however, credit this method for
-motivating the hybrid OE-TMMC method in Section \ref{sec:oetmmc}.
+The algorithm provided in \cite{optimized_ensemble} to construct
+weights given by (\ref{eq:optimized_weights}) finds the diffusivity
+$\alpha$ by solving for it in
+(\ref{eq:walker_current_definition}). This procedure accepts flat
+histogram weights, or equivalently the density of states, as an input,
+and is not meant to construct weights from scratch. We therefore will
+not be comparing the optimized ensemble to other methods in this
+paper. We do, however, credit \cite{optimized_ensemble} for motivating
+the hybrid OE-TMMC method in Section \ref{sec:oetmmc}.
 
@@
 
-By the central limit theorem, the probability distribution $P\p{E\to
-  E+\Delta E}$ of Monte Carlo moves which changes the energy of a
-system by $\Delta E$ is
-\begin{align}
-  P\p{E\to E+\Delta E}=\f1{\sqrt{2\pi\sigma^2}}
-  \exp\p{-\f{\Delta E^2}{2\sigma^2}}
-\end{align}
-for some variance
-\begin{align}
-  \sigma^2=\sigma_E\p{\Delta E}^2=\bk{\Delta E^2}_E-\bk{\Delta E}_E
-\end{align}
-which generally depends on the energy $E$.
-
-\vspace{3cm}
-
-\begin{align}
-  \bk{\Delta E}_E=\f{\sum_{\Delta E}\Delta E~\tilde T\p{E,\Delta E}
-    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
-    w\p{E+\Delta E'}/w\p{E}}
-\end{align}
-
-\begin{align}
-  \bk{\Delta E^2}_E=\f{\sum_{\Delta E}\Delta E^2~\tilde T\p{E,\Delta E}
-    w\p{E+\Delta E}/w\p{E}}{\sum_{\Delta E'}\tilde T\p{E,\Delta E'}
-    w\p{E+\Delta E'}/w\p{E}}
-\end{align}
-
-\begin{align}
-  \alpha=\bk{\Delta E^2}_E-\bk{\Delta E}_E^2=\sigma\p{E}^2
-\end{align}
-
-\fixme{Explain, provide algorithm. This method is our own synthesis of
-  other work.}
-
-\begin{algorithm}[H]
-  \caption{Modifying transition weights via the optimized ensemble}
-  \label{alg:oetmmc}
-\end{algorithm}
-
-\subsection{Identifying the energy range of interest}
+As well as the density of states, the transition matrix $T\p{E_i,E_j}$
+determines the local diffusivity $\alpha\p{E}$ for any given weights
+$w\p{E}$. Denoting by $P_b\p{E_i\to E_j}$ the actual, biased
+probability of transitioning from $E_i$ to $E_j$ in a single Monte
+Carlo move, we can say that
+\begin{align}
+  P_b\p{E_i\to E_f}=P\p{E_i\to E_f}P_m\p{E_i\to E_f}
+  =T_{ji}\max\set{\f{w\p{E_f}}{w\p{E_i}},1},
+\end{align}
+where $P\p{E_i\to E_j}$ is the probability of attempting the move
+$E_i\to E_j$ and $P_m\p{E_i\to E_j}$ is the probability of accepting
+that move, should it occur. In terms of a transition histogram $\tilde
+T\p{E,\Delta E}$ with $E=E_i$ and $\Delta E=E_f-E_i$,
+\begin{align}
+  P_b\p{E\to E+\Delta E} =\f{\tilde T_d\p{E,\Delta E}w\p{E+\Delta
+      E}/w\p{E}} {\sum_{\Delta E'}\tilde T_d\p{E,\Delta
+      E'}w\p{E+\Delta E'}/w\p{E}}.
+  \label{eq:actual_transition_prob}
+\end{align}
+We now define
+\begin{align}
+  \bk{\Delta E}_E=\sum_{\Delta E}\Delta E P_b\p{E\to E+\Delta E}
+  =\f{\sum_{\Delta E}\Delta E~\tilde T_d\p{E,\Delta E}
+    \max\set{w\p{E+\Delta E}/w\p{E},1}} {\sum_{\Delta E'}\tilde
+    T_d\p{E,\Delta E'} \max\set{w\p{E+\Delta E'}/w\p{E},1}}
+  \label{eq:<de>}
+\end{align}
+and
+\begin{align}
+  \bk{\Delta E^2}_E=\sum_{\Delta E}\Delta E^2 P_b\p{E\to E+\Delta E}
+  =\f{\sum_{\Delta E}\Delta E~\tilde T_d\p{E,\Delta E}
+    \max\set{w\p{E+\Delta E}/w\p{E},1}} {\sum_{\Delta E'}\tilde
+    T_d\p{E,\Delta E'} \max\set{w\p{E+\Delta E'}/w\p{E},1}},
+  \label{eq:<de^2>}
+\end{align}
+in terms of which the diffusivity $\alpha\p{E}$ is
+\begin{align}
+  \alpha\p{E}=\bk{\Delta E^2}_E -\bk{\Delta E}_E^2.
+  \label{eq:microscopic_diffusitivy}
+\end{align}
+Diffusivity $\alpha\p{E}$ is thus the variance of the change in energy
+$\Delta E$ from as single move out of $E$. Reconciling this definition
+with that we provide in (\ref{eq:walker_current_definition}) is
+considered outside the scope of this paper.
+
+The hybrid OE-TMMC method thus first runs Algorithm \ref{alg:tmmc} to
+collect a transition histogram and compute flat histogram weights,
+then computes the diffusivity $\alpha\p{E}$ via
+(\ref{eq:<de>})-(\ref{eq:microscopic_diffusitivy}), and finally
+computes optimized weights as given by
+(\ref{eq:optimized_weights}). Unlike the optimized ensemble, OE-TMMC
+does not require any extra simulation routines, and is therefore
+straightforward to implement and compare to other histogram methods.
+\fixme{provide iterative algorithm?}
+
+\subsection{Loose ends}
+\label{sec:loose_ends}
+
+Throughout the discussion and development of histogram methods, we
+have thus far neglected three subjects: initialization end conditions
+(i.e. the means by which we decide when an algorithm to construct
+weights is finished), the energy range about which we care, and the
+minimum temperature $T_{\t{min}}$ which we sometimes make reference
+to. The last two of these subjects turn out to be related, as the
+minimum temperature $T_{\t{min}}$ chosen for any given simulation
+determines the minimum energy of interest.
+
+\subsubsection{End conditions}
+\label{sec:end_conditions}
+
+While the Wang-Landau method comes equipped with an end condition of
+its own, most histogram methods require some external means of
+deciding when they are finished initializing. Wang-Landau's
+``natural'' end condition does not advantage it over other methods,
+however, as the WL end condition makes use of a free parameter, which
+any end condition will require. In order to compare all methods
+fairly, we need a method-independent end condition. The most
+straightforward of such end conditions is to simply to stop
+initializing after some fixed number of Monte Carlo iterations, and
+appropriately finish up any other calculations in the algorithm before
+terminating it. We thus initialize systems of $N$ spheres for
+$10^5N^2$ iterations; the choice of $10^5N^2$ we will address in
+Section \ref{sec:results}\fixme{address in results section}. In
+general, however, one might want an end condition which is independent
+of system size. We came up with a four such end conditions as
+examples, though we do not use these end conditions in this paper.
+
+Let $E_+$ be the energy at which the density of states $D\p{E}$ and
+entropy $S\p{E}$ are maximal, and let $E_-$ be the minimum energy of
+interest. We define the pessimistic sample count $s_p\p{E_0}$ as the
+number of times which a system has traveled from $E_+$ to $E_0$, in
+any number of moves. This count represents the most pessimistic
+estimate of the number of independent, uncorrelated observations of
+the energy $E_0$, as going to the state of maximal entropy between
+samples wipes all correlations between them. We similarly define an
+optimistic sample count $s_o\p{E_0}$ as the number of times which the
+system has transitioned into $E_0$ from any energy $E>E_0$. This count
+represents the most optimistic estimate of the number of independent
+observations of the energy $E_0$, as the system had merely to go to
+any energy $E>E_0$, where the density of states may be only marginally
+larger, between each sample.
+
+In general, any initialization routine should sample the entire range
+$R=\sp{E_-,E_+}$ before terminating. Furthermore, samples at the end
+of initialization should be numerous enough to be statistically
+significant.\fixme{include mean sample error?} Given that any samples
+of $E_-$ necessarily require first sampling all other energies in $R$
+on the way down from $E_+$, this energy ($E_-$) will generally have
+the fewest number of samples. As an end condition, one can therefore
+enforce some minimum number of optimistic or pessimistic samples at
+$E_-$, i.e. $s_o\p{E_-}$ or $s_p\p{E_-}$. A more sophisticated end
+condition would be to enforce some minimum value to the mean
+fractional error $\bk{\delta_s}_{E,T_{\t{tmin}}}$ in either of the
+sample counts $s_o\p{E}$ or $s_p\p{E}$ at the minimum temperature
+$T_{\t{min}}$. These mean errors are
+\begin{align}
+  \bk{\delta_s}_{E,T_{\t{tmin}}}=\bk{s^{-1/2}}_{E,T_{\t{min}}}
+  =\sum_Es\p{E}^{-1/2}P\p{E,T_{\t{min}}}
+  =\f{\sum_Es\p{E}^{-1/2}D\p{E}e^{-E/kT_{\t{min}}}}
+  {\sum_{E'}D\p{E'}e^{-E'/kT_{\t{min}}}},
+\end{align}
+where we use the fact that the fractional error $\delta_G$ in any
+histogram $G$ collected via random Monte Carlo sampling is $G^{-1/2}$.
+
+\subsubsection{Determining the energy range of interest}
 \label{sec:energy_range}
+
+While some particular systems, such as the Ising model, have
+well-defined and easily determined minimum and maximum energies, the
+same is not true of all systems in general.
 
@@
 
-\subsubsection{State of maximal entropy}
-\label{sec:max_entropy}
-
 \fixme{The energy at which the density of states is maximal, i.e. the
@@
   in which appropriate weights are not obvious. Explain why.}
-
-\subsubsection{Minimum important energy}
-\label{sec:min_energy}
 
@@
   the minimum important energy after all initializations.}
-
-\subsubsection{Computing the density of states}
-\label{sec:dos}
-
-\fixme{This section might just become a part of the section on the
-  minimum important energy. (\ref{eq:dos}) gives us an explicit
-  formula for the density of states if we are confident in our energy
-  histogram. We can also compute the density of states from the
-  transition matrix, which is sometimes more appropriate to do.}
-
-\subsection{End conditions}
-\label{sec:end_conditions}
-
-% \fixme{Talk about how we want method-independent end conditions, if
-% possible. Introduce the concept of statistically independent energy
-% samples, and explain two measures (one optimistic, and one
-% pessimistic) for determining the number of statistically independent
-% times we have samples a given energy.}
-
-% \fixme{Talk about two end conditions: a fixed number of times which
-% we require to independently sample the minimum important energy, and
-% an enforced fractional error in the number of times we have
-% independently sampled any energy (at a given minimum temperature).}
 

### 2015-05-06 papers/histogram/thesis/thesis.tex
first completed draft of methods section

@@
 
+\fixme{Make all mention of what we do in present, rather than past
+  tense}
+
 \newpage
@@
 determine properties $X\p{T\ne T_0}$, and are thus referred to as
-``fixed temperature'' simulations.
+``fixed temperature'' simulations. Incidentally, Metropolis Monte
+Carlo simulations can be thought of as infinite-temperature
+simulations, as they are equivalent to simulations with canonical
+weights
+\begin{align}
+  w\p{E}=\lim_{T\to\infty}e^{-E/kT}=1.
+  \label{eq:inf_temp_weights}
+\end{align}
+This observation will be used in Section \ref{sec:min_energy} to
+identify a maximum energy of interest.
 
@@
 for portions of all weight arrays, which we discuss in Section
-\ref{sec:energy_range}.
+\ref{sec:min_energy}.
 
@@
 D\p{E}e^{-E/kT}$ dominates the sum in (\ref{eq:XT}). Section
-\ref{sec:energy_range} provides more discussion on this subject,
-namely on identifying the energies which are important for computing
-system properties at a given temperature.
+\ref{sec:min_energy} discusses identifying the energies which are
+important for computing system properties at a given temperature.
 
@@
 but all simulations in this paper require a choice of minimum
-temperature anyways (discussed in Section \ref{sec:energy_range}). In
+temperature anyways (discussed in Section \ref{sec:min_energy}). In
 this paper, we will always use $kT_{\t{min}}=0.2\epsilon$.
@@
 commonly suggested flatness condition is that the minimum histogram
-value be at least some fixed proportion of the mean histogram value,
-i.e. $C\sp{H}:\min\sp{H\p{E}}\ge x\bk{H}_E$, and several papers report
-using $x=0.95$. We used the same condition with $x=0.25$.
+value in some specified range of energies be at least some fixed
+proportion of the mean histogram value, i.e.
+$C\sp{H}:\min\sp{H\p{E}}\ge x\bk{H}_E$, and several papers report
+using $x=0.95$. We used the same condition with $x=0.25$. Determining
+the range of energies over which to evaluate $C\sp{H}$ is the subject
+of Section \ref{sec:min_energy}.
 
@@
 At equilibrium with a flat histogram, meaning $n$ is constant, the
-down-going walker current $j\p{E}$ is defined in terms of the
+down-going walker current $j\p{E}$ can be expressed in terms of the
 diffusivity $\alpha\p{E}$ by
@@
 \end{align}
-where a constant $n$ means $dn/dE=0$. We can rearrange
+where a constant $n$ implies $dn/dE=0$. We can rearrange
 (\ref{eq:walker_current}) and integrate over the energy range as
@@
 Throughout the discussion and development of histogram methods, we
-have thus far neglected three subjects: initialization end conditions
-(i.e. the means by which we decide when an algorithm to construct
-weights is finished), the energy range about which we care, and the
-minimum temperature $T_{\t{min}}$ which we sometimes make reference
-to. The last two of these subjects turn out to be related, as the
+have thus far neglected a few important subjects:
+\begin{enumerate*}
+\item initialization end conditions i.e. the means by which we decide
+  when an algorithm to construct weights is finished,
+\item the minimum temperature $T_{\t{min}}$ which we have sometimes
+  mentioned,
+\item the energy range over which we are interested in, and
+\item what to do with the weight array outside of the energy range of
+  interest, or at in which it is not defined.
+\end{enumerate*}
+The last three of these subjects turn out to be related, as the
 minimum temperature $T_{\t{min}}$ chosen for any given simulation
-determines the minimum energy of interest.
+determines the minimum energy $E_-$ of interest, as well as the
+weights $w\p{E}$ at energies $E<E_-$.
 
@@
 In general, any initialization routine should sample the entire range
-$R=\sp{E_-,E_+}$ before terminating. Furthermore, samples at the end
-of initialization should be numerous enough to be statistically
-significant.\fixme{include mean sample error?} Given that any samples
-of $E_-$ necessarily require first sampling all other energies in $R$
-on the way down from $E_+$, this energy ($E_-$) will generally have
-the fewest number of samples. As an end condition, one can therefore
-enforce some minimum number of optimistic or pessimistic samples at
-$E_-$, i.e. $s_o\p{E_-}$ or $s_p\p{E_-}$. A more sophisticated end
-condition would be to enforce some minimum value to the mean
-fractional error $\bk{\delta_s}_{E,T_{\t{tmin}}}$ in either of the
-sample counts $s_o\p{E}$ or $s_p\p{E}$ at the minimum temperature
-$T_{\t{min}}$. These mean errors are
-\begin{align}
-  \bk{\delta_s}_{E,T_{\t{tmin}}}=\bk{s^{-1/2}}_{E,T_{\t{min}}}
-  =\sum_Es\p{E}^{-1/2}P\p{E,T_{\t{min}}}
-  =\f{\sum_Es\p{E}^{-1/2}D\p{E}e^{-E/kT_{\t{min}}}}
-  {\sum_{E'}D\p{E'}e^{-E'/kT_{\t{min}}}},
+$R=\sp{E_-,E_+}$ of interest for simulation before terminating.
+Furthermore, samples at the end of initialization should be numerous
+enough to be statistically significant. Given that any samples of
+$E_-$ necessarily require first sampling all other energies in $R$ on
+the way down from $E_+$, this energy ($E_-$) will generally have the
+fewest number of samples. As an end condition, therefore, one can
+simply enforce some minimum number of optimistic or pessimistic
+samples at $E_-$, i.e. $s_o\p{E_-}$ or $s_p\p{E_-}$. A more
+sophisticated end condition would be to choose a minimum temperature
+of interest $T_{\t{min}}$ and initialize until the mean fractional
+error $\bk{\delta_s}_{E,T_{\t{tmin}}}$ in either of the sample counts
+$s_o\p{E}$ or $s_p\p{E}$ reaches some minimum value. These mean errors
+are
+\begin{align}
+  \bk{\delta_s}_{E,T}=\bk{s^{-1/2}}_{E,T}=\sum_Es\p{E}^{-1/2}P\p{E,T}
+  =\f{\sum_Es\p{E}^{-1/2}D\p{E}e^{-E/kT}}{\sum_{E'}D\p{E'}e^{-E'/kT}},
+  \label{eq:fractional_sample_error}
 \end{align}
@@
 
-\subsubsection{Determining the energy range of interest}
-\label{sec:energy_range}
-
-While some particular systems, such as the Ising model, have
-well-defined and easily determined minimum and maximum energies, the
-same is not true of all systems in general.
-
-Histogram methods typically rely on knowledge of the minimum and
-maximum energies of the system in question \fixme{how?}, which are
-unknown for the square-well fluid. Worse still, some square-well fluid
-energies have such an incredibly low density of states that one cannot
-reasonably expect to ever observe them via Monte Carlo, biased or
-otherwise. As such, we need to identify an energy range of interest,
-bounded by minimum and maximum ``important'' energies. \fixme{Rewrite,
-  expand}
-
-\fixme{The energy at which the density of states is maximal, i.e. the
-  state of maximal entropy, sets an upper bound for the energy range
-  in which appropriate weights are not obvious. Explain why.}
-
-\fixme{Determining the minimum important energy is considerably
-  trickier. Punchline: we have to choose a minimum temperature of
-  interest, $T_{\t{min}}$, and the minimum important energy is the
-  energy at which the slope of the density of states is
-  $\exp\p{kT_{\t{min}}/\epsilon}/\epsilon$. This condition for the
-  minimum important energy is directly related to canonical weights at
-  a temperature of $T_{\t{min}}$.}
-
-\fixme{Mention that in all initializations which simulate for some
-  time with unchanging weights (i.e. simple flat), the weight array
-  should be set to canonical weights below the minimum important
-  energy, as soon as it is identified.}
-
-\fixme{Mention that weights should be set to canonical weights below
-  the minimum important energy after all initializations.}
+\subsubsection{The energy range of interest}
+\label{sec:min_energy}
+
+Using histogram methods generally requires knowing the energy range of
+interest for the system at hand. The Wang-Landau method, for example,
+periodically checks a flatness condition $C\sp{H}$ on the energy
+histogram $H\p{E}$. In order for such a condition to be well-defined,
+it needs have a specified energy range over which to check the
+flatness of $H$. While some systems, such as the Ising model, have
+well-defined and easily computable energy range, the same is not true
+of most systems. In the case of the square-well fluid, not only is the
+minimum energy unknown, some energies have such an incredibly low
+state density $D\p{E}$ that one cannot reasonably expect to ever
+observe them via Monte Carlo, biased or otherwise. We therefore need
+to identify an energy range of interest, specified by a maximum and
+minimum energies of interest.
+
+Identifying a maximum energy of interest is straightforward: this is
+simply the energy $E_+$ at which the density of states $D\p{E}$ is
+maximal. Though high energies become more important for determining
+system properties $\bk{X}_T$ at higher temperatures, the relative
+importance of energies $E>E_+$ to the energy $E_+$ approaches a
+maximum as $T\to\infty$. There is therefore a maximum weight $w\p{E}$
+which is appropriate to assign energies $E>E_+$ relative to the weight
+$w\p{E_+}$, which is namely the ratio of these weights as
+$T\to\infty$. As observed in (\ref{eq:inf_temp_weights}), the weight
+array becomes flat as $T\to\infty$, which means that the ratio of
+$w\p{E>E_+}$ to $w\p{E}$ appropriate for any given temperature $T$
+maximally approaches unity. We can therefore always set
+$w\p{E>E_+}=w\p{E_+}$, which will ensure that energies $E>E_+$ are not
+sampled any more than is necessary for any positive temperature. The
+maximum of $D\p{E}$ can be found by simulating without weights for a
+short while, after which the energy histogram $H\p{E}$ should be
+maximal at $E_+$.
+
+The minimum important energy $E_-$ for a system is less obvious than
+$E_+$. Just as finding $E_+$ naturally required using a ``maximal''
+temperature $T\to\infty$, determining $E_-$ requires a choice of some
+minimum temperature $T_{\t{min}}$. The minimum important energy $E_-$
+is then the energy at which the partition function terms
+$D\p{E}e^{-E/kT_{\t{min}}}$ are maximal. Lower energy terms grow
+increasingly less important with decreasing energy, and are
+sufficiently sampled with canonical weights
+$w\p{E<E_-}=e^{-E/kT_{\t{min}}}$ to determine any property
+$\bk{X}_{T\ge T_{\t{min}}}$. To maximize $D\p{E}e^{-E/kT_{\t{min}}}$,
+we need
+\begin{align}
+  \f{d}{dE}\p{De^{-E/kT_{\t{min}}}}
+  =\f{dD}{dE}e^{-E/kT_{\t{min}}}+De^{-E/kT_{\t{min}}}\p{-\f{1}{kT_{\t{min}}}}=0,
+  \label{eq:min_E_setup}
+\end{align}
+which means
+\begin{align}
+  \f1{D}\f{dD}{dE}=\f{d\ln D}{dE}=\f{1}{kT_{\t{min}}}.
+  \label{eq:min_E_condition}
+\end{align}
+Though we do not always know the density of states $D\p{E}$, in
+practice we compute the best available estimate $D\p{E}$ every time we
+need to know $E_-$. If we do not find any energy satisfying
+(\ref{eq:min_E_condition}), we use the minimum energy the simulation
+has ever observed in place of $E_-$ to define our energy range of
+interest.
+
+Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
+in an initialization routine for any histogram method, we effectively
+forbid transitions to energies $E<E_-$ by tentatively setting weights
+$w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ for some temperature $T\ll
+T_{\t{min}}$. These weights prevent the initialization routine from
+wasting time on energies at which we will override the weight array
+after initialization anyways. A special case was made for Wang-Landau,
+which was designed assuming knowledge of the energy range of interest
+a priori. In order to fairly compare Wang-Landau to other methods, we
+therefore identified $E_-$ for each simulated system using other
+methods, and hard-coded the minimum important energies for each system
+into the Wang-Landau initialization routines\fixme{talk about how bad
+  this is}. We always set $w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ after
+initialization of all histogram method.
 
@@
 
-\subsection{Implementation and overview}
-\label{sec:implementation}
-
-\fixme{Talk about the relative difficulty of implementing each method,
-  as well as the relative difficulty of optimizing or tuning each
-  method. Free parameters are generally bad, and especially so if they
-  come without ``obvious'' or prescribed values. Justify our
-  modifications to Wang-Landau by showing that we did not make it
-  worse. Include mention of how memory requirements for transition
-  matrix Monte Carlo scale poorly with system size, as well as energy
-  resolution for systems with continuous energies.}
-
-\subsection{Simulation results and errors}
-\label{sec:results_and_errors}
-
-\fixme{Talk about the energy histogram, weights, and density of states
-  yielded by the methods. These should all be fairly similar. Talk
-  about why this is so.}
+\subsection{Histograms and arrays}
+
+\fixme{Give examples of energy histograms for one system, density of
+  states for several $N$, and perhaps a weight array.}
 
@@
   \centering
+  \caption[Density of states]{Density of states}
+  \label{fig:density_of_states}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
   \caption[Energy weights]{Energy weights}
@@
 
-\begin{figure}[H]
-  \centering
-  \caption[Normalized densities of states]{Normalized densities of
-    states}
-  \label{fig:density_of_states}
-\end{figure}
+\subsection{Sample rates}
 
@@
   size}
-
-\begin{figure}[H]
-  \centering
-  \caption[Initialization iterations vs. system size]{Initialization
-    iterations vs. system size}
-  \label{fig:scaling}
-\end{figure}
-
-\fixme{Discuss energy initialization times, sampling rates, and errors
-  in computed thermodynamic properties. Explain that errors are given
-  relative to a simulation which ran much longer, and talk about how
-  we ran that simulation (which method did we use?). Show that all
-  methods converge on the same answer as they simulate for longer.}
 
@@
 
+\subsection{Errors in $u$, $c_V$, and $s_{\t{config}}$}
+
+\fixme{Discuss errors in computed thermodynamic properties. Introduce
+  Converged flat weights.}
+
 \begin{figure}[H]
   \centering
-  \caption[Initialization iterations vs. data quality]{Initialization
-    iterations vs. data quality}
-  \label{fig:quality}
+  \caption[Internal energy]{Specific internal energy}
+  \label{fig:u_error}
 \end{figure}
@@
   \centering
-  \caption[Specific internal energy]{Specific internal energy}
-  \label{fig:internal_energy}
+  \caption[Heat capacity]{Specific heat capacity}
+  \label{fig:cv_error}
 \end{figure}
@@
   \centering
-  \caption[Specific heat capacity]{Specific heat capacity}
-  \label{fig:heat_capacity}
+  \caption[Configurational entropy]{Specific configurational
+    entropy}
+  \label{fig:s_error}
 \end{figure}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\nocite{*} \bibliography{thesis}
+\bibliography{thesis}
 

### 2015-05-06 papers/histogram/thesis/thesis.tex

@@
 those methods which are most efficient and effective at sampling
-otherwise improbable system states.
+otherwise improbable system states.\fixme{make this an abstract,
+  rather than a proposal...}
 
@@
 
-\fixme{Make all mention of what we do in present, rather than past
-  tense}
+\fixme{Always talk about what we do in our own code in present, rather
+  than past tense}
 
@@
 
-\fixme{Give an overview of the methods section}
-
-\fixme{Make a note on storing the logarithm of weights and
-  periodically rescaling them (resetting the zero on $\ln w$) to
-  reduce numerical error.}
+Having covered the basics of the square-well fluid as well as
+Metropolis and biased Monte Carlo simulations, we can now discuss
+several histogram methods for constructing energy weights.\fixme{more
+  overview of the methods section?}
 
@@
 after initialization anyways. A special case was made for Wang-Landau,
-which was designed assuming knowledge of the energy range of interest
-a priori. In order to fairly compare Wang-Landau to other methods, we
-therefore identified $E_-$ for each simulated system using other
-methods, and hard-coded the minimum important energies for each system
-into the Wang-Landau initialization routines\fixme{talk about how bad
-  this is}. We always set $w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ after
-initialization of all histogram method.
+as the design of this method assumes knowledge of the energy range of
+interest a priori. In order to fairly compare Wang-Landau to other
+methods, we therefore identified $E_-$ for each simulated system using
+other methods, and hard-coded the minimum important energies for each
+system into the Wang-Landau initialization routines\fixme{talk about
+  how bad this is}. We always set
+$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
+histogram methods with $kT_{\t{min}}=0.2\epsilon$. We chose this
+temperature because it seems to be right below the temperature at
+which the square-well fluid condenses from a gas into a
+liquid (see Section \ref{sec:errors}). \fixme{what about pressure?}
 
@@
 \subsection{Histograms and arrays}
+\label{sec:histograms}
 
@@
 
-\subsection{Sample rates}
-
-\fixme{Talk about initialization times, and how they scale with system
-  size}
+\subsection{Sampling rates}
+\label{sec:sampling_rates}
 
@@
 \subsection{Errors in $u$, $c_V$, and $s_{\t{config}}$}
+\label{sec:errors}
 
 \fixme{Discuss errors in computed thermodynamic properties. Introduce
-  Converged flat weights.}
+  converged flat weights.}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\bibliography{thesis}
+\nocite{*} \bibliography{thesis}
 

### 2015-05-07 papers/histogram/thesis/thesis.tex
added a sample figure and an overview of the methods section

@@
   \includegraphics[width=0.6\textwidth]{figs/dos-sample.pdf}
-  \caption[Sample density of states]{Sample density of states for
-    square-well fluid with a well width $\lambda=1.3$, filling
-    fraction $\eta=0.3$, and $N=20$ spheres, as computed by several
-    histogram methods. The density of states has been normalized to
-    have a maximum value of $1$.}
+  \caption[Density of states]{Sample density of states for square-well
+    fluid with a well width $\lambda=1.3$, filling fraction
+    $\eta=0.3$, and $N=20$ spheres, as computed by several histogram
+    methods. The density of states has been normalized to have a
+    maximum value of $1$.}
   \label{fig:dos_sample}
@@
 Metropolis and biased Monte Carlo simulations, we can now discuss
-several histogram methods for constructing energy weights.\fixme{more
-  overview of the methods section?}
+several histogram methods for constructing energy weights. In Section
+\ref{sec:flat_histogram}, we will introduce the flat histogram method,
+which is not directly applicable to the study of most systems, but
+which forms the basis of many other methods. We will then introduce,
+in Section \ref{sec:simple_flat}, the ``simple flat'' method, which we
+designed to be the simplest viable histogram method, and against which
+we can compare more complicated methods. Next, in Sections
+\ref{sec:wang_landau}--\ref{sec:optimized_ensemble} we introduce three
+published methods: Wang-Landau, transition matrix Monte Carlo, and the
+optimized ensemble. The last of these methods we will not actually
+implement and analyze because, unlike most histogram methods, it was
+designed to optimize weights rather than construct them. We will,
+however, borrow ideas from the optimized ensemble to design a
+histogram method of our own, a hybrid of the optimized ensemble and
+transition matrix Monte Carlo, in Section \ref{sec:oetmmc}. Finally,
+we will cover some general implementation details and loose ends in
+Section \ref{sec:loose_ends}
 
@@
 
-\begin{algorithm}[!b]
+\begin{algorithm}[tb]
   \caption{The simple flat method}
@@
 
-\fixme{find more sources to cite for some of the claims made in this
-  section}
-
 The Wang-Landau method is the first published histogram method we
@@
 \end{align}
-To compute $T$ from $\tilde T$, one must therefore normalize each row
+Computing $T$ from $\tilde T$ therefore requires normalizing each row
 of $\tilde T$ independently.
@@
 
-\begin{algorithm}[!b]
+\begin{algorithm}[tb]
   \caption{Transition matrix Monte Carlo initialization}
@@
 transition matrix $T_d\p{E,\Delta E}$ for a particular square-well
-fluid. The yellow band at $\Delta E=0$ indicates that most moves
-transition from an energy to itself. Low energies are more likely to
-transition up in energy, with $\Delta E>0$, than down in energy, with
-$\Delta E<0$. The high energies which are more likely to transition
-down in energy than up are those above the state of maximum entropy.
+fluid. The band at $\Delta E=0$ indicates that most moves transition
+from an energy to itself. Low energies are more likely to transition
+up in energy, with $\Delta E>0$, than down in energy, with $\Delta
+E<0$. The high energies which are more likely to transition down in
+energy than up are those above the state of maximum entropy.
 
@@
   \includegraphics[width=0.6\textwidth]{figs/transitions-sample.pdf}
-  \caption[Sample transition matrix]{Sample transition matrix
-    $T_d\p{E,\Delta E}$ for square-well fluid with a well width
-    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=20$ spheres.}
+  \caption[Transition matrix]{Sample transition matrix $T_d\p{E,\Delta
+      E}$ for square-well fluid with a well width $\lambda=1.3$,
+    filling fraction $\eta=0.3$, and $N=20$ spheres.  The band at
+    $\Delta E=0$ indicates that any given state is most likely to
+    transition into another state with the same energy.}
   \label{fig:transitions_sample}
@@
 histogram methods with $kT_{\t{min}}=0.2\epsilon$. We chose this
-temperature because it seems to be right below the temperature at
-which the square-well fluid condenses from a gas into a
-liquid (see Section \ref{sec:errors}). \fixme{what about pressure?}
+temperature because it is right below the temperature at which the
+square-well fluid condenses from a gas into a liquid (see Section
+\ref{sec:errors}). \fixme{what about pressure?}
 
@@
 
+Having developed and implemented all of the histogram methods
+discussed in Section \ref{sec:methods}, we will now look at the
+results from simulating with these methods.
+
 \subsection{Histograms and arrays}
@@
   \centering
-  \caption[Energy histograms]{Energy histograms}
-  \label{fig:histograms}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \caption[Density of states]{Density of states}
-  \label{fig:density_of_states}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \caption[Energy weights]{Energy weights}
-  \label{fig:weights}
+  \includegraphics[width=\textwidth]{figs/array-samples.pdf}
+  \caption[Sample arrays]{Sample arrays computed by various histogram
+    methods for a square-well fluid with a well width $\lambda=1.3$,
+    filling fraction $\eta=0.3$, and $N=20$ spheres after simulating
+    for $10^5N^2$ iterations: (top left) energy histogram, (top right)
+    weight array, (bottom left) density of states.}
+  \label{fig:arrays}
 \end{figure}

### 2015-05-07 papers/histogram/thesis/thesis.tex
changed to using example results from the more interesting N=25 simulations

@@
 have a maximum value of $1$ in order to emphasize the logarithmic span
-of $D\p{E}$. This logarithmic span ($\sim10^{45}$) is proportional to
+of $D\p{E}$. This logarithmic span ($\sim10^{60}$) is proportional to
 the number of moves which one must simulate via Metropolis Monte Carlo
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figs/dos-sample.pdf}
+  \includegraphics[width=0.6\textwidth]{figs/dos-example.pdf}
   \caption[Density of states]{Sample density of states for square-well
     fluid with a well width $\lambda=1.3$, filling fraction
-    $\eta=0.3$, and $N=20$ spheres, as computed by several histogram
+    $\eta=0.3$, and $N=25$ spheres, as computed by several histogram
     methods. The density of states has been normalized to have a
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figs/transitions-sample.pdf}
+  \includegraphics[width=0.6\textwidth]{figs/transitions-example.pdf}
   \caption[Transition matrix]{Sample transition matrix $T_d\p{E,\Delta
       E}$ for square-well fluid with a well width $\lambda=1.3$,
-    filling fraction $\eta=0.3$, and $N=20$ spheres.  The band at
+    filling fraction $\eta=0.3$, and $N=25$ spheres.  The band at
     $\Delta E=0$ indicates that any given state is most likely to
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/array-samples.pdf}
-  \caption[Sample arrays]{Sample arrays computed by various histogram
-    methods for a square-well fluid with a well width $\lambda=1.3$,
-    filling fraction $\eta=0.3$, and $N=20$ spheres after simulating
-    for $10^5N^2$ iterations: (top left) energy histogram, (top right)
-    weight array, (bottom left) density of states.}
+  \includegraphics[width=\textwidth]{figs/array-example.pdf}
+  \caption[Example arrays]{Example arrays computed by various
+    histogram methods for a square-well fluid with a well width
+    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres
+    after simulating for $10^5N^2$ iterations: (top left) energy
+    histogram, (top right) weight array, (bottom left) density of
+    states.}
   \label{fig:arrays}
@@
 
-\subsection{Sampling rates}
-\label{sec:sampling_rates}
-
 \begin{figure}[H]
   \centering
-  \caption[Optimistic energy sampling rates]{Optimistic energy
-    sampling rates}
-  \label{fig:opt_sample_rate}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \caption[Pessimistic energy sampling rates]{Pessimistic energy
-    sampling rates}
-  \label{fig:pes_sample_rate}
+  \includegraphics[width=0.75\textwidth]{figs/sample-rate-example.pdf}
+  \caption[Example energy sampling rates]{Example pessimistic energy
+    sampling rates computed by various histogram methods for a
+    square-well fluid with a well width $\lambda=1.3$, filling
+    fraction $\eta=0.3$, and $N=25$ spheres after simulating for
+    $10^5N^2$ iterations.}
+  \label{fig:sample_rates}
 \end{figure}

### 2015-05-07 papers/histogram/thesis/figs/plot-sample-rate-sample.py

@@ -1,0 +1,65 @@
+#!/usr/bin/python2
+import matplotlib, sys
+# matplotlib.use('Agg')
+import matplotlib.pyplot as plt
+import numpy
+import sys
+
+sys.path.append('../figs')
+import styles
+
+matplotlib.rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+matplotlib.rc('text', usetex=True)
+
+fig_size = numpy.array([6,5])
+
+if len(sys.argv) != 6:
+    print 'useage: %s ww ff N seed method' % sys.argv[0]
+    exit(1)
+
+ww = float(sys.argv[1])
+#arg ww = [1.3]
+
+ff = float(sys.argv[2])
+#arg ff = [0.3]
+
+N = int(sys.argv[3])
+#arg N = [20]
+
+seed = int(sys.argv[4])
+#arg seed = [0]
+
+methods = eval(sys.argv[5])
+#arg methods = [["nw","simple_flat","vanilla_wang_landau","wang_landau","tmmc","oetmmc","cfw"]]
+
+# input: ["../data/s%03d/periodic-ww%04.2f-ff%04.2f-N%i-%s-%s.dat" % (seed, ww, ff, N, method, data) for method in methods for data in ["ps","os"]]
+
+plt.figure('ps',figsize=fig_size)
+ax_ps = plt.subplot(111)
+
+for method in methods:
+    data_file = "../data/s%03d/periodic-ww%04.2f-ff%04.2f-N%i-%s-ps.dat" \
+                % (seed, ww, ff, N, method)
+    with open(data_file,'r') as file_handle:
+        for line in file_handle:
+            entries = line.split()
+            if 'iterations:' in entries:
+                iterations = int(entries[entries.index('iterations:')+1].replace(',',''))
+                continue
+    data = numpy.loadtxt(data_file)
+    energy = -data[:,0]/N
+    round_trips = data[:,1]
+    energy = energy[round_trips != 0]
+    round_trips = round_trips[round_trips != 0]
+    if sum(round_trips) > 0:
+        plt.semilogy(energy, iterations/round_trips, styles.dots(method),
+                     label=styles.title(method))
+
+plt.xlabel('$E/N\epsilon$')
+plt.ylabel('Iterations per pessimistic sample')
+plt.legend(loc='upper right')
+plt.tight_layout(pad=0.2)
+ax_ps.legend(bbox_to_anchor=(1.16,1))
+plt.savefig("figs/sample-rate-sample.pdf")
+
+plt.show()

### 2015-05-07 papers/histogram/thesis/figs/plot-transitions-example.py

@@ -21,7 +21,7 @@ ff = float(sys.argv[2])
 #arg ff = [0.3]
 
 N = int(sys.argv[3])
-#arg N = [20]
+#arg N = [25]
 
 method = sys.argv[4]
 #arg method = ['tmmc-golden']
@@ -68,4 +68,4 @@ plt.xlabel('$\Delta E/\epsilon$')
 plt.ylabel('$E/\epsilon$')
 
 plt.tight_layout()
-plt.savefig("figs/transitions-sample.pdf")
+plt.savefig("figs/transitions-example.pdf")

### 2015-05-07 papers/histogram/thesis/thesis.tex
fixed N=25 nw data and added u, hs, and s figures to thesis
some tweaks to figures

@@
   \includegraphics[width=0.47\textwidth]{figs/square-well.pdf}
-  \caption[The square-well pair potential]{The square-well pair
-    potential can be used to model short range forces to first
-    order. The infinite potential at $r<\sigma$ simply enforces the
-    condition that spheres cannot overlap. At $r>\lambda\sigma$, the
-    potential is null, and makes no contribution to the fluid's
-    internal energy.}
+  \caption[The square-well pair potential]
+  {The square-well pair potential can be used to model short range
+    forces to first order. The infinite potential at $r<\sigma$ simply
+    enforces the condition that spheres cannot overlap. At
+    $r>\lambda\sigma$, the potential is null, and makes no
+    contribution to the fluid's internal energy.}
   \label{fig:pair_potential}
@@
   \includegraphics[width=0.6\textwidth]{figs/dos-example.pdf}
-  \caption[Density of states]{Sample density of states for square-well
-    fluid with a well width $\lambda=1.3$, filling fraction
-    $\eta=0.3$, and $N=25$ spheres, as computed by several histogram
-    methods. The density of states has been normalized to have a
-    maximum value of $1$.}
+  \caption[Density of states]
+  {Sample density of states for square-well fluid with a well width
+    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres, as
+    computed by several histogram methods. The density of states has
+    been normalized to have a maximum value of $1$.}
   \label{fig:dos_sample}
@@
   \includegraphics[width=0.6\textwidth]{figs/transitions-example.pdf}
-  \caption[Transition matrix]{Sample transition matrix $T_d\p{E,\Delta
-      E}$ for square-well fluid with a well width $\lambda=1.3$,
-    filling fraction $\eta=0.3$, and $N=25$ spheres.  The band at
-    $\Delta E=0$ indicates that any given state is most likely to
-    transition into another state with the same energy.}
+  \caption[Transition matrix]
+  {Sample transition matrix $T_d\p{E,\Delta E}$ for square-well fluid
+    with a well width $\lambda=1.3$, filling fraction $\eta=0.3$, and
+    $N=25$ spheres.  The band at $\Delta E=0$ indicates that any given
+    state is most likely to transition into another state with the
+    same energy.}
   \label{fig:transitions_sample}
@@
   \includegraphics[width=\textwidth]{figs/array-example.pdf}
-  \caption[Example arrays]{Example arrays computed by various
+  \caption[Example arrays]
+  {Example arrays computed by various histogram methods for a
+    square-well fluid with a well width $\lambda=1.3$, filling
+    fraction $\eta=0.3$, and $N=25$ spheres after simulating for
+    $10^5N^2$ iterations: (top left) energy histogram, (top right)
+    weight array, (bottom left) density of states.}
+  \label{fig:arrays}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.8\textwidth]{figs/sample-rate-example.pdf}
+  \caption[Example energy sampling rates]
+  {Example pessimistic energy sampling rates computed by various
     histogram methods for a square-well fluid with a well width
     $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres
-    after simulating for $10^5N^2$ iterations: (top left) energy
-    histogram, (top right) weight array, (bottom left) density of
-    states.}
-  \label{fig:arrays}
+    after simulating for $10^5N^2$ iterations.}
+  \label{fig:sample_rates}
 \end{figure}
+
+\subsection{Thermodynamic properties and errors: $u$, $c_V$, and
+  $s_{\t{config}}$}
+\label{sec:errors}
+
+\fixme{Discuss errors in computed thermodynamic properties. Introduce
+  converged flat weights.}
 
@@
   \centering
-  \includegraphics[width=0.75\textwidth]{figs/sample-rate-example.pdf}
-  \caption[Example energy sampling rates]{Example pessimistic energy
-    sampling rates computed by various histogram methods for a
-    square-well fluid with a well width $\lambda=1.3$, filling
-    fraction $\eta=0.3$, and $N=25$ spheres after simulating for
-    $10^5N^2$ iterations.}
-  \label{fig:sample_rates}
+    \includegraphics[width=0.7\textwidth]{figs/internal-energy.pdf}
+  \caption[Internal energy]
+  {Specific internal energy}
+  \label{fig:internal_energy}
 \end{figure}
-
-\subsection{Errors in $u$, $c_V$, and $s_{\t{config}}$}
-\label{sec:errors}
-
-\fixme{Discuss errors in computed thermodynamic properties. Introduce
-  converged flat weights.}
 
@@
   \centering
-  \caption[Internal energy]{Specific internal energy}
-  \label{fig:u_error}
+  \includegraphics[width=0.75\textwidth]{figs/heat-capacity.pdf}
+  \caption[Heat capacity]
+  {Specific heat capacity}
+  \label{fig:heat_capacity}
 \end{figure}
@@
   \centering
-  \caption[Heat capacity]{Specific heat capacity}
-  \label{fig:cv_error}
+  \includegraphics[width=0.7\textwidth]{figs/config-entropy.pdf}
+  \caption[Configurational entropy]
+  {Specific configurational entropy}
+  \label{fig:config_entropy}
 \end{figure}
@@
   \centering
-  \caption[Configurational entropy]{Specific configurational
-    entropy}
-  \label{fig:s_error}
+  \includegraphics[width=0.7\textwidth]
+  {figs/internal-energy-err.pdf}
+  \caption[Errors in internal energy]
+  {Errors in specific internal energy}
+  \label{fig:internal_energy_err}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/heat-capacity-err.pdf}
+  \caption[Errors in heat capacity]
+  {Errors in specific heat capacity}
+  \label{fig:heat_capacity_err}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/config-entropy-err.pdf}
+  \caption[Errors in configurational entropy]
+  {Errors in specific configurational entropy}
+  \label{fig:config_entropy_err}
 \end{figure}

### 2015-05-07 papers/histogram/thesis/thesis.tex
more figure tweaks

@@
 
-\subsection{Thermodynamic properties and errors: $u$, $c_V$, and
-  $s_{\t{config}}$}
+\subsection{Thermodynamic properties and errors}
 \label{sec:errors}
@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\textwidth]
+  {figs/internal-energy-comps.pdf}
+  \caption[Internal energy error comparisons]
+  {Internal energy error comparisons}
+  \label{fig:u_comps}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\textwidth]
+  {figs/heat-capacity-comps.pdf}
+  \caption[Heat capacity error comparisons]
+  {Heat capacity error comparisons}
+  \label{fig:u_comps}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\textwidth]
+  {figs/config-entropy-comps.pdf}
+  \caption[Configurational entropy error comparisons]
+  {Configurational entropy error comparisons}
+  \label{fig:u_comps}
+\end{figure}
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2015-05-07 papers/histogram/thesis/figs/plot-uhc.py

@@ -1,11 +1,8 @@
-#!/usr/bin/python2
+#!/usr/bin/env python2
 import matplotlib, sys, os
 matplotlib.use('Agg')
 import matplotlib.pyplot as plt
 import numpy
-
-matplotlib.rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
-matplotlib.rc('text', usetex=True)
 
 sys.path.append('../figs')
 import styles
@@ -13,6 +10,10 @@ import readandcompute
 
 thesis_dir = os.getcwd()
 os.chdir('../')
+
+matplotlib.rc('font', **{'family': 'serif', 'serif': ['Computer Modern']})
+matplotlib.rc('text', usetex=True)
+
 
 if len(sys.argv) != 6:
     print 'useage: %s ww ff N methods seed' % sys.argv[0]

### 2015-05-07 papers/histogram/thesis/thesis.tex
fix build issue in histogram/thesis Tested on bennet. Took 0:27.63.

@@
   \centering
-  \includegraphics[width=0.75\textwidth]
-  {figs/internal-energy-comps.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/internal-energy-comps.pdf}
   \caption[Internal energy error comparisons]
@@
   \centering
-  \includegraphics[width=0.75\textwidth]
-  {figs/heat-capacity-comps.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/heat-capacity-comps.pdf}
   \caption[Heat capacity error comparisons]
@@
   \centering
-  \includegraphics[width=0.75\textwidth]
-  {figs/config-entropy-comps.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/config-entropy-comps.pdf}
   \caption[Configurational entropy error comparisons]

### 2015-05-08 papers/histogram/thesis/thesis.tex

@@
 finding the eigenvector of the transition matrix which has a
-corresponding unit eigenvalue\fixme{why is this eigenvector
-  unique?}. Once one has determined the density of states of a system,
-one can initialize weights via the flat histogram method,
+corresponding unit eigenvalue. \fixme{why is this eigenvector unique?}
+Once one has determined the density of states of a system, one can
+initialize weights via the flat histogram method,
 i.e. (\ref{eq:flat_weights}).
@@
 transition matrix $T$, we use Algorithm \ref{alg:tmmc}, which is a
-modified version of a routine provided in \cite{tmmc_prl}\fixme{is
-  this citation correct?}. The primary motivation behind this
-algorithm is such: as lower energies are always more difficult to
-sample than higher energies, transitions to lower energies should
-always be accepted, while transitions to higher energies should pass a
+modified version of a routine provided in \cite{tmmc_prl}. \fixme{is
+  this citation correct?} The primary motivation behind this algorithm
+is such: as lower energies are always more difficult to sample than
+higher energies, transitions to lower energies should always be
+accepted, while transitions to higher energies should pass a
 probabilistic acceptance test. To understand the probability in step
@@
 $10^5N^2$ iterations; the choice of $10^5N^2$ we will address in
-Section \ref{sec:results}\fixme{address in results section}. In
+Section \ref{sec:results}. \fixme{address in results section} In
 general, however, one might want an end condition which is independent
@@
 other methods, and hard-coded the minimum important energies for each
-system into the Wang-Landau initialization routines\fixme{talk about
-  how bad this is}. We always set
+system into the Wang-Landau initialization routines. \fixme{talk about
+  how bad this is} We always set
 $w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
@@
 square-well fluid condenses from a gas into a liquid (see Section
-\ref{sec:errors}). \fixme{what about pressure?}
+\ref{sec:properties}). \fixme{what about pressure?}
 
@@
 discussed in Section \ref{sec:methods}, we will now look at the
-results from simulating with these methods.
+results from simulating with these methods. Due to time constraints,
+we simulated only one system: the well width $\lambda=1.3$ and filling
+fraction $\eta=0.3$ square-well fluid, for all sizes (i.e. number of
+spheres) $N\in\sp{5,25}$. We initialized histogram methods for
+$10^5N^2$ iterations ($10^5N^3$ moves), and simulated for just as long
+afterwards. To check for consistency in simulation results, we
+simulated each combination of histogram method and system size for 30
+different random number generator seeds.
 
@@
 \label{sec:histograms}
+
+Three arrays contain the raw data for all analysis performed in this
+paper:
 
@@
 
+\subsection{Sample rates}
+\label{sec:sample_rates}
+
 \begin{figure}[H]
@@
     $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres
-    after simulating for $10^5N^2$ iterations.}
+    after simulating for $10^5N^2$ iterations. Sample rates for each
+    method are averaged over 30 simulations with different random
+    number seeds.}
   \label{fig:sample_rates}
@@
 \subsection{Thermodynamic properties and errors}
-\label{sec:errors}
+\label{sec:properties}
 
@@
   converged flat weights.}
+
+\subsubsection{Example curves}
+\label{sec:property_examples}
 
@@
 
+\subsubsection{Error scaling and comparisons}
+\label{sec:errors}
+
 \begin{figure}[H]
   \centering
-  \includegraphics[width=0.75\textwidth]{figs/internal-energy-comps.pdf}
+    \includegraphics[width=0.7\textwidth]{figs/u-scaling.pdf}
+  \caption[Internal energy error scaling]
+  {Specific internal energy error scaling}
+  \label{fig:u_scaling}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/cv-scaling.pdf}
+  \caption[Heat capacity error scaling]
+  {Specific heat capacity error scaling}
+  \label{fig:cv_scaling}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/s-scaling.pdf}
+  \caption[Configurational entropy error scaling]
+  {Specific configurational entropy error scaling}
+  \label{fig:s_scaling}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\textwidth]
+  {figs/internal-energy-comps.pdf}
   \caption[Internal energy error comparisons]

### 2015-05-08 papers/histogram/thesis/thesis.tex
removing wang landau from thesis for now

@@
   Z\p{T}=\sum_ED\p{E}e^{-E/kT}.
+  \label{eq:Z_norm}
 \end{align}
@@
   \tilde Z\p{T}=\sum_E\tilde D\p{E}e^{-E/kT},
+  \label{eq:Z}
 \end{align}
@@
 \end{align}
-and the value of a thermodynamic property $X\p{T_0}$
-\begin{align}
-  X\p{T_0}=\f1{\tilde Z\p{T_0}}\sum_E\bk{X}_E\tilde D\p{E}e^{-E/kT_0}
-  =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}},
+and the value of a thermodynamic property $\bk{X}_{T_0}$
+\begin{align}
+  \bk{X}_{T_0}=\f1{\tilde Z\p{T_0}}\sum_E\bk{X}_E\tilde
+  D\p{E}e^{-E/kT_0} =\f{\sum_E\bk{X}_EH\p{E}}{\sum_EH\p{E}},
   \label{eq:X_canonical}
@@
 temperatures. As a consequence, such simulations should not be used to
-determine properties $X\p{T\ne T_0}$, and are thus referred to as
+determine properties $\bk{X}_{T\ne T_0}$, and are thus referred to as
 ``fixed temperature'' simulations. Incidentally, Metropolis Monte
@@
 reporting a system-independent $n=10^3$\cite{wang_landau_mod} or
-$10^5$\cite{wang_landau}.  \fixme{find more citations for this last
-  claim}
+$10^5$\cite{wang_landau}.
 
@@
 simulations to take an exceedingly long time to converge. Lacking any
-heuristics for assigning $f_0$, literature typically suggests $f_0=e$
-\cite{wang_landau}. After some experimentation with the free
-parameters in the Wang-Landau method, we decided on $f_0=e^{1/8}$. In
-practice, one usually works with $\ln f$, making the commonly
-suggested value $\ln f_0=1$ and our own $\ln f_0=1/8$.
+heuristics for assigning $f_0$, we use $f_0=e$ (more precisely, $\ln
+f_0=1$), a value commonly suggested by literature\cite{wang_landau}.
 
@@
 values such as $u=2$ \cite{wang_landau, wang_landau_mod} and $u=10$
-\cite{wang_landau_analysis}; we use the former, conservatively
-favoring more initialization iterations by decreasing $f$ by a smaller
-amount on each reset.
+\cite{wang_landau_analysis}; we use the former value.
 
@@
 $C\sp{H}:\min\sp{H\p{E}}\ge x\bk{H}_E$, and several papers report
-using $x=0.95$. We used the same condition with $x=0.25$. Determining
-the range of energies over which to evaluate $C\sp{H}$ is the subject
-of Section \ref{sec:min_energy}.
+using $x=0.95$. We use this condition with the same value of
+$x$. Determining the range of energies over which to evaluate
+$C\sp{H}$ is the subject of Section \ref{sec:min_energy}.
 
@@
 other methods, and hard-coded the minimum important energies for each
-system into the Wang-Landau initialization routines. \fixme{talk about
-  how bad this is} We always set
+system into the Wang-Landau initialization routines\fixme{talk about
+  how bad this is}. We always set
 $w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
@@
 square-well fluid condenses from a gas into a liquid (see Section
-\ref{sec:properties}). \fixme{what about pressure?}
+\ref{sec:errors}). \fixme{what about pressure?}
 
@@
 results from simulating with these methods. Due to time constraints,
-we simulated only one system: the well width $\lambda=1.3$ and filling
-fraction $\eta=0.3$ square-well fluid, for all sizes (i.e. number of
-spheres) $N\in\sp{5,25}$. We initialized histogram methods for
+we simulated only one system, the square-well fluid with well width
+$\lambda=1.3$ and filling fraction $\eta=0.3$, for sizes (i.e. number
+of spheres) $N\in\sp{5,25}$. We initialized histogram methods for
 $10^5N^2$ iterations ($10^5N^3$ moves), and simulated for just as long
@@
 
+\subsection{Controls}
+\label{sec:controls}
+
+Our analysis will involve comparing errors in system properties
+computed from simulation results. As we do not know the true values of
+these system properties, we will compare standard simulation results
+against those from ``gold standard'' simulations, which run for
+significantly longer than the standard $10^5N^2$ iterations. For our
+gold standard simulations, we use the TMMC method and initialize until
+it collects $10^4$ optimistic samples of the minimum important energy
+(i.e. until $s_o\p{E_-}\ge 10^4$). To increase confidence in
+low-temperature gold standard results, we initialized them with
+$kT_{\t{min}}=0.1\epsilon$. After initialization, the gold standard
+runs simulate for $10^{10}$ iterations ($10^{10}N$ moves). While the
+entire initialization and simulation process takes around two hours
+for the largest regular simulations, the largest golden simulations
+initialize for $\sim11$ hours, and simulate for a week. To control for
+iteration number in the quality of results, we compare against a
+``converged flat'' simulation. Converged flat simulations use the
+golden (flat histogram) weight array, but simulates for the same
+number of iterations as the standard simulations.
+
 \subsection{Histograms and arrays}
@@
 
-Three arrays contain the raw data for all analysis performed in this
-paper:
-
-\fixme{Give examples of energy histograms for one system, density of
-  states for several $N$, and perhaps a weight array.}
-
-\begin{figure}[H]
+\begin{figure}[!b]
   \centering
@@
     $10^5N^2$ iterations: (top left) energy histogram, (top right)
-    weight array, (bottom left) density of states.}
+    weight array, (bottom left) density of states. Vertical reference
+    lines denote the minimum important energy $E_-$.}
   \label{fig:arrays}
@@
 
-\subsection{Sample rates}
-\label{sec:sample_rates}
-
-\begin{figure}[H]
+Most of our analysis will use quantities computed from two arrays
+given by simulations: the energy histogram $H\p{E}$ and the weights
+$w\p{E}$. To get a sense of what these arrays look like, Figure
+\ref{fig:arrays} shows the energy histogram and weights from an $N=25$
+simulation, along with the density of states computed via $\tilde
+D=H/w$ and normalized to have a maximum value of 1. The vertical
+reference line in these plots denotes the minimum important energy
+$E_-$. These plots demonstrate that, as expected, different methods
+can accumulate different histograms $H$ and weights $w$, yet agree on
+system properties such as the density of states $D$.
+
+Though high energy histogram counts at low energies mean that a
+simulation has collected many statistics on these energies, the energy
+histogram says nothing about how correlated different energy
+observations are. To determine how effective different simulations are
+at collecting uncorrelated statistics, we can look at the pessimistic
+sample rate $r_p\p{E}$, which is the mean number of simulation
+iterations required to sample a given energy $E$. This rate is given
+in terms of the number of iterations $I$ in a simulation (where we
+always have $I=10^5N^2$) and the pessimistic sample rate $s_p\p{E}$ by
+$r_p\p{E}=I/s_p\p{E}$. Figure \ref{fig:sample_rates} shows pessimistic
+sample rates for all methods averaged over 30 random number seeds for
+$N=25$ simulations. Lower pessimistic sample rates mean that a
+simulation takes less time to collect statistics which are guaranteed
+to be uncorrelated.
+
+\fixme{talk about plots}
+
+\begin{figure}[!b]
   \centering
-  \includegraphics[width=0.8\textwidth]{figs/sample-rate-example.pdf}
+  \includegraphics[width=0.8\textwidth]{figs/sample-rates.pdf}
   \caption[Example energy sampling rates]
   {Example pessimistic energy sampling rates computed by various
-    histogram methods for a square-well fluid with a well width
-    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres
-    after simulating for $10^5N^2$ iterations. Sample rates for each
-    method are averaged over 30 simulations with different random
-    number seeds.}
+    histogram methods. Sample rates for each method are averaged over
+    30 simulations with different random number seeds. The vertical
+    reference line denotes the minimum important energy $E_-$.}
   \label{fig:sample_rates}
@@
 
+\newpage
+
 \subsection{Thermodynamic properties and errors}
@@
 
-\fixme{Discuss errors in computed thermodynamic properties. Introduce
-  converged flat weights.}
-
-\subsubsection{Example curves}
-\label{sec:property_examples}
-
-\begin{figure}[H]
+We can use the formula in (\ref{eq:XT}) to compute the temperature
+dependence of a few system properties directly from the density of
+states, such as the internal energy
+\begin{align}
+  U=\bk{E}_T,
+  \label{eq:U}
+\end{align}
+the heat capacity
+\begin{align}
+  C_V\p{T}=\f{\partial U}{\partial T}
+  =\bk{\p{\f{E}{kT}}^2}_T-\bk{\f{E}{kT}}_T^2,
+  \label{eq:CV}
+\end{align}
+and the configurational entropy\fixme{how is this different from
+  regular entropy?}
+\begin{align}
+  S_{\t{config}}\p{T}=k\bk{\f{E}{kT}+\ln Z}_T.
+  \label{eq:S}
+\end{align}
+Figures \ref{fig:internal_energy}--\ref{fig:config_entropy} show these
+properties as computed via all methods for a single $N=25$ simulation
+each. When looking at these figures, one should remember that results
+are only trusted down to $kT=0.2\epsilon$ for regular simulations, and
+$kT=0.1\epsilon$ for golden ones. As with the density of states, all
+histogram methods agree fairly well on the physical properties of the
+square-well fluid. Though Figures
+\ref{fig:internal_energy}--\ref{fig:config_entropy} are interesting
+for studying the square-well fluid, they do not help us resolve the
+efficacy of different histogram methods. These figures do, however,
+show that the square-well fluid has a phase transition around
+$kT\approx0.5\epsilon$, which justifies our choice of
+$kT_{\t{min}}=0.2\epsilon$ and ensures that the histogram methods are
+performing the nontrivial task of exploring the liquid-gas
+phase boundary of the square-well fluid. For the purposes of this
+paper, we will neglect to demonstrate that this phase transition is
+indeed between liquid and gas, and take this fact as given.
+
+Having a sense of how $U$, $C_V$ and $S_{\t{config}}$ depend on
+temperature, we can now look at the how the errors $\Delta U$, $\Delta
+C_V$ and $\Delta S_{\t{config}}$ relative to the TMMC golden
+calculations. These are shown for a single $N=25$ simulation in
+Figures \ref{fig:internal_energy_err}--\ref{fig:config_entropy_err}.
+
+\begin{figure}[b]
   \centering
@@
   \caption[Internal energy]
-  {Specific internal energy}
+  {Specific internal energy curves from single $N=25$ simulations.}
   \label{fig:internal_energy}
@@
 
-\begin{figure}[H]
+\begin{figure}
   \centering
@@
   \caption[Heat capacity]
-  {Specific heat capacity}
+  {Specific heat capacity curves from single $N=25$ simulations.}
   \label{fig:heat_capacity}
@@
   \caption[Configurational entropy]
-  {Specific configurational entropy}
+  {Specific configurational entropy curves from single $N=25$
+    simulations.}
   \label{fig:config_entropy}
@@
   {figs/internal-energy-err.pdf}
-  \caption[Errors in internal energy]
-  {Errors in specific internal energy}
+  \caption[Internal energy error scaling]
+  {Errors in specific internal energy from single $N=25$ simulations.}
   \label{fig:internal_energy_err}
@@
 
-\begin{figure}[H]
+\begin{figure}
   \centering
   \includegraphics[width=0.7\textwidth]{figs/heat-capacity-err.pdf}
-  \caption[Errors in heat capacity]
-  {Errors in specific heat capacity}
+  \caption[Heat capacity error scaling]
+  {Errors in specific heat capacity from single $N=25$ simulations.}
   \label{fig:heat_capacity_err}
@@
 
-\begin{figure}[H]
+\begin{figure}
   \centering
   \includegraphics[width=0.7\textwidth]{figs/config-entropy-err.pdf}
-  \caption[Errors in configurational entropy]
-  {Errors in specific configurational entropy}
+  \caption[Configurational entropy error scaling]
+  {Errors in specific configurational entropy from single $N=25$
+    simulations.}
   \label{fig:config_entropy_err}
@@
 
-\subsubsection{Error scaling and comparisons}
-\label{sec:errors}
-
-\begin{figure}[H]
+\subsubsection{Error scaling}
+\label{sec:scaling}
+
+\begin{figure}
   \centering
@@
   \caption[Internal energy error scaling]
-  {Specific internal energy error scaling}
+  {Errors in internal energy as computed by various histogram methods,
+    each averaged over 30 simulations with different random number
+    seeds.}
   \label{fig:u_scaling}
@@
   \caption[Heat capacity error scaling]
-  {Specific heat capacity error scaling}
+  {Errors in specific heat capacity as computed by various histogram
+    methods, each averaged over 30 simulations with different random
+    number seeds.}
   \label{fig:cv_scaling}
@@
   \caption[Configurational entropy error scaling]
-  {Specific configurational entropy error scaling}
+  {Errors in specific configurational entropy as computed by various
+    histogram methods, each averaged over 30 simulations with
+    different random number seeds.}
   \label{fig:s_scaling}
@@
 
-\begin{figure}[H]
+\subsubsection{Error comparisons}
+\label{sec:errors}
+
+\begin{figure}
   \centering

### 2015-05-08 papers/histogram/thesis/thesis.tex
added back wang landau, and finished first complete draft of thesis

@@
 traditional models for testing and benchmarking such numerical
-techniques (e.g. the Ising model). We propose to implement these
-methods in Monte Carlo simulations of the square-well fluid, determine
-appropriate performance metrics for their comparison, and identify
-those methods which are most efficient and effective at sampling
-otherwise improbable system states.\fixme{make this an abstract,
-  rather than a proposal...}
-
-\fixme{Rewrite after the rest of the thesis is completed}
-
-\fixme{Always talk about what we do in our own code in present, rather
-  than past tense}
+techniques (e.g. the Ising model). We implement the Wang-Landau
+method, transition matrix Monte Carlo, as well as two methods of our
+own. The first of our methods is novel, while the second is a hybrid
+of two published methods (transition matrix Monte Carlo and the
+optimized ensemble).
+
+We find that Wang-Landau is capable of performing the best of all
+methods, but fails catastrophically without a predefined energy range
+for the square-well fluid. Our first method gives results comparable
+to successful Wang-Landau simulations, but with some anomalous
+cases. Our implementation of the optimized ensemble using a transition
+matrix results in worse performance over the straightforward
+transition matrix Monte Carlo method.
+
+\fixme{edit when I'm awake}
 
@@
 but all simulations in this paper require a choice of minimum
-temperature anyways (discussed in Section \ref{sec:min_energy}). In
-this paper, we will always use $kT_{\t{min}}=0.2\epsilon$.
+temperature anyways (discussed in Section \ref{sec:min_energy}).
 
@@
 wasting time on energies at which we will override the weight array
-after initialization anyways. A special case was made for Wang-Landau,
-as the design of this method assumes knowledge of the energy range of
-interest a priori. In order to fairly compare Wang-Landau to other
-methods, we therefore identified $E_-$ for each simulated system using
-other methods, and hard-coded the minimum important energies for each
-system into the Wang-Landau initialization routines\fixme{talk about
-  how bad this is}. We always set
+after initialization anyways. We always set
 $w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
-histogram methods with $kT_{\t{min}}=0.2\epsilon$. We chose this
+histogram methods with $kT_{\t{min}}/\epsilon=0.2$. We chose this
 temperature because it is right below the temperature at which the
 square-well fluid condenses from a gas into a liquid (see Section
-\ref{sec:errors}). \fixme{what about pressure?}
+\ref{sec:properties}). \fixme{what about pressure?}
+
+We made a special case for Wang-Landau, as the design of this method
+assumes knowledge of the energy range of interest a priori. In order
+to more fairly compare Wang-Landau to other methods, we therefore
+simulated two cases: one in which the minimum important energy $E_-$
+is identified via other methods and hard-coded into the Wang-Landau
+initialization routine, and one in which Wang-Landau simply assumes
+that the lowest energy a given simulation has ever observed is the
+minimum important energy. The simulations in which $E_-$ is hard-coded
+we will refer to as ``vanilla Wang-Landau'' simulations.
 
@@
 it collects $10^4$ optimistic samples of the minimum important energy
-(i.e. until $s_o\p{E_-}\ge 10^4$). To increase confidence in
-low-temperature gold standard results, we initialized them with
-$kT_{\t{min}}=0.1\epsilon$. After initialization, the gold standard
-runs simulate for $10^{10}$ iterations ($10^{10}N$ moves). While the
-entire initialization and simulation process takes around two hours
-for the largest regular simulations, the largest golden simulations
-initialize for $\sim11$ hours, and simulate for a week. To control for
-iteration number in the quality of results, we compare against a
-``converged flat'' simulation. Converged flat simulations use the
-golden (flat histogram) weight array, but simulates for the same
-number of iterations as the standard simulations.
+(i.e. until $s_o\p{E_-}\ge 10^4$). \fixme{why TMMC?} To increase
+confidence in low-temperature gold standard results, we initialized
+them with $kT_{\t{min}}/\epsilon=0.1$. After initialization, the gold
+standard runs simulate for $10^{10}$ iterations ($10^{10}N$
+moves). While the entire initialization and simulation process takes
+around two hours for the largest regular simulations, the largest
+golden simulations initialize for $\sim11$ hours, and simulate for a
+week. To control for iteration number in the quality of results, we
+compare against ``converged flat'' simulations. Converged flat
+simulations read in the transition matrix from golden simulations
+(which accumulates statistics in both during and after initialization)
+to construct a weight array, but simulate for the same number of
+iterations as the standard simulations.
 
@@
 
-\begin{figure}[!b]
+Most of our analysis will use quantities computed from two arrays
+given by simulations: the energy histogram $H\p{E}$ and the weights
+$w\p{E}$. To get a sense of what these arrays look like, Figure
+\ref{fig:arrays} shows the energy histogram and weights from an $N=25$
+simulation, along with the density of states computed via $\tilde
+D=H/w$ and normalized to have a maximum value of 1. The vertical
+reference line in these plots denotes the minimum important energy
+$E_-$. For reference, we also provide the histogram produced by an
+infinite temperature (no weights) simulation.
+
+The most immediate observation in these plots is the failure of
+Wang-Landau to produce meaningful results without a predefined energy
+range. This failure occurs because the Wang-Landau algorithm does not
+have a mechanism for properly setting and adjusting weights on newly
+observed energies. After several reset cycles of the Wang-Landau
+algorithm, the weight-modifying factor $f_k$ gets exceedingly small,
+causing Wang-Landau to get stuck at newly encountered low energies
+trying to fix their weights with a factor $f_k$ whose magnitude is
+appropriate for fine tuning, rather than heavy adjustment. If
+Wang-Landau knows the energy range of interest to begin with, it will
+not decrease $f_k$ until it has observed all important energies, and
+adjusted their weights to some reasonable value. Due to the drastic
+failure of Wang-Landau, will omit its results, which are nonsensical,
+from many figures.
+
+The energy histogram in Figure \ref{fig:arrays} shows that TMMC and
+OE-TMMC spend considerably less time at important low energies than
+the simple flat and vanilla Wang-Landau methods. At face value, this
+behavior of TMMC and OE-TMMC is undesirable.
+
+High energy histogram counts at low energies mean that a simulation
+has collected many statistics on these energies, but the energy
+histogram contains no information about how correlated different
+energy observations are. To determine how effective different
+simulations are at collecting uncorrelated statistics, we can look at
+the pessimistic sample rate $r_p\p{E}$, which is the mean number of
+simulation iterations required to sample a given energy $E$. This rate
+is given in terms of the number of iterations $I$ in a simulation
+(where we always have $I=10^5N^2$) and the pessimistic sample rate
+$s_p\p{E}$ by $r_p\p{E}=I/s_p\p{E}$. Figure \ref{fig:sample_rates}
+shows pessimistic sample rates for all methods except Wang-Landau
+averaged over 30 random number generator seeds for $N=25$
+simulations. Lower pessimistic sample rates mean that a simulation
+takes less time to collect statistics which are guaranteed to be
+uncorrelated. By this metric, TMMC and OE-TMMC again perform worse
+than simple flat and vanilla Wang-Landau.\fixme{comment on OETMMC
+  doing worse than TMMC}
+
+\begin{figure}
   \centering
@@
 
-Most of our analysis will use quantities computed from two arrays
-given by simulations: the energy histogram $H\p{E}$ and the weights
-$w\p{E}$. To get a sense of what these arrays look like, Figure
-\ref{fig:arrays} shows the energy histogram and weights from an $N=25$
-simulation, along with the density of states computed via $\tilde
-D=H/w$ and normalized to have a maximum value of 1. The vertical
-reference line in these plots denotes the minimum important energy
-$E_-$. These plots demonstrate that, as expected, different methods
-can accumulate different histograms $H$ and weights $w$, yet agree on
-system properties such as the density of states $D$.
-
-Though high energy histogram counts at low energies mean that a
-simulation has collected many statistics on these energies, the energy
-histogram says nothing about how correlated different energy
-observations are. To determine how effective different simulations are
-at collecting uncorrelated statistics, we can look at the pessimistic
-sample rate $r_p\p{E}$, which is the mean number of simulation
-iterations required to sample a given energy $E$. This rate is given
-in terms of the number of iterations $I$ in a simulation (where we
-always have $I=10^5N^2$) and the pessimistic sample rate $s_p\p{E}$ by
-$r_p\p{E}=I/s_p\p{E}$. Figure \ref{fig:sample_rates} shows pessimistic
-sample rates for all methods averaged over 30 random number seeds for
-$N=25$ simulations. Lower pessimistic sample rates mean that a
-simulation takes less time to collect statistics which are guaranteed
-to be uncorrelated.
-
-\fixme{talk about plots}
-
-\begin{figure}[!b]
+\begin{figure}
   \centering
@@
 
-\newpage
-
 \subsection{Thermodynamic properties and errors}
@@
 
-We can use the formula in (\ref{eq:XT}) to compute the temperature
-dependence of a few system properties directly from the density of
-states, such as the internal energy
+Histograms and sampling rates let us speculate about the quality of a
+simulations, but the most important task for any simulation is to
+yield accurate physical results. We therefore compute, via the formula
+in (\ref{eq:XT}), the temperature dependence of the square-well
+fluid's potential energy
 \begin{align}
@@
 \begin{align}
-  S_{\t{config}}\p{T}=k\bk{\f{E}{kT}+\ln Z}_T.
+  S_{\t{config}}\p{T}=k\bk{\f{E}{kT}+\ln Z}_T
+  -k\bk{\f{E}{kT}+\ln Z}_{T\to\infty}.
   \label{eq:S}
 \end{align}
-Figures \ref{fig:internal_energy}--\ref{fig:config_entropy} show these
-properties as computed via all methods for a single $N=25$ simulation
-each. When looking at these figures, one should remember that results
-are only trusted down to $kT=0.2\epsilon$ for regular simulations, and
-$kT=0.1\epsilon$ for golden ones. As with the density of states, all
-histogram methods agree fairly well on the physical properties of the
-square-well fluid. Though Figures
-\ref{fig:internal_energy}--\ref{fig:config_entropy} are interesting
-for studying the square-well fluid, they do not help us resolve the
-efficacy of different histogram methods. These figures do, however,
-show that the square-well fluid has a phase transition around
-$kT\approx0.5\epsilon$, which justifies our choice of
-$kT_{\t{min}}=0.2\epsilon$ and ensures that the histogram methods are
-performing the nontrivial task of exploring the liquid-gas
+All of these properties are directly computable from the density of
+states $D\p{E}$, and are shown in Figures
+\ref{fig:internal_energy}--\ref{fig:config_entropy} for $N=25$
+simulations, averaged over 30 random number generator seeds. When
+looking at these figures, one should remember that we are only
+confident in results down to $kT/\epsilon=0.2$ for regular
+simulations, and $kT/\epsilon=0.1$ for golden ones. With the exception
+of Wang-Landau, which has a heavy low-energy bias, all simulations
+agree fairly well on these properties of the square-well fluid.
+
+Though Figures \ref{fig:internal_energy}--\ref{fig:config_entropy} are
+interesting for studying the square-well fluid, they do not help us
+resolve the efficacy of different histogram methods. These figures do,
+however, show that the square-well fluid has a phase transition around
+$kT/\epsilon\approx0.5$, which justifies our choice of
+$kT_{\t{min}}/\epsilon=0.2$ and ensures that the histogram methods are
+indeed performing the nontrivial task of exploring the liquid-gas
 phase boundary of the square-well fluid. For the purposes of this
 paper, we will neglect to demonstrate that this phase transition is
-indeed between liquid and gas, and take this fact as given.
-
-Having a sense of how $U$, $C_V$ and $S_{\t{config}}$ depend on
-temperature, we can now look at the how the errors $\Delta U$, $\Delta
-C_V$ and $\Delta S_{\t{config}}$ relative to the TMMC golden
-calculations. These are shown for a single $N=25$ simulation in
-Figures \ref{fig:internal_energy_err}--\ref{fig:config_entropy_err}.
-
-\begin{figure}[b]
+indeed between liquid and gas phases, and take this fact as given.
+
+\begin{figure}[!b]
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/internal-energy.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/internal-energy.pdf}
   \caption[Internal energy]
-  {Specific internal energy curves from single $N=25$ simulations.}
+  {Specific internal energy curves averaged over 30 $N=25$
+    simulations.}
   \label{fig:internal_energy}
@@
   \caption[Heat capacity]
-  {Specific heat capacity curves from single $N=25$ simulations.}
+  {Specific heat capacity curves averaged over 30 $N=25$ simulations.}
   \label{fig:heat_capacity}
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/config-entropy.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/config-entropy.pdf}
   \caption[Configurational entropy]
-  {Specific configurational entropy curves from single $N=25$
+  {Specific configurational entropy curves averaged over 30 $N=25$
     simulations.}
@@
   \centering
-  \includegraphics[width=0.7\textwidth]
-  {figs/internal-energy-err.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/u-err.pdf}
   \caption[Internal energy error scaling]
-  {Errors in specific internal energy from single $N=25$ simulations.}
+  {Errors in specific internal energy, averaged over 30 $N=25$
+    simulations.}
   \label{fig:internal_energy_err}
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/heat-capacity-err.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/cv-err.pdf}
   \caption[Heat capacity error scaling]
-  {Errors in specific heat capacity from single $N=25$ simulations.}
+  {Errors in specific heat capacity, averaged over 30 $N=25$
+    simulations.}
   \label{fig:heat_capacity_err}
@@
   \centering
-  \includegraphics[width=0.7\textwidth]{figs/config-entropy-err.pdf}
+  \includegraphics[width=0.75\textwidth]{figs/s-err.pdf}
   \caption[Configurational entropy error scaling]
-  {Errors in specific configurational entropy from single $N=25$
+  {Errors in specific configurational entropy, averaged over 30 $N=25$
     simulations.}
@@
 
-\subsubsection{Error scaling}
-\label{sec:scaling}
-
-\begin{figure}
+Having a sense of how $U$, $C_V$ and $S_{\t{config}}$ behave with
+respect to temperature, we can now look at the how the errors $\Delta
+U$, $\Delta C_V$ and $\Delta S_{\t{config}}$ relative to the TMMC
+golden calculations for each method (except Wang-Landau). Figures
+\ref{fig:internal_energy_err}--\ref{fig:config_entropy_err} show these
+errors. Much like the arrays in Figure \ref{fig:arrays}, we provide
+these figures only for a sense of how error curves behave. Errors (at
+$kT/\epsilon\ge0.2$) typically peak near the phase transition at
+$kT/\epsilon\approx0.5$, and drop back down.
+
+Figures \ref{fig:u_scaling}--\ref{fig:s_scaling} provide a more
+comprehensive picture of errors from each method. These figures show
+the average maximum error in $U$, $C_V$, and $S$ for different system
+sizes. Though errors jump around considerably in these figures, the
+fact that they are roughly level between $N=15$ to $25$ justifies the
+choice $N^2$ scaling for our number of initialization iterations.
+Future studies of larger systems might use this scaling for initial
+estimates of initialization times.
+
+\begin{figure}[p]
   \centering
@@
   \caption[Internal energy error scaling]
-  {Errors in internal energy as computed by various histogram methods,
-    each averaged over 30 simulations with different random number
-    seeds.}
+  {Specific internal energy error scaling.}
   \label{fig:u_scaling}
@@
 
-\begin{figure}
+\begin{figure}[p]
   \centering
@@
   \caption[Heat capacity error scaling]
-  {Errors in specific heat capacity as computed by various histogram
-    methods, each averaged over 30 simulations with different random
-    number seeds.}
+  {Specific heat capacity error scaling.}
   \label{fig:cv_scaling}
@@
   \caption[Configurational entropy error scaling]
-  {Errors in specific configurational entropy as computed by various
-    histogram methods, each averaged over 30 simulations with
-    different random number seeds.}
+  {Specific configurational entropy error scaling.}
   \label{fig:s_scaling}
@@
 
-\subsubsection{Error comparisons}
-\label{sec:errors}
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/u-comps.pdf}
+  \caption[Internal energy error comparisons]
+  {Specific internal energy errors, plotted against those from
+    converged flat histogram weight simulations.}
+  \label{fig:u_comps}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.7\textwidth]{figs/cv-comps.pdf}
+  \caption[Heat capacity error comparisons]
+  {Specific heat capacity errors, plotted against those from converged
+    flat histogram weight simulations.}
+  \label{fig:cv_comps}
+\end{figure}
 
@@
   \centering
-  \includegraphics[width=0.75\textwidth]
-  {figs/internal-energy-comps.pdf}
-  \caption[Internal energy error comparisons]
-  {Internal energy error comparisons}
-  \label{fig:u_comps}
+  \includegraphics[width=0.7\textwidth]{figs/s-comps.pdf}
+  \caption[Configurational entropy error comparisons]
+  {Specific configurational entropy errors, plotted against those from
+    converged flat histogram weight simulations.}
+  \label{fig:s_comps}
 \end{figure}
 
-\begin{figure}
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/heat-capacity-comps.pdf}
-  \caption[Heat capacity error comparisons]
-  {Heat capacity error comparisons}
-  \label{fig:u_comps}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/config-entropy-comps.pdf}
-  \caption[Configurational entropy error comparisons]
-  {Configurational entropy error comparisons}
-  \label{fig:u_comps}
-\end{figure}
+It is more meaningful to compare errors against errors, rather than
+system size. \fixme{wording} Figures
+\ref{fig:u_comps}--\ref{fig:s_comps} show the errors of each method
+plotted against errors of the converged flat histogram weight
+simulations. The fact that some points fall below the diagonal
+reference line is strange, and indicates that some regular simulations
+are outperforming the converged flat ones. We do not understand why
+such events occur, as converged flat simulations should be using
+high-quality golden data to construct weight arrays, and should
+therefore outperform all standard simulations. These events are thus
+worthwhile looking into in more detail.
+
+Figures \ref{fig:u_scaling}--\ref{fig:s_comps} make clear, again, the
+inability of Wang-Landau to sample a system without a predefined
+energy range. These figures also show that the performances of simple
+flat method, vanilla Wang Landau, and TMMC are about the same, except
+for an anomalous case with $N=21$ in which simple flat yields
+simulation results with large errors. We do not know the reason for
+this anomalous case, which might be worth investigating. Finally,
+these error figures further confirm earlier suspicions of OE-TMMC's
+underperformance relative to TMMC, despite the fact that it is
+supposed to provide an improvement over the TMMC weight array.
 
@@
 
-\fixme{What have we learned? What recommendations can we make? When is
-  it appropriate to use which histogram methods?}
+While the results in this paper are not strongly conclusive, they may
+be used as guidelines for future studies of histogram methods,
+especially as applied to the square-well fluid. After implementing
+several histogram methods and applying them to the $\lambda=1.3$,
+$\eta=0.3$ square-well fluid, we found that Wang-Landau, together with
+a predefined energy range, consistently yields the smallest errors and
+highest low-energy sampling rates. Without prior knowledge of the
+minimum important energy of a system, however, Wang-Landau can get
+stuck at low energies, and is incapable of recovering. Wang-Landau is
+therefore most appropriate for studying systems in which the entire
+energy range is known a priori, or in which only a given energy range
+is of particular interest.
+
+The simple flat method and TMMC methods seem to perform roughly on par
+with vanilla Wang-Landau (i.e. Wang-Landau with a specified energy
+range), except an anomalous case of failure worth investigating on the
+part of simple flat. Our implementation of the optimized ensemble
+using a transition matrix, which is meant to improve the TMMC weight
+array, seemed to only make the weight array worse.
+
+\fixme{fix when I'm awake}
 

### 2015-05-09 papers/histogram/thesis/thesis.tex
made a fix in non-vanilla wang landau for fair comparison

@@
 \renewcommand{\title}{Optimizing Monte Carlo simulation of the
-  square-well fluid} \renewcommand{\author}{Michael A. Perlin}
+  square-well fluid}
+\renewcommand{\author}{Michael A. Perlin}
 \newcommand{\institution}{Oregon State University}
@@
 
-We find that Wang-Landau is capable of performing the best of all
-methods, but fails catastrophically without a predefined energy range
-for the square-well fluid. Our first method gives results comparable
-to successful Wang-Landau simulations, but with some anomalous
-cases. Our implementation of the optimized ensemble using a transition
-matrix results in worse performance over the straightforward
-transition matrix Monte Carlo method.
+We find that Wang-Landau potentially performs better than other
+histogram methods, but fails catastrophically without a predefined
+energy range for the square-well fluid. Our first method gives results
+comparable to successful Wang-Landau simulations, but can have
+anomalous cases with large error. The transition matrix Monte Carlo
+consistently yields results whose quality is marginally worse than
+functional Wang-Landau. Finally, our implementation of the optimized
+ensemble using a transition matrix results in worse performance over
+the straightforward transition matrix Monte Carlo method.
 
@@
 Once we identify an energy $E_-$ satisfying (\ref{eq:min_E_condition})
-in an initialization routine for any histogram method, we effectively
-forbid transitions to energies $E<E_-$ by tentatively setting weights
-$w\p{E<E_-}=e^{-\p{E-E_-}/kT}$ for some temperature $T\ll
-T_{\t{min}}$. These weights prevent the initialization routine from
-wasting time on energies at which we will override the weight array
-after initialization anyways. We always set
-$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ after initialization of all
-histogram methods with $kT_{\t{min}}/\epsilon=0.2$. We chose this
-temperature because it is right below the temperature at which the
+in an initialization routine for any histogram method, we set weights
+$w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ with a temperature
+$kT_{\t{min}}/\epsilon=0.2$. These weights prevent the initialization
+routine from wasting time on energies at which we will override the
+weight array after initialization anyways. We chose this temperature
+because it is comfortably below the temperature at which the
 square-well fluid condenses from a gas into a liquid (see Section
@@
 
-We made a special case for Wang-Landau, as the design of this method
-assumes knowledge of the energy range of interest a priori. In order
-to more fairly compare Wang-Landau to other methods, we therefore
-simulated two cases: one in which the minimum important energy $E_-$
-is identified via other methods and hard-coded into the Wang-Landau
-initialization routine, and one in which Wang-Landau simply assumes
-that the lowest energy a given simulation has ever observed is the
-minimum important energy. The simulations in which $E_-$ is hard-coded
-we will refer to as ``vanilla Wang-Landau'' simulations.
+We made a special case for Wang-Landau treatment of the energy range
+of interest in, as the design of this method assumes a priori
+knowledge of $E_-$. In order to more fairly compare Wang-Landau to
+other methods, we therefore also determined the appropriate energy
+$E_-$ for each system size $N$ using other histogram methods, and
+simulated via Wang-Landau with the a lower energy bound hard-coded. We
+will distinguish the two sets of Wang-Landau simulations by referring
+to those in which $E_-$ is hard-coded as ``vanilla Wang-Landau.''
 

### 2015-05-10 papers/histogram/thesis/thesis.tex

@@
 Throughout the discussion and development of histogram methods, we
-have thus far neglected a few important subjects:
-\begin{enumerate*}
+have thus far neglected expanding on a few important subjects:
+\begin{enumerate*}[label=\roman*)]
 \item initialization end conditions i.e. the means by which we decide
@@
 \item what to do with the weight array outside of the energy range of
-  interest, or at in which it is not defined.
+  interest, or at energies where it is not defined.
 \end{enumerate*}
@@
 ``natural'' end condition does not advantage it over other methods,
-however, as the WL end condition makes use of a free parameter, which
-any end condition will require. In order to compare all methods
-fairly, we need a method-independent end condition. The most
-straightforward of such end conditions is to simply to stop
-initializing after some fixed number of Monte Carlo iterations, and
-appropriately finish up any other calculations in the algorithm before
-terminating it. We thus initialize systems of $N$ spheres for
-$10^5N^2$ iterations; the choice of $10^5N^2$ we will address in
-Section \ref{sec:results}. \fixme{address in results section} In
-general, however, one might want an end condition which is independent
-of system size. We came up with a four such end conditions as
-examples, though we do not use these end conditions in this paper.
-
-Let $E_+$ be the energy at which the density of states $D\p{E}$ and
-entropy $S\p{E}$ are maximal, and let $E_-$ be the minimum energy of
-interest. We define the pessimistic sample count $s_p\p{E_0}$ as the
-number of times which a system has traveled from $E_+$ to $E_0$, in
-any number of moves. This count represents the most pessimistic
-estimate of the number of independent, uncorrelated observations of
-the energy $E_0$, as going to the state of maximal entropy between
-samples wipes all correlations between them. We similarly define an
-optimistic sample count $s_o\p{E_0}$ as the number of times which the
-system has transitioned into $E_0$ from any energy $E>E_0$. This count
-represents the most optimistic estimate of the number of independent
-observations of the energy $E_0$, as the system had merely to go to
-any energy $E>E_0$, where the density of states may be only marginally
-larger, between each sample.
+however, as the WL end condition makes use of a free parameter (the
+cutoff $c$), which any end condition will require. In order to compare
+all methods fairly, we need a method-independent end condition to
+initialization. The most straightforward of such end conditions is to
+simply to stop initializing after some fixed number of Monte Carlo
+iterations, and finish up any other calculations in the algorithm if
+appropriate. We thus initialize systems of $N$ spheres for $10^5N^2$
+iterations; the choice of $10^5N^2$ we address in Section
+\ref{sec:results}. In general, however, one might want an end
+condition which is independent of system size. We came up with a four
+such end conditions as examples, though we do not use them in
+simulations analyzed in this paper.
+
+Our end conditions require knowledge of the minimum and maximum
+energies of interest, which we denote respectively by $E_-$ and $E_+$.
+We define the pessimistic sample count $s_p\p{E_0}$ as the number of
+times which a system has traveled from $E_+$ to $E_0$. The trip from
+$E_+$ to $E_0$ may be made in any number of moves, but cannot include
+intermediate visits to $E_-$ or $E_0$. The count $s_p\p{E_0}$ provides
+the most pessimistic estimate for the number of independent,
+uncorrelated observations of the energy $E_0$, assuming that a visit
+to $E_+$ wipes all correlations between samples of any other energy;
+this assumption will be justified by our definition of $E_+$ as the
+energy at which the density of states $D\p{E}$, and by extension the
+entropy $S\p{E}$, is maximal. We similarly define an optimistic sample
+count $s_o\p{E_0}$ as the number of times which the system has
+transitioned into $E_0$ from any energy $E>E_0$ in a single move. This
+count represents an optimistic estimate of the number of independent
+observations of $E_0$, as the system had merely to go to any energy
+$E>E_0$, where the density of states may be only marginally larger,
+between each sample.
 
 In general, any initialization routine should sample the entire range
-$R=\sp{E_-,E_+}$ of interest for simulation before terminating.
-Furthermore, samples at the end of initialization should be numerous
-enough to be statistically significant. Given that any samples of
-$E_-$ necessarily require first sampling all other energies in $R$ on
+$R=\sp{E_-,E_+}$ of interest for a given simulation before
+terminating. Furthermore, samples at the end of initialization should
+be numerous enough to be statistically significant. Given that any
+samples of $E_-$ require first sampling all other energies in $R$ on
 the way down from $E_+$, this energy ($E_-$) will generally have the
@@
 simply enforce some minimum number of optimistic or pessimistic
-samples at $E_-$, i.e. $s_o\p{E_-}$ or $s_p\p{E_-}$. A more
-sophisticated end condition would be to choose a minimum temperature
-of interest $T_{\t{min}}$ and initialize until the mean fractional
-error $\bk{\delta_s}_{E,T_{\t{tmin}}}$ in either of the sample counts
-$s_o\p{E}$ or $s_p\p{E}$ reaches some minimum value. These mean errors
-are
+samples at $E_-$, i.e. minimum values of $s_o\p{E_-}$ or
+$s_p\p{E_-}$. A more sophisticated end condition would be to choose a
+minimum temperature of interest $T_{\t{min}}$ and initialize until the
+mean fractional error $\bk{\delta_s}_{E,T_{\t{tmin}}}$ in one of the
+sample counts reaches some minimum value. These mean errors are
 \begin{align}
   \bk{\delta_s}_{E,T}=\bk{s^{-1/2}}_{E,T}=\sum_Es\p{E}^{-1/2}P\p{E,T}
-  =\f{\sum_Es\p{E}^{-1/2}D\p{E}e^{-E/kT}}{\sum_{E'}D\p{E'}e^{-E'/kT}},
+  =\f{\sum_Es\p{E}^{-1/2}\tilde D\p{E}e^{-E/kT}}{\sum_{E'}\tilde
+    D\p{E'}e^{-E'/kT}},
   \label{eq:fractional_sample_error}
@@
 
-Using histogram methods generally requires knowing the energy range of
-interest for the system at hand. The Wang-Landau method, for example,
-periodically checks a flatness condition $C\sp{H}$ on the energy
-histogram $H\p{E}$. In order for such a condition to be well-defined,
-it needs have a specified energy range over which to check the
-flatness of $H$. While some systems, such as the Ising model, have
-well-defined and easily computable energy range, the same is not true
-of most systems. In the case of the square-well fluid, not only is the
-minimum energy unknown, some energies have such an incredibly low
-state density $D\p{E}$ that one cannot reasonably expect to ever
-observe them via Monte Carlo, biased or otherwise. We therefore need
-to identify an energy range of interest, specified by a maximum and
-minimum energies of interest.
-
-Identifying a maximum energy of interest is straightforward: this is
-simply the energy $E_+$ at which the density of states $D\p{E}$ is
+Using histogram methods generally requires specifying, whether
+explicitly or implicitly, the energy range of interest for a given
+simulation. The Wang-Landau method, for example, periodically checks a
+flatness condition $C\sp{H}$ on the energy histogram $H\p{E}$. In
+order for such a condition to be well-defined, it needs have a
+specified energy range over which to evaluate $C\sp{H}$. While some
+systems, such as the Ising model, have a well-defined and easily
+computable energy range, the same is not true of most systems. In the
+case of the square-well fluid, not only is the minimum energy unknown,
+but some energies have such an incredibly low state density $D\p{E}$
+that one cannot reasonably expect to ever observe them via Monte
+Carlo, biased or otherwise. We therefore need to identify an energy
+range of interest, bounded by a minimum $E_-$ and maximum $E_+$.
+
+Identifying a maximum energy $E_+$ of interest is straightforward:
+this is simply the energy at which the density of states $D\p{E}$ is
 maximal. Though high energies become more important for determining
 system properties $\bk{X}_T$ at higher temperatures, the relative
-importance of energies $E>E_+$ to the energy $E_+$ approaches a
-maximum as $T\to\infty$. There is therefore a maximum weight $w\p{E}$
-which is appropriate to assign energies $E>E_+$ relative to the weight
-$w\p{E_+}$, which is namely the ratio of these weights as
+importance of energies $E>E_+$ to that of the energy $E_+$ approaches
+a maximum as $T\to\infty$. There is therefore a maximum weight
+$w\p{E}$ which is appropriate to assign energies $E>E_+$ relative to
+the weight $w\p{E_+}$, which is given by the ratio of these weights as
 $T\to\infty$. As observed in (\ref{eq:inf_temp_weights}), the weight
@@
 The minimum important energy $E_-$ for a system is less obvious than
-$E_+$. Just as finding $E_+$ naturally required using a ``maximal''
+$E_+$. Just as determining $E_+$ involved reference to a ``maximal''
 temperature $T\to\infty$, determining $E_-$ requires a choice of some
 minimum temperature $T_{\t{min}}$. The minimum important energy $E_-$
-is then the energy at which the partition function terms
-$D\p{E}e^{-E/kT_{\t{min}}}$ are maximal. Lower energy terms grow
-increasingly less important with decreasing energy, and are
-sufficiently sampled with canonical weights
-$w\p{E<E_-}=e^{-E/kT_{\t{min}}}$ to determine any property
-$\bk{X}_{T\ge T_{\t{min}}}$. To maximize $D\p{E}e^{-E/kT_{\t{min}}}$,
-we need
+is then the energy at which the term $D\p{E}e^{-E/kT_{\t{min}}}$ in
+the partition function is maximal. Lower energy terms in the partition
+function grow increasingly less important as the distance
+$\abs{E-E_-}$ grows, and are sufficiently sampled with canonical
+weights $w\p{E<E_-}=w\p{E_-}e^{-E/kT_{\t{min}}}$ to determine any
+property $\bk{X}_{T\ge T_{\t{min}}}$. To maximize
+$D\p{E}e^{-E/kT_{\t{min}}}$, we need
 \begin{align}
@@
 Though we do not always know the density of states $D\p{E}$, in
-practice we compute the best available estimate $D\p{E}$ every time we
-need to know $E_-$. If we do not find any energy satisfying
+practice we compute the best available estimate of $D\p{E}$ every time
+we need to know $E_-$. If we do not find any energy satisfying
 (\ref{eq:min_E_condition}), we use the minimum energy the simulation
@@
 $w\p{E<E_-}=e^{-\p{E-E_-}/kT_{\t{min}}}$ with a temperature
-$kT_{\t{min}}/\epsilon=0.2$. These weights prevent the initialization
+$kT_{\t{min}}=0.2\epsilon$. These weights prevent the initialization
 routine from wasting time on energies at which we will override the
@@
 We made a special case for Wang-Landau treatment of the energy range
-of interest in, as the design of this method assumes a priori
-knowledge of $E_-$. In order to more fairly compare Wang-Landau to
-other methods, we therefore also determined the appropriate energy
-$E_-$ for each system size $N$ using other histogram methods, and
-simulated via Wang-Landau with the a lower energy bound hard-coded. We
-will distinguish the two sets of Wang-Landau simulations by referring
-to those in which $E_-$ is hard-coded as ``vanilla Wang-Landau.''
+of interest, as the design of this method assumes knowledge of the
+energy range over which to evaluate $C\sp{H}$. In order to more fairly
+compare Wang-Landau to other methods, as well as having Wang-Landau
+determine $E_-$ in the manner described above, we also simulated via
+Wang-Landau with $E_-$ hard-coded for each system size $N$. We will
+distinguish the two sets of Wang-Landau simulations by referring to
+those in which $E_-$ is hard-coded as ``vanilla Wang-Landau.''
 
@@
 we simulated only one system, the square-well fluid with well width
-$\lambda=1.3$ and filling fraction $\eta=0.3$, for sizes (i.e. number
-of spheres) $N\in\sp{5,25}$. We initialized histogram methods for
+$\lambda=1.3$ and filling fraction $\eta=0.3$, for sizes (number of
+spheres) $N\in\sp{5,25}$. We initialized histogram methods for
 $10^5N^2$ iterations ($10^5N^3$ moves), and simulated for just as long
 afterwards. To check for consistency in simulation results, we
-simulated each combination of histogram method and system size for 30
+simulated each combination of histogram method and system size with 30
 different random number generator seeds.
@@
 confidence in low-temperature gold standard results, we initialized
-them with $kT_{\t{min}}/\epsilon=0.1$. After initialization, the gold
+them with $kT_{\t{min}}=0.1\epsilon$. After initialization, the gold
 standard runs simulate for $10^{10}$ iterations ($10^{10}N$
@@
 compare against ``converged flat'' simulations. Converged flat
-simulations read in the transition matrix from golden simulations
-(which accumulates statistics in both during and after initialization)
-to construct a weight array, but simulate for the same number of
+simulations read in the transition matrix from golden simulations to
+construct a weight array, and simulate for the same number of
 iterations as the standard simulations.
@@
 
-Most of our analysis will use quantities computed from two arrays
-given by simulations: the energy histogram $H\p{E}$ and the weights
-$w\p{E}$. To get a sense of what these arrays look like, Figure
-\ref{fig:arrays} shows the energy histogram and weights from an $N=25$
-simulation, along with the density of states computed via $\tilde
-D=H/w$ and normalized to have a maximum value of 1. The vertical
-reference line in these plots denotes the minimum important energy
-$E_-$. For reference, we also provide the histogram produced by an
-infinite temperature (no weights) simulation.
+Most of our analysis will use quantities computed from two arrays: the
+energy histogram $H\p{E}$ and the weights $w\p{E}$. To get a sense of
+what these arrays look like, Figure \ref{fig:arrays} shows the energy
+histogram and weights from $N=25$ simulations, along with the density
+of states computed via $\tilde D=H/w$ and normalized to have a maximum
+value of 1. The vertical reference line in these plots denotes the
+minimum important energy $E_-$. For reference, we also provide the
+histogram produced by an infinite temperature (no weights) simulation.
 
@@
 observed energies. After several reset cycles of the Wang-Landau
-algorithm, the weight-modifying factor $f_k$ gets exceedingly small,
-causing Wang-Landau to get stuck at newly encountered low energies
-trying to fix their weights with a factor $f_k$ whose magnitude is
+algorithm, the weight-modifying factor $f_k$ gets exceedingly small
+relative to its initial value. When Wang-Landau finds new low
+energies, therefore, the simulation gets stuck trying to fix the
+weights at these energies with a factor $f_k$ whose magnitude is
 appropriate for fine tuning, rather than heavy adjustment. If
@@
 adjusted their weights to some reasonable value. Due to the drastic
-failure of Wang-Landau, will omit its results, which are nonsensical,
-from many figures.
+failure of Wang-Landau, will omit its nonsensical results from some
+figures. \fixme{which figures?}
 
@@
 the simple flat and vanilla Wang-Landau methods. At face value, this
-behavior of TMMC and OE-TMMC is undesirable.
+behavior of TMMC and OE-TMMC is undesirable, as these are the most
+interesting energies for low temperatures, and the most difficult
+energies to sample.
 
@@
 has collected many statistics on these energies, but the energy
-histogram contains no information about how correlated different
-energy observations are. To determine how effective different
-simulations are at collecting uncorrelated statistics, we can look at
-the pessimistic sample rate $r_p\p{E}$, which is the mean number of
-simulation iterations required to sample a given energy $E$. This rate
-is given in terms of the number of iterations $I$ in a simulation
-(where we always have $I=10^5N^2$) and the pessimistic sample rate
-$s_p\p{E}$ by $r_p\p{E}=I/s_p\p{E}$. Figure \ref{fig:sample_rates}
-shows pessimistic sample rates for all methods except Wang-Landau
-averaged over 30 random number generator seeds for $N=25$
-simulations. Lower pessimistic sample rates mean that a simulation
-takes less time to collect statistics which are guaranteed to be
-uncorrelated. By this metric, TMMC and OE-TMMC again perform worse
-than simple flat and vanilla Wang-Landau.\fixme{comment on OETMMC
-  doing worse than TMMC}
+histogram contains no information about how correlated these
+statistics are. To determine how effective different simulations are
+at collecting uncorrelated statistics, we can look at the pessimistic
+sample rate $r_p\p{E}$, which is the mean number of simulation
+iterations required to sample a given energy $E$. This rate is roughly
+equal to the number of iterations $I$ in a simulation and the
+pessimistic sample rate $s_p\p{E}$ by $r_p\p{E}=I/s_p\p{E}$; the
+greater the number of iterations $I$, the better this
+approximation. Figure \ref{fig:sample_rates} shows pessimistic sample
+rates averaged over 30 simulations with $N=25$. Lower pessimistic
+sample rates mean that a simulation takes less time to collect
+statistics which are guaranteed to be uncorrelated. By this metric,
+TMMC and OE-TMMC again perform worse than simple flat and vanilla
+Wang-Landau.\fixme{comment on why it is interesting that OETMMC doing
+  worse than TMMC?}
 
@@
 Histograms and sampling rates let us speculate about the quality of a
-simulations, but the most important task for any simulation is to
-yield accurate physical results. We therefore compute, via the formula
-in (\ref{eq:XT}), the temperature dependence of the square-well
-fluid's potential energy
+simulations, but the most important deliverable of any simulation is
+accurate physical results. We therefore compute, via the formula in
+(\ref{eq:XT}), the temperature dependence of the square-well fluid's
+potential energy
 \begin{align}
@@
 \end{align}
-and the configurational entropy\fixme{how is this different from
-  regular entropy?}
+and the configurational entropy \fixme{is this correct?}
 \begin{align}
@@
 \ref{fig:internal_energy}--\ref{fig:config_entropy} for $N=25$
-simulations, averaged over 30 random number generator seeds. When
-looking at these figures, one should remember that we are only
-confident in results down to $kT/\epsilon=0.2$ for regular
-simulations, and $kT/\epsilon=0.1$ for golden ones. With the exception
-of Wang-Landau, which has a heavy low-energy bias, all simulations
-agree fairly well on these properties of the square-well fluid.
+simulations, averaged over 30 simulations. When looking at these
+figures, one should remember that we are only confident in results
+down to $kT=0.2\epsilon$ for regular simulations, and $kT=0.1\epsilon$
+for golden ones. With the exception of Wang-Landau, which has a heavy
+low-energy bias, all simulations agree fairly well on these properties
+of the square-well fluid.
 
@@
 however, show that the square-well fluid has a phase transition around
-$kT/\epsilon\approx0.5$, which justifies our choice of
-$kT_{\t{min}}/\epsilon=0.2$ and ensures that the histogram methods are
+$kT\approx0.5\epsilon$, which justifies our choice of
+$kT_{\t{min}}=0.2\epsilon$ and ensures that the histogram methods are
 indeed performing the nontrivial task of exploring the liquid-gas
@@
 paper, we will neglect to demonstrate that this phase transition is
-indeed between liquid and gas phases, and take this fact as given.
+indeed between liquid and gas phases, and take this fact as
+given. \fixme{cite instead of stating as fact?}
 
@@
 U$, $\Delta C_V$ and $\Delta S_{\t{config}}$ relative to the TMMC
-golden calculations for each method (except Wang-Landau). Figures
-\ref{fig:internal_energy_err}--\ref{fig:config_entropy_err} show these
-errors. Much like the arrays in Figure \ref{fig:arrays}, we provide
-these figures only for a sense of how error curves behave. Errors (at
-$kT/\epsilon\ge0.2$) typically peak near the phase transition at
-$kT/\epsilon\approx0.5$, and drop back down.
+golden calculations, shown in Figures
+\ref{fig:internal_energy_err}--\ref{fig:config_entropy_err}. Much like
+the arrays in Figure \ref{fig:arrays}, we provide these figures for a
+sense of how error curves behave. Errors (at $kT\ge0.2\epsilon$)
+typically peak near the phase transition at $kT\approx0.5\epsilon$,
+and drop back down.
 
@@
 comprehensive picture of errors from each method. These figures show
-the average maximum error in $U$, $C_V$, and $S$ for different system
-sizes. Though errors jump around considerably in these figures, the
-fact that they are roughly level between $N=15$ to $25$ justifies the
-choice $N^2$ scaling for our number of initialization iterations.
-Future studies of larger systems might use this scaling for initial
-estimates of initialization times.
+the average maximum error in $U$, $C_V$, and $S_{\t{config}}$ for
+different system sizes. Though errors jump around considerably in
+these figures, the fact that they are roughly stable (logarithmically)
+between $N=15$ to $25$ justifies the choice $N^2$ scaling for our
+number of initialization iterations.  Future studies might use this
+scaling to estimate appropriate initialization time for larger
+systems.
 
@@
 
-It is more meaningful to compare errors against errors, rather than
-system size. \fixme{wording} Figures
-\ref{fig:u_comps}--\ref{fig:s_comps} show the errors of each method
-plotted against errors of the converged flat histogram weight
+Figures \ref{fig:u_comps}--\ref{fig:s_comps} provide a different
+visualization of the same information as in Figures
+\ref{fig:u_scaling}--\ref{fig:s_scaling}, plotting the errors of all
+histogram methods against those of converged flat histogram weight
 simulations. The fact that some points fall below the diagonal
@@
 high-quality golden data to construct weight arrays, and should
-therefore outperform all standard simulations. These events are thus
-worthwhile looking into in more detail.
-
-Figures \ref{fig:u_scaling}--\ref{fig:s_comps} make clear, again, the
+therefore outperform all standard simulations. \fixme{comment on why
+  some methods might do better than CFW?}
+
+As with the arrays, the error comparison figures demonstrate the
 inability of Wang-Landau to sample a system without a predefined
-energy range. These figures also show that the performances of simple
-flat method, vanilla Wang Landau, and TMMC are about the same, except
-for an anomalous case with $N=21$ in which simple flat yields
-simulation results with large errors. We do not know the reason for
-this anomalous case, which might be worth investigating. Finally,
-these error figures further confirm earlier suspicions of OE-TMMC's
+energy range. The simple flat method, vanilla Wang Landau, and TMMC
+seem to give comparable errors, except for some anomalous cases with
+$N=18$ and $N=21$ in which simple flat has large errors. We do not
+know the reason for these anomalies, but they could occur if the
+method terminates before finding energies near $E_-$, in which case it
+would get stuck in simulation at low energies with canonical weights
+for the minimum temperature $kT_{\t{min}}=0.2\epsilon$. Finally, these
+error figures further confirm earlier suspicions of OE-TMMC's
 underperformance relative to TMMC, despite the fact that it is
@@
 
-While the results in this paper are not strongly conclusive, they may
-be used as guidelines for future studies of histogram methods,
-especially as applied to the square-well fluid. After implementing
-several histogram methods and applying them to the $\lambda=1.3$,
-$\eta=0.3$ square-well fluid, we found that Wang-Landau, together with
-a predefined energy range, consistently yields the smallest errors and
-highest low-energy sampling rates. Without prior knowledge of the
-minimum important energy of a system, however, Wang-Landau can get
-stuck at low energies, and is incapable of recovering. Wang-Landau is
-therefore most appropriate for studying systems in which the entire
-energy range is known a priori, or in which only a given energy range
-is of particular interest.
-
-The simple flat method and TMMC methods seem to perform roughly on par
-with vanilla Wang-Landau (i.e. Wang-Landau with a specified energy
-range), except an anomalous case of failure worth investigating on the
-part of simple flat. Our implementation of the optimized ensemble
-using a transition matrix, which is meant to improve the TMMC weight
-array, seemed to only make the weight array worse.
-
-\fixme{fix when I'm awake}
+After implementing several histogram methods and applying them to the
+$\lambda=1.3$, $\eta=0.3$ square-well fluid, we found that
+Wang-Landau, if given a predefined energy range, consistently yields
+some of the smallest errors and highest low-energy sampling rates of
+all histogram methods, with comparable results from TMMC and the
+simple flat method. Without prior knowledge of the minimum important
+energy of a system, however, Wang-Landau can get stuck at low
+energies, and is incapable of recovering. Wang-Landau is therefore
+most appropriate for studying systems in which the entire energy range
+is known a priori, or in which only a given energy range is of
+particular interest. The simple flat method also sometimes fails to
+initialize; if gone unnoticed, such failures can be costly in terms of
+computation time and resources. Our implementation of the optimized
+ensemble using the transition matrix resulted in lower sampling rates
+and higher errors than the flat histogram TMMC weights. We suspect
+that the optimized ensemble is failing to improve TMMC because its
+theory neglects the discretization of energies, which may be
+non-negligible for the systems we simulated.
+
+\fixme{make the conclusion more satisfying?}
 

### 2015-05-11 papers/histogram/thesis/thesis.tex
minor changes and an acknowledgments statement

@@
 method, transition matrix Monte Carlo, as well as two methods of our
-own. The first of our methods is novel, while the second is a hybrid
-of two published methods (transition matrix Monte Carlo and the
-optimized ensemble).
+own. The first of our methods, the simple flat method, is novel, while
+the second is a hybrid of two published methods (transition matrix
+Monte Carlo and the optimized ensemble).
 
@@
 histogram methods, but fails catastrophically without a predefined
-energy range for the square-well fluid. Our first method gives results
-comparable to successful Wang-Landau simulations, but can have
-anomalous cases with large error. The transition matrix Monte Carlo
-consistently yields results whose quality is marginally worse than
-functional Wang-Landau. Finally, our implementation of the optimized
-ensemble using a transition matrix results in worse performance over
-the straightforward transition matrix Monte Carlo method.
+energy range for the square-well fluid. The simple flat method and
+transition matrix Monte Carlo give results comparable to successful
+Wang-Landau simulations, but simple flat has some anomalous cases with
+large errors. Finally, our implementation of the optimized ensemble
+using a transition matrix results in worse performance over the
+straightforward transition matrix Monte Carlo method.
 
@@
 
-\subsection{Monte Carlo fluid simulations}
+\subsection{Monte Carlo fluid simulation}
 \label{sec:mc_sim}
 
-While model systems are powerful tools for understanding complex
-physical systems, they do not themselves exist in the real
-world. Consequently, direct experimental tests of theories for model
-system (e.g. square-well density functional theories) are not
-possible. For this reason, model system theories are commonly tested
-against Monte Carlo simulations. Proper implementation of Monte Carlo
-methods to study completely characterized systems ensures that
-statistical results from simulations converge on the exact properties
-of the simulated system in the infinite simulation time
-limit. Furthermore, uncertainties in quantities computed via Monte
-Carlo simulations are typically well-defined, monotonically decreasing
-functions of simulation time, allowing one to run simulations to the
-desired level of accuracy.
+Model systems such as the square-well fluid are powerful tools for
+understanding complex physical systems, but they do not themselves
+exist in the real world. Consequently, direct experimental tests of
+theories for model system (e.g. square-well density functional
+theories) are not possible. For this reason, model system theories are
+commonly tested against Monte Carlo simulations. Proper implementation
+of Monte Carlo methods to study completely characterized systems
+ensures that statistical results from simulations converge on the
+exact properties of the simulated system in the infinite simulation
+time limit. Furthermore, uncertainties in quantities computed via
+Monte Carlo simulations are typically well-defined, monotonically
+decreasing functions of simulation time, allowing one to run
+simulations to the desired level of accuracy.
 
@@
 Algorithm \ref{alg:metropolis} provides a sketch of unbiased,
-Metropolis Monte Carlo fluid simulations. Such an algorithm is
+Metropolis Monte Carlo fluid simulation. Such an algorithm is
 ``unbiased'' in the sense that it collects statistics (i.e. data) on
@@
 that move, should it occur. In terms of a transition histogram $\tilde
-T\p{E,\Delta E}$ with $E=E_i$ and $\Delta E=E_f-E_i$,
+T_d\p{E,\Delta E}$ with $E=E_i$ and $\Delta E=E_f-E_i$,
 \begin{align}
@@
 \end{align}
-Diffusivity $\alpha\p{E}$ is thus the variance of the change in energy
-$\Delta E$ from as single move out of $E$. Reconciling this definition
-with that we provide in (\ref{eq:walker_current_definition}) is
-considered outside the scope of this paper.
+The diffusivity $\alpha\p{E}$ is thus the variance of the change in
+energy $\Delta E$ from as single move out of $E$. Reconciling this
+definition with that we provide in
+(\ref{eq:walker_current_definition}) is considered outside the scope
+of this paper. \fixme{say something more convincing? cite?}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Acknowledgments}
+\label{sec:acknowledgements}
+
+I wish to thank, first and foremost, my advisor Dr. David Roundy, for
+mentoring me and providing me a wealth of knowledge and expertise in
+various subjects of physics, numerical methods, and programming
+nuances. I would also like thank Paho Lurie-Gregg, with whom I
+frequently exchange and discuss codes and ideas, and whose hard
+polyhedra Monte Carlo simulations I used as the template for my own
+codes. Lastly, I would like to thank thank Eva Morgun, who prompted me
+to {\it act} on my textbook enthusiasm for physics, and get involved
+in the fun and interesting world of scientific research.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \nocite{*} \bibliography{thesis}

### 2015-05-11 papers/histogram/thesis/poster.tex
added beginnings of poster on my project

\documentclass[paperwidth=48in,paperheight=36in,
fontscale=0.27,margin=0.75in]{baposter}

\definecolor{bordercol}{RGB}{40,40,40}
\definecolor{headercol1}{RGB}{186,215,230}
\definecolor{headercol2}{RGB}{80,80,80}
\definecolor{headerfontcol}{RGB}{0,0,0}
\definecolor{boxcolor}{RGB}{186,215,230}

\frenchspacing

%%% enumerating
\usepackage{enumitem} % include for \setlist{}, use below
\setlist{nolistsep} % more compact spacing between environments
\setlist[itemize]{leftmargin=*} % nice margins for itemize ...

%%% math and physics
\usepackage{amsmath} % math packages
\renewcommand{\t}{\text} % for text in math environment
\usepackage[boldvectors,braket]{physymb} % physics packages
\renewcommand{\v}{\vec} % bold vectors
\newcommand{\p}[1]{\left(#1\right)} % bold vectors

%%% figures
\usepackage{graphicx,grffile,float,subcaption} % floats, etc.
\usepackage{multirow} % for multirow entries in tables

%% misc.
\usepackage[font=normalsize,labelfont=bf]{caption} % caption text

\renewcommand{\title}{Optimizing Monte Carlo Simulation of the
  Square-Well Fluid}
\renewcommand{\author}{\LARGE Michael A. Perlin, David Roundy}
\newcommand{\location}{Oregon State University, Department of Physics}

\begin{document}

\begin{poster}{
% options
    %% colors and appearance
    borderColor=bordercol, %
    headerColorOne=headercol1, %
    headerColorTwo=headercol2, %
    headerFontColor=headerfontcol, %
    boxColorOne=boxcolor, %
    textborder=rectangle, %
    headerborder=open, %
    background=none, %
    %% headers
    headerfont=\Large\sf\bf, %
    headerheight=0.11\textheight, %
    headershape=roundedright, %
    %% other
    columns=3 %
  }%
  {% top left logo
  }%
  {% title
    \LARGE\title
  }%
  {% authors
    \Large\author \\[1mm]
    \location
  }%
  {% top right logo
    % \includegraphics[width=2.5cm]{figs/osu.pdf}
    % \hspace{5mm}
  }%

  \headerbox{Motivation}
  {name=intro,column=0,row=0}{%

    The critical point of a substance is the point in phase space at
    which the distinction between liquid and gas phases becomes
    ill-defined. Understanding the behavior of fluids near the
    critical point is a major challenge in the field of classical
    density functional theory.

    % fixme: figure out how to properly do paragraph spacing
    \vspace{6pt}

    The square-well fluid is the simplest system with a liquid-vapor
    phase transition, and is therefore of interest for studying
    critical point behavior of fluids. We wish to study the
    square-well fluid via numerical methods so as to test theories of
    fluid behavior.

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/phase-diagram.pdf}
      \caption{Generic phase diagram.}
      \label{fig:phase_diagram}
    \end{figure}

  }

  \headerbox{The Square-Well Fluid}
  {name=sw_fluid,column=0,row=0,below=intro,above=bottom}{%

    % The square-well fluid is a model system used to capture low order
    % effects of short-range attractive forces, such as the van Der
    % Waals force. The fluid is composed of spheres with diameter
    % $\sigma$ and the pair potential shown graphically in Figure
    % \ref{fig:pair_potential}.

    % \vspace{6pt}

    % The parameters $\lambda$ and $\epsilon$ are called the well width
    % and well depth, respectively. The first ($r<\sigma$) part of the
    % pair potential forbids spheres from overlapping, whereas the
    % second ($\sigma<r<\lambda\sigma$) associates an energy $-\epsilon$
    % with each pair of spheres whose centers are within distance of
    % $\lambda\sigma$ of each other.

    \begin{itemize}
    \item Model system used to capture low order effects of
      short-range attractive forces, e.g. the van Der Waals force
    \item Composed of spheres with diameter $\sigma$ and the pair
      potential given in Figure \ref{fig:pair_potential}
    \item $\lambda$ is called the well width, and $\epsilon$ the well
      depth
    % \item Natural distance scale $\sigma$ and energy scale $\epsilon$
    \item First part of potential, $v_{sw}\p{r<\sigma}=\infty$,
      forbids spheres from overlapping
    \item Second part of potential,
      $v_{sw}\p{r:\sigma<r<\lambda\sigma}=-\epsilon$, associates an
      energy $-\epsilon$ with each pair of spheres whose centers are
      within distance of $\lambda\sigma$ of each other
    \item Spheres whose centers are more than $\lambda\sigma$ apart do
      not interact
    \end{itemize}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.45\columnwidth]{figs/square-well.pdf}
      \caption{Square-well pair potential.}
      \label{fig:pair_potential}
    \end{figure}

  }

  \headerbox{Monte Carlo Fluid Simulation}
  {name=mc_sims,column=1,row=0}{%

    % Model systems such as the square-well fluid are powerful tools for
    % understanding complex physical systems, but they do not themselves
    % exist in the real world. Consequently, direct experimental tests
    % of theories for model system are not possible. For this reason,
    % model system theories are commonly tested against Monte Carlo
    % simulations.

    % \vspace{6pt}

    % Standard Monte Carlo simulations of a fluid sample the fluid's
    % configuration space by randomly moving fluid elements
    % (i.e. spheres), and collecting statistics on observed fluid
    % properties (e.g. internal energy, density, etc.). Studying the
    % square-well fluid's critical point, however, requires sampling
    % states with highly localized distributions of spheres, which are
    % extremely unlikely to be found in a random sampling of all
    % possible fluid configurations. In the case of the square-well
    % fluid, the search for localized distributions of spheres is
    % equivalent to the search for low energy states.

    \begin{itemize}
    \item Model systems are powerful tools for understanding complex
      physical systems, but they do not exist in the real world
    \item Direct experimental tests are not possible, so we test
      theories against Monte Carlo simulations instead
    \item Monte Carlo fluid simulations: moving spheres randomly to
      collect statistics on properties (e.g. internal energy, density,
      etc.) of valid fluid configurations
    \item Studying critical point requires sampling states with highly
      localized distribution of spheres, which is unlikely to occur by
      randomly sampling of all possible fluid configurations (second
      law of thermodynamics)
    \item For the square-well fluid, localized distribution of spheres
      is synonymous with a low energy state
    \end{itemize}

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/dos-example.pdf}
      \caption{Sample density of states of a particular square-well
        fluid ($\lambda=1.3$, $\eta=0.3$), as computed using several
        methods. The logarithmic span of the density of states is
        proportional to the number of moves one must simulate via
        standard Monte Carlo in order to observe the energies with the
        lowest state densities.}
      \label{fig:dos}
    \end{figure}

  }

  \headerbox{Histogram Methods}
  {name=histogram_methods,column=1,row=0,below=mc_sims,above=bottom} {%

    % Histogram methods are a means for biasing Monte Carlo simulations
    % in favor of some regions of state space, i.e. toward low
    % energies. The idea, as applied to the problem at hand, is simple:
    % accept all Monte Carlo moves into fluid states with a lower
    % energy, and reject, with some probability, moves into fluid states
    % with a higher energy.

    % \vspace{6pt}

    % There are various published histogram methods, but little to no
    % literature exists on these methods' systematic comparison, or on
    % their application to the square-well fluid. We implement several
    % methods, including Wang-Landau, transition matrix Monte Carlo, and
    % a modified version of the so-called optimized ensemble, and
    % determine several metrics for their comparison.

    \begin{itemize}
    \item Means for biasing Monte Carlo simulations in favor of lower
      energy states
    \item Accept all moves in simulation which lower the system's
      energy, probabilistically reject moves to higher energy
    \item Details of rejection process define a histogram method
    \item There are several published histogram methods, but little to
      no literature on their systematic comparison, or their
      application to the square-well fluid
    \item We implement several methods: our own ``simple flat''
      method, Wang-Landau, transition matrix Monte Carlo (TMMC), and a
      hybrid of TMMC with a so-called ``optimized ensemble'' method
    \end{itemize}

  }

  \headerbox{Comparison}
  {name=comparison,column=2}{%

    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/sample-rates.pdf}
      \caption{Sample rates}
      \label{fig:sample_rates}
    \end{figure}
    \begin{figure}[H]
      \centering
      \includegraphics[width=0.5\columnwidth]{figs/u-comps.pdf}
      \caption{Errors}
      \label{fig:errors}
    \end{figure}

  }

  \headerbox{Conclusions}
  {name=conclusions,column=2,below=comparison}{%

  }

  \headerbox{Acknowledgments}
  {name=thanks,column=2,below=conclusions,above=bottom}{%

  }

\end{poster}

\end{document}

### 2015-05-11 papers/histogram/thesis/thesis.tex
fixed some mistakes in figure labels; changed sample density of states figure in thesis; inserted a small joke in my thesis

@@
 have a maximum value of $1$ in order to emphasize the logarithmic span
-of $D\p{E}$. This logarithmic span ($\sim10^{60}$) is proportional to
+of $D\p{E}$. This logarithmic span ($\sim10^{300}$) is proportional to
 the number of moves which one must simulate via Metropolis Monte Carlo
 in order to observe the energies with the lowest state densities. For
-reference, the world's current fastest supercomputers can perform
-$\sim10^{16}$ floating-point operations per second.
+reference, there are approximately $10^{80}$ atoms in the observable
+universe, and the world's current fastest supercomputers can perform
+about $10^{16}$ floating-point operations per second.
 
@@
   {Sample density of states for square-well fluid with a well width
-    $\lambda=1.3$, filling fraction $\eta=0.3$, and $N=25$ spheres, as
-    computed by several histogram methods. The density of states has
-    been normalized to have a maximum value of $1$.}
+    $\lambda=1.3$, filling fraction $\eta=0.1$, and $N=100$ spheres,
+    computed via the transition matrix Monte Carlo method. The density
+    of states has been normalized to have a maximum value of $1$.}
   \label{fig:dos_sample}
@@
 
-Histogram methods provide a means to address Metropolis Monte Carlo's
-inability to sufficiently sample low entropy states by introducing a
-bias into the otherwise random sampling of state space. A weighting
-function $w$ is introduced, whose domain is the value of some property
-$X\p{s}$ which depends on the system state $s$. An additional
-condition is then added to step \ref{alg:metropolis_move} of Algorithm
-\ref{alg:metropolis} in order to accept an attempted move: the weights
-$w\p{X_i}$ and $w\p{X_f}$ of the initial (pre-move) state $s_i$ and
-final (post-move) state $s_f$, respectively, are used to determine the
-probability $P_m\p{s_i\to s_f}$ of accepting an otherwise valid move
-via
+We do not have time to count all of the atoms in the observable
+universe, and we certainly do not have time to wait for Metropolis
+Monte Carlo to sample (and worse, collect decent statistics on) low
+energy square-well fluid states. Histogram methods provide a means to
+address Metropolis Monte Carlo's inability to sufficiently sample low
+energy states by introducing a bias into the otherwise random
+sampling of state space. A weighting function $w$ is introduced, whose
+domain is the value of some property $X\p{s}$ which depends on the
+system state $s$. An additional condition is then added to step
+\ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis} in order
+to accept an attempted move: the weights $w\p{X_i}$ and $w\p{X_f}$ of
+the initial (pre-move) state $s_i$ and final (post-move) state $s_f$,
+respectively, are used to determine the probability $P_m\p{s_i\to
+  s_f}$ of accepting an otherwise valid move via
 \begin{align}
@@
 E<0$. The high energies which are more likely to transition down in
-energy than up are those above the state of maximum entropy.
+energy than up are those above the state of maximum entropy, i.e. the
+state at which $D\p{E}$ is maximal.
 
@@
 to {\it act} on my textbook enthusiasm for physics, and get involved
-in the fun and interesting world of scientific research.
+in the fun and interesting world of scientific research. \fixme{Thank
+  Jansen and McIntyre}
 

### 2015-05-11 papers/histogram/thesis/poster.tex
updates to poster

@@
       \caption{Sample density of states of a particular square-well
-        fluid ($\lambda=1.3$, $\eta=0.3$), as computed using several
-        methods. The logarithmic span of the density of states is
-        proportional to the number of moves one must simulate via
-        standard Monte Carlo in order to observe the energies with the
-        lowest state densities.}
+        fluid ($\lambda=1.3$, $\eta=0.3$). The logarithmic span of the
+        density of states is proportional to the number of moves one
+        must simulate via standard Monte Carlo in order to observe the
+        energies with the lowest state densities.}
       \label{fig:dos}
@@
     % methods, including Wang-Landau, transition matrix Monte Carlo, and
-    % a modified version of the so-called optimized ensemble, and
-    % determine several metrics for their comparison.
+    % a modified version of the so-called optimized ensemble. We compare
+    % low energy sampling rates for all methods, as well as relative
+    % errors in thermodynamic properties computed via each method.
 
@@
       hybrid of TMMC with a so-called ``optimized ensemble'' method
+    \item Comparison metrics: low energy sampling rates, and relative
+      errors in thermodynamic properties computed using results from
+      simulating via each method
     \end{itemize}
@@
 
-  \headerbox{Comparison}
-  {name=comparison,column=2}{%
+  \headerbox{Results}
+  {name=results,column=2}{%
 
@@
       \centering
-      \includegraphics[width=0.5\columnwidth]{figs/sample-rates.pdf}
-      \caption{Sample rates}
+      \includegraphics[width=0.6\columnwidth]{figs/sample-rates.pdf}
+      \caption{Iterations per statistically independent energy
+        sample.}
       \label{fig:sample_rates}
@@
       \centering
-      \includegraphics[width=0.5\columnwidth]{figs/u-comps.pdf}
-      \caption{Errors}
+      \includegraphics[width=0.6\columnwidth]{figs/u-comps.pdf}
+      \caption{Maximum errors in a set of simulations, relative to a
+        method which controls for system size and simulation time.}
       \label{fig:errors}
@@
   \headerbox{Conclusions}
-  {name=conclusions,column=2,below=comparison}{%
+  {name=conclusions,column=2,below=results}{%
 

### 2015-05-12 papers/histogram/thesis/poster.tex

@@
 
-\definecolor{bordercol}{RGB}{40,40,40}
-\definecolor{headercol1}{RGB}{186,215,230}
-\definecolor{headercol2}{RGB}{80,80,80}
-\definecolor{headerfontcol}{RGB}{0,0,0}
-\definecolor{boxcolor}{RGB}{186,215,230}
+\definecolor{border_color}{RGB}{40,40,40}
+\definecolor{header_color_left}{RGB}{186,215,230}
+\definecolor{header_color_right}{RGB}{80,80,80}
+\definecolor{header_font_color}{RGB}{0,0,0}
+\definecolor{box_color}{RGB}{186,215,230}
 
@@
     %% colors and appearance
-    borderColor=bordercol, %
-    headerColorOne=headercol1, %
-    headerColorTwo=headercol2, %
-    headerFontColor=headerfontcol, %
-    boxColorOne=boxcolor, %
+    borderColor=border_color, %
+    headerColorOne=header_color_left, %
+    headerColorTwo=header_color_right, %
+    headerFontColor=header_font_color, %
+    boxColorOne=box_color, %
     textborder=rectangle, %
@@
       physical systems, but they do not exist in the real world
-    \item Direct experimental tests are not possible, so we test
-      theories against Monte Carlo simulations instead
+    \item Theories tested against Monte Carlo simulations
     \item Monte Carlo fluid simulations: moving spheres randomly to
-      collect statistics on properties (e.g. internal energy, density,
+      collect statistics on properties (e.g. potential energy, density,
       etc.) of valid fluid configurations

### 2015-05-13 papers/histogram/thesis/poster.tex
some changes to plots; added presentation

@@
 
-%%% math and physics
-\usepackage{amsmath} % math packages
-\renewcommand{\t}{\text} % for text in math environment
-\usepackage[boldvectors,braket]{physymb} % physics packages
-\renewcommand{\v}{\vec} % bold vectors
-\newcommand{\p}[1]{\left(#1\right)} % bold vectors
+%%% math and figures
+\usepackage{amsmath,graphicx,float}
 
-%%% figures
-\usepackage{graphicx,grffile,float,subcaption} % floats, etc.
-\usepackage{multirow} % for multirow entries in tables
-
-%% misc.
-\usepackage[font=normalsize,labelfont=bf]{caption} % caption text
+%% caption text
+\usepackage[font=normalsize,labelfont=bf]{caption}
 
@@
 \begin{poster}{
-% options
+% poster options
     %% colors and appearance
@@
       depth
-    % \item Natural distance scale $\sigma$ and energy scale $\epsilon$
-    \item First part of potential, $v_{sw}\p{r<\sigma}=\infty$,
+    \item First part of potential, $v_{sw}(r<\sigma)=\infty$,
       forbids spheres from overlapping
     \item Second part of potential,
-      $v_{sw}\p{r:\sigma<r<\lambda\sigma}=-\epsilon$, associates an
+      $v_{sw}(r:\sigma<r<\lambda\sigma)=-\epsilon$, associates an
       energy $-\epsilon$ with each pair of spheres whose centers are
@@
       physical systems, but they do not exist in the real world
-    \item Theories tested against Monte Carlo simulations
-    \item Monte Carlo fluid simulations: moving spheres randomly to
-      collect statistics on properties (e.g. potential energy, density,
-      etc.) of valid fluid configurations
-    \item Studying critical point requires sampling states with highly
-      localized distribution of spheres, which is unlikely to occur by
-      randomly sampling of all possible fluid configurations (second
-      law of thermodynamics)
+    \item Theories of model systems are tested against Monte Carlo
+      simulations
+    \item Metropolis Monte Carlo fluid simulations: random walk in
+      configuration space
+    \item Collect histograms of observed system properties
+      (e.g. potential energy,
+    \item Critical point: highly localized distributions of spheres;
+      unlikely to be observed in random samples of all possible fluid
+      configurations (second law of thermodynamics)
     \item For the square-well fluid, localized distribution of spheres
-      is synonymous with a low energy state
+      implies a low energy
     \end{itemize}
@@
       \centering
-      \includegraphics[width=0.5\columnwidth]{figs/dos-example.pdf}
+      \includegraphics[width=0.55\columnwidth]
+      {figs/dos-poster-example.pdf}
       \caption{Sample density of states of a particular square-well
-        fluid ($\lambda=1.3$, $\eta=0.3$). The logarithmic span of the
-        density of states is proportional to the number of moves one
-        must simulate via standard Monte Carlo in order to observe the
-        energies with the lowest state densities.}
+        fluid ($\lambda=1.3$, $\eta=0.3$), as computed via several
+        methods. The logarithmic span of the density of states is
+        proportional to the number of moves one must simulate via
+        standard Monte Carlo in order to observe the energies with the
+        lowest state densities.}
       \label{fig:dos}
@@
     \item Details of rejection process define a histogram method
-    \item There are several published histogram methods, but little to
-      no literature on their systematic comparison, or their
-      application to the square-well fluid
+    \item Little to no literature on systematic comparison of
+      published histogram methods, or on application to the
+      square-well fluid
     \item We implement several methods: our own ``simple flat''
@@
       hybrid of TMMC with a so-called ``optimized ensemble'' method
-    \item Comparison metrics: low energy sampling rates, and relative
-      errors in thermodynamic properties computed using results from
-      simulating via each method
+    \item Comparison metrics:
+      \begin{itemize}
+      \item low energy sampling rates
+      \item errors in fluid properties computed from simulation
+        results
+      \end{itemize}
     \end{itemize}
@@
 
+    \begin{itemize}
+    \item Errors computed relative to long, ``gold standard''
+      simulations
+    \end{itemize}
+
     \begin{figure}[H]
@@
       \includegraphics[width=0.6\columnwidth]{figs/sample-rates.pdf}
-      \caption{Iterations per statistically independent energy
-        sample.}
+      \caption{Simulation iterations per independent energy sample.}
       \label{fig:sample_rates}
@@
       \centering
-      \includegraphics[width=0.6\columnwidth]{figs/u-comps.pdf}
-      \caption{Maximum errors in a set of simulations, relative to a
-        method which controls for system size and simulation time.}
+      \includegraphics[width=0.5\columnwidth]{figs/u-comps.pdf}
+      \caption{Maximum errors in a series of simulations, relative to
+        a method which controls for system size and simulation time.}
       \label{fig:errors}
     \end{figure}
+
+    \begin{itemize}
+    \item Simple flat method: performs decently in most cases, but has
+      a few outlying cases of failure
+    \item Wang-Landau: performs well if provided with a predefined
+      energy range, otherwise fails catastrophically
+    \item TMMC: decent performance; no failures; does not require
+      predefined energy range
+    \item Optimized ensemble applied to TMMC: worsens performance (by
+      all metrics)
+    \end{itemize}
 
@@
   \headerbox{Conclusions}
-  {name=conclusions,column=2,below=results}{%
+  {name=conclusions,column=2,below=results,above=bottom}{%
 
-  }
-
-  \headerbox{Acknowledgments}
-  {name=thanks,column=2,below=conclusions,above=bottom}{%
+    \begin{itemize}
+    \item Simple flat method possibly unreliable, though may be worth
+      investigating further (current algorithm is minimalistic)
+    \item For systems with a known energy range, or a particular
+      energy range of interest, Wang-Landau is best for exploring all
+      energies
+    \item TMMC: most reliable for systems without a known energy range
+    \item Optimized ensemble: designed to optimize standard strategy
+      for histogram methods, but only did damage; this strange result
+      is worth further investigation
+    \end{itemize}
 

### 2015-05-18 papers/histogram/thesis/thesis.tex

@@
 \newcommand{\supervisor}{Dr. David Roundy}
-\renewcommand{\date}{08 May 2015}
+\renewcommand{\date}{15 May 2015}
 
@@
 incapable of sampling the low energy fluid states which dominate the
-partition function at low (i.e. liquid-state) temperatures. Previous
-works have developed several generic Monte Carlo histogram methods for
+partition function at low liquid-state temperatures. Previous works
+have developed several generic Monte Carlo histogram methods for
 collecting statistics on low energy system states, but little or no
@@
 traditional models for testing and benchmarking such numerical
-techniques (e.g. the Ising model). We implement the Wang-Landau
-method, transition matrix Monte Carlo, as well as two methods of our
-own. The first of our methods, the simple flat method, is novel, while
-the second is a hybrid of two published methods (transition matrix
-Monte Carlo and the optimized ensemble).
-
-We find that Wang-Landau potentially performs better than other
-histogram methods, but fails catastrophically without a predefined
-energy range for the square-well fluid. The simple flat method and
-transition matrix Monte Carlo give results comparable to successful
-Wang-Landau simulations, but simple flat has some anomalous cases with
-large errors. Finally, our implementation of the optimized ensemble
-using a transition matrix results in worse performance over the
+techniques (e.g. the Ising model).
+
+We implement our own ``simple flat'' method, the Wang-Landau method,
+transition matrix Monte Carlo, and a modified version of the optimized
+ensemble. Performance of each method is determined by low energy
+sampling rates and maximum errors in several computed system
+properties. We find that Wang-Landau potentially performs better than
+other histogram methods, but fails catastrophically without a
+predetermined energy range. The simple flat method and transition
+matrix Monte Carlo give results comparable to successful Wang-Landau
+simulations, but simple flat has some anomalous cases with large
+errors. Finally, our implementation of the optimized ensemble using a
+transition matrix results in worse performance over the
 straightforward transition matrix Monte Carlo method.
 
-\fixme{edit when I'm awake}
+\fixme{edit}
 
@@
 
-\fixme{give more intro about phase transitions and the critical
-  point?}
-
-Understanding the behavior of fluids near the critical point, or the
-point at which the distinction between liquid and gas ceases to become
-well-defined, is a major challenge in the field of classical density
-functional theory. Studying the critical point of a fluid via
-simulations requires exploring the boundary between its liquid and
-gaseous phases. The square-well fluid is the simplest system with a
+The critical point of a system is the point in phase space at which
+the distinction between liquid and gaseous phases ceases to be
+well-defined. As shown in Figure \ref{fig:phase_diagram}, a fluid need
+not boil or condense to transition between liquid to gaseous phases:
+it can smoothly transition between the two by circumventing the
+critical point. Understanding the behavior of fluids near the critical
+point is a major challenge in the field of classical density
+functional theory. The square-well fluid is the simplest system with a
 liquid-vapor phase transition, and is therefore of interest for
-studying the critical point of fluids. For a generic square-well fluid
-to be observed in a liquid state, it must have a localized
-distribution of the spheres which make up the fluid. Such a
-distribution implies a low energy fluid state, as all interactions
-between spheres are attractive. Conversely, a low energy liquid state
-of a square-well fluid otherwise observable in both liquid and gaseous
-phases implies a localized distribution of spheres. The search for
-liquid states of the square-well fluid is thus equivalent to the
-search for low energy states.
+studying the critical point. For a generic square-well fluid to be
+observed in a liquid state, it must have a localized distribution of
+the spheres which make up the fluid. Such a distribution implies a low
+energy fluid state, as all interactions between spheres are
+attractive. Conversely, a low energy liquid state of a square-well
+fluid otherwise observable in both liquid and gaseous phases implies a
+localized distribution of spheres. The search for liquid states of the
+square-well fluid is thus equivalent to the search for low energy
+states.
 
@@
 straightforward implementation of Monte Carlo fluid simulations,
-however, highly localized distributions of fluid elements are
-extremely unlikely to occur, making such simulations impractical for
-studying the critical point of the square-well fluid. Several generic
+however, highly localized distributions of spheres are extremely
+unlikely to occur, making such simulations impractical for studying
+the critical point of the square-well fluid. Several generic
 computational methods, called histogram methods, exist for dealing
@@
 state space cannot be sampled via standard Monte Carlo methods in a
-reasonable amount of time. There is no common consensus, however, on
-the relative efficacy of these methods. Furthermore, some methods,
-such as the Wang-Landau method\cite{wang_landau}, require tuning
-several free parameters that lack both ``obvious'' values and
-systematic means for assigning them. This work will thus attempt to
-quantify and compare the effectiveness of a few such methods as
-applied to the square-well fluid.
+reasonable amount of time. There is little consensus, however, on the
+relative efficacy of these methods, and few have applied them to
+simulation of the square-well fluid.
+
+In this paper, we first review what the square-well fluid is, as well
+as how to simulate it via standard (Metropolis) Monte Carlo (Sections
+\ref{sec:sw_fluid}--\ref{sec:monte_carlo}). We then address the
+shortcomings of this means to study the square-well fluid, and
+consider how one can modify simulations to overcome these shortcomings
+in Section \ref{sec:histogram_methods}. We develop a thorough
+understanding of several histogram methods in Sections
+\ref{sec:flat_histogram}--\ref{sec:oetmmc}, and discuss some details
+pertinent to the actual implementation of these methods in Section
+\ref{sec:loose_ends}. Finally, in Section \ref{sec:results} we analyze
+the performance of a few histogram methods in simulations of a
+particular square-well fluid.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[width=0.47\textwidth]{figs/phase-diagram.pdf}
+  \caption[Phase diagram]
+  {A generic phase diagram showing the boundaries between solid,
+    liquid, and gaseous phases.}
+  \label{fig:phase_diagram}
+\end{figure}
 
@@
 \label{sec:sw_fluid}
-
-\begin{figure}[tb]
-  \centering
-  \includegraphics[width=0.47\textwidth]{figs/square-well.pdf}
-  \caption[The square-well pair potential]
-  {The square-well pair potential can be used to model short range
-    forces to first order. The infinite potential at $r<\sigma$ simply
-    enforces the condition that spheres cannot overlap. At
-    $r>\lambda\sigma$, the potential is null, and makes no
-    contribution to the fluid's internal energy.}
-  \label{fig:pair_potential}
-\end{figure}
 
@@
 
+\begin{figure}[!b]
+  \centering
+  \includegraphics[width=0.47\textwidth]{figs/square-well.pdf}
+  \caption[The square-well pair potential]
+  {The square-well pair potential can be used to model short range
+    forces to first order. The infinite potential at $r<\sigma$ simply
+    enforces the condition that spheres cannot overlap. At
+    $r>\lambda\sigma$, the potential is null, and makes no
+    contribution to the fluid's internal energy.}
+  \label{fig:pair_potential}
+\end{figure}
+
 \subsection{Monte Carlo fluid simulation}
-\label{sec:mc_sim}
-
-Model systems such as the square-well fluid are powerful tools for
+\label{sec:monte_carlo}
+
+Model systems, such as the square-well fluid, are powerful tools for
 understanding complex physical systems, but they do not themselves
@@
 
-\begin{algorithm}[tb]
+\begin{algorithm}[!b]
   \caption{Metropolis Monte Carlo fluid simulation}
@@
   \item Construct an initial ``typical'' (i.e. non-ordered) fluid
-    configuration. \fixme{explain how}
+    configuration. \fixme{explain how?}
 
@@
 than once every $\chi\p{N}$ moves, where $\chi\p{N}$ may be an
-arbitrary function of $N$ void of constant prefactors, e.g. $N\log N$,
-or $2^N$. These collection rules ensure that scaling up simulations
-does not cause them to asymptotically spend all computation time only
-collecting statistics, or only simulating the fluid. Collected
-statistics are used to find thermodynamic properties of the simulated
-fluid.
+arbitrary function of $N$, e.g. $N\log N$, or $2^N$. These collection
+rules ensure that scaling up simulations does not cause them to
+asymptotically spend all computation time only collecting statistics,
+or only simulating the fluid. Collected statistics are used to find
+thermodynamic properties of the simulated fluid.
 
@@
   \centering
-  \includegraphics[width=0.6\textwidth]{figs/dos-example.pdf}
+  \includegraphics[width=0.6\textwidth]{figs/dos-thesis-example.pdf}
   \caption[Density of states]
@@
 This observation will be used in Section \ref{sec:min_energy} to
-identify a maximum energy of interest.
+identify a maximum energy of interest in simulations.
 
@@
   systems? cite}, and is the standard method\fixme{cite} to achieve
-broad energy sampling via Monte Carlo simulations. \fixme{more intro
-  on WL}
+broad energy sampling via Monte Carlo simulations.
 
@@
 histogram\cite{tmmc}. Unlike the previous histogram methods, we desire
-a flat histogram during initialization via Algorithm \ref{alg:tmmc}
-not because we are constructing weights $w\p{E}$, but because we wish
-to sample the transition histogram $\tilde T_d\p{E,\Delta E}$ equally
-at all energies $E$. \fixme{briefly explain
-  (\ref{eq:tmmc_prob_simplified}) by considering a two-level system,
-  and extending to larger systems}.
+a flat histogram during initialization via Algorithm \ref{alg:tmmc} in
+order to broadly sample the transition histogram $\tilde
+T_d\p{E,\Delta E}$, rather than to construct weights $w\p{E}$
+directly.
 
@@
 straightforward to implement and compare to other histogram methods.
-\fixme{provide iterative algorithm?}
 
@@
 golden simulations initialize for $\sim11$ hours, and simulate for a
-week. To control for iteration number in the quality of results, we
-compare against ``converged flat'' simulations. Converged flat
-simulations read in the transition matrix from golden simulations to
-construct a weight array, and simulate for the same number of
-iterations as the standard simulations.
+week. To control for iteration number in the results from any given
+set of simulations, we compare against ``converged flat''
+simulations. Converged flat simulations read in the transition matrix
+from golden simulations to construct a weight array, and simulate for
+the same number of iterations as the standard simulations.
 
@@
 failure of Wang-Landau, will omit its nonsensical results from some
-figures. \fixme{which figures?}
+figures, as these results are distracting and unfit for drawing any
+conclusions.
 
@@
 TMMC and OE-TMMC again perform worse than simple flat and vanilla
-Wang-Landau.\fixme{comment on why it is interesting that OETMMC doing
-  worse than TMMC?}
+Wang-Landau.
 
@@
 phase boundary of the square-well fluid. For the purposes of this
-paper, we will neglect to demonstrate that this phase transition is
-indeed between liquid and gas phases, and take this fact as
-given. \fixme{cite instead of stating as fact?}
+paper, we will neglect to prove that this phase transition is indeed
+between liquid and gas phases, and take this fact as given. \fixme{say
+  more convincingly}
 
@@
 
-After implementing several histogram methods and applying them to the
-$\lambda=1.3$, $\eta=0.3$ square-well fluid, we found that
-Wang-Landau, if given a predefined energy range, consistently yields
-some of the smallest errors and highest low-energy sampling rates of
-all histogram methods, with comparable results from TMMC and the
-simple flat method. Without prior knowledge of the minimum important
-energy of a system, however, Wang-Landau can get stuck at low
-energies, and is incapable of recovering. Wang-Landau is therefore
-most appropriate for studying systems in which the entire energy range
-is known a priori, or in which only a given energy range is of
-particular interest. The simple flat method also sometimes fails to
-initialize; if gone unnoticed, such failures can be costly in terms of
+After implementing several histogram methods and applying them to
+simulation of the $\lambda=1.3$, $\eta=0.3$ square-well fluid, we
+found that Wang-Landau, if given a predefined energy range,
+consistently yields some of the smallest errors and highest low-energy
+sampling rates of all histogram methods, with comparable results from
+TMMC and the simple flat method. Without prior knowledge of the
+minimum important energy of a system, however, Wang-Landau can get
+stuck at low energies, in which case it is incapable of
+recovering. Wang-Landau is therefore most appropriate for studying
+systems in which the entire energy range is known a priori, or in
+which only a given energy range is of particular interest. The simple
+flat method, while generally performing well, sometimes fails to
+initialize. If gone unnoticed, such failures can be costly in terms of
 computation time and resources. Our implementation of the optimized
@@
 and higher errors than the flat histogram TMMC weights. We suspect
-that the optimized ensemble is failing to improve TMMC because its
-theory neglects the discretization of energies, which may be
-non-negligible for the systems we simulated.
-
-\fixme{make the conclusion more satisfying?}
+that the optimized ensemble is failing to improve TMMC because the
+theory behind it neglects the discretization of energies.
+
+\fixme{make conclusions more satisfying?}
 

### 2015-05-27 papers/histogram/thesis/thesis.tex
updating thesis and presentation

@@
 
-%%% standard header
+%%% general typesetting rules
 \usepackage[margin=1in]{geometry} % one inch margins
@@
 \usepackage{hyperref} % for linking references
-\newcommand{\psl}{6pt} % store spacing as command for consistency
-\setlength{\parindent}{0cm} % don't indent new paragraphs...
-\parskip \psl % ... place a space between paragraphs instead
-\usepackage[inline]{enumitem} % include for \setlist{}, use below
+\usepackage{indentfirst}
+\usepackage[inline]{enumitem} % for inline enumeration
 \frenchspacing % add a single space after a period
@@
 interface is an interesting regime for testing density functional
-theories, but unbiased Metropolis Monte Carlo simulations are
-incapable of sampling the low energy fluid states which dominate the
-partition function at low liquid-state temperatures. Previous works
-have developed several generic Monte Carlo histogram methods for
-collecting statistics on low energy system states, but little or no
-literature exists on these methods' systematic comparison, as well as
-their application to the square-well fluid. The square-well fluid in
+theories, but unbiased Monte Carlo simulations are incapable of
+sampling the low energy fluid states which dominate the partition
+function at low liquid-state temperatures. Previous works have
+developed several generic Monte Carlo histogram methods for collecting
+statistics on low energy system states, but little or no literature
+exists on these methods' systematic comparison, as well as their
+application to the square-well fluid. The square-well fluid in
 particular introduces application challenges not manifest in
@@
 straightforward transition matrix Monte Carlo method.
-
-\fixme{edit}
 
@@
 In this paper, we first review what the square-well fluid is, as well
-as how to simulate it via standard (Metropolis) Monte Carlo (Sections
+as how to simulate it via standard, unbiased Monte Carlo (Sections
 \ref{sec:sw_fluid}--\ref{sec:monte_carlo}). We then address the
-shortcomings of this means to study the square-well fluid, and
-consider how one can modify simulations to overcome these shortcomings
-in Section \ref{sec:histogram_methods}. We develop a thorough
-understanding of several histogram methods in Sections
-\ref{sec:flat_histogram}--\ref{sec:oetmmc}, and discuss some details
-pertinent to the actual implementation of these methods in Section
-\ref{sec:loose_ends}. Finally, in Section \ref{sec:results} we analyze
-the performance of a few histogram methods in simulations of a
+shortcomings of these simulations as a means for studying the
+square-well fluid, and consider how one can modify simulations to
+overcome these shortcomings in Section \ref{sec:histogram_methods}. We
+develop a thorough understanding of several histogram methods in
+Sections \ref{sec:flat_histogram}--\ref{sec:oetmmc}, and discuss some
+details pertinent to the actual implementation of these methods in
+Section \ref{sec:loose_ends}. Finally, in Section \ref{sec:results} we
+analyze the performance of a few histogram methods in simulations of a
 particular square-well fluid.
@@
 \begin{algorithm}[!b]
-  \caption{Metropolis Monte Carlo fluid simulation}
+  \caption{Unbiased Monte Carlo fluid simulation}
   \label{alg:metropolis}
@@
 
-Algorithm \ref{alg:metropolis} provides a sketch of unbiased,
-Metropolis Monte Carlo fluid simulation. Such an algorithm is
-``unbiased'' in the sense that it collects statistics (i.e. data) on
-all valid system configurations with equal probability. Statistics
-whose collection time scales as $\mathcal O\p{1}$ with system size,
-meaning that increasing the number of spheres $N$ in the simulation
-does not affect collection time, can be collected after every move
-(defined in the algorithm). Statistics whose collection time scales as
-$\mathcal O\p{N}$, meaning that doubling $N$ doubles collection time,
-can be collected after each iteration. In general, collection with
-$\mathcal O\sp{\chi\p{N}}$ time scaling should not occur more often
-than once every $\chi\p{N}$ moves, where $\chi\p{N}$ may be an
-arbitrary function of $N$, e.g. $N\log N$, or $2^N$. These collection
-rules ensure that scaling up simulations does not cause them to
+Algorithm \ref{alg:metropolis} provides a sketch of unbiased Monte
+Carlo fluid simulation. Such an algorithm is ``unbiased'' in the sense
+that it collects statistics (i.e. data) on all valid system
+configurations with equal probability. Statistics whose collection
+time scales as $\mathcal O\p{1}$ with system size, meaning that
+increasing the number of spheres $N$ in the simulation does not affect
+collection time, can be collected after every move (defined in the
+algorithm). Statistics whose collection time scales as $\mathcal
+O\p{N}$, meaning that doubling $N$ doubles collection time, can be
+collected after each iteration. In general, collection with $\mathcal
+O\sp{\chi\p{N}}$ time scaling should not occur more often than once
+every $\chi\p{N}$ moves, where $\chi\p{N}$ may be an arbitrary
+function of $N$, e.g. $N\log N$, or $2^N$. These collection rules
+ensure that scaling up simulations does not cause them to
 asymptotically spend all computation time only collecting statistics,
@@
 
-Due to the fact that Metropolis Monte Carlo simulations sample all of
-state space randomly and without preference, a histogram $H\p{X}$ of
-observations of some system property $X$ is directly proportional to
-the density of states $D\p{X}$; that is, the number of observations
-$H\p{X_0}$ of the system with $X\p{s}=X_0$ is proportional to the
-total number of states $s$ for which $X\p{s}=X_0$. It is sometimes the
-case, however, that the density of states in some range $R$ of
-possible $X$ is so low that it is practically impossible (via
-Metropolis Monte Carlo) to sufficiently sample $R$, that is, to
-accumulate a statistically significant histogram $H\p{X\in R}$, in any
-reasonable amount of time. Given that the entropy $S\p{X}$ can be
-expressed in terms of the number of microstates $\Omega\p{X}$ as
-$S\p{X}=k\ln\Omega\p{X}\propto\ln D\p{X}$, we may refer to regions
-with low state densities $D\p{X}$ as low entropy states.
+Due to the fact that unbiased Monte Carlo simulations sample all of
+state space randomly, a histogram $H\p{X}$ of observations of some
+system property $X$ is directly proportional to the density of states
+$D\p{X}$; that is, the number of observations $H\p{X_0}$ of the system
+with $X\p{s}=X_0$ is proportional to the total number of states $s$
+for which $X\p{s}=X_0$. It is sometimes the case, however, that the
+density of states in some range $R$ of possible $X$ is so low that it
+is practically impossible (via unbiased Monte Carlo) to sufficiently
+sample $R$, that is, to accumulate a statistically significant
+histogram $H\p{X\in R}$, in any reasonable amount of time. Given that
+the entropy $S\p{X}$ can be expressed in terms of the number of
+microstates $\Omega\p{X}$ as $S\p{X}=k\ln\Omega\p{X}\propto\ln
+D\p{X}$, we may refer to regions with low state densities $D\p{X}$ as
+low entropy states.
 
@@
 of $D\p{E}$. This logarithmic span ($\sim10^{300}$) is proportional to
-the number of moves which one must simulate via Metropolis Monte Carlo
+the number of moves which one must simulate via unbiased Monte Carlo
 in order to observe the energies with the lowest state densities. For
@@
 We do not have time to count all of the atoms in the observable
-universe, and we certainly do not have time to wait for Metropolis
-Monte Carlo to sample (and worse, collect decent statistics on) low
-energy square-well fluid states. Histogram methods provide a means to
-address Metropolis Monte Carlo's inability to sufficiently sample low
-energy states by introducing a bias into the otherwise random
-sampling of state space. A weighting function $w$ is introduced, whose
-domain is the value of some property $X\p{s}$ which depends on the
-system state $s$. An additional condition is then added to step
+universe, and we certainly do not have time to wait for unbiased Monte
+Carlo to sample (and worse, collect decent statistics on) low energy
+square-well fluid states. Histogram methods provide a means to address
+unbiased Monte Carlo's inability to sufficiently sample low energy
+states by introducing a bias into the otherwise random sampling of
+state space. A weighting function $w$ is introduced, whose domain is
+the value of some property $X\p{s}$ which depends on the system state
+$s$. An additional condition is then added to step
 \ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis} in order
@@
 determine properties $\bk{X}_{T\ne T_0}$, and are thus referred to as
-``fixed temperature'' simulations. Incidentally, Metropolis Monte
-Carlo simulations can be thought of as infinite-temperature
-simulations, as they are equivalent to simulations with canonical
-weights
+``fixed temperature'' simulations. Incidentally, unbiased Monte Carlo
+simulations can be thought of as infinite-temperature simulations, as
+they are equivalent to simulations with canonical weights
 \begin{align}
@@
 
-Having covered the basics of the square-well fluid as well as
-Metropolis and biased Monte Carlo simulations, we can now discuss
-several histogram methods for constructing energy weights. In Section
+Having covered the basics of the square-well fluid as well as unbiased
+and biased Monte Carlo simulations, we can now discuss several
+histogram methods for constructing energy weights. In Section
 \ref{sec:flat_histogram}, we will introduce the flat histogram method,
@@
 that histogram methods are generally necessary when some range of
-energies $R$ is inaccessible via Metropolis Monte Carlo
-simulations. In such a case, the energy histogram $H\p{E\in R}$ after
-simulating for any reasonable amount of time will be statistically
-insignificant or null, invalidating the proportionality $H\p{E}\propto
-D\p{E}$. This method will therefore only ``flatten'' the histogram at
-energies $E$ which have been sufficiently sampled for $H\p{E}$ to be
-statistically significant. The end goal in implementing histogram
-methods, however, is precisely to sample those energies that cannot be
-sufficiently sampled without employing clever schemes to bias Monte
-Carlo simulations. Therefore, this method does not directly help us
-study the square-well fluid. This method does, however, motivate the
-method in Section \ref{sec:simple_flat}, and provides some of the
-theory behind the methods in Sections \ref{sec:tmmc} and
-\ref{sec:oetmmc}.
+energies $R$ is inaccessible via unbiased Monte Carlo simulations. In
+such a case, the energy histogram $H\p{E\in R}$ after simulating for
+any reasonable amount of time will be statistically insignificant or
+null, invalidating the proportionality $H\p{E}\propto D\p{E}$. This
+method will therefore only ``flatten'' the histogram at energies $E$
+which have been sufficiently sampled for $H\p{E}$ to be statistically
+significant. The end goal in implementing histogram methods, however,
+is precisely to sample those energies that cannot be sufficiently
+sampled without employing clever schemes to bias Monte Carlo
+simulations. Therefore, this method does not directly help us study
+the square-well fluid. This method does, however, motivate the method
+in Section \ref{sec:simple_flat}, and provides some of the theory
+behind the methods in Sections \ref{sec:tmmc} and \ref{sec:oetmmc}.
 
@@
 The transition matrix is trivially square and positive-definite, but
-it is not symmetric. Given that Metropolis Monte Carlo samples all of
+it is not symmetric. Given that unbiased Monte Carlo samples all of
 state-space randomly, meaning all possible system states are equally
@@
 properties, the system's density of states. Consider an ensemble of
-Metropolis Monte Carlo simulations with a distribution of energies at
-a time $t$ given by $\hat D^{t}\p{E}$, such that the probability of
-any given simulation to have an energy $E_i$ at a time $t$ is $\hat
+unbiased Monte Carlo simulations with a distribution of energies at a
+time $t$ given by $\hat D^{t}\p{E}$, such that the probability of any
+given simulation to have an energy $E_i$ at a time $t$ is $\hat
 D_i^{t}=\hat D^{t}\p{E_i}$. The probability $\hat D_i^{t}$ can be
 expressed in terms of the distribution $\hat D^{t-1}\p{E}$ one time
-step (i.e. one Metropolis Monte Carlo move) prior by
+step (i.e. one Monte Carlo move) prior by
 \begin{align}
@@
 meaning $\hat D^{t}=\hat D^{t-1}$. Furthermore, the distribution of
-states in an equilibrium ensemble of Metropolis Monte Carlo
-simulations is precisely the density of states $D\p{E}$, which means
-$D_i=T_{ij}D_j$.  Simultaneously expressing this condition for all
-energies,
+states in an equilibrium ensemble of unbiased Monte Carlo simulations
+is precisely the density of states $D\p{E}$, which means
+$D_i=\sum_jT_{ij}D_j$.  Expressing this condition for all energies
+with a state density vector $D$ and transition matrix $T$,
 \begin{align}
@@
 histogram $\tilde T_{ij}=\tilde T\p{E_i,E_j}$ of the energy
-transitions after each Metropolis Monte Carlo move, i.e. after step
-\ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis}. A
+transitions after each unbiased Monte Carlo move attempt, i.e. after
+step \ref{alg:metropolis_move} of Algorithm \ref{alg:metropolis}. A
 histogram of transitions $\tilde T\p{E_i,\bar E_j}$ from a fixed
@@
 transition matrix $T$, we use Algorithm \ref{alg:tmmc}, which is a
-modified version of a routine provided in \cite{tmmc_prl}. \fixme{is
-  this citation correct?} The primary motivation behind this algorithm
-is such: as lower energies are always more difficult to sample than
-higher energies, transitions to lower energies should always be
-accepted, while transitions to higher energies should pass a
-probabilistic acceptance test. To understand the probability in step
-\ref{alg:tmmc_prob} of Algorithm \ref{alg:tmmc} of accepting a move
-$s_i\to s_f$ for which $E_f>E_i$, first consider the case of $c=0$ and
-$T_{\t{min}}\to0$, in which case the probability becomes
+modified version of a routine provided in \cite{tmmc}. The primary
+motivation behind this algorithm is such: as lower energies are always
+more difficult to sample than higher energies, transitions to lower
+energies should always be accepted, while transitions to higher
+energies should pass a probabilistic acceptance test. To understand
+the probability in step \ref{alg:tmmc_prob} of Algorithm
+\ref{alg:tmmc} of accepting a move $s_i\to s_f$ for which $E_f>E_i$,
+first consider the case of $c=0$ and $T_{\t{min}}\to0$, in which case
+the probability becomes
 \begin{align}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\nocite{*} \bibliography{thesis}
+\nocite{apsrev41Control} \bibliography{thesis}
 

### 2015-05-27 papers/histogram/thesis/thesis.tex
minor thesis revisions

@@
 theories, but unbiased Monte Carlo simulations are incapable of
-sampling the low energy fluid states which dominate the partition
-function at low liquid-state temperatures. Previous works have
-developed several generic Monte Carlo histogram methods for collecting
-statistics on low energy system states, but little or no literature
-exists on these methods' systematic comparison, as well as their
-application to the square-well fluid. The square-well fluid in
+effectively sampling the low energy fluid states which dominate the
+partition function at low liquid-state temperatures. Previous works
+have developed several generic Monte Carlo histogram methods for
+collecting statistics on low energy system states, but little or no
+literature exists on these methods' systematic comparison, as well as
+their application to the square-well fluid. The square-well fluid in
 particular introduces application challenges not manifest in
@@
 
-Employing biased Monte Carlo simulations, sketched out in Algorithm
-\ref{alg:biased_MC}, allows one to construct weight functions that
-favor some region of state space over others, as transitions to states
-with higher weights are always accepted, whereas transitions to states
-with lower weights may be rejected, artificially preventing the
-simulated system from leaving interesting regions of state
-space. Crucially, the bias introduced by weights can be reversed when
-computing system properties from sampling statistics, as the weight of
-a particular state is directly proportional to the probability bias of
-that state; that is, a state with a weight of 2 will be sampled twice
-as often as it would have been with a weight of 1. If we wish to
-convert a histogram $H\p{X}$ of observations in a biased Monte Carlo
-simulation into a numerical unnormalized density of states $\tilde
-D\p{X}$ in $X$, we therefore divide the histogram by the corresponding
-weights $w\p{X}$, i.e.
+This sort of biased simulation is called Metropolis-Hastings Monte
+Carlo sampling. Employing Metropolis-Hastings Monte Carlo simulations,
+sketched out in Algorithm \ref{alg:biased_MC}, allows one to construct
+weight functions that favor some region of state space over others, as
+transitions to states with higher weights are always accepted, whereas
+transitions to states with lower weights may be rejected, artificially
+preventing the simulated system from leaving interesting regions of
+state space. Crucially, the bias introduced by weights can be reversed
+when computing system properties from sampling statistics, as the
+weight of a particular state is directly proportional to the
+probability bias of that state; that is, a state with a weight of 2
+will be sampled twice as often as it would have been with a weight of
+1. If we wish to convert a histogram $H\p{X}$ of observations in a
+Metropolis-Hastings Monte Carlo simulation into a numerical
+unnormalized density of states $\tilde D\p{X}$ in $X$, we therefore
+divide the histogram by the corresponding weights $w\p{X}$, i.e.
 \begin{align}
@@
 \begin{align}
-  D\p{X}=\f{\tilde D\p{X}}{\sum_X\tilde D\p{X}}
-  =\f{H\p{X}/w\p{X}}{\sum_XH\p{X}/w\p{X}}.
+  D\p{X}=\f{\tilde D\p{X}}{\sum_{X'}\tilde D\p{X'}}
+  =\f{H\p{X}/w\p{X}}{\sum_{X'}H\p{X'}/w\p{X'}}.
   \label{eq:dos_norm}
@@
 \begin{algorithm}[tb]
-  \caption{Biased Monte Carlo fluid simulation}
+  \caption{Metropolis-Hastings Monte Carlo fluid simulation}
   \label{alg:biased_MC}
@@
 The most common weight array $w\p{E}$ used by physicists for what is
-called a canonical Monte Carlo simulation involves choosing a
-particular temperature $T_0$, and using weights proportional to the
-Boltzmann factor at that temperature, i.e.
+called a canonical (or Metropolis) Monte Carlo simulation involves
+choosing a particular temperature $T_0$, and using weights
+proportional to the Boltzmann factor at that temperature, i.e.
 \begin{align}
@@
 
-Having covered the basics of the square-well fluid as well as unbiased
-and biased Monte Carlo simulations, we can now discuss several
-histogram methods for constructing energy weights. In Section
-\ref{sec:flat_histogram}, we will introduce the flat histogram method,
-which is not directly applicable to the study of most systems, but
-which forms the basis of many other methods. We will then introduce,
-in Section \ref{sec:simple_flat}, the ``simple flat'' method, which we
-designed to be the simplest viable histogram method, and against which
-we can compare more complicated methods. Next, in Sections
+Having covered the basics of the square-well fluid as well as Monte
+Carlo fluid simulation, we can now discuss several histogram methods
+for constructing energy weights. In Section \ref{sec:flat_histogram},
+we will introduce the flat histogram method, which is not directly
+applicable to the study of most systems, but which forms the basis of
+many other methods. We will then introduce, in Section
+\ref{sec:simple_flat}, the ``simple flat'' method, which we designed
+to be the simplest viable histogram method, and against which we can
+compare more complicated methods. Next, in Sections
 \ref{sec:wang_landau}--\ref{sec:optimized_ensemble} we introduce three
@@
 implement and analyze because, unlike most histogram methods, it was
-designed to optimize weights rather than construct them. We will,
-however, borrow ideas from the optimized ensemble to design a
-histogram method of our own, a hybrid of the optimized ensemble and
-transition matrix Monte Carlo, in Section \ref{sec:oetmmc}. Finally,
-we will cover some general implementation details and loose ends in
-Section \ref{sec:loose_ends}
+designed to optimize pre-existing weights rather than construct
+them. We will, however, borrow ideas from the optimized ensemble to
+design a histogram method of our own, a hybrid of the optimized
+ensemble and transition matrix Monte Carlo, in Section
+\ref{sec:oetmmc}. Finally, we will cover some general implementation
+details and loose ends in Section \ref{sec:loose_ends}
 
@@
 Unlike the previous histogram methods, the transition matrix Monte
-Carlo method does not use weights, and merely compute the density of
+Carlo method does not use weights, and merely computes the density of
 states $D\p{E}$ of a system\cite{tmmc}. The density of states can in
@@
 finding the eigenvector of the transition matrix which has a
-corresponding unit eigenvalue. \fixme{why is this eigenvector unique?}
-Once one has determined the density of states of a system, one can
-initialize weights via the flat histogram method,
-i.e. (\ref{eq:flat_weights}).
+corresponding unit eigenvalue. Once one has determined the density of
+states of a system, one can initialize weights via the flat histogram
+method, i.e. (\ref{eq:flat_weights}).
 
@@
 equally. This motivation considers the quantity of statistics, but not
-their quality. Biased simulations collect statistics on low energy
-states by first getting into these states, and then rejecting moves
-which move the system into higher energy states. Statistics on low
-energy states will therefore generally be highly correlated, as they
-are based on many samples of only a few low energy regions of a
-system's energy landscape. The optimized
+their quality. Metropolis-Hastings Monte Carlo simulations collect
+statistics on low energy states by first getting into these states,
+and then rejecting moves which move the system into higher energy
+states. Statistics on low energy states will therefore generally be
+highly correlated, as they are based on many samples of only a few low
+energy regions of a system's energy landscape. The optimized
 ensemble\cite{optimized_ensemble} attempts to address the
@@
 paper. We do, however, credit \cite{optimized_ensemble} for motivating
-the hybrid OE-TMMC method in Section \ref{sec:oetmmc}.
-
-\subsection{The hybrid OE-TMMC method}
+the hybrid DOTMMC method in Section \ref{sec:oetmmc}.
+
+\subsection{The diffusion-optimized transition matrix Monte Carlo
+  (DOTMMC) method}
 \label{sec:oetmmc}
@@
 
-The hybrid OE-TMMC method thus first runs Algorithm \ref{alg:tmmc} to
-collect a transition histogram and compute flat histogram weights,
-then computes the diffusivity $\alpha\p{E}$ via
+The diffusion-optimized transition matrix Monte Carlo (DOTMMC) method,
+a hybrid of TMMC and the optimized ensemble, thus first runs Algorithm
+\ref{alg:tmmc} to collect a transition histogram and compute flat
+histogram weights, then computes the diffusivity $\alpha\p{E}$ via
 (\ref{eq:<de>})-(\ref{eq:microscopic_diffusitivy}), and finally
 computes optimized weights as given by
-(\ref{eq:optimized_weights}). Unlike the optimized ensemble, OE-TMMC
+(\ref{eq:optimized_weights}). Unlike the optimized ensemble, DOTMMC
 does not require any extra simulation routines, and is therefore
@@
 it collects $10^4$ optimistic samples of the minimum important energy
-(i.e. until $s_o\p{E_-}\ge 10^4$). \fixme{why TMMC?} To increase
-confidence in low-temperature gold standard results, we initialized
-them with $kT_{\t{min}}=0.1\epsilon$. After initialization, the gold
-standard runs simulate for $10^{10}$ iterations ($10^{10}N$
-moves). While the entire initialization and simulation process takes
-around two hours for the largest regular simulations, the largest
-golden simulations initialize for $\sim11$ hours, and simulate for a
-week. To control for iteration number in the results from any given
-set of simulations, we compare against ``converged flat''
-simulations. Converged flat simulations read in the transition matrix
-from golden simulations to construct a weight array, and simulate for
-the same number of iterations as the standard simulations.
+(i.e. until $s_o\p{E_-}\ge 10^4$). To increase confidence in
+low-temperature gold standard results, we initialized them with
+$kT_{\t{min}}=0.1\epsilon$. After initialization, the gold standard
+runs simulate for $10^{10}$ iterations ($10^{10}N$ moves). While the
+entire initialization and simulation process takes around two hours
+for the largest regular simulations, the largest golden simulations
+initialize for $\sim11$ hours, and simulate for a week. To control for
+iteration number in the results from any given set of simulations, we
+compare against ``converged flat'' simulations. Converged flat
+simulations read in the transition matrix from golden simulations to
+construct a weight array, and simulate for the same number of
+iterations as the standard simulations.
 
@@
 The energy histogram in Figure \ref{fig:arrays} shows that TMMC and
-OE-TMMC spend considerably less time at important low energies than
+DOTMMC spend considerably less time at important low energies than
 the simple flat and vanilla Wang-Landau methods. At face value, this
-behavior of TMMC and OE-TMMC is undesirable, as these are the most
+behavior of TMMC and DOTMMC is undesirable, as these are the most
 interesting energies for low temperatures, and the most difficult
@@
 statistics which are guaranteed to be uncorrelated. By this metric,
-TMMC and OE-TMMC again perform worse than simple flat and vanilla
+TMMC and DOTMMC again perform worse than simple flat and vanilla
 Wang-Landau.
@@
 \end{align}
-and the configurational entropy \fixme{is this correct?}
-\begin{align}
-  S_{\t{config}}\p{T}=k\bk{\f{E}{kT}+\ln Z}_T
-  -k\bk{\f{E}{kT}+\ln Z}_{T\to\infty}.
+and the configurational entropy
+\begin{align}
+  S_{\t{config}}\p{T}=S\p{T}-S\p{T\to\infty}
+  =k\bk{\f{E}{kT}+\ln Z}_T-k\bk{\f{E}{kT}+\ln Z}_{T\to\infty}.
   \label{eq:S}
@@
 paper, we will neglect to prove that this phase transition is indeed
-between liquid and gas phases, and take this fact as given. \fixme{say
-  more convincingly}
-
-\begin{figure}[!b]
+between liquid and gas phases, and take this fact as given.
+
+\begin{figure}[p]
   \centering
@@
   \label{fig:config_entropy}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/u-err.pdf}
-  \caption[Internal energy error scaling]
-  {Errors in specific internal energy, averaged over 30 $N=25$
-    simulations.}
-  \label{fig:internal_energy_err}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/cv-err.pdf}
-  \caption[Heat capacity error scaling]
-  {Errors in specific heat capacity, averaged over 30 $N=25$
-    simulations.}
-  \label{fig:heat_capacity_err}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/s-err.pdf}
-  \caption[Configurational entropy error scaling]
-  {Errors in specific configurational entropy, averaged over 30 $N=25$
-    simulations.}
-  \label{fig:config_entropy_err}
 \end{figure}
@@
 
+\begin{figure}[!b]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/u-err.pdf}
+  \caption[Internal energy error scaling]
+  {Errors in specific internal energy, averaged over 30 $N=25$
+    simulations.}
+  \label{fig:internal_energy_err}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/cv-err.pdf}
+  \caption[Heat capacity error scaling]
+  {Errors in specific heat capacity, averaged over 30 $N=25$
+    simulations.}
+  \label{fig:heat_capacity_err}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/s-err.pdf}
+  \caption[Configurational entropy error scaling]
+  {Errors in specific configurational entropy, averaged over 30 $N=25$
+    simulations.}
+  \label{fig:config_entropy_err}
+\end{figure}
+
 Figures \ref{fig:u_scaling}--\ref{fig:s_scaling} provide a more
@@
 high-quality golden data to construct weight arrays, and should
-therefore outperform all standard simulations. \fixme{comment on why
-  some methods might do better than CFW?}
+therefore outperform all standard simulations.
 
@@
 for the minimum temperature $kT_{\t{min}}=0.2\epsilon$. Finally, these
-error figures further confirm earlier suspicions of OE-TMMC's
+error figures further confirm earlier suspicions of DOTMMC's
 underperformance relative to TMMC, despite the fact that it is
@@
 computation time and resources. Our implementation of the optimized
-ensemble using the transition matrix resulted in lower sampling rates
-and higher errors than the flat histogram TMMC weights. We suspect
-that the optimized ensemble is failing to improve TMMC because the
-theory behind it neglects the discretization of energies.
-
-\fixme{make conclusions more satisfying?}
+ensemble using the transition matrix, i.e. DOTMMC, resulted in lower
+sampling rates and higher errors than the flat histogram TMMC
+weights. Though we have not completely ruled out that our
+implementation of DOTMMC contains bugs, we suspect that the optimized
+ensemble is failing to improve TMMC either because
+\begin{enumerate*}[label=\roman*)]
+\item the theory behind the optimized ensemble neglects the
+  discretization of energies, or
+\item DOTMMC is highly sensitive to sampling error in the transition
+  matrix.
+\end{enumerate*}
+The first of these suspicions can be tested by simulating larger
+systems with a denser specific energy spectrum, and for which
+discretization should become asymptotically negligible; the second
+suspicion can be tested simply by initializing via TMMC for longer,
+thereby reducing errors in the transition matrix, before computing
+diffusion-optimized weights.
 
@@
 
-I wish to thank, first and foremost, my advisor Dr. David Roundy, for
+I wish to thank, first and foremost my advisor, David Roundy, for
 mentoring me and providing me a wealth of knowledge and expertise in
@@
 polyhedra Monte Carlo simulations I used as the template for my own
-codes. Lastly, I would like to thank thank Eva Morgun, who prompted me
-to {\it act} on my textbook enthusiasm for physics, and get involved
-in the fun and interesting world of scientific research. \fixme{Thank
-  Jansen and McIntyre}
+square-well codes. Lastly, I would like to thank thank Eva Morgun, who
+prompted me to {\it act} on my textbook enthusiasm for physics, and
+get involved in the fun and interesting world of scientific
+research. \fixme{Thank Jansen and McIntyre}
 

### 2015-05-28 papers/histogram/thesis/thesis.tex

@@
 
-\subsection{The Wang-Landau (WL) method}
+\subsection{The Wang-Landau method}
 \label{sec:wang_landau}
@@
       \begin{align*}
-        P_m\p{s_i\to s_f} =\max\set{\f{\p{\tilde T_d\p{E_i,\Delta
-                E}+c}/\p{n_i+c}} {\p{\tilde T_d\p{E_f,-\Delta
-                E}+c}/\p{n_f+c}}, \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
+        P_m\p{s_i\to s_f}=\max\set{\f{\p{\tilde T_d\p{E_i,\Delta E}+c}
+            /\p{n_i+c}} {\p{\tilde T_d\p{E_f,-\Delta E}+c}/\p{n_f+c}},
+          \exp\p{-\f{\Delta E}{kT_{\t{min}}}}}
       \end{align*}
@@
 step \ref{alg:tmmc_prob} to modify (\ref{eq:tmmc_prob_simplified}).
-This parameter makes the rejection of transitions more conservative
-when there are too few statistics in the histogram $\tilde T_d$ to be
-confident in the current estimate of the transition matrix. In order
-to make the effects of $c$ negligible as $\tilde T_d$ collects more
-statistics throughout initialization, $c$ should be of order unity,
-though its optimal value is generally system-dependent. In our
-implementation of Algorithm \ref{alg:tmmc}, we use $c=16$.
+This parameter, which is our own addition to the algorithm provided in
+\cite{tmmc}, makes the rejection of transitions more conservative when
+there are too few statistics in the histogram $\tilde T_d$ to be
+confident in the current estimate of the transition matrix. Without
+this parameter, the TMMC initialization routine can get stuck at
+energies which, due to poor collection of statistics in $\tilde T_d$,
+appear to be more important than they actually are. In order to make
+the effects of $c$ negligible as $\tilde T_d$ collects more statistics
+throughout initialization, $c$ should be of order unity, though its
+optimal value is generally system-dependent. In our implementation of
+Algorithm \ref{alg:tmmc}, we use $c=16$.
 
@@
 ``natural'' end condition does not advantage it over other methods,
-however, as the WL end condition makes use of a free parameter (the
-cutoff $c$), which any end condition will require. In order to compare
-all methods fairly, we need a method-independent end condition to
-initialization. The most straightforward of such end conditions is to
-simply to stop initializing after some fixed number of Monte Carlo
-iterations, and finish up any other calculations in the algorithm if
-appropriate. We thus initialize systems of $N$ spheres for $10^5N^2$
-iterations; the choice of $10^5N^2$ we address in Section
-\ref{sec:results}. In general, however, one might want an end
+however, as the Wang-Landau end condition makes use of a free
+parameter (the cutoff $c$), which any end condition will require. In
+order to compare all methods fairly, we need a method-independent end
+condition to initialization. The most straightforward of such end
+conditions is to simply to stop initializing after some fixed number
+of Monte Carlo iterations, and finish up any other calculations in the
+algorithm if appropriate. We thus initialize systems of $N$ spheres
+for $10^5N^2$ iterations; the choice of $10^5N^2$ we address in
+Section \ref{sec:results}. In general, however, one might want an end
 condition which is independent of system size. We came up with a four
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[width=\textwidth]{figs/array-example.pdf}
+  \caption[Example arrays]
+  {Example arrays computed by various histogram methods for a
+    square-well fluid with a well width $\lambda=1.3$, filling
+    fraction $\eta=0.3$, and $N=25$ spheres after simulating for
+    $10^5N^2$ iterations: (top left) energy histogram, (top right)
+    weight array, (bottom left) density of states. Vertical reference
+    lines denote the minimum important energy $E_-$.}
+  \label{fig:arrays}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.8\textwidth]{figs/sample-rates.pdf}
+  \caption[Example energy sampling rates]
+  {Example pessimistic energy sampling rates computed by various
+    histogram methods. Sample rates for each method are averaged over
+    30 simulations with different random number seeds. The vertical
+    reference line denotes the minimum important energy $E_-$.}
+  \label{fig:sample_rates}
+\end{figure}
+
 Most of our analysis will use quantities computed from two arrays: the
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[width=\textwidth]{figs/array-example.pdf}
-  \caption[Example arrays]
-  {Example arrays computed by various histogram methods for a
-    square-well fluid with a well width $\lambda=1.3$, filling
-    fraction $\eta=0.3$, and $N=25$ spheres after simulating for
-    $10^5N^2$ iterations: (top left) energy histogram, (top right)
-    weight array, (bottom left) density of states. Vertical reference
-    lines denote the minimum important energy $E_-$.}
-  \label{fig:arrays}
-\end{figure}
-
-\begin{figure}
-  \centering
-  \includegraphics[width=0.8\textwidth]{figs/sample-rates.pdf}
-  \caption[Example energy sampling rates]
-  {Example pessimistic energy sampling rates computed by various
-    histogram methods. Sample rates for each method are averaged over
-    30 simulations with different random number seeds. The vertical
-    reference line denotes the minimum important energy $E_-$.}
-  \label{fig:sample_rates}
-\end{figure}
-
 \subsection{Thermodynamic properties and errors}
@@
 of the square-well fluid.
+
+\begin{figure}[!b]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/internal-energy.pdf}
+  \caption[Internal energy]
+  {Specific internal energy curves averaged over 30 $N=25$
+    simulations.}
+  \label{fig:internal_energy}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/heat-capacity.pdf}
+  \caption[Heat capacity]
+  {Specific heat capacity curves averaged over 30 $N=25$ simulations.}
+  \label{fig:heat_capacity}
+\end{figure}
+
+\begin{figure}[p]
+  \centering
+  \includegraphics[width=0.75\textwidth]{figs/config-entropy.pdf}
+  \caption[Configurational entropy]
+  {Specific configurational entropy curves averaged over 30 $N=25$
+    simulations.}
+  \label{fig:config_entropy}
+\end{figure}
 
@@
 
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/internal-energy.pdf}
-  \caption[Internal energy]
-  {Specific internal energy curves averaged over 30 $N=25$
-    simulations.}
-  \label{fig:internal_energy}
-\end{figure}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/heat-capacity.pdf}
-  \caption[Heat capacity]
-  {Specific heat capacity curves averaged over 30 $N=25$ simulations.}
-  \label{fig:heat_capacity}
-\end{figure}
-
-\begin{figure}[p]
-  \centering
-  \includegraphics[width=0.75\textwidth]{figs/config-entropy.pdf}
-  \caption[Configurational entropy]
-  {Specific configurational entropy curves averaged over 30 $N=25$
-    simulations.}
-  \label{fig:config_entropy}
-\end{figure}
-
 Having a sense of how $U$, $C_V$ and $S_{\t{config}}$ behave with
@@
 
-\begin{figure}[p]
+\begin{figure}[!p]
   \centering
@@
 are outperforming the converged flat ones. We do not understand why
-such events occur, as converged flat simulations should be using
-high-quality golden data to construct weight arrays, and should
-therefore outperform all standard simulations.
+such events occur, as converged flat simulations should be using high
+quality golden data to construct weight arrays, and should therefore
+outperform all standard simulations.
 

### 2015-05-30 papers/histogram/thesis/thesis.tex
changed bibliography referencing style, and formatted thesis for the honors college

@@
   \renewcommand{\contentsname}{Table of Contents}
-  \renewcommand\refname{\bf References}}
+  \renewcommand\refname{References}}
 \usepackage{tocloft}
@@
 
+\usepackage{xspace}
+
 \renewcommand{\title}{Optimizing Monte Carlo simulation of the
-  square-well fluid}
-\renewcommand{\author}{Michael A. Perlin}
-\newcommand{\institution}{Oregon State University}
-\newcommand{\department}{Department of Physics}
-\newcommand{\supervisor}{Dr. David Roundy}
-\renewcommand{\date}{15 May 2015}
+  square-well fluid\xspace}
+\renewcommand{\author}{Michael A. Perlin\xspace}
+\newcommand{\degree}{Honors Baccalaureate of Science
+  in Physics\xspace}
+\renewcommand{\date}{02 June 2015\xspace}
+\newcommand{\email}{\href{mailto:mika.perlin@gmail.com}{mika.perlin@gmail.com}\xspace}
 
@@
 
-\fancypagestyle{abstract}{
-  \rhead{}
-  \lhead{}
+\renewcommand{\titlepage}{
+  { \centering
+
+    \title \\
+    ~\\~\\
+    by \\
+    \author \\
+    ~\\~\\~\\~\\~\\~\\~\\~\\
+    A PROJECT \\
+    ~\\~\\
+    submitted to \\
+    ~\\
+    Oregon State University \\
+    ~\\
+    University Honors College \\
+    ~\\~\\~\\~\\~\\~\\
+    in partial fulfillment of \\
+    the requirements for the \\
+    degree of \\
+    ~\\~\\
+    \degree \\
+    (Honors Scholar) \\
+    ~\\~\\~\\~\\~\\
+    Presented \date \\
+    Commencement June 2015 \\
   }
+  \newpage
+}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{titlepage}
-
-  \newcommand{\HRule}{\rule{\linewidth}{0.5mm}}
-
-  \center
-
-  \textsc{\LARGE \institution}\\[1.5cm]
-  \textsc{\Large \department}\\[0.5cm]
-
-  \HRule \\[0.4cm]
-  { \huge \bfseries \title}\\[0.4cm]
-  \HRule \\[1.5cm]
-
-  \begin{minipage}{0.4\textwidth}
-    \begin{flushleft} \large
-      \emph{Author:}\\
-      \author
-    \end{flushleft}
-  \end{minipage}
-  ~
-  \begin{minipage}{0.4\textwidth}
-    \begin{flushright} \large
-      \emph{Supervisor:} \\
-      \supervisor
-    \end{flushright}
-  \end{minipage}\\[4cm]
-
-  {\large\date}
-
-  \vfill
-
-\end{titlepage}
-
-\thispagestyle{empty}
-
-\newpage
-
-\tableofcontents
-
-\thispagestyle{empty}
-
-\newpage
-
-\listof{algorithm}{\Large List of Algorithms}
-
-\listof{figure}{\Large List of Figures}
-
-\thispagestyle{empty}
-
-\newpage
-
-\setcounter{page}{1}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section*{Abstract}
-\label{sec:abstract}
-
-\thispagestyle{abstract}
+\setlength\parindent{0mm}
+\pagestyle{empty}
+
+\titlepage
+
+~\newpage
+
+\begin{center}
+  AN ABSTRACT OF THE THESIS OF
+\end{center}
+
+\author for the degree of \underline{\degree} presented on
+\date. Title: \underline{\title}.
+
+~\\~\\~\\~\\
+
+Abstract approved:
+
+~\\
+\hrule
+\begin{center}
+  David Roundy
+\end{center}
+
+~\\
 
@@
 
+~
+
 We implement our own ``simple flat'' method, the Wang-Landau method,
@@
 
+~\\
+
+Key words: Monte Carlo methods, Square-well fluid, histogram methods
+
+~
+
+Corresponding e-mail address: \email
+
 \newpage
+
+~\vspace{3cm}
+
+\begin{center}
+  \textcopyright Copyright by \author
+
+  \date
+
+  All Rights Reserved
+\end{center}
+
+\newpage
+
+\titlepage
+
+\underline{\degree} project of \author presented on \date.
+
+\vspace{2cm}
+
+APPROVED:
+
+\vspace{2cm}
+
+\hrule ~\\
+David Roundy, Mentor, representing the Department of Physics
+
+\vspace{1.5cm}
+
+\hrule ~\\
+Henri J. F. Jansen, Committee Member, representing the Department of
+Physics
+
+\vspace{1.5cm}
+
+\hrule ~\\
+David H. McIntyre, Committee Member, representing the Department of
+Physics
+
+\vspace{2.5cm}
+
+\hrule ~\\
+Toni Doolen, Dean, University Honors College
+
+\vspace{3cm}
+
+I understand that my project will become part of the permanent
+collection of Oregon State University, University Honors College.  My
+signature below authorizes release of my project to any reader upon
+request.
+
+\vspace{2cm}
+
+\hrule
+\begin{center}
+  \author, Author
+\end{center}
+
+\newpage
+
+\tableofcontents
+
+\thispagestyle{empty}
+
+\newpage
+
+\listof{algorithm}{\Large List of Algorithms}
+
+\listof{figure}{\Large List of Figures}
+
+\thispagestyle{empty}
+
+\newpage
+
+\pagestyle{fancy}
+\setlength\parindent{5mm}
+\setcounter{page}{1}
 
@@
   \item Construct an initial ``typical'' (i.e. non-ordered) fluid
-    configuration. \fixme{explain how?}
+    configuration.
 
@@
 The Wang-Landau method is the first published histogram method we
-discuss\cite{wang_landau, wang_landau_mod, wang_landau_analysis}. This
-method has been used to study a wide variety of systems\fixme{which
-  systems? cite}, and is the standard method\fixme{cite} to achieve
-broad energy sampling via Monte Carlo simulations.
+discuss\cite{wang2001determining, belardinelli2007fast,
+  zhou2005understanding, landau2004new}. This method has been used to
+study a wide variety of systems, and is the standard method to compute
+the density of states of a system via Monte Carlo
+simulations\cite{landau2004new, yamaguchi2001three,
+  parsons2006globule, parsons2006off}.
 
@@
 works typically neglect providing any heuristic or formula for $n$,
-reporting a system-independent $n=10^3$\cite{wang_landau_mod} or
-$10^5$\cite{wang_landau}.
+reporting a system-independent $n=10^3$\cite{belardinelli2007fast} or
+$10^5$\cite{wang2001determining}.
 
@@
 heuristics for assigning $f_0$, we use $f_0=e$ (more precisely, $\ln
-f_0=1$), a value commonly suggested by literature\cite{wang_landau}.
+f_0=1$), a value commonly suggested by
+literature\cite{wang2001determining}.
 
@@
 to satisfy these requirements. Previous works have reported using
-values such as $u=2$ \cite{wang_landau, wang_landau_mod} and $u=10$
-\cite{wang_landau_analysis}; we use the former value.
+values such as $u=2$ \cite{wang2001determining, belardinelli2007fast}
+and $u=10$ \cite{zhou2005understanding}; we use the former value.
 
@@
 efficient and effective implementation of this method. Some previous
-works\cite{wang_landau} suggest that $C\sp{H}$ should be made as
-stringent as possible, cautioning only that some simulations might not
-satisfy too strict of a flatness condition in any reasonable amount of
-time. These works fail to recognize the reliance of Wang-Landau on the
-{\it failure} of $C\sp{H}$ to enforce a perfectly flat histogram. A
-commonly suggested flatness condition is that the minimum histogram
-value in some specified range of energies be at least some fixed
-proportion of the mean histogram value, i.e.
+works\cite{wang2001determining} suggest that $C\sp{H}$ should be made
+as stringent as possible, cautioning only that some simulations might
+not satisfy too strict of a flatness condition in any reasonable
+amount of time. These works fail to recognize the reliance of
+Wang-Landau on the {\it failure} of $C\sp{H}$ to enforce a perfectly
+flat histogram. A commonly suggested flatness condition is that the
+minimum histogram value in some specified range of energies be at
+least some fixed proportion of the mean histogram value, i.e.
 $C\sp{H}:\min\sp{H\p{E}}\ge x\bk{H}_E$, and several papers report
@@
 Carlo method does not use weights, and merely computes the density of
-states $D\p{E}$ of a system\cite{tmmc}. The density of states can in
-turn be used to determine flat histogram weights via
+states $D\p{E}$ of a system\cite{wang2002transition}. The density of
+states can in turn be used to determine flat histogram weights via
 (\ref{eq:flat_weights}). TMMC introduces a new object: the energy
@@
 transition matrix $T$, we use Algorithm \ref{alg:tmmc}, which is a
-modified version of a routine provided in \cite{tmmc}. The primary
-motivation behind this algorithm is such: as lower energies are always
-more difficult to sample than higher energies, transitions to lower
-energies should always be accepted, while transitions to higher
-energies should pass a probabilistic acceptance test. To understand
-the probability in step \ref{alg:tmmc_prob} of Algorithm
-\ref{alg:tmmc} of accepting a move $s_i\to s_f$ for which $E_f>E_i$,
-first consider the case of $c=0$ and $T_{\t{min}}\to0$, in which case
-the probability becomes
+modified version of a routine provided in
+\cite{wang2002transition}. The primary motivation behind this
+algorithm is such: as lower energies are always more difficult to
+sample than higher energies, transitions to lower energies should
+always be accepted, while transitions to higher energies should pass a
+probabilistic acceptance test. To understand the probability in step
+\ref{alg:tmmc_prob} of Algorithm \ref{alg:tmmc} of accepting a move
+$s_i\to s_f$ for which $E_f>E_i$, first consider the case of $c=0$ and
+$T_{\t{min}}\to0$, in which case the probability becomes
 \begin{align}
@@
 via Algorithm \ref{alg:tmmc} to yield a flat energy
-histogram\cite{tmmc}. Unlike the previous histogram methods, we desire
-a flat histogram during initialization via Algorithm \ref{alg:tmmc} in
-order to broadly sample the transition histogram $\tilde
-T_d\p{E,\Delta E}$, rather than to construct weights $w\p{E}$
-directly.
+histogram\cite{wang2002transition}. Unlike the previous histogram
+methods, we desire a flat histogram during initialization via
+Algorithm \ref{alg:tmmc} in order to broadly sample the transition
+histogram $\tilde T_d\p{E,\Delta E}$, rather than to construct weights
+$w\p{E}$ directly.
 
@@
 This parameter, which is our own addition to the algorithm provided in
-\cite{tmmc}, makes the rejection of transitions more conservative when
-there are too few statistics in the histogram $\tilde T_d$ to be
-confident in the current estimate of the transition matrix. Without
-this parameter, the TMMC initialization routine can get stuck at
-energies which, due to poor collection of statistics in $\tilde T_d$,
-appear to be more important than they actually are. In order to make
-the effects of $c$ negligible as $\tilde T_d$ collects more statistics
-throughout initialization, $c$ should be of order unity, though its
-optimal value is generally system-dependent. In our implementation of
-Algorithm \ref{alg:tmmc}, we use $c=16$.
+\cite{wang2002transition}, makes the rejection of transitions more
+conservative when there are too few statistics in the histogram
+$\tilde T_d$ to be confident in the current estimate of the transition
+matrix. Without this parameter, the TMMC initialization routine can
+get stuck at energies which, due to poor collection of statistics in
+$\tilde T_d$, appear to be more important than they actually are. In
+order to make the effects of $c$ negligible as $\tilde T_d$ collects
+more statistics throughout initialization, $c$ should be of order
+unity, though its optimal value is generally system-dependent. In our
+implementation of Algorithm \ref{alg:tmmc}, we use $c=16$.
 
@@
 energy regions of a system's energy landscape. The optimized
-ensemble\cite{optimized_ensemble} attempts to address the
+ensemble\cite{trebst2004optimizing} attempts to address the
 autocorrelation between low energy samples by finding weights to
@@
 assumption that $\alpha\p{E}$ does not strongly depend on the weights
-$w\p{E}$, the authors of \cite{optimized_ensemble} find that the
+$w\p{E}$, the authors of \cite{trebst2004optimizing} find that the
 optimal walker density $n$ is
@@
 
-The algorithm provided in \cite{optimized_ensemble} to construct
+The algorithm provided in \cite{trebst2004optimizing} to construct
 weights given by (\ref{eq:optimized_weights}) finds the diffusivity
@@
 not be comparing the optimized ensemble to other methods in this
-paper. We do, however, credit \cite{optimized_ensemble} for motivating
-the hybrid DOTMMC method in Section \ref{sec:oetmmc}.
-
-\subsection{The diffusion-optimized transition matrix Monte Carlo
-  (DOTMMC) method}
+paper. We do, however, credit \cite{trebst2004optimizing} for
+motivating the hybrid OETMMC method in Section \ref{sec:oetmmc}.
+
+\subsection{The hybrid OETMMC method}
 \label{sec:oetmmc}
@@
 (\ref{eq:walker_current_definition}) is considered outside the scope
-of this paper. \fixme{say something more convincing? cite?}
-
-The diffusion-optimized transition matrix Monte Carlo (DOTMMC) method,
-a hybrid of TMMC and the optimized ensemble, thus first runs Algorithm
-\ref{alg:tmmc} to collect a transition histogram and compute flat
-histogram weights, then computes the diffusivity $\alpha\p{E}$ via
-(\ref{eq:<de>})-(\ref{eq:microscopic_diffusitivy}), and finally
-computes optimized weights as given by
-(\ref{eq:optimized_weights}). Unlike the optimized ensemble, DOTMMC
+of this paper.
+
+The OETMMC method, a hybrid of TMMC and the optimized ensemble, first
+runs Algorithm \ref{alg:tmmc} to collect a transition histogram and
+compute flat histogram weights, then computes the diffusivity
+$\alpha\p{E}$ via (\ref{eq:<de>})-(\ref{eq:microscopic_diffusitivy}),
+and finally computes optimized weights as given by
+(\ref{eq:optimized_weights}). Unlike the optimized ensemble, OETMMC
 does not require any extra simulation routines, and is therefore
@@
 because it is comfortably below the temperature at which the
-square-well fluid condenses from a gas into a liquid (see Section
-\ref{sec:properties}). \fixme{what about pressure?}
+square-well fluid condenses from a gas into a liquid in simulation
+(see Section \ref{sec:properties}).
 
@@
 The energy histogram in Figure \ref{fig:arrays} shows that TMMC and
-DOTMMC spend considerably less time at important low energies than
-the simple flat and vanilla Wang-Landau methods. At face value, this
-behavior of TMMC and DOTMMC is undesirable, as these are the most
+OETMMC spend considerably less time at important low energies than the
+simple flat and vanilla Wang-Landau methods. At face value, this
+behavior of TMMC and OETMMC is undesirable, as these are the most
 interesting energies for low temperatures, and the most difficult
@@
 statistics which are guaranteed to be uncorrelated. By this metric,
-TMMC and DOTMMC again perform worse than simple flat and vanilla
+TMMC and OETMMC again perform worse than simple flat and vanilla
 Wang-Landau.
@@
 \begin{align}
-  S_{\t{config}}\p{T}=S\p{T}-S\p{T\to\infty}
-  =k\bk{\f{E}{kT}+\ln Z}_T-k\bk{\f{E}{kT}+\ln Z}_{T\to\infty}.
+  S_{\t{config}}\p{T}=S\p{T}-S\p{T\to\infty} =k\bk{\f{E}{kT}+\ln
+    Z}_T-k\bk{\f{E}{kT}+\ln Z}_{T\to\infty}.
   \label{eq:S}
@@
 for the minimum temperature $kT_{\t{min}}=0.2\epsilon$. Finally, these
-error figures further confirm earlier suspicions of DOTMMC's
+error figures further confirm earlier suspicions of OETMMC's
 underperformance relative to TMMC, despite the fact that it is
@@
 computation time and resources. Our implementation of the optimized
-ensemble using the transition matrix, i.e. DOTMMC, resulted in lower
+ensemble using the transition matrix, i.e. OETMMC, resulted in lower
 sampling rates and higher errors than the flat histogram TMMC
 weights. Though we have not completely ruled out that our
-implementation of DOTMMC contains bugs, we suspect that the optimized
+implementation of OETMMC contains bugs, we suspect that the optimized
 ensemble is failing to improve TMMC either because
@@
   discretization of energies, or
-\item DOTMMC is highly sensitive to sampling error in the transition
+\item OETMMC is highly sensitive to sampling error in the transition
   matrix.
@@
 get involved in the fun and interesting world of scientific
-research. \fixme{Thank Jansen and McIntyre}
+research.
 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\nocite{apsrev41Control} \bibliography{thesis}
+\nocite{apsrev41Control} \bibliography{thesis} \lhead{}
 

### 2015-06-01 papers/histogram/thesis/thesis.tex
trimming a couple dependencies for thesis.tex
shortened abstract for the honors college; final version of thesis

@@
 %%% figures
-\usepackage{graphicx,float,subcaption} % floats, etc.
-\usepackage{footnote} % for footnotes in floats
+\usepackage{graphicx,float} % floats, etc.
 \usepackage[font=small,labelfont=bf]{caption} % caption text options
@@
 \renewcommand{\date}{02 June 2015\xspace}
-\newcommand{\email}{\href{mailto:mika.perlin@gmail.com}{mika.perlin@gmail.com}\xspace}
+\newcommand{\email}{\href{mailto:mika.perlin@gmail.com}
+  {mika.perlin@gmail.com}\xspace}
 
@@
 
-\author for the degree of \underline{\degree} presented on
-\date. Title: \underline{\title}.
+\underline{\author} for the degree of \underline{\degree} presented on
+\underline{\date}. Title: \underline{\title}.
 
@@
 
-We identify and develop efficient numerical methods for determining
+We identify and develop efficient Monte Carlo methods for determining
 thermodynamic properties of the square-well fluid in order to test
-square-well density functional theories. The liquid-vapor phase
-interface is an interesting regime for testing density functional
-theories, but unbiased Monte Carlo simulations are incapable of
-effectively sampling the low energy fluid states which dominate the
-partition function at low liquid-state temperatures. Previous works
-have developed several generic Monte Carlo histogram methods for
-collecting statistics on low energy system states, but little or no
-literature exists on these methods' systematic comparison, as well as
-their application to the square-well fluid. The square-well fluid in
-particular introduces application challenges not manifest in
-traditional models for testing and benchmarking such numerical
-techniques (e.g. the Ising model).
+square-well density functional theories near the critical
+point. Previous works have developed generic so-called histogram
+methods for collecting statistics on low energy system states, but
+little or no literature exists on these methods' systematic
+comparison, as well as their application to the square-well fluid. The
+square-well fluid in particular introduces application challenges not
+manifest in traditional models for testing and benchmarking such
+numerical techniques (e.g. the Ising model).
 
@@
 We implement our own ``simple flat'' method, the Wang-Landau method,
-transition matrix Monte Carlo, and a modified version of the optimized
-ensemble. Performance of each method is determined by low energy
-sampling rates and maximum errors in several computed system
+transition matrix Monte Carlo (TMMC), and a modified version of the
+optimized ensemble. The performance of each method is measured by low
+energy sampling rates and maximum errors in computed system
 properties. We find that Wang-Landau potentially performs better than
 other histogram methods, but fails catastrophically without a
-predetermined energy range. The simple flat method and transition
-matrix Monte Carlo give results comparable to successful Wang-Landau
-simulations, but simple flat has some anomalous cases with large
-errors. Finally, our implementation of the optimized ensemble using a
-transition matrix results in worse performance over the
-straightforward transition matrix Monte Carlo method.
+predetermined energy range. The simple flat method and TMMC give
+results comparable to successful Wang-Landau simulations, but simple
+flat has some anomalous cases with large errors. Finally, our
+implementation of the optimized ensemble using a transition matrix
+results in worse performance over the straightforward TMMC.
 
@@
 
-Key words: Monte Carlo methods, Square-well fluid, histogram methods
+Key words: Monte Carlo methods, Square-well fluid, Histogram methods
 
@@
 
-\underline{\degree} project of \author presented on \date.
+\underline{\degree} project of \underline{\author} presented on
+\underline{\date}.
 
@@
 \hrule ~\\
-David Roundy, Mentor, representing the Department of Physics
+David Roundy, Mentor, representing Physics
 
@@
 \hrule ~\\
-Henri J. F. Jansen, Committee Member, representing the Department of
-Physics
+Henri J. F. Jansen, Committee Member, representing Physics
 
@@
 \hrule ~\\
-David H. McIntyre, Committee Member, representing the Department of
-Physics
+David H. McIntyre, Committee Member, representing Physics
 
@@
 
+\listof{figure}{\Large List of Figures}
+
 \listof{algorithm}{\Large List of Algorithms}
-
-\listof{figure}{\Large List of Figures}
 
@@
 
-Algorithm \ref{alg:wang_landau} spells out the Wang-Landau
-method. This algorithm includes four free parameters which affect
-initialization process: the number of iterations $n$ to run in step
-\ref{alg:wang_landau_sim}, the initial factor $f_0$, the update factor
-$u$, and the flatness condition $C\sp{H}$. The cutoff $c$ determines
-the condition for Algorithm \ref{alg:wang_landau} to terminate.
+Algorithm \ref{alg:wang_landau} walks through the Wang-Landau
+initialization procedure. This algorithm includes four free parameters
+which affect initialization process: the number of iterations $n$ to
+run in step \ref{alg:wang_landau_sim}, the initial factor $f_0$, the
+update factor $u$, and the flatness condition $C\sp{H}$. The cutoff
+$c$ determines the condition for Algorithm \ref{alg:wang_landau} to
+terminate.
 

### 2015-07-14 papers/square-well-fluid/paper.tex
Added radial distribution plotting for square well fluid. Data is currently junk. Tested on knightley. Took 0:19.59.
add homogeneous p-vs-eta plot for sw dft Tested on knightley. Took 0:24.96.
include coexistence plot in sw paper Tested on darcy. Took 0:59.90.
show g(r) fit figure in sw paper Tested on darcy. Took 1:00.46.
first guess for scalefunction, and add some discussion to sw paper Tested on darcy. Took 2:51.96.

@@
 \usepackage{color}
+
+\usepackage{dcolumn}
+\newcolumntype{d}[1]{D{.}{.}{#1}}
 
@@
 
-\author{Student One}
+\author{Eric Krebs}
 \author{Michael A. Perlin}
@@
 \section{Theory}
+
+\subsection{FIXME fit for g(r)}
+
+\begin{figure}
+  \begin{center}
+    \includegraphics[width=\columnwidth]{figs/short-range-ghs-alt}
+  \end{center}
+  \caption{g(r) fit.}
+  \label{fig:gr}
+\end{figure}
+
+We have found a problem with the pair distribution function developed
+in reference~\citenum{lurie2014approach}, which is that this pair
+distribution function can diverge as the contact value of the pair
+distribution function $g_\sigma$ diverges, as happens at high
+densities.  While the published pair distribution function does
+provide a good approximation at reasonable densities, this divergence
+causes functionals constructed with it to exhibit non-physical
+densities.  We have fixed this problem by modifying the
+pair distribution function at densities at which the hard-sphere fluid
+is crystalline.
+
+\input{figs/short-range-fit-parameters}
+
+\begin{table}
+  \begin{align*}
+    \kappa &= \kappatable
+  \end{align*}
+  \caption{The fitted $\kappa_{ij}$ matrix.  Note that the first
+    column of $\kappa$ corresponds to $j=1$ while the first row
+    corresponds to $i=1$ in Eq.~\ref{eq:kappa-defined}.  }\label{tab:kappa}
+\end{table}
+
+I should write more about the form of the fit, or move this where
+appropriate.  We have performed a modified fit to the Monte Carlo
+radial distribution function using the following functional form
+\begin{align}
+  g_S(r;g_\sigma) &=  1 + \sum_i \gamma_i(g_\sigma) \left(\frac{r}{\sigma}-1\right)^i,
+  \label{eq:new-fit-form}
+\end{align}
+where the polynomial coefficients in $g_S(r;g_\sigma)$ have the
+following dependency on $g_\sigma$:
+\begin{equation}
+  \gamma_i(g_\sigma) = f(g_\sigma)
+  \begin{cases}
+    \displaystyle
+    g_\sigma - 1 & i = 0 \\
+    \displaystyle
+    \sum_{j=1}^{4} (g_\sigma
+    - 1)^j \kappa_{ji} & i > 0 \label{eq:kappa-defined}
+  \end{cases}
+\end{equation}
+with
+\begin{equation}
+  f(g_\sigma) = \frac{1 - \tanh(4 \frac{h_\sigma -
+      \hsigmastop}{\hsigmastop - \hsigmastart})}{2}
+\end{equation}
+This function is constructed to quickly approach 1 in the fluid for
+where $\eta < \etastart$ and to approach 0 in the crystalline form
+where $\eta > \etastop$.  The largest error in this fit (an error of
+$\maxerr$) is at the maximum density considered with $\eta =
+\etamaxerr$, and is at a distance $r = \rmaxerr R$.  As is
+Ref.~\citenum{lurie2014approach}, we constrained our fit to radii less
+than $\maxrfit R$, so the maximum error appears at close to the
+maximum allowed distance.  We show the fitted values of $\kappa_{ji}$
+in Table~\ref{tab:kappa}.
 
@@
   \frac{1}{k^2}\left[ \sin (k\lambda\sigma) - k\lambda\sigma
-    \cos (k\lambda\sigma) - \sin (k\sigma) + k\sigma \cos (k-sigma) \right]\\
+    \cos (k\lambda\sigma) - \sin (k\sigma) + k\sigma \cos (k\sigma) \right]\\
   \int_\sigma^{\lambda \sigma} r
@@
  \sigma^2(\lambda-1))\cos(k\lambda\sigma)-k\sigma(\sin(k\sigma)
- +(1-2\lambda) \sin(k\lambda\sigma)) -\cos(k\sigma) \right] \\
+ +(1-2\lambda) \sin(k\lambda\sigma)) -2\cos(k\sigma) \right] \\
   \int_\sigma^{\lambda \sigma} r
@@
    &+24\cos(k\sigma) + (6k^2\sigma^2(\lambda-1)(2\lambda-1)-\lambda
-   k^4\sigma^4(\lambda-1)^3)\cos(k\lambda\sigma)\big] \\
+   k^4\sigma^4(\lambda-1)^3-24)\cos(k\lambda\sigma)\big] \\
   \int_\sigma^{\lambda \sigma} r
@@
    (24(5\lambda-4)-4k^2\sigma^2(\lambda-1)^2 (5\lambda-2) +
-   \lambda k^4\sigma^4(\lambda-1)^4))\cos(k\lambda\sigma) \big]
+   \lambda k^4\sigma^4(\lambda-1)^4)\cos(k\lambda\sigma)) \big]
 \end{align}
@@
 
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/coexistence-ww1_3}
+\end{center}
+\caption{Pressure.}
+\label{fig:coexistencec}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/p-vs-eta}
+\end{center}
+\caption{Pressure.}
+\label{fig:p-vs-eta}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/radial-sw-20}
+\end{center}
+\caption{RDF.}
+\label{fig:homogeneous-20}
+\end{figure}
+
+\begin{figure}[h]
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/radial-sw-30}
+\end{center}
+\caption{RDF.}
+\label{fig:homogeneous-30}
+\end{figure}
+
 \fixme{Here we present the comparison between various theories and in

### 2015-07-14 papers/square-well-fluid/figs/short-range-ghs.py

@@ -40,8 +40,14 @@ positive_variables = ['kappa_11', 'kappa_12', 'kappa_13', 'kappa_14',
 
 stupidconstant = 0.25
 
+def cs_hsigma(eta):
+  return pylab.around((1-eta/2)/(1-eta)**3 - 1, 2)
+
+start = 0.494
+stop = 0.545
 def scalefunction(hsig): # ~1 at 0.494, and ~0 at 0.545
-  return 1.0/(1 + (stupidconstant*v['h_sigma'])**4)
+  x = (hsig - cs_hsigma(stop))/(cs_hsigma(stop)-cs_hsigma(start))
+  return 0.5 - sympy.tanh(4*x)/2
 
 expressions = [
   ('g_HS', lambda:
@@ -394,6 +400,14 @@ outfile.write(r"""
 }
 \newcommand\alphaval{%(alpha)g}
 """ % locals())
+outfile.write(r"""
+
+\newcommand\hsigmastart{%g}
+\newcommand\hsigmastop{%g}
+\newcommand\etastart{%g}
+\newcommand\etastop{%g}
+
+""" % (cs_hsigma(start), cs_hsigma(stop), start, stop ))
 outfile.close()
 
 

### 2015-08-05 papers/square-well-fluid/paper.tex
add caption for sw coexistence plot (and update SConstruct which was broken) Tested on darcy. Took 0:52.08.
new SW homogeneous plots and add TODO for SW stuff Tested on darcy. Took 1:29.47.
add reference to nice review of modern monte carlo methods Tested on darcy. Took 0:51.98.

@@
 \end{center}
-\caption{Pressure.}
+\caption{Liquid-vapor coexistence curve for the square-well fluid as
+  predicted using thermodynamic perturbation theory.}
 \label{fig:coexistencec}
@@
 \label{fig:p-vs-eta}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=\columnwidth]{figs/free-energy-vs-eta}
+\end{center}
+\caption{Free energy.}
+\label{fig:free-energy-vs-eta}
 \end{figure}
@@
 
+\section{EXTRA}
+
+Here is a nice review article that introduces numerous modern Monte
+Carlo methods for fluids~\cite{dubbeldam2013inner}.
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2015-09-08 papers/fuzzy-fmt/figs/homogeneous.py
bring over changes from Eric's dissertation Tested on darcy. Took 1:31.41.

@@ -246,6 +246,7 @@ plot([],[], 'k--', label='Monte Carlo')
 plot([],[], 'k:', label='BH')
 legend(loc='best', frameon=False)
 
+plt.tight_layout();
 savefig('figs/p-vs-T.pdf', bbox_inches=0)
 
 show()

### 2016-04-18 papers/thesis-vischer/Thesis.tex

\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\setlength{\parindent}{0cm} % don't indent new paragraphs...
%\parskip \psl % ... place a space between paragraphs instead
\usepackage{graphicx, float}
\usepackage[font=small, labelfont=bf]{caption}
%\usepackage[english]{babel}
\usepackage{tikz}
%\usepackage{biblatex}
\usepackage{amsmath}
\usepackage[inline]{enumitem}
%\usepackage{float}
\usetikzlibrary{automata, positioning}
\newcommand{\ignore}[1]{}
\newcommand*{\diff}{\mathrm{d}}
\newcommand{\some}{{\color{red} something}}
\renewcommand{\vec}[1]{\mathbf{#1}}

\floatstyle{plaintop}
\newfloat{algorithm}{thb}{lop}
\floatname{algorithm}{Algorithm}
\newenvironment{alg}{\hrulefill\begin{enumerate}}{\end{enumerate}\hrulefill}

\usepackage[sort&compress,numbers]{natbib}
\bibliographystyle{apsrev4-1}
\usepackage[english]{babel}
\addto\captionsenglish{
    \renewcommand{\contentsname}{Table of Contents}
    \renewcommand{\refname}{References}
}
\usepackage{tocloft}
%\renewcommand{\cftsecleader}{\cftdotfil{\cftdotsep}}

% Commands go here
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\renewcommand{\title}{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}

\renewcommand{\author}{Brenden Vischer}

\renewcommand{\date}{02 June 2015}

\renewcommand{\titlepage}{
    {\centering
        \vspace*{4cm}

        \title

        \vspace{1.5cm}

        \author \\
        \text{Advisor: David Roundy}

        \vfill

        Department of Physics\\
        Oregon State University\\
        \today 
        \newpage}       
}

\begin{document}

\titlepage

\input{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\tableofcontents
\listoffigures

% Note: include new graphic about probability of move acceptance in Thermo. Int. (see austin thesis)

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%           INTRO                              %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
\subsection{Motivation}
Fluids have long garnered interest\cite{theorysimpleliquids}\cite{theoryofcrits}, particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes. A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor are indistinguishable and, more importantly, large fluctuations of fluid density occur. Most traditional fluid models assume the density across the cell to be nearly constant. As a result, these methods tend to poorly handle these fluctuations near criticality. 
Using the well-established model of the square-well fluid, we seek to determine the applicability of a particular generalized renormalization-group (``GRG'') formalism to find the absolute Helmholtz free energy of systems near criticality.\\

Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success, particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length. GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating a vast number of cells, varying the parameter each time, to find the most well-suited value for a particular fluid. We present a non-fitted approach that is GRG-minded from the ground up. We exploit the periodicity of Monte-Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths. We then carefully couple simulations of different ``doubling regimes'', each of which considers a different maximum wavelength of fluctuations, where the number density remains fixed but the volume and number have increased. Summing the difference in absolute Helmholtz free energy $\Delta F^{\text{abs}}$ between doubling regimes will then approximate the absolute free energy of the full cell, including near criticality. Finding absolute free energies is a somewhat challenging problem, but is certainly surmountable with more Monte Carlo. \\
We will begin by outlining the necessary models and assumptions/approximations for this work. We will then detail the algorithms used to compute the quantities with the the discussion on results following. 
\subsection{Hard Sphere}
The hard sphere model approximates monoatomic fluids as having no interactions other than collisions. These spheres exhibit no attraction or repulsion, and consequently each carries kinetic energy $T = 3/2 k_b T$. There is also a harder-to-calculate ``excess'' energy to be considered (which, in fact, contains most of the interesting behavior) which will be discussed further on. The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic. For simulation purposes, the system is required to have periodic boundary conditions - that is, when a sphere would classically collide with the wall of the cell it instead breaches the wall and then reappears on the other side of the cell with the same velocity. \\
\ignore{The hard sphere model is incredibly useful; as high temperatures are approached, energy associated with sphere location tends to asymptote for most models [are there any that this isn't true for?] and the kinetic term consequently dominates.}

\subsection{Square-Well Liquid}
The square-well liquid is a ubiquitous approximation that sees extensive use throughout computational fluid research. This simple fluid model consists of hard, immutable spheres that are subject to an intermolecular pair potential - that is, an attractive force between spheres. This model consists of several key parameters: sphere size $\sigma$, well width $\lambda$, and well depth $\epsilon$. These parameters can be fit to model virtually any monoatomic fluid (and some diatomic fluids, such as water). 
 As a hard-sphere model, the spheres are not allowed to overlap. Any arrangement of spheres - hereafter referred to as the ``state'' of the system - that involves either two or more spheres overlapping or a sphere outside the allowed area - ``cell'' - is defined to be invalid. Modifying the radius of the spheres changes the allowed locations in the cell, which in turn affects the partition function of the system. The ``square'' refers to the piecewise intermolecular pair potential $\Phi_{12}$, shown mathematically in Eq. \ref{phi12} and graphically in Figure ~\ref{sw_phi}.

\begin{equation} 
\Phi_{12} = \begin{cases}\infty & |\vec{r}|\leq \sigma\\ -\epsilon & \sigma \leq |\vec{r}| \leq \sigma\lambda\\ 0 & |\vec{r}| > \sigma\lambda \end{cases}
\label{phi12}
\end{equation}

The energy of the fluid is then always an integer multiple of the well depth $\epsilon$. This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation. We then define dimensionless temperatures $T/\epsilon$, energies $E/\epsilon$, and distances $r/\sigma$. Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ (as well as any other physical constants) to one.

\begin{figure}
    \centering
    \includegraphics[width=.75\textwidth]{SWF-E.png}
    \caption{Demonstration of Square-Well approximation.}
    \label{sw_phi}
\end{figure}
\subsection{Critical Point} 
The critical point is the temperature and pressure (or any other thermodynamic pairing) on a phase diagram (shown in Figure ~\ref{phase}) where the line between liquid and vapor terminates - liquid and vapor become indistinguishable and are assimilated into the ``fluid'' phase. Many interesting effects occur here, including the well-researched phenomena of critical opalescence. The correlation length, naively interpreted here as the maximum existing wavelength of the density fluctuations within the fluid, diverges to infinity near criticality. As mentioned previously, the diverging of the correlation length invalidates SAFT and many of its variants which assume constant density. 
\begin{figure}
    \centering
    \includegraphics[width=.75\textwidth]{phase.png}
    \caption{Phase transitions of water with critical point and triple point identified.}
    \label{phase}
\end{figure}
\\
The theory of this project is largely predicated upon the behavior of critical systems, namely the phenomena of self-similarity. Non-critical systems exhibit different characteristics when viewed at different length scales. Ever-present small-scale fluctuations in density are relevant, if not dominant, when viewing the system at very small scales, but typically become less important at larger scales and nearly irrelevant near the thermodynamic limit. Near the critical point all wavelengths of fluctuation are present, and so the system looks the same regardless of length scale chosen. This self-similarity allows for the decomposition of the large critical cell to smaller constituent cells. The details of the procedure are more carefully discussed later on.

\subsection{Free energy}
The free energy we are interested in calculating is the ``absolute'' Helmholtz free energy $F^{\text{abs}}$. The Helmholtz free energy is a Legendre transform of the internal energy $U$:
\begin{align}
    F &= U - TS
    \label{Fdef}
\end{align}
and can be calculated given the ever-pervasive partition function $Z$:
\begin{align}
    F &= -T \ln Z
    \label{FdefSM}
\end{align}
With $F$ calculated, simple partial derivatives yield all other thermodynamic quantities. Though the Gibbs free energy yields the same information, the Helmholtz free energy is particularly useful because of the relationship with entropy:
\ignore{\begin{align}
    S &= \pdd{F}{T}{V}
\end{align}}
\\EQ HERE\\

Equations ~\ref{Fdef} and ~\ref{FdefSM} together bridge classical thermodynamics and statistical mechanics. Due to the nature of this project, we will primarily be concerned with the statistical mechanics definition (that is, Eq. ~\ref{FdefSM}) of the free energy. We are, then, ultimately concerned with determining the partition function for any system we are concerned with. The canonical ensemble partition function can be compactly represented by a summation:
\begin{align}
    Z &= \sum^{\text{energies}}~e^{-\beta E_i}
\end{align}
where $\beta =  T^{-1}$. Of course, it is typically impractical to sum over the individual energies of the system, especially if there are expected degeneracies. Instead we sum over states:
\begin{align}
    Z &= \sum^{\text{states}}~D(E)~e^{-\beta E_i}
\end{align}
where we have now introduced the density of states $D(E)$ which is a simple function that returns the number of states which have a given energy $E$. This introduction is now the focus of calculation - if we can accurately determine the density of states of the system and have knowledge of the available energies, we can calculate the Helmholtz free energy and, from there, any other  \\   
Since $F$ contains a natural log term, we can separate and characterize the individual contributions of kinetic energy (``ideal'') and energy associated with the configuration of the system (``excess'').  
A derivation of the two energy contributions for a monoatomic fluid follows. \\

{\color{red} [This should probably go in an appendix. Further: replace $Z_C$ with $Z_{exc}$; fix the to the N on the $Z_{exc}$ term; include the $N!$ into the ideal term; fix the $V(r_1, r_2, ...)$]}

We begin with the canonical form of the canonical ensemble partition function $Z$:
\begin{align}
    Z &= \sum_i e^{-\beta E_i}
\end{align}
where we sum over individual energies of the system $E_i$. However, this method doesn't account for degeneracies in the system, so instead we sum over the states $s$ (to avoid including the density of states, which is difficult to find for most systems):
\begin{align}
    Z &= \sum_s e^{-\beta E_s}
\end{align}

Further, the energy of each state is simply the sum of the excess and kinetic energies $V$ and $K$, so

\begin{align}
    E_s &= V_s(\mathbf{r}) + K_s(\mathbf{p})
\end{align}

So now we have

\begin{align}
    Z &= \sum_S e^{-\beta (V_s(\mathbf{r}) + K_s(\mathbf{p}))}\\
    &= \sum_S e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}
\end{align}

Each state is specified by a particular point in phase space (parameterized by $\mathbf{r}$ and $\mathbf{p}$). Even though this phase space is discrete, we can approximate the space as continuous by recognizing that the {\color{red}spacing between individual states is small compared to the total range of the states}, so that the summation becomes an integral:

\begin{align}
    Z &= \int_s~ d\mathbf{r}~ d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}
    %&= \sum_s e^{\beta V(\mathbf{r})}e^{-\beta T(\mathbf{p})}
\end{align} 
However, the above is only partially correct, as it is only for one particle. We can account for multiple particles by requiring that each particle has the same partition function. This is equivalent to integrating $N$ times for $N$ particles:

\begin{align}
    Z &= \int_{s_1} \int_{s_2} ... \int_{s_N} d\mathbf{r_1}~ d\mathbf{p_1} ~d\mathbf{r_2} ~d\mathbf{p_2} ~... ~d\mathbf{r_N}~ d\mathbf{p_N} ~e^{-\beta(V(\mathbf{r_1}) + V(\mathbf{r_2}) ... V(\mathbf{r_N}))}~e^{-\beta(K(\mathbf{p_1}) + K(\mathbf{p_2}) ... K(\mathbf{p_N}))}\\
    &= \left[\int_s d\mathbf{r} ~ d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}\right]^{N}
\end{align}
Additionally, for a proper quantum mechanical treatment we must assume that the particles are indistinguishable which means the above expression over-counts by a factor of $N!$. Our expression for the partition function is then

\begin{align}
    Z &= \frac{1}{N!}\left[\int_s d\mathbf{r} d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}\right]^{N}
\end{align}

Furthermore, we can separate the integrals completely because the exponentials are each attached to only one differential element:

\begin{align}
    Z &= \frac{1}{N!}\left[\int_s d\mathbf{r} ~e^{-\beta V(\mathbf{r})}\right]^{N}\left[\int_s d\mathbf{p} ~e^{-\beta K(\mathbf{p})}\right]^{N}
    \label{Z_full} 
\end{align}

We then define the two terms obvious in Eq. \ref{Z_full}, the excess term dependent on the potential energy $V(\mathbf{r})$ and the ideal term dependent on the kinetic energy $K(\mathbf{p})$:

\begin{align}
    Z &= (Z_{exc})(Z_{id}) \,,
\end{align}
where $Z_{id}$ contains the $N!$ term by convention. The excess term $Z_{exc}$ is system dependent because $V(\mathbf{r})$ is buried inside the integral, but the ideal term is the same for any system where the total energy separates in the preceding fashion. We judiciously decide to leave the excess term alone and work specifically with the ideal term. We then have

\begin{align}
    Z_{id} &=  \frac{1}{N!}\left[\int_s d\mathbf{p} ~e^{-\beta T(\mathbf{p})}\right]^{N}
\end{align}

We begin by imagining a cell of length $L$ and volume $L^3$ in which lives a particle. Then, we quantize the allowed wavelengths of the particle within the cell by defining the wavevector $\mathbf{k}$ as
\begin{align}
    \mathbf{k} &= \frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}})
\end{align}
It should be noted that this definition of $k$ introduces three-fold degeneracy.\\
Remembering that kinetic energy can be expressed as $T = \frac{\hbar^2 k^2}{2m}$, and using our definition of $\mathbf{k}$:
\begin{align}
    Z_{id} &= \frac{1}{N!}\left[\int_s d\mathbf{p} ~e^{-\beta \frac{\hbar^2 k^2}{2m}}\right]^{N}
    \label{Zt_def}
\end{align}
This unfortunately introduces a mass quantity into the partition function which adds a parameter that needs fitting. [more] 
If we then allow the cell to become very large, then the we can easily see in Eq. \ref{Zt_def} that the spacing between occupants in phase space becomes small, so we can integrate instead $\diff ^3 n$ over all possible values of $n$:
\begin{align}
    Z_{id} &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{-\frac{\beta \hbar^2}{2m} \left(\frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}}\right)^2}\right]^{N} = \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{ \frac{-2\beta \hbar^2\pi^2}{mL^2}(n_x^2 + n_y^2 + n_z^2)}\right]^{N}
\end{align}
Evaluating this integral yields
\begin{align}
    Z_{id} &= \left(\frac{V^N}{N!\Lambda^{3N}}\right) \,, \\
    \Lambda &= \frac{\hbar\sqrt{2\pi\beta}}{\sqrt{m}}
\end{align}
where we have introduced the thermal deBroglie wavelength $\Lambda$, loosely interpreted as [physical thing]. The full partition function for any monoatomic system is then
\begin{align}
    Z &= Z_{exc} \left(\frac{V}{\Lambda^{3}}\right)^{N} \,.
\end{align}
For liquids, the $Z_{exc}$ term tends to be difficult to evaluate; the hard sphere liquid, the simplest model for a fluid, does not have an analytic representation for the excess free energy. It follows that a considerable amount of work is required to determine the excess free energy for a given system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                            METHODS                                  %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Methods}
This work requires the intersection of three methods to calculate the total absolute free energy $F^{abs}_{tot}$ of the bulk cell. A preliminary value of the free energy of the bulk can be obtained from Monte-Carlo simulations. This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the excess free energy of a system. [more] 
\subsection{Monte-Carlo Methods}
Monte-Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations. Traditional random sampling requires that the current step, or ``move'', is irrespective of previous steps. True random movement is an ideal simulation characteristic for the canonical ``drunken rambler'' scenario, in which a intoxicated man attempts to make his way home by taking random steps. Since the drunken man chooses his next move arbitrarily, his movement is considered ``random''; it follows that true random sampling is desirable for this physical system. Statistical mechanics, however, does not exhibit this independence. Each time the man makes a move there is an equal probability that he will move in any particular direction. Rather than all movements having equal probability, statistical mechanics stipulates that the probability of a particular move occurring depends explicitly upon the energy difference of the current and resulting configurations. Instead of traditional Monte Carlo methods, Markov chain (specifically the Metropolis-Hastings algorithm) Monte Carlo is the natural decision.
\subsubsection{Markov Chain}
Markov chain methods eschew the true randomness of regular Monte Carlo by taking the current configuration into account before moving. 
This method can lead to unsavory correlations in samples, but this danger is far outweighed by the usefulness of modeling systems where direct sampling is nigh impossible.
\begin{figure}
    \begin{tikzpicture}
    \end{tikzpicture}
\end{figure}

\begin{figure}
    \caption{Monte-Carlo contrasted with Markov-Chain Monte-Carlo (This might be difficult...)}
\end{figure}
\ignore{\subsubsection{Histogram Methods}
Histograms are graphical representation of ``counting'' data. Monte Carlo histogram methods simply track the number of times the system has visited a particular energy. If the transition matrix 
\begin{figure}
    \caption{Sample histogram.}
\end{figure}}
\subsection{Broad Histogram Monte-Carlo}
[Things about broad histogram methods]
Broad histograms are focused entirely on determining the density of states $D(E)$ for a given system. [more]

{\color{red} succeeding text may be worthless.}

The method of the transition matrix has existed for over a decade and has seen several varying implementations. The canonical form involves knowing the exact number of states, and more importantly the relative transition probabilities, available to the system {\it a priori}, which is simply not possible for cells as large as we are interested in~\cite{perlinthesis}. Other methods spends the majority of run time initializing - that is, constructing the weighting arrays, which then are used to form the transition matrix [possibly incorrect]. Once the transition matrix is complete, we need only wait to gather sufficient statistics for the energies we are interested in. In the pursuit of time-saving, {\color{red} we have implemented a particular type of transition matrix construction that spends this initialization time both constructing the transition matrix on the fly and gathering statistics. Further, our method outputs the density of states as its sole output, decreasing the file storage burden of the simulations.}
\begin{figure}
    \caption{Sample transition matrix/demonstration of transition matrix.}
\end{figure}
\subsubsection{Algorithm}

\subsection{Renormalization Group Theory}
Renormalization Group attempts to describe how a particular system's properties are altered when the system is viewed from varying length scales. Near the critical point, the fluid becomes self similar, and so fluctuations of all wavelengths are present. We can then compare quantities of different, yet related, systems as they all exhibit the same properties.\\
We begin with a periodic cubic cell of length $L_0$ and volume $L_0^3$. 
This cell has periodic boundary conditions and consequently allows for fluctuations up to wavelength $2L_0$. These fluctuations contribute some absolute free energy $F_0$ to the absolute free energy of the full cell. We then double the length of the cell to allow for fluctuations of up to wavelength $4L_0$. We continue this process until all relevant wavelengths have been included. Adding up the change in absolute free energy calculated in this manner then yields the total free energy of the critical system, which can then be compared to accepted simulation methods. Algorithm~\ref{alg_renorm} details this process.

\begin{algorithm}[tb]
    \caption{Renormalization-esque calculation of total free energy}
    \label{alg_renorm}
    \hrulefill
    \begin{enumerate}
        \item Simulate smallest cell of length $L_0$.
        \item Calculate absolute free energy from simulation output.
        \item Simulate new cell with same density but octupled volume of previous cell.
        \item Calculate absolute free energy from simulation output.
        \item Calculate change in absolute free energy between the current regime and the last, add to running $\Delta F^{\text{abs}}_{\text{tot}}$
        \item Repeat 3-4 for each desired doubling regime.
    \end{enumerate}
    \hrulefill
\end{algorithm}

Eq. ~\ref{GRGT-F1} details the mathematics of this coupling procedure.

\begin{align}
    \label{GRGT-F1}
    F^{\text{abs}}_{\text{tot}} = \sum_{i=1}^{k}\Delta F^{\text{abs}} &= \lim_{k\to\infty}\sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}\\
    &\approx \sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}
    \label{GRGT-F2}
\end{align}

Strictly speaking, a formal limit is required for equality in Eq.~\ref{GRGT-F1}, rather than only including wavelengths up to $2^k\lambda_0$. Though theoretically all wavelengths are present in the critical system we posit that it is only necessary to include the first several. Given approximation, we choose instead to use Eq.~\ref{GRGT-F2}. Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely TMMC and Thermodynamic Integration. A graphical representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\

%%%%%%% TIKZZZZZ%%%%%%%

\begin{figure}
\centering
    \resizebox{.75\textwidth}{!}{
     \begin{tikzpicture}
            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (0,0) rectangle (1,1);
            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (2,0) rectangle (4,2);
            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (5,0) rectangle (9,4);
            %\draw[black, ultra thick, fill=yellow, fill opacity=.2] (10,0) rectangle (14,4);
            \draw[<-] (5.15,2) to (5.65,2);
            \draw[->] (7,3.35) to (7,3.85);
            \draw[<-] (7,.15) to (7,.65);
            \draw[->] (8.35,2) to (8.85,2);
            \node[above] at (.5,1) {\tiny $g_0$: $L=L_0$, $N=1$};
            \node[above] at (3,2) {\tiny $g_1$: $L=2L_0$, $N=4$};
            %\node[above] at (7,4) {\tiny $g_2$: $L=4L_0$, $N=8$};
            \node[above] at (7,4) {\tiny $g_n$: $L=2^nL_0$, $N=(2^n)(2^n)=4^n$};

            % Balls!
            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];
            % Second box
            \draw[ultra thick, fill=red] (3.25,.5) circle [radius=.25];
            \draw[ultra thick, fill=red] (3.5,1) circle [radius=.25];
            \draw[ultra thick, fill=red] (2.375,1.5) circle [radius=.25];
            \draw[ultra thick, fill=red] (2.5,.75) circle [radius=.25];
            % third
            \draw[ultra thick, fill=red] (5.5,1.5) circle [radius=.25];
            \draw[ultra thick, fill=red] (5.375,.375) circle [radius=.25];
            \draw[ultra thick, fill=red] (6,3.5) circle [radius=.25];
            \draw[ultra thick, fill=red] (7.5,3.25) circle [radius=.25];
            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];    
            \draw[ultra thick, fill=red] (7,2) circle [radius=.25];        
            \draw[ultra thick, fill=red] (5.75,3) circle [radius=.25];        
            \draw[ultra thick, fill=red] (6.5,2.75) circle [radius=.25];        
            \draw[ultra thick, fill=red] (8, 3) circle [radius=.25];        
            \draw[ultra thick, fill=red] (7.5,1.25) circle [radius=.25];        
            \draw[ultra thick, fill=red] (7.75,1.75) circle [radius=.25];        
            \draw[ultra thick, fill=red] (8.5,1.5) circle [radius=.25];
            \draw[ultra thick, fill=red] (6.5,1.25) circle [radius=.25];                
     \end{tikzpicture}}
\caption{2-D Demonstration of MC-GRGT doubling method.}
\label{GRTG-demo} 
\end{figure}

%%%%%% TIKZZZZ %%%%%%

\subsection{Thermodynamic Integration}
\subsubsection{General Process}
Thermodynamic Integration is a unique way to determine the ``excess'' free energy of the system. Using hard sphere Monte-Carlo, we transition from a filling fraction of zero (note that this statement is identical to assuming that the cell begins at infinite temperature, since the total energy does not depend on where the spheres are placed) to a near-close-packed system with filling fraction $\eta_max$, calculating the change in free energy at each step. When we have reached the desired final filling fraction we have found the free energy that was necessary to create that configuration of particles - this is exactly the absolute free energy of the system.\\
\subsubsection{Algorithm}
The hard sphere fluid is simple to work with since its partition function simply counts the number of microstates in the system\cite{valeskethesis}. Since we are calculating the change in free energy as we transition to higher and higher filling fractions, we need only be concerned with finding the ratio of partition functions between each step. Since the hard sphere partition function is a ``counting'' mechanism, we can determine this ratio quite easily using Monte Carlo simulations. Given a starting state (defined by the arrangement of particles and filling fraction) we shrink the volume and intermolecular distances by a somewhat arbitrary scaling factor and then see if the resulting configuration of particles is valid. If the configuration is valid, we accept this is a working move and then return to the original configuration and test another scaling factor. If the resulting configuration is invalid, we count the move as failed and then repeat. The ``golden'' acceptance ratio is fifty percent - when we have found a scaling factor that yields a ratio near the desired value\dots\\

As with the Monte Carlo these simulations are written in $C++$ for efficiency.

\begin{algorithm}[!b]
\caption{Thermodynamic integration.}
\label{alg:ti}
\begin{alg}
\item Construct cell with desired number of spheres with filling fraction of zero.
\item Move spheres random fractions of a predefined translation distance. \label{alg-ti-move}
\item Repeat step \ref{alg-ti-move} for [desired number of iterations before small cell check].
\item After [desired number of iterations], check the ``small cell'' by shrinking the volume and checking whether or not the configuration is still valid.
\item If configuration is still valid, valid checks ++.
\item If configuration is invalid, failed checks ++. \label{alg-ti-move-end}
\item Repeat steps \ref{alg-ti-move} - \ref{alg-ti-move-end} until desired iterations reached.

\end{alg}      
\end{algorithm}

\ignore{Should I detail the algorithm itself and string simulations together or consider the entire process $\eta=0\rightarrow\eta_max$ to be the algorithm?}

\ignore{is the distinction between infinite case and regular necessary? we always use infinite...}

\begin{figure} \centering
\resizebox{.75\textwidth}{!}{
\begin{tikzpicture}[scale=.8,  auto]
\draw[blue, ultra thick] (10,3) rectangle (14,7);
\draw[blue, ultra thick, dashed] (0,1) rectangle (8,9);
\draw[blue, ultra thick] (16,4.375) rectangle (17.25,5.625);
\draw[->] (4,8) to (4, 8.85);
\draw[->] (7,5) to (7.85, 5);
\draw[->] (1,5) to (.15, 5);
\draw[->] (4,2) to (4, 1.15);
\draw[ultra thick, ->] (8.5,5) to (9.5,5);
\node [above] at (9,5.25) {\small $k$ steps};
\draw[ultra thick, ->] (14.5,5) to (15.5,5);
\node[align=center] [above] at (15, 5.25) {\small $n-k$ \\[-1em] steps};
\draw[ultra thick, fill=red] (3.5,4.25) circle [radius=.25];
\draw[ultra thick, fill=red] (6,3.75) circle [radius=.25];
\draw[ultra thick, fill=red] (5.5,6) circle [radius=.25];
\draw[ultra thick, fill=red] (10.5,4.5) circle [radius=.25];
\draw[ultra thick, fill=red] (13,3.5) circle [radius=.25];
\draw[ultra thick, fill=red] (11,6) circle [radius=.25];
\draw[ultra thick, fill=red] (16.3,5.3) circle [radius=.25];
\draw[ultra thick, fill=red] (17,4.75) circle [radius=.25];
\draw[ultra thick, fill=red] (16.25,4.65) circle [radius=.25];
\node [above] at
 (4,9) {$\eta_0 \approx 0$};
\node [above] at (12,7) {$\eta_k$};
\node [above] at (16.675,5.625) {$\eta_f$}; 
\end{tikzpicture}}
\caption{Demonstration of filling fraction increase during thermodynamic integration from $\eta=0$ to $\eta=\eta_f=\eta_{max}$.}
\end{figure}

\subsubsection{Carnahan-Starling Equation of State}
In order to determine the filling fraction step size during thermodynamic integration we must find changes in filling fraction that yield a suitably-sized change in the system's absolute free energy. To do this we use the Carnahan-Starling equation of state, shown in Eq.~\ref{C-S-eq}.
\begin{align}
    F(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2}   
    \label{C-S-eq}  
\end{align} 
where $\eta$ is the filling fraction. Combining Eq.~\ref{C-S-eq} with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy. We determined that steps yielding an approximate change in specific absolute free energy of $\Delta F^{abs}/NT = \ln(2)$ through trial and error - smaller step sizes unnecessarily increased simulation time and larger step sizes tended to shrink too quickly causing the simulation to miss relevant states in the system.

\ignore{NB: is there a more physical reason we chose this filling fraction increase?}

\subsection{Data Storage}
All simulation output is saved in text files for plotting and visualization ease. The question of organization becomes important with this amount of files being generated. Each different system simulation (length $L = \sqrt2 \sigma, 5,10$) has data regarding each doubling regime ($i=0,1,2,3?$) and each doubling regime contains data for each possible sphere count ($N=1,2,3\dots N_{max}$, where $N_{max}$ is typically limited by maximal sphere packing of $N_{max}/L^3\approx.741$). The file structure mirrors the above symmetry: 
\begin{align*}
                &\rightarrow \text{i=0} &\rightarrow \text{N=0}\\
                & &\rightarrow \text{N=1}\\
                && \vdots\\
\text{System} & \rightarrow \text{i=1} &\rightarrow \text{N=0}\\
                && \rightarrow \text{N=1}\\
                &&\vdots\\
                &\rightarrow \text{i=2} &\rightarrow \text{N=0}\\
                && \rightarrow \text{N=1}\\
                &&\vdots\\
\end{align*}
Each sphere count subfolder contains a single text file from TMMC output and another subfolder where all of the Thermodynamic Integration output is stored. Each Thermodynamic Integration simulation outputs a single text file upon completion of each filling fraction step. On average, a single TI calculation requires approximately $175$ steps. \\
This structure was chosen both for elegant parallelism and for plotting accessibility.

\subsection{Calculating the Absolute Change in Free Energy}
It is straightforward the compute a Helmholtz free energy through the Monte-Carlo simulations; the difficulty in this method lies in computing the correct free energy. TMMC, as mentioned previously, yields a free energy accurate only up to a unknown constant. Since we need to add up changes in free energy across multiple simulations, we need each calculated free energy to be ``absolute'' - that is, we must eliminate this unknown constant. \\

At infinite temperature, the square well fluid behaves exactly as the hard sphere. The correct inclusion, then, is to enforce that the free energy calculated via TMMC at $T=\infty$ is equal to the free energy calculated through thermodynamic integration. We do this by exploiting the infinite temperature behavior of the canonical ensemble:
\begin{align}
    Z(T=\infty) &= \left.\sum_{i=0}^{\text{\tiny Energies}} D(E) e^{-\beta E}\right|_{\beta = 0}\\
    &= \sum_{i=0}^{\text{\tiny Energies}} D(E)\\
    &= \#\text{~states}
\end{align}

\ignore{NB: cough cough the HS energy probs wrong}
The partition function now calculates the total number of states in the system which we expected due to the equipartition theorem. The hard sphere free energy [more]

\section{Results}
An important remaining question is whether there is a preferred choice of minimum cell length. Physically we would [might?] expect that there is some dominant minimum fluctuation in the bulk cell - this would then lend a natural decision for the minimum cell length $\lambda_0$. Unfortunately we do not know this {\it a priori}, so we are reduced to guessing. Using the coupling procedure outlined previously, we calculated changes in free energy up to the $n^{\text{th}}$ doubling regime for several different cell lengths. We are primarily concerned with the ``scrunched'' case where the cell length is $\sqrt2\sigma$.

Figures ~\ref{F-T} and ~\ref{F-eta} show preliminary free energy calculations.
\begin{figure}[H]
\centering
    \includegraphics[width=.75\textwidth]{ww130-L03-i1_F-T.pdf}
    \caption{Preliminary results for absolute specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$. The vertical line at $T/\epsilon = 0.3$ is the lowest temperature at which our statistics are trustworthy.}
    \label{F-T}
\end{figure}
\begin{figure}
\centering
    \includegraphics[width=.75\textwidth]{ww130-L03-F_i1_fixed_i.pdf}
    \caption{Preliminary results for free energy as a function of density.}
    \label{F-eta}
\end{figure}

\subsection{Primitive Cell}
The ``scrunched'' system allows exactly four spheres to be packed -- at maximal filling fraction $\eta_{max} = \approx .741$ -- in the original non-doubled cell. This case corresponds to the unit cell a face-centered cubic lattice, the simplest sphere packing possible. We chose to focus on the scrunched case for two reasons. Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length

\subsubsection{Seed Disparity}
Because of the random/psuedorandom nature of Monte-Carlo sampling, we tested over many different initial sphere placements (``seeds'') to see if there was consistent agreement. 
\subsection{Other Cells?}
In addition to the ``scrunched'' primitive cell, we tested [hopefully one or two] other minimum correlation lengths corresponding to other natural parameters in the system.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                       Discussion                            %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Discussion}
If the decomposition procedure outlined previously proves successful, the absolute free energy yielded by this renormalization method ought to agree with the free energy yielded from other accepted theories. [more]
\subsection{Change in Free Energy Over Multiple Regimes}
[more]
\subsection{Comparison With A Modified Statistical Associating Fluid Theory}
[more]

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                             CONCLUSION                         %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Conclusion}

\subsection{Future Research}
There are many other models that GRGT might be applicable to, such as the soft-sphere liquid or soft polyhedra.
%\bibliographystyle{plain}
\bibliography{Thesis}
\end{document}

### 2016-04-18 papers/thesis-vischer/abstract.tex

\thispagestyle{plain}
\begin{center}
    \Large
    \textbf{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}

    \vspace{0.8cm}
    \large
    A Computational Project

    \vspace{0.8cm}
    \textbf{Brenden Vischer}

    \vspace{1.2cm}
    \textbf{Abstract}
\end{center}

You take a box, stuff it with spheres, move them around, make it bigger, do it again. Win grant.
\clearpage

### 2016-04-18 papers/thesis-vischer/kinetic_partition.tex

\documentclass[10pt]{article}

%%% standard header
\usepackage[margin=1in]{geometry} % one inch margins
\usepackage{color}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{fancyhdr} % easier header and footer management
\pagestyle{fancyplain} % page formatting style
\usepackage{hyperref} % for linking references
\newcommand{\psl}{6pt} % store spacing as command for consistency
\setlength{\parindent}{0cm} % don't indent new paragraphs...
\parskip \psl % ... place a space between paragraphs instead
\usepackage[inline]{enumitem} % include for \setlist{}, use below
%\setlist{nolistsep} % more compact spacing between environments
\setlist[itemize]{leftmargin=*} % nice margins for itemize ...
\setlist[enumerate]{leftmargin=*} % ... and enumerate environments
\frenchspacing % add a single space after a period
\usepackage{lastpage} % for referencing last page
\cfoot{\thepage~of \pageref{LastPage}} % "x of y" page labeling

\renewcommand{\headrulewidth}{0.5pt} % horizontal line in header
\lhead{Brenden Vischer}
\chead{March 9th, 2015}
\rhead{Kinetic Energy Partition Function}

\begin{document}
\begin{center}
  \large \bf Derivation of Kinetic Contribution to Partition Function
\end{center}

\section*{Partition Function}

### 2016-04-20 papers/thesis-vischer/Thesis.tex
Thesis work, wrote a loose abstract Tests skipped!

@@
 As a result, these methods tend to poorly handle these fluctuations near criticality.
-Using the well-established model of the square-well fluid, we seek to determine the applicability of a particular generalized renormalization-group (``GRG'') formalism to find the absolute Helmholtz free energy of systems near criticality.\\
+Using the well-established model of the square-well fluid, we seek to determine the applicability of a particular generalized renormalization-group formalism (``GRG'') to find the absolute Helmholtz free energy of systems near criticality.\\
 
@@
     \centering
-    \includegraphics[width=.75\textwidth]{SWF-E.png}
+    %\includegraphics[width=.75\textwidth]{SWF-E.png}
     \caption{Demonstration of Square-Well approximation.}
@@
     \centering
-    \includegraphics[width=.75\textwidth]{phase.png}
+    %\includegraphics[width=.75\textwidth]{phase.png}
     \caption{Phase transitions of water with critical point and triple point identified.}
@@
 \centering
-    \includegraphics[width=.75\textwidth]{ww130-L03-i1_F-T.pdf}
+    %\includegraphics[width=.75\textwidth]{ww130-L03-i1_F-T.pdf}
     \caption{Preliminary results for absolute specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$.
@@
 \centering
-    \includegraphics[width=.75\textwidth]{ww130-L03-F_i1_fixed_i.pdf}
+    %\includegraphics[width=.75\textwidth]{ww130-L03-F_i1_fixed_i.pdf}
     \caption{Preliminary results for free energy as a function of density.}

### 2016-04-20 papers/thesis-vischer/abstract.tex
Thesis work, wrote a loose abstract Tests skipped!

@@
     \textbf{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}
+    \\or\\
+    \textbf{Decomposition of the Critical Square-Well Liquid Into Consitutent Cells}    \\or\\
+    \textbf{Free Energy Decomposition of the Critical Square-Well Liquid Using a Renormalization Group Method}
 
@@
 
-You take a box, stuff it with spheres, move them around, make it bigger, do it again. Win grant.
+The free energy of a bulk liquid is a known quantity [cite], near and far from the critical region [cite.
+true?].
+Computing the free energy of a liquid requires lengthy, difficult computations that [something].
+These computations generally involve renormalization methods which [do things].
+Analytic models using renormalization methods for fluids have been [time scale] developed [cite], but computational models [something].
+[something about including all lengths, not knowning what each contributes].
+We present a method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Monte-Carlo simultations.
+We further detail a method by which to compute the ``absolute'' free energy, with both the ideal gas free energy and the excess free energy, for the square-well liquid.
+[more] We find that the square-well liquid near the critical region is well-characterized by a base cell length of $\sqrt2\sigma$, the side length of an FCC lattice.
 \clearpage

### 2016-04-25 papers/thesis-vischer/Thesis.tex
Thesis work - cleaned up free energy derivation, fixed some language throughout. More to come! Tests skipped!

@@
 The square-well liquid is a ubiquitous approximation that sees extensive use throughout computational fluid research.
-This simple fluid model consists of hard, immutable spheres that are subject to an intermolecular pair potential - that is, an attractive force between spheres.
+This simple fluid model consists of hard spheres that are subject to an intermolecular pair potential - that is, an attractive force between spheres.
 This model consists of several key parameters: sphere size $\sigma$, well width $\lambda$, and well depth $\epsilon$.
-These parameters can be fit to model virtually any monoatomic fluid (and some diatomic fluids, such as water).
+\\
  As a hard-sphere model, the spheres are not allowed to overlap.
@@
 Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ (as well as any other physical constants) to one.
+This then allows us to resonably compare energies to temperatures since they have the same units.
 
@@
 Though the Gibbs free energy yields the same information, the Helmholtz free energy is particularly useful because of the relationship with entropy:
-\ignore{\begin{align}
-    S &= \pdd{F}{T}{V}
-\end{align}}
-\\EQ HERE\\
+\begin{align}
+    S &= \left(\frac{\partial F}{\partial T}\right)_{V}
+\end{align}
+Since we are defining cells of fixed volume this partial derivative will be simple to evaulate for any temperature $T$.\\
 
@@
 
-{\color{red} [This should probably go in an appendix. Further: replace $Z_C$ with $Z_{exc}$; fix the to the N on the $Z_{exc}$ term; include the $N!$ into the ideal term; fix the $V(r_1, r_2, ...)$]}
+{\color{red} [This should probably go in an appendix.
+Further: replace $Z_C$ with $Z_{exc}$; fix the to the N on the $Z_{exc}$ term; include the $N!$ into the ideal term; fix the $V(r_1, r_2, ...)$||| All done!
+NOW: fix the dimensions on the $Z_{id}$ term!]}
 
@@
 \end{align}
-
+noting that $V(\mathbf{r})$ is poorly defined for a single particle, but we digress. [FIXME]. 
 So now we have
@@
 \begin{align}
-    Z &= \int_s~ d\mathbf{r}~ d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}
+    Z &= \int_s~ \diff\mathbf{r}~ \diff\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}
     %&= \sum_s e^{\beta V(\mathbf{r})}e^{-\beta T(\mathbf{p})}
@@
 However, the above is only partially correct, as it is only for one particle.
-We can account for multiple particles by requiring that each particle has the same partition function.
-This is equivalent to integrating $N$ times for $N$ particles:
+We can account for multiple particles by requiring that each particle has the same partition function contribution for kinetic energy; however, the potential energy does not seperate as nicely since it depends on the position of all particles.
+For the kinetic energy, we are simply integrating $N$ times for $N$ particles:
 
 \begin{align}
-    Z &= \int_{s_1} \int_{s_2} ...
-\int_{s_N} d\mathbf{r_1}~ d\mathbf{p_1} ~d\mathbf{r_2} ~d\mathbf{p_2} ~...
-~d\mathbf{r_N}~ d\mathbf{p_N} ~e^{-\beta(V(\mathbf{r_1}) + V(\mathbf{r_2}) ...
-V(\mathbf{r_N}))}~e^{-\beta(K(\mathbf{p_1}) + K(\mathbf{p_2}) ...
+    Z &= \int_{s_1} \int_{s_2} \dots \int_{s_N} \diff\mathbf{r_1}~ \diff\mathbf{p_1} ~\diff\mathbf{r_2} ~\diff\mathbf{p_2} ~\dots ~\diff\mathbf{r_N}~ \diff\mathbf{p_N} ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})}~e^{-\beta(K(\mathbf{p_1}) + K(\mathbf{p_2}) ...
 K(\mathbf{p_N}))}\\
-    &= \left[\int_s d\mathbf{r} ~ d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}\right]^{N}
+    &= \int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
 \end{align}
+where we have separated the integrals by recognizing that the integrands are independant.\\
 Additionally, for a proper quantum mechanical treatment we must assume that the particles are indistinguishable which means the above expression over-counts by a factor of $N!$.
@@
 \begin{align}
-    Z &= \frac{1}{N!}\left[\int_s d\mathbf{r} d\mathbf{p} ~e^{-\beta V(\mathbf{r})}~e^{-\beta K(\mathbf{p})}\right]^{N}
-\end{align}
-
-Furthermore, we can separate the integrals completely because the exponentials are each attached to only one differential element:
-
-\begin{align}
-    Z &= \frac{1}{N!}\left[\int_s d\mathbf{r} ~e^{-\beta V(\mathbf{r})}\right]^{N}\left[\int_s d\mathbf{p} ~e^{-\beta K(\mathbf{p})}\right]^{N}
+    Z &= \frac{1}{N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
     \label{Z_full}
@@
 \begin{align}
-    Z_{id} &=  \frac{1}{N!}\left[\int_s d\mathbf{p} ~e^{-\beta T(\mathbf{p})}\right]^{N}
+    Z_{id} &=  \frac{1}{N!}\left[\int_s \diff\mathbf{p} ~e^{-\beta K(\mathbf{p})}\right]^{N}
 \end{align}
@@
 It should be noted that this definition of $k$ introduces three-fold degeneracy.\\
-Remembering that kinetic energy can be expressed as $T = \frac{\hbar^2 k^2}{2m}$, and using our definition of $\mathbf{k}$:
+Remembering that kinetic energy can be expressed as $K = \frac{\hbar^2 k^2}{2m}$, and using our definition of $\mathbf{k}$:
 \begin{align}
-    Z_{id} &= \frac{1}{N!}\left[\int_s d\mathbf{p} ~e^{-\beta \frac{\hbar^2 k^2}{2m}}\right]^{N}
+    Z_{id} &= \frac{1}{N!}\left[\int_s \diff\mathbf{p} ~e^{-\beta \frac{\hbar^2 k^2}{2m}}\right]^{N}
     \label{Zt_def}
@@
 \begin{align}
-    Z_{id} &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{-\frac{\beta \hbar^2}{2m} \left(\frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}}\right)^2}\right]^{N} = \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{ \frac{-2\beta \hbar^2\pi^2}{mL^2}(n_x^2 + n_y^2 + n_z^2)}\right]^{N}
+    Z_{id} &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{-\frac{\beta \hbar^2}{2m} \left(\frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}}\right)^2}\right]^{N} \\
+    &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{ \frac{-2\beta \hbar^2\pi^2}{mL^2}(n_x^2 + n_y^2 + n_z^2)}\right]^{N}
 \end{align}
@@
 where $\eta$ is the filling fraction.
+The Carnahan-Starling equation of state is an excellent approximation for low-density hard spheres.
+Close to solidification, near $\eta_{freeze} \approx .5$, this equation of state fails.
+Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\ \ignore{BLAH}
 Combining Eq.~\ref{C-S-eq} with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy.
@@
 \centering
-    %\includegraphics[width=.75\textwidth]{ww130-L03-i1_F-T.pdf}
-    \caption{Preliminary results for absolute specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$.
+    \includegraphics[width=.75\textwidth]{Fexc-vs-T.pdf}
+    \caption{Preliminary results for excess specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$.
 The vertical line at $T/\epsilon = 0.3$ is the lowest temperature at which our statistics are trustworthy.}
@@
 \end{figure}
+
 \begin{figure}
 \centering
-    %\includegraphics[width=.75\textwidth]{ww130-L03-F_i1_fixed_i.pdf}
-    \caption{Preliminary results for free energy as a function of density.}
+    \includegraphics[width=.75\textwidth]{Shs-vs-eta.pdf}
+    \caption{Hard sphere entropy as a function of density.}
+    \label{Shs-eta}
+\end{figure}
+\ignore{Filling fraction or density? Value in distinction? We never use number density for the sake of units...}
+
+\begin{figure}
+\centering
+    \includegraphics[width=.75\textwidth]{Uexc-vs-eta.pdf}
+    \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}
+    \label{U-eta}
+\end{figure}
+
+\begin{figure}
+\centering
+    \includegraphics[width=.75\textwidth]{Fexc-vs-eta.pdf}
+    \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}
     \label{F-eta}

### 2016-04-26 papers/thesis-vischer/Thesis.tex
More plot discussion, more information on Carnahan Starling Tests skipped!

@@
 where $\eta$ is the filling fraction.
-The Carnahan-Starling equation of state is an excellent approximation for low-density hard spheres.
-Close to solidification, near $\eta_{freeze} \approx .5$, this equation of state fails.
-Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\ \ignore{BLAH}
+The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit.
+We are very much concerned with cells significantly far from the thermodynamic limit.
+Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\
 Combining Eq.~\ref{C-S-eq} with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy.
 We determined that steps yielding an approximate change in specific absolute free energy of $\Delta F^{abs}/NT = \ln(2)$ through trial and error - smaller step sizes unnecessarily increased simulation time and larger step sizes tended to shrink too quickly causing the simulation to miss relevant states in the system.
+This step size corresponds to changing the entropy by $1/2$ each step.
 
@@
 
-Figures ~\ref{F-T} and ~\ref{F-eta} show preliminary free energy calculations.
+Figures ~\ref{F-T} through ~\ref{F-eta} show several computed thermodynamic quantities for the scrunched system.
+Each  It is worth noting that the two succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
+The intervening points have been filled in using the Carnahan-Starling equation of state.
 \begin{figure}[H]
@@
     \includegraphics[width=.75\textwidth]{Fexc-vs-T.pdf}
-    \caption{Preliminary results for excess specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$.
-The vertical line at $T/\epsilon = 0.3$ is the lowest temperature at which our statistics are trustworthy.}
+    \caption{Excess specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$. }
     \label{F-T}

### 2016-04-29 papers/thesis-vischer/Thesis.tex
add data for Brenden's thesis and make fac latex it

@@
 \centering
-    \includegraphics[width=.75\textwidth]{Fexc-vs-T.pdf}
+    \includegraphics[width=.75\textwidth]{../renormalization/figs/Fexc-vs-T.pdf}
     \caption{Excess specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$. }
@@
 \centering
-    \includegraphics[width=.75\textwidth]{Shs-vs-eta.pdf}
+    \includegraphics[width=.75\textwidth]{../renormalization/figs/Shs-vs-eta.pdf}
     \caption{Hard sphere entropy as a function of density.}
@@
     \centering
-    \includegraphics[width=.75\textwidth]{Uexc-vs-eta.pdf}
+    \includegraphics[width=.75\textwidth]{../renormalization/figs/Uexc-vs-eta.pdf}
     \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}
@@
 \centering
-    \includegraphics[width=.75\textwidth]{Fexc-vs-eta.pdf}
+    \includegraphics[width=.75\textwidth]{../renormalization/figs/Fexc-vs-eta.pdf}
     \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}

### 2016-04-29 papers/renormalization/figs/gatherandcalculate.py

@@ -28,13 +28,13 @@ def lndos_energy_Ns(dbase):
 def Sexc_hardsphere_Ns(dbase):
     Ns = []
     S = []
-    for N in range(2,100): # maybe go higher?
+    for N in range(2,600): # maybe go higher?
         fbase = '%s/N%03d/absolute/' % (dbase, N)
         # we only try to add this N value if we have one .dat file,
         # and our number of .dat files is the same as our number of
         # .out files.  If the latter is not true, we probably have not
         # finished running the absolute simulations.
-        if os.path.isfile(fbase+'00000.dat') and len(glob.glob(fbase+'*.dat')) == len(glob.glob(fbase+'*.out')):
+        if os.path.isfile(fbase+'Sexc.dat') or (os.path.isfile(fbase+'00000.dat') and len(glob.glob(fbase+'*.dat')) == len(glob.glob(fbase+'*.out'))):
             try:
                 thisS = Sexc_hardsphere(dbase, N)
                 S.append(thisS)
@@ -45,6 +45,10 @@ def Sexc_hardsphere_Ns(dbase):
 
 def Sexc_hardsphere(dbase, N):
     fbase = '%s/N%03d/absolute/' % (dbase, N)
+    if os.path.isfile(fbase+'Sexc.dat'):
+        foo = np.loadtxt(fbase+'Sexc.dat')
+        print foo, foo.shape
+        return foo
     S = 0
     # loop over files in the ./absolute/ directory
     j = 0

### 2016-04-30 papers/thesis-vischer/Thesis.tex
expanded absolute free energy derivation and added calculating quantities from simulation Tests skipped!

@@
 
-\subsection{Calculating the Absolute Change in Free Energy}
+\subsection{Calculating the Absolute Free Energy}
 It is straightforward the compute a Helmholtz free energy through the Monte-Carlo simulations; the difficulty in this method lies in computing the correct free energy.
@@
 
-At infinite temperature, the square well fluid behaves exactly as the hard sphere.
-The correct inclusion, then, is to enforce that the free energy calculated via TMMC at $T=\infty$ is equal to the free energy calculated through thermodynamic integration.
+At infinite temperature, the square well fluid has approximately the same behavior as the hard sphere fluid.
+We assert that the correct inclusion, then, is to enforce that the excess free energy calculated via TMMC at $T=\infty$ is equal to the free energy calculated through thermodynamic integration.
 We do this by exploiting the infinite temperature behavior of the canonical ensemble:
@@
     &= \sum_{i=0}^{\text{\tiny Energies}} D(E)\\
-    &= \#\text{~states}
+    &= \#\text{~states}\;.
 \end{align}
 
-\ignore{NB: cough cough the HS energy probs wrong}
-The partition function now calculates the total number of states in the system which we expected due to the equipartition theorem. The hard sphere free energy [more]
+The partition function now calculates the total number of states in the system which we expected due to the equipartition theorem. The excess free energy is simply
+\begin{align*}
+    F_{SW,~exc} &= U_{SW,~exc} - T\,S_{SW,~exc}     
+\end{align*} 
+and, at infinite temperature, the internal energy vanishes, leaving us with
+\begin{align}
+    F_{SW,~exc} &= -T\,S_{SW, ~exc} \;.
+\end{align}
+
+The hard sphere system has no excess internal energy, so the excess free energy is merely
+\begin{align*}
+     F_{HS,~exc} &= -T\,S_{HS,~exc} 
+\end{align*}
+
+Since the excess hard sphere free energy is really just the entropy (scaled by temperature), the statement that the square well and hard sphere excess free energies need be equal in the infinite temperature limit is really a constraint on the {\it entropy} of the two systems.
+We then conclude that
+\begin{align*}
+    S_{SW,~exc} \Bigm\lvert_{T=\infty} &= S_{HS,~exc}
+\end{align*}
+
+\subsection{Calculating Thermodynamic Quantities from Simulation Output}
+Given the density of states output $\ln[D(E)]$ from the Monte-Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can easily calculate, for instance, the excess internal energy:
+\begin{align*}
+     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E))] - \beta E_i}\\
+     &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i \exp\left(-\beta E_i\right)}{Z}
+\end{align*} 
+where, as usual, $\beta = T^{-1}$.
+Taking the exponential of $\ln(D(E))$, unfortunately, tends to introduce overflow or underflow errors during computation.
+We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots} from each element of $D(E)$, and then adding the maximum back in after the exponential.
+We calculate the excess free energy by using the entropy equality developed earlier:
+\begin{align*}
+     S_{SW,~exc}\Bigm\lvert_{T=\infty} &= S_{HS,~exc}\\
+     \implies F_{SW,~exc} &= -TS_{HS,~exc} + F_{inf}\\
+     F_{SW,~exc} &= -TS_{HS,~exc} - T\ln\left(\frac{Z}{Z_{inf}}\right)
+\end{align*} 
+The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration and both the partition functions are calculated the usual way [MORE ABOUT $Z_inf$].
+Once we have calculated both the excess free energy and excess internal energy, we can directly find the excess entropy:
+\begin{align*}
+    S_{SW, ~exc} &= \frac{1}{T}\left(U_{SW,~exc} - F_{SW,~exc} \right)
+\end{align*}
 

### 2016-05-02 papers/thesis-vischer/Thesis.tex
added to markov chains, sample conclusion, sample results discussion,  beginning of broad hist section to thesis Tests skipped!

@@
 \section{Methods}
-This work requires the intersection of three methods to calculate the total absolute free energy $F^{abs}_{tot}$ of the bulk cell.
+This work requires the intersection of three methods to calculate the absolute free energy $F^{abs}_{tot}$ of the bulk cell.
 A preliminary value of the free energy of the bulk can be obtained from Monte-Carlo simulations.
 This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the excess free energy of a system.
-[more]
+[more !fixme]
 \subsection{Monte-Carlo Methods}
@@
 Markov chain methods eschew the true randomness of regular Monte Carlo by taking the current configuration into account before moving.
-This method can lead to unsavory correlations in samples, but this danger is far outweighed by the usefulness of modeling systems where direct sampling is nigh impossible.
-\begin{figure}
-    \begin{tikzpicture}
-    \end{tikzpicture}
-\end{figure}
+If we define the initial state of the system by a vector column $\mathbf{q}$ (where the components refer to general coordinates, typically taken to be position of each sphere), and the resulting state $\mathbf{p}$ after a single time step, the relationship
+\begin{align*}
+    \mathbf{q} &= \mathcal{P}\, \mathbf{p}\\
+\end{align*}
+defines a ``transition'' of the system.
+$\mathcal{P}$ is a necessarily square matrix (as long as the number of coordinates remain fixed) that contains information about these transitions.
+We call this construction $\mathcal{P}$ the {\it transition matrix}.
+The transition matrix's $i^{\text{th}}$ eigenvalue the probability of the $i^{\text{th}}$ state occurring which, for the canonical ensemble.
+If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
+A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
+\begin{align*}
+    \mathcal{P}\,\mathbf{q}_{eq} &= \mathbf{q}_{eq}
+\end{align*}
+The transition matrix thusly embodies the physics of the simulations [more?]
 
-\begin{figure}
-    \caption{Monte-Carlo contrasted with Markov-Chain Monte-Carlo (This might be difficult...)}
-\end{figure}
 \ignore{\subsubsection{Histogram Methods}
@@
 \subsection{Broad Histogram Monte-Carlo}
-[Things about broad histogram methods]
-Broad histograms are focused entirely on determining the density of states $D(E)$ for a given system. [more]
+If one records the energy of the system at each time step, it is straightforward to construct a histogram of energies.
 
@@
 Figures ~\ref{F-T} through ~\ref{F-eta} show several computed thermodynamic quantities for the scrunched system.
-Each  It is worth noting that the two succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
-The intervening points have been filled in using the Carnahan-Starling equation of state.
-\begin{figure}[H]
+Each plot contains calculations from the first doubling regime ($L=2\lambda_0$) represented by solid lines and second doubling regime ($L = 4\lambda_0$) in dashes.
+It is worth noting that the succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
+The intervening points have been filled in using the Carnahan-Starling equation of state.\\
+\begin{figure}  
 \centering
@@
 
+Figure \ref{F-T} shows expected behavior.
+As a function of temperature, the excess free energy monotonically increases for all densities tested.
+We see a positive free energy because $F = U -TS$; as temperature increases without bound, entropy remains negative and [blah blah more].\\
+
 \begin{figure}
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Shs-vs-eta.pdf}
-    \caption{Hard sphere entropy as a function of density.}
+    \caption{Excess hard sphere entropy as a function of density.}
     \label{Shs-eta}
@@
 \ignore{Filling fraction or density? Value in distinction? We never use number density for the sake of units...}
+Figure \ref{Shs-eta} shows the computed excess entropy as a function of density.
+We see that the both the first and second doubling regime closely match the Carnahan-Starling prediction for the entropy.
+After freezing, the first doubling regime veers from the prediction by the Carnahan-Starling equation of state.\\
 
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Uexc-vs-eta.pdf}
-    \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}
+    \caption{Excess internal energy as a function of density. The $x$ markers connected by dashed lines delineate second doubling regime densities that were simulated.}
     \label{U-eta}
 \end{figure}
+We see in Figure~\ref{U-eta} that the excess internal energy, as we would expect, becomes more negative as the density increases.
+The lower the temperature, the more ``clumped'' the system becomes, resulting in lower internal energies.
 
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Fexc-vs-eta.pdf}
-    \caption{Preliminary results for free energy as a function of density. Solid line represents first doubling regime, [more]}
+    \caption{Excess specific free energy as a function of density.}
     \label{F-eta}
@@
 \section{Conclusion}
-
-\subsection{Future Research}
-There are many other models that GRGT might be applicable to, such as the soft-sphere liquid or soft polyhedra.
+We have presented a method by which the properties of bulk fluids can be estimated by decomposing the bulk into individual cells in a renormalization-minded manner by exploiting the critical behavior of fluids.
+The free energy contribution of each length scale can then be investigated and compared, both to determine the most energetic density fluctuations and to provide calculations that serve as a intermediate comparison with SAFT methods.
+This method is immediately applicable to any system where the energy of a configuration is straightforward to compute (that is, relatively inexpensive computationally).
+We presented an application for a system that obeys the canonical ensemble, but the extension to the grand canonical ensemble can be done with little modification.
 %\bibliographystyle{plain}

### 2016-05-04 papers/thesis-vischer/Thesis.tex
more cleanup, revision, content to methods Tests skipped!

@@
 %\usepackage{float}
+
+% Header definition
+\usepackage{fancyhdr}
+\pagestyle{fancy}
+\fancyhead{}
+%\renewcommand{\sectionmark}[1]{\markright{\thesection.\ #1}}
+
+\fancyhead[LE, RO]{\slshape \rightmark}
+\fancyhead[RE, LO]{\slshape Chapter \leftmark}
+
+% end header
+
 \usetikzlibrary{automata, positioning}
@@
 
-\renewcommand{\title}{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}
+\renewcommand{\title}{Free Energy Decomposition of the Critical Square-Well Liquid Using a Renormalization Group Method}
 
@@
 Fluids have long garnered interest\cite{theorysimpleliquids}\cite{theoryofcrits}, particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
-A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor are indistinguishable and, more importantly, large fluctuations of fluid density occur.
+A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor become indistinguishable and, more importantly, large fluctuations of fluid density occur.
 Most traditional fluid models assume the density across the cell to be nearly constant.
@@
 
-Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success, particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
-GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating a vast number of cells, varying the parameter each time, to find the most well-suited value for a particular fluid.
+Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success [cite me], particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
+GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating a vast number of cells, varying the parameter each time, to find the most well-suited value for a particular fluid [!fixme not quite].
 We present a non-fitted approach that is GRG-minded from the ground up.
@@
 The hard sphere model approximates monoatomic fluids as having no interactions other than collisions.
-These spheres exhibit no attraction or repulsion, and consequently each carries kinetic energy $T = 3/2 k_b T$.
-There is also a harder-to-calculate ``excess'' energy to be considered (which, in fact, contains most of the interesting behavior) which will be discussed further on.
+These spheres exhibit no attraction, and consequently each carries kinetic energy $K = 3/2 k_b T$.
+There is also a harder-to-calculate [!fixme] ``excess'' energy to be considered (which, in fact, contains most of the interesting behavior) which will be discussed further on.
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
@@
 \subsection{Square-Well Liquid}
-The square-well liquid is a ubiquitous approximation that sees extensive use throughout computational fluid research.
+The square-well liquid is a common approximation that sees extensive use throughout computational fluid research.
 This simple fluid model consists of hard spheres that are subject to an intermolecular pair potential - that is, an attractive force between spheres.
@@
 
-The energy of the fluid is then always an integer multiple of the well depth $\epsilon$.
+The potential energy of the fluid is then always an integer multiple of the well depth $\epsilon$.
 This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
@@
 \begin{align}
-    Z &= \sum^{\text{energies}}~e^{-\beta E_i}
+    Z &= \sum^{\text{states}}~e^{-\beta E_i}
 \end{align}
-where $\beta =  T^{-1}$. Of course, it is typically impractical to sum over the individual energies of the system, especially if there are expected degeneracies. Instead we sum over states:
+where $\beta =  T^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$.
+Of course, it is typically impractical to sum over the individual states of the system, especially if there are expected degeneracies.
+Instead we sum over energies:
 \begin{align}
-    Z &= \sum^{\text{states}}~D(E)~e^{-\beta E_i}
+    Z &= \sum^{\text{energies}}~D(E)~e^{-\beta E_i}
 \end{align}
@@
 \begin{align}
-    Z &= \sum_i e^{-\beta E_i}
+    Z &= \sum_i D(E) e^{-\beta E_i}
 \end{align}
-where we sum over individual energies of the system $E_i$.
-However, this method doesn't account for degeneracies in the system, so instead we sum over the states $s$ (to avoid including the density of states, which is difficult to find for most systems):
+where we sum over individual energies of the system $E_i$. [reason that we sum over states] , so instead we sum over the states $s$:
 \begin{align}
@@
 However, the above is only partially correct, as it is only for one particle.
-We can account for multiple particles by requiring that each particle has the same partition function contribution for kinetic energy; however, the potential energy does not seperate as nicely since it depends on the position of all particles.
+We can account for multiple particles by requiring that each particle has the same partition function contribution for kinetic energy; however, the potential energy does not separate as nicely since it depends on the position of all particles.
 For the kinetic energy, we are simply integrating $N$ times for $N$ particles:
@@
     &= \int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
+    \label{Z-intermed}
 \end{align}
-where we have separated the integrals by recognizing that the integrands are independant.\\
+where we have separated the integrals by recognizing that the integrands are independent.\\
 Additionally, for a proper quantum mechanical treatment we must assume that the particles are indistinguishable which means the above expression over-counts by a factor of $N!$.
+Additionally, we will multiply Eq.
+\ref{Z-intermed} by an overall factor of $\frac{V^N}{V^N}$ to preserve units and extensivity.
 Our expression for the partition function is then
@@
 \begin{align}
-    Z &= \frac{1}{N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
+    Z &= \frac{1}{V^N N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[V \int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
     \label{Z_full}
@@
 \end{align}
-This unfortunately introduces a mass quantity into the partition function which adds a parameter that needs fitting. [more] 
+This unfortunately introduces a mass quantity into the partition function which adds an additional parameter that needs to be determined. [more] 
 If we then allow the cell to become very large, then the we can easily see in Eq.
@@
 \end{align}
-where we have introduced the thermal deBroglie wavelength $\Lambda$, loosely interpreted as [physical thing]. The full partition function for any monoatomic system is then
+where we have introduced the thermal deBroglie wavelength $\Lambda$ which defines the de Broglie wavelength of an ideal gas at a temperature $T$.
+The full partition function for any monoatomic system is then
 \begin{align}
@@
 \end{align}
-For liquids, the $Z_{exc}$ term tends to be difficult to evaluate; the hard sphere liquid, the simplest model for a fluid, does not have an analytic representation for the excess free energy.
+For liquids [fluids?], the $Z_{exc}$ term is difficult to ever evaluate; the hard sphere liquid, the simplest model for a dense fluid, does not have an analytic representation for the excess free energy.
 It follows that a considerable amount of work is required to determine the excess free energy for a given system.
@@
 Statistical mechanics, however, does not exhibit this independence.
-Each time the man makes a move there is an equal probability that he will move in any particular direction.
+Each time the drunken man makes a move there is an equal probability that he will move in any particular direction.
 Rather than all movements having equal probability, statistical mechanics stipulates that the probability of a particular move occurring depends explicitly upon the energy difference of the current and resulting configurations.
-Instead of traditional Monte Carlo methods, Markov chain (specifically the Metropolis-Hastings algorithm) Monte Carlo is the natural decision.
+Instead of true random Monte Carlo methods, we clearly need some way to control the probability of transitions from a given state.
+In pursuit of this control, we employ the method of Markov chains, specifically Metropolis-Hastings Monte Carlo.
 \subsubsection{Markov Chain}
+[cite me pls]
 Markov chain methods eschew the true randomness of regular Monte Carlo by taking the current configuration into account before moving.
@@
 \begin{align*}
-    \mathbf{q} &= \mathcal{P}\, \mathbf{p}\\
+    \mathbf{p} &= \mathcal{P}\, \mathbf{q}\\
 \end{align*}
@@
 If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
-A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
+[This is the best I could synthesize Hansen, following statement could be wrong!] A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
 \begin{align*}
@@
 \end{align*}
-The transition matrix thusly embodies the physics of the simulations [more?]
+The transition matrix thus embodies the physics of the simulation. [more?]
 
@@
 If one records the energy of the system at each time step, it is straightforward to construct a histogram of energies.
+Typical implementations of histogram methods exhibit histograms that sharply peak near a central value $E_0$ which then symmetrically decays.
+Of course, at higher temperatures, the histogram broadens and flattens out as more energies become easily accessible and, in the limit that temperature becomes infinite, the histogram becomes completely flat since all states are equally likely.
+If the simulation is allowed to run for an infinite amount of time, the histogram will be precisely detail the system.
+x The tails of the histogram, however, contain spurious statistics because the system rarely visits these energies.
+Gathering sufficient statistics
 
@@
 We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots} from each element of $D(E)$, and then adding the maximum back in after the exponential.
-We calculate the excess free energy by using the entropy equality developed earlier:
+We calculate the excess free energy by using the entropy equality developed previously:
 \begin{align*}
@@
 We see that the both the first and second doubling regime closely match the Carnahan-Starling prediction for the entropy.
-After freezing, the first doubling regime veers from the prediction by the Carnahan-Starling equation of state.\\
+After freezing, the first doubling regime veers from the prediction by the Carnahan-Starling equation of state; since the Carhan-Starling approximation is accurate only for low density fluids, we expect this behavior.\\
 
@@
 This method is immediately applicable to any system where the energy of a configuration is straightforward to compute (that is, relatively inexpensive computationally).
-We presented an application for a system that obeys the canonical ensemble, but the extension to the grand canonical ensemble can be done with little modification.
+The application presented was a system that obeys the canonical ensemble, but the extension to the grand canonical ensemble can be done with little modification.
+Future work can include stuff [more].
 %\bibliographystyle{plain}
+\cleardoublepage
 \bibliography{Thesis}

### 2016-05-05 papers/thesis-vischer/Thesis.tex
added more derivation things, corrections, more about computing quantities Tests skipped!

@@
     \centering
-    %\includegraphics[width=.75\textwidth]{SWF-E.png}
+    \includegraphics[width=.75\textwidth]{SWF-E.png}
     \caption{Demonstration of Square-Well approximation.}
@@
     \centering
-    %\includegraphics[width=.75\textwidth]{phase.png}
-    \caption{Phase transitions of water with critical point and triple point identified.}
+    \includegraphics[width=.75\textwidth]{phase.png}
+    \caption{Phase diagram of water with critical point and triple point identified.}
     \label{phase}
@@
 \begin{align}
-    Z &= \frac{1}{V^N N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[V \int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
+    Z &= \frac{V^N}{V^N N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
     \label{Z_full}
@@
     Z_{id} &= \left(\frac{V^N}{N!\Lambda^{3N}}\right) \,, \\
-    \Lambda &= \frac{\hbar\sqrt{2\pi\beta}}{\sqrt{m}}
+    \Lambda &= \frac{\hbar\sqrt{2\pi}}{\sqrt{mT}}
 \end{align}
@@
 \begin{align}
-    Z &= Z_{exc} \left(\frac{V}{\Lambda^{3}}\right)^{N} \,.
+    Z &= Z_{exc} \left(\frac{V}{N!\Lambda^{3}}\right)^{N} \,.
 \end{align}
@@
 It follows that a considerable amount of work is required to determine the excess free energy for a given system.
-
+The absolute free energy, by way of Eq.
+\ref{FdefSM}, now has the form
+\begin{align}
+    F_{abs} &= -T\,\ln\left(Z_{exc} \frac{V^N}{N!\Lambda^{3N}}\right)\\
+    &= -T\ln\left(Z_{exc}\right)\, -T\left(\frac{V}{(N!)^{1/N}\Lambda^3} \right)^N\\
+    &= -T\ln\left(Z_{exc}\right) - NT\ln\left(\frac{V}{\Lambda^3}\right) + NT\ln\left(N!^{1/N}\right)\\ 
+\end{align}
+Redefining the term including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at
+\begin{align}
+    F_{abs} &= F_{exc} - NT\ln\left(\frac{V}{\Lambda^3}\right) + NT\ln\left(N\right)
+\end{align}
+At a specified volume and number of particles, the ideal gas contribution is constant. Consequently, for a given temperature we only need to calculate the excess free energy. 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Methods}
-This work requires the intersection of three methods to calculate the absolute free energy $F^{abs}_{tot}$ of the bulk cell.
+This work requires the intersection of three methods to calculate the absolute free energy $F_{abs}$ of the bulk cell.
 A preliminary value of the free energy of the bulk can be obtained from Monte-Carlo simulations.
-This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the excess free energy of a system.
-[more !fixme]
+This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the  free energy of the hard-sphere system.
+The resulting absolute free energies for each fluctuation length can be summed to yield the absolute free energy of the bulk.
+[more]
 \subsection{Monte-Carlo Methods}
@@
 Typical implementations of histogram methods exhibit histograms that sharply peak near a central value $E_0$ which then symmetrically decays.
-Of course, at higher temperatures, the histogram broadens and flattens out as more energies become easily accessible and, in the limit that temperature becomes infinite, the histogram becomes completely flat since all states are equally likely.
+Of course, at higher temperatures, the histogram broadens and flattens out as more energies become easily accessible and, in the limit that temperature becomes infinite, the histogram becomes completely flat since all states are equally likely [flat?
+what about $D(E)$].
 If the simulation is allowed to run for an infinite amount of time, the histogram will be precisely detail the system.
@@
 \begin{align}
-    Z(T=\infty) &= \left.\sum_{i=0}^{\text{\tiny Energies}} D(E) e^{-\beta E}\right|_{\beta = 0}\\
+    Z_{inf} &= \left.\sum_{i=0}^{\text{\tiny Energies}} D(E) e^{-\beta E}\right|_{\beta = 0}\\
     &= \sum_{i=0}^{\text{\tiny Energies}} D(E)\\
@@
 
-\subsection{Calculating Thermodynamic Quantities from Simulation Output}
-Given the density of states output $\ln[D(E)]$ from the Monte-Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can easily calculate, for instance, the excess internal energy:
+\subsection{Thermodynamic Quantities from Simulation}
+Given the density of states output $\ln[D(E)]$ from the Monte-Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic , for instance, the excess internal energy:
 \begin{align*}
@@
 \section{Discussion}
-If the decomposition procedure outlined previously proves successful, the absolute free energy yielded by this renormalization method ought to agree with the free energy yielded from other accepted theories.
-[more]
-\subsection{Change in Free Energy Over Multiple Regimes}
-[more]
-\subsection{Comparison With A Modified Statistical Associating Fluid Theory}
-[more]
+If the decomposition procedure outlined previously proves successful, the absolute free energy ought to obey known physics about the system.
+We particularly care about identifying liquid-vapor coexistence within the fluid. 
+\ignore{\subsection{Comparison With A Modified Statistical Associating Fluid Theory}
+[more]}
 

### 2016-05-05 papers/thesis-vischer/abstract.tex
added more derivation things, corrections, more about computing quantities Tests skipped!

@@
 Analytic models using renormalization methods for fluids have been [time scale] developed [cite], but computational models [something].
-[something about including all lengths, not knowning what each contributes].
-We present a method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Monte-Carlo simultations.
+[something about including all lengths, not knowing what each contributes].
+We present a method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Monte-Carlo simulations.
 We further detail a method by which to compute the ``absolute'' free energy, with both the ideal gas free energy and the excess free energy, for the square-well liquid.
-[more] We find that the square-well liquid near the critical region is well-characterized by a base cell length of $\sqrt2\sigma$, the side length of an FCC lattice.
+We determine that the square-well liquid near the critical region is well-characterized by a base cell length of $\sqrt2\sigma$, the side length of an FCC lattice, by establishing liquid-vapor coexistence between scaled temperatures $.5 < T < 1$.
 \clearpage

### 2016-05-06 papers/thesis-vischer/Thesis.tex
corrections, more discussion. Tests skipped!

@@
 Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success [cite me], particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
-GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating a vast number of cells, varying the parameter each time, to find the most well-suited value for a particular fluid [!fixme not quite].
+GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid [!fixme not quite].
 We present a non-fitted approach that is GRG-minded from the ground up.
 We exploit the periodicity of Monte-Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
-We then carefully couple simulations of different ``doubling regimes'', each of which considers a different maximum wavelength of fluctuations, where the number density remains fixed but the volume and number have increased.
-Summing the difference in absolute Helmholtz free energy $\Delta F^{\text{abs}}$ between doubling regimes will then approximate the absolute free energy of the full cell, including near criticality.
-Finding absolute free energies is a somewhat challenging problem, but is certainly surmountable with more Monte Carlo.
-\\
+We then carefully couple simulations of different ``doubling regimes'', each of which considers a different maximum wavelength of fluctuations, where the fluid density $\eta$ remains fixed but the volume $V$ and number of particles $N$ have increased.
+Summing the absolute Helmholtz free energy $F^{\text{abs}}$ of each doubling regime will then approximate the absolute free energy of the full cell, especially near the critical point.
 We will begin by outlining the necessary models and assumptions/approximations for this work.
@@
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
-For simulation purposes, the system is required to have periodic boundary conditions - that is, when a sphere would classically collide with the wall of the cell it instead breaches the wall and then reappears on the other side of the cell with the same velocity.
 \\
 \ignore{The hard sphere model is incredibly useful; as high temperatures are approached, energy associated with sphere location tends to asymptote for most models [are there any that this isn't true for?] and the kinetic term consequently dominates.}
+
+\ignore{[move me | For simulation purposes, the system is required to have periodic boundary conditions - that is, when a sphere would classically collide with the wall of the cell it instead breaches the wall and then reappears on the other side of the cell with the same velocity.]}
 
@@
 Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ (as well as any other physical constants) to one.
-This then allows us to resonably compare energies to temperatures since they have the same units.
+This then allows us to reasonably compare energies and temperatures since they have the same units.
 
@@
 \end{figure}
+
 \subsection{Critical Point} 
-The critical point is the temperature and pressure (or any other thermodynamic pairing) on a phase diagram (shown in Figure ~\ref{phase}) where the line between liquid and vapor terminates - liquid and vapor become indistinguishable and are assimilated into the ``fluid'' phase.
+The critical point is the temperature and pressure (or any other thermodynamic pairing) on a phase diagram (shown in Figure ~\ref{phase}) where the line between liquid and vapor terminates - liquid and vapor become indistinguishable and are assimilated into the ``supercritical fluid'' phase.
 Many interesting effects occur here, including the well-researched phenomena of critical opalescence.
@@
 This self-similarity allows for the decomposition of the large critical cell to smaller constituent cells.
-The details of the procedure are more carefully discussed later on.
+The details of the procedure are more carefully discussed in Chapter {\color{red} [blah]}.
 
 \subsection{Free energy}
-The free energy we are interested in calculating is the ``absolute'' Helmholtz free energy $F^{\text{abs}}$. The Helmholtz free energy is a Legendre transform of the internal energy $U$:
+The energy we are interested in calculating is the ``absolute'' Helmholtz free energy $F^{\text{abs}}$. The Helmholtz free energy is a Legendre transform of the internal energy $U$
 \begin{align}
-    F &= U - TS
+    F &= U - TS\, ,
     \label{Fdef}
@@
 \begin{align}
-    F &= -T \ln Z
+    F &= -T \ln Z\, .
     \label{FdefSM}
 \end{align}
-With $F$ calculated, simple partial derivatives yield all other thermodynamic quantities.
-Though the Gibbs free energy yields the same information, the Helmholtz free energy is particularly useful because of the relationship with entropy:
+With $F$ calculated, partial derivatives yield all other interesting thermodynamic quantities.
+Though the Gibbs free energy yields the same information, the Helmholtz free energy is particularly useful because of the relationship with entropy, namely
 \begin{align}
-    S &= \left(\frac{\partial F}{\partial T}\right)_{V}
+    S &= \left(\frac{\partial F}{\partial T}\right)_{V}\, .
 \end{align}
-Since we are defining cells of fixed volume this partial derivative will be simple to evaulate for any temperature $T$.\\
+Further, if we allow the number of particles in the system to vary, we have the grand free energy $\Phi$, defined as
+\begin{align}
+    \Phi &= U-TS-\mu N\\
+    &= F - \mu N
+    \label{Phidef}
+\end{align}
+where $\mu$ is the chemical potential of the fluid, which is related to the tendency of particles to flow into or out of the system.\\
 
-Equations ~\ref{Fdef} and ~\ref{FdefSM} together bridge classical thermodynamics and statistical mechanics.
-Due to the nature of this project, we will primarily be concerned with the statistical mechanics definition (that is, Eq.
-~\ref{FdefSM}) of the free energy.
-We are, then, ultimately concerned with determining the partition function for any system we are concerned with.
+Equations~\ref{Fdef} and~\ref{FdefSM} together bridge classical thermodynamics and statistical mechanics.
+Due to the nature of this project, we will primarily be concerned with the statistical mechanics definition (that is, Eq.~\ref{FdefSM}) of the free energy.
+We are, then, ultimately concerned with determining the partition function for any given system.
 The canonical ensemble partition function can be compactly represented by a summation:
@@
 \end{align}
-where $\beta =  T^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$.
-Of course, it is typically impractical to sum over the individual states of the system, especially if there are expected degeneracies.
+where $\beta =  T^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$ state.
+Of course, it is impractical to sum over the individual states of the system because the number of states are generally innumerable.
 Instead we sum over energies:
@@
 where we have now introduced the density of states $D(E)$ which is a simple function that returns the number of states which have a given energy $E$.
-This introduction is now the focus of calculation - if we can accurately determine the density of states of the system and have knowledge of the available energies, we can calculate the Helmholtz free energy and, from there, any other  \\
+This introduction is now the focus of calculation - if we can accurately determine the density of states of the system and have knowledge of the available energies, we can calculate the Helmholtz free energy and, from there, any other thermodynamic quantity we wish.
+\\
 Since $F$ contains a natural log term, we can separate and characterize the individual contributions of kinetic energy (``ideal'') and energy associated with the configuration of the system (``excess'').
@@
 \end{align}
-At a specified volume and number of particles, the ideal gas contribution is constant. Consequently, for a given temperature we only need to calculate the excess free energy. 
+At a specified volume and number of particles, the ideal gas contribution is constant. Consequently, for a given temperature we only need to compute the excess free energy.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \ignore{\subsubsection{Histogram Methods}
-Histograms are graphical representation of ``counting'' data. Monte Carlo histogram methods simply track the number of times the system has visited a particular energy. If the transition matrix 
+Histograms are graphical representatio
+
+n of ``counting'' data. Monte Carlo histogram methods simply track the number of times the system has visited a particular energy. If the transition matrix 
 \begin{figure}
@@
 
-Since the excess hard sphere free energy is really just the entropy (scaled by temperature), the statement that the square well and hard sphere excess free energies need be equal in the infinite temperature limit is really a constraint on the {\it entropy} of the two systems.
+Since the excess hard sphere free energy is really just the entropy[cite] (scaled by temperature), the statement that the square well and hard sphere excess free energies need be equal in the infinite temperature limit is really a constraint on the {\it entropy} of the two systems.
 We then conclude that
@@
 \begin{align*}
-     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E))] - \beta E_i}\\
+     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E)] - \beta E_i})\\
      &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i \exp\left(-\beta E_i\right)}{Z}
@@
 Taking the exponential of $\ln(D(E))$, unfortunately, tends to introduce overflow or underflow errors during computation.
-We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots} from each element of $D(E)$, and then adding the maximum back in after the exponential.
+We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots}from each element of $D(E)$, and then adding the maximum back in after the exponential.
 We calculate the excess free energy by using the entropy equality developed previously:
@@
 Using the coupling procedure outlined previously, we calculated changes in free energy up to the $n^{\text{th}}$ doubling regime for several different cell lengths.
-We are primarily concerned with the ``scrunched'' case where the cell length is $\sqrt2\sigma$.
+We are primarily concerned with the ``scrunched'' case where the cell length is $\sqrt2\sigma$.\\
 
@@
 The lower the temperature, the more ``clumped'' the system becomes, resulting in lower internal energies.
+[correct?]
 
@@
 We chose to focus on the scrunched case for two reasons.
-Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length
+Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length [necessary?]
 
+\ignore{ Might never need these sections. |
 \subsubsection{Seed Disparity}
@@
 \subsection{Other Cells?}
-In addition to the ``scrunched'' primitive cell, we tested [hopefully one or two] other minimum correlation lengths corresponding to other natural parameters in the system.
+In addition to the ``scrunched'' primitive cell, we tested [hopefully one or two] other minimum correlation lengths corresponding to other natural parameters in the system.}
 
@@
 If the decomposition procedure outlined previously proves successful, the absolute free energy ought to obey known physics about the system.
-We particularly care about identifying liquid-vapor coexistence within the fluid. 
+We particularly care about identifying liquid-vapor coexistence within the fluid. \\
+
+Finding coexistence requires use of the grand free energy $\Phi$.
+Liquid and vapor are simultaneously present in the fluid at densities that cause local minima in the grand free energy.
+Referring back to Eq.~\ref{Phidef}, we can compute an absolute grand free energy given the absolute free energy and number of spheres.
+The chemical potential $\mu$ is a somewhat unintuitive quantity that, for our purposes, we will treat as a free parameter.
+Tweaking the value of the chemical potential will allow us to easily identify coexistence in the fluid.
+\begin{figure}
+\centering
+    \includegraphics[width=.75\textwidth]{../renormalization/figs/Phiabs-vs-eta.pdf}
+    \caption{Absolute specific free energy as a function of density.}
+    \label{Phi-eta}
+\end{figure}
+
 \ignore{\subsection{Comparison With A Modified Statistical Associating Fluid Theory}

### 2016-05-06 papers/thesis-vischer/Thesis.tex
Broad histogram section information Tests skipped!

@@
 We present a non-fitted approach that is GRG-minded from the ground up.
-We exploit the periodicity of Monte-Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
+We exploit the periodicity of Monte Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
 We then carefully couple simulations of different ``doubling regimes'', each of which considers a different maximum wavelength of fluctuations, where the fluid density $\eta$ remains fixed but the volume $V$ and number of particles $N$ have increased.
@@
 This work requires the intersection of three methods to calculate the absolute free energy $F_{abs}$ of the bulk cell.
-A preliminary value of the free energy of the bulk can be obtained from Monte-Carlo simulations.
+A preliminary value of the free energy of the bulk can be obtained from Monte Carlo simulations.
 This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the  free energy of the hard-sphere system.
@@
 [more]
-\subsection{Monte-Carlo Methods}
-Monte-Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
+\subsection{Monte Carlo Methods}
+Monte Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
 Traditional random sampling requires that the current step, or ``move'', is irrespective of previous steps.
@@
 \subsubsection{Markov Chain}
-[cite me pls]
+\ignore{[I think this is framed incorrectly... markov chains are used to generate the transition matrix, no? That's the procedure outlined by de Oleveira, reinforced by michael...]}
 Markov chain methods eschew the true randomness of regular Monte Carlo by taking the current configuration into account before moving.
@@
 If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
-[This is the best I could synthesize Hansen, following statement could be wrong!] A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
-\begin{align*}
+A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
+\begin{align}
     \mathcal{P}\,\mathbf{q}_{eq} &= \mathbf{q}_{eq}
-\end{align*}
+    \label{thermo-steady-state}
+\end{align}
 The transition matrix thus embodies the physics of the simulation. [more?]
 
-\ignore{\subsubsection{Histogram Methods}
-Histograms are graphical representatio
+\subsection{Broad Histogram Monte Carlo}
+If the energy of the system at each time step in a Monte Carlo simulation is recorded, it is straightforward to construct a histogram of energies.
+That is, a histogram with a bin for each narrow $\Delta E$ around a value $E$ that displays the number of times the system visited that energy range.
+Typical implementations of histogram methods follow the ``canonical histogram'' procedure outlined by Ferrenberg and Swendsen \cite{ferrenberg1988histogram} -- given a histogram $h_T(E)$ extracted from a simulation at some temperature $T_0$, a reweighting scheme can then extend the histogram to any other temperature $T$ via
+\begin{align}
+    h_{T}(E) &= h_{T_0}(E) \exp\left(E/T - E/T_0 \right)
+\end{align}
+and then normalizing appropriately.
+This method unfortunately loses substantial accuracy far from the original temperature and causes an untoward spike near the average value of the original histogram.
+Further, this reweighting detracts from the accuracy of the tails of the histogram.
+We choose instead to implement a similar approach to that of P.M.C.
+de Oleveira and coworkers\cite{de1998broad}, the ``broad histogram''.
+[more] This process is outlined in Algorithm~\ref{alg_tmmc}.\\
 
-n of ``counting'' data. Monte Carlo histogram methods simply track the number of times the system has visited a particular energy. If the transition matrix 
-\begin{figure}
-    \caption{Sample histogram.}
-\end{figure}}
-\subsection{Broad Histogram Monte-Carlo}
-If one records the energy of the system at each time step, it is straightforward to construct a histogram of energies.
-Typical implementations of histogram methods exhibit histograms that sharply peak near a central value $E_0$ which then symmetrically decays.
-Of course, at higher temperatures, the histogram broadens and flattens out as more energies become easily accessible and, in the limit that temperature becomes infinite, the histogram becomes completely flat since all states are equally likely [flat?
-what about $D(E)$].
-If the simulation is allowed to run for an infinite amount of time, the histogram will be precisely detail the system.
-x The tails of the histogram, however, contain spurious statistics because the system rarely visits these energies.
-Gathering sufficient statistics
+[stuff: markovian walk along energies known a priori, accept state always if $\Delta E< 0$, potentially accept state if\dots what?
+We can't directly enumerate states.
+Must be in Michael thesis somewhere.
+Check wang-landau paper?
+or [8] on michael thesis]
 
-{\color{red} succeeding text may be worthless.}
+\begin{algorithm}[tb]
+    \caption{Transition Matrix Monte Carlo}
+    \label{alg_tmmc}
+    \hrulefill
+    \begin{enumerate}
+        \item hi
+    \end{enumerate}
+    \hrulefill
+\end{algorithm}
 
-The method of the transition matrix has existed for over a decade and has seen several varying implementations.
-The canonical form involves knowing the exact number of states, and more importantly the relative transition probabilities, available to the system {\it a priori}, which is simply not possible for cells as large as we are interested in~\cite{perlinthesis}.
-Other methods spends the majority of run time initializing - that is, constructing the weighting arrays, which then are used to form the transition matrix [possibly incorrect].
-Once the transition matrix is complete, we need only wait to gather sufficient statistics for the energies we are interested in.
-In the pursuit of time-saving, {\color{red} we have implemented a particular type of transition matrix construction that spends this initialization time both constructing the transition matrix on the fly and gathering statistics.
-Further, our method outputs the density of states as its sole output, decreasing the file storage burden of the simulations.}
-\begin{figure}
-    \caption{Sample transition matrix/demonstration of transition matrix.}
-\end{figure}
-\subsubsection{Algorithm}
+Calculating the density of states, given the transition matrix constructed via the markov chain process, is a trivial matter.
+The equilibrium condition given in Eq.~\ref{thermo-steady-state} is really a statement about the density of states; for a Monte Carlo simulation, the equilibrium state $q_{eq}$ is exactly the density of states.
+Solving the steady state condition as an eigenvalue problem for $q_{eq}$ then yields the density of states.
 
 \subsection{Renormalization Group Theory}
-Renormalization Group attempts to describe how a particular system's properties are altered when the system is viewed from varying length scales.
+Renormalization group methods attempt to describe how a particular system's properties are altered when the system is viewed from varying length scales.
 Near the critical point, the fluid becomes self similar, and so fluctuations of all wavelengths are present.
-We can then compare quantities of different, yet related, systems as they all exhibit the same properties.\\
+We can then compare quantities of related yet physically different systems as they all exhibit the similar properties.\\
 We begin with a periodic cubic cell of length $L_0$ and volume $L_0^3$. 
@@
 Thermodynamic Integration is a unique way to determine the ``excess'' free energy of the system.
-Using hard sphere Monte-Carlo, we transition from a filling fraction of zero (note that this statement is identical to assuming that the cell begins at infinite temperature, since the total energy does not depend on where the spheres are placed) to a near-close-packed system with filling fraction $\eta_max$, calculating the change in free energy at each step.
+Using hard sphere Monte Carlo, we transition from a filling fraction of zero (note that this statement is identical to assuming that the cell begins at infinite temperature, since the total energy does not depend on where the spheres are placed) to a near-close-packed system with filling fraction $\eta_max$, calculating the change in free energy at each step.
 When we have reached the desired final filling fraction we have found the free energy that was necessary to create that configuration of particles - this is exactly the absolute free energy of the system.\\
@@
 \subsection{Calculating the Absolute Free Energy}
-It is straightforward the compute a Helmholtz free energy through the Monte-Carlo simulations; the difficulty in this method lies in computing the correct free energy.
+It is straightforward the compute a Helmholtz free energy through the Monte Carlo simulations; the difficulty in this method lies in computing the correct free energy.
 TMMC, as mentioned previously, yields a free energy accurate only up to a unknown constant.
@@
 \subsection{Thermodynamic Quantities from Simulation}
-Given the density of states output $\ln[D(E)]$ from the Monte-Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic , for instance, the excess internal energy:
+Given the density of states output $\ln[D(E)]$ from the Monte Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic , for instance, the excess internal energy:
 \begin{align*}
@@
 \subsubsection{Seed Disparity}
-Because of the random/psuedorandom nature of Monte-Carlo sampling, we tested over many different initial sphere placements (``seeds'') to see if there was consistent agreement. 
+Because of the random/psuedorandom nature of Monte Carlo sampling, we tested over many different initial sphere placements (``seeds'') to see if there was consistent agreement. 
 \subsection{Other Cells?}

### 2016-05-06 papers/thesis-vischer/Thesis.tex
Finished broad hist section Tests skipped!

@@
 As a result, these methods tend to poorly handle these fluctuations near criticality.
-Using the well-established model of the square-well fluid, we seek to determine the applicability of a particular generalized renormalization-group formalism (``GRG'') to find the absolute Helmholtz free energy of systems near criticality.\\
+Using the well-established model of the square-well fluid, we seek to determine the applicability of a particular generalized renormalization group-like decomposition procedure to find the absolute Helmholtz free energy of systems near criticality.\\
 
 Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success [cite me], particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
-GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid [!fixme not quite].
-We present a non-fitted approach that is GRG-minded from the ground up.
+GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid [!fixme not quite].\\
+
+We present here a method to calculate free energy of a liquid that largely avoids this fitting problem.
 We exploit the periodicity of Monte Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
@@
 We will then detail the algorithms used to compute the quantities with the the discussion on results following.
+
 \subsection{Hard Sphere}
@@
 These spheres exhibit no attraction, and consequently each carries kinetic energy $K = 3/2 k_b T$.
-There is also a harder-to-calculate [!fixme] ``excess'' energy to be considered (which, in fact, contains most of the interesting behavior) which will be discussed further on.
+Further, with no attraction, there is no energy associated with [more].
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
 \\
-\ignore{The hard sphere model is incredibly useful; as high temperatures are approached, energy associated with sphere location tends to asymptote for most models [are there any that this isn't true for?] and the kinetic term consequently dominates.}
 
@@
 [more]
+
 \subsection{Monte Carlo Methods}
+
 Monte Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
@@
 Instead of true random Monte Carlo methods, we clearly need some way to control the probability of transitions from a given state.
-In pursuit of this control, we employ the method of Markov chains, specifically Metropolis-Hastings Monte Carlo.
+In pursuit of this control, we employ the method of Markov chains, specifically Broad histogram transition matrix Monte Carlo.
 \subsubsection{Markov Chain}
-\ignore{[I think this is framed incorrectly... markov chains are used to generate the transition matrix, no? That's the procedure outlined by de Oleveira, reinforced by michael...]}
-Markov chain methods eschew the true randomness of regular Monte Carlo by taking the current configuration into account before moving.
+
+Markov chain Monte Carlo methods eschew the true randomness of traditional Monte Carlo by taking the current configuration into account before moving.
 If we define the initial state of the system by a vector column $\mathbf{q}$ (where the components refer to general coordinates, typically taken to be position of each sphere), and the resulting state $\mathbf{p}$ after a single time step, the relationship
@@
 defines a ``transition'' of the system.
-$\mathcal{P}$ is a necessarily square matrix (as long as the number of coordinates remain fixed) that contains information about these transitions.
+$\mathcal{P}$ is a necessarily square matrix that contains information about these transitions.
 We call this construction $\mathcal{P}$ the {\it transition matrix}.
-The transition matrix's $i^{\text{th}}$ eigenvalue the probability of the $i^{\text{th}}$ state occurring which, for the canonical ensemble.
+This matrix is $L\times L$, where $L$ is the number of energy levels in system.
+The transition matrix's $i^{\text{th}}$ eigenvalue the probability of the $i^{\text{th}}$ state occurring in the system.
 If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
@@
 \end{align}
-The transition matrix thus embodies the physics of the simulation. [more?]
+The transition matrix thus embodies the physics of the simulation.
 
@@
 Further, this reweighting detracts from the accuracy of the tails of the histogram.
-We choose instead to implement a similar approach to that of P.M.C.
-de Oleveira and coworkers\cite{de1998broad}, the ``broad histogram''.
-[more] This process is outlined in Algorithm~\ref{alg_tmmc}.\\
+We choose instead to implement a similar approach to that of de Oleveira and coworkers\cite{de1998broad}, the ``broad histogram''.
+\\
 
-[stuff: markovian walk along energies known a priori, accept state always if $\Delta E< 0$, potentially accept state if\dots what?
-We can't directly enumerate states.
-Must be in Michael thesis somewhere.
-Check wang-landau paper?
-or [8] on michael thesis]
+To construct the broad histogram we need to perform a Markovian walk along the energy levels of the system.
+Throughout the walk we count the number of times we transitioned from a given initial state to to a given resulting state.
+If we walk across the entire energy range, we can construct histogram of energy transitions $h(E_i, E_j)$ for all $i, j$.
+The walk goes as follows: starting with a state $\mathbf{q_i}$ with energy $E_i$, we attempt perform a random move to a state $\mathbf{q_f}$ with energy $E_f$.
+Assuming the resulting configuration is valid, the move is always accepted if the energy decreases, i.e, $E_f - E_i <0$.
+Otherwise, we accept the move with a probability related to the ratio of counts of the transition occurring ``forwards'' to the transition occurring the opposite direction.
+Since each state is reachable within a finite amount of time in a Markovian process, a broader, more accurate histogram merely requires more computation time.
+This process is outlined in Algorithm~\ref{alg_tmmc}.\\
 
 \begin{algorithm}[tb]
-    \caption{Transition Matrix Monte Carlo}
+    \caption{Transition Matrix Monte Carlo Initialization}
     \label{alg_tmmc}
@@
     \begin{enumerate}
-        \item hi
+        \item Begin state $\mathbf{q_0}$ and a zero transition matrix.
+        \item Perform a random walk, transitioning from an initial state $\mathbf{q_i}$ to a final state $\mathbf{q_{f}}$.
+Initially accept the state if the configuration is valid.
+\label{alg-tmmc-start}
+        \item Increment the transition histogram $h(E_i, E_f)$.
+        \item Accept move unconditionally if $\Delta E_f - E_i = E < 0$.
+        \item Else, accept move with a probability proportional to $\frac{h_(E_i, E_f )}{h_(E_f, E_i)}$.
+        \item If not end condition has been met, return to step \ref{alg-tmmc-start}.   
     \end{enumerate}
@@
 
-Calculating the density of states, given the transition matrix constructed via the markov chain process, is a trivial matter.
+Calculating the density of states, given the transition matrix constructed via the process in Algorithm~\ref{alg_tmmc}, is a trivial matter.
 The equilibrium condition given in Eq.~\ref{thermo-steady-state} is really a statement about the density of states; for a Monte Carlo simulation, the equilibrium state $q_{eq}$ is exactly the density of states.
 Solving the steady state condition as an eigenvalue problem for $q_{eq}$ then yields the density of states.
+It bares mentioning that most implementations of transition matrix methods consider the above procedure the ``initialization'' process.
+The canonical ``simulation'' period refers to gathering statistics for quantities that cannot be calculated from the density of states, such as the density of the fluid or the pair distribution function.
+Given that we only are concerned with quantities that can be computed given the density of states, we never actually ``simulate''.
+With the density of states computed, we can now calculate the partition function and thereby the free energy.
 

### 2016-05-06 papers/thesis-vischer/Thesis.tex
More cleanup, more plot discussion. Tests skipped!

@@
 
-Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success [cite me], particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
+Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success \cite{white2001global}, particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
 GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid [!fixme not quite].\\
@@
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
-\\
+
+\subsubsection{Carnahan-Starling Equation of State}
+In order to determine the filling fraction step size during thermodynamic integration we must find changes in filling fraction that yield a suitably-sized change in the system's absolute free energy.
+To do this we use the Carnahan-Starling equation of state, shown in Eq.~\ref{C-S-eq}.
+\begin{align}
+    F(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2}   
+    \label{C-S-eq}  
+\end{align} 
+where $\eta$ is the filling fraction.
+The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit.
+We are very much concerned with cells significantly far from the thermodynamic limit.
+Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\
 
@@
 Since the drunken man chooses his next move arbitrarily, his movement is considered ``random''; it follows that true random sampling is desirable for this physical system.
+\\
+
 Statistical mechanics, however, does not exhibit this independence.
@@
 The equilibrium condition given in Eq.~\ref{thermo-steady-state} is really a statement about the density of states; for a Monte Carlo simulation, the equilibrium state $q_{eq}$ is exactly the density of states.
-Solving the steady state condition as an eigenvalue problem for $q_{eq}$ then yields the density of states.
+Solving the steady state condition as an eigenvalue problem for $q_{eq}$ then yields the density of states\cite{perlinthesis}.
+\\
+
 It bares mentioning that most implementations of transition matrix methods consider the above procedure the ``initialization'' process.
@@
 
-\subsubsection{Carnahan-Starling Equation of State}
-In order to determine the filling fraction step size during thermodynamic integration we must find changes in filling fraction that yield a suitably-sized change in the system's absolute free energy.
-To do this we use the Carnahan-Starling equation of state, shown in Eq.~\ref{C-S-eq}.
-\begin{align}
-    F(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2}   
-    \label{C-S-eq}  
-\end{align} 
-where $\eta$ is the filling fraction.
-The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit.
-We are very much concerned with cells significantly far from the thermodynamic limit.
-Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\
 Combining Eq.~\ref{C-S-eq} with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy.
@@
 
-Since the excess hard sphere free energy is really just the entropy[cite] (scaled by temperature), the statement that the square well and hard sphere excess free energies need be equal in the infinite temperature limit is really a constraint on the {\it entropy} of the two systems.
+Since the excess hard sphere free energy is really just the entropy \cite{valeskethesis} (scaled by temperature), the statement that the square well and hard sphere excess free energies need be equal in the infinite temperature limit is really a constraint on the {\it entropy} of the two systems.
 We then conclude that
@@
 \end{align*} 
-The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration and both the partition functions are calculated the usual way [MORE ABOUT $Z_inf$].
+The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration, $Z$ is calculated using  .
 Once we have calculated both the excess free energy and excess internal energy, we can directly find the excess entropy:
@@
 Liquid and vapor are simultaneously present in the fluid at densities that cause local minima in the grand free energy.
+We know that the grand free energy has a guaranteed minimum at the origin because [things]; we then only need to identify the other minima.
+\\
+
 Referring back to Eq.~\ref{Phidef}, we can compute an absolute grand free energy given the absolute free energy and number of spheres.
 The chemical potential $\mu$ is a somewhat unintuitive quantity that, for our purposes, we will treat as a free parameter.
-Tweaking the value of the chemical potential will allow us to easily identify coexistence in the fluid.
+Tweaking the value of the chemical potential will allow us to easily identify second minima in the density curve.
 \begin{figure}
@@
 
-\ignore{\subsection{Comparison With A Modified Statistical Associating Fluid Theory}
-[more]}
-
+Figure \ref{Phi-eta}, 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \cleardoublepage
-\bibliography{Thesis}
+\bibliography{./Thesis}
 \end{document}

### 2016-05-15 papers/thesis-vischer/Thesis.tex
minor edits - language cleanup, unit fixing, etc. Tests skipped!

@@
 Fluids have long garnered interest\cite{theorysimpleliquids}\cite{theoryofcrits}, particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
-A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor become indistinguishable and, more importantly, large fluctuations of fluid density occur.
-Most traditional fluid models assume the density across the cell to be nearly constant.
+A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor become indistinguishable and, more importantly, fluctuations of fluid density at large length scales occur.
+Most traditional fluid models assume the fluid density does not fluctuate beyond the molecular scale.
 As a result, these methods tend to poorly handle these fluctuations near criticality.
@@
 Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success \cite{white2001global}, particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
-GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid [!fixme not quite].\\
+GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid \ignore{[!fixme not quite]}.\\
 
@@
 These spheres exhibit no attraction, and consequently each carries kinetic energy $K = 3/2 k_b T$.
-Further, with no attraction, there is no energy associated with [more].
+Further, with no attraction, there is excess internal energy, implying the excess free energy is purely entropic.
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
@@
 \begin{align}
-    F(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2}   
+    F_{exc}(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2} N k_b T  
     \label{C-S-eq}  
 \end{align} 
-where $\eta$ is the filling fraction.
+where $\eta$ is the filling fraction, $N/V$.
 The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit.
 We are very much concerned with cells significantly far from the thermodynamic limit.
-Thankfully, we are only concerned with filling fractions slightly or significantly below this value, and so the Carnahan-Starling equation of state is an invaluable tool.\\
+Thankfully, the Carnahan-Starling equation of state is still an excellent approximation at low to moderate filling fractions.\\
 
@@
 We then define dimensionless temperatures $T/\epsilon$, energies $E/\epsilon$, and distances $r/\sigma$.
-Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ (as well as any other physical constants) to one.
+Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ to one.
 This then allows us to reasonably compare energies and temperatures since they have the same units.
@@
 Many interesting effects occur here, including the well-researched phenomena of critical opalescence.
-The correlation length, naively interpreted here as the maximum existing wavelength of the density fluctuations within the fluid, diverges to infinity near criticality.
-As mentioned previously, the diverging of the correlation length invalidates SAFT and many of its variants which assume constant density.
+The correlation length, loosely interpreted here as the maximum wavelength of the density fluctuations within the fluid, diverges to infinity near criticality.
+As mentioned previously, the diverging of the correlation length invalidates SAFT and many of its variants which assume constant density beyond small length scales.
 \begin{figure}
@@
     \includegraphics[width=.75\textwidth]{phase.png}
-    \caption{Phase diagram of water with critical point and triple point identified.}
+    \caption{Phase diagram of a liquid with critical point and triple point identified.}
     \label{phase}
@@
 Non-critical systems exhibit different characteristics when viewed at different length scales.
-Ever-present small-scale fluctuations in density are relevant, if not dominant, when viewing the system at very small scales, but typically become less important at larger scales and nearly irrelevant near the thermodynamic limit.
-Near the critical point all wavelengths of fluctuation are present, and so the system looks the same regardless of length scale chosen.
+Ever-present small-scale correlations (that is, fluctuations in density) are relevant, if not dominant, when viewing the system at very small scales, but typically become less important at larger scales and nearly irrelevant near the thermodynamic limit.
+Near the critical point all correlation wavelengths are present, and so the system looks the same regardless of length scale chosen.
 This self-similarity allows for the decomposition of the large critical cell to smaller constituent cells.
-The details of the procedure are more carefully discussed in Chapter {\color{red} [blah]}.
+The details of the procedure are more thoroughly discussed in Chapter.
 
@@
 \end{align}
-With $F$ calculated, partial derivatives yield all other interesting thermodynamic quantities.
+With $F$ calculated, its partial derivatives yield all other interesting thermodynamic quantities.
 Though the Gibbs free energy yields the same information, the Helmholtz free energy is particularly useful because of the relationship with entropy, namely
@@
 \end{align}
-Further, if we allow the number of particles in the system to vary, we have the grand free energy $\Phi$, defined as
+Entropy is a very similar partial derivative of the Gibbs free energy, but with pressure held fixed instead of volume.
+We typically will want to hold volume fixed, so the Helmholtz free energy is the natural choice.
+Further, if we allow the number of particles in the system to vary, we will use the grand free energy $\Phi$, defined as
 \begin{align}
@@
 where $\beta =  T^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$ state.
-Of course, it is impractical to sum over the individual states of the system because the number of states are generally innumerable.
+Of course, it is impractical to sum over the individual states of the system because the number of states of a classical system are generally innumerable.
 Instead we sum over energies:
@@
 where we have now introduced the density of states $D(E)$ which is a simple function that returns the number of states which have a given energy $E$.
-This introduction is now the focus of calculation - if we can accurately determine the density of states of the system and have knowledge of the available energies, we can calculate the Helmholtz free energy and, from there, any other thermodynamic quantity we wish.
+The density of states is now the focus of calculation - if we can accurately determine the density of states of the system, we can calculate the Helmholtz free energy and, from there, any other thermodynamic quantity we wish.
 \\
+
 Since $F$ contains a natural log term, we can separate and characterize the individual contributions of kinetic energy (``ideal'') and energy associated with the configuration of the system (``excess'').
@@
 
-{\color{red} [This should probably go in an appendix.
+\ignore{{\color{red} [This should probably go in an appendix.
 Further: replace $Z_C$ with $Z_{exc}$; fix the to the N on the $Z_{exc}$ term; include the $N!$ into the ideal term; fix the $V(r_1, r_2, ...)$||| All done!
-NOW: fix the dimensions on the $Z_{id}$ term!]}
+NOW: fix the dimensions on the $Z_{id}$ term!]}}
 
@@
 \begin{align}
-    Z &= \sum_i D(E) e^{-\beta E_i}
+    Z &= \sum_E D(E) e^{-\beta E}
 \end{align}
-where we sum over individual energies of the system $E_i$. [reason that we sum over states] , so instead we sum over the states $s$:
+where we sum over individual energies of the system $E_i$. However, to provide a general expression, we will instead sum over the states $s$:
 \begin{align}
@@
 \end{align}
-noting that $V(\mathbf{r})$ is poorly defined for a single particle, but we digress. [FIXME]. 
+
 So now we have
@@
 Each state is specified by a particular point in phase space (parameterized by $\mathbf{r}$ and $\mathbf{p}$).
-Even though this phase space is discrete, we can approximate the space as continuous by recognizing that the {\color{red}spacing between individual states is small compared to the total range of the states}, so that the summation becomes an integral:
+Even though quantum mechanically this phase space is discrete, we can approximate the space as continuous by recognizing that the spacing between individual states is small compared to the total range of the states, so that the summation becomes an integral:
 
@@
 \begin{align}
-    Z &= \frac{V^N}{V^N N!}\int_s \diff\mathbf{r}^N ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s \diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
+    Z &= \frac{1}{ N!}\int_s \frac{\diff\mathbf{r}^N}{V^N} ~e^{-\beta V(\mathbf{r_1}, \mathbf{r_2}, \dots \mathbf{r_N})} \left[\int_s V\,\diff\mathbf{p}~e^{-\beta K(\mathbf{p})}\right]^{N}
     \label{Z_full}
@@
 
-We then define the two terms obvious in Eq. \ref{Z_full}, the excess term dependent on the potential energy $V(\mathbf{r})$ and the ideal term dependent on the kinetic energy $K(\mathbf{p})$:
+We then define the two factors obvious in Eq. \ref{Z_full}, the excess factor dependent on the potential energy $V(\mathbf{r})$ and the ideal factor dependent on the kinetic energy $K(\mathbf{p})$:
 
@@
 where $Z_{id}$ contains the $N!$ term by convention.
-The excess term $Z_{exc}$ is system dependent because $V(\mathbf{r})$ is buried inside the integral, but the ideal term is the same for any system where the total energy separates in the preceding fashion.
-We judiciously decide to leave the excess term alone and work specifically with the ideal term.
+The excess term $Z_{exc}$ is system dependent because $V(\mathbf{r})$ is buried inside the integral, but the ideal factor is the same for any system where the total energy separates in the preceding fashion.
+We judiciously decide to leave the excess factor alone, for now, and work specifically with the ideal term.
 We then have
@@
 \end{align}
-It should be noted that this definition of $k$ introduces three-fold degeneracy.\\
 Remembering that kinetic energy can be expressed as $K = \frac{\hbar^2 k^2}{2m}$, and using our definition of $\mathbf{k}$:
 \begin{align}
-    Z_{id} &= \frac{1}{N!}\left[\int_s \diff\mathbf{p} ~e^{-\beta \frac{\hbar^2 k^2}{2m}}\right]^{N}
+    Z_{id} &= \frac{1}{N!}\left[\int_s V\, \diff\mathbf{p} ~e^{-\beta \frac{\hbar^2 k^2}{2m}}\right]^{N}
     \label{Zt_def}
 \end{align}
-This unfortunately introduces a mass quantity into the partition function which adds an additional parameter that needs to be determined. [more] 
+This unfortunately introduces a mass quantity into the partition function which adds an additional parameter that needs to be determined to properly account for quantum mechanical effects.
 If we then allow the cell to become very large, then the we can easily see in Eq.
@@
 \end{align}
-Evaluating this integral yields
+where extra factors of $\hbar $and$ \pi$ were gained from the change of variables. Evaluating this integral yields
 \begin{align}
@@
 
-Monte Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
+Unbiased Monte Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
 Traditional random sampling requires that the current step, or ``move'', is irrespective of previous steps.
@@
 \subsubsection{Markov Chain}
-
-Markov chain Monte Carlo methods eschew the true randomness of traditional Monte Carlo by taking the current configuration into account before moving.
-If we define the initial state of the system by a vector column $\mathbf{q}$ (where the components refer to general coordinates, typically taken to be position of each sphere), and the resulting state $\mathbf{p}$ after a single time step, the relationship
+Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving.
+If we define the equilibrium state of the system by a vector column $\mathbf{q}$ (where the components refer to general coordinates, typically taken to be position of each sphere), and the resulting state $\mathbf{p}$ after a single time step, the relationship
 \begin{align*}
@@
 This matrix is $L\times L$, where $L$ is the number of energy levels in system.
-The transition matrix's $i^{\text{th}}$ eigenvalue the probability of the $i^{\text{th}}$ state occurring in the system.
 If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
@@
 We chose to focus on the scrunched case for two reasons.
-Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length [necessary?]
-
-\ignore{ Might never need these sections. |
-\subsubsection{Seed Disparity}
-Because of the random/psuedorandom nature of Monte Carlo sampling, we tested over many different initial sphere placements (``seeds'') to see if there was consistent agreement. 
-\subsection{Other Cells?}
-In addition to the ``scrunched'' primitive cell, we tested [hopefully one or two] other minimum correlation lengths corresponding to other natural parameters in the system.}
+Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length.
 
@@
 
-Figure \ref{Phi-eta}, 
+Figure \ref{Phi-eta}, \dots.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 The application presented was a system that obeys the canonical ensemble, but the extension to the grand canonical ensemble can be done with little modification.
-Future work can include stuff [more].
+Future work can include investing further doubling regimes ($i\geq 3$).
 %\bibliographystyle{plain}

### 2016-05-20 papers/thesis-vischer/Thesis.tex
Corrected science and language throughout Tests skipped!

@@
 \end{align}
-For liquids [fluids?], the $Z_{exc}$ term is difficult to ever evaluate; the hard sphere liquid, the simplest model for a dense fluid, does not have an analytic representation for the excess free energy.
+For liquids, the $Z_{exc}$ factor is difficult to evaluate; the hard sphere liquid, the simplest model for a dense fluid, does not have an analytic representation for the excess free energy.
 It follows that a considerable amount of work is required to determine the excess free energy for a given system.
@@
 \end{align}
-Redefining the term including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at
+Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at
 \begin{align}
-    F_{abs} &= F_{exc} - NT\ln\left(\frac{V}{\Lambda^3}\right) + NT\ln\left(N\right)
+    F_{abs} &= F_{exc} - NT\left(\ln\left(\frac{V}{\Lambda^3}\right) - \ln\left(N\right)\right)
 \end{align}
-At a specified volume and number of particles, the ideal gas contribution is constant. Consequently, for a given temperature we only need to compute the excess free energy.
+At a specified volume and number of particles, the ideal gas contribution is only a function of temperature. Consequently, for a given temperature we only need to compute the excess free energy.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 A preliminary value of the free energy of the bulk can be obtained from Monte Carlo simulations.
-This value can then be made absolute by carefully comparing the results with Thermodynamic Integration, a method of determining the  free energy of the hard-sphere system.
+This value can then be made absolute by carefully comparing the results with thermodynamic integration, a method of determining the free energy of the hard-sphere system.
 The resulting absolute free energies for each fluctuation length can be summed to yield the absolute free energy of the bulk.
-[more]
+{\color{red}[deal with the problems here]}
 
@@
 In pursuit of this control, we employ the method of Markov chains, specifically Broad histogram transition matrix Monte Carlo.
-\subsubsection{Markov Chain}
+\subsubsection{Metropolis-Hastings}
 Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving.
-If we define the equilibrium state of the system by a vector column $\mathbf{q}$ (where the components refer to general coordinates, typically taken to be position of each sphere), and the resulting state $\mathbf{p}$ after a single time step, the relationship
+If we define the ensemble of states of the system in equilibrium by a vector column $\mathbf{q}$, and the resulting state $\mathbf{p}$ after a single time step, the relationship
 \begin{align*}
@@
 This matrix is $L\times L$, where $L$ is the number of energy levels in system.
-If all elements of the transition matrix are nonzero, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
-A physically judicious choice of transition matrix is one that satisfies the thermodynamic steady state condition; given the equilibrium state $\mathbf{q}_{eq}$, we enforce that
+If all rows of the transition matrix are , we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
+A physically correct simulation will generate a transition matrix that satisfies the thermodynamic steady state condition; given the ensemble of equilibrium states $\mathbf{q}_{eq}$, we enforce that
 \begin{align}
@@
 If the energy of the system at each time step in a Monte Carlo simulation is recorded, it is straightforward to construct a histogram of energies.
-That is, a histogram with a bin for each narrow $\Delta E$ around a value $E$ that displays the number of times the system visited that energy range.
+That is, a histogram with a bin for each narrow $\Delta E$ around a value $E$ that counts the number of times the system visited that energy range.
 Typical implementations of histogram methods follow the ``canonical histogram'' procedure outlined by Ferrenberg and Swendsen \cite{ferrenberg1988histogram} -- given a histogram $h_T(E)$ extracted from a simulation at some temperature $T_0$, a reweighting scheme can then extend the histogram to any other temperature $T$ via
@@
 To construct the broad histogram we need to perform a Markovian walk along the energy levels of the system.
-Throughout the walk we count the number of times we transitioned from a given initial state to to a given resulting state.
+Throughout the walk we count the number of times we transitioned from a given initial energy to to a given resulting energy.
 If we walk across the entire energy range, we can construct histogram of energy transitions $h(E_i, E_j)$ for all $i, j$.
@@
         \item Accept move unconditionally if $\Delta E_f - E_i = E < 0$.
-        \item Else, accept move with a probability proportional to $\frac{h_(E_i, E_f )}{h_(E_f, E_i)}$.
+        \item Else, accept move with a probability proportional to $\frac{h(E_i, E_f )}{h(E_f, E_i)}$.
         \item If not end condition has been met, return to step \ref{alg-tmmc-start}.   
@@
 
-It bares mentioning that most implementations of transition matrix methods consider the above procedure the ``initialization'' process.
+It bears mentioning that most implementations of transition matrix methods consider the above procedure the ``initialization'' process.
 The canonical ``simulation'' period refers to gathering statistics for quantities that cannot be calculated from the density of states, such as the density of the fluid or the pair distribution function.
@@
 We begin with a periodic cubic cell of length $L_0$ and volume $L_0^3$. 
-This cell has periodic boundary conditions and consequently allows for fluctuations up to wavelength $2L_0$.
+This cell has periodic boundary conditions and consequently allows for fluctuations up to wavelength $L_0$.
 These fluctuations contribute some absolute free energy $F_0$ to the absolute free energy of the full cell.
-We then double the length of the cell to allow for fluctuations of up to wavelength $4L_0$.
+We then double the length of the cell to allow for fluctuations of up to wavelength $L_0$.
 We continue this process until all relevant wavelengths have been included.
@@
 Strictly speaking, a formal limit is required for equality in Eq.~\ref{GRGT-F1}, rather than only including wavelengths up to $2^k\lambda_0$.
-Though theoretically all wavelengths are present in the critical system we posit that it is only necessary to include the first several.
+Though theoretically all wavelengths are present at the critical point it is only necessary to include the first several to accurately describe systems moderately close to the critical point.
 Given approximation, we choose instead to use Eq.~\ref{GRGT-F2}.
-Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely TMMC and Thermodynamic Integration.
+Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely Metropolis-Hastings Monte Carlo and thermodynamic integration.
 A graphical representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
@@
 \subsubsection{General Process}
-Thermodynamic Integration is a unique way to determine the ``excess'' free energy of the system.
-Using hard sphere Monte Carlo, we transition from a filling fraction of zero (note that this statement is identical to assuming that the cell begins at infinite temperature, since the total energy does not depend on where the spheres are placed) to a near-close-packed system with filling fraction $\eta_max$, calculating the change in free energy at each step.
+Thermodynamic integration is a unique way to determine the ``excess'' free energy of the system.
+Using hard sphere Monte Carlo, we transition from a filling fraction of zero to a dense system with filling fraction $\eta_{max}$, calculating the change in free energy at each step.
 When we have reached the desired final filling fraction we have found the free energy that was necessary to create that configuration of particles - this is exactly the absolute free energy of the system.\\
@@
 \end{align*}
-Each sphere count subfolder contains a single text file from TMMC output and another subfolder where all of the Thermodynamic Integration output is stored.
-Each Thermodynamic Integration simulation outputs a single text file upon completion of each filling fraction step.
-On average, a single TI calculation requires approximately $175$ steps.
-\\
+Each sphere count subfolder contains a single text file from TMMC output and another subfolder where all of the thermodynamic integration output is stored.
+Each thermodynamic integration simulation outputs a single text file upon completion of each filling fraction step.
+On average, a single calculation requires approximately $175$ steps.
 This structure was chosen both for elegant parallelism and for plotting accessibility.
@@
 \subsection{Calculating the Absolute Free Energy}
+{\color{red} please rewrite me. This paragraph has been terrible since it was written.}\\
 It is straightforward the compute a Helmholtz free energy through the Monte Carlo simulations; the difficulty in this method lies in computing the correct free energy.
@@
 At infinite temperature, the square well fluid has approximately the same behavior as the hard sphere fluid.
-We assert that the correct inclusion, then, is to enforce that the excess free energy calculated via TMMC at $T=\infty$ is equal to the free energy calculated through thermodynamic integration.
+We assert that the correct inclusion, then, is to enforce that the excess entropy calculated via TMMC at $T=\infty$ is equal to the hard sphere entropy calculated through thermodynamic integration.
 We do this by exploiting the infinite temperature behavior of the canonical ensemble:
 \begin{align}
-    Z_{inf} &= \left.\sum_{i=0}^{\text{\tiny Energies}} D(E) e^{-\beta E}\right|_{\beta = 0}\\
+    Z_{\infty} &= \left.\sum_{i=0}^{\text{\tiny Energies}} D(E) e^{-\beta E}\right|_{\beta = 0}\\
     &= \sum_{i=0}^{\text{\tiny Energies}} D(E)\\
@@
 \end{align*} 
-and, at infinite temperature, the internal energy vanishes, leaving us with
+and, at infinite temperature, the internal energy asymptotes to a value independent of temperature, leaving us with
 \begin{align}
-    F_{SW,~exc} &= -T\,S_{SW, ~exc} \;.
+    F_{SW,~exc} &= -T\,S_{SW, ~exc} + U_{exc}\;.
 \end{align}
@@
 \begin{align*}
-     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E)] - \beta E_i})\\
+     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E)] - \beta E_i)}\\
      &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i \exp\left(-\beta E_i\right)}{Z}
@@
 where, as usual, $\beta = T^{-1}$.
-Taking the exponential of $\ln(D(E))$, unfortunately, tends to introduce overflow or underflow errors during computation.
+unfortunately, tends to introduce overflow or underflow errors during computation.
 We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots}from each element of $D(E)$, and then adding the maximum back in after the exponential.
@@
 Using the coupling procedure outlined previously, we calculated changes in free energy up to the $n^{\text{th}}$ doubling regime for several different cell lengths.
-We are primarily concerned with the ``scrunched'' case where the cell length is $\sqrt2\sigma$.\\
+We choose to focus on the ``scrunched'' case where the cell length is $\sqrt2\sigma$.\\
 

### 2016-05-20 papers/thesis-vischer/Thesis.tex
edits from meeting with Brenden

@@
 The partition function now calculates the total number of states in the system which we expected due to the equipartition theorem. The excess free energy is simply
-\begin{align*}
+\begin{align}
     F_{SW,~exc} &= U_{SW,~exc} - T\,S_{SW,~exc}     
-\end{align*} 
+\end{align} 
 and, at infinite temperature, the internal energy asymptotes to a value independent of temperature, leaving us with
@@
 The hard sphere system has no excess internal energy, so the excess free energy is merely
-\begin{align*}
+\begin{align}
      F_{HS,~exc} &= -T\,S_{HS,~exc} 
-\end{align*}
+\end{align}
 
@@
 We then conclude that
-\begin{align*}
+\begin{align}
     S_{SW,~exc} \Bigm\lvert_{T=\infty} &= S_{HS,~exc}
-\end{align*}
+\end{align}
 
@@
 Given the density of states output $\ln[D(E)]$ from the Monte Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic , for instance, the excess internal energy:
-\begin{align*}
-     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp\left(\ln[D(E)] - \beta E_i\right)}{\sum_{i=0}^{\text{\tiny Energies}} \exp[\ln(D(E)] - \beta E_i)}\\
+\begin{align}
+     U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp(\ln[D(E)] - \beta E_i)}{\sum_{i=0}^{\text{\tiny Energies}} \exp(\ln[D(E)] - \beta E_i)}\\
      &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i \exp\left(-\beta E_i\right)}{Z}
-\end{align*} 
+\end{align} 
 where, as usual, $\beta = T^{-1}$.
@@
 We calculate the excess free energy by using the entropy equality developed previously:
-\begin{align*}
+\begin{align}
      S_{SW,~exc}\Bigm\lvert_{T=\infty} &= S_{HS,~exc}\\
-     \implies F_{SW,~exc} &= -TS_{HS,~exc} + F_{inf}\\
-     F_{SW,~exc} &= -TS_{HS,~exc} - T\ln\left(\frac{Z}{Z_{inf}}\right)
-\end{align*} 
+     \implies F_{SW,~exc}\Bigm\lvert_{T=\infty} &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} \\
+     \implies F_{SW,~exc}(T) &= F_{SW,~exc}\Bigm\lvert_{T=\infty} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
+     \\
+     \implies F_{SW,~exc}(T) &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
+     \\
+     F_{SW,~exc} &= U_{\infty} -TS_{HS,~exc} - T\ln\left(\frac{Z}{Z_{\infty}}\right)
+\end{align} 
 The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration, $Z$ is calculated using  .
 Once we have calculated both the excess free energy and excess internal energy, we can directly find the excess entropy:
-\begin{align*}
+\begin{align}
     S_{SW, ~exc} &= \frac{1}{T}\left(U_{SW,~exc} - F_{SW,~exc} \right)
-\end{align*}
+\end{align}
 

### 2016-05-21 papers/thesis-vischer/Thesis.tex
add note about Stirling

@@
 \end{align}
-Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at
+Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at \textcolor{red}{BRENDEN FIX: $N\ln N - N$, plus I think we may want to add to the excess free energy (probably to the excess HS entropy?) the difference?}
 \begin{align}

### 2016-05-27 papers/thesis-vischer/Thesis.tex
cleanup all around, fixes, slight content expansion for results and TI Tests skipped!

--- 2016-05-20 papers/thesis-vischer/Thesis.tex
@@
 \subsection{Hard Sphere}
-The hard sphere model approximates monoatomic fluids as having no interactions other than collisions.
+The hard sphere model approximates fluids as having no interactions other than collisions.
 These spheres exhibit no attraction, and consequently each carries kinetic energy $K = 3/2 k_b T$.
-Further, with no attraction, there is excess internal energy, implying the excess free energy is purely entropic.
+Further, with no attraction, there is no excess internal energy, implying the excess free energy is purely entropic.
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
@@
 \begin{align}
-    F_{exc}(\eta) &= \frac{4\eta - 3\eta^2}{(1-\eta)^2} N k_b T  
+    F_{exc}(\eta) &= N k_B T\,\frac{4\eta - 3\eta^2}{(1-\eta)^2}   
     \label{C-S-eq}  
@@
 
-The potential energy of the fluid is then always an integer multiple of the well depth $\epsilon$.
 This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
@@
     F_{abs} &= -T\,\ln\left(Z_{exc} \frac{V^N}{N!\Lambda^{3N}}\right)\\
-    &= -T\ln\left(Z_{exc}\right)\, -T\left(\frac{V}{(N!)^{1/N}\Lambda^3} \right)^N\\
+    &= -T\ln\left(Z_{exc}\right)\, -T\ln\left(\frac{V}{(N!)^{1/N}\Lambda^3} \right)^N\\
     &= -T\ln\left(Z_{exc}\right) - NT\ln\left(\frac{V}{\Lambda^3}\right) + NT\ln\left(N!^{1/N}\right)\\ 
 \end{align}
-Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N$ for large N), we arrive at
+Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N - N$ for large N), we arrive at \textcolor{red}{BRENDEN FIX: $N\ln N - N$, plus I think we may want to add to the excess free energy (probably to the excess HS entropy?) the difference?}
 \begin{align}
-    F_{abs} &= F_{exc} - NT\left(\ln\left(\frac{V}{\Lambda^3}\right) - \ln\left(N\right)\right)
+    F_{abs} &= F_{exc} - NT\left(\ln\left(\frac{V}{\Lambda^3}\right) - \ln\left(N\right) + 1\right)
 \end{align}
@@
 Instead of true random Monte Carlo methods, we clearly need some way to control the probability of transitions from a given state.
-In pursuit of this control, we employ the method of Markov chains, specifically Broad histogram transition matrix Monte Carlo.
+In pursuit of this control, we employ the method of Markov chains, specifically broad histogram transition matrix Monte Carlo.
 \subsubsection{Metropolis-Hastings}
@@
 Since we are calculating the change in free energy as we transition to higher and higher filling fractions, we need only be concerned with finding the ratio of partition functions between each step.
-Since the hard sphere partition function is a ``counting'' mechanism, we can determine this ratio quite easily using Monte Carlo simulations.
-Given a starting state (defined by the arrangement of particles and filling fraction) we shrink the volume and intermolecular distances by a somewhat arbitrary scaling factor and then see if the resulting configuration of particles is valid.
-If the configuration is valid, we accept this is a working move and then return to the original configuration and test another scaling factor.
-If the resulting configuration is invalid, we count the move as failed and then repeat.
-The ``golden'' acceptance ratio is fifty percent - when we have found a scaling factor that yields a ratio near the desired value\dots\\
+Since the hard sphere partition function is a ``counting'' mechanism, we can determine this ratio using Monte Carlo simulations.
+Given a starting state (defined by the arrangement of particles and filling fraction) we shrink the volume and intermolecular distances by a scaling factor and then see if the resulting configuration of particles is valid.
+If the configuration is valid, we accept it as also valid in the small cell and then return to the original configuration.
+If the resulting configuration is invalid, we count the move as failed.
+This process continues until a predefined number of iterations are completed.\\
 
-As with the Monte Carlo these simulations are written in $C++$ for efficiency.
+Once a count of valid and failed moves for a given filling fraction step has been established, the entropy of the system can be determined.
+Give $N_i$ total checks of the small cell during the $i^{\text{th}}$ step, $M_i$ of which were valid, the entropy calculated integrating from a filling fraction of $\eta=0$ to some $\eta_f$ in $K$ steps is
+\begin{align}
+    S_{HS,~exc} &= \sum_i^K \ln\left(\frac{M_i}{N_i-M_i}\right)
+\end{align}
 
@@
 \item Move spheres random fractions of a predefined translation distance. \label{alg-ti-move}
-\item Repeat step \ref{alg-ti-move} for [desired number of iterations before small cell check].
-\item After [desired number of iterations], check the ``small cell'' by shrinking the volume and checking whether or not the configuration is still valid.
-\item If configuration is still valid, valid checks ++.
-\item If configuration is invalid, failed checks ++. \label{alg-ti-move-end}
+\item Repeat step \ref{alg-ti-move} for desired number of iterations before checking the small cell.
+\item Check the ``small cell'' by shrinking the volume and determine whether or not the configuration is still valid.
+\item If configuration is still valid, increment the count of valid checks.
+\item If configuration is invalid, increment the count of failed checks. \label{alg-ti-move-end}
 \item Repeat steps \ref{alg-ti-move} - \ref{alg-ti-move-end} until desired iterations reached.
@@
 
-\ignore{Should I detail the algorithm itself and string simulations together or consider the entire process $\eta=0\rightarrow\eta_max$ to be the algorithm?}
-
-\ignore{is the distinction between infinite case and regular necessary? we always use infinite...}
-
+%TIKZ
 \begin{figure} \centering
@@
     \caption{Demonstration of filling fraction increase during thermodynamic integration from $\eta=0$ to $\eta=\eta_f=\eta_{max}$.}
+    \label{TI-demo}
 \end{figure}
@@
 Figures ~\ref{F-T} through ~\ref{F-eta} show several computed thermodynamic quantities for the scrunched system.
-Each plot contains calculations from the first doubling regime ($L=2\lambda_0$) represented by solid lines and second doubling regime ($L = 4\lambda_0$) in dashes.
+Each plot contains calculated quantities for either the first two doubling regimes or the first three.
+Calculations from the first doubling regime ($L=\lambda_0$) are solid lines, with the second doubling regime ($L = 2\lambda_0$) in dashes and the third ($L = 4\lambda_0$), when present, in circles .
 It is worth noting that the succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Fexc-vs-T.pdf}
-    \caption{Excess specific free energy as a function of temperature with each quantity scaled by well depth $\epsilon$. }
+    \caption{Excess specific free energy as a function of temperature.}
     \label{F-T}
@@
 Liquid and vapor are simultaneously present in the fluid at densities that cause local minima in the grand free energy.
-We know that the grand free energy has a guaranteed minimum at the origin because [things]; we then only need to identify the other minima.
+We know that the grand free energy has a guaranteed minimum at the origin because [things]; we then only need to identify the other minimum.
 \\
@@
 The chemical potential $\mu$ is a somewhat unintuitive quantity that, for our purposes, we will treat as a free parameter.
-Tweaking the value of the chemical potential will allow us to easily identify second minima in the density curve.
+Tweaking the value of the chemical potential will allow us to easily identify second minima in the grand free energy, and thereby the densities of the liquid and vapor present in the system.
 \begin{figure}
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Phiabs-vs-eta.pdf}
-    \caption{Absolute specific free energy as a function of density.}
+    \caption{Absolute grand free energy as a function of density.}
     \label{Phi-eta}
 \end{figure}
-
-Figure \ref{Phi-eta}, \dots.
+Figure \ref{Phi-eta} 
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-05-27 papers/thesis-vischer/abstract.tex
cleanup all around, fixes, slight content expansion for results and TI Tests skipped!

--- 2016-04-18 papers/thesis-vischer/abstract.tex
@@
     \Large
-    \textbf{Application of Generalized Renormalization Group Methods to the Square-Well Liquid Near Criticality}
+    \textbf{Free Energy Decomposition of the Critical Square-Well Liquid Using a Renormalization Group Method}
 
@@
 
-You take a box, stuff it with spheres, move them around, make it bigger, do it again. Win grant.
+Analytic models using renormalization methods to approximate the free energy of fluids have been developed, but there is currently no way to investigate the intermediate steps of the approximation.
+We present a computational method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Metropolis-Hastings Monte-Carlo simulations.
+We further detail a method by which to compute the ``absolute'' free energy, including both the ideal gas free energy and the excess free energy.
+We further present an application of this method to the well-research square-well liquid near the critical region.
+We find that the square-well liquid is well-characterized by a base cell length of $\sqrt2\sigma$, the side length of an FCC lattice, by establishing liquid-vapor coexistence between scaled temperatures $0.5 \leq T < .75$.
 \clearpage

### 2016-05-31 papers/thesis-vischer/Thesis.tex
Most of thesis work for nearly final draft. Small edits to results tomorrow morning. Tests skipped!

@@
 % Commands go here
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%                                 TO DO                               %        
+% Add to citations: C-S paper, GNI Clark water (wertheim), Metropolis,%
+% maybe others?                                                       %
+% Add to content: more on TI, move RG to critical point discussion(RIP%
+% content), more results discussion, potentially. More information    %
+% on grand free calculation and results                               %
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-We present here a method to calculate free energy of a liquid that largely avoids this fitting problem.
+We present here a method to calculate the free energy of a liquid that largely avoids this fitting problem.
 We exploit the periodicity of Monte Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
-We then carefully couple simulations of different ``doubling regimes'', each of which considers a different maximum wavelength of fluctuations, where the fluid density $\eta$ remains fixed but the volume $V$ and number of particles $N$ have increased.
-Summing the absolute Helmholtz free energy $F^{\text{abs}}$ of each doubling regime will then approximate the absolute free energy of the full cell, especially near the critical point.
-We will begin by outlining the necessary models and assumptions/approximations for this work.
-We will then detail the algorithms used to compute the quantities with the the discussion on results following.
+We then perform simulations of different doubling regimes, each of which considers a different maximum wavelength of fluctuations, where the fluid density $\eta$ remains fixed but the volume $V$ and number of particles $N$ have increased.
+We can calculate a so-called``absolute'' free energy, including the ideal and excess free energy contributions, for each doubling regime by using two Monte Carlo methods in concert: Metropolis-Hastings Monte Carlo, and thermodynamic integration.
+Summing the difference in absolute free energy of each doubling regime will then approximate the absolute free energy of the full cell, especially near the critical point.
+\\
 
-\subsection{Hard Sphere}
-The hard sphere model approximates fluids as having no interactions other than collisions.
-These spheres exhibit no attraction, and consequently each carries kinetic energy $K = 3/2 k_b T$.
-Further, with no attraction, there is no excess internal energy, implying the excess free energy is purely entropic.
-The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
-
-\subsubsection{Carnahan-Starling Equation of State}
-In order to determine the filling fraction step size during thermodynamic integration we must find changes in filling fraction that yield a suitably-sized change in the system's absolute free energy.
-To do this we use the Carnahan-Starling equation of state, shown in Eq.~\ref{C-S-eq}.
-\begin{align}
-    F_{exc}(\eta) &= N k_B T\,\frac{4\eta - 3\eta^2}{(1-\eta)^2}   
-    \label{C-S-eq}  
-\end{align} 
-where $\eta$ is the filling fraction, $N/V$.
-The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit.
-We are very much concerned with cells significantly far from the thermodynamic limit.
-Thankfully, the Carnahan-Starling equation of state is still an excellent approximation at low to moderate filling fractions.\\
-
-\ignore{[move me | For simulation purposes, the system is required to have periodic boundary conditions - that is, when a sphere would classically collide with the wall of the cell it instead breaches the wall and then reappears on the other side of the cell with the same velocity.]}
-
-\subsection{Square-Well Liquid}
-The square-well liquid is a common approximation that sees extensive use throughout computational fluid research.
-This simple fluid model consists of hard spheres that are subject to an intermolecular pair potential - that is, an attractive force between spheres.
-This model consists of several key parameters: sphere size $\sigma$, well width $\lambda$, and well depth $\epsilon$.
-\\
-As a hard-sphere model, the spheres are not allowed to overlap.
-Any arrangement of spheres - hereafter referred to as the ``state'' of the system - that involves either two or more spheres overlapping or a sphere outside the allowed area - ``cell'' - is defined to be invalid.
-Modifying the radius of the spheres changes the allowed locations in the cell, which in turn affects the partition function of the system.
-The ``square'' refers to the piecewise intermolecular pair potential $\Phi_{12}$, shown mathematically in Eq.
-\ref{phi12} and graphically in Figure ~\ref{sw_phi}.
-
-\begin{equation} 
-\Phi_{12} = \begin{cases}\infty & |\vec{r}|\leq \sigma\\ -\epsilon & \sigma \leq |\vec{r}| \leq \sigma\lambda\\ 0 & |\vec{r}| > \sigma\lambda \end{cases}
-\label{phi12}
-\end{equation}
-
-This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
-We then define dimensionless temperatures $T/\epsilon$, energies $E/\epsilon$, and distances $r/\sigma$.
-Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ to one.
-This then allows us to reasonably compare energies and temperatures since they have the same units.
-
-\begin{figure}
-    \centering
-    \includegraphics[width=.75\textwidth]{SWF-E.png}
-    \caption{Demonstration of Square-Well approximation.}
-    \label{sw_phi}
-\end{figure}
-
-\subsection{Critical Point} 
-The critical point is the temperature and pressure (or any other thermodynamic pairing) on a phase diagram (shown in Figure ~\ref{phase}) where the line between liquid and vapor terminates - liquid and vapor become indistinguishable and are assimilated into the ``supercritical fluid'' phase.
-Many interesting effects occur here, including the well-researched phenomena of critical opalescence.
-The correlation length, loosely interpreted here as the maximum wavelength of the density fluctuations within the fluid, diverges to infinity near criticality.
-As mentioned previously, the diverging of the correlation length invalidates SAFT and many of its variants which assume constant density beyond small length scales.
-\begin{figure}
-    \centering
-    \includegraphics[width=.75\textwidth]{phase.png}
-    \caption{Phase diagram of a liquid with critical point and triple point identified.}
-    \label{phase}
-\end{figure}
-\\
-The theory of this project is largely predicated upon the behavior of critical systems, namely the phenomena of self-similarity.
-Non-critical systems exhibit different characteristics when viewed at different length scales.
-Ever-present small-scale correlations (that is, fluctuations in density) are relevant, if not dominant, when viewing the system at very small scales, but typically become less important at larger scales and nearly irrelevant near the thermodynamic limit.
-Near the critical point all correlation wavelengths are present, and so the system looks the same regardless of length scale chosen.
-This self-similarity allows for the decomposition of the large critical cell to smaller constituent cells.
-The details of the procedure are more thoroughly discussed in Chapter.
+We will begin by outlining the necessary models for this work, and then in detail discuss the theoretical underpinnings of our method.
+We will then describe the algorithms used to compute the concerned quantities with the the discussion on results following.
 
 \subsection{Free energy}
-The energy we are interested in calculating is the ``absolute'' Helmholtz free energy $F^{\text{abs}}$. The Helmholtz free energy is a Legendre transform of the internal energy $U$
+The Helmholtz free energy is a Legendre transform of the internal energy $U$
 \begin{align}
@@
 \end{align}
+The Helmholtz free energy can also be separated into the contributions:
+\begin{align}
+    F &= F_{\text{id}} + F_{\text{exc}}
+\end{align} 
+The $F_{\text{id}}$ term is the energy from the ideal gas behavior of the fluid.
+The excess free energy $F_{\text{exc}}$ contains the remaining energy of the fluid after including the ideal gas energy, and is therefore related to the energy of the configuration of the system.\\
+
 With $F$ calculated, its partial derivatives yield all other interesting thermodynamic quantities.
@@
 
+\subsection{Hard Sphere}
+The ``hard sphere'' is the simplest fluid model -- a hard sphere approximates a fluid particle as having no interactions other than collision.
+The only parameter for the hard sphere model is the sphere radius $\sigma$.
+The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
+Two spheres are not allowed to overlap, and else contribute nothing to the internal energy of the system.
+The energy between spheres $1$ and $2$, called the intermolecular pair potential, is then
+\begin{equation} 
+    \Phi_{12}(|\vec{r}|) = \begin{cases}\infty & |\vec{r}|\leq \sigma\\ 0 & |\vec{r}| > \sigma \end{cases}
+\label{phi12hs}
+\end{equation}
+No attraction in Eq.~\ref{phi12hs}, and consequently no excess internal energy, implies that the excess free energy is purely entropic: 
+\begin{align}
+    F_{exc} &= -T S_{exc}
+    \label{hs-fexc}
+\end{align}
+Eq.~\ref{hs-Fexc} showcases the beauty of the hard sphere system: one need only calculate the excess entropy of the hard sphere system to find the excess free energy.   
+\subsubsection{Carnahan-Starling Equation of State}
+We will need to compute the free energy of the a hard sphere system innumerable times throughout this work, and so we must make use of the Carnahan-Starling equation of state.
+shown in Eq.~\ref{C-S-eq}.
+\begin{align}
+    F_{exc}(\eta) &= N k_B T\,\frac{4\eta - 3\eta^2}{(1-\eta)^2}
+    \label{C-S-eq}  
+\end{align} 
+where $\eta$ is the filling fraction, $N/V$.
+The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit \cite{cs-eos}.
+We are very much concerned with cells significantly far from the thermodynamic limit.
+Thankfully, the Carnahan-Starling equation of state is still an excellent approximation at low to moderate filling fractions.\\
+
+\ignore{[move me | For simulation purposes, the system is required to have periodic boundary conditions - that is, .]}
+
+\subsection{Square-Well Liquid}
+As a hard-sphere model, the spheres are not allowed to overlap.
+Any arrangement of spheres - hereafter referred to as the ``state'' of the system - that involves either two or more spheres overlapping or a sphere outside the allowed area - ``cell'' - is defined to be invalid.
+The ``square'' refers to the piecewise intermolecular pair potential $\Phi_{12}$, shown mathematically in Eq.
+\ref{phi12} and graphically in Figure~\ref{sw_phi}.
+
+\begin{equation} 
+\Phi_{12}(\vec{r}) = \begin{cases}\infty & |\vec{r}|\leq \sigma\\ -\epsilon & \sigma \leq |\vec{r}| \leq \sigma\lambda\\ 0 & |\vec{r}| > \sigma\lambda \end{cases}
+\label{phi12}
+\end{equation}
+
+This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
+We then define dimensionless temperatures $T/\epsilon$, energies $E/\epsilon$, and distances $r/\sigma$.
+Moreover, we make the decision to work in geometric units for coding simplicity, setting the Boltzmann constant $k_B$ to one.
+This then allows us to reasonably compare energies and temperatures since they have the same units.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=.75\textwidth]{SWF-E.png}
+    \caption{Demonstration of Square-Well approximation.}
+    \label{sw_phi}
+\end{figure}
+
+\subsection{Critical Point} 
+The critical point is the temperature and pressure (or any other thermodynamic pairing) on a phase diagram (shown in Figure ~\ref{phase}) where the line between liquid and vapor terminates - liquid and vapor become indistinguishable and are assimilated into the ``supercritical fluid'' phase.
+Many interesting effects occur here, including the well-researched phenomena of critical opalescence.
+The correlation length, loosely interpreted here as the maximum wavelength of the density fluctuations within the fluid, diverges to infinity near criticality.
+As mentioned previously, the diverging of the correlation length invalidates SAFT and many of its variants which assume constant density beyond small length scales.
+\begin{figure}
+    \centering
+    \includegraphics[width=.75\textwidth]{phase.png}
+    \caption{Phase diagram of a liquid with critical point and triple point identified.}
+    \label{phase}
+\end{figure}
+\\
+
+The theory of this project is largely predicated upon the behavior of critical systems, namely the phenomena of self-similarity.
+Non-critical systems exhibit different characteristics when viewed at different length scales.
+Ever-present small-scale correlations (that is, fluctuations in density) are relevant, if not dominant, when viewing the system at very small scales, but typically become less important at larger scales and nearly irrelevant near the thermodynamic limit.
+Near the critical point correlations are present at all scales, and so the system looks the same regardless of length scale chosen.
+This self-similarity allows for the decomposition of the large critical system to smaller constituent cells.
+The details of the procedure are more thoroughly discussed in the next section.
+
+\subsubsection{Criticality and Renormalization Group Theory}
+Renormalization group methods attempt to describe how a particular system's properties are altered when the system is viewed from varying length scales.
+Near the critical point, the fluid becomes self similar, and so fluctuations of all wavelengths are present.
+We can then compare quantities of related yet physically separate systems as they all exhibit the similar properties.\\
+
+Imagine a cubic cell of length $L_0$ and number of particles $N_0$ with a corresponding filling fraction $\eta_0 = 4/3\pi\sigma^3N_0/L_0^3$.
+We enforce that this cell has periodic boundary conditions -- that is, when a sphere would normally collide with the wall of the cell it instead reappears on the other side of the cell with the same velocity.
+This implementation of periodicity truncates the allowed fluctuations to wavelength $\lambda_0 = L_0$.
+Fluctuations at this scale contribute some absolute free energy $F_0$ to the absolute free energy of the full critical system.
+\\
+
+Now consider a ``doubled'' version of the original cell, with length and allowed fluctuations $L_1 = \lambda_1 = 2L_0$, that contains $n_1 = 8n_0$ particles such that the original cell's density is intact.
+This new cell is, in a sense,  the original cell viewed on a different length scale.
+There is a Helmholtz free energy difference $\delta F_1$ between the original cell and the new doubled cell.
+If we continue to double the cell in this manner, we can find the absolute Helmholtz free energy $F^{\text{abs}}_{\text{tot}}$ of the bulk:
+\begin{align}
+    \label{GRGT-F1}
+    F^{\text{abs}}_{\text{tot}} = \sum_{i=1}^{k}\delta F^{\text{abs}} &= \lim_{k\to\infty}\sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}\\
+    &\approx \sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}
+    \label{GRGT-F2}
+\end{align}
+Strictly speaking, a formal limit is required for equality in Eq.~\ref{GRGT-F1}, rather than only including wavelengths up to $2^k\lambda_0$.
+Though theoretically all wavelengths are present at the critical point, it is only necessary to include the first several to accurately describe systems moderately close to the critical point, so we instead choose to use Eq.~\ref{GRGT-F2}.
+Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely Metropolis-Hastings Monte Carlo and thermodynamic integration.
+%A graphical representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
+
+\ignore{\begin{algorithm}[tb]
+    \caption{Renormalization-esque calculation of total free energy}
+    \label{alg_renorm}
+    \hrulefill
+    \begin{enumerate}
+        \item Simulate smallest cell of length $L_0$.
+        \item Calculate absolute free energy from simulation output.
+        \item Simulate new cell with same density but octupled volume of previous cell.
+        \item Calculate absolute free energy from simulation output.
+        \item Calculate change in absolute free energy between the current regime and the last, add to running $\Delta F^{\text{abs}}_{\text{tot}}$
+        \item Repeat 3-4 for each desired doubling regime.
+    \end{enumerate}
+    \hrulefill
+\end{algorithm}
+
+%%%%%%% TIKZZZZZ%%%%%%%
+
+\begin{figure}
+\centering
+    \resizebox{.75\textwidth}{!}{
+     \begin{tikzpicture}
+            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (0,0) rectangle (1,1);
+            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (2,0) rectangle (4,2);
+            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (5,0) rectangle (9,4);
+            %\draw[black, ultra thick, fill=yellow, fill opacity=.2] (10,0) rectangle (14,4);
+            \draw[<-] (5.15,2) to (5.65,2);
+            \draw[->] (7,3.35) to (7,3.85);
+            \draw[<-] (7,.15) to (7,.65);
+            \draw[->] (8.35,2) to (8.85,2);
+            \node[above] at (.5,1) {\tiny $g_0$: $L=L_0$, $N=1$};
+            \node[above] at (3,2) {\tiny $g_1$: $L=2L_0$, $N=4$};
+            %\node[above] at (7,4) {\tiny $g_2$: $L=4L_0$, $N=8$};
+            \node[above] at (7,4) {\tiny $g_n$: $L=2^nL_0$, $N=(2^n)(2^n)=4^n$};
+
+            % Balls!
+            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];
+            % Second box
+            \draw[ultra thick, fill=red] (3.25,.5) circle [radius=.25];
+            \draw[ultra thick, fill=red] (3.5,1) circle [radius=.25];
+            \draw[ultra thick, fill=red] (2.375,1.5) circle [radius=.25];
+            \draw[ultra thick, fill=red] (2.5,.75) circle [radius=.25];
+            % third
+            \draw[ultra thick, fill=red] (5.5,1.5) circle [radius=.25];
+            \draw[ultra thick, fill=red] (5.375,.375) circle [radius=.25];
+            \draw[ultra thick, fill=red] (6,3.5) circle [radius=.25];
+            \draw[ultra thick, fill=red] (7.5,3.25) circle [radius=.25];
+            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];    
+            \draw[ultra thick, fill=red] (7,2) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (5.75,3) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (6.5,2.75) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (8, 3) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (7.5,1.25) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (7.75,1.75) circle [radius=.25];        
+            \draw[ultra thick, fill=red] (8.5,1.5) circle [radius=.25];
+            \draw[ultra thick, fill=red] (6.5,1.25) circle [radius=.25];                
+     \end{tikzpicture}}
+\caption{2-D Demonstration of MC-GRGT doubling method.}
+\label{GRTG-demo} 
+\end{figure} 
+}
+
+%%%%%% TIKZZZZ %%%%%%
+
+\section{Separation and Characterization of Free Energy}
+
 Since $F$ contains a natural log term, we can separate and characterize the individual contributions of kinetic energy (``ideal'') and energy associated with the configuration of the system (``excess'').
-A derivation of the two energy contributions for a monoatomic fluid follows. \\
+A derivation and characterization of the two energy contributions for a monoatomic fluid follows.
+\\
 
@@
 \section{Methods}
-This work requires the intersection of three methods to calculate the absolute free energy $F_{abs}$ of the bulk cell.
-A preliminary value of the free energy of the bulk can be obtained from Monte Carlo simulations.
-This value can then be made absolute by carefully comparing the results with thermodynamic integration, a method of determining the free energy of the hard-sphere system.
-The resulting absolute free energies for each fluctuation length can be summed to yield the absolute free energy of the bulk.
-{\color{red}[deal with the problems here]}
-
+To compute the absolute Helmholtz free energy $F^{abs}$ of the cell, we must make use of two Monte Carlo methods.
+The first, broad-histogram Monte Carlo, is a square well simulation that yields the density of states of the system.
+Unfortunately, the given density of states has an arbitrary scaling that is system-dependent, which leaves us unable to compare results from different doubling regimes.
+The second, thermodynamic integration, is a hard sphere simulation that calculates the excess entropy of the system.
+A careful equality between sets of results from these simulations will allow us to compute the absolute free energy.
+Thought another way, if we view the absolute Helmholtz free energy has being a function of both temperature and density, we can think of broad-histogram Monte Carlo as integration over temperature and thermodynamic integration as an integration over density.
 \subsection{Monte Carlo Methods}
 
-Unbiased Monte Carlo methods utilize random or pseudorandom sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
+Unbiased Monte Carlo methods use random sampling and check for certain conditions or measure certain quantities after a specified number of iterations.
 Traditional random sampling requires that the current step, or ``move'', is irrespective of previous steps.
@@
 \subsubsection{Metropolis-Hastings}
-Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving.
+Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving \cite{metro-hastings}.
 If we define the ensemble of states of the system in equilibrium by a vector column $\mathbf{q}$, and the resulting state $\mathbf{p}$ after a single time step, the relationship
@@
 Given that we only are concerned with quantities that can be computed given the density of states, we never actually ``simulate''.
-With the density of states computed, we can now calculate the partition function and thereby the free energy.
-
-\subsection{Renormalization Group Theory}
-Renormalization group methods attempt to describe how a particular system's properties are altered when the system is viewed from varying length scales.
-Near the critical point, the fluid becomes self similar, and so fluctuations of all wavelengths are present.
-We can then compare quantities of related yet physically different systems as they all exhibit the similar properties.\\
-We begin with a periodic cubic cell of length $L_0$ and volume $L_0^3$. 
-This cell has periodic boundary conditions and consequently allows for fluctuations up to wavelength $L_0$.
-These fluctuations contribute some absolute free energy $F_0$ to the absolute free energy of the full cell.
-We then double the length of the cell to allow for fluctuations of up to wavelength $L_0$.
-We continue this process until all relevant wavelengths have been included.
-Adding up the change in absolute free energy calculated in this manner then yields the total free energy of the critical system, which can then be compared to accepted simulation methods.
-Algorithm~\ref{alg_renorm} details this process.
-
-\begin{algorithm}[tb]
-    \caption{Renormalization-esque calculation of total free energy}
-    \label{alg_renorm}
-    \hrulefill
-    \begin{enumerate}
-        \item Simulate smallest cell of length $L_0$.
-        \item Calculate absolute free energy from simulation output.
-        \item Simulate new cell with same density but octupled volume of previous cell.
-        \item Calculate absolute free energy from simulation output.
-        \item Calculate change in absolute free energy between the current regime and the last, add to running $\Delta F^{\text{abs}}_{\text{tot}}$
-        \item Repeat 3-4 for each desired doubling regime.
-    \end{enumerate}
-    \hrulefill
-\end{algorithm}
-
-Eq. ~\ref{GRGT-F1} details the mathematics of this coupling procedure.
-
-\begin{align}
-    \label{GRGT-F1}
-    F^{\text{abs}}_{\text{tot}} = \sum_{i=1}^{k}\Delta F^{\text{abs}} &= \lim_{k\to\infty}\sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}\\
-    &\approx \sum_{i=1}^{k} F^{\text{abs}}_{i} - F^{\text{abs}}_{i-1}
-    \label{GRGT-F2}
-\end{align}
-
-Strictly speaking, a formal limit is required for equality in Eq.~\ref{GRGT-F1}, rather than only including wavelengths up to $2^k\lambda_0$.
-Though theoretically all wavelengths are present at the critical point it is only necessary to include the first several to accurately describe systems moderately close to the critical point.
-Given approximation, we choose instead to use Eq.~\ref{GRGT-F2}.
-Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely Metropolis-Hastings Monte Carlo and thermodynamic integration.
-A graphical representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
-
-%%%%%%% TIKZZZZZ%%%%%%%
-
-\begin{figure}
-\centering
-    \resizebox{.75\textwidth}{!}{
-     \begin{tikzpicture}
-            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (0,0) rectangle (1,1);
-            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (2,0) rectangle (4,2);
-            \draw[black, ultra thick, fill=yellow, fill opacity=.2] (5,0) rectangle (9,4);
-            %\draw[black, ultra thick, fill=yellow, fill opacity=.2] (10,0) rectangle (14,4);
-            \draw[<-] (5.15,2) to (5.65,2);
-            \draw[->] (7,3.35) to (7,3.85);
-            \draw[<-] (7,.15) to (7,.65);
-            \draw[->] (8.35,2) to (8.85,2);
-            \node[above] at (.5,1) {\tiny $g_0$: $L=L_0$, $N=1$};
-            \node[above] at (3,2) {\tiny $g_1$: $L=2L_0$, $N=4$};
-            %\node[above] at (7,4) {\tiny $g_2$: $L=4L_0$, $N=8$};
-            \node[above] at (7,4) {\tiny $g_n$: $L=2^nL_0$, $N=(2^n)(2^n)=4^n$};
-
-            % Balls!
-            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];
-            % Second box
-            \draw[ultra thick, fill=red] (3.25,.5) circle [radius=.25];
-            \draw[ultra thick, fill=red] (3.5,1) circle [radius=.25];
-            \draw[ultra thick, fill=red] (2.375,1.5) circle [radius=.25];
-            \draw[ultra thick, fill=red] (2.5,.75) circle [radius=.25];
-            % third
-            \draw[ultra thick, fill=red] (5.5,1.5) circle [radius=.25];
-            \draw[ultra thick, fill=red] (5.375,.375) circle [radius=.25];
-            \draw[ultra thick, fill=red] (6,3.5) circle [radius=.25];
-            \draw[ultra thick, fill=red] (7.5,3.25) circle [radius=.25];
-            \draw[ultra thick, fill=red] (.5,.5) circle [radius=.25];    
-            \draw[ultra thick, fill=red] (7,2) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (5.75,3) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (6.5,2.75) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (8, 3) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (7.5,1.25) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (7.75,1.75) circle [radius=.25];        
-            \draw[ultra thick, fill=red] (8.5,1.5) circle [radius=.25];
-            \draw[ultra thick, fill=red] (6.5,1.25) circle [radius=.25];                
-     \end{tikzpicture}}
-\caption{2-D Demonstration of MC-GRGT doubling method.}
-\label{GRTG-demo} 
-\end{figure}
-
-%%%%%% TIKZZZZ %%%%%%
+With the density of states computed, we can now calculate the system's partition function and thereby the Helmholtz free energy by Eq.~\ref{FdefSM}.
 
@@
 \subsubsection{General Process}
-Thermodynamic integration is a unique way to determine the ``excess'' free energy of the system.
-Using hard sphere Monte Carlo, we transition from a filling fraction of zero to a dense system with filling fraction $\eta_{max}$, calculating the change in free energy at each step.
-When we have reached the desired final filling fraction we have found the free energy that was necessary to create that configuration of particles - this is exactly the absolute free energy of the system.\\
+Thermodynamic integration is a hard sphere simulation that computes the excess entropy of a system with a given filling fraction and number of spheres.
+Using hard-sphere Monte Carlo, we transition from a filling fraction of zero to the desired value, calculating the change in entropy at each step.
+
 \subsubsection{Algorithm}
@@
     \draw[ultra thick, ->] (14.5,5) to (15.5,5);
-    \node[align=center] [above] at (15, 5.25) {\small $n-k$ \\[-1em] steps};
+    \node[align=center] [above] at (15, 5.25) {\small $n-k$ \\[0 em] steps};
     \draw[ultra thick, fill=red] (3.5,4.25) circle [radius=.25];
@@
 
-Combining Eq.~\ref{C-S-eq} with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy.
+Combining the Carnahan-Starling equation of state (Eq.~\ref{C-S-eq}) with a simple bisection method allows for control of the filling fraction step size given a desired change in free energy.
 We determined that steps yielding an approximate change in specific absolute free energy of $\Delta F^{abs}/NT = \ln(2)$ through trial and error - smaller step sizes unnecessarily increased simulation time and larger step sizes tended to shrink too quickly causing the simulation to miss relevant states in the system.
@@
 An important remaining question is whether there is a preferred choice of minimum cell length.
-Physically we would [might?] expect that there is some dominant minimum fluctuation in the bulk cell - this would then lend a natural decision for the minimum cell length $\lambda_0$.
+Physically we would expect that there is some dominant minimum fluctuation in the bulk cell - this would then lend a natural decision for the minimum cell length $\lambda_0$.
 Unfortunately we do not know this {\it a priori}, so we are reduced to guessing.
@@
 
-\subsection{Primitive Cell}
-The ``scrunched'' system allows exactly four spheres to be packed -- at maximal filling fraction $\eta_{max} = \approx .741$ -- in the original non-doubled cell.
-This case corresponds to the unit cell a face-centered cubic lattice, the simplest sphere packing possible.
-We chose to focus on the scrunched case for two reasons.
-Firstly, if any physical length of the system is likely to be the correct choice for minimum fluctuation length.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 Liquid and vapor are simultaneously present in the fluid at densities that cause local minima in the grand free energy.
-We know that the grand free energy has a guaranteed minimum at the origin because [things]; we then only need to identify the other minimum.
+We know that the grand free energy has a guaranteed minimum at the origin; we then only need to identify the other minimum.
 \\
@@
 Future work can include investing further doubling regimes ($i\geq 3$).
+
+\section{Acknowledgments}
+I would to thank David Roundy for both the project idea, the theoretical understanding, and countless time spent helping ferret out bugs.
+Further, thanks to both Michael Perlin and Austin Valeske for explanations of their code.
+Finally, thanks to my collective peers for feedback and encouragement.
 %\bibliographystyle{plain}
 \cleardoublepage
-\bibliography{./Thesis}
+\bibliography{Thesis}
 \end{document}

### 2016-05-31 papers/thesis-vischer/abstract.tex
Most of thesis work for nearly final draft. Small edits to results tomorrow morning. Tests skipped!

@@
 
-Analytic models using renormalization methods to approximate the free energy of fluids have been developed, but there is currently no way to investigate the intermediate steps of the approximation.
-We present a computational method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Metropolis-Hastings Monte-Carlo simulations.
-We further detail a method by which to compute the ``absolute'' free energy, including both the ideal gas free energy and the excess free energy.
-We further present an application of this method to the well-research square-well liquid near the critical region.
-We find that the square-well liquid is well-characterized by a base cell length of $\sqrt2\sigma$, the side length of an FCC lattice, by establishing liquid-vapor coexistence between scaled temperatures $0.5 \leq T < .75$.
 \clearpage

### 2016-05-31 papers/thesis-vischer/Thesis.tex
basically finished for perusal. finished many content areas, cleaned up units and text, added more results discussion. Tests skipped!

@@
     F &= F_{\text{id}} + F_{\text{exc}}
+    \label{F_abs}
 \end{align} 
@@
 Thankfully, the Carnahan-Starling equation of state is still an excellent approximation at low to moderate filling fractions.\\
-
-\ignore{[move me | For simulation purposes, the system is required to have periodic boundary conditions - that is, .]}
 
@@
 Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely Metropolis-Hastings Monte Carlo and thermodynamic integration.
-%A graphical representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
+A graphical, two-dimensional representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
 
@@
 \end{algorithm}
+}
 
@@
 \end{figure}
-}
 
@@
 \begin{align}
-    Z_{id} &=  \frac{1}{N!}\left[\int_s \diff\mathbf{p} ~e^{-\beta K(\mathbf{p})}\right]^{N}
+    Z_{id} &= \frac{1}{N!}\left[\int_s V \diff\mathbf{p} ~e^{-\beta K(\mathbf{p})}\right]^{N}
 \end{align}
@@
 \begin{align}
-    Z_{id} &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{-\frac{\beta \hbar^2}{2m} \left(\frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}}\right)^2}\right]^{N} \\
-    &= \frac{1}{N!}\left[\int_0^\infty \diff^3 n ~e^{ \frac{-2\beta \hbar^2\pi^2}{mL^2}(n_x^2 + n_y^2 + n_z^2)}\right]^{N}
+    Z_{id} &= \frac{(2\pi \hbar)^{3N}}{V^N N!}\left[\int_0^\infty \diff^3 n ~e^{-\frac{\beta \hbar^2}{2m} \left(\frac{2\pi}{L}(n_x\mathbf{\hat{x}} + n_y\mathbf{\hat{y}} + n_z\mathbf{\hat{z}}\right)^2}\right]^{N} \\
+    &= \frac{h^{3N}}{N!}\left[\int_0^\infty \diff^3 n ~e^{ \frac{-2\beta \hbar^2\pi^2}{mL^2}(n_x^2 + n_y^2 + n_z^2)}\right]^{N}
 \end{align}
-where extra factors of $\hbar $and$ \pi$ were gained from the change of variables. Evaluating this integral yields
+where extra factors of $\hbar $ and volume were gained from the change of variables. Evaluating this integral yields
 \begin{align}
-    Z_{id} &= \left(\frac{V^N}{N!\Lambda^{3N}}\right) \,, \\
+    Z_{id} &= \left(\frac{h^{3N} V^N}{N!\Lambda^{3N}}\right) \,, \\
     \Lambda &= \frac{\hbar\sqrt{2\pi}}{\sqrt{mT}}
@@
 \begin{align}
-    Z &= Z_{exc} \left(\frac{V}{N!\Lambda^{3}}\right)^{N} \,.
+    Z &= Z_{exc} \left(\frac{h^3 V}{N!\Lambda^{3}}\right)^{N} \,.
 \end{align}
@@
 \begin{align}
-    F_{abs} &= -T\,\ln\left(Z_{exc} \frac{V^N}{N!\Lambda^{3N}}\right)\\
-    &= -T\ln\left(Z_{exc}\right)\, -T\ln\left(\frac{V}{(N!)^{1/N}\Lambda^3} \right)^N\\
-    &= -T\ln\left(Z_{exc}\right) - NT\ln\left(\frac{V}{\Lambda^3}\right) + NT\ln\left(N!^{1/N}\right)\\ 
+    F_{abs} &= -T\,\ln\left(Z_{exc} \frac{h^{3N} V^N}{N!\Lambda^{3N}}\right)\\
+    &= -T\ln\left(Z_{exc}\right)\, -T\ln\left(\frac{h^3 V}{(N!)^{1/N}\Lambda^3} \right)^N\\
+    &= -T\ln\left(Z_{exc}\right) - NT\ln\left(\frac{h^3 V}{\Lambda^3}\right) + NT\ln\left(N!^{1/N}\right)\\     
 \end{align}
-Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N - N$ for large N), we arrive at \textcolor{red}{BRENDEN FIX: $N\ln N - N$, plus I think we may want to add to the excess free energy (probably to the excess HS entropy?) the difference?}
+Redefining the first factor including $Z_{exc}$ to be $F_{exc}$, and then using Sterling's approximation ($\ln(N!) = N\ln N - N$ for large N), we arrive at \ignore{\textcolor{red}{BRENDEN FIX: $N\ln N - N$, plus I think we may want to add to the excess free energy (probably to the excess HS entropy?) the difference?}}
 \begin{align}
-    F_{abs} &= F_{exc} - NT\left(\ln\left(\frac{V}{\Lambda^3}\right) - \ln\left(N\right) + 1\right)
+    F_{abs} &= F_{exc} - NT\left(\ln\left(\frac{h^3 V}{\Lambda^3}\right) - \ln\left(N\right) + 1\right)
 \end{align}
@@
 The question of organization becomes important with this amount of files being generated.
-Each different system simulation (length $L = \sqrt2 \sigma, 5,10$) has data regarding each doubling regime ($i=0,1,2,3?$) and each doubling regime contains data for each possible sphere count ($N=1,2,3\dots N_{max}$, where $N_{max}$ is typically limited by maximal sphere packing of $N_{max}/L^3\approx.741$).
+Each different system simulation (length $L = \sqrt2 \sigma, 5,10$) has data regarding each doubling regime ($i=0,1, 2, 3$) and each doubling regime contains data for each possible sphere count ($N=1, 2, 3\dots N_{max}$, where $N_{max}$ is typically limited by maximal sphere packing of $N_{max}V_{\text{sphere}}/L^3\approx.741$).
 The file structure mirrors the above symmetry:
@@
 \subsection{Calculating the Absolute Free Energy}
-{\color{red} please rewrite me. This paragraph has been terrible since it was written.}\\
-It is straightforward the compute a Helmholtz free energy through the Monte Carlo simulations; the difficulty in this method lies in computing the correct free energy.
-TMMC, as mentioned previously, yields a free energy accurate only up to a unknown constant.
-Since we need to add up changes in free energy across multiple simulations, we need each calculated free energy to be ``absolute'' - that is, we must eliminate this unknown constant.
-\\
+As mentioned previously, the broad histogram Monte Carlo simulations have an arbitrary scaling factor in the density of states output.
+Reasonable comparisons between different doubling regimes will require  the removal of this arbitrary scaling.
+The key to doing so lies in the infinite temperature limit of the square well fluid.\\
 
@@
 \subsection{Thermodynamic Quantities from Simulation}
-Given the density of states output $\ln[D(E)]$ from the Monte Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic , for instance, the excess internal energy:
+Given the density of states output $\ln[D(E)]$ from the Monte Carlo simulation, combined with {\it a priori} knowledge that the system obeys a canonical ensemble distribution of states, we can calculate any thermodynamic quantity we wish given only the density of states.
+For instance, we can compute the the excess internal energy:
 \begin{align}
@@
 where, as usual, $\beta = T^{-1}$.
-unfortunately, tends to introduce overflow or underflow errors during computation.
-We combat this by subtracting the maximum value of the natural log of the density of states \ignore{subtracting maximum ought to make it more negative\dots}from each element of $D(E)$, and then adding the maximum back in after the exponential.
+tends to introduce overflow or underflow errors during computation.
+We combat this by subtracting the maximum value of the natural log of the density of states from each element of $D(E)$, and then adding the maximum back in after the exponential.
 We calculate the excess free energy by using the entropy equality developed previously:
@@
      S_{SW,~exc}\Bigm\lvert_{T=\infty} &= S_{HS,~exc}\\
-     \implies F_{SW,~exc}\Bigm\lvert_{T=\infty} &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} \\
-     \implies F_{SW,~exc}(T) &= F_{SW,~exc}\Bigm\lvert_{T=\infty} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
+     F_{SW,~exc}\Bigm\lvert_{T=\infty} &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} \\
+     F_{SW,~exc}(T) &= F_{SW,~exc}\Bigm\lvert_{T=\infty} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
      \\
-     \implies F_{SW,~exc}(T) &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
+     F_{SW,~exc}(T) &= U_{SW,~exc}\Bigm\lvert_{T=\infty} -TS_{HS,~exc} + \left(F_{SW,~exc}(T) - F_{SW,~exc}\Bigm\lvert_{T=\infty}\right)
      \\
@@
 \end{align} 
-The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration, $Z$ is calculated using  .
-Once we have calculated both the excess free energy and excess internal energy, we can directly find the excess entropy:
+The excess hard sphere entropy $T_{HS,~exc}$ is calculated directly from thermodynamic integration.
+Both partition functions and the excess internal energy can be obtained from broad histogram results.
+Once we have calculated both the excess free energy and excess internal energy, we can find the excess entropy for the square well system:
 \begin{align}
@@
 \end{align}
+Moreover, with the excess free energy computed, we merely need to include the ideal gas energy to find the absolute free energy of the system by way of Eq.~\ref{F_abs}.
 
@@
 Using the coupling procedure outlined previously, we calculated changes in free energy up to the $n^{\text{th}}$ doubling regime for several different cell lengths.
-We choose to focus on the ``scrunched'' case where the cell length is $\sqrt2\sigma$.\\
+We choose to focus on the ``scrunched'' case where the base cell length is $\sqrt2\sigma$.\\
 
-Figures ~\ref{F-T} through ~\ref{F-eta} show several computed thermodynamic quantities for the scrunched system.
+Figures ~\ref{Shs-eta} through ~\ref{F-eta} show several computed thermodynamic quantities for the scrunched system.
 Each plot contains calculated quantities for either the first two doubling regimes or the first three.
-Calculations from the first doubling regime ($L=\lambda_0$) are solid lines, with the second doubling regime ($L = 2\lambda_0$) in dashes and the third ($L = 4\lambda_0$), when present, in circles .
+Calculations from the first doubling regime ($L=\lambda_0$) are solid lines, with the second doubling regime ($L = 2\lambda_0$) in dashes and the third ($L = 4\lambda_0$), when present, in circles.
 It is worth noting that the succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
 The intervening points have been filled in using the Carnahan-Starling equation of state.\\
-\begin{figure}  
+\ignore{\begin{figure}  
 \centering
@@
     \label{F-T}
-\end{figure}
-
-Figure \ref{F-T} shows expected behavior.
-As a function of temperature, the excess free energy monotonically increases for all densities tested.
-We see a positive free energy because $F = U -TS$; as temperature increases without bound, entropy remains negative and [blah blah more].\\
+\end{figure}}
 
@@
 Figure \ref{Shs-eta} shows the computed excess entropy as a function of density.
-We see that the both the first and second doubling regime closely match the Carnahan-Starling prediction for the entropy.
+We see that both the first and second doubling regime closely match the Carnahan-Starling prediction for the entropy.
 After freezing, the first doubling regime veers from the prediction by the Carnahan-Starling equation of state; since the Carhan-Starling approximation is accurate only for low density fluids, we expect this behavior.\\
-
 \begin{figure}
@@
     \includegraphics[width=.75\textwidth]{../renormalization/figs/Uexc-vs-eta.pdf}
-    \caption{Excess internal energy as a function of density. The $x$ markers connected by dashed lines delineate second doubling regime densities that were simulated.}
+    \caption{Excess internal energy as a function of density. .}
     \label{U-eta}
 \end{figure}
+
 We see in Figure~\ref{U-eta} that the excess internal energy, as we would expect, becomes more negative as the density increases.
-The lower the temperature, the more ``clumped'' the system becomes, resulting in lower internal energies.
-[correct?]
-
-\begin{figure}
+The more ``clumped'' the system becomes, the more spheres are within one another's attractive wells, resulting in more negative system energies.
+\begin{figure}[h]
 \centering
@@
 \end{figure}
+We see in Figure~\ref{F-eta} laregly positive free energies because $F = U -TS$; as the density increases, the entropy becomes more negative.
+For a given temperature, the increasingly negative entropy results in a positive free energy.
+We also see that higher temperatures have a higher free energy, which results from $T$ scaling the strength of the entropy contribution to the free energy.\\
 
@@
 Tweaking the value of the chemical potential will allow us to easily identify second minima in the grand free energy, and thereby the densities of the liquid and vapor present in the system.
-\begin{figure}
+\begin{figure}[H]
 \centering
@@
 \end{figure}
-Figure \ref{Phi-eta} 
+Figure \ref{Phi-eta} shows the first three doubling regimes for three temperatures.
+For each temperature, the first doubling regime showcases a grand free energy minimum around the freezing point $\eta_{\text{freeze}\approx 0.5}$.
+The second doubling regime calculations show minima at $\eta_{\text{coex},~T_1} = 0.29$, $\eta_{\text{coex},~T_2} = 0.35$, and $\eta_{\text{coex},~T_3} = 0.4$.
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2016-05-31 papers/thesis-vischer/Thesis.tex
fix reference bug in Brenden's thesis

@@
 \end{align}
-Eq.~\ref{hs-Fexc} showcases the beauty of the hard sphere system: one need only calculate the excess entropy of the hard sphere system to find the excess free energy.   
+Eq.~\ref{hs-fexc} showcases the beauty of the hard sphere system: one need only calculate the excess entropy of the hard sphere system to find the excess free energy.
 \subsubsection{Carnahan-Starling Equation of State}

### 2016-05-31 papers/thesis-vischer/Thesis.tex
fix an annoying line break with Brenden

@@
 
-\begin{algorithm}[!b]
+\begin{algorithm}
 \caption{Thermodynamic integration.}
@@
 The question of organization becomes important with this amount of files being generated.
-Each different system simulation (length $L = \sqrt2 \sigma, 5,10$) has data regarding each doubling regime ($i=0,1, 2, 3$) and each doubling regime contains data for each possible sphere count ($N=1, 2, 3\dots N_{max}$, where $N_{max}$ is typically limited by maximal sphere packing of $N_{max}V_{\text{sphere}}/L^3\approx.741$).
+Each different system simulation (length $L = \sqrt2 \sigma, 5,10$) has data regarding each doubling regime ($i=\nobreak 0,1, 2, 3$) and each doubling regime contains data for each and every possible sphere count $N=\nobreak 1, 2, 3\dots N_{max}$, where $N_{max}$ is typically limited by maximal sphere packing of $N_{max}V_{\text{sphere}}/L^3\approx\nobreak 0.741$.
 The file structure mirrors the above symmetry:

### 2016-06-01 papers/thesis-vischer/Thesis.tex
fixed all things from thesis scribbles from david edits, sans 2 citations to be added tomorrow morning Tests skipped!

@@
 \subsection{Motivation}
-Fluids have long garnered interest\cite{theorysimpleliquids}\cite{theoryofcrits}, particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
+Fluids have long garnered interest[verlet, van der waals, move Th.
+o'crits], particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
 A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor become indistinguishable and, more importantly, fluctuations of fluid density at large length scales occur.
@@
 Past efforts to incorporate renormalization group methods into standard Statistical Associating Fluid Theory (``SAFT'') methods have seen success \cite{white2001global}, particularly the treatment offered by Forte et al.~\cite{forte2011application} with the inclusion of a fitted parameter related to the correlation length.
-GRG's history in fluid simulation has largely been a fitted one - that is, determining a free parameter and then simulating many times to find the most well-suited value for a particular fluid \ignore{[!fixme not quite]}.\\
+Rather than include a fitted parameter, we elect to use a renormalization-inspired method to calculate the free energy of a liquid that is {\it predicated} upon critifcal behavior rather than invalidated by it.
+\\
 
-We present here a method to calculate the free energy of a liquid that largely avoids this fitting problem.
-We exploit the periodicity of Monte Carlo methods to decompose large cells in the thermodynamic limit into small cells that only allow certain fluctuation wavelengths.
+We exploit the critical phenomena of self-similarity to decompose large cells near the thermodynamic limit into small cells that only allow certain fluctuation wavelengths, enforced by the periodicity of Monte Carlo methods.
 We then perform simulations of different doubling regimes, each of which considers a different maximum wavelength of fluctuations, where the fluid density $\eta$ remains fixed but the volume $V$ and number of particles $N$ have increased.
-We can calculate a so-called``absolute'' free energy, including the ideal and excess free energy contributions, for each doubling regime by using two Monte Carlo methods in concert: Metropolis-Hastings Monte Carlo, and thermodynamic integration.
+We can calculate a so-called ``absolute'' free energy, including the ideal and excess free energy contributions, for each doubling regime by using two Monte Carlo methods in concert: Metropolis-Hastings Monte Carlo, and thermodynamic integration.
 Summing the difference in absolute free energy of each doubling regime will then approximate the absolute free energy of the full cell, especially near the critical point.
@@
 \begin{align}
-    F &= -T \ln Z\, .
+    F &= -k_B T \ln Z\, .
     \label{FdefSM}
@@
 The $F_{\text{id}}$ term is the energy from the ideal gas behavior of the fluid.
-The excess free energy $F_{\text{exc}}$ contains the remaining energy of the fluid after including the ideal gas energy, and is therefore related to the energy of the configuration of the system.\\
+The excess free energy $F_{\text{exc}}$ contains the remaining free energy of the fluid after including the ideal gas energy, and is therefore related to the free energy of the configuration of the system.\\
 
@@
 \end{align}
-where $\beta =  T^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$ state.
+where $\beta = (k_B T)^{-1}$ and $E_i$ is the energy of the $i^{\text{th}}$ state.
 Of course, it is impractical to sum over the individual states of the system because the number of states of a classical system are generally innumerable.
@@
 \begin{align}
-    Z &= \sum^{\text{energies}}~D(E)~e^{-\beta E_i}
+    Z &= \sum^{\text{energies}}~D(E)~e^{-\beta E}
 \end{align}
@@
 The system is defined to have zero total linear momentum, and any collision between two or more spheres is perfectly elastic.
-Two spheres are not allowed to overlap, and else contribute nothing to the internal energy of the system.
+Two spheres are not allowed to overlap, and otherwise contribute nothing to the internal energy of the system.
 The energy between spheres $1$ and $2$, called the intermolecular pair potential, is then
@@
 \subsubsection{Carnahan-Starling Equation of State}
-We will need to compute the free energy of the a hard sphere system innumerable times throughout this work, and so we must make use of the Carnahan-Starling equation of state.
-shown in Eq.~\ref{C-S-eq}.
+We will need to compute the free energy of the a hard sphere system innumerable times throughout this work, and so we make use of the Carnahan-Starling equation of state. shown in Eq.~\ref{C-S-eq}.
 \begin{align}
@@
 \subsection{Square-Well Liquid}
-As a hard-sphere model, the spheres are not allowed to overlap.
+The square well liquid is the simplest model for liquids\cite{theorysimpleliquids}.
+The square well approximation adds a constant short-range attractive potential to the hard sphere model.
+As in the hard-sphere model, the spheres are not allowed to overlap.
 Any arrangement of spheres - hereafter referred to as the ``state'' of the system - that involves either two or more spheres overlapping or a sphere outside the allowed area - ``cell'' - is defined to be invalid.
@@
 \begin{align}
-    Z &= \sum_s e^{-\beta E_s}
+    Z &= \sum_s^{\text{states}} e^{-\beta E_s}
 \end{align}
 
-Further, the energy of each state is simply the sum of the excess and kinetic energies $V$ and $K$, so
+Further, the energy of each state is simply the sum of the potential and kinetic energies $V$ and $K$, so
 
@@
 where $Z_{id}$ contains the $N!$ term by convention.
-The excess term $Z_{exc}$ is system dependent because $V(\mathbf{r})$ is buried inside the integral, but the ideal factor is the same for any system where the total energy separates in the preceding fashion.
+The excess term $Z_{exc}$ is system dependent because $V(\mathbf{r})$ is buried inside the integral, but the ideal gas factor is the same for any system where the total energy separates in the preceding fashion.
 We judiciously decide to leave the excess factor alone, for now, and work specifically with the ideal term.
@@
 This unfortunately introduces a mass quantity into the partition function which adds an additional parameter that needs to be determined to properly account for quantum mechanical effects.
-If we then allow the cell to become very large, then the we can easily see in Eq.
+If we then allow the cell to become very large, then the we can see in Eq.
 \ref{Zt_def} that the spacing between occupants in phase space becomes small, so we can integrate instead $\diff ^3 n$ over all possible values of $n$:
@@
 \begin{align}
-    Z &= Z_{exc} \left(\frac{h^3 V}{N!\Lambda^{3}}\right)^{N} \,.
+    Z &= \frac{Z_{exc}}{N!}\left(\frac{h^3 V}{\Lambda^{3}}\right)^{N} \,.
 \end{align}
-For liquids, the $Z_{exc}$ factor is difficult to evaluate; the hard sphere liquid, the simplest model for a dense fluid, does not have an analytic representation for the excess free energy.
+For liquids, the $Z_{exc}$ factor is difficult to evaluate; the hard sphere fluid, the simplest model for a dense fluid, does not have an analytic representation for the excess free energy.
 It follows that a considerable amount of work is required to determine the excess free energy for a given system.
@@
 Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving \cite{metro-hastings}.
-If we define the ensemble of states of the system in equilibrium by a vector column $\mathbf{q}$, and the resulting state $\mathbf{p}$ after a single time step, the relationship
+If we define the ensemble of states of the system in by a vector column $\mathbf{q}$, and the resulting ensemble of states after a step by $\mathbf{p}$, the relationship
 \begin{align*}
@@
 This matrix is $L\times L$, where $L$ is the number of energy levels in system.
-If all rows of the transition matrix are , we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
+If all rows of the transition matrix contain at least one nonzero entry, we can see that any state of the system is reachable after a finite number of transitions from any given initial state.
 A physically correct simulation will generate a transition matrix that satisfies the thermodynamic steady state condition; given the ensemble of equilibrium states $\mathbf{q}_{eq}$, we enforce that
@@
 If we walk across the entire energy range, we can construct histogram of energy transitions $h(E_i, E_j)$ for all $i, j$.
+This histogram is then proportional to the transition matrix we wish to construct.
 The walk goes as follows: starting with a state $\mathbf{q_i}$ with energy $E_i$, we attempt perform a random move to a state $\mathbf{q_f}$ with energy $E_f$.
@@
 Given that we only are concerned with quantities that can be computed given the density of states, we never actually ``simulate''.
+Further, we actually store the natural log of the density of states as output to help reduce overflow and underflow errors when calculating quantities.
 With the density of states computed, we can now calculate the system's partition function and thereby the Helmholtz free energy by Eq.~\ref{FdefSM}.
@@
 \begin{align}
-    S_{HS,~exc} &= \sum_i^K \ln\left(\frac{M_i}{N_i-M_i}\right)
+    S_{HS,~exc} &= \sum_i^K \ln\left(\frac{M_i}{N_i}\right)
 \end{align}
@@
 \end{align*}
-Each sphere count subfolder contains a single text file from TMMC output and another subfolder where all of the thermodynamic integration output is stored.
+Each sphere count subfolder contains a single text file from broad histogram output and another subfolder where all of the thermodynamic integration output is stored.
 Each thermodynamic integration simulation outputs a single text file upon completion of each filling fraction step.
@@
 At infinite temperature, the square well fluid has approximately the same behavior as the hard sphere fluid.
-We assert that the correct inclusion, then, is to enforce that the excess entropy calculated via TMMC at $T=\infty$ is equal to the hard sphere entropy calculated through thermodynamic integration.
+We assert that the correct inclusion, then, is to enforce that the excess entropy calculated via broad histogram at $T=\infty$ is equal to the hard sphere entropy calculated through thermodynamic integration.
 We do this by exploiting the infinite temperature behavior of the canonical ensemble:
@@
      U_{exc}(T) &= \frac{\sum_{i=0}^{\text{\tiny Energies}}\,E_i\, \exp(\ln[D(E)] - \beta E_i)}{\sum_{i=0}^{\text{\tiny Energies}} \exp(\ln[D(E)] - \beta E_i)}\\
-     &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i \exp\left(-\beta E_i\right)}{Z}
+     &= \frac{\sum_{i=0}^{\text{\tiny Energies}} E_i D(E)\exp\left(-\beta E_i\right)}{Z}
 \end{align} 
 where, as usual, $\beta = T^{-1}$.
-tends to introduce overflow or underflow errors during computation.
-We combat this by subtracting the maximum value of the natural log of the density of states from each element of $D(E)$, and then adding the maximum back in after the exponential.
+There are still overflow and underflow errors present, even with our choice to compute the natural log of the density of states.
+To further combat this issue, we subtract off the maximum of the natural log of the density of states before taking the exponential, and then add it back in afterwards -- this greatly reduces the likelihood that a quantity becomes erroneously large or small.
 We calculate the excess free energy by using the entropy equality developed previously:
@@
 Calculations from the first doubling regime ($L=\lambda_0$) are solid lines, with the second doubling regime ($L = 2\lambda_0$) in dashes and the third ($L = 4\lambda_0$), when present, in circles.
-It is worth noting that the succeeding plots with respect to density have been interpolated; with a fixed cell size, it is only possible to simulate discrete filling fractions.
-The intervening points have been filled in using the Carnahan-Starling equation of state.\\
-\ignore{\begin{figure}  
-\centering
-    \includegraphics[width=.75\textwidth]{../renormalization/figs/Fexc-vs-T.pdf}
-    \caption{Excess specific free energy as a function of temperature.}
-    \label{F-T}
-\end{figure}}
 
@@
 Figure \ref{Shs-eta} shows the computed excess entropy as a function of density.
+It is worth noting that this plot has been interpolated; the thermodynamic integration calculations tend to be very lengthly for larger cells and, in the pursuit of both time and CPU time, we chose to simulate only specific filling fractions.
+The intervening points have been filled in using the Carnahan-Starling equation of state.
 We see that both the first and second doubling regime closely match the Carnahan-Starling prediction for the entropy.
@@
 \end{figure}
-We see in Figure~\ref{F-eta} laregly positive free energies because $F = U -TS$; as the density increases, the entropy becomes more negative.
+We see in Figure~\ref{F-eta} laregly positive free energies because $F = U -TS$; as the density increases, the excess entropy becomes more negative.
 For a given temperature, the increasingly negative entropy results in a positive free energy.
@@
 If the decomposition procedure outlined previously proves successful, the absolute free energy ought to obey known physics about the system.
-We particularly care about identifying liquid-vapor coexistence within the fluid. \\
+We showcase here preliminary, and by no means concrete, results for a calculation of interest.\\
 
+One of the more well-known uses of SAFT is generating curves that demonstrate under what conditions both the liquid and gas phase are present in the fluid.
+Identifying the densities at which the system exhibits liquid-vapor coexistence will allow us to partially determine whether the decomposition method presented here has merit.
 Finding coexistence requires use of the grand free energy $\Phi$.
-Liquid and vapor are simultaneously present in the fluid at densities that cause local minima in the grand free energy.
-We know that the grand free energy has a guaranteed minimum at the origin; we then only need to identify the other minimum.
+Liquid and vapor are simultaneously present in the fluid at densities that have local minima in the grand free energy.
+We know from theory that the grand free energy has a guaranteed minimum at the origin\cite{theoryofcrits}; we then only need to identify the other minimum.
 \\
@@
 The chemical potential $\mu$ is a somewhat unintuitive quantity that, for our purposes, we will treat as a free parameter.
-Tweaking the value of the chemical potential will allow us to easily identify second minima in the grand free energy, and thereby the densities of the liquid and vapor present in the system.
+Tweaking the value of the chemical potential will allow us to identify second minima in the grand free energy, and thereby the densities of the liquid and vapor present in the system.
 \begin{figure}[H]

### 2016-06-01 papers/thesis-vischer/Thesis.tex
added last citations, thesis should be finished product Tests skipped!

@@
 \begin{document}
+\thispagestyle{empty}
 
@@
 \subsection{Motivation}
-Fluids have long garnered interest[verlet, van der waals, move Th.
-o'crits], particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
+Fluids have long garnered interest\cite{verlet67}\cite{vanderwaals}, particularly among computationally-minded researchers, as an area in which current models are unequipped to robustly predict behavior across multiple regimes.
 A largely troublesome \cite{white2001global} regime is near the so-called ``critical point'' where liquid and vapor become indistinguishable and, more importantly, fluctuations of fluid density at large length scales occur.
@@
 where $\eta$ is the filling fraction, $N/V$.
-The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit \cite{cs-eos}.
+The Carnahan-Starling equation of state was developed to model hard spheres at low densities in the thermodynamic limit \cite{cseos}.
 We are very much concerned with cells significantly far from the thermodynamic limit.
@@
 Each $F^{\text{abs}}_i$ can be calculated using a combination of different Monte Carlo methods, namely Metropolis-Hastings Monte Carlo and thermodynamic integration.
-A graphical, two-dimensional representation of this process can be shown in Figure ~\ref{GRTG-demo}.\\
+A graphical, two-dimensional representation of this process is shown in Figure ~\ref{GRTG-demo} for the case of $L=L_0$ with a single ball in the original cell.\\
 
@@
      \end{tikzpicture}}
-\caption{2-D Demonstration of MC-GRGT doubling method.}
+\caption{2-D Demonstration of doubling method.}
 \label{GRTG-demo} 
@@
 \subsubsection{Metropolis-Hastings}
-Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving \cite{metro-hastings}.
+Metropolis-Hastings Monte Carlo methods eschew the true randomness of unbiased Monte Carlo by taking the current configuration into account before moving \cite{metrohastings}.
 If we define the ensemble of states of the system in by a vector column $\mathbf{q}$, and the resulting ensemble of states after a step by $\mathbf{p}$, the relationship

### 2016-06-01 papers/thesis-vischer/Thesis.tex
citation fixing, added small content to square well. NOW FINSHED. Tests skipped!

@@
 
-This leads us to choose $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
+This leads us to choose well depth $\epsilon$ and $\sigma$ as natural scales for the quantities of our simulation.
+For reference, recent\cite{clarkwater} work estimates that the best square well approximation for water is a sphere radius of $\sigma = 3.03 \AA$, a well depth of $\epsilon = 21.5433 \text{m eV}$, and a range of interaction of $\lambda = 1.7889$.
 We then define dimensionless temperatures $T/\epsilon$, energies $E/\epsilon$, and distances $r/\sigma$.

### 2016-08-15 papers/thesis-scheirer/final/project report/Project_tex/chapters/chapter04.tex
added folder containing near final versions of all scripts and a working copy of the project paper

\section{Overview}

In this chapter, I wish to highlight some of the fun computational techniques I employed while working towards implementing SnAFT and GRG.

\section{Common Tangent}

Consider the generic function

\begin{equation}
g(x) = x^4 - 4 x^3 - 5 x^2
\end{equation}

It is possible to draw a straight line, that only touches the function at two points and is tangent to the function at both points.  Such a line is call a common tangent line and is illustrated in figure (INSERT FIGURE REFERENCE HERE)

\begin{figure}[h]
	\centering
	\includegraphics[scale=0.5]{generic_common_tangent.png}
	\caption{The cyan line shows the common tangent line of the generic function of x shown in orange.  The black and red dots are the locations of the two points along the generic function that the common tangent is tangent to.}
	\label{fig:generic_common_tangent}
\end{figure}

Mathematically, we can construct this common tangent by constructing two constraints:

\quad The derivative at the right location $x_L$ is equal to the derivative at the left location $x_R$.

\begin{equation}
\frac{dg(x)}{dx}\Bigr|_{x_{\text{L}}} = \frac{dg(x)}{dx}\Bigr|_{x_{\text{R}}}
\end{equation}

\quad The tangent line is a linear line, thus the average slope between the two points is the same as the value of the derivative at either of the two points.

\begin{equation}
\frac{\Delta y}{\Delta x} = \frac{dg(x)}{dx}\Bigr|_{x_{\text{L}}}
\end{equation}

As it turns out, our Helmholtz free energy per volume has a similar shape to the generic function outlined above.  So lets take a look at what happens when we demand our two constraints for finding a common tangent.

Recall our Helmholtz free energy per volume is just one Legendre transform away from internal energy, leaving us with

\begin{equation}
df = - S \, dT - P \, dV + \mu \, dn
\end{equation}

If we hold volume and number density constant

\begin{equation}
\left( \frac{df}{dn} \right)_{T,V} = \mu
\end{equation}

So our first constraint in finding common tangents for our Helmholtz free energy per volume leads us to

\begin{equation}
\left( \frac{df}{dn}\Bigr|_{n_{\text{L}}} \right)_{T,V} = \left( \frac{df}{dn}\Bigr|_{n_{\text{R}}} \right)_{T,V}
\end{equation}

\begin{equation}
\mu_{\text{L}} = \mu_{\text{R}}
\end{equation}

Thus the first constraint tells us that our chemical potentials at both points must be equal.

Now lets look at the second constraint for finding the common tangent.  Consider the Helmholtz free energy.

\begin{equation}
dF = - S \, dT - P \, dV + \mu \, dN 
\end{equation}

We can find the pressure with

\begin{equation}
\left( \frac{dF}{dV} \right)_{T,N} = - P
\end{equation}

But we want to work with free energies per volume.

\begin{equation}
F(T,V,N)  \rightarrow \frac{F}{V}(T,\frac{N}{V}) \rightarrow F(T,V,N) = f(T,\text{n}) \cdot V
\end{equation}

Thus our pressure is

%\begin{equation}
%-P = \left( \frac{df(T,\text{n}) \cdot V}{dV} \right)_{T,N} \rightarrow -P = f(T,\text{n}) + V \left( \frac{df(T,\text{n})}{dV} \right)_{T,N}
%\end{equation}
%
%
%\begin{equation}
%- P = -P = f(T,\text{n}) + V  \frac{df(T,\text{n})}{d\text{n}} \frac{d\text{n}}{dV}
%\end{equation}
%
%
%\begin{equation}
%-P = f(T,\text{n}) - V \, \mu \, \frac{N}{V^2}
%\end{equation}

\begin{equation}
-P = f(T,\text{n}) - \mu \, \text{n}
\end{equation}

Going back to the second constraint

\begin{equation}
\frac{\Delta y}{\Delta x} = \frac{dg}{dx} \rightarrow  \frac{f(\text{n}_R) - f(\text{n}_L)}{\text{n}_R-\text{n}_L} = \frac{df}{dn} = \mu
\end{equation}

Using the expression we found for pressure in equation 4.11, the following pops out

\begin{equation}
P_L = P_R
\end{equation}

Which says our pressure at both points must be the same.

Finally, notice how the two constraints for the common tangent of the Helmholtz free energy lead to an equilibrium state between liquid and vapor!  So all we need to do is find the two densities for which a common tangent exists in our Helmholtz free energy and we found out liquid and vapor equilibrium locations.

The next subsection looks at how we computationally do this.

\subsection{fsolve}

In python, I use the root finding function found in the scipy.optimize library.  For a given free energy, I have fsolve find the roots of the two constraints described in the common tangent section.  Unfortunately, some pre-knowledge must exists to make an initial guess as to what the solutions are in order for fsolve to find the correct locations.  Luckily, it doesn't take too much time to quickly guess two solutions for fsolve to work.

Remember our Helmholtz free energy per volume is for a fixed temperature.  To find the common tangent for all temperatures I use the solution from fsolve as an initial guess of the next temperature.

\section{Overflow}
Explain dealing with large numbers... Emax was factored out then reintroduced.

\section{Reuse}
RG calculates free energy, then saves data.  Then loads and reuses data.

Provide computation methods and data of what I have done so far.

This is how you can refer to fig \ref{fig:liq vap co}. % Note: Have to build the latex document twice for it to work.

### 2016-08-17 papers/thesis-scheirer/final/project report/Project_tex/chapters/chapter01.tex

\begin{section}{Overview}

Matter exists in many different phases.  These phases can be thought of as the states of matter, each with distinctive properties.  For example, fluids can exist in the vapor, liquid, or solid state.  Transitions between phases can be characterized by how the properties of the matter quantitatively change while transitioning between phases.  For a fluid, the property of interest is density, and it is discontinuous when transitioning between liquid and vapor phases while below the critical point.  When the property of interest is discontinuous while transitioning between phases, it is described as a $1^{st}$ order phase transition.  Phase transitions that involve a continuous behavior of a property are $2^{nd}$ order phase transitions, such as density for supercritical fluids.  Of interest here, is to be able to accurately describe the liquid-vapor phase coexistence curve as shown by the cyan curve in figure (INSERT FIG REF HERE).

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.5]{p_vs_t_h20.png}
	\caption{Shown above is a typical pressure vs temperature curve for $H_2O$.  The liquid-vapor coexistence curve is shown in cyan, which starts from the triple point in purple and terminates at the critical point in red. }
	\label{fig:p_vs_t_h20}
\end{figure}

In this chapter, I hope to provide sufficient general motivation to help clarify why anyone would consider applying computational statistical physics to predict phase coexistence curves.  I then introduce how some basic models are used when attempting to determine these phase coexistence curves.  While highlighting the early models, I hope to illustrate the underlying shortcoming of them.  This will lead us into chapter $2$ which will focus explaining the details and limitations of Statistical non-Associating Fluid Theory (SnAFT).  Chapter $3$ will introduce how Generalized Renormalization Group Theory (GRGT) can be used in conjunction with SnAFT to overcome the limitations of previous models.  In chapter $4$ I will present my computational code along with pretty figures and results that will highlight much of what was discussed in previous chapters.  Chapter $5$ will be a reflection and discussion of my research with hints as to what can be done going forward.

\end{section}

\begin{section}{Motivation}

The chemical industry relies on accurate predictions of phase coexistence for fluids in order to ensure safe, efficient, and profitable operations.  Models predicting phase coexistence far from criticality have been well established for over a century [REFERENCE?].  More recently it has been of interest to accurately predict phase coexistence near the critical point of fluids because of the useful applications of supercritical fluids.  Our everyday experience with supercritical fluids is limited if not seemingly non-existent.  However, the use of supercritical fluids spans a vast majority of applications directly and indirectly related to our lives.  For instance, super critical carbon dioxide is a non-toxic replacement for the conventional tetrachloroethylene used for dry cleaning.  Supercritical carbon dioxide is also used in the decaffeination process of coffee beans.  For both examples, the pure supercritical fluid is used because of its miscibility.  Carbon dioxide enhanced oil recovery also utilizes this miscible property of the supercriticial fluid.  The process of the oil recovery results in a complicated mixture of fluids all under extreme pressures and temperatures.  Thus the supercritical fluid is no longer a pure fluid.

For both pure fluids and mixtures, phase coexistence curves can be readily found through experiments focusing on macroscopic quantities like pressure, temperature, and density.  Consider radiator fluid, a mixture, which contains both water and ethylene glycol.  It can be shown experimentally that the phase coexistence depends on the ratio of the water to ethylene glycol in radiator fluid.  If the radiator fluid was mixed at a different ratio of water to ethylene glycol, another experiment must be done in order to find this new mixture's phase coexistence curves.  What if we did yet another ratio of water to ethylene glycol?  Many experiments would need to be done in order to get a wide range of ratios.  On the other had, we can do one experiment at a fixed ratio of water and ethylene glycol, then use information about this experiment to computationally determine what will happen at all ratios.  Basically the computational model depends on arbitrary parameters which can be fit empirically to recreate the phase coexistence at the fixed ratio which the experiment was carried out.  These parameters can then be used to find the phase coexistence at all ratios.  The benefit to this method is that the phase coexistence of mixtures can be explored for a near continuum of ratios, compared to the discrete nature of using only experiments to determine phase coexistence curves.

It should be noted that simulations can also predict phase coexistence curves just as well as actual experiments for a wide variety of fluids.  The simulations are of two overarching categories; molecular dynamics, and Monte Carlo methods.  Molecular dynamics uses equations of motion for each molecule to collect information about thermodynamic properties.  Monte Carlo methods also extract thermodynamic properties by looking at the changes in energy that could occur if the fluid configuration is altered by moving a single molecule within the system.  With only a few hundred molecules in a system, both molecular dynamics and Monte Carlo methods can produce accurate results of phase coexistence far from the critical point.  In theory, both could accurately describe phase coexistence near the critical point if enough molecules are incorporated into the simulations.  As it turn out, increasing the number of molecules to a sufficient number, to incorporate near critical point information, dramatically affects the time it takes to run the simulations.  The time becomes on the order of somewhere between, not practical and ridiculous.

Computational statistical physics bridges the gap between experiment and simulations, by incorporating relevant physical models in conjunction with statistical thermodynamics to describe fluid properties.  This allows us to come up with clever models, which have physical significance, that can better describe fluid behavior near the critical point in computation time that is feasible.

\end{section}

\begin{section}{Model}

Phase coexistence inherently depends on the model used.  For example the ideal gas equation of state predicts a gas all the way down to absolute zero.  The ideal gas has the familiar form:

\begin{equation}
pV=Nk_bT
\end{equation}

  This clearly is wrong since oxygen condenses to a liquid at roughly 90K.  The first model with the ability to predict phase coexistence was the van der Waals equation of state, which has the form:

\begin{equation}
\left( p + \left( \frac{N}{N_A} \right)^2 \frac{a}{V^2} \right) \left( V - \frac{N}{N_A}  b \right) = N k_B T
\end{equation}

The van der Waals equation of state is considered to be a cubic equation of state, since expanding it leads to a lingering $V^3$ term.  Many other cubic equation of states have been proposed [REFERENCE?].  Unfortunately they all suffer from the same limitation of van der Waals equation of state; they are unable to accurately predict phase coexistence near the critical point.

At this point, it would be wise to ask the question why all of these models fail at accurately describing phase coexistence near the critical point.  Perhaps the largest contribution to their downfall stems from the appearance of density fluctuations on large length scales when near the critical point.  A fantastic example of these large density fluctuations can be seen in the phenomena known as critical opalescence, where density fluctuations are large enough in the fluid to scatter visible light.

The van der Waals model can also be considered a mean field theory model, because it assumes that the density of the fluid is a constant.  Thus, density fluctuations are not accounted for in the van der Waals model, and many other cubic equation of state models.  So it would seem that the absence of incorporating density fluctuations in a model yields results that are not accurate near the critical point.

In the following chapter, I will discuss the model known as Statistical non-Associating Fluid Theory (SnAFT), which does a much better job near the critical point than cubic equations of state because it attempts to account for local density fluctuations.  However, as one might imagine, local approximations are insufficient when density fluctuations occur at all length scales as the system nears the critical point; therein enter Generalized Renormalization Group Theory, outlined in chapter 3.

\end{section}

### 2017-01-05 papers/pair-correlation/figs/plot-path-triplet-contact.py
changes to enable compilation on debian stretch

@@ -128,7 +128,7 @@ twod_plot.set_ylim(-rmax, rmax)
 #fig.subplots_adjust(hspace=0.001)
 
 twod_plot.set_aspect('equal')
-g3mc = read_triplet(ff, 'mc')[:, center/dx:-1]
+g3mc = read_triplet(ff, 'mc')[:, int(center/dx):-1]
 rpoints = len(g3mc[:,0])
 zpoints = len(g3mc[0,:])
 r = arange(0, rpoints*dx, dx)

### 2017-01-05 papers/thesis-scheirer/final/project report/Project_tex/abstract.tex

\thispagestyle{plain}
\begin{center}
	\Large
	\textbf{Exploring Phase Equilibrium with Statistical non-Associating Fluid Theory:}

	\vspace{0.4cm}
	\large
	A Generalized Renormalization Group Theory Approach.

	\vspace{0.4cm}
	\textbf{Ryan M. Scheirer}

	\vspace{0.9cm}
	\textbf{Abstract}
\end{center}

Through the use of computational modeling we seek an accurate prediction of liquid-vapor coexistence curves.  At the moment, phase equilibrium far from the critical point can be modeled using statistical non-associating fluid theory (SnAFT).  However, our goal is to create better phase coexistence curves near the critical point by incorporating generalized renormalization group theory (GRGT) with SnAFT.  Within GRGT there is an arbitrary parameter "L".  We seek to fit GRGT to SnAFT far from the critical point in such a way that "L" will no longer be fit to the critical behavior.

### 2017-01-05 papers/thesis-scheirer/final/project report/Project_tex/chapters/chapter02.tex

%\pagestyle{plain}   %Use if you do not want the fancy headers and footers... 
					 %({plain} -> just page number, {empty} -> nothing)...
					 % this will change all pages that come after this tex file...
					 % so make sure to reset the pagestyle to {fancy} at the end of this file

...Splain stuff here...

\section{Introduction}

Recall, our interest lies in the ability to be able to accurately describe the liquid-vapor phase equilibrium conditions for fluids.  As mentioned in chapter 1, many models fall short of an accurate description of phase equilibrium for fluids near the critical point.  Incorporating finer details about the interactions within the fluid will result in better experimental agreement.  SnAFT is one such model which modifies the Helmholtz free energy to include these finer details for interactions.  In this chapter we will explore some of the important features of (SnAFT).

Naturally, before tackling real fluids, we will do what physics teaches us best, simplify the system into a more digestible problem.  We will simplify real fluids, like water, by considering all molecules as spherical interacting objects, subjected to a square well potential as shown in figure (INSERT FIGURE REF HERE).  This type of fluid is known as the square well fluid.

\begin{figure}[H]
	\centering
	\includegraphics[scale=0.2]{square_well_fluid_potential.png}
	\caption{Shown above is the intermolecular pair potential for the square well fluid.  Each sphere has a diameter of $\sigma$.  The interaction range can be adjusted by varying the a scaling factor, $\lambda$.  The strength of the interaction is adjusted by $\epsilon$}
	\label{fig:p_vs_t_h20}
\end{figure}

Mathematically, the shape of the square well potential is described by

\[
  \phi(\textbf{r};\lambda) = \left\{\def\arraystretch{1.2}%
  \begin{array}{@{}c@{\quad}l@{}}
    \infty & \text{if} \indent |\textbf{r}| \leq \sigma \\
    - \epsilon & \text{if} \indent \sigma < |\textbf{r}| \leq \lambda \sigma \\
    0 & \text{if} \indent |\textbf{r}| > \lambda \sigma \\

  \end{array}\right.
\]

where $\textbf{r}$ is the separation distance between the center of each spherical molecule of diameter $\sigma$.  The range of interaction is adjusted by varying the parameter $\lambda$, while the strength of the interaction is characterized by $\epsilon$.

To summarize what we have thus far;  we are looking at a system of particles that are interacting with each other, described by the square well potential, all contained in a region of constant volume.  Since the particles are in random motion, the interaction energy will constantly be changing as particles move closer and further away from each other.  Thus the total energy of our system will fluctuate.  Now we can make one more simplification, by surrounding our system in a heat bath, keeping the temperature constant.  With this physical model in mind, it is naturally ideal to use the canonical ensemble to statistically analyze our system.

The classical partition function for a canonical ensemble containing $N$ indistinguishable spheres is the following,

\begin{equation}
Z = \frac{1}{h^{3N} N!} \int d\textbf{r}^N d\textbf{p}^N e^{-\beta \mathcal{H}_N(\textbf{r}^N,\textbf{p}^N)}
\end{equation}

where $h$ is Plank's constant and $\mathcal{H}_N$ is the classical $N$ sphere Hamiltonian with the functional dependence of  $(\textbf{r}^N,\textbf{p}^N)$, the $3N$ coordinates of position and momentum respectively.  Note that the $N!$ takes care of over counting, and Plank's constant is there to ensure our partition function is dimensionless.

With our square well potential, our classical $N$ sphere Hamiltonian takes the form

\begin{equation}
\mathcal{H}_N = \sum^{N}_{i}  \frac{| \textbf{p}_i|^2}{2m} + \frac{1}{2}  \sum^{N}_{ij} \phi_{ij} ( | \textbf{r}_j - \textbf{r}_i | ) 
\end{equation}

where the first term is the kinetic energy $K_N$, which is only a function of momentum.  The second term is the potential $\phi_N$.  Note that our potential is dependent on the separation distance between two spheres, yielding in the square well coefficient $\phi_{ij}$.  To simplify the Hamiltonian, we write it as

\begin{equation}
\mathcal{H}_N = K_N + \phi_N
\end{equation}

With our Hamiltonian defined, we can update our partition function.

\begin{equation}
Z = \frac{1}{h^{3N} N!} \int d\textbf{r}^N d\textbf{p}^N e^{-\beta K_N + \phi_N }
\end{equation}

Expanding the exponential, and realizing that we are integrating separable momenta for $N$ indistinguishable spheres, the partition function is simplified.

\begin{equation}
Z = \left( \int d\textbf{r}^N e^{\beta \phi_N} \right)  \left( \frac{1}{h^{3N} N!}  \left[ \int d\textbf{p} e^{-\beta K } \right]^N \right)
\end{equation}

Notice the strange grouping of terms above.  This sets us up for a clever trick; multiplying by a fancy one, in the form of volume divided by volume, $\frac{V^N}{V^N}$.  Leading us to a partition function of the form

\begin{equation}
Z = \left( \frac{1}{V^N} \int d\textbf{r}^N e^{\beta \phi_N} \right)  \left( \frac{V^N}{h^{3N} N!}  \left[ \int d\textbf{p} e^{-\beta K } \right]^N \right)
\end{equation}

Where we can now define the following:  The excesses partition function $Z_{\text{exc}}$, and the ideal gas partition function $Z_{\text{id}}$.

Finally our total partition function looks like

\begin{equation}
Z = (Z_{\text{exc}})(Z_{\text{id}})
\end{equation}

The momentum of the spheres are well behaved functions, giving rise to an analytical expression for the idea gas partition function which will be explored in section 2 of this chapter.  Believe it or not, even with our simple square well potential, there is no analytical expression for the excess partition function.  The integral that shows up in the excess partition function is often referred to as the configurational integral;  it is here where much work is done trying to approximate the integrand into a form for computational methods to take care of.

If we rely on Boltzmann, we can take things one step further and write our total Helmholtz free energy as

\begin{equation}
A = -N k_B T \, \text{ln}(Z_{\text{exc}}Z_{\text{id}})
\end{equation}

Expanding the total Helmholtz free energy looks like

\begin{equation}
A = -N k_B T \, \text{ln}(Z_{\text{exc}}) -N k_B T \, \text{ln}(Z_{\text{id}})
\end{equation}

Simply redefining the first and second terms in the total Helmholtz free energy above gives us

\begin{equation}
A =  A_{\text{exc}} +  A_{\text{id}}
\end{equation}

Finally, we are at a spot where we can introduce Statistical Associating Fluid Theory (SAFT).  Recall, there exists no analytical solution for the excess partition function.  Likewise, we are out of luck if we look for an analytical expression for the excess Helmholtz free energy.  SAFT is a model which attempts to split this excess Helmholtz free energy up into physically sound components, which we will describe in the following sections.  For completeness, SAFT writes the total Helmholtz free energy as

\begin{equation}
A =  A_{\text{mono}} +  A_{\text{chain}} +  A_{\text{assoc}} + A_{\text{id}}
\end{equation}

As seen in the above expression for the SAFT Helmholtz free energy, the excess Helmholtz free energy is split up into three terms: mono, chain, and association.  The chain free energy is the contribution due to $m$ number of spheres forming a chain.  The association free energy is the contribution from chains of spheres interacting with one another via binding sites.  Going forward we will not consider the chain and association free energy terms, thus the creation of Statistical non-Associating Fluid Theory (SnAFT).  Forthermore, we will split the mono term into two parts, the hard sphere ($A_{\text{HS}}$) free energy and the dispersion ($A_{\text{disp}}$) free energy.  So now, our final definition for how SnAFT describes the total Helmholtz free energy is

\begin{equation}
A = A_{\text{id}} +  A_{\text{HS}} +  A_{\text{disp}}
\end{equation}

In the following sections, we will explore how each term is constructed based off of physical arguments.

\section{Ideal Gas Free Energy}

Recall the idea gas partition function

\begin{equation}
Z_{\text{id}} =    \frac{V^N}{h^{3N} N!}  \left[ \int d\textbf{p} e^{-\beta K } \right]^N 
\end{equation}

The kinetic energy is a function of momentum,

\begin{equation}
K = \frac{| \textbf{p}|^2}{2m}
\end{equation}

which means the integral that shows up in the idea gas partition function is just a Gaussian integral.  After integrating, the idea gas partition function has the analytical form

\begin{equation}
Z_{\text{id}} = \frac{V^N}{N! \Lambda^{3N}}
\end{equation}

where $\Lambda$ is the thermal de Broglie wavelength

\begin{equation}
\Lambda = \sqrt{\frac{2 \pi \hbar^2}{m k_B T}}
\end{equation}

Now our ideal gas free energy is

\begin{equation}
A_{\text{id}} = - N k_B T \text{ln} \left(  \frac{V^N}{N! \Lambda^{3N}}  \right)
\end{equation}

With the help of Stirling's approximation, our idea free energy has the form

\begin{equation}
A_{\text{id}} = N k_B T [ \text{ln} ( \text{n} \Lambda^3) -1 ]
\end{equation}

where $\text{n}$ is the number density of spheres, $\frac{N}{V}$.

\section{Thermodynamic Perturbation Theory}

I left out a vital piece of information back in section 1 of this chapter when I mentioned that SnAFT splits the mono free energy term into a hard sphere and dispersion term.  More formally, a high temperature perturbation theory is utilized in order to expand the mono term in terms of inverse temperature, $\beta$.

\begin{equation}
A_{\text{mono}} = A_{\text{HS}} + \beta a_1 + \beta^2 a_2 + ...
\end{equation}

Here, the hard sphere system is taken as the known system, while the attractive potential part of the square well potential $- \epsilon$ acts as a perturbation.  The second two terms in the expansion above are what encompass what we are calling the dispersion free energy.

In the next subsections we will take a closer look at the hard sphere and dispersion terms.

\subsection{Hard Sphere Free Energy}

The hard sphere free energy is found via a Fundamental Measure Theory, with the results highlighted in Huges' paper (INSERT REFERENCE HERE).  In her paper, Huges' splits the hard sphere free energy into three terms that are added together and scaled by $N k_B T$ to construct the total hard sphere free energy.  The three terms are as follows:

\begin{equation}
\Phi_1 = \text{n} \, \text{ln} (1-\eta)
\end{equation}

\begin{equation}
\Phi_2 = \frac{3 \text{n} \eta}{(1-\eta)}
\end{equation}

\begin{equation}
\Phi_3 = \frac{\text{n}[\eta + (1-\eta)^2 \, \text{ln}(1-\eta)]}{(1-\eta)^2}
\end{equation}

The resulting hard sphere free energy then takes the form

\begin{equation}
A_{HS} = N k_B T [ \Phi_1 + \Phi_2 + \Phi_3 ]
\end{equation}

\subsection{Dispersion Free Energy}

Recall we defined the dispersion free energy as the first and second correction term to the reference hard sphere system,

\begin{equation}
A_{\text{disp}} = \beta a_1 + \beta^2 a_2
\end{equation}

The term $a_1$ can be thought of as the mean attractive energy, given by

\begin{equation}
a_1 = 2 \pi \text{n} \int^{\sigma}_{\infty} r^2 \phi(r) g^{\text{HS}}(r) dr
\end{equation}

where $r$ is the separation distance between the center of the interacting sphere, $\sigma$ is the sphere's diameter, and $g^{\text{HS}}$ is the pair correlation function.  Great care was taken into consideration when attempting to determine the pair correlation function.  A good approximation is to use an effective filling fraction, $\eta_{\text{eff}}$ combined with the Carnahan and Starling equation of state.  This yields a pair correlation function roughly equal to

\begin{equation}
g^{\text{HS}}_{\text{eff}} \approx \frac{\left( 1 - \frac{\eta_{\text{eff}}}{2} \right)}{(1-\eta_{\text{eff}})^3}
\end{equation}

Notice the lack of separation distance in this definition of the correlation function.  This greatly simplifies our mean attractive energy since we do not need to integrate over it.  Using our square well potential, the mean attractive energy term reduces down to

\begin{equation}
a_1 = - 2 \, \pi \, \text{n} \, g^{\text{HS}}_{\text{eff}} \, \epsilon \, \int^{\lambda \sigma}_{\sigma} r^2 \, dr
\end{equation}

\begin{equation}
a_1 = - 2 \, \pi \, \text{n} \, g^{\text{HS}}_{\text{eff}} \, \epsilon \, \sigma^3 (\lambda^3 -1)
\end{equation}

Some algebraic manipulation and we get

\begin{equation}
a_1 = -4 \, \eta \, \epsilon \, (\lambda^3 -1) g^{\text{HS}}_{\text{eff}}
\end{equation}

Moving on to the second correction term, $a_2$ is a bit more difficult.  Hand waving tells us that we can consider $a_2$ as a term that account for fluctuations of energies.  If a local group of spheres are within each others interaction well, the fluid would look somewhat compressed in that region.  So the $a_2$ term accounts for these local compression regions, i.e. local density fluctuations. It takes the form

\begin{equation}
a_2 = \frac{1}{2} \epsilon \, \kappa^{\text{HS}} \, \eta \frac{\partial a_1}{\partial \eta}
\end{equation}

where $\kappa^{\text{HS}}$ is the isothermal compressibility.

\begin{equation}
\kappa^{\text{HS}} = \frac{(1-\eta)^4}{1+4 \eta + 4 \eta^2}
\end{equation}

Putting it all together, our dispersion free energy term looks like

\begin{equation}
A_{\text{disp}} = - \beta \, N k_B T \left( 4 \, \eta \, \epsilon \, (\lambda^3 -1) g^{\text{HS}}_{\text{eff}} \right) + \beta^2 \, N k_B T \left( \frac{1}{2} \epsilon \, \kappa^{\text{HS}} \, \eta \frac{\partial a_1}{\partial \eta} \right) 
\end{equation}

%\pagestyle{fancy}   % Reset all pages after this file to fnacy headers

### 2017-01-05 papers/thesis-vischer/abstract.tex

@@
 \end{center}
-
+Analytic models using renormalization methods to approximate the free energy of fluids have been developed, but there is currently no way to investigate the intermediate steps of the approximation.
+We present a computational method to determine the free energy contributions of each length scale by exploiting both the behavior of the critical fluid and the periodicity of Metropolis-Hastings Monte-Carlo simulations.
+We further detail a method by which to compute the “absolute” free energy, including both the ideal gas free energy and the excess free energy.
+We further present an application of this method to the well-research square-well liquid near the critical region.
+We find that the square-well liquid is well-characterized by a base cell length of $L=\sqrt2 \sigma$ by establishing liquid-vapor coexistence between scaled temperatures $0.5<T\leq0.75$.
 \clearpage

### 2017-05-09 src/Monte-Carlo/square-well.cpp
tweak tmi and toe algorithms to be more clever about statistics and believing the DOS

@@ -1261,21 +1261,41 @@ op       difference is that we compute the diffusivity here *directly*
 // update the weight array using transitions
 void sw_simulation::update_weights_using_transitions() {
   double *ln_dos = compute_ln_dos(transition_dos);
-  // Flatten the ln_dos at energies lower than the "converged-to state",
-  // since we don't trust ln_dos values down here.  The converged-to
-  // state is defined as where we have 10 pessimistic samples, but we
-  // may want to revisit this later!  FIXME
-  int converged = converged_to_state();
-  for (int i=converged; i<energy_levels; i++) {
-	  ln_dos[i] = ln_dos[converged];
+  if (false) {
+    // This is the slightly older hokey version, which should be
+    // deleted FIXME, when we are confident the new code actually
+    // works.  Note that we should also delete code declared in a
+    // FIXME below when we delete this.
+
+    // Flatten the ln_dos at energies lower than the "converged-to state",
+    // since we don't trust ln_dos values down here.  The converged-to
+    // state is defined as where we have 10 pessimistic samples, but we
+    // may want to revisit this later!  FIXME
+    int converged = converged_to_state();
+    for (int i=converged; i<energy_levels; i++) {
+      ln_dos[i] = ln_dos[converged];
+    }
   }
   // Above the max_entropy_state we level out the weights.
-  for (int i = 0; i < max_entropy_state; i++) {
+  for (int i = 0; i <= max_entropy_state; i++) {
     ln_energy_weights[i] = -ln_dos[max_entropy_state];
   }
   // Down to the min_important_energy we use the DOS for the weights.
   for (int i = max_entropy_state; i <= min_important_energy; i++) {
+    // FIXME the following is unused and should be deleted when we
+    // delete things based on the above FIXME.
     ln_energy_weights[i] = -ln_dos[i];
+
+    // The following is new code added to make the code more
+    // intelligently conservative when it comes to believing the
+    // density of states when there are poor statistics.
+
+    // the following is ln of 1/sqrt(ps), which is a fractional
+    // uncertainty in our count at energy i.
+    double ln_uncertainty = -0.5*log(pessimistic_samples[i]);
+    double ln_dos_ratio = ln_dos[i] - ln_dos[i-1];
+    ln_energy_weights[i] = ln_energy_weights[i-1] - min(ln_dos_ratio,
+                                                        -ln_uncertainty);
   }
   // At lower energies, we use Boltzmann weights with the minimum
   // temperature we are interested in, except in cases where the

### 2018-02-19 papers/histogram/sad.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommand{\tssays}[1]{{\color{red} [\cyan{Tanner:} \emph{#1}]}}

\begin{document}
\title{Dynamical Stochastic Approximation Monte Carlo}

\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
\author{Michael A. Perlin} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  We present several histogram methods and compare the performance and efficiency at treating the square-well fluid.
\end{abstract}

\maketitle

\section{Introduction}
Over the past several decades a number of advanced Monte-Carlo
simulation algorithms have been developed which calculate the
properties of a thermal system over a range of temperatures.  This
idea began with the original histogram method, which used a single
canonical Monte Carlo simulation to predict properties for nearby
temperatures~\cite{ferrenberg1988new}.  For large simulations this
approach is limited to a narrow temperature range because a single
canonical simulation explores only a small range of energies.  This
led to a variety of ``broad'' (or ``flat'') histogram
methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
  wang2001determining, wang2001efficient, trebst2004optimizing}, which
attempt to explore a wide range of energies.  These methods also
benefit in that they are unlikely to be trapped in a local energy
minimum.

Wang and Landau introduced an algorithm that uses an update
factor and a statistical histogram to compute the density of states of
a given system~\cite{wang2001determining, wang2001efficient}.  While
the method is incredibly powerful, it has several user-defined
parameters.  This can make its application to a variety of systems
something of an art-form since the user needs to determine the ideal
parameters for his particular system.  Also, detailed balance is
violated (although for only short periods of time), yet the method
converges in most situations.  This raises several questions such as
when does the method converges, what rate will the method converge to
the correct density of states, and how to appropriately decide what
parameters to choose so that the algorithm solves the problem in the
best way. Transition Matrix Monte Carlo~\cite{wang1999transition,
swendsen1999transition, fitzgerald2000monte} became an attractive
complimentary simulation algorithm to Wang-Landau in that, if detailed
balance is ensured, the system is guaranteed to converge (something
which is not guaranteed for the Wang-Landau Algorithm).  Unfortunately,
the algorithm can be considerably more difficult to implement when
compared to Wang-Landau: in particular, the infinite temperature
transition matrix~\cite{wang2002transition}.  Also, the time necessary
for the density of states to converge could be considerable
due to the algorithm spending too much time exploring low energy states.
~\tssays{(note to self) The above paragraph could be condensed a bit}

Shell et al.~\cite{shell2003improved, shell2004flat} originially
implemented Wang-Landau Transition Matrix Monte Carlo (WL-TMMC) in an
effort to quickly explore the energies of a given system using WL then
switch over to TMMC to guarantee convergence. Considerable effort has
been spent trying to determine the flatness criteria and cutoff for the
WL portion, but there remains no rigorous systematic way to determine
these parameters~\cite{rane2013monte}.  The algorithm is first run for
several trial simulations and/or `experience' is used to determine the
best parameters for a users particular system of
study~\cite{siderius2013use}.  Around the same time, the convergence
rate for Wang-Landau was being explored~\cite{zhou2005understanding}.
It was found that utilizing modification factors that decrease faster
than $1/t^2$ leads to nonconvergence~\cite{belardinelli2007fast}.  This
lead to the so-called $1/t$-WL algorithm which ensures that excess CPU
time is not wasted by continuing to perfrom calculations once the error
in the density of states becomes
saturated~\cite{belardinelli2008analysis}. An effort to reduce the
number of user defined parameters of WL and formulate a theory for why
the method converges despite detailed balance being violated although
infrequently was also being udertaken.  Liang, Liu, and Carrol began to
consider whether WL could be considered a special case of Monte Carlo
whose convergence could be mathematically
proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
al.~\cite{liang2007stochastic} argued that WL can be considered a form
of Stochastic Approximation Monte Carlo (SAMC).  While SAMC can
guarantee convergence, the method still has a system specific
user-defined variable which makes applying this algorithm to arbitrary
systems difficult.

In our work, we have developed a novel algorithm based on SAMC that
does not require user-defined inputs and therefore should be easily
applicable to a given system.  We call this method SAD (Dynamical
Stochastic Approximation), and will discuss it in detail in the methods
section. We compare it along with several histogram methods including TMMC,
WL, WL-TMMC, and SAMC.

We consider the square-well fluid i.e. a system of particles whose
interactions are governed by a square-well
potential~\cite{singh2003surface, barker2004perturbationSW}.  The
square-well potential is an ideal test-bed as it is the simplest model
that ensures both attractive and repulsive forces are experienced by
interacting particles~\cite{barker1967-SW-perturbation, vega1992phase}.
The potential $U(\textbf{r})$ for such a system is given by
\begin{equation}
 U(\textbf{r})=\begin{cases} \infty &
 \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
 \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
 \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
\end{equation}
where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
reduced range of the potential well, and $\epsilon$ is its depth.

The organization of this paper is as follows: In Section II, we
describe in detail several broad-histogram methods used to calculate
the density of states for the liquid-vapor system.  Section III
contains the results and comparison of the methods applied to the
square-well fluid.  In Section IV, we discuss the performance of the
various methods applied to the model system.

\section{Methods}

Here we employ a variety of broad-histogram methods.  We outline the
general workings of algorithm we developed in detail and summarize
algorithms that can be found elsewhere.  The following algorithms are
introduced and applied to the square-well fluid:  Transition
Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau Transition
Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation Monte-Carlo
Monte-Carlo (SAMC), and Dynamical Stochastic-Approximation (SAD).

\subsection{TMMC Algorithm}
The Transition Matrix Monte-Carlo (TMMC) method as introduced by Fitzgerald et al.
consists of three primary steps.  The first two steps are common to the Metropolis
algorithm while the third makes use of the actual transition
probabilities~\cite{fitzgerald2000monte}.

\subsubsection{define a transition probability}
Assuming that the system under study is in a state denoted old and a
proposed transition is made to a new state denoted new, the probability
is defined as
\begin{align}
  p_{old \rightarrow new} = p_{new \rightarrow old}
\end{align}
where we have allowed the probabilies to be equal for simplicity of presentation.

\subsubsection{define an acceptance probability}
The probability for accepting a transition from an old state to a new one can be
defined
\begin{align}
  p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
  {\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
\end{align}
where $\alpha_{new\rightarrow old}$ is the probability of generating a
new system configuration the old.  For conventional Monte-Carlo moves
$\alpha_{old \rightarrow new} =\alpha_{new\rightarrow old}$ although
for advanced Monte-Carlo moves this would not be the
case~\cite{paluch2008comparing, siepmann1990method}.  The probabily of
observing the system in the old state can be found based on the
ensemble constraints placed on the system.  Fitzgerald et al.
considered the system of interest using the canonical ensemble however
extension to other ensembles is quite natural.
\begin{align}
  \pi_{old} = \frac{e^{-\beta E_{old}}}{Z}
\end{align}
The acceptance probability given for TMMC is idential to that introduced by
Swendsen and Wang~\cite{wang2002transition}.
\begin{align}
  p_{a} = \min\bigg[1,e^{-\frac{\delta\epsilon}{k_{B} T}}\bigg]
\end{align}
\subsubsection{define a bookkeeping step}
The third step in the TMMC algorithm consists of recording the data in a collection
matrix (analogous to the transition histogram matrix).
\begin{align}
  C_{N,\delta} = C_{N,\delta} + p_{a}\\
  C_{N,0} = C_{N,0} +(1 - p_{a})
\end{align} If $\delta=0$, that is, the energy state has not been
changed, then only $C_{N,0}$ is incremented by unity. Fitzgerald et
al.~\cite{fitzgerald2000monte} show that defining the third step this
way results in a uniform improvement over using histogram estimators to
determine the particle number probability distribution (PNPD).

\subsection{Wang-Landau Algorithm}

The Wang-Landau Algorithm developed by Wang and Landau proposes a
random walk in energy space in order to estimate the density of
states~\cite{wang2001efficient}. While methods such as Metropolis
Sampling and Swendsen-Wang cluster
flipping~\cite{swendsen1987nonuniversal} generate a narrow distribution
and require sampling at individual temperatures, the Wang-Landau
algorithm uses a flat histogram and performs a random walk in energy
space to determine the density of states~\cite{wang2001determining,
landau2014guide}. Wang-Landau's major tenant is that when counting the
histogram $H(\epsilon)$, the energy occurences should form a flat
distribution.  The criteria for flat sampling is given by
\begin{equation}
	\frac{\min_{\epsilon} H(\epsilon)}
	{\big\langle H(\epsilon)\big\rangle }
	> \eta
\end{equation} where $\eta$ is usually set to $0.80$ and determines how
many times each energy is sampled relative to the mean number of
visits.  Difficulties can occur with this flatness criteria due to the
fact that some energies on a given energy range might never be
sampled~\cite{haber2014transition}.  Over the past decade much
attention was directed at examining the rate of convergence and error
saturation~\cite{lee2006convergence, belardinelli2007wang}.  The
algorithm is briefly outlined below where $w$ represents the weights:
\begin{equation}
	\ln{w_{k+1}(\epsilon,N)}=\ln{w_{k}(\epsilon,N)}
	+\gamma
\end{equation}
\begin{equation}
	\ln{\gamma_{k+1}}=\frac{u}{2}\ln{\gamma_{k}}
\end{equation}
for $u$ typically greater than 1 and $\gamma$ is the modification factor for
the density of states.  We then accept a new configuration for the density of
states with a probability given by the following
\begin{equation}
	\mathcal{P}(\epsilon_i \rightarrow \epsilon_j)
	= \min[1,e^{\ln{w(\epsilon_i)}-\ln{w(\epsilon_j)}}]
\end{equation}
Zhou and Bhatt explored the convergence of the Wang-Landau algorithm and
determined that choosing a large modification factor $\gamma_0 = e^{4}$ and
rapidly reducing the factor by 10 during each step resulted in reduced statistical
error~\cite{zhou2005understanding}.

\subsection{WL-TMMC Algorithm}

Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
advantage of the ease of implementation and applicability to a variety
of systems, while making use of a transition matrix to address WL
weaknesses such as a limited statistical accuracy plateau which is not
improved with additional Monte-Carlo steps.  This WL version is more
accurate for a given number of simulation steps.

\paragraph{Initializing Wang-Landau:} We begin the simulation by
defining a minimum important energy and a maximum entropy state. In
traditional Wang-Landau the WL factor begins at 1.0 and the factor is
modified by 2.0 until the simulation is cutoff typically around
$10^{-10}$.  Rane et. al. published an implemenation of
WL-TMMC~\cite{rane2013monte} that suggested using a WL factor between 1.0 and
$10^{-2}$ with a cutoff $<10^{-5}$. In our implementation of WL-TMMC,
we set the WL factor to be 1 and the cutoff at $10^{-10}$. A single
sweep is defined as each macrostate being sampled a number of times in
this case 10,000 (Siderius et. al. does 100). Wang-Landau sets a
flatness criteria~\cite{wang2001determining, wang2001efficient,
hatch2015computational, mahynski2017predicting} for the accumulated
histogram usually between 0.75 to 0.90.  For our simulation, it is
sufficient that each macrostate is visited at least once (each
histogram bin has one entry)~\cite{shell2003improved}.  The goal of
this WL initialization is to prefill the transition matrix for the next
portion of the simulation.
\paragraph{Initializing TMMC:} The purpose of prefilling the transition
matrix is apparent because if numerous zeros existed in the collection
matrix, the infinite temperature transition matrix would be
ill-defined.  Sampling over large ranges of density of states would
therefore take an unacceptable length of time to
complete~\cite{shell2003improved, shen2014elucidating}.  The transition
matrix in terms of the collection matrix is given as follows:
\begin{align}
\widetilde{T}_{\infty}(I\rightarrow J) = \frac{C(I,J)}
{\sum_{K} C(I,K)}
\end{align}
We now run the TMMC portion of the simulation.  Siderius does this for
20 additional sweeps (does about 8 sweeps for WL).  We do this until we
reach a specified temperature of interest.
We have not yet implemented a biasing function to control acceptance
rate of trial moves but we probably could/should? at some point.
\begin{align}
  P_{acc}(i\rightarrow j) = \min\bigg[1,\frac{\widetilde{T}_{\infty}(J\rightarrow I)}
  {\widetilde{T}_{\infty}(I\rightarrow J)}\bigg]
\end{align}
this is only used to control actual transitions not update the collection
matrices.

\subsection{SAD Algorithm}
The SAD (Dynamic Stochastic Approximation) is a version of the SAMC
Algorithm that attempts to dynamically choose the modification factor
rather than relying on system dependent parameters such as $t_0$ or
$\gamma_0$.  There is an immediate advantage of such an algorithm where
parameters are chosen independent of system size or type. SAMC can be
applied as a dynamic importance sampling algorithm that uses an update
factor to explore energy space~\cite{liang2007stochastic,
werlich2015stochastic, schneider2017convergence}.  The update factor is
defined in the original implementation~\cite{liang2007stochastic} in
terms of an arbitrary tunable parameter $t_0$.
\begin{align}
\gamma_{t}^{\text{SA}} = \frac{t_0}{\max(t_0,t)}
\end{align}
We argue that it is possible to set this parameter to a system
independent parameter such that $t_0 = N_E^2/t$ where $N_E$ is the
number of energies that have been found.  One of the energy found terms
can be thought of as a change in the entropy $\Delta S$.  The change in
the entropy can be thought of as a slope of the triangle made by
connecting the minimum important energy and the maximum entropy state
divided by the minimum temperature:
\begin{align}
\Delta S = \frac{E_{\text{max}}-E_{\text{min}}}{T_{\text{min}}}
\end{align}
This gives an update factor as follows:
\begin{align}
\gamma_{t}^{\text{SA}} = \frac{N_E\Delta S}{t} =
\frac{N_E}{t}\frac{E_{\text{max}}-E_{\text{min}}}{T_{\text{min}}}
\end{align}
The update factor has the same $1/t$ behavior as the original SAMC
algorithm; however now every time a new energy is discovered during the
course of the simulation, the update factor experiences a jump.
~\jpsays{Add in additional material for SAD!}

\section{Results}

For the square-well fluid, our simulations treat systems with a
well-width of $\text{ww} = 1.30$ and a filling fraction of either
$\text{ff} = 0.10, 0.20$ or $\text{ww} = 0.20$.  The simulation explores the
energy space of the system down to a reduced temperature of
$T_{\text{min}} = 0.2$.  All simulations lead to the minimum important
energy $E_{\text{min}}$ and maximum entropy state $S_{\text{max}}$
being calculated (with the exception of the WL-TMMC method where both
of these parameters are needed a priori).  In the sections below, we
examine systems with different particles numbers and filling fractions.

~\jpsays{Add N50 ff0.1 and N50 ff0.3 and remove sticky-wall!}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-30x4-max-entropy-error.pdf}
  \caption{The average entropy error for each method is shown with
  SAD serving as the `golden'
  standard.}\label{fig:avgS-30x4}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-60x8-max-entropy-error.pdf}
  \caption{The average entropy error for each method is shown with SAD serving as
  the `golden' standard.}\label{fig:avgS-60x8}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-100x10-max-entropy-error}
  \caption{The average entropy error for each method is shown with
    WL-TMMC using a cutoff of 10$^{-10}$ serving as the
    standard.}\label{fig:avgS-100x10}
\end{figure}
\begin{figure}
  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-max-entropy-error.pdf}
  \caption{The average entropy error is shown with TMI acting as the
  `golden' standard.}\label{fig:N50-mean-error}
\end{figure}
\begin{figure}
\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-max-entropy-error.pdf}
  \caption{The average entropy error is shown with TMI acting as the
  `golden' standard.}
\end{figure}

\subsubsection{A $30\times4\times4$ $(N = 25)$ system with a sticky
wall}
In order to determine how the various methods scale with size,
simulations are run for systems with different particle numbers.  For
the $N = 25$ system, a sticky-wall is added to ensure that we see the
liquid and vapor phase separation.

A system is configured with dimensions $x = y = 4\sigma$ and $z =
30\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
yields $N = 25$ particles in the rectangular system.  These particles
are randomly inserted into the system and random moves are proposed
(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
inserted along the $z$-axis at one end to promote phase separation.
~\jpsays{should we include a sticky-wall.py plot or something else or
are we not going to talk about phase separation in this paper?}

We use the entropy error vs iterations as a metric to compare
simulation runtimes.  Simulations that take more iterations to reach
the same entropy error as a different method take longer to converge.
This behavior is seen in Fig.~\ref{fig:avgS-30x4}.
~\jpsays{SAD appears to do pretty well as does WL-TMMC for a cutoff of $10^{-10}$
although this seems to depend on what is used as the `standard'.}

\subsubsection{A periodic system with 20\% filling fraction and $N = 50$}

Figure~\ref{fig:N50-mean-error} shows a system consisting of 50 atoms
are placed in a periodic simple cubic unit cell with lattice constant
chosen to give a filling fraction of 0.20.  These particles are
randomly inserted into the system and random moves are proposed (with
a translation scale $\delta_0 = 0.05$).

\subsubsection{A $60\times8\times8$ $(N = 202)$ system with a sticky
wall}

A system is configured with dimensions $x = y = 8\sigma$ and $z =
60\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
yields $N = 202$ particles in the rectangular system.  These particles
are randomly inserted into the system and random moves are proposed
(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
inserted along the $z$-axis at one end to promote phase separation.

\subsubsection{A $100\times10\times10$ $(N = 525)$ system with a sticky
wall}

A system is configured with dimensions $x = y = 10\sigma$ and $z =
100\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
yields $N = 525$ particles in the rectangular system.  These particles
are randomly inserted into the system and random moves are proposed
(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
inserted along the $z$-axis at one end to promote phase separation.

\subsubsection{A Tanner dimensions? $(N = 500)$ system with no sticky
wall}

A system is configured with dimensions $x = y = z = 19\sigma?$.  A
filling fraction $\text{ff} = 0.30$ is chosen which yields $N = 500$
particles in a cubical system.  These particles are randomly inserted
into the system and random moves are proposed (with a translation scale
$\delta_0 = 0.05$). The boundaries in $x$ and $y$ are periodic while
the $z$-axis is not periodic.  A sticky-wall is not inserted along the
$z$-axis.

\section{Conclusions}

We introduced a new algorithm SAD (which is Stochastic Approximation
with the modification factor dynamically chosen) that effectively
samples the entire energy space for a variety of system sizes.  Also,
SAD proves effective in comparison with other Monte-Carlo methods at
converging to the correct density of states.  The ease of
implementation and the absence of user-defined parameters makes SAD an
ideal algorithm.

~\jpsays{We need about a paragraph here that says exactly how much
better SAD performed than other methods and or how easy it was to
implement (Likely best added once simulations run a little longer).}

We note that SAD should be tested on additional systems, and the
accuracy of the method should be examined.  Also, a rigorous
mathematical justification for our choice of the update factor is
necessary to ensure the most optimal convergence of the algorithm.

\section{Acknowledgement}

J.K. Pommerenck would like to gratefully acknowledge D.W. Siderius and
J.R. Errington for helpful communications regarding the implementation
of WL-TMMC.

% \jpsays{Everything below this is TMI1 and GC-TMMC}

%\subsection{GC-TMMC Algorithm}
%Grand Canonical Transition Matrix Monte-Carlo (GC-TMMC) allows the energy and
%particle number of the system to fluctuate while holding the chemical potential,
%volume, and temperature constant~\cite{chen2001aggregation, chen2002simulating,
%errington2003direct, fenwick2006accurate, fenwick2008direct, errington2005direct}.
%The algorithm boasts a lower statistical noise than conventional GCMC due to its
%incorporation of a transition matrix to estimate ensemble averages~\cite{siderius2013use,
%paluch2008comparing, grzelak2008computation}.  The steps for the algorithm are
%outlined below:

%\subsubsection{Define particle insertion/deletion}
%In the system of consideration, the number of particles $\mathcal{N}$ can change
%by an amount $\delta$.  It is assumed that in the system either no particles can be
%inserted or deleted, a single insertion takes place, or a single deletion takes place.
%This yields a condition $\delta={-1,0,+1}$ for particle insertion.

%\subsubsection{Define an acceptance probability}
%Now it is necessary to define criteria for determining whether a particle insertion/deletion
%is allowed.  This criteria is called the acceptance probability and is denoted as
%\begin{align}
  %p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
%\end{align}
%where $\alpha_{new\rightarrow old}$ is the probability of generating an old system
%configuration from the new.  For conventional Monte-Carlo moves $\alpha_{old \rightarrow new}
%=\alpha_{new\rightarrow old}$ although for advanced Monte-Carlo moves this would not
%be the case~\cite{siepmann1990method}.  The probabilies of actually observing the
%system in the old state is given by
%\begin{align}
  %\pi_{old} = \frac{1}{\Xi}\frac{V^{N_{old}}}{\Lambda^{3N_{old}}N_{old}!}
  %e^{-\beta(\epsilon_{old}-\mu N_{old})}
%\end{align}
%where $\Xi$ is the grand canonical partition function and $\Lambda$ is the De Broglie
%wavelength.

%\subsubsection{Define a Collection Matrix}
%In order to calculate an overall transition probability $\mathcal{P}_{N,\delta}$, we
%must first define how the system procession is to be recorded.  A Collection Matrix
%can be defined, such that for any one move, two elements are updated.
%\begin{align}
  %{C}_{N,\delta} = {C}_{N,\delta} + p_{a}\\
  %{C}_{N,0} = {C}_{N,0} +(1 - p_{a})
%\end{align}
%If $\delta=0$, that is, the number of particles has not been changed, then only
%$\mathcal{C}_{N,0}$ is incremented by unity.  The overal transition probability can
%now be defined as
%\begin{align}
  %\mathcal{P}_{N,\delta} = \frac{{C}_{N,\delta}}
  %{\sum_{j=-1}^{+1} {C}_{N,j}}
%\end{align}

%\subsubsection{Define the Particle Number Probability Distribution}
%The Particle Number Probability Distribution (PNPD) $\Pi(N;\mu,V,T)$ can be found
%by enforcing detailed balance in the system.
%\begin{align}
  %\ln(\Pi_{N+1}) = \ln(\Pi_{N}) + \ln\bigg(\frac{\mathcal{P}_{N\rightarrow N+1}}
  %{\mathcal{P}_{N+1\rightarrow N}}\bigg)
%\end{align}
%This equation will hold as long as N-space can be sampled sufficiently.  In an
%implemented algorithm $N$, $N_{min}$, and $N_{max}$ are specified for macrostate domain
%sampling.

%\subsubsection{Define a biasing function}
%The primary difference between TMMC as defined by Fitzgerald et al. and GC-TMMC as
%pioneered by JR Errington et al. is the use of a biasing function to control the acceptance
%rate of trial moves~\cite{siderius2013use, fitzgerald2000monte}.  This type of
%multi-canonical sampling augments the acceptance probability in the following way
%\begin{align}
  %p_{\eta,old\rightarrow new} = \min\bigg[1,\frac{e^{\eta(N_{new})}}
  %{e^{\eta(N_{old})}}\frac{\alpha_{new\rightarrow old}}
  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
%\end{align}

%Although $p_{\eta}$ is used to control the actual transitions for trial configurations
%in GC-TMMC, $p_{a}$ is still used to update the Collection Matrices.  Ideally the biasing
%function would be given by
%\begin{align}
  %\eta(N) = -\ln\Pi(N;\mu,V,T)
%\end{align}
%Since $\Pi$ is not usually known a priori, the biasing function is usually set to some
%arbitrary constant for initialization.  The goal of this biasing function is to achieve
%uniform sampling (a common feature shared by the flat histogram Monte-Carlo family).

%\subsubsection{Histogram reweighting for the PNPD}
%Once the PNPD has beeen collected for a given chemical potential $\mu_0$, histogram
%reweighting is used to determine thermodynamic properties at other chemic potentials.
%Since the canonical partition function $Z(N,V,T)$ does not depend on the chemical
%potential, we can write
%\begin{align}
%\begin{split}
  %\Pi(N;\mu,V,T) = \frac{e^{\beta\mu N}Z(N,V,T)}{\Xi(\mu,V,T)}\\
  %\ln\Pi(N;\mu,V,T) = \ln\Pi(N;\mu_0,V,T) \\
  %+ \beta N(\mu-\mu_0) + C
%\end{split}
%\end{align}
%where $C$ is a normalization constant independent of the number of particles.

\bibliography{paper}% Produces the bibliography via BibTeX.

\end{document}

### 2018-02-20 papers/histogram/sad.tex

@@
 
-\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
-\author{Michael A. Perlin} \author{David Roundy}
+\author{Jordan K. Pommerenck} \author{Michael A. Perlin}
+\author{Tanner T. Simpson} \author{David Roundy}
 \affiliation{Department of Physics, Oregon State University,
@@
 
-In our work, we have developed a novel algorithm based on SAMC that
-does not require user-defined inputs and therefore should be easily
-applicable to a given system.  We call this method SAD (Dynamical
-Stochastic Approximation), and will discuss it in detail in the methods
-section. We compare it along with several histogram methods including TMMC,
-WL, WL-TMMC, and SAMC.
+In our work, we have developed several novel algorithms that do not
+require user-defined inputs and therefore should be easily applicable
+to a given system.  These methods, each of which will be discussed in
+detail in the methods section, can also be made to satisfy theoretical
+requirements, guaranteeing convergence. These algorithms are Transition
+Matrix Initialization (TMI), Dynamic Stochastic Approximation (SAD),
+and Stochastic Approximation Transition Matrix Monte Carlo (SA-TMMC).
+TMI works by using logical conditionals about changes in the logrithm
+of the density of states to control the weights of the simulation.  SAD
+operates effectively the same as standard SA; however, we claim that
+the user defined parameter can be set independent of the system type or
+size.  SA-TMMC works by performing SA for a system-independent amount
+of time in order to explore the entire energy space.  The simulation
+then switches over to TMMC for the remainder of the simulation.  All of
+these methods have no parameter inputs that depend on the type of
+system or size of system under study.
 
@@
 algorithms that can be found elsewhere.  The following algorithms are
-introduced and applied to the square-well fluid:  Transition
+introduced and applied to the square-well fluid: Transition Matrix
+Initialization (TMI), Transition Optimized Ensemble (TOE), Transition
 Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau Transition
-Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation Monte-Carlo
-Monte-Carlo (SAMC), and Dynamical Stochastic-Approximation (SAD).
+Matrix Monte-Carlo (WL-TMMC), Dynamic Stochastic Approximation
+Monte-Carlo (SAD), and Stochastic-Approximation Transition Matrix
+Monte-Carlo (SA-TMMC).
+
+\subsection{TMI2 Algorithm}
+
+The second version of transition matrix initialization algorithm works
+by looking at the logarithm of the density of states.  We begin at
+high energies ($\epsilon_{S_{\max}}$ the energy that maximizes the
+entropy) and work our way
+downward.
+
+\paragraph{Reaching $T_{\min}$:}
+If the change in the logarithm of the density of states is greater
+than the reciprocal of the minimum temperature, i.e. if
+\begin{align}
+  \ln\frac{D(\epsilon_i)}{D(\epsilon_{i-1})} &>
+  \frac{\epsilon_{i-1} - \epsilon_i}{k_BT_{\min}}
+  \\
+  \frac{D(\epsilon_i)}{D(\epsilon_{i-1})} &>
+  e^{\beta_{\max}(\epsilon_{i-1} - \epsilon_i)}
+\end{align}
+then the algorithm has reached the minimum temperature of interest.
+In this case, we set the log weights at all lower energies using a
+Boltzmann ratio relative to this energy.
+\paragraph{Confident at $\epsilon_i$:} We are confident in our density
+of states at energy $\epsilon_i$ provided two conditions are true.
+\begin{enumerate}
+\item The density of states must be decreasing as the energy decreases
+  from its next-higher energy.
+\item The change in density of states (relative to the next higher
+  energy) must be smaller than a pessimistic estimate of the counting
+  uncertainty at the energy we are examining:
+  \begin{align}
+    \frac{D(\epsilon_i)}{D(\epsilon_{i-1})} < \frac1{\sqrt{N_i}}
+  \end{align}
+  where $N_i$ is the number of ``pessimistic
+  samples'' at $\epsilon_i$, the lower of the two energies under
+  consideration.
+\end{enumerate}
+If both of these are true, we believe we know the density of states at
+this energy well, and the weights are set by
+\begin{align}
+  \ln w(\epsilon_i) = -\ln{D}(\epsilon_i)
+\end{align}
+\paragraph{Reached $T_\text{converged}$:}
+If this is not the case, $\epsilon_i$ is the lowest energy that is
+adequately converged.  We set the weights for energies lower than this
+to a Boltzmann ratio with a temperature determined using the slope of
+a secant line on the $\ln w$ versus $\epsilon$ graph.
+\begin{align}
+  \beta_\text{secant} = \bigg(\frac{\ln{D}(\epsilon_{S_{\max}}) - \ln{D}
+  (\epsilon_\text{pivot})}{\epsilon_{S_{\max}}-\epsilon_\text{pivot}}\bigg)
+\end{align}
+This ``slope'' is a positive quantity.  The weights at energies below
+$\epsilon_\text{pivot}$ are given by
+\begin{align}
+  \ln w(\epsilon) = \ln w(\epsilon_\text{pivot}) -
+  \beta_\text{secant}(\epsilon - \epsilon_\text{pivot})
+\end{align}
+indicating that there are higher weights at lower energies, since
+$\beta$ is positive.  We have one additional caveat: we examine the
+lower energies for which we have samples (``pessimistic'' samples),
+and if we find any energy at which the density of states is greater
+than the inverse of the weight, we adjust $\beta$ to cause it to pass
+through that point.  This ensures that the product of the weight with
+the estimated density of states is always less than or equal one for
+energies below $\epsilon_\text{converged}$.
+
+\subsection{TMI3 Algorithm}
+The TMI$_3$ algorithm is like the TMI$_2$ algorithm, with one
+difference, which is that instead of a secant line being used to
+estimate the converged temperature, we use a tangent line, estimated
+using the lowest two ``converged'' energies.
+
+\subsection{TOE3 Algorithm}
+The TOE$_3$ method desribed by Trebst et.
+al.~\cite{trebst2004optimizing} uses the metric of round trips defined
+by a walker visiting the extreme energies for the density of states.
+The rate for the walker is dependent of the diffusivity $D(\epsilon)$
+at a given energy $E$.  The goal of the method is to maximize the
+number of round trips that the Monte-Carlo simulation makes.  The
+optimized density of states is given by the following:
+\begin{align}
+  n_{w}^{(opt)} = \frac{1}{\sqrt{D(E)}}
+\end{align}
+where unlike in the original paper detailing the method, we compute the
+diffusivity directly from the transition matrix normalized with
+canonical weights.  The optimal ensemble therefore relates the density
+of states to the inverse of the square root of the diffusivity.
+\begin{align}
+  D(E) = \mid (T(\Delta\epsilon)^2 - (T(\Delta\epsilon))^2) \mid
+\end{align}
 
@@
 course of the simulation, the update factor experiences a jump.
-~\jpsays{Add in additional material for SAD!}
+
+\subsection{SA-TMMC Algorithm} We introduce a new algorithm consisting
+of the merger of the Stochastic Approximation (SA) method proposed by
+Liang et al.~\cite{liang2007stochastic} and the TMMC Algorithm with
+suitable adaptations.  While the Stochastic Approximation method has
+recently been implemented in comparison with the Wang-Landau
+Algorithm~\cite{werlich2015stochastic, schneider2017convergence}, it
+seems quite natural to create a new Monte-Carlo method which we will
+call SA-TMMC.  Like in the original implementation, we define an update
+factor where $t_0$ is the number of Monte-Carlo moves and $t$ is the
+current iteration.
+\begin{align} \gamma_{t}^{\text{SA}} =
+\frac{t_0}{\max(t_0,t)}
+\end{align}
+One potential problem of merging SA and TMMC is that SA provides no
+nice transition to the later method due to the lack of a
+user-controlled cutoff.  Also for SA Monte-Carlo, the choice of
+parameter $t_0$ proves difficult particularly for complex systems.
+Finally, we enforce a minimum temperature probability which prevents
+the system from sampling states below the minimum temperature. We
+discuss here how we resolve all of the potential issues we encountered
+when integrating both SA and TMMC.
+
+\subsubsection{Define an adaptive parameter $t_0$}
+It has been shown that if $F_{t}^{\text{SA}}$ satisfies two conditions
+the estimator is proven to converge to the real
+DOS~\cite{liang2006theory, liang2007stochastic}.
+\begin{align}
+\sum_{t=1}^\infty \gamma_{t}^{\text{SA}} = \infty \quad\textrm{and}\quad
+\sum_{t=1}^\infty (\gamma_{t}^{\text{SA}})^\zeta < \infty
+\end{align}
+where $\zeta \in \{1,2\}$.  For practical implementation with TMMC, we
+find it necessary to reset $t_0$ everytime a new energy state is found.
+This ensures that all energy states are sampled before TMMC move
+probabilities begin to be favored.
+
+\subsubsection{Define a statistical weight to control the transition between SA and TMMC}
+Our algorithm can alternate between SA and TMMC on the fly using
+weights to determine the probability that a move between adjacent
+energy states is allowed.  We define the transition probability
+$\text{SA}_{\text{Weight}}$ as
+\begin{align}
+\delta_{\text{TMMC}}^2 =\frac{1}{n_\downarrow}+\frac{1}{n_\uparrow}, \quad\quad
+\delta_{\text{SA}}^2 = \frac{1}{N_E}
+\end{align}
+\begin{align}
+\text{SA}_{\text{Weight}} =\frac{\delta_{\text{TMMC}}^2}{\delta_{\text{TMMC}}^2 + \delta_{\text{SA}}^2}
+\end{align}
+where $N_\text{found}$ is the number of energy states that have been
+found i.e. visited, and $n_{\uparrow\downarrow}$ are the number of
+upward or downward moves between a pair of energy states respectively.
+The rational behind the transition probability is that the transition
+weight can be related to the fractional uncertainty in upward and
+downward moves assuming there is no system correlation (the moves are
+independent).  Although this is not necessarily the case, we find that
+this definition for the transition weight is satifactory. The
+probability of a move is then the geometic average of either using the
+SA or TMMC Algorithm.
+
+\begin{align}
+\mathcal{P}_\text{move} = \mathcal{P}_\text{move}^{\text{SA}_\text{Weight}} \mathcal{P}_\text{move}^{1-\text{SA}_\text{Weight}}
+\end{align}
+
+\subsubsection{Enforce a minimum temperature probability}
+One potentially damaging aspect to the convergence of the Stochastic
+Approximation method is the need to enforce a minimum temperature
+probability.  We define this probability as
+\begin{align}
+P_{\text{min}} = e^{\beta_{\text{max}}\Delta\epsilon}
+\end{align}
+where $\beta_{\text{max}} = \nicefrac{1}{k_B T_\text{min}}$.  If the
+transition probability is ever below the minimum temperature transition
+probability, we set the former equal to the later.
 
@@
 well-width of $\text{ww} = 1.30$ and a filling fraction of either
-$\text{ff} = 0.10, 0.20$ or $\text{ww} = 0.20$.  The simulation explores the
+$\text{ff} = 0.22$ or $\text{ww} = 0.30$.  The simulation explores the
 energy space of the system down to a reduced temperature of
@@
 of these parameters are needed a priori).  In the sections below, we
-examine systems with different particles numbers and filling fractions.
+examine systems with different particles sizes, system dimensions, and
+sticky-wall.
 
-~\jpsays{Add N50 ff0.1 and N50 ff0.3 and remove sticky-wall!}
+\subsubsection{Sticky-wall configuration}
+A sticky-wall is an attractive wall that impacts particle moves.  Only
+one such wall is inserted into a simulation and the intent is to assist
+in visualizing phase separation.  As the temperature of the system
+decreases, an increase in particle aggregation near the sticky-wall
+becomes evident. As the energy of the particles is decreased, the
+particles are unlikely to be able to escape the attractive sticky-wall
+potential.  By examining the particle number density vs the system
+length (z-axis), phase transitions to a liquid-vapor system from a
+vapor dominated system become evident.
+
 \begin{figure}

### 2018-02-22 papers/histogram/sad.tex
data update with sad paper figure change.

--- 2018-02-19 papers/histogram/sad.tex
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-30x4-max-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-max-entropy-error.pdf}
   \caption{The average entropy error for each method is shown with
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-60x8-max-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-max-entropy-error.pdf}
   \caption{The average entropy error for each method is shown with SAD serving as

### 2018-02-26 papers/histogram/sad.tex
updated paper and changed colors.py to order/include methods when plotting.

@@
 well-width of $\text{ww} = 1.30$ and a filling fraction of either
-$\text{ff} = 0.10, 0.20$ or $\text{ww} = 0.20$.  The simulation explores the
+$\text{ff} = 0.10, 0.20,$ or $0.30$.  The simulation explores the
 energy space of the system down to a reduced temperature of
@@
 
-~\jpsays{Add N50 ff0.1 and N50 ff0.3 and remove sticky-wall!}
 \begin{figure}
@@
   SAD serving as the `golden'
-  standard.}\label{fig:avgS-30x4}
+  standard.}\label{fig:N50-ff0.3-max-error}
 \end{figure}
@@
   \caption{The average entropy error for each method is shown with SAD serving as
-  the `golden' standard.}\label{fig:avgS-60x8}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/lv/ww1_30-ff0_22-100x10-max-entropy-error}
-  \caption{The average entropy error for each method is shown with
-    WL-TMMC using a cutoff of 10$^{-10}$ serving as the
-    standard.}\label{fig:avgS-100x10}
+  the `golden' standard.}\label{fig:N50-ff0.1-max-error}
 \end{figure}
@@
   \caption{The average entropy error is shown with TMI acting as the
-  `golden' standard.}\label{fig:N50-mean-error}
+  `golden' standard.}\label{fig:N50-ff0.2-max-error}
 \end{figure}
@@
 
-\subsubsection{A $30\times4\times4$ $(N = 25)$ system with a sticky
-wall}
+\subsubsection{A periodic system with 10\% filling fraction and $(N = 50)$}
 In order to determine how the various methods scale with size,
-simulations are run for systems with different particle numbers.  For
-the $N = 25$ system, a sticky-wall is added to ensure that we see the
-liquid and vapor phase separation.
+simulations are run for systems with different particle numbers. A filling
+fraction $\text{ff} = 0.10$ is chosen which yields $N = 50$ particles
+in the cubical system.  These particles are randomly inserted into the
+system and random moves are proposed (with a translation scale
+$\delta_0 = 0.05$). The boundaries in $x$ and $y$ are periodic while
+the $z$-axis is not periodic.
 
-A system is configured with dimensions $x = y = 4\sigma$ and $z =
-30\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
-yields $N = 25$ particles in the rectangular system.  These particles
-are randomly inserted into the system and random moves are proposed
-(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
-$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
-inserted along the $z$-axis at one end to promote phase separation.
-~\jpsays{should we include a sticky-wall.py plot or something else or
-are we not going to talk about phase separation in this paper?}
-
-We use the entropy error vs iterations as a metric to compare
-simulation runtimes.  Simulations that take more iterations to reach
-the same entropy error as a different method take longer to converge.
-This behavior is seen in Fig.~\ref{fig:avgS-30x4}.
-~\jpsays{SAD appears to do pretty well as does WL-TMMC for a cutoff of $10^{-10}$
-although this seems to depend on what is used as the `standard'.}
+We use the maximum entropy error vs iterations as a metric to compare
+simulation runtimes and overall convergence.  Simulations that take
+more iterations to reach the same entropy error as a different method
+take longer to converge. This behavior is seen in
+Fig.~\ref{fig:N50-ff0.1-max-error}.
 
@@
 
-Figure~\ref{fig:N50-mean-error} shows a system consisting of 50 atoms
-are placed in a periodic simple cubic unit cell with lattice constant
+Figure~\ref{fig:N50-ff0.2-max-error} shows a system consisting of 50 atoms
+that are placed in a periodic simple cubic unit cell with lattice constant
 chosen to give a filling fraction of 0.20.  These particles are
@@
 
-\subsubsection{A $60\times8\times8$ $(N = 202)$ system with a sticky
-wall}
+\subsubsection{A periodic system with 30\% filling fraction and $N = 50$}
 
-A system is configured with dimensions $x = y = 8\sigma$ and $z =
-60\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
-yields $N = 202$ particles in the rectangular system.  These particles
-are randomly inserted into the system and random moves are proposed
-(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
-$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
-inserted along the $z$-axis at one end to promote phase separation.
+Figure~\ref{fig:N50-ff0.3-max-error} shows a system consisting of 50 atoms
+that are placed in a periodic simple cubic unit cell with lattice constant
+chosen to give a filling fraction of 0.30.  These particles are
+randomly inserted into the system and random moves are proposed (with
+a translation scale $\delta_0 = 0.05$).
 
-\subsubsection{A $100\times10\times10$ $(N = 525)$ system with a sticky
-wall}
+\subsubsection{A periodic system with 30\% filling fraction and $N = 500$}
 
-A system is configured with dimensions $x = y = 10\sigma$ and $z =
-100\sigma$.  A filling fraction $\text{ff} = 0.22$ is chosen which
-yields $N = 525$ particles in the rectangular system.  These particles
-are randomly inserted into the system and random moves are proposed
-(with a translation scale $\delta_0 = 0.05$). The boundaries in $x$ and
-$y$ are periodic while the $z$-axis is not periodic.  A sticky-wall is
-inserted along the $z$-axis at one end to promote phase separation.
-
-\subsubsection{A Tanner dimensions? $(N = 500)$ system with no sticky
-wall}
-
-A system is configured with dimensions $x = y = z = 19\sigma?$.  A
+A system is configured with dimensions $x = y = z = 19\sigma$.  A
 filling fraction $\text{ff} = 0.30$ is chosen which yields $N = 500$
@@
 $\delta_0 = 0.05$). The boundaries in $x$ and $y$ are periodic while
-the $z$-axis is not periodic.  A sticky-wall is not inserted along the
-$z$-axis.
+the $z$-axis is not periodic.
 

### 2018-04-11 papers/histogram/sad.tex
include gamma plot in sad paper (first draft of plot)

@@
 \section{Introduction}
-Over the past several decades a number of advanced Monte-Carlo
+Over the past several decades a number of broad histogram Monte-Carlo
 simulation algorithms have been developed which calculate the
-properties of a thermal system over a range of temperatures.  This
-idea began with the original histogram method, which used a single
-canonical Monte Carlo simulation to predict properties for nearby
-temperatures~\cite{ferrenberg1988new}.  For large simulations this
-approach is limited to a narrow temperature range because a single
-canonical simulation explores only a small range of energies.  This
-led to a variety of ``broad'' (or ``flat'') histogram
+thermodynamic properties of various systems for all temperatures.  The
+development began with the original histogram method, which used a
+single canonical Monte Carlo simulation to predict properties for
+nearby temperatures~\cite{ferrenberg1988new}.  For large simulations
+this approach is limited to a narrow temperature range because a single
+canonical simulation explores only a small range of energies.  This led
+to a variety of ``broad'' (or ``flat'') histogram
 methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
@@
 attempt to explore a wide range of energies.  These methods also
-benefit in that they are unlikely to be trapped in a local energy
-minimum.
+benefit in that they are unlikely to be trapped in a low enetropy state.
 
-Wang and Landau introduced an algorithm that uses an update
-factor and a statistical histogram to compute the density of states of
-a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has several user-defined
-parameters.  This can make its application to a variety of systems
-something of an art-form since the user needs to determine the ideal
-parameters for his particular system.  Also, detailed balance is
-violated (although for only short periods of time), yet the method
-converges in most situations.  This raises several questions such as
-when does the method converges, what rate will the method converge to
-the correct density of states, and how to appropriately decide what
+Wang and Landau introduced an algorithm that uses an update factor and
+a statistical histogram to compute the density of states of a given
+system~\cite{wang2001determining, wang2001efficient}.  While the method
+is incredibly powerful, it has several user-defined parameters.  This
+can make its application to a variety of systems something of an
+art-form since the user needs to determine the ideal parameters for the
+particular system being studied.  Also, detailed balance is violated
+(although only for short periods of time), ensuring that convergence is
+not guaranteed.  This raises several questions such as when the method
+does converge, what rate will the method converge to the correct
+density of states. Also, how can the user appropriately decide what
 parameters to choose so that the algorithm solves the problem in the
-best way. Transition Matrix Monte Carlo~\cite{wang1999transition,
+most ideal way.
+
+Transition Matrix Monte Carlo~\cite{wang1999transition,
 swendsen1999transition, fitzgerald2000monte} became an attractive
-complimentary simulation algorithm to Wang-Landau in that, if detailed
-balance is ensured, the system is guaranteed to converge (something
-which is not guaranteed for the Wang-Landau Algorithm).  Unfortunately,
+complimentary simulation algorithm to Wang-Landau, in that, if detailed
+balance is ensured, the system is guaranteed to converge.  Unfortunately,
 the algorithm can be considerably more difficult to implement when
-compared to Wang-Landau: in particular, the infinite temperature
+compared to Wang-Landau: in particular, due to the infinite temperature
 transition matrix~\cite{wang2002transition}.  Also, the time necessary
@@
 due to the algorithm spending too much time exploring low energy states.
-~\tssays{(note to self) The above paragraph could be condensed a bit}
+Unlike WL, TMMC does not require a prior knowledge of the energy range of
+the system.
 
@@
 best parameters for a users particular system of
-study~\cite{siderius2013use}.  Around the same time, the convergence
-rate for Wang-Landau was being explored~\cite{zhou2005understanding}.
-It was found that utilizing modification factors that decrease faster
-than $1/t^2$ leads to nonconvergence~\cite{belardinelli2007fast}.  This
-lead to the so-called $1/t$-WL algorithm which ensures that excess CPU
-time is not wasted by continuing to perfrom calculations once the error
-in the density of states becomes
-saturated~\cite{belardinelli2008analysis}. An effort to reduce the
-number of user defined parameters of WL and formulate a theory for why
-the method converges despite detailed balance being violated although
-infrequently was also being udertaken.  Liang, Liu, and Carrol began to
-consider whether WL could be considered a special case of Monte Carlo
-whose convergence could be mathematically
-proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
-al.~\cite{liang2007stochastic} argued that WL can be considered a form
-of Stochastic Approximation Monte Carlo (SAMC).  While SAMC can
-guarantee convergence, the method still has a system specific
+study~\cite{siderius2013use}.
+
+Around the same time, the convergence rate for WL was being
+explored~\cite{zhou2005understanding,lee2006convergence,
+belardinelli2007wang}. It was found that utilizing modification factors
+that decrease faster than $1/t^2$ leads to
+nonconvergence~\cite{belardinelli2007fast}.  This lead to the so-called
+$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
+continuing to perfrom calculations once the error in the density of
+states becomes saturated~\cite{belardinelli2008analysis}. An effort to
+reduce the number of user defined parameters of WL and formulate a
+theory for why the method converges despite detailed balance being
+violated although infrequently was also being udertaken.
+
+Liang, Liu, and Carrol began to consider whether WL could be considered
+a special case of Stochastic Approximation whose convergence could be
+mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
+2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
+considered a form of Stochastic Approximation Monte Carlo (SAMC). While
+SAMC can guarantee convergence, the method still has a system specific
 user-defined variable which makes applying this algorithm to arbitrary
@@
 
-In our work, we have developed a novel algorithm based on SAMC that
+In this work, we have developed a novel algorithm based on SAMC that
 does not require user-defined inputs and therefore should be easily
@@
 Stochastic Approximation), and will discuss it in detail in the methods
-section. We compare it along with several histogram methods including TMMC,
-WL, WL-TMMC, and SAMC.
+section. We compare it along with several broad histogram methods
+which include TMMC, WL, WL-TMMC, and SAMC.
 
@@
 
-Here we employ a variety of broad-histogram methods.  We outline the
-general workings of algorithm we developed in detail and summarize
-algorithms that can be found elsewhere.  The following algorithms are
-introduced and applied to the square-well fluid:  Transition
-Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau Transition
-Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation Monte-Carlo
-Monte-Carlo (SAMC), and Dynamical Stochastic-Approximation (SAD).
+In this work, We compare a variety of broad-histogram methods.  We
+outline the general workings of each algorithm that we developed in
+detail while summarizing algorithms that can be found elsewhere.  The
+following methods are introduced and applied to the square-well fluid:
+Transition Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau
+Transition Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation
+Monte-Carlo (SAMC), and Dynamical Stochastic-Approximation
+(SAD).
 
 \subsection{TMMC Algorithm}
-The Transition Matrix Monte-Carlo (TMMC) method as introduced by Fitzgerald et al.
-consists of three primary steps.  The first two steps are common to the Metropolis
-algorithm while the third makes use of the actual transition
-probabilities~\cite{fitzgerald2000monte}.
+The Transition Matrix Monte-Carlo (TMMC) method as introduced by
+Fitzgerald et al. consists of three primary steps.  The first two steps
+are common to the Metropolis algorithm while the third makes use of the
+actual transition probabilities~\cite{fitzgerald2000monte}.
 
 \subsubsection{define a transition probability}
-Assuming that the system under study is in a state denoted old and a
+Assuming that the system under study is in a state denoted old, and a
 proposed transition is made to a new state denoted new, the probability
@@
 \end{align}
-where we have allowed the probabilies to be equal for simplicity of presentation.
+where we have allowed the probabilies to be equal for simplicity of
+presentation.
 
 \subsubsection{define an acceptance probability}
-The probability for accepting a transition from an old state to a new one can be
-defined
+The probability for accepting a transition from an old state to a new
+one can be defined
 \begin{align}
@@
 ensemble constraints placed on the system.  Fitzgerald et al.
-considered the system of interest using the canonical ensemble however
-extension to other ensembles is quite natural.
+considered the system of interest using the canonical ensemble;
+however, the extension to other ensembles is quite natural.
 \begin{align}
@@
 \end{align}
-The acceptance probability given for TMMC is idential to that introduced by
-Swendsen and Wang~\cite{wang2002transition}.
+The acceptance probability given for TMMC is idential to that
+introduced by Swendsen and Wang~\cite{wang2002transition}.
 \begin{align}
@@
 \subsubsection{define a bookkeeping step}
-The third step in the TMMC algorithm consists of recording the data in a collection
-matrix (analogous to the transition histogram matrix).
+The third step in the TMMC algorithm consists of recording the data in
+a collection matrix (the unormalized transition histogram matrix).
 \begin{align}
@@
 changed, then only $C_{N,0}$ is incremented by unity. Fitzgerald et
-al.~\cite{fitzgerald2000monte} show that defining the third step this
+al.~\cite{fitzgerald2000monte} showed that defining the third step this
 way results in a uniform improvement over using histogram estimators to
@@
 random walk in energy space in order to estimate the density of
-states~\cite{wang2001efficient}. While methods such as Metropolis
-Sampling and Swendsen-Wang cluster
-flipping~\cite{swendsen1987nonuniversal} generate a narrow distribution
-and require sampling at individual temperatures, the Wang-Landau
-algorithm uses a flat histogram and performs a random walk in energy
-space to determine the density of states~\cite{wang2001determining,
-landau2014guide}. Wang-Landau's major tenant is that when counting the
-histogram $H(\epsilon)$, the energy occurences should form a flat
-distribution.  The criteria for flat sampling is given by
+states~\cite{wang2001efficient,wang2001determining, landau2014guide}.
+Wang-Landau's major tenant is that when counting the histogram
+$H(\epsilon)$, the energy occurences should form a flat distribution.
+The criteria for flat sampling is given by
 \begin{equation}
@@
 	> \eta
-\end{equation} where $\eta$ is usually set to $0.80$ and determines how
-many times each energy is sampled relative to the mean number of
-visits.  Difficulties can occur with this flatness criteria due to the
-fact that some energies on a given energy range might never be
-sampled~\cite{haber2014transition}.  Over the past decade much
-attention was directed at examining the rate of convergence and error
-saturation~\cite{lee2006convergence, belardinelli2007wang}.  The
-algorithm is briefly outlined below where $w$ represents the weights:
+\end{equation}
+where $\eta$ is usually set to $0.80$ and determines how many times
+each energy is sampled relative to the mean number of visits.
+Difficulties can occur with this flatness criteria due to the fact that
+some energies on a given energy range might never be
+sampled~\cite{haber2014transition}. The algorithm is briefly outlined
+below where $w$ represents the weights:
 \begin{equation}
-	\ln{w_{k+1}(\epsilon,N)}=\ln{w_{k}(\epsilon,N)}
+	\ln{D_{i+1}(\epsilon,N)}=\ln{D_{i}(\epsilon,N)}
 	+\gamma
@@
 \begin{equation}
-	\ln{\gamma_{k+1}}=\frac{u}{2}\ln{\gamma_{k}}
+	\ln{\gamma_{i+1}}=\frac{u}{2}\ln{\gamma_{i}}
 \end{equation}
-for $u$ typically greater than 1 and $\gamma$ is the modification factor for
-the density of states.  We then accept a new configuration for the density of
-states with a probability given by the following
+for $u$ typically greater than 1 and $\gamma$ is the modification
+factor for the density of states.  A new configuration for the density
+of states with a probability given by the following is chosen:
 \begin{equation}
-	\mathcal{P}(\epsilon_i \rightarrow \epsilon_j)
-	= \min[1,e^{\ln{w(\epsilon_i)}-\ln{w(\epsilon_j)}}]
+	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
+	= \min[1,e^{\ln{D(\epsilon_\text{old})}-\ln{D(\epsilon_\text{new})}}]
 \end{equation}
-Zhou and Bhatt explored the convergence of the Wang-Landau algorithm and
-determined that choosing a large modification factor $\gamma_0 = e^{4}$ and
-rapidly reducing the factor by 10 during each step resulted in reduced statistical
-error~\cite{zhou2005understanding}.
+The entire process is repeated until a desired cutoff is reached.
 
@@
 $10^{-10}$.  Rane et. al. published an implemenation of
-WL-TMMC~\cite{rane2013monte} that suggested using a WL factor between 1.0 and
-$10^{-2}$ with a cutoff $<10^{-5}$. In our implementation of WL-TMMC,
-we set the WL factor to be 1 and the cutoff at $10^{-10}$. A single
-sweep is defined as each macrostate being sampled a number of times in
-this case 10,000 (Siderius et. al. does 100). Wang-Landau sets a
-flatness criteria~\cite{wang2001determining, wang2001efficient,
+WL-TMMC~\cite{rane2013monte} that suggested using a WL factor between
+1.0 and $10^{-2}$ with a cutoff $<10^{-5}$. In our implementation of
+WL-TMMC, we set the WL factor to be 1 and the cutoff to be either
+$10^{-4}$ or $10^{-10}$. A single sweep is defined as each macrostate
+being sampled a number of times in this case 10,000 (although 100 is
+often sufficient for small systems). Wang-Landau sets a flatness
+criteria~\cite{wang2001determining, wang2001efficient,
 hatch2015computational, mahynski2017predicting} for the accumulated
-histogram usually between 0.75 to 0.90.  For our simulation, it is
-sufficient that each macrostate is visited at least once (each
-histogram bin has one entry)~\cite{shell2003improved}.  The goal of
-this WL initialization is to prefill the transition matrix for the next
-portion of the simulation.
+histogram usually between 0.75 to 0.90.  In the original
+implementation, it was found to be sufficient that each macrostate is
+visited at least once (each histogram bin has one
+entry)~\cite{shell2003improved}; however, for other systems this is
+rarely sufficient.  The goal of this WL initialization is to prefill
+the transition matrix for the next portion of the simulation.
 \paragraph{Initializing TMMC:} The purpose of prefilling the transition
@@
 complete~\cite{shell2003improved, shen2014elucidating}.  The transition
-matrix in terms of the collection matrix is given as follows:
+matrix in terms of the collection matrix is given as
+follows:
 \begin{align}
@@
 \end{align}
-We now run the TMMC portion of the simulation.  Siderius does this for
-20 additional sweeps (does about 8 sweeps for WL).  We do this until we
-reach a specified temperature of interest.
-We have not yet implemented a biasing function to control acceptance
-rate of trial moves but we probably could/should? at some point.
+We now run the TMMC portion of the simulation.
+%Siderius does this for 20 additional sweeps (does about 8 sweeps for
+%WL).  We do this until we reach a specified temperature of interest. We
+%have not yet implemented a biasing function to control acceptance rate
+%of trial moves but we probably could/should? at some point.
+The acceptance probability is written in terms of the infinite temperature
+transition matrix.
 \begin{align}
@@
 \end{align}
-this is only used to control actual transitions not update the collection
-matrices.
+This probability is only used to control actual transitions not update
+the collection matrices.
 
-\subsection{SAD Algorithm}
+\subsection{SAMC Algorithm}
+SAMC can be applied as a dynamic importance sampling algorithm that
+uses an update factor to explore energy
+space~\cite{liang2007stochastic, werlich2015stochastic,
+schneider2017convergence}.  The update factor is defined in the
+original implementation~\cite{liang2007stochastic} in terms of an
+arbitrary tunable parameter $t_0$ and a scale factor $\gamma_0$.
+\begin{align}
+\gamma_{t}^{\text{SA}} =\gamma_0 \frac{t_0}{\max(t_0,t)}
+\end{align}
+SAMC offers several advantages over other broad histogram methods
+including proven convergence. It has been shown that if
+$F_{t}^{\text{SA}}$ satisfies two conditions the estimator is proven to
+converge to the real DOS~\cite{liang2006theory, liang2007stochastic}.
+\begin{align}
+\sum_{t=1}^\infty \gamma_{t}^{\text{SA}} = \infty \quad\textrm{and}\quad
+\sum_{t=1}^\infty (\gamma_{t}^{\text{SA}})^\zeta < \infty
+\end{align}
+where $\zeta \in \{1,2\}$.  Unlike WL and WL-TMMC, the energy range need
+not be known a priori.  The time to converge depends only on the choice of
+parameters $t_0$ and $\gamma_0$.
+Unfortunately, only by performing many computational runs might the user
+determine what SAMC arbitrary parameters to choose.  This has lead us
+to develop a systematic way (which we will discuss in the next section)
+for the simulation to determine what $t_0$ should be.
+
+\section{SAD Algorithm}
+\subsubsection{The theory behind SAD}
 The SAD (Dynamic Stochastic Approximation) is a version of the SAMC
@@
 $\gamma_0$.  There is an immediate advantage of such an algorithm where
-parameters are chosen independent of system size or type. SAMC can be
-applied as a dynamic importance sampling algorithm that uses an update
-factor to explore energy space~\cite{liang2007stochastic,
-werlich2015stochastic, schneider2017convergence}.  The update factor is
-defined in the original implementation~\cite{liang2007stochastic} in
-terms of an arbitrary tunable parameter $t_0$.
+parameters are chosen independent of system size or type. While for
+SAMC the update factor is defined in the original
+implementation~\cite{liang2007stochastic} in terms of an arbitrary
+tunable parameter $t_0$,
 \begin{align}
@@
 \end{align}
-We argue that it is possible to set this parameter to a system
-independent parameter such that $t_0 = N_E^2/t$ where $N_E$ is the
-number of energies that have been found.  One of the energy found terms
-can be thought of as a change in the entropy $\Delta S$.  The change in
-the entropy can be thought of as a slope of the triangle made by
-connecting the minimum important energy and the maximum entropy state
-divided by the minimum temperature:
+The update factor $\gamma_{t}^{\text{SA}}$ can be thought of as
+$\nicefrac{\text{d}S}{\text{d}t}$. This tells us that $t_0$ should have
+dimensions of entropy.  We argue that we can write this entropy `term'
+geometrically as an area bounded by the $\ln{D(E)}$ i.e. the entropy.  This
+gives the following
 \begin{align}
-\Delta S = \frac{E_{\text{max}}-E_{\text{min}}}{T_{\text{min}}}
+\gamma_{t}^{\text{SA}} =
+\frac{N_E}{t}\frac{E_{\text{max}}-E_{\text{min}}}{3T_{\text{min}}}
 \end{align}
-This gives an update factor as follows:
-\begin{align}
-\gamma_{t}^{\text{SA}} = \frac{N_E\Delta S}{t} =
-\frac{N_E}{t}\frac{E_{\text{max}}-E_{\text{min}}}{T_{\text{min}}}
-\end{align}
-The update factor has the same $1/t$ behavior as the original SAMC
-algorithm; however now every time a new energy is discovered during the
-course of the simulation, the update factor experiences a jump.
-~\jpsays{Add in additional material for SAD!}
+where $N_E$ is the number of energies that have been found. The change
+in the entropy the minimum temperature can be thought of as a slope of
+the triangle made by connecting the minimum important energy and the
+maximum entropy state .  The factor of $3$ comes from the area bounded
+by the $\ln{D(E)}$ The update factor has the same $1/t$ behavior as the
+original SAMC algorithm; however now every time a new energy is
+discovered during the course of the simulation, the
+update factor experiences a jump.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/gamma-ww1_30-ff0_22-100x10}
+  \caption{The update factor $\gamma$ versus iteration for three
+    different methods: WL, SAMC, and SAD.
+    method.}\label{fig:gamma-vs-t}
+\end{figure}
+
+Fig.~\ref{fig:gamma-vs-t}. shows the monotonic decrease of the update
+factor for SAMC.  The update factor for WL decreases like a step
+function while the update factor for SAD dynamically fluctuates around
+a value less than 1 for early MC moves (while finding new energy
+states).  Ultimately once all energy states have been found, the update
+factor for SAD decreases in the same monotonic fashion as SAMC.
+
+\subsubsection{Energy binning with SAD}
+
+Energy binning is an important consideration for complex continous
+systems. Coarse binning yields fewer energy states for the walker to
+explore but more time spent in the individual bin. Fine binning yields
+more energy states to explore but less time spent in each bin.  Some
+methods such as AdaWL~\cite{koh2013dynamically} (discussed elsewhere)
+employ a tunable mechanism for controlling the binning for low entropic
+states in order to ensure the exploration of all energies. A
+significant advantage of SAD over other broad histogram methods is that
+energy binning is solved dynamically. SAD should perform independently
+for a reasonable range of binning because $\gamma \propto
+\nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$
+increases (fine binning), the time spent $t$ in each bin will decrease.
+This argument explains why SAD should perform well for complex systems
+in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
+similar way but this adds yet another parameter for the user to choose.
+
+\subsubsection{Scaling, step size, and SAD}
+Methods such as WL can fail to converge for complex systems in part
+because the same random MC moves are performed for the entire energy
+range.  The main problem lies in trying to determine how to optimize
+the step size such that the time to explore all energies is reasonable.
+Small step sizes lead to improved sampling at the expense of slow
+convergence. Koh et al. proposed multiplying $\gamma$ by an energy
+dependent factor in order to improve the converge of WL in regards to
+convergence and scaling~\cite{koh2013dynamically}.  Because the update
+factor for SAD has an energy dependent factor already present (and
+dynamically adjusted), we expect that SAD will perform consitently
+scaling free for a reasonable choice of step size.  SAD should perform
+extremely well compared with other broad histogram methods when applied
+to complex continuous systems such as polymers/proteins and spin models.
 

### 2018-04-25 papers/histogram/sad.tex
made preliminary changes to methods section

@@
 \begin{document}
-\title{Dynamical Stochastic Approximation Monte Carlo}
+\title{Dynamic Stochastic Approximation Monte Carlo}
 
@@
 does not require user-defined inputs and therefore should be easily
-applicable to a given system.  We call this method SAD (Dynamical
+applicable to a given system.  We call this method SAD (Dynamic
 Stochastic Approximation), and will discuss it in detail in the methods
@@
 
-In this work, We compare a variety of broad-histogram methods.  We
+In this work, We compare a variety of broad histogram methods.  We
 outline the general workings of each algorithm that we developed in
-detail while summarizing algorithms that can be found elsewhere.  The
+detail while summarizing algorithms that originated elsewhere.  The
 following methods are introduced and applied to the square-well fluid:
@@
 Transition Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation
-Monte-Carlo (SAMC), and Dynamical Stochastic-Approximation
+Monte-Carlo (SAMC), and Dynamic Stochastic Approximation 
 (SAD).
@@
 The acceptance probability given for TMMC is idential to that
-introduced by Swendsen and Wang~\cite{wang2002transition}.
+introduced by Swendsen and Wang~\cite{wang2002transition}. \tssays{Isn't
+this not quite right? Not sure if needed.}
 \begin{align}
@@
 \end{equation}
-for $u$ typically greater than 1 and $\gamma$ is the modification
-factor for the density of states.  A new configuration for the density
-of states with a probability given by the following is chosen:
+\tssays{Should this be just $\gamma$, not $\ln{\gamma}$?}for $u$ typically 
+greater than 1 and $\gamma$ is the modification factor for the density of 
+states.  A new configuration for the density of states with a probability 
+given by the following is chosen:
 \begin{equation}
@@
 of systems, while making use of a transition matrix to address WL
-weaknesses such as a limited statistical accuracy plateau which is not
-improved with additional Monte-Carlo steps.  This WL version is more
-accurate for a given number of simulation steps.
+weaknesses, such as its limited statistical accuracy plateau.  
+This WL version is more accurate for a given number of simulation steps.
 
@@
 the transition matrix for the next portion of the simulation.
-\paragraph{Initializing TMMC:} The purpose of prefilling the transition
-matrix is apparent because if numerous zeros existed in the collection
+\paragraph{Initializing TMMC:} Prefilling the transition
+matrix is necessary because if numerous zeros existed in the collection
 matrix, the infinite temperature transition matrix would be
@@
 the triangle made by connecting the minimum important energy and the
-maximum entropy state .  The factor of $3$ comes from the area bounded
-by the $\ln{D(E)}$ The update factor has the same $1/t$ behavior as the
-original SAMC algorithm; however now every time a new energy is
-discovered during the course of the simulation, the
-update factor experiences a jump.
+maximum entropy state.  The factor of $3$ comes from the ratio of areas bounded
+by the $\ln{D(E)}$ and the minimum temperature line.
+The update factor has the same $1/t$ behavior as the original SAMC algorithm; 
+however now every time a new energy is discovered during the course 
+of the simulation, the update factor experiences a jump.
 

### 2018-04-25 papers/histogram/sad.tex
changed to plot average entropy error in sad.

--- 2018-04-11 papers/histogram/sad.tex
@@
 \begin{abstract}
-  We present several histogram methods and compare the performance and efficiency at treating the square-well fluid.
+  We present several histogram methods and compare the performance and
+  efficiency at treating the square-well fluid.
 \end{abstract}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-max-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
   \caption{The average entropy error for each method is shown with
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-max-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-entropy-error.pdf}
   \caption{The average entropy error for each method is shown with SAD serving as
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-max-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-entropy-error.pdf}
   \caption{The average entropy error is shown with TMI acting as the
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-max-entropy-error.pdf}
+\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
   \caption{The average entropy error is shown with TMI acting as the

### 2018-05-02 papers/histogram/sad.tex
use n500 for gamma plot in sad paper
added discussion section to paper and edits.

@@
 \end{align}
-The acceptance probability given for TMMC is idential to that
-introduced by Swendsen and Wang~\cite{wang2002transition}. \tssays{Isn't
-this not quite right? Not sure if needed.}
-\begin{align}
-  p_{a} = \min\bigg[1,e^{-\frac{\delta\epsilon}{k_{B} T}}\bigg]
-\end{align}
 \subsubsection{define a bookkeeping step}
@@
 \begin{equation}
-	\ln{\gamma_{i+1}}=\frac{u}{2}\ln{\gamma_{i}}
+	\gamma_{i+1}=\frac{u}{2} \gamma_{i}
 \end{equation}
-\tssays{Should this be just $\gamma$, not $\ln{\gamma}$?}for $u$ typically 
-greater than 1 and $\gamma$ is the modification factor for the density of 
-states.  A new configuration for the density of states with a probability 
-given by the following is chosen:
+for $u$ typically greater than 1 and $\gamma$ is the modification
+factor for the density of states.  A new configuration for the density
+of states with a probability given by the following is chosen:
 \begin{equation}
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-ww1_30-ff0_22-100x10}
-  \caption{The update factor $\gamma$ versus iteration for three
-    different methods: WL, SAMC, and SAD.
+  \includegraphics[width=\columnwidth]{figs/gamma-n500.pdf}
+  \caption{The update factor $\gamma$ for N = 500 versus iteration for
+  three different methods: WL, SAMC, and SAD.
     method.}\label{fig:gamma-vs-t}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown with
-  SAD serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-max-error}
+  \caption{The average entropy error for each method is shown for a filling fraction of 0.30 with
+  SAMC followed by TMMC serving as the `golden'
+  standard.}\label{fig:N50-ff0.3-avg-error}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown with SAD serving as
-  the `golden' standard.}\label{fig:N50-ff0.1-max-error}
+  \caption{The average entropy error for each method is shown for a filling fraction of 0.10 with
+  SAMC followed by TMMC serving as the `golden' standard.}\label{fig:N50-ff0.1-avg-error}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-entropy-error.pdf}
-  \caption{The average entropy error is shown with TMI acting as the
-  `golden' standard.}\label{fig:N50-ff0.2-max-error}
+  \caption{The average entropy error is shown for a filling fraction of 0.20 with TMI acting as the
+  `golden' standard.}\label{fig:N50-ff0.2-avg-error}
 \end{figure}
@@
 \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
-  \caption{The average entropy error is shown with TMI acting as the
-  `golden' standard.}
+  \caption{The average entropy error is shown for a filling fraction of 0.30 with TMI acting as the
+  `golden' standard.}\label{fig:n500-ff0.2}
 \end{figure}
@@
 
-We use the maximum entropy error vs iterations as a metric to compare
-simulation runtimes and overall convergence.  Simulations that take
-more iterations to reach the same entropy error as a different method
-take longer to converge. This behavior is seen in
-Fig.~\ref{fig:N50-ff0.1-max-error}.
+We use the average entropy error vs iterations~\jpsays{we should
+explain how we do this.} as a metric to compare simulation runtimes and
+overall convergence.  Simulations that take more iterations to reach
+the same entropy error as a different method take longer to converge.
+This behavior is seen in Fig.~\ref{fig:N50-ff0.1-avg-error}.
 
@@
 
-Figure~\ref{fig:N50-ff0.2-max-error} shows a system consisting of 50 atoms
+Figure~\ref{fig:N50-ff0.2-avg-error} shows a system consisting of 50 atoms
 that are placed in a periodic simple cubic unit cell with lattice constant
@@
 
-Figure~\ref{fig:N50-ff0.3-max-error} shows a system consisting of 50 atoms
+Figure~\ref{fig:N50-ff0.3-avg-error} shows a system consisting of 50 atoms
 that are placed in a periodic simple cubic unit cell with lattice constant
@@
 the $z$-axis is not periodic.
+
+\subsubsection{Discussion on convergence and update factor}
+
+A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$ for the 
+Wang-Landau algorithm decreases much more rapidly than $\nicefrac{1}{\sqrt{t}}$.
+It can be observed that all the methods from this family struggle to converge if
+the update factor decreases faster than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.2}
+shows both WL and SAMC with inappropriate choices of $t_0$ i.e. less than 
+$\nicefrac{10^5}{\sqrt{t}}$ failing to converge in a reasonable amount of time.
+Thus, the update factor $\gamma$, shown as a function of simulation moves
+in Fig.~\ref{fig:gamma-vs-t}, proves useful in determining which methods will 
+converge in a reasonable amount of time.
+We compare each method to a `golden' standard which we define to be an ideal choice
+of SAMC for $10^11$ moves followed by TMMC.  This gives an unbiased comparison with
+our developed method SAD.
 

### 2018-05-10 papers/histogram/sad.tex

@@
 %Siderius does this for 20 additional sweeps (does about 8 sweeps for
-%WL).  We do this until we reach a specified temperature of interest. We
+%WL).  We do this until we reach a defined number of iterations. We
 %have not yet implemented a biasing function to control acceptance rate
-%of trial moves but we probably could/should? at some point.
+%of trial moves but we could? at some point.
 The acceptance probability is written in terms of the infinite temperature
@@
 however now every time a new energy is discovered during the course
-of the simulation, the update factor experiences a jump.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n500.pdf}
-  \caption{The update factor $\gamma$ for N = 500 versus iteration for
-  three different methods: WL, SAMC, and SAD.
-    method.}\label{fig:gamma-vs-t}
-\end{figure}
+of the simulation, the update factor experiences a jump. This behavior allows
+SAD to dynamically prevent the update factor from decreasing too rapidly.
 
@@
 factor for SAD decreases in the same monotonic fashion as SAMC.
+
+\subsubsection{High and Low energy racheting}
+Since SAD explores all energy states it is imperative that the method have
+a reasonable idea about where the minimum important energy and maximum
+entropy state are located in energy space. Since these are not user input
+parameters like for methods suchas WL-TMMC, the simulation is responsible
+for determining and updating these energy locations. In order to prevent,
+the method from experiencing errors near the energy extrema, we employ a 
+high energy and a low energy to define the range over which the energy 
+histogram is made flat. The high energy is the highest ever value that the
+maximum entropy state has taken while the low energy is the lowest ever value
+that the minimum important energy has taken. Together these two energies serve
+as a kind of ratcheting effect to determine the energy landscape boundaries.
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/gamma-n500}
+  \caption{The update factor $\gamma$ for $N=500$ versus iteration for three
+    different methods: WL, SAMC, and SAD.
+    method.}\label{fig:gamma-vs-t}
+\end{figure}
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown for a filling fraction of 0.30 with
-  SAMC followed by TMMC serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-avg-error}
-\end{figure}
+We use the average entropy error vs moves a metric to compare 
+simulation runtimes and overall convergence. The overall accuracy
+is determined by examining the fractional error of a particular method to
+a `golden' standard. For each simulation, the `golden' standard is
+SAMC for the best choice of $t_0$ for $10^{11}$ moves followed by TMMC.  
+\begin{align}
+\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{golden}}\bigg\rvert
+\end{align}
+Simulations that take more iterations to reach the same entropy error as a 
+different method take longer to converge. This behavior is seen in 
+Fig.~\ref{fig:N50-ff0.1-avg-error}.
+
 \begin{figure}
@@
 \begin{figure}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
+  \caption{The average entropy error for each method is shown for a filling fraction of 0.30 with
+  SAMC followed by TMMC serving as the `golden'
+  standard.}\label{fig:N50-ff0.3-avg-error}
+\end{figure}
+\begin{figure}
 \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
@@
 In order to determine how the various methods scale with size,
-simulations are run for systems with different particle numbers. A filling
-fraction $\text{ff} = 0.10$ is chosen which yields $N = 50$ particles
-in the cubical system.  These particles are randomly inserted into the
-system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05$). The boundaries in $x$ and $y$ are periodic while
-the $z$-axis is not periodic.
-
-We use the average entropy error vs iterations~\jpsays{we should
-explain how we do this.} as a metric to compare simulation runtimes and
-overall convergence.  Simulations that take more iterations to reach
-the same entropy error as a different method take longer to converge.
-This behavior is seen in Fig.~\ref{fig:N50-ff0.1-avg-error}.
+simulations are run for systems with different particle numbers. 
+Figure~\ref{fig:N50-ff0.1-avg-error} shows the average entropy vs moves for
+a system consisting of 50 atoms that are placed in a periodic simple cubic unit
+cell with lattice constant chosen to give a filling fraction of 0.10.  These particles
+are randomly inserted into the system and random moves are proposed (with
+a translation scale $\delta_0 = 0.05$).
 
@@
 
-Figure~\ref{fig:N50-ff0.2-avg-error} shows a system consisting of 50 atoms
-that are placed in a periodic simple cubic unit cell with lattice constant
-chosen to give a filling fraction of 0.20.  These particles are
-randomly inserted into the system and random moves are proposed (with
+Figure~\ref{fig:N50-ff0.2-avg-error} shows the average entropy vs moves for
+a system consisting of 50 atoms that are placed in a periodic simple cubic unit
+cell with lattice constant chosen to give a filling fraction of 0.20.  These particles
+are randomly inserted into the system and random moves are proposed (with
 a translation scale $\delta_0 = 0.05$).
@@
 
-Figure~\ref{fig:N50-ff0.3-avg-error} shows a system consisting of 50 atoms
-that are placed in a periodic simple cubic unit cell with lattice constant
-chosen to give a filling fraction of 0.30.  These particles are
-randomly inserted into the system and random moves are proposed (with
+Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves for
+a system consisting of 50 atoms that are placed in a periodic simple cubic unit
+cell with lattice constant chosen to give a filling fraction of 0.30.  These particles
+are randomly inserted into the system and random moves are proposed (with
 a translation scale $\delta_0 = 0.05$).
@@
 into the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05$). The boundaries in $x$ and $y$ are periodic while
-the $z$-axis is not periodic.
+$\delta_0 = 0.05$).
 
@@
 
-A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$ for the 
-Wang-Landau algorithm decreases much more rapidly than $\nicefrac{1}{\sqrt{t}}$.
-It can be observed that all the methods from this family struggle to converge if
-the update factor decreases faster than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.2}
-shows both WL and SAMC with inappropriate choices of $t_0$ i.e. less than 
-$\nicefrac{10^5}{\sqrt{t}}$ failing to converge in a reasonable amount of time.
-Thus, the update factor $\gamma$, shown as a function of simulation moves
-in Fig.~\ref{fig:gamma-vs-t}, proves useful in determining which methods will 
-converge in a reasonable amount of time.
-We compare each method to a `golden' standard which we define to be an ideal choice
-of SAMC for $10^11$ moves followed by TMMC.  This gives an unbiased comparison with
-our developed method SAD.
+A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
+for the Wang-Landau algorithm decreases much more rapidly than
+$\nicefrac{1}{\sqrt{t}}$. It can be observed that all the methods from
+this family struggle to converge if the update factor decreases faster
+than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.2} shows both
+WL and SAMC with inappropriate choices of $t_0$ i.e. less than
+$\nicefrac{10^5}{\sqrt{t}}$ failing to converge in a reasonable amount
+of time. Thus, the update factor $\gamma$, shown as a function of
+simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves useful in
+determining which methods will converge in a reasonable amount of time.
+We compare each method to a `golden' standard which we define to be an
+ideal choice of SAMC for $10^{11}$ moves followed by TMMC.  This gives an
+unbiased comparison with our developed method SAD. The dynamic version
+of SAMC which we call SAD performs as well as SAMC (assuming the user
+can correctly identify an appropriate choice of $t_0$).  SAD
+outperforms WL and WL-TMMC for the same given number of moves. The
+absence of user defined parameters makes SAD an exciting alternative to
+methods where user experience or successive implementation iteration
+are needed to achieve reasonable results.
 
@@
 
-We introduced a new algorithm SAD (which is Stochastic Approximation
+We have introduced a new algorithm SAD (which is Stochastic Approximation
 with the modification factor dynamically chosen) that effectively
@@
 implementation and the absence of user-defined parameters makes SAD an
-ideal algorithm.
+ideal algorithm. SAD converges as quickly as SAMC (in the case of an ideally
+chosen $t_0$ for our simulations. SAD doesn't suffer from the weakness of SAMC,
+i.e. the user choosing parameters that could lead to the simulation not
+converging in a reasonable amount of time. SAD also converges more quickly
+than TMMC (the reader will remember that SAMC followed by TMMC is what each
+method is compared against) and much more quickly than WL-TMMC.
 
-~\jpsays{We need about a paragraph here that says exactly how much
-better SAD performed than other methods and or how easy it was to
-implement (Likely best added once simulations run a little longer).}
+SAD decreases $\gamma$ in a controlled way that prevents the algorithm
+from taking too long to converge (as in the case of SAMC with a poor choice
+of $t_0$) or failing to converge such as is the case with WL for complex
+systems. A major advantage is that the user does not need to determine a
+priori how quickly the update factor $\gamma$ should be decreased as the
+method SAD dynamically determines this.
 
 We note that SAD should be tested on additional systems, and the
-accuracy of the method should be examined.  Also, a rigorous
+convergence of the method should be examined.  Also, a rigorous
 mathematical justification for our choice of the update factor is

### 2018-05-16 papers/histogram/sad.tex
start adding a figure comparing the entropy for N500
adding text to paper sad.

@@
 \subsubsection{The theory behind SAD}
+~\jpsays{We need to define or talk about Tmin.  This is a disadvantage
+of of SAD. WL requires an energy range. SAMC removes the energy range
+requirement but requires simulating every possible energy.  For SAD,
+Tmax set to $\infty$ is fine.  Easy to simulate this way. SAD focuses
+on energies between Tmin and infinity.}
+
 The SAD (Dynamic Stochastic Approximation) is a version of the SAMC
@@
 the triangle made by connecting the minimum important energy and the
-maximum entropy state.  The factor of $3$ comes from the ratio of areas bounded
-by the $\ln{D(E)}$ and the minimum temperature line.
-The update factor has the same $1/t$ behavior as the original SAMC algorithm;
-however now every time a new energy is discovered during the course
-of the simulation, the update factor experiences a jump. This behavior allows
-SAD to dynamically prevent the update factor from decreasing too rapidly.
-
-Fig.~\ref{fig:gamma-vs-t}. shows the monotonic decrease of the update
-factor for SAMC.  The update factor for WL decreases like a step
-function while the update factor for SAD dynamically fluctuates around
-a value less than 1 for early MC moves (while finding new energy
-states).  Ultimately once all energy states have been found, the update
-factor for SAD decreases in the same monotonic fashion as SAMC.
+maximum entropy state.  The factor of $3$ comes from the ratio of areas
+bounded by the $\ln{D(E)}$ and the minimum temperature line. The update
+factor has the same $1/t$ behavior as the original SAMC algorithm;
+however now every time a new energy is discovered during the course of
+the simulation, the update factor experiences a jump. This behavior
+allows SAD to dynamically prevent the update factor from decreasing too
+rapidly. In contrast to SAD, Fig.~\ref{fig:gamma-vs-t}. shows the
+monotonic decrease of the update factor for SAMC.  The update factor
+for WL decreases like a step function while the update factor for SAD
+dynamically fluctuates around a value less than 1 for early MC moves
+(while finding new energy states).  Ultimately once all energy states
+have been found, the update factor for SAD decreases in the same
+monotonic fashion as SAMC.
 
 \subsubsection{High and Low energy racheting}
-Since SAD explores all energy states it is imperative that the method have
-a reasonable idea about where the minimum important energy and maximum
-entropy state are located in energy space. Since these are not user input
-parameters like for methods suchas WL-TMMC, the simulation is responsible
-for determining and updating these energy locations. In order to prevent,
-the method from experiencing errors near the energy extrema, we employ a 
-high energy and a low energy to define the range over which the energy 
-histogram is made flat. The high energy is the highest ever value that the
-maximum entropy state has taken while the low energy is the lowest ever value
-that the minimum important energy has taken. Together these two energies serve
-as a kind of ratcheting effect to determine the energy landscape boundaries.
+~\jpsays{change name to $\epsilon_H$ and $\epsilon_L$. Include formula
+for updating weights above and below high/low energy.  Explain how we
+compute the maximum entropy state and minimum important energy.  Then
+we are in a position to talk about racheting.} Since SAD explores all
+energy states it is imperative that the method have a reasonable idea
+about where the minimum important energy and maximum entropy state are
+located in energy space. Since these are not user input parameters like
+for methods suchas WL-TMMC, the simulation is responsible for
+determining and updating these energy locations. In order to prevent,
+the method from experiencing errors near the energy extrema, we employ
+a high energy $\epsilon_H$ and a low energy $\epsilon_L$ to define the
+range over which the energy histogram is made flat. $\epsilon_H$ is the
+highest ever value that the maximum entropy state has taken while
+$\epsilon_L$ is the lowest ever value that the minimum important energy
+has taken. SAD uses these two energies to control the weights $w(E)$
+above $\epsilon_H$ and $\epsilon_L$. If the energy $E$ is less than
+$\epsilon_H$ ~\jpsays{greater than sign backward} then the weights can
+be calculated as ~\jpsays{equation formatting problems!}
+\begin{equation}
+\begin{aligned}
+\ln w(E) = \min(\ln w(E) - \gamma, \\
+    \ln w(\epsilon_H) - \log(e^{\ln w(\epsilon_H) -w(E)}))
+\end{aligned}
+\end{equation}
+If the energy $E$ is greater than $\epsilon_L$ ~\jpsays{greater than
+sign backward} then the weights can be calculated as
+\begin{equation}
+\begin{aligned}
+\ln w(E) = \min(\ln w(E) - \gamma, \ln w(\epsilon_L) \\
+    + \frac{E-\epsilon_L}{T_\text{min}}
+    - \log(\gamma + e^{\ln w(\epsilon_L) -w(E)
+    + \frac{E-\epsilon_L}{T_\text{min}}}))
+\end{aligned}
+\end{equation}
+Sad sets the maximum entropy state equal to the energy if either of the
+following is true.~\jpsays{less than sign backward}
+\begin{align}
+\ln w(E) < \ln w(S_\text{max}) \; \text{or} \; H(S_\text{max}) = 0
+\end{align}
+Thus if the method is beyond the maximum entropy state or has collected no
+statistics at the maximum entropy state then it is set equal to the energy.
+Similarly, Sad sets the minimum important energy equal to the energy if
+either of the following is true.~\jpsays{check less than sign}
+~\jpsays{equation formatting problems!}
+\begin{align}
+\ln w(E) < \ln w(E_\text{min}) + \frac{E-E_\text{min}}{T_\text{min}} 
+\; \text{or} \; H(E_\text{min}) = 0
+\end{align}
+Thus if the method is beyond the minimum important energy or has collected no
+statistics at the minimum important energy then it is set equal to the energy.
+
+Together these two
+energies serve as a kind of ratcheting effect to determine the energy
+landscape boundaries.
+~\jpsays{So outside the range set by the weights $w(E)$, the maximum entropy
+state is set to the energy $E$. Likewise with minimum important energy.}
 

### 2018-05-21 papers/histogram/sad.tex
added some text on Tmin and fixed some SAD formulae.

--- 2018-05-10 papers/histogram/sad.tex
@@
 \maketitle
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/N500-lndos-comparison}
+  \caption{Put this in the right place}
+\end{figure}
 
@@
 $\gamma_0$.  There is an immediate advantage of such an algorithm where
-parameters are chosen independent of system size or type. While for
-SAMC the update factor is defined in the original
-implementation~\cite{liang2007stochastic} in terms of an arbitrary
-tunable parameter $t_0$,
-\begin{align}
-\gamma_{t}^{\text{SA}} = \frac{t_0}{\max(t_0,t)}
-\end{align}
-The update factor $\gamma_{t}^{\text{SA}}$ can be thought of as
-$\nicefrac{\text{d}S}{\text{d}t}$. This tells us that $t_0$ should have
-dimensions of entropy.  We argue that we can write this entropy `term'
-geometrically as an area bounded by the $\ln{D(E)}$ i.e. the entropy.  This
-gives the following
+parameters are chosen independent of system size or type. Each
+Broad-Histogram method has unique advantages and disadvantages.
+Wang-Landau requires an energy range for initialization.  SAMC removes
+this energy range requirement but requires simulating every possible
+energy. Our proposed method SAD requires the user to input
+$T_\text{min}$ which is an immediate disadvantage of the method;
+however, this allows the simulating to define an energy simulation
+range between $T_\text{min}$ and $\infty$. Setting a physical parameter
+such as a minimum temperature $T_\text{min}$ is intuitive vs. a user
+needing to know a prior either an energy range or some unphysical
+parameter $t_0$.
+
+While for SAMC the update factor is defined in the original
+implementation, for SAD the update factor $\gamma_{t}^{\text{SA}}$ can
+be thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
+$t_0$ should have dimensions of entropy.  We argue that we can write
+this entropy `term' geometrically as an area bounded by the $\ln{D(E)}$
+i.e. the entropy. This gives the following
 \begin{align}
@@
 the triangle made by connecting the minimum important energy and the
-maximum entropy state.  The factor of $3$ comes from the ratio of areas bounded
-by the $\ln{D(E)}$ and the minimum temperature line.
-The update factor has the same $1/t$ behavior as the original SAMC algorithm;
-however now every time a new energy is discovered during the course
-of the simulation, the update factor experiences a jump. This behavior allows
-SAD to dynamically prevent the update factor from decreasing too rapidly.
-
-Fig.~\ref{fig:gamma-vs-t}. shows the monotonic decrease of the update
-factor for SAMC.  The update factor for WL decreases like a step
-function while the update factor for SAD dynamically fluctuates around
-a value less than 1 for early MC moves (while finding new energy
-states).  Ultimately once all energy states have been found, the update
-factor for SAD decreases in the same monotonic fashion as SAMC.
+maximum entropy state.  The factor of $3$ comes from the ratio of areas
+bounded by the $\ln{D(E)}$ and the minimum temperature line. The update
+factor has the same $1/t$ behavior as the original SAMC algorithm;
+however now every time a new energy is discovered during the course of
+the simulation, the update factor experiences a jump. This behavior
+allows SAD to dynamically prevent the update factor from decreasing too
+rapidly. In contrast to SAD, Fig.~\ref{fig:gamma-vs-t}. shows the
+monotonic decrease of the update factor for SAMC.  The update factor
+for WL decreases like a step function while the update factor for SAD
+dynamically fluctuates around a value less than 1 for early MC moves
+(while finding new energy states).  Ultimately once all energy states
+have been found, the update factor for SAD decreases in the same
+monotonic fashion as SAMC.
 
 \subsubsection{High and Low energy racheting}
-Since SAD explores all energy states it is imperative that the method have
-a reasonable idea about where the minimum important energy and maximum
-entropy state are located in energy space. Since these are not user input
-parameters like for methods suchas WL-TMMC, the simulation is responsible
-for determining and updating these energy locations. In order to prevent,
-the method from experiencing errors near the energy extrema, we employ a 
-high energy and a low energy to define the range over which the energy 
-histogram is made flat. The high energy is the highest ever value that the
-maximum entropy state has taken while the low energy is the lowest ever value
-that the minimum important energy has taken. Together these two energies serve
-as a kind of ratcheting effect to determine the energy landscape boundaries.
+Since SAD explores all energy states it is imperative that the method
+have a reasonable idea about where the minimum important energy and
+maximum entropy state are located in energy space. Since these are not
+user input parameters such as they are for methods like WL-TMMC, the
+simulation is responsible for determining and updating these energy
+locations. In order to prevent, the method from experiencing errors
+near the energy extrema, we employ a high energy $\epsilon_H$ and a low
+energy $\epsilon_L$ to define the range over which the energy histogram
+is made flat. $\epsilon_H$ is the highest ever value that the maximum
+entropy state has taken while $\epsilon_L$ is the lowest ever value
+that the minimum important energy has taken. SAD uses these two
+energies to control the weights $w(E)$ above $\epsilon_H$ and
+$\epsilon_L$. If the energy $E$ is greater than $\epsilon_H$ then the
+weights can be calculated as ~\jpsays{equation formatting problems!}
+\begin{equation}
+\begin{aligned}
+\ln w(E) = \max(\ln w(E) + \gamma, \\
+    \ln w(\epsilon_H) + \log(e^{w(E)-\ln w(\epsilon_H)}))
+\end{aligned}
+\end{equation}
+If the energy $E$ is less than $\epsilon_L$ then the weights can be
+calculated as
+\begin{equation}
+\begin{aligned}
+\ln w(E) = \max(\ln w(E) + \gamma, \ln w(\epsilon_L) \\
+    + \frac{\epsilon_L - E}{T_\text{min}}
+    + \log(\gamma + e^{w(E) -\ln w(\epsilon_L)
+    + \frac{\epsilon_L -E}{T_\text{min}}}))
+\end{aligned}
+\end{equation}
+Sad sets the maximum entropy state equal to the energy if either of the
+following is true.
+\begin{align}
+\ln w(E) > \ln w(S_\text{max}) \; \text{or} \; H(S_\text{max}) = 0
+\end{align}
+Thus if the method is beyond the maximum entropy state or has collected no
+statistics at the maximum entropy state then it is set equal to the energy.
+Similarly, Sad sets the minimum important energy equal to the energy if
+either of the following is true.
+~\jpsays{equation formatting problems!}
+\begin{align}
+\ln w(E) > \ln w(E_\text{min}) - \frac{E_\text{min}-E}{T_\text{min}}
+\; \text{or} \; H(E_\text{min}) = 0
+\end{align}
+Thus if the method is beyond the minimum important energy or has collected no
+statistics at the minimum important energy then it is set equal to the energy.
+Together these two energies serve as a kind of ratcheting effect to
+determine the energy landscape boundaries. Outside the range set by the
+weights $w(E)$, the maximum entropy state is set to the energy $E$. The
+minimum important energy is also set to the energy when outside the
+calculated range.
 

### 2018-06-25 papers/histogram/sad.tex
spell checks in paper

@@
 that decrease faster than $1/t^2$ leads to
-nonconvergence~\cite{belardinelli2007fast}.  This lead to the so-called
+nonconvergence~\cite{belardinelli2007fast}.  This leads to the so-called
 $1/t$-WL algorithm which ensures that excess CPU time is not wasted by
-continuing to perfrom calculations once the error in the density of
+continuing to perform calculations once the error in the density of
 states becomes saturated~\cite{belardinelli2008analysis}. An effort to
@@
 theory for why the method converges despite detailed balance being
-violated although infrequently was also being udertaken.
+violated although infrequently was also being undertaken.
 
@@
 Wang-Landau's major tenant is that when counting the histogram
-$H(\epsilon)$, the energy occurences should form a flat distribution.
+$H(\epsilon)$, the energy occurrences should form a flat distribution.
 The criteria for flat sampling is given by
@@
 \subsubsection{The theory behind SAD}
-The SAD (Dynamic Stochastic Approximation) is a version of the SAMC
+The SAD (Dynamic Stochastic Approximation) method is a version of the SAMC
 Algorithm that attempts to dynamically choose the modification factor
@@
 $T_\text{min}$ which is an immediate disadvantage of the method;
-however, this allows the simulating to define an energy simulation
+however, this allows the user to define an energy simulation
 range between $T_\text{min}$ and $\infty$. Setting a physical parameter

### 2018-06-25 papers/histogram/sad.tex
simplify sad algorithm at extreme energies (MUST RERUN ALL SAD DATA!)

@@
 \ln w(E) = \max(\ln w(E) + \gamma, \\
-    \ln w(\epsilon_H) + \log(e^{w(E)-\ln w(\epsilon_H)}))
+    \ln w(\epsilon_H) + \log(\gamma + e^{\ln w(E)-\ln w(\epsilon_H)}))
 \end{aligned}
@@
     + \frac{\epsilon_L - E}{T_\text{min}}
-    + \log(\gamma + e^{w(E) -\ln w(\epsilon_L)
+    + \log(\gamma + e^{\ln w(E) -\ln w(\epsilon_L)
     + \frac{\epsilon_L -E}{T_\text{min}}}))
@@
 calculated range.
+
+\paragraph{Accepting moves above and below}
+We do \emph{not} use the weights to determine move acceptance outside
+the middle range.  FIXME WRITE THIS SECTION.
+
+\paragraph{Above $\epsilon_H$}
+Each energy $E>\epsilon_H$, should be visited at a rate that
+is proportional to the density of states at that energy, as in a
+simulation with constant weights.  Thus the weight at each energy
+needs to be proportional to the number of times that weight has been
+visited.  In order to ensure that $w(E) \propto D(E)$, the change in
+weight with each visit should be equal to the change in weight at
+the border energy $E=\epsilon_H$.
+\begin{align}
+  \Delta w(E) &= \Delta w(\epsilon_H) \\
+  w_{\text{new}}(E)-w_{\text{old}}(E)
+    &= e^\gamma w(\epsilon_H) - w(\epsilon_H) \\
+    &= (e^\gamma-1) w(\epsilon_H)
+\end{align}
+This ensures that although we visit each energy above $\epsilon_H$ only
+in proportion to its density of states, the weights at these energies
+will approach the entropy as the simulation progresses.
+
+\paragraph{Below $\epsilon_L$}
+For energies $E<\epsilon_L$, we will visit each state in proportion to
+the Boltzmann distribution at temperature $T_{\min}$.  This ensures that
+we have adequate statistics to describe the system at temperature
+$T_{\min}$ while not wasting time simulating extremely low energies with
+correspondingly low entropy.  This means that we need to shift the
+weights at lower energies by a smaller amount in proportion to the
+Boltzmann factor.
+\begin{align}
+  \Delta w(E) &= \Delta w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}} \\
+  w_{\text{new}}(E)-w_{\text{old}}(E)
+    %&= \left(e^\gamma w(\epsilon_L) - w(\epsilon_L)\right)e^{(E-\epsilon_L)/T_{\min}} \\
+    &= (e^\gamma-1) w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}}
+\end{align}
+We note that when updating the natural logarithms of the weights, care
+is required in order to ensure that overflow does not occur.
 

### 2018-06-25 papers/histogram/figs/ising.cpp
changed gamma to use ratcheting energies.

@@ -58,6 +58,9 @@ struct energy {
   double operator*(double other) const { return value*other; }
   double operator/(double other) const { return value/other; }
 };
+double operator*(double other, energy e) {
+  return other*e.value;
+}
 
 enum dos_types { histogram_dos, transition_dos, weights_dos };
 
@@ -269,8 +272,8 @@ void ising_simulation::end_flip_updates(){
   }
 
   if (param.sa_t0 || param.use_sad) {
-    if (param.use_sad && energies_found > 1) {
-      gamma = param.sa_prefactor*energies_found*(max_energy.value-min_energy.value)
+    if (param.use_sad && too_hi_energy > too_lo_energy) {
+      gamma = param.sa_prefactor*energies_found*(too_hi_energy-too_lo_energy)
         /(3*param.minT*moves);
     } else {
       gamma = param.sa_prefactor*param.sa_t0/max(param.sa_t0, moves);

### 2018-06-25 papers/histogram/sad.tex
changed gamma to use ratcheting energies.
edits to paper (should we merge methods with history bits of introduction?)

@@
 
-\section{Methods}
+\davidsays{Fix this, it was moved when I reorganized the methods
+  section} In this work, We compare a variety of broad histogram
+methods.  We outline the general workings of each algorithm that we
+developed in detail while summarizing algorithms that originated
+elsewhere.  The following methods are introduced and applied to the
+square-well fluid: Transition Matrix Monte-Carlo (TMMC), Wang-Landau
+(WL), Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC), Stochastic
+Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
+(SAD).
 
-In this work, We compare a variety of broad histogram methods.  We
-outline the general workings of each algorithm that we developed in
-detail while summarizing algorithms that originated elsewhere.  The
-following methods are introduced and applied to the square-well fluid:
-Transition Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau
-Transition Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation
-Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
-(SAD).
+\section{Broad histogram methods}
+
+The goal of broad histogram methods (also called \emph{flat histogram}
+or \emph{multicanonical} methods) is to simulate each energy with
+similar accuracy, so as to accurately determine the density of states
+over a broad range of energies.  Once the density of states is known,
+a number of thermodynamic quantities---such as heat capacity or
+internal energy---can be easily computed for any temperature.
+Properties that require more information---such as a spatial
+correlation function or a response function---can still be computed
+for any temperature, provided statistics are collected for each
+individual energy, which can then be reweighted for any temperature.
+
+Each broad histogram Monte Carlo method uses a given set of ``moves''
+which change a system and should satisfy detailed balance.  Each
+algorithm differs in how it determines the probability of accepting a
+move, and in what additional statistics (if any) must be collected in
+order decide on that probability.  The methods we will compare fall in
+two categories:  weight-based methods, which only require a set of
+weights $w(E)$ be collected, and transition-matrix methods that
+require that statistics be collected for any transitions between pairs
+of energies.
+
+\davidsays{We should discuss continuous energies and binning somewhere
+  here...}
+
+\section{Weight-based methods}
+
+We will begin by introducing two related weight-based methods which
+rely on a weight function $w(E)$, noting that the method proposed in
+this paper (discussed in Section~\ref{sec:sad}) also falls in this
+category.  In these algorithms, the probability of accepting a move is
+given by
+\begin{equation}
+	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
+	= \min\left[1,\frac{w(\epsilon_\text{old})}{w(\epsilon_\text{new})}\right]
+\end{equation}
+This biases the simulation in favor of energies with low weights.  A
+set of weights that are proportional to the density of states of the
+system $D(\epsilon)$ will result in an entirely flat histogram.  Thus
+the goal of each of these methods is to achieve a flat histogram.  For
+numerical reasons, it is common to store and work with the natural
+logarithm of the weights, since the weights may vary over more than a
+few hundred orders of magnitude.  Note that the logarithm of the
+density of states is the entropy at that energy, so we can think of
+the logarithm of the weights as an approximation of the entropy
+\davidsays{make this less colloquial}.
+
+Each weighted density approach use a random walk in energy space to
+estimate the density of states.  The core of these approaches
+is to continuously update the weights at each step of the simulation
+\begin{equation}
+	\ln{w_{t+1}(\epsilon)}=\ln{w_{t}(\epsilon)}
+	+\gamma
+\end{equation}
+where $t$ is number of the current move, and $\gamma$ is an update
+factor.  This update causes the random walk to avoid energies that
+have been frequently sampled, leading to a rapid exploration of energy
+space.  This approach, however, violates detailed balance, due to the
+acceptance probabilities changing with each move.  The severity of
+this violation decreased as we decrease $\gamma$.  The various
+weight-based methods discussed here differ in how the schedule the
+decreasing of $\gamma$.
+
+\subsection{Wang-Landau}
+
+Wang-Landau's approach~\cite{wang2001efficient,wang2001determining,
+  landau2014guide} begins with $\gamma=1$, and then decreases $\gamma$
+in discrete stages.  We track the number of steps at each energy during
+each stage in a histogram.  When that histogram is sufficiently flat,
+$\gamma$ is decreased by a specified factor, which is usually
+$\frac12$.  The flatness is defined by the ratio between the minimum
+value of the histogram and its average value.  When this flatness
+reaches a specified threshold (typically 0.8), the $\gamma$ value is
+decreased.  This approach requires that the energy range of interest
+be known in advance, and difficulties can occur with this flatness
+criteria due to the fact that some energies in this energy range might
+never be sampled~\cite{haber2014transition}.  The entire process is
+repeated until $\gamma$ reaches a desired cutoff is reached.
+
+The Wang-Landau approach thus has three parameters that need be
+specified: the factor by which to decrease $\gamma$ when flatness is
+acheived, the flatness criterion, and the cutoff that determines when
+the computation is complete.  In addition, an energy range (or more
+specifically, a set of energies) must be supplied, so that the
+flatness criterion can be defined.
+
+\davidsays{We need to look up these references.}  While this approach
+is very efficient and has been widely uses, it suffers a few
+shortcomings.  Firstly, the set of energies must be specified, which
+may require multiple simulations.  Secondly, while Wang-Landau
+converges quickly, it does not in general converge to the true density
+of states.  It can and does decrease $\gamma$ so quickly that it will
+never (for any cutoff value) decrease the error in the density of
+states beyond a given value.
+
+\subsection{SAMC}
+The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses
+the lack of convergence of Wang-Landau's approach with a simple
+schedule by which the update factor $\gamma$ is continuously
+decreased~\cite{liang2007stochastic, werlich2015stochastic,
+  schneider2017convergence}.  The update factor is defined in the
+original implementation~\cite{liang2007stochastic} in terms of an
+arbitrary tunable parameter $t_0$ and a scale factor
+$\gamma_0$. \davidsays{Did it originally have a scale factor?}
+\begin{align}
+\gamma_{t}^{\text{SA}} =\gamma_0 \frac{t_0}{\max(t_0,t)}
+\end{align}
+where as above $t$ is the number of moves that have been attempted.
+The major advantage that SAMC offers is its proven convergence.
+Provided the update factor satisfies
+\begin{align}
+\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
+\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
+\end{align}
+where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
+to converge to the true density of states~\cite{liang2006theory,
+  liang2007stochastic}.  In addition, the energy range need not be
+known a priori.  The time to converge depends only on the choice of
+parameters $t_0$ and $\gamma_0$.  Unfortunately, there is no
+prescription for finding acceptable values for these parameters, and
+while the algorithm formally converges, for a poor choice of $t_0$
+that convergence may be far too slow to be practical.  This work
+proposes a means for dynamically setting the update factor $\gamma_t$
+that should robustly converge on a variety of systems.
+
+\section{Transition-matrix methods}
 
@@
 determine the particle number probability distribution (PNPD).
-
-\subsection{Wang-Landau Algorithm}
-
-The Wang-Landau Algorithm developed by Wang and Landau proposes a
-random walk in energy space in order to estimate the density of
-states~\cite{wang2001efficient,wang2001determining, landau2014guide}.
-Wang-Landau's major tenant is that when counting the histogram
-$H(\epsilon)$, the energy occurrences should form a flat distribution.
-The criteria for flat sampling is given by
-\begin{equation}
-	\frac{\min_{\epsilon} H(\epsilon)}
-	{\big\langle H(\epsilon)\big\rangle }
-	> \eta
-\end{equation}
-where $\eta$ is usually set to $0.80$ and determines how many times
-each energy is sampled relative to the mean number of visits.
-Difficulties can occur with this flatness criteria due to the fact that
-some energies on a given energy range might never be
-sampled~\cite{haber2014transition}. The algorithm is briefly outlined
-below where $w$ represents the weights:
-\begin{equation}
-	\ln{D_{i+1}(\epsilon,N)}=\ln{D_{i}(\epsilon,N)}
-	+\gamma
-\end{equation}
-\begin{equation}
-	\gamma_{i+1}=\frac{u}{2} \gamma_{i}
-\end{equation}
-for $u$ typically greater than 1 and $\gamma$ is the modification
-factor for the density of states.  A new configuration for the density
-of states with a probability given by the following is chosen:
-\begin{equation}
-	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
-	= \min[1,e^{\ln{D(\epsilon_\text{old})}-\ln{D(\epsilon_\text{new})}}]
-\end{equation}
-The entire process is repeated until a desired cutoff is reached.
 
@@
 
-\subsection{SAMC Algorithm}
-SAMC can be applied as a dynamic importance sampling algorithm that
-uses an update factor to explore energy
-space~\cite{liang2007stochastic, werlich2015stochastic,
-schneider2017convergence}.  The update factor is defined in the
-original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$ and a scale factor $\gamma_0$.
-\begin{align}
-\gamma_{t}^{\text{SA}} =\gamma_0 \frac{t_0}{\max(t_0,t)}
-\end{align}
-SAMC offers several advantages over other broad histogram methods
-including proven convergence. It has been shown that if
-$F_{t}^{\text{SA}}$ satisfies two conditions the estimator is proven to
-converge to the real DOS~\cite{liang2006theory, liang2007stochastic}.
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t}^{\text{SA}} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty (\gamma_{t}^{\text{SA}})^\zeta < \infty
-\end{align}
-where $\zeta \in \{1,2\}$.  Unlike WL and WL-TMMC, the energy range need
-not be known a priori.  The time to converge depends only on the choice of
-parameters $t_0$ and $\gamma_0$.
-Unfortunately, only by performing many computational runs might the user
-determine what SAMC arbitrary parameters to choose.  This has lead us
-to develop a systematic way (which we will discuss in the next section)
-for the simulation to determine what $t_0$ should be.
-
-\section{SAD Algorithm}
+\section{SAD Algorithm}\label{sec:sad}
 \subsubsection{The theory behind SAD}
@@
 \gamma_{t}^{\text{SA}} =
-\frac{N_E}{t}\frac{E_{\text{max}}-E_{\text{min}}}{3T_{\text{min}}}
+\frac{N_E}{t}\frac{\epsilon_{H}-\epsilon_{L}}{3T_{\text{min}}}
 \end{align}

### 2018-06-28 papers/histogram/sad.tex
added citations and addressed problems in sad paper.

@@
 
-The organization of this paper is as follows: In Section II, we
-describe in detail several broad-histogram methods used to calculate
-the density of states for the liquid-vapor system.  Section III
-contains the results and comparison of the methods applied to the
-square-well fluid.  In Section IV, we discuss the performance of the
-various methods applied to the model system.
+The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
+describe in detail broad-histogram methods.  Section~\ref{sec:weight} outlines
+weight-based histogram methods used in this work. Section~\ref{sec:transition}
+details the workings of the transition-matrix methods studied in this work.
+Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
+In Section~\ref{sec:results}, we discuss the performance of the various methods 
+applied to the model system.
 
-\davidsays{Fix this, it was moved when I reorganized the methods
-  section} In this work, We compare a variety of broad histogram
+In this work, We compare a variety of broad histogram
 methods.  We outline the general workings of each algorithm that we
-developed in detail while summarizing algorithms that originated
-elsewhere.  The following methods are introduced and applied to the
+developed in detail while summarizing algorithms that were developed in other
+works.  The following methods are introduced and applied to the
 square-well fluid: Transition Matrix Monte-Carlo (TMMC), Wang-Landau
@@
 
-\section{Broad histogram methods}
+\section{Broad histogram methods}\label{sec:histogram}
 
@@
   here...}
+\jpsays{I put some text about this but it was under a sad subsection also
+may need reworked and/or edited.}
 
-\section{Weight-based methods}
+\section{Weight-based methods}\label{sec:weight}
 
@@
 \end{equation}
-This biases the simulation in favor of energies with low weights.  A
-set of weights that are proportional to the density of states of the
+the probability biases the simulation in favor of energies with low weights.  
+A set of weights that are proportional to the density of states of the
 system $D(\epsilon)$ will result in an entirely flat histogram.  Thus
-the goal of each of these methods is to achieve a flat histogram.  For
-numerical reasons, it is common to store and work with the natural
-logarithm of the weights, since the weights may vary over more than a
-few hundred orders of magnitude.  Note that the logarithm of the
-density of states is the entropy at that energy, so we can think of
-the logarithm of the weights as an approximation of the entropy
-\davidsays{make this less colloquial}.
+flat histogram is a criteria for convergence for these methods.  To avoid 
+overflow error , since the weights may vary over more than a
+few hundred orders of magnitude, the natural logarithm of the weights are used 
+in this work.  Since $D(\epsilon) = S(\epsilon)$, the logarithm of the weights 
+can be thought of as an approximation of the entropy.
 
-Each weighted density approach use a random walk in energy space to
+Each weighted density approach uses a random walk in energy space to
 estimate the density of states.  The core of these approaches
@@
 this violation decreased as we decrease $\gamma$.  The various
-weight-based methods discussed here differ in how the schedule the
-decreasing of $\gamma$.
+weight-based methods discussed here differ in how the decreasing of $\gamma$
+is scheduled.
 
@@
 never be sampled~\cite{haber2014transition}.  The entire process is
-repeated until $\gamma$ reaches a desired cutoff is reached.
+repeated until $\gamma$ reaches a desired cutoff.
 
@@
 
-\davidsays{We need to look up these references.}  While this approach
-is very efficient and has been widely uses, it suffers a few
-shortcomings.  Firstly, the set of energies must be specified, which
-may require multiple simulations.  Secondly, while Wang-Landau
-converges quickly, it does not in general converge to the true density
-of states.  It can and does decrease $\gamma$ so quickly that it will
-never (for any cutoff value) decrease the error in the density of
-states beyond a given value.
+While this approach
+is very efficient and has been widely used, it suffers a few
+shortcomings.  Firstly, the set of energies must be specified
+~\cite{wang2001efficient, schulz2003avoiding, yan2003fast}, which may 
+require multiple simulations. Secondly, while Wang-Landau converges quickly, 
+it does not in general converge to the true density of states
+~\cite{belardinelli2008analysis, zhou2008optimal}.  It can and does decrease
+$\gamma$ so quickly that it will never (for any cutoff value) decrease the 
+error in the density of states beyond a given value.
 
@@
 original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$ and a scale factor
-$\gamma_0$. \davidsays{Did it originally have a scale factor?}
+arbitrary tunable parameter $t_0$ and in future works~\cite{werlich2015stochastic} with a scale factor
+$\gamma_0$.
 \begin{align}
@@
 
-\section{Transition-matrix methods}
+\section{Transition-matrix methods}\label{sec:transition}
 
@@
 
-\section{Results}
+\section{Results}\label{sec:results}
 

### 2018-06-29 papers/grand/paper.tex
add embryonic 2d grand paper

--- 2018-06-25 papers/histogram/sad.tex
@@
 \begin{document}
-\title{Dynamic Stochastic Approximation Monte Carlo}
+\title{Two-dimensional Broad Histogram Monte Carlo}
 
-\author{Jordan K. Pommerenck} \author{Tanner T. Simpson}
-\author{Michael A. Perlin} \author{David Roundy}
+\author{Jordan K. Pommerenck}
+\author{Cade MIDDLE INITIAL? Trotter}
+\author{David Roundy}
 \affiliation{Department of Physics, Oregon State University,
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/N500-lndos-comparison}
-  \caption{Put this in the right place}
-\end{figure}
-
-\section{Introduction}
-Over the past several decades a number of broad histogram Monte-Carlo
-simulation algorithms have been developed which calculate the
-thermodynamic properties of various systems for all temperatures.  The
-development began with the original histogram method, which used a
-single canonical Monte Carlo simulation to predict properties for
-nearby temperatures~\cite{ferrenberg1988new}.  For large simulations
-this approach is limited to a narrow temperature range because a single
-canonical simulation explores only a small range of energies.  This led
-to a variety of ``broad'' (or ``flat'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient, trebst2004optimizing}, which
-attempt to explore a wide range of energies.  These methods also
-benefit in that they are unlikely to be trapped in a low enetropy state.
-
-Wang and Landau introduced an algorithm that uses an update factor and
-a statistical histogram to compute the density of states of a given
-system~\cite{wang2001determining, wang2001efficient}.  While the method
-is incredibly powerful, it has several user-defined parameters.  This
-can make its application to a variety of systems something of an
-art-form since the user needs to determine the ideal parameters for the
-particular system being studied.  Also, detailed balance is violated
-(although only for short periods of time), ensuring that convergence is
-not guaranteed.  This raises several questions such as when the method
-does converge, what rate will the method converge to the correct
-density of states. Also, how can the user appropriately decide what
-parameters to choose so that the algorithm solves the problem in the
-most ideal way.
-
-Transition Matrix Monte Carlo~\cite{wang1999transition,
-swendsen1999transition, fitzgerald2000monte} became an attractive
-complimentary simulation algorithm to Wang-Landau, in that, if detailed
-balance is ensured, the system is guaranteed to converge.  Unfortunately,
-the algorithm can be considerably more difficult to implement when
-compared to Wang-Landau: in particular, due to the infinite temperature
-transition matrix~\cite{wang2002transition}.  Also, the time necessary
-for the density of states to converge could be considerable
-due to the algorithm spending too much time exploring low energy states.
-Unlike WL, TMMC does not require a prior knowledge of the energy range of
-the system.
-
-Shell et al.~\cite{shell2003improved, shell2004flat} originially
-implemented Wang-Landau Transition Matrix Monte Carlo (WL-TMMC) in an
-effort to quickly explore the energies of a given system using WL then
-switch over to TMMC to guarantee convergence. Considerable effort has
-been spent trying to determine the flatness criteria and cutoff for the
-WL portion, but there remains no rigorous systematic way to determine
-these parameters~\cite{rane2013monte}.  The algorithm is first run for
-several trial simulations and/or `experience' is used to determine the
-best parameters for a users particular system of
-study~\cite{siderius2013use}.
-
-Around the same time, the convergence rate for WL was being
-explored~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}. It was found that utilizing modification factors
-that decrease faster than $1/t^2$ leads to
-nonconvergence~\cite{belardinelli2007fast}.  This leads to the so-called
-$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
-continuing to perform calculations once the error in the density of
-states becomes saturated~\cite{belardinelli2008analysis}. An effort to
-reduce the number of user defined parameters of WL and formulate a
-theory for why the method converges despite detailed balance being
-violated although infrequently was also being undertaken.
-
-Liang, Liu, and Carrol began to consider whether WL could be considered
-a special case of Stochastic Approximation whose convergence could be
-mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
-2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
-considered a form of Stochastic Approximation Monte Carlo (SAMC). While
-SAMC can guarantee convergence, the method still has a system specific
-user-defined variable which makes applying this algorithm to arbitrary
-systems difficult.
-
-In this work, we have developed a novel algorithm based on SAMC that
-does not require user-defined inputs and therefore should be easily
-applicable to a given system.  We call this method SAD (Dynamic
-Stochastic Approximation), and will discuss it in detail in the methods
-section. We compare it along with several broad histogram methods
-which include TMMC, WL, WL-TMMC, and SAMC.
-
-We consider the square-well fluid i.e. a system of particles whose
-interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The
-square-well potential is an ideal test-bed as it is the simplest model
-that ensures both attractive and repulsive forces are experienced by
-interacting particles~\cite{barker1967-SW-perturbation, vega1992phase}.
-The potential $U(\textbf{r})$ for such a system is given by
-\begin{equation}
- U(\textbf{r})=\begin{cases} \infty &
- \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
- \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
- \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
-\end{equation}
-where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
-reduced range of the potential well, and $\epsilon$ is its depth.
-
-The organization of this paper is as follows: In Section II, we
-describe in detail several broad-histogram methods used to calculate
-the density of states for the liquid-vapor system.  Section III
-contains the results and comparison of the methods applied to the
-square-well fluid.  In Section IV, we discuss the performance of the
-various methods applied to the model system.
-
-\section{Methods}
-
-In this work, We compare a variety of broad histogram methods.  We
-outline the general workings of each algorithm that we developed in
-detail while summarizing algorithms that originated elsewhere.  The
-following methods are introduced and applied to the square-well fluid:
-Transition Matrix Monte-Carlo (TMMC), Wang-Landau (WL), Wang-Landau
-Transition Matrix Monte-Carlo (WL-TMMC), Stochastic Approximation
-Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
-(SAD).
-
-\subsection{TMMC Algorithm}
-The Transition Matrix Monte-Carlo (TMMC) method as introduced by
-Fitzgerald et al. consists of three primary steps.  The first two steps
-are common to the Metropolis algorithm while the third makes use of the
-actual transition probabilities~\cite{fitzgerald2000monte}.
-
-\subsubsection{define a transition probability}
-Assuming that the system under study is in a state denoted old, and a
-proposed transition is made to a new state denoted new, the probability
-is defined as
-\begin{align}
-  p_{old \rightarrow new} = p_{new \rightarrow old}
-\end{align}
-where we have allowed the probabilies to be equal for simplicity of
-presentation.
-
-\subsubsection{define an acceptance probability}
-The probability for accepting a transition from an old state to a new
-one can be defined
-\begin{align}
-  p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
-  {\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
-\end{align}
-where $\alpha_{new\rightarrow old}$ is the probability of generating a
-new system configuration the old.  For conventional Monte-Carlo moves
-$\alpha_{old \rightarrow new} =\alpha_{new\rightarrow old}$ although
-for advanced Monte-Carlo moves this would not be the
-case~\cite{paluch2008comparing, siepmann1990method}.  The probabily of
-observing the system in the old state can be found based on the
-ensemble constraints placed on the system.  Fitzgerald et al.
-considered the system of interest using the canonical ensemble;
-however, the extension to other ensembles is quite natural.
-\begin{align}
-  \pi_{old} = \frac{e^{-\beta E_{old}}}{Z}
-\end{align}
-\subsubsection{define a bookkeeping step}
-The third step in the TMMC algorithm consists of recording the data in
-a collection matrix (the unormalized transition histogram matrix).
-\begin{align}
-  C_{N,\delta} = C_{N,\delta} + p_{a}\\
-  C_{N,0} = C_{N,0} +(1 - p_{a})
-\end{align} If $\delta=0$, that is, the energy state has not been
-changed, then only $C_{N,0}$ is incremented by unity. Fitzgerald et
-al.~\cite{fitzgerald2000monte} showed that defining the third step this
-way results in a uniform improvement over using histogram estimators to
-determine the particle number probability distribution (PNPD).
-
-\subsection{Wang-Landau Algorithm}
-
-The Wang-Landau Algorithm developed by Wang and Landau proposes a
-random walk in energy space in order to estimate the density of
-states~\cite{wang2001efficient,wang2001determining, landau2014guide}.
-Wang-Landau's major tenant is that when counting the histogram
-$H(\epsilon)$, the energy occurrences should form a flat distribution.
-The criteria for flat sampling is given by
-\begin{equation}
-	\frac{\min_{\epsilon} H(\epsilon)}
-	{\big\langle H(\epsilon)\big\rangle }
-	> \eta
-\end{equation}
-where $\eta$ is usually set to $0.80$ and determines how many times
-each energy is sampled relative to the mean number of visits.
-Difficulties can occur with this flatness criteria due to the fact that
-some energies on a given energy range might never be
-sampled~\cite{haber2014transition}. The algorithm is briefly outlined
-below where $w$ represents the weights:
-\begin{equation}
-	\ln{D_{i+1}(\epsilon,N)}=\ln{D_{i}(\epsilon,N)}
-	+\gamma
-\end{equation}
-\begin{equation}
-	\gamma_{i+1}=\frac{u}{2} \gamma_{i}
-\end{equation}
-for $u$ typically greater than 1 and $\gamma$ is the modification
-factor for the density of states.  A new configuration for the density
-of states with a probability given by the following is chosen:
-\begin{equation}
-	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
-	= \min[1,e^{\ln{D(\epsilon_\text{old})}-\ln{D(\epsilon_\text{new})}}]
-\end{equation}
-The entire process is repeated until a desired cutoff is reached.
-
-\subsection{WL-TMMC Algorithm}
-
-Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
-by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
-advantage of the ease of implementation and applicability to a variety
-of systems, while making use of a transition matrix to address WL
-weaknesses, such as its limited statistical accuracy plateau.
-This WL version is more accurate for a given number of simulation steps.
-
-\paragraph{Initializing Wang-Landau:} We begin the simulation by
-defining a minimum important energy and a maximum entropy state. In
-traditional Wang-Landau the WL factor begins at 1.0 and the factor is
-modified by 2.0 until the simulation is cutoff typically around
-$10^{-10}$.  Rane et. al. published an implemenation of
-WL-TMMC~\cite{rane2013monte} that suggested using a WL factor between
-1.0 and $10^{-2}$ with a cutoff $<10^{-5}$. In our implementation of
-WL-TMMC, we set the WL factor to be 1 and the cutoff to be either
-$10^{-4}$ or $10^{-10}$. A single sweep is defined as each macrostate
-being sampled a number of times in this case 10,000 (although 100 is
-often sufficient for small systems). Wang-Landau sets a flatness
-criteria~\cite{wang2001determining, wang2001efficient,
-hatch2015computational, mahynski2017predicting} for the accumulated
-histogram usually between 0.75 to 0.90.  In the original
-implementation, it was found to be sufficient that each macrostate is
-visited at least once (each histogram bin has one
-entry)~\cite{shell2003improved}; however, for other systems this is
-rarely sufficient.  The goal of this WL initialization is to prefill
-the transition matrix for the next portion of the simulation.
-\paragraph{Initializing TMMC:} Prefilling the transition
-matrix is necessary because if numerous zeros existed in the collection
-matrix, the infinite temperature transition matrix would be
-ill-defined.  Sampling over large ranges of density of states would
-therefore take an unacceptable length of time to
-complete~\cite{shell2003improved, shen2014elucidating}.  The transition
-matrix in terms of the collection matrix is given as
-follows:
-\begin{align}
-\widetilde{T}_{\infty}(I\rightarrow J) = \frac{C(I,J)}
-{\sum_{K} C(I,K)}
-\end{align}
-We now run the TMMC portion of the simulation.
-%Siderius does this for 20 additional sweeps (does about 8 sweeps for
-%WL).  We do this until we reach a defined number of iterations. We
-%have not yet implemented a biasing function to control acceptance rate
-%of trial moves but we could? at some point.
-The acceptance probability is written in terms of the infinite temperature
-transition matrix.
-\begin{align}
-  P_{acc}(i\rightarrow j) = \min\bigg[1,\frac{\widetilde{T}_{\infty}(J\rightarrow I)}
-  {\widetilde{T}_{\infty}(I\rightarrow J)}\bigg]
-\end{align}
-This probability is only used to control actual transitions not update
-the collection matrices.
-
-\subsection{SAMC Algorithm}
-SAMC can be applied as a dynamic importance sampling algorithm that
-uses an update factor to explore energy
-space~\cite{liang2007stochastic, werlich2015stochastic,
-schneider2017convergence}.  The update factor is defined in the
-original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$ and a scale factor $\gamma_0$.
-\begin{align}
-\gamma_{t}^{\text{SA}} =\gamma_0 \frac{t_0}{\max(t_0,t)}
-\end{align}
-SAMC offers several advantages over other broad histogram methods
-including proven convergence. It has been shown that if
-$F_{t}^{\text{SA}}$ satisfies two conditions the estimator is proven to
-converge to the real DOS~\cite{liang2006theory, liang2007stochastic}.
-\begin{align}
-\sum_{t=1}^\infty \gamma_{t}^{\text{SA}} = \infty \quad\textrm{and}\quad
-\sum_{t=1}^\infty (\gamma_{t}^{\text{SA}})^\zeta < \infty
-\end{align}
-where $\zeta \in \{1,2\}$.  Unlike WL and WL-TMMC, the energy range need
-not be known a priori.  The time to converge depends only on the choice of
-parameters $t_0$ and $\gamma_0$.
-Unfortunately, only by performing many computational runs might the user
-determine what SAMC arbitrary parameters to choose.  This has lead us
-to develop a systematic way (which we will discuss in the next section)
-for the simulation to determine what $t_0$ should be.
-
-\section{SAD Algorithm}
-\subsubsection{The theory behind SAD}
-The SAD (Dynamic Stochastic Approximation) method is a version of the SAMC
-Algorithm that attempts to dynamically choose the modification factor
-rather than relying on system dependent parameters such as $t_0$ or
-$\gamma_0$.  There is an immediate advantage of such an algorithm where
-parameters are chosen independent of system size or type. Each
-Broad-Histogram method has unique advantages and disadvantages.
-Wang-Landau requires an energy range for initialization.  SAMC removes
-this energy range requirement but requires simulating every possible
-energy. Our proposed method SAD requires the user to input
-$T_\text{min}$ which is an immediate disadvantage of the method;
-however, this allows the user to define an energy simulation
-range between $T_\text{min}$ and $\infty$. Setting a physical parameter
-such as a minimum temperature $T_\text{min}$ is intuitive vs. a user
-needing to know a prior either an energy range or some unphysical
-parameter $t_0$.
-
-While for SAMC the update factor is defined in the original
-implementation, for SAD the update factor $\gamma_{t}^{\text{SA}}$ can
-be thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
-$t_0$ should have dimensions of entropy.  We argue that we can write
-this entropy `term' geometrically as an area bounded by the $\ln{D(E)}$
-i.e. the entropy. This gives the following
-\begin{align}
-\gamma_{t}^{\text{SA}} =
-\frac{N_E}{t}\frac{E_{\text{max}}-E_{\text{min}}}{3T_{\text{min}}}
-\end{align}
-where $N_E$ is the number of energies that have been found. The change
-in the entropy the minimum temperature can be thought of as a slope of
-the triangle made by connecting the minimum important energy and the
-maximum entropy state.  The factor of $3$ comes from the ratio of areas
-bounded by the $\ln{D(E)}$ and the minimum temperature line. The update
-factor has the same $1/t$ behavior as the original SAMC algorithm;
-however now every time a new energy is discovered during the course of
-the simulation, the update factor experiences a jump. This behavior
-allows SAD to dynamically prevent the update factor from decreasing too
-rapidly. In contrast to SAD, Fig.~\ref{fig:gamma-vs-t}. shows the
-monotonic decrease of the update factor for SAMC.  The update factor
-for WL decreases like a step function while the update factor for SAD
-dynamically fluctuates around a value less than 1 for early MC moves
-(while finding new energy states).  Ultimately once all energy states
-have been found, the update factor for SAD decreases in the same
-monotonic fashion as SAMC.
-
-\subsubsection{High and Low energy racheting}
-Since SAD explores all energy states it is imperative that the method
-have a reasonable idea about where the minimum important energy and
-maximum entropy state are located in energy space. Since these are not
-user input parameters such as they are for methods like WL-TMMC, the
-simulation is responsible for determining and updating these energy
-locations. In order to prevent, the method from experiencing errors
-near the energy extrema, we employ a high energy $\epsilon_H$ and a low
-energy $\epsilon_L$ to define the range over which the energy histogram
-is made flat. $\epsilon_H$ is the highest ever value that the maximum
-entropy state has taken while $\epsilon_L$ is the lowest ever value
-that the minimum important energy has taken. SAD uses these two
-energies to control the weights $w(E)$ above $\epsilon_H$ and
-$\epsilon_L$. If the energy $E$ is greater than $\epsilon_H$ then the
-weights can be calculated as ~\jpsays{equation formatting problems!}
-\begin{equation}
-\begin{aligned}
-\ln w(E) = \max(\ln w(E) + \gamma, \\
-    \ln w(\epsilon_H) + \log(\gamma + e^{\ln w(E)-\ln w(\epsilon_H)}))
-\end{aligned}
-\end{equation}
-If the energy $E$ is less than $\epsilon_L$ then the weights can be
-calculated as
-\begin{equation}
-\begin{aligned}
-\ln w(E) = \max(\ln w(E) + \gamma, \ln w(\epsilon_L) \\
-    + \frac{\epsilon_L - E}{T_\text{min}}
-    + \log(\gamma + e^{\ln w(E) -\ln w(\epsilon_L)
-    + \frac{\epsilon_L -E}{T_\text{min}}}))
-\end{aligned}
-\end{equation}
-Sad sets the maximum entropy state equal to the energy if either of the
-following is true.
-\begin{align}
-\ln w(E) > \ln w(S_\text{max}) \; \text{or} \; H(S_\text{max}) = 0
-\end{align}
-Thus if the method is beyond the maximum entropy state or has collected no
-statistics at the maximum entropy state then it is set equal to the energy.
-Similarly, Sad sets the minimum important energy equal to the energy if
-either of the following is true.
-~\jpsays{equation formatting problems!}
-\begin{align}
-\ln w(E) > \ln w(E_\text{min}) - \frac{E_\text{min}-E}{T_\text{min}}
-\; \text{or} \; H(E_\text{min}) = 0
-\end{align}
-Thus if the method is beyond the minimum important energy or has collected no
-statistics at the minimum important energy then it is set equal to the energy.
-Together these two energies serve as a kind of ratcheting effect to
-determine the energy landscape boundaries. Outside the range set by the
-weights $w(E)$, the maximum entropy state is set to the energy $E$. The
-minimum important energy is also set to the energy when outside the
-calculated range.
-
-\paragraph{Accepting moves above and below}
-We do \emph{not} use the weights to determine move acceptance outside
-the middle range.  FIXME WRITE THIS SECTION.
-
-\paragraph{Above $\epsilon_H$}
-Each energy $E>\epsilon_H$, should be visited at a rate that
-is proportional to the density of states at that energy, as in a
-simulation with constant weights.  Thus the weight at each energy
-needs to be proportional to the number of times that weight has been
-visited.  In order to ensure that $w(E) \propto D(E)$, the change in
-weight with each visit should be equal to the change in weight at
-the border energy $E=\epsilon_H$.
-\begin{align}
-  \Delta w(E) &= \Delta w(\epsilon_H) \\
-  w_{\text{new}}(E)-w_{\text{old}}(E)
-    &= e^\gamma w(\epsilon_H) - w(\epsilon_H) \\
-    &= (e^\gamma-1) w(\epsilon_H)
-\end{align}
-This ensures that although we visit each energy above $\epsilon_H$ only
-in proportion to its density of states, the weights at these energies
-will approach the entropy as the simulation progresses.
-
-\paragraph{Below $\epsilon_L$}
-For energies $E<\epsilon_L$, we will visit each state in proportion to
-the Boltzmann distribution at temperature $T_{\min}$.  This ensures that
-we have adequate statistics to describe the system at temperature
-$T_{\min}$ while not wasting time simulating extremely low energies with
-correspondingly low entropy.  This means that we need to shift the
-weights at lower energies by a smaller amount in proportion to the
-Boltzmann factor.
-\begin{align}
-  \Delta w(E) &= \Delta w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}} \\
-  w_{\text{new}}(E)-w_{\text{old}}(E)
-    %&= \left(e^\gamma w(\epsilon_L) - w(\epsilon_L)\right)e^{(E-\epsilon_L)/T_{\min}} \\
-    &= (e^\gamma-1) w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}}
-\end{align}
-We note that when updating the natural logarithms of the weights, care
-is required in order to ensure that overflow does not occur.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n500}
-  \caption{The update factor $\gamma$ for $N=500$ versus iteration for three
-    different methods: WL, SAMC, and SAD.
-    method.}\label{fig:gamma-vs-t}
-\end{figure}
-
-\subsubsection{Energy binning with SAD}
-
-Energy binning is an important consideration for complex continous
-systems. Coarse binning yields fewer energy states for the walker to
-explore but more time spent in the individual bin. Fine binning yields
-more energy states to explore but less time spent in each bin.  Some
-methods such as AdaWL~\cite{koh2013dynamically} (discussed elsewhere)
-employ a tunable mechanism for controlling the binning for low entropic
-states in order to ensure the exploration of all energies. A
-significant advantage of SAD over other broad histogram methods is that
-energy binning is solved dynamically. SAD should perform independently
-for a reasonable range of binning because $\gamma \propto
-\nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$
-increases (fine binning), the time spent $t$ in each bin will decrease.
-This argument explains why SAD should perform well for complex systems
-in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
-similar way but this adds yet another parameter for the user to choose.
-
-\subsubsection{Scaling, step size, and SAD}
-Methods such as WL can fail to converge for complex systems in part
-because the same random MC moves are performed for the entire energy
-range.  The main problem lies in trying to determine how to optimize
-the step size such that the time to explore all energies is reasonable.
-Small step sizes lead to improved sampling at the expense of slow
-convergence. Koh et al. proposed multiplying $\gamma$ by an energy
-dependent factor in order to improve the converge of WL in regards to
-convergence and scaling~\cite{koh2013dynamically}.  Because the update
-factor for SAD has an energy dependent factor already present (and
-dynamically adjusted), we expect that SAD will perform consitently
-scaling free for a reasonable choice of step size.  SAD should perform
-extremely well compared with other broad histogram methods when applied
-to complex continuous systems such as polymers/proteins and spin models.
-
-\section{Results}
-
-For the square-well fluid, our simulations treat systems with a
-well-width of $\text{ww} = 1.30$ and a filling fraction of either
-$\text{ff} = 0.10, 0.20,$ or $0.30$.  The simulation explores the
-energy space of the system down to a reduced temperature of
-$T_{\text{min}} = 0.2$.  All simulations lead to the minimum important
-energy $E_{\text{min}}$ and maximum entropy state $S_{\text{max}}$
-being calculated (with the exception of the WL-TMMC method where both
-of these parameters are needed a priori).  In the sections below, we
-examine systems with different particles numbers and filling fractions.
-
-We use the average entropy error vs moves a metric to compare
-simulation runtimes and overall convergence. The overall accuracy
-is determined by examining the fractional error of a particular method to
-a `golden' standard. For each simulation, the `golden' standard is
-SAMC for the best choice of $t_0$ for $10^{11}$ moves followed by TMMC.
-\begin{align}
-\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{golden}}\bigg\rvert
-\end{align}
-Simulations that take more iterations to reach the same entropy error as a
-different method take longer to converge. This behavior is seen in
-Fig.~\ref{fig:N50-ff0.1-avg-error}.
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown for a filling fraction of 0.10 with
-  SAMC followed by TMMC serving as the `golden' standard.}\label{fig:N50-ff0.1-avg-error}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-entropy-error.pdf}
-  \caption{The average entropy error is shown for a filling fraction of 0.20 with TMI acting as the
-  `golden' standard.}\label{fig:N50-ff0.2-avg-error}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown for a filling fraction of 0.30 with
-  SAMC followed by TMMC serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-avg-error}
-\end{figure}
-\begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
-  \caption{The average entropy error is shown for a filling fraction of 0.30 with TMI acting as the
-  `golden' standard.}\label{fig:n500-ff0.2}
-\end{figure}
-
-\subsubsection{A periodic system with 10\% filling fraction and $(N = 50)$}
-In order to determine how the various methods scale with size,
-simulations are run for systems with different particle numbers.
-Figure~\ref{fig:N50-ff0.1-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.10.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
-
-\subsubsection{A periodic system with 20\% filling fraction and $N = 50$}
-
-Figure~\ref{fig:N50-ff0.2-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.20.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
-
-\subsubsection{A periodic system with 30\% filling fraction and $N = 50$}
-
-Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.30.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
-
-\subsubsection{A periodic system with 30\% filling fraction and $N = 500$}
-
-A system is configured with dimensions $x = y = z = 19\sigma$.  A
-filling fraction $\text{ff} = 0.30$ is chosen which yields $N = 500$
-particles in a cubical system.  These particles are randomly inserted
-into the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05$).
-
-\subsubsection{Discussion on convergence and update factor}
-
-A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
-for the Wang-Landau algorithm decreases much more rapidly than
-$\nicefrac{1}{\sqrt{t}}$. It can be observed that all the methods from
-this family struggle to converge if the update factor decreases faster
-than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.2} shows both
-WL and SAMC with inappropriate choices of $t_0$ i.e. less than
-$\nicefrac{10^5}{\sqrt{t}}$ failing to converge in a reasonable amount
-of time. Thus, the update factor $\gamma$, shown as a function of
-simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves useful in
-determining which methods will converge in a reasonable amount of time.
-We compare each method to a `golden' standard which we define to be an
-ideal choice of SAMC for $10^{11}$ moves followed by TMMC.  This gives an
-unbiased comparison with our developed method SAD. The dynamic version
-of SAMC which we call SAD performs as well as SAMC (assuming the user
-can correctly identify an appropriate choice of $t_0$).  SAD
-outperforms WL and WL-TMMC for the same given number of moves. The
-absence of user defined parameters makes SAD an exciting alternative to
-methods where user experience or successive implementation iteration
-are needed to achieve reasonable results.
-
-\section{Conclusions}
-
-We have introduced a new algorithm SAD (which is Stochastic Approximation
-with the modification factor dynamically chosen) that effectively
-samples the entire energy space for a variety of system sizes.  Also,
-SAD proves effective in comparison with other Monte-Carlo methods at
-converging to the correct density of states.  The ease of
-implementation and the absence of user-defined parameters makes SAD an
-ideal algorithm. SAD converges as quickly as SAMC (in the case of an ideally
-chosen $t_0$ for our simulations. SAD doesn't suffer from the weakness of SAMC,
-i.e. the user choosing parameters that could lead to the simulation not
-converging in a reasonable amount of time. SAD also converges more quickly
-than TMMC (the reader will remember that SAMC followed by TMMC is what each
-method is compared against) and much more quickly than WL-TMMC.
-
-SAD decreases $\gamma$ in a controlled way that prevents the algorithm
-from taking too long to converge (as in the case of SAMC with a poor choice
-of $t_0$) or failing to converge such as is the case with WL for complex
-systems. A major advantage is that the user does not need to determine a
-priori how quickly the update factor $\gamma$ should be decreased as the
-method SAD dynamically determines this.
-
-We note that SAD should be tested on additional systems, and the
-convergence of the method should be examined.  Also, a rigorous
-mathematical justification for our choice of the update factor is
-necessary to ensure the most optimal convergence of the algorithm.
-
-\section{Acknowledgement}
-
-J.K. Pommerenck would like to gratefully acknowledge D.W. Siderius and
-J.R. Errington for helpful communications regarding the implementation
-of WL-TMMC.
-
-% \jpsays{Everything below this is TMI1 and GC-TMMC}
-
-%\subsection{GC-TMMC Algorithm}
-%Grand Canonical Transition Matrix Monte-Carlo (GC-TMMC) allows the energy and
-%particle number of the system to fluctuate while holding the chemical potential,
-%volume, and temperature constant~\cite{chen2001aggregation, chen2002simulating,
-%errington2003direct, fenwick2006accurate, fenwick2008direct, errington2005direct}.
-%The algorithm boasts a lower statistical noise than conventional GCMC due to its
-%incorporation of a transition matrix to estimate ensemble averages~\cite{siderius2013use,
-%paluch2008comparing, grzelak2008computation}.  The steps for the algorithm are
-%outlined below:
-
-%\subsubsection{Define particle insertion/deletion}
-%In the system of consideration, the number of particles $\mathcal{N}$ can change
-%by an amount $\delta$.  It is assumed that in the system either no particles can be
-%inserted or deleted, a single insertion takes place, or a single deletion takes place.
-%This yields a condition $\delta={-1,0,+1}$ for particle insertion.
-
-%\subsubsection{Define an acceptance probability}
-%Now it is necessary to define criteria for determining whether a particle insertion/deletion
-%is allowed.  This criteria is called the acceptance probability and is denoted as
-%\begin{align}
-  %p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
-  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
-%\end{align}
-%where $\alpha_{new\rightarrow old}$ is the probability of generating an old system
-%configuration from the new.  For conventional Monte-Carlo moves $\alpha_{old \rightarrow new}
-%=\alpha_{new\rightarrow old}$ although for advanced Monte-Carlo moves this would not
-%be the case~\cite{siepmann1990method}.  The probabilies of actually observing the
-%system in the old state is given by
-%\begin{align}
-  %\pi_{old} = \frac{1}{\Xi}\frac{V^{N_{old}}}{\Lambda^{3N_{old}}N_{old}!}
-  %e^{-\beta(\epsilon_{old}-\mu N_{old})}
-%\end{align}
-%where $\Xi$ is the grand canonical partition function and $\Lambda$ is the De Broglie
-%wavelength.
-
-%\subsubsection{Define a Collection Matrix}
-%In order to calculate an overall transition probability $\mathcal{P}_{N,\delta}$, we
-%must first define how the system procession is to be recorded.  A Collection Matrix
-%can be defined, such that for any one move, two elements are updated.
-%\begin{align}
-  %{C}_{N,\delta} = {C}_{N,\delta} + p_{a}\\
-  %{C}_{N,0} = {C}_{N,0} +(1 - p_{a})
-%\end{align}
-%If $\delta=0$, that is, the number of particles has not been changed, then only
-%$\mathcal{C}_{N,0}$ is incremented by unity.  The overal transition probability can
-%now be defined as
-%\begin{align}
-  %\mathcal{P}_{N,\delta} = \frac{{C}_{N,\delta}}
-  %{\sum_{j=-1}^{+1} {C}_{N,j}}
-%\end{align}
-
-%\subsubsection{Define the Particle Number Probability Distribution}
-%The Particle Number Probability Distribution (PNPD) $\Pi(N;\mu,V,T)$ can be found
-%by enforcing detailed balance in the system.
-%\begin{align}
-  %\ln(\Pi_{N+1}) = \ln(\Pi_{N}) + \ln\bigg(\frac{\mathcal{P}_{N\rightarrow N+1}}
-  %{\mathcal{P}_{N+1\rightarrow N}}\bigg)
-%\end{align}
-%This equation will hold as long as N-space can be sampled sufficiently.  In an
-%implemented algorithm $N$, $N_{min}$, and $N_{max}$ are specified for macrostate domain
-%sampling.
-
-%\subsubsection{Define a biasing function}
-%The primary difference between TMMC as defined by Fitzgerald et al. and GC-TMMC as
-%pioneered by JR Errington et al. is the use of a biasing function to control the acceptance
-%rate of trial moves~\cite{siderius2013use, fitzgerald2000monte}.  This type of
-%multi-canonical sampling augments the acceptance probability in the following way
-%\begin{align}
-  %p_{\eta,old\rightarrow new} = \min\bigg[1,\frac{e^{\eta(N_{new})}}
-  %{e^{\eta(N_{old})}}\frac{\alpha_{new\rightarrow old}}
-  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
-%\end{align}
-
-%Although $p_{\eta}$ is used to control the actual transitions for trial configurations
-%in GC-TMMC, $p_{a}$ is still used to update the Collection Matrices.  Ideally the biasing
-%function would be given by
-%\begin{align}
-  %\eta(N) = -\ln\Pi(N;\mu,V,T)
-%\end{align}
-%Since $\Pi$ is not usually known a priori, the biasing function is usually set to some
-%arbitrary constant for initialization.  The goal of this biasing function is to achieve
-%uniform sampling (a common feature shared by the flat histogram Monte-Carlo family).
-
-%\subsubsection{Histogram reweighting for the PNPD}
-%Once the PNPD has beeen collected for a given chemical potential $\mu_0$, histogram
-%reweighting is used to determine thermodynamic properties at other chemic potentials.
-%Since the canonical partition function $Z(N,V,T)$ does not depend on the chemical
-%potential, we can write
-%\begin{align}
-%\begin{split}
-  %\Pi(N;\mu,V,T) = \frac{e^{\beta\mu N}Z(N,V,T)}{\Xi(\mu,V,T)}\\
-  %\ln\Pi(N;\mu,V,T) = \ln\Pi(N;\mu_0,V,T) \\
-  %+ \beta N(\mu-\mu_0) + C
-%\end{split}
-%\end{align}
-%where $C$ is a normalization constant independent of the number of particles.
-
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2018-06-30 papers/grand/paper.tex
added some text with mathematical reasoning to 2D broad histogram

@@
 \usepackage{color}
-\usepackage{nicefrac} % Include for inline fractions
+\usepackage{nicefrac} % Include for in-line fractions
+\usepackage{harpoon}% <---
 
@@
 \newcommand{\tssays}[1]{{\color{red} [\cyan{Tanner:} \emph{#1}]}}
+
+\newcommand{\Int}{\int\limits}
 
@@
 
+In component notation, the free energy can be broken down into its constituent components.
+\begin{align}
+\overrightharp{g} = \bigg(\frac{\partial S}{\partial U}\bigg)_N \hat{U} + \bigg(\frac{\partial S}{\partial S}\bigg)_U \hat{N}
+\end{align}
+\begin{align}
+\overrightharp{g} = \frac{1}{T_{\min}} \hat{U} - \frac{\mu_{\max}}{T_{\min}}  \hat{N}
+\end{align}
+The gradient of the free energy can be written as follows (thinking of a $\Delta S$ near the minimum entropy $S_{\min}$:
+\begin{align}
+\Delta S = \left|\overrightharp{g} \cdot \overrightharp{r} \right| = \left|\frac{E_{\min}}{T_{\min}} \right| + \left|\frac{\mu_{\max}N_{\max}}{T_{\min}} \right|
+\end{align}
+\begin{align}
+\langle\Delta S \rangle = \frac{1}{3}\bigg( \left|\frac{E_{\min}}{T_{\min}} \right| + \left|\frac{\mu_{\max}N_{\max}}{T_{\min}} \right| \bigg)
+\end{align}
+We can write the entropy as a parabolic function with free parameters $A$ and $B$ in terms of the energy and particle number of the
+system.
+\begin{align}
+S = -A E^2 - B N^2
+\end{align}
+By taking to partial derivatives of the entropy, we can solve for the free parameters.
+\begin{align}
+\bigg(\frac{\partial S}{\partial E}\bigg)_N = \frac{1}{T_{\min}} = - 2AE
+\end{align}
+\begin{align}
+\bigg(\frac{\partial S}{\partial N}\bigg)_E = - \frac{\mu_{\max}}{T_{\min}} = - 2BN
+\end{align}
+Therefore the parabolic equation can now be written in terms of known parameters.
+\begin{align}
+S = \bigg( \frac{1}{2 T_{\min} E}\bigg) E^2 - \bigg( \frac{\mu_{\max}}{2 T_{\min} N}\bigg) N^2 \\
+S = \bigg( \frac{1}{2 T_{\min}}\bigg) E - \bigg( \frac{\mu_{\max}}{2 T_{\min}}\bigg) N
+\end{align}
+Now we integrate this functional form of entropy in terms of the internal energy $E$ and particle number $N$.
+\begin{align}
+\Int_{0}^{N_{\max}} \Int_{E_{\min}}^{0} \bigg( S - S_{\min}\bigg) \,dE\,dN
+\end{align}
+Now we insert our derived expressions into the integral and obtain a functional form for the update factor $\gamma$.
+\begin{align}
+\Int_{0}^{N_{\max}} \Int_{E_{\min}}^{0} \bigg( \bigg( \frac{1}{2 T_{\min}}\bigg) E - \bigg( \frac{\mu_{\max}}{2 T_{\min}}\bigg) N \\- \frac{1}{3}\bigg( \frac{E_{\min}}{T_{\min}}  + \frac{\mu_{\max}N_{\max}}{T_{\min}} \bigg) \,dE\,dN
+\end{align}
+We then find that the result of the integral is equal to the following:
+\begin{align}
+\gamma =\frac{1}{t} \frac{E_{\min}N_{\max}}{T_{\min}}\bigg(E_{\min} + 7 \mu_{\max}N_{\max} \bigg)
+\end{align}
+
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2018-07-12 papers/histogram/sad.tex
edited results to reflect our analysis of translation scale and minT.

@@
 For the square-well fluid, our simulations treat systems with a
-well-width of $\text{ww} = 1.30$ and a filling fraction of either
-$\text{ff} = 0.10, 0.20,$ or $0.30$.  The simulation explores the
-energy space of the system down to a reduced temperature of
-$T_{\text{min}} = 0.2$.  All simulations lead to the minimum important
-energy $E_{\text{min}}$ and maximum entropy state $S_{\text{max}}$
-being calculated (with the exception of the WL-TMMC method where both
+well-width of $\lambda = 1.30$ and a filling fraction of
+$\eta = 0.30$. We also explore various translation scales $\delta_0 = 0.005,
+0.05,$ and $0.5$. The simulation explores the
+energy space of the system at a reduced temperature of
+$T_{\text{min}} = 0.2$ and $0.5$.  All simulations lead to the minimum important
+energy $E_{\min}$ and maximum entropy energy $E_{\max}$
+being calculated (with the exception of the WL methods where both
 of these parameters are needed a priori).  In the sections below, we
-examine systems with different particles numbers and filling fractions.
+examine systems parameters on the convergence of the methods.
 
@@
 different method take longer to converge. This behavior is seen in
-Fig.~\ref{fig:N50-ff0.1-avg-error}.
+Fig.~\ref{fig:N50-ff0.3-avg-error}.
 
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_10-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown for a filling fraction of 0.10 with
-  SAMC followed by TMMC serving as the `golden' standard.}\label{fig:N50-ff0.1-avg-error}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_20-N50-entropy-error.pdf}
-  \caption{The average entropy error is shown for a filling fraction of 0.20 with TMI acting as the
-  `golden' standard.}\label{fig:N50-ff0.2-avg-error}
-\end{figure}
-\begin{figure}
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
-  \caption{The average entropy error for each method is shown for a filling fraction of 0.30 with
+  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
   SAMC followed by TMMC serving as the `golden'
@@
 \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
-  \caption{The average entropy error is shown for a filling fraction of 0.30 with TMI acting as the
-  `golden' standard.}\label{fig:n500-ff0.2}
+  \caption{The average entropy error is shown for $\lambda = 0.30$ with TMI acting as the
+  `golden' standard.}\label{fig:n500-ff0.3}
 \end{figure}
 
-\subsubsection{A periodic system with 10\% filling fraction and $(N = 50)$}
-In order to determine how the various methods scale with size,
-simulations are run for systems with different particle numbers.
-Figure~\ref{fig:N50-ff0.1-avg-error} shows the average entropy vs moves for
+~\jpsays{We have discussed including a figure for the $\lambda = 30\%$ and $N = 50$
+showing translation scale impact. Also a figure for the $\lambda = 30\%$ and $N = 50$
+showing the impact of choice of minimum temperature.}
+
+\subsection{A periodic system with $\lambda = 30\%$ and $N = 50$}
+
+Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves for
 a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.10.  These particles
+cell with lattice constant chosen to give a $\lambda = 0.30$.  These particles
 are randomly inserted into the system and random moves are proposed (with
@@
 
-\subsubsection{A periodic system with 20\% filling fraction and $N = 50$}
+\subsubsection{translation scale}
+For the $\lambda = 30\%$ and $N = 50$ system, we explore the choice of $\delta_0$
+as a parameter. Smaller translation scales should cause the given method to take
+additional time to explore all energies.  A larger translation scale could result 
+in the simulation taking shorter or longer (because longer distance moves may end up
+not being allowed). Ideal methods should scale linearly with the $\delta_0$, that is,
+a decrease in the translation scale should result in a proportional increase in convergence
+time. Some methods such as WL may not be able to handle arbitrary $\delta_0$.
 
-Figure~\ref{fig:N50-ff0.2-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.20.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
+\subsubsection{minimum temperature}
+For the $\lambda = 30\%$ and $N = 50$ system, we explore the choice of $T_{\min}$
+as a parameter. We have theoretically postulated that a dynamic version of SAMC which
+we call SAD should depend upon this physically meaningful parameter. We thus explore
+the behaviour of SAD for two different reduced temperatures.
 
-\subsubsection{A periodic system with 30\% filling fraction and $N = 50$}
-
-Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of 0.30.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
-
-\subsubsection{A periodic system with 30\% filling fraction and $N = 500$}
+\subsection{A periodic system with $\lambda = 30\%$ and $N = 500$}
 
 A system is configured with dimensions $x = y = z = 19\sigma$.  A
-filling fraction $\text{ff} = 0.30$ is chosen which yields $N = 500$
-particles in a cubical system.  These particles are randomly inserted
+filling fraction $\lambda = 0.30$ is chosen which yields $N = 500$
+particles in a cubical lattice system.  These particles are randomly inserted
 into the system and random moves are proposed (with a translation scale
@@
 this family struggle to converge if the update factor decreases faster
-than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.2} shows both
+than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.3} shows both
 WL and SAMC with inappropriate choices of $t_0$ i.e. less than

### 2018-07-16 papers/histogram/sad.tex
start adding slow/fast output and adjusting scripts to plot it

@@
 \end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/gamma-n50}
+  \caption{The update factor $\gamma$ for $N=50$ versus iteration for three
+    different methods: WL, SAMC, and SAD.  This is exciting because we
+    have different MC moves (slow, etc).}
+\end{figure}
 
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
   \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error.pdf}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
+  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
+  SAMC followed by TMMC serving as the `golden'
+  standard.}\label{fig:N50-ff0.3-avg-error}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-fast.pdf}
+  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
+  SAMC followed by TMMC serving as the `golden'
+  standard.}\label{fig:N50-ff0.3-avg-error}
+\end{figure}
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
   \caption{The average entropy error is shown for $\lambda = 0.30$ with TMI acting as the

### 2018-07-23 papers/histogram/sad.tex
update paper for change below too_low_energy

@@
 \begin{align}
-\gamma_{t}^{\text{SA}} =
+\gamma_{t}^{\text{SAD}} =
 \frac{N_E}{t}\frac{\epsilon_{H}-\epsilon_{L}}{3T_{\text{min}}}
@@
 
-\subsubsection{High and Low energy racheting}
-Since SAD explores all energy states it is imperative that the method
-have a reasonable idea about where the minimum important energy and
-maximum entropy state are located in energy space. Since these are not
-user input parameters such as they are for methods like WL-TMMC, the
-simulation is responsible for determining and updating these energy
-locations. In order to prevent, the method from experiencing errors
-near the energy extrema, we employ a high energy $\epsilon_H$ and a low
-energy $\epsilon_L$ to define the range over which the energy histogram
-is made flat. $\epsilon_H$ is the highest ever value that the maximum
-entropy state has taken while $\epsilon_L$ is the lowest ever value
-that the minimum important energy has taken. SAD uses these two
-energies to control the weights $w(E)$ above $\epsilon_H$ and
-$\epsilon_L$. If the energy $E$ is greater than $\epsilon_H$ then the
-weights can be calculated as ~\jpsays{equation formatting problems!}
-\begin{equation}
-\begin{aligned}
-\ln w(E) = \max(\ln w(E) + \gamma, \\
-    \ln w(\epsilon_H) + \log(\gamma + e^{\ln w(E)-\ln w(\epsilon_H)}))
-\end{aligned}
-\end{equation}
-If the energy $E$ is less than $\epsilon_L$ then the weights can be
-calculated as
-\begin{equation}
-\begin{aligned}
-\ln w(E) = \max(\ln w(E) + \gamma, \ln w(\epsilon_L) \\
-    + \frac{\epsilon_L - E}{T_\text{min}}
-    + \log(\gamma + e^{\ln w(E) -\ln w(\epsilon_L)
-    + \frac{\epsilon_L -E}{T_\text{min}}}))
-\end{aligned}
-\end{equation}
-Sad sets the maximum entropy state equal to the energy if either of the
-following is true.
+\subsubsection{The ``interesting'' energy range}
+Since SAD explores all energy states it needs to have a reasonable idea
+of what energy range corresponds to the temperature range of interest,
+defined by $T_{\min}<T<\infty$.
+The
+simulation is responsible for determining and updating this energy
+range.
+Given the true entropy $S(E)$, we can define the interesting energy
+range as
 \begin{align}
-\ln w(E) > \ln w(S_\text{max}) \; \text{or} \; H(S_\text{max}) = 0
+  E(T_{\min}) <E< E(T=\infty)
 \end{align}
-Thus if the method is beyond the maximum entropy state or has collected no
-statistics at the maximum entropy state then it is set equal to the energy.
-Similarly, Sad sets the minimum important energy equal to the energy if
-either of the following is true.
-~\jpsays{equation formatting problems!}
+where $E(T)$ is the energy that maximizes $S-E/T$.  In the course of the
+simulation this precise energy may be challenging to evaluate accurately.
+In order to ensure that we sample this entire energy range adequately,
+we define two energy limits:  a high energy $E_H$ and a low
+energy $E_L$ to define the range over which the energy histogram
+is made flat. $E_H$ is chosen to be the greatest energy that
+maximized the entropy at any time during the simulation.  Similarly,
+$E_L$ is the lowest energy that ever maximized $S(E)-E/T_{\min}$ during
+the simulation.  This creates a ``ratcheting'' effect, in which $E_H$
+may only increase, while $E_L$ may only decrease over the course of the
+simulation, which results in a conservative estimate of the range of
+energies that need be sampled.
+
+Outside of the energy range $E_L \le E \le E_H$, the weights are ignored
+when computing acceptance probabilities, and the weights $w(E_L)$ or
+$w(E_H)$ are used instead.  This means that the histogram is \emph{not}
+flat in this region, but instead is proportional to the density of
+states.  In order to ensure that $w(E) \propto D(E)$, the change in
+weight with each visit should be equal to the change in weight at
+the border energy $E=E_{H/L}$.
 \begin{align}
-\ln w(E) > \ln w(E_\text{min}) - \frac{E_\text{min}-E}{T_\text{min}}
-\; \text{or} \; H(E_\text{min}) = 0
+  w_{t+1}(E)-w_t(E) &= w_{t+1}(E_{H/L}) -w_t(E_{H/L}) \\
+    &= e^\gamma w(E_{H/L}) - w(E_{H/L}) \\
+    &= (e^\gamma-1) w(E_{H/L})
 \end{align}
-Thus if the method is beyond the minimum important energy or has collected no
-statistics at the minimum important energy then it is set equal to the energy.
-Together these two energies serve as a kind of ratcheting effect to
-determine the energy landscape boundaries. Outside the range set by the
-weights $w(E)$, the maximum entropy state is set to the energy $E$. The
-minimum important energy is also set to the energy when outside the
-calculated range.
-
-\paragraph{Accepting moves above and below}
-We do \emph{not} use the weights to determine move acceptance outside
-the middle range.  FIXME WRITE THIS SECTION.
-
-\paragraph{Above $\epsilon_H$}
-Each energy $E>\epsilon_H$, should be visited at a rate that
-is proportional to the density of states at that energy, as in a
-simulation with constant weights.  Thus the weight at each energy
-needs to be proportional to the number of times that weight has been
-visited.  In order to ensure that $w(E) \propto D(E)$, the change in
-weight with each visit should be equal to the change in weight at
-the border energy $E=\epsilon_H$.
-\begin{align}
-  \Delta w(E) &= \Delta w(\epsilon_H) \\
-  w_{\text{new}}(E)-w_{\text{old}}(E)
-    &= e^\gamma w(\epsilon_H) - w(\epsilon_H) \\
-    &= (e^\gamma-1) w(\epsilon_H)
-\end{align}
-This ensures that although we visit each energy above $\epsilon_H$ only
+This ensures that although we visit each energy above $E_H$ or
+below $E_L$ only
 in proportion to its density of states, the weights at these energies
 will approach the entropy as the simulation progresses.
-
-\paragraph{Below $\epsilon_L$}
-For energies $E<\epsilon_L$, we will visit each state in proportion to
-the Boltzmann distribution at temperature $T_{\min}$.  This ensures that
-we have adequate statistics to describe the system at temperature
-$T_{\min}$ while not wasting time simulating extremely low energies with
-correspondingly low entropy.  This means that we need to shift the
-weights at lower energies by a smaller amount in proportion to the
-Boltzmann factor.
-\begin{align}
-  \Delta w(E) &= \Delta w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}} \\
-  w_{\text{new}}(E)-w_{\text{old}}(E)
-    %&= \left(e^\gamma w(\epsilon_L) - w(\epsilon_L)\right)e^{(E-\epsilon_L)/T_{\min}} \\
-    &= (e^\gamma-1) w(\epsilon_L)e^{(E-\epsilon_L)/T_{\min}}
-\end{align}
-We note that when updating the natural logarithms of the weights, care
-is required in order to ensure that overflow does not occur.
 

### 2018-07-23 papers/histogram/sad.tex
add red text describing our new idea and why it is so

@@
 
-\subsubsection{The ``interesting'' energy range}
+{\color{red}
+While for SAMC the update factor is defined in the original
+implementation, for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
+thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
+$t_0$ should have dimensions of entropy.
+We begin with an estimate of the average value of the entropy (relative
+to the lowest entropy at $T_{\min}$).  If we assume a quadratic
+dependence on energy, this is given by
+\begin{align}
+\langle S\rangle \approx \frac13 \frac{E({T=\infty}) - E(T_{\min})}{T_{\min}}
+\end{align}
+We approximate this energy difference by the energy difference
+$E_H -E_L$ where these energies are defined below in the following
+section.  Early in the simulation, when the energy range is still
+expanding, this average entropy is a good choice
+for $t_0$.  However, as the simulation progresses, our entropy factor
+needs to scale with the total number of energy states that have been
+found $N_S$, since the product $N_S\langle S\rangle$ corresponds to the
+total change of entropy required, given that each of the $N_S$ states
+will has a separate weight $w(E)$.  Finally, at \emph{long} times, after
+all the energies have been found and the weights are being refined
+we wish for a lower $t_0$ in order to reduce the fluctuations in the
+weights that result from large values for $\gamma_t$.  We enable this
+by tracking the last time we encountered a new energy $t_L$, and
+gradually transition to lower $t_0$ (but still asymptotically scaling
+as $\gamma_t \propto 1/t$ to ensure eventual convergence).  The SAD
+expression for $\gamma_t$ which incorporates these ideas is:
+\begin{align}
+  \gamma_{t}^{\text{SAD}} =
+     \frac{E_{H}-E_{L}}{3T_{\text{min}}t}
+    \frac{
+      E_S^2 + E_S t + \left(t/t_L-1\right)t
+    }{
+      E_S^2 + t + \left(t/t_L-1\right)t
+    }
+\end{align}
+This factor asymptotically has the same $1/t$ behavior as the original
+SAMC algorithm;
+however now every time a new energy is discovered during the course of
+the simulation, the update factor experiences a jump. This behavior
+allows SAD to dynamically prevent the update factor from decreasing too
+rapidly. Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for SAD, WL,
+and SAMC.  The update factor
+for WL remains high for many iterations, and then decreases very rapidly,
+while the SAMC $\gamma_t$ remains constant before dropping as $1/t$.
+The update factor for SAD in contrast fluctuates
+dynamically around a value less than 1 for early MC moves
+(while finding new energy states), before dropping in a \davidsays{similar
+to SAD, but described after we see the plot}.
+}
+
+\subsubsection{The ``interesting'' energy range}\label{sec:EHL}
 Since SAD explores all energy states it needs to have a reasonable idea
@@
 range as
-\begin{align}
-  E(T_{\min}) <E< E(T=\infty)
-\end{align}
+%\begin{align}
+  $E(T_{\min}) <E< E(T=\infty)$
+%\end{align}
 where $E(T)$ is the energy that maximizes $S-E/T$.  In the course of the

### 2018-07-23 papers/histogram/sad.tex
clean up equation and use \gamma_t more consistently

@@
   w_{t+1}(E)-w_t(E) &= w_{t+1}(E_{H/L}) -w_t(E_{H/L}) \\
-    &= e^\gamma w(E_{H/L}) - w(E_{H/L}) \\
-    &= (e^\gamma-1) w(E_{H/L})
+    &= (e^{\gamma_t}-1) w_t(E_{H/L})
 \end{align}
@@
   \includegraphics[width=\columnwidth]{figs/gamma-n500}
-  \caption{The update factor $\gamma$ for $N=500$ versus iteration for three
+  \caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
     different methods: WL, SAMC, and SAD.
@@
   \includegraphics[width=\columnwidth]{figs/gamma-n50}
-  \caption{The update factor $\gamma$ for $N=50$ versus iteration for three
+  \caption{The update factor $\gamma_t$ for $N=50$ versus iteration for three
     different methods: WL, SAMC, and SAD.  This is exciting because we

### 2018-07-23 papers/ising/ising-sad.tex
move ising to new paper directory

--- 2018-06-29 papers/grand/paper.tex
@@
 
+\newcommand{\Int}{\int}
+\newcommand\Emin{E_{\min}}
+\newcommand\Nmax{N_{\max}}
+\newcommand\Tmin{T_{\min}}
+\newcommand\mumax{\mu_{\max}}
+
 \begin{document}
-\title{Two-dimensional Broad Histogram Monte Carlo}
+\title{SAD Monte Carlo}
 
 \author{Jordan K. Pommerenck}
-\author{Cade MIDDLE INITIAL? Trotter}
 \author{David Roundy}
@@
 We present several histogram methods and compare the performance and
-  efficiency at treating the square-well fluid.
+efficiency at treating the Ising model.
 \end{abstract}

### 2018-07-23 papers/histogram/sad.tex
broad -> flat histogram and other edits.

@@
 \section{Introduction}
-Over the past several decades a number of broad histogram Monte-Carlo
+Over the past several decades a number of flat histogram Monte-Carlo
 simulation algorithms have been developed which calculate the
@@
 Stochastic Approximation), and will discuss it in detail in the methods
-section. We compare it along with several broad histogram methods
+section. We compare it along with several flat histogram methods
 which include TMMC, WL, WL-TMMC, and SAMC.
@@
 The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
-describe in detail broad-histogram methods.  Section~\ref{sec:weight} outlines
+describe in detail flat-histogram methods.  Section~\ref{sec:weight} outlines
 weight-based histogram methods used in this work. Section~\ref{sec:transition}
@@
 
-In this work, We compare a variety of broad histogram
+In this work, We compare a variety of flat histogram
 methods.  We outline the general workings of each algorithm that we
@@
 
-\section{Broad histogram methods}\label{sec:histogram}
+\section{Flat histogram methods}\label{sec:histogram}
 
-The goal of broad histogram methods (also called \emph{flat histogram}
+The goal of flat histogram methods (also called \emph{broad histogram}
 or \emph{multicanonical} methods) is to simulate each energy with
@@
 
-Each broad histogram Monte Carlo method uses a given set of ``moves''
+Each flat histogram Monte Carlo method uses a given set of ``moves''
 which change a system and should satisfy detailed balance.  Each
@@
 \begin{equation}
-	\mathcal{P}(\epsilon_\text{old} \rightarrow \epsilon_\text{new})
-	= \min\left[1,\frac{w(\epsilon_\text{old})}{w(\epsilon_\text{new})}\right]
+	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
+	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
 \end{equation}
@@
 A set of weights that are proportional to the density of states of the
-system $D(\epsilon)$ will result in an entirely flat histogram.  Thus
+system $D(E)$ will result in an entirely flat histogram.  Thus
 flat histogram is a criteria for convergence for these methods.  To avoid 
@@
 few hundred orders of magnitude, the natural logarithm of the weights are used
-in this work.  Since $D(\epsilon) = S(\epsilon)$, the logarithm of the weights 
+in this work.  Since $D(E) = S(E)$, the logarithm of the weights
 can be thought of as an approximation of the entropy.
@@
 \begin{equation}
-	\ln{w_{t+1}(\epsilon)}=\ln{w_{t}(\epsilon)}
+	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
 	+\gamma
@@
 original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$ and in future works~\cite{werlich2015stochastic} with a scale factor
-$\gamma_0$.
+arbitrary tunable parameter $t_0$.
 \begin{align}
-\gamma_{t}^{\text{SA}} =\gamma_0 \frac{t_0}{\max(t_0,t)}
+\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
 \end{align}
@@
 known a priori.  The time to converge depends only on the choice of
-parameters $t_0$ and $\gamma_0$.  Unfortunately, there is no
-prescription for finding acceptable values for these parameters, and
+parameters $t_0$.  Unfortunately, there is no
+prescription for finding an acceptable value for $t_0$, and
 while the algorithm formally converges, for a poor choice of $t_0$
-that convergence may be far too slow to be practical.  This work
-proposes a means for dynamically setting the update factor $\gamma_t$
-that should robustly converge on a variety of systems.
+that convergence can be far too slow to be practical.
+
+Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t$ by a factor
+$\gamma_0$.  While this may result in an improved rate of convergence,
+it adds yet another parameter that must be empirically determined.
+
+%This work proposes a means for dynamically setting the update factor
+%$\gamma_t$ that should robustly converge on a variety of systems.
 
@@
 parameters are chosen independent of system size or type. Each
-Broad-Histogram method has unique advantages and disadvantages.
+Flat-Histogram method has unique advantages and disadvantages.
 Wang-Landau requires an energy range for initialization.  SAMC removes
@@
 \gamma_{t}^{\text{SAD}} =
-\frac{N_E}{t}\frac{\epsilon_{H}-\epsilon_{L}}{3T_{\text{min}}}
+\frac{N_E}{t}\frac{E_{H}-E_{L}}{3T_{\text{min}}}
 \end{align}
@@
 systems. Coarse binning yields fewer energy states for the walker to
-explore but more time spent in the individual bin. Fine binning yields
-more energy states to explore but less time spent in each bin.  Some
+explore but more time spent in each individual bin. Fine binning yields
+more energy states to explore with less time spent in each bin.
+\davidsays{move this discussion earlier.} Some
 methods such as AdaWL~\cite{koh2013dynamically} (discussed elsewhere)
@@
 states in order to ensure the exploration of all energies. A
-significant advantage of SAD over other broad histogram methods is that
+significant advantage of SAD over other flat histogram methods is that
 energy binning is solved dynamically. SAD should perform independently
@@
 scaling free for a reasonable choice of step size.  SAD should perform
-extremely well compared with other broad histogram methods when applied
+extremely well compared with other flat histogram methods when applied
 to complex continuous systems such as polymers/proteins and spin models.
@@
 
-J.K. Pommerenck would like to gratefully acknowledge D.W. Siderius and
+We gratefully acknowledge D.W. Siderius and
 J.R. Errington for helpful communications regarding the implementation

### 2018-07-25 papers/histogram/sad.tex
added some more recent text on TMMC section

@@
 \subsection{TMMC Algorithm}
-The Transition Matrix Monte-Carlo (TMMC) method as introduced by
+{\color{red}
+The flat-histogram Wang  and  Landau method utilizes a so-called
+``visited states'' approach to calculate  thermodynamic
+properties~\cite{errington2003direct}. The algorithm is fundamentally
+based on the number of times a given state is visited while the
+system is being sampled. The Transition Matrix Monte Carlo (TMMC) method operates
+differently than the visited states approach~\cite{fitzgerald2000monte}. The method
+uses attempted transitions between states to obtain an estimate of transition probabilities.
+
+Say something about detailed balance here?...
+
+For problems that sample states where the relative probabilities
+vary by many orders of magnitude, all of the states will
+not be sampled efficiently within reasonable simulation
+time without introducing a bias in the sampling distribution.
+
+While the current algorithm is highly efficient, a weighting function $\eta_t$
+is introduced to ensure that the system samples all states
+evenly~\cite{mahynski2017multivariable, blanco2016effect}.
+As for the Flat-histogram approach, the weighting function is
+set inversely proportional to the current macrostate probability.
+\begin{align}
+  \eta(I) = - \ln(D(I))
+\end{align}
+The probability is defined in terms of this biasing function:
+\begin{align}
+  p(I\rightarrow J) = \min\bigg[1,\frac{\eta(J)}
+  {\eta(I)} e^{\beta (E(I)-E(J))}\bigg]
+\end{align}
+The TMMC algorithm consists of recording the data in a collection matrix
+(the unormalized transition histogram matrix).
+\begin{align}
+  C(I \rightarrow J) = C(I \rightarrow J) + p(I \rightarrow J)\\
+  C(I \rightarrow I) = C(I \rightarrow I) +(1 - p(I \rightarrow J))
+\end{align}
+If the energy state has not been changed, then $C$ is still incremented by
+unity. Fitzgerald et al.~\cite{fitzgerald2000monte} showed that defining
+the third step this way results in a uniform improvement over using
+histogram estimators to determine the particle number probability
+distribution (PNPD). The algorithm is highly efficient due to the ability
+to bias the simulation while keeping the criteria used to update the
+collection matrix $C(I\rightarrow J)$.
+}
+
+The Transition Matrix Monte-Carlo (TMMC) method as outlined by
 Fitzgerald et al. consists of three primary steps.  The first two steps

### 2018-07-27 papers/histogram/sad.tex
added section on binning and made a plotting fix

@@
 \begin{document}
-\title{Dynamic Stochastic Approximation Monte Carlo}
+\title{Dynamic Stochastic Approximation Monte Carlo
+\jpsays{DSA for stochastic optimization (not Monte Carlo) is actually a 
+thing (1965, 2017) although unrealated to what we are doing}
+}
 
@@
 
-\davidsays{We should discuss continuous energies and binning somewhere
-  here...}
-\jpsays{I put some text about this but it was under a sad subsection also
-may need reworked and/or edited.}
+{\color{red} Both weight-based and transition matrix methods have
+calculate the density of states $D(E)$ for discrete energy
+levels~\cite{haber2018performance}. Energy binning becomes an important
+consideration for complex continous systems for this reason. Energy
+bins are typically of uniform size for the entire energy
+continuum~\cite{fasnacht2004adaptive}. Coarse binning yields fewer
+energy states to explore but more time spent in each individual bin.
+Fine binning yields more energy states to explore with less time spent
+in each bin. Some methods such as AdaWL~\cite{koh2013dynamically}
+employ a tunable mechanism for controlling the binning for low entropic
+states in order to ensure the exploration of all energies.
+}
 
@@
 
+Say something about collecting statistics vs WL?
 Say something about detailed balance here?...
@@
 
-While the current algorithm is highly efficient, a weighting function $\eta_t$
+While the current algorithm is highly efficient, a weighting function $\eta(I)$
 is introduced to ensure that the system samples all states

### 2018-08-03 papers/histogram/sad.tex
reworked wltmmc section in the paper.

--- 2018-07-23 papers/histogram/sad.tex
@@
 \begin{document}
-\title{Dynamic Stochastic Approximation Monte Carlo}
+\title{Stochastic Approximation Monte Carlo with Dynamic Update
+Factor (SAD)
+}
 
@@
 
-\davidsays{We should discuss continuous energies and binning somewhere
-  here...}
-\jpsays{I put some text about this but it was under a sad subsection also
-may need reworked and/or edited.}
+{\color{red} Both weight-based and transition matrix methods calculate
+the density of states $D(E)$ for discrete energy
+levels~\cite{haber2018performance}. For this reason, energy binning
+becomes an important consideration for complex continous systems.
+Energy bins are typically of uniform size for the entire energy
+continuum~\cite{fasnacht2004adaptive}. Coarse binning yields fewer
+energy states to explore but more time spent in each individual bin.
+Fine binning yields more energy states to explore with less time spent
+in each bin. Some methods such as AdaWL~\cite{koh2013dynamically}
+employ a tunable mechanism for controlling the binning for low entropic
+states in order to ensure the exploration of all energies. }
 
@@
 \subsection{TMMC Algorithm}
-The Transition Matrix Monte-Carlo (TMMC) method as introduced by
-Fitzgerald et al. consists of three primary steps.  The first two steps
-are common to the Metropolis algorithm while the third makes use of the
-actual transition probabilities~\cite{fitzgerald2000monte}.
+The previously described weight-based flat-histogram methods utilizes a so-called
+``visited states'' approach to calculate  thermodynamic
+properties~\cite{errington2003direct}. The algorithm is fundamentally
+based on the number of times a given state is visited while the
+system is being sampled. The Transition Matrix Monte Carlo (TMMC) method operates
+differently than the visited states approach~\cite{fitzgerald2000monte}.
+The method uses attempted transitions between states to obtain an estimate of
+the probability that a given transition will be proposed.  The density
+of states is computed from the resulting transition matrix.  An advantage
+of collecting the transition matrix is that this matrix may be collected
+while another method is employed to compute acceptance probabilities, as
+in Wang-Landau TMMC, discussed below.
 
-\subsubsection{define a transition probability}
-Assuming that the system under study is in a state denoted old, and a
-proposed transition is made to a new state denoted new, the probability
-is defined as
+The TMMC algorithm records the transition data in a \emph{collection matrix},
+i.e. the unnormalized transition histogram matrix, which stores the number
+of times a transition between two energy states has been proposed.
 \begin{align}
-  p_{old \rightarrow new} = p_{new \rightarrow old}
+  C(E_{\text{old}} \rightarrow E_{\text{new}},t+1)
+     = C(E_{\text{old}} \rightarrow E_{\text{new}},t) + 1
 \end{align}
-where we have allowed the probabilies to be equal for simplicity of
-presentation.
-
-\subsubsection{define an acceptance probability}
-The probability for accepting a transition from an old state to a new
-one can be defined
+The transition matrix is computed from the collection matrix via
 \begin{align}
-  p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
-  {\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
+  T(E_{\text{old}} \rightarrow E_{\text{new}})
+   = \frac{C(E_{\text{old}} \rightarrow E_{\text{new}})}{
+     \sum_E C(E_{\text{old}} \rightarrow E)
+  }
 \end{align}
-where $\alpha_{new\rightarrow old}$ is the probability of generating a
-new system configuration the old.  For conventional Monte-Carlo moves
-$\alpha_{old \rightarrow new} =\alpha_{new\rightarrow old}$ although
-for advanced Monte-Carlo moves this would not be the
-case~\cite{paluch2008comparing, siepmann1990method}.  The probabily of
-observing the system in the old state can be found based on the
-ensemble constraints placed on the system.  Fitzgerald et al.
-considered the system of interest using the canonical ensemble;
-however, the extension to other ensembles is quite natural.
+The acceptance probability is defined in terms of this biasing function:
 \begin{align}
-  \pi_{old} = \frac{e^{-\beta E_{old}}}{Z}
+  \mathcal{P}(E_{\text{old}} \rightarrow E_{\text{new}}) =
+  \max\left(1,\frac{T(E_{\text{old}} \rightarrow E_{\text{new}})}{
+       T(E_{\text{new}} \rightarrow E_{\text{old}})}\right)
 \end{align}
-\subsubsection{define a bookkeeping step}
-The third step in the TMMC algorithm consists of recording the data in
-a collection matrix (the unormalized transition histogram matrix).
-\begin{align}
-  C_{N,\delta} = C_{N,\delta} + p_{a}\\
-  C_{N,0} = C_{N,0} +(1 - p_{a})
-\end{align} If $\delta=0$, that is, the energy state has not been
-changed, then only $C_{N,0}$ is incremented by unity. Fitzgerald et
-al.~\cite{fitzgerald2000monte} showed that defining the third step this
-way results in a uniform improvement over using histogram estimators to
-determine the particle number probability distribution (PNPD).
+where we note that the transition matrix is updated to reflect the
+transition under consideration \emph{prior} to computing the acceptance
+probability.  The TMMC method suffers from poor convergence primarily
+because it is sensitive to poor statistics in less probable energy
+transitions.  Like the weight-based methods above, the TMMC method violates
+detailed balance, with the level of violation decreasing as the simulation
+progresses and a difference of 1 in the collection matrix has a smaller
+impact on the acceptance probabilities.
 
 \subsection{WL-TMMC Algorithm}
-
 Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
 by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
-advantage of the ease of implementation and applicability to a variety
-of systems, while making use of a transition matrix to address WL
-weaknesses, such as its limited statistical accuracy plateau.
-This WL version is more accurate for a given number of simulation steps.
+advantage of the WL methods ease of implementation, while making use of
+a transition matrix to address WL weaknesses, such as its limited
+statistical accuracy convergence.
 
-\paragraph{Initializing Wang-Landau:} We begin the simulation by
-defining a minimum important energy and a maximum entropy state. In
-traditional Wang-Landau the WL factor begins at 1.0 and the factor is
-modified by 2.0 until the simulation is cutoff typically around
-$10^{-10}$.  Rane et. al. published an implemenation of
-WL-TMMC~\cite{rane2013monte} that suggested using a WL factor between
-1.0 and $10^{-2}$ with a cutoff $<10^{-5}$. In our implementation of
-WL-TMMC, we set the WL factor to be 1 and the cutoff to be either
-$10^{-4}$ or $10^{-10}$. A single sweep is defined as each macrostate
-being sampled a number of times in this case 10,000 (although 100 is
-often sufficient for small systems). Wang-Landau sets a flatness
-criteria~\cite{wang2001determining, wang2001efficient,
+\paragraph{Initializing Wang-Landau:}
+We set the energy range along which the method may sample as
+$E(T_{\min}) <E< E(T=\infty)$. Conventional Wang-Landau prescribes
+updating the WL factor $\gamma$ by starting at 1.0 and reducing by
+$\frac12$ until the simulation is cutoff typically around $10^{-10}$.
+While variants of this approach appear in the literature, Rane et. al.
+published an implemenation of WL-TMMC~\cite{rane2013monte} that
+suggested using a WL factor between 1.0 and $10^{-2}$ with a cutoff
+$<10^{-5}$, We implement WL-TMMC by setting the WL factor to be 1.0 and
+the cutoff to be either $10^{-4}$ or $10^{-10}$.
+
+A single sweep ``roundtrip'' is defined as each macrostate being
+sampled a number of times in this case 100,000 times. Wang-Landau sets
+a flatness criteria~\cite{wang2001determining, wang2001efficient,
 hatch2015computational, mahynski2017predicting} for the accumulated
-histogram usually between 0.75 to 0.90.  In the original
-implementation, it was found to be sufficient that each macrostate is
-visited at least once (each histogram bin has one
-entry)~\cite{shell2003improved}; however, for other systems this is
-rarely sufficient.  The goal of this WL initialization is to prefill
-the transition matrix for the next portion of the simulation.
-\paragraph{Initializing TMMC:} Prefilling the transition
-matrix is necessary because if numerous zeros existed in the collection
-matrix, the infinite temperature transition matrix would be
-ill-defined.  Sampling over large ranges of density of states would
-therefore take an unacceptable length of time to
-complete~\cite{shell2003improved, shen2014elucidating}.  The transition
-matrix in terms of the collection matrix is given as
-follows:
-\begin{align}
-\widetilde{T}_{\infty}(I\rightarrow J) = \frac{C(I,J)}
-{\sum_{K} C(I,K)}
-\end{align}
-We now run the TMMC portion of the simulation.
-%Siderius does this for 20 additional sweeps (does about 8 sweeps for
-%WL).  We do this until we reach a defined number of iterations. We
-%have not yet implemented a biasing function to control acceptance rate
-%of trial moves but we could? at some point.
-The acceptance probability is written in terms of the infinite temperature
-transition matrix.
-\begin{align}
-  P_{acc}(i\rightarrow j) = \min\bigg[1,\frac{\widetilde{T}_{\infty}(J\rightarrow I)}
-  {\widetilde{T}_{\infty}(I\rightarrow J)}\bigg]
-\end{align}
-This probability is only used to control actual transitions not update
-the collection matrices.
+histogram usually to be 0.8.  In the original implementation, it was
+found to be sufficient that each macrostate is visited at least once
+(each histogram bin has one entry)~\cite{shell2003improved}; however,
+for other systems this is rarely sufficient~\cite{bhar2009computer}.
+The goal of this WL initialization is to prefill the transition matrix
+for the next portion of the simulation.
+
+\paragraph{Initializing TMMC:} Prefilling the transition matrix is
+necessary because if numerous zeros existed in the collection matrix,
+the infinite temperature transition matrix would be ill-defined.
+Sampling over large ranges of density of states would therefore take an
+unacceptable length of time to complete~\cite{shell2003improved,
+shen2014elucidating}.  The TMMC portion of the algorithm is run only
+after WL has reached the specified cutoff. An important note is that
+the acceptatnce probability for TMMC is only used to control actual
+transitions not update the collection matrices.
 
 \section{SAD Algorithm}\label{sec:sad}
-\subsubsection{The theory behind SAD}
-The SAD (Dynamic Stochastic Approximation) method is a version of the SAMC
+The Stochastic Approximation with Dynamic update factor (SAD) method
+is a variant of the SAMC
 Algorithm that attempts to dynamically choose the modification factor
@@
 
-\subsubsection{The ``interesting'' energy range}\label{sec:EHL}
-Since SAD explores all energy states it needs to have a reasonable idea
+Since SAD does not explore \emph{all} energy states it needs to have a
+reasonable idea
 of what energy range corresponds to the temperature range of interest,
@@
 
-\subsubsection{Energy binning with SAD}
+\paragraph{Energy binning with SAD}
 
@@
 
-\subsubsection{Scaling, step size, and SAD}
+\paragraph{Scaling, step size, and SAD}
 Methods such as WL can fail to converge for complex systems in part

### 2018-08-06 papers/histogram/sad.tex
update paper with new sad

@@
 for $t_0$.  However, as the simulation progresses, our entropy factor
-needs to scale with the total number of energy states that have been
+needs to scale with the total number of \davidsays{interesting} energy states that have been
 found $N_S$, since the product $N_S\langle S\rangle$ corresponds to the
@@
   \gamma_{t}^{\text{SAD}} =
-     \frac{E_{H}-E_{L}}{3T_{\text{min}}t}
-    \frac{
-      E_S^2 + E_S t + \left(t/t_L-1\right)t
+     N_S \frac{
+       \frac{E_{H}-E_{L}}{3T_{\text{min}}} + \frac{t}{t_L}
      }{
-      E_S^2 + t + \left(t/t_L-1\right)t
+       N_S^2 + \frac{t^2}{t_L}
      }
@@
 SAMC algorithm;
-however now every time a new energy is discovered during the course of
+however now every time a new energy is found in the
+\davidsays{interesting energy range} during the course of
 the simulation, the update factor experiences a jump. This behavior

### 2018-08-08 papers/histogram/sad.tex
added some text on SAD

@@
 that convergence can be far too slow to be practical.
+
+{\color{red}
+It is for this reason that we propose a way to determine the update factor
+``dynamically'' while still satisfying the criteria for convergence.
+For the update factor to scale well across both discrete and continuous systems,
+it must include the number of states $N_S$. We define the average entropy as
+$\bar S$ and the time since the simulation last found a new energy state $t_L$.
+We then can write the following expression for the \emph{proposed} update factor.
+\begin{align}
+\gamma = N_S \frac{\bar S + \nicefrac{t}{t_L}}{N_{S}^2 + \nicefrac{t^2}{t_L}}
+\end{align}
+We now consider the following limiting cases:
+\subsubsection{$t \propto N_S \approx t_L$ and $ t << N_S$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{\bar S}{N_S} \approx \frac{\bar S}{t} 
+\end{align}
+Because $t \propto N_S$, the update factor scales as 1/t which is consistent
+with the requirement for SAMC convergence.
+\subsubsection{$t >> N_{S}^2$ and $ t\approx t_L$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{N_S \bar S t_L}{t^2} \approx \frac{N_S \bar S}{t} 
+\end{align}
+\subsubsection{$t >> t_L \bar S$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{N_S}{t}
+\end{align}
+We now integrate (sum) the update factor from the last time an energy state was found
+to the final simulation time to ensure that convergence is guaranteed.
+\begin{align}
+\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[ \int_{t_L}^{t_f} 
+\frac{\bar S -1}{t^2} dt + \frac{1}{t_L}\int_{t_L}^{t_f} \frac{1}{t} dt \right]
+\end{align}
+From this we obtain
+\begin{align}
+\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[\left(\bar S -1\right)\left(\frac{1}{t_L}-\frac{1}{t_f}\right) + \frac{1}{t_L} \ln\left(\frac{t_f}{t_L}\right) \right]
+\end{align}
+Simplifying, we see that as $t_f \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
+the requirement for SAMC convergence.
+\begin{align}
+\gamma \approx \left(1-\frac{t_L}{t_f}\right)N_S\left(\bar S -1\right) + N_S \ln\left(\frac{t_f}{t_L}\right)
+\end{align}
+}
 

### 2018-08-13 papers/histogram/sad.tex
describe sad from last friday.

@@
 
-Outside of the energy range $E_L \le E \le E_H$, the weights are ignored
-when computing acceptance probabilities, and the weights $w(E_L)$ or
-$w(E_H)$ are used instead.  This means that the histogram is \emph{not}
-flat in this region, but instead is proportional to the density of
-states.  In order to ensure that $w(E) \propto D(E)$, the change in
-weight with each visit should be equal to the change in weight at
-the border energy $E=E_{H/L}$.
+Outside of the energy range $E_L \le E \le E_H$, the weights are
+determined from the accumulated energy histogram. For each energy $E_i
+> E_H$, the energy weights are calculated
 \begin{align}
-  w_{t+1}(E)-w_t(E) &= w_{t+1}(E_{H/L}) -w_t(E_{H/L}) \\
-    &= (e^{\gamma_t}-1) w_t(E_{H/L})
+w(E_i)_{t+1} = w(E_H)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right)
 \end{align}
-This ensures that although we visit each energy above $E_H$ or
-below $E_L$ only
-in proportion to its density of states, the weights at these energies
-will approach the entropy as the simulation progresses.
+assuming of course that the simulation has collected statistics at that energy.
+For each energy $E_i
+< E_L$, the energy weights are calculated
+\begin{align}
+w(E_i)_{t+1} = w(E_L)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right) +
+\frac{E_i - E_L}{T_{\min}}
+\end{align}
 
@@
 
-Energy binning is an important consideration for complex continous
-systems. Coarse binning yields fewer energy states for the walker to
-explore but more time spent in each individual bin. Fine binning yields
-more energy states to explore with less time spent in each bin.
-\davidsays{move this discussion earlier.} Some
-methods such as AdaWL~\cite{koh2013dynamically} (discussed elsewhere)
-employ a tunable mechanism for controlling the binning for low entropic
-states in order to ensure the exploration of all energies. A
-significant advantage of SAD over other flat histogram methods is that
-energy binning is solved dynamically. SAD should perform independently
-for a reasonable range of binning because $\gamma \propto
+A significant advantage of SAD over other flat histogram methods is
+that energy binning is solved dynamically. SAD should perform
+independently for a reasonable range of binning because $\gamma \propto
 \nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$

### 2018-08-18 papers/histogram/sad.tex

@@
 While for SAMC the update factor is defined in the original
-implementation, for SAD the update factor $\gamma_{t}^{\text{SA}}$ can
-be thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
-$t_0$ should have dimensions of entropy.  We argue that we can write
-this entropy `term' geometrically as an area bounded by the $\ln{D(E)}$
-i.e. the entropy. This gives the following
-\begin{align}
-\gamma_{t}^{\text{SAD}} =
-\frac{N_E}{t}\frac{E_{H}-E_{L}}{3T_{\text{min}}}
-\end{align}
-where $N_E$ is the number of energies that have been found. The change
-in the entropy the minimum temperature can be thought of as a slope of
-the triangle made by connecting the minimum important energy and the
-maximum entropy state.  The factor of $3$ comes from the ratio of areas
-bounded by the $\ln{D(E)}$ and the minimum temperature line. The update
-factor has the same $1/t$ behavior as the original SAMC algorithm;
-however now every time a new energy is discovered during the course of
-the simulation, the update factor experiences a jump. This behavior
-allows SAD to dynamically prevent the update factor from decreasing too
-rapidly. In contrast to SAD, Fig.~\ref{fig:gamma-vs-t}. shows the
-monotonic decrease of the update factor for SAMC.  The update factor
-for WL decreases like a step function while the update factor for SAD
-dynamically fluctuates around a value less than 1 for early MC moves
-(while finding new energy states).  Ultimately once all energy states
-have been found, the update factor for SAD decreases in the same
-monotonic fashion as SAMC.
-
-{\color{red}
-While for SAMC the update factor is defined in the original
 implementation, for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
 thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
+the SAMC parameter
 $t_0$ should have dimensions of entropy.
@@
 \end{align}
-We approximate this energy difference by the energy difference
-$E_H -E_L$ where these energies are defined below in the following
-section.  Early in the simulation, when the energy range is still
-expanding, this average entropy is a good choice
-for $t_0$.  However, as the simulation progresses, our entropy factor
-needs to scale with the total number of \davidsays{interesting} energy states that have been
-found $N_S$, since the product $N_S\langle S\rangle$ corresponds to the
-total change of entropy required, given that each of the $N_S$ states
-will has a separate weight $w(E)$.  Finally, at \emph{long} times, after
-all the energies have been found and the weights are being refined
-we wish for a lower $t_0$ in order to reduce the fluctuations in the
-weights that result from large values for $\gamma_t$.  We enable this
+We approximate this energy difference by
+$E_H -E_L$ where $E_H$ and $E_L$ are defined below.
+The entropy numerator of the update factor
+in general should scale with the total number of interesting
+energy states $N_S$, since updates to the weights are distributed between
+that many energy states.  The product $N_S\langle S\rangle$ is the
+total change of entropy required (starting from constant weights) to find
+the true entropy, and puts a lower bound on the convergence time.
+After \emph{long} times, when
+all the energies have been long ago found,
+we wish for a lower $t_0$ in order to more rapidly refine the remaining
+error in entropy.  We enable this
 by tracking the last time we encountered a new energy $t_L$, and
 gradually transition to lower $t_0$ (but still asymptotically scaling
-as $\gamma_t \propto 1/t$ to ensure eventual convergence).  The SAD
+as $\gamma_t \propto 1/t$ to ensure eventual convergence).  Finally,
+we wish for an update factor that is \emph{never} greater than 1, because
+a very large update factor could introduce errors in entropy that may take
+many iterations to remove.  The SAD
 expression for $\gamma_t$ which incorporates these ideas is:
@@
   \gamma_{t}^{\text{SAD}} =
-     N_S \frac{
+     \frac{
        \frac{E_{H}-E_{L}}{3T_{\text{min}}} + \frac{t}{t_L}
      }{
-       N_S^2 + \frac{t^2}{t_L}
+       \frac{E_{H}-E_{L}}{3T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
      }
@@
 SAMC algorithm;
-however now every time a new energy is found in the
-\davidsays{interesting energy range} during the course of
-the simulation, the update factor experiences a jump. This behavior
+however now every time a new energy is found to be important
+during the course of
+the simulation, the update factor will experience a jump. This behavior
 allows SAD to dynamically prevent the update factor from decreasing too
-rapidly. Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for SAD, WL,
+rapidly.
+
+Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for SAD, WL,
 and SAMC.  The update factor
@@
 to SAD, but described after we see the plot}.
-}
 
-Since SAD does not explore \emph{all} energy states it needs to have a
-reasonable idea
-of what energy range corresponds to the temperature range of interest,
+Since SAD does not explore \emph{all} energy states it needs determine
+what energy range corresponds to the temperature range of interest,
 defined by $T_{\min}<T<\infty$.
@@
 where $E(T)$ is the energy that maximizes $S-E/T$.  In the course of the
-simulation this precise energy may be challenging to evaluate accurately.
+simulation this precise energy is challenging to evaluate accurately.
 In order to ensure that we sample this entire energy range adequately,
 we define two energy limits:  a high energy $E_H$ and a low
-energy $E_L$ to define the range over which the energy histogram
-is made flat. $E_H$ is chosen to be the greatest energy that
-maximized the entropy at any time during the simulation.  Similarly,
-$E_L$ is the lowest energy that ever maximized $S(E)-E/T_{\min}$ during
-the simulation.  This creates a ``ratcheting'' effect, in which $E_H$
+energy $E_L$, which define the range over which the energy histogram
+is made flat. At move $t$, $E_H$ and $E_L$ are the greatest and lowest
+energy that has prior to that move
+had the highest histogram value (i.e. been visited the most times) in
+the course of the simulation.
+This definition results in a ``ratcheting'' effect, in which $E_H$
 may only increase, while $E_L$ may only decrease over the course of the
@@
 
-Outside of the energy range $E_L \le E \le E_H$, the weights are
-determined from the accumulated energy histogram. For each energy $E_i
-> E_H$, the energy weights are calculated
-\begin{align}
-w(E_i)_{t+1} = w(E_H)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right)
+During the simulation when considering a move outside of the energy
+range $E_L \le E \le E_H$, Boltzmann weights are used.  If $E\ge E_H$,
+the weight is taken to be
+\begin{align}\label{eq:highw}
+  w(E>E_H) &= w(E_H),
 \end{align}
-assuming of course that the simulation has collected statistics at that energy.
-For each energy $E_i
-< E_L$, the energy weights are calculated
-\begin{align}
-w(E_i)_{t+1} = w(E_L)_{t} + \ln\left(\frac{H(E_i)}{H_{\max}}\right) +
-\frac{E_i - E_L}{T_{\min}}
+which corresponds to an infinite temperature.
+This choice ensures that
+if the maximum in entropy is at an energy $E_{\max}>E_H$, then the energy
+$E_{\max}$
+will eventually have the highest number of counts and the ratcheting will
+result in $E_H\ge E_{\max}$ .
+At lower energies, Boltzmann weights corresponding to the minimum temperature
+are used:
+\begin{align}\label{eq:loww}
+  w(E<E_L) &= w(E_L)e^{-\frac{E_L-E}{T_{\min}}}.
 \end{align}
+This choice has the result that if the energy $E_{\min}$ at which the free
+energy at $T_{\min}$ is minimized is less than $E_L$, the lower energy
+limit will ratchet down to include $E_{\min}$.
+Each time we change the value of $E_H$ or $E_L$, the weights within the
+new portion of the interesting energy range are set to the expressions
+in Equations~\ref{eq:highw} and~\ref{eq:loww}.
 

### 2018-08-20 papers/histogram/sad.tex
more text on convergence

@@
 
-{\color{red}
-It is for this reason that we propose a way to determine the update factor
-``dynamically'' while still satisfying the criteria for convergence.
-For the update factor to scale well across both discrete and continuous systems,
-it must include the number of states $N_S$. We define the average entropy as
-$\bar S$ and the time since the simulation last found a new energy state $t_L$.
-We then can write the following expression for the \emph{proposed} update factor.
-\begin{align}
-\gamma = N_S \frac{\bar S + \nicefrac{t}{t_L}}{N_{S}^2 + \nicefrac{t^2}{t_L}}
-\end{align}
-We now consider the following limiting cases:
-\subsubsection{$t \propto N_S \approx t_L$ and $ t << N_S$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{\bar S}{N_S} \approx \frac{\bar S}{t}
-\end{align}
-Because $t \propto N_S$, the update factor scales as 1/t which is consistent
-with the requirement for SAMC convergence.
-\subsubsection{$t >> N_{S}^2$ and $ t\approx t_L$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{N_S \bar S t_L}{t^2} \approx \frac{N_S \bar S}{t}
-\end{align}
-\subsubsection{$t >> t_L \bar S$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{N_S}{t}
-\end{align}
-We now integrate (sum) the update factor from the last time an energy state was found
-to the final simulation time to ensure that convergence is guaranteed.
-\begin{align}
-\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[ \int_{t_L}^{t_f}
-\frac{\bar S -1}{t^2} dt + \frac{1}{t_L}\int_{t_L}^{t_f} \frac{1}{t} dt \right]
-\end{align}
-From this we obtain
-\begin{align}
-\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[\left(\bar S -1\right)\left(\frac{1}{t_L}-\frac{1}{t_f}\right) + \frac{1}{t_L} \ln\left(\frac{t_f}{t_L}\right) \right]
-\end{align}
-Simplifying, we see that as $t_f \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
-the requirement for SAMC convergence.
-\begin{align}
-\gamma \approx \left(1-\frac{t_L}{t_f}\right)N_S\left(\bar S -1\right) + N_S \ln\left(\frac{t_f}{t_L}\right)
-\end{align}
-}
-
 Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t$ by a factor
@@
 %$\gamma_t$ that should robustly converge on a variety of systems.
+
+\subsubsection{Convergence}
+The difficulty in using the SAMC method lies in identifying an appropriate
+value for $t_0$.  A value that is either too high or too low will result
+in slow convergence to the true entropy.  It is instructive to consider
+separately values of $t_0$ that are too low or too high.
+
+We can place a rigorous \emph{lower} bound $t_{\min}$ on the number of
+moves required to find the true energy by considering the total change
+that needs to be made to the entropy.
+\begin{align}
+  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}
+\end{align}
+The minimum number of moves that could converge the entropy is
+the number of moves that will enable the above total entropy change,
+which we can approximate using an integral:
+\begin{align}
+   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t \\
+  &= \int_0^{t_{\min}} \gamma(t)dt \\
+  &= t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
+  \\
+  &= t_0\left(1 + \ln\left(\frac{t_{\min}}{t_0}\right)\right)
+\end{align}
+Solving for ${t_{\min}}$ we find that
+\begin{align}
+  {t_{\min}} &= t_0 e^{\frac{\Delta S_{\text{tot}}}{t_0} - 1}
+\end{align}
+which means that the minimum time to convergence grows exponentially
+as $t_0$ is made smaller.  You \emph{seriously} don't want to underestimate
+$t_0$!
+
+One might reasonably choose to err on the high side when selecting a $t_0$.
+The rate of convergence is harder to estimate when $t_0$ is large, but
+in general $\gamma_t$ itself forms a lower bound on the accuracy with which
+the entropy may be known, with an unknown prefactor which is roughly
+the coherence time of the Monte Carlo simulation.  Since $\gamma_t$ is
+given by $t_0/t$, the time to converge to a given accuracy is increased
+in proportion to the ratio by which we overestimate $t_0$.  Thus, while
+it is exponentially painful to underestimate $t_0$, overestimating by
+several orders of magnitude is also not acceptable.  We should note
+that these extreme limiting cases do not preclude the possibility that there
+is a wide range of $t_0$ values that lead to an acceptable convergence
+rate.
 
@@
 
+\subsection{Convergence}
+{\color{red}
+It is for this reason that we propose a way to determine the update factor
+``dynamically'' while still satisfying the criteria for convergence.
+For the update factor to scale well across both discrete and continuous systems,
+it must include the number of states $N_S$. We define the average entropy as
+$\bar S$ and the time since the simulation last found a new energy state $t_L$.
+We then can write the following expression for the \emph{proposed} update factor.
+\begin{align}
+\gamma = N_S \frac{\bar S + \nicefrac{t}{t_L}}{N_{S}^2 + \nicefrac{t^2}{t_L}}
+\end{align}
+We now consider the following limiting cases:
+\subsubsection{$t \propto N_S \approx t_L$ and $ t << N_S$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{\bar S}{N_S} \approx \frac{\bar S}{t}
+\end{align}
+Because $t \propto N_S$, the update factor scales as 1/t which is consistent
+with the requirement for SAMC convergence.
+\subsubsection{$t >> N_{S}^2$ and $ t\approx t_L$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{N_S \bar S t_L}{t^2} \approx \frac{N_S \bar S}{t}
+\end{align}
+\subsubsection{$t >> t_L \bar S$}
+In this limiting case, we find that the update factor reduces to
+\begin{align}
+\gamma \approx \frac{N_S}{t}
+\end{align}
+We now integrate (sum) the update factor from the last time an energy state was found
+to the final simulation time to ensure that convergence is guaranteed.
+\begin{align}
+\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[ \int_{t_L}^{t_f}
+\frac{\bar S -1}{t^2} dt + \frac{1}{t_L}\int_{t_L}^{t_f} \frac{1}{t} dt \right]
+\end{align}
+From this we obtain
+\begin{align}
+\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[\left(\bar S -1\right)\left(\frac{1}{t_L}-\frac{1}{t_f}\right) + \frac{1}{t_L} \ln\left(\frac{t_f}{t_L}\right) \right]
+\end{align}
+Simplifying, we see that as $t_f \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
+the requirement for SAMC convergence.
+\begin{align}
+\gamma \approx \left(1-\frac{t_L}{t_f}\right)N_S\left(\bar S -1\right) + N_S \ln\left(\frac{t_f}{t_L}\right)
+\end{align}
+}
+
 \section{Results}\label{sec:results}

### 2018-08-22 papers/histogram/sad.tex
cut one line from equations
made changes to convergence section for SAD.

@@
    \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t \\
-  &= \int_0^{t_{\min}} \gamma(t)dt \\
+  % &= \int_0^{t_{\min}} \gamma(t)dt \\
   &= t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
@@
 {\color{red}
-It is for this reason that we propose a way to determine the update factor
-``dynamically'' while still satisfying the criteria for convergence.
-For the update factor to scale well across both discrete and continuous systems,
-it must include the number of states $N_S$. We define the average entropy as
-$\bar S$ and the time since the simulation last found a new energy state $t_L$.
-We then can write the following expression for the \emph{proposed} update factor.
+Although SAD determines the update factor ``dynamically'', the method must still 
+satisfy the criteria for convergence. Similar to the discussion that was presented
+on the convergence for SAMC, we approximate the total entropy change using an integral.
+The update factor is integrated from the last time an energy has been discovered $t_L$ to
+some minimum time $t_{\min}$ necessary for convergence.
 \begin{align}
-\gamma = N_S \frac{\bar S + \nicefrac{t}{t_L}}{N_{S}^2 + \nicefrac{t^2}{t_L}}
+\Delta S_{\text{tot}} = \int_{t_L}^{t_{\min}} \gamma dt \approx t_L + t_L \ln\left(\frac{t_{\min}}{t_L}\right) \\
++ N_S t_L \Delta S_{\text{tot}} \left(\frac{1}{t_{\min}}-\frac{1}{t_L}\right) 
 \end{align}
+
+Simplifying, we see that as $t_{\min} \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
+the requirement for SAMC convergence. We can solve for the minimum time
+\begin{align}
+\Delta S_{\text{tot}} = \Delta S_{\text{SAMC}} + N_S t_L \Delta S_{\text{tot}} \left(\frac{1}{t_{\min}}-\frac{1}{t_L}\right)
+\end{align}
+Solving for the minimum time necessary $t_{\min}$, we find that
+\begin{align}
+t_{\min} = t_L \left( \frac{N_S \Delta S_{\text{tot}}}{\Delta S_{\text{tot}} + \Delta S_{\text{SAMC}}} + 1\right)
+\end{align}
+Immediately we see that the minimum time necessary is roughly proportional
+to the last time the simulation has found an energy times a ratio of entropies.
+
+\jpsays{Are these right??? Check!}
 We now consider the following limiting cases:
@@
 \end{align}
-We now integrate (sum) the update factor from the last time an energy state was found
-to the final simulation time to ensure that convergence is guaranteed.
-\begin{align}
-\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[ \int_{t_L}^{t_f}
-\frac{\bar S -1}{t^2} dt + \frac{1}{t_L}\int_{t_L}^{t_f} \frac{1}{t} dt \right]
-\end{align}
-From this we obtain
-\begin{align}
-\int_{t_L}^{t_f} \gamma dt \approx t_L N_S \left[\left(\bar S -1\right)\left(\frac{1}{t_L}-\frac{1}{t_f}\right) + \frac{1}{t_L} \ln\left(\frac{t_f}{t_L}\right) \right]
-\end{align}
-Simplifying, we see that as $t_f \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
-the requirement for SAMC convergence.
-\begin{align}
-\gamma \approx \left(1-\frac{t_L}{t_f}\right)N_S\left(\bar S -1\right) + N_S \ln\left(\frac{t_f}{t_L}\right)
-\end{align}
+
 }

### 2018-08-29 papers/histogram/sad.tex
update the paper and figure colors (ready for total edit!)

@@
 \subsection{Convergence}
-{\color{red}
 Although SAD determines the update factor ``dynamically'', the method must still
@@
 \begin{align}
-\Delta S_{\text{tot}} = \int_{t_L}^{t_{\min}} \gamma dt \approx t_L + t_L \ln\left(\frac{t_{\min}}{t_L}\right) \\
-+ N_S t_L \Delta S_{\text{tot}} \left(\frac{1}{t_{\min}}-\frac{1}{t_L}\right) 
+\Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}} \gamma dt \\
+&\approx N_S \ln\left(\frac{t_{\min}}{t_L}\right)
+   + \Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
 \end{align}
-
 Simplifying, we see that as $t_{\min} \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
@@
 \begin{align}
-\Delta S_{\text{tot}} = \Delta S_{\text{SAMC}} + N_S t_L \Delta S_{\text{tot}} \left(\frac{1}{t_{\min}}-\frac{1}{t_L}\right)
+\Delta S_{\text{tot}} \approx N_S \ln\left(\frac{t_{\min}}{t_L}\right)
++ \Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
 \end{align}
@@
 \begin{align}
-t_{\min} = t_L \left( \frac{N_S \Delta S_{\text{tot}}}{\Delta S_{\text{tot}} + \Delta S_{\text{SAMC}}} + 1\right)
+\frac{t_{\min}}{t_L} = e^{\frac{\langle S\rangle t_L}{t_{\min}}}
 \end{align}
-Immediately we see that the minimum time necessary is roughly proportional
-to the last time the simulation has found an energy times a ratio of entropies.
+The solution for $t_{\min}$ can be written in terms of the Lambert-W function
+\begin{align}
+t_{\min} = \frac{3\langle S\rangle}{W(3\langle S\rangle e^{2\langle S\rangle})} t_L
+\end{align}
+For $\langle S\rangle \rightarrow 0$, we find that the minimum time
+$t_{\min} \rightarrow t_L$ and for $\langle S\rangle \rightarrow \infty$,
+we find that the minimum time $t_{\min} \rightarrow \frac{3}{2} t_L$.
 
-\jpsays{Are these right??? Check!}
-We now consider the following limiting cases:
-\subsubsection{$t \propto N_S \approx t_L$ and $ t << N_S$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{\bar S}{N_S} \approx \frac{\bar S}{t}
-\end{align}
-Because $t \propto N_S$, the update factor scales as 1/t which is consistent
-with the requirement for SAMC convergence.
-\subsubsection{$t >> N_{S}^2$ and $ t\approx t_L$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{N_S \bar S t_L}{t^2} \approx \frac{N_S \bar S}{t}
-\end{align}
-\subsubsection{$t >> t_L \bar S$}
-In this limiting case, we find that the update factor reduces to
-\begin{align}
-\gamma \approx \frac{N_S}{t}
-\end{align}
-
-}
+We can examine the limiting cases for SAD and see that the behavior at
+short times is analogous to the weight-based family algorithms $\gamma \approx 1$.
+In the limit that the number of moves $t >> t_L$, $\gamma \approx \nicefrac{N_S}{t}$
+which has a familiar 1/t form.
 
@@
 well-width of $\lambda = 1.30$ and a filling fraction of
-$\eta = 0.30$. We also explore various translation scales $\delta_0 = 0.005,
-0.05,$ and $0.5$. The simulation explores the
-energy space of the system at a reduced temperature of
+$\eta = 0.30$. We also explore various translation scales $\delta_0 = 0.005 and
+0.05$. The simulations explore the
+energy space of the systems at reduced temperatures of
 $T_{\text{min}} = 0.2$ and $0.5$.  All simulations lead to the minimum important
@@
 of these parameters are needed a priori).  In the sections below, we
-examine systems parameters on the convergence of the methods.
+examine the systems parameters impact on the convergence of the methods.
 
-We use the average entropy error vs moves a metric to compare
+We use the average entropy error vs moves as a metric to compare
 simulation runtimes and overall convergence. The overall accuracy
@@
 a `golden' standard. For each simulation, the `golden' standard is
-SAMC for the best choice of $t_0$ for $10^{11}$ moves followed by TMMC.
+SAMC for the best choice of $t_0$.
 \begin{align}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
-  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
-  SAMC followed by TMMC serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-avg-error}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.2$ with
+  SAMC $t_0 = 10^5$ serving as the `golden' standard.}\label{fig:N50-ff0.3-avg-error}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
-  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
-  SAMC followed by TMMC serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-avg-error}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.2$ with
+  SAMC $t_0 = 10^5$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
+  serving as the `golden' standard.}\label{fig:N50-ff0.3-avg-error-slow}
 \end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-fast.pdf}
-  \caption{The average entropy error for each method is shown for $\lambda = 0.30$ with
-  SAMC followed by TMMC serving as the `golden'
-  standard.}\label{fig:N50-ff0.3-avg-error}
-\end{figure}
+%\begin{figure}
+  %\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-fast.pdf}
+  %\caption{The average entropy error for each method is shown for $\eta = 0.30$ with
+  %SAMC followed by TMMC serving as the `golden'
+  %standard.}\label{fig:N50-ff0.3-avg-error-fast}
+%\end{figure}
 \begin{figure}
 \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
-  \caption{The average entropy error is shown for $\lambda = 0.30$ with TMI acting as the
-  `golden' standard.}\label{fig:n500-ff0.3}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.5$ with
+  SAMC $t_0 = 10^6$ serving as the `golden' standard.}\label{fig:n500-ff0.3}
 \end{figure}
 
-~\jpsays{We have discussed including a figure for the $\lambda = 30\%$ and $N = 50$
-showing translation scale impact. Also a figure for the $\lambda = 30\%$ and $N = 50$
-showing the impact of choice of minimum temperature.}
+\subsection{A periodic system with $\eta = 30\%$ and $N = 50$}
 
-\subsection{A periodic system with $\lambda = 30\%$ and $N = 50$}
+Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves
+for a system consisting of 50 atoms that are placed in a periodic
+simple cubic unit cell with lattice constant chosen to give a $\eta =
+0.30$.  These particles are randomly inserted into the system and
+random moves are proposed (with a translation scale $\delta_0 = 0.05$).
+For the $\eta = 30\%$ and $N = 50$ system, we explore the choice of
+$\delta_0$ as a parameter. Smaller translation scales cause the given
+method to take additional time to explore all energies. Ideal methods
+should scale linearly with the $\delta_0$, that is, a decrease in the
+translation scale should result in a proportional increase in
+convergence time. However, as can be seen from
+Fig.~\ref{fig:N50-ff0.3-avg-error-slow} SAMC methods that rapidly
+converged for $\delta = 0.05$ do not proportionaly converge for a
+translation scale of $\delta = 0.005$. We see that SAD handles
+arbitrary translation scales and proportionaly converges based on the
+translation scale that is chosen.
 
-Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves for
-a system consisting of 50 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a $\lambda = 0.30$.  These particles
-are randomly inserted into the system and random moves are proposed (with
-a translation scale $\delta_0 = 0.05$).
-
-\subsubsection{translation scale}
-For the $\lambda = 30\%$ and $N = 50$ system, we explore the choice of $\delta_0$
-as a parameter. Smaller translation scales should cause the given method to take
-additional time to explore all energies.  A larger translation scale could result
-in the simulation taking shorter or longer (because longer distance moves may end up
-not being allowed). Ideal methods should scale linearly with the $\delta_0$, that is,
-a decrease in the translation scale should result in a proportional increase in convergence
-time. Some methods such as WL may not be able to handle arbitrary $\delta_0$.
-
-\subsubsection{minimum temperature}
-For the $\lambda = 30\%$ and $N = 50$ system, we explore the choice of $T_{\min}$
-as a parameter. We have theoretically postulated that a dynamic version of SAMC which
-we call SAD should depend upon this physically meaningful parameter. We thus explore
-the behaviour of SAD for two different reduced temperatures.
-
-\subsection{A periodic system with $\lambda = 30\%$ and $N = 500$}
+\subsection{A periodic system with $\eta = 30\%$ and $N = 500$}
 
 A system is configured with dimensions $x = y = z = 19\sigma$.  A
-filling fraction $\lambda = 0.30$ is chosen which yields $N = 500$
+filling fraction $\eta = 0.30$ is chosen which yields $N = 500$
 particles in a cubical lattice system.  These particles are randomly inserted
@@
 $\delta_0 = 0.05$).
+
+For the larger system, SAMC parameter values of $t_0$ that gave rapid convergence for
+the $N = 50$ system with $\delta = 0.05$ are too small.  Several additional
+larger $t_0$ parameters are chosen to compare with SAD.  SAD has significantly smaller
+fluctuations in the average entropy error for a given number of moves.
 
@@
 for the Wang-Landau algorithm decreases much more rapidly than
-$\nicefrac{1}{\sqrt{t}}$. It can be observed that all the methods from
-this family struggle to converge if the update factor decreases faster
-than $\nicefrac{10^5}{\sqrt{t}}$. Fig.~\ref{fig:n500-ff0.3} shows both
+$\nicefrac{1}{\sqrt{t}}$. It can be observed  for the simulations that were carried out that all the methods from
+the weight-based family struggle to converge if the update factor decreases too rapidlly. Fig.~\ref{fig:n500-ff0.3} shows
 WL and SAMC with inappropriate choices of $t_0$ i.e. less than
@@
 determining which methods will converge in a reasonable amount of time.
-We compare each method to a `golden' standard which we define to be an
-ideal choice of SAMC for $10^{11}$ moves followed by TMMC.  This gives an
+We compare each method to a `golden' standard SAMC simulation for an ideal choice of $t_0$.  This gives an
 unbiased comparison with our developed method SAD. The dynamic version
 of SAMC which we call SAD performs as well as SAMC (assuming the user
-can correctly identify an appropriate choice of $t_0$).  SAD
+can correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
 outperforms WL and WL-TMMC for the same given number of moves. The
@@
 
-We have introduced a new algorithm SAD (which is Stochastic Approximation
-with the modification factor dynamically chosen) that effectively
+We have introduced a new algorithm SAD (Stochastic Approximation
+with a Dynamic $\gamma$) that effectively
 samples the entire energy space for a variety of system sizes.  Also,
@@
 converging in a reasonable amount of time. SAD also converges more quickly
-than TMMC (the reader will remember that SAMC followed by TMMC is what each
-method is compared against) and much more quickly than WL-TMMC.
+than Wl, TMMC, and WL-TMMC.
 

### 2018-09-28 papers/histogram/sad.tex
edit paper with Jordan

@@
 
-\subsubsection{Convergence}
+\subsection{SAMC Convergence}\label{sec:samc-convergence}
 The difficulty in using the SAMC method lies in identifying an appropriate
@@
      \frac{
-       \frac{E_{H}-E_{L}}{3T_{\text{min}}} + \frac{t}{t_L}
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
      }{
-       \frac{E_{H}-E_{L}}{3T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
      }
@@
 \subsection{Convergence}
-Although SAD determines the update factor ``dynamically'', the method must still
-satisfy the criteria for convergence. Similar to the discussion that was presented
-on the convergence for SAMC, we approximate the total entropy change using an integral.
-The update factor is integrated from the last time an energy has been discovered $t_L$ to
-some minimum time $t_{\min}$ necessary for convergence.
+Although SAD determines the update factor ``dynamically'', the method
+must still satisfy the SAMC convergence criteria. As in
+Section~\ref{sec:samc-convergence} on the convergence for SAMC, we
+will approximate the total entropy change using an integral.  We
+integrate update factor from the last time an energy has been
+discovered $t_L$ to some minimum time $t_{\min}$ necessary for
+convergence.  This gives us a lower bound on the total entropy change
+that can be acheived in $t_{\min}$ moves.
 \begin{align}
-\Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}} \gamma dt \\
+\Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
+     \frac{
+       3\langle S\rangle + \frac{t}{t_L}
+     }{
+       3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
+     } dt \\
 &\approx N_S \ln\left(\frac{t_{\min}}{t_L}\right)
@@
 \end{align}
-Simplifying, we see that as $t_{\min} \rightarrow \infty$ then $\gamma \rightarrow \infty$ satisfying
-the requirement for SAMC convergence. We can solve for the minimum time
+where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
+that it takes a long time to find all the interesting energy states.
+Simplifying, we see that as $t_{\min} \rightarrow \infty$ then $\gamma
+\rightarrow \infty$ satisfying the requirement for SAMC
+convergence. We can solve for the minimum time
 \begin{align}

### 2018-09-28 papers/histogram/sad.tex
abstract update for paper.

--- 2018-08-29 papers/histogram/sad.tex
@@
 \begin{document}
-\title{Stochastic Approximation Monte Carlo with Dynamic Update
+\title{Stochastic Approximation Monte Carlo with a Dynamic Update
 Factor (SAD)
@@
 \begin{abstract}
-  We present several histogram methods and compare the performance and
-  efficiency at treating the square-well fluid.
+  We present a novel Monte-Carlo algorithm based on Stochastic
+  Approximation Monte Carlo (SAMC) algorithm for directly calculating
+  the density of states. The proposed method is Stochastic
+  Approximation with a dynamic update factor (SAD-$\gamma$ or SAD)
+  dynamically adjusts the update factor $\gamma$ during the course of
+  the simulation. We test this MC method on the square-well fluid and
+  compare the convergence time and entropy error for a variety of weight-based
+  and transition matrix Monte-Carlo methods. SAD rapidly converges to the
+  correct density of states without the need for the user to specify an
+  arbitrary tunable parameter $t_0$ as in the case of SAMC.
 \end{abstract}

### 2018-11-07 papers/histogram/sad.tex

@@
 \maketitle
-
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/N500-lndos-comparison}
-  \caption{Put this in the right place}
-\end{figure}
 
@@
 
-{\color{red} Both weight-based and transition matrix methods calculate
+Both weight-based and transition matrix methods calculate
 the density of states $D(E)$ for discrete energy
@@
 Energy bins are typically of uniform size for the entire energy
-continuum~\cite{fasnacht2004adaptive}. Coarse binning yields fewer
-energy states to explore but more time spent in each individual bin.
-Fine binning yields more energy states to explore with less time spent
-in each bin. Some methods such as AdaWL~\cite{koh2013dynamically}
+continuum~\cite{fasnacht2004adaptive}. Some methods such as
+AdaWL~\cite{koh2013dynamically}
 employ a tunable mechanism for controlling the binning for low entropic
-states in order to ensure the exploration of all energies. }
+states in order to ensure the exploration of all energies.  The method
+introduced in this paper is designed to scale appropriately as bin size
+is changed, but we do not test this scaling, as we use a system with
+discrete energy levels.
 
@@
 that convergence can be far too slow to be practical.
+Liang et al. give a rule of thumb in which $t_0$
+is chosen in the range from $2N_S$ to $100N_S$ where $N_S$ is the number
+of energy bins~\cite{liang2007stochastic}.  Schneider et al. found
+that for the
+Ising model this heuristic is helpful for small spin
+systems, but that larger systems require an even higher $t_0$
+value~\cite{schneider2017convergence}.  We find that for the square-well
+fluid also $t_0$ in some cases needs to be two orders of magnitude higher
+than the rule of thumb of $100N_S$.
 
@@
 $\gamma_0$.  While this may result in an improved rate of convergence,
-it adds yet another parameter that must be empirically determined.
+it adds yet another parameter that must be empirically determined, and
+we have not explored this degree of freedom.
 
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/N50-lndos-comparison}
+  \caption{The entropy of a square well fluid with 50 atoms and filling fraction
+        0.3 as the green line.  The green hatched area reflects the
+        minimum entropy change needed to converge to the true value.
+        The light blue area is the quadratic approximation
+        for the change in entropy.  The vertical dotted lines represent
+        the energy corresponding to $T=0.2$ and $T=\infty$.}
+\end{figure}
+
 While for SAMC the update factor is defined in the original
@@
 The update factor for SAD in contrast fluctuates
-dynamically around a value less than 1 for early MC moves
-(while finding new energy states), before dropping in a \davidsays{similar
-to SAD, but described after we see the plot}.
+dynamically around a value less than 1 for early MC moves.
+{\color{red} For intermediate MC moves, $\gamma_t$ decreases as
+$1/t$ while continuing to fluctuate whenever a new important energy
+is discovered.  Once all the important energy states are discovered, the
+update factor decreases faster than $1/t$ until asymptoting to $N_S/t$.
+}
 
@@
   \caption{The update factor $\gamma_t$ for $N=50$ versus iteration for three
-    different methods: WL, SAMC, and SAD.  This is exciting because we
-    have different MC moves (slow, etc).}
+    different methods: WL, SAMC, and SAD.}
 \end{figure}
@@
 independently for a reasonable range of binning because $\gamma \propto
-\nicefrac{N_{E}}{t}$.  As the number of energy states found $N_{E}$
+\nicefrac{N_S}{t}$.  As the number of energy states found $N_S$
 increases (fine binning), the time spent $t$ in each bin will decrease.
@@
 \subsection{Convergence}
-Although SAD determines the update factor ``dynamically'', the method
-must still satisfy the SAMC convergence criteria. As in
-Section~\ref{sec:samc-convergence} on the convergence for SAMC, we
-will approximate the total entropy change using an integral.  We
-integrate update factor from the last time an energy has been
-discovered $t_L$ to some minimum time $t_{\min}$ necessary for
-convergence.  This gives us a lower bound on the total entropy change
-that can be acheived in $t_{\min}$ moves.
+Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
+convergence criteria.  However, we also want to ensure that SAD avoids
+the exponentially poor convergence time that SAMC suffers when $t_0$ is
+small as discussed in Section~\ref{sec:samc-convergence}.
+To place a lower bound on the convergence of SAD, we will again
+approximate the total entropy change using an integral.  To sidestep
+the dynamical nature of SAD, we ensure that \emph{after discovering the
+last important energy} at time $t_L$ the entropy change
+$\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
+Thus we integrate the update factor from $t_L$ to some minimum time
+$t_{\min}$ necessary for convergence.  Technically this is not a lower
+bound as in the case of SAMC because the algorithm will already have a
+reasonable approximation for the entropy.
 \begin{align}
@@
      } dt \\
-&\approx N_S \ln\left(\frac{t_{\min}}{t_L}\right)
-   + \Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
+&\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
 \end{align}
@@
 that it takes a long time to find all the interesting energy states.
-Simplifying, we see that as $t_{\min} \rightarrow \infty$ then $\gamma
-\rightarrow \infty$ satisfying the requirement for SAMC
-convergence. We can solve for the minimum time
+Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
+of order unity.  This omission only makes our bound more conservative,
+and enables us to solve to find that
 \begin{align}
-\Delta S_{\text{tot}} \approx N_S \ln\left(\frac{t_{\min}}{t_L}\right)
-+ \Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
+  t_{\min} &= \frac32 t_L
 \end{align}
-Solving for the minimum time necessary $t_{\min}$, we find that
-\begin{align}
-\frac{t_{\min}}{t_L} = e^{\frac{\langle S\rangle t_L}{t_{\min}}}
-\end{align}
-The solution for $t_{\min}$ can be written in terms of the Lambert-W function
-\begin{align}
-t_{\min} = \frac{3\langle S\rangle}{W(3\langle S\rangle e^{2\langle S\rangle})} t_L
-\end{align}
-For $\langle S\rangle \rightarrow 0$, we find that the minimum time
-$t_{\min} \rightarrow t_L$ and for $\langle S\rangle \rightarrow \infty$,
-we find that the minimum time $t_{\min} \rightarrow \frac{3}{2} t_L$.
-
-We can examine the limiting cases for SAD and see that the behavior at
-short times is analogous to the weight-based family algorithms $\gamma \approx 1$.
-In the limit that the number of moves $t >> t_L$, $\gamma \approx \nicefrac{N_S}{t}$
-which has a familiar 1/t form.
+which means that we only need to run fifty percent longer after we
+discover the final interesting energy state in order to have made
+sufficient updates to \emph{possibly} converge to the true entropy.
 
@@
 For the square-well fluid, our simulations treat systems with a
-well-width of $\lambda = 1.30$ and a filling fraction of
-$\eta = 0.30$. We also explore various translation scales $\delta_0 = 0.005 and
-0.05$. The simulations explore the
+well-width of $\lambda = 1.3$ and a filling fraction of
+$\eta = 0.3$. We also explore various translation scales $\delta_0 = 0.005R$ and
+$0.05R$. The simulations explore the
 energy space of the systems at reduced temperatures of
@@
 is determined by examining the fractional error of a particular method to
-a `golden' standard. For each simulation, the `golden' standard is
-SAMC for the best choice of $t_0$.
+a precise reference system. For each simulation, the reference system
+is chosen to be the final output of the SAMC simulation with the best
+choice of $t_0$.
 \begin{align}
-\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{golden}}\bigg\rvert
+\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{ref}}\bigg\rvert
 \end{align}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.2$ with
-  SAMC $t_0 = 10^5$ serving as the `golden' standard.}\label{fig:N50-ff0.3-avg-error}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
+  SAMC $t_0 = 10^5$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
 \end{figure}
@@
   \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.2$ with
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
   SAMC $t_0 = 10^5$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
-  serving as the `golden' standard.}\label{fig:N50-ff0.3-avg-error-slow}
+  serving as the reference.}\label{fig:N50-ff0.3-avg-error-slow}
 \end{figure}
@@
   %\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-fast.pdf}
-  %\caption{The average entropy error for each method is shown for $\eta = 0.30$ with
-  %SAMC followed by TMMC serving as the `golden'
-  %standard.}\label{fig:N50-ff0.3-avg-error-fast}
+  %\caption{The average entropy error for each method is shown for $\eta = 0.3$ with
+  %SAMC followed by TMMC serving as the reference system.}\label{fig:N50-ff0.3-avg-error-fast}
 %\end{figure}
-\begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.30$ and $T_{\min} = 0.5$ with
-  SAMC $t_0 = 10^6$ serving as the `golden' standard.}\label{fig:n500-ff0.3}
-\end{figure}
 
-\subsection{A periodic system with $\eta = 30\%$ and $N = 50$}
+\subsection{A periodic system with 50 atoms}
 
@@
 for a system consisting of 50 atoms that are placed in a periodic
-simple cubic unit cell with lattice constant chosen to give a $\eta =
-0.30$.  These particles are randomly inserted into the system and
-random moves are proposed (with a translation scale $\delta_0 = 0.05$).
-For the $\eta = 30\%$ and $N = 50$ system, we explore the choice of
-$\delta_0$ as a parameter. Smaller translation scales cause the given
-method to take additional time to explore all energies. Ideal methods
-should scale linearly with the $\delta_0$, that is, a decrease in the
-translation scale should result in a proportional increase in
-convergence time. However, as can be seen from
+simple cubic unit cell with lattice constant chosen to give a filling
+fraction of $\eta = 0.3$.  These particles are randomly inserted into
+the system and random moves are proposed (with a translation scale
+$\delta_0 = 0.05R$). For the $\eta = 30\%$ and $N = 50$ system, we
+explore the choice of $\delta_0$ as a parameter.
+
+{\color{red}
+$S_{\max} = 120$ and $S_{\min} = 265$ for this system.  The number of
+important energy states for this system is then $N_S = S_{\min} - S_{\max} = 145$.
+The reduced temperature of $1/3$ is chosen in order to examine the convergence
+properties of SAD in a reasonable simulation time.
+}
+
+Smaller translation
+scales cause the given method to take additional time to explore all
+energies. Ideal methods should scale linearly with the $\delta_0$, that
+is, a decrease in the translation scale should result in a proportional
+increase in convergence time. However, as can be seen from
 Fig.~\ref{fig:N50-ff0.3-avg-error-slow} SAMC methods that rapidly
-converged for $\delta = 0.05$ do not proportionaly converge for a
-translation scale of $\delta = 0.005$. We see that SAD handles
+converged for $\delta = 0.05R$ do not proportionaly converge for a
+translation scale of $\delta = 0.005R$. We see that SAD handles
 arbitrary translation scales and proportionaly converges based on the
@@
 
-\subsection{A periodic system with $\eta = 30\%$ and $N = 500$}
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/2$ with
+  SAMC $t_0 = 10^6$ serving as the reference.}\label{fig:n500-ff0.3}
+\end{figure}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/N500-lndos-comparison}
+  \caption{Use this to create description of the system.}\label{fig:n500-lndos-comparison}
+\end{figure}
+
+\subsection{A periodic system with 500 atoms}
 
 A system is configured with dimensions $x = y = z = 19\sigma$.  A
-filling fraction $\eta = 0.30$ is chosen which yields $N = 500$
-particles in a cubical lattice system.  These particles are randomly inserted
-into the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05$).
+filling fraction $\eta = 0.3$ is chosen which yields $N = 500$
+particles in a cubical lattice system.  These particles are randomly
+inserted into the system and random moves are proposed (with a
+translation scale $\delta_0 = 0.05R$).
 
-For the larger system, SAMC parameter values of $t_0$ that gave rapid convergence for
-the $N = 50$ system with $\delta = 0.05$ are too small.  Several additional
-larger $t_0$ parameters are chosen to compare with SAD.  SAD has significantly smaller
-fluctuations in the average entropy error for a given number of moves.
+{\color{red}
+$S_{\max} = 1206$ and $S_{\min} = 2606$ for this system.  The number of
+important energy states for this system is then $N_S = S_{\min} -
+S_{\max} = 1400$. The reduced temperature of $1/2$ is chosen for the
+larger system in order to examine the convergence properties of SAD in
+a reasonable simulation time.
+}
+
+For the larger system, SAMC parameter values of $t_0$ that gave rapid
+convergence for the $N = 50$ system with $\delta = 0.05R$ do not
+converge rapidly.  Several additional larger $t_0$ parameters are
+chosen to compare with SAD.  SAD has significantly smaller fluctuations
+in the average entropy error for a given number of moves.
+
+{\color{red}
+Figure~\ref{fig:n500-lndos-comparison} shows the calculated entropy for
+the $N = 500$ system. The grey dash lines indicate the $S_{\max} = 1206$
+point and the $S_{\min} = 2606$. Only the reference method is shown.
+}
 
@@
 for the Wang-Landau algorithm decreases much more rapidly than
-$\nicefrac{1}{\sqrt{t}}$. It can be observed  for the simulations that were carried out that all the methods from
-the weight-based family struggle to converge if the update factor decreases too rapidlly. Fig.~\ref{fig:n500-ff0.3} shows
-WL and SAMC with inappropriate choices of $t_0$ i.e. less than
-$\nicefrac{10^5}{\sqrt{t}}$ failing to converge in a reasonable amount
-of time. Thus, the update factor $\gamma$, shown as a function of
-simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves useful in
-determining which methods will converge in a reasonable amount of time.
-We compare each method to a `golden' standard SAMC simulation for an ideal choice of $t_0$.  This gives an
-unbiased comparison with our developed method SAD. The dynamic version
-of SAMC which we call SAD performs as well as SAMC (assuming the user
-can correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
+$\nicefrac{1}{\sqrt{t}}$. It can be observed  for the simulations that
+were carried out that all the methods from the weight-based family
+struggle to converge if the update factor decreases too rapidlly.
+Fig.~\ref{fig:n500-ff0.3} shows WL and SAMC with inappropriate choices
+of $t_0$ i.e. less than $\nicefrac{10^5}{\sqrt{t}}$ failing to converge
+in a reasonable amount of time. Thus, the update factor $\gamma$, shown
+as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
+useful in determining which methods will converge in a reasonable
+amount of time. We compare each method to a reference SAMC
+simulation for an ideal choice of $t_0$.  This gives an unbiased
+comparison with our developed method SAD. The dynamic version of SAMC
+which we call SAD performs as well as SAMC (assuming the user can
+correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
 outperforms WL and WL-TMMC for the same given number of moves. The
@@
 
-We have introduced a new algorithm SAD (Stochastic Approximation
-with a Dynamic $\gamma$) that effectively
-samples the entire energy space for a variety of system sizes.  Also,
-SAD proves effective in comparison with other Monte-Carlo methods at
-converging to the correct density of states.  The ease of
-implementation and the absence of user-defined parameters makes SAD an
-ideal algorithm. SAD converges as quickly as SAMC (in the case of an ideally
-chosen $t_0$ for our simulations. SAD doesn't suffer from the weakness of SAMC,
-i.e. the user choosing parameters that could lead to the simulation not
-converging in a reasonable amount of time. SAD also converges more quickly
-than Wl, TMMC, and WL-TMMC.
+We have introduced a new algorithm SAD (Stochastic Approximation with a
+Dynamic $\gamma$) that effectively samples the entire energy space for
+a variety of system sizes.  Also, SAD proves effective in comparison
+with other Monte-Carlo methods at converging to the correct density of
+states.  The ease of implementation and the absence of user-defined
+parameters makes SAD an ideal algorithm. SAD converges as quickly as
+SAMC (in the case of an ideally chosen $t_0$ for our simulations. SAD
+doesn't suffer from the weakness of SAMC, i.e. the user choosing
+parameters that could lead to the simulation not converging in a
+reasonable amount of time. SAD also converges more quickly than Wl,
+TMMC, and WL-TMMC.
 

### 2018-11-09 papers/ising/ising-sad.tex
added some text to the ising paper

@@
 \begin{abstract}
-We present several histogram methods and compare the performance and
-efficiency at treating the Ising model.
+In this work, we finish writing everything before the abstract~\jpsays{FIX AT END}
 \end{abstract}
+
+\maketitle
+
+\section{Introduction}
+In the early 2000s, Wang and Landau introduced a novel flat-histogram
+Monte-Carlo algorithm that accurately determined the density of states
+(DOS) for a statistical
+system~\cite{wang2001determining,wang2001efficient}. For all of its
+power, the method unfortunately requires a priori knowledge of several
+user-defined parameters. Thus, for any given system under study, the
+user needs to determine the ideal parameters in order to apply the
+method. The Wang-Landau algorithm is also known to violate detailed
+balance (although only for brief time intervals)~\cite{yan2003fast,
+shell2002generalization}. With the violation of detailed balance,
+convergence of the algorithm is not guaranteed. The convergence rate
+for WL is well known~\cite{zhou2005understanding,lee2006convergence,
+belardinelli2007wang}. Modification factors that decrease faster than
+$1/t$ lead to non-convergence~\cite{belardinelli2007fast}.  These
+findings lead to the formation of the $1/t$-WL algorithm.  The ideal
+choice of decreasing the modification factor prevents CPU time from
+being wasted by continuing to perform calculations after the error in
+the density of states is saturated~\cite{belardinelli2008analysis}.
+Liang independently considered whether WL could be treated as a special
+case of Stochastic Approximation whose convergence could be
+mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
+2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
+considered a form of Stochastic Approximation Monte Carlo (SAMC).
+Unlike WL, SAMC can guarantee convergence (if certain conditions are
+met). Despite the added benefit of guaranteed convergence, the method
+still has a system specific user-defined variable. Such variables often
+create difficulty when applying Monte-Carlo methods across arbitrary
+systems.  The SAD (Dynamic Stochastic Approximation)
+method~\jpsays{outlined by Pommerenck et.al include David and my paper
+on SAD square well in citations!} is special version of the SAMC
+Algorithm that dynamically chooses the modification factor rather than
+relying on system dependent parameters. SAD shares the same convergence
+properties with SAMC while replacing unphysical user-defined parameters
+with the algorithms dynamic choice.  In this work, we apply the family
+of Monte-Carlo algorithms (WL, SAMC, SAD) to the 2D Potts Model. We
+explore the convergence properties of the three algorithms in an effort
+to demonstrate the system scalability of SAD.
+
+\section{Ising Model}
+The 2D ten state Potts Model was used as a test system for the Wang 
+Landau method during its inception~\cite{wang2001determining,wang2001efficient}.
+
+Apparently Wang and Landau cite~\cite{wu1982potts} because of exact solutions 
+and extensive simulation data are available. Since they use a Q = 10 2D Potts
+Model I assume they are comparing 1 of 2 things.
+(1) The latent heat of the system which can be obtained by plotting U vs T
+and looking at how much U drops at Tc.
+(2) Looking at the value for the critical temperature Tc.
+Or it is possible that they clarify in a later work~\cite{landau2004new} how
+to extract the correct density of states. 
+\begin{align}
+\sum_E g_N\left( E\right)=Q^N
+\end{align}
+Also the number of ground states (where $E = -2N$) is $Q$. He then rescales
+the density of states $g_N\left( E\right)$ and writes
+\begin{align}
+\ln \left[ g_N\left(E\right)\right]=\ln \left[ g\left( E\right)\right]
+-\ln \left[ g\left(E=-2N\right)\right]+\ln\left[ Q \right]
+\end{align}
+He calls this the normallized density of states (for his work Q = 10).
+
+\section{Methods}
+\subsection{Wang-Landau: WL}
+
+\subsection{Stochastic Approximation Monte Carlo: SAMC}
+
+\subsection{Dynamic Stochastic Approximation: SAD}
+
+\section{Results}
+\section{Conclusion}
+\section{Acknowledgements}
 

### 2018-11-30 papers/histogram/sad.tex

@@
 well-width of $\lambda = 1.3$ and a filling fraction of
-$\eta = 0.3$. We also explore various translation scales $\delta_0 = 0.005R$ and
-$0.05R$. The simulations explore the
+$\eta = 0.3$. We also explore various translation scales $\delta_0 = 0.005\sigma$ and
+$0.05\sigma$. The simulations explore the
 energy space of the systems at reduced temperatures of
@@
 the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05R$). For the $\eta = 30\%$ and $N = 50$ system, we
+$\delta_0 = 0.05\sigma$). For the $\eta = 30\%$ and $N = 50$ system, we
 explore the choice of $\delta_0$ as a parameter.
@@
 {\color{red}
-$S_{\max} = 120$ and $S_{\min} = 265$ for this system.  The number of
+$S_{\max} = 120$ and $S_{\min} = 248$ for this system.  The number of
 important energy states for this system is then $N_S = S_{\min} - S_{\max} = 145$.
@@
 properties of SAD in a reasonable simulation time.
+
+As a side note, WL-TMMC does not converge well for $\delta_0 = 0.05\sigma$ while
+it does perform well for $\delta_0 = 0.005\sigma$. The WL portion of the algorithm
+spends nearly 100x longer for the smaller translation scale ensuring that the algorithm
+converges quickly upon transitioning to TMMC.
 }
@@
 Fig.~\ref{fig:N50-ff0.3-avg-error-slow} SAMC methods that rapidly
-converged for $\delta = 0.05R$ do not proportionaly converge for a
-translation scale of $\delta = 0.005R$. We see that SAD handles
+converged for $\delta = 0.05\sigma$ do not proportionaly converge for a
+translation scale of $\delta = 0.005\sigma$. We see that SAD handles
 arbitrary translation scales and proportionaly converges based on the
@@
 
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-defualt.pdf}
+  \caption{The average entropy error for each MC method is shown for $\eta = 0.17$ and $T_{\min} = 1$ with
+  SAD serving as the reference.}\label{fig:n256}
+\end{figure}
 \begin{figure}
@@
 inserted into the system and random moves are proposed (with a
-translation scale $\delta_0 = 0.05R$).
+translation scale $\delta_0 = 0.05\sigma$).
 
@@
 For the larger system, SAMC parameter values of $t_0$ that gave rapid
-convergence for the $N = 50$ system with $\delta = 0.05R$ do not
+convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
 converge rapidly.  Several additional larger $t_0$ parameters are

### 2018-12-12 papers/histogram/sad.tex
made remaining changes to the SAD paper

@@
 systems, but that larger systems require an even higher $t_0$
-value~\cite{schneider2017convergence}.  We find that for the square-well
-fluid also $t_0$ in some cases needs to be two orders of magnitude higher
+value~\cite{schneider2017convergence}.  Also, we find that for the square-well
+fluid $t_0$ in some cases needs to be two orders of magnitude higher
 than the rule of thumb of $100N_S$.
@@
         for the change in entropy.  The vertical dotted lines represent
-        the energy corresponding to $T=0.2$ and $T=\infty$.}
+        the energy corresponding to $T=0.\overline{3}$ and $T=\infty$.}
 \end{figure}
@@
 
+%\begin{figure}
+  %\includegraphics[width=\columnwidth]{figs/gamma-n500}
+  %\caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
+    %different methods: WL, SAMC, and SAD.
+    %method.}\label{fig:gamma-vs-t}
+%\end{figure}
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n500}
-  \caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
+  \includegraphics[width=\columnwidth]{figs/gamma-n256}
+  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for three
     different methods: WL, SAMC, and SAD.
@@
 For the square-well fluid, our simulations treat systems with a
-well-width of $\lambda = 1.3$ and a filling fraction of
-$\eta = 0.3$. We also explore various translation scales $\delta_0 = 0.005\sigma$ and
+well-width of $\lambda = 1.3$ and $\lambda = 1.5$ and a filling fraction of
+$\eta = 0.3$ and $\eta = 0.17$. We also explore various translation scales $\delta_0 = 0.005\sigma$ and
 $0.05\sigma$. The simulations explore the
 energy space of the systems at reduced temperatures of
-$T_{\text{min}} = 0.2$ and $0.5$.  All simulations lead to the minimum important
+$T_{\text{min}} = 0.\overline{3}$ and $1.0$.  All simulations lead to the minimum important
 energy $E_{\min}$ and maximum entropy energy $E_{\max}$
@@
 \end{figure}
-%\begin{figure}
-  %\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-fast.pdf}
-  %\caption{The average entropy error for each method is shown for $\eta = 0.3$ with
-  %SAMC followed by TMMC serving as the reference system.}\label{fig:N50-ff0.3-avg-error-fast}
-%\end{figure}
 
@@
 
-Figure~\ref{fig:N50-ff0.3-avg-error} shows the average entropy vs moves
-for a system consisting of 50 atoms that are placed in a periodic
+Figures~\ref{fig:N50-ff0.3-avg-error}
+and~\ref{fig:N50-ff0.3-avg-error-slow} show the average entropy vs
+moves for a system consisting of 50 atoms that are placed in a periodic
 simple cubic unit cell with lattice constant chosen to give a filling
@@
 the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05\sigma$). For the $\eta = 30\%$ and $N = 50$ system, we
-explore the choice of $\delta_0$ as a parameter.
+$\delta_0 = 0.05\sigma$ and $\delta_0 = 0.005\sigma$ respectively). For
+the $\eta = 30\%$ and $N = 50$ system, we explore the choice of
+$\delta_0$ as a parameter.
 
@@
 is, a decrease in the translation scale should result in a proportional
-increase in convergence time. However, as can be seen from
-Fig.~\ref{fig:N50-ff0.3-avg-error-slow} SAMC methods that rapidly
+increase in convergence time. As can be seen from
+Fig.~\ref{fig:N50-ff0.3-avg-error-slow}, SAMC methods that rapidly
 converged for $\delta = 0.05\sigma$ do not proportionaly converge for a
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-defualt.pdf}
+\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
   \caption{The average entropy error for each MC method is shown for $\eta = 0.17$ and $T_{\min} = 1$ with
@@
 \end{figure}
+%\begin{figure}
+%\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
+  %\caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/2$ with
+  %SAMC $t_0 = 10^6$ serving as the reference.}\label{fig:n500-ff0.3}
+%\end{figure}
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/2$ with
-  SAMC $t_0 = 10^6$ serving as the reference.}\label{fig:n500-ff0.3}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/N500-lndos-comparison}
-  \caption{Use this to create description of the system.}\label{fig:n500-lndos-comparison}
+  \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
+  \caption{Use this to create description of the system.}\label{fig:n256-lndos-comparison}
 \end{figure}
 
-\subsection{A periodic system with 500 atoms}
+\subsection{A periodic system with 256 atoms}
 
-A system is configured with dimensions $x = y = z = 19\sigma$.  A
-filling fraction $\eta = 0.3$ is chosen which yields $N = 500$
-particles in a cubical lattice system.  These particles are randomly
-inserted into the system and random moves are proposed (with a
-translation scale $\delta_0 = 0.05\sigma$).
+Figure~\ref{fig:n256} shows the average entropy vs moves for a system
+consisting of 256 atoms that are placed in a periodic simple cubic unit
+cell with lattice constant chosen to give a filling fraction of $\eta =
+0.17$.   These particles are randomly inserted into the system and
+random moves are proposed (with a translation scale $\delta_0 =
+0.05\sigma$).
 
 {\color{red}
-$S_{\max} = 1206$ and $S_{\min} = 2606$ for this system.  The number of
+$S_{\max} = 509$ and $S_{\min} = 915$ for this system.  The number of
 important energy states for this system is then $N_S = S_{\min} -
-S_{\max} = 1400$. The reduced temperature of $1/2$ is chosen for the
-larger system in order to examine the convergence properties of SAD in
-a reasonable simulation time.
+S_{\max} = 406$. The reduced temperature of $1.0$ and well-width of $\lambda = 1.5$
+are chosen in order to ensure that the simulation encounters liquid vapor coexistence
+and that the results are comparable with NIST.
 }
@@
 convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
-converge rapidly.  Several additional larger $t_0$ parameters are
-chosen to compare with SAD.  SAD has significantly smaller fluctuations
+converge rapidly.  Larger $t_0$ parameters are simulated for
+comparison with SAD.  SAD has significantly smaller fluctuations
 in the average entropy error for a given number of moves.
@@
 {\color{red}
-Figure~\ref{fig:n500-lndos-comparison} shows the calculated entropy for
-the $N = 500$ system. The grey dash lines indicate the $S_{\max} = 1206$
-point and the $S_{\min} = 2606$. Only the reference method is shown.
+Figure~\ref{fig:n256-lndos-comparison} shows the calculated entropy for
+the $N = 256$ system. The grey dash lines indicate the $S_{\max} = 509$
+point and the $S_{\min} = 915$. Only the reference method is shown.
 }
@@
 struggle to converge if the update factor decreases too rapidlly.
-Fig.~\ref{fig:n500-ff0.3} shows WL and SAMC with inappropriate choices
+Fig.~\ref{fig:N50-ff0.3-avg-error-slow} shows SAMC with inappropriate choices
 of $t_0$ i.e. less than $\nicefrac{10^5}{\sqrt{t}}$ failing to converge

### 2019-01-09 papers/histogram/sad.tex
update gamma plot and some text about reference comparison choices.

@@
 Fig.~\ref{fig:N50-ff0.3-avg-error}.
-
+\jpsays{We use a SAMC $t_0 = 10^3$ with a fixed energy range $S_{\max} = 120$ and $S_{\min} = 248$ as the reference.
+This is noted because this is not strictly SAMC method. We use this as a reference to increase convergence time of the
+reference.}
 \begin{figure}
@@
   \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-  SAMC $t_0 = 10^5$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
+  a fixed energy range SAMC $t_0 = 10^3$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
 \end{figure}
@@
   \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-  SAMC $t_0 = 10^5$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
+  a fixed energy range SAMC $t_0 = 10^3$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
   serving as the reference.}\label{fig:N50-ff0.3-avg-error-slow}

### 2019-01-23 papers/histogram/sad.tex

@@
 
-Transition Matrix Monte Carlo~\cite{wang1999transition,
-swendsen1999transition, fitzgerald2000monte} became an attractive
-complimentary simulation algorithm to Wang-Landau, in that, if detailed
-balance is ensured, the system is guaranteed to converge.  Unfortunately,
-the algorithm can be considerably more difficult to implement when
-compared to Wang-Landau: in particular, due to the infinite temperature
-transition matrix~\cite{wang2002transition}.  Also, the time necessary
-for the density of states to converge could be considerable
-due to the algorithm spending too much time exploring low energy states.
-Unlike WL, TMMC does not require a prior knowledge of the energy range of
-the system.
-
-Shell et al.~\cite{shell2003improved, shell2004flat} originially
-implemented Wang-Landau Transition Matrix Monte Carlo (WL-TMMC) in an
-effort to quickly explore the energies of a given system using WL then
-switch over to TMMC to guarantee convergence. Considerable effort has
-been spent trying to determine the flatness criteria and cutoff for the
-WL portion, but there remains no rigorous systematic way to determine
-these parameters~\cite{rane2013monte}.  The algorithm is first run for
-several trial simulations and/or `experience' is used to determine the
-best parameters for a users particular system of
-study~\cite{siderius2013use}.
-
+%Transition Matrix Monte Carlo~\cite{wang1999transition,
+%swendsen1999transition, fitzgerald2000monte} became an attractive
+%complimentary simulation algorithm to Wang-Landau, in that, if detailed
+%balance is ensured, the system is guaranteed to converge.  Unfortunately,
+%the algorithm can be considerably more difficult to implement when
+%compared to Wang-Landau: in particular, due to the infinite temperature
+%transition matrix~\cite{wang2002transition}.  Also, the time necessary
+%for the density of states to converge could be considerable
+%due to the algorithm spending too much time exploring low energy states.
+%Unlike WL, TMMC does not require a prior knowledge of the energy range of
+%the system.
+%
+%Shell et al.~\cite{shell2003improved, shell2004flat} originially
+%implemented Wang-Landau Transition Matrix Monte Carlo (WL-TMMC) in an
+%effort to quickly explore the energies of a given system using WL then
+%switch over to TMMC to guarantee convergence. Considerable effort has
+%been spent trying to determine the flatness criteria and cutoff for the
+%WL portion, but there remains no rigorous systematic way to determine
+%these parameters~\cite{rane2013monte}.  The algorithm is first run for
+%several trial simulations and/or `experience' is used to determine the
+%best parameters for a users particular system of
+%study~\cite{siderius2013use}.
+\jpsays{I should look at some more 1/t-WL citations. I know a few more
+that probably should be included here...}
 Around the same time, the convergence rate for WL was being
@@
 section. We compare it along with several flat histogram methods
-which include TMMC, WL, WL-TMMC, and SAMC.
+which include WL, 1/t-WL, and SAMC.
+%which include TMMC, WL, WL-TMMC, and SAMC.
 
@@
 describe in detail flat-histogram methods.  Section~\ref{sec:weight} outlines
-weight-based histogram methods used in this work. Section~\ref{sec:transition}
-details the workings of the transition-matrix methods studied in this work.
+weight-based histogram methods used in this work.
+%Section~\ref{sec:transition}
+%details the workings of the transition-matrix methods studied in this work.
+
 Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
@@
 works.  The following methods are introduced and applied to the
-square-well fluid: Transition Matrix Monte-Carlo (TMMC), Wang-Landau
-(WL), Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC), Stochastic
-Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
-(SAD).
+square-well fluid: Wang-Landau (WL), 1/t-Wang-Landau (1/t-WL), Stochastic
+Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation (SAD).
+
+%Transition Matrix Monte-Carlo (TMMC), Wang-Landau
+%(WL), Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC), Stochastic
+%Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
+%(SAD).
 
@@
 
-\section{Transition-matrix methods}\label{sec:transition}
-
-\subsection{TMMC Algorithm}
-The previously described weight-based flat-histogram methods utilizes a so-called
-``visited states'' approach to calculate  thermodynamic
-properties~\cite{errington2003direct}. The algorithm is fundamentally
-based on the number of times a given state is visited while the
-system is being sampled. The Transition Matrix Monte Carlo (TMMC) method operates
-differently than the visited states approach~\cite{fitzgerald2000monte}.
-The method uses attempted transitions between states to obtain an estimate of
-the probability that a given transition will be proposed.  The density
-of states is computed from the resulting transition matrix.  An advantage
-of collecting the transition matrix is that this matrix may be collected
-while another method is employed to compute acceptance probabilities, as
-in Wang-Landau TMMC, discussed below.
-
-The TMMC algorithm records the transition data in a \emph{collection matrix},
-i.e. the unnormalized transition histogram matrix, which stores the number
-of times a transition between two energy states has been proposed.
-\begin{align}
-  C(E_{\text{old}} \rightarrow E_{\text{new}},t+1)
-     = C(E_{\text{old}} \rightarrow E_{\text{new}},t) + 1
-\end{align}
-The transition matrix is computed from the collection matrix via
-\begin{align}
-  T(E_{\text{old}} \rightarrow E_{\text{new}})
-   = \frac{C(E_{\text{old}} \rightarrow E_{\text{new}})}{
-     \sum_E C(E_{\text{old}} \rightarrow E)
-  }
-\end{align}
-The acceptance probability is defined in terms of this biasing function:
-\begin{align}
-  \mathcal{P}(E_{\text{old}} \rightarrow E_{\text{new}}) =
-  \max\left(1,\frac{T(E_{\text{old}} \rightarrow E_{\text{new}})}{
-       T(E_{\text{new}} \rightarrow E_{\text{old}})}\right)
-\end{align}
-where we note that the transition matrix is updated to reflect the
-transition under consideration \emph{prior} to computing the acceptance
-probability.  The TMMC method suffers from poor convergence primarily
-because it is sensitive to poor statistics in less probable energy
-transitions.  Like the weight-based methods above, the TMMC method violates
-detailed balance, with the level of violation decreasing as the simulation
-progresses and a difference of 1 in the collection matrix has a smaller
-impact on the acceptance probabilities.
-
-\subsection{WL-TMMC Algorithm}
-Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
-by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
-advantage of the WL methods ease of implementation, while making use of
-a transition matrix to address WL weaknesses, such as its limited
-statistical accuracy convergence.
-
-\paragraph{Initializing Wang-Landau:}
-We set the energy range along which the method may sample as
-$E(T_{\min}) <E< E(T=\infty)$. Conventional Wang-Landau prescribes
-updating the WL factor $\gamma$ by starting at 1.0 and reducing by
-$\frac12$ until the simulation is cutoff typically around $10^{-10}$.
-While variants of this approach appear in the literature, Rane et. al.
-published an implemenation of WL-TMMC~\cite{rane2013monte} that
-suggested using a WL factor between 1.0 and $10^{-2}$ with a cutoff
-$<10^{-5}$, We implement WL-TMMC by setting the WL factor to be 1.0 and
-the cutoff to be either $10^{-4}$ or $10^{-10}$.
-
-A single sweep ``roundtrip'' is defined as each macrostate being
-sampled a number of times in this case 100,000 times. Wang-Landau sets
-a flatness criteria~\cite{wang2001determining, wang2001efficient,
-hatch2015computational, mahynski2017predicting} for the accumulated
-histogram usually to be 0.8.  In the original implementation, it was
-found to be sufficient that each macrostate is visited at least once
-(each histogram bin has one entry)~\cite{shell2003improved}; however,
-for other systems this is rarely sufficient~\cite{bhar2009computer}.
-The goal of this WL initialization is to prefill the transition matrix
-for the next portion of the simulation.
-
-\paragraph{Initializing TMMC:} Prefilling the transition matrix is
-necessary because if numerous zeros existed in the collection matrix,
-the infinite temperature transition matrix would be ill-defined.
-Sampling over large ranges of density of states would therefore take an
-unacceptable length of time to complete~\cite{shell2003improved,
-shen2014elucidating}.  The TMMC portion of the algorithm is run only
-after WL has reached the specified cutoff. An important note is that
-the acceptatnce probability for TMMC is only used to control actual
-transitions not update the collection matrices.
+%\section{Transition-matrix methods}\label{sec:transition}
+%
+%\subsection{TMMC Algorithm}
+%The previously described weight-based flat-histogram methods utilizes a so-called
+%``visited states'' approach to calculate  thermodynamic
+%properties~\cite{errington2003direct}. The algorithm is fundamentally
+%based on the number of times a given state is visited while the
+%system is being sampled. The Transition Matrix Monte Carlo (TMMC) method operates
+%differently than the visited states approach~\cite{fitzgerald2000monte}.
+%The method uses attempted transitions between states to obtain an estimate of
+%the probability that a given transition will be proposed.  The density
+%of states is computed from the resulting transition matrix.  An advantage
+%of collecting the transition matrix is that this matrix may be collected
+%while another method is employed to compute acceptance probabilities, as
+%in Wang-Landau TMMC, discussed below.
+%
+%The TMMC algorithm records the transition data in a \emph{collection matrix},
+%i.e. the unnormalized transition histogram matrix, which stores the number
+%of times a transition between two energy states has been proposed.
+%\begin{align}
+%  C(E_{\text{old}} \rightarrow E_{\text{new}},t+1)
+%     = C(E_{\text{old}} \rightarrow E_{\text{new}},t) + 1
+%\end{align}
+%The transition matrix is computed from the collection matrix via
+%\begin{align}
+%  T(E_{\text{old}} \rightarrow E_{\text{new}})
+%   = \frac{C(E_{\text{old}} \rightarrow E_{\text{new}})}{
+%     \sum_E C(E_{\text{old}} \rightarrow E)
+%  }
+%\end{align}
+%The acceptance probability is defined in terms of this biasing function:
+%\begin{align}
+%  \mathcal{P}(E_{\text{old}} \rightarrow E_{\text{new}}) =
+%  \max\left(1,\frac{T(E_{\text{old}} \rightarrow E_{\text{new}})}{
+%       T(E_{\text{new}} \rightarrow E_{\text{old}})}\right)
+%\end{align}
+%where we note that the transition matrix is updated to reflect the
+%transition under consideration \emph{prior} to computing the acceptance
+%probability.  The TMMC method suffers from poor convergence primarily
+%because it is sensitive to poor statistics in less probable energy
+%transitions.  Like the weight-based methods above, the TMMC method violates
+%detailed balance, with the level of violation decreasing as the simulation
+%progresses and a difference of 1 in the collection matrix has a smaller
+%impact on the acceptance probabilities.
+%
+%\subsection{WL-TMMC Algorithm}
+%Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
+%by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
+%advantage of the WL methods ease of implementation, while making use of
+%a transition matrix to address WL weaknesses, such as its limited
+%statistical accuracy convergence.
+%
+%\paragraph{Initializing Wang-Landau:}
+%We set the energy range along which the method may sample as
+%$E(T_{\min}) <E< E(T=\infty)$. Conventional Wang-Landau prescribes
+%updating the WL factor $\gamma$ by starting at 1.0 and reducing by
+%$\frac12$ until the simulation is cutoff typically around $10^{-10}$.
+%While variants of this approach appear in the literature, Rane et. al.
+%published an implemenation of WL-TMMC~\cite{rane2013monte} that
+%suggested using a WL factor between 1.0 and $10^{-2}$ with a cutoff
+%$<10^{-5}$, We implement WL-TMMC by setting the WL factor to be 1.0 and
+%the cutoff to be either $10^{-4}$ or $10^{-10}$.
+%
+%A single sweep ``roundtrip'' is defined as each macrostate being
+%sampled a number of times in this case 100,000 times. Wang-Landau sets
+%a flatness criteria~\cite{wang2001determining, wang2001efficient,
+%hatch2015computational, mahynski2017predicting} for the accumulated
+%histogram usually to be 0.8.  In the original implementation, it was
+%found to be sufficient that each macrostate is visited at least once
+%(each histogram bin has one entry)~\cite{shell2003improved}; however,
+%for other systems this is rarely sufficient~\cite{bhar2009computer}.
+%The goal of this WL initialization is to prefill the transition matrix
+%for the next portion of the simulation.
+%
+%\paragraph{Initializing TMMC:} Prefilling the transition matrix is
+%necessary because if numerous zeros existed in the collection matrix,
+%the infinite temperature transition matrix would be ill-defined.
+%Sampling over large ranges of density of states would therefore take an
+%unacceptable length of time to complete~\cite{shell2003improved,
+%shen2014elucidating}.  The TMMC portion of the algorithm is run only
+%after WL has reached the specified cutoff. An important note is that
+%the acceptatnce probability for TMMC is only used to control actual
+%transitions not update the collection matrices.
 
@@
 
-As a side note, WL-TMMC does not converge well for $\delta_0 = 0.05\sigma$ while
-it does perform well for $\delta_0 = 0.005\sigma$. The WL portion of the algorithm
-spends nearly 100x longer for the smaller translation scale ensuring that the algorithm
-converges quickly upon transitioning to TMMC.
+%As a side note, WL-TMMC does not converge well for $\delta_0 = 0.05\sigma$ while
+%it does perform well for $\delta_0 = 0.005\sigma$. The WL portion of the algorithm
+%spends nearly 100x longer for the smaller translation scale ensuring that the algorithm
+%converges quickly upon transitioning to TMMC.
 }
@@
 correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
-outperforms WL and WL-TMMC for the same given number of moves. The
+outperforms WL and 1/t-WL? for the same given number of moves. The
 absence of user defined parameters makes SAD an exciting alternative to
@@
 parameters that could lead to the simulation not converging in a
-reasonable amount of time. SAD also converges more quickly than Wl,
-TMMC, and WL-TMMC.
+reasonable amount of time. 
+%SAD also converges more quickly than Wl,
+%TMMC, and WL-TMMC.
 
@@
 
-We gratefully acknowledge D.W. Siderius and
-J.R. Errington for helpful communications regarding the implementation
-of WL-TMMC.
+%We gratefully acknowledge D.W. Siderius and
+%J.R. Errington for helpful communications regarding the implementation
+%of WL-TMMC.
 

### 2019-01-23 papers/histogram/sad.tex

@@
 system~\cite{wang2001determining, wang2001efficient}.  While the method
-is incredibly powerful, it has several user-defined parameters.  This
-can make its application to a variety of systems something of an
-art-form since the user needs to determine the ideal parameters for the
-particular system being studied.  Also, detailed balance is violated
-(although only for short periods of time), ensuring that convergence is
-not guaranteed.  This raises several questions such as when the method
-does converge, what rate will the method converge to the correct
-density of states. Also, how can the user appropriately decide what
-parameters to choose so that the algorithm solves the problem in the
-most ideal way.
+is incredibly powerful, it has several user-defined
+parameters~\cite{landau2004new}.  This can make its application to a
+variety of systems something of an art-form since the user needs to
+determine the ideal parameters for the particular system being studied.
+Also, detailed balance is violated (although only for short periods of
+time), ensuring that convergence is not guaranteed.  This raises
+several questions such as when the method does converge, what rate will
+the method converge to the correct density of states. Also, how can the
+user appropriately decide what parameters to choose so that the
+algorithm solves the problem in the most ideal way.
 
@@
 %study~\cite{siderius2013use}.
-\jpsays{I should look at some more 1/t-WL citations. I know a few more
-that probably should be included here...}
 Around the same time, the convergence rate for WL was being
@@
 
+\subsection{1/t Wang-Landau}
+The Wang-Landau algorithm's error saturation can be corrected by
+modifying the update factor such that it does not decrease too quickly.
+Belardinelli and Pereyra raised the question as to whether the update
+factor should be decrease by $\nicefrac12$ or some other factor for
+each update~\cite{belardinelli2007fast}. They implemented a schedule
+that enforced if the update factor $\gamma$ is less than
+$\nicefrac{1}{t}$ then the update factor is modified by
+$\nicefrac{1}{t}$ and the histogram is no longer tracked. Employing
+this scheduler, they found that the error saturation is avoided since
+the correct density of states is approached asymptotically as
+$\nicefrac{1}{t}$~\cite{belardinelli2008analysis}. Zhou et. al further
+confirmed that the WL algorithm never converges exponentially and
+succesfully bounded the statistical error between
+$\nicefrac{1}{t^{1/2}}$ and $\nicefrac{1}{t}$~\cite{zhou2008optimal}.
+
+Schneider et. al outlines minor refinements to the 1/t WL algorithm
+including $\nicefrac{N_E}{t}$ scaling and switching from standard WL to
+1/t WL when the update factor $\gamma <
+\nicefrac{N_E}{t}$~\cite{schneider2017convergence}. As per the original
+1/t implementation~\cite{belardinelli2007fast}, all energy states $N_E$
+are required to be visited at least once $H(E) \neq 0$ effectively
+avoiding the concept of `flatness'.
+
 \subsection{SAMC}
@@
 we have not explored this degree of freedom.
-
-%This work proposes a means for dynamically setting the update factor
-%$\gamma_t$ that should robustly converge on a variety of systems.
 

### 2019-02-07 papers/fuzzy-fmt/plot-isotherm.py
plot-isotherm.py: added --tar 1 for the tarazona tensor; paper.tex: did some edits; and more data files

@@ -1,4 +1,7 @@
 #!/usr/bin/python2
+#NOTE: Run this plot script from directory deft/papers/fuzzy-fmt 
+#with comand ./plot-isotherm.py --kT [temp] --tar [optional]
+
 from __future__ import print_function, division
 
 import numpy as np
@@ -12,6 +15,9 @@ parser = argparse.ArgumentParser(description='Creates data for a FE vs gw plot.'
 parser.add_argument('--kT', metavar='temperature', type=float,
                     help='reduced temperature - REQUIRED')
 
+parser.add_argument('--tar', metavar='tarazona', type=float,
+                    help='--tar 1 for use tarazona tensor')
+
 args=parser.parse_args()
 
 kT=args.kT
@@ -20,7 +26,10 @@ n = []
 gw = []
 fe_difference = []
 
-files = sorted(list(glob.glob('crystallization/kT%.3f_n*_best.dat' % kT)))
+if args.tar :
+    files = sorted(list(glob.glob('crystallization/kT%.3f_n*_best_tara.dat' % kT)))    
+else :
+    files = sorted(list(glob.glob('crystallization/kT%.3f_n*_best.dat' % kT)))
 for f in files:
     data = np.loadtxt(f)
     n.append(data[1])

### 2019-02-07 papers/histogram/sad.tex

--- 2019-01-23 papers/histogram/sad.tex
@@
 
+\subsection{$1/t$ Wang-Landau}
+The Wang-Landau algorithm's error saturation can be corrected by
+modifying the update factor such that it does not decrease too quickly.
+Belardinelli and Pereyra raised the question as to whether the update
+factor should be decrease by $\nicefrac12$ or some other factor for
+each update~\cite{belardinelli2007fast}. They implemented a schedule
+that enforced if the update factor $\gamma$ is less than
+$\nicefrac{N_S}{t}$ then the update factor is set to
+$\nicefrac{N_S}{t}$ and the histogram is no longer tracked. Employing
+this scheduler, they found that the error saturation is avoided since
+the correct density of states is approached asymptotically as
+$t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al further
+confirmed that the WL algorithm never converges exponentially and
+succesfully bounded the statistical error between
+$t^{-\frac12}$ and $\nicefrac{1}{t}$~\cite{zhou2008optimal}.
+
+Schneider et. al outlines minor refinements to the $1/t$ WL algorithm
+including $\nicefrac{N_E}{t}$ scaling and switching from standard WL to
+$1/t$ WL when the update factor $\gamma <
+\nicefrac{N_E}{t}$~\cite{schneider2017convergence}. As per the original
+$1/t$ implementation~\cite{belardinelli2007fast}, all energy states $N_E$
+are required to be visited at least once $H(E) \neq 0$ effectively
+avoiding the concept of `flatness'.
+
 \subsection{SAMC}

### 2019-02-11 papers/histogram/sad.tex

@@
 
-Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for SAD, WL,
-and SAMC.  The update factor
-for WL remains high for many iterations, and then decreases very rapidly,
-while the SAMC $\gamma_t$ remains constant before dropping as $1/t$.
-The update factor for SAD in contrast fluctuates
-dynamically around a value less than 1 for early MC moves.
-{\color{red} For intermediate MC moves, $\gamma_t$ decreases as
-$1/t$ while continuing to fluctuate whenever a new important energy
-is discovered.  Once all the important energy states are discovered, the
-update factor decreases faster than $1/t$ until asymptoting to $N_S/t$.
-}
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/gamma-n256}
+  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for
+    four different methods: WL, $1/t$-WL, SAMC, and SAD}
+    \label{fig:gamma-vs-t}
+\end{figure}
+Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related methods
+SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains constant
+before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for many iterations,
+and then decreases very rapidly, with $1/t$-WL behaving similarly before
+transitioning to a $1/t$ behavior.
+The update factor for SAD fluctuates
+dynamically around a value less than 1 for early MC moves, and then
+decreases as approximately $1/t$ while continuing to fluctuate as 
+new energies are found to be important.  At late times, the SAD $\gamma_t$
+decreases as approximately $1/t^2$ until asymptoting to $N_S/t$, which
+is the same as $1/t$-WL.
 
@@
 %\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n256}
-  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for three
-    different methods: WL, SAMC, and SAD.
-    method.}\label{fig:gamma-vs-t}
-\end{figure}
 \begin{figure}

### 2019-02-11 papers/histogram/sad.tex
minor edits to sad paper

@@
         for the change in entropy.  The vertical dotted lines represent
-        the energy corresponding to $T=0.\overline{3}$ and $T=\infty$.}
+        the energy corresponding to $T=1/3$ and $T=\infty$.}
+  \label{fig:entropy-cartoon}
 \end{figure}
@@
 to the lowest entropy at $T_{\min}$).  If we assume a quadratic
-dependence on energy, this is given by
+dependence on energy (see Fig.~\ref{fig:entropy-cartoon}), this is given by
 \begin{align}
@@
 
+Our smaller system consists of 50 square-well atoms with a well width
+of $\lambda=1.3$, in a volume corresponding to a filling fraction of
+$\eta=0.3$.  For this simulation, we chose a minimum reduced
+temperature of $1/3$, which corresponds to an interesting energy range
+from -248 to -120.  The entropy of this system is shown in
+Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
+energy range the entropy differs by 198.
+
 Figures~\ref{fig:N50-ff0.3-avg-error}
-and~\ref{fig:N50-ff0.3-avg-error-slow} show the average entropy vs
+and~\ref{fig:N50-ff0.3-avg-error-slow} show the average error in the entropy vs
 moves for a system consisting of 50 atoms that are placed in a periodic
@@
 
-{\color{red}
+The interesting range of energies for this system is from
+{\color{blue}
 $S_{\max} = 120$ and $S_{\min} = 248$ for this system.  The number of

### 2019-02-11 papers/histogram/sad.tex
edits to paper

@@
 reference.}
+
+\davidsays{Add this info to the section above.}  The SAMC simulations
+computed the density of states for the entire range of possible
+energies.  The SAD simulations determined the energy range of interest
+dynamically as described above, based on a specified $T_{\min}$.  For
+the WL and $1/t$-WL simulations, we constrained the simulation to
+remain in the energy range corresponding to $T_{\min} < T < \infty$,
+as determined by a previous SAMC simulation.  Thus the WL and $1/t$-WL
+simulations were given extra information that in practice would not be
+available without additional computational effort.
+
+\subsection{A periodic system with 50 atoms}
+
+\davidsays{The following paragraph duplicates some of what is written
+  in the previous.  Please move things up as seems best!}
+
+Our smaller system consists of 50 square-well atoms with a well width
+of $\lambda=1.3$, in a volume corresponding to a filling fraction of
+$\eta=0.3$.  For this simulation, we chose a minimum reduced
+temperature of $1/3$, which corresponds to an interesting energy range
+from -248 to -120.  The entropy of this system is shown in
+Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
+energy range the entropy differs by 198, corresponding to a ratio of
+$10^{86}$ between the highest and lowest density of states.
+
+In order to explore the effect of simulation details on convergence,
+we consider two values for the mean distance by which atoms are moved
+during a Monte Carlo step.  We began with a reasonable displacement
+distance of $\delta_0 = 0.05\sigma$, which corresponds to an
+acceptance rate of 38\%.  We further ran simulations with a much
+smaller displacement distance of $\delta_0 = 0.005\sigma$, which
+resulted in an acceptance rate of 86\%, which we expected converge
+more slowly.
+
 \begin{figure}
@@
 
-\subsection{A periodic system with 50 atoms}
+Figure~\ref{fig:N50-ff0.3-avg-error} shows the average error in the
+entropy as a function of time for this system with the reasonable
+displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
+represent the average of the absolute value of the error in the
+entropy averaged over four simulations using different random number
+seeds.  The range of average errors for each simulation is shown as a
+shaded region around its mean error.  By the time $10^8$ moves have
+been made all but the SAMC simulation with the shortest $t_0$ have
+begun to converge as $1/\sqrt{t}$.  We then see the WL error begin to
+saturate around $10^{10}$ moves.
 
-Our smaller system consists of 50 square-well atoms with a well width
-of $\lambda=1.3$, in a volume corresponding to a filling fraction of
-$\eta=0.3$.  For this simulation, we chose a minimum reduced
-temperature of $1/3$, which corresponds to an interesting energy range
-from -248 to -120.  The entropy of this system is shown in
-Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
-energy range the entropy differs by 198.
-
-Figures~\ref{fig:N50-ff0.3-avg-error}
-and~\ref{fig:N50-ff0.3-avg-error-slow} show the average error in the entropy vs
-moves for a system consisting of 50 atoms that are placed in a periodic
-simple cubic unit cell with lattice constant chosen to give a filling
-fraction of $\eta = 0.3$.  These particles are randomly inserted into
-the system and random moves are proposed (with a translation scale
-$\delta_0 = 0.05\sigma$ and $\delta_0 = 0.005\sigma$ respectively). For
-the $\eta = 30\%$ and $N = 50$ system, we explore the choice of
-$\delta_0$ as a parameter.
-
-The interesting range of energies for this system is from
-{\color{blue}
-$S_{\max} = 120$ and $S_{\min} = 248$ for this system.  The number of
-important energy states for this system is then $N_S = S_{\min} - S_{\max} = 145$.
-The reduced temperature of $1/3$ is chosen in order to examine the convergence
-properties of SAD in a reasonable simulation time.
-
-%As a side note, WL-TMMC does not converge well for $\delta_0 = 0.05\sigma$ while
-%it does perform well for $\delta_0 = 0.005\sigma$. The WL portion of the algorithm
-%spends nearly 100x longer for the smaller translation scale ensuring that the algorithm
-%converges quickly upon transitioning to TMMC.
-}
-
-Smaller translation
-scales cause the given method to take additional time to explore all
-energies. Ideal methods should scale linearly with the $\delta_0$, that
-is, a decrease in the translation scale should result in a proportional
-increase in convergence time. As can be seen from
-Fig.~\ref{fig:N50-ff0.3-avg-error-slow}, SAMC methods that rapidly
-converged for $\delta = 0.05\sigma$ do not proportionaly converge for a
-translation scale of $\delta = 0.005\sigma$. We see that SAD handles
-arbitrary translation scales and proportionaly converges based on the
-translation scale that is chosen.
+Figure~\ref{fig:N50-ff0.3-avg-error-slow} shows the average error in
+the entropy as a function of time for this system with the
+unreasonably small displacement distance of $\delta_0 = 0.005\sigma$.
+%
+The smaller translation scale causes all methods to take additional
+time to explore all energies. Based on random walk scaling, a ideal
+method should scale roughly as $\delta_0^{-2}$ in the limit of small
+$\delta_0$, that is, one order of magnitude in the displacement
+distance should result in two order of magnitude increase in
+convergence time. SAMC simulations with a $t_0$ value that rapidly
+converged for $\delta = 0.05\sigma$ do not converge at all for a
+translation scale of $\delta = 0.005\sigma$. SAD, WL, and $1/t$-WL
+handle the shift in displacement distance and converge roughly as
+expected.
 

### 2019-02-11 papers/histogram/sad.tex
a few more davidsays in the paper

@@
   \includegraphics[width=\columnwidth]{figs/gamma-n256}
-  \caption{The update factor $\gamma_t$ for $N=256$ versus iteration for
-    four different methods: WL, $1/t$-WL, SAMC, and SAD}
+  \caption{The update factor $\gamma_t$ for $N=256$ versus
+    iteration for four different methods: WL, $1/t$-WL, SAMC, and SAD}
     \label{fig:gamma-vs-t}
@@
 %\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n50}
-  \caption{The update factor $\gamma_t$ for $N=50$ versus iteration for three
-    different methods: WL, SAMC, and SAD.}
-\end{figure}
-
 \paragraph{Energy binning with SAD}
@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/gamma-n50}
+  \caption{\davidsays{We need to make a distinction between SAD and
+      SAD both in figure and in caption.  I don't know what to call
+      the slow SAD.  Also, should this move down to results, and do we
+      want slow WL and slow-1/t WL?} The update factor $\gamma_t$ for
+    $N=50$ versus iteration for three different methods: WL, SAMC, and
+    SAD.}\label{fig:fast-slow-gamma}
+\end{figure}
+
 Figure~\ref{fig:N50-ff0.3-avg-error-slow} shows the average error in
@@
 
+To see how SAD, WL, and $1/t$-WL compensate for the smaller
+displacement distance, \davidsays{see Fig.~\ref{fig:fast-slow-gamma}
+  and discuss this}
+
 \begin{figure}
@@
   \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
-  \caption{Use this to create description of the system.}\label{fig:n256-lndos-comparison}
+  \caption{\davidsays{This could be prettier, and do we need it? We
+      definitely need a caption if we keep it.} Use this to create
+    description of the system.}\label{fig:n256-lndos-comparison}
 \end{figure}
@@
 \subsection{A periodic system with 256 atoms}
+
+\davidsays{Make this section match the previous one.}
 

### 2019-02-13 papers/histogram/sad.tex
made some paper additions.

@@
 
-%Transition Matrix Monte Carlo~\cite{wang1999transition,
-%swendsen1999transition, fitzgerald2000monte} became an attractive
-%complimentary simulation algorithm to Wang-Landau, in that, if detailed
-%balance is ensured, the system is guaranteed to converge.  Unfortunately,
-%the algorithm can be considerably more difficult to implement when
-%compared to Wang-Landau: in particular, due to the infinite temperature
-%transition matrix~\cite{wang2002transition}.  Also, the time necessary
-%for the density of states to converge could be considerable
-%due to the algorithm spending too much time exploring low energy states.
-%Unlike WL, TMMC does not require a prior knowledge of the energy range of
-%the system.
-%
-%Shell et al.~\cite{shell2003improved, shell2004flat} originially
-%implemented Wang-Landau Transition Matrix Monte Carlo (WL-TMMC) in an
-%effort to quickly explore the energies of a given system using WL then
-%switch over to TMMC to guarantee convergence. Considerable effort has
-%been spent trying to determine the flatness criteria and cutoff for the
-%WL portion, but there remains no rigorous systematic way to determine
-%these parameters~\cite{rane2013monte}.  The algorithm is first run for
-%several trial simulations and/or `experience' is used to determine the
-%best parameters for a users particular system of
-%study~\cite{siderius2013use}.
 Around the same time, the convergence rate for WL was being
@@
 section. We compare it along with several flat histogram methods
-which include WL, 1/t-WL, and SAMC.
-%which include TMMC, WL, WL-TMMC, and SAMC.
+which include WL, $1/t$-WL, and SAMC.
 
@@
 weight-based histogram methods used in this work.
-%Section~\ref{sec:transition}
-%details the workings of the transition-matrix methods studied in this work.
 
@@
 works.  The following methods are introduced and applied to the
-square-well fluid: Wang-Landau (WL), 1/t-Wang-Landau (1/t-WL), Stochastic
+square-well fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
 Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation (SAD).
-
-%Transition Matrix Monte-Carlo (TMMC), Wang-Landau
-%(WL), Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC), Stochastic
-%Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation
-%(SAD).
 
@@
 
-\subsection{$1/t$ Wang-Landau}
+\subsection{$1/t$-Wang-Landau}
 The Wang-Landau algorithm's error saturation can be corrected by
@@
 succesfully bounded the statistical error between
-$t^{-\frac12}$ and $\nicefrac{1}{t}$~\cite{zhou2008optimal}.
+$t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
 
-Schneider et. al outlines minor refinements to the $1/t$ WL algorithm
+Schneider et. al outlines minor refinements to the $1/t$-WL algorithm
 including $\nicefrac{N_E}{t}$ scaling and switching from standard WL to
-$1/t$ WL when the update factor $\gamma <
+$1/t$-WL when the update factor $\gamma <
 \nicefrac{N_E}{t}$~\cite{schneider2017convergence}. As per the original
@@
 
-%\section{Transition-matrix methods}\label{sec:transition}
-%
-%\subsection{TMMC Algorithm}
-%The previously described weight-based flat-histogram methods utilizes a so-called
-%``visited states'' approach to calculate  thermodynamic
-%properties~\cite{errington2003direct}. The algorithm is fundamentally
-%based on the number of times a given state is visited while the
-%system is being sampled. The Transition Matrix Monte Carlo (TMMC) method operates
-%differently than the visited states approach~\cite{fitzgerald2000monte}.
-%The method uses attempted transitions between states to obtain an estimate of
-%the probability that a given transition will be proposed.  The density
-%of states is computed from the resulting transition matrix.  An advantage
-%of collecting the transition matrix is that this matrix may be collected
-%while another method is employed to compute acceptance probabilities, as
-%in Wang-Landau TMMC, discussed below.
-%
-%The TMMC algorithm records the transition data in a \emph{collection matrix},
-%i.e. the unnormalized transition histogram matrix, which stores the number
-%of times a transition between two energy states has been proposed.
-%\begin{align}
-%  C(E_{\text{old}} \rightarrow E_{\text{new}},t+1)
-%     = C(E_{\text{old}} \rightarrow E_{\text{new}},t) + 1
-%\end{align}
-%The transition matrix is computed from the collection matrix via
-%\begin{align}
-%  T(E_{\text{old}} \rightarrow E_{\text{new}})
-%   = \frac{C(E_{\text{old}} \rightarrow E_{\text{new}})}{
-%     \sum_E C(E_{\text{old}} \rightarrow E)
-%  }
-%\end{align}
-%The acceptance probability is defined in terms of this biasing function:
-%\begin{align}
-%  \mathcal{P}(E_{\text{old}} \rightarrow E_{\text{new}}) =
-%  \max\left(1,\frac{T(E_{\text{old}} \rightarrow E_{\text{new}})}{
-%       T(E_{\text{new}} \rightarrow E_{\text{old}})}\right)
-%\end{align}
-%where we note that the transition matrix is updated to reflect the
-%transition under consideration \emph{prior} to computing the acceptance
-%probability.  The TMMC method suffers from poor convergence primarily
-%because it is sensitive to poor statistics in less probable energy
-%transitions.  Like the weight-based methods above, the TMMC method violates
-%detailed balance, with the level of violation decreasing as the simulation
-%progresses and a difference of 1 in the collection matrix has a smaller
-%impact on the acceptance probabilities.
-%
-%\subsection{WL-TMMC Algorithm}
-%Wang-Landau Transition Matrix Monte-Carlo (WL-TMMC) method as proposed
-%by Shell and coworkers~\cite{shell2003improved,shell2004flat} takes
-%advantage of the WL methods ease of implementation, while making use of
-%a transition matrix to address WL weaknesses, such as its limited
-%statistical accuracy convergence.
-%
-%\paragraph{Initializing Wang-Landau:}
-%We set the energy range along which the method may sample as
-%$E(T_{\min}) <E< E(T=\infty)$. Conventional Wang-Landau prescribes
-%updating the WL factor $\gamma$ by starting at 1.0 and reducing by
-%$\frac12$ until the simulation is cutoff typically around $10^{-10}$.
-%While variants of this approach appear in the literature, Rane et. al.
-%published an implemenation of WL-TMMC~\cite{rane2013monte} that
-%suggested using a WL factor between 1.0 and $10^{-2}$ with a cutoff
-%$<10^{-5}$, We implement WL-TMMC by setting the WL factor to be 1.0 and
-%the cutoff to be either $10^{-4}$ or $10^{-10}$.
-%
-%A single sweep ``roundtrip'' is defined as each macrostate being
-%sampled a number of times in this case 100,000 times. Wang-Landau sets
-%a flatness criteria~\cite{wang2001determining, wang2001efficient,
-%hatch2015computational, mahynski2017predicting} for the accumulated
-%histogram usually to be 0.8.  In the original implementation, it was
-%found to be sufficient that each macrostate is visited at least once
-%(each histogram bin has one entry)~\cite{shell2003improved}; however,
-%for other systems this is rarely sufficient~\cite{bhar2009computer}.
-%The goal of this WL initialization is to prefill the transition matrix
-%for the next portion of the simulation.
-%
-%\paragraph{Initializing TMMC:} Prefilling the transition matrix is
-%necessary because if numerous zeros existed in the collection matrix,
-%the infinite temperature transition matrix would be ill-defined.
-%Sampling over large ranges of density of states would therefore take an
-%unacceptable length of time to complete~\cite{shell2003improved,
-%shen2014elucidating}.  The TMMC portion of the algorithm is run only
-%after WL has reached the specified cutoff. An important note is that
-%the acceptatnce probability for TMMC is only used to control actual
-%transitions not update the collection matrices.
-
 \section{SAD Algorithm}\label{sec:sad}
@@
         the energy corresponding to $T=1/3$ and $T=\infty$.}
+        ~\jpsays{I fixed this to use samc-1e4 reference. Remove comment}
   \label{fig:entropy-cartoon}
@@
 
-%\begin{figure}
-  %\includegraphics[width=\columnwidth]{figs/gamma-n500}
-  %\caption{The update factor $\gamma_t$ for $N=500$ versus iteration for three
-    %different methods: WL, SAMC, and SAD.
-    %method.}\label{fig:gamma-vs-t}
-%\end{figure}
 \paragraph{Energy binning with SAD}
@@
 
-For the square-well fluid, our simulations treat systems with a
-well-width of $\lambda = 1.3$ and $\lambda = 1.5$ and a filling fraction of
-$\eta = 0.3$ and $\eta = 0.17$. We also explore various translation scales $\delta_0 = 0.005\sigma$ and
-$0.05\sigma$. The simulations explore the
+~\jpsays{I edited this section}
+For the square-well fluid, our simulations treat systems with a particle number
+of $50$ and $256$, a well-width of $\lambda = 1.3$ and $\lambda = 1.5$, and a volume
+corresponding to a
+filling fraction of $\eta = 0.3$ and $\eta = 0.17$ respectively. The simulations
+also explore a reasonable displacement distance $\delta_0 = 0.05\sigma$ and
+an unreasonably small displacement distance of $0.005\sigma$. The simulations explore the
 energy space of the systems at reduced temperatures of
-$T_{\text{min}} = 0.\overline{3}$ and $1.0$.  All simulations lead to the minimum important
+$T_{\text{min}} = 1/3$ and $1.0$.  All simulations lead to the minimum important
 energy $E_{\min}$ and maximum entropy energy $E_{\max}$
@@
 
-We use the average entropy error vs moves as a metric to compare
-simulation runtimes and overall convergence. The overall accuracy
-is determined by examining the fractional error of a particular method to
-a precise reference system. For each simulation, the reference system
-is chosen to be the final output of the SAMC simulation with the best
-choice of $t_0$.
-\begin{align}
-\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{ref}}\bigg\rvert
-\end{align}
-Simulations that take more iterations to reach the same entropy error as a
-different method take longer to converge. This behavior is seen in
-Fig.~\ref{fig:N50-ff0.3-avg-error}.
-\jpsays{We use a SAMC $t_0 = 10^3$ with a fixed energy range $S_{\max} = 120$ and $S_{\min} = 248$ as the reference.
-This is noted because this is not strictly SAMC method. We use this as a reference to increase convergence time of the
-reference.}
-
-\davidsays{Add this info to the section above.}  The SAMC simulations
+The SAMC simulations
 computed the density of states for the entire range of possible
@@
 
+We use the average entropy error vs moves as a metric to compare
+simulation runtimes and overall convergence. The overall accuracy
+is determined by examining the fractional error of a particular method to
+a precise reference system. For each simulation, the reference system
+is chosen to be the final output of a SAMC simulation with a fixed interesting energy range.
+We note this because SAMC does not require an energy range as an input parameter; however,
+by fixing the energy range to for SAMC, a smaller $t_0$ can be chosen considerably reducing the
+simulation time.
+\begin{align}
+\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{ref}}\bigg\rvert
+\end{align}
+Simulations that take more iterations to reach the same entropy error as a
+different method take longer to converge. This behavior is seen in
+Fig.~\ref{fig:N50-ff0.3-avg-error}.
+
 \subsection{A periodic system with 50 atoms}
 
-\davidsays{The following paragraph duplicates some of what is written
-  in the previous.  Please move things up as seems best!}
-
-Our smaller system consists of 50 square-well atoms with a well width
-of $\lambda=1.3$, in a volume corresponding to a filling fraction of
-$\eta=0.3$.  For this simulation, we chose a minimum reduced
+For this simulation, we chose a minimum reduced
 temperature of $1/3$, which corresponds to an interesting energy range
-from -248 to -120.  The entropy of this system is shown in
+from $-248$ to $-120$.  The entropy of this system is shown in
 Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
@@
 represent the average of the absolute value of the error in the
-entropy averaged over four simulations using different random number
+entropy averaged over eight simulations using different random number
 seeds.  The range of average errors for each simulation is shown as a
@@
 %\begin{figure}
-%\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N500-entropy-error-default.pdf}
-  %\caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/2$ with
-  %SAMC $t_0 = 10^6$ serving as the reference.}\label{fig:n500-ff0.3}
+%  \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
+%  \caption{\davidsays{This could be prettier, and do we need it? We
+%      definitely need a caption if we keep it.} Use this to create
+%    description of the system.}\label{fig:n256-lndos-comparison}
 %\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
-  \caption{\davidsays{This could be prettier, and do we need it? We
-      definitely need a caption if we keep it.} Use this to create
-    description of the system.}\label{fig:n256-lndos-comparison}
-\end{figure}
 
 \subsection{A periodic system with 256 atoms}
+For this simulation, we chose a minimum reduced
+temperature of $1.0$, which corresponds to an interesting energy range
+from $-915$ to $-509$.  The number of important energy states for this
+system is then $N_S = S_{\min} - S_{\max} = 406$.
+The entropy of this system is has an entropy that
+differs by 275, corresponding to a ratio of
+$10^{119}$ between the highest and lowest density of states.
 
-\davidsays{Make this section match the previous one.}
-
-Figure~\ref{fig:n256} shows the average entropy vs moves for a system
-consisting of 256 atoms that are placed in a periodic simple cubic unit
-cell with lattice constant chosen to give a filling fraction of $\eta =
-0.17$.   These particles are randomly inserted into the system and
-random moves are proposed (with a translation scale $\delta_0 =
-0.05\sigma$).
-
-{\color{red}
-$S_{\max} = 509$ and $S_{\min} = 915$ for this system.  The number of
-important energy states for this system is then $N_S = S_{\min} -
-S_{\max} = 406$. The reduced temperature of $1.0$ and well-width of $\lambda = 1.5$
-are chosen in order to ensure that the simulation encounters liquid vapor coexistence
-and that the results are comparable with NIST.
-}
+Figure~\ref{fig:N256} shows the average error in the
+entropy as a function of time for this system with the reasonable
+displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
+represent the average of the absolute value of the error in the
+entropy averaged over eight simulations using different random number
+seeds.  The range of average errors for each simulation is shown as a
+shaded region around its mean error.  By the time $10^8$ moves have
+been made all but the SAMC simulation with the shortest $t_0$ have
+begun to converge as $1/\sqrt{t}$.  We then see the WL error begin to
+saturate around $10^{10}$ moves.
 
@@
 reasonable amount of time.
-%SAD also converges more quickly than Wl,
-%TMMC, and WL-TMMC.
 
@@
 
-% \jpsays{Everything below this is TMI1 and GC-TMMC}
-
-%\subsection{GC-TMMC Algorithm}
-%Grand Canonical Transition Matrix Monte-Carlo (GC-TMMC) allows the energy and
-%particle number of the system to fluctuate while holding the chemical potential,
-%volume, and temperature constant~\cite{chen2001aggregation, chen2002simulating,
-%errington2003direct, fenwick2006accurate, fenwick2008direct, errington2005direct}.
-%The algorithm boasts a lower statistical noise than conventional GCMC due to its
-%incorporation of a transition matrix to estimate ensemble averages~\cite{siderius2013use,
-%paluch2008comparing, grzelak2008computation}.  The steps for the algorithm are
-%outlined below:
-
-%\subsubsection{Define particle insertion/deletion}
-%In the system of consideration, the number of particles $\mathcal{N}$ can change
-%by an amount $\delta$.  It is assumed that in the system either no particles can be
-%inserted or deleted, a single insertion takes place, or a single deletion takes place.
-%This yields a condition $\delta={-1,0,+1}$ for particle insertion.
-
-%\subsubsection{Define an acceptance probability}
-%Now it is necessary to define criteria for determining whether a particle insertion/deletion
-%is allowed.  This criteria is called the acceptance probability and is denoted as
-%\begin{align}
-  %p_{a} = \min\bigg[1,\frac{\alpha_{new\rightarrow old}}
-  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
-%\end{align}
-%where $\alpha_{new\rightarrow old}$ is the probability of generating an old system
-%configuration from the new.  For conventional Monte-Carlo moves $\alpha_{old \rightarrow new}
-%=\alpha_{new\rightarrow old}$ although for advanced Monte-Carlo moves this would not
-%be the case~\cite{siepmann1990method}.  The probabilies of actually observing the
-%system in the old state is given by
-%\begin{align}
-  %\pi_{old} = \frac{1}{\Xi}\frac{V^{N_{old}}}{\Lambda^{3N_{old}}N_{old}!}
-  %e^{-\beta(\epsilon_{old}-\mu N_{old})}
-%\end{align}
-%where $\Xi$ is the grand canonical partition function and $\Lambda$ is the De Broglie
-%wavelength.
-
-%\subsubsection{Define a Collection Matrix}
-%In order to calculate an overall transition probability $\mathcal{P}_{N,\delta}$, we
-%must first define how the system procession is to be recorded.  A Collection Matrix
-%can be defined, such that for any one move, two elements are updated.
-%\begin{align}
-  %{C}_{N,\delta} = {C}_{N,\delta} + p_{a}\\
-  %{C}_{N,0} = {C}_{N,0} +(1 - p_{a})
-%\end{align}
-%If $\delta=0$, that is, the number of particles has not been changed, then only
-%$\mathcal{C}_{N,0}$ is incremented by unity.  The overal transition probability can
-%now be defined as
-%\begin{align}
-  %\mathcal{P}_{N,\delta} = \frac{{C}_{N,\delta}}
-  %{\sum_{j=-1}^{+1} {C}_{N,j}}
-%\end{align}
-
-%\subsubsection{Define the Particle Number Probability Distribution}
-%The Particle Number Probability Distribution (PNPD) $\Pi(N;\mu,V,T)$ can be found
-%by enforcing detailed balance in the system.
-%\begin{align}
-  %\ln(\Pi_{N+1}) = \ln(\Pi_{N}) + \ln\bigg(\frac{\mathcal{P}_{N\rightarrow N+1}}
-  %{\mathcal{P}_{N+1\rightarrow N}}\bigg)
-%\end{align}
-%This equation will hold as long as N-space can be sampled sufficiently.  In an
-%implemented algorithm $N$, $N_{min}$, and $N_{max}$ are specified for macrostate domain
-%sampling.
-
-%\subsubsection{Define a biasing function}
-%The primary difference between TMMC as defined by Fitzgerald et al. and GC-TMMC as
-%pioneered by JR Errington et al. is the use of a biasing function to control the acceptance
-%rate of trial moves~\cite{siderius2013use, fitzgerald2000monte}.  This type of
-%multi-canonical sampling augments the acceptance probability in the following way
-%\begin{align}
-  %p_{\eta,old\rightarrow new} = \min\bigg[1,\frac{e^{\eta(N_{new})}}
-  %{e^{\eta(N_{old})}}\frac{\alpha_{new\rightarrow old}}
-  %{\alpha_{old \rightarrow new}}\frac{\pi_{new}}{\pi_{old}}\bigg]
-%\end{align}
-
-%Although $p_{\eta}$ is used to control the actual transitions for trial configurations
-%in GC-TMMC, $p_{a}$ is still used to update the Collection Matrices.  Ideally the biasing
-%function would be given by
-%\begin{align}
-  %\eta(N) = -\ln\Pi(N;\mu,V,T)
-%\end{align}
-%Since $\Pi$ is not usually known a priori, the biasing function is usually set to some
-%arbitrary constant for initialization.  The goal of this biasing function is to achieve
-%uniform sampling (a common feature shared by the flat histogram Monte-Carlo family).
-
-%\subsubsection{Histogram reweighting for the PNPD}
-%Once the PNPD has beeen collected for a given chemical potential $\mu_0$, histogram
-%reweighting is used to determine thermodynamic properties at other chemic potentials.
-%Since the canonical partition function $Z(N,V,T)$ does not depend on the chemical
-%potential, we can write
-%\begin{align}
-%\begin{split}
-  %\Pi(N;\mu,V,T) = \frac{e^{\beta\mu N}Z(N,V,T)}{\Xi(\mu,V,T)}\\
-  %\ln\Pi(N;\mu,V,T) = \ln\Pi(N;\mu_0,V,T) \\
-  %+ \beta N(\mu-\mu_0) + C
-%\end{split}
-%\end{align}
-%where $C$ is a normalization constant independent of the number of particles.
-
 \bibliography{paper}% Produces the bibliography via BibTeX.

### 2019-02-25 papers/histogram/sad.tex
paper additions and figure formatting.

@@
         the energy corresponding to $T=1/3$ and $T=\infty$.}
-        ~\jpsays{I fixed this to use samc-1e4 reference. Remove comment}
+        ~\jpsays{I fixed this to use samc-1e4-s1. Remove comment}
   \label{fig:entropy-cartoon}
@@
 
+\begin{figure*}
+  \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
+\includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
+  \caption{(a) The average entropy error for each MC method is shown for $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
+               a fixed energy range reference as discussed in the text. The update factor $\gamma_t$ for
+               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.
+           (b) The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
+               the same fixed energy range reference used in (a). The update factor $\gamma_t$ for
+               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.}\label{fig:fast-slow-gamma}
+\end{figure*}
+
 ~\jpsays{I edited this section}
@@
 different method take longer to converge. This behavior is seen in
-Fig.~\ref{fig:N50-ff0.3-avg-error}.
+Fig.~\ref{fig:fast-slow-gamma}.
 
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-  a fixed energy range SAMC $t_0 = 10^3$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
-\end{figure}
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-  a fixed energy range SAMC $t_0 = 10^3$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
-  serving as the reference.}\label{fig:N50-ff0.3-avg-error-slow}
-\end{figure}
+%\begin{figure}
+%  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
+%  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
+%  a fixed energy range SAMC $t_0 = 10^3$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
+%\end{figure}
+%\begin{figure}
+%  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
+%  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
+%  a fixed energy range SAMC $t_0 = 10^3$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
+%  serving as the reference.}\label{fig:N50-ff0.3-avg-error-slow}
+%\end{figure}
 
-Figure~\ref{fig:N50-ff0.3-avg-error} shows the average error in the
+Figure~\ref{fig:fast-slow-gamma}a shows the average error in the
 entropy as a function of time for this system with the reasonable
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n50}
-  \caption{\davidsays{We need to make a distinction between SAD and
-      SAD both in figure and in caption.  I don't know what to call
-      the slow SAD.  Also, should this move down to results, and do we
-      want slow WL and slow-1/t WL?} The update factor $\gamma_t$ for
-    $N=50$ versus iteration for three different methods: WL, SAMC, and
-    SAD.}\label{fig:fast-slow-gamma}
-\end{figure}
-
-Figure~\ref{fig:N50-ff0.3-avg-error-slow} shows the average error in
+Figure~\ref{fig:fast-slow-gamma}b shows the average error in
 the entropy as a function of time for this system with the
@@
 
-To see how SAD, WL, and $1/t$-WL compensate for the smaller
-displacement distance, \davidsays{see Fig.~\ref{fig:fast-slow-gamma}
-  and discuss this}
+The methods SAD, WL, and $1/t$-WL compensate for the smaller
+displacement distance as can be seen from Figure~\ref{fig:fast-slow-gamma}.
+The update factors take approximately 10x longer to reach steady-state for
+the slower translation scale. Because of this update behavior, these methods
+do not suffer in converging to the correct density of states from having a
+static parameter such as SAMC $t_0$ in the update factor. SAMC methods that converge
+in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not 
+necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
+
+Sad performs similarly regardless of the displacement distance chosen although
+it transitions to $1/\sqrt{t}$ convergence after
 
@@
 from $-915$ to $-509$.  The number of important energy states for this
-system is then $N_S = S_{\min} - S_{\max} = 406$.
+system is then $N_S = E_{\min} - E_{\max} = 406$.
 The entropy of this system is has an entropy that
@@
 
-Figure~\ref{fig:N256} shows the average error in the
-entropy as a function of time for this system with the reasonable
+Figure~\ref{fig:n256} shows the average error in the
+entropy as a function of moves for this system with the reasonable
 displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
@@
 
-{\color{red}
-Figure~\ref{fig:n256-lndos-comparison} shows the calculated entropy for
-the $N = 256$ system. The grey dash lines indicate the $S_{\max} = 509$
-point and the $S_{\min} = 915$. Only the reference method is shown.
-}
+%{\color{red}
+%Figure~\ref{fig:n256-lndos-comparison} shows the calculated entropy for
+%the $N = 256$ system. The grey dash lines indicate the $S_{\max} = 509$
+%point and the $S_{\min} = 915$. Only the reference method is shown.
+%}
 

### 2019-03-04 papers/histogram/sad.tex
edits during meeting

@@
         the energy corresponding to $T=1/3$ and $T=\infty$.}
-        ~\jpsays{I fixed this to use samc-1e4-s1. Remove comment}
   \label{fig:entropy-cartoon}
@@
 \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
-  \caption{(a) The average entropy error for each MC method is shown for $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
-               a fixed energy range reference as discussed in the text. The update factor $\gamma_t$ for
-               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.
-           (b) The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
+  \caption{(a) The entropy error for each MC method for $N=50$, 
+               $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
+               as a function of number of iterations run.  The error is
+               averaged over 8 independent simulations, and the best
+               and worst simulations for each method are shown as a
+               semi-transparent shaded area.
+              (b) The update factor $\gamma_t$ versus iteration number
+               for the same simulations.
+           (c) \davidsays{Make this similar} The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
                the same fixed energy range reference used in (a). The update factor $\gamma_t$ for
@@
 
-~\jpsays{I edited this section}
-For the square-well fluid, our simulations treat systems with a particle number
-of $50$ and $256$, a well-width of $\lambda = 1.3$ and $\lambda = 1.5$, and a volume
-corresponding to a
-filling fraction of $\eta = 0.3$ and $\eta = 0.17$ respectively. The simulations
-also explore a reasonable displacement distance $\delta_0 = 0.05\sigma$ and
-an unreasonably small displacement distance of $0.005\sigma$. The simulations explore the
-energy space of the systems at reduced temperatures of
-$T_{\text{min}} = 1/3$ and $1.0$.  All simulations lead to the minimum important
+For the square-well fluid, our simulations treat two systems.  The first is
+a smaller simulation with a particle number
+of 50, a well-width of $\lambda = 1.3$, and a volume
+corresponding to a filling fraction of $\eta = 0.3$. The second system
+is larger, with a particle number
+of 256, a well-width of $\lambda = 1.5$, and a volume
+corresponding to a filling fraction of $\eta = 0.17$.
+For each system we use a reasonable displacement distance $\delta_0 = 0.05\sigma$,
+and for the smaller system we also use an unreasonably small displacement
+distance of $0.005\sigma$. The simulations explore the
+energy space of the systems with minimum reduced temperatures of
+$T_{\text{min}} = 1/3$ for simulations of the smaller system, and $T_{\min}=1$
+for the larger system.
+All simulations lead to the minimum important
 energy $E_{\min}$ and maximum entropy energy $E_{\max}$
@@
 simulations were given extra information that in practice would not be
-available without additional computational effort.
+available without additional computational effort, and the SAMC simulations
+computed the entropy over the entire range of possible energies, which
+required more effort.
 

### 2019-03-04 papers/histogram/sad.tex
made some paper changes.

@@
 \begin{abstract}
-  We present a novel Monte-Carlo algorithm based on Stochastic
+  We present a novel Monte-Carlo algorithm based on the Stochastic
   Approximation Monte Carlo (SAMC) algorithm for directly calculating
@@
   Approximation with a dynamic update factor (SAD-$\gamma$ or SAD)
-  dynamically adjusts the update factor $\gamma$ during the course of
+  which dynamically adjusts the update factor $\gamma$ during the course of
   the simulation. We test this MC method on the square-well fluid and
-  compare the convergence time and entropy error for a variety of weight-based
-  and transition matrix Monte-Carlo methods. SAD rapidly converges to the
+  compare the convergence time and average entropy error for a variety of weight-based
+  Monte-Carlo methods. SAD rapidly converges to the
   correct density of states without the need for the user to specify an
@@
 time), ensuring that convergence is not guaranteed.  This raises
-several questions such as when the method does converge, what rate will
-the method converge to the correct density of states. Also, how can the
+several questions such as when the method does converge, at what rate will
+the method converge to the correct density of states. Also, can the
 user appropriately decide what parameters to choose so that the
@@
 does not require user-defined inputs and therefore should be easily
-applicable to a given system.  We call this method SAD (Dynamic
-Stochastic Approximation), and will discuss it in detail in the methods
+applicable to a given system.  We call this method SAD (Stochastic Approximation
+with a dynamic update factor), and will discuss it in detail in the methods
 section. We compare it along with several flat histogram methods
@@
 square-well fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
-Approximation Monte-Carlo (SAMC), and Dynamic Stochastic Approximation (SAD).
+Approximation Monte-Carlo (SAMC), and Stochastic Approximation with a
+dynamic $\gamma$ (SAD).
 
@@
 move, and in what additional statistics (if any) must be collected in
-order decide on that probability.  The methods we will compare fall in
-two categories:  weight-based methods, which only require a set of
-weights $w(E)$ be collected, and transition-matrix methods that
-require that statistics be collected for any transitions between pairs
-of energies.
+order decide on that probability.  The methods we compare belong to the
+family of weight-based methods, which only require a set of
+weights $w(E)$ be collected.
 
-Both weight-based and transition matrix methods calculate
+Weight-based methods calculate
 the density of states $D(E)$ for discrete energy
@@
 
-We will begin by introducing two related weight-based methods which
+We will begin by introducing three related weight-based methods which
 rely on a weight function $w(E)$, noting that the method proposed in
@@
                for the same simulations.
-           (c) \davidsays{Make this similar} The average entropy error for each MC method is shown for $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$ with
-               the same fixed energy range reference used in (a). The update factor $\gamma_t$ for
-               $N=50$ versus iteration for the four different methods: WL, $1/t$-WL, SAMC, and SAD is shown beneath.}\label{fig:fast-slow-gamma}
+           (c) The entropy error for each MC method for $N=50$,
+               $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
+               as a function of number of iterations run. The error is
+               determined exactly as in (a).  
+           (d) The update factor $\gamma_t$ versus iteration number
+               for the same simulations.
+  }
 \end{figure*}

### 2019-03-07 papers/histogram/sad.tex
edit to two paragraphs of sad paper

@@
 single canonical Monte Carlo simulation to predict properties for
-nearby temperatures~\cite{ferrenberg1988new}.  For large simulations
-this approach is limited to a narrow temperature range because a single
-canonical simulation explores only a small range of energies.  This led
-to a variety of ``broad'' (or ``flat'') histogram
+nearby temperatures~\cite{ferrenberg1988new}.  For large systems this
+approach is limited to a narrow temperature range because a single
+canonical simulation explores only a small range of energies.  This
+led to a variety of ``broad'' (or ``flat'') histogram
 methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
   wang2001determining, wang2001efficient, trebst2004optimizing}, which
-attempt to explore a wide range of energies.  These methods also
-benefit in that they are unlikely to be trapped in a low enetropy state.
+attempt to explore a wide range of energies.  Another benefit, in
+contrast with low-temperature canonical Monte Carlo, is that these
+approaches cannot be trapped in a local energy minimum.
 
-Wang and Landau introduced an algorithm that uses an update factor and
-a statistical histogram to compute the density of states of a given
-system~\cite{wang2001determining, wang2001efficient}.  While the method
-is incredibly powerful, it has several user-defined
-parameters~\cite{landau2004new}.  This can make its application to a
-variety of systems something of an art-form since the user needs to
-determine the ideal parameters for the particular system being studied.
-Also, detailed balance is violated (although only for short periods of
-time), ensuring that convergence is not guaranteed.  This raises
-several questions such as when the method does converge, at what rate will
-the method converge to the correct density of states. Also, can the
-user appropriately decide what parameters to choose so that the
-algorithm solves the problem in the most ideal way.
+Wang and Landau introduced an algorithm (WL) that uses an update
+factor and a statistical histogram to compute the density of states of
+a given system~\cite{wang2001determining, wang2001efficient}.  While
+the method is incredibly powerful, it has a few disadvantage.
+Firstly, it requires the user to select a number of parameters,
+including the range of energies to be studied~\cite{landau2004new}.
+This adds an additional hurdle to its application to a variety of
+systems.  The simulation violates detailed balance, with the size of
+the violation decreasing with time, which complicates convergence
+analysis.  In fact, the error in a WL computation the error has been
+demonstrated to saturate at a non-zero value, i.e. the method does not
+converge to the true density of states~\davidsays{Find as many
+  citations as we can! Presumably some from the following paragraph?}.
 

### 2019-03-15 papers/histogram/sad.tex

@@
 analysis.  In fact, the error in a WL computation the error has been
-demonstrated to saturate at a non-zero value, i.e. the method does not
-converge to the true density of states~\davidsays{Find as many
-  citations as we can! Presumably some from the following paragraph?}.
+demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method does not
+converge to the true density of states~\cite{belardinelli2007wang,
+belardinelli2008analysis, belardinelli2014intrinsic,
+singh2012density}.
 
@@
 energy. Our proposed method SAD requires the user to input
-$T_\text{min}$ which is an immediate disadvantage of the method;
-however, this allows the user to define an energy simulation
-range between $T_\text{min}$ and $\infty$. Setting a physical parameter
-such as a minimum temperature $T_\text{min}$ is intuitive vs. a user
-needing to know a prior either an energy range or some unphysical
+$T_\text{min}$, the lowest temperature of interest,
+which is an immediate disadvantage of the method. However,
+setting a physical parameter such as a minimum temperature
+$T_\text{min}$ is possibly easier for a user
+than determining in advance either an energy range or some unphysical
 parameter $t_0$.
@@
 all the energies have been long ago found,
-we wish for a lower $t_0$ in order to more rapidly refine the remaining
-error in entropy.  We enable this
-by tracking the last time we encountered a new energy $t_L$, and
-gradually transition to lower $t_0$ (but still asymptotically scaling
+we wish for a lower update factor in order to more rapidly refine the remaining
+error in entropy.  We track the time at which we first visted each possible
+energy.  We define $t_L$ to be the last time that we encountered an energy
+that we currently believe is in the energy range of interest, so a $t\gg t_L$
+we feel confident that we have established the true energy range of interest. We
+gradually transition to a lower update factor (but still asymptotically scaling
 as $\gamma_t \propto 1/t$ to ensure eventual convergence).  Finally,
@@
 \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
-  \caption{(a) The entropy error for each MC method for $N=50$,
+  \caption{(a) The average entropy error for each MC method for $N=50$,
                $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
@@
                and worst simulations for each method are shown as a
-               semi-transparent shaded area.
-           (b) The update factor $\gamma_t$ versus iteration number
+               semi-transparent shaded area, and
+           (b) the update factor $\gamma_t$ versus iteration number
                for the same simulations.
-           (c) The entropy error for each MC method for $N=50$,
+           (c) The average entropy error for each MC method for $N=50$,
                $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run. The error is
-               determined exactly as in (a).  
-           (d) The update factor $\gamma_t$ versus iteration number
+               as a function of number of iterations run, and
+           (d) the update factor $\gamma_t$ versus iteration number
                for the same simulations.
-  }
+  }\label{fig:fast-slow-gamma}
 \end{figure*}
@@
 \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
-  \caption{The average entropy error for each MC method is shown for $\eta = 0.17$ and $T_{\min} = 1$ with
-  SAD serving as the reference.}\label{fig:n256}
+  \caption{
+  The average entropy error for each MC method for $N=256$,
+               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
+               as a function of number of iterations run.  The error is
+               averaged over 8 independent simulations, and the best
+               and worst simulations for each method are shown as a
+               semi-transparent shaded area.  The update factor for this
+               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
 \end{figure}

### 2019-03-19 papers/histogram/sad.tex
a little more sad writing

@@
 analysis.  In fact, the error in a WL computation the error has been
-demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method does not
-converge to the true density of states~\cite{belardinelli2007wang,
+demonstrated to saturate at a non-zero value~\cite{yan2003fast},
+i.e. the method does not converge to the true density of
+states~\cite{belardinelli2007wang, belardinelli2007fast,
   belardinelli2008analysis, belardinelli2014intrinsic,
   singh2012density}.  
-
-Around the same time, the convergence rate for WL was being
-explored~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}. It was found that utilizing modification factors
-that decrease faster than $1/t^2$ leads to
-nonconvergence~\cite{belardinelli2007fast}.  This leads to the so-called
-$1/t$-WL algorithm which ensures that excess CPU time is not wasted by
-continuing to perform calculations once the error in the density of
-states becomes saturated~\cite{belardinelli2008analysis}. An effort to
-reduce the number of user defined parameters of WL and formulate a
-theory for why the method converges despite detailed balance being
-violated although infrequently was also being undertaken.
+%
+Belardinelli and Pereyra demonstrated that allowing the update factor
+to decrease faster than $1/t$ leads to
+nonconvergence~\cite{belardinelli2007wang}.
+%% Around the same time, the convergence rate for WL was being
+%% explored~\cite{zhou2005understanding,lee2006convergence,
+%%   belardinelli2007wang}.
+%
+This leads to their $1/t$-WL algorithm which ensures that the error
+continues to decrease asymptotically as
+$1/\sqrt{t}$~\cite{belardinelli2008analysis}. \davidsays{I don't
+  understand the following: An effort to reduce the number of user
+  defined parameters of WL and formulate a theory for why the method
+  converges despite detailed balance being violated although
+  infrequently was also being undertaken.  CUT THIS MAYBE?}
 

### 2019-04-19 papers/histogram/sad.tex
cut hard-to-understand sentence
fixed a few references to figures in the paper.

@@
 continues to decrease asymptotically as
-$1/\sqrt{t}$~\cite{belardinelli2008analysis}. \davidsays{I don't
-  understand the following: An effort to reduce the number of user
-  defined parameters of WL and formulate a theory for why the method
-  converges despite detailed balance being violated although
-  infrequently was also being undertaken.  CUT THIS MAYBE?}
+$1/\sqrt{t}$~\cite{belardinelli2008analysis}.
 
@@
 struggle to converge if the update factor decreases too rapidlly.
-Fig.~\ref{fig:N50-ff0.3-avg-error-slow} shows SAMC with inappropriate choices
-of $t_0$ i.e. less than $\nicefrac{10^5}{\sqrt{t}}$ failing to converge
+Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
+of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
 in a reasonable amount of time. Thus, the update factor $\gamma$, shown

### 2019-04-19 papers/histogram/sad.tex

@@
 The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
-describe in detail flat-histogram methods.  Section~\ref{sec:weight} outlines
-weight-based histogram methods used in this work.
-
+describe in detail flat-histogram methods.
+%
 Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
@@
 similar accuracy, so as to accurately determine the density of states
-over a broad range of energies.  Once the density of states is known,
-a number of thermodynamic quantities---such as heat capacity or
-internal energy---can be easily computed for any temperature.
+over a broad range of energies---and thus to determine the thermodynamic
+quantities such as heat capacity or internal energy
+over a broad range of temperatures.
 Properties that require more information---such as a spatial
@@
 algorithm differs in how it determines the probability of accepting a
-move, and in what additional statistics (if any) must be collected in
-order decide on that probability.  The methods we compare belong to the
-family of weight-based methods, which only require a set of
-weights $w(E)$ be collected.
+move, and in what additional statistics must be collected in
+order decide on that probability.
 
-Weight-based methods calculate
-the density of states $D(E)$ for discrete energy
-levels~\cite{haber2018performance}. For this reason, energy binning
-becomes an important consideration for complex continous systems.
+Flat histogram methods calculate
+the density of states $D(E)$ for a discrete set of
+energies~\cite{haber2018performance}. For this reason, energy binning
+becomes an important consideration for systems with a continuum of
+energies possible.
 Energy bins are typically of uniform size for the entire energy
@@
 
-\section{Weight-based methods}\label{sec:weight}
-
-We will begin by introducing three related weight-based methods which
-rely on a weight function $w(E)$, noting that the method proposed in
+In this section we will introduce four related flat histogram methods
+which each rely on a weight function $w(E)$, noting that the method proposed in
 this paper (discussed in Section~\ref{sec:sad}) also falls in this
@@
 
-Each weighted density approach uses a random walk in energy space to
+Each approach uses a random walk in energy space to
 estimate the density of states.  The core of these approaches
@@
 known a priori.  The time to converge depends only on the choice of
-parameters $t_0$.  Unfortunately, there is no
+parameter $t_0$.  Unfortunately, there is no
 prescription for finding an acceptable value for $t_0$, and
@@
 
-\paragraph{Energy binning with SAD}
+%~ \paragraph{Energy binning with SAD}
 
 A significant advantage of SAD over other flat histogram methods is
-that energy binning is solved dynamically. SAD should perform
+that energy binning is handled dynamically. SAD should perform
 independently for a reasonable range of binning because $\gamma \propto
 \nicefrac{N_S}{t}$.  As the number of energy states found $N_S$
-increases (fine binning), the time spent $t$ in each bin will decrease.
-This argument explains why SAD should perform well for complex systems
+increases (fine binning), the time spent $t$ in each bin will decrease
+with the effect that the convergence should be roughly independent of
+the bin size chosen.
+This argument explains why SAD should perform well for continuum systems
 in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
@@
 
-\paragraph{Scaling, step size, and SAD}
-Methods such as WL can fail to converge for complex systems in part
-because the same random MC moves are performed for the entire energy
-range.  The main problem lies in trying to determine how to optimize
-the step size such that the time to explore all energies is reasonable.
-Small step sizes lead to improved sampling at the expense of slow
-convergence. Koh et al. proposed multiplying $\gamma$ by an energy
-dependent factor in order to improve the converge of WL in regards to
-convergence and scaling~\cite{koh2013dynamically}.  Because the update
-factor for SAD has an energy dependent factor already present (and
-dynamically adjusted), we expect that SAD will perform consitently
-scaling free for a reasonable choice of step size.  SAD should perform
-extremely well compared with other flat histogram methods when applied
-to complex continuous systems such as polymers/proteins and spin models.
+%~ \paragraph{Scaling, step size, and SAD}
+%~ Methods such as WL can fail to converge for complex systems in part
+%~ because the same random MC moves are performed for the entire energy
+%~ range.  The main problem lies in trying to determine how to optimize
+%~ the step size such that the time to explore all energies is reasonable.
+%~ Small step sizes lead to improved sampling at the expense of slow
+%~ convergence. Koh et al. proposed multiplying $\gamma$ by an energy
+%~ dependent factor in order to improve the converge of WL in regards to
+%~ convergence and scaling~\cite{koh2013dynamically}.  Because the update
+%~ factor for SAD has an energy dependent factor already present (and
+%~ dynamically adjusted), we expect that SAD will perform consitently
+%~ scaling free for a reasonable choice of step size.  SAD should perform
+%~ extremely well compared with other flat histogram methods when applied
+%~ to complex continuous systems such as polymers/proteins and spin models.
 
-\subsection{Convergence}
-Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
-convergence criteria.  However, we also want to ensure that SAD avoids
-the exponentially poor convergence time that SAMC suffers when $t_0$ is
-small as discussed in Section~\ref{sec:samc-convergence}.
-To place a lower bound on the convergence of SAD, we will again
-approximate the total entropy change using an integral.  To sidestep
-the dynamical nature of SAD, we ensure that \emph{after discovering the
-last important energy} at time $t_L$ the entropy change
-$\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
-Thus we integrate the update factor from $t_L$ to some minimum time
-$t_{\min}$ necessary for convergence.  Technically this is not a lower
-bound as in the case of SAMC because the algorithm will already have a
-reasonable approximation for the entropy.
-\begin{align}
-\Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
-     \frac{
-       3\langle S\rangle + \frac{t}{t_L}
-     }{
-       3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
-     } dt \\
-&\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
-\end{align}
-where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
-that it takes a long time to find all the interesting energy states.
-Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
-of order unity.  This omission only makes our bound more conservative,
-and enables us to solve to find that
-\begin{align}
-  t_{\min} &= \frac32 t_L
-\end{align}
-which means that we only need to run fifty percent longer after we
-discover the final interesting energy state in order to have made
-sufficient updates to \emph{possibly} converge to the true entropy.
+%~ \subsection{Convergence}
+%~ Because SAD scales as $1/t$ for large $t$, it satisfies the SAMC
+%~ convergence criteria.  However, we also want to ensure that SAD avoids
+%~ the exponentially poor convergence time that SAMC suffers when $t_0$ is
+%~ small as discussed in Section~\ref{sec:samc-convergence}.
+%~ To place a lower bound on the convergence of SAD, we will again
+%~ approximate the total entropy change using an integral.  To sidestep
+%~ the dynamical nature of SAD, we ensure that \emph{after discovering the
+%~ last important energy} at time $t_L$ the entropy change
+%~ $\Delta S_{\text{tot}}$ can be made in a reasonable number of iterations.
+%~ Thus we integrate the update factor from $t_L$ to some minimum time
+%~ $t_{\min}$ necessary for convergence.  Technically this is not a lower
+%~ bound as in the case of SAMC because the algorithm will already have a
+%~ reasonable approximation for the entropy.
+%~ \begin{align}
+%~ \Delta S_{\text{tot}} &= \int_{t_L}^{t_{\min}}
+     %~ \frac{
+       %~ 3\langle S\rangle + \frac{t}{t_L}
+     %~ }{
+       %~ 3\langle S\rangle + \frac{t}{N_S}\frac{t}{t_L}
+     %~ } dt \\
+%~ &\approx 3\Delta S_{\text{tot}} \left(1-\frac{t_L}{t_{\min}}\right)
+%~ \end{align}
+%~ where we assumed that $t_L\gg N_S\langle S\rangle$, which is to say
+%~ that it takes a long time to find all the interesting energy states.
+%~ Furthermore, we omitted the $\frac{t}{t_L}$ in the numerator, which is
+%~ of order unity.  This omission only makes our bound more conservative,
+%~ and enables us to solve to find that
+%~ \begin{align}
+  %~ t_{\min} &= \frac32 t_L
+%~ \end{align}
+%~ which means that we only need to run fifty percent longer after we
+%~ discover the final interesting energy state in order to have made
+%~ sufficient updates to \emph{possibly} converge to the true entropy.
 
@@
 a precise reference system. For each simulation, the reference system
-is chosen to be the final output of a SAMC simulation with a fixed interesting energy range.
-We note this because SAMC does not require an energy range as an input parameter; however,
-by fixing the energy range to for SAMC, a smaller $t_0$ can be chosen considerably reducing the
-simulation time.
-\begin{align}
-\epsilon_\text{avg} = 1 - \bigg\lvert\frac{S_\text{method}}{S_\text{ref}}\bigg\rvert
-\end{align}
-Simulations that take more iterations to reach the same entropy error as a
-different method take longer to converge. This behavior is seen in
-Fig.~\ref{fig:fast-slow-gamma}.
+is chosen to be the final output of a SAMC simulation with a fixed energy range
+corresponding to the temperature range of interest.
+Although SAMC does not require an energy range as an input parameter,
+we find that by limiting the simulation to this energy range, we can
+acheive much faster convergence with a smaller $t_0$.  We compute an average
+of the error by averaging the error in the entropy over the 
+interesting energy range, and then averaging \emph{this} error over several
+simulations run with different random number seeds.
+%~ \begin{align}
+%~ \epsilon_\text{avg} =
+  %~ \frac{1}{N_S \text{num seeds}}\sum_s^{\text{num seeds}}\sum_{i}^{N_S} \left|S_\text{ref}(E_i) - S_\text{method}(E_i)\right|
+%~ \end{align}
 
@@
 the slower translation scale. Because of this update behavior, these methods
-do not suffer in converging to the correct density of states from having a
-static parameter such as SAMC $t_0$ in the update factor. SAMC methods that converge
-in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not 
-necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
+are able to converge regardless of the displacement distance used in Monte
+Carlo moves.
 
-Sad performs similarly regardless of the displacement distance chosen although
-it transitions to $1/\sqrt{t}$ convergence after
+%~ SAMC methods that converge
+%~ in a reasonable amount of moves in Figure~\ref{fig:fast-slow-gamma}a do not 
+%~ necessarily do so for Figure~\ref{fig:fast-slow-gamma}b.
+%
+%~ Sad performs similarly regardless of the displacement distance chosen although
+%~ it transitions to $1/\sqrt{t}$ convergence after
 
@@
 \end{figure}
-%\begin{figure}
-%  \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
-%  \caption{\davidsays{This could be prettier, and do we need it? We
-%      definitely need a caption if we keep it.} Use this to create
-%    description of the system.}\label{fig:n256-lndos-comparison}
-%\end{figure}
+%~ \begin{figure}
+  %~ \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
+  %~ \caption{\davidsays{This could be prettier, and do we need it? We
+      %~ definitely need a caption if we keep it.} Use this to create
+    %~ description of the system.}\label{fig:n256-lndos-comparison}
+%~ \end{figure}
 
 \subsection{A periodic system with 256 atoms}
-For this simulation, we chose a minimum reduced
+Next we will introduce a considerably larger simulation containing
+256 atoms which has a total entropy range of around 1500.  This makes
+exploring the entire range of energies extremely expensive, and strongly
+favors methods that restrict the energy (or temperature) range of
+interest.
+For this simulation, we chose a much higher minimum reduced
 temperature of $1.0$, which corresponds to an interesting energy range
 from $-915$ to $-509$.  The number of important energy states for this
-system is then $N_S = E_{\min} - E_{\max} = 406$.
+system is then $N_S = E_{\min} - E_{\max} + 1 = 407$.
 The entropy of this system is has an entropy that
@@
 $10^{119}$ between the highest and lowest density of states.
+\davidsays{Also give the entropy difference, looking at the data.}
+\jpsays{$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
+gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'}
 
@@
 been made all but the SAMC simulation with the shortest $t_0$ have
-begun to converge as $1/\sqrt{t}$.  We then see the WL error begin to
+begun to converge as $1/\sqrt{t}$.  We then see the WL error
 saturate around $10^{10}$ moves.
 
-For the larger system, SAMC parameter values of $t_0$ that gave rapid
-convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
-converge rapidly.  Larger $t_0$ parameters are simulated for
-comparison with SAD.  SAD has significantly smaller fluctuations
-in the average entropy error for a given number of moves.
+%~ For this larger system, SAMC parameter values of $t_0$ that gave rapid
+%~ convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
+%~ converge rapidly.
 
@@
 
-\subsubsection{Discussion on convergence and update factor}
+%~ \subsection{Discussion on convergence and update factor}
 
-A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
-for the Wang-Landau algorithm decreases much more rapidly than
-$\nicefrac{1}{\sqrt{t}}$. It can be observed  for the simulations that
-were carried out that all the methods from the weight-based family
-struggle to converge if the update factor decreases too rapidlly.
-Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
-of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
-in a reasonable amount of time. Thus, the update factor $\gamma$, shown
-as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
-useful in determining which methods will converge in a reasonable
-amount of time. We compare each method to a reference SAMC
-simulation for an ideal choice of $t_0$.  This gives an unbiased
-comparison with our developed method SAD. The dynamic version of SAMC
-which we call SAD performs as well as SAMC (assuming the user can
-correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
-outperforms WL and 1/t-WL? for the same given number of moves. The
-absence of user defined parameters makes SAD an exciting alternative to
-methods where user experience or successive implementation iteration
-are needed to achieve reasonable results.
+%~ A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
+%~ for the Wang-Landau algorithm decreases much more rapidly than
+%~ $\nicefrac{1}{t}$. It can be observed  for the simulations that
+%~ were carried out that all the methods
+%~ struggle to converge if the update factor decreases too rapidlly.
+%~ Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
+%~ of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
+%~ in a reasonable amount of time. Thus, the update factor $\gamma$, shown
+%~ as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
+%~ useful in determining which methods will converge in a reasonable
+%~ amount of time. We compare each method to a reference SAMC
+%~ simulation for an ideal choice of $t_0$.  This gives an unbiased
+%~ comparison with our developed method SAD. The dynamic version of SAMC
+%~ which we call SAD performs as well as SAMC (assuming the user can
+%~ correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
+%~ outperforms WL and 1/t-WL? for the same given number of moves. The
+%~ absence of user defined parameters makes SAD an exciting alternative to
+%~ methods where user experience or successive implementation iteration
+%~ are needed to achieve reasonable results.
 
 \section{Conclusions}
+
+\davidsays{We find both 1/tWL and sad do really well and very similarly.
+SAD wins if you want to specify a temperature range rather than an energy.
+1/tWL cheated because you had to run a prior simulation in order to find
+the energy range of interest.  Yay good stuff.}
+
+\jpsays{Both SAD and 1/t-WL converge more rapidly than Wl or SAMC. SAD requires
+the user to specify a temperature range rather than an energy range as 1/t-WL
+does. In addition to requiring the energy range,
+1/t-WL requires a prior simulation in order to determine the energy range
+of interest. WL saturates in convergence as discussion in the introduction and
+requires the user to optimize the choice of parameters to avoid this such as
+increasing the flatness criterion.  SAMC converges for a reasonable choice of
+$t_0$ but this parameter can be difficult to tune especially across different
+energy bins, displacement distances, or systems. Additionaly, SAMC does not
+converge as rapidly as either SAD or 1/t-WL even for the best choice of $t_0$.}
 

### 2019-04-26 papers/histogram/sad.tex
these are the paper edits and fixed horizontal plotting range to be consistent for figures.

@@
 \usepackage{nicefrac} % Include for inline fractions
+
+\usepackage{xargs}                      % Use more than one optional parameter in a new commands
+\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
+\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
+\usepackage{mdframed}
+
+% define colors for comments
+\definecolor{dark-gray}{gray}{0.10}
+\definecolor{light-gray}{gray}{0.70}
 
@@
 \newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
-\newcommand{\tssays}[1]{{\color{red} [\cyan{Tanner:} \emph{#1}]}}
+\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
+\begin{document}
 
-\begin{document}
 \title{Stochastic Approximation Monte Carlo with a Dynamic Update
@@
 
-Wang and Landau introduced an algorithm (WL) that uses an update
-factor and a statistical histogram to compute the density of states of
-a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has a few disadvantage.
-Firstly, it requires the user to select a number of parameters,
-including the range of energies to be studied~\cite{landau2004new}.
-This adds an additional hurdle to its application to a variety of
-systems.  The simulation violates detailed balance, with the size of
-the violation decreasing with time, which complicates convergence
-analysis.  In fact, the error in a WL computation the error has been
-demonstrated to saturate at a non-zero value~\cite{yan2003fast},
-i.e. the method does not converge to the true density of
-states~\cite{belardinelli2007wang, belardinelli2007fast,
-  belardinelli2008analysis, belardinelli2014intrinsic,
-  singh2012density}.  
+Wang and Landau introduced an algorithm (WL) that uses an update factor
+and a statistical histogram to compute the density of states of a given
+system~\cite{wang2001determining, wang2001efficient}.  While the method
+is incredibly powerful, it has a few disadvantages. Firstly, it
+requires the user to select a number of parameters, including the range
+of energies to be studied~\cite{landau2004new}. This adds an additional
+hurdle to its application to a variety of systems because the user must
+simulate until the energy range of interest is known.  The simulation
+violates detailed balance albiet briefly as the size of the violation decreases
+with time, which complicates convergence analysis.  In fact, the error
+in a WL computation the error has been demonstrated to saturate at a
+non-zero value~\cite{yan2003fast}, i.e. the method does not converge to
+the true density of states~\cite{belardinelli2007wang,
+belardinelli2007fast, belardinelli2008analysis,
+belardinelli2014intrinsic, singh2012density}.
 %
@@
 
-Liang, Liu, and Carrol began to consider whether WL could be considered
+Liang, Liu, and Carrol began to consider that WL could be considered
 a special case of Stochastic Approximation whose convergence could be
@@
 In this work, we have developed a novel algorithm based on SAMC that
-does not require user-defined inputs and therefore should be easily
-applicable to a given system.  We call this method SAD (Stochastic Approximation
-with a dynamic update factor), and will discuss it in detail in the methods
-section. We compare it along with several flat histogram methods
-which include WL, $1/t$-WL, and SAMC.
+does not require an array of non-physical, user-defined inputs and
+therefore should be easily applicable to a given system. The method
+does require the user to define a temperature range ($T_\text{min}$ to $T_\infty$).
+We call this method SAD (Stochastic Approximation with a dynamic update
+factor), and will discuss it in detail in the methods section. We
+compare it along with several flat histogram methods which include WL,
+$1/t$-WL, and SAMC.
 
@@
 
-The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
-describe in detail flat-histogram methods.
-%
-Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
-In Section~\ref{sec:results}, we discuss the performance of the various methods
-applied to the model system.
+%The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
+%describe in detail flat-histogram methods.
+%Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
+%In Section~\ref{sec:results}, we discuss the performance of the various methods
+%applied to the model system.
 
@@
 developed in detail while summarizing algorithms that were developed in other
-works.  The following methods are introduced and applied to the
+works.  The following methods are discussed and simulated for the
 square-well fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
@@
 for any temperature, provided statistics are collected for each
-individual energy, which can then be reweighted for any temperature.
+individual energy, which can then be reweighted for any
+temperature~\cite{errington2003direct}.
 
 Each flat histogram Monte Carlo method uses a given set of ``moves''
-which change a system and should satisfy detailed balance.  Each
+which change the system and should satisfy detailed balance.  Each
 algorithm differs in how it determines the probability of accepting a
-move, and in what additional statistics must be collected in
+move and in what additional statistics must be collected in
 order decide on that probability.
@@
 acceptance probabilities changing with each move.  The severity of
-this violation decreased as we decrease $\gamma$.  The various
-weight-based methods discussed here differ in how the decreasing of $\gamma$
-is scheduled.
+this violation decreases as we decrease $\gamma$.  The various
+weight-based methods discussed here differ in how they schedule the
+decreasing of $\gamma$.
 
@@
 $\gamma$ so quickly that it will never (for any cutoff value) decrease the
-error in the density of states beyond a given value.
+error in the density of states beyond a given nonzero value.
 
@@
 each update~\cite{belardinelli2007fast}. They implemented a schedule
-that enforced if the update factor $\gamma$ is less than
+that enforced that if the update factor $\gamma$ is less than
 $\nicefrac{N_S}{t}$ then the update factor is set to
-$\nicefrac{N_S}{t}$ and the histogram is no longer tracked. Employing
-this scheduler, they found that the error saturation is avoided since
-the correct density of states is approached asymptotically as
+$\nicefrac{N_S}{t}$ where $N_S$ is the number of energy states and the
+histogram is no longer tracked. Employing this scheduler, they found
+that the error saturation is avoided since the correct density of
+states is approached asymptotically as
 $t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al further
 confirmed that the WL algorithm never converges exponentially and
-succesfully bounded the statistical error between
-$t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
+succesfully bounded the statistical error between $t^{-\frac12}$ and
+$1/t$~\cite{zhou2008optimal}.
 
@@
 original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$.
+arbitrary tunable parameter $t_0$,
 \begin{align}
@@
 to converge to the true density of states~\cite{liang2006theory,
-  liang2007stochastic}.  In addition, the energy range need not be
-known a priori.  The time to converge depends only on the choice of
-parameter $t_0$.  Unfortunately, there is no
-prescription for finding an acceptable value for $t_0$, and
-while the algorithm formally converges, for a poor choice of $t_0$
-that convergence can be far too slow to be practical.
-Liang et al. give a rule of thumb in which $t_0$
-is chosen in the range from $2N_S$ to $100N_S$ where $N_S$ is the number
-of energy bins~\cite{liang2007stochastic}.  Schneider et al. found
-that for the
-Ising model this heuristic is helpful for small spin
-systems, but that larger systems require an even higher $t_0$
-value~\cite{schneider2017convergence}.  Also, we find that for the square-well
-fluid $t_0$ in some cases needs to be two orders of magnitude higher
-than the rule of thumb of $100N_S$.
+liang2007stochastic}.  In addition, the energy range need not be known
+a priori.  The time to converge depends only on the choice of parameter
+$t_0$.  Unfortunately, there is no prescription for finding an
+acceptable value for $t_0$, and while the algorithm formally converges,
+for a poor choice of $t_0$ that convergence can be far too slow to be
+practical. Liang et al. give a rule of thumb in which $t_0$ is chosen
+in the range from $2N_S$ to $100N_S$ where $N_S$ is the number of
+energy bins~\cite{liang2007stochastic}.  Schneider et al. found that
+for the Ising model this heuristic is helpful for small spin systems,
+but that larger systems require an even higher $t_0$
+value~\cite{schneider2017convergence}.  Also we find that for one case
+with a small displacement distance for the square-well fluid, $t_0$
+needs to be as much as two orders of magnitude higher than the rule of
+thumb of $100N_S$ in order to converge in $10^{12}$ moves.
 
 Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t$ by a factor
-$\gamma_0$.  While this may result in an improved rate of convergence,
-it adds yet another parameter that must be empirically determined, and
-we have not explored this degree of freedom.
+$\gamma_0$~\cite{werlich2015stochastic}.  While this may result in an
+improved rate of convergence, it adds yet another parameter that must
+be empirically determined, and we have not explored this degree of
+freedom.
 
 \subsection{SAMC Convergence}\label{sec:samc-convergence}
-The difficulty in using the SAMC method lies in identifying an appropriate
+A primary difficulty in using the SAMC method lies in identifying an appropriate
 value for $t_0$.  A value that is either too high or too low will result
-in slow convergence to the true entropy.  It is instructive to consider
+in slow convergence to the true entropy of the system.  It is instructive to consider
 separately values of $t_0$ that are too low or too high.
@@
 We can place a rigorous \emph{lower} bound $t_{\min}$ on the number of
-moves required to find the true energy by considering the total change
+moves required to find the true entropy by considering the total change
 that needs to be made to the entropy.
@@
 
-One might reasonably choose to err on the high side when selecting a $t_0$.
+One might reasonably choose to err by selecting a large $t_0$.
 The rate of convergence is harder to estimate when $t_0$ is large, but
@@
 
-While for SAMC the update factor is defined in the original
+While for SAMC, the update factor is defined in the original
 implementation, for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
@@
 \end{align}
-We approximate this energy difference by
-$E_H -E_L$ where $E_H$ and $E_L$ are defined below.
-The entropy numerator of the update factor
-in general should scale with the total number of interesting
-energy states $N_S$, since updates to the weights are distributed between
-that many energy states.  The product $N_S\langle S\rangle$ is the
-total change of entropy required (starting from constant weights) to find
-the true entropy, and puts a lower bound on the convergence time.
-After \emph{long} times, when
-all the energies have been long ago found,
-we wish for a lower update factor in order to more rapidly refine the remaining
-error in entropy.  We track the time at which we first visted each possible
-energy.  We define $t_L$ to be the last time that we encountered an energy
-that we currently believe is in the energy range of interest, so a $t\gg t_L$
-we feel confident that we have established the true energy range of interest. We
-gradually transition to a lower update factor (but still asymptotically scaling
-as $\gamma_t \propto 1/t$ to ensure eventual convergence).  Finally,
-we wish for an update factor that is \emph{never} greater than 1, because
-a very large update factor could introduce errors in entropy that may take
-many iterations to remove.  The SAD
-expression for $\gamma_t$ which incorporates these ideas is:
+We approximate this energy difference by $E_H -E_L$ where $E_H$ and
+$E_L$ are defined below. The entropy numerator of the update factor in
+general should scale with the total number of interesting energy states
+$N_S$, since updates to the weights are distributed between that many
+energy states.  The product $N_S\langle S\rangle$ is the total change
+of entropy required (starting from constant weights) to find the true
+entropy, and puts a lower bound on the convergence time. After
+\emph{long} times, when all the energies have been a long time ago, we
+wish for a lower update factor in order to more rapidly refine the
+remaining error in entropy.  We track the time at which we first visted
+each possible energy.  We define $t_L$ to be the last time that we
+encountered an energy that we currently believe is in the energy range
+of interest, so a $t\gg t_L$ we feel confident that we have established
+the true energy range of interest. We gradually transition to a lower
+update factor (but still asymptotically scaling as $\gamma_t \propto
+1/t$ to ensure eventual convergence).  Finally, we wish for an update
+factor that is \emph{never} greater than 1, because a very large update
+factor could introduce errors in entropy that may take many iterations
+to remove.  The SAD expression for $\gamma_t$ which incorporates these
+ideas is:
 \begin{align}
@@
 
-Since SAD does not explore \emph{all} energy states it needs determine
-what energy range corresponds to the temperature range of interest,
-defined by $T_{\min}<T<\infty$.
-The
-simulation is responsible for determining and updating this energy
-range.
-Given the true entropy $S(E)$, we can define the interesting energy
-range as
+Since SAD does not explore \emph{all} energy states, it needs to
+determine what energy range corresponds to the temperature range of
+interest defined by $T_{\min}<T<\infty$. The simulation is responsible
+for determining and updating this energy range. Given the true entropy
+$S(E)$, we can define the interesting energy range as
 %\begin{align}
@@
 %\end{align}
-where $E(T)$ is the energy that maximizes $S-E/T$.  In the course of the
-simulation this precise energy is challenging to evaluate accurately.
+where $E(T)$ is the energy that maximizes $S-E/T$.  During the course of the
+simulation, this precise energy is challenging to evaluate accurately.
 In order to ensure that we sample this entire energy range adequately,
@@
 is made flat. At move $t$, $E_H$ and $E_L$ are the greatest and lowest
-energy that has prior to that move
-had the highest histogram value (i.e. been visited the most times) in
+energy that prior to that move
+had the highest histogram value (i.e. been visited the most times) during
 the course of the simulation.
@@
 
-During the simulation when considering a move outside of the energy
+During the simulation when considering a move inside of the energy
 range $E_L \le E \le E_H$, Boltzmann weights are used.  If $E\ge E_H$,
@@
 independently for a reasonable range of binning because $\gamma \propto
-\nicefrac{N_S}{t}$.  As the number of energy states found $N_S$
+\nicefrac{N_S}{t}$.  As the number of energy states $N_S$ found
 increases (fine binning), the time spent $t$ in each bin will decrease
@@
 in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
-similar way but this adds yet another parameter for the user to choose.
+similar way~\cite{werlich2015stochastic} but this adds yet another parameter for the user to choose.
 
@@
                $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run.  The error is
+               as a function of number of iterations run. The number of important
+               energies is $129$ of about $280$ energies.  The error is
                averaged over 8 independent simulations, and the best
@@
                $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run, and
+               as a function of number of iterations run. The number of important
+               energies is $407$ of about $1800$ energies, and
            (d) the update factor $\gamma_t$ versus iteration number
@@
 temperature of $1/3$, which corresponds to an interesting energy range
-from $-248$ to $-120$.  The entropy of this system is shown in
+from $-248$ to $-120$.  The number of important energy states
+for this system is therefore $N_S = 129$  The entropy of this system is shown in
 Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
@@
 In order to explore the effect of simulation details on convergence,
-we consider two values for the mean distance by which atoms are moved
+we consider two values for the displacement distance by which atoms are moved
 during a Monte Carlo step.  We began with a reasonable displacement
@@
 smaller displacement distance of $\delta_0 = 0.005\sigma$, which
-resulted in an acceptance rate of 86\%, which we expected converge
+resulted in an acceptance rate of 86\%, which as expected converge
 more slowly.
@@
 entropy as a function of time for this system with the reasonable
-displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
+displacement distance of $\delta_0 = 0.05\sigma$.  The solid/dashed lines
 represent the average of the absolute value of the error in the
@@
 been made all but the SAMC simulation with the shortest $t_0$ have
-begun to converge as $1/\sqrt{t}$.  We then see the WL error begin to
+begun to converge as $1/\sqrt{t}$.  We then see the WL error
 saturate around $10^{10}$ moves.
@@
 convergence time. SAMC simulations with a $t_0$ value that rapidly
-converged for $\delta = 0.05\sigma$ do not converge at all for a
-translation scale of $\delta = 0.005\sigma$. SAD, WL, and $1/t$-WL
-handle the shift in displacement distance and converge roughly as
-expected.
+converged for $\delta = 0.05\sigma$ do not converge at all in $10^{12}$
+moves for a translation scale of $\delta = 0.005\sigma$. It is also
+worth noting that for the smaller displacement distance, the SAMC rule
+of thumb of choosing $t_0$ to be approximately $100N_S$ is no longer
+valid. SAD, WL, and $1/t$-WL handle the shift in displacement distance
+and converge roughly as expected.
 
@@
 from $-915$ to $-509$.  The number of important energy states for this
-system is then $N_S = E_{\min} - E_{\max} + 1 = 407$.
-The entropy of this system is has an entropy that
-differs by 275, corresponding to a ratio of
-$10^{119}$ between the highest and lowest density of states.
-\davidsays{Also give the entropy difference, looking at the data.}
-\jpsays{$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
-gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'}
+system is then $N_S = 407$.
+The entropy of this system differs by 395, corresponding to a ratio of
+$10^{118}$ between the highest and lowest density of states.
+
+%$395.2833013702813 - 4.754731035740479e^{-8}$ = $395.2833$ so $2^{396.2833}$
+%gives us about $10^{118}$ from 'samc-1e4-256-s8.yaml'
 
@@
 begun to converge as $1/\sqrt{t}$.  We then see the WL error
-saturate around $10^{10}$ moves.
+saturates around $10^{10}$ moves.
 
@@
 \section{Conclusions}
-
-\davidsays{We find both 1/tWL and sad do really well and very similarly.
-SAD wins if you want to specify a temperature range rather than an energy.
-1/tWL cheated because you had to run a prior simulation in order to find
-the energy range of interest.  Yay good stuff.}
-
-\jpsays{Both SAD and 1/t-WL converge more rapidly than Wl or SAMC. SAD requires
+Both SAD and 1/t-WL converge more rapidly than Wl or SAMC. SAD requires
 the user to specify a temperature range rather than an energy range as 1/t-WL
-does. In addition to requiring the energy range,
-1/t-WL requires a prior simulation in order to determine the energy range
-of interest. WL saturates in convergence as discussion in the introduction and
+does. Both WL and 1/t-WL require a prior simulation in order to determine the energy range
+of interest. WL saturates in convergence as discussed in the introduction and
 requires the user to optimize the choice of parameters to avoid this such as
@@
 energy bins, displacement distances, or systems. Additionaly, SAMC does not
-converge as rapidly as either SAD or 1/t-WL even for the best choice of $t_0$.}
+converge as rapidly as either SAD or 1/t-WL even for the best choice of $t_0$.
 
@@
 parameters that could lead to the simulation not converging in a
-reasonable amount of time.
+reasonable amount of moves.
 
@@
 systems. A major advantage is that the user does not need to determine a
-priori how quickly the update factor $\gamma$ should be decreased as the
+priori how quickly the update factor $\gamma$ should decrease as the
 method SAD dynamically determines this.

### 2019-05-09 papers/histogram/sad.tex
start editing sad paper

@@
 simulation algorithms have been developed which calculate the
-thermodynamic properties of various systems for all temperatures.  The
+thermodynamic properties of various systems for over a range of temperatures.  This
 development began with the original histogram method, which used a
@@
 canonical simulation explores only a small range of energies.  This
-led to a variety of ``broad'' (or ``flat'') histogram
+led to a variety of ``flat'' (or ``broad'') histogram
 methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
   wang2001determining, wang2001efficient, trebst2004optimizing}, which
-attempt to explore a wide range of energies.  Another benefit, in
+explore a wider range of energies.  Another benefit, in
 contrast with low-temperature canonical Monte Carlo, is that these
@@
 
-Wang and Landau introduced an algorithm (WL) that uses an update factor
+Wang and Landau introduced the most widely used flat histogram
+algorithm (WL) that uses an update factor
 and a statistical histogram to compute the density of states of a given
@@
 is incredibly powerful, it has a few disadvantages. Firstly, it
-requires the user to select a number of parameters, including the range
+requires the user to select the range
 of energies to be studied~\cite{landau2004new}. This adds an additional
-hurdle to its application to a variety of systems because the user must
-simulate until the energy range of interest is known.  The simulation
+hurdle to its application to systems for which the interesting range
+of energies is not known \emph{a priori}.  The simulation
 violates detailed balance albiet briefly as the size of the violation decreases
@@
 
-Liang, Liu, and Carrol began to consider that WL could be considered
-a special case of Stochastic Approximation whose convergence could be
-mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
-2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
-considered a form of Stochastic Approximation Monte Carlo (SAMC). While
+Liang began to consider that WL could be considered
+a special case of the Stochastic Approximation approach, whose
+convergence could be
+mathematically proven~\cite{liang2006theory}. In
+2007, Liang, Liu, and Carrol et al. argued that WL can be
+considered a form of Stochastic Approximation Monte Carlo
+(SAMC)~\cite{liang2007stochastic}. \davidsays{I'm here} While
 SAMC can guarantee convergence, the method still has a system specific

### 2019-05-10 papers/histogram/sad.tex
edits of sad paper

@@
 \begin{abstract}
+  \davidsays{Let's look at the abstract together.}
   We present a novel Monte-Carlo algorithm based on the Stochastic
@@
 
-Liang began to consider that WL could be considered
-a special case of the Stochastic Approximation approach, whose
-convergence could be
-mathematically proven~\cite{liang2006theory}. In
-2007, Liang, Liu, and Carrol et al. argued that WL can be
-considered a form of Stochastic Approximation Monte Carlo
-(SAMC)~\cite{liang2007stochastic}. \davidsays{I'm here} While
-SAMC can guarantee convergence, the method still has a system specific
-user-defined variable which makes applying this algorithm to arbitrary
-systems difficult.
+Liang began to consider that WL could be considered a special case of
+the Stochastic Approximation approach, whose convergence could be
+mathematically proven~\cite{liang2006theory}. In 2007, Liang, Liu, and
+Carrol \emph{et al.} developed Stochastic Approximation Monte Carlo
+(SAMC)~\cite{liang2007stochastic}, and proved its convergence,
+although the method still has a system specific user-defined parameter
+which must be tuned when applying this algorithm to a new system.
 
-In this work, we have developed a novel algorithm based on SAMC that
-does not require an array of non-physical, user-defined inputs and
-therefore should be easily applicable to a given system. The method
-does require the user to define a temperature range ($T_\text{min}$ to $T_\infty$).
-We call this method SAD (Stochastic Approximation with a dynamic update
-factor), and will discuss it in detail in the methods section. We
-compare it along with several flat histogram methods which include WL,
-$1/t$-WL, and SAMC.
+In this work, we have developed an improved algorithm based on SAMC
+that does not require an array of non-physical, user-defined inputs
+and therefore should be easily applicable to any system. The method
+does require the user to define a temperature range of ($T_\text{min}$
+to $T_\infty$).  We call this method SAD (Stochastic Approximation
+with a Dynamic update factor), and will discuss it in detail in the
+methods section. We compare its convergence properties with three
+existing flat histogram methods: WL, $1/t$-WL, and SAMC.
 
-We consider the square-well fluid i.e. a system of particles whose
-interactions are governed by a square-well
+As a test case, we consider the square-well fluid i.e. a system of
+particles whose interactions are governed by a square-well
 potential~\cite{singh2003surface, barker2004perturbationSW}.  The
-square-well potential is an ideal test-bed as it is the simplest model
-that ensures both attractive and repulsive forces are experienced by
-interacting particles~\cite{barker1967-SW-perturbation, vega1992phase}.
-The potential $U(\textbf{r})$ for such a system is given by
+square-well potential is an ideal test-bed as it is a simple model for
+a liquid, which includes both attractive and repulsive
+interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The
+potential $U(\textbf{r})$ for such a system is given by
 \begin{equation}
@@
 reduced range of the potential well, and $\epsilon$ is its depth.
+This model has the further advantage that because the energy is
+discrete, binning is not required.
 
@@
 
-In this work, We compare a variety of flat histogram
-methods.  We outline the general workings of each algorithm that we
-developed in detail while summarizing algorithms that were developed in other
-works.  The following methods are discussed and simulated for the
-square-well fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
-Approximation Monte-Carlo (SAMC), and Stochastic Approximation with a
-dynamic $\gamma$ (SAD).
+In this work, we compare four flat histogram methods.  We outline the
+general workings of each algorithm that we developed in detail while
+summarizing algorithms that were developed in other works.  The
+following methods are discussed and simulated for the square-well
+fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
+Approximation Monte-Carlo (SAMC), and SAD.
 
@@
 
-Each flat histogram Monte Carlo method uses a given set of ``moves''
-which change the system and should satisfy detailed balance.  Each
-algorithm differs in how it determines the probability of accepting a
-move and in what additional statistics must be collected in
-order decide on that probability.
+All the flat histogram Monte Carlo methods begin with randomly chosen
+``moves'' which change the state of the system and must satisfy
+detailed balance.  Each algorithm differs in how it determines the
+probability of accepting a move and in what additional statistics must
+be collected in order decide on that probability.
 
-Flat histogram methods calculate
-the density of states $D(E)$ for a discrete set of
-energies~\cite{haber2018performance}. For this reason, energy binning
-becomes an important consideration for systems with a continuum of
-energies possible.
-Energy bins are typically of uniform size for the entire energy
-continuum~\cite{fasnacht2004adaptive}. Some methods such as
-AdaWL~\cite{koh2013dynamically}
-employ a tunable mechanism for controlling the binning for low entropic
-states in order to ensure the exploration of all energies.  The method
-introduced in this paper is designed to scale appropriately as bin size
-is changed, but we do not test this scaling, as we use a system with
-discrete energy levels.
+Flat histogram methods calculate the density of states $D(E)$ for a
+discrete set of energies~\cite{haber2018performance}. For this reason,
+energy binning becomes an important consideration for systems with a
+continuum of possible energies.  Energy bins are typically of uniform
+size for the entire energy continuum~\cite{fasnacht2004adaptive}. Some
+methods such as AdaWL~\cite{koh2013dynamically} employ a tunable
+mechanism for controlling the binning for low entropic states in order
+to ensure the exploration of all energies.  The method introduced in
+this paper is designed to scale appropriately as bin size is changed,
+but we do not test this scaling, as we use a system with discrete
+energy levels.
 
-In this section we will introduce four related flat histogram methods
-which each rely on a weight function $w(E)$, noting that the method proposed in
-this paper (discussed in Section~\ref{sec:sad}) also falls in this
-category.  In these algorithms, the probability of accepting a move is
-given by
+In this section we will introduce four closely related flat histogram
+methods which each rely on a weight function $w(E)$.  In these
+algorithms, the probability of accepting a move is given by
 \begin{equation}
@@
 \end{equation}
-the probability biases the simulation in favor of energies with low weights.
-A set of weights that are proportional to the density of states of the
-system $D(E)$ will result in an entirely flat histogram.  Thus
-flat histogram is a criteria for convergence for these methods.  To avoid
-overflow error , since the weights may vary over more than a
-few hundred orders of magnitude, the natural logarithm of the weights are used
-in this work.  Since $D(E) = S(E)$, the logarithm of the weights
-can be thought of as an approximation of the entropy.
+which biases the simulation in favor of energies with low weights.  A
+set of weights that are proportional to the density of states of the
+system $D(E)$ will result in an entirely flat histogram.  Thus flat
+histogram is a criteria for convergence for these methods.  To avoid
+overflow error, since the weights may vary over more than a few
+hundred orders of magnitude, the natural logarithm of the weights are
+stored.  Since in the microcanonical ensemble the entropy is defined
+as $S(E) \equiv k_B\ln(D(E))$, the logarithm of the weights can be
+thought of as an approximation of the entropy.
 
@@
 	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
-	+\gamma
+	+\gamma_t
 \end{equation}
-where $t$ is number of the current move, and $\gamma$ is an update
-factor.  This update causes the random walk to avoid energies that
-have been frequently sampled, leading to a rapid exploration of energy
-space.  This approach, however, violates detailed balance, due to the
+where $t$ is number of the current move, $\gamma(t)$ is a
+move-dependent update factor, and $E$ is the current energy.  This
+update causes the random walk to avoid energies that have been
+frequently sampled, leading to a rapid exploration of energy space.
+This approach, however, violates detailed balance, due to the
 acceptance probabilities changing with each move.  The severity of
-this violation decreases as we decrease $\gamma$.  The various
-weight-based methods discussed here differ in how they schedule the
-decreasing of $\gamma$.
+this violation decreases as we decrease $\gamma_t$.  The four methods
+differ primarily in how they schedule the decreasing of $\gamma_t$.
 
@@
 
-Wang-Landau's approach~\cite{wang2001efficient,wang2001determining,
+The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
   landau2014guide} begins with $\gamma=1$, and then decreases $\gamma$
-in discrete stages.  We track the number of steps at each energy during
-each stage in a histogram.  When that histogram is sufficiently flat,
-$\gamma$ is decreased by a specified factor, which is usually
-$\frac12$.  The flatness is defined by the ratio between the minimum
-value of the histogram and its average value.  When this flatness
-reaches a specified threshold (typically 0.8), the $\gamma$ value is
-decreased.  This approach requires that the energy range of interest
-be known in advance, and difficulties can occur with this flatness
-criteria due to the fact that some energies in this energy range might
-never be sampled~\cite{haber2014transition}.  The entire process is
-repeated until $\gamma$ reaches a desired cutoff.
+in discrete stages.  We track the number of moves ending at each
+energy in a histogram.  When that histogram is sufficiently flat,
+$\gamma$ is decreased by a specified factor of $\frac12$.  The
+flatness is defined by the ratio between the minimum value of the
+histogram and its average value.  When this flatness reaches a
+specified threshold (typically 0.8), the $\gamma$ value is decreased
+and the histogram is reset to zero.  This approach requires that the
+energy range of interest be known in advance, and difficulties can
+occur with this flatness criteria due to the fact that some energies
+in this energy range might never be
+sampled~\cite{haber2014transition}.  The entire process is repeated
+until $\gamma$ reaches a desired cutoff.
 
@@
 acheived, the flatness criterion, and the cutoff that determines when
-the computation is complete.  In addition, an energy range (or more
-specifically, a set of energies) must be supplied, so that the
+the computation is complete.  In addition, an energy range (or in
+general, a \emph{set} of energies) must be supplied, so that the
 flatness criterion can be defined.
 
-While this approach
-is very efficient and has been widely used, it suffers a few
-shortcomings.  Firstly, the set of energies must be specified
-~\cite{wang2001efficient, schulz2003avoiding, yan2003fast}, which may
-require multiple simulations. Secondly, while Wang-Landau converges quickly,
-it does not in general converge to the true density of states
-~\cite{belardinelli2008analysis, zhou2008optimal}.  It can and does decrease
-$\gamma$ so quickly that it will never (for any cutoff value) decrease the
-error in the density of states beyond a given nonzero value.
+While this approach is very efficient and has been widely used, it
+suffers a few shortcomings.  Firstly, the set of energies must be
+specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
+which may require multiple simulations. Secondly, while Wang-Landau
+converges quickly, it does not in general converge to the true density
+of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
+and does decrease $\gamma$ so quickly that it will never (for any
+cutoff value) decrease the error in the density of states beyond a
+given nonzero value.
 
 \subsection{$1/t$-Wang-Landau}
-The Wang-Landau algorithm's error saturation can be corrected by
-modifying the update factor such that it does not decrease too quickly.
-Belardinelli and Pereyra raised the question as to whether the update
-factor should be decrease by $\nicefrac12$ or some other factor for
-each update~\cite{belardinelli2007fast}. They implemented a schedule
-that enforced that if the update factor $\gamma$ is less than
-$\nicefrac{N_S}{t}$ then the update factor is set to
-$\nicefrac{N_S}{t}$ where $N_S$ is the number of energy states and the
-histogram is no longer tracked. Employing this scheduler, they found
-that the error saturation is avoided since the correct density of
-states is approached asymptotically as
+The error saturation of Wang-Landau algorithm algorithm can be
+corrected by modifying the update factor such that it does not
+decrease too quickly.  Belardinelli and Pereyra raised the question as
+to whether the update factor should be decrease by $\nicefrac12$ or
+some other factor for each update~\cite{belardinelli2007fast}. They
+implemented a schedule that enforced that if the update factor
+$\gamma$ is ever less than $\nicefrac{N_S}{t}$ then the update factor
+is set to $\nicefrac{N_S}{t}$ where $N_S$ is the number of energy
+states and the histogram is no longer tracked. Employing this
+scheduler, they found that the error saturation is avoided since the
+correct density of states is approached asymptotically as
 $t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al further
@@
 Schneider et. al outlines minor refinements to the $1/t$-WL algorithm
-including $\nicefrac{N_E}{t}$ scaling and switching from standard WL to
-$1/t$-WL when the update factor $\gamma <
-\nicefrac{N_E}{t}$~\cite{schneider2017convergence}. As per the original
-$1/t$ implementation~\cite{belardinelli2007fast}, all energy states $N_E$
-are required to be visited at least once $H(E) \neq 0$ effectively
-avoiding the concept of `flatness'.
+including $\nicefrac{N_S}{t}$ scaling and switching from standard WL
+to $1/t$-WL when the update factor $\gamma <
+\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. \davidsays{Above
+  we say that 1/t-WL does this.  One must be wrong?} As per the
+original $1/t$ implementation~\cite{belardinelli2007fast}, the update
+factor is decreased once all $N_S$ energy states are visited at least
+once (i.e. $H(E) \neq 0$) effectively avoiding the concept of
+`flatness'.
 
@@
 where as above $t$ is the number of moves that have been attempted.
-The major advantage that SAMC offers is its proven convergence.
+SAMC offers extreme simplicity, combined with is proven convergence.
 Provided the update factor satisfies
@@
 to converge to the true density of states~\cite{liang2006theory,
-liang2007stochastic}.  In addition, the energy range need not be known
-a priori.  The time to converge depends only on the choice of parameter
-$t_0$.  Unfortunately, there is no prescription for finding an
-acceptable value for $t_0$, and while the algorithm formally converges,
-for a poor choice of $t_0$ that convergence can be far too slow to be
-practical. Liang et al. give a rule of thumb in which $t_0$ is chosen
-in the range from $2N_S$ to $100N_S$ where $N_S$ is the number of
-energy bins~\cite{liang2007stochastic}.  Schneider et al. found that
-for the Ising model this heuristic is helpful for small spin systems,
-but that larger systems require an even higher $t_0$
-value~\cite{schneider2017convergence}.  Also we find that for one case
-with a small displacement distance for the square-well fluid, $t_0$
-needs to be as much as two orders of magnitude higher than the rule of
-thumb of $100N_S$ in order to converge in $10^{12}$ moves.
+  liang2007stochastic}.  In addition, the energy range need not be
+known a priori.  The time to converge depends only on the choice of
+parameter $t_0$.  Unfortunately, there is no prescription for finding
+an acceptable value for $t_0$, and while the algorithm formally
+converges, for a poor choice of $t_0$ that convergence can be far too
+slow to be practical. Liang \emph{et al.} give a rule of thumb in
+which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
+is the number of energy bins~\cite{liang2007stochastic}.  Schneider
+\emph{et al.} found that for the Ising model this heuristic is helpful
+for small spin systems, but that larger systems require an even higher
+$t_0$ value~\cite{schneider2017convergence}.  We will describe below
+one case we examined, in which $t_0$ needs to be as much as two orders
+of magnitude higher than the rule of thumb of $100N_S$ in order to
+converge in $10^{12}$ moves.
 
@@
 
-\subsection{SAMC Convergence}\label{sec:samc-convergence}
-A primary difficulty in using the SAMC method lies in identifying an appropriate
-value for $t_0$.  A value that is either too high or too low will result
-in slow convergence to the true entropy of the system.  It is instructive to consider
-separately values of $t_0$ that are too low or too high.
+\subsection{SAMC convergence time}\label{sec:samc-convergence}
+A primary difficulty in using the SAMC method lies in identifying an
+appropriate value for $t_0$.  Although SAMC is proven to formally
+converge regardless of the $t_0$ value, a choice that is either too
+high or too low will result in prohibitively slow convergence to the true
+entropy of the system.  It is instructive to consider separately
+values of $t_0$ that are too low or too high.
 
@@
 The Stochastic Approximation with Dynamic update factor (SAD) method
-is a variant of the SAMC
-Algorithm that attempts to dynamically choose the modification factor
-rather than relying on system dependent parameters such as $t_0$ or
-$\gamma_0$.  There is an immediate advantage of such an algorithm where
-parameters are chosen independent of system size or type. Each
-Flat-Histogram method has unique advantages and disadvantages.
-Wang-Landau requires an energy range for initialization.  SAMC removes
-this energy range requirement but requires simulating every possible
-energy. Our proposed method SAD requires the user to input
-$T_\text{min}$, the lowest temperature of interest,
-which is an immediate disadvantage of the method. However,
-setting a physical parameter such as a minimum temperature
-$T_\text{min}$ is possibly easier for a user
-than determining in advance either an energy range or some unphysical
-parameter $t_0$.
+is a variant of the SAMC Algorithm that attempts to dynamically choose
+the modification factor rather than relying on system dependent
+parameters such as $t_0$ or $\gamma_0$.  There is an immediate
+advantage of such an algorithm where parameters are chosen independent
+of system size or type. Each flat-histogram method has unique
+advantages and disadvantages.  Wang-Landau and $1/t$-WL require an
+energy range for initialization.  SAMC removes this energy range
+requirement but requires simulating every possible energy. Our
+proposed method SAD requires the user to input $T_\text{min}$, the
+lowest temperature of interest, which is an immediate disadvantage of
+the method. However, identifying a minimum temperature of interest
+$T_\text{min}$ may be easier for a user than determining in advance
+an energy rang of intereste or the unphysical parameter $t_0$.
 
@@
 factor that is \emph{never} greater than 1, because a very large update
-factor could introduce errors in entropy that may take many iterations
+factor could introduce very large errors in entropy that may take many iterations
 to remove.  The SAD expression for $\gamma_t$ which incorporates these
@@
 \end{align}
-This factor asymptotically has the same $1/t$ behavior as the original
-SAMC algorithm;
-however now every time a new energy is found to be important
-during the course of
-the simulation, the update factor will experience a jump. This behavior
-allows SAD to dynamically prevent the update factor from decreasing too
-rapidly.
+where $E_H$ and $E_L$ are the current estimates for the highest and
+lowest energies of interest as defined below.  This factor
+asymptotically has the same $1/t$ behavior as the original SAMC
+algorithm and with the same $N_S$ prefactor used by the $1/t$-WL
+method; however for earlier values of $t$, the update factor drops as
+$1/t^2$ and jumps every time a new energy is determined to be of
+interest.  This behavior allows SAD to dynamically prevent the update
+factor from decreasing too rapidly.
 
@@
   \includegraphics[width=\columnwidth]{figs/gamma-n256}
-  \caption{The update factor $\gamma_t$ for $N=256$ versus
-    iteration for four different methods: WL, $1/t$-WL, SAMC, and SAD}
+  \caption{The update factor $\gamma_t$ versus iteration for $N=256$
+    square-well atoms, for four different methods: WL, $1/t$-WL, SAMC,
+    and SAD}
     \label{fig:gamma-vs-t}
 \end{figure}
-Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related methods
-SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains constant
-before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for many iterations,
-and then decreases very rapidly, with $1/t$-WL behaving similarly before
-transitioning to a $1/t$ behavior.
-The update factor for SAD fluctuates
-dynamically around a value less than 1 for early MC moves, and then
-decreases as approximately $1/t$ while continuing to fluctuate as
-new energies are found to be important.  At late times, the SAD $\gamma_t$
-decreases as approximately $1/t^2$ until asymptoting to $N_S/t$, which
-is the same as $1/t$-WL.
+Figure~\ref{fig:gamma-vs-t} compares $\gamma_t$ for the related
+methods SAD, WL, $1/t$-WL, and SAMC.  For SAMC, $\gamma_t$ remains
+constant before dropping as $1/t$.  WL $\gamma_t$ remains at 1 for
+many iterations, and then decreases very rapidly, with $1/t$-WL
+behaving similarly but decreasing more aggressively before
+transitioning to a more conservative $1/t$ behavior.  The update
+factor for SAD fluctuates dynamically around a value less than 1 for
+early MC moves, and then decreases as approximately $1/t$ while
+continuing to fluctuate as new energies are found to be important.  At
+intermediate times, the SAD $\gamma_t$ decreases as $1/t^2$ before
+asymptoting to $N_S/t$, which is the same as $1/t$-WL.
 
@@
 During the simulation when considering a move inside of the energy
-range $E_L \le E \le E_H$, Boltzmann weights are used.  If $E\ge E_H$,
-the weight is taken to be
+range of interest $E_L \le E \le E_H$, the weights are used as in the
+three methods already described.  If $E\ge E_H$, the weight is taken
+to be
 \begin{align}\label{eq:highw}
@@
 \end{align}
-which corresponds to an infinite temperature.
-This choice ensures that
-if the maximum in entropy is at an energy $E_{\max}>E_H$, then the energy
-$E_{\max}$
-will eventually have the highest number of counts and the ratcheting will
-result in $E_H\ge E_{\max}$ .
-At lower energies, Boltzmann weights corresponding to the minimum temperature
+which corresponds to an infinite temperature.  This choice ensures
+that if the maximum in entropy is at an energy $E_{\max}>E_H$, then
+the energy $E_{\max}$ will eventually have the highest number of
+counts and the ratcheting will result in $E_H\ge E_{\max}$ .  At lower
+energies, Boltzmann weights corresponding to the minimum temperature
 are used:
@@
 
-A significant advantage of SAD over other flat histogram methods is
-that energy binning is handled dynamically. SAD should perform
-independently for a reasonable range of binning because $\gamma \propto
-\nicefrac{N_S}{t}$.  As the number of energy states $N_S$ found
-increases (fine binning), the time spent $t$ in each bin will decrease
-with the effect that the convergence should be roughly independent of
-the bin size chosen.
-This argument explains why SAD should perform well for continuum systems
-in terms of binning. SAMC has a prefactor $\gamma_0$ to aid in a
-similar way~\cite{werlich2015stochastic} but this adds yet another parameter for the user to choose.
+A significant advantage of SAD over SAMC---which the $1/t$-WL and WL
+methods share \emph{after} they have found all the energies---is that
+the schedule for $\gamma$ automatically responds to the choice of bin
+size. SAD should perform similarly over a wide range of bin sizes
+because $\gamma \propto \nicefrac{N_S}{t}$.  As the number of energy
+states $N_S$ found increases (fine binning), the time spent $t$ in
+each bin will decrease with the effect that the convergence should be
+roughly independent of the bin size chosen.  SAMC could be used with a
+prefactor $\gamma_0$ to aid in a similar
+way~\cite{werlich2015stochastic} but this adds yet another parameter
+for the user to choose.
 
@@
                $\delta_0 = 0.05\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run. The number of important
-               energies is $129$ of about $280$ energies.  The error is
+               as a function of number of iterations run. The error is
                averaged over 8 independent simulations, and the best
@@
                for the same simulations.
-           (c) The average entropy error for each MC method for $N=50$,
-               $\delta_0 = 0.005\sigma$, $\eta = 0.3$, and $T_{\min} = 1/3$
-               as a function of number of iterations run. The number of important
-               energies is $407$ of about $1800$ energies, and
+           (c) The average entropy error for each MC method for the
+               same physical system with a smaller displacement distance
+               $\delta_0 = 0.005\sigma$,
+               as a function of number of iterations run, and
            (d) the update factor $\gamma_t$ versus iteration number
@@
 
-For the square-well fluid, our simulations treat two systems.  The first is
-a smaller simulation with a particle number
-of 50, a well-width of $\lambda = 1.3$, and a volume
-corresponding to a filling fraction of $\eta = 0.3$. The second system
-is larger, with a particle number
-of 256, a well-width of $\lambda = 1.5$, and a volume
-corresponding to a filling fraction of $\eta = 0.17$.
-For each system we use a reasonable displacement distance $\delta_0 = 0.05\sigma$,
-and for the smaller system we also use an unreasonably small displacement
-distance of $0.005\sigma$. The simulations explore the
-energy space of the systems with minimum reduced temperatures of
-$T_{\text{min}} = 1/3$ for simulations of the smaller system, and $T_{\min}=1$
-for the larger system.
-All simulations lead to the minimum important
-energy $E_{\min}$ and maximum entropy energy $E_{\max}$
-being calculated (with the exception of the WL methods where both
-of these parameters are needed a priori).  In the sections below, we
-examine the systems parameters impact on the convergence of the methods.
+We tested the algorithms on two square-well fluid systems.  The first
+is a smaller simulation with a particle number of 50, a well-width of
+$\lambda = 1.3$, and a volume corresponding to a filling fraction
+(defined as the fraction of volume filled by atoms) of $\eta =
+0.3$. The second system is larger, with a particle number of 256, a
+well-width of $\lambda = 1.5$, and a volume corresponding to a filling
+fraction of $\eta = 0.17$.  For each system we use a reasonable
+root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
+proposed moves, and for the smaller system we also use an unreasonably
+small displacement distance of $0.005\sigma$. The simulations explore
+the energy space of the systems with minimum reduced temperatures of
+$T_{\text{min}} = 1/3$ for simulations of the smaller system, and
+$T_{\min}=1$ for the larger system.  All simulations lead to the
+minimum important energy $E_{\min}$ and maximum entropy energy
+$E_{\max}$ being calculated (with the exception of the WL methods
+where both of these parameters are needed a priori).
 
@@
 
-We use the average entropy error vs moves as a metric to compare
+We use the average entropy error versus moves as a metric to compare
 simulation runtimes and overall convergence. The overall accuracy
@@
 from $-248$ to $-120$.  The number of important energy states
-for this system is therefore $N_S = 129$  The entropy of this system is shown in
+for this system is therefore $N_S = 129$.  The entropy of this system is shown in
 Fig.~\ref{fig:entropy-cartoon} above, which shows that over this
@@
 In order to explore the effect of simulation details on convergence,
-we consider two values for the displacement distance by which atoms are moved
-during a Monte Carlo step.  We began with a reasonable displacement
-distance of $\delta_0 = 0.05\sigma$, which corresponds to an
-acceptance rate of 38\%.  We further ran simulations with a much
-smaller displacement distance of $\delta_0 = 0.005\sigma$, which
-resulted in an acceptance rate of 86\%, which as expected converge
-more slowly.
+we consider two values for the displacement distance by which atoms
+are moved during a Monte Carlo step.  We began with a reasonable
+displacement distance of $\delta_0 = 0.05\sigma$, which corresponds to
+an acceptance rate of proposed moves of 38\%.  We further ran
+simulations with a much smaller displacement distance of $\delta_0 =
+0.005\sigma$, which resulted in an acceptance rate of 86\%, which
+converged more slowly.
 
@@
 
-Figure~\ref{fig:fast-slow-gamma}b shows the average error in
+Figure~\ref{fig:fast-slow-gamma}c shows the average error in
 the entropy as a function of time for this system with the
@@
 The methods SAD, WL, and $1/t$-WL compensate for the smaller
-displacement distance as can be seen from Figure~\ref{fig:fast-slow-gamma}.
-The update factors take approximately 10x longer to reach steady-state for
-the slower translation scale. Because of this update behavior, these methods
-are able to converge regardless of the displacement distance used in Monte
-Carlo moves.
+displacement distance by reducing $\gamma$ more slowly, as can be seen from
+Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.
+The update factors take approximately 10x longer to reach steady-state
+for the slower translation scale. Because of this update behavior,
+these methods are less sensitive to the choice of displacement
+distance than SAMC is.  \davidsays{This paragraph could benefit from
+  going through again.}
 
@@
 \subsection{A periodic system with 256 atoms}
-Next we will introduce a considerably larger simulation containing
-256 atoms which has a total entropy range of around 1500.  This makes
-exploring the entire range of energies extremely expensive, and strongly
-favors methods that restrict the energy (or temperature) range of
-interest.
-For this simulation, we chose a much higher minimum reduced
-temperature of $1.0$, which corresponds to an interesting energy range
-from $-915$ to $-509$.  The number of important energy states for this
-system is then $N_S = 407$.
-The entropy of this system differs by 395, corresponding to a ratio of
-$10^{118}$ between the highest and lowest density of states.
+Next we will introduce a considerably larger simulation containing 256
+atoms which has a maximum entropy about 1500 greater than its minimum.
+This makes exploring the entire range of energies extremely expensive,
+and strongly favors the methods that restrict the energy (or
+temperature) range of interest.  For this simulation, we chose a much
+higher minimum reduced temperature of $1.0$, which corresponds to an
+interesting energy range from $-915$ to $-509$.  The number of
+important energy states for this system is then $N_S = 407$.  The
+minimum entropy over this energy range is just 395 less than the
+maximum, corresponding to a ratio of only $10^{118}$ between the
+highest and lowest density of states.
 
@@
 
-Figure~\ref{fig:n256} shows the average error in the
-entropy as a function of moves for this system with the reasonable
-displacement distance of $\delta_0 = 0.05\sigma$.  The solid lines
-represent the average of the absolute value of the error in the
-entropy averaged over eight simulations using different random number
-seeds.  The range of average errors for each simulation is shown as a
-shaded region around its mean error.  By the time $10^8$ moves have
-been made all but the SAMC simulation with the shortest $t_0$ have
-begun to converge as $1/\sqrt{t}$.  We then see the WL error
-saturates around $10^{10}$ moves.
+Figure~\ref{fig:n256} shows the average error in the entropy as a
+function of moves for this system with the reasonable displacement
+distance of $\delta_0 = 0.05\sigma$.  The solid lines represent the
+average of the absolute value of the error in the entropy averaged
+over eight simulations using different random number seeds.  The range
+of average errors for each simulation is shown as a shaded region
+around its mean error.  By the time $10^8$ moves have been made all
+but the SAMC simulation with the shortest $t_0$ shown have begun to
+converge as $1/\sqrt{t}$.  We then see the WL error saturates around
+$10^{10}$ moves.  Once again, the convergence of SAD is essentially
+the same as that of the $1/t$-WL method.
+
+\davidsays{I got to here.}
 

### 2019-05-10 papers/histogram/sad.tex
final edit of paper

@@
 \begin{abstract}
-  \davidsays{Let's look at the abstract together.}
-  We present a novel Monte-Carlo algorithm based on the Stochastic
+  We present a new Monte-Carlo algorithm based on the Stochastic
   Approximation Monte Carlo (SAMC) algorithm for directly calculating
   the density of states. The proposed method is Stochastic
-  Approximation with a dynamic update factor (SAD-$\gamma$ or SAD)
+  Approximation with a Dynamic update factor (SAD)
   which dynamically adjusts the update factor $\gamma$ during the course of
-  the simulation. We test this MC method on the square-well fluid and
-  compare the convergence time and average entropy error for a variety of weight-based
-  Monte-Carlo methods. SAD rapidly converges to the
+  the simulation. We test this method on the square-well fluid and
+  compare the convergence time and average entropy error for several
+  related
+  Monte-Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
+  methods rapidly converge to the
   correct density of states without the need for the user to specify an
-  arbitrary tunable parameter $t_0$ as in the case of SAMC.
+  arbitrary tunable parameter $t_0$ as in the case of SAMC.  SAD requires
+  as input the temperature range of interst, in contrast to
+  $1/t$-WL, which requires that the user identify the interesting range
+  of energies.  Thus SAD is more convenient when the range of energies is
+  not known in advance.
 \end{abstract}
@@
 implemented a schedule that enforced that if the update factor
-$\gamma$ is ever less than $\nicefrac{N_S}{t}$ then the update factor
-is set to $\nicefrac{N_S}{t}$ where $N_S$ is the number of energy
-states and the histogram is no longer tracked. Employing this
+$\gamma$ is ever less than $\nicefrac{1}{t}$ then the update factor
+is set to $\nicefrac{1}{t}$ and the histogram is no longer tracked. Employing this
 scheduler, they found that the error saturation is avoided since the
@@
 to $1/t$-WL when the update factor $\gamma <
-\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. \davidsays{Above
-  we say that 1/t-WL does this.  One must be wrong?} As per the
+\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the
 original $1/t$ implementation~\cite{belardinelli2007fast}, the update
@@
 once (i.e. $H(E) \neq 0$) effectively avoiding the concept of
-`flatness'.
+`flatness'.  In this work, we refer to this refined algorithm as $1/t$-WL,
+with the update factor asymptoting to $\nicefrac{N_S}{t}$.
 
@@
 Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.
-The update factors take approximately 10x longer to reach steady-state
-for the slower translation scale. Because of this update behavior,
+The update factors take approximately 10$\times$ longer to reach steady-state
+for the smaller displacement distance. Because of this update behavior,
 these methods are less sensitive to the choice of displacement
-distance than SAMC is.  \davidsays{This paragraph could benefit from
-  going through again.}
+distance than SAMC is.
 
@@
 
-\davidsays{I got to here.}
-
 %~ For this larger system, SAMC parameter values of $t_0$ that gave rapid
@@
 \section{Conclusions}
-Both SAD and 1/t-WL converge more rapidly than Wl or SAMC. SAD requires
-the user to specify a temperature range rather than an energy range as 1/t-WL
-does. Both WL and 1/t-WL require a prior simulation in order to determine the energy range
-of interest. WL saturates in convergence as discussed in the introduction and
-requires the user to optimize the choice of parameters to avoid this such as
-increasing the flatness criterion.  SAMC converges for a reasonable choice of
-$t_0$ but this parameter can be difficult to tune especially across different
-energy bins, displacement distances, or systems. Additionaly, SAMC does not
-converge as rapidly as either SAD or 1/t-WL even for the best choice of $t_0$.
+We have introduced a new algorithm that effectively samples the
+energy space corresponding to a desired range of temperatures for a
+variety of system sizes. 
+%
+We find that both SAD and $1/t$-WL demonstrate excellent and robust
+convergence.
+They both converge more rapidly than SAMC, and unlike WL consistently
+converge to the correct density of states. SAD requires
+the user to specify a temperature range of interest rather than an
+energy range of interest as $1/t$-WL does.  For use cases in which
+a range of desired temperatures is known, this will make the SAD method
+considerably more convenient.
 
-We have introduced a new algorithm SAD (Stochastic Approximation with a
-Dynamic $\gamma$) that effectively samples the entire energy space for
-a variety of system sizes.  Also, SAD proves effective in comparison
-with other Monte-Carlo methods at converging to the correct density of
-states.  The ease of implementation and the absence of user-defined
-parameters makes SAD an ideal algorithm. SAD converges as quickly as
-SAMC (in the case of an ideally chosen $t_0$ for our simulations. SAD
-doesn't suffer from the weakness of SAMC, i.e. the user choosing
-parameters that could lead to the simulation not converging in a
-reasonable amount of moves.
+We find that SAMC converges for a reasonable choice of
+$t_0$ but this parameter can be difficult to tune especially across
+significantly differing systems.  We find that even simple
+changes to the Monte Carlo moves can have a dramatic effect on the
+range of practical $t_0$ values.  Additionaly, SAMC does not
+converge as rapidly as either SAD or $1/t$-WL even for the best choice of $t_0$,
+when a relatively small range of energies is required, because it always
+simulates all possible energies.
 
-SAD decreases $\gamma$ in a controlled way that prevents the algorithm
-from taking too long to converge (as in the case of SAMC with a poor choice
-of $t_0$) or failing to converge such as is the case with WL for complex
-systems. A major advantage is that the user does not need to determine a
-priori how quickly the update factor $\gamma$ should decrease as the
-method SAD dynamically determines this.
+%We note that SAD should be tested on additional systems, and the
+%convergence of the method should be examined.  Also, a rigorous
+%mathematical justification for our choice of the update factor is
+%necessary to ensure the most optimal convergence of the algorithm.
 
-We note that SAD should be tested on additional systems, and the
-convergence of the method should be examined.  Also, a rigorous
-mathematical justification for our choice of the update factor is
-necessary to ensure the most optimal convergence of the algorithm.
-
-\section{Acknowledgement}
+%\section{Acknowledgement}
 

### 2019-05-15 papers/histogram/ising.tex
added aps sad paper and new ising paper in the right directory for scripts to work well.

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Ising Sad paper
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
In this work, we finish writing everything before the abstract~\jpsays{FIX AT END}
\end{abstract}

\maketitle

\section{Introduction}
\jpsays{this needs a total rewrite}
In the early 2000s, Wang and Landau introduced a novel flat-histogram
Monte-Carlo algorithm that accurately determined the density of states
(DOS) for a statistical
system~\cite{wang2001determining,wang2001efficient}. For all of its
power, the method unfortunately requires a priori knowledge of several
user-defined parameters. Thus, for any given system under study, the
user needs to determine the ideal parameters in order to apply the
method. The Wang-Landau algorithm is also known to violate detailed
balance (although only for brief time intervals)~\cite{yan2003fast,
shell2002generalization}. With the violation of detailed balance,
convergence of the algorithm is not guaranteed. The convergence rate
for WL is well known~\cite{zhou2005understanding,lee2006convergence,
belardinelli2007wang}. Modification factors that decrease faster than
$1/t$ lead to non-convergence~\cite{belardinelli2007fast}.  These
findings lead to the formation of the $1/t$-WL algorithm.  The ideal
choice of decreasing the modification factor prevents CPU time from
being wasted by continuing to perform calculations after the error in
the density of states is saturated~\cite{belardinelli2008analysis}.
Liang independently considered whether WL could be treated as a special
case of Stochastic Approximation whose convergence could be
mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
considered a form of Stochastic Approximation Monte Carlo (SAMC).
Unlike WL, SAMC can guarantee convergence (if certain conditions are
met). Despite the added benefit of guaranteed convergence, the method
still has a system specific user-defined variable. Such variables often
create difficulty when applying Monte-Carlo methods across arbitrary
systems.  The SAD (Stochastic Approximation with Dynamic $\gamma$)
method~\jpsays{outlined by Pommerenck et.al include David and my paper
on SAD square well in citations!} is special version of the SAMC
algorithm that dynamically chooses the modification factor rather than
relying on system dependent parameters. SAD shares the same convergence
properties with SAMC while replacing un-physical user-defined parameters
with the algorithms dynamic choice.  In this work, we apply the family
of Monte-Carlo algorithms (WL, 1/t-WL, SAMC, SAD) to the 2D Potts andbinning of discrete energies.

%\section{Discrete Spin Systems} In this work, we compare the
%convergence properties of the family of Weight-Based Broad Histogram
%Methods applied to the Q-State Potts model and Ising model. The two
%classical spin systems are ideal testbeds for comparing the convergence
%of Monte-Carlo methods. The Potts and Ising models have 1st and 2nd
%order phase transitions respectively and will both require energy
%binning due to the discrete energy levels of the systems. Convergence
%scaling for the family of Weight-Based Broad Histogram methods can
%easily be explored on the two non-continuous systems.

%\subsection{Potts Model}
%We test the convergence of several Broad Histogram methods
%on the periodic 2D square lattice ($Q=8$) Potts model with nearest
%neighbor interactions~\cite{landau2004new}. The $N\times N$ spin system can
%take on values of $q_i = 1,2,3...Q$ for the Potts spin at site $i$. We
%can write the Hamiltonian as follows:

%\begin{align}
%\mathcal{H} = \sum_{<i,j>} \delta(q_i,q_j)
%\end{align}

%The 2D ten state (Q=10) Potts Model was used as a test system for the Wang
%Landau method during its inception~\cite{wang2001determining,wang2001efficient}.
%Some of the popularity of the model is due to knowledge of exact solutions
%and access to extensive simulation data~\cite{wu1982potts}.
%\jpsays{
%(1) The latent heat of the system which can be obtained by plotting U vs T
%and looking at how much U drops at Tc.
%(2) Looking at the value for the critical temperature Tc.
%Or it is possible that they clarify in a later work~\cite{landau2004new} how
%to extract the correct density of states (Q = 8) in this work.
%\begin{align}
%\sum_E g_N\left( E\right)=Q^N
%\end{align}
%Also the number of ground states (where $E = -2N$) is $Q$. He then re-scales
%the density of states $g_N\left( E\right)$ and writes
%\begin{align}
%\ln \left[ g_N\left(E\right)\right]=\ln \left[ g\left( E\right)\right]
%-\ln \left[ g\left(E=-2N\right)\right]+\ln\left[ Q \right]
%\end{align}
%He calls this the normalized density of states (for his work Q = 10). The
%transition temperature for a 2D lattice $k_B T_C\left(L\right)$ can be directly
%calculated~\cite{landau2004new} and compared with the known value,~\cite{wu1982potts}
%\begin{align}
%k_B T_C = \frac{1}{\ln\left(1+(Q)^{0.5}\right)}
%\end{align}
%Mention: 1st order phase transition nearest neighbor and Hamiltonian
%}

\subsection{Ising Model}
% Flat Histogram Method of Wang-Landau and N-Fold Way (Schulz Binder Muller)
% https://doi.org/10.1142/S0129183102003243
\jpsays{talk about papers that covered Ising model in context of flat histogram
as an introduction to this.}
We test the convergence of several flat histogram methods
on the periodic 2D square lattice ferromagnetic Ising model with nearest
neighbor interactions~\cite{landau2004new}. The $N\times N$ spin system can
take on values of $\sigma_i = \pm 1$ for up or down spins respectively. We
can write the Hamiltonian as follows:
\begin{align}
\mathcal{H} = -\sum_{<i,j>} \sigma_i \sigma_j
\end{align}
where the sum is over all enumerated spin sites. The 2D Ising spin-lattice
system is widely used as a testbed when benchmarking or comparing Monte-Carlo
methods~\cite{ferdinand1969bounded, wang1999transition}. The
2nd order phase transition behavior and the ability to directly calculate the
exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently
interesting for such theoretical comparisons. It is also of important note that
direct comparison of the other methods can be made with WL as its original implementation
was done on this system~\cite{wang2001determining,wang2001efficient}.

\subsection{Wang-Landau}
Wang-Landau here.

\subsection{$1/t$-Wang-Landau}
1/t WL here.

\subsection{SAMC}
Stochastic Approximation Monte Carlo here.

\subsection{SAD Algorithm}\label{sec:sad}
Stochastic Approximation w/ Dynamic $\gamma$ here.

\section{Results}

\subsection{The 16 $\times$ 16 Ising model}
Something small...
\subsection{128 $\times$ 128 Ising system}
Something big (WL uses 256 by 256 in their paper but only one simulation and
no comparison while more recent comparison type papers use 128 by 128 as max)...

\section{Conclusion}

\section{Acknowledgements}

\bibliography{ising} % Produces the bibliography via BibTeX.

\end{document}

### 2019-05-15 papers/histogram/sad.tex
shorten title
added aps sad paper and new ising paper in the right directory for scripts to work well.

@@
 \title{Stochastic Approximation Monte Carlo with a Dynamic Update
-Factor (SAD)
+Factor
 }
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/N50-lndos-comparison}
+  \includegraphics[width=\columnwidth]{N50-lndos-comparison}
   \caption{The entropy of a square well fluid with 50 atoms and filling fraction
@@
 \begin{figure}
-  \includegraphics[width=\columnwidth]{figs/gamma-n256}
+  \includegraphics[width=\columnwidth]{gamma-n256}
   \caption{The update factor $\gamma_t$ versus iteration for $N=256$
@@
 \begin{figure*}
-  \includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
-\includegraphics[width=0.5\textwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
+  \includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-default.pdf}%
+\includegraphics[width=0.5\textwidth]{periodic-ww1_30-ff0_30-N50-gamma-and-entropy-error-slow.pdf}
   \caption{(a) The average entropy error for each MC method for $N=50$,
@@
 
-%\begin{figure}
-%  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-default.pdf}
-%  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-%  a fixed energy range SAMC $t_0 = 10^3$ serving as the reference.}\label{fig:N50-ff0.3-avg-error}
-%\end{figure}
-%\begin{figure}
-%  \includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_30-ff0_30-N50-entropy-error-slow.pdf}
-%  \caption{The average entropy error for each MC method is shown for $\eta = 0.3$ and $T_{\min} = 1/3$ with
-%  a fixed energy range SAMC $t_0 = 10^3$ from Fig.~\ref{fig:N50-ff0.3-avg-error}
-%  serving as the reference.}\label{fig:N50-ff0.3-avg-error-slow}
-%\end{figure}
-
 Figure~\ref{fig:fast-slow-gamma}a shows the average error in the
@@
 \begin{figure}
-\includegraphics[width=\columnwidth]{figs/s000/periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
+\includegraphics[width=\columnwidth]{periodic-ww1_50-ff0_17-N256-entropy-error-default.pdf}
   \caption{
@@
 \end{figure}
-%~ \begin{figure}
-  %~ \includegraphics[width=\columnwidth]{figs/N256-lndos-comparison}
-  %~ \caption{\davidsays{This could be prettier, and do we need it? We
-      %~ definitely need a caption if we keep it.} Use this to create
-    %~ description of the system.}\label{fig:n256-lndos-comparison}
-%~ \end{figure}
 

### 2019-05-20 papers/histogram/ising.tex
working on writing the ising paper.

@@
 \usepackage{mdframed}
+\usepackage{braket}
 
@@
 \subsection{Ising Model}
-% Flat Histogram Method of Wang-Landau and N-Fold Way (Schulz Binder Muller)
-% https://doi.org/10.1142/S0129183102003243
-\jpsays{talk about papers that covered Ising model in context of flat histogram
-as an introduction to this.}
+The 2D Ising spin-lattice system is widely used as a testbed when
+benchmarking or comparing Monte-Carlo
+methods~\cite{ferdinand1969bounded, wang1999transition}. The 2nd order
+phase transition behavior and the ability to directly calculate the
+exact solution for finite lattices~\cite{beale1996exact} make the
+system sufficiently interesting for such theoretical comparisons. It is
+also of important note that direct comparison of the other methods can
+be made with WL as its original implementation was done on this
+system~\cite{wang2001determining,wang2001efficient}.
 We test the convergence of several flat histogram methods
 on the periodic 2D square lattice ferromagnetic Ising model with nearest
-neighbor interactions~\cite{landau2004new}. The $N\times N$ spin system can
-take on values of $\sigma_i = \pm 1$ for up or down spins respectively. We
-can write the Hamiltonian as follows:
+neighbor interactions~\cite{landau2004new}.
 \begin{align}
-\mathcal{H} = -\sum_{<i,j>} \sigma_i \sigma_j
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
 \end{align}
-where the sum is over all enumerated spin sites. The 2D Ising spin-lattice
-system is widely used as a testbed when benchmarking or comparing Monte-Carlo
-methods~\cite{ferdinand1969bounded, wang1999transition}. The
-2nd order phase transition behavior and the ability to directly calculate the
-exact solution for finite lattices~\cite{beale1996exact} make the system sufficiently
-interesting for such theoretical comparisons. It is also of important note that
-direct comparison of the other methods can be made with WL as its original implementation
-was done on this system~\cite{wang2001determining,wang2001efficient}.
+The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
+for up or down spins respectively. In the infinite volume limit ($h =
+0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
+kaufman1949crystal}:
+\begin{align}
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
+\end{align}
+where the sum is over all enumerated spin sites. Beale showed that for finite
+lattices the Density of States could directly be calculated from the partition
+function~\cite{beale1996exact}
+\begin{align}
+Z = \sum_E g(E) e^{-{\beta E}}
+\end{align}
+where $g(E)$ is the multiplicity of the system which is identical to
+$D(E)$. We can estimate the average deviation from the exact solution
+using the relation~\cite{schneider2017convergence, shakirov2018convergence,
+barash2017control}:
+\begin{align}
+\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
+\end{align}
+As per the implementation by Wang and
+Landau~\cite{wang2001determining}, the total number of available energy
+states are $N-1$.
+\jpsays{Also the necessary condition that the ground state is
+$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
+density of states at low temperatures. If $2^N$ states is the condition
+used for the ferromagnetic ising model, the rescaled factor is
+determined primarily by the maximum density of states impacing the
+accuracy of the energy levels at or near the ground state.}
 
 \subsection{Wang-Landau}
-Wang-Landau here.
+The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
+landau2014guide} explores energy space by initially setting $\gamma=1$,
+and then decreases $\gamma$ in discrete stages.  We track the number of
+moves ending at each energy and store the counts in a histogram.  For a
+sufficiently flat energy histogram, $\gamma$ is decreased by a specified factor of
+$\frac12$.  Flatness is defined by the ratio between the minimum
+value of the histogram and its average value.  If this flatness
+reaches a specified threshold (typically 0.8), the $\gamma$ value is
+decreased and the histogram is reset to zero. The entire process is repeated
+until $\gamma$ reaches a desired cutoff.
+
+%The Wang-Landau approach thus has three parameters that need be
+%specified: the factor by which to decrease $\gamma$ when flatness is
+%acheived, the flatness criterion, and the cutoff that determines when
+%the computation is complete.  In addition, an energy range (or in
+%general, a \emph{set} of energies) must be supplied, so that the
+%flatness criterion can be defined.
+
+Although the algorithm is both efficient and widely used , it
+suffers from a few shortcomings.  Frist, an energy range of interest must be
+specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
+which may require multiple simulations. Second, while Wang-Landau
+converges quickly, it does not in general converge to the true density
+of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
+and does decrease $\gamma$ so quickly that it will never (for any
+cutoff value) decrease the error in the density of states beyond a
+given nonzero value.
 

### 2019-05-20 papers/histogram/ising.tex
working on adding some text about methods to ising paper.

@@
 
-%The Wang-Landau approach thus has three parameters that need be
-%specified: the factor by which to decrease $\gamma$ when flatness is
-%acheived, the flatness criterion, and the cutoff that determines when
-%the computation is complete.  In addition, an energy range (or in
-%general, a \emph{set} of energies) must be supplied, so that the
-%flatness criterion can be defined.
-
 Although the algorithm is both efficient and widely used , it
@@
 \subsection{$1/t$-Wang-Landau}
-1/t WL here.
+The $1/t$-WL method is a refinement of the original implementation that
+addresses the error saturation of WL. Belardinelli and Pereyra
+implemented a schedule~\cite{belardinelli2007fast} that enforced that
+if $\gamma < \nicefrac{1}{t}$ then the update factor is set to
+$\nicefrac{1}{t}$ and the histogram is no longer tracked. Using this
+metric, the error saturation is avoided since the correct density of
+states is approached asymptotically as
+$t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al also showed
+that the WL algorithm never converges exponentially and succesfully
+bounded the statistical error between $t^{-\frac12}$ and
+$1/t$~\cite{zhou2008optimal}.
+
+Schneider $\text{et. al}$ outlines minor refinements to the $1/t$-WL algorithm 
+including $\nicefrac{N_S}{t}$ scaling and switching from standard WL to 
+$1/t$-WL when the update factor $\gamma < 
+\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the original 
+$1/t$ implementation~\cite{belardinelli2007fast}, the update factor is 
+decreased once all $N_S$ energy states are visited at least once (i.e. 
+$H(E) \neq 0$) effectively avoiding the concept of `flatness'.  In this 
+work, we refer to this refined algorithm as $1/t$-WL, with the update 
+factor asymptoting to $\nicefrac{N_S}{t}$.
 
 \subsection{SAMC}
-Stochastic Approximation Monte Carlo here.
+The Stochastic Approximation Monte Carlo (SAMC) algorithm has a simple
+schedule by which the update factor $\gamma$ is continuously
+decreased~\cite{liang2007stochastic, werlich2015stochastic,
+schneider2017convergence}.  The update factor is defined in the
+original implementation~\cite{liang2007stochastic} in terms of an
+arbitrary tunable parameter $t_0$,
+\begin{align}
+\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
+\end{align}
+where as above $t$ is the number of moves that have been attempted.
+SAMC offers extreme simplicity and proven convergence.
+Provided the update factor satisfies
+\begin{align}
+\sum_{t=1}^\infty \gamma_{t} = \infty \quad\textrm{and}\quad
+\sum_{t=1}^\infty \gamma_{t}^\zeta < \infty
+\end{align}
+where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
+to converge to the true density of states~\cite{liang2006theory,
+liang2007stochastic}.  Unlike WL methods, the energy range need not be
+known \emph{a priori.}  The convergenc time depends only on the choice of
+parameter $t_0$.  Liang \emph{et al.} give a rule of thumb in
+which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
+is the number of energy bins~\cite{liang2007stochastic}.  Schneider
+\emph{et al.} found that for the Ising model this heuristic is helpful
+for small spin systems, but that larger systems require an even higher
+$t_0$ value~\cite{schneider2017convergence}.
 
 \subsection{SAD Algorithm}\label{sec:sad}
+In our previous work we discuss this in detail so here we only highlight...
 Stochastic Approximation w/ Dynamic $\gamma$ here.

### 2019-05-29 papers/histogram/ising.tex
some paper updates still working on intro

@@
 
-\title{Ising Sad paper
+\title{Flat histogram method comparison on 2D Ising Model
 }
@@
 \section{Introduction}
-\jpsays{this needs a total rewrite}
+\jpsays{What is a flat histogram}
+Flat histogram Monte-Carlo simulation algorithms calculate the
+thermodynamic properties of various systems for over a range of
+temperatures.  The first histogram method used a single canonical Monte
+Carlo simulation to predict properties for nearby
+temperatures~\cite{ferrenberg1988new}. This method is unable to
+effectively sample large energy ranges where the simulation is limited
+to a narrow temperature range.
+
+This led to a
+variety of ``flat'' (or ``broad'') histogram
+methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
+wang2001determining, wang2001efficient, trebst2004optimizing}, which
+explore a wider range of energies.  Another benefit, in contrast with
+low-temperature canonical Monte Carlo, is that these approaches cannot
+be trapped in a local energy minimum.
+
+\jpsays{What is WL}
 In the early 2000s, Wang and Landau introduced a novel flat-histogram
@@
 the density of states is saturated~\cite{belardinelli2008analysis}.
+
+\jpsays{What is SAMC}
 Liang independently considered whether WL could be treated as a special
@@
 with the algorithms dynamic choice.  In this work, we apply the family
-of Monte-Carlo algorithms (WL, 1/t-WL, SAMC, SAD) to the 2D Potts andbinning of discrete energies.
+of Monte-Carlo algorithms (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
 
@@
 %}
+\section{Flat histogram methods}\label{sec:histogram}
+Flat histogram methods determine the density of states $D(E)$ over a broad range
+of energies by simulating each energy with equivalent accuracy.
+
+Flat histogram Monte Carlo methods propose randomly chosen ``moves''
+which change the state of the system and must satisfy detailed balance.  Each
+algorithm differs in how it determines the probability of accepting a move and
+in what additional statistics must be collected in order decide on that
+probability.
+
+We describe four closely related flat histogram methods which each rely on a
+weight function $w(E)$ to determine $D(E)$.  For these algorithms, the
+probability of accepting a move is given by
+\begin{equation}
+	\mathcal{P}(E_\text{old} \rightarrow E_\text{new})
+	= \min\left[1,\frac{w(E_\text{old})}{w(E_\text{new})}\right]
+\end{equation}
+which biases the simulation in favor of energies with low weights. The result of
+weights $w(E)$ that are proportional to $D(E)$ is an entirely flat histogram. We
+can relate the entropy to the weights in the microcanonical ensemble, since the
+entropy is defined as $S(E) \equiv k_B\ln(D(E)) \approx \ln w(E)$.
+
+Each approach uses a random walk in energy space to estimate $D(E)$.  At the
+core of each of these approaches is to continuously update the weights at each
+step of the simulation
+\begin{equation}
+	\ln{w_{t+1}(E)}=\ln{w_{t}(E)}
+	+\gamma_t
+\end{equation}
+where $t$ is number of the current move, $\gamma(t)$ is a move-dependent update
+factor, and $E$ is the current energy.  This update causes the random walk to
+avoid energies that have been frequently sampled, leading to a rapid exploration
+of energy space. The four methods differ primarily in how they schedule the
+decreasing of $\gamma_t$.
 

### 2019-06-13 papers/histogram/ising.tex
added some text to paper and sh script to generate data.

@@
 
-\subsection{The 16 $\times$ 16 Ising model}
+\subsection{The 32 $\times$ 32 Ising model}
 Something small...
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/N32-entropy-error-default.pdf}
+  \caption{
+  The average entropy error for each MC method for $N=256$,
+               $\delta_0 = 0.05\sigma$, $\eta = 0.17$, and $T_{\min} = 1$
+               as a function of number of iterations run.  The error is
+               averaged over 8 independent simulations, and the best
+               and worst simulations for each method are shown as a
+               semi-transparent shaded area.  The update factor for this
+               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
+\end{figure}
+
 \subsection{128 $\times$ 128 Ising system}

### 2019-06-25 papers/histogram/sad.tex
sad paper edit and aps figure update script.

@@
 
-Schneider et. al outlines minor refinements to the $1/t$-WL algorithm
+Schneider et. al outline minor refinements to the $1/t$-WL algorithm
 including $\nicefrac{N_S}{t}$ scaling and switching from standard WL

### 2019-06-25 papers/histogram/ising.tex
add text to ising paper and gamma plot figs.

--- 2019-05-29 papers/histogram/ising.tex
@@
 \section{Introduction}
-\jpsays{What is a flat histogram}
-Flat histogram Monte-Carlo simulation algorithms calculate the
-thermodynamic properties of various systems for over a range of
-temperatures.  The first histogram method used a single canonical Monte
-Carlo simulation to predict properties for nearby
-temperatures~\cite{ferrenberg1988new}. This method is unable to
-effectively sample large energy ranges where the simulation is limited
-to a narrow temperature range.
+Flat histogram Monte-Carlo simulation algorithms calculate the thermodynamic
+properties of various systems over a range of temperatures.  The first histogram
+method used a single canonical Monte Carlo simulation to predict properties for
+nearby temperatures~\cite{ferrenberg1988new}. While the method effectively
+samples a narrow energy range given limited by a narrow temperature range, it
+proves computationally inefficient at sampling large energy ranges.
 
-This led to a
-variety of ``flat'' (or ``broad'') histogram
+Due to a canonical Monte Carlo simulation's inefficient sampling of large
+energies, a variety of ``flat'' (or ``broad'') histogram
 methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-wang2001determining, wang2001efficient, trebst2004optimizing}, which
-explore a wider range of energies.  Another benefit, in contrast with
-low-temperature canonical Monte Carlo, is that these approaches cannot
-be trapped in a local energy minimum.
+wang2001determining, wang2001efficient} were developed to explore a wider range
+of energies.  In addition to obtaining thermodynamic information for the entire
+energy range for a single simulation, these approaches cannot be easily trapped
+in a local energy minimum like a canonical simulation.
 
-\jpsays{What is WL}
-In the early 2000s, Wang and Landau introduced a novel flat-histogram
-Monte-Carlo algorithm that accurately determined the density of states
-(DOS) for a statistical
-system~\cite{wang2001determining,wang2001efficient}. For all of its
-power, the method unfortunately requires a priori knowledge of several
-user-defined parameters. Thus, for any given system under study, the
-user needs to determine the ideal parameters in order to apply the
-method. The Wang-Landau algorithm is also known to violate detailed
-balance (although only for brief time intervals)~\cite{yan2003fast,
-shell2002generalization}. With the violation of detailed balance,
-convergence of the algorithm is not guaranteed. The convergence rate
-for WL is well known~\cite{zhou2005understanding,lee2006convergence,
-belardinelli2007wang}. Modification factors that decrease faster than
-$1/t$ lead to non-convergence~\cite{belardinelli2007fast}.  These
-findings lead to the formation of the $1/t$-WL algorithm.  The ideal
-choice of decreasing the modification factor prevents CPU time from
-being wasted by continuing to perform calculations after the error in
-the density of states is saturated~\cite{belardinelli2008analysis}.
+Wang and Landau introduced one of the most widely used flat-histogram
+Monte-Carlo algorithms that accurately determined the density of states (DOS) for
+a statistical system~\cite{wang2001determining,wang2001efficient}. For all of
+its power, the method unfortunately requires a priori knowledge of several
+user-defined parameters. Thus, for any given system under study, the user needs
+to determine the ideal parameters in order to apply the method. The Wang-Landau
+algorithm is also known to violate detailed balance (although only for brief
+time intervals)~\cite{yan2003fast, shell2002generalization}. With the violation
+of detailed balance, convergence of the algorithm is not guaranteed.
 
-\jpsays{What is SAMC}
-Liang independently considered whether WL could be treated as a special
-case of Stochastic Approximation whose convergence could be
-mathematically proven~\cite{liang2006theory, liang2007stochastic}. In
-2007, Liang et al.~\cite{liang2007stochastic} argued that WL can be
-considered a form of Stochastic Approximation Monte Carlo (SAMC).
-Unlike WL, SAMC can guarantee convergence (if certain conditions are
-met). Despite the added benefit of guaranteed convergence, the method
-still has a system specific user-defined variable. Such variables often
-create difficulty when applying Monte-Carlo methods across arbitrary
-systems.  The SAD (Stochastic Approximation with Dynamic $\gamma$)
-method~\jpsays{outlined by Pommerenck et.al include David and my paper
-on SAD square well in citations!} is special version of the SAMC
-algorithm that dynamically chooses the modification factor rather than
-relying on system dependent parameters. SAD shares the same convergence
-properties with SAMC while replacing un-physical user-defined parameters
-with the algorithms dynamic choice.  In this work, we apply the family
-of Monte-Carlo algorithms (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
+Because of the uncertainty of convergence for WL, many studies were undertaken
+to understand how the modification (or update) factor $\gamma$ impacts the
+convergence~\cite{zhou2005understanding,lee2006convergence,
+belardinelli2007wang}. The ideal choice of decreasing the modification factor
+helps to prevent CPU time from being wasted by continuing to perform
+calculations after the error in the density of states is
+saturated~\cite{belardinelli2008analysis}. Modification factors that decrease
+faster than $1/t$ were found to lead to
+non-convergence~\cite{belardinelli2007fast}.  These findings led to the
+formation of the $1/t$-WL algorithm.
 
+Liang independently considered whether WL could be treated as a special case of
+Stochastic Approximation whose convergence could be mathematically
+proven~\cite{liang2006theory, liang2007stochastic}. In 2007, Liang et
+al.~\cite{liang2007stochastic} argued that WL can be considered a form of
+Stochastic Approximation Monte Carlo (SAMC). Unlike WL, SAMC can guarantee
+convergence (if certain conditions are met). Despite the added benefit of
+guaranteed convergence, the method still has a system specific user-defined
+variable. Such variables often create difficulty when applying Monte-Carlo
+methods across arbitrary systems.  The SAD (Stochastic Approximation with
+Dynamic $\gamma$) method~\jpsays{outlined by Pommerenck et.al include David and
+my paper on SAD square well in citations!} is a special version of the SAMC
+algorithm that dynamically chooses the modification factor rather than relying
+on system dependent parameters. SAD shares the same convergence properties with
+SAMC while replacing un-physical user-defined parameters with the algorithms
+dynamic choice.  In this work, we apply the family of weight-based flat
+histogram Monte Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
+
+\section{Ising Model}
+The 2D Ising spin-lattice system is widely used as a testbed when
+benchmarking or comparing Monte-Carlo
+methods~\cite{ferdinand1969bounded, wang1999transition, trebst2004optimizing}. The 2nd order
+phase transition behavior and the ability to directly calculate the
+exact solution for finite lattices~\cite{beale1996exact} make the
+system sufficiently interesting for such theoretical comparisons. It is
+also of important note that direct comparison of the other methods can
+be made with WL as its original implementation was done on this
+system~\cite{wang2001determining,wang2001efficient}.
+We test the convergence of several flat histogram methods
+on the periodic 2D square lattice ferromagnetic Ising model with nearest
+neighbor interactions~\cite{landau2004new}.
+\begin{align}
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
+\end{align}
+The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
+for up or down spins respectively. In the infinite volume limit ($h =
+0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
+kaufman1949crystal}:
+\begin{align}
+\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
+\end{align}
+where the sum is over all enumerated spin sites. Beale showed that for finite
+lattices the Density of States could directly be calculated from the partition
+function~\cite{beale1996exact}
+\begin{align}
+Z = \sum_E g(E) e^{-{\beta E}}
+\end{align}
+where $g(E)$ is the multiplicity of the system which is identical to
+$D(E)$. We can estimate the average deviation from the exact solution
+using the relation~\cite{schneider2017convergence, shakirov2018convergence,
+barash2017control}:
+\begin{align}
+\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
+\end{align}
+As per the implementation by Wang and
+Landau~\cite{wang2001determining}, the total number of available energy
+states are $N-1$.
+\jpsays{Also the necessary condition that the ground state is
+$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
+density of states at low temperatures. If $2^N$ states is the condition
+used for the ferromagnetic ising model, the rescaled factor is
+determined primarily by the maximum density of states impacing the
+accuracy of the energy levels at or near the ground state.}
 %\section{Discrete Spin Systems} In this work, we compare the
@@
 Flat histogram methods determine the density of states $D(E)$ over a broad range
-of energies by simulating each energy with equivalent accuracy.
-
-Flat histogram Monte Carlo methods propose randomly chosen ``moves''
-which change the state of the system and must satisfy detailed balance.  Each
-algorithm differs in how it determines the probability of accepting a move and
-in what additional statistics must be collected in order decide on that
-probability.
+of energies by simulating each energy with equivalent accuracy. Flat histogram
+Monte Carlo methods propose randomly chosen ``moves'' which change the state of
+the system and must satisfy detailed balance.  Each algorithm differs in how it
+determines the probability of accepting a move and in what additional statistics
+must be collected in order decide on that probability.
 
@@
 
-\subsection{Ising Model}
-The 2D Ising spin-lattice system is widely used as a testbed when
-benchmarking or comparing Monte-Carlo
-methods~\cite{ferdinand1969bounded, wang1999transition}. The 2nd order
-phase transition behavior and the ability to directly calculate the
-exact solution for finite lattices~\cite{beale1996exact} make the
-system sufficiently interesting for such theoretical comparisons. It is
-also of important note that direct comparison of the other methods can
-be made with WL as its original implementation was done on this
-system~\cite{wang2001determining,wang2001efficient}.
-We test the convergence of several flat histogram methods
-on the periodic 2D square lattice ferromagnetic Ising model with nearest
-neighbor interactions~\cite{landau2004new}.
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j - h \sum_i s_i
-\end{align}
-The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
-for up or down spins respectively. In the infinite volume limit ($h =
-0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
-kaufman1949crystal}:
-\begin{align}
-\mathcal{H} = -\sum_{\braket{i,j}} \sigma_i \sigma_j
-\end{align}
-where the sum is over all enumerated spin sites. Beale showed that for finite
-lattices the Density of States could directly be calculated from the partition
-function~\cite{beale1996exact}
-\begin{align}
-Z = \sum_E g(E) e^{-{\beta E}}
-\end{align}
-where $g(E)$ is the multiplicity of the system which is identical to
-$D(E)$. We can estimate the average deviation from the exact solution
-using the relation~\cite{schneider2017convergence, shakirov2018convergence,
-barash2017control}:
-\begin{align}
-\braket{\epsilon(t)}_E = \frac{1}{N_E - 1} \sum_E \left( \frac{S(E,t) - S_{\text{Beale}(E)}}{S_{\text{Beale}(E)}} \right)
-\end{align}
-As per the implementation by Wang and
-Landau~\cite{wang2001determining}, the total number of available energy
-states are $N-1$.
-\jpsays{Also the necessary condition that the ground state is
-$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
-density of states at low temperatures. If $2^N$ states is the condition
-used for the ferromagnetic ising model, the rescaled factor is
-determined primarily by the maximum density of states impacing the
-accuracy of the energy levels at or near the ground state.}
+The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
+landau2014guide} explores energy space by initially setting $\gamma=1$, and then
+decreases $\gamma$ in discrete stages. An energy range of interest must be
+specified~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
+which can require multiple simulations if unknown.
 
-\subsection{Wang-Landau}
-The Wang-Landau algorithm~\cite{wang2001efficient,wang2001determining,
-landau2014guide} explores energy space by initially setting $\gamma=1$,
-and then decreases $\gamma$ in discrete stages.  We track the number of
-moves ending at each energy and store the counts in a histogram.  For a
-sufficiently flat energy histogram, $\gamma$ is decreased by a specified factor of
-$\frac12$.  Flatness is defined by the ratio between the minimum
-value of the histogram and its average value.  If this flatness
-reaches a specified threshold (typically 0.8), the $\gamma$ value is
-decreased and the histogram is reset to zero. The entire process is repeated
-until $\gamma$ reaches a desired cutoff.
+The number (``counts'') of moves ending at each energy are stored in a
+histogram.  For a sufficiently flat energy histogram (specified by the user to
+typically be 0.8), $\gamma$ is decreased by a specified factor of $\frac12$ and
+the histogram is reset to zero. The entire process is repeated until $\gamma$
+reaches a desired cutoff.
 
-Although the algorithm is both efficient and widely used , it
-suffers from a few shortcomings.  Frist, an energy range of interest must be
-specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
-which may require multiple simulations. Second, while Wang-Landau
-converges quickly, it does not in general converge to the true density
-of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
-and does decrease $\gamma$ so quickly that it will never (for any
-cutoff value) decrease the error in the density of states beyond a
-given nonzero value.
+Zhou et. al also showed that the WL algorithm never converges
+exponentially~\cite{zhou2008optimal} and succesfully bounded the statistical
+error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}. He also suggests
+that a modification factor of $1/t$ is optimal for WL.
 
-\subsection{$1/t$-Wang-Landau}
 The $1/t$-WL method is a refinement of the original implementation that
@@
 states is approached asymptotically as
-$t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al also showed
-that the WL algorithm never converges exponentially and succesfully
-bounded the statistical error between $t^{-\frac12}$ and
-$1/t$~\cite{zhou2008optimal}.
+$t^{-\frac12}$~\cite{belardinelli2008analysis}.
 
-Schneider $\text{et. al}$ outlines minor refinements to the $1/t$-WL algorithm
+\jpsays{I am here!}
+
+Schneider $\text{et. al}$ outline minor refinements to the $1/t$-WL algorithm
 including $\nicefrac{N_S}{t}$ scaling and switching from standard WL to
@@
 \section{Results}
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/gamma-n32.pdf}
+  \caption{
+  gamma plot}
+\end{figure}
 
-\subsection{The 16 $\times$ 16 Ising model}
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
+  \caption{
+  gamma plot}
+\end{figure}
+
+\subsection{The 32 $\times$ 32 Ising model}
 Something small...
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/N32-entropy-error-default.pdf}
+  \caption{
+  The average entropy error for each MC method for $N=32 \times 32$,
+               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+               as a function of number of iterations run.  The error is
+               averaged over 8 independent simulations, and the best
+               and worst simulations for each method are shown as a
+               semi-transparent shaded area.}\label{fig:n32}
+\end{figure}
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/N128-entropy-error-default.pdf}
+  \caption{
+  The average entropy error for each MC method for $N=128 \times 128$,
+               $\delta_0 = 0.05\sigma$, and $T_{\min} = 1$
+               as a function of number of iterations run.  The error is
+               averaged over 8 independent simulations, and the best
+               and worst simulations for each method are shown as a
+               semi-transparent shaded ar.}\label{fig:n128}
+\end{figure}
+
 \subsection{128 $\times$ 128 Ising system}

### 2019-07-03 papers/histogram/ising.tex
working on ising paper currently on sad section.

@@
 methods across arbitrary systems.  The SAD (Stochastic Approximation with
-Dynamic $\gamma$) method~\jpsays{outlined by Pommerenck et.al include David and
-my paper on SAD square well in citations!} is a special version of the SAMC
-algorithm that dynamically chooses the modification factor rather than relying
-on system dependent parameters. SAD shares the same convergence properties with
-SAMC while replacing un-physical user-defined parameters with the algorithms
-dynamic choice.  In this work, we apply the family of weight-based flat
-histogram Monte Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
+Dynamic $\gamma$) method outlined by Pommerenck
+et.al~\cite{pommerenck2019stochastic} is a special version of the SAMC algorithm
+that dynamically chooses the modification factor rather than relying on system
+dependent parameters. SAD shares the same convergence properties with SAMC while
+replacing un-physical user-defined parameters with the algorithms dynamic
+choice.  In this work, we apply the family of weight-based flat histogram Monte
+Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
 
@@
 $t^{-\frac12}$~\cite{belardinelli2008analysis}.
-
-\jpsays{I am here!}
-
 Schneider $\text{et. al}$ outline minor refinements to the $1/t$-WL algorithm
@@
 decreased once all $N_S$ energy states are visited at least once (i.e.
-$H(E) \neq 0$) effectively avoiding the concept of `flatness'.  In this
-work, we refer to this refined algorithm as $1/t$-WL, with the update
-factor asymptoting to $\nicefrac{N_S}{t}$.
+$H(E) \neq 0$) effectively avoiding the concept of `flatness'.
 
-\subsection{SAMC}
-The Stochastic Approximation Monte Carlo (SAMC) algorithm has a simple
+Another weight-based flat histogram method is
+the Stochastic Approximation Monte Carlo (SAMC) algorithm. SAMC has a simple
 schedule by which the update factor $\gamma$ is continuously
@@
 where as above $t$ is the number of moves that have been attempted.
+
 SAMC offers extreme simplicity and proven convergence.
@@
 known \emph{a priori.}  The convergenc time depends only on the choice of
-parameter $t_0$.  Liang \emph{et al.} give a rule of thumb in
+parameter $t_0$.
+The parameter $t_0$ can unfortunately be difficult to chose in advance
+for arbitrary systems.
+Liang \emph{et al.} give a rule of thumb in
 which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
@@
 for small spin systems, but that larger systems require an even higher
-$t_0$ value~\cite{schneider2017convergence}.
+$t_0$ value~\cite{schneider2017convergence}. Our work also confirms this.
 
-\subsection{SAD Algorithm}\label{sec:sad}
-In our previous work we discuss this in detail so here we only highlight...
-Stochastic Approximation w/ Dynamic $\gamma$ here.
+Pommerenck et al. propose a refinement~\cite{pommerenck2019stochastic} to SAMC where
+the update factor is determined dynamically rather than by the user.
+Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to provide the
+lowest temperature of interest $T_{\min}$. This is analogous to WL requiring a priori
+an energy range of interest; however, this may easier to identify and is more physical
+than the SAMC parameter $t_0$.
+The update factor can be written in terms of the current estimates for the highest $E_H$ and
+lowest $E_L$ energies of interest.
+\begin{align}
+  \gamma_{t}^{\text{SAD}} =
+     \frac{
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{t_L}
+     }{
+       \frac{E_{H}-E_{L}}{T_{\text{min}}} + \frac{t}{N_S}\frac{t}{t_L}
+     }
+\end{align}
 
@@
 
-\section{Acknowledgements}
-
 \bibliography{ising} % Produces the bibliography via BibTeX.

### 2019-07-03 papers/histogram/run-ising-comparisons.sh

@@ -7,11 +7,13 @@ python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-32-s1 32 isi
 python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-32-s1 32 ising-wl-inv-t-32
 
 # This is the reference for 128 X 128 system
-
+python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-128-s1 128 ising-wl-128
+python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-sad-128-s1 128 ising-sad-128
+python figs/parse-ising-out.py /home/jordan/ising-cp-data/ising-wl-inv-t-128-s1 128 ising-wl-inv-t-128
 
 # This is the reference for 32 X 32 system
 python figs/yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/deft/papers/histogram/data/ising-32-reference-lndos.dat ising/N32 32 2040 48 true 8 ising-sad-32 ising-samc-1e5-32 ising-samc-1e6-32 ising-samc-1e7-32 ising-wl-32 ising-wl-inv-t-32
 # This is the reference for 128 X 128 system
-python figs/yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/deft/papers/histogram/data/ising-sad-128-s1-reference-lndos.dat ising/N128 128 32760 100 true 8 ising-sad-128 ising-samc-1e5-128 ising-samc-1e6-128 ising-samc-1e7-128 ising-wl-128 ising-wl-inv-t-128
+python figs/yaml-multi-comparison.py /home/jordan/ising-cp-data/ /home/jordan/deft/papers/histogram/data/ising-128-reference-lndos.dat ising/N128 128 32760 100 true 8 ising-sad-128 ising-samc-1e7-128 ising-wl-128 ising-wl-inv-t-128 ising-samc-1e8-128 ising-samc-1e9-128
 
 # ising-cp-data not weniger-cp-data

### 2019-07-10 papers/histogram/ising.tex
some final ising paper additions before a full edit.

@@
 
-Pommerenck et al. propose a refinement~\cite{pommerenck2019stochastic} to SAMC where
-the update factor is determined dynamically rather than by the user.
-Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to provide the
-lowest temperature of interest $T_{\min}$. This is analogous to WL requiring a priori
-an energy range of interest; however, this may easier to identify and is more physical
-than the SAMC parameter $t_0$.
-The update factor can be written in terms of the current estimates for the highest $E_H$ and
-lowest $E_L$ energies of interest.
+Pommerenck et al. propose a refinement~\cite{pommerenck2019stochastic} to SAMC
+where the update factor is determined dynamically rather than by the user.
+Stochastic Approximation with a dynamic $\gamma$ (SAD) requires the user to
+provide the lowest temperature of interest $T_{\min}$. This is analogous to WL
+requiring a priori an energy range of interest; however, this may easier to
+identify and is more physical than the SAMC parameter $t_0$. The update factor
+can be written in terms of the current estimates for the highest $E_H$ and
+lowest $E_L$ energies of interest and the last time that an energy in the range
+of interest is encountered $t_L$.
 \begin{align}
@@
 \end{align}
+SAD only explores the energy range of interest as specified by the minimum
+temperature of interest $T_{\min} < T < \infty$. During the simulation the two
+energies $E_H$ and $E_L$, are refined such that the range of energies are conservatively
+estimated. The weights are calculated for each energy region according to the original
+prescription.
+\begin{enumerate}
+\item {$E < E_L$:} $w(E>E_H) = w(E_H)$
+\item {$E_L < E < E_H$:} moves are handled the same as other weight-based
+methods that are mentioned
+\item {$E > E_H$:} $w(E<E_L) = w(E_L)e^{-\frac{E_L-E}{T_{\min}}}$
+\end{enumerate}
+Each time the simulation changes the value of $E_H$ or $E_L$, the weights
+within the new portion of the interesting energy range are updated.
 
 \section{Results}
+We tested the algorithms on two 2D Ising model systems.  The first is a smaller
+simulation with a lattic size of 32 and the second has a lattice size of 128.
+For each system we use a reasonable root-mean-square displacement distance
+$\delta_0 = 0.05\sigma$ for proposed moves. The simulations explore the energy
+space of the systems with minimum reduced temperatures of $T_{\text{min}} = 1$
+for simulations All simulations lead to the minimum important energy $E_{\min}$
+and maximum entropy energy $E_{\max}$ being calculated (with the exception of
+the WL methods where both of these parameters are needed a priori).
+
 \begin{figure}
@@
   \caption{
-  gamma plot}
+  The update factor $\gamma_t$ versus the iteration number for the $N=32 \times 32$
+  system.}
 \end{figure}
 
-\begin{figure}
-\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
-  \caption{
-  gamma plot}
-\end{figure}
-
-\subsection{The 32 $\times$ 32 Ising model}
-Something small...
 \begin{figure}
@@
 \end{figure}
+
+\subsection{The 32 $\times$ 32 Ising model}
+For this simulation, we chose a minimum reduced temperature of $1$, which
+corresponds to an interesting energy range from $-2048$ to $50$.  The number of
+important energy states for this system is $N_S = 524$. \jpsays{The entropy of
+this system is shown in Fig.~\ref{fig:entropy-cartoon} above, which shows that
+over this energy range the entropy differs by 198, corresponding to a ratio of
+$10^{86}$ between the highest and lowest density of states.}
+
+Fig.~\ref{fig:n32} shows the average error in the entropy as a function of time for
+this system with the displacement distance of $0 = 0.05$. The solid/dashed lines
+represent the average of the absolute value of the error in the entropy averaged
+over eight simulations using different random number seeds. The range of average
+errors for each simulation is shown as a shaded region around its mean error. By
+the time $10^8$ moves have been made all but the WL simulation have begun to
+converge as $1/\sqrt{t}$. We then see the WL error saturate around $10^9$ moves.
+
+\begin{figure}
+\includegraphics[width=\columnwidth]{figs/ising/gamma-n128.pdf}
+  \caption{
+  The update factor $\gamma_t$ versus the iteration number for the $N=128 \times 128$
+  system.}
+\end{figure}
+
 \begin{figure}
@@
 \subsection{128 $\times$ 128 Ising system}
-Something big (WL uses 256 by 256 in their paper but only one simulation and
-no comparison while more recent comparison type papers use 128 by 128 as max)...
+For this simulation, we chose a minimum reduced temperature of $1$, which
+corresponds to an interesting energy range from $-30268$ to $100$.  The number of
+important energy states for this system is $N_S = 524$.
+
+Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
+for this system with the displacement distance of $0 = 0.05$. The solid/dashed
+lines represent the average of the absolute value of the error in the entropy
+averaged over eight simulations using different random number seeds. The range
+of average errors for each simulation is shown as a shaded region around its
+mean error. By the time $10^12$ moves have been made all but the WL and SAD
+simulations have begun to converge as $1/\sqrt{t}$. We then see the WL error
+saturate around $10^12$ moves. The error in the SAD method will continue to
+converge faster than $1/\sqrt{t}$ until the update factor approaches that of
+1/t-WL at which point both methods will converge equivalently.
 
 \section{Conclusion}
+We have examined the convergence of SAD versus other widely used weight-based
+histogram methods  We have shown that Sad samples the energy space corresponding
+to a desired range of temperatures for a variety of system sizes. We find that
+both SAD and 1/t-WL demonstrate excellent and robust convergence. They both
+converge more rapidly than SAMC, and unlike WL consistently converge to the
+correct density of states. We find that for larger systems SAD reduces the updated
+factor more slowly (and conservatively) than 1/t-WL. This means that SAD will take
+proportionately more moves to converge to the same value as 1/t-WL as system size is
+increased.
+SAD requires the user to specify a temperature range of interest rather than an
+energy range of interest as 1/t-WL does. For use cases in which a range of
+desired temperatures is known, this will make the SAD method considerably more
+convenient.
 

### 2019-07-11 papers/histogram/ising.tex
fixed some ising paper errors, made the ground state part of the data plot consideration, and edited the run script to account for it.

@@
 states are $N-1$.
-\jpsays{Also the necessary condition that the ground state is
-$2$ for the 2D Ising model is enforced to guarantee the accuracy of the
-density of states at low temperatures. If $2^N$ states is the condition
-used for the ferromagnetic ising model, the rescaled factor is
-determined primarily by the maximum density of states impacing the
-accuracy of the energy levels at or near the ground state.}
+% \jpsays{Also the necessary condition that the ground state is
+% $2$ for the 2D Ising model is enforced to guarantee the accuracy of the
+% density of states at low temperatures. If $2^N$ states is the condition
+% used for the ferromagnetic ising model, the rescaled factor is
+% determined primarily by the maximum density of states impacing the
+% accuracy of the energy levels at or near the ground state.}
 %\section{Discrete Spin Systems} In this work, we compare the
@@
 For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-2048$ to $50$.  The number of
-important energy states for this system is $N_S = 524$. \jpsays{The entropy of
-this system is shown in Fig.~\ref{fig:entropy-cartoon} above, which shows that
-over this energy range the entropy differs by 198, corresponding to a ratio of
-$10^{86}$ between the highest and lowest density of states.}
+corresponds to an interesting energy range from $-2048$ to $48$.  The number of
+important energy states for this system is $N_S = 529$.
 
@@
 For this simulation, we chose a minimum reduced temperature of $1$, which
-corresponds to an interesting energy range from $-30268$ to $100$.  The number of
-important energy states for this system is $N_S = 524$.
-
+corresponds to an interesting energy range from $-32768$ to $100$.  The number of
+important energy states for this system is $N_S = 8214$.
 Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
@@
 of average errors for each simulation is shown as a shaded region around its
-mean error. By the time $10^12$ moves have been made all but the WL and SAD
+mean error. By the time $10^{12}$ moves have been made all but the WL and SAD
 simulations have begun to converge as $1/\sqrt{t}$. We then see the WL error
-saturate around $10^12$ moves. The error in the SAD method will continue to
+saturate around $10^{12}$ moves. The error in the SAD method will continue to
 converge faster than $1/\sqrt{t}$ until the update factor approaches that of

### 2019-07-11 papers/histogram/figs/yaml-multi-comparison.py

@@ -53,7 +53,6 @@ for f in filename:
                 print('unable to read file', filename_location + name)
                 raise ValueError("%s isn't a file!" % (filename_location + name))
 
-            #print(data_loaded)
             data = yaml_data
             data['bins']['histogram'] = np.array(data['bins']['histogram'])
             data['bins']['lnw'] = np.array(data['bins']['lnw'])
@@ -105,13 +104,26 @@ for f in filename:
                 if yamlRef:
                     # if using yaml as a reference the range is from 0 to len while for C++ the range is
                     # from maxref to minref + 1
-                    norm_factor = np.mean(lndos[i][maxyaml:minyaml+1]) - np.mean(lndosref[0:minref-maxref+1])
-                    doserror = lndos[i][maxyaml:minyaml+1][::-1] - lndosref[0:minref-maxref+1] - norm_factor
+                    if 'ising' in filebase:
+                        ising_norm = lndos[i][maxyaml:minyaml+1] # remove impossible state
+                        # the states are counted backward hence the second to last state would be at index = 1
+                        new_ising_norm = np.delete(ising_norm,[1])
+
+                        ising_lndos = lndos[i][maxyaml:minyaml+1][::-1] # remove impossible state
+                        new_ising_lndos = np.delete(ising_lndos,[len(ising_lndos)-2])
+
+                        norm_factor = np.mean(new_ising_norm) - np.mean(lndosref[0:minref-maxref+1])
+                        doserror = new_ising_lndos - lndosref[0:minref-maxref+1] - norm_factor
+                    else:
+                        norm_factor = np.mean(lndos[i][maxyaml:minyaml+1]) - np.mean(lndosref[0:minref-maxref+1])
+                        doserror = lndos[i][maxyaml:minyaml+1][::-1] - lndosref[0:minref-maxref+1] - norm_factor
                 else:
                     norm_factor = np.mean(lndos[i][maxyaml:minyaml+1]) - np.mean(lndosref[maxref:minref+1])
                     doserror = lndos[i][maxyaml:minyaml+1][::-1] - lndosref[maxref:minref+1] - norm_factor
-                errorinentropy[i] = np.sum(abs(doserror))/len(doserror)
+
+                errorinentropy[i] = np.sum(abs(doserror))/len(doserror) #- np.mean(doserror)
                 maxerror[i] = np.amax(doserror) - np.amin(doserror)
+
 
             # remove N from moves in yaml file because N is added back in the
             # comparison-plot script

### 2019-07-17 papers/histogram/sad.tex
made all of Michael Perlins suggestions that seemed relavent.

@@
 \begin{abstract}
-  We present a new Monte-Carlo algorithm based on the Stochastic
+  We present a new Monte Carlo algorithm based on the Stochastic
   Approximation Monte Carlo (SAMC) algorithm for directly calculating
@@
   Approximation with a Dynamic update factor (SAD)
-  which dynamically adjusts the update factor $\gamma$ during the course of
+  which dynamically adjusts the update factor $\gamma_t$ during the course of
   the simulation. We test this method on the square-well fluid and
@@
   related
-  Monte-Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
+  Monte Carlo methods. We find that both the SAD and $1/t$-Wang-Landau ($1/t$-WL)
   methods rapidly converge to the
@@
 \section{Introduction}
-Over the past several decades a number of flat histogram Monte-Carlo
-simulation algorithms have been developed which calculate the
-thermodynamic properties of various systems for over a range of temperatures.  This
-development began with the original histogram method, which used a
-single canonical Monte Carlo simulation to predict properties for
-nearby temperatures~\cite{ferrenberg1988new}.  For large systems this
-approach is limited to a narrow temperature range because a single
-canonical simulation explores only a small range of energies.  This
-led to a variety of ``flat'' (or ``broad'') histogram
-methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
-  wang2001determining, wang2001efficient, trebst2004optimizing}, which
-explore a wider range of energies.  Another benefit, in
-contrast with low-temperature canonical Monte Carlo, is that these
+Over the past several decades a number of flat histogram Monte Carlo simulation
+algorithms have been developed which calculate the thermodynamic properties of
+various systems for over a range of temperatures.  This development began with
+the original histogram method, which used a single canonical Monte Carlo
+simulation to predict properties for nearby
+temperatures~\cite{ferrenberg1988new}.  For large systems this approach is
+limited to a narrow temperature range because a single canonical simulation
+explores only a small range of energies.  This led to a variety of ``flat'' (or
+``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
+swendsen1999transition, wang2001determining, wang2001efficient,
+trebst2004optimizing}, which explore a wider range of energies.  Another
+benefit, in contrast with low-temperature canonical Monte Carlo, is that these
 approaches cannot be trapped in a local energy minimum.
 
-Wang and Landau introduced the most widely used flat histogram
-algorithm (WL) that uses an update factor
-and a statistical histogram to compute the density of states of a given
-system~\cite{wang2001determining, wang2001efficient}.  While the method
-is incredibly powerful, it has a few disadvantages. Firstly, it
-requires the user to select the range
-of energies to be studied~\cite{landau2004new}. This adds an additional
-hurdle to its application to systems for which the interesting range
-of energies is not known \emph{a priori}.  The simulation
-violates detailed balance albiet briefly as the size of the violation decreases
-with time, which complicates convergence analysis.  In fact, the error
-in a WL computation the error has been demonstrated to saturate at a
-non-zero value~\cite{yan2003fast}, i.e. the method does not converge to
-the true density of states~\cite{belardinelli2007wang,
-belardinelli2007fast, belardinelli2008analysis,
-belardinelli2014intrinsic, singh2012density}.
+Wang and Landau introduced the most widely used flat histogram algorithm (WL)
+that uses an update factor and a statistical histogram to compute the density of
+states of a given system~\cite{wang2001determining, wang2001efficient}.  While
+the method is incredibly powerful, it has a few disadvantages. Firstly, it
+requires the user to select the range of energies to be
+studied~\cite{landau2004new}. This adds an additional hurdle to its application
+to systems for which the interesting range of energies is not known \emph{a
+priori}.  The simulation violates detailed balance, albeit briefly as the size of
+the violation decreases with time, which complicates convergence analysis.  In
+fact, the error in a WL computation has been demonstrated to saturate
+at a non-zero value~\cite{yan2003fast}, i.e. the method does not converge to the
+true density of states~\cite{belardinelli2007wang, belardinelli2007fast,
+belardinelli2008analysis, belardinelli2014intrinsic, singh2012density}.
 %
-Belardinelli and Pereyra demonstrated that allowing the update factor
-to decrease faster than $1/t$ leads to
-nonconvergence~\cite{belardinelli2007wang}.
-%% Around the same time, the convergence rate for WL was being
-%% explored~\cite{zhou2005understanding,lee2006convergence,
-%%   belardinelli2007wang}.
+Belardinelli and Pereyra demonstrated that allowing the update factor to
+decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang}
+where $t$ corresponds to the number of moves.
 %
-This leads to their $1/t$-WL algorithm which ensures that the error
-continues to decrease asymptotically as
-$1/\sqrt{t}$~\cite{belardinelli2008analysis}.
+This leads to their $1/t$-WL algorithm which ensures that the error continues to
+decrease asymptotically as $1/\sqrt{t}$~\cite{belardinelli2008analysis}.
 
-Liang began to consider that WL could be considered a special case of
-the Stochastic Approximation approach, whose convergence could be
-mathematically proven~\cite{liang2006theory}. In 2007, Liang, Liu, and
-Carrol \emph{et al.} developed Stochastic Approximation Monte Carlo
-(SAMC)~\cite{liang2007stochastic}, and proved its convergence,
-although the method still has a system specific user-defined parameter
-which must be tuned when applying this algorithm to a new system.
+Liang began to consider that WL could be considered a special case of the
+Stochastic Approximation approach, whose convergence could be mathematically
+proven~\cite{liang2006theory}. In 2007, Liang, Liu, and Carrol \emph{et al.}
+developed Stochastic Approximation Monte Carlo
+(SAMC)~\cite{liang2007stochastic}, and proved its convergence, although the
+method still has a system specific user-defined parameter which must be tuned
+when applying this algorithm to a new system.
 
-In this work, we have developed an improved algorithm based on SAMC
-that does not require an array of non-physical, user-defined inputs
-and therefore should be easily applicable to any system. The method
-does require the user to define a temperature range of ($T_\text{min}$
-to $T_\infty$).  We call this method SAD (Stochastic Approximation
-with a Dynamic update factor), and will discuss it in detail in the
-methods section. We compare its convergence properties with three
-existing flat histogram methods: WL, $1/t$-WL, and SAMC.
+In this work, we have developed an improved algorithm based on SAMC that does
+not require an array of non-physical, user-defined inputs and therefore should
+be easily applicable to any system. The method does require the user to define a
+temperature range of interest ($T_\text{min}$ to $T_\infty$).  We call this
+method SAD (Stochastic Approximation with a Dynamic update factor), and will
+discuss it in detail in the methods section. We compare its convergence
+properties with three existing flat histogram methods: WL, $1/t$-WL, and SAMC.
 
-As a test case, we consider the square-well fluid i.e. a system of
-particles whose interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The
-square-well potential is an ideal test-bed as it is a simple model for
-a liquid, which includes both attractive and repulsive
-interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The
-potential $U(\textbf{r})$ for such a system is given by
+As a test case, we consider the square-well fluid i.e. a system of particles
+whose interactions are governed by a square-well
+potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well
+potential is an ideal test-bed as it is a simple model for a liquid, which
+includes both attractive and repulsive
+interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The potential
+$U(\textbf{r})$ for such a system is given by
 \begin{equation}
@@
 where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
-reduced range of the potential well, and $\epsilon$ is its depth.
-This model has the further advantage that because the energy is
-discrete, binning is not required.
-
-%The organization of this paper is as follows: In Section~\ref{sec:histogram}, we
-%describe in detail flat-histogram methods.
-%Section~\ref{sec:sad} introduces the dynamic stochastic approximation method.
-%In Section~\ref{sec:results}, we discuss the performance of the various methods
-%applied to the model system.
+reduced range of the potential well, and $\epsilon$ is its depth. This model has
+the further advantage that binning is not required because the energy is
+discrete.
 
@@
 fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
-Approximation Monte-Carlo (SAMC), and SAD.
+Approximation Monte Carlo (SAMC), and SAD.
 
@@
 stored.  Since in the microcanonical ensemble the entropy is defined
-as $S(E) \equiv k_B\ln(D(E))$, the logarithm of the weights can be
+as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights can be
 thought of as an approximation of the entropy.
@@
 \end{equation}
-where $t$ is number of the current move, $\gamma(t)$ is a
+where $t$ is number of the current move, $\gamma_t(t)$ is a
 move-dependent update factor, and $E$ is the current energy.  This
@@
 The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
-  landau2014guide} begins with $\gamma=1$, and then decreases $\gamma$
+  landau2014guide} begins with $\gamma_t=1$, and then decreases $\gamma$
 in discrete stages.  We track the number of moves ending at each
 energy in a histogram.  When that histogram is sufficiently flat,
-$\gamma$ is decreased by a specified factor of $\frac12$.  The
+$\gamma_t$ is decreased by a specified factor of $\frac12$.  The
 flatness is defined by the ratio between the minimum value of the
 histogram and its average value.  When this flatness reaches a
-specified threshold (typically 0.8), the $\gamma$ value is decreased
+specified threshold (typically 0.8), the $\gamma_t$ value is decreased
 and the histogram is reset to zero.  This approach requires that the
@@
 sampled~\cite{haber2014transition}.  The entire process is repeated
-until $\gamma$ reaches a desired cutoff.
+until $\gamma_t$ reaches a desired cutoff.
 
 The Wang-Landau approach thus has three parameters that need be
-specified: the factor by which to decrease $\gamma$ when flatness is
+specified: the factor by which to decrease $\gamma_t$ when flatness is
 acheived, the flatness criterion, and the cutoff that determines when
@@
 of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
-and does decrease $\gamma$ so quickly that it will never (for any
+and does decrease $\gamma_t$ so quickly that it will never (for any
 cutoff value) decrease the error in the density of states beyond a
@@
 implemented a schedule that enforced that if the update factor
-$\gamma$ is ever less than $\nicefrac{1}{t}$ then the update factor
+$\gamma_t$ is ever less than $\nicefrac{1}{t}$ then the update factor
 is set to $\nicefrac{1}{t}$ and the histogram is no longer tracked. Employing this
@@
 including $\nicefrac{N_S}{t}$ scaling and switching from standard WL
-to $1/t$-WL when the update factor $\gamma <
+to $1/t$-WL when the update factor $\gamma_t <
 \nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the
@@
 the lack of convergence of Wang-Landau's approach with a simple
-schedule by which the update factor $\gamma$ is continuously
+schedule by which the update factor $\gamma_t$ is continuously
 decreased~\cite{liang2007stochastic, werlich2015stochastic,
@@
 \begin{align}
-\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}
+\gamma_{t}^{\text{SA}} =\frac{t_0}{\max(t_0,t)}\label{eq:1}
 \end{align}
@@
 \begin{align}
-   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t \\
+   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{SA} \\
   % &= \int_0^{t_{\min}} \gamma(t)dt \\
@@
 $T_\text{min}$ may be easier for a user than determining in advance
-an energy rang of intereste or the unphysical parameter $t_0$.
+an energy range of interest or the unphysical parameter $t_0$.
 
@@
 While for SAMC, the update factor is defined in the original
-implementation, for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
+implementation (see equation~\eqref{eq:1}), for SAD the update factor $\gamma_{t}^{\text{SAD}}$ is
 thought of as $\nicefrac{\text{d}S}{\text{d}t}$. This tells us that
@@
 entropy, and puts a lower bound on the convergence time. After
-\emph{long} times, when all the energies have been a long time ago, we
+\emph{long} times, when all the energies have been explored a long time ago, we
 wish for a lower update factor in order to more rapidly refine the
@@
 energy that prior to that move
-had the highest histogram value (i.e. been visited the most times) during
-the course of the simulation.
+had the highest histogram value (i.e. been visited the most times) at some point
+during the course of the simulation.
 This definition results in a ``ratcheting'' effect, in which $E_H$
@@
 limit will ratchet down to include $E_{\min}$.
-Each time we change the value of $E_H$ or $E_L$, the weights within the
+Each time we change the value of $E_H$ or $E_L$, the weights outside the
 new portion of the interesting energy range are set to the expressions
@@
 methods share \emph{after} they have found all the energies---is that
-the schedule for $\gamma$ automatically responds to the choice of bin
+the schedule for $\gamma_t$ automatically responds to the choice of bin
 size. SAD should perform similarly over a wide range of bin sizes
-because $\gamma \propto \nicefrac{N_S}{t}$.  As the number of energy
+because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
 states $N_S$ found increases (fine binning), the time spent $t$ in
@@
 %
-The smaller translation scale causes all methods to take additional
-time to explore all energies. Based on random walk scaling, a ideal
-method should scale roughly as $\delta_0^{-2}$ in the limit of small
-$\delta_0$, that is, one order of magnitude in the displacement
-distance should result in two order of magnitude increase in
-convergence time. SAMC simulations with a $t_0$ value that rapidly
-converged for $\delta = 0.05\sigma$ do not converge at all in $10^{12}$
-moves for a translation scale of $\delta = 0.005\sigma$. It is also
-worth noting that for the smaller displacement distance, the SAMC rule
-of thumb of choosing $t_0$ to be approximately $100N_S$ is no longer
-valid. SAD, WL, and $1/t$-WL handle the shift in displacement distance
-and converge roughly as expected.
+The smaller translation scale causes all methods to take additional time to
+explore all energies. Based on random walk scaling, the convergence time of an
+ideal method should scale roughly as $\delta_0^{-2}$ in the limit of small
+$\delta_0$, that is, one order of magnitude decrease in the displacement
+distance should result in two order of magnitude increase in convergence time.
+SAMC simulations with a $t_0$ value that rapidly converged for $\delta =
+0.05\sigma$ do not converge at all in $10^{12}$ moves for a translation scale of
+$\delta = 0.005\sigma$. It is also worth noting that for the smaller
+displacement distance, the SAMC rule of thumb of choosing $t_0$ to be
+approximately $100N_S$ is no longer valid. SAD, WL, and $1/t$-WL handle the
+shift in displacement distance and converge roughly as expected.
 
 The methods SAD, WL, and $1/t$-WL compensate for the smaller
-displacement distance by reducing $\gamma$ more slowly, as can be seen from
+displacement distance by reducing $\gamma_t$ more slowly, as can be seen from
 Figure~\ref{fig:fast-slow-gamma}b and ~\ref{fig:fast-slow-gamma}d.

### 2019-07-19 papers/histogram/sad-referee-response.md
made some additions to the paper citations and referee response.

Dear editor,

...

Response to first referee

The first referee acknowledges that our paper is well written and that the field
is an important one, but does not believe that our work is a substantial
contribution to the field.  The referee has three major criticisms of the value
of our work (aside from issues with the introduction section).

Firstly, the referee criticizes our paper for testing the method on a single
physical system (the square well fluid).  The referee gives as an argument an
example of a paper which tests a method on three different physical systems.  We
give as counter examples references **look them up** (including references
**look them up** which the referee asked us to add) which each use a single
physical system as the test case.  We agree that our method should be tested on
additional different physical systems but dispute that this is necessary for the
first paper introducing this method.  **Decide whether to just throw in the
Ising...**

Secondly, the referee does not see much value in providing a temperature range
(rather than an energy range) as input to the method.  We believe that a
temperature range is *far* more physical and convenient, and this is the primary
value of our paper.  **We have modified the abstract to better highlight this
contribution.**

Thirdly, because we find that 1/t-WL converges at the same rate as our method,
the referee concludes that our method is not a substantial improvement. We note
first (as mentioned in the paper in **...**) that the 1/t-WL method is solving a
significantly easier problem, because it is told in advance the range of
energies that are of interest.  We agree, however, that the benefit of our
method lies specifically in the fact that a temperature range is given as an
input.  Achieving the same result as the SAD method with 1/t-WL would either
require multiple simulations over a wider range of energies, or would require
guessing in advance a wider range of energies than are needed.  In either case,
the convergence would be worse than our tests indicate.

1. **Existence of the SAMC method.** We state in the paper "In this work, we
have developed an improved algorithm based on SAMC."  This is again reinforced
in Secion III.

2. We have added a section detailing the important impact of multicanonical
methods on influencing methods like WL etc... referencing the original authors
to the introduction.

3.

4.

5.

Minor details:

6.

7.

8. *Proofreading* **Note all changes made under proof reading, as well as what was
suggested.**
In the introduction, we improve the clarity of the paper by defining t as the number
of moves when first introducing 1/t-WL.
As the first referee suggests we edited phrases such as Monte Carlo and the
update factor for consistency.
We corrected several typographical errors throughout the paper.
References were checked for completeness.

The second referee was far more positive on the value of our work, but did
suggest several places where the paper could be made more clear.

1. The referee wrote that the description of the flat histogram methods in
section II was confusing, and should be rewritten to highlight the differences
between methods.  We have added ...

2. The referee shared that the entropy argument in Section II.D is unclear, and
wished for the relationship between entropy and the update factor to be more
fully described, as well as the basis for Equation 7.

3. The referee asked for a better choice of line styles to highlight the SAD
method in our plots.

4. The referee finally suggested that it might be interesting to try out our
method with a Go protein model.  We agree that this would be interesting, and
will consider this suggestion for a future paper.

### 2019-07-19 papers/histogram/sad.tex
made some additions to the paper citations and referee response.

@@
 limited to a narrow temperature range because a single canonical simulation
-explores only a small range of energies.  This led to a variety of ``flat'' (or
+explores only a small range of energies.~\jpsays{
+Berg and Neuhaus developed multicanonical
+methods that improved the sampling of configuration space and attempted to
+overcome free energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+These works led to an explosion in the interest of}
+``flat'' (or
 ``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
@@
 
+~\jpsays{STMC uses a dynamic microcanonical temperature $\tilde{T}$ to
+accelerates convergence by refining the temperature estimate rather than the
+density of states estimate~\cite{kim2006statistical, kim2007statistical}. STMC
+applied to WL requires a temperature range rather than an energy range.
+From what I can tell the user needs to input a temperature range $T_l$ and $T_h$
+in general and an energy range for continuum systems $E_l$ and $E_h$.
+}
+
+~\jpsays{The benefits of Monte Carlo methods such as WL cannot easily be
+exploited on systems where MC moves are not easily
+defined~\cite{kim2006statistical}, such as in the case of molecular simulations,
+or where for continuum systems where energy bins increase extensively with
+system size~\cite{kim2009replica, junghans2014molecular}.}
+
 As a test case, we consider the square-well fluid i.e. a system of particles
@@
 individual energy, which can then be reweighted for any
-temperature~\cite{errington2003direct}.
+temperature~\cite{panagiotopoulos1998phase, panagiotopoulos2000monte,
+errington2003direct}.
 
@@
 
-Flat histogram methods calculate the density of states $D(E)$ for a
-discrete set of energies~\cite{haber2018performance}. For this reason,
-energy binning becomes an important consideration for systems with a
-continuum of possible energies.  Energy bins are typically of uniform
-size for the entire energy continuum~\cite{fasnacht2004adaptive}. Some
-methods such as AdaWL~\cite{koh2013dynamically} employ a tunable
-mechanism for controlling the binning for low entropic states in order
-to ensure the exploration of all energies.  The method introduced in
-this paper is designed to scale appropriately as bin size is changed,
-but we do not test this scaling, as we use a system with discrete
-energy levels.
+Flat histogram methods calculate the density of states $D(E)$ for a discrete set
+of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
+trebst2004optimizing}. For this reason, energy binning becomes an important
+consideration for systems with a continuum of possible energies.  Energy bins
+are typically of uniform size for the entire energy
+continuum~\cite{fasnacht2004adaptive}. Some methods such as
+AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
+binning for low entropic states in order to ensure the exploration of all
+energies.  The method introduced in this paper is designed to scale
+appropriately as bin size is changed, but we do not test this scaling, as we use
+a system with discrete energy levels.
 

### 2019-07-23 papers/histogram/sad-referee-response.md
added some significant text to work with as we write the response.

@@
 
+***Jordan Says:*** Computational Physics Papers in Physical Review E
+Appropriate papers for this section are those that report new advances in
+computational physics. This includes papers that describe new developments of
+computational methods or substantial improvements of existing computational
+methods, provided they are applicable to significant physics problems.
+In many cases such a paper will improve its impact by explicitly demonstrating
+the usefulness of the described method, i.e., by applying it to some physics
+problem (old or new). However, it should be noted that such an application is
+not a necessary condition for eligibility. Rather, the requirement that "papers
+must contain new results in physics" should be interpreted as a requirement of
+novelty with respect to the method(s), the physics, or both.
+*PRE only even loosely requires that new methods be tested on a single (old) or new system.
+Simply arguing that our method is distinct from WL (the referees argument) 'it
+is a new version of SAMC!' should be sufficient based on PRE guidelines.*
+
 Firstly, the referee criticizes our paper for testing the method on a single
@@
 Ising...**
+
+**look them up**
+It is quite common for papers detailing comparison of convergence for various
+methods to only examine a single physical or sometimes even an unphysical (as in
+the case of integrals) system. When examining the convergence and/or comparing
+various computational methods it cannot therefore be expected that multiple
+systems should be considered in the same work. The authors provide a list of
+such works that they directly cite (many accepted by PRE) that prove their point:
+
+Analysis of the convergence of the 1/t and Wang-Landau algorithms in the calculation
+of multidimensional integrals (accepted by PRE)
+This paper compares WL and 1/t-WL on several different integrals (not even a real physical system because the exact answer is known)
+
+Avoiding boundary effects in Wang-Landau sampling (accepted by PRE)
+This paper with DP Landau as an author only considers the 2D ising model with
+the novelty of the paper to avoid boundary effects in WL which is succinctly
+stated in the abstract: 'A simple modification of the ‘‘Wang-Landau sampling’’
+algorithm...'
+
+Optimal modification factor and convergence of the Wang-Landau algorithm (accepted by PRE)
+This paper examines the optimal modification factor and tests only on the 2D
+ising model.
+
+Convergence of Stochastic Approximation Monte Carlo and modified
+Wang–Landau algorithms: Tests for the Ising model (accepted by Computer Physics Communications)
+This paper compares SAMC, WL, and 1/t-WL on a single system: the 2D ising model
+
+Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model (accepted by Computer Physics Communications)
+SAMC and WL are compared for a single continuum model with a square well type potential.
+
+The referee even suggested to cite some of the original multicannonical literature
+Multicanonical Ensemble: A New Approach to Simulate First-Order Phase Transitions (accepted by PRL)
+This important work considers a single system in the 2D 10 state Potts model.
 
@@
 
-3.
+3. As per the referees request, we have briefly reviewed methods the Statistical
+temperature Monte Carlo method.
 
-4.
+4.  As per the referees request, we have briefly reviewed methods used in the
+metadynamics communities in the introduction.
 
-5.
+5. more test cases
 
@@
 
-6.
+6. Out of curiosity: The authors say a temperature range T_min < T <
+infinity should be chosen. Why not T_min < T < T_max?
 
-7.
+7. On page 5 the authors say "SAD should perform similarly ...". Is
+there any evidence?
 

### 2019-07-25 papers/histogram/sad-referee-response.md
added more supporting citations and text to referee response for trimming.

--- 2019-07-19 papers/histogram/sad-referee-response.md
@@
 
+***Jordan Says:*** Computational Physics Papers in Physical Review E
+Appropriate papers for this section are those that report new advances in
+computational physics. This includes papers that describe new developments of
+computational methods or substantial improvements of existing computational
+methods, provided they are applicable to significant physics problems.
+In many cases such a paper will improve its impact by explicitly demonstrating
+the usefulness of the described method, i.e., by applying it to some physics
+problem (old or new). _However, it should be noted that such an application is
+not a necessary condition for eligibility_. Rather, the requirement that "papers
+must contain new results in physics" should be interpreted as a requirement of
+novelty with respect to the method(s), the physics, or both.
+*PRE only even loosely requires that new methods be tested on a single (old) or new system.
+Simply arguing that our method is distinct from WL (the referees argument) 'it
+is a new version of SAMC!' should be sufficient based on PRE guidelines.*
+
 Firstly, the referee criticizes our paper for testing the method on a single
@@
 example of a paper which tests a method on three different physical systems.  We
-give as counter examples references **look them up** (including references
-**look them up** which the referee asked us to add) which each use a single
+give as counter examples references **look them up**  which each use a single
 physical system as the test case.  We agree that our method should be tested on
@@
 Ising...**
+It is quite common for papers detailing comparison of convergence for various
+methods to only examine a single physical or sometimes even an unphysical (as in
+the case of integrals) system. When examining the convergence and/or comparing
+various computational methods it cannot therefore be expected that multiple
+systems should be considered in the same work. The authors provide a list of
+such works that they directly cite (many accepted by PRE) that prove their point:
+
+Analysis of the convergence of the 1/t and Wang-Landau algorithms in the calculation
+of multidimensional integrals *(accepted by PRE)*
+This paper compares WL and 1/t-WL on several different integrals (not even a
+real physical system because the exact answer is known)
+
+Avoiding boundary effects in Wang-Landau sampling *(accepted by PRE)*
+This paper with DP Landau as an author only considers the 2D ising model with
+the novelty of the paper to avoid boundary effects in WL which is succinctly
+stated in the abstract: 'A simple modification of the ‘‘Wang-Landau sampling’’
+algorithm...'
+
+Optimal modification factor and convergence of the Wang-Landau algorithm *(accepted by PRE)*
+This paper examines the optimal modification factor and tests on a single
+system: the 2D ising model.
+
+Convergence of Stochastic Approximation Monte Carlo and modified
+Wang–Landau algorithms: Tests for the Ising model *(accepted by Computer Physics Communications)*
+This paper compares SAMC, WL, and 1/t-WL on a single system: the 2D ising model
+
+Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model *(accepted by Computer Physics Communications)*
+SAMC and WL are compared for a single continuum model with a square well type potential.
 
@@
 the referee concludes that our method is not a substantial improvement. We note
-first (as mentioned in the paper in **...**) that the 1/t-WL method is solving a
+first (as mentioned in the paper in **...** The results section we could highlight again in the conclusion?) that the 1/t-WL method is solving a
 significantly easier problem, because it is told in advance the range of
@@
 
+Multidimensional stochastic approximation Monte Carlo
+*(accepted by PRE)*
+"the Wang-Landau MC [3] scheme and its mathematical generalization, stochastic approximation
+MC (SAMC) [4–6].""
+
+Convergence estimation of flat-histogram algorithms based on simulation results
+*(accepted by Computer Physics Communications)*
+"At the same time, Liang suggested a generalized simulation method:stochastic
+approximation Monte Carlo (SAMC) [7,8] including the 1/t-WL as a particular case"
+
+Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a continuum polymer model
+*(accepted by Computer Physics Communications)*
+"Liang et al.[30] showed that WLMC could be seen as a version of SAMC and using
+the mathematical background of stochastic approximation methods they proved the
+convergence of SAMC."
+
 2. We have added a section detailing the important impact of multicanonical

### 2019-07-31 papers/histogram/sad-referee-response.md
added significant updates to the sad-referee response

--- 2019-07-19 papers/histogram/sad-referee-response.md
@@
 The first referee acknowledges that our paper is well written and that the field
-is an important one, but does not believe that our work is a substantial
-contribution to the field.  The referee has three major criticisms of the value
-of our work (aside from issues with the introduction section).
+is an important one, but does not believe that our work is a *'substantial
+contribution*'' to the field.  The referee has three major criticisms of the
+value of our work (aside from issues with the introduction section).
+
+**We argue that our method is based on SAMC (as we state) and is a substantial
+improvement on the method (convergence is of 1 - 2 orders of magnitude superior
+as can be seen from Fig. 3 and Fig. 4)**
+
+**In addition, we also provide a detailed comparison of 4 different methods on a
+system that would potentially be of value to those most likely to implement our
+method (such as the polymer community has implemented SAMC)**
+
+**Furthermore, the guidlines that the referee is using come from Computational
+Physics Papers in the Physical Review E (2013). This specifically states
+'Rather, the requirement that "papers must contain new results in physics"
+should be interpreted as a requirement of novelty with respect to the
+method(s), the physics, or both.' As stated above it is only necessary that the
+method is novel OR the physical problem is novel. The paper below accepted by
+PRE confirms this thought.**
+
+**Analysis of the convergence of the 1/t and Wang-Landau algorithms in the
+calculation of multidimensional integrals** *(accepted by PRE)*
+This paper compares the performance of two methods (WL and 1/t-WL) to calculate
+different integrals (not a real physical system because the exact answer is
+known).
 
@@
 example of a paper which tests a method on three different physical systems.  We
-give as counter examples references **look them up** (including references
-**look them up** which the referee asked us to add) which each use a single
+give as counter examples *references that we cite* which each use a single
 physical system as the test case.  We agree that our method should be tested on
 additional different physical systems but dispute that this is necessary for the
-first paper introducing this method.  **Decide whether to just throw in the
-Ising...**
+first paper introducing this method.
+
+**Analysis of the convergence of the 1/t and Wang-Landau algorithms in the
+calculation of multidimensional integrals** *(accepted by PRE)*
+This paper compares the performance of two methods (WL and 1/t-WL) to calculate
+different integrals (not a real physical system because the exact answer is
+known).
+
+**Avoiding boundary effects in Wang-Landau sampling** *(accepted by PRE)*
+This paper with DP Landau as an author considers a single system (the 2D Ising
+model). The novelty of the paper is providing a scheme to avoid boundary effects
+in WL.
+
+**Optimal modification factor and convergence of the Wang-Landau algorithm**
+*(accepted by PRE)*
+This paper examines the optimal modification factor for the WL algorithm and
+tests on a single system (the 2D Ising model).
+
+**Convergence of Stochastic Approximation Monte Carlo and modified Wang–Landau
+algorithms: Tests for the Ising model**
+*(accepted by Computer Physics Communications)*
+This paper compares SAMC, WL, and 1/t-WL on a single system (the 2D Ising model).
+
+**Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a
+continuum polymer model** *(accepted by Computer Physics Communications)*
+SAMC and WL are compared for a single continuum model with a square well type
+potential.
+
+These papers help to highlight that the convergence properties of the various
+Monte Carlo methods are of importance (while the system(s) are not necessarily).
+In our case, we compare the convergence *four* different methods, one of which
+(SAD) is an improvement on the existing algorithm (SAMC), on a single system
+(square well fluid).
 
@@
 the referee concludes that our method is not a substantial improvement. We note
-first (as mentioned in the paper in **...**) that the 1/t-WL method is solving a
-significantly easier problem, because it is told in advance the range of
-energies that are of interest.  We agree, however, that the benefit of our
+first (as mentioned in the paper in *Section IV* that the 1/t-WL method is
+solving a significantly easier problem, because it is told in advance the range
+of energies that are of interest.  We agree, however, that the benefit of our
 method lies specifically in the fact that a temperature range is given as an
@@
 guessing in advance a wider range of energies than are needed.  In either case,
-the convergence would be worse than our tests indicate.
+the convergence of 1/t-WL would be worse than our tests indicate.
+*Jordan Says we could add* **We have modified the conclusion to better emphasize
+this important distication**
 
-1. **Existence of the SAMC method.** We state in the paper "In this work, we
-have developed an improved algorithm based on SAMC."  This is again reinforced
-in Secion III.
+**----------------------------------------------------------------------------**
 
-2. We have added a section detailing the important impact of multicanonical
-methods on influencing methods like WL etc... referencing the original authors
-to the introduction.
+1. The referee does not acknowledge that our method is based on SAMC (the
+mathematical generalization of WL). In fact, the referee does not seem to
+acknowledge *the existence of the SAMC method* which we find curious. We clearly
+state in the paper in section I *'In this work, we have developed an improved
+algorithm based on SAMC.'*  This is again reinforced in Secion III.
 
-3.
+There are many authors that acknowledge SAMC as a mathematical generalization of
+WL. We list a few here in our response.
 
-4.
+**Multidimensional stochastic approximation Monte Carlo**
+*(accepted by PRE)*
+"the Wang-Landau MC [3] scheme and its mathematical generalization, stochastic
+approximation MC (SAMC) [4–6].""
 
-5.
+**Convergence estimation of flat-histogram algorithms based on simulation results**
+*(accepted by Computer Physics Communications)*
+"At the same time, Liang suggested a generalized simulation method:stochastic
+approximation Monte Carlo (SAMC) [7,8] including the 1/t-WL as a particular case"
+
+**Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a
+continuum polymer model**
+*(accepted by Computer Physics Communications)*
+"Liang et al.[30] showed that WLMC could be seen as a version of SAMC and using
+the mathematical background of stochastic approximation methods they proved the
+convergence of SAMC."
+
+The referee further disputes the use of language *'we have introduced a new
+algorithm'*. In response, we argue that this language is commonly used to
+describe MC methods that have their foundation in other methods (SAD is a
+derivative of SAMC). Our choice of language is further validated by other
+researchers in the field that are published in PRE.
+
+**Dynamically optimized Wang-Landau sampling with adaptive trial moves and
+modification factors**
+*(accepted by PRE)*
+This work introduces an improved update scheme as the referee would call it.
+The authors still use language such as *'To test our new algorithm...'*
+
+Still we have changed the language to say *'we have introduced a new variant of
+SAMC called SAD'"'* to better highlight the subtle distinction.
+
+**----------------------------------------------------------------------------**
+
+2. The referee suggests that we might include references to the originators of
+the multicanonical methods.
+**We have added a paragraph to Section I detailing the important impact of
+multicanonical methods on influencing flat histogram methods. We have referenced
+the original authors Berg and Neuhaus.**
+
+**----------------------------------------------------------------------------**
+
+3. The referee suggests that that it would be helpful to include references to
+the Statistical *(not Stochastic)* Temperature Monte Carlo methods.
+**We have added a paragraph to Section I detailing the important impact of STMC
+and RESTMC and its improvement on the WL method and Replica Exchange
+*respectively.**
+
+**----------------------------------------------------------------------------**
+
+4. The referee suggests that that it would be valuable to include references to
+STMD and metadynamics.
+**We have added a paragraph to Section I detailing the important impact of STMD
+forging a connection between WL and metadynamics.**
+
+**----------------------------------------------------------------------------**
+
+5. The referee argues that our method should be tested on multiple other systems.
+The referee gives for example a single paper accepted by *PRL* that compares STMC
+to WL and introduces STMD.
+
+**----------------------------------------------------------------------------**
 
@@
 
-6.
+6. The referee asks why we chose a maximum temperature of infinity.  We have
+added a paragraph to Section III explaining the reasoning, which is that
+infinite temperature allows the system to randomize more reliably, and costs us
+essentially nothing.  We do note that for some systems or situations in which
+the energy of maximum entropy is very high (or worse, infinite) we would need
+to introduce a maximum temperature.
 
-7.
+7. On page 5 the authors say "SAD should perform similarly ...". Is
+there any evidence?
 
@@
 References were checked for completeness.
+
+**----------------------------------------------------------------------------**
 

### 2019-08-01 papers/histogram/sad-referee-response.md
finished the sad-referee response. now just need to implement changes in the paper.

@@
 improvement on the method (convergence is of 1 - 2 orders of magnitude superior
-as can be seen from Fig. 3 and Fig. 4)**
+as can be seen from Fig. 3 and Fig. 4).
 
-**In addition, we also provide a detailed comparison of 4 different methods on a
-system that would potentially be of value to those most likely to implement our
-method (such as the polymer community has implemented SAMC)**
+In addition, we also provide a detailed comparison of 4 different methods on the
+square well fluid.
 
-**Furthermore, the guidlines that the referee is using come from Computational
-Physics Papers in the Physical Review E (2013). This specifically states
-'Rather, the requirement that "papers must contain new results in physics"
-should be interpreted as a requirement of novelty with respect to the
+Furthermore, the guidelines that the referee is using presumably come from
+Computational Physics Papers in the Physical Review E (2013). This specifically
+states '...Rather, the requirement that "papers must contain new results in
+physics" should be interpreted as a requirement of novelty with respect to the
 method(s), the physics, or both.' As stated above it is only necessary that the
@@
 These papers help to highlight that the convergence properties of the various
-Monte Carlo methods are of importance (while the system(s) are not necessarily).
-In our case, we compare the convergence *four* different methods, one of which
-(SAD) is an improvement on the existing algorithm (SAMC), on a single system
-(square well fluid).
+Monte Carlo methods are of importance (while the system(s) are often not
+necessarily). In our case, we compare the convergence of *four* different
+methods, one of which (SAD) is an improvement on the existing algorithm (SAMC),
+on a single system (square well fluid).
 
@@
 *Jordan Says we could add* **We have modified the conclusion to better emphasize
-this important distication**
+this important distinction (also our method is based on SAMC and is a signifcant
+improvement on that method)**
 
@@
 to WL and introduces STMD.
+**We provide as argument multiple counter examples of papers accepted by PRE
+(the journal to which we are submitting) that test on a single system (not
+necessarily even a system of physical interest). The novelty of our method as
+well as the in depth comparison with WL, 1/t-WL, and SAMC make our paper a
+considerable and 'substantial' addition to the field.
+
+Additionally it is quite obvious that our method (SAD) is far superior from Fig.
+3 and 4 to the method that it is based on (SAMC). Our method is related to
+1/t-WL but hardly an extension of the method as the referee claims; however, it
+performs as well as 1/t-WL and SAD is solving a significantly harder problem.
+Also, SAD has fewer  parameters to chose (only one) and many researchers will
+find temperature easier to determine than the energy range of the system.
+**
 
@@
 
-6. The referee asks why we chose a maximum temperature of infinity.  We have
-added a paragraph to Section III explaining the reasoning, which is that
+6. The referee asks why we chose a maximum temperature of infinity.  *We have
+added a paragraph to Section III explaining the reasoning*, which is that
 infinite temperature allows the system to randomize more reliably, and costs us
@@
 there any evidence?
+*Jordan Says "How do we want to address this" Probably easiest just to remove?*
 
@@
 suggested.**
-In the introduction, we improve the clarity of the paper by defining t as the number
-of moves when first introducing 1/t-WL.
+In the introduction, we improve the clarity of the paper by defining t as the
+number of moves when first introducing 1/t-WL.
 As the first referee suggests we edited phrases such as Monte Carlo and the
@@
 suggest several places where the paper could be made more clear.
+**The referee clearly acknowledges SAMC and that our method is SAD converges
+faster than SAMC. The referee also notes that the paper makes an important
+contribution.**
+
+**----------------------------------------------------------------------------**
 
@@
 between methods.  We have added ...
+**We have better described the implementation of each method in more detail in
+section II to highlight the major differences in implementation between the
+methods. The update factor is now shown for each method so that it can be seen
+how each method differs.**
+
+**----------------------------------------------------------------------------**
 
@@
 fully described, as well as the basis for Equation 7.
+**We have added a better description of the relationship between entropy and the
+update factor and added a paragraph describing the basis of equation 7.**
+
+**----------------------------------------------------------------------------**
 
@@
 method in our plots.
+**We have edited the line styles to better highlight the SAD method in our plots.**
+
+**----------------------------------------------------------------------------**
 

### 2019-08-01 papers/histogram/sad.tex
various paper(s) edits.

--- 2019-07-17 papers/histogram/sad.tex
@@
 limited to a narrow temperature range because a single canonical simulation
-explores only a small range of energies.  This led to a variety of ``flat'' (or
+explores only a small range of energies. Berg and Neuhaus developed multicanonical
+methods that introduced a weight function to enable flat histogram sampling
+which improved the exploration of configuration space and allows the system
+to
+overcome free energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
+These works led to an explosion in the interest of
+``flat'' (or
 ``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
@@
 %
+
 Belardinelli and Pereyra demonstrated that allowing the update factor to
@@
 This leads to their $1/t$-WL algorithm which ensures that the error continues to
-decrease asymptotically as $1/\sqrt{t}$~\cite{belardinelli2008analysis}.
+decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids the
+error saturation and asymptotically approaches the true density of
+states~\cite{belardinelli2008analysis}.
+Zhou et. al further
+confirmed that the original WL algorithm never converges exponentially and
+succesfully bounded the statistical error between $t^{-\frac12}$ and
+$1/t$~\cite{zhou2008optimal}.
+Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
+including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
 
@@
 
+Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and
+the related Statistical Temperature Molecular Dynamics (STMD) which is an
+improvement on the WL method which approximates the entropy (or log density of states)
+as a piecewise
+linear function, which improves convergence for systems with a continuously varying
+energy~\cite{kim2006statistical, kim2007statistical}. STMC
+applied to WL requires a temperature range be specified rather than an energy
+range.  Kim \emph{et al.} extended this work as Replica Exchange Statistical
+Temperature Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping
+STMC simulations to improve convergence~\cite{kim2009replica}.
+More recently, Junghans \emph{et al.} have demonstrated a close
+connection between metadynamics, which was introduced by Laio and
+Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo
+methods, with STMD forging the connection~\cite{junghans2014molecular}.
+
 In this work, we have developed an improved algorithm based on SAMC that does
 not require an array of non-physical, user-defined inputs and therefore should
-be easily applicable to any system. The method does require the user to define a
+be easily applicable to any system. The method (like STMC above) does require the user to define a
 temperature range of interest ($T_\text{min}$ to $T_\infty$).  We call this
@@
 \section{Flat histogram methods}\label{sec:histogram}
+\jpsays{REFEREE 2: The description of flat histogram methods in section II was a bit hard
+to understand at first and may need to be edited to better highlight
+the differences between the methods. Also, the entropy argument in
+section II.D is unclear. The relationship between entropy, the
+histogram, and the update factor needs to be more fully described and
+the basis for equation 7 (and the approximations made there) needs to
+be made clearer.}
 
@@
 
-Flat histogram methods calculate the density of states $D(E)$ for a
-discrete set of energies~\cite{haber2018performance}. For this reason,
-energy binning becomes an important consideration for systems with a
-continuum of possible energies.  Energy bins are typically of uniform
-size for the entire energy continuum~\cite{fasnacht2004adaptive}. Some
-methods such as AdaWL~\cite{koh2013dynamically} employ a tunable
-mechanism for controlling the binning for low entropic states in order
-to ensure the exploration of all energies.  The method introduced in
-this paper is designed to scale appropriately as bin size is changed,
-but we do not test this scaling, as we use a system with discrete
-energy levels.
+Flat histogram methods calculate the density of states $D(E)$ for a discrete set
+of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
+trebst2004optimizing}. For this reason, energy binning becomes an important
+consideration for systems with a continuum of possible energies.  Energy bins
+are typically of uniform size for the entire energy
+continuum~\cite{fasnacht2004adaptive}. Some methods such as
+AdaWL~\cite{koh2013dynamically} employ a tunable mechanism for controlling the
+binning for low entropic states in order to ensure the exploration of all
+energies.  The method introduced in this paper is designed to scale
+appropriately as bin size is changed, but we do not test this scaling, as we use
+a system with discrete energy levels.
 
 In this section we will introduce four closely related flat histogram
-methods which each rely on a weight function $w(E)$.  In these
+methods each of which rely on a weight function $w(E)$.  In these
 algorithms, the probability of accepting a move is given by
@@
 set of weights that are proportional to the density of states of the
-system $D(E)$ will result in an entirely flat histogram.  Thus flat
-histogram is a criteria for convergence for these methods.  To avoid
-overflow error, since the weights may vary over more than a few
-hundred orders of magnitude, the natural logarithm of the weights are
-stored.  Since in the microcanonical ensemble the entropy is defined
-as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights can be
-thought of as an approximation of the entropy.
+system $D(E)$ will result in an entirely flat histogram, which means that the
+weights should converge to being proportional to the density of states.  The
+natural logarithm of the weights are typically stored, since the density of
+states will often vary over a few hundred orders of magnitude.
+In the microcanonical ensemble the entropy is defined
+as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights
+is an approximation of the entropy.
 
@@
 estimate the density of states.  The core of these approaches
-is to continuously update the weights at each step of the simulation
+is to update the weights at each step of the simulation
 \begin{equation}
@@
 where $t$ is number of the current move, $\gamma_t(t)$ is a
-move-dependent update factor, and $E$ is the current energy.  This
+update factor that varies over the simulation, and $E$ is the current energy.  This
 update causes the random walk to avoid energies that have been
-frequently sampled, leading to a rapid exploration of energy space.
-This approach, however, violates detailed balance, due to the
-acceptance probabilities changing with each move.  The severity of
+frequently sampled, enabling a rapid exploration of energy space.
+This approach violates detailed balance, due to the
+acceptance probabilities changing with each move, but the severity of
 this violation decreases as we decrease $\gamma_t$.  The four methods
-differ primarily in how they schedule the decreasing of $\gamma_t$.
+differ primarily in how they schedule the decrease of $\gamma_t$.
 
@@
   landau2014guide} begins with $\gamma_t=1$, and then decreases $\gamma$
-in discrete stages.  We track the number of moves ending at each
-energy in a histogram.  When that histogram is sufficiently flat,
+in discrete stages.  When the energy histogram is sufficiently flat,
 $\gamma_t$ is decreased by a specified factor of $\frac12$.  The
@@
 in this energy range might never be
-sampled~\cite{haber2014transition}.  The entire process is repeated
+sampled~\cite{troster2005wang}.  The entire process is repeated
 until $\gamma_t$ reaches a desired cutoff.
@@
 
+\jpsays{remove and add to intro as this does not convey how to implement WL!}
 While this approach is very efficient and has been widely used, it
@@
 \subsection{$1/t$-Wang-Landau}
-The error saturation of Wang-Landau algorithm algorithm can be
-corrected by modifying the update factor such that it does not
-decrease too quickly.  Belardinelli and Pereyra raised the question as
-to whether the update factor should be decrease by $\nicefrac12$ or
-some other factor for each update~\cite{belardinelli2007fast}. They
-implemented a schedule that enforced that if the update factor
-$\gamma_t$ is ever less than $\nicefrac{1}{t}$ then the update factor
-is set to $\nicefrac{1}{t}$ and the histogram is no longer tracked. Employing this
-scheduler, they found that the error saturation is avoided since the
-correct density of states is approached asymptotically as
-$t^{-\frac12}$~\cite{belardinelli2008analysis}. Zhou et. al further
-confirmed that the WL algorithm never converges exponentially and
-succesfully bounded the statistical error between $t^{-\frac12}$ and
-$1/t$~\cite{zhou2008optimal}.
 
-Schneider et. al outline minor refinements to the $1/t$-WL algorithm
-including $\nicefrac{N_S}{t}$ scaling and switching from standard WL
-to $1/t$-WL when the update factor $\gamma_t <
-\nicefrac{N_S}{t}$~\cite{schneider2017convergence}. As per the
-original $1/t$ implementation~\cite{belardinelli2007fast}, the update
-factor is decreased once all $N_S$ energy states are visited at least
-once (i.e. $H(E) \neq 0$) effectively avoiding the concept of
-`flatness'.  In this work, we refer to this refined algorithm as $1/t$-WL,
-with the update factor asymptoting to $\nicefrac{N_S}{t}$.
+The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
+from dropping below $N_S/t$~\cite{belardinelli2008analysis, schneider2017convergence}.
+The method follows the standard WL algorithm with two modifications.  Firstly, the histogram
+is considered flat, and $\gamma_t$ is decreased by a factor of two, when every
+energy state has been visited once, i.e. when the WL ``flatness'' becomes nonzero.
+Secondly, when the $\gamma_t$ becomes less than $N_S/t$, the algorithm switches
+to use
+\begin{align}
+  \gamma_t = \begin{cases}
+     \gamma_{WL} & \gamma_{WL} > \frac{N_S}{t} \\
+     \frac{N_S}{t} & \text{forever more}
+ \end{cases}
+\end{align}
+where $t$ is the number of moves and $N_S$ is the number of energy bins.
+\jpsays{David and I got to here!}
 
@@
 the entropy may be known, with an unknown prefactor which is roughly
-the coherence time of the Monte Carlo simulation.  Since $\gamma_t$ is
+the coherence time of the Monte Carlo simulation \jpsays{Michael Perlin asked about this in his comments and my understanding was that we meant the time for $\gamma_t$ to reach 1/t behavior.
+Since this is part of a section needing editing I thought we should clarify}.
+Since $\gamma_t$ is
 given by $t_0/t$, the time to converge to a given accuracy is increased
@@
 an energy range of interest or the unphysical parameter $t_0$.
+
+We set the maximum temperature of interest as $T=\infty$ for several reasons.
+Firstly, when this temperature is reached, the system is maximally randomized.
+Allowing infinite temperature further ensures that th system can surmount any
+size of energy barrier.  Finally, the high temperatures are ``easy'' to simulate,
+and tend to converge very quickly.  We note, however, that this is under the
+assumption that infinite temperature corresponds to a finite energy, as is the
+case for configurational sampling.  Were we to include kinetic energy in a
+simulation (as in molecular dynamics), or to examine a system such as a high
+density Lennard-Jones fluid which might have an energy of maximum entropy far
+from the energy range of physical interest, then we would need to apply a maximum
+temperature, which could slow down the rate at which state space is explored.
 
@@
 \section{Conclusions}
+\jpsays{We have introduced SAD, a new variant of SAMC, that}
 We have introduced a new algorithm that effectively samples the

### 2019-08-06 papers/histogram/sad-referee-response.md
made some minor changes and added notes to fix paper where we respond to referee(s).

@@
 method(s), the physics, or both.' As stated above it is only necessary that the
-method is novel OR the physical problem is novel. The paper below accepted by
-PRE confirms this thought.**
+method is novel OR the physical problem is novel. As example, the paper below
+accepted by PRE confirms this thought.**
 
@@
 This paper compares the performance of two methods (WL and 1/t-WL) to calculate
-different integrals (not a real physical system because the exact answer is
+different integrals (not a real physical system chosen because the exact answer is
 known).
@@
 **We provide as argument multiple counter examples of papers accepted by PRE
-(the journal to which we are submitting) that test on a single system (not
+(the journal to which we are submitting not PRL) that test on a single system (not
 necessarily even a system of physical interest). The novelty of our method as
@@
 Also, SAD has fewer  parameters to chose (only one) and many researchers will
-find temperature easier to determine than the energy range of the system.
+find temperature more convenient than the energy range of the system.
 **
@@
 suggest several places where the paper could be made more clear.
-**The referee clearly acknowledges SAMC and that our method is SAD converges
-faster than SAMC. The referee also notes that the paper makes an important
-contribution.**
+**The referee clearly acknowledges SAMC and that our method SAD converges
+significantly faster than SAMC. The referee also notes that the paper makes an
+important contribution.**
 

### 2019-08-06 papers/histogram/ising.tex

@@
 Fig.~\ref{fig:n32} shows the average error in the entropy as a function of time for
-this system with the displacement distance of $0 = 0.05$. The solid/dashed lines
+this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed lines
 represent the average of the absolute value of the error in the entropy averaged
@@
 Fig.~\ref{fig:n128} shows the average error in the entropy as a function of time
-for this system with the displacement distance of $0 = 0.05$. The solid/dashed
+for this system with the displacement distance of $\delta_0 = 0.05$. The solid/dashed
 lines represent the average of the absolute value of the error in the entropy
@@
 We have examined the convergence of SAD versus other widely used weight-based
-histogram methods  We have shown that Sad samples the energy space corresponding
-to a desired range of temperatures for a variety of system sizes. We find that
-both SAD and 1/t-WL demonstrate excellent and robust convergence. They both
-converge more rapidly than SAMC, and unlike WL consistently converge to the
-correct density of states. We find that for larger systems SAD reduces the updated
-factor more slowly (and conservatively) than 1/t-WL. This means that SAD will take
-proportionately more moves to converge to the same value as 1/t-WL as system size is
-increased.
-SAD requires the user to specify a temperature range of interest rather than an
-energy range of interest as 1/t-WL does. For use cases in which a range of
-desired temperatures is known, this will make the SAD method considerably more
-convenient.
+histogram methods.  We have shown that Sad samples the energy space
+corresponding to a desired range of temperatures for a variety of system sizes.
+We find that both SAD and 1/t-WL demonstrate excellent and robust convergence.
+They both converge more rapidly than SAMC, and unlike WL do not suffer from
+error saturation. We find that for larger systems SAD reduces the updated factor
+more slowly (and conservatively) than 1/t-WL. This means that SAD will take
+proportionately more moves to converge to the same value as 1/t-WL as system
+size is increased. SAD requires the user to specify a temperature range of
+interest rather than an energy range of interest as 1/t-WL does. For use cases
+in which a range of desired temperatures is known, this will make the SAD method
+considerably more convenient.
 

### 2019-08-06 papers/histogram/sad.tex

@@
   of energies.  Thus SAD is more convenient when the range of energies is
-  not known in advance.
+  not known in advance. \jpsays{We believe that a
+  temperature range is \emph{far} more physical and convenient, and this is the primary
+  value of our paper.  We have modified the abstract to better highlight this
+  contribution.}
 \end{abstract}
@@
 \subsection{SAMC convergence time}\label{sec:samc-convergence}
+\jpsays{a better description of the relationship between entropy and the
+update factor and added a paragraph describing the basis of equation 7!}
 A primary difficulty in using the SAMC method lies in identifying an
@@
     square-well atoms, for four different methods: WL, $1/t$-WL, SAMC,
-    and SAD}
+    and SAD \jpsays{edit linestyles to highlight SAD}}
     \label{fig:gamma-vs-t}
@@
 the schedule for $\gamma_t$ automatically responds to the choice of bin
-size. SAD should perform similarly over a wide range of bin sizes
+size. \jpsays{Do we just remove this?} SAD should perform similarly over a wide range of bin sizes
 because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
@@
                for the same simulations.
+           \jpsays{edit linestyle to highlight SAD!}
   }\label{fig:fast-slow-gamma}
@@
                semi-transparent shaded area.  The update factor for this
-               system is in Fig.~\ref{fig:gamma-vs-t} above.}\label{fig:n256}
+               system is in Fig.~\ref{fig:gamma-vs-t} above. \jpsays{edit linestyle to highlight SAD!}}\label{fig:n256}
 \end{figure}

### 2019-08-07 papers/histogram/sad-referee-response.md
changes to paper in response to referees

@@
 
-**We argue that our method is based on SAMC (as we state) and is a substantial
-improvement on the method (convergence is of 1 - 2 orders of magnitude superior
-as can be seen from Fig. 3 and Fig. 4).
-
-In addition, we also provide a detailed comparison of 4 different methods on the
-square well fluid.
+Our method is based on SAMC (as we clearly state in the paper) and is a substantial
+improvement on this method (convergence is of 1 - 2 orders of magnitude superior
+as can be seen from Fig. 3 and Fig. 4).  In addition, we also provide a detailed
+comparison with 4 different methods on the square well system.
 
@@
 method is novel OR the physical problem is novel. As example, the paper below
-accepted by PRE confirms this thought.**
+accepted by PRE confirms this thought.
 
@@
 The referee further disputes the use of language *'we have introduced a new
-algorithm'*. In response, we argue that this language is commonly used to
-describe MC methods that have their foundation in other methods (SAD is a
-derivative of SAMC). Our choice of language is further validated by other
-researchers in the field that are published in PRE.
-
-**Dynamically optimized Wang-Landau sampling with adaptive trial moves and
-modification factors**
-*(accepted by PRE)*
-This work introduces an improved update scheme as the referee would call it.
-The authors still use language such as *'To test our new algorithm...'*
-
-Still we have changed the language to say *'we have introduced a new variant of
-SAMC called SAD'"'* to better highlight the subtle distinction.
+algorithm'*. This language is commonly used to describe MC methods that have
+their foundation in other methods (as SAD is a derivative of SAMC). However,
+we did revise the paper to clarify in the conclusion that our algorithm is a
+variant of the SAMC method.
 
@@
 there any evidence?
-*Jordan Says "How do we want to address this" Probably easiest just to remove?*
+**We have edited the paper to make clear that we have not tested and confirmed
+the behavior of SAD in systems with a continuum of energies.**
 
@@
 fully described, as well as the basis for Equation 7.
-**We have added a better description of the relationship between entropy and the
-update factor and added a paragraph describing the basis of equation 7.**
+**We have added an additional description of the relationship between entropy,
+ln w, and the update factor immediately after equation 7 to make more clear how
+the total change in entropy connects with the update factor schedule.**
 
@@
 method in our plots.
-**We have edited the line styles to better highlight the SAD method in our plots.**
+**We found that by thickening the SAD line style and thinning SAMC lines we
+could make it considerably easier to distinguish between the different methods
+and to better highlight the SAD method in our plots.**
 

### 2019-08-07 papers/histogram/sad.tex
changes to paper in response to referees

--- 2019-08-01 papers/histogram/sad.tex
@@
   $1/t$-WL, which requires that the user identify the interesting range
-  of energies.  Thus SAD is more convenient when the range of energies is
-  not known in advance.
+  of energies.  Thus SAD is more powerful in the common case in which the range
+  of energies is not known in advance.
 \end{abstract}
@@
 ``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
-swendsen1999transition, wang2001determining, wang2001efficient,
-trebst2004optimizing}, which explore a wider range of energies.  Another
+swendsen1999transition, wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding, yan2003fast,
+trebst2004optimizing,
+belardinelli2007wang, belardinelli2007fast,
+belardinelli2008analysis, belardinelli2014intrinsic, singh2012density, zhou2008optimal,
+schneider2017convergence, liang2006theory, liang2007stochastic,
+kim2006statistical,
+kim2007statistical, kim2009replica, junghans2014molecular}, which explore a wider range of energies.  Another
 benefit, in contrast with low-temperature canonical Monte Carlo, is that these
@@
 requires the user to select the range of energies to be
-studied~\cite{landau2004new}. This adds an additional hurdle to its application
+studied~\cite{landau2004new, wang2001efficient, schulz2003avoiding, yan2003fast}. This adds an additional hurdle to its application
 to systems for which the interesting range of energies is not known \emph{a
@@
 true density of states~\cite{belardinelli2007wang, belardinelli2007fast,
-belardinelli2008analysis, belardinelli2014intrinsic, singh2012density}.
+belardinelli2008analysis, belardinelli2014intrinsic, singh2012density, zhou2008optimal}.
 %
@@
 error saturation and asymptotically approaches the true density of
-states~\cite{belardinelli2008analysis}.
-Zhou et. al further
-confirmed that the original WL algorithm never converges exponentially and
-succesfully bounded the statistical error between $t^{-\frac12}$ and
-$1/t$~\cite{zhou2008optimal}.
+states~\cite{belardinelli2008analysis}. Zhou et. al further confirmed that the
+original WL algorithm never converges exponentially and succesfully bounded the
+statistical error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
 Schneider \emph{et al.} outline minor refinements to the $1/t$-WL algorithm
-including scaling the update factor with the number of energy bins~\cite{schneider2017convergence}.
+including scaling the update factor with the number of energy
+bins~\cite{schneider2017convergence}.
 
@@
 method still has a system specific user-defined parameter which must be tuned
-when applying this algorithm to a new system.
+when applying this algorithm to a new system.  In
+contrast to the WL-based methods, the SAMC method does not require users to
+determine the energy range of interest \emph{a priori}.
 
-Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and
-the related Statistical Temperature Molecular Dynamics (STMD) which is an
-improvement on the WL method which approximates the entropy (or log density of states)
-as a piecewise
-linear function, which improves convergence for systems with a continuously varying
-energy~\cite{kim2006statistical, kim2007statistical}. STMC
-applied to WL requires a temperature range be specified rather than an energy
-range.  Kim \emph{et al.} extended this work as Replica Exchange Statistical
-Temperature Monte Carlo (RESTMC), which uses replica exchange of multiple overlapping
-STMC simulations to improve convergence~\cite{kim2009replica}.
-More recently, Junghans \emph{et al.} have demonstrated a close
-connection between metadynamics, which was introduced by Laio and
-Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo
-methods, with STMD forging the connection~\cite{junghans2014molecular}.
+Kim \emph{et al.} introduced Statistical Temperature Monte Carlo (STMC) and the
+related Statistical Temperature Molecular Dynamics (STMD) which is an
+adaption of the WL method which approximates the entropy (or log density of
+states) as a piecewise linear function, which improves convergence for systems
+with a continuously varying energy~\cite{kim2006statistical,
+kim2007statistical}. STMC applied to WL requires a temperature range be
+specified rather than an energy range.  Kim \emph{et al.} extended this work as
+Replica Exchange Statistical Temperature Monte Carlo (RESTMC), which uses
+replica exchange of multiple overlapping STMC simulations to improve
+convergence~\cite{kim2009replica}. More recently, Junghans \emph{et al.} have
+demonstrated a close connection between metadynamics, which was introduced by
+Laio and Parinello~\cite{laio2002escaping}, and WL-based Monte Carlo methods,
+with STMD forging the connection~\cite{junghans2014molecular}.
 
@@
 not require an array of non-physical, user-defined inputs and therefore should
-be easily applicable to any system. The method (like STMC above) does require the user to define a
-temperature range of interest ($T_\text{min}$ to $T_\infty$).  We call this
-method SAD (Stochastic Approximation with a Dynamic update factor), and will
-discuss it in detail in the methods section. We compare its convergence
-properties with three existing flat histogram methods: WL, $1/t$-WL, and SAMC.
+be easily applicable to any system. The method (like STMC above) does require
+the user to define a temperature range of interest ($T_\text{min}$ to
+$T_\infty$).  We call this method SAD (Stochastic Approximation with a Dynamic
+update factor), and will discuss it in detail in the methods section. We compare
+its convergence properties with three existing flat histogram methods: WL,
+$1/t$-WL, and SAMC.
 
@@
 \section{Flat histogram methods}\label{sec:histogram}
-\jpsays{REFEREE 2: The description of flat histogram methods in section II was a bit hard
-to understand at first and may need to be edited to better highlight
-the differences between the methods. Also, the entropy argument in
-section II.D is unclear. The relationship between entropy, the
-histogram, and the update factor needs to be more fully described and
-the basis for equation 7 (and the approximations made there) needs to
-be made clearer.}
-
 The goal of flat histogram methods (also called \emph{broad histogram}
@@
 The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
-  landau2014guide} begins with $\gamma_t=1$, and then decreases $\gamma$
+  landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases $\gamma^{\text{WL}}$
 in discrete stages.  When the energy histogram is sufficiently flat,
-$\gamma_t$ is decreased by a specified factor of $\frac12$.  The
+$\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$.  The
 flatness is defined by the ratio between the minimum value of the
 histogram and its average value.  When this flatness reaches a
-specified threshold (typically 0.8), the $\gamma_t$ value is decreased
+specified threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased
 and the histogram is reset to zero.  This approach requires that the
@@
 sampled~\cite{troster2005wang}.  The entire process is repeated
-until $\gamma_t$ reaches a desired cutoff.
+until $\gamma^{\text{WL}}$ reaches a desired cutoff.
 
 The Wang-Landau approach thus has three parameters that need be
-specified: the factor by which to decrease $\gamma_t$ when flatness is
+specified: the factor by which to decrease $\gamma^{\text{WL}}$ when flatness is
 acheived, the flatness criterion, and the cutoff that determines when
@@
 flatness criterion can be defined.
-
-\jpsays{remove and add to intro as this does not convey how to implement WL!}
-While this approach is very efficient and has been widely used, it
-suffers a few shortcomings.  Firstly, the set of energies must be
-specified ~\cite{wang2001efficient, schulz2003avoiding, yan2003fast},
-which may require multiple simulations. Secondly, while Wang-Landau
-converges quickly, it does not in general converge to the true density
-of states ~\cite{belardinelli2008analysis, zhou2008optimal}.  It can
-and does decrease $\gamma_t$ so quickly that it will never (for any
-cutoff value) decrease the error in the density of states beyond a
-given nonzero value.
 
@@
 energy state has been visited once, i.e. when the WL ``flatness'' becomes nonzero.
-Secondly, when the $\gamma_t$ becomes less than $N_S/t$, the algorithm switches
-to use
+Secondly, when $\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches
+to use  $\gamma_t = N_S/t$ for the remainder of the simulation:
 \begin{align}
-  \gamma_t = \begin{cases}
-     \gamma_{WL} & \gamma_{WL} > \frac{N_S}{t} \\
-     \frac{N_S}{t} & \text{forever more}
+  \gamma_t^{1/t\text{-WL}} = \begin{cases}
+     \gamma^{\text{WL}}_t & \gamma^{\text{WL}}_t > \frac{N_S}{t} \\
+     \frac{N_S}{t} & t \ge t_0
  \end{cases}
 \end{align}
-where $t$ is the number of moves and $N_S$ is the number of energy bins.
-\jpsays{David and I got to here!}
+where $t$ is the number of moves, $\gamma^{\text{WL}}_t$ is the Wang-Landau update factor
+at move $t$, and $N_S$ is the number of energy bins.
 
@@
 original implementation~\cite{liang2007stochastic} in terms of an
-arbitrary tunable parameter $t_0$,
+tunable parameter $t_0$,
 \begin{align}
@@
 \end{align}
-where $\zeta \in \{1,2\}$, Liang has shown that the weights are proven
-to converge to the true density of states~\cite{liang2006theory,
+where $\zeta \in \{1,2\}$, Liang has proven that the weights converge to the
+true density of states~\cite{liang2006theory,
   liang2007stochastic}.  In addition, the energy range need not be
@@
 improved rate of convergence, it adds yet another parameter that must
-be empirically determined, and we have not explored this degree of
+be empirically determined, and we have not explored this additional degree of
 freedom.
@@
 \end{align}
-The minimum number of moves that could converge the entropy is
-the number of moves that will enable the above total entropy change,
+This connects with $\gamma_t$ because the $\ln w(E)$ (our approximation for entropy)
+has $\gamma_t$ added to it on each move.  Thus a minimum number of moves
+that could possibly result in the true entropy $S(E)$ starting with a flat set of
+weights is determined by the total entropy change required.  We can estimate
+this number of moves required by summing the $\gamma_t$,
 which we can approximate using an integral:
@@
   % &= \int_0^{t_{\min}} \gamma(t)dt \\
-  &= t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
+  &\approx t_0 + \int_{t_0}^{t_{\min}} \frac{t_0}{t}dt
   \\
@@
 \end{align}
-which means that the minimum time to convergence grows exponentially
+which means that the minimum time to converge grows exponentially
 as $t_0$ is made smaller.  You \emph{seriously} don't want to underestimate
@@
 in general $\gamma_t$ itself forms a lower bound on the accuracy with which
-the entropy may be known, with an unknown prefactor which is roughly
-the coherence time of the Monte Carlo simulation \jpsays{Michael Perlin asked about this in his comments and my understanding was that we meant the time for $\gamma_t$ to reach 1/t behavior.
-Since this is part of a section needing editing I thought we should clarify}.
-Since $\gamma_t$ is
+the entropy may be known, with an unknown prefactor which is related to
+the coherence time of the Monte Carlo simulation.  Since $\gamma_t$ is
 given by $t_0/t$, the time to converge to a given accuracy is increased
@@
     square-well atoms, for four different methods: WL, $1/t$-WL, SAMC,
-    and SAD}
+    and SAD.}
     \label{fig:gamma-vs-t}
@@
 the schedule for $\gamma_t$ automatically responds to the choice of bin
-size. SAD should perform similarly over a wide range of bin sizes
+size. Although we have not yet tested this,
+SAD should perform similarly over a wide range of bin sizes
 because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
@@
 \section{Conclusions}
-\jpsays{We have introduced SAD, a new variant of SAMC, that}
-We have introduced a new algorithm that effectively samples the
+We have introduced a new algorithm, a variant of the SAMC method, which
+effectively samples the
 energy space corresponding to a desired range of temperatures for a

### 2019-08-08 papers/histogram/sad-referee-response.md
added more examples of papers that test on only a single system.

@@
 
-Furthermore, the guidelines that the referee is using presumably come from
-Computational Physics Papers in the Physical Review E (2013). This specifically
-states '...Rather, the requirement that "papers must contain new results in
-physics" should be interpreted as a requirement of novelty with respect to the
-method(s), the physics, or both.' As stated above it is only necessary that the
-method is novel OR the physical problem is novel. As example, the paper below
-accepted by PRE confirms this thought.
-
-**Analysis of the convergence of the 1/t and Wang-Landau algorithms in the
-calculation of multidimensional integrals** *(accepted by PRE)*
-This paper compares the performance of two methods (WL and 1/t-WL) to calculate
-different integrals (not a real physical system chosen because the exact answer is
-known).
-
 Firstly, the referee criticizes our paper for testing the method on a single
@@
 additional different physical systems but dispute that this is necessary for the
-first paper introducing this method.
+first paper introducing an improved method.
+
+For instance, our citations
+[10,15,18,19,22] are all papers (the first three published in PRE **?**) which
+examine the convergence properties of flat histogram methods using a single
+physical system.
+
+**Avoiding boundary effects in Wang-Landau sampling** *(accepted by PRE)*
+This paper with DP Landau as an author considers a single system (the 2D Ising
+model). The novelty of the paper is providing a scheme to avoid boundary effects
+in WL. [10]
+
+**Fast Calculation of the Density of States of a Fluid by Monte Carlo Simulations**
+*(accepted by PRL)*
+This paper compares two new methods with WL on the truncated Lennard-Jones system. [11]
+
+**Optimizing the ensemble for equilibration in broad-histogram Monte Carlo simulations**
+*(accepted by PRE)*
+This paper presents an adaptive algorithm and compares with WL on 2D Ising systems. [12]
 
@@
 different integrals (not a real physical system because the exact answer is
-known).
-
-**Avoiding boundary effects in Wang-Landau sampling** *(accepted by PRE)*
-This paper with DP Landau as an author considers a single system (the 2D Ising
-model). The novelty of the paper is providing a scheme to avoid boundary effects
-in WL.
+known). [15]
 
@@
 This paper examines the optimal modification factor for the WL algorithm and
-tests on a single system (the 2D Ising model).
+tests on a single system (the 2D Ising model). [18]
 
@@
 *(accepted by Computer Physics Communications)*
-This paper compares SAMC, WL, and 1/t-WL on a single system (the 2D Ising model).
+This paper compares SAMC, WL, and 1/t-WL on a single system (the 2D Ising model). [19]
 
@@
 SAMC and WL are compared for a single continuum model with a square well type
-potential.
+potential. [22]
+
+**Direct calculation of liquid–vapor phase equilibria from transition
+matrix Monte Carlo simulation** *J. Chem. Phys.*
+This paper introduces a new method (adaptation of Fitzgerald et al.) TMMC and tests
+on Lennard-Jones fluid. [34]
+
+**Performance Limitations of Flat-Histogram Methods**
+*(accepted by PRL)*
+This paper examines optimal scaling of flat histogram methods on 2D Ising systems. [35]
+
+**Flat Histogram Methods for Quantum Systems: Algorithms to Overcome Tunneling Problems
+and Calculate the Free Energy**
+*(accepted by PRL)*
+This paper examines the generalization of WL algorithm on a spin-1/2 Heisenberg model. [36]
+
+**Adaptive integration method for Monte Carlo simulations**
+*(accepted by PRE)*
+This paper introduces an adaptive sampling method and tests on Lennard-Jones. [37]
+
+**Dynamically optimized Wang-Landau sampling with adaptive trial moves and
+modification factors**
+*(accepted by PRE)*
+This paper introduces dynamically optimized WL sampling and tests on XY-Model. [38]
 

### 2019-08-08 papers/histogram/sad.tex
added more examples of papers that test on only a single system.

@@
 belardinelli2008analysis, belardinelli2014intrinsic, singh2012density, zhou2008optimal,
-schneider2017convergence, liang2006theory, liang2007stochastic,
+schneider2017convergence, liang2006theory, liang2007stochastic, werlich2015stochastic,
 kim2006statistical,
@@
 determine the energy range of interest \emph{a priori}.
+Werlich \emph{et al.} proposed the introduction of an additional tuning parameter
+into SAMC~\cite{werlich2015stochastic}.
 

### 2019-08-20 papers/histogram/sad.tex
made paper corrections as comments by Tanner and Michael.

@@
 algorithms have been developed which calculate the thermodynamic properties of
-various systems for over a range of temperatures.  This development began with
-the original histogram method, which used a single canonical Monte Carlo
-simulation to predict properties for nearby
-temperatures~\cite{ferrenberg1988new}.  For large systems this approach is
-limited to a narrow temperature range because a single canonical simulation
-explores only a small range of energies. Berg and Neuhaus developed multicanonical
-methods that introduced a weight function to enable flat histogram sampling
-which improved the exploration of configuration space and allows the system
-to
-overcome free energy barriers~\cite{berg1991multicanonical, berg1992multicanonical}.
-These works led to an explosion in the interest of
-``flat'' (or
-``broad'') histogram methods~\cite{penna1996broad, penna1998broad,
-swendsen1999transition, wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding, yan2003fast,
-trebst2004optimizing,
-belardinelli2007wang, belardinelli2007fast,
-belardinelli2008analysis, belardinelli2014intrinsic, singh2012density, zhou2008optimal,
-schneider2017convergence, liang2006theory, liang2007stochastic, werlich2015stochastic,
-kim2006statistical,
-kim2007statistical, kim2009replica, junghans2014molecular}, which explore a wider range of energies.  Another
-benefit, in contrast with low-temperature canonical Monte Carlo, is that these
-approaches cannot be trapped in a local energy minimum.
+various systems over a range of temperatures.  This development began with the
+original histogram method, which used a single canonical Monte Carlo simulation
+to predict properties for nearby temperatures~\cite{ferrenberg1988new}.  For
+large systems this approach is limited to a narrow temperature range because a
+single canonical simulation explores only a small range of energies. Berg and
+Neuhaus developed multicanonical methods that introduced a weight function to
+enable flat histogram sampling which improved the exploration of configuration
+space and allows the system to overcome free energy
+barriers~\cite{berg1991multicanonical, berg1992multicanonical}. These works led
+to an explosion in the interest of ``flat'' (or ``broad'') histogram
+methods~\cite{penna1996broad, penna1998broad, swendsen1999transition,
+wang2001determining, wang2001efficient, landau2004new, schulz2003avoiding,
+yan2003fast, trebst2004optimizing, belardinelli2007wang, belardinelli2007fast,
+belardinelli2008analysis, belardinelli2014intrinsic, singh2012density,
+zhou2008optimal, schneider2017convergence, liang2006theory, liang2007stochastic,
+liang2009improving, werlich2015stochastic, kim2006statistical,
+kim2007statistical, kim2009replica, junghans2014molecular}, which explore a
+wider range of energies.  \jpsays{In addition to simulating a range of
+temperatures}, in contrast with low-temperature canonical Monte Carlo, is that
+these approaches \jpsays{avoid being} trapped in a local energy minimum.
 
@@
 states of a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has a few disadvantages. Firstly, it
-requires the user to select the range of energies to be
-studied~\cite{landau2004new, wang2001efficient, schulz2003avoiding, yan2003fast}. This adds an additional hurdle to its application
-to systems for which the interesting range of energies is not known \emph{a
-priori}.  The simulation violates detailed balance, albeit briefly as the size of
-the violation decreases with time, which complicates convergence analysis.  In
-fact, the error in a WL computation has been demonstrated to saturate
-at a non-zero value~\cite{yan2003fast}, i.e. the method does not converge to the
-true density of states~\cite{belardinelli2007wang, belardinelli2007fast,
-belardinelli2008analysis, belardinelli2014intrinsic, singh2012density, zhou2008optimal}.
-%
+the method is incredibly powerful, it has a few disadvantages. \jpsays{The most
+significant disadvantage is that the method} requires the user to select the
+range of energies to be studied~\cite{landau2004new, wang2001efficient,
+schulz2003avoiding, yan2003fast}. This \jpsays{requirement} adds an additional
+hurdle to its application to systems for which the interesting range of energies
+is not known \emph{a priori}.  The simulation violates detailed balance, albeit
+briefly as the size of the violation decreases with time, which complicates
+convergence analysis.  In fact, the error in a WL computation has been
+demonstrated to saturate at a non-zero value~\cite{yan2003fast}, i.e. the method
+does not converge to the true density of states~\cite{belardinelli2007wang,
+belardinelli2007fast, belardinelli2008analysis, belardinelli2014intrinsic,
+singh2012density, zhou2008optimal}.
 
 Belardinelli and Pereyra demonstrated that allowing the update factor to
-decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang}
+decrease faster than $1/t$ leads to nonconvergence~\cite{belardinelli2007wang},
 where $t$ corresponds to the number of moves.
-%
+
 This leads to their $1/t$-WL algorithm which ensures that the error continues to
@@
 
-Liang began to consider that WL could be considered a special case of the
-Stochastic Approximation approach, whose convergence could be mathematically
+Liang began to consider \jpsays{Stochastic Approximation as a mathematical
+generalization of the WL approach}, whose convergence could be mathematically
 proven~\cite{liang2006theory}. In 2007, Liang, Liu, and Carrol \emph{et al.}
-developed Stochastic Approximation Monte Carlo
-(SAMC)~\cite{liang2007stochastic}, and proved its convergence, although the
-method still has a system specific user-defined parameter which must be tuned
-when applying this algorithm to a new system.  In
-contrast to the WL-based methods, the SAMC method does not require users to
-determine the energy range of interest \emph{a priori}.
-Werlich \emph{et al.} proposed the introduction of an additional tuning parameter
-into SAMC~\cite{werlich2015stochastic}.
+developed Stochastic Approximation Monte Carlo (SAMC)~\cite{liang2007stochastic,
+liang2009improving}, and proved its convergence, although the method still has a
+system specific user-defined parameter which must be tuned when applying this
+algorithm to a new system.  In contrast to the WL-based methods, the SAMC method
+does not require users to determine the energy range of interest \emph{a
+priori}. Werlich \emph{et al.} proposed the introduction of an additional tuning
+parameter into SAMC~\cite{werlich2015stochastic}.
 
@@
 related Statistical Temperature Molecular Dynamics (STMD) which is an
-adaption of the WL method which approximates the entropy (or log density of
+adaption of the WL method \jpsays{that} approximates the entropy (or log density of
 states) as a piecewise linear function, which improves convergence for systems
@@
 \end{equation}
-which biases the simulation in favor of energies with low weights.  A
-set of weights that are proportional to the density of states of the
-system $D(E)$ will result in an entirely flat histogram, which means that the
-weights should converge to being proportional to the density of states.  The
-natural logarithm of the weights are typically stored, since the density of
-states will often vary over a few hundred orders of magnitude.
-In the microcanonical ensemble the entropy is defined
-as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the logarithm of the weights
-is an approximation of the entropy.
+which biases the simulation in favor of energies with low weights.  A set of
+weights that are proportional to the density of states of the system $D(E)$ will
+result in an entirely flat histogram, which means that the weights should
+converge to being proportional to the density of states.  The natural logarithm
+of the weights are typically stored, since the density of states will often vary
+over a few hundred orders of magnitude. In the microcanonical ensemble the
+entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
+logarithm of the weights is an approximation of the entropy.
 
-Each approach uses a random walk in energy space to
-estimate the density of states.  The core of these approaches
-is to update the weights at each step of the simulation
+Each approach uses a random walk in energy space to estimate the density of
+states.  The core of these approaches is to update the weights at each step of
+the simulation
 \begin{equation}
@@
 \end{equation}
-where $t$ is number of the current move, $\gamma_t(t)$ is a
-update factor that varies over the simulation, and $E$ is the current energy.  This
-update causes the random walk to avoid energies that have been
-frequently sampled, enabling a rapid exploration of energy space.
-This approach violates detailed balance, due to the
-acceptance probabilities changing with each move, but the severity of
-this violation decreases as we decrease $\gamma_t$.  The four methods
-differ primarily in how they schedule the decrease of $\gamma_t$.
+where $t$ is number of the current move, \jpsays{$\gamma_t$} is an update factor
+that varies over the simulation, and $E$ is the current energy.  This update
+causes the random walk to avoid energies that have been frequently sampled,
+enabling a rapid exploration of energy space. This approach violates detailed
+balance, due to the acceptance probabilities changing with each move, but the
+severity of this violation decreases as we decrease $\gamma_t$.  The four
+methods differ primarily in how they schedule the decrease of $\gamma_t$.
 
@@
 The Wang-Landau approach~\cite{wang2001efficient,wang2001determining,
-  landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases $\gamma^{\text{WL}}$
-in discrete stages.  When the energy histogram is sufficiently flat,
-$\gamma^{\text{WL}}$ is decreased by a specified factor of $\frac12$.  The
-flatness is defined by the ratio between the minimum value of the
-histogram and its average value.  When this flatness reaches a
-specified threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased
-and the histogram is reset to zero.  This approach requires that the
-energy range of interest be known in advance, and difficulties can
-occur with this flatness criteria due to the fact that some energies
-in this energy range might never be
-sampled~\cite{troster2005wang}.  The entire process is repeated
-until $\gamma^{\text{WL}}$ reaches a desired cutoff.
+landau2014guide} begins with $\gamma^{\text{WL}}_{t=0}=1$, and then decreases
+$\gamma^{\text{WL}}$ in discrete stages.  When the energy histogram is
+sufficiently flat, $\gamma^{\text{WL}}$ is decreased by a specified factor of
+$\frac12$.  The flatness is defined by the ratio between the minimum value of
+the histogram and its average value.  When this flatness reaches a specified
+threshold (typically 0.8), the $\gamma^{\text{WL}}$ value is decreased and the
+histogram is reset to zero.  This approach requires that the energy range of
+interest be known in advance, and difficulties can occur with this flatness
+criteria due to the fact that some energies in this energy range might never be
+sampled~\cite{troster2005wang}.  The entire process is repeated until
+$\gamma^{\text{WL}}$ reaches a desired cutoff.
 
@@
 The $1/t$-WL algorithm ensures convergence by preventing the $\gamma_t$ factor
-from dropping below $N_S/t$~\cite{belardinelli2008analysis, schneider2017convergence}.
-The method follows the standard WL algorithm with two modifications.  Firstly, the histogram
-is considered flat, and $\gamma_t$ is decreased by a factor of two, when every
-energy state has been visited once, i.e. when the WL ``flatness'' becomes nonzero.
-Secondly, when $\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches
-to use  $\gamma_t = N_S/t$ for the remainder of the simulation:
+from dropping below $N_S/t$~\cite{belardinelli2008analysis,
+schneider2017convergence}. The method follows the standard WL algorithm with two
+modifications.  Firstly, the histogram is considered flat, and $\gamma_t$ is
+decreased by a factor of two, when every energy state has been visited once,
+i.e. when the WL ``flatness'' becomes nonzero. Secondly, when
+$\gamma^{\text{WL}} < N_S/t$ at time $t_0$, the algorithm switches to use
+$\gamma_t = N_S/t$ for the remainder of the simulation:
+\jpsays{Michael prefers removing $t_0$ since he found it confusing with SAMC later.
+replace the formula with this:
+\begin{align}
+\gamma_t^{1/t\text{-WL}} = \min \left( \gamma_t^{WL}, N_S/t \right)
+\end{align}
+This has problems with not defining when all the states have been visited once!}
+
 \begin{align}
@@
 \subsection{SAMC}
-The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses
-the lack of convergence of Wang-Landau's approach with a simple
-schedule by which the update factor $\gamma_t$ is continuously
-decreased~\cite{liang2007stochastic, werlich2015stochastic,
-  schneider2017convergence}.  The update factor is defined in the
-original implementation~\cite{liang2007stochastic} in terms of an
-tunable parameter $t_0$,
+The Stochastic Approximation Monte Carlo (SAMC) algorithm addresses the lack of
+convergence of Wang-Landau's approach with a simple schedule by which the update
+factor $\gamma_t$ is continuously decreased~\cite{liang2007stochastic,
+werlich2015stochastic, schneider2017convergence}.  The update factor is defined
+in the original implementation~\cite{liang2007stochastic} in terms of an tunable
+parameter $t_0$,
 \begin{align}
@@
 \end{align}
-where $\zeta \in \{1,2\}$, Liang has proven that the weights converge to the
-true density of states~\cite{liang2006theory,
-  liang2007stochastic}.  In addition, the energy range need not be
-known a priori.  The time to converge depends only on the choice of
-parameter $t_0$.  Unfortunately, there is no prescription for finding
-an acceptable value for $t_0$, and while the algorithm formally
-converges, for a poor choice of $t_0$ that convergence can be far too
-slow to be practical. Liang \emph{et al.} give a rule of thumb in
-which $t_0$ is chosen in the range from $2N_S$ to $100N_S$ where $N_S$
-is the number of energy bins~\cite{liang2007stochastic}.  Schneider
-\emph{et al.} found that for the Ising model this heuristic is helpful
-for small spin systems, but that larger systems require an even higher
-$t_0$ value~\cite{schneider2017convergence}.  We will describe below
-one case we examined, in which $t_0$ needs to be as much as two orders
-of magnitude higher than the rule of thumb of $100N_S$ in order to
-converge in $10^{12}$ moves.
+where $\zeta > 1$, Liang has proven that the weights converge to the true
+density of states~\cite{liang2006theory, liang2007stochastic,
+liang2009improving}.  In addition, the energy range need not be known a priori.
+The time to converge depends only on the choice of parameter $t_0$.
+Unfortunately, there is no prescription for finding an acceptable value for
+$t_0$, and while the algorithm formally converges, for a poor choice of $t_0$
+that convergence can be far too slow to be practical. Liang \emph{et al.} give a
+rule of thumb in which $t_0$ is chosen in the range from $2N_S$ to $100N_S$
+where $N_S$ is the number of energy bins~\cite{liang2007stochastic}. Schneider
+\emph{et al.} found that for the Ising model this heuristic is helpful for small
+spin systems, but that larger systems require an even higher $t_0$
+value~\cite{schneider2017convergence}.  We will describe below one case we
+examined, in which $t_0$ needs to be as much as two orders of magnitude higher
+than the rule of thumb of $100N_S$ in order to converge in $10^{12}$ moves.
 
-Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t$ by a factor
-$\gamma_0$~\cite{werlich2015stochastic}.  While this may result in an
-improved rate of convergence, it adds yet another parameter that must
-be empirically determined, and we have not explored this additional degree of
+Werlich \emph{et al.} proposed scaling the SAMC \jpsays{$\gamma_t^{\text{SA}}$}
+by a factor $\gamma_0$~\cite{werlich2015stochastic}.  While this may result in
+an improved rate of convergence, it adds yet another parameter that must be
+empirically determined, and we have not explored this additional degree of
 freedom.
@@
 moves required to find the true entropy by considering the total change
-that needs to be made to the entropy.
+that needs to be made to the entropy:
 \begin{align}
-  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}
+  \Delta S_{\text{tot}} \equiv \sum_E S(E) - S_{\min}.
 \end{align}
@@
 
-One might reasonably choose to err by selecting a large $t_0$.
-The rate of convergence is harder to estimate when $t_0$ is large, but
-in general $\gamma_t$ itself forms a lower bound on the accuracy with which
-the entropy may be known, with an unknown prefactor which is related to
-the coherence time of the Monte Carlo simulation.  Since $\gamma_t$ is
-given by $t_0/t$, the time to converge to a given accuracy is increased
-in proportion to the ratio by which we overestimate $t_0$.  Thus, while
-it is exponentially painful to underestimate $t_0$, overestimating by
-several orders of magnitude is also not acceptable.  We should note
-that these extreme limiting cases do not preclude the possibility that there
-is a wide range of $t_0$ values that lead to an acceptable convergence
-rate.
+One might reasonably choose to err by selecting a large $t_0$. The rate of
+convergence is harder to estimate when $t_0$ is large, but in general $\gamma_t$
+itself forms a lower bound on the accuracy with which the entropy may be known,
+with an unknown prefactor which is related to the coherence time \jpsays{define
+coherence time!} of the Monte Carlo simulation.  Since $\gamma_t$ is given by
+$t_0/t$, the time to converge to a given accuracy is increased in proportion to
+the ratio by which we overestimate $t_0$.  Thus, while it is exponentially
+painful to underestimate $t_0$, overestimating by several orders of magnitude is
+also not acceptable.  We should note that these extreme limiting cases do not
+preclude the possibility that there is a wide range of $t_0$ values that lead to
+an acceptable convergence rate.
 
@@
 Firstly, when this temperature is reached, the system is maximally randomized.
-Allowing infinite temperature further ensures that th system can surmount any
-size of energy barrier.  Finally, the high temperatures are ``easy'' to simulate,
-and tend to converge very quickly.  We note, however, that this is under the
-assumption that infinite temperature corresponds to a finite energy, as is the
-case for configurational sampling.  Were we to include kinetic energy in a
-simulation (as in molecular dynamics), or to examine a system such as a high
-density Lennard-Jones fluid which might have an energy of maximum entropy far
-from the energy range of physical interest, then we would need to apply a maximum
-temperature, which could slow down the rate at which state space is explored.
+Allowing infinite temperature further ensures that the system can surmount
+\jpsays{an energy barrier of any size}.  Finally, the high temperatures are
+``easy'' to simulate, and tend to converge very quickly.  We note, however, that
+this is under the assumption that infinite temperature corresponds to a finite
+energy, as is the case for configurational sampling.  Were we to include kinetic
+energy in a simulation (as in molecular dynamics), or to examine a system such
+as a high density Lennard-Jones fluid \jpsays{with an entropy maximum at an
+energy that is far from the energy range of physical interest} which might have
+an energy of maximum entropy far from the energy range of physical interest,
+then we would need to apply a maximum temperature, which could slow down the
+rate at which state space is explored.
 
@@
 each possible energy.  We define $t_L$ to be the last time that we
+\jpsays{Perlin asks: Did you mean to define $t_L$ as the last time you encountered a new energy of interest.}
 encountered an energy that we currently believe is in the energy range
@@
 fraction of $\eta = 0.17$.  For each system we use a reasonable
-root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
+root-mean-square displacement distance \jpsays{what type gaussian?} $\delta_0 = 0.05\sigma$ for
 proposed moves, and for the smaller system we also use an unreasonably

### 2019-08-21 papers/histogram/sad-referee-response.md
made paper corrections as comments by Tanner and Michael.
edits to referee response

--- 2019-08-07 papers/histogram/sad-referee-response.md
@@
 Dear editor,
+
+We appreciate the referees' careful reading of our paper, and feel that their
+feedback has led to a significant improvement.  We have made some significant
+edits to the paper in response to their comments.  These edits were focused on
+providing a broader background with regard to previous work, increasing clarity,
+and general proofreading.  We believe that we have addressed all of the
+referees' concerns with the exception of the unreasonable request by the first
+referee that we present results on additional new systems.
 
@@
 as can be seen from Fig. 3 and Fig. 4).  In addition, we also provide a detailed
-comparison with 4 different methods on the square well system.
-
-Furthermore, the guidelines that the referee is using presumably come from
-Computational Physics Papers in the Physical Review E (2013). This specifically
-states '...Rather, the requirement that "papers must contain new results in
-physics" should be interpreted as a requirement of novelty with respect to the
-method(s), the physics, or both.' As stated above it is only necessary that the
-method is novel OR the physical problem is novel. As example, the paper below
-accepted by PRE confirms this thought.
-
-**Analysis of the convergence of the 1/t and Wang-Landau algorithms in the
-calculation of multidimensional integrals** *(accepted by PRE)*
-This paper compares the performance of two methods (WL and 1/t-WL) to calculate
-different integrals (not a real physical system chosen because the exact answer is
-known).
+comparison with 4 different flat histogram methods on the square well system.
 
@@
 example of a paper which tests a method on three different physical systems.  We
-give as counter examples *references that we cite* which each use a single
-physical system as the test case.  We agree that our method should be tested on
-additional different physical systems but dispute that this is necessary for the
-first paper introducing this method.
-
-**Analysis of the convergence of the 1/t and Wang-Landau algorithms in the
-calculation of multidimensional integrals** *(accepted by PRE)*
-This paper compares the performance of two methods (WL and 1/t-WL) to calculate
-different integrals (not a real physical system because the exact answer is
-known).
-
-**Avoiding boundary effects in Wang-Landau sampling** *(accepted by PRE)*
-This paper with DP Landau as an author considers a single system (the 2D Ising
-model). The novelty of the paper is providing a scheme to avoid boundary effects
-in WL.
-
-**Optimal modification factor and convergence of the Wang-Landau algorithm**
-*(accepted by PRE)*
-This paper examines the optimal modification factor for the WL algorithm and
-tests on a single system (the 2D Ising model).
-
-**Convergence of Stochastic Approximation Monte Carlo and modified Wang–Landau
-algorithms: Tests for the Ising model**
-*(accepted by Computer Physics Communications)*
-This paper compares SAMC, WL, and 1/t-WL on a single system (the 2D Ising model).
-
-**Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a
-continuum polymer model** *(accepted by Computer Physics Communications)*
-SAMC and WL are compared for a single continuum model with a square well type
-potential.
-
-These papers help to highlight that the convergence properties of the various
-Monte Carlo methods are of importance (while the system(s) are often not
-necessarily). In our case, we compare the convergence of *four* different
-methods, one of which (SAD) is an improvement on the existing algorithm (SAMC),
-on a single system (square well fluid).
+give as counter examples twelve of our citations [10-12, 15, 18, 19, 23, 35-39],
+which are all papers (six of which were published in PRE) which examine flat
+histogram methods using a single physical system. Many of these papers only
+present the convergence properties of a single Monte Carlo method. In our case,
+we compare the convergence of *four* different methods, one of which (SAD) is an
+improvement on the existing algorithm (SAMC), on a single system (square well
+fluid).  We also note that this work is the only instance of a test of any flat
+histogram method to the square well fluid.  We agree that our method should be
+tested on additional different physical systems but dispute that this is
+necessary for the first paper introducing an improved method.
 
@@
 (rather than an energy range) as input to the method.  We believe that a
-temperature range is *far* more physical and convenient, and this is the primary
-value of our paper.  **We have modified the abstract to better highlight this
-contribution.**
+temperature range is *far* more physical and convenient, and this is one of the
+primary contributions of our paper.  We have modified the abstract to better
+highlight this contribution.
 
@@
 the referee concludes that our method is not a substantial improvement. We note
-first (as mentioned in the paper in *Section IV* that the 1/t-WL method is
+first (as mentioned in the paper in *Section IV*) that the 1/t-WL method is
 solving a significantly easier problem, because it is told in advance the range
-of energies that are of interest.  We agree, however, that the benefit of our
-method lies specifically in the fact that a temperature range is given as an
-input.  Achieving the same result as the SAD method with 1/t-WL would either
-require multiple simulations over a wider range of energies, or would require
-guessing in advance a wider range of energies than are needed.  In either case,
-the convergence of 1/t-WL would be worse than our tests indicate.
-*Jordan Says we could add* **We have modified the conclusion to better emphasize
-this important distinction (also our method is based on SAMC and is a signifcant
-improvement on that method)**
+of energies that are of interest.  We do agree that the benefit of our method
+lies specifically in the fact that a temperature range is given as an input.
+Achieving the same result as the SAD method with 1/t-WL would either require
+multiple simulations over a wider range of energies, or would require guessing
+in advance a wider range of energies than are needed.  In either case, the
+convergence of 1/t-WL would be worse than our tests indicate. We have modified
+the conclusion to better emphasize this important distinction.
 
@@
 mathematical generalization of WL). In fact, the referee does not seem to
-acknowledge *the existence of the SAMC method* which we find curious. We clearly
-state in the paper in section I *'In this work, we have developed an improved
-algorithm based on SAMC.'*  This is again reinforced in Secion III.
+acknowledge *the existence of the SAMC method* which we find curious. The
+referee states that "the original WL paper does not specify a scheme for
+updating the update factor."  This is false.  Both of the 2001 WL papers specify
+the same scheme that is generally used in when the WL method is used (gamma
+drops by a factor of 2).  It is true that the papers do not *mandate* this
+schedule.  However, neither our method nor SAMC can be considered an
+implementation of the WL method, because the WL method *does* mandate that the
+update factor is only decreased when the histogram has become sufficiently flat.
 
-There are many authors that acknowledge SAMC as a mathematical generalization of
-WL. We list a few here in our response.
-
-**Multidimensional stochastic approximation Monte Carlo**
-*(accepted by PRE)*
-"the Wang-Landau MC [3] scheme and its mathematical generalization, stochastic
-approximation MC (SAMC) [4–6].""
-
-**Convergence estimation of flat-histogram algorithms based on simulation results**
-*(accepted by Computer Physics Communications)*
-"At the same time, Liang suggested a generalized simulation method:stochastic
-approximation Monte Carlo (SAMC) [7,8] including the 1/t-WL as a particular case"
-
-**Stochastic approximation Monte Carlo and Wang–Landau Monte Carlo applied to a
-continuum polymer model**
-*(accepted by Computer Physics Communications)*
-"Liang et al.[30] showed that WLMC could be seen as a version of SAMC and using
-the mathematical background of stochastic approximation methods they proved the
-convergence of SAMC."
+Our method clearly does *not* implement a new update scheme for the WL method.
+We also do not simply implement a new update scheme for SAMC, since SAMC
+specifies an update schedule which is predetermined.  As stated in Section I
+*'In this work, we have developed an improved algorithm based on SAMC.'*  This
+idea is again reinforced in Secion III.
 
@@
 algorithm'*. This language is commonly used to describe MC methods that have
-their foundation in other methods (as SAD is a derivative of SAMC). However,
-we did revise the paper to clarify in the conclusion that our algorithm is a
-variant of the SAMC method.
+their foundation in other methods (as SAD is a derivative of SAMC, and SAMC is
+itself inspired by WL). However, we did revise the paper to clarify in the
+conclusion that our algorithm is a variant of the SAMC method.
 
@@
 2. The referee suggests that we might include references to the originators of
-the multicanonical methods.
-**We have added a paragraph to Section I detailing the important impact of
-multicanonical methods on influencing flat histogram methods. We have referenced
-the original authors Berg and Neuhaus.**
+the multicanonical methods.  We have added a paragraph to Section I detailing
+the important impact of multicanonical methods on influencing flat histogram
+methods. We have referenced the original authors Berg and Neuhaus.
 
@@
 3. The referee suggests that that it would be helpful to include references to
-the Statistical *(not Stochastic)* Temperature Monte Carlo methods.
-**We have added a paragraph to Section I detailing the important impact of STMC
-and RESTMC and its improvement on the WL method and Replica Exchange
-*respectively.**
+the Statistical *(not Stochastic)* Temperature Monte Carlo methods.  We have
+added a paragraph to Section I detailing the important impact of STMC and RESTMC
+and its improvement on the WL method and Replica Exchange respectively.
 
@@
 4. The referee suggests that that it would be valuable to include references to
-STMD and metadynamics.
-**We have added a paragraph to Section I detailing the important impact of STMD
-forging a connection between WL and metadynamics.**
+STMD and metadynamics. We have added a paragraph to Section I detailing the
+important impact of STMD forging a connection between WL and metadynamics.
 
@@
 
-5. The referee argues that our method should be tested on multiple other systems.
-The referee gives for example a single paper accepted by *PRL* that compares STMC
-to WL and introduces STMD.
-**We provide as argument multiple counter examples of papers accepted by PRE
-(the journal to which we are submitting not PRL) that test on a single system (not
-necessarily even a system of physical interest). The novelty of our method as
-well as the in depth comparison with WL, 1/t-WL, and SAMC make our paper a
-considerable and 'substantial' addition to the field.
+5. The referee argues that our method should be tested on multiple other
+systems. The referee gives for example a single paper accepted by *PRL* that
+compares STMC to WL and introduces STMD.  The paper cited by the referee is
+unusual in that it introduces *two* new methods (STMC and STMD), each of which
+is valuable with a different type of system.  As we state above, we do not
+believe this is a reasonable or ordinary expectation.  We cite six PRE papers
+(as well as six other papers) that examine flat histogram methods using a single
+system.
 
-Additionally it is quite obvious that our method (SAD) is far superior from Fig.
-3 and 4 to the method that it is based on (SAMC). Our method is related to
-1/t-WL but hardly an extension of the method as the referee claims; however, it
-performs as well as 1/t-WL and SAD is solving a significantly harder problem.
-Also, SAD has fewer  parameters to chose (only one) and many researchers will
-find temperature more convenient than the energy range of the system.
-**
+The referee mentions the possibility of mentioning ground states.  The goal of
+our method is explicitly to *not* explore the entire energy range, which in most
+cases will mean not either seeking or finding the ground state.
 
@@
 
-Minor details:
+Minor details (according to the referee):
 
@@
 
-7. On page 5 the authors say "SAD should perform similarly ...". Is
-there any evidence?
-**We have edited the paper to make clear that we have not tested and confirmed
-the behavior of SAD in systems with a continuum of energies.**
+7. The referee points out that on page 5 we say "SAD should perform
+similarly..." and asks if there is any evidence? We have edited the paper to
+make clear that we have not tested and confirmed the behavior of SAD in systems
+with a continuum of energies.
 
-8. *Proofreading* **Note all changes made under proof reading, as well as what was
-suggested.**
+8. The referee asks that we do more proofreading of the paper.  We have done so.
 In the introduction, we improve the clarity of the paper by defining t as the
-number of moves when first introducing 1/t-WL.
-As the first referee suggests we edited phrases such as Monte Carlo and the
-update factor for consistency.
-We corrected several typographical errors throughout the paper.
-References were checked for completeness.
+number of moves when first introducing 1/t-WL. As the first referee suggests we
+edited phrases such as Monte Carlo and the update factor for consistency. We
+corrected several typographical errors throughout the paper. References were
+checked for completeness.
 
@@
 
-The second referee was far more positive on the value of our work, but did
-suggest several places where the paper could be made more clear.
-**The referee clearly acknowledges SAMC and that our method SAD converges
-significantly faster than SAMC. The referee also notes that the paper makes an
-important contribution.**
+The second referee states that our paper "makes an important contribution to
+improving algorithms for this type of system."  Furthermore, the second referee
+recognizes the difficulty of setting the timescale parameter in SAMC, and
+acknowledges that our algorithm addresses this challenge.  The referee did have
+several specific questions, which we have numbered below.
 
@@
 section II was confusing, and should be rewritten to highlight the differences
-between methods.  We have added ...
-**We have better described the implementation of each method in more detail in
-section II to highlight the major differences in implementation between the
-methods. The update factor is now shown for each method so that it can be seen
-how each method differs.**
+between methods.  We have edited this section to better describe the
+implementation of each method in more detail in section II, and to highlight the
+major differences between the methods. The update factor is now shown in display
+math for each method so that it can more easily be seen how each method differs.
 
@@
 
-2. The referee shared that the entropy argument in Section II.D is unclear, and
+2. The referee felt that the entropy argument in Section II.D was unclear, and
 wished for the relationship between entropy and the update factor to be more
-fully described, as well as the basis for Equation 7.
-**We have added an additional description of the relationship between entropy,
-ln w, and the update factor immediately after equation 7 to make more clear how
-the total change in entropy connects with the update factor schedule.**
+fully described, as well as the basis for Equation 7. We have added an
+additional description of the relationship between entropy, ln w, and the update
+factor immediately after equation 7 to make more clear how the total change in
+entropy connects with the update factor schedule.
 
@@
 3. The referee asked for a better choice of line styles to highlight the SAD
-method in our plots.
-**We found that by thickening the SAD line style and thinning SAMC lines we
-could make it considerably easier to distinguish between the different methods
-and to better highlight the SAD method in our plots.**
+method in our plots. We changed our line styles to make the comparison plots
+easier to read, to be more effective when printed without color, and to
+highlight the SAD method. We found that by thickening the SAD line style and
+thinning the SAMC lines we could make it considerably easier to distinguish
+between the different methods and to better highlight the SAD method in our
+plots.
 
@@
 4. The referee finally suggested that it might be interesting to try out our
-method with a Go protein model.  We agree that this would be interesting, and
+method with a Go protein model.  We agree that this could be interesting, and
 will consider this suggestion for a future paper.

### 2019-08-21 papers/histogram/sad-referee-response.md
made a number of proofread edits before resubmission.

@@
 is an important one, but does not believe that our work is a *'substantial
-contribution*'' to the field.  The referee has three major criticisms of the
+contribution'* to the field.  The referee has three major criticisms of the
 value of our work (aside from issues with the introduction section).
@@
 *'In this work, we have developed an improved algorithm based on SAMC.'*  This
-idea is again reinforced in Secion III.
+idea is again reinforced in Section III.
 

### 2019-08-21 papers/histogram/sad.tex
made a number of proofread edits before resubmission.

@@
   arbitrary tunable parameter $t_0$ as in the case of SAMC.  SAD requires
-  as input the temperature range of interst, in contrast to
+  as input the temperature range of interest, in contrast to
   $1/t$-WL, which requires that the user identify the interesting range
@@
 kim2007statistical, kim2009replica, junghans2014molecular}, which explore a
-wider range of energies.  \jpsays{In addition to simulating a range of
-temperatures}, in contrast with low-temperature canonical Monte Carlo, is that
-these approaches \jpsays{avoid being} trapped in a local energy minimum.
+wider range of energies. In addition to simulating a range of temperatures, in
+contrast with low-temperature canonical Monte Carlo, is that these approaches
+avoid being trapped in a local energy minimum.
 
@@
 states of a given system~\cite{wang2001determining, wang2001efficient}.  While
-the method is incredibly powerful, it has a few disadvantages. \jpsays{The most
-significant disadvantage is that the method} requires the user to select the
+the method is incredibly powerful, it has a few disadvantages. The most
+significant disadvantage is that the method requires the user to select the
 range of energies to be studied~\cite{landau2004new, wang2001efficient,
-schulz2003avoiding, yan2003fast}. This \jpsays{requirement} adds an additional
+schulz2003avoiding, yan2003fast}. This requirement adds an additional
 hurdle to its application to systems for which the interesting range of energies
@@
 states~\cite{belardinelli2008analysis}. Zhou et. al further confirmed that the
-original WL algorithm never converges exponentially and succesfully bounded the
+original WL algorithm never converges exponentially and successfully bounded the
 statistical error between $t^{-\frac12}$ and $1/t$~\cite{zhou2008optimal}.
@@
 
-Liang began to consider \jpsays{Stochastic Approximation as a mathematical
-generalization of the WL approach}, whose convergence could be mathematically
+Liang began to consider Stochastic Approximation as a mathematical
+generalization of the WL approach, whose convergence could be mathematically
 proven~\cite{liang2006theory}. In 2007, Liang, Liu, and Carrol \emph{et al.}
@@
 related Statistical Temperature Molecular Dynamics (STMD) which is an
-adaption of the WL method \jpsays{that} approximates the entropy (or log density of
+adaption of the WL method that approximates the entropy (or log density of
 states) as a piecewise linear function, which improves convergence for systems
@@
 \end{equation}
-where $t$ is number of the current move, \jpsays{$\gamma_t$} is an update factor
+where $t$ is number of the current move, $\gamma_t$ is an update factor
 that varies over the simulation, and $E$ is the current energy.  This update
@@
 specified: the factor by which to decrease $\gamma^{\text{WL}}$ when flatness is
-acheived, the flatness criterion, and the cutoff that determines when
+achieved, the flatness criterion, and the cutoff that determines when
 the computation is complete.  In addition, an energy range (or in
@@
 $\gamma_t = N_S/t$ for the remainder of the simulation:
-\jpsays{Michael prefers removing $t_0$ since he found it confusing with SAMC later.
-replace the formula with this:
-\begin{align}
-\gamma_t^{1/t\text{-WL}} = \min \left( \gamma_t^{WL}, N_S/t \right)
-\end{align}
-This has problems with not defining when all the states have been visited once!}
-
 \begin{align}
@@
 
-Werlich \emph{et al.} proposed scaling the SAMC \jpsays{$\gamma_t^{\text{SA}}$}
+Werlich \emph{et al.} proposed scaling the SAMC $\gamma_t^{\text{SA}}$
 by a factor $\gamma_0$~\cite{werlich2015stochastic}.  While this may result in
@@
 \begin{align}
-   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{SA} \\
+   \Delta S_{\text{tot}} &= \sum_{t=0}^{t_{\min}} \gamma_t^{\text{SA}} \\
   % &= \int_0^{t_{\min}} \gamma(t)dt \\
@@
 One might reasonably choose to err by selecting a large $t_0$. The rate of
-convergence is harder to estimate when $t_0$ is large, but in general $\gamma_t$
-itself forms a lower bound on the accuracy with which the entropy may be known,
-with an unknown prefactor which is related to the coherence time \jpsays{define
-coherence time!} of the Monte Carlo simulation.  Since $\gamma_t$ is given by
-$t_0/t$, the time to converge to a given accuracy is increased in proportion to
-the ratio by which we overestimate $t_0$.  Thus, while it is exponentially
-painful to underestimate $t_0$, overestimating by several orders of magnitude is
-also not acceptable.  We should note that these extreme limiting cases do not
-preclude the possibility that there is a wide range of $t_0$ values that lead to
-an acceptable convergence rate.
+convergence is harder to estimate when $t_0$ is large, but in general
+$\gamma_t^{\text{SA}}$ itself forms a lower bound on the accuracy with which the
+entropy may be known, with an unknown prefactor which is related to the
+coherence time ($t = t_0$) of the Monte Carlo simulation.  Since
+$\gamma_t^{\text{SA}}$ is given by $t_0/t$, the time to converge to a given
+accuracy is increased in proportion to the ratio by which we overestimate $t_0$.
+Thus, while it is exponentially painful to underestimate $t_0$, overestimating
+by several orders of magnitude is also not acceptable.  We should note that
+these extreme limiting cases do not preclude the possibility that there is a
+wide range of $t_0$ values that lead to an acceptable convergence rate.
 
@@
 Firstly, when this temperature is reached, the system is maximally randomized.
-Allowing infinite temperature further ensures that the system can surmount
-\jpsays{an energy barrier of any size}.  Finally, the high temperatures are
-``easy'' to simulate, and tend to converge very quickly.  We note, however, that
-this is under the assumption that infinite temperature corresponds to a finite
-energy, as is the case for configurational sampling.  Were we to include kinetic
-energy in a simulation (as in molecular dynamics), or to examine a system such
-as a high density Lennard-Jones fluid \jpsays{with an entropy maximum at an
-energy that is far from the energy range of physical interest} which might have
-an energy of maximum entropy far from the energy range of physical interest,
-then we would need to apply a maximum temperature, which could slow down the
-rate at which state space is explored.
+Allowing infinite temperature further ensures that the system can surmount an
+energy barrier of any size.  Finally, the high temperatures are ``easy'' to
+simulate, and tend to converge very quickly.  We note, however, that this is
+under the assumption that infinite temperature corresponds to a finite energy,
+as is the case for configurational sampling.  Were we to include kinetic energy
+in a simulation (as in molecular dynamics), or to examine a system such as a
+high density Lennard-Jones fluid with an entropy maximum at an energy that is
+far from the energy range of physical interest, then we would need to apply a
+maximum temperature, which could slow down the rate at which state space is
+explored.
 
@@
 wish for a lower update factor in order to more rapidly refine the
-remaining error in entropy.  We track the time at which we first visted
+remaining error in entropy.  We track the time at which we first visited
 each possible energy.  We define $t_L$ to be the last time that we
-\jpsays{Perlin asks: Did you mean to define $t_L$ as the last time you encountered a new energy of interest.}
 encountered an energy that we currently believe is in the energy range
@@
 fraction of $\eta = 0.17$.  For each system we use a reasonable
-root-mean-square displacement distance \jpsays{what type gaussian?} $\delta_0 = 0.05\sigma$ for
+root-mean-square displacement distance $\delta_0 = 0.05\sigma$ for
 proposed moves, and for the smaller system we also use an unreasonably
@@
 we find that by limiting the simulation to this energy range, we can
-acheive much faster convergence with a smaller $t_0$.  We compute an average
+achieve much faster convergence with a smaller $t_0$.  We compute an average
 of the error by averaging the error in the entropy over the
@@
 %~ were carried out that all the methods
-%~ struggle to converge if the update factor decreases too rapidlly.
+%~ struggle to converge if the update factor decreases too rapidly.
 %~ Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
@@
 changes to the Monte Carlo moves can have a dramatic effect on the
-range of practical $t_0$ values.  Additionaly, SAMC does not
+range of practical $t_0$ values.  Additionally, SAMC does not
 converge as rapidly as either SAD or $1/t$-WL even for the best choice of $t_0$,

### 2019-08-21 papers/histogram/sad-referee-response.md
fixed references to conform to aps and paper changes.

@@
 improvement on the existing algorithm (SAMC), on a single system (square well
-fluid).  We also note that this work is the only instance of a test of any flat
-histogram method to the square well fluid.  We agree that our method should be
-tested on additional different physical systems but dispute that this is
-necessary for the first paper introducing an improved method.
+fluid).  We also note that to our knowledge this work is the only instance of a
+test of any flat histogram method to the square well fluid.  We agree that our
+method should be tested on additional different physical systems but dispute
+that this is necessary for the first paper introducing an improved method.
 
@@
 updating the update factor."  This is false.  Both of the 2001 WL papers specify
-the same scheme that is generally used in when the WL method is used (gamma
+the same scheme that is generally used for the WL method (the update factor
 drops by a factor of 2).  It is true that the papers do not *mandate* this

### 2019-11-07 papers/histogram/sad.tex
start work collecting error data for LJ31
work on making lj CV error plot work
added some editing notes to self.

@@
 \begin{abstract}
+  \jpsays{edit to include LJ clusters}
   We present a new Monte Carlo algorithm based on the Stochastic
@@
 
+\jpcom{edit some here!}
 As a test case, we consider the square-well fluid i.e. a system of particles
@@
 
+\jpcom{Describe Lennard-Jones clusters we test SAD ability to handle complex
+energy landscapes. We choose 31 and 38 due to complex funnels In the particular
+case of n=38 there exist two funnels at low energies representing octahedral (Oh)
+and icosahedral (ico) energy minima [5] leading to a pronounced size-specific
+behavior at low temperatures, which is especially difficult to simulate. The
+first numerical success}
+
 In this work, we compare four flat histogram methods.  We outline the
@@
 Approximation Monte Carlo (SAMC), and SAD.
+\jpcom{In addition, we test the ability of SAD to accurately and efficiently
+samples low-temperature solid-solid transitions for Lennard-Jones clusters.}
 
@@
 
-%~ For this larger system, SAMC parameter values of $t_0$ that gave rapid
-%~ convergence for the $N = 50$ system with $\delta = 0.05\sigma$ do not
-%~ converge rapidly.
+\subsection{Lennard-Jones clusters with 31 atoms}
 
-%{\color{red}
-%Figure~\ref{fig:n256-lndos-comparison} shows the calculated entropy for
-%the $N = 256$ system. The grey dash lines indicate the $S_{\max} = 509$
-%point and the $S_{\min} = 915$. Only the reference method is shown.
-%}
-
-%~ \subsection{Discussion on convergence and update factor}
-
-%~ A close examination of Fig.~\ref{fig:gamma-vs-t} shows that $\gamma$
-%~ for the Wang-Landau algorithm decreases much more rapidly than
-%~ $\nicefrac{1}{t}$. It can be observed  for the simulations that
-%~ were carried out that all the methods
-%~ struggle to converge if the update factor decreases too rapidly.
-%~ Fig.~\ref{fig:n256} shows SAMC with inappropriate choices
-%~ of $t_0$ i.e. less than $\nicefrac{10^4}{\sqrt{t}}$ failing to converge
-%~ in a reasonable amount of time. Thus, the update factor $\gamma$, shown
-%~ as a function of simulation moves in Fig.~\ref{fig:gamma-vs-t}, proves
-%~ useful in determining which methods will converge in a reasonable
-%~ amount of time. We compare each method to a reference SAMC
-%~ simulation for an ideal choice of $t_0$.  This gives an unbiased
-%~ comparison with our developed method SAD. The dynamic version of SAMC
-%~ which we call SAD performs as well as SAMC (assuming the user can
-%~ correctly identify an appropriate choice of $t_0$ for SAMC).  SAD
-%~ outperforms WL and 1/t-WL? for the same given number of moves. The
-%~ absence of user defined parameters makes SAD an exciting alternative to
-%~ methods where user experience or successive implementation iteration
-%~ are needed to achieve reasonable results.
+\subsection{Lennard-Jones clusters with 38 atoms}
 
@@
 
-%We note that SAD should be tested on additional systems, and the
-%convergence of the method should be examined.  Also, a rigorous
-%mathematical justification for our choice of the update factor is
-%necessary to ensure the most optimal convergence of the algorithm.
+\jpcom{mention something about LJ clusters and what we find!}
 

### 2019-11-08 papers/histogram/sad.tex
change to plot maximum error rather than mean for LJ31
add plot of C_V with converged result

@@
 
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/lj-cv-error}
+  \caption{The maximum error in the heat capacity of the LJ31 cluster
+    in the temperature range from 0.01~$\epsilon/k_B$ to
+    0.1~$\epsilon/k_B$.}
+  \label{fig:lj-cv-error}
+\end{figure}
+
+\begin{figure}
+  \includegraphics[width=\columnwidth]{figs/lj-cv}
+  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
+    as calculated with each method.}
+  \label{fig:lj-cv}
+\end{figure}
+
 \subsection{Lennard-Jones clusters with 31 atoms}
 
-\subsection{Lennard-Jones clusters with 38 atoms}
+The heat capacity changes significantly even at these low temperatures
+when the size of the cavity is increased.  We use a 2.5 cavity because
+``everyone else'' does.
 

### 2019-11-08 papers/histogram/sad.tex
a bit more sloppy text about what we need to write regarding the LJ31 cluster

@@
 when the size of the cavity is increased.  We use a 2.5 cavity because
-``everyone else'' does.
+``everyone else'' does.  For purposes of testing the convergence of
+the algorithm, we don't need to solve for the true heat capacity.
+Furthermore, the heat capacity at fixed volume is not the most
+meaningful quantity to predict for a cluster.
+
+Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
+heat capacity converges just as well for bin widths varying over two
+orders of magnitude.  WL and $1/t$-WL show similar behavior, so this
+is not new, but is encouraging.
+
+Specifying temperature range rather than energy range is beneficial.
+Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
+chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
+that showed them not converging because it chose too low a minimum
+energy, proving that it is \emph{not} a trivial procedure to determine
+an appropriate energy range.  Specifying the temperature range
+\emph{does} matter! We had to do a whole bunch of $1/t$-WL
+computations in order to determine what would be a sufficient
+low-energy cutoff.
 

### 2019-11-09 papers/histogram/sad.tex
start intro to lj31 section

@@
 
+Clusters of Lennard-Jones atoms have been frequently used for testing
+Monte Carlo algorithms~\davidsays{add a few citations}.  The 31-atom
+cluster is popular because it features a two-funnel energy landscape
+with a significant barrier between the two low-energy
+states~\davidsays{add citations}.
+
+Poulain \emph{et al.} found \cite{poulain2006performances}
+
 The heat capacity changes significantly even at these low temperatures

### 2019-11-12 papers/histogram/sad.tex
some more data, and add more caption information to sad.tex

@@
     in the temperature range from 0.01~$\epsilon/k_B$ to
-    0.1~$\epsilon/k_B$.}
+    0.1~$\epsilon/k_B$.  Note that the value of $C_V$ ranges from
+    90$k_B$ to 120$k_B$, so the smallest (maximum) errors seen are
+    around 1\%.  All WL methods shown were run with a histogram bin
+    width of $\Delta E=0.01\epsilon$.  Their convergence at other bin
+    widths was similar to that of SAD. }
   \label{fig:lj-cv-error}

### 2019-11-12 papers/histogram/sad.tex
plot a few more C_V (which still need refining) and work more on LJ31 text.

@@
 
-Poulain \emph{et al.} found \cite{poulain2006performances}
+Poulain \emph{et al.} found very poor convergence for a variety of
+modified WL methods on the LJ31
+cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
+energy range or bin size that was used in their simulation, so their
+results cannot be reproduced.  They do mention that they use 1000
+energy bins, which if they used an energy range of
+$[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
+E\approx 0.13\epsilon$, which we find to be too large to accurately
+predict the heat capacity peak at $0.03\epsilon/k_B$.
 
@@
 when the size of the cavity is increased.  We use a 2.5 cavity because
-``everyone else'' does.  For purposes of testing the convergence of
-the algorithm, we don't need to solve for the true heat capacity.
-Furthermore, the heat capacity at fixed volume is not the most
-meaningful quantity to predict for a cluster.
+``everyone else'' does~\cite{poulain2006performances}.  For purposes
+of testing the convergence of the algorithm, we don't need to solve
+for the true heat capacity.  Furthermore, the heat capacity at fixed
+volume is not the most meaningful quantity to predict for a cluster.
 

### 2019-11-13 papers/histogram/sad.tex
try to figure out what Poulaine et al. actually did, requires a script
clarify Poulaine paragraph
major edits on LJ31 bits, totally new figure

@@
 
-\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
+\newcommand{\davidsays}[1]{{\color{blue} [\green{David:} \emph{#1}]}}
 \newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
@@
   the simulation. We test this method on the square-well fluid and
+  the 31-atom Lennard-Jones cluster and
   compare the convergence time and average entropy error for several
@@
 
-\jpcom{edit some here!}
-As a test case, we consider the square-well fluid i.e. a system of particles
-whose interactions are governed by a square-well
-potential~\cite{singh2003surface, barker2004perturbationSW}.  The square-well
-potential is an ideal test-bed as it is a simple model for a liquid, which
-includes both attractive and repulsive
-interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The potential
-$U(\textbf{r})$ for such a system is given by
-\begin{equation}
- U(\textbf{r})=\begin{cases} \infty &
- \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
- \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
- \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
-\end{equation}
-where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
-reduced range of the potential well, and $\epsilon$ is its depth. This model has
-the further advantage that binning is not required because the energy is
-discrete.
-
-\jpcom{Describe Lennard-Jones clusters we test SAD ability to handle complex
-energy landscapes. We choose 31 and 38 due to complex funnels In the particular
-case of n=38 there exist two funnels at low energies representing octahedral (Oh)
-and icosahedral (ico) energy minima [5] leading to a pronounced size-specific
-behavior at low temperatures, which is especially difficult to simulate. The
-first numerical success}
-
 In this work, we compare four flat histogram methods.  We outline the
@@
 following methods are discussed and simulated for the square-well
-fluid: Wang-Landau (WL), $1/t$-Wang-Landau ($1/t$-WL), Stochastic
-Approximation Monte Carlo (SAMC), and SAD.
-\jpcom{In addition, we test the ability of SAD to accurately and efficiently
-samples low-temperature solid-solid transitions for Lennard-Jones clusters.}
+fluid and the 31-atom Lennard-Jones cluster: Wang-Landau (WL),
+$1/t$-Wang-Landau ($1/t$-WL), Stochastic Approximation Monte Carlo
+(SAMC), and SAD.
 
@@
 
+\subsection{Square-well fluid}
+\davidsays{Edit this paragraph: introduce novelty of square well fluid
+in particular, and importance.}
+As our first test case, we consider the square-well fluid i.e. a
+system of particles whose interactions are governed by a square-well
+potential~\cite{singh2003surface, barker2004perturbationSW}.  The
+square-well potential is an ideal test-bed as it is a simple model for
+a liquid, which includes both attractive and repulsive
+interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The
+potential $U(\textbf{r})$ for such a system is given by
+\begin{equation}
+ U(\textbf{r})=\begin{cases} \infty &
+ \lvert\textbf{r}\rvert< \sigma\\-\epsilon &
+ \sigma<\lvert\textbf{r}\rvert<\lambda\sigma\\0 &
+ \lvert\textbf{r}\rvert > \lambda\sigma\end{cases}
+\end{equation}
+where $\sigma$ is the hard-sphere diameter of the particle, $\lambda$ is the
+reduced range of the potential well, and $\epsilon$ is its depth. This model has
+the further advantage that binning is not required because the energy is
+discrete.
+
 \begin{figure*}
@@
 
-\subsection{A periodic system with 50 atoms}
-
+\subsubsection{50 square-well atoms}
+\davidsays{check this periodic}
 For this simulation, we chose a minimum reduced
@@
 
-\subsection{A periodic system with 256 atoms}
+\subsubsection{256 square-well atoms}
+\davidsays{check this periodic}
 Next we will introduce a considerably larger simulation containing 256
@@
 \begin{figure}
+  \includegraphics[width=\columnwidth]{figs/lj-cv}
+  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
+    as calculated with each method.}
+  \label{fig:lj-cv}
+\end{figure}
+
+\subsection{31-atom Lennard-Jones cluster}
+
+In order to explore both the bin-size behavior and low-temperature
+effectiveness of our algorithm, we chose to study the 31-atom cluster
+(LJ31).
+Clusters of Lennard-Jones atoms have been frequently used for testing
+Monte Carlo algorithms~\davidsays{add a few citations}.  The LJ31
+cluster is commonly used for testing convergence of algorithms
+because it features a two-funnel energy landscape with a significant
+barrier between the two low-energy states~\davidsays{add citations}.
+In particular, the LJ31 cluster features a solid-solid phase
+transition at a temperature of around $T=0.03\epsilon/k_B$ that is
+challenging to converge~\davidsays{cite}.  Poulain \emph{et al.}
+tested a variety of variants of WL on LJ31 and found
+\davidsays{summarize}~\cite{poulain2006performances}.
+
+\jpcom{This \P\ should be commented once summarized.
+Poulain \emph{et al.} found very poor convergence for a variety of
+modified WL methods on the LJ31
+cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
+energy range or bin size that was used in their simulation, so their
+results cannot be reproduced.  They do mention that they use 1000
+energy bins, which if they used an energy range of
+$[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
+E\approx 0.13\epsilon$, which we find to be too large to accurately
+predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
+only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}
+
+Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster as
+a function of temperature.  There are two peaks, a large peak at 0.2?
+corresponding to melting temperature, and a small peak around 0.03 for
+the solid-solid transition from a mackay to antimackay
+wahatever~\davidsays{cite}.  We keep our atoms in a spherical box with
+radius 2.5$\sigma$, as others do~\davidsays{cite}.  We find that using
+a larger box (with radius $5\sigma$) has a large effect on the melting
+peak, and a smaller but still significant effect on the solid-solid
+transition.  Since our focus is on the convergence rate, we focused on
+the smaller box size.  We further restricted our simulations to
+negative energies, as the microcanonical energy at infinite
+temperature is rather high for such a small box.
+
+\begin{figure}
   \includegraphics[width=\columnwidth]{figs/lj-cv-error}
@@
 
-\begin{figure}
-  \includegraphics[width=\columnwidth]{figs/lj-cv}
-  \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
-    as calculated with each method.}
-  \label{fig:lj-cv}
-\end{figure}
-
-\subsection{Lennard-Jones clusters with 31 atoms}
-
-Clusters of Lennard-Jones atoms have been frequently used for testing
-Monte Carlo algorithms~\davidsays{add a few citations}.  The 31-atom
-cluster is popular because it features a two-funnel energy landscape
-with a significant barrier between the two low-energy
-states~\davidsays{add citations}.
-
-Poulain \emph{et al.} found very poor convergence for a variety of
-modified WL methods on the LJ31
-cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
-energy range or bin size that was used in their simulation, so their
-results cannot be reproduced.  They do mention that they use 1000
-energy bins, which if they used an energy range of
-$[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
-E\approx 0.13\epsilon$, which we find to be too large to accurately
-predict the heat capacity peak at $0.03\epsilon/k_B$.
-
 The heat capacity changes significantly even at these low temperatures
@@
 heat capacity converges just as well for bin widths varying over two
-orders of magnitude.  WL and $1/t$-WL show similar behavior, so this
-is not new, but is encouraging.
+orders of magnitude, although for bins as large as $\Delta
+E=0.1\epsilon$ the error in the heat capacity is significant, as is
+apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
+more slowly for a smaller $\Delta E$, but when summing over energies
+the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
+show similar behavior, so this is not new, but is encouraging.
 
@@
 low-energy cutoff.
+
+SAMC is a major pain.  The $t_0$ value needs to be adjusted when
+$\Delta E$ is changed or when the energy range is changed.  The energy
+range has the same downsides as the WL methods, although it manifests
+differently, with failure to converge as the problem, rather than
+failure for $\gamma_t$ to decrease.
 

### 2019-11-13 papers/histogram/figs/lj-comparison-plot.py

@@ -115,10 +115,10 @@ for i in np.arange(-8, 19, 1.0):
 #         data = np.loadtxt(datadir+fname+'-cv-error.txt')
 #         colors.loglog(data[:,0], data[:,2], method=method)
 
-plt.xlabel(r'$\textrm{Moves}$')
-plt.ylabel(r'$\textrm{Maximum Error in }C_V/k_B$')
+plt.xlabel(r'Moves')
+plt.ylabel(r'Maximum Error in $C_V/k_B$')
 
-plt.xlim(1e7, 4e12)
+plt.xlim(1e7, 2e12)
 plt.ylim(3e-1,1e4)
 
 colors.legend()
@@ -129,27 +129,70 @@ plt.savefig('figs/' + 'lj-cv-error.pdf')
 plt.figure('cv')
 
 data = np.loadtxt(datadir+'bench-cv.txt')
-colors.plot(data[:,0], data[:,1], method='bench')
+
+wideax = plt.gca();
+# inset axes....
+inset_rect = [0.08, 0.43, 0.46, 0.57]
+zoomax = wideax.inset_axes(inset_rect)
+
+zoomax.set_xlim(data[0,0], data[-1,0])
+
+zoom_ymin = 89
+zoom_ymax = 125
+zoomax.set_ylim(zoom_ymin, zoom_ymax)
+colors.plot(data[:,0], data[:,1], method='bench', axes=zoomax)
+data = np.loadtxt(datadir+'wide-cv.txt')
+wideax.set_xlim(0, data[-1,0])
+wide_ymin = 89
+wide_ymax = 190
+wideax.set_ylim(wide_ymin, wide_ymax)
+colors.plot(data[:,0], data[:,1], method='bench', axes=wideax)
 
 for method in files:
         fname = files[method][0.01]
         data = np.loadtxt(datadir+fname+'-cv.txt')
+        widedata = np.loadtxt(datadir+fname+'-wide-cv.txt')
         if len(data) == 0:
                 continue
-        colors.plot(data[:,0], data[:,1], method=method)
+        if 'sad' in method:
+                colors.plot(data[:,0], data[:,1], method=r'SAD $\Delta E=0.01\epsilon$', axes=zoomax)
+                colors.plot(widedata[:,0], widedata[:,1], method=r'SAD $\Delta E=0.01\epsilon$', axes=wideax)
+                fname = files[method][0.001]
+                if os.path.exists(datadir+fname+'-cv.txt'):
[…]

### 2019-11-14 papers/histogram/sad.tex

@@
 \begin{abstract}
-  \jpsays{edit to include LJ clusters}
+  \jpsays{edit to include LJ clusters when paper finished!}
   We present a new Monte Carlo algorithm based on the Stochastic
@@
   \includegraphics[width=\columnwidth]{N50-lndos-comparison}
-  \caption{The entropy of a square well fluid with 50 atoms and filling fraction
+  \caption{The entropy of a square-well fluid with 50 atoms and filling fraction
         0.3 as the green line.  The green hatched area reflects the
@@
   \includegraphics[width=\columnwidth]{gamma-n256}
-  \caption{The update factor $\gamma_t$ versus iteration for $N=256$
-    square-well atoms, for four different methods: WL, $1/t$-WL, SAMC,
+  \caption{The update factor $\gamma_t$ versus iteration for the
+    square-well fluid with 256 atoms, for four different methods: WL, $1/t$-WL, SAMC,
     and SAD.}
@@
 \subsection{Square-well fluid}
-\davidsays{Edit this paragraph: introduce novelty of square well fluid
-in particular, and importance.}
 As our first test case, we consider the square-well fluid i.e. a
@@
 a liquid, which includes both attractive and repulsive
-interactions~\cite{barker1967-SW-perturbation, vega1992phase}.  The
+interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
+\jpcom{In addition to providing an ideal test-bed for new algorithm development,
+to the best of the knowledge of the authors, there are \emph{not} any direct convergence comparison tests in the literature for the variety of flat histogram methods presented here applied to the square-well fluid.}
+The
 potential $U(\textbf{r})$ for such a system is given by
@@
 \subsubsection{50 square-well atoms}
-\davidsays{check this periodic}
-For this simulation, we chose a minimum reduced
+For this 50 atom simulation, we chose a minimum reduced
 temperature of $1/3$, which corresponds to an interesting energy range
@@
 \subsubsection{256 square-well atoms}
-\davidsays{check this periodic}
-Next we will introduce a considerably larger simulation containing 256
+Next we introduce a considerably larger simulation containing 256
 atoms which has a maximum entropy about 1500 greater than its minimum.
@@
 effectiveness of our algorithm, we chose to study the 31-atom cluster
-(LJ31).
+(LJ31) because it is the smallest Lennard-Jones cluster that exhibits broken ergoidcity
+and a solid-solid phase transition at a low temperature~\cite{doye1998thermodynamics, martiniani2014superposition}.
 Clusters of Lennard-Jones atoms have been frequently used for testing
-Monte Carlo algorithms~\davidsays{add a few citations}.  The LJ31
+Monte Carlo algorithms~\cite{wales1997global, neirotti2000phase, frantsuzov2005size, mandelshtam2006multiple}.  The LJ31
 cluster is commonly used for testing convergence of algorithms
 because it features a two-funnel energy landscape with a significant
-barrier between the two low-energy states~\davidsays{add citations}.
+barrier between the two low-energy states~\cite{calvo2000entropic, calvo2000phase, poulain2006performances, martiniani2014superposition}.
 In particular, the LJ31 cluster features a solid-solid phase
-transition at a temperature of around $T=0.03\epsilon/k_B$ that is
-challenging to converge~\davidsays{cite}.  Poulain \emph{et al.}
-tested a variety of variants of WL on LJ31 and found
+transition at a temperature of around $T \approx 0.027\epsilon/k_B$ that is
+challenging to converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
+tested a number of variants of WL on LJ31 and found
 \davidsays{summarize}~\cite{poulain2006performances}.
@@
 
-Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster as
-a function of temperature.  There are two peaks, a large peak at 0.2?
-corresponding to melting temperature, and a small peak around 0.03 for
-the solid-solid transition from a mackay to antimackay
-wahatever~\davidsays{cite}.  We keep our atoms in a spherical box with
-radius 2.5$\sigma$, as others do~\davidsays{cite}.  We find that using
-a larger box (with radius $5\sigma$) has a large effect on the melting
-peak, and a smaller but still significant effect on the solid-solid
-transition.  Since our focus is on the convergence rate, we focused on
-the smaller box size.  We further restricted our simulations to
-negative energies, as the microcanonical energy at infinite
-temperature is rather high for such a small box.
+Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster as a function
+of temperature.  There are two peaks, a large peak around $T \approx
+0.3\epsilon/k_B$? corresponding to melting temperature, and a small peak around
+$T \approx 0.027\epsilon/k_B$ for the solid-solid transition from a
+Mackay-to-anti-Mackay (M $\rightarrow$ aM) transition~\cite{doye1997thermally,
+hendy2001molecular, mandelshtam2006multiple}.  We keep our atoms in a spherical
+box with radius 2.5$\sigma$, as is common in the
+literature~\cite{poulain2006performances, mandelshtam2006multiple}.  We find
+that using a larger box (with radius $5\sigma$) has a large effect on the
+melting peak, and a smaller but still significant effect on the solid-solid
+transition.  Since our focus is on the convergence rate, we focused on the
+smaller box size.  We further restricted our simulations to negative energies,
+as the microcanonical energy at infinite temperature is rather high for such a
+small box.
 
@@
 
+\jpcom{this is a basically a repeat of above...
+
 The heat capacity changes significantly even at these low temperatures
@@
 volume is not the most meaningful quantity to predict for a cluster.
+}
 

### 2019-11-15 papers/histogram/sad.tex
edits on LJ section

@@
   \caption{The heat capacity of the LJ31 cluster after $10^{12}$ moves
-    as calculated with each method.}
+    as calculated with each method.  Each simulation used a bin size
+    of $\Delta E=0.01\epsilon$.  The dotted black line labeled
+    ``converged result'' was a $1/t$-WL simulation with the energy
+    constrained to a range $-133.53\epsilon\le E\le -110\epsilon$,
+    all other simulations had a maximum allowed energy of zero.  The
+    ``converged result'' therefore is only converged at low temperatures.}
   \label{fig:lj-cv}
@@
 In order to explore both the bin-size behavior and low-temperature
-effectiveness of our algorithm, we chose to study the 31-atom cluster
-(LJ31) because it is the smallest Lennard-Jones cluster that exhibits broken ergoidcity
-and a solid-solid phase transition at a low temperature~\cite{doye1998thermodynamics, martiniani2014superposition}.
-Clusters of Lennard-Jones atoms have been frequently used for testing
-Monte Carlo algorithms~\cite{wales1997global, neirotti2000phase, frantsuzov2005size, mandelshtam2006multiple}.  The LJ31
-cluster is commonly used for testing convergence of algorithms
+convergence of our algorithm, we chose to study the 31-atom cluster
+(LJ31), which is the smallest Lennard-Jones cluster that exhibits
+broken ergoidcity and a solid-solid phase transition at a low
+temperature~\cite{doye1998thermodynamics,
+  martiniani2014superposition}.  Clusters of Lennard-Jones atoms have
+been frequently used for testing Monte Carlo
+algorithms~\cite{wales1997global, neirotti2000phase,
+  frantsuzov2005size, mandelshtam2006multiple}.  The LJ31 cluster
+specifically is commonly used for testing convergence of algorithms
 because it features a two-funnel energy landscape with a significant
-barrier between the two low-energy states~\cite{calvo2000entropic, calvo2000phase, poulain2006performances, martiniani2014superposition}.
-In particular, the LJ31 cluster features a solid-solid phase
-transition at a temperature of around $T \approx 0.027\epsilon/k_B$ that is
-challenging to converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
-tested a number of variants of WL on LJ31 and found
-\davidsays{summarize}~\cite{poulain2006performances}.
+barrier between the two low-energy states~\cite{calvo2000entropic,
+  calvo2000phase, poulain2006performances,
+  martiniani2014superposition}.  In particular, the LJ31 cluster
+features a solid-solid phase transition at a temperature of around $T
+\approx 0.027\epsilon/k_B$ that is challenging to
+converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
+tested a number of variants of the WL algorithm on LJ31 and failed to
+converge the spike in heat capacity corresponding to the solid-solid
+phase transition~\cite{poulain2006performances}.
 
-\jpcom{This \P\ should be commented once summarized.
-Poulain \emph{et al.} found very poor convergence for a variety of
-modified WL methods on the LJ31
-cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
-energy range or bin size that was used in their simulation, so their
-results cannot be reproduced.  They do mention that they use 1000
-energy bins, which if they used an energy range of
-$[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
-E\approx 0.13\epsilon$, which we find to be too large to accurately
-predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
-only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}
+%% \jpcom{This \P\ should be commented once summarized.
+%% Poulain \emph{et al.} found very poor convergence for a variety of
+%% modified WL methods on the LJ31
+%% cluster~\cite{poulain2006performances}.  Sadly, they don't mention the
+%% energy range or bin size that was used in their simulation, so their
+%% results cannot be reproduced.  They do mention that they use 1000
+%% energy bins, which if they used an energy range of
+%% $[-133.58\epsilon,0]$ would correspond to a bin size of $\Delta
+%% E\approx 0.13\epsilon$, which we find to be too large to accurately
+%% predict the heat capacity peak at $0.03\epsilon/k_B$.  Looks like they
+%% only ran $\sim 10^{10}$ moves. We can't converge in that few moves either.}
 
-Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster as a function
-of temperature.  There are two peaks, a large peak around $T \approx
-0.3\epsilon/k_B$? corresponding to melting temperature, and a small peak around
-$T \approx 0.027\epsilon/k_B$ for the solid-solid transition from a
-Mackay-to-anti-Mackay (M $\rightarrow$ aM) transition~\cite{doye1997thermally,
-hendy2001molecular, mandelshtam2006multiple}.  We keep our atoms in a spherical
-box with radius 2.5$\sigma$, as is common in the
-literature~\cite{poulain2006performances, mandelshtam2006multiple}.  We find
-that using a larger box (with radius $5\sigma$) has a large effect on the
-melting peak, and a smaller but still significant effect on the solid-solid
-transition.  Since our focus is on the convergence rate, we focused on the
-smaller box size.  We further restricted our simulations to negative energies,
-as the microcanonical energy at infinite temperature is rather high for such a
-small box.
+Figure~\ref{fig:lj-cv} shows the heat capacity of the LJ31 cluster
+over the temperature range from $k_BT=0.01\epsilon$ to
+$k_BT=0.40\epsilon$.  There are two peaks, a large peak around $k_BT
+\approx 0.3\epsilon$? corresponding to melting temperature, and a
+small peak around $T \approx 0.027\epsilon/k_B$ for the solid-solid
+transition from a Mackay-to-anti-Mackay (M $\rightarrow$ aM)
+transition~\cite{doye1997thermally, hendy2001molecular,
+  mandelshtam2006multiple}.  We constrain our atoms within a spherical
+box of radius 2.5$\sigma$, as is common in the
+literature~\cite{poulain2006performances, mandelshtam2006multiple}.
+We note that using a larger box (with radius $5\sigma$) has a large
+effect on the melting peak, and a smaller but still significant effect
+on the solid-solid transition.  Since our focus is on the convergence
+rate, we focused on the smaller box size.  We further restricted all
+our simulations to negative energies.
 
@@
   \caption{The maximum error in the heat capacity of the LJ31 cluster
-    in the temperature range from 0.01~$\epsilon/k_B$ to
-    0.1~$\epsilon/k_B$.  Note that the value of $C_V$ ranges from
-    90$k_B$ to 120$k_B$, so the smallest (maximum) errors seen are
-    around 1\%.  All WL methods shown were run with a histogram bin
-    width of $\Delta E=0.01\epsilon$.  Their convergence at other bin
-    widths was similar to that of SAD. }
+    in the temperature range 0.01~$\epsilon\le k_BT \le 0.05\epsilon$.
+    Note that the value of $C_V$ ranges from 90$k_B$ to 120$k_B$, so
+    the smallest (maximum) errors seen are around 1\%.  All WL methods
+    shown were run with a histogram bin width of $\Delta
+    E=0.01\epsilon$.  Their convergence at other bin widths was
+    similar to that of SAD. }
   \label{fig:lj-cv-error}
@@
 
-\jpcom{this is a basically a repeat of above...
+In order to test the low-temperature convergence, we focused on the
+heat capacity in the temperature range $0.01\epsilon \le k_BT \le
+0.05\epsilon$, which contains the solid-solid phase transition.
+%
+We compute the heat capacity from the entropy using a standard
+canonical analysis, which gives
+\begin{align}
+  C_V &= 31\times\frac32k_B + \frac1{k_BT^2}\left< (E- \langle
+  E\rangle)^2\right>
+  \label{eq:cv}
+\end{align}
+where the first term represents the kinetic energy contribution, and
+the second term is the potential energy contribution, with canonical
+averages using our microcanonical entropy to compute multiplicities.
+%
+For our reference heat capacity, we ran a $1/t$-WL simulation with the
+energy constrained to a range $-133.53\epsilon\le E\le -110\epsilon$.
+We chose this energy range through experimentation, as we found the
+convergence rate to be very sensitive to the minimum energy chosen.
 
-The heat capacity changes significantly even at these low temperatures
-when the size of the cavity is increased.  We use a 2.5 cavity because
-``everyone else'' does~\cite{poulain2006performances}.  For purposes
-of testing the convergence of the algorithm, we don't need to solve
-for the true heat capacity.  Furthermore, the heat capacity at fixed
-volume is not the most meaningful quantity to predict for a cluster.
-}
+%% \jpcom{this is a basically a repeat of above...
+%% The heat capacity changes significantly even at these low temperatures
+%% when the size of the cavity is increased.  We use a 2.5 cavity because
+%% ``everyone else'' does~\cite{poulain2006performances}.  For purposes
+%% of testing the convergence of the algorithm, we don't need to solve
+%% for the true heat capacity.  Furthermore, the heat capacity at fixed
+%% volume is not the most meaningful quantity to predict for a cluster.
+%% }
+We ran simulations with bin widths of $\Delta E=0.1\epsilon$,
+$0.01\epsilon$, and $0.001\epsilon$ using each algorithm we studied.
+When examining the converged results, we saw that each bin size
+measureably different heat capacity at low temperatures. The largest
+bin size of $0.1\epsilon$ introducing a maximum error in heat capacity
+of more than 20$k_B$ while the intermediate bin size of $0.01\epsilon$
+introduced an error of less than $k_B$ when compared with the
+converged result at our smallest bin size.  Therefore we chose to
+compare our simulations primarily at the intermediate bin size.  We
+used a bin size of $0.01\epsilon$ for our reference in all plots in
+this paper (including the reference), except where otherwise
+specified.
 
-Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
-heat capacity converges just as well for bin widths varying over two
-orders of magnitude, although for bins as large as $\Delta
-E=0.1\epsilon$ the error in the heat capacity is significant, as is
-apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
-more slowly for a smaller $\Delta E$, but when summing over energies
-the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
-show similar behavior, so this is not new, but is encouraging.
+Figure~\ref{fig:lj-cv-error} shows the largest error in heat capacity
+at low temperatures as a function of the number of Monte Carlo moves
+performed.
+%
+We explored the convergence behavior of each algorithm as a function
+of bin width, but show only the results for SAD.  Both the WL and
+$1/t$-WL algorithms showed similar behavior to SAD.  In each of these
+algorithms, the rate of convergence of the heat capacity is roughly
+independent of the bin width.  In contrast (but not shown) the entropy
+itself shows significantly larger statistical errors with smaller bin
+sizes.  The canonical averaging process in Eq.~\ref{eq:cv} averages
+out those statistical fluctuations, largely eliminating the penalty of
+using a smaller bin size.
 
-Specifying temperature range rather than energy range is beneficial.
-Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
-chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
-that showed them not converging because it chose too low a minimum
-energy, proving that it is \emph{not} a trivial procedure to determine
-an appropriate energy range.  Specifying the temperature range
-\emph{does} matter! We had to do a whole bunch of $1/t$-WL
-computations in order to determine what would be a sufficient
-low-energy cutoff.
+%% Dependence on bin width.  Figure~\ref{fig:lj-cv-error} shows that the
+%% heat capacity converges just as well for bin widths varying over two
+%% orders of magnitude, although for bins as large as $\Delta
+%% E=0.1\epsilon$ the error in the heat capacity is significant, as is
+%% apparent in Fig~\ref{fig:lj-cv}.  The entropy itself converges much
+%% more slowly for a smaller $\Delta E$, but when summing over energies
+%% the statistical fluctuations in the entropy cancel. WL and $1/t$-WL
+%% show similar behavior, so this is not new, but is encouraging.
 
-SAMC is a major pain.  The $t_0$ value needs to be adjusted when
-$\Delta E$ is changed or when the energy range is changed.  The energy
-range has the same downsides as the WL methods, although it manifests
-differently, with failure to converge as the problem, rather than
-failure for $\gamma_t$ to decrease.
+We found the convergence of the WL and $1/t$-WL to be incredibly
+sensitve to the minimum energy chosen, complicating the process of
+obtaining a workable result.  The known ground state energy of this
+cluster is \davidsays{look up and cite}.  With a minimum energy of
+$-133.58\epsilon$, our simulation did not converge in 10$^{12}$ moves.
+With minimum energies of $-133.54\epsilon$, $-133.53\epsilon$, and
+$-133.52\epsilon$, the simulations do seem to converge, although the
+lowest energy case was significantly slower and never reached as small
+an error.  This result highlights the advantage of being able to
+specify a temperature range rather than an energy range when computing
+properties as a function of temperature.  It is not only more
+convenient, but also far more efficient due the the fact that only a
+single simulation need be performed.
+
+%% Specifying temperature range rather than energy range is beneficial.
+%% Both WL and $1/t$-WL are incredibly sensitive to the minimum energy
+%% chosen, as seen in Fig.~\ref{fig:lj-cv-error}.  Cite previous work
+%% that showed them not converging because it chose too low a minimum
+%% energy, proving that it is \emph{not} a trivial procedure to determine
+%% an appropriate energy range.  Specifying the temperature range
+%% \emph{does} matter! We had to do a whole bunch of $1/t$-WL
+%% computations in order to determine what would be a sufficient
+%% low-energy cutoff.
+
+Finally, we want to mention our results for SAMC, which are not shown
+in our plots in order to reduce clutter.  With the right value for
+$t_0$ and with the right energy range chosen, SAMC converges as
+rapidly as either $1/t$-WL or SAD.  However, with a wrong $t_0$
+parameter, the convergence rate is very poor.  In addition, the
+optimal $t_0$ parameter depends on the bin width as well as the energy
+range chosen, further complicating matters.
+
+%% SAMC is a major pain.  The $t_0$ value needs to be adjusted when
+%% $\Delta E$ is changed or when the energy range is changed.  The energy
+%% range has the same downsides as the WL methods, although it manifests
+%% differently, with failure to converge as the problem, rather than
+%% failure for $\gamma_t$ to decrease.
 

### 2019-11-20 papers/histogram/sad.tex
added in the citations for global minimum.
edit paper to remove smaller bin size

@@
     as calculated with each method.  Each simulation used a bin size
-    of $\Delta E=0.01\epsilon$.  The dotted black line labeled
-    ``converged result'' was a $1/t$-WL simulation with the energy
-    constrained to a range $-133.53\epsilon\le E\le -110\epsilon$,
-    all other simulations had a maximum allowed energy of zero.  The
-    ``converged result'' therefore is only converged at low temperatures.}
+    of $\Delta E=0.01\epsilon$.}
   \label{fig:lj-cv}
@@
     shown were run with a histogram bin width of $\Delta
-    E=0.01\epsilon$.  Their convergence at other bin widths was
-    similar to that of SAD. }
+    E=0.01\epsilon$.  The converged reference heat capacity was
+    computed with a $1/t$-WL simulation with the energy constrained to
+    a range $-133.53\epsilon\le E\le -110\epsilon$.
+  }
   \label{fig:lj-cv-error}
@@
 
-Figure~\ref{fig:lj-cv-error} shows the largest error in heat capacity
-at low temperatures as a function of the number of Monte Carlo moves
-performed.
+Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
+in the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
+function of the number of Monte Carlo moves performed.
 %
-We explored the convergence behavior of each algorithm as a function
-of bin width, but show only the results for SAD.  Both the WL and
-$1/t$-WL algorithms showed similar behavior to SAD.  In each of these
-algorithms, the rate of convergence of the heat capacity is roughly
+We explored the convergence behavior of each algorithm at two energy
+bin widths ($\Delta E=0.01\epsilon$ and $\Delta E=0.1\epsilon$).  We
+were unable to converge the WL or $1/t$-WL algorithms with the larger
+bin size.
+%
+The rate of convergence for the heat capacity seems to be roughly
 independent of the bin width.  In contrast (but not shown) the entropy
-itself shows significantly larger statistical errors with smaller bin
+itself shows much larger statistical fluctuations with smaller bin
 sizes.  The canonical averaging process in Eq.~\ref{eq:cv} averages
-out those statistical fluctuations, largely eliminating the penalty of
-using a smaller bin size.
+out those statistical fluctuations, which reduces the penalty of using
+a smaller bin size.
 
@@
 We found the convergence of the WL and $1/t$-WL to be incredibly
-sensitve to the minimum energy chosen, complicating the process of
+sensitive to the minimum energy chosen, complicating the process of
 obtaining a workable result.  The known ground state energy of this
-cluster is \davidsays{look up and cite}.  With a minimum energy of
-$-133.58\epsilon$, our simulation did not converge in 10$^{12}$ moves.
-With minimum energies of $-133.54\epsilon$, $-133.53\epsilon$, and
+cluster is $-133.586422\epsilon$~\cite{northby1987structure,
+  wales1997global}.  With a minimum energy of $-133.58\epsilon$, none
+of our simulations converged in 10$^{12}$ moves.  With minimum
+energies of $-133.54\epsilon$, $-133.53\epsilon$, and
 $-133.52\epsilon$, the simulations do seem to converge, although the
@@
 
-\jpcom{mention something about LJ clusters and what we find!}
+We found that for examining low-energy structures of a Lennard-Jones
+cluster the WL and $1/t$-WL algorithms---versions of which had
+previously found to be unsuccessful on this problem~\cite{}---have
+convergence properties that are highly dependent on the choice of
+energy range to be examined.  This is particularly problematic when
+examining the low-temperature heat capacity, where it is difficult to
+determine the lowest energy that will have a significant impact at the
+temperatures of interest.
 

### 2019-11-20 papers/histogram/sad-referee-response-2.md
read through printed paper and made necessary edits. now working on response.

Dear editor,

We appreciate the referees' careful reading of our paper, and feel that their
feedback has led to a significant improvement.  We have made some significant
edits to the paper in response to their comments.  These edits were focused on
providing a broader background with regard to previous work, increasing clarity,
and general proofreading.  We believe that we have addressed all of the
referees' concerns with the exception of the unreasonable request by the first
referee that we present results on additional new systems.

...

Response to first referee
# -----------------------------------------------------------------------------#
In the first round of review I raised the general concern that this
work might not present "substantial improvements of existing
computational methods, provided they are applicable to significant
physics problems," as required for papers to be published in Physical
Review. While the authors have addressed all of my concerns, the
question the editor asks is "Do you find the authors' response
persuasive?" I have thought about this question a lot and tend to say
no, the response is not persuasive. The authors mostly reiterate their
(strong) opinion on the strength of their work (which every author
would do, of course), but did not really convince me that I was very
mistaken in my initial review. For example, I was wondering if the
authors could show results for more than a test system to prove
applicability to "significant physics problems" (as required by PRE).
The authors just state this would be "unreasonable", arguing based on
precedence by referring to six PRE papers that only study a single
system. I could certainly find another six papers, where multiple
systems were studied with a "new Monte Carlo algorithm" that even
showed new physics results. And so on. But the question remains: Do
the authors show evidence for applicability to "significant physics
problems"? Even though the method looks sound and promising and is
well described, I tend to say no. Even though the authors assume this
to be the case, and they are probably right, this evidence is not
shown in the paper.

On the other hand, it is true that the revised version of the paper
contains significant improvements, "providing a broader background
with regard to previous work, increasing clarity, and general
proofreading." I agree. The paper is very well written. It is also
true that the authors addressed all of my previous concerns in a
satisfactory manner. Except for my main concern stated above.

The authors show results for a square-well fluid system. I assume it
would be straightforward to modify the code to simulate a
Lennard-Jones fluid. Since SAD estimates the density of states it
should be possible to show specific heat curves, for example. The
authors could, for example, show that SAD correctly (and much faster
and/or reliably?) finds and samples low-temperature solid-solid
transitions in specific Lennard-Jones clusters (for example, LJ135; J.
Chem. Phys. 124, 204511 (2006)). If that or similar data would be in
the paper there would be no doubt whatsoever that SAD is a
"substantial improvement of existing computational methods".

### 2019-11-20 papers/histogram/sad.tex
read through printed paper and made necessary edits. now working on response.

@@
 \begin{abstract}
-  \jpsays{edit to include LJ clusters when paper finished!}
   We present a new Monte Carlo algorithm based on the Stochastic
@@
   $1/t$-WL, which requires that the user identify the interesting range
-  of energies.  Thus SAD is more powerful in the common case in which the range
+  of energies.  Thus, SAD is more powerful in the common case in which the range
   of energies is not known in advance.
@@
 \section{Introduction}
-Over the past several decades a number of flat histogram Monte Carlo simulation
+Over the past several decades, a number of flat histogram Monte Carlo simulation
 algorithms have been developed which calculate the thermodynamic properties of
@@
 to predict properties for nearby temperatures~\cite{ferrenberg1988new}.  For
-large systems this approach is limited to a narrow temperature range because a
+large systems, this approach is limited to a narrow temperature range because a
 single canonical simulation explores only a small range of energies. Berg and
@@
 enable flat histogram sampling which improved the exploration of configuration
-space and allows the system to overcome free energy
+space and allowed the system to overcome free energy
 barriers~\cite{berg1991multicanonical, berg1992multicanonical}. These works led
@@
 wider range of energies. In addition to simulating a range of temperatures, in
-contrast with low-temperature canonical Monte Carlo, is that these approaches
+contrast with low-temperature canonical Monte Carlo, these approaches
 avoid being trapped in a local energy minimum.
@@
 where $t$ corresponds to the number of moves.
-
 This leads to their $1/t$-WL algorithm which ensures that the error continues to
-decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids the
+decrease asymptotically as $1/\sqrt{t}$, which they demonstrated avoids
 error saturation and asymptotically approaches the true density of
-states~\cite{belardinelli2008analysis}. Zhou et. al further confirmed that the
+states~\cite{belardinelli2008analysis}. Zhou \emph{et al.} further confirmed that the
 original WL algorithm never converges exponentially and successfully bounded the
@@
 Liang began to consider Stochastic Approximation as a mathematical
-generalization of the WL approach, whose convergence could be mathematically
+generalization of the WL approach, with convergence that could be mathematically
 proven~\cite{liang2006theory}. In 2007, Liang, Liu, and Carrol \emph{et al.}
@@
 related Statistical Temperature Molecular Dynamics (STMD) which is an
-adaption of the WL method that approximates the entropy (or log density of
+adaption of the WL method that approximates the entropy (or natural logarithm of the density of
 states) as a piecewise linear function, which improves convergence for systems
@@
 the user to define a temperature range of interest ($T_\text{min}$ to
-$T_\infty$).  We call this method SAD (Stochastic Approximation with a Dynamic
+$T_\infty$) which we explain in Section~\ref{sec:sad}.  We call this method SAD (Stochastic Approximation with a Dynamic
 update factor), and will discuss it in detail in the methods section. We compare
@@
 or \emph{multicanonical} methods) is to simulate each energy with
-similar accuracy, so as to accurately determine the density of states
+similar accuracy so as to accurately determine the density of states
 over a broad range of energies---and thus to determine the thermodynamic
@@
 binning for low entropic states in order to ensure the exploration of all
-energies.  The method introduced in this paper is designed to scale
-appropriately as bin size is changed, but we do not test this scaling, as we use
-a system with discrete energy levels.
+energies.  We test the methods on both a discrete and continuum energy system.
 
@@
 which biases the simulation in favor of energies with low weights.  A set of
-weights that are proportional to the density of states of the system $D(E)$ will
+weights that are proportional to the density of states $D(E)$ of the system will
 result in an entirely flat histogram, which means that the weights should
@@
 density of states~\cite{liang2006theory, liang2007stochastic,
-liang2009improving}.  In addition, the energy range need not be known a priori.
+liang2009improving}.  In addition, the energy range need not be known \emph{a priori}.
 The time to converge depends only on the choice of parameter $t_0$.
@@
 A significant advantage of SAD over SAMC---which the $1/t$-WL and WL
-methods share \emph{after} they have found all the energies---is that
+methods share \emph{after} they have discovered all the energies---is that
 the schedule for $\gamma_t$ automatically responds to the choice of bin
-size. Although we have not yet tested this,
-SAD should perform similarly over a wide range of bin sizes
+size. Although we have not yet thoroughly tested this over a wide range of bin sizes,
+SAD should perform similarly over a reasonable range of bin sizes
 because $\gamma_t \propto \nicefrac{N_S}{t}$.  As the number of energy
@@
 interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
-\jpcom{In addition to providing an ideal test-bed for new algorithm development,
-to the best of the knowledge of the authors, there are \emph{not} any direct convergence comparison tests in the literature for the variety of flat histogram methods presented here applied to the square-well fluid.}
+In addition to providing an ideal test-bed for new algorithm development,
+to the best of the knowledge of the authors, there are \emph{not} any direct convergence comparison tests in the literature for the variety of flat histogram methods presented here applied to the square-well fluid.
 The
@@
 $E_{\max}$ being calculated (with the exception of the WL methods
-where both of these parameters are needed a priori).
+where both of these parameters are needed \emph{a priori}).
 
@@
 
-In order to explore both the bin-size behavior and low-temperature
+In order to explore both the \jpsays{remove? bin-size behavior} and low-temperature
 convergence of our algorithm, we chose to study the 31-atom cluster
@@
   frantsuzov2005size, mandelshtam2006multiple}.  The LJ31 cluster
-specifically is commonly used for testing convergence of algorithms
+is commonly used for testing convergence of algorithms
 because it features a two-funnel energy landscape with a significant
@@
 $k_BT=0.40\epsilon$.  There are two peaks, a large peak around $k_BT
-\approx 0.3\epsilon$? corresponding to melting temperature, and a
+\approx 0.3\epsilon$ corresponding to melting temperature, and a
 small peak around $T \approx 0.027\epsilon/k_B$ for the solid-solid
@@
 We ran simulations with bin widths of $\Delta E=0.1\epsilon$,
-$0.01\epsilon$, and $0.001\epsilon$ using each algorithm we studied.
+$0.01\epsilon$, and \jpsays{should we mention if we dont show data? $0.001\epsilon$} using each algorithm we studied.
 When examining the converged results, we saw that each bin size
@@
 
-Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
+Figure~\ref{fig:lj-cv-error} (inset) shows the maximum error in heat capacity
 in the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
@@
 cluster the WL and $1/t$-WL algorithms---versions of which had
-previously found to be unsuccessful on this problem~\cite{}---have
+previously found to be unsuccessful on this problem~\cite{poulain2006performances}---have
 convergence properties that are highly dependent on the choice of
@@
 determine the lowest energy that will have a significant impact at the
-temperatures of interest.
+temperatures of interest. This is an example of how SAD is more powerful when the range
+of energies is not known in advance.
 

### 2019-11-20 papers/histogram/sad-referee-response-2.md
I have written a rough draft of the referee response.

@@
 We appreciate the referees' careful reading of our paper, and feel that their
-feedback has led to a significant improvement.  We have made some significant
-edits to the paper in response to their comments.  These edits were focused on
-providing a broader background with regard to previous work, increasing clarity,
-and general proofreading.  We believe that we have addressed all of the
-referees' concerns with the exception of the unreasonable request by the first
-referee that we present results on additional new systems.
+feedback has led to a significant improvement. We have added new data and a
+section for the 31 atom Lennard-Jones cluster. We show specific heat curves for
+this test system and show the convergence SAD vs the other method quantitatively
+(a quantitative comparison of this kind has not been done in the literature to
+our knowledge). We also added a note to the Square-well fluid section
+highlighting that this system has not been explored using broad histogram
+methods. We believe that we have addressed all of the referees' concerns and per
+the first referee's own words 'If that or similar data would be in the paper
+there would be no doubt whatsoever that SAD is a "substantial improvement of
+existing computational methods".'
 
@@
 # -----------------------------------------------------------------------------#
-In the first round of review I raised the general concern that this
-work might not present "substantial improvements of existing
-computational methods, provided they are applicable to significant
-physics problems," as required for papers to be published in Physical
-Review. While the authors have addressed all of my concerns, the
-question the editor asks is "Do you find the authors' response
-persuasive?" I have thought about this question a lot and tend to say
-no, the response is not persuasive. The authors mostly reiterate their
-(strong) opinion on the strength of their work (which every author
-would do, of course), but did not really convince me that I was very
-mistaken in my initial review. For example, I was wondering if the
-authors could show results for more than a test system to prove
-applicability to "significant physics problems" (as required by PRE).
-The authors just state this would be "unreasonable", arguing based on
-precedence by referring to six PRE papers that only study a single
-system. I could certainly find another six papers, where multiple
-systems were studied with a "new Monte Carlo algorithm" that even
-showed new physics results. And so on. But the question remains: Do
-the authors show evidence for applicability to "significant physics
-problems"? Even though the method looks sound and promising and is
-well described, I tend to say no. Even though the authors assume this
-to be the case, and they are probably right, this evidence is not
-shown in the paper.
 
-On the other hand, it is true that the revised version of the paper
-contains significant improvements, "providing a broader background
-with regard to previous work, increasing clarity, and general
-proofreading." I agree. The paper is very well written. It is also
-true that the authors addressed all of my previous concerns in a
-satisfactory manner. Except for my main concern stated above.
+In the previous review, the first referee acknowledged that  'The paper is very
+well written. It is also true that the authors addressed all of my previous
+concerns in a satisfactory manner.'
 
-The authors show results for a square-well fluid system. I assume it
-would be straightforward to modify the code to simulate a
-Lennard-Jones fluid. Since SAD estimates the density of states it
-should be possible to show specific heat curves, for example. The
-authors could, for example, show that SAD correctly (and much faster
-and/or reliably?) finds and samples low-temperature solid-solid
-transitions in specific Lennard-Jones clusters (for example, LJ135; J.
-Chem. Phys. 124, 204511 (2006)). If that or similar data would be in
-the paper there would be no doubt whatsoever that SAD is a
-"substantial improvement of existing computational methods".
+The first referee asked that we simulate (either 'faster and/or reliably')
+low-temperature solid-solid transitions in specific Lennard-Jones clusters. The
+referee gave LJ135 as an example, but stated that if specific heat capacity
+curves were in the paper for LJ135 or 'similar' (we did LJ31 for reasons we
+state in the paper) then 'there would be no doubt whatsoever that SAD is a
+"substantial improvement of existing computational methods".'
+We believe that we have completely satisfied the first referee and thank the referee
+for pointing our attention toward simulating Lennard-Jones clusters.
+
+Response to second referee
+# -----------------------------------------------------------------------------#
+All of the second referee's comments were addressed previously.

### 2019-11-22 papers/histogram/sad.tex
edits in our meeting

@@
   the 31-atom Lennard-Jones cluster and
-  compare the convergence time and average entropy error for several
+  compare the convergence behavior of several
   related
@@
   $1/t$-WL, which requires that the user identify the interesting range
-  of energies.  Thus, SAD is more powerful in the common case in which the range
+  of energies.
+  %
+  The convergence of the $1/t$-WL method is very sensitive to the energy
+  range chosen for the low-temperature heat capacity of the
+  Lennard-Jones cluster.
+  %
+  Thus, SAD is more powerful in the common case in which the range
   of energies is not known in advance.
@@
 binning for low entropic states in order to ensure the exploration of all
-energies.  We test the methods on both a discrete and continuum energy system.
+energies.  The method introduced in this paper is designed to scale appropriately
+as bin size is changed.
 
@@
 the schedule for $\gamma_t$ automatically responds to the choice of bin
-size. Although we have not yet thoroughly tested this over a wide range of bin sizes,
+size.
 SAD should perform similarly over a reasonable range of bin sizes
@@
 interactions~\cite{barker1967-SW-perturbation, vega1992phase}.
-In addition to providing an ideal test-bed for new algorithm development,
-to the best of the knowledge of the authors, there are \emph{not} any direct convergence comparison tests in the literature for the variety of flat histogram methods presented here applied to the square-well fluid.
-The
-potential $U(\textbf{r})$ for such a system is given by
+To date, there have not been any published direct convergence comparison tests
+for flat histogram methods applied to the square-well fluid.
+The potential $U(\textbf{r})$ for such a system is given by
 \begin{equation}
@@
 
-In order to explore both the \jpsays{remove? bin-size behavior} and low-temperature
+In order to further explore both the
 convergence of our algorithm, we chose to study the 31-atom cluster
@@
   martiniani2014superposition}.  In particular, the LJ31 cluster
-features a solid-solid phase transition at a temperature of around $T
-\approx 0.027\epsilon/k_B$ that is challenging to
+features a low-temperature solid-solid phase transition that is challenging to
 converge~\cite{mandelshtam2006multiple}.  Poulain \emph{et al.}
-tested a number of variants of the WL algorithm on LJ31 and failed to
+tested a number of variants on the WL algorithm on LJ31 and failed to
 converge the spike in heat capacity corresponding to the solid-solid
@@
 \approx 0.3\epsilon$ corresponding to melting temperature, and a
-small peak around $T \approx 0.027\epsilon/k_B$ for the solid-solid
+small peak around $k_BT \approx 0.027\epsilon$ for the solid-solid
 transition from a Mackay-to-anti-Mackay (M $\rightarrow$ aM)
@@
 on the solid-solid transition.  Since our focus is on the convergence
-rate, we focused on the smaller box size.  We further restricted all
-our simulations to negative energies.
+rate, we will restrict ourselves to the smaller box size.  In addition, we
+further restricted all our simulations to negative total potential energies.
 
@@
 heat capacity in the temperature range $0.01\epsilon \le k_BT \le
-0.05\epsilon$, which contains the solid-solid phase transition.
+0.05\epsilon$ (see inset in Fig.~\ref{fig:lj-cv}), which contains
+the solid-solid phase transition.
 %
@@
 the second term is the potential energy contribution, with canonical
-averages using our microcanonical entropy to compute multiplicities.
+averages using the microcanonical entropy to compute multiplicities.
 %
@@
 %% }
-We ran simulations with bin widths of $\Delta E=0.1\epsilon$,
-$0.01\epsilon$, and \jpsays{should we mention if we dont show data? $0.001\epsilon$} using each algorithm we studied.
-When examining the converged results, we saw that each bin size
-measureably different heat capacity at low temperatures. The largest
-bin size of $0.1\epsilon$ introducing a maximum error in heat capacity
-of more than 20$k_B$ while the intermediate bin size of $0.01\epsilon$
-introduced an error of less than $k_B$ when compared with the
-converged result at our smallest bin size.  Therefore we chose to
-compare our simulations primarily at the intermediate bin size.  We
-used a bin size of $0.01\epsilon$ for our reference in all plots in
-this paper (including the reference), except where otherwise
-specified.
+We ran simulations with several bin widths using each algorithm we studied
+and show results for $\Delta E=0.1\epsilon$ and
+$0.01\epsilon$.
+When examining the converged results, we saw that each bin size yielded a
+measureably different low temperature heat capacity. The largest
+bin size of $0.1\epsilon$ introduced a maximum error in heat capacity
+of more than $20k_B$ compared with the smaller bin size.  We will focus on
+the smaller bin size, and
+used a bin size of $0.01\epsilon$ for our converged reference heat capacity.
 
-Figure~\ref{fig:lj-cv-error} (inset) shows the maximum error in heat capacity
-in the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
+Figure~\ref{fig:lj-cv-error} shows the maximum error in heat capacity
+over the temperature range $0.01\epsilon\le k_BT\le 0.05\epsilon$ as a
 function of the number of Monte Carlo moves performed.
@@
 %
-The rate of convergence for the heat capacity seems to be roughly
+The rate of convergence of SAD for the heat capacity seems to be roughly
 independent of the bin width.  In contrast (but not shown) the entropy
@@
 
-We found the convergence of the WL and $1/t$-WL to be incredibly
-sensitive to the minimum energy chosen, complicating the process of
+The convergence behavior of the WL and $1/t$-WL is incredibly
+sensitive to the minimum energy chosen, which complicates the process of
 obtaining a workable result.  The known ground state energy of this
@@
   wales1997global}.  With a minimum energy of $-133.58\epsilon$, none
-of our simulations converged in 10$^{12}$ moves.  With minimum
+of our WL or $1/t$-WL simulations converged in 10$^{12}$ moves.  With minimum
 energies of $-133.54\epsilon$, $-133.53\epsilon$, and
 $-133.52\epsilon$, the simulations do seem to converge, although the
-lowest energy case was significantly slower and never reached as small
-an error.  This result highlights the advantage of being able to
+lowest energy case was significantly slower.
+This result highlights the advantage of being able to
 specify a temperature range rather than an energy range when computing
@@
 Finally, we want to mention our results for SAMC, which are not shown
-in our plots in order to reduce clutter.  With the right value for
-$t_0$ and with the right energy range chosen, SAMC converges as
-rapidly as either $1/t$-WL or SAD.  However, with a wrong $t_0$
-parameter, the convergence rate is very poor.  In addition, the
+in our plots in order to reduce clutter.  With a well-chosen power of ten value for
+$t_0$, SAMC can converge to the correct result, but less rapidly than
+either $1/t$-WL or SAD.  However, with a poor $t_0$
+parameter, the method cannot converge in practice.  In addition, the
 optimal $t_0$ parameter depends on the bin width as well as the energy
@@
 energy space corresponding to a desired range of temperatures for a
-variety of system sizes.
+few systems.
 %
-We find that both SAD and $1/t$-WL demonstrate excellent and robust
-convergence.
-They both converge more rapidly than SAMC, and unlike WL consistently
-converge to the correct density of states. SAD requires
+We find that both SAD and $1/t$-WL converge more rapidly than SAMC, and
+unlike WL consistently converge to the correct density of states. SAD requires
 the user to specify a temperature range of interest rather than an
@@
 
-We found that for examining low-energy structures of a Lennard-Jones
+Finally, we find that for examining a Lennard-Jones
 cluster the WL and $1/t$-WL algorithms---versions of which had

### 2019-11-22 papers/histogram/sad.tex
edits from my dad

@@
 generalization of the WL approach, with convergence that could be mathematically
-proven~\cite{liang2006theory}. In 2007, Liang, Liu, and Carrol \emph{et al.}
+proven~\cite{liang2006theory}. In 2007, Liang \emph{et al.}
 developed Stochastic Approximation Monte Carlo (SAMC)~\cite{liang2007stochastic,
@@
 probability of accepting a move and in what additional statistics must
-be collected in order decide on that probability.
+be collected in order to decide on that probability.
 
@@
 of energies~\cite{wang2001determining, dayal2004performance, troyer2003flat,
-trebst2004optimizing}. For this reason, energy binning becomes an important
+trebst2004optimizing}. Therefore, energy binning becomes an important
 consideration for systems with a continuum of possible energies.  Energy bins
@@
 converge to being proportional to the density of states.  The natural logarithm
-of the weights are typically stored, since the density of states will often vary
-over a few hundred orders of magnitude. In the microcanonical ensemble the
+of the weight is typically stored, since the density of states will often vary
+over a few hundred orders of magnitude. In the microcanonical ensemble, the
 entropy is defined as $S(E) \equiv \ln(D(E))$ (where $k_B \equiv 1$), the
-logarithm of the weights is an approximation of the entropy.
+logarithm of the weight is an approximation of the entropy.
 
@@
 \end{equation}
-where $t$ is number of the current move, $\gamma_t$ is an update factor
+where $t$ is the number of the current move, $\gamma_t$ is an update factor
 that varies over the simulation, and $E$ is the current energy.  This update
@@
 
-We set the maximum temperature of interest as $T=\infty$ for several reasons.
-Firstly, when this temperature is reached, the system is maximally randomized.
-Allowing infinite temperature further ensures that the system can surmount an
-energy barrier of any size.  Finally, the high temperatures are ``easy'' to
-simulate, and tend to converge very quickly.  We note, however, that this is
-under the assumption that infinite temperature corresponds to a finite energy,
-as is the case for configurational sampling.  Were we to include kinetic energy
-in a simulation (as in molecular dynamics), or to examine a system such as a
-high density Lennard-Jones fluid with an entropy maximum at an energy that is
-far from the energy range of physical interest, then we would need to apply a
-maximum temperature, which could slow down the rate at which state space is
-explored.
+We set the maximum temperature of interest as $T=\infty$ for a couple
+of reasons.  First, when this temperature is reached, the system is
+maximally randomized, so allowing infinite temperature further ensures
+that the system can surmount an energy barrier of any size.  Second,
+the high temperatures tend to be ``easy'' to simulate, and tend to
+converge very quickly.  We note, however, that this is under the
+assumption that infinite temperature corresponds to a finite energy,
+as is the case for configurational sampling.  Were we to include
+kinetic energy in a simulation (as in molecular dynamics), or to
+examine a system such as a high density Lennard-Jones fluid with an
+entropy maximum at an energy that is far from the energy range of
+physical interest, then we would need to apply either a maximum
+temperature or a maximum energy, either of which could slow down the
+rate at which state space is explored.
 
@@
 
-During the simulation when considering a move inside of the energy
+During the simulation when considering a move inside the energy
 range of interest $E_L \le E \le E_H$, the weights are used as in the
@@
 
-In order to further explore both the
-convergence of our algorithm, we chose to study the 31-atom cluster
+We also examined convergence of each algorithm on the 31-atom cluster
 (LJ31), which is the smallest Lennard-Jones cluster that exhibits
-broken ergoidcity and a solid-solid phase transition at a low
+broken ergodicity and a solid-solid phase transition at a low
 temperature~\cite{doye1998thermodynamics,
@@
 
-Finally, we find that for examining a Lennard-Jones
-cluster the WL and $1/t$-WL algorithms---versions of which had
-previously found to be unsuccessful on this problem~\cite{poulain2006performances}---have
+Finally, we find that for examining a Lennard-Jones cluster the WL and
+$1/t$-WL algorithms---versions of which had previously been found to
+be unsuccessful on this problem~\cite{poulain2006performances}---have
 convergence properties that are highly dependent on the choice of
@@
 determine the lowest energy that will have a significant impact at the
-temperatures of interest. This is an example of how SAD is more powerful when the range
-of energies is not known in advance.
+temperatures of interest. This is an example of how SAD is more
+powerful when the range of energies is not known in advance.
 

### 2019-11-23 papers/histogram/sad-referee-response-2.md

@@
 this test system and show the convergence SAD vs the other method quantitatively
-(a quantitative comparison of this kind has not been done in the literature to
+(a quantitative comparison of convergence on the LJ31 system has not been done in the literature to
 our knowledge).  We also added a note to the Square-well fluid section
@@
 
+We were surprised by how poorly the 1/t-WL algorithm performed on the LJ31 system,
+in particular the convergence was very strongly affected by the lower limit of
+the energy range chosen.  This example thus strongly highlights the benefits
+SAD provides when examining the heat capacity over a chosen temperature range.
+
 Response to second referee

### 2020-01-24 papers/histogram/sad-referee-response-2.md
fixed a spelling error in paper.

--- 2019-11-20 papers/histogram/sad-referee-response-2.md
@@
 
-We appreciate the referees' careful reading of our paper, and feel that their
-feedback has led to a significant improvement. We have added new data and a
-section for the 31 atom Lennard-Jones cluster. We show specific heat curves for
-this test system and show the convergence SAD vs the other method quantitatively
-(a quantitative comparison of this kind has not been done in the literature to
-our knowledge). We also added a note to the Square-well fluid section
-highlighting that this system has not been explored using broad histogram
-methods. We believe that we have addressed all of the referees' concerns and per
-the first referee's own words 'If that or similar data would be in the paper
-there would be no doubt whatsoever that SAD is a "substantial improvement of
-existing computational methods".'
+We appreciate the referees' careful reading of our paper, and feel
+that their feedback has led to a significant improvement. We have
+added new data and a section for the 31 atom Lennard-Jones cluster. We
+compute the low-temperature specific heat for this test system and
+show the convergence SAD vs the other methods quantitatively. A
+quantitative comparison of convergence on the LJ31 system has not been
+done in the literature to our knowledge.  We also added a note to the
+Square-well fluid section highlighting that this system has not
+previously been explored using broad histogram methods. We believe
+that we have addressed all of the referees' concerns and per the first
+referee's own words 'If that or similar data would be in the paper
+there would be no doubt whatsoever that SAD is a "substantial
+improvement of existing computational methods".'
 
@@
 
-The first referee asked that we simulate (either 'faster and/or reliably')
-low-temperature solid-solid transitions in specific Lennard-Jones clusters. The
-referee gave LJ135 as an example, but stated that if specific heat capacity
-curves were in the paper for LJ135 or 'similar' (we did LJ31 for reasons we
-state in the paper) then 'there would be no doubt whatsoever that SAD is a
-"substantial improvement of existing computational methods".'
-We believe that we have completely satisfied the first referee and thank the referee
-for pointing our attention toward simulating Lennard-Jones clusters.
+The first referee asked that we simulate (either 'faster and/or
+reliably') low-temperature solid-solid transitions in specific
+Lennard-Jones clusters. The referee gave LJ135 as an example, but
+stated that if specific heat capacity curves were in the paper for
+LJ135 or 'similar' then 'there would be no doubt whatsoever that SAD
+is a "substantial improvement of existing computational methods".' We
+chose to examine the LJ31 cluster for reasons we state in the paper,
+specifically that this cluster has a clear solid-solid phase
+transition at very low temperature that has been shown in the past to
+be challenging to simulate using flat histogram methods.  We believe
+that we have completely addressed the request by the first referee and
+thank the referee for pointing our attention toward simulating
+Lennard-Jones clusters.
+
+We were surprised by how poorly the 1/t-WL algorithm performed on the
+LJ31 system.  In particular, the convergence was very strongly
+affected by the lower limit of the energy range chosen.  This example
+thus strongly highlights the benefits SAD provides when examining the
+low-temperature heat capacity, where the temperature of interest is
+known, but the range of energies that affect that temperature are hard
+to determine.
 
@@
 # -----------------------------------------------------------------------------#
-All of the second referee's comments were addressed previously.
+All of the second referee's comments have been addressed previously.

### 2020-01-24 papers/histogram/ising.tex

@@
 replacing un-physical user-defined parameters with the algorithms dynamic
-choice.  In this work, we apply the family of weight-based flat histogram Monte
+choice.
+
+In this work, we apply the family of weight-based flat histogram Monte
 Carlo methods (WL, 1/t-WL, SAMC, SAD) to the 2D Ising model.
@@
 system sufficiently interesting for such theoretical comparisons. It is
-also of important note that direct comparison of the other methods can
+also important to note that direct comparison of the other methods can
 be made with WL as its original implementation was done on this
@@
 The $N\times N$ spin system can take on values of $\sigma_i = \pm 1$
-for up or down spins respectively. In the infinite volume limit ($h =
+for up or down spins respectively. In the absence of a magnetic field ($h =
 0$), We can write the Hamiltonian as follows~\cite{onsager1944crystal,
@@
 \end{align}
-where the sum is over all enumerated spin sites. Beale showed that for finite
+where the sum is over nearest neighbor spin sites. Beale showed that for finite
 lattices the Density of States could directly be calculated from the partition
@@
 \end{align}
-where $g(E)$ is the multiplicity of the system which is identical to
+where $g(E)$ is the multiplicity of the system which is proportional to
 $D(E)$. We can estimate the average deviation from the exact solution
@@
 
+\davidsays{The results feel a bit slim.  I wonder whether we can look
+  at the heat capacity also? Or perhaps there is a way that we can see
+  something ``interesting'' arise out of the computation? I also
+  wonder if we can run longer and get a bit more accuracy.}
+
 \begin{figure}

### 2020-01-24 papers/histogram/sad2d.tex

\documentclass[letterpaper,twocolumn,amsmath,amssymb,pre,aps,10pt]{revtex4-1}
\usepackage{graphicx} % Include figure files
\usepackage{color}
\usepackage{nicefrac} % Include for inline fractions

\usepackage{xargs}                      % Use more than one optional parameter in a new commands
\usepackage[pdftex,dvipsnames]{xcolor}  % Coloured text etc.
\usepackage[colorinlistoftodos,prependcaption,textsize=normalsize]{todonotes}
\usepackage{mdframed}

% define colors for comments
\definecolor{dark-gray}{gray}{0.10}
\definecolor{light-gray}{gray}{0.70}

\newcommand{\red}[1]{{\bf \color{red} #1}}
\newcommand{\green}[1]{{\bf \color{green} #1}}
\newcommand{\blue}[1]{{\bf \color{blue} #1}}
\newcommand{\cyan}[1]{{\bf \color{cyan} #1}}

\newcommand{\davidsays}[1]{{\color{red} [\green{David:} \emph{#1}]}}
\newcommand{\jpsays}[1]{{\color{red} [\blue{Jordan:} \emph{#1}]}}
\newcommandx{\jpcom}[2][1=inline]{\todo[linecolor=gray,backgroundcolor=light-gray,bordercolor=dark-gray,#1]{\textbf{Jordan says:} #2} }
\begin{document}

\title{Notes on 2D sad
}

\author{Jordan K. Pommerenck} \author{David Roundy}
\affiliation{Department of Physics, Oregon State University,
  Corvallis, OR 97331}

\begin{abstract}
  Notes for a possible 2D sad algorithm
\end{abstract}

\maketitle

\section{The basics}

\begin{align}
  dU &= TdS - pdV + \mu dN \\
  dE_{exc} &= TdS_{exc} - p_{exc}dV + \mu_{exc} dN
\end{align}
This tells us that
\begin{align}
  \frac{1}{T} &= \frac{S_{exc}(N,E+\Delta E) - S_{exc}(N,E)}{\Delta E}
\end{align}
and
\begin{align}
  \mu_{exc} &= \left(\frac{\partial E_{exc}}{\partial N}\right)_{S_{exc}}
  \\
  &= -\frac{
    \left(\frac{\partial E_{exc}}{\partial S_{exc}}\right)_{N}
  }{
    \left(\frac{\partial N}{\partial S_{exc}}\right)_{E_{exc}}
  }
  \\
  &= -\frac{
    \left(\frac{\partial S_{exc}}{\partial N}\right)_{E_{exc}}
  }{
    \left(\frac{\partial S_{exc}}{\partial E_{exc}}\right)_{N}
  }
  \\
  &= -T\left(\frac{\partial S_{exc}}{\partial N}\right)_{E_{exc}}
\end{align}
where I used the cyclic chain rule to make the derivatives feasible.

\section{Moving forward}

\begin{align}
  U &= ST - pV + \mu N \\
  E_{exc} &= S_{exc}T - p_{exc}V + \mu_{exc}N
\end{align}
We want a maximum total pressure $p_{\max}$, which makes for a maximum excess pressure of
$p_{\max} - p_{id} = p_{\max} - NkT/V$.  Thus at higher $N$ values, we have
\begin{align}
  E_{exc} &= S_{exc}T - \left(p_{\max}-\frac{NkT}{V}\right)V - NT\left(\frac{\partial S_{exc}}{\partial N}\right)_{E}
  \\
  &= S_{exc}T - p_{\max}V + NkT - NT\left(\frac{\partial S_{exc}}{\partial N}\right)_{E}
\end{align}
solving for the rate of change, we see
\begin{align}
  \left(\frac{\partial S_{exc}}{\partial N}\right)_{E} &= \frac{S_{exc}T - E_{exc} - p_{\max}V + NkT}{NT}
  \\
  &= \frac{S_{exc}}{N} - \frac{E_{exc}+p_{\max}V}{NT} + k
\end{align}
So at a finite value of $N>N_{hi}$ let's start with an ansatz:
\begin{align}
  S_{exc}(N,E) &= A e^{BN} + C \\
  \left(\frac{\partial S_{exc}}{\partial N}\right)_{E} &= AB e^{BN} \\
  &= BS_{exc}(N,E) - BC
\end{align}
thus we can see that our ansatz was nonsense.

\end{document}

### 2020-09-01 papers/thesis-kirstie/MyThesisinLaTeX.tex
thesis

\documentclass[12pt]{article}
\usepackage{setspace}
\doublespacing
%\usepackage[top=1.2in, bottom=1.2in]{geometry}
%\usepackage[top=1.5in]{geometry}
\usepackage[bottom=1.2in]{geometry}
\usepackage{fancyhdr}
\usepackage{yhmath}
\pagestyle{fancy}
\fancyhf{}
\renewcommand{\headrulewidth}{0pt}
\fancyhead[R]{\thepage}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{color}
\usepackage{enumitem}
\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
\author{by Kirstie Finster}
\date{August 2020}
\begin{document}
  \maketitle
  \thispagestyle{fancy}

\noindent IA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)

Real fluids and their behavior are very complex from a physicist's point of view. Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules, and their interaction with charges from other molecules make it difficult to study a fluid composed of water molecules directly. The task becomes even more daunting considering the complicated molecular motion of each water molecule, and the sheer number of molecules that must be taken into account. 
%Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules and their interaction with charges from other molecules, the complicated molecular motion of a collection of water molecules, and the sheer number of molecules that must be taken into account make it difficult to study a fluid composed of water molecules directly.
The problem of understanding the behavoir of real liquids is better addressed by using models which reduce a complex picture to a simple one. Modifications, or perturbations, can then be incorporated into the model making it progressively more complex until it mimics the behavior of a real liquid as closely as possible.

In developing a model for a fluid, simplifications can be made regarding the composition, type, shape, and interatomic interaction of the molecules. Rather than dealing with any particular molecule or atom, the model only deals with generic "particles", which may also be referred to as "atoms". The shape of the particles, or atoms will be the simplest shape possible - a sphere. The spheres in the simplest models do not exhibit quantum interference effects as the mass of the spheres, and the temperature are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. The fluid can then be treated as a classical fluid where the spheres interact with each other according to their electrostatic interatomic potential.

There are many interatomic potentials which can be used to model a fluid each of which describes the interaction between two real atoms to some degree of accuracy. The simplest, and oldest, model used to describe a fluid is the hard-sphere model. The hard-sphere model envisions a fluid to be composed of hard spheres like, for example, a collection of marbles. This model came about by considering the near incompressibilty of liquids indicating that the particles making up the liquid strongly resist being pushed closer together. This behavior can be described by an interatomic potential energy that is zero when two spheres are not in contact, but goes instantly to infinity when contact is made. The inablilty to push real atoms closer together is due to the repulsion of the electrons in each atom, and is the dominating interparticle electrostatic interaction. Thus, the hard-sphere model - which is the simplest, but still effective representation of this replusive behavior that dominates the interaction between real atoms, is currently the most widely used model for fluids.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
    %\includegraphics[height=3.5cm]{plot_HardSphere_Potential.png}
    %\caption{Hard Sphere Potential}
    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
    \label{fig:HardSphere_potential}
  \end{figure}
\[{}\]
\[{}\]
\[{}\]
\[{}\]

Another fairly simple, but relatively accurate interatomic potential widely used to represent generic, neutral atoms is the Leonard-Jones potential. The Leonard-Jones potential describes both the long-range attraction, and the short-range replusion that occurs between two nuetral atoms. The attraction between two nuetral atoms is due to fleeting dipoles that form spontaneously when electrons are redistributed as atoms come into close proximty. These are often characterized as Van der Walls, or London dispersion forces. The attraction grows stronger as two atoms move closer together until their atomic orbitals begin to overlap. %At this point the electrons in each atom begin to repel the electrons in the other atom and oppose the overlap of the two atoms.
 At this point the electrons in each atom begin to oppose the overlap of the two atoms through their replusion. The repulsion increases rapidly and approaches infinity as the distance between the atoms conintues to decrease.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential.png}
    \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential}
  \end{figure}

The combined attractive and repulsive effects produce the overall shape of the Leonard-Jones interatomic potential shown in Figure \ref{fig:LJ_potential}. The potential is zero when the distance r between the centers of two atoms is equal to $\sigma$ (shown in the figure at r/$\sigma$=1). At this point the attraction bewteen the atoms just cancels their repulsion. As the distance between the centers of the two atoms decreases below $\sigma$, the potential rapidly approaches positive infinity corresponding to infinite repulsion. Above $\sigma$ the potential becomes negative indicating a net attraction between the atoms. The net attraction is maximized at a distance of r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere given by $R={2^{-\frac{5}{6}}}\sigma$. At this point the depth of the potential well, given by epsilon $\epsilon$, is a maximum. As r increases further, the net attraction decreases and the potential approaches zero asymptotically. The attractive and repulsive effects can be separated into two potentials, shown with different colors in Figure \ref{fig:LJ_potential_2parts}, which when added together regenerate the full Leonard-Jones potential.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{LJ_Potential_2part.png}
    \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
    \label{fig:LJ_potential_2parts}
  \end{figure}

%\[{}\]
%\[{}\]

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{WCA_Potential.png}
    \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
    \label{fig:WCA_potential}
  \end{figure}

%\[{}\]
%\[{}\]

The repulsive portion of the Leonard-Jones potential forms the potential for a Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. The WCA potential is described by the equation: \begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}

\noindent As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases from infinity as the distance between the centers of the atoms r increases, and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. As with the Lenard-Jones potential, the repulsive WCA potential increases rapidly and approches infinity as the two spheres begin to overlap. This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise. Because the WCA spheres have no hard boundary in their interatomic potential, they are called "soft spheres" and their gradually fading potential makes them slightly "squishy".

\begin{figure}[h!]
    \centering
   % \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=3.5cm]{TwoSpheres.png}
    %\end{subfigure} 
    %\begin{subfigure}[b]{0.4\linewidth}
    %  \includegraphics[height=3.5cm]{2spheres.png}
    %\end{subfigure} 
    \caption{(Left) The WCA potential is a minimum at r=2R. 
             (Right) The WCA potential increases rapidly for $r<2R$.}
    \label{fig:TwoSpheres}
    \end{figure} 
%\[\]

The WCA potential is not well suited for implementing computationally.  Therefore, an erf potential which closely approximates the WCA potential at temperatures of interest will be used instead of the WCA potential. The erf potential is given by: \begin{equation}{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} and is given by \begin{equation}{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}\end{equation}  The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} and is given by: \begin{equation}{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}\end{equation}

\noindent These conditions, along with V$_{erf}=$V$_{WCA}$ at $r=\alpha$, help fit the curve of the erf potential to the curve of the WCA potential.

Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$. Figure \ref{fig:alphaXivsT} shows how $\alpha$ and $\Xi$ vary with temperature when $\sigma=1$ and $\epsilon=1$, and Figure \ref{fig:erf_potential} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential. %As seen in Figure \ref{fig:erf_potential}, the erf potentials closely match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, and for r/$\sigma$ less than ???. % so only values of $V_{erf}$ that are much less than $k_BT$/$\epsilon$ ??? will be considered. 
% In these regions $\alpha$ decreases with increasing temperature, and $\Xi$ increases with increasing temperature.

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=4.5cm]{plot_alpha_vs_KbT.png}
      %\includegraphics[height=5cm]{plot_alpha_vs_KbT.png}
      \caption{$\alpha$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=4.5cm]{plot_xi_vs_KbT.png}
      %\includegraphics[height=5cm]{plot_xi_vs_KbT.png}
      \caption{$\Xi$ versus $k_BT/\epsilon$}
    \end{subfigure} 
    \caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, show how the parameters $\alpha$ and $\Xi$ vary with temperature. (a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then gradually decreases.}
    \label{fig:alphaXivsT}
    \end{figure}

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{Verf.png}
    %\includegraphics[height=5cm]{Verf.png}
    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
    \label{fig:erf_potential}
  \end{figure}

\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
\[\]
\[\]

Another way to find values for $\Xi(T)$ is to find computationally the value $\Xi$ for which the second virial coefficient evaluated for the error function potential $V_{erf}(r)$ is equal to the value for the second virial coefficent evalued for the WCA potential $V_{WCA}$, with $\alpha(T)$ plugged into the equation for $B_{2erf}$ at each temperature T. A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:XifromB2vsT}.
\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
\begin{equation}B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]\end{equation}
and 
\begin{equation}B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} \end{equation}

\[\]
\[\]
\[\]

\begin{figure}[h!]
    \centering
    %\includegraphics[height=7cm]{Xi_from_B2.png}
    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
    \label{fig:XifromB2vsT}
  \end{figure}
\color{red}
\noindent !ADD PLOT of Verf !!!
\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
\color{black}
%\[\]
%\[\]

%----------------BEGIN Virial Expansion--------------
The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
\begin{equation}PV=TS-(U-\mu{N})\end{equation}
\begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
Equation~\ref{Euler_rearranged} then becomes
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\sum{P_i\beta(E_i-\mu
{n_i}})\end{equation}
\begin{equation}\frac{PV}{k_BT}=-\sum{P_i}\left[\ln{P_i}+\beta(E_i-\mu
{n_i})\right]\end{equation}
For a Grand Canonical Ensemble the probablity of a state given by
\begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
results in 
\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[\ln{\left(\frac{ e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}

\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}

\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}\left(\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}} \left[-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\right)\end{equation}

\begin{equation}\frac{PV}{k_BT}=\left[\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\left(\frac{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right) \end{equation}

\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)} \end{equation}

\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \left(e^{\beta\mu}\right)^n\sum_{s(n)}e^{-\beta{E}_{s(n)}}\right)} \end{equation}

\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_n\right)} \end{equation}

\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_{on}Q_n\right)} \end{equation}
where $Q_n$ is the configuration integral.
\color{red}For a system of N particles only, the sum collapses to one term, that for n=N, which gives \color{black}
\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
%=\ln\lambda^n + \ln{Z}_{on}+\ln{Q}_n
=\beta\mu{N} + \ln{Z}_{oN} +\ln{Q}_N\end{equation}
\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN} + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
Noting that $Q_N=1$ for an ideal gas since $V(r)=0$ for an ideal gas
\begin{equation}\frac{P_{ideal}V}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN}=N\end{equation}
%\begin{equation}P_{ideal}=\mu{N} + k_BT\ln{Z}_{oN}\end{equation}
\begin{equation}\frac{P_{excess}}{k_BT}= \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
and Equation~\ref{PV/kBT_reduces_for_N} becomes
\begin{equation}\frac{PV}{k_BT}=N + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
Plugging in the expression for $F_{excess}$ in terms of the mayer function for low density fluids
\begin{equation}F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
gives
\begin{equation}\frac{PV}{k_BT}=N + \left(-\frac{N}{V}\int{f(r)d\vec{r}}\right){N} + \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
%\begin{equation}\frac{PV}{k_BT}=N - \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
\begin{equation}\frac{PV}{k_BT}=N\left(1-\frac{N}{V}\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)\right) \end{equation}
where $B_2(T)$ is given by
\begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(r)d\vec{r} \end{equation}

\color{red}This result was for low densities where higher terms in the expression for the excess free energy were neglected. The full virial equation is given by\color{black}
\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}

%-----------------END Virial Expansion---------------
%\[\]

%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a classical fluid.
%\\

%\[\]
%\[\]
\[\]
\noindent IB. The Theory of Freezing and Maxwell's Double Tangent Construction

Freezing is governed by entropy, which the universe, or any closed system, seeks to maximize according to the second law of Thermodynamics. When the entropy of a pure substance in its liquid state is lower than the entropy that it would have in its solid state at the same temperature (necessary for thermal equilibrium), the same pressure (necessary for mechanical equilibrium), and the same chemical potential (necessary for diffusive equilibrium), an abrupt phase transition occurs, that is, the liquid freezes forming a crystalline structure. The particular crystalline structure and density that result are affected by the interatomic potential. When repulsive, soft sphere fluids (that are not too soft) freeze, they form lattice structures composed of Face Centered Cubic (FCC) crystal cells, as do hard spheres. %This structure is the three-dimensional equivilant to the two-dimensional structure that one would see by filling the bottom of a box with a single layer of marbles. 
This structure is what one would see by filling a cube with marbles. It results in the highest density achievable for hard, and nearly-hard, spheres that are purely repulsive.

%move this paragraph to methods?
Although the entire crystal lattice can be constructed with FCC cells, FCC cells are not the smallest cells with which the lattice can be built. The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell_firsttime} is the smallest structure that, when repeated, re-creates the entire crystal lattice. The parallepiped primitive cell takes up one-fourth the volume of an FCC cell, includes portions of eight atoms for a total of one full atom, and has eight lattice sites convenietly reachable by moving along one of three different lattice vectors to step from one atom to the next. The three lattice vectors will be useful in creating the lattice structure computationally, and will also make the program run faster.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell_firsttime}
  \end{figure}
%-------------------------------

For pure substances, a phase transition occurs along a line on a Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is representative of some arbitrary, unknown substance composed of atoms, or molecules which exibit both attraction and repulsion. However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term "liquid" or "gas" is not used, but simply "fluid". Points along the line that separates the solid and the fluid phases indicate a phase transition, and show that the solid and the fluid both have the same temperature and the same pressure at a phase transistion.

%For each particular temperature, the pressure of the liquid and the solid are the same at the point of transistion. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=5cm]{P-T_Diagram.png}
    \caption{Conceptual Pressure-Temperature Phase Diagram}
    \label{fig:P-T_Diagram}
  \end{figure}
The pressure can be related to the Helmholtz free energy F, which is minimized for a system at fixed temperature, volume and number of particles as the entropy of the system and its surroundings is maximized. The negative of the partial derivative of the Helmholtz free energy with respect to volume V at a fixed temperature T and number of particles N equals the pressure. 
\begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
\noindent Dividing both the Helmholtz free energy and the volume by the number of particles, the partial derivative of the Helmholtz free energy with respect to volume becomes the partial derivative of the Helmholtz free energy per atom with respect to the inverse number density. \begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.

%In a process called Maxwell's Double Tangent construction, two curves are used to find the densities of the liquid and the solid at the point of transistion. Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for the solid both at a fixed temperature. Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope.%
Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid and one for the solid at a fixed temperature. Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope. This process is called Maxwell's Double Tangent Construction.

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{MDTC.png}
    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
    \label{fig:MaxwellDT}
  \end{figure}

To facilitate finding the pressure at the phase transition, %, which occurs at the points which have the same slope,
it is helpful to consider the Gibbs free energy given by G=F+PV. Rearranging the terms and dividing by N gives an  expression for the Helmholtz free energy per atom $f$ in terms of the Gibbs free energy per atom  $g$ (which is also the chemical potential):  %From the Gibbs free energy an expression can be derived which relates  ....:

%\begin{displaymath}{F=-PV+G}\end{displaymath} 
\begin{displaymath}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{displaymath} 
\begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
\noindent This equation has the form of a line:
\begin{displaymath}{y=mx+y_o}\end{displaymath}
%\begin{displaymath}{G_1=G_2}\end{displaymath} 
%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
\noindent and describes the red line shown in Figure \ref{fig:MaxwellDT} with the pressure given by the negative of the slope of the line, and the Gibb's free energy per atom given by the y-intercept. At the point of phase transition:
\begin{displaymath}{g_L=g_S}\end{displaymath} 
\begin{displaymath}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{displaymath}

The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid. 
\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{MDTC-2.png}
    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
    \label{fig:GibbsvsP}
  \end{figure}
%CURVE FITTING! - put in methods/analysis

Once the pressure is known for the transition point, the density of the liquid at the transition point can be found from a plot of the pressure versus the inverse of the density for the liquid. The density of the solid at the transition point can be found the same way, as shown in Figure \ref{fig:Pvsinvn}.
%\[\]

\begin{figure}[h!]
    \centering
    \includegraphics[height=7cm]{MDTC-3.png}
    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
    \label{fig:Pvsinvn}
  \end{figure}

A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this procedure for various temperatures. %The lowest density the crystal exhibts and the highest density the fluid exhibits at the phase transistion can then be identified for each temperature.

\begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{T-n_Diagram.png}
    \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion. The gray region is where the liquid and the solid coexist, and are therefore not pure states.}
    \label{fig:T-n_Diagram}
  \end{figure} 
%One way of implementing this practically is to determine what the Helmholtz free energy would be for both the liquid and solid phases at a particular temperature and density. Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in. 
%\[\]
\[\]
IC. Classical Density Functional Theory (cDFT)

The key idea behind classical Density Functional Theory (cDFT) is that free energy can be written as a functional (a function of a function) of a number density profile $n(\vec{r})$ which describes the way the number density, or atoms per volume, varies spatially. 
\begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}

This, coupled with the fact that the free energy of a system tends toward a minimum at equilibrium means that the equilibrium free energy of the system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying $n(\vec{r})$ through all possible spatial profiles until the free energy is minimized.

\begin{displaymath}{f[n(}\vec{r}{)]}\rightarrow{f_{equil}}{[}\rho{(}\vec{r}{)]}  \mbox{ as $f$ is minimized} \end{displaymath}
\begin{displaymath}{\mbox{where }  n(\vec{r})=\rho{(}\vec{r})  \mbox{ at equilibrium}}\end{displaymath}

The reason this works is because cDFT satisfies the conditions of the variational principle, the same mathematical principle used to determine that the shortest distance between two points is a straight line by varying the path between the two points until the distance is minimized.

A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
\begin{equation}\Phi=F-\mu{N}\end{equation}
%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
which can then be expressed in terms of a varrying number density profile $n(\vec r)$
%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and
\begin{equation}F_{int}= F -\int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume. But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest. 
\color{black}

%*********************** DFT minimize F/N *********************
--------------------------------------------------------
\color{red}
Starting with Euler's equation
\begin{equation}U=TS-PV+\mu{N}\end{equation}
\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)+d\mu\end{equation}
and using the Gibbs-Duhem relation which follows from Euler's equation
\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
which gives
\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
Equation~\ref{d(U/N)long} becomes
\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
\color{black}
The Helmholtz free energy F is defined as 
\begin{equation}F=U-TS\end{equation}
\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
After plugging in Equation~\ref{d(U/N)}, and setting $f_N=\frac{F}{N}$ this becomes
\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed.
\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
Thus the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.

The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will expreience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$. Here "surroundings" means the entire universe minus the system. This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.  How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium? Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
where conservation of energy was also applied. Using Eq~\ref{d(U/N)} for the surroundings, but fixing the volume and the number of particles so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$ results in
\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative. (reference schroder)

More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium. Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
\begin{equation}F=\text{min}_{n(\vec r)}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} An extremum is reached when 
\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=0\end{equation}
Taking this derivative gives  \color{red} CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL \color{black}
\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\left(\frac{1}{T}\right)\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
\begin{equation}=0\end{equation}
As shown earlier, the extremum that is reached at equilibrium is a minimum. (reference shethna) \color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids). \color{black} It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.

------------------------------------------------------
%*************************END DFT minimize F/N ****************

%Before looking at the details of cDFT, let us first look at the free energy. 
%To see how cDFT is adventageous, consider what is involved in a straightforward calculation of the free energy without using Density Functional Theory.

The Helmholtz free energy can be written as
\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$:
\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}\exp^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~where~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~where~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}

Breaking the Hamiltonians H into its component energies:
\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
where the translational kinetic energy of the N atoms is given by: \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
the potential energy of the total interaction of each atom with the other N-1 atoms is given by:
\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position. Often $V_{ext}$ is set to zero, as it is in this paper.

Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts: 
\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume defined by: \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
the equation can be written:
\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)
}\end{equation}  and the partitian function can be expressed as:
\begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
where:
\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another. The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential. The free energy can also be broken into two parts using Eq~\ref{Zmultiply}:
%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}

The ideal Helmholtz free energy is
\begin{equation}{F_{ideal}= -k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation} where 
%the integral can easily be evaluated:

\begin{displaymath}{\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}=\frac{1}{h^{3}}\int{d\vec{p}_1}\exp^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}\exp^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}\exp^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}}\end{displaymath}
%Each term on the right 
These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$:
\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}\exp^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
Rewriting equation (20) in terms of $\Lambda$ gives:
\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes:
\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$: \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
\begin{equation}{F_{ideal}= k_BT[n\ln(n\Lambda^3)-n]V}\label{eq:Fideal}\end{equation}   
For a spatially varying number density, the ideal free energy in Eq~\ref{eq:Fideal} becomes:
\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation} 
\[\]
xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx

Now consider the excess free energy which is written as:
\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation} 
This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes: \begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in equation \ref{Fexcess} becomes:

\begin{displaymath}{\exp^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=\exp^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}\exp^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}\exp^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. %\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    %$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
The function $f_{ij}$ is called a Mayer function, and is given by\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=\exp^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} Multiplying the Mayer function terms, the excess free energy can be written as
\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}
At low densities, it is not likely to find two or more spheres in close proximity, and so only interactions between pairs of spheres are considered. In this case the equation reduces to
\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation} Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ we get \begin{equation}{\int{...}\int{dQ}=V^N}\end{equation}\begin{equation}{\int{...}\int{dQ}\sum{f_{ij}}=\frac{1}{2}N(N-1)V^{N-1}\int{f(r)}{d\vec{r}}}\end{equation} \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}}=\frac{1}{8}N(N-1)(N-2)(N-3)V^{N-2}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation} %which becomes \begin{equation}{F_{excess}=-k_BT\ln{\left(1+ \frac{1}{2}\frac{N(N-1)}{V^{N}}V^{N-2}\int{...}\int{dQ}{f_{ij}} + .....\sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...\right)}}\end{equation}% 
%FINISH!!!!! \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}f_{mn}}=\frac{1}{48}N(N-1)(N-2)(N-3)(N-4)(N-5)(N-6)V^{N-3}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation} 
and so on. In the low-density limit, the terms in the series get smaller, and the excess free energy can then be written as %begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{8}\frac{N^4}{V^{2}}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}+ ...\right)}\end{equation}%
\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ for a homogeneous density n=$\frac{N}{V}$ this becomes
\begin{equation}{F_{excess}=-k_BT\ln{\left(exp^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} \begin{equation}{F_{excess}=-k_BT\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\end{equation}  %\begin{equation}{F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation}%
Thus the free excess energy per atom in the low-density limit is  \begin{equation}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} where the quantity in parenthesis can be recognized as the second virial expanision coefficient typically called $B_2$. %where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
It is important that the expression for the excess Helmholtz free energy correctly reduces to this result in the low-density limit.
%Although we will not be restricting ourselves to the low-density limit, it is important that the theory we use correctly reduces to this same result in the low-density limit.

%(FINISH! Take P=-dF/dV and set equal to virial expansion for P to get virial coefficients in terms of the mayer functions....CLARIFY why cDFT is better than this route!)

Another way to form an expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ is to use Fundamental Measure Theory (FMT). FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory (SPT). Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.

%NEW-------Scaled Partical Theory-----------------
--------------------------------------------------

--------------------------------------------------

Scaled Partical Theory (SPT) gives an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
\begin{equation}\label{SPT_FE}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
where
\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}

%\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
%\begin{equation}{\eta = \left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}}\end{equation}
%\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
From this expression, a quantity $\zeta_{3}$ can be defined as follows
\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=n\zeta_{3}}\end{equation}
where $\zeta_{3}$ is the volume of a sphere. Likewise,
\begin{equation}{\zeta_{2}=4\pi{R^2}}\end{equation}
\begin{equation}{\zeta_{1}=R}\end{equation}
\begin{equation}{\zeta_{0}=1}\end{equation}
where $\zeta_{2}$ is the surface area of a sphere, and $\zeta_{1}$ is the radius of a sphere. The quantities $\zeta_{3}, \zeta_{2}, \zeta_{1},$ and $\zeta_{0}$ are fundamental geometric measures of a sphere, and they will prove useful by defining a set of weighted densities $n_{i}=n\zeta_{i}$ where $n=\frac{N}{V}$. 
\begin{equation}\label{n3}{n_{3}=n\zeta_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
\begin{equation}\label{n2}{n_{2}=n\zeta_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
\begin{equation}\label{n1}{n_{1}=n\zeta_{1}=\left(\frac{N}{V}\right)R}\end{equation}
\begin{equation}\label{n0}{n_{0}=n\zeta_{0}=\left(\frac{N}{V}\right)}\end{equation}

\noindent The excess free energy in Eq~\ref{SPT_FE} can be written in terms of the number density $\frac{N}{V}$
\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
\begin{equation}\label{FexfromSPT}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V}\end{equation}
which has the form 
\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
where 
\begin{align}
   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2} 
\end{align}

The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted densities given the relation $n_{i}=n\zeta_{i}$, where the weights are given by $\zeta_{0},\zeta_{1},\zeta_{2},\zeta_{3}$ respectively. These weights are valid for determining the excess free energy of a system of hard spheres with a homogeneous number density n. But these weights, and their related  weighted densities, can be written in a more general form allowing for their extention to a system with a spacially varying number density, that is, an inhomogeneous fluid.

Using a backwards Heavyside step function defined as
\begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
the following expressions are obtained
\begin{equation}{\zeta_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
\begin{equation}{\zeta_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}

%\begin{equation}{\zeta_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
\begin{equation}{\zeta_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
%\begin{equation}{\zeta_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
\begin{equation}{\zeta_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}

The weighted densities can now be expressed as
\begin{equation}{CHECK!n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} with weight functions
\begin{equation}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
\begin{equation}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}

%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
where the weight functions $w_{2}$ and $w_{3}$ have the relation
\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}

The weight functions can be related to the mayer function \begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} where r is the distance between the centers of two hard spheres. For a hard sphere fluid the interatomic potential $V(r)$ is given by \begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
and so the mayer function for a hard sphere is given by \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
The mayer function can be written in terms of weight functions if two additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  where the vector weights are defined as \begin{equation}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
\begin{equation}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}

Taking the derivative of the mayer function in Eq~\ref{f(r)step} will prove useful, and is given by
\begin{equation}\label{mayerder=w2convolution}{\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}}\end{equation} 
where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
\begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
Using the general relation for convolution, which is equal to the inverse Fourier Transform of a product of Fourier Transforms \begin{equation}{f \otimes g = \text{IFT[ FT}[f]\text{ x FT[g] }]}\end{equation} with
\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
\begin{equation}{\text{Fourier Transform}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
the right side of Equation~\ref{mayerder=w2convolution} gives
\begin{equation}{w_2 \otimes w_2 = \text{Inverse Fourier Transform}[e^{-ikR}e^{-ikR}]}\end{equation} 
\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
which matches Equation~\ref{mayerder}.

Incorporating into Equation~\ref{FexfromSPT} the two vector weights which were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated expression for $\Phi_3$ developed by Tarazona, the excess free energy becomes
\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation} 
where the tensor weight is given by
\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
%and $\Phi_3$ is now
and
\begin{align}
\Phi_1 &= -n_{0}ln(1-n_{3}) \\
\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  
\end{align}

The expression for the weighted densities can be written in a way that accomodates both homogenious fluids where the number denstiy n is constant, and inhomogeneous fluids where the number density becomes a function of the position $n(\vec{r})$.
\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
%The equatons above were derived for a uniform fluid where the number density n is constant, but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
For a spatially varying number density $n(\vec{r})$, the excess Free Energy becomes a functional of the number density profile $n(\vec{r})$ :
\begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}

The equations developed so far have been for a hard sphere fluid. All of the weight function equations, as well as the equation for the deriviative of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists of a delta function characteristic of hard spheres. Soft Fundamental Measure Theory (SFMT) extends the Fundamental Measure Theory discussed so far to fluids with soft potenials by keeping the relationships between the weight functions, and that the weight functions satisfy the same, but replacing $w_{2}(r)$ with a weight function appropriate for a soft sphere potential. A new set of weight functions can then be derived for a system of soft spheres.

In this paper,  $w_{2}(r)$ will be constructed using the following expresssion for $w_3(r)$ proposed by Schmidt$^{(?)}$ 
\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
 where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ and $w_3(\infty)=0$. 
\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
Using Equation~\ref{w2_w3_relation}, shown again below, 
\begin{displaymath}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{displaymath}
and solving for $w_2(r)$
\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
gives
\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
which can be integrated to obtain the Mayer function
\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
The Mayer function for the error function potential $V_{erf}(r)$ is thus given by
\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} %This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
\begin{equation}{V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} The two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), and are determined to make $V_{erf}(r)$ match $V_{WCA}(r)$ as closely as possible by compensating for the unwanted temperature dependence that shows up in the error function potential $V_{erf}(r)$.

--------------------------------------------------
%END NEW------------------------------------------

%In summary, it is the aim of this paper to use these expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.

In summary, classical Density Functional Theory (cDFT) used in conjunction with Soft Fundamental Measure Theory (SFMT) provides a powerful tool in modeling the thermodynamic properties of fluids. With SFMT an expression for the excess free energy in terms of the number density can be found simply and directly without taking into account the interaction each sphere, or atom, has with every other sphere. For a spatially varying number density, as would be expected for a crystal, the free energy can then be written as a functional of the number density profile $n(\vec{r})$ 
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
To implement cDFT, the number density profile $n(\vec{r})$ is varied until the free energy is minimized. The minimum free energy and its corresponding number density profile are then the equilibrium values for that temperature. \color{red} $\leftarrow$CAREFUL! \color{black} Thus to use cDFT all that is needed are number density profiles, such as Gaussian profiles that depend on position and which can be varied, for example, by changing the width of the Gaussian. 
\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation}

For a number density n that is constant everywhere, as would be expected for a homogeneous fluid in equilbrium, the free energy can be written as a function of n where the ideal Helmholtz free energy can be determined analytically, and SFMT is used once again to form an expresssion for the excess Helmholtz free energy
\begin{equation}{F(n)=F_{ideal}(n) + F_{excess}(n)}\end{equation}

Once the Helmholtz free energy for both the crystal and liquid states of the fluid are found, they can be compared to find the resulting state of the fluid. Outside the region of coexistence, the fluid will be completely in a crystalline state if the crystal free energy is lower than the liquid free energy, and it will be completely in a liquid state if the liquid free energy is lower than the crystal free energy as was discussed in Section IB. 
%The complexity is greatly reduced making cDFT a powerful tool in modeling fluid behavior.
\[{}\]
II. METHODS

To implement cDFT computationally, we begin by picturing one single crystal cell. Although the crystal structure is composed of Face Centered Cubic (FCC) cells, the entire lattice can be generated with a cell one-fourth the size called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}. The primitive cell has the shape of a parallelpiped, and contains the equivalent of one atom.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
    \label{fig:primitivecell}
  \end{figure}

  \begin{figure}[h!]
    \centering
    \includegraphics[height=4.5cm]{Simple_Cell.png}
    \caption{A Simplified Crystal Cell}
    \label{fig:simplecell}
  \end{figure}

This cell is rather complicated for the purposes of this discussion, so consider instead a simplified crystal cell with one atom in its center as shown in Figure \ref{fig:simplecell}. 
The Helmholtz free energy is minimized for a fixed volume (and temperature and number of particles).  A fixed cubical volume called the "bold-box" volume, $V_{bold-box}$ will be used for illustration. The number density n for a collection of simplified cells is simply 1 atom per fixed bold-box volume.
\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per "box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density). Reduced number density is given by n* and is equal to the number of atoms in a "box" of size $\sigma^3$.
\begin{displaymath}{n*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}

In a large collection of simple cells, the atom at the center of each cell will not realistically reside exactly at the lattice point in the center of the cell. Statistically there will typically be a normal distribution of positions about the lattice point as pictured in the far right box in Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian distribution is shown in Figure \ref{fig:Gaus_plot}. A 3-dimenstional Gaussian distribution centered at a lattice point can be used to represent a number density profile $n(\vec{r})$  which describes how the atoms are statistically distrubuted over a collection of cells.

  \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Gaus}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{Gaussian_Plot.png}
    \caption{A 2D Gaussian Distribution of width 2$\sigma$}
    \label{fig:Gaus_plot}
  \end{figure}

It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per "bold box" volume.

\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}

In a large collection of cells, not all lattice sites will be occupied. Vacancies can be taken into account while keeping the volume of the bold box, and thus the overall number density, the same. This is done by making the cells smaller. A large collection of smaller cells creates a new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} with eight smaller Gaussians of one-eigth the height.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
    \caption{A Gaussian Distribution Number Density Profile}
    \label{fig:Ensemble_Smallcells}
  \end{figure}

Instead of thinking of an atom in one of the eight small cells, the same Gaussian distribution can be obtained from thinking of one-eigth of an atom in each small cell. This makes it convenient to describe the new reduced number density in terms of the fraction of vacancies, $f_v$:

\begin{displaymath}{ n^* = (1-f_v){n_{no.vacancies}^*}}\end{displaymath}

\begin{figure}[h!]
    \centering
    \includegraphics[height=2.5cm]{SameStatPic.png}
    \caption{Front views of a cube with eight smaller cells. The two views show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which give rise to the same Gaussian distribution statistically}
    \label{fig:SameStatPic}
  \end{figure}

By varying the width of the Gaussians and the fraction of vacancies, new number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.

\begin{figure}[h!]
    \centering
    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
    \caption{Several Gaussian Distribution Number Density Profiles all with a number density of 1 atom per $V_{bold box}$}
    \label{fig:Ensemble_vary}
  \end{figure}

Once again, it is important to note that the number density is the same for all these number density profiles $n(\vec{r})$. There is still a total of one atom per "bold box" volume. This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.

 \begin{figure}[h!]
    \centering
    %\includegraphics[height=4cm]{HomogeneousDen.png}
    \includegraphics[height=4cm]{Homogeneous_bold-box.png}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

A simplified flowchart of the program created to investigate the freezing of a WCA fluid is shown on pages x-x. The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.The number density profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found for that combination of n* and t*, which is known from cDFT to be unique. The process is then repeated to find the free energies at other number densities and temperatures.

In general, if the crystal free energy at some n* and t* is lower than the homogeneous free energy of a liquid at that same n* and t*, then the crystaline state dominates. But if the crystal free energy is higher than the homogeneous free energy, the liquid state dominates. However, there is a region of densities between the liquid and crystal states that are forbidden, and so the actual state can not be determined just by which energy is lower at that particular n* and t*. Maxwell's Double Tangent Construction must be implemented using the date collected to determine the actual crystal and liquid states, and construct phase diagrams. 
\[{}\]
\textbf{Program Implementation}

The program computes both the crystal and the homogeneous Helmholtz free energies. The Helmholtz free energy is minimized for a system with a fixed temperature, volume, and number density. The temperature, T is an input to the program. The number density, which is also an input to the program, is used to calculate the volume of one primitive cell over which the free energy is computed. The number of atoms in one primitive cell is one. To compute the free energy, it is necessary to compute both the ideal and excess free energies and add them, $F = F_{ideal} + F_{excess}$.
%\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
%\begin{displaymath}F_{crystal} = F_{ideal.crystal} + F_{excess.crystal}\end{displaymath}
%\newline
%\newline\noindent Inputs: 
%t*, n*, range of gw values , range of fv values
%\newline Outputs: Homogeneous (liquid)  Helmholtz free energy (at the given T,V,N) 
%\newline\indent\indent\indent Crystal Helmholtz free energy (at the given T,V,N)

%\noindent Input:
\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
%\noindent Set:
%\begin{displaymath}\text{N = 1}\end{displaymath}
\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
%Outer Loop:  %set fv to one of the input values
Loop: 
\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
%\indent\indent\indent Inner Loop : %set gw to one of the input values
\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}

%Compare:
\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
%\noindent Output:
\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
%\begin{displaymath}\left(\frac{F}{N}\right)_{homogeneous}\end{displaymath}
%lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
\[\]

The Liquid (homogeneous) ideal free energy $hF_{ideal}$ is computed analytically for a given n* and t*, where $\sigma=1$ so that n*=n$\sigma^3$=n, and $\epsilon=1$ so that t*=$\frac{k_BT}{\epsilon}=k_BT$. 
%\begin{displaymath}{\frac{hF_{ideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
\begin{displaymath}{\frac{hF_{ideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} where

\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  and
\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units!\color{black}\end{displaymath} 
%\begin{displaymath}{m=}\end{displaymath}  
%\begin{displaymath}{\hbar=}\end{displaymath} 
%\begin{displaymath}{k_B=}\end{displaymath}

The Liquid (homogeneous) excess free energy $hF_{excess}$ is computed analytically for a given n* and t* using SFMT with a uniform number density profile that does not depend on $\vec r$. 
%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
\begin{displaymath}{\frac{hF_{excess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
For the case of a homogeneous number density n, $\Phi_1$, $\Phi_2$, and $\Phi_3$ simplify to:
\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
where \begin{displaymath}{n_i=nw_i}\end{displaymath} 
%\begin{displaymath}{n_0=nw_0}\end{displaymath} 
%\begin{displaymath}{n_1=nw_1}\end{displaymath}
%\begin{displaymath}{n_2=nw_2}\end{displaymath}
%\begin{displaymath}{n_3=nw_3}\end{displaymath}
and the weights are as usual 
\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
\begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation}
\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
%\begin{displaymath}{w_0=?}\end{displaymath} 
%\begin{displaymath}{w_1=\frac{\alpha}{2}}\end{displaymath}
%\begin{displaymath}{w_2=6.2...\Xi^2 + \pi\alpha^2}\end{displaymath}
%\begin{displaymath}{w_3=\alpha(\pi\Xi^2 + 0.52...\alpha^2)}\end{displaymath} \[\]
\[\]

The Crystal ideal free energy $cF_{ideal}$ is computed in two different ways from the equation 
\begin{displaymath}{F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV}\end{displaymath}
with the varying number density profile given by 
\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}}\end{displaymath}
\newline $\bullet$ For small gw values it is assumed that the Guassians do not overlap, in which case the equation for the free energy can be simplified and solved analytically with the result
\begin{displaymath}{cF_{ideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)}\end{displaymath}

%\begin{displaymath}{cF_{ideal}=(1-f_v)k_BT\left(\ln(\frac{(1-fv)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)-DELETE-EQN!}\end{displaymath}

$\bullet$ For larger gw values it is assumed that the Gaussians do overlap, and so the free energy is caluculated computationally. First dx is scaled by gw
\begin{displaymath}{dx_{scaled}=(dx_{input})(gw)}\end{displaymath}
then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the primitive cell taking into account the contribution of Gaussians from many (5x5x5) nearby cells whose centers are given by $R_i$.
%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
The contribution to the free energy of $n(\vec{r})$ at the one position is then added to the accumulating sum.
%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
This process is repeated for all positions within one primitive cell to get the crystal ideal free energy for the primitive cell. This energy is then divided by the volume of the primitive cell to get the crystal ideal free energy per volume.
%\[\]

The crystal excess free energy $cF_{excess}$ is computed computationally using SFMT which gives
%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
%Then the crystal, or inhomogeneous, excess free energy is found from
\begin{displaymath}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} and using Monte-Carlo integration to compute the weighted densities $n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$ used to form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.

The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ within the primitive cell, and a vector $\vec{R}_j$ to the center of one of the Gaussians is selected that is within the primitive cell, or one of the neighboring cells within a prudently picked inclusion radius as Gaussians farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).

 \begin{figure}[h!]
    \centering
    \includegraphics[height=5cm]{InclusionRadius.png}
    \caption{The Gaussians at each lattice point are of width gw. After a radius of few gw's the Gaussians go to zero and can be neglected.}
    \label{fig:InclusionRadius}
  \end{figure}

The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, $n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions $w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to some point in space. Monte-Carlo integration works by randomly generating $\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + \vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ is the gaussian number density profile for no vacancies, the equation:

\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} implemented computationally in the form:
\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} becomes:
\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ (the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of $w_i(|\vec{r}-\vec{r}'|)$ will occur. 
Using the relation:
\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} $n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}

The result can be made more accurate by increasing NumPoints which is the number of vectors randomly generated. In order to further reduce the error in the average, antithetic variates are used. Rather than computing an average of a set of random points, each random point in the set is paired with a point along the same line as the first point, but opposite in direction, as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order error in the average by filtering out an erroneous offset of values along the line connecting the two points by averaging the two results, one of which may be erroneously higher than it should be while the other is erronesously lower than it should be.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{AntitheticVariate.png}
    \caption{Antithetic variates reduce first order error in an average. Each point in the average is paired with another point from along the same line as the first, and at an equal, but opposite distance.}
    \label{fig:AntitheticVariate}
  \end{figure} 
Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians. 
\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
Then the entire process repeated for each vector $\vec{r}_k$ to get the total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.

The standard error of the mean (SEM) is given by:
\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives a measure of the error in $n_3=\overline{w}_3$:
\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} NumPoints, the number of vectors randomly generated, is increased until:
\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum value that $n_3$ can be without error is 1.

Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions $w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, $w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid sphere up to a radius of ?FIX? after which it tapers off over a distance of ?FIX?.

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{weight_functions.png}
    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the z-axis for a reduced temperature of 2}
    \label{fig:weight_functions}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{W2.png}
      \caption{A crossectional view of $w_2$}
    \end{subfigure} 
    \begin{subfigure}[b]{0.4\linewidth}
      \includegraphics[height=5cm]{W3.png}
      \caption{A crossectional view of $w_3$}
    \end{subfigure} 
    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
    \label{fig:W0andW3}
    \end{figure}

Figure \ref{fig:GaussandW2_actual} ......

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{GaussandW2_actual.png}
    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ is generated randomly according to a Gaussian distribution.} 
  \label{fig:GaussandW2_actual}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  and fixed in place as $\vec{dr}$ varies randomly.} 
  \label{fig:GaussandW2_thinkas}
  \end{figure}

The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which become wider in general as the temperature increases. The radial distance to the peak decreases as the temperature increases, that is, the shells shrink, as does the sphere. The functions can be treated as neglible outside of a distance of about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of the weight function.

\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
III. RESULTS
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]
\[{}\]

APPENDIX
\begin{enumerate}

\item Fourier Transforms of Weight Functions $w_0(r), w_1(r), w_2(r), w_3(r)$
\item Fourier Transforms of Vector Weight Functions $\vec{w}_1(\vec{r}), \vec{w}_2(\vec{r})$  
\item Fourier Transform of Tensor Weight $\overleftrightarrow{w}_{m2}(r)$ in coordinate system (x',y',z')
%\item The tensor weight components in the coordinate system (x,y,z) 
%\item Show transform of tensor weight $\overleftrightarrow{w}_{m2}(r)$ goes to zero as k goes to 0
\item Integrals used to Compute Fourier Transforms of Weight Functions
\item Deriviation of Tensor Density $n_{m2ij}(\vec{r})$ 
\item Deriviation of the Tensor version of $\Phi_3$ 
\item Equations related to the Error Function Potential $V_{erf}$ ....
\item Second Virial Coefficient $B_{2}$ for the Error Function Potential 
\item Third Virial Coefficient $B_{3}$
\end{enumerate}

\textbf{1a. Fourier Transform of Weight Function $w_0(r)$:}
\begin{equation}{w_0(r)=\frac{w_2(r)}{4{\pi}r^2}}\end{equation}
where
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
temporarily set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_0(r)=\frac{1}{4{\pi}\sqrt{\pi}a}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}
%\[{}\]
\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
\[{}\]
Calculate $\widetilde{w}_0(k)$ 
\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
\[{}\]
\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
\[{}\]
Set the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero. 
\begin{equation}{\widetilde{w}_0(k)=\frac{1}{\sqrt{\pi}ka}\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}}\end{equation}
\[{}\]
\begin{equation}{\widetilde{w}_0(k)=\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
\[{}\] 
Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_0(k)$ becomes
\begin{equation}{\widetilde{w}_0(k)=\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}

\[{}\]
%\[{}\]
\textbf{1b. Fourier Transform of Weight Function $w_{1}(r)$:}
\begin{equation}{w_1(r)=\frac{w_2(r)}{4{\pi}r}}\end{equation}
where
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
temporarily set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_1(r)=\frac{1}{4{\pi}\sqrt{\pi}a}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}

\begin{equation}{\widetilde{w}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_1(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

\noindent Calculate $\widetilde{w}_1(k)$ 
\begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
\[{}\]
\begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
\[{}\]
\begin{equation}{\widetilde{w}_1(k)=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\frac{2\sin(kr)}{kr}d{r}}\end{equation}
Set the lower limit to $-\infty$ since the integrand is nearly zero by the time r is zero. 
\begin{equation}{\widetilde{w}_1(k)=\frac{1}{\sqrt{\pi}ka}\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}}\end{equation}
After doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$ this becomes
\begin{equation}{\widetilde{w}_1(k)=\frac{1}{k}\sin\left(\frac{k\alpha}{2}\right)e^{-\left(\frac{k^2a^2}{4}\right)}}\end{equation}
Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{w}_1(k)$ becomes
\begin{equation}
    \widetilde{w}_1(k)=\frac{1}{k}e^{-\frac{k^2\Xi^2}{8}}\sin\left(\frac{k\alpha}{2}\right)
\end{equation}

\[{}\]
%\[{}\]
\textbf{1c. Fourier Transform of Weight Function $w_{2}(r)$:}
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_2(r)=\frac{1}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}

\begin{equation}{\widetilde{w}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_2(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

\noindent Calculate $\widetilde{w}_2(k)$ 
\begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
\[{}\]
\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
\begin{equation}{\widetilde{w}_2(k)=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\frac{2\sin(kr)}{kr}d{r}}\end{equation}
Set the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero.
\begin{equation}{\widetilde{w}_2(k)=\frac{4\sqrt{\pi}}{ka}\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}}\end{equation}
After doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$ this becomes
\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{k}e^{-\left(\frac{k^2a^2}{4}\right)}\left(ka^2\cos\left(\frac{k\alpha}{2}\right)+\alpha\sin\left(\frac{k\alpha}{2}\right)\right)}\end{equation}
Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{w}_2(k)$ becomes
\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{k}e^{-\frac{k^2\Xi^2}{8}}\left[\frac{k\Xi^2}{2}\cos\left(\frac{k\alpha}{2}\right)+\alpha\sin\left(\frac{k\alpha}{2}\right)\right]}\end{equation}

\[{}\]
%\[{}\]
\textbf{1d. Fourier Transform of Weight Function $w_{3}(r)$:}
\begin{equation}{w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]}\end{equation}
temporarily set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]}\end{equation}

\begin{equation}{\widetilde{w}_3(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}$ 
\begin{equation}{\widetilde{w}_3(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

\noindent Calculate $\widetilde{w}_3(k)$ 
\begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
\[{}\]

\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}

\begin{equation}{\widetilde{w}_3(k)=\pi\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\frac{2\sin(kr)}{kr}d{r}}\end{equation}

\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{k}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r\sin(kr)d{r}}\end{equation}
Integrating by parts with 
\begin{displaymath}{f=1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)}\end{displaymath}
\begin{displaymath}{g=\frac{\sin(kr)-kr\cos(kr)}{k^2}}\end{displaymath}

\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{k}\left[\left(1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right)\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\bigg|^{\infty}_0\color{white}\right]}\end{equation} \color{black}
\begin{displaymath}{-\int_{0}^{\infty}\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\left(\frac{2}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\right)d{r}}\end{displaymath}
Evaluating the integral by doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$, and setting the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero this becomes
\begin{equation}{\widetilde{w}_3(k)=\frac{4\pi}{k^3}e^{-\frac{k^2a^2}{4}}\left[\left(1+\frac{k^2a^2}{2}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{k\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]}\end{equation}
Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{w}_3(k)$ becomes
\begin{equation}{\widetilde{w}_3(k)=\frac{4\pi}{k^3}e^{-\frac{k^2\Xi^2}{8}}\left[\left(1+\frac{k^2\Xi^2}{4}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{k\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]}\end{equation}

\[{}\]
%\[{}\]
\textbf{2a. Fourier Transform of Vector Weight Function $\vec{w}_{1}(r)$:}
\begin{equation}{\vec{w}_1(\vec{r})=w_1(r)\frac{\vec{r}}{r}}\end{equation}
where
\begin{equation}{w_1(r)=\frac{w_2(r)}{4{\pi}r}}\end{equation}
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
temporarily set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_1(r)=\frac{1}{4{\pi}\sqrt{\pi}a}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}\vec{w}_1(\vec{r})e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate system $(x',y'z')$ 
\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
where 
\begin{displaymath}{r_{x'}=r\sin\theta'\cos\phi'}\end{displaymath}
\begin{displaymath}{r_{y'}=r\sin\theta'\sin\phi'}\end{displaymath}
\begin{displaymath}{r_{z'}=r\cos\theta'}\end{displaymath} 
The components in the $x'$ and $y'$ directions vanish due to symmetry about $z'$. That leaves
\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right]{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{\vec{w}}_1(k)$ becomes
\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}

\[{}\]
%\[{}\]
\textbf{2b. Fourier Transform of Vector Weight Function $\vec{w}_{2}(r)$:}
\begin{equation}{\vec{w}_2(\vec{r})=w_2(r)\frac{\vec{r}}{r}}\end{equation}
where
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
temporarily set 
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{w_2(r)=\frac{1}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate system $(x',y'z')$ 
\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
where 
\begin{displaymath}{r_{x'}=r\sin\theta'\cos\phi'}\end{displaymath}
\begin{displaymath}{r_{y'}=r\sin\theta'\sin\phi'}\end{displaymath}
\begin{displaymath}{r_{z'}=r\cos\theta'}\end{displaymath} 
The components in the $x'$ and $y'$ directions vanish due to symmetry about $z'$. That leaves
\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{w}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(-\frac{2i}{kr}\cos(kr) + \frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4i\sqrt{\pi}}{ak}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r} + \frac{4i\sqrt{\pi}}{ak^2}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}{~~}\hat{z}'}\end{equation}

\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4\sqrt{\pi}i}{ak}\left[\frac{a\alpha\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]}\end{equation} 
\begin{displaymath}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~~}\hat{z}'}\end{displaymath}
\[{}\]
\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation} 
Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{\vec{w}}_2(k)$ becomes
%\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation} 
\[{}\]
\textbf{3. Fourier Transform of Tensor Weight $\overleftrightarrow{w}_{m2}(r)$:}
\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation}
where
\begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
Form the outer product
\begin{equation}{\vec{r}\vec{r}=\left(\begin{array}{c} r_x \\ r_y \\ r_z \end{array} \right) \left(\begin{array}{rrr} r_x & r_y & r_z \end{array} \right)=\left(\begin{array}{ccc} {r^2}_x & r_xr_y & r_xr_z \\ r_yr_x & {r_y}^2 & r_yr_z \\ r_zr_x & r_zr_y & {r_z}^2 \end{array}\right)}\end{equation}
and identity matrix
\begin{equation}{I=\left(\begin{array}{ccc} 1 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 1 \end{array}\right)}\end{equation}
The tensor weight matrix elements are then given by
\begin{equation}{w_{m2_{ij}}(\vec{r})=\frac{1}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)}\end{equation}
with temporarily setting
\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
The fourier transform of the tensor weight matrix elements are given by
\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
%\mbox{~%\scriptsize\textbullet~}
Since the integration is over all space, the integral can be simplified by setting 
\begin{equation}{\vec{k}=k\hat{z}}\end{equation}
\begin{equation}{\vec{k}\cdot\vec{r}=kr\cos\theta\hat{z}}\end{equation}
Instead of constraining vector $\vec{k}$ to be in the z direction, however, a rotated coordinate system $(x',y',z')$ is used which is rotated from the (x,y,z) coordinate system in such a way that $\hat{z}'$ points in the direction of vector $\vec{k}$. After solving for the elements of the transformed weight function in coordinate system $(x',y',z')$, they will be expressed again in terms of vector $\vec{k}$ as it appears in coordinate system (x,y,z) to get the elements of the transformed weight function in coordinate system (x,y,z).%The weight components of interest are those corresponding to the component of vector $\vec{r}$ parallel to vector $\vec{k}$ which will be called $r_{z'}$, and to the two components of $\vec{r}$ perpendicular to vector $\vec{k}$ (and to each other) which will be called $r_{x'}$ and $r_{y'}$. 
\begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
%\begin{displaymath}{r_x=r\sin(\theta)\cos(\phi) ~~~~~ r_y=r\sin(\theta)\sin(\phi)~~~~~ r_z=r\cos(\theta)}\end{displaymath}
where $r=r'$ and
\begin{displaymath}{r_{x'}=r\sin\theta\cos\phi}\end{displaymath}
\begin{displaymath}{r_{y'}=r\sin\theta\sin\phi}\end{displaymath}
\begin{displaymath}{r_{z'}=r\cos\theta}\end{displaymath} 
with
\begin{displaymath}{\delta_{ij}=\left\{ \begin{array}{rc} 1 & i = j \\ 0  & i\neq j \end{array}\right.}\end{displaymath}
\[{}\]

%\[{}\]
%\[{}\]
Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
%where \begin{displaymath}{r_x=r\sin(\theta)\cos(\phi)}\end{displaymath}
%\begin{displaymath}{r_y=r\sin(\theta)\sin(\phi)}\end{displaymath}
\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{y'})}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$

Calculate $\widetilde{w}_{{m2}_{x'z'}}(k)$ 
\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{z'})}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
\[{}\]
Calculate $\widetilde{w}_{{m2}_{y'z'}}(k)$ 
\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{z'})}{r^2}-\frac{\delta_{yz}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
\[{}\]
Calculate $\widetilde{w}_{{m2}_{x'x'}}(k)$ 
\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{x'})}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\[{}\]
%\[{}\]
%\begin{displaymath}{\widetilde{w}_{{m2}_{xx}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\sin^2(\theta)e^{-ikr\cos(\theta)}\left(\int_{0}^{2\pi}\cos^2(\phi)d{\phi}\right)d{\cos(\theta)}\right]d{r}}\end{displaymath} 
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
\[{}\]
\color{blue}
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
\color{black}
\[{}\]
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}-\frac{4\sin(kr)}{k^3r^3}\right)d{r}}\end{displaymath} 
\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\frac{2\sin(kr)}{kr}\right)d{r}}\end{equation}
%\[{}\]
Set the lower limit to -infinity since the integrand is nearly zero by the time r is zero. 
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{4\sqrt{\pi}}{3ka}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}\right)}\end{displaymath} 
\begin{displaymath}{-\frac{4\sqrt{\pi}}{k^2a}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}\right)}\end{displaymath} 
\begin{equation}{+\frac{4\sqrt{\pi}}{k^3a}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right)}\end{equation} 
\[{}\]
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{4\sqrt{\pi}}{3ka}\left(\frac{ka^3}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})+\frac{a\alpha}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})\right)}\end{displaymath} 
\begin{displaymath}{-\frac{4\sqrt{\pi}}{k^2a}\left(a\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})\right)}\end{displaymath} 
\begin{equation}{+\frac{4\sqrt{\pi}}{k^3a}\left(\frac{\pi}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]\right)}\end{equation} 
\[{}\]
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-2\pi{a}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{displaymath} 
\begin{equation}{+\frac{2\pi\sqrt{\pi}}{k^3a}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation} 
Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_{{m2}_{xx}}(k)$ becomes
\[{}\] \color{green}
%\color{red}
%ERROR\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=4{\pi}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\left[\left(\frac{\Xi}{3k\sqrt{2}}-\frac{1}{k^2}\right)\cos(\frac{k\alpha}{2})+\frac{\Xi\alpha}{12*\sqrt{2}}\sin(\frac{k\alpha}{2})\right]}\end{displaymath} 
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath} 
\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation} 
\color{black}
\[{}\]
Calculate $\widetilde{w}_{{m2}_{y'y'}}(k)$ 
\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\[{}\]
\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath} 
\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}

\color{blue}
\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
\color{black} 
\[{}\]
This is the same as equation \ref{compare_equation}, and so \begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\widetilde{w}_{{m2}_{x'x'}}(k)}\end{equation}
Calculate $\widetilde{w}_{{m2}_{z'z'}}(k)$ 
\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
\[{}\]
\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath} 
\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}

\color{blue}
\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}

\begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
\color{black}

This is the same as -2 times equation \ref{compare_equation}, and so \begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=-2\widetilde{w}_{{m2}_{x'x'}}(k)}\end{equation}

%\subsection{The tensor weight components in the coordinate system (x,y,z)}
\noindent The tensor weight components in the coordinate system $(x',y',z')$ are:
\begin{equation}\label{tensorcomp}{\widetilde{w'}_{m2}=\left(\begin{array}{ccc} \widetilde{w}_{{m2}_{x'x'}} & 0 & 0 \\ 0 & \widetilde{w}_{{m2}_{y'y'}} & 0 \\ 0 & 0 & \widetilde{w}_{{m2}_{z'z'}} \end{array}\right)'}\end{equation}

\begin{displaymath}{=\widetilde{w}_{{m2}_{x'x'}}\left(\begin{array}{ccc} 1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 0 \end{array}\right)'+ \widetilde{w}_{{m2}_{y'y'}}\left(\begin{array}{ccc} 0 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0 \end{array}\right)' + \widetilde{w}_{{m2}_{z'z'}}\left(\begin{array}{ccc} 0 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 1 \end{array}\right)'}\end{displaymath}

\begin{equation}{=\widetilde{w}_{{m2}_{x'x'}}\hat{x'}\hat{x'}+\widetilde{w}_{{m2}_{y'y'}}\hat{y'}\hat{y'}+\widetilde{w}_{{m2}_{z'z'}}\hat{z'}\hat{z'}}\end{equation}
written in terms of the outer products

\begin{displaymath}{\hat{x}'\hat{x}'= \left(\begin{array}{c} 1 \\ 0 \\ 0 \end{array}\right)'\left(\begin{array}{ccc} 1 & 0 & 0 \end{array}\right)'=\left(\begin{array}{ccc} 1 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 0 \end{array}\right)'}\end{displaymath}

\begin{displaymath}{\hat{y}'\hat{y}'= \left(\begin{array}{c} 0 \\ 1 \\ 0 \end{array}\right)'\left(\begin{array}{ccc} 0 & 1 & 0 \end{array}\right)'=\left(\begin{array}{ccc} 0 & 0 & 0 \\ 0 & 1 & 0 \\ 0 & 0 & 0 \end{array}\right)'}\end{displaymath}

\begin{displaymath}{\hat{z}'\hat{z}'= \left(\begin{array}{c} 0 \\ 0 \\ 1 \end{array}\right)'\left(\begin{array}{ccc} 0 & 0 & 1 \end{array}\right)'=\left(\begin{array}{ccc} 0 & 0 & 0 \\ 0 & 0 & 0 \\ 0 & 0 & 1 \end{array}\right)'}\end{displaymath}

%\newline $k_{x'}= k_{x'_x} \hat{x} + k_{x'_y} \hat{y} + k_{x'_z} \hat{z}$
%\newline $k_{y'}= k_{y'_x} \hat{x} + k_{y'_y} \hat{y} + k_{y'_z} \hat{z}$
%\newline $k_{z'}= k_{z'_x} \hat{x} + k_{z'_y} \hat{y} + k_{z'_z} \hat{z}$
%\newline \begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(\vec{k})=\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{displaymath}
\noindent Using the relation
%\begin{displaymath}{\hat{k_{x'}}\hat{k_{x'}} + \hat{k_{y'}}\hat{k_{y'}} + \hat{k_{z'}}\hat{k_{z'}} = 1}\end{displaymath} 
\begin{displaymath}{\text{I} = \hat{x'}\hat{x'} + \hat{y'}\hat{y'} + \hat{z'}\hat{z'}}\end{displaymath}
\begin{displaymath}{\text{I} -\hat{z'}\hat{z'} = \hat{x'}\hat{x'} + \hat{y'}\hat{y'}}\end{displaymath}
and the results from earlier
\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}=\widetilde{w}_{{m2}_{x'x'}}}\end{equation}
\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}=-2\widetilde{w}_{{m2}_{x'x'}}}\end{equation}
Equation \ref{tensorcomp} then becomes 
\begin{equation}{\widetilde{w}_{m2}(\vec{k})= (\text{I}-3\hat{z}'\hat{z}')\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
\noindent In the $(x',y',z')$ coordinate system, vector $\vec{k}$ had been aligned with the $\hat{z}'$ direction so that 
\begin{displaymath}{\hat{k}=\hat{z'}}\end{displaymath}
Replacing $\hat{z}'$ with $\hat{k}$ gives
\begin{equation}{\widetilde{w}_{m2}(\vec{k})= (\text{I}-3\hat{k}\hat{k})\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
Now  $\vec{k}$ can be expressed in the $(x,y,z)$ coordinate system, $\vec{k}=k_x\hat{x} + k_y\hat{y} + k_z\hat{z}$. The components of the tensor weight in the $(x,y,z)$ coordinate system are then given by
\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})= (\delta{ij}-3\frac{k_ik_j}{k^2})\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
\[{}\]
%\subsection{Transform of tensor weight ${w}_{m2}(r)$ goes to zero as k goes to zero}
\noindent Check that the transform of tensor weight ${w}_{m2}(r)$ goes to zero as k goes to zero:
\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{k^2\Xi^2}{8}\right)}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\sin(\frac{k\alpha}{2})}\end{displaymath} 
\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
\[{}\]
Term 1: 
%\color{green}
\begin{equation}{\left(\frac{-\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{k^2\Xi^2}{8}\right)}\cos(\frac{k\alpha}{2})}\end{equation}
\begin{displaymath}{\approx\left(-\frac{\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)\left(1-\frac{k^2\Xi^2}{8}\right)\left(1-\left(\frac{1}{2}\right)\frac{k^2\alpha^2}{4}\right)}\end{displaymath} 
\begin{displaymath}{\approx-\frac{4\pi}{k^2}+\frac{\pi\alpha^2}{2}+\frac{\pi\Xi^2}{6}}\end{displaymath} 
\color{black}
Term 2:
%\color{blue}
\begin{equation}{-\frac{2\pi\alpha}{3k}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\sin(\frac{k\alpha}{2})}\end{equation} 
\begin{displaymath}{\approx\left(-\frac{2\pi\alpha}{3k}\right)\left(1-\frac{k^2\Xi^2}{8}\right)\left(\frac{k\alpha}{2}\right)}\end{displaymath} 
\begin{displaymath}{\approx}-\frac{\pi\alpha^2}{3}\end{displaymath}
\color{black} 
Term 3:
%\color{red}
\begin{equation}{\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
\begin{displaymath}{\approx\frac{\left(2\pi\right)^\frac{2}{3}}{\Xi{k}^3}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[2\operatorname{Re}\left(\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)\right)\right]}\end{displaymath} 
\begin{displaymath}{\approx\frac{\left(2\pi\right)^\frac{2}{3}}{\Xi{k}^3}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[2e^{\left(\frac{\alpha^2}{2\Xi^2}\right)}\frac{2}{\sqrt{\pi}}   \left(\frac{k\Xi}{2\sqrt{2}}   -\frac{2\left(\frac{\alpha^2}{2\Xi^2}\right)+1}{3}\left(\frac{k\Xi}{2\sqrt{2}}\right)^3\right)\right]}\end{displaymath}
\begin{displaymath}{\approx\frac{4\pi}{k^2}-\frac{\pi}{6}\left(\alpha^2-\Xi^2\right)}\end{displaymath}  
\color{black}
Adding the three terms gives:
\begin{displaymath}{\approx-\frac{4\pi}{k^2}+\frac{\pi\alpha^2}{2}+\frac{\pi\Xi^2}{6}-\frac{\pi\alpha^2}{3}+\frac{4\pi}{k^2}-\frac{\pi\alpha^2}{6}-\frac{\pi\Xi^2}{6}}\end{displaymath} 
\begin{displaymath}{\approx0}\end{displaymath} 
\[{}\]
%\subsection{}
\textbf{4. Integrals used to Compute Fourier Transforms of Weight Functions}
\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
%\tiny
\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation} 
\[{}\]
\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}+\frac{4\sin(kr)}{k^3r^3}}\end{equation} 
\[{}\]
\[{}\]
\begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}}}\end{equation}
\begin{displaymath}{=a\int_{-\infty}^{\infty}{e^{-u^2}\left[\cos(kau)\cos(\frac{k\alpha}{2})-\sin(kau)\sin(\frac{k\alpha}{2})\right]d{u}}}\end{displaymath}  
\begin{equation}{=a\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})}\end{equation} 
\[{}\]
\[{}\]
\begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r}}}\end{equation}
\begin{equation}{=\frac{a\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\int_{-\infty}^{\infty}{e^{-u^2}\cos(kau)d{u}} -a^2\sin\left(\frac{k\alpha}{2}\right)\int_{-\infty}^{\infty}{e^{-u^2}u\sin(kau)d{u}}}\end{equation}
\begin{equation}{=\frac{a\alpha\sqrt{\pi}}{2}e^{-\left(\frac{ka}{2}\right)^2}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\left(\frac{ka}{2}\right)^2}\sin\left(\frac{k\alpha}{2}\right)}\end{equation}
\[{}\]
\begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}}}\end{equation} 
\begin{displaymath}{=a\int_{-\infty}^{\infty}{e^{-u^2}\left[\sin(kau)\cos(\frac{k\alpha}{2})+\cos(kau)\sin(\frac{k\alpha}{2})\right]d{u}}}\end{displaymath}  
\begin{equation}{=a\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{equation} 
\[{}\]
\[{}\]
\begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}}}\end{equation} 
\begin{displaymath}{=a\int_{-\infty}^{\infty}{e^{-u^2}(au+\frac{\alpha}{2})\sin(k(au+\frac{\alpha}{2}))d{u}}}\end{displaymath} 
\begin{displaymath}{=a^2\cos(\frac{k\alpha}{2})\int_{-\infty}^{\infty}{e^{-u^2}u\sin(kau)d{u} +\frac{a\alpha}{2}\sin(\frac{k\alpha}{2})\int_{-\infty}^{\infty}e^{-u^2}\cos(kau)d{u}}}\end{displaymath} 
\begin{equation}{=\frac{ka^3}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})+\frac{a\alpha}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{equation} 
%\[{}\]
\[{}\]
\begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}}}\end{equation} 
%Noting that \begin{equation}{\frac{\sin(kr)}{r}=\int_{0}^{k}{\cos(kr)d{k}}}\end{equation} 
%this becomes
\begin{displaymath}{=\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{0}^{k}{\cos(kr)d{k}}\right)d{r}}}\end{displaymath} 
\begin{displaymath}{=\int_{0}^{k}\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}{\cos(kr)d{r}}d{k}}}\end{displaymath} 
\begin{displaymath}{=a\sqrt{\pi}\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2}\cos\left(\frac{k\alpha}{2}\right)d{k} }\end{displaymath} 
%\begin{displaymath}{=a\sqrt{\pi}\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2}\left(\frac{e^{i\frac{k\alpha}{2}}+e^{-i\frac{k\alpha}{2}}}{2}\right)d{k} }\end{displaymath} 
%\begin{displaymath}{=\frac{a\sqrt{\pi}}{2}\left[\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2 + i\frac{k\alpha}{2}}d{k}+\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2-i\frac{k\alpha}{2}}d{k}\right]}\end{displaymath} 
\begin{displaymath}{=\frac{a\sqrt{\pi}}{2}\left[\int_{-k}^{k}e^{-\left(\frac{ka}{2}\right)^2 + i\frac{k\alpha}{2}}d{k}\right]}\end{displaymath} 
%\begin{displaymath}{=\frac{a\sqrt{\pi}}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\int_{-k}^{k}e^{-\left(\frac{ka}{2} + i\frac{\alpha}{2a}\right)^2}d{k}\right]}\end{displaymath} 
\begin{displaymath}{=\sqrt{\pi}e^{-\left(\frac{\alpha}{2a}\right)^2}\int_{-\frac{ka}{2}+i\frac{\alpha}{2a}}^{\frac{ka}{2}+i\frac{\alpha}{2a}}e^{-u^2}d{u}}\end{displaymath} 
\begin{equation}{=\frac{\pi}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation} 
%\[{}\]
\[{}\]
\[{}\]
\[{}\]
\textbf{5. Deriviation of Tensor Density $n_{m2ij}(\vec{r})$}
\begin{equation}{\overleftrightarrow{n}_{m2}(\vec{r})=\int_{allspace}n(\vec{r'})\overleftrightarrow{w}_{m2}(\vec{r}-\vec{r'})d{\vec{r'}}}\end{equation} 
\begin{equation}{n_{m2ij}(\vec{r})=\int_{allspace}{n(\vec{r}')w_{m2ij}(\vec{r}-\vec{r}'){}d{\vec{r}'}}}\end{equation}
Using the definintion of the delta function
\begin{equation}{\delta}((\vec{r}-\vec{r}')-\vec{r}'')={ \frac{1}{\left(2\pi\right)^3}\int e^{i\vec k\cdot ((\vec r-\vec r')-\vec{r}'')}d\vec{k}}\end{equation} 
$w_{m2ij}(\vec{r}-\vec{r}')$ can be expressed as
\begin{align}
    w_{m2ij}(\vec{r}-\vec{r}') &= \int{\delta((\vec{r}-\vec{r}')-\vec{r}'')w_{m2ij}(\vec{r}''){}d{\vec{r}''}} \\
    &= \int{\left(\frac{1}{\left(2\pi\right)^3}\int e^{i\vec k\cdot((\vec r-\vec r')-\vec{r}'')}d\vec{k}\right){~}w_{m2ij}(\vec{r}''){}d{\vec{r}''}} \\
    &= \frac{1}{\left(2\pi\right)^3}\int{\left(\int w_{m2ij}(\vec{r}'')e^{-i\vec k\cdot\vec{r}''}d\vec{r}''\right)e^{i\vec k\cdot(\vec r-\vec r')}{~}{}d{\vec{k}}} \\
    &= \frac{1}{\left(2\pi\right)^3}\int{w_{m2ij}(\vec k)e^{i\vec k\cdot(\vec r-\vec r')}{~}{}d{\vec{k}}} 
  \end{align} 
Putting this into the expression for $n_{m2ij}(\vec{r})$ gives 
\begin{align}
%    n_{m2ij}(\vec r) &= \int n(\vec r') w_{m2ij}(\vec r- \vec r')d\vec r' \\
%    w_{m2ij}(\vec r) &= \frac1{2\pi}\int d\vec k w_{m2ij}(\vec k)e^{i\vec k\cdot \vec r}\\
%    n_{m2ij}(\vec{r})=\int_{allspace{n(\vec{r}')w_{m2ij}(\vec{r}-\vec{r}'){}d{\vec{r}'}}\\
    n_{m2ij}(\vec r) &= \int n(\vec r') \left(\frac{1}{\left(2\pi\right)^3}\int w_{m2ij}(\vec k)e^{i\vec k\cdot (\vec r-\vec r')}d\vec{k}\right)d\vec r' \\
    &= \int d\vec r' \left(\frac{1}{\left(2\pi\right)^3}\int d\vec k' n(\vec k')e^{i\vec k'\cdot \vec r'}\right) \left(\frac{1}{\left(2\pi\right)^3}\int d\vec k w_{m2ij}(\vec k)e^{i\vec k\cdot (\vec r-\vec r')}\right) \\
    &=  \frac{1}{\left(2\pi\right)^3}\int d\vec k' n(\vec k') \int d\vec k w_{m2ij}(\vec k)
    e^{i\vec k\cdot \vec r}\left(\frac{1}{\left(2\pi\right)^3}\int d\vec r'e^{i(\vec k'-\vec k)\cdot \vec r'}\right)
    \\
   &= \frac{1}{\left(2\pi\right)^3}\int d\vec k' n(\vec k') \int d\vec k w_{m2ij}(\vec k)e^{i\vec k\cdot \vec r}\delta(\vec k'-\vec k) \\
    &= \frac{1}{\left(2\pi\right)^3}\int d\vec k \left(\int d\vec k' n(\vec k')\delta(\vec k'-\vec k)\right) w_{m2ij}(\vec k)e^{i\vec k\cdot \vec r}
    \\
    &= \frac{1}{\left(2\pi\right)^3}\int d\vec k\, n(\vec k) w_{m2ij}(\vec k)e^{i\vec k\cdot \vec r}
  \end{align} 
\[{}\]
\[{}\]
%\[{}\]
\textbf{6. Deriviation of $\Phi_3$}
\[{}\]
\begin{equation}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left[\overrightarrow{\operatorname{V}}\cdot\overleftrightarrow{\operatorname{T}}\cdot\overrightarrow{\operatorname{V}}-n\overrightarrow{\operatorname{V}}\cdot\overrightarrow{\operatorname{V}}-\operatorname{Tr}\left(\overleftrightarrow{\operatorname{T}}^3\right)+n\operatorname{Tr}\left(\overleftrightarrow{\operatorname{T}}^2\right)\right]}\end{equation}
\[{}\]
\begin{equation}{\overleftrightarrow{\operatorname{T}}=\frac{{\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}}{4\pi{R}^2}}\end{equation}
\[{}\]
\begin{equation}{\overrightarrow{\operatorname{V}}=\frac{\vec{n}_2}{4\pi{R}^2}}\end{equation}
\[{}\]
\begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}
\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
\[{}\]
\begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2\vec{n}_{v2}\cdot\vec{n}_{v2}-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left((\overleftrightarrow{n}_{m2}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
\[{}\]
\begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2|\vec{n}_{v2}|^2-\operatorname{Tr}({\overleftrightarrow{n}_{m2}^3})
-n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)\color{white}\right)\color{black}}\end{displaymath} \begin{equation}{\color{white}\left(\color{black}-\frac{1}{3}{n_2}^2\operatorname{Tr}(\overleftrightarrow{n}_{m2})-\frac{1}{9}{n_2}^3+n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)+\frac{1}{3}{n_2}^3+\frac{2}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation} 
\[{}\]
\begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation} 
\[{}\]
\begin{equation}{\operatorname{Tr}(\overleftrightarrow{n}_{m2})=0}\end{equation} 
\[{}\]
\begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2|\vec{n}_{v2}|^2+\frac{9}{2}(\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}})-\frac{9}{2}\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})}{24\pi(1-\eta)^2}}\end{equation} 
\[{}\]
\begin{equation}{\eta=n_3}\end{equation} 
\[{}\]
\begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}}\end{equation} 
\[{}\]
\[{}\]

%\textbf{7. Equations related to the Error Function Potential $V_{erf}$ ....:}
%\[{}\]
%\[{}\]

\textbf{8. Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$:}
\begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}
The mayer function for the Error Function potential $V_{erf}$ is
\begin{equation}f(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation} from which it can be seen that $f(r)$ is approximately 0 for $r>r_{max}$ where $r_{max}$ is some maximum value of $r$.
\begin{align}
 B_2 &= -\frac{1}{2}\int_0^{\pi}\int_0^{2\pi}\int_0^\infty\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2\sin{\theta}d{\theta}d{\phi}dr \\
     &= -\frac{1}{2}4\pi\frac{1}{2}\int_{0}^{r_{max}}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2dr \\
     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \pi\int_0^{r_{max}}r^2dr   \\
     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \frac{\pi}{3}r_{max}^3   
\end{align}

\begin{align}
  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}r_{max}^3\right)\operatorname{erf}\left(\frac{\alpha-r_{max}}{\Xi}\right) \\
      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha r_{max}-\Xi^3-\Xi r_{max}^2\right)e^{-{\left(\frac{\alpha-r_{max}}{\Xi}\right)^2}} \\
      &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}r_{max}^3 
\end{align}
Letting $r_{max}\rightarrow\infty$ this becomes
\begin{align}
  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}\left(\infty\right)^3\right)\left(-1\right) \\
      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha\left(\infty\right)-\Xi^3-\Xi r_{max}^2\right)\left(0\right) \\
      &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}\left(\infty\right)^3 
\end{align}
\begin{equation}B_2 = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]\end{equation}

\[{}\]
\textbf{9. Third Virial Coefficient $B_{3}$:}

\begin{equation}\label{B3}B_3=-\frac{1}{3}\int_{allspace}\int_{allspace}f(\vec{r})f(\vec{r'})f(\vec{r}-\vec{r'})d\vec rd\vec r'\end{equation}

First find
\begin{equation}f(\vec{r}-\vec{r'})=\int_{allspace}f(\vec{r''})\delta(\vec{r}''-(\vec{r}-\vec{r}'))d\vec r'' \end{equation}

where
\begin{equation}{\delta}(\vec{r}''-(\vec{r}-\vec{r}'))={ \frac{1}{\left(2\pi\right)^3}\int_{allspace} e^{i\vec k\cdot (\vec{r}''-(\vec r-\vec r'))}d\vec{k}}\end{equation}

\begin{equation}f(\vec{r}-\vec{r'})=\int{f(\vec{r''})\left( \frac{1}{\left(2\pi\right)^3}\int e^{i\vec k\cdot (\vec{r}''-(\vec r-\vec r'))}d\vec{k} \right) d\vec r''} \end{equation}

\begin{equation}f(\vec{r}-\vec{r'})=\int{ e^{-i\vec k\cdot (\vec r-\vec r')}\left(\frac{1}{\left(2\pi\right)^3}\int{f(\vec{r''}) e^{i\vec k\cdot (\vec{r}'')}d\vec{r''}} \right) d\vec k} \end{equation}

\begin{equation}f(\vec{r}-\vec{r'})=\int{ e^{-i\vec k\cdot \vec r}e^{i\vec k\cdot \vec r}\widetilde{f}(-\vec k) d\vec k} \end{equation}

Putting this result into Eq~\ref{B3} gives
\begin{equation}B_3=-\frac{1}{3}\int{\int{f(\vec{r})f(\vec{r'})\left(\int{ e^{-i\vec k\cdot \vec r}e^{i\vec k\cdot \vec r}\widetilde{f}(-\vec k) d\vec k}\right)d\vec rd\vec r'}}\end{equation}

\begin{equation}B_3=-\frac{1}{3}\int{\widetilde{f}(-\vec k)\left(\int{f(\vec{r})e^{-i\vec k\cdot \vec r}}d\vec r\int{f(\vec{r'})e^{i\vec k\cdot \vec r'} d\vec r'}\right)d\vec k}\end{equation}

\begin{equation}B_3=-\frac{1}{3}\int{\widetilde{f}(-\vec k)\widetilde{f}(\vec k)\widetilde{f}(-\vec k)d\vec k}\end{equation}

\begin{equation}B_3=-\frac{1}{3}\int_0^{2\pi}\int_0^{\pi}\int_0^{\infty}{|\widetilde{f}(\vec k)|^3d\vec k}\end{equation}

Since the Mayer function does not depend on $\phi$ and $\theta$, this reduced to
\begin{equation}B_3=-\frac{4\pi}{3}\int{|\widetilde{f}(\vec k)|^3k^2dk}\end{equation}

The Mayer function for the Error Function potential $V_{erf}$, and its Fourier Transform are
\begin{equation}f_{erf}(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation}

\begin{equation}\widetilde{f}_{erf}(k)=-\frac{4\pi}{k^3}e^{-\frac{k^2\Xi^2}{4}}\left[\left(1+\frac{k^2\Xi^2}{2}\right)\sin(k\alpha)-k\alpha\cos(k\alpha)\right]\end{equation}

The Mayer function for the WCA potential $V_{WCA}$, and its Fourier Transform are

%\begin{equation}f(r)=e^{-\frac{1}{k_BT}\left[4\epsilon\left(\left(\frac{\sigma}{r}\right)^{12}\left(-\frac{\sigma}{r}\right)^6\right)+\epsilon\right]}-1\end{equation}

\begin{equation}f_{WCA}(r)=\operatorname{exp}\left(-\frac{1}{k_BT}\left[4\epsilon\left(\sigma^{12}r^{-12}-\sigma^{6}r^{-6}\right)+\epsilon\right]\color{white}\frac{1}{1}\color{black}\right)-1\end{equation}

\begin{equation}\widetilde{f}_{WCA}(k)=\frac{4\pi}{k}\int_0^{\infty}{\left(e^{-\frac{1}{k_BT}\left[4\epsilon\left(\sigma^{12}r^{-12}-\sigma^{6}r^{-6}\right)+\epsilon\right]}-1\right) r\sin(kr)dr}\end{equation}

%\begin{equation}B_3=-\frac{4\pi}{3}\int{|\widetilde{f}(-\vec k)|^3k^2dk}\end{equation}

\[{}\]
\noindent \textbf{Derivation of the Excess Free Energy from Scaled Particle Theory}

The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R. 
\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
This result is obtained by considering the difference in the excess Helmholtz free energy when one atom is added to the system. It was shown by Widom to be related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ of a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
%Equation~\ref{widom} is called Widom's Insertion Formula. 
\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
%In a randomly positioned (I think it has to be homogeneous!)
\indent In a collection of hard spheres making up a hard sphere fluid, no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere. But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.

\begin{figure}[h!]
    \centering
    \includegraphics[height=3cm]{P_overlap.png}
    \caption{No energy is needed to place a sphere at a position where there is no overlap such as shown by the black dot at the center of the dotted region near the top. But it is not possible to place a sphere at a position where an overlap occurs.}
    \label{fig:p_overlap}
  \end{figure}

For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
\color{red}Thus, considering cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}

%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}

Using this result, scaled particle theory seeks a more general expression of the excess chemical potential of a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_o$ where the size of $R_o$ falls into three different cases: $R_o >> R$, $R_o > R$, and $R_o < R$.

\begin{equation}\label{3_parts_to_mu_ex}{\mu_{ex}(R_o)=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_o^3 & R_o >> R \\ P\frac{4\pi}{3}R_o^3 + c_2R_o^2 + c_1R_o^1 + c_0R_o^0  & R_o > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3)) & R_o < R\end{array}\right.}\end{equation}

%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  must tend toward in the limit that $R_o$ goes to infinity. By expanding the logrythmic term (ln(1-x)= ...) in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.

%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_o$, and so it is assumed that the middle function, which is valid for values of $R_o$ in between the two extremes, can also be expressed in terms of powers of $R_o$.  The coeficients $c_2$, $c_1$ and $c_0$ in the middle function can be determnined be considering that this function (valid for $R_o > R$) must be continuous with the last function (valid for $R_o < R$) along with their first and second derivatives. The middle function must also tend toward the first function in the limit that $R_o$ goes to infinity. The chemical potential then becomes
\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
where $\eta$ is the packing fraction given by
\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}

%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$. From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation 
%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
%From the expression for pressure given by

The pressure can be obtained from \color{red}
\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
%The excess Hemlholtz free energy is then found from the relation
and the Helmholz free energy from
\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
which gives
%\begin{displaymath}F_{ex}=\end{displaymath}
\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}

\end{document}

--------------------------------------------------------------------------------
--------------------------------------------------------------------------------
 \begin{figure}[h!]
    \centering
    \includegraphics[height=6cm]{SimpleCell_Gaus.png}
    \caption{A 3D Gaussian Distribution}
    \label{fig:SimpleCell_Gaus}
  \end{figure}

The number density profile is varied, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found, which is known from cDFT to be unique. This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure ?? .

\[{}\]
Program Implemtation
\[{}\]
A simplified flowchart of the program created to investigate the freezing of a WCA fluid is shown on pages x-x. The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.  Given values for n* and t*, the program varys gw and fv and compares the free energies of the different number density profiles to find the minimum free energy for that combination of n* and t*.

******startA
In the Monte-Carlo Integration vectors are generated randomly according to a Gaussian distribution, the weighted density is computed for each vector, and the average of all the results gives the final weighted density. .....
ABOUT R!!
\begin{displaymath}{n_i += \frac{(1-fv)w_i}{NumPoints}}\end{displaymath}

The result can be made more accurate by increasing NumPoints which is the number of vectors randomly generated. In order to further reduce error, antithetic variates are used. An atithetic variate is an equal, but opposite point at which a function is evaluated. The final result is then given by the average of the two results. This reduces first order error by filtering out....

\begin{displaymath}{n_i += \frac{(1-fv)\left[\frac{w_i(\vec{r}) +w_i(-\vec{r})}{2}\right]}{NumPoints}}\end{displaymath}

(...........Find weights $w_0(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$,  $w_1(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$, $w_2(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$, $w_3(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$ and compute weighted densities.........)
******endA

***startB  old - before corrected center of weight function 
For a vector $\vec{r}=\vec{r}_k$ within the primitive cell, a vector $\vec{R}_j$ to the center of one of the Gaussians is selected that is within the primitive cell, or one of the neighboring cells within a prudently picked inclusion radius as Gaussians farther away will not contriubute much. The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, $n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions $w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to the center of the weight function. Monte-Carlo integration works by randomly generating $\vec{dr}$ according to a Gaussian distribution.
***endB

\begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{SimpleCell_Gaus.png}
    \caption{A homogeneouse number density profile $n(\vec{r})$}
    \label{fig:homogen_denisty}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{W0.png}
    \caption{A crossectional view of weight function $w_2$ which is a spherical shell with soft edges in 3D. Weight functions $w_0$ and $w_1$ have a similar shape.}
    \label{fig:W0}
    \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4cm]{W3.png}
    \caption{A crossectional view of weight function $w_3$ which is a sphere with soft edges in 3D.}
    \label{fig:W3}
    \end{figure}

////////alpha and Xi vary with temp:
\begin{figure}[h!]
    \centering
    \includegraphics[height=4.7cm]{plot_alpha_vs_KbT.png}
    \caption{Shows how $\alpha$ varies with temperature}
    \label{fig:alpha_vs_T}
  \end{figure}

 \begin{figure}[h!]
    \centering
    \includegraphics[height=4.7cm]{plot_xi_vs_KbT.png}
    \caption{Shows how $\Xi$ varies with temperature}
    \label{fig:xi_vs_T}
  \end{figure}

/////weight equations
\begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} and the shifted weighting functions $w_0$, $w_1$, $w_2$, $w_3$ are given by:
\begin{equation}{w_0(|\vec{r}-\vec{r}'|)= \frac{w_2}{4\pi(|\vec{r}-\vec{r}'|)^2}}\end{equation} 
\begin{equation}{w_1(|\vec{r}-\vec{r}'|)= \frac{w_2}{4\pi(|\vec{r}-\vec{r}'|)}}\end{equation} 
\begin{equation}{\vec{w}_1(|\vec{r}-\vec{r}'|)= w_1\frac{\vec{r}-\vec{r}'}{|\vec{r}-\vec{r}'|}}\end{equation} 
\begin{equation}{}\end{equation} 
\begin{equation}{\vec{w}_2(|\vec{r}-\vec{r}'|)=w_2\frac{\vec{r}-\vec{r}'}{|\vec{r}-\vec{r}'|}}\end{equation}
\begin{equation}{w_3(|\vec{r}-\vec{r}'|)={\int_r}^{\infty}{w_2(|\vec{r}-\vec{r}'|)}{d}\vec{r}'}\end{equation} 
In this paper, w2 is given by ....
\begin{equation}{w_2(|\vec{r}-\vec{r}'|)=\frac{1}{\Xi\sqrt\pi}\exp^-\frac{(|\vec{r}-\vec{r}'|-\frac{\alpha}{2})^2}{\Xi^2}}\end{equation}
which gives:
\begin{equation}{w_3(|\vec{r}-\vec{r}'|)=\frac{1}{2}}\left(1-erf(\frac{(|\vec{r}-\vec{r}'|)-\frac{\alpha}{2}}{\Xi})\right)\end{equation} where the two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4).

%BEGIN SFMT--------------------------------BEGIN SFMT----------------------------------
%Unlike a standard weighted density approximation, 
Soft Fundamental Measure Theory (SFMT) is used to obtain an expression for $F_{excess}$ in terms of a set of functions $\Phi_1$, $\Phi_2$, and $\Phi_3$ 
\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation} \color{red} which in turn depend on a set of weighted densities $n_0, n_1, n_2, n_3$ that arise naturally in the expression for the excess free energy derived from scaled particle theory. %The weighted densities $n_0, n_1, n_2, n_3$ arise naturally in the expression for the excess free energy derived from scaled particle theory.

%SFMT uses a set of weighted densities $n_0, n_1, n_2, n_3$ that arise naturally in the expression for the excess free energy derived from scaled particle theory. 
 %There is no one single weight function, however, that can effectively describe the coarse density. Rather, a set of weights will do the job, each giving rise to its own "weighted density". 
%\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}

\begin{equation}{F_{excess}=k_BT\left(-n_0ln(1-n_3)+\frac{n_1n_2}{1-n_3}+\frac{n_2^3}{24\pi(1-n_3)^2}\right)V}\end{equation}
where \begin{equation}n_i=nw_i\end{equation}
\begin{equation}w_0=\frac{w_2}{4\pi{R^2}}=1\end{equation}
\begin{equation}w_1=\frac{w_2}{4\pi{R}}=R\end{equation}
\begin{equation}w_2=4\pi{R}^2\end{equation}
\begin{equation}w_3=\frac{4}{3}\pi{R}^3\end{equation}
\color{black}
For a spatially varying number density, $F_{excess}$ becomes a functional of the number density profile $n(\vec{r})$ :
%SFMT extends these equations to a spatially varying number density
\begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
\begin{equation}{F_{excess}[n(\vec{r})]=k_BT\int\left(-n_0ln(1-n_3)+\frac{n_1n_2-\vec{n_1}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{1-n_3}+  \frac{n_2^3\left(1-{(\frac{\vec{n_2}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{{n_2}^2})}\right)^3}{24\pi(1-n_3)^2 }\right){d\vec{r}}}\end{equation}
where \begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation}
%which reduces for a uniform number density to
\color{red}

%///////Compare this to what is involved with a calculation of the free energy using classical Density Functional Theory.//////

\begin{equation}{\Phi_1= -n_0\ln(1-n_3)}\end{equation} 
\begin{equation}{\Phi_2= \frac{n_1n_2-\vec{n_1}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{1-n_3}}\end{equation} 
\begin{equation}{\Phi_3=\frac{n_2^3\left(1-{(\frac{\vec{n_2}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{{n_2}^2})}\right)^3}{24\pi(1-n_3)^2}}\end{equation} %where the weighted densities $n_0$, $n_1$, $n_2$, $n_3$ are given by:

%\begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} and the weighting functions $w_0$, $w_1$, $w_2$, $w_3$ are given by:

\begin{equation}{w_0(r)= \frac{w_2(r)}{4\pi{r^2}}}\end{equation} 
\begin{equation}{w_1(r)= \frac{w_2(r)}{4\pi{r}}}\end{equation}
\begin{equation}{w_2(r)=-\frac{\partial{w_3(r)}}{\partial{r}}}\end{equation}  
\begin{equation}{\vec{w}_1(r)= w_1(r)\frac{\vec{r}}{|\vec{r}|}=\frac{w_2(r)}{4\pi{r}}\frac{\vec{r}}{|\vec{r}|}}\end{equation} 
\begin{equation}{\vec{w}_2(r)=w_2(r)\frac{\vec{r}}{|\vec{r}|}}\end{equation}

%\begin{equation}{w_3(r)={\int_r}^{\infty}{w_2(r')}{dr'}}\end{equation} 
\color{black}
In this paper, $w_3(r)$ is given by 
\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-erf\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
as proposed by Schmidt$^{(?)}$ where $w_3$ is subject to the boundary conditions $w_3(0)\approx{1}$ and $w_3(inf)=0$. These boundary conditions are satisfied if $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. Solving equation (??) for $w_2(r)$
%\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{infinity}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
\begin{equation}{\frac{df(r)}{dr}=\int_{-inf}^{inf}{w_2(r')w_2(r-r')dr'}}\end{equation}  Solving equation (??) for the Mayer function
\begin{displaymath}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{displaymath} \begin{displaymath}{f(r)=\int_{inf}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{displaymath} we get
\begin{equation}{f(r)=-\frac{1}{2}\left[1-erf\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} This gives an interatomic potential of \begin{displaymath}{V_{erf}=-k_BTln(f(r)-1)}\end{displaymath} \begin{equation}{V_{erf}=-k_BTln\left[\frac{1}{2}\left(erf\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} where the two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), and are determined to make $V_{erf}$ match $V_{WCA}$ as closely as possible.

%The SFMT equations and expressions for $w_2$, $\Xi$ and $\alpha$ are developed in prior work. It is not the aim of this paper to enter into the details of SFMT, but rather to take these given expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.

%END SFMT--------------------------------END SFMT----------------------------------

\color{red} %Virial Expansion
%\begin{equation}U=TS-PV+\mu{N}\end{equation}
%%%\begin{equation}U-\mu{N}=TS-PV\end{equation}
%\begin{equation}e^{-\beta(U-\mu{N})}=e^{-\beta(TS-PV)}\end{equation}
%%%\begin{equation}e^{-\beta(U-\mu{N})}=e^{-\frac{TS}{k_BT}+\frac{PV}{k_BT}}\end{equation}
%\begin{equation}e^{-\frac{(U-\mu{N})}{k_BT}}e^{\frac{S}{k_B}}=e^{\frac{PV}{k_BT}}%\end{equation}
%%\begin{equation}e^{-\beta(U-\mu{N})}e^{\ln\Omega}=e^{\frac{PV}{k_BT}}\end{equation}
%\begin{equation}e^{-\beta(U-\mu{N})}e^{-\sum{P_i}\ln{P_i}}=e^{\frac{PV}{k_BT}}\end{equation}
%%%\begin{equation}e^{-\beta(U-\mu{N})}\Omega=e^{\frac{PV}{k_BT}}\end{equation}
%%%\begin{equation}\ln(e^{-\beta(U-\mu{N})}\Omega)=\ln(e^{\frac{PV}{k_BT}})\end{equation}
%%\begin{equation}\ln(e^{-\beta(U-\mu{N})}\Omega)=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
%%%\begin{equation}\ln(Z_{Grand})+\Omega=\frac{PV}{k_BT}\end{equation}
%%\begin{equation}\ln((e^{\beta\mu})^Ne^{-\beta{U}}\Omega)=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
%\begin{equation}\ln((e^{\beta\mu})^Ne^{-\beta{U_N}-\sum{P_i}\ln{P_i}})=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
\color{black}

%*********************** DFT minimize F/V *********************
--------------------------------------------------------

Starting with Euler's equation
\begin{equation}U=TS-PV+\mu{N}\end{equation}
\begin{equation}\frac{U}{V}=\frac{TS}{V}-\frac{PV}{V}+\frac{\mu{N}}{V}\end{equation}
\begin{equation}\label{d(U/V)long}d\left(\frac{U}{V}\right)=\frac{T}{V}dS+\frac{S}{V}dT+TSd\left(\frac{1}{V}\right)-dP+\mu{dn}+nd\mu\end{equation}
and using the following relation from Euler's equation
\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
which gives
\begin{equation}\frac{S}{V}dT-dP+nd\mu=0\end{equation}
Equation~\ref{d(U/V)long} becomes
\begin{equation}\label{d(U/V)}d\left(\frac{U}{V}\right)=\frac{T}{V}dS+TSd\left(\frac{1}{V}\right)+\mu{dn}\end{equation}
%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
The Helmholtz free energy F is defined as 
\begin{equation}F=U-TS\end{equation}
\begin{equation}\frac{F}{V}=\frac{U}{V}-\frac{TS}{V}\end{equation}
\begin{equation}d\left(\frac{F}{V}\right)=d\left(\frac{U}{V}\right)-d\left(\frac{TS}{V}\right)\end{equation}
\begin{equation}d\left(\frac{F}{V}\right)=d\left(\frac{U}{V}\right)-\frac{T}{V}dS-\frac{S}{V}dT-TS d\left(\frac{1}{V}\right)\end{equation}
After plugging in Equation~\ref{d(U/V)}, and setting $f_V=\frac{F}{V}$ this becomes
\begin{equation}df_V=-\frac{S}{V}dT+\mu{dn}\end{equation}
This result shows that the natural variables of the Helmholtz free energy per volume $f_V$ are temperature and number density
\begin{equation}f_V(T,n)\end{equation}
It is the Helmholtz free energy per volume $f_V$ that will be minimized in a system that has a temperature and number density that are fixed.

------------------------------------------------------
%*************************END DFT minimize F/V ****************

-------------------------this blue/red part basically garbage------------------:
\color{blue}
A faster, and much easier way to find the excess free energy is to use classical Density Functional Theory (cDFT). As stated earlier, in cDFT the free energy is expressed as a functional of the number density $F[n(\vec{r})]$. 
%The ideal free energy in Eq~\ref{eq:Fideal} can be expressed in terms of a spatially varying number density $n(\vec{r})$.

%For a spatially varying number density, the ideal free energy in Eq~\ref{eq:Fideal} becomes:
%\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}

To form an expression for the excess free energy as a functional of the number density, Soft Fundamental Measure theory (SFMT) is used. The excess free energy of a fluid can be expressed in general as an integral over an excess free energy density $\frac{F}{V}$.
\begin{displaymath}{F_{excess}=\int{\frac{F}{V}{dV}}}\end{displaymath}
Rewriting this in terms of the excess free energy per particle $\phi=\frac{F}{N}$, or in terms of $\Phi=\frac{1}{k_BT}\frac{F}{N}\frac{N}{V}=\frac{\phi{n}}{k_BT}$ we get
%$\frac{F}{V}=\frac{F}{N}\frac{N}{V}=\phi{n}$  
\begin{displaymath}{F_{excess}=\int{\frac{F}{N}\frac{N}{V}{dV}}=\int{\phi{n}{dV}}=k_BT\int\Phi{dV}}\end{displaymath} Showing explicitly the dependence of $\phi$ on the density at each point this becomes
\begin{displaymath}{F_{excess}=\int{\phi(n(\vec{r}))n(\vec{r}){d\vec{r}}}=k_BT\int\Phi(n(\vec{r})){d\vec{r}}}\end{displaymath}
where $n(\vec{r})$ is a smoothly varying function. \color{red} However, there is no real smoothly varying function when looking closeup. The real density is coarse, being zero where there are no atoms, or spheres located in a volume, and having a maximum of one at the center of each sphere over a volume of one sphere. In other places, with various portions of a sphere within a volume of one sphere, it is somewhere in between. To account for, and describe this coarse density as a function of position, the smoothly varying density function is muliplied by a weight function which is centered about each point in space. The weighted density as a function of position then becomes
\begin{displaymath}{n_{weighted}(\vec{r})=\int{n(\vec{r})w(|\vec{r}-\vec{r}'|){d\vec{r}}}}\end{displaymath}
%In a weighted density approximation, the excess free energy per particle $\phi$ is expressed in terms of the weighted density \begin{displaymath}{F_{excess}\approx\int{\phi(n_{weighted}(\vec{r}))n(\vec{r}){d\vec{r}}}}\end{displaymath} or, setting $\Phi=\frac{\phi{n}}{k_BT}$ \begin{displaymath}{F_{excess}\approx{k_BT}\int{\Phi(\vec{r}){d\vec{r}}}}\end{displaymath}

%(CAREFUL- probability density!)
%(CAREFUL- weighted density is a weighted average!)
\color{black}
-------------------------------end basically garbage------------------

Another way to form an expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ is to use Soft Fundamental Measure theory (SFMT). SFMT extends Fundamental Measure Theory (FMT) developed for hard sphere fluids to those with soft potentials like that of the WCA fluid.

%In summary, it is the aim of this paper to use these expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.

In summary, classical Density Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT) provides a method by which the free energy can be found simply and directly without taking into account the interaction each sphere, or atom, has with every other sphere. For a number density n that is constant everywhere, as would be expected for a homogeneous fluid in equilbrium, the free energy can be written as a function of n where the ideal Helmholtz free energy can be determined analytically, and SFMT is used to form an expresssion for the excess Helmholtz free energy
\begin{equation}{F(n)=F_{ideal}(n) + F_{excess}(n)}\end{equation} 
For a spatially varying number density, as would be expected for a crystal, the free energy can be written as a functional of $n(\vec{r})$ where again SFMT is used to form the expression for the excess Helmholtz free energy.
\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
To solve for the crystal free energy $F[n(\vec{r})]$ using cDFT all that is needed are number density profiles, such as Gaussian profiles that depend only on $\vec{r}$. 
\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} The profiles are varied until the free energy is minimized. The minimum free energy and its corresponding number density profile are the equilibrium values for that temperature. 
%The complexity is greatly reduced making cDFT a powerful tool in modeling fluid behavior.

### 2020-09-01 papers/thesis-kirstie/thesis.tex
rename to thesis and make it draft mode
comment out graphics from thesis for now

--- 2020-09-01 papers/thesis-kirstie/MyThesisinLaTeX.tex
@@
     \centering
-    \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
-    %\includegraphics[height=3.5cm]{plot_HardSphere_Potential.png}
-    %\caption{Hard Sphere Potential}
+    % FIXME \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
     \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
@@
     \centering
-    \includegraphics[height=7cm]{LJ_Potential.png}
+    % FIXME \includegraphics[height=7cm]{LJ_Potential.png}
     \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    \includegraphics[height=7cm]{LJ_Potential_2part.png}
+    % FIXME \includegraphics[height=7cm]{LJ_Potential_2part.png}
     \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    \includegraphics[height=7cm]{WCA_Potential.png}
+    % FIXME \includegraphics[height=7cm]{WCA_Potential.png}
     \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
@@
    % \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=3.5cm]{TwoSpheres.png}
+      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.png}
     %\end{subfigure} 
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=4.5cm]{plot_alpha_vs_KbT.png}
+      % FIXME \includegraphics[height=4.5cm]{plot_alpha_vs_KbT.png}
       %\includegraphics[height=5cm]{plot_alpha_vs_KbT.png}
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=4.5cm]{plot_xi_vs_KbT.png}
+      % FIXME \includegraphics[height=4.5cm]{plot_xi_vs_KbT.png}
       %\includegraphics[height=5cm]{plot_xi_vs_KbT.png}
@@
     \centering
-    \includegraphics[height=7cm]{Verf.png}
+    % FIXME \includegraphics[height=7cm]{Verf.png}
     %\includegraphics[height=5cm]{Verf.png}
@@
     \centering
-    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
+    % FIXME \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
     \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
@@
     \centering
-    \includegraphics[height=5cm]{P-T_Diagram.png}
+    % FIXME \includegraphics[height=5cm]{P-T_Diagram.png}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
     \centering
-    \includegraphics[height=7cm]{MDTC.png}
+    % FIXME \includegraphics[height=7cm]{MDTC.png}
     \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
@@
     \centering
-    \includegraphics[height=7cm]{MDTC-2.png}
+    % FIXME \includegraphics[height=7cm]{MDTC-2.png}
     \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
@@
     \centering
-    \includegraphics[height=7cm]{MDTC-3.png}
+    % FIXME \includegraphics[height=7cm]{MDTC-3.png}
     \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
@@
     \centering
-    \includegraphics[height=4cm]{T-n_Diagram.png}
+    % FIXME \includegraphics[height=4cm]{T-n_Diagram.png}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion.
@@
     \centering
-    \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
+    % FIXME \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
     \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
@@
     \centering
-    \includegraphics[height=4.5cm]{Simple_Cell.png}
+    % FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
     \caption{A Simplified Crystal Cell}
@@
     \centering
-    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
+    % FIXME \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
     \caption{A Gaussian Distribution Number Density Profile}
@@
     \centering
-    \includegraphics[height=6cm]{Gaussian_Plot.png}
+    % FIXME \includegraphics[height=6cm]{Gaussian_Plot.png}
     \caption{A 2D Gaussian Distribution of width 2$\sigma$}
@@
     \centering
-    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
+    % FIXME \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
     \caption{A Gaussian Distribution Number Density Profile}
@@
     \centering
-    \includegraphics[height=2.5cm]{SameStatPic.png}
+    % FIXME \includegraphics[height=2.5cm]{SameStatPic.png}
     \caption{Front views of a cube with eight smaller cells.
@@
     \centering
-    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
+    % FIXME \includegraphics[height=8cm]{VaryWidthandVacancies.png}
     \caption{Several Gaussian Distribution Number Density Profiles all with a number density of 1 atom per $V_{bold box}$}
@@
     %\includegraphics[height=4cm]{HomogeneousDen.png}
-    \includegraphics[height=4cm]{Homogeneous_bold-box.png}
+    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.png}
     \caption{A homogeneouse number density profile $n(\vec{r})$}
@@
     \centering
-    \includegraphics[height=5cm]{InclusionRadius.png}
+    % FIXME \includegraphics[height=5cm]{InclusionRadius.png}
     \caption{The Gaussians at each lattice point are of width gw. After a radius of few gw's the Gaussians go to zero and can be neglected.}
@@
     \centering
-    \includegraphics[height=4cm]{AntitheticVariate.png}
+    % FIXME \includegraphics[height=4cm]{AntitheticVariate.png}
     \caption{Antithetic variates reduce first order error in an average.
@@
     \centering
-    \includegraphics[height=6cm]{weight_functions.png}
+    % FIXME \includegraphics[height=6cm]{weight_functions.png}
     \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the z-axis for a reduced temperature of 2}
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W2.png}
+      % FIXME % FIXME \includegraphics[height=5cm]{W2.png}
       \caption{A crossectional view of $w_2$}
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W3.png}
+      % FIXME \includegraphics[height=5cm]{W3.png}
       \caption{A crossectional view of $w_3$}
@@
     \centering
-    \includegraphics[height=6cm]{GaussandW2_actual.png}
+    % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
     \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional view of $w_2$ centered at $\vec{r}'$ is shown in purple.
@@
     \centering
-    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
     \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  and fixed in place as $\vec{dr}$ varies randomly.}
@@
     \centering
-    \includegraphics[height=3cm]{P_overlap.png}
+    % FIXME \includegraphics[height=3cm]{P_overlap.png}
     \caption{No energy is needed to place a sphere at a position where there is no overlap such as shown by the black dot at the center of the dotted region near the top.
@@
     \centering
-    \includegraphics[height=6cm]{SimpleCell_Gaus.png}
+    % FIXME \includegraphics[height=6cm]{SimpleCell_Gaus.png}
     \caption{A 3D Gaussian Distribution}
@@
     \centering
-    \includegraphics[height=4cm]{SimpleCell_Gaus.png}
+    % FIXME \includegraphics[height=4cm]{SimpleCell_Gaus.png}
     \caption{A homogeneouse number density profile $n(\vec{r})$}
@@
     \centering
-    \includegraphics[height=4cm]{W0.png}
+    % FIXME \includegraphics[height=4cm]{W0.png}
     \caption{A crossectional view of weight function $w_2$ which is a spherical shell with soft edges in 3D. Weight functions $w_0$ and $w_1$ have a similar shape.}
@@
     \centering
-    \includegraphics[height=4cm]{W3.png}
+    % FIXME \includegraphics[height=4cm]{W3.png}
     \caption{A crossectional view of weight function $w_3$ which is a sphere with soft edges in 3D.}
@@
     \centering
-    \includegraphics[height=4.7cm]{plot_alpha_vs_KbT.png}
+    % FIXME \includegraphics[height=4.7cm]{plot_alpha_vs_KbT.png}
     \caption{Shows how $\alpha$ varies with temperature}
@@
     \centering
-    \includegraphics[height=4.7cm]{plot_xi_vs_KbT.png}
+    % FIXME \includegraphics[height=4.7cm]{plot_xi_vs_KbT.png}
     \caption{Shows how $\Xi$ varies with temperature}

### 2020-09-01 papers/thesis-kirstie/figs/plot_WCA_Potential_fixed.py
make first four plots build with fac

@@ -8,8 +8,6 @@ import numpy as np
 import matplotlib.pyplot as plt
 import math
 
-plot1="plot_WCA_Potential_fixed.png"
-
 #Plot WCA Potential vs r 
 R=1/1.781797436  #for a sigma=1   DOESN"T WORK!!  graph wrong shape!
 epsilon=1
@@ -37,9 +35,9 @@ plt.xlabel('r/$\sigma$')
 plt.ylabel('$V(r)/\epsilon$')
 plt.title('WCA Potential')
 #plt.legend()
-plt.savefig(plot1)
+plt.savefig('WCA_Potential.png')
 
-plt.show()
+# plt.show()
 
 
 

### 2020-09-01 papers/thesis-kirstie/figs/plot_w3_w2_w1_w0.py
add weights figure to thesis

@@ -8,10 +8,6 @@ import numpy as np
 import matplotlib.pyplot as plt
 import math
 
-
-#plot1="plot_w0_w1_w2_w3_vs_KbT.png"
-plot2="plot_w0_w1_w2_w3_vs_rz_at_kT2newcolors.png"
-
 sigma=1
 epsilon=1
 PI=np.pi
@@ -67,6 +63,6 @@ plt.xlabel('rz/(alpha/2)')
 plt.ylabel('weight function')
 plt.title('Weight Functions at KbT=2')
 plt.legend()
-plt.savefig(plot2)
+plt.savefig('weight_functions.pdf')
 
-plt.show()
+# plt.show()

### 2020-09-03 papers/thesis-kirstie/thesis.tex
switch to standard build for papers
make first four plots build with fac
add weights figure to thesis
added more plots to thesis

@@
     \centering
-    % FIXME \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
+    \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
     \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{LJ_Potential.png}
+    \includegraphics[height=7cm]{LJ_Potential.png}
     \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{LJ_Potential_2part.png}
+    \includegraphics[height=7cm]{LJ_Potential_2part.png}
     \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{WCA_Potential.png}
+    \includegraphics[height=7cm]{WCA_Potential.png}
     \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
@@
 Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$.
-Figure \ref{fig:alphaXivsT} shows how $\alpha$ and $\Xi$ vary with temperature when $\sigma=1$ and $\epsilon=1$, and Figure \ref{fig:erf_potential} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.
+Figure \ref{fig:alphaXivsT} shows how $\alpha$ and $\Xi$ vary with temperature when $\sigma=1$ and $\epsilon=1$.
+% and Figure \ref{fig:erf_Potential} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.
 %As seen in Figure \ref{fig:erf_potential}, the erf potentials closely match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, and for r/$\sigma$ less than ???.
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=4.5cm]{plot_alpha_vs_KbT.png}
-      %\includegraphics[height=5cm]{plot_alpha_vs_KbT.png}
+      % FIXME \includegraphics[height=4.5cm]{plot_alpha.png}
+      \includegraphics[height=5cm]{plot_alpha.pdf}
       \caption{$\alpha$ versus $k_BT/\epsilon$}
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=4.5cm]{plot_xi_vs_KbT.png}
-      %\includegraphics[height=5cm]{plot_xi_vs_KbT.png}
+      % FIXME \includegraphics[height=4.5cm]{plot_xi.png}
+      \includegraphics[height=5cm]{plot_xi.pdf}
       \caption{$\Xi$ versus $k_BT/\epsilon$}
@@
 
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=7cm]{Verf.png}
-    %\includegraphics[height=5cm]{Verf.png}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
-    \label{fig:erf_potential}
-  \end{figure}
-
-\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
-\[\]
-\[\]
+%\begin{figure}[h!]
+%    \centering
+    %\includegraphics[height=7cm]{erf_Potential.pdf}
+%    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
+%    \label{fig:erf_Potential}
+%  \end{figure}
+
+%\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
 
 Another way to find values for $\Xi(T)$ is to find computationally the value $\Xi$ for which the second virial coefficient evaluated for the error function potential $V_{erf}(r)$ is equal to the value for the second virial coefficent evalued for the WCA potential $V_{WCA}$, with $\alpha(T)$ plugged into the equation for $B_{2erf}$ at each temperature T.
-A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:XifromB2vsT}.
 \begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
@@
 \begin{equation}B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} \end{equation}
-
-\[\]
-\[\]
-\[\]
-
-\begin{figure}[h!]
-    \centering
-    %\includegraphics[height=7cm]{Xi_from_B2.png}
+A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{xi_fromB2.pdf}
     \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
-    \label{fig:XifromB2vsT}
+    \label{fig:xi_fromB2vsT}
   \end{figure}
 \color{red}
-\noindent !ADD PLOT of Verf !!!
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{erf_Potential_xifromB2.pdf}
+    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
+    \label{fig:erf_Potential_xifromB2}
+  \end{figure}
+
 \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
@@
     \centering
-    % FIXME \includegraphics[height=6cm]{Gaussian_Plot.png}
+    \includegraphics[height=6cm]{Gaussian}
     \caption{A 2D Gaussian Distribution of width 2$\sigma$}
@@
     \centering
-    % FIXME \includegraphics[height=6cm]{weight_functions.png}
+    \includegraphics[height=6cm]{weight_functions.pdf}
     \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the z-axis for a reduced temperature of 2}
@@
 
+\cite{fake}
+
+\bibliography{thesis}
+\bibliographystyle{plain}
+
 \end{document}

### 2020-09-04 papers/thesis-kirstie/thesis.tex
updated more figure references
updated another figure reference
modified discussion on B2 and moved stuff to appendix

@@
     \centering
-    \includegraphics[height=7cm]{plot_HardSphere_Potential.png}
+    \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
     \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
@@
     \centering
-    \includegraphics[height=7cm]{LJ_Potential.png}
+    \includegraphics[height=7cm]{LJ_Potential.pdf}
     \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    \includegraphics[height=7cm]{LJ_Potential_2part.png}
+    \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
     \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-    \includegraphics[height=7cm]{WCA_Potential.png}
+    \includegraphics[height=7cm]{WCA_Potential.pdf}
     \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
@@
     \centering
-   % \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.png}
-    %\end{subfigure} 
-    %\begin{subfigure}[b]{0.4\linewidth}
-    %  \includegraphics[height=3.5cm]{2spheres.png}
-    %\end{subfigure} 
+      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
     \caption{(Left) The WCA potential is a minimum at r=2R. 
@@
     \end{figure}
-%\[\]
-
-The WCA potential is not well suited for implementing computationally.
+
+The WCA potential is not well suited for implementing Density Functional Theory.
 Therefore, an erf potential which closely approximates the WCA potential at temperatures of interest will be used instead of the WCA potential.
-The erf potential is given by: \begin{equation}{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} and is given by \begin{equation}{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}\end{equation}  The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} and is given by: \begin{equation}{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}\end{equation}
+The erf potential is given by
+\begin{equation}
+	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
+\end{equation} 
+where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ 
+\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
+and is given by 
+\begin{equation}\label{alphaT}
+	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
+\end{equation} 
+The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:
+\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} 
+and is given by: 
+\begin{equation}
+	{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}
+\end{equation}
 
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=4.5cm]{plot_alpha.png}
       \includegraphics[height=5cm]{plot_alpha.pdf}
@@
     \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=4.5cm]{plot_xi.png}
       \includegraphics[height=5cm]{plot_xi.pdf}
@@
 
-Another way to find values for $\Xi(T)$ is to find computationally the value $\Xi$ for which the second virial coefficient evaluated for the error function potential $V_{erf}(r)$ is equal to the value for the second virial coefficent evalued for the WCA potential $V_{WCA}$, with $\alpha(T)$ plugged into the equation for $B_{2erf}$ at each temperature T.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
-\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
-\begin{equation}B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]\end{equation}
-and 
-\begin{equation}B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} \end{equation}
-A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
-    \label{fig:xi_fromB2vsT}
-  \end{figure}
-\color{red}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
-    \label{fig:erf_Potential_xifromB2}
-  \end{figure}
-
-\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
-\color{black}
-%\[\]
-%\[\]
-
-%----------------BEGIN Virial Expansion--------------
-The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
-\begin{equation}PV=TS-(U-\mu{N})\end{equation}
-\begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
-Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
+Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
+\begin{equation}\label{virialequation}
+    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
+\end{equation}
+where 
+\begin{equation}
+	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
+\end{equation}
+The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
+\begin{equation}U=TS-PV+\mu{N}\end{equation}
+using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
 \begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
@@
 \begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
-%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
-Equation~\ref{Euler_rearranged} then becomes
-\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\sum{P_i\beta(E_i-\mu
-{n_i}})\end{equation}
-\begin{equation}\frac{PV}{k_BT}=-\sum{P_i}\left[\ln{P_i}+\beta(E_i-\mu
-{n_i})\right]\end{equation}
-For a Grand Canonical Ensemble the probablity of a state given by
-\begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
-results in 
-\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[\ln{\left(\frac{ e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}\left(\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}} \left[-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\right)\end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=\left[\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\left(\frac{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right) \end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)} \end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \left(e^{\beta\mu}\right)^n\sum_{s(n)}e^{-\beta{E}_{s(n)}}\right)} \end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_n\right)} \end{equation}
-
-\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_{on}Q_n\right)} \end{equation}
-where $Q_n$ is the configuration integral.
-\color{red}For a system of N particles only, the sum collapses to one term, that for n=N, which gives \color{black}
-\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
-%=\ln\lambda^n + \ln{Z}_{on}+\ln{Q}_n
-=\beta\mu{N} + \ln{Z}_{oN} +\ln{Q}_N\end{equation}
-\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN} + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
-Noting that $Q_N=1$ for an ideal gas since $V(r)=0$ for an ideal gas
-\begin{equation}\frac{P_{ideal}V}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN}=N\end{equation}
-%\begin{equation}P_{ideal}=\mu{N} + k_BT\ln{Z}_{oN}\end{equation}
-\begin{equation}\frac{P_{excess}}{k_BT}= \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
-and Equation~\ref{PV/kBT_reduces_for_N} becomes
-\begin{equation}\frac{PV}{k_BT}=N + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
+which gives
 \begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Plugging in the expression for $F_{excess}$ in terms of the mayer function for low density fluids
-\begin{equation}F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
-gives
-\begin{equation}\frac{PV}{k_BT}=N + \left(-\frac{N}{V}\int{f(r)d\vec{r}}\right){N} + \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
-%\begin{equation}\frac{PV}{k_BT}=N - \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
-\begin{equation}\frac{PV}{k_BT}=N\left(1-\frac{N}{V}\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
-\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)\right) \end{equation}
-where $B_2(T)$ is given by
-\begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(r)d\vec{r} \end{equation}
-
-\color{red}This result was for low densities where higher terms in the expression for the excess free energy were neglected. The full virial equation is given by\color{black}
-\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}
-
-%-----------------END Virial Expansion---------------
-%\[\]
+Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.
+
+The value of $\Xi$ at a given temperature, T, can be found computationally by equating the second virial coefficient evaluated for the error function potential $V_{erf}(r)$, and the second virial coefficent evalued for the WCA potential $V_{WCA}$.
+\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+Evaluating $B_{2}$ for the erf potential gives
+%\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
+\begin{equation}
+	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
+\end{equation}
+where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, $B_{2WCA}$ can be found computationally from 
+\begin{equation}
+	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
+\end{equation}
+
+A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at a few different temperatures alongside the WCA potential.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
+    \label{fig:xi_fromB2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{erf_Potential_xifromB2.pdf}
+    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
+    \label{fig:erf_Potential_xifromB2}
+  \end{figure}
+
+%\color{red}
+% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
+%\color{black}
 
@@
 Thus, the WCA fluid can be treated as a classical fluid.
-%\\
-
-%\[\]
-%\[\]
+
 \[\]
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{MDTC.png}
+     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
     \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{MDTC-2.png}
+     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
     \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
@@
     \centering
-    % FIXME \includegraphics[height=7cm]{MDTC-3.png}
+     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
     \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
@@
     \centering
-    %\includegraphics[height=4cm]{HomogeneousDen.png}
-    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.png}
+    %\includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
+    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
     \caption{A homogeneouse number density profile $n(\vec{r})$}
@@
 
+\noindent \textbf{Derivation of the Virial Equation}
+%----------------BEGIN Virial Expansion--------------
+The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
+\begin{equation}PV=TS-(U-\mu{N})\end{equation}
+\begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
+Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
+\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
+\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
+\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
+%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
+%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
+Equation~\ref{Euler_rearranged} then becomes
+\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\sum{P_i\beta(E_i-\mu
+{n_i}})\end{equation}
+\begin{equation}\frac{PV}{k_BT}=-\sum{P_i}\left[\ln{P_i}+\beta(E_i-\mu
+{n_i})\right]\end{equation}
+For a Grand Canonical Ensemble the probablity of a state given by
+\begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
+results in 
+\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[\ln{\left(\frac{ e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}\left(\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}} \left[-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\right)\end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=\left[\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\right]\left(\frac{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\right) \end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)} \end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \left(e^{\beta\mu}\right)^n\sum_{s(n)}e^{-\beta{E}_{s(n)}}\right)} \end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_n\right)} \end{equation}
+
+\begin{equation}\frac{PV}{k_BT}=\ln{\left(\sum_{n=0}^\infty \lambda^nZ_{on}Q_n\right)} \end{equation}
+where $Q_n$ is the configuration integral.
+\color{red}For a system of N particles only, the sum collapses to one term, that for n=N, which gives \color{black}
+\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
+%=\ln\lambda^n + \ln{Z}_{on}+\ln{Q}_n
+=\beta\mu{N} + \ln{Z}_{oN} +\ln{Q}_N\end{equation}
+\begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN} + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
+Noting that $Q_N=1$ for an ideal gas since $V(r)=0$ for an ideal gas
+\begin{equation}\frac{P_{ideal}V}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN}=N\end{equation}
+%\begin{equation}P_{ideal}=\mu{N} + k_BT\ln{Z}_{oN}\end{equation}
+\begin{equation}\frac{P_{excess}}{k_BT}= \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
+and Equation~\ref{PV/kBT_reduces_for_N} becomes
+\begin{equation}\frac{PV}{k_BT}=N + \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
+\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
+Plugging in the expression for $F_{excess}$ in terms of the mayer function for low density fluids
+\begin{equation}F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
+gives
+\begin{equation}\frac{PV}{k_BT}=N + \left(-\frac{N}{V}\int{f(r)d\vec{r}}\right){N} + \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
+%\begin{equation}\frac{PV}{k_BT}=N - \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
+\begin{equation}\frac{PV}{k_BT}=N\left(1-\frac{N}{V}\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
+\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)\right) \end{equation}
+where $B_2(T)$ is given by
+\begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(r)d\vec{r} \end{equation}
+
+\color{red}This result was for low densities where higher terms in the expression for the excess free energy were neglected. The full virial equation is given by\color{black}
+\begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}
+
+%-----------------END Virial Expansion---------------
+%\[\]
+
 \cite{fake}

### 2020-09-05 papers/thesis-kirstie/thesis.tex
more cleanups

@@
 \begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-which gives
+which, for a canonical ensemble, gives
 \begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
@@
     \centering
-     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
+     %\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
+     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
     \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
@@
     \centering
-     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
+     %\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
+     %FIXME\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
     \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
@@
     \centering
-     % FIXME\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
+    %\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
+    % FIXME\includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
     \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
@@
 %The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
+
 The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume.
 But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
-\color{black}
-
 %*********************** DFT minimize F/N *********************
---------------------------------------------------------
-\color{red}
-Starting with Euler's equation
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
-\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)+d\mu\end{equation}
-and using the Gibbs-Duhem relation which follows from Euler's equation
-\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
-which gives
-\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
-Equation~\ref{d(U/N)long} becomes
-\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
+%--------------------------------------------------------
+%\color{red}
+%Starting with Euler's equation
+%%\begin{equation}U=TS-PV+\mu{N}\end{equation}
+%\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
+%\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}%dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)%+d\mu\end{equation}
+%and using the Gibbs-Duhem relation which follows from Euler's equation
+%\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
+%which gives
+%\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
+%Equation~\ref{d(U/N)long} becomes
+%\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
 %\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
-\color{black}
+%\color{black}
 The Helmholtz free energy F is defined as 
@@
 \begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
-After plugging in Equation~\ref{d(U/N)}, and setting $f_N=\frac{F}{N}$ this becomes
+Setting $f_N=\frac{F}{N}$ this becomes
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
 This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed.
-\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
-Thus the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
+It is the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
+\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_N(T,\frac{V}{N})\end{equation}
 
@@
 \begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
-where conservation of energy was also applied. Using Eq~\ref{d(U/N)} for the surroundings, but fixing the volume and the number of particles so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$ results in
+where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
 \begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
@@
 Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
-\begin{equation}F=\text{min}_{n(\vec r)}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} An extremum is reached when 
-\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=0\end{equation}
-Taking this derivative gives  \color{red} CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL \color{black}
+\begin{equation}F=\text{min}_{U}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation}
+An extremum is reached when
+\begin{equation}\frac{\partial{F}}{\partial{U}}=0\end{equation}
+\color{red} 
+\begin{equation}F=\text{min}_{n(\vec{r})}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} 
+\color{red} 
+\begin{equation}\frac{\delta{F}}{\delta{n(\vec r)}}\bigg|_{n(\vec{r})_{equil}}=0\end{equation}
+Taking this derivative gives  ??????? CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL 
+%\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
 \begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
@@
 \begin{equation}=0\end{equation}
+\color{black}
 As shown earlier, the extremum that is reached at equilibrium is a minimum.
@@
 
-------------------------------------------------------
+%------------------------------------------------------
 %*************************END DFT minimize F/N ****************
@@
 \begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
-where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$:
-\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}\exp^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
+where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$
+\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
 \begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~where~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
@@
 
-Breaking the Hamiltonians H into its component energies:
+Breaking the Hamiltonians H into its component energies
 \begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
-where the translational kinetic energy of the N atoms is given by: \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
-the potential energy of the total interaction of each atom with the other N-1 atoms is given by:
+where the translational kinetic energy of the N atoms is given by \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
+the potential energy of the total interaction of each atom with the other N-1 atoms is given by
 \begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position.
@@
 
-Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts:
-\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume defined by: \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation}
-the equation can be written:
-\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)
-}\end{equation}  and the partitian function can be expressed as:
+Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts
+\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation}
+%the equation can be written
+%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}
+and the partitian function can be expressed as
 \begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
-where:
-\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
-\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another.
+where
+\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
+\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another.
 The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential.
-The free energy can also be broken into two parts using Eq~\ref{Zmultiply}:
+The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
 %\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
@@
 The ideal Helmholtz free energy is
-\begin{equation}{F_{ideal}= -k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation} where
-%the integral can easily be evaluated:
-
-\begin{displaymath}{\frac{1}{h^{3N}}\int{...}\int{dP}\exp^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}=\frac{1}{h^{3}}\int{d\vec{p}_1}\exp^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}\exp^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}\exp^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}}\end{displaymath}
-%Each term on the right 
-These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$:
-\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}\exp^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
+\begin{displaymath}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{displaymath}
+
+\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
+These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
+\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
 where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
-Rewriting equation (20) in terms of $\Lambda$ gives:
-\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes:
-\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$: \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
+Rewriting Equation~\ref{Fidealmultipleint} in terms of $\Lambda$ gives
+\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes
+\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
+Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$ gives \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
 \begin{equation}{F_{ideal}= k_BT[n\ln(n\Lambda^3)-n]V}\label{eq:Fideal}\end{equation}   
-For a spatially varying number density, the ideal free energy in Eq~\ref{eq:Fideal} becomes:
+For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
 \begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}
-\[\]
-xxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxxx
-
-Now consider the excess free energy which is written as:
-\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\exp^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
+
+Now consider the excess free energy which is written as
+\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
 This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system.
-In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes: \begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation}
+In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes
+\begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
 %\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
@@
 %\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
-Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in equation \ref{Fexcess} becomes:
-
-\begin{displaymath}{\exp^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=\exp^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}\exp^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}\exp^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
+Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in equation \ref{Fexcess} becomes
+
+\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.
 %\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    %$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%
-The function $f_{ij}$ is called a Mayer function, and is given by\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=\exp^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} Multiplying the Mayer function terms, the excess free energy can be written as
+The function $f_{ij}$ is called a Mayer function, and is given by\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
+Multiplying the Mayer function terms, the excess free energy can be written as
 \begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}
 At low densities, it is not likely to find two or more spheres in close proximity, and so only interactions between pairs of spheres are considered. In this case the equation reduces to
-\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...
-\right]\right)  }}\end{equation} Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ we get \begin{equation}{\int{...}\int{dQ}=V^N}\end{equation}\begin{equation}{\int{...}\int{dQ}\sum{f_{ij}}=\frac{1}{2}N(N-1)V^{N-1}\int{f(r)}{d\vec{r}}}\end{equation} \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}}=\frac{1}{8}N(N-1)(N-2)(N-3)V^{N-2}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation} %which becomes \begin{equation}{F_{excess}=-k_BT\ln{\left(1+ \frac{1}{2}\frac{N(N-1)}{V^{N}}V^{N-2}\int{...}\int{dQ}{f_{ij}} + .....\sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...\right)}}\end{equation}%
+%\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...
+\right]\right)  }}\end{equation}
+\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +... \right]\right)  }}\end{equation} 
+Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ we get \begin{equation}{\int{...}\int{dQ}=V^N}\end{equation}\begin{equation}{\int{...}\int{dQ}\sum{f_{ij}}=\frac{1}{2}N(N-1)V^{N-1}\int{f(r)}{d\vec{r}}}\end{equation} \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}}=\frac{1}{8}N(N-1)(N-2)(N-3)V^{N-2}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation}
+%which becomes \begin{equation}{F_{excess}=-k_BT\ln{\left(1+ \frac{1}{2}\frac{N(N-1)}{V^{N}}V^{N-2}\int{...}\int{dQ}{f_{ij}} + .....\sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...\right)}}\end{equation}%
 %FINISH!!!!! \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}f_{mn}}=\frac{1}{48}N(N-1)(N-2)(N-3)(N-4)(N-5)(N-6)V^{N-3}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation} 
@@
 Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ for a homogeneous density n=$\frac{N}{V}$ this becomes
-\begin{equation}{F_{excess}=-k_BT\ln{\left(exp^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} \begin{equation}{F_{excess}=-k_BT\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\end{equation}  %\begin{equation}{F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation}%
+\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} \begin{equation}{F_{excess}=-k_BT\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\end{equation}  %\begin{equation}{F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation}%
 Thus the free excess energy per atom in the low-density limit is  \begin{equation}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} where the quantity in parenthesis can be recognized as the second virial expanision coefficient typically called $B_2$.

### 2020-09-05 papers/thesis-kirstie/thesis.tex
more cleanups

@@
 
-Breaking the Hamiltonians H into its component energies
+Breaking the Hamiltonian, H into its component energies
 \begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
@@
 %\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}
-and the partitian function can be expressed as
+the partitian function can be expressed as
 \begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
@@
 
-\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
+\begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
 These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
@@
 \begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
-Multiplying by $\frac{V}{V}$ gives an expression in terms of the number density $n=\frac{N}{V}$ gives \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N\Lambda^3}{V}\right)}-\frac{N}{V}\right]}V\end{equation}
-\begin{equation}{F_{ideal}= k_BT[n\ln(n\Lambda^3)-n]V}\label{eq:Fideal}\end{equation}   
+Rewriting the equation in terms of the number density $n=\frac{N}{V}$, and the quantum volume, $n_Q=\frac{1}{\lambda^3}$ gives
+\begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N}{V}\frac{1}{n_Q}\right)}-\frac{N}{V}\right]}V\end{equation}
+\begin{equation}{F_{ideal}= k_BT[n\ln(n/n_Q)-n]V}\label{eq:Fideal}\end{equation}   
 For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
-\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}
-
-Now consider the excess free energy which is written as
+\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/n_Q)-n(\vec{r})]d\vec{r}}\end{equation}
+
+Now consider the excess free energy which is written most generally as
 \begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
@@
 %\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
-Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in equation \ref{Fexcess} becomes
+Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in Equation~\ref{Fexcess} becomes
 

### 2020-09-05 papers/thesis-kirstie/thesis.tex
more cleanup

@@
 \begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
-These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$
-\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}\end{equation} 
+These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$, or the quantum volume $\text{n}_\text{Q}$
+\begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}=\text{n}_\text{Q}\end{equation} 
 where \begin{equation}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{equation} 
@@
 \begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
-Rewriting the equation in terms of the number density $n=\frac{N}{V}$, and the quantum volume, $n_Q=\frac{1}{\lambda^3}$ gives
-\begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N}{V}\frac{1}{n_Q}\right)}-\frac{N}{V}\right]}V\end{equation}
-\begin{equation}{F_{ideal}= k_BT[n\ln(n/n_Q)-n]V}\label{eq:Fideal}\end{equation}   
+Rewriting the equation in terms of the number density $n=\frac{N}{V}$, and the quantum volume, $\text{n}_\text{Q}$ gives
+\begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N}{V}\frac{1}{\text{n}_\text{Q}}\right)}-\frac{N}{V}\right]}V\end{equation}
+\begin{equation}{F_{ideal}= k_BT[n\ln(n/\text{n}_\text{Q})-n]V}\label{eq:Fideal}\end{equation}   
 For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
-\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/n_Q)-n(\vec{r})]d\vec{r}}\end{equation}
+\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]d\vec{r}}\end{equation}
 

### 2020-09-07 papers/thesis-kirstie/thesis.tex
more modifications and removed a bunch of junk

@@
 It is the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
-\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_N(T,\frac{V}{N})\end{equation}
+%\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_N(T,n)\end{equation}
+\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
 
@@
 As shown earlier, the extremum that is reached at equilibrium is a minimum.
-(reference shethna) \color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids).
-\color{black} It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.
+(reference shethna) %\color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids).
+\color{black}
+\color{red} To show that F is unique ... \color{black} 
+It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.
 
@@
 Now consider the excess free energy which is written most generally as
-\begin{equation}\label{Fexcess}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
+\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
 This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system.
@@
 %\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    %$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%
-The function $f_{ij}$ is called a Mayer function, and is given by\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
+The function $f_{ij}$ is called a Mayer function, and is given by
+\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
 Multiplying the Mayer function terms, the excess free energy can be written as
-\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)  }}\end{equation}
-At low densities, it is not likely to find two or more spheres in close proximity, and so only interactions between pairs of spheres are considered. In this case the equation reduces to
-%\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...
+\begin{equation}\label{Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +...
 \right]\right)  }}\end{equation}
-\begin{equation}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +... \right]\right)  }}\end{equation} 
-Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ we get \begin{equation}{\int{...}\int{dQ}=V^N}\end{equation}\begin{equation}{\int{...}\int{dQ}\sum{f_{ij}}=\frac{1}{2}N(N-1)V^{N-1}\int{f(r)}{d\vec{r}}}\end{equation} \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}}=\frac{1}{8}N(N-1)(N-2)(N-3)V^{N-2}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation}
-%which becomes \begin{equation}{F_{excess}=-k_BT\ln{\left(1+ \frac{1}{2}\frac{N(N-1)}{V^{N}}V^{N-2}\int{...}\int{dQ}{f_{ij}} + .....\sum_{i\neq{j}\neq{k}\neq{l}}{f_{ij}f_{kl}} +\sum_{i\neq{j}\neq{k}\neq{l}\neq{m}\neq{n}}{f_{ij}f_{kl}f_{mn}} +...\right)}}\end{equation}%
-%FINISH!!!!! \begin{equation}{\int{...}\int{dQ}\sum{f_{ij}f_{kl}f_{mn}}=\frac{1}{48}N(N-1)(N-2)(N-3)(N-4)(N-5)(N-6)V^{N-3}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}}\end{equation} 
-and so on.
-In the low-density limit, the terms in the series get smaller, and the excess free energy can then be written as %begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{8}\frac{N^4}{V^{2}}\int{f(r)}{d\vec{r}}\int{f(r)}{d\vec{r}}+ ...\right)}\end{equation}%
-\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
+
+  \begin{figure}[h!]
+    \centering
+    %\includegraphics[height=7cm]{diagrammic.pdf}
+    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations between individual pairs of atoms are shown in green. 
+             Interactions between pairs of atoms with an atom shared in common are shown in red. 
+             At low densities only the interactions shown in green are considered, since atoms are not likely to be clustered.}
+    \label{fig:diagrammic}
+  \end{figure}
+
+At low densities, it is not likely to find two or more atoms in close proximity, and so only interactions between individual pairs of spheres, such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
+Each pair can then be described by a function of the distance between the two atoms in the pair. Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ for each pair we get 
+\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
+\color{black}
+With a careful counting of the resulting number of such pairs, the excess free energy can then be written as 
 where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
 Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ for a homogeneous density n=$\frac{N}{V}$ this becomes
-\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} \begin{equation}{F_{excess}=-k_BT\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\end{equation}  %\begin{equation}{F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation}%
-Thus the free excess energy per atom in the low-density limit is  \begin{equation}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} where the quantity in parenthesis can be recognized as the second virial expanision coefficient typically called $B_2$.
-%where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.%
+\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
+Thus the free excess energy per atom in the low-density limit is given by 
+\begin{equation}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} where the quantity in parenthesis can be recognized 
+as the second virial expanision coefficient typically called $B_2$. %where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
 It is important that the expression for the excess Helmholtz free energy correctly reduces to this result in the low-density limit. 
 %Although we will not be restricting ourselves to the low-density limit, it is important that the theory we use correctly reduces to this same result in the low-density limit. 
+Outside of the low density limit, it is quite difficult to evaluate the excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}. 
+Another way to form an expression for the excess Helmholtz free energy in terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
 
 %(FINISH! Take P=-dF/dV and set equal to virial expansion for P to get virial coefficients in terms of the mayer functions....CLARIFY why cDFT is better than this route!)
-
-Another way to form an expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ is to use Fundamental Measure Theory (FMT).
-FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory (SPT).
-Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.
 
@@
 
+Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid 
+which can then be integrated to find the excess Helmholtz free energy.
+\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+It is an extension of the work done by Widom who showed that the excess chemical potential of a homogenesous fluid corredsponds to the reversible work 
+of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R.
+
+Widom's result was obtained by considering the difference in the excess Helmholtz free energy when one atom is added to an otherwise closed system. 
+He showed that the excess chemcial potential is related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
+between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
+\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
+\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+%Equation~\ref{widom} is called Widom's Insertion Formula.
+
+%%COME BACK TO THIS POINT!
+>>>>>\color{red}For a system...
+$<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widoms-insertion-formula} becomes \begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
+
+%In a randomly positioned (I think it has to be homogeneous!)
+\noindent To evaluate Equation~\ref{widoms-insertion-formula} for a Hard Sphere fluid consider that
+no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere. 
+But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; 
+the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widoms-insertion-formula} becomes
+\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
+where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
+
+\begin{figure}[h!]
+    \centering
+    %\includegraphics[height=5cm]{Probability-of-overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
+             as shown by the black dot at the center of the dotted region near the top. 
+             But it is not possible to place a sphere at a position where an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
+
+For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
+by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
+that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
+%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
+\color{red}Thus, for cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
+%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
+\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+\color{red} ???More generally, for Ro $<$ R this becomes???
+\begin{equation}???\mu_{ex}(Ro < R)=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))???\end{equation}
+\color{black}
+
+NOW DO UPPER LIMIT  Ro $>>$ R
+
+For Ro $>>$ R ...
+
 --------------------------------------------------
+%-----SFMT -------
+
+An expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ can be found using Fundamental Measure Theory (FMT).
+FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of 
+inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory (SPT). 
+Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.
 
@@
 \item Deriviation of the Tensor version of $\Phi_3$ 
-\item Equations related to the Error Function Potential $V_{erf}$ ....
 \item Second Virial Coefficient $B_{2}$ for the Error Function Potential 
 \item Third Virial Coefficient $B_{3}$
+\item Derivation of the Virial Equation
 \end{enumerate}
@@
 %\[{}\]
-\textbf{6. Deriviation of $\Phi_3$}
+\textbf{6. Deriviation of the Tensor version of $\Phi_3$ }
 \[{}\]
@@
 
-%\textbf{7. Equations related to the Error Function Potential $V_{erf}$ ....:}
-%\[{}\]
-%\[{}\]
-
-\textbf{8. Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$:}
+\textbf{7. Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$:}
 \begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}
@@
 \[{}\]
-\textbf{9. Third Virial Coefficient $B_{3}$:}
+\textbf{8. Third Virial Coefficient $B_{3}$:}
 
@@
 \[{}\]
-\noindent \textbf{Derivation of the Excess Free Energy from Scaled Particle Theory}
-
-The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R. 
-\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
-This result is obtained by considering the difference in the excess Helmholtz free energy when one atom is added to the system.
-It was shown by Widom to be related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ of a system with N atoms and one with N+1 atoms.
-The resulting equation is known as Widom's Insertion Formula.
-%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
-\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
-\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-%Equation~\ref{widom} is called Widom's Insertion Formula. 
-\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
-\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
-%In a randomly positioned (I think it has to be homogeneous!)
-\indent In a collection of hard spheres making up a hard sphere fluid, no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere.
-But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; the energy required to do that would be infinite.
-Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
-\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
-
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=3cm]{P_overlap.png}
-    \caption{No energy is needed to place a sphere at a position where there is no overlap such as shown by the black dot at the center of the dotted region near the top.
-But it is not possible to place a sphere at a position where an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
-
-For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere.
-%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
-\color{red}Thus, considering cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
-%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-Using this result, scaled particle theory seeks a more general expression of the excess chemical potential of a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_o$ where the size of $R_o$ falls into three different cases: $R_o >> R$, $R_o > R$, and $R_o < R$.
-
-\begin{equation}\label{3_parts_to_mu_ex}{\mu_{ex}(R_o)=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_o^3 & R_o >> R \\ P\frac{4\pi}{3}R_o^3 + c_2R_o^2 + c_1R_o^1 + c_0R_o^0  & R_o > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3)) & R_o < R\end{array}\right.}\end{equation}
-
-%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  must tend toward in the limit that $R_o$ goes to infinity.
-By expanding the logrythmic term (ln(1-x)= ...) in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.
-
-%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
-The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_o$, and so it is assumed that the middle function, which is valid for values of $R_o$ in between the two extremes, can also be expressed in terms of powers of $R_o$.
-The coeficients $c_2$, $c_1$ and $c_0$ in the middle function can be determnined be considering that this function (valid for $R_o > R$) must be continuous with the last function (valid for $R_o < R$) along with their first and second derivatives.
-The middle function must also tend toward the first function in the limit that $R_o$ goes to infinity.
-The chemical potential then becomes
-\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
-where $\eta$ is the packing fraction given by
-\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}
-
-%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$.
-From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation
-%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
-%From the expression for pressure given by
-
-The pressure can be obtained from \color{red}
-\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
-%The excess Hemlholtz free energy is then found from the relation
-and the Helmholz free energy from
-\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
-which gives
-%\begin{displaymath}F_{ex}=\end{displaymath}
-\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}
-
-\noindent \textbf{Derivation of the Virial Equation}
+
+\noindent \textbf{9. Derivation of the Virial Equation}
 %----------------BEGIN Virial Expansion--------------
@@
 \begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
-Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
+Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the 
+thermodynamic quantities as ensemble averages
 \begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
@@
 
+\[{}\]
+\noindent \textbf{10. Derivation of the Excess Free Energy from Scaled Particle Theory}
+
+The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R 
+into a fluid consisting of hard spheres of radius R. 
+\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
+This result is obtained by considering the difference in the excess Helmholtz free energy when one atom is added to the system. 
+It was shown by Widom to be related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
+of a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
+\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
+\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+%Equation~\ref{widom} is called Widom's Insertion Formula. 
+\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
+\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
+%In a randomly positioned (I think it has to be homogeneous!)
+\indent In a collection of hard spheres making up a hard sphere fluid, no energy would be required to place a hard sphere at points 
+surrounded by an empty space the size of a hard sphere. But a hard sphere cannot be placed at points where a hard sphere would 
+overlap another hard sphere; the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
+\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
+where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
+
+\begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=3cm]{Probability-of-overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
+             as shown by the black dot at the center of the dotted region near the top. 
+             But it is not possible to place a sphere at a position where an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
+
+For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
+by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
+that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
+%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
+\color{red}Thus, considering cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
+%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
+\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+
+%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+
+Using this result, scaled particle theory seeks a more general expression of the excess chemical potential of 
+a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_o$ 
+where the size of $R_o$ falls into three different cases: $R_o >> R$, $R_o > R$, and $R_o < R$.
+
+\begin{equation}\label{3_parts_to_mu_ex}{\mu_{ex}(R_o)=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_o^3 & R_o >> R 
+\\ P\frac{4\pi}{3}R_o^3 + c_2R_o^2 + c_1R_o^1 + c_0R_o^0  & R_o > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3)) & R_o < R\end{array}\right.}\end{equation}
+
+%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  
+%must tend toward in the limit that $R_o$ goes to infinity. By expanding the logrythmic term (ln(1-x)= ...) 
+%in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.
+
+%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
+The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_o$, 
+and so it is assumed that the middle function, which is valid for values of $R_o$ in between the two extremes, 
+can also be expressed in terms of powers of $R_o$.  
+The coeficients $c_2$, $c_1$ and $c_0$ in the middle function can be determnined be considering that this function (valid for $R_o > R$) 
+must be continuous with the last function (valid for $R_o < R$) along with their first and second derivatives. 
+The middle function must also tend toward the first function in the limit that $R_o$ goes to infinity. The chemical potential then becomes
+\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
+where $\eta$ is the packing fraction given by
+\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}
+
+%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$. 
+%From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation 
+%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
+%From the expression for pressure given by
+
+The pressure can be obtained from \color{red}
+\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
+%The excess Hemlholtz free energy is then found from the relation
+and the Helmholz free energy from
+\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
+which gives
+%\begin{displaymath}F_{ex}=\end{displaymath}
+\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}
+
 %-----------------END Virial Expansion---------------
-%\[\]
 
@@
 \end{document}
-
---------------------------------------------------------------------------------
---------------------------------------------------------------------------------
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=6cm]{SimpleCell_Gaus.png}
-    \caption{A 3D Gaussian Distribution}
-    \label{fig:SimpleCell_Gaus}
-  \end{figure}
-
-The number density profile is varied, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found, which is known from cDFT to be unique.
-This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure ??
-.
-
-\[{}\]
-Program Implemtation
-\[{}\]
-A simplified flowchart of the program created to investigate the freezing of a WCA fluid is shown on pages x-x.
-The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.
-Given values for n* and t*, the program varys gw and fv and compares the free energies of the different number density profiles to find the minimum free energy for that combination of n* and t*.
-
-******startA
-In the Monte-Carlo Integration vectors are generated randomly according to a Gaussian distribution, the weighted density is computed for each vector, and the average of all the results gives the final weighted density.
-.....
-ABOUT R!!
-\begin{displaymath}{n_i += \frac{(1-fv)w_i}{NumPoints}}\end{displaymath}
-
-The result can be made more accurate by increasing NumPoints which is the number of vectors randomly generated.
-In order to further reduce error, antithetic variates are used.
-An atithetic variate is an equal, but opposite point at which a function is evaluated.
-The final result is then given by the average of the two results.
-This reduces first order error by filtering out....
-
-\begin{displaymath}{n_i += \frac{(1-fv)\left[\frac{w_i(\vec{r}) +w_i(-\vec{r})}{2}\right]}{NumPoints}}\end{displaymath}
-
-(...........Find weights $w_0(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$,  $w_1(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$, $w_2(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$, $w_3(|(\vec{r}-\vec{R_j}) - \vec{r'}|)$ and compute weighted densities.........)
-******endA
-
-***startB  old - before corrected center of weight function 
-For a vector $\vec{r}=\vec{r}_k$ within the primitive cell, a vector $\vec{R}_j$ to the center of one of the Gaussians is selected that is within the primitive cell, or one of the neighboring cells within a prudently picked inclusion radius as Gaussians farther away will not contriubute much.
-The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, $n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions $w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to the center of the weight function.
-Monte-Carlo integration works by randomly generating $\vec{dr}$ according to a Gaussian distribution.
-***endB
-
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4cm]{SimpleCell_Gaus.png}
-    \caption{A homogeneouse number density profile $n(\vec{r})$}
-    \label{fig:homogen_denisty}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4cm]{W0.png}
-    \caption{A crossectional view of weight function $w_2$ which is a spherical shell with soft edges in 3D. Weight functions $w_0$ and $w_1$ have a similar shape.}
-    \label{fig:W0}
-    \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4cm]{W3.png}
-    \caption{A crossectional view of weight function $w_3$ which is a sphere with soft edges in 3D.}
-    \label{fig:W3}
-    \end{figure}
-
-////////alpha and Xi vary with temp:
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4.7cm]{plot_alpha_vs_KbT.png}
-    \caption{Shows how $\alpha$ varies with temperature}
-    \label{fig:alpha_vs_T}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4.7cm]{plot_xi_vs_KbT.png}
-    \caption{Shows how $\Xi$ varies with temperature}
-    \label{fig:xi_vs_T}
-  \end{figure}
-
-/////weight equations
-\begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} and the shifted weighting functions $w_0$, $w_1$, $w_2$, $w_3$ are given by:
-\begin{equation}{w_0(|\vec{r}-\vec{r}'|)= \frac{w_2}{4\pi(|\vec{r}-\vec{r}'|)^2}}\end{equation} 
-\begin{equation}{w_1(|\vec{r}-\vec{r}'|)= \frac{w_2}{4\pi(|\vec{r}-\vec{r}'|)}}\end{equation} 
-\begin{equation}{\vec{w}_1(|\vec{r}-\vec{r}'|)= w_1\frac{\vec{r}-\vec{r}'}{|\vec{r}-\vec{r}'|}}\end{equation} 
-\begin{equation}{}\end{equation} 
-\begin{equation}{\vec{w}_2(|\vec{r}-\vec{r}'|)=w_2\frac{\vec{r}-\vec{r}'}{|\vec{r}-\vec{r}'|}}\end{equation}
-\begin{equation}{w_3(|\vec{r}-\vec{r}'|)={\int_r}^{\infty}{w_2(|\vec{r}-\vec{r}'|)}{d}\vec{r}'}\end{equation} 
-In this paper, w2 is given by ....
-\begin{equation}{w_2(|\vec{r}-\vec{r}'|)=\frac{1}{\Xi\sqrt\pi}\exp^-\frac{(|\vec{r}-\vec{r}'|-\frac{\alpha}{2})^2}{\Xi^2}}\end{equation}
-which gives:
-\begin{equation}{w_3(|\vec{r}-\vec{r}'|)=\frac{1}{2}}\left(1-erf(\frac{(|\vec{r}-\vec{r}'|)-\frac{\alpha}{2}}{\Xi})\right)\end{equation} where the two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4).
-
-%BEGIN SFMT--------------------------------BEGIN SFMT----------------------------------
-%Unlike a standard weighted density approximation, 
-Soft Fundamental Measure Theory (SFMT) is used to obtain an expression for $F_{excess}$ in terms of a set of functions $\Phi_1$, $\Phi_2$, and $\Phi_3$ 
-\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation} \color{red} which in turn depend on a set of weighted densities $n_0, n_1, n_2, n_3$ that arise naturally in the expression for the excess free energy derived from scaled particle theory.
-%The weighted densities $n_0, n_1, n_2, n_3$ arise naturally in the expression for the excess free energy derived from scaled particle theory.
-
-%SFMT uses a set of weighted densities $n_0, n_1, n_2, n_3$ that arise naturally in the expression for the excess free energy derived from scaled particle theory. 
- %There is no one single weight function, however, that can effectively describe the coarse density. Rather, a set of weights will do the job, each giving rise to its own "weighted density". 
-%\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
-
-\begin{equation}{F_{excess}=k_BT\left(-n_0ln(1-n_3)+\frac{n_1n_2}{1-n_3}+\frac{n_2^3}{24\pi(1-n_3)^2}\right)V}\end{equation}
-where \begin{equation}n_i=nw_i\end{equation}
-\begin{equation}w_0=\frac{w_2}{4\pi{R^2}}=1\end{equation}
-\begin{equation}w_1=\frac{w_2}{4\pi{R}}=R\end{equation}
-\begin{equation}w_2=4\pi{R}^2\end{equation}
-\begin{equation}w_3=\frac{4}{3}\pi{R}^3\end{equation}
-\color{black}
-For a spatially varying number density, $F_{excess}$ becomes a functional of the number density profile $n(\vec{r})$ :
-%SFMT extends these equations to a spatially varying number density
-\begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
-\begin{equation}{F_{excess}[n(\vec{r})]=k_BT\int\left(-n_0ln(1-n_3)+\frac{n_1n_2-\vec{n_1}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{1-n_3}+  \frac{n_2^3\left(1-{(\frac{\vec{n_2}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{{n_2}^2})}\right)^3}{24\pi(1-n_3)^2 }\right){d\vec{r}}}\end{equation}
-where \begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation}
-%which reduces for a uniform number density to
-\color{red}
-
-%///////Compare this to what is involved with a calculation of the free energy using classical Density Functional Theory.//////
-
-\begin{equation}{\Phi_1= -n_0\ln(1-n_3)}\end{equation} 
-\begin{equation}{\Phi_2= \frac{n_1n_2-\vec{n_1}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{1-n_3}}\end{equation} 
-\begin{equation}{\Phi_3=\frac{n_2^3\left(1-{(\frac{\vec{n_2}\mbox{~\scriptsize\textbullet~}\vec{n_2}}{{n_2}^2})}\right)^3}{24\pi(1-n_3)^2}}\end{equation} %where the weighted densities $n_0$, $n_1$, $n_2$, $n_3$ are given by:
-
-%\begin{equation}{n_i(\vec{r})= \int{n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} and the weighting functions $w_0$, $w_1$, $w_2$, $w_3$ are given by:
-
-\begin{equation}{w_0(r)= \frac{w_2(r)}{4\pi{r^2}}}\end{equation} 
-\begin{equation}{w_1(r)= \frac{w_2(r)}{4\pi{r}}}\end{equation}
-\begin{equation}{w_2(r)=-\frac{\partial{w_3(r)}}{\partial{r}}}\end{equation}  
-\begin{equation}{\vec{w}_1(r)= w_1(r)\frac{\vec{r}}{|\vec{r}|}=\frac{w_2(r)}{4\pi{r}}\frac{\vec{r}}{|\vec{r}|}}\end{equation} 
-\begin{equation}{\vec{w}_2(r)=w_2(r)\frac{\vec{r}}{|\vec{r}|}}\end{equation}
-
-%\begin{equation}{w_3(r)={\int_r}^{\infty}{w_2(r')}{dr'}}\end{equation} 
-\color{black}
-In this paper, $w_3(r)$ is given by 
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-erf\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
-as proposed by Schmidt$^{(?)}$ where $w_3$ is subject to the boundary conditions $w_3(0)\approx{1}$ and $w_3(inf)=0$.
-These boundary conditions are satisfied if $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$.
-Solving equation (??) for $w_2(r)$
-%\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{infinity}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
-we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
-\begin{equation}{\frac{df(r)}{dr}=\int_{-inf}^{inf}{w_2(r')w_2(r-r')dr'}}\end{equation}  Solving equation (??) for the Mayer function
-\begin{displaymath}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{displaymath} \begin{displaymath}{f(r)=\int_{inf}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{displaymath} we get
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-erf\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} This gives an interatomic potential of \begin{displaymath}{V_{erf}=-k_BTln(f(r)-1)}\end{displaymath} \begin{equation}{V_{erf}=-k_BTln\left[\frac{1}{2}\left(erf\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} where the two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), and are determined to make $V_{erf}$ match $V_{WCA}$ as closely as possible.
-
-%The SFMT equations and expressions for $w_2$, $\Xi$ and $\alpha$ are developed in prior work.
-It is not the aim of this paper to enter into the details of SFMT, but rather to take these given expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.
-
-%END SFMT--------------------------------END SFMT----------------------------------
-
-\color{red} %Virial Expansion
-%\begin{equation}U=TS-PV+\mu{N}\end{equation}
-%%%\begin{equation}U-\mu{N}=TS-PV\end{equation}
-%\begin{equation}e^{-\beta(U-\mu{N})}=e^{-\beta(TS-PV)}\end{equation}
-%%%\begin{equation}e^{-\beta(U-\mu{N})}=e^{-\frac{TS}{k_BT}+\frac{PV}{k_BT}}\end{equation}
-%\begin{equation}e^{-\frac{(U-\mu{N})}{k_BT}}e^{\frac{S}{k_B}}=e^{\frac{PV}{k_BT}}%\end{equation}
-%%\begin{equation}e^{-\beta(U-\mu{N})}e^{\ln\Omega}=e^{\frac{PV}{k_BT}}\end{equation}
-%\begin{equation}e^{-\beta(U-\mu{N})}e^{-\sum{P_i}\ln{P_i}}=e^{\frac{PV}{k_BT}}\end{equation}
-%%%\begin{equation}e^{-\beta(U-\mu{N})}\Omega=e^{\frac{PV}{k_BT}}\end{equation}
-%%%\begin{equation}\ln(e^{-\beta(U-\mu{N})}\Omega)=\ln(e^{\frac{PV}{k_BT}})\end{equation}
-%%\begin{equation}\ln(e^{-\beta(U-\mu{N})}\Omega)=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
-%%%\begin{equation}\ln(Z_{Grand})+\Omega=\frac{PV}{k_BT}\end{equation}
-%%\begin{equation}\ln((e^{\beta\mu})^Ne^{-\beta{U}}\Omega)=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
-%\begin{equation}\ln((e^{\beta\mu})^Ne^{-\beta{U_N}-\sum{P_i}\ln{P_i}})=\frac{PV}{k_BT}~~~~~~\text{for N particles in microcanonical ensemble?}\end{equation}
-\color{black}
-
-%*********************** DFT minimize F/V *********************
---------------------------------------------------------
-
-Starting with Euler's equation
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-\begin{equation}\frac{U}{V}=\frac{TS}{V}-\frac{PV}{V}+\frac{\mu{N}}{V}\end{equation}
-\begin{equation}\label{d(U/V)long}d\left(\frac{U}{V}\right)=\frac{T}{V}dS+\frac{S}{V}dT+TSd\left(\frac{1}{V}\right)-dP+\mu{dn}+nd\mu\end{equation}
-and using the following relation from Euler's equation
-\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
-which gives
-\begin{equation}\frac{S}{V}dT-dP+nd\mu=0\end{equation}
-Equation~\ref{d(U/V)long} becomes
-\begin{equation}\label{d(U/V)}d\left(\frac{U}{V}\right)=\frac{T}{V}dS+TSd\left(\frac{1}{V}\right)+\mu{dn}\end{equation}
-%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
-The Helmholtz free energy F is defined as 
-\begin{equation}F=U-TS\end{equation}
-\begin{equation}\frac{F}{V}=\frac{U}{V}-\frac{TS}{V}\end{equation}
-\begin{equation}d\left(\frac{F}{V}\right)=d\left(\frac{U}{V}\right)-d\left(\frac{TS}{V}\right)\end{equation}
-\begin{equation}d\left(\frac{F}{V}\right)=d\left(\frac{U}{V}\right)-\frac{T}{V}dS-\frac{S}{V}dT-TS d\left(\frac{1}{V}\right)\end{equation}
-After plugging in Equation~\ref{d(U/V)}, and setting $f_V=\frac{F}{V}$ this becomes
-\begin{equation}df_V=-\frac{S}{V}dT+\mu{dn}\end{equation}
-This result shows that the natural variables of the Helmholtz free energy per volume $f_V$ are temperature and number density
-\begin{equation}f_V(T,n)\end{equation}
-It is the Helmholtz free energy per volume $f_V$ that will be minimized in a system that has a temperature and number density that are fixed.
-
-------------------------------------------------------
-%*************************END DFT minimize F/V ****************
-
--------------------------this blue/red part basically garbage------------------:
-\color{blue}
-A faster, and much easier way to find the excess free energy is to use classical Density Functional Theory (cDFT).
-As stated earlier, in cDFT the free energy is expressed as a functional of the number density $F[n(\vec{r})]$.
-%The ideal free energy in Eq~\ref{eq:Fideal} can be expressed in terms of a spatially varying number density $n(\vec{r})$.
-
-%For a spatially varying number density, the ideal free energy in Eq~\ref{eq:Fideal} becomes:
-%\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]d\vec{r}}\end{equation}
-
-To form an expression for the excess free energy as a functional of the number density, Soft Fundamental Measure theory (SFMT) is used.
-The excess free energy of a fluid can be expressed in general as an integral over an excess free energy density $\frac{F}{V}$.
-\begin{displaymath}{F_{excess}=\int{\frac{F}{V}{dV}}}\end{displaymath}
-Rewriting this in terms of the excess free energy per particle $\phi=\frac{F}{N}$, or in terms of $\Phi=\frac{1}{k_BT}\frac{F}{N}\frac{N}{V}=\frac{\phi{n}}{k_BT}$ we get
-%$\frac{F}{V}=\frac{F}{N}\frac{N}{V}=\phi{n}$  
-\begin{displaymath}{F_{excess}=\int{\frac{F}{N}\frac{N}{V}{dV}}=\int{\phi{n}{dV}}=k_BT\int\Phi{dV}}\end{displaymath} Showing explicitly the dependence of $\phi$ on the density at each point this becomes
-\begin{displaymath}{F_{excess}=\int{\phi(n(\vec{r}))n(\vec{r}){d\vec{r}}}=k_BT\int\Phi(n(\vec{r})){d\vec{r}}}\end{displaymath}
-where $n(\vec{r})$ is a smoothly varying function.
-\color{red} However, there is no real smoothly varying function when looking closeup.
-The real density is coarse, being zero where there are no atoms, or spheres located in a volume, and having a maximum of one at the center of each sphere over a volume of one sphere.
-In other places, with various portions of a sphere within a volume of one sphere, it is somewhere in between.
-To account for, and describe this coarse density as a function of position, the smoothly varying density function is muliplied by a weight function which is centered about each point in space.
-The weighted density as a function of position then becomes
-\begin{displaymath}{n_{weighted}(\vec{r})=\int{n(\vec{r})w(|\vec{r}-\vec{r}'|){d\vec{r}}}}\end{displaymath}
-%In a weighted density approximation, the excess free energy per particle $\phi$ is expressed in terms of the weighted density \begin{displaymath}{F_{excess}\approx\int{\phi(n_{weighted}(\vec{r}))n(\vec{r}){d\vec{r}}}}\end{displaymath} or, setting $\Phi=\frac{\phi{n}}{k_BT}$ \begin{displaymath}{F_{excess}\approx{k_BT}\int{\Phi(\vec{r}){d\vec{r}}}}\end{displaymath}
-
-%(CAREFUL- probability density!)
-%(CAREFUL- weighted density is a weighted average!)
-\color{black}
--------------------------------end basically garbage------------------
-
-Another way to form an expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ is to use Soft Fundamental Measure theory (SFMT).
-SFMT extends Fundamental Measure Theory (FMT) developed for hard sphere fluids to those with soft potentials like that of the WCA fluid.
-
-%In summary, it is the aim of this paper to use these expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.
-
-In summary, classical Density Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT) provides a method by which the free energy can be found simply and directly without taking into account the interaction each sphere, or atom, has with every other sphere.
-For a number density n that is constant everywhere, as would be expected for a homogeneous fluid in equilbrium, the free energy can be written as a function of n where the ideal Helmholtz free energy can be determined analytically, and SFMT is used to form an expresssion for the excess Helmholtz free energy
-\begin{equation}{F(n)=F_{ideal}(n) + F_{excess}(n)}\end{equation} 
-For a spatially varying number density, as would be expected for a crystal, the free energy can be written as a functional of $n(\vec{r})$ where again SFMT is used to form the expression for the excess Helmholtz free energy.
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-To solve for the crystal free energy $F[n(\vec{r})]$ using cDFT all that is needed are number density profiles, such as Gaussian profiles that depend only on $\vec{r}$. 
-\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} The profiles are varied until the free energy is minimized.
-The minimum free energy and its corresponding number density profile are the equilibrium values for that temperature.
-%The complexity is greatly reduced making cDFT a powerful tool in modeling fluid behavior.

### 2020-09-07 papers/thesis-kirstie/thesis.tex
some more cleanups

@@
 
-%------------------------------------------------------
 %*************************END DFT minimize F/N ****************
-
-%Before looking at the details of cDFT, let us first look at the free energy. 
-%To see how cDFT is adventageous, consider what is involved in a straightforward calculation of the free energy without using Density Functional Theory.
 
@@
 
-%In summary, it is the aim of this paper to use these expressions, and apply cDFT to see how well these expressions and this tool predict the freezing phase transistion of a WCA fluid.
-
 In summary, classical Density Functional Theory (cDFT) used in conjunction with Soft Fundamental Measure Theory (SFMT) provides a powerful tool in modeling the thermodynamic properties of fluids.
@@
 Outside the region of coexistence, the fluid will be completely in a crystalline state if the crystal free energy is lower than the liquid free energy, and it will be completely in a liquid state if the liquid free energy is lower than the crystal free energy as was discussed in Section IB.
-%The complexity is greatly reduced making cDFT a powerful tool in modeling fluid behavior.
+
 \[{}\]
@@
 Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{\vec{w}}_2(k)$ becomes
-%\begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
@@
 \begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
-%\mbox{~%\scriptsize\textbullet~}
 Since the integration is over all space, the integral can be simplified by setting 
@@
 \[{}\] \color{green}
-%\color{red}
-%ERROR\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=4{\pi}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\left[\left(\frac{\Xi}{3k\sqrt{2}}-\frac{1}{k^2}\right)\cos(\frac{k\alpha}{2})+\frac{\Xi\alpha}{12*\sqrt{2}}\sin(\frac{k\alpha}{2})\right]}\end{displaymath}
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath}
@@
 \[{}\]
-%\subsection{}
+
 \textbf{4. Integrals used to Compute Fourier Transforms of Weight Functions}
 \begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
-%\tiny
 \begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation} 
@@
 \begin{equation}{=\frac{ka^3}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})+\frac{a\alpha}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{equation} 
-%\[{}\]
 \[{}\]
@@
 \begin{equation}{=\frac{\pi}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
-%\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+%\[{}\]
+
 \textbf{5. Deriviation of Tensor Density $n_{m2ij}(\vec{r})$}
@@
 %\[{}\]
+
 \textbf{6. Deriviation of the Tensor version of $\Phi_3$ }
@@
 \[{}\]
+%\[{}\]
 
@@
 \[{}\]
+%\[{}\]
+
 \textbf{8. Third Virial Coefficient $B_{3}$:}
@@
 \[{}\]
+%\[{}\]
 
 \noindent \textbf{9. Derivation of the Virial Equation}
-%----------------BEGIN Virial Expansion--------------
 The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
@@
 \[{}\]
+%\[{}\]
+
 \noindent \textbf{10. Derivation of the Excess Free Energy from Scaled Particle Theory}
-
 The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R 
@@
 
-%-----------------END Virial Expansion---------------
-
 \cite{fake}

### 2020-09-08 papers/thesis-kirstie/thesis.tex
simplified

@@
 But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
-%*********************** DFT minimize F/N *********************
-%--------------------------------------------------------
-%\color{red}
-%Starting with Euler's equation
-%%\begin{equation}U=TS-PV+\mu{N}\end{equation}
-%\begin{equation}\frac{U}{N}=\frac{TS}{N}-\frac{PV}{N}+\mu\end{equation}
-%\begin{equation}\label{d(U/N)long}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+\frac{S}{N}%dT+TSd\left(\frac{1}{N}\right)-\frac{V}{N}dP-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)%+d\mu\end{equation}
-%and using the Gibbs-Duhem relation which follows from Euler's equation
-%\begin{equation}SdT-VdP+Nd\mu=0\end{equation}
-%which gives
-%\begin{equation}\frac{S}{N}dT-\frac{V}{N}dP+d\mu=0\end{equation}
-%Equation~\ref{d(U/N)long} becomes
-%\begin{equation}\label{d(U/N)}d\left(\frac{U}{N}\right)=\frac{T}{N}dS+TSd\left(\frac{1}{N}\right)-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
-%\begin{equation}dU=TdS+SdT-PdV-VdP+\mu{dN}+Nd\mu\end{equation}
-%\color{black}
+
 The Helmholtz free energy F is defined as 
@@
 \begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
-
-The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will expreience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
-Here "surroundings" means the entire universe minus the system.
-This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.
-How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?
-Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
-and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
-\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
-where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
-\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
-\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative.
-(reference schroder)
-
-More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium.
-Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
-\begin{equation}F=\text{min}_{U}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation}
-An extremum is reached when
-\begin{equation}\frac{\partial{F}}{\partial{U}}=0\end{equation}
-\color{red} 
-\begin{equation}F=\text{min}_{n(\vec{r})}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation} 
-\color{red} 
-\begin{equation}\frac{\delta{F}}{\delta{n(\vec r)}}\bigg|_{n(\vec{r})_{equil}}=0\end{equation}
-Taking this derivative gives  ??????? CHECK PARTIAL DERVIVATIVE OF FUNCTIONAL 
-%\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
-\begin{equation}\frac{\partial{(U-TS)}}{\partial{n(\vec r)}}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\frac{\partial{S}}{\partial{U}}\bigg|_{V,N}\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
-\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-T\left(\frac{1}{T}\right)\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
-\begin{equation}=\frac{\partial{U}}{\partial{n(\vec r)}}-\frac{\partial{U}}{\partial{n(\vec r)}}\end{equation}
-\begin{equation}=0\end{equation}
-\color{black}
-As shown earlier, the extremum that is reached at equilibrium is a minimum.
-(reference shethna) %\color{red} To show that there is only one minimum, that the extremum is unique, I refer the reader to ...(reference fluids).
-\color{black}
-\color{red} To show that F is unique ... \color{black} 
-It remains to be shown that the Helmholtz free energy can be expressed as a functional of the number density profile.
-
-%*************************END DFT minimize F/N ****************
 

### 2020-09-08 papers/thesis-kirstie/thesis.tex
a bit more simplified

@@
 \begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
-
 This, coupled with the fact that the free energy of a system tends toward a minimum at equilibrium means that the equilibrium free energy of the system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying $n(\vec{r})$ through all possible spatial profiles until the free energy is minimized.
@@
 
-The reason this works is because cDFT satisfies the conditions of the variational principle, the same mathematical principle used to determine that the shortest distance between two points is a straight line by varying the path between the two points until the distance is minimized.
-
-A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
-\begin{equation}\Phi=F-\mu{N}\end{equation}
-%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
-which can then be expressed in terms of a varrying number density profile $n(\vec r)$
-%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
-\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and
-\begin{equation}F_{int}= F -\int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
-%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$.
-%The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
-
-The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume.
-But this paper will be examining a system with a fixed temperature, volume, and number of particles in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
-
+%The reason this works is because cDFT satisfies the conditions of the variational principle, the same mathematical principle used to determine that the shortest distance between two points is a straight line by varying the path between the two points until the distance is minimized.
+
+In this paper a system with a fixed temperature, volume, and number of particles will be examined in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
 The Helmholtz free energy F is defined as
@@
 With a careful counting of the resulting number of such pairs, the excess free energy can then be written as 
+\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
 where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 

### 2020-09-08 papers/thesis-kirstie/thesis.tex
much more simplified

@@
 Outside of the low density limit, it is quite difficult to evaluate the excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}.
+
 Another way to form an expression for the excess Helmholtz free energy in terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
-
-%(FINISH! Take P=-dF/dV and set equal to virial expansion for P to get virial coefficients in terms of the mayer functions....CLARIFY why cDFT is better than this route!)
-
-%NEW-------Scaled Partical Theory-----------------
---------------------------------------------------
-
-Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid 
-which can then be integrated to find the excess Helmholtz free energy.
+Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid which can then be integrated to find the excess Helmholtz free energy.
 \begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
@@
 of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R. 
-
 Widom's result was obtained by considering the difference in the excess Helmholtz free energy when one atom is added to an otherwise closed system. 
@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-%Equation~\ref{widom} is called Widom's Insertion Formula.
-
-%%COME BACK TO THIS POINT!
->>>>>\color{red}For a system...
-$<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widoms-insertion-formula} becomes \begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
-
-%In a randomly positioned (I think it has to be homogeneous!)
-\noindent To evaluate Equation~\ref{widoms-insertion-formula} for a Hard Sphere fluid consider that
-no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere. 
-But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; 
-the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widoms-insertion-formula} becomes
-\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
-
-\begin{figure}[h!]
-    \centering
-    %\includegraphics[height=5cm]{Probability-of-overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
-             as shown by the black dot at the center of the dotted region near the top. 
-             But it is not possible to place a sphere at a position where an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
-
-For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
-by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
-that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
-%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
-\color{red}Thus, for cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
-%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-\color{red} ???More generally, for Ro $<$ R this becomes???
-\begin{equation}???\mu_{ex}(Ro < R)=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))???\end{equation}
-\color{black}
-
-NOW DO UPPER LIMIT  Ro $>>$ R
-
-For Ro $>>$ R ...
-
---------------------------------------------------
-%-----SFMT -------
-
-An expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ can be found using Fundamental Measure Theory (FMT).
-FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of 
-inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory (SPT). 
-Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.
-
-Scaled Partical Theory (SPT) gives an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
-\begin{equation}\label{SPT_FE}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
+\begin{equation}=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))\end{equation}
+Forming a more general result for different sizes of cavities in the fluid, Scaled Partical Theory arrives at an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$
+\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
 where
@@
 
+An expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ can be found using Fundamental Measure Theory (FMT).
+FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory.
+Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.
+
 %\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
@@
 %\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
-From this expression, a quantity $\zeta_{3}$ can be defined as follows
+From Equation~\ref{Fexcess-SPT}, a quantity $\zeta_{3}$ can be defined as follows
 \begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=n\zeta_{3}}\end{equation}

### 2020-09-08 papers/thesis-kirstie/thesis.tex
smoothed some stuff out

@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-\begin{equation}=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))\end{equation}
+\begin{equation}=-k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity}+R)^3))\end{equation}
+where the radius of the cavity $R_{cavity}$ is equal to the radius of a hard sphere, R.
 Forming a more general result for different sizes of cavities in the fluid, Scaled Partical Theory arrives at an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$
@@
 
-\noindent The excess free energy in Eq~\ref{SPT_FE} can be written in terms of the number density $\frac{N}{V}$
+\noindent The excess free energy in Eq~\ref{Fexcess-SPT} can be written in terms of the number density $\frac{N}{V}$
 \begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
@@
  where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ and $w_3(\infty)=0$. 
-\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
+%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
 Using Equation~\ref{w2_w3_relation}, shown again below, 
@@
 
---------------------------------------------------
-%END NEW------------------------------------------
-
-In summary, classical Density Functional Theory (cDFT) used in conjunction with Soft Fundamental Measure Theory (SFMT) provides a powerful tool in modeling the thermodynamic properties of fluids.
-With SFMT an expression for the excess free energy in terms of the number density can be found simply and directly without taking into account the interaction each sphere, or atom, has with every other sphere.
-For a spatially varying number density, as would be expected for a crystal, the free energy can then be written as a functional of the number density profile $n(\vec{r})$
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-To implement cDFT, the number density profile $n(\vec{r})$ is varied until the free energy is minimized.
+In summary, classical Density Functional Theory (cDFT) is used in conjunction with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free energy of a system, and the corresponding equilibrium number density profile.
+In particular, SFMT is used to form the expression for the excess free energy without having to take into account the interaction each sphere, or atom, has with every other sphere.
+The excess free energy is then added to the ideal free energy calculated analytically to find the resulting Free energy for the system.
+\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+For a spatially varying number density, as would be expected for a crystal, the free energy is a functional of the number density profile $n(\vec{r})$  
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} To implement cDFT, the number density profile $n(\vec{r})$ is varied with the temperature fixed until the free energy is minimized.
 The minimum free energy and its corresponding number density profile are then the equilibrium values for that temperature.
-\color{red} $\leftarrow$CAREFUL!
-\color{black} Thus to use cDFT all that is needed are number density profiles, such as Gaussian profiles that depend on position and which can be varied, for example, by changing the width of the Gaussian.
-\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation}
-
-For a number density n that is constant everywhere, as would be expected for a homogeneous fluid in equilbrium, the free energy can be written as a function of n where the ideal Helmholtz free energy can be determined analytically, and SFMT is used once again to form an expresssion for the excess Helmholtz free energy
-\begin{equation}{F(n)=F_{ideal}(n) + F_{excess}(n)}\end{equation}
 

### 2020-09-08 papers/thesis-kirstie/thesis.tex
a couple cleanups

@@
 In particular, SFMT is used to form the expression for the excess free energy without having to take into account the interaction each sphere, or atom, has with every other sphere.
-The excess free energy is then added to the ideal free energy calculated analytically to find the resulting Free energy for the system.
+The excess free energy is then added to the ideal free energy calculated analytically to find the resulting free energy for the system.
 \begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
@@
 Once the Helmholtz free energy for both the crystal and liquid states of the fluid are found, they can be compared to find the resulting state of the fluid.
-Outside the region of coexistence, the fluid will be completely in a crystalline state if the crystal free energy is lower than the liquid free energy, and it will be completely in a liquid state if the liquid free energy is lower than the crystal free energy as was discussed in Section IB.
+Outside the region of coexistence, the fluid will be completely in a crystalline state if the crystal free energy is lower than the liquid free energy, and it will be completely in a liquid state if the liquid free energy is lower than the crystal free energy.
 

### 2020-09-28 papers/thesis-kirstie/thesis.tex
cleanup
attempt to generate pdf from odg files
working on references

@@
 \title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
-\author{by Kirstie Finster}
+\author{Kirstie L. Finster}
 \date{August 2020}
@@
 
-The key idea behind classical Density Functional Theory (cDFT) is that free energy can be written as a functional (a function of a function) of a number density profile $n(\vec{r})$ which describes the way the number density, or atoms per volume, varies spatially.
+The key idea behind classical Density Functional Theory (cDFT) is that free energy can be written as a functional (a function of a function) of a number density profile $n(\vec{r})$ which describes the way the number density, or atoms per volume, varies spatially on average.
 \begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
@@
 
-In this paper a system with a fixed temperature, volume, and number of particles will be examined in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
-The Helmholtz free energy F is defined as
-\begin{equation}F=U-TS\end{equation}
-\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
-\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
-\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
-Setting $f_N=\frac{F}{N}$ this becomes
+A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
+\begin{equation}\Phi=F-\mu{N}\end{equation}
+%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
+which can then be expressed in terms of a varrying number density profile $n(\vec r)$
+%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
+\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
+where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and the Helmholtz free energy is expressed as
+\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
+%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$.
+%The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
+
+The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume.
+But in this paper a system with a fixed temperature, volume, and number of particles will be of interest in which case
+\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
+and so 
+\begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
+The dependence on $n(\vec r)$ is now only present in the expression for the Helmholtz free energy, and so it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium.
+%In this paper a system with a fixed temperature, volume, and number of particles will be examined in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
+More specifically, in this paper, the Helmholtz free energy per particle, or atom,  $f_N=\frac{F}{N}$ will be minimized which is given by
+\begin{equation}f_N=\frac{U}{N}-\frac{TS}{N}\end{equation}
+with differential
+\begin{equation}\label{usetoshowmin}df_N=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
+%\begin{equation}df_N=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
 This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed.
-It is the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
-%\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_N(T,n)\end{equation}
+%Thus, the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
+%\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_V(T,n)\end{equation}
 \begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
+
+The minimization can be illustrated by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
+Here "surroundings" means the entire universe minus the system.
+This, together with the conservation of energy (the first law of thermodynamics) which says that energy lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that changes in the free energy will be negative for spontaneous processes when the natural variables of the free energy are held fixed.
+%And so the free energy tends toward a minimum.
+%How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?
+
+Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
+\begin{equation}df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}\end{equation}
+and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
+\begin{equation}df_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
+where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
+\begin{equation}df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
+\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative, and thus the free energy tends toward a minimum.
+(reference schroder) This illustrates the tendancy toward minimization for a system near equilibrium.
+A general proof of minimization can be found in (reference simple fluids).
 
@@
 \begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
-where the paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$
+which results from equating the expressions for the Helmholtz free energy from thermodynamics to those of statistical mechanics.
+The paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$ is given by
 \begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
@@
 Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid which can then be integrated to find the excess Helmholtz free energy.
-\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
+\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
 It is an extension of the work done by Widom who showed that the excess chemical potential of a homogenesous fluid corredsponds to the reversible work 
@@
 He showed that the excess chemcial potential is related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
-between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula given by
 %\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-\begin{equation}=-k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity}+R)^3))\end{equation}
-where the radius of the cavity $R_{cavity}$ is equal to the radius of a hard sphere, R.
+\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
+where the radius of the cavity $R_{cavity}$ is equal to the radius of a hard sphere, R, and the cavity is either filled or not.
+\color{black}%For cavities that are much larger than one of the hard spheres,
 Forming a more general result for different sizes of cavities in the fluid, Scaled Partical Theory arrives at an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$
@@
 The weighted densities can now be expressed as
-\begin{equation}{CHECK!n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} with weight functions
+\begin{equation}{\color{red} CHECK!~~\color{black}n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} 
+with weight functions
 \begin{equation}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
@@
 %\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+\begin{equation}\label{mu_excess}\mu_{ex}(R_{cavity}=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
 
@@
 Using this result, scaled particle theory seeks a more general expression of the excess chemical potential of 
-a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_o$ 
-where the size of $R_o$ falls into three different cases: $R_o >> R$, $R_o > R$, and $R_o < R$.
-
-\begin{equation}\label{3_parts_to_mu_ex}{\mu_{ex}(R_o)=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_o^3 & R_o >> R 
-\\ P\frac{4\pi}{3}R_o^3 + c_2R_o^2 + c_1R_o^1 + c_0R_o^0  & R_o > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3)) & R_o < R\end{array}\right.}\end{equation}
+a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
+where the size of $R_{cavity}$ falls into three different cases: $R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
+
+\begin{equation}\label{3_parts_to_mu_ex}{W_{rev}(R_{cavity})=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_{cavity}^3 & R_{cavity} >> R 
+\\ P\frac{4\pi}{3}R_{cavity}^3 + c_2R_{cavity}^2 + c_1R_{cavity}^1 + c_0R_{cavity}^0  & R_{cavity} > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity})^3)) & R_{cavity} \leq R\end{array}\right.}\end{equation}
 
@@
 %The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
-The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_o$, 
-and so it is assumed that the middle function, which is valid for values of $R_o$ in between the two extremes, 
-can also be expressed in terms of powers of $R_o$.  
-The coeficients $c_2$, $c_1$ and $c_0$ in the middle function can be determnined be considering that this function (valid for $R_o > R$) 
-must be continuous with the last function (valid for $R_o < R$) along with their first and second derivatives. 
-The middle function must also tend toward the first function in the limit that $R_o$ goes to infinity. The chemical potential then becomes
+The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_{cavity}$, 
+and so it is assumed that the middle function, which is valid for values of $R_{cavity}$ in between the two extremes, 
+can also be expressed in terms of powers of $R_{cavity}$.  
+The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
 \begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
@@
 
-\cite{fake}
+%------------------n(r)--move to "back" later---------------------
+------------------------------------------------------------------
+
+The actual positions of atoms in a microstate  \color{red}?at time t, or averaged over time?:\color{black}
+\begin{equation}\rho_{s}(\vec r)=\sum_{i=1}^N\delta(\vec r - \vec r_i)\end{equation}
+The equilibrium single number density distribution equals the ensemble average over all microstates, and is independent of time:
+\begin{equation}n(\vec r)=\rho^{(1)}(\vec r)=~<\rho_s(\vec r)>~\end{equation}
+The reduced equilibrium phase-space probablity density of n particles in volume $dr^ndp^n$:
+\begin{equation}f_o^{(n)}(\vec r^n, \vec p^n)=\rho^{(n)}(\vec r^n)f_M^{(n)}(\vec p^n)\end{equation}
+The relation between the  reduced equilibrium phase-space probablity density and the equilibrium phase-space probablity density:
+\begin{equation}f_o^{(n)}(\vec r^n, \vec p^n)=\frac{N!}{(N-n)!}\int\dot~\dot~\dot~\int f_o^{[N]}(\vec r^N, \vec p^N) d\vec r^{N-n} d\vec p^{N-n}\end{equation}
+The canonical equilibrium phase-space probablity density:
+\begin{equation}f_o^{[N]}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}
+where
+\begin{equation}{partian function}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
+The equilibrium phase-space probablity density integrated over all continuous microstates $(\vec r^N,\vec p^N)$ is 1:
+\begin{equation}\int\dot~\dot~\dot~\int f_o^{[N]}(\vec r^N, \vec p^N) d\vec r^{N} d\vec p^{N}=1\end{equation}
+The equilibrium phase-space probablity density is independent of time:
+\begin{equation}f_o^{[N]}(\vec r^N, \vec p^N) = f_o^{[N]}(\vec r^N, \vec p^N; t)\end{equation}
+The Helmholtz free energy is a functional of the equilibrium phase-space probability density:
+\begin{equation}U = <E> = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N}\end{equation}
+\begin{equation}TS = -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln\left(N!~h^{3N}f_o^{[N]}\right) d\vec r^{N} d\vec p^{N} \end{equation}
+\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N} \end{equation}
+\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[f_o^{[N]}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right] \end{equation}
+\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[f_o^{[N]}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right] \end{equation}
+\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln(Z)d\vec r^{N} d\vec p^{N}\end{equation}
+\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}d\vec r^{N} d\vec p^{N}\right]\end{equation}
+\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[1\right]\end{equation}
+and so 
+\begin{equation}F = U - TS\end{equation}
+\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N}-\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} -k_BTln(Z)\end{equation}
+\begin{equation}F = -k_BTln(Z)\end{equation}
+
+----------------------------------------------
+
+\cite{rosenfeld1989}
+\cite{tarazona2000}
+\cite{schmidt1999density}
+\cite{schmidt2000fluid}
+
+\cite{santos2012phi3}
+
+\cite{TschoppBHextended}
+
+%1-21
+\cite{lucretius}
+\cite{rowlinson1964statistical}
+\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
+  lafitte2013accurate}
+\cite{carnahan1969equation}
+\cite{rosenfeld1989, rosenfeld1997,
+  roth2002whitebear}
+\cite{rosenfeld1989}
+%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
+\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
+%\cite{schmidt1999density}
+\cite{schmidt2000density, groh2001density, kim2001adsorption,
+  sweatman2002fundamental}
+%\cite{schmidt2000fluid, sweatman2002fundamental}
+\cite{weeks1971}
+%\cite{rosenfeld1989}
+%\cite{schmidt1999density}
+%%\cite{tarazonaFIXME,santos2012phi3}
+%\cite{schmidt1999density} 
+%\cite{schmidt2000fluid}
+%\cite{schmidt1999density, schmidt2000fluid}
+%\cite{schmidt1999density}
+%\citenum{schmidt2000fluid}
+%\citenum{schmidt2000fluid}
+%a Face Centered Cubic (FCC) lattice is formed~\cite{reference!}
+\cite{barker1967perturbation}
+\cite{roth2002whitebear} 
+%\cite{roth2002whitebear}
+\cite{mikolaj2004structure, eggert2002quantitative,
+yarnell1973structure}
+\cite{verlet1967computer}
 
@@
 \end{document}
+
+%-------------Minimization-----------------
+%The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
+Here "surroundings" means the entire universe minus the system.
+This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.
+How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?
+Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
+%\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
+%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
+%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
+%where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
+%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
+%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative.
+(reference schroder)
+
+%More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium.
+Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
+%\begin{equation}F=\text{min}_{U}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation}
+%An extremum is reached when
+%\begin{equation}\frac{\partial{F}}{\partial{U}}=0\end{equation}
+
+%\begin{equation}F=U-TS\end{equation}
+%\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
+%\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
+%\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
+%Setting $f_N=\frac{F}{N}$ this becomes
+%\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
+
+%-------Scaled Partical Theory-----------------
+Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid 
+which can then be integrated to find the excess Helmholtz free energy.
+\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+It is an extension of the work done by Widom who showed that the excess chemical potential of a homogenesous fluid corredsponds to the reversible work 
+of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R.
+
+Widom's result was obtained by considering the difference in the excess Helmholtz free energy when one atom is added to an otherwise closed system. 
+He showed that the excess chemcial potential is related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
+between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
+\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
+\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+%Equation~\ref{widom} is called Widom's Insertion Formula.
+
+%%COME BACK TO THIS POINT!
+>>>>>\color{red}For a system...
+$<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widoms-insertion-formula} becomes \begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
+
+%In a randomly positioned (I think it has to be homogeneous!)
+\noindent To evaluate Equation~\ref{widoms-insertion-formula} for a Hard Sphere fluid consider that
+no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere. 
+But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; 
+the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widoms-insertion-formula} becomes
+\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
+\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
+where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
+
+\begin{figure}[h!]
+    \centering
+    %\includegraphics[height=5cm]{Probability-of-overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
+             as shown by the black dot at the center of the dotted region near the top. 
+             But it is not possible to place a sphere at a position where an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
+
+For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
+by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
+that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
+%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
+\color{red}Thus, for cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
+%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
+\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+\color{red} ???More generally, for Ro $<$ R this becomes???
+\begin{equation}???\mu_{ex}(Ro < R)=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))???\end{equation}
+\color{black}
+
+NOW DO UPPER LIMIT  Ro $>>$ R
+
+For Ro $>>$ R ...
+--------------------------------------------------

### 2020-10-13 papers/thesis-kirstie/thesis.tex
more updates

@@
 
-\noindent IA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)
+\noindent I. Introduction
+
+Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
+with one another like billiard balls~\cite{lucretius}. This model, called the Hard-Sphere fluid, has served as 
+the primary reference system for fluids. This is due to the fact that the repulsion between atoms is 
+much stronger than their attraction, and so the repulsive potential, which is descriptive of hard spheres, 
+dominates the interactions between atoms.
+
+%Van der Waal took into account both the attraction, and the respulsion 
+%of the impenetrable spheres in devopling his equation of state for an ideal gas in 1xxx.
+
+In the late 1950s, Scaled Particle Theory (SPT) was developed which provided an expression for the 
+excess Helmholtz free energy of a homogeneous Hard-Sphere fluid. 
+%This .... allowed for ... 
+In the 1960s Percus-Yevick applied Density Functional Theory (DFT) to find an expression 
+for the excess Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. In the late 1960s, and early 1970s, 
+Barker and Henderson adapted the DFT method to use with fluids composed of 
+spheres with softly fading potential functions rather than the hard edges of the spheres 
+in the Hard-Sphere fluid. This was done by making the radius of the soft spheres temperature dependent, 
+so that at low temperatures the fluid could be modeled by hard spheres with large radii, and at high 
+temperatures the fluid could be modeled by hard spheres with smaller radii. This is because at higher 
+temperatures, the soft-spheres would penetrate into each other farther when they collide. 
+Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential, 
+which became known as the Weeks-Chandler-Anderson model fluid.
+
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT) to form an expression for the excess 
+Helmholtz free energy of an inhomogeneous hard sphere fluid based on the results from SPT. In 
+19xx Schmidt adopted FMT for soft-sphere atoms to form Soft Fundamental Measure Theory (SFMT). In this paper, 
+Density Functional Theory will be used to predict the freezing of a Weeks-Chandler-Anderson model fluid 
+by applying SFMT to form the expression for the excess Helmholtz free energy, and the results will be 
+compared to those generated from Monte-Carlo simulations.
+%\[{}\]
+\newline
+\newline
+\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)
 
@@
 \[\]
-\noindent IB. The Theory of Freezing and Maxwell's Double Tangent Construction
+\noindent IIB. The Theory of Freezing and Maxwell's Double Tangent Construction
 
@@
     \centering
-     %\includegraphics[height=7cm]{MaxwellDTC-Fig1.pdf}
+     %\includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
      % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
@@
 \[\]
-IC. Classical Density Functional Theory (cDFT)
+\noindent IIC. Classical Density Functional Theory (cDFT)
 
@@
 \[{}\]
-II. METHODS
+III. METHODS
 
@@
 from the center of the weight function.
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-III. RESULTS
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
+%\[{}\]
+
+Monte-Carlo integration is employed to more efficiently 
+compute the weighted densities used to calculate the excess free energy which is then 
+added to the ideal free energy (computed analytically) to get the total free energy 
+which is minimized.
+
+\newpage
+
+\noindent IV. Results and Discussion
+
+Phase diagrams for the WCA fluid are shown below. In all cases $\sigma=1$ and $\epsilon=1$.
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+\begin{figure}
+ \centering
+  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
+\caption{Phase Diagram of the reduced termperature verses the reduced density. Green lines indicate the region of coexistence from Monte-Carlo simulations.}
+\label{fig:Phase_Diagram_of_T_vs_n}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
+ \end{center}
+\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The green line marks the boundary from Monte-Carlo simulations.}
+\label{fig:Phase_Diagram_P_vs_T}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
+ \end{center}
+\caption{Phase Diagram of the reduced temperature verses the reduced density. The grey region indicates coexistence.}
+\label{fig:T-vs-n_at_fixed_P}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
+ \end{center}
+\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The lines bending downward indicate freezing.}
+\label{fig:p-vs-T_at_fixed_density}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
+ \end{center}
+\caption{Phase Diagram of the reduced pressure verses the reduced volume. The grey region indicates coexistence.}
+\label{fig:p-vs-V_at_fixed_T}
+\end{figure}
+
+\begin{figure}
+ \begin{center}
+  %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
+ \end{center}
+\caption{Phase Diagram of the reduced pressure verses the reduced density. The grey region indicates coexistence.}
+\label{fig:p-vs-n_at_fixed_T}
+\end{figure}
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\[{}\]
+\[{}\]
+\[{}\]
+\[{}\]
+
+\noindent V. Conclusion
+
 \[{}\]

### 2020-10-15 papers/thesis-kirstie/thesis.tex
start getting ready to add more figures
more updates

@@
 Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
-with one another like billiard balls~\cite{lucretius}. This model, called the Hard-Sphere fluid, has served as 
+with one another like billiard balls%~\cite{kinetictheory}
+. This model, called the Hard-Sphere fluid, has served as 
 the primary reference system for fluids. This is due to the fact that the repulsion between atoms is 
@@
 
-In the late 1950s, Scaled Particle Theory (SPT) was developed which provided an expression for the 
-excess Helmholtz free energy of a homogeneous Hard-Sphere fluid. 
-%This .... allowed for ... 
-In the 1960s Percus-Yevick applied Density Functional Theory (DFT) to find an expression 
-for the excess Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. In the late 1960s, and early 1970s, 
-Barker and Henderson adapted the DFT method to use with fluids composed of 
+In the late 1950s, Scaled Particle Theory (SPT) was developed~\cite{Reiss Frich Lebowitz} from which an equation of state, and
+an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be obtained~\cite{Hansen}. 
+%Percus and Yevick took another approach to the Hard-Sphere fluid.
+In 1961 Morita and Hiroike introduced a density functional approach to finding an expression for the excess Helmholtz free energy of inhomogeneous fluids~\cite{MoritaDFT}.
+% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
+In 1967 %the late 1960s, and early 1970s, 
+Barker and Henderson~\cite{barker1967perturbation} adapted perturbation methods to use with fluids composed of 
 spheres with softly fading potential functions rather than the hard edges of the spheres 
-in the Hard-Sphere fluid. This was done by making the radius of the soft spheres temperature dependent, 
-so that at low temperatures the fluid could be modeled by hard spheres with large radii, and at high 
-temperatures the fluid could be modeled by hard spheres with smaller radii. This is because at higher 
-temperatures, the soft-spheres would penetrate into each other farther when they collide. 
-Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential, 
+in the Hard-Sphere fluid. This was done by making the radius of the spheres temperature dependent, 
+so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
+and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
+%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other when they collide at higher temperatures.
+%The different size radii coorespond to the different depths to which two soft spheres would penetrate into each other when they collide. 
+This reflects the idea that soft-spheres penetrate more deeply into each other when they collide at higher temperatures. 
+Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
 which became known as the Weeks-Chandler-Anderson model fluid.
 
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT) to form an expression for the excess 
-Helmholtz free energy of an inhomogeneous hard sphere fluid based on the results from SPT. In 
-19xx Schmidt adopted FMT for soft-sphere atoms to form Soft Fundamental Measure Theory (SFMT). In this paper, 
+DFT Evans 1979
+
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} which provides an expression for the excess 
+Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
+Tarazona later improved on the method by introducing a tensor version of the functional~\cite{tarazonaphi3}. 
+In 1999 Schmidt adopted FMT for soft spheres forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. In this paper, 
 Density Functional Theory will be used to predict the freezing of a Weeks-Chandler-Anderson model fluid 
-by applying SFMT to form the expression for the excess Helmholtz free energy, and the results will be 
-compared to those generated from Monte-Carlo simulations.
+by applying SFMT with Schmidt's proposed erf model\cite{schmidt2000fluid}. %for generating the weight funcions, 
+%and Tarazona's tensor functional. 
+Results will be compared to those generated from Monte-Carlo simulations.
 %\[{}\]

### 2020-10-15 papers/thesis-kirstie/thesis.tex
more edits

@@
 Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
-with one another like billiard balls%~\cite{kinetictheory}
-. This model, called the Hard-Sphere fluid, has served as 
-the primary reference system for fluids. This is due to the fact that the repulsion between atoms is 
-much stronger than their attraction, and so the repulsive potential, which is descriptive of hard spheres, 
-dominates the interactions between atoms.
+with one another like billiard balls. In the simplest case, the spheres are treated as having 
+only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
+reference system for fluids. This is due to the fact that the repulsion between atoms is much 
+stronger than their attraction, and so the repulsive potential, which is descriptive of 
+hard spheres, dominates the interactions between atoms.
 
@@
 %Percus and Yevick took another approach to the Hard-Sphere fluid.
-In 1961 Morita and Hiroike introduced a density functional approach to finding an expression for the excess Helmholtz free energy of inhomogeneous fluids~\cite{MoritaDFT}.
-% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
+In 1961 Morita and Hiroike introduced a density functional approach to finding an expression for the 
+excess Helmholtz free energy of inhomogeneous fluids~\cite{MoritaDFT}.
+% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
+Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
 In 1967 %the late 1960s, and early 1970s, 
@@
     \centering
-      % FIXME \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
+    %\includegraphics[height=3.5cm]{TwoSpheres.pdf} 
     \caption{(Left) The WCA potential is a minimum at r=2R. 
@@
 
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
 \begin{figure}
@@
 \begin{figure}
- \begin{center}
+ \centering
  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
- \end{center}
 \caption{Phase Diagram of the reduced pressure verses the reduced temperature. The green line marks the boundary from Monte-Carlo simulations.}

### 2020-10-15 papers/thesis-kirstie/thesis.tex
more edits

@@
 
-In the late 1950s, Scaled Particle Theory (SPT) was developed~\cite{Reiss Frich Lebowitz} from which an equation of state, and
+In the late 1950s, Scaled Particle Theory (SPT) was developed~\cite{ReissSPT} from which an equation of state, and
 an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be obtained~\cite{Hansen}. 

### 2020-10-16 papers/thesis-kirstie/thesis.tex
edits on intro

@@
 hard spheres, dominates the interactions between atoms. 
-
-%Van der Waal took into account both the attraction, and the respulsion 
-%of the impenetrable spheres in devopling his equation of state for an ideal gas in 1xxx.
-
-In the late 1950s, Scaled Particle Theory (SPT) was developed~\cite{ReissSPT} from which an equation of state, and
-an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be obtained~\cite{Hansen}. 
+In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
+created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
+fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
+modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+This reflects the idea that soft-spheres penetrate more deeply into each other 
+when they collide at higher temperatures. 
+%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
+%and thus be represented by spheres with smaller radii. 
+In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
+with Rowlinson's idea of using temperature-dependent radii to model spheres with 
+soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
+where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
+larger radii at low temperatures, and smaller radii at high temperatures. 
+%
+%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
+% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
+%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
+%
+%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
+%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
+%in the Hard-Sphere fluid. 
+%This was done by making the radius of the spheres temperature dependent, 
+%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
+%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
+%
+%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
+%when they collide at higher temperatures.
+%The different size radii coorespond to the different depths to which two soft 
+%spheres would penetrate into each other when they collide. 
+%This reflects the idea that soft-spheres penetrate more deeply into each other when 
+%they collide at higher temperatures. 
+In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
+radii to model spheres with the more realistic softly fading repulsive portion of the 
+Leonard-Jones potential~\cite{andersen1971relationship}, which became known as the 
+Weeks-Chandler-Anderson model fluid.
+
+While models for particular fluids were being created, powerful methods to describe the properties 
+of fluids in general were also being developed. In the late 1950s, Scaled Particle 
+Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
+an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
+obtained~\cite{Hansen}. 
 %Percus and Yevick took another approach to the Hard-Sphere fluid.
-In 1961 Morita and Hiroike introduced a density functional approach to finding an expression for the 
-excess Helmholtz free energy of inhomogeneous fluids~\cite{MoritaDFT}.
-% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
-Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
-In 1967 %the late 1960s, and early 1970s, 
-Barker and Henderson~\cite{barker1967perturbation} adapted perturbation methods to use with fluids composed of 
-spheres with softly fading potential functions rather than the hard edges of the spheres 
-in the Hard-Sphere fluid. This was done by making the radius of the spheres temperature dependent, 
-so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
-and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
-%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other when they collide at higher temperatures.
-%The different size radii coorespond to the different depths to which two soft spheres would penetrate into each other when they collide. 
-This reflects the idea that soft-spheres penetrate more deeply into each other when they collide at higher temperatures. 
-Weeks, Chandler, and Anderson studied the soft repulsive portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-which became known as the Weeks-Chandler-Anderson model fluid.
-
-DFT Evans 1979
-
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} which provides an expression for the excess 
-Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
-Tarazona later improved on the method by introducing a tensor version of the functional~\cite{tarazonaphi3}. 
-In 1999 Schmidt adopted FMT for soft spheres forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. In this paper, 
-Density Functional Theory will be used to predict the freezing of a Weeks-Chandler-Anderson model fluid 
-by applying SFMT with Schmidt's proposed erf model\cite{schmidt2000fluid}. %for generating the weight funcions, 
+In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy, 
+or the equilibrium Helmholtz free energy in the case of a canonical ensemble, and the cooresponding equilibrium 
+number density profile for inhomogeneous fluids. To use this approach, a functional must be 
+constructed for a particular fluid.
+%DFT Evans 1979
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+to construct a functional for 
+%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
+the Hard-Sphere fluid. 
+Tarazona later improved on the method by introducing a tensor version of the 
+functional~\cite{tarazonaphi3}, and in 1999 Schmidt adopted FMT for soft spheres 
+forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
+
+In this paper, Density Functional Theory (DFT) will be used to predict the freezing of a 
+Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
+erf model\cite{schmidt2000fluid} to form the functional. %for generating the weight funcions, 
 %and Tarazona's tensor functional. 

### 2020-10-17 papers/thesis-kirstie/thesis.tex
more edits on intro

@@
 an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
-obtained~\cite{Hansen}. 
+obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an alternative method...OZ... \color{black}
 %Percus and Yevick took another approach to the Hard-Sphere fluid.
-In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy, 
-or the equilibrium Helmholtz free energy in the case of a canonical ensemble, and the cooresponding equilibrium 
+Led by the consistency bewteen results from SPT and Percus-Yevick, in 1961 Morita and Hiroike 
+introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
+(the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
 number density profile for inhomogeneous fluids. To use this approach, a functional must be 
@@
 In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-to construct a functional for 
+(which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 
 %which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
@@
 
-In this paper, Density Functional Theory (DFT) will be used to predict the freezing of a 
+Schmidt applied his Soft Fundamental Measure Theory to various fluids with soft repulsive 
+and soft attractive potentials, and developed an erf model for 
+spheres with a steep, short-range soft repulsive potential~\cite{schmidt2000fluid}. In this paper, 
+Density Functional Theory (DFT) will be used to predict the freezing of a 
 Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
-erf model\cite{schmidt2000fluid} to form the functional. %for generating the weight funcions, 
-%and Tarazona's tensor functional. 
+erf model to form the functional for the excess Helmholtz free energy. 
+%for generating the weight funcions, and Tarazona's tensor functional. 
 Results will be compared to those generated from Monte-Carlo simulations.

### 2020-10-17 papers/thesis-kirstie/thesis.tex
more updates

@@
 radii to model spheres with the more realistic softly fading repulsive portion of the 
-Leonard-Jones potential~\cite{andersen1971relationship}, which became known as the 
+Leonard-Jones potential~\cite{andersen1971relationship}, which is known as the 
 Weeks-Chandler-Anderson model fluid.
@@
 Tarazona later improved on the method by introducing a tensor version of the 
-functional~\cite{tarazonaphi3}, and in 1999 Schmidt adopted FMT for soft spheres 
+functional~\cite{tarazonaphi3, santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
 forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
@@
 and soft attractive potentials, and developed an erf model for 
-spheres with a steep, short-range soft repulsive potential~\cite{schmidt2000fluid}. In this paper, 
+spheres with steep, short-range soft repulsive potentials~\cite{schmidt2000fluid}. In this paper, 
 Density Functional Theory (DFT) will be used to predict the freezing of a 
@@
 
-Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions $w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis.
-In three dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$.
-As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, $w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving the shell.
-The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid sphere up to a radius of ?FIX?
-after which it tapers off over a distance of ?FIX?.
+Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions 
+$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+sphere up to a radius of ?FIX? after which it tapers off over a distance of ?FIX?.
 
@@
     \includegraphics[height=6cm]{weight_functions.pdf}
-    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the z-axis for a reduced temperature of 2}
+    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2}
     \label{fig:weight_functions}
@@
     \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions.
-(a) $w_2$ is a spherical shell in 3D with soft edges.
-Weight functions $w_0$ and $w_1$ have a similar shape.
-(b) $w_3$ is a sphere with soft edges in 3D.}
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
     \label{fig:W0andW3}
@@
     % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
-    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional view of $w_2$ centered at $\vec{r}'$ is shown in purple.
-Here, the weight function is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ is generated randomly according to a Gaussian distribution.}
+    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    is generated randomly according to a Gaussian distribution.} 
   \label{fig:GaussandW2_actual}
@@
     % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  and fixed in place as $\vec{dr}$ varies randomly.}
+    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
+    and fixed in place as $\vec{dr}$ varies randomly.} 
   \label{fig:GaussandW2_thinkas}
@@
 
-The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in Figure \ref{fig:alphaXivsT}.
-$\Xi$ is related to the width of the functions which become wider in general as the temperature increases.
-The radial distance to the peak decreases as the temperature increases, that is, the shells shrink, as does the sphere.
-The functions can be treated as neglible outside of a distance of about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!!
-from the center of the weight function.
+The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in 
+Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which 
+become wider in general as the temperature increases. The radial distance to the 
+peak decreases as the temperature increases, that is, the shells shrink, as does 
+the sphere. The functions can be treated as neglible outside of a distance of 
+about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
+the weight function.
 %\[{}\]
@@
 
-Phase diagrams for the WCA fluid are shown below. In all cases $\sigma=1$ and $\epsilon=1$.
+Phase diagrams for the WCA fluid are shown in Figures \ref{fig:Phase_Diagram_of_T_vs_n} 
+and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
+In Figure \ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
+is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
+The red area is fluid, and the blue area is crystal. The gray area that separates the fluid, 
+and the crystal regions is where some fluid and crystal exist together at varying porportions. 
+The region of coexistence slants slightly to the right toward higher density as the temperature 
+increases. This is expected since at higher temperatures, the fluid would have to be under 
+greater compression before forming a solid. The white lines show results from Monte-Carlo 
+simulations of the WCA fluid. Overall, the general shape, and position of the lines correspond
+fairly closely to the edges of the region of coexistance predicted from the DFT method, 
+but the width of the coexistance region is about one-third larger for the Monte-Carlo results 
+for most temperatures. The white line on the right, representing the density of the solid at 
+each temperature, is in fairly good aggreement with our DFT results for all temperatures. 
+However, the white line on the left, representing the density of the liquid at each temperature, 
+is consistenty farther left than the liquid transition densities our method predicted.
+
+Figure \ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
+plotted against the reduced temperature for temperatures of kT = 0.5 to 3.0. Again, the red area 
+is fluid, and the blue area is crystal. The general trend of the values for the transition 
+temperatures and pressures is what would be expected, with the WCA fluid becoming a solid at
+a higher pressure for a higher temperatuer, or becoming a liquid at a higher temperature 
+for a higher pressure. But the transition pressures and temperatures are consistently higher 
+than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
+the temperature increases.
 
@@
   %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
-\caption{Phase Diagram of the reduced termperature verses the reduced density. Green lines indicate the region of coexistence from Monte-Carlo simulations.}
+  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
+  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  White lines show the liquid and solid transition densities at each temperature from 
+  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
   \label{fig:Phase_Diagram_of_T_vs_n}
@@
   %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
-\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The green line marks the boundary from Monte-Carlo simulations.}
+  \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
+  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
+  The green line marks the boundary from Monte-Carlo simulations.}
   \label{fig:Phase_Diagram_P_vs_T}
@@
 
+%Plots of
+
 \begin{figure}
- \begin{center}
+  \centering
   %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
- \end{center}
-\caption{Phase Diagram of the reduced temperature verses the reduced density. The grey region indicates coexistence.}
+  \caption{Plot of the reduced temperature verses the reduced density. 
+  The grey region indicates coexistence.}
   \label{fig:T-vs-n_at_fixed_P}
@@
 \begin{figure}
- \begin{center}
+  \centering
   %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
- \end{center}
-\caption{Phase Diagram of the reduced pressure verses the reduced temperature. The lines bending downward indicate freezing.}
+  \caption{Plot of the reduced pressure verses the reduced temperature. 
+  The lines bending downward indicate freezing.}
   \label{fig:p-vs-T_at_fixed_density}
@@
 \begin{figure}
- \begin{center}
+  \centering
   %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
- \end{center}
-\caption{Phase Diagram of the reduced pressure verses the reduced volume. The grey region indicates coexistence.}
+  \caption{Plot of the reduced pressure verses the reduced volume. 
+  The grey region indicates coexistence.}
   \label{fig:p-vs-V_at_fixed_T}
@@
 \begin{figure}
- \begin{center}
+  \centering
   %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
- \end{center}
-\caption{Phase Diagram of the reduced pressure verses the reduced density. The grey region indicates coexistence.}
+  \caption{Plot of the reduced pressure verses the reduced density. 
+  The grey region indicates coexistence.}
   \label{fig:p-vs-n_at_fixed_T}
@@
 
-\cite{rosenfeld1989}
-\cite{tarazona2000}
-\cite{schmidt1999density}
-\cite{schmidt2000fluid}
-
-\cite{santos2012phi3}
-
-\cite{TschoppBHextended}
+  %\cite{TschoppBHextended}
 
 	%1-21	
-\cite{lucretius}
-\cite{rowlinson1964statistical}
-\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
-  lafitte2013accurate}
-\cite{carnahan1969equation}
-\cite{rosenfeld1989, rosenfeld1997,
-  roth2002whitebear}
-\cite{rosenfeld1989}
+	%\cite{lucretius}
+%\cite{rowlinson1964statistical}
+%\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
+%  lafitte2013accurate}
+%\cite{carnahan1969equation}
+%\cite{rosenfeld1989, rosenfeld1997,
+%  roth2002whitebear}
+%\cite{rosenfeld1989}
 	%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
-\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
+%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
 	%\cite{schmidt1999density}
-\cite{schmidt2000density, groh2001density, kim2001adsorption,
-  sweatman2002fundamental}
+%\cite{schmidt2000density, groh2001density, kim2001adsorption,
+%  sweatman2002fundamental}
 	%\cite{schmidt2000fluid, sweatman2002fundamental}
-\cite{weeks1971}
+%\cite{weeks1971}
 	%\cite{rosenfeld1989}
@@
 	%a Face Centered Cubic (FCC) lattice is formed~\cite{reference!}
-\cite{barker1967perturbation}
-\cite{roth2002whitebear} 
+%\cite{barker1967perturbation}
 %\cite{roth2002whitebear} 
-\cite{mikolaj2004structure, eggert2002quantitative,
-yarnell1973structure}
-\cite{verlet1967computer}
+	%\cite{roth2002whitebear}
+%\cite{mikolaj2004structure, eggert2002quantitative,
+%yarnell1973structure}
+	%\cite{verlet1967computer}
 

### 2020-10-20 papers/thesis-kirstie/thesis.tex
fix quote marks

@@
 In developing a model for a fluid, simplifications can be made regarding the composition, type, shape, and interatomic interaction of the molecules.
-Rather than dealing with any particular molecule or atom, the model only deals with generic "particles", which may also be referred to as "atoms".
+Rather than dealing with any particular molecule or atom, the model only deals with generic ``particles", which may also be referred to as ``atoms''.
 The shape of the particles, or atoms will be the simplest shape possible - a sphere.
@@
 This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise.
-Because the WCA spheres have no hard boundary in their interatomic potential, they are called "soft spheres" and their gradually fading potential makes them slightly "squishy".
+Because the WCA spheres have no hard boundary in their interatomic potential, they are called ``soft spheres'' and their gradually fading potential makes them slightly ``squishy''.
 
@@
 This phase diagram is representative of some arbitrary, unknown substance composed of atoms, or molecules which exibit both attraction and repulsion.
-However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term "liquid" or "gas" is not used, but simply "fluid".
+However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term ``liquid" or ``gas" is not used, but simply ``fluid".
 Points along the line that separates the solid and the fluid phases indicate a phase transition, and show that the solid and the fluid both have the same temperature and the same pressure at a phase transistion.
@@
 The minimization can be illustrated by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
-Here "surroundings" means the entire universe minus the system.
+Here ``surroundings" means the entire universe minus the system.
 This, together with the conservation of energy (the first law of thermodynamics) which says that energy lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that changes in the free energy will be negative for spontaneous processes when the natural variables of the free energy are held fixed.
@@
 The Helmholtz free energy is minimized for a fixed volume (and temperature and number of particles).
-A fixed cubical volume called the "bold-box" volume, $V_{bold-box}$ will be used for illustration.
+A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will be used for illustration.
 The number density n for a collection of simplified cells is simply 1 atom per fixed bold-box volume.
-\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per "box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density).
-Reduced number density is given by n* and is equal to the number of atoms in a "box" of size $\sigma^3$.
+\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per ``box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density).
+Reduced number density is given by n* and is equal to the number of atoms in a ``box" of size $\sigma^3$.
 \begin{displaymath}{n*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
@@
 
-It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per "bold box" volume.
+It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per ``bold box" volume.
 
@@
 Once again, it is important to note that the number density is the same for all these number density profiles $n(\vec{r})$.
-There is still a total of one atom per "bold box" volume.
+There is still a total of one atom per ``bold box" volume.
 This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.
@@
 
-%------------------n(r)--move to "back" later---------------------
+%------------------n(r)--move to ``back" later---------------------
 ------------------------------------------------------------------
@@
 %The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
-Here "surroundings" means the entire universe minus the system.
+Here ``surroundings" means the entire universe minus the system.
 This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.

### 2020-10-20 papers/thesis-kirstie/thesis.tex
fix problematic uses of "*"

@@
 \begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per ``box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density).
-Reduced number density is given by n* and is equal to the number of atoms in a ``box" of size $\sigma^3$.
-\begin{displaymath}{n*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
+Reduced number density is given by $n^*$ and is equal to the number of atoms in a ``box" of size $\sigma^3$.
+\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
 \begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}
@@
 A simplified flowchart of the program created to investigate the freezing of a WCA fluid is shown on pages x-x.
-The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.The number density profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found for that combination of n* and t*, which is known from cDFT to be unique.
+The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.The number density profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found for that combination of $n^*$ and $t^*$, which is known from cDFT to be unique.
 The process is then repeated to find the free energies at other number densities and temperatures.
 
-In general, if the crystal free energy at some n* and t* is lower than the homogeneous free energy of a liquid at that same n* and t*, then the crystaline state dominates.
+In general, if the crystal free energy at some $n^*$ and $t^*$ is lower than the homogeneous free energy of a liquid at that same $n^*$ and $t^*$, then the crystaline state dominates.
 But if the crystal free energy is higher than the homogeneous free energy, the liquid state dominates.
-However, there is a region of densities between the liquid and crystal states that are forbidden, and so the actual state can not be determined just by which energy is lower at that particular n* and t*.
+However, there is a region of densities between the liquid and crystal states that are forbidden, and so the actual state can not be determined just by which energy is lower at that particular $n^*$ and $t^*$.
 Maxwell's Double Tangent Construction must be implemented using the date collected to determine the actual crystal and liquid states, and construct phase diagrams.
@@
 
-The Liquid (homogeneous) ideal free energy $hF_{ideal}$ is computed analytically for a given n* and t*, where $\sigma=1$ so that n*=n$\sigma^3$=n, and $\epsilon=1$ so that t*=$\frac{k_BT}{\epsilon}=k_BT$.
+The Liquid (homogeneous) ideal free energy $hF_{ideal}$ is computed analytically for a given $n^*$ and $t^*$, where $\sigma=1$ so that $n^*=n\sigma^3$=n, and $\epsilon=1$ so that $t^*=\frac{k_BT}{\epsilon}=k_BT$.
 %\begin{displaymath}{\frac{hF_{ideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
@@
 
-The Liquid (homogeneous) excess free energy $hF_{excess}$ is computed analytically for a given n* and t* using SFMT with a uniform number density profile that does not depend on $\vec r$. 
+The Liquid (homogeneous) excess free energy $hF_{excess}$ is computed analytically for a given $n^*$ and $t^*$ using SFMT with a uniform number density profile that does not depend on $\vec r$. 
 %done in hf.energy() note- does not depend on fv or gw but does use SFMT):
@@
 Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_0(k)$ becomes
-\begin{equation}{\widetilde{w}_0(k)=\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+\begin{equation}{\widetilde{w}_0(k)=\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 
@@
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath}
-\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 \color{black}
@@
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{k^2\Xi^2}{8}\right)}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\sin(\frac{k\alpha}{2})}\end{displaymath}
-\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 \[{}\]
@@
 %\color{red}
-\begin{equation}{\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+\begin{equation}{\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 \begin{displaymath}{\approx\frac{\left(2\pi\right)^\frac{2}{3}}{\Xi{k}^3}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[2\operatorname{Re}\left(\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)\right)\right]}\end{displaymath}

### 2020-10-20 papers/thesis-kirstie/thesis.tex
make some latex more readable by introducing newlines

@@
 Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-\begin{equation}df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}\end{equation}
+\begin{equation}
+  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
+\end{equation}
 and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
-\begin{equation}df_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
+\begin{equation}
+  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
+\end{equation}
 where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
-\begin{equation}df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
+\begin{equation}
+  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+\end{equation}
+which, for fixed N, is the same as
 \begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative, and thus the free energy tends toward a minimum.

### 2020-10-20 papers/thesis-kirstie/thesis.tex
fix misspelling
another edit

@@
 %Percus and Yevick took another approach to the Hard-Sphere fluid.
-Led by the consistency bewteen results from SPT and Percus-Yevick, in 1961 Morita and Hiroike 
-introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
+In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
 (the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
@@
 %DFT Evans 1979
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+Led by the consistency bewteen results from SPT and Percus-Yevick, in 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
 (which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 

### 2020-10-20 papers/thesis-kirstie/thesis.tex
some more edits

@@
 
-\noindent I. Introduction
+\section I. Chapter 1 - Introduction
 
@@
 \newline
-\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)
+\section{Chapter 2 - Theory}
+\subsection{Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)}
 
@@
     \centering
-    %\includegraphics[height=3.5cm]{TwoSpheres.pdf} 
+    \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
     \caption{(Left) The WCA potential is a minimum at r=2R. 
@@
 The WCA potential is not well suited for implementing Density Functional Theory.  
-Therefore, an erf potential which closely approximates the WCA potential at temperatures of interest will be used instead of the WCA potential.
-The erf potential is given by
+Therefore, an erf potential which closely approximates the WCA potential at temperatures 
+of interest will be used instead of the WCA potential. The erf potential is given by 
 \begin{equation}
@@
 \end{equation} 
-where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ 
+where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
+potentail when r equals $\alpha$ 
 \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
@@
 \end{equation}
-The parameter $\Xi$ is derived by matching the slope of the erf potential to the slope of the WCA potential at r=$\alpha$:
-\begin{displaymath}{\frac{dV_{erf}(r=\alpha)}{dr}=\frac{dV_{WCA}(r=\alpha)}{dr}}\end{displaymath} 
-and is given by: 
-\begin{equation}
-	{\Xi=\frac{\alpha}{6\sqrt{\pi}\left(\sqrt{\frac{\epsilon}{k_BT}\ln(2)}+\ln(2)\right)}}
-\end{equation}
-
-\noindent These conditions, along with V$_{erf}=$V$_{WCA}$ at $r=\alpha$, help fit the curve of the erf potential to the curve of the WCA potential.
+
+%Deleted old Xi method
+%move 3-9 to methods
 
@@
 \[\]
-\noindent IIB. The Theory of Freezing and Maxwell's Double Tangent Construction
+\subsection{The Theory of Freezing and Maxwell's Double Tangent Construction}
 
@@
 \[\]
-\noindent IIC. Classical Density Functional Theory (cDFT)
+\subsection{Classical Density Functional Theory (cDFT)}
 
@@
 \[{}\]
-III. METHODS
+\section{Chapter 3 - Methods}
 
@@
 
-\noindent IV. Results and Discussion
+\section{Chapter 4 - Results and Discussion}
 
@@
 
-\noindent V. Conclusion
+\section{Chapter 5 -  Conclusion}
 

### 2020-10-20 papers/thesis-kirstie/thesis.tex
switch to beavtex for thesis-kirstie

--- 2020-10-20 papers/thesis-kirstie/thesis.tex
@@
-\documentclass[12pt]{article}
-\usepackage{setspace}
-\doublespacing
-%\usepackage[top=1.2in, bottom=1.2in]{geometry}
-%\usepackage[top=1.5in]{geometry}
-\usepackage[bottom=1.2in]{geometry}
-\usepackage{fancyhdr}
-\usepackage{yhmath}
-\pagestyle{fancy}
-\fancyhf{}
-\renewcommand{\headrulewidth}{0pt}
-\fancyhead[R]{\thepage}
+\documentclass[double,12pt]{beavtex}
 \usepackage{graphicx}
-\usepackage{subcaption}
-\usepackage{color}
-\usepackage{enumitem}
+\usepackage{rotating} %Package added to allow the rotation of figures and chart on a page, {sidewaysfigure} command
+\usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
 \title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
 \author{Kirstie L. Finster}
+\degree{Master of Science}
+\doctype{Thesis}
+\department{Physics}
+\depttype{Department}
+\depthead{Head}
+\major{Physics}
+\advisor{David Roundy}
+\submitdate{October ??, 2020}
+\commencementyear{2020}
+
+\usepackage{color}
+\usepackage{amsmath}
+\usepackage{subcaption}
+
 \date{August 2020}
+
+\abstract{
+  This is a test
+}
+
 \begin{document}
    \maketitle
-   \thispagestyle{fancy}
-
-\noindent I. Introduction
+   \mainmatter
+
+\chapter{Introduction}
 
@@
 %\[{}\]
-\newline
-\newline
-\noindent IIA. Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)
+
+\section{Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)}
+
+\subsection{Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)}
 
@@
 \end{equation} 
-where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potentail when r equals $\alpha$ 
+where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potential when r equals $\alpha$ 
 \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
@@
 and the weights are as usual 
-\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
-\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
-\begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation}
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
+\begin{align}
+  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
+  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
+  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
+  \\
+  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
+\end{align}
 %\begin{displaymath}{w_0=?}\end{displaymath} 
@@
 %\begin{displaymath}{w_3=\alpha(\pi\Xi^2 + 0.52...\alpha^2)}\end{displaymath} \[\]
-\[\]
 
@@
 with the varying number density profile given by 
-\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}}\end{displaymath}
-\newline $\bullet$ For small gw values it is assumed that the Guassians do not overlap, in which case the equation for the free energy can be simplified and solved analytically with the result
-\begin{displaymath}{cF_{ideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)}\end{displaymath}
-
-%\begin{displaymath}{cF_{ideal}=(1-f_v)k_BT\left(\ln(\frac{(1-fv)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)-DELETE-EQN!}\end{displaymath}
-
-$\bullet$ For larger gw values it is assumed that the Gaussians do overlap, and so the free energy is caluculated computationally. First dx is scaled by gw
-\begin{displaymath}{dx_{scaled}=(dx_{input})(gw)}\end{displaymath}
+\begin{displaymath}
+  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}}
+\end{displaymath}
+\begin{itemize}
+  \item For small gw values it is assumed that the Guassians do not overlap, in
+  which case the equation for the free energy can be simplified and solved
+  analytically with the result
+  \begin{align}
+    cF_{ideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
+  \end{align}
+\item For larger gw values it is assumed that the Gaussians do overlap, and so the free energy is caluculated computationally. First dx is scaled by gw
+\begin{align}
+  dx_{scaled} &= (dx_{input})(gw)
+\end{align}
 then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the primitive cell taking into account the contribution of Gaussians from many (5x5x5) nearby cells whose centers are given by $R_i$.
 %\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
-The contribution to the free energy of $n(\vec{r})$ at the one position is then added to the accumulating sum.
+The contribution to the free energy of $n(\vec{r})$ at the one position is then
+added to the accumulating sum.
 %\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
 %\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
-This process is repeated for all positions within one primitive cell to get the crystal ideal free energy for the primitive cell.
-This energy is then divided by the volume of the primitive cell to get the crystal ideal free energy per volume.
+This process is repeated for all positions within one primitive cell to get the
+crystal ideal free energy for the primitive cell. This energy is then divided by
+the volume of the primitive cell to get the crystal ideal free energy per
+volume.
 %\[\]
+\end{itemize}
 
@@
 
-APPENDIX
+\backmatter
+
+\chapter{Appendix}
 \begin{enumerate}
@@
 
-\textbf{1a. Fourier Transform of Weight Function $w_0(r)$:}
+\section{Fourier Transform of Weight Function $w_0(r)$:}
 \begin{equation}{w_0(r)=\frac{w_2(r)}{4{\pi}r^2}}\end{equation}
@@
 %\[{}\]
-\textbf{1b. Fourier Transform of Weight Function $w_{1}(r)$:}
+\section{Fourier Transform of Weight Function $w_{1}(r)$:}
 \begin{equation}{w_1(r)=\frac{w_2(r)}{4{\pi}r}}\end{equation}
@@
 %\[{}\]
-\textbf{1c. Fourier Transform of Weight Function $w_{2}(r)$:}
+\section{Fourier Transform of Weight Function $w_{2}(r)$:}
 \begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
@@
 %\[{}\]
-\textbf{1d. Fourier Transform of Weight Function $w_{3}(r)$:}
+\section{Fourier Transform of Weight Function $w_{3}(r)$:}
 \begin{equation}{w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]}\end{equation}
@@
 
-\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{k}\left[\left(1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right)\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\bigg|^{\infty}_0\color{white}\right]}\end{equation} \color{black}
-\begin{displaymath}{-\int_{0}^{\infty}\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\left(\frac{2}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\right)d{r}}\end{displaymath}
+\begin{multline}
+  \widetilde{w}_3(k)=\frac{2\pi}{k}\Bigg[\left(1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right)\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\bigg|^{\infty}_0
+   \\
+  -\int_{0}^{\infty}\left(\frac{\sin(kr)-kr\cos(kr)}{k^2}\right)\left(\frac{2}{a\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\right)dr\Bigg]
+\end{multline}
 Evaluating the integral by doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$, and setting the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero this becomes
@@
 %\[{}\]
-\textbf{2a. Fourier Transform of Vector Weight Function $\vec{w}_{1}(r)$:}
+\section{Fourier Transform of Vector Weight Function $\vec{w}_{1}(r)$:}
 \begin{equation}{\vec{w}_1(\vec{r})=w_1(r)\frac{\vec{r}}{r}}\end{equation}
@@
 %\[{}\]
-\textbf{2b. Fourier Transform of Vector Weight Function $\vec{w}_{2}(r)$:}
+\section{Fourier Transform of Vector Weight Function $\vec{w}_{2}(r)$:}
 \begin{equation}{\vec{w}_2(\vec{r})=w_2(r)\frac{\vec{r}}{r}}\end{equation}
@@
 \begin{displaymath}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~~}\hat{z}'}\end{displaymath}
-\[{}\]
+
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
@@
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
-\[{}\]
-\textbf{3. Fourier Transform of Tensor Weight $\overleftrightarrow{w}_{m2}(r)$:}
+
+\section{Fourier Transform of Tensor Weight $w_{m2}$}
 \begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation}
@@
 \bibliography{thesis}
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 

### 2020-10-20 papers/thesis-kirstie/thesis.tex

@@
 Results will be compared to those generated from Monte-Carlo simulations.
-%\[{}\]
+
+\chapter{Theory}
 
 \section{Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)}
-
-\subsection{Model Fluids and the Weeks-Chandler-Anderson Fluid (WCA)}
 
@@
   \end{figure}
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
 
@@
 
-%\[{}\]
-%\[{}\]
-
 \begin{figure}[h!]
@@
   \end{figure}
-
-%\[{}\]
-%\[{}\]
 
@@
 \end{equation} 
-where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA potential when r equals $\alpha$ 
+where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
+potentail when r equals $\alpha$ 
 \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 

### 2020-10-20 papers/thesis-kirstie/thesis.tex
more figures and edits

@@
 
-%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres.
-Thus, the WCA fluid can be treated as a classical fluid.
-
-\[\]
-\subsection{The Theory of Freezing and Maxwell's Double Tangent Construction}
+%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
+%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
+%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
+%classical fluid.
+
+\section{The Theory of Freezing and Maxwell's Double Tangent Construction}
 
@@
     \centering
-    % FIXME \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
+    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
     \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
@@
     \centering
-    % FIXME \includegraphics[height=5cm]{P-T_Diagram.png}
+    \includegraphics[height=5cm]{P-T_Diagram.png}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
     \centering
-     %\includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
+     \includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
      % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
@@
     \centering
-     %\includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
-     %FIXME\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
+     \includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
+     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
     \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
@@
 
-Once the pressure is known for the transition point, the density of the liquid at the transition point can be found from a plot of the pressure versus the inverse of the density for the liquid.
-The density of the solid at the transition point can be found the same way, as shown in Figure \ref{fig:Pvsinvn}.
-%\[\]
-
-\begin{figure}[h!]
-    \centering
-    %\includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
-    % FIXME\includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
+Once the pressure is known for the transition point, the density of the liquid at the 
+transition point can be found from a plot of the pressure versus the inverse of the 
+density for the liquid. The density of the solid at the transition point can be found 
+the same way, as shown in Figure \ref{fig:Pvsinvn}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
+    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
     \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
@@
     \centering
-    % FIXME \includegraphics[height=4cm]{T-n_Diagram.png}
+    \includegraphics[height=4cm]{T-n_Diagram.png}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion.
@@
 Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in.
-%\[\]
-\[\]
-\subsection{Classical Density Functional Theory (cDFT)}
+
+\section{Classical Density Functional Theory (cDFT)}
 
@@
 
-\[{}\]
-\section{Chapter 3 - Methods}
+\chapter{Methods}
 
@@
 
-\section{Chapter 4 - Results and Discussion}
+\chapter{Results and Discussion}
 
@@
 
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
-
-\section{Chapter 5 -  Conclusion}
-
-\[{}\]
-\[{}\]
-\[{}\]
-\[{}\]
+\chapter{Conclusion}
 
@@
 \chapter{Appendix}
-\begin{enumerate}
-
-\item Fourier Transforms of Weight Functions $w_0(r), w_1(r), w_2(r), w_3(r)$
-\item Fourier Transforms of Vector Weight Functions $\vec{w}_1(\vec{r}), \vec{w}_2(\vec{r})$  
-\item Fourier Transform of Tensor Weight $\overleftrightarrow{w}_{m2}(r)$ in coordinate system (x',y',z')
-%\item The tensor weight components in the coordinate system (x,y,z) 
-%\item Show transform of tensor weight $\overleftrightarrow{w}_{m2}(r)$ goes to zero as k goes to 0
-\item Integrals used to Compute Fourier Transforms of Weight Functions
-\item Deriviation of Tensor Density $n_{m2ij}(\vec{r})$ 
-\item Deriviation of the Tensor version of $\Phi_3$ 
-\item Second Virial Coefficient $B_{2}$ for the Error Function Potential 
-\item Third Virial Coefficient $B_{3}$
-\item Derivation of the Virial Equation
-\end{enumerate}
-
-\section{Fourier Transform of Weight Function $w_0(r)$:}
+
+\section{Fourier Transform of Weight Function $w_0(r)$}
 \begin{equation}{w_0(r)=\frac{w_2(r)}{4{\pi}r^2}}\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-\section{Fourier Transform of Weight Function $w_{1}(r)$:}
+\section{Fourier Transform of Weight Function $w_{1}(r)$}
 \begin{equation}{w_1(r)=\frac{w_2(r)}{4{\pi}r}}\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-\section{Fourier Transform of Weight Function $w_{2}(r)$:}
+\section{Fourier Transform of Weight Function $w_{2}(r)$}
 \begin{equation}{w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)^2}}\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-\section{Fourier Transform of Weight Function $w_{3}(r)$:}
+\section{Fourier Transform of Weight Function $w_{3}(r)$}
 \begin{equation}{w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]}\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-\section{Fourier Transform of Vector Weight Function $\vec{w}_{1}(r)$:}
+\section{Fourier Transform of Vector Weight Function $\vec{w}_{1}(r)$}
 \begin{equation}{\vec{w}_1(\vec{r})=w_1(r)\frac{\vec{r}}{r}}\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-\section{Fourier Transform of Vector Weight Function $\vec{w}_{2}(r)$:}
+\section{Fourier Transform of Vector Weight Function $\vec{w}_{2}(r)$}
 \begin{equation}{\vec{w}_2(\vec{r})=w_2(r)\frac{\vec{r}}{r}}\end{equation}
@@
 
-\[{}\]
-\[{}\]
-\[{}\]
-%\[{}\]
-
-\textbf{5. Deriviation of Tensor Density $n_{m2ij}(\vec{r})$}
+\section{Deriviation of Tensor Density $n_{m2ij}(\vec{r})$}
 \begin{equation}{\overleftrightarrow{n}_{m2}(\vec{r})=\int_{allspace}n(\vec{r'})\overleftrightarrow{w}_{m2}(\vec{r}-\vec{r'})d{\vec{r'}}}\end{equation} 
@@
   \end{align}
-\[{}\]
-\[{}\]
-%\[{}\]
-
-\textbf{6. Deriviation of the Tensor version of $\Phi_3$ }
-\[{}\]
+
+\section{Deriviation of the Tensor version of $\Phi_3$}
+
 \begin{equation}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left[\overrightarrow{\operatorname{V}}\cdot\overleftrightarrow{\operatorname{T}}\cdot\overrightarrow{\operatorname{V}}-n\overrightarrow{\operatorname{V}}\cdot\overrightarrow{\operatorname{V}}-\operatorname{Tr}\left(\overleftrightarrow{\operatorname{T}}^3\right)+n\operatorname{Tr}\left(\overleftrightarrow{\operatorname{T}}^2\right)\right]}\end{equation}
-\[{}\]
+
 \begin{equation}{\overleftrightarrow{\operatorname{T}}=\frac{{\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}}{4\pi{R}^2}}\end{equation}
-\[{}\]
+
 \begin{equation}{\overrightarrow{\operatorname{V}}=\frac{\vec{n}_2}{4\pi{R}^2}}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}
 \begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2\vec{n}_{v2}\cdot\vec{n}_{v2}-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left((\overleftrightarrow{n}_{m2}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2|\vec{n}_{v2}|^2-\operatorname{Tr}({\overleftrightarrow{n}_{m2}^3})
 -n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)\color{white}\right)\color{black}}\end{displaymath} \begin{equation}{\color{white}\left(\color{black}-\frac{1}{3}{n_2}^2\operatorname{Tr}(\overleftrightarrow{n}_{m2})-\frac{1}{9}{n_2}^3+n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)+\frac{1}{3}{n_2}^3+\frac{2}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
-\[{}\]
+
 \begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
-\[{}\]
+
 \begin{equation}{\operatorname{Tr}(\overleftrightarrow{n}_{m2})=0}\end{equation}
-\[{}\]
+
 \begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2|\vec{n}_{v2}|^2+\frac{9}{2}(\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}})-\frac{9}{2}\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})}{24\pi(1-\eta)^2}}\end{equation}
-\[{}\]
+
 \begin{equation}{\eta=n_3}\end{equation}
-\[{}\]
+
 \begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}}\end{equation}
-\[{}\]
-\[{}\]
-%\[{}\]
-
-\textbf{7. Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$:}
+
+\section{Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$}
 \begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}
@@
 
-\[{}\]
-%\[{}\]
-
-\textbf{8. Third Virial Coefficient $B_{3}$:}
+\section{Third Virial Coefficient $B_{3}$}
 
@@
 \begin{equation}f(\vec{r}-\vec{r'})=\int_{allspace}f(\vec{r''})\delta(\vec{r}''-(\vec{r}-\vec{r}'))d\vec r'' \end{equation}
-
 where
@@
 
-\[{}\]
-%\[{}\]
-
-\noindent \textbf{9. Derivation of the Virial Equation}
+\section{Derivation of the Virial Equation}
 The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
@@
 
-\[{}\]
-%\[{}\]
-
-\noindent \textbf{10. Derivation of the Excess Free Energy from Scaled Particle Theory}
+\section{Derivation of the Excess Free Energy from Scaled Particle Theory}
 The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R 

### 2020-10-22 papers/thesis-kirstie/thesis.tex
updates and figures

@@
 \abstract{
-  This is a test
+  We use Classical Density Functional Theory (cDFT) to predict the 
+  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
+  consists of soft spheres that exhibit the repulsive portion of the 
+  Leonard-Jones potential widely used to model the interaction of two 
+  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
+  Theory (SFMT) with Tarazona's tensor expression for Phi3, and Schmidt's 
+  erf model in which we have incorporated our temperature dependent 
+  parameters to fit the results to those predicted by Monte-Carlo simulations. 
+  Our functional was successful in predicting the freezing of the WCA fluid 
+  with a fairly good match to Monte-Carlo simulations. This result suggests 
+  that SFMT may provide a solid foundation for future density functionals 
+  that will model real fluids.
 }
@@
 
-Classically, fluids have been modeled by a collection of small, impenetrable spheres colliding 
-with one another like billiard balls. In the simplest case, the spheres are treated as having 
-only a replulsive potential. This model, called the Hard-Sphere fluid, has served as the primary 
-reference system for fluids. This is due to the fact that the repulsion between atoms is much 
-stronger than their attraction, and so the repulsive potential, which is descriptive of 
-hard spheres, dominates the interactions between atoms. 
-In 1954 Zwanzig introduced the Square-well fluid which is a perturbation of the Hard-Sphere fluid 
-created by adding a hard-edge attractive potential to the repulsive potential of the Hard-Sphere 
-fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied spheres with softly fading replusive potentials which he 
-modeled by making the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
-This reflects the idea that soft-spheres penetrate more deeply into each other 
-when they collide at higher temperatures. 
-%At higher temperatures the spheres would penetrate more deeply into each other when they collide, 
-%and thus be represented by spheres with smaller radii. 
+Classically, fluids have been modeled by a collection of small, 
+impenetrable spheres colliding with one another like billiard balls. 
+In the simplest case, the spheres are treated as having only a replulsive 
+potential. This model, called the Hard-Sphere fluid, has served as the 
+primary reference system for fluids. This is due to the fact that the 
+repulsion between atoms is much stronger than their attraction, and so 
+the repulsive potential, which is descriptive of hard spheres, dominates 
+the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
+fluid which is a perturbation of the Hard-Sphere fluid created by adding 
+a hard-edge attractive potential to the repulsive potential of the 
+Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
+spheres with softly fading replusive potentials which he modeled by making
+the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+This reflects the idea that soft-spheres penetrate more deeply into 
+each other when they collide at higher temperatures.
+
 In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
-with Rowlinson's idea of using temperature-dependent radii to model spheres with 
-soft repulsive potentials to create the Barker-Henderson model fluid~\cite{barker1967perturbation}, 
-where a soft-sphere fluid by is modeled by using a Hard-Sphere fluid with hard spheres of 
-larger radii at low temperatures, and smaller radii at high temperatures. 
-%
-%an expression for the excess Helmholtz free energy of inhomogeneous fluids.
-% ?  Percus-Yevick applied Density Functional Theory (DFT) to find an expression for the excess 
-%Helmholtz free energy of an inhomogeneous Hard-Sphere fluid. 
-%
-%In 1967 Barker and Henderson~\cite{barker1967perturbation} adapted the perturbation method of Zwanzig 
-%to use with fluids composed of spheres with softly fading potentials rather than the hard edges of the spheres 
-%in the Hard-Sphere fluid. 
-%This was done by making the radius of the spheres temperature dependent, 
-%so that at low temperatures the Soft-Sphere fluid could be modeled by hard spheres with large radii, 
-%and at high temperatures the Soft-Sphere fluid could be modeled by hard spheres with smaller radii. 
-%
-%A larger radii cooresponds to a deeper level of penetration of two soft spheres into each other 
-%when they collide at higher temperatures.
-%The different size radii coorespond to the different depths to which two soft 
-%spheres would penetrate into each other when they collide. 
-%This reflects the idea that soft-spheres penetrate more deeply into each other when 
-%they collide at higher temperatures. 
-In 1971 Weeks, Chandler, and Anderson used temperature and density dependent hard-sphere 
-radii to model spheres with the more realistic softly fading repulsive portion of the 
-Leonard-Jones potential~\cite{andersen1971relationship}, which is known as the 
-Weeks-Chandler-Anderson model fluid.
-
-While models for particular fluids were being created, powerful methods to describe the properties 
-of fluids in general were also being developed. In the late 1950s, Scaled Particle 
-Theory (SPT)~\cite{ReissSPT} was introduced from which an equation of state, and
-an expression for the excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
-obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an alternative method...OZ... \color{black}
+with Rowlinson's idea of using temperature-dependent radii to model spheres 
+with soft repulsive potentials to create the Barker-Henderson model 
+fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
+modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
+at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
+Chandler, and Anderson used temperature and density dependent hard-sphere 
+radii to model spheres with the more realistic softly fading repulsive 
+portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
+which is known as the Weeks-Chandler-Anderson model fluid.
+
+While models for particular fluids were being created, powerful methods 
+to describe the properties of fluids in general were also being developed. 
+In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
+introduced from which an equation of state, and an expression for the 
+excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
+obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
+alternative method...OZ... \color{black}
 %Percus and Yevick took another approach to the Hard-Sphere fluid.
-In 1961 Morita and Hiroike introduced a density functional approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
-(the equilibrium Helmholtz free energy in the case of a canonical ensemble), and the cooresponding equilibrium 
-number density profile for inhomogeneous fluids. To use this approach, a functional must be 
+In 1961 Morita and Hiroike introduced a density functional 
+approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
+(the equilibrium Helmholtz free energy in the case of a canonical 
+ensemble), and the cooresponding equilibrium number density profile for 
+inhomogeneous fluids. To use this approach, a functional must be 
 constructed for a particular fluid.
 %DFT Evans 1979
-Led by the consistency bewteen results from SPT and Percus-Yevick, in 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-(which can be constructed from SPT~\cite{santos2012phi3}) to form a functional for 
-%which provides an expression for the excess Helmholtz free energy of an inhomogeneous 
-the Hard-Sphere fluid. 
-Tarazona later improved on the method by introducing a tensor version of the 
-functional~\cite{tarazonaphi3, santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
+%Led by the consistency bewteen results from SPT and Percus-Yevick, 
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
+functional for the Hard-Sphere fluid. Tarazona later improved on the 
+method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
+santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
 forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
 
-Schmidt applied his Soft Fundamental Measure Theory to various fluids with soft repulsive 
-and soft attractive potentials, and developed an erf model for 
-spheres with steep, short-range soft repulsive potentials~\cite{schmidt2000fluid}. In this paper, 
-Density Functional Theory (DFT) will be used to predict the freezing of a 
-Weeks-Chandler-Anderson model fluid by applying SFMT with Schmidt's proposed 
-erf model to form the functional for the excess Helmholtz free energy. 
+Schmidt applied his Soft Fundamental Measure Theory to various fluids 
+with soft repulsive and soft attractive potentials, and developed an erf 
+model for spheres with steep, short-range soft repulsive 
+potentials~\cite{schmidt2000fluid}. In this paper, Density Functional 
+Theory (DFT) will be used to predict the freezing of a Weeks-Chandler-Anderson 
+model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
+dependent parameters to form the functional for the excess Helmholtz free energy. 
 %for generating the weight funcions, and Tarazona's tensor functional. 
@@
 
-Real fluids and their behavior are very complex from a physicist's point of view.
-Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules, and their interaction with charges from other molecules make it difficult to study a fluid composed of water molecules directly.
-The task becomes even more daunting considering the complicated molecular motion of each water molecule, and the sheer number of molecules that must be taken into account.
-%Although a water molecule, for example, is composed of only three atoms, the ever-changing redistribution of charges within the molecules and their interaction with charges from other molecules, the complicated molecular motion of a collection of water molecules, and the sheer number of molecules that must be taken into account make it difficult to study a fluid composed of water molecules directly.
-The problem of understanding the behavoir of real liquids is better addressed by using models which reduce a complex picture to a simple one.
-Modifications, or perturbations, can then be incorporated into the model making it progressively more complex until it mimics the behavior of a real liquid as closely as possible.
-
-In developing a model for a fluid, simplifications can be made regarding the composition, type, shape, and interatomic interaction of the molecules.
-Rather than dealing with any particular molecule or atom, the model only deals with generic ``particles", which may also be referred to as ``atoms''.
-The shape of the particles, or atoms will be the simplest shape possible - a sphere.
-The spheres in the simplest models do not exhibit quantum interference effects as the mass of the spheres, and the temperature are taken to be large enough that the de Broglie wavelength is smaller than the mean distance between the spheres.
-The fluid can then be treated as a classical fluid where the spheres interact with each other according to their electrostatic interatomic potential.
-
-There are many interatomic potentials which can be used to model a fluid each of which describes the interaction between two real atoms to some degree of accuracy.
-The simplest, and oldest, model used to describe a fluid is the hard-sphere model.
-The hard-sphere model envisions a fluid to be composed of hard spheres like, for example, a collection of marbles.
-This model came about by considering the near incompressibilty of liquids indicating that the particles making up the liquid strongly resist being pushed closer together.
-This behavior can be described by an interatomic potential energy that is zero when two spheres are not in contact, but goes instantly to infinity when contact is made.
-The inablilty to push real atoms closer together is due to the repulsion of the electrons in each atom, and is the dominating interparticle electrostatic interaction.
-Thus, the hard-sphere model - which is the simplest, but still effective representation of this replusive behavior that dominates the interaction between real atoms, is currently the most widely used model for fluids.
+Real fluids and their behavior are very complex from a physicist's point 
+of view. Although a water molecule, for example, is composed of only three
+atoms, the ever-changing redistribution of charges within the molecules, 
+and their interaction with charges from other molecules make it difficult 
+to study a fluid composed of water molecules directly. The task becomes 
+even more daunting considering the complicated molecular motion of each 
+water molecule, and the sheer number of molecules that must be taken 
+into account.
+
+The problem of understanding the behavoir of real liquids is better addressed 
+by using models which reduce a complex picture to a simple one. Modifications, 
+or perturbations, can then be incorporated into the model making it 
+progressively more complex until it mimics the behavior of a real 
+liquid as closely as possible.
+
+In developing a model for a fluid, simplifications can be made regarding 
+the composition, type, shape, and interatomic interaction of the molecules. 
+Rather than dealing with any particular molecule or atom, the model only 
+deals with generic ``particles", which may also be referred to as ``atoms''. 
+The shape of the particles, or atoms will be the simplest shape possible - 
+a sphere. The spheres in the simplest models do not exhibit quantum 
+interference effects as the mass of the spheres, and the temperature 
+are taken to be large enough that the de Broglie wavelength is smaller 
+than the mean distance between the spheres. The fluid can then be treated 
+as a classical fluid where the spheres interact with each other according 
+to their electrostatic interatomic potential.
+
+There are many interatomic potentials which can be used to model a fluid 
+each of which describes the interaction between two real atoms to some 
+degree of accuracy. The simplest, and oldest, model used to describe a 
+fluid is the hard-sphere model. The hard-sphere model envisions a fluid 
+to be composed of hard spheres like, for example, a collection of marbles. 
+This model came about by considering the near incompressibilty of liquids 
+indicating that the particles making up the liquid strongly resist being 
+pushed closer together. This behavior can be described by an interatomic 
+potential energy that is zero when two spheres are not in contact, but 
+goes instantly to infinity when contact is made. The inablilty to push 
+real atoms closer together is due to the repulsion of the electrons in 
+each atom, and is the dominating interparticle electrostatic interaction. 
+Thus, the hard-sphere model - which is the simplest, but still effective 
+representation of this replusive behavior that dominates the interaction 
+between real atoms, is currently the most widely used model for fluids.
 
@@
     \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
-    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity for $r\leq{2R}$ which is approximately 1.225 in this example.}
+    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity 
+    for $r\leq{2R}$ which is approximately 1.225 in this example.}
     \label{fig:HardSphere_potential}
@@
 
-Another fairly simple, but relatively accurate interatomic potential widely used to represent generic, neutral atoms is the Leonard-Jones potential.
-The Leonard-Jones potential describes both the long-range attraction, and the short-range replusion that occurs between two nuetral atoms.
-The attraction between two nuetral atoms is due to fleeting dipoles that form spontaneously when electrons are redistributed as atoms come into close proximty.
-These are often characterized as Van der Walls, or London dispersion forces.
-The attraction grows stronger as two atoms move closer together until their atomic orbitals begin to overlap.
-%At this point the electrons in each atom begin to repel the electrons in the other atom and oppose the overlap of the two atoms.
-At this point the electrons in each atom begin to oppose the overlap of the two atoms through their replusion.
-The repulsion increases rapidly and approaches infinity as the distance between the atoms conintues to decrease.
+Another fairly simple, but relatively accurate interatomic potential widely 
+used to represent generic, neutral atoms is the Leonard-Jones potential. 
+The Leonard-Jones potential describes both the long-range attraction, 
+and the short-range replusion that occurs between two nuetral atoms. 
+The attraction between two nuetral atoms is due to fleeting dipoles that 
+form spontaneously when electrons are redistributed as atoms come into 
+close proximty. These are often characterized as Van der Walls, or London 
+dispersion forces. The attraction grows stronger as two atoms move closer 
+together until their atomic orbitals begin to overlap. 
+%At this point the electrons in each atom begin to repel the electrons 
+%in the other atom and oppose the overlap of the two atoms.
+At this point the electrons in each atom begin to oppose the overlap of 
+the two atoms through their replusion. The repulsion increases rapidly and 
+approaches infinity as the distance between the atoms conintues to decrease.
 
@@
 
-The combined attractive and repulsive effects produce the overall shape of the Leonard-Jones interatomic potential shown in Figure \ref{fig:LJ_potential}.
-The potential is zero when the distance r between the centers of two atoms is equal to $\sigma$ (shown in the figure at r/$\sigma$=1).
-At this point the attraction bewteen the atoms just cancels their repulsion.
-As the distance between the centers of the two atoms decreases below $\sigma$, the potential rapidly approaches positive infinity corresponding to infinite repulsion.
-Above $\sigma$ the potential becomes negative indicating a net attraction between the atoms.
-The net attraction is maximized at a distance of r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere given by $R={2^{-\frac{5}{6}}}\sigma$.
-At this point the depth of the potential well, given by epsilon $\epsilon$, is a maximum.
-As r increases further, the net attraction decreases and the potential approaches zero asymptotically.
-The attractive and repulsive effects can be separated into two potentials, shown with different colors in Figure \ref{fig:LJ_potential_2parts}, which when added together regenerate the full Leonard-Jones potential.
+The combined attractive and repulsive effects produce the overall shape of 
+the Leonard-Jones interatomic potential shown in Figure \ref{fig:LJ_potential}. 
+The potential is zero when the distance r between the centers of two atoms 
+is equal to $\sigma$ (shown in the figure at r/$\sigma$=1). At this point 
+the attraction bewteen the atoms just cancels their repulsion. As the distance 
+between the centers of the two atoms decreases below $\sigma$, the potential 
+rapidly approaches positive infinity corresponding to infinite repulsion. 
+Above $\sigma$ the potential becomes negative indicating a net attraction 
+between the atoms. The net attraction is maximized at a distance of 
+r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of 
+the sphere given by $R={2^{-\frac{5}{6}}}\sigma$. At this point the depth of
+the potential well, given by epsilon $\epsilon$, is a maximum. As r increases 
+further, the net attraction decreases and the potential approaches zero
+asymptotically. The attractive and repulsive effects can be separated into two
+potentials, shown with different colors in Figure \ref{fig:LJ_potential_2parts}, 
+which when added together regenerate the full Leonard-Jones potential.
 
@@
 
-The repulsive portion of the Leonard-Jones potential forms the potential for a Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}.
-The WCA potential is described by the equation: \begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}
-
-\noindent As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases from infinity as the distance between the centers of the atoms r increases, and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}.
-As with the Lenard-Jones potential, the repulsive WCA potential increases rapidly and approches infinity as the two spheres begin to overlap.
-This is a more realistic repulsive potential than that of the Hard Sphere model which goes instantly to infinity when the two spheres make contact (at r=2R), and is zero otherwise.
-Because the WCA spheres have no hard boundary in their interatomic potential, they are called ``soft spheres'' and their gradually fading potential makes them slightly ``squishy''.
+The repulsive portion of the Leonard-Jones potential forms the potential for a 
+Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. 
+The WCA potential is described by the equation: 
+\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}
+As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
+from infinity as the distance between the centers of the atoms r increases, 
+and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
+where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
+As with the Lenard-Jones potential, the repulsive WCA potential increases 
+rapidly and approches infinity as the two spheres begin to overlap. This 
+is a more realistic repulsive potential than that of the Hard Sphere model 
+which goes instantly to infinity when the two spheres make contact (at r=2R), 
+and is zero otherwise. Because the WCA spheres have no hard boundary in 
+their interatomic potential, they are called ``soft spheres'' and their 
+gradually fading potential makes them slightly ``squishy''.
 
@@
 
-The WCA potential is not well suited for implementing Density Functional Theory.  
-Therefore, an erf potential which closely approximates the WCA potential at temperatures 
-of interest will be used instead of the WCA potential. The erf potential is given by 
-\begin{equation}
-	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
-\end{equation} 
-where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
-potentail when r equals $\alpha$ 
-\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
-and is given by 
-\begin{equation}\label{alphaT}
-	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
-\end{equation}
-
-%Deleted old Xi method
-%move 3-9 to methods
-
-Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$.
-Figure \ref{fig:alphaXivsT} shows how $\alpha$ and $\Xi$ vary with temperature when $\sigma=1$ and $\epsilon=1$.
-% and Figure \ref{fig:erf_Potential} shows the erf potential at reduced temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and $\frac{k_BT}{\epsilon}=3$ alongside the WCA potential.
-%As seen in Figure \ref{fig:erf_potential}, the erf potentials closely match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, and for r/$\sigma$ less than ???.
-% so only values of $V_{erf}$ that are much less than $k_BT$/$\epsilon$ ???
-will be considered.
-% In these regions $\alpha$ decreases with increasing temperature, and $\Xi$ increases with increasing temperature.
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{plot_alpha.pdf}
-      \caption{$\alpha$ versus $k_BT/\epsilon$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{plot_xi.pdf}
-      \caption{$\Xi$ versus $k_BT/\epsilon$}
-    \end{subfigure} 
-    \caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, show how the parameters $\alpha$ and $\Xi$ vary with temperature.
-(a) $\alpha$ decreases with increasing temperature.
-(b) $\Xi$ rapidly increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then gradually decreases.}
-    \label{fig:alphaXivsT}
-    \end{figure}
-
-%\begin{figure}[h!]
-%    \centering
-    %\includegraphics[height=7cm]{erf_Potential.pdf}
-%    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
-%    \label{fig:erf_Potential}
-%  \end{figure}
-
-%\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
-
-Another way to find values for $\Xi(T)$ is to use the second virial coefficent $B_{2}$ from the equation of state expressed in terms of the virial expansion. 
-\begin{equation}\label{virialequation}
-    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
-\end{equation}
-where 
-\begin{equation}
-	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
-\end{equation}
-The second virial coefficient can be derived from Euler's equation for the internal energy of a system, 
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-using the relations from statistical thermophysics which views the thermodynamic quantities as ensemble averages
-\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
-\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
-\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-which, for a canonical ensemble, gives
-\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Expressing the Helmholtz free energy in terms of the pair potential $V(r)$ results in the equation of state expressed in terms of the virial expansion as given by Equation~\ref{virialequation}.
-
-The value of $\Xi$ at a given temperature, T, can be found computationally by equating the second virial coefficient evaluated for the error function potential $V_{erf}(r)$, and the second virial coefficent evalued for the WCA potential $V_{WCA}$.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
-Evaluating $B_{2}$ for the erf potential gives
-%\begin{equation}B_{2erf}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{erf}(r)}-1\right)d\vec{r} \end{equation}
-\begin{equation}
-	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
-\end{equation}
-where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, $B_{2WCA}$ can be found computationally from 
-\begin{equation}
-	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
-\end{equation}
-
-A plot of $\Xi(T)$ obtained through this method is shown in Figure \ref{fig:xi_fromB2vsT}, and Figure \ref{fig:erf_Potential_xifromB2} shows the erf potential at a few different temperatures alongside the WCA potential.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $\Xi$ generated from matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ at each temperature.}
-    \label{fig:xi_fromB2vsT}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
-    \label{fig:erf_Potential_xifromB2}
-  \end{figure}
-
-%\color{red}
-% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
-%\color{black}
-
-%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
-%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
-%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
-%classical fluid.
+%MOVE to METHODS start >>>
+%........................
+%MOVE to METHODS  end  >>>
 
@@
 
-Freezing is governed by entropy, which the universe, or any closed system, seeks to maximize according to the second law of Thermodynamics.
-When the entropy of a pure substance in its liquid state is lower than the entropy that it would have in its solid state at the same temperature (necessary for thermal equilibrium), the same pressure (necessary for mechanical equilibrium), and the same chemical potential (necessary for diffusive equilibrium), an abrupt phase transition occurs, that is, the liquid freezes forming a crystalline structure.
-The particular crystalline structure and density that result are affected by the interatomic potential.
-When repulsive, soft sphere fluids (that are not too soft) freeze, they form lattice structures composed of Face Centered Cubic (FCC) crystal cells, as do hard spheres.
-%This structure is the three-dimensional equivilant to the two-dimensional structure that one would see by filling the bottom of a box with a single layer of marbles.
-This structure is what one would see by filling a cube with marbles. It results in the highest density achievable for hard, and nearly-hard, spheres that are purely repulsive.
+Freezing is governed by entropy, which the universe, or any closed system, 
+seeks to maximize according to the second law of Thermodynamics. When the 
+entropy of a pure substance in its liquid state is lower than the entropy 
+that it would have in its solid state at the same temperature (necessary 
+for thermal equilibrium), the same pressure (necessary for mechanical 
+equilibrium), and the same chemical potential (necessary for diffusive 
+equilibrium), an abrupt phase transition occurs, that is, the liquid freezes 
+forming a crystalline structure. The particular crystalline structure and 
+density that result are affected by the interatomic potential. When repulsive, 
+soft sphere fluids (that are not too soft) freeze, they form lattice 
+structures composed of Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, 
+as do hard spheres. 
+%This structure is the three-dimensional equivilant to the two-dimensional 
+%structure that one would see by filling the bottom of a box with a single layer of marbles. 
+This structure is what one would see by filling a cube with marbles. It 
+results in the highest density achievable for hard, and nearly-hard, spheres 
+that are purely repulsive.
 
 %move this paragraph to methods?
-Although the entire crystal lattice can be constructed with FCC cells, FCC cells are not the smallest cells with which the lattice can be built.
-The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell_firsttime} is the smallest structure that, when repeated, re-creates the entire crystal lattice.
-The parallepiped primitive cell takes up one-fourth the volume of an FCC cell, includes portions of eight atoms for a total of one full atom, and has eight lattice sites convenietly reachable by moving along one of three different lattice vectors to step from one atom to the next.
-The three lattice vectors will be useful in creating the lattice structure computationally, and will also make the program run faster.
+Although the entire crystal lattice can be constructed with FCC cells, 
+FCC cells are not the smallest cells with which the lattice can be built. 
+The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell_firsttime} 
+is the smallest structure that, when repeated, re-creates the entire crystal 
+lattice. The parallepiped primitive cell takes up one-fourth the volume 
+of an FCC cell, includes portions of eight atoms for a total of one full 
+atom, and has eight lattice sites convenietly reachable by moving along 
+one of three different lattice vectors to step from one atom to the next. 
+The three lattice vectors will be useful in creating the lattice structure 
+computationally, and will also make the program run faster.
 
@@
 
-For pure substances, a phase transition occurs along a line on a Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase diagram shown in Figure \ref{fig:P-T_Diagram}.
-This phase diagram is representative of some arbitrary, unknown substance composed of atoms, or molecules which exibit both attraction and repulsion.
-However, since the atoms in a WCA fliud do not exhibit attraction there will not be a dividing line between the gas and liquid phases, and so the term ``liquid" or ``gas" is not used, but simply ``fluid".
-Points along the line that separates the solid and the fluid phases indicate a phase transition, and show that the solid and the fluid both have the same temperature and the same pressure at a phase transistion.
-
-%For each particular temperature, the pressure of the liquid and the solid are the same at the point of transistion. 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=5cm]{P-T_Diagram.png}
+For pure substances, a phase transition occurs along a line on a 
+Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
+diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is 
+representative of some arbitrary, unknown substance composed of atoms, 
+or molecules which exibit both attraction and repulsion. However, since 
+the atoms in a WCA fliud do not exhibit attraction there will not be a 
+dividing line between the gas and liquid phases, and so the term ``liquid" 
+or ``gas" is not used, but simply ``fluid". Points along the line that 
+separates the solid and the fluid phases indicate a phase transition, 
+and show that the solid and the fluid both have the same temperature and 
+the same pressure at a phase transistion.
+
+%For each particular temperature, the pressure of the liquid and the 
+%solid are the same at the point of transistion. 
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{P-T_Diagram.png}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
   \end{figure}
-The pressure can be related to the Helmholtz free energy F, which is minimized for a system at fixed temperature, volume and number of particles as the entropy of the system and its surroundings is maximized.
-The negative of the partial derivative of the Helmholtz free energy with respect to volume V at a fixed temperature T and number of particles N equals the pressure.
+The pressure can be related to the Helmholtz free energy F, which is minimized 
+for a system at fixed temperature, volume and number of particles as the 
+entropy of the system and its surroundings is maximized. The negative of 
+the partial derivative of the Helmholtz free energy with respect to volume V 
+at a fixed temperature T and number of particles N equals the pressure. 
 \begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
-\noindent Dividing both the Helmholtz free energy and the volume by the number of particles, the partial derivative of the Helmholtz free energy with respect to volume becomes the partial derivative of the Helmholtz free energy per atom with respect to the inverse number density.
+\noindent Dividing both the Helmholtz free energy and the volume by the 
+number of particles, the partial derivative of the Helmholtz free energy 
+with respect to volume becomes the partial derivative of the Helmholtz 
+free energy per atom with respect to the inverse number density. 
 \begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.
 
-%In a process called Maxwell's Double Tangent construction, two curves are used to find the densities of the liquid and the solid at the point of transistion.
-Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for the solid both at a fixed temperature.
-Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope.%
-Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid and one for the solid at a fixed temperature.
-Since the pressure of the liquid is the same as the pressure of the solid at the point of transistion, the density of the liquid and the density of the solid at the point of freezing can be found from identifying the inverse densties at which the tangents to each curve have the same slope.
-This process is called Maxwell's Double Tangent Construction.
-
-\begin{figure}[h!]
-    \centering
-     \includegraphics[height=7cm]{figs/MaxwellDTC-Fig1.pdf}
+%In a process called Maxwell's Double Tangent construction, two curves are 
+%used to find the densities of the liquid and the solid at the point of transistion. 
+%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
+%the solid both at a fixed temperature. Since the pressure of the liquid 
+%is the same as the pressure of the solid at the point of transistion, the 
+%density of the liquid and the density of the solid at the point of freezing 
+%can be found from identifying the inverse densties at which the tangents 
+%to each curve have the same slope.%
+Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid 
+and one for the solid at a fixed temperature. Since the pressure of the 
+liquid is the same as the pressure of the solid at the point of transistion, 
+the density of the liquid and the density of the solid at the point of 
+freezing can be found from identifying the inverse densties at which the 
+tangents to each curve have the same slope. This process is called 
+Maxwell's Double Tangent Construction.
+
+\begin{figure}[h!]
+    \centering
+     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
      % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
-    \caption{Identifying densities at the solid-liquid transition using Maxwell's Double Tangent Construction}
+    \caption{Identifying densities at the solid-liquid transition using 
+    Maxwell's Double Tangent Construction}
     \label{fig:MaxwellDT}
@@
 
-To facilitate finding the pressure at the phase transition, %, which occurs at the points which have the same slope,
+To facilitate finding the pressure at the phase transition, 
+%, which occurs at the points which have the same slope,
 it is helpful to consider the Gibbs free energy given by G=F+PV. 
-Rearranging the terms and dividing by N gives an  expression for the Helmholtz free energy per atom $f$ in terms of the Gibbs free energy per atom  $g$ (which is also the chemical potential):  %From the Gibbs free energy an expression can be derived which relates  ....:
+Rearranging the terms and dividing by N gives an  expression for the 
+Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
+per atom  $g$ (which is also the chemical potential):  
+%From the Gibbs free energy an expression can be derived which relates  ....:
 
@@
 \begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
-\noindent This equation has the form of a line:
+This equation has the form of a line:
 \begin{displaymath}{y=mx+y_o}\end{displaymath}
@@
 %\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
-\noindent and describes the red line shown in Figure \ref{fig:MaxwellDT} with the pressure given by the negative of the slope of the line, and the Gibb's free energy per atom given by the y-intercept.
-At the point of phase transition:
+and describes the red line shown in Figure \ref{fig:MaxwellDT} with the 
+pressure given by the negative of the slope of the line, and the Gibb's 
+free energy per atom given by the y-intercept. At the point of phase transition:
 \begin{displaymath}{g_L=g_S}\end{displaymath} 
@@
 
-The pressure at the transistion point can be obtained from the point of intersection that occurs when the Gibbs free energy per atom is plotted against the pressure for both the liquid and the solid at fixed temperature, as shown in Figure \ref{fig:GibbsvsP}.
-These curves are constructed from the slope and cooresponding y-intercept data collected from lines tangent along the curves on the the Helmholtz free energy per atom verses inverse density plots for both the liquid and the solid.
-\begin{figure}[h!]
-    \centering
-     \includegraphics[height=7cm]{MaxwellDTC-Fig2.pdf}
+The pressure at the transistion point can be obtained from the point of 
+intersection that occurs when the Gibbs free energy per atom is plotted 
+against the pressure for both the liquid and the solid at fixed temperature, 
+as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
+the slope and cooresponding y-intercept data collected from lines tangent 
+along the curves on the the Helmholtz free energy per atom verses inverse 
+density plots for both the liquid and the solid. 
+\begin{figure}[h!]
+    \centering
+     \includegraphics[height=8cm]{MaxwellDTC-Fig2.pdf}
      %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
@@
 
-Once the pressure is known for the transition point, the density of the liquid at the 
-transition point can be found from a plot of the pressure versus the inverse of the 
-density for the liquid. The density of the solid at the transition point can be found 
-the same way, as shown in Figure \ref{fig:Pvsinvn}.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{MaxwellDTC-Fig3.pdf}
+Once the pressure is known for the transition point, the density of the 
+liquid at the transition point can be found from a plot of the pressure 
+versus the inverse of the density for the liquid. The density of the solid 
+at the transition point can be found the same way, as shown in Figure 
+\ref{fig:Pvsinvn}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{MaxwellDTC-Fig3.pdf}
     % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
-    \caption{The liquid and solid densities at the point of transistion are found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
+    \caption{The liquid and solid densities at the point of transistion are 
+    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
     \label{fig:Pvsinvn}
@@
 
-A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this procedure for various temperatures.
-%The lowest density the crystal exhibts and the highest density the fluid exhibits at the phase transistion can then be identified for each temperature.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=4cm]{T-n_Diagram.png}
-    \caption{Conceptual Temperature-Density Phase Diagram used to identify the number densities of the liquid and the solid during a phase transistion.
-The gray region is where the liquid and the solid coexist, and are therefore not pure states.}
+A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
+can be generated by repeating this procedure for various temperatures. 
+%The lowest density the crystal exhibts and the highest density the fluid 
+%exhibits at the phase transistion can then be identified for each temperature.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{T-n_Diagram.png}
+    \caption{Conceptual Temperature-Density Phase Diagram used to identify 
+    the number densities of the liquid and the solid during a phase 
+    transistion. The gray region is where the liquid and the solid coexist, 
+    and are therefore not pure states.}
     \label{fig:T-n_Diagram}
   \end{figure} 
-%One way of implementing this practically is to determine what the Helmholtz free energy would be for both the liquid and solid phases at a particular temperature and density.
-Whichever phase has the lower Helmholtz free energy will be the phase that the substance will be in.
+%One way of implementing this practically is to determine what the Helmholtz 
+%free energy would be for both the liquid and solid phases at a particular 
+%temperature and density. Whichever phase has the lower Helmholtz free 
+%energy will be the phase that the substance will be in.
 
@@
 
-The key idea behind classical Density Functional Theory (cDFT) is that free energy can be written as a functional (a function of a function) of a number density profile $n(\vec{r})$ which describes the way the number density, or atoms per volume, varies spatially on average.
+The key idea behind classical Density Functional Theory (cDFT) is that 
+free energy can be written as a functional (a function of a function) of 
+a number density profile $n(\vec{r})$ which describes the way the number 
+density, or atoms per volume, varies spatially on average. 
 \begin{displaymath}{f[n(}\vec{r}{)]}\end{displaymath}
-This, coupled with the fact that the free energy of a system tends toward a minimum at equilibrium means that the equilibrium free energy of the system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying $n(\vec{r})$ through all possible spatial profiles until the free energy is minimized.
+This, coupled with the fact that the free energy of a system tends toward 
+a minimum at equilibrium means that the equilibrium free energy of the 
+system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
+number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying 
+$n(\vec{r})$ through all possible spatial profiles until the free energy 
+is minimized.
 
@@
 
-%The reason this works is because cDFT satisfies the conditions of the variational principle, the same mathematical principle used to determine that the shortest distance between two points is a straight line by varying the path between the two points until the distance is minimized.
-
-A description of cDFT ordinarily begins with considering the Grand Free Energy defined as
+%The reason this works is because cDFT satisfies the conditions of the 
+%variational principle, the same mathematical principle used to determine 
+%that the shortest distance between two points is a straight line by 
+%varying the path between the two points until the distance is minimized.
+
+A description of cDFT ordinarily begins with considering the Grand Free 
+Energy defined as
 \begin{equation}\Phi=F-\mu{N}\end{equation}
 %\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
-which can then be expressed in terms of a varrying number density profile $n(\vec r)$
+which can then be expressed in terms of a varrying number density 
+profile $n(\vec r)$
 %\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
 \begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally applied potential energy, and the Helmholtz free energy is expressed as
+where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally 
+applied potential energy, and the Helmholtz free energy is expressed as
 \begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
-%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces to the Helmholz free energy $F$.
-%The second and third terms in Eq~\ref{GrandFE} are both functionals of $n(\vec{r})$, so it remains to show that $F$, and thus $\Phi$, is a functional of $n(\vec{r})$.
-
-The Grand free energy is applicable to systems with a varying number of particles and fixed chemical potential, temperature and volume.
-But in this paper a system with a fixed temperature, volume, and number of particles will be of interest in which case
+%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces 
+%to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} 
+%are both functionals of $n(\vec{r})$, so it remains to show that $F$, and 
+%thus $\Phi$, is a functional of $n(\vec{r})$.
+
+The Grand free energy is applicable to systems with a varying number of 
+particles and fixed chemical potential, temperature and volume. But in 
+this paper a system with a fixed temperature, volume, and number of 
+particles will be of interest in which case
 \begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
@@
 \begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
-The dependence on $n(\vec r)$ is now only present in the expression for the Helmholtz free energy, and so it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium.
-%In this paper a system with a fixed temperature, volume, and number of particles will be examined in which case it is the Helmholtz energy as a functional of the number density profile $F[n(\vec r)]$ that is of interest.
-More specifically, in this paper, the Helmholtz free energy per particle, or atom,  $f_N=\frac{F}{N}$ will be minimized which is given by
+The dependence on $n(\vec r)$ is now only present in the expression for 
+the Helmholtz free energy, and so it is the Helmholtz energy as a functional 
+of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium. 
+%In this paper a system with a fixed temperature, volume, and number of 
+%particles will be examined in which case it is the Helmholtz energy as a 
+%functional of the number density profile $F[n(\vec r)]$ that is of interest. 
+More specifically, in this paper, the Helmholtz free energy per particle, 
+or atom,  $f_N=\frac{F}{N}$ will be minimized which is given by
 \begin{equation}f_N=\frac{U}{N}-\frac{TS}{N}\end{equation}
@@
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
-This result shows that the natural variables of the Helmholtz free energy per particle $f_N$ are temperature, volume and one over the number of particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed.
-%Thus, the Helmholtz free energy will be minimized in a system that has a temperature, volume, and number of particles that are fixed.
+This result shows that the natural variables of the Helmholtz free energy 
+per particle $f_N$ are temperature, volume and one over the number of 
+particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
+%Thus, the Helmholtz free energy will be minimized in a system that has 
+%a temperature, volume, and number of particles that are fixed.
 %\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_V(T,n)\end{equation}
@@
 
-The minimization can be illustrated by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
-Here ``surroundings" means the entire universe minus the system.
-This, together with the conservation of energy (the first law of thermodynamics) which says that energy lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that changes in the free energy will be negative for spontaneous processes when the natural variables of the free energy are held fixed.
+The minimization can be illustrated by considering that according to the 
+second law of thermodynamics, a system that is not in equilibrium will 
+exprience spontanteous processes that increase the total entropy of a 
+system and its surroundings until equilibrium is reached, 
+$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
+the entire universe minus the system. This, together with the conservation 
+of energy (the first law of thermodynamics) which says that energy lost 
+(or gained) by the system is gained (or lost) by the surroundings 
+$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
+changes in the free energy will be negative for spontaneous processes 
+when the natural variables of the free energy are held fixed. 
 %And so the free energy tends toward a minimum. 
-%How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?
+%How will a system at a fixed temperature, volume, and number of 
+%particles respond to a small offset from equilibrium?
 
@@
 \end{equation}
-and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
+and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
+$dT=0$ gives
 \begin{equation}
@@
 \end{equation}
-where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
+where conservation of energy was also applied. Fixing the volume and the 
+number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
+results in
 \begin{equation}
@@
 which, for fixed N, is the same as
-\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative, and thus the free energy tends toward a minimum.
-(reference schroder) This illustrates the tendancy toward minimization for a system near equilibrium.
-A general proof of minimization can be found in (reference simple fluids).
+\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
+Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
+of particles are fixed, the change in the free energy of the system will 
+always be negative, and thus the free energy tends toward a minimum. 
+(reference schroder) This illustrates the tendancy toward minimization 
+for a system near equilibrium. A general proof of minimization can be 
+found in (reference simple fluids).
 
@@
 \begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
-which results from equating the expressions for the Helmholtz free energy from thermodynamics to those of statistical mechanics.
-The paritian function Z is integrated over all of phase space for N identical particles each with a momentum $\vec{p}$ and a position $\vec{r}$ is given by
+which results from equating the expressions for the Helmholtz free energy 
+from thermodynamics to those of statistical mechanics. The paritian function 
+Z is integrated over all of phase space for N identical particles each with 
+a momentum $\vec{p}$ and a position $\vec{r}$ is given by
 \begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
@@
 \begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
-where the translational kinetic energy of the N atoms is given by \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
-the potential energy of the total interaction of each atom with the other N-1 atoms is given by
-\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} is an externally applied potential that varies spatially, such that its effects on each atom depend on the atom's position.
-Often $V_{ext}$ is set to zero, as it is in this paper.
-
-Noting that the kinetic energy depends only on momentum while the total interatomic and external potential energies depend only on position, the integral in equation (8) can be separated into two parts
+where the translational kinetic energy of the N atoms is given by 
+\begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
+the potential energy of the total interaction of each atom with the other 
+N-1 atoms is given by
+\begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
+and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
+is an externally applied potential that varies spatially, such that its 
+effects on each atom depend on the atom's position. Often $V_{ext}$ is 
+set to zero, as it is in this paper.
+
+Noting that the kinetic energy depends only on momentum while the total 
+interatomic and external potential energies depend only on position, 
+the integral in equation (8) can be separated into two parts 
 \begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation}
@@
 \begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
-These integrals can conveniently be written in terms of the DeBroglie wavelength $\Lambda$, or the quantum volume $\text{n}_\text{Q}$
+These integrals can conveniently be written in terms of the DeBroglie 
+wavelength $\Lambda$, or the quantum volume $\text{n}_\text{Q}$
 \begin{equation}{\frac{1}{h^{3}}\int{d\vec{p}}~e^\frac{-\frac{|\vec{p}|^2}{2m}}{k_BT}=\left(\sqrt{\frac{k_BTm}{2\pi\hbar^2}}\right)^3=\frac{1}{\Lambda^{3}}}=\text{n}_\text{Q}\end{equation} 
@@
 Rewriting Equation~\ref{Fidealmultipleint} in terms of $\Lambda$ gives
-\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}Using Stirling's approximation \begin{displaymath}{\ln(N!)=N\ln{N}-N}\end{displaymath} this becomes
+\begin{equation}{F_{ideal}= -k_BT[\ln(V^N)-\ln(N!) - \ln(\Lambda^{3N})]}\end{equation}
+Using Stirling's approximation \begin{equation}{\ln(N!)=N\ln{N}-N}\end{equation} 
+this becomes
 \begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
-Rewriting the equation in terms of the number density $n=\frac{N}{V}$, and the quantum volume, $\text{n}_\text{Q}$ gives
+Rewriting the equation in terms of the number density $n=\frac{N}{V}$, 
+and the quantum volume, $\text{n}_\text{Q}$ gives
 \begin{equation}{F_{ideal}= k_BT\left[\frac{N}{V}\ln{\left(\frac{N}{V}\frac{1}{\text{n}_\text{Q}}\right)}-\frac{N}{V}\right]}V\end{equation}
-\begin{equation}{F_{ideal}= k_BT[n\ln(n/\text{n}_\text{Q})-n]V}\label{eq:Fideal}\end{equation}   
-For a spatially varying number density, the ideal free energy in Equation~\ref{eq:Fideal} becomes
-\begin{equation}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]d\vec{r}}\end{equation}
+\begin{equation}\label{eq:Fideal}{F_{ideal}= k_BT[n\ln(n/\text{n}_\text{Q})-n]V}\end{equation}   
+For a spatially varying number density, the ideal free energy in 
+Equation~\ref{eq:Fideal} becomes
+\begin{equation}\label{eq:Fideal-n(r)}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]d\vec{r}}\end{equation}
 
@@
 \begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
-This integral can be greatly simplified if it can be assumed that the potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the potential energies between pairs of atoms, where each atom is paired in turn with each of the other atoms in the system.
-In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes
+This integral can be greatly simplified if it can be assumed that the 
+potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
+potential energies between pairs of atoms, where each atom is paired in 
+turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ 
+becomes 
 \begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
@@
     %\includegraphics[height=7cm]{diagrammic.pdf}
-    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations between individual pairs of atoms are shown in green. 
-             Interactions between pairs of atoms with an atom shared in common are shown in red. 
-             At low densities only the interactions shown in green are considered, since atoms are not likely to be clustered.}
+    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations 
+    between individual pairs of atoms are shown in green. Interactions 
+    between pairs of atoms with an atom shared in common are shown in red. 
+    At low densities only the interactions shown in green are considered, 
+    since atoms are not likely to be clustered.}
     \label{fig:diagrammic}
@@
 
-At low densities, it is not likely to find two or more atoms in close proximity, and so only interactions between individual pairs of spheres, such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
-Each pair can then be described by a function of the distance between the two atoms in the pair. Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ for each pair we get 
+At low densities, it is not likely to find two or more atoms in close 
+proximity, and so only interactions between individual pairs of spheres, 
+such as those shown in green in Figure~\ref{fig:diagrammic}, are considered.
+Each pair can then be described by a function of the distance between the 
+two atoms in the pair. Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ 
+for each pair we get 
 \begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
 \color{black}
-With a careful counting of the resulting number of such pairs, the excess free energy can then be written as 
+With a careful counting of the resulting number of such pairs, the excess 
+free energy can then be written as 
 \begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
 where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
-Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ for a homogeneous density n=$\frac{N}{V}$ this becomes
+Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
+for a homogeneous density n=$\frac{N}{V}$ this becomes
 \begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
 Thus the free excess energy per atom in the low-density limit is given by 
-\begin{equation}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} where the quantity in parenthesis can be recognized 
-as the second virial expanision coefficient typically called $B_2$. %where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
-It is important that the expression for the excess Helmholtz free energy correctly reduces to this result in the low-density limit. 
-%Although we will not be restricting ourselves to the low-density limit, it is important that the theory we use correctly reduces to this same result in the low-density limit. 
-Outside of the low density limit, it is quite difficult to evaluate the excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}.
-
-Another way to form an expression for the excess Helmholtz free energy in terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
-Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid which can then be integrated to find the excess Helmholtz free energy.
+\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
+where the quantity in parenthesis can be recognized 
+as the second virial expanision coefficient typically called $B_2$. 
+%where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
+It is important that the expression for the excess Helmholtz free energy 
+correctly reduces to this result in the low-density limit. 
+%Although we will not be restricting ourselves to the low-density limit, 
+%it is important that the theory we use correctly reduces to this same 
+%result in the low-density limit. 
+Outside of the low density limit, it is quite difficult to evaluate the 
+excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}.
+
+Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
+\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+is a function of the number density n. For a spatially varying number density, 
+as would be expected for a crystal, the free energy is a functional of 
+the number density profile $n(\vec{r})$  
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
+where the ideal Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}.
+To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
+the temperature fixed until the free energy is minimized. The minimum free 
+energy and its corresponding number density profile are then the equilibrium 
+values for that temperature.
+
+\section{Soft Fundamental Measure Theory (SFMT)}
+Another way to form an expression for the excess Helmholtz free energy in 
+terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
+Scaled particle theory seeks a general expression for the excess chemical 
+potential of a homogeneous Hard Sphere fluid which can then be integrated 
+to find the excess Helmholtz free energy.
 %\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
@@
 \begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
-It is an extension of the work done by Widom who showed that the excess chemical potential of a homogenesous fluid corredsponds to the reversible work 
-of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R. 
-Widom's result was obtained by considering the difference in the excess Helmholtz free energy when one atom is added to an otherwise closed system. 
-He showed that the excess chemcial potential is related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
-between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula given by
+It is an extension of the work done by Widom who showed that the excess 
+chemical potential of a homogenesous fluid corredsponds to the reversible 
+work of inserting a sphere of radius R into a fluid consisting of hard 
+spheres of radius R. Widom's result was obtained by considering the difference 
+in the excess Helmholtz free energy when one atom is added to an otherwise 
+closed system. He showed that the excess chemcial potential is related to 
+the ensemble average of the Boltzman factor for the potential energy 
+difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
+The resulting equation is known as Widom's Insertion Formula given by
 %\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
@@
 \color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
-where the radius of the cavity $R_{cavity}$ is equal to the radius of a hard sphere, R, and the cavity is either filled or not.
-\color{black}%For cavities that are much larger than one of the hard spheres,
-Forming a more general result for different sizes of cavities in the fluid, Scaled Partical Theory arrives at an expression for the excess Helmholtz free energy of a homogeneous fluid in terms of the packing fraction $\eta$
+where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
+hard sphere, R, and the cavity is either filled or not. \color{black}
+%For cavities that are much larger than one of the hard spheres, 
+Forming a more general result for different sizes of cavities in the fluid, 
+Scaled Partical Theory arrives at an expression for the excess Helmholtz 
+free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
 \begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
@@
 
-An expression for the excess Helmholtz free energy as a functional of the number density profile $n(\vec r)$ can be found using Fundamental Measure Theory (FMT).
-FMT was developed for hard sphere fluids, and has been very sucessful at predicting the thermodynamic properties of inhomogeneous hard sphere fluids starting from an expression for the excess Helmholtz free energy obtained from Scaled Partical Theory.
-Soft Fundamental Measure Theory (SFMT), which will be addressed later, extends FMT to fluids with soft potentials like that of the WCA fluid.
+An expression for the excess Helmholtz free energy as a functional of the 
+number density profile $n(\vec r)$ can be found using Fundamental Measure 
+Theory (FMT). FMT was developed for hard sphere fluids, and has been very 
+sucessful at predicting the thermodynamic properties of inhomogeneous hard 
+sphere fluids starting from an expression for the excess Helmholtz free 
+energy obtained from Scaled Partical Theory. Soft Fundamental Measure 
+Theory (SFMT), which will be addressed later, extends FMT to fluids with 
+soft potentials like that of the WCA fluid.
 
@@
 \begin{equation}{\zeta_{0}=1}\end{equation}
-where $\zeta_{2}$ is the surface area of a sphere, and $\zeta_{1}$ is the radius of a sphere.
-The quantities $\zeta_{3}, \zeta_{2}, \zeta_{1},$ and $\zeta_{0}$ are fundamental geometric measures of a sphere, and they will prove useful by defining a set of weighted densities $n_{i}=n\zeta_{i}$ where $n=\frac{N}{V}$.
+where $\zeta_{2}$ is the surface area of a sphere, and $\zeta_{1}$ is the 
+radius of a sphere. The quantities $\zeta_{3}, \zeta_{2}, \zeta_{1},$ 
+and $\zeta_{0}$ are fundamental geometric measures of a sphere, and they 
+will prove useful by defining a set of weighted densities $n_{i}=n\zeta_{i}$ 
+where $n=\frac{N}{V}$. 
 \begin{equation}\label{n3}{n_{3}=n\zeta_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
@@
 
-\noindent The excess free energy in Eq~\ref{Fexcess-SPT} can be written in terms of the number density $\frac{N}{V}$
+\noindent The excess free energy in Eq~\ref{Fexcess-SPT} can be written 
+in terms of the number density $\frac{N}{V}$
 \begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
@@
 
-The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted densities given the relation $n_{i}=n\zeta_{i}$, where the weights are given by $\zeta_{0},\zeta_{1},\zeta_{2},\zeta_{3}$ respectively.
-These weights are valid for determining the excess free energy of a system of hard spheres with a homogeneous number density n.
-But these weights, and their related  weighted densities, can be written in a more general form allowing for their extention to a system with a spacially varying number density, that is, an inhomogeneous fluid.
+The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
+densities given the relation $n_{i}=n\zeta_{i}$, where the weights are 
+given by $\zeta_{0},\zeta_{1},\zeta_{2},\zeta_{3}$ respectively. These 
+weights are valid for determining the excess free energy of a system of 
+hard spheres with a homogeneous number density n. But these weights, and 
+their related  weighted densities, can be written in a more general form 
+allowing for their extention to a system with a spacially varying number 
+density, that is, an inhomogeneous fluid.
 
@@
 and so the mayer function for a hard sphere is given by \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
-The mayer function can be written in terms of weight functions if two additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
+The mayer function can be written in terms of weight functions if two 
+additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
 \begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  where the vector weights are defined as \begin{equation}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
@@
 
-Taking the derivative of the mayer function in Eq~\ref{f(r)step} will prove useful, and is given by
+Taking the derivative of the mayer function in Eq~\ref{f(r)step} will 
+prove useful, and is given by
 \begin{equation}\label{mayerder=w2convolution}{\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}}\end{equation} 
@@
 \begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
-Using the general relation for convolution, which is equal to the inverse Fourier Transform of a product of Fourier Transforms \begin{equation}{f \otimes g = \text{IFT[ FT}[f]\text{ x FT[g] }]}\end{equation} with
+Using the general relation for convolution, which is equal to the inverse 
+Fourier Transform of a product of Fourier Transforms 
+\begin{equation}{f \otimes g = \text{IFT[ FT}[f]\text{ x FT[g] }]}\end{equation} 
+with
 \begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
@@
 
-Incorporating into Equation~\ref{FexfromSPT} the two vector weights which were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated expression for $\Phi_3$ developed by Tarazona, the excess free energy becomes
+Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
+were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
+expression for $\Phi_3$ developed by Tarazona, the excess free energy becomes
 \begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
@@
 
-The expression for the weighted densities can be written in a way that accomodates both homogenious fluids where the number denstiy n is constant, and inhomogeneous fluids where the number density becomes a function of the position $n(\vec{r})$.
+The expression for the weighted densities can be written in a way that 
+accomodates both homogenious fluids where the number denstiy n is constant, 
+and inhomogeneous fluids where the number density becomes a function of 
+the position $n(\vec{r})$.
 \begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-%The equatons above were derived for a uniform fluid where the number density n is constant, but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
-For a spatially varying number density $n(\vec{r})$, the excess Free Energy becomes a functional of the number density profile $n(\vec{r})$ :
+%The equatons above were derived for a uniform fluid where the number density n is constant, 
+%but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
+For a spatially varying number density $n(\vec{r})$, the excess Free 
+Energy becomes a functional of the number density profile $n(\vec{r})$ :
 \begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
 
-The equations developed so far have been for a hard sphere fluid.
-All of the weight function equations, as well as the equation for the deriviative of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists of a delta function characteristic of hard spheres.
-Soft Fundamental Measure Theory (SFMT) extends the Fundamental Measure Theory discussed so far to fluids with soft potenials by keeping the relationships between the weight functions, and that the weight functions satisfy the same, but replacing $w_{2}(r)$ with a weight function appropriate for a soft sphere potential.
+The equations developed so far have been for a hard sphere fluid. All of 
+the weight function equations, as well as the equation for the deriviative 
+of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
+of a delta function characteristic of hard spheres. Soft Fundamental Measure 
+Theory (SFMT) extends the Fundamental Measure Theory discussed so far to 
+fluids with soft potenials by keeping the relationships between the weight 
+functions, and that the weight functions satisfy the same, but replacing 
+$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
 A new set of weight functions can then be derived for a system of soft spheres.
 
-In this paper,  $w_{2}(r)$ will be constructed using the following expresssion for $w_3(r)$ proposed by Schmidt$^{(?)}$ 
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
- where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ and $w_3(\infty)=0$. 
-%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
-Using Equation~\ref{w2_w3_relation}, shown again below, 
-\begin{displaymath}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{displaymath}
-and solving for $w_2(r)$
-\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
-we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
-Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
-\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
-gives
-\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
-which can be integrated to obtain the Mayer function
-\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
-The Mayer function for the error function potential $V_{erf}(r)$ is thus given by
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} %This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation}
-gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
-\begin{equation}{V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} The two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), and are determined to make $V_{erf}(r)$ match $V_{WCA}(r)$ as closely as possible by compensating for the unwanted temperature dependence that shows up in the error function potential $V_{erf}(r)$.
-
-In summary, classical Density Functional Theory (cDFT) is used in conjunction with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free energy of a system, and the corresponding equilibrium number density profile.
-In particular, SFMT is used to form the expression for the excess free energy without having to take into account the interaction each sphere, or atom, has with every other sphere.
-The excess free energy is then added to the ideal free energy calculated analytically to find the resulting free energy for the system.
+%MOVE to METHODS start2 >>>
+%........................
+%MOVE to METHODS end >>>
+
+%MOVE to METHODS start3 >>>
+%........................
+%MOVE to METHODS end >>>
+
+\chapter{Methods}
+
+%MOVE to METHODS start3 >>>
+
+We use classical Density Functional Theory (cDFT) in conjunction 
+with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
+energy of a system, and the corresponding equilibrium number density profile. 
+In particular, SFMT is used to form the expression for the excess free energy 
+without having to take into account the interaction each sphere, or atom, 
+has with every other sphere. The excess free energy is then added to the 
+ideal free energy calculated analytically to find the resulting free energy 
+for the system. 
 \begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-For a spatially varying number density, as would be expected for a crystal, the free energy is a functional of the number density profile $n(\vec{r})$  
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} To implement cDFT, the number density profile $n(\vec{r})$ is varied with the temperature fixed until the free energy is minimized.
-The minimum free energy and its corresponding number density profile are then the equilibrium values for that temperature.
-
-Once the Helmholtz free energy for both the crystal and liquid states of the fluid are found, they can be compared to find the resulting state of the fluid.
-Outside the region of coexistence, the fluid will be completely in a crystalline state if the crystal free energy is lower than the liquid free energy, and it will be completely in a liquid state if the liquid free energy is lower than the crystal free energy.
-
-\chapter{Methods}
-
-To implement cDFT computationally, we begin by picturing one single crystal cell.
-Although the crystal structure is composed of Face Centered Cubic (FCC) cells, the entire lattice can be generated with a cell one-fourth the size called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}.
-The primitive cell has the shape of a parallelpiped, and contains the equivalent of one atom.
-
-  \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
+For a spatially varying number density, as would be expected for a crystal, 
+the free energy is a functional of the number density profile $n(\vec{r})$  
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+
+To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
+the temperature fixed until the free energy is minimized. The minimum free 
+energy and its corresponding number density profile are then the equilibrium 
+values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
+the fluid are found, they can be compared to find the resulting state of 
+the fluid. Outside the region of coexistence, the fluid will be completely 
+in a crystalline state if the crystal free energy is lower than the liquid 
+free energy, and it will be completely in a liquid state if the liquid 
+free energy is lower than the crystal free energy.
+%MOVE to METHODS end3 >>>
+
+%MOVE start4 >>>>
+
+***** DFT uses n(r) for BOTH Fideal and Fexcess ******
+
+To implement cDFT computationally, we begin by picturing one single crystal 
+cell. Although the crystal structure is composed of Face Centered Cubic (FCC) 
+cells, the entire lattice can be generated with a cell one-fourth the size 
+called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}. 
+The primitive cell has the shape of a parallelpiped, and contains the 
+equivalent of one atom.
+
+  \begin{figure}[h!]
+    \centering
+    \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
     \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
@@
 
-This cell is rather complicated for the purposes of this discussion, so consider instead a simplified crystal cell with one atom in its center as shown in Figure \ref{fig:simplecell}. 
-The Helmholtz free energy is minimized for a fixed volume (and temperature and number of particles).
-A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will be used for illustration.
-The number density n for a collection of simplified cells is simply 1 atom per fixed bold-box volume.
-\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} It is easier to think of 1 atom per ``box" (a reduced number density) than 1 atom per some length$^3$ (the actual number density).
-Reduced number density is given by $n^*$ and is equal to the number of atoms in a ``box" of size $\sigma^3$.
-\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  has a volume of $\sigma^3$, the reduced number density is just the number of atoms in the fixed bold-box volume. 
+This cell is rather complicated for the purposes of this discussion, so 
+consider instead a simplified crystal cell with one atom in its center 
+as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
+minimized for a fixed volume (and temperature and number of particles).  
+A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
+be used for illustration. The number density n for a collection of 
+simplified cells is simply 1 atom per fixed bold-box volume.
+\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
+It is easier to think of 1 atom per ``box" (a reduced number density) 
+than 1 atom per some length$^3$ (the actual number density). Reduced number 
+density is given by $n^*$ and is equal to the number of atoms in a ``box" 
+of size $\sigma^3$.
+\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
+has a volume of $\sigma^3$, the reduced number density is just the number 
+of atoms in the fixed bold-box volume. 
 \begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}
 
-In a large collection of simple cells, the atom at the center of each cell will not realistically reside exactly at the lattice point in the center of the cell.
-Statistically there will typically be a normal distribution of positions about the lattice point as pictured in the far right box in Figure \ref{fig:Ensemble_Gaus}.
-A plot of a 2-dimensional Gaussian distribution is shown in Figure \ref{fig:Gaus_plot}.
-A 3-dimenstional Gaussian distribution centered at a lattice point can be used to represent a number density profile $n(\vec{r})$  which describes how the atoms are statistically distrubuted over a collection of cells.
-
-  \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
+In a large collection of simple cells, the atom at the center of each cell 
+will not realistically reside exactly at the lattice point in the center 
+of the cell. Statistically there will typically be a normal distribution 
+of positions about the lattice point as pictured in the far right box in 
+Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
+distribution is shown in Figure \ref{fig:Gaus_plot}. A 3-dimenstional 
+Gaussian distribution centered at a lattice point can be used to represent 
+a number density profile $n(\vec{r})$  which describes how the atoms are 
+statistically distrubuted over a collection of cells.
+
+  \begin{figure}[h!]
+    \centering
+    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
     \caption{A Gaussian Distribution Number Density Profile}
@@
 
-It is important to note that the number density does not change when introducing a number density profile. There is still a total of one atom per ``bold box" volume.
+It is important to note that the number density does not change when 
+introducing a number density profile. There is still a total of one atom 
+per ``bold box" volume.
 
@@
 In a large collection of cells, not all lattice sites will be occupied. 
-Vacancies can be taken into account while keeping the volume of the bold box, and thus the overall number density, the same.
-This is done by making the cells smaller.
-A large collection of smaller cells creates a new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} with eight smaller Gaussians of one-eigth the height.
+Vacancies can be taken into account while keeping the volume of the 
+bold box, and thus the overall number density, the same. This is done by 
+making the cells smaller. A large collection of smaller cells creates a 
+new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
+with eight smaller Gaussians of one-eigth the height.
 
@@
 
-Instead of thinking of an atom in one of the eight small cells, the same Gaussian distribution can be obtained from thinking of one-eigth of an atom in each small cell.
-This makes it convenient to describe the new reduced number density in terms of the fraction of vacancies, $f_v$:
+Instead of thinking of an atom in one of the eight small cells, the same 
+Gaussian distribution can be obtained from thinking of one-eigth of an 
+atom in each small cell. This makes it convenient to describe the new 
+reduced number density in terms of the fraction of vacancies, $f_v$:
 
@@
     % FIXME \includegraphics[height=2.5cm]{SameStatPic.png}
-    \caption{Front views of a cube with eight smaller cells.
-The two views show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which give rise to the same Gaussian distribution statistically}
+    \caption{Front views of a cube with eight smaller cells. The two views 
+    show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
+    give rise to the same Gaussian distribution statistically}
     \label{fig:SameStatPic}
@@
 
-By varying the width of the Gaussians and the fraction of vacancies, new number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.
-
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=8cm]{VaryWidthandVacancies.png}
-    \caption{Several Gaussian Distribution Number Density Profiles all with a number density of 1 atom per $V_{bold box}$}
+By varying the width of the Gaussians and the fraction of vacancies, new 
+number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
+    \caption{Several Gaussian Distribution Number Density Profiles all 
+    with a number density of 1 atom per $V_{bold box}$}
     \label{fig:Ensemble_vary}
@@
 
-Once again, it is important to note that the number density is the same for all these number density profiles $n(\vec{r})$.
-There is still a total of one atom per ``bold box" volume.
-This prodcedure is used only to find the crystal free energy as number density profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.
+Once again, it is important to note that the number density is the same 
+for all these number density profiles $n(\vec{r})$. There is still a total 
+of one atom per ``bold box" volume. This prodcedure is used only to find 
+the crystal free energy as number density profile for a liquid is 
+homogeneous, as shown in Figure \ref{fig:homogen_denisty}.
 
@@
 
-A simplified flowchart of the program created to investigate the freezing of a WCA fluid is shown on pages x-x.
-The program takes four inputs: reduced number density n*, reduced temperature t*, a range of Gaussian widths gw, and a range of fraction of vacancies fv.The number density profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ is calculated for each profile until the minimum free energy is found for that combination of $n^*$ and $t^*$, which is known from cDFT to be unique.
-The process is then repeated to find the free energies at other number densities and temperatures.
-
-In general, if the crystal free energy at some $n^*$ and $t^*$ is lower than the homogeneous free energy of a liquid at that same $n^*$ and $t^*$, then the crystaline state dominates.
-But if the crystal free energy is higher than the homogeneous free energy, the liquid state dominates.
-However, there is a region of densities between the liquid and crystal states that are forbidden, and so the actual state can not be determined just by which energy is lower at that particular $n^*$ and $t^*$.
-Maxwell's Double Tangent Construction must be implemented using the date collected to determine the actual crystal and liquid states, and construct phase diagrams.
-\[{}\]
+A simplified flowchart of the program created to investigate the freezing 
+of a WCA fluid is shown on %pages x-x. 
+The program takes four inputs: reduced number density n*, reduced temperature t*, 
+a range of Gaussian 
+widths gw, and a range of fraction of vacancies fv.The number density 
+profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
+is calculated for each profile until the minimum free energy is found for 
+that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
+The process is then repeated to find the free energies at other number 
+densities and temperatures. 
+%MOVE end4 >>>>
+
 \textbf{Program Implementation}
 
-The program computes both the crystal and the homogeneous Helmholtz free energies.
-The Helmholtz free energy is minimized for a system with a fixed temperature, volume, and number density.
-The temperature, T is an input to the program.
-The number density, which is also an input to the program, is used to calculate the volume of one primitive cell over which the free energy is computed.
-The number of atoms in one primitive cell is one.
-To compute the free energy, it is necessary to compute both the ideal and excess free energies and add them, $F = F_{ideal} + F_{excess}$.
+The program computes both the crystal and the homogeneous Helmholtz free 
+energies. The Helmholtz free energy is minimized for a system with a fixed 
+temperature, volume, and number density. The temperature, T is an input to 
+the program. The number density, which is also an input to the program, 
+is used to calculate the volume of one primitive cell over which the free 
+energy is computed. The number of atoms in one primitive cell is one. To 
+compute the free energy, it is necessary to compute both the ideal and 
+excess free energies and add them, $F = F_{ideal} + F_{excess}$.
 %\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
@@
 %lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
-\[\]
-
-The Liquid (homogeneous) ideal free energy $hF_{ideal}$ is computed analytically for a given $n^*$ and $t^*$, where $\sigma=1$ so that $n^*=n\sigma^3$=n, and $\epsilon=1$ so that $t^*=\frac{k_BT}{\epsilon}=k_BT$.
-%\begin{displaymath}{\frac{hF_{ideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-\begin{displaymath}{\frac{hF_{ideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} where
+\newline
+\newline
+*****   start stuff?  ******
+
+In general, if the crystal free energy at some $n^*$ and $t^*$ is lower 
+than the homogeneous free energy of a liquid at that same $n^*$ and $t^*$, 
+then the crystaline state dominates. But if the crystal free energy is 
+higher than the homogeneous free energy, the liquid state dominates. 
+However, there is a region of densities between the liquid and crystal 
+states that are forbidden, and so the actual state can not be determined 
+just by which energy is lower at that particular $n^*$ and $t^*$. Maxwell's 
+Double Tangent Construction must be implemented using the date collected 
+to determine the actual crystal and liquid states, and construct phase 
+diagrams.
+
+*****  end  stuff?  ******
+\newline
+\newline
+
+**** Fideal - liquid and crystal  ****
+
+The Liquid (homogeneous) ideal free energy $F_{Hideal}$ is computed analytically for a given $n^*$ and $t^*$, where $\sigma=1$ so that $n^*=n\sigma^3$=n, and $\epsilon=1$ so that $t^*=\frac{k_BT}{\epsilon}=k_BT$.
+%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
+\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} where
 
@@
 
-The Liquid (homogeneous) excess free energy $hF_{excess}$ is computed analytically for a given $n^*$ and $t^*$ using SFMT with a uniform number density profile that does not depend on $\vec r$. 
-%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
-\begin{displaymath}{\frac{hF_{excess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
-For the case of a homogeneous number density n, $\Phi_1$, $\Phi_2$, and $\Phi_3$ simplify to:
-\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
-\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
-\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
-where \begin{displaymath}{n_i=nw_i}\end{displaymath} 
-%\begin{displaymath}{n_0=nw_0}\end{displaymath} 
-%\begin{displaymath}{n_1=nw_1}\end{displaymath}
-%\begin{displaymath}{n_2=nw_2}\end{displaymath}
-%\begin{displaymath}{n_3=nw_3}\end{displaymath}
-and the weights are as usual 
-\begin{align}
-  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
-  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
-  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
-  \\
-  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
-\end{align}
-%\begin{displaymath}{w_0=?}\end{displaymath} 
-%\begin{displaymath}{w_1=\frac{\alpha}{2}}\end{displaymath}
-%\begin{displaymath}{w_2=6.2...\Xi^2 + \pi\alpha^2}\end{displaymath}
-%\begin{displaymath}{w_3=\alpha(\pi\Xi^2 + 0.52...\alpha^2)}\end{displaymath} \[\]
-
-The Crystal ideal free energy $cF_{ideal}$ is computed in two different ways from the equation 
-\begin{displaymath}{F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV}\end{displaymath}
+The Crystal ideal free energy $F_{Cideal}$ is computed in two different ways from the equation 
+\begin{displaymath}{F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV}\end{displaymath}
 with the varying number density profile given by 
@@
   \begin{align}
-    cF_{ideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
+    F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
   \end{align}
-\item For larger gw values it is assumed that the Gaussians do overlap, and so the free energy is caluculated computationally. First dx is scaled by gw
+\item For larger gw values it is assumed that the Gaussians do overlap, 
+and so the free energy is caluculated computationally. First dx is scaled by gw
 \begin{align}
@@
 volume.
-%\[\]
 \end{itemize}
 
-The crystal excess free energy $cF_{excess}$ is computed computationally using SFMT which gives
-%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
-%Then the crystal, or inhomogeneous, excess free energy is found from
-\begin{displaymath}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} and using Monte-Carlo integration to compute the weighted densities $n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$ used to form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
-
-The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ within the primitive cell, and a vector $\vec{R}_j$ to the center of one of the Gaussians is selected that is within the primitive cell, or one of the neighboring cells within a prudently picked inclusion radius as Gaussians farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=5cm]{InclusionRadius.png}
-    \caption{The Gaussians at each lattice point are of width gw. After a radius of few gw's the Gaussians go to zero and can be neglected.}
-    \label{fig:InclusionRadius}
-  \end{figure}
-
-The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, $n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions $w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to some point in space.
-Monte-Carlo integration works by randomly generating $\vec{dr}$ according to a Gaussian distribution.
-Setting $\vec{r}'=\vec{R}_j + \vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ is the gaussian number density profile for no vacancies, the equation:
-
-\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} implemented computationally in the form:
-\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} becomes:
-\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ (the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of $w_i(|\vec{r}-\vec{r}'|)$ will occur.
-Using the relation:
-\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} $n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
-
-The result can be made more accurate by increasing NumPoints which is the number of vectors randomly generated.
-In order to further reduce the error in the average, antithetic variates are used.
-Rather than computing an average of a set of random points, each random point in the set is paired with a point along the same line as the first point, but opposite in direction, as shown in Figure \ref{fig:AntitheticVariate}.
-This reduces first order error in the average by filtering out an erroneous offset of values along the line connecting the two points by averaging the two results, one of which may be erroneously higher than it should be while the other is erronesously lower than it should be.
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4cm]{AntitheticVariate.png}
-    \caption{Antithetic variates reduce first order error in an average.
-Each point in the average is paired with another point from along the same line as the first, and at an equal, but opposite distance.}
-    \label{fig:AntitheticVariate}
-  \end{figure} 
-Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
-\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-Then the entire process repeated for each vector $\vec{r}_k$ to get the total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
-The standard error of the mean (SEM) is given by:
-\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives a measure of the error in $n_3=\overline{w}_3$:
-\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} NumPoints, the number of vectors randomly generated, is increased until:
-\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum value that $n_3$ can be without error is 1.
+******  Fexcess depends on V(r) ... use SFMT for liquid and crystal ******
+
+%MOVE to METHODS start1 >>>
+The WCA potential is not well suited for implementing Density Functional 
+Theory.  Therefore, an erf potential which closely approximates the WCA 
+potential at temperatures of interest will be used instead of the WCA 
+potential. The erf potential is given by 
+\begin{equation}
+	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
+\end{equation} 
+where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
+potentail when r equals $\alpha$ 
+\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
+and is given by 
+\begin{equation}\label{alphaT}
+	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
+\end{equation}
+
+Unlike the WCA potential, the erf potential depends on temperature through 
+the parameters $\alpha$ and $\Xi$. Figure \ref{fig:alphaXivsT} shows how 
+$\alpha$ varies with temperature when $\sigma=1$ and $\epsilon=1$. 
+% and Figure \ref{fig:erf_Potential} shows the erf potential at reduced 
+%temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and 
+%$\frac{k_BT}{\epsilon}=3$ alongside the WCA potential. 
+%As seen in Figure \ref{fig:erf_potential}, the erf potentials closely 
+%match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, 
+%and for r/$\sigma$ less than ???. % so only values of $V_{erf}$ that are 
+%much less than $k_BT$/$\epsilon$ ??? will be considered. 
+% In these regions $\alpha$ decreases with increasing temperature, 
+%and $\Xi$ increases with increasing temperature.
+ %\begin{figure}[h!]
+    %\centering
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{plot_alpha.pdf}
+      %\caption{$\alpha$ versus $k_BT/\epsilon$}
+    %\end{subfigure} 
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{plot_xi.pdf}
+      %\caption{$\Xi$ versus $k_BT/\epsilon$}
+    %\end{subfigure} 
+    %\caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, 
+    %show how the parameters $\alpha$ and $\Xi$ vary with temperature. 
+    %(a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly 
+    %increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then 
+    %gradually decreases.}
+    %\label{fig:alphaXivsT}
+    %\end{figure}     
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{plot_alpha.pdf}
+    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
+    for $\sigma=1$ and $\epsilon=1$}
+    \label{fig:alphaXivsT}
+  \end{figure}
+%\begin{figure}[h!]
+%    \centering
+    %\includegraphics[height=7cm]{erf_Potential.pdf}
+%    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
+%with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
+%    \label{fig:erf_Potential}
+%  \end{figure}
+%\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
+To find values for $\Xi(T)$ we use the second virial 
+coefficent $B_{2}$ from the equation of state expressed in terms of 
+the virial expansion. 
+\begin{equation}\label{virialequation}
+    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
+\end{equation}
+where 
+\begin{equation}
+	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
+\end{equation}
+The second virial coefficient can be derived from Euler's equation for 
+the internal energy of a system, 
+\begin{equation}U=TS-PV+\mu{N}\end{equation}
+using the relations from statistical thermophysics which views the 
+thermodynamic quantities as ensemble averages
+\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
+\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
+\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
+which, for a canonical ensemble, gives
+\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
+Expressing the Helmholtz free energy in terms of the pair potential 
+$V(r)$ results in the equation of state expressed in terms of the virial 
+expansion as given by Equation~\ref{virialequation}.
+
+The value of $\Xi$ at a given temperature, T, can be found computationally 
+by equating the second virial coefficient evaluated for the error function 
+potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
+WCA potential $V_{WCA}$.
+\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+Evaluating $B_{2}$ for the erf potential gives
+\begin{equation}
+	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
+\end{equation}
+where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, 
+$B_{2WCA}$ can be found computationally from 
+\begin{equation}
+	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
+\end{equation}
+
+A plot of the resulting second virial coefficient $B_2(T)$  with respect 
+to temperature is given in Figure~\ref{B2-vs-T} and a plot of $\Xi(T)$ obtained 
+through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
+Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential at a 
+few different temperatures alongside the WCA potential.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{B2_vs_T_with_xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $B_2$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:B2-vs-T}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $\Xi$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:xi_fromB2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{erf_Potential_xifromB2.pdf}
+    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
+    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
+    shown in black.}
+    \label{fig:erf_Potential_xifromB2}
+  \end{figure}
+
+%\color{red}
+% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
+%\color{black}
+
+%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
+%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
+%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
+%classical fluid.
+
+%MOVE to METHODS end1 >>>
+
+******  SFMT Functional ******
+
+%MOVE to METHODS start2 >>>
+In this paper,  $w_{2}(r)$ will be constructed using the following expresssion 
+for $w_3(r)$ proposed by Schmidt\cite{schmidt2000fluid}.
+\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
+where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
+and $w_3(\infty)=0$. 
+%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
+We replace Shcmidt's parameters $\sigma$ and $a$ with our temperature 
+dependent parameters $\alpha(T)$ and $\Xi(T)$ which were defined above.
+\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
+This gives the expression
+\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
+Using Equation~\ref{w2_w3_relation}, shown again below, 
+\begin{displaymath}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{displaymath}
+and solving for $w_2(r)$
+\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
+we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
+Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
+\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
+gives
+\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
+which can be integrated to obtain the Mayer function
+\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
+The Mayer function for the error function potential $V_{erf}(r)$ is thus 
+given by
+\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
+Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} 
+%This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
+gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
+\begin{equation}{V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} 
+The two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), 
+and are used to compensate for the temperature dependence that enters 
+in through the expression for the error function potential $V_{erf}(r)$.
+%determined to make $V_{erf}(r)$ match $V_{WCA}(r)$ as closely as 
+%possible by compensating for the unwanted temperature dependence that shows 
+%up in the error function potential $V_{erf}(r)$.
+%MOVE to METHODS end2 >>>
+
+**** FUNCTIONAL HERE! *****
+\begin{align}
+\Phi_1 &= -n_{0}ln(1-n_{3}) \\
+\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
+\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
+\end{align}
+
+\begin{align}\label{eq:weights}
+  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
+  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
+  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
+  \\
+  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
+\end{align}
 
@@
 
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{weight_functions.pdf}
-    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2}
-    \label{fig:weight_functions}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME % FIXME \includegraphics[height=5cm]{W2.png}
-      \caption{A crossectional view of $w_2$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      % FIXME \includegraphics[height=5cm]{W3.png}
-      \caption{A crossectional view of $w_3$}
-    \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    \label{fig:W0andW3}
-    \end{figure}
-
-Figure \ref{fig:GaussandW2_actual} ......
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
-    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
-    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
-    and fixed in place as $\vec{dr}$ varies randomly.} 
-  \label{fig:GaussandW2_thinkas}
-  \end{figure}
-
 The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in 
@@
 the weight function.
-%\[{}\]
-
-Monte-Carlo integration is employed to more efficiently 
-compute the weighted densities used to calculate the excess free energy which is then 
-added to the ideal free energy (computed analytically) to get the total free energy 
-which is minimized.
-
-\newpage
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{weight_functions.pdf}
+    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2}
+    \label{fig:weight_functions}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W2.png}
+      \caption{A crossectional view of $w_2$}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W3.png}
+      \caption{A crossectional view of $w_3$}
+    \end{subfigure} 
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    \label{fig:W0andW3}
+    \end{figure}
+
+%MOVE start4 >>>>
+%...............
+%MOVE end4 >>>>
+
+The Liquid (homogeneous) excess free energy $F_{Hexcess}$ is computed 
+analytically for a given $n^*$ and $t^*$ using SFMT with a uniform number 
+density profile, n that does not depend on $\vec r$. 
+%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
+\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
+For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
+%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
+\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
+\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
+The weighted densities are found from
+\begin{displaymath}{n_i=nw_i}\end{displaymath} 
+where the weights $w_i$ are given by Equations~\ref{eq:weights}.
+
+The crystal excess free energy $F_{Cexcess}$ is computed computationally 
+using SFMT with a number density profile $n(\vec r)$ that varies spacially.
+%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
+%Then the crystal, or inhomogeneous, excess free energy is found from
+\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
+Monte-Carlo integration is used to compute the weighted densities 
+$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
+%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
+
+The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
+within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
+the Gaussians is selected that is within the primitive cell, or one of the 
+neighboring cells within a prudently picked inclusion radius as Gaussians 
+farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=5cm]{InclusionRadius.png}
+    \caption{The Gaussians at each lattice point are of width gw. 
+    After a radius of few gw's the Gaussians go to zero and can be neglected.}
+    \label{fig:InclusionRadius}
+  \end{figure}
+
+The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
+$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
+$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
+$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
+where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
+some point in space. Monte-Carlo integration works by randomly generating 
+$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
+\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
+is the gaussian number density profile for no vacancies, the equation:
+
+\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+implemented computationally in the form:
+\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+becomes:
+\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
+where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
+(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
+which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
+$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
+Using the relation:
+\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
+$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
+\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
+
+The result can be made more accurate by increasing NumPoints which is the 
+number of vectors randomly generated. In order to further reduce the error 
+in the average, antithetic variates are used. Rather than computing an average 
+of a set of random points, each random point in the set is paired with a 
+point along the same line as the first point, but opposite in direction, 
+as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
+error in the average by filtering out an erroneous offset of values along 
+the line connecting the two points by averaging the two results, one of 
+which may be erroneously higher than it should be while the other is 
+erronesously lower than it should be.
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=4cm]{AntitheticVariate.png}
+    \caption{Antithetic variates reduce first order error in an average. 
+    Each point in the average is paired with another point from along 
+    the same line as the first, and at an equal, but opposite distance.}
+    \label{fig:AntitheticVariate}
+  \end{figure} 
+Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
+\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
+\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
+Then the entire process repeated for each vector $\vec{r}_k$ to get the 
+total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
+
+The standard error of the mean (SEM) is given by:
+\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
+a measure of the error in $n_3=\overline{w}_3$:
+\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
+NumPoints, the number of vectors randomly generated, is increased until:
+\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
+value that $n_3$ can be without error is 1.
+
+Figure \ref{fig:GaussandW2_actual} ......
+
+ \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
+    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    is generated randomly according to a Gaussian distribution.} 
+  \label{fig:GaussandW2_actual}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
+    and fixed in place as $\vec{dr}$ varies randomly.} 
+  \label{fig:GaussandW2_thinkas}
+  \end{figure}
+
+%Monte-Carlo integration is employed to more efficiently 
+%compute the weighted densities used to calculate the excess free energy which is then 
+%added to the ideal free energy (computed analytically) to get the total free energy 
+%which is minimized.
 
@@
 
-Phase diagrams for the WCA fluid are shown in Figures \ref{fig:Phase_Diagram_of_T_vs_n} 
-and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
-In Figure \ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
+Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
+through \ref{fig:Phase_Diagram_P_vs_T_AS}. In all cases $\sigma=1$ and $\epsilon=1$. 
+In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
 is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
@@
 
-Figure \ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
+Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
+range of reduced temperatures from kT=0.5 to 38. The white lines show the results
+of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
+As can be seen in the figure, the region of coexistence in our simulations 
+falls within the region of coexistance predicted by Ahmed and Sadus at almost 
+all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
+the width of the coexistance regions are nearly the same, while moving to 
+higher temperatures,the width of our coexistance region ranges from about 
+one-half to one-forth the size. Although, our coexistence region is only 
+about one-forth the size at kT=28, it is centered within the coexistence 
+region predicted by Ahmed and Sadus.
+
+\begin{figure}
+  \centering
+  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
+  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
+  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  White lines show the liquid and solid transition densities at each temperature from 
+  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
+  \label{fig:Phase_Diagram_of_T_vs_n}
+\end{figure}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
+  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
+  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  White lines show the liquid and solid transition densities at each temperature from 
+  simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
+  \label{fig:Phase_Diagram_of_T_vs_n_AS}
+\end{figure}
+
+Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
 plotted against the reduced temperature for temperatures of kT = 0.5 to 3.0. Again, the red area 
@@
 than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
-the temperature increases.
-
-\begin{figure}
-  \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
-  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
-  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
-  White lines show the liquid and solid transition densities at each temperature from 
-  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
-  \label{fig:Phase_Diagram_of_T_vs_n}
-\end{figure}
+the temperature increases. Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
+range of reduced temperatures from kT=0.5 to 38, compated with the results
+from Ahmed and Sadus's molecular simulations. Here again, our transition 
+pressures and temperatures are consistently higher.
 
@@
   \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
-  The green line marks the boundary from Monte-Carlo simulations.}
+  The white line marks the boundary from Monte-Carlo simulations.}
   \label{fig:Phase_Diagram_P_vs_T}
@@
 
-%Plots of
-
 \begin{figure}
   \centering
-  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P}
-  \caption{Plot of the reduced temperature verses the reduced density. 
-  The grey region indicates coexistence.}
-  \label{fig:T-vs-n_at_fixed_P}
-\end{figure}
+  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
+  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T_AS}
+  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
+  The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
+  \label{fig:Phase_Diagram_P_vs_T_AS}
+\end{figure}
+
+Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
+reduced pressure versus the reduced temperature at fixed reduced densities 
+from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
+\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
+the WCA fluid transistions into a solid. These downward bends align to 
+form the curved interface separating the blue and red areas in 
+Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
 
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n}
+  \includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
   \caption{Plot of the reduced pressure verses the reduced temperature. 
@@
 
-\begin{figure}
-  \centering
-  %\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
-  \caption{Plot of the reduced pressure verses the reduced volume. 
-  The grey region indicates coexistence.}
-  \label{fig:p-vs-V_at_fixed_T}
-\end{figure}
-
-\begin{figure}
-  \centering
-  %\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
-  \caption{Plot of the reduced pressure verses the reduced density. 
-  The grey region indicates coexistence.}
-  \label{fig:p-vs-n_at_fixed_T}
-\end{figure}
+%More plots...
+
+%\begin{figure}
+  %\centering
+  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P_moreTemp}
+  %\caption{Plot of the reduced temperature verses the reduced density. 
+  %The grey region indicates coexistence.}
+  %\label{fig:T-vs-n_at_fixed_P}
+%\end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
+  %\caption{Plot of the reduced pressure verses the reduced volume. 
+  %The grey region indicates coexistence.}
+  %\label{fig:p-vs-V_at_fixed_T}
+%\end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
+  %\caption{Plot of the reduced pressure verses the reduced density. 
+  %The grey region indicates coexistence.}
+  %\label{fig:p-vs-n_at_fixed_T}
+%\end{figure}
 
 \chapter{Conclusion}
+In this paper we were sucessful in predicting the freezing of a 
+Weeks-Chandler-Anderson fluid using our SFMT functional built by 
+forming temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ to use 
+in Schmidt's erf model to better fit our results to those expected for a
+WCA fluid. We were able to generate a temperature-density 
+phase diagram that yielded coexistence regions that approximately 
+matched the Monte-Carlo and molecular simulation results for 
+values of kT=0.5 to 38. We are also able to generate a pressure-temperature
+phase diagram with a shape that is expected, but the transition 
+temperatures and pressures were too high.
 
@@
 \begin{equation}{w_0(r)=\frac{1}{4{\pi}\sqrt{\pi}a}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}}\end{equation}
-%\[{}\]
+
 \begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
@@
 \begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
-\[{}\]
+
 Calculate $\widetilde{w}_0(k)$ 
 \begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-\[{}\]
+
 \begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
-\[{}\]
+
 Set the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero. 
 \begin{equation}{\widetilde{w}_0(k)=\frac{1}{\sqrt{\pi}ka}\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}}\end{equation}
-\[{}\]
+
 \begin{equation}{\widetilde{w}_0(k)=\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
-\[{}\] 
+
 Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_0(k)$ becomes
@@
 \begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-\[{}\]
+
 \begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
-\[{}\]
+
 \begin{equation}{\widetilde{w}_1(k)=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\frac{2\sin(kr)}{kr}d{r}}\end{equation}
@@
 \begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-\[{}\]
+
 \begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
@@
 \begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-\[{}\]
 
@@
 \begin{displaymath}{\delta_{ij}=\left\{ \begin{array}{rc} 1 & i = j \\ 0  & i\neq j \end{array}\right.}\end{displaymath}
-\[{}\]
-
-%\[{}\]
-%\[{}\]
+
 Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
@@
 $~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
-\[{}\]
+
 Calculate $\widetilde{w}_{{m2}_{y'z'}}(k)$ 
@@
 $~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
-\[{}\]
+
 Calculate $\widetilde{w}_{{m2}_{x'x'}}(k)$ 
@@
 \begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\[{}\]
-%\[{}\]
+
 %\begin{displaymath}{\widetilde{w}_{{m2}_{xx}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\sin^2(\theta)e^{-ikr\cos(\theta)}\left(\int_{0}^{2\pi}\cos^2(\phi)d{\phi}\right)d{\cos(\theta)}\right]d{r}}\end{displaymath}
@@
 \begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
-\[{}\]
+
 \color{blue}
@@
 \color{black}
-\[{}\]
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}-\frac{4\sin(kr)}{k^3r^3}\right)d{r}}\end{displaymath}
 \begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\frac{2\sin(kr)}{kr}\right)d{r}}\end{equation}
-%\[{}\]
+
 Set the lower limit to -infinity since the integrand is nearly zero by the time r is zero. 
@@
 \begin{equation}{+\frac{4\sqrt{\pi}}{k^3a}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right)}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{4\sqrt{\pi}}{3ka}\left(\frac{ka^3}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})+\frac{a\alpha}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})\right)}\end{displaymath}
@@
 \begin{equation}{+\frac{4\sqrt{\pi}}{k^3a}\left(\frac{\pi}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]\right)}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-2\pi{a}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{displaymath}
@@
 Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_{{m2}_{xx}}(k)$ becomes
-\[{}\] \color{green}
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath}
 \begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
-\color{black}
-\[{}\]
+
 Calculate $\widetilde{w}_{{m2}_{y'y'}}(k)$ 
@@
 \begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
@@
 
-\color{blue}
 \begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
 \begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
-\color{black} 
-\[{}\]
+
 This is the same as equation \ref{compare_equation}, and so \begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\widetilde{w}_{{m2}_{x'x'}}(k)}\end{equation}
@@
 \begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\[{}\]
+
 \begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
@@
 
-\color{blue}
 \begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}
@@
 \begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
-\color{black}
 
@@
 \begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})= (\delta{ij}-3\frac{k_ik_j}{k^2})\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
-\[{}\]
+
 %\subsection{Transform of tensor weight ${w}_{m2}(r)$ goes to zero as k goes to zero}
@@
 \begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
-\[{}\]
+
 Term 1: 
-%\color{green}
 \begin{equation}{\left(\frac{-\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{k^2\Xi^2}{8}\right)}\cos(\frac{k\alpha}{2})}\end{equation}
@@
 \begin{displaymath}{\approx-\frac{4\pi}{k^2}+\frac{\pi\alpha^2}{2}+\frac{\pi\Xi^2}{6}}\end{displaymath}
-\color{black}
+
 Term 2:
-%\color{blue}
 \begin{equation}{-\frac{2\pi\alpha}{3k}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\sin(\frac{k\alpha}{2})}\end{equation} 
@@
 \begin{displaymath}{\approx}-\frac{\pi\alpha^2}{3}\end{displaymath}
-\color{black} 
+
 Term 3:
-%\color{red}
 \begin{equation}{\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
@@
 \begin{displaymath}{\approx\frac{4\pi}{k^2}-\frac{\pi}{6}\left(\alpha^2-\Xi^2\right)}\end{displaymath}
-\color{black}
+
 Adding the three terms gives:
@@
 \begin{displaymath}{\approx0}\end{displaymath}
-\[{}\]
 
@@
 \begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation}
-\[{}\]
+
 \begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}+\frac{4\sin(kr)}{k^3r^3}}\end{equation}
-\[{}\]
-\[{}\]
+
 \begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}}}\end{equation}
@@
 \begin{equation}{=a\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})}\end{equation}
-\[{}\]
-\[{}\]
+
 \begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r}}}\end{equation}
@@
 \begin{equation}{=\frac{a\alpha\sqrt{\pi}}{2}e^{-\left(\frac{ka}{2}\right)^2}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\left(\frac{ka}{2}\right)^2}\sin\left(\frac{k\alpha}{2}\right)}\end{equation}
-\[{}\]
+
 \begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}}}\end{equation} 
@@
 \begin{equation}{=a\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{equation}
-\[{}\]
-\[{}\]
+
 \begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}}}\end{equation} 
@@
 \begin{equation}{=\frac{ka^3}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\cos(\frac{k\alpha}{2})+\frac{a\alpha}{2}\sqrt{\pi}e^{-\left(\frac{ka}{2}\right)^2}\sin(\frac{k\alpha}{2})}\end{equation}
-\[{}\]
+
 \begin{equation}{\int_{-\infty}^{\infty}{e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}}}\end{equation} 
@@
 \section{Derivation of the Virial Equation}
-The second virial coefficient can be derived from Euler's equation for the internal energy of a system.
+The second virial coefficient can be derived from Euler's equation for 
+the internal energy of a system.
 \begin{equation}PV=TS-(U-\mu{N})\end{equation}
 \begin{equation}\label{Euler_rearranged}\frac{PV}{k_BT}=\frac{S}{k_B}-\frac{1}{k_BT}(U-\mu{N})\end{equation}
-Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical thermophysics which views the 
-thermodynamic quantities as ensemble averages
+Setting $\beta=\frac{1}{k_BT}$, and putting in the relations from statistical 
+thermophysics which views the thermodynamic quantities as ensemble averages
 \begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
@@
 where $Q_n$ is the configuration integral.
-\color{red}For a system of N particles only, the sum collapses to one term, that for n=N, which gives \color{black}
+\color{red}For a system of N particles only, the sum collapses to one term, 
+that for n=N, which gives \color{black}
 \begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
@@
 \begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Plugging in the expression for $F_{excess}$ in terms of the mayer function for low density fluids
+Plugging in the expression for $F_{excess}$ in terms of the mayer function 
+for low density fluids
 \begin{equation}F_{excess}=k_BT\frac{N^2}{V}\left(-\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
@@
 
-\color{red}This result was for low densities where higher terms in the expression for the excess free energy were neglected. The full virial equation is given by\color{black}
+\color{red}This result was for low densities where higher terms in the 
+expression for the excess free energy were neglected. The full virial 
+equation is given by\color{black}
 \begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}
@@
 \section{Derivation of the Excess Free Energy from Scaled Particle Theory}
-The excess chemical potential of a homogeneous fluid corresdponds to the reversible work of inserting a sphere of radius R 
+The excess chemical potential of a homogeneous fluid corresdponds to the 
+reversible work of inserting a sphere of radius R 
 into a fluid consisting of hard spheres of radius R. 
 \begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
-This result is obtained by considering the difference in the excess Helmholtz free energy when one atom is added to the system. 
-It was shown by Widom to be related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
-of a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+This result is obtained by considering the difference in the excess Helmholtz 
+free energy when one atom is added to the system. It was shown by Widom 
+to be related to the ensemble average of the Boltzman factor for the 
+potential energy difference $V_{diff}$ of a system with N atoms and one 
+with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
 %\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
@@
 %In a randomly positioned (I think it has to be homogeneous!)
-\indent In a collection of hard spheres making up a hard sphere fluid, no energy would be required to place a hard sphere at points 
-surrounded by an empty space the size of a hard sphere. But a hard sphere cannot be placed at points where a hard sphere would 
-overlap another hard sphere; the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
+\indent In a collection of hard spheres making up a hard sphere fluid, 
+no energy would be required to place a hard sphere at points 
+surrounded by an empty space the size of a hard sphere. But a hard sphere 
+cannot be placed at points where a hard sphere would 
+overlap another hard sphere; the energy required to do that would be infinite. 
+Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
 \begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
@@
 \begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
+where $P_{NoOverlap}$ is the probablity of positions existing in the hard 
+fluid that are surrounded by an empty space the size of a hard sphere.
 
@@
     % FIXME \includegraphics[height=3cm]{Probability-of-overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
-             as shown by the black dot at the center of the dotted region near the top. 
-             But it is not possible to place a sphere at a position where an overlap occurs.}
+    \caption{No energy is needed to place a sphere at a position where 
+    there is no overlap such as shown by the black dot at the center of 
+    the dotted region near the top. But it is not possible to place a 
+    sphere at a position where an overlap occurs.}
     \label{fig:p_overlap}
@@
 
-For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
-by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
-that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
-%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
-\color{red}Thus, considering cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
+For a homogeneous system, the probablity of positions existing in the hard 
+fluid that are surrounded by an empty space the size of a hard sphere, 
+$P_{NoOverlap}$, can be found by considering 
+that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the 
+number density times the volume of one sphere. 
+%\color{red}Thus the chemical potential for a fluid with cavities having 
+%a radius equal to the radius R of a hard sphere is \color{black}
+\color{red}Thus, considering cavities having a radius Ro equal to the 
+radius R of a hard sphere, the chemical potential is found to be \color{black}
 %\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
@@
 
-Using this result, scaled particle theory seeks a more general expression of the excess chemical potential of 
-a hard sphere fluid by considering the work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
-where the size of $R_{cavity}$ falls into three different cases: $R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
+Using this result, scaled particle theory seeks a more general expression 
+of the excess chemical potential of a hard sphere fluid by considering the 
+work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
+where the size of $R_{cavity}$ falls into three different cases: 
+$R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
 
@@
 can also be expressed in terms of powers of $R_{cavity}$.  
-The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
+The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at 
+the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
 \begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
@@
 ----------------------------------------------
-
   %\cite{TschoppBHextended}
 
-	%1-21	
-	%\cite{lucretius}
 %\cite{rowlinson1964statistical}
@@
 %  roth2002whitebear}
-%\cite{rosenfeld1989}
+
 	%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
 %\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
-	%\cite{schmidt1999density}
 %\cite{schmidt2000density, groh2001density, kim2001adsorption,
 %  sweatman2002fundamental}
-	%\cite{schmidt2000fluid, sweatman2002fundamental}
+
 %\cite{weeks1971}
 	%\cite{rosenfeld1989}
-	%\cite{schmidt1999density}
 	%%\cite{tarazonaFIXME,santos2012phi3}
@@
 	%\cite{schmidt2000fluid}
-	%\cite{schmidt1999density, schmidt2000fluid}
-	%\cite{schmidt1999density}
-	%\citenum{schmidt2000fluid}
-	%\citenum{schmidt2000fluid}
-	%a Face Centered Cubic (FCC) lattice is formed~\cite{reference!}
+
 %\cite{barker1967perturbation}
@@
 	%\cite{roth2002whitebear}
+
 %\cite{mikolaj2004structure, eggert2002quantitative,

### 2020-10-23 papers/thesis-kirstie/thesis.tex
edits

@@
     \centering
-    \includegraphics[height=3.5cm]{TwoSpheres.pdf} 
+    \includegraphics[height=3.5cm]{figs/TwoSpheres.pdf} 
     \caption{(Left) The WCA potential is a minimum at r=2R. 
@@
 
-%MOVE to METHODS start >>>
-%........................
-%MOVE to METHODS  end  >>>
-
+\newpage
 \section{The Theory of Freezing and Maxwell's Double Tangent Construction}
@@
 
-%move this paragraph to methods?
-Although the entire crystal lattice can be constructed with FCC cells, 
-FCC cells are not the smallest cells with which the lattice can be built. 
-The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell_firsttime} 
-is the smallest structure that, when repeated, re-creates the entire crystal 
-lattice. The parallepiped primitive cell takes up one-fourth the volume 
-of an FCC cell, includes portions of eight atoms for a total of one full 
-atom, and has eight lattice sites convenietly reachable by moving along 
-one of three different lattice vectors to step from one atom to the next. 
-The three lattice vectors will be useful in creating the lattice structure 
-computationally, and will also make the program run faster.
-
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
-    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
-    \label{fig:primitivecell_firsttime}
-  \end{figure}
-%-------------------------------
-
 For pure substances, a phase transition occurs along a line on a 
@@
     \centering
-    \includegraphics[height=7cm]{P-T_Diagram.png}
+    \includegraphics[height=6cm]{P-T_Diagram.png}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
     \centering
-     \includegraphics[height=8cm]{MaxwellDTC-Fig2.pdf}
+     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
      %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
@@
 at the transition point can be found the same way, as shown in Figure 
-\ref{fig:Pvsinvn}.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{MaxwellDTC-Fig3.pdf}
+\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
+can be generated by repeating this procedure for various temperatures. 
+%The lowest density the crystal exhibts and the highest density the fluid 
+%exhibits at the phase transistion can then be identified for each temperature.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
     % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
@@
 
-A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
-can be generated by repeating this procedure for various temperatures. 
-%The lowest density the crystal exhibts and the highest density the fluid 
-%exhibits at the phase transistion can then be identified for each temperature.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{T-n_Diagram.png}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{T-n_Diagram.png}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
@@
 of particles are fixed, the change in the free energy of the system will 
-always be negative, and thus the free energy tends toward a minimum. 
-(reference schroder) This illustrates the tendancy toward minimization 
+always be negative, and thus the free energy tends toward a 
+minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
 for a system near equilibrium. A general proof of minimization can be 
-found in (reference simple fluids).
+found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
 
@@
 %\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
-Using this potential, and setting the external potential $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in Equation~\ref{Fexcess} becomes
+Using this potential, and setting the external potential 
+$V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
+Equation~\ref{Fexcess} becomes
 
@@
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$. 
-%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    %$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%
+%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
+%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
 The function $f_{ij}$ is called a Mayer function, and is given by
@@
 Multiplying the Mayer function terms, the excess free energy can be written as
-\begin{equation}\label{Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +...
+\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +...
 \right]\right)  }}\end{equation}
@@
     \centering
-    %\includegraphics[height=7cm]{diagrammic.pdf}
+    \includegraphics[height=7cm]{figs/diagrammic.pdf}
     \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations 
@@
 Each pair can then be described by a function of the distance between the 
-two atoms in the pair. Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ 
-for each pair we get 
-\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
-\color{black}
+two atoms in the pair. 
+%Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ for each pair we get 
+%\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
 With a careful counting of the resulting number of such pairs, the excess 
-free energy can then be written as 
-\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
-where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
-Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
-for a homogeneous density n=$\frac{N}{V}$ this becomes
-\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
-Thus the free excess energy per atom in the low-density limit is given by 
+free energy can then be written~\cite{schroeder} as
+%\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
+%where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
+%Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
+%for a homogeneous density n=$\frac{N}{V}$ this becomes
+%\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
+%Thus the free excess energy per atom in the low-density limit is given by 
 \begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
+in the thermodynamic limit of large N 
 where the quantity in parenthesis can be recognized 
@@
 Outside of the low density limit, it is quite difficult to evaluate the 
-excess Helmholtz free energy given in Equation~\ref{Fexcess-simplified}.
+excess Helmholtz free energy given in Equation~\ref{eq:Fexcess-simplified}.
 
@@
 is a function of the number density n. For a spatially varying number density, 
-as would be expected for a crystal, the free energy is a functional of 
-the number density profile $n(\vec{r})$  
+as would be expected for a crystal, the free energy becomes a functional of 
+the number density profile $n(\vec{r})$.  
 \begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-where the ideal Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}.
-To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
-the temperature fixed until the free energy is minimized. The minimum free 
-energy and its corresponding number density profile are then the equilibrium 
-values for that temperature.
+The functional for the ideal Helmholtz free energy is given by 
+Equation~\ref{eq:Fideal-n(r)}, and the functional for the excess Helmholtz 
+free energy depends on the particular fluid as defined by the spheres' 
+interatomic potential V(r). To implement cDFT, the number density profile 
+$n(\vec{r})$ is varied with the temperature fixed until the free energy 
+is minimized. The minimum free energy and its corresponding number density 
+profile are then the equilibrium values for that temperature.
 
 \section{Soft Fundamental Measure Theory (SFMT)}
-Another way to form an expression for the excess Helmholtz free energy in 
-terms of a homogeneous number density is to use Scaled Particle Theory (SPT).
-Scaled particle theory seeks a general expression for the excess chemical 
+In order to use cDFT, a functional for the excess Helmholtz free
+energy for the particular fluid being modeled must be formed. But forming 
+an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
+For the Hard Sphere fluid, there is another way to form an expression for 
+the excess Helmholtz free energy in terms of a homogeneous number density 
+and, that is to use Scaled Particle Theory (SPT).
+
+Scaled Particle Theory seeks a general expression for the excess chemical 
 potential of a homogeneous Hard Sphere fluid which can then be integrated 
@@
 
-%MOVE to METHODS start2 >>>
-%........................
-%MOVE to METHODS end >>>
-
-%MOVE to METHODS start3 >>>
-%........................
-%MOVE to METHODS end >>>
-
 \chapter{Methods}
 
-%MOVE to METHODS start3 >>>
-
+\section{Overview}
 We use classical Density Functional Theory (cDFT) in conjunction 
@@
 free energy is lower than the crystal free energy.
-%MOVE to METHODS end3 >>>
-
-%MOVE start4 >>>>
 
@@
 
-  \begin{figure}[h!]
-    \centering
-    \includegraphics[height=4.5cm]{PrimitiveCellLightBlue.png}
+%move this paragraph to methods?
+-----------*****-------------
+
+Although the entire crystal lattice can be constructed with FCC cells, 
+FCC cells are not the smallest cells with which the lattice can be built. 
+The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell} 
+is the smallest structure that, when repeated, re-creates the entire crystal 
+lattice. The parallepiped primitive cell takes up one-fourth the volume 
+of an FCC cell, includes portions of eight atoms for a total of one full 
+atom, and has eight lattice sites convenietly reachable by moving along 
+one of three different lattice vectors to step from one atom to the next. 
+The three lattice vectors will be useful in creating the lattice structure 
+computationally, and will also make the program run faster.
+
+----------****--------------
+%-------------------------------
+
+  \begin{figure}[h!]
+    \centering
+    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
     \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
@@
     \centering
-    %\includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
+    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
     % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
@@
 densities and temperatures.
-%MOVE end4 >>>>
 
@@
 *****  end  stuff?  ******
-\newline
-\newline
-
-**** Fideal - liquid and crystal  ****
-
-The Liquid (homogeneous) ideal free energy $F_{Hideal}$ is computed analytically for a given $n^*$ and $t^*$, where $\sigma=1$ so that $n^*=n\sigma^3$=n, and $\epsilon=1$ so that $t^*=\frac{k_BT}{\epsilon}=k_BT$.
+
+\section{Computing the Liquid and Crystal Ideal Helmholtz free energies}
+
+The liquid ideal Helmholtz free energy $F_{Hideal}$ is computed 
+analytically for a given homogeneous reduced density $n^*$, and reduced 
+temperature $t^*$ from the equation 
 %\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} where
-
-\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  and
-\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units!\color{black}\end{displaymath} 
+\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} 
+where $\sigma=1$ (which gives $n^*=n\sigma^3$=n), and $\epsilon=1$ (which 
+gives $t^*=\frac{k_BT}{\epsilon}=k_BT$). The quantity $\Lambda$ is the DeBroglie 
+wavelength given by
+\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
+where in our program we have set
+\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
 %\begin{displaymath}{m=}\end{displaymath}  
@@
 
-The Crystal ideal free energy $F_{Cideal}$ is computed in two different ways from the equation 
+The crystal ideal Helmholtz free energy $F_{Cideal}$ is computed from the equation 
 \begin{displaymath}{F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV}\end{displaymath}
@@
 \end{displaymath}
-\begin{itemize}
-  \item For small gw values it is assumed that the Guassians do not overlap, in
-  which case the equation for the free energy can be simplified and solved
+For small gw values it is assumed that the Guassians do not overlap, 
+in which case the equation for the free energy can be simplified, and solved
 analytically with the result
-  \begin{align}
+\begin{equation}
  F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
-  \end{align}
-\item For larger gw values it is assumed that the Gaussians do overlap, 
-and so the free energy is caluculated computationally. First dx is scaled by gw
-\begin{align}
-  dx_{scaled} &= (dx_{input})(gw)
-\end{align}
-then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the primitive cell taking into account the contribution of Gaussians from many (5x5x5) nearby cells whose centers are given by $R_i$.
+\end{equation}
+For larger gw values it is assumed that the Gaussians do overlap, 
+and so the free energy is caluculated computationally. First dx is scaled by gw,
+\begin{equation}
+  dx_{scaled} = (dx_{input})(gw)
+\end{equation}
+and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
+primitive cell taking into account the contribution of Gaussians from 
+many (5x5x5) nearby cells whose centers are given by $R_i$.
 %\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
@@
 volume.
-\end{itemize}
-
-******  Fexcess depends on V(r) ... use SFMT for liquid and crystal ******
-
+
+\section{Our Functional for the Excess Helmholtz free energy}
+
+The excess Helmholtz free energy is particular to the fluid being modeled, 
+as it depends on the interatomic potential V(r). 
 %MOVE to METHODS start1 >>>
@@
 \end{equation}
-where the parameter $\alpha$ is derived by setting the erf potential equal to the WCA 
-potentail when r equals $\alpha$ 
+
+Unlike the WCA potential, the erf potential depends on temperature, but 
+we temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ to compensate for the
+variance with temperature. The parameter $\alpha(T)$ is derived by setting 
+the erf potential equal to the WCA potential when r equals $\alpha$ 
 \begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
@@
 \begin{equation}\label{alphaT}
-	{\alpha=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
+	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
 \end{equation} 
-
-Unlike the WCA potential, the erf potential depends on temperature through 
-the parameters $\alpha$ and $\Xi$. Figure \ref{fig:alphaXivsT} shows how 
+%%%%Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$. 
+Figure \ref{fig:alphaXivsT} shows how 
 $\alpha$ varies with temperature when $\sigma=1$ and $\epsilon=1$. 
@@
   \end{figure}
-%\begin{figure}[h!]
-%    \centering
-    %\includegraphics[height=7cm]{erf_Potential.pdf}
-%    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
-%with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ shown in black.}
-%    \label{fig:erf_Potential}
-%  \end{figure}
-%\color{red}\noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!\color{black}
+
 To find values for $\Xi(T)$ we use the second virial 
@@
 \end{equation}
-
 A plot of the resulting second virial coefficient $B_2(T)$  with respect 
-to temperature is given in Figure~\ref{B2-vs-T} and a plot of $\Xi(T)$ obtained 
+to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
 through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
-Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential at a 
-few different temperatures alongside the WCA potential.
+Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
+$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
+potential which is not temperature dependent.
 
@@
     at each temperature.}
-    \label{fig:B2-vs-T}
+    \label{fig:B2vsT}
   \end{figure}
@@
 
-%MOVE start4 >>>>
-%...............
-%MOVE end4 >>>>
-
-The Liquid (homogeneous) excess free energy $F_{Hexcess}$ is computed 
-analytically for a given $n^*$ and $t^*$ using SFMT with a uniform number 
-density profile, n that does not depend on $\vec r$. 
+\section{Computing the Liquid and Crystal Excess Helmholtz free energies}
+
+The liquid excess free energy $F_{Hexcess}$ is computed analytically for 
+a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
+using SFMT with a uniform number density profile, n that does not depend 
+on $\vec r$. 
 %done in hf.energy() note- does not depend on fv or gw but does use SFMT):

### 2020-10-26 papers/thesis-kirstie/thesis.tex
edits

@@
 minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
-for a system near equilibrium. A general proof of minimization can be 
-found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
+for a system near equilibrium. A general proof of minimization 
+can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
 
@@
 
-Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
-\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-is a function of the number density n. For a spatially varying number density, 
-as would be expected for a crystal, the free energy becomes a functional of 
-the number density profile $n(\vec{r})$.  
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
+%Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
+%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+%is a function of the number density n. For a spatially varying number density, 
+%as would be expected for a crystal, the free energy becomes a functional of 
+%the number density profile $n(\vec{r})$.
+
 The functional for the ideal Helmholtz free energy is given by 
@@
 $n(\vec{r})$ is varied with the temperature fixed until the free energy 
-is minimized. The minimum free energy and its corresponding number density 
+is minimized. 
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+The minimum free energy and its corresponding number density 
 profile are then the equilibrium values for that temperature.
@@
 A simplified flowchart of the program created to investigate the freezing 
-of a WCA fluid is shown on %pages x-x. 
-The program takes four inputs: reduced number density n*, reduced temperature t*, 
-a range of Gaussian 
-widths gw, and a range of fraction of vacancies fv.The number density 
+of a WCA fluid is shown below. The program takes four inputs: reduced number 
+density n*, reduced temperature t*, a range of Gaussian 
+widths gw, and a range of fraction of vacancies fv. The number density 
 profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
@@
 
-\textbf{Program Implementation}
-
+%Program Implementation
 The program computes both the crystal and the homogeneous Helmholtz free 
@@
 \newline
-*****   start stuff?  ******
+*****   start stuff  ******
 
@@
 
-*****  end  stuff?  ******
+*****  end  stuff  ******
 
@@
 
-The excess Helmholtz free energy is particular to the fluid being modeled, 
-as it depends on the interatomic potential V(r). 
-%MOVE to METHODS start1 >>>
-The WCA potential is not well suited for implementing Density Functional 
-Theory.  Therefore, an erf potential which closely approximates the WCA 
-potential at temperatures of interest will be used instead of the WCA 
-potential. The erf potential is given by 
-\begin{equation}
-	{V_{erf}=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}
-\end{equation}
-
-Unlike the WCA potential, the erf potential depends on temperature, but 
-we temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ to compensate for the
-variance with temperature. The parameter $\alpha(T)$ is derived by setting 
-the erf potential equal to the WCA potential when r equals $\alpha$ 
-\begin{displaymath}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{displaymath} 
-and is given by 
-\begin{equation}\label{alphaT}
-	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
-\end{equation} 
-%%%%Unlike the WCA potential, the erf potential depends on temperature through the parameters $\alpha$ and $\Xi$. 
-Figure \ref{fig:alphaXivsT} shows how 
-$\alpha$ varies with temperature when $\sigma=1$ and $\epsilon=1$. 
-% and Figure \ref{fig:erf_Potential} shows the erf potential at reduced 
-%temperatures $\frac{k_BT}{\epsilon}=1$, $\frac{k_BT}{\epsilon}=2$, and 
-%$\frac{k_BT}{\epsilon}=3$ alongside the WCA potential. 
-%As seen in Figure \ref{fig:erf_potential}, the erf potentials closely 
-%match the WCA potential for reduced temperatures below $\frac{k_BT}{\epsilon}=2$, 
-%and for r/$\sigma$ less than ???. % so only values of $V_{erf}$ that are 
-%much less than $k_BT$/$\epsilon$ ??? will be considered. 
-% In these regions $\alpha$ decreases with increasing temperature, 
-%and $\Xi$ increases with increasing temperature.
- %\begin{figure}[h!]
-    %\centering
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{plot_alpha.pdf}
-      %\caption{$\alpha$ versus $k_BT/\epsilon$}
-    %\end{subfigure} 
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{plot_xi.pdf}
-      %\caption{$\Xi$ versus $k_BT/\epsilon$}
-    %\end{subfigure} 
-    %\caption{These plots, generated with $\sigma=1$ and $\epsilon=1$, 
-    %show how the parameters $\alpha$ and $\Xi$ vary with temperature. 
-    %(a) $\alpha$ decreases with increasing temperature. (b) $\Xi$ rapidly 
-    %increases for temperatures below $\frac{k_BT}{\epsilon}=52$ and then 
-    %gradually decreases.}
-    %\label{fig:alphaXivsT}
-    %\end{figure}     
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=9cm]{plot_alpha.pdf}
-    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
-    for $\sigma=1$ and $\epsilon=1$}
-    \label{fig:alphaXivsT}
-  \end{figure}
-
-To find values for $\Xi(T)$ we use the second virial 
-coefficent $B_{2}$ from the equation of state expressed in terms of 
-the virial expansion. 
-\begin{equation}\label{virialequation}
-    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
-\end{equation}
+%The excess Helmholtz free energy is particular to the fluid being modeled, 
+%as it depends on the interatomic potential V(r). 
+We use Soft Fundamental Measure Theory to find the excess Helmholtz free energy.
+The functional is written in general form as 
+\begin{displaymath}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
 where
-\begin{equation}
-	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
-\end{equation}
-The second virial coefficient can be derived from Euler's equation for 
-the internal energy of a system, 
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-using the relations from statistical thermophysics which views the 
-thermodynamic quantities as ensemble averages
-\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
-\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
-\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-which, for a canonical ensemble, gives
-\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Expressing the Helmholtz free energy in terms of the pair potential 
-$V(r)$ results in the equation of state expressed in terms of the virial 
-expansion as given by Equation~\ref{virialequation}.
-
-The value of $\Xi$ at a given temperature, T, can be found computationally 
-by equating the second virial coefficient evaluated for the error function 
-potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
-WCA potential $V_{WCA}$.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
-Evaluating $B_{2}$ for the erf potential gives
-\begin{equation}
-	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
-\end{equation}
-where $\alpha(T)$ is given by Equation~\ref{alphaT} as before. Similarly, 
-$B_{2WCA}$ can be found computationally from 
-\begin{equation}
-	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
-\end{equation}
-A plot of the resulting second virial coefficient $B_2(T)$  with respect 
-to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
-through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
-Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
-$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
-potential which is not temperature dependent.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=9cm]{B2_vs_T_with_xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $B_2$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:B2vsT}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=9cm]{xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $\Xi$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:xi_fromB2vsT}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=9cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
-    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
-    shown in black.}
-    \label{fig:erf_Potential_xifromB2}
-  \end{figure}
-
-%\color{red}
-% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
-%\color{black}
-
-%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
-%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
-%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
-%classical fluid.
-
-%MOVE to METHODS end1 >>>
-
-******  SFMT Functional ******
-
-%MOVE to METHODS start2 >>>
-In this paper,  $w_{2}(r)$ will be constructed using the following expresssion 
-for $w_3(r)$ proposed by Schmidt\cite{schmidt2000fluid}.
+\begin{align}
+\Phi_1 &= -n_{0}ln(1-n_{3}) \\
+\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
+\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
+\end{align} 
+and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
+for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
+\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
+where the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
+\begin{align}\label{eq:weights}
+  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
+  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
+  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
+  \\
+  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
+\end{align}
+The weight fucntions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
+Schmidt's proposed erf model~\cite{schmidt2000fluid} in which he set
 \begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
@@
 %\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
-We replace Shcmidt's parameters $\sigma$ and $a$ with our temperature 
-dependent parameters $\alpha(T)$ and $\Xi(T)$ which were defined above.
+Our functional is  created by replacing Schmidt's parameters $\sigma$ 
+and $a$ with our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$.
 \begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
-This gives the expression
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]\end{equation} 
+
+The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
+temperature dependence that enters in through the expression for the 
+error function potential $V_{erf}(r)$ which is used to approximate the
+non-temperature dependent WCA potential that is more difficult to 
+implement computationally. The erf potential emerges from the relationships
+that the weight functions must satisfy.
+%potential at temperatures of interest will be used instead of the WCA potential. 
 Using Equation~\ref{w2_w3_relation}, shown again below, 
@@
 gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
-\begin{equation}{V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]}\end{equation} 
-The two parameters $\alpha$ and $\Xi$ are given in equations (3) and (4), 
-and are used to compensate for the temperature dependence that enters 
-in through the expression for the error function potential $V_{erf}(r)$.
-%determined to make $V_{erf}(r)$ match $V_{WCA}(r)$ as closely as 
-%possible by compensating for the unwanted temperature dependence that shows 
-%up in the error function potential $V_{erf}(r)$.
-%MOVE to METHODS end2 >>>
-
-**** FUNCTIONAL HERE! *****
-\begin{align}
-\Phi_1 &= -n_{0}ln(1-n_{3}) \\
-\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
-\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
-\end{align}
-
-\begin{align}\label{eq:weights}
-  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
-  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
-  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
-  \\
-  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
-\end{align}
-
+\begin{equation}
+	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
+\end{equation}
+
+Expressions for $\alpha(T)$ and $\Xi(T)$  can be found as follows. 
+The parameter $\alpha(T)$ is derived by setting 
+the erf potential equal to the WCA potential when r equals $\alpha$ 
+\begin{equation}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
+and is given by 
+\begin{equation}\label{alphaT}
+	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
+\end{equation} 
+Figure \ref{fig:alphaXivsT} shows how $\alpha$ varies with temperature 
+when $\sigma=1$ and $\epsilon=1$.    
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{plot_alpha.pdf}
+    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
+    for $\sigma=1$ and $\epsilon=1$}
+    \label{fig:alphaXivsT}
+  \end{figure}
+The value of $\Xi$ at a given temperature, T, can be found computationally 
+by equating the second virial coefficient evaluated for the error function 
+potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
+WCA potential $V_{WCA}$.
+\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+The second virial coefficient can be derived from Euler's equation for 
+the internal energy of a system, 
+\begin{equation}U=TS-PV+\mu{N}\end{equation}
+using the relations from statistical thermophysics which views the 
+thermodynamic quantities as ensemble averages
+\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
+\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
+\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
+which, for a canonical ensemble, gives
+\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
+Expressing the Helmholtz free energy in terms of the pair potential 
+$V(r)$ results in an equation of state expressed in terms of the virial expansion 
+\begin{equation}\label{virialequation}
+    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
+\end{equation}
+where 
+\begin{equation}
+	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
+\end{equation}
+Evaluating $B_{2}$ for the erf potential gives
+\begin{equation}
+	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
+\end{equation}
+where $\alpha(T)$ is given by Equation~\ref{alphaT}. Similarly, 
+$B_{2WCA}$ can be found computationally from 
+\begin{equation}
+	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
+\end{equation}
+
+A plot of the resulting second virial coefficient $B_2(T)$  with respect 
+to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
+through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
+Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
+$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
+potential which is not temperature dependent.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{B2_vs_T_with_xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $B_2$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:B2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $\Xi$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:xi_fromB2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{erf_Potential_xifromB2.pdf}
+    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
+    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
+    shown in black.}
+    \label{fig:erf_Potential_xifromB2}
+  \end{figure}
+
+%\color{red}
+% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
+%\color{black}
+
+%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
+%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
+%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
+%classical fluid.
+
+\color{red}
 Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions 
@@
     \includegraphics[height=6cm]{weight_functions.pdf}
-    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    \caption{\color{red}RERUN for NEW XI!\color{black}Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
     z-axis for a reduced temperature of 2}
@@
     \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    \caption{\color{red}RERUN for NEW XI!\color{black}Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
     spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
@@
     \end{figure} 
+\color{black}
 
@@
 \begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}
-
-\section{Derivation of the Excess Free Energy from Scaled Particle Theory}
-The excess chemical potential of a homogeneous fluid corresdponds to the 
-reversible work of inserting a sphere of radius R 
-into a fluid consisting of hard spheres of radius R. 
-\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
-This result is obtained by considering the difference in the excess Helmholtz 
-free energy when one atom is added to the system. It was shown by Widom 
-to be related to the ensemble average of the Boltzman factor for the 
-potential energy difference $V_{diff}$ of a system with N atoms and one 
-with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
-%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
-\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
-\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-%Equation~\ref{widom} is called Widom's Insertion Formula. 
-\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
-\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
-%In a randomly positioned (I think it has to be homogeneous!)
-\indent In a collection of hard spheres making up a hard sphere fluid, 
-no energy would be required to place a hard sphere at points 
-surrounded by an empty space the size of a hard sphere. But a hard sphere 
-cannot be placed at points where a hard sphere would 
-overlap another hard sphere; the energy required to do that would be infinite. 
-Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
-\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-where $P_{NoOverlap}$ is the probablity of positions existing in the hard 
-fluid that are surrounded by an empty space the size of a hard sphere.
-
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=3cm]{Probability-of-overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where 
-    there is no overlap such as shown by the black dot at the center of 
-    the dotted region near the top. But it is not possible to place a 
-    sphere at a position where an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
-
-For a homogeneous system, the probablity of positions existing in the hard 
-fluid that are surrounded by an empty space the size of a hard sphere, 
-$P_{NoOverlap}$, can be found by considering 
-that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the 
-number density times the volume of one sphere. 
-%\color{red}Thus the chemical potential for a fluid with cavities having 
-%a radius equal to the radius R of a hard sphere is \color{black}
-\color{red}Thus, considering cavities having a radius Ro equal to the 
-radius R of a hard sphere, the chemical potential is found to be \color{black}
-%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-\begin{equation}\label{mu_excess}\mu_{ex}(R_{cavity}=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-Using this result, scaled particle theory seeks a more general expression 
-of the excess chemical potential of a hard sphere fluid by considering the 
-work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
-where the size of $R_{cavity}$ falls into three different cases: 
-$R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
-
-\begin{equation}\label{3_parts_to_mu_ex}{W_{rev}(R_{cavity})=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_{cavity}^3 & R_{cavity} >> R 
-\\ P\frac{4\pi}{3}R_{cavity}^3 + c_2R_{cavity}^2 + c_1R_{cavity}^1 + c_0R_{cavity}^0  & R_{cavity} > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity})^3)) & R_{cavity} \leq R\end{array}\right.}\end{equation}
-
-%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  
-%must tend toward in the limit that $R_o$ goes to infinity. By expanding the logrythmic term (ln(1-x)= ...) 
-%in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.
-
-%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
-The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_{cavity}$, 
-and so it is assumed that the middle function, which is valid for values of $R_{cavity}$ in between the two extremes, 
-can also be expressed in terms of powers of $R_{cavity}$.  
-The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at 
-the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
-\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
-where $\eta$ is the packing fraction given by
-\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}
-
-%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$. 
-%From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation 
-%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
-%From the expression for pressure given by
-
-The pressure can be obtained from \color{red}
-\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
-%The excess Hemlholtz free energy is then found from the relation
-and the Helmholz free energy from
-\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
-which gives
-%\begin{displaymath}F_{ex}=\end{displaymath}
-\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}
 
@@
 ----------------------------------------------
+
+\bibliography{thesis}
+\bibliographystyle{unsrt}
+
+\end{document}
+
   %\cite{TschoppBHextended}
@@
 	%\cite{verlet1967computer}
-
-\bibliography{thesis}
-\bibliographystyle{unsrt}
-
-\end{document}
 
@@
 --------------------------------------------------
+
+%\section{Derivation of the Excess Free Energy from Scaled Particle Theory}
+%The excess chemical potential of a homogeneous fluid corresdponds to the 
+%reversible work of inserting a sphere of radius R 
+%into a fluid consisting of hard spheres of radius R. 
+%\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
+%This result is obtained by considering the difference in the excess Helmholtz 
+%free energy when one atom is added to the system. It was shown by Widom 
+%to be related to the ensemble average of the Boltzman factor for the 
+%potential energy difference $V_{diff}$ of a system with N atoms and one 
+%with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
+%%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
+%\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
+%\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+%%Equation~\ref{widom} is called Widom's Insertion Formula. 
+%\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
+%\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
+%%In a randomly positioned (I think it has to be homogeneous!)
+%\indent In a collection of hard spheres making up a hard sphere fluid, 
+%no energy would be required to place a hard sphere at points 
+%surrounded by an empty space the size of a hard sphere. But a hard sphere 
+%cannot be placed at points where a hard sphere would 
+%overlap another hard sphere; the energy required to do that would be infinite. 
+%Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
+%\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
+%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
+%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
+%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
+%where $P_{NoOverlap}$ is the probablity of positions existing in the hard 
+%fluid that are surrounded by an empty space the size of a hard sphere.
+
+%\begin{figure}[h!]
+    %\centering
+    %% FIXME \includegraphics[height=3cm]{Probability-of-overlap.pdf}
+    %\caption{No energy is needed to place a sphere at a position where 
+    %there is no overlap such as shown by the black dot at the center of 
+    %the dotted region near the top. But it is not possible to place a 
+    %sphere at a position where an overlap occurs.}
+    %\label{fig:p_overlap}
+  %\end{figure}
+
+%For a homogeneous system, the probablity of positions existing in the hard 
+%fluid that are surrounded by an empty space the size of a hard sphere, 
+%$P_{NoOverlap}$, can be found by considering 
+%that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the 
+%number density times the volume of one sphere. 
+%%\color{red}Thus the chemical potential for a fluid with cavities having 
+%%a radius equal to the radius R of a hard sphere is \color{black}
+%\color{red}Thus, considering cavities having a radius Ro equal to the 
+%radius R of a hard sphere, the chemical potential is found to be \color{black}
+%%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
+%\begin{equation}\label{mu_excess}\mu_{ex}(R_{cavity}=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+
+%%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
+
+%Using this result, scaled particle theory seeks a more general expression 
+%of the excess chemical potential of a hard sphere fluid by considering the 
+%work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
+%where the size of $R_{cavity}$ falls into three different cases: 
+%$R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
+
+%\begin{equation}\label{3_parts_to_mu_ex}{W_{rev}(R_{cavity})=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_{cavity}^3 & R_{cavity} >> R 
+%\\ P\frac{4\pi}{3}R_{cavity}^3 + c_2R_{cavity}^2 + c_1R_{cavity}^1 + c_0R_{cavity}^0  & R_{cavity} > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity})^3)) & R_{cavity} \leq R\end{array}\right.}\end{equation}
+
+%%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  
+%%must tend toward in the limit that $R_o$ goes to infinity. By expanding the logrythmic term (ln(1-x)= ...) 
+%%in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.
+
+%%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
+%The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_{cavity}$, 
+%and so it is assumed that the middle function, which is valid for values of $R_{cavity}$ in between the two extremes, 
+%can also be expressed in terms of powers of $R_{cavity}$.  
+%The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at 
+%the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
+%\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
+%where $\eta$ is the packing fraction given by
+%\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}
+
+%%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$. 
+%%From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation 
+%%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
+%%From the expression for pressure given by
+
+%The pressure can be obtained from \color{red}
+%\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
+%%The excess Hemlholtz free energy is then found from the relation
+%and the Helmholz free energy from
+%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
+%which gives
+%%\begin{displaymath}F_{ex}=\end{displaymath}
+%\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}

### 2020-10-27 papers/thesis-kirstie/thesis.tex
add equation numbers

@@
 %\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)}\end{displaymath} 
+\begin{align}
+  \frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
+\end{align}
 where $\sigma=1$ (which gives $n^*=n\sigma^3$=n), and $\epsilon=1$ (which 
@@
 The crystal ideal Helmholtz free energy $F_{Cideal}$ is computed from the equation 
-\begin{displaymath}{F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV}\end{displaymath}
+\begin{align}
+  F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
+\end{align}
 with the varying number density profile given by 

### 2020-10-27 papers/thesis-kirstie/thesis.tex
new Xi in the plot

@@
     \includegraphics[height=6cm]{weight_functions.pdf}
-    \caption{\color{red}RERUN for NEW XI!\color{black}Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2}
+    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2.}
     \label{fig:weight_functions}
@@
       \includegraphics[height=5cm]{W2.png}
+      % \includegraphics[width=5cm]{figs/w2-ring}
       \caption{A crossectional view of $w_2$}
@@
     \end{subfigure} 
-    \caption{\color{red}RERUN for NEW XI!\color{black}Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
     spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 

### 2020-10-29 papers/thesis-kirstie/thesis.tex
latex cleanup
edits

@@
 energy of a system, and the corresponding equilibrium number density profile. 
-In particular, SFMT is used to form the expression for the excess free energy 
-without having to take into account the interaction each sphere, or atom, 
-has with every other sphere. The excess free energy is then added to the 
-ideal free energy calculated analytically to find the resulting free energy 
-for the system. 
+In particular, SFMT is used to form the expression for the excess Helmholtz 
+free energy without having to take into account the interaction each sphere, 
+or atom, has with every other sphere. The excess free energy is then added 
+to the ideal free energy calculated analytically to find the resulting 
+free energy for the system. 
 \begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
@@
 
-***** DFT uses n(r) for BOTH Fideal and Fexcess ******
-
-To implement cDFT computationally, we begin by picturing one single crystal 
-cell. Although the crystal structure is composed of Face Centered Cubic (FCC) 
-cells, the entire lattice can be generated with a cell one-fourth the size 
-called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}. 
-The primitive cell has the shape of a parallelpiped, and contains the 
-equivalent of one atom.
-
-%move this paragraph to methods?
------------*****-------------
-
 Although the entire crystal lattice can be constructed with FCC cells, 
 FCC cells are not the smallest cells with which the lattice can be built. 
-The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell} 
+The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
 is the smallest structure that, when repeated, re-creates the entire crystal 
-lattice. The parallepiped primitive cell takes up one-fourth the volume 
-of an FCC cell, includes portions of eight atoms for a total of one full 
-atom, and has eight lattice sites convenietly reachable by moving along 
-one of three different lattice vectors to step from one atom to the next. 
-The three lattice vectors will be useful in creating the lattice structure 
-computationally, and will also make the program run faster.
-
-----------****--------------
-%-------------------------------
+lattice. This parallepiped primitive cell takes up one-fourth the volume 
+of an FCC cell, and contains the equivalent of one atom.
+It includes portions of eight atoms each located on a lattice site convenietly 
+reachable by moving along one of three different lattice vectors to step 
+from one atom to the next. The three lattice vectors will be useful in 
+creating the lattice structure computationally, and will also make the 
+program run faster.
 
@@
 
-  \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
-    \caption{A Simplified Crystal Cell}
-    \label{fig:simplecell}
-  \end{figure}
-
-This cell is rather complicated for the purposes of this discussion, so 
-consider instead a simplified crystal cell with one atom in its center 
-as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
-minimized for a fixed volume (and temperature and number of particles).  
-A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
-be used for illustration. The number density n for a collection of 
-simplified cells is simply 1 atom per fixed bold-box volume.
-\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
-It is easier to think of 1 atom per ``box" (a reduced number density) 
-than 1 atom per some length$^3$ (the actual number density). Reduced number 
-density is given by $n^*$ and is equal to the number of atoms in a ``box" 
-of size $\sigma^3$.
-\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
-has a volume of $\sigma^3$, the reduced number density is just the number 
-of atoms in the fixed bold-box volume. 
-\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}
-
-In a large collection of simple cells, the atom at the center of each cell 
+  %\begin{figure}[h!]
+    %\centering
+    %% FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
+    %%\caption{A Simplified Crystal Cell}
+    %\label{fig:simplecell}
+  %\end{figure}
+
+%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
+%consider instead a simplified crystal cell with one atom in its center 
+%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
+%minimized for a fixed volume (and temperature and number of particles).  
+%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
+%be used for illustration. The number density n for a collection of 
+%simplified cells is simply 1 atom per fixed bold-box volume.
+%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
+%It is easier to think of 1 atom per ``box" (a reduced number density) 
+%than 1 atom per some length$^3$ (the actual number density). Reduced number 
+%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
+%of size $\sigma^3$.
+%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
+%has a volume of $\sigma^3$, the reduced number density is just the number 
+%of atoms in the fixed bold-box volume. 
+%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
+%}
+In a large collection of cells, the atom at each lattice site of 
 will not realistically reside exactly at the lattice point in the center 
@@
 Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
-distribution is shown in Figure \ref{fig:Gaus_plot}. A 3-dimenstional 
-Gaussian distribution centered at a lattice point can be used to represent 
+distribution is shown in Figure \ref{fig:Gaus_plot}. 
+A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
 a number density profile $n(\vec{r})$  which describes how the atoms are 
 statistically distrubuted over a collection of cells. 
-
+%In 3-dimensions, the Gaussian function takes on the form
+%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
+It is important to note that the number density does not change when introducing a number 
+density profile. There are still the same number of atoms over some fixed volume. 
+%There is still a total of one atom per ``bold box" volume.
+%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
+\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
   \begin{figure}[h!]
@@
 
-It is important to note that the number density does not change when 
-introducing a number density profile. There is still a total of one atom 
-per ``bold box" volume.
-
-\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
-
-In a large collection of cells, not all lattice sites will be occupied. 
-Vacancies can be taken into account while keeping the volume of the 
-bold box, and thus the overall number density, the same. This is done by 
+It may be that not all lattice sites are occupied. Vacancies can be taken 
+into account while keeping %the volume of the bold box, and thus 
+the overall number density the same. This is done by 
 making the cells smaller. A large collection of smaller cells creates a 
@@
 with eight smaller Gaussians of one-eigth the height. 
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{A Gaussian Distribution Number Density Profile}
-    \label{fig:Ensemble_Smallcells}
-  \end{figure}
-
 Instead of thinking of an atom in one of the eight small cells, the same 
@@
 atom in each small cell. This makes it convenient to describe the new 
-reduced number density in terms of the fraction of vacancies, $f_v$:
-
-\begin{displaymath}{ n^* = (1-f_v){n_{no.vacancies}^*}}\end{displaymath}
-
-\begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=2.5cm]{SameStatPic.png}
+number density in terms of the fraction of vacancies, $f_v$:
+
+\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
+%\begin{align}
+  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
+%\end{align}
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
+    \caption{A Gaussian Distribution Number Density Profile}
+    \label{fig:Ensemble_Smallcells}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=2.5cm]{SameStatPic.png}
     \caption{Front views of a cube with eight smaller cells. The two views 
@@
 By varying the width of the Gaussians and the fraction of vacancies, new 
-number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.
+number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
+Once again, it is important to note that the number density is the same 
+for all these number density profiles $n(\vec{r})$. 
+%There is still a total of one atom per ``bold box" volume. 
+%This prodcedure is used only to find the crystal free energy as the number density 
+%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
+A homogeneous number density profile is shown in Figure \ref{fig:homogen_denisty}.
 
@@
     \caption{Several Gaussian Distribution Number Density Profiles all 
-    with a number density of 1 atom per $V_{bold box}$}
+    with the same number density, that is, there is one atom over the fixed volume.}
     \label{fig:Ensemble_vary}
@@
 
-Once again, it is important to note that the number density is the same 
-for all these number density profiles $n(\vec{r})$. There is still a total 
-of one atom per ``bold box" volume. This prodcedure is used only to find 
-the crystal free energy as number density profile for a liquid is 
-homogeneous, as shown in Figure \ref{fig:homogen_denisty}.
-
  \begin{figure}[h!]
@@
     \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
-    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
     \caption{A homogeneouse number density profile $n(\vec{r})$}
@@
 
-A simplified flowchart of the program created to investigate the freezing 
-of a WCA fluid is shown below. The program takes four inputs: reduced number 
-density n*, reduced temperature t*, a range of Gaussian 
-widths gw, and a range of fraction of vacancies fv. The number density 
-profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
-is calculated for each profile until the minimum free energy is found for 
-that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
-The process is then repeated to find the free energies at other number 
-densities and temperatures.
-
-%Program Implementation
-The program computes both the crystal and the homogeneous Helmholtz free 
-energies. The Helmholtz free energy is minimized for a system with a fixed 
-temperature, volume, and number density. The temperature, T is an input to 
-the program. The number density, which is also an input to the program, 
-is used to calculate the volume of one primitive cell over which the free 
-energy is computed. The number of atoms in one primitive cell is one. To 
-compute the free energy, it is necessary to compute both the ideal and 
-excess free energies and add them, $F = F_{ideal} + F_{excess}$.
+%The number density profile is varied by varying gw and fv, and the 
+For a given combination of number density and temperature, a suite of number 
+density profiles are generated.
+% like those represented by Figure~\ref{fig:Ensemble_vary}.
+The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number density profile,
+and also for the homogeneous number denisty profile. 
+Both the ideal and excess free energies are found, and added together to get the total Helmholtz 
+free energy.
+The Helmholtz free energy is minimized for a system with a fixed temperature, 
+volume, and number of particles. Setting the number of particles to one, 
+the volume can be found from the number density.
+The process is then repeated to find the free energies at other combinations 
+of number densities and temperatures.
+
+%The number of atoms in one primitive cell is one.
+
+In general, if at some number density and temperature the crystal free energy 
+is lower than the homogeneous free energy of the liquid then the fluid will 
+be in the crystalline state. If the crystal free energy is higher than 
+the homogeneous free energy, then the fluid will be in the liquid state. 
+However, there is a region of densities between the liquid and crystal 
+states that are forbidden, and so the actual state can not be determined 
+just by which energy is lower at that particular number density and
+temperature. Maxwell's Double Tangent Construction must be implemented 
+using the date collected to determine the actual crystal and liquid states, 
+and construct phase diagrams.
+
 %\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
@@
 
-%\noindent Input:
-\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\noindent Set:
-%\begin{displaymath}\text{N = 1}\end{displaymath}
-\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
-%Outer Loop:  %set fv to one of the input values
-Loop: 
-\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
-\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
-\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
-%\indent\indent\indent Inner Loop : %set gw to one of the input values
-\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
-\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
-
-%Compare:
-\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\noindent Output:
-\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+%PROGRAM FLOWCHART ---------------------------
+%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
+%Loop: 
+%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
+%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
+%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
+%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
+%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
+%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
+%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+
 %\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
@@
 %lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
-\newline
-\newline
-*****   start stuff  ******
-
-In general, if the crystal free energy at some $n^*$ and $t^*$ is lower 
-than the homogeneous free energy of a liquid at that same $n^*$ and $t^*$, 
-then the crystaline state dominates. But if the crystal free energy is 
-higher than the homogeneous free energy, the liquid state dominates. 
-However, there is a region of densities between the liquid and crystal 
-states that are forbidden, and so the actual state can not be determined 
-just by which energy is lower at that particular $n^*$ and $t^*$. Maxwell's 
-Double Tangent Construction must be implemented using the date collected 
-to determine the actual crystal and liquid states, and construct phase 
-diagrams.
-
-*****  end  stuff  ******
-
-\section{Computing the Liquid and Crystal Ideal Helmholtz free energies}
-
-The liquid ideal Helmholtz free energy $F_{Hideal}$ is computed 
-analytically for a given homogeneous reduced density $n^*$, and reduced 
-temperature $t^*$ from the equation 
-%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-\begin{align}
-  \frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
-\end{align}
-where $\sigma=1$ (which gives $n^*=n\sigma^3$=n), and $\epsilon=1$ (which 
-gives $t^*=\frac{k_BT}{\epsilon}=k_BT$). The quantity $\Lambda$ is the DeBroglie 
-wavelength given by
-\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
-where in our program we have set
-\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
-%\begin{displaymath}{m=}\end{displaymath}  
-%\begin{displaymath}{\hbar=}\end{displaymath} 
-%\begin{displaymath}{k_B=}\end{displaymath}
-
-The crystal ideal Helmholtz free energy $F_{Cideal}$ is computed from the equation 
-\begin{align}
-  F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
-\end{align}
-with the varying number density profile given by 
-\begin{displaymath}
-  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}}
-\end{displaymath}
-For small gw values it is assumed that the Guassians do not overlap, 
-in which case the equation for the free energy can be simplified, and solved
-analytically with the result
-\begin{equation}
- F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
-\end{equation}
-For larger gw values it is assumed that the Gaussians do overlap, 
-and so the free energy is caluculated computationally. First dx is scaled by gw,
-\begin{equation}
-  dx_{scaled} = (dx_{input})(gw)
-\end{equation}
-and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
-primitive cell taking into account the contribution of Gaussians from 
-many (5x5x5) nearby cells whose centers are given by $R_i$.
-%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
-The contribution to the free energy of $n(\vec{r})$ at the one position is then
-added to the accumulating sum.
-%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
-%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
-This process is repeated for all positions within one primitive cell to get the
-crystal ideal free energy for the primitive cell. This energy is then divided by
-the volume of the primitive cell to get the crystal ideal free energy per
-volume.
 
@@
 The functional is written in general form as 
-\begin{displaymath}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
+\begin{align}
+  F_{excess}[n(\vec{r})]=k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}
+\end{align}
 where
 \begin{align}
-\Phi_1 &= -n_{0}ln(1-n_{3}) \\
+\Phi_1 &= -n_{0}\ln(1-n_{3}) \\
 \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
@@
 \begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-where the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
+where the number density profile $n(\vec r)$ within the integral is given by a Gaussian function,
+%\begin{align}
+  %n(\vec r)=(1-f_v)\frac{\exp^{-\frac{r^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}
+%\end{align}
+and the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
 \begin{align}\label{eq:weights}
@@
 
-\section{Computing the Liquid and Crystal Excess Helmholtz free energies}
-
-The liquid excess free energy $F_{Hexcess}$ is computed analytically for 
-a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
+\section{Computing the Liquid and Crystal Helmholtz free energies}
+
+The liquid ideal Helmholtz free energy is computed analytically 
+from Equation~\ref{} 
+for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
+%from the equation 
+%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
+%\begin{align}
+  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
+%\end{align}
+%The quantity $\Lambda$ is the DeBroglie wavelength given by
+%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
+%where in our program we have set
+%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
+%\begin{displaymath}{m=}\end{displaymath}  
+%\begin{displaymath}{\hbar=}\end{displaymath} 
+%\begin{displaymath}{k_B=}\end{displaymath} 
+The crystal ideal Helmholtz free energy is computed computationally 
+from Equation~\ref{} with the varying number density profile given by
+%\begin{align}
+  %F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
+%\end{align} 
+\begin{align}
+  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
+\end{align}
+%For small gw values it is assumed that the Guassians do not overlap, 
+%in which case the equation for the free energy can be simplified, and solved
+%analytically with the result
+%\begin{equation}
+ %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
+%\end{equation}
+%For larger gw values it is assumed that the Gaussians do overlap, 
+%and so the free energy is caluculated computationally. First dx is scaled by gw,
+%\begin{equation}
+  %dx_{scaled} = (dx_{input})(gw)
+%\end{equation}
+%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
+%primitive cell taking into account the contribution of Gaussians from 
+%many (5x5x5) nearby cells whose centers are given by $R_i$.
+%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
+%The contribution to the free energy of $n(\vec{r})$ at the one position is then
+%added to the accumulating sum.
+%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
+%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
+%This process is repeated for all positions within one primitive cell to get the
+%crystal ideal free energy for the primitive cell. This energy is then divided by
+%the volume of the primitive cell to get the crystal ideal free energy per
+%volume.
+
+The liquid excess Helmholtz free energy is computed analytically  
+%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
 using SFMT with a uniform number density profile, n that does not depend 
@@
 %done in hf.energy() note- does not depend on fv or gw but does use SFMT):
-\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
-For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
+%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
+%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
 %\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
-\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
-\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
-The weighted densities are found from
-\begin{displaymath}{n_i=nw_i}\end{displaymath} 
-where the weights $w_i$ are given by Equations~\ref{eq:weights}.
-
-The crystal excess free energy $F_{Cexcess}$ is computed computationally 
+%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
+%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
+%The weighted densities are found from
+%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
+%where the weights $w_i$ are given by Equations~\ref{eq:weights}.
+
+The crystal excess Helmholtz free energy is computed computationally 
 using SFMT with a number density profile $n(\vec r)$ that varies spacially.
@@
 %Then the crystal, or inhomogeneous, excess free energy is found from
-\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
+%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
 Monte-Carlo integration is used to compute the weighted densities 

### 2020-10-29 papers/thesis-kirstie/thesis.tex
edits

@@
 
-Classically, fluids have been modeled by a collection of small, 
-impenetrable spheres colliding with one another like billiard balls. 
-In the simplest case, the spheres are treated as having only a replulsive 
-potential. This model, called the Hard-Sphere fluid, has served as the 
-primary reference system for fluids. This is due to the fact that the 
-repulsion between atoms is much stronger than their attraction, and so 
-the repulsive potential, which is descriptive of hard spheres, dominates 
-the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
-fluid which is a perturbation of the Hard-Sphere fluid created by adding 
-a hard-edge attractive potential to the repulsive potential of the 
-Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-spheres with softly fading replusive potentials which he modeled by making
-the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
-This reflects the idea that soft-spheres penetrate more deeply into 
-each other when they collide at higher temperatures.
-
-In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
-with Rowlinson's idea of using temperature-dependent radii to model spheres 
-with soft repulsive potentials to create the Barker-Henderson model 
-fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
-modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
-at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
-Chandler, and Anderson used temperature and density dependent hard-sphere 
-radii to model spheres with the more realistic softly fading repulsive 
-portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-which is known as the Weeks-Chandler-Anderson model fluid.
-
-While models for particular fluids were being created, powerful methods 
-to describe the properties of fluids in general were also being developed. 
-In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
-introduced from which an equation of state, and an expression for the 
-excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
-obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
-alternative method...OZ... \color{black}
-%Percus and Yevick took another approach to the Hard-Sphere fluid.
-In 1961 Morita and Hiroike introduced a density functional 
-approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
-(the equilibrium Helmholtz free energy in the case of a canonical 
-ensemble), and the cooresponding equilibrium number density profile for 
-inhomogeneous fluids. To use this approach, a functional must be 
-constructed for a particular fluid.
-%DFT Evans 1979
-%Led by the consistency bewteen results from SPT and Percus-Yevick, 
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
-functional for the Hard-Sphere fluid. Tarazona later improved on the 
-method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
-santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
-forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
-
+The goal of this project is to generate temperature-density and 
+pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
+model fluid using a computer program. 
+To do this we need to find the Helmholtz free energies that the fluid
+would have in a liquid state, and in a crystalline solid state 
+at a given temperature and density. 
+%Once the Helmholtz free energy for both the crystal and liquid states of 
+%the fluid are found for a given temperature and density, they can be 
+%compared to find the resulting state of the fluid. 
+Outside the region of 
+coexistence which exists between the liquid and crystal states, 
+the fluid will be completely in a crystalline state if the 
+crystal free energy is lower than the liquid free energy, and it will be 
+completely in a liquid state if the liquid free energy is lower than 
+the crystal free energy. Maxwell's Double Tangent Construction is used 
+to determine the actual number densities of the crystal and liquid states
+at the boundaries of the coexistance region for a given temperature from 
+which the phase diagrams can be constructed.
+
+%In general, if at some number density and temperature the crystal free energy 
+%is lower than the homogeneous free energy of the liquid, then the fluid will 
+%be in the crystalline state. If the homogeneous liquid free energy is lower 
+%than the crystal free energy, then the fluid will be in the liquid state.
+
+%To find the equilibrium Helmholtz free energies we use classical Density 
+%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
+%%the corresponding equilibrium number density profile. 
+%In particular,
+
+In general, the Helmholtz free energy can be written as a functional of 
+the number density profile $n(\vec{r})$ which describes the way the average number 
+density varies spatially. 
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
+Classical Density Functional Theory (cDFT) is used to find the equilbrium 
+Helmholtz free energy of the system 
+by varying the number density profile $n(\vec{r})$ until the free energy 
+is minimized. The minimum free energy, and its corresponding number 
+density profile are then the equilibrium values of the Helmholtz free energy
+of the system at the given temperature and number density.
+
+To find the Helmholtz free energy both the ideal Helmholtz free energy
+and the excess Helmholtz free energy must be found and added together.
+The ideal Helmholtz free energy is the free energy the fluid would have in
+the absence of any interactomic potential energy $V(r)$ between the spheres, 
+or atoms that make up the fluid. The excess Helmholtz free energy is that 
+additional energy which results from the presence of an interatomic potential 
+energy $V(r)$ between the spheres.
+%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+%For a spatially varying number density, as would be expected for a crystal, 
+%the free energy is a functional of the number density profile $n(\vec{r})$  
+%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
+%analytically to find the resulting free energy for the system. 
+Soft Fundamental Measure Theory (SFMT) is a method used to form an expression 
+for the excess Helmholtz free energy of a particular model fluid without 
+having to take explicitly into account the interaction each sphere, or atom, 
+has with every other sphere. 
+%Classically, fluids have been modeled by a collection of small, 
+%impenetrable spheres colliding with one another like billiard balls. 
+%In the simplest case, the spheres are treated as having only a replulsive 
+%potential. This model, called the Hard-Sphere fluid, has served as the 
+%primary reference system for fluids. This is due to the fact that the 
+%repulsion between atoms is much stronger than their attraction, and so 
+%the repulsive potential, which is descriptive of hard spheres, dominates 
+%the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
+%fluid which is a perturbation of the Hard-Sphere fluid created by adding 
+%a hard-edge attractive potential to the repulsive potential of the 
+%Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
+%spheres with softly fading replusive potentials which he modeled by making
+%the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+%This reflects the idea that soft-spheres penetrate more deeply into 
+%each other when they collide at higher temperatures. 
+%
+%In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
+%with Rowlinson's idea of using temperature-dependent radii to model spheres 
+%with soft repulsive potentials to create the Barker-Henderson model 
+%fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
+%modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
+%at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
+%Chandler, and Anderson used temperature and density dependent hard-sphere 
+%radii to model spheres with the more realistic softly fading repulsive 
+%portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
+%which is known as the Weeks-Chandler-Anderson model fluid. 
+%
+%While models for particular fluids were being created, powerful methods 
+%to describe the properties of fluids in general were also being developed. 
+%In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
+%introduced from which an equation of state, and an expression for the 
+%excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
+%obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
+%alternative method...OZ... \color{black}
+%%Percus and Yevick took another approach to the Hard-Sphere fluid.
+%In 1961 Morita and Hiroike introduced a density functional 
+%approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
+%(the equilibrium Helmholtz free energy in the case of a canonical 
+%ensemble), and the cooresponding equilibrium number density profile for 
+%inhomogeneous fluids. To use this approach, a functional must be 
+%constructed for a particular fluid.
+%%DFT Evans 1979
+%%Led by the consistency bewteen results from SPT and Percus-Yevick, 
+%In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+%(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
+%functional for the Hard-Sphere fluid. Tarazona later improved on the 
+%method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
+%santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
+%forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
 Schmidt applied his Soft Fundamental Measure Theory to various fluids 
@@
 model for spheres with steep, short-range soft repulsive 
-potentials~\cite{schmidt2000fluid}. In this paper, Density Functional 
-Theory (DFT) will be used to predict the freezing of a Weeks-Chandler-Anderson 
+potentials~\cite{schmidt2000fluid}.
+% like that exhibited by a WCA fluid.
+
+In this paper, classical Density Functional 
+Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
 model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
@@
 \section{Overview}
-We use classical Density Functional Theory (cDFT) in conjunction 
-with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
-energy of a system, and the corresponding equilibrium number density profile. 
-In particular, SFMT is used to form the expression for the excess Helmholtz 
-free energy without having to take into account the interaction each sphere, 
-or atom, has with every other sphere. The excess free energy is then added 
-to the ideal free energy calculated analytically to find the resulting 
-free energy for the system. 
-\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-For a spatially varying number density, as would be expected for a crystal, 
-the free energy is a functional of the number density profile $n(\vec{r})$  
+The goal of this project is to generate phase diagrams for a 
+Weeks-Chandler-Anderson model fluid using a computer program. 
+To do this we find the Helmholtz free energies that the fluid
+would have at a given temperature and density in a  
+liquid state, and in a crystalline solid state. 
+%Once the Helmholtz free energy for both the crystal and liquid states of 
+%the fluid are found for a given temperature and density, they can be 
+%compared to find the resulting state of the fluid. 
+Outside the region of 
+coexistence which exists between the liquid and crystal states, 
+the fluid will be completely in a crystalline state if the 
+crystal free energy is lower than the liquid free energy, and it will be 
+completely in a liquid state if the liquid free energy is lower than 
+the crystal free energy. Maxwell's Double Tangent Construction is used 
+to determine the actual number densities of the crystal and liquid states
+at the boundaries of the coexistance region for a given temperature, 
+and construct phase diagrams.
+
+%In general, if at some number density and temperature the crystal free energy 
+%is lower than the homogeneous free energy of the liquid, then the fluid will 
+%be in the crystalline state. If the homogeneous liquid free energy is lower 
+%than the crystal free energy, then the fluid will be in the liquid state.
+
+%To find the equilibrium Helmholtz free energies we use classical Density 
+%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
+%%the corresponding equilibrium number density profile. 
+%In particular,
+
+In general, the Helmholtz free energy can be written as a functional of 
+the number density profile $n(\vec{r})$ which describes the way the average number 
+density varies spatially. 
 \begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-
-To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
-the temperature fixed until the free energy is minimized. The minimum free 
-energy and its corresponding number density profile are then the equilibrium 
-values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
-the fluid are found, they can be compared to find the resulting state of 
-the fluid. Outside the region of coexistence, the fluid will be completely 
-in a crystalline state if the crystal free energy is lower than the liquid 
-free energy, and it will be completely in a liquid state if the liquid 
-free energy is lower than the crystal free energy.
+Classical Density Functional Theory (cDFT) is used to find the equilbrium 
+Helmholtz free energy of the system at a given temperature and density
+by varying the number density profile $n(\vec{r})$ until the free energy 
+is minimized. The minimum free energy, and its corresponding number 
+density profile are then the equilibrium values for that temperature 
+and number density.
+
+The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
+and the excess Helmholtz free energy.
+The ideal Helmholtz free energy is the free energy the fluid would have in
+the absence of any interactomic potential energy $V(r)$ between the spheres, 
+or atoms that make up the fluid. The excess Helmholtz free energy is that 
+additional energy which results from the presence of an interatmoic potential 
+energy $V(r)$ between the spheres.
+%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+%For a spatially varying number density, as would be expected for a crystal, 
+%the free energy is a functional of the number density profile $n(\vec{r})$  
+%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
+%analytically to find the resulting free energy for the system. 
+Soft Fundamental Measure Theory (SFMT) is used to form the expression 
+for the excess Helmholtz free energy without having to take explicitly 
+into account the interaction each sphere, or atom, has with every other 
+sphere.
+
+ ******
 
@@
 
-In general, if at some number density and temperature the crystal free energy 
-is lower than the homogeneous free energy of the liquid then the fluid will 
-be in the crystalline state. If the crystal free energy is higher than 
-the homogeneous free energy, then the fluid will be in the liquid state. 
-However, there is a region of densities between the liquid and crystal 
-states that are forbidden, and so the actual state can not be determined 
-just by which energy is lower at that particular number density and
-temperature. Maxwell's Double Tangent Construction must be implemented 
-using the date collected to determine the actual crystal and liquid states, 
-and construct phase diagrams.
-
 %\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
@@
 \section{Our Functional for the Excess Helmholtz free energy}
-
+In order to apply classical Density Functional Theory, the Helmholtz energy
+must be found. Computing the ideal portion of the Helmholtz free energy is
+straightforward, but computing the excess portion of the Helmholtz free energy 
+is difficult since it depends on the interatomic potential energy $V(r)$ between 
+each sphere with all the other spheres. Soft Fundamental Measure Theory is a method
+developed to form an expression for the excess Helmholtz free energy without
+explicitly taking into account the interaction of each sphere with all the others, 
+but it must be taylored to fit a particular fluid. 
 %The excess Helmholtz free energy is particular to the fluid being modeled, 
 %as it depends on the interatomic potential V(r).
-We use Soft Fundamental Measure Theory to find the excess Helmholtz free energy.
-The functional is written in general form as 
+
+We use Soft Fundamental Measure Theory to form a functional for the excess 
+Helmholtz free energy appropriate for the WCA fluid by introducing our 
+temperature dependent parameters $\alpha(T)$ and $\Xi(T)$. The functional 
+is written in general form as 
 \begin{align}
@@
 The result can be made more accurate by increasing NumPoints which is the 
-number of vectors randomly generated. In order to further reduce the error 
-in the average, antithetic variates are used. Rather than computing an average 
-of a set of random points, each random point in the set is paired with a 
-point along the same line as the first point, but opposite in direction, 
-as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
-error in the average by filtering out an erroneous offset of values along 
-the line connecting the two points by averaging the two results, one of 
-which may be erroneously higher than it should be while the other is 
-erronesously lower than it should be.
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=4cm]{AntitheticVariate.png}
-    \caption{Antithetic variates reduce first order error in an average. 
-    Each point in the average is paired with another point from along 
-    the same line as the first, and at an equal, but opposite distance.}
-    \label{fig:AntitheticVariate}
-  \end{figure} 
-Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
-\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-Then the entire process repeated for each vector $\vec{r}_k$ to get the 
-total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
-The standard error of the mean (SEM) is given by:
-\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
-a measure of the error in $n_3=\overline{w}_3$:
-\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
-NumPoints, the number of vectors randomly generated, is increased until:
-\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
-value that $n_3$ can be without error is 1.
-
-Figure \ref{fig:GaussandW2_actual} ......
-
- \begin{figure}[h!]
-    \centering
-    % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
+number of vectors randomly generated. 
+%-------Antithetic-------------------
+%In order to further reduce the error 
+%in the average, antithetic variates are used. Rather than computing an average 
+%of a set of random points, each random point in the set is paired with a 
+%point along the same line as the first point, but opposite in direction, 
+%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
+%error in the average by filtering out an erroneous offset of values along 
+%the line connecting the two points by averaging the two results, one of 
+%which may be erroneously higher than it should be while the other is 
+%erronesously lower than it should be.
+
+ %\begin{figure}[h!]
+    %\centering
+    %\includegraphics[height=4cm]{AntitheticVariate.png}
+    %\caption{Antithetic variates reduce first order error in an average. 
+    %Each point in the average is paired with another point from along 
+    %the same line as the first, and at an equal, but opposite distance.}
+    %\label{fig:AntitheticVariate}
+  %\end{figure} 
+%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
+%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
+%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
+%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
+%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
+
+%The standard error of the mean (SEM) is given by:
+%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
+%a measure of the error in $n_3=\overline{w}_3$:
+%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
+%NumPoints, the number of vectors randomly generated, is increased until:
+%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
+%value that $n_3$ can be without error is 1.
+
+%Figure \ref{fig:GaussandW2_actual} ......
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{GaussandW2_actual.png}
     \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
@@
     \centering
-    % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
-    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
-    and fixed in place as $\vec{dr}$ varies randomly.} 
+    %\includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    %\caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    %the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
+    %and fixed in place as $\vec{dr}$ varies randomly.} 
   \label{fig:GaussandW2_thinkas}
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
+  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
   %\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
-  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
-  White lines show the liquid and solid transition densities at each temperature from 
-  simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
+  %\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  %White lines show the liquid and solid transition densities at each temperature from 
+  %simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
   \label{fig:Phase_Diagram_of_T_vs_n_AS}
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
+  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
   %\includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T_AS}
-  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
-  The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
+  %\caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
+  %The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
   \label{fig:Phase_Diagram_P_vs_T_AS}
@@
 
-Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
-reduced pressure versus the reduced temperature at fixed reduced densities 
-from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
-\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
-the WCA fluid transistions into a solid. These downward bends align to 
-form the curved interface separating the blue and red areas in 
-Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
+%Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
+%reduced pressure versus the reduced temperature at fixed reduced densities 
+%from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
+%\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
+%the WCA fluid transistions into a solid. These downward bends align to 
+%form the curved interface separating the blue and red areas in 
+%Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
 
@@
   \centering
-  \includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
-  \caption{Plot of the reduced pressure verses the reduced temperature. 
-  The lines bending downward indicate freezing.}
+  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
+  %\caption{Plot of the reduced pressure verses the reduced temperature. 
+  %The lines bending downward indicate freezing.}
   \label{fig:p-vs-T_at_fixed_density}

### 2020-10-29 papers/thesis-kirstie/thesis.tex
edit-intro

@@
 %compared to find the resulting state of the fluid. 
-Outside the region of 
-coexistence which exists between the liquid and crystal states, 
-the fluid will be completely in a crystalline state if the 
-crystal free energy is lower than the liquid free energy, and it will be 
-completely in a liquid state if the liquid free energy is lower than 
-the crystal free energy. Maxwell's Double Tangent Construction is used 
-to determine the actual number densities of the crystal and liquid states
-at the boundaries of the coexistance region for a given temperature from 
-which the phase diagrams can be constructed.
-
+%
+%save:
+%Outside the region of 
+%coexistence which exists between the liquid and crystal states, and which 
+%consists of various amounts of liquid and crystal existing together,
+%the fluid will be completely in a crystalline state if the 
+%crystal free energy is lower than the liquid free energy, and it will be 
+%completely in a liquid state if the liquid free energy is lower than 
+%the crystal free energy. Maxwell's Double Tangent Construction is used 
+%to determine the actual number densities of the crystal and liquid states
+%at the boundaries of the coexistance region for a given temperature from 
+%which the phase diagrams can be constructed.
+%end save
+%
 %In general, if at some number density and temperature the crystal free energy 
@@
 %than the crystal free energy, then the fluid will be in the liquid state. 
-
+%
 %To find the equilibrium Helmholtz free energies we use classical Density 
@@
 %In particular, 
-
-In general, the Helmholtz free energy can be written as a functional of 
+In general, the Helmholtz free energy, 
+which consists of an ideal part and an excess part ,
+can be written as a functional of 
 the number density profile $n(\vec{r})$ which describes the way the average number 
@@
 \begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-Classical Density Functional Theory (cDFT) is used to find the equilbrium 
+%
+
+In this paper, classical Density Functional Theory (cDFT)
+will be used to find the equilibrium Helmholtz free energy, and 
+Soft Fundamental Measure Theory (SFMT) will be used to form the functional for the excess Helmholtz 
+free energy. 
+%to predict the freezing of a Weeks-Chandler-Anderson 
+%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
+%dependent parameters to form the functional for the excess Helmholtz free energy. 
+%for generating the weight funcions, and Tarazona's tensor functional. 
+%
+%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
+According to classical Density Functional Theory (cDFT) the equilbrium 
 Helmholtz free energy of the system 
+can be found
 by varying the number density profile $n(\vec{r})$ until the free energy 
-is minimized. The minimum free energy, and its corresponding number 
-density profile are then the equilibrium values of the Helmholtz free energy
-of the system at the given temperature and number density.
-
-To find the Helmholtz free energy both the ideal Helmholtz free energy
-and the excess Helmholtz free energy must be found and added together.
-The ideal Helmholtz free energy is the free energy the fluid would have in
-the absence of any interactomic potential energy $V(r)$ between the spheres, 
-or atoms that make up the fluid. The excess Helmholtz free energy is that 
-additional energy which results from the presence of an interatomic potential 
-energy $V(r)$ between the spheres.
+is minimized. 
+%The minimum free energy, and its corresponding number 
+%density profile are then the equilibrium values of the Helmholtz free energy
+%of the system at the given temperature and number density. 
+%
+%To find the Helmholtz free energy both the ideal Helmholtz free energy
+%and the excess Helmholtz free energy must be found and added together.
+%The ideal Helmholtz free energy is the free energy the fluid would have in
+%the absence of any interactomic potential energy $V(r)$ between the spheres, 
+%or atoms that make up the fluid. 
+%
+%The excess Helmholtz free energy is that 
+%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
+%energy $V(r)$ between the spheres.
+%
 %\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
@@
 %analytically to find the resulting free energy for the system. 
-Soft Fundamental Measure Theory (SFMT) is a method used to form an expression 
-for the excess Helmholtz free energy of a particular model fluid without 
-having to take explicitly into account the interaction each sphere, or atom, 
-has with every other sphere. 
+%
+Soft Fundamental Measure Theory (SFMT) makes the task of forming an expression
+%is a method used to form an expression 
+for the excess Helmholtz free energy of a particular model fluid 
+much easier since the interaction that each sphere, or atom, 
+has with every other sphere does not have to be taken explicitly into account.
+%without 
+%having to take explicitly into account the interaction each sphere, or atom, 
+%has with every other sphere which would make the task very difficult, 
+%or even prohibtive.
+%
 %Classically, fluids have been modeled by a collection of small, 
@@
 %portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-%which is known as the Weeks-Chandler-Anderson model fluid. 
+%which is known as the Weeks-Chandler-Anderson model fluid. f
 %
@@
 %forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
-Schmidt applied his Soft Fundamental Measure Theory to various fluids 
-with soft repulsive and soft attractive potentials, and developed an erf 
-model for spheres with steep, short-range soft repulsive 
-potentials~\cite{schmidt2000fluid}.
+%
+%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
+%with soft repulsive and soft attractive potentials, and developed an erf 
+%model for spheres with steep, short-range soft repulsive 
+%potentials~\cite{schmidt2000fluid}.
 % like that exhibited by a WCA fluid. 
-
-In this paper, classical Density Functional 
-Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
-model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
-dependent parameters to form the functional for the excess Helmholtz free energy. 
-%for generating the weight funcions, and Tarazona's tensor functional. 
-Results will be compared to those generated from Monte-Carlo simulations.
+%
+Results will be compared to those from Monte-Carlo simulations.
 

### 2020-10-30 papers/thesis-kirstie/thesis.tex
edits

@@
 
-The goal of this project is to generate temperature-density and 
-pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
-model fluid using a computer program. 
+The goal of this project is to generate computationally temperature-density 
+and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
+model fluid using classical Density Functional Theory (cDFT). 
 To do this we need to find the Helmholtz free energies that the fluid
@@
 
-In this paper, classical Density Functional Theory (cDFT)
-will be used to find the equilibrium Helmholtz free energy, and 
-Soft Fundamental Measure Theory (SFMT) will be used to form the functional for the excess Helmholtz 
-free energy. 
+Classical Density Functional Theory is used to find the equilibrium 
+Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
+be used to form the functional for the excess Helmholtz free energy. 
 %to predict the freezing of a Weeks-Chandler-Anderson 
@@
 %Classical Density Functional Theory (cDFT) is used to find the equilbrium 
-According to classical Density Functional Theory (cDFT) the equilbrium 
+According to classical Density Functional Theory the equilbrium 
 Helmholtz free energy of the system 
@@
 %
-Soft Fundamental Measure Theory (SFMT) makes the task of forming an expression
+SFMT makes the task of forming an expression
 %is a method used to form an expression 
@@
 changes in the free energy will be negative for spontaneous processes 
-when the natural variables of the free energy are held fixed. 
+when the natural variables of the free energy are held fixed~\cite{schroeder}. 
 %And so the free energy tends toward a minimum. 
@@
 %particles respond to a small offset from equilibrium? 
-
-Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-\begin{equation}
-  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
-\end{equation}
-and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
-$dT=0$ gives
-\begin{equation}
-  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
-\end{equation}
-where conservation of energy was also applied. Fixing the volume and the 
-number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
-results in
-\begin{equation}
-  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
-\end{equation}
-which, for fixed N, is the same as
-\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
-Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
-of particles are fixed, the change in the free energy of the system will 
-always be negative, and thus the free energy tends toward a 
-minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
-for a system near equilibrium. A general proof of minimization 
+%
+%Save for me:
+%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
+%\begin{equation}
+  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
+%\end{equation}
+%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
+%$dT=0$ gives
+%\begin{equation}
+  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
+%\end{equation}
+%where conservation of energy was also applied. Fixing the volume and the 
+%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
+%results in
+%\begin{equation}
+  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+%\end{equation}
+%which, for fixed N, is the same as
+%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
+%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
+%of particles are fixed, the change in the free energy of the system will 
+%always be negative, and thus the free energy tends toward a 
+%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
+%for a system near equilibrium. 
+%End save for me.
+A general proof of minimization 
 can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
@@
 of a delta function characteristic of hard spheres. Soft Fundamental Measure 
-Theory (SFMT) extends the Fundamental Measure Theory discussed so far to 
+Theory extends the Fundamental Measure Theory discussed so far to 
 fluids with soft potenials by keeping the relationships between the weight 
@@
 \section{Overview}
-The goal of this project is to generate phase diagrams for a 
-Weeks-Chandler-Anderson model fluid using a computer program. 
-To do this we find the Helmholtz free energies that the fluid
-would have at a given temperature and density in a  
-liquid state, and in a crystalline solid state. 
-%Once the Helmholtz free energy for both the crystal and liquid states of 
-%the fluid are found for a given temperature and density, they can be 
-%compared to find the resulting state of the fluid. 
-Outside the region of 
-coexistence which exists between the liquid and crystal states, 
-the fluid will be completely in a crystalline state if the 
-crystal free energy is lower than the liquid free energy, and it will be 
-completely in a liquid state if the liquid free energy is lower than 
-the crystal free energy. Maxwell's Double Tangent Construction is used 
-to determine the actual number densities of the crystal and liquid states
-at the boundaries of the coexistance region for a given temperature, 
-and construct phase diagrams.
-
-%In general, if at some number density and temperature the crystal free energy 
-%is lower than the homogeneous free energy of the liquid, then the fluid will 
-%be in the crystalline state. If the homogeneous liquid free energy is lower 
-%than the crystal free energy, then the fluid will be in the liquid state.
-
-%To find the equilibrium Helmholtz free energies we use classical Density 
-%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
-%%the corresponding equilibrium number density profile. 
-%In particular,
-
-In general, the Helmholtz free energy can be written as a functional of 
-the number density profile $n(\vec{r})$ which describes the way the average number 
-density varies spatially. 
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-Classical Density Functional Theory (cDFT) is used to find the equilbrium 
-Helmholtz free energy of the system at a given temperature and density
-by varying the number density profile $n(\vec{r})$ until the free energy 
-is minimized. The minimum free energy, and its corresponding number 
-density profile are then the equilibrium values for that temperature 
-and number density.
-
-The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
-and the excess Helmholtz free energy.
-The ideal Helmholtz free energy is the free energy the fluid would have in
-the absence of any interactomic potential energy $V(r)$ between the spheres, 
-or atoms that make up the fluid. The excess Helmholtz free energy is that 
-additional energy which results from the presence of an interatmoic potential 
-energy $V(r)$ between the spheres.
-%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%For a spatially varying number density, as would be expected for a crystal, 
-%the free energy is a functional of the number density profile $n(\vec{r})$  
-%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
-%analytically to find the resulting free energy for the system. 
-Soft Fundamental Measure Theory (SFMT) is used to form the expression 
-for the excess Helmholtz free energy without having to take explicitly 
-into account the interaction each sphere, or atom, has with every other 
-sphere.
-
- ******
-
-Although the entire crystal lattice can be constructed with FCC cells, 
+%
+%The goal of this project is to generate phase diagrams for a 
+%Weeks-Chandler-Anderson model fluid using a computer program. 
+%To do this we find the Helmholtz free energies that the fluid
+%would have at a given temperature and density in a  
+%liquid state, and in a crystalline solid state. 
+%%Once the Helmholtz free energy for both the crystal and liquid states of 
+%%the fluid are found for a given temperature and density, they can be 
+%%compared to find the resulting state of the fluid. 
+%Outside the region of 
+%coexistence which exists between the liquid and crystal states, 
+%the fluid will be completely in a crystalline state if the 
+%crystal free energy is lower than the liquid free energy, and it will be 
+%completely in a liquid state if the liquid free energy is lower than 
+%the crystal free energy. Maxwell's Double Tangent Construction is used 
+%to determine the actual number densities of the crystal and liquid states
+%at the boundaries of the coexistance region for a given temperature, 
+%and construct phase diagrams.
+
+%%In general, if at some number density and temperature the crystal free energy 
+%%is lower than the homogeneous free energy of the liquid, then the fluid will 
+%%be in the crystalline state. If the homogeneous liquid free energy is lower 
+%%than the crystal free energy, then the fluid will be in the liquid state. 
+%
+%%To find the equilibrium Helmholtz free energies we use classical Density 
+%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
+%%%the corresponding equilibrium number density profile. 
+%%In particular, 
+%
+%In general, the Helmholtz free energy can be written as a functional of 
+%the number density profile $n(\vec{r})$ which describes the way the average number 
+%density varies spatially. 
+%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
+%Classical Density Functional Theory is used to find the equilbrium 
+%Helmholtz free energy of the system at a given temperature and density
+%by varying the number density profile $n(\vec{r})$ until the free energy 
+%is minimized. The minimum free energy, and its corresponding number 
+%density profile are then the equilibrium values for that temperature 
+%and number density. 
+%
+%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
+%and the excess Helmholtz free energy.
+%The ideal Helmholtz free energy is the free energy the fluid would have in
+%the absence of any interactomic potential energy $V(r)$ between the spheres, 
+%or atoms that make up the fluid. The excess Helmholtz free energy is that 
+%additional energy which results from the presence of an interatmoic potential 
+%energy $V(r)$ between the spheres.
+%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+%%For a spatially varying number density, as would be expected for a crystal, 
+%%the free energy is a functional of the number density profile $n(\vec{r})$  
+%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
+%%analytically to find the resulting free energy for the system. 
+%Soft Fundamental Measure Theory is used to form the expression 
+%for the excess Helmholtz free energy without having to take explicitly 
+%into account the interaction each sphere, or atom, has with every other 
+%sphere.
+
+Our goal is to generate computationally temperature-density and pressure-
+temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
+classical Density Functional Theory. 
+A temperature-density phase diagram for a fluid that does not
+have attraction between the particles that make up the fluid will show
+two discinct regions - one at lower densities, and one at higher densities,
+seperated by a coexistence region in which there is found a mixture of the 
+fluid at some fixed lower density, and at some fixed higher density. An 
+abrupt change from some maximum lower density to some minimum higher density 
+occurs for certian combinations of pressure and temperature which form
+a line on a pressure-temperature phase diagram. This line distinguishes 
+the region in which the fluid exhibits lower densities from the region 
+in which the fluid exhbits higher densities. These two regions are known
+as the liquid state and the solid state of the fluid.
+
+It is known that a fluid with a soft potential like the WCA fluid will 
+form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
+The transition between the liquid 
+and solid states is abruptly inhomogeneous, and we study the behavior 
+of the fluid by treating the liquid as homogeneous, and by constructing 
+a crystal lattice in real space for the solid. Although the entire 
+crystal lattice can be constructed with FCC cells, 
 FCC cells are not the smallest cells with which the lattice can be built. 
@@
 number density in terms of the fraction of vacancies, $f_v$: 
-
 \begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
@@
 
-By varying the width of the Gaussians and the fraction of vacancies, new 
+By varying the width of the Gaussians and the fraction of vacancies, various
 number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
@@
 %profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
-A homogeneous number density profile is shown in Figure \ref{fig:homogen_denisty}.
+A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.
 
@@
 
+\section{Computing the Liquid and Crystal Helmholtz free energies}
+%-----------Comuting the Liquid and Crystal free energies ------
+
+The liquid ideal Helmholtz free energy is computed analytically 
+using Equation~\ref{eq:Fideal} 
+for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
+%from the equation 
+%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
+%\begin{align}
+  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
+%\end{align}
+%The quantity $\Lambda$ is the DeBroglie wavelength given by
+%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
+%where in our program we have set
+%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
+%\begin{displaymath}{m=}\end{displaymath}  
+%\begin{displaymath}{\hbar=}\end{displaymath} 
+%\begin{displaymath}{k_B=}\end{displaymath} 
+The crystal ideal Helmholtz free energy is computed computationally 
+from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
+%\begin{align}
+  %F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
+%\end{align} 
+\begin{align}
+  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
+\end{align}
+%For small gw values it is assumed that the Guassians do not overlap, 
+%in which case the equation for the free energy can be simplified, and solved
+%analytically with the result
+%\begin{equation}
+ %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
+%\end{equation}
+%For larger gw values it is assumed that the Gaussians do overlap, 
+%and so the free energy is caluculated computationally. First dx is scaled by gw,
+%\begin{equation}
+  %dx_{scaled} = (dx_{input})(gw)
+%\end{equation}
+%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
+%primitive cell taking into account the contribution of Gaussians from 
+%many (5x5x5) nearby cells whose centers are given by $R_i$.
+%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
+%The contribution to the free energy of $n(\vec{r})$ at the one position is then
+%added to the accumulating sum.
+%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
+%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
+%This process is repeated for all positions within one primitive cell to get the
+%crystal ideal free energy for the primitive cell. This energy is then divided by
+%the volume of the primitive cell to get the crystal ideal free energy per
+%volume.
+%
+The liquid excess Helmholtz free energy is computed analytically  
+%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
+using SFMT with a uniform number density profile, n that does not depend 
+on $\vec r$. 
+%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
+%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
+%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
+%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
+%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
+%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
+%The weighted densities are found from
+%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
+%where the weights $w_i$ are given by Equations~\ref{eq:weights}.
+
+The crystal excess Helmholtz free energy is computed computationally 
+using SFMT with a number density profile $n(\vec r)$ that varies spacially.
+%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
+%Then the crystal, or inhomogeneous, excess free energy is found from
+%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
+Monte-Carlo integration is used to compute the weighted densities 
+$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
+%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
+%
+The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
+within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
+the Gaussians is selected that is within the primitive cell, or one of the 
+neighboring cells within a prudently picked inclusion radius as Gaussians 
+farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=5cm]{InclusionRadius.png}
+    \caption{The Gaussians at each lattice point are of width gw. 
+    After a radius of few gw's the Gaussians go to zero and can be neglected.}
+    \label{fig:InclusionRadius}
+  \end{figure}
+
+The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
+$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
+$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
+$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
+where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
+some point in space. Monte-Carlo integration works by randomly generating 
+$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
+\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
+is the gaussian number density profile for no vacancies, the equation:
+
+\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+implemented computationally in the form:
+\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+becomes:
+\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
+where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
+(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
+which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
+$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
+Using the relation:
+\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
+$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
+\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
+The result can be made more accurate by increasing NumPoints which is the 
+number of vectors randomly generated. 
+%-------Antithetic-------------------
+%In order to further reduce the error 
+%in the average, antithetic variates are used. Rather than computing an average 
+%of a set of random points, each random point in the set is paired with a 
+%point along the same line as the first point, but opposite in direction, 
+%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
+%error in the average by filtering out an erroneous offset of values along 
+%the line connecting the two points by averaging the two results, one of 
+%which may be erroneously higher than it should be while the other is 
+%erronesously lower than it should be.
+
+ %\begin{figure}[h!]
+    %\centering
+    %\includegraphics[height=4cm]{AntitheticVariate.png}
+    %\caption{Antithetic variates reduce first order error in an average. 
+    %Each point in the average is paired with another point from along 
+    %the same line as the first, and at an equal, but opposite distance.}
+    %\label{fig:AntitheticVariate}
+  %\end{figure} 
+%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
+%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
+%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
+%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
+%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
+
+%The standard error of the mean (SEM) is given by:
+%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
+%a measure of the error in $n_3=\overline{w}_3$:
+%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
+%NumPoints, the number of vectors randomly generated, is increased until:
+%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
+%value that $n_3$ can be without error is 1.
+
+%Figure \ref{fig:GaussandW2_actual} ......
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{GaussandW2_actual.png}
+    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    is generated randomly according to a Gaussian distribution.} 
+  \label{fig:GaussandW2_actual}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
+    and fixed in place as $\vec{dr}$ varies randomly.} 
+  \label{fig:GaussandW2_thinkas}
+  \end{figure}
+
+%Monte-Carlo integration is employed to more efficiently 
+%compute the weighted densities used to calculate the excess free energy which is then 
+%added to the ideal free energy (computed analytically) to get the total free energy 
+%which is minimized.
+
 \section{Our Functional for the Excess Helmholtz free energy}
@@
 \begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-where the number density profile $n(\vec r)$ within the integral is given by a Gaussian function,
+where the number density profile $n(\vec r)$ within the integral is given 
+by a Gaussian function centered at a lattice point,
 %\begin{align}
@@
 \end{align}
+%
+ \begin{figure}[h!]
+    \centering
+    %\includegraphics[height=6cm]{GaussandW2_actual.png}
+    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    %view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    %is generated randomly according to a Gaussian distribution.} 
+  \label{fig:GaussandW2_actual}
+  \end{figure}
+
+Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
+$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off. 
+$\Xi$ is related to the width of the functions which 
+become wider in general as the temperature increases. 
+The radial distance 
+to the peak is related to $\alpha$, and decreases as the temperature 
+increases, that is, the shells shrink, as does 
+the sphere.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{weight_functions.pdf}
+    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2.}
+    \label{fig:weight_functions}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W2.png}
+      % \includegraphics[width=5cm]{figs/w2-ring}
+      \caption{A crossectional view of $w_2$}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W3.png}
+      \caption{A crossectional view of $w_3$}
+    \end{subfigure} 
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    \label{fig:W0andW3}
+    \end{figure}
+
 The weight fucntions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
@@
 
-\color{red}
-Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions 
-$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-sphere up to a radius of ?FIX? after which it tapers off over a distance of ?FIX?.
-
-The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in 
-Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which 
-become wider in general as the temperature increases. The radial distance to the 
-peak decreases as the temperature increases, that is, the shells shrink, as does 
-the sphere. The functions can be treated as neglible outside of a distance of 
-about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
-the weight function.
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{weight_functions.pdf}
-    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2.}
-    \label{fig:weight_functions}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W2.png}
-      % \includegraphics[width=5cm]{figs/w2-ring}
-      \caption{A crossectional view of $w_2$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W3.png}
-      \caption{A crossectional view of $w_3$}
-    \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    \label{fig:W0andW3}
-    \end{figure} 
-\color{black}
-
-\section{Computing the Liquid and Crystal Helmholtz free energies}
-
-The liquid ideal Helmholtz free energy is computed analytically 
-from Equation~\ref{} 
-for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
-%from the equation 
-%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-%\begin{align}
-  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
-%\end{align}
-%The quantity $\Lambda$ is the DeBroglie wavelength given by
-%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
-%where in our program we have set
-%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
-%\begin{displaymath}{m=}\end{displaymath}  
-%\begin{displaymath}{\hbar=}\end{displaymath} 
-%\begin{displaymath}{k_B=}\end{displaymath} 
-The crystal ideal Helmholtz free energy is computed computationally 
-from Equation~\ref{} with the varying number density profile given by
-%\begin{align}
-  %F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
-%\end{align} 
-\begin{align}
-  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
-\end{align}
-%For small gw values it is assumed that the Guassians do not overlap, 
-%in which case the equation for the free energy can be simplified, and solved
-%analytically with the result
-%\begin{equation}
- %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
-%\end{equation}
-%For larger gw values it is assumed that the Gaussians do overlap, 
-%and so the free energy is caluculated computationally. First dx is scaled by gw,
-%\begin{equation}
-  %dx_{scaled} = (dx_{input})(gw)
-%\end{equation}
-%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
-%primitive cell taking into account the contribution of Gaussians from 
-%many (5x5x5) nearby cells whose centers are given by $R_i$.
-%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
-%The contribution to the free energy of $n(\vec{r})$ at the one position is then
-%added to the accumulating sum.
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
-%This process is repeated for all positions within one primitive cell to get the
-%crystal ideal free energy for the primitive cell. This energy is then divided by
-%the volume of the primitive cell to get the crystal ideal free energy per
-%volume.
-
-The liquid excess Helmholtz free energy is computed analytically  
-%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
-using SFMT with a uniform number density profile, n that does not depend 
-on $\vec r$. 
-%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
-%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
-%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
-%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
-%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
-%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
-%The weighted densities are found from
-%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
-%where the weights $w_i$ are given by Equations~\ref{eq:weights}.
-
-The crystal excess Helmholtz free energy is computed computationally 
-using SFMT with a number density profile $n(\vec r)$ that varies spacially.
-%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
-%Then the crystal, or inhomogeneous, excess free energy is found from
-%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
-Monte-Carlo integration is used to compute the weighted densities 
-$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
-%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
-
-The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
-within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
-the Gaussians is selected that is within the primitive cell, or one of the 
-neighboring cells within a prudently picked inclusion radius as Gaussians 
-farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=5cm]{InclusionRadius.png}
-    \caption{The Gaussians at each lattice point are of width gw. 
-    After a radius of few gw's the Gaussians go to zero and can be neglected.}
-    \label{fig:InclusionRadius}
-  \end{figure}
-
-The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
-$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
-$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
-$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
-where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
-some point in space. Monte-Carlo integration works by randomly generating 
-$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
-\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
-is the gaussian number density profile for no vacancies, the equation:
-
-\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
-implemented computationally in the form:
-\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
-becomes:
-\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
-(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
-which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
-$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
-Using the relation:
-\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
-$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
-
-The result can be made more accurate by increasing NumPoints which is the 
-number of vectors randomly generated. 
-%-------Antithetic-------------------
-%In order to further reduce the error 
-%in the average, antithetic variates are used. Rather than computing an average 
-%of a set of random points, each random point in the set is paired with a 
-%point along the same line as the first point, but opposite in direction, 
-%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
-%error in the average by filtering out an erroneous offset of values along 
-%the line connecting the two points by averaging the two results, one of 
-%which may be erroneously higher than it should be while the other is 
-%erronesously lower than it should be.
+%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
+%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
+%The temperature dependence of parameters $\alpha$ and $\Xi$ 
+%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
+%can be related 
+%to the weight functions. 
+%$\Xi$ is related to the width of the functions which 
+%become wider in general as the temperature increases. 
+%The radial distance 
+%to the peak is related to $\alpha$, and decreases as the temperature 
+%increases, that is, the shells shrink, as does 
+%the sphere.
+%The functions can be treated as neglible outside of a distance of 
+%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
+%the weight function.
 
@@
     %\centering
-    %\includegraphics[height=4cm]{AntitheticVariate.png}
-    %\caption{Antithetic variates reduce first order error in an average. 
-    %Each point in the average is paired with another point from along 
-    %the same line as the first, and at an equal, but opposite distance.}
-    %\label{fig:AntitheticVariate}
+    %\includegraphics[height=7cm]{weight_functions.pdf}
+    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    %z-axis for a reduced temperature of 2.}
+    %\label{fig:weight_functions}
   %\end{figure}
-%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
-%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
-%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
-%The standard error of the mean (SEM) is given by:
-%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
-%a measure of the error in $n_3=\overline{w}_3$:
-%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
-%NumPoints, the number of vectors randomly generated, is increased until:
-%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
-%value that $n_3$ can be without error is 1.
-
-%Figure \ref{fig:GaussandW2_actual} ......
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{GaussandW2_actual.png}
-    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    %\caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
-    %the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
-    %and fixed in place as $\vec{dr}$ varies randomly.} 
-  \label{fig:GaussandW2_thinkas}
-  \end{figure}
-
-%Monte-Carlo integration is employed to more efficiently 
-%compute the weighted densities used to calculate the excess free energy which is then 
-%added to the ideal free energy (computed analytically) to get the total free energy 
-%which is minimized.
+
+ %\begin{figure}[h!]
+    %\centering
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{W2.png}
+      %% \includegraphics[width=5cm]{figs/w2-ring}
+      %\caption{A crossectional view of $w_2$}
+    %\end{subfigure} 
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{W3.png}
+      %\caption{A crossectional view of $w_3$}
+    %\end{subfigure} 
+    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    %\label{fig:W0andW3}
+    %\end{figure}
 
@@
 Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
-through \ref{fig:Phase_Diagram_P_vs_T_AS}. In all cases $\sigma=1$ and $\epsilon=1$. 
+and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
 In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
@@
 
-Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
-range of reduced temperatures from kT=0.5 to 38. The white lines show the results
-of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
-As can be seen in the figure, the region of coexistence in our simulations 
-falls within the region of coexistance predicted by Ahmed and Sadus at almost 
-all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
-the width of the coexistance regions are nearly the same, while moving to 
-higher temperatures,the width of our coexistance region ranges from about 
-one-half to one-forth the size. Although, our coexistence region is only 
-about one-forth the size at kT=28, it is centered within the coexistence 
-region predicted by Ahmed and Sadus.
+%Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
+%range of reduced temperatures from kT=0.5 to 38. The white lines show the results
+%of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
+%As can be seen in the figure, the region of coexistence in our simulations 
+%falls within the region of coexistance predicted by Ahmed and Sadus at almost 
+%all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
+%the width of the coexistance regions are nearly the same, while moving to 
+%higher temperatures,the width of our coexistance region ranges from about 
+%one-half to one-forth the size. Although, our coexistence region is only 
+%about one-forth the size at kT=28, it is centered within the coexistence 
+%region predicted by Ahmed and Sadus.
 
@@
 than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
-the temperature increases. Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
-range of reduced temperatures from kT=0.5 to 38, compated with the results
-from Ahmed and Sadus's molecular simulations. Here again, our transition 
-pressures and temperatures are consistently higher.
+the temperature increases. 
+%Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
+%range of reduced temperatures from kT=0.5 to 38, compated with the results
+%from Ahmed and Sadus's molecular simulations. Here again, our transition 
+%pressures and temperatures are consistently higher.
 
@@
 phase diagram that yielded coexistence regions that approximately 
-matched the Monte-Carlo and molecular simulation results for 
-values of kT=0.5 to 38. We are also able to generate a pressure-temperature
+matched the Monte-Carlo
+% and molecular simulation 
+results. 
+%for values of kT=0.5 to 3. 
+%kT=0.5 to 38
+We are also able to generate a pressure-temperature
 phase diagram with a shape that is expected, but the transition 

### 2020-10-30 papers/thesis-kirstie/thesis.tex
edits

@@
 \submitdate{October ??, 2020}
-\commencementyear{2020}
+\commencementyear{2021}
 
@@
 
-\date{August 2020}
+\date{October 2020}
 
@@
 The repulsive portion of the Leonard-Jones potential forms the potential for a 
-Weeks-Chandler-Anderson (WCA) model fluid, shown in Figure \ref{fig:WCA_potential}. 
+Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
+shown in Figure \ref{fig:WCA_potential}. 
 The WCA potential is described by the equation: 
@@
 from infinity as the distance between the centers of the atoms r increases, 
-and becomes, and remains, essentially zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
+and becomes, and remains, essentially zero when 
+r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
 where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
@@
 $n(\vec{r})$ through all possible spatial profiles until the free energy 
-is minimized.
+is minimized~\cite{MoritaDFT}.
 
@@
 the excess Helmholtz free energy in terms of a homogeneous number density 
-and, that is to use Scaled Particle Theory (SPT).
+and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
 
@@
 Scaled Partical Theory arrives at an expression for the excess Helmholtz 
-free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
+free energy of a homogeneous fluid~\cite{Hansen} in terms of the packing fraction $\eta$ 
 \begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
@@
 number density profile $n(\vec r)$ can be found using Fundamental Measure 
-Theory (FMT). FMT was developed for hard sphere fluids, and has been very 
+Theory (FMT)~\cite{rosenfeld1989}. FMT was developed for hard sphere fluids, and has been very 
 sucessful at predicting the thermodynamic properties of inhomogeneous hard 
 sphere fluids starting from an expression for the excess Helmholtz free 
-energy obtained from Scaled Partical Theory. Soft Fundamental Measure 
-Theory (SFMT), which will be addressed later, extends FMT to fluids with 
-soft potentials like that of the WCA fluid.
+energy obtained from Scaled Partical Theory~\cite{santos2012phi3}. 
+Soft Fundamental Measure Theory (SFMT), which will be addressed later, 
+extends FMT to fluids with soft potentials like that of the WCA fluid.
 
@@
 were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
-expression for $\Phi_3$ developed by Tarazona, the excess free energy becomes
+expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
+the excess free energy becomes
 \begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
@@
 fluids with soft potenials by keeping the relationships between the weight 
-functions, and that the weight functions satisfy the same, but replacing 
+functions, and which the weight functions satisfy the same, but replacing 
 $w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
-A new set of weight functions can then be derived for a system of soft spheres.
+A new set of weight functions can then be derived for a system of soft 
+spheres which is what we do in forming our functional for the 
+Weeks-Chandler-Anderson fluid as detailed in the chapter on Methods.
 
@@
 
-\section{Overview}
+\section{The Overall Approach}
 %
@@
 \begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
+where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time:
+\begin{equation}n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~\end{equation}
+
   \begin{figure}[h!]
@@
 
+%Classical Density Functional Theory is used to find the equilbrium 
+%Helmholtz free energy of the system at a given temperature and density
+%by varying the number density profile $n(\vec{r})$ until the free energy 
+%is minimized. The minimum free energy, and its corresponding number 
+%density profile are then the equilibrium values for that temperature 
+%and number density.
+
 %The number density profile is varied by varying gw and fv, and the 
-For a given combination of number density and temperature, a suite of number 
-density profiles are generated.
-% like those represented by Figure~\ref{fig:Ensemble_vary}.
-The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number density profile,
-and also for the homogeneous number denisty profile. 
-Both the ideal and excess free energies are found, and added together to get the total Helmholtz 
-free energy.
-The Helmholtz free energy is minimized for a system with a fixed temperature, 
-volume, and number of particles. Setting the number of particles to one, 
-the volume can be found from the number density.
-The process is then repeated to find the free energies at other combinations 
-of number densities and temperatures.
+For a given combination of temperature and number density, a suite of number 
+density profiles are generated similiar to those represented in the example 
+set of profiles shown in Figure~\ref{fig:Ensemble_vary}. 
+The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
+density profile, and also for the homogeneous number denisty profile 
+characteristic of a liquid. Both the ideal and excess free energies are 
+found, and added together to get the total Helmholtz free energy.
+According to classical Density Functional Theory, the minimum free energy, 
+and its corresponding number density profile are then the equilibrium values 
+for that temperature and number density.
+%The Helmholtz free energy is minimized for a system with a fixed temperature, 
+%volume, and number of particles. Setting the number of particles to one, 
+%the volume can be found from the number density.
+The process is repeated to find the free energies at other combinations 
+of temperatures and number densities.
+
+%Once the Helmholtz free energy for both the crystal and liquid states of 
+%the fluid are found for a given temperature and density, they can be 
+%compared to find the resulting state of the fluid. 
+Outside the region of coexistence which exists between the liquid and 
+crystal states, the fluid will be completely in a crystalline state if the 
+crystal free energy is lower than the liquid free energy, and it will be 
+completely in a liquid state if the liquid free energy is lower than 
+the crystal free energy. Maxwell's Double Tangent Construction is used 
+to determine the actual number densities of the crystal and liquid states
+at the boundaries of the coexistance region for a given temperature, 
+and construct phase diagrams as described in the section on The Theory of
+Freezing and Maxwell's Double Tangent Construction.
 
@@
 \begin{align}
-  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
+  n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
 \end{align}
@@
 explicitly taking into account the interaction of each sphere with all the others, 
-but it must be taylored to fit a particular fluid. 
+but it must be taylored to fit a particular fluid\cite{schmidt1999density}. 
 %The excess Helmholtz free energy is particular to the fluid being modeled, 
@@
     \centering
-    \includegraphics[height=9cm]{B2_vs_T_with_xi_fromB2.pdf}
+    \includegraphics[height=8cm]{B2_vs_T_with_xi_fromB2.pdf}
     \caption{Shows the temperature dependence of $B_2$ generated from 
@@
     \centering
-    \includegraphics[height=9cm]{xi_fromB2.pdf}
+    \includegraphics[height=8cm]{xi_fromB2.pdf}
     \caption{Shows the temperature dependence of $\Xi$ generated from 
@@
     \centering
-    \includegraphics[height=9cm]{erf_Potential_xifromB2.pdf}
+    \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
     \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
@@
 
-\color{blue}
+%\color{blue}
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
 \begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
-\color{black}
+%\color{black}
 
@@
 where $Q_n$ is the configuration integral.
-\color{red}For a system of N particles only, the sum collapses to one term, 
-that for n=N, which gives \color{black}
+For a system of N particles only, the sum collapses to one term, 
+that for n=N, which gives 
 \begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
@@
 
-\color{red}This result was for low densities where higher terms in the 
+This result was for low densities where higher terms in the 
 expression for the excess free energy were neglected. The full virial 
-equation is given by\color{black}
+equation is given by
 \begin{equation}\frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) \end{equation}
 
-%------------------n(r)--move to ``back" later---------------------
-------------------------------------------------------------------
-
-The actual positions of atoms in a microstate  \color{red}?at time t, or averaged over time?:\color{black}
-\begin{equation}\rho_{s}(\vec r)=\sum_{i=1}^N\delta(\vec r - \vec r_i)\end{equation}
-The equilibrium single number density distribution equals the ensemble average over all microstates, and is independent of time:
-\begin{equation}n(\vec r)=\rho^{(1)}(\vec r)=~<\rho_s(\vec r)>~\end{equation}
-The reduced equilibrium phase-space probablity density of n particles in volume $dr^ndp^n$:
-\begin{equation}f_o^{(n)}(\vec r^n, \vec p^n)=\rho^{(n)}(\vec r^n)f_M^{(n)}(\vec p^n)\end{equation}
-The relation between the  reduced equilibrium phase-space probablity density and the equilibrium phase-space probablity density:
-\begin{equation}f_o^{(n)}(\vec r^n, \vec p^n)=\frac{N!}{(N-n)!}\int\dot~\dot~\dot~\int f_o^{[N]}(\vec r^N, \vec p^N) d\vec r^{N-n} d\vec p^{N-n}\end{equation}
-The canonical equilibrium phase-space probablity density:
-\begin{equation}f_o^{[N]}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}
+\section{Derivation of $F = -k_BTln(Z)$}
+The canonical equilibrium phase-space probablity density is given by
+\begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}
+\begin{equation}{partian function}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
+The Helmholtz free energy is given by 
+\begin{equation}\label{F=U-TS}F = U - TS\end{equation}
 where
-\begin{equation}{partian function}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
-The equilibrium phase-space probablity density integrated over all continuous microstates $(\vec r^N,\vec p^N)$ is 1:
-\begin{equation}\int\dot~\dot~\dot~\int f_o^{[N]}(\vec r^N, \vec p^N) d\vec r^{N} d\vec p^{N}=1\end{equation}
-The equilibrium phase-space probablity density is independent of time:
-\begin{equation}f_o^{[N]}(\vec r^N, \vec p^N) = f_o^{[N]}(\vec r^N, \vec p^N; t)\end{equation}
-The Helmholtz free energy is a functional of the equilibrium phase-space probability density:
-\begin{equation}U = <E> = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N}\end{equation}
-\begin{equation}TS = -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln\left(N!~h^{3N}f_o^{[N]}\right) d\vec r^{N} d\vec p^{N} \end{equation}
-\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N} \end{equation}
-\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[f_o^{[N]}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right] \end{equation}
-\begin{equation}  =-k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[f_o^{[N]}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right] \end{equation}
-\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}ln(Z)d\vec r^{N} d\vec p^{N}\end{equation}
-\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}d\vec r^{N} d\vec p^{N}\right]\end{equation}
-\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[1\right]\end{equation}
-and so 
-\begin{equation}F = U - TS\end{equation}
-\begin{equation} = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N}-\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int f_o^{[N]}~H~d\vec r^{N} d\vec p^{N} -k_BTln(Z)\end{equation}
+\begin{equation}U = <E> = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N}\end{equation}
+\begin{align}
+ TS &= T\left[-k_B\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(N!~h^{3N}P_{eq}\right) d\vec r^{N} d\vec p^{N}\right]   \\
+	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N}   \\
+	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]  \\
+    &= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]
+\end{align}  
+\begin{align}  
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}ln(Z)d\vec r^{N} d\vec p^{N}     \\
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}d\vec r^{N} d\vec p^{N}\right]   \\
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[1\right]   
+\end{align}
+Plugging these into Equation~\ref{F=U-TS} gives
+%\begin{equation}F = U - TS\end{equation}
+\begin{equation} F = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N}-\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} -k_BTln(Z)\end{equation}
 \begin{equation}F = -k_BTln(Z)\end{equation}
-
-----------------------------------------------
 

### 2020-10-31 papers/thesis-kirstie/thesis.tex
edits

@@
 at a given temperature and density. 
-%Once the Helmholtz free energy for both the crystal and liquid states of 
-%the fluid are found for a given temperature and density, they can be 
-%compared to find the resulting state of the fluid. 
-%
-%save:
-%Outside the region of 
-%coexistence which exists between the liquid and crystal states, and which 
-%consists of various amounts of liquid and crystal existing together,
-%the fluid will be completely in a crystalline state if the 
-%crystal free energy is lower than the liquid free energy, and it will be 
-%completely in a liquid state if the liquid free energy is lower than 
-%the crystal free energy. Maxwell's Double Tangent Construction is used 
-%to determine the actual number densities of the crystal and liquid states
-%at the boundaries of the coexistance region for a given temperature from 
-%which the phase diagrams can be constructed.
-%end save
-%
-%In general, if at some number density and temperature the crystal free energy 
-%is lower than the homogeneous free energy of the liquid, then the fluid will 
-%be in the crystalline state. If the homogeneous liquid free energy is lower 
-%than the crystal free energy, then the fluid will be in the liquid state. 
-%
-%To find the equilibrium Helmholtz free energies we use classical Density 
-%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
-%%the corresponding equilibrium number density profile. 
-%In particular, 
-In general, the Helmholtz free energy, 
-which consists of an ideal part and an excess part ,
-can be written as a functional of 
-the number density profile $n(\vec{r})$ which describes the way the average number 
+In general, the Helmholtz free energy, which consists of an ideal part 
+and an excess part, can be written as a functional of the number density 
+profile $n(\vec{r})$ which describes the way the average number 
 density varies spatially. 
 \begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
-%
 
@@
 be used to form the functional for the excess Helmholtz free energy. 
-%to predict the freezing of a Weeks-Chandler-Anderson 
-%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
-%dependent parameters to form the functional for the excess Helmholtz free energy. 
-%for generating the weight funcions, and Tarazona's tensor functional. 
-%
-%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
 According to classical Density Functional Theory the equilbrium 
-Helmholtz free energy of the system 
-can be found
-by varying the number density profile $n(\vec{r})$ until the free energy 
-is minimized. 
-%The minimum free energy, and its corresponding number 
-%density profile are then the equilibrium values of the Helmholtz free energy
-%of the system at the given temperature and number density. 
-%
-%To find the Helmholtz free energy both the ideal Helmholtz free energy
-%and the excess Helmholtz free energy must be found and added together.
-%The ideal Helmholtz free energy is the free energy the fluid would have in
-%the absence of any interactomic potential energy $V(r)$ between the spheres, 
-%or atoms that make up the fluid. 
-%
-%The excess Helmholtz free energy is that 
-%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
-%energy $V(r)$ between the spheres.
-%
-%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%For a spatially varying number density, as would be expected for a crystal, 
-%the free energy is a functional of the number density profile $n(\vec{r})$  
-%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
-%analytically to find the resulting free energy for the system. 
-%
+Helmholtz free energy of the system can be found by varying the number 
+density profile $n(\vec{r})$ until the free energy is minimized. 
 SFMT makes the task of forming an expression
-%is a method used to form an expression 
 for the excess Helmholtz free energy of a particular model fluid 
@@
 has with every other sphere does not have to be taken explicitly into account.
-%without 
-%having to take explicitly into account the interaction each sphere, or atom, 
-%has with every other sphere which would make the task very difficult, 
-%or even prohibtive.
-%
-%Classically, fluids have been modeled by a collection of small, 
-%impenetrable spheres colliding with one another like billiard balls. 
-%In the simplest case, the spheres are treated as having only a replulsive 
-%potential. This model, called the Hard-Sphere fluid, has served as the 
-%primary reference system for fluids. This is due to the fact that the 
-%repulsion between atoms is much stronger than their attraction, and so 
-%the repulsive potential, which is descriptive of hard spheres, dominates 
-%the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
-%fluid which is a perturbation of the Hard-Sphere fluid created by adding 
-%a hard-edge attractive potential to the repulsive potential of the 
-%Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-%spheres with softly fading replusive potentials which he modeled by making
-%the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
-%This reflects the idea that soft-spheres penetrate more deeply into 
-%each other when they collide at higher temperatures. 
-%
-%In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
-%with Rowlinson's idea of using temperature-dependent radii to model spheres 
-%with soft repulsive potentials to create the Barker-Henderson model 
-%fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
-%modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
-%at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
-%Chandler, and Anderson used temperature and density dependent hard-sphere 
-%radii to model spheres with the more realistic softly fading repulsive 
-%portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-%which is known as the Weeks-Chandler-Anderson model fluid. f
-%
-%While models for particular fluids were being created, powerful methods 
-%to describe the properties of fluids in general were also being developed. 
-%In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
-%introduced from which an equation of state, and an expression for the 
-%excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
-%obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
-%alternative method...OZ... \color{black}
-%%Percus and Yevick took another approach to the Hard-Sphere fluid.
-%In 1961 Morita and Hiroike introduced a density functional 
-%approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
-%(the equilibrium Helmholtz free energy in the case of a canonical 
-%ensemble), and the cooresponding equilibrium number density profile for 
-%inhomogeneous fluids. To use this approach, a functional must be 
-%constructed for a particular fluid.
-%%DFT Evans 1979
-%%Led by the consistency bewteen results from SPT and Percus-Yevick, 
-%In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-%(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
-%functional for the Hard-Sphere fluid. Tarazona later improved on the 
-%method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
-%santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
-%forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
-%
-%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
-%with soft repulsive and soft attractive potentials, and developed an erf 
-%model for spheres with steep, short-range soft repulsive 
-%potentials~\cite{schmidt2000fluid}.
-% like that exhibited by a WCA fluid. 
-%
 Results will be compared to those from Monte-Carlo simulations.
@@
 together until their atomic orbitals begin to overlap. 
-%At this point the electrons in each atom begin to repel the electrons 
-%in the other atom and oppose the overlap of the two atoms.
 At this point the electrons in each atom begin to oppose the overlap of 
@@
 as do hard spheres. 
-%This structure is the three-dimensional equivilant to the two-dimensional 
-%structure that one would see by filling the bottom of a box with a single layer of marbles. 
 This structure is what one would see by filling a cube with marbles. It 
@@
 Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
-diagram shown in Figure \ref{fig:P-T_Diagram}. This phase diagram is 
+diagram shown in Figure~\ref{fig:P-T_Diagram}. This phase diagram is 
 representative of some arbitrary, unknown substance composed of atoms, 
@@
 the atoms in a WCA fliud do not exhibit attraction there will not be a 
-dividing line between the gas and liquid phases, and so the term ``liquid" 
-or ``gas" is not used, but simply ``fluid". Points along the line that 
+dividing line between the gas and liquid phases, and so the term "liquid" 
+or "gas" is not used, but simply "fluid". Points along the line that 
 separates the solid and the fluid phases indicate a phase transition, 
@@
 %to each curve have the same slope.%
-Two such curves are shown in Figure \ref{fig:MaxwellDT}, one for the liquid 
+Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
 and one for the solid at a fixed temperature. Since the pressure of the 
@@
 at the transition point can be found the same way, as shown in Figure 
-\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as that in Figure \ref{fig:T-n_Diagram}, 
-can be generated by repeating this procedure for various temperatures. 
+\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as 
+that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this 
+procedure for various temperatures. 
 %The lowest density the crystal exhibts and the highest density the fluid 
@@
 when the natural variables of the free energy are held fixed~\cite{schroeder}. 
-%And so the free energy tends toward a minimum. 
-%How will a system at a fixed temperature, volume, and number of 
-%particles respond to a small offset from equilibrium? 
-%
-%Save for me:
-%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-%\begin{equation}
-  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
-%\end{equation}
-%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
-%$dT=0$ gives
-%\begin{equation}
-  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
-%\end{equation}
-%where conservation of energy was also applied. Fixing the volume and the 
-%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
-%results in
-%\begin{equation}
-  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
-%\end{equation}
-%which, for fixed N, is the same as
-%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
-%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
-%of particles are fixed, the change in the free energy of the system will 
-%always be negative, and thus the free energy tends toward a 
-%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
-%for a system near equilibrium. 
-%End save for me.
 A general proof of minimization 
@@
 potential energies between pairs of atoms, where each atom is paired in 
-turn with each of the other atoms in the system. In this case $V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ 
-becomes 
+turn with each of the other atoms in the system. In this case 
+$V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ becomes 
 \begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
@@
 $V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)$ to zero, the exponent in 
-Equation~\ref{Fexcess} becomes
+Equation~\ref{Fexcess-mostgeneral} becomes
 
@@
 
-%The number of atoms in one primitive cell is one.
-
-%\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
-%\begin{displaymath}F_{crystal} = F_{ideal.crystal} + F_{excess.crystal}\end{displaymath}
-%\newline
-%\newline\noindent Inputs: 
-%t*, n*, range of gw values , range of fv values
-%\newline Outputs: Homogeneous (liquid)  Helmholtz free energy (at the given T,V,N) 
-%\newline\indent\indent\indent Crystal Helmholtz free energy (at the given T,V,N)
-
-%PROGRAM FLOWCHART ---------------------------
-%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
-%Loop: 
-%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
-%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
-%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
-%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
-%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
-%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
-
-%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
-%\begin{displaymath}\left(\frac{F}{N}\right)_{homogeneous}\end{displaymath}
-%lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
-
 \section{Computing the Liquid and Crystal Helmholtz free energies}
-%-----------Comuting the Liquid and Crystal free energies ------
 
@@
 number of vectors randomly generated.
-%-------Antithetic-------------------
-%In order to further reduce the error 
-%in the average, antithetic variates are used. Rather than computing an average 
-%of a set of random points, each random point in the set is paired with a 
-%point along the same line as the first point, but opposite in direction, 
-%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
-%error in the average by filtering out an erroneous offset of values along 
-%the line connecting the two points by averaging the two results, one of 
-%which may be erroneously higher than it should be while the other is 
-%erronesously lower than it should be.
-
- %\begin{figure}[h!]
-    %\centering
-    %\includegraphics[height=4cm]{AntitheticVariate.png}
-    %\caption{Antithetic variates reduce first order error in an average. 
-    %Each point in the average is paired with another point from along 
-    %the same line as the first, and at an equal, but opposite distance.}
-    %\label{fig:AntitheticVariate}
-  %\end{figure} 
-%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
-%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
-%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
-%The standard error of the mean (SEM) is given by:
-%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
-%a measure of the error in $n_3=\overline{w}_3$:
-%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
-%NumPoints, the number of vectors randomly generated, is increased until:
-%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
-%value that $n_3$ can be without error is 1.
-
-%Figure \ref{fig:GaussandW2_actual} ......
 
@@
     \includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 
     the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
@@
 \begin{equation}U=TS-PV+\mu{N}\end{equation}
-using the relations from statistical thermophysics which views the 
-thermodynamic quantities as ensemble averages
+using the relations
 \begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
@@
 greater compression before forming a solid. The white lines show results from Monte-Carlo 
-simulations of the WCA fluid. Overall, the general shape, and position of the lines correspond
+simulations of the WCA fluid~\cite{May}. Overall, the general shape, and position of the lines correspond
 fairly closely to the edges of the region of coexistance predicted from the DFT method, 
@@
 Plugging these into Equation~\ref{F=U-TS} gives
-%\begin{equation}F = U - TS\end{equation}
 \begin{equation} F = \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N}-\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} -k_BTln(Z)\end{equation}
@@
 
+%\section{Program Flowchart}
+
+%\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
+%\begin{displaymath}F_{crystal} = F_{ideal.crystal} + F_{excess.crystal}\end{displaymath}
+%\newline
+
+%\newline\noindent Inputs: 
+%t*, n*, range of gw values , range of fv values
+%\newline Outputs: Homogeneous (liquid)  Helmholtz free energy (at the given T,V,N) 
+%\newline\indent\indent\indent Crystal Helmholtz free energy (at the given T,V,N)
+
+%PROGRAM FLOWCHART:
+%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
+%Loop: 
+%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
+%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
+%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
+%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
+%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
+%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
+%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+
+%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
+%\begin{displaymath}\left(\frac{F}{N}\right)_{homogeneous}\end{displaymath}
+%lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
+
+%The number of atoms in one primitive cell is one.
+
 \bibliography{thesis}
@@
 \end{document}
-
-  %\cite{TschoppBHextended}
-
-%\cite{rowlinson1964statistical}
-%\cite{gil-villegas-1997-SAFT-VR, clark2006developing,
-%  lafitte2013accurate}
-%\cite{carnahan1969equation}
-%\cite{rosenfeld1989, rosenfeld1997,
-%  roth2002whitebear}
-
-	%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density, hughes2013classical, krebs2014improved}
-%\cite{cuesta1997dimensional, hansen2009fundamental, marechal2013density}
-%\cite{schmidt2000density, groh2001density, kim2001adsorption,
-%  sweatman2002fundamental}
-
-%\cite{weeks1971}
-	%\cite{rosenfeld1989}
-	%%\cite{tarazonaFIXME,santos2012phi3}
-	%\cite{schmidt1999density} 
-	%\cite{schmidt2000fluid}
-
-%\cite{barker1967perturbation}
-%\cite{roth2002whitebear} 
-	%\cite{roth2002whitebear}
-
-%\cite{mikolaj2004structure, eggert2002quantitative,
-%yarnell1973structure}
-	%\cite{verlet1967computer}
-
-%-------------Minimization-----------------
-%The minimization can be shown by considering that according to the second law of thermodynamics, a system that is not in equilibrium will exprience spontanteous processes that increase the total entropy of a system and its surroundings until equilibrium is reached, $\Delta{S}_{System + Surroundings} \geq 0$.
-Here ``surroundings" means the entire universe minus the system.
-This, together with the conservation of energy (the first law of thermodynamics) which says that whatever energy is lost (or gained) by the system is gained (or lost) by the surroundings $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that any change in the free energy will be negative with each spontaneous process until a minimum is reached if the natural variables of the free energy are held fixed.
-How will a system at a fixed temperature, volume, and number of particles respond to a small offset from equilibrium?
-Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-%\begin{equation}d\left(\frac{F}{N}\right)_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\frac{S}{N}dT_{System}\end{equation}
-%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that $dT=0$ gives
-%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}\end{equation}
-%where conservation of energy was also applied. Fixing the volume and the number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, results in
-%\begin{equation}d\left(\frac{F}{N}\right)_{System}=-\frac{T}{N}dS_{System+Surroundings} \end{equation}which, for fixed N, is the same as
-%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number of particles are fixed, the change in the free energy of the system will always be negative.
-(reference schroder)
-
-%More generally, the Helmholtz free energy is the value assumed by U-TS for a system that has a fixed volume, number of particles, and temperature that equals the temperature of the reservoir, or surroundings as it settles spontaneously to equilibrium.
-Under these conditions, the quantity U-TS settles to a minimum at equilibrium with respect to unconstrained variables, such as the number density profile $n(\vec r)$.
-%\begin{equation}F=\text{min}_{U}(U_{Not.equilibrium} - TS_{Not.equilibrium}) = U_{equilibrium}-TS_{equilibrium}\end{equation}
-%An extremum is reached when
-%\begin{equation}\frac{\partial{F}}{\partial{U}}=0\end{equation}
-
-%\begin{equation}F=U-TS\end{equation}
-%\begin{equation}\frac{F}{N}=\frac{U}{N}-\frac{TS}{N}\end{equation}
-%\begin{equation}\label{usetoshowmin}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
-%\begin{equation}d\left(\frac{F}{N}\right)=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
-%Setting $f_N=\frac{F}{N}$ this becomes
-%\begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
-
-%-------Scaled Partical Theory-----------------
-Scaled particle theory seeks a general expression for the excess chemical potential of a homogeneous Hard Sphere fluid 
-which can then be integrated to find the excess Helmholtz free energy.
-\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
-It is an extension of the work done by Widom who showed that the excess chemical potential of a homogenesous fluid corredsponds to the reversible work 
-of inserting a sphere of radius R into a fluid consisting of hard spheres of radius R.
-
-Widom's result was obtained by considering the difference in the excess Helmholtz free energy when one atom is added to an otherwise closed system. 
-He showed that the excess chemcial potential is related to the ensemble average of the Boltzman factor for the potential energy difference $V_{diff}$ 
-between a system with N atoms and one with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
-%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
-\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
-\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-%Equation~\ref{widom} is called Widom's Insertion Formula.
-
-%%COME BACK TO THIS POINT!
->>>>>\color{red}For a system...
-$<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widoms-insertion-formula} becomes \begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
-
-%In a randomly positioned (I think it has to be homogeneous!)
-\noindent To evaluate Equation~\ref{widoms-insertion-formula} for a Hard Sphere fluid consider that
-no energy would be required to place a hard sphere at points surrounded by an empty space the size of a hard sphere. 
-But a hard sphere cannot be placed at points where a hard sphere would overlap another hard sphere; 
-the energy required to do that would be infinite. Thus, for a Hard Sphere fluid, Equation~\ref{widoms-insertion-formula} becomes
-\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
-\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-where $P_{NoOverlap}$ is the probablity of positions existing in the hard fluid that are surrounded by an empty space the size of a hard sphere.
-
-\begin{figure}[h!]
-    \centering
-    %\includegraphics[height=5cm]{Probability-of-overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where there is no overlap such 
-             as shown by the black dot at the center of the dotted region near the top. 
-             But it is not possible to place a sphere at a position where an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
-
-For a homogeneous system, the probablity of positions existing in the hard fluid that are surrounded 
-by an empty space the size of a hard sphere, $P_{NoOverlap}$, can be found by considering 
-that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the number density times the volume of one sphere. 
-%\color{red}Thus the chemical potential for a fluid with cavities having a radius equal to the radius R of a hard sphere is \color{black}
-\color{red}Thus, for cavities having a radius Ro equal to the radius R of a hard sphere, the chemical potential is found to be \color{black}
-%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-\begin{equation}\mu_{ex}(Ro=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-\color{red} ???More generally, for Ro $<$ R this becomes???
-\begin{equation}???\mu_{ex}(Ro < R)=-k_BT(\ln(1-n\frac{4\pi}{3}(R_o+R)^3))???\end{equation}
-\color{black}
-
-NOW DO UPPER LIMIT  Ro $>>$ R
-
-For Ro $>>$ R ...
---------------------------------------------------
-
-%\section{Derivation of the Excess Free Energy from Scaled Particle Theory}
-%The excess chemical potential of a homogeneous fluid corresdponds to the 
-%reversible work of inserting a sphere of radius R 
-%into a fluid consisting of hard spheres of radius R. 
-%\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
-%This result is obtained by considering the difference in the excess Helmholtz 
-%free energy when one atom is added to the system. It was shown by Widom 
-%to be related to the ensemble average of the Boltzman factor for the 
-%potential energy difference $V_{diff}$ of a system with N atoms and one 
-%with N+1 atoms. The resulting equation is known as Widom's Insertion Formula.
-%%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
-%\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
-%\begin{equation}\label{widom}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-%%Equation~\ref{widom} is called Widom's Insertion Formula. 
-%\color{red}For a system... $<e^{-\beta{V_{diff}}}> \rightarrow e^{-\beta{V_{diff}}}$, and Equation~\ref{widom} becomes 
-%\begin{equation}\mu_{ex}=-k_BT\ln\left(e^{-\beta{V_{diff}}}\right)=-k_BT(-\beta{V}_{diff})=V_{diff}=W_{rev}\end{equation}\color{black}
-%%In a randomly positioned (I think it has to be homogeneous!)
-%\indent In a collection of hard spheres making up a hard sphere fluid, 
-%no energy would be required to place a hard sphere at points 
-%surrounded by an empty space the size of a hard sphere. But a hard sphere 
-%cannot be placed at points where a hard sphere would 
-%overlap another hard sphere; the energy required to do that would be infinite. 
-%Thus, for a Hard Sphere fluid, Equation~\ref{widom} becomes
-%\begin{equation}\mu_{ex}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{V}_{NoOverlap}}+P_{Overlap}e^{-\beta{V}_{Overlap}}\right)\end{equation}
-%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}e^{-\beta{0}}+P_{Overlap}e^{-\beta\infty}\right)\end{equation}
-%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}(1)+P_{Overlap}(0)\right)\end{equation}
-%\begin{equation}=-k_BT\ln\left(P_{NoOverlap}\right)\end{equation}
-%where $P_{NoOverlap}$ is the probablity of positions existing in the hard 
-%fluid that are surrounded by an empty space the size of a hard sphere.
-
-%\begin{figure}[h!]
-    %\centering
-    %% FIXME \includegraphics[height=3cm]{Probability-of-overlap.pdf}
-    %\caption{No energy is needed to place a sphere at a position where 
-    %there is no overlap such as shown by the black dot at the center of 
-    %the dotted region near the top. But it is not possible to place a 
-    %sphere at a position where an overlap occurs.}
-    %\label{fig:p_overlap}
-  %\end{figure}
-
-%For a homogeneous system, the probablity of positions existing in the hard 
-%fluid that are surrounded by an empty space the size of a hard sphere, 
-%$P_{NoOverlap}$, can be found by considering 
-%that $P_{NoOverlap}~=1-~P_{Overlap}$ where $P_{Overlap}$ is simply the 
-%number density times the volume of one sphere. 
-%%\color{red}Thus the chemical potential for a fluid with cavities having 
-%%a radius equal to the radius R of a hard sphere is \color{black}
-%\color{red}Thus, considering cavities having a radius Ro equal to the 
-%radius R of a hard sphere, the chemical potential is found to be \color{black}
-%%\begin{equation}P_{Overlap}=nV_{sphere}\right)\end{equation}
-%\begin{equation}\label{mu_excess}\mu_{ex}(R_{cavity}=R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-%%\begin{equation}\mu_{ex}(R)=-k_BT\ln\left(1-nV_{sphere}\right)\end{equation}
-
-%Using this result, scaled particle theory seeks a more general expression 
-%of the excess chemical potential of a hard sphere fluid by considering the 
-%work of inserting a hard sphere into a cavity of radius $R_{cavity}$ 
-%where the size of $R_{cavity}$ falls into three different cases: 
-%$R_{cavity} >> R$, $R_{cavity} > R$, and $R_{cavity} \leq R$.
-
-%\begin{equation}\label{3_parts_to_mu_ex}{W_{rev}(R_{cavity})=\left\{ \begin{array}{rc} P\frac{4\pi}{3}R_{cavity}^3 & R_{cavity} >> R 
-%\\ P\frac{4\pi}{3}R_{cavity}^3 + c_2R_{cavity}^2 + c_1R_{cavity}^1 + c_0R_{cavity}^0  & R_{cavity} > R \\ -k_BT(\ln(1-n\frac{4\pi}{3}(R_{cavity})^3)) & R_{cavity} \leq R\end{array}\right.}\end{equation}
-
-%%The first function, valid for $R_o >> R$, yields the dominant term to which the second function, valid for $R_o > R$,  
-%%must tend toward in the limit that $R_o$ goes to infinity. By expanding the logrythmic term (ln(1-x)= ...) 
-%%in the last function, valid for $R_o < R$, the function can be expressed in terms of powers of $R_o$.
-
-%%The first and last functions in Equation~\ref{3_parts_to_mu_ex} are derived from theory, and since they 
-%The first and last functions in Equation~\ref{3_parts_to_mu_ex} can be expressed in terms of powers of $R_{cavity}$, 
-%and so it is assumed that the middle function, which is valid for values of $R_{cavity}$ in between the two extremes, 
-%can also be expressed in terms of powers of $R_{cavity}$.  
-%The coeficients $c_2$, $c_1$ and $c_0$ are determnined by considering that the three functions must be conintuous at 
-%the boundaries along with their first and second derivatives, and using Equation~\ref{mu_excess}.The chemical potential then becomes
-%\begin{equation}\mu_{ex}=-k_BT\left(\ln(1-\eta) + \frac{6\eta}{1-\eta}+\frac{9\eta^2}{2(1-\eta)^2}+\frac{PV\eta}{k_BTN}\right)\end{equation}
-%where $\eta$ is the packing fraction given by
-%\begin{displaymath}\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}=\frac{N\frac{4}{3}\pi{R}^3}{V}\end{displaymath}
-
-%%The last two functions, valid for $R_o > R$, and $R_o < R$, and their first and second derivatives must be continuous at $R_o=R$. 
-%%From these considerations, factors $c_2$, $c_1$ and $c_0$ can be determnined, and using the relation 
-%%...ADD!...P, an expression is found for the reversible work done when $R_o=R$ which is equal to the excess chemical potential. 
-%%From the expression for pressure given by
-
-%The pressure can be obtained from \color{red}
-%\begin{equation}{\frac{\partial{P}}{\partial{N}}\bigg|_{T,V}V=N\frac{\partial\mu}{\partial{N}}\bigg|_{T,V}}\end{equation} \color{black}
-%%The excess Hemlholtz free energy is then found from the relation
-%and the Helmholz free energy from
-%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\text{or}{~~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{displaymath}
-%which gives
-%%\begin{displaymath}F_{ex}=\end{displaymath}
-%\begin{equation}F_{excess}=k_BTN\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)\end{equation}

### 2020-11-02 papers/thesis-kirstie/thesis.tex
edit from last week

--- 2020-10-30 papers/thesis-kirstie/thesis.tex
@@
 
-The goal of this project is to generate computationally temperature-density 
-and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
-model fluid using classical Density Functional Theory (cDFT). 
-To do this we need to find the Helmholtz free energies that the fluid
-would have in a liquid state, and in a crystalline solid state 
-at a given temperature and density. 
-%Once the Helmholtz free energy for both the crystal and liquid states of 
-%the fluid are found for a given temperature and density, they can be 
-%compared to find the resulting state of the fluid. 
-%
-%save:
-%Outside the region of 
-%coexistence which exists between the liquid and crystal states, and which 
-%consists of various amounts of liquid and crystal existing together,
-%the fluid will be completely in a crystalline state if the 
-%crystal free energy is lower than the liquid free energy, and it will be 
-%completely in a liquid state if the liquid free energy is lower than 
-%the crystal free energy. Maxwell's Double Tangent Construction is used 
-%to determine the actual number densities of the crystal and liquid states
-%at the boundaries of the coexistance region for a given temperature from 
-%which the phase diagrams can be constructed.
-%end save
-%
-%In general, if at some number density and temperature the crystal free energy 
-%is lower than the homogeneous free energy of the liquid, then the fluid will 
-%be in the crystalline state. If the homogeneous liquid free energy is lower 
-%than the crystal free energy, then the fluid will be in the liquid state. 
-%
-%To find the equilibrium Helmholtz free energies we use classical Density 
-%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
-%%the corresponding equilibrium number density profile. 
-%In particular, 
-In general, the Helmholtz free energy, 
-which consists of an ideal part and an excess part ,
-can be written as a functional of 
-the number density profile $n(\vec{r})$ which describes the way the average number 
-density varies spatially. 
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-%
-
-Classical Density Functional Theory is used to find the equilibrium 
-Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
-be used to form the functional for the excess Helmholtz free energy. 
-%to predict the freezing of a Weeks-Chandler-Anderson 
-%model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
-%dependent parameters to form the functional for the excess Helmholtz free energy. 
+Classically, fluids have been modeled by a collection of small, 
+impenetrable spheres colliding with one another like billiard balls. 
+In the simplest case, the spheres are treated as having only a replulsive 
+potential. This model, called the Hard-Sphere fluid, has served as the 
+primary reference system for fluids. This is due to the fact that the 
+repulsion between atoms is much stronger than their attraction, and so 
+the repulsive potential, which is descriptive of hard spheres, dominates 
+the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
+fluid which is a perturbation of the Hard-Sphere fluid created by adding 
+a hard-edge attractive potential to the repulsive potential of the 
+Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
+spheres with softly fading replusive potentials which he modeled by making
+the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+This reflects the idea that soft-spheres penetrate more deeply into 
+each other when they collide at higher temperatures.
+
+In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
+with Rowlinson's idea of using temperature-dependent radii to model spheres 
+with soft repulsive potentials to create the Barker-Henderson model 
+fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
+modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
+at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
+Chandler, and Anderson used temperature and density dependent hard-sphere 
+radii to model spheres with the more realistic softly fading repulsive 
+portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
+which is known as the Weeks-Chandler-Anderson model fluid.
+
+While models for particular fluids were being created, powerful methods 
+to describe the properties of fluids in general were also being developed. 
+In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
+introduced from which an equation of state, and an expression for the 
+excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
+obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
+alternative method...OZ... \color{black}
+%Percus and Yevick took another approach to the Hard-Sphere fluid.
+In 1961 Morita and Hiroike introduced a density functional 
+approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
+(the equilibrium Helmholtz free energy in the case of a canonical 
+ensemble), and the cooresponding equilibrium number density profile for 
+inhomogeneous fluids. To use this approach, a functional must be 
+constructed for a particular fluid.
+%DFT Evans 1979
+%Led by the consistency bewteen results from SPT and Percus-Yevick, 
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
+functional for the Hard-Sphere fluid. Tarazona later improved on the 
+method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
+santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
+forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
+
+Schmidt applied his Soft Fundamental Measure Theory to various fluids 
+with soft repulsive and soft attractive potentials, and developed an erf 
+model for spheres with steep, short-range soft repulsive 
+potentials~\cite{schmidt2000fluid}. In this paper, Density Functional 
+Theory (DFT) will be used to predict the freezing of a Weeks-Chandler-Anderson 
+model fluid by applying SFMT with Schmidt's proposed erf model, and our temperature
+dependent parameters to form the functional for the excess Helmholtz free energy. 
 %for generating the weight funcions, and Tarazona's tensor functional. 
-%
-%Classical Density Functional Theory (cDFT) is used to find the equilbrium 
-According to classical Density Functional Theory the equilbrium 
-Helmholtz free energy of the system 
-can be found
-by varying the number density profile $n(\vec{r})$ until the free energy 
-is minimized. 
-%The minimum free energy, and its corresponding number 
-%density profile are then the equilibrium values of the Helmholtz free energy
-%of the system at the given temperature and number density. 
-%
-%To find the Helmholtz free energy both the ideal Helmholtz free energy
-%and the excess Helmholtz free energy must be found and added together.
-%The ideal Helmholtz free energy is the free energy the fluid would have in
-%the absence of any interactomic potential energy $V(r)$ between the spheres, 
-%or atoms that make up the fluid. 
-%
-%The excess Helmholtz free energy is that 
-%additional energy beyond the ideal Helmholtz free energy which results from the presence of an interatomic potential 
-%energy $V(r)$ between the spheres.
-%
-%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%For a spatially varying number density, as would be expected for a crystal, 
-%the free energy is a functional of the number density profile $n(\vec{r})$  
-%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
-%analytically to find the resulting free energy for the system. 
-%
-SFMT makes the task of forming an expression
-%is a method used to form an expression 
-for the excess Helmholtz free energy of a particular model fluid 
-much easier since the interaction that each sphere, or atom, 
-has with every other sphere does not have to be taken explicitly into account.
-%without 
-%having to take explicitly into account the interaction each sphere, or atom, 
-%has with every other sphere which would make the task very difficult, 
-%or even prohibtive.
-%
-%Classically, fluids have been modeled by a collection of small, 
-%impenetrable spheres colliding with one another like billiard balls. 
-%In the simplest case, the spheres are treated as having only a replulsive 
-%potential. This model, called the Hard-Sphere fluid, has served as the 
-%primary reference system for fluids. This is due to the fact that the 
-%repulsion between atoms is much stronger than their attraction, and so 
-%the repulsive potential, which is descriptive of hard spheres, dominates 
-%the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
-%fluid which is a perturbation of the Hard-Sphere fluid created by adding 
-%a hard-edge attractive potential to the repulsive potential of the 
-%Hard-Sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-%spheres with softly fading replusive potentials which he modeled by making
-%the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
-%This reflects the idea that soft-spheres penetrate more deeply into 
-%each other when they collide at higher temperatures. 
-%
-%In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
-%with Rowlinson's idea of using temperature-dependent radii to model spheres 
-%with soft repulsive potentials to create the Barker-Henderson model 
-%fluid~\cite{barker1967perturbation}, where a soft-sphere fluid by is 
-%modeled by using a Hard-Sphere fluid with hard spheres of larger radii 
-%at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
-%Chandler, and Anderson used temperature and density dependent hard-sphere 
-%radii to model spheres with the more realistic softly fading repulsive 
-%portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-%which is known as the Weeks-Chandler-Anderson model fluid. f
-%
-%While models for particular fluids were being created, powerful methods 
-%to describe the properties of fluids in general were also being developed. 
-%In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
-%introduced from which an equation of state, and an expression for the 
-%excess Helmholtz free energy of a homogeneous Hard-Sphere fluid can be 
-%obtained~\cite{Hansen}. \color{red}ADD Percus-Yevick introduced an 
-%alternative method...OZ... \color{black}
-%%Percus and Yevick took another approach to the Hard-Sphere fluid.
-%In 1961 Morita and Hiroike introduced a density functional 
-%approach~\cite{MoritaDFT} to finding the equilibrium grand free energy 
-%(the equilibrium Helmholtz free energy in the case of a canonical 
-%ensemble), and the cooresponding equilibrium number density profile for 
-%inhomogeneous fluids. To use this approach, a functional must be 
-%constructed for a particular fluid.
-%%DFT Evans 1979
-%%Led by the consistency bewteen results from SPT and Percus-Yevick, 
-%In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-%(which can be constructed from SPT~\cite{santos2012phi3}) to form a 
-%functional for the Hard-Sphere fluid. Tarazona later improved on the 
-%method by introducing a tensor version of the functional~\cite{tarazonaphi3, 
-%santos2012phi3}, and in 1999 Schmidt adopted FMT for soft spheres 
-%forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}. 
-%
-%Schmidt applied his Soft Fundamental Measure Theory to various fluids 
-%with soft repulsive and soft attractive potentials, and developed an erf 
-%model for spheres with steep, short-range soft repulsive 
-%potentials~\cite{schmidt2000fluid}.
-% like that exhibited by a WCA fluid. 
-%
-Results will be compared to those from Monte-Carlo simulations.
+Results will be compared to those generated from Monte-Carlo simulations.
 
@@
 changes in the free energy will be negative for spontaneous processes 
-when the natural variables of the free energy are held fixed~\cite{schroeder}. 
+when the natural variables of the free energy are held fixed. 
 %And so the free energy tends toward a minimum. 
@@
 %particles respond to a small offset from equilibrium?
-%
-%Save for me:
-%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
-%\begin{equation}
-  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
-%\end{equation}
-%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
-%$dT=0$ gives
-%\begin{equation}
-  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
-%\end{equation}
-%where conservation of energy was also applied. Fixing the volume and the 
-%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
-%results in
-%\begin{equation}
-  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
-%\end{equation}
-%which, for fixed N, is the same as
-%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
-%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
-%of particles are fixed, the change in the free energy of the system will 
-%always be negative, and thus the free energy tends toward a 
-%minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
-%for a system near equilibrium. 
-%End save for me.
-A general proof of minimization 
+
+Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
+\begin{equation}
+  df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
+\end{equation}
+and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
+$dT=0$ gives
+\begin{equation}
+  f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
+\end{equation}
+where conservation of energy was also applied. Fixing the volume and the 
+number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
+results in
+\begin{equation}
+  df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+\end{equation}
+which, for fixed N, is the same as
+\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
+Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
+of particles are fixed, the change in the free energy of the system will 
+always be negative, and thus the free energy tends toward a 
+minimum~\cite{schroeder}. This illustrates the tendancy toward minimization 
+for a system near equilibrium. A general proof of minimization 
 can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
@@
 of a delta function characteristic of hard spheres. Soft Fundamental Measure 
-Theory extends the Fundamental Measure Theory discussed so far to 
+Theory (SFMT) extends the Fundamental Measure Theory discussed so far to 
 fluids with soft potenials by keeping the relationships between the weight 
@@
 \section{Overview}
-%
-%The goal of this project is to generate phase diagrams for a 
-%Weeks-Chandler-Anderson model fluid using a computer program. 
-%To do this we find the Helmholtz free energies that the fluid
-%would have at a given temperature and density in a  
-%liquid state, and in a crystalline solid state. 
-%%Once the Helmholtz free energy for both the crystal and liquid states of 
-%%the fluid are found for a given temperature and density, they can be 
-%%compared to find the resulting state of the fluid. 
-%Outside the region of 
-%coexistence which exists between the liquid and crystal states, 
-%the fluid will be completely in a crystalline state if the 
-%crystal free energy is lower than the liquid free energy, and it will be 
-%completely in a liquid state if the liquid free energy is lower than 
-%the crystal free energy. Maxwell's Double Tangent Construction is used 
-%to determine the actual number densities of the crystal and liquid states
-%at the boundaries of the coexistance region for a given temperature, 
-%and construct phase diagrams.
-
-%%In general, if at some number density and temperature the crystal free energy 
-%%is lower than the homogeneous free energy of the liquid, then the fluid will 
-%%be in the crystalline state. If the homogeneous liquid free energy is lower 
-%%than the crystal free energy, then the fluid will be in the liquid state. 
-%
-%%To find the equilibrium Helmholtz free energies we use classical Density 
-%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
-%%%the corresponding equilibrium number density profile. 
-%%In particular, 
-%
-%In general, the Helmholtz free energy can be written as a functional of 
-%the number density profile $n(\vec{r})$ which describes the way the average number 
-%density varies spatially. 
-%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-%Classical Density Functional Theory is used to find the equilbrium 
-%Helmholtz free energy of the system at a given temperature and density
-%by varying the number density profile $n(\vec{r})$ until the free energy 
-%is minimized. The minimum free energy, and its corresponding number 
-%density profile are then the equilibrium values for that temperature 
-%and number density. 
-%
-%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
-%and the excess Helmholtz free energy.
-%The ideal Helmholtz free energy is the free energy the fluid would have in
-%the absence of any interactomic potential energy $V(r)$ between the spheres, 
-%or atoms that make up the fluid. The excess Helmholtz free energy is that 
-%additional energy which results from the presence of an interatmoic potential 
-%energy $V(r)$ between the spheres.
-%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%%For a spatially varying number density, as would be expected for a crystal, 
-%%the free energy is a functional of the number density profile $n(\vec{r})$  
-%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
-%%analytically to find the resulting free energy for the system. 
-%Soft Fundamental Measure Theory is used to form the expression 
-%for the excess Helmholtz free energy without having to take explicitly 
-%into account the interaction each sphere, or atom, has with every other 
-%sphere.
-
-Our goal is to generate computationally temperature-density and pressure-
-temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
-classical Density Functional Theory. 
-A temperature-density phase diagram for a fluid that does not
-have attraction between the particles that make up the fluid will show
-two discinct regions - one at lower densities, and one at higher densities,
-seperated by a coexistence region in which there is found a mixture of the 
-fluid at some fixed lower density, and at some fixed higher density. An 
-abrupt change from some maximum lower density to some minimum higher density 
-occurs for certian combinations of pressure and temperature which form
-a line on a pressure-temperature phase diagram. This line distinguishes 
-the region in which the fluid exhibits lower densities from the region 
-in which the fluid exhbits higher densities. These two regions are known
-as the liquid state and the solid state of the fluid.
-
-It is known that a fluid with a soft potential like the WCA fluid will 
-form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
-The transition between the liquid 
-and solid states is abruptly inhomogeneous, and we study the behavior 
-of the fluid by treating the liquid as homogeneous, and by constructing 
-a crystal lattice in real space for the solid. Although the entire 
-crystal lattice can be constructed with FCC cells, 
+We use classical Density Functional Theory (cDFT) in conjunction 
+with Soft Fundamental Measure Theory (SFMT) to find the equilibrium free 
+energy of a system, and the corresponding equilibrium number density profile. 
+In particular, SFMT is used to form the expression for the excess free energy 
+without having to take into account the interaction each sphere, or atom, 
+has with every other sphere. The excess free energy is then added to the 
+ideal free energy calculated analytically to find the resulting free energy 
+for the system. 
+\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
+For a spatially varying number density, as would be expected for a crystal, 
+the free energy is a functional of the number density profile $n(\vec{r})$  
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+
+To implement cDFT, the number density profile $n(\vec{r})$ is varied with 
+the temperature fixed until the free energy is minimized. The minimum free 
+energy and its corresponding number density profile are then the equilibrium 
+values for that temperature. Once the Helmholtz free energy for both the crystal and liquid states of 
+the fluid are found, they can be compared to find the resulting state of 
+the fluid. Outside the region of coexistence, the fluid will be completely 
+in a crystalline state if the crystal free energy is lower than the liquid 
+free energy, and it will be completely in a liquid state if the liquid 
+free energy is lower than the crystal free energy.
+
+***** DFT uses n(r) for BOTH Fideal and Fexcess ******
+
+To implement cDFT computationally, we begin by picturing one single crystal 
+cell. Although the crystal structure is composed of Face Centered Cubic (FCC) 
+cells, the entire lattice can be generated with a cell one-fourth the size 
+called a primitive cell shown by the shaded region in Figure \ref{fig:primitivecell}. 
+The primitive cell has the shape of a parallelpiped, and contains the 
+equivalent of one atom.
+
+%move this paragraph to methods?
+-----------*****-------------
+
+Although the entire crystal lattice can be constructed with FCC cells, 
 FCC cells are not the smallest cells with which the lattice can be built. 
-The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
+The shaded parallepiped shown within a FCC cell in Figure \ref{fig:primitivecell} 
 is the smallest structure that, when repeated, re-creates the entire crystal 
-lattice. This parallepiped primitive cell takes up one-fourth the volume 
-of an FCC cell, and contains the equivalent of one atom.
-It includes portions of eight atoms each located on a lattice site convenietly 
-reachable by moving along one of three different lattice vectors to step 
-from one atom to the next. The three lattice vectors will be useful in 
-creating the lattice structure computationally, and will also make the 
-program run faster.
+lattice. The parallepiped primitive cell takes up one-fourth the volume 
+of an FCC cell, includes portions of eight atoms for a total of one full 
+atom, and has eight lattice sites convenietly reachable by moving along 
+one of three different lattice vectors to step from one atom to the next. 
+The three lattice vectors will be useful in creating the lattice structure 
+computationally, and will also make the program run faster.
+
+----------****--------------
+%-------------------------------
 
@@
 
-  %\begin{figure}[h!]
-    %\centering
-    %% FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
-    %%\caption{A Simplified Crystal Cell}
-    %\label{fig:simplecell}
-  %\end{figure}
-
-%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
-%consider instead a simplified crystal cell with one atom in its center 
-%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
-%minimized for a fixed volume (and temperature and number of particles).  
-%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
-%be used for illustration. The number density n for a collection of 
-%simplified cells is simply 1 atom per fixed bold-box volume.
-%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
-%It is easier to think of 1 atom per ``box" (a reduced number density) 
-%than 1 atom per some length$^3$ (the actual number density). Reduced number 
-%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
-%of size $\sigma^3$.
-%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
-%has a volume of $\sigma^3$, the reduced number density is just the number 
-%of atoms in the fixed bold-box volume. 
-%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
-%}
-In a large collection of cells, the atom at each lattice site of 
+  \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
+    \caption{A Simplified Crystal Cell}
+    \label{fig:simplecell}
+  \end{figure}
+
+This cell is rather complicated for the purposes of this discussion, so 
+consider instead a simplified crystal cell with one atom in its center 
+as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
+minimized for a fixed volume (and temperature and number of particles).  
+A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
+be used for illustration. The number density n for a collection of 
+simplified cells is simply 1 atom per fixed bold-box volume.
+\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
+It is easier to think of 1 atom per ``box" (a reduced number density) 
+than 1 atom per some length$^3$ (the actual number density). Reduced number 
+density is given by $n^*$ and is equal to the number of atoms in a ``box" 
+of size $\sigma^3$.
+\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
+has a volume of $\sigma^3$, the reduced number density is just the number 
+of atoms in the fixed bold-box volume. 
+\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath}
+
+In a large collection of simple cells, the atom at the center of each cell 
 will not realistically reside exactly at the lattice point in the center 
@@
 Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
-distribution is shown in Figure \ref{fig:Gaus_plot}. 
-A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
+distribution is shown in Figure \ref{fig:Gaus_plot}. A 3-dimenstional 
+Gaussian distribution centered at a lattice point can be used to represent 
 a number density profile $n(\vec{r})$  which describes how the atoms are 
 statistically distrubuted over a collection of cells.
-%In 3-dimensions, the Gaussian function takes on the form
-%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-It is important to note that the number density does not change when introducing a number 
-density profile. There are still the same number of atoms over some fixed volume. 
-%There is still a total of one atom per ``bold box" volume.
-%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
-\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
+
   \begin{figure}[h!]
@@
 
-It may be that not all lattice sites are occupied. Vacancies can be taken 
-into account while keeping %the volume of the bold box, and thus 
-the overall number density the same. This is done by 
+It is important to note that the number density does not change when 
+introducing a number density profile. There is still a total of one atom 
+per ``bold box" volume.
+
+\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
+
+In a large collection of cells, not all lattice sites will be occupied. 
+Vacancies can be taken into account while keeping the volume of the 
+bold box, and thus the overall number density, the same. This is done by 
 making the cells smaller. A large collection of smaller cells creates a 
@@
 with eight smaller Gaussians of one-eigth the height.
+
+ \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
+    \caption{A Gaussian Distribution Number Density Profile}
+    \label{fig:Ensemble_Smallcells}
+  \end{figure}
+
 Instead of thinking of an atom in one of the eight small cells, the same 
@@
 atom in each small cell. This makes it convenient to describe the new 
-number density in terms of the fraction of vacancies, $f_v$: 
-\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath} 
-%\begin{align}
-  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
-%\end{align}
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{A Gaussian Distribution Number Density Profile}
-    \label{fig:Ensemble_Smallcells}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=2.5cm]{SameStatPic.png}
+reduced number density in terms of the fraction of vacancies, $f_v$:
+
+\begin{displaymath}{ n^* = (1-f_v){n_{no.vacancies}^*}}\end{displaymath}
+
+\begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=2.5cm]{SameStatPic.png}
     \caption{Front views of a cube with eight smaller cells. The two views 
@@
 
-By varying the width of the Gaussians and the fraction of vacancies, various
-number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
-Once again, it is important to note that the number density is the same 
-for all these number density profiles $n(\vec{r})$. 
-%There is still a total of one atom per ``bold box" volume. 
-%This prodcedure is used only to find the crystal free energy as the number density 
-%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
-A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.
+By varying the width of the Gaussians and the fraction of vacancies, new 
+number density profiles can be generated as shown in Figure \ref{fig:Ensemble_vary}.
 
@@
     \caption{Several Gaussian Distribution Number Density Profiles all 
-    with the same number density, that is, there is one atom over the fixed volume.}
+    with a number density of 1 atom per $V_{bold box}$}
     \label{fig:Ensemble_vary}
@@
 
+Once again, it is important to note that the number density is the same 
+for all these number density profiles $n(\vec{r})$. There is still a total 
+of one atom per ``bold box" volume. This prodcedure is used only to find 
+the crystal free energy as number density profile for a liquid is 
+homogeneous, as shown in Figure \ref{fig:homogen_denisty}.
+
  \begin{figure}[h!]
@@
     \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
+    % FIXME \includegraphics[height=4cm]{Homogeneous_bold-box.pdf}
     \caption{A homogeneouse number density profile $n(\vec{r})$}
@@
 
-%The number density profile is varied by varying gw and fv, and the 
-For a given combination of number density and temperature, a suite of number 
-density profiles are generated.
-% like those represented by Figure~\ref{fig:Ensemble_vary}.
-The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number density profile,
-and also for the homogeneous number denisty profile. 
-Both the ideal and excess free energies are found, and added together to get the total Helmholtz 
-free energy.
-The Helmholtz free energy is minimized for a system with a fixed temperature, 
-volume, and number of particles. Setting the number of particles to one, 
-the volume can be found from the number density.
-The process is then repeated to find the free energies at other combinations 
-of number densities and temperatures.
-
-%The number of atoms in one primitive cell is one.
-
+A simplified flowchart of the program created to investigate the freezing 
+of a WCA fluid is shown below. The program takes four inputs: reduced number 
+density n*, reduced temperature t*, a range of Gaussian 
+widths gw, and a range of fraction of vacancies fv. The number density 
+profile is varied by varying gw and fv, and the free energy $F[n(\vec{r})]$ 
+is calculated for each profile until the minimum free energy is found for 
+that combination of $n^*$ and $t^*$, which is known from cDFT to be unique. 
+The process is then repeated to find the free energies at other number 
+densities and temperatures.
+
+%Program Implementation
+The program computes both the crystal and the homogeneous Helmholtz free 
+energies. The Helmholtz free energy is minimized for a system with a fixed 
+temperature, volume, and number density. The temperature, T is an input to 
+the program. The number density, which is also an input to the program, 
+is used to calculate the volume of one primitive cell over which the free 
+energy is computed. The number of atoms in one primitive cell is one. To 
+compute the free energy, it is necessary to compute both the ideal and 
+excess free energies and add them, $F = F_{ideal} + F_{excess}$.
 %\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
@@
 
-%PROGRAM FLOWCHART ---------------------------
-%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
-%Loop: 
-%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
-%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
-%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
-%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
-%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
-%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
-
+%\noindent Input:
+\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+%\noindent Set:
+%\begin{displaymath}\text{N = 1}\end{displaymath}
+\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
+%Outer Loop:  %set fv to one of the input values
+Loop: 
+\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
+\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
+\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
+%\indent\indent\indent Inner Loop : %set gw to one of the input values
+\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
+\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
+\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
+
+%Compare:
+\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+%\noindent Output:
+\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
 %\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
@@
 %lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
-
-\section{Computing the Liquid and Crystal Helmholtz free energies}
-%-----------Comuting the Liquid and Crystal free energies ------
-
-The liquid ideal Helmholtz free energy is computed analytically 
-using Equation~\ref{eq:Fideal} 
-for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
-%from the equation 
+\newline
+\newline
+*****   start stuff  ******
+
+In general, if the crystal free energy at some $n^*$ and $t^*$ is lower 
+than the homogeneous free energy of a liquid at that same $n^*$ and $t^*$, 
+then the crystaline state dominates. But if the crystal free energy is 
+higher than the homogeneous free energy, the liquid state dominates. 
+However, there is a region of densities between the liquid and crystal 
+states that are forbidden, and so the actual state can not be determined 
+just by which energy is lower at that particular $n^*$ and $t^*$. Maxwell's 
+Double Tangent Construction must be implemented using the date collected 
+to determine the actual crystal and liquid states, and construct phase 
+diagrams.
+
+*****  end  stuff  ******
+
+\section{Computing the Liquid and Crystal Ideal Helmholtz free energies}
+
+The liquid ideal Helmholtz free energy $F_{Hideal}$ is computed 
+analytically for a given homogeneous reduced density $n^*$, and reduced 
+temperature $t^*$ from the equation 
 %\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-%\begin{align}
-  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
-%\end{align}
-%The quantity $\Lambda$ is the DeBroglie wavelength given by
-%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
-%where in our program we have set
-%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
+\begin{align}
+  \frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
+\end{align}
+where $\sigma=1$ (which gives $n^*=n\sigma^3$=n), and $\epsilon=1$ (which 
+gives $t^*=\frac{k_BT}{\epsilon}=k_BT$). The quantity $\Lambda$ is the DeBroglie 
+wavelength given by
+\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
+where in our program we have set
+\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
 %\begin{displaymath}{m=}\end{displaymath}  
@@
 %\begin{displaymath}{k_B=}\end{displaymath}
-The crystal ideal Helmholtz free energy is computed computationally 
-from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
-%\begin{align}
-  %F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
-%\end{align} 
+
+The crystal ideal Helmholtz free energy $F_{Cideal}$ is computed from the equation 
 \begin{align}
-  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}}
+  F_{Cideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
 \end{align}
-%For small gw values it is assumed that the Guassians do not overlap, 
-%in which case the equation for the free energy can be simplified, and solved
-%analytically with the result
-%\begin{equation}
- %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
-%\end{equation}
-%For larger gw values it is assumed that the Gaussians do overlap, 
-%and so the free energy is caluculated computationally. First dx is scaled by gw,
-%\begin{equation}
-  %dx_{scaled} = (dx_{input})(gw)
-%\end{equation}
-%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
-%primitive cell taking into account the contribution of Gaussians from 
-%many (5x5x5) nearby cells whose centers are given by $R_i$.
-%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
-%The contribution to the free energy of $n(\vec{r})$ at the one position is then
-%added to the accumulating sum.
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
-%This process is repeated for all positions within one primitive cell to get the
-%crystal ideal free energy for the primitive cell. This energy is then divided by
-%the volume of the primitive cell to get the crystal ideal free energy per
-%volume.
-%
-The liquid excess Helmholtz free energy is computed analytically  
-%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
-using SFMT with a uniform number density profile, n that does not depend 
-on $\vec r$. 
-%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
-%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
-%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
-%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
-%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
-%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
-%The weighted densities are found from
-%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
-%where the weights $w_i$ are given by Equations~\ref{eq:weights}.
-
-The crystal excess Helmholtz free energy is computed computationally 
-using SFMT with a number density profile $n(\vec r)$ that varies spacially.
-%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
-%Then the crystal, or inhomogeneous, excess free energy is found from
-%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
-Monte-Carlo integration is used to compute the weighted densities 
-$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
-%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
-%
-The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
-within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
-the Gaussians is selected that is within the primitive cell, or one of the 
-neighboring cells within a prudently picked inclusion radius as Gaussians 
-farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=5cm]{InclusionRadius.png}
-    \caption{The Gaussians at each lattice point are of width gw. 
-    After a radius of few gw's the Gaussians go to zero and can be neglected.}
-    \label{fig:InclusionRadius}
-  \end{figure}
-
-The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
-$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
-$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
-$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
-where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
-some point in space. Monte-Carlo integration works by randomly generating 
-$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
-\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
-is the gaussian number density profile for no vacancies, the equation:
-
-\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
-implemented computationally in the form:
-\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
-becomes:
-\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
-(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
-which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
-$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
-Using the relation:
-\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
-$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
-The result can be made more accurate by increasing NumPoints which is the 
-number of vectors randomly generated. 
-%-------Antithetic-------------------
-%In order to further reduce the error 
-%in the average, antithetic variates are used. Rather than computing an average 
-%of a set of random points, each random point in the set is paired with a 
-%point along the same line as the first point, but opposite in direction, 
-%as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
-%error in the average by filtering out an erroneous offset of values along 
-%the line connecting the two points by averaging the two results, one of 
-%which may be erroneously higher than it should be while the other is 
-%erronesously lower than it should be.
-
- %\begin{figure}[h!]
-    %\centering
-    %\includegraphics[height=4cm]{AntitheticVariate.png}
-    %\caption{Antithetic variates reduce first order error in an average. 
-    %Each point in the average is paired with another point from along 
-    %the same line as the first, and at an equal, but opposite distance.}
-    %\label{fig:AntitheticVariate}
-  %\end{figure} 
-%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
-%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
-%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
-%The standard error of the mean (SEM) is given by:
-%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
-%a measure of the error in $n_3=\overline{w}_3$:
-%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
-%NumPoints, the number of vectors randomly generated, is increased until:
-%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
-%value that $n_3$ can be without error is 1.
-
-%Figure \ref{fig:GaussandW2_actual} ......
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{GaussandW2_actual.png}
-    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
-    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
-    and fixed in place as $\vec{dr}$ varies randomly.} 
-  \label{fig:GaussandW2_thinkas}
-  \end{figure}
-
-%Monte-Carlo integration is employed to more efficiently 
-%compute the weighted densities used to calculate the excess free energy which is then 
-%added to the ideal free energy (computed analytically) to get the total free energy 
-%which is minimized.
+with the varying number density profile given by 
+\begin{displaymath}
+  {n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}}
+\end{displaymath}
+For small gw values it is assumed that the Guassians do not overlap, 
+in which case the equation for the free energy can be simplified, and solved
+analytically with the result
+\begin{equation}
+ F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
+\end{equation}
+For larger gw values it is assumed that the Gaussians do overlap, 
+and so the free energy is caluculated computationally. First dx is scaled by gw,
+\begin{equation}
+  dx_{scaled} = (dx_{input})(gw)
+\end{equation}
+and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
+primitive cell taking into account the contribution of Gaussians from 
+many (5x5x5) nearby cells whose centers are given by $R_i$.
+%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
+The contribution to the free energy of $n(\vec{r})$ at the one position is then
+added to the accumulating sum.
+%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
+%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
+This process is repeated for all positions within one primitive cell to get the
+crystal ideal free energy for the primitive cell. This energy is then divided by
+the volume of the primitive cell to get the crystal ideal free energy per
+volume.
 
 \section{Our Functional for the Excess Helmholtz free energy}
-In order to apply classical Density Functional Theory, the Helmholtz energy
-must be found. Computing the ideal portion of the Helmholtz free energy is
-straightforward, but computing the excess portion of the Helmholtz free energy 
-is difficult since it depends on the interatomic potential energy $V(r)$ between 
-each sphere with all the other spheres. Soft Fundamental Measure Theory is a method
-developed to form an expression for the excess Helmholtz free energy without
-explicitly taking into account the interaction of each sphere with all the others, 
-but it must be taylored to fit a particular fluid. 
+
 %The excess Helmholtz free energy is particular to the fluid being modeled, 
 %as it depends on the interatomic potential V(r). 
-
-We use Soft Fundamental Measure Theory to form a functional for the excess 
-Helmholtz free energy appropriate for the WCA fluid by introducing our 
-temperature dependent parameters $\alpha(T)$ and $\Xi(T)$. The functional 
-is written in general form as 
+We use Soft Fundamental Measure Theory to find the excess Helmholtz free energy.
+The functional is written in general form as 
 \begin{align}
@@
 and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
-for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
+for $\Phi_3$. The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
 \begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-where the number density profile $n(\vec r)$ within the integral is given 
-by a Gaussian function centered at a lattice point,
-%\begin{align}
-  %n(\vec r)=(1-f_v)\frac{\exp^{-\frac{r^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}
-%\end{align}
-and the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
+where the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
 \begin{align}\label{eq:weights}
@@
 \end{align}
-%
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_actual.png}
-    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    %view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    %is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
-Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off. 
-$\Xi$ is related to the width of the functions which 
-become wider in general as the temperature increases. 
-The radial distance 
-to the peak is related to $\alpha$, and decreases as the temperature 
-increases, that is, the shells shrink, as does 
-the sphere.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{weight_functions.pdf}
-    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2.}
-    \label{fig:weight_functions}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W2.png}
-      % \includegraphics[width=5cm]{figs/w2-ring}
-      \caption{A crossectional view of $w_2$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W3.png}
-      \caption{A crossectional view of $w_3$}
-    \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    \label{fig:W0andW3}
-    \end{figure}
-
 The weight fucntions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
@@
 
-%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
-%The temperature dependence of parameters $\alpha$ and $\Xi$ 
-%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
-%can be related 
-%to the weight functions. 
-%$\Xi$ is related to the width of the functions which 
-%become wider in general as the temperature increases. 
-%The radial distance 
-%to the peak is related to $\alpha$, and decreases as the temperature 
-%increases, that is, the shells shrink, as does 
-%the sphere.
-%The functions can be treated as neglible outside of a distance of 
-%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
-%the weight function.
-
- %\begin{figure}[h!]
-    %\centering
-    %\includegraphics[height=7cm]{weight_functions.pdf}
-    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    %z-axis for a reduced temperature of 2.}
-    %\label{fig:weight_functions}
-  %\end{figure}
-
- %\begin{figure}[h!]
-    %\centering
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W2.png}
-      %% \includegraphics[width=5cm]{figs/w2-ring}
-      %\caption{A crossectional view of $w_2$}
-    %\end{subfigure} 
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W3.png}
-      %\caption{A crossectional view of $w_3$}
-    %\end{subfigure} 
-    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    %\label{fig:W0andW3}
-    %\end{figure}
+\color{red}
+Figure \ref{fig:weight_functions} shows a 1 dimensional view of the weight functions 
+$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+sphere up to a radius of ?FIX? after which it tapers off over a distance of ?FIX?.
+
+The parameters $\alpha$ and $\Xi$ are temperature dependent, as shown earlier in 
+Figure \ref{fig:alphaXivsT}. $\Xi$ is related to the width of the functions which 
+become wider in general as the temperature increases. The radial distance to the 
+peak decreases as the temperature increases, that is, the shells shrink, as does 
+the sphere. The functions can be treated as neglible outside of a distance of 
+about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
+the weight function.
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=6cm]{weight_functions.pdf}
+    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2.}
+    \label{fig:weight_functions}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W2.png}
+      % \includegraphics[width=5cm]{figs/w2-ring}
+      \caption{A crossectional view of $w_2$}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W3.png}
+      \caption{A crossectional view of $w_3$}
+    \end{subfigure} 
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    \label{fig:W0andW3}
+    \end{figure} 
+\color{black}
+
+\section{Computing the Liquid and Crystal Excess Helmholtz free energies}
+
+The liquid excess free energy $F_{Hexcess}$ is computed analytically for 
+a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
+using SFMT with a uniform number density profile, n that does not depend 
+on $\vec r$. 
+%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
+\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
+For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
+%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
+\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
+\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
+The weighted densities are found from
+\begin{displaymath}{n_i=nw_i}\end{displaymath} 
+where the weights $w_i$ are given by Equations~\ref{eq:weights}.
+
+The crystal excess free energy $F_{Cexcess}$ is computed computationally 
+using SFMT with a number density profile $n(\vec r)$ that varies spacially.
+%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
+%Then the crystal, or inhomogeneous, excess free energy is found from
+\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
+Monte-Carlo integration is used to compute the weighted densities 
+$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
+%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
+
+The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
+within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
+the Gaussians is selected that is within the primitive cell, or one of the 
+neighboring cells within a prudently picked inclusion radius as Gaussians 
+farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=5cm]{InclusionRadius.png}
+    \caption{The Gaussians at each lattice point are of width gw. 
+    After a radius of few gw's the Gaussians go to zero and can be neglected.}
+    \label{fig:InclusionRadius}
+  \end{figure}
+
+The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
+$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
+$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
+$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
+where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
+some point in space. Monte-Carlo integration works by randomly generating 
+$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
+\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
+is the gaussian number density profile for no vacancies, the equation:
+
+\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+implemented computationally in the form:
+\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+becomes:
+\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
+where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
+(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
+which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
+$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
+Using the relation:
+\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
+$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
+\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
+
+The result can be made more accurate by increasing NumPoints which is the 
+number of vectors randomly generated. In order to further reduce the error 
+in the average, antithetic variates are used. Rather than computing an average 
+of a set of random points, each random point in the set is paired with a 
+point along the same line as the first point, but opposite in direction, 
+as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
+error in the average by filtering out an erroneous offset of values along 
+the line connecting the two points by averaging the two results, one of 
+which may be erroneously higher than it should be while the other is 
+erronesously lower than it should be.
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=4cm]{AntitheticVariate.png}
+    \caption{Antithetic variates reduce first order error in an average. 
+    Each point in the average is paired with another point from along 
+    the same line as the first, and at an equal, but opposite distance.}
+    \label{fig:AntitheticVariate}
+  \end{figure} 
+Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
+\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}   The process is repeated for all the other $\vec{R}_j$'s to get the total weighted densities at vector $\vec{r}_k$ due to all the Gaussians.
+\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
+Then the entire process repeated for each vector $\vec{r}_k$ to get the 
+total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
+
+The standard error of the mean (SEM) is given by:
+\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} which gives 
+a measure of the error in $n_3=\overline{w}_3$:
+\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{NumPoints}}}\end{equation} 
+NumPoints, the number of vectors randomly generated, is increased until:
+\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation} where the maximum 
+value that $n_3$ can be without error is 1.
+
+Figure \ref{fig:GaussandW2_actual} ......
+
+ \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=6cm]{GaussandW2_actual.png}
+    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    is generated randomly according to a Gaussian distribution.} 
+  \label{fig:GaussandW2_actual}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    % FIXME \includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    \caption{An alternative way of viewing Figure \ref{fig:GaussandW2_actual} where 
+    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
+    and fixed in place as $\vec{dr}$ varies randomly.} 
+  \label{fig:GaussandW2_thinkas}
+  \end{figure}
+
+%Monte-Carlo integration is employed to more efficiently 
+%compute the weighted densities used to calculate the excess free energy which is then 
+%added to the ideal free energy (computed analytically) to get the total free energy 
+%which is minimized.
 
@@
 Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
-and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
+through \ref{fig:Phase_Diagram_P_vs_T_AS}. In all cases $\sigma=1$ and $\epsilon=1$. 
 In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
@@
 
-%Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
-%range of reduced temperatures from kT=0.5 to 38. The white lines show the results
-%of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
-%As can be seen in the figure, the region of coexistence in our simulations 
-%falls within the region of coexistance predicted by Ahmed and Sadus at almost 
-%all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
-%the width of the coexistance regions are nearly the same, while moving to 
-%higher temperatures,the width of our coexistance region ranges from about 
-%one-half to one-forth the size. Although, our coexistence region is only 
-%about one-forth the size at kT=28, it is centered within the coexistence 
-%region predicted by Ahmed and Sadus.
+Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
+range of reduced temperatures from kT=0.5 to 38. The white lines show the results
+of molecular simulations of the WCA fluid made by Ahmed and Sadus~\cite{ahmedsadus}.
+As can be seen in the figure, the region of coexistence in our simulations 
+falls within the region of coexistance predicted by Ahmed and Sadus at almost 
+all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
+the width of the coexistance regions are nearly the same, while moving to 
+higher temperatures,the width of our coexistance region ranges from about 
+one-half to one-forth the size. Although, our coexistence region is only 
+about one-forth the size at kT=28, it is centered within the coexistence 
+region predicted by Ahmed and Sadus.
 
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
+  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
   %\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
-  %\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
-  %White lines show the liquid and solid transition densities at each temperature from 
-  %simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
+  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  White lines show the liquid and solid transition densities at each temperature from 
+  simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
   \label{fig:Phase_Diagram_of_T_vs_n_AS}
@@
 than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
-the temperature increases. 
-%Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
-%range of reduced temperatures from kT=0.5 to 38, compated with the results
-%from Ahmed and Sadus's molecular simulations. Here again, our transition 
-%pressures and temperatures are consistently higher.
+the temperature increases. Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
+range of reduced temperatures from kT=0.5 to 38, compated with the results
+from Ahmed and Sadus's molecular simulations. Here again, our transition 
+pressures and temperatures are consistently higher.
 
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
+  \includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T_AS}
   %\includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T_AS}
-  %\caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
-  %The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
+  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
+  The white line marks the boundary from simulations by Ahmed and Sadus~\cite{ahmedsadus}.}
   \label{fig:Phase_Diagram_P_vs_T_AS}
@@
 
-%Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
-%reduced pressure versus the reduced temperature at fixed reduced densities 
-%from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
-%\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
-%the WCA fluid transistions into a solid. These downward bends align to 
-%form the curved interface separating the blue and red areas in 
-%Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
+Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
+reduced pressure versus the reduced temperature at fixed reduced densities 
+from $n^*$ = 0.7 to 1.2. As can be seen by comparison with Figure 
+\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
+the WCA fluid transistions into a solid. These downward bends align to 
+form the curved interface separating the blue and red areas in 
+Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
 
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
-  %\caption{Plot of the reduced pressure verses the reduced temperature. 
-  %The lines bending downward indicate freezing.}
+  \includegraphics[width=\columnwidth]{figs/p-vs-T_at_fixed_n_noMC}
+  \caption{Plot of the reduced pressure verses the reduced temperature. 
+  The lines bending downward indicate freezing.}
   \label{fig:p-vs-T_at_fixed_density}
@@
 phase diagram that yielded coexistence regions that approximately 
-matched the Monte-Carlo
-% and molecular simulation 
-results. 
-%for values of kT=0.5 to 3. 
-%kT=0.5 to 38
-We are also able to generate a pressure-temperature
+matched the Monte-Carlo and molecular simulation results for 
+values of kT=0.5 to 38. We are also able to generate a pressure-temperature
 phase diagram with a shape that is expected, but the transition 

### 2020-11-03 papers/thesis-kirstie/thesis.tex
some comments and fixes in section 3.3

--- 2020-10-31 papers/thesis-kirstie/thesis.tex
@@
 \commencementyear{2021}
+
+\newcommand\davidsays[1]{\textcolor{red}{[\it D: #1]}}
 
@@
 and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
-for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
+for $\Phi_3$. The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
 \begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
@@
 centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, with a width determined 
 by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
@@
 
-The weight fucntions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
+\davidsays{Maybe say this before presenting the weights? But also, I don't think
+there is any benefit to introducing $a$ and $\sigma$ here.} The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed
+using 
 Schmidt's proposed erf model~\cite{schmidt2000fluid} in which he set
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
+\begin{equation}
+  w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]
+\end{equation} 
 where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
@@
 Our functional is  created by replacing Schmidt's parameters $\sigma$ 
-and $a$ with our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$.
+and $a$ with temperature-dependent parameters $\alpha(T)$ and $\Xi(T)$.
 \begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
@@
 temperature dependence that enters in through the expression for the 
-error function potential $V_{erf}(r)$ which is used to approximate the
+error function potential $V_{\operatorname{erf}}(r)$ which is used to approximate the
 non-temperature dependent WCA potential that is more difficult to 
 implement computationally. The erf potential emerges from the relationships
-that the weight functions must satisfy.
+that the weight functions must satisfy\davidsays{In order for what?}.
 %potential at temperatures of interest will be used instead of the WCA potential. 
 Using Equation~\ref{w2_w3_relation}, shown again below, 
-\begin{displaymath}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{displaymath}
+\begin{align}
+  w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+\end{align}
 and solving for $w_2(r)$
-\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
+\begin{align}
+  \frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)
+\end{align} 
 we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
@@
 \begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
-The Mayer function for the error function potential $V_{erf}(r)$ is thus 
+The Mayer function for the error function potential $V_{\operatorname{erf}}(r)$ is thus 
 given by
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
+\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{\operatorname{erf}}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
 Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} 
-%This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
-gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
+%This gives an interatomic potential of \begin{equation}{V_{\operatorname{erf}}=-k_BT\ln(f(r)-1)}\end{equation} 
+gives an expression for the for the error function potential  $V(r)=V_{\operatorname{erf}}(r)$ 
 \begin{equation}
-	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
+	V_{\operatorname{erf}}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
 \end{equation}
@@
 The parameter $\alpha(T)$ is derived by setting 
-the erf potential equal to the WCA potential when r equals $\alpha$ 
-\begin{equation}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
-and is given by 
+the erf potential equal to the WCA potential when $r=\alpha$, i.e. at the
+maximum of $w_2(r)$,
+% \begin{equation}{V_{\operatorname{erf}}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
+which gives us
 \begin{equation}\label{alphaT}
@@
   \end{figure}
-The value of $\Xi$ at a given temperature, T, can be found computationally 
+We determine value of $\Xi$ at a given temperature $T$,
 by equating the second virial coefficient evaluated for the error function 
-potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
+potential $V_{\operatorname{erf}}(r)$, and the second virial coefficent evalued for the 
 WCA potential $V_{WCA}$.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+\begin{equation}B_{2\operatorname{erf}}(T) =B_{2WCA}(T)\end{equation}
+\davidsays{I think we'd benefit from cutting much of the following, since you
+show the standard equations, but skip the bit where we put in the pair potential,
+which is the tricky step.}
 The second virial coefficient can be derived from Euler's equation for 
@@
 \begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
-\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
-\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
+\begin{equation}U=\langle E\rangle=\sum_{i=0}^\infty{P_iE_i}\end{equation}
+\begin{equation}N=\langle n\rangle=\sum_{i=0}^\infty{P_in_i}\end{equation}
 which, for a canonical ensemble, gives
@@
 \begin{equation}
-	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
+	B_{2\operatorname{erf}}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{\operatorname{erf}}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
 \end{equation}
@@
 
+\davidsays{Some of this is worth showing, but I think you're giving too many
+plots here.
+In general, we'd like to only show plots if we have something to say about them,
+so either we want more words, or fewer figures.  e.g. what does the reader learn
+from looking at your plot of $B_2$ versus $T$? If there is something they should
+learn from it, then you should write that down, so they don't have to figure it
+out on their own.  If there isn't anything to say, then probably you can skip
+it. There \emph{are} times to just show a plot, but it's usually if it's an
+prediction, and you want to say \emph{something} about the plot beyond its
+existence.}
 A plot of the resulting second virial coefficient $B_2(T)$  with respect 
@@
     \caption{Shows the temperature dependence of $B_2$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    matching the second virial coeficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
     at each temperature.}
@@
     \caption{Shows the temperature dependence of $\Xi$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    matching the second virial coeficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
     at each temperature.}
@@
     \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
+    \caption{Error Function Potential $V_{\operatorname{erf}}(r)$ at several temperatures 
     with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
@@
 
-\section{Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{erf}$}
+\section{Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{\operatorname{erf}}$}
 \begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}
-The mayer function for the Error Function potential $V_{erf}$ is
+The mayer function for the Error Function potential $V_{\operatorname{erf}}$ is
 \begin{equation}f(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation} from which it can be seen that $f(r)$ is approximately 0 for $r>r_{max}$ where $r_{max}$ is some maximum value of $r$.
@@
 
-The Mayer function for the Error Function potential $V_{erf}$, and its Fourier Transform are
-\begin{equation}f_{erf}(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation}
-
-\begin{equation}\widetilde{f}_{erf}(k)=-\frac{4\pi}{k^3}e^{-\frac{k^2\Xi^2}{4}}\left[\left(1+\frac{k^2\Xi^2}{2}\right)\sin(k\alpha)-k\alpha\cos(k\alpha)\right]\end{equation}
+The Mayer function for the Error Function potential $V_{\operatorname{erf}}$, and its Fourier Transform are
+\begin{equation}f_{\operatorname{erf}}(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation}
+
+\begin{equation}\widetilde{f}_{\operatorname{erf}}(k)=-\frac{4\pi}{k^3}e^{-\frac{k^2\Xi^2}{4}}\left[\left(1+\frac{k^2\Xi^2}{2}\right)\sin(k\alpha)-k\alpha\cos(k\alpha)\right]\end{equation}
 

### 2020-11-07 papers/thesis-kirstie/thesis.tex
edits

--- 2020-10-31 papers/thesis-kirstie/thesis.tex
@@
 
-The goal of this project is to generate computationally temperature-density 
+The goal of this project is to generate temperature-density 
 and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
 model fluid using classical Density Functional Theory (cDFT). 
-To do this we need to find the Helmholtz free energies that the fluid
-would have in a liquid state, and in a crystalline solid state 
-at a given temperature and density. 
+To do this, the Helmholtz free energies of the fluid
+in a liquid state and in a crystalline solid state must be found.
 In general, the Helmholtz free energy, which consists of an ideal part 
@@
 
-Classical Density Functional Theory is used to find the equilibrium 
-Helmholtz free energy, and Soft Fundamental Measure Theory (SFMT) will 
-be used to form the functional for the excess Helmholtz free energy. 
-According to classical Density Functional Theory the equilbrium 
+Soft Fundamental Measure Theory (SFMT) will be used to form the 
+functional for the excess Helmholtz free energy. 
+SFMT is a method that makes the task of 
+forming an expression for the excess Helmholtz free energy of a particular 
+model fluid much easier since the interaction that each sphere, or atom, 
+has with every other sphere does not have to be taken explicitly into account.
+Once the functional for the Helmholtz free energy is formed, classical 
+Density Functional Theory is used to find the equilibrium 
+Helmholtz free energy at a particular temperature and density. 
+According to cDFT the equilbrium 
 Helmholtz free energy of the system can be found by varying the number 
 density profile $n(\vec{r})$ until the free energy is minimized. 
-SFMT makes the task of forming an expression
-for the excess Helmholtz free energy of a particular model fluid 
-much easier since the interaction that each sphere, or atom, 
-has with every other sphere does not have to be taken explicitly into account.
 Results will be compared to those from Monte-Carlo simulations.
@@
 into account. 
-
 The problem of understanding the behavoir of real liquids is better addressed 
@@
     \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity 
-    for $r\leq{2R}$ which is approximately 1.225 in this example.}
+    for $r\leq{2R}$ where R is the radius of a hard sphere (which is 
+    approximately 1.225/2 in this example).}
     \label{fig:HardSphere_potential}
@@
 \begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}
-As seen in Figure \ref{fig:WCA_potential}, the WCA potential decreases 
+As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
 from infinity as the distance between the centers of the atoms r increases, 
@@
 r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
-where R is the radius of the sphere as shown in Figure \ref{fig:TwoSpheres}. 
+where R is the radius of the sphere.
+
 As with the Lenard-Jones potential, the repulsive WCA potential increases 
@@
 their interatomic potential, they are called ``soft spheres'' and their 
-gradually fading potential makes them slightly ``squishy''.
+gradually fading potential makes them slightly ``squishy'', as illustrated in
+Figure~\ref{fig:TwoSpheres}.
 
@@
 with respect to volume becomes the partial derivative of the Helmholtz 
-free energy per atom with respect to the inverse number density. 
+free energy per atom, $f$,  with respect to the inverse number density. 
 \begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.
@@
 Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
-and one for the solid at a fixed temperature. Since the pressure of the 
+and one for the solid at a fixed temperature. The slope of the curves is 
+related to the pressure. Since the pressure of the 
 liquid is the same as the pressure of the solid at the point of transistion, 
@@
 %\begin{displaymath}{F=-PV+G}\end{displaymath} 
-\begin{displaymath}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{displaymath} 
-\begin{displaymath}{f=-P\frac{1}{n}+g}\end{displaymath}
+\begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
+\begin{align}{f=-P\frac{1}{n}+g}\end{align}
 This equation has the form of a line:
-\begin{displaymath}{y=mx+y_o}\end{displaymath}
+\begin{align}{y=mx+y_o}\end{align}
 %\begin{displaymath}{G_1=G_2}\end{displaymath} 
@@
 %\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
-and describes the red line shown in Figure \ref{fig:MaxwellDT} with the 
+and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
 pressure given by the negative of the slope of the line, and the Gibb's 
-free energy per atom given by the y-intercept. At the point of phase transition:
-\begin{displaymath}{g_L=g_S}\end{displaymath} 
-\begin{displaymath}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{displaymath}
+free energy per atom given by the y-intercept. 
+At the point of phase transition the chemical potential of the liquid is 
+the same as the chemical potential of the solid.
+\begin{align}{g_L=g_S}\end{align} 
+\begin{align}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{align}
 
@@
 against the pressure for both the liquid and the solid at fixed temperature, 
-as shown in Figure \ref{fig:GibbsvsP}. These curves are constructed from 
+as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
 the slope and cooresponding y-intercept data collected from lines tangent 
 along the curves on the the Helmholtz free energy per atom verses inverse 
-density plots for both the liquid and the solid. 
+density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
+liquid and the solid at a given temperature. 
 \begin{figure}[h!]
@@
 The function $f_{ij}$ is called a Mayer function, and is given by
-\begin{equation}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
+\begin{equation}\label{eq:mayerfunction}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
 Multiplying the Mayer function terms, the excess free energy can be written as
@@
 with weight functions
-\begin{equation}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
-\begin{equation}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
+\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
+\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
 
 %\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
-\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
+\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
 %\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
-\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
+\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
 where the weight functions $w_{2}$ and $w_{3}$ have the relation
@@
 
-The weight functions can be related to the mayer function \begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} where r is the distance between the centers of two hard spheres.
-For a hard sphere fluid the interatomic potential $V(r)$ is given by \begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
-and so the mayer function for a hard sphere is given by \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
+The weight functions can be related to the mayer function 
+\begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} 
+where r is the distance between the centers of two hard spheres. 
+For a hard sphere fluid the interatomic potential $V(r)$ is given by 
+\begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
+and so the mayer function for a hard sphere is given by 
+\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
 The mayer function can be written in terms of weight functions if two 
 additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
-\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  where the vector weights are defined as \begin{equation}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
-\begin{equation}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}
+\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  
+where the vector weights are defined as 
+\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
+\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation}
 
@@
 where the tensor weight is given by
-\begin{equation}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
+\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
 %and $\Phi_3$ is now
@@
 \begin{align}
-\Phi_1 &= -n_{0}ln(1-n_{3}) \\
-\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
-\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
+\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
+\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
+\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
 \end{align}
@@
 Energy becomes a functional of the number density profile $n(\vec{r})$ :
-\begin{equation}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
+\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
 
@@
 
-Our goal is to generate computationally temperature-density and pressure-
-temperature phase diagrams for a Weeks-Chandler-Anderson model fluid using 
+Our goal is to generate temperature-density and pressure-temperature 
+phase diagrams for a Weeks-Chandler-Anderson model fluid using 
 classical Density Functional Theory. 
@@
 abrupt change from some maximum lower density to some minimum higher density 
-occurs for certian combinations of pressure and temperature which form
+occurs for certain combinations of pressure and temperature which form
 a line on a pressure-temperature phase diagram. This line distinguishes 
-the region in which the fluid exhibits lower densities from the region 
-in which the fluid exhbits higher densities. These two regions are known
-as the liquid state and the solid state of the fluid.
-
+the liquid region which exhibits lower densities from the 
+solid which exhbits higher densities.
+
+The transition between the liquid 
+and solid state is abruptly inhomogeneous in terms of the number density, 
+and we study the behavior 
+of the fluid by treating the liquid as having a homogeneous number density, 
+and by constructing a crystal lattice in real space for the solid. 
 It is known that a fluid with a soft potential like the WCA fluid will 
 form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
-The transition between the liquid 
-and solid states is abruptly inhomogeneous, and we study the behavior 
-of the fluid by treating the liquid as homogeneous, and by constructing 
-a crystal lattice in real space for the solid. Although the entire 
+Although the entire 
 crystal lattice can be constructed with FCC cells, 
@@
 %}
-In a large collection of cells, the atom at each lattice site of 
-will not realistically reside exactly at the lattice point in the center 
-of the cell. Statistically there will typically be a normal distribution 
+In a large collection of cells, the atom at each lattice site
+will not realistically reside exactly at the lattice point. 
+Statistically there will typically be a normal, or Gaussian distribution 
 of positions about the lattice point as pictured in the far right box in 
-Figure \ref{fig:Ensemble_Gaus}. A plot of a 2-dimensional Gaussian 
-distribution is shown in Figure \ref{fig:Gaus_plot}. 
+Figure \ref{fig:Ensemble_Gaus}. 
+A plot of a Gaussian distribution in 2-dimensions is shown in Figure~\ref{fig:Gaus_plot}
+where the width of the Guassian is described by $\sigma$.  
 A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
@@
 \begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
-where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time:
+where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time.
 \begin{equation}n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~\end{equation}
@@
 into account while keeping %the volume of the bold box, and thus 
-the overall number density the same. This is done by 
+the overall number density the same by 
 making the cells smaller. A large collection of smaller cells creates a 
-new number density profile shown in Figure \ref{fig:Ensemble_Smallcells} 
-with eight smaller Gaussians of one-eigth the height. 
+new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
+with eight smaller Gaussians each of which is now one-eigth the height. 
 Instead of thinking of an atom in one of the eight small cells, the same 
-Gaussian distribution can be obtained from thinking of one-eigth of an 
-atom in each small cell. This makes it convenient to describe the new 
+Gaussian distribution can be obtained from thinking of one-eighth of an 
+atom in each small cell as shown in Figure~\ref{fig:SameStatPic}. 
+This makes it convenient to describe the new 
 number density in terms of the fraction of vacancies, $f_v$: 
@@
 found, and added together to get the total Helmholtz free energy.
-According to classical Density Functional Theory, the minimum free energy, 
-and its corresponding number density profile are then the equilibrium values 
+According to classical Density Functional Theory, of all the Helmholtz 
+free energies calculated is the minimum Helmholtz free energy, and its 
+corresponding number density profile that are the actual equilibrium values 
 for that temperature and number density.
@@
 
-%Once the Helmholtz free energy for both the crystal and liquid states of 
-%the fluid are found for a given temperature and density, they can be 
+Once the Helmholtz free energy for both the crystal and liquid states of 
+the fluid are found for a given temperature and density, they can be 
+used to construct phase diagrams.
 %compared to find the resulting state of the fluid. 
@@
 Freezing and Maxwell's Double Tangent Construction.
+
+\section{Our Functional for the Excess Helmholtz free energy}
+In order to apply classical Density Functional Theory, the Helmholtz free 
+energy must be found. Computing the ideal portion of the Helmholtz free 
+energy is straightforward, but computing the excess portion of the 
+Helmholtz free energy is difficult since it depends on the interatomic 
+potential energy $V(r)$ between each sphere with all the other spheres, 
+as is shown in Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
+Soft Fundamental Measure Theory is a method
+developed to form an expression for the excess Helmholtz free energy without
+explicitly taking into account the interaction of each sphere with all the others, 
+but it must be taylored to fit a particular fluid\cite{schmidt1999density}. 
+%The excess Helmholtz free energy is particular to the fluid being modeled, 
+%as it depends on the interatomic potential V(r). 
+It is based on the expression for the excess Helmholtz 
+free energy of a homogeneous Hard-sphere fluid derived 
+from Scaled Particle Theory, and given in Equation~\ref{FexfromSPT}. 
+The resulting functional 
+%(given in Equations~\ref{eq:F_ex-n-of-r} - \ref{eq:Phi3} and repeated here) 
+is written in general form as 
+\begin{align}\label{eq:Fexfunctional}
+  F_{excess}[n(\vec{r})]=k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}
+\end{align}
+where
+\begin{align}
+\Phi_1 &= -n_{0}\ln(1-n_{3}) \\
+\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
+\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
+\end{align} 
+and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
+for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
+\begin{align}\label{eq:numdenprofile}
+	n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
+\end{align}
+where the number density profile $n(\vec r)$ within the integral will be 
+given in our program by a Gaussian function centered at a lattice point, 
+and $w(r)$ is a weight function.
+%\begin{align}
+  %n(\vec r)=(1-f_v)\frac{\exp^{-\frac{r^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}
+%\end{align}
+%
+There are four scaler weight functions, two vector weight functions, and 
+one tensor weight function all of which 
+%The weight functions $w_0(r), w_1(r), w_2(r)$ 
+depend on the expression for $w_3(r)$ 
+which in turn depends on the interatomic potential $V(r)$ that exists 
+between the spheres, and which defines the particular fluid being modeled.
+%FMT models the Hard-sphere fluid, and SFMT models fluids with soft potentials.
+We introduce our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ 
+to fit the functional to a WCA fluid.
+The scalor weight functions are given by
+\begin{align}\label{eq:weights}
+  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
+  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
+  %w_{2}(r) &=-\frac{\partial{w_3(r)}}{\partial{r}}   \\
+  %w_{3}(r) &=~?
+  w_2(r) &=-\frac{\partial{w_3(r)}}{\partial{r}} \\
+  &= \frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}  \\
+  w_3(r) &=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]  
+  %\vec{w_{1}} &= w_{1}\frac{\vec{r}}{r}   \\
+  %\vec{w_{2}} &= w_{2}\frac{\vec{r}}{r}  \\
+  %\overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)
+\end{align}
+%
+The vector weights are given by 
+\begin{equation}\label{eq:w_v2}{  \vec{w_{2}}=w_{2}\frac{\vec{r}}{r}  }\end{equation}
+\begin{equation}\label{eq:w_v1}{  \vec{w_{1}}=w_{1}\frac{\vec{r}}{r}  }\end{equation} 
+The tensor weight is given by
+\begin{equation}\label{eq:tensorweight-methods}\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)\end{equation}
+%
+ \begin{figure}[h!]
+    \centering
+    %\includegraphics[height=6cm]{GaussandW2_actual.png}
+    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    %view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
+    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    %is generated randomly according to a Gaussian distribution.} 
+  \label{fig:GaussandW2_actual}
+  \end{figure} 
+%We use Soft Fundamental Measure Theory to form a functional for the excess 
+%Helmholtz free energy appropriate for the WCA fluid by introducing our 
+%
+%Equation 2.94  general form of Fex for n(r)    
+%\ref{eq:F_ex-n-of-r}, 
+%phis 1-3   2.90, 2.91, 2.92
+%\ref{eq:Phi1}, 
+%\ref{eq:Phi2}, 
+%\ref{eq:Phi3}, 							
+%weights 2.73, 2.74, 2.89  2.79, 2.80  replace-> 2.71, 2.72		
+%\ref{eq:w0}, 
+%\ref{eq:w1}, 
+%\ref{eq:w2}, 
+%\ref{eq:w3}, 
+%\ref{eq:w_v2}, 
+%\ref{eq:w_v1}, 
+%\ref{eq:tensorweight}, 
+%%weight relations 2.75, 2.81
+%\ref{w2_w3_relation}, 
+%\ref{mayerder=w2convolution},
+
+The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
+Schmidt's proposed erf model~\cite{schmidt2000fluid} formed to model a steep, 
+soft potential in which he set
+\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
+where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
+and $w_3(\infty)=0$. 
+%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
+Our functional is  created by replacing Schmidt's parameters $\sigma$ 
+and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
+\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
+%
+The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
+temperature dependence that enters in through the expression for the 
+error function potential $V_{erf}(r)$ which 
+%is used to approximate the non-temperature dependent WCA potential that 
+%is more difficult to implement computationally. The erf potential 
+emerges from the relationships
+that the weight functions must satisfy.
+%potential at temperatures of interest will be used instead of the WCA potential. 
+Using the relation
+%Equation~\ref{w2_w3_relation}, shown again below, 
+\begin{align}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{align}
+and solving for $w_2(r)$
+\begin{align}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{align} 
+we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
+Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
+\begin{align}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{align}
+gives
+\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
+which can be integrated to obtain the Mayer function
+\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
+The Mayer function for the error function potential $V_{erf}(r)$ is thus 
+given by
+\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
+Combining this result with the definition of the Mayer function 
+\begin{align}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{align} 
+%This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
+gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
+\begin{equation}
+	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
+\end{equation}
+
+Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
+$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off. 
+$\Xi$ is related to the width of the functions which 
+become wider in general as the temperature increases. 
+The radial distance 
+to the peak is related to $\alpha$, and decreases as the temperature 
+increases, that is, the shells shrink, as does 
+the sphere.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{weight_functions.pdf}
+    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2.}
+    \label{fig:weight_functions}
+  \end{figure}
+
+ \begin{figure}[h!]
+    \centering
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W2.png}
+      % \includegraphics[width=5cm]{figs/w2-ring}
+      \caption{A crossectional view of $w_2$}
+    \end{subfigure} 
+    \begin{subfigure}[b]{0.4\linewidth}
+      \includegraphics[height=5cm]{W3.png}
+      \caption{A crossectional view of $w_3$}
+    \end{subfigure} 
+    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    \label{fig:W0andW3}
+    \end{figure}
+
+Expressions for $\alpha(T)$ and $\Xi(T)$  can be found as follows. 
+The parameter $\alpha(T)$ is derived by setting 
+the erf potential equal to the WCA potential when r equals $\alpha$ 
+\begin{equation}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
+and is given by 
+\begin{equation}\label{alphaT}
+	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
+\end{equation} 
+Figure \ref{fig:alphaXivsT} shows how $\alpha$ varies with temperature 
+when $\sigma=1$ and $\epsilon=1$.    
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=9cm]{plot_alpha.pdf}
+    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
+    for $\sigma=1$ and $\epsilon=1$}
+    \label{fig:alphaXivsT}
+  \end{figure}
+The value of $\Xi$ at a given temperature, T, can be found computationally 
+by equating the second virial coefficient, $B_2$, evaluated for the error function 
+potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
+WCA potential $V_{WCA}$.
+\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+The second virial coefficient can be derived from Euler's equation for 
+the internal energy of a system, 
+\begin{equation}U=TS-PV+\mu{N}\end{equation}
+using the relations (discrete versions are shown here for clarity)
+\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
+\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
+\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
+which, for a canonical ensemble, gives
+\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
+Expressing the Helmholtz free energy in terms of the pair potential 
+$V(r)$ results in an equation of state expressed in terms of the virial expansion 
+\begin{equation}\label{virialequation}
+    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
+\end{equation}
+where 
+\begin{equation}
+	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
+\end{equation}
+Evaluating $B_{2}$ for the erf potential gives
+\begin{equation}
+	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
+\end{equation}
+where $\alpha(T)$ is given by Equation~\ref{alphaT}. Similarly, 
+$B_{2WCA}$ can be found computationally from 
+\begin{equation}
+	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
+\end{equation}
+
+A plot of the resulting second virial coefficient $B_2(T)$  with respect 
+to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
+through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
+Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
+$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
+potential which is not temperature dependent.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{B2_vs_T_with_xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $B_2$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:B2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $\Xi$ generated from 
+    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:xi_fromB2vsT}
+  \end{figure}
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
+    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
+    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
+    shown in black.}
+    \label{fig:erf_Potential_xifromB2}
+  \end{figure}
+
+%\color{red}
+% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
+%\color{black}
+
+%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
+%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
+%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
+%classical fluid.
+
+%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
+%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
+%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
+%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
+%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
+%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
+%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
+%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
+%The temperature dependence of parameters $\alpha$ and $\Xi$ 
+%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
+%can be related 
+%to the weight functions. 
+%$\Xi$ is related to the width of the functions which 
+%become wider in general as the temperature increases. 
+%The radial distance 
+%to the peak is related to $\alpha$, and decreases as the temperature 
+%increases, that is, the shells shrink, as does 
+%the sphere.
+%The functions can be treated as neglible outside of a distance of 
+%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
+%the weight function.
+
+ %\begin{figure}[h!]
+    %\centering
+    %\includegraphics[height=7cm]{weight_functions.pdf}
+    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    %z-axis for a reduced temperature of 2.}
+    %\label{fig:weight_functions}
+  %\end{figure}
+
+ %\begin{figure}[h!]
+    %\centering
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{W2.png}
+      %% \includegraphics[width=5cm]{figs/w2-ring}
+      %\caption{A crossectional view of $w_2$}
+    %\end{subfigure} 
+    %\begin{subfigure}[b]{0.4\linewidth}
+      %\includegraphics[height=5cm]{W3.png}
+      %\caption{A crossectional view of $w_3$}
+    %\end{subfigure} 
+    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
+    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
+    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
+    %\label{fig:W0andW3}
+    %\end{figure}
 
@@
 \begin{align}
-  n(|\vec{r}-\vec{R_i}|)=\sum_i{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
+  n(|\vec{r}-\vec{R_j}|)=\sum_j{(1-f_v)\frac{e^{-\frac{{|\vec{r}-\vec{R_j}|}^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}}
 \end{align}
+where $\vec R_j$ is the position vector of the jth lattice site.
 %For small gw values it is assumed that the Guassians do not overlap, 
@@
 %for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
-using SFMT with a uniform number density profile, n that does not depend 
-on $\vec r$. 
+using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
+with a uniform number density profile, n that does not depend on $\vec r$. 
 %done in hf.energy() note- does not depend on fv or gw but does use SFMT):
@@
 
-The crystal excess Helmholtz free energy is computed computationally 
-using SFMT with a number density profile $n(\vec r)$ that varies spacially.
+The crystal excess Helmholtz free energy is computed  
+using SFMT with a number density profile 
+%$n(|\vec r'-\vec R_j|)$ 
+%used in Equation~\ref{eq:numdenprofile} 
+%that is composed of Guassian distributions centered at each lattice site.
+that consists of Guassian distributions 
+%$n(|\vec r'-\vec R_j|)$ 
+centered at each lattice site.
+%specified by position vectors $\vec R$.
+The weight functions $w_i(|\vec r - \vec r'|)$ are centered at $\vec r'$,
+as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}. 
+But because the weight function is a function of the magnitude of the 
+difference between $\vec r$ and $\vec r'$, it can be thought of as 
+centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_thinkas}. 
+Keeping the weight function centered at $\vec r$, 
+%, and setting $\vec{dr}=\vec r' - \vec{R}_j$, 
+% and 
+%a vector $\vec dr$ is defined to be the difference between $\vec R$ and $\vec r'$.
+% $\vec dr = \vec r' - \vec R$.
 %First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
@@
 %\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
-Monte-Carlo integration is used to compute the weighted densities 
-$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$.
+Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
+%$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$
 %form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
 %
-The center of with weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
-within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
-the Gaussians is selected that is within the primitive cell, or one of the 
-neighboring cells within a prudently picked inclusion radius as Gaussians 
-farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).
-
 \begin{figure}[h!]
@@
 \end{figure} 
-
-The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
-$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
-$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
-$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
-where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
-some point in space. Monte-Carlo integration works by randomly generating 
-$\vec{dr}$ according to a Gaussian distribution. Setting $\vec{r}'=\vec{R}_j + 
-\vec{dr}$, and $n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
-is the gaussian number density profile for no vacancies, the equation:
-
-\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
-implemented computationally in the form:
-\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n_g(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
-becomes:
-\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{\vec{dr}}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-where NumPoints is the number of vectors $\vec{r'}$ randomly generated, and $n_g(\vec{r'})$ 
-(the value of the Gaussian at at radius of $|\vec{r'}|$) is equal to a probability density 
-which when multiplied by dV gives the probability $P_{\vec{dr}}$ at which the value of 
+%
+%\textcolor{red}{The center of the weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
+%within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
+%the Gaussians
+%is selected
+%that is within the primitive cell, or one of the 
+%neighboring cells within an inclusion radius as Gaussians 
+%farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).}
+%
+%The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
+%$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
+%$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
+%$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
+%where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
+%some point in space.
+%Monte-Carlo integration works 
+by randomly generating vectors
+$\vec{r'}$ according to a Gaussian distribution. 
+%Setting $\vec{r}'=\vec{R}_j + \vec{dr}$, and 
+%Setting $\vec{dr}=\vec r' - \vec{R}_j$,  and 
+%$n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
+%where $n_g(\vec{r'})$ is the gaussian number density profile for no vacancies. 
+Only lengths for $\vec{r'}$ that are within an inclusion radius
+about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
+will be taken into account as Gaussians 
+farther away will not contriubute much.
+The equation for the contribution to the weighted density from a 
+Gaussian centered about a lattice site is given by
+%lattice point located at $\vec R_j$ is given by
+\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+and is implemented computationally in the form
+\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+which can be written as
+\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
+where NumPoints is the number of vectors $\vec{r'}$ randomly generated, 
+%and $n_g(\vec{r'} - \vec R_j)$ 
+%(the value of the Gaussian at a radius of $|\vec{r'}|$) 
+%(the value of the Gaussian at a radius of $|\vec{dr}|$) 
+and $n(\vec r')$
+is equal to a probability density 
+which when multiplied by dV gives the probability $P_{r'}$ at which the value of 
 $w_i(|\vec{r}-\vec{r}'|)$ will occur. 
-Using the relation:
+Using the relation
 \begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
 $n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{NumPoints}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{NumPoints}}\end{equation}
+\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{\text{NumPoints}}}\end{equation}
+This process is repeated to get the contributions to the weighted density 
+from the Gaussians around each lattice point which are then added together 
+to get the final value of the weighted density. 
 The result can be made more accurate by increasing NumPoints which is the 
@@
     \centering
-    \includegraphics[height=6cm]{GaussandW2_actual.png}
-    \caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    is generated randomly according to a Gaussian distribution.} 
+    %\includegraphics[height=6cm]{GaussandW2_actual.png}
+    \includegraphics[height=7cm]{weightandGuass1.pdf}
+    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
+    %view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
+    \caption{A Gaussian distribution about a lattice site is shown in blue, and a cross-sectional 
+    view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
+    %Here, the weight function 
+    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
+    %is generated randomly according to a Gaussian distribution.} 
   \label{fig:GaussandW2_actual}
@@
     \centering
-    \includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    %\includegraphics[height=6cm]{GaussandW2_thinkas.png}
+    \includegraphics[height=7cm]{weightandGuass2.pdf}
     \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 
-    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$  
-    and fixed in place as $\vec{dr}$ varies randomly.} 
+    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
+    and fixed in place as $\vec r'$
+    %as $\vec{dr}=\vec{r}'-\vec{R}$ 
+    varies randomly with a probability according to a Gaussian distribution.} 
   \label{fig:GaussandW2_thinkas}
@@
 
-\section{Our Functional for the Excess Helmholtz free energy}
-In order to apply classical Density Functional Theory, the Helmholtz energy
-must be found. Computing the ideal portion of the Helmholtz free energy is
-straightforward, but computing the excess portion of the Helmholtz free energy 
-is difficult since it depends on the interatomic potential energy $V(r)$ between 
-each sphere with all the other spheres. Soft Fundamental Measure Theory is a method
-developed to form an expression for the excess Helmholtz free energy without
-explicitly taking into account the interaction of each sphere with all the others, 
-but it must be taylored to fit a particular fluid\cite{schmidt1999density}. 
-%The excess Helmholtz free energy is particular to the fluid being modeled, 
-%as it depends on the interatomic potential V(r).
-
-We use Soft Fundamental Measure Theory to form a functional for the excess 
-Helmholtz free energy appropriate for the WCA fluid by introducing our 
-temperature dependent parameters $\alpha(T)$ and $\Xi(T)$. The functional 
-is written in general form as 
-\begin{align}
-  F_{excess}[n(\vec{r})]=k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}
-\end{align}
-where
-\begin{align}
-\Phi_1 &= -n_{0}\ln(1-n_{3}) \\
-\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
-\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
-\end{align} 
-and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
-for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
-\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-where the number density profile $n(\vec r)$ within the integral is given 
-by a Gaussian function centered at a lattice point,
-%\begin{align}
-  %n(\vec r)=(1-f_v)\frac{\exp^{-\frac{r^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}
-%\end{align}
-and the weight functions $w_0, w_1, w_2,$ and $w_3$ are given by
-\begin{align}\label{eq:weights}
-  w_{0}(r) &=\frac{w_{2}}{4\pi{r}^2} \\
-  w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
-  w_2(r)&=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2}
-  \\
-  w_3(r)&=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]
-\end{align}
-%
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_actual.png}
-    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    %view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    %is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
-Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off. 
-$\Xi$ is related to the width of the functions which 
-become wider in general as the temperature increases. 
-The radial distance 
-to the peak is related to $\alpha$, and decreases as the temperature 
-increases, that is, the shells shrink, as does 
-the sphere.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{weight_functions.pdf}
-    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2.}
-    \label{fig:weight_functions}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W2.png}
-      % \includegraphics[width=5cm]{figs/w2-ring}
-      \caption{A crossectional view of $w_2$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W3.png}
-      \caption{A crossectional view of $w_3$}
-    \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    \label{fig:W0andW3}
-    \end{figure}
-
-The weight fucntions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
-Schmidt's proposed erf model~\cite{schmidt2000fluid} in which he set
-\begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
-where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
-and $w_3(\infty)=0$. 
-%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
-Our functional is  created by replacing Schmidt's parameters $\sigma$ 
-and $a$ with our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$.
-\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
-
-The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
-temperature dependence that enters in through the expression for the 
-error function potential $V_{erf}(r)$ which is used to approximate the
-non-temperature dependent WCA potential that is more difficult to 
-implement computationally. The erf potential emerges from the relationships
-that the weight functions must satisfy.
-%potential at temperatures of interest will be used instead of the WCA potential. 
-Using Equation~\ref{w2_w3_relation}, shown again below, 
-\begin{displaymath}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{displaymath}
-and solving for $w_2(r)$
-\begin{displaymath}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{displaymath} 
-we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
-Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
-\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
-gives
-\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
-which can be integrated to obtain the Mayer function
-\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
-The Mayer function for the error function potential $V_{erf}(r)$ is thus 
-given by
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
-Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} 
-%This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
-gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
-\begin{equation}
-	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
-\end{equation}
-
-Expressions for $\alpha(T)$ and $\Xi(T)$  can be found as follows. 
-The parameter $\alpha(T)$ is derived by setting 
-the erf potential equal to the WCA potential when r equals $\alpha$ 
-\begin{equation}{V_{erf}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
-and is given by 
-\begin{equation}\label{alphaT}
-	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
-\end{equation} 
-Figure \ref{fig:alphaXivsT} shows how $\alpha$ varies with temperature 
-when $\sigma=1$ and $\epsilon=1$.    
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=9cm]{plot_alpha.pdf}
-    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
-    for $\sigma=1$ and $\epsilon=1$}
-    \label{fig:alphaXivsT}
-  \end{figure}
-The value of $\Xi$ at a given temperature, T, can be found computationally 
-by equating the second virial coefficient evaluated for the error function 
-potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
-WCA potential $V_{WCA}$.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
-The second virial coefficient can be derived from Euler's equation for 
-the internal energy of a system, 
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-using the relations
-\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
-\begin{equation}U=<E>=\sum_{i=0}^\infty{P_iE_i}\end{equation}
-\begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-which, for a canonical ensemble, gives
-\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Expressing the Helmholtz free energy in terms of the pair potential 
-$V(r)$ results in an equation of state expressed in terms of the virial expansion 
-\begin{equation}\label{virialequation}
-    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
-\end{equation}
-where 
-\begin{equation}
-	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
-\end{equation}
-Evaluating $B_{2}$ for the erf potential gives
-\begin{equation}
-	B_{2erf}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{erf}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
-\end{equation}
-where $\alpha(T)$ is given by Equation~\ref{alphaT}. Similarly, 
-$B_{2WCA}$ can be found computationally from 
-\begin{equation}
-	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
-\end{equation}
-
-A plot of the resulting second virial coefficient $B_2(T)$  with respect 
-to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
-through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
-Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
-$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
-potential which is not temperature dependent.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{B2_vs_T_with_xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $B_2$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:B2vsT}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $\Xi$ generated from 
-    matching the second virial coeficients for $V_{erf}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:xi_fromB2vsT}
-  \end{figure}
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{erf}(r)$ at several temperatures 
-    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
-    shown in black.}
-    \label{fig:erf_Potential_xifromB2}
-  \end{figure}
-
-%\color{red}
-% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
-%\color{black}
-
-%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
-%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
-%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
-%classical fluid.
-
-%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
-%The temperature dependence of parameters $\alpha$ and $\Xi$ 
-%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
-%can be related 
-%to the weight functions. 
-%$\Xi$ is related to the width of the functions which 
-%become wider in general as the temperature increases. 
-%The radial distance 
-%to the peak is related to $\alpha$, and decreases as the temperature 
-%increases, that is, the shells shrink, as does 
-%the sphere.
-%The functions can be treated as neglible outside of a distance of 
-%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
-%the weight function.
-
- %\begin{figure}[h!]
-    %\centering
-    %\includegraphics[height=7cm]{weight_functions.pdf}
-    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    %z-axis for a reduced temperature of 2.}
-    %\label{fig:weight_functions}
-  %\end{figure}
-
- %\begin{figure}[h!]
-    %\centering
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W2.png}
-      %% \includegraphics[width=5cm]{figs/w2-ring}
-      %\caption{A crossectional view of $w_2$}
-    %\end{subfigure} 
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W3.png}
-      %\caption{A crossectional view of $w_3$}
-    %\end{subfigure} 
-    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    %\label{fig:W0andW3}
-    %\end{figure}
+%-------Antithetic-------------------
+In order to further reduce the error 
+in the average, antithetic variates are used. Rather than computing an average 
+of a set of random points, each random point in the set is paired with a 
+point along the same line as the first point, but opposite in direction, 
+as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
+error in the average by filtering out an erroneous offset of values along 
+the line connecting the two points by averaging the two results, one of 
+which may be erroneously higher than it should be while the other is 
+erronesously lower than it should be.
+
+ \begin{figure}[h!]
+    \centering
+    \includegraphics[height=4cm]{AntitheticVariate.png}
+    \caption{Antithetic variates reduce first order error in an average. 
+    Each point in the average is paired with another point from along 
+    the same line as the first, and at an equal, but opposite distance.}
+    \label{fig:AntitheticVariate}
+  \end{figure} 
+%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
+%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}
+%The process is repeated for all the other $\vec{R}_j$'s to get the total 
+%weighted densities at vector $\vec{r}_k$ due to all the Gaussians. 
+%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
+%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
+%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
+
+The standard error of the mean (SEM) is given in general by
+\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
+%which gives a measure of the error in $n_3=\overline{w}_3$:
+To compute the error, the standard deviation of weight function $w_3$ is used 
+\begin{equation}SD=\sqrt{\overline{w_3^2}-(\overline{w}_3)^2}\end{equation} 
+\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{\text{NumPoints}}}}\end{equation} 
+NumPoints, the number of vectors randomly generated, is increased until
+\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
+where the maximum value that $n_3$ (which equals $\overline{w}_3$) can have without error is 1.
+
+%Figure \ref{fig:GaussandW2_actual} ......
 
@@
 
-%\section{Program Flowchart}
+\section{Program Outline}
 
@@
 
-%PROGRAM FLOWCHART:
-%\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
-%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
-%Loop: 
-%\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
-%\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
-%\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
-%\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
-%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-%\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{homogeneous}=\frac{hFideal_{primitive.cell} + hFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-%\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{homogeneous}\end{equation}
-%\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-%\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{homogeneous}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+\begin{displaymath}\text{Input:~~~~~~~~~~~~~~~~~~~~~~~~~~~T,~~ n,~~Range of fv,~~Range of gw~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+\begin{displaymath}\text{Set:~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~N~=~1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath} 
+\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
+Loop: 
+\begin{displaymath}\text{fv = one of the fv values from input}\end{displaymath}
+\begin{equation}N_{primitive.cell}=(1-\text{fv})\text{N}\end{equation}
+\begin{equation}V_{primitive.cell}=\frac{\text{1}}{\text{n}}(1-\text{fv})\text{N}\end{equation}
+\begin{displaymath}\sigma = \text{one of the gw values from input}\end{displaymath}
+\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
+\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{liquid}=\frac{lFideal_{primitive.cell} + lFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{liquid}\end{equation}
+\begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+\begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
+\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{liquid}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
 
@@
 %\begin{displaymath}\left(\frac{F}{N}\right)_{homogeneous}\end{displaymath}
+
 %lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)

### 2020-11-07 papers/thesis-kirstie/thesis.tex
edits

@@
 %$n(|\vec r'-\vec R_j|)$ 
-centered at each lattice site.
+centered at lattice sites specified by position vectors $\vec R$. 
+%where j indicates the jth lattice site.
 %specified by position vectors $\vec R$.
@@
 by randomly generating vectors
-$\vec{r'}$ according to a Gaussian distribution. 
+$\vec r' -\vec R$ according to a Gaussian distribution. 
 %Setting $\vec{r}'=\vec{R}_j + \vec{dr}$, and 
@@
 %where $n_g(\vec{r'})$ is the gaussian number density profile for no vacancies. 
-Only lengths for $\vec{r'}$ that are within an inclusion radius
+Only lengths for $\vec r'- \vec R$ that are within an inclusion radius
 about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
@@
 The equation for the contribution to the weighted density from a 
-Gaussian centered about a lattice site is given by
+Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
 %lattice point located at $\vec R_j$ is given by
-\begin{equation}{n_i(\vec{r})= \int_{r'}{(1-f_v)n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r' - \vec R_j)w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
 and is implemented computationally in the form
-\begin{equation}{n_i(\vec{r})= \sum_{r'}{(1-f_v)n(\vec{r'})w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r' -\vec R_j)w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
 which can be written as
-\begin{equation}{n_i(\vec{r})= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-where NumPoints is the number of vectors $\vec{r'}$ randomly generated, 
+\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
+NumPoints is the number of vectors $\vec r' - \vec R_j$ randomly generated about the jth lattice site, 
 %and $n_g(\vec{r'} - \vec R_j)$ 
@@
 %(the value of the Gaussian at a radius of $|\vec{dr}|$) 
-and $n(\vec r')$
+and $n(\vec r' - \vec R_j)$
 is equal to a probability density 
@@
 This process is repeated to get the contributions to the weighted density 
-from the Gaussians around each lattice point which are then added together 
-to get the final value of the weighted density. 
-The result can be made more accurate by increasing NumPoints which is the 
-number of vectors randomly generated.
-
+$n_i(\vec r)$ from the Guassians around each lattice point which begin to 
+merge together as the width of the Guassians increases.
+All the contributions are then added together 
+to get the final value of the weighted density at some point $\vec r$.
+
+The result can be made more accurate by increasing NumPoints (the 
+number of vectors randomly generated), and by using antithetic variates.
+%
  \begin{figure}[h!]
@@
     %view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
-    \caption{A Gaussian distribution about a lattice site is shown in blue, and a cross-sectional 
+    \caption{A Gaussian distribution about a lattice site located at 
+    $\vec R$ is shown in blue, and a cross-sectional 
     view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
@@
   \end{figure} 
-
+%
  \begin{figure}[h!]
@@
     the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
-    and fixed in place as $\vec r'$
+    and fixed in place as $\vec r' - \vec R$
     %as $\vec{dr}=\vec{r}'-\vec{R}$ 
@@
   \end{figure} 
-
+%
 %Monte-Carlo integration is employed to more efficiently 
@@
 %which is minimized. 
-
+%
 %-------Antithetic-------------------
-In order to further reduce the error 
-in the average, antithetic variates are used. Rather than computing an average 
-of a set of random points, each random point in the set is paired with a 
-point along the same line as the first point, but opposite in direction, 
-as shown in Figure \ref{fig:AntitheticVariate}. This reduces first order 
-error in the average by filtering out an erroneous offset of values along 
-the line connecting the two points by averaging the two results, one of 
+%To further reduce the error in the average, antithetic variates are used. 
+%Rather than computing an average of a set of random points, 
+%antithetic variates can be used to reduce error. 
+To use antithetic variates,
+each randomly generated  vector is paired with a 
+vector of equal magnitude, but opposite in direction. 
+%as shown in Figure \ref{fig:AntitheticVariate}. 
+%The quantity that is a function of the randomly generated vector
+%now has two values that can be averaged to get a more accurate value. 
+First order error is reduced
+%in the quantity that is a function of the randomly generated vector
+%average by filtering out an erroneous offset of values 
+%along the line connecting the two points 
+by averaging the two results, one of 
 which may be erroneously higher than it should be while the other is 
@@
     \centering
-    \includegraphics[height=4cm]{AntitheticVariate.png}
-    \caption{Antithetic variates reduce first order error in an average. 
-    Each point in the average is paired with another point from along 
-    the same line as the first, and at an equal, but opposite distance.}
+    %\includegraphics[height=4cm]{AntitheticVariate.png}
+    %\caption{Antithetic variates reduce first order error in an average. 
+    %Each point in the average is paired with another point from along 
+    %the same line as the first, and at an equal, but opposite distance.}
     \label{fig:AntitheticVariate}
@@
 \begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
-where the maximum value that $n_3$ (which equals $\overline{w}_3$) can have without error is 1.
+where the maximum value that the weighted density $n_3$ 
+%(which equals $\overline{w}_3$) 
+can properly have is 1.
 

### 2020-11-07 papers/thesis-kirstie/thesis.tex
edits

@@
 %along the line connecting the two points 
-by averaging the two results, one of 
+by averaging the two results that depend on the vectors, one of 
 which may be erroneously higher than it should be while the other is 

### 2020-11-08 papers/thesis-kirstie/thesis.tex
edits

@@
 
-A description of cDFT ordinarily begins with considering the Grand Free 
+A description of cDFT often begins with considering the Grand Free 
 Energy defined as
@@
 \begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi=\frac{V_{ext}}{N}$, $V_{ext}(\vec r)$ is an externally 
-applied potential energy, and the Helmholtz free energy is expressed as
+where $\phi=\frac{V_{ext}}{N}$ 
+is the externally 
+applied potential energy $V_{ext}(\vec r)$ per particle, 
+and the Helmholtz free energy is expressed as
 \begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}  
@@
 
-The functional for the ideal Helmholtz free energy is given by 
-Equation~\ref{eq:Fideal-n(r)}, and the functional for the excess Helmholtz 
-free energy depends on the particular fluid as defined by the spheres' 
-interatomic potential V(r). To implement cDFT, the number density profile 
-$n(\vec{r})$ is varied with the temperature fixed until the free energy 
-is minimized. 
+%According to classical Density Functional Theory,
+
+%A number density profile is a time-invariant average 
+%of the spacial distribution of the particles that make up the fluid, and
+%There will be one, and only one 
+%that minimizes the Helmholtz free energy.
+%The minimum free energy, 
+%is the value of the Helmholtz free energy that
+%the system will have when it is in equilibrium at that temperature, 
+%and its corresponding number density profile 
+%are the equilibrium values the system will have at that temperature.
+%is the equilibrium number density profile. 
+%
+In this section 
+%To form the functional for the Helmholtz free energy, expressions for 
+a functional for the Helmholtz free energy was formed by adding 
+the ideal and the excess portions of the free 
+energy.
 \begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
-The minimum free energy and its corresponding number density 
-profile are then the equilibrium values for that temperature.
+The functional for the ideal portion of the Helmholtz free energy is given by 
+Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
+free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
+form for the excess Helmholtz free energy depends on the particular 
+fluid being modeled as defined by the  
+interatomic potential energy V(r) between the spheres, which can make
+forming an expression difficult. 
+Once an expression for the Helmholtz free energy as a functional of the
+number density profile $n(\vec r)$ is formed then, according to classical 
+Density Functional Theory, the equilibrium value of the 
+Helmholtz energy, and its cooresonding equilibrium density profile can be
+found by varying the number density profile
+%can be varied 
+through all possible configurations at some fixed temperature
+until the free energy is minimized.
 
 \section{Soft Fundamental Measure Theory (SFMT)}
-In order to use cDFT, a functional for the excess Helmholtz free
+In order to use classical Density Functional Theory, a functional for 
+the excess Helmholtz free
 energy for the particular fluid being modeled must be formed. But forming 

### 2020-11-10 papers/thesis-kirstie/thesis.tex
edits

@@
 %
-In this section 
-%To form the functional for the Helmholtz free energy, expressions for 
-a functional for the Helmholtz free energy was formed by adding 
-the ideal and the excess portions of the free 
-energy.
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
-The functional for the ideal portion of the Helmholtz free energy is given by 
-Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
-free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
-form for the excess Helmholtz free energy depends on the particular 
-fluid being modeled as defined by the  
-interatomic potential energy V(r) between the spheres, which can make
-forming an expression difficult. 
 Once an expression for the Helmholtz free energy as a functional of the
@@
 until the free energy is minimized.  
+In this section 
+%To form the functional for the Helmholtz free energy, expressions for 
+a functional for the Helmholtz free energy was formed by finding 
+general expressions for the ideal and the excess portions of the free 
+energy, and adding them together.
+\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+The functional for the ideal portion of the Helmholtz free energy is given by 
+Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
+free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
+form for the excess Helmholtz free energy depends on the particular 
+fluid being modeled as defined by the  
+interatomic potential energy V(r) between the spheres, which can make
+forming an expression difficult. In this paper, Soft Fundamental Measure
+Theory will be used to form an epxression for the excess Helmholtz free 
+energy, as explained in the next section.
 
 \section{Soft Fundamental Measure Theory (SFMT)}
-In order to use classical Density Functional Theory, a functional for 
-the excess Helmholtz free
-energy for the particular fluid being modeled must be formed. But forming 
-an expression based on Equation~\ref{eq:Fexcess-simplified} can be challenging. 
-For the Hard Sphere fluid, there is another way to form an expression for 
-the excess Helmholtz free energy in terms of a homogeneous number density 
-and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
+In order to apply classical Density Functional Theory to a system that has
+a fixed temperature, volume, and nubmer of particles, a functional for 
+the Helmholtz free energy must first be formed. An expression for the ideal 
+portion of the Helmnoltz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
+but forming an expression for the excess Helmholtz free
+energy for the particular fluid being modeled 
+from Equation~\ref{eq:Fexcess-simplified} can be challenging. 
+Soft Fundamental Measure Theory (SFMT) is a method for forming an 
+expression for the excess Helmholtz free energy of a Soft-sphere fluid,
+and will be adapted in the Methods section of this paper to fit the 
+soft-sphere WCA fluid.
+SFMT is an extension of Fundamental Measure Theory (FMT)
+that was developed for the Hard-sphere fluid, and which can be derived from
+Scaled Particle Theory (SPT).  
+%For the Hard Sphere fluid, there is another way to form an expression for 
+%the excess Helmholtz free energy in terms of a homogeneous number density 
+%and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
 
@@
 potential of a homogeneous Hard Sphere fluid which can then be integrated 
-to find the excess Helmholtz free energy.
+to find the excess Helmholtz free energy according to the relation
 %\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
@@
 work of inserting a sphere of radius R into a fluid consisting of hard 
-spheres of radius R. Widom's result was obtained by considering the difference 
+spheres of radius R. 
+\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
+Widom's result was obtained by considering the difference 
 in the excess Helmholtz free energy when one atom is added to an otherwise 
@@
 difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
-The resulting equation is known as Widom's Insertion Formula given by
+The resulting equation is known as Widom's Insertion Formula.
 %\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-\color{red} \begin{equation}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
-where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
-hard sphere, R, and the cavity is either filled or not. \color{black}
+Considering that at each point in the fluid, there is a probability that 
+either a hard sphere of radius R could be inserted, or not, and using 
+these two probablities to evaluate $<e^{-\beta{V_{diff}}}>$ this becomes
+\begin{equation}{~}=-k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)\end{equation}
+\begin{equation}{~}=V_{diff}=W_{rev}(R)\end{equation}
+Returning to Widom's insertion formula Equation~\ref{widoms-insertion-formula},
+and considering that it takes no energy for a hard sphere to be placed at a point
+in the fluid where it does not overlap with any other spheres, but would require
+infinite energy to place a sphere where it would overlap another sphere, 
+Equation~\ref{widoms-insertion-formula} becomes
+\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
+where the probability of a position existing in a homogeneous Hard-sphere fluid 
+with number density n that a hard sphere of radius R can be placed without 
+overlapping another sphere is given by the number density times the volume 
+of the sphere, $nV_{Sphere}$.
+
+%where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
+%hard sphere, R, and $<e^{-\beta{V_{diff}}}>$ is evaluated using the 
+%probablities that the cavity is either filled or not filled~\cite{Hansen}.
 %For cavities that are much larger than one of the hard spheres, 
@@
 the excess free energy becomes
-\begin{displaymath}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}}\color{white}\right)}\end{displaymath}
-\begin{equation}{\color{white}\left(\color{black}+\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}\right)\color{black}V}\end{equation}
-where the tensor weight is given by
-\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
-%and $\Phi_3$ is now
-and
+\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
+where
 \begin{align}
@@
 \end{align}
+and the tensor weight is given by
+\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation}
 

### 2020-11-10 papers/thesis-kirstie/thesis.tex
edits and odg file

@@
 but forming an expression for the excess Helmholtz free
-energy for the particular fluid being modeled 
-from Equation~\ref{eq:Fexcess-simplified} can be challenging. 
+energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
+being modeled can be challenging. 
 Soft Fundamental Measure Theory (SFMT) is a method for forming an 
-expression for the excess Helmholtz free energy of a Soft-sphere fluid,
-and will be adapted in the Methods section of this paper to fit the 
-soft-sphere WCA fluid.
+expression for the excess Helmholtz free energy of a Soft-sphere fluid
+that does not depend on  explicitly taking into account the interaction 
+that each sphere with every other sphere in the fluid as required by 
+Equation~\ref{eq:Fexcess-simplified}. In the next chapter on methods, we
+will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
+
 SFMT is an extension of Fundamental Measure Theory (FMT)
@@
 %and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
-
 Scaled Particle Theory seeks a general expression for the excess chemical 
@@
 spheres of radius R. 
-\begin{displaymath}\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}\end{displaymath}
+\begin{align}
+	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
+\end{align}
 Widom's result was obtained by considering the difference 
@@
 difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
-The resulting equation is known as Widom's Insertion Formula.
+The resulting equation, shown below, is known as Widom's Insertion 
+Formula~\cite{Hansen}.
 %\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-Considering that at each point in the fluid, there is a probability that 
-either a hard sphere of radius R could be inserted, or not, and using 
-these two probablities to evaluate $<e^{-\beta{V_{diff}}}>$ this becomes
-\begin{equation}{~}=-k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)\end{equation}
-\begin{equation}{~}=V_{diff}=W_{rev}(R)\end{equation}
-Returning to Widom's insertion formula Equation~\ref{widoms-insertion-formula},
-and considering that it takes no energy for a hard sphere to be placed at a point
-in the fluid where it does not overlap with any other spheres, but would require
-infinite energy to place a sphere where it would overlap another sphere, 
-Equation~\ref{widoms-insertion-formula} becomes
-\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3)\end{equation}
-where the probability of a position existing in a homogeneous Hard-sphere fluid 
-with number density n that a hard sphere of radius R can be placed without 
-overlapping another sphere is given by the number density times the volume 
-of the sphere, $nV_{Sphere}$.
+The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering
+that at each point in the fluid, there is a probability that 
+a hard sphere of radius R has space available to exist at that point
+without overlapping with other spheres, or it does not, as shown in 
+Figure~\ref{fig:p_overlap}.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{P_overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where 
+    there is no overlap such as shown by the black dot at the center of 
+    the dotted region near the top. But it is not possible to place a 
+    sphere at a position where an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
+
+Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
+\begin{align}
+	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
+	         &= V_{diff} \\
+             &= W_{rev}(R)
+\end{align}
+The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
+that it takes no energy for a hard sphere to be placed at a point
+in the fluid where it does not overlap with any other spheres 
+($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
+to place a sphere where it would overlap another sphere. 
+In this case Equation~\ref{widoms-insertion-formula} becomes
+%\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3) &  \mbox{for} & R_{\cavity}=R\end{equation}
+\begin{align}
+  \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
+\end{align}
+where the radius of the cavity $R_{cavity}$ equals the radius of a hard sphere R,
+and $nV_{Sphere}$ is the probability that a sphere cannot be placed at
+some random point in the homogeneous fluid of number density, $n$, without
+overlapping with another sphere.
 
@@
 %For cavities that are much larger than one of the hard spheres, 
-Forming a more general result for different sizes of cavities in the fluid, 
-Scaled Partical Theory arrives at an expression for the excess Helmholtz 
-free energy of a homogeneous fluid~\cite{Hansen} in terms of the packing fraction $\eta$ 
+Forming a more general result by requiring 
+continuity at the boundaries between equations for fluids with 
+cavities of a radius $\operatorname{R}_{cavity}$, that is larger, or very 
+much larger than the hard sphere radius R, Scaled Partical Theory arrives at an 
+expression~\cite{Hansen} for the excess Helmholtz 
+free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
 \begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 

### 2020-11-10 papers/thesis-kirstie/thesis.tex
edits

@@
 ($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
-to place a sphere where it would overlap another sphere. 
+to place a sphere where it would overlap another sphere, as described 
+in Figure~\ref{fig:p_overlap}.
 In this case Equation~\ref{widoms-insertion-formula} becomes
@@
 Forming a more general result by requiring 
-continuity at the boundaries between equations for fluids with 
+continuity at the boundaries between equations written for fluids with 
 cavities of a radius $\operatorname{R}_{cavity}$, that is larger, or very 

### 2020-11-10 papers/thesis-kirstie/thesis.tex

@@
 \end{align}
-where the number density profile $n(\vec r)$ within the integral will be 
-given in our program by a Gaussian function centered at a lattice point, 
-and $w(r)$ is a weight function.
+%where the number density profile $n(\vec r)$ within the integral will be 
+%given in our program by a Gaussian function centered at a lattice point, 
+%and $w(r)$ is a weight function.
+where the number density profile $n(\vec r)$ within the integral is given 
+by a Gaussian function centered at a lattice point.
 %\begin{align}
@@
 
+%<<<<<<< HEAD
+%\davidsays{Maybe say this before presenting the weights? But also, I don't think
+%there is any benefit to introducing $a$ and $\sigma$ here.} 
+The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
+Schmidt's proposed erf model~\cite{schmidt2000fluid} in which he set
+\begin{equation}
+  w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]
+\end{equation} 
+where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
+and $w_3(\infty)=0$. 
+%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
+Our functional is  created by replacing Schmidt's parameters $\sigma$ 
+and $a$ with temperature-dependent parameters $\alpha(T)$ and $\Xi(T)$.
+\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
+
+The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
+temperature dependence that enters in through the expression for the 
+error function potential $V_{\operatorname{erf}}(r)$ which is used to approximate the
+non-temperature dependent WCA potential that is more difficult to 
+implement computationally. The erf potential emerges from the relationships
+that the weight functions must satisfy\davidsays{In order for what?}.
+%potential at temperatures of interest will be used instead of the WCA potential. 
+Using Equation~\ref{w2_w3_relation}, shown again below, 
+\begin{align}
+  w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+\end{align}
+and solving for $w_2(r)$
+\begin{align}
+  \frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)
+\end{align} 
+we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
+Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
+\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
+gives
+\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
+which can be integrated to obtain the Mayer function
+\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
+The Mayer function for the error function potential $V_{\operatorname{erf}}(r)$ is thus 
+given by
+\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{\operatorname{erf}}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
+Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} 
+%This gives an interatomic potential of \begin{equation}{V_{\operatorname{erf}}=-k_BT\ln(f(r)-1)}\end{equation} 
+gives an expression for the for the error function potential  $V(r)=V_{\operatorname{erf}}(r)$ 
+\begin{equation}
+	V_{\operatorname{erf}}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
+\end{equation}
+
 Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
@@
   \end{figure}
-The value of $\Xi$ at a given temperature, T, can be found computationally 
-by equating the second virial coefficient, $B_2$, evaluated for the error function 
-potential $V_{erf}(r)$, and the second virial coefficent evalued for the 
+
+We determine value of $\Xi$ at a given temperature $T$,
+by equating the second virial coefficient evaluated for the error function 
+potential $V_{\operatorname{erf}}(r)$, and the second virial coefficent evaluated for the 
 WCA potential $V_{WCA}$.
-\begin{equation}B_{2erf}(T) =B_{2WCA}(T)\end{equation}
+\begin{equation}B_{2\operatorname{erf}}(T) =B_{2WCA}(T)\end{equation}
+\davidsays{I think we'd benefit from cutting much of the following, since you
+show the standard equations, but skip the bit where we put in the pair potential,
+which is the tricky step.}
 The second virial coefficient can be derived from Euler's equation for 

### 2020-11-10 papers/thesis-kirstie/thesis.tex
a couple edits in the abstract

@@
   neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
-  Theory (SFMT) with Tarazona's tensor expression for Phi3, and Schmidt's 
-  erf model in which we have incorporated our temperature dependent 
-  parameters to fit the results to those predicted by Monte-Carlo simulations. 
+  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
+  erf model in which we have incorporated temperature dependent
+  parameters to
+  \davidsays{fit the results to those predicted by Monte-Carlo simulations}.
   Our functional was successful in predicting the freezing of the WCA fluid 

### 2020-11-20 papers/thesis-kirstie/thesis.tex
spelling
fixed fig reference
more eidts

@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+
 The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering
@@
 Figure~\ref{fig:p_overlap}. 
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{P_overlap.pdf}
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=7cm]{figs/P_overlap.pdf}
     \caption{No energy is needed to place a sphere at a position where 
@@
   \end{figure}
-
 Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
@@
 	         &= V_{diff} \\
-             &= W_{rev}(R)
+             &= W_{rev}(R)  \label{workrev}
 \end{align}
@@
 %\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3) &  \mbox{for} & R_{\cavity}=R\end{equation}
-\begin{align}
+\begin{align} \label{muexcess}
   \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
@@
 some random point in the homogeneous fluid of number density, $n$, without
-overlapping with another sphere.
-
+overlapping with another sphere. Combining  Equations~\ref{workrev} and \ref{muexcess} gives
+\begin{align}
+  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
+\end{align}
 %where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
@@
 %For cavities that are much larger than one of the hard spheres, 
-Forming a more general result by requiring 
-continuity at the boundaries between equations written for fluids with 
-cavities of a radius $\operatorname{R}_{cavity}$, that is larger, or very 
-much larger than the hard sphere radius R, Scaled Partical Theory arrives at an 
-expression~\cite{Hansen} for the excess Helmholtz 
-free energy of a homogeneous fluid in terms of the packing fraction $\eta$ 
+Forming a more general result by requiring continuity of the functions, 
+along with their first and second derivatives, for the reversible work
+written for fluids with a cavitiy radius 
+%$\operatorname{R}_{cavity}$ 
+equal to, larger than, or very 
+much larger than the hard sphere radius R, Scaled Partical Theory arrives 
+at an expression~\cite{Hansen} for the excess Helmholtz free energy of 
+a homogeneous Hard-Sphere fluid in terms of the packing fraction $\eta$ 
 \begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
@@
 
-An expression for the excess Helmholtz free energy as a functional of the 
-number density profile $n(\vec r)$ can be found using Fundamental Measure 
-Theory (FMT)~\cite{rosenfeld1989}. FMT was developed for hard sphere fluids, and has been very 
-sucessful at predicting the thermodynamic properties of inhomogeneous hard 
-sphere fluids starting from an expression for the excess Helmholtz free 
-energy obtained from Scaled Partical Theory~\cite{santos2012phi3}. 
-Soft Fundamental Measure Theory (SFMT), which will be addressed later, 
-extends FMT to fluids with soft potentials like that of the WCA fluid.
-
+%for the excess Helmholtz 
+%free energy obtained from Scaled Partical Theory for a homogeneous 
+%Hard-Sphere fluid, 
+Fundamental Measure Theory~\cite{rosenfeld1989} (FMT) 
+can be constructed from Equation~\ref{Fexcess-SPT}~\cite{santos2012phi3},  
+and provides an expression for the excess Helmholtz free energy of an 
+inhomogeneous 
+Hard-Sphere fluid as a functional of the number density profile $n(\vec r)$. 
+%FMT was developed for hard sphere fluids, and has been very sucessful 
+%at predicting the thermodynamic properties of inhomogeneous hard sphere fluids. 
+Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
+is an extension of FMT and provides an expression for the excess Helmoholtz
+free energy of fluids with soft potentials like that of the WCA fluid. 
+It will be helpful to continue looking at FMT before addressing SFMT.
+
+Fundamental Mearsure Theory begins by noticing that 
 %\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
@@
 %\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
-From Equation~\ref{Fexcess-SPT}, a quantity $\zeta_{3}$ can be defined as follows
-\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=n\zeta_{3}}\end{equation}
-where $\zeta_{3}$ is the volume of a sphere. Likewise,
-\begin{equation}{\zeta_{2}=4\pi{R^2}}\end{equation}
-\begin{equation}{\zeta_{1}=R}\end{equation}
-\begin{equation}{\zeta_{0}=1}\end{equation}
-where $\zeta_{2}$ is the surface area of a sphere, and $\zeta_{1}$ is the 
-radius of a sphere. The quantities $\zeta_{3}, \zeta_{2}, \zeta_{1},$ 
-and $\zeta_{0}$ are fundamental geometric measures of a sphere, and they 
-will prove useful by defining a set of weighted densities $n_{i}=n\zeta_{i}$ 
-where $n=\frac{N}{V}$. 
-\begin{equation}\label{n3}{n_{3}=n\zeta_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
-\begin{equation}\label{n2}{n_{2}=n\zeta_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
-\begin{equation}\label{n1}{n_{1}=n\zeta_{1}=\left(\frac{N}{V}\right)R}\end{equation}
-\begin{equation}\label{n0}{n_{0}=n\zeta_{0}=\left(\frac{N}{V}\right)}\end{equation}
-
-\noindent The excess free energy in Eq~\ref{Fexcess-SPT} can be written 
-in terms of the number density $\frac{N}{V}$
-\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
-and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
-\begin{equation}\label{FexfromSPT}{F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V}\end{equation}
+from Equation~\ref{Fexcess-SPT}, a quantity $G_{3}$ can be defined as follows
+\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
+where $G_{3}$ is the volume of a sphere. Likewise, by defining
+quantities
+\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
+\begin{equation}{G_{1}=R}\end{equation}
+\begin{equation}{G_{0}=1}\end{equation}
+where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
+radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
+%reproduced by introducing a set of weighted densities $n_{i}=nG_{i}$.
+\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
+The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
+geometric measures of a sphere, and are what give rise to the name
+Fundamental Measure Theory. Introducing a set of weighted densities $n_{i}=nG_{i}$ 
+given by
+%and they will prove useful by introducing a set of weighted densities 
+%$n_{i}=nG_{i}$ where $n=\frac{N}{V}$. 
+\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
+\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
+\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
+\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
+Equation~\ref{FexfromSPTinthermsofG} can be written as
+%Writing Eq~\ref{Fexcess-SPT} in terms of the number density $n=\frac{N}{V}$
+%\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
+%and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
+\begin{align}\label{FexfromSPT}
+    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
+\end{align}
 which has the form 
-\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
+\begin{align}
+    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
+\end{align}
+or for a spatially varying number density $n(\vec{r})$, 
+\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
 where 
@@
    \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
-   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2} 
+   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
 \end{align}
@@
 The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
-densities given the relation $n_{i}=n\zeta_{i}$, where the weights are 
-given by $\zeta_{0},\zeta_{1},\zeta_{2},\zeta_{3}$ respectively. These 
-weights are valid for determining the excess free energy of a system of 
+densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
+\begin{align}
+    n_{i}=nG_{i}
+\end{align}
+These weights are valid for determining the excess free energy of a system of 
 hard spheres with a homogeneous number density n. But these weights, and 
@@
 allowing for their extention to a system with a spacially varying number 
-density, that is, an inhomogeneous fluid.
-
-Using a backwards Heavyside step function defined as
+density, that is, an inhomogeneous fluid. Using a backwards Heavyside step function defined as
 \begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
 the following expressions are obtained
-\begin{equation}{\zeta_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
-\begin{equation}{\zeta_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
-
-%\begin{equation}{\zeta_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
-\begin{equation}{\zeta_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
-%\begin{equation}{\zeta_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
-\begin{equation}{\zeta_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
-
-The weighted densities can now be expressed as
-\begin{equation}{\color{red} CHECK!~~\color{black}n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}}\end{equation} 
+\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
+\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
+%\begin{equation}{G_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
+\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
+%\begin{equation}{G_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
+\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
+where the weighted densities can now be expressed as
+%CHECK
+\begin{align}
+    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
+\end{align} 
+or for a spatially varying number density $n(\vec{r})$,
+\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
 with weight functions
@@
 \begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
-
 %\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
@@
 \begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
-where the weight functions $w_{2}$ and $w_{3}$ have the relation
+The weight functions $w_{2}$ and $w_{3}$ have the relation
 \begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
+which will be important to keep in tact when moving to SFMT.
 
 The weight functions can be related to the mayer function 
-\begin{equation}{f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1}\end{equation} 
-where r is the distance between the centers of two hard spheres. 
-For a hard sphere fluid the interatomic potential $V(r)$ is given by 
-\begin{displaymath}{V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.}\end{displaymath}
-and so the mayer function for a hard sphere is given by 
+\begin{align}
+     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+\end{align} 
+where r is the distance between the centers of two hard spheres, and 
+$V(r)$ is the interatomic potential between two hard spheres given by 
+\begin{align}
+    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+Evaluating Equation with $V(r)$ given by Equation,  
+the mayer function for a Hard-sphere fluid is found to be
 \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
-The mayer function can be written in terms of weight functions if two 
+which can be written in terms of the weight functions if two 
 additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
-\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}  
-where the vector weights are defined as 
 \begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
 \begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
-
-Taking the derivative of the mayer function in Eq~\ref{f(r)step} will 
-prove useful, and is given by
-\begin{equation}\label{mayerder=w2convolution}{\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}}\end{equation} 
-where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
+The mayor function can then be written in terms of the weight functions~\cite{Hansen}
+\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
+
+Taking the derivative of the mayer function in Eq~\ref{f(r)step} provides 
+another relation that will be important to keep in tact when moving to SFMT.
+\begin{align}\label{mayerder=w2convolution}
+     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
+\end{align} 
+This can easily be checked by evalating the left side for a Hard-sphere fluid as follows,
+%where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
 \begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
 \begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
-Using the general relation for convolution, which is equal to the inverse 
-Fourier Transform of a product of Fourier Transforms 
-\begin{equation}{f \otimes g = \text{IFT[ FT}[f]\text{ x FT[g] }]}\end{equation} 
-with
-\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
-\begin{equation}{\text{Fourier Transform}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
-the right side of Equation~\ref{mayerder=w2convolution} gives
-\begin{equation}{w_2 \otimes w_2 = \text{Inverse Fourier Transform}[e^{-ikR}e^{-ikR}]}\end{equation} 
+and evaluating the right side by using the general relation for convolution which is equal to the Inverse 
+Fourier Transform of a product of Fourier Transforms. 
+\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
+Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
+%\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
+gives
+\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
+and the right side of Equation~\ref{mayerder=w2convolution} becomes
+\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
 \begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
@@
 
-Incorporating into Equation~\ref{FexfromSPT} the two vector weights which 
-were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
-expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
-the excess free energy becomes
-\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
-where
+Tarazona later improved upon the original FMT equations by replacing the
+expression for $\Phi3$ given by Equation~\ref{oldPhi3equation} with the
+more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
+%
+%The final FMT expression for the excess Helmholtz free energy is given by
+%
+%Incorporating the two vector weights which 
+%were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
+%expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
+%into Equation~\ref{FexfromSPT} the excess free energy becomes
+%\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
+%where
 \begin{align}
-\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
-\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
+%\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
+%\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
 \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
 \end{align}
-and the tensor weight is given by
+in which a tensor weight is incoporated given by
 \begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
-
-The expression for the weighted densities can be written in a way that 
-accomodates both homogenious fluids where the number denstiy n is constant, 
-and inhomogeneous fluids where the number density becomes a function of 
-the position $n(\vec{r})$.
-\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
+All the other FMT equations remain the same.
+
+%The expression for the weighted densities can be written in a way that 
+%accomodates both homogenious fluids where the number denstiy n is constant, 
+%and inhomogeneous fluids where the number density becomes a function of 
+%the position $n(\vec{r})$.
+%\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
 %The equatons above were derived for a uniform fluid where the number density n is constant, 
 %but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
-For a spatially varying number density $n(\vec{r})$, the excess Free 
-Energy becomes a functional of the number density profile $n(\vec{r})$ :
-\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
-
-The equations developed so far have been for a hard sphere fluid. All of 
-the weight function equations, as well as the equation for the deriviative 
-of the mayer function, can be expressed in terms of $w_{2}(r)$ which consists 
-of a delta function characteristic of hard spheres. Soft Fundamental Measure 
-Theory extends the Fundamental Measure Theory discussed so far to 
-fluids with soft potenials by keeping the relationships between the weight 
-functions, and which the weight functions satisfy the same, but replacing 
-$w_{2}(r)$ with a weight function appropriate for a soft sphere potential. 
-A new set of weight functions can then be derived for a system of soft 
-spheres which is what we do in forming our functional for the 
-Weeks-Chandler-Anderson fluid as detailed in the chapter on Methods.
+%For a spatially varying number density $n(\vec{r})$, the excess Free 
+%Energy becomes a functional of the number density profile $n(\vec{r})$ 
+%\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
+
+The FMT equations developed so far have been for a Hard-sphere fluid.
+All of the weight functions, the relations between them, and the equation 
+for the deriviative of the mayer function, can be expressed in terms of 
+$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function 
+characteristic of hard spheres. 
+Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
+with soft potenials by keeping the relationships between the weight 
+functions the same, but replacing $w_{2}(r)$ with a weight function 
+appropriate for a soft sphere potential. This is what we do to
+%A new set of weight functions can then be derived for a system of soft spheres  
+form a functional for the 
+Weeks-Chandler-Anderson fluid as detailed in the next chapter on Methods.
 
@@
     %\includegraphics[height=6cm]{GaussandW2_actual.png}
-    \includegraphics[height=7cm]{weightandGuass1.pdf}
+    \includegraphics[height=7cm]{figs/weightandGauss1.pdf}
     %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
@@
     %\includegraphics[height=6cm]{GaussandW2_thinkas.png}
-    \includegraphics[height=7cm]{weightandGuass2.pdf}
+    \includegraphics[height=7cm]{figs/weightandGauss2.pdf}
     \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 

### 2020-11-24 papers/thesis-kirstie/thesis.tex
more edits

@@
   erf model in which we have incorporated temperature dependent
-  parameters to
-  \davidsays{fit the results to those predicted by Monte-Carlo simulations}.
+  parameters to fit the results to those predicted by Monte-Carlo simulations.
   Our functional was successful in predicting the freezing of the WCA fluid 
@@
 The WCA potential is described by the equation: 
-\begin{equation}{V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.}\end{equation}
+\begin{align} \label{eq:VWCA}
+    V_{WCA}=\left\{\begin{array}{rcl} {4\epsilon{\left[\left(\frac{\sigma}{r}\right)^{12} - \left(\frac{\sigma}{r}\right)^6 \right]}+\epsilon} & \mbox{for} & 0<r<{2R} \\ 0 & \mbox{for} & r>2R \end{array}\right.
+\end{align} 
 As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
@@
 
-\newpage
+%\newpage
 \section{The Theory of Freezing and Maxwell's Double Tangent Construction}
@@
 
-%For each particular temperature, the pressure of the liquid and the 
-%solid are the same at the point of transistion. 
 \begin{figure}[h!]
@@
 
-%In a process called Maxwell's Double Tangent construction, two curves are 
-%used to find the densities of the liquid and the solid at the point of transistion. 
-%Figure \ref{fig:MaxwellDT} shows a curve for the liquid, and a curve for 
-%the solid both at a fixed temperature. Since the pressure of the liquid 
-%is the same as the pressure of the solid at the point of transistion, the 
-%density of the liquid and the density of the solid at the point of freezing 
-%can be found from identifying the inverse densties at which the tangents 
-%to each curve have the same slope.%
 Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-     % FIXME\includegraphics[height=7cm]{plot-pressure_Fig1.pdf}
     \caption{Identifying densities at the solid-liquid transition using 
@@
 To facilitate finding the pressure at the phase transition, 
-%, which occurs at the points which have the same slope,
 it is helpful to consider the Gibbs free energy given by G=F+PV. 
@@
 Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
-per atom  $g$ (which is also the chemical potential):  
-%From the Gibbs free energy an expression can be derived which relates  ....:
-
-%\begin{displaymath}{F=-PV+G}\end{displaymath} 
+per atom  $g$ (which is also the chemical potential)  
 \begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
@@
 \begin{align}{y=mx+y_o}\end{align}
-%\begin{displaymath}{G_1=G_2}\end{displaymath} 
-%\begin{displaymath}{F_1+PV_1=F_2+PV_2}\end{displaymath} 
-%\begin{displaymath}{\frac{F_1}{N}+P\frac{V_1}{N}=\frac{F_2}{N}+P\frac{V_2}{N}}\end{displaymath} 
 and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
@@
 liquid and the solid at a given temperature.
+
 \begin{figure}[h!]
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
-     %\includegraphics[height=7cm]{plot-pressure_Fig2.pdf}
     \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
@@
   \end{figure}
-%CURVE FITTING! - put in methods/analysis
 
@@
 procedure for various temperatures.
-%The lowest density the crystal exhibts and the highest density the fluid 
-%exhibits at the phase transistion can then be identified for each temperature.
 
@@
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
-    % \includegraphics[height=7cm]{plot-pressure_Fig3.pdf}
     \caption{The liquid and solid densities at the point of transistion are 
@@
   \end{figure}
-%One way of implementing this practically is to determine what the Helmholtz 
-%free energy would be for both the liquid and solid phases at a particular 
-%temperature and density. Whichever phase has the lower Helmholtz free 
-%energy will be the phase that the substance will be in.
 
@@
 
-%The reason this works is because cDFT satisfies the conditions of the 
-%variational principle, the same mathematical principle used to determine 
-%that the shortest distance between two points is a straight line by 
-%varying the path between the two points until the distance is minimized.
-
 A description of cDFT often begins with considering the Grand Free 
@@
 \begin{equation}\Phi=F-\mu{N}\end{equation}
-%\begin{equation}\int\frac{N}{V}\frac{V_{ext}}{N}dV\end{equation}
 which can then be expressed in terms of a varrying number density 
 profile $n(\vec r)$
-%\begin{equation}\Phi=F-\int n(\vec{r})\phi{(\vec r)}d\vec{r} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu{N}\end{equation}
 \begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
@@
 \begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}
-%If $\mu=0$ and $V_{ext}=0$, then the Grand free energy $\Phi$ reduces 
-%to the Helmholz free energy $F$. %The second and third terms in Eq~\ref{GrandFE} 
-%are both functionals of $n(\vec{r})$, so it remains to show that $F$, and 
-%thus $\Phi$, is a functional of $n(\vec{r})$.
 
@@
 of the number density profile $F[n(\vec r)]$ that is minimized at equilibrium. 
-%In this paper a system with a fixed temperature, volume, and number of 
-%particles will be examined in which case it is the Helmholtz energy as a 
-%functional of the number density profile $F[n(\vec r)]$ that is of interest. 
 More specifically, in this paper, the Helmholtz free energy per particle, 
@@
 \begin{equation}\label{usetoshowmin}df_N=d\left(\frac{U}{N}\right)-d\left(\frac{TS}{N}\right)\end{equation}
-%\begin{equation}df_N=d\left(\frac{U}{N}\right)-\frac{T}{N}dS-\frac{S}{N}dT-TS d\left(\frac{1}{N}\right)\end{equation}
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
@@
 particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
-%Thus, the Helmholtz free energy will be minimized in a system that has 
-%a temperature, volume, and number of particles that are fixed.
-%\begin{equation}f_N(T,V, \frac{1}{N})\mbox{~~~~or~~~~}f_V(T,n)\end{equation}
 \begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
@@
 \begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation}
-%the equation can be written
-%\begin{equation}{Z=\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}e^\frac{-(\frac{|\vec{p}_2|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)\left(\frac{1}{V^N}\int{...}\int{dQ}e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}\right)}\end{equation}
 the partitian function can be expressed as
@@
 \begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
-\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} The ideal partitian function $Z_{ideal}$ is associated with an ideal gas which exhibits no interaction between the atoms - a condition approached by a gas at low density where the atoms rarely encounter one another.
-The excess partitian function $Z_{excess}$ accounts for the interaction of all the atoms according to their interatomic potential.
-The free energy can also be broken into two parts using Equation~\ref{Zmultiply} giving
-%\begin{equation}{F=-k_{B}T\ln(Z)}\end{equation}
+\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} 
+The ideal partitian function $Z_{ideal}$ is associated with an ideal gas 
+which exhibits no interaction between the atoms - a condition approached 
+by a gas at low density where the atoms rarely encounter one another. 
+he excess partitian function $Z_{excess}$ accounts for the interaction 
+of all the atoms according to their interatomic potential. The free energy 
+can also be broken into two parts using Equation~\ref{Zmultiply} giving
 \begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
@@
 \begin{equation}{V(\vec{r_1},\vec{r_2},...\vec{r_N})=\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}\end{equation} 
-%\begin{displaymath}{=V_{12}(\vec{r_1},\vec{r_2})+V_{13}(\vec{r_1},\vec{r_3}) + V_{14}(\vec{r_1},\vec{r_4}) + ...}\end{displaymath}
-%\begin{displaymath}{+ V_{23}(\vec{r_2},\vec{r_3})+V_{24}(\vec{r_2},\vec{r_4})+V_{25}(\vec{r_2},\vec{r_5})+ ...}\end{displaymath} 
-%\begin{displaymath}{+ V_{34}(\vec{r_3},\vec{r_4})+V_{35}(\vec{r_3},\vec{r_5})+V_{36}(\vec{r_3},\vec{r_6})+...}\end{displaymath}
 Using this potential, and setting the external potential 
@@
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
-%\begin{equation}{\exp^{-V_{ij}(\vec{r_i},\vec{r_j})}\approx 1+f_{ij}(\vec{r_i},\vec{r_j})}\end{equation}%    
-%$\exp^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}...$.%  
 The function $f_{ij}$ is called a Mayer function, and is given by
@@
 two atoms in the pair. 
-%Making a change of variables $\vec{r}=\vec{r}_i-\vec{r}_j $ for each pair we get 
-%\begin{equation}f_{ij}(\vec{r_i}-\vec{r_j})=f(\vec{r})=e^{-\frac{V(\vec{r})}{k_BT}}-1\end{equation} 
 With a careful counting of the resulting number of such pairs, the excess 
 free energy can then be written~\cite{schroeder} as
-%\begin{equation}{F_{excess}=-k_BT\ln\left(1+\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}+\frac{1}{2}\left(\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}\right)^2+ ...\right)}\end{equation}
-%where in the thermodynamic limit (large N), $N-1, N-2, ...\approx{N}$. 
-%Written in terms of an exponential $e^x=1+x+\frac{x^2}{2!}+\frac{x^3}{3!}+ ...$ 
-%for a homogeneous density n=$\frac{N}{V}$ this becomes
-%\begin{equation}{F_{excess}=-k_BT\ln{\left(e^{\frac{1}{2}\frac{N^2}{V}\int{f(r)}{d\vec{r}}}\right)}}\end{equation} 
-%Thus the free excess energy per atom in the low-density limit is given by 
 \begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
@@
 as the second virial expanision coefficient typically called $B_2$. 
-%where $x=\frac{1}{2}\frac{N(N-1)}{V^{2}}\int{...}\int{dQ}{f_{12}}$.% 
 It is important that the expression for the excess Helmholtz free energy 
 correctly reduces to this result in the low-density limit. 
-%Although we will not be restricting ourselves to the low-density limit, 
-%it is important that the theory we use correctly reduces to this same 
-%result in the low-density limit. 
 Outside of the low density limit, it is quite difficult to evaluate the 
@@
 
-%Equations~\ref{eq:Fideal} and \ref{eq:Fexcess} show that the Helmholtz free energy 
-%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%is a function of the number density n. For a spatially varying number density, 
-%as would be expected for a crystal, the free energy becomes a functional of 
-%the number density profile $n(\vec{r})$.
-
-%According to classical Density Functional Theory,
-
-%A number density profile is a time-invariant average 
-%of the spacial distribution of the particles that make up the fluid, and
-%There will be one, and only one 
-%that minimizes the Helmholtz free energy.
-%The minimum free energy, 
-%is the value of the Helmholtz free energy that
-%the system will have when it is in equilibrium at that temperature, 
-%and its corresponding number density profile 
-%are the equilibrium values the system will have at that temperature.
-%is the equilibrium number density profile. 
-%
 Once an expression for the Helmholtz free energy as a functional of the
@@
 found by varying the number density profile
-%can be varied 
 through all possible configurations at some fixed temperature
@@
 In this section 
-%To form the functional for the Helmholtz free energy, expressions for 
 a functional for the Helmholtz free energy was formed by finding 
@@
 Scaled Particle Theory (SPT).  
-%For the Hard Sphere fluid, there is another way to form an expression for 
-%the excess Helmholtz free energy in terms of a homogeneous number density 
-%and, that is to use Scaled Particle Theory (SPT)~\cite{ReissSPT}.
 Scaled Particle Theory seeks a general expression for the excess chemical 
@@
 to find the excess Helmholtz free energy according to the relation
-%\begin{equation}F_{ex}=\int{\mu_{ex}dN}\end{equation}
-%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}\end{equation}
 \begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
@@
 Formula~\cite{Hansen}.
-%\begin{displaymath}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~~}\end{displaymath}
 \begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
@@
 In this case Equation~\ref{widoms-insertion-formula} becomes
-%\begin{equation}\mu_{excess}=-k_BT\ln(1-n\frac{4\pi}{3}R_{cavity}^3) &  \mbox{for} & R_{\cavity}=R\end{equation}
 \begin{align} \label{muexcess}
@@
 \end{align}
-%where the radius of the cavity $R_{cavity}$ is equal to the radius of a 
-%hard sphere, R, and $<e^{-\beta{V_{diff}}}>$ is evaluated using the 
-%probablities that the cavity is either filled or not filled~\cite{Hansen}.
-%For cavities that are much larger than one of the hard spheres, 
 Forming a more general result by requiring continuity of the functions, 
@@
 written for fluids with a cavitiy radius 
-%$\operatorname{R}_{cavity}$ 
 equal to, larger than, or very 
@@
 
-%for the excess Helmholtz 
-%free energy obtained from Scaled Partical Theory for a homogeneous 
-%Hard-Sphere fluid, 
 Fundamental Measure Theory~\cite{rosenfeld1989} (FMT) 
@@
 Hard-Sphere fluid as a functional of the number density profile $n(\vec r)$. 
-%FMT was developed for hard sphere fluids, and has been very sucessful 
-%at predicting the thermodynamic properties of inhomogeneous hard sphere fluids. 
 Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
@@
 Fundamental Mearsure Theory begins by noticing that 
-%\begin{equation}{\eta = \frac{NR_{3}}{V}}\end{equation}
-%\begin{equation}{\eta = \left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}}\end{equation}
-%\begin{equation}{\eta = n\frac{4}{3}\pi{R^3}=nR_{3}=n_{3}}\end{equation}
 from Equation~\ref{Fexcess-SPT}, a quantity $G_{3}$ can be defined as follows
@@
 radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
-%reproduced by introducing a set of weighted densities $n_{i}=nG_{i}$.
 \begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
@@
 given by
-%and they will prove useful by introducing a set of weighted densities 
-%$n_{i}=nG_{i}$ where $n=\frac{N}{V}$. 
 \begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
@@
 Equation~\ref{FexfromSPTinthermsofG} can be written as
-%Writing Eq~\ref{Fexcess-SPT} in terms of the number density $n=\frac{N}{V}$
-%\begin{equation}{F_{excess}=k_{B}T\left(-\frac{N}{V}ln(1-\eta)+\frac{N}{V}\frac{3\eta}{1-\eta}+\frac{N}{V}\frac{3{\eta}^2}{2(1-\eta)^2}\right)V}\end{equation}
-%and using the expressions in Equations~\ref{n3} through~\ref{n0}, this becomes
 \begin{align}\label{FexfromSPT}
@@
 \begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
-%\begin{equation}{G_{1}=\frac{R_{2}}{4\pi{R}}=R}\end{equation}
 \begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
-%\begin{equation}{G_{0}=\frac{R_{2}}{4\pi{R}^2}=1}\end{equation}
 \begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
 where the weighted densities can now be expressed as
-%CHECK
 \begin{align}
@@
 \begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
-%\begin{equation}{w_{1}=\frac{w_{2}}{4\pi{R}}}\end{equation}
 \begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
-%\begin{equation}{w_{0}=\frac{w_{2}}{4\pi{R}^2}}\end{equation}
 \begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
@@
 This can easily be checked by evalating the left side for a Hard-sphere fluid as follows,
-%where, for a hard sphere fluid the left side of Equation~\ref{mayerder=w2convolution} gives
 \begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
@@
 Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
-%\begin{equation}{w_2(r)=\delta(r-R)}\end{equation} 
 gives
@@
 more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
-%
-%The final FMT expression for the excess Helmholtz free energy is given by
-%
-%Incorporating the two vector weights which 
-%were added to satisfy Equation~\ref{mayer_deconvolution}, and an updated 
-%expression for $\Phi_3$ developed by Tarazona~\cite{tarazonaphi3, santos2012phi3}, 
-%into Equation~\ref{FexfromSPT} the excess free energy becomes
-%\begin{equation}{F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V}\end{equation}
-%where
 \begin{align}
-%\Phi_1 &= -n_{0}ln(1-n_{3}) \label{eq:Phi1}\\
-%\Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \label{eq:Phi2} \\
 \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
@@
 
-%The expression for the weighted densities can be written in a way that 
-%accomodates both homogenious fluids where the number denstiy n is constant, 
-%and inhomogeneous fluids where the number density becomes a function of 
-%the position $n(\vec{r})$.
-%\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-%The equatons above were derived for a uniform fluid where the number density n is constant, 
-%but they can be extended for inhomogenious fluids where the number density becomes a function of r, $n(\vec{r})$.
-%For a spatially varying number density $n(\vec{r})$, the excess Free 
-%Energy becomes a functional of the number density profile $n(\vec{r})$ 
-%\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation}
-
 The FMT equations developed so far have been for a Hard-sphere fluid.
@@
 for the deriviative of the mayer function, can be expressed in terms of 
-$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function 
-characteristic of hard spheres. 
+$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function.
+The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+Equation~\ref{w2_w3_relation}, which is a step function for the Hard-sphere fluid.
 Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
@@
 appropriate for a soft sphere potential. This is what we do to
-%A new set of weight functions can then be derived for a system of soft spheres  
 form a functional for the 
@@
 \section{The Overall Approach}
-%
-%The goal of this project is to generate phase diagrams for a 
-%Weeks-Chandler-Anderson model fluid using a computer program. 
-%To do this we find the Helmholtz free energies that the fluid
-%would have at a given temperature and density in a  
-%liquid state, and in a crystalline solid state. 
-%%Once the Helmholtz free energy for both the crystal and liquid states of 
-%%the fluid are found for a given temperature and density, they can be 
-%%compared to find the resulting state of the fluid. 
-%Outside the region of 
-%coexistence which exists between the liquid and crystal states, 
-%the fluid will be completely in a crystalline state if the 
-%crystal free energy is lower than the liquid free energy, and it will be 
-%completely in a liquid state if the liquid free energy is lower than 
-%the crystal free energy. Maxwell's Double Tangent Construction is used 
-%to determine the actual number densities of the crystal and liquid states
-%at the boundaries of the coexistance region for a given temperature, 
-%and construct phase diagrams.
-
-%%In general, if at some number density and temperature the crystal free energy 
-%%is lower than the homogeneous free energy of the liquid, then the fluid will 
-%%be in the crystalline state. If the homogeneous liquid free energy is lower 
-%%than the crystal free energy, then the fluid will be in the liquid state. 
-%
-%%To find the equilibrium Helmholtz free energies we use classical Density 
-%%Functional Theory (cDFT) in conjunction with Soft Fundamental Measure Theory (SFMT).
-%%%the corresponding equilibrium number density profile. 
-%%In particular, 
-%
-%In general, the Helmholtz free energy can be written as a functional of 
-%the number density profile $n(\vec{r})$ which describes the way the average number 
-%density varies spatially. 
-%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation} 
-%Classical Density Functional Theory is used to find the equilbrium 
-%Helmholtz free energy of the system at a given temperature and density
-%by varying the number density profile $n(\vec{r})$ until the free energy 
-%is minimized. The minimum free energy, and its corresponding number 
-%density profile are then the equilibrium values for that temperature 
-%and number density. 
-%
-%The Helmholtz free energy consists of two parts: the ideal Helmholtz free energy
-%and the excess Helmholtz free energy.
-%The ideal Helmholtz free energy is the free energy the fluid would have in
-%the absence of any interactomic potential energy $V(r)$ between the spheres, 
-%or atoms that make up the fluid. The excess Helmholtz free energy is that 
-%additional energy which results from the presence of an interatmoic potential 
-%energy $V(r)$ between the spheres.
-%%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation} 
-%%For a spatially varying number density, as would be expected for a crystal, 
-%%the free energy is a functional of the number density profile $n(\vec{r})$  
-%%The Helmholtz free energy excess free energy is then added to the ideal free energy calculated 
-%%analytically to find the resulting free energy for the system. 
-%Soft Fundamental Measure Theory is used to form the expression 
-%for the excess Helmholtz free energy without having to take explicitly 
-%into account the interaction each sphere, or atom, has with every other 
-%sphere.
 
@@
 
-  %\begin{figure}[h!]
-    %\centering
-    %% FIXME \includegraphics[height=4.5cm]{Simple_Cell.png}
-    %%\caption{A Simplified Crystal Cell}
-    %\label{fig:simplecell}
-  %\end{figure}
-
-%\textcolor{red}{This cell is rather complicated for the purposes of this discussion, so 
-%consider instead a simplified crystal cell with one atom in its center 
-%as shown in Figure \ref{fig:simplecell}. The Helmholtz free energy is 
-%minimized for a fixed volume (and temperature and number of particles).  
-%A fixed cubical volume called the ``bold-box" volume, $V_{bold-box}$ will 
-%be used for illustration. The number density n for a collection of 
-%simplified cells is simply 1 atom per fixed bold-box volume.
-%\begin{displaymath}{n = \frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{displaymath} 
-%It is easier to think of 1 atom per ``box" (a reduced number density) 
-%than 1 atom per some length$^3$ (the actual number density). Reduced number 
-%density is given by $n^*$ and is equal to the number of atoms in a ``box" 
-%of size $\sigma^3$.
-%\begin{displaymath}{n^*=n\sigma^3}\end{displaymath} If $V_{bold-box}$  
-%has a volume of $\sigma^3$, the reduced number density is just the number 
-%of atoms in the fixed bold-box volume. 
-%\begin{displaymath}\text{Reduced Number Density of the simple cell = 1 atom}\end{displaymath} 
-%}
 In a large collection of cells, the atom at each lattice site
@@
 statistically distrubuted over a collection of cells. 
-%In 3-dimensions, the Gaussian function takes on the form
-%\begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int{e}^{-\frac{r^2}{2\sigma^2}}d\vec{r}}\end{equation} 
 It is important to note that the number density does not change when introducing a number 
 density profile. There are still the same number of atoms over some fixed volume. 
-%There is still a total of one atom per ``bold box" volume.
-%\begin{equation}{n=<n(\vec{r})>=\frac{1}{V_{bold-box}}\overbrace{\int_{bold-box}{n(\vec{r})}{d\vec{r}}}=\frac{N}{V}=\frac{1~\text{atom}}{V_{bold-box}}}\end{equation}
 \begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
@@
 It may be that not all lattice sites are occupied. Vacancies can be taken 
-into account while keeping %the volume of the bold box, and thus 
+into account while keeping 
 the overall number density the same by 
@@
 \begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath}
-%\begin{align}
-  %n(\vec{r})=(1-f_v)\frac{\exp^{-\frac{{\vec{r}}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}
-%\end{align}
 
@@
 for all these number density profiles $n(\vec{r})$. 
-%There is still a total of one atom per ``bold box" volume. 
-%This prodcedure is used only to find the crystal free energy as the number density 
-%profile for a liquid is homogeneous, as shown in Figure \ref{fig:homogen_denisty}.  
 A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.
@@
 
-%Classical Density Functional Theory is used to find the equilbrium 
-%Helmholtz free energy of the system at a given temperature and density
-%by varying the number density profile $n(\vec{r})$ until the free energy 
-%is minimized. The minimum free energy, and its corresponding number 
-%density profile are then the equilibrium values for that temperature 
-%and number density.
-
-%The number density profile is varied by varying gw and fv, and the 
 For a given combination of temperature and number density, a suite of number 
@@
 for that temperature and number density.
-%The Helmholtz free energy is minimized for a system with a fixed temperature, 
-%volume, and number of particles. Setting the number of particles to one, 
-%the volume can be found from the number density.
 The process is repeated to find the free energies at other combinations 
@@
 used to construct phase diagrams.
-%compared to find the resulting state of the fluid. 
 Outside the region of coexistence which exists between the liquid and 
@@
 \section{Our Functional for the Excess Helmholtz free energy}
-In order to apply classical Density Functional Theory, the Helmholtz free 
-energy must be found. Computing the ideal portion of the Helmholtz free 
-energy is straightforward, but computing the excess portion of the 
-Helmholtz free energy is difficult since it depends on the interatomic 
-potential energy $V(r)$ between each sphere with all the other spheres, 
+In order to apply classical Density Functional Theory, an expression 
+for the Helmholtz free energy as a functional of a number density profile 
+must be found. Computing the ideal portion of the 
+Helmholtz free energy is straightforward, but computing the excess portion 
+of the Helmholtz free energy is difficult since it depends on the interatomic 
+potential energy $V(r)$ that exists between each sphere with all the other spheres, 
 as is shown in Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
-Soft Fundamental Measure Theory is a method
+Soft Fundamental Measure Theory (SFMT) is a method
 developed to form an expression for the excess Helmholtz free energy without
@@
 but it must be taylored to fit a particular fluid\cite{schmidt1999density}.  
-%The excess Helmholtz free energy is particular to the fluid being modeled, 
-%as it depends on the interatomic potential V(r). 
-It is based on the expression for the excess Helmholtz 
-free energy of a homogeneous Hard-sphere fluid derived 
-from Scaled Particle Theory, and given in Equation~\ref{FexfromSPT}. 
-The resulting functional 
-%(given in Equations~\ref{eq:F_ex-n-of-r} - \ref{eq:Phi3} and repeated here) 
+Our SFMT functional is taylored to fit a WCA fluid, and 
 is written in general form as 
@@
 \end{align} 
-and Tarazona's tensor expression~\cite{tarazonaphi3, santos2012phi3} is used 
-for $\Phi_3$. The weigthed densities $n_0, n_1, n_2,$ and $n_3$ are found from
+The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
 \begin{align}\label{eq:numdenprofile}
@@
 \end{align}
-%where the number density profile $n(\vec r)$ within the integral will be 
-%given in our program by a Gaussian function centered at a lattice point, 
-%and $w(r)$ is a weight function.
 where the number density profile $n(\vec r)$ within the integral is given 
-by a Gaussian function centered at a lattice point.
-%\begin{align}
-  %n(\vec r)=(1-f_v)\frac{\exp^{-\frac{r^2}{2{\sigma}^2}}}{\left(\sqrt{2\pi}\sigma\right)^3}
-%\end{align}
-%
-There are four scaler weight functions, two vector weight functions, and 
+by Gaussian functions centered at each lattice point.
+There are four scalor weight functions, two vector weight functions, and 
 one tensor weight function all of which 
-%The weight functions $w_0(r), w_1(r), w_2(r)$ 
 depend on the expression for $w_3(r)$ 
@@
 between the spheres, and which defines the particular fluid being modeled.
-%FMT models the Hard-sphere fluid, and SFMT models fluids with soft potentials.
 We introduce our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ 
@@
   w_{1}(r) &=\frac{w_{2}}{4\pi{r}} \\
-  %w_{2}(r) &=-\frac{\partial{w_3(r)}}{\partial{r}}   \\
-  %w_{3}(r) &=~?
   w_2(r) &=-\frac{\partial{w_3(r)}}{\partial{r}} \\
@@
   w_3(r) &=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{\frac{\Xi}{\sqrt{2}}}\right)\right]  
-  %\vec{w_{1}} &= w_{1}\frac{\vec{r}}{r}   \\
-  %\vec{w_{2}} &= w_{2}\frac{\vec{r}}{r}  \\
-  %\overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)
-\end{align}
-%
+\end{align}
 The vector weights are given by 
@@
 \begin{equation}\label{eq:tensorweight-methods}\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)\end{equation}
-%
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_actual.png}
-    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    %view of $w_2$ centered at $\vec{r}'$ is shown in purple. Here, the weight function 
-    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    %is generated randomly according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure} 
-%We use Soft Fundamental Measure Theory to form a functional for the excess 
-%Helmholtz free energy appropriate for the WCA fluid by introducing our 
-%
-%Equation 2.94  general form of Fex for n(r)    
-%\ref{eq:F_ex-n-of-r}, 
-%phis 1-3   2.90, 2.91, 2.92
-%\ref{eq:Phi1}, 
-%\ref{eq:Phi2}, 
-%\ref{eq:Phi3}, 							
-%weights 2.73, 2.74, 2.89  2.79, 2.80  replace-> 2.71, 2.72		
-%\ref{eq:w0}, 
-%\ref{eq:w1}, 
-%\ref{eq:w2}, 
-%\ref{eq:w3}, 
-%\ref{eq:w_v2}, 
-%\ref{eq:w_v1}, 
-%\ref{eq:tensorweight}, 
-%%weight relations 2.75, 2.81
-%\ref{w2_w3_relation}, 
-%\ref{mayerder=w2convolution},
 
 The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
-Schmidt's proposed erf model~\cite{schmidt2000fluid} formed to model a steep, 
+Schmidt's proposed erf model~\cite{schmidt2000fluid} designed to model a steep, 
 soft potential in which he set
 \begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
-where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
+subject to the boundary conditions $w_3(0)\approx{1}$ 
 and $w_3(\infty)=0$. 
-%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
 Our functional is  created by replacing Schmidt's parameters $\sigma$ 
 and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
-\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
-%
-The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
-temperature dependence that enters in through the expression for the 
-error function potential $V_{erf}(r)$ which 
-%is used to approximate the non-temperature dependent WCA potential that 
-%is more difficult to implement computationally. The erf potential 
-emerges from the relationships
-that the weight functions must satisfy.
-%potential at temperatures of interest will be used instead of the WCA potential. 
-Using the relation
-%Equation~\ref{w2_w3_relation}, shown again below, 
-\begin{align}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{align}
+The two parameters $\alpha$ and $\Xi$ make temperature-dependent error 
+function potential $V_{erf}(r)$ approximate in the region of interest 
+of the temperature-independent WCA potential which is difficult 
+to implement computionally. $\Xi$ is determined by setting the second virial 
+coeficeint $B_2$ of the error function potential $V_{erf}(r)$ and of the
+WCA potential equal.  
+The error function potential emerges from $w_{3}(r)$, and the relationships 
+in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
+Using Equation~\ref{w2_w3_relation},
+\begin{displaymath}
+     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+\end{displaymath}
 and solving for $w_2(r)$
@@
 Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
-\begin{align}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{align}
+\begin{displaymath}
+    \frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}
+\end{displaymath}
 gives
@@
 given by
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
+\begin{align}  \label{eq:VerfMayerfunction}
+     f(r)=-\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)\right]
+\end{align} 
 Combining this result with the definition of the Mayer function 
 \begin{align}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{align} 
-%This gives an interatomic potential of \begin{equation}{V_{erf}=-k_BT\ln(f(r)-1)}\end{equation} 
 gives an expression for the for the error function potential  $V(r)=V_{erf}(r)$ 
-\begin{equation}
+\begin{equation} \label{eq:Verf}
 	V_{erf}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
@@
 
-%<<<<<<< HEAD
-%\davidsays{Maybe say this before presenting the weights? But also, I don't think
-%there is any benefit to introducing $a$ and $\sigma$ here.} 
-The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
-Schmidt's proposed erf model~\cite{schmidt2000fluid} in which he set
-\begin{equation}
-  w_3(r)=\frac{1}{2}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]
-\end{equation} 
-where $w_3(r)$ is subject to the boundary conditions $w_3(0)\approx{1}$ 
-and $w_3(\infty)=0$. 
-%\color{red}These boundary conditions are satisfied for $\Xi{\sqrt{2}<<\alpha}$, or $k_BT << 2.21$. \color{black} 
-Our functional is  created by replacing Schmidt's parameters $\sigma$ 
-and $a$ with temperature-dependent parameters $\alpha(T)$ and $\Xi(T)$.
-\begin{equation}\sigma = \alpha(T)\text{~~~~~~~and~~~~~~~} a = \Xi(T)\end{equation}
-
-The two parameters $\alpha$ and $\Xi$ are used to compensate for the 
-temperature dependence that enters in through the expression for the 
-error function potential $V_{\operatorname{erf}}(r)$ which is used to approximate the
-non-temperature dependent WCA potential that is more difficult to 
-implement computationally. The erf potential emerges from the relationships
-that the weight functions must satisfy\davidsays{In order for what?}.
-%potential at temperatures of interest will be used instead of the WCA potential. 
-Using Equation~\ref{w2_w3_relation}, shown again below, 
-\begin{align}
-  w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
-\end{align}
-and solving for $w_2(r)$
-\begin{align}
-  \frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)
-\end{align} 
-we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
-Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
-\begin{displaymath}{\frac{df(r)}{dr}=\int_{-\infty}^{\infty}{w_2(r')w_2(r-r')dr'}}\end{displaymath}
-gives
-\begin{equation}{\frac{df(r)}{dr}=\frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r-\alpha}{\Xi}\right)^2}}\end{equation} 
-which can be integrated to obtain the Mayer function
-\begin{equation}{f(r)=\int_{\infty}^r{ \frac{1}{\Xi\sqrt{\pi}}e^{-\left(\frac{r'-\alpha}{\Xi}\right)^2}{dr'}}}\end{equation} 
-The Mayer function for the error function potential $V_{\operatorname{erf}}(r)$ is thus 
-given by
-\begin{equation}{f(r)=-\frac{1}{2}\left[1-\operatorname{\operatorname{erf}}\left(\frac{r-\alpha}{\Xi}\right)\right]}\end{equation} 
-Combining this result with the definition of the Mayer function \begin{displaymath}f(r)=e^{-\frac{V(r)}{K_BT}}-1\end{displaymath} 
-%This gives an interatomic potential of \begin{equation}{V_{\operatorname{erf}}=-k_BT\ln(f(r)-1)}\end{equation} 
-gives an expression for the for the error function potential  $V(r)=V_{\operatorname{erf}}(r)$ 
-\begin{equation}
-	V_{\operatorname{erf}}(r)=-k_BT\ln\left[\frac{1}{2}\left(\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)+1\right)\right]
-\end{equation}
-
-Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, with a width determined 
-by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off. 
-$\Xi$ is related to the width of the functions which 
-become wider in general as the temperature increases. 
-The radial distance 
-to the peak is related to $\alpha$, and decreases as the temperature 
-increases, that is, the shells shrink, as does 
-the sphere.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{weight_functions.pdf}
-    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2.}
-    \label{fig:weight_functions}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W2.png}
-      % \includegraphics[width=5cm]{figs/w2-ring}
-      \caption{A crossectional view of $w_2$}
-    \end{subfigure} 
-    \begin{subfigure}[b]{0.4\linewidth}
-      \includegraphics[height=5cm]{W3.png}
-      \caption{A crossectional view of $w_3$}
-    \end{subfigure} 
-    \caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    \label{fig:W0andW3}
-    \end{figure}
-
-Expressions for $\alpha(T)$ and $\Xi(T)$  can be found as follows. 
 The parameter $\alpha(T)$ is derived by setting 
-the erf potential equal to the WCA potential when $r=\alpha$, i.e. at the
-maximum of $w_2(r)$,
-% \begin{equation}{V_{\operatorname{erf}}(r=\alpha)=V_{WCA}(r=\alpha)}\end{equation} 
-which gives us
+the error function potential $V_{erf}(r)$ given by Equation~\ref{eq:Verf} 
+equal to the WCA potential given by Equation~\ref{eq:VWCA} when $r=\alpha$. 
+This gives
 \begin{equation}\label{alphaT}
@@
     \centering
-    \includegraphics[height=9cm]{plot_alpha.pdf}
+    \includegraphics[height=8cm]{plot_alpha.pdf}
     \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
@@
 
-We determine value of $\Xi$ at a given temperature $T$,
+We determine the value of $\Xi$ at a given temperature $T$,
 by equating the second virial coefficient evaluated for the error function 
@@
 \begin{equation}B_{2\operatorname{erf}}(T) =B_{2WCA}(T)\end{equation}
-\davidsays{I think we'd benefit from cutting much of the following, since you
-show the standard equations, but skip the bit where we put in the pair potential,
-which is the tricky step.}
 The second virial coefficient can be derived from Euler's equation for 
-the internal energy of a system, 
-\begin{equation}U=TS-PV+\mu{N}\end{equation}
-using the relations (discrete versions are shown here for clarity)
-\begin{equation}S=-k_B\sum_{i=0}^\infty{P_i\ln{P_i}}\end{equation}
-\begin{equation}U=\langle E\rangle=\sum_{i=0}^\infty{P_iE_i}\end{equation}
-\begin{equation}N=\langle n\rangle=\sum_{i=0}^\infty{P_in_i}\end{equation}
-which, for a canonical ensemble, gives
-\begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Expressing the Helmholtz free energy in terms of the pair potential 
-$V(r)$ results in an equation of state expressed in terms of the virial expansion 
-\begin{equation}\label{virialequation}
-    \frac{PV}{k_BT}=N\left(1+\frac{N}{V}B_2(T)+\left(\frac{N}{V}\right)^2B_3(T)+ \cdot\cdot\cdot\right) 
-\end{equation}
-where 
-\begin{equation}
-	B_{2}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}(r)}-1\right)d\vec{r} 
-\end{equation}
-Evaluating $B_{2}$ for the erf potential gives
-\begin{equation}
+the internal energy of a system and is given by
+\begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(\vec{r})d\vec r\end{equation}
+As shown in detail in the Appendix, evaluating the second virial coefficient 
+$B_{2}$ for the error function potential $V_{erf}$ with the mayer function 
+given by Equation~\ref{eq:VerfMayerfunction} 
+gives 
+\begin{equation} \label{eq:B2erf}
 	B_{2\operatorname{erf}}(T) = \frac{\pi}{3}\Xi^3\left[\left(\frac{\alpha^3}{\Xi^3}+\frac{3\alpha}{2\Xi}\right)\left(1+\operatorname{\operatorname{erf}}\left(\frac{\alpha}{\Xi}\right)\right)+\frac{1}{\sqrt{\pi}}\left(\frac{\alpha^2}{\Xi^2}+1\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}\right]
@@
 $B_{2WCA}$ can be found computationally from 
-\begin{equation}
+\begin{equation} \label{eq:B2wca}
 	B_{2WCA}=-\frac{1}{2}\int_{Volume}\left(e^{-\beta{V}_{WCA}(r)}-1\right)d\vec{r} 
 \end{equation}
-
-\davidsays{Some of this is worth showing, but I think you're giving too many
-plots here.
-In general, we'd like to only show plots if we have something to say about them,
-so either we want more words, or fewer figures.  e.g. what does the reader learn
-from looking at your plot of $B_2$ versus $T$? If there is something they should
-learn from it, then you should write that down, so they don't have to figure it
-out on their own.  If there isn't anything to say, then probably you can skip
-it. There \emph{are} times to just show a plot, but it's usually if it's an
-prediction, and you want to say \emph{something} about the plot beyond its
-existence.}
-A plot of the resulting second virial coefficient $B_2(T)$  with respect 
-to temperature is given in Figure~\ref{fig:B2vsT}, and a plot of $\Xi(T)$ obtained 
-through this method is shown in Figure~\ref{fig:xi_fromB2vsT}. 
-Figure~\ref{fig:erf_Potential_xifromB2} shows the erf potential generated with
-$\alpha(T)$ and $\Xi(T)$ at a few different temperatures alongside the WCA 
-potential which is not temperature dependent.
-
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{B2_vs_T_with_xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $B_2$ generated from 
-    matching the second virial coeficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:B2vsT}
-  \end{figure}
-
+The parameter $\Xi$ is found to be the value that makes Equations~\ref{eq:B2erf} 
+and \ref{eq:B2wca} equal each other for a given temperature T. 
+A plot of $\Xi(T)$ obtained through this method is shown in 
+Figure~\ref{fig:xi_fromB2vsT}.
 \begin{figure}[h!]
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{erf_Potential_xifromB2.pdf}
-    \caption{Error Function Potential $V_{\operatorname{erf}}(r)$ at several temperatures 
-    with $\sigma=1$, $\epsilon=1$, compared to the WCA potential $V_{WCA}(r)$ 
-    shown in black.}
-    \label{fig:erf_Potential_xifromB2}
-  \end{figure}
-
-%\color{red}
-% FIXME \noindent!ADD PLOT of df/dr=w2*w2 !!! BUT not here!
-%\color{black}
-
-%The spheres in the WCA model fluid do not exhibit quantum interference effects as the mass of 
-%the spheres, and the temperatures are taken to be large enough that the de Broglie wavelength 
-%is smaller than the mean distance between the spheres. Thus, the WCA fluid can be treated as a 
-%classical fluid.
-
-%Figure \ref{fig:weight_functions} shows a one dimensional view of the weight functions 
-%$w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance along the z-axis. In three 
-%dimensions, the weighting functions $w_0$, $w_1$, $w_2$ are spherical shells 
-%centered at $\vec{r}$. As shown in the cross-sectional view in Figure \ref{fig:W0andW3}a, 
-%$w_2$ peaks at a radius of $|\vec{r}-\vec{r'}|=\frac{\alpha}{2}$, a width determined 
-%by $\frac{\Xi}{\sqrt{2}}$, and edges that tapers off on either side of the peak giving 
-%the shell. The weighting function $w_3$, shown in Figure \ref{fig:W0andW3}b is a solid 
-%sphere up to $\frac{\alpha-4\Xi}{2}$ after which it tapers off.
-%The temperature dependence of parameters $\alpha$ and $\Xi$ 
-%%shown in Figures~\ref{fig:alphaXivsT} and \ref{fig:xi_fromB2vsT} 
-%can be related 
-%to the weight functions. 
-%$\Xi$ is related to the width of the functions which 
-%become wider in general as the temperature increases. 
-%The radial distance 
-%to the peak is related to $\alpha$, and decreases as the temperature 
-%increases, that is, the shells shrink, as does 
-%the sphere.
-%The functions can be treated as neglible outside of a distance of 
-%about $3(\frac{\Xi}{\sqrt{2}}) + \frac{\alpha}{2}$ CHECK!! from the center of 
-%the weight function.
-
- %\begin{figure}[h!]
-    %\centering
-    %\includegraphics[height=7cm]{weight_functions.pdf}
-    %\caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    %z-axis for a reduced temperature of 2.}
-    %\label{fig:weight_functions}
-  %\end{figure}
-
- %\begin{figure}[h!]
-    %\centering
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W2.png}
-      %% \includegraphics[width=5cm]{figs/w2-ring}
-      %\caption{A crossectional view of $w_2$}
-    %\end{subfigure} 
-    %\begin{subfigure}[b]{0.4\linewidth}
-      %\includegraphics[height=5cm]{W3.png}
-      %\caption{A crossectional view of $w_3$}
-    %\end{subfigure} 
-    %\caption{Cross-sectional views of the 3D weight functions. (a) $w_2$ is a 
-    %spherical shell in 3D with soft edges. Weight functions $w_0$ and $w_1$ 
-    %have a similar shape. (b) $w_3$ is a sphere with soft edges in 3D.}
-    %\label{fig:W0andW3}
-    %\end{figure}
+Figure \ref{fig:weight_functions} shows a one dimensional view of the 
+weight functions $w_0$, $w_1$, $w_2$, $w_3$ as they vary with distance 
+along the z-axis. In three dimensions, the weighting functions 
+$w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$, 
+and $w_3$ is a solid sphere with an edge that tapers off.
+$\Xi$ is related to the width of the weight functions which 
+become thicker, or taper off longer as the temperature increases. 
+$\alpha$ is related to the radius of the shell, or sphere, and decreases 
+as the temperature increases.
+
+\begin{figure}[h!]
+    \centering
+    \includegraphics[height=8cm]{weight_functions.pdf}
+    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
+    z-axis for a reduced temperature of 2.}
+    \label{fig:weight_functions}
+  \end{figure}
 
@@
 for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
-%from the equation 
-%\begin{displaymath}{\frac{F_{Hideal}}{V}=k_BTn(\ln(\frac{n*2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)-DELETE-EQN!}\end{displaymath} 
-%\begin{align}
-  %\frac{F_{Hideal}}{V}=k_BTn(\ln(n\Lambda^3)-1.0)
-%\end{align}
-%The quantity $\Lambda$ is the DeBroglie wavelength given by
-%\begin{displaymath}{\Lambda =\sqrt{\frac{2\pi\hbar^2}{k_BTm}}}\end{displaymath}  
-%where in our program we have set
-%\begin{displaymath}\left(\sqrt{\frac{2\hbar^2}{m}}\right)^3=4.752731840013864\text{x}10^{-7}\color{red}units?\color{black}\end{displaymath} 
-%\begin{displaymath}{m=}\end{displaymath}  
-%\begin{displaymath}{\hbar=}\end{displaymath} 
-%\begin{displaymath}{k_B=}\end{displaymath} 
 The crystal ideal Helmholtz free energy is computed computationally 
 from Equation~\ref{eq:Fideal-n(r)} with the varying number density profile given by
-%\begin{align}
-  %F_{ideal}[n(\vec{r})]= k_BT\int[n(\vec{r})\ln(n(\vec{r})\Lambda^3)-n(\vec{r})]dV
-%\end{align} 
 \begin{align}
@@
 where $\vec R_j$ is the position vector of the jth lattice site.
-%For small gw values it is assumed that the Guassians do not overlap, 
-%in which case the equation for the free energy can be simplified, and solved
-%analytically with the result
-%\begin{equation}
- %F_{Cideal}=(1-f_v)k_BT\left(\ln((1-f_v)\Lambda^3)-3\ln(\sqrt{2\pi{gw}})-\frac{5.0}{2}\right)
-%\end{equation}
-%For larger gw values it is assumed that the Gaussians do overlap, 
-%and so the free energy is caluculated computationally. First dx is scaled by gw,
-%\begin{equation}
-  %dx_{scaled} = (dx_{input})(gw)
-%\end{equation}
-%and then $n(\vec{r})$ is computed at one position of $\vec{r}$ within the 
-%primitive cell taking into account the contribution of Gaussians from 
-%many (5x5x5) nearby cells whose centers are given by $R_i$.
-%%\begin{displaymath}{n(|\vec{r}-\vec{R_i}|)+=\frac{(1-f_v)\exp^{-\frac{{|\vec{r}-\vec{R_i}|}^2}{2{(gw)}^2}}}{\left(\sqrt{2\pi}gw\right)^3}}\end{displaymath}
-%The contribution to the free energy of $n(\vec{r})$ at the one position is then
-%added to the accumulating sum.
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(ln(\frac{(n)2.646476976618268x10^{-6}}{\sqrt{(k_BT)^3}})-1.0)dV-DELETE-EQN!}\end{displaymath}
-%%\begin{displaymath}{cF_{ideal}+=k_BTn(\ln(n\Lambda^3)-1.0)dV}\end{displaymath}
-%This process is repeated for all positions within one primitive cell to get the
-%crystal ideal free energy for the primitive cell. This energy is then divided by
-%the volume of the primitive cell to get the crystal ideal free energy per
-%volume.
-%
 The liquid excess Helmholtz free energy is computed analytically  
-%for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$ 
 using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
 with a uniform number density profile, n that does not depend on $\vec r$.
-%done in hf.energy() note- does not depend on fv or gw but does use SFMT):
-%\begin{displaymath}{\frac{F_{Hexcess}}{V} = k_BT(\Phi_1+\Phi_2+\Phi_3)}\end{displaymath} 
-%For the case of a homogeneous number density n, $\Phi_2$, and $\Phi_3$ simplify to:
-%\begin{displaymath}{\Phi_1= -n_0\ln(1-n_3)}\end{displaymath} 
-%\begin{displaymath}{\Phi_2= \frac{n_1n_2}{1-n_3}}\end{displaymath} 
-%\color{red}\begin{displaymath}{\Phi_3=\frac{n_2^3}{24\pi(1-n_3)^2}}\end{displaymath}\color{black}
-%The weighted densities are found from
-%\begin{displaymath}{n_i=nw_i}\end{displaymath} 
-%where the weights $w_i$ are given by Equations~\ref{eq:weights}.
 
@@
 using SFMT with a number density profile 
-%$n(|\vec r'-\vec R_j|)$ 
-%used in Equation~\ref{eq:numdenprofile} 
-%that is composed of Guassian distributions centered at each lattice site.
 that consists of Guassian distributions 
-%$n(|\vec r'-\vec R_j|)$ 
 centered at lattice sites specified by position vectors $\vec R$. 
-%where j indicates the jth lattice site.
-%specified by position vectors $\vec R$.
 The weight functions $w_i(|\vec r - \vec r'|)$ are centered at $\vec r'$,
@@
 Keeping the weight function centered at $\vec r$, 
-%, and setting $\vec{dr}=\vec r' - \vec{R}_j$, 
-% and 
-%a vector $\vec dr$ is defined to be the difference between $\vec R$ and $\vec r'$.
-% $\vec dr = \vec r' - \vec R$.
-%First dx is scaled by $\Xi$ or gw, whichever is larger \begin{displaymath}{dx_{scaled}=(dx_{input})(\Xi + gw)}\end{displaymath}
-%Then the crystal, or inhomogeneous, excess free energy is found from
-%\begin{displaymath}{F_{Cexcess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{displaymath} 
 Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
-%$n_0(\vec{r})$, $n_1(\vec{r})$, $n_2(\vec{r})$, $n_3(\vec{r})$
-%form $\Phi_1(\vec{r})$, $\Phi_2(\vec{r})$, and $\Phi_3(\vec{r})$.
-%
 \begin{figure}[h!]
@@
 \end{figure} 
-%
-%\textcolor{red}{The center of the weighting function is placed at some vector $\vec{r}=\vec{r}_k$ 
-%within the primitive cell, and a vector $\vec{R}_j$ to the center of one of 
-%the Gaussians
-%is selected
-%that is within the primitive cell, or one of the 
-%neighboring cells within an inclusion radius as Gaussians 
-%farther away will not contriubute much (see Figure \ref{fig:InclusionRadius}).}
-%
-%The weighted densities $n_0(\vec{r}_k)$, $n_1(\vec{r}_k)$, $n_2(\vec{r}_k)$, 
-%$n_3(\vec{r}_k)$ at $\vec{r}_k$ are then computed using the weight functions 
-%$w_0(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_1(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, 
-%$w_2(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$, $w_3(|\vec{r}_k-\vec{R_j}-\vec{dr}|)$ 
-%where $\vec{dr}$ is a vector from the Gaussian centered at $\vec{R}_j$ to 
-%some point in space.
-%Monte-Carlo integration works 
 by randomly generating vectors
 $\vec r' -\vec R$ according to a Gaussian distribution. 
-%Setting $\vec{r}'=\vec{R}_j + \vec{dr}$, and 
-%Setting $\vec{dr}=\vec r' - \vec{R}_j$,  and 
-%$n(\vec{r'})=(1-f_v)n_g(\vec{r'})$ where $n_g(\vec{r'})$ 
-%where $n_g(\vec{r'})$ is the gaussian number density profile for no vacancies. 
 Only lengths for $\vec r'- \vec R$ that are within an inclusion radius
@@
 Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
-%lattice point located at $\vec R_j$ is given by
 \begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r' - \vec R_j)w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
@@
 NumPoints is the number of vectors $\vec r' - \vec R_j$ randomly generated about the jth lattice site, 
-%and $n_g(\vec{r'} - \vec R_j)$ 
-%(the value of the Gaussian at a radius of $|\vec{r'}|$) 
-%(the value of the Gaussian at a radius of $|\vec{dr}|$) 
 and $n(\vec r' - \vec R_j)$
@@
 number of vectors randomly generated), and by using antithetic variates.
-%
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_actual.png}
+ \begin{figure}[h!]
+    \centering
     \includegraphics[height=7cm]{figs/weightandGauss1.pdf}
-    %\caption{A Gaussian centered at $\vec{R}$ is shown in blue, and a cross-sectional 
-    %view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
     \caption{A Gaussian distribution about a lattice site located at 
@@
     view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
-    %Here, the weight function 
-    %is evaluated at $\vec{r}-\vec{r}'$ where $\vec{r}'=\vec{R}+\vec{dr}$ and $\vec{dr}$ 
-    %is generated randomly according to a Gaussian distribution.} 
   \label{fig:GaussandW2_actual}
   \end{figure}
-%
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=6cm]{GaussandW2_thinkas.png}
+
+ \begin{figure}[h!]
+    \centering
     \includegraphics[height=7cm]{figs/weightandGauss2.pdf}
@@
     and fixed in place as $\vec r' - \vec R$
-    %as $\vec{dr}=\vec{r}'-\vec{R}$ 
     varies randomly with a probability according to a Gaussian distribution.} 
@@
   \end{figure} 
-%
-%Monte-Carlo integration is employed to more efficiently 
-%compute the weighted densities used to calculate the excess free energy which is then 
-%added to the ideal free energy (computed analytically) to get the total free energy 
-%which is minimized. 
-%
-%-------Antithetic-------------------
-%To further reduce the error in the average, antithetic variates are used. 
-%Rather than computing an average of a set of random points, 
-%antithetic variates can be used to reduce error. 
 To use antithetic variates,
@@
 vector of equal magnitude, but opposite in direction. 
-%as shown in Figure \ref{fig:AntitheticVariate}. 
-%The quantity that is a function of the randomly generated vector
-%now has two values that can be averaged to get a more accurate value. 
 First order error is reduced
-%in the quantity that is a function of the randomly generated vector
-%average by filtering out an erroneous offset of values 
-%along the line connecting the two points 
 by averaging the two results that depend on the vectors, one of 
@@
 
- \begin{figure}[h!]
-    \centering
-    %\includegraphics[height=4cm]{AntitheticVariate.png}
-    %\caption{Antithetic variates reduce first order error in an average. 
-    %Each point in the average is paired with another point from along 
-    %the same line as the first, and at an equal, but opposite distance.}
-    \label{fig:AntitheticVariate}
-  \end{figure} 
-%Incorporating antithetic variates $-\vec{dr}$ for every $\vec{dr}$ equation ?? becomes:
-%\begin{displaymath}{n_i(\vec{r}_k, \vec{R}_j)=\sum_{NumPoints}\frac{1}{2}\frac{(1-f_v)\left[w_i(|\vec{r}-\vec{R_j}-\vec{dr}|) +w_i(|\vec{r}-\vec{R_j}+\vec{dr}|)\right]}{NumPoints}}\end{displaymath}
-%The process is repeated for all the other $\vec{R}_j$'s to get the total 
-%weighted densities at vector $\vec{r}_k$ due to all the Gaussians. 
-%\begin{displaymath}{n_i(\vec{r}_k)=\sum_j{n_i(\vec{r}_k, \vec{R}_j)}}\end{displaymath}
-%Then the entire process repeated for each vector $\vec{r}_k$ to get the 
-%total weighted densities $n_i(\vec{r})$ as a function of the position $\vec{r}$.
-
 The standard error of the mean (SEM) is given in general by
 \begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
-%which gives a measure of the error in $n_3=\overline{w}_3$:
 To compute the error, the standard deviation of weight function $w_3$ is used 
@@
 where the maximum value that the weighted density $n_3$ 
-%(which equals $\overline{w}_3$) 
 can properly have is 1.
-
-%Figure \ref{fig:GaussandW2_actual} ......
 
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n}
   \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
@@
   \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_P_vs_T}
   \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
@@
 \end{figure}
-
-%More plots...
-
-%\begin{figure}
-  %\centering
-  %\includegraphics[width=\columnwidth]{figs/T-vs-n_at_fixed_P_moreTemp}
-  %\caption{Plot of the reduced temperature verses the reduced density. 
-  %The grey region indicates coexistence.}
-  %\label{fig:T-vs-n_at_fixed_P}
-%\end{figure}
-
-%\begin{figure}
-  %\centering
-  %%\includegraphics[width=\columnwidth]{figs/p-vs-V_at_fixed_T}
-  %\caption{Plot of the reduced pressure verses the reduced volume. 
-  %The grey region indicates coexistence.}
-  %\label{fig:p-vs-V_at_fixed_T}
-%\end{figure}
-
-%\begin{figure}
-  %\centering
-  %%\includegraphics[width=\columnwidth]{figs/p-vs-n_at_fixed_T}
-  %\caption{Plot of the reduced pressure verses the reduced density. 
-  %The grey region indicates coexistence.}
-  %\label{fig:p-vs-n_at_fixed_T}
-%\end{figure}
 

### 2020-11-25 papers/thesis-kirstie/thesis.tex
cleanup
fixed spelling and removed white borders around plots

@@
   consists of soft spheres that exhibit the repulsive portion of the 
-  Leonard-Jones potential widely used to model the interaction of two 
+  Lennard-Jones potential widely used to model the interaction of two 
   neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
@@
 Helmholtz free energy at a particular temperature and density. 
-According to cDFT the equilbrium 
+According to cDFT the equilibrium 
 Helmholtz free energy of the system can be found by varying the number 
@@
 into account. 
-The problem of understanding the behavoir of real liquids is better addressed 
+The problem of understanding the behavior of real liquids is better addressed 
 by using models which reduce a complex picture to a simple one. Modifications, 
@@
 to be composed of hard spheres like, for example, a collection of marbles. 
-This model came about by considering the near incompressibilty of liquids 
+This model came about by considering the near incompressibility of liquids 
 indicating that the particles making up the liquid strongly resist being 
@@
 potential energy that is zero when two spheres are not in contact, but 
-goes instantly to infinity when contact is made. The inablilty to push 
+goes instantly to infinity when contact is made. The inability to push 
 real atoms closer together is due to the repulsion of the electrons in 
@@
 Thus, the hard-sphere model - which is the simplest, but still effective 
-representation of this replusive behavior that dominates the interaction 
+representation of this repulsive behavior that dominates the interaction 
 between real atoms, is currently the most widely used model for fluids.
@@
 Another fairly simple, but relatively accurate interatomic potential widely 
-used to represent generic, neutral atoms is the Leonard-Jones potential. 
-The Leonard-Jones potential describes both the long-range attraction, 
-and the short-range replusion that occurs between two nuetral atoms. 
-The attraction between two nuetral atoms is due to fleeting dipoles that 
+used to represent generic, neutral atoms is the Lennard-Jones potential. 
+The Lennard-Jones potential describes both the long-range attraction, 
+and the short-range repulsion that occurs between two neutral atoms. 
+The attraction between two neutral atoms is due to fleeting dipoles that 
 form spontaneously when electrons are redistributed as atoms come into 
-close proximty. These are often characterized as Van der Walls, or London 
+close proximity. These are often characterized as Van der Walls, or London 
 dispersion forces. The attraction grows stronger as two atoms move closer 
@@
 At this point the electrons in each atom begin to oppose the overlap of 
-the two atoms through their replusion. The repulsion increases rapidly and 
-approaches infinity as the distance between the atoms conintues to decrease.
+the two atoms through their repulsion. The repulsion increases rapidly and 
+approaches infinity as the distance between the atoms continues to decrease.
 
@@
     \includegraphics[height=7cm]{LJ_Potential.pdf}
-    \caption{Leonard-Jones Potential for $\sigma=1$, $\epsilon=1$}
+    \caption{Lennard-Jones Potential for $\sigma=1$, $\epsilon=1$}
     \label{fig:LJ_potential}
@@
 The combined attractive and repulsive effects produce the overall shape of 
-the Leonard-Jones interatomic potential shown in Figure \ref{fig:LJ_potential}. 
+the Lennard-Jones interatomic potential shown in Figure \ref{fig:LJ_potential}. 
 The potential is zero when the distance r between the centers of two atoms 
 is equal to $\sigma$ (shown in the figure at r/$\sigma$=1). At this point 
-the attraction bewteen the atoms just cancels their repulsion. As the distance 
+the attraction between the atoms just cancels their repulsion. As the distance 
 between the centers of the two atoms decreases below $\sigma$, the potential 
@@
 potentials, shown with different colors in Figure \ref{fig:LJ_potential_2parts}, 
-which when added together regenerate the full Leonard-Jones potential.
+which when added together regenerate the full Lennard-Jones potential.
 
@@
     \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
-    \caption{Leanard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
+    \caption{Lennard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
     \label{fig:LJ_potential_2parts}
@@
 
-The repulsive portion of the Leonard-Jones potential forms the potential for a 
+The repulsive portion of the Lennard-Jones potential forms the potential for a 
 Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
@@
 As with the Lenard-Jones potential, the repulsive WCA potential increases 
-rapidly and approches infinity as the two spheres begin to overlap. This 
+rapidly and approaches infinity as the two spheres begin to overlap. This 
 is a more realistic repulsive potential than that of the Hard Sphere model 
@@
 
-%\newpage
 \section{The Theory of Freezing and Maxwell's Double Tangent Construction}
@@
 representative of some arbitrary, unknown substance composed of atoms, 
-or molecules which exibit both attraction and repulsion. However, since 
-the atoms in a WCA fliud do not exhibit attraction there will not be a 
+or molecules which exhibit both attraction and repulsion. However, since 
+the atoms in a WCA fluid do not exhibit attraction there will not be a 
 dividing line between the gas and liquid phases, and so the term "liquid" 
@@
 and show that the solid and the fluid both have the same temperature and 
-the same pressure at a phase transistion.
+the same pressure at a phase transition.
 
@@
 related to the pressure. Since the pressure of the 
-liquid is the same as the pressure of the solid at the point of transistion, 
+liquid is the same as the pressure of the solid at the point of transition, 
 the density of the liquid and the density of the solid at the point of 
-freezing can be found from identifying the inverse densties at which the 
+freezing can be found from identifying the inverse densities at which the 
 tangents to each curve have the same slope. This process is called 
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Identifying densities at the solid-liquid transition using 
+    \caption{Densities at the solid-liquid transition are identified using 
     Maxwell's Double Tangent Construction}
@@
 
-The pressure at the transistion point can be obtained from the point of 
+The pressure at the transition point can be obtained from the point of 
 intersection that occurs when the Gibbs free energy per atom is plotted 
@@
 as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
-the slope and cooresponding y-intercept data collected from lines tangent 
+the slope and corresponding y-intercept data collected from lines tangent 
 along the curves on the the Helmholtz free energy per atom verses inverse 
@@
     the number densities of the liquid and the solid during a phase 
-    transistion. The gray region is where the liquid and the solid coexist, 
+    transition. The gray region is where the liquid and the solid coexist, 
     and are therefore not pure states.}
@@
 \begin{equation}\Phi=F-\mu{N}\end{equation}
-which can then be expressed in terms of a varrying number density 
+which can then be expressed in terms of a varying number density 
 profile $n(\vec r)$
@@
 second law of thermodynamics, a system that is not in equilibrium will 
-exprience spontanteous processes that increase the total entropy of a 
+experience spontaneous processes that increase the total entropy of a 
 system and its surroundings until equilibrium is reached, 
@@
 \begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
-These integrals can conveniently be written in terms of the DeBroglie 
+These integrals can conveniently be written in terms of the de Broglie 
 wavelength $\Lambda$, or the quantum volume $\text{n}_\text{Q}$
@@
     \includegraphics[height=7cm]{figs/diagrammic.pdf}
-    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interations 
+    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interactions 
     between individual pairs of atoms are shown in green. Interactions 
@@
 where the quantity in parenthesis can be recognized 
-as the second virial expanision coefficient typically called $B_2$. 
+as the second virial expansion coefficient typically called $B_2$. 
 It is important that the expression for the excess Helmholtz free energy 
@@
 Density Functional Theory, the equilibrium value of the 
-Helmholtz energy, and its cooresonding equilibrium density profile can be
+Helmholtz energy, and its corresponding equilibrium density profile can be
 found by varying the number density profile
@@
 forming an expression difficult. In this paper, Soft Fundamental Measure
-Theory will be used to form an epxression for the excess Helmholtz free 
+Theory will be used to form an expression for the excess Helmholtz free 
 energy, as explained in the next section.
@@
 In order to apply classical Density Functional Theory to a system that has
-a fixed temperature, volume, and nubmer of particles, a functional for 
+a fixed temperature, volume, and number of particles, a functional for 
 the Helmholtz free energy must first be formed. An expression for the ideal 
-portion of the Helmnoltz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
+portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
 but forming an expression for the excess Helmholtz free
@@
 It is an extension of the work done by Widom who showed that the excess 
-chemical potential of a homogenesous fluid corredsponds to the reversible 
+chemical potential of a homogenesous fluid corresponds to the reversible 
 work of inserting a sphere of radius R into a fluid consisting of hard 
@@
 in the excess Helmholtz free energy when one atom is added to an otherwise 
-closed system. He showed that the excess chemcial potential is related to 
+closed system. He showed that the excess chemical potential is related to 
 the ensemble average of the Boltzman factor for the potential energy 
@@
 along with their first and second derivatives, for the reversible work
-written for fluids with a cavitiy radius 
+written for fluids with a cavity radius 
 equal to, larger than, or very 
-much larger than the hard sphere radius R, Scaled Partical Theory arrives 
+much larger than the hard sphere radius R, Scaled Particle Theory arrives 
 at an expression~\cite{Hansen} for the excess Helmholtz free energy of 
@@
 Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
-is an extension of FMT and provides an expression for the excess Helmoholtz
+is an extension of FMT and provides an expression for the excess Helmholtz
 free energy of fluids with soft potentials like that of the WCA fluid. 
@@
 
-Fundamental Mearsure Theory begins by noticing that 
+Fundamental Measure Theory begins by noticing that 
 from Equation~\ref{Fexcess-SPT}, a quantity $G_{3}$ can be defined as follows
@@
 their related  weighted densities, can be written in a more general form 
-allowing for their extention to a system with a spacially varying number 
+allowing for their extension to a system with a spatially varying number 
 density, that is, an inhomogeneous fluid. Using a backwards Heavyside step function defined as
@@
 \begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
-which will be important to keep in tact when moving to SFMT.
+which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
 
@@
 Taking the derivative of the mayer function in Eq~\ref{f(r)step} provides 
-another relation that will be important to keep in tact when moving to SFMT.
+another relation that will be important to keep in tact when moving to 
+SFMT~\cite{schmidt2000fluid}.
 \begin{align}\label{mayerder=w2convolution}
@@
 \end{align} 
-This can easily be checked by evalating the left side for a Hard-sphere fluid as follows,
+This can easily be checked by evaluating the left side for a Hard-sphere fluid as follows,
 \begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
@@
 \end{align}
-in which a tensor weight is incoporated given by
+in which a tensor weight is incorporated given by
 \begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
@@
 All of the weight functions, the relations between them, and the equation 
-for the deriviative of the mayer function, can be expressed in terms of 
+for the derivative of the mayer function, can be expressed in terms of 
 $w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function.
 The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
-Equation~\ref{w2_w3_relation}, which is a step function for the Hard-sphere fluid.
+Equation~\ref{w2_w3_relation}.
 Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
-with soft potenials by keeping the relationships between the weight 
-functions the same, but replacing $w_{2}(r)$ with a weight function 
-appropriate for a soft sphere potential. This is what we do to
-form a functional for the 
-Weeks-Chandler-Anderson fluid as detailed in the next chapter on Methods.
+with soft potentials by keeping the relationships between the weight 
+functions the same, but replacing $w_{3}(r)$ with a weight function 
+appropriate for a soft sphere potential~\cite{schmidt1999density}. 
+This is what we do to form a functional for the Weeks-Chandler-Anderson 
+fluid as detailed in the next chapter on Methods.
 
@@
 have attraction between the particles that make up the fluid will show
-two discinct regions - one at lower densities, and one at higher densities,
-seperated by a coexistence region in which there is found a mixture of the 
+two distinct regions - one at lower densities, and one at higher densities,
+separated by a coexistence region in which there is found a mixture of the 
 fluid at some fixed lower density, and at some fixed higher density. An 
@@
 the liquid region which exhibits lower densities from the 
-solid which exhbits higher densities.
+solid which exhibits higher densities.
 
@@
 of an FCC cell, and contains the equivalent of one atom.
-It includes portions of eight atoms each located on a lattice site convenietly 
+It includes portions of eight atoms each located on a lattice site conveniently 
 reachable by moving along one of three different lattice vectors to step 
@@
     \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
-    \caption{The Primivitive Crystal Cell is shown by the shaded parallelpiped}
+    \caption{The Primitive Crystal Cell is shown by the shaded parallelpiped}
     \label{fig:primitivecell}
@@
 A plot of a Gaussian distribution in 2-dimensions is shown in Figure~\ref{fig:Gaus_plot}
-where the width of the Guassian is described by $\sigma$.  
+where the width of the Gaussian is described by $\sigma$.  
 A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
 a number density profile $n(\vec{r})$  which describes how the atoms are 
-statistically distrubuted over a collection of cells. 
+statistically distributed over a collection of cells. 
 It is important to note that the number density does not change when introducing a number 
@@
 It may be that not all lattice sites are occupied. Vacancies can be taken 
-into account while keeping 
-the overall number density the same by 
+into account while keeping the overall number density the same by 
 making the cells smaller. A large collection of smaller cells creates a 
 new number density profile shown in Figure~\ref{fig:Ensemble_Smallcells} 
-with eight smaller Gaussians each of which is now one-eigth the height. 
+with eight smaller Gaussians each of which is now one-eighth the height. 
 Instead of thinking of an atom in one of the eight small cells, the same 
@@
     \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{A Gaussian Distribution Number Density Profile}
+    \caption{A Gaussian Distribution Number Density Profile.}
     \label{fig:Ensemble_Smallcells}
@@
     show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
-    give rise to the same Gaussian distribution statistically}
+    give rise to the same Gaussian distribution statistically.}
     \label{fig:SameStatPic}
@@
     \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
-    \caption{A homogeneouse number density profile $n(\vec{r})$}
+    \caption{A homogeneous number density profile $n(\vec{r})$.}
     \label{fig:homogen_denisty}
@@
 For a given combination of temperature and number density, a suite of number 
-density profiles are generated similiar to those represented in the example 
+density profiles are generated similar to those represented in the example 
 set of profiles shown in Figure~\ref{fig:Ensemble_vary}. 
 The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
-density profile, and also for the homogeneous number denisty profile 
+density profile, and also for the homogeneous number density profile 
 characteristic of a liquid. Both the ideal and excess free energies are 
@@
 to determine the actual number densities of the crystal and liquid states
-at the boundaries of the coexistance region for a given temperature, 
+at the boundaries of the coexistence region for a given temperature, 
 and construct phase diagrams as described in the section on The Theory of
@@
 explicitly taking into account the interaction of each sphere with all the others, 
-but it must be taylored to fit a particular fluid\cite{schmidt1999density}.  
+but it must be tailored to fit a particular fluid\cite{schmidt1999density}.  
 Our SFMT functional is taylored to fit a WCA fluid, and 
@@
 by Gaussian functions centered at each lattice point.
-There are four scalor weight functions, two vector weight functions, and 
+There are four scalar weight functions, two vector weight functions, and 
 one tensor weight function all of which 
@@
 to fit the functional to a WCA fluid.
-The scalor weight functions are given by
+The scalar weight functions are given by
 \begin{align}\label{eq:weights}
@@
 The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
-Schmidt's proposed erf model~\cite{schmidt2000fluid} designed to model a steep, 
-soft potential in which he set
+Schmidt's proposed error function model~\cite{schmidt2000fluid} designed 
+to model a steep, soft potential in which he set
 \begin{equation}{w_3(r)=\frac{1}{2}}\left[1-\operatorname{erf}\left(\frac{r-\frac{\sigma}{2}}{\frac{\text{a}}{\sqrt{2}}}\right)\right]\end{equation} 
@@
 of the temperature-independent WCA potential which is difficult 
-to implement computionally. $\Xi$ is determined by setting the second virial 
-coeficeint $B_2$ of the error function potential $V_{erf}(r)$ and of the
+to implement computationally. $\Xi$ is determined by setting the second virial 
+coefficient $B_2$ of the error function potential $V_{erf}(r)$ and of the
 WCA potential equal.  
@@
 \begin{align}{\frac{dw_3(r)}{dr}\bigg|_{\infty}-\frac{dw_3(r)}{dr}\bigg|_{r}=w_2(r)}\end{align} 
-we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} %From the deconvolution equation \begin{equation}{-\frac{1}{2}f(r)=w_0*w_3 + w_1*w_2 -\vec{w}_{V1}*\vec{w}_{V2}}\end{equation} it follows that $w_2$ must satisfy the relation
+we get \begin{equation}{ w_2(r)=\frac{\sqrt{2}}{\Xi\sqrt\pi}\exp^{-\left(\frac{r-\frac{\alpha}{2}}{\Xi/\sqrt{2}}\right)^2} }\end{equation} 
 Putting $w_2(r)$ into Equation~\ref{mayerder=w2convolution}, shown again below, 
@@
     \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
-    for $\sigma=1$ and $\epsilon=1$}
+    for $\sigma=1$ and $\epsilon=1$.}
     \label{fig:alphaXivsT}
@@
 We determine the value of $\Xi$ at a given temperature $T$,
-by equating the second virial coefficient evaluated for the error function 
-potential $V_{\operatorname{erf}}(r)$, and the second virial coefficent evaluated for the 
-WCA potential $V_{WCA}$.
+by equating the second virial coefficient evaluated for the error 
+function potential $V_{\operatorname{erf}}(r)$, and the second virial 
+coefficient evaluated for the WCA potential $V_{WCA}$.
 \begin{equation}B_{2\operatorname{erf}}(T) =B_{2WCA}(T)\end{equation}
@@
     \caption{Shows the temperature dependence of $\Xi$ generated from 
-    matching the second virial coeficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
+    matching the second virial coefficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
     at each temperature.}
@@
 using SFMT with a number density profile 
-that consists of Guassian distributions 
+that consists of Gaussian distributions 
 centered at lattice sites specified by position vectors $\vec R$. 
@@
 will be taken into account as Gaussians 
-farther away will not contriubute much.
+farther away will not contribute much.
 The equation for the contribution to the weighted density from a 
@@
 This process is repeated to get the contributions to the weighted density 
-$n_i(\vec r)$ from the Guassians around each lattice point which begin to 
-merge together as the width of the Guassians increases.
+$n_i(\vec r)$ from the Gaussians around each lattice point which begin to 
+merge together as the width of the Gaussians increases.
 All the contributions are then added together 
@@
 number of vectors randomly generated), and by using antithetic variates.
+To use antithetic variates, each randomly generated  vector is paired 
+with a vector of equal magnitude, but opposite in direction. 
+First order error is reduced by averaging the two results that depend on 
+the vectors, one of which may be erroneously higher than it should be 
+while the other is erroneously lower than it should be.
+
  \begin{figure}[h!]
@@
   \end{figure}
-To use antithetic variates,
-each randomly generated  vector is paired with a 
-vector of equal magnitude, but opposite in direction. 
-First order error is reduced
-by averaging the two results that depend on the vectors, one of 
-which may be erroneously higher than it should be while the other is 
-erronesously lower than it should be.
 
@@
 and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
-In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature $T^*=\frac{kT}{\epsilon}$ 
-is plotted against the reduced number density $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
-The red area is fluid, and the blue area is crystal. The gray area that separates the fluid, 
-and the crystal regions is where some fluid and crystal exist together at varying porportions. 
-The region of coexistence slants slightly to the right toward higher density as the temperature 
-increases. This is expected since at higher temperatures, the fluid would have to be under 
-greater compression before forming a solid. The white lines show results from Monte-Carlo 
-simulations of the WCA fluid~\cite{May}. Overall, the general shape, and position of the lines correspond
-fairly closely to the edges of the region of coexistance predicted from the DFT method, 
-but the width of the coexistance region is about one-third larger for the Monte-Carlo results 
-for most temperatures. The white line on the right, representing the density of the solid at 
-each temperature, is in fairly good aggreement with our DFT results for all temperatures. 
-However, the white line on the left, representing the density of the liquid at each temperature, 
-is consistenty farther left than the liquid transition densities our method predicted.
+In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
+$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
+$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. The red area is fluid, 
+and the blue area is crystal. The gray area that separates the fluid, 
+and the crystal regions is where some fluid and crystal exist together 
+at varying proportions. The region of coexistence slants slightly to the 
+right toward higher density as the temperature increases. 
+This is expected since at higher temperatures, the fluid would have to 
+be under greater compression before forming a solid. The white lines show 
+results from Monte-Carlo simulations of the WCA fluid~\cite{May}. 
+Overall, the general shape, and position of the lines correspond
+fairly closely to the edges of the region of coexistance predicted from 
+the DFT method, but the width of the coexistence region is about one-third 
+larger for the Monte-Carlo results for most temperatures. The white line 
+on the right, representing the density of the solid at each temperature, 
+is in fairly good agreement with our DFT results for all temperatures. 
+However, the white line on the left, representing the density of the 
+liquid at each temperature, is consistently farther left than the liquid 
+transition densities our method predicted.
 
@@
   \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
-  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
   White lines show the liquid and solid transition densities at each temperature from 
@@
 temperatures and pressures is what would be expected, with the WCA fluid becoming a solid at
-a higher pressure for a higher temperatuer, or becoming a liquid at a higher temperature 
+a higher pressure for a higher temperature, or becoming a liquid at a higher temperature 
 for a higher pressure. But the transition pressures and temperatures are consistently higher 
-than the Monte-Carlo simulations shown by the white line, and the descripancy grows as 
+than the Monte-Carlo simulations shown by the white line, and the discrepancy grows as 
 the temperature increases. 
@@
 \chapter{Conclusion}
-In this paper we were sucessful in predicting the freezing of a 
+In this paper we were successful in predicting the freezing of a 
 Weeks-Chandler-Anderson fluid using our SFMT functional built by 
@@
 \begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-%\begin{displaymath}{r_x=r\sin(\theta)\cos(\phi) ~~~~~ r_y=r\sin(\theta)\sin(\phi)~~~~~ r_z=r\cos(\theta)}\end{displaymath}
 where $r=r'$ and
@@
 Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
-%where \begin{displaymath}{r_x=r\sin(\theta)\cos(\phi)}\end{displaymath}
-%\begin{displaymath}{r_y=r\sin(\theta)\sin(\phi)}\end{displaymath}
 \begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{y'})}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
@@
 
-%\begin{displaymath}{\widetilde{w}_{{m2}_{xx}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\sin^2(\theta)e^{-ikr\cos(\theta)}\left(\int_{0}^{2\pi}\cos^2(\phi)d{\phi}\right)d{\cos(\theta)}\right]d{r}}\end{displaymath}
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
@@
 
-%\newline $k_{x'}= k_{x'_x} \hat{x} + k_{x'_y} \hat{y} + k_{x'_z} \hat{z}$
-%\newline $k_{y'}= k_{y'_x} \hat{x} + k_{y'_y} \hat{y} + k_{y'_z} \hat{z}$
-%\newline $k_{z'}= k_{z'_x} \hat{x} + k_{z'_y} \hat{y} + k_{z'_z} \hat{z}$
-%\newline \begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(\vec{k})=\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{displaymath}
 \noindent Using the relation
-%\begin{displaymath}{\hat{k_{x'}}\hat{k_{x'}} + \hat{k_{y'}}\hat{k_{y'}} + \hat{k_{z'}}\hat{k_{z'}} = 1}\end{displaymath} 
 \begin{displaymath}{\text{I} = \hat{x'}\hat{x'} + \hat{y'}\hat{y'} + \hat{z'}\hat{z'}}\end{displaymath}
@@
 \begin{displaymath}{=a\sqrt{\pi}\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2}\cos\left(\frac{k\alpha}{2}\right)d{k} }\end{displaymath} 
-%\begin{displaymath}{=a\sqrt{\pi}\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2}\left(\frac{e^{i\frac{k\alpha}{2}}+e^{-i\frac{k\alpha}{2}}}{2}\right)d{k} }\end{displaymath} 
-%\begin{displaymath}{=\frac{a\sqrt{\pi}}{2}\left[\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2 + i\frac{k\alpha}{2}}d{k}+\int_{0}^{k}e^{-\left(\frac{ka}{2}\right)^2-i\frac{k\alpha}{2}}d{k}\right]}\end{displaymath}
 \begin{displaymath}{=\frac{a\sqrt{\pi}}{2}\left[\int_{-k}^{k}e^{-\left(\frac{ka}{2}\right)^2 + i\frac{k\alpha}{2}}d{k}\right]}\end{displaymath} 
-%\begin{displaymath}{=\frac{a\sqrt{\pi}}{2}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\int_{-k}^{k}e^{-\left(\frac{ka}{2} + i\frac{\alpha}{2a}\right)^2}d{k}\right]}\end{displaymath} 
 \begin{displaymath}{=\sqrt{\pi}e^{-\left(\frac{\alpha}{2a}\right)^2}\int_{-\frac{ka}{2}+i\frac{\alpha}{2a}}^{\frac{ka}{2}+i\frac{\alpha}{2a}}e^{-u^2}d{u}}\end{displaymath} 
@@
 
-\section{Deriviation of Tensor Density $n_{m2ij}(\vec{r})$}
+\section{Derivation of Tensor Density $n_{m2ij}(\vec{r})$}
 \begin{equation}{\overleftrightarrow{n}_{m2}(\vec{r})=\int_{allspace}n(\vec{r'})\overleftrightarrow{w}_{m2}(\vec{r}-\vec{r'})d{\vec{r'}}}\end{equation} 
@@
 \begin{align}
-%    n_{m2ij}(\vec r) &= \int n(\vec r') w_{m2ij}(\vec r- \vec r')d\vec r' \\
-%    w_{m2ij}(\vec r) &= \frac1{2\pi}\int d\vec k w_{m2ij}(\vec k)e^{i\vec k\cdot \vec r}\\
-%    n_{m2ij}(\vec{r})=\int_{allspace{n(\vec{r}')w_{m2ij}(\vec{r}-\vec{r}'){}d{\vec{r}'}}\\
     n_{m2ij}(\vec r) &= \int n(\vec r') \left(\frac{1}{\left(2\pi\right)^3}\int w_{m2ij}(\vec k)e^{i\vec k\cdot (\vec r-\vec r')}d\vec{k}\right)d\vec r' \\
@@
 
-\section{Deriviation of the Tensor version of $\Phi_3$}
+\section{Derivation of the Tensor version of $\Phi_3$}
 
@@
 
-%\begin{equation}f(r)=e^{-\frac{1}{k_BT}\left[4\epsilon\left(\left(\frac{\sigma}{r}\right)^{12}\left(-\frac{\sigma}{r}\right)^6\right)+\epsilon\right]}-1\end{equation}
-
 \begin{equation}f_{WCA}(r)=\operatorname{exp}\left(-\frac{1}{k_BT}\left[4\epsilon\left(\sigma^{12}r^{-12}-\sigma^{6}r^{-6}\right)+\epsilon\right]\color{white}\frac{1}{1}\color{black}\right)-1\end{equation}
@@
 \begin{equation}\widetilde{f}_{WCA}(k)=\frac{4\pi}{k}\int_0^{\infty}{\left(e^{-\frac{1}{k_BT}\left[4\epsilon\left(\sigma^{12}r^{-12}-\sigma^{6}r^{-6}\right)+\epsilon\right]}-1\right) r\sin(kr)dr}\end{equation}
-
-%\begin{equation}B_3=-\frac{4\pi}{3}\int{|\widetilde{f}(-\vec k)|^3k^2dk}\end{equation}
 
@@
 \begin{equation}N=<n>=\sum_{i=0}^\infty{P_in_i}\end{equation}
-%\begin{equation}\beta=\frac{1}{k_BT}\end{equation}
-%\begin{equation}\frac{PV}{k_BT}=-\sum{P_i\ln{P_i}}-\frac{1}{k_BT}(\sum{P_iE_i}-\mu\sum{P_iN})\end{equation}
 Equation~\ref{Euler_rearranged} then becomes
@@
 \begin{equation}\label{PV/kBT_reduces_for_N}\frac{PV}{k_BT}=\ln{\left(\lambda^NZ_{oN}Q_N\right)}
-%=\ln\lambda^n + \ln{Z}_{on}+\ln{Q}_n
 =\beta\mu{N} + \ln{Z}_{oN} +\ln{Q}_N\end{equation}
@@
 \begin{equation}\frac{P_{ideal}V}{k_BT}=\beta\mu_{ideal}{N} + \ln{Z}_{oN}=N\end{equation}
-%\begin{equation}P_{ideal}=\mu{N} + k_BT\ln{Z}_{oN}\end{equation}
 \begin{equation}\frac{P_{excess}}{k_BT}= \beta\mu_{excess}{N} + \ln{Q}_N\end{equation}
@@
 \begin{equation}\frac{PV}{k_BT}=N + \left(-\frac{N}{V}\int{f(r)d\vec{r}}\right){N} + \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
-%\begin{equation}\frac{PV}{k_BT}=N - \frac{N^2}{V}\frac{1}{2}\int{f(r)d\vec{r}}\end{equation}
 \begin{equation}\frac{PV}{k_BT}=N\left(1-\frac{N}{V}\frac{1}{2}\int{f(r)d\vec{r}}\right)\end{equation}
@@
 \section{Program Outline}
-
-%\begin{displaymath}F_{homogeneous} = F_{ideal.homogeneous} + F_{excess.homogeneous}\end{displaymath}
-%\begin{displaymath}F_{crystal} = F_{ideal.crystal} + F_{excess.crystal}\end{displaymath}
-%\newline
-
-%\newline\noindent Inputs: 
-%t*, n*, range of gw values , range of fv values
-%\newline Outputs: Homogeneous (liquid)  Helmholtz free energy (at the given T,V,N) 
-%\newline\indent\indent\indent Crystal Helmholtz free energy (at the given T,V,N)
 
@@
 
-%\begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{crystal}\end{displaymath}
-%\begin{displaymath}\left(\frac{F}{N}\right)_{homogeneous}\end{displaymath}
-
 %lattice_constant= pow(4*(1-fv)/reduced_density, 1.0/3)
-
 %The number of atoms in one primitive cell is one.

### 2020-11-25 papers/thesis-kirstie/figs/plot_WCA_Potential_fixed.py

@@ -30,6 +30,8 @@ r=np.linspace(1.11, 2.5, 2000)
 V=(r/r) - 1
 plt.plot(r/sigma, V/epsilon, color='blue')
 
+plt.xlim(left=0.91,right=2.5)
+plt.ylim(top=5)
 
 plt.xlabel('r/$\sigma$')
 plt.ylabel('$V(r)/\epsilon$')

### 2020-11-25 papers/thesis-kirstie/figs/plot_w3_w2_w1_w0.py

@@ -62,8 +62,7 @@ plt.plot(rz/(alpha/2),w0, label = 'w0', color='red')
 plt.plot(rz/(alpha/2),w1, label = 'w1', color='blue')
 plt.plot(rz/(alpha/2),w2(rz), label = 'w2', color='purple')
 plt.plot(rz/(alpha/2),w3, label = 'w3', color='green')
-#plt.xlim(0, 2)
-#plt.ylim(0, 6)
+plt.xlim(0, 2)
 plt.xlabel('rz/(alpha/2)')
 plt.ylabel('weight function')
 plt.title('Weight Functions at KbT=2')

### 2020-11-27 papers/thesis-kirstie/thesis.tex
more edits

@@
 \advisor{David Roundy}
-\submitdate{October ??, 2020}
+\submitdate{December 10, 2020}
 \commencementyear{2021}
@@
 
-\date{October 2020}
+\date{December 2020}
 
@@
   with a fairly good match to Monte-Carlo simulations. This result suggests 
-  that SFMT may provide a solid foundation for future density functionals 
+  that SFMT may provide a strong foundation for future density functionals 
   that will model real fluids.
@@
 has with every other sphere does not have to be taken explicitly into account.
-Once the functional for the Helmholtz free energy is formed, classical 
-Density Functional Theory is used to find the equilibrium 
-Helmholtz free energy at a particular temperature and density. 
-According to cDFT the equilibrium 
-Helmholtz free energy of the system can be found by varying the number 
+Once the functional for the Helmholtz free energy is formed, then 
+according to cDFT, the equilibrium
+Helmholtz free energy of the system can be found for a particular 
+temperature and density by varying the number 
 density profile $n(\vec{r})$ until the free energy is minimized. 
+Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
 Results will be compared to those from Monte-Carlo simulations.
@@
 with soft potentials by keeping the relationships between the weight 
-functions the same, but replacing $w_{3}(r)$ with a weight function 
-appropriate for a soft sphere potential~\cite{schmidt1999density}. 
-This is what we do to form a functional for the Weeks-Chandler-Anderson 
+functions (given by Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},
+~\ref{eq:w_v2},~\ref{w2_w3_relation},~\ref{eq:tensorweight}) and the 
+equation for the derivative of the mayer function 
+(given by Equation~\ref{mayerder=w2convolution}) the same, 
+but replacing $w_{3}(r)$ with a weight function 
+appropriate for a soft sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+This is what we do in forming a functional for the Weeks-Chandler-Anderson 
 fluid as detailed in the next chapter on Methods.
@@
 Our goal is to generate temperature-density and pressure-temperature 
-phase diagrams for a Weeks-Chandler-Anderson model fluid using 
-classical Density Functional Theory. 
+phase diagrams for a Weeks-Chandler-Anderson model fluid. 
 A temperature-density phase diagram for a fluid that does not
 have attraction between the particles that make up the fluid will show
-two distinct regions - one at lower densities, and one at higher densities,
-separated by a coexistence region in which there is found a mixture of the 
-fluid at some fixed lower density, and at some fixed higher density. An 
-abrupt change from some maximum lower density to some minimum higher density 
-occurs for certain combinations of pressure and temperature which form
-a line on a pressure-temperature phase diagram. This line distinguishes 
-the liquid region which exhibits lower densities from the 
-solid which exhibits higher densities.
-
-The transition between the liquid 
-and solid state is abruptly inhomogeneous in terms of the number density, 
-and we study the behavior 
-of the fluid by treating the liquid as having a homogeneous number density, 
-and by constructing a crystal lattice in real space for the solid. 
+two distinct regions - a "liquid" region at lower densities, and a solid 
+region at higher densities. 
+A phase transition between a liquid and a solid is accompanied by an 
+abrupt change in the number density, and so there is a gap between the 
+purely liquid and the purely solid regions. Within the gap, or coexistence
+region, is found mixtures of various amounts of liquid and solid.  
+Within the coexistence region, the liquid and solid have number densities 
+that are fixed at the values they had at the phase transition. To construct
+the phase diagram, these liquid and solid phase transition number densities
+must be found. Maxwell's double tangent construction can be used to find 
+these number densities given plots of the liquid and the solid Helmholtz 
+free energies verses the inverse density at a given temperature. 
+To construct these plots we use classical Density Functional Theory 
+to find the Helmholtz free energies of the liquid and the solid. 
+A functional for the ideal portion of the
+Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, and a 
+functional for 
+the excess Helmholtz free energy, which depends on the interatomic potential
+between particles, will be developed in the next section using SFMT. 
+Once a functional is constructed, the Helmholtz free energy can be found - 
+but not directly. The equilibrium number density profile must first be 
+identified. According to cDFT, the equilibrium number density profile, 
+and its corresponding equilibrium Helmholtz free energy can be found 
+by trying many different number densities, and identifying the one that 
+minimizes the Helmholtz free energy. In this section, we look at how to 
+generate various number density profiles that can be used to implement cDFT.
+
 It is known that a fluid with a soft potential like the WCA fluid will 
@@
 FCC cells are not the smallest cells with which the lattice can be built. 
-The shaded parallepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
+The shaded parallelepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
 is the smallest structure that, when repeated, re-creates the entire crystal 
-lattice. This parallepiped primitive cell takes up one-fourth the volume 
+lattice. This parallelepiped primitive cell takes up one-fourth the volume 
 of an FCC cell, and contains the equivalent of one atom.
 It includes portions of eight atoms each located on a lattice site conveniently 
-reachable by moving along one of three different lattice vectors to step 
-from one atom to the next. The three lattice vectors will be useful in 
+reachable by moving along one of three different lattice vectors. 
+The three lattice vectors will be useful in 
 creating the lattice structure computationally, and will also make the 
@@
     \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
-    \caption{The Primitive Crystal Cell is shown by the shaded parallelpiped}
+    \caption{The Primitive Crystal Cell is shown by the shaded parallelepiped}
     \label{fig:primitivecell}
@@
 
-In a large collection of cells, the atom at each lattice site
-will not realistically reside exactly at the lattice point. 
+An atom will not realistically reside exactly at a lattice point. 
 Statistically there will typically be a normal, or Gaussian distribution 
-of positions about the lattice point as pictured in the far right box in 
-Figure \ref{fig:Ensemble_Gaus}. 
-A plot of a Gaussian distribution in 2-dimensions is shown in Figure~\ref{fig:Gaus_plot}
-where the width of the Gaussian is described by $\sigma$.  
-A 3-dimensional Gaussian distribution centered at a lattice point can be used to represent 
-a number density profile $n(\vec{r})$  which describes how the atoms are 
-statistically distributed over a collection of cells. 
-It is important to note that the number density does not change when introducing a number 
-density profile. There are still the same number of atoms over some fixed volume. 
-\begin{equation}{n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}=\frac{1~\text{atom}}{V}}\end{equation}
-where $n(\vec r)$ is itself an ensemble average over all microstates, and is independent of time.
-\begin{equation}n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~\end{equation}
+of positions about the lattice point. 
+Of course, atoms are always in motion, so when talking about "the location" 
+of an atom, we are talking about a time averaged position, so that it 
+does not change with time.
+Figure \ref{fig:Ensemble_Gaus} shows examples 
+of four possible positions an atom (shown in blue) might have in a crystal 
+cell that has one lattice site (represented by a black dot the center of 
+the cell). The atom is slightly off the lattice site in each example, and 
+a sampling of many possibilities - more than just the four given 
+as examples - would create an average profile that looks like a Gaussian
+distribution, as shown on the right.
+A plot of a Gaussian distribution in 2-dimensions is shown in 
+Figure~\ref{fig:Gaus_plot} where the width of the Gaussian is described 
+by $\sigma$ (not to be confused with the $\sigma$ related to the 
+Lennard-Jones model).
+
+3-dimensional Gaussian distributions centered at lattice points can be 
+used to represent a number density profile $n(\vec{r})$  which describes 
+how the atoms are distributed on average over a collection of cells.
 
@@
     \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
-    \caption{A Gaussian Distribution Number Density Profile}
+    \caption{Four examples of possible positions an atom (blue sphere) 
+    might have within a cell that has one lattice site (black dot) at 
+    the center are shown on the left. A large number of such 
+    possibilities would give rise to a Gaussian distribution, 
+    as shown on the right.}
     \label{fig:Ensemble_Gaus}
@@
 
+There can be many different number density profiles for a given number
+density. The number density is the number of atoms per volume, whereas
+the number density profile describes how much "atom" will be found at various 
+points in the cell. 
+It can be thought of as an average over a large collection of actual cells
+of various possible spatial distributions of the atoms within a cell, or
+more precisely, as an ensemble average over all possible microstates.
+\begin{align}
+    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
+\end{align} 
+By varying the width of the Gaussian distributions, a number of different 
+number density profiles can be generated for a given number density.
+It is important to note that the number density does not change when 
+introducing a number density profile.  
+There are still the same number of atoms over some fixed volume. 
+Taking the average of the number density profile over some volume 
+gives the number density. 
+\begin{align}
+    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
+\end{align}
+
+Besides changing the width of the Gaussians, there is another way to 
+generate different number density profiles for a given number density.
 It may be that not all lattice sites are occupied. Vacancies can be taken 
@@
     \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{A Gaussian Distribution Number Density Profile.}
+    \caption{Four examples of possible positions an atom (blue sphere) might 
+    have within a fixed volume containing eight cells are shown on the left. 
+    Each cell has one lattice site at the center represented by a black dot. 
+    A large number of such possibilities would give rise to the Gaussian 
+    distributions shown on the right.}
     \label{fig:Ensemble_Smallcells}
@@
 
-By varying the width of the Gaussians and the fraction of vacancies, various
-number density profiles can be generated as shown in Figure~\ref{fig:Ensemble_vary}.
+By varying the width of the Gaussians and the fraction of vacancies, 
+various number density profiles can be generated for a given combination 
+of temperature and number density, as shown in Figure~\ref{fig:Ensemble_vary}.
+A homogeneous number density profile is used for a liquid is shown in 
+Figure \ref{fig:homogen_denisty}.
 Once again, it is important to note that the number density is the same 
 for all these number density profiles $n(\vec{r})$. 
-A homogeneous number density profile is used for a liquid is shown in Figure \ref{fig:homogen_denisty}.
-
 \begin{figure}[h!]
@@
     \caption{Several Gaussian Distribution Number Density Profiles all 
-    with the same number density, that is, there is one atom over the fixed volume.}
+    with the same number density of one atom over the fixed volume.}
     \label{fig:Ensemble_vary}
   \end{figure}  
-
  \begin{figure}[h!]
@@
   \end{figure}   
-
-For a given combination of temperature and number density, a suite of number 
-density profiles are generated similar to those represented in the example 
-set of profiles shown in Figure~\ref{fig:Ensemble_vary}. 
 The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
@@
 at the boundaries of the coexistence region for a given temperature, 
-and construct phase diagrams as described in the section on The Theory of
-Freezing and Maxwell's Double Tangent Construction.
+and construct phase diagrams as described in Section 2.2.
 
@@
 $T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
-$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. The red area is fluid, 
-and the blue area is crystal. The gray area that separates the fluid, 
-and the crystal regions is where some fluid and crystal exist together 
-at varying proportions. The region of coexistence slants slightly to the 
-right toward higher density as the temperature increases. 
-This is expected since at higher temperatures, the fluid would have to 
-be under greater compression before forming a solid. The white lines show 
-results from Monte-Carlo simulations of the WCA fluid~\cite{May}. 
-Overall, the general shape, and position of the lines correspond
-fairly closely to the edges of the region of coexistance predicted from 
-the DFT method, but the width of the coexistence region is about one-third 
-larger for the Monte-Carlo results for most temperatures. The white line 
-on the right, representing the density of the solid at each temperature, 
-is in fairly good agreement with our DFT results for all temperatures. 
-However, the white line on the left, representing the density of the 
-liquid at each temperature, is consistently farther left than the liquid 
-transition densities our method predicted.
-
+$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
+It is seen that in general, a liquid exists at lower number densities, 
+and a solid exists at higher number densities. 
+Phase transitions between the liquid and solid phases will occur when the liquid 
+and solid are at the same temperature, pressure and chemical potential. 
+Looking at one temperature, it is seen that as the density of the 
+liquid increases, a maximum number density is reached for the liquid. 
+This is the liquid phase transition number density, and 
+marks the beginning of a coexistence region where various amounts 
+of liquid and solid can exist together in equilibrium.
+Moving on toward higher number densities, the end of the coexistence 
+region is marked by the minimum number density that the solid can have.
+This is the solid phase transition number density. At a phase transition 
+there is an abrupt change in the number density as seen here where the
+density changes suddenly, and abruptly from the liquid 
+phase transition number density to the solid phase transition number density.
+Within the coexistence region, the number density of the liquid is fixed 
+at the liquid phase transition number density, and the number density of 
+the solid is fixed at the solid phase transition number density, while
+varying amounts of liquid and solid combine to give the infinite number 
+of possible densities seen within the coexistence region.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=9cm]{figs/Phase_Diagram_of_T_vs_n}
+  \caption{Phase Diagram of the reduced temperature verses the reduced 
+  density for the WCA fluid. The red region is liquid, and the blue region is 
+  solid. The gray area that separates the liquid and solid regions 
+  is the region of coexistence where some liquid and solid exist together 
+  at varying proportions. White lines show the boundaries of the coexistence 
+  region generated from Monte-Carlo of the WCA fluid.}
+  \label{fig:Phase_Diagram_of_T_vs_n}
+\end{figure}
+
+Looking at the coexistence region as a whole for all temperatures, 
+it is seen to extend over a narrow range of number densities which 
+gradually shifts towards higher densities as the temperature increases. 
+This is expected since at higher temperatures the fluid would have to 
+be under greater compression before forming a solid. 
+White lines superimposed on the phase diagram show the boundaries of the 
+coexistence region as predicted by Monte-Carlo simulations~\cite{May} 
+of the WCA fluid. 
+Overall, the general shape, and position of the two coexistence regions 
+correspond fairly closely, but the Monte-Carlo region is a little wider
+and shifted a bit toward smaller densities. For example, at a temperature 
+of kT=1, the cDFT liquid transition reduced number density is 0.96, 
+and the solid transition reduced number density is 1.02, giving a spread 
+of 0.06, whereas the Monte-Carlo results give a liquid transition 
+reduced number density of 0.92, and a solid transition reduced number 
+density of 1.00, giving a spread of 0.08. 
+%The solid phase transition number densities predicted by each method 
+%are very nearly the same along all temperatures, but the liquid phase 
+%transition number densities predicted by cDFT are notably higher than 
+%those predicted by the Monte-Carlo simulations, making the width of the 
+%Monte-Carlo region about one-third larger for most temperatures above KT=1. 
 %Figure~\ref{fig:Phase_Diagram_of_T_vs_n_AS} shows our results for a wider 
@@
 %region predicted by Ahmed and Sadus. 
+%
+%
+%
+%\begin{figure}
+  %\centering
+  %%\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
+  %%\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
+  %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  %%White lines show the liquid and solid transition densities at each temperature from 
+  %%simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
+  %\label{fig:Phase_Diagram_of_T_vs_n_AS}
+%\end{figure}
 
@@
   \centering
-  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
-  \caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
-  White lines show the liquid and solid transition densities at each temperature from 
-  Monte-Carlo simulations of the WCA fluid. The gray region is the region of coexistence.}
-  \label{fig:Phase_Diagram_of_T_vs_n}
-\end{figure}
-
-\begin{figure}
-  \centering
-  %\includegraphics[width=\columnwidth]{figs/Phase_Diagram_of_T_vs_n_AS}
-  %\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
-  %\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
-  %White lines show the liquid and solid transition densities at each temperature from 
-  %simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
-  \label{fig:Phase_Diagram_of_T_vs_n_AS}
-\end{figure}
-
-Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure $p^*=\frac{P\sigma^3}{\epsilon}$ 
-plotted against the reduced temperature for temperatures of kT = 0.5 to 3.0. Again, the red area 
-is fluid, and the blue area is crystal. The general trend of the values for the transition 
-temperatures and pressures is what would be expected, with the WCA fluid becoming a solid at
-a higher pressure for a higher temperature, or becoming a liquid at a higher temperature 
-for a higher pressure. But the transition pressures and temperatures are consistently higher 
-than the Monte-Carlo simulations shown by the white line, and the discrepancy grows as 
-the temperature increases. 
+  \includegraphics[height=9cm]{figs/Phase_Diagram_of_P_vs_T}
+  \caption{Phase Diagram of the reduced pressure verses the reduced 
+  temperature. The blue region is solid, and the red region is liquid.
+  The white line marks the solid-liquid boundary as predicted by 
+  Monte-Carlo simulations.}
+  \label{fig:Phase_Diagram_P_vs_T}
+\end{figure}
+
+Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
+$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
+for temperatures of kT = 0.5 to 3.0. 
+The black line marks the phase boundary between the solid and liquid regions, 
+and indicates the various temperature-pressure combinations at which a 
+phase transition occurs, and where various amounts of solid and liquid 
+can coexist. 
+It is seen that the higher the temperature, the higher the pressure must be
+to undergo a phase transition from a liquid to a solid, as would be expected. 
+The slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
+\begin{align} 
+    \frac{dP}{dT}=\frac{S_l-S_s}{V_l-V_s}
+\end{align}
+Since the entropy and the volume of the liquid is greater than the entropy
+and the volume of the solid, the slope should be positive.
+Comparing the slope of the phase boundary predicted by cDFT to that
+from Monte-Carlo simulations shown by the white line, it is seen that
+the cDFT transition temperatures are consistently higher, and the 
+discrepancy grows as the temperature increases. For example, at a 
+temperature of kT=1, the phase transition reduced pressure predicted by 
+cDFT is 15.24, whereas the Monte-Carlo simulation predicted a value of 9.99,
+and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
+Monte-Carlo simulation predicts 43.80.
+This may indicate that either the difference between the liquid 
+and solid phase transition entropies predicted by cDFT is too high, or 
+the difference in the volumes is too low. 
+%
 %Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
@@
 %pressures and temperatures are consistently higher.
-
-\begin{figure}
-  \centering
-  \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
-  \caption{Phase Diagram of the reduced pressure verses the reduced temperature. 
-  The white line marks the boundary from Monte-Carlo simulations.}
-  \label{fig:Phase_Diagram_P_vs_T}
-\end{figure}
-
+%
+%
 \begin{figure}
@@
 \end{figure}
-
+%
 %Figure~\ref{fig:p-vs-T_at_fixed_density} shows individual plots of the 
@@
 %Figures~\ref{fig:Phase_Diagram_P_vs_T} and \ref{fig:Phase_Diagram_P_vs_T_AS}.
-
+%
 \begin{figure}
@@
 \chapter{Conclusion}
-In this paper we were successful in predicting the freezing of a 
-Weeks-Chandler-Anderson fluid using our SFMT functional built by 
-forming temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ to use 
-in Schmidt's erf model to better fit our results to those expected for a
-WCA fluid. We were able to generate a temperature-density 
+We were successful in predicting the freezing of a Weeks-Chandler-Anderson 
+fluid using classical Density Functional Theory. To construct a Helmholtz 
+free energy functional $F[n(r)]$, we used SFMT with Schmidt's error 
+function model, and defined two temperature dependent parameters, $\alpha$ 
+and $\Xi$, to model a WCA fluid. 
+$\alpha$ was found by setting the error function potential equal to the 
+WCA potential at $r=\alpha$, and $\Xi$ was found for each temperature by 
+setting the second virial coefficients for the error function potential 
+and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$.
+To implement cDFT we used Gaussian distribution functions centered at 
+lattice points to represent the number density profile which was then varied
+by changing the width of the Gaussians, and the fraction of vacant lattice
+sites until the Helmholtz free energy was minimized. Once the equilibrium 
+Helmholtz free energy for the liquid and solid states were found for a given
+temperature and number density, Maxwell's Double Tangent Construction 
+was used to find the number densities of the liquid and the solid at 
+the phase transition.
+Using this information, we generated a temperature-density 
 phase diagram that yielded coexistence regions that approximately 
@@
 % and molecular simulation 
-results. 
+results, though shifted a bit toward higher densities. 
 %for values of kT=0.5 to 3. 
 %kT=0.5 to 38
-We are also able to generate a pressure-temperature
-phase diagram with a shape that is expected, but the transition 
-temperatures and pressures were too high.
+We also generated a pressure-temperature
+phase diagram, but the slope was notably steeper than that
+predicted from Monte-Carlo simulations, giving transition 
+temperatures and pressures that were too high.
 
@@
 
-\section{Third Virial Coefficient $B_{3}$}
+\section{Third Virial Coefficient $B_{3}$ General Approach}
 
@@
 
-\section{Derivation of the Virial Equation}
+\section{General Derivation of the Virial Equation}
 The second virial coefficient can be derived from Euler's equation for 
@@
 
-\section{Derivation of $F = -k_BTln(Z)$}
+\section{Derivation of $F = -k_BTln(Z)$ for Continuous Systems}
 The canonical equilibrium phase-space probablity density is given by

### 2020-11-27 papers/thesis-kirstie/thesis.tex
modified abstract

@@
   We use Classical Density Functional Theory (cDFT) to predict the 
-  freezing of a Weeks-Chandler-Anderson (WCA) fluid. A WCA fluid 
-  consists of soft spheres that exhibit the repulsive portion of the 
-  Lennard-Jones potential widely used to model the interaction of two 
-  neutral atoms. Our cDFT functional is constructed using Soft Fundamental 
-  Theory (SFMT) with Tarazona's tensor expression for $\Phi_3$, and Schmidt's
-  erf model in which we have incorporated temperature dependent
-  parameters to fit the results to those predicted by Monte-Carlo simulations.
-  Our functional was successful in predicting the freezing of the WCA fluid 
-  with a fairly good match to Monte-Carlo simulations. This result suggests 
-  that SFMT may provide a strong foundation for future density functionals 
-  that will model real fluids.
+  freezing of a Weeks-Chandler-Anderson (WCA) fluid, and construct
+  pressure-temperature and temperature-density phase diagrams. 
+  A WCA fluid consists of soft spheres that exhibit the repulsive 
+  portion of the Lennard-Jones potential widely used to model the 
+  interaction of two neutral atoms. Our cDFT functional is constructed 
+  using Soft Fundamental Theory (SFMT) with Tarazona's tensor expression 
+  for $\Phi_3$, and Schmidt's error function model in which we have 
+  incorporated our temperature dependent parameters $\alpha$ and $\Xi$. 
+  $\Xi$ is found by making the second virial coefficient for the error 
+  function potential equal to the second virial coefficient for the WCA 
+  potential. Our functional was successful in predicting the freezing of 
+  the WCA fluid with a fairly good match to Monte-Carlo simulations. 
+  This result suggests that SFMT may provide a strong foundation for 
+  future density functionals that will model real fluids.
 }

### 2020-11-28 papers/thesis-kirstie/thesis.tex
fixed abstract again
more edits

@@
   incorporated our temperature dependent parameters $\alpha$ and $\Xi$. 
-  $\Xi$ is found by making the second virial coefficient for the error 
-  function potential equal to the second virial coefficient for the WCA 
-  potential. Our functional was successful in predicting the freezing of 
-  the WCA fluid with a fairly good match to Monte-Carlo simulations. 
+  We find $\Xi$ by making the second virial coefficients of the error 
+  function potential and the WCA potential equal. Our functional was 
+  successful in predicting the freezing of the WCA fluid with a fairly 
+  good match to Monte-Carlo simulations. 
   This result suggests that SFMT may provide a strong foundation for 
@@
 potential energy that is zero when two spheres are not in contact, but 
-goes instantly to infinity when contact is made. The inability to push 
+goes instantly to infinity when contact is made, as shown in 
+Figure~\ref{fig:HardSphere_potential}. The inability to push 
 real atoms closer together is due to the repulsion of the electrons in 
@@
 The repulsive portion of the Lennard-Jones potential forms the potential for a 
-Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}, 
-shown in Figure \ref{fig:WCA_potential}. 
+Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}. 
 The WCA potential is described by the equation: 
@@
 equilibrium), and the same chemical potential (necessary for diffusive 
-equilibrium), an abrupt phase transition occurs, that is, the liquid freezes 
-forming a crystalline structure. The particular crystalline structure and 
-density that result are affected by the interatomic potential. When repulsive, 
-soft sphere fluids (that are not too soft) freeze, they form lattice 
-structures composed of Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, 
-as do hard spheres. 
+equilibrium), an abrupt phase transition occurs, that is, the liquid 
+freezes. The particular structure and density that result are affected 
+by the interatomic potential. When repulsive, soft sphere fluids 
+(that are not too soft) freeze, they form lattice structures composed of 
+Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
 This structure is what one would see by filling a cube with marbles. It 
@@
 the atoms in a WCA fluid do not exhibit attraction there will not be a 
-dividing line between the gas and liquid phases, and so the term "liquid" 
-or "gas" is not used, but simply "fluid". Points along the line that 
-separates the solid and the fluid phases indicate a phase transition, 
-and show that the solid and the fluid both have the same temperature and 
-the same pressure at a phase transition.
+dividing line between the gas and liquid phases, there will just simply
+be a fluid, which might be referred to as a "liquid". Points along the 
+line that separates the solid and the liquid phases indicate a phase 
+transition, and show that the solid and the liquid both have the same 
+temperature and the same pressure at a phase transition.
 
@@
 Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
-and one for the solid at a fixed temperature. The slope of the curves is 
+and one for the solid at a fixed temperature. Each curve consists of a 
+collection of data points $(\frac{F}{N}, \frac{1}{n})$ for the liquid and
+the solid respectively.
+We will use cDFT (discussed in the next section) to find the Helmholtz
+free energies of the liquid and the solid at a given temperature and 
+number density needed to create these curves. 
+The slope of the curves is 
 related to the pressure. Since the pressure of the 
@@
 the density of the liquid and the density of the solid at the point of 
-freezing can be found from identifying the inverse densities at which the 
+freezing can be found from identifying the points at which the 
 tangents to each curve have the same slope. This process is called 
@@
 \begin{align}{f=-P\frac{1}{n}+g}\end{align}
-This equation has the form of a line:
+This equation has the form of a line
 \begin{align}{y=mx+y_o}\end{align}
@@
 At the point of phase transition the chemical potential of the liquid is 
-the same as the chemical potential of the solid.
-\begin{align}{g_L=g_S}\end{align} 
-\begin{align}{f_L+P\frac{1}{n_L}=f_S+P\frac{1}{n_S}}\end{align}
-
-The pressure at the transition point can be obtained from the point of 
-intersection that occurs when the Gibbs free energy per atom is plotted 
-against the pressure for both the liquid and the solid at fixed temperature, 
-as shown in Figure~\ref{fig:GibbsvsP}. These curves are constructed from 
-the slope and corresponding y-intercept data collected from lines tangent 
+the same as the chemical potential of the solid, as is the pressure.
+\begin{align}{g_L=g_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}}\end{align} 
+The pressure at the phase transition can be obtained from the point of 
+intersection that occurs when the Gibbs free energy per atom, or chemical
+potential, is plotted against the pressure for both the liquid and the 
+solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
+These curves are constructed from data collected from lines tangent 
 along the curves on the the Helmholtz free energy per atom verses inverse 
@@
 liquid and the solid at a given temperature. 
+The slope gives -P and corresponding y-intercept gives $g$. The 
+curves in Figure~\ref{fig:GibbsvsP} consist of a collection of data 
+points ($g$,P) for the liquid and the solid respectively.
 
@@
 liquid at the transition point can be found from a plot of the pressure 
-versus the inverse of the density for the liquid. The density of the solid 
-at the transition point can be found the same way, as shown in Figure 
-\ref{fig:Pvsinvn}. A Temperature-Density (T-n) phase diagram, such as 
+versus the inverse of the density for the liquid, as shown in Figure 
+\ref{fig:Pvsinvn}. The density of the solid at the transition point can 
+be found the same way. A Temperature-Density (T-n) phase diagram, such as 
 that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this 
@@
 a minimum at equilibrium means that the equilibrium free energy of the 
-system ${f_{equil}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
-number density profile $\rho{(}\vec{r}{)}$, can be found simply by varying 
+system ${f_{eq}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
+number density profile $n_{eq}(\vec r)$, can be found simply by varying 
 $n(\vec{r})$ through all possible spatial profiles until the free energy 
 is minimized~\cite{MoritaDFT}. 
-
-\begin{displaymath}{f[n(}\vec{r}{)]}\rightarrow{f_{equil}}{[}\rho{(}\vec{r}{)]}  \mbox{ as $f$ is minimized} \end{displaymath}
-\begin{displaymath}{\mbox{where }  n(\vec{r})=\rho{(}\vec{r})  \mbox{ at equilibrium}}\end{displaymath}
+\begin{displaymath}f[n(\vec r)]\rightarrow f_{eq}[n_{eq}(\vec r)]  \mbox{ as $f$ is minimized} \end{displaymath}
+\begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
+%The reason this works is because DFT satisfies the conditions of the 
+%variational principle, the same mathematical principle used to determine 
+%that the shortest distance between two points is a straight line by 
+%varying the path between the two points until the distance is minimized.
 
@@
 \begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi=\frac{V_{ext}}{N}$ 
-is the externally 
-applied potential energy $V_{ext}(\vec r)$ per particle, 
-and the Helmholtz free energy is expressed as
-\begin{equation}F = F_{int} + \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}
+where $\phi=\frac{V_{ext}}{N}$ is an externally applied potential energy 
+$V_{ext}(\vec r)$ per particle, and an intrinsic Helmholtz free energy 
+is defined as
+\begin{equation}F_{int} = F - \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}
 
@@
 \begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
-and so 
+and so the Grand free energy is constrained
 \begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
@@
 
-The minimization can be illustrated by considering that according to the 
+The minimization can be illustrated for a system near 
+equilibrium by considering that according to the 
 second law of thermodynamics, a system that is not in equilibrium will 
@@
 when the natural variables of the free energy are held fixed~\cite{schroeder}. 
+%Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
+%\begin{equation}
+  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
+%\end{equation}
+%and holding the temperature fixed ($T_{System}=T_{Surroundings}$) so that 
+%$dT=0$ gives
+%\begin{equation}
+  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
+%\end{equation}
+%where conservation of energy was also applied. Fixing the volume and the 
+%number of particles, so that $dV=0$ and $d\left(\frac{1}{N}\right)=0$, 
+%results in
+%\begin{equation}
+  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+%\end{equation}
+%which, for fixed N, is the same as
+%\begin{equation}dF_{System}=-TdS_{System+Surroundings} \end{equation} 
+%Since $dS_{System+Surroundings} \geq{0}$, and the temperature and number 
+%of particles are fixed, the change in the free energy of the system will 
+%always be negative, and thus the free energy tends toward a minimum.
+%~\cite{schroeder}. 
 A general proof of minimization 
@@
 by a gas at low density where the atoms rarely encounter one another. 
-he excess partitian function $Z_{excess}$ accounts for the interaction 
-of all the atoms according to their interatomic potential. The free energy 
-can also be broken into two parts using Equation~\ref{Zmultiply} giving
+The excess partitian function $Z_{excess}$ accounts for the interaction 
+of all the atoms according to their interatomic potential. Using 
+Equation~\ref{Zmultiply}, the free energy can also be broken into two parts.
 \begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
@@
 
-Now consider the excess free energy which is written most generally as
+The excess Helmholtz free energy is written most generally as
 \begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
@@
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
-The function $f_{ij}$ is called a Mayer function, and is given by
-\begin{equation}\label{eq:mayerfunction}{f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1}\end{equation} 
-Multiplying the Mayer function terms, the excess free energy can be written as
+The function $f_{ij}$ the Mayer function given by
+\begin{align}\label{eq:mayerfunction}
+     f_{ij}(\vec{r_i},\vec{r_j})=e^{-\frac{V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}-1
+\end{align} 
+Multiplying the Mayer function terms, the excess free energy can be 
+written as~\cite{schroeder}
 \begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +...
@@
 \begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
-in the thermodynamic limit of large N 
-where the quantity in parenthesis can be recognized 
+in the thermodynamic limit of large N. 
+The quantity in parenthesis can be recognized 
 as the second virial expansion coefficient typically called $B_2$. 
-It is important that the expression for the excess Helmholtz free energy 
+It is important that an expression for the excess Helmholtz free energy 
 correctly reduces to this result in the low-density limit. 
@@
 excess Helmholtz free energy given in Equation~\ref{eq:Fexcess-simplified}. 
+SFMT, discussed in the next section, provides a way around this problem
+for some types of fluids.
 
@@
 until the free energy is minimized.  
-In this section 
-a functional for the Helmholtz free energy was formed by finding 
-general expressions for the ideal and the excess portions of the free 
-energy, and adding them together.
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
-The functional for the ideal portion of the Helmholtz free energy is given by 
-Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
-free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
-form for the excess Helmholtz free energy depends on the particular 
-fluid being modeled as defined by the  
-interatomic potential energy V(r) between the spheres, which can make
-forming an expression difficult. In this paper, Soft Fundamental Measure
-Theory will be used to form an expression for the excess Helmholtz free 
-energy, as explained in the next section.
+%In this section 
+%a functional for the Helmholtz free energy was formed by finding 
+%general expressions for the ideal and the excess portions of the free 
+%energy, and adding them together.
+%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+%The functional for the ideal portion of the Helmholtz free energy is given by 
+%Equation~\ref{eq:Fideal-n(r)}, and a functional for the excess portion of the Helmholtz 
+%free energy is given in Equation~\ref{eq:Fexcess-simplified}. The specific 
+%form for the excess Helmholtz free energy depends on the particular 
+%fluid being modeled as defined by the  
+%interatomic potential energy V(r) between the spheres, which can make
+%forming an expression difficult. In this paper, Soft Fundamental Measure
+%Theory will be used to form an expression for the excess Helmholtz free 
+%energy, as explained in the next section.
 

### 2020-11-28 papers/thesis-kirstie/thesis.tex
more edits

@@
     \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
-    \caption{A Hard Sphere Potential is zero for $r>2R$, and infinity 
+    \caption{A hard-sphere Potential is zero for $r>2R$, and infinity 
     for $r\leq{2R}$ where R is the radius of a hard sphere (which is 
@@
 The potential is zero when the distance r between the centers of two atoms 
-is equal to $\sigma$ (shown in the figure at r/$\sigma$=1). At this point 
+is equal to $\sigma$. At this point 
 the attraction between the atoms just cancels their repulsion. As the distance 
@@
 rapidly and approaches infinity as the two spheres begin to overlap. This 
-is a more realistic repulsive potential than that of the Hard Sphere model 
+is a more realistic repulsive potential than that of the hard-sphere model 
 which goes instantly to infinity when the two spheres make contact (at r=2R), 
@@
 dividing line between the gas and liquid phases, there will just simply
-be a fluid, which might be referred to as a "liquid". Points along the 
-line that separates the solid and the liquid phases indicate a phase 
-transition, and show that the solid and the liquid both have the same 
-temperature and the same pressure at a phase transition.
+be a fluid, which we will refer to as a liquid in this paper.
+%which may be referred to as a "liquid", since at the phase boundary the fluid is dense. 
+Points along the line that separates the solid and the liquid phases 
+indicate a phase transition, and show that the solid and the liquid 
+both have the same temperature and the same pressure at a phase transition.
 
@@
 free energy per atom, $f$,  with respect to the inverse number density. 
-\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} Thus, the pressure corresponds to the slope of the curve on a plot of the Helmholtz free energy per atom versus the inverse density at fixed temperature.
+\begin{equation}{P=-\frac{\partial{\frac{F}{N}}}{\partial{\frac{V}{N}}}\bigg|_{T,N} = -\frac{\partial{f}}{\partial{\frac{1}{n}}}\bigg|_{T,N}}\end{equation} 
+Thus, the pressure corresponds to the negative of the slope of the curve 
+on a plot of the Helmholtz free energy per atom versus the inverse 
+density at fixed temperature.
 
@@
 collection of data points $(\frac{F}{N}, \frac{1}{n})$ for the liquid and
-the solid respectively.
-We will use cDFT (discussed in the next section) to find the Helmholtz
-free energies of the liquid and the solid at a given temperature and 
-number density needed to create these curves. 
-The slope of the curves is 
-related to the pressure. Since the pressure of the 
+the solid respectively, which will be obtained by using cDFT to find the 
+Helmholtz free energies of the liquid and the solid for various 
+temperatures and number densities. 
+Since the pressure of the 
 liquid is the same as the pressure of the solid at the point of transition, 
@@
 
+Classical Density Fucntional Theory will be used to find the 
+Helmholtz free energies of the liquid and the solid for various 
+temperatures and number densities, so that phase diagrams can be 
+constructed using Maxwell's Double Tangent Construction. 
 The key idea behind classical Density Functional Theory (cDFT) is that 
@@
 \begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
-%The reason this works is because DFT satisfies the conditions of the 
-%variational principle, the same mathematical principle used to determine 
-%that the shortest distance between two points is a straight line by 
-%varying the path between the two points until the distance is minimized.
 
@@
 
+%The reason the minimization works is because DFT satisfies the conditions of the 
+%variational principle, the same mathematical principle used to determine 
+%that the shortest distance between two points is a straight line by 
+%varying the path between the two points until the distance is minimized.
 The minimization can be illustrated for a system near 
@@
 when the natural variables of the free energy are held fixed~\cite{schroeder}. 
+And so over time teh free energy settles to a minimum. 
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
@@
 %~\cite{schroeder}. 
-A general proof of minimization 
+A proof of minimization in general
 can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
@@
 Z is integrated over all of phase space for N identical particles each with 
-a momentum $\vec{p}$ and a position $\vec{r}$ is given by
+a momentum $\vec{p}$ and a position $\vec{r}$, and is given by
 \begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
-\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~where~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
-\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~where~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
-
-Breaking the Hamiltonian, H into its component energies
+where
+\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~and~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
+\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~and~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
+Writing the Hamiltonian, H in terms of its component kinetic and potential energies gives
 \begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
@@
 \begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
-the potential energy of the total interaction of each atom with the other 
+and the potential energy of the total interaction of each atom with the other 
 N-1 atoms is given by
 \begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
-and \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
+and where \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
 is an externally applied potential that varies spatially, such that its 
@@
 the integral in equation (8) can be separated into two parts 
-\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume satisfying \begin{equation}{V^N=}\int{...}\int{dQ}\end{equation}
+\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
+Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume, and 
+\begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
 the partitian function can be expressed as
@@
 where
-\begin{equation}{Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}}\end{equation}
-\begin{equation}{Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}}\end{equation} 
+\begin{align} \label{eq:Zideal}
+    Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}
+\end{align}
+\begin{align} \label{eq:Zexcess}
+    Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
+\end{align} 
 The ideal partitian function $Z_{ideal}$ is associated with an ideal gas 
@@
 \begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
-\begin{equation}{F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})}\end{equation}
+\begin{align} \label{eq:Finparts}
+    F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})
+\end{align}
 \begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}
 
-The ideal Helmholtz free energy is
+Putting Equation~\ref{eq:Zideal} into Equation~\ref{eq:Finparts}, 
+the ideal Helmholtz free energy is
 \begin{displaymath}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{displaymath}
@@
 
-The excess Helmholtz free energy is written most generally as
+Putting Equation~\ref{eq:Zexcess} into Equation~\ref{eq:Finparts}, 
+the excess Helmholtz free energy is written most generally as
 \begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
@@
 written as~\cite{schroeder}
-\begin{equation}\label{eq:Fexcess-simplified}{F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +...
-\right]\right)  }}\end{equation}
+\begin{align} \label{eq:Fexcess-simplified}
+    F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)}
+\end{align}
 
@@
 
+Equation~\ref{eq:Fexcess-simplified} is difficult to solve in general, 
+but it can be solved for low densities, and this provides a way to check
+a more general solution.
 At low densities, it is not likely to find two or more atoms in close 
@@
 Soft Fundamental Measure Theory (SFMT) is a method for forming an 
-expression for the excess Helmholtz free energy of a Soft-sphere fluid
+expression for the excess Helmholtz free energy of a soft-sphere fluid
 that does not depend on  explicitly taking into account the interaction 
-that each sphere with every other sphere in the fluid as required by 
-Equation~\ref{eq:Fexcess-simplified}. In the next chapter on methods, we
+that each sphere with every other sphere in the fluid (as required by 
+Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
 will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
 
-SFMT is an extension of Fundamental Measure Theory (FMT)
-that was developed for the Hard-sphere fluid, and which can be derived from
-Scaled Particle Theory (SPT).  
+SFMT is an extension of Fundamental Measure Theory (FMT)~\cite{rosenfeld1989}
+that was developed for the hard-sphere fluid. FMT can be derived from
+Scaled Particle Theory (SPT)~\cite{santos2012phi3}.  
 Scaled Particle Theory seeks a general expression for the excess chemical 
-potential of a homogeneous Hard Sphere fluid which can then be integrated 
+potential of a homogeneous hard-sphere fluid which can then be integrated 
 to find the excess Helmholtz free energy according to the relation
@@
 
-The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering
-that at each point in the fluid, there is a probability that 
-a hard sphere of radius R has space available to exist at that point
-without overlapping with other spheres, or it does not, as shown in 
-Figure~\ref{fig:p_overlap}. 
+The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
+there will be a change in the potential energy of $V_{diff}$ if one sphere
+is added, but $V_{diff}$ will be zero if no sphere is added. 
+%at each point in the fluid, there is a probability that 
+%a hard sphere of radius R has space available to exist at that point
+%without overlapping with other spheres, or it does not, as shown in 
+Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
+\begin{align} 
+	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
+	         &= V_{diff} \\
+             &= W_{rev}(R)  \label{workrev}
+\end{align}
+The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
+that it takes no energy for a hard sphere to be placed at a point
+in the fluid where it does not overlap with any other spheres 
+($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
+to place a sphere where it would overlap another sphere, as described 
+in Figure~\ref{fig:p_overlap}. 
 \begin{figure}[h!]
@@
   \end{figure}
-Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
-\begin{align} 
-	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
-	         &= V_{diff} \\
-             &= W_{rev}(R)  \label{workrev}
-\end{align}
-The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
-that it takes no energy for a hard sphere to be placed at a point
-in the fluid where it does not overlap with any other spheres 
-($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
-to place a sphere where it would overlap another sphere, as described 
-in Figure~\ref{fig:p_overlap}.
 In this case Equation~\ref{widoms-insertion-formula} becomes
@@
 \end{align}
-where the radius of the cavity $R_{cavity}$ equals the radius of a hard sphere R,
-and $nV_{Sphere}$ is the probability that a sphere cannot be placed at
-some random point in the homogeneous fluid of number density, $n$, without
-overlapping with another sphere. Combining  Equations~\ref{workrev} and \ref{muexcess} gives
-\begin{align}
-  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
-\end{align}
+where the radius of the cavity $R_{cavity}$ equals the radius of a 
+hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
+occupied by the spheres divided by the total volumet avaliable, 
+is the probability that a sphere cannot be placed at some random
+point in the homogeneous fluid of number density, $n$, without
+overlapping with another sphere. Combining  Equations~\ref{workrev} 
+and \ref{muexcess} gives
+\begin{align}
+  %W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3)
+  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-\eta) 
+\end{align}
+where $\eta$ is the packing fraction defined as
+\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
+\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
 Forming a more general result by requiring continuity of the functions, 
 along with their first and second derivatives, for the reversible work
-written for fluids with a cavity radius 
-equal to, larger than, or very 
+written for fluids with a cavity radius equal to, larger than, or very 
 much larger than the hard sphere radius R, Scaled Particle Theory arrives 
-at an expression~\cite{Hansen} for the excess Helmholtz free energy of 
-a homogeneous Hard-Sphere fluid in terms of the packing fraction $\eta$ 
-\begin{equation}\label{Fexcess-SPT}{\frac{\beta{F_{excess}}}{N}=-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}}\end{equation} 
-where
-\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
-\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
-
-Fundamental Measure Theory~\cite{rosenfeld1989} (FMT) 
-can be constructed from Equation~\ref{Fexcess-SPT}~\cite{santos2012phi3},  
-and provides an expression for the excess Helmholtz free energy of an 
-inhomogeneous 
-Hard-Sphere fluid as a functional of the number density profile $n(\vec r)$. 
-Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
-is an extension of FMT and provides an expression for the excess Helmholtz
-free energy of fluids with soft potentials like that of the WCA fluid. 
-It will be helpful to continue looking at FMT before addressing SFMT.
-
-Fundamental Measure Theory begins by noticing that 
-from Equation~\ref{Fexcess-SPT}, a quantity $G_{3}$ can be defined as follows
+at an expression for the excess Helmholtz free energy of 
+a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
+For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
+\begin{align}\label{Fexcess-SPT}
+    F_{excess}=k_BT\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)N
+\end{align}
+
+Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
+%and provides an expression for the excess Helmholtz free energy of an 
+%inhomogeneous 
+%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
+%
+%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
+%is an extension of FMT and provides an expression for the excess Helmholtz
+%free energy of fluids with soft potentials like that of the WCA fluid. 
+%It will be helpful to continue looking at FMT before addressing SFMT.
+%FMT begins 
+by noticing that a quantity $G_{3}$ can be defined such that
 \begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
-where $G_{3}$ is the volume of a sphere. Likewise, by defining
-quantities
+where $G_{3}$ is the volume of a sphere. 
+Likewise, by defining quantities
 \begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
@@
 geometric measures of a sphere, and are what give rise to the name
-Fundamental Measure Theory. Introducing a set of weighted densities $n_{i}=nG_{i}$ 
+Fundamental Measure Theory. Introducing a set of variables
+%weighted densities $n_{i}=nG_{i}$ 
 given by
@@
 \end{align}
-or for a spatially varying number density $n(\vec{r})$, 
-\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
 where 
@@
 allowing for their extension to a system with a spatially varying number 
-density, that is, an inhomogeneous fluid. Using a backwards Heavyside step function defined as
+density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
+excess Helmholtz free energy is written as
+\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
+Using a backwards Heavyside step function defined as
 \begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
@@
 Evaluating Equation with $V(r)$ given by Equation,  
-the mayer function for a Hard-sphere fluid is found to be
+the mayer function for a hard-sphere fluid is found to be
 \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
@@
 \end{align} 
-This can easily be checked by evaluating the left side for a Hard-sphere fluid as follows,
+This relation can easily be checked for a hard-sphere fluid where
 \begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
 \begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
-and evaluating the right side by using the general relation for convolution which is equal to the Inverse 
-Fourier Transform of a product of Fourier Transforms. 
+The right side can be evaluated by using the general relation for 
+convolution which is equal to the Inverse Fourier Transform 
+of a product of Fourier Transforms. 
 \begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
@@
 Tarazona later improved upon the original FMT equations by replacing the
-expression for $\Phi3$ given by Equation~\ref{oldPhi3equation} with the
+expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
 more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
@@
 
-The FMT equations developed so far have been for a Hard-sphere fluid.
+The FMT equations developed so far have been for a hard-sphere fluid.
 All of the weight functions, the relations between them, and the equation 
 for the derivative of the mayer function, can be expressed in terms of 
-$w_{2}(r)$ which, for a Hard-sphere fluid, consists of a delta function.
+$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
 The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
@@
 but replacing $w_{3}(r)$ with a weight function 
-appropriate for a soft sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+appropriate for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
 This is what we do in forming a functional for the Weeks-Chandler-Anderson 

### 2020-11-29 papers/thesis-kirstie/thesis.tex
more edits

@@
 purely liquid and the purely solid regions. Within the gap, or coexistence
-region, is found mixtures of various amounts of liquid and solid.  
-Within the coexistence region, the liquid and solid have number densities 
-that are fixed at the values they had at the phase transition. To construct
+region, is found mixtures of various amounts of liquid and solid each of
+which have number densities that are fixed at the values they had at the 
+phase transition. To construct
 the phase diagram, these liquid and solid phase transition number densities
@@
 these number densities given plots of the liquid and the solid Helmholtz 
-free energies verses the inverse density at a given temperature. 
-To construct these plots we use classical Density Functional Theory 
+free energies verses the inverse density at a given temperature, as
+discussed in Section 2.2. We use classical Density Functional Theory 
 to find the Helmholtz free energies of the liquid and the solid. 
@@
 form a Face Centered Cubic (FCC) crystal in its solid state~\cite{Hansen}. 
-Although the entire 
+Although an entire 
 crystal lattice can be constructed with FCC cells, 
-FCC cells are not the smallest cells with which the lattice can be built. 
+they are not the smallest cells with which the lattice can be built. 
 The shaded parallelepiped shown within the FCC cell in Figure \ref{fig:primitivecell} 
@@
 The three lattice vectors will be useful in 
-creating the lattice structure computationally, and will also make the 
+creating the lattice structure computationally, and will also make our 
 program run faster.
@@
 of positions about the lattice point. 
-Of course, atoms are always in motion, so when talking about "the location" 
+Atoms are always in motion, so when talking about "the location" 
 of an atom, we are talking about a time averaged position, so that it 
@@
 of four possible positions an atom (shown in blue) might have in a crystal 
-cell that has one lattice site (represented by a black dot the center of 
-the cell). The atom is slightly off the lattice site in each example, and 
+cell that has one lattice site at its center (represented by a black dot). 
+The atom is slightly off the lattice site in each example, and 
 a sampling of many possibilities - more than just the four given 
-as examples - would create an average profile that looks like a Gaussian
-distribution, as shown on the right.
+as examples - would create an average distribution profile that looks 
+like a Gaussian distribution, as shown on the right.
 A plot of a Gaussian distribution in 2-dimensions is shown in 
@@
 Lennard-Jones model). 
-
-3-dimensional Gaussian distributions centered at lattice points can be 
+Three-dimensional Gaussian distributions centered at lattice points can be 
 used to represent a number density profile $n(\vec{r})$  which describes 
@@
 of various possible spatial distributions of the atoms within a cell, or
-more precisely, as an ensemble average over all possible microstates.
+more precisely, as an ensemble average over all possible microstates, 
+each of which consists of a unique set of positions for the atoms in 
+that occupy the cell. Mathematically, the microstates are represented 
+by a set of delta functions.
 \begin{align}
@@
 This makes it convenient to describe the new 
-number density in terms of the fraction of vacancies, $f_v$: 
-\begin{displaymath}{ n = (1-f_v){n_{no.vacancies}}}\end{displaymath}
+number density in terms of the fraction of vacancies, $f_v$. 
+The resulting number density is then given by
+\begin{align}
+    n = (1-f_v){n_{no.vacancies}}
+\end{align}
 
@@
 
-By varying the width of the Gaussians and the fraction of vacancies, 
-various number density profiles can be generated for a given combination 
-of temperature and number density, as shown in Figure~\ref{fig:Ensemble_vary}.
-A homogeneous number density profile is used for a liquid is shown in 
+To implement classical Density Functional Theory, 
+a set of number density profiles is generated by varying the Gaussian width 
+and the fraction of vacancies at a given combination of temperature and 
+number density, as shown by the example in Figure~\ref{fig:Ensemble_vary}.
+A homogeneous number density profile is used for a liquid, as shown in 
 Figure \ref{fig:homogen_denisty}.
@@
 The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
-density profile, and also for the homogeneous number density profile 
+density profile in the set, and also for the homogeneous number density profile 
 characteristic of a liquid. Both the ideal and excess free energies are 
 found, and added together to get the total Helmholtz free energy.
-According to classical Density Functional Theory, of all the Helmholtz 
-free energies calculated is the minimum Helmholtz free energy, and its 
-corresponding number density profile that are the actual equilibrium values 
-for that temperature and number density.
-The process is repeated to find the free energies at other combinations 
+According to cDFT, the actual equilibrium Helmholtz free energy at
+the given temperature and number density will be the one
+that corresponds to the number density profile that minimizes the functional. 
+This process is repeated to find the free energies at other combinations 
 of temperatures and number densities.
@@
 potential energy $V(r)$ that exists between each sphere with all the other spheres, 
-as is shown in Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
-Soft Fundamental Measure Theory (SFMT) is a method
+as is shown by Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
+Soft Fundamental Measure Theory is a method
 developed to form an expression for the excess Helmholtz free energy without
@@
 but it must be tailored to fit a particular fluid\cite{schmidt1999density}.  
-Our SFMT functional is taylored to fit a WCA fluid, and 
+Our SFMT functional is tailored to fit a WCA fluid, and 
 is written in general form as 
@@
 by Gaussian functions centered at each lattice point.
-There are four scalar weight functions, two vector weight functions, and 
-one tensor weight function all of which 
-depend on the expression for $w_3(r)$ 
-which in turn depends on the interatomic potential $V(r)$ that exists 
-between the spheres, and which defines the particular fluid being modeled.
-We introduce our temperature dependent parameters $\alpha(T)$ and $\Xi(T)$ 
-to fit the functional to a WCA fluid.
-The scalar weight functions are given by
+There are four scalar weight functions given by
 \begin{align}\label{eq:weights}
@@
 \end{align}
-The vector weights are given by 
+two vector weight functions given by 
 \begin{equation}\label{eq:w_v2}{  \vec{w_{2}}=w_{2}\frac{\vec{r}}{r}  }\end{equation}
 \begin{equation}\label{eq:w_v1}{  \vec{w_{1}}=w_{1}\frac{\vec{r}}{r}  }\end{equation} 
-The tensor weight is given by
+and one tensor weight function which is given by
 \begin{equation}\label{eq:tensorweight-methods}\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)\end{equation}
+All of the weight functions depend on the expression for $w_3(r)$ 
+which in turn depends on the interatomic potential $V(r)$ that exists 
+between the spheres, and which defines the particular fluid being modeled.
 
@@
 and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
-The two parameters $\alpha$ and $\Xi$ make temperature-dependent error 
-function potential $V_{erf}(r)$ approximate in the region of interest 
-of the temperature-independent WCA potential which is difficult 
-to implement computationally. $\Xi$ is determined by setting the second virial 
-coefficient $B_2$ of the error function potential $V_{erf}(r)$ and of the
-WCA potential equal.  
+The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
+function potential $V_{erf}(r)$ approximate, in the region of interest, 
+the temperature-independent WCA potential which is difficult 
+to implement computationally. 
+%$\Xi(T)$ is determined by setting the second virial coefficient $B_2$ of 
+%the error function potential $V_{erf}(r)$ equal to that of the
+%WCA potential, $B_{2erf}(T)=B_{2wca}(T)$ for a given temperature.  
 The error function potential emerges from $w_{3}(r)$, and the relationships 
@@
 The parameter $\alpha(T)$ is derived by setting 
-the error function potential $V_{erf}(r)$ given by Equation~\ref{eq:Verf} 
-equal to the WCA potential given by Equation~\ref{eq:VWCA} when $r=\alpha$. 
+the error function potential $V_{erf}(r)$ 
+equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$. 
+\begin{align}
+    V_{erf}(\alpha)=V_{WCA}(\alpha)
+\end{align}
 This gives
-\begin{equation}\label{alphaT}
-	{\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}}
-\end{equation} 
-Figure \ref{fig:alphaXivsT} shows how $\alpha$ varies with temperature 
+\begin{equation}\label{alphaT}\alpha(T)=\sigma\left(\frac{2}{1+\sqrt{\frac{k_BT}{\epsilon}\ln(2)}}\right)^\frac{1}{6}\end{equation} 
+Figure \ref{fig:alphaXivsT} shows how $\alpha(T)$ varies with temperature 
 when $\sigma=1$ and $\epsilon=1$.    
@@
 
-We determine the value of $\Xi$ at a given temperature $T$,
-by equating the second virial coefficient evaluated for the error 
-function potential $V_{\operatorname{erf}}(r)$, and the second virial 
-coefficient evaluated for the WCA potential $V_{WCA}$.
+We determine the value of $\Xi(T)$ at a given temperature $T$,
+by setting the second virial coefficient evaluated for the error 
+function potential $V_{\operatorname{erf}}(r)$ equal to the second virial 
+coefficient evaluated for the WCA potential $V_{WCA}(r)$.
 \begin{equation}B_{2\operatorname{erf}}(T) =B_{2WCA}(T)\end{equation}
-The second virial coefficient can be derived from Euler's equation for 
-the internal energy of a system and is given by
+The second virial coefficient is given by
 \begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(\vec{r})d\vec r\end{equation}
-As shown in detail in the Appendix, evaluating the second virial coefficient 
-$B_{2}$ for the error function potential $V_{erf}$ with the mayer function 
-given by Equation~\ref{eq:VerfMayerfunction} 
-gives 
+Using the mayer function for the error function potential given by 
+Equation~\ref{eq:VerfMayerfunction} we found (see Appendix) the second 
+virial coefficient $B_{2}$ for the error function potential $V_{erf}(r)$ 
+to be
 \begin{equation} \label{eq:B2erf}
@@
 \end{equation}
-The parameter $\Xi$ is found to be the value that makes Equations~\ref{eq:B2erf} 
+The parameter $\Xi(T)$ is found to be the value that makes Equations~\ref{eq:B2erf} 
 and \ref{eq:B2wca} equal each other for a given temperature T. 

### 2020-11-30 papers/thesis-kirstie/thesis.tex
more edits and fixed figure

@@
 where the translational kinetic energy of the N atoms is given by 
-\begin{displaymath}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{displaymath}
+\begin{equation}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{equation}
 and the potential energy of the total interaction of each atom with the other 
@@
 the ideal Helmholtz free energy is
-\begin{displaymath}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{displaymath}
+\begin{equation}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation}
 
@@
 
-\begin{displaymath}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{displaymath}    \begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
+\begin{equation}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{equation}
+\begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
@@
 Using a backwards Heavyside step function defined as
-\begin{displaymath}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{displaymath}
+\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
 the following expressions are obtained
@@
 This relation can easily be checked for a hard-sphere fluid where
-\begin{displaymath}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{displaymath}
+\begin{equation}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{equation}
 \begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
@@
 with soft potentials by keeping the relationships between the weight 
-functions (given by Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},
-~\ref{eq:w_v2},~\ref{w2_w3_relation},~\ref{eq:tensorweight}) and the 
-equation for the derivative of the mayer function 
-(given by Equation~\ref{mayerder=w2convolution}) the same, 
-but replacing $w_{3}(r)$ with a weight function 
-appropriate for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+functions, and the equation for the derivative of the mayer function 
+the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
+~\ref{w2_w3_relation},~\ref{eq:tensorweight},~\ref{mayerder=w2convolution}), 
+ but replacing $w_{3}(r)$ with a weight function appropriate 
+for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
 This is what we do in forming a functional for the Weeks-Chandler-Anderson 
@@
 which have number densities that are fixed at the values they had at the 
-phase transition. To construct
-the phase diagram, these liquid and solid phase transition number densities
+phase transition.
+
+To construct the phase diagram, the liquid and solid phase transition number densities
 must be found. Maxwell's double tangent construction can be used to find 
@@
 in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
-Using Equation~\ref{w2_w3_relation},
+Using Equation~\ref{w2_w3_relation} repeated here,
 \begin{displaymath}
@@
 \begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-NumPoints is the number of vectors $\vec r' - \vec R_j$ randomly generated about the jth lattice site, 
-and $n(\vec r' - \vec R_j)$
-is equal to a probability density 
-which when multiplied by dV gives the probability $P_{r'}$ at which the value of 
-$w_i(|\vec{r}-\vec{r}'|)$ will occur. 
-Using the relation
-\begin{displaymath}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{displaymath} 
+NumPoints is the number of vectors $\vec r' - \vec R_j$ randomly 
+generated about the jth lattice site, and $n(\vec r' - \vec R_j)$
+is equal to a probability density which when multiplied by dV gives 
+the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}'|)$ 
+will occur. Using the relation
+\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
 $n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
@@
 and a solid exists at higher number densities. 
-Phase transitions between the liquid and solid phases will occur when the liquid 
-and solid are at the same temperature, pressure and chemical potential. 
-Looking at one temperature, it is seen that as the density of the 
+For a given temperature, it is seen that as the density of the 
 liquid increases, a maximum number density is reached for the liquid. 
-This is the liquid phase transition number density, and 
-marks the beginning of a coexistence region where various amounts 
-of liquid and solid can exist together in equilibrium.
-Moving on toward higher number densities, the end of the coexistence 
-region is marked by the minimum number density that the solid can have.
-This is the solid phase transition number density. At a phase transition 
-there is an abrupt change in the number density as seen here where the
-density changes suddenly, and abruptly from the liquid 
-phase transition number density to the solid phase transition number density.
-Within the coexistence region, the number density of the liquid is fixed 
-at the liquid phase transition number density, and the number density of 
-the solid is fixed at the solid phase transition number density, while
-varying amounts of liquid and solid combine to give the infinite number 
-of possible densities seen within the coexistence region.
-
-\begin{figure}
-  \centering
-  \includegraphics[height=9cm]{figs/Phase_Diagram_of_T_vs_n}
-  \caption{Phase Diagram of the reduced temperature verses the reduced 
-  density for the WCA fluid. The red region is liquid, and the blue region is 
-  solid. The gray area that separates the liquid and solid regions 
-  is the region of coexistence where some liquid and solid exist together 
-  at varying proportions. White lines show the boundaries of the coexistence 
-  region generated from Monte-Carlo of the WCA fluid.}
-  \label{fig:Phase_Diagram_of_T_vs_n}
-\end{figure}
-
 Looking at the coexistence region as a whole for all temperatures, 
@@
 density of 1.00, giving a spread of 0.08.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n}
+  \caption{Phase Diagram of the reduced temperature verses the reduced 
+  density for the WCA fluid. The red region is liquid, and the blue region is 
+  solid. The gray area that separates the liquid and solid regions 
+  is the region of coexistence where some liquid and solid exist together 
+  at varying proportions. White lines show the boundaries of the coexistence 
+  region generated from Monte-Carlo of the WCA fluid.}
+  \label{fig:Phase_Diagram_of_T_vs_n}
+\end{figure}
+
 %The solid phase transition number densities predicted by each method 
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[height=9cm]{figs/Phase_Diagram_of_P_vs_T}
-  \caption{Phase Diagram of the reduced pressure verses the reduced 
-  temperature. The blue region is solid, and the red region is liquid.
-  The white line marks the solid-liquid boundary as predicted by 
-  Monte-Carlo simulations.}
-  \label{fig:Phase_Diagram_P_vs_T}
-\end{figure}
-
 Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
@@
 for temperatures of kT = 0.5 to 3.0. 
-The black line marks the phase boundary between the solid and liquid regions, 
-and indicates the various temperature-pressure combinations at which a 
-phase transition occurs, and where various amounts of solid and liquid 
-can coexist. 
+The black line marks the phase boundary between the solid and liquid regions. 
 It is seen that the higher the temperature, the higher the pressure must be
 to undergo a phase transition from a liquid to a solid, as would be expected. 
-The slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
+Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
 \begin{align} 
@@
 the difference in the volumes is too low.
+
+\begin{figure}
+  \centering
+  \includegraphics[height=10cm]{figs/Phase_Diagram_of_P_vs_T}
+  \caption{Phase Diagram of the reduced pressure verses the reduced 
+  temperature. The blue region is solid, and the red region is liquid.
+  The white line marks the solid-liquid boundary as predicted by 
+  Monte-Carlo simulations.}
+  \label{fig:Phase_Diagram_P_vs_T}
+\end{figure}
 %
@@
 \end{multline}
-Evaluating the integral by doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$, and setting the lower limit to $-\infty$  since the integrand is nearly zero by the time r is zero this becomes
+Evaluating the integral by doing a u-substitution with $u=\frac{r-\frac{\alpha}{2}}{a}$, 
+and setting the lower limit to $-\infty$  since the integrand is 
+nearly zero by the time r is zero this becomes
 \begin{equation}{\widetilde{w}_3(k)=\frac{4\pi}{k^3}e^{-\frac{k^2a^2}{4}}\left[\left(1+\frac{k^2a^2}{2}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{k\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]}\end{equation}
-Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{w}_3(k)$ becomes
+Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
+$\widetilde{w}_3(k)$ becomes
 \begin{equation}{\widetilde{w}_3(k)=\frac{4\pi}{k^3}e^{-\frac{k^2\Xi^2}{8}}\left[\left(1+\frac{k^2\Xi^2}{4}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{k\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]}\end{equation}
@@
 \begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
-Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate system $(x',y'z')$ 
+Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated 
+coordinate system $(x',y'z')$ 
 \begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
@@
 \begin{displaymath}{r_{z'}=r\cos\theta'}\end{displaymath} 
-The components in the $x'$ and $y'$ directions vanish due to symmetry about $z'$. That leaves
+The components in the $x'$ and $y'$ directions vanish due to symmetry 
+about $z'$. That leaves
 \begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
@@
 \begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
-Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{\vec{w}}_1(k)$ becomes
+Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
+$\widetilde{\vec{w}}_1(k)$ becomes
 \begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
@@
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
-Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate system $(x',y'z')$ 
+Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate 
+system $(x',y'z')$ 
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
@@
 \begin{displaymath}{r_{z'}=r\cos\theta'}\end{displaymath} 
-The components in the $x'$ and $y'$ directions vanish due to symmetry about $z'$. That leaves
+The components in the $x'$ and $y'$ directions vanish due to symmetry 
+about $z'$. That leaves
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
@@
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
-Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for $\widetilde{\vec{w}}_2(k)$ becomes
+Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
+$\widetilde{\vec{w}}_2(k)$ becomes
 \begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
@@
 \begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
-Since the integration is over all space, the integral can be simplified by setting 
+Since the integration is over all space, the integral can be simplified 
+by setting 
 \begin{equation}{\vec{k}=k\hat{z}}\end{equation}
 \begin{equation}{\vec{k}\cdot\vec{r}=kr\cos\theta\hat{z}}\end{equation}
-Instead of constraining vector $\vec{k}$ to be in the z direction, however, a rotated coordinate system $(x',y',z')$ is used which is rotated from the (x,y,z) coordinate system in such a way that $\hat{z}'$ points in the direction of vector $\vec{k}$.
-After solving for the elements of the transformed weight function in coordinate system $(x',y',z')$, they will be expressed again in terms of vector $\vec{k}$ as it appears in coordinate system (x,y,z) to get the elements of the transformed weight function in coordinate system (x,y,z).%The weight components of interest are those corresponding to the component of vector $\vec{r}$ parallel to vector $\vec{k}$ which will be called $r_{z'}$, and to the two components of $\vec{r}$ perpendicular to vector $\vec{k}$ (and to each other) which will be called $r_{x'}$ and $r_{y'}$.
+Instead of constraining vector $\vec{k}$ to be in the z direction, 
+owever, a rotated coordinate system $(x',y',z')$ is used which is 
+rotated from the (x,y,z) coordinate system in such a way that $\hat{z}'$ 
+points in the direction of vector $\vec{k}$. After solving for the 
+elements of the transformed weight function in coordinate system 
+$(x',y',z')$, they will be expressed again in terms of vector $\vec{k}$ 
+as it appears in coordinate system (x,y,z) to get the elements of the 
+transformed weight function in coordinate system (x,y,z). 
 \begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
@@
 \begin{equation}{+\frac{2\pi\sqrt{\pi}}{k^3a}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
-Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for $\widetilde{w}_{{m2}_{xx}}(k)$ becomes
+Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for 
+$\widetilde{w}_{{m2}_{xx}}(k)$ becomes
 
@@
 %\subsection{The tensor weight components in the coordinate system (x,y,z)}
-\noindent The tensor weight components in the coordinate system $(x',y',z')$ are:
+\noindent The tensor weight components in the coordinate system 
+$(x',y',z')$ are:
 \begin{equation}\label{tensorcomp}{\widetilde{w'}_{m2}=\left(\begin{array}{ccc} \widetilde{w}_{{m2}_{x'x'}} & 0 & 0 \\ 0 & \widetilde{w}_{{m2}_{y'y'}} & 0 \\ 0 & 0 & \widetilde{w}_{{m2}_{z'z'}} \end{array}\right)'}\end{equation}
@@
 \begin{equation}{\widetilde{w}_{m2}(\vec{k})= (\text{I}-3\hat{z}'\hat{z}')\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
-\noindent In the $(x',y',z')$ coordinate system, vector $\vec{k}$ had been aligned with the $\hat{z}'$ direction so that 
+\noindent In the $(x',y',z')$ coordinate system, vector $\vec{k}$ had 
+been aligned with the $\hat{z}'$ direction so that 
 \begin{displaymath}{\hat{k}=\hat{z'}}\end{displaymath}
@@
 \begin{equation}{\widetilde{w}_{m2}(\vec{k})= (\text{I}-3\hat{k}\hat{k})\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
-Now  $\vec{k}$ can be expressed in the $(x,y,z)$ coordinate system, $\vec{k}=k_x\hat{x} + k_y\hat{y} + k_z\hat{z}$.
-The components of the tensor weight in the $(x,y,z)$ coordinate system are then given by
+Now  $\vec{k}$ can be expressed in the $(x,y,z)$ coordinate system, 
+$\vec{k}=k_x\hat{x} + k_y\hat{y} + k_z\hat{z}$. The components of the 
+tensor weight in the $(x,y,z)$ coordinate system are then given by
 \begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})= (\delta{ij}-3\frac{k_ik_j}{k^2})\widetilde{w}_{{m2}_{x'x'}}(\vec{k})}\end{equation}
@@
 %\subsection{Transform of tensor weight ${w}_{m2}(r)$ goes to zero as k goes to zero}
-\noindent Check that the transform of tensor weight ${w}_{m2}(r)$ goes to zero as k goes to zero:
+\noindent Check that the transform of tensor weight ${w}_{m2}(r)$ 
+goes to zero as k goes to zero:
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi\Xi^2}{3}-\frac{4\pi}{k^2}\right)e^{-\left(\frac{k^2\Xi^2}{8}\right)}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\left(\frac{k^2\Xi^2}{8}\right)}\sin(\frac{k\alpha}{2})}\end{displaymath}
@@
 
-\begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2\vec{n}_{v2}\cdot\vec{n}_{v2}-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left((\overleftrightarrow{n}_{m2}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
+\begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2\vec{n}_{v2}\cdot\vec{n}_{v2}-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}
+\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left((\overleftrightarrow{n}_{m2}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
 
 \begin{displaymath}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\frac{2}{3}n_2|\vec{n}_{v2}|^2-\operatorname{Tr}({\overleftrightarrow{n}_{m2}^3})
--n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)\color{white}\right)\color{black}}\end{displaymath} \begin{equation}{\color{white}\left(\color{black}-\frac{1}{3}{n_2}^2\operatorname{Tr}(\overleftrightarrow{n}_{m2})-\frac{1}{9}{n_2}^3+n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)+\frac{1}{3}{n_2}^3+\frac{2}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
+-n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)\color{white}\right)\color{black}}\end{displaymath} 
+\begin{equation}{\color{white}\left(\color{black}-\frac{1}{3}{n_2}^2\operatorname{Tr}(\overleftrightarrow{n}_{m2})-\frac{1}{9}{n_2}^3+n_2\operatorname{Tr}(\overleftrightarrow{n}_{m2}^2)+\frac{1}{3}{n_2}^3+\frac{2}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
 
@@
 The mayer function for the Error Function potential $V_{\operatorname{erf}}$ is
-\begin{equation}f(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation} from which it can be seen that $f(r)$ is approximately 0 for $r>r_{max}$ where $r_{max}$ is some maximum value of $r$.
+\begin{equation}f(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation} 
+from which it can be seen that $f(r)$ is approximately 0 for $r>r_{max}$ 
+where $r_{max}$ is some maximum value of $r$.
 \begin{align}
@@
 
-The Mayer function for the Error Function potential $V_{\operatorname{erf}}$, and its Fourier Transform are
+The Mayer function for the Error Function potential $V_{\operatorname{erf}}$, 
+and its Fourier Transform are
 \begin{equation}f_{\operatorname{erf}}(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation}

### 2020-11-30 papers/thesis-kirstie/thesis.tex
updated figure
edits

@@
     \includegraphics[height=7cm]{plot_HardSphere_Potential.pdf}
-    \caption{A hard-sphere Potential is zero for $r>2R$, and infinity 
-    for $r\leq{2R}$ where R is the radius of a hard sphere (which is 
+    \caption{A hard-sphere potential is zero for $r>2R$, and infinity 
+    for $r<{2R}$ where R is the radius of a hard sphere (which is 
     approximately 1.225/2 in this example).}
@@
     \includegraphics[height=7cm]{LJ_Potential.pdf}
-    \caption{Lennard-Jones Potential for $\sigma=1$, $\epsilon=1$}
+    \caption{Lennard-Jones potential for $\sigma=1$, $\epsilon=1$}
     \label{fig:LJ_potential}
@@
     \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
-    \caption{Lennard-Jones Potential broken into two parts for $\sigma=1$, $\epsilon=1$}
+    \caption{Lennard-Jones potential broken into two parts for $\sigma=1$, $\epsilon=1$}
     \label{fig:LJ_potential_2parts}
@@
     \includegraphics[height=7cm]{WCA_Potential.pdf}
-    \caption{WCA Potential for $\sigma=1$, $\epsilon=1$}
+    \caption{WCA potential for $\sigma=1$, $\epsilon=1$}
     \label{fig:WCA_potential}
@@
     \centering
-    \includegraphics[height=6cm]{P-T_Diagram.png}
+    \includegraphics[height=6cm]{P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Densities at the solid-liquid transition are identified using 
-    Maxwell's Double Tangent Construction}
+    \caption{Liquid and solid densities at the solid-liquid phase 
+    transition are identified using Maxwell's Double Tangent Construction}
     \label{fig:MaxwellDT}
@@
 the same as the chemical potential of the solid, as is the pressure.
-\begin{align}{g_L=g_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}}\end{align} 
+\begin{align}{\mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}}\end{align} 
 The pressure at the phase transition can be obtained from the point of 
@@
 liquid and the solid at a given temperature. 
-The slope gives -P and corresponding y-intercept gives $g$. The 
+The slope gives the negative of the pressure, and the corresponding
+y-intercept gives the chemical potential. The 
 curves in Figure~\ref{fig:GibbsvsP} consist of a collection of data 
-points ($g$,P) for the liquid and the solid respectively.
+points (P,~$\mu$) for the liquid and the solid respectively.
 
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
-    \caption{Pressure is found at point of intersection which occurs when $g_L=g_S$}
+    \caption{Pressure is found at point of intersection which occurs 
+    when $\mu_L=\mu_S$.}
     \label{fig:GibbsvsP}
@@
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
-    \caption{The liquid and solid densities at the point of transistion are 
+    \caption{The liquid and solid densities at the phase transistion are 
     found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
@@
 when the natural variables of the free energy are held fixed~\cite{schroeder}. 
-And so over time teh free energy settles to a minimum. 
+And so over time the free energy settles to a minimum. 
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
@@
 where the translational kinetic energy of the N atoms is given by 
-\begin{equation}{\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{equation}
+\begin{equation}{\text{KE}=\sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)}\end{equation}
 and the potential energy of the total interaction of each atom with the other 
@@
     \includegraphics[height=7cm]{figs/diagrammic.pdf}
-    \caption{Two-pair Interactions $f_{ij}f_{kl}$ for N=4 atoms. Interactions 
-    between individual pairs of atoms are shown in green. Interactions 
-    between pairs of atoms with an atom shared in common are shown in red. 
-    At low densities only the interactions shown in green are considered, 
-    since atoms are not likely to be clustered.}
+    \caption{Two-pair interactions $f_{ij}f_{kl}$ are shown for N=4 atoms. 
+    Interactions between individual pairs of atoms are shown in green. 
+    Interactions between pairs of atoms with an atom shared in common 
+    are shown in red. At low densities only the interactions shown in 
+    green are considered, since atoms are not likely to be clustered.}
     \label{fig:diagrammic}
@@
     \caption{No energy is needed to place a sphere at a position where 
-    there is no overlap such as shown by the black dot at the center of 
-    the dotted region near the top. But it is not possible to place a 
-    sphere at a position where an overlap occurs.}
+    there is no overlap, such as that shown by the black dot near the 
+    top. But it is not possible to place a sphere at a position where 
+    an overlap occurs.}
     \label{fig:p_overlap}
@@
 hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
-occupied by the spheres divided by the total volumet avaliable, 
+occupied by the spheres divided by the total volume avaliable, 
 is the probability that a sphere cannot be placed at some random
@@
     \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
-    \caption{The Primitive Crystal Cell is shown by the shaded parallelepiped}
+    \caption{A primitive crystal cell is shown by the shaded parallelepiped.}
     \label{fig:primitivecell}
@@
     \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
-    \caption{Four examples of possible positions an atom (blue sphere) 
-    might have within a cell that has one lattice site (black dot) at 
-    the center are shown on the left. A large number of such 
+    \caption{On the left are shown four examples of possible positions 
+    that an atom (blue sphere) might have within a cell that has one 
+    lattice site (black dot) at the center. A large number of such 
     possibilities would give rise to a Gaussian distribution, 
@@
     \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{Four examples of possible positions an atom (blue sphere) might 
-    have within a fixed volume containing eight cells are shown on the left. 
-    Each cell has one lattice site at the center represented by a black dot. 
-    A large number of such possibilities would give rise to the Gaussian 
-    distributions shown on the right.}
+    \caption{On the left are shown four examples of possible positions 
+    that an atom (blue sphere) might have within a fixed volume 
+    containing eight cells. Each cell has one lattice site at the center 
+    represented by a black dot. A large number of such possibilities 
+    would give rise to the Gaussian distributions shown on the right.}
     \label{fig:Ensemble_Smallcells}
@@
     show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
-    give rise to the same Gaussian distribution statistically.}
+    give rise to the same Gaussian distribution statistically, 
+    and the same number density.}
     \label{fig:SameStatPic}
@@
     \includegraphics[height=8cm]{VaryWidthandVacancies.png}
-    \caption{Several Gaussian Distribution Number Density Profiles all 
+    \caption{Shows an example set of Gaussian distribution number density 
+    profiles $n(vec r)$ all with the same number density.
     with the same number density of one atom over the fixed volume.}
@@
     \includegraphics[height=8cm]{plot_alpha.pdf}
-    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT/{\epsilon}$ 
+    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
     for $\sigma=1$ and $\epsilon=1$.}
@@
     \includegraphics[height=8cm]{weight_functions.pdf}
-    \caption{The weight functions $w_0$, $w_1$, $w_2$, $w_3$ vary along the 
-    z-axis for a reduced temperature of 2.}
+    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ 
+    vary along the z-axis for a reduced temperature of $K_BT=2$.}
     \label{fig:weight_functions}
@@
 centered at lattice sites specified by position vectors $\vec R$. 
-The weight functions $w_i(|\vec r - \vec r'|)$ are centered at $\vec r'$,
+The weight functions $w_i(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
 as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}. 
 But because the weight function is a function of the magnitude of the 
-difference between $\vec r$ and $\vec r'$, it can be thought of as 
+difference between $\vec r$ and $\vec r~'$, it can be thought of as 
 centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_thinkas}. 
@@
    \includegraphics[height=5cm]{InclusionRadius.png}
-   \caption{The Gaussians at each lattice point are of width gw. 
-   After a radius of few gw's the Gaussians go to zero and can be neglected.}
+   \caption{Gaussians are centered at each lattice point. After a radius 
+   of a few Gaussian widths about each lattice point, the Gaussians go to 
+   zero and can be neglected.}
    \label{fig:InclusionRadius}
@@
 by randomly generating vectors
-$\vec r' -\vec R$ according to a Gaussian distribution. 
-Only lengths for $\vec r'- \vec R$ that are within an inclusion radius
+$\vec r~' -\vec R$ according to a Gaussian distribution. 
+Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
 about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
@@
 Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
-\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r' - \vec R_j)w_i(|\vec{r}-\vec{r}'|)} {d}\vec{r}'}\end{equation} 
+\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_i(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
 and is implemented computationally in the form
-\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r' -\vec R_j)w_i(|\vec{r}-\vec{r}'|)dV}}\end{equation} 
+\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_i(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
 which can be written as
-\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}'|)}\end{equation} 
-NumPoints is the number of vectors $\vec r' - \vec R_j$ randomly 
-generated about the jth lattice site, and $n(\vec r' - \vec R_j)$
+\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}~'|)}\end{equation} 
+NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
+generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
 is equal to a probability density which when multiplied by dV gives 
-the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}'|)$ 
+the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}~'|)$ 
 will occur. Using the relation
 \begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
-$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}'|)}{\text{NumPoints}}}\end{equation}
+$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}~'|)$:
+\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
 This process is repeated to get the contributions to the weighted density 
@@
     $\vec R$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec{r}'$ is shown in purple.} 
+    view of $w_2$ centered at $\vec r~'$ is shown in purple.} 
   \label{fig:GaussandW2_actual}
@@
     the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
-    and fixed in place as $\vec r' - \vec R$
+    and fixed in place as $\vec r~' - \vec R$
     varies randomly with a probability according to a Gaussian distribution.} 
@@
   at varying proportions. White lines show the boundaries of the coexistence 
-  region generated from Monte-Carlo of the WCA fluid.}
+  region generated from Monte-Carlo simulations of the WCA fluid.}
   \label{fig:Phase_Diagram_of_T_vs_n}
@@
   \caption{Phase Diagram of the reduced pressure verses the reduced 
-  temperature. The blue region is solid, and the red region is liquid.
-  The white line marks the solid-liquid boundary as predicted by 
-  Monte-Carlo simulations.}
+  temperature. The blue region is solid, the red region is liquid,
+  and the black line shows the phase boundary where the solid and liquid
+  phases coexist. The white line marks the solid-liquid phase boundary  
+  as predicted by Monte-Carlo simulations.}
   \label{fig:Phase_Diagram_P_vs_T}

### 2020-11-30 papers/thesis-kirstie/thesis.tex
more edits

@@
 
-\section{The Theory of Freezing and Maxwell's Double Tangent Construction}
+\section{Freezing and Maxwell's Double Tangent Construction}
 
@@
 \begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
+%The reason the minimization works is because DFT satisfies the conditions of the 
+%variational principle, the same mathematical principle used to determine 
+%that the shortest distance between two points is a straight line by 
+%varying the path between the two points until the distance is minimized.
 
@@
 
-%The reason the minimization works is because DFT satisfies the conditions of the 
-%variational principle, the same mathematical principle used to determine 
-%that the shortest distance between two points is a straight line by 
-%varying the path between the two points until the distance is minimized.
 The minimization can be illustrated for a system near 
@@
 $\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
-changes in the free energy will be negative for spontaneous processes 
-when the natural variables of the free energy are held fixed~\cite{schroeder}. 
-And so over time the free energy settles to a minimum. 
+changes in the free energy 
+will be negative for spontaneous processes 
+when the natural variables of the free energy are held fixed. 
+For our case, with a fixed temperature, volume, and number of particles
+we get
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
 %\begin{equation}
-  %df_{N-System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
+  %df_{System}=d\left(\frac{U}{N}\right)_{System}-Td\left(\frac{S}{N}\right)_{System}+\left(\frac{S}{N}\right)dT_{System}
 %\end{equation}
@@
 %\begin{equation}
-  %f_{N-System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
+  %f_{System}=-d\left(\frac{U}{N}\right)_{Surroundings}-Td\left(\frac{S}{N}\right)_{System}
 %\end{equation}
@@
 %results in
-%\begin{equation}
-  %df_{N-System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
-%\end{equation}
+\begin{equation}
+  df_{System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+\end{equation}
 %which, for fixed N, is the same as
@@
 %~\cite{schroeder}. 
-A proof of minimization in general
-can be found in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
+and so the free energy tends toward a minimum. 
+Details of this procedure be found in Thermal Physics by Schroeder~\cite{schroeder}, 
+and a proof of minimization in general can be found in 
+Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.
 
@@
 SFMT is an extension of Fundamental Measure Theory (FMT)~\cite{rosenfeld1989}
-that was developed for the hard-sphere fluid. FMT can be derived from
+that was developed for the hard-sphere fluid which can be derived from
 Scaled Particle Theory (SPT)~\cite{santos2012phi3}.  
@@
 Formula~\cite{Hansen}.
-\begin{equation}\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}\end{equation}
+\begin{align} \label{mu-der-int}
+    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
+\end{align}
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
@@
 there will be a change in the potential energy of $V_{diff}$ if one sphere
-is added, but $V_{diff}$ will be zero if no sphere is added. 
+is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
 %at each point in the fluid, there is a probability that 
@@
 %without overlapping with other spheres, or it does not, as shown in 
-Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, then becomes
+Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, 
+then becomes~\cite{Hansen}
 \begin{align} 
@@
   \end{figure}
-In this case Equation~\ref{widoms-insertion-formula} becomes
+In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
 \begin{align} \label{muexcess}
@@
 \begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
-Forming a more general result by requiring continuity of the functions, 
-along with their first and second derivatives, for the reversible work
-written for fluids with a cavity radius equal to, larger than, or very 
-much larger than the hard sphere radius R, Scaled Particle Theory arrives 
-at an expression for the excess Helmholtz free energy of 
-a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
-For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
+Scaled Particle Theory forms a more general result for the reversible work 
+as a function of the radius of the cavity by considering additional 
+cases where the cavity is larger, and much larger than the radius of 
+a hard sphere, and requiring continuity of the functions for the 
+different cases and their derivatives (first and second). 
+By setting the radius of the cavity to the hard sphere radius, the 
+more general work equation becomes equal to the chemical potential. 
+Details of this procedure can be found 
+in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.  
+An expression for the excess Helmholtz free energy in terms of the
+packing fraction $\eta$ can then be obtained through 
+Equation~\ref{mu-der-int}. 
+%
+%By requiring continuity of the functions valid for different cavity 
+%sizes, along with their first and second derivatives, an equation for 
+%the work is obtained which can then be used to form 
+%set equal to the excess chemical 
+%potential by setting the cavity radius to the radius of hard sphere. 
+%
+%Forming a more general result by requiring continuity of the functions, 
+%along with their first and second derivatives, for the reversible work
+%written for fluids with a cavity radius equal to, larger than, or very 
+%much larger than the hard sphere radius R, 
+%
+%Scaled Particle Theory arrives at 
+%an expression for the excess Helmholtz free energy of 
+%a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
+%For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
 \begin{align}\label{Fexcess-SPT}
@@
 
-The FMT equations developed so far have been for a hard-sphere fluid.
+The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
 All of the weight functions, the relations between them, and the equation 

### 2020-12-01 papers/thesis-kirstie/thesis.tex
edits

@@
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
-    \caption{The liquid and solid densities at the phase transistion are 
-    found from a plot of pressure vs inverse denisty for both the liquid and the solid.}
+    \caption{The liquid and solid densities at the phase transition are 
+    found from a plot of pressure vs inverse density for both the liquid and the solid.}
     \label{fig:Pvsinvn}
@@
 
-Classical Density Fucntional Theory will be used to find the 
+Classical Density Functional Theory will be used to find the 
 Helmholtz free energies of the liquid and the solid for various 
@@
 \begin{displaymath}{\mbox{where }  n(\vec{r})=n_{eq}(\vec r)  \mbox{ at equilibrium}}\end{displaymath}
-%The reason the minimization works is because DFT satisfies the conditions of the 
-%variational principle, the same mathematical principle used to determine 
+The reason the minimization works is because DFT satisfies the conditions 
+of the variational principle. 
+%the same mathematical principle used to determine 
 %that the shortest distance between two points is a straight line by 
@@
 It is an extension of the work done by Widom who showed that the excess 
-chemical potential of a homogenesous fluid corresponds to the reversible 
+chemical potential of a homogeneous fluid corresponds to the reversible 
 work of inserting a sphere of radius R into a fluid consisting of hard 
@@
 hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
-occupied by the spheres divided by the total volume avaliable, 
+occupied by the spheres divided by the total volume available, 
 is the probability that a sphere cannot be placed at some random
@@
 
-The weight functions can be related to the mayer function 
+The weight functions can be related to the Mayer function 
 \begin{align}
@@
 Evaluating Equation with $V(r)$ given by Equation,  
-the mayer function for a hard-sphere fluid is found to be
+the Mayer function for a hard-sphere fluid is found to be
 \begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
@@
 
-Taking the derivative of the mayer function in Eq~\ref{f(r)step} provides 
+Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
 another relation that will be important to keep in tact when moving to 
@@
 All of the weight functions, the relations between them, and the equation 
-for the derivative of the mayer function, can be expressed in terms of 
+for the derivative of the Mayer function, can be expressed in terms of 
 $w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
@@
 with soft potentials by keeping the relationships between the weight 
-functions, and the equation for the derivative of the mayer function 
+functions, and the equation for the derivative of the Mayer function 
 the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
@@
 \begin{equation}B_2=-\frac{1}{2}\int_{Volume}f(\vec{r})d\vec r\end{equation}
-Using the mayer function for the error function potential given by 
+Using the Mayer function for the error function potential given by 
 Equation~\ref{eq:VerfMayerfunction} we found (see Appendix) the second 
@@
 %As can be seen in the figure, the region of coexistence in our simulations 
-%falls within the region of coexistance predicted by Ahmed and Sadus at almost 
+%falls within the region of coexistence predicted by Ahmed and Sadus at almost 
 %all temperatures from approximately kT=0.5 to 28. At the low temperature of kT=0.5,
-%the width of the coexistance regions are nearly the same, while moving to 
-%higher temperatures,the width of our coexistance region ranges from about 
+%the width of the coexistence regions are nearly the same, while moving to 
+%higher temperatures,the width of our coexistence region ranges from about 
 %one-half to one-forth the size. Although, our coexistence region is only 
@@
   %%\includegraphics[height=10cm]{figs/Phase_Diagram_of_T_vs_n_AS}
-  %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fuild. 
+  %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
   %%White lines show the liquid and solid transition densities at each temperature from 
@@
 %Figure~\ref{fig:Phase_Diagram_P_vs_T_AS} shows our results for a larger 
-%range of reduced temperatures from kT=0.5 to 38, compated with the results
+%range of reduced temperatures from kT=0.5 to 38, compared with the results
 %from Ahmed and Sadus's molecular simulations. Here again, our transition 
@@
 %\ref{fig:Phase_Diagram_P_vs_T_AS}, the plotted lines display kinks when 
-%the WCA fluid transistions into a solid. These downward bends align to 
+%the WCA fluid transitions into a solid. These downward bends align to 
 %form the curved interface separating the blue and red areas in 
@@
 \begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}
-The mayer function for the Error Function potential $V_{\operatorname{erf}}$ is
+The Mayer function for the Error Function potential $V_{\operatorname{erf}}$ is
 \begin{equation}f(r)=\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]\end{equation} 
@@
 \begin{equation}\frac{PV}{k_BT}=N + \beta\frac{\partial{F}_{excess}}{\partial{N}}{N} - \frac{F_{excess}}{k_BT}\end{equation}
-Plugging in the expression for $F_{excess}$ in terms of the mayer function 
+Plugging in the expression for $F_{excess}$ in terms of the Mayer function 
 for low density fluids

### 2020-12-01 papers/thesis-kirstie/thesis.tex
fixed spelling

@@
 \begin{equation}{a=\frac{\Xi}{\sqrt{2}}}\end{equation}
-The fourier transform of the tensor weight matrix elements are given by
+The Fourier transform of the tensor weight matrix elements are given by
 \begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
@@
 Instead of constraining vector $\vec{k}$ to be in the z direction, 
-owever, a rotated coordinate system $(x',y',z')$ is used which is 
+however, a rotated coordinate system $(x',y',z')$ is used which is 
 rotated from the (x,y,z) coordinate system in such a way that $\hat{z}'$ 
@@
 \begin{equation}{n_{m2ij}(\vec{r})=\int_{allspace}{n(\vec{r}')w_{m2ij}(\vec{r}-\vec{r}'){}d{\vec{r}'}}}\end{equation}
-Using the definintion of the delta function
+Using the definition of the delta function
 \begin{equation}{\delta}((\vec{r}-\vec{r}')-\vec{r}'')={ \frac{1}{\left(2\pi\right)^3}\int e^{i\vec k\cdot ((\vec r-\vec r')-\vec{r}'')}d\vec{k}}\end{equation} 
@@
 {n_i})\right]\end{equation}
-For a Grand Canonical Ensemble the probablity of a state given by
+For a Grand Canonical Ensemble the probability of a state given by
 \begin{equation}P_{n,s(n)}=\frac{e^{-\beta(E_{s(n)}-\mu{n})}}{\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}}\end{equation}
@@
 \section{Derivation of $F = -k_BTln(Z)$ for Continuous Systems}
-The canonical equilibrium phase-space probablity density is given by
+The canonical equilibrium phase-space probability density is given by
 \begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}

### 2020-12-01 papers/thesis-kirstie/thesis.tex
edit
figs

@@
     \centering
-    \includegraphics[height=6cm]{T-n_Diagram.png}
+    \includegraphics[height=6cm]{T-n_Diagram.pdf}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
@@
 Scaled Particle Theory (SPT)~\cite{santos2012phi3}. 
-Scaled Particle Theory seeks a general expression for the excess chemical 
+Scaled Particle Theory~\cite{ReissSPT} 
+seeks a general expression for the excess chemical 
 potential of a homogeneous hard-sphere fluid which can then be integrated 
@@
 \begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-
+where 
+\begin{align}
+    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
+\end{align}
+and
+\begin{align}
+    Z=\sum_s{e^{-\beta V_s}}
+\end{align}
 The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
@@
 is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
+In this case, $P_s=\frac{e^{-\beta V_s}}{Z}$.
 %at each point in the fluid, there is a probability that 
@@
 point in the homogeneous fluid of number density, $n$, without
-overlapping with another sphere. Combining  Equations~\ref{workrev} 
+overlapping with another sphere. $P_s$ would then have values of
+$nV_{Sphere}$ and $1-nV_{Sphere}$.
+Combining  Equations~\ref{workrev} 
 and \ref{muexcess} gives
@@
 
-\section{Second Virial Coefficient $B_{2}$ for the Error Function Potential $V_{\operatorname{erf}}$}
+\section{Second Virial Coefficient $B_{2}$ for the Error Function Potential}
 \begin{equation}B_2=-\frac{1}{2}\int_{allspace}f(\vec{r})d\vec r\end{equation}

### 2020-12-01 papers/thesis-kirstie/thesis.tex
use generated figurs
remove figure placement tweaks

@@
 
-\begin{figure}[h!]
+\begin{figure}
     \centering
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{P-T_Diagram.pdf}
+\begin{figure}
+    \centering
+    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=6cm]{T-n_Diagram.pdf}
+\begin{figure}
+    \centering
+    \includegraphics[height=6cm]{figs/T-n_Diagram.pdf}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
@@
 
-  \begin{figure}[h!]
+  \begin{figure}
     \centering
@@
 in Figure~\ref{fig:p_overlap}. 
-\begin{figure}[h!]
+\begin{figure}
     \centering
@@
 
- \begin{figure}[h!]
+ \begin{figure}
     \centering
@@
 for all these number density profiles $n(\vec{r})$. 
-\begin{figure}[h!]
+\begin{figure}
     \centering
@@
   \end{figure}  
- \begin{figure}[h!]
+ \begin{figure}
     \centering
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[height=8cm]{plot_alpha.pdf}
+  \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
+  for $\sigma=1$ and $\epsilon=1$.}
+  \label{fig:alphaXivsT}
+\end{figure}
 The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
@@
 \end{equation}  
-
 The parameter $\alpha(T)$ is derived by setting 
@@
 when $\sigma=1$ and $\epsilon=1$.
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{plot_alpha.pdf}
-    \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
-    for $\sigma=1$ and $\epsilon=1$.}
-    \label{fig:alphaXivsT}
+
+  \begin{figure}
+    \centering
+    \includegraphics[height=8cm]{xi_fromB2.pdf}
+    \caption{Shows the temperature dependence of $\Xi$ generated from 
+    matching the second virial coefficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
+    at each temperature.}
+    \label{fig:xi_fromB2vsT}
   \end{figure}
@@
 Figure~\ref{fig:xi_fromB2vsT}.
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{xi_fromB2.pdf}
-    \caption{Shows the temperature dependence of $\Xi$ generated from 
-    matching the second virial coefficients for $V_{\operatorname{erf}}(r)$ and $V_{WCA}(r)$ 
-    at each temperature.}
-    \label{fig:xi_fromB2vsT}
-  \end{figure}
-
+
+\begin{figure}
+  \centering
+  \includegraphics[height=8cm]{weight_functions.pdf}
+  \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ 
+  vary along the z-axis for a reduced temperature of $K_BT=2$.}
+  \label{fig:weight_functions}
+\end{figure}
 Figure \ref{fig:weight_functions} shows a one dimensional view of the 
@@
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[height=8cm]{weight_functions.pdf}
-    \caption{Shows how the weight functions $w_0$, $w_1$, $w_2$, $w_3$ 
-    vary along the z-axis for a reduced temperature of $K_BT=2$.}
-    \label{fig:weight_functions}
-  \end{figure}
-
 \section{Computing the Liquid and Crystal Helmholtz free energies}
@@
 with a uniform number density profile, n that does not depend on $\vec r$. 
+\begin{figure}
+  \centering
+  (a) \includegraphics[height=7cm]{figs/weightandGauss1.pdf}\\
+  (b) \includegraphics[height=7cm]{figs/weightandGauss2.pdf}
+  \caption{(a) A Gaussian distribution about a lattice site located at 
+  $\vec R$ is shown in blue, and a cross-sectional 
+  view of $w_2$ centered at $\vec r~'$ is shown in purple.
+  (b) An alternative way of viewing the top figure, where 
+  the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
+  and fixed in place as $\vec r~' - \vec R$
+  varies randomly with a probability according to a Gaussian distribution.}
+\label{fig:GaussandW2_actual}
+\end{figure}
+
+% \begin{figure}
+%   \centering
+%   \includegraphics[height=7cm]{figs/weightandGauss2.pdf}
+%   \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 
+%   the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
+%   and fixed in place as $\vec r~' - \vec R$
+%   varies randomly with a probability according to a Gaussian distribution.} 
+% \label{fig:GaussandW2_thinkas}
+% \end{figure}
 
@@
 The weight functions $w_i(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
-as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}. 
+as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}a. 
 But because the weight function is a function of the magnitude of the 
 difference between $\vec r$ and $\vec r~'$, it can be thought of as 
-centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_thinkas}. 
+centered at $\vec r$, as shown in Figure~\ref{fig:GaussandW2_actual}b.
 Keeping the weight function centered at $\vec r$, 
 Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
-\begin{figure}[h!]
+\begin{figure}
    \centering
@@
 while the other is erroneously lower than it should be.
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{figs/weightandGauss1.pdf}
-    \caption{A Gaussian distribution about a lattice site located at 
-    $\vec R$ is shown in blue, and a cross-sectional 
-    view of $w_2$ centered at $\vec r~'$ is shown in purple.} 
-  \label{fig:GaussandW2_actual}
-  \end{figure}
-
- \begin{figure}[h!]
-    \centering
-    \includegraphics[height=7cm]{figs/weightandGauss2.pdf}
-    \caption{An alternative way of viewing Figure~\ref{fig:GaussandW2_actual} where 
-    the weight function, shown by the purple dashed lines, is centered at $\vec{r}$,  
-    and fixed in place as $\vec r~' - \vec R$
-    varies randomly with a probability according to a Gaussian distribution.} 
-  \label{fig:GaussandW2_thinkas}
-  \end{figure}
 

### 2020-12-02 papers/thesis-kirstie/thesis.tex
skeleton for beamer presentation for Kirstie
thesis-kirstie/thesis.bib

@@
   interaction of two neutral atoms. Our cDFT functional is constructed 
-  using Soft Fundamental Theory (SFMT) with Tarazona's tensor expression 
-  for $\Phi_3$, and Schmidt's error function model in which we have 
-  incorporated our temperature dependent parameters $\alpha$ and $\Xi$. 
-  We find $\Xi$ by making the second virial coefficients of the error 
-  function potential and the WCA potential equal. Our functional was 
+  using Soft Fundamental Theory (SFMT) Schmidt's error function model 
+  in which we have incorporated temperature dependent parameters to make
+  %$\alpha$ and $\Xi$. We find $\Xi$ by making the second virial 
+  %coefficients of the error function potential and the WCA potential equal. 
+  the second virial coefficient correct. Our functional was 
   successful in predicting the freezing of the WCA fluid with a fairly 
-  good match to Monte-Carlo simulations. 
-  This result suggests that SFMT may provide a strong foundation for 
-  future density functionals that will model real fluids.
+  good match to Monte-Carlo simulations. This result suggests that SFMT 
+  may provide a strong foundation for future density functionals that 
+  will model real fluids.
 }
@@
 
-SFMT is an extension of Fundamental Measure Theory (FMT)~\cite{rosenfeld1989}
-that was developed for the hard-sphere fluid which can be derived from
+SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
+for soft-spheres, and is an extension of Fundamental Measure Theory (FMT). 
+FMT was developed 
+by Rosenfeld~\cite{rosenfeld1989, rosenfeld1994,fmtfromsptandsignerror}, 
+for the hard-sphere fluid,
+%, and which can be 
+and derived from 
+%by expanding 
 Scaled Particle Theory (SPT)~\cite{santos2012phi3}. 
+%to inhomogeneous, or non-uniform hard-sphere fluids.
 Scaled Particle Theory~\cite{ReissSPT} 

### 2020-12-03 papers/thesis-kirstie/thesis.tex
a few more edits

--- 2020-12-01 papers/thesis-kirstie/thesis.tex
@@
   interaction of two neutral atoms. Our cDFT functional is constructed 
-  using Soft Fundamental Theory (SFMT) with Tarazona's tensor expression 
-  for $\Phi_3$, and Schmidt's error function model in which we have 
-  incorporated our temperature dependent parameters $\alpha$ and $\Xi$. 
-  We find $\Xi$ by making the second virial coefficients of the error 
-  function potential and the WCA potential equal. Our functional was 
+  using Soft Fundamental Theory (SFMT) with Schmidt's error function 
+  model in which we have incorporated temperature dependent parameters 
+  to make the second virial coefficient correct. Our functional was 
   successful in predicting the freezing of the WCA fluid with a fairly 
-  good match to Monte-Carlo simulations. 
-  This result suggests that SFMT may provide a strong foundation for 
-  future density functionals that will model real fluids.
+  good match to Monte-Carlo simulations. This result suggests that SFMT 
+  may provide a strong foundation for future density functionals that 
+  will model real fluids.
 }
@@
 The goal of this project is to generate temperature-density 
-and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson (WCA) 
-model fluid using classical Density Functional Theory (cDFT). 
+and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
+(WCA) model fluid using classical Density Functional Theory (cDFT). 
 To do this, the Helmholtz free energies of the fluid
@@
     \caption{A hard-sphere potential is zero for $r>2R$, and infinity 
-    for $r<{2R}$ where R is the radius of a hard sphere (which is 
-    approximately 1.225/2 in this example).}
+    for $r<{2R}$ where R is the radius of a hard sphere 
+    (R$=\frac{r}{2}\approx$1.22/2=0.61 in this example).}
     \label{fig:HardSphere_potential}
@@
 The repulsive portion of the Lennard-Jones potential forms the potential for a 
-Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship}. 
-The WCA potential is described by the equation: 
+Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship},
+shown in Figure \ref{fig:WCA_potential}. 
+The WCA potential is described by the equation 
 \begin{align} \label{eq:VWCA}
@@
 \end{align} 
-As seen in Figure \ref{fig:WCA_potential} the WCA potential decreases 
-from infinity as the distance between the centers of the atoms r increases, 
-and becomes, and remains, essentially zero when 
-r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ 
-where R is the radius of the sphere.
-
+The WCA potential decreases from infinity as the distance between the 
+centers of the atoms r increases, and becomes, and remains, essentially 
+zero when r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the 
+radius of the sphere. 
 As with the Lenard-Jones potential, the repulsive WCA potential increases 
@@
 Face Centered Cubic (FCC) crystal cells~\cite{Hansen}, as do hard spheres. 
-This structure is what one would see by filling a cube with marbles. It 
+This structure is what one would see by filling a box with marbles. It 
 results in the highest density achievable for hard, and nearly-hard, spheres 
 that are purely repulsive.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
+    \caption{Conceptual Pressure-Temperature Phase Diagram}
+    \label{fig:P-T_Diagram}
+\end{figure}
 
@@
 Points along the line that separates the solid and the liquid phases 
-indicate a phase transition, and show that the solid and the liquid 
-both have the same temperature and the same pressure at a phase transition.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
-    \caption{Conceptual Pressure-Temperature Phase Diagram}
-    \label{fig:P-T_Diagram}
-  \end{figure}
+indicate a phase transition where the solid and liquid phases coexist, 
+and show that both the solid and the liquid have the same temperature 
+and the same pressure at a phase transition.
+
 The pressure can be related to the Helmholtz free energy F, which is minimized 
-for a system at fixed temperature, volume and number of particles as the 
+for a system at fixed temperature, volume and number of particles, as the 
 entropy of the system and its surroundings is maximized. The negative of 
-the partial derivative of the Helmholtz free energy with respect to volume V 
-at a fixed temperature T and number of particles N equals the pressure. 
+the partial derivative of the Helmholtz free energy with respect to volume V, 
+and at a fixed temperature T and number of particles N, equals the pressure. 
 \begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
@@
 density at fixed temperature.
+
+ \begin{figure}
+    \centering
+     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
+    \caption{Liquid and solid densities at the solid-liquid phase 
+    transition are identified using Maxwell's Double Tangent Construction}
+    \label{fig:MaxwellDT}
+  \end{figure}
 
@@
 freezing can be found from identifying the points at which the 
-tangents to each curve have the same slope. This process is called 
-Maxwell's Double Tangent Construction.
-
-\begin{figure}
-    \centering
-     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Liquid and solid densities at the solid-liquid phase 
-    transition are identified using Maxwell's Double Tangent Construction}
-    \label{fig:MaxwellDT}
-  \end{figure}
-
-To facilitate finding the pressure at the phase transition, 
+tangents to each curve have the same slope, as shown by the black line. 
+This process is called Maxwell's Double Tangent Construction.
+
+To facilitate finding the pressure at the phase transition computationally, 
 it is helpful to consider the Gibbs free energy given by G=F+PV. 
@@
 per atom  $g$ (which is also the chemical potential)  
-\begin{align}{\frac{F}{N}=-P\frac{V}{N}+\frac{G}{N}}\end{align} 
-\begin{align}{f=-P\frac{1}{n}+g}\end{align}
+\begin{align}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{align}
+\begin{align}f=-P\frac{1}{n}+g\end{align}
 This equation has the form of a line
-\begin{align}{y=mx+y_o}\end{align}
+\begin{equation}
+    y=mx+y_o
+\end{equation}
 and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
 pressure given by the negative of the slope of the line, and the Gibb's 
-free energy per atom given by the y-intercept. 
+free energy per atom, or chemical potential given by the y-intercept. 
 At the point of phase transition the chemical potential of the liquid is 
 the same as the chemical potential of the solid, as is the pressure.
-\begin{align}{\mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}}\end{align} 
-The pressure at the phase transition can be obtained from the point of 
-intersection that occurs when the Gibbs free energy per atom, or chemical
-potential, is plotted against the pressure for both the liquid and the 
-solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
-These curves are constructed from data collected from lines tangent 
-along the curves on the the Helmholtz free energy per atom verses inverse 
-density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
-liquid and the solid at a given temperature. 
-The slope gives the negative of the pressure, and the corresponding
-y-intercept gives the chemical potential. The 
-curves in Figure~\ref{fig:GibbsvsP} consist of a collection of data 
-points (P,~$\mu$) for the liquid and the solid respectively.
+\begin{equation}
+    \mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}
+\end{equation}
 
@@
 
-Once the pressure is known for the transition point, the density of the 
-liquid at the transition point can be found from a plot of the pressure 
-versus the inverse of the density for the liquid, as shown in Figure 
-\ref{fig:Pvsinvn}. The density of the solid at the transition point can 
-be found the same way. A Temperature-Density (T-n) phase diagram, such as 
-that in Figure \ref{fig:T-n_Diagram}, can be generated by repeating this 
-procedure for various temperatures.
-
 \begin{figure}
@@
   \end{figure}
+
+The pressure at the phase transition can be obtained from the point of 
+intersection that occurs when the chemical potential 
+is plotted against the pressure for both the liquid and the 
+solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
+This plot is constructed from data collected from lines tangent 
+along the curves on the Helmholtz free energy per atom verses inverse 
+density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
+liquid and the solid at a given temperature. 
+The slope of the tangent line at a given point gives the negative of the 
+pressure at that point, and the corresponding y-intercept of the tangent 
+line gives the chemical potential at that point. Collecting this data 
+(P,~$\mu$) at each point for the liquid and the solid respectively, 
+produces the curves in Figure~\ref{fig:GibbsvsP}.
 
@@
   \end{figure}
+
+Once the pressure is known at the point of phase transition, for a given 
+temperature and number density, the liquid and solid number densities at 
+the phase transition can be found from plots of the pressure versus the 
+inverse of the number density for both the solid and the liquid, 
+as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
+find the liquid and solid number densities at the phase transition 
+for various temperatures, a Temperature-Density (T-n) phase diagram 
+can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
+but without the gray hump that is characteristic of fluids that have 
+both attractive and repulsive interatomic potentials.
 
@@
     Interactions between pairs of atoms with an atom shared in common 
-    are shown in red. At low densities only the interactions shown in 
-    green are considered, since atoms are not likely to be clustered.}
+    are shown in red. At low densities only the green interactions
+    are considered, since atoms are not likely to be clustered.}
     \label{fig:diagrammic}
@@
 \section{Soft Fundamental Measure Theory (SFMT)}
-In order to apply classical Density Functional Theory to a system that has
-a fixed temperature, volume, and number of particles, a functional for 
-the Helmholtz free energy must first be formed. An expression for the ideal 
-portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
-but forming an expression for the excess Helmholtz free
-energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
-being modeled can be challenging. 
 Soft Fundamental Measure Theory (SFMT) is a method for forming an 
-expression for the excess Helmholtz free energy of a soft-sphere fluid
+expression for the excess Helmholtz free energy of a soft-sphere fluid.
+In the next chapter, 
+we will adapt an SFMT functional to fit the soft-sphere WCA fluid.
+Recall that in order to apply classical Density Functional Theory 
+to a system that has a fixed temperature, volume, and number of particles, 
+a functional for the Helmholtz free energy must first be formed. 
+The ideal portion of the Helmholtz free energy is given by 
+Equation~\ref{eq:Fideal-n(r)}, but forming an expression for the excess 
+Helmholtz free energy from Equation~\ref{eq:Fexcess-simplified} for the 
+particular fluid being modeled can be challenging. 
+SFMT is a method for forming an 
+expression for the excess Helmholtz free energy
 that does not depend on explicitly taking into account the interaction 
 that each sphere with every other sphere in the fluid (as required by 
-Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
-will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
-
-SFMT is an extension of Fundamental Measure Theory (FMT)~\cite{rosenfeld1989}
-that was developed for the hard-sphere fluid which can be derived from
-Scaled Particle Theory (SPT)~\cite{santos2012phi3}. 
-Scaled Particle Theory~\cite{ReissSPT} 
+Equation~\ref{eq:Fexcess-simplified}).
+%In the next chapter on methods, we will use an SFMT functional that we 
+%adapt to fit the soft-sphere WCA fluid.
+SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
+for soft-spheres, and is an extension of Fundamental Measure Theory (FMT) 
+originally developed by 
+Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
+for the hard-sphere fluid. FMT in turn extented  
+Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
+developed for homogeneous hard-sphere fluids, 
+to inhomogeneous, or non-uniform hard-sphere fluids.
+
+%In order to apply classical Density Functional Theory to a system that has
+%a fixed temperature, volume, and number of particles, a functional for 
+%the Helmholtz free energy must first be formed. An expression for the ideal 
+%portion of the Helmholtz free energy is given by Equation~\ref{eq:Fideal-n(r)}, 
+%but forming an expression for the excess Helmholtz free
+%energy from Equation~\ref{eq:Fexcess-simplified} for the particular fluid 
+%being modeled can be challenging. 
+%Soft Fundamental Measure Theory (SFMT) is a method for forming an 
+%expression for the excess Helmholtz free energy of a soft-sphere fluid
+%that does not depend on explicitly taking into account the interaction 
+%that each sphere with every other sphere in the fluid (as required by 
+%Equation~\ref{eq:Fexcess-simplified}). In the next chapter on methods, we
+%will use an SFMT functional that we adapt to fit the soft-sphere WCA fluid.
+%SFMT was developed by Schmidt~\cite{schmidt1999density, schmidt2000fluid} 
+%for soft-spheres. It is an extension of Fundamental Measure Theory (FMT) 
+%originally developed by 
+%Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
+%for the hard-sphere fluid, and which expanded 
+%Scaled Particle Theory (SPT)~\cite{santos2012phi3} 
+%to inhomogeneous, or non-uniform hard-sphere fluids.
+
+Scaled Particle Theory
 seeks a general expression for the excess chemical 
@@
 \end{align}
+
 The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
@@
 \end{align}
+\begin{figure}
+    \centering
+    \includegraphics[height=7cm]{figs/P_overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where 
+    there is no overlap, such as that shown by the black dot near the 
+    top. But it is not possible to place a sphere at a position where 
+    an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
 The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
@@
 in Figure~\ref{fig:p_overlap}. 
-\begin{figure}
-    \centering
-    \includegraphics[height=7cm]{figs/P_overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where 
-    there is no overlap, such as that shown by the black dot near the 
-    top. But it is not possible to place a sphere at a position where 
-    an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
 In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
@@
 \begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
+
 Scaled Particle Theory forms a more general result for the reversible work 
@@
 
+\begin{figure}
+  \centering
+  \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
+  \caption{A primitive crystal cell is shown by the shaded parallelepiped.}
+  \label{fig:primitivecell}
+\end{figure}
+
 It is known that a fluid with a soft potential like the WCA fluid will 
@@
   \centering
-    \includegraphics[height=3.5cm]{PrimitiveCellLightBlue.png}
-    \caption{A primitive crystal cell is shown by the shaded parallelepiped.}
-    \label{fig:primitivecell}
+  \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
+  \caption{On the left are shown four examples of possible positions 
+  that an atom (blue sphere) might have within a cell that has one 
+  lattice site (black dot) at the center. A large number of such 
+  possibilities would give rise to a Gaussian distribution, 
+  as shown on the right.}
+  \label{fig:Ensemble_Gaus}
 \end{figure}
@@
   \centering
-    \includegraphics[height=2.5cm]{Ensemble_Gaussian.png}
-    \caption{On the left are shown four examples of possible positions 
-    that an atom (blue sphere) might have within a cell that has one 
-    lattice site (black dot) at the center. A large number of such 
-    possibilities would give rise to a Gaussian distribution, 
-    as shown on the right.}
-    \label{fig:Ensemble_Gaus}
-  \end{figure}
-
- \begin{figure}
-    \centering
   \includegraphics[height=6cm]{Gaussian}
@@
 \end{align}
+
+\begin{figure}
+   \centering
+   \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
+   \caption{On the left are shown four examples of possible positions 
+   that an atom (blue sphere) might have within a fixed volume 
+   containing eight cells. Each cell has one lattice site at the center 
+   represented by a black dot. A large number of such possibilities 
+   would give rise to the Gaussian distributions shown on the right.}
+   \label{fig:Ensemble_Smallcells}
+\end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[height=2.5cm]{SameStatPic.png}
+   \caption{Front views of a cube with eight smaller cells. The two views 
+   show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
+   give rise to the same Gaussian distribution statistically, 
+   and the same number density.}
+   \label{fig:SameStatPic}
+\end{figure}
 
@@
    \centering
-    \includegraphics[height=2.5cm]{Ensemble_Smallcells.png}
-    \caption{On the left are shown four examples of possible positions 
-    that an atom (blue sphere) might have within a fixed volume 
-    containing eight cells. Each cell has one lattice site at the center 
-    represented by a black dot. A large number of such possibilities 
-    would give rise to the Gaussian distributions shown on the right.}
-    \label{fig:Ensemble_Smallcells}
-  \end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[height=2.5cm]{SameStatPic.png}
-    \caption{Front views of a cube with eight smaller cells. The two views 
-    show two ways of picturing a fraction of vacancies, $f_v$ of 7/8 which 
-    give rise to the same Gaussian distribution statistically, 
-    and the same number density.}
-    \label{fig:SameStatPic}
+   \includegraphics[height=8cm]{VaryWidthandVacancies.png}
+   \caption{Shows an example set of Gaussian distribution number density 
+   profiles $n(vec r)$ all with the same number density.
+   with the same number density of one atom over the fixed volume.}
+   \label{fig:Ensemble_vary}
+\end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
+   \caption{A homogeneous number density profile $n(\vec{r})$.}
+   \label{fig:homogen_denisty}
 \end{figure}
@@
 for all these number density profiles $n(\vec{r})$.  
-\begin{figure}
-    \centering
-    \includegraphics[height=8cm]{VaryWidthandVacancies.png}
-    \caption{Shows an example set of Gaussian distribution number density 
-    profiles $n(vec r)$ all with the same number density.
-    with the same number density of one atom over the fixed volume.}
-    \label{fig:Ensemble_vary}
-  \end{figure}  
- \begin{figure}
-    \centering
-    \includegraphics[height=4cm]{figs/homogeneous_bold-box.pdf}
-    \caption{A homogeneous number density profile $n(\vec{r})$.}
-    \label{fig:homogen_denisty}
-  \end{figure}  
 The Helmholtz free energy $F[n(\vec{r})]$ is calculated for each number 
@@
 but it must be tailored to fit a particular fluid\cite{schmidt1999density}.
+
 Our SFMT functional is tailored to fit a WCA fluid, and 
-is written in general form as 
+is written as 
 \begin{align}\label{eq:Fexfunctional}
@@
 \end{align}
-where
+%where 
 \begin{align}
@@
 
-\begin{figure}
-  \centering
-  \includegraphics[height=8cm]{plot_alpha.pdf}
-  \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
-  for $\sigma=1$ and $\epsilon=1$.}
-  \label{fig:alphaXivsT}
-\end{figure}
 The weight functions $w_{2}(r)$ and $w_{3}(r)$ are constructed using 
@@
 \end{equation}
+
+\begin{figure}
+  \centering
+  \includegraphics[height=8cm]{plot_alpha.pdf}
+  \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
+  for $\sigma=1$ and $\epsilon=1$.}
+  \label{fig:alphaXivsT}
+\end{figure}
+
 The parameter $\alpha(T)$ is derived by setting 
@@
 \end{figure}
+
 Figure \ref{fig:weight_functions} shows a one dimensional view of the 
@@
 with a uniform number density profile, n that does not depend on $\vec r$.
+
 \begin{figure}
@@
 % \end{figure}
+
+\begin{figure}
+   \centering
+   \includegraphics[height=5cm]{InclusionRadius.png}
+   \caption{Gaussians are centered at each lattice point. After a radius 
+   of a few Gaussian widths about each lattice point, the Gaussians go to 
+   zero and can be neglected.}
+   \label{fig:InclusionRadius}
+\end{figure}
 
@@
 Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
-\begin{figure}
-   \centering
-   \includegraphics[height=5cm]{InclusionRadius.png}
-   \caption{Gaussians are centered at each lattice point. After a radius 
-   of a few Gaussian widths about each lattice point, the Gaussians go to 
-   zero and can be neglected.}
-   \label{fig:InclusionRadius}
-\end{figure} 
 by randomly generating vectors
@@
 be under greater compression before forming a solid.
+
 White lines superimposed on the phase diagram show the boundaries of the 
@@
 and the volume of the solid, the slope should be positive.
+
 Comparing the slope of the phase boundary predicted by cDFT to that

### 2020-12-03 papers/thesis-kirstie/thesis.tex
spelling

@@
 Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
-for the hard-sphere fluid. FMT in turn extented  
+for the hard-sphere fluid. FMT in turn extended  
 Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 

### 2020-12-04 papers/thesis-kirstie/thesis.tex
fixed typos

@@
    \caption{Shows an example set of Gaussian distribution number density 
-   profiles $n(vec r)$ all with the same number density.
-   with the same number density of one atom over the fixed volume.}
+   profiles $n(\vec r)$ all with the same number density, n.}
    \label{fig:Ensemble_vary}

### 2020-12-05 papers/thesis-kirstie/thesis.tex
edits
edit

@@
 atom in each small cell as shown in Figure~\ref{fig:SameStatPic}.  
-This makes it convenient to describe the new 
-number density in terms of the fraction of vacancies, $f_v$. 
-The resulting number density is then given by
-\begin{align}
-    n = (1-f_v){n_{no.vacancies}}
+The number of particles within a primitive cell is then given by  
+\begin{align}
+    N= (1-f_v){N_{no.vacancies}}
 \end{align}
@@
 \begin{displaymath}\text{Best}\left(\frac{F}{N}\right)_{diff}= 1e100\end{displaymath}
+\begin{displaymath}\text{lattice constant}=\left(\frac{4(1-f_v)}{n}\right)^{\frac{1}{3}}\end{displaymath}
+The number of atoms in one primitive cell is one, $N_{primitive.cell}=1$.
+
 Loop: 

### 2020-12-05 papers/thesis-kirstie/thesis.tex
edit

@@
 \begin{displaymath}\text{lattice constant}=\left(\frac{4(1-f_v)}{n}\right)^{\frac{1}{3}}\end{displaymath}
-The number of atoms in one primitive cell is one, $N_{primitive.cell}=1$.
+The number of atoms in one primitive cell is one when there are no vacancies.
 

### 2020-12-07 papers/thesis-kirstie/thesis.tex
added math to make a point explicit, edit pres

@@
 \begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
+Taking the differential
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT +\frac{\partial f_N}{V}dV+ \frac{\partial f_N}{\partial \left(\frac{1}{N}\right)}d\left(\frac{1}{N}\right)\end{displaymath}
+and using the relation
+\begin{displaymath}d\left(\frac{1}{n}\right)=Vd\left(\frac{1}{N}\right)+\frac{1}{N}dV\end{displaymath}
+we get
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT + \frac{\partial f_N}{\partial \frac{1}{n}}d\left(\frac{1}{n}\right)\end{displaymath}
+
+\begin{equation}f_N(T,\frac{1}{n})\end{equation}
 
@@
 when the natural variables of the free energy are held fixed.
-For our case, with a fixed temperature, volume, and number of particles
-we get
+
+For our case, with a fixed temperature and number density we get
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
@@
 \end{align}
-where 
+%where 
 \begin{align}

### 2020-12-13 papers/thesis-kirstie/thesis.tex
pres edits
went back to previous intro with more detail on FMT

@@
 \usepackage{tablefootnote} %Packaged added to allow footnotes in the tabular environment, use \tablefootnote command
-\title{Freezing of a Weeks-Chandler-Anderson Fluid using classical Density Functional Theory}
+\title{Freezing of a Weeks-Chandler-Anderson Fluid using Classical Density Functional Theory}
 \author{Kirstie L. Finster}
@@
 
-The goal of this project is to generate temperature-density 
-and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
-(WCA) model fluid using classical Density Functional Theory (cDFT). 
-To do this, the Helmholtz free energies of the fluid
-in a liquid state and in a crystalline solid state must be found.
-In general, the Helmholtz free energy, which consists of an ideal part 
-and an excess part, can be written as a functional of the number density 
-profile $n(\vec{r})$ which describes the way the average number 
-density varies spatially. 
-\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
-
-Soft Fundamental Measure Theory (SFMT) will be used to form the 
-functional for the excess Helmholtz free energy. 
-SFMT is a method that makes the task of 
-forming an expression for the excess Helmholtz free energy of a particular 
-model fluid much easier since the interaction that each sphere, or atom, 
-has with every other sphere does not have to be taken explicitly into account.
-Once the functional for the Helmholtz free energy is formed, then 
-according to cDFT, the equilibrium
-Helmholtz free energy of the system can be found for a particular 
-temperature and density by varying the number 
-density profile $n(\vec{r})$ until the free energy is minimized. 
-Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
-Results will be compared to those from Monte-Carlo simulations.
+Classically, fluids have been modeled by a collection of small, 
+impenetrable spheres colliding with one another like billiard balls. 
+In the simplest case, the spheres are treated as having only a replulsive 
+potential. This model, called the hard-sphere fluid, has served as the 
+primary reference system for fluids. This is due to the fact that the 
+repulsion between atoms is much stronger than their attraction, and so 
+the repulsive potential, which is descriptive of hard spheres, dominates 
+the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
+fluid which is a perturbation of the hard-sphere fluid created by adding 
+a hard-edge attractive potential to the repulsive potential of the 
+hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
+spheres with softly fading replusive potentials which he modeled by making
+the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+This reflects the idea that soft-spheres penetrate more deeply into 
+each other when they collide at higher temperatures.
+
+In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
+with Rowlinson's idea of using temperature-dependent radii to model spheres 
+with soft repulsive potentials to create the Barker-Henderson model 
+fluid~\cite{barker1967perturbation}. In this model, a soft-sphere fluid is 
+modeled by using a hard-sphere fluid with hard spheres of larger radii 
+at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
+Chandler, and Anderson used temperature and density dependent hard-sphere 
+radii to model spheres with the more realistic softly fading repulsive 
+portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
+which is known as the Weeks-Chandler-Anderson model fluid.
+
+ADD MORE HERE TO COME UP TO DATE!!
+
+While models for particular fluids were being created, powerful methods 
+to describe the properties of fluids were also being developed. 
+In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
+introduced from which an equation of state, and an expression for the 
+excess Helmholtz free energy of a homogeneous hard-Sphere fluid can be 
+obtained~\cite{Hansen}. 
+%\textcolor{red}{
+In 1958, Percus-Yevick introduced an alternative method 
+for the hard-sphere fluid based on an approximate closure to the 
+Ornstein-Zernike integral equation~\cite{PercusYevick}.
+%Percus and Yevick took another approach to the hard-Sphere fluid.
+In 1961 Morita and Hiroike introduced a density functional 
+approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
+and the cooresponding equilibrium number density profile for 
+inhomogeneous fluids. To use this classical Density Functional Theory (cDFT) 
+approach, an expression for the free energy 
+as a functional of an average number density that varies spatially 
+must be constructed for a particular fluid.
+%DFT Evans 1979
+%Led by the consistency bewteen results from SPT and Percus-Yevick, 
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+(which can be constructed from SPT~\cite{rosenfeld1989,Hansen,santos2012phi3}) 
+to form a functional for the hard-sphere fluid.
+FMT is based on the idea that for a hard-sphere fluid the
+lowest term in the virial expansion of the excess chemical potential can 
+be expressed in terms of a sum of weighted densities using a weight 
+function of the form of a backward Heavyside step function (shown below) 
+that can be expressed in terms of a sum of convolutions involving 
+a set of weight functions that can be related to the geometry of a sphere.
+\begin{displaymath}
+   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = \end{displaymath}
+\begin{equation} w_i^{(3)} x w_j^{(0)} + w_j^{(3)} x w_i^{(0)} + w_i^{(1)} x w_j^{(2)} + w_j^{(1)} x w_i^{(2)} + \vec w_i^{(1)} x \vec w_j^{(2)} + \vec w_j^{(1)} x \vec w_i^{(2)}
+\end{equation}
+Tarazona later improved on the 
+method by introducing a tensor version of one of the energy density terms 
+in the functional~\cite{tarazonaphi3, 
+santos2012phi3}, and in 1999 Schmidt adopted FMT for spheres with soft 
+interatomic potential energies
+(which rise to infinity over a distance, rather than instantaneously 
+like the potential energy between hard spheres)
+forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
+
+%by replacing $w^{(3)}$ with a weight function that would result in  
+%soft interatomic potentials (which rise to infinity over a distance, 
+%rather than instantaneously).
+
+Schmidt applied his Soft Fundamental Measure Theory to various fluids 
+with soft repulsive and soft attractive potentials, and developed an error function  
+model 
+%based on a Gaussian weight function 
+for spheres with steep, short-range soft repulsive 
+potentials~\cite{schmidt2000fluid}.
+%, similar to that exhibited by a WCA fluid. 
+In this paper, clasical Density Functional 
+Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
+model fluid by applying SFMT with Schmidt's proposed error function model, 
+in which we have incorporated our temperature dependent parameters to fit the 
+functional to the soft-sphere WCA fluid.
+%form the functional for the excess Helmholtz free energy. 
+%for generating the weight funcions, and Tarazona's tensor functional. 
+Results will be compared to those generated from Monte-Carlo simulations.
+
+%Results will be compared to those from Monte-Carlo simulations.
+
+%Newer Intro....
+%The goal of this project is to generate temperature-density 
+%and pressure-temperature phase diagrams for a Weeks-Chandler-Anderson 
+%(WCA) model fluid using classical Density Functional Theory (cDFT). 
+%To do this, the Helmholtz free energies of the fluid
+%in a liquid state and in a crystalline solid state must be found.
+%In general, the Helmholtz free energy, which consists of an ideal part 
+%and an excess part, can be written as a functional of the number density 
+%profile $n(\vec{r})$ which describes the way the average number 
+%density varies spatially. 
+%\begin{equation}{F[n(\vec{r})]=F_{ideal}[n(\vec{r})] + F_{excess}[n(\vec{r})]}\end{equation}
+
+%Soft Fundamental Measure Theory (SFMT) will be used to form the 
+%functional for the excess Helmholtz free energy. 
+%SFMT is a method that makes the task of 
+%forming an expression for the excess Helmholtz free energy of a particular 
+%model fluid much easier since the interaction that each sphere, or atom, 
+%has with every other sphere does not have to be taken explicitly into account.
+%Once the functional for the Helmholtz free energy is formed, then 
+%according to cDFT, the equilibrium
+%Helmholtz free energy of the system can be found for a particular 
+%temperature and density by varying the number 
+%density profile $n(\vec{r})$ until the free energy is minimized. 
+%Phase diagrams can then be constructed using Maxwell's Double Tangent Construction. 
+%Results will be compared to those from Monte-Carlo simulations.
 

### 2020-12-13 papers/thesis-kirstie/thesis.tex
edits

@@
 each other when they collide at higher temperatures. 
-
 In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
@@
 In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
-(which can be constructed from SPT~\cite{rosenfeld1989,Hansen,santos2012phi3}) 
-to form a functional for the hard-sphere fluid.
-FMT is based on the idea that for a hard-sphere fluid the
-lowest term in the virial expansion of the excess chemical potential can 
+to form a functional for the hard-sphere fluid based on the idea that 
+the lowest term in the virial expansion 
+of the excess chemical potential for a hard-sphere fluid can 
 be expressed in terms of a sum of weighted densities using a weight 
@@
 that can be expressed in terms of a sum of convolutions involving 
-a set of weight functions that can be related to the geometry of a sphere.
+a set of weight functions related to the geometry of a sphere.
 \begin{displaymath}
@@
 \end{equation}
+FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
+SPT (from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3}),
+and for the direct correlation function as Percus-Yevick.
 Tarazona later improved on the 
@@
 The potential is zero when the distance r between the centers of two atoms 
-is equal to $\sigma$. At this point 
-the attraction between the atoms just cancels their repulsion. As the distance 
+is equal to $\sigma$. 
+%At this point the attraction between the atoms just cancels their repulsion. 
+As the distance 
 between the centers of the two atoms decreases below $\sigma$, the potential 

### 2020-12-14 papers/thesis-kirstie/thesis.tex
edits

@@
 ADD MORE HERE TO COME UP TO DATE!!
+\newline
 
@@
 introduced from which an equation of state, and an expression for the 
-excess Helmholtz free energy of a homogeneous hard-Sphere fluid can be 
+excess Helmholtz free energy of a homogeneous hard-sphere fluid can be 
 obtained~\cite{Hansen}. 
@@
 Ornstein-Zernike integral equation~\cite{PercusYevick}.
-%Percus and Yevick took another approach to the hard-Sphere fluid.
+%Percus and Yevick took another approach to the hard-sphere fluid.
 In 1961 Morita and Hiroike introduced a density functional 
@@
 %Led by the consistency bewteen results from SPT and Percus-Yevick, 
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989} 
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
 to form a functional for the hard-sphere fluid based on the idea that 
@@
 \begin{displaymath}
-   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = \end{displaymath}
-\begin{equation} w_i^{(3)} x w_j^{(0)} + w_j^{(3)} x w_i^{(0)} + w_i^{(1)} x w_j^{(2)} + w_j^{(1)} x w_i^{(2)} + \vec w_i^{(1)} x \vec w_j^{(2)} + \vec w_j^{(1)} x \vec w_i^{(2)}
+   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
+\begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
 \end{equation}
 FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
-SPT (from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3}),
-and for the direct correlation function as Percus-Yevick.
+SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
+and for the direct correlation function given by Percus-Yevick.
 Tarazona later improved on the 
@@
 Helmholtz free energy per atom $f$ in terms of the Gibbs free energy 
-per atom  $g$ (which is also the chemical potential)  
-\begin{align}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{align}
+per atom  $g$ (which is the chemical potential)  
+\begin{equation}\frac{F}{N} = -P\frac{V}{N}+\frac{G}{N} \end{equation}
 \begin{align}f=-P\frac{1}{n}+g\end{align}
@@
     the number densities of the liquid and the solid during a phase 
-    transition. The gray region is where the liquid and the solid coexist, 
-    and are therefore not pure states.}
+    transition. The gray region is where different phases coexist in 
+    varying amounts at their phase transition number densities.}
     \label{fig:T-n_Diagram}
@@
 can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
-but without the gray hump that is characteristic of fluids that have 
+but without the gray hump characteristic of fluids that have 
 both attractive and repulsive interatomic potentials.

### 2020-12-15 papers/thesis-kirstie/thesis.tex
edits

@@
 impenetrable spheres colliding with one another like billiard balls. 
-In the simplest case, the spheres are treated as having only a replulsive 
+In the simplest case, the spheres are treated as having only a repulsive 
 potential. This model, called the hard-sphere fluid, has served as the 
@@
 hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-spheres with softly fading replusive potentials which he modeled by making
+spheres with softly fading repulsive potentials which he modeled by making
 the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
@@
 each other when they collide at higher temperatures. 
-In 1967 Barker and Henderson combined the perturabtion approach of Zwanzig 
+In 1967 Barker and Henderson combined the perturbation approach of Zwanzig 
 with Rowlinson's idea of using temperature-dependent radii to model spheres 
@@
 approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
-and the cooresponding equilibrium number density profile for 
+and the corresponding equilibrium number density profile for 
 inhomogeneous fluids. To use this classical Density Functional Theory (cDFT) 
@@
 %DFT Evans 1979
-%Led by the consistency bewteen results from SPT and Percus-Yevick, 
+%Led by the consistency between results from SPT and Percus-Yevick, 
 In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
@@
 SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
-and for the direct correlation function given by Percus-Yevick.
+and for the direct correlation function approximation given by Percus-Yevick.
 Tarazona later improved on the 
@@
 %, similar to that exhibited by a WCA fluid. 
-In this paper, clasical Density Functional 
+In this paper, classical Density Functional 
 Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
@@
 %form the functional for the excess Helmholtz free energy. 
-%for generating the weight funcions, and Tarazona's tensor functional. 
+%for generating the weight functions, and Tarazona's tensor functional. 
 Results will be compared to those generated from Monte-Carlo simulations.
@@
     \centering
-    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
+    \includegraphics[height=8cm]{figs/P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
     \centering
-    \includegraphics[height=6cm]{figs/T-n_Diagram.pdf}
+    \includegraphics[height=8cm]{figs/T-n_Diagram.pdf}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 

### 2020-12-15 papers/thesis-kirstie/thesis.tex
edits

@@
 
-\section{Freezing and Maxwell's Double Tangent Construction}
-
-Freezing is governed by entropy, which the universe, or any closed system, 
+\section{Freezing and the Construction of Phase Diagrams}
+
+Freezing is governed by entropy, which the universe %, or any closed system, 
 seeks to maximize according to the second law of Thermodynamics. When the 
@@
 temperatures and number densities, so that phase diagrams can be 
-constructed using Maxwell's Double Tangent Construction. 
+constructed. 
 The key idea behind classical Density Functional Theory (cDFT) is that 
@@
 To construct the phase diagram, the liquid and solid phase transition number densities
-must be found. Maxwell's double tangent construction can be used to find 
-these number densities given plots of the liquid and the solid Helmholtz 
-free energies verses the inverse density at a given temperature, as
-discussed in Section 2.2. We use classical Density Functional Theory 
+are found
+
+%Maxwell's double tangent construction can be used to find 
+%these number densities given plots of
+
+from the liquid and the solid Helmholtz 
+free energies and inverse densities at a given temperature, as
+discussed in Section 2.2.
+
+We use classical Density Functional Theory 
 to find the Helmholtz free energies of the liquid and the solid. 
@@
 completely in a liquid state if the liquid free energy is lower than 
-the crystal free energy. Maxwell's Double Tangent Construction is used 
-to determine the actual number densities of the crystal and liquid states
+the crystal free energy. 
+We determine the actual number densities of the crystal and liquid states
 at the boundaries of the coexistence region for a given temperature, 
@@
 Helmholtz free energy for the liquid and solid states were found for a given
-temperature and number density, Maxwell's Double Tangent Construction 
-was used to find the number densities of the liquid and the solid at 
-the phase transition.
+temperature and number density, 
+we used this data 
+to find the number densities of the liquid and the solid at the phase transition.
 Using this information, we generated a temperature-density 

### 2020-12-15 papers/thesis-kirstie/thesis.tex

@@
 and the corresponding equilibrium number density profile for 
-inhomogeneous fluids. To use this classical Density Functional Theory (cDFT) 
+inhomogeneous fluids. To use the classical Density Functional Theory (cDFT) 
 approach, an expression for the free energy 
@@
 \end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[height=8cm]{figs/w2-w3-radius-thickness.png}
+  %%\caption{}
+  %\label{fig:weight_functions}
+%\end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[height=8cm]{figs/df-dr_withXi_fromB2.pdf}
+  %%\caption{}
+%\end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[height=8cm]{figs/erf_Potential_overkT-xifromB2.pdf}
+  %%\caption{}
+%\end{figure}
+
+%\begin{figure}
+  %\centering
+  %%\includegraphics[height=8cm]{figs/B2_vs_T_with_xi_fromB2.pdf}
+  %%\caption{}
+%\end{figure}
 

### 2020-12-16 papers/thesis-kirstie/thesis.tex
edits

@@
 Schmidt applied his Soft Fundamental Measure Theory to various fluids 
-with soft repulsive and soft attractive potentials, and developed an error function  
-model 
+with soft, short-range repulsive and attractive potentials, and developed 
+an error function model 
 %based on a Gaussian weight function 

### 2020-12-23 papers/thesis-kirstie/thesis.tex
edits and figure

@@
 \begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
-The mayor function can then be written in terms of the weight functions~\cite{Hansen}
+The Mayer function can then be written in terms of the weight functions~\cite{Hansen}
 \begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
@@
 
-\textbf{4. Integrals used to Compute Fourier Transforms of Weight Functions}
+\section{Integrals used to Compute Weight Function Fourier Transforms}
 \begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
@@
 \begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}}\end{equation}
+
+\section{Deconvolution of the Hard-Sphere Mayer Function}
+
+The Mayer function written in terms of convolutions of weight functions is
+\begin{align}
+    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})  
+\end{align}
+where the mayer function of the hard-sphere fluid is 
+\begin{align}
+   f(r) &= e^{\frac{V(r)}{kBT}}-1 \\
+        &= -\frac{\theta(|\vec r|-2R)}{2}
+\end{align}
+and $V(r)$ is given by
+\begin{align}
+    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+$\theta(|\vec r|-2R)$ is a backwards Heavyside step function 
+given by 
+\begin{align}
+     \Theta(|\vec{r}|-2R)=\left\{ \begin{array}{rc} 1 & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+We can verfy that 
+\begin{align}  \label{equation-to-verify}
+     \frac{\Theta(|\vec{r}|-2R)}{2}= w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}
+\end{align}
+by using the relation
+\begin{align}
+    f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]
+\end{align} 
+\begin{align}\label{FTdecon}
+    F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
+        &= F.T.\left[w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}\right]  \\                                                  
+        &= F.T.\left[I.F.T\left(\widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1\right)\right]  \\
+        &= \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
+\end{align}
+
+Setting $\hat k$ in the $\hat z'$ direction, the right hand of Equation~\ref{FTdecon} gives
+\begin{align} \label{leftside}
+    F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
+       &=\frac{1}{2}\int_0^\pi \int_{-1}^1\int_0^{2R}(1)e^{-ikr\cos \theta'}r^2~dr~d\cos \theta'~d\phi'   \\
+       &=\frac{2\pi}{2}\int_0^{2R} r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}~d\cos \theta'~d\phi'   \\
+       &=\pi\int_0^{2R} r^2\left(\frac{2\sin (kr)}{kr}\right)~dr  \\
+       &=\frac{2\pi}{k}\int_0^{2R} r\sin (kr)~dr  \\
+       &=\frac{2\pi}{k}\left[\frac{\sin (kr)-kr\cos (kr))}{k^2}\right] \bigg|_0^{2R} \\
+       &= \frac{2\pi}{k^3}\sin (2kR)-\frac{4\pi R}{k^2}\cos (2kR)  \\
+      \label{leftside-solved} &= \frac{4\pi}{k^3}\sin (kR)\cos (kR) - \frac{4 \pi R}{k^2}\cos^2 (kR) + \frac{4\pi R}{k^2} \sin^2 (kR)
+\end{align}
+
+The Fourier transforms of the hard-sphere weight functions give
+\begin{align}
+    \widetilde{w}_2 &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'} r^2~dr~d\cos \theta'~d\phi   \\
+                    &= 2\pi\int\delta(r-R)r^2 \frac{2\sin (kr)}{kr}~dr  \\
+                    &= 2\pi R^2 \frac{2\sin (kR)}{kR}  \\
+                    &= \frac{4\pi R}{k} \sin (kR)
+\end{align}
+
+\begin{align}
+    \widetilde{w}_1 &= \frac{\widetilde{w}_2}{4\pi R}= \frac{\sin (kR)}{k} 
+                    %&= \frac{\sin (kR)}{k} 
+\end{align}
+\begin{align}
+    \widetilde{w}_0 &= \frac{\widetilde{w}_2}{4\pi R^2}= \frac{\sin (kR)}{kR}
+                   % &= \frac{\sin (kR)}{kR} 
+\end{align}
+
+\begin{align}
+    \widetilde{w}_3 &= \int\int\int\theta(r-R)e^{-ikr\cos \theta'}r^2~dr~d\cos\theta'~d\phi  \\
+                    &= 2\pi\int\theta(r-R)\left(\frac{2\sin (kr)}{kr}\right)r^2~dr   \\
+                    &= \frac{4\pi}{k}\int_0^R (1)r\sin (kr)~dr   \\
+                    &= \frac{4\pi}{k}\left[\frac{\sin (kr) - kr\cos (kr)}{k^2}\right]\bigg|_0^R   \\
+                    &= \frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]
+\end{align}
+
+\begin{align}
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r^2~\frac{\vec r}{r}~dr~d\cos\theta'~d\phi  \\
+                          &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r_{x'} \hat x' + r_{y'} \hat y'+ r_{z'} \hat z')~dr~d\cos\theta'~d\phi  
+\end{align}
+Again $\vec{k}$ is set in the direction of $\hat{z}'$ of a rotated coordinate 
+system $(x',y'z')$ where
+\begin{align}
+     r_{x'} &= r\sin\theta'\cos\phi'  \\
+     r_{y'} &= r\sin\theta'\sin\phi'   \\
+     r_{z'} &= r\cos\theta'  \\
+     \text{and~~~} r &=|\vec r| = |\vec r'|
+\end{align}
+The terms in the $\hat x'$ and $\hat y'$ direction give zero due to symmetry ($\int_0^{2\pi} \cos \phi ~d\phi= 0$ and $\int_0^{2\pi} \sin \phi~d\phi= 0$), 
+and so only the term in the $\hat z'$ direction remains.
+
+\begin{align}
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r\cos \theta')~dr~d\cos\theta'~d\phi~\hat k   \\
+                          &= 2\pi\int\delta(r-R)r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}\cos \theta'~d\cos\theta'~\hat k   \\
+                          &= 2\pi\int\delta(r-R)r^2 \left(\frac{-2i}{kr}\cos (kr)+ \frac{2i}{k^2r^2}\sin (kr)\right)dr~\hat k   \\
+                          &= 2\pi\left[\frac{-2iR}{k}\cos (kR)+ \frac{2i}{k^2}\sin (kR)\right]~\hat k   \\
+                          &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k 
+\end{align}
+
+\begin{align}
+    \widetilde{\vec w}_1  &= \frac{\widetilde{\vec w}_2}{4\pi R}   \\
+                          &= \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right)\hat k
+\end{align}
+The terms on the right side of Equation~\ref{FTdecon} are then
+\begin{align}
+    \widetilde{w}_0\widetilde{w}_3 &= \left(\frac{\sin (kR)}{kR}\right)\left(\frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]\right)  \\
+                                   &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR)
+\end{align}
+
+\begin{align}
+    \widetilde{w}_1\widetilde{w}_2 &= \left(\frac{\sin (kR)}{k}\right)\left(\frac{4\pi R}{k} \sin (kR)\right)  \\                                
+                                   &= \frac{4\pi R}{k^2}\sin^2 (kR)
+\end{align}
+
+\begin{align}
+   \widetilde{\vec w}_1\cdot\widetilde{\vec w}_2   
+    &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k \cdot \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right) \hat k   \\
+    &= -\frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) - \frac{4\pi}{Rk^4}\sin^2(kR)
+\end{align}
+Adding the terms together gives
+\begin{align}
+    \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
+    &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR) \\
+    &+ \frac{4\pi R}{k^2}\sin^2 (kR) \\
+    &- \frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) -  \frac{4\pi}{Rk^4}\sin^2(kR)  \\
+    &= \left(\frac{4\pi}{k^4R}+\frac{4\pi R}{k^2}-\frac{4\pi}{k^4R}\right)\sin^2(kR)\\
+    &-\frac{4\pi R}{k^2}\cos^2(kR) +\left(-\frac{4\pi}{k^3} + \frac{8\pi}{k^3}\right)\sin(kR)\cos(kR)   \\
+    &=\frac{4\pi R}{k^2}\sin^2(kR)-\frac{4\pi R}{k^2}\cos^2(kR)+\frac{4\pi}{k^3}\sin(kR)\cos(kR)
+\end{align}
+which equals Equation~\ref{leftside-solved}, and so
+%\begin{align}
+     %\Theta(|\vec{r}|-2R)= 2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+%\end{align}
+Equation~\ref{equation-to-verify} is satisfied.
 

### 2020-12-24 papers/thesis-kirstie/thesis.tex
fixing margins

--- 2020-12-16 papers/thesis-kirstie/thesis.tex
@@
 to undergo a phase transition from a liquid to a solid, as would be expected. 
-Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
+Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) 
+is given by
 \begin{align} 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right]{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
+%\begin{align}
+    %\widetilde{\vec{w}}_1(\vec{k}) &= \frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'  \\
+     %&= \frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'   \\
+     %&= \frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'   \\
+     %&= \frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~~}\hat{z}' \\
+     %&= \frac{i}{k\sqrt{\pi}a}\left[-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right]{~~}\hat{z}'
+%\end{align}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~}\hat{z}'}\end{equation}
+
+%KEEP%\begin{equation}\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left(-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right){~}\hat{z}'\end{equation}
+\begin{equation}
+   \widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left(-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}
+   \cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}
+   \left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)\color{white}
+   \right]\color{white}\right)\color{black}  \nonumber
+\end{equation} 
+\begin{equation}
+   +\color{white}\left(\left[\color{black}\operatorname{erf}
+   \left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\color{black}\right)
+   \color{black}{~}\hat{z}'
+\end{equation}
+
+%KEEP%\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]\vec{k}}\end{equation}
+\begin{equation}
+   \widetilde{\vec{w}}_1(\vec{k}) = \frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}
+   e^{-\left(\frac{\alpha}{2a}\right)^2}\color{white}\right]\color{black}
+   \left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)
+   \color{white}\right]\color{black}  \nonumber
+\end{equation} 
+\begin{equation}  
+   \color{white}\left[\color{black} +\operatorname{erf}\left(\frac{ka}{2}
+   -\frac{i\alpha}{2a}\right)\right]\color{white}\left[\color{black}
+   -e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]\color{black}{~}\vec k 
+\end{equation}
+
 Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
 $\widetilde{\vec{w}}_1(k)$ becomes
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
+%KEEP%\begin{equation}\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}\end{equation}
+\begin{equation}
+   \widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}
+   e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}
+   \left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)
+   +\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}
+   \right)\right]\color{white}\right]\color{black}  \nonumber
+\end{equation}
+\begin{equation}
+   \color{white}\left[\color{black}-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\color{black}\vec{k}
+\end{equation}
 
@@
 about $z'$. That leaves
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{w}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(-\frac{2i}{kr}\cos(kr) + \frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4i\sqrt{\pi}}{ak}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r} + \frac{4i\sqrt{\pi}}{ak^2}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}{~~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4\sqrt{\pi}i}{ak}\left[\frac{a\alpha\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]}\end{equation}
-\begin{displaymath}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~~}\hat{z}'}\end{displaymath}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{w}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(-\frac{2i}{kr}\cos(kr) + \frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4i\sqrt{\pi}}{ak}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r} + \frac{4i\sqrt{\pi}}{ak^2}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4\sqrt{\pi}i}{ak}\left[\frac{a\alpha\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]} \nonumber \end{equation}
+\begin{equation}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\vec{k}}\end{equation}
 Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
 $\widetilde{\vec{w}}_2(k)$ becomes
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\vec{k}}\end{equation}
 
@@
 Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{y'})}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{y'}}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{z'})}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{z'}}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{y'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{z'})}{r^2}-\frac{\delta_{yz}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{z'}}{r^2}-\frac{\delta_{yz}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'x'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{x'})}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{x'}}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
 \begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
@@
 Calculate $\widetilde{w}_{{m2}_{y'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
+\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
 \begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
@@
 Calculate $\widetilde{w}_{{m2}_{z'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 \begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
@@
 
-\begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}
-\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
+\begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2\color{white}\right]\color{black}}\end{displaymath}
+\begin{equation}{\color{white}\left[\color{black} -\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)   +n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
 
@@
 
-\begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
+%KEEP%\begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
+\begin{equation}
+     \Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)  \color{white}\right)\color{black}\nonumber
+\end{equation}
+\begin{equation}
+      +\color{white}\left(\color{black}\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)\color{black} 
+\end{equation}
 
@@
 \begin{equation}{\Phi_3=\frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}}\end{equation}
+
+\section{Deconvolution Equation for the Hard-Sphere Mayer Function}
+
+The Mayer function written in terms of convolutions of weight functions is
+\begin{align}
+    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})  
+\end{align}
+where the mayer function of the hard-sphere fluid is 
+\begin{align}
+   f(r) &= e^{\frac{V(r)}{kBT}}-1 \\
+        &= -\frac{\theta(|\vec r|-2R)}{2}
+\end{align}
+and $V(r)$ is given by
+\begin{align}
+    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+$\theta(|\vec r|-2R)$ is a backwards Heavyside step function 
+given by 
+\begin{align}
+     \Theta(|\vec{r}|-2R)=\left\{ \begin{array}{rc} 1 & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+We can verfy that 
+\begin{align}  \label{equation-to-verify}
+     \frac{\Theta(|\vec{r}|-2R)}{2}= w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}
+\end{align}
+by using the relation
+\begin{align}
+    f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]
+\end{align} 
+\begin{align}\label{FTdecon}
+    F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
+        &= F.T.\left[w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}\right]  \nonumber \\                                                  
+        &= F.T.\left[I.F.T\left(\widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1\right)\right]  \nonumber \\
+        &= \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
+\end{align}
+
+Setting $\hat k$ in the $\hat z'$ direction, the right hand of Equation~\ref{FTdecon} gives
+\begin{align} \label{leftside} 
+    F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
+       &=\frac{1}{2}\int_0^\pi \int_{-1}^1\int_0^{2R}(1)e^{-ikr\cos \theta'}r^2~dr~d\cos \theta'~d\phi'  \nonumber  \\
+       &=\frac{2\pi}{2}\int_0^{2R} r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}~d\cos \theta'~d\phi' \nonumber   \\
+       &=\pi\int_0^{2R} r^2\left(\frac{2\sin (kr)}{kr}\right)~dr \nonumber  \\
+       &=\frac{2\pi}{k}\int_0^{2R} r\sin (kr)~dr \nonumber  \\
+       &=\frac{2\pi}{k}\left[\frac{\sin (kr)-kr\cos (kr))}{k^2}\right] \bigg|_0^{2R} \nonumber  \\
+       &= \frac{2\pi}{k^3}\sin (2kR)-\frac{4\pi R}{k^2}\cos (2kR) \nonumber  \\
+       &= \frac{4\pi}{k^3}\sin (kR)\cos (kR) - \frac{4 \pi R}{k^2}\cos^2 (kR) + \frac{4\pi R}{k^2} \sin^2 (kR)
+\end{align}
+
+The Fourier transforms of the hard-sphere weight functions give
+\begin{align}
+    \widetilde{w}_2 &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'} r^2~dr~d\cos \theta'~d\phi  \nonumber  \\
+                    &= 2\pi\int\delta(r-R)r^2 \frac{2\sin (kr)}{kr}~dr \nonumber  \\
+                    &= 2\pi R^2 \frac{2\sin (kR)}{kR} \nonumber  \\
+                    &= \frac{4\pi R}{k} \sin (kR)
+\end{align}
+
+\begin{align}
+    \widetilde{w}_1 &= \frac{\widetilde{w}_2}{4\pi R}= \frac{\sin (kR)}{k} 
+                    %&= \frac{\sin (kR)}{k} 
+\end{align}
+\begin{align}
+    \widetilde{w}_0 &= \frac{\widetilde{w}_2}{4\pi R^2}= \frac{\sin (kR)}{kR}
+                   % &= \frac{\sin (kR)}{kR} 
+\end{align}
+
+\begin{align}
+    \widetilde{w}_3 &= \int\int\int\theta(r-R)e^{-ikr\cos \theta'}r^2~dr~d\cos\theta'~d\phi  \nonumber \\
+                    &= 2\pi\int\theta(r-R)\left(\frac{2\sin (kr)}{kr}\right)r^2~dr  \nonumber  \\
+                    &= \frac{4\pi}{k}\int_0^R (1)r\sin (kr)~dr  \nonumber  \\
+                    &= \frac{4\pi}{k}\left[\frac{\sin (kr) - kr\cos (kr)}{k^2}\right]\bigg|_0^R  \nonumber  \\
+                    &= \frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]
+\end{align}
+
+\begin{align}
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r^2~\frac{\vec r}{r}~dr~d\cos\theta'~d\phi \nonumber  \\
+                          &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r_{x'} \hat x' + r_{y'} \hat y'+ r_{z'} \hat z')~dr~d\cos\theta'~d\phi  
+\end{align}
+Again $\vec{k}$ is set in the direction of $\hat{z}'$ of a rotated coordinate 
+system $(x',y'z')$ where
+\begin{align}
+     r_{x'} &= r\sin\theta'\cos\phi'  \\
+     r_{y'} &= r\sin\theta'\sin\phi'   \\
+     r_{z'} &= r\cos\theta'  \\
+     \text{and~~~} r &=|\vec r| = |\vec r'|
+\end{align}
+The terms in the $\hat x'$ and $\hat y'$ direction give zero due to symmetry ($\int_0^{2\pi} \cos \phi ~d\phi= 0$ and $\int_0^{2\pi} \sin \phi~d\phi= 0$), 
+and so only the term in the $\hat z'$ direction remains.
+
+\begin{align}
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r\cos \theta')~dr~d\cos\theta'~d\phi~\hat k  \nonumber  \\
+                          &= 2\pi\int\delta(r-R)r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}\cos \theta'~d\cos\theta'~\hat k  \nonumber  \\
+                          &= 2\pi\int\delta(r-R)r^2 \left(\frac{-2i}{kr}\cos (kr)+ \frac{2i}{k^2r^2}\sin (kr)\right)dr~\hat k  \nonumber  \\
+                          &= 2\pi\left[\frac{-2iR}{k}\cos (kR)+ \frac{2i}{k^2}\sin (kR)\right]~\hat k \nonumber   \\
+                          &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k 
+\end{align}
+
+\begin{align}
+    \widetilde{\vec w}_1  &= \frac{\widetilde{\vec w}_2}{4\pi R}  \nonumber  \\
+                          &= \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right)\hat k
+\end{align}
+The terms on the right side of Equation~\ref{FTdecon} are then
+\begin{align}
+    \widetilde{w}_0\widetilde{w}_3 &= \left(\frac{\sin (kR)}{kR}\right)\left(\frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]\right) \nonumber  \\
+                                   &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR)
+\end{align}
+
+\begin{align}
+    \widetilde{w}_1\widetilde{w}_2 &= \left(\frac{\sin (kR)}{k}\right)\left(\frac{4\pi R}{k} \sin (kR)\right) \nonumber  \\                                
+                                   &= \frac{4\pi R}{k^2}\sin^2 (kR)
+\end{align}
+
+\begin{align}
+   \widetilde{\vec w}_1\cdot\widetilde{\vec w}_2   
+    &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k \cdot \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right) \hat k  \nonumber  \\
+    &= -\frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) - \frac{4\pi}{Rk^4}\sin^2(kR)
+\end{align}
+Adding the terms together gives
+\begin{align}
+    \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
+    &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR) \nonumber  \\
+    &+ \frac{4\pi R}{k^2}\sin^2 (kR) \nonumber  \\
+    &- \frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) -  \frac{4\pi}{Rk^4}\sin^2(kR) \nonumber  \\
+    &= \left(\frac{4\pi}{k^4R}+\frac{4\pi R}{k^2}-\frac{4\pi}{k^4R}\right)\sin^2(kR) \nonumber   \\
+    &-\frac{4\pi R}{k^2}\cos^2(kR) +\left(-\frac{4\pi}{k^3} + \frac{8\pi}{k^3}\right)\sin(kR)\cos(kR) \nonumber   \\
+    &=\frac{4\pi R}{k^2}\sin^2(kR)-\frac{4\pi R}{k^2}\cos^2(kR)+\frac{4\pi}{k^3}\sin(kR)\cos(kR)
+\end{align}
+which equals Equation~\ref{leftside}, and so
+%\begin{align}
+     %\Theta(|\vec{r}|-2R)= 2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+%\end{align}
+Equation~\ref{equation-to-verify} is satisfied.
 
@@
 \begin{align}
- B_2 &= -\frac{1}{2}\int_0^{\pi}\int_0^{2\pi}\int_0^\infty\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2\sin{\theta}d{\theta}d{\phi}dr \\
-     &= -\frac{1}{2}4\pi\frac{1}{2}\int_{0}^{r_{max}}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2dr \\
-     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \pi\int_0^{r_{max}}r^2dr   \\
+ B_2 &= -\frac{1}{2}\int_0^{\pi}\int_0^{2\pi}\int_0^\infty\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2\sin{\theta}d{\theta}d{\phi}dr \nonumber  \\
+     &= -\frac{1}{2}4\pi\frac{1}{2}\int_{0}^{r_{max}}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2dr \nonumber  \\
+     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \pi\int_0^{r_{max}}r^2dr  \nonumber  \\
      &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \frac{\pi}{3}r_{max}^3   
@@
 \begin{align}
-  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}r_{max}^3\right)\operatorname{erf}\left(\frac{\alpha-r_{max}}{\Xi}\right) \\
-      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
-      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha r_{max}-\Xi^3-\Xi r_{max}^2\right)e^{-{\left(\frac{\alpha-r_{max}}{\Xi}\right)^2}} \\
+  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}r_{max}^3\right)\operatorname{erf}\left(\frac{\alpha-r_{max}}{\Xi}\right) \nonumber \\
+      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \nonumber \\
+      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha r_{max}-\Xi^3-\Xi r_{max}^2\right)e^{-{\left(\frac{\alpha-r_{max}}{\Xi}\right)^2}} \nonumber \\
       &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}r_{max}^3 
@@
 \begin{align}
-  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}\left(\infty\right)^3\right)\left(-1\right) \\
-      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
-      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha\left(\infty\right)-\Xi^3-\Xi r_{max}^2\right)\left(0\right) \\
+  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}\left(\infty\right)^3\right)\left(-1\right) \nonumber \\
+      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \nonumber \\
+      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha\left(\infty\right)-\Xi^3-\Xi r_{max}^2\right)\left(0\right) \nonumber \\
       &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}\left(\infty\right)^3 
@@
 
-\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
+%KEEP%\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
+\begin{equation}
+   \frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\color{white}\right]\color{black} \nonumber
+\end{equation}
+\begin{equation}
+   \color{white}\left[-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\color{black}+\beta(E_{s(n)}-\mu{n})\right]\color{black}
+\end{equation}
 
@@
 The canonical equilibrium phase-space probability density is given by
-\begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}
-\begin{equation}{partian function}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
+\begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{\text{partian function}}\end{equation}
+\begin{equation}{\text{partian function}}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
 The Helmholtz free energy is given by 
@@
 \begin{align}
- TS &= T\left[-k_B\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(N!~h^{3N}P_{eq}\right) d\vec r^{N} d\vec p^{N}\right]   \\
-	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N}   \\
-	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]  \\
-    &= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]
-\end{align}  
-\begin{align}  
-    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}ln(Z)d\vec r^{N} d\vec p^{N}     \\
-    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}d\vec r^{N} d\vec p^{N}\right]   \\
+ TS &= T\left[-k_B\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(N!~h^{3N}P_{eq}\right) d\vec r^{N} d\vec p^{N}\right]  \nonumber \\
+	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N}  \nonumber \\
+	%KEEP%&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H-\ln\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right]\right) d\vec r^{N} d\vec p^{N}\right]
+    &= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)\color{white}\right]\color{black} \nonumber \\ 
+    &- \color{white}\left[\color{black} P_{eq}\ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) \color{black}d\vec r^{N} d\vec p^{N}\right] \nonumber \\
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}ln(Z)d\vec r^{N} d\vec p^{N}   \nonumber 
+\end{align}  
+\begin{align}   
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}d\vec r^{N} d\vec p^{N}\right]  \nonumber \\
     &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[1\right]   

### 2020-12-30 papers/thesis-kirstie/thesis.tex
edits

@@
 the repulsive potential, which is descriptive of hard spheres, dominates 
-the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
-fluid which is a perturbation of the hard-sphere fluid created by adding 
+the interactions between atoms.
+
+In more recent times, other model fluids have been developed.
+Zwanzig introduced the square-well 
+fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
 a hard-edge attractive potential to the repulsive potential of the 
@@
 with Rowlinson's idea of using temperature-dependent radii to model spheres 
-with soft repulsive potentials to create the Barker-Henderson model 
-fluid~\cite{barker1967perturbation}. In this model, a soft-sphere fluid is 
-modeled by using a hard-sphere fluid with hard spheres of larger radii 
-at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
+with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
+In this model, 
+%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
+hard spheres are given larger radii 
+at low temperatures, and smaller radii at high temperatures. 
+In 1971 Weeks, 
 Chandler, and Anderson used temperature and density dependent hard-sphere 
@@
 which is known as the Weeks-Chandler-Anderson model fluid. 
-
-ADD MORE HERE TO COME UP TO DATE!!
-\newline
+Models continue to be developed to characterize fluids with particular 
+properties, or to address particular problems like surface tension, 
+interfaces, adsorption, wetting or confinment.
+\textcolor{red}{A review of model fluids, and problems studied more 
+recently is given in ...}
 
 While models for particular fluids were being created, powerful methods 
-to describe the properties of fluids were also being developed. 
+to predict the properties of fluids were also being developed. 
 In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
@@
 obtained~\cite{Hansen}. 
-%\textcolor{red}{
-In 1958, Percus-Yevick introduced an alternative method 
-for the hard-sphere fluid based on an approximate closure to the 
+In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
+of the hard-sphere fluid based on an approximate closure to the 
 Ornstein-Zernike integral equation~\cite{PercusYevick}.
@@
 In 1961 Morita and Hiroike introduced a density functional 
-approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
-and the corresponding equilibrium number density profile for 
-inhomogeneous fluids. To use the classical Density Functional Theory (cDFT) 
-approach, an expression for the free energy 
-as a functional of an average number density that varies spatially 
-must be constructed for a particular fluid.
+approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
+%, and the corresponding equilibrium number density profile 
+of a classical inhomogeneous fluid. 
+Density functional theory was further developed for inhomogeneous classical fluids
+by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
+Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
+To use classical Density Functional Theory (cDFT), an expression for the free energy 
+as a functional of a spatially varying average single-particle number density 
+%that varies spatially 
+is constructed for a particular fluid.
 %DFT Evans 1979
 %Led by the consistency between results from SPT and Percus-Yevick, 
-In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
+\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
 to form a functional for the hard-sphere fluid based on the idea that 
-the lowest term in the virial expansion 
-of the excess chemical potential for a hard-sphere fluid can 
-be expressed in terms of a sum of weighted densities using a weight 
-function of the form of a backward Heavyside step function (shown below) 
-that can be expressed in terms of a sum of convolutions involving 
+%the lowest term in the virial expansion of the excess chemical potential 
+the lowest order term in the virial expansion of the excess chemical 
+potential of a hard-sphere fluid can 
+be expressed in terms of a weighted density using a weight function 
+%using a weight 
+%function 
+of the form of 
+a backward Heaviside step function}
+%\begin{equation}
+   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
+%\end{equation}
+that can itself be expressed in terms of a sum of convolutions involving 
 a set of weight functions related to the geometry of a sphere. 
+For a hard-sphere fluid with spheres of two different sized radii, 
+$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
 \begin{displaymath}
@@
 and for the direct correlation function approximation given by Percus-Yevick. 
-Tarazona later improved on the 
-method by introducing a tensor version of one of the energy density terms 
-in the functional~\cite{tarazonaphi3, 
-santos2012phi3}, and in 1999 Schmidt adopted FMT for spheres with soft 
+However, it did not predict freezing due to one term in the expression that
+diverged when the fluid was highly confined.  
+Tarazona later solved this problem by introducing a tensor version of 
+one of the energy density terms 
+in the functional~\cite{tarazonaphi3, santos2012phi3} 
+which took into account the non-spherical, and thus 
+non-symmetric, shape of the cavity in which the fluid is confined 
+while also yielding the correct excess free energy for the fluid as its 
+confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
+to a two-dimensional plane (as in a thin-film), to a one-dimensional 
+line (as in a fluid confined to a capillary tube), to a zero-dimensional 
+dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.
+
+In 1999 Schmidt adopted FMT for spheres with soft 
 interatomic potential energies
@@
 like the potential energy between hard spheres)
-forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
-
+forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
 %by replacing $w^{(3)}$ with a weight function that would result in  
@@
 %rather than instantaneously).
-
 Schmidt applied his Soft Fundamental Measure Theory to various fluids 
@@
 %, similar to that exhibited by a WCA fluid. 
+\textcolor{red}{For more recent developments in classical 
+Density Functional Theory see (2020), Evan's review~\cite{evansreview} (2016), 
+and Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013).}
+
+FREEZING in literature :
+\textcolor{red}{Classical density functional theory was used
+to ... (paper from 2010), but not until recently has there been good success 
+(other 23 papers...), and in all these the hard-sphere fluid has been 
+used as the reference system.} 
 In this paper, classical Density Functional 
@@
 
-Freezing is governed by entropy, which the universe %, or any closed system, 
+Freezing of simple fluids is governed by entropy, which the universe
 seeks to maximize according to the second law of Thermodynamics. When the 
@@
 equilibrium), and the same chemical potential (necessary for diffusive 
-equilibrium), an abrupt phase transition occurs, that is, the liquid 
-freezes. The particular structure and density that result are affected 
+equilibrium), a phase transition occurs, that is, the liquid 
+freezes. 
+Both the entropy and the volume (and thus the density) change abruptly 
+at the phase transition.
+The particular structure and density that result are affected 
 by the interatomic potential. When repulsive, soft sphere fluids 
@@
 
+\textcolor{blue}{
+Rosenfeld begins the development of FMT by considering the lowest order 
+term in the virial expansion of the excess chemical potential.
+Taking the partial derivative of the free energy in the low density limit given 
+by Equation~\ref{eq:Fexcess} gives
+\begin{align}
+   \mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
+            &= k_BT\int -f(r)~n~d\vec r
+\end{align}
+For a hard-sphere fluid (simplified in this account by considering only 
+equal sized spheres), the Mayer function, $f(r)$ can be written in terms 
+of a backward Heaviside step function
+\begin{align} 
+    f(r) = -\theta(|\vec r|-2R)
+\end{align}
+where and $\vec r$ is the distance between the centers of two spheres.
+The low-density excess free energy given by Equation~\ref{eq:Fexcess} 
+can then be written as a weighted density.
+\begin{align}
+     F_{excess}=k_BT\left(\int n^2(\vec r)\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
+\end{align} 
+where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
+Rosenfeld recognized that the Heaviside function describes mathmetically 
+the exclusion volume of a pair of spheres, one of the two variables 
+(the other one being the "overlap volume") in which the total 
+excess chemical potential can be expanded. He also noted that the 
+Heaviside function can uniquely be deconvolved into a set of weight functions
+related to the geometry of a sphere.
+\begin{align}
+    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+\end{align} 
+where again this expression is simplified in this account by considering only 
+equal sized spheres. 
+The excess free energy in the low density limit then takes the form
+\begin{align}
+     F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
+\end{align}
+%which can now be expressed in terms of a set of weighted densities
+\begin{align}
+     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
+\end{align}
+}
 %In order to apply classical Density Functional Theory to a system that has
@@
 \begin{align}
-   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
+   \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
    \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
@@
 \begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
-Using a backwards Heavyside step function defined as
+Using a backwards Heaviside step function defined as
 \begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
@@
 phase transition. 
-
 To construct the phase diagram, the liquid and solid phase transition number densities
 are found 
-
 %Maxwell's double tangent construction can be used to find 
 %these number densities given plots of 
-
 from the liquid and the solid Helmholtz 
@@
 
-\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 Calculate $\widetilde{w}_0(k)$ 
-\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{w}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_1(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_1(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_1(k)$ 
-\begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{w}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_2(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_2(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_2(k)$ 
-\begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 \begin{equation}{\widetilde{w}_2(k)=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\frac{2\sin(kr)}{kr}d{r}}\end{equation}
@@
 
-\begin{equation}{\widetilde{w}_3(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_3(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_3(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_3(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_3(k)$ 
-\begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}\vec{w}_1(\vec{r})e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}\vec{w}_1(\vec{r})e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated 
 coordinate system $(x',y'z')$ 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where
@@
 about $z'$. That leaves
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
+%\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate 
 system $(x',y'z')$ 
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where 
@@
 about $z'$. That leaves
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
 
@@
 The Fourier transform of the tensor weight matrix elements are given by
-\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
+\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
 Since the integration is over all space, the integral can be simplified 
@@
 transformed weight function in coordinate system (x,y,z). 
-\begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where $r=r'$ and
@@
 Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{y'}}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{y'}}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{z'}}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{z'}}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{y'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{z'}}{r^2}-\frac{\delta_{yz}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{z'}}{r^2}-\frac{\delta_{yz}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'x'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{x'}}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{x'}}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
 
 %\color{blue}
-\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
 %\color{black}
@@
 
-Set the lower limit to -infinity since the integrand is nearly zero by the time r is zero. 
+Set the lower limit to $-\infty$ as the integrand is nearly zero by the time r is zero. 
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{4\sqrt{\pi}}{3ka}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}\right)}\end{displaymath} 
@@
 \begin{equation}{+\frac{2\pi\sqrt{\pi}}{k^3a}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
-Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for 
-$\widetilde{w}_{{m2}_{xx}}(k)$ becomes
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath}
-\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+
+Putting in $a=\frac{\Xi}{\sqrt{2}}$, the final equation for $\widetilde{w}_{{m2}_{xx}}(k)$ becomes
+
+\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\nonumber\end{equation}
+\begin{equation}{+\frac{{(2\pi)}^{\frac{3}{2}}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2^\frac{3}{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2^\frac{3}{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 
 Calculate $\widetilde{w}_{{m2}_{y'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
 
@@
 Calculate $\widetilde{w}_{{m2}_{z'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}
-
-\begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}
+
+\begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
 
@@
 \section{Integrals used to Compute Weight Function Fourier Transforms}
-\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
-\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation}
-
-\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}+\frac{4\sin(kr)}{k^3r^3}}\end{equation}
+\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
+\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation}
+
+\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}-\frac{4\sin(kr)}{k^3r^3}}\end{equation}
 
@@
 \end{align}
-$\theta(|\vec r|-2R)$ is a backwards Heavyside step function 
+$\theta(|\vec r|-2R)$ is a backwards Heaviside step function 
 given by 

### 2021-01-06 papers/thesis-kirstie/thesis.tex
edits

@@
 interfaces, adsorption, wetting or confinment.
-\textcolor{red}{A review of model fluids, and problems studied more 
-recently is given in ...}
+\textcolor{red}{A review of some model fluids and problems that have been
+studied more recently is given by L\"{o}wen~\cite{lowen}.}
 
@@
 a set of weight functions related to the geometry of a sphere. 
-For a hard-sphere fluid with spheres of two different sized radii, 
+For two spheres with radii, 
 $R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
 \begin{displaymath}
-   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
+   \theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
 \begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
@@
 %, similar to that exhibited by a WCA fluid. 
-\textcolor{red}{For more recent developments in classical 
-Density Functional Theory see (2020), Evan's review~\cite{evansreview} (2016), 
-and Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013).}
-
-FREEZING in literature :
-\textcolor{red}{Classical density functional theory was used
-to ... (paper from 2010), but not until recently has there been good success 
-(other 23 papers...), and in all these the hard-sphere fluid has been 
-used as the reference system.} 
-In this paper, classical Density Functional 
-Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
-model fluid by applying SFMT with Schmidt's proposed error function model, 
-in which we have incorporated our temperature dependent parameters to fit the 
-functional to the soft-sphere WCA fluid.
+\textcolor{red}{For some more recent developments, see the review of new developments in 
+classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
+Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
+and a recent generalization of the Barker-Henderson model to inhomogeneous 
+fluids by Tschopp, Vuijk, Sharma and Brader~\cite{BHextended} (2020).}
+
+\textcolor{red}{Classical density functional theory has also been applied 
+to phase transitions, and freezing, but not until recently has there
+been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
+crystal as a highly inhomogeneous liquid~\cite{} (1979). Oettel and fellow 
+researchers studied hard-sphere crystals with ~\cite{} (2010). Mortazavifar and Oettel
+studied the fluid and crystal phases of the Asakura-Oosawa model fluid (2016). 
+Lutsko and Lam studied studied crystallization of the Leonard-Jones fluid~\cite{lutsko} (2018),
+and Lutsko and Schoonen looked at cDFT applied to the solid state~\cite{2020}.
+In all these studies the hard-sphere fluid has been used as the reference system,
+which is implemented with FMT, and to which other terms are added to create 
+the full functional for the fluid being studied.}
+In this paper, classical Density Functional Theory will be used to predict 
+the freezing of a Weeks-Chandler-Anderson model fluid directly by applying 
+SFMT with Schmidt's proposed error function model, 
+in which we have incorporated our temperature dependent parameters to 
+fit the functional to the soft-sphere WCA fluid. 
 %form the functional for the excess Helmholtz free energy. 
@@
 
-Freezing of simple fluids is governed by entropy, which the universe
-seeks to maximize according to the second law of Thermodynamics. When the 
-entropy of a pure substance in its liquid state is lower than the entropy 
-that it would have in its solid state at the same temperature (necessary 
-for thermal equilibrium), the same pressure (necessary for mechanical 
-equilibrium), and the same chemical potential (necessary for diffusive 
-equilibrium), a phase transition occurs, that is, the liquid 
-freezes. 
+%Freezing of simple fluids is governed by entropy, which the universe
+%seeks to maximize according to the second law of Thermodynamics. When the 
+%entropy of a pure substance in its liquid state is lower than the entropy 
+%that it would have in its solid state at the same temperature (necessary 
+%for thermal equilibrium with the environment), the same pressure (necessary for mechanical 
+%equilibrium with the environment), and the same chemical potential (necessary for diffusive 
+%equilibrium with the environment), a phase transition occurs, that is, the liquid 
+%freezes. 
+%%Along the phase boundary between pure liquid and pure solid states, both 
+%%liquid and solid states can coexist in equilibrium. 
+%Freezing of simple fluids can be seen when the entropy that a pure substance 
+%would have in a liquid state is lower than the entropy 
+%that it would have in its solid state at the same temperature (necessary 
+%for thermal equilibrium), the same pressure (necessary for mechanical 
+%equilibrium), and the same chemical potential (necessary for diffusive 
+%equilibrium).
+
+A liquid will eventually freeze as heat is removed, or the liquid is compressed. 
+The state of a simple fluid can be characterized by its entropy per particle~\cite{entropyfingerprint}.
+The entropy of a pure substance in a crystalline state is higher than the entropy 
+that it would have in its liquid state at a given temperature and pressure, 
+since there is more freedom for a particle to move between atoms  
+at lattice sites than in a dense liquid with irregular and blocked pathways~\cite{Hansen}.
 Both the entropy and the volume (and thus the density) change abruptly 
@@
 
-For pure substances, a phase transition occurs along a line on a 
+\textcolor{red}{A phase transition occurs along a line on a 
 Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
-diagram shown in Figure~\ref{fig:P-T_Diagram}. This phase diagram is 
+diagram shown in Figure~\ref{fig:P-T_Diagram}. 
+%
+Points along the line that separates the solid and the liquid phases 
+indicate a phase transition where the solid and liquid phases coexist, 
+and show that both the solid and the liquid have the same temperature (necessary 
+for thermal equilibrium)
+and the same pressure (necessary for mechanical 
+equilibrium)at a phase transition.
+%
+, and the same chemical potential (necessary for diffusive equilibrium). 
+%
+This phase diagram is 
 representative of some arbitrary, unknown substance composed of atoms, 
@@
 dividing line between the gas and liquid phases, there will just simply
-be a fluid, which we will refer to as a liquid in this paper.
+be a fluid, which we will refer to as a liquid in this paper.}
 %which may be referred to as a "liquid", since at the phase boundary the fluid is dense.
-Points along the line that separates the solid and the liquid phases 
-indicate a phase transition where the solid and liquid phases coexist, 
-and show that both the solid and the liquid have the same temperature 
-and the same pressure at a phase transition.
 
@@
 \textcolor{blue}{
-Rosenfeld begins the development of FMT by considering the lowest order 
-term in the virial expansion of the excess chemical potential.
-Taking the partial derivative of the free energy in the low density limit given 
-by Equation~\ref{eq:Fexcess} gives
-\begin{align}
-   \mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
-            &= k_BT\int -f(r)~n~d\vec r
-\end{align}
-For a hard-sphere fluid (simplified in this account by considering only 
-equal sized spheres), the Mayer function, $f(r)$ can be written in terms 
-of a backward Heaviside step function
-\begin{align} 
-    f(r) = -\theta(|\vec r|-2R)
-\end{align}
-where and $\vec r$ is the distance between the centers of two spheres.
-The low-density excess free energy given by Equation~\ref{eq:Fexcess} 
-can then be written as a weighted density.
-\begin{align}
-     F_{excess}=k_BT\left(\int n^2(\vec r)\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
-\end{align} 
-where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
-Rosenfeld recognized that the Heaviside function describes mathmetically 
-the exclusion volume of a pair of spheres, one of the two variables 
-(the other one being the "overlap volume") in which the total 
-excess chemical potential can be expanded. He also noted that the 
-Heaviside function can uniquely be deconvolved into a set of weight functions
-related to the geometry of a sphere.
+%Rosenfeld begins the development of FMT by considering the lowest order 
+%term in the virial expansion of the excess chemical potential.
+%Taking the partial derivative of the free energy in the low density limit given 
+%by Equation~\ref{eq:Fexcess} gives
+%\begin{align}
+   %\mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
+            %&= k_BT\int -f(r)~n~d\vec r
+%\end{align}
+Rosenfeld, who was studying mixtures, begins the development of FMT by considering that 
+for a hard-sphere fluid the Mayer function, $f(r)$, can be written in terms 
+of a backward Heaviside step function which, for a hard-sphere fluid 
+composed of a mixture of spheres of radius $R_i$ and $R_j$, is
+written as
+\begin{align} 
+    f(r) = -\theta(|\vec r_i - \vec r_j|-(R_i + R_j))
+\end{align}
+where and $|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
+%
+%The low-density excess free energy given by Equation~\ref{eq:Fexcess}
+%The lowest order term of the excess chemical potential  
+%can then be written as a weighted density.
+%\begin{align}
+     %\mu_{excess}=k_BT\int n\theta(|\vec r|-2R)~d\vec r
+      %%F_{excess}=k_BT\left(\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
+%\end{align} 
+%where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
+%The excess Helmholtz free energy given in Equation~\ref{eq:Fexcess} 
+%can then be written as
+%\begin{align}
+      %F_{excess} = k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r  
+%\end{align} 
+Rosenfeld recognized that the integral of the Heaviside step function
+is equal to the volume of a sphere with a radius equal to that of the radii 
+of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$
+this becomes
+\begin{align}
+       \int_{Volume} \theta(|\vec r_i- \vec r_j|-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \\
+       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + (4\pi R_j^2)R_i + \frac{4\pi}{3}R_j^3  \nonumber \\
+       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)
+\end{align}  
+where the volume, surface area, radius, and the number 1 can be written as
+\begin{align}
+        V &= \int_{Volume} \theta(r-R)d\vec r \\
+        S &= \int_{Volume} \delta(r-R) d\vec r \\
+        R &= \int_{Volume} \frac{\delta(r-R)}{4\pi r} d\vec r \\
+                1 &= \int_{Volume} \frac{\delta(r-R)}{4\pi r^2} d\vec r   
+\end{align}
+Limiting this discussion to a homogeneous hard-sphere fluid composed of hard spheres that have the same
+radius R, the lowest order term of the virial expansion of the excess Helmholtz free energy which is given in Equation~\ref{eq:Fexcess} 
+can now be written as
+\begin{align}
+      F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
+                 &= k_BTn^2\left(\int \frac{\theta(|\vec r|-2R)}{2}~d\vec r\right)V  \\
+                 &= k_BTn^2(2V(1) + 2SR)V  \\
+                 &= k_BT\left[\left(\int n\theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
+                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
+                 &= k_BT(n_3n_0 + n_2n_1)V \\
+                 &= k_BT(\Phi_1 + \Phi_2)V \\
+                 &= k_BT\Phi(\{n_i\})V
+\end{align} 
+where $n_i$ are weighted densities of the form
+\begin{align}
+     n_i=\int nw_i(r) d\vec r
+\end{align}
+with weight functions defined as
+\begin{align} 
+      w_3 &= \theta(r-R)  \label{scalor-weight3} \\
+      w_2 &= \delta(r-R)  \\
+      w_1 &= \frac{w_2}{4\pi r}  \\
+      w_0 &= \frac{w_2}{4\pi r^2} \label{scalor-weight0}
+\end{align}
+and dimensionless energy densities $\Phi_1$ and $\Phi_2$ are defined as
+\begin{align}
+      \Phi_1 = n_3n_0  \\
+      \Phi_2 = n_1n_2
+\end{align}
+The excess Helmholtz free energy can also be expressed in terms of the
+packing fraction $\eta=nV$ as
+\begin{align}
+     F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
+                &= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
+                &= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
+                &= k_BT4\eta^2
+\end{align} 
+%
+For the inhomogeneous case where the number density varies spatially,
+$n\rightarrow n(\vec r)$, Equation~\ref{eq:Fexcess} becomes
+\begin{align}
+     F_{excess} = k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}
+\end{align}
+where $\vec r_a$ is a position vector of one sphere, $\vec r_b$ is a position vector to another sphere, and 
+$|\vec r_a - \vec r_b|$ is the distance between the centers of the two spheres.
+Rosenfeld recognized that the Heaviside function can be written as 
+a convolution of the weight functions, 
 \begin{align}
@@
 \end{align} 
-where again this expression is simplified in this account by considering only 
-equal sized spheres. 
-The excess free energy in the low density limit then takes the form
-\begin{align}
-     F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
-\end{align}
-%which can now be expressed in terms of a set of weighted densities
-\begin{align}
-     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
-\end{align}
+where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
+and the two vector weights are given by
+\begin{align}
+      \vec {w}_2 &= w_2~\hat r \\
+      \vec {w}_1 &= w_1~\hat r
+\end{align}
+which are needed for the relation to be satisfied (although other expressions 
+have been developed without using vector weights). 
+%a uniquely deconvolved 
+%into a set of weight functions related to the geometry of a sphere, and that it
+%describes mathmetically 
+%the exclusion volume of a pair of spheres, one of the two variables 
+%(the other one being the "overlap volume") in which the total 
+%excess chemical potential can be expanded. 
+%
+He then noted the following relation
+\begin{align}
+     w_\alpha(\vec r_a - \vec r_b) \otimes w_\gamma(\vec r_a - \vec r_b) &= \int w_\alpha(\vec r_a - \vec x) w_\gamma(\vec r_b - \vec x) d\vec x \\
+                                                                         &= \int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
+\end{align}
+which simplifies the equation for the excess Helmholtz free energy
+\begin{align}
+     F_{excess} &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}  \\
+                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~(w_0\otimes w_3+ w_1 \otimes w_2+\vec {w}_1\otimes \vec {w}_2)  \\
+                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)\left(\int w_3(\vec x - \vec r_a)w_0(\vec x - \vec r_b) \vec dx \color{white}\right)  \nonumber \\
+                &+ \color{white}\left(\color{black}\int w_1(\vec x - \vec r_a)w_2(\vec x - \vec r_b) d\vec x  + \int \vec w_1(\vec x - \vec r_a)\cdot\vec w_2(\vec x - \vec r_b) d\vec x\right)    \\
+                &= k_BT\int \left[\int n(\vec r_a) w_3(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_0(\vec x - \vec r_b)~d\vec r_b \color{white}\right]  \nonumber \\
+                &+ \int n(\vec r_a) w_2(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_1(\vec x - \vec r_b)~d\vec r_b  \nonumber \\
+                &+ \color{white}\left[\color{black} \int n(\vec r_a) \vec {w}_2(\vec x - \vec r_a)~d\vec r_a \cdot \int n(\vec r_b) \vec {w}_1(\vec x - \vec r_b)~d\vec r_b \right]\color{black}d\vec x  \\
+                &= k_BT\int n_3n_0 + n_2n_1 + \vec n_1\cdot\vec n_2~d\vec x  \\
+                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec x  \\
+                & =k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r               
+\end{align}
+where $\vec n_1\cdot\vec n_2$=0 for the homogeneous case, 
+and the weighted densities $n_i$ are defined as
+\begin{align}
+    n_i(\vec r)=\int n(\vec r')w_i(\vec r-\vec r')~d\vec r'
+\end{align} 
+%The excess free energy in the low density limit then takes the form
+%\begin{align}
+%    F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
+%\end{align}
+%%which can now be expressed in terms of a set of weighted densities
+%\begin{align}
+%     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
+%\end{align}
+%For a homogeneous fluid in the low density limit, the excess Helmholtz free
+%energy can then be expressed as
+%\begin{align}
+      %F_{excess}=k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r
+%\end{align} 
+%
+%Rosenfeld arrived at an expression for the excess Helmholtz free energy 
+%based on a set of weighted densities
+%\begin{align}
+    %F_{excess}=k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r
+%\end{align} 
+%
+%The resulting expression for the excess Helmholtz free energy which is
+%applicable to the more general case of an inhomogeneous fluid, reduces 
+%to the same result obtained from Scaled Particle Theory for a homogeneous 
+%hard-sphere fluid. 
+}\textcolor{red}{
+Lifting the restriction to the low density limit,
+The expression for the excess Helmholtz free energy derived from FMT
+is applicable to the more general case of an inhomogeneous fluid, but reduces 
+to the same result obtained from Scaled Particle Theory for a homogeneous 
+hard-sphere fluid. 
 }
+
 %In order to apply classical Density Functional Theory to a system that has
@@
 %to inhomogeneous, or non-uniform hard-sphere fluids.
-
+%
 Scaled Particle Theory

### 2021-01-09 papers/thesis-kirstie/thesis.tex
edits

@@
 \begin{align} 
-    f(r) = -\theta(|\vec r_i - \vec r_j|-(R_i + R_j))
-\end{align}
-where and $|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
+    f(r) = -\Theta(r-(R_i + R_j))
+\end{align}
+where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
 %
@@
 is equal to the volume of a sphere with a radius equal to that of the radii 
-of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$
+of two hard spheres added together. For a mixture of hard spheres of radius $R_a$ and $R_b$
 this becomes
 \begin{align}
-       \int_{Volume} \theta(|\vec r_i- \vec r_j|-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \\
-       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + (4\pi R_j^2)R_i + \frac{4\pi}{3}R_j^3  \nonumber \\
+       \int_{Volume} \Theta(r-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\
+       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + R_i(4\pi R_j^2) + \frac{4\pi}{3}R_j^3  \nonumber \\
        &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)
@@
 \begin{align}
-        V &= \int_{Volume} \theta(r-R)d\vec r \\
+        V &= \int_{Volume} \Theta(r-R)d\vec r \\
         S &= \int_{Volume} \delta(r-R) d\vec r \\
@@
       F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
-                 &= k_BTn^2\left(\int \frac{\theta(|\vec r|-2R)}{2}~d\vec r\right)V  \\
+                 &= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\
                  &= k_BTn^2(2V(1) + 2SR)V  \\
-                 &= k_BT\left[\left(\int n\theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
-                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
-                 &= k_BT(n_3n_0 + n_2n_1)V \\
+                 &= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
+                 &+ \color{white}\left[\color{blue}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{blue}V   \\          
+                 &= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
                  &= k_BT(\Phi_1 + \Phi_2)V \\
-                 &= k_BT\Phi(\{n_i\})V
+                 &= k_BT\Phi(\{n_\alpha\})V
+\end{align} 
+The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
+with a dimensionless energy composed 
+of the sum of $\Phi_1$ and $\Phi_2$ which are defined as
+\begin{align}
+      \Phi_1 = n_3n_0  \\
+      \Phi_2 = n_1n_2
 \end{align}
@@
 \begin{align}
-     n_i=\int nw_i(r) d\vec r
+     n_\alpha=\int nw_\alpha(r) d\vec r
 \end{align}
@@
 \begin{align} 
-      w_3 &= \theta(r-R)  \label{scalor-weight3} \\
+      w_3 &= \Theta(r-R)  \label{scalor-weight3} \\
       w_2 &= \delta(r-R)  \\
@@
 \end{align}
-and dimensionless energy densities $\Phi_1$ and $\Phi_2$ are defined as
-\begin{align}
-      \Phi_1 = n_3n_0  \\
-      \Phi_2 = n_1n_2
-\end{align}
-The excess Helmholtz free energy can also be expressed in terms of the
-packing fraction $\eta=nV$ as
-\begin{align}
-     F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
-                &= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
-                &= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
-                &= k_BT4\eta^2
-\end{align} 
+%The excess Helmholtz free energy can also be expressed in terms of the
+%packing fraction $\eta=nV$ as
+%\begin{align}
+     %F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
+                %&= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
+                %&= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
+                %&= k_BT4\eta^2
+%\end{align} 
 %
@@
 \begin{align}
-     F_{excess} = k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}
-\end{align}
-where $\vec r_a$ is a position vector of one sphere, $\vec r_b$ is a position vector to another sphere, and 
-$|\vec r_a - \vec r_b|$ is the distance between the centers of the two spheres.
+     F_{excess} = k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
+\end{align}
+where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
+position vector to another sphere, and $|\vec r_i - \vec r_j|$ 
+is the distance between the centers of the two spheres.
 Rosenfeld recognized that the Heaviside function can be written as 
-a convolution of the weight functions, 
-\begin{align}
-    f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
-\end{align} 
-where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
-and the two vector weights are given by
-\begin{align}
-      \vec {w}_2 &= w_2~\hat r \\
-      \vec {w}_1 &= w_1~\hat r
-\end{align}
-which are needed for the relation to be satisfied (although other expressions 
-have been developed without using vector weights). 
+a convolution of the weight functions which, for a fluid composed of 
+spheres of radius R, is given by
+\begin{align}
+    f(|\vec r_i-\vec r_j|)=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+\end{align} 
+%where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
+where the two vector weights are given by
+\begin{align}
+      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
+      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
+\end{align}
+which are added for the relation to be satisfied, 
+and the other weights take the form 
+\begin{align} 
+      w_3 &= \Theta(|\vec r_i-\vec r_j|-R)  \label{scalor-weight3} \\
+      w_2 &= \delta(|\vec r_i-\vec r_j|-R)  \\
+      w_1 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|}  \\
+      w_0 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|^2} \label{scalor-weight0}
+\end{align}
 %a uniquely deconvolved 
@@
 %
-He then noted the following relation
-\begin{align}
-     w_\alpha(\vec r_a - \vec r_b) \otimes w_\gamma(\vec r_a - \vec r_b) &= \int w_\alpha(\vec r_a - \vec x) w_\gamma(\vec r_b - \vec x) d\vec x \\
-                                                                         &= \int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
-\end{align}
-which simplifies the equation for the excess Helmholtz free energy
-\begin{align}
-     F_{excess} &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~\frac{\theta(|\vec r_a-\vec r_b|-2R)}{2}  \\
-                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)~(w_0\otimes w_3+ w_1 \otimes w_2+\vec {w}_1\otimes \vec {w}_2)  \\
-                &= k_BT\int d\vec r_a \int d\vec r_b~n(\vec r_a)~n(\vec r_b)\left(\int w_3(\vec x - \vec r_a)w_0(\vec x - \vec r_b) \vec dx \color{white}\right)  \nonumber \\
-                &+ \color{white}\left(\color{black}\int w_1(\vec x - \vec r_a)w_2(\vec x - \vec r_b) d\vec x  + \int \vec w_1(\vec x - \vec r_a)\cdot\vec w_2(\vec x - \vec r_b) d\vec x\right)    \\
-                &= k_BT\int \left[\int n(\vec r_a) w_3(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_0(\vec x - \vec r_b)~d\vec r_b \color{white}\right]  \nonumber \\
-                &+ \int n(\vec r_a) w_2(\vec x - \vec r_a)~d\vec r_a \int n(\vec r_b) w_1(\vec x - \vec r_b)~d\vec r_b  \nonumber \\
-                &+ \color{white}\left[\color{black} \int n(\vec r_a) \vec {w}_2(\vec x - \vec r_a)~d\vec r_a \cdot \int n(\vec r_b) \vec {w}_1(\vec x - \vec r_b)~d\vec r_b \right]\color{black}d\vec x  \\
-                &= k_BT\int n_3n_0 + n_2n_1 + \vec n_1\cdot\vec n_2~d\vec x  \\
-                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec x  \\
-                & =k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r               
-\end{align}
-where $\vec n_1\cdot\vec n_2$=0 for the homogeneous case, 
-and the weighted densities $n_i$ are defined as
-\begin{align}
-    n_i(\vec r)=\int n(\vec r')w_i(\vec r-\vec r')~d\vec r'
-\end{align} 
+The convolution of the weight functions can be written as
+\begin{align}
+     w_\alpha\otimes w_\gamma(\vec r_i - \vec r_j) &= \int w_\alpha(\vec r') w_\gamma(\vec r_i -\vec r_j - \vec r') d\vec r'   \\                                                             
+                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i -\vec r_j -(\vec u - \vec r_j))d\vec u     \\
+                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i - \vec u) d\vec u  \\  
+                                                   %&= \int w_\gamma(\vec r_a - \vec u)w_\alpha(\vec u - \vec r_b) d\vec u  \\                                                              
+                                                   &= \int w_\gamma(-(\vec u - \vec r_i))w_\alpha(\vec u - \vec r_j))  d\vec u  
+\end{align}
+For the scalor weights this is the same as
+\begin{align} 
+                                                   &= \int w_\gamma(\vec u - \vec r_i)w_\alpha(\vec u - \vec r_j)  d\vec u  
+\end{align}
+and for the vector weights this becomes
+\begin{align} 
+                                                   &= -\int \vec {w}_\gamma(\vec u - \vec r_i)\cdot\vec {w}_\alpha(\vec u - \vec r_j)  d\vec u  
+\end{align}
+%
+%which can also be written as
+%\begin{align}
+     %\color{white}w_\alpha\otimes w_\gamma(\vec r_a - \vec r_b)\color{blue} &=\int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
+%\end{align}
+%since the weight functions involve the distances $|\vec r_a - \vec x|=|\vec x - \vec r_a|$.
+The lowest order term of the excess Helmholtz free energy can then be written as
+\begin{align}
+     F_{excess} &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  \\
+                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}  \\
+                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
+                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
+                &+ \color{white}\left(\color{blue}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec x ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
+                &= k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
+                &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
+                &+ \color{white}\left[\color{blue} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{blue}d\vec u  \\
+                &= k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\
+                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
+                &= k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
+                &\text{or replacing u notation with r} \nonumber \\
+                &= k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r             
+\end{align}
+where the weighted densities $n_\alpha$ have the form
+\begin{align}
+    n_\alpha(\vec r)=\int n(\vec r')w_\alpha(\vec r-\vec r')~d\vec r'  
+    %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
+\end{align}  
+For the homogenesous case $\vec n_1\cdot\vec n_2$=0, and Equation~\ref{Fexlowden-inhomo} reduces to Equation~\ref{Fexlowden-homo}.
 %The excess free energy in the low density limit then takes the form
@@
 %hard-sphere fluid. 
-}\textcolor{red}{
-Lifting the restriction to the low density limit,
-The expression for the excess Helmholtz free energy derived from FMT
-is applicable to the more general case of an inhomogeneous fluid, but reduces 
-to the same result obtained from Scaled Particle Theory for a homogeneous 
-hard-sphere fluid. 
+%
+}\textcolor{blue}{
+Lifting the restriction to the low density limit, Rosenfeld derived an
+expression for the excess Helmholtz free energy by noting that the higher
+order terms in the virial expansion can all be written in terms of an 
+"exclusion volume" which is given by Equation~\ref{exclusion-volume}, 
+and an "overlap volume".
+The exclusion volume is the spherical volume about the center of one sphere 
+which the center of another sphere can not enter, and is used to describe
+terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. The overlap volume is 
+the volume of intersection between two spheres that overlap, and is used in conjunction 
+with the exclusion volume to describe higher order terms in the expansion~\cite{rosenfeld1988}. 
+%
+The resulting expression for the excess Helmholtz free energy is given in 
+Equation~\ref{FexfromSPT} with $\Phi$'s defined in Equations~\ref{Phi1equation} 
+through~\ref{oldPhi3equation}, and
+and weighted densities defined in Equation~\ref{weighted-density}, 
+and reduces to the same result obtained from Scaled Particle Theory for a 
+homogeneous hard-sphere fluid. 
 }
-
+%
 %In order to apply classical Density Functional Theory to a system that has
@@
 %to inhomogeneous, or non-uniform hard-sphere fluids.
-%
+
 Scaled Particle Theory
@@
 \begin{align}
-   \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
+   \Phi_1 &= -n_{0}\ln(1-n_{3}) \label{Phi1equation} \\
    \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
@@
 or for a spatially varying number density $n(\vec{r})$,
-\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
+\begin{equation}\label{weighted-density}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
 with weight functions
@@
 \begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
-All the other FMT equations remain the same.
+%All the other FMT equations remain the same.
 
@@
 \end{document}
+
+%my notes
+Definition of a convolution:
+\begin{align}
+    w_\alpha \otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r')w_\gamma(\vec r-\vec r')d\vec r'
+\end{align}
+
+\begin{align}
+    w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r')w_\gamma(\vec r'-\vec r)d\vec r
+\end{align}
+
+Set $\vec r' = \vec r_i-\vec r$
+\begin{align}
+     w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-\vec r-\vec r)d\vec r
+\end{align}
+
+Put in $\color{blue}\vec r=\vec r_i-\vec r_j\color{black}$
+\begin{align}
+     w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-(\color{blue}\vec r_i-\vec r_j\color{black})-r)d\vec r  \\
+                                    &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-\vec r_i+\vec r_j-\vec r)d\vec r \\
+                                    &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_j-\vec r)d\vec r
+\end{align}

### 2021-01-10 papers/thesis-kirstie/thesis.tex
edits

@@
 
-\textcolor{blue}{
+\textcolor{red}{FROM HERE}
+%\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
 %Rosenfeld begins the development of FMT by considering the lowest order 
@@
 composed of a mixture of spheres of radius $R_i$ and $R_j$, is
-written as
+given by
 \begin{align} 
@@
 where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
+%This can easily be shown by putting the interatomic potential between two hard spheres
+%given by 
+%\begin{align}
+    %V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+%\end{align}
+%into the equation for the Mayer function
+%\begin{align}
+     %f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+%\end{align} 
 %
@@
                  &= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
-                 &+ \color{white}\left[\color{blue}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{blue}V   \\          
+                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
                  &= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
@@
 \end{align}
-which are added for the relation to be satisfied, 
+which are needed for the relation to be satisfied, 
 and the other weights take the form 
@@
                 &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
-                &+ \color{white}\left(\color{blue}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec x ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
+                &+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec x ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
                 &= k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
                 &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
-                &+ \color{white}\left[\color{blue} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{blue}d\vec u  \\
+                &+ \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u  \\
                 &= k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\
@@
                 &= k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
-                &\text{or replacing u notation with r} \nonumber \\
-                &= k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r             
+                %&\text{or replacing u notation with r} \nonumber \\
+                &= k_BT\int \Phi(n(\vec r))~d\vec r             
 \end{align}
@@
 \begin{align}
-    n_\alpha(\vec r)=\int n(\vec r')w_\alpha(\vec r-\vec r')~d\vec r'  
+    n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
     %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
@@
 %
-}\textcolor{blue}{
+
 Lifting the restriction to the low density limit, Rosenfeld derived an
@@
 %
-The resulting expression for the excess Helmholtz free energy is given in 
-Equation~\ref{FexfromSPT} with $\Phi$'s defined in Equations~\ref{Phi1equation} 
-through~\ref{oldPhi3equation}, and
-and weighted densities defined in Equation~\ref{weighted-density}, 
-and reduces to the same result obtained from Scaled Particle Theory for a 
-homogeneous hard-sphere fluid. 
-}
+The resulting expression for the excess Helmholtz free energy with an 
+updated version of $\Phi_3$ that was later constructed by Tarazona is given by
+\begin{align}
+    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
+    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
+    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
+    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
+    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
+    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
+    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
+\end{align}
+where in addition to the weight functions already defined, a tensor weight function is given by
+\begin{align}\label{eq:tensorweight-methods}
+     \overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
+\end{align}
+The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
+all depend on the spatially varying number density profile $n(\vec r)$.
+\begin{align}%\label{eq:numdenprofile}
+	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
+\end{align}
+This result reduces to same result obtained from Scaled Particle Theory 
+for a homogeneous hard-sphere fluid.
+
+The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
+All of the weight functions can be expressed in terms of 
+$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
+The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+the relation
+\begin{align}\label{w2_w3_relation}
+    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+\end{align}
+%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
+Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
+another important relation expressed in terms of $w_{2}(r)$ 
+%that will be important 
+%to keep in tact when moving to SFMT~\cite{schmidt2000fluid}.
+\begin{align}\label{mayerder=w2convolution}
+     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
+\end{align} 
+Soft Fundamental Measure Theory extends FMT to fluids 
+with soft potentials by keeping the relationships between the weight 
+functions, and the equation for the derivative of the Mayer function 
+the same, 
+but replacing $w_{3}(r)$ with a weight function appropriate 
+for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+This is what we do in forming a functional for the Weeks-Chandler-Anderson 
+fluid as detailed in the next chapter on Methods.
+
+%\textcolor{red}{TO HERE ----------------------------------}
 %
@@
 
-Scaled Particle Theory
-seeks a general expression for the excess chemical 
-potential of a homogeneous hard-sphere fluid which can then be integrated 
-to find the excess Helmholtz free energy according to the relation
-\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
-It is an extension of the work done by Widom who showed that the excess 
-chemical potential of a homogeneous fluid corresponds to the reversible 
-work of inserting a sphere of radius R into a fluid consisting of hard 
-spheres of radius R. 
-\begin{align}
-	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
-\end{align}
-Widom's result was obtained by considering the difference 
-in the excess Helmholtz free energy when one atom is added to an otherwise 
-closed system. He showed that the excess chemical potential is related to 
-the ensemble average of the Boltzman factor for the potential energy 
-difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
-The resulting equation, shown below, is known as Widom's Insertion 
-Formula~\cite{Hansen}.
-\begin{align} \label{mu-der-int}
-    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
-\end{align}
-\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
-where 
-\begin{align}
-    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
-\end{align}
-and
-\begin{align}
-    Z=\sum_s{e^{-\beta V_s}}
-\end{align}
-
-The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
-there will be a change in the potential energy of $V_{diff}$ if one sphere
-is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
-In this case, $P_s=\frac{e^{-\beta V_s}}{Z}$.
-%at each point in the fluid, there is a probability that 
-%a hard sphere of radius R has space available to exist at that point
-%without overlapping with other spheres, or it does not, as shown in 
-Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, 
-then becomes~\cite{Hansen}
-\begin{align} 
-	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
-	         &= V_{diff} \\
-             &= W_{rev}(R)  \label{workrev}
-\end{align}
-\begin{figure}
-    \centering
-    \includegraphics[height=7cm]{figs/P_overlap.pdf}
-    \caption{No energy is needed to place a sphere at a position where 
-    there is no overlap, such as that shown by the black dot near the 
-    top. But it is not possible to place a sphere at a position where 
-    an overlap occurs.}
-    \label{fig:p_overlap}
-  \end{figure}
-The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
-that it takes no energy for a hard sphere to be placed at a point
-in the fluid where it does not overlap with any other spheres 
-($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
-to place a sphere where it would overlap another sphere, as described 
-in Figure~\ref{fig:p_overlap}. 
-In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
-\begin{align} \label{muexcess}
-  \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
-\end{align}
-where the radius of the cavity $R_{cavity}$ equals the radius of a 
-hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
-occupied by the spheres divided by the total volume available, 
-is the probability that a sphere cannot be placed at some random
-point in the homogeneous fluid of number density, $n$, without
-overlapping with another sphere. $P_s$ would then have values of
-$nV_{Sphere}$ and $1-nV_{Sphere}$.
-Combining  Equations~\ref{workrev} 
-and \ref{muexcess} gives
-\begin{align}
-  %W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3)
-  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-\eta) 
-\end{align}
-where $\eta$ is the packing fraction defined as
-\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
-\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
-
-Scaled Particle Theory forms a more general result for the reversible work 
-as a function of the radius of the cavity by considering additional 
-cases where the cavity is larger, and much larger than the radius of 
-a hard sphere, and requiring continuity of the functions for the 
-different cases and their derivatives (first and second). 
-By setting the radius of the cavity to the hard sphere radius, the 
-more general work equation becomes equal to the chemical potential. 
-Details of this procedure can be found 
-in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.  
-An expression for the excess Helmholtz free energy in terms of the
-packing fraction $\eta$ can then be obtained through 
-Equation~\ref{mu-der-int}. 
-%
-%By requiring continuity of the functions valid for different cavity 
-%sizes, along with their first and second derivatives, an equation for 
-%the work is obtained which can then be used to form 
-%set equal to the excess chemical 
-%potential by setting the cavity radius to the radius of hard sphere. 
-%
-%Forming a more general result by requiring continuity of the functions, 
-%along with their first and second derivatives, for the reversible work
-%written for fluids with a cavity radius equal to, larger than, or very 
-%much larger than the hard sphere radius R, 
-%
-%Scaled Particle Theory arrives at 
-%an expression for the excess Helmholtz free energy of 
-%a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
-%For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
-\begin{align}\label{Fexcess-SPT}
-    F_{excess}=k_BT\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)N
-\end{align}
-
-Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
-%and provides an expression for the excess Helmholtz free energy of an 
-%inhomogeneous 
-%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
-%
-%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
-%is an extension of FMT and provides an expression for the excess Helmholtz
-%free energy of fluids with soft potentials like that of the WCA fluid. 
-%It will be helpful to continue looking at FMT before addressing SFMT.
-%FMT begins 
-by noticing that a quantity $G_{3}$ can be defined such that
-\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
-where $G_{3}$ is the volume of a sphere. 
-Likewise, by defining quantities
-\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
-\begin{equation}{G_{1}=R}\end{equation}
-\begin{equation}{G_{0}=1}\end{equation}
-where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
-radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
-\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
-The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
-geometric measures of a sphere, and are what give rise to the name
-Fundamental Measure Theory. Introducing a set of variables
-%weighted densities $n_{i}=nG_{i}$ 
-given by
-\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
-\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
-\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
-\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
-Equation~\ref{FexfromSPTinthermsofG} can be written as
-\begin{align}\label{FexfromSPT}
-    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
-\end{align}
-which has the form 
-\begin{align}
-    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
-\end{align}
+%\textcolor{red}{BEGIN COMMENTED OUT SECTION...} %########################
+
+%Scaled Particle Theory
+%seeks a general expression for the excess chemical 
+%potential of a homogeneous hard-sphere fluid which can then be integrated 
+%to find the excess Helmholtz free energy according to the relation
+%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+%It is an extension of the work done by Widom who showed that the excess 
+%chemical potential of a homogeneous fluid corresponds to the reversible 
+%work of inserting a sphere of radius R into a fluid consisting of hard 
+%spheres of radius R. 
+%\begin{align}
+	%\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
+%\end{align}
+%Widom's result was obtained by considering the difference 
+%in the excess Helmholtz free energy when one atom is added to an otherwise 
+%closed system. He showed that the excess chemical potential is related to 
+%the ensemble average of the Boltzman factor for the potential energy 
+%difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
+%The resulting equation, shown below, is known as Widom's Insertion 
+%Formula~\cite{Hansen}.
+%\begin{align} \label{mu-der-int}
+    %\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
+%\end{align}
+%\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
 %where 
-\begin{align}
-   \Phi_1 &= -n_{0}\ln(1-n_{3}) \label{Phi1equation} \\
-   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
-   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
-\end{align}
-
-The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
-densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
-\begin{align}
-    n_{i}=nG_{i}
-\end{align}
-These weights are valid for determining the excess free energy of a system of 
-hard spheres with a homogeneous number density n. But these weights, and 
-their related  weighted densities, can be written in a more general form 
-allowing for their extension to a system with a spatially varying number 
-density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
-excess Helmholtz free energy is written as
-\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
-Using a backwards Heaviside step function defined as
-\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
-the following expressions are obtained
-\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
-\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
-\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
-\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
-where the weighted densities can now be expressed as
-\begin{align}
-    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
-\end{align} 
-or for a spatially varying number density $n(\vec{r})$,
-\begin{equation}\label{weighted-density}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-with weight functions
-\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
-\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
-\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
-\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
-The weight functions $w_{2}$ and $w_{3}$ have the relation
-\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
-which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-
-The weight functions can be related to the Mayer function 
-\begin{align}
-     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
-\end{align} 
-where r is the distance between the centers of two hard spheres, and 
-$V(r)$ is the interatomic potential between two hard spheres given by 
-\begin{align}
-    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
-\end{align}
-Evaluating Equation with $V(r)$ given by Equation,  
-the Mayer function for a hard-sphere fluid is found to be
-\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
-which can be written in terms of the weight functions if two 
-additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
-\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
-\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
-The Mayer function can then be written in terms of the weight functions~\cite{Hansen}
-\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
-
-Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
-another relation that will be important to keep in tact when moving to 
-SFMT~\cite{schmidt2000fluid}.
-\begin{align}\label{mayerder=w2convolution}
-     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
-\end{align} 
-This relation can easily be checked for a hard-sphere fluid where
-\begin{equation}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{equation}
-\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
-The right side can be evaluated by using the general relation for 
-convolution which is equal to the Inverse Fourier Transform 
-of a product of Fourier Transforms. 
-\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
-Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
-gives
-\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
-and the right side of Equation~\ref{mayerder=w2convolution} becomes
-\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
-\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
-which matches Equation~\ref{mayerder}.
-
-Tarazona later improved upon the original FMT equations by replacing the
-expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
-more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
-\begin{align}
-\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
-\end{align}
-in which a tensor weight is incorporated given by
-\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
-%All the other FMT equations remain the same.
-
-The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
-All of the weight functions, the relations between them, and the equation 
-for the derivative of the Mayer function, can be expressed in terms of 
-$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
-The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
-Equation~\ref{w2_w3_relation}.
-Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
-with soft potentials by keeping the relationships between the weight 
-functions, and the equation for the derivative of the Mayer function 
-the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
-~\ref{w2_w3_relation},~\ref{eq:tensorweight},~\ref{mayerder=w2convolution}), 
-but replacing $w_{3}(r)$ with a weight function appropriate 
-for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
-This is what we do in forming a functional for the Weeks-Chandler-Anderson 
-fluid as detailed in the next chapter on Methods.
+%\begin{align}
+    %<e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
+%\end{align}
+%and
+%\begin{align}
+    %Z=\sum_s{e^{-\beta V_s}}
+%\end{align}
+
+%The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
+%there will be a change in the potential energy of $V_{diff}$ if one sphere
+%is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
+%In this case, $P_s=\frac{e^{-\beta V_s}}{Z}$.
+%%at each point in the fluid, there is a probability that 
+%%a hard sphere of radius R has space available to exist at that point
+%%without overlapping with other spheres, or it does not, as shown in 
+%Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, 
+%then becomes~\cite{Hansen}
+%\begin{align} 
+	%\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
+	         %&= V_{diff} \\
+             %&= W_{rev}(R)  \label{workrev}
+%\end{align}
+%\begin{figure}
+    %\centering
+    %\includegraphics[height=7cm]{figs/P_overlap.pdf}
+    %\caption{No energy is needed to place a sphere at a position where 
+    %there is no overlap, such as that shown by the black dot near the 
+    %top. But it is not possible to place a sphere at a position where 
+    %an overlap occurs.}
+    %\label{fig:p_overlap}
+  %\end{figure}
+%The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
+%that it takes no energy for a hard sphere to be placed at a point
+%in the fluid where it does not overlap with any other spheres 
+%($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
+%to place a sphere where it would overlap another sphere, as described 
+%in Figure~\ref{fig:p_overlap}. 
+%In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
+%\begin{align} \label{muexcess}
+  %\mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
+%\end{align}
+%where the radius of the cavity $R_{cavity}$ equals the radius of a 
+%hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
+%occupied by the spheres divided by the total volume available, 
+%is the probability that a sphere cannot be placed at some random
+%point in the homogeneous fluid of number density, $n$, without
+%overlapping with another sphere. $P_s$ would then have values of
+%$nV_{Sphere}$ and $1-nV_{Sphere}$.
+%Combining  Equations~\ref{workrev} 
+%and \ref{muexcess} gives
+%\begin{align}
+  %%W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3)
+  %W_{rev}(R_{cavity}=R)=-k_BT\ln(1-\eta) 
+%\end{align}
+%where $\eta$ is the packing fraction defined as
+%\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
+%\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
+
+%Scaled Particle Theory forms a more general result for the reversible work 
+%as a function of the radius of the cavity by considering additional 
+%cases where the cavity is larger, and much larger than the radius of 
+%a hard sphere, and requiring continuity of the functions for the 
+%different cases and their derivatives (first and second). 
+%By setting the radius of the cavity to the hard sphere radius, the 
+%more general work equation becomes equal to the chemical potential. 
+%Details of this procedure can be found 
+%in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.  
+%An expression for the excess Helmholtz free energy in terms of the
+%packing fraction $\eta$ can then be obtained through 
+%Equation~\ref{mu-der-int}.
+
+%%By requiring continuity of the functions valid for different cavity 
+%%sizes, along with their first and second derivatives, an equation for 
+%%the work is obtained which can then be used to form 
+%%set equal to the excess chemical 
+%%potential by setting the cavity radius to the radius of hard sphere. 
+%%
+%%Forming a more general result by requiring continuity of the functions, 
+%%along with their first and second derivatives, for the reversible work
+%%written for fluids with a cavity radius equal to, larger than, or very 
+%%much larger than the hard sphere radius R, 
+%%
+%%Scaled Particle Theory arrives at 
+%%an expression for the excess Helmholtz free energy of 
+%%a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
+%%For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
+%\begin{align}\label{Fexcess-SPT}
+    %F_{excess}=k_BT\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)N
+%\end{align}
+
+%Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
+%%and provides an expression for the excess Helmholtz free energy of an 
+%%inhomogeneous 
+%%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
+%%
+%%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
+%%is an extension of FMT and provides an expression for the excess Helmholtz
+%%free energy of fluids with soft potentials like that of the WCA fluid. 
+%%It will be helpful to continue looking at FMT before addressing SFMT.
+%%FMT begins 
+%by noticing that a quantity $G_{3}$ can be defined such that
+%\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
+%where $G_{3}$ is the volume of a sphere. 
+%Likewise, by defining quantities
+%\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
+%\begin{equation}{G_{1}=R}\end{equation}
+%\begin{equation}{G_{0}=1}\end{equation}
+%where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
+%radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
+%\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
+%The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
+%geometric measures of a sphere, and are what give rise to the name
+%Fundamental Measure Theory. Introducing a set of variables
+%%weighted densities $n_{i}=nG_{i}$ 
+%given by
+%\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
+%\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
+%\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
+%\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
+%Equation~\ref{FexfromSPTinthermsofG} can be written as
+%\begin{align}\label{FexfromSPT}
+    %F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
+%\end{align}
+%which has the form 
+%\begin{align}
+    %F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
+%\end{align}
+%%where 
+%\begin{align}
+   %\Phi_1 &= -n_{0}\ln(1-n_{3}) \label{Phi1equation} \\
+   %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
+   %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
+%\end{align}
+
+%The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
+%densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
+%\begin{align}
+    %n_{i}=nG_{i}
+%\end{align}
+%These weights are valid for determining the excess free energy of a system of 
+%hard spheres with a homogeneous number density n. But these weights, and 
+%their related  weighted densities, can be written in a more general form 
+%allowing for their extension to a system with a spatially varying number 
+%density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
+%excess Helmholtz free energy is written as
+%\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
+%Using a backwards Heaviside step function defined as
+%\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
+%the following expressions are obtained
+%\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
+%\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
+%\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
+%\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
+%where the weighted densities can now be expressed as
+%\begin{align}
+    %n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
+%\end{align} 
+%or for a spatially varying number density $n(\vec{r})$,
+%\begin{equation}\label{weighted-density}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
+%with weight functions
+%\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
+%\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
+%\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
+%\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
+
+%The weight functions $w_{2}$ and $w_{3}$ have the relation
+%\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
+%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
+
+%The weight functions can be related to the Mayer function 
+%\begin{align}
+     %f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+%\end{align} 
+%where r is the distance between the centers of two hard spheres, and 
+%$V(r)$ is the interatomic potential between two hard spheres given by 
+%\begin{align}
+    %V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+%\end{align}
+%Evaluating Equation with $V(r)$ given by Equation,  
+%the Mayer function for a hard-sphere fluid is found to be
+%\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
+%which can be written in terms of the weight functions if two 
+%additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
+%\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
+%\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
+%The Mayer function can then be written in terms of the weight functions~\cite{Hansen}
+%\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
+
+%Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
+%another relation that will be important to keep in tact when moving to 
+%SFMT~\cite{schmidt2000fluid}.
+%\begin{align}\label{mayerder=w2convolution}
+     %\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
+%\end{align} 
+%This relation can easily be checked for a hard-sphere fluid where
+%\begin{equation}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{equation}
+%\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
+%The right side can be evaluated by using the general relation for 
+%convolution which is equal to the Inverse Fourier Transform 
+%of a product of Fourier Transforms. 
+%\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
+%Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
+%gives
+%\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
+%and the right side of Equation~\ref{mayerder=w2convolution} becomes
+%\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
+%\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
+%which matches Equation~\ref{mayerder}.
+
+%Tarazona later improved upon the original FMT equations by replacing the
+%expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
+%more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
+%\begin{align}
+%\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
+%\end{align}
+%in which a tensor weight is incorporated given by
+%\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
+%%All the other FMT equations remain the same.
+
+%The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
+%All of the weight functions, the relations between them, and the equation 
+%for the derivative of the Mayer function, can be expressed in terms of 
+%$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
+%The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+%Equation~\ref{w2_w3_relation}.
+%Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
+%with soft potentials by keeping the relationships between the weight 
+%functions, and the equation for the derivative of the Mayer function 
+%the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
+%~\ref{w2_w3_relation},~\ref{eq:tensorweight},~\ref{mayerder=w2convolution}), 
+%but replacing $w_{3}(r)$ with a weight function appropriate 
+%for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+%This is what we do in forming a functional for the Weeks-Chandler-Anderson 
+%fluid as detailed in the next chapter on Methods.
 

### 2021-01-10 papers/thesis-kirstie/thesis.tex
edits

@@
 %To do this, the Helmholtz free energies of the fluid
-%in a liquid state and in a crystalline solid state must be found.
+%in a fluid state and in a crystalline solid state must be found.
 %In general, the Helmholtz free energy, which consists of an ideal part 
@@
 
-A liquid will eventually freeze as heat is removed, or the liquid is compressed. 
-The state of a simple fluid can be characterized by its entropy per particle~\cite{entropyfingerprint}.
-The entropy of a pure substance in a crystalline state is higher than the entropy 
-that it would have in its liquid state at a given temperature and pressure, 
-since there is more freedom for a particle to move between atoms  
-at lattice sites than in a dense liquid with irregular and blocked pathways~\cite{Hansen}.
-Both the entropy and the volume (and thus the density) change abruptly 
-at the phase transition.
+A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
+the density will change abruptly at the phase transition.
 The particular structure and density that result are affected 
@@
     \centering
-    \includegraphics[height=8cm]{figs/P-T_Diagram.pdf}
+    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
 
-\textcolor{red}{A phase transition occurs along a line on a 
+A phase transition occurs along a line on a 
 Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
@@
 %
-Points along the line that separates the solid and the liquid phases 
-indicate a phase transition where the solid and liquid phases coexist, 
-and show that both the solid and the liquid have the same temperature (necessary 
-for thermal equilibrium)
-and the same pressure (necessary for mechanical 
-equilibrium)at a phase transition.
-%
-, and the same chemical potential (necessary for diffusive equilibrium). 
+%Points along the phase boundary that separates the solid and the fluid phases 
+%indicate a phase transition where the solid and fluid phases coexist, 
+%and show that both the solid and the fluid have the same temperature (necessary 
+%for thermal equilibrium)
+%and the same pressure (necessary for mechanical 
+%equilibrium)at a phase transition, and the same chemical potential 
+%(necessary for diffusive equilibrium). 
 %
 This phase diagram is 
-representative of some arbitrary, unknown substance composed of atoms, 
-or molecules which exhibit both attraction and repulsion. However, since 
+representative of some substance 
+which exhibits both attraction and repulsion. However, since 
 the atoms in a WCA fluid do not exhibit attraction there will not be a 
 dividing line between the gas and liquid phases, there will just simply
-be a fluid, which we will refer to as a liquid in this paper.}
-%which may be referred to as a "liquid", since at the phase boundary the fluid is dense.
+be a fluid.\textcolor{red}{liquid in this paper.}
+The phase diagram indicates the equilibrium state a substance will be in at
+different temperatures and pressures.
+At a phase tranition, both the fluid and the solid have 
+the same temperature (necessary 
+for thermal equilibrium), the same pressure (necessary for mechanical 
+equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
+%The temperature , pressure, and chemical potential will be the same for both
+%the fluid and the solid at a fluid-solid phase transition.
 
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Liquid and solid densities at the solid-liquid phase 
+    \caption{Fluid and solid densities at the solid-fluid phase 
     transition are identified using Maxwell's Double Tangent Construction}
@@
 
-Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
-and one for the solid at a fixed temperature. Each curve consists of a 
-collection of data points $(\frac{F}{N}, \frac{1}{n})$ for the liquid and
-the solid respectively, which will be obtained by using cDFT to find the 
-Helmholtz free energies of the liquid and the solid for various 
-temperatures and number densities. 
+Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the fluid
+and one for the solid at a fixed temperature. 
+%Each curve consists of a collection of data points $(\frac{F}{N}, \frac{1}{n})$ 
+The data for the curves 
+%and the solid respectively, which 
+will be obtained by using cDFT. 
+%to find the Helmholtz free energies of the fluid and the solid for various 
+%temperatures and number densities. 
 Since the pressure of the 
-liquid is the same as the pressure of the solid at the point of transition, 
-the density of the liquid and the density of the solid at the point of 
+fluid is the same as the pressure of the solid at the phase transition, 
+the density of the fluid and the density of the solid at 
 freezing can be found from identifying the points at which the 
@@
 To facilitate finding the pressure at the phase transition computationally, 
-it is helpful to consider the Gibbs free energy given by G=F+PV. 
+it is helpful to take different approach by considering the Gibbs free energy given by G=F+PV. 
 Rearranging the terms and dividing by N gives an  expression for the 
@@
 free energy per atom, or chemical potential given by the y-intercept. 
-At the point of phase transition the chemical potential of the liquid is 
+At the point of phase transition the chemical potential of the fluid is 
 the same as the chemical potential of the solid, as is the pressure.
 \begin{equation}
-    \mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}
+    \mu_F=\mu_S\mbox{~~~~and~~~~}P_F=P_S\mbox{~~~~at a phase transition}
 \end{equation}
@@
     \caption{Pressure is found at point of intersection which occurs 
-    when $\mu_L=\mu_S$.}
+    when $\mu_F=\mu_S$.}
     \label{fig:GibbsvsP}
@@
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
-    \caption{The liquid and solid densities at the phase transition are 
-    found from a plot of pressure vs inverse density for both the liquid and the solid.}
+    \caption{The fluid and solid densities at the phase transition are 
+    found from a plot of pressure vs inverse density for both the fluid and the solid.}
     \label{fig:Pvsinvn}
@@
 intersection that occurs when the chemical potential 
-is plotted against the pressure for both the liquid and the 
+is plotted against the pressure for both the fluid and the 
 solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
@@
 density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
-liquid and the solid at a given temperature. 
+fluid and the solid at a given temperature. 
 The slope of the tangent line at a given point gives the negative of the 
@@
 line gives the chemical potential at that point. Collecting this data 
-(P,~$\mu$) at each point for the liquid and the solid respectively, 
+(P,~$\mu$) at each point for the fluid and the solid respectively, 
 produces the curves in Figure~\ref{fig:GibbsvsP}.
@@
     \centering
-    \includegraphics[height=8cm]{figs/T-n_Diagram.pdf}
+    \includegraphics[height=7cm]{figs/T-n_Diagram.pdf}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
-    the number densities of the liquid and the solid during a phase 
+    the number densities of the fluid and the solid during a phase 
     transition. The gray region is where different phases coexist in 
@@
 Once the pressure is known at the point of phase transition, for a given 
-temperature and number density, the liquid and solid number densities at 
+temperature and number density, the fluid and solid number densities at 
 the phase transition can be found from plots of the pressure versus the 
-inverse of the number density for both the solid and the liquid, 
+inverse of the number density for both the solid and the fluid, 
 as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
-find the liquid and solid number densities at the phase transition 
+find the fluid and solid number densities at the phase transition 
 for various temperatures, a Temperature-Density (T-n) phase diagram 
@@
 but without the gray hump characteristic of fluids that have 
-both attractive and repulsive interatomic potentials.
+attractive interatomic potentials, and which represents the region where
+vapor coexists with liquid as evidenced by an interface.
 
@@
 Classical Density Functional Theory will be used to find the 
-Helmholtz free energies of the liquid and the solid for various 
+Helmholtz free energies of the fluid and the solid for various 
 temperatures and number densities, so that phase diagrams can be 
@@
 until the free energy is minimized.
+
 %In this section 
@@
 Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
-for the hard-sphere fluid. FMT in turn extended  
-Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
-developed for homogeneous hard-sphere fluids, 
-to inhomogeneous, or non-uniform hard-sphere fluids.
-
-\textcolor{red}{FROM HERE}
+for the hard-sphere fluid. 
+%FMT in turn extended  
+%Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
+%developed for homogeneous hard-sphere fluids, 
+%to inhomogeneous, or non-uniform hard-sphere fluids.
+
+%\textcolor{red}{FROM HERE}
 %\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
@@
 where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
-%This can easily be shown by putting the interatomic potential between two hard spheres
-%given by 
-%\begin{align}
-    %V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
-%\end{align}
-%into the equation for the Mayer function
-%\begin{align}
-     %f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
-%\end{align} 
+This can easily be shown by putting the interatomic potential between two hard spheres
+$V(r)$
+into the equation for the Mayer function
+\begin{align}
+     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+\end{align} 
+where
+\begin{align}
+    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq R_i+R_j \\ 0  & r> R_i+R_j \end{array}\right.
+\end{align}
 %
@@
 the volume of intersection between two spheres that overlap, and is used in conjunction 
-with the exclusion volume to describe higher order terms in the expansion~\cite{rosenfeld1988}. 
+with the exclusion volume to describe higher order terms in the expansion
+in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988}. 
 %
@@
 %which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
+Taking the derivative of the Mayer function provides 
 another important relation expressed in terms of $w_{2}(r)$ 
@@
 have attraction between the particles that make up the fluid will show
-two distinct regions - a "liquid" region at lower densities, and a solid 
+two distinct regions - a fluid region at lower densities, and a solid 
 region at higher densities. 
-A phase transition between a liquid and a solid is accompanied by an 
+A phase transition between a fluid and a solid is accompanied by an 
 abrupt change in the number density, and so there is a gap between the 
-purely liquid and the purely solid regions. Within the gap, or coexistence
-region, is found mixtures of various amounts of liquid and solid each of
+purely fluid and the purely solid regions. Within the gap, or coexistence
+region, is found mixtures of various amounts of fluid and solid each of
 which have number densities that are fixed at the values they had at the 
 phase transition. 
-To construct the phase diagram, the liquid and solid phase transition number densities
-are found 
+To construct the phase diagram, the fluid and solid phase transition 
+number densities are found 
 %Maxwell's double tangent construction can be used to find 
 %these number densities given plots of 
-from the liquid and the solid Helmholtz 
+from the fluid and the solid Helmholtz 
 free energies and inverse densities at a given temperature, as
@@
 We use classical Density Functional Theory 
-to find the Helmholtz free energies of the liquid and the solid. 
+to find the Helmholtz free energies of the fluid and the solid. 
 A functional for the ideal portion of the
@@
 number density, as shown by the example in Figure~\ref{fig:Ensemble_vary}.
-A homogeneous number density profile is used for a liquid, as shown in 
+A homogeneous number density profile is used for a fluid, as shown in 
 Figure \ref{fig:homogen_denisty}.
@@
 density profile in the set, and also for the homogeneous number density profile 
-characteristic of a liquid. Both the ideal and excess free energies are 
+characteristic of a fluid. Both the ideal and excess free energies are 
 found, and added together to get the total Helmholtz free energy.
@@
 
-Once the Helmholtz free energy for both the crystal and liquid states of 
+Once the Helmholtz free energy for both the crystal and fluid states of 
 the fluid are found for a given temperature and density, they can be 
 used to construct phase diagrams.
-Outside the region of coexistence which exists between the liquid and 
+Outside the region of coexistence which exists between the fluid and 
 crystal states, the fluid will be completely in a crystalline state if the 
-crystal free energy is lower than the liquid free energy, and it will be 
-completely in a liquid state if the liquid free energy is lower than 
+crystal free energy is lower than the fluid free energy, and it will be 
+completely in a fluid state if the fluid free energy is lower than 
 the crystal free energy. 
-We determine the actual number densities of the crystal and liquid states
+We determine the actual number densities of the crystal and fluid states
 at the boundaries of the coexistence region for a given temperature, 
@@
 
-\section{Computing the Liquid and Crystal Helmholtz free energies}
-
-The liquid ideal Helmholtz free energy is computed analytically 
+\section{Computing the Fluid and Crystal Helmholtz free energies}
+
+The fluid ideal Helmholtz free energy is computed analytically 
 using Equation~\ref{eq:Fideal} 
@@
 where $\vec R_j$ is the position vector of the jth lattice site.
-The liquid excess Helmholtz free energy is computed analytically  
+The fluid excess Helmholtz free energy is computed analytically  
 using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
@@
 $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
-It is seen that in general, a liquid exists at lower number densities, 
+It is seen that in general, a fluid exists at lower number densities, 
 and a solid exists at higher number densities. 
 For a given temperature, it is seen that as the density of the 
-liquid increases, a maximum number density is reached for the liquid. 
+fluid increases, a maximum number density is reached for the fluid. 
 Looking at the coexistence region as a whole for all temperatures, 
@@
 and shifted a bit toward smaller densities. For example, at a temperature 
-of kT=1, the cDFT liquid transition reduced number density is 0.96, 
+of kT=1, the cDFT fluid transition reduced number density is 0.96, 
 and the solid transition reduced number density is 1.02, giving a spread 
-of 0.06, whereas the Monte-Carlo results give a liquid transition 
+of 0.06, whereas the Monte-Carlo results give a fluid transition 
 reduced number density of 0.92, and a solid transition reduced number 
@@
   \caption{Phase Diagram of the reduced temperature verses the reduced 
-  density for the WCA fluid. The red region is liquid, and the blue region is 
-  solid. The gray area that separates the liquid and solid regions 
-  is the region of coexistence where some liquid and solid exist together 
+  density for the WCA fluid. The red region is fluid, and the blue region is 
+  solid. The gray area that separates the fluid and solid regions 
+  is the region of coexistence where some fluid and solid exist together 
   at varying proportions. White lines show the boundaries of the coexistence 
@@
 %The solid phase transition number densities predicted by each method 
-%are very nearly the same along all temperatures, but the liquid phase 
+%are very nearly the same along all temperatures, but the fluid phase 
 %transition number densities predicted by cDFT are notably higher than 
@@
   %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
-  %%White lines show the liquid and solid transition densities at each temperature from 
+  %%White lines show the fluid and solid transition densities at each temperature from 
   %%simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
@@
 for temperatures of kT = 0.5 to 3.0. 
-The black line marks the phase boundary between the solid and liquid regions. 
+The black line marks the phase boundary between the solid and fluid regions. 
 It is seen that the higher the temperature, the higher the pressure must be
-to undergo a phase transition from a liquid to a solid, as would be expected. 
+to undergo a phase transition from a fluid to a solid, as would be expected. 
 Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) 
@@
 \end{align}
-Since the entropy and the volume of the liquid is greater than the entropy
+Since the entropy and the volume of the fluid is greater than the entropy
 and the volume of the solid, the slope should be positive.
@@
 Monte-Carlo simulation predicts 43.80.
-This may indicate that either the difference between the liquid 
+This may indicate that either the difference between the fluid 
 and solid phase transition entropies predicted by cDFT is too high, or 
@@
   \caption{Phase Diagram of the reduced pressure verses the reduced 
-  temperature. The blue region is solid, the red region is liquid,
-  and the black line shows the phase boundary where the solid and liquid
-  phases coexist. The white line marks the solid-liquid phase boundary  
+  temperature. The blue region is solid, the red region is fluid,
+  and the black line shows the phase boundary where the solid and fluid
+  phases coexist. The white line marks the solid-fluid phase boundary  
   as predicted by Monte-Carlo simulations.}
@@
 sites until the Helmholtz free energy was minimized. Once the equilibrium 
-Helmholtz free energy for the liquid and solid states were found for a given
+Helmholtz free energy for the fluid and solid states were found for a given
 temperature and number density, 
 we used this data 
-to find the number densities of the liquid and the solid at the phase transition.
+to find the number densities of the fluid and the solid at the phase transition.
 Using this information, we generated a temperature-density 
@@
 \begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{liquid}=\frac{lFideal_{primitive.cell} + lFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{liquid}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{crystalFideal_{primitive.cell} + crystalFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{fluid}=\frac{fluidFideal_{primitive.cell} + fluidFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{fluid}\end{equation}
 \begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
 \begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{liquid}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{fluid}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
 

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

@@
 primary reference system for fluids. This is due to the fact that the 
-repulsion between atoms is much stronger than their attraction, and so 
-the repulsive potential, which is descriptive of hard spheres, dominates 
-the interactions between atoms.
+short-range repulsion between atoms is much stronger than their long-range 
+attraction, and so the repulsive potential, which is descriptive of 
+hard spheres, dominates the interactions between atoms.
 
@@
 fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
-a hard-edge attractive potential to the repulsive potential of the 
+a hard-edge short-range attractive potential to the repulsive potential of the 
 hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-spheres with softly fading repulsive potentials which he modeled by making
+spheres with softly fading short-range repulsive potentials which he modeled by making
 the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
@@
 Chandler, and Anderson used temperature and density dependent hard-sphere 
-radii to model spheres with the more realistic softly fading repulsive 
-portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
-which is known as the Weeks-Chandler-Anderson model fluid. 
+radii to model spheres with a more realistic softly fading repulsive 
+potential similar to that between neutral atoms
+%portion of the Leonard-Jones potential
+which is known as the Weeks-Chandler-Anderson model fluid~\cite{andersen1971relationship} . 
 Models continue to be developed to characterize fluids with particular 
 properties, or to address particular problems like surface tension, 
-interfaces, adsorption, wetting or confinment.
-\textcolor{red}{A review of some model fluids and problems that have been
-studied more recently is given by L\"{o}wen~\cite{lowen}.}
+interfaces, adsorption, wetting and confinement.
+A review of some model fluids and problems that have been
+studied more recently is given by L\"{o}wen~\cite{lowen}.
 
@@
 by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
-Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium}.
+Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium} amoung others. A review
+of the development of cDFT is found in Fundamentals of Inhomogeneous fluids~\cite{inhomogeneousfluids}.
+
 To use classical Density Functional Theory (cDFT), an expression for the free energy 
@@
 %Led by the consistency between results from SPT and Percus-Yevick, 
-\textcolor{red}{In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
+In 1989 Rosenfeld introduced Fundamental Measure Theory (FMT)~\cite{rosenfeld1989,rosenfeld1988} 
 to form a functional for the hard-sphere fluid based on the idea that 
@@
 of the form of 
-a backward Heaviside step function}
+a backward Heaviside step function
 %\begin{equation}
@@
 while also yielding the correct excess free energy for the fluid as its 
-confinment is reduced from a three-dimensional volume (as in a bulk fluid), 
+confinement is reduced from a three-dimensional volume (as in a bulk fluid), 
 to a two-dimensional plane (as in a thin-film), to a one-dimensional 
@@
 %, similar to that exhibited by a WCA fluid. 
-\textcolor{red}{For some more recent developments, see the review of new developments in 
+For some more recent developments, see the review of new developments in 
 classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
 Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
-and a recent generalization of the Barker-Henderson model to inhomogeneous 
-fluids by Tschopp, Vuijk, Sharma and Brader~\cite{BHextended} (2020).}
-
-\textcolor{red}{Classical density functional theory has also been applied 
+and a generalization of the Barker-Henderson model to inhomogeneous 
+fluids by Tschopp and coworkers~\cite{BHextended} (2020).
+
+Classical density functional theory has also been applied 
 to phase transitions, and freezing, but not until recently has there
 been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
-crystal as a highly inhomogeneous liquid~\cite{} (1979). Oettel and fellow 
-researchers studied hard-sphere crystals with ~\cite{} (2010). Mortazavifar and Oettel
-studied the fluid and crystal phases of the Asakura-Oosawa model fluid (2016). 
-Lutsko and Lam studied studied crystallization of the Leonard-Jones fluid~\cite{lutsko} (2018),
-and Lutsko and Schoonen looked at cDFT applied to the solid state~\cite{2020}.
-In all these studies the hard-sphere fluid has been used as the reference system,
-which is implemented with FMT, and to which other terms are added to create 
-the full functional for the fluid being studied.}
+crystal as a highly inhomogeneous liquid~\cite{ramakrishnanandyussouff} (1979). 
+Oettel and fellow researchers studied hard-sphere crystals with fully developed FMT~\cite{Oettel2010} (2010). 
+Mortazavifar and Oettel studied the fluid and crystal phases of the 
+Asakura-Oosawa model fluid with FMT~\cite{asakuraoosawa} (2016). 
+Lutsko and Lam studied studied crystallization of the Leonard-Jones 
+fluid~\cite{lutsko} (2018), and Lutsko and Schoonen looked at cDFT applied 
+to the solid state~\cite{lutskoschoonen2020} (2020).
+In all these studies the hard-sphere fluid has been used as the reference 
+system to which other terms are added to create 
+the full functional for the fluid being studied.
+
 In this paper, classical Density Functional Theory will be used to predict 
@@
 Results will be compared to those generated from Monte-Carlo simulations.
-
-%Results will be compared to those from Monte-Carlo simulations.
 
@@
 \end{align} 
-The WCA potential decreases from infinity as the distance between the 
+The WCA potential decreases from as the distance between the 
 centers of the atoms r increases, and becomes, and remains, essentially 
@@
 their interatomic potential, they are called ``soft spheres'' and their 
-gradually fading potential makes them slightly ``squishy'', as illustrated in
+gradually fading potential makes them slightly ``squishy'', as pictured in
 Figure~\ref{fig:TwoSpheres}.
@@
 
-A phase transition occurs along a line on a 
-Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
-diagram shown in Figure~\ref{fig:P-T_Diagram}. 
+A phase transition occurs along a phase boundary line on a phase diagram
+such as can be seen in the pressure-temperature (P-T) phase diagram
+shown in Figure~\ref{fig:P-T_Diagram}. 
 %
@@
 dividing line between the gas and liquid phases, there will just simply
-be a fluid.\textcolor{red}{liquid in this paper.}
+be a fluid.
 The phase diagram indicates the equilibrium state a substance will be in at
 different temperatures and pressures.
-At a phase tranition, both the fluid and the solid have 
+At a phase transition, both the fluid and the solid have 
 the same temperature (necessary 
@@
 The pressure can be related to the Helmholtz free energy F, which is minimized 
-for a system at fixed temperature, volume and number of particles, as the 
-entropy of the system and its surroundings is maximized. The negative of 
-the partial derivative of the Helmholtz free energy with respect to volume V, 
-and at a fixed temperature T and number of particles N, equals the pressure. 
+for a system at fixed temperature, volume, and number of particles as the 
+entropy of the system and its surroundings added together is maximized. The negative of 
+the partial derivative of the Helmholtz free energy with respect to the volume V 
+at a fixed temperature T and number of particles N, equals the pressure. 
 \begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
@@
 \end{equation}
-and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
-pressure given by the negative of the slope of the line, and the Gibb's 
-free energy per atom, or chemical potential given by the y-intercept. 
-At the point of phase transition the chemical potential of the fluid is 
+and describes the black line shown in Figure~\ref{fig:MaxwellDT} where the 
+pressure is given by the negative of the slope of the line, and the Gibb's 
+free energy per atom, or chemical potential is given by the y-intercept. 
+At the phase transition the chemical potential of the fluid is 
 the same as the chemical potential of the solid, as is the pressure.
@@
 This plot is constructed from data collected from lines tangent 
-along the curves on the Helmholtz free energy per atom verses inverse 
+along the curves on Helmholtz free energy per atom verses inverse 
 density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
-fluid and the solid at a given temperature. 
-The slope of the tangent line at a given point gives the negative of the 
-pressure at that point, and the corresponding y-intercept of the tangent 
-line gives the chemical potential at that point. Collecting this data 
-(P,~$\mu$) at each point for the fluid and the solid respectively, 
-produces the curves in Figure~\ref{fig:GibbsvsP}.
+fluid and the solid at a given temperature where 
+the slope of the tangent line at a given point gives the negative of the 
+pressure, and the corresponding y-intercept of the tangent 
+line gives the chemical potential at a given point. 
+%Collecting this data 
+%(P,~$\mu$) at each point for the fluid and the solid respectively, 
+%produces the curves in Figure~\ref{fig:GibbsvsP}.
 
@@
 find the fluid and solid number densities at the phase transition 
-for various temperatures, a Temperature-Density (T-n) phase diagram 
-can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
-but without the gray hump characteristic of fluids that have 
-attractive interatomic potentials, and which represents the region where
-vapor coexists with liquid as evidenced by an interface.
+for various temperatures, a temperature-density (T-n) phase diagram 
+can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}. 
+However, for a WCA fluid there will be no gray hump which is characteristic 
+of fluids that have attractive interatomic potentials, and which display
+an interface between vapor and liquid at temperatures below the critical point.
 
@@
 free energy can then be written~\cite{schroeder} as
-\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
+\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}\approx k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
 in the thermodynamic limit of large N. 
@@
 \begin{align} 
-    f(r) = -\Theta(r-(R_i + R_j))
-\end{align}
-where $r=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres.
+    f(r_{ij}) = -\Theta(r_{ij}-(R_i + R_j))
+\end{align}
+where $r_{ij}=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres, 
+and $\Theta$ is 1 for $r_{ij}<R_i+R_j$ and 0 for $r_{ij}>R_i+R_j$.
 This can easily be shown by putting the interatomic potential between two hard spheres
-$V(r)$
+$V(r_{ij})$
 into the equation for the Mayer function
 \begin{align}
-     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+     f(r_{ij})=\exp^{-\frac{V(r_{ij})}{k_{B}T}}-1
 \end{align} 
@@
 \begin{align}
-    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq R_i+R_j \\ 0  & r> R_i+R_j \end{array}\right.
+    V(r_{ij})=\left\{ \begin{array}{rc} \infty & 0<r_{ij} \leq R_i+R_j \\ 0  & r_{ij}> R_i+R_j \end{array}\right.
 \end{align}
@@
 is equal to the volume of a sphere with a radius equal to that of the radii 
-of two hard spheres added together. For a mixture of hard spheres of radius $R_a$ and $R_b$
+of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$
 this becomes
 \begin{align}
-       \int_{Volume} \Theta(r-(R_i+R_j))~d\vec r &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\
+       \int_{Volume} \Theta(r_{ij}-(R_i+R_j))~d\vec r_{ij} &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\
        &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + R_i(4\pi R_j^2) + \frac{4\pi}{3}R_j^3  \nonumber \\
-       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)
+       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)  \label{VSR1sum}
 \end{align}  
@@
 \begin{align}
-        V &= \int_{Volume} \Theta(r-R)d\vec r \\
-        S &= \int_{Volume} \delta(r-R) d\vec r \\
-        R &= \int_{Volume} \frac{\delta(r-R)}{4\pi r} d\vec r \\
-                1 &= \int_{Volume} \frac{\delta(r-R)}{4\pi r^2} d\vec r   
-\end{align}
-Limiting this discussion to a homogeneous hard-sphere fluid composed of hard spheres that have the same
-radius R, the lowest order term of the virial expansion of the excess Helmholtz free energy which is given in Equation~\ref{eq:Fexcess} 
-can now be written as
-\begin{align}
-      F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
-                 &= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\
-                 &= k_BTn^2(2V(1) + 2SR)V  \\
-                 &= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
-                 &+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
-                 &= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
-                 &= k_BT(\Phi_1 + \Phi_2)V \\
-                 &= k_BT\Phi(\{n_\alpha\})V
-\end{align} 
-The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
-with a dimensionless energy composed 
-of the sum of $\Phi_1$ and $\Phi_2$ which are defined as
-\begin{align}
-      \Phi_1 = n_3n_0  \\
-      \Phi_2 = n_1n_2
-\end{align}
-where $n_i$ are weighted densities of the form
-\begin{align}
-     n_\alpha=\int nw_\alpha(r) d\vec r
-\end{align}
-with weight functions defined as
-\begin{align} 
-      w_3 &= \Theta(r-R)  \label{scalor-weight3} \\
-      w_2 &= \delta(r-R)  \\
-      w_1 &= \frac{w_2}{4\pi r}  \\
-      w_0 &= \frac{w_2}{4\pi r^2} \label{scalor-weight0}
-\end{align}
+        V &= \int_{Volume} \Theta(r_{ij}-R)d\vec r_{ij} \\
+        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  \\
+        R &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}} d\vec r_{ij} \\
+        1 &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}^2} d\vec r_{ij}    
+\end{align}
+Rosenfeld also recognized that the Heaviside function can be written as 
+a convolution of the weight functions which, for a fluid composed of 
+spheres of radius R, is given by
+\begin{align}
+    f(r_{ij})=-\Theta(r_{ij}-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+\end{align} 
+where the weights are  
+\begin{align} 
+      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} \\
+      w_2 &= \delta(r_{ij}-R)  \\
+      w_1 &= \frac{w_2}{4\pi r_{ij}}  \\
+      w_0 &= \frac{w_2}{4\pi r_{ij}^2} \label{scalar-weight0}
+\end{align}
+and the two vector weights are
+\begin{align}
+      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{r_{ij}} \\
+      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{r_{ij}}
+\end{align}
+which are needed for the relation to be satisfied.
+%%SAVE the following - connects to convolution! ------------------------
+%Rewriting the first term V(1) in Equation~\ref{VSR1sum} gives
+%\begin{align}
+       %V(1) &=\left(\int\Theta(r_{ij}-R)d\vec r_{ij}\right)\left(\int \frac{\delta(r_{ij}-R)}{4\pi r_{ij}^2} d\vec r_{ij}\right) \\
+           %%&=\left(\int_{Volume} \Theta(r'-R)d\vec r~'\right)\left(\int_{Volume} \frac{\delta(r"-R)}{4\pi {r"}^2} d\vec r~"\right)   \\
+            %&=\left(\int \Theta(|\vec r~'|-R)d\vec r~'\right)\left(\int \frac{\delta(|\vec r~"|-R)}{4\pi {|\vec r~"|}^2} d\vec r~"\right) 
+           %%&=\int_{Volume} \Theta(|\vec r~'|-R)\left(\int_{Volume} \frac{\delta(|\vec r_{ij}-\vec r~'|-R)}{4\pi {|\vec r_{ij}-\vec r~'|}^2} d\vec r_{ij}\right) d\vec r~'
+%\end{align}
+%Setting $\vec r~"=\vec r-\vec r~'$ and treating $\vec r~'$ as a 
+%constant gives $d\vec r~" = d\vec r$ which is moved to the outside to give 
+%\begin{align}
+       %V(1) &=\int \Theta(|\vec r~'|-R)\left(\int\frac{\delta(|\vec r-\vec r~'|-R)}{4\pi {|\vec r-\vec r~'|}^2} d\vec r\right) d\vec r~'  \\
+            %&=\int \left(\Theta(|\vec r~'|-R)\int \frac{\delta(|\vec r-\vec r~'|-R)}{4\pi {|\vec r-\vec r~'|}^2} d\vec r~'\right) d\vec r  \\
+            %&=\int \left(\int w_3(\vec r~')w_0(\vec r-\vec r~')~d\vec r~'\right) d\vec r  \\
+            %&=\int (w_3\otimes w_0(\vec r)) d\vec r
+%\end{align}
+%%end SAVE -------------------------------------------------------------
+%
+%Limiting this discussion to a fluid composed of hard spheres that have the same
+%radius R, the lowest order term of the virial expansion of the excess Helmholtz 
+%free energy of a homogeneous fluid given in Equation~\ref{eq:Fexcess} 
+%can now be written as
+%\begin{align}
+      %F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
+                 %&= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\
+                 %&= k_BTn^2(2V(1) + 2SR)V  \\
+                 %&= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
+                 %&+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
+                 %&= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
+                 %&= k_BT(\Phi_1 + \Phi_2)V \\
+                 %&= k_BT\Phi(\{n_\alpha\})V
+%\end{align} 
+%The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
+%with a dimensionless energy composed 
+%of the sum of $\Phi_1$ and $\Phi_2$ which are defined as
+%\begin{align}
+      %\Phi_1 = n_3n_0  \\
+      %\Phi_2 = n_1n_2
+%\end{align}
+%where $n_i$ are weighted densities of the form
+%\begin{align}
+     %n_\alpha=\int nw_\alpha(r) d\vec r
+%\end{align}
+%The other terms can be written in a similar way with weight functions defined as
+%\begin{align} 
+      %w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
+      %w_2 &= \delta(r-R)  \\
+      %w_1 &= \frac{w_2}{4\pi r}  \\
+      %w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}
+%\end{align}
+%The Heaviside function can then be written as 
+%a convolution of the weight functions which, for a fluid composed of 
+%spheres of radius R, is given by
+%\begin{align}
+    %f(r_{ij})=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+%\end{align} 
+%where the two vector weights are given by
+%\begin{align}
+      %\vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
+      %\vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
+%\end{align}
+%which are needed for the relation to be satisfied.
 %The excess Helmholtz free energy can also be expressed in terms of the
@@
 %
-For the inhomogeneous case where the number density varies spatially,
-$n\rightarrow n(\vec r)$, Equation~\ref{eq:Fexcess} becomes
-\begin{align}
-     F_{excess} = k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
-\end{align}
-where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
-position vector to another sphere, and $|\vec r_i - \vec r_j|$ 
-is the distance between the centers of the two spheres.
-Rosenfeld recognized that the Heaviside function can be written as 
-a convolution of the weight functions which, for a fluid composed of 
-spheres of radius R, is given by
-\begin{align}
-    f(|\vec r_i-\vec r_j|)=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
-\end{align} 
-%where the weights are the same ones given in Equations~\ref{scalor-weight3} through~\ref{scalor-weight0}, 
-where the two vector weights are given by
-\begin{align}
-      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
-      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
-\end{align}
-which are needed for the relation to be satisfied, 
-and the other weights take the form 
-\begin{align} 
-      w_3 &= \Theta(|\vec r_i-\vec r_j|-R)  \label{scalor-weight3} \\
-      w_2 &= \delta(|\vec r_i-\vec r_j|-R)  \\
-      w_1 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|}  \\
-      w_0 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|^2} \label{scalor-weight0}
-\end{align}
+%For the inhomogeneous case where the number density varies spatially,
+%$n\rightarrow n(\vec r)$, Equation%~\ref{eq:Fexcess} 
+%becomes (for a zero external potential)
+%\begin{align}
+     %F_{excess} = k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
+%\end{align}
+%where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
+%position vector to another sphere, and $r_{ij}=|\vec r_i - \vec r_j|$ 
+%is the distance between the centers of the two spheres.
+%
+%For the inhomogeneous case where the number density varies spatially,
+%$n\rightarrow n(\vec r)$, Equation%~\ref{eq:Fexcess} 
+%becomes (for a zero external potential)
+%\begin{align}
+     %F_{excess} &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  \\
+                %&= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
+%\end{align}
+%where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
+%position vector to another sphere, and $r_{ij}=|\vec r_i - \vec r_j|$ 
+%is the distance between the centers of the two spheres.
+%Rosenfeld recognized that the Heaviside function can be written as 
+%a convolution of the weight functions which, for a fluid composed of 
+%spheres of radius R, is given by
+%\begin{align}
+    %f(r_{ij})=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
+%\end{align} 
+%where the weights are the same ones given in Equations~\ref{scalar-weight3} through~\ref{scalar-weight0}, 
+%where the two vector weights are given by
+%\begin{align}
+      %\vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
+      %\vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
+%\end{align}
+%which are needed for the relation to be satisfied, 
+%and the other weights take the form 
+%The weights take the form
+%\begin{align} 
+      %w_3 &= \Theta(|\vec r_i-\vec r_j|-R)  \label{scalar-weight3} \\
+      %w_2 &= \delta(|\vec r_i-\vec r_j|-R)  \\
+      %w_1 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|}  \\
+      %w_0 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|^2} \label{scalar-weight0}
+%\end{align}
 %a uniquely deconvolved 
 %into a set of weight functions related to the geometry of a sphere, and that it
-%describes mathmetically 
+%describes mathematically 
 %the exclusion volume of a pair of spheres, one of the two variables 
@@
 \end{align}
-For the scalor weights this is the same as
-\begin{align} 
-                                                   &= \int w_\gamma(\vec u - \vec r_i)w_\alpha(\vec u - \vec r_j)  d\vec u  
+For the scalar weights this is the same as
+\begin{align} 
+     w_\alpha\otimes w_\gamma(\vec r_i - \vec r_j) &= \int w_\gamma(\vec u - \vec r_i)w_\alpha(\vec u - \vec r_j)  d\vec u  
 \end{align}
@@
 \begin{align} 
-                                                   &= -\int \vec {w}_\gamma(\vec u - \vec r_i)\cdot\vec {w}_\alpha(\vec u - \vec r_j)  d\vec u  
+     \vec w_\alpha\otimes \vec w_\gamma(\vec r_i - \vec r_j) &= -\int \vec {w}_\gamma(\vec u - \vec r_i)\cdot\vec {w}_\alpha(\vec u - \vec r_j)  d\vec u  
 \end{align}
@@
 %since the weight functions involve the distances $|\vec r_a - \vec x|=|\vec x - \vec r_a|$.
-The lowest order term of the excess Helmholtz free energy can then be written as
-\begin{align}
-     F_{excess} &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  \\
-                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}  \\
-                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
-                &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
-                &+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec x ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
-                &= k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
+The lowest order term of the virial expansion of the excess Helmholtz 
+free energy (for a zero external potential) is given by
+%The lowest order term of the excess Helmholtz free energy can then be written as
+\begin{align}
+     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  
+\end{align} 
+can then be written as 
+\begin{align}      
+     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}  \\
+                &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
+                &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
+                &+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec u ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)    \\
+                &\approx k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
                 &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
-                &+ \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u  \\
-                &= k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\
-                &= k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
-                &= k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
+                &- \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u
+\end{align} 
+\begin{align}                
+                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\                
+                &\approx k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
+                &\approx k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
                 %&\text{or replacing u notation with r} \nonumber \\
-                &= k_BT\int \Phi(n(\vec r))~d\vec r             
-\end{align}
-where the weighted densities $n_\alpha$ have the form
-\begin{align}
-    n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
+                &\approx k_BT\int \Phi(\vec r)~d\vec r             
+\end{align}
+where the weighted densities $n_\alpha$ all depend on the spatially varying 
+number density profile $n(\vec r)$, and have the form
+\begin{align}
+    n_\alpha(\vec u)=\int n(\vec {r}~')w_\alpha(\vec u-\vec {r}~')~d\vec {r}~'  
     %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
 \end{align}  
-For the homogenesous case $\vec n_1\cdot\vec n_2$=0, and Equation~\ref{Fexlowden-inhomo} reduces to Equation~\ref{Fexlowden-homo}.
+and  $\vec n_1\cdot\vec n_2$=0 for the homogenesous case where $n(\vec r)\rightarrow n$.
+%Equation~\ref{Fexlowden-inhomo} reduces to Equation~\ref{Fexlowden-homo}.
 %The excess free energy in the low density limit then takes the form
@@
 \end{align}
-where in addition to the weight functions already defined, a tensor weight function is given by
+where in addition to the weight functions already defined,  
+\begin{align} 
+      w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
+      w_2 &= \delta(r-R)  \\
+      w_1 &= \frac{w_2}{4\pi r}  \\
+      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  \\
+      \vec {w}_2 &= w_2~\hat r \\
+      \vec {w}_1 &= w_1~\hat r
+\end{align}
+a tensor weight function is given by
 \begin{align}\label{eq:tensorweight-methods}
@@
 \end{align}
-The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
-all depend on the spatially varying number density profile $n(\vec r)$.
-\begin{align}%\label{eq:numdenprofile}
-	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
-\end{align}
-This result reduces to same result obtained from Scaled Particle Theory 
+%The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
+%all depend on the spatially varying number density profile $n(\vec r)$.
+%\begin{align}%\label{eq:numdenprofile}
+	%n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
+%\end{align}
+This result reduces to the same result obtained from Scaled Particle Theory 
 for a homogeneous hard-sphere fluid.
@@
 \begin{align}\label{w2_w3_relation}
-    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
 \end{align}
@@
 \end{align} 
+Thus, all the weight functions, and the above relation are determined
+by $w_{3}(r)$.
 Soft Fundamental Measure Theory extends FMT to fluids 
 with soft potentials by keeping the relationships between the weight 
-functions, and the equation for the derivative of the Mayer function 
-the same, 
-but replacing $w_{3}(r)$ with a weight function appropriate 
+functions, and Equation~\ref{mayerder=w2convolution} 
+for the derivative of the Mayer function 
+the same, but replacing $w_{3}(r)$ with a weight function appropriate 
 for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
@@
 %The weight functions $w_{2}$ and $w_{3}$ have the relation
-%\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
+%\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
 %which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
@@
 \begin{displaymath}
-     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
 \end{displaymath}

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

@@
 It can be thought of as an average over a large collection of actual cells
-of various possible spatial distributions of the atoms within a cell, or
+with various possible spatial distributions of the atoms within the cell, or
 more precisely, as an ensemble average over all possible microstates, 
-each of which consists of a unique set of positions for the atoms in 
+each of which consists of a unique set of positions for the atoms 
 that occupy the cell. Mathematically, the microstates are represented 
@@
 
-Besides changing the width of the Gaussians, there is another way to 
-generate different number density profiles for a given number density.
+There is another way to generate different number density profiles for 
+a given number density besides changing the width of the Gaussians. 
 It may be that not all lattice sites are occupied. Vacancies can be taken 
@@
 
-Once the Helmholtz free energy for both the crystal and fluid states of 
-the fluid are found for a given temperature and density, they can be 
+Once the Helmholtz free energy for both the crystal and fluid 
+are found for a given temperature and density, they can be 
 used to construct phase diagrams.
-Outside the region of coexistence which exists between the fluid and 
-crystal states, the fluid will be completely in a crystalline state if the 
-crystal free energy is lower than the fluid free energy, and it will be 
-completely in a fluid state if the fluid free energy is lower than 
+Outside the region of coexistence, a purely crystalline state will exist 
+if the crystal free energy is lower than the fluid free energy, and 
+a purely fluid state will exist if the fluid free energy is lower than 
 the crystal free energy. 
-We determine the actual number densities of the crystal and fluid states
-at the boundaries of the coexistence region for a given temperature, 
-and construct phase diagrams as described in Section 2.2.
+After the actual number densities of the crystal and fluid states
+at the boundaries of the coexistence region are determined for a given 
+temperature, phase diagrams can be constructed as described in Section 2.2.
 
@@
 of the Helmholtz free energy is difficult since it depends on the interatomic 
-potential energy $V(r)$ that exists between each sphere with all the other spheres, 
+potential energy $V(r)$ that exists between each sphere and all the other spheres, 
 as is shown by Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
@@
 \end{align} 
-The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
+The weighted densities $n_0, n_1, n_2, n_3, \vec {n}_1, \vec {n}_2$ and $\overleftrightarrow{n}_{m2}$ are found from
 \begin{align}\label{eq:numdenprofile}
-	n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
+	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
 \end{align}
@@
 and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
-The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
-function potential $V_{erf}(r)$ approximate, in the region of interest, 
-the temperature-independent WCA potential which is difficult 
+%The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
+%function potential $V_{erf}(r)$ approximate, in the region of interest, 
+%the temperature-independent WCA potential which is difficult 
+%to implement computationally. 
+The two parameters $\alpha$ and $\Xi$ are designed to make the functional
+produce results which approximate, in the region of interest, that of  
+a WCA fluid using the temperature-dependent error function potential $V_{erf}(r)$ 
+instead of the temperature-independent WCA potential which is difficult 
 to implement computationally. 
@@
 in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
-Using Equation~\ref{w2_w3_relation} repeated here,
+Using Equation~\ref{w2_w3_relation}(repeated below),
 \begin{displaymath}
@@
 along the z-axis. In three dimensions, the weighting functions 
-$w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$, 
+$w_0$, $w_1$, $w_2$ are spherical shells, 
 and $w_3$ is a solid sphere with an edge that tapers off.
 $\Xi$ is related to the width of the weight functions which 
-become thicker, or taper off longer as the temperature increases. 
+become thicker, or taper off longer as the temperature increases, and 
 $\alpha$ is related to the radius of the shell, or sphere, and decreases 
@@
 \section{Computing the Fluid and Crystal Helmholtz free energies}
-
+The Helmholtz free energy is composed of two parts: an ideal Helmholtz free energy 
+and an excess Helmholtz free energy that are added together to get the total
+Helmholtz free energy, $F=F_{ideal} + F_{excess}$. 
+And so the ideal and excess Helmholtz free energies
+must be computed for both the fluid and the crystal states.
 The fluid ideal Helmholtz free energy is computed analytically 
-using Equation~\ref{eq:Fideal} 
-for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
+using Equation~\ref{eq:Fideal} for a given homogeneous reduced 
+(or dimensionless) number density $n^*$ and temperature $t^*$. 
 The crystal ideal Helmholtz free energy is computed computationally 
@@
 where $\vec R_j$ is the position vector of the jth lattice site.
-The fluid excess Helmholtz free energy is computed analytically  
-using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
-with a uniform number density profile, n that does not depend on $\vec r$.
 
@@
 
+The fluid excess Helmholtz free energy is computed analytically  
+using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
+with a uniform number density profile, n that does not depend on $\vec r$. 
 The crystal excess Helmholtz free energy is computed  
@@
 centered at lattice sites specified by position vectors $\vec R$. 
-The weight functions $w_i(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
+The weight functions $w_\alpha(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
 as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}a. 
@@
 Keeping the weight function centered at $\vec r$, 
-Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
+Monte-Carlo integration can be used to compute the weighted densities $n_\alpha(\vec r)$
 by randomly generating vectors
@@
 Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
-\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_i(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
+\begin{equation}{n_\alpha(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
 and is implemented computationally in the form
-\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_i(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
+\begin{equation}{n_\alpha(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
 which can be written as
-\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}~'|)}\end{equation} 
+\begin{equation}{n_\alpha(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_\alpha(|\vec{r}-\vec{r}~'|)}\end{equation} 
 NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
@@
 is equal to a probability density which when multiplied by dV gives 
-the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}~'|)$ 
+the probability $P_{r'}$ at which the value of $w_\alpha(|\vec{r}-\vec{r}~'|)$ 
 will occur. Using the relation
 \begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
-$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}~'|)$:
-\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
+$n_\alpha(\vec{r})$ can be expressed as the average value of $(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)$:
+\begin{equation}{n_\alpha(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
 This process is repeated to get the contributions to the weighted density 
-$n_i(\vec r)$ from the Gaussians around each lattice point which begin to 
+$n_\alpha(\vec r)$ from the Gaussians around each lattice point which begin to 
 merge together as the width of the Gaussians increases.
@@
 Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
-and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
+and \ref{fig:Phase_Diagram_P_vs_T}. 
 In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
@@
 $n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
+In all cases $\sigma=1$ and $\epsilon=1$. 
 It is seen that in general, a fluid exists at lower number densities, 
 and a solid exists at higher number densities. 
-For a given temperature, it is seen that as the density of the 
-fluid increases, a maximum number density is reached for the fluid. 
+For a given temperature, it is seen that as the density increases, 
+a maximum number density is reached for the fluid. 
 Looking at the coexistence region as a whole for all temperatures, 
@@
 \begin{align} 
-    \frac{dP}{dT}=\frac{S_l-S_s}{V_l-V_s}
+    \frac{dP}{dT}=\frac{S_f-S_s}{V_f-V_s}
 \end{align}
@@
 from Monte-Carlo simulations shown by the white line, it is seen that
-the cDFT transition temperatures are consistently higher, and the 
+the phase transition pressures and temperatures predicted by cDFT 
+are consistently higher, and the 
 discrepancy grows as the temperature increases. For example, at a 
 temperature of kT=1, the phase transition reduced pressure predicted by 
-cDFT is 15.24, whereas the Monte-Carlo simulation predicted a value of 9.99,
+cDFT is 15.24, whereas the Monte-Carlo simulation predicted a reduced 
+pressure of 9.99,
 and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
@@
 setting the second virial coefficients for the error function potential 
-and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$.
+and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$ 
+at each temperature.
 To implement cDFT we used Gaussian distribution functions centered at 

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

@@
 In the simplest case, the spheres are treated as having only a repulsive 
-potential. This model, called the hard-sphere fluid, has served as the 
+potential that rises instantly to infinity when the spheres make contact. 
+This model, called the hard-sphere fluid, has served as the 
 primary reference system for fluids. This is due to the fact that the 
 short-range repulsion between atoms is much stronger than their long-range 
-attraction, and so the repulsive potential, which is descriptive of 
-hard spheres, dominates the interactions between atoms.
+attraction, and thus dominates the interactions between atoms.
 
@@
 radii to model spheres with a more realistic softly fading repulsive 
-potential similar to that between neutral atoms
+potential similar to that between neutral atoms.
 %portion of the Leonard-Jones potential
-which is known as the Weeks-Chandler-Anderson model fluid~\cite{andersen1971relationship} . 
+Their is known as the Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship} . 
 Models continue to be developed to characterize fluids with particular 
-properties, or to address particular problems like surface tension, 
+properties, or to address problems like surface tension, 
 interfaces, adsorption, wetting and confinement.
@@
 by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
-Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium} amoung others. A review
+Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium} among others. A review
 of the development of cDFT is found in Fundamentals of Inhomogeneous fluids~\cite{inhomogeneousfluids}.
@@
 To use classical Density Functional Theory (cDFT), an expression for the free energy 
-as a functional of a spatially varying average single-particle number density 
+as a functional of a spatially varying average single-particle number density $n(\vec r)$
 %that varies spatially 
@@
 \begin{displaymath}
-   \theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
+   \Theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
 \begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
 \end{equation}
+
 FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
@@
 diverged when the fluid was highly confined.  
-Tarazona later solved this problem by introducing a tensor version of 
+Tarazona later solved this problem by introducing a tensor version of $\Phi_3$,
 one of the energy density terms 
-in the functional~\cite{tarazonaphi3, santos2012phi3} 
+in the functional,~\cite{tarazonaphi3, santos2012phi3} 
 which took into account the non-spherical, and thus 
@@
 %rather than instantaneously).
-Schmidt applied his Soft Fundamental Measure Theory to various fluids 
-with soft, short-range repulsive and attractive potentials, and developed 
+Schmidt applied his Soft Fundamental Measure Theory to various fluids, 
+%with soft, short-range repulsive and attractive potentials, 
+and developed 
 an error function model 
@@
 %, similar to that exhibited by a WCA fluid. 
-For some more recent developments, see the review of new developments in 
-classical Density Functional Theory by Evans ~\cite{evansreview} (2016), 
+For some more recent developments in classical Density Functional Theory, 
+see the review by Evans ~\cite{evansreview} (2016), the book 
 Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
-and a generalization of the Barker-Henderson model to inhomogeneous 
-fluids by Tschopp and coworkers~\cite{BHextended} (2020).
+and the paper by Tschopp and coworkers~\cite{BHextended} (2020) in which 
+they extend the Barker-Henderson model to inhomogeneous 
+fluids from first principles.
 
@@
 to phase transitions, and freezing, but not until recently has there
-been good progress. Ramakrishnan and Yussouff studied freezing by viewing a 
-crystal as a highly inhomogeneous liquid~\cite{ramakrishnanandyussouff} (1979). 
-Oettel and fellow researchers studied hard-sphere crystals with fully developed FMT~\cite{Oettel2010} (2010). 
-Mortazavifar and Oettel studied the fluid and crystal phases of the 
-Asakura-Oosawa model fluid with FMT~\cite{asakuraoosawa} (2016). 
+been promising progress. In 1979 Ramakrishnan and Yussouff studied 
+freezing by viewing a 
+crystal as a highly inhomogeneous liquid~\cite{ramakrishnanandyussouff}. 
+In 2010 Oettel and fellow researchers studied hard-sphere crystals with 
+fully developed FMT~\cite{Oettel2010}, and 
+in 2016 FMT was used to find the fluid and crystal phases of the 
+Asakura-Oosawa model fluid~\cite{asakuraoosawa}. 
 Lutsko and Lam studied studied crystallization of the Leonard-Jones 
-fluid~\cite{lutsko} (2018), and Lutsko and Schoonen looked at cDFT applied 
-to the solid state~\cite{lutskoschoonen2020} (2020).
+fluid~\cite{lutsko} in 2018, and Lutsko and Schoonen looked at cDFT applied 
+to the solid state~\cite{lutskoschoonen2020} in 2020.
 In all these studies the hard-sphere fluid has been used as the reference 
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Fluid and solid densities at the solid-fluid phase 
+    \caption{Fluid and solid densities at the fluid-solid phase 
     transition are identified using Maxwell's Double Tangent Construction}
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
-    \caption{Pressure is found at point of intersection which occurs 
-    when $\mu_F=\mu_S$.}
+    \caption{Pressure is found at the point of intersection where $\mu_F=\mu_S$.}
     \label{fig:GibbsvsP}
@@
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
-    the number densities of the fluid and the solid during a phase 
+    the number densities of the fluid and the solid at a phase 
     transition. The gray region is where different phases coexist in 
@@
 that does not depend on explicitly taking into account the interaction 
-that each sphere with every other sphere in the fluid (as required by 
+that each sphere has with every other sphere in the fluid (as required by 
 Equation~\ref{eq:Fexcess-simplified}).
@@
 for a hard-sphere fluid the Mayer function, $f(r)$, can be written in terms 
-of a backward Heaviside step function which, for a hard-sphere fluid 
+of a backward Heaviside step function $\Theta$ which, for a hard-sphere fluid 
 composed of a mixture of spheres of radius $R_i$ and $R_j$, is
@@
 spheres of radius R, is given by
-\begin{align}
+\begin{align} \label{mayerconvolution}
     f(r_{ij})=-\Theta(r_{ij}-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
@@
 \end{align}
-which are needed for the relation to be satisfied.
+which are needed for the relation in Equation~\ref{mayerconvolution} to be satisfied.
 %%SAVE the following - connects to convolution! ------------------------
@@
 %
-The convolution of the weight functions can be written as
+The convolution of two weight functions can be written as
 \begin{align}
@@
 \end{align} 
-can then be written as 
+which can be written as 
 \begin{align}      
@@
 order terms in the virial expansion can all be written in terms of an 
-"exclusion volume" which is given by Equation~\ref{exclusion-volume}, 
-and an "overlap volume".
+"exclusion volume" (given by Equation~\ref{exclusion-volume}), 
+and an "overlap volume" which is the volume of intersection between two 
+overlapping spherical regions.
 The exclusion volume is the spherical volume about the center of one sphere 
 which the center of another sphere can not enter, and is used to describe
-terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. The overlap volume is 
-the volume of intersection between two spheres that overlap, and is used in conjunction 
-with the exclusion volume to describe higher order terms in the expansion
-in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988}. 
+terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. Used in conjunction 
+with the overlap volume, the higher order terms in the expansion
+in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988} can be described. 
 %
@@
    \caption{On the left are shown four examples of possible positions 
-   that an atom (blue sphere) might have within a fixed volume 
+   that an atom (blue sphere) might have within a volume 
    containing eight cells. Each cell has one lattice site at the center 
@@
    give rise to the same Gaussian distribution statistically, 
-   and the same number density.}
+   and the same number density. In reality, the fraction of vacancies will
+   be very small.}
    \label{fig:SameStatPic}
@@
   is the region of coexistence where some fluid and solid exist together 
-  at varying proportions. White lines show the boundaries of the coexistence 
-  region generated from Monte-Carlo simulations of the WCA fluid.}
+  in varying proportions. White lines show the boundaries of the coexistence 
+  region generated from Monte-Carlo simulations.}
   \label{fig:Phase_Diagram_of_T_vs_n}
@@
   \caption{Phase Diagram of the reduced pressure verses the reduced 
-  temperature. The blue region is solid, the red region is fluid,
-  and the black line shows the phase boundary where the solid and fluid
+  temperature. The blue region is solid, the red region is fluid.
+  The black line shows the phase boundary where the solid and fluid
   phases coexist. The white line marks the solid-fluid phase boundary  

### 2021-01-12 papers/thesis-kirstie/thesis.tex
edits

@@
 a minimum at equilibrium means that the equilibrium free energy of the 
-system ${f_{eq}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
+system $f_{eq}[n_{eq}(\vec r)]$, and its corresponding equilibrium 
 number density profile $n_{eq}(\vec r)$, can be found simply by varying 
@@
 \end{align}
-Rosenfeld also recognized that the Heaviside function can be written as 
-a convolution of the weight functions which, for a fluid composed of 
-spheres of radius R, is given by
+where R is either $R_i$ or $R_j$ accordingly. Rosenfeld also recognized that the 
+Heaviside function can be written as a convolution of the weight functions 
+which, for a fluid composed only of spheres of radius R, is given by
 \begin{align} \label{mayerconvolution}
@@
                 &\approx k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
-                %&\text{or replacing u notation with r} \nonumber \\
+                &\text{Or, replacing u notation with r} \nonumber \\
                 &\approx k_BT\int \Phi(\vec r)~d\vec r             
 \end{align}
-where the weighted densities $n_\alpha$ all depend on the spatially varying 
+The weighted densities $n_\alpha$ all depend on the spatially varying 
 number density profile $n(\vec r)$, and have the form
 \begin{align}
-    n_\alpha(\vec u)=\int n(\vec {r}~')w_\alpha(\vec u-\vec {r}~')~d\vec {r}~'  
+    n_\alpha(\vec r) &= \int n(\vec {r}~')w_\alpha(\vec r-\vec {r}~')~d\vec {r}~'  \label{weighted_densities}  \\
+                     &= n\otimes w_\alpha(\vec r)
     %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
@@
 %
-The resulting expression for the excess Helmholtz free energy with an 
-updated version of $\Phi_3$ that was later constructed by Tarazona is given by
+The resulting expression for the excess Helmholtz free energy with Tarazona's $\Phi_3$
+(see Introduction) is given by
 \begin{align}
@@
 \end{align}
-where in addition to the weight functions already defined,  
+%where in addition to the weight functions already defined,  
+where the weighted densities are given by Equation~\ref{weighted_densities} with 
+weight functions
 \begin{align} 
@@
 \end{align}
-a tensor weight function is given by
+and a tensor weight function given by
 \begin{align}\label{eq:tensorweight-methods}

### 2021-01-13 papers/thesis-kirstie/thesis.tex
edits

@@
     \centering
-    \includegraphics[height=6cm]{figs/P-T_Diagram.pdf}
+    \includegraphics[height=7cm]{figs/P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
 constructed. 
-The key idea behind classical Density Functional Theory (cDFT) is that 
+The key idea behind classical Density Functional Theory is that 
 free energy can be written as a functional (a function of a function) of 
@@
 profile $n(\vec r)$
-\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi=\frac{V_{ext}}{N}$ is an externally applied potential energy 
-$V_{ext}(\vec r)$ per particle, and an intrinsic Helmholtz free energy 
+\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
+where $\phi_{ext}=\frac{V_{ext}}{N}$ is an externally applied potential energy 
+$V_{ext}(\vec r)$ per particle, an intrinsic Helmholtz free energy 
 is defined as
-\begin{equation}F_{int} = F - \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}
-
+\begin{equation}F_{int} = F - \int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}\end{equation}  
+and 
+\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}%
 The Grand free energy is applicable to systems with a varying number of 
@@
 particles will be of interest in which case
-\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
-and so the Grand free energy is constrained
+%\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation} and so
+the Grand free energy is constrained
 \begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
@@
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
-This result shows that the natural variables of the Helmholtz free energy 
-per particle $f_N$ are temperature, volume and one over the number of 
-particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
-\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
-Taking the differential
+%This result shows that the natural variables of the Helmholtz free energy 
+%per particle $f_N$ are temperature, volume and one over the number of 
+%particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
+%\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
+%Taking the differential
+which can be set equal to 
 \begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT +\frac{\partial f_N}{V}dV+ \frac{\partial f_N}{\partial \left(\frac{1}{N}\right)}d\left(\frac{1}{N}\right)\end{displaymath}
-and using the relation
+Using the relation
 \begin{displaymath}d\left(\frac{1}{n}\right)=Vd\left(\frac{1}{N}\right)+\frac{1}{N}dV\end{displaymath}
@@
 \begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT + \frac{\partial f_N}{\partial \frac{1}{n}}d\left(\frac{1}{n}\right)\end{displaymath}
-
 \begin{equation}f_N(T,\frac{1}{n})\end{equation}
+This result shows that temperature and the inverse number density, 
+or volume per particle, are natural variables of the Helmholtz free energy 
+per particle $f_N$.
+%which will thus be minimized when these natural variables are held fixed.
 
@@
 experience spontaneous processes that increase the total entropy of a 
-system and its surroundings until equilibrium is reached, 
-$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
+system and its surroundings until equilibrium is reached
+\begin{align}
+    \Delta{S}_{System + Surroundings} \geq 0 
+\end{align}
+Here "surroundings" means 
 the entire universe minus the system. This, together with the conservation 
@@
 (or gained) by the system is gained (or lost) by the surroundings 
-$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
+\begin{align}
+    \Delta{U}_{System}=-\Delta{U}_{Surroundings} 
+\end{align}
+gives the result that 
 changes in the free energy 
@@
 when the natural variables of the free energy are held fixed. 
-
-For our case, with a fixed temperature and number density we get
+For our case, the temperature and inverse number density are the natural 
+variables, and so holding these fixed we get
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
@@
 \begin{equation}
-  df_{System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+  df_{N~System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
 \end{equation}
@@
 %~\cite{schroeder}. 
-and so the free energy tends toward a minimum. 
+Since the changes in the total entropy are always positive, the changes
+in the free energy are negative and the free energy tends toward a minimum. 
 Details of this procedure be found in Thermal Physics by Schroeder~\cite{schroeder}, 
@@
 a momentum $\vec{p}$ and a position $\vec{r}$, and is given by
-\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
+\begin{align}  \label{Z_total}
+     Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}\int{dQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}
+\end{align}
 where
-\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~and~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
-\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~and~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
+\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~and~each~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
+\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~and~each~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
 Writing the Hamiltonian, H in terms of its component kinetic and potential energies gives
@@
 interatomic and external potential energies depend only on position, 
-the integral in equation (8) can be separated into two parts 
-\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
+the integral in Equation~\ref{Z_total} can be separated into two parts 
+\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
 Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume, and 
-\begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
-the partitian function can be expressed as
+\begin{equation}V^N=\int dQ\end{equation} 
+the partitian function can then be expressed as
 \begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
@@
 \begin{align} \label{eq:Zideal}
-    Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}
+    Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}
 \end{align}
 \begin{align} \label{eq:Zexcess}
-    Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
+    Z_{excess}=\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
 \end{align} 
@@
 The excess partitian function $Z_{excess}$ accounts for the interaction 
-of all the atoms according to their interatomic potential. Using 
-Equation~\ref{Zmultiply}, the free energy can also be broken into two parts.
-\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
+of all the atoms according to their interatomic potential.
+
+Using Equation~\ref{Zmultiply}, the free energy can also be broken into two parts.
+%\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
 \begin{align} \label{eq:Finparts}
-    F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})
-\end{align}
-\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}
-
-Putting Equation~\ref{eq:Zideal} into Equation~\ref{eq:Finparts}, 
-the ideal Helmholtz free energy is
-\begin{equation}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation}
-
+     F &= -k_{B}T\ln(Z_{ideal}Z_{excess})  \\
+       &= -k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess}) \\
+       &= F_{ideal} + F_{excess}
+\end{align}
+%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}
+
+The ideal Helmholtz free energy is given by 
+%(put Equation~\ref{eq:Zideal} into Equation~\ref{eq:Finparts})
+\begin{align}
+    F_{ideal} &= -k_{B}T\ln(Z_{ideal}) \\
+              &= -k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\right)} 
+\end{align}  
 \begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
@@
 this becomes
-\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
+\begin{equation}F_{ideal}= k_BT[N\ln\left(\frac{N\Lambda^{3}}{V}\right)-N]\end{equation} 
 Rewriting the equation in terms of the number density $n=\frac{N}{V}$, 
@@
 Equation~\ref{eq:Fideal} becomes
-\begin{equation}\label{eq:Fideal-n(r)}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]d\vec{r}}\end{equation}
-
-Putting Equation~\ref{eq:Zexcess} into Equation~\ref{eq:Finparts}, 
-the excess Helmholtz free energy is written most generally as
-\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
+\begin{equation}\label{eq:Fideal-n(r)}{F_{ideal}[n(\vec{r})]= k_BT\int_{Volume}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]~d\vec{r}}\end{equation}
+
+%Putting Equation~\ref{eq:Zexcess} into Equation~\ref{eq:Finparts}, 
+The excess Helmholtz free energy is written most generally as
+\begin{align}
+      F_{excess} &= -k_{B}T\ln(Z_{excess}) \\
+                 &= -k_BT\ln{\left(\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)} \label{Fexcess-mostgeneral}
+\end{align} 
 This integral can be greatly simplified if it can be assumed that the 
@@
 where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
-The function $f_{ij}$ the Mayer function given by
+The function $f_{ij}$ is the Mayer function given by
 \begin{align}\label{eq:mayerfunction}
@@
 \begin{align} \label{eq:Fexcess-simplified}
-    F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)}
+    F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)}
 \end{align}
@@
     \centering
-    \includegraphics[height=7cm]{figs/diagrammic.pdf}
+    \includegraphics[height=6.5cm]{figs/diagrammic.pdf}
     \caption{Two-pair interactions $f_{ij}f_{kl}$ are shown for N=4 atoms. 
@@
 two atoms in the pair. 
-With a careful counting of the resulting number of such pairs, the excess 
-free energy can then be written~\cite{schroeder} as
-\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}\approx k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
+With a careful counting of the resulting number of such pairs, the 
+low-density excess free energy for a homogeneous fluid can then be 
+written~\cite{schroeder} as
+\begin{align}\label{eq:Fexcess}
+    F_{excess} &\approx k_BTn^2\int d\vec {r}~'\int d\vec {r}~"\left(-\frac{1}{2}f(|\vec {r}~'-\vec {r}~"|)\right) \\
+               &\approx k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V
+    %\frac{F_{excess}}{N} &\approx k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)
+\end{align} 
 in the thermodynamic limit of large N. 
@@
 as the second virial expansion coefficient typically called $B_2$. 
-It is important that an expression for the excess Helmholtz free energy 
-correctly reduces to this result in the low-density limit. 
+A correct expression for the excess Helmholtz free energy should 
+reduce to this result in the low-density limit. 
 Outside of the low density limit, it is quite difficult to evaluate the 
@@
 found by varying the number density profile
-through all possible configurations at some fixed temperature
-until the free energy is minimized.
+at some fixed temperature until the free energy is minimized.
 
@@
 $V(r_{ij})$
+\begin{align}
+    V(r_{ij})=\left\{ \begin{array}{rc} \infty & 0<r_{ij} \leq R_i+R_j \\ 0  & r_{ij}> R_i+R_j \end{array}\right.
+\end{align}
 into the equation for the Mayer function
@@
 \end{align} 
-where
-\begin{align}
-    V(r_{ij})=\left\{ \begin{array}{rc} \infty & 0<r_{ij} \leq R_i+R_j \\ 0  & r_{ij}> R_i+R_j \end{array}\right.
-\end{align}
+%where
+
 %
@@
         V &= \int_{Volume} \Theta(r_{ij}-R)d\vec r_{ij} \\
-        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  \\
+        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  
+\end{align}
+\begin{align}
         R &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}} d\vec r_{ij} \\
@@
 \end{align} 
-where the weights are  
+where the scalar weights are given by
 \begin{align} 
@@
 %excess chemical potential can be expanded.
-%
+
 The convolution of two weight functions can be written as
@@
 \begin{align}      
-     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}  \\
+     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}                                                              
+\end{align} 
+\begin{align} 
                 &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
@@
                 &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
-                &- \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u
-\end{align} 
-\begin{align}                
-                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}\\                
+                &- \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u  \\
+                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}  \\                
                 &\approx k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
@@
                 &\text{Or, replacing u notation with r} \nonumber \\
+                &\approx k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r  \\
                 &\approx k_BT\int \Phi(\vec r)~d\vec r             
@@
       w_1 &= \frac{w_2}{4\pi r}  \\
-      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  \\
+      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  
+\end{align}
+\begin{align}      
       \vec {w}_2 &= w_2~\hat r \\
-      \vec {w}_1 &= w_1~\hat r
-\end{align}
-and a tensor weight function given by
-\begin{align}\label{eq:tensorweight-methods}
-     \overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
-\end{align}
+      \vec {w}_1 &= w_1~\hat r \\
+      \overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
+\end{align}
+%and a tensor weight function given by
+%\begin{align}\label{eq:tensorweight-methods}
+     %\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
+%\end{align}
 %The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
@@
 $w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
-The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
-the relation
-\begin{align}\label{w2_w3_relation}
-    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
-\end{align}
-%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-Taking the derivative of the Mayer function provides 
+The derivative of the Mayer function provides 
 another important relation expressed in terms of $w_{2}(r)$ 
@@
 \end{align} 
+The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+the relation
+\begin{align}\label{w2_w3_relation}
+    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+\end{align}
+%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
 Thus, all the weight functions, and the above relation are determined
@@
 functions, and Equation~\ref{mayerder=w2convolution} 
-for the derivative of the Mayer function 
+%for the derivative of the Mayer function 
 the same, but replacing $w_{3}(r)$ with a weight function appropriate 
@@
   \centering
-  \includegraphics[height=8cm]{plot_alpha.pdf}
+  \includegraphics[height=7cm]{plot_alpha.pdf}
   \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
@@
   \centering
-  \includegraphics[height=8cm]{xi_fromB2.pdf}
+  \includegraphics[height=7cm]{xi_fromB2.pdf}
   \caption{Shows the temperature dependence of $\Xi$ generated from 

### 2021-01-13 papers/thesis-kirstie/thesis.tex
edits

@@
 
-A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
-the density will change abruptly at the phase transition.
-The particular structure and density that result are affected 
+%A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
+A liquid will freeze when its free energy is lower than the free energy
+that a solid would have under the same conditions.
+At the phase transition the density will change abruptly, and
+the particular structure and density that result will be affected 
 by the interatomic potential. When repulsive, soft sphere fluids 

### 2021-01-14 papers/thesis-kirstie/thesis.tex
edits

@@
 spheres with softly fading short-range repulsive potentials which he modeled by making
-the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
+the radius of hard spheres temperature dependent~\cite{rowlinson1964statistical}. 
 This reflects the idea that soft-spheres penetrate more deeply into 
@@
 
-FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
-SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
-and for the direct correlation function approximation given by Percus-Yevick. 
+FMT gives the same result for the excess Helmholtz free energy of the 
+hard-sphere fluid as SPT does, from which it can be 
+constructed~\cite{rosenfeld1989,Hansen,santos2012phi3}. It also gives the same 
+result for the direct correlation function approximation given by Percus-Yevick. 
 However, it did not predict freezing due to one term in the expression that
@@
 Classical density functional theory has also been applied 
-to phase transitions, and freezing, but not until recently has there
-been promising progress. In 1979 Ramakrishnan and Yussouff studied 
+to phase transitions, and freezing. In 1979 Ramakrishnan and Yussouff studied 
 freezing by viewing a 
@@
 as a classical fluid where the spheres interact with each other according 
-to their electrostatic interatomic potential.
+to their interatomic potential.
 
@@
 Above $\sigma$ the potential becomes negative indicating a net attraction 
-between the atoms. The net attraction is maximized at a distance of 
+between the atoms. The net attractive energy is maximized at a distance of 
 r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of 
 the sphere given by $R={2^{-\frac{5}{6}}}\sigma$. At this point the depth of
-the potential well, given by epsilon $\epsilon$, is a maximum. As r increases 
+the potential well, given by epsilon $\epsilon$, is a maximum, and the force
+between the spheres is zero. As r increases 
 further, the net attraction decreases and the potential approaches zero
@@
     \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
-    \caption{Lennard-Jones potential broken into two parts for $\sigma=1$, $\epsilon=1$}
+    \caption{Lennard-Jones potential broken into two parts}
     \label{fig:LJ_potential_2parts}
@@
     \includegraphics[height=7cm]{WCA_Potential.pdf}
-    \caption{WCA potential for $\sigma=1$, $\epsilon=1$}
+    \caption{WCA potential}
     \label{fig:WCA_potential}
@@
 \end{align} 
-The WCA potential decreases from as the distance between the 
+The WCA potential decreases as the distance between the 
 centers of the atoms r increases, and becomes, and remains, essentially 
@@
     \includegraphics[height=3.5cm]{figs/TwoSpheres.pdf} 
-    \caption{(Left) The WCA potential is a minimum at r=2R. 
+    \caption{(Left) The WCA potential is zero at r=2R. 
              (Right) The WCA potential increases rapidly for $r<2R$.}
@@
 %A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
-A liquid will freeze when its free energy is lower than the free energy
-that a solid would have under the same conditions.
+A liquid will freeze when the free energy it would have as a solid is lower 
+than the free energy that it has as a liquid at the same density.
 At the phase transition the density will change abruptly, and
@@
 the atoms in a WCA fluid do not exhibit attraction there will not be a 
-dividing line between the gas and liquid phases, there will just simply
+dividing line between the gas and liquid phases, there will simply
 be a fluid.
@@
 The pressure can be related to the Helmholtz free energy F, which is minimized 
-for a system at fixed temperature, volume, and number of particles as the 
+for a system at fixed temperature, volume, and number of particles, as the 
 entropy of the system and its surroundings added together is maximized. The negative of 
@@
 which can be set equal to 
-\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT +\frac{\partial f_N}{V}dV+ \frac{\partial f_N}{\partial \left(\frac{1}{N}\right)}d\left(\frac{1}{N}\right)\end{displaymath}
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}\bigg|_{V,\frac{1}{N}}dT 
++\frac{\partial f_N}{V}\bigg|_{T,\frac{1}{N}}dV + \frac{\partial f_N}
+{\partial \left(\frac{1}{N}\right)}\bigg|_{T,V}d\left(\frac{1}{N}\right)\end{displaymath}
 Using the relation
@@
 we get
-\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT + \frac{\partial f_N}{\partial \frac{1}{n}}d\left(\frac{1}{n}\right)\end{displaymath}
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}\bigg|_{\frac{1}{n}}dT 
++ \frac{\partial f_N}{\partial \frac{1}{n}}\bigg|_{T}d\left(\frac{1}{n}\right)\end{displaymath}
 \begin{equation}f_N(T,\frac{1}{n})\end{equation}
@@
 \end{align} 
-The ideal partitian function $Z_{ideal}$ is associated with an ideal gas 
+The ideal partitian function $Z_{ideal}$ is that of an ideal gas 
 which exhibits no interaction between the atoms - a condition approached 
@@
 \end{align} 
-This integral can be greatly simplified if it can be assumed that the 
+This integral can be somewhat simplified if it can be assumed that the 
 potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
@@
 
-\begin{equation}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{equation}
-\begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
-where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
-The function $f_{ij}$ is the Mayer function given by
+\begin{equation}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},
+\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}
+(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{equation}
+\begin{equation}{=-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
+%where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
+where function $f_{ij}$ is the Mayer function given by
 \begin{align}\label{eq:mayerfunction}
@@
 a more general solution.
-At low densities, it is not likely to find two or more atoms in close 
+At low densities, it is not likely to find more than two atoms in close 
 proximity, and so only interactions between individual pairs of spheres, 
@@
 Soft Fundamental Measure Theory (SFMT) is a method for forming an 
-expression for the excess Helmholtz free energy of a soft-sphere fluid.
+approximation for the excess Helmholtz free energy of a soft-sphere fluid.
 In the next chapter, 
@@
 order terms in the virial expansion can all be written in terms of an 
-"exclusion volume" (given by Equation~\ref{exclusion-volume}), 
-and an "overlap volume" which is the volume of intersection between two 
+``exclusion volume" (given by Equation~\ref{exclusion-volume}), 
+and an ``overlap volume" which is the volume of intersection between two 
 overlapping spherical regions.
@@
 of positions about the lattice point. 
-Atoms are always in motion, so when talking about "the location" 
+Atoms are always in motion, so when talking about ``the location" 
 of an atom, we are talking about a time averaged position, so that it 
@@
 density. The number density is the number of atoms per volume, whereas
-the number density profile describes how much "atom" will be found at various 
+the number density profile describes how much ``atom" will be found at various 
 points in the cell. 
@@
 \begin{align}
-    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
+    n(\vec r)=~\left<\sum_{i=1}^N\delta(\vec r - \vec r_i)\right>
 \end{align} 
@@
 \begin{align}
-    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
+    n=\left<n(\vec{r})\right>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
 \end{align}
@@
 
-\begin{figure}
-   \centering
-   \includegraphics[height=5cm]{InclusionRadius.png}
-   \caption{Gaussians are centered at each lattice point. After a radius 
-   of a few Gaussian widths about each lattice point, the Gaussians go to 
-   zero and can be neglected.}
-   \label{fig:InclusionRadius}
-\end{figure}
+%\begin{figure}
+   %\centering
+   %\includegraphics[height=5cm]{InclusionRadius.png}
+   %\caption{Gaussians are centered at each lattice point. After a radius 
+   %of a few Gaussian widths about each lattice point, the Gaussians go to 
+   %zero and can be neglected.}
+   %\label{fig:InclusionRadius}
+%\end{figure}
 
@@
 using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
-with a uniform number density profile, n that does not depend on $\vec r$. 
+with a uniform number density profile, $n$ that does not depend on $\vec r$. 
 The crystal excess Helmholtz free energy is computed  
@@
 $\vec r~' -\vec R$ according to a Gaussian distribution.
-Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
-about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
-will be taken into account as Gaussians 
-farther away will not contribute much.
-The equation for the contribution to the weighted density from a 
-Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
-\begin{equation}{n_\alpha(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
-and is implemented computationally in the form
-\begin{equation}{n_\alpha(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
-which can be written as
-\begin{equation}{n_\alpha(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_\alpha(|\vec{r}-\vec{r}~'|)}\end{equation} 
-NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
-generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
-is equal to a probability density which when multiplied by dV gives 
-the probability $P_{r'}$ at which the value of $w_\alpha(|\vec{r}-\vec{r}~'|)$ 
-will occur. Using the relation
-\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
-$n_\alpha(\vec{r})$ can be expressed as the average value of $(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)$:
-\begin{equation}{n_\alpha(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
-This process is repeated to get the contributions to the weighted density 
-$n_\alpha(\vec r)$ from the Gaussians around each lattice point which begin to 
-merge together as the width of the Gaussians increases.
-All the contributions are then added together 
-to get the final value of the weighted density at some point $\vec r$.
-
-The result can be made more accurate by increasing NumPoints (the 
-number of vectors randomly generated), and by using antithetic variates.
-To use antithetic variates, each randomly generated  vector is paired 
-with a vector of equal magnitude, but opposite in direction. 
-First order error is reduced by averaging the two results that depend on 
-the vectors, one of which may be erroneously higher than it should be 
-while the other is erroneously lower than it should be.
-
-The standard error of the mean (SEM) is given in general by
-\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
-To compute the error, the standard deviation of weight function $w_3$ is used 
-\begin{equation}SD=\sqrt{\overline{w_3^2}-(\overline{w}_3)^2}\end{equation} 
-\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{\text{NumPoints}}}}\end{equation} 
-NumPoints, the number of vectors randomly generated, is increased until
-\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
-where the maximum value that the weighted density $n_3$ 
-can properly have is 1.
+%
+%ERRORS IN HERE... 
+%Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
+%about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
+%will be taken into account as Gaussians 
+%farther away will not contribute much.
+%The equation for the contribution to the weighted density from a 
+%Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
+%\begin{equation}{n_\alpha(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
+%and is implemented computationally in the form
+%\begin{equation}{n_\alpha(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
+%which can be written as
+%\begin{equation}{n_\alpha(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_\alpha(|\vec{r}-\vec{r}~'|)}\end{equation} 
+%NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
+%generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
+%is equal to a probability density which when multiplied by dV gives 
+%the probability $P_{r'}$ at which the value of $w_\alpha(|\vec{r}-\vec{r}~'|)$ 
+%will occur. Using the relation
+%\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
+%$n_\alpha(\vec{r})$ can be expressed as the average value of $(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)$:
+%\begin{equation}{n_\alpha(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
+%This process is repeated to get the contributions to the weighted density 
+%$n_\alpha(\vec r)$ from the Gaussians around each lattice point which begin to 
+%merge together as the width of the Gaussians increases.
+%All the contributions are then added together 
+%to get the final value of the weighted density at some point $\vec r$.
+
+%The result can be made more accurate by increasing NumPoints (the 
+%number of vectors randomly generated), and by using antithetic variates.
+%To use antithetic variates, each randomly generated  vector is paired 
+%with a vector of equal magnitude, but opposite in direction. 
+%First order error is reduced by averaging the two results that depend on 
+%the vectors, one of which may be erroneously higher than it should be 
+%while the other is erroneously lower than it should be.
+
+%The standard error of the mean (SEM) is given in general by
+%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
+%To compute the error, the standard deviation of weight function $w_3$ is used 
+%\begin{equation}SD=\sqrt{\overline{w_3^2}-(\overline{w}_3)^2}\end{equation} 
+%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{\text{NumPoints}}}}\end{equation} 
+%NumPoints, the number of vectors randomly generated, is increased until
+%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
+%where the maximum value that the weighted density $n_3$ 
+%can properly have is 1.
+%end ERRORONEOUS SECTION
 
@@
 In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
-$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
-$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
-In all cases $\sigma=1$ and $\epsilon=1$. 
+$T^*\equiv\frac{kT}{\epsilon}$ is plotted against the reduced number density 
+$n^*\equiv n\sigma^3$ for temperatures of $T^*$ = 0.5 to 3.0. 
+%In all cases $\sigma=1$ and $\epsilon=1$. 
 It is seen that in general, a fluid exists at lower number densities, 
@@
 and shifted a bit toward smaller densities. For example, at a temperature 
-of kT=1, the cDFT fluid transition reduced number density is 0.96, 
+of $T^*$=1, the cDFT fluid transition reduced number density is 0.96, 
 and the solid transition reduced number density is 1.02, giving a spread 
@@
 Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
-$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
-for temperatures of kT = 0.5 to 3.0. 
+$p^*\equiv\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
+for temperatures of $T^*$ = 0.5 to 3.0. 
 The black line marks the phase boundary between the solid and fluid regions. 
@@
 discrepancy grows as the temperature increases. For example, at a 
-temperature of kT=1, the phase transition reduced pressure predicted by 
+temperature of $T^*$=1, the phase transition reduced pressure predicted by 
 cDFT is 15.24, whereas the Monte-Carlo simulation predicted a reduced 
 pressure of 9.99,
-and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
+and at a temperature of $T^*$=3, cDFT gives a value of 77.27 whereas the 
 Monte-Carlo simulation predicts 43.80.

### 2021-01-15 papers/thesis-kirstie/thesis.tex
edits from david

@@
 Energy defined as
-\begin{equation}\Phi=F-\mu{N}\end{equation}
+\begin{equation}\Omega=F-\mu{N}\end{equation}
 which can then be expressed in terms of a varying number density 
 profile $n(\vec r)$
-\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where $\phi_{ext}=\frac{V_{ext}}{N}$ is an externally applied potential energy 
-$V_{ext}(\vec r)$ per particle, an intrinsic Helmholtz free energy 
-is defined as
+\begin{equation}\label{GrandFE}\Omega= F_{int} +\int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
+where
+\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}%
+and the intrinsic Helmholtz free energy is defined as
 \begin{equation}F_{int} = F - \int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}\end{equation}  
-and 
-\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}%
+%The quantity $\phi_{ext}$
+where $\phi_{ext}=\frac{V_{ext}}{N}$ and $V_{ext}(\vec r)$
+is the externally applied potential energy given in Equation~\ref{Hamiltonian}. 
 The Grand free energy is applicable to systems with a varying number of 
@@
 the Grand free energy is constrained
-\begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
+\begin{equation}\label{GrandFE}\Omega_{constrained}= F[n(\vec r)]-\mu N\end{equation}
 The dependence on $n(\vec r)$ is now only present in the expression for 
@@
 Writing the Hamiltonian, H in terms of its component kinetic and potential energies gives
-\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
+\begin{equation}\label{Hamiltonian}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
 where the translational kinetic energy of the N atoms is given by 
@@
 \end{align} 
-The ideal partitian function $Z_{ideal}$ is that of an ideal gas 
-which exhibits no interaction between the atoms - a condition approached 
-by a gas at low density where the atoms rarely encounter one another. 
+The ideal partitian function $Z_{ideal}$ is that of an ideal gas.
+%an ideal gas which exhibits no interaction between the atoms - a condition approached 
+%by a gas at low density where the atoms rarely encounter one another. 
 The excess partitian function $Z_{excess}$ accounts for the interaction 
@@
                 &- \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u  \\
-                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}  \\                
-                &\approx k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
-                &\approx k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
-                &\text{Or, replacing u notation with r} \nonumber \\
-                &\approx k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r  \\
-                &\approx k_BT\int \Phi(\vec r)~d\vec r             
+                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}                  
+                %&\approx k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
+                %&\approx k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
+                %&\text{Or, replacing u notation with r} \nonumber \\
+                %&\approx k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r  \\
+                %&\approx k_BT\int \Phi(\vec r)~d\vec r             
 \end{align}
@@
 The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
-All of the weight functions can be expressed in terms of 
+To model a fluid with "soft" spheres which have an interatomic potential 
+that decreases smoothly from infinity over a distance of 2R rather than instanteously 
+at 2R like that of hard spheres, Schmidt noted that 
+all of the weight functions can be expressed in terms of 
 $w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
-The derivative of the Mayer function provides 
-another important relation expressed in terms of $w_{2}(r)$ 
-%that will be important 
-%to keep in tact when moving to SFMT~\cite{schmidt2000fluid}.
+He also noted that the derivative of the Mayer function 
+%provides another important relation 
+can be expressed in terms of $w_{2}(r)$, 
 \begin{align}\label{mayerder=w2convolution}
-     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
-\end{align} 
-The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+     \frac{df(r)}{dr} = w_2 \otimes w_2(r) = \int{w_2(r')w_2(r-r')dr'}
+\end{align} 
+and that the weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
 the relation
@@
 %which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-Thus, all the weight functions, and the above relation are determined
-by $w_{3}(r)$.
-Soft Fundamental Measure Theory extends FMT to fluids 
+%Thus, all the weight functions, and the above relation are determined by $w_{3}(r)$.
+Schmidt developed Soft Fundamental Measure Theory (SFMT) which extends FMT to fluids 
 with soft potentials by keeping the relationships between the weight 
@@
 
+Unlike the WCA potential, the error function potential $V_{erf}(r)$ is temperature dependent.
+To compensate for this temperature dependence so that our functional better approximates
+a WCA potential, we make the parameters $\alpha$ and $\Xi$ functions of temperature as well. 
 The parameter $\alpha(T)$ is derived by setting 
 the error function potential $V_{erf}(r)$ 
-equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$. 
+equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$, since 
+the derivative of the mayer function peaks at this value. 
 \begin{align}

### 2021-01-15 papers/thesis-kirstie/thesis.tex
edits

@@
 constructed~\cite{rosenfeld1989,Hansen,santos2012phi3}. It also gives the same 
-result for the direct correlation function approximation given by Percus-Yevick. 
+result for the direct correlation function found by Percus-Yevick. 
 However, it did not predict freezing due to one term in the expression that
@@
 rapidly approaches positive infinity corresponding to infinite repulsion. 
-Above $\sigma$ the potential becomes negative indicating a net attraction 
-between the atoms. The net attractive energy is maximized at a distance of 
+Above $\sigma$ the potential becomes negative indicating a net attractive energy 
+between the atoms, even though the force may be repulsive. 
+The net attractive energy is maximized at a distance of 
 r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of 
@@
 between the spheres is zero. As r increases 
-further, the net attraction decreases and the potential approaches zero
-asymptotically. The attractive and repulsive effects can be separated into two
+further, the net attractive energy decreases and the potential approaches zero
+asymptotically. The attractive and repulsive forces can be separated into two
 potentials, shown with different colors in Figure \ref{fig:LJ_potential_2parts}, 
@@
 %A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
-A liquid will freeze when the free energy it would have as a solid is lower 
-than the free energy that it has as a liquid at the same density.
+A system will be solid when the free energy it would have as a solid is lower 
+than the free energy that it has as a liquid at the same temperature and density.
 At the phase transition the density will change abruptly, and
@@
 and the intrinsic Helmholtz free energy is defined as
-\begin{equation}F_{int} = F - \int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}\end{equation}  
+\begin{equation}F_{int} = F - \int n(\vec{r})V_{ext}{(\vec r)}d\vec{r}\end{equation}  
 %The quantity $\phi_{ext}$
-where $\phi_{ext}=\frac{V_{ext}}{N}$ and $V_{ext}(\vec r)$
-is the externally applied potential energy given in Equation~\ref{Hamiltonian}. 
+where $V_{ext}$ is an externally applied potential. 
 The Grand free energy is applicable to systems with a varying number of 
@@
 Writing the Hamiltonian, H in terms of its component kinetic and potential energies gives
-\begin{equation}\label{Hamiltonian}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
+\begin{equation}\label{Hamiltonian}{H = KE + V = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
 where the translational kinetic energy of the N atoms is given by 
@@
 \begin{displaymath}{V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
-and where \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
-is an externally applied potential that varies spatially, such that its 
-effects on each atom depend on the atom's position. Often $V_{ext}$ is 
-set to zero, as it is in this paper.
+%and where \begin{displaymath}{V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{displaymath} 
+%is an externally applied potential that varies spatially, such that its 
+%effects on each atom depend on the atom's position. Often $V_{ext}$ is 
+%set to zero, as it is in this paper.
 
@@
 the integral in Equation~\ref{Z_total} can be separated into two parts 
-\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
+\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}
++\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\int{dQ}
+~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}  
 Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume, and 
@@
 \begin{align} \label{eq:Zexcess}
-    Z_{excess}=\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
+    Z_{excess}=\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}
 \end{align} 
@@
       F_{excess} &= -k_{B}T\ln(Z_{excess}) \\
-                 &= -k_BT\ln{\left(\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)} \label{Fexcess-mostgeneral}
+                 &= -k_BT\ln{\left(\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)} \label{Fexcess-mostgeneral}
 \end{align} 
@@
 The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
-To model a fluid with "soft" spheres which have an interatomic potential 
-that decreases smoothly from infinity over a distance of 2R rather than instanteously 
+To model a fluid with ``soft" spheres which have an interatomic potential 
+that decreases smoothly from infinity to zero over a distance of 2R rather than instanteously 
 at 2R like that of hard spheres, Schmidt noted that 
@@
 \begin{align}\label{mayerder=w2convolution}
-     \frac{df(r)}{dr} = w_2 \otimes w_2(r) = \int{w_2(r')w_2(r-r')dr'}
+     \frac{df(r)}{dr} = \int_{-\infty}^\infty{w_2(r')w_2(r-r')dr'}
 \end{align} 
@@
 \begin{align}\label{w2_w3_relation}
-    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+    w_{3}=\int_{r}^{\infty}{w_{2}(r)dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
 \end{align}
@@
 \end{equation}  
-
 \begin{figure}
@@
 \end{figure}
-
-Unlike the WCA potential, the error function potential $V_{erf}(r)$ is temperature dependent.
-To compensate for this temperature dependence so that our functional better approximates
-a WCA potential, we make the parameters $\alpha$ and $\Xi$ functions of temperature as well. 
-The parameter $\alpha(T)$ is derived by setting 
-the error function potential $V_{erf}(r)$ 
+Unlike physical potential energies, the error function potential $V_{erf}(r)$ is temperature dependent.
+We construct temperature-dependent parameters $\alpha$ and $\Xi$ 
+to model a WCA potential. 
+The parameter $\alpha(T)$ is derived by setting the error function potential $V_{erf}(r)$ 
 equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$, since 
-the derivative of the mayer function peaks at this value. 
+the derivative of the Mayer function peaks at this value. 
 \begin{align}
@@
 from Monte-Carlo simulations shown by the white line, it is seen that
-the phase transition pressures and temperatures predicted by cDFT 
+the phase transition pressures predicted by cDFT 
 are consistently higher, and the 
-discrepancy grows as the temperature increases. For example, at a 
+discrepancy grows as the pressure increases. For example, at a 
 temperature of $T^*$=1, the phase transition reduced pressure predicted by 
@@
 predicted from Monte-Carlo simulations, giving transition 
-temperatures and pressures that were too high.
+pressures that were too high.
 

### 2021-01-15 papers/thesis-kirstie/thesis.tex
edits

@@
 
-A phase transition occurs along a phase boundary line on a phase diagram
+A phase transition occurs along a phase boundary line %on a phase diagram
 such as can be seen in the pressure-temperature (P-T) phase diagram
@@
         V &= \int_{Volume} \Theta(r_{ij}-R)d\vec r_{ij} \\
-        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  
-\end{align}
-\begin{align}
+        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  \\
+%\end{align}
+%\begin{align}
         R &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}} d\vec r_{ij} \\
@@
 \begin{align} 
-      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} \\
+      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} 
+\end{align} 
+\begin{align} 
       w_2 &= \delta(r_{ij}-R) \\
@@
 \begin{align}      
-     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}                                                              
-\end{align} 
-\begin{align} 
+     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}         \\ 
                 &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
@@
                 &+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec u ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right) \\
+%\end{align} 
+%\begin{align} 
+                %&\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
+                %&\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
+                %&+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec u ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)  \\
                 &\approx k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
@@
       w_1 &= \frac{w_2}{4\pi r}  \\
-      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  
-\end{align}
-\begin{align}      
+      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  \\
+%\end{align}
+%\begin{align}      
       \vec {w}_2 &= w_2~\hat r \\

### 2021-02-02 papers/thesis-kirstie/thesis.tex

--- 2021-01-15 papers/thesis-kirstie/thesis.tex
@@
 In the simplest case, the spheres are treated as having only a repulsive 
-potential that rises instantly to infinity when the spheres make contact. 
-This model, called the hard-sphere fluid, has served as the 
+potential. This model, called the hard-sphere fluid, has served as the 
 primary reference system for fluids. This is due to the fact that the 
-short-range repulsion between atoms is much stronger than their long-range 
-attraction, and thus dominates the interactions between atoms.
-
-In more recent times, other model fluids have been developed.
-Zwanzig introduced the square-well 
-fluid in 1954 which is a perturbation of the hard-sphere fluid created by adding 
-a hard-edge short-range attractive potential to the repulsive potential of the 
+repulsion between atoms is much stronger than their attraction, and so 
+the repulsive potential, which is descriptive of hard spheres, dominates 
+the interactions between atoms. In 1954 Zwanzig introduced the Square-well 
+fluid which is a perturbation of the hard-sphere fluid created by adding 
+a hard-edge attractive potential to the repulsive potential of the 
 hard-sphere fluid~\cite{ZwanzigSqrWell}. In 1964 Rowlinson studied 
-spheres with softly fading short-range repulsive potentials which he modeled by making
-the radius of hard spheres temperature dependent~\cite{rowlinson1964statistical}. 
+spheres with softly fading repulsive potentials which he modeled by making
+the radius of the spheres temperature dependent~\cite{rowlinson1964statistical}. 
 This reflects the idea that soft-spheres penetrate more deeply into 
@@
 with Rowlinson's idea of using temperature-dependent radii to model spheres 
-with soft repulsive potentials using hard-spheres~\cite{barker1967perturbation}. 
-In this model, 
-%a soft-sphere fluid is modeled by using a hard-sphere fluid with 
-hard spheres are given larger radii 
-at low temperatures, and smaller radii at high temperatures. 
-In 1971 Weeks, 
+with soft repulsive potentials to create the Barker-Henderson model 
+fluid~\cite{barker1967perturbation}. In this model, a soft-sphere fluid is 
+modeled by using a hard-sphere fluid with hard spheres of larger radii 
+at low temperatures, and smaller radii at high temperatures. In 1971 Weeks, 
 Chandler, and Anderson used temperature and density dependent hard-sphere 
-radii to model spheres with a more realistic softly fading repulsive 
-potential similar to that between neutral atoms.
-%portion of the Leonard-Jones potential
-Their is known as the Weeks-Chandler-Anderson (WCA) model fluid~\cite{andersen1971relationship} . 
-Models continue to be developed to characterize fluids with particular 
-properties, or to address problems like surface tension, 
-interfaces, adsorption, wetting and confinement.
-A review of some model fluids and problems that have been
-studied more recently is given by L\"{o}wen~\cite{lowen}.
+radii to model spheres with the more realistic softly fading repulsive 
+portion of the Leonard-Jones potential~\cite{andersen1971relationship}, 
+which is known as the Weeks-Chandler-Anderson model fluid.
+
+ADD MORE HERE TO COME UP TO DATE!!
+\newline
 
 While models for particular fluids were being created, powerful methods 
-to predict the properties of fluids were also being developed. 
+to describe the properties of fluids were also being developed. 
 In the late 1950s, Scaled Particle Theory (SPT)~\cite{ReissSPT} was 
@@
 obtained~\cite{Hansen}. 
-In 1958, Percus-Yevick introduced a method to describe the thermodynamics 
-of the hard-sphere fluid based on an approximate closure to the 
+%\textcolor{red}{
+In 1958, Percus-Yevick introduced an alternative method 
+for the hard-sphere fluid based on an approximate closure to the 
 Ornstein-Zernike integral equation~\cite{PercusYevick}.
@@
 In 1961 Morita and Hiroike introduced a density functional 
-approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies  
-%, and the corresponding equilibrium number density profile 
-of a classical inhomogeneous fluid. 
-Density functional theory was further developed for inhomogeneous classical fluids
-by Stillinger and Buff (1962)~\cite{stillingerandbuff}, Lebowitz and 
-Percus (1963)~\cite{lobowitzandpercus}, and Mermin (1965)~\cite{merium} among others. A review
-of the development of cDFT is found in Fundamentals of Inhomogeneous fluids~\cite{inhomogeneousfluids}.
-
-To use classical Density Functional Theory (cDFT), an expression for the free energy 
-as a functional of a spatially varying average single-particle number density $n(\vec r)$
-%that varies spatially 
-is constructed for a particular fluid.
+approach~\cite{MoritaDFT} to finding the equilibrium Grand and Helmholtz free energies, 
+and the corresponding equilibrium number density profile for 
+inhomogeneous fluids. To use the classical Density Functional Theory (cDFT) 
+approach, an expression for the free energy 
+as a functional of an average number density that varies spatially 
+must be constructed for a particular fluid.
 %DFT Evans 1979
@@
 to form a functional for the hard-sphere fluid based on the idea that 
-%the lowest term in the virial expansion of the excess chemical potential 
-the lowest order term in the virial expansion of the excess chemical 
-potential of a hard-sphere fluid can 
-be expressed in terms of a weighted density using a weight function 
-%using a weight 
-%function 
-of the form of 
-a backward Heaviside step function
-%\begin{equation}
-   %F_{ex}=k_BT\int_{volume} n(\vec r) \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) d\vec r
-%\end{equation}
-that can itself be expressed in terms of a sum of convolutions involving 
+the lowest term in the virial expansion 
+of the excess chemical potential for a hard-sphere fluid can 
+be expressed in terms of a sum of weighted densities using a weight 
+function of the form of a backward Heavyside step function (shown below) 
+that can be expressed in terms of a sum of convolutions involving 
 a set of weight functions related to the geometry of a sphere.
-For two spheres with radii, 
-$R_i$  and $R_j$, the convolution is written as~\cite{rosenfeld1989}
 \begin{displaymath}
-   \Theta(|\vec r - \vec r'|-(R_i + R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
+   \theta(|\vec r - \vec r'|-(\vec R_i + \vec R_j)) = w_i^{(3)} \otimes w_j^{(0)} + w_j^{(3)} \otimes w_i^{(0)} + w_i^{(1)} \otimes w_j^{(2)} + w_j^{(1)} \otimes w_i^{(2)}\end{displaymath}
 \begin{equation}   + \vec w_i^{(1)} \otimes \vec w_j^{(2)} + \vec w_j^{(1)} \otimes \vec w_i^{(2)}
 \end{equation}
-
-FMT gives the same result for the excess Helmholtz free energy of the 
-hard-sphere fluid as SPT does, from which it can be 
-constructed~\cite{rosenfeld1989,Hansen,santos2012phi3}. It also gives the same 
-result for the direct correlation function approximation given by Percus-Yevick. 
-However, it did not predict freezing due to one term in the expression that
-diverged when the fluid was highly confined.  
-Tarazona later solved this problem by introducing a tensor version of $\Phi_3$,
-one of the energy density terms 
-in the functional,~\cite{tarazonaphi3, santos2012phi3} 
-which took into account the non-spherical, and thus 
-non-symmetric, shape of the cavity in which the fluid is confined 
-while also yielding the correct excess free energy for the fluid as its 
-confinement is reduced from a three-dimensional volume (as in a bulk fluid), 
-to a two-dimensional plane (as in a thin-film), to a one-dimensional 
-line (as in a fluid confined to a capillary tube), to a zero-dimensional 
-dot, as in an atom located at a lattice site within a crystal~\cite{rothreview}.
-
-In 1999 Schmidt adopted FMT for spheres with soft 
+FMT gives the same result for the excess Helmholtz free energy of the hard-sphere fluid as 
+SPT does, from which it can be constructed~\cite{rosenfeld1989,Hansen,santos2012phi3},
+and for the direct correlation function approximation given by Percus-Yevick.
+Tarazona later improved on the 
+method by introducing a tensor version of one of the energy density terms 
+in the functional~\cite{tarazonaphi3, 
+santos2012phi3}, and in 1999 Schmidt adopted FMT for spheres with soft 
 interatomic potential energies
@@
 like the potential energy between hard spheres)
-forming Soft Fundamental Measure Theory (SFMT)~\cite{schmidt1999density}. 
+forming Soft Fundamental Measure Theory (SFMT)\cite{schmidt1999density}.
+
 %by replacing $w^{(3)}$ with a weight function that would result in  
@@
 %rather than instantaneously).
-Schmidt applied his Soft Fundamental Measure Theory to various fluids, 
-%with soft, short-range repulsive and attractive potentials, 
-and developed 
+
+Schmidt applied his Soft Fundamental Measure Theory to various fluids 
+with soft, short-range repulsive and attractive potentials, and developed 
 an error function model 
@@
 %, similar to that exhibited by a WCA fluid. 
-For some more recent developments in classical Density Functional Theory, 
-see the review by Evans ~\cite{evansreview} (2016), the book 
-Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen} (2013), 
-and the paper by Tschopp and coworkers~\cite{BHextended} (2020) in which 
-they extend the Barker-Henderson model to inhomogeneous 
-fluids from first principles.
-
-Classical density functional theory has also been applied 
-to phase transitions, and freezing. In 1979 Ramakrishnan and Yussouff studied 
-freezing by viewing a 
-crystal as a highly inhomogeneous liquid~\cite{ramakrishnanandyussouff}. 
-In 2010 Oettel and fellow researchers studied hard-sphere crystals with 
-fully developed FMT~\cite{Oettel2010}, and 
-in 2016 FMT was used to find the fluid and crystal phases of the 
-Asakura-Oosawa model fluid~\cite{asakuraoosawa}. 
-Lutsko and Lam studied studied crystallization of the Leonard-Jones 
-fluid~\cite{lutsko} in 2018, and Lutsko and Schoonen looked at cDFT applied 
-to the solid state~\cite{lutskoschoonen2020} in 2020.
-In all these studies the hard-sphere fluid has been used as the reference 
-system to which other terms are added to create 
-the full functional for the fluid being studied.
-
-In this paper, classical Density Functional Theory will be used to predict 
-the freezing of a Weeks-Chandler-Anderson model fluid directly by applying 
-SFMT with Schmidt's proposed error function model, 
-in which we have incorporated our temperature dependent parameters to 
-fit the functional to the soft-sphere WCA fluid. 
+In this paper, classical Density Functional 
+Theory will be used to predict the freezing of a Weeks-Chandler-Anderson 
+model fluid by applying SFMT with Schmidt's proposed error function model, 
+in which we have incorporated our temperature dependent parameters to fit the 
+functional to the soft-sphere WCA fluid.
 %form the functional for the excess Helmholtz free energy. 
@@
 Results will be compared to those generated from Monte-Carlo simulations.
+
+%Results will be compared to those from Monte-Carlo simulations.
 
@@
 %To do this, the Helmholtz free energies of the fluid
-%in a fluid state and in a crystalline solid state must be found.
+%in a liquid state and in a crystalline solid state must be found.
 %In general, the Helmholtz free energy, which consists of an ideal part 
@@
 as a classical fluid where the spheres interact with each other according 
-to their interatomic potential.
+to their electrostatic interatomic potential.
 
@@
 Above $\sigma$ the potential becomes negative indicating a net attraction 
-between the atoms. The net attractive energy is maximized at a distance of 
+between the atoms. The net attraction is maximized at a distance of 
 r=$2^\frac{1}{6}\sigma=2R\approx{1.122}\sigma$ where R is the radius of 
 the sphere given by $R={2^{-\frac{5}{6}}}\sigma$. At this point the depth of
-the potential well, given by epsilon $\epsilon$, is a maximum, and the force
-between the spheres is zero. As r increases 
+the potential well, given by epsilon $\epsilon$, is a maximum. As r increases 
 further, the net attraction decreases and the potential approaches zero
@@
     \includegraphics[height=7cm]{LJ_Potential_2part.pdf}
-    \caption{Lennard-Jones potential broken into two parts}
+    \caption{Lennard-Jones potential broken into two parts for $\sigma=1$, $\epsilon=1$}
     \label{fig:LJ_potential_2parts}
@@
     \includegraphics[height=7cm]{WCA_Potential.pdf}
-    \caption{WCA potential}
+    \caption{WCA potential for $\sigma=1$, $\epsilon=1$}
     \label{fig:WCA_potential}
@@
 \end{align} 
-The WCA potential decreases as the distance between the 
+The WCA potential decreases from infinity as the distance between the 
 centers of the atoms r increases, and becomes, and remains, essentially 
@@
 their interatomic potential, they are called ``soft spheres'' and their 
-gradually fading potential makes them slightly ``squishy'', as pictured in
+gradually fading potential makes them slightly ``squishy'', as illustrated in
 Figure~\ref{fig:TwoSpheres}.
@@
     \includegraphics[height=3.5cm]{figs/TwoSpheres.pdf} 
-    \caption{(Left) The WCA potential is zero at r=2R. 
+    \caption{(Left) The WCA potential is a minimum at r=2R. 
              (Right) The WCA potential increases rapidly for $r<2R$.}
@@
 
-%Freezing of simple fluids is governed by entropy, which the universe
-%seeks to maximize according to the second law of Thermodynamics. When the 
-%entropy of a pure substance in its liquid state is lower than the entropy 
-%that it would have in its solid state at the same temperature (necessary 
-%for thermal equilibrium with the environment), the same pressure (necessary for mechanical 
-%equilibrium with the environment), and the same chemical potential (necessary for diffusive 
-%equilibrium with the environment), a phase transition occurs, that is, the liquid 
-%freezes. 
-%%Along the phase boundary between pure liquid and pure solid states, both 
-%%liquid and solid states can coexist in equilibrium. 
-%Freezing of simple fluids can be seen when the entropy that a pure substance 
-%would have in a liquid state is lower than the entropy 
-%that it would have in its solid state at the same temperature (necessary 
-%for thermal equilibrium), the same pressure (necessary for mechanical 
-%equilibrium), and the same chemical potential (necessary for diffusive 
-%equilibrium).
-
-%A liquid will eventually freeze as heat is removed, or the liquid is compressed, and 
-A liquid will freeze when the free energy it would have as a solid is lower 
-than the free energy that it has as a liquid at the same density.
-At the phase transition the density will change abruptly, and
-the particular structure and density that result will be affected 
+Freezing is governed by entropy, which the universe %, or any closed system, 
+seeks to maximize according to the second law of Thermodynamics. When the 
+entropy of a pure substance in its liquid state is lower than the entropy 
+that it would have in its solid state at the same temperature (necessary 
+for thermal equilibrium), the same pressure (necessary for mechanical 
+equilibrium), and the same chemical potential (necessary for diffusive 
+equilibrium), an abrupt phase transition occurs, that is, the liquid 
+freezes. The particular structure and density that result are affected 
 by the interatomic potential. When repulsive, soft sphere fluids 
@@
     \centering
-    \includegraphics[height=7cm]{figs/P-T_Diagram.pdf}
+    \includegraphics[height=8cm]{figs/P-T_Diagram.pdf}
     \caption{Conceptual Pressure-Temperature Phase Diagram}
@@
 
-A phase transition occurs along a phase boundary line on a phase diagram
-such as can be seen in the pressure-temperature (P-T) phase diagram
-shown in Figure~\ref{fig:P-T_Diagram}. 
-%
-%Points along the phase boundary that separates the solid and the fluid phases 
-%indicate a phase transition where the solid and fluid phases coexist, 
-%and show that both the solid and the fluid have the same temperature (necessary 
-%for thermal equilibrium)
-%and the same pressure (necessary for mechanical 
-%equilibrium)at a phase transition, and the same chemical potential 
-%(necessary for diffusive equilibrium). 
-%
-This phase diagram is 
-representative of some substance 
-which exhibits both attraction and repulsion. However, since 
+For pure substances, a phase transition occurs along a line on a 
+Pressure-Temperature (P-T) phase diagram, such as can be seen in the phase 
+diagram shown in Figure~\ref{fig:P-T_Diagram}. This phase diagram is 
+representative of some arbitrary, unknown substance composed of atoms, 
+or molecules which exhibit both attraction and repulsion. However, since 
 the atoms in a WCA fluid do not exhibit attraction there will not be a 
-dividing line between the gas and liquid phases, there will simply
-be a fluid.
-The phase diagram indicates the equilibrium state a substance will be in at
-different temperatures and pressures.
-At a phase transition, both the fluid and the solid have 
-the same temperature (necessary 
-for thermal equilibrium), the same pressure (necessary for mechanical 
-equilibrium) and the same chemical potential (necessary for diffusive equilibrium).
-%The temperature , pressure, and chemical potential will be the same for both
-%the fluid and the solid at a fluid-solid phase transition.
+dividing line between the gas and liquid phases, there will just simply
+be a fluid, which we will refer to as a liquid in this paper.
+%which may be referred to as a "liquid", since at the phase boundary the fluid is dense. 
+Points along the line that separates the solid and the liquid phases 
+indicate a phase transition where the solid and liquid phases coexist, 
+and show that both the solid and the liquid have the same temperature 
+and the same pressure at a phase transition.
 
 The pressure can be related to the Helmholtz free energy F, which is minimized 
-for a system at fixed temperature, volume, and number of particles, as the 
-entropy of the system and its surroundings added together is maximized. The negative of 
-the partial derivative of the Helmholtz free energy with respect to the volume V 
-at a fixed temperature T and number of particles N, equals the pressure. 
+for a system at fixed temperature, volume and number of particles, as the 
+entropy of the system and its surroundings is maximized. The negative of 
+the partial derivative of the Helmholtz free energy with respect to volume V, 
+and at a fixed temperature T and number of particles N, equals the pressure. 
 \begin{equation}{P=-\frac{\partial{F}}{\partial{V}}\bigg|_{T,N}}\end{equation}
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig1.pdf}
-    \caption{Fluid and solid densities at the fluid-solid phase 
+    \caption{Liquid and solid densities at the solid-liquid phase 
     transition are identified using Maxwell's Double Tangent Construction}
@@
 
-Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the fluid
-and one for the solid at a fixed temperature. 
-%Each curve consists of a collection of data points $(\frac{F}{N}, \frac{1}{n})$ 
-The data for the curves 
-%and the solid respectively, which 
-will be obtained by using cDFT. 
-%to find the Helmholtz free energies of the fluid and the solid for various 
-%temperatures and number densities. 
+Two such curves are shown in Figure~\ref{fig:MaxwellDT}, one for the liquid 
+and one for the solid at a fixed temperature. Each curve consists of a 
+collection of data points $(\frac{F}{N}, \frac{1}{n})$ for the liquid and
+the solid respectively, which will be obtained by using cDFT to find the 
+Helmholtz free energies of the liquid and the solid for various 
+temperatures and number densities. 
 Since the pressure of the 
-fluid is the same as the pressure of the solid at the phase transition, 
-the density of the fluid and the density of the solid at 
+liquid is the same as the pressure of the solid at the point of transition, 
+the density of the liquid and the density of the solid at the point of 
 freezing can be found from identifying the points at which the 
@@
 To facilitate finding the pressure at the phase transition computationally, 
-it is helpful to take different approach by considering the Gibbs free energy given by G=F+PV. 
+it is helpful to consider the Gibbs free energy given by G=F+PV. 
 Rearranging the terms and dividing by N gives an  expression for the 
@@
 \end{equation}
-and describes the black line shown in Figure~\ref{fig:MaxwellDT} where the 
-pressure is given by the negative of the slope of the line, and the Gibb's 
-free energy per atom, or chemical potential is given by the y-intercept. 
-At the phase transition the chemical potential of the fluid is 
+and describes the black line shown in Figure~\ref{fig:MaxwellDT} with the 
+pressure given by the negative of the slope of the line, and the Gibb's 
+free energy per atom, or chemical potential given by the y-intercept. 
+At the point of phase transition the chemical potential of the liquid is 
 the same as the chemical potential of the solid, as is the pressure.
 \begin{equation}
-    \mu_F=\mu_S\mbox{~~~~and~~~~}P_F=P_S\mbox{~~~~at a phase transition}
+    \mu_L=\mu_S\mbox{~~~~and~~~~}P_L=P_S\mbox{~~~~at a phase transition}
 \end{equation}
@@
      \includegraphics[height=8cm]{figs/MaxwellDTC-Fig2.pdf}
-    \caption{Pressure is found at the point of intersection where $\mu_F=\mu_S$.}
+    \caption{Pressure is found at point of intersection which occurs 
+    when $\mu_L=\mu_S$.}
     \label{fig:GibbsvsP}
@@
     \includegraphics[height=8cm]{figs/MaxwellDTC-Fig3.pdf}
-    \caption{The fluid and solid densities at the phase transition are 
-    found from a plot of pressure vs inverse density for both the fluid and the solid.}
+    \caption{The liquid and solid densities at the phase transition are 
+    found from a plot of pressure vs inverse density for both the liquid and the solid.}
     \label{fig:Pvsinvn}
@@
 intersection that occurs when the chemical potential 
-is plotted against the pressure for both the fluid and the 
+is plotted against the pressure for both the liquid and the 
 solid at fixed temperature, as shown in Figure~\ref{fig:GibbsvsP}. 
 This plot is constructed from data collected from lines tangent 
-along the curves on Helmholtz free energy per atom verses inverse 
+along the curves on the Helmholtz free energy per atom verses inverse 
 density plots (like that of Figure~\ref{fig:MaxwellDT}) for both the 
-fluid and the solid at a given temperature where 
-the slope of the tangent line at a given point gives the negative of the 
-pressure, and the corresponding y-intercept of the tangent 
-line gives the chemical potential at a given point. 
-%Collecting this data 
-%(P,~$\mu$) at each point for the fluid and the solid respectively, 
-%produces the curves in Figure~\ref{fig:GibbsvsP}.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=7cm]{figs/T-n_Diagram.pdf}
+liquid and the solid at a given temperature. 
+The slope of the tangent line at a given point gives the negative of the 
+pressure at that point, and the corresponding y-intercept of the tangent 
+line gives the chemical potential at that point. Collecting this data 
+(P,~$\mu$) at each point for the liquid and the solid respectively, 
+produces the curves in Figure~\ref{fig:GibbsvsP}.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=8cm]{figs/T-n_Diagram.pdf}
     \caption{Conceptual Temperature-Density Phase Diagram used to identify 
-    the number densities of the fluid and the solid at a phase 
+    the number densities of the liquid and the solid during a phase 
     transition. The gray region is where different phases coexist in 
@@
 Once the pressure is known at the point of phase transition, for a given 
-temperature and number density, the fluid and solid number densities at 
+temperature and number density, the liquid and solid number densities at 
 the phase transition can be found from plots of the pressure versus the 
-inverse of the number density for both the solid and the fluid, 
+inverse of the number density for both the solid and the liquid, 
 as shown in Figure \ref{fig:Pvsinvn}. By repeating this procedure to 
-find the fluid and solid number densities at the phase transition 
-for various temperatures, a temperature-density (T-n) phase diagram 
-can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}. 
-However, for a WCA fluid there will be no gray hump which is characteristic 
-of fluids that have attractive interatomic potentials, and which display
-an interface between vapor and liquid at temperatures below the critical point.
+find the liquid and solid number densities at the phase transition 
+for various temperatures, a Temperature-Density (T-n) phase diagram 
+can be generated, similar to that shown in Figure \ref{fig:T-n_Diagram}, 
+but without the gray hump characteristic of fluids that have 
+both attractive and repulsive interatomic potentials.
 
@@
 Classical Density Functional Theory will be used to find the 
-Helmholtz free energies of the fluid and the solid for various 
+Helmholtz free energies of the liquid and the solid for various 
 temperatures and number densities, so that phase diagrams can be 
 constructed. 
-The key idea behind classical Density Functional Theory is that 
+The key idea behind classical Density Functional Theory (cDFT) is that 
 free energy can be written as a functional (a function of a function) of 
@@
 a minimum at equilibrium means that the equilibrium free energy of the 
-system $f_{eq}[n_{eq}(\vec r)]$, and its corresponding equilibrium 
+system ${f_{eq}}{[}\rho{(}\vec{r}{)]}$, and its corresponding equilibrium 
 number density profile $n_{eq}(\vec r)$, can be found simply by varying 
@@
 Energy defined as
-\begin{equation}\Omega=F-\mu{N}\end{equation}
+\begin{equation}\Phi=F-\mu{N}\end{equation}
 which can then be expressed in terms of a varying number density 
 profile $n(\vec r)$
-\begin{equation}\label{GrandFE}\Omega= F_{int} +\int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
-where
-\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}%
-and the intrinsic Helmholtz free energy is defined as
-\begin{equation}F_{int} = F - \int n(\vec{r})\phi_{ext}{(\vec r)}d\vec{r}\end{equation}  
-%The quantity $\phi_{ext}$
-where $\phi_{ext}=\frac{V_{ext}}{N}$ and $V_{ext}(\vec r)$
-is the externally applied potential energy given in Equation~\ref{Hamiltonian}. 
+\begin{equation}\label{GrandFE}\Phi= F_{int} +\int n(\vec{r})\phi{(\vec r)}d\vec{r}-\mu\int n(\vec r)d\vec{r}\end{equation}
+where $\phi=\frac{V_{ext}}{N}$ is an externally applied potential energy 
+$V_{ext}(\vec r)$ per particle, and an intrinsic Helmholtz free energy 
+is defined as
+\begin{equation}F_{int} = F - \int n(\vec{r})\phi{(\vec r)}d\vec{r}\end{equation}
+
 The Grand free energy is applicable to systems with a varying number of 
@@
 particles will be of interest in which case
-%\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation} and so
-the Grand free energy is constrained
-\begin{equation}\label{GrandFE}\Omega_{constrained}= F[n(\vec r)]-\mu N\end{equation}
+\begin{equation}\int n(\vec r)d\vec{r}=N\end{equation}
+and so the Grand free energy is constrained
+\begin{equation}\label{GrandFE}\Phi_{constrained}= F[n(\vec r)]-\mu N\end{equation}
 The dependence on $n(\vec r)$ is now only present in the expression for 
@@
 \begin{equation}df_N=-\frac{S}{N}dT-\frac{P}{N}dV-PVd\left(\frac{1}{N}\right)\end{equation}
-%This result shows that the natural variables of the Helmholtz free energy 
-%per particle $f_N$ are temperature, volume and one over the number of 
-%particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
-%\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
-%Taking the differential
-which can be set equal to 
-\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}\bigg|_{V,\frac{1}{N}}dT 
-+\frac{\partial f_N}{V}\bigg|_{T,\frac{1}{N}}dV + \frac{\partial f_N}
-{\partial \left(\frac{1}{N}\right)}\bigg|_{T,V}d\left(\frac{1}{N}\right)\end{displaymath}
-Using the relation
+This result shows that the natural variables of the Helmholtz free energy 
+per particle $f_N$ are temperature, volume and one over the number of 
+particles, which in turn also holds the number density $n=\frac{V}{N}$ fixed. 
+\begin{equation}f_N(T,V, \frac{1}{N})\end{equation}
+Taking the differential
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT +\frac{\partial f_N}{V}dV+ \frac{\partial f_N}{\partial \left(\frac{1}{N}\right)}d\left(\frac{1}{N}\right)\end{displaymath}
+and using the relation
 \begin{displaymath}d\left(\frac{1}{n}\right)=Vd\left(\frac{1}{N}\right)+\frac{1}{N}dV\end{displaymath}
 we get
-\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}\bigg|_{\frac{1}{n}}dT 
-+ \frac{\partial f_N}{\partial \frac{1}{n}}\bigg|_{T}d\left(\frac{1}{n}\right)\end{displaymath}
+\begin{displaymath}df_N=\frac{\partial f_N}{\partial T}dT + \frac{\partial f_N}{\partial \frac{1}{n}}d\left(\frac{1}{n}\right)\end{displaymath}
+
 \begin{equation}f_N(T,\frac{1}{n})\end{equation}
-This result shows that temperature and the inverse number density, 
-or volume per particle, are natural variables of the Helmholtz free energy 
-per particle $f_N$.
-%which will thus be minimized when these natural variables are held fixed.
 
@@
 experience spontaneous processes that increase the total entropy of a 
-system and its surroundings until equilibrium is reached
-\begin{align}
-    \Delta{S}_{System + Surroundings} \geq 0 
-\end{align}
-Here "surroundings" means 
+system and its surroundings until equilibrium is reached, 
+$\Delta{S}_{System + Surroundings} \geq 0$. Here ``surroundings" means 
 the entire universe minus the system. This, together with the conservation 
@@
 (or gained) by the system is gained (or lost) by the surroundings 
-\begin{align}
-    \Delta{U}_{System}=-\Delta{U}_{Surroundings} 
-\end{align}
-gives the result that 
+$\Delta{U}_{System}=-\Delta{U}_{Surroundings}$, gives the result that 
 changes in the free energy 
@@
 when the natural variables of the free energy are held fixed.
-For our case, the temperature and inverse number density are the natural 
-variables, and so holding these fixed we get
+
+For our case, with a fixed temperature and number density we get
 %Rewriting Eq~\ref{usetoshowmin} to show the system terms explicitly
@@
 \begin{equation}
-  df_{N~System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
+  df_{System}=-\left(\frac{T}{N}\right)dS_{System+Surroundings}
 \end{equation}
@@
 %~\cite{schroeder}. 
-Since the changes in the total entropy are always positive, the changes
-in the free energy are negative and the free energy tends toward a minimum. 
+and so the free energy tends toward a minimum. 
 Details of this procedure be found in Thermal Physics by Schroeder~\cite{schroeder}, 
@@
 a momentum $\vec{p}$ and a position $\vec{r}$, and is given by
-\begin{align}  \label{Z_total}
-     Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}\int{dQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}
-\end{align}
+\begin{equation}{Z=}\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dPdQ}~e^\frac{-H(\vec{p}_1,\vec{p}_2,...\vec{p}_N,\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\end{equation}
 where
-\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~and~each~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
-\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~and~each~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
+\begin{displaymath}{dP=d\vec{p}_1d\vec{p}_2...d\vec{p}_N \mbox{~~~~and~~~~} d\vec{p}=dp_xdp_ydp_z}\end{displaymath}
+\begin{displaymath}{dQ=d\vec{r}_1d\vec{r}_2...d\vec{r}_N \mbox{~~~~and~~~~} d\vec{r}=dxdydz\mbox{~~~~}}\end{displaymath}
 Writing the Hamiltonian, H in terms of its component kinetic and potential energies gives
-\begin{equation}\label{Hamiltonian}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
+\begin{equation}{H = KE + V + V_{ext} = \sum_{i=1}^N\left(\frac{|\vec{p}_i|^2}{2m}\right)+V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)+V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}\end{equation}
 where the translational kinetic energy of the N atoms is given by 
@@
 interatomic and external potential energies depend only on position, 
-the integral in Equation~\ref{Z_total} can be separated into two parts 
-\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
+the integral in equation (8) can be separated into two parts 
+\begin{equation}{Z=\frac{1}{N!}\frac{1}{h^{3N}}\int...\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+\frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\int...\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)}{k_BT}}\end{equation}
 Multiplying by $\frac{V^N}{V^N}$ where V is the real space volume, and 
-\begin{equation}V^N=\int dQ\end{equation} 
-the partitian function can then be expressed as
+\begin{equation}{V^N=}\int{...}\int{dQ}\end{equation} 
+the partitian function can be expressed as
 \begin{equation}\label{Zmultiply}{Z=Z_{ideal}Z_{excess}}\end{equation}
@@
 \begin{align} \label{eq:Zideal}
-    Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}
+    Z_{ideal}=V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}
 \end{align}
 \begin{align} \label{eq:Zexcess}
-    Z_{excess}=\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
-\end{align} 
-The ideal partitian function $Z_{ideal}$ is that of an ideal gas.
-%an ideal gas which exhibits no interaction between the atoms - a condition approached 
-%by a gas at low density where the atoms rarely encounter one another. 
+    Z_{excess}=\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N)-V_{ext}(\vec{r}_1,\vec{r}_2,{...},\vec{r}_N )}{k_BT}
+\end{align} 
+The ideal partitian function $Z_{ideal}$ is associated with an ideal gas 
+which exhibits no interaction between the atoms - a condition approached 
+by a gas at low density where the atoms rarely encounter one another. 
 The excess partitian function $Z_{excess}$ accounts for the interaction 
-of all the atoms according to their interatomic potential.
-
-Using Equation~\ref{Zmultiply}, the free energy can also be broken into two parts.
-%\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
+of all the atoms according to their interatomic potential. Using 
+Equation~\ref{Zmultiply}, the free energy can also be broken into two parts.
+\begin{equation}{F=-k_{B}T\ln(Z_{ideal}Z_{excess})}\end{equation}
 \begin{align} \label{eq:Finparts}
-     F &= -k_{B}T\ln(Z_{ideal}Z_{excess})  \\
-       &= -k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess}) \\
-       &= F_{ideal} + F_{excess}
-\end{align}
-%\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}
-
-The ideal Helmholtz free energy is given by 
-%(put Equation~\ref{eq:Zideal} into Equation~\ref{eq:Finparts})
-\begin{align}
-    F_{ideal} &= -k_{B}T\ln(Z_{ideal}) \\
-              &= -k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{dP}~e^\frac{-\left(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m}\right)}{k_BT}\right)} 
-\end{align}  
+    F=-k_{B}T\ln(Z_{ideal})-k_{B}T\ln(Z_{excess})
+\end{align}
+\begin{equation}{F=F_{ideal} + F_{excess}}\end{equation}
+
+Putting Equation~\ref{eq:Zideal} into Equation~\ref{eq:Finparts}, 
+the ideal Helmholtz free energy is
+\begin{equation}{F_{ideal}=-k_BT\ln{\left(V^N\frac{1}{N!}\frac{1}{h^{3N}}\int{...}\int{dP}~e^\frac{-(\frac{|\vec{p}_1|^2}{2m}+ \frac{|\vec{p}_2|^2}{2m}+...\frac{|\vec{p}_N|^2}{2m})}{k_BT}\right)}}\end{equation}
+
 \begin{equation}{\label{Fidealmultipleint}}{=-k_BT\ln\left(V^N\frac{1}{N!}\frac{1}{h^{3}}\int{d\vec{p}_1}~e^\frac{-\frac{|\vec{p}_1|^2}{2m}}{k_BT}\frac{1}{h^{3}}\int{d\vec{p}_2}~e^\frac{-\frac{|\vec{p}_2|^2}{2m}}{k_BT}{...}\frac{1}{h^{3}}\int{d\vec{p}_N}~e^\frac{-\frac{|\vec{p}_N|^2}{2m}}{k_BT}\right)}\end{equation}
@@
 this becomes
-\begin{equation}F_{ideal}= k_BT[N\ln\left(\frac{N\Lambda^{3}}{V}\right)-N]\end{equation} 
+\begin{equation}{F_{ideal}= k_BT[N\ln(\frac{N\Lambda^{3}}{V}{)-N}]}\end{equation} 
 Rewriting the equation in terms of the number density $n=\frac{N}{V}$, 
@@
 Equation~\ref{eq:Fideal} becomes
-\begin{equation}\label{eq:Fideal-n(r)}{F_{ideal}[n(\vec{r})]= k_BT\int_{Volume}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]~d\vec{r}}\end{equation}
-
-%Putting Equation~\ref{eq:Zexcess} into Equation~\ref{eq:Finparts}, 
-The excess Helmholtz free energy is written most generally as
-\begin{align}
-      F_{excess} &= -k_{B}T\ln(Z_{excess}) \\
-                 &= -k_BT\ln{\left(\frac{1}{V^N}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)} \label{Fexcess-mostgeneral}
-\end{align} 
-This integral can be somewhat simplified if it can be assumed that the 
+\begin{equation}\label{eq:Fideal-n(r)}{F_{ideal}[n(\vec{r})]= k_BT\int_{Vol}[n(\vec{r})\ln(n(\vec{r})/\text{n}_\text{Q})-n(\vec{r})]d\vec{r}}\end{equation}
+
+Putting Equation~\ref{eq:Zexcess} into Equation~\ref{eq:Finparts}, 
+the excess Helmholtz free energy is written most generally as
+\begin{equation}\label{Fexcess-mostgeneral}{F_{excess}= -k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}~e^\frac{-V(\vec{r}_1, \vec{r}_2,...\vec{r}_N)-V_{ext}(\vec{r}_1, \vec{r}_2,...\vec{r}_N)}{k_BT}\right)}}\end{equation}
+This integral can be greatly simplified if it can be assumed that the 
 potential energy $V(\vec{r_1},\vec{r_2},...\vec{r_N})$ is a sum of the 
@@
 
-\begin{equation}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},
-\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}
-(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{equation}
-\begin{equation}{=-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
-%where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
-where function $f_{ij}$ is the Mayer function given by
+\begin{equation}{e^{-\frac{\frac{1}{2}\sum^N_i\sum^N_{j\neq{i}}V_{ij}(\vec{r_i},\vec{r_j})}{k_BT}}=e^{\frac{-V_{12}(\vec{r_1},\vec{r_2})}{k_BT}}e^{-\frac{V_{13}(\vec{r_1},\vec{r_3})}{k_BT}}e^{-\frac{V_{14}(\vec{r_1},\vec{r_4})}{k_BT}}...}\end{equation}
+\begin{equation}{\approx-(1+f_{12})(1+f_{13})(1+f_{14})...}\end{equation}
+where each exponential term is simplified by a Taylor expansion $e^x\approx{1+x}$.  
+The function $f_{ij}$ the Mayer function given by
 \begin{align}\label{eq:mayerfunction}
@@
 \begin{align} \label{eq:Fexcess-simplified}
-    F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)}
-\end{align}
-
-  \begin{figure}
-    \centering
-    \includegraphics[height=6.5cm]{figs/diagrammic.pdf}
+    F_{excess}=-k_BT\ln{\left(\frac{1}{V^N}\int{...}\int{dQ}\left[1 + \sum{f_{ij}} + \sum{f_{ij}f_{kl}} +\sum{f_{ij}f_{kl}f_{mn}} +... \right]\right)}
+\end{align}
+
+  \begin{figure}
+    \centering
+    \includegraphics[height=7cm]{figs/diagrammic.pdf}
     \caption{Two-pair interactions $f_{ij}f_{kl}$ are shown for N=4 atoms. 
@@
 a more general solution.
-At low densities, it is not likely to find more than two atoms in close 
+At low densities, it is not likely to find two or more atoms in close 
 proximity, and so only interactions between individual pairs of spheres, 
@@
 two atoms in the pair. 
-With a careful counting of the resulting number of such pairs, the 
-low-density excess free energy for a homogeneous fluid can then be 
-written~\cite{schroeder} as
-\begin{align}\label{eq:Fexcess}
-    F_{excess} &\approx k_BTn^2\int d\vec {r}~'\int d\vec {r}~"\left(-\frac{1}{2}f(|\vec {r}~'-\vec {r}~"|)\right) \\
-               &\approx k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V
-    %\frac{F_{excess}}{N} &\approx k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)
-\end{align} 
+With a careful counting of the resulting number of such pairs, the excess 
+free energy can then be written~\cite{schroeder} as
+\begin{equation}\label{eq:Fexcess}{\frac{F_{excess}}{N}=k_BTn\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)}\end{equation} 
 in the thermodynamic limit of large N. 
@@
 as the second virial expansion coefficient typically called $B_2$. 
-A correct expression for the excess Helmholtz free energy should 
-reduce to this result in the low-density limit. 
+It is important that an expression for the excess Helmholtz free energy 
+correctly reduces to this result in the low-density limit. 
 Outside of the low density limit, it is quite difficult to evaluate the 
@@
 found by varying the number density profile
-at some fixed temperature until the free energy is minimized.
-
+through all possible configurations at some fixed temperature
+until the free energy is minimized.  
 %In this section 
@@
 Soft Fundamental Measure Theory (SFMT) is a method for forming an 
-approximation for the excess Helmholtz free energy of a soft-sphere fluid.
+expression for the excess Helmholtz free energy of a soft-sphere fluid.
 In the next chapter, 
@@
 that does not depend on explicitly taking into account the interaction 
-that each sphere has with every other sphere in the fluid (as required by 
+that each sphere with every other sphere in the fluid (as required by 
 Equation~\ref{eq:Fexcess-simplified}).
@@
 Rosenfeld~\cite{rosenfeld1989, rosenfeld1994, rosenfeld1996, fmtfromsptandsignerror} 
-for the hard-sphere fluid. 
-%FMT in turn extended  
-%Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
-%developed for homogeneous hard-sphere fluids, 
-%to inhomogeneous, or non-uniform hard-sphere fluids.
-
-%\textcolor{red}{FROM HERE}
-%\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
-%Rosenfeld begins the development of FMT by considering the lowest order 
-%term in the virial expansion of the excess chemical potential.
-%Taking the partial derivative of the free energy in the low density limit given 
-%by Equation~\ref{eq:Fexcess} gives
-%\begin{align}
-   %\mu_{ex} &= \frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V} \nonumber \\
-            %&= k_BT\int -f(r)~n~d\vec r
-%\end{align}
-Rosenfeld, who was studying mixtures, begins the development of FMT by considering that 
-for a hard-sphere fluid the Mayer function, $f(r)$, can be written in terms 
-of a backward Heaviside step function $\Theta$ which, for a hard-sphere fluid 
-composed of a mixture of spheres of radius $R_i$ and $R_j$, is
-given by
-\begin{align} 
-    f(r_{ij}) = -\Theta(r_{ij}-(R_i + R_j))
-\end{align}
-where $r_{ij}=|\vec r_i - \vec r_j|$ is the distance between the centers of two spheres, 
-and $\Theta$ is 1 for $r_{ij}<R_i+R_j$ and 0 for $r_{ij}>R_i+R_j$.
-This can easily be shown by putting the interatomic potential between two hard spheres
-$V(r_{ij})$
-\begin{align}
-    V(r_{ij})=\left\{ \begin{array}{rc} \infty & 0<r_{ij} \leq R_i+R_j \\ 0  & r_{ij}> R_i+R_j \end{array}\right.
-\end{align}
-into the equation for the Mayer function
-\begin{align}
-     f(r_{ij})=\exp^{-\frac{V(r_{ij})}{k_{B}T}}-1
-\end{align} 
-%where
-
-%
-%The low-density excess free energy given by Equation~\ref{eq:Fexcess}
-%The lowest order term of the excess chemical potential  
-%can then be written as a weighted density.
-%\begin{align}
-     %\mu_{excess}=k_BT\int n\theta(|\vec r|-2R)~d\vec r
-      %%F_{excess}=k_BT\left(\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r\right) V
-%\end{align} 
-%where $n \rightarrow n(\vec r)$ for an inhomogeneous fluid.
-%The excess Helmholtz free energy given in Equation~\ref{eq:Fexcess} 
-%can then be written as
-%\begin{align}
-      %F_{excess} = k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r  
-%\end{align} 
-Rosenfeld recognized that the integral of the Heaviside step function
-is equal to the volume of a sphere with a radius equal to that of the radii 
-of two hard spheres added together. For a mixture of hard spheres of radius $R_i$ and $R_j$
-this becomes
-\begin{align}
-       \int_{Volume} \Theta(r_{ij}-(R_i+R_j))~d\vec r_{ij} &=\frac{4\pi}{3}(R_i+R_j)^3  \label{exclusion-volume}\\
-       &= \frac{4\pi}{3}R_i^3 + (4\pi R_i^2)R_j + R_i(4\pi R_j^2) + \frac{4\pi}{3}R_j^3  \nonumber \\
-       &= V_i(1) + S_iR_j + R_iS_j+ V_j(1)  \label{VSR1sum}
-\end{align}  
-where the volume, surface area, radius, and the number 1 can be written as
-\begin{align}
-        V &= \int_{Volume} \Theta(r_{ij}-R)d\vec r_{ij} \\
-        S &= \int_{Volume} \delta(r_{ij} -R) d\vec r_{ij}  
-\end{align}
-\begin{align}
-        R &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}} d\vec r_{ij} \\
-        1 &= \int_{Volume} \frac{\delta(r_{ij} -R)}{4\pi r_{ij}^2} d\vec r_{ij}    
-\end{align}
-where R is either $R_i$ or $R_j$ accordingly. Rosenfeld also recognized that the 
-Heaviside function can be written as a convolution of the weight functions 
-which, for a fluid composed only of spheres of radius R, is given by
-\begin{align} \label{mayerconvolution}
-    f(r_{ij})=-\Theta(r_{ij}-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
-\end{align} 
-where the scalar weights are given by
-\begin{align} 
-      w_3 &= \Theta(r_{ij}-R)  \label{scalar-weight3} \\
-      w_2 &= \delta(r_{ij}-R)  \\
-      w_1 &= \frac{w_2}{4\pi r_{ij}}  \\
-      w_0 &= \frac{w_2}{4\pi r_{ij}^2} \label{scalar-weight0}
-\end{align}
-and the two vector weights are
-\begin{align}
-      \vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{r_{ij}} \\
-      \vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{r_{ij}}
-\end{align}
-which are needed for the relation in Equation~\ref{mayerconvolution} to be satisfied.
-%%SAVE the following - connects to convolution! ------------------------
-%Rewriting the first term V(1) in Equation~\ref{VSR1sum} gives
-%\begin{align}
-       %V(1) &=\left(\int\Theta(r_{ij}-R)d\vec r_{ij}\right)\left(\int \frac{\delta(r_{ij}-R)}{4\pi r_{ij}^2} d\vec r_{ij}\right) \\
-           %%&=\left(\int_{Volume} \Theta(r'-R)d\vec r~'\right)\left(\int_{Volume} \frac{\delta(r"-R)}{4\pi {r"}^2} d\vec r~"\right)   \\
-            %&=\left(\int \Theta(|\vec r~'|-R)d\vec r~'\right)\left(\int \frac{\delta(|\vec r~"|-R)}{4\pi {|\vec r~"|}^2} d\vec r~"\right) 
-           %%&=\int_{Volume} \Theta(|\vec r~'|-R)\left(\int_{Volume} \frac{\delta(|\vec r_{ij}-\vec r~'|-R)}{4\pi {|\vec r_{ij}-\vec r~'|}^2} d\vec r_{ij}\right) d\vec r~'
-%\end{align}
-%Setting $\vec r~"=\vec r-\vec r~'$ and treating $\vec r~'$ as a 
-%constant gives $d\vec r~" = d\vec r$ which is moved to the outside to give 
-%\begin{align}
-       %V(1) &=\int \Theta(|\vec r~'|-R)\left(\int\frac{\delta(|\vec r-\vec r~'|-R)}{4\pi {|\vec r-\vec r~'|}^2} d\vec r\right) d\vec r~'  \\
-            %&=\int \left(\Theta(|\vec r~'|-R)\int \frac{\delta(|\vec r-\vec r~'|-R)}{4\pi {|\vec r-\vec r~'|}^2} d\vec r~'\right) d\vec r  \\
-            %&=\int \left(\int w_3(\vec r~')w_0(\vec r-\vec r~')~d\vec r~'\right) d\vec r  \\
-            %&=\int (w_3\otimes w_0(\vec r)) d\vec r
-%\end{align}
-%%end SAVE -------------------------------------------------------------
-%
-%Limiting this discussion to a fluid composed of hard spheres that have the same
-%radius R, the lowest order term of the virial expansion of the excess Helmholtz 
-%free energy of a homogeneous fluid given in Equation~\ref{eq:Fexcess} 
-%can now be written as
-%\begin{align}
-      %F_{excess} &= k_BTn^2\left(-\frac{1}{2}\int{f(r)}{d\vec{r}}\right)V   \\
-                 %&= k_BTn^2\left(\int \frac{\Theta(r-2R)}{2}~d\vec r\right)V  \\
-                 %&= k_BTn^2(2V(1) + 2SR)V  \\
-                 %&= k_BT\left[\left(\int n\Theta(r-R)d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R^2} d\vec r \right)\color{white} \right]   \nonumber \\
-                 %&+ \color{white}\left[\color{black}\left(\int n\delta(r-R) d\vec r\right)\left(\int n\frac{\delta(r-R)}{4\pi R} d\vec r \right) \right]\color{black}V   \\          
-                 %&= k_BT(n_3n_0 + n_2n_1)V \label{Fexlowden-homo} \\
-                 %&= k_BT(\Phi_1 + \Phi_2)V \\
-                 %&= k_BT\Phi(\{n_\alpha\})V
-%\end{align} 
-%The quantity $\Phi$ has dimensions of $V^{-1}$ and is an energy density 
-%with a dimensionless energy composed 
-%of the sum of $\Phi_1$ and $\Phi_2$ which are defined as
-%\begin{align}
-      %\Phi_1 = n_3n_0  \\
-      %\Phi_2 = n_1n_2
-%\end{align}
-%where $n_i$ are weighted densities of the form
-%\begin{align}
-     %n_\alpha=\int nw_\alpha(r) d\vec r
-%\end{align}
-%The other terms can be written in a similar way with weight functions defined as
-%\begin{align} 
-      %w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
-      %w_2 &= \delta(r-R)  \\
-      %w_1 &= \frac{w_2}{4\pi r}  \\
-      %w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}
-%\end{align}
-%The Heaviside function can then be written as 
-%a convolution of the weight functions which, for a fluid composed of 
-%spheres of radius R, is given by
-%\begin{align}
-    %f(r_{ij})=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
-%\end{align} 
-%where the two vector weights are given by
-%\begin{align}
-      %\vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
-      %\vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
-%\end{align}
-%which are needed for the relation to be satisfied.
-%The excess Helmholtz free energy can also be expressed in terms of the
-%packing fraction $\eta=nV$ as
-%\begin{align}
-     %F_{excess} &= k_BT(n_3n_0 + n_2n_1)V  \\
-                %&= k_BT\left(\left(\eta\right)\left(n\right) + \left(\frac{3\eta}{R}\right)\left(\frac{3\eta}{4\pi R^2}\right)\right)V  \\      
-                %&= k_BT\left(\eta^2+3\eta^2\frac{3}{4\pi R^3}V\right)  \\
-                %&= k_BT4\eta^2
-%\end{align} 
-%
-%For the inhomogeneous case where the number density varies spatially,
-%$n\rightarrow n(\vec r)$, Equation%~\ref{eq:Fexcess} 
-%becomes (for a zero external potential)
-%\begin{align}
-     %F_{excess} = k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
-%\end{align}
-%where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
-%position vector to another sphere, and $r_{ij}=|\vec r_i - \vec r_j|$ 
-%is the distance between the centers of the two spheres.
-%
-%For the inhomogeneous case where the number density varies spatially,
-%$n\rightarrow n(\vec r)$, Equation%~\ref{eq:Fexcess} 
-%becomes (for a zero external potential)
-%\begin{align}
-     %F_{excess} &= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  \\
-                %&= k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}
-%\end{align}
-%where $\vec r_i$ is a position vector of one sphere, $\vec r_j$ is a 
-%position vector to another sphere, and $r_{ij}=|\vec r_i - \vec r_j|$ 
-%is the distance between the centers of the two spheres.
-%Rosenfeld recognized that the Heaviside function can be written as 
-%a convolution of the weight functions which, for a fluid composed of 
-%spheres of radius R, is given by
-%\begin{align}
-    %f(r_{ij})=-\Theta(|\vec r_i-\vec r_j|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})
-%\end{align} 
-%where the weights are the same ones given in Equations~\ref{scalar-weight3} through~\ref{scalar-weight0}, 
-%where the two vector weights are given by
-%\begin{align}
-      %\vec {w}_2 &= w_2~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|} \\
-      %\vec {w}_1 &= w_1~\frac{\vec r_i-\vec r_j}{|\vec r_i-\vec r_j|}
-%\end{align}
-%which are needed for the relation to be satisfied, 
-%and the other weights take the form 
-%The weights take the form
-%\begin{align} 
-      %w_3 &= \Theta(|\vec r_i-\vec r_j|-R)  \label{scalar-weight3} \\
-      %w_2 &= \delta(|\vec r_i-\vec r_j|-R)  \\
-      %w_1 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|}  \\
-      %w_0 &= \frac{w_2}{4\pi |\vec r_i-\vec r_j|^2} \label{scalar-weight0}
-%\end{align}
-%a uniquely deconvolved 
-%into a set of weight functions related to the geometry of a sphere, and that it
-%describes mathematically 
-%the exclusion volume of a pair of spheres, one of the two variables 
-%(the other one being the "overlap volume") in which the total 
-%excess chemical potential can be expanded.
-
-The convolution of two weight functions can be written as
-\begin{align}
-     w_\alpha\otimes w_\gamma(\vec r_i - \vec r_j) &= \int w_\alpha(\vec r') w_\gamma(\vec r_i -\vec r_j - \vec r') d\vec r'   \\                                                             
-                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i -\vec r_j -(\vec u - \vec r_j))d\vec u     \\
-                                                   &= \int w_\alpha(\vec u - \vec r_j) w_\gamma(\vec r_i - \vec u) d\vec u  \\  
-                                                   %&= \int w_\gamma(\vec r_a - \vec u)w_\alpha(\vec u - \vec r_b) d\vec u  \\                                                              
-                                                   &= \int w_\gamma(-(\vec u - \vec r_i))w_\alpha(\vec u - \vec r_j))  d\vec u  
-\end{align}
-For the scalar weights this is the same as
-\begin{align} 
-     w_\alpha\otimes w_\gamma(\vec r_i - \vec r_j) &= \int w_\gamma(\vec u - \vec r_i)w_\alpha(\vec u - \vec r_j)  d\vec u  
-\end{align}
-and for the vector weights this becomes
-\begin{align} 
-     \vec w_\alpha\otimes \vec w_\gamma(\vec r_i - \vec r_j) &= -\int \vec {w}_\gamma(\vec u - \vec r_i)\cdot\vec {w}_\alpha(\vec u - \vec r_j)  d\vec u  
-\end{align}
-%
-%which can also be written as
-%\begin{align}
-     %\color{white}w_\alpha\otimes w_\gamma(\vec r_a - \vec r_b)\color{blue} &=\int w_\alpha(\vec x - \vec r_a) w_\gamma(\vec x - \vec r_b) d\vec x
-%\end{align}
-%since the weight functions involve the distances $|\vec r_a - \vec x|=|\vec x - \vec r_a|$.
-The lowest order term of the virial expansion of the excess Helmholtz 
-free energy (for a zero external potential) is given by
-%The lowest order term of the excess Helmholtz free energy can then be written as
-\begin{align}
-     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\left(\frac{-f(|\vec r_i-\vec r_j|)}{2}\right)  
-\end{align} 
-which can be written as 
-\begin{align}      
-     F_{excess} &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~\frac{\Theta(|\vec r_i-\vec r_j|-2R)}{2}                                                              
-\end{align} 
-\begin{align} 
-                &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)~(w_3\otimes w_0+ w_2 \otimes w_1+\vec {w}_2\otimes \vec {w}_1)  \\
-                &\approx k_BT\int d\vec r_i \int d\vec r_j~n(\vec r_i)~n(\vec r_j)\left(\int w_0(\vec u - \vec r_i)w_3(\vec u - \vec r_j) d\vec u \color{white}\right)  \nonumber \\
-                &+ \color{white}\left(\color{black}\int w_1(\vec u - \vec r_i)w_2(\vec u - \vec r_j) d\vec u ~-\int \vec {w}_1(\vec u - \vec r_i)\cdot\vec {w}_2(\vec u - \vec r_j) d\vec u\right)  \\  
-                &\approx k_BT\int \left[\int n(\vec r_i) w_0(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_3(\vec u - \vec r_j)~d\vec r_j \color{white}\right]  \nonumber \\
-                &+ \int n(\vec r_i) w_1(\vec u - \vec r_i)~d\vec r_i \int n(\vec r_j) w_2(\vec u - \vec r_j)~d\vec r_j  \nonumber \\
-                &- \color{white}\left[\color{black} \int n(\vec r_i) \vec {w}_1(\vec u - \vec r_i)~d\vec r_i \cdot \int n(\vec r_j) \vec {w}_2(\vec u - \vec r_j)~d\vec r_j \right]\color{black}d\vec u  \\
-                &\approx k_BT\int n_0n_3 + n_1n_2 - \vec n_1\cdot\vec n_2~d\vec u  \label{Fexlowden-inhomo}                  
-                %&\approx k_BT\int \Phi_1 + \Phi_2 + \Phi_3~d\vec u  \\
-                %&\approx k_BT\int \Phi(\{n_\alpha(\vec u)\})~d\vec u  \\
-                %&\text{Or, replacing u notation with r} \nonumber \\
-                %&\approx k_BT\int \Phi(\{n_\alpha(\vec r)\})~d\vec r  \\
-                %&\approx k_BT\int \Phi(\vec r)~d\vec r             
-\end{align}
-The weighted densities $n_\alpha$ all depend on the spatially varying 
-number density profile $n(\vec r)$, and have the form
-\begin{align}
-    n_\alpha(\vec r) &= \int n(\vec {r}~')w_\alpha(\vec r-\vec {r}~')~d\vec {r}~'  \label{weighted_densities}  \\
-                     &= n\otimes w_\alpha(\vec r)
-    %n_\alpha(\vec u)=\int n(\vec r')w_\alpha(\vec u-\vec r')~d\vec r'  
-\end{align}  
-and  $\vec n_1\cdot\vec n_2$=0 for the homogenesous case where $n(\vec r)\rightarrow n$.
-%Equation~\ref{Fexlowden-inhomo} reduces to Equation~\ref{Fexlowden-homo}.
-%The excess free energy in the low density limit then takes the form
-%\begin{align}
-%    F_{excess}=k_BT\left(\int n^2(\vec r)(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})~d\vec r\right)V
-%\end{align}
-%%which can now be expressed in terms of a set of weighted densities
-%\begin{align}
-%     F_{excess}=k_BT\left(\int n(\vec r)w_3 \otimes n(\vec r)w_0 + n(\vec r)w_2 \otimes n(\vec r)w_1 + n(\vec r)\vec{w_2} \otimes n(\vec r)\vec{w_1}~d\vec r\right)V
-%\end{align}
-%For a homogeneous fluid in the low density limit, the excess Helmholtz free
-%energy can then be expressed as
-%\begin{align}
-      %F_{excess}=k_BTV\int n^2\frac{\theta(|\vec r|-2R)}{2}~d\vec r
-%\end{align} 
-%
-%Rosenfeld arrived at an expression for the excess Helmholtz free energy 
-%based on a set of weighted densities
-%\begin{align}
-    %F_{excess}=k_BT\int \Phi(\{n_i(\vec r)\})~d\vec r
-%\end{align} 
-%
-%The resulting expression for the excess Helmholtz free energy which is
-%applicable to the more general case of an inhomogeneous fluid, reduces 
-%to the same result obtained from Scaled Particle Theory for a homogeneous 
-%hard-sphere fluid. 
-%
-
-Lifting the restriction to the low density limit, Rosenfeld derived an
-expression for the excess Helmholtz free energy by noting that the higher
-order terms in the virial expansion can all be written in terms of an 
-``exclusion volume" (given by Equation~\ref{exclusion-volume}), 
-and an ``overlap volume" which is the volume of intersection between two 
-overlapping spherical regions.
-The exclusion volume is the spherical volume about the center of one sphere 
-which the center of another sphere can not enter, and is used to describe
-terms with $f_{ij}=f(|\vec r_i - \vec r_j|)$ as has been seen. Used in conjunction 
-with the overlap volume, the higher order terms in the expansion
-in Equation~\ref{eq:Fexcess-simplified}~\cite{rosenfeld1988} can be described. 
-%
-The resulting expression for the excess Helmholtz free energy with Tarazona's $\Phi_3$
-(see Introduction) is given by
-\begin{align}
-    F_{excess}[n(\vec{r})] &= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r} \\
-    %\Phi_1 &= -n_{0}ln(1-n_{3})  \\
-    %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
-    %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}
-    \Phi_1 &= -n_{0}\ln(1-n_{3}) \\
-    \Phi_2 &= \frac{n_{1}n_{2}-\vec{n_{1}}\cdot\vec{n_{2}}}{1-n_{3}} \\
-    \Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}
-\end{align}
-%where in addition to the weight functions already defined,  
-where the weighted densities are given by Equation~\ref{weighted_densities} with 
-weight functions
-\begin{align} 
-      w_3 &= \Theta(r-R)  \label{scalar-weight3} \\
-      w_2 &= \delta(r-R)  \\
-      w_1 &= \frac{w_2}{4\pi r}  \\
-      w_0 &= \frac{w_2}{4\pi r^2} \label{scalar-weight0}  
-\end{align}
-\begin{align}      
-      \vec {w}_2 &= w_2~\hat r \\
-      \vec {w}_1 &= w_1~\hat r \\
-      \overleftrightarrow{w}_{m2}(\vec{r}) &= w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
-\end{align}
-%and a tensor weight function given by
-%\begin{align}\label{eq:tensorweight-methods}
-     %\overleftrightarrow{w}_{m2}(\vec{r}) = w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right) 
-%\end{align}
-%The weighted densities $n_0, n_1, n_2, n_3, \vec{n}_1, \vec{n}_2$ and $\overleftrightarrow{n}_{m2}$
-%all depend on the spatially varying number density profile $n(\vec r)$.
-%\begin{align}%\label{eq:numdenprofile}
-	%n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(\vec{r}-\vec{r'})d{\vec{r'}}}
-%\end{align}
-This result reduces to the same result obtained from Scaled Particle Theory 
-for a homogeneous hard-sphere fluid.
-
-The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
-To model a fluid with "soft" spheres which have an interatomic potential 
-that decreases smoothly from infinity over a distance of 2R rather than instanteously 
-at 2R like that of hard spheres, Schmidt noted that 
-all of the weight functions can be expressed in terms of 
-$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
-He also noted that the derivative of the Mayer function 
-%provides another important relation 
-can be expressed in terms of $w_{2}(r)$, 
-\begin{align}\label{mayerder=w2convolution}
-     \frac{df(r)}{dr} = w_2 \otimes w_2(r) = \int{w_2(r')w_2(r-r')dr'}
-\end{align} 
-and that the weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
-the relation
-\begin{align}\label{w2_w3_relation}
-    w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
-\end{align}
-%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-%Thus, all the weight functions, and the above relation are determined by $w_{3}(r)$.
-Schmidt developed Soft Fundamental Measure Theory (SFMT) which extends FMT to fluids 
-with soft potentials by keeping the relationships between the weight 
-functions, and Equation~\ref{mayerder=w2convolution} 
-%for the derivative of the Mayer function 
-the same, but replacing $w_{3}(r)$ with a weight function appropriate 
-for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
-This is what we do in forming a functional for the Weeks-Chandler-Anderson 
-fluid as detailed in the next chapter on Methods.
-
-%\textcolor{red}{TO HERE ----------------------------------}
-%
+for the hard-sphere fluid. FMT in turn extended  
+Scaled Particle Theory (SPT)~\cite{ReissSPT, santos2012phi3}, which was 
+developed for homogeneous hard-sphere fluids, 
+to inhomogeneous, or non-uniform hard-sphere fluids.
+
 %In order to apply classical Density Functional Theory to a system that has
@@
 
-%\textcolor{red}{BEGIN COMMENTED OUT SECTION...} %########################
-
-%Scaled Particle Theory
-%seeks a general expression for the excess chemical 
-%potential of a homogeneous hard-sphere fluid which can then be integrated 
-%to find the excess Helmholtz free energy according to the relation
-%\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
-%It is an extension of the work done by Widom who showed that the excess 
-%chemical potential of a homogeneous fluid corresponds to the reversible 
-%work of inserting a sphere of radius R into a fluid consisting of hard 
-%spheres of radius R. 
-%\begin{align}
-	%\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
-%\end{align}
-%Widom's result was obtained by considering the difference 
-%in the excess Helmholtz free energy when one atom is added to an otherwise 
-%closed system. He showed that the excess chemical potential is related to 
-%the ensemble average of the Boltzman factor for the potential energy 
-%difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
-%The resulting equation, shown below, is known as Widom's Insertion 
-%Formula~\cite{Hansen}.
-%\begin{align} \label{mu-der-int}
-    %\mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
-%\end{align}
-%\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+Scaled Particle Theory
+seeks a general expression for the excess chemical 
+potential of a homogeneous hard-sphere fluid which can then be integrated 
+to find the excess Helmholtz free energy according to the relation
+\begin{equation}\mu_{ex}=\frac{\partial{F_{ex}}}{\partial{N}}\bigg|_{T,V}{~~~~~}\rightarrow{~~~~~}F_{ex}=\int{\mu_{ex}dN}\end{equation}
+It is an extension of the work done by Widom who showed that the excess 
+chemical potential of a homogeneous fluid corresponds to the reversible 
+work of inserting a sphere of radius R into a fluid consisting of hard 
+spheres of radius R. 
+\begin{align}
+	\mu_{ex}=W_{rev}(R){~~~~}\text{work to insert a hard sphere of radius R}
+\end{align}
+Widom's result was obtained by considering the difference 
+in the excess Helmholtz free energy when one atom is added to an otherwise 
+closed system. He showed that the excess chemical potential is related to 
+the ensemble average of the Boltzman factor for the potential energy 
+difference $V_{diff}$ between a system with N atoms and one with N+1 atoms. 
+The resulting equation, shown below, is known as Widom's Insertion 
+Formula~\cite{Hansen}.
+\begin{align} \label{mu-der-int}
+    \mu_{ex}=\frac{F_{ex}(N+1)-F_{ex}(N)}{\Delta{N}}\bigg|_{T,V}
+\end{align}
+\begin{equation}\label{widoms-insertion-formula}{~}=-k_BT\ln\left(<e^{-\beta{V_{diff}}}>\right)\end{equation}
+where 
+\begin{align}
+    <e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
+\end{align}
+and
+\begin{align}
+    Z=\sum_s{e^{-\beta V_s}}
+\end{align}
+
+The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
+there will be a change in the potential energy of $V_{diff}$ if one sphere
+is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
+In this case, $P_s=\frac{e^{-\beta V_s}}{Z}$.
+%at each point in the fluid, there is a probability that 
+%a hard sphere of radius R has space available to exist at that point
+%without overlapping with other spheres, or it does not, as shown in 
+Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, 
+then becomes~\cite{Hansen}
+\begin{align} 
+	\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
+	         &= V_{diff} \\
+             &= W_{rev}(R)  \label{workrev}
+\end{align}
+\begin{figure}
+    \centering
+    \includegraphics[height=7cm]{figs/P_overlap.pdf}
+    \caption{No energy is needed to place a sphere at a position where 
+    there is no overlap, such as that shown by the black dot near the 
+    top. But it is not possible to place a sphere at a position where 
+    an overlap occurs.}
+    \label{fig:p_overlap}
+  \end{figure}
+The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
+that it takes no energy for a hard sphere to be placed at a point
+in the fluid where it does not overlap with any other spheres 
+($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
+to place a sphere where it would overlap another sphere, as described 
+in Figure~\ref{fig:p_overlap}. 
+In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
+\begin{align} \label{muexcess}
+  \mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
+\end{align}
+where the radius of the cavity $R_{cavity}$ equals the radius of a 
+hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
+occupied by the spheres divided by the total volume available, 
+is the probability that a sphere cannot be placed at some random
+point in the homogeneous fluid of number density, $n$, without
+overlapping with another sphere. $P_s$ would then have values of
+$nV_{Sphere}$ and $1-nV_{Sphere}$.
+Combining  Equations~\ref{workrev} 
+and \ref{muexcess} gives
+\begin{align}
+  %W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3)
+  W_{rev}(R_{cavity}=R)=-k_BT\ln(1-\eta) 
+\end{align}
+where $\eta$ is the packing fraction defined as
+\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
+\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
+
+Scaled Particle Theory forms a more general result for the reversible work 
+as a function of the radius of the cavity by considering additional 
+cases where the cavity is larger, and much larger than the radius of 
+a hard sphere, and requiring continuity of the functions for the 
+different cases and their derivatives (first and second). 
+By setting the radius of the cavity to the hard sphere radius, the 
+more general work equation becomes equal to the chemical potential. 
+Details of this procedure can be found 
+in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.  
+An expression for the excess Helmholtz free energy in terms of the
+packing fraction $\eta$ can then be obtained through 
+Equation~\ref{mu-der-int}. 
+%
+%By requiring continuity of the functions valid for different cavity 
+%sizes, along with their first and second derivatives, an equation for 
+%the work is obtained which can then be used to form 
+%set equal to the excess chemical 
+%potential by setting the cavity radius to the radius of hard sphere. 
+%
+%Forming a more general result by requiring continuity of the functions, 
+%along with their first and second derivatives, for the reversible work
+%written for fluids with a cavity radius equal to, larger than, or very 
+%much larger than the hard sphere radius R, 
+%
+%Scaled Particle Theory arrives at 
+%an expression for the excess Helmholtz free energy of 
+%a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
+%For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
+\begin{align}\label{Fexcess-SPT}
+    F_{excess}=k_BT\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)N
+\end{align}
+
+Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
+%and provides an expression for the excess Helmholtz free energy of an 
+%inhomogeneous 
+%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
+%
+%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
+%is an extension of FMT and provides an expression for the excess Helmholtz
+%free energy of fluids with soft potentials like that of the WCA fluid. 
+%It will be helpful to continue looking at FMT before addressing SFMT.
+%FMT begins 
+by noticing that a quantity $G_{3}$ can be defined such that
+\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
+where $G_{3}$ is the volume of a sphere. 
+Likewise, by defining quantities
+\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
+\begin{equation}{G_{1}=R}\end{equation}
+\begin{equation}{G_{0}=1}\end{equation}
+where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
+radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
+\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
+The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
+geometric measures of a sphere, and are what give rise to the name
+Fundamental Measure Theory. Introducing a set of variables
+%weighted densities $n_{i}=nG_{i}$ 
+given by
+\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
+\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
+\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
+\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
+Equation~\ref{FexfromSPTinthermsofG} can be written as
+\begin{align}\label{FexfromSPT}
+    F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
+\end{align}
+which has the form 
+\begin{align}
+    F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
+\end{align}
 %where 
-%\begin{align}
-    %<e^{-\beta{V_{diff}}}>=\sum_s{P_se^{-\beta V_s}}
-%\end{align}
-%and
-%\begin{align}
-    %Z=\sum_s{e^{-\beta V_s}}
-%\end{align}
-
-%The quantity $<e^{-\beta{V_{diff}}}>$ can be evaluated by considering that
-%there will be a change in the potential energy of $V_{diff}$ if one sphere
-%is added, but no change in potential energy if no sphere is added ($V_{diff}=0$). 
-%In this case, $P_s=\frac{e^{-\beta V_s}}{Z}$.
-%%at each point in the fluid, there is a probability that 
-%%a hard sphere of radius R has space available to exist at that point
-%%without overlapping with other spheres, or it does not, as shown in 
-%Widom's Insertion formula Equation~\ref{widoms-insertion-formula}, 
-%then becomes~\cite{Hansen}
-%\begin{align} 
-	%\mu_{ex} &= -k_BT\ln\left(e^{-\beta{V_{diff}}}+1\right)  \\
-	         %&= V_{diff} \\
-             %&= W_{rev}(R)  \label{workrev}
-%\end{align}
-%\begin{figure}
-    %\centering
-    %\includegraphics[height=7cm]{figs/P_overlap.pdf}
-    %\caption{No energy is needed to place a sphere at a position where 
-    %there is no overlap, such as that shown by the black dot near the 
-    %top. But it is not possible to place a sphere at a position where 
-    %an overlap occurs.}
-    %\label{fig:p_overlap}
-  %\end{figure}
-%The quantity $<e^{-\beta{V_{diff}}}>$ can also be evaluated by considering
-%that it takes no energy for a hard sphere to be placed at a point
-%in the fluid where it does not overlap with any other spheres 
-%($V_{diff}=0$), but it would require infinite energy ($V_{diff}=\infty$)
-%to place a sphere where it would overlap another sphere, as described 
-%in Figure~\ref{fig:p_overlap}. 
-%In this case Equation~\ref{widoms-insertion-formula} becomes~\cite{Hansen}
-%\begin{align} \label{muexcess}
-  %\mu_{excess}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3) 
-%\end{align}
-%where the radius of the cavity $R_{cavity}$ equals the radius of a 
-%hard sphere R, and $nV_{Sphere}$, which is equal to the total volume 
-%occupied by the spheres divided by the total volume available, 
-%is the probability that a sphere cannot be placed at some random
-%point in the homogeneous fluid of number density, $n$, without
-%overlapping with another sphere. $P_s$ would then have values of
-%$nV_{Sphere}$ and $1-nV_{Sphere}$.
-%Combining  Equations~\ref{workrev} 
-%and \ref{muexcess} gives
-%\begin{align}
-  %%W_{rev}(R_{cavity}=R)=-k_BT\ln(1-n\frac{4\pi}{3}R^3)
-  %W_{rev}(R_{cavity}=R)=-k_BT\ln(1-\eta) 
-%\end{align}
-%where $\eta$ is the packing fraction defined as
-%\begin{equation}{\eta = \frac{\mbox{Volume occupied by the spheres}}{\mbox{Total Volume}}}\end{equation}
-%\begin{equation}{\eta = \frac{N\frac{4}{3}\pi{R^3}}{V}}\end{equation}
-
-%Scaled Particle Theory forms a more general result for the reversible work 
-%as a function of the radius of the cavity by considering additional 
-%cases where the cavity is larger, and much larger than the radius of 
-%a hard sphere, and requiring continuity of the functions for the 
-%different cases and their derivatives (first and second). 
-%By setting the radius of the cavity to the hard sphere radius, the 
-%more general work equation becomes equal to the chemical potential. 
-%Details of this procedure can be found 
-%in Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}.  
-%An expression for the excess Helmholtz free energy in terms of the
-%packing fraction $\eta$ can then be obtained through 
-%Equation~\ref{mu-der-int}.
-
-%%By requiring continuity of the functions valid for different cavity 
-%%sizes, along with their first and second derivatives, an equation for 
-%%the work is obtained which can then be used to form 
-%%set equal to the excess chemical 
-%%potential by setting the cavity radius to the radius of hard sphere. 
-%%
-%%Forming a more general result by requiring continuity of the functions, 
-%%along with their first and second derivatives, for the reversible work
-%%written for fluids with a cavity radius equal to, larger than, or very 
-%%much larger than the hard sphere radius R, 
-%%
-%%Scaled Particle Theory arrives at 
-%%an expression for the excess Helmholtz free energy of 
-%%a homogeneous hard-sphere fluid in terms of the packing fraction $\eta$.
-%%For details, see Theory of Simple Fluids by Hansen and McDonald~\cite{Hansen}. 
-%\begin{align}\label{Fexcess-SPT}
-    %F_{excess}=k_BT\left(-ln(1-\eta)+\frac{3\eta}{1-\eta}+\frac{3{\eta}^2}{2(1-\eta)^2}\right)N
-%\end{align}
-
-%Fundamental Measure Theory can be constructed from Equation~\ref{Fexcess-SPT}  
-%%and provides an expression for the excess Helmholtz free energy of an 
-%%inhomogeneous 
-%%hard-sphere fluid as a functional of the number density profile $n(\vec r)$. 
-%%
-%%Soft Fundamental Measure Theory (SFMT), which will be used in this paper, 
-%%is an extension of FMT and provides an expression for the excess Helmholtz
-%%free energy of fluids with soft potentials like that of the WCA fluid. 
-%%It will be helpful to continue looking at FMT before addressing SFMT.
-%%FMT begins 
-%by noticing that a quantity $G_{3}$ can be defined such that
-%\begin{equation}{\eta =\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=nG_{3}}\end{equation}
-%where $G_{3}$ is the volume of a sphere. 
-%Likewise, by defining quantities
-%\begin{equation}{G_{2}=4\pi{R^2}}\end{equation}
-%\begin{equation}{G_{1}=R}\end{equation}
-%\begin{equation}{G_{0}=1}\end{equation}
-%where $G_{2}$ is the surface area of a sphere, and $G_{1}$ is the 
-%radius of a sphere, Equation~\ref{Fexcess-SPT} can be written as
-%\begin{equation}\label{FexfromSPTinthermsofG}{F_{excess}=k_{B}T\left(-nG_{0}ln(1-nG_{3})+\frac{nG_{1}nG_{2}}{1-nG_{3}}+\frac{nG_{2}^3}{24\pi(1-Gn_{3})^2}\right)V}\end{equation}
-%The quantities $G_{3}, G_{2}, G_{1},$ and $G_{0}$ are fundamental 
-%geometric measures of a sphere, and are what give rise to the name
-%Fundamental Measure Theory. Introducing a set of variables
-%%weighted densities $n_{i}=nG_{i}$ 
-%given by
-%\begin{equation}\label{n3}{n_{3}=nG_{3}=\left(\frac{N}{V}\right)\frac{4}{3}\pi{R^3}=\eta}\end{equation}
-%\begin{equation}\label{n2}{n_{2}=nG_{2}=\left(\frac{N}{V}\right)4\pi{R^2}=\frac{3\eta}{R}}\end{equation}
-%\begin{equation}\label{n1}{n_{1}=nG_{1}=\left(\frac{N}{V}\right)R}\end{equation}
-%\begin{equation}\label{n0}{n_{0}=nG_{0}=\left(\frac{N}{V}\right)}\end{equation}
-%Equation~\ref{FexfromSPTinthermsofG} can be written as
-%\begin{align}\label{FexfromSPT}
-    %F_{excess}=k_{B}T\left(-n_{0}ln(1-n_{3})+\frac{n_{1}n_{2}}{1-n_{3}}+\frac{n_{2}^3}{24\pi(1-n_{3})^2}\right)V
-%\end{align}
-%which has the form 
-%\begin{align}
-    %F_{excess} = k_BT(\Phi_1+\Phi_2+\Phi_3)V
-%\end{align}
-%%where 
-%\begin{align}
-   %\Phi_1 &= -n_{0}\ln(1-n_{3}) \label{Phi1equation} \\
-   %\Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
-   %\Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
-%\end{align}
-
-%The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
-%densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
-%\begin{align}
-    %n_{i}=nG_{i}
-%\end{align}
-%These weights are valid for determining the excess free energy of a system of 
-%hard spheres with a homogeneous number density n. But these weights, and 
-%their related  weighted densities, can be written in a more general form 
-%allowing for their extension to a system with a spatially varying number 
-%density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
-%excess Helmholtz free energy is written as
-%\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
-%Using a backwards Heaviside step function defined as
-%\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
-%the following expressions are obtained
-%\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
-%\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
-%\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
-%\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
-%where the weighted densities can now be expressed as
-%\begin{align}
-    %n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
-%\end{align} 
-%or for a spatially varying number density $n(\vec{r})$,
-%\begin{equation}\label{weighted-density}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
-%with weight functions
-%\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
-%\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
-%\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
-%\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
-
-%The weight functions $w_{2}$ and $w_{3}$ have the relation
-%\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
-%which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
-
-%The weight functions can be related to the Mayer function 
-%\begin{align}
-     %f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
-%\end{align} 
-%where r is the distance between the centers of two hard spheres, and 
-%$V(r)$ is the interatomic potential between two hard spheres given by 
-%\begin{align}
-    %V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
-%\end{align}
-%Evaluating Equation with $V(r)$ given by Equation,  
-%the Mayer function for a hard-sphere fluid is found to be
-%\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
-%which can be written in terms of the weight functions if two 
-%additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
-%\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
-%\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
-%The Mayer function can then be written in terms of the weight functions~\cite{Hansen}
-%\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
-
-%Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
-%another relation that will be important to keep in tact when moving to 
-%SFMT~\cite{schmidt2000fluid}.
-%\begin{align}\label{mayerder=w2convolution}
-     %\frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
-%\end{align} 
-%This relation can easily be checked for a hard-sphere fluid where
-%\begin{equation}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{equation}
-%\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
-%The right side can be evaluated by using the general relation for 
-%convolution which is equal to the Inverse Fourier Transform 
-%of a product of Fourier Transforms. 
-%\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
-%Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
-%gives
-%\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
-%and the right side of Equation~\ref{mayerder=w2convolution} becomes
-%\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
-%\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
-%which matches Equation~\ref{mayerder}.
-
-%Tarazona later improved upon the original FMT equations by replacing the
-%expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
-%more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
-%\begin{align}
-%\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
-%\end{align}
-%in which a tensor weight is incorporated given by
-%\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
-%%All the other FMT equations remain the same.
-
-%The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
-%All of the weight functions, the relations between them, and the equation 
-%for the derivative of the Mayer function, can be expressed in terms of 
-%$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
-%The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
-%Equation~\ref{w2_w3_relation}.
-%Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
-%with soft potentials by keeping the relationships between the weight 
-%functions, and the equation for the derivative of the Mayer function 
-%the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
-%~\ref{w2_w3_relation},~\ref{eq:tensorweight},~\ref{mayerder=w2convolution}), 
-%but replacing $w_{3}(r)$ with a weight function appropriate 
-%for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
-%This is what we do in forming a functional for the Weeks-Chandler-Anderson 
-%fluid as detailed in the next chapter on Methods.
+\begin{align}
+   \Phi_1 &= -n_{0}ln(1-n_{3}) \\
+   \Phi_2 &= \frac{n_{1}n_{2}}{1-n_{3}} \\
+   \Phi_3 &= \frac{n_{2}^3}{24\pi(1-n_{3})^2}   \label{oldPhi3equation}
+\end{align}
+
+The quantities $n_{0},n_{1},n_{2},n_{3}$ can be considered as weighted 
+densities where the weights are given by $G_{0},G_{1},G_{2},G_{3}$ respectively.
+\begin{align}
+    n_{i}=nG_{i}
+\end{align}
+These weights are valid for determining the excess free energy of a system of 
+hard spheres with a homogeneous number density n. But these weights, and 
+their related  weighted densities, can be written in a more general form 
+allowing for their extension to a system with a spatially varying number 
+density $n(\vec{r})$, that is, an inhomogeneous fluid, in which case, the
+excess Helmholtz free energy is written as
+\begin{equation}\label{eq:F_ex-n-of-r}{F_{excess}[n(\vec{r})]= k_BT\int(\Phi_1(\vec{r})+\Phi_2(\vec{r})+\Phi_3(\vec{r}{)) d}\vec{r}}\end{equation} 
+Using a backwards Heavyside step function defined as
+\begin{equation}{\Theta(|\vec{r}|-R)=\left\{ \begin{array}{rc} 1 & 0<r \leq R \\ 0  & r>R \end{array}\right.}\end{equation}
+the following expressions are obtained
+\begin{equation}{G_{3}=\int_{Volume}{\Theta(|\vec{r}|-R)d{\vec{r}}} = \frac{4}{3}\pi{R^3}}\end{equation}
+\begin{equation}{G_{2}=\int_{Volume}{\delta(|\vec{r}|-R)d{\vec{r}}} = 4\pi{R^2}}\end{equation}
+\begin{equation}{G_{1}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r}d{\vec{r}}} = R}\end{equation}
+\begin{equation}{G_{0}=\int_{Volume}{\frac{\delta(|\vec{r}|-R)}{r^2}d{\vec{r}}} = 1}\end{equation}
+where the weighted densities can now be expressed as
+\begin{align}
+    n_{i}=\int_{Volume}{nw_{i}(r)}d{\vec{r}}
+\end{align} 
+or for a spatially varying number density $n(\vec{r})$,
+\begin{equation}{n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}}\end{equation}
+with weight functions
+\begin{equation}\label{eq:w3}{w_{3}=\Theta(|\vec{r}|-R)}\end{equation}
+\begin{equation}\label{eq:w2}{w_{2}=\delta(|\vec{r}|-R)}\end{equation}
+\begin{equation}\label{eq:w1}{w_{1}=\frac{w_{2}}{4\pi{r}}}\end{equation}
+\begin{equation}\label{eq:w0}{w_{0}=\frac{w_{2}}{4\pi{r}^2}}\end{equation}
+The weight functions $w_{2}$ and $w_{3}$ have the relation
+\begin{equation}\label{w2_w3_relation}{w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)}\end{equation}
+which will be important to keep in tact when moving to SFMT~\cite{schmidt1999density}.
+
+The weight functions can be related to the Mayer function 
+\begin{align}
+     f(r)=\exp^{-\frac{V(r)}{k_{B}T}}-1
+\end{align} 
+where r is the distance between the centers of two hard spheres, and 
+$V(r)$ is the interatomic potential between two hard spheres given by 
+\begin{align}
+    V(r)=\left\{ \begin{array}{rc} \infty & 0<r \leq 2R \\ 0  & r>2R \end{array}\right.
+\end{align}
+Evaluating Equation with $V(r)$ given by Equation,  
+the Mayer function for a hard-sphere fluid is found to be
+\begin{equation}\label{f(r)step}{f(r)=-\Theta(|\vec{r}|-2R)}\end{equation} 
+which can be written in terms of the weight functions if two 
+additional vector weight functions $\vec{w_2}$ and $\vec{w_1}$ are introduced
+\begin{equation}\label{eq:w_v2}{\vec{w_{2}}=w_{2}\frac{\vec{r}}{r}}\end{equation}
+\begin{equation}\label{eq:w_v1}{\vec{w_{1}}=w_{1}\frac{\vec{r}}{r}}\end{equation} 
+The Mayer function can then be written in terms of the weight functions~\cite{Hansen}
+\begin{equation}\label{mayer_deconvolution}{f(r)=-\Theta(|\vec{r}|-2R)= -2(w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1})}\end{equation}
+
+Taking the derivative of the Mayer function in Eq~\ref{f(r)step} provides 
+another relation that will be important to keep in tact when moving to 
+SFMT~\cite{schmidt2000fluid}.
+\begin{align}\label{mayerder=w2convolution}
+     \frac{df(r)}{dr} = w_2 \otimes w_2 = \int{w_2(r')w_2(r-r')dr'}
+\end{align} 
+This relation can easily be checked for a hard-sphere fluid where
+\begin{equation}{\frac{df(r)}{dr}=\left\{ \begin{array}{rc} \frac{d}{dr}(-1)=0 & r<2R \\\frac{d}{dr}(\text{vertical line})=+\infty & r=2R \\ \frac{d}{dr}(0)=0  & r>2R \end{array}\right.}\end{equation}
+\begin{equation}\label{mayerder}{\frac{df(r)}{dr} = \delta(r-2R)}\end{equation} 
+The right side can be evaluated by using the general relation for 
+convolution which is equal to the Inverse Fourier Transform 
+of a product of Fourier Transforms. 
+\begin{equation}{f \otimes g = \text{I.F.T.[ F.T.}[f]\text{ x F.T.[g] }]}\end{equation} 
+Taking the Fourier Transform of the weight function $w_2(r)=\delta(r-R)$ 
+gives
+\begin{equation}{\text{F.T.}[w_2]=\int_{-\infty}^{\infty}\delta(r-R)e^{-ikr}dr=e^{-ikR}}\end{equation} 
+and the right side of Equation~\ref{mayerder=w2convolution} becomes
+\begin{equation}{w_2 \otimes w_2 = \text{I.F.T.}[e^{-ikR}e^{-ikR}]}\end{equation} 
+\begin{equation}\label{w2convolution}{=\int_{-\infty}^{\infty}e^{-ik2R}e^{ikr}dk=\int_{-\infty}^{\infty}e^{ik(r-2R)}dk=\delta(r-2R)}\end{equation} 
+which matches Equation~\ref{mayerder}.
+
+Tarazona later improved upon the original FMT equations by replacing the
+expression for $\Phi_3$ given by Equation~\ref{oldPhi3equation} with the
+more accurate expression~\cite{tarazonaphi3, santos2012phi3} 
+\begin{align}
+\Phi_3 &= \frac{{n_2}^3-3n_2\vec{n}_{v2}\cdot\vec{n}_{v2}+\frac{9}{2}[\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}^3_{m2}})]}{24\pi(1-n_3)^2}  \label{eq:Phi3}
+\end{align}
+in which a tensor weight is incorporated given by
+\begin{equation}\label{eq:tensorweight}{\overleftrightarrow{w}_{m2}(\vec{r})=w_2(r)\left(\frac{\vec{r}\vec{r}}{r^2}-\frac{I}{3}\right)}\end{equation} 
+All the other FMT equations remain the same.
+
+The FMT equations shown so far have been for a hard-sphere fluid~\cite{Hansen}.
+All of the weight functions, the relations between them, and the equation 
+for the derivative of the Mayer function, can be expressed in terms of 
+$w_{2}(r)$ which, for a hard-sphere fluid, consists of a delta function.
+The weight function $w_{2}(r)$ is in turn related to $w_{3}(r)$ through 
+Equation~\ref{w2_w3_relation}.
+Soft Fundamental Measure Theory extends FMT discussed so far to fluids 
+with soft potentials by keeping the relationships between the weight 
+functions, and the equation for the derivative of the Mayer function 
+the same (Equations~\ref{eq:w0},~\ref{eq:w1},~\ref{eq:w_v1},~\ref{eq:w_v2},
+~\ref{w2_w3_relation},~\ref{eq:tensorweight},~\ref{mayerder=w2convolution}), 
+but replacing $w_{3}(r)$ with a weight function appropriate 
+for a soft-sphere potential~\cite{schmidt1999density, schmidt2000fluid}. 
+This is what we do in forming a functional for the Weeks-Chandler-Anderson 
+fluid as detailed in the next chapter on Methods.
 
@@
 have attraction between the particles that make up the fluid will show
-two distinct regions - a fluid region at lower densities, and a solid 
+two distinct regions - a "liquid" region at lower densities, and a solid 
 region at higher densities. 
-A phase transition between a fluid and a solid is accompanied by an 
+A phase transition between a liquid and a solid is accompanied by an 
 abrupt change in the number density, and so there is a gap between the 
-purely fluid and the purely solid regions. Within the gap, or coexistence
-region, is found mixtures of various amounts of fluid and solid each of
+purely liquid and the purely solid regions. Within the gap, or coexistence
+region, is found mixtures of various amounts of liquid and solid each of
 which have number densities that are fixed at the values they had at the 
 phase transition.
-To construct the phase diagram, the fluid and solid phase transition 
-number densities are found 
+
+To construct the phase diagram, the liquid and solid phase transition number densities
+are found
+
 %Maxwell's double tangent construction can be used to find 
 %these number densities given plots of
-from the fluid and the solid Helmholtz 
+
+from the liquid and the solid Helmholtz 
 free energies and inverse densities at a given temperature, as
@@
 We use classical Density Functional Theory 
-to find the Helmholtz free energies of the fluid and the solid. 
+to find the Helmholtz free energies of the liquid and the solid. 
 A functional for the ideal portion of the
@@
 of positions about the lattice point. 
-Atoms are always in motion, so when talking about ``the location" 
+Atoms are always in motion, so when talking about "the location" 
 of an atom, we are talking about a time averaged position, so that it 
@@
 density. The number density is the number of atoms per volume, whereas
-the number density profile describes how much ``atom" will be found at various 
+the number density profile describes how much "atom" will be found at various 
 points in the cell. 
 It can be thought of as an average over a large collection of actual cells
-with various possible spatial distributions of the atoms within the cell, or
+of various possible spatial distributions of the atoms within a cell, or
 more precisely, as an ensemble average over all possible microstates, 
-each of which consists of a unique set of positions for the atoms 
+each of which consists of a unique set of positions for the atoms in 
 that occupy the cell. Mathematically, the microstates are represented 
@@
 \begin{align}
-    n(\vec r)=~\left<\sum_{i=1}^N\delta(\vec r - \vec r_i)\right>
+    n(\vec r)=~<\sum_{i=1}^N\delta(\vec r - \vec r_i)>~
 \end{align} 
@@
 \begin{align}
-    n=\left<n(\vec{r})\right>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
+    n=<n(\vec{r})>=\frac{1}{V}\int{n(\vec{r})}{d\vec{r}}=\frac{N}{V}
 \end{align}
@@
    \caption{On the left are shown four examples of possible positions 
-   that an atom (blue sphere) might have within a volume 
+   that an atom (blue sphere) might have within a fixed volume 
    containing eight cells. Each cell has one lattice site at the center 
@@
    give rise to the same Gaussian distribution statistically, 
-   and the same number density. In reality, the fraction of vacancies will
-   be very small.}
+   and the same number density.}
    \label{fig:SameStatPic}
@@
 
-There is another way to generate different number density profiles for 
-a given number density besides changing the width of the Gaussians. 
+Besides changing the width of the Gaussians, there is another way to 
+generate different number density profiles for a given number density.
 It may be that not all lattice sites are occupied. Vacancies can be taken 
@@
 number density, as shown by the example in Figure~\ref{fig:Ensemble_vary}.
-A homogeneous number density profile is used for a fluid, as shown in 
+A homogeneous number density profile is used for a liquid, as shown in 
 Figure \ref{fig:homogen_denisty}.
@@
 density profile in the set, and also for the homogeneous number density profile 
-characteristic of a fluid. Both the ideal and excess free energies are 
+characteristic of a liquid. Both the ideal and excess free energies are 
 found, and added together to get the total Helmholtz free energy.
@@
 
-Once the Helmholtz free energy for both the crystal and fluid 
-are found for a given temperature and density, they can be 
+Once the Helmholtz free energy for both the crystal and liquid states of 
+the fluid are found for a given temperature and density, they can be 
 used to construct phase diagrams.
-Outside the region of coexistence, a purely crystalline state will exist 
-if the crystal free energy is lower than the fluid free energy, and 
-a purely fluid state will exist if the fluid free energy is lower than 
+Outside the region of coexistence which exists between the liquid and 
+crystal states, the fluid will be completely in a crystalline state if the 
+crystal free energy is lower than the liquid free energy, and it will be 
+completely in a liquid state if the liquid free energy is lower than 
 the crystal free energy. 
-After the actual number densities of the crystal and fluid states
-at the boundaries of the coexistence region are determined for a given 
-temperature, phase diagrams can be constructed as described in Section 2.2.
+We determine the actual number densities of the crystal and liquid states
+at the boundaries of the coexistence region for a given temperature, 
+and construct phase diagrams as described in Section 2.2.
 
@@
 of the Helmholtz free energy is difficult since it depends on the interatomic 
-potential energy $V(r)$ that exists between each sphere and all the other spheres, 
+potential energy $V(r)$ that exists between each sphere with all the other spheres, 
 as is shown by Equations~\ref{eq:mayerfunction} and \ref{eq:Fexcess-simplified}.
@@
 \end{align} 
-The weighted densities $n_0, n_1, n_2, n_3, \vec {n}_1, \vec {n}_2$ and $\overleftrightarrow{n}_{m2}$ are found from
+The weighted densities $n_0, n_1, n_2,$ and $n_3$ are found from
 \begin{align}\label{eq:numdenprofile}
-	n_\alpha(\vec{r})=\int_{Volume}{n(\vec{r'})w_\alpha(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
+	n_i(\vec{r})=\int_{Volume}{n(\vec{r'})w_i(|\vec{r}-\vec{r'}|)d{\vec{r'}}}
 \end{align}
@@
 and $a$ with our temperature dependent parameters $\alpha$ and $\Xi$.
-%The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
-%function potential $V_{erf}(r)$ approximate, in the region of interest, 
-%the temperature-independent WCA potential which is difficult 
-%to implement computationally. 
-The two parameters $\alpha$ and $\Xi$ are designed to make the functional
-produce results which approximate, in the region of interest, that of  
-a WCA fluid using the temperature-dependent error function potential $V_{erf}(r)$ 
-instead of the temperature-independent WCA potential which is difficult 
+The two parameters $\alpha$ and $\Xi$ make the temperature-dependent error 
+function potential $V_{erf}(r)$ approximate, in the region of interest, 
+the temperature-independent WCA potential which is difficult 
 to implement computationally. 
@@
 in Equations~\ref{w2_w3_relation} and~\ref{mayerder=w2convolution}.
-Using Equation~\ref{w2_w3_relation}(repeated below),
+Using Equation~\ref{w2_w3_relation} repeated here,
 \begin{displaymath}
-     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivalently,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
+     w_{3}=\int_{r}^{\infty}{w_{2}(\vec{r})dr}\mbox{~~~~or equivilantly,~~~~}-\frac{\partial{w_3(r)}}{\partial{r}}=w_2(r)
 \end{displaymath}
@@
   \centering
-  \includegraphics[height=7cm]{plot_alpha.pdf}
+  \includegraphics[height=8cm]{plot_alpha.pdf}
   \caption{Shows the temperature dependence of $\alpha$ vs $k_BT$ 
@@
 
-Unlike the WCA potential, the error function potential $V_{erf}(r)$ is temperature dependent.
-To compensate for this temperature dependence so that our functional better approximates
-a WCA potential, we make the parameters $\alpha$ and $\Xi$ functions of temperature as well. 
 The parameter $\alpha(T)$ is derived by setting 
 the error function potential $V_{erf}(r)$ 
-equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$, since 
-the derivative of the mayer function peaks at this value. 
+equal to the WCA potential given by Equation~\ref{eq:VWCA} at $r=\alpha$. 
 \begin{align}
@@
   \centering
-  \includegraphics[height=7cm]{xi_fromB2.pdf}
+  \includegraphics[height=8cm]{xi_fromB2.pdf}
   \caption{Shows the temperature dependence of $\Xi$ generated from 
@@
 along the z-axis. In three dimensions, the weighting functions 
-$w_0$, $w_1$, $w_2$ are spherical shells, 
+$w_0$, $w_1$, $w_2$ are spherical shells centered at $\vec{r}$, 
 and $w_3$ is a solid sphere with an edge that tapers off.
 $\Xi$ is related to the width of the weight functions which 
-become thicker, or taper off longer as the temperature increases, and 
+become thicker, or taper off longer as the temperature increases. 
 $\alpha$ is related to the radius of the shell, or sphere, and decreases 
@@
 
-\section{Computing the Fluid and Crystal Helmholtz free energies}
-The Helmholtz free energy is composed of two parts: an ideal Helmholtz free energy 
-and an excess Helmholtz free energy that are added together to get the total
-Helmholtz free energy, $F=F_{ideal} + F_{excess}$. 
-And so the ideal and excess Helmholtz free energies
-must be computed for both the fluid and the crystal states.
-The fluid ideal Helmholtz free energy is computed analytically 
-using Equation~\ref{eq:Fideal} for a given homogeneous reduced 
-(or dimensionless) number density $n^*$ and temperature $t^*$. 
+\section{Computing the Liquid and Crystal Helmholtz free energies}
+
+The liquid ideal Helmholtz free energy is computed analytically 
+using Equation~\ref{eq:Fideal} 
+for a given homogeneous reduced density $n^*$, and reduced temperature $t^*$. 
 The crystal ideal Helmholtz free energy is computed computationally 
@@
 where $\vec R_j$ is the position vector of the jth lattice site.
+The liquid excess Helmholtz free energy is computed analytically  
+using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
+with a uniform number density profile, n that does not depend on $\vec r$.
 
@@
 
-%\begin{figure}
-   %\centering
-   %\includegraphics[height=5cm]{InclusionRadius.png}
-   %\caption{Gaussians are centered at each lattice point. After a radius 
-   %of a few Gaussian widths about each lattice point, the Gaussians go to 
-   %zero and can be neglected.}
-   %\label{fig:InclusionRadius}
-%\end{figure}
-
-The fluid excess Helmholtz free energy is computed analytically  
-using SFMT Equations~\ref{eq:Fexfunctional} through \ref{eq:tensorweight-methods} 
-with a uniform number density profile, $n$ that does not depend on $\vec r$. 
+\begin{figure}
+   \centering
+   \includegraphics[height=5cm]{InclusionRadius.png}
+   \caption{Gaussians are centered at each lattice point. After a radius 
+   of a few Gaussian widths about each lattice point, the Gaussians go to 
+   zero and can be neglected.}
+   \label{fig:InclusionRadius}
+\end{figure}
+
 The crystal excess Helmholtz free energy is computed  
@@
 centered at lattice sites specified by position vectors $\vec R$. 
-The weight functions $w_\alpha(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
+The weight functions $w_i(|\vec r - \vec r~'|)$ are centered at $\vec r'$,
 as shown for $w_2$ in Figure~\ref{fig:GaussandW2_actual}a. 
@@
 Keeping the weight function centered at $\vec r$, 
-Monte-Carlo integration can be used to compute the weighted densities $n_\alpha(\vec r)$
+Monte-Carlo integration can be used to compute the weighted densities $n_i(\vec r)$
 by randomly generating vectors
 $\vec r~' -\vec R$ according to a Gaussian distribution. 
-%
-%ERRORS IN HERE... 
-%Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
-%about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
-%will be taken into account as Gaussians 
-%farther away will not contribute much.
-%The equation for the contribution to the weighted density from a 
-%Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
-%\begin{equation}{n_\alpha(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
-%and is implemented computationally in the form
-%\begin{equation}{n_\alpha(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_\alpha(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
-%which can be written as
-%\begin{equation}{n_\alpha(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_\alpha(|\vec{r}-\vec{r}~'|)}\end{equation} 
-%NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
-%generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
-%is equal to a probability density which when multiplied by dV gives 
-%the probability $P_{r'}$ at which the value of $w_\alpha(|\vec{r}-\vec{r}~'|)$ 
-%will occur. Using the relation
-%\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
-%$n_\alpha(\vec{r})$ can be expressed as the average value of $(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)$:
-%\begin{equation}{n_\alpha(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_\alpha(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
-%This process is repeated to get the contributions to the weighted density 
-%$n_\alpha(\vec r)$ from the Gaussians around each lattice point which begin to 
-%merge together as the width of the Gaussians increases.
-%All the contributions are then added together 
-%to get the final value of the weighted density at some point $\vec r$.
-
-%The result can be made more accurate by increasing NumPoints (the 
-%number of vectors randomly generated), and by using antithetic variates.
-%To use antithetic variates, each randomly generated  vector is paired 
-%with a vector of equal magnitude, but opposite in direction. 
-%First order error is reduced by averaging the two results that depend on 
-%the vectors, one of which may be erroneously higher than it should be 
-%while the other is erroneously lower than it should be.
-
-%The standard error of the mean (SEM) is given in general by
-%\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
-%To compute the error, the standard deviation of weight function $w_3$ is used 
-%\begin{equation}SD=\sqrt{\overline{w_3^2}-(\overline{w}_3)^2}\end{equation} 
-%\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{\text{NumPoints}}}}\end{equation} 
-%NumPoints, the number of vectors randomly generated, is increased until
-%\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
-%where the maximum value that the weighted density $n_3$ 
-%can properly have is 1.
-%end ERRORONEOUS SECTION
+Only lengths for $\vec r~'- \vec R$ that are within an inclusion radius
+about a lattice site, like that shown in Figure \ref{fig:InclusionRadius}, 
+will be taken into account as Gaussians 
+farther away will not contribute much.
+The equation for the contribution to the weighted density from a 
+Gaussian centered at the jth lattice site located at $\vec R_j$ is given by
+\begin{equation}{n_i(\vec r)= \int_{r'}{(1-f_v)n(\vec r~' - \vec R_j)w_i(|\vec{r}-\vec{r}~'|)} {d}\vec{r}~'}\end{equation} 
+and is implemented computationally in the form
+\begin{equation}{n_i(\vec r)= \sum_{r'}{(1-f_v)n(\vec r~' -\vec R_j)w_i(|\vec{r}-\vec{r}~'|)dV}}\end{equation} 
+which can be written as
+\begin{equation}{n_i(\vec r)= \sum_{NumPoints}(1-f_v)P_{r'}w_i(|\vec{r}-\vec{r}~'|)}\end{equation} 
+NumPoints is the number of vectors $\vec r~' - \vec R_j$ randomly 
+generated about the jth lattice site, and $n(\vec r~' - \vec R_j)$
+is equal to a probability density which when multiplied by dV gives 
+the probability $P_{r'}$ at which the value of $w_i(|\vec{r}-\vec{r}~'|)$ 
+will occur. Using the relation
+\begin{equation}{f_{average}=\sum_i^N{P_if_i}=\sum_i{\frac{f_i}{N}}}\end{equation} 
+$n_i(\vec{r})$ can be expressed as the average value of $(1-f_v)w_i(|\vec{r}-\vec{r}~'|)$:
+\begin{equation}{n_i(\vec{r})=\sum_{\text{NumPoints}}\frac{(1-f_v)w_i(|\vec{r}-\vec{r}~'|)}{\text{NumPoints}}}\end{equation}
+This process is repeated to get the contributions to the weighted density 
+$n_i(\vec r)$ from the Gaussians around each lattice point which begin to 
+merge together as the width of the Gaussians increases.
+All the contributions are then added together 
+to get the final value of the weighted density at some point $\vec r$.
+
+The result can be made more accurate by increasing NumPoints (the 
+number of vectors randomly generated), and by using antithetic variates.
+To use antithetic variates, each randomly generated  vector is paired 
+with a vector of equal magnitude, but opposite in direction. 
+First order error is reduced by averaging the two results that depend on 
+the vectors, one of which may be erroneously higher than it should be 
+while the other is erroneously lower than it should be.
+
+The standard error of the mean (SEM) is given in general by
+\begin{equation}{SD_{mean}=\frac{SD}{\sqrt{N}}}\end{equation} 
+To compute the error, the standard deviation of weight function $w_3$ is used 
+\begin{equation}SD=\sqrt{\overline{w_3^2}-(\overline{w}_3)^2}\end{equation} 
+\begin{equation}{Error=\sqrt{\frac{\overline{w_3^2}-(\overline{w}_3)^2}{\text{NumPoints}}}}\end{equation} 
+NumPoints, the number of vectors randomly generated, is increased until
+\begin{equation}{Error<\frac{|1-n_3|}{4}}\end{equation}
+where the maximum value that the weighted density $n_3$ 
+can properly have is 1.
 
@@
 Phase diagrams for the WCA fluid are shown in Figures~\ref{fig:Phase_Diagram_of_T_vs_n} 
-and \ref{fig:Phase_Diagram_P_vs_T}. 
+and \ref{fig:Phase_Diagram_P_vs_T}. In all cases $\sigma=1$ and $\epsilon=1$. 
 In Figure~\ref{fig:Phase_Diagram_of_T_vs_n} the reduced temperature 
-$T^*\equiv\frac{kT}{\epsilon}$ is plotted against the reduced number density 
-$n^*\equiv n\sigma^3$ for temperatures of $T^*$ = 0.5 to 3.0. 
-%In all cases $\sigma=1$ and $\epsilon=1$. 
-It is seen that in general, a fluid exists at lower number densities, 
+$T^*=\frac{kT}{\epsilon}$ is plotted against the reduced number density 
+$n^*=n\sigma^3$ for temperatures of kT = 0.5 to 3.0. 
+It is seen that in general, a liquid exists at lower number densities, 
 and a solid exists at higher number densities. 
-For a given temperature, it is seen that as the density increases, 
-a maximum number density is reached for the fluid. 
+For a given temperature, it is seen that as the density of the 
+liquid increases, a maximum number density is reached for the liquid. 
 Looking at the coexistence region as a whole for all temperatures, 
@@
 and shifted a bit toward smaller densities. For example, at a temperature 
-of $T^*$=1, the cDFT fluid transition reduced number density is 0.96, 
+of kT=1, the cDFT liquid transition reduced number density is 0.96, 
 and the solid transition reduced number density is 1.02, giving a spread 
-of 0.06, whereas the Monte-Carlo results give a fluid transition 
+of 0.06, whereas the Monte-Carlo results give a liquid transition 
 reduced number density of 0.92, and a solid transition reduced number 
@@
   \caption{Phase Diagram of the reduced temperature verses the reduced 
-  density for the WCA fluid. The red region is fluid, and the blue region is 
-  solid. The gray area that separates the fluid and solid regions 
-  is the region of coexistence where some fluid and solid exist together 
-  in varying proportions. White lines show the boundaries of the coexistence 
-  region generated from Monte-Carlo simulations.}
+  density for the WCA fluid. The red region is liquid, and the blue region is 
+  solid. The gray area that separates the liquid and solid regions 
+  is the region of coexistence where some liquid and solid exist together 
+  at varying proportions. White lines show the boundaries of the coexistence 
+  region generated from Monte-Carlo simulations of the WCA fluid.}
   \label{fig:Phase_Diagram_of_T_vs_n}
@@
 %The solid phase transition number densities predicted by each method 
-%are very nearly the same along all temperatures, but the fluid phase 
+%are very nearly the same along all temperatures, but the liquid phase 
 %transition number densities predicted by cDFT are notably higher than 
@@
   %%\caption{Phase Diagram of the reduced temperature verses the reduced density for the WCA fluid. 
-  %%White lines show the fluid and solid transition densities at each temperature from 
+  %%White lines show the liquid and solid transition densities at each temperature from 
   %%simulations of a WCA fluid by Ahmed and Sadus~\cite{ahmedsadus}. The gray region is the region of coexistence.}
@@
 Figure~\ref{fig:Phase_Diagram_P_vs_T} shows the reduced pressure 
-$p^*\equiv\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
-for temperatures of $T^*$ = 0.5 to 3.0. 
-The black line marks the phase boundary between the solid and fluid regions. 
+$p^*=\frac{P\sigma^3}{\epsilon}$ plotted against the reduced temperature 
+for temperatures of kT = 0.5 to 3.0. 
+The black line marks the phase boundary between the solid and liquid regions. 
 It is seen that the higher the temperature, the higher the pressure must be
-to undergo a phase transition from a fluid to a solid, as would be expected. 
-Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) 
-is given by
-\begin{align} 
-    \frac{dP}{dT}=\frac{S_f-S_s}{V_f-V_s}
-\end{align}
-Since the entropy and the volume of the fluid is greater than the entropy
+to undergo a phase transition from a liquid to a solid, as would be expected. 
+Ideally, the slope of the phase boundary $\frac{dP}{dT}$ at some point (P,T) is given by
+\begin{align} 
+    \frac{dP}{dT}=\frac{S_l-S_s}{V_l-V_s}
+\end{align}
+Since the entropy and the volume of the liquid is greater than the entropy
 and the volume of the solid, the slope should be positive.
@@
 from Monte-Carlo simulations shown by the white line, it is seen that
-the phase transition pressures and temperatures predicted by cDFT 
-are consistently higher, and the 
+the cDFT transition temperatures are consistently higher, and the 
 discrepancy grows as the temperature increases. For example, at a 
-temperature of $T^*$=1, the phase transition reduced pressure predicted by 
-cDFT is 15.24, whereas the Monte-Carlo simulation predicted a reduced 
-pressure of 9.99,
-and at a temperature of $T^*$=3, cDFT gives a value of 77.27 whereas the 
+temperature of kT=1, the phase transition reduced pressure predicted by 
+cDFT is 15.24, whereas the Monte-Carlo simulation predicted a value of 9.99,
+and at a temperature of kT=3, cDFT gives a value of 77.27 whereas the 
 Monte-Carlo simulation predicts 43.80.
-This may indicate that either the difference between the fluid 
+This may indicate that either the difference between the liquid 
 and solid phase transition entropies predicted by cDFT is too high, or 
@@
   \caption{Phase Diagram of the reduced pressure verses the reduced 
-  temperature. The blue region is solid, the red region is fluid.
-  The black line shows the phase boundary where the solid and fluid
-  phases coexist. The white line marks the solid-fluid phase boundary  
+  temperature. The blue region is solid, the red region is liquid,
+  and the black line shows the phase boundary where the solid and liquid
+  phases coexist. The white line marks the solid-liquid phase boundary  
   as predicted by Monte-Carlo simulations.}
@@
 setting the second virial coefficients for the error function potential 
-and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$ 
-at each temperature.
+and the WCA potential equal to each other, $B_{2erf}(T)=B_{2wca}(T)$.
 To implement cDFT we used Gaussian distribution functions centered at 
@@
 sites until the Helmholtz free energy was minimized. Once the equilibrium 
-Helmholtz free energy for the fluid and solid states were found for a given
+Helmholtz free energy for the liquid and solid states were found for a given
 temperature and number density, 
 we used this data 
-to find the number densities of the fluid and the solid at the phase transition.
+to find the number densities of the liquid and the solid at the phase transition.
 Using this information, we generated a temperature-density 
@@
 
-\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_0(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_0(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_0(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 Calculate $\widetilde{w}_0(k)$ 
-\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_0(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r^2}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_0(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{w}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_1(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_1(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_1(k)$ 
-\begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_1(k)=\frac{1}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}\left(\frac{1}{r}\right)e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right]d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_1(k)=\frac{2\pi}{4{\pi}\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{w}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_2(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_2(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_2(k)$ 
-\begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_2(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_2(k)=\frac{2\pi}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 \begin{equation}{\widetilde{w}_2(k)=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\frac{2\sin(kr)}{kr}d{r}}\end{equation}
@@
 
-\begin{equation}{\widetilde{w}_3(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_3(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}$ 
-\begin{equation}{\widetilde{w}_3(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{w}_3(k)=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_3(r)e^{-ikr\cos\theta}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 
 \noindent Calculate $\widetilde{w}_3(k)$ 
-\begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
-
-\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_3(k)=\frac{1}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\left(\int_{0}^{2\pi}d{\phi}\right)\right)d{r}}\end{equation}
+
+\begin{equation}{\widetilde{w}_3(k)=\frac{2\pi}{2}\int_{0}^{\infty}\left[1-\operatorname{erf}\left(\frac{r-\frac{\alpha}{2}}{a}\right)\right]r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta}d{\cos\theta}\right)d{r}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}\vec{w}_1(\vec{r})e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}\vec{w}_1(\vec{r})e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated 
 coordinate system $(x',y'z')$ 
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where
@@
 about $z'$. That leaves
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-%\begin{align}
-    %\widetilde{\vec{w}}_1(\vec{k}) &= \frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'  \\
-     %&= \frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'   \\
-     %&= \frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'   \\
-     %&= \frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~~}\hat{z}' \\
-     %&= \frac{i}{k\sqrt{\pi}a}\left[-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right]{~~}\hat{z}'
-%\end{align}
-
-%\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~}\hat{z}'}\end{equation}
-
-%KEEP%\begin{equation}\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left(-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right){~}\hat{z}'\end{equation}
-\begin{equation}
-   \widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left(-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}
-   \cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}
-   \left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)\color{white}
-   \right]\color{white}\right)\color{black}  \nonumber
-\end{equation} 
-\begin{equation}
-   +\color{white}\left(\left[\color{black}\operatorname{erf}
-   \left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\color{black}\right)
-   \color{black}{~}\hat{z}'
-\end{equation}
-
-%KEEP%\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]\vec{k}}\end{equation}
-\begin{equation}
-   \widetilde{\vec{w}}_1(\vec{k}) = \frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}
-   e^{-\left(\frac{\alpha}{2a}\right)^2}\color{white}\right]\color{black}
-   \left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)
-   \color{white}\right]\color{black}  \nonumber
-\end{equation} 
-\begin{equation}  
-   \color{white}\left[\color{black} +\operatorname{erf}\left(\frac{ka}{2}
-   -\frac{i\alpha}{2a}\right)\right]\color{white}\left[\color{black}
-   -e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]\color{black}{~}\vec k 
-\end{equation}
-
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_1(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{1}{4\pi\sqrt{\pi}a}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos{\theta}'e^{-ikr\cos\theta'}d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{2}{4\sqrt{\pi}a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\left(-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\cos(kr)d{r}+\frac{1}{k}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\frac{\sin(kr)}{r}d{r}\right]{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k\sqrt{\pi}a}\left[-a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)+\frac{\pi}{2k}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]\right]{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{2ka}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+\frac{i\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-\frac{i\alpha}{2a}\right)\right]-e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
 Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
 $\widetilde{\vec{w}}_1(k)$ becomes
-%KEEP%\begin{equation}\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}\end{equation}
-\begin{equation}
-   \widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}
-   e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}
-   \left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)
-   +\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}
-   \right)\right]\color{white}\right]\color{black}  \nonumber
-\end{equation}
-\begin{equation}
-   \color{white}\left[\color{black}-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\color{black}\vec{k}
-\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_1(\vec{k})=\frac{i}{k^2}\left[\frac{\sqrt{\pi}}{\sqrt{2}k\Xi}e^{-\left(\frac{\alpha}{\sqrt{2}\Xi}\right)^2}\left[\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}+\frac{i\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{(\sqrt{2})^3}-\frac{i\alpha}{\sqrt{2}\Xi}\right)\right]-e^{-\frac{k^2\Xi^2}{8}}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
 
@@
 
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r){~}\frac{\vec{r}}{r}{~}e^{-i\vec{k}\cdot{\vec{r}}}r^2d{r}d{\cos\theta}d{\phi}}\end{equation}
 Set $\vec{k}$ in the direction of $\hat{z}'$ of a rotated coordinate 
 system $(x',y'z')$ 
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\frac{(r_{x'}\hat{x}'+r_{y'}\hat{y}'+r_{z'}\hat{z}')}{r}e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where 
@@
 about $z'$. That leaves
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{w}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(-\frac{2i}{kr}\cos(kr) + \frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4i\sqrt{\pi}}{ak}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r} + \frac{4i\sqrt{\pi}}{ak^2}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4\sqrt{\pi}i}{ak}\left[\frac{a\alpha\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]} \nonumber \end{equation}
-\begin{equation}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~}\hat{z}'}\end{equation}
-
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\vec{k}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}w_2(r)\cos{\theta}'e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(\int_{-1}^{1}e^{-ikr\cos\theta'}\cos{\theta}'d{\cos\theta'}\right)d{r}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{w}_2(\vec{k})=\frac{2\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left(-\frac{2i}{kr}\cos(kr) + \frac{2i}{k^2r^2}\sin(kr)\right)d{r}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4i\sqrt{\pi}}{ak}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\cos(kr)d{r} + \frac{4i\sqrt{\pi}}{ak^2}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\sin(kr)d{r}{~~}\hat{z}'}\end{equation}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{-4\sqrt{\pi}i}{ak}\left[\frac{a\alpha\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\cos\left(\frac{k\alpha}{2}\right)-\frac{a^3k\sqrt{\pi}}{2}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]}\end{equation}
+\begin{displaymath}{+{~~}\frac{4\sqrt{\pi}i}{ak^2}\left[a\sqrt{\pi}e^{-\frac{k^2a^2}{4}}\sin\left(\frac{k\alpha}{2}\right)\right]{~~}\hat{z}'}\end{displaymath}
+
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2a^2}{4}}\left[\left(\frac{a^2k}{2}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
 Putting $a=\frac{\Xi}{\sqrt{2}}$ back in, the final equation for 
 $\widetilde{\vec{w}}_2(k)$ becomes
-\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~}\vec{k}}\end{equation}
+\begin{equation}{\widetilde{\vec{w}}_2(\vec{k})=\frac{4\pi{i}}{k^2}e^{-\frac{k^2\Xi^2}{8}}\left[\left(\frac{\Xi^2k}{4}+\frac{1}{k}\right)\sin\left(\frac{k\alpha}{2}\right)-\frac{\alpha}{2}\cos\left(\frac{k\alpha}{2}\right)\right]{~~}\vec{k}}\end{equation}
 
@@
 The Fourier transform of the tensor weight matrix elements are given by
-\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
+\begin{equation}{\widetilde{w}_{m2_{ij}}(\vec{k})=\int_{allspace}{w_{{m2}_{ij}}}(\vec{r})e^{-i\vec{k}\cdot\vec{r}}d{\vec{r}}}\end{equation}
 Since the integration is over all space, the integral can be simplified 
@@
 transformed weight function in coordinate system (x,y,z). 
-\begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{m2_{ij}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_ir_j}{r^2}-\frac{\delta_{ij}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
 where $r=r'$ and
@@
 Calculate $\widetilde{w}_{{m2}_{x'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{y'}}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{y'})}{r^2}-\frac{\delta_{x'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{x'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin^2\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'\sin{\phi'}~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{z'}}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{z'})}{r^2}-\frac{\delta_{x'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{x'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\cos\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{y'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{z'}}{r^2}-\frac{\delta_{yz}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
-$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
+\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{z'})}{r^2}-\frac{\delta_{yz}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{y'z'}}(\vec{k})=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2e^{-ikr\cos\theta'}\sin\theta'\cos\theta'\underbrace{\left(\int_{0}^{2\pi}\sin\phi'~d{\phi'}\right)}d{r}d{\cos\theta'}=0}\end{equation}
+$~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~0$
 
 Calculate $\widetilde{w}_{{m2}_{x'x'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{x'}r_{x'}}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{x'})(r_{x'})}{r^2}-\frac{\delta_{x'x'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\cos^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta'\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\cos^2\phi'~d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
 
 %\color{blue}
-\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}\label{compare_equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
 %\color{black}
@@
 
-Set the lower limit to $-\infty$ as the integrand is nearly zero by the time r is zero. 
+Set the lower limit to -infinity since the integrand is nearly zero by the time r is zero. 
 \begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=-\frac{4\sqrt{\pi}}{3ka}\left(\int_{-\infty}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r\sin(kr)d{r}\right)}\end{displaymath} 
@@
 \begin{equation}{+\frac{2\pi\sqrt{\pi}}{k^3a}e^{-\left(\frac{\alpha}{2a}\right)^2}\left[\operatorname{erf}\left(\frac{ka}{2}+i\frac{\alpha}{2a}\right)+\operatorname{erf}\left(\frac{ka}{2}-i\frac{\alpha}{2a}\right)\right]}\end{equation}
-
-Putting in $a=\frac{\Xi}{\sqrt{2}}$, the final equation for $\widetilde{w}_{{m2}_{xx}}(k)$ becomes
-
-\begin{equation}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\nonumber\end{equation}
-\begin{equation}{+\frac{{(2\pi)}^{\frac{3}{2}}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2^\frac{3}{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2^\frac{3}{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
+Putting in $a=\frac{\Xi}{\sqrt{2}}$ the final equation for 
+$\widetilde{w}_{{m2}_{xx}}(k)$ becomes
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{x'x'}}(k)=\left(\frac{-\pi{\Xi}^2}{3}-\frac{4\pi}{k^2}\right)e^{-\frac{k^2\Xi^2}{8}}\cos(\frac{k\alpha}{2})-\frac{2\pi\alpha}{3k}e^{-\frac{k^2\Xi^2}{8}}\sin(\frac{k\alpha}{2})}\end{displaymath}
+\begin{equation}{+\frac{2\pi\sqrt{2\pi}}{k^3\Xi}e^{-\left(\frac{\alpha^2}{2\Xi^2}\right)}\left[\operatorname{erf}\left(\frac{k\Xi}{2*\sqrt{2}}+i\frac{\alpha}{\sqrt{2}\Xi}\right)+\operatorname{erf}\left(\frac{k\Xi}{2\sqrt{2}}-i\frac{\alpha}{\sqrt{2}\Xi}\right)\right]}\end{equation}
 
 Calculate $\widetilde{w}_{{m2}_{y'y'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{y'y'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\sin^2\theta'\sin^2\phi'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\left(1-\cos^2\theta\right)e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}\sin^2\phi~d{\phi}\right)d{\cos\theta}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{y'y'}}(k)=-\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta~e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{+\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}~d{\cos\theta'}\right]d{r}}\end{equation}
 
@@
 Calculate $\widetilde{w}_{{m2}_{z'z'}}(k)$ 
-\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{r_{y'}r_{y'}}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
-\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
-
-\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}
-
-\begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\frac{(r_{y'})(r_{y'})}{r^2}-\frac{\delta_{z'z'}}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+\begin{equation}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}\int_{-1}^{1}\int_{0}^{2\pi}e^{\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}\left(\cos^2\theta'-\frac{1}{3}\right)e^{-ikr\cos\theta'}r^2d{r}d{\cos\theta'}d{\phi'}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=\frac{1}{a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{displaymath}
+\begin{equation}{-\frac{1}{3a\sqrt{\pi}}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}\left(\int_{0}^{2\pi}d{\phi'}\right)d{\cos\theta'}\right]d{r}}\end{equation}
+
+\begin{displaymath}{\widetilde{w}_{{m2}_{z'z'}}(k)=2\frac{\sqrt{\pi}}{a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}\cos^2\theta'~e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{displaymath}
+
+\begin{equation}{-2\frac{\sqrt{\pi}}{3a}\int_{0}^{\infty}e^{-\left(\frac{r-\frac{\alpha}{2}}{a}\right)^2}r^2\left[\int_{-1}^{1}e^{-ikr\cos\theta'}d{\cos\theta'}\right]d{r}}\end{equation}
 
@@
 \section{Integrals used to Compute Weight Function Fourier Transforms}
-\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
-\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation}
-
-\begin{equation}{\int_{-1}^{1}{e^{ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}-\frac{4\sin(kr)}{k^3r^3}}\end{equation}
+\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}}\end{equation} 
+\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos{\theta}{~}d{\cos{\theta}}}=-\frac{2i}{kr}\cos(kr)+\frac{2i}{k^2r^2}\sin(kr)}\end{equation}
+
+\begin{equation}{\int_{-1}^{1}{e^{-ikr\cos{\theta}}\cos^2(\theta)~d{\cos{\theta}}}=\frac{2\sin(kr)}{kr}+\frac{4\cos(kr)}{k^2r^2}+\frac{4\sin(kr)}{k^3r^3}}\end{equation}
 
@@
 
-\begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2\color{white}\right]\color{black}}\end{displaymath}
-\begin{equation}{\color{white}\left[\color{black} -\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)   +n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
+\begin{displaymath}{\Phi_3=\frac{12\pi^2R^6}{(1-\eta)^2}\left(\frac{1}{4\pi{R}^2}\right)^3\left[\vec{n}_{2v}\cdot\left({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}}\right)\cdot\vec{n}_{2v}-n_2|\vec{n}_{2v}|^2-\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^3\right)\color{white}\right]\color{black}}\end{displaymath}
+\begin{equation}{\color{white}\left[\color{black}+n_2\operatorname{Tr}\left(({\overleftrightarrow{n}_{m2}}+\frac{n_2}{3}\hat{\operatorname{I}})^2\right)\right]}\end{equation}
 
@@
 
-%KEEP%\begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
-\begin{equation}
-     \Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)  \color{white}\right)\color{black}\nonumber
-\end{equation}
-\begin{equation}
-      +\color{white}\left(\color{black}\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)\color{black} 
-\end{equation}
+\begin{equation}{\Phi_3=\frac{3}{16\pi(1-\eta)^2}\left(\frac{2}{9}{n_2}^3-\frac{2}{3}n_2|\vec{n}_{v2}|^2+\vec{n}_{v2}\cdot{\overleftrightarrow{n}_{m2}}\cdot{\vec{n}_{v2}}-\operatorname{Tr}({\overleftrightarrow{n}_{m2}}^3)+\frac{1}{3}n^2_2\operatorname{Tr}(\overleftrightarrow{n}_{m2})\right)}\end{equation}
 
@@
 \end{align}
-$\theta(|\vec r|-2R)$ is a backwards Heaviside step function 
+$\theta(|\vec r|-2R)$ is a backwards Heavyside step function 
 given by 
@@
     F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
-        &= F.T.\left[w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}\right]  \nonumber \\                                                  
-        &= F.T.\left[I.F.T\left(\widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1\right)\right]  \nonumber \\
+        &= F.T.\left[w_3 \otimes w_0 + w_2 \otimes w_1 + \vec{w_2} \otimes \vec{w_1}\right]  \\                                                  
+        &= F.T.\left[I.F.T\left(\widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1\right)\right]  \\
         &= \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
@@
     F.T.\left(\frac{\Theta(|\vec{r}|-2R)}{2}\right) 
-       &=\frac{1}{2}\int_0^\pi \int_{-1}^1\int_0^{2R}(1)e^{-ikr\cos \theta'}r^2~dr~d\cos \theta'~d\phi'  \nonumber  \\
-       &=\frac{2\pi}{2}\int_0^{2R} r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}~d\cos \theta'~d\phi' \nonumber   \\
-       &=\pi\int_0^{2R} r^2\left(\frac{2\sin (kr)}{kr}\right)~dr \nonumber  \\
-       &=\frac{2\pi}{k}\int_0^{2R} r\sin (kr)~dr \nonumber  \\
-       &=\frac{2\pi}{k}\left[\frac{\sin (kr)-kr\cos (kr))}{k^2}\right] \bigg|_0^{2R} \nonumber  \\
-       &= \frac{2\pi}{k^3}\sin (2kR)-\frac{4\pi R}{k^2}\cos (2kR) \nonumber  \\
-       &= \frac{4\pi}{k^3}\sin (kR)\cos (kR) - \frac{4 \pi R}{k^2}\cos^2 (kR) + \frac{4\pi R}{k^2} \sin^2 (kR)
+       &=\frac{1}{2}\int_0^\pi \int_{-1}^1\int_0^{2R}(1)e^{-ikr\cos \theta'}r^2~dr~d\cos \theta'~d\phi'   \\
+       &=\frac{2\pi}{2}\int_0^{2R} r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}~d\cos \theta'~d\phi'   \\
+       &=\pi\int_0^{2R} r^2\left(\frac{2\sin (kr)}{kr}\right)~dr  \\
+       &=\frac{2\pi}{k}\int_0^{2R} r\sin (kr)~dr  \\
+       &=\frac{2\pi}{k}\left[\frac{\sin (kr)-kr\cos (kr))}{k^2}\right] \bigg|_0^{2R} \\
+       &= \frac{2\pi}{k^3}\sin (2kR)-\frac{4\pi R}{k^2}\cos (2kR)  \\
+      \label{leftside-solved} &= \frac{4\pi}{k^3}\sin (kR)\cos (kR) - \frac{4 \pi R}{k^2}\cos^2 (kR) + \frac{4\pi R}{k^2} \sin^2 (kR)
 \end{align}
@@
 \begin{align}
-    \widetilde{w}_2 &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'} r^2~dr~d\cos \theta'~d\phi  \nonumber  \\
-                    &= 2\pi\int\delta(r-R)r^2 \frac{2\sin (kr)}{kr}~dr \nonumber  \\
-                    &= 2\pi R^2 \frac{2\sin (kR)}{kR} \nonumber  \\
+    \widetilde{w}_2 &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'} r^2~dr~d\cos \theta'~d\phi   \\
+                    &= 2\pi\int\delta(r-R)r^2 \frac{2\sin (kr)}{kr}~dr  \\
+                    &= 2\pi R^2 \frac{2\sin (kR)}{kR}  \\
                     &= \frac{4\pi R}{k} \sin (kR)
@@
 \begin{align}
-    \widetilde{w}_3 &= \int\int\int\theta(r-R)e^{-ikr\cos \theta'}r^2~dr~d\cos\theta'~d\phi  \nonumber \\
-                    &= 2\pi\int\theta(r-R)\left(\frac{2\sin (kr)}{kr}\right)r^2~dr  \nonumber  \\
-                    &= \frac{4\pi}{k}\int_0^R (1)r\sin (kr)~dr  \nonumber  \\
-                    &= \frac{4\pi}{k}\left[\frac{\sin (kr) - kr\cos (kr)}{k^2}\right]\bigg|_0^R  \nonumber  \\
+    \widetilde{w}_3 &= \int\int\int\theta(r-R)e^{-ikr\cos \theta'}r^2~dr~d\cos\theta'~d\phi  \\
+                    &= 2\pi\int\theta(r-R)\left(\frac{2\sin (kr)}{kr}\right)r^2~dr   \\
+                    &= \frac{4\pi}{k}\int_0^R (1)r\sin (kr)~dr   \\
+                    &= \frac{4\pi}{k}\left[\frac{\sin (kr) - kr\cos (kr)}{k^2}\right]\bigg|_0^R   \\
                     &= \frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]
@@
 \begin{align}
-    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r^2~\frac{\vec r}{r}~dr~d\cos\theta'~d\phi \nonumber  \\
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r^2~\frac{\vec r}{r}~dr~d\cos\theta'~d\phi  \\
                           &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r_{x'} \hat x' + r_{y'} \hat y'+ r_{z'} \hat z')~dr~d\cos\theta'~d\phi  
@@
 \begin{align}
-    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r\cos \theta')~dr~d\cos\theta'~d\phi~\hat k  \nonumber  \\
-                          &= 2\pi\int\delta(r-R)r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}\cos \theta'~d\cos\theta'~\hat k  \nonumber  \\
-                          &= 2\pi\int\delta(r-R)r^2 \left(\frac{-2i}{kr}\cos (kr)+ \frac{2i}{k^2r^2}\sin (kr)\right)dr~\hat k  \nonumber  \\
-                          &= 2\pi\left[\frac{-2iR}{k}\cos (kR)+ \frac{2i}{k^2}\sin (kR)\right]~\hat k \nonumber   \\
+    \widetilde{\vec w}_2  &= \int\int\int\delta(r-R)e^{-ikr\cos \theta'}r~(r\cos \theta')~dr~d\cos\theta'~d\phi~\hat k   \\
+                          &= 2\pi\int\delta(r-R)r^2~dr\int_{-1}^1 e^{-ikr\cos \theta'}\cos \theta'~d\cos\theta'~\hat k   \\
+                          &= 2\pi\int\delta(r-R)r^2 \left(\frac{-2i}{kr}\cos (kr)+ \frac{2i}{k^2r^2}\sin (kr)\right)dr~\hat k   \\
+                          &= 2\pi\left[\frac{-2iR}{k}\cos (kR)+ \frac{2i}{k^2}\sin (kR)\right]~\hat k   \\
                           &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k 
@@
 \begin{align}
-    \widetilde{\vec w}_1  &= \frac{\widetilde{\vec w}_2}{4\pi R}  \nonumber  \\
+    \widetilde{\vec w}_1  &= \frac{\widetilde{\vec w}_2}{4\pi R}   \\
                           &= \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right)\hat k
@@
 \begin{align}
-    \widetilde{w}_0\widetilde{w}_3 &= \left(\frac{\sin (kR)}{kR}\right)\left(\frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]\right) \nonumber  \\
+    \widetilde{w}_0\widetilde{w}_3 &= \left(\frac{\sin (kR)}{kR}\right)\left(\frac{4\pi}{k^3}\left[\sin (kR) - kR\cos (kR)\right]\right)  \\
                                    &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR)
@@
 \begin{align}
-    \widetilde{w}_1\widetilde{w}_2 &= \left(\frac{\sin (kR)}{k}\right)\left(\frac{4\pi R}{k} \sin (kR)\right) \nonumber  \\                                
+    \widetilde{w}_1\widetilde{w}_2 &= \left(\frac{\sin (kR)}{k}\right)\left(\frac{4\pi R}{k} \sin (kR)\right)  \\                                
                                    &= \frac{4\pi R}{k^2}\sin^2 (kR)
@@
    \widetilde{\vec w}_1\cdot\widetilde{\vec w}_2   
-    &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k \cdot \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right) \hat k  \nonumber  \\
+    &= \left(\frac{-4\pi iR}{k}\cos (kR)+ \frac{4\pi i}{k^2}\sin (kR)\right)\hat k \cdot \left(\frac{-i}{k}\cos (kR)+ \frac{i}{Rk^2}\sin (kR)\right) \hat k   \\
     &= -\frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) - \frac{4\pi}{Rk^4}\sin^2(kR)
@@
     \widetilde{w}_3\widetilde{w}_0 + \widetilde{w}_2\widetilde{w}_1 + \widetilde{\vec w}_2 \cdot \widetilde{\vec w}_1 
-    &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR) \nonumber  \\
-    &+ \frac{4\pi R}{k^2}\sin^2 (kR) \nonumber  \\
-    &- \frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) -  \frac{4\pi}{Rk^4}\sin^2(kR) \nonumber  \\
-    &= \left(\frac{4\pi}{k^4R}+\frac{4\pi R}{k^2}-\frac{4\pi}{k^4R}\right)\sin^2(kR) \nonumber   \\
-    &-\frac{4\pi R}{k^2}\cos^2(kR) +\left(-\frac{4\pi}{k^3} + \frac{8\pi}{k^3}\right)\sin(kR)\cos(kR) \nonumber   \\
+    &= \frac{4\pi}{k^4R}\sin^2 (kR) - \frac{4\pi}{k^3}\sin(kR)\cos(kR) \\
+    &+ \frac{4\pi R}{k^2}\sin^2 (kR) \\
+    &- \frac{4\pi R}{k^2}\cos^2(kR) + \frac{8\pi}{k^3}\sin(kR)\cos(kR) -  \frac{4\pi}{Rk^4}\sin^2(kR)  \\
+    &= \left(\frac{4\pi}{k^4R}+\frac{4\pi R}{k^2}-\frac{4\pi}{k^4R}\right)\sin^2(kR)\\
+    &-\frac{4\pi R}{k^2}\cos^2(kR) +\left(-\frac{4\pi}{k^3} + \frac{8\pi}{k^3}\right)\sin(kR)\cos(kR)   \\
     &=\frac{4\pi R}{k^2}\sin^2(kR)-\frac{4\pi R}{k^2}\cos^2(kR)+\frac{4\pi}{k^3}\sin(kR)\cos(kR)
 \end{align}
-which equals Equation~\ref{leftside}, and so
+which equals Equation~\ref{leftside-solved}, and so
 %\begin{align}
@@
 \begin{align}
- B_2 &= -\frac{1}{2}\int_0^{\pi}\int_0^{2\pi}\int_0^\infty\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2\sin{\theta}d{\theta}d{\phi}dr \nonumber  \\
-     &= -\frac{1}{2}4\pi\frac{1}{2}\int_{0}^{r_{max}}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2dr \nonumber  \\
-     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \pi\int_0^{r_{max}}r^2dr  \nonumber  \\
+ B_2 &= -\frac{1}{2}\int_0^{\pi}\int_0^{2\pi}\int_0^\infty\frac{1}{2}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2\sin{\theta}d{\theta}d{\phi}dr \\
+     &= -\frac{1}{2}4\pi\frac{1}{2}\int_{0}^{r_{max}}\left[\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)-1\right]r^2dr \\
+     &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \pi\int_0^{r_{max}}r^2dr   \\
      &= -\pi\int_{0}^{r_{max}}\operatorname{erf}\left(\frac{r-\alpha}{\Xi}\right)r^2dr {~~}+{~~} \frac{\pi}{3}r_{max}^3   
@@
 \begin{align}
-  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}r_{max}^3\right)\operatorname{erf}\left(\frac{\alpha-r_{max}}{\Xi}\right) \nonumber \\
-      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \nonumber \\
-      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha r_{max}-\Xi^3-\Xi r_{max}^2\right)e^{-{\left(\frac{\alpha-r_{max}}{\Xi}\right)^2}} \nonumber \\
+  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}r_{max}^3\right)\operatorname{erf}\left(\frac{\alpha-r_{max}}{\Xi}\right) \\
+      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
+      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha r_{max}-\Xi^3-\Xi r_{max}^2\right)e^{-{\left(\frac{\alpha-r_{max}}{\Xi}\right)^2}} \\
       &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}r_{max}^3 
@@
 \begin{align}
-  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}\left(\infty\right)^3\right)\left(-1\right) \nonumber \\
-      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \nonumber \\
-      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha\left(\infty\right)-\Xi^3-\Xi r_{max}^2\right)\left(0\right) \nonumber \\
+  B_2 &= \left(-\frac{\pi}{3}\alpha^3-\frac{\pi}{2}\alpha\Xi^2+\frac{\pi}{3}\left(\infty\right)^3\right)\left(-1\right) \\
+      &+ \left(\frac{\pi}{3}\alpha^3+\frac{\pi}{2}\alpha\Xi^2\right)\operatorname{erf}\left(\frac{\alpha}{\Xi}\right) \\
+      &+ \frac{\sqrt{\pi}}{3}\left(-\Xi\alpha^2-\Xi\alpha\left(\infty\right)-\Xi^3-\Xi r_{max}^2\right)\left(0\right) \\
       &+ \frac{\sqrt{\pi}}{3}\left(\Xi\alpha^2+\Xi^3\right)e^{-\left(\frac{\alpha}{\Xi}\right)^2}+\frac{\pi}{3}\left(\infty\right)^3 
@@
 
-%KEEP%\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
-\begin{equation}
-   \frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\color{white}\right]\color{black} \nonumber
-\end{equation}
-\begin{equation}
-   \color{white}\left[-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}\color{black}+\beta(E_{s(n)}-\mu{n})\right]\color{black}
-\end{equation}
+\begin{equation}\frac{PV}{k_BT}=-\sum_{n=0}^\infty\sum_{s(n)}P_{n,s(n)}\left[-\beta(E_{s(n)}-\mu{n})-\ln{\left(\sum_{n=0}^\infty\sum_{s(n)}e^{-\beta(E_{s(n)}-\mu{n})}\right)}+\beta(E_{s(n)}-\mu{n})\right]\end{equation}
 
@@
 The canonical equilibrium phase-space probability density is given by
-\begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{\text{partian function}}\end{equation}
-\begin{equation}{\text{partian function}}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
+\begin{equation}P_{eq}(\vec r^N, \vec p^N)=\frac{1}{N!}\frac{1}{h^{3N}}\frac{e^{-\beta H}}{partian function}\end{equation}
+\begin{equation}{partian function}=\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\end{equation}
 The Helmholtz free energy is given by 
@@
 \begin{align}
- TS &= T\left[-k_B\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(N!~h^{3N}P_{eq}\right) d\vec r^{N} d\vec p^{N}\right]  \nonumber \\
-	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N}  \nonumber \\
-	%KEEP%&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H-\ln\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right]\right) d\vec r^{N} d\vec p^{N}\right]
-    &= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)\color{white}\right]\color{black} \nonumber \\ 
-    &- \color{white}\left[\color{black} P_{eq}\ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) \color{black}d\vec r^{N} d\vec p^{N}\right] \nonumber \\
-    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}ln(Z)d\vec r^{N} d\vec p^{N}   \nonumber 
-\end{align}  
-\begin{align}   
-    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}d\vec r^{N} d\vec p^{N}\right]  \nonumber \\
+ TS &= T\left[-k_B\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(N!~h^{3N}P_{eq}\right) d\vec r^{N} d\vec p^{N}\right]   \\
+	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}\operatorname{ln}\left(\frac{e^{-\beta H}}{\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}}\right) d\vec r^{N} d\vec p^{N}   \\
+	&= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]  \\
+    &= -k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int \left[P_{eq}\left(\frac{-1}{k_BT}H\right)-ln\left(\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int e^{-\beta H} d\vec r^{N} d\vec p^{N}\right) d\vec r^{N} d\vec p^{N}\right]
+\end{align}  
+\begin{align}  
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BT\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}ln(Z)d\vec r^{N} d\vec p^{N}     \\
+    &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[\frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}d\vec r^{N} d\vec p^{N}\right]   \\
     &= \frac{1}{N!}\frac{1}{h^{3N}}\int\dot~\dot~\dot~\int P_{eq}~H~d\vec r^{N} d\vec p^{N} +k_BTln(Z)\left[1\right]   
@@
 \begin{equation}{n(r)=\frac{1}{\left(\sqrt{2\pi}\sigma\right)^3}\int\exp^{-\frac{|\vec{r}-\vec{R}|^2}{2\sigma^2}}d\vec{r}}\end{equation} 
-\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{crystalFideal_{primitive.cell} + crystalFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{fluid}=\frac{fluidFideal_{primitive.cell} + fluidFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
-\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{fluid}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{crystal}=\frac{cFideal_{primitive.cell} + cFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{liquid}=\frac{lFideal_{primitive.cell} + lFexcess_{primitive.cell}}{N_{primitive.cell}}\end{equation}
+\begin{equation}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{crystal}-{~~}\left(\frac{F}{N}\right)_{liquid}\end{equation}
 \begin{displaymath}\text{If~~~~~~}\left(\frac{F}{N}\right)_{diff}<\text{~~~~Best}\left(\frac{F}{N}\right)_{diff}\end{displaymath}
 \begin{displaymath}\text{Then~~~~~~}\text{Best}\left(\frac{F}{N}\right)_{diff}=\left(\frac{F}{N}\right)_{diff}\end{displaymath}
-\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{fluid}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
+\begin{displaymath}\text{Output:~~~~~~~~~~~~ Best}\left(\frac{F}{N}\right)_{crystal}\text{~~and~~~~~~}\left(\frac{F}{N}\right)_{liquid}\text{~~~~for the given T, n ~~~~~~~~~~~~~~~~~~~~~~~~~}\end{displaymath}
 
@@
 \end{document}
-
-%my notes
-Definition of a convolution:
-\begin{align}
-    w_\alpha \otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r')w_\gamma(\vec r-\vec r')d\vec r'
-\end{align}
-
-\begin{align}
-    w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r')w_\gamma(\vec r'-\vec r)d\vec r
-\end{align}
-
-Set $\vec r' = \vec r_i-\vec r$
-\begin{align}
-     w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-\vec r-\vec r)d\vec r
-\end{align}
-
-Put in $\color{blue}\vec r=\vec r_i-\vec r_j\color{black}$
-\begin{align}
-     w_\alpha\otimes w_\gamma(\vec r) &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-(\color{blue}\vec r_i-\vec r_j\color{black})-r)d\vec r  \\
-                                    &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_i-\vec r_i+\vec r_j-\vec r)d\vec r \\
-                                    &= \int w_\alpha(\vec r_i-\vec r)w_\gamma(\vec r_j-\vec r)d\vec r
-\end{align}

### 2021-02-02 papers/thesis-kirstie/thesis.tex

--- 2021-01-15 papers/thesis-kirstie/thesis.tex
@@
   interaction of two neutral atoms. Our cDFT functional is constructed 
-  using Soft Fundamental Theory (SFMT) with Schmidt's error function 
+  using Soft Fundamental Measure Theory (SFMT) with Schmidt's error function 
   model in which we have incorporated temperature dependent parameters 

### 2021-08-31 papers/fuzzy-fmt/figs/new-melting.cpp
looking for pressure bug

@@ -407,13 +407,14 @@ data find_energy_new(double temp, double reduced_density, double fv, double gwid
           phi_3 =( n_2*(uipow(n_2,2) - 3.0*nv_2.dot(nv_2)) + (9/2.0)*(nv_2.dot(nm_2.dot(nv_2)) - traceof_nm_2cubed))
             /(24*M_PI*uipow(1-n_3,2));         // Santos (2012) - yes
           //phi_3 = (nv_2.dot(nm_2.dot(nv_2))-n_2*nv_2.dot(nv_2) - traceof_nm_2cubed + n_2*traceof_nm_2squared)/((16/3.0)*M_PI*uipow(1.0-n_3,2));      // Sweatman, Groh & Schmidt  (2001) - same as Santos with different w_nm
-          mu_of_r = -log(1.0-n_3)+(M_PI*uipow(Xi,2)*alpha/2)*(n_0/(1.0-n_3))
-          +(n_1*M_PI*(uipow(Xi,2)+uipow(alpha,2))+n_2*(alpha/2))/(1.0-n_3)
-          +((n_1*n_2-nv_1.dot(nv_2))*(M_PI*uipow(Xi,2)*alpha/2))/(uipow((1.0-n_3),2))         
-          +(n_2*(uipow(n_2,2) - 3.0*nv_2.dot(nv_2))+(9/2.0)*(nv_2.dot(nm_2.dot(nv_2)) 
-          -traceof_nm_2cubed))*((uipow(Xi,2)*alpha)/(24*(1.0-n_3)*uipow((1.0-n_3),2)))
-          +(1/(24*uipow((1.0-n_3),2)))*(uipow(n_2,2)*M_PI*(uipow(Xi,2)
-          +uipow(alpha,2))-3*M_PI*(uipow(Xi,2)+uipow(alpha,2))*nv_2.dot(nv_2));
+          mu_of_r = -0*log(1-n_3)
+                    +0*(M_PI*uipow(Xi,2)*alpha/2)*(n_0/(1-n_3))
+            +0*(n_1*M_PI*(uipow(Xi,2)+uipow(alpha,2))+n_2*(alpha/2))/(1-n_3)
+            +0*((n_1*n_2-nv_1.dot(nv_2))*(M_PI*uipow(Xi,2)*alpha/2))/uipow((1-n_3),2)
[…]

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm stuff

\documentclass[double,12pt]{revtex4-2}
\usepackage{graphicx}
\usepackage{color}
\usepackage{amsmath}
\usepackage{subcaption}

\begin{document}
%WORK ALL SECTIONS in Parallel!

\title{RESEARCH PROPOSAL}
\maketitle
\large
\begin{center}Toward completion of the Doctor of Philosophy degree in Physics 
~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~at Oregon State University for Kirstie Finster
\end{center}

\normalsize

\section{Part 1: Overview}
My research involves creating models for use in the study of liquids. 
Although there are many approaches that can be taken, the simplest and 
most useful approach is through the use of the perturbation theory first
applied to the study of liquids by Zwanzig in 1954. 
Zwanzig demonstrated that a real liquid can be modeled by using the well 
known and highly developed hard sphere model as a reference fluid for the 
repulsive portion of the particle interaction to which 
a small attractive portion could be added as a perturbation. 
In 1964, Rowlinson introduced the idea of making the diameter of the hard sphere 
fluid temperature-dependent to better fit certain classes of soft interatomic potentials
to the behavior that would be expected as the temperature changes and particles 
come ever closer together during collisions at increasing temperatures.
%SHOW FIGURE
In 1967, Barker and Henderson made a monumental break-through 
by generalizing 
Rowlinson's use of the temperature-dependent hard sphere diameter and 
developing a method by which \textit{any} soft repulsive can be modeled 
by mapping it to a hard-sphere fluid with a temperature-dependent diameter. %\textcolor{red}{(cite paper)} 
This highly successful model, known as the 
Barker-Henderson fluid, has been in use now for around fifty years and
is the standard used to model fluids today. %\textcolor{red}{(cite papers)}.
But there is a problem.
The hard-sphere fluid model around which the Barker-Henderson fluid is 
constructed uses delta functions which makes the model difficult and 
inconvenient to apply in computer simulations. 
My research aims at constructing a model for fluids around a reference 
fluid that has a smooth function instead of the sharply discontinuous 
repulsive hard-sphere interatomic potential that is zero when the 
particles are not in contact and infinity upon contact.

In selecting a soft interatomic potential to use as a reference fluid, 
there is one choice to consider that stands out above others, and that is 
the repulsive portion of the interatomic potential of the 
Weeks-Chandler-Andersen (WCA) fluid.
The WCA potential reproduces the repulsive force of the Lennard-Jones 
interatomic potential that is used to model neutral atoms. 
The thermodynamic properties of the fluid can be obtained from the 
Helmholtz free energy which
can be expressed as a functional of the average number 
density profile $n(\vec r)$ 
which gives the probability of finding a particle at each position $\vec r$. 
The Helmholtz free energy, along with the corresponding equilibrium number density profile,
can be found using density functional theory by 
varying the number density profile until the free energy is minimized.
For classical fluids, Rosenfeld formulated a highly successful 
density functional theory called Fundamental Measure Theory (FMT)
by noting that the Mayer function of the hard-sphere fluid is a step function
that can be decomposed into a set of weight functions that can be used 
to express the Helmholtz free energy functional in terms of a set of weighted densities. 
Schmidt created Soft Fundamental Measure Theory (SFMT) for soft sphere 
fluids which is based on FMT but reformulated for soft potentials by 
meeting the limiting conditions for a 0-dimensional fluid (where a sphere 
is confined to a cavity large enough for only one sphere) 
and for low-density. 
Emerging from his work is an "error-function" potential he formed to 
demonstrate the application of SFMT to a soft-sphere fluid and which is
similar to the repulsive potential of the WCA fluid. However, 
Schmidt's error function potential can only be made similar to the WCA fluid 
at one temperature -
it will not work for a variety of temperatures. 
Therefore, we use the error function potential but make
its parameters temperature-dependent (similar to the way Barker-Henderson made the
diameter of the hard sphere fluid temperature-dependent) and derive them by 
making the second virial coefficient of our error function potential match the second virial 
coefficient of a WCA fluid. Our approach is easily extensible to other soft
repulsive potentials.

\section{Part 2: Research Done}
We have a working model of a soft-sphere reference fluid that depends 
on a smooth potential function instead of on the discontinuous hard-sphere potential 
which has been the standard in use for the last fifty years. 
In our paper prepared for publication we have: \\
$\bullet$ Demonstrated good agreement between our WCA fluid and Barker-Henderson fluid \\
$\bullet$ Demonstrated good agreement between our WCA fluid and Monte-Carlo simulations \\
$\bullet$ Demonstrated good agreement between our WCA fluid and experiments with Argon \\

\section{Part 3: Research To Do}
\noindent Toward a better model and a better understanding of our model, I propose to: \\
$\bullet$ Find the internal energy for our WCA fluid model \\
$\bullet$ Improve results at low temperature, high density \\ %pressure calculations  \\
%$\bullet$ Possibly derive temperature-dependent parameter with B3 and B2  \\
%4- another functional for better pressure in freezing with improved pressure calculation\\
%4- another function (instead of Verf) for better pressure in freezing \\
$\bullet$ Examine and demonstrate the generality and usefulness of our 
method by applying it to other soft repulsive fluid potentials (Dr. Roundy famous!) \\
%$\bullet$ Apply our model  (ie. show by example) to demonstrate usefulness \\ %and market \\
%better understand and meet the needs of those who would benefit from our approach and market \\

\section{Part 4: Means/timeline/program}
\noindent Requirements of the PhD program \\
$\bullet$ One class left: PH575 Spring 2023 \\
$\bullet$ Allowed time left: 4 years \\
$\bullet$ Resources needed: Bingley computer in Dr. Roundy's lab \\
$\bullet$ Funding: Currently I am a TA, but I'm going to apply for a grant from NSF \\

\end{document}

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
OralExm edits

@@
 $\bullet$ Demonstrated good agreement between our WCA fluid and Monte-Carlo simulations \\
-$\bullet$ Demonstrated good agreement between our WCA fluid and experiments with Argon \\
+$\bullet$ Demonstrated good agreement between our WCA fluid and experiments with Argon under high pressure (so attractive forces can be ignored) \\
 
 \section{Part 3: Research To Do}
-\noindent Toward a better model and a better understanding of our model, I propose to: \\
-$\bullet$ Find the internal energy for our WCA fluid model \\
-$\bullet$ Improve results at low temperature, high density \\ %pressure calculations  \\
+\noindent %Toward a better model and a better understanding of our model, 
+I propose some possible paths for wrapping up my research: \\
+$\bullet$ Finish up the paper we have and graduate. \\
+$\bullet$ Write a follow-on paper and graduate. In a second paper, I would 
+like to look at finding the internal energy for our WCA fluid model using 
+classical Density Functional Theory and our SFMT functional 
+for the Helmholtz free energy that we developed in our first paper 
+\begin{align}
+  U[n(\vec r), s(\vec r)] = F[n(\vec r)] + T\int s(\vec r)
+\end{align}
+ If more is needed, or time allows, I could also: \\
+$\bullet$ Look at how to improve results at low temperature, high density 
+by adjusting our functional and/or potential and our pressure calculations.\\ %pressure calculations  \\
 %$\bullet$ Possibly derive temperature-dependent parameter with B3 and B2  \\

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm another edit
oralexm edit

@@
 $\bullet$ Write a follow-on paper and graduate. In a second paper, I would 
-like to look at finding the internal energy for our WCA fluid model using 
+like to look at finding the internal energy for our WCA fluid using 
 classical Density Functional Theory and our SFMT functional 
@@
 \begin{align}
-  U[n(\vec r), s(\vec r)] = F[n(\vec r)] + T\int s(\vec r)
+  U[n(\vec r), s(\vec r)] = F[n(\vec r)] + T\int s(\vec r) d\vec r
 \end{align}
- If more is needed, or time allows, I could also: \\
+ Alternatively (or if more is needed, or time allows), I could: \\
 $\bullet$ Look at how to improve results at low temperature, high density 

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm edit

@@
 its parameters temperature-dependent (similar to the way Barker-Henderson made the
-diameter of the hard sphere fluid temperature-dependent) and derive them by 
-making the second virial coefficient of our error function potential match the second virial 
-coefficient of a WCA fluid. Our approach is easily extensible to other soft
-repulsive potentials.
+diameter of the hard sphere fluid temperature-dependent).
+One parameter makes the softness of the error function potential comparable 
+with the softness (steepness) of the WCA potential, while the other parameter 
+sets the length scale of the particle interaction and
+is derived by 
+setting the second virial coefficient of our error function potential equal to the second virial 
+coefficient of a WCA fluid. 
+Our approach is easily extensible to other soft repulsive potentials.
 
@@
 
+\noindent Papers would be submitted to Physical Review E.
+
 \section{Part 4: Means/timeline/program}
@@
 $\bullet$ Resources needed: Bingley computer in Dr. Roundy's lab \\
-$\bullet$ Funding: Currently I am a TA, but I'm going to apply for a grant from NSF \\
+$\bullet$ Funding: Currently I am a TA, but I'm going to try applying for a grant from NSF \\
 

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm edit

@@
 with the softness (steepness) of the WCA potential, while the other parameter 
-sets the length scale of the particle interaction and
+sets the length-scale of the particle interaction and
 is derived by 

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm edits

@@
 with the softness (steepness) of the WCA potential, while the other parameter 
-sets the length-scale of the particle interaction and
-is derived by 
-setting the second virial coefficient of our error function potential equal to the second virial 
-coefficient of a WCA fluid. 
+matches the length-scale of the particle interaction and is
+derived by setting the second virial coefficient of our error function 
+potential equal to the second virial coefficient of a WCA fluid. 
 Our approach is easily extensible to other soft repulsive potentials.

### 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
oralexm edits

--- 2022-05-24 papers/fuzzy-fmt/OralExm-researchproposal.tex
@@
 diameter of the hard sphere fluid temperature-dependent). 
-One parameter makes the softness of the error function potential comparable 
-with the softness (steepness) of the WCA potential, while the other parameter 
-sets the length-scale of the particle interaction and
-is derived by 
-setting the second virial coefficient of our error function potential equal to the second virial 
-coefficient of a WCA fluid. 
+%One parameter makes the softness of the error function potential comparable 
+%with the softness (steepness) of the WCA potential, while the other parameter 
+%sets the length-scale of the particle interaction and is
+%derived by setting 
+%We design the parameters make
+Our parameters are derived in part by making 
+the second virial coefficient of our error function 
+potential equal to the second virial coefficient of a WCA fluid. 
 Our approach is easily extensible to other soft repulsive potentials.
